### 2020-05-06 manuscript.Rmd
initial commit

---
title: "Are elevation gradients competition gradients for bumble bees? Yes, no, and sort of."
author: "Sponsler, [...], Steffan-Dewenter"
output: pdf_document
bibliography: refs.bibtex
---

**Introduction**
Bumble bee communities are a classic model system for the study of community assembly processes...elevation gradients offer unique opportunity to study community/interaction dynamics over short spatial scales, controlling for confounding factors...the basic interest in bumble bee communities as a model system for theoretical inquiries has morphed into an applied interest in BB communities as conservation priorities, perhaps particularly in high elevation environments likely to see strong effects of climate change.

Note that we are steering clear of predictions concerning trait-dispersion, since the forces shapign trait dispersion would operate and time scales imcompatible with our intra-seasonal meta-hypothesis.

*H1: No competition.* The hypothesis that floral resource competition is not a significant process in bumble bee community assembly could be taken as a "null hypothesis" when competition *per se* is the phenomenon in question, but it is not "null" in the sense of mere negation of mechanistic explanation. To the contrary, at least two plausible and ecologically meaningful mechanisms can be proposed that would account for a lack of competition in bumble bee communities [@Ranta1981-vx]. First, it is possible that a super-abundance of floral resources could saturate the foraging capacity of a local bumble bee population. Indeed, this may be a common (though transient) phenomenon associated with the bloom of mass flowering crops []. Alternatively, competition could be attenuated when and where abiotic stress, such as low temperature or high elevation, suppress bumble bee populations below competition-inducing densities. This pattern would be a prediction of the "stress-gradient hypothesis" [@Bertness1994-zx], and while it has not yet, to our knowledge, been documented for bumble bees specifically, a recent study of whole bee communities (including bumble bees) found evidence of dampened competition at high elevation [@Hoiss2012-sn].

Under either mechanism, the most straightforward prediction is that bumble bee population density would be decoupled from resource density, since the latter would be non-limiting. For the same reason, the absence of competition would also predict that bumble bee species' dietary niche overlap would not exceed that predicted by a random null model. Finally, under a no-competition scenario, bumble bee species would suffer no displacement from their optimal floral resource species. Thus, holding other aspects of floral resource attractiveness constant (e.g. abundance, nectar volume/concentration), a lack of competition would predict that morphological trait-matching between bumble bee species and their chosen host flowers would exceed that predicted by a random null model.

*H2: Interspecific competition.* The similarity of bumble bee species, both in terms of traits and resource needs, has made them intuitive model organisms for the study of interspecific resource competition, and a rich literature on the topic has revealed a complex and still rather obscure ecological architecture. It is evident that sympatric bumble bees *can* compete for floral resources, and strongly so. @Heinrich1976-wc found that a bumble bee community can, at least under some circumstances, deplete the daily standing nectar pool of a given locality by more than 90%, leaving little room doubt regarding the potential for resource limitation. Other studies have found local bumble bee communities to be over-dispersed with respect to proboscis length [e.g. @Inouye1977-kl], an important matching trait corresponding to floral corolla depth, or phylogeny [@Pellissier2013-pk], which can be interpreted as a catch-all for unmeasured trait variation. Nevertheless, it seems equally evident that symaptric bumble bee species do not *always* compete for floral resources to the point of competitive exclusion, since bumble bee communities with an apparent excess of species richness and deficit of trait dispersion are not uncommon [@Goulson2008-zy].

*H3: Intraspecific competition.* As noted by @Ranta1981-vx, theory predicts that strong intraspecific competition can promote species coexistence through the relief of interspecific competition [@MacArthur1984-qz], and this could, as an alternative to the "no competition" hypothesis, account for cases in which the observed diversity of sympatric bumble bee species exceeds that predicted by trait-based niche models [@Goulson2008-zy].

+-------------------------------+----------------------------------------------------+
| Hypothesis                    | Prediction                                         |
+===============================+====================================================+
| H1: No competition            | 1. BB density !~ FL density                        |
|                               | 2. BB niche overlap = null expectation             |
|                               | 3. BB-FL trait-matching > null expectation         |
+-------------------------------+----------------------------------------------------+
| H2: Intraspecific competition | 1. BB density ~ FL density                         |
|                               | 2. BB diet breadth ~ BB conspecific-packing        |
|                               | 3. BB-FL trait-matching ~ 1/BB conspecific-packing |
+-------------------------------+----------------------------------------------------+
| H3: Intraspecific competition | 1. BB density ~ FL density                         |
|                               | 2.                                                 |
|                               | 3.                                                 |
+-------------------------------+----------------------------------------------------+

*Seasonal-Elevational metahypothesis.*

It is likely that which of these processes prevails -- and which species are favored by the prevailing process -- varies in both space and time, and this heterogenity of process (arising from heterogenity in distribution of bumble bees and their floral resources) could explain the co-occurrence of species that might exclude one another under static conditions [@Ranta1981-vx].

Thus, we formulate a "metahypothesis". Competition need not be frequent to be a powerful driver of community composition [@MacArthur1984-qz].

```{r echo = FALSE, include = FALSE}
library(tidyverse)
```

```{r echo = FALSE}
Time <- c(0,1,2,3)
Elevation <- c(0,1,2,3)
tab <- cbind(Time, Elevation) %>%
  as_tibble()

ggplot(tab, aes(Time, Elevation)) +
  geom_vline(xintercept = c(1,2), color = "white") +
  geom_hline(yintercept = c(1,2), color = "white") +
  scale_x_continuous(expand = c(0,0),
                     breaks = c(0.5, 1.5, 2.5), 
                     labels = c("founding", "build-up", "reproductive"),
                     limits = c(0,3)) +
  scale_y_continuous(expand = c(0,0),
                     breaks = c(0.5, 1.5, 2.5), 
                     labels = c("low", "mid", "high"),
                     limits = c(0,3)) +
  #geom_text() +
  annotate("text", x = 0.5, y = 2.5, label = "H1") +
  annotate("text", x = 1.5, y = 2.5, label = "H1") +
  annotate("text", x = 2.5, y = 2.5, label = "H1") +
  annotate("text", x = 0.5, y = 1.5, label = "H1") +
  annotate("text", x = 0.5, y = 0.5, label = "H1") +
  annotate("text", x = 1.5, y = 1.5, label = "H2 | H3") +
  annotate("text", x = 2.5, y = 1.5, label = "H2 | H3") +
  annotate("text", x = 1.5, y = 0.5, label = "H2 | H3") +
  annotate("text", x = 2.5, y = 0.5, label = "H2 | H3") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

Intraspecific competition should increase diet breadth and decrease trait-matching relative to no competition.

Interspecific competition should decrease diet breath and decrease trait-matching relative to no competition.

Under no competition, foragers would never be displaced, either by conspecifics or heterospecifics, from their optimal resource in a given location and during a given time. This should allow maximal trait-matching (if we ignore potential differences in abundance and/or floral reward level across floral morphologies).

Actually, this might probaby be better framed with respect to phenotypes than to species. Focusing on tongue length, we would predict that competition within a tongue-length class would lead to the broadening of diet breadth, since foragers would tend to be displaced from their optimal resource by competition with other foragers of similar tongue length (whether conspecifics or heterospecifics). Competition between tongue-length classes, however, would tend to narrow diet breadth within tongue-length classes, since a given tongue-length class would be strongly outcompeted on resource that is a better fit for longer- or shorter-tongued competitor. So, intra-guild and inter-guild competition should have opposing effects, the former pushing in the direction of increased diet breadth and decreased trait-matching and the latter in the direction of decreased diet breadth and increased trait-matching. So, if it can be established by independent lines of evidence that competition is indeed occurring, the prevailing type of competition could be inferred by comparing these predictions. If, indeed, my analysis of the relationship between bumble bee density and floral area is a valid indication of competition, I could analyze the deviation of diet breadth and trait matching from null expectations to infer which type of competiton prevails at different points along my elevation gradient (or, perhaps, at different points within the year, though this becomes problematic due to sample size limitations).

I guess from an evolutionary perspective, though, high intra-guild competition may not be a stable situation. So, wide niches and high niche overlap across species within a guild may be more indicative of no competition that of high competition. The difference, I suppose, is that under a no competition scenario, trait-matching would still be strong, whereas under an intra-guild competition scenario, trait-matching would be attenuated due to foragers being displaced from their optimal resources.

"Altitudinal niche-breadth hypothesis" [@Rasmann2014-zb]: Analogous to the latitudinal niche-breadth hypothesis, organisms at high elevations tend to have broader niches due to higher environmental variability, and possibly other reasons. Interestingly, @Rasmann2014-zb do not invoke the concept of competition release as a reason for increased niche breadth at higher elevations.

Bumble bees (Hymenoptera: *Bombus* spp.) are classic model organisms for the study of community assembly, yet there remains a longstanding tension in the literature concerning whether and to what degree bumble bee communities are governed by interspecific competition for floral resources. Conflicting evidence has accumulated, largely along continental lines, with North American studies tending to show a stronger role of competition than their European counterparts [@Goulson2008-zy].

One explanation for this apparent inconsistency, proposed by @Ranta1981-vx, is that the importance of competition varies inversely with that of environmental filtering in response to variation in environmental severity. Under this hypothesis, favorable environmental conditions allow bumble bee populations to increase to the point of floral resource carrying capacity (K), inducing interspecific competition as a constitutive driver of community assembly. In constrast, severe environmental conditions constrain bumble bee populations to sub-K densities at which interspecific competition can be, at most, only a sporadic product of the spatiotemporal heterogeneity of bumble bee and flower distribution. Thus, gradients of environmental severity, such as montane-to-alpine elevation clines, are joint-gradients of community assembly processes, with competition and environmental filtering prevailing, respectively, under mild and severe environmental conditions.

A more generalized form of this hypothesis, i.e. that plant-pollinator networks and/or their constituent communities are governed by competition under mild conditions and environmental filtering under severe conditions, has received ample empirical support, particularly from studies using elevation gradients as a proxy for environmental severity (). Yet there is reason to suspect that bumble bees, despite being the model organisms for

the unusual physiological ecology of bumble bees

for which this hypothesis was first formulated might, ironically, be among the rare exceptions to it. While the vast majority of taxa exhibit an inverse response of diversity to latitude (), bumble bees peak in diversity at higher latitudes (and higher elevations), reflecting their initial diversification in the cooling environment of the Eocene-Oligocene boundary [@Hines2008-er]. In this

\

**Methods**

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
### Prepare environment
library(bipartite)
library(lme4)
library(lmerTest)
library(tidyverse)
library(lubridate)

### Load site data
site_data <- read_csv("./processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten"))) # turn elav.class into an ordered factor

elev_classification <- site_data %>% 
  select(elev.mean, elev.class) %>% 
  unique() %>% 
  arrange(-elev.mean)

### Load network and survey data
bb_traits <- read_csv("./processed_data/bb_traits.csv") %>%
  select(subgenus, bb.sp, pbl.w, pbl.w.class)

net <- read_csv("./processed_data/network.csv") %>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  unite(site.date, site, date, sep = "_", remove = FALSE) %>%
  mutate(year = factor(year)) %>% # convert year to factor
  full_join(bb_traits, by = c("bb.sp.lat" = "bb.sp")) %>%
  left_join(site_data, by = "site")

survey <- read_csv("./processed_data/floral_survey.csv") %>%
  mutate(year = factor(year)) %>% # convert year to factor
  semi_join(net, by = c("site", "plant.sp")) # Consider only plants visited at a given site-date

### Per-transect Bombus density
bb_dens <- net %>%
  group_by(site, date, elev.class, elev.mean, year) %>%
  summarize(bb.dens = n())

### Per-transect floral density
fl_dens <- survey %>%
  group_by(site, date, year) %>%
  summarize(fl.dens = sum(flower.cover),
            log.fl.dens = log(sum(flower.cover)))

### Join BB density and floral cover
density <- bb_dens %>%
  full_join(fl_dens) %>% # Note that there is not perfect alignment in site-dates between the network observations and the floral surveying, so this join will cause some NAs that will need to be removed, causing some transects to drop out of the dataset
  na.omit() %>%
  ungroup %>%
  mutate(bb_per_fl = bb.dens/fl.dens,
         site = factor(site),
         yday = yday(date),
         week = week(date),
         month = month(date))

density_month.agg <- bb_dens %>%
  full_join(fl_dens) %>% # Note that there is not perfect alignment in site-dates between the network observations and the floral surveying, so this join will cause some NAs that will need to be removed, causing some transects to drop out of the dataset
  na.omit() %>%
  ungroup %>%
  mutate(site = factor(site),
         yday = yday(date),
         week = week(date),
         month = month(date)) %>%
  group_by(year, month, site, elev.mean, elev.class) %>%
  summarize(mean.bb.dens = mean(bb.dens),
            mean.fl.dens = mean(fl.dens)) %>%
  mutate(log.mean.fl.dens = log(mean.fl.dens),
         bb.per.fl = mean.bb.dens/mean.fl.dens)
```

*Study sites.*

*Field sampling.*

*Data analysis.* 
\

**Results**

Per-transect bumble bee abundance is positively correlated with per-transect floral cover (log-transformed). To the extent, therefore, that our
metrics of bumble bee and floral density reflect true site-scale abundance and not just co-aggregation, this pattern suggests that bumble bee
density is resource limited, particularly at sites with very low floral density. Importantly, this approach does not distinguish between inter- and intraspecies (or even intra-colony) competition [@Ranta1981-vx].

Note, moreover, that this relationship varies seasonally. For now, we will use Julian month as a stand-in for GDD, which we'll use once we have analyzed our weather station data. Early in the year, the relationship between BB density and floral density is weak; this makes sense because colonies are just getting established during the first couple of months, and BB density is more limited by the autocorrelation component of colony size (bigger colonies can raise more brood) than by floral resource availability. In mid- to late-season, however, BB density becomes strongly dependent on floral resource density. Interestingly, in September, the relationship is strongest and floral density is also lowest.  
```{r eval=FALSE, echo=FALSE}
ggplot(density_month.agg, aes(log.mean.fl.dens, mean.bb.dens, color = year)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~month) +
  ylab("Mean BB density") +
  xlab("log(Mean floral density)")
```

So, we have at least circumstantial evidence that bumble bee density is limited by floral density (again, the key assumption is that our
metrics of bumble bee and floral density reflect true site-scale abundance). What we might as next is whether the degree of dependence of bumble bee density upon floral density vary with elevation. It's hard to draw conclusions about the lowest elevation class, but it seems pretty clear that the slope of bumble bee density on floral density is steep at the highest elevation class than at the middle elevation class; moreover, the relationship is tighter (lower variance) at high elevation. It appears that, contrary to the predictions of the Ranta hypothesis, bumble bees are more dependent on floral resource availability at higher elevations.

```{r eval=FALSE,, echo=FALSE}
ggplot(density_month.agg, aes(log.mean.fl.dens, mean.bb.dens)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(elev.class ~ month) +
  ylab("Mean BB density") +
  xlab("log(Mean floral density)")
```

But let's look at this another way. By dividing bumble bee density by floral density, we can derive an index that represents "bees per unit of floral abundance" (which we willlog-transform for this analysis to correct for skewed distribution). We'll call this our "bee-packing" index, and we'll understand it as an estimate of predicted competition intensity. When we plot this against elevation, parsed by year, we see that "bee-packing" seems to increase at high elevation for 2011, and possibly also in 2010. But the relationship seems to disappear in 2012. This is worth thinking about. 2012 was a year of very low floral density. Perhaps this increased resource limitation across the whole elevation gradient, erasing the effect of elevation.
```{r eval=FALSE}
ggplot(density_month.agg, aes(elev.mean, log(bb.per.fl))) +
  geom_point() +
  geom_smooth(se = TRUE) +
  facet_wrap(~year) +
  ylab("log(bee-packing)") +
  xlab("Elevation") 
```

Some pretty neat patterns emerge when we break our data down by month. For most months, bee-packing increases with elevation, though sometimes the relationship is fairly flat. In Septmeber, however, bee-packing decreases with elevation, presumably due to the elevational staggering of phenology; by that point in the year, only the colder sites still have appreciable floral density.
```{r eval=FALSE, echo=FALSE}
ggplot(density_month.agg, aes(elev.mean, log(bb.per.fl))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(year~month, scales = "free") +
  ylab("log(bee-packing)") +
  xlab("Elevation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Another way to probe the question of competition is to examine network structure, though we will have to sacrifice our temporal resolution to do so. An advantage of this analysis is that we will no longer rely on the estimate of floral density using our survey data, so we can shed the somewhat vulnerable assumption that the correlation between bumble bee density and floral density is not due merely to co-aggregation. To do this, we will focus on the group-level metrics of niche overlap and C-score. Niche overlap would be expected to vary inversely with competition, while C-score -- a measure of ... -- would be expected to vary directly with competition. For this analysis to be sound, though, we need to rarefy our networks. Otherwise, niche overlap and C-score could be driven by uneven sampling effort (less intensively sampled sites will appear to have lower niche overlap and higher C-score).
```{r eval=FALSE}
### Inspect the evenness (or lack thereof) of sampling intensity
net_balance <- net %>%
  group_by(year, elev.mean, elev.class) %>%
  summarize(observations = n(),
            transects = length(unique(date)))

ggplot(net_balance, aes(elev.mean, transects)) +
  geom_point() +
  facet_wrap(~year, scales = "free")
```

A tedious manual rarefaction
```{r eval=FALSE}
### Determine the minimum transect count per site*year to set rarefaction sampling
transects <- net %>%
  select(site, year,date) %>%
  unique()

transects_per_siteyear <- transects %>%
  group_by(site, year) %>%
  summarize(transects = n())

min_transects_per_year <- transects_per_siteyear %>%
  group_by(year) %>%
  summarize(min_transects = min(transects))

### Create rare_net function
rare_net <- function(net, n, iter) {

  out <- data.frame()

  for(i in 1:iter) {
    temp <- net %>%
      select(site, date) %>%
      unique() %>%
      group_by(site) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(grouplevel, weighted = TRUE) %>% # We probably care mainly about the BB level of the network
      data.frame() %>%
      t() %>%
      data.frame() %>%
      rownames_to_column() %>%
      dplyr::select(site = rowname, everything()) %>%
      as_tibble() %>%
      mutate(iteration = rep(i, n())) %>%
      left_join(site_data)

    out <- bind_rows(out, temp)
  } 
  return(as_tibble(out))
}

### Create yearly subsets to feed to rare_net
net_2012 <- filter(net, year == "2012")
net_2011 <- filter(net, year == "2011")
net_2010 <- filter(net, year == "2010")

#### Create rarified networks and tabulate grouplevel metrics
groupmet_rare_2012 <- rare_net(net_2012, n = 9, iter = 10)
groupmet_rare_2011 <- rare_net(net_2011, n = 4, iter = 10)
groupmet_rare_2010 <- rare_net(net_2010, n = 3, iter = 10)

### Plot
ggplot(groupmet_rare_2012, aes(elev.mean, niche.overlap.HL)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~iteration)

ggplot(groupmet_rare_2011, aes(elev.mean, niche.overlap.HL)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~iteration)

ggplot(groupmet_rare_2010, aes(elev.mean, niche.overlap.HL)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~iteration)

ggplot(groupmet_rare_2012, aes(elev.mean, C.score.HL)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~iteration)

ggplot(groupmet_rare_2011, aes(elev.mean, C.score.HL)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~iteration)

ggplot(groupmet_rare_2010, aes(elev.mean, C.score.HL)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~iteration)
```

**Discussion**
Early studies of bumble bee community ecology were framed as basic studies of community assembly rules. In the 20th century, the framing of these studies changed markedly, from the basic ecology of community assembly to the applied ecology of bumble bee convservation.

Competition need not be frequent to be a powerful driver of community composition [@MacArthur1984-qz]. Presumably, this is true at various temporal scales, whether the annual ones that MacArthur mentions or season, daily, etc...

**References**

### 2020-05-06 data_proc.R

@@ -1,0 +1,71 @@
+library(tidyverse)
+library(lubridate)
+
+# Get the data cleaned up and properly organized
+### Floral survey data
+fl_survey <- read_delim("./raw_data/data_plantcover_forDoug_2020.01.17.csv", delim = "\t", 
+                           locale = locale(decimal_mark = ",", grouping_mark = ".")) %>%
+  separate(col = species_plant, into = c("genus", "species", "w", "x", "y", "z"), sep = " ") %>%
+  unite(species_plant, genus, species, sep = " ") %>%
+  select(-c(w,x,y,z)) %>%
+  mutate(species_plant = str_replace_all(species_plant, 
+                                         c("Arabis caerula" = "Arabis caerulea", # correct nomenclature typos and synonyms
+                                           "Arctostaphylos alpina" = "Arctous alpina",
+                                           "Calluna vulagris" = "Calluna vulgaris",
+                                           "Deschampsia caespitosa" = "Deschampsia cespitosa",
+                                           "Gymnadenia bifolia" = "Platanthera bifolia"))) %>%
+  separate(species_plant, c("genus", "species"), remove = FALSE) %>%
+  dplyr::select(year, dayofyear, site = site_name, snowcover, 
+                plant.sp = species_plant, flower.cover = cover_all_flowers) %>%
+  mutate(temp.date = as_date(dayofyear),
+         day = day(temp.date),
+         month = month(temp.date),
+         date = make_date(year, month, day))
+  
+write_csv(fl_survey, "./processed_data/floral_survey.csv")
+
+### Visitation data
+network <- read_delim("./raw_data/data_bumblebee_forDoug_2020.01.17.csv", delim = ";", 
+                           locale = locale(decimal_mark = ",", grouping_mark = ".")) %>%
+  separate(col = plant_sp_latin, into = c("genus", "species", "w", "x", "y", "z"), sep = " ") %>%
+  unite(plant_sp_latin, genus, species, sep = " ") %>%
+  dplyr::select(-c(w,x,y,z)) %>%
+  separate(plant_sp_latin, c("genus", "species"), remove = FALSE) %>%
+  dplyr::select(year, dayofyear, site = site_name, trap.time = trapping_time, pollen, 
+                bb.sp = bumb_species, bb.sp.lat = bumb_sp_latin, plant.sp.abb = forage_plant, 
+                plant.sp = plant_sp_latin) %>%
[…]

### 2020-05-07 manuscript.Rmd
writing

@@
 ---
-title: "Are elevation gradients competition gradients for bumble bees? Yes, no, and sort of."
-author: "Sponsler, [...], Steffan-Dewenter"
+title: "Competitive processes among bumble bees are spatiotemporally heterogeneous"
+author: "Douglas B. Sponsler, Katharina Kallnik, Fabrice Requier, Alice Claßen, A. Fabienne Maihoff, Ingolf Steffan-Dewenter"
 output: pdf_document
@@
 **Introduction**
-Bumble bee communities are a classic model system for the study of community assembly processes...elevation gradients offer unique opportunity to study community/interaction dynamics over short spatial scales, controlling for confounding factors...the basic interest in bumble bee communities as a model system for theoretical inquiries has morphed into an applied interest in BB communities as conservation priorities, perhaps particularly in high elevation environments likely to see strong effects of climate change.
 
-Note that we are steering clear of predictions concerning trait-dispersion, since the forces shapign trait dispersion would operate and time scales imcompatible with our intra-seasonal meta-hypothesis.
+Bumble bees (Hymenoptera: *Bombus* spp.) are classic model organisms for the study of resource competition in biological communities.
+Nevertheless, the role of resource competition in structuring bumble bee communities has proven difficult to generalize.
+It is evident that sympatric bumble bees *can* compete for floral resources, and strongly so.
+@Heinrich1976-wc found that a bumble bee community can deplete the daily standing nectar pool of a given locality by more than 90%, leaving little room to doubt the potential for resource limitation.
+Other studies have found local bumble bee communities to be over-dispersed in terms of tongue length [e.g.
+@Inouye1977-kl], a matching trait [sensu @Schleuning2020-vq] corresponding to floral corolla depth, and phylogeny [@Pellissier2013-pk], which can be interpreted as a catch-all for unmeasured trait variation.
+Nevertheless, it is equally evident that sympatric bumble bee species do not *always* compete for floral resources to the point of competitive exclusion, since bumble bee communities with an apparent excess of species richness and deficit of trait dispersion are commonly observed [@Goulson2008-zy].
 
-*H1: No competition.* The hypothesis that floral resource competition is not a significant process in bumble bee community assembly could be taken as a "null hypothesis" when competition *per se* is the phenomenon in question, but it is not "null" in the sense of mere negation of mechanistic explanation.
-To the contrary, at least two plausible and ecologically meaningful mechanisms can be proposed that would account for a lack of competition in bumble bee communities [@Ranta1981-vx].
-First, it is possible that a super-abundance of floral resources could saturate the foraging capacity of a local bumble bee population.
-Indeed, this may be a common (though transient) phenomenon associated with the bloom of mass flowering crops [].
-Alternatively, competition could be attenuated when and where abiotic stress, such as low temperature or high elevation, suppress bumble bee populations below competition-inducing densities.
-This pattern would be a prediction of the "stress-gradient hypothesis" [@Bertness1994-zx], and while it has not yet, to our knowledge, been documented for bumble bees specifically, a recent study of whole bee communities (including bumble bees) found evidence of dampened competition at high elevation [@Hoiss2012-sn].
-
-Under either mechanism, the most straightforward prediction is that bumble bee population density would be decoupled from resource density, since the latter would be non-limiting.
-For the same reason, the absence of competition would also predict that bumble bee species' dietary niche overlap would not exceed that predicted by a random null model.
-Finally, under a no-competition scenario, bumble bee species would suffer no displacement from their optimal floral resource species.
-Thus, holding other aspects of floral resource attractiveness constant (e.g.
-abundance, nectar volume/concentration), a lack of competition would predict that morphological trait-matching between bumble bee species and their chosen host flowers would exceed that predicted by a random null model.
-
-*H2: Interspecific competition.* The similarity of bumble bee species, both in terms of traits and resource needs, has made them intuitive model organisms for the study of interspecific resource competition, and a rich literature on the topic has revealed a complex and still rather obscure ecological architecture.
-It is evident that sympatric bumble bees *can* compete for floral resources, and strongly so.
-@Heinrich1976-wc found that a bumble bee community can, at least under some circumstances, deplete the daily standing nectar pool of a given locality by more than 90%, leaving little room doubt regarding the potential for resource limitation.
-Other studies have found local bumble bee communities to be over-dispersed with respect to proboscis length [e.g.
-@Inouye1977-kl], an important matching trait corresponding to floral corolla depth, or phylogeny [@Pellissier2013-pk], which can be interpreted as a catch-all for unmeasured trait variation.
-Nevertheless, it seems equally evident that symaptric bumble bee species do not *always* compete for floral resources to the point of competitive exclusion, since bumble bee communities with an apparent excess of species richness and deficit of trait dispersion are not uncommon [@Goulson2008-zy].
-
-*H3: Intraspecific competition.* As noted by @Ranta1981-vx, theory predicts that strong intraspecific competition can promote species coexistence through the relief of interspecific competition [@MacArthur1984-qz], and this could, as an alternative to the "no competition" hypothesis, account for cases in which the observed diversity of sympatric bumble bee species exceeds that predicted by trait-based niche models [@Goulson2008-zy].
+To account for both the importance and variability of competitive processes in bumble bee communities, it is helpful to distinguish three processes that could obtain to varying degrees: no competition, interspecific competition, and intraspecific competiion (**Table 1**).
+We will refer to these processes as hypotheses, since their realization in any given community is matter of inference.
+Below, we elaborate each and formulate corresponding empirical predictons.
 
@@
 | H1: No competition            | 1. BB density !~ FL density                        |
-|                               | 2. BB niche overlap = null expectation             |
+|                               | 2. BB trait dispersion = null expectation          |
 |                               | 3. BB-FL trait-matching > null expectation         |
 +-------------------------------+----------------------------------------------------+
-| H2: Intraspecific competition | 1. BB density ~ FL density                         |
+| H2: Interspecific competition | 1. BB density ~ FL density                         |
+|                               | 2. BB trait dispersion > null expectation          |
+|                               | 3. BB-FL trait-matching > null expectation         |
++-------------------------------+----------------------------------------------------+
+| H3: Intraspecific competition | 1. BB density ~ FL density                         |
 |                               | 2. BB diet breadth ~ BB conspecific-packing        |
@@
 +-------------------------------+----------------------------------------------------+
-| H3: Intraspecific competition | 1. BB density ~ FL density                         |
-|                               | 2.                                                 |
-|                               | 3.                                                 |
-+-------------------------------+----------------------------------------------------+
 
-*Seasonal-Elevational metahypothesis.*
+**Table 1:** Hypothesized competitive processes with empirical predictions. BB = bumble bee; FL = floral. \
 
-It is likely that which of these processes prevails -- and which species are favored by the prevailing process -- varies in both space and time, and this heterogenity of process (arising from heterogenity in distribution of bumble bees and their floral resources) could explain the co-occurrence of species that might exclude one another under static conditions [@Ranta1981-vx].
+*H1: No competition.* The hypothesis that floral resource competition is not a significant process governing a bumble bee community functions as a "null hypothesis" when competition *per se* is the phenomenon in question, but it is not "null" in the sense of mere negation of mechanistic explanation.
+To the contrary, at least two plausible and ecologically meaningful mechanisms can be proposed that would account for a lack of competition in bumble bee communities [@Ranta1981-vx].
+First, it is possible that a super-abundance of floral resources could saturate the foraging capacity of a local bumble bee population.
+Indeed, this may be a common (though transient) phenomenon associated with the bloom of mass flowering crops [@Ranta1981-vx].
+Alternatively, competition could be attenuated when and where abiotic stress, such as low temperature, suppresses bumble bee populations below competition-inducing densities.
+This pattern would be a prediction of the "stress-gradient hypothesis" [@Bertness1994-zx], and while it has not, to our knowledge, been documented for bumble bees specifically, a recent study of whole bee communities (of which bumble bees were a part) found evidence of dampened competition at high elevation sires along a montane-to-alpine gradient [@Hoiss2012-sn].
+Under either mechanism, the most straightforward prediction is that bumble bee population density would be decoupled from resource density, since the latter would be non-limiting.
+For the same reason, the absence of competition would also predict that the trait dispersion of sympatric bumble bee species would not exceed that predicted by a random null model.
+Third, under a no-competition scenario, bumble bee species would suffer no displacement from their optimal floral resource species.
+Thus, holding other aspects of floral resource attractiveness constant (e.g.
+abundance, nectar volume/concentration), a lack of competition would predict that morphological trait-matching between bumble bee species and their chosen host flowers would exceed that predicted by a random null model.
 
-Thus, we formulate a "metahypothesis". Competition need not be frequent to be a powerful driver of community composition [@MacArthur1984-qz].
+*H2: Interspecific competition.* Resource partitioning by bumble bees is mediated by variation in matching traits, the most well-studied of these being tongue length.
+Bumble bee tongue lengths vary between species, and the efficiency with which a bumble bee can collect nectar from a given flower depends largely on the closeness of match between the tongue length of the bee and the corolla depth of the flower [@Inouye1980-mv; @Ranta1980-ok].
+Thus, under conditions of interspecific competition, a species' "defensible" niche space (i.e.
+the space in which it is a superior competitor) should center on the floral trait space that corresponds to its tongue length, and the number of species that can be "packed" into a given locality should be limited by some minimal differentiation of tongue length [@Inouye1977-kl; @Pyke1982-gn] within the range of variation supported by the trait space of available flora.
+Since resource limitation is a necessary condition for resource competition, a key prediction distiguishing interspecific competition from a lack of competition is that bumble bee density should vary in proportion to floral resource density.
+A bumble bee community under interspecific competition should also exhibit higher tongue length dispersion and stronger trait matching relative to the predictions of a random null model [@Inouye1977-kl].
+
+*H3: Intraspecific competition.* As noted by @Ranta1981-vx, theory predicts that intraspecific competition can promote species coexistence through the relief of interspecific competition [@MacArthur1984-qz], and could thus, as an alternative to the "no competition" hypothesis, account for cases in which the observed diversity of sympatric bumble bee species exceeds that predicted by trait-based niche models.
+Under the pressure of intraspecific competition, individuals face the depletion of preferred resources and may respond by foraging more opportunistically within the constraints of their fundamental niche space [].
+Alternatively, a population may relieve intraspecific competition by subdividing, on the basis of intraspecific trait heterogeneity, into differently specialized "guilds" [@Araujo2008-bp].
+In either case, the population-level effect would be an increase in the species' realized niche breadth [though see @Parent2014-hx].
+Thus, the predictions of intraspecific competition can be formulated as follows.
+First, as in interspecific competition, resources must be limiting, so bumble bee density should vary in proportion to floral density.
+In the case of intraspecific competition, though, we can refine this prediction to say that the density of a given bumble bee species should vary in proportion to the density of the portion of the floral resource community belonging to its dietary niche, since the effects of intraspecific competition are realized per-species.
+Second, the dietary breadth of a given species should vary in proportion to the density of conspecifics per density of floral resources (we will refer to this as a "bee-packing" index).
+Conversely, a broadening of diet should entail a relaxation of trait-matching, so the degree of trait-matching should vary inversely with the density of conspecifics per density of floral resources.
+
+*Seasonal-Elevational metahypothesis.* That any of the processes hypothesized above should obtain universally is both empirically unsupported and theoretically doubtful [@Ranta1981-vx].
+More plausibly, which of these processes prevails -- and which species are favored by the prevailing process -- vary in both space and time, and this heterogenity of process could explain the idiosyncrasies of observed bumble bee communities [@Ranta1981-vx].
+Rather than viewing this heterogeneity as an obstacle to inferring some universal rule of bumble bee community assembly, our aim in the present study is to investigate the heterogeneity itself as an ecological phenomenon, characterizing the variation in competitive processes (or lack thereof) in bumble bee communities through time and space.
+
+Our study is set in the Berchtesgadener Alps of southwestern Germany, where a steep elevation gradient permits the study of wide environmental variability while minimizing the confounding effects of geographical distance [e.g.
+@Hoiss2015-gv].
+In this context, we formulate a "metahypothesis" concerning the variation of hypothesized competition processes through elevation and time (**Figure 1**).
+Consistent with the stress-gradient hypothesis [@Bertness1994-zx] and with previous empirical work in our study system [@Hoiss2012-sn], we hypothesize an inverse relationship between competition and elevation.
+Along a temporal axis, we hypothesize that the strength of competition is structured by the seasonal life history pattern of bumble bee colonies.
+Early in the season, during the colony founding phase, foraging bees consist only of foundress queens, and population densities are low.
+With the emergence of the first cohort of worker brood, a colony enters the build-up phase, with workers taking over the majority of foraging and population densities increasing.
+...reproductive phase.
+
+We propose no *a priori* hypothesis concerning which competitive process, interspecific or intraspecific, will predominate with respect to our spatiotemporal gradient.
+Indeed, it is likely that interspecific competition cannot occur in the absence of intraspecific competition.
+
+In addition to clarifying the basic ecology of bumble bee community assembly, parsing competitive processes in this way has an applied significance for the conservation of bumble bees in the changing environments of the Anthropocene.
+Under the forcing of environmental change, such as climate warming or land use conversion, bumble bee communities might not be affected only in terms of range shifts, but also by a restructuring of competitive interactions, the nature and consequences of which may be difficult to predict.
 
@@
 
-Intraspecific competition should increase diet breadth and decrease trait-matching relative to no competition.
-
-Interspecific competition should decrease diet breath and decrease trait-matching relative to no competition.
-
-Under no competition, foragers would never be displaced, either by conspecifics or heterospecifics, from their optimal resource in a given location and during a given time.
-This should allow maximal trait-matching (if we ignore potential differences in abundance and/or floral reward level across floral morphologies).
-
-Actually, this might probaby be better framed with respect to phenotypes than to species.
-Focusing on tongue length, we would predict that competition within a tongue-length class would lead to the broadening of diet breadth, since foragers would tend to be displaced from their optimal resource by competition with other foragers of similar tongue length (whether conspecifics or heterospecifics).
-Competition between tongue-length classes, however, would tend to narrow diet breadth within tongue-length classes, since a given tongue-length class would be strongly outcompeted on resource that is a better fit for longer- or shorter-tongued competitor.
-So, intra-guild and inter-guild competition should have opposing effects, the former pushing in the direction of increased diet breadth and decreased trait-matching and the latter in the direction of decreased diet breadth and increased trait-matching.
-So, if it can be established by independent lines of evidence that competition is indeed occurring, the prevailing type of competition could be inferred by comparing these predictions.
-If, indeed, my analysis of the relationship between bumble bee density and floral area is a valid indication of competition, I could analyze the deviation of diet breadth and trait matching from null expectations to infer which type of competiton prevails at different points along my elevation gradient (or, perhaps, at different points within the year, though this becomes problematic due to sample size limitations).
-
-I guess from an evolutionary perspective, though, high intra-guild competition may not be a stable situation.
-So, wide niches and high niche overlap across species within a guild may be more indicative of no competition that of high competition.
-The difference, I suppose, is that under a no competition scenario, trait-matching would still be strong, whereas under an intra-guild competition scenario, trait-matching would be attenuated due to foragers being displaced from their optimal resources.
-
-"Altitudinal niche-breadth hypothesis" [@Rasmann2014-zb]: Analogous to the latitudinal niche-breadth hypothesis, organisms at high elevations tend to have broader niches due to higher environmental variability, and possibly other reasons.
-Interestingly, @Rasmann2014-zb do not invoke the concept of competition release as a reason for increased niche breadth at higher elevations.
-
-Bumble bees (Hymenoptera: *Bombus* spp.) are classic model organisms for the study of community assembly, yet there remains a longstanding tension in the literature concerning whether and to what degree bumble bee communities are governed by interspecific competition for floral resources.
-Conflicting evidence has accumulated, largely along continental lines, with North American studies tending to show a stronger role of competition than their European counterparts [@Goulson2008-zy].
-
-One explanation for this apparent inconsistency, proposed by @Ranta1981-vx, is that the importance of competition varies inversely with that of environmental filtering in response to variation in environmental severity.
-Under this hypothesis, favorable environmental conditions allow bumble bee populations to increase to the point of floral resource carrying capacity (K), inducing interspecific competition as a constitutive driver of community assembly.
-In constrast, severe environmental conditions constrain bumble bee populations to sub-K densities at which interspecific competition can be, at most, only a sporadic product of the spatiotemporal heterogeneity of bumble bee and flower distribution.
-Thus, gradients of environmental severity, such as montane-to-alpine elevation clines, are joint-gradients of community assembly processes, with competition and environmental filtering prevailing, respectively, under mild and severe environmental conditions.
-
-A more generalized form of this hypothesis, i.e.
-that plant-pollinator networks and/or their constituent communities are governed by competition under mild conditions and environmental filtering under severe conditions, has received ample empirical support, particularly from studies using elevation gradients as a proxy for environmental severity ().
-Yet there is reason to suspect that bumble bees, despite being the model organisms for
-
-the unusual physiological ecology of bumble bees
-
-for which this hypothesis was first formulated might, ironically, be among the rare exceptions to it.
-While the vast majority of taxa exhibit an inverse response of diversity to latitude (), bumble bees peak in diversity at higher latitudes (and higher elevations), reflecting their initial diversification in the cooling environment of the Eocene-Oligocene boundary [@Hines2008-er].
-In this
+**Figure 1:** Hypothesized variation in prevailing competitive processes through time and elevation. \
 
@@
 
-**Methods**
+<!-- **Methods**  -->
 
@@
 
-*Study sites.*
+<!-- *Study sites.*  -->
 
-*Field sampling.*
+<!-- *Field sampling.*  -->
 
-*Data analysis.* 
-\
+<!-- *Data analysis.*  -->
+<!-- \ -->
 
-**Results**
+<!-- **Results** -->
 
-Per-transect bumble bee abundance is positively correlated with per-transect floral cover (log-transformed). To the extent, therefore, that our
-metrics of bumble bee and floral density reflect true site-scale abundance and not just co-aggregation, this pattern suggests that bumble bee
-density is resource limited, particularly at sites with very low floral density.
+<!-- Per-transect bumble bee abundance is positively correlated with per-transect floral cover (log-transformed). To the extent, therefore, that our -->
+<!-- metrics of bumble bee and floral density reflect true site-scale abundance and not just co-aggregation, this pattern suggests that bumble bee -->
+<!-- density is resource limited, particularly at sites with very low floral density.
 Importantly, this approach does not distinguish between inter- and intraspecies (or even intra-colony) competition [@Ranta1981-vx].
+-->
 
-Note, moreover, that this relationship varies seasonally.
+<!-- Note, moreover, that this relationship varies seasonally.
 For now, we will use Julian month as a stand-in for GDD, which we'll use once we have analyzed our weather station data.
@@
 Interestingly, in September, the relationship is strongest and floral density is also lowest.
+-->
 ```{r eval=FALSE, echo=FALSE}
@@
 
-So, we have at least circumstantial evidence that bumble bee density is limited by floral density (again, the key assumption is that our
-metrics of bumble bee and floral density reflect true site-scale abundance).
+<!-- So, we have at least circumstantial evidence that bumble bee density is limited by floral density (again, the key assumption is that our -->
+<!-- metrics of bumble bee and floral density reflect true site-scale abundance).
 What we might as next is whether the degree of dependence of bumble bee density upon floral density vary with elevation.
@@
 It appears that, contrary to the predictions of the Ranta hypothesis, bumble bees are more dependent on floral resource availability at higher elevations.
+-->
 
-```{r eval=FALSE,, echo=FALSE}
+```{r eval=FALSE, echo=FALSE}
 ggplot(density_month.agg, aes(log.mean.fl.dens, mean.bb.dens)) +
@@
 
-But let's look at this another way.
+<!-- But let's look at this another way.
 By dividing bumble bee density by floral density, we can derive an index that represents "bees per unit of floral abundance" (which we willlog-transform for this analysis to correct for skewed distribution).
@@
 Perhaps this increased resource limitation across the whole elevation gradient, erasing the effect of elevation.
-```{r eval=FALSE}
+-->
+```{r eval=FALSE, echo = FALSE}
 ggplot(density_month.agg, aes(elev.mean, log(bb.per.fl))) +
@@
 
-Some pretty neat patterns emerge when we break our data down by month.
+<!-- Some pretty neat patterns emerge when we break our data down by month.
 For most months, bee-packing increases with elevation, though sometimes the relationship is fairly flat.
 In Septmeber, however, bee-packing decreases with elevation, presumably due to the elevational staggering of phenology; by that point in the year, only the colder sites still have appreciable floral density.
+-->
 ```{r eval=FALSE, echo=FALSE}
@@
 
-Another way to probe the question of competition is to examine network structure, though we will have to sacrifice our temporal resolution to do so.
+<!-- Another way to probe the question of competition is to examine network structure, though we will have to sacrifice our temporal resolution to do so.
 An advantage of this analysis is that we will no longer rely on the estimate of floral density using our survey data, so we can shed the somewhat vulnerable assumption that the correlation between bumble bee density and floral density is not due merely to co-aggregation.
@@
 Otherwise, niche overlap and C-score could be driven by uneven sampling effort (less intensively sampled sites will appear to have lower niche overlap and higher C-score).
-```{r eval=FALSE}
+-->
+```{r eval=FALSE, echo = FALSE}
 ### Inspect the evenness (or lack thereof) of sampling intensity
@@
 
-A tedious manual rarefaction
-```{r eval=FALSE}
+<!-- A tedious manual rarefaction -->
+```{r eval=FALSE, echo = FALSE}
 ### Determine the minimum transect count per site*year to set rarefaction sampling
@@
 
-**Discussion**
-Early studies of bumble bee community ecology were framed as basic studies of community assembly rules.
-In the 20th century, the framing of these studies changed markedly, from the basic ecology of community assembly to the applied ecology of bumble bee convservation.
+<!-- **Discussion** -->
 
-Competition need not be frequent to be a powerful driver of community composition [@MacArthur1984-qz].
-Presumably, this is true at various temporal scales, whether the annual ones that MacArthur mentions or season, daily, etc...
+<!-- Under conditions of resource limitation, intraspecific competition may be a stronger constraint on bumble bee density (though not on diversity) than interspecific competition, since conspecific are limited in their ability to partition further their potential niche space.
+-->
+
+<!-- "Altitudinal niche-breadth hypothesis" [@Rasmann2014-zb]: Analogous to the latitudinal niche-breadth hypothesis, organisms at high elevations tend to have broader niches due to higher environmental variability, and possibly other reasons.
+Interestingly, @Rasmann2014-zb do not invoke the concept of competition release as a reason for increased niche breadth at higher elevations.
+-->
+
+<!-- It is important to clarify what we are and are not sampling.
+We are not sampling bumble bee communities; we are sampling spatial (and temporal) subsets of the floral resource landscape.
+So, the questions we are able to answer concern not the composition of the community as such, however one chooses to delimit it, but rather the composition of foragers within an arbitrarily delimited (but standardized) segment of the resource landscape (i.e.
+the sampling plot).
+-->
+
+<!-- Importantly, though, competition need not be universal, or even frequent, to be a powerful driver of community assembly [@MacArthur1984-qz]. -->
+
+<!-- Moreover, trait matching under interspecific competition might also exceed that predicted under no competition, since interspecific competition would limit a species' ability to "adaptively mismatch" to exploit a rich but poorly matched resource, since the resource would likely be monopolized by a better-matched (and therefore more competitive) species.
+-->
+
+<!-- Am I right in thinking that it is impossible to have interspecific competition without also having intraspecific competition?
+These processes would have opposite effects on niche breadth.
+By itself, interspecific competition would favor the maximal contraction of niche breadth until no overlap remained.
+But intraspecific competition would have the opposite effect, forcing conspecifics to broaden their niches to minimize competition with one another.
+Presumably, observed niche breadth and niche overlap represent a dynamic equilibrium of these opposing forces.
+-->
 

### 2020-05-11 manuscript.Rmd
added glm breakpoint analysis

@@
 ---
-title: "Competitive processes among bumble bees are spatiotemporally heterogeneous"
-author: "Douglas B. Sponsler, Katharina Kallnik, Fabrice Requier, Alice Claßen, A. Fabienne Maihoff, Ingolf Steffan-Dewenter"
-output: pdf_document
-bibliography: refs.bibtex
+title: Bumble bee communities are structured by a spatiotemporal mosaic of competitive
+  processes
+author:
+- Douglas B. Sponsler
+- Katharina Kallnik
+- Fabrice Requier
+- Alice Claßen
+- A. Fabienne Maihoff
+- Ingolf Steffan-Dewenter
+output:
+  word_document:
+    reference_docx: docx_template.docx
+bibliography: refs.bib
+
 ---
 
-**Introduction**
-
-Bumble bees (Hymenoptera: *Bombus* spp.) are classic model organisms for the study of resource competition in biological communities.
-Nevertheless, the role of resource competition in structuring bumble bee communities has proven difficult to generalize.
-It is evident that sympatric bumble bees *can* compete for floral resources, and strongly so.
-@Heinrich1976-wc found that a bumble bee community can deplete the daily standing nectar pool of a given locality by more than 90%, leaving little room to doubt the potential for resource limitation.
-Other studies have found local bumble bee communities to be over-dispersed in terms of tongue length [e.g.
-@Inouye1977-kl], a matching trait [sensu @Schleuning2020-vq] corresponding to floral corolla depth, and phylogeny [@Pellissier2013-pk], which can be interpreted as a catch-all for unmeasured trait variation.
-Nevertheless, it is equally evident that sympatric bumble bee species do not *always* compete for floral resources to the point of competitive exclusion, since bumble bee communities with an apparent excess of species richness and deficit of trait dispersion are commonly observed [@Goulson2008-zy].
-
-To account for both the importance and variability of competitive processes in bumble bee communities, it is helpful to distinguish three processes that could obtain to varying degrees: no competition, interspecific competition, and intraspecific competiion (**Table 1**).
-We will refer to these processes as hypotheses, since their realization in any given community is matter of inference.
-Below, we elaborate each and formulate corresponding empirical predictons.
-
-+-------------------------------+----------------------------------------------------+
-| Hypothesis                    | Prediction                                         |
-+===============================+====================================================+
-| H1: No competition            | 1. BB density !~ FL density                        |
-|                               | 2. BB trait dispersion = null expectation          |
-|                               | 3. BB-FL trait-matching > null expectation         |
-+-------------------------------+----------------------------------------------------+
-| H2: Interspecific competition | 1. BB density ~ FL density                         |
-|                               | 2. BB trait dispersion > null expectation          |
-|                               | 3. BB-FL trait-matching > null expectation         |
-+-------------------------------+----------------------------------------------------+
-| H3: Intraspecific competition | 1. BB density ~ FL density                         |
-|                               | 2. BB diet breadth ~ BB conspecific-packing        |
-|                               | 3. BB-FL trait-matching ~ 1/BB conspecific-packing |
-+-------------------------------+----------------------------------------------------+
-
-**Table 1:** Hypothesized competitive processes with empirical predictions. BB = bumble bee; FL = floral. \
-
-*H1: No competition.* The hypothesis that floral resource competition is not a significant process governing a bumble bee community functions as a "null hypothesis" when competition *per se* is the phenomenon in question, but it is not "null" in the sense of mere negation of mechanistic explanation.
-To the contrary, at least two plausible and ecologically meaningful mechanisms can be proposed that would account for a lack of competition in bumble bee communities [@Ranta1981-vx].
-First, it is possible that a super-abundance of floral resources could saturate the foraging capacity of a local bumble bee population.
-Indeed, this may be a common (though transient) phenomenon associated with the bloom of mass flowering crops [@Ranta1981-vx].
-Alternatively, competition could be attenuated when and where abiotic stress, such as low temperature, suppresses bumble bee populations below competition-inducing densities.
-This pattern would be a prediction of the "stress-gradient hypothesis" [@Bertness1994-zx], and while it has not, to our knowledge, been documented for bumble bees specifically, a recent study of whole bee communities (of which bumble bees were a part) found evidence of dampened competition at high elevation sires along a montane-to-alpine gradient [@Hoiss2012-sn].
-Under either mechanism, the most straightforward prediction is that bumble bee population density would be decoupled from resource density, since the latter would be non-limiting.
-For the same reason, the absence of competition would also predict that the trait dispersion of sympatric bumble bee species would not exceed that predicted by a random null model.
-Third, under a no-competition scenario, bumble bee species would suffer no displacement from their optimal floral resource species.
-Thus, holding other aspects of floral resource attractiveness constant (e.g.
-abundance, nectar volume/concentration), a lack of competition would predict that morphological trait-matching between bumble bee species and their chosen host flowers would exceed that predicted by a random null model.
-
-*H2: Interspecific competition.* Resource partitioning by bumble bees is mediated by variation in matching traits, the most well-studied of these being tongue length.
-Bumble bee tongue lengths vary between species, and the efficiency with which a bumble bee can collect nectar from a given flower depends largely on the closeness of match between the tongue length of the bee and the corolla depth of the flower [@Inouye1980-mv; @Ranta1980-ok].
-Thus, under conditions of interspecific competition, a species' "defensible" niche space (i.e.
-the space in which it is a superior competitor) should center on the floral trait space that corresponds to its tongue length, and the number of species that can be "packed" into a given locality should be limited by some minimal differentiation of tongue length [@Inouye1977-kl; @Pyke1982-gn] within the range of variation supported by the trait space of available flora.
-Since resource limitation is a necessary condition for resource competition, a key prediction distiguishing interspecific competition from a lack of competition is that bumble bee density should vary in proportion to floral resource density.
-A bumble bee community under interspecific competition should also exhibit higher tongue length dispersion and stronger trait matching relative to the predictions of a random null model [@Inouye1977-kl].
-
-*H3: Intraspecific competition.* As noted by @Ranta1981-vx, theory predicts that intraspecific competition can promote species coexistence through the relief of interspecific competition [@MacArthur1984-qz], and could thus, as an alternative to the "no competition" hypothesis, account for cases in which the observed diversity of sympatric bumble bee species exceeds that predicted by trait-based niche models.
-Under the pressure of intraspecific competition, individuals face the depletion of preferred resources and may respond by foraging more opportunistically within the constraints of their fundamental niche space [].
-Alternatively, a population may relieve intraspecific competition by subdividing, on the basis of intraspecific trait heterogeneity, into differently specialized "guilds" [@Araujo2008-bp].
-In either case, the population-level effect would be an increase in the species' realized niche breadth [though see @Parent2014-hx].
-Thus, the predictions of intraspecific competition can be formulated as follows.
-First, as in interspecific competition, resources must be limiting, so bumble bee density should vary in proportion to floral density.
-In the case of intraspecific competition, though, we can refine this prediction to say that the density of a given bumble bee species should vary in proportion to the density of the portion of the floral resource community belonging to its dietary niche, since the effects of intraspecific competition are realized per-species.
-Second, the dietary breadth of a given species should vary in proportion to the density of conspecifics per density of floral resources (we will refer to this as a "bee-packing" index).
-Conversely, a broadening of diet should entail a relaxation of trait-matching, so the degree of trait-matching should vary inversely with the density of conspecifics per density of floral resources.
-
-*Seasonal-Elevational metahypothesis.* That any of the processes hypothesized above should obtain universally is both empirically unsupported and theoretically doubtful [@Ranta1981-vx].
-More plausibly, which of these processes prevails -- and which species are favored by the prevailing process -- vary in both space and time, and this heterogenity of process could explain the idiosyncrasies of observed bumble bee communities [@Ranta1981-vx].
-Rather than viewing this heterogeneity as an obstacle to inferring some universal rule of bumble bee community assembly, our aim in the present study is to investigate the heterogeneity itself as an ecological phenomenon, characterizing the variation in competitive processes (or lack thereof) in bumble bee communities through time and space.
-
-Our study is set in the Berchtesgadener Alps of southwestern Germany, where a steep elevation gradient permits the study of wide environmental variability while minimizing the confounding effects of geographical distance [e.g.
-@Hoiss2015-gv].
-In this context, we formulate a "metahypothesis" concerning the variation of hypothesized competition processes through elevation and time (**Figure 1**).
-Consistent with the stress-gradient hypothesis [@Bertness1994-zx] and with previous empirical work in our study system [@Hoiss2012-sn], we hypothesize an inverse relationship between competition and elevation.
-Along a temporal axis, we hypothesize that the strength of competition is structured by the seasonal life history pattern of bumble bee colonies.
-Early in the season, during the colony founding phase, foraging bees consist only of foundress queens, and population densities are low.
-With the emergence of the first cohort of worker brood, a colony enters the build-up phase, with workers taking over the majority of foraging and population densities increasing.
-...reproductive phase.
-
-We propose no *a priori* hypothesis concerning which competitive process, interspecific or intraspecific, will predominate with respect to our spatiotemporal gradient.
-Indeed, it is likely that interspecific competition cannot occur in the absence of intraspecific competition.
-
-In addition to clarifying the basic ecology of bumble bee community assembly, parsing competitive processes in this way has an applied significance for the conservation of bumble bees in the changing environments of the Anthropocene.
-Under the forcing of environmental change, such as climate warming or land use conversion, bumble bee communities might not be affected only in terms of range shifts, but also by a restructuring of competitive interactions, the nature and consequences of which may be difficult to predict.
-
 ```{r echo = FALSE, include = FALSE}
-library(tidyverse)
-```
-
-```{r echo = FALSE}
-Time <- c(0,1,2,3)
-Elevation <- c(0,1,2,3)
-tab <- cbind(Time, Elevation) %>%
-  as_tibble()
-
-ggplot(tab, aes(Time, Elevation)) +
-  geom_vline(xintercept = c(1,2), color = "white") +
-  geom_hline(yintercept = c(1,2), color = "white") +
-  scale_x_continuous(expand = c(0,0),
-                     breaks = c(0.5, 1.5, 2.5), 
-                     labels = c("founding", "build-up", "reproductive"),
-                     limits = c(0,3)) +
-  scale_y_continuous(expand = c(0,0),
-                     breaks = c(0.5, 1.5, 2.5), 
-                     labels = c("low", "mid", "high"),
-                     limits = c(0,3)) +
-  #geom_text() +
-  annotate("text", x = 0.5, y = 2.5, label = "H1") +
-  annotate("text", x = 1.5, y = 2.5, label = "H1") +
-  annotate("text", x = 2.5, y = 2.5, label = "H1") +
-  annotate("text", x = 0.5, y = 1.5, label = "H1") +
-  annotate("text", x = 0.5, y = 0.5, label = "H1") +
-  annotate("text", x = 1.5, y = 1.5, label = "H2 | H3") +
-  annotate("text", x = 2.5, y = 1.5, label = "H2 | H3") +
-  annotate("text", x = 1.5, y = 0.5, label = "H2 | H3") +
-  annotate("text", x = 2.5, y = 0.5, label = "H2 | H3") +
-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
-```
-
-**Figure 1:** Hypothesized variation in prevailing competitive processes through time and elevation. \
-
-\
-
-<!-- **Methods**  -->
-
-```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
 ### Prepare environment
@@
 site_data <- read_csv("./processed_data/site_data.csv") %>%
-  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max)) %>% # drop these variables
-  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten"))) # turn elav.class into an ordered factor
+  #dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
+  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
+         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) # turn elav.class into an ordered factor
 
@@
   arrange(-elev.mean)
+
+### Climate data
+gdd <- read_csv("./processed_data/climate.csv") %>%
+  mutate(year = factor(year)) %>%
+  select(site, date, gdd.cum)
 
@@
   full_join(bb_traits, by = c("bb.sp.lat" = "bb.sp")) %>%
-  left_join(site_data, by = "site")
+  left_join(site_data, by = "site") %>%
+  left_join(gdd, by = c("site", "date")) %>%
+  arrange(site, date)
 
@@
   mutate(year = factor(year)) %>% # convert year to factor
-  semi_join(net, by = c("site", "plant.sp")) # Consider only plants visited at a given site-date
+  semi_join(net, by = c("site", "plant.sp")) %>% # Consider only plants visited at a given site-date
+  left_join(gdd, by = c("site", "date")) %>%
+  arrange(site, date)
 
@@
 
+**Introduction**
+
+Bumble bees (Hymenoptera: *Bombus* spp.) are classic model organisms for the study of resource competition in biological communities.
+Nevertheless, the role of resource competition in structuring bumble bee communities has proven difficult to generalize.
+It is evident that sympatric bumble bees *can* compete for floral resources, and strongly so.
+@Heinrich1976-wc found that a bumble bee community can deplete the daily standing nectar pool of a given locality by more than 90%, and starvation has been implicated in the demise of colonies in isolated subalpine meadows [@Bowers1985-df], leaving little room to doubt the potential for resource limitation.
+Other studies have found local bumble bee communities to be over-dispersed in terms of tongue length [e.g.
+@Inouye1977-kl], a matching trait [sensu @Schleuning2020-vq] corresponding to floral corolla depth, and phylogeny [@Pellissier2013-pk], which can be interpreted as a catch-all for unmeasured trait variation.
+Nevertheless, it is equally evident that sympatric bumble bee species do not *always* compete for floral resources to the point of competitive exclusion, since bumble bee communities with an apparent excess of species richness and deficit of trait dispersion are commonly observed [@Goulson2008-zy].
+
+To account for both the importance and variability of competitive processes in bumble bee communities, it is helpful to distinguish three processes that could obtain to varying degrees: no competition, interspecific competition, and intraspecific competition (**Table 1**).
+We will refer to these processes as hypotheses, since their realization in any given community is matter of inference.
+Below, we elaborate each and formulate corresponding empirical predictons.
+
++-------------------------------+----------------------------------------------------+
+| Hypothesis                    | Prediction                                         |
++===============================+====================================================+
+| H1: No competition            | 1. BB density !~ FL density                        |
+|                               | 2. BB trait dispersion = null expectation          |
+|                               | 3. BB-FL trait-matching > null expectation         |
++-------------------------------+----------------------------------------------------+
+| H2: Interspecific competition | 1. BB density ~ FL density                         |
+|                               | 2. BB trait dispersion > null expectation          |
+|                               | 3. BB-FL trait-matching > null expectation         |
++-------------------------------+----------------------------------------------------+
+| H3: Intraspecific competition | 1. BB density ~ FL density                         |
+|                               | 2. BB diet breadth ~ BB conspecific-packing        |
+|                               | 3. BB-FL trait-matching ~ 1/BB conspecific-packing |
++-------------------------------+----------------------------------------------------+
+
+**Table 1:** Hypothesized competitive processes with empirical predictions. BB = bumble bee; FL = floral. \
+
+*H1: No competition.*    The hypothesis that floral resource competition is not a significant process governing a bumble bee community functions as a "null hypothesis" when competition *per se* is the phenomenon in question, but it is not "null" in the sense of mere negation of mechanistic explanation.
+To the contrary, at least two plausible and ecologically meaningful mechanisms can be proposed that would account for a lack of competition in bumble bee communities [@Ranta1981-vx].
+First, it is possible that a super-abundance of floral resources could saturate the foraging capacity of a local bumble bee population.
+Indeed, this may be a common (though transient) phenomenon associated with the bloom of mass flowering crops [@Ranta1981-vx].
+Alternatively, competition could be attenuated if abiotic stress, such as low temperature, suppresses bumble bee populations below competition-inducing densities.
+This pattern would be a prediction of the "stress-gradient hypothesis" [@Bertness1994-zx], and while it has not, to our knowledge, been documented for bumble bees specifically, a recent study of whole bee communities (of which bumble bees were a part) found evidence of dampened competition at high elevation sites along a montane-to-alpine gradient [@Hoiss2012-sn].
+
+Under either mechanism, the most straightforward prediction is that bumble bee population density would be decoupled from resource density, since the latter would be non-limiting.
+For the same reason, the absence of competition would also predict that the trait dispersion of sympatric bumble bee species would not exceed that predicted by a random null model.
+Third, under a no-competition scenario, bumble bee species would suffer no displacement from their optimal floral resource species.
+Thus, holding other aspects of floral resource value constant (e.g.
+abundance, nectar volume/concentration), a lack of competition would predict that morphological trait-matching between bumble bee species and their chosen host flowers would exceed that predicted by a random null model.
+
+*H2: Interspecific competition.*    Resource partitioning by bumble bees is mediated by variation in matching traits, the most well-studied of these being tongue length.
+Bumble bee tongue lengths vary between species, and the efficiency with which a bumble bee can collect nectar from a given flower depends largely on the closeness of match between the tongue length of the bee and the corolla depth of the flower [@Inouye1980-mv; @Ranta1980-ok].
+Thus, under conditions of interspecific competition, a species' "defensible" niche space (i.e.
+the space in which it is a superior competitor) should center on the floral trait space that corresponds to its tongue length, and the number of species that can be "packed" into a given locality should be limited by some minimal differentiation of tongue length [@Inouye1977-kl; @Pyke1982-gn] within the range of variation supported by the trait space of available flora.
+
+Since resource limitation is a necessary condition for resource competition, a key prediction distiguishing interspecific competition from a lack of competition is that bumble bee density should vary in proportion to floral resource density.
+A bumble bee community under interspecific competition should also exhibit higher tongue length dispersion and stronger trait matching relative to the predictions of a random null model [@Inouye1977-kl].
+
+*H3: Intraspecific competition.*    As noted by @Ranta1981-vx, theory predicts that intraspecific competition can promote species coexistence through the relief of interspecific competition [@MacArthur1984-qz], and could thus, as an alternative to the "no competition" hypothesis, account for cases in which the observed diversity of sympatric bumble bee species exceeds that predicted by trait-based niche models.
+Under the pressure of intraspecific competition, individuals face the depletion of preferred resources and may respond by foraging more opportunistically within the constraints of their fundamental niche space [].
+Alternatively, a population may relieve intraspecific competition by subdividing, on the basis of intraspecific trait heterogeneity, into differently specialized "guilds" [@Araujo2008-bp].
+In either case, the population-level effect would be an increase in the species' realized niche breadth [though see @Parent2014-hx].
+
+Thus, the predictions of intraspecific competition can be formulated as follows.
+First, as in interspecific competition, resources must be limiting, so bumble bee density should vary in proportion to floral density.
+In the case of intraspecific competition, though, we can refine this prediction to say that the density of a given bumble bee species should vary in proportion to the density of the portion of the floral resource community belonging to its dietary niche, since the effects of intraspecific competition are realized on a per-species basis.
+Second, the dietary breadth of a given species should vary in proportion to the density of conspecifics per density of floral resources (we will refer to this as a "conspecific-packing" index).
+Conversely, a broadening of diet should entail a relaxation of trait-matching, so the strength of trait-matching should vary inversely with conspecific-packing.
+
+*Seasonal-Elevational metahypothesis.*    That any of the processes hypothesized above should obtain universally is both empirically and theoretically doubtful [@Ranta1981-vx].
+More plausibly, which of these processes prevails -- and which species are favored by the prevailing process -- vary in both space and time, and this heterogenity of process could explain the structural idiosyncrasies of observed bumble bee communities [@Ranta1981-vx].
+Rather than viewing this heterogeneity as an obstacle to inferring some universal rule of bumble bee community assembly, our aim in the present study is to investigate the heterogeneity itself as an ecological phenomenon, characterizing the variation in competitive processes (or lack thereof) in bumble bee communities through time and space.
+
+Our study is set in the Berchtesgadener Alps of southwestern Germany, where a steep elevation gradient permits the study of wide environmental variability while minimizing the confounding effects of geographical distance [e.g.
+@Hoiss2015-gv].
+In this context, we formulate a "metahypothesis" concerning the variation of hypothesized competition processes through elevation and time (**Figure 1**).
+Consistent with the stress-gradient hypothesis [@Bertness1994-zx] and with previous empirical work in our study system [@Hoiss2012-sn], we hypothesize an inverse relationship between competition and elevation.
+Along a temporal axis, we hypothesize that the strength of competition is structured by the seasonal life history pattern of bumble bee colonies [@Goulson2010-lo].
+During the solitary phase, when the bumble bee community consists only of foundress queens, competition should be low.
+Both inter- and intraspecific competition should intensify as colonies increase in size during the social phase.
+After the switch point, when colonies begin producing gynes and males instead of workers, intraspecific comeptition should decrease due to the strong morphological differences between queens, males, and workers within each species, but interspecific competition might intensify due to the effective "over-packing" of trait space wih the introduction of multiple morphological castes per species.
+
+In addition to clarifying the basic ecology of bumble bee community assembly, parsing competitive processes in this way has an applied significance for the conservation of bumble bees in the changing environments of the Anthropocene.
+Under the forcing of environmental change, such as climate warming or land use conversion, bumble bee communities might not be affected only in terms of range shifts, but also by a restructuring of competitive interactions, the nature and consequences of which may be difficult to predict.
+
+```{r echo = FALSE, figu}
+Time <- c(0,1,2,3)
+Elevation <- c(0,1,2,3)
+tab <- cbind(Time, Elevation) %>%
+  as_tibble()
+
+ggplot(tab, aes(Time, Elevation)) +
+  geom_vline(xintercept = c(1,2), color = "white") +
+  geom_hline(yintercept = c(1,2), color = "white") +
+  scale_x_continuous(expand = c(0,0),
+                     breaks = c(0.5, 1.5, 2.5), 
+                     labels = c("founding", "build-up", "reproductive"),
+                     limits = c(0,3)) +
+  scale_y_continuous(expand = c(0,0),
+                     breaks = c(0.5, 1.5, 2.5), 
+                     labels = c("low", "mid", "high"),
+                     limits = c(0,3)) +
+  #geom_text() +
+  annotate("text", x = 0.5, y = 2.5, label = "H1") +
+  annotate("text", x = 1.5, y = 2.5, label = "H1") +
+  annotate("text", x = 2.5, y = 2.5, label = "H1") +
+  annotate("text", x = 0.5, y = 1.5, label = "H1") +
+  annotate("text", x = 0.5, y = 0.5, label = "H1") +
+  annotate("text", x = 1.5, y = 1.5, label = "H2 | H3") +
+  annotate("text", x = 2.5, y = 1.5, label = "H2 < H3") +
+  annotate("text", x = 1.5, y = 0.5, label = "H2 | H3") +
+  annotate("text", x = 2.5, y = 0.5, label = "H2 < H3") +
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
+```
+
+**Figure 1:** Hypothesized variation in prevailing competitive processes through time and elevation. \
+
+**Methods**\
+
+*Field sampling*
+
+*Degree day modeling.*
+
+*Inferring reosurce limitation*
+
+**Results**\
+
+*Colony phase*
+```{r}
+### Get caste counts
+castes <- read_csv("./processed_data/network.csv") %>%
+  group_by(site, date, year, caste) %>%
+  summarize(caste.count = n()) %>%
+  spread(caste, caste.count) %>%
+  replace(is.na(.), 0) %>%
+  mutate(q.prop = queen/(queen+worker)) %>%
+  left_join(gdd, by = c("site", "date")) %>%
+  left_join(site_data, by = c("site")) %>%
+  arrange(gdd.cum)
+
+### Logistic regression
+glm.qw.2012.high <- glm(q.prop ~ gdd.cum,
+                        family = quasibinomial,
+                        data = filter(castes, year == "2012",
+                                      elev.class2 == "high"))
+
+glm.qw.2012.mid <- glm(q.prop ~ gdd.cum,
+                        family = quasibinomial,
+                        data = filter(castes, year == "2012",
+                                      elev.class2 == "mid"))
+
+glm.qw.2012.low <- glm(q.prop ~ gdd.cum,
+                        family = quasibinomial,
+                        data = filter(castes, year == "2012",
+                                      elev.class2 == "low"))
+
+### Extract gdd corresponding to desired proportion; based on https://stackoverflow.com/questions/32040504/regression-logistic-in-r-finding-x-value-predictor-for-a-particular-y-value
+#### Note that this only works for logistic regression!
+getX <- function(x, y) {
+  (log(y/(1-y)) - coef(x)[1])/coef(x)[2]
+}
+
+break1_2012_high <- getX(glm.qw.2012.high, 0.5) %>%
+  data.frame() %>%
+  select(break1 = 1) %>%
+  mutate(elev.class2 = "high",
+         year = 2012)
+
+break1_2012_mid <- getX(glm.qw.2012.mid, 0.5)  %>%
+  data.frame() %>%
+  select(break1 = 1) %>%
+  mutate(elev.class2 = "mid",
+         year = 2012)
+
+break1_2012_low <- getX(glm.qw.2012.low, 0.5)  %>%
+  data.frame() %>%
+  select(break1 = 1) %>%
+  mutate(elev.class2 = "low",
+         year = 2012)
+
+break1_2012 <- bind_rows(break1_2012_high, break1_2012_mid, break1_2012_low) %>%
+  mutate(elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))
+
+# Add break points to castes data frame
+castes_2012 <- castes %>%
+  filter(year == "2012") %>%
+  left_join(break1_2012, by = c("year", "elev.class2"))
+
+# Plot
+ggplot(castes_2012, aes(gdd.cum, q.prop)) +
+  geom_point() +
+  geom_smooth(method = "glm", method.args=list(family="quasibinomial")) +
+  geom_vline(aes(xintercept = break1), color = "red") +
+  facet_wrap(~elev.class2, ncol = 1)
+
+%>%
+  group_by(site, date, year) %>%
+  #mutate(caste_prop = caste_count/sum(caste_count)) %>%
+  spread(caste, caste_count) %>%
+  replace(is.na(.), 0) %>%
+  mutate(queen = oldqueen + queen,
+         reproductive = male + youngqueen) %>%
+  select(-oldqueen, -male, -youngqueen) %>%
+  group_by(site, year) %>%
+  mutate(worker.cum = cumsum(worker),
+         repro.cum = cumsum(reproductive)) %>%
+  left_join(gdd, by = c("site", "date")) %>%
+  left_join(site_data, by = c("site")) %>%
+  select(site, date, year, queen, worker, reproductive, gdd.cum, elev.class, elev.class2, elev.mean) %>%
+  mutate(q.w = queen/worker)
+
+  gather(caste, count, -c(site, date, year, gdd.cum, elev.class, elev.class2, elev.mean)) %>%
+  group_by(site, date) %>%
+  mutate(prop = count/sum(count))
+
+glm_queen_2012 <- glm(prop ~ gdd.cum, 
+                      family = quasibinomial(),
+                      data = filter(castes, year == "2012" & caste == "queen" & elev.class2 == "mid"))
+
+beta_queen_2012 <- betareg(prop ~ gdd.cum, 
+                      data = filter(castes, year == "2012" & caste == "queen" & elev.class2 == "mid"))
+
+p <- predict(glm_queen_2012, type = "response")
+
+founding_break <- castes %>%
+  spread(caste, count) %>%
+  replace(is.na(.), 0) %>%
+  select(site, date, year, gdd.cum, queen, worker, reproductive) %>% 
+  filter(worker > queen & queen > 0) %>%
+  group_by(site, year) %>%
+  filter(date == min(date)) %>%
+  select(site, year, queen, worker, reproductive, founding.break = gdd.cum)
+
+repro_break <- castes %>%
+  spread(caste, count) %>%
+  replace(is.na(.), 0) %>%
+  select(site, date, year, gdd.cum, queen, worker, reproductive) %>% 
+  filter(reproductive / (reproductive + worker) >= 0.1) %>%
+  group_by(site, year) %>%
+  filter(date == min(date)) %>%
+  select(site, year, queen, worker, reproductive, repro.break = gdd.cum)
+
+phases <- castes %>%
+  full_join(founding_break, by = c("site", "year")) %>%
+  full_join(repro_break, by = c("site", "year"))
+
+# ggplot(phase, aes(gdd.cum, prop, color = caste)) +
+#   geom_point(alpha = 0.25) +
+#   geom_smooth() +
+#   theme_gray(10) +
+#   facet_grid(elev.class2~year, scales = "free") +
+#   xlab("Cumulative growing degree days") +
+#   ylab("Proportional abundance")
+# 
+# ggsave("./output/gdd_phase.pdf", width = 6, height = 4)
+
+ggplot(filter(castes, year == "2012"), aes(gdd.cum, q.w)) +
+  geom_point() +
+  geom_smooth() +
+  #geom_vline(aes(xintercept = founding.break), linetype = "dotted") +
+  facet_wrap(~elev.class2)
+
+ggplot(filter(phases, year == "2011"), aes(gdd.cum, prop, color = caste)) +
+  geom_point() +
+  geom_smooth() +
+  #geom_vline(aes(xintercept = founding.break), linetype = "dotted") +
+  facet_wrap(~elev.class2, scales = "free")
+
+ggplot(filter(phases, year == "2010"), aes(gdd.cum, prop, color = caste)) +
+  geom_point() +
+  geom_smooth() +
+  #geom_vline(aes(xintercept = founding.break), linetype = "dotted") +
+  facet_wrap(~elev.class2, scales = "free")
+
+```
+
+*Resource limitation*.
+Per-transect bumble bee abundance is positively correlated with per-transect floral cover (log-transformed).
+To the extent, therefore, that our metrics of bumble bee and floral density reflect true site-scale abundance and not just co-aggregation, this pattern suggests that bumble bee
+density is resource limited, particularly at sites with very low floral density.
+Importantly, this approach does not distinguish between inter- and intraspecies (or even intra-colony) competition [@Ranta1981-vx].
+
+Note, moreover, that this relationship varies seasonally.
+For now, we will use Julian month as a stand-in for GDD, which we'll use once we have analyzed our weather station data.
+Early in the year, the relationship between BB density and floral density is weak; this makes sense because colonies are just getting established during the first couple of months, and BB density is more limited by the autocorrelation component of colony size (bigger colonies can raise more brood) than by floral resource availability.
+In mid- to late-season, however, BB density becomes strongly dependent on floral resource density.
+Interestingly, in September, the relationship is strongest and floral density is also lowest.
+
+```{r eval=FALSE, echo=FALSE}
+ggplot(density_month.agg, aes(log.mean.fl.dens, mean.bb.dens, color = year)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE) +
+  facet_wrap(~month) +
+  ylab("Mean BB density") +
+  xlab("log(Mean floral density)")
+```
+
+**Discussion**\
+
 <!-- *Study sites.*  -->
@@
 <!-- **Results** -->
-
-<!-- Per-transect bumble bee abundance is positively correlated with per-transect floral cover (log-transformed). To the extent, therefore, that our -->
-<!-- metrics of bumble bee and floral density reflect true site-scale abundance and not just co-aggregation, this pattern suggests that bumble bee -->
-<!-- density is resource limited, particularly at sites with very low floral density.
-Importantly, this approach does not distinguish between inter- and intraspecies (or even intra-colony) competition [@Ranta1981-vx].
--->
-
-<!-- Note, moreover, that this relationship varies seasonally.
-For now, we will use Julian month as a stand-in for GDD, which we'll use once we have analyzed our weather station data.
-Early in the year, the relationship between BB density and floral density is weak; this makes sense because colonies are just getting established during the first couple of months, and BB density is more limited by the autocorrelation component of colony size (bigger colonies can raise more brood) than by floral resource availability.
-In mid- to late-season, however, BB density becomes strongly dependent on floral resource density.
-Interestingly, in September, the relationship is strongest and floral density is also lowest.
--->
-```{r eval=FALSE, echo=FALSE}
-ggplot(density_month.agg, aes(log.mean.fl.dens, mean.bb.dens, color = year)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +
-  facet_wrap(~month) +
-  ylab("Mean BB density") +
-  xlab("log(Mean floral density)")
-```
 

### 2020-05-11 data_proc.R

@@ -25,13 +25,13 @@ fl_survey <- read_delim("./raw_data/data_plantcover_forDoug_2020.01.17.csv", del
 write_csv(fl_survey, "./processed_data/floral_survey.csv")
 
 ### Visitation data
-network <- read_delim("./raw_data/data_bumblebee_forDoug_2020.01.17.csv", delim = ";", 
+network <- read_delim("./raw_data/data_bumblebee_forDoug_2020.04.17.csv", delim = ";", 
                            locale = locale(decimal_mark = ",", grouping_mark = ".")) %>%
   separate(col = plant_sp_latin, into = c("genus", "species", "w", "x", "y", "z"), sep = " ") %>%
   unite(plant_sp_latin, genus, species, sep = " ") %>%
   dplyr::select(-c(w,x,y,z)) %>%
   separate(plant_sp_latin, c("genus", "species"), remove = FALSE) %>%
-  dplyr::select(year, dayofyear, site = site_name, trap.time = trapping_time, pollen, 
+  dplyr::select(year, dayofyear, site = site_name, trap.time = trapping_time, caste, pollen, 
                 bb.sp = bumb_species, bb.sp.lat = bumb_sp_latin, plant.sp.abb = forage_plant, 
                 plant.sp = plant_sp_latin) %>%
   mutate(temp.date = as_date(dayofyear),
@@ -51,15 +51,22 @@ network <- read_delim("./raw_data/data_bumblebee_forDoug_2020.01.17.csv", delim 
 
 write_csv(network, "./processed_data/network.csv")
 
-site_data <- read_tsv("./ancillary_from_Fabrice/SiteLocationFabriceUpdate2020.tsv",
+### Site data
+site_data <- read_tsv("./raw_data/SiteLocationFabriceUpdate2020.tsv",
                       locale = locale(decimal_mark = ",", grouping_mark = ".")) %>%
   dplyr::select(site = ID, elev.class = Altitude, management, temp.mean = mean.temp, elev.mean = Mean_altitude, 
                 elev.class = Altitude, transect = Transekt, slope.calc = Inklination_degree_calc, slope.est = Inklination_degree_estim,
                 elev.min = Min_altitude, elev.max  = Max_altitude, lat = Lat, lon = Long) %>%
[…]

### 2020-05-13 manuscript.Rmd
functionalized breakpoints

@@
 ---
-title: Bumble bee communities are structured by a spatiotemporal mosaic of competitive
-  processes
+title: 
 author:
@@
 bibliography: refs.bib
+---
 
----
+Conversation with Ingolf
+- Add opening paragraph setting context of clinate change and community ecology
+- Look for divergent population trajectories of species as evidence of competition with temporal lag on effects
+- Where does network analysis fit in?
+A legit question -- no need to jump right to fancy network metrics if simpler approaches work.
+But so far, none of my empirical predictions involve network metrics.
+- Ingolf had a REALLY nice title -- ask him to send it to me.
+-  Ingolf said that he does not think that competition from syrphids at high elevation would be significant, but competition from honey bees at low elevation could be
+- Switch Fabrice and Kalli's order
+- Cite Ingolf's recent papers showing the effect of pulses of mass-flowering crop resource on bumble bee health
+- He liked my colony phase figures
 
@@
 site_data <- read_csv("./processed_data/site_data.csv") %>%
-  #dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
+  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
   mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
@@
   replace(is.na(.), 0) %>%
-  mutate(q.prop = queen/(queen+worker)) %>%
+  mutate(queen = oldqueen + queen,
+         reproductive = male + youngqueen,
+         q.prop = queen/(queen+worker+reproductive),
+         r.prop = reproductive/(queen+worker+reproductive)) %>%
+  select(-oldqueen, -youngqueen, -male) %>%
   left_join(gdd, by = c("site", "date")) %>%
@@
 
-### Logistic regression
-glm.qw.2012.high <- glm(q.prop ~ gdd.cum,
+### Logistic regression for first break point (founding to buildup)
+breakpoints <- function(x, year) {
+
+  data <- filter(x, year == !!year)
+
+  # Logistic regression for first break point (founding to buildup)
+  glm1.low <- glm(q.prop ~ gdd.cum,
                   family = quasibinomial,
-                        data = filter(castes, year == "2012",
-                                      elev.class2 == "high"))
+                  data = filter(data, elev.class2 == "low"))
 
-glm.qw.2012.mid <- glm(q.prop ~ gdd.cum,
+  glm1.mid <- glm(q.prop ~ gdd.cum,
                   family = quasibinomial,
-                        data = filter(castes, year == "2012",
-                                      elev.class2 == "mid"))
+                  data = filter(data, elev.class2 == "mid"))
 
-glm.qw.2012.low <- glm(q.prop ~ gdd.cum,
+  glm1.high <- glm(q.prop ~ gdd.cum,
                    family = quasibinomial,
-                        data = filter(castes, year == "2012",
-                                      elev.class2 == "low"))
+                   data = filter(data, elev.class2 == "high"))
+
+  # Logistic regression for second break point (buildup to reproductive)
+  glm2.low <- glm(r.prop ~ gdd.cum,
+                  family = quasibinomial,
+                  data = filter(data, elev.class2 == "low"))
+
+  glm2.mid <- glm(r.prop ~ gdd.cum,
+                  family = quasibinomial,
+                  data = filter(data, elev.class2 == "mid"))
+
+  glm2.high <- glm(r.prop ~ gdd.cum,
+                   family = quasibinomial,
+                   data = filter(data, elev.class2 == "high"))
 
   ### Extract gdd corresponding to desired proportion; based on https://stackoverflow.com/questions/32040504/regression-logistic-in-r-finding-x-value-predictor-for-a-particular-y-value
-#### Note that this only works for logistic regression!
+
   getX <- function(x, y) {
@@
 
-break1_2012_high <- getX(glm.qw.2012.high, 0.5) %>%
+  break1.low <- getX(glm1.low, 0.5) %>%
     data.frame() %>%
-  select(break1 = 1) %>%
+    select(gdd = 1) %>%
+    mutate(elev.class2 = "low",
+           breakID = "break1")
+
+  break1.mid <- getX(glm1.mid, 0.5) %>%
+    data.frame() %>%
+    select(gdd = 1) %>%
+    mutate(elev.class2 = "mid",
+           breakID = "break1")
+
+  break1.high <- getX(glm1.high, 0.5) %>%
+    data.frame() %>%
+    select(gdd = 1) %>%
     mutate(elev.class2 = "high",
-         year = 2012)
+           breakID = "break1")
 
-break1_2012_mid <- getX(glm.qw.2012.mid, 0.5)  %>%
+  break2.low <- getX(glm2.low, 0.2) %>%
     data.frame() %>%
-  select(break1 = 1) %>%
+    select(gdd = 1) %>%
+    mutate(elev.class2 = "low",
+           breakID = "break2")
+
+  break2.mid <- getX(glm2.mid, 0.2) %>%
+    data.frame() %>%
+    select(gdd = 1) %>%
     mutate(elev.class2 = "mid",
-         year = 2012)
+           breakID = "break2")
 
-break1_2012_low <- getX(glm.qw.2012.low, 0.5)  %>%
+  break2.high <- getX(glm2.high, 0.2) %>%
     data.frame() %>%
-  select(break1 = 1) %>%
-  mutate(elev.class2 = "low",
-         year = 2012)
+    select(gdd = 1) %>%
+    mutate(elev.class2 = "high",
+           breakID = "break2")
 
-break1_2012 <- bind_rows(break1_2012_high, break1_2012_mid, break1_2012_low) %>%
-  mutate(elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))
+  # Join into data frame, convert elev.class2 to ordered factor, spread
+  breaks <- full_join(break1.low, break1.mid) %>%
+    full_join(break1.high) %>%
+    full_join(break2.low) %>%
+    full_join(break2.mid) %>%
+    full_join(break2.high) %>%
+    mutate(elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) %>%
+    spread(breakID, gdd)
 
-# Add break points to castes data frame
-castes_2012 <- castes %>%
-  filter(year == "2012") %>%
-  left_join(break1_2012, by = c("year", "elev.class2"))
+  #
+  out <- data %>%
+    full_join(breaks) %>%
+    gather(caste, count, -c(site, date, year, q.prop, r.prop,
+                          gdd.cum, elev.class, elev.class2, 
+                          management, transect, elev.mean, 
+                          lat, lon, break1, break2)) %>%
+  group_by(site, date) %>%
+  mutate(prop = count/sum(count))
+}
+
+breaks_2012 <- breakpoints(castes, 2012)
+breaks_2011 <- breakpoints(castes, 2011)
+breaks_2010 <- breakpoints(castes, 2010)
 
 # Plot
-ggplot(castes_2012, aes(gdd.cum, q.prop)) +
+binomial_smooth <- function(...) {
+  geom_smooth(method = "glm", method.args = list(family = "quasibinomial"), se = FALSE, ...)
+}
+
+facet_names <- c(`low` = "Low (641-1000 m)",
+                 `mid` = "Mid (1000-1499 m)",
+                 `high` = "High (1500-2032 m)")
+
+ggplot(breaks_2012, aes(gdd.cum, prop, color = caste)) +
   geom_point() +
-  geom_smooth(method = "glm", method.args=list(family="quasibinomial")) +
-  geom_vline(aes(xintercept = break1), color = "red") +
-  facet_wrap(~elev.class2, ncol = 1)
+  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
+  geom_vline(aes(xintercept = break1), linetype = "dashed") +
+  geom_vline(aes(xintercept = break2), linetype = "dashed") +
+  facet_wrap(~elev.class2, ncol = 1, labeller = as_labeller(facet_names)) +
+  xlab("Cumulative growing degree days") +
+  ylab("Proportional abundance")
 
-%>%
-  group_by(site, date, year) %>%
-  #mutate(caste_prop = caste_count/sum(caste_count)) %>%
-  spread(caste, caste_count) %>%
-  replace(is.na(.), 0) %>%
-  mutate(queen = oldqueen + queen,
-         reproductive = male + youngqueen) %>%
-  select(-oldqueen, -male, -youngqueen) %>%
-  group_by(site, year) %>%
-  mutate(worker.cum = cumsum(worker),
-         repro.cum = cumsum(reproductive)) %>%
-  left_join(gdd, by = c("site", "date")) %>%
-  left_join(site_data, by = c("site")) %>%
-  select(site, date, year, queen, worker, reproductive, gdd.cum, elev.class, elev.class2, elev.mean) %>%
-  mutate(q.w = queen/worker)
+ggsave("./output/phases2012.pdf")
 
-  gather(caste, count, -c(site, date, year, gdd.cum, elev.class, elev.class2, elev.mean)) %>%
-  group_by(site, date) %>%
-  mutate(prop = count/sum(count))
+ggplot(breaks_2011, aes(gdd.cum, prop, color = caste)) +
+  geom_point() +
+  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
+  geom_vline(aes(xintercept = break1), linetype = "dashed") +
+  geom_vline(aes(xintercept = break2), linetype = "dashed") +
+  facet_wrap(~elev.class2, ncol = 1, labeller = as_labeller(facet_names)) +
+  xlab("Cumulative growing degree days") +
+  ylab("Proportional abundance")
 
-glm_queen_2012 <- glm(prop ~ gdd.cum, 
-                      family = quasibinomial(),
-                      data = filter(castes, year == "2012" & caste == "queen" & elev.class2 == "mid"))
+ggsave("./output/phases2011.pdf")
 
-beta_queen_2012 <- betareg(prop ~ gdd.cum, 
-                      data = filter(castes, year == "2012" & caste == "queen" & elev.class2 == "mid"))
+ggplot(breaks_2010, aes(gdd.cum, prop, color = caste)) +
+  geom_point() +
+  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
+  geom_vline(aes(xintercept = break1), linetype = "dashed") +
+  geom_vline(aes(xintercept = break2), linetype = "dashed") +
+  facet_wrap(~elev.class2, ncol = 1, labeller = as_labeller(facet_names)) +
+  xlab("Cumulative growing degree days") +
+  ylab("Proportional abundance")
 
-p <- predict(glm_queen_2012, type = "response")
-
-founding_break <- castes %>%
-  spread(caste, count) %>%
-  replace(is.na(.), 0) %>%
-  select(site, date, year, gdd.cum, queen, worker, reproductive) %>% 
-  filter(worker > queen & queen > 0) %>%
-  group_by(site, year) %>%
-  filter(date == min(date)) %>%
-  select(site, year, queen, worker, reproductive, founding.break = gdd.cum)
-
-repro_break <- castes %>%
-  spread(caste, count) %>%
-  replace(is.na(.), 0) %>%
-  select(site, date, year, gdd.cum, queen, worker, reproductive) %>% 
-  filter(reproductive / (reproductive + worker) >= 0.1) %>%
-  group_by(site, year) %>%
-  filter(date == min(date)) %>%
-  select(site, year, queen, worker, reproductive, repro.break = gdd.cum)
-
-phases <- castes %>%
-  full_join(founding_break, by = c("site", "year")) %>%
-  full_join(repro_break, by = c("site", "year"))
-
-# ggplot(phase, aes(gdd.cum, prop, color = caste)) +
-#   geom_point(alpha = 0.25) +
-#   geom_smooth() +
-#   theme_gray(10) +
-#   facet_grid(elev.class2~year, scales = "free") +
-#   xlab("Cumulative growing degree days") +
-#   ylab("Proportional abundance")
-# 
-# ggsave("./output/gdd_phase.pdf", width = 6, height = 4)
-
-ggplot(filter(castes, year == "2012"), aes(gdd.cum, q.w)) +
-  geom_point() +
-  geom_smooth() +
-  #geom_vline(aes(xintercept = founding.break), linetype = "dotted") +
-  facet_wrap(~elev.class2)
-
-ggplot(filter(phases, year == "2011"), aes(gdd.cum, prop, color = caste)) +
-  geom_point() +
-  geom_smooth() +
-  #geom_vline(aes(xintercept = founding.break), linetype = "dotted") +
-  facet_wrap(~elev.class2, scales = "free")
-
-ggplot(filter(phases, year == "2010"), aes(gdd.cum, prop, color = caste)) +
-  geom_point() +
-  geom_smooth() +
-  #geom_vline(aes(xintercept = founding.break), linetype = "dotted") +
-  facet_wrap(~elev.class2, scales = "free")
-
+ggsave("./output/phases2010.pdf")
 ```
@@
 
+<!-- Bumble bee communities are structured by a spatiotemporal mosaic of competitive -->
+<!--   processes -->
+
 <!-- Under conditions of resource limitation, intraspecific competition may be a stronger constraint on bumble bee density (though not on diversity) than interspecific competition, since conspecific are limited in their ability to partition further their potential niche space.

### 2020-05-13 data_proc.R

@@ -91,3 +91,6 @@ climate <- read_delim("./raw_data/climate2.txt", delim = "\t") %>%
   mutate(gdd.cum = cumsum(gdd))
 
 write_csv(climate, "./processed_data/climate.csv")
+
+### TRY data
+try <- read_delim("./raw_data/TRY_13_05_2020.txt", delim = "\t")

### 2020-05-15 manuscript.Rmd
refocusing introo on the distinction between competition as a community assembly process and a behavior process

@@
 ---
-title: 
+title: Bumble bee competition revisited under global climate change
 author:
 - Douglas B. Sponsler
+- Fabrice Requier
 - Katharina Kallnik
-- Fabrice Requier
 - Alice Claßen
@@
 ---
-
-Conversation with Ingolf
-- Add opening paragraph setting context of clinate change and community ecology
-- Look for divergent population trajectories of species as evidence of competition with temporal lag on effects
-- Where does network analysis fit in?
-A legit question -- no need to jump right to fancy network metrics if simpler approaches work.
-But so far, none of my empirical predictions involve network metrics.
-- Ingolf had a REALLY nice title -- ask him to send it to me.
--  Ingolf said that he does not think that competition from syrphids at high elevation would be significant, but competition from honey bees at low elevation could be
-- Switch Fabrice and Kalli's order
-- Cite Ingolf's recent papers showing the effect of pulses of mass-flowering crop resource on bumble bee health
-- He liked my colony phase figures
 
@@
          bb.per.fl = mean.bb.dens/mean.fl.dens)
+
+###############################
+######### Define functions
+##############################
+
+# Find breakpoints on GDD scale
+breakpoints_gdd <- function(x, year, t1, t2) {
+
+  data <- filter(x, year == !!year)
+
+  # Logistic regression for first break point (founding to buildup)
+  glm1.low <- glm(q.prop ~ gdd.cum,
+                  family = quasibinomial,
+                  data = filter(data, elev.class2 == "low"))
+
+  glm1.mid <- glm(q.prop ~ gdd.cum,
+                  family = quasibinomial,
+                  data = filter(data, elev.class2 == "mid"))
+
+  glm1.high <- glm(q.prop ~ gdd.cum,
+                   family = quasibinomial,
+                   data = filter(data, elev.class2 == "high"))
+
+  # Logistic regression for second break point (buildup to reproductive)
+  glm2.low <- glm(r.prop ~ gdd.cum,
+                  family = quasibinomial,
+                  data = filter(data, elev.class2 == "low"))
+
+  glm2.mid <- glm(r.prop ~ gdd.cum,
+                  family = quasibinomial,
+                  data = filter(data, elev.class2 == "mid"))
+
+  glm2.high <- glm(r.prop ~ gdd.cum,
+                   family = quasibinomial,
+                   data = filter(data, elev.class2 == "high"))
+
+  ### Extract gdd corresponding to desired proportion; based on https://stackoverflow.com/questions/32040504/regression-logistic-in-r-finding-x-value-predictor-for-a-particular-y-value
+
+  getX <- function(x, y) {
+    (log(y/(1-y)) - coef(x)[1])/coef(x)[2]
+    }
+
+  break1.low <- getX(glm1.low, t1) %>%
+    data.frame() %>%
+    select(gdd = 1) %>%
+    mutate(elev.class2 = "low",
+           breakID = "break1")
+
+  break1.mid <- getX(glm1.mid, t1) %>%
+    data.frame() %>%
+    select(gdd = 1) %>%
+    mutate(elev.class2 = "mid",
+           breakID = "break1")
+
+  break1.high <- getX(glm1.high, t1) %>%
+    data.frame() %>%
+    select(gdd = 1) %>%
+    mutate(elev.class2 = "high",
+           breakID = "break1")
+
+  break2.low <- getX(glm2.low, t2) %>%
+    data.frame() %>%
+    select(gdd = 1) %>%
+    mutate(elev.class2 = "low",
+           breakID = "break2")
+
+  break2.mid <- getX(glm2.mid, t2) %>%
+    data.frame() %>%
+    select(gdd = 1) %>%
+    mutate(elev.class2 = "mid",
+           breakID = "break2")
+
+  break2.high <- getX(glm2.high, t2) %>%
+    data.frame() %>%
+    select(gdd = 1) %>%
+    mutate(elev.class2 = "high",
+           breakID = "break2")
+
+  # Join into data frame, convert elev.class2 to ordered factor, spread
+  breaks <- full_join(break1.low, break1.mid) %>%
+    full_join(break1.high) %>%
+    full_join(break2.low) %>%
+    full_join(break2.mid) %>%
+    full_join(break2.high) %>%
+    mutate(elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) %>%
+    spread(breakID, gdd)
+
+  #
+  out <- data %>%
+    full_join(breaks) %>%
+    gather(caste, count, -c(site, date, year, q.prop, r.prop,
+                          gdd.cum, yday, elev.class, elev.class2, 
+                          management, transect, elev.mean, 
+                          lat, lon, break1, break2)) %>%
+  group_by(site, date) %>%
+  mutate(prop = count/sum(count))
+}
+
+# Find breakpoints on Julian scale
+breakpoints_yday <- function(x, year, t1, t2) {
+
+  data <- filter(x, year == !!year)
+
+  # Logistic regression for first break point (founding to buildup)
+  glm1.low <- glm(q.prop ~ yday,
+                  family = quasibinomial,
+                  data = filter(data, elev.class2 == "low"))
+
+  glm1.mid <- glm(q.prop ~ yday,
+                  family = quasibinomial,
+                  data = filter(data, elev.class2 == "mid"))
+
+  glm1.high <- glm(q.prop ~ yday,
+                   family = quasibinomial,
+                   data = filter(data, elev.class2 == "high"))
+
+  # Logistic regression for second break point (buildup to reproductive)
+  glm2.low <- glm(r.prop ~ yday,
+                  family = quasibinomial,
+                  data = filter(data, elev.class2 == "low"))
+
+  glm2.mid <- glm(r.prop ~ yday,
+                  family = quasibinomial,
+                  data = filter(data, elev.class2 == "mid"))
+
+  glm2.high <- glm(r.prop ~ yday,
+                   family = quasibinomial,
+                   data = filter(data, elev.class2 == "high"))
+
+  ### Extract gdd corresponding to desired proportion; based on https://stackoverflow.com/questions/32040504/regression-logistic-in-r-finding-x-value-predictor-for-a-particular-y-value
+
+  getX <- function(x, y) {
+    (log(y/(1-y)) - coef(x)[1])/coef(x)[2]
+    }
+
+  break1.low <- getX(glm1.low, t1) %>%
+    data.frame() %>%
+    select(day = 1) %>%
+    mutate(elev.class2 = "low",
+           breakID = "break1")
+
+  break1.mid <- getX(glm1.mid, t1) %>%
+    data.frame() %>%
+    select(day = 1) %>%
+    mutate(elev.class2 = "mid",
+           breakID = "break1")
+
+  break1.high <- getX(glm1.high, t1) %>%
+    data.frame() %>%
+    select(day = 1) %>%
+    mutate(elev.class2 = "high",
+           breakID = "break1")
+
+  break2.low <- getX(glm2.low, t2) %>%
+    data.frame() %>%
+    select(day = 1) %>%
+    mutate(elev.class2 = "low",
+           breakID = "break2")
+
+  break2.mid <- getX(glm2.mid, t2) %>%
+    data.frame() %>%
+    select(day = 1) %>%
+    mutate(elev.class2 = "mid",
+           breakID = "break2")
+
+  break2.high <- getX(glm2.high, t2) %>%
+    data.frame() %>%
+    select(day = 1) %>%
+    mutate(elev.class2 = "high",
+           breakID = "break2")
+
+  # Join into data frame, convert elev.class2 to ordered factor, spread
+  breaks <- full_join(break1.low, break1.mid) %>%
+    full_join(break1.high) %>%
+    full_join(break2.low) %>%
+    full_join(break2.mid) %>%
+    full_join(break2.high) %>%
+    mutate(elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) %>%
+    spread(breakID, day)
+
+  #
+  out <- data %>%
+    full_join(breaks) %>%
+    gather(caste, count, -c(site, date, year, q.prop, r.prop,
+                          gdd.cum, yday, elev.class, elev.class2, 
+                          management, transect, elev.mean, 
+                          lat, lon, break1, break2)) %>%
+  group_by(site, date) %>%
+  mutate(prop = count/sum(count))
+}
+
+# Find Juian breakpoints across all elevations
+breakpoints_yday_all <- function(x, year, t1, t2) {
+
+  data <- filter(x, year == !!year)
+
+  # Logistic regression for first break point (founding to buildup)
+  glm1 <- glm(q.prop ~ yday,
+                  family = quasibinomial,
+                  data = data)
+
+  # Logistic regression for second break point (buildup to reproductive)
+  glm2 <- glm(r.prop ~ yday,
+                  family = quasibinomial,
+                  data = data)
+
+  ### Extract day corresponding to desired proportion; based on https://stackoverflow.com/questions/32040504/regression-logistic-in-r-finding-x-value-predictor-for-a-particular-y-value
+
+  getX <- function(x, y) {
+    (log(y/(1-y)) - coef(x)[1])/coef(x)[2]
+    }
+
+  break1 <- getX(glm1, t1) %>%
+    data.frame() %>%
+    select(day = 1) %>%
+    mutate(elev.class2 = "low",
+           breakID = "break1")
+
+  break2 <- getX(glm2, t2) %>%
+    data.frame() %>%
+    select(day = 1) %>%
+    mutate(elev.class2 = "low",
+           breakID = "break2")
+
+  # Join into data frame, convert elev.class2 to ordered factor, spread
+  breaks <- full_join(break1, break2) %>%
+    spread(breakID, day)
+
+  #
+  out <- data %>%
+    full_join(breaks) %>%
+    gather(caste, count, -c(site, date, year, q.prop, r.prop,
+                          gdd.cum, yday, elev.class, elev.class2, 
+                          management, transect, elev.mean, 
+                          lat, lon, break1, break2)) %>%
+  group_by(site, date) %>%
+  mutate(prop = count/sum(count))
+}
+
+# Binomial ggplot smoother
+binomial_smooth <- function(...) {
+  geom_smooth(method = "glm", method.args = list(family = "quasibinomial"), se = FALSE, ...)
+}
 ```
@@
 
+The nascent reality of climate change has already revealed the fragile contingency of ecological paradigms that rely on the implicit assumption of climatic stability [@Hannah2002-nv; @Walther2002-mq].
+Thus, one of the central questions facing ecologists today is how climate change will affect the future fitness and distribution of species and, thereby, the composition and function of ecological communities.
+While it is intuitive that climate change should affect species and communities through environmental filtering, it is now recognized that climate change can also restructure biotic interactions [@Tylianakis2008-fi], such as competition and facilitation [@Ettinger2017-pk], altering the fitness landscape that governs community assembly and species coexistence [@Chesson2000-aw].
+
+Bumble bees (Hymenoptera: *Bombus* spp.), due to their well-studied natural history and amenability to field observation, have become a model system for the study of biotic interactions in ecological communities.
+In particular, the broad and overlapping distributions of bumble bee species, together with their gross similarity in morphology, behavior, and dietary niche, have attracted the attention of researchers interested in questions of interspecific competition and, more broadly, the ecological constraints of species coexistence [@Plowright1984-lc; @Goulson2008-zy].
+In this regard, however, bumble bees have proven somewhat enigmatic [@Ranta1981-vx].
+On the one hand, there is evidence that resource limitation, a necessary precondition for competition, can be a significant constraint on local bumble populations [@Heinrich1976-wc], even to the point of widespread colony failure due to starvation [@Bowers1985-df].
+Moreover, the species of certain bumble bee communties have been shown to be over-dispersed in terms of tongue length (a matching trait corresponding to floral corolla depth) or phyologeny (understood to encompass unmeasured trait variation), providing evidence for competition as a community assembly process [e.g.
+@Inouye1977-kl; @Pellissier2013-pk].
+Nevertheless, it is equally evident that sympatric bumble bee species do not *always* compete for floral resources to the point of competitive exclusion, since bumble bee communities with an apparent excess of species richness and deficit of trait (or phylogenetic) dispersion are commonly observed [@Goulson2008-zy].
+
+To account for this inconsistency, @Ranta1981-vx proposed that bumble bees communities are governed by a spatiotemporally heterogeneous "mosaic" of competitive conditions, arising from the interaction of bumble bee species' traits and nest locations with spatiotemporal variation in floral resource abundance and composition, such that competitive inequalities between species are sufficiently variable (and fine-grained relative to the scale of observation) to allow more species to coexist than predicted by trait-based competition models.
+A complementary explanation, based on the core/satellite species hypothesis [@Hanski1982-hs], is that competitive forces are significant only among regionally abundant and widespread "core species", and the effects of competition are more pronounced in the structure of regional species pools than in that of local assemblages [@Hanski1982-xd].
+
+As pointed out by @Williams1989-vr, however, the question of competition as a process of community assembly, relevant to evolutionary and biogeographic scales of inquiry, differs from that of competition as a behavioral phenomenon within a given community, irrespective of the forces responsible for its assembly.
+
+"This is not to deny that interspecific competition affects patterns of foraging by bumble bees, which has been demonstrated by the effect of competitor removal  (e.g.
+Inouye, 1978; Plowright, Pendrel & McLaren, 1978; Plowright & Rodd, 1980; Pleasants,  1981).
+It is  concluded only that interspecific competition does not appear to govern the composition of local  assemblages of bumble bee species in the manner predicted by  the  theory of species-packing." [@Williams1989-vr]
+
+Bring up Hanksi's core/satellite hypothesis.
+
+Steep elevation gradients can serve as space-for-time proxies in the study of climate effects, capturing a broad range of climate variation  while minimizing the confounding effects of geographic distance [e.g.
+@Hoiss2015-gv].
+
+One of the most well-studied hypotheses linking climate and biotic interactions is the "stress gradient hypothesis", which posits a that the relative significance of competition and faciitation vary predictably along gradients of abiotic (e.g.
+climatic) stress, with low stress environments characterized by competition and high stress environments by facilitation [@Bertness1994-zx].
+
+When phylogenetic/trait dispersion are used as indices of competition, the relevant question is whether competition played a role in community assembly.
+These indices do not answer the question of whether, right now, competition is happening.
+Trait-matching, niche overlap, specialization (e.g.
+d') based indices are better suited for the question of immediate competiion.
+It could be that competition was not important over the time scales relevant to community assembly, but now it has become a significant phenomenon due to some change in the system.
+Or, it is possible that competition is spatiotemporally transient, and its net effect is not sufficient to cause a trait or phylogenetic signal.
+
 Bumble bees (Hymenoptera: *Bombus* spp.) are classic model organisms for the study of resource competition in biological communities.
 Nevertheless, the role of resource competition in structuring bumble bee communities has proven difficult to generalize.
-It is evident that sympatric bumble bees *can* compete for floral resources, and strongly so.
-@Heinrich1976-wc found that a bumble bee community can deplete the daily standing nectar pool of a given locality by more than 90%, and starvation has been implicated in the demise of colonies in isolated subalpine meadows [@Bowers1985-df], leaving little room to doubt the potential for resource limitation.
+Resource limitation, a precondition for competition,
+
+@Heinrich1976-wc found that a bumble bee community can deplete the daily standing nectar pool of a given locality by more than 90%, and starvation has been implicated in the demise of colonies in isolated subalpine meadows [@Bowers1985-df], in.
 Other studies have found local bumble bee communities to be over-dispersed in terms of tongue length [e.g.
@@
 First, it is possible that a super-abundance of floral resources could saturate the foraging capacity of a local bumble bee population.
-Indeed, this may be a common (though transient) phenomenon associated with the bloom of mass flowering crops [@Ranta1981-vx].
+Indeed, this may be a common (though transient) phenomenon associated with the bloom of mass flowering crops, such as red clover [@Ranta1981-vx] and oilseed rape [@Westphal2009-sb].
 Alternatively, competition could be attenuated if abiotic stress, such as low temperature, suppresses bumble bee populations below competition-inducing densities.
@@
 
+<!-- - Look for divergent population trajectories of species as evidence of competition with temporal lag on effects -->
+<!-- - Where does network analysis fit in?
+A legit question -- no need to jump right to fancy network metrics if simpler approaches work.
+But so far, none of my empirical predictions involve network metrics.
+-->
+
 *Field sampling*
@@
   left_join(site_data, by = c("site")) %>%
-  arrange(gdd.cum)
+  arrange(gdd.cum) %>%
+  mutate(yday = yday(date))
 
-### Logistic regression for first break point (founding to buildup)
-breakpoints <- function(x, year) {
+breaks_2012_gdd <- breakpoints_gdd(castes, 2012, t1 = 0.5, t2 = 0.1)
+breaks_2011_gdd <- breakpoints_gdd(castes, 2011, t1 = 0.5, t2 = 0.1)
+breaks_2010_gdd <- breakpoints_gdd(castes, 2010, t1 = 0.5, t2 = 0.1)
 
-  data <- filter(x, year == !!year)
+breaks_2012_yday <- breakpoints_yday(castes, 2012, t1 = 0.5, t2 = 0.1)
+breaks_2011_yday <- breakpoints_yday(castes, 2011, t1 = 0.5, t2 = 0.1)
+breaks_2010_yday <- breakpoints_yday(castes, 2010, t1 = 0.5, t2 = 0.1)
 
-  # Logistic regression for first break point (founding to buildup)
-  glm1.low <- glm(q.prop ~ gdd.cum,
-                  family = quasibinomial,
-                  data = filter(data, elev.class2 == "low"))
-
-  glm1.mid <- glm(q.prop ~ gdd.cum,
-                  family = quasibinomial,
-                  data = filter(data, elev.class2 == "mid"))
-
-  glm1.high <- glm(q.prop ~ gdd.cum,
-                   family = quasibinomial,
-                   data = filter(data, elev.class2 == "high"))
-
-  # Logistic regression for second break point (buildup to reproductive)
-  glm2.low <- glm(r.prop ~ gdd.cum,
-                  family = quasibinomial,
-                  data = filter(data, elev.class2 == "low"))
-
-  glm2.mid <- glm(r.prop ~ gdd.cum,
-                  family = quasibinomial,
-                  data = filter(data, elev.class2 == "mid"))
-
-  glm2.high <- glm(r.prop ~ gdd.cum,
-                   family = quasibinomial,
-                   data = filter(data, elev.class2 == "high"))
-
-  ### Extract gdd corresponding to desired proportion; based on https://stackoverflow.com/questions/32040504/regression-logistic-in-r-finding-x-value-predictor-for-a-particular-y-value
-
-  getX <- function(x, y) {
-    (log(y/(1-y)) - coef(x)[1])/coef(x)[2]
-    }
-
-  break1.low <- getX(glm1.low, 0.5) %>%
-    data.frame() %>%
-    select(gdd = 1) %>%
-    mutate(elev.class2 = "low",
-           breakID = "break1")
-
-  break1.mid <- getX(glm1.mid, 0.5) %>%
-    data.frame() %>%
-    select(gdd = 1) %>%
-    mutate(elev.class2 = "mid",
-           breakID = "break1")
-
-  break1.high <- getX(glm1.high, 0.5) %>%
-    data.frame() %>%
-    select(gdd = 1) %>%
-    mutate(elev.class2 = "high",
-           breakID = "break1")
-
-  break2.low <- getX(glm2.low, 0.2) %>%
-    data.frame() %>%
-    select(gdd = 1) %>%
-    mutate(elev.class2 = "low",
-           breakID = "break2")
-
-  break2.mid <- getX(glm2.mid, 0.2) %>%
-    data.frame() %>%
-    select(gdd = 1) %>%
-    mutate(elev.class2 = "mid",
-           breakID = "break2")
-
-  break2.high <- getX(glm2.high, 0.2) %>%
-    data.frame() %>%
-    select(gdd = 1) %>%
-    mutate(elev.class2 = "high",
-           breakID = "break2")
-
-  # Join into data frame, convert elev.class2 to ordered factor, spread
-  breaks <- full_join(break1.low, break1.mid) %>%
-    full_join(break1.high) %>%
-    full_join(break2.low) %>%
-    full_join(break2.mid) %>%
-    full_join(break2.high) %>%
-    mutate(elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) %>%
-    spread(breakID, gdd)
-
-  #
-  out <- data %>%
-    full_join(breaks) %>%
-    gather(caste, count, -c(site, date, year, q.prop, r.prop,
-                          gdd.cum, elev.class, elev.class2, 
-                          management, transect, elev.mean, 
-                          lat, lon, break1, break2)) %>%
-  group_by(site, date) %>%
-  mutate(prop = count/sum(count))
-}
-
-breaks_2012 <- breakpoints(castes, 2012)
-breaks_2011 <- breakpoints(castes, 2011)
-breaks_2010 <- breakpoints(castes, 2010)
+breaks_2012_yday_all <- breakpoints_yday_all(castes, 2012, t1 = 0.5, t2 = 0.1)
+breaks_2011_yday_all <- breakpoints_yday_all(castes, 2011, t1 = 0.5, t2 = 0.1)
+breaks_2010_yday_all <- breakpoints_yday_all(castes, 2010, t1 = 0.5, t2 = 0.1)
 
 # Plot
-binomial_smooth <- function(...) {
-  geom_smooth(method = "glm", method.args = list(family = "quasibinomial"), se = FALSE, ...)
-}
-
 facet_names <- c(`low` = "Low (641-1000 m)",
@@
 
-ggplot(breaks_2012, aes(gdd.cum, prop, color = caste)) +
+ggplot(breaks_2012_gdd, aes(gdd.cum, prop, color = caste)) +
   geom_point() +
@@
 
-ggsave("./output/phases2012.pdf")
+ggsave("./output/phases2012_gdd.pdf")
 
-ggplot(breaks_2011, aes(gdd.cum, prop, color = caste)) +
+ggplot(breaks_2012_yday, aes(yday, prop, color = caste)) +
+  geom_point() +
+  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
+  geom_vline(aes(xintercept = break1), linetype = "dashed") +
+  geom_vline(aes(xintercept = break2), linetype = "dashed") +
+  facet_wrap(~elev.class2, ncol = 1, labeller = as_labeller(facet_names)) +
+  xlab("Julian day") +
+  ylab("Proportional abundance")
+
+ggsave("./output/phases2012_yday.pdf")
+
+ggplot(breaks_2011_gdd, aes(gdd.cum, prop, color = caste)) +
   geom_point() +
@@
 
-ggsave("./output/phases2011.pdf")
+ggsave("./output/phases2011_gdd.pdf")
 
-ggplot(breaks_2010, aes(gdd.cum, prop, color = caste)) +
+ggplot(breaks_2011_yday, aes(yday, prop, color = caste)) +
   geom_point() +
@@
 
-ggsave("./output/phases2010.pdf")
+ggsave("./output/phases2011_yday.pdf")
+
+ggplot(breaks_2010_gdd, aes(gdd.cum, prop, color = caste)) +
+  geom_point() +
+  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
+  geom_vline(aes(xintercept = break1), linetype = "dashed") +
+  geom_vline(aes(xintercept = break2), linetype = "dashed") +
+  facet_wrap(~elev.class2, ncol = 1, labeller = as_labeller(facet_names)) +
+  xlab("Cumulative growing degree days") +
+  ylab("Proportional abundance")
+
+ggsave("./output/phases2010_gdd.pdf")
+
+ggplot(breaks_2010_yday, aes(yday, prop, color = caste)) +
+  geom_point() +
+  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
+  geom_vline(aes(xintercept = break1), linetype = "dashed") +
+  geom_vline(aes(xintercept = break2), linetype = "dashed") +
+  facet_wrap(~elev.class2, ncol = 1, labeller = as_labeller(facet_names)) +
+  xlab("Cumulative growing degree days") +
+  ylab("Proportional abundance")
+
+ggsave("./output/phases2010_yday.pdf")
+
+ggplot(breaks_2012_yday_all, aes(yday, prop, color = caste)) +
+  geom_point() +
+  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
+  geom_vline(aes(xintercept = break1), linetype = "dashed") +
+  geom_vline(aes(xintercept = break2), linetype = "dashed") +
+  xlab("Cumulative growing degree days") +
+  ylab("Proportional abundance")
+
+ggsave("./output/phases2012_yday_all.pdf")
+
+ggplot(breaks_2011_yday_all, aes(yday, prop, color = caste)) +
+  geom_point() +
+  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
+  geom_vline(aes(xintercept = break1), linetype = "dashed") +
+  geom_vline(aes(xintercept = break2), linetype = "dashed") +
+  xlab("Cumulative growing degree days") +
+  ylab("Proportional abundance")
+
+ggsave("./output/phases2011_yday_all.pdf")
+
+ggplot(breaks_2010_yday_all, aes(yday, prop, color = caste)) +
+  geom_point() +
+  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
+  geom_vline(aes(xintercept = break1), linetype = "dashed") +
+  geom_vline(aes(xintercept = break2), linetype = "dashed") +
+  xlab("Cumulative growing degree days") +
+  ylab("Proportional abundance")
+
+ggsave("./output/phases2010_yday_all.pdf")
 ```
@@
 
+<!-- -  Ingolf said that he does not think that competition from syrphids at high elevation would be significant, but competition from honey bees at low elevation could be -->
+
 <!-- Bumble bee communities are structured by a spatiotemporal mosaic of competitive -->
 <!--   processes -->
+
+<!-- The two processes of environmental filtering and biotic interactions may be difficult to disentangle [@Kraft2015-gh].  -->
+
+<!-- The nascent reality of climate change, and the grave expectation of its progression in the coming decades, have revealed the fragile contingency of conservation paradigms that rely on the implicit assumption of climatic stability [@Hannah2002-nv].
+-->
 

### 2020-05-23 analysis.Rmd
oh yeah

---
title: "analysis"
output: html_document
---

### Prepare environment
```{r setup, include=FALSE}
library(bipartite)
library(lme4)
library(lmerTest)
library(tidyverse)
library(lubridate)
library(GGally)
```

### Define functions
```{r}
# Find breakpoints on GDD scale
breakpoints_gdd <- function(x, year, t1, t2) {

  data <- filter(x, year == !!year)

  # Logistic regression for first break point (founding to buildup)
  glm1.low <- glm(q.prop ~ gdd.cum,
                  family = quasibinomial,
                  data = filter(data, elev.class2 == "low"))

  glm1.mid <- glm(q.prop ~ gdd.cum,
                  family = quasibinomial,
                  data = filter(data, elev.class2 == "mid"))

  glm1.high <- glm(q.prop ~ gdd.cum,
                   family = quasibinomial,
                   data = filter(data, elev.class2 == "high"))

  # Logistic regression for second break point (buildup to reproductive)
  glm2.low <- glm(r.prop ~ gdd.cum,
                  family = quasibinomial,
                  data = filter(data, elev.class2 == "low"))

  glm2.mid <- glm(r.prop ~ gdd.cum,
                  family = quasibinomial,
                  data = filter(data, elev.class2 == "mid"))

  glm2.high <- glm(r.prop ~ gdd.cum,
                   family = quasibinomial,
                   data = filter(data, elev.class2 == "high"))

  ### Extract gdd corresponding to desired proportion; based on https://stackoverflow.com/questions/32040504/regression-logistic-in-r-finding-x-value-predictor-for-a-particular-y-value

  getX <- function(x, y) {
    (log(y/(1-y)) - coef(x)[1])/coef(x)[2]
    }

  break1.low <- getX(glm1.low, t1) %>%
    data.frame() %>%
    select(gdd = 1) %>%
    mutate(elev.class2 = "low",
           breakID = "break1")

  break1.mid <- getX(glm1.mid, t1) %>%
    data.frame() %>%
    select(gdd = 1) %>%
    mutate(elev.class2 = "mid",
           breakID = "break1")

  break1.high <- getX(glm1.high, t1) %>%
    data.frame() %>%
    select(gdd = 1) %>%
    mutate(elev.class2 = "high",
           breakID = "break1")

  break2.low <- getX(glm2.low, t2) %>%
    data.frame() %>%
    select(gdd = 1) %>%
    mutate(elev.class2 = "low",
           breakID = "break2")

  break2.mid <- getX(glm2.mid, t2) %>%
    data.frame() %>%
    select(gdd = 1) %>%
    mutate(elev.class2 = "mid",
           breakID = "break2")

  break2.high <- getX(glm2.high, t2) %>%
    data.frame() %>%
    select(gdd = 1) %>%
    mutate(elev.class2 = "high",
           breakID = "break2")

  # Join into data frame, convert elev.class2 to ordered factor, spread
  breaks <- full_join(break1.low, break1.mid) %>%
    full_join(break1.high) %>%
    full_join(break2.low) %>%
    full_join(break2.mid) %>%
    full_join(break2.high) %>%
    mutate(elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) %>%
    spread(breakID, gdd)

  #
  out <- data %>%
    full_join(breaks) %>%
    gather(caste, count, -c(site, date, year, q.prop, r.prop,
                          gdd.cum, yday, elev.class, elev.class2, 
                          management, transect, elev.mean, 
                          lat, lon, break1, break2)) %>%
  group_by(site, date) %>%
  mutate(prop = count/sum(count))
}

# Find breakpoints on Julian scale
breakpoints_yday <- function(x, year, t1, t2) {

  data <- filter(x, year == !!year)

  # Logistic regression for first break point (founding to buildup)
  glm1.low <- glm(q.prop ~ yday,
                  family = quasibinomial,
                  data = filter(data, elev.class2 == "low"))

  glm1.mid <- glm(q.prop ~ yday,
                  family = quasibinomial,
                  data = filter(data, elev.class2 == "mid"))

  glm1.high <- glm(q.prop ~ yday,
                   family = quasibinomial,
                   data = filter(data, elev.class2 == "high"))

  # Logistic regression for second break point (buildup to reproductive)
  glm2.low <- glm(r.prop ~ yday,
                  family = quasibinomial,
                  data = filter(data, elev.class2 == "low"))

  glm2.mid <- glm(r.prop ~ yday,
                  family = quasibinomial,
                  data = filter(data, elev.class2 == "mid"))

  glm2.high <- glm(r.prop ~ yday,
                   family = quasibinomial,
                   data = filter(data, elev.class2 == "high"))

  ### Extract gdd corresponding to desired proportion; based on https://stackoverflow.com/questions/32040504/regression-logistic-in-r-finding-x-value-predictor-for-a-particular-y-value

  getX <- function(x, y) {
    (log(y/(1-y)) - coef(x)[1])/coef(x)[2]
    }

  break1.low <- getX(glm1.low, t1) %>%
    data.frame() %>%
    select(day = 1) %>%
    mutate(elev.class2 = "low",
           breakID = "break1")

  break1.mid <- getX(glm1.mid, t1) %>%
    data.frame() %>%
    select(day = 1) %>%
    mutate(elev.class2 = "mid",
           breakID = "break1")

  break1.high <- getX(glm1.high, t1) %>%
    data.frame() %>%
    select(day = 1) %>%
    mutate(elev.class2 = "high",
           breakID = "break1")

  break2.low <- getX(glm2.low, t2) %>%
    data.frame() %>%
    select(day = 1) %>%
    mutate(elev.class2 = "low",
           breakID = "break2")

  break2.mid <- getX(glm2.mid, t2) %>%
    data.frame() %>%
    select(day = 1) %>%
    mutate(elev.class2 = "mid",
           breakID = "break2")

  break2.high <- getX(glm2.high, t2) %>%
    data.frame() %>%
    select(day = 1) %>%
    mutate(elev.class2 = "high",
           breakID = "break2")

  # Join into data frame, convert elev.class2 to ordered factor, spread
  breaks <- full_join(break1.low, break1.mid) %>%
    full_join(break1.high) %>%
    full_join(break2.low) %>%
    full_join(break2.mid) %>%
    full_join(break2.high) %>%
    mutate(elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) %>%
    spread(breakID, day)

  #
  out <- data %>%
    full_join(breaks) %>%
    gather(caste, count, -c(site, date, year, q.prop, r.prop,
                          gdd.cum, yday, elev.class, elev.class2, 
                          management, transect, elev.mean, 
                          lat, lon, break1, break2)) %>%
  group_by(site, date) %>%
  mutate(prop = count/sum(count))
}

# Find Juian breakpoints across all elevations
breakpoints_yday_all <- function(x, year, t1, t2) {

  data <- filter(x, year == !!year)

  # Logistic regression for first break point (founding to buildup)
  glm1 <- glm(q.prop ~ yday,
                  family = quasibinomial,
                  data = data)

  # Logistic regression for second break point (buildup to reproductive)
  glm2 <- glm(r.prop ~ yday,
                  family = quasibinomial,
                  data = data)

  ### Extract day corresponding to desired proportion; based on https://stackoverflow.com/questions/32040504/regression-logistic-in-r-finding-x-value-predictor-for-a-particular-y-value

  getX <- function(x, y) {
    (log(y/(1-y)) - coef(x)[1])/coef(x)[2]
    }

  break1 <- getX(glm1, t1) %>%
    data.frame() %>%
    select(day = 1) %>%
    mutate(elev.class2 = "low",
           breakID = "break1")

  break2 <- getX(glm2, t2) %>%
    data.frame() %>%
    select(day = 1) %>%
    mutate(elev.class2 = "low",
           breakID = "break2")

  # Join into data frame, convert elev.class2 to ordered factor, spread
  breaks <- full_join(break1, break2) %>%
    spread(breakID, day)

  #
  out <- data %>%
    full_join(breaks) %>%
    gather(caste, count, -c(site, date, year, q.prop, r.prop,
                          gdd.cum, yday, elev.class, elev.class2, 
                          management, transect, elev.mean, 
                          lat, lon, break1, break2)) %>%
  group_by(site, date) %>%
  mutate(prop = count/sum(count))
}

# Binomial ggplot smoother
binomial_smooth <- function(...) {
  geom_smooth(method = "glm", method.args = list(family = "quasibinomial"), se = FALSE, ...)
}

# Rare_net function: rarefies networks by iteratively subsampling more frequently sampled networks to match the transect count of the least frequently sampled one
## group level
rare_net_gp <- function(net, n, iter) {

  out <- data.frame()

  for(i in 1:iter) {
    temp <- net %>%
      select(site, date) %>%
      unique() %>%
      group_by(site) %>%
      sample_n(n) %>% # drawn n sites at random
      left_join(net) %>%
      group_by(site, bb.sp, plant.sp) %>% # group by site, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(grouplevel, weighted = TRUE) %>% # We probably care mainly about the BB level of the network
      data.frame() %>%
      t() %>%
      data.frame() %>%
      rownames_to_column() %>%
      dplyr::select(site = rowname, everything()) %>%
      as_tibble() %>%
      mutate(iteration = rep(i, n()))

    out <- bind_rows(out, temp)
  }

  out <- out %>%
      gather(metric, value, -c(site, iteration)) %>% # get mean value across iterations for each metric
      group_by(site, metric) %>%
      summarize(value = mean(value)) %>%
      left_join(site_data)
}

# species level
rare_net_sp <- function(net, n, iter) {

  out <- data.frame()

  for(i in 1:iter) {
    temp <- net %>%
      select(site, date) %>%
      unique() %>%
      group_by(site) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(specieslevel, level = "higher") %>%
      map(rownames_to_column) %>%
      map(as_tibble) %>%
      bind_rows(.id = "site") %>% # Collapse list into big data frame
      rename(sp = rowname) %>%
      mutate(iteration = rep(i, n()))

    out <- bind_rows(out, temp)
  } 
  out <- out %>%
    gather(metric, value, -c(sp, site, iteration)) %>% # get mean value across iterations for each metric
    group_by(sp, site, metric) %>%
    summarize(value = mean(value)) %>%
    left_join(site_data) %>%
    left_join(bb_traits, by = c("sp" = "bb.sp"))
}

# network level
rare_net_nt <- function(net, n, iter) {

  out <- data.frame()

  for(i in 1:iter) {
    temp <- net %>%
      select(site, date) %>%
      unique() %>%
      group_by(site) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(networklevel, weighted = TRUE) %>% # We probably care mainly about the BB level of the network
      data.frame() %>%
      t() %>%
      data.frame() %>%
      rownames_to_column() %>%
      dplyr::select(site = rowname, everything()) %>%
      as_tibble() %>%
      mutate(iteration = rep(i, n()))

    out <- bind_rows(out, temp)
  } 
  out <- out  %>%
      gather(metric, value, -c(site, iteration)) %>% # get mean value across iterations for each metric
      group_by(site, metric) %>%
      summarize(value = mean(value)) %>%
      left_join(site_data)
}
```

### Load site data
```{r}
site_data <- read_csv("./processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) # turn elav.class into an ordered factor
```

### Climate data
```{r}
gdd <- read_csv("./processed_data/climate.csv") %>%
  mutate(year = factor(year)) %>%
  select(site, date, gdd.cum)
```

### Colony phase breakpoint analysis
```{r}
castes <- read_csv("./processed_data/network.csv") %>%
  group_by(site, date, year, caste) %>%
  summarize(caste.count = n()) %>%
  spread(caste, caste.count) %>%
  replace(is.na(.), 0) %>%
  mutate(queen = oldqueen + queen,
         reproductive = male + youngqueen,
         q.prop = queen/(queen+worker+reproductive),
         r.prop = reproductive/(queen+worker+reproductive)) %>%
  select(-oldqueen, -youngqueen, -male) %>%
  left_join(gdd, by = c("site", "date")) %>%
  left_join(site_data, by = c("site")) %>%
  arrange(gdd.cum) %>%
  mutate(yday = yday(date))

breaks_2012_gdd <- breakpoints_gdd(castes, 2012, t1 = 0.75, t2 = 0.1)
breaks_2011_gdd <- breakpoints_gdd(castes, 2011, t1 = 0.75, t2 = 0.1)
breaks_2010_gdd <- breakpoints_gdd(castes, 2010, t1 = 0.75, t2 = 0.1)

breaks_2012_yday <- breakpoints_yday(castes, 2012, t1 = 0.75, t2 = 0.1)
breaks_2011_yday <- breakpoints_yday(castes, 2011, t1 = 0.75, t2 = 0.1)
breaks_2010_yday <- breakpoints_yday(castes, 2010, t1 = 0.75, t2 = 0.1)

breaks_gdd_summary <- full_join(breaks_2012_gdd, 
                                 breaks_2011_gdd) %>%
  full_join(breaks_2010_gdd)

write_csv(breaks_gdd_summary, "./output/breaks_gdd.csv")

breaks_yday_summary <- full_join(breaks_2012_yday, 
                                 breaks_2011_yday) %>%
  full_join(breaks_2010_yday)

write_csv(breaks_yday_summary, "./output/breaks_yday.csv")

breaks_yday <- bind_rows(breaks_2012_yday, breaks_2011_yday, breaks_2010_yday) %>%
  ungroup() %>%
  select(site, year, break1.yday = break1, break2.yday = break2) %>%
  unique()

breaks_gdd <- bind_rows(breaks_2012_gdd, breaks_2011_gdd, breaks_2010_gdd) %>%
  ungroup() %>%
  select(site, year, break1.gdd = break1, break2.gdd = break2) %>%
  unique()

breaks <- full_join(breaks_yday, breaks_gdd, by = c("site", "year")) %>%
  mutate(year = factor(year))
```

### Plot phase breakpoints
```{r}
ggplot(breaks_yday_summary, aes(yday, prop, color = caste)) +
  geom_point() +
  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
  geom_vline(aes(xintercept = break1), linetype = "dashed") +
  geom_vline(aes(xintercept = break2), linetype = "dashed") +
  facet_grid(elev.class2 ~ year) +
  xlab("Julian day") +
  ylab("Proportional abundance")

ggplot(breaks_gdd_summary, aes(gdd.cum, prop, color = caste)) +
  geom_point() +
  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
  geom_vline(aes(xintercept = break1), linetype = "dashed") +
  geom_vline(aes(xintercept = break2), linetype = "dashed") +
  facet_grid(elev.class2 ~ year) +
  xlab("GDD") +
  ylab("Proportional abundance")
```

### Load network and survey data
```{r}
bb_traits <- read_csv("./processed_data/bb_traits.csv") %>%
  select(subgenus, bb.sp, pbl.w, pbl.w.class)

net <- read_csv("./processed_data/network.csv") %>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  unite(site.date, site, date, sep = "_", remove = FALSE) %>%
  mutate(year = factor(year)) %>% # convert year to factor
  full_join(bb_traits, by = c("bb.sp" = "bb.sp")) %>%
  left_join(site_data, by = "site") %>%
  left_join(gdd, by = c("site", "date")) %>%
  left_join(breaks, by = c("site", "year")) %>%
  ungroup() %>%
  mutate(phase = case_when(
    gdd.cum < break1.gdd ~ "founding",
    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
    gdd.cum > break2.gdd ~ "reproductive"
  )) %>%
  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
  arrange(site, date)

survey <- read_csv("./processed_data/floral_survey.csv") %>%
  mutate(year = factor(year)) %>% # convert year to factor
  semi_join(net, by = c("site", "plant.sp")) %>% # Consider only plants visited at a given site-date
  left_join(gdd, by = c("site", "date")) %>%
  left_join(breaks, by = c("site", "year")) %>%
  ungroup() %>%
  mutate(phase = case_when(
    gdd.cum < break1.gdd ~ "founding",
    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
    gdd.cum > break2.gdd ~ "reproductive"
  )) %>%
  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
  arrange(site, date)
```

### BB density, diversity, and packing
```{r eval=FALSE, echo=FALSE}
### Per-transect Bombus density
bb_dens <- net %>%
  group_by(site, date, gdd.cum, year, elev.class2, elev.mean, break1.gdd, break2.gdd) %>%
  summarize(bb.dens = n())

### Per-transect Bombus richness
bb_rich <- net %>%
  group_by(site, date, bb.sp, gdd.cum, year, elev.class2, elev.mean, break1.gdd, break2.gdd) %>%
  summarize(bb.rich = n())

### Per-transect floral density
fl_dens <- survey %>%
  group_by(site, date, gdd.cum, year, break1.gdd, break2.gdd) %>%
  summarize(fl.dens = sum(flower.cover),
            log.fl.dens = log(sum(flower.cover)))

### Per-transect floral richness
fl_rich <- survey %>%
  filter(flower.cover > 0) %>%
  group_by(site, date, gdd.cum, year, break1.gdd, break2.gdd) %>%
  summarize(fl.rich = n())

### Join BB density and floral cover
dens_div <- bb_dens %>%
  full_join(bb_rich) %>%
  full_join(fl_dens) %>% # Note that there is not perfect alignment in site-dates between the network observations and the floral surveying, so this join will cause some NAs that will need to be removed, causing some transects to drop out of the dataset
  full_join(fl_rich) %>%
  na.omit() %>%
  ungroup %>%
  mutate(bb.per.fl = bb.dens/fl.dens,
         site = factor(site),
         yday = yday(date),
         week = week(date),
         month = month(date)) %>%
  ungroup() %>%
  mutate(phase = case_when(
    gdd.cum < break1.gdd ~ "founding",
    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
    gdd.cum > break2.gdd ~ "reproductive"
  )) %>%
  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive")))

ggplot(dens_div, aes(log.fl.dens, bb.dens, color = year)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(elev.class2 ~ phase, scales = "free") +
  ylab("BB density") +
  xlab("log(floral density)")

ggplot(dens_div, aes(log.fl.dens, bb.rich, color = year)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(elev.class2 ~ phase, scales = "free") +
  ylab("BB richness") +
  xlab("log(floral density)")

ggplot(dens_div, aes(fl.rich, bb.dens, color = year)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(elev.class2 ~ phase, scales = "free") +
  ylab("BB density") +
  xlab("Floral richness")

ggplot(dens_div, aes(fl.rich, bb.rich, color = year)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(elev.class2 ~ phase, scales = "free") +
  ylab("BB richness") +
  xlab("Floral richness")

ggplot(dens_div, aes(fl.rich, log.fl.dens, color = year)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(elev.class2 ~ phase, scales = "free") +
  ylab("log(floral density") +
  xlab("Floral richness")

ggplot(dens_div, aes(fl.rich, bb.per.fl, color = year)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  facet_grid(elev.class2 ~ phase, scales = "free") +
  ylab("BB packing") +
  xlab("Florla richness")

ggplot(dens_div, aes(elev.mean, bb.per.fl, color = year)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  facet_wrap(~phase, scales = "free") +
  ylab("BB packing") +
  xlab("Elevation")

ggplot(dens_div, aes(gdd.cum, bb.per.fl)) +
  geom_point() +
  #geom_smooth(method = "lm", se = TRUE) +
  geom_smooth() +
  facet_grid(year~elev.class2, scales = "free") +
  ylab("BB packing") +
  xlab("GDD")

ggplot(dens_div, aes(gdd.cum, sqrt(bb.per.fl))) +
  geom_point() +
  #geom_smooth(method = "lm", se = TRUE) +
  geom_smooth() +
  facet_grid(elev.class2 ~ year, scales = "free") +
  ylab("BB packing") +
  xlab("GDD")

ggplot(dens_div, aes(yday, sqrt(bb.per.fl))) +
  geom_point() +
  #geom_smooth(method = "lm", se = TRUE) +
  geom_smooth() +
  facet_grid(elev.class2 ~ year, scales = "free") +
  ylab("BB packing") +
  xlab("Julian day")

ggplot(dens_div, aes(elev.mean, sqrt(bb.rich))) +
  geom_point() +
  geom_smooth() +
  facet_grid(year ~ phase)

ggplot(dens_div, aes(elev.mean, bb.dens)) +
  geom_point() +
  geom_smooth() +
  facet_grid(year ~ phase)
```

### Rarefied network analysis
```{r}
### Determine the minimum transect count per site*year to set rarefaction sampling
transects <- net %>%
  select(site, year, date) %>%
  unique()

transects_per_siteyear <- transects %>%
  group_by(site, year) %>%
  summarize(transects = n())

min_transects_per_year <- transects_per_siteyear %>%
  group_by(year) %>%
  summarize(min_transects = min(transects))

### Create yearly subsets to feed to rare_net
net_2012 <- filter(net, year == "2012")
net_2011 <- filter(net, year == "2011")
net_2010 <- filter(net, year == "2010")
```

### group level
```{r}
#### Create rarified networks and tabulate grouplevel metrics
groupmet_rare_2012 <- rare_net_gp(net_2012, n = 9, iter = 20)
groupmet_rare_2011 <- rare_net_gp(net_2011, n = 4, iter = 20)
groupmet_rare_2010 <- rare_net_gp(net_2010, n = 3, iter = 20)

### Plot
ggplot(filter(groupmet_rare_2012, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2011, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2010, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2012, metric == "C.score.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2011, metric == "C.score.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2010, metric == "C.score.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2012, metric == "V.ratio.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2011, metric == "V.ratio.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2010, metric == "V.ratio.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2012, metric == "discrepancy.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2011, metric == "discrepancy.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2010, metric == "discrepancy.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2012, metric == "generality.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2011, metric == "generality.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2010, metric == "generality.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()
```

### species level
```{r}
### Species level
spmet_2012 <- rare_net_sp(net_2012, 9, 20) 
spmet_2012_mean <- spmet_2012 %>%
  group_by(site, elev.mean, metric) %>%
  summarize(mean.value = mean(value))

spmet_2011 <- rare_net_sp(net_2011, 4, 20) %>%
  group_by(site, elev.mean, iteration) %>%
  summarize(mean.d = mean(d))
spmet_2010 <- rare_net_sp(net_2010, 3, 20) %>%
  group_by(site, elev.mean, iteration) %>%
  summarize(mean.d = mean(d))

ggplot(filter(spmet_2012_mean, metric == "d"), aes(elev.mean, mean.value)) +
  geom_point() +
  geom_smooth()

ggplot(spmet_2011, aes(elev.mean, mean.d)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~iteration)

ggplot(spmet_2010, aes(elev.mean, mean.d)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~iteration)

```

### network level
```{r}
#### Create rarified networks and tabulate grouplevel metrics
netmet_rare_2012 <- rare_net_nt(net_2012, n = 9, iter = 20)
netmet_rare_2011 <- rare_net_nt(net_2011, n = 4, iter = 20)
netmet_rare_2010 <- rare_net_nt(net_2010, n = 3, iter = 20)

### Plot
ggplot(filter(netmet_rare_2012, metric == "H2"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(netmet_rare_2011, metric == "H2"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(netmet_rare_2010, metric == "H2"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(netmet_rare_2012, aes(elev.mean, number.of.compartments)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~iteration)

ggplot(netmet_rare_2011, aes(elev.mean, number.of.compartments)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~iteration)

ggplot(netmet_rare_2010, aes(elev.mean, number.of.compartments)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~iteration)

ggplot(netmet_rare_2012, aes(elev.mean, web.asymmetry)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~iteration)

ggplot(netmet_rare_2011, aes(elev.mean, web.asymmetry)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~iteration)

ggplot(netmet_rare_2010, aes(elev.mean, web.asymmetry)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~iteration)
```

### Let's look at how abundance varies on a per-species basis
```{r}
species <- net %>%
  group_by(site, date, year, dayofyear, gdd.cum, phase, elev.mean, elev.class2, bb.sp, subgenus, pbl.w, pbl.w.class) %>%
  summarize(abund = n()) %>%
  group_by(site, date) %>%
  mutate(prop = abund/sum(abund))

subgenera <- net %>%
  group_by(site, date, year, dayofyear, gdd.cum, phase, elev.mean, elev.class2, subgenus, pbl.w, pbl.w.class) %>%
  summarize(abund = n()) %>%
  group_by(site, date) %>%
  mutate(prop = abund/sum(abund))

ggplot(species, aes(dayofyear, abund)) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  facet_grid(year ~ bb.sp)

ggplot(species, aes(dayofyear, prop)) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  facet_grid(year ~ bb.sp)

ggplot(species, aes(elev.mean, abund)) +
  geom_point(alpha = 0.1) +
  geom_smooth(se = FALSE, span = 1) +
  facet_grid(year ~ bb.sp, scales = "free")

ggplot(species, aes(elev.mean, prop)) +
  geom_point(alpha = 0.1) +
  geom_smooth(se = FALSE, span = 1) +
  facet_grid(year ~ bb.sp, scales = "free")

ggplot(subgenera, aes(dayofyear, abund)) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  facet_grid(year ~ subgenus)

ggplot(subgenera, aes(dayofyear, prop)) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  facet_grid(year ~ subgenus)

ggplot(subgenera, aes(elev.mean, abund)) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  facet_grid(year ~ subgenus)

ggplot(subgenera, aes(elev.mean, prop)) +
  geom_point(alpha = 0.1) +
  geom_smooth(se = FALSE, span = 1) +
  facet_grid(year ~ subgenus)

### Pairs analysis
species_pairs <- species %>%
  select(site, year, date, bb.sp, abund) %>%
  ungroup() %>%
  spread(bb.sp, abund) %>%
  replace(is.na(.), 0)

ggplot(species_pairs, aes(pasc, prat)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~year)

ggpairs(species_pairs, columns = c("bss", "camp", "flav", 
                                   "gers", "hort", "humi",
                                   "hypn", "jone", "lapi",
                                   "mend", "mont", "muci",
                                   "pasc", "prat", "psyt",
                                   "pyre", "soro", "wurf"))

ggpairs(species, columns = c("bb.sp", "abund"),
        cardinality_threshold = 20)

```

### Plot some dang networks
```{r}
web <- net %>%
  group_by(year, elev.class2, phase, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
  summarize(freq = n()) %>%
  unite(webID, c(year, elev.class2, phase), sep = "_") %>% # calculate interaction frequency per species pair within each site
  dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs()  # convert to to bipartite web

visweb(web[[1]])
visweb(web[[2]])
```

### 2020-05-23 manuscript.Rmd
oh yeah

@@
 
-Bumble bees (Hymenoptera: *Bombus* spp.), due to their well-studied natural history and amenability to field observation, have become a model system for the study of biotic interactions in ecological communities.
-In particular, the broad and overlapping distributions of bumble bee species, together with their gross similarity in morphology, behavior, and dietary niche, have attracted the attention of researchers interested in questions of interspecific competition and, more broadly, the ecological constraints of species coexistence [@Plowright1984-lc; @Goulson2008-zy].
-In this regard, however, bumble bees have proven somewhat enigmatic [@Ranta1981-vx].
-On the one hand, there is evidence that resource limitation, a necessary precondition for competition, can be a significant constraint on local bumble populations [@Heinrich1976-wc], even to the point of widespread colony failure due to starvation [@Bowers1985-df].
-Moreover, the species of certain bumble bee communties have been shown to be over-dispersed in terms of tongue length (a matching trait corresponding to floral corolla depth) or phyologeny (understood to encompass unmeasured trait variation), providing evidence for competition as a community assembly process [e.g.
+Bumble bees (Hymenoptera: *Bombus* spp.) have become a model system for the study of biotic interactions in ecological communities.
+In particular, the broad and overlapping distributions of bumble bee species, together with their gross similarity in morphology, behavior, and dietary niche, have attracted the attention of researchers interested in questions of interspecific competition and, more broadly, the rules of community assembly and species coexistence [@Plowright1984-lc; @Goulson2008-zy].
+In this regard, however, bumble bees have proven somewhat enigmatic.
+On the one hand, there is evidence that resource limitation can be a significant constraint on local bumble populations [@Heinrich1976-wc; @Bowers1985-df], and certain bumble bee communties have been shown to be over-dispersed in terms of tongue length (a matching trait corresponding to floral corolla depth) or phyologeny (understood to encompass unmeasured trait variation), providing evidence for competition as a community assembly process [e.g.
 @Inouye1977-kl; @Pellissier2013-pk].
-Nevertheless, it is equally evident that sympatric bumble bee species do not *always* compete for floral resources to the point of competitive exclusion, since bumble bee communities with an apparent excess of species richness and deficit of trait (or phylogenetic) dispersion are commonly observed [@Goulson2008-zy].
-
-To account for this inconsistency, @Ranta1981-vx proposed that bumble bees communities are governed by a spatiotemporally heterogeneous "mosaic" of competitive conditions, arising from the interaction of bumble bee species' traits and nest locations with spatiotemporal variation in floral resource abundance and composition, such that competitive inequalities between species are sufficiently variable (and fine-grained relative to the scale of observation) to allow more species to coexist than predicted by trait-based competition models.
+Nevertheless, bumble bee communities with an apparent excess of species richness and deficit of trait (or phylogenetic) dispersion are commonly observed [@Ranta1981-vx].
+To account for this inconsistency, @Ranta1981-vx proposed that bumble bees communities are governed by a spatiotemporally heterogeneous "mosaic" of competitive conditions, such that competitive inequalities between species are sufficiently variable to allow more species to coexist than predicted by trait-based competition models.
 A complementary explanation, based on the core/satellite species hypothesis [@Hanski1982-hs], is that competitive forces are significant only among regionally abundant and widespread "core species", and the effects of competition are more pronounced in the structure of regional species pools than in that of local assemblages [@Hanski1982-xd].
 
-As pointed out by @Williams1989-vr, however, the question of competition as a process of community assembly, relevant to evolutionary and biogeographic scales of inquiry, differs from that of competition as a behavioral phenomenon within a given community, irrespective of the forces responsible for its assembly.
+As pointed out by @Williams1989-vr, however, the question of competition as a biogeographic/evolutionary-scale driver of communty assembly differs from that of competition as a local/behavioral-scale driver of species' interactions.
+While the role of competition in bumble bee community assembly remains somewhat obscure, there is broad empirical evidence that bumble bees facultatively partition resources on the basis of trait-matching and possibly dominance hierarchy [e.g @Inouye1978-qi; @Morse1982-bh; @Ishii2013-sm; @Brosi2013-mu].
+This distinction between spatiotemporal scales of competition, echoing the logic of @Ranta1981-vx, becomes especially important when the scope of inquiry switches from retrospective to prospective, and all the more so when there is reason to think that the ecological context of competition is undergoing rapid and unprecedented change.
+Since the height of bumble bee competition research in the 1980s, such a shift has indeed occurred, with retrospective questions of community assembly eclipsed by the urgency of conservation in the face of global changes in climate and land use [@Goulson2010-lo].
+Given the uncertain spatiotemporal lags and exogenous drivers of community assembly, inferring the past role of competition through the study of bumble bee community structure may be of limited use in predicting the future fitness and distribution of bumble bee species, together with the ecological functions and services they provide.
+Studying local/behavioral-scale competition in bumble bee communities along climate gradients provides a vital complement to the classical study of bumble bee community structure.
 
-"This is not to deny that interspecific competition affects patterns of foraging by bumble bees, which has been demonstrated by the effect of competitor removal  (e.g.
-Inouye, 1978; Plowright, Pendrel & McLaren, 1978; Plowright & Rodd, 1980; Pleasants,  1981).
-It is  concluded only that interspecific competition does not appear to govern the composition of local  assemblages of bumble bee species in the manner predicted by  the  theory of species-packing." [@Williams1989-vr]
+The study of competition at local/behavioral scales, however, entails two empirical challenges.
+First, it is not enough simply to know which species occur and their similarity in traits/phyologeny; one must record their behavior with respect to their shared resources, and at a frequency appropriate to the temporal scale of inquiry.
+Second, climate variation must be decoupled from confounding spatial effects.
+The use of species interaction networks along steep elevation gradients provides an elegant solution to both of these challenges.
 
-Bring up Hanksi's core/satellite hypothesis.
+Interaction networks provide a tractable analytical framework for the study of competition and resource partitioning [], and steep elevation gradients can serve as space-for-time proxies in the study of climate effects, capturing a broad range of climate variation while minimizing the confounding effects of geographic distance [e.g.
+@Hoiss2015-gv].
 
-Steep elevation gradients can serve as space-for-time proxies in the study of climate effects, capturing a broad range of climate variation  while minimizing the confounding effects of geographic distance [e.g.
+Our study is set in the Berchtesgadener Alps of southwestern Germany, where a steep elevation gradient permits the study of wide environmental variability while minimizing the confounding effects of geographical distance [e.g.
 @Hoiss2015-gv].
+In this context, we formulate a "metahypothesis" concerning the variation of hypothesized competition processes through elevation and time (**Figure 1**).
+Consistent with the stress-gradient hypothesis [@Bertness1994-zx] and with previous empirical work in our study system [@Hoiss2012-sn], we hypothesize an inverse relationship between competition and elevation.
+Along a temporal axis, we hypothesize that the strength of competition is structured by the seasonal life history pattern of bumble bee colonies [@Goulson2010-lo].
+During the solitary phase, when the bumble bee community consists only of foundress queens, competition should be low.
+Both inter- and intraspecific competition should intensify as colonies increase in size during the social phase.
+After the switch point, when colonies begin producing gynes and males instead of workers, intraspecific comeptition should decrease due to the strong morphological differences between queens, males, and workers within each species, but interspecific competition might intensify due to the effective "over-packing" of trait space wih the introduction of multiple morphological castes per species.
+
+In addition to clarifying the basic ecology of bumble bee community assembly, parsing competitive processes in this way has an applied significance for the conservation of bumble bees in the changing environments of the Anthropocene.
+Under the forcing of environmental change, such as climate warming or land use conversion, bumble bee communities might not be affected only in terms of range shifts, but also by a restructuring of competitive interactions, the nature and consequences of which may be difficult to predict.
 
@@
 climatic) stress, with low stress environments characterized by competition and high stress environments by facilitation [@Bertness1994-zx].
-
-When phylogenetic/trait dispersion are used as indices of competition, the relevant question is whether competition played a role in community assembly.
-These indices do not answer the question of whether, right now, competition is happening.
-Trait-matching, niche overlap, specialization (e.g.
-d') based indices are better suited for the question of immediate competiion.
-It could be that competition was not important over the time scales relevant to community assembly, but now it has become a significant phenomenon due to some change in the system.
-Or, it is possible that competition is spatiotemporally transient, and its net effect is not sufficient to cause a trait or phylogenetic signal.
-
-Bumble bees (Hymenoptera: *Bombus* spp.) are classic model organisms for the study of resource competition in biological communities.
-Nevertheless, the role of resource competition in structuring bumble bee communities has proven difficult to generalize.
-Resource limitation, a precondition for competition,
-
-@Heinrich1976-wc found that a bumble bee community can deplete the daily standing nectar pool of a given locality by more than 90%, and starvation has been implicated in the demise of colonies in isolated subalpine meadows [@Bowers1985-df], in.
-Other studies have found local bumble bee communities to be over-dispersed in terms of tongue length [e.g.
-@Inouye1977-kl], a matching trait [sensu @Schleuning2020-vq] corresponding to floral corolla depth, and phylogeny [@Pellissier2013-pk], which can be interpreted as a catch-all for unmeasured trait variation.
-Nevertheless, it is equally evident that sympatric bumble bee species do not *always* compete for floral resources to the point of competitive exclusion, since bumble bee communities with an apparent excess of species richness and deficit of trait dispersion are commonly observed [@Goulson2008-zy].
 
@@
 Rather than viewing this heterogeneity as an obstacle to inferring some universal rule of bumble bee community assembly, our aim in the present study is to investigate the heterogeneity itself as an ecological phenomenon, characterizing the variation in competitive processes (or lack thereof) in bumble bee communities through time and space.
-
-Our study is set in the Berchtesgadener Alps of southwestern Germany, where a steep elevation gradient permits the study of wide environmental variability while minimizing the confounding effects of geographical distance [e.g.
-@Hoiss2015-gv].
-In this context, we formulate a "metahypothesis" concerning the variation of hypothesized competition processes through elevation and time (**Figure 1**).
-Consistent with the stress-gradient hypothesis [@Bertness1994-zx] and with previous empirical work in our study system [@Hoiss2012-sn], we hypothesize an inverse relationship between competition and elevation.
-Along a temporal axis, we hypothesize that the strength of competition is structured by the seasonal life history pattern of bumble bee colonies [@Goulson2010-lo].
-During the solitary phase, when the bumble bee community consists only of foundress queens, competition should be low.
-Both inter- and intraspecific competition should intensify as colonies increase in size during the social phase.
-After the switch point, when colonies begin producing gynes and males instead of workers, intraspecific comeptition should decrease due to the strong morphological differences between queens, males, and workers within each species, but interspecific competition might intensify due to the effective "over-packing" of trait space wih the introduction of multiple morphological castes per species.
-
-In addition to clarifying the basic ecology of bumble bee community assembly, parsing competitive processes in this way has an applied significance for the conservation of bumble bees in the changing environments of the Anthropocene.
-Under the forcing of environmental change, such as climate warming or land use conversion, bumble bee communities might not be affected only in terms of range shifts, but also by a restructuring of competitive interactions, the nature and consequences of which may be difficult to predict.
 

### 2020-05-23 data_proc.R

@@ -39,7 +39,7 @@ network <- read_delim("./raw_data/data_bumblebee_forDoug_2020.04.17.csv", delim 
          month = month(temp.date),
          date = make_date(year, month, day)) %>%
   mutate(bb.sp = str_replace_all(bb.sp, c("luco" = "bss","terr" = "bss", "telu" = "bss", "cryp" = "bss", 
-                                          "sylv" = "psyt", "quad" = "psyt")),
+                                          "sylv" = "psit", "quad" = "psit", "camp" = "psit", "flav" = "psit", "psyt" = "psit")),
          bb.sp.lat = str_replace_all(bb.sp.lat, c("Bombus terrestris/lucorum" = "Bombus ss",
                                         "Bombus terrestris" = "Bombus ss",
                                         "Bombus lucorum" = "Bombus ss",
@@ -47,7 +47,9 @@ network <- read_delim("./raw_data/data_bumblebee_forDoug_2020.04.17.csv", delim 
                                         "Bombus cryptarum" = "Bombus ss",
                                         "B. barbutellus, B. bohemicus, B. sylvestris" = "Psithyrus",
                                         "Bombus sylvestris" = "Psithyrus",
-                                        "Bombus quadricolor" = "Psithyrus")))
+                                        "Bombus quadricolor" = "Psithyrus",
+                                        "Bombus campestris" = "Psithyrus",
+                                        "Bombus flavidus" = "Psithyrus")))
 
 write_csv(network, "./processed_data/network.csv")
 

### 2020-05-25 manuscript_v2.Rmd
added ms v2 on desktop

--- 2020-05-13 manuscript.Rmd
@@
 ---
-title: 
+title: "Bumble bee competition revisited under global climate change"
 author:
 - Douglas B. Sponsler
+- Fabrice Requier
 - Katharina Kallnik
-- Fabrice Requier
 - Alice Claßen
@@
     reference_docx: docx_template.docx
+  pdf_document: default
 bibliography: refs.bib
@@
 
-Conversation with Ingolf
-- Add opening paragraph setting context of clinate change and community ecology
-- Look for divergent population trajectories of species as evidence of competition with temporal lag on effects
-- Where does network analysis fit in?
-A legit question -- no need to jump right to fancy network metrics if simpler approaches work.
-But so far, none of my empirical predictions involve network metrics.
-- Ingolf had a REALLY nice title -- ask him to send it to me.
--  Ingolf said that he does not think that competition from syrphids at high elevation would be significant, but competition from honey bees at low elevation could be
-- Switch Fabrice and Kalli's order
-- Cite Ingolf's recent papers showing the effect of pulses of mass-flowering crop resource on bumble bee health
-- He liked my colony phase figures
-
 ```{r echo = FALSE, include = FALSE}
-### Prepare environment
-library(bipartite)
-library(lme4)
-library(lmerTest)
 library(tidyverse)
-library(lubridate)
-
-### Load site data
-site_data <- read_csv("./processed_data/site_data.csv") %>%
-  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
-  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
-         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) # turn elav.class into an ordered factor
-
-elev_classification <- site_data %>% 
-  select(elev.mean, elev.class) %>% 
-  unique() %>% 
-  arrange(-elev.mean)
-
-### Climate data
-gdd <- read_csv("./processed_data/climate.csv") %>%
-  mutate(year = factor(year)) %>%
-  select(site, date, gdd.cum)
-
-### Load network and survey data
-bb_traits <- read_csv("./processed_data/bb_traits.csv") %>%
-  select(subgenus, bb.sp, pbl.w, pbl.w.class)
-
-net <- read_csv("./processed_data/network.csv") %>%
-  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
-  unite(site.date, site, date, sep = "_", remove = FALSE) %>%
-  mutate(year = factor(year)) %>% # convert year to factor
-  full_join(bb_traits, by = c("bb.sp.lat" = "bb.sp")) %>%
-  left_join(site_data, by = "site") %>%
-  left_join(gdd, by = c("site", "date")) %>%
-  arrange(site, date)
-
-survey <- read_csv("./processed_data/floral_survey.csv") %>%
-  mutate(year = factor(year)) %>% # convert year to factor
-  semi_join(net, by = c("site", "plant.sp")) %>% # Consider only plants visited at a given site-date
-  left_join(gdd, by = c("site", "date")) %>%
-  arrange(site, date)
-
-### Per-transect Bombus density
-bb_dens <- net %>%
-  group_by(site, date, elev.class, elev.mean, year) %>%
-  summarize(bb.dens = n())
-
-### Per-transect floral density
-fl_dens <- survey %>%
-  group_by(site, date, year) %>%
-  summarize(fl.dens = sum(flower.cover),
-            log.fl.dens = log(sum(flower.cover)))
-
-### Join BB density and floral cover
-density <- bb_dens %>%
-  full_join(fl_dens) %>% # Note that there is not perfect alignment in site-dates between the network observations and the floral surveying, so this join will cause some NAs that will need to be removed, causing some transects to drop out of the dataset
-  na.omit() %>%
-  ungroup %>%
-  mutate(bb_per_fl = bb.dens/fl.dens,
-         site = factor(site),
-         yday = yday(date),
-         week = week(date),
-         month = month(date))
-
-density_month.agg <- bb_dens %>%
-  full_join(fl_dens) %>% # Note that there is not perfect alignment in site-dates between the network observations and the floral surveying, so this join will cause some NAs that will need to be removed, causing some transects to drop out of the dataset
-  na.omit() %>%
-  ungroup %>%
-  mutate(site = factor(site),
-         yday = yday(date),
-         week = week(date),
-         month = month(date)) %>%
-  group_by(year, month, site, elev.mean, elev.class) %>%
-  summarize(mean.bb.dens = mean(bb.dens),
-            mean.fl.dens = mean(fl.dens)) %>%
-  mutate(log.mean.fl.dens = log(mean.fl.dens),
-         bb.per.fl = mean.bb.dens/mean.fl.dens)
 ```
@@
 
-Bumble bees (Hymenoptera: *Bombus* spp.) are classic model organisms for the study of resource competition in biological communities.
-Nevertheless, the role of resource competition in structuring bumble bee communities has proven difficult to generalize.
-It is evident that sympatric bumble bees *can* compete for floral resources, and strongly so.
-@Heinrich1976-wc found that a bumble bee community can deplete the daily standing nectar pool of a given locality by more than 90%, and starvation has been implicated in the demise of colonies in isolated subalpine meadows [@Bowers1985-df], leaving little room to doubt the potential for resource limitation.
-Other studies have found local bumble bee communities to be over-dispersed in terms of tongue length [e.g.
-@Inouye1977-kl], a matching trait [sensu @Schleuning2020-vq] corresponding to floral corolla depth, and phylogeny [@Pellissier2013-pk], which can be interpreted as a catch-all for unmeasured trait variation.
-Nevertheless, it is equally evident that sympatric bumble bee species do not *always* compete for floral resources to the point of competitive exclusion, since bumble bee communities with an apparent excess of species richness and deficit of trait dispersion are commonly observed [@Goulson2008-zy].
+The nascent reality of climate change has already revealed the fragile contingency of ecological paradigms that rely on the implicit assumption of climate stability [@Hannah2002-nv; @Walther2002-mq].
+Thus, one of the central questions facing ecologists today is how climate change will affect the future fitness and distribution of species and, thereby, the composition and function of ecological communities.
+While it is intuitive that climate change should affect species and communities through environmental filtering, it is now recognized that climate change can also restructure biotic interactions [@Tylianakis2008-fi], such as competition and facilitation [@Ettinger2017-pk], altering the fitness landscape that governs community assembly and species coexistence [@Chesson2000-aw].
 
-To account for both the importance and variability of competitive processes in bumble bee communities, it is helpful to distinguish three processes that could obtain to varying degrees: no competition, interspecific competition, and intraspecific competition (**Table 1**).
-We will refer to these processes as hypotheses, since their realization in any given community is matter of inference.
-Below, we elaborate each and formulate corresponding empirical predictons.
+Bumble bees (Hymenoptera: *Bombus* spp.) have become a model system for the study of biotic interactions in ecological communities.
+In particular, the broad and overlapping distributions of bumble bee species, together with their gross similarity in morphology, behavior, and dietary niche, have attracted the attention of researchers interested in questions of interspecific competition and, more broadly, the rules of community assembly and species coexistence [@Plowright1984-lc; @Goulson2008-zy].
+In this regard, however, bumble bees have proven somewhat enigmatic.
+On the one hand, there is evidence that resource limitation can be a significant constraint on local bumble populations [@Heinrich1976-wc; @Bowers1985-df], and certain bumble bee communties have been shown to be over-dispersed in terms of tongue length (a matching trait corresponding to floral corolla depth) or phyologeny (understood to encompass unmeasured trait variation), providing evidence for competition as a community assembly process [e.g.
+@Inouye1977-kl; @Pellissier2013-pk].
+Nevertheless, bumble bee communities with an apparent excess of species richness and deficit of trait (or phylogenetic) dispersion are commonly observed [@Ranta1981-vx].
+To account for this inconsistency, @Ranta1981-vx proposed that bumble bees communities are governed by a spatiotemporally heterogeneous mosaic of competitive conditions, such that competitive inequalities between species are sufficiently variable to allow more species to coexist than predicted by trait-based competition models.
+A complementary explanation, based on the core/satellite species hypothesis [@Hanski1982-hs], is that competitive forces are significant only among regionally abundant and widespread "core species", and the effects of competition are more pronounced in the structure of regional species pools than in that of local assemblages [@Hanski1982-xd].
 
-+-------------------------------+----------------------------------------------------+
+As pointed out by @Williams1989-vr, however, the question of competition as a biogeographic/evolutionary-scale driver of communty assembly differs from that of competition as a local/behavioral-scale driver of species' interactions.
+While the role of competition in bumble bee community assembly remains somewhat obscure, there is broad empirical evidence that bumble bees facultatively partition resources to minimize competition [e.g @Inouye1978-qi; @Morse1982-bh; @Ishii2013-sm; @Brosi2013-mu; @Brosi2017-oc].
+This distinction between spatiotemporal scales of competition, echoing the logic of @Ranta1981-vx, becomes especially important when the scope of inquiry switches from retrospective to prospective, and all the more so when there is reason to think that the ecological context of competition is undergoing rapid and unprecedented change.
+Since the height of bumble bee competition research in the 1980s, such a shift has indeed occurred, with retrospective questions of community assembly eclipsed by the urgency of conservation in the face of global changes in climate and land use [@Goulson2010-lo].
+Given the uncertain spatiotemporal lags and exogenous drivers of community assembly, inferring the past role of competition through the study of bumble bee community structure may be of limited use in predicting the future fitness and distribution of bumble bee species.
+In this regard, the study of competition via behavioral patterns of floral resource partitioning that respond in real-time to variation in ecological context [@Spiesman2016-nf] provides a vital complement to the classical study of community assembly.
+
+The study of competition as a behavioral phenomenon, however, is empirically and analytically demanding, since species interactions, not merely community composition, must be measured and interpreted.
+The difficulty is compounded when the goal is to infer the effects of environmental variation, since observations must be replicated across gradients of interest while minimizing the influence of confounding spatial or temporal effects.
+An elegant solution to both these challenges is the use of species interaction networks along steep elevation gradients [e.g.
+@Hoiss2015-gv; @Albrecht2018-gn].
+Interaction network analysis provides a tractable framework for the study of resource partitioning [@Bluthgen2006-if], and steep elevation gradients can serve as space-for-time proxies in the study of climate effects, capturing a broad range of climate variation while minimizing the confounding effects of geographic distance [@Blois2013-wr].
+To our knowledge, however, the analysis of plant-pollinator interaction networks along elevational climate gradients has never been applied specifically to the question of competition in bumble bees and its relation to climate change.
+
+In this study, we analyze patterns of resource partitioning among bumble bees along an elevation gradient in the Berchtesgadener Alps of southeastern Germany.
+Our analysis has a hierarchical structure.
+First, we test the hypothesis that competition is of negligible significance in our study system.
+At least two plausible mechanisms can be proposed that would account for a lack of competition in bumble bee communities [@Ranta1981-vx].
+First, it is possible that a super-abundance of floral resources could saturate the foraging capacity of a local bumble bee population.
+Indeed, this may be a common (though transient) phenomenon associated with the bloom of mass flowering crops, such as red clover [@Ranta1981-vx] and oilseed rape [@Westphal2009-sb].
+Alternatively, competition could be attenuated if abiotic stress, such as low temperature, suppresses bumble bee populations below competition-inducing densities.
+This pattern would be a prediction of the "stress-gradient hypothesis" [@Bertness1994-zx], and while it has not, to our knowledge, been documented for bumble bees specifically, a recent study of whole bee communities in the Berchtesgaden system (of which bumble bees were a part) found evidence of dampened competition at high elevation sites along a montane-to-alpine gradient [@Hoiss2012-sn].
+Under either mechanism, the most straightforward prediction is that bumble bee population density would be decoupled from resource density, since the latter would be non-limiting.
+For the same reason, the absence of competition would also predict that resource partitioning would be constrained only by trait-matching, not by niche overlap; thus, niche partitioning among species with similar traits should not exceed that predicted by a random null model.
+
+Where the hypothesis of no competition can be rejected, we seek to differentiate inter- from intraspecific competition.
+These two forms of competition share the necessary precondition of resource limitation, but they imply distinct predictions regarding resource partitioning and trait-matching.
+Resource partitioning by bumble bees is mediated by variation in matching traits, the most well-studied of these being tongue length.
+Bumble bee tongue lengths vary between species, and the efficiency with which a bumble bee can collect nectar from a given flower depends largely on the closeness of match between the tongue length of the bee and the corolla depth of the flower [@Inouye1980-mv; @Ranta1980-ok].
+Since tongue length also varies within species, and literature values for a given species vary, it is common to average available values and bin species into tongue-length "guilds" [@Pellissier2013-pk].
+
+Where interspecific competition is strong, a species' dietary specialization should increase in proportion to the density of other species within the same tongue length guild per density of floral resources shared by that guild (hereafter referred to as "guild-packing").
+Moreover, species within a tongue length guild should tend to exhibit divergent abundance dynamics, since they are reciprocally limiting.
+Under the pressure of intraspecific competition, however, individuals may respond by foraging more opportunistically within the constraints of their fundamental niche space [].
+Alternatively, a species may relieve intraspecific competition by subdividing, on the basis of intraspecific trait heterogeneity, into differently specialized populations [@Araujo2008-bp].
+In either case, the species-level effect would be an increase in realized niche breadth (i.e.
+a decrease in specialization) in proportion to conspecific-packing (density of conspecifics per density of guild-level floral resources) [though see @Parent2014-hx].
+
+These hypotheses and predictions are summarized in Table 1.
+
++--------------------------------+--------------------------------------------------------+
 | Hypothesis                     | Prediction                                             |
-+===============================+====================================================+
++================================+========================================================+
 | H1: No competition             | 1. BB density !~ FL density                            |
-|                               | 2. BB trait dispersion = null expectation          |
-|                               | 3. BB-FL trait-matching > null expectation         |
-+-------------------------------+----------------------------------------------------+
-| H2: Interspecific competition | 1. BB density ~ FL density                         |
-|                               | 2. BB trait dispersion > null expectation          |
-|                               | 3. BB-FL trait-matching > null expectation         |
-+-------------------------------+----------------------------------------------------+
-| H3: Intraspecific competition | 1. BB density ~ FL density                         |
-|                               | 2. BB diet breadth ~ BB conspecific-packing        |
-|                               | 3. BB-FL trait-matching ~ 1/BB conspecific-packing |
-+-------------------------------+----------------------------------------------------+
++--------------------------------+--------------------------------------------------------+
+| H2a: Interspecific competition | 1. specialization ~ guild-packing                      |
+|                                | 2. within-guild abundance divergence                   | 
++--------------------------------+--------------------------------------------------------+
+| H2b: Intraspecific competition | 1. specialization ~ 1/conspecific-packing              |
+|                                | 2. trait-matching ~ 1/conspecific-packing              |
++--------------------------------+--------------------------------------------------------+
 
@@
 
-*H1: No competition.*    The hypothesis that floral resource competition is not a significant process governing a bumble bee community functions as a "null hypothesis" when competition *per se* is the phenomenon in question, but it is not "null" in the sense of mere negation of mechanistic explanation.
-To the contrary, at least two plausible and ecologically meaningful mechanisms can be proposed that would account for a lack of competition in bumble bee communities [@Ranta1981-vx].
-First, it is possible that a super-abundance of floral resources could saturate the foraging capacity of a local bumble bee population.
-Indeed, this may be a common (though transient) phenomenon associated with the bloom of mass flowering crops [@Ranta1981-vx].
-Alternatively, competition could be attenuated if abiotic stress, such as low temperature, suppresses bumble bee populations below competition-inducing densities.
-This pattern would be a prediction of the "stress-gradient hypothesis" [@Bertness1994-zx], and while it has not, to our knowledge, been documented for bumble bees specifically, a recent study of whole bee communities (of which bumble bees were a part) found evidence of dampened competition at high elevation sites along a montane-to-alpine gradient [@Hoiss2012-sn].
-
-Under either mechanism, the most straightforward prediction is that bumble bee population density would be decoupled from resource density, since the latter would be non-limiting.
-For the same reason, the absence of competition would also predict that the trait dispersion of sympatric bumble bee species would not exceed that predicted by a random null model.
-Third, under a no-competition scenario, bumble bee species would suffer no displacement from their optimal floral resource species.
-Thus, holding other aspects of floral resource value constant (e.g.
-abundance, nectar volume/concentration), a lack of competition would predict that morphological trait-matching between bumble bee species and their chosen host flowers would exceed that predicted by a random null model.
-
-*H2: Interspecific competition.*    Resource partitioning by bumble bees is mediated by variation in matching traits, the most well-studied of these being tongue length.
-Bumble bee tongue lengths vary between species, and the efficiency with which a bumble bee can collect nectar from a given flower depends largely on the closeness of match between the tongue length of the bee and the corolla depth of the flower [@Inouye1980-mv; @Ranta1980-ok].
-Thus, under conditions of interspecific competition, a species' "defensible" niche space (i.e.
-the space in which it is a superior competitor) should center on the floral trait space that corresponds to its tongue length, and the number of species that can be "packed" into a given locality should be limited by some minimal differentiation of tongue length [@Inouye1977-kl; @Pyke1982-gn] within the range of variation supported by the trait space of available flora.
-
-Since resource limitation is a necessary condition for resource competition, a key prediction distiguishing interspecific competition from a lack of competition is that bumble bee density should vary in proportion to floral resource density.
-A bumble bee community under interspecific competition should also exhibit higher tongue length dispersion and stronger trait matching relative to the predictions of a random null model [@Inouye1977-kl].
-
-*H3: Intraspecific competition.*    As noted by @Ranta1981-vx, theory predicts that intraspecific competition can promote species coexistence through the relief of interspecific competition [@MacArthur1984-qz], and could thus, as an alternative to the "no competition" hypothesis, account for cases in which the observed diversity of sympatric bumble bee species exceeds that predicted by trait-based niche models.
-Under the pressure of intraspecific competition, individuals face the depletion of preferred resources and may respond by foraging more opportunistically within the constraints of their fundamental niche space [].
-Alternatively, a population may relieve intraspecific competition by subdividing, on the basis of intraspecific trait heterogeneity, into differently specialized "guilds" [@Araujo2008-bp].
-In either case, the population-level effect would be an increase in the species' realized niche breadth [though see @Parent2014-hx].
-
-Thus, the predictions of intraspecific competition can be formulated as follows.
-First, as in interspecific competition, resources must be limiting, so bumble bee density should vary in proportion to floral density.
-In the case of intraspecific competition, though, we can refine this prediction to say that the density of a given bumble bee species should vary in proportion to the density of the portion of the floral resource community belonging to its dietary niche, since the effects of intraspecific competition are realized on a per-species basis.
-Second, the dietary breadth of a given species should vary in proportion to the density of conspecifics per density of floral resources (we will refer to this as a "conspecific-packing" index).
-Conversely, a broadening of diet should entail a relaxation of trait-matching, so the strength of trait-matching should vary inversely with conspecific-packing.
-
-*Seasonal-Elevational metahypothesis.*    That any of the processes hypothesized above should obtain universally is both empirically and theoretically doubtful [@Ranta1981-vx].
-More plausibly, which of these processes prevails -- and which species are favored by the prevailing process -- vary in both space and time, and this heterogenity of process could explain the structural idiosyncrasies of observed bumble bee communities [@Ranta1981-vx].
-Rather than viewing this heterogeneity as an obstacle to inferring some universal rule of bumble bee community assembly, our aim in the present study is to investigate the heterogeneity itself as an ecological phenomenon, characterizing the variation in competitive processes (or lack thereof) in bumble bee communities through time and space.
-
-Our study is set in the Berchtesgadener Alps of southwestern Germany, where a steep elevation gradient permits the study of wide environmental variability while minimizing the confounding effects of geographical distance [e.g.
-@Hoiss2015-gv].
 In this context, we formulate a "metahypothesis" concerning the variation of hypothesized competition processes through elevation and time (**Figure 1**).
@@
 
-In addition to clarifying the basic ecology of bumble bee community assembly, parsing competitive processes in this way has an applied significance for the conservation of bumble bees in the changing environments of the Anthropocene.
-Under the forcing of environmental change, such as climate warming or land use conversion, bumble bee communities might not be affected only in terms of range shifts, but also by a restructuring of competitive interactions, the nature and consequences of which may be difficult to predict.
+*Seasonal-Elevational metahypothesis.*    That any of the processes hypothesized above should obtain universally is both empirically and theoretically doubtful [@Ranta1981-vx].
+More plausibly, which of these processes prevails -- and which species are favored by the prevailing process -- vary in both space and time, and this heterogenity of process could explain the structural idiosyncrasies of observed bumble bee communities [@Ranta1981-vx].
+Rather than viewing this heterogeneity as an obstacle to inferring some universal rule of bumble bee community assembly, our aim in the present study is to investigate the heterogeneity itself as an ecological phenomenon, characterizing the variation in competitive processes (or lack thereof) in bumble bee communities through time and space.
 
-```{r echo = FALSE, figu}
+```{r echo = FALSE, fig.cap = "**Figure 1:** Hypothesized variation in prevailing competitive processes through time and elevation."}
 Time <- c(0,1,2,3)
@@
 ```
-
-**Figure 1:** Hypothesized variation in prevailing competitive processes through time and elevation. \
+\
 
@@
 
-*Field sampling*
+<!-- - Look for divergent population trajectories of species as evidence of competition with temporal lag on effects -->
+<!-- - Where does network analysis fit in?
+A legit question -- no need to jump right to fancy network metrics if simpler approaches work.
+But so far, none of my empirical predictions involve network metrics.
+-->
+
+*Field sampling.*  Our study is set in the Berchtesgadener Alps of southwestern Germany, a system in which both plant and pollinator communities have been well-studied [@Hoiss2012-sn; Hoiss2012-kr; Hoiss2015-gv].
+Our study sites are the same as those  ranging in elevation from 641 to 2032 m above sea level.
 
@@
 
-*Inferring reosurce limitation*
+*Colony life cycle analysis.* Bumble bee colonies exhibit three distinct life history phases: a "founding" phase during which colonies are initiated by solitary queens, a "build-up" phase during which workers are produced, and a "reproductive" phase where colonies produce new queens and males [@Goulson2010-lo].
+A formal delimitation of these phases serves both as a temporal framework for analyzing bumble bee and floral dynamics and as a check that different elevational subsets of our data capture the same life history interval and are thus comparable in terms of network structure.
+
+To delimit colony phases, we fit binomial regression models to the proportional abundance of foundress queens (Q) and new reproductives (R) (males + new queens), respectively.
+We set the breakpoint between the founding and build-up stages at Q = 0.75, and the breakpoint between buildup and reproductive stages at R = 0.1.
+These cutoffs are somewhat arbitrary points in a continuous process, but their primary purpose is to facilitate comparison across sites.
+This analysis was performed on both a GDD and Julian temporal scale.
+All operations were done in R.
+
+*Inferring resource limitation.* During the founding phase, it would be hard to detect resource limitation even if it were occurring, since the population would not be able to respond to it except by actual starvation of queens.
+In other words, until workers emerge, the relationship between bumble bee density and floral density would be expected to be null.
+At least if competition is the proposed mechanism.
+If mere concentration of foragers is the mechanism, than you would expect to see BB density respond positively to floral density, and the fact that it does not in our data suggests that, at least during the founding phase, forager concentration is not an important mechanism.
+That's good news.
 
 **Results**\
+*Colony life cycle analysis.*
 
-*Colony phase*
-```{r}
-### Get caste counts
-castes <- read_csv("./processed_data/network.csv") %>%
-  group_by(site, date, year, caste) %>%
-  summarize(caste.count = n()) %>%
-  spread(caste, caste.count) %>%
-  replace(is.na(.), 0) %>%
-  mutate(queen = oldqueen + queen,
-         reproductive = male + youngqueen,
-         q.prop = queen/(queen+worker+reproductive),
-         r.prop = reproductive/(queen+worker+reproductive)) %>%
-  select(-oldqueen, -youngqueen, -male) %>%
-  left_join(gdd, by = c("site", "date")) %>%
-  left_join(site_data, by = c("site")) %>%
-  arrange(gdd.cum)
+```{r, echo = FALSE, message = FALSE, fig.cap = "**Figure 2:** Colony life cycle phases delimited by the proportional abundances of foundress queens and new reproductives.
+Dashed lines indicate where foundress queen and reproductive proportional abundances cross the thresholds of 0.75 and 0.10, respectively.
+Binomial regression curves are fit to the proportional abundances, with a spline term added to capture the unimodal pattern of worker abundance (not used in calculating breakpoints)."}
+breaks_yday_summary <- read_csv("./output/breaks_yday.csv") %>%
+  mutate(elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))
+breaks_gdd_summary <- read_csv("./output/breaks_gdd.csv") %>%
+  mutate(elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))
 
-### Logistic regression for first break point (founding to buildup)
-breakpoints <- function(x, year) {
-
-  data <- filter(x, year == !!year)
-
-  # Logistic regression for first break point (founding to buildup)
-  glm1.low <- glm(q.prop ~ gdd.cum,
-                  family = quasibinomial,
-                  data = filter(data, elev.class2 == "low"))
-
-  glm1.mid <- glm(q.prop ~ gdd.cum,
-                  family = quasibinomial,
-                  data = filter(data, elev.class2 == "mid"))
-
-  glm1.high <- glm(q.prop ~ gdd.cum,
-                   family = quasibinomial,
-                   data = filter(data, elev.class2 == "high"))
-
-  # Logistic regression for second break point (buildup to reproductive)
-  glm2.low <- glm(r.prop ~ gdd.cum,
-                  family = quasibinomial,
-                  data = filter(data, elev.class2 == "low"))
-
-  glm2.mid <- glm(r.prop ~ gdd.cum,
-                  family = quasibinomial,
-                  data = filter(data, elev.class2 == "mid"))
-
-  glm2.high <- glm(r.prop ~ gdd.cum,
-                   family = quasibinomial,
-                   data = filter(data, elev.class2 == "high"))
-
-  ### Extract gdd corresponding to desired proportion; based on https://stackoverflow.com/questions/32040504/regression-logistic-in-r-finding-x-value-predictor-for-a-particular-y-value
-
-  getX <- function(x, y) {
-    (log(y/(1-y)) - coef(x)[1])/coef(x)[2]
-    }
-
-  break1.low <- getX(glm1.low, 0.5) %>%
-    data.frame() %>%
-    select(gdd = 1) %>%
-    mutate(elev.class2 = "low",
-           breakID = "break1")
-
-  break1.mid <- getX(glm1.mid, 0.5) %>%
-    data.frame() %>%
-    select(gdd = 1) %>%
-    mutate(elev.class2 = "mid",
-           breakID = "break1")
-
-  break1.high <- getX(glm1.high, 0.5) %>%
-    data.frame() %>%
-    select(gdd = 1) %>%
-    mutate(elev.class2 = "high",
-           breakID = "break1")
-
-  break2.low <- getX(glm2.low, 0.2) %>%
-    data.frame() %>%
-    select(gdd = 1) %>%
-    mutate(elev.class2 = "low",
-           breakID = "break2")
-
-  break2.mid <- getX(glm2.mid, 0.2) %>%
-    data.frame() %>%
-    select(gdd = 1) %>%
-    mutate(elev.class2 = "mid",
-           breakID = "break2")
-
-  break2.high <- getX(glm2.high, 0.2) %>%
-    data.frame() %>%
-    select(gdd = 1) %>%
-    mutate(elev.class2 = "high",
-           breakID = "break2")
-
-  # Join into data frame, convert elev.class2 to ordered factor, spread
-  breaks <- full_join(break1.low, break1.mid) %>%
-    full_join(break1.high) %>%
-    full_join(break2.low) %>%
-    full_join(break2.mid) %>%
-    full_join(break2.high) %>%
-    mutate(elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) %>%
-    spread(breakID, gdd)
-
-  #
-  out <- data %>%
-    full_join(breaks) %>%
-    gather(caste, count, -c(site, date, year, q.prop, r.prop,
-                          gdd.cum, elev.class, elev.class2, 
-                          management, transect, elev.mean, 
-                          lat, lon, break1, break2)) %>%
-  group_by(site, date) %>%
-  mutate(prop = count/sum(count))
-}
-
-breaks_2012 <- breakpoints(castes, 2012)
-breaks_2011 <- breakpoints(castes, 2011)
-breaks_2010 <- breakpoints(castes, 2010)
-
-# Plot
+# Binomial ggplot smoother
 binomial_smooth <- function(...) {
@@
 
-facet_names <- c(`low` = "Low (641-1000 m)",
-                 `mid` = "Mid (1000-1499 m)",
-                 `high` = "High (1500-2032 m)")
-
-ggplot(breaks_2012, aes(gdd.cum, prop, color = caste)) +
+ggplot(breaks_yday_summary, aes(yday, prop, color = caste)) +
   geom_point() +
@@
   geom_vline(aes(xintercept = break2), linetype = "dashed") +
-  facet_wrap(~elev.class2, ncol = 1, labeller = as_labeller(facet_names)) +
-  xlab("Cumulative growing degree days") +
+  facet_grid(elev.class2 ~ year) +
+  xlab("Julian day") +
   ylab("Proportional abundance")
 
-ggsave("./output/phases2012.pdf")
-
-ggplot(breaks_2011, aes(gdd.cum, prop, color = caste)) +
-  geom_point() +
-  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
-  geom_vline(aes(xintercept = break1), linetype = "dashed") +
-  geom_vline(aes(xintercept = break2), linetype = "dashed") +
-  facet_wrap(~elev.class2, ncol = 1, labeller = as_labeller(facet_names)) +
-  xlab("Cumulative growing degree days") +
-  ylab("Proportional abundance")
-
-ggsave("./output/phases2011.pdf")
-
-ggplot(breaks_2010, aes(gdd.cum, prop, color = caste)) +
-  geom_point() +
-  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
-  geom_vline(aes(xintercept = break1), linetype = "dashed") +
-  geom_vline(aes(xintercept = break2), linetype = "dashed") +
-  facet_wrap(~elev.class2, ncol = 1, labeller = as_labeller(facet_names)) +
-  xlab("Cumulative growing degree days") +
-  ylab("Proportional abundance")
-
-ggsave("./output/phases2010.pdf")
+# ggplot(breaks_gdd_summary, aes(gdd.cum, prop, color = caste)) +
+#   geom_point() +
+#   binomial_smooth(formula = y ~ splines::ns(x, 2)) +
+#   geom_vline(aes(xintercept = break1), linetype = "dashed") +
+#   geom_vline(aes(xintercept = break2), linetype = "dashed") +
+#   facet_grid(elev.class2 ~ year) +
+#   xlab("GDD") +
+#   ylab("Proportional abundance")
 ```
+The colony life cycle intervals captured by our sampling varied across years, but were consistent across elevational classes within years, validating the comparison of foraging patterns along our elevation gradient **(Figure 2)**.
+In 2010, all three colony phases are represented in our samples, but the founding phase appears somewhat truncated, suggesting that sampling began after colony founding was already underway.
+In 2011, the founding phase was missed entirely, with sampling beginning mid-way through the build-up phase.
+In 2012, we achieved the longest and densest sampling, yielding good representation of all three colony phases.
+It is worth noting that as elevation increases, colony lifecycle is delayed in onset and contracts in duration, but the phase breakpoints remain approximately equal (on a Julian time scale) across elevation classes.
 
@@
 
-```{r eval=FALSE, echo=FALSE}
-ggplot(density_month.agg, aes(log.mean.fl.dens, mean.bb.dens, color = year)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +
-  facet_wrap(~month) +
-  ylab("Mean BB density") +
-  xlab("log(Mean floral density)")
-```
+*Colony life cycle analysis.*
 
@@
 
-```{r eval=FALSE, echo=FALSE}
-ggplot(density_month.agg, aes(log.mean.fl.dens, mean.bb.dens)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +
-  facet_grid(elev.class ~ month) +
-  ylab("Mean BB density") +
-  xlab("log(Mean floral density)")
-```
-
 <!-- But let's look at this another way.
@@
 -->
-```{r eval=FALSE, echo = FALSE}
-ggplot(density_month.agg, aes(elev.mean, log(bb.per.fl))) +
-  geom_point() +
-  geom_smooth(se = TRUE) +
-  facet_wrap(~year) +
-  ylab("log(bee-packing)") +
-  xlab("Elevation") 
-```
 
@@
 -->
-```{r eval=FALSE, echo=FALSE}
-ggplot(density_month.agg, aes(elev.mean, log(bb.per.fl))) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +
-  facet_grid(year~month, scales = "free") +
-  ylab("log(bee-packing)") +
-  xlab("Elevation") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-```
 
@@
 -->
-```{r eval=FALSE, echo = FALSE}
-### Inspect the evenness (or lack thereof) of sampling intensity
-net_balance <- net %>%
-  group_by(year, elev.mean, elev.class) %>%
-  summarize(observations = n(),
-            transects = length(unique(date)))
-
-ggplot(net_balance, aes(elev.mean, transects)) +
-  geom_point() +
-  facet_wrap(~year, scales = "free")
-```
-
-<!-- A tedious manual rarefaction -->
-```{r eval=FALSE, echo = FALSE}
-### Determine the minimum transect count per site*year to set rarefaction sampling
-transects <- net %>%
-  select(site, year,date) %>%
-  unique()
-
-transects_per_siteyear <- transects %>%
-  group_by(site, year) %>%
-  summarize(transects = n())
-
-min_transects_per_year <- transects_per_siteyear %>%
-  group_by(year) %>%
-  summarize(min_transects = min(transects))
-
-### Create rare_net function
-rare_net <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      select(site, date) %>%
-      unique() %>%
-      group_by(site) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(grouplevel, weighted = TRUE) %>% # We probably care mainly about the BB level of the network
-      data.frame() %>%
-      t() %>%
-      data.frame() %>%
-      rownames_to_column() %>%
-      dplyr::select(site = rowname, everything()) %>%
-      as_tibble() %>%
-      mutate(iteration = rep(i, n())) %>%
-      left_join(site_data)
-
-    out <- bind_rows(out, temp)
-  } 
-  return(as_tibble(out))
-}
-
-### Create yearly subsets to feed to rare_net
-net_2012 <- filter(net, year == "2012")
-net_2011 <- filter(net, year == "2011")
-net_2010 <- filter(net, year == "2010")
-
-#### Create rarified networks and tabulate grouplevel metrics
-groupmet_rare_2012 <- rare_net(net_2012, n = 9, iter = 10)
-groupmet_rare_2011 <- rare_net(net_2011, n = 4, iter = 10)
-groupmet_rare_2010 <- rare_net(net_2010, n = 3, iter = 10)
-
-### Plot
-ggplot(groupmet_rare_2012, aes(elev.mean, niche.overlap.HL)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_wrap(~iteration)
-
-ggplot(groupmet_rare_2011, aes(elev.mean, niche.overlap.HL)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_wrap(~iteration)
-
-ggplot(groupmet_rare_2010, aes(elev.mean, niche.overlap.HL)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_wrap(~iteration)
-
-ggplot(groupmet_rare_2012, aes(elev.mean, C.score.HL)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_wrap(~iteration)
-
-ggplot(groupmet_rare_2011, aes(elev.mean, C.score.HL)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_wrap(~iteration)
-
-ggplot(groupmet_rare_2010, aes(elev.mean, C.score.HL)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_wrap(~iteration)
-```
 

### 2020-05-25 manuscript_v2.Rmd
draft 2 of intro

@@
 output:
+  pdf_document: default
   word_document:
     reference_docx: docx_template.docx
-  pdf_document: default
 bibliography: refs.bib
@@
 
-The nascent reality of climate change has already revealed the fragile contingency of ecological paradigms that rely on the implicit assumption of climate stability [@Hannah2002-nv; @Walther2002-mq].
+The nascent reality of climate change has already revealed the fragile contingency of ecological paradigms that rely on the assumption of climate stability [@Hannah2002-nv; @Walther2002-mq].
 Thus, one of the central questions facing ecologists today is how climate change will affect the future fitness and distribution of species and, thereby, the composition and function of ecological communities.
@@
 In this regard, however, bumble bees have proven somewhat enigmatic.
-On the one hand, there is evidence that resource limitation can be a significant constraint on local bumble populations [@Heinrich1976-wc; @Bowers1985-df], and certain bumble bee communties have been shown to be over-dispersed in terms of tongue length (a matching trait corresponding to floral corolla depth) or phyologeny (understood to encompass unmeasured trait variation), providing evidence for competition as a community assembly process [e.g.
+On the one hand, there is evidence that resource limitation can be a significant constraint on local bumble populations [@Heinrich1976-wc; @Bowers1985-df], and certain bumble bee communities have been shown to be over-dispersed in terms of tongue length (a matching trait corresponding to floral corolla depth) or phyloogeny (understood to encompass unmeasured trait variation), providing evidence for competition as a community assembly process [e.g.
 @Inouye1977-kl; @Pellissier2013-pk].
@@
 
-As pointed out by @Williams1989-vr, however, the question of competition as a biogeographic/evolutionary-scale driver of communty assembly differs from that of competition as a local/behavioral-scale driver of species' interactions.
+As pointed out by @Williams1989-vr, however, the question of competition as a biogeographic/evolutionary-scale driver of community assembly differs from that of competition as a local/behavioral-scale driver of species' interactions.
 While the role of competition in bumble bee community assembly remains somewhat obscure, there is broad empirical evidence that bumble bees facultatively partition resources to minimize competition [e.g @Inouye1978-qi; @Morse1982-bh; @Ishii2013-sm; @Brosi2013-mu; @Brosi2017-oc].
 This distinction between spatiotemporal scales of competition, echoing the logic of @Ranta1981-vx, becomes especially important when the scope of inquiry switches from retrospective to prospective, and all the more so when there is reason to think that the ecological context of competition is undergoing rapid and unprecedented change.
-Since the height of bumble bee competition research in the 1980s, such a shift has indeed occurred, with retrospective questions of community assembly eclipsed by the urgency of conservation in the face of global changes in climate and land use [@Goulson2010-lo].
+Since the height of bumble bee competition research in the 1980s, such a shift has indeed occurred, with retrospective questions of community assembly all but eclipsed by the urgency of conservation in the face of global changes in climate and land use [@Goulson2010-lo].
 Given the uncertain spatiotemporal lags and exogenous drivers of community assembly, inferring the past role of competition through the study of bumble bee community structure may be of limited use in predicting the future fitness and distribution of bumble bee species.
@@
 The study of competition as a behavioral phenomenon, however, is empirically and analytically demanding, since species interactions, not merely community composition, must be measured and interpreted.
-The difficulty is compounded when the goal is to infer the effects of environmental variation, since observations must be replicated across gradients of interest while minimizing the influence of confounding spatial or temporal effects.
-An elegant solution to both these challenges is the use of species interaction networks along steep elevation gradients [e.g.
+The difficulty is compounded when the goal is to infer the effects of climate variation, since observations must be replicated across gradients of interest while minimizing the influence of confounding spatial or temporal effects.
+An elegant solution to both these challenges is the use of species interaction network analysis along steep elevation gradients [e.g.
 @Hoiss2015-gv; @Albrecht2018-gn].
 Interaction network analysis provides a tractable framework for the study of resource partitioning [@Bluthgen2006-if], and steep elevation gradients can serve as space-for-time proxies in the study of climate effects, capturing a broad range of climate variation while minimizing the confounding effects of geographic distance [@Blois2013-wr].
-To our knowledge, however, the analysis of plant-pollinator interaction networks along elevational climate gradients has never been applied specifically to the question of competition in bumble bees and its relation to climate change.
 
-In this study, we analyze patterns of resource partitioning among bumble bees along an elevation gradient in the Berchtesgadener Alps of southeastern Germany.
-Our analysis has a hierarchical structure.
-First, we test the hypothesis that competition is of negligible significance in our study system.
-At least two plausible mechanisms can be proposed that would account for a lack of competition in bumble bee communities [@Ranta1981-vx].
-First, it is possible that a super-abundance of floral resources could saturate the foraging capacity of a local bumble bee population.
-Indeed, this may be a common (though transient) phenomenon associated with the bloom of mass flowering crops, such as red clover [@Ranta1981-vx] and oilseed rape [@Westphal2009-sb].
-Alternatively, competition could be attenuated if abiotic stress, such as low temperature, suppresses bumble bee populations below competition-inducing densities.
-This pattern would be a prediction of the "stress-gradient hypothesis" [@Bertness1994-zx], and while it has not, to our knowledge, been documented for bumble bees specifically, a recent study of whole bee communities in the Berchtesgaden system (of which bumble bees were a part) found evidence of dampened competition at high elevation sites along a montane-to-alpine gradient [@Hoiss2012-sn].
-Under either mechanism, the most straightforward prediction is that bumble bee population density would be decoupled from resource density, since the latter would be non-limiting.
-For the same reason, the absence of competition would also predict that resource partitioning would be constrained only by trait-matching, not by niche overlap; thus, niche partitioning among species with similar traits should not exceed that predicted by a random null model.
+Previous studies of pollinator and plant communities along elevation gradients have tended to find decreased specialization and/or increased niche overlap at high elevation, which is often interpreted as an indication of reduced interspecific competition, though other mechanisms cannot be excluded [@Hoiss2012-sn; @Miller-Struttmann2014-wz; @Hoiss2015-gv; @Classen2020-nh].
+If, in fact, competition in pollinator communities is attenuated at high elevation, this would be broadly consistent with the "stress-gradient hypothesis", which posits that prevailing biotic interactions shift from negative (e.g.
+competition) to positive (e.g.
+facilitation) along ascending gradients of environmental stress [@Bertness1994-zx].
+As applied to pollinator communities, the logic would be that environmental stresses at high elevation, such as low temperatures and short growing seasons, suppress pollinator densities more strongly than the densities of their shared resources, thus causing a per-capita increase in food availability and a corresponding relaxation of competition.
 
-Where the hypothesis of no competition can be rejected, we seek to differentiate inter- from intraspecific competition.
-These two forms of competition share the necessary precondition of resource limitation, but they imply distinct predictions regarding resource partitioning and trait-matching.
-Resource partitioning by bumble bees is mediated by variation in matching traits, the most well-studied of these being tongue length.
-Bumble bee tongue lengths vary between species, and the efficiency with which a bumble bee can collect nectar from a given flower depends largely on the closeness of match between the tongue length of the bee and the corolla depth of the flower [@Inouye1980-mv; @Ranta1980-ok].
-Since tongue length also varies within species, and literature values for a given species vary, it is common to average available values and bin species into tongue-length "guilds" [@Pellissier2013-pk].
-
-Where interspecific competition is strong, a species' dietary specialization should increase in proportion to the density of other species within the same tongue length guild per density of floral resources shared by that guild (hereafter referred to as "guild-packing").
-Moreover, species within a tongue length guild should tend to exhibit divergent abundance dynamics, since they are reciprocally limiting.
-Under the pressure of intraspecific competition, however, individuals may respond by foraging more opportunistically within the constraints of their fundamental niche space [].
-Alternatively, a species may relieve intraspecific competition by subdividing, on the basis of intraspecific trait heterogeneity, into differently specialized populations [@Araujo2008-bp].
-In either case, the species-level effect would be an increase in realized niche breadth (i.e.
-a decrease in specialization) in proportion to conspecific-packing (density of conspecifics per density of guild-level floral resources) [though see @Parent2014-hx].
-
-These hypotheses and predictions are summarized in Table 1.
-
-+--------------------------------+--------------------------------------------------------+
-| Hypothesis                     | Prediction                                             |
-+================================+========================================================+
-| H1: No competition             | 1. BB density !~ FL density                            |
-+--------------------------------+--------------------------------------------------------+
-| H2a: Interspecific competition | 1. specialization ~ guild-packing                      |
-|                                | 2. within-guild abundance divergence                   | 
-+--------------------------------+--------------------------------------------------------+
-| H2b: Intraspecific competition | 1. specialization ~ 1/conspecific-packing              |
-|                                | 2. trait-matching ~ 1/conspecific-packing              |
-+--------------------------------+--------------------------------------------------------+
-
-**Table 1:** Hypothesized competitive processes with empirical predictions. BB = bumble bee; FL = floral. \
-
-In this context, we formulate a "metahypothesis" concerning the variation of hypothesized competition processes through elevation and time (**Figure 1**).
-Consistent with the stress-gradient hypothesis [@Bertness1994-zx] and with previous empirical work in our study system [@Hoiss2012-sn], we hypothesize an inverse relationship between competition and elevation.
-Along a temporal axis, we hypothesize that the strength of competition is structured by the seasonal life history pattern of bumble bee colonies [@Goulson2010-lo].
-During the solitary phase, when the bumble bee community consists only of foundress queens, competition should be low.
-Both inter- and intraspecific competition should intensify as colonies increase in size during the social phase.
-After the switch point, when colonies begin producing gynes and males instead of workers, intraspecific comeptition should decrease due to the strong morphological differences between queens, males, and workers within each species, but interspecific competition might intensify due to the effective "over-packing" of trait space wih the introduction of multiple morphological castes per species.
-
-*Seasonal-Elevational metahypothesis.*    That any of the processes hypothesized above should obtain universally is both empirically and theoretically doubtful [@Ranta1981-vx].
-More plausibly, which of these processes prevails -- and which species are favored by the prevailing process -- vary in both space and time, and this heterogenity of process could explain the structural idiosyncrasies of observed bumble bee communities [@Ranta1981-vx].
-Rather than viewing this heterogeneity as an obstacle to inferring some universal rule of bumble bee community assembly, our aim in the present study is to investigate the heterogeneity itself as an ecological phenomenon, characterizing the variation in competitive processes (or lack thereof) in bumble bee communities through time and space.
-
-```{r echo = FALSE, fig.cap = "**Figure 1:** Hypothesized variation in prevailing competitive processes through time and elevation."}
-Time <- c(0,1,2,3)
-Elevation <- c(0,1,2,3)
-tab <- cbind(Time, Elevation) %>%
-  as_tibble()
-
-ggplot(tab, aes(Time, Elevation)) +
-  geom_vline(xintercept = c(1,2), color = "white") +
-  geom_hline(yintercept = c(1,2), color = "white") +
-  scale_x_continuous(expand = c(0,0),
-                     breaks = c(0.5, 1.5, 2.5), 
-                     labels = c("founding", "build-up", "reproductive"),
-                     limits = c(0,3)) +
-  scale_y_continuous(expand = c(0,0),
-                     breaks = c(0.5, 1.5, 2.5), 
-                     labels = c("low", "mid", "high"),
-                     limits = c(0,3)) +
-  #geom_text() +
-  annotate("text", x = 0.5, y = 2.5, label = "H1") +
-  annotate("text", x = 1.5, y = 2.5, label = "H1") +
-  annotate("text", x = 2.5, y = 2.5, label = "H1") +
-  annotate("text", x = 0.5, y = 1.5, label = "H1") +
-  annotate("text", x = 0.5, y = 0.5, label = "H1") +
-  annotate("text", x = 1.5, y = 1.5, label = "H2 | H3") +
-  annotate("text", x = 2.5, y = 1.5, label = "H2 < H3") +
-  annotate("text", x = 1.5, y = 0.5, label = "H2 | H3") +
-  annotate("text", x = 2.5, y = 0.5, label = "H2 < H3") +
-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
-```
-\
+In this study, we analyze patterns of resource partitioning in a bumble- bee-wildflower network sampled along an elevation gradient in the Berchtesgadener Alps of southeastern Germany.
+We first analyze species-level and group-level metrics of specialization and niche overlap to determine whether the bumble bees in our study system exihibit the expected pattern of decreased specialization and increased niche overlap at high elevation.
+To evaluate the plausibility of intepreting specialization and niche overlap as indices of competition, we then use independent floral surveying data to determine whether observed patterns in specialization and niche overlap correspond to variation in per-capita resource availability.
+Finally, we use the within-year temporal resolution of our dataset to test for reciprocal population dynamics of species within sites as a direct indication of interspecific competition.
+From these results, we derive predictions concerning the effects of climate change on competition as a constraint on bumble bee species and communities.
 
 **Methods**\
-
-<!-- - Look for divergent population trajectories of species as evidence of competition with temporal lag on effects -->
-<!-- - Where does network analysis fit in?
-A legit question -- no need to jump right to fancy network metrics if simpler approaches work.
-But so far, none of my empirical predictions involve network metrics.
--->
 
@@
 
+<!-- Our analysis has a hierarchical structure.
+First, we test the hypothesis that competition is of negligible significance in our study system.
+At least two plausible mechanisms can be proposed that would account for a lack of competition in bumble bee communities [@Ranta1981-vx].
+First, it is possible that a super-abundance of floral resources could saturate the foraging capacity of a local bumble bee population.
+Indeed, this may be a common (though transient) phenomenon associated with the bloom of mass flowering crops, such as red clover [@Ranta1981-vx] and oilseed rape [@Westphal2009-sb].
+Alternatively, competition could be attenuated if abiotic stress, such as low temperature, suppresses bumble bee populations below competition-inducing densities.
+This pattern would be a prediction of the "stress-gradient hypothesis" [@Bertness1994-zx], and while it has not, to our knowledge, been documented for bumble bees specifically, a recent study of whole bee communities in the Berchtesgaden system (of which bumble bees were a part) found evidence of dampened competition at high elevation sites along a montane-to-alpine gradient [@Hoiss2012-sn].
+Under either mechanism, the most straightforward prediction is that bumble bee population density would be decoupled from resource density, since the latter would be non-limiting.
+For the same reason, the absence of competition would also predict that resource partitioning would be constrained only by trait-matching, not by niche overlap; thus, niche partitioning among species with similar traits should not exceed that predicted by a random null model.
+-->
+
+<!-- Where the hypothesis of no competition can be rejected, we seek to differentiate inter- from intraspecific competition.
+These two forms of competition share the necessary precondition of resource limitation, but they imply distinct predictions regarding resource partitioning and trait-matching.
+Resource partitioning by bumble bees is mediated by variation in matching traits, the most well-studied of these being tongue length.
+Bumble bee tongue lengths vary between species, and the efficiency with which a bumble bee can collect nectar from a given flower depends largely on the closeness of match between the tongue length of the bee and the corolla depth of the flower [@Inouye1980-mv; @Ranta1980-ok].
+Since tongue length also varies within species, and literature values for a given species vary, it is common to average available values and bin species into tongue-length "guilds" [@Pellissier2013-pk].
+-->
+
+<!-- Where interspecific competition is strong, a species' dietary specialization should increase in proportion to the density of other species within the same tongue length guild per density of floral resources shared by that guild (hereafter referred to as "guild-packing").
+Moreover, species within a tongue length guild should tend to exhibit divergent abundance dynamics, since they are reciprocally limiting.
+Under the pressure of intraspecific competition, however, individuals may respond by foraging more opportunistically within the constraints of their fundamental niche space [].
+Alternatively, a species may relieve intraspecific competition by subdividing, on the basis of intraspecific trait heterogeneity, into differently specialized populations [@Araujo2008-bp].
+In either case, the species-level effect would be an increase in realized niche breadth (i.e.
+a decrease in specialization) in proportion to conspecific-packing (density of conspecifics per density of guild-level floral resources) [though see @Parent2014-hx].
+-->
+
+<!-- These hypotheses and predictions are summarized in Table 1. -->
+
+<!-- +--------------------------------+--------------------------------------------------------+ -->
+<!-- | Hypothesis                     | Prediction                                             | -->
+<!-- +================================+========================================================+ -->
+<!-- | H1: No competition             | 1. BB density !~ FL density                            | -->
+<!-- +--------------------------------+--------------------------------------------------------+ -->
+<!-- | H2a: Interspecific competition | 1. specialization ~ guild-packing                      | -->
+<!-- |                                | 2. within-guild abundance divergence                   |  -->
+<!-- +--------------------------------+--------------------------------------------------------+ -->
+<!-- | H2b: Intraspecific competition | 1. specialization ~ 1/conspecific-packing              | -->
+<!-- |                                | 2. trait-matching ~ 1/conspecific-packing              | -->
+<!-- +--------------------------------+--------------------------------------------------------+ -->
+
+<!-- **Table 1:** Hypothesized competitive processes with empirical predictions. BB = bumble bee; FL = floral. \ -->
+
+<!-- In this context, we formulate a "metahypothesis" concerning the variation of hypothesized competition processes through elevation and time (**Figure 1**).
+Consistent with the stress-gradient hypothesis [@Bertness1994-zx] and with previous empirical work in our study system [@Hoiss2012-sn], we hypothesize an inverse relationship between competition and elevation.
+Along a temporal axis, we hypothesize that the strength of competition is structured by the seasonal life history pattern of bumble bee colonies [@Goulson2010-lo].
+During the solitary phase, when the bumble bee community consists only of foundress queens, competition should be low.
+Both inter- and intraspecific competition should intensify as colonies increase in size during the social phase.
+After the switch point, when colonies begin producing gynes and males instead of workers, intraspecific comeptition should decrease due to the strong morphological differences between queens, males, and workers within each species, but interspecific competition might intensify due to the effective "over-packing" of trait space wih the introduction of multiple morphological castes per species.
+-->
+
+<!-- *Seasonal-Elevational metahypothesis.*    That any of the processes hypothesized above should obtain universally is both empirically and theoretically doubtful [@Ranta1981-vx].
+More plausibly, which of these processes prevails -- and which species are favored by the prevailing process -- vary in both space and time, and this heterogenity of process could explain the structural idiosyncrasies of observed bumble bee communities [@Ranta1981-vx].
+Rather than viewing this heterogeneity as an obstacle to inferring some universal rule of bumble bee community assembly, our aim in the present study is to investigate the heterogeneity itself as an ecological phenomenon, characterizing the variation in competitive processes (or lack thereof) in bumble bee communities through time and space.
+-->
+
+<!-- ```{r echo = FALSE, fig.cap = "**Figure 1:** Hypothesized variation in prevailing competitive processes through time and elevation."} -->
+<!-- Time <- c(0,1,2,3) -->
+<!-- Elevation <- c(0,1,2,3) -->
+<!-- tab <- cbind(Time, Elevation) %>% -->
+<!--   as_tibble() -->
+
+<!-- ggplot(tab, aes(Time, Elevation)) + -->
+<!--   geom_vline(xintercept = c(1,2), color = "white") + -->
+<!--   geom_hline(yintercept = c(1,2), color = "white") + -->
+<!--   scale_x_continuous(expand = c(0,0), -->
+<!--                      breaks = c(0.5, 1.5, 2.5),  -->
+<!--                      labels = c("founding", "build-up", "reproductive"), -->
+<!--                      limits = c(0,3)) + -->
+<!--   scale_y_continuous(expand = c(0,0), -->
+<!--                      breaks = c(0.5, 1.5, 2.5),  -->
+<!--                      labels = c("low", "mid", "high"), -->
+<!--                      limits = c(0,3)) + -->
+<!--   #geom_text() + -->
+<!--   annotate("text", x = 0.5, y = 2.5, label = "H1") + -->
+<!--   annotate("text", x = 1.5, y = 2.5, label = "H1") + -->
+<!--   annotate("text", x = 2.5, y = 2.5, label = "H1") + -->
+<!--   annotate("text", x = 0.5, y = 1.5, label = "H1") + -->
+<!--   annotate("text", x = 0.5, y = 0.5, label = "H1") + -->
+<!--   annotate("text", x = 1.5, y = 1.5, label = "H2 | H3") + -->
+<!--   annotate("text", x = 2.5, y = 1.5, label = "H2 < H3") + -->
+<!--   annotate("text", x = 1.5, y = 0.5, label = "H2 | H3") + -->
+<!--   annotate("text", x = 2.5, y = 0.5, label = "H2 < H3") + -->
+<!--   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) -->
+<!-- ``` -->
+<!-- \ -->
+
 **References**

### 2020-05-27 analysis.Rmd
dang

@@
 # Rare_net function: rarefies networks by iteratively subsampling more frequently sampled networks to match the transect count of the least frequently sampled one
-## group level
-rare_net_gp <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      select(site, date) %>%
-      unique() %>%
-      group_by(site) %>%
-      sample_n(n) %>% # drawn n sites at random
-      left_join(net) %>%
-      group_by(site, bb.sp, plant.sp) %>% # group by site, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(grouplevel, weighted = TRUE) %>% # We probably care mainly about the BB level of the network
-      data.frame() %>%
-      t() %>%
-      data.frame() %>%
-      rownames_to_column() %>%
-      dplyr::select(site = rowname, everything()) %>%
-      as_tibble() %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  }
-
-  out <- out %>%
-      gather(metric, value, -c(site, iteration)) %>% # get mean value across iterations for each metric
-      group_by(site, metric) %>%
-      summarize(value = mean(value)) %>%
-      left_join(site_data)
-}
-
 # species level
@@
       summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
+      filter(freq > 1) %>% # remove singletons
       dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
@@
       frame2webs() %>% # convert to to bipartite web
-      map(specieslevel, level = "higher") %>%
+      map(specieslevel, level = "higher", index = c("d", "species specificity")) %>%
       map(rownames_to_column) %>%
@@
       summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
+      #filter(freq > 1) %>%
       dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
@@
       frame2webs() %>% # convert to to bipartite web
-      map(networklevel, weighted = TRUE) %>% # We probably care mainly about the BB level of the network
+      map(networklevel, weighted = TRUE, level = "higher", index = c("web asymmetry", "weighted NODF", "niche overlap", "C score")) %>% # We probably care mainly about the BB level of the network
       data.frame() %>%
@@
 ```
+### Floral turnover ~ elevation
+```{r}
+fl_sp_elev <- survey %>%
+  group_by(site, plant.sp) %>%
+  summarize(abund = sum(flower.cover)) %>%
+  left_join(site_data) %>%
+  semi_join(net, by = "plant.sp") %>%
+  group_by(plant.sp) %>%
+  mutate(elev.floor = min(elev.mean),
+         elev.ceiling = max(elev.mean),
+         elev.range = elev.ceiling - elev.floor)
+
+fl_sp_use_elev <- net %>%
+  group_by(site, elev.mean, plant.sp) %>%
+  summarize(abund = n()) %>%
+  group_by(plant.sp) %>%
+  mutate(elev.floor = min(elev.mean),
+         elev.ceiling = max(elev.mean),
+         elev.range = elev.ceiling - elev.floor)
+
+ggplot(fl_sp_elev, aes(elev.mean, log(abund))) +
+  geom_point() +
+  #geom_smooth() +
+  geom_vline(aes(xintercept = elev.floor)) +
+  geom_vline(aes(xintercept = elev.ceiling)) +
+  facet_wrap(~reorder(plant.sp, elev.range), scales = "free_y")
+
+ggplot(fl_sp_use_elev, aes(elev.mean, log(abund))) +
+  geom_point() +
+  #geom_smooth() +
+  geom_vline(aes(xintercept = elev.floor)) +
+  geom_vline(aes(xintercept = elev.ceiling)) +
+  facet_wrap(~reorder(plant.sp, elev.range), scales = "free_y")
+```
 
@@
 
-ggplot(dens_div, aes(elev.mean, bb.per.fl, color = year)) +
+ggplot(dens_div, aes(elev.mean, bb.per.fl)) +
   geom_point() +
-  geom_smooth(method = "lm", se = TRUE) +
-  facet_wrap(~phase, scales = "free") +
+  #geom_smooth(method = "lm", se = TRUE) +
+  geom_smooth() +
+  facet_grid(year~phase, scales = "free") +
   ylab("BB packing") +
@@
 net_2010 <- filter(net, year == "2010")
-```
 
-### group level
-```{r}
-#### Create rarified networks and tabulate grouplevel metrics
-groupmet_rare_2012 <- rare_net_gp(net_2012, n = 9, iter = 20)
-groupmet_rare_2011 <- rare_net_gp(net_2011, n = 4, iter = 20)
-groupmet_rare_2010 <- rare_net_gp(net_2010, n = 3, iter = 20)
+### Demonstrate that the distribution of interaction frequencies does not differ across elevation classes; in other words, no systematic bias is introduced to the network analysis due to uneven frequency of species with few observations
+net_freq <- net %>%
+  group_by(site, dayofyear, year, bb.sp, elev.mean, elev.class2) %>%
+  summarize(freq = n())
+
+ggplot(net_freq, aes(elev.class2, freq)) +
+  geom_boxplot() +
+  facet_wrap(~year)
+
+### Demonstrate that, while community composition may vary across sites, there is not evidence that any species is completely excluded from any point along our elevation gradient.
+bb_elev <- net %>% 
+  group_by(bb.sp, elev.mean) %>%
+  summarize(abund = n())
+
+ggplot(bb_elev, aes(elev.mean, bb.sp, color = log(abund))) +
+  geom_point()
+
+ggplot(bb_elev, aes(elev.mean, abund)) +
+  geom_point() +
+  facet_wrap(~bb.sp, scales = "free_y")
+
+### Species level
+spmet_2012 <- rare_net_sp(net_2012, 9, 20) %>%
+  mutate(year = factor(rep("2012", n())))
+spmet_2011 <- rare_net_sp(net_2011, 4, 20) %>%
+  mutate(year = factor(rep("2011", n())))  
+spmet_2010 <- rare_net_sp(net_2010, 3, 20) %>%
+  mutate(year = factor(rep("2010", n())))
+
+spmet <- bind_rows(spmet_2012, spmet_2011, spmet_2010)
+
+### Network- and group-level
+netmet_rare_2012 <- rare_net_nt(net_2012, n = 9, iter = 20) %>%
+  mutate(year = factor(rep("2012", n())))
+netmet_rare_2011 <- rare_net_nt(net_2011, n = 4, iter = 20)  %>%
+  mutate(year = factor(rep("2011", n())))
+netmet_rare_2010 <- rare_net_nt(net_2010, n = 3, iter = 20) %>%
+  mutate(year = factor(rep("2010", n())))
+
+netmet <- bind_rows(netmet_rare_2012, 
+                    netmet_rare_2011, 
+                    netmet_rare_2010) %>%
+  mutate(metric = factor(metric, levels = c("web.asymmetry",
+                                            "weighted.NODF",
+                                            "niche.overlap.HL", 
+                                            "C.score.HL")))
 
 ### Plot
-ggplot(filter(groupmet_rare_2012, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
+ggplot(spmet_2012, aes(elev.mean, value, color = metric)) +
   geom_point() +
-  geom_smooth()
+  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  geom_smooth() +
+  facet_wrap(~sp, scales = "free")
 
-ggplot(filter(groupmet_rare_2011, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
+ggplot(spmet_2011, aes(elev.mean, value, color = metric)) +
   geom_point() +
-  geom_smooth()
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_wrap(~sp, scales = "free")
 
-ggplot(filter(groupmet_rare_2010, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
+ggplot(spmet_2010, aes(elev.mean, value, color = metric)) +
   geom_point() +
-  geom_smooth()
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_wrap(~sp, scales = "free")
 
-ggplot(filter(groupmet_rare_2012, metric == "C.score.HL"), aes(elev.mean, value)) +
+ggplot(netmet, aes(elev.mean, value)) +
   geom_point() +
-  geom_smooth()
-
-ggplot(filter(groupmet_rare_2011, metric == "C.score.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth()
-
-ggplot(filter(groupmet_rare_2010, metric == "C.score.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth()
-
-ggplot(filter(groupmet_rare_2012, metric == "V.ratio.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth()
-
-ggplot(filter(groupmet_rare_2011, metric == "V.ratio.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth()
-
-ggplot(filter(groupmet_rare_2010, metric == "V.ratio.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth()
-
-ggplot(filter(groupmet_rare_2012, metric == "discrepancy.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth()
-
-ggplot(filter(groupmet_rare_2011, metric == "discrepancy.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth()
-
-ggplot(filter(groupmet_rare_2010, metric == "discrepancy.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth()
-
-ggplot(filter(groupmet_rare_2012, metric == "generality.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth()
-
-ggplot(filter(groupmet_rare_2011, metric == "generality.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth()
-
-ggplot(filter(groupmet_rare_2010, metric == "generality.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth()
-```
-
-### species level
-```{r}
-### Species level
-spmet_2012 <- rare_net_sp(net_2012, 9, 20) 
-spmet_2012_mean <- spmet_2012 %>%
-  group_by(site, elev.mean, metric) %>%
-  summarize(mean.value = mean(value))
-
-spmet_2011 <- rare_net_sp(net_2011, 4, 20) %>%
-  group_by(site, elev.mean, iteration) %>%
-  summarize(mean.d = mean(d))
-spmet_2010 <- rare_net_sp(net_2010, 3, 20) %>%
-  group_by(site, elev.mean, iteration) %>%
-  summarize(mean.d = mean(d))
-
-ggplot(filter(spmet_2012_mean, metric == "d"), aes(elev.mean, mean.value)) +
-  geom_point() +
-  geom_smooth()
-
-ggplot(spmet_2011, aes(elev.mean, mean.d)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~iteration)
-
-ggplot(spmet_2010, aes(elev.mean, mean.d)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~iteration)
-
-```
-
-### network level
-```{r}
-#### Create rarified networks and tabulate grouplevel metrics
-netmet_rare_2012 <- rare_net_nt(net_2012, n = 9, iter = 20)
-netmet_rare_2011 <- rare_net_nt(net_2011, n = 4, iter = 20)
-netmet_rare_2010 <- rare_net_nt(net_2010, n = 3, iter = 20)
-
-### Plot
-ggplot(filter(netmet_rare_2012, metric == "H2"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth()
-
-ggplot(filter(netmet_rare_2011, metric == "H2"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth()
-
-ggplot(filter(netmet_rare_2010, metric == "H2"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth()
-
-ggplot(netmet_rare_2012, aes(elev.mean, number.of.compartments)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~iteration)
-
-ggplot(netmet_rare_2011, aes(elev.mean, number.of.compartments)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~iteration)
-
-ggplot(netmet_rare_2010, aes(elev.mean, number.of.compartments)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~iteration)
-
-ggplot(netmet_rare_2012, aes(elev.mean, web.asymmetry)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~iteration)
-
-ggplot(netmet_rare_2011, aes(elev.mean, web.asymmetry)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~iteration)
-
-ggplot(netmet_rare_2010, aes(elev.mean, web.asymmetry)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~iteration)
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_grid(metric~year, scales = "free")
 ```

### 2020-05-27 manuscript_v2.Rmd
dang

@@
 In this study, we analyze patterns of resource partitioning in a bumble- bee-wildflower network sampled along an elevation gradient in the Berchtesgadener Alps of southeastern Germany.
-We first analyze species-level and group-level metrics of specialization and niche overlap to determine whether the bumble bees in our study system exihibit the expected pattern of decreased specialization and increased niche overlap at high elevation.
+We first analyze species- and group-level network metrics of specialization and niche overlap to determine whether the bumble bees in our study system exihibit the expected pattern of decreased specialization and increased niche overlap at high elevation.
 To evaluate the plausibility of intepreting specialization and niche overlap as indices of competition, we then use independent floral surveying data to determine whether observed patterns in specialization and niche overlap correspond to variation in per-capita resource availability.
@@
 *Colony life cycle analysis.*  
-
 ```{r, echo = FALSE, message = FALSE, fig.cap = "**Figure 2:** Colony life cycle phases delimited by the proportional abundances of foundress queens and new reproductives.
@@
 
+*Specialization and niche overlap*.
+
 *Resource limitation*.
@@
 Interestingly, in September, the relationship is strongest and floral density is also lowest.
+
+*Reciprocal population dynamics*.
 

### 2020-05-29 analysis.Rmd
built rarefied pxe network analysis, expanded abund/div analysis

@@
 ### Define functions
+#### Colony phase breakpoint functions
 ```{r}
@@
 }
+```
 
+#### Rarefied network analysis
+```{r}
 # Rare_net function: rarefies networks by iteratively subsampling more frequently sampled networks to match the transect count of the least frequently sampled one
@@
       left_join(site_data)
+}
+
+# network level version 2
+rare_net_nt2 <- function(net, n, iter) {
+
+  out <- data.frame()
+
+  t_filter <- net %>%
+    select(site, date, phase, year) %>%
+    unique() %>%
+    group_by(site, phase, year) %>%
+    mutate(transects = n()) %>%
+    filter(transects >= n)
+
+  for(i in 1:iter) {
+    temp <- net %>%
+      semi_join(t_filter) %>%
+      select(site, date, phase, year) %>%
+      unique() %>%
+      group_by(site, phase, year) %>%
+      sample_n(n) %>%
+      left_join(net) %>%
+      group_by(site, phase, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
+      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
+      #filter(freq > 1) %>%
+      unite(webID, c(site, phase, year), sep = "_") %>%
+      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
+      data.frame() %>% # convert to data frame
+      frame2webs() %>% # convert to to bipartite web
+      map(networklevel, weighted = TRUE, level = "higher",
+          index = c("web asymmetry", "weighted NODF", "H2", "C score", "niche overlap", "generality")) %>% 
+      data.frame() %>%
+      t() %>%
+      data.frame() %>%
+      rownames_to_column() %>%
+      dplyr::select(id = rowname, everything()) %>%
+      as_tibble() %>%
+      mutate(iteration = rep(i, n()))
+
+    out <- bind_rows(out, temp)
+  } 
+  out <- out  %>%
+      gather(metric, value, -c(id, iteration)) %>% # get mean value across iterations for each metric
+      group_by(id, metric) %>%
+      summarize(value = mean(value)) %>%
+      separate(id, sep = "_", into = c("site", "phase", "year")) %>%
+      mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
+      left_join(site_data)
+}
+
+# species level version 2
+rare_net_sp2 <- function(net, n, iter) {
+
+  out <- data.frame()
+
+  t_filter <- net %>%
+    select(site, date, phase, year) %>%
+    unique() %>%
+    group_by(site, phase, year) %>%
+    mutate(transects = n()) %>%
+    filter(transects >= n)
+
+  for(i in 1:iter) {
+    temp <- net %>%
+      semi_join(t_filter) %>%
+      select(site, date, phase, year) %>%
+      unique() %>%
+      group_by(site, phase, year) %>%
+      sample_n(n) %>%
+      left_join(net) %>%
+      group_by(site, phase, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
+      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
+      #filter(freq > 1) %>%
+      unite(webID, c(site, phase, year), sep = "_") %>%
+      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
+      data.frame() %>% # convert to data frame
+      frame2webs() %>% # convert to to bipartite web
+      map(specieslevel, level = "higher", index = c("d", "species strength")) %>%
+      map(rownames_to_column) %>%
+      map(as_tibble) %>%
+      bind_rows(.id = "id") %>% # Collapse list into big data frame
+      rename(sp = rowname) %>%
+      mutate(iteration = rep(i, n()))
+
+    out <- bind_rows(out, temp)
+  } 
+  out <- out %>%
+    gather(metric, value, -c(sp, id, iteration)) %>% # get mean value across iterations for each metric
+    group_by(sp, id, metric) %>%
+    summarize(value = mean(value)) %>%
+    separate(id, sep = "_", into = c("site", "phase", "year")) %>%
+    mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
+    left_join(site_data) %>%
+    left_join(bb_traits, by = c("sp" = "bb.sp"))
 }
@@
   mutate(year = factor(year)) %>% # convert year to factor
-  semi_join(net, by = c("site", "plant.sp")) %>% # Consider only plants visited at a given site-date
+  #semi_join(net, by = c("site", "plant.sp")) %>% # Consider only plants visited at a given site-date
   left_join(gdd, by = c("site", "date")) %>%
@@
 ```
-### Floral turnover ~ elevation
+### BB and floral turnover ~ elevation
 ```{r}
-fl_sp_elev <- survey %>%
+bb_range <- net %>%
+  group_by(site, elev.mean, bb.sp) %>%
+  summarize(abund = n()) %>%
+  group_by(bb.sp) %>%
+  mutate(elev.floor = min(elev.mean),
+         elev.ceiling = max(elev.mean),
+         elev.range = elev.ceiling - elev.floor,
+         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))
+
+fl_survey_elev <- survey %>%
+  filter(flower.cover > 0) %>%
   group_by(site, plant.sp) %>%
@@
   left_join(site_data) %>%
-  semi_join(net, by = "plant.sp") %>%
+  semi_join(net, by = "plant.sp") %>% # constrain to plant species that were actually visited
   group_by(plant.sp) %>%
@@
          elev.ceiling = max(elev.mean),
-         elev.range = elev.ceiling - elev.floor)
+         elev.range = elev.ceiling - elev.floor,
+         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2),
+         sites = length(unique(site)))
 
-fl_sp_use_elev <- net %>%
+fl_net_elev <- net %>%
   group_by(site, elev.mean, plant.sp) %>%
@@
          elev.ceiling = max(elev.mean),
-         elev.range = elev.ceiling - elev.floor)
+         elev.range = elev.ceiling - elev.floor,
+         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2),
+         sites = length(unique(site)))
 
-ggplot(fl_sp_elev, aes(elev.mean, log(abund))) +
-  geom_point() +
-  #geom_smooth() +
-  geom_vline(aes(xintercept = elev.floor)) +
-  geom_vline(aes(xintercept = elev.ceiling)) +
-  facet_wrap(~reorder(plant.sp, elev.range), scales = "free_y")
+ggplot(filter(fl_survey_elev, sites > 0), 
+       aes(reorder(plant.sp, elev.med), elev.mean)) +
+  geom_line(size = 1, color = "gray60") +
+  geom_point(aes(size = abund), alpha = 0.5) +
+  theme_bw() +
+  theme(axis.text.x = element_blank())
 
-ggplot(fl_sp_use_elev, aes(elev.mean, log(abund))) +
-  geom_point() +
-  #geom_smooth() +
-  geom_vline(aes(xintercept = elev.floor)) +
-  geom_vline(aes(xintercept = elev.ceiling)) +
-  facet_wrap(~reorder(plant.sp, elev.range), scales = "free_y")
+ggplot(filter(fl_net_elev, sites > 0), 
+       aes(reorder(plant.sp, elev.med), elev.mean)) +
+  geom_line(size = 1, color = "gray60") +
+  geom_point(aes(size = abund), alpha = 0.5) +
+  theme_bw() +
+  theme(axis.text.x = element_blank())
+
+ggplot(bb_range, aes(reorder(bb.sp, elev.med), elev.mean)) +
+  geom_line(size = 3, color = "gray60") +
+  geom_point(aes(size = abund), alpha = 0.5) +
+  theme_bw()
 ```
 
-### BB density, diversity, and packing
+### BB abundance, diversity, and packing
 ```{r eval=FALSE, echo=FALSE}
 ### Per-transect Bombus density
-bb_dens <- net %>%
-  group_by(site, date, gdd.cum, year, elev.class2, elev.mean, break1.gdd, break2.gdd) %>%
-  summarize(bb.dens = n())
+bb_abund <- net %>%
+  group_by(site, date, gdd.cum, year, elev.class2, 
+           elev.mean, break1.gdd, break2.gdd) %>%
+  summarize(bb.abund = n())
 
@@
 bb_rich <- net %>%
-  group_by(site, date, bb.sp, gdd.cum, year, elev.class2, elev.mean, break1.gdd, break2.gdd) %>%
-  summarize(bb.rich = n())
+  group_by(site, date, gdd.cum, year, elev.class2,
+           elev.mean, break1.gdd, break2.gdd) %>%
+  summarize(bb.rich = length(unique(bb.sp)))
 
-### Per-transect floral density
-fl_dens <- survey %>%
+### Per-transect floral abund
+fl_abund <- survey %>%
   group_by(site, date, gdd.cum, year, break1.gdd, break2.gdd) %>%
-  summarize(fl.dens = sum(flower.cover),
-            log.fl.dens = log(sum(flower.cover)))
+  summarize(fl.abund = sum(flower.cover),
+            log.fl.abund = log(sum(flower.cover)))
 
@@
 ### Join BB density and floral cover
-dens_div <- bb_dens %>%
+abund_div <- bb_abund %>%
   full_join(bb_rich) %>%
-  full_join(fl_dens) %>% # Note that there is not perfect alignment in site-dates between the network observations and the floral surveying, so this join will cause some NAs that will need to be removed, causing some transects to drop out of the dataset
+  full_join(fl_abund) %>% # Note that there is not perfect alignment in site-dates between the network observations and the floral surveying, so this join will cause some NAs that will need to be removed, causing some transects to drop out of the dataset
   full_join(fl_rich) %>%
@@
   ungroup %>%
-  mutate(bb.per.fl = bb.dens/fl.dens,
+  mutate(fl.per.bb = fl.abund/bb.abund,
          site = factor(site),
@@
   )) %>%
-  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive")))
+  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive")),
+         total.rich = bb.rich + fl.rich,
+         web.asymmetry = (bb.rich - fl.rich) / (bb.rich + fl.rich))
 
-ggplot(dens_div, aes(log.fl.dens, bb.dens, color = year)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +
+### BB abundance ~ floral abundance
+ggplot(abund_div, aes(log.fl.abund, bb.abund, color = year)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(method = "lm", se = TRUE) +
   facet_grid(elev.class2 ~ phase, scales = "free") +
-  ylab("BB density") +
-  xlab("log(floral density)")
+  ylab("BB abundance") +
+  xlab("log(floral cover)") +
+  theme_bw()
 
-ggplot(dens_div, aes(log.fl.dens, bb.rich, color = year)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +
+### BB richness ~ floral abundance
+ggplot(abund_div, aes(log.fl.abund, bb.rich, color = year)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(method = "lm", se = TRUE) +
   facet_grid(elev.class2 ~ phase, scales = "free") +
   ylab("BB richness") +
-  xlab("log(floral density)")
+  xlab("log(floral cover)") +
+  theme_bw()
 
-ggplot(dens_div, aes(fl.rich, bb.dens, color = year)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +
+### BB abundance ~ floral richness
+ggplot(abund_div, aes(fl.rich, bb.abund, color = year)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(method = "lm", se = TRUE) +
   facet_grid(elev.class2 ~ phase, scales = "free") +
-  ylab("BB density") +
+  ylab("BB abundance") +
   xlab("Floral richness")
 
-ggplot(dens_div, aes(fl.rich, bb.rich, color = year)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +
+### BB richness ~ floral richness
+ggplot(abund_div, aes(fl.rich, bb.rich, color = year)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(method = "lm", se = TRUE) +
   facet_grid(elev.class2 ~ phase, scales = "free") +
@@
 
-ggplot(dens_div, aes(fl.rich, log.fl.dens, color = year)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +
-  facet_grid(elev.class2 ~ phase, scales = "free") +
-  ylab("log(floral density") +
-  xlab("Floral richness")
-
-ggplot(dens_div, aes(fl.rich, bb.per.fl, color = year)) +
-  geom_point() +
+### Floral richness ~ floral abundance
+ggplot(abund_div, aes(log.fl.abund, fl.rich, color = year)) +
+  geom_point(alpha = 0.5) +
   geom_smooth(method = "lm", se = TRUE) +
   facet_grid(elev.class2 ~ phase, scales = "free") +
-  ylab("BB packing") +
-  xlab("Florla richness")
+  ylab("Floral richness") +
+  xlab("log(floral cover)")
 
-ggplot(dens_div, aes(elev.mean, bb.per.fl)) +
-  geom_point() +
-  #geom_smooth(method = "lm", se = TRUE) +
+### Per-capita floral resource abundance ~ elevation
+ggplot(abund_div, aes(elev.mean, log(fl.per.bb))) +
+  geom_point(alpha = 0.5) +
   geom_smooth() +
   facet_grid(year~phase, scales = "free") +
-  ylab("BB packing") +
-  xlab("Elevation")
+  ylab("log(fl-per-bb)") +
+  xlab("Elevation") +
+  theme_light()
 
-ggplot(dens_div, aes(gdd.cum, bb.per.fl)) +
-  geom_point() +
-  #geom_smooth(method = "lm", se = TRUE) +
-  geom_smooth() +
-  facet_grid(year~elev.class2, scales = "free") +
-  ylab("BB packing") +
-  xlab("GDD")
+### Per-capita floral resource abundance ~ GDD
+ggplot(abund_div, aes(gdd.cum, log(fl.per.bb))) +
+  geom_point(alpha = 0.5, aes(color = phase)) +
+  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
+  facet_grid(elev.class2 ~ year, scales = "free") +
+  ylab("log(fl-per-bb)") +
+  xlab("GDD") +
+  theme_light()
 
-ggplot(dens_div, aes(gdd.cum, sqrt(bb.per.fl))) +
-  geom_point() +
-  #geom_smooth(method = "lm", se = TRUE) +
-  geom_smooth() +
+### Per-capita floral resource abundance ~ Julian day
+ggplot(abund_div, aes(yday, log(fl.per.bb))) +
+  geom_point(alpha = 0.5, aes(color = phase)) +
+  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
   facet_grid(elev.class2 ~ year, scales = "free") +
-  ylab("BB packing") +
-  xlab("GDD")
+  ylab("log(fl-per-bb)") +
+  xlab("Julian day") +
+  theme_light()
 
-ggplot(dens_div, aes(yday, sqrt(bb.per.fl))) +
-  geom_point() +
-  #geom_smooth(method = "lm", se = TRUE) +
-  geom_smooth() +
-  facet_grid(elev.class2 ~ year, scales = "free") +
-  ylab("BB packing") +
-  xlab("Julian day")
-
-ggplot(dens_div, aes(elev.mean, sqrt(bb.rich))) +
+### BB richness ~ elevation
+ggplot(abund_div, aes(elev.mean, bb.rich)) +
   geom_point() +
@@
 
-ggplot(dens_div, aes(elev.mean, bb.dens)) +
+### BB richness ~ GDD
+ggplot(abund_div, aes(gdd.cum, bb.rich)) +
+  geom_point() +
+  geom_smooth() +
+  facet_grid(elev.class2 ~ year)
+
+### BB richness ~ Julian day
+ggplot(abund_div, aes(yday, bb.rich)) +
+  geom_point() +
+  geom_smooth() +
+  facet_grid(elev.class2 ~ year)
+
+### BB abundance ~ elevation
+ggplot(abund_div, aes(elev.mean, bb.abund)) +
+  geom_point() +
+  geom_smooth() +
+  facet_grid(year ~ phase, scales = "free")
+
+### BB abundance ~ GDD
+ggplot(abund_div, aes(gdd.cum, bb.abund)) +
+  geom_point() +
+  geom_smooth() +
+  facet_grid(elev.class2 ~ year)
+
+### BB abundance ~ Julian day
+ggplot(abund_div, aes(yday, bb.abund)) +
+  geom_point() +
+  geom_smooth() +
+  facet_grid(elev.class2 ~ year)
+
+### Floral richness ~ elevation
+ggplot(abund_div, aes(elev.mean, fl.rich)) +
   geom_point() +
@@
   facet_grid(year ~ phase)
+
+### Floral richness ~ GDD
+ggplot(abund_div, aes(gdd.cum, fl.rich)) +
+  geom_point() +
+  geom_smooth() +
+  facet_grid(elev.class2 ~ year)
+
+### Floral richness ~ Julian day
+ggplot(abund_div, aes(yday, fl.rich)) +
+  geom_point() +
+  geom_smooth() +
+  facet_grid(elev.class2 ~ year)
+
+### Floral abundance ~ elevation
+ggplot(abund_div, aes(elev.mean, log(fl.abund))) +
+  geom_point(alpha = 0.5) +
+  geom_smooth() +
+  facet_grid(year ~ phase, scales = "free") +
+  theme_light()
+
+### Floral abundance ~ GDD
+ggplot(abund_div, aes(gdd.cum, log(fl.abund))) +
+  geom_point(aes(color = phase), alpha =  0.5) +
+  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
+  facet_grid(elev.class2 ~ year, scales = "free") +
+  theme_light()
+
+### Floral abundance ~ Julian day
+ggplot(abund_div, aes(yday, log(fl.abund))) +
+  geom_point() +
+  geom_smooth() +
+  facet_grid(elev.class2 ~ year)
+
+### Total (BB = FL) richness ~ elevation
+ggplot(abund_div, aes(elev.mean, total.rich)) +
+  geom_point() +
+  geom_smooth() +
+  facet_grid(year ~ phase, scales = "free")
+
+### Web asymmetry ~ elevation
+ggplot(abund_div, aes(elev.mean, web.asymmetry)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth() +
+  facet_grid(year ~ phase, scales = "free")
+
+### Web asymmetry ~ floral abundance
+ggplot(abund_div, aes(log(fl.abund), web.asymmetry)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth() +
+  facet_grid(year ~ phase, scales = "free")
+
+### Web asymmetry ~ BB abundance
+ggplot(abund_div, aes(bb.abund, web.asymmetry)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth() +
+  facet_grid(year ~ phase, scales = "free")
+
+### Web asymmetry ~ Per-capita floral resource abundance
+ggplot(abund_div, aes(log(fl.per.bb), web.asymmetry)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth() +
+  facet_grid(year ~ phase, scales = "free")
+```
+
+### Phase x elev network analysis
+```{r}
+### Prepare pxe abundance and diversity table
+pxe_abund_div <- abund_div %>%
+  select(-gdd.cum, -yday, -week, -month, -date, -elev.class2) %>%
+  gather(ad.metric, ad.value, -c(site, year, phase, elev.mean)) %>%
+  group_by(site, phase, year, ad.metric) %>%
+  summarize(ad.value = mean(ad.value))
+
+pxe_net <- rare_net_nt2(net, 4, 20) %>%
+  full_join((pxe_abund_div))
+
+ggplot(filter(pxe_net, 
+              metric == "H2" & ad.metric == "fl.abund"),
+       aes(log(ad.value), value)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_grid(year ~ phase, scales = "free")
+
+ggplot(filter(pxe_net, 
+              metric == "H2" & ad.metric == "fl.rich"),
+       aes(log(ad.value), value)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_grid(year ~ phase, scales = "free")
+
+ggplot(filter(pxe_net, 
+              metric == "H2" & ad.metric == "bb.abund"),
+       aes(log(ad.value), value)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_grid(year ~ phase, scales = "free")
+
+ggplot(filter(pxe_net, 
+              metric == "H2" & ad.metric == "bb.rich"),
+       aes(ad.value, value)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_grid(year ~ phase, scales = "free")
+
+ggplot(filter(pxe_net, metric == "weighted.NODF"), aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_grid(year ~ phase, scales = "free")
+
+ggplot(filter(pxe_net, metric == "H2"), aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_grid(year ~ phase, scales = "free")
+
+ggplot(filter(pxe_net, metric == "C.score.HL"), aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_grid(year ~ phase, scales = "free")
+
+ggplot(filter(pxe_net, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_grid(year ~ phase, scales = "free")
+
+ggplot(filter(pxe_net, metric == "generality.HL"), aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_grid(year ~ phase, scales = "free")
+
+ggplot(filter(pxe_net, metric == "web.asymmetry"), aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_grid(year ~ phase, scales = "free")
+
+pxe_sp <- rare_net_sp2(net, 4, 20)  %>%
+  full_join((pxe_abund_div))
+
+ggplot(filter(pxe_sp, 
+              metric == "d" &
+                sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit") &
+                ad.metric == "fl.rich"), 
+       aes(ad.value, value, color = year)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_grid(sp ~ phase, scales = "free") +
+  ylim(c(0,1))
+
+pxe_sp_mean <- pxe_sp %>%
+  filter(sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")) %>%
+  group_by(site, year, phase, elev.mean, metric) %>%
+  summarize(value = mean(value))
+
+ggplot(filter(pxe_sp_mean, metric == "d"), aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_grid(year ~ phase, scales = "free")
+
+ggplot(filter(pxe_sp_mean, metric == "species.strength"), aes(elev.mean, value, color = metric)) +
+  geom_point() +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_grid(year ~ phase, scales = "free")
+
+ggplot(filter(pxe_sp, 
+              metric == "species.strength",
+              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
+       aes(elev.mean, value, color = year)) +
+  geom_point() +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_grid(sp ~ phase, scales = "free")
+
+ggplot(filter(pxe_sp, 
+              metric == "d",
+              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
+       aes(elev.mean, value, color = year)) +
+  geom_point() +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_grid(sp ~ phase, scales = "free") +
+  ylim(c(0,1))
+
+ggplot(filter(pxe_sp, 
+              metric == "d",
+              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
+       aes(elev.mean, value, color = phase)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_grid(year ~ sp, scales = "free") +
+  theme_light()
 ```

### 2020-05-29 analysis.Rmd
added GAM surface analysis

@@
   geom_point(alpha = 0.5) +
+  #stat_summary(aes(y = log(fl.per.bb), group=1), fun.y=mean, colour="black", geom="point",group=1) +
   geom_smooth() +
@@
 ```
+
+### GAM surface plots of abundance and diversity metrics (just for fun)
+```{r}
+library(mgcv)
+library(mgcViz)
+
+percap_surf <- bam(fl.per.bb ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                   data = filter(abund_div, year == 2012),
+                   #family = "poisson", # I think Poisson is the right call since richness data = count data
+                   method = "REML")
+
+webass_surf <- bam(web.asymmetry ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                   data = filter(abund_div, year == 2012),
+                   #family = "poisson", # I think Poisson is the right call since richness data = count data
+                   method = "REML")
+
+bb.rich_surf <- bam(bb.rich ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                   data = filter(abund_div, year == 2012),
+                   #family = "poisson", # I think Poisson is the right call since richness data = count data
+                   method = "REML")
+
+bb.abund_surf <- bam(bb.abund ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                   data = filter(abund_div, year == 2012),
+                   #family = "poisson", # I think Poisson is the right call since richness data = count data
+                   method = "REML")
+
+fl.rich_surf <- bam(fl.rich ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                   data = filter(abund_div, year == 2012),
+                   #family = "poisson", # I think Poisson is the right call since richness data = count data
+                   method = "REML")
+
+fl.abund_surf <- bam(fl.abund ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                   data = filter(abund_div, year == 2012),
+                   #family = "poisson", # I think Poisson is the right call since richness data = count data
+                   method = "REML")
+
+s1 <- getViz(percap_surf)
+plot(s1)
+
+s2 <- getViz(webass_surf)
+plot(s2)
+
+s3 <- getViz(bb.rich_surf)
+plot(s3)
+
+s4 <- getViz(bb.abund_surf)
+plot(s4)
+
+s5 <- getViz(fl.rich_surf)
+plot(s5)
+
+s6 <- getViz(fl.abund_surf)
+plot(s6)
+```

### 2020-05-29 analysis.Rmd
refined GAM analysis

--- 2020-05-29 analysis.Rmd
@@
 library(bipartite)
+library(mgcv)
+library(mgcViz)
 library(lme4)
@@
 ggplot(abund_div, aes(elev.mean, web.asymmetry)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth() +
+  facet_grid(year ~ phase, scales = "free")
+
+### Web asymmetry ~ gdd
+ggplot(abund_div, aes(gdd.cum, web.asymmetry)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth() +
+  facet_grid(year ~ phase, scales = "free")
+
+### Web asymmetry ~ yday
+ggplot(abund_div, aes(yday, web.asymmetry)) +
   geom_point(alpha = 0.5) +
@@
 ```
+
+### GAM surface plots of abundance and diversity metrics (just for fun)
+#### 2012
+```{r}
+### Per-capita resource availability
+percap_yday <- gam(log(fl.per.bb) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                   data = filter(abund_div, year == 2012),
+                   method = "REML")
+
+percap_yday.vis <- getViz(percap_yday)
+check.gamViz(percap_yday.vis)
+plot(sm(percap_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Per-capita FL (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+percap_gdd <- gam(log(fl.per.bb) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                  data = filter(abund_div, year == 2012),
+                  method = "REML")
+
+percap_gdd.vis <- getViz(percap_gdd)
+check.gamViz(percap_gdd.vis)
+plot(sm(percap_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Per-capita FL (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+### Web asymmetry
+#Create a version of abund_div in which web.asymmetry has been forced into postive values (to enable transformation)
+abund_div_trans <- abund_div %>%
+  mutate(web.asymmetry = web.asymmetry - min(web.asymmetry - 0.1))
+
+ggplot(abund_div_trans, aes(web.asymmetry)) +
+  geom_histogram()
+
+webass_yday <- gam(log(web.asymmetry) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                   data = filter(abund_div_trans, year == 2012),
+                   method = "REML")
+
+webass_yday.vis <- getViz(webass_yday)
+check.gamViz(webass_yday.vis)
+plot(sm(webass_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Web asymmetry (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+webass_gdd <- gam(log(web.asymmetry) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                  data = filter(abund_div_trans, year == 2012),
+                  method = "REML")
+
+webass_gdd.vis <- getViz(webass_gdd)
+check.gamViz(webass_gdd.vis)
+plot(sm(webass_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Web asymmetry (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+### Floral abundance
+fl_abund_yday <- gam(log(fl.abund) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2012),
+                    method = "REML")
+
+fl_abund_yday.vis <- getViz(fl_abund_yday)
+check.gamViz(fl_abund_yday.vis)
+plot(sm(fl_abund_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("FL abundance (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+fl_abund_gdd <- gam(log(fl.abund) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2012),
+                    method = "REML")
+
+fl_abund_gdd.vis <- getViz(fl_abund_gdd)
+check.gamViz(fl_abund_gdd.vis)
+plot(sm(fl_abund_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("FL abundance (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+### BB abundance
+bb_abund_yday <- gam(log(bb.abund) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2012),
+                    method = "REML")
+
+bb_abund_yday.vis <- getViz(bb_abund_yday)
+check.gamViz(bb_abund_yday.vis)
+plot(sm(bb_abund_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("BB abundance (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+bb_abund_gdd <- gam(log(bb.abund) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2012),
+                    method = "REML")
+
+bb_abund_gdd.vis <- getViz(bb_abund_gdd)
+check.gamViz(bb_abund_gdd.vis)
+plot(sm(bb_abund_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("BB abundance (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+### BB richness
+bb_rich_yday <- gam(log(bb.rich) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2012),
+                    method = "REML")
+
+bb_rich_yday.vis <- getViz(bb_rich_yday)
+check.gamViz(bb_rich_yday.vis)
+plot(sm(bb_rich_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("BB richness (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+bb_rich_gdd <- gam(log(bb.rich) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2012),
+                    method = "REML")
+
+bb_rich_gdd.vis <- getViz(bb_rich_gdd)
+check.gamViz(bb_rich_gdd.vis)
+plot(sm(bb_rich_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("BB richness (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+```

### 2020-05-29 analysis.Rmd
added 2010 and 2011 GAM analysis

@@
 ```
+#### 2011
+```{r}
+### Per-capita resource availability
+percap_yday <- gam(log(fl.per.bb) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                   data = filter(abund_div, year == 2011),
+                   method = "REML")
+
+percap_yday.vis <- getViz(percap_yday)
+check.gamViz(percap_yday.vis)
+plot(sm(percap_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Per-capita FL (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+percap_gdd <- gam(log(fl.per.bb) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                  data = filter(abund_div, year == 2011),
+                  method = "REML")
+
+percap_gdd.vis <- getViz(percap_gdd)
+check.gamViz(percap_gdd.vis)
+plot(sm(percap_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Per-capita FL (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+### Web asymmetry
+#Create a version of abund_div in which web.asymmetry has been forced into postive values (to enable transformation)
+abund_div_trans <- abund_div %>%
+  mutate(web.asymmetry = web.asymmetry - min(web.asymmetry - 0.1))
+
+ggplot(abund_div_trans, aes(web.asymmetry)) +
+  geom_histogram()
+
+webass_yday <- gam(log(web.asymmetry) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                   data = filter(abund_div_trans, year == 2011),
+                   method = "REML")
+
+webass_yday.vis <- getViz(webass_yday)
+check.gamViz(webass_yday.vis)
+plot(sm(webass_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Web asymmetry (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+webass_gdd <- gam(log(web.asymmetry) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                  data = filter(abund_div_trans, year == 2011),
+                  method = "REML")
+
+webass_gdd.vis <- getViz(webass_gdd)
+check.gamViz(webass_gdd.vis)
+plot(sm(webass_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Web asymmetry (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+### Floral abundance
+fl_abund_yday <- gam(log(fl.abund) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2011),
+                    method = "REML")
+
+fl_abund_yday.vis <- getViz(fl_abund_yday)
+check.gamViz(fl_abund_yday.vis)
+plot(sm(fl_abund_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("FL abundance (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+fl_abund_gdd <- gam(log(fl.abund) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2011),
+                    method = "REML")
+
+fl_abund_gdd.vis <- getViz(fl_abund_gdd)
+check.gamViz(fl_abund_gdd.vis)
+plot(sm(fl_abund_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("FL abundance (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+### BB abundance
+bb_abund_yday <- gam(log(bb.abund) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2011),
+                    method = "REML")
+
+bb_abund_yday.vis <- getViz(bb_abund_yday)
+check.gamViz(bb_abund_yday.vis)
+plot(sm(bb_abund_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("BB abundance (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+bb_abund_gdd <- gam(log(bb.abund) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2011),
+                    method = "REML")
+
+bb_abund_gdd.vis <- getViz(bb_abund_gdd)
+check.gamViz(bb_abund_gdd.vis)
+plot(sm(bb_abund_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("BB abundance (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+### BB richness
+bb_rich_yday <- gam(log(bb.rich) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2011),
+                    method = "REML")
+
+bb_rich_yday.vis <- getViz(bb_rich_yday)
+check.gamViz(bb_rich_yday.vis)
+plot(sm(bb_rich_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("BB richness (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+bb_rich_gdd <- gam(log(bb.rich) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2011),
+                    method = "REML")
+
+bb_rich_gdd.vis <- getViz(bb_rich_gdd)
+check.gamViz(bb_rich_gdd.vis)
+plot(sm(bb_rich_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("BB richness (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+```
+#### 2010
+```{r}
+### Per-capita resource availability
+percap_yday <- gam(log(fl.per.bb) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                   data = filter(abund_div, year == 2010),
+                   method = "REML")
+
+percap_yday.vis <- getViz(percap_yday)
+check.gamViz(percap_yday.vis)
+plot(sm(percap_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Per-capita FL (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+percap_gdd <- gam(log(fl.per.bb) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                  data = filter(abund_div, year == 2010),
+                  method = "REML")
+
+percap_gdd.vis <- getViz(percap_gdd)
+check.gamViz(percap_gdd.vis)
+plot(sm(percap_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Per-capita FL (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+### Web asymmetry
+#Create a version of abund_div in which web.asymmetry has been forced into postive values (to enable transformation)
+abund_div_trans <- abund_div %>%
+  mutate(web.asymmetry = web.asymmetry - min(web.asymmetry - 0.1))
+
+ggplot(abund_div_trans, aes(web.asymmetry)) +
+  geom_histogram()
+
+webass_yday <- gam(log(web.asymmetry) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                   data = filter(abund_div_trans, year == 2010),
+                   method = "REML")
+
+webass_yday.vis <- getViz(webass_yday)
+check.gamViz(webass_yday.vis)
+plot(sm(webass_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Web asymmetry (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+webass_gdd <- gam(log(web.asymmetry) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                  data = filter(abund_div_trans, year == 2010),
+                  method = "REML")
+
+webass_gdd.vis <- getViz(webass_gdd)
+check.gamViz(webass_gdd.vis)
+plot(sm(webass_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Web asymmetry (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+### Floral abundance
+fl_abund_yday <- gam(log(fl.abund) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2010),
+                    method = "REML")
+
+fl_abund_yday.vis <- getViz(fl_abund_yday)
+check.gamViz(fl_abund_yday.vis)
+plot(sm(fl_abund_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("FL abundance (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+fl_abund_gdd <- gam(log(fl.abund) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2010),
+                    method = "REML")
+
+fl_abund_gdd.vis <- getViz(fl_abund_gdd)
+check.gamViz(fl_abund_gdd.vis)
+plot(sm(fl_abund_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("FL abundance (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+### BB abundance
+bb_abund_yday <- gam(log(bb.abund) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2010),
+                    method = "REML")
+
+bb_abund_yday.vis <- getViz(bb_abund_yday)
+check.gamViz(bb_abund_yday.vis)
+plot(sm(bb_abund_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("BB abundance (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+bb_abund_gdd <- gam(log(bb.abund) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2010),
+                    method = "REML")
+
+bb_abund_gdd.vis <- getViz(bb_abund_gdd)
+check.gamViz(bb_abund_gdd.vis)
+plot(sm(bb_abund_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("BB abundance (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+### BB richness
+bb_rich_yday <- gam(log(bb.rich) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2010),
+                    method = "REML")
+
+bb_rich_yday.vis <- getViz(bb_rich_yday)
+check.gamViz(bb_rich_yday.vis)
+plot(sm(bb_rich_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("BB richness (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+bb_rich_gdd <- gam(log(bb.rich) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2010),
+                    method = "REML")
+
+bb_rich_gdd.vis <- getViz(bb_rich_gdd)
+check.gamViz(bb_rich_gdd.vis)
+plot(sm(bb_rich_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("BB richness (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+```

### 2020-06-03 analysis.Rmd
working on HGAM analysis

@@
 library(mgcViz)
-library(lme4)
-library(lmerTest)
+#library(lme4)
+#library(lmerTest)
 library(tidyverse)
@@
       summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
+      group_by(site) %>%
+      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
+      ungroup() %>%
       filter(freq > 1) %>% # remove singletons
@@
       summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      #filter(freq > 1) %>%
+      group_by(site) %>%
+      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
+      ungroup() %>%
       dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
@@
       frame2webs() %>% # convert to to bipartite web
-      map(networklevel, weighted = TRUE, level = "higher", index = c("web asymmetry", "weighted NODF", "niche overlap", "C score")) %>% # We probably care mainly about the BB level of the network
+      map(networklevel, weighted = TRUE, level = "higher", index = c("H2", "weighted NODF", "niche overlap", "C score", "generality.HL")) %>% # We probably care mainly about the BB level of the network
       data.frame() %>%
@@
 
-# network level version 2
-rare_net_nt2 <- function(net, n, iter) {
+# network level PxE
+rare_net_nt_pxe <- function(net, n, iter) {
 
@@
       summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      #filter(freq > 1) %>%
+      group_by(site, phase, year) %>%
+      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
+      ungroup() %>%
       unite(webID, c(site, phase, year), sep = "_") %>%
@@
 castes <- read_csv("./processed_data/network.csv") %>%
+  filter(bb.sp != "psit") %>% # We do not want to consider Psithyrus, since these parasitic species do not have a worker caste
   group_by(site, date, year, caste) %>%
@@
 
-breaks_2012_gdd <- breakpoints_gdd(castes, 2012, t1 = 0.75, t2 = 0.1)
-breaks_2011_gdd <- breakpoints_gdd(castes, 2011, t1 = 0.75, t2 = 0.1)
-breaks_2010_gdd <- breakpoints_gdd(castes, 2010, t1 = 0.75, t2 = 0.1)
-
-breaks_2012_yday <- breakpoints_yday(castes, 2012, t1 = 0.75, t2 = 0.1)
-breaks_2011_yday <- breakpoints_yday(castes, 2011, t1 = 0.75, t2 = 0.1)
-breaks_2010_yday <- breakpoints_yday(castes, 2010, t1 = 0.75, t2 = 0.1)
+breaks_2012_gdd <- breakpoints_gdd(castes, 2012, t1 = 0.5, t2 = 0.15)
+breaks_2011_gdd <- breakpoints_gdd(castes, 2011, t1 = 0.5, t2 = 0.15)
+breaks_2010_gdd <- breakpoints_gdd(castes, 2010, t1 = 0.5, t2 = 0.15)
+
+breaks_2012_yday <- breakpoints_yday(castes, 2012, t1 = 0.75, t2 = 0.15)
+breaks_2011_yday <- breakpoints_yday(castes, 2011, t1 = 0.75, t2 = 0.15)
+breaks_2010_yday <- breakpoints_yday(castes, 2010, t1 = 0.75, t2 = 0.15)
 
@@
 ```{r}
+breaks_summary <- full_join(breaks_gdd_summary, breaks_yday_summary) %>%
+  select(site, date, year, elev.class2, yday, gdd.cum, break1, break2, caste, prop) %>%
+  gather(time.metric, value, -c(site, date, year, elev.class2, break1, break2, caste, prop))
+
 ggplot(breaks_yday_summary, aes(yday, prop, color = caste)) +
-  geom_point() +
+  geom_point(alpha = 0.5) +
   binomial_smooth(formula = y ~ splines::ns(x, 2)) +
@@
 ggplot(breaks_gdd_summary, aes(gdd.cum, prop, color = caste)) +
-  geom_point() +
+  geom_point(alpha = 0.5) +
   binomial_smooth(formula = y ~ splines::ns(x, 2)) +
@@
   xlab("GDD") +
+  ylab("Proportional abundance")
+
+ggplot(breaks_summary, aes(value, prop, color = caste)) +
+  geom_point(alpha = 0.25) +
+  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
+  facet_grid(elev.class2~ time.metric, scales = "free") +
+  xlab("Julian day") +
   ylab("Proportional abundance") 
@@
   left_join(breaks, by = c("site", "year")) %>%
+  left_join(site_data, by = "site") %>%
   ungroup() %>%
@@
   geom_line(size = 1, color = "gray60") +
-  geom_point(aes(size = abund), alpha = 0.5) +
-  theme_bw() +
-  theme(axis.text.x = element_blank())
+  geom_point(alpha = 0.5) +
+  theme_light(14) +
+  theme(axis.text.x = element_blank()) +
+  xlab("Plant species") +
+  ylab("Elevation")
 
@@
   geom_point(aes(size = abund), alpha = 0.5) +
-  theme_bw()
+  theme_light(14) +
+  xlab("BB species") +
+  ylab("Elevation") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 ```
 
-### BB abundance, diversity, and packing
+### BB abundance, diversity, and per-capita resource availability
 ```{r eval=FALSE, echo=FALSE}
@@
 bb_abund <- net %>%
-  group_by(site, date, gdd.cum, year, elev.class2, 
+  group_by(site, date, dayofyear, gdd.cum, year, elev.class2, 
            elev.mean, break1.gdd, break2.gdd) %>%
@@
 bb_rich <- net %>%
-  group_by(site, date, gdd.cum, year, elev.class2,
+  group_by(site, date, dayofyear, gdd.cum, year, elev.class2,
            elev.mean, break1.gdd, break2.gdd) %>%
@@
 fl_abund <- survey %>%
-  group_by(site, date, gdd.cum, year, break1.gdd, break2.gdd) %>%
+  group_by(site, date, dayofyear, gdd.cum, year, elev.class2, 
+           elev.mean, break1.gdd, break2.gdd) %>%
   summarize(fl.abund = sum(flower.cover),
@@
   filter(flower.cover > 0) %>%
-  group_by(site, date, gdd.cum, year, break1.gdd, break2.gdd) %>%
+  group_by(site, date, dayofyear, gdd.cum, year, elev.class2, 
+           elev.mean, break1.gdd, break2.gdd) %>%
   summarize(fl.rich = n())
@@
   full_join(fl_rich) %>%
-  na.omit() %>%
+  #na.omit() %>%
   ungroup %>%
@@
          site = factor(site),
-         yday = yday(date),
+         yday = dayofyear,
          week = week(date),
@@
 
+abund_div_mean <- abund_div %>%
+  select(-dayofyear, -week, -month, -break1.gdd, -break2.gdd) %>%
+  gather(metric, value, -c(site, date, gdd.cum, year, elev.class2, elev.mean, phase)) %>%
+  group_by(site, year, elev.class2, elev.mean, metric) %>%
+  summarize(value = mean(value, na.rm = TRUE))
+
 ### BB abundance ~ floral abundance
@@
 
+ggplot(abund_div, aes(elev.mean, log(fl.per.bb))) +
+  geom_point(alpha = 0.5, aes(color = phase)) +
+  #stat_summary(aes(y = log(fl.per.bb), group=1), fun.y=mean, colour="black", geom="point",group=1) +
+  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
+  facet_grid(year~., scales = "free") +
+  ylab("log(fl-per-bb)") +
+  xlab("Elevation") +
+  theme_light()
+
+ggplot(abund_div, aes(elev.mean, log(fl.per.bb))) +
+  geom_point(alpha = 0.5, aes(color = phase)) +
+  #stat_summary(aes(y = log(fl.per.bb), group=1), fun.y=mean, colour="black", geom="point",group=1) +
+  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
+  ylab("log(fl-per-bb)") +
+  xlab("Elevation") +
+  theme_light()
+
+ggplot(abund_div, aes(elev.mean, fl.per.bb)) +
+  geom_point(alpha = 0.5, aes(color = phase)) +
+  #stat_summary(aes(y = log(fl.per.bb), group=1), fun.y=mean, colour="black", geom="point",group=1) +
+  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
+  ylab("log(fl-per-bb)") +
+  xlab("Elevation") +
+  theme_light()
+
+ggplot(filter(abund_div_mean, metric == "fl.per.bb"), aes(elev.mean, log(value))) +
+  geom_point(alpha = 1) +
+  #geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) +
+  ylab("log(fl-per-bb)") +
+  xlab("Elevation") +
+  facet_wrap(~year) +
+  theme_light()
+
+lm1 <- lm(value ~ elev.mean * year, data = filter(abund_div_mean, metric == "fl.per.bb"))
+summary(lm1)
+
+ggplot(filter(abund_div_mean, metric == "fl.per.bb"), aes(elev.mean, value)) +
+  geom_point(alpha = 1) +
+  #geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
+  geom_smooth() +
+  ylab("fl-per-bb") +
+  xlab("Elevation") +
+  theme_light()
+
 ### Per-capita floral resource abundance ~ GDD
@@
 
+ggplot(abund_div, aes(gdd.cum, log(fl.per.bb))) +
+  geom_point(alpha = 0.5, aes(color = phase)) +
+  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
+  ylab("log(fl-per-bb)") +
+  xlab("GDD") +
+  theme_light()
+
 ### Per-capita floral resource abundance ~ Julian day
@@
 
+ggplot(filter(abund_div_mean, metric == "web.asymmetry"), aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth() +
+  facet_grid(year ~ ., scales = "free")
+
 ### Web asymmetry ~ gdd
@@
   geom_smooth() +
-  facet_grid(year ~ phase, scales = "free")
+  facet_grid(year ~.)
 
@@
   facet_grid(year ~ phase, scales = "free")
+
+ggplot(abund_div, aes(log(fl.per.bb), web.asymmetry)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth() +
+  facet_grid(year ~ .)
 ```
 
-### Phase x elev network analysis
+### GAM surface plots of abundance and diversity metrics
+#### HGAM elevation x time x year (I would really like someone -- maybe my future self -- to check that I'm doing this right)
+##### Floral abundance
 ```{r}
-### Prepare pxe abundance and diversity table
-pxe_abund_div <- abund_div %>%
-  select(-gdd.cum, -yday, -week, -month, -date, -elev.class2) %>%
-  gather(ad.metric, ad.value, -c(site, year, phase, elev.mean)) %>%
-  group_by(site, phase, year, ad.metric) %>%
-  summarize(ad.value = mean(ad.value))
-
-pxe_net <- rare_net_nt2(net, 4, 20) %>%
-  full_join((pxe_abund_div))
-
-ggplot(filter(pxe_net, 
-              metric == "H2" & ad.metric == "fl.abund"),
-       aes(log(ad.value), value)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_grid(year ~ phase, scales = "free")
-
-ggplot(filter(pxe_net, 
-              metric == "H2" & ad.metric == "fl.rich"),
-       aes(log(ad.value), value)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_grid(year ~ phase, scales = "free")
-
-ggplot(filter(pxe_net, 
-              metric == "H2" & ad.metric == "bb.abund"),
-       aes(log(ad.value), value)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_grid(year ~ phase, scales = "free")
-
-ggplot(filter(pxe_net, 
-              metric == "H2" & ad.metric == "bb.rich"),
-       aes(ad.value, value)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_grid(year ~ phase, scales = "free")
-
-ggplot(filter(pxe_net, metric == "weighted.NODF"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(year ~ phase, scales = "free")
-
-ggplot(filter(pxe_net, metric == "H2"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(year ~ phase, scales = "free")
-
-ggplot(filter(pxe_net, metric == "C.score.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(year ~ phase, scales = "free")
-
-ggplot(filter(pxe_net, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(year ~ phase, scales = "free")
-
-ggplot(filter(pxe_net, metric == "generality.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(year ~ phase, scales = "free")
-
-ggplot(filter(pxe_net, metric == "web.asymmetry"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(year ~ phase, scales = "free")
-
-pxe_sp <- rare_net_sp2(net, 4, 20)  %>%
-  full_join((pxe_abund_div))
-
-ggplot(filter(pxe_sp, 
-              metric == "d" &
-                sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit") &
-                ad.metric == "fl.rich"), 
-       aes(ad.value, value, color = year)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_grid(sp ~ phase, scales = "free") +
-  ylim(c(0,1))
-
-pxe_sp_mean <- pxe_sp %>%
-  filter(sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")) %>%
-  group_by(site, year, phase, elev.mean, metric) %>%
-  summarize(value = mean(value))
-
-ggplot(filter(pxe_sp_mean, metric == "d"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(year ~ phase, scales = "free")
-
-ggplot(filter(pxe_sp_mean, metric == "species.strength"), aes(elev.mean, value, color = metric)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(year ~ phase, scales = "free")
-
-ggplot(filter(pxe_sp, 
-              metric == "species.strength",
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
-       aes(elev.mean, value, color = year)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(sp ~ phase, scales = "free")
-
-ggplot(filter(pxe_sp, 
-              metric == "d",
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
-       aes(elev.mean, value, color = year)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(sp ~ phase, scales = "free") +
-  ylim(c(0,1))
-
-ggplot(filter(pxe_sp, 
-              metric == "d",
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
-       aes(elev.mean, value, color = phase)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(year ~ sp, scales = "free") +
-  theme_light()
+fl_abund_yday <- gam(fl.abund ~ year +
+                    te(yday, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
+                    data = filter(abund_div),
+                    family = "tw",
+                    method = "REML")
+
+fl_abund_yday.vis <- getViz(fl_abund_yday)
+check.gamViz(fl_abund_yday.vis)
+summary(fl_abund_yday.vis)
+plot(sm(fl_abund_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("FL abundance 2010 (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+plot(sm(fl_abund_yday.vis, 2)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("FL abundance 2011 (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+plot(sm(fl_abund_yday.vis, 3)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("FL abundance 2012 (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+fl_abund_gdd <- gam(fl.abund ~ year +
+                    te(gdd.cum, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
+                    data = filter(abund_div),
+                    family = "tw",
+                    method = "REML")
+
+fl_abund_gdd.vis <- getViz(fl_abund_gdd)
+check.gamViz(fl_abund_gdd.vis)
+summary(fl_abund_gdd.vis)
+plot(sm(fl_abund_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("FL abundance 2010 (gdd)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+plot(sm(fl_abund_gdd.vis, 2)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("FL abundance 2011 (gdd)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+plot(sm(fl_abund_gdd.vis, 3)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("FL abundance 2012 (gdd)") +
+  xlab("Day of year") +
+  ylab("Elevation")
 ```
 
-### Rarefied network analysis
+##### BB abundance
+```{r}
+bb_abund_yday <- gam(bb.abund ~ year +
+                    te(yday, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
+                    data = filter(abund_div),
+                    family = "scat",
+                    method = "REML")
+
+bb_abund_yday.vis <- getViz(bb_abund_yday)
+check.gamViz(bb_abund_yday.vis)
+summary(bb_abund_yday.vis)
+plot(sm(bb_abund_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("bb abundance 2010 (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+plot(sm(bb_abund_yday.vis, 2)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("bb abundance 2011 (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+plot(sm(bb_abund_yday.vis, 3)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("bb abundance 2012 (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+bb_abund_gdd <- gam(bb.abund ~ year +
+                    te(gdd.cum, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
+                    data = filter(abund_div),
+                    method = "REML")
+
+bb_abund_gdd.vis <- getViz(bb_abund_gdd)
+check.gamViz(bb_abund_gdd.vis)
+summary(bb_abund_gdd.vis)
+plot(sm(bb_abund_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("bb abundance 2010 (gdd)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+plot(sm(bb_abund_gdd.vis, 2)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("bb abundance 2011 (gdd)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+plot(sm(bb_abund_gdd.vis, 3)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("bb abundance 2012 (gdd)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+```
+
+##### Per-capita floral resource abundance
+```{r}
+percap_yday <- gam(fl.per.bb ~ year +
+                    te(yday, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
+                    data = abund_div,
+                    family = "tw",
+                    method = "REML")
+
+percap_yday.vis <- getViz(percap_yday)
+check.gamViz(percap_yday.vis)
+summary(percap_yday.vis)
+plot(sm(percap_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Per-capita floral abundance 2010 (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+plot(sm(percap_yday.vis, 2)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Per-capita floral abundance 2011 (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+plot(sm(percap_yday.vis, 3)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Per-capita floral abundance 2012 (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+percap_gdd <- gam(fl.per.bb ~ year +
+                    te(gdd.cum, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
+                    data = abund_div,
+                    family = "tw",
+                    method = "REML")
+
+percap_gdd.vis <- getViz(percap_gdd)
+check.gamViz(percap_gdd.vis)
+summary(percap_gdd.vis)
+plot(sm(percap_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Per-capita floral abundance 2010 (gdd)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+plot(sm(percap_gdd.vis, 2)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Per-capita floral abundance 2011 (gdd)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+plot(sm(percap_gdd.vis, 3)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Per-capita floral abundance 2012 (gdd)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+```
+
+##### Web asymmetry
+```{r}
+webass_yday <- gam(web.asymmetry ~ year +
+                    te(yday, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
+                    data = abund_div,
+                    family = "tw",
+                    method = "REML")
+
+percap_yday.vis <- getViz(percap_yday)
+check.gamViz(percap_yday.vis)
+summary(percap_yday.vis)
+plot(sm(percap_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Per-capita floral abundance 2010 (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+plot(sm(percap_yday.vis, 2)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Per-capita floral abundance 2011 (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+plot(sm(percap_yday.vis, 3)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Per-capita floral abundance 2012 (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+percap_gdd <- gam(fl.per.bb ~ year +
+                    te(gdd.cum, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
+                    data = abund_div,
+                    family = "tw",
+                    method = "REML")
+
+percap_gdd.vis <- getViz(percap_gdd)
+check.gamViz(percap_gdd.vis)
+summary(percap_gdd.vis)
+plot(sm(percap_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Per-capita floral abundance 2010 (gdd)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+plot(sm(percap_gdd.vis, 2)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Per-capita floral abundance 2011 (gdd)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+plot(sm(percap_gdd.vis, 3)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Per-capita floral abundance 2012 (gdd)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+```
+
+#### 2012
+```{r}
+### Per-capita resource availability
+percap_yday <- gam(log(fl.per.bb) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                   data = filter(abund_div, year == 2012),
+                   method = "REML")
+
+percap_yday.vis <- getViz(percap_yday)
+check.gamViz(percap_yday.vis)
+summary(percap_yday.vis)
+plot(sm(percap_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Per-capita FL (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+percap_gdd <- gam(log(fl.per.bb) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                  data = filter(abund_div, year == 2012),
+                  method = "REML")
+
+percap_gdd.vis <- getViz(percap_gdd)
+check.gamViz(percap_gdd.vis)
+plot(sm(percap_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Per-capita FL (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+### Web asymmetry
+#Create a version of abund_div in which web.asymmetry has been forced into postive values (to enable transformation)
+abund_div_trans <- abund_div %>%
+  mutate(web.asymmetry = web.asymmetry - min(web.asymmetry - 0.1, na.rm = TRUE))
+
+ggplot(abund_div_trans, aes(web.asymmetry)) +
+  geom_histogram()
+
+webass_yday <- gam(log(web.asymmetry) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                   data = filter(abund_div_trans, year == 2012),
+                   method = "REML")
+
+webass_yday.vis <- getViz(webass_yday)
+check.gamViz(webass_yday.vis)
+plot(sm(webass_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Web asymmetry (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+webass_gdd <- gam(log(web.asymmetry) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                  data = filter(abund_div_trans, year == 2012),
+                  method = "REML")
+
+webass_gdd.vis <- getViz(webass_gdd)
+check.gamViz(webass_gdd.vis)
+plot(sm(webass_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Web asymmetry (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+### Floral abundance
+fl_abund_yday <- gam(log(fl.abund) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2012),
+                    method = "REML")
+
+fl_abund_yday.vis <- getViz(fl_abund_yday)
+check.gamViz(fl_abund_yday.vis)
+plot(sm(fl_abund_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("FL abundance (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+fl_abund_gdd <- gam(fl.abund ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2012 & !is.na(fl.abund)),
+                    family = "tw",
+                    method = "REML")
+
+fl_abund_gdd.vis <- getViz(fl_abund_gdd)
+check.gamViz(fl_abund_gdd.vis)
+summary(fl_abund_gdd)
+plot(sm(fl_abund_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("FL abundance (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+### BB abundance
+bb_abund_yday <- gam(bb.abund ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2012),
+                    family = "tw",
+                    method = "REML")
+
+bb_abund_yday.vis <- getViz(bb_abund_yday)
+check.gamViz(bb_abund_yday.vis)
+plot(sm(bb_abund_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("BB abundance (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+bb_abund_gdd <- gam(log(bb.abund) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2012),
+                    method = "REML")
+
+bb_abund_gdd.vis <- getViz(bb_abund_gdd)
+check.gamViz(bb_abund_gdd.vis)
+plot(sm(bb_abund_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("BB abundance (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+### BB richness
+bb_rich_yday <- gam(log(bb.rich) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2012),
+                    method = "REML")
+
+bb_rich_yday.vis <- getViz(bb_rich_yday)
+check.gamViz(bb_rich_yday.vis)
+plot(sm(bb_rich_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("BB richness (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+bb_rich_gdd <- gam(log(bb.rich) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2012),
+                    method = "REML")
+
+bb_rich_gdd.vis <- getViz(bb_rich_gdd)
+check.gamViz(bb_rich_gdd.vis)
+plot(sm(bb_rich_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("BB richness (GDD)") +
+  xlab("GDD") +
+  ylab("Elevation")
+```
+#### 2011
+```{r}
+### Per-capita resource availability
+percap_yday <- gam(log(fl.per.bb) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                   data = filter(abund_div, year == 2011),
+                   method = "REML")
+
+percap_yday.vis <- getViz(percap_yday)
+check.gamViz(percap_yday.vis)
+plot(sm(percap_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Per-capita FL (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+percap_gdd <- gam(log(fl.per.bb) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                  data = filter(abund_div, year == 2011),
+                  method = "REML")
+
+percap_gdd.vis <- getViz(percap_gdd)
+check.gamViz(percap_gdd.vis)
+plot(sm(percap_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Per-capita FL (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+### Web asymmetry
+#Create a version of abund_div in which web.asymmetry has been forced into postive values (to enable transformation)
+abund_div_trans <- abund_div %>%
+  mutate(web.asymmetry = web.asymmetry - min(web.asymmetry - 0.1, na.rm = TRUE))
+
+ggplot(abund_div_trans, aes(web.asymmetry)) +
+  geom_histogram()
+
+webass_yday <- gam(log(web.asymmetry) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                   data = filter(abund_div_trans, year == 2011),
+                   method = "REML")
+
+webass_yday.vis <- getViz(webass_yday)
+check.gamViz(webass_yday.vis)
+plot(sm(webass_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Web asymmetry (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+webass_gdd <- gam(log(web.asymmetry) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                  data = filter(abund_div_trans, year == 2011),
+                  method = "REML")
+
+webass_gdd.vis <- getViz(webass_gdd)
+check.gamViz(webass_gdd.vis)
+plot(sm(webass_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Web asymmetry (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+### Floral abundance
+fl_abund_yday <- gam(log(fl.abund) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2011),
+                    method = "REML")
+
+fl_abund_yday.vis <- getViz(fl_abund_yday)
+check.gamViz(fl_abund_yday.vis)
+plot(sm(fl_abund_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("FL abundance (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+fl_abund_gdd <- gam(log(fl.abund) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2011),
+                    method = "REML")
+
+fl_abund_gdd.vis <- getViz(fl_abund_gdd)
+check.gamViz(fl_abund_gdd.vis)
+plot(sm(fl_abund_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("FL abundance (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+### BB abundance
+bb_abund_yday <- gam(log(bb.abund) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2011),
+                    method = "REML")
+
+bb_abund_yday.vis <- getViz(bb_abund_yday)
+check.gamViz(bb_abund_yday.vis)
+plot(sm(bb_abund_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("BB abundance (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+bb_abund_gdd <- gam(log(bb.abund) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2011),
+                    method = "REML")
+
+bb_abund_gdd.vis <- getViz(bb_abund_gdd)
+check.gamViz(bb_abund_gdd.vis)
+plot(sm(bb_abund_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("BB abundance (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+### BB richness
+bb_rich_yday <- gam(log(bb.rich) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2011),
+                    method = "REML")
+
+bb_rich_yday.vis <- getViz(bb_rich_yday)
+check.gamViz(bb_rich_yday.vis)
+plot(sm(bb_rich_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("BB richness (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+bb_rich_gdd <- gam(log(bb.rich) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2011),
+                    method = "REML")
+
+bb_rich_gdd.vis <- getViz(bb_rich_gdd)
+check.gamViz(bb_rich_gdd.vis)
+plot(sm(bb_rich_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("BB richness (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+```
+#### 2010
+```{r}
+### Per-capita resource availability
+percap_yday <- gam(log(fl.per.bb) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                   data = filter(abund_div, year == 2010),
+                   method = "REML")
+
+percap_yday.vis <- getViz(percap_yday)
+check.gamViz(percap_yday.vis)
+plot(sm(percap_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Per-capita FL (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+percap_gdd <- gam(log(fl.per.bb) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                  data = filter(abund_div, year == 2010),
+                  method = "REML")
+
+percap_gdd.vis <- getViz(percap_gdd)
+check.gamViz(percap_gdd.vis)
+plot(sm(percap_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Per-capita FL (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+### Web asymmetry
+#Create a version of abund_div in which web.asymmetry has been forced into postive values (to enable transformation)
+abund_div_trans <- abund_div %>%
+  mutate(web.asymmetry = web.asymmetry - min(web.asymmetry - 0.1, na.rm = TRUE))
+
+ggplot(abund_div_trans, aes(web.asymmetry)) +
+  geom_histogram()
+
+webass_yday <- gam(log(web.asymmetry) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                   data = filter(abund_div_trans, year == 2010),
+                   method = "REML")
+
+webass_yday.vis <- getViz(webass_yday)
+check.gamViz(webass_yday.vis)
+plot(sm(webass_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Web asymmetry (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+webass_gdd <- gam(log(web.asymmetry) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                  data = filter(abund_div_trans, year == 2010),
+                  method = "REML")
+
+webass_gdd.vis <- getViz(webass_gdd)
+check.gamViz(webass_gdd.vis)
+plot(sm(webass_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("Web asymmetry (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+### Floral abundance
+fl_abund_yday <- gam(log(fl.abund) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2010),
+                    method = "REML")
+
+fl_abund_yday.vis <- getViz(fl_abund_yday)
+check.gamViz(fl_abund_yday.vis)
+plot(sm(fl_abund_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("FL abundance (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+fl_abund_gdd <- gam(log(fl.abund) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2010),
+                    method = "REML")
+
+fl_abund_gdd.vis <- getViz(fl_abund_gdd)
+check.gamViz(fl_abund_gdd.vis)
+plot(sm(fl_abund_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("FL abundance (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+### BB abundance
+bb_abund_yday <- gam(log(bb.abund) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2010),
+                    method = "REML")
+
+bb_abund_yday.vis <- getViz(bb_abund_yday)
+check.gamViz(bb_abund_yday.vis)
+plot(sm(bb_abund_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("BB abundance (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+bb_abund_gdd <- gam(log(bb.abund) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2010),
+                    method = "REML")
+
+bb_abund_gdd.vis <- getViz(bb_abund_gdd)
+check.gamViz(bb_abund_gdd.vis)
+plot(sm(bb_abund_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("BB abundance (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+### BB richness
+bb_rich_yday <- gam(log(bb.rich) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2010),
+                    method = "REML")
+
+bb_rich_yday.vis <- getViz(bb_rich_yday)
+check.gamViz(bb_rich_yday.vis)
+plot(sm(bb_rich_yday.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("BB richness (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+bb_rich_gdd <- gam(log(bb.rich) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
+                    data = filter(abund_div, year == 2010),
+                    method = "REML")
+
+bb_rich_gdd.vis <- getViz(bb_rich_gdd)
+check.gamViz(bb_rich_gdd.vis)
+plot(sm(bb_rich_gdd.vis, 1)) +
+  l_fitRaster() +
+  l_fitContour() +
+  l_points() +
+  ggtitle("BB richness (GDD)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+```
+
+### Rarefied network analysis (not broken down by phase)
 ```{r}
@@
 
-ggplot(net_freq, aes(elev.class2, freq)) +
+ggplot(net_freq, aes(elev.class2, log(freq))) +
   geom_boxplot() +
   facet_wrap(~year)
-
-### Demonstrate that, while community composition may vary across sites, there is not evidence that any species is completely excluded from any point along our elevation gradient.
-bb_elev <- net %>% 
-  group_by(bb.sp, elev.mean) %>%
-  summarize(abund = n())
-
-ggplot(bb_elev, aes(elev.mean, bb.sp, color = log(abund))) +
-  geom_point()
-
-ggplot(bb_elev, aes(elev.mean, abund)) +
-  geom_point() +
-  facet_wrap(~bb.sp, scales = "free_y")
 
@@
 
-spmet <- bind_rows(spmet_2012, spmet_2011, spmet_2010)
+spmet <- bind_rows(spmet_2012, spmet_2011, spmet_2010) %>%
+  filter(metric == "d")
+
+spmet_mean <- spmet %>%
+  group_by(site, year, elev.mean, metric) %>%
+  summarize(value = mean(value))
 
@@
                     netmet_rare_2010) %>%
-  mutate(metric = factor(metric, levels = c("web.asymmetry",
+  mutate(metric = factor(metric, levels = c("H2",
                                             "weighted.NODF",
@@
 ### Plot
-ggplot(spmet_2012, aes(elev.mean, value, color = metric)) +
+#### Per species d'
+ggplot(filter(spmet_2012,
+              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
+       aes(elev.mean, value)) +
   geom_point() +
@@
 
-ggplot(spmet_2011, aes(elev.mean, value, color = metric)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+ggplot(filter(spmet_2011,
+              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
+       aes(elev.mean, value)) +
+  geom_point() +
+  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  geom_smooth() +
   facet_wrap(~sp, scales = "free")
 
-ggplot(spmet_2010, aes(elev.mean, value, color = metric)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+ggplot(filter(spmet_2010,
+              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
+       aes(elev.mean, value)) +
+  geom_point() +
+  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  geom_smooth() +
   facet_wrap(~sp, scales = "free")
 
+#### Community mean d'
+ggplot(spmet_mean, aes(elev.mean, value)) +
+  geom_point() +
+  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  geom_smooth() +
+  facet_wrap(~year)
+
+#### Network level
 ggplot(netmet, aes(elev.mean, value)) +
   geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  geom_smooth() +
   facet_grid(metric~year, scales = "free")
@@
 
-### GAM surface plots of abundance and diversity metrics (just for fun)
-#### 2012
+######## Old analyses that I probably don't care to use but don't want to delete yet #########
+
+### Per-transect network analysis -- is it worth it? I would have to filter out the transects for which the networks are too small.
+#### Network level
 ```{r}
-### Per-capita resource availability
-percap_yday <- gam(log(fl.per.bb) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                   data = filter(abund_div, year == 2012),
-                   method = "REML")
-
-percap_yday.vis <- getViz(percap_yday)
-check.gamViz(percap_yday.vis)
-plot(sm(percap_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Per-capita FL (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-percap_gdd <- gam(log(fl.per.bb) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                  data = filter(abund_div, year == 2012),
-                  method = "REML")
-
-percap_gdd.vis <- getViz(percap_gdd)
-check.gamViz(percap_gdd.vis)
-plot(sm(percap_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Per-capita FL (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### Web asymmetry
-#Create a version of abund_div in which web.asymmetry has been forced into postive values (to enable transformation)
-abund_div_trans <- abund_div %>%
-  mutate(web.asymmetry = web.asymmetry - min(web.asymmetry - 0.1))
-
-ggplot(abund_div_trans, aes(web.asymmetry)) +
-  geom_histogram()
-
-webass_yday <- gam(log(web.asymmetry) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                   data = filter(abund_div_trans, year == 2012),
-                   method = "REML")
-
-webass_yday.vis <- getViz(webass_yday)
-check.gamViz(webass_yday.vis)
-plot(sm(webass_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Web asymmetry (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-webass_gdd <- gam(log(web.asymmetry) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                  data = filter(abund_div_trans, year == 2012),
-                  method = "REML")
-
-webass_gdd.vis <- getViz(webass_gdd)
-check.gamViz(webass_gdd.vis)
-plot(sm(webass_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Web asymmetry (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### Floral abundance
-fl_abund_yday <- gam(log(fl.abund) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2012),
-                    method = "REML")
-
-fl_abund_yday.vis <- getViz(fl_abund_yday)
-check.gamViz(fl_abund_yday.vis)
-plot(sm(fl_abund_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("FL abundance (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-fl_abund_gdd <- gam(log(fl.abund) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2012),
-                    method = "REML")
-
-fl_abund_gdd.vis <- getViz(fl_abund_gdd)
-check.gamViz(fl_abund_gdd.vis)
-plot(sm(fl_abund_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("FL abundance (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### BB abundance
-bb_abund_yday <- gam(log(bb.abund) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2012),
-                    method = "REML")
-
-bb_abund_yday.vis <- getViz(bb_abund_yday)
-check.gamViz(bb_abund_yday.vis)
-plot(sm(bb_abund_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("BB abundance (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-bb_abund_gdd <- gam(log(bb.abund) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2012),
-                    method = "REML")
-
-bb_abund_gdd.vis <- getViz(bb_abund_gdd)
-check.gamViz(bb_abund_gdd.vis)
-plot(sm(bb_abund_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("BB abundance (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### BB richness
-bb_rich_yday <- gam(log(bb.rich) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2012),
-                    method = "REML")
-
-bb_rich_yday.vis <- getViz(bb_rich_yday)
-check.gamViz(bb_rich_yday.vis)
-plot(sm(bb_rich_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("BB richness (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-bb_rich_gdd <- gam(log(bb.rich) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2012),
-                    method = "REML")
-
-bb_rich_gdd.vis <- getViz(bb_rich_gdd)
-check.gamViz(bb_rich_gdd.vis)
-plot(sm(bb_rich_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("BB richness (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
+per.transect_netmet <- net %>%
+  group_by(site, dayofyear, gdd.cum, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
+  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
+  group_by(site, dayofyear, gdd.cum, year) %>%
+  mutate(size = sum(freq),
+         bb.rich = length(unique(bb.sp)),
+         fl.rich = length(unique(plant.sp)),
+         total.rich = bb.rich + fl.rich) %>%
+  filter(bb.rich > 1 & fl.rich > 1) %>%
+  unite(webID, c(site, dayofyear, gdd.cum, year, size), sep = "_") %>%
+  dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
+  data.frame() %>% # convert to data frame
+  frame2webs() %>% # convert to to bipartite web
+  map(networklevel, weighted = TRUE, level = "higher", index = c("web asymmetry", "weighted NODF", "niche overlap", "C score", "H2")) %>% # We probably care mainly about the BB level of the network
+  data.frame() %>%
+  t() %>%
+  data.frame() %>%
+  rownames_to_column() %>%
+  dplyr::select(id = rowname, everything()) %>%
+  as_tibble() %>%
+  gather(metric, value, -id) %>%
+  separate(id, into = c("site", "yday", "gdd.cum", "year", "size"), sep = "_") %>%
+  mutate(size = as.numeric(size),
+         year = factor(year),
+         gdd.cum = as.numeric(gdd.cum),
+         yday = as.numeric(yday),
+         site = as.factor(site)) %>%
+  left_join(site_data) %>%
+  left_join(abund_div)
+
+### Can we explain network metrics with abundance and diversity data?
+# BB rich
+ggplot(per.transect_netmet, aes(bb.rich, value)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_grid(metric ~ year, scales = "free")
+# FL rich
+ggplot(per.transect_netmet, aes(fl.rich, value)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_grid(metric ~ year, scales = "free")
+# BB abund
+ggplot(per.transect_netmet, aes(log(bb.abund), value)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_grid(metric ~ year, scales = "free")
+# FL abund
+ggplot(per.transect_netmet, aes(log(fl.abund), value)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_grid(metric ~ year, scales = "free")
+# Per-capita FL abund
+ggplot(per.transect_netmet, aes(log(fl.per.bb), value)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_grid(metric ~ year, scales = "free")
+
+### Can we explain network metrics with elevation and/or time?
+ggplot(per.transect_netmet, aes(elev.mean, value, color = log(size))) +
+  geom_point() +
+  geom_smooth() +
+  facet_grid(metric ~ year, scales = "free")
+
+ggplot(per.transect_netmet, aes(log(size), value, color = log(size))) +
+  geom_point() +
+  geom_smooth() +
+  facet_grid(metric ~ year, scales = "free")
+
+ggplot(filter(per.transect_netmet, size > 1), 
+       aes(yday, value)) +
+  geom_point() +
+  geom_smooth() +
+  facet_grid(metric ~ year, scales = "free")
+
+ggplot(filter(per.transect_netmet, size > 10), 
+       aes(gdd.cum, value)) +
+  geom_point() +
+  geom_smooth() +
+  facet_grid(metric ~ year, scales = "free")
+
+ggplot(filter(per.transect_netmet, size > 10), 
+       aes(yday, elev.mean)) +
+  geom_point(aes(color = value)) +
+  geom_smooth() +
+  facet_grid(metric ~ year, scales = "free")
 ```
-#### 2011
+
+#### Species level
 ```{r}
-### Per-capita resource availability
-percap_yday <- gam(log(fl.per.bb) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                   data = filter(abund_div, year == 2011),
-                   method = "REML")
-
-percap_yday.vis <- getViz(percap_yday)
-check.gamViz(percap_yday.vis)
-plot(sm(percap_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Per-capita FL (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-percap_gdd <- gam(log(fl.per.bb) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                  data = filter(abund_div, year == 2011),
-                  method = "REML")
-
-percap_gdd.vis <- getViz(percap_gdd)
-check.gamViz(percap_gdd.vis)
-plot(sm(percap_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Per-capita FL (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### Web asymmetry
-#Create a version of abund_div in which web.asymmetry has been forced into postive values (to enable transformation)
-abund_div_trans <- abund_div %>%
-  mutate(web.asymmetry = web.asymmetry - min(web.asymmetry - 0.1))
-
-ggplot(abund_div_trans, aes(web.asymmetry)) +
-  geom_histogram()
-
-webass_yday <- gam(log(web.asymmetry) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                   data = filter(abund_div_trans, year == 2011),
-                   method = "REML")
-
-webass_yday.vis <- getViz(webass_yday)
-check.gamViz(webass_yday.vis)
-plot(sm(webass_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Web asymmetry (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-webass_gdd <- gam(log(web.asymmetry) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                  data = filter(abund_div_trans, year == 2011),
-                  method = "REML")
-
-webass_gdd.vis <- getViz(webass_gdd)
-check.gamViz(webass_gdd.vis)
-plot(sm(webass_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Web asymmetry (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### Floral abundance
-fl_abund_yday <- gam(log(fl.abund) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2011),
-                    method = "REML")
-
-fl_abund_yday.vis <- getViz(fl_abund_yday)
-check.gamViz(fl_abund_yday.vis)
-plot(sm(fl_abund_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("FL abundance (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-fl_abund_gdd <- gam(log(fl.abund) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2011),
-                    method = "REML")
-
-fl_abund_gdd.vis <- getViz(fl_abund_gdd)
-check.gamViz(fl_abund_gdd.vis)
-plot(sm(fl_abund_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("FL abundance (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### BB abundance
-bb_abund_yday <- gam(log(bb.abund) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2011),
-                    method = "REML")
-
-bb_abund_yday.vis <- getViz(bb_abund_yday)
-check.gamViz(bb_abund_yday.vis)
-plot(sm(bb_abund_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("BB abundance (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-bb_abund_gdd <- gam(log(bb.abund) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2011),
-                    method = "REML")
-
-bb_abund_gdd.vis <- getViz(bb_abund_gdd)
-check.gamViz(bb_abund_gdd.vis)
-plot(sm(bb_abund_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("BB abundance (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### BB richness
-bb_rich_yday <- gam(log(bb.rich) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2011),
-                    method = "REML")
-
-bb_rich_yday.vis <- getViz(bb_rich_yday)
-check.gamViz(bb_rich_yday.vis)
-plot(sm(bb_rich_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("BB richness (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-bb_rich_gdd <- gam(log(bb.rich) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2011),
-                    method = "REML")
-
-bb_rich_gdd.vis <- getViz(bb_rich_gdd)
-check.gamViz(bb_rich_gdd.vis)
-plot(sm(bb_rich_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("BB richness (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
+per.transect_spmet <- net %>%
+  group_by(site, dayofyear, gdd.cum, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
+  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
+  group_by(site, dayofyear, gdd.cum, year) %>%
+  mutate(size = sum(freq),
+         bb.rich = length(unique(bb.sp)),
+         fl.rich = length(unique(plant.sp)),
+         total.rich = bb.rich + fl.rich) %>%
+  filter(bb.rich > 1 & fl.rich > 1) %>%
+  unite(webID, c(site, dayofyear, gdd.cum, year, size), sep = "_") %>%
+  dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
+  data.frame() %>% # convert to data frame
+  frame2webs() %>% # convert to to bipartite web
+  map(specieslevel, level = "higher", index = "d") %>% # We probably care mainly about the BB level of the network
+  map(rownames_to_column) %>%
+  map(as_tibble) %>%
+  bind_rows(.id = "id") %>% # Collapse list into big data frame
+  rename(sp = rowname) %>%
+  gather(metric, value, -c(sp, id)) %>% # get mean value across iterations for each metric
+  separate(id, sep = "_", into = c("site", "yday", "gdd.cum", "year", "size")) %>%
+  mutate(size = as.numeric(size),
+         year = factor(year),
+         gdd.cum = as.numeric(gdd.cum),
+         yday = as.numeric(yday)) %>%
+  left_join(site_data) %>%
+  left_join(bb_traits, by = c("sp" = "bb.sp")) %>%
+  left_join(abund_div)
+
+per.transect_spmet.mean <- per.transect_spmet %>%
+  group_by(site, yday, gdd.cum, year) %>%
+  summarize(value = mean(value)) %>%
+  left_join(site_data)
+
+ggplot(filter(per.transect_spmet), 
+       aes(log(fl.abund), value)) +
+  geom_point() +
+  geom_smooth() +
+  facet_grid(. ~ year, scales = "free")
+
+ggplot(filter(per.transect_spmet), 
+       aes(elev.mean, size)) +
+  geom_point() +
+  geom_smooth() +
+  facet_grid(. ~ year, scales = "free")
+
+ggplot(filter(per.transect_spmet.mean), 
+       aes(yday, value)) +
+  geom_point() +
+  geom_smooth() +
+  facet_grid(. ~ year, scales = "free")
+
+ggplot(filter(per.transect_spmet), 
+       aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth() +
+  facet_grid(. ~ year, scales = "free")
+
+ggplot(filter(per.transect_spmet, size >= 10), 
+       aes(gdd.cum, value, color = log(size))) +
+  geom_point() +
+  geom_smooth() +
+  facet_grid(metric ~ year, scales = "free")
+
+ggplot(filter(per.transect_spmet, size >= 10), 
+       aes(yday, value, color = log(size))) +
+  geom_point() +
+  geom_smooth() +
+  facet_grid(metric ~ year, scales = "free")
 ```
-#### 2010
+
+### Phase x elev network analysis
 ```{r}
-### Per-capita resource availability
-percap_yday <- gam(log(fl.per.bb) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                   data = filter(abund_div, year == 2010),
-                   method = "REML")
-
-percap_yday.vis <- getViz(percap_yday)
-check.gamViz(percap_yday.vis)
-plot(sm(percap_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Per-capita FL (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-percap_gdd <- gam(log(fl.per.bb) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                  data = filter(abund_div, year == 2010),
-                  method = "REML")
-
-percap_gdd.vis <- getViz(percap_gdd)
-check.gamViz(percap_gdd.vis)
-plot(sm(percap_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Per-capita FL (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### Web asymmetry
-#Create a version of abund_div in which web.asymmetry has been forced into postive values (to enable transformation)
-abund_div_trans <- abund_div %>%
-  mutate(web.asymmetry = web.asymmetry - min(web.asymmetry - 0.1))
-
-ggplot(abund_div_trans, aes(web.asymmetry)) +
-  geom_histogram()
-
-webass_yday <- gam(log(web.asymmetry) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                   data = filter(abund_div_trans, year == 2010),
-                   method = "REML")
-
-webass_yday.vis <- getViz(webass_yday)
-check.gamViz(webass_yday.vis)
-plot(sm(webass_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Web asymmetry (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-webass_gdd <- gam(log(web.asymmetry) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                  data = filter(abund_div_trans, year == 2010),
-                  method = "REML")
-
-webass_gdd.vis <- getViz(webass_gdd)
-check.gamViz(webass_gdd.vis)
-plot(sm(webass_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Web asymmetry (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### Floral abundance
-fl_abund_yday <- gam(log(fl.abund) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2010),
-                    method = "REML")
-
-fl_abund_yday.vis <- getViz(fl_abund_yday)
-check.gamViz(fl_abund_yday.vis)
-plot(sm(fl_abund_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("FL abundance (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-fl_abund_gdd <- gam(log(fl.abund) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2010),
-                    method = "REML")
-
-fl_abund_gdd.vis <- getViz(fl_abund_gdd)
-check.gamViz(fl_abund_gdd.vis)
-plot(sm(fl_abund_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("FL abundance (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### BB abundance
-bb_abund_yday <- gam(log(bb.abund) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2010),
-                    method = "REML")
-
-bb_abund_yday.vis <- getViz(bb_abund_yday)
-check.gamViz(bb_abund_yday.vis)
-plot(sm(bb_abund_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("BB abundance (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-bb_abund_gdd <- gam(log(bb.abund) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2010),
-                    method = "REML")
-
-bb_abund_gdd.vis <- getViz(bb_abund_gdd)
-check.gamViz(bb_abund_gdd.vis)
-plot(sm(bb_abund_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("BB abundance (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### BB richness
-bb_rich_yday <- gam(log(bb.rich) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2010),
-                    method = "REML")
-
-bb_rich_yday.vis <- getViz(bb_rich_yday)
-check.gamViz(bb_rich_yday.vis)
-plot(sm(bb_rich_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("BB richness (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-bb_rich_gdd <- gam(log(bb.rich) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2010),
-                    method = "REML")
-
-bb_rich_gdd.vis <- getViz(bb_rich_gdd)
-check.gamViz(bb_rich_gdd.vis)
-plot(sm(bb_rich_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("BB richness (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
+### Prepare pxe abundance and diversity table
+pxe_abund_div <- abund_div %>%
+  select(-gdd.cum, -yday, -week, -month, -date, -elev.class2) %>%
+  gather(ad.metric, ad.value, -c(site, year, phase, elev.mean)) %>%
+  group_by(site, phase, year, ad.metric) %>%
+  summarize(ad.value = mean(ad.value, na.rm = TRUE))
+
+pxe_net <- rare_net_nt2(net, 4, 20) %>%
+  full_join((pxe_abund_div))
+
+ggplot(filter(pxe_net, 
+              metric == "C.score.HL" & ad.metric == "fl.abund"),
+       aes(log(ad.value), value)) +
+  geom_point() +
+  geom_smooth(method = "lm")  +
+  facet_grid(year ~ phase, scales = "free") +
+  xlab("log(floral abundance)") +
+  ylab("C.score")
+
+ggplot(filter(pxe_net, 
+              metric == "C.score.HL" & ad.metric == "fl.per.bb"),
+       aes(log(ad.value), value)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_grid(year ~ phase, scales = "free") +
+  xlab("log(per-capita floral abundance)") +
+  ylab("C.score")
+
+ggplot(filter(pxe_net, 
+              metric == "H2" & ad.metric == "fl.rich"),
+       aes(ad.value, value)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_grid(year ~ phase, scales = "free") +
+  xlab("Floral richness") +
+  ylab("H2")
+
+ggplot(filter(pxe_net, 
+              metric == "H2" & ad.metric == "bb.abund"),
+       aes(log(ad.value), value)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_grid(year ~ phase, scales = "free") +
+  xlab("BB abund") +
+  ylab("H2")
+
+ggplot(filter(pxe_net, 
+              metric == "H2" & ad.metric == "bb.rich"),
+       aes(ad.value, value)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_grid(year ~ phase, scales = "free") +
+  xlab("BB richness") +
+  ylab("H2")
+
+ggplot(filter(pxe_net, metric == "weighted.NODF"), aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_grid(year ~ phase, scales = "free") +
+  xlab("Elevation") +
+  ylab("Weighted NODF")
+
+ggplot(filter(pxe_net, metric == "H2"), aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_grid(year ~ phase, scales = "free") +
+  xlab("Elevation") +
+  ylab("H2")
+
+ggplot(filter(pxe_net, metric == "C.score.HL"), aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_grid(year ~ phase, scales = "free") +
+  xlab("Elevation") +
+  ylab("C.score")
+
+ggplot(filter(pxe_net, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_grid(year ~ phase, scales = "free") +
+  xlab("Elevation") +
+  ylab("Niche overlap")
+
+ggplot(filter(pxe_net, metric == "generality.HL"), aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_grid(year ~ phase, scales = "free") +
+  xlab("Elevation") +
+  ylab("Generality")
+
+ggplot(filter(pxe_net, metric == "web.asymmetry"), aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_grid(year ~ phase, scales = "free") +
+  xlab("Elevation") +
+  ylab("Web asymmetry")
+
+pxe_sp <- rare_net_sp2(net, 4, 20)  %>%
+  full_join((pxe_abund_div))
+
+ggplot(filter(pxe_sp, 
+              metric == "d" &
+                sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit") &
+                ad.metric == "fl.rich"), 
+       aes(ad.value, value, color = year)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE) +
+  facet_grid(sp ~ phase, scales = "free") +
+  ylim(c(0,1)) +
+  xlab("Floral richness") +
+  ylab("d'")
+
+pxe_sp_mean <- pxe_sp %>%
+  filter(sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")) %>%
+  group_by(site, year, phase, elev.mean, metric) %>%
+  summarize(value = mean(value, na.rm = TRUE))
+
+ggplot(filter(pxe_sp_mean, metric == "d"), aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_grid(year ~ phase, scales = "free") +
+  xlab("Elevation") +
+  ylab("Mean d'")
+
+ggplot(filter(pxe_sp_mean, metric == "species.strength"), aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_grid(year ~ phase, scales = "free") +
+  xlab("Elevation") +
+  ylab("Mean species strength'")
+
+ggplot(filter(pxe_sp, 
+              metric == "species.strength",
+              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
+       aes(elev.mean, value, color = year)) +
+  geom_point() +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_grid(sp ~ phase, scales = "free") +
+  xlab("Elevation") +
+  ylab("Species strength")
+
+ggplot(filter(pxe_sp, 
+              metric == "d",
+              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
+       aes(elev.mean, value, color = year)) +
+  geom_point() +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_grid(sp ~ phase, scales = "free") +
+  ylim(c(0,1)) +
+  xlab("Elevation") +
+  ylab("d'")
+
+ggplot(filter(pxe_sp, 
+              metric == "d",
+              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
+       aes(elev.mean, value)) +
+  geom_point(aes(color = year)) +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_grid(sp ~ phase, scales = "free") +
+  ylim(c(0,1)) +
+  xlab("Elevation") +
+  ylab("d'")
+
+ggplot(filter(pxe_sp, 
+              metric == "d",
+              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
+       aes(elev.mean, value, color = phase)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_grid(year ~ sp, scales = "free") +
+  theme_light() +
+  xlab("Elevation") +
+  ylab("d'")
 ```

### 2020-06-03 data_proc.R

@@ -90,7 +90,7 @@ climate <- read_delim("./raw_data/climate2.txt", delim = "\t") %>%
     pred_Tmean_day - 10 <= 0 ~ 0)
     ) %>%
   group_by(site, year) %>%
-  mutate(gdd.cum = cumsum(gdd))
+  mutate(gdd.cum = round(cumsum(gdd)))
 
 write_csv(climate, "./processed_data/climate.csv")
 

### 2020-06-04 analysis.Rmd
added GAM modeling for network metrics ~ elevation, fixed a bug in data proc that left a bunch of NAs from the raw data

@@
       frame2webs() %>% # convert to to bipartite web
-      map(specieslevel, level = "higher", index = c("d", "species specificity")) %>%
+      map(specieslevel, level = "higher", index = c("d", "species strength", "proportional generality")) %>%
       map(rownames_to_column) %>%
@@
       frame2webs() %>% # convert to to bipartite web
-      map(networklevel, weighted = TRUE, level = "higher", index = c("H2", "weighted NODF", "niche overlap", "C score", "generality.HL")) %>% # We probably care mainly about the BB level of the network
+      map(networklevel, weighted = TRUE, level = "higher", index = c("H2", "weighted NODF", "niche overlap", "C score", "generality", "web asymmetry", "weighted connectance")) %>% # We probably care mainly about the BB level of the network
       data.frame() %>%
@@
            elev.mean, break1.gdd, break2.gdd) %>%
-  summarize(fl.abund = sum(flower.cover),
-            log.fl.abund = log(sum(flower.cover)))
+  summarize(fl.abund = sum(flower.cover))
 
@@
 ### BB abundance ~ floral abundance
-ggplot(abund_div, aes(log.fl.abund, bb.abund, color = year)) +
+ggplot(abund_div, aes(log(fl.abund), bb.abund, color = year)) +
   geom_point(alpha = 0.5) +
@@
 ### BB richness ~ floral abundance
-ggplot(abund_div, aes(log.fl.abund, bb.rich, color = year)) +
+ggplot(abund_div, aes(log(fl.abund), bb.rich, color = year)) +
   geom_point(alpha = 0.5) +
@@
 ### Floral richness ~ floral abundance
-ggplot(abund_div, aes(log.fl.abund, fl.rich, color = year)) +
+ggplot(abund_div, aes(log(fl.abund), fl.rich, color = year)) +
   geom_point(alpha = 0.5) +
@@
 summary(fl_abund_gdd.vis)
-plot(sm(fl_abund_gdd.vis, 1)) +
+
+plot(fl_abund_gdd.vis, allTerms = TRUE) +
+  ggtitle("FL abundance (gdd)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+```
+
+##### Floral richness
+```{r}
+fl_rich_yday <- gam(fl.rich ~ year +
+                    te(yday, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
+                    data = filter(abund_div),
+                    method = "REML")
+
+fl_rich_yday.vis <- getViz(fl_rich_yday)
+check.gamViz(fl_rich_yday.vis)
+summary(fl_rich_yday.vis)
+plot(sm(fl_rich_yday.vis, 1)) +
   l_fitRaster() +
@@
   l_points() +
-  ggtitle("FL abundance 2010 (gdd)") +
+  ggtitle("FL richness 2010 (yday)") +
   xlab("Day of year") +
   ylab("Elevation")
-plot(sm(fl_abund_gdd.vis, 2)) +
+plot(sm(fl_rich_yday.vis, 2)) +
   l_fitRaster() +
@@
   l_points() +
-  ggtitle("FL abundance 2011 (gdd)") +
+  ggtitle("FL richness 2011 (yday)") +
   xlab("Day of year") +
   ylab("Elevation")
-plot(sm(fl_abund_gdd.vis, 3)) +
+plot(sm(fl_rich_yday.vis, 3)) +
   l_fitRaster() +
@@
   l_points() +
-  ggtitle("FL abundance 2012 (gdd)") +
+  ggtitle("FL richness 2012 (yday)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+fl_rich_gdd <- gam(fl.rich ~ year +
+                    te(gdd.cum, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
+                    data = filter(abund_div),
+                    method = "REML")
+
+fl_rich_gdd.vis <- getViz(fl_rich_gdd)
+check.gamViz(fl_rich_gdd.vis)
+summary(fl_rich_gdd.vis)
+
+plot(fl_rich_gdd.vis, allTerms = TRUE) +
+  ggtitle("FL richness (gdd)") +
   xlab("Day of year") +
@@
                     data = filter(abund_div),
-                    family = "scat",
                     method = "REML")
@@
 summary(bb_abund_gdd.vis)
-plot(sm(bb_abund_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("bb abundance 2010 (gdd)") +
+
+plot(bb_abund_gdd.vis, allTerms = TRUE) +
+  ggtitle("BB abundance (gdd)") +
   xlab("Day of year") +
   ylab("Elevation")
-plot(sm(bb_abund_gdd.vis, 2)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("bb abundance 2011 (gdd)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-plot(sm(bb_abund_gdd.vis, 3)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("bb abundance 2012 (gdd)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-```
-
-##### Per-capita floral resource abundance
+```
+
+##### BB richness
 ```{r}
-percap_yday <- gam(fl.per.bb ~ year +
+bb_rich_yday <- gam(bb.rich ~ year +
                     te(yday, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
-                    data = abund_div,
-                    family = "tw",
-                    method = "REML")
-
-percap_yday.vis <- getViz(percap_yday)
-check.gamViz(percap_yday.vis)
-summary(percap_yday.vis)
-plot(sm(percap_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Per-capita floral abundance 2010 (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-plot(sm(percap_yday.vis, 2)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Per-capita floral abundance 2011 (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-plot(sm(percap_yday.vis, 3)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Per-capita floral abundance 2012 (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-percap_gdd <- gam(fl.per.bb ~ year +
-                    te(gdd.cum, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
-                    data = abund_div,
-                    family = "tw",
-                    method = "REML")
-
-percap_gdd.vis <- getViz(percap_gdd)
-check.gamViz(percap_gdd.vis)
-summary(percap_gdd.vis)
-plot(sm(percap_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Per-capita floral abundance 2010 (gdd)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-plot(sm(percap_gdd.vis, 2)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Per-capita floral abundance 2011 (gdd)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-plot(sm(percap_gdd.vis, 3)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Per-capita floral abundance 2012 (gdd)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-```
-
-##### Web asymmetry
-```{r}
-webass_yday <- gam(web.asymmetry ~ year +
-                    te(yday, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
-                    data = abund_div,
-                    family = "tw",
-                    method = "REML")
-
-percap_yday.vis <- getViz(percap_yday)
-check.gamViz(percap_yday.vis)
-summary(percap_yday.vis)
-plot(sm(percap_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Per-capita floral abundance 2010 (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-plot(sm(percap_yday.vis, 2)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Per-capita floral abundance 2011 (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-plot(sm(percap_yday.vis, 3)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Per-capita floral abundance 2012 (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-percap_gdd <- gam(fl.per.bb ~ year +
-                    te(gdd.cum, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
-                    data = abund_div,
-                    family = "tw",
-                    method = "REML")
-
-percap_gdd.vis <- getViz(percap_gdd)
-check.gamViz(percap_gdd.vis)
-summary(percap_gdd.vis)
-plot(sm(percap_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Per-capita floral abundance 2010 (gdd)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-plot(sm(percap_gdd.vis, 2)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Per-capita floral abundance 2011 (gdd)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-plot(sm(percap_gdd.vis, 3)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Per-capita floral abundance 2012 (gdd)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-```
-
-#### 2012
-```{r}
-### Per-capita resource availability
-percap_yday <- gam(log(fl.per.bb) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                   data = filter(abund_div, year == 2012),
-                   method = "REML")
-
-percap_yday.vis <- getViz(percap_yday)
-check.gamViz(percap_yday.vis)
-summary(percap_yday.vis)
-plot(sm(percap_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Per-capita FL (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-percap_gdd <- gam(log(fl.per.bb) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                  data = filter(abund_div, year == 2012),
-                  method = "REML")
-
-percap_gdd.vis <- getViz(percap_gdd)
-check.gamViz(percap_gdd.vis)
-plot(sm(percap_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Per-capita FL (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### Web asymmetry
-#Create a version of abund_div in which web.asymmetry has been forced into postive values (to enable transformation)
-abund_div_trans <- abund_div %>%
-  mutate(web.asymmetry = web.asymmetry - min(web.asymmetry - 0.1, na.rm = TRUE))
-
-ggplot(abund_div_trans, aes(web.asymmetry)) +
-  geom_histogram()
-
-webass_yday <- gam(log(web.asymmetry) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                   data = filter(abund_div_trans, year == 2012),
-                   method = "REML")
-
-webass_yday.vis <- getViz(webass_yday)
-check.gamViz(webass_yday.vis)
-plot(sm(webass_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Web asymmetry (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-webass_gdd <- gam(log(web.asymmetry) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                  data = filter(abund_div_trans, year == 2012),
-                  method = "REML")
-
-webass_gdd.vis <- getViz(webass_gdd)
-check.gamViz(webass_gdd.vis)
-plot(sm(webass_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Web asymmetry (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### Floral abundance
-fl_abund_yday <- gam(log(fl.abund) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2012),
-                    method = "REML")
-
-fl_abund_yday.vis <- getViz(fl_abund_yday)
-check.gamViz(fl_abund_yday.vis)
-plot(sm(fl_abund_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("FL abundance (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-fl_abund_gdd <- gam(fl.abund ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2012 & !is.na(fl.abund)),
-                    family = "tw",
-                    method = "REML")
-
-fl_abund_gdd.vis <- getViz(fl_abund_gdd)
-check.gamViz(fl_abund_gdd.vis)
-summary(fl_abund_gdd)
-plot(sm(fl_abund_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("FL abundance (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### BB abundance
-bb_abund_yday <- gam(bb.abund ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2012),
-                    family = "tw",
-                    method = "REML")
-
-bb_abund_yday.vis <- getViz(bb_abund_yday)
-check.gamViz(bb_abund_yday.vis)
-plot(sm(bb_abund_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("BB abundance (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-bb_abund_gdd <- gam(log(bb.abund) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2012),
-                    method = "REML")
-
-bb_abund_gdd.vis <- getViz(bb_abund_gdd)
-check.gamViz(bb_abund_gdd.vis)
-plot(sm(bb_abund_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("BB abundance (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### BB richness
-bb_rich_yday <- gam(log(bb.rich) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2012),
+                    data = filter(abund_div),
                     method = "REML")
@@
 check.gamViz(bb_rich_yday.vis)
+summary(bb_rich_yday.vis)
 plot(sm(bb_rich_yday.vis, 1)) +
@@
   l_points() +
-  ggtitle("BB richness (yday)") +
+  ggtitle("bb richness 2010 (yday)") +
   xlab("Day of year") +
   ylab("Elevation")
-
-bb_rich_gdd <- gam(log(bb.rich) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2012),
-                    method = "REML")
-
-bb_rich_gdd.vis <- getViz(bb_rich_gdd)
-check.gamViz(bb_rich_gdd.vis)
-plot(sm(bb_rich_gdd.vis, 1)) +
+plot(sm(bb_rich_yday.vis, 2)) +
   l_fitRaster() +
@@
   l_points() +
-  ggtitle("BB richness (GDD)") +
-  xlab("GDD") +
+  ggtitle("bb richness 2011 (yday)") +
+  xlab("Day of year") +
   ylab("Elevation")
-```
-#### 2011
-```{r}
-### Per-capita resource availability
-percap_yday <- gam(log(fl.per.bb) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                   data = filter(abund_div, year == 2011),
-                   method = "REML")
-
-percap_yday.vis <- getViz(percap_yday)
-check.gamViz(percap_yday.vis)
-plot(sm(percap_yday.vis, 1)) +
+plot(sm(bb_rich_yday.vis, 3)) +
   l_fitRaster() +
@@
   l_points() +
-  ggtitle("Per-capita FL (yday)") +
+  ggtitle("bb richness 2012 (yday)") +
   xlab("Day of year") +
@@
 
-percap_gdd <- gam(log(fl.per.bb) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                  data = filter(abund_div, year == 2011),
-                  method = "REML")
-
-percap_gdd.vis <- getViz(percap_gdd)
-check.gamViz(percap_gdd.vis)
-plot(sm(percap_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Per-capita FL (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### Web asymmetry
-#Create a version of abund_div in which web.asymmetry has been forced into postive values (to enable transformation)
-abund_div_trans <- abund_div %>%
-  mutate(web.asymmetry = web.asymmetry - min(web.asymmetry - 0.1, na.rm = TRUE))
-
-ggplot(abund_div_trans, aes(web.asymmetry)) +
-  geom_histogram()
-
-webass_yday <- gam(log(web.asymmetry) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                   data = filter(abund_div_trans, year == 2011),
-                   method = "REML")
-
-webass_yday.vis <- getViz(webass_yday)
-check.gamViz(webass_yday.vis)
-plot(sm(webass_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Web asymmetry (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-webass_gdd <- gam(log(web.asymmetry) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                  data = filter(abund_div_trans, year == 2011),
-                  method = "REML")
-
-webass_gdd.vis <- getViz(webass_gdd)
-check.gamViz(webass_gdd.vis)
-plot(sm(webass_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Web asymmetry (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### Floral abundance
-fl_abund_yday <- gam(log(fl.abund) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2011),
-                    method = "REML")
-
-fl_abund_yday.vis <- getViz(fl_abund_yday)
-check.gamViz(fl_abund_yday.vis)
-plot(sm(fl_abund_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("FL abundance (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-fl_abund_gdd <- gam(log(fl.abund) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2011),
-                    method = "REML")
-
-fl_abund_gdd.vis <- getViz(fl_abund_gdd)
-check.gamViz(fl_abund_gdd.vis)
-plot(sm(fl_abund_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("FL abundance (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### BB abundance
-bb_abund_yday <- gam(log(bb.abund) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2011),
-                    method = "REML")
-
-bb_abund_yday.vis <- getViz(bb_abund_yday)
-check.gamViz(bb_abund_yday.vis)
-plot(sm(bb_abund_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("BB abundance (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-bb_abund_gdd <- gam(log(bb.abund) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2011),
-                    method = "REML")
-
-bb_abund_gdd.vis <- getViz(bb_abund_gdd)
-check.gamViz(bb_abund_gdd.vis)
-plot(sm(bb_abund_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("BB abundance (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### BB richness
-bb_rich_yday <- gam(log(bb.rich) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2011),
-                    method = "REML")
-
-bb_rich_yday.vis <- getViz(bb_rich_yday)
-check.gamViz(bb_rich_yday.vis)
-plot(sm(bb_rich_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("BB richness (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-bb_rich_gdd <- gam(log(bb.rich) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2011),
+bb_rich_gdd <- gam(bb.rich ~ year +
+                    te(gdd.cum, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
+                    data = filter(abund_div),
                     method = "REML")
@@
 check.gamViz(bb_rich_gdd.vis)
-plot(sm(bb_rich_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("BB richness (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-```
-#### 2010
-```{r}
-### Per-capita resource availability
-percap_yday <- gam(log(fl.per.bb) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                   data = filter(abund_div, year == 2010),
-                   method = "REML")
-
-percap_yday.vis <- getViz(percap_yday)
-check.gamViz(percap_yday.vis)
-plot(sm(percap_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Per-capita FL (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-percap_gdd <- gam(log(fl.per.bb) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                  data = filter(abund_div, year == 2010),
-                  method = "REML")
-
-percap_gdd.vis <- getViz(percap_gdd)
-check.gamViz(percap_gdd.vis)
-plot(sm(percap_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Per-capita FL (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### Web asymmetry
-#Create a version of abund_div in which web.asymmetry has been forced into postive values (to enable transformation)
-abund_div_trans <- abund_div %>%
-  mutate(web.asymmetry = web.asymmetry - min(web.asymmetry - 0.1, na.rm = TRUE))
-
-ggplot(abund_div_trans, aes(web.asymmetry)) +
-  geom_histogram()
-
-webass_yday <- gam(log(web.asymmetry) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                   data = filter(abund_div_trans, year == 2010),
-                   method = "REML")
-
-webass_yday.vis <- getViz(webass_yday)
-check.gamViz(webass_yday.vis)
-plot(sm(webass_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Web asymmetry (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-webass_gdd <- gam(log(web.asymmetry) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                  data = filter(abund_div_trans, year == 2010),
-                  method = "REML")
-
-webass_gdd.vis <- getViz(webass_gdd)
-check.gamViz(webass_gdd.vis)
-plot(sm(webass_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("Web asymmetry (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### Floral abundance
-fl_abund_yday <- gam(log(fl.abund) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2010),
-                    method = "REML")
-
-fl_abund_yday.vis <- getViz(fl_abund_yday)
-check.gamViz(fl_abund_yday.vis)
-plot(sm(fl_abund_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("FL abundance (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-fl_abund_gdd <- gam(log(fl.abund) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2010),
-                    method = "REML")
-
-fl_abund_gdd.vis <- getViz(fl_abund_gdd)
-check.gamViz(fl_abund_gdd.vis)
-plot(sm(fl_abund_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("FL abundance (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### BB abundance
-bb_abund_yday <- gam(log(bb.abund) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2010),
-                    method = "REML")
-
-bb_abund_yday.vis <- getViz(bb_abund_yday)
-check.gamViz(bb_abund_yday.vis)
-plot(sm(bb_abund_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("BB abundance (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-bb_abund_gdd <- gam(log(bb.abund) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2010),
-                    method = "REML")
-
-bb_abund_gdd.vis <- getViz(bb_abund_gdd)
-check.gamViz(bb_abund_gdd.vis)
-plot(sm(bb_abund_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("BB abundance (GDD)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### BB richness
-bb_rich_yday <- gam(log(bb.rich) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2010),
-                    method = "REML")
-
-bb_rich_yday.vis <- getViz(bb_rich_yday)
-check.gamViz(bb_rich_yday.vis)
-plot(sm(bb_rich_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("BB richness (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-bb_rich_gdd <- gam(log(bb.rich) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
-                    data = filter(abund_div, year == 2010),
-                    method = "REML")
-
-bb_rich_gdd.vis <- getViz(bb_rich_gdd)
-check.gamViz(bb_rich_gdd.vis)
-plot(sm(bb_rich_gdd.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("BB richness (GDD)") +
+summary(bb_rich_gdd.vis)
+
+plot(bb_rich_gdd.vis, allTerms = TRUE) +
+  ggtitle("BB richness (gdd)") +
   xlab("Day of year") +
@@
 
-spmet <- bind_rows(spmet_2012, spmet_2011, spmet_2010) %>%
-  filter(metric == "d")
+spmet <- bind_rows(spmet_2012, spmet_2011, spmet_2010)
 
@@
                                             "niche.overlap.HL", 
-                                            "C.score.HL")))
+                                            "C.score.HL",
+                                            "generality.HL",
+                                            "web.asymmetry",
+                                            "weighted.connectance")))
 
 ### Plot
-#### Per species d'
+#### Per species values
 ggplot(filter(spmet_2012,
@@
   geom_smooth() +
-  facet_wrap(~sp, scales = "free")
+  facet_grid(metric~sp, scales = "free")
 
@@
   geom_smooth() +
-  facet_wrap(~sp, scales = "free")
-
-#### Community mean d'
-ggplot(spmet_mean, aes(elev.mean, value)) +
-  geom_point() +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth() +
-  facet_wrap(~year)
+  facet_grid(metric~sp, scales = "free")
+
+#### Community mean d
+ggplot(filter(spmet_mean, metric == "d"), aes(elev.mean, value)) +
+  geom_point(aes(color = year)) +
+  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
+  xlab("Elevation") +
+  ylab("Community mean d'") +
+  theme_light(16)
+
+gam_d <- gam(value ~ year + s(elev.mean, k = 10), 
+             data = filter(spmet, metric == "d"),
+             method = "REML")
+
+gam_d.vis <- getViz(gam_d)
+check.gamViz(gam_d.vis)
+summary(gam_d.vis)
+
+plot(gam_d.vis, allTerms = TRUE) +
+  xlab("Elevation")
 
@@
   facet_grid(metric~year, scales = "free")
+
+##### H2
+ggplot(filter(netmet, metric == "H2"), aes(elev.mean, value)) +
+  geom_point(aes(color = year)) +
+  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
+  xlab("Elevation") +
+  ylab("H2'") +
+  theme_light(16)
+
+gam_H2 <- gam(value ~ year + s(elev.mean), 
+              data = filter(netmet, metric == "H2"),
+              method = "REML")
+
+gam_H2.vis <- getViz(gam_H2)
+check.gamViz(gam_H2.vis)
+summary(gam_H2.vis)
+
+plot(gam_H2.vis, allTerms = TRUE) +
+  xlab("Elevation")
+
+##### niche overlap
+ggplot(filter(netmet, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
+  geom_point(aes(color = year)) +
+  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
+  #geom_smooth() +
+  xlab("Elevation") +
+  ylab("Niche overlap") +
+  theme_light(16)
+
+gam_niche.overlap <- gam(value ~ year + s(elev.mean), 
+                         data = filter(netmet, metric == "niche.overlap.HL"),
+                         method = "REML")
+
+gam_niche.overlap.vis <- getViz(gam_niche.overlap)
+check.gamViz(gam_niche.overlap.vis)
+summary(gam_niche.overlap.vis)
+
+plot(gam_niche.overlap.vis, allTerms = TRUE) +
+  xlab("Elevation")
+
+##### C.score
+ggplot(filter(netmet, metric == "C.score.HL"), aes(elev.mean, value)) +
+  geom_point(aes(color = year)) +
+  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
+  #geom_smooth() +
+  xlab("Elevation") +
+  ylab("C score") +
+  theme_light(16)
+
+gam_cscore <- gam(value ~ year + s(elev.mean), 
+                  data = filter(netmet, metric == "C.score.HL"),
+                  method = "REML")
+
+gam_cscore.vis <- getViz(gam_cscore)
+check.gamViz(gam_cscore.vis)
+summary(gam_cscore.vis)
+
+plot(gam_cscore.vis, allTerms = TRUE) +
+  xlab("Elevation")
+
+##### Connectance
+ggplot(filter(netmet, metric == "weighted.connectance"), aes(elev.mean, value)) +
+  geom_point(aes(color = year)) +
+  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
+  #geom_smooth() +
+  xlab("Elevation") +
+  ylab("Connectance (weighted)") +
+  theme_light(16)
+
+gam_connectance <- gam(value ~ year + s(elev.mean), 
+                  data = filter(netmet, metric == "weighted.connectance"),
+                  method = "REML")
+
+gam_connectance.vis <- getViz(gam_connectance)
+check.gamViz(gam_connectance.vis)
+summary(gam_connectance.vis)
+
+plot(gam_connectance.vis, allTerms = TRUE) +
+  xlab("Elevation")
+
+##### NODF
+ggplot(filter(netmet, metric == "weighted.NODF"), aes(elev.mean, value)) +
+  geom_point(aes(color = year)) +
+  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
+  #geom_smooth() +
+  xlab("Elevation") +
+  ylab("NODF (weighted)") +
+  theme_light(16)
+
+gam_NODF <- gam(value ~ year + s(elev.mean), 
+                data = filter(netmet, metric == "weighted.NODF"),
+                method = "REML")
+
+gam_NODF.vis <- getViz(gam_NODF)
+check.gamViz(gam_NODF.vis)
+summary(gam_NODF.vis)
+
+plot(gam_NODF.vis, allTerms = TRUE) +
+  xlab("Elevation")
+
+##### Web asymmetry
+ggplot(filter(netmet, metric == "web.asymmetry"), aes(elev.mean, value)) +
+  geom_point(aes(color = year)) +
+  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
+  #geom_smooth() +
+  xlab("Elevation") +
+  ylab("Web asymmetry") +
+  theme_light(16)
+
+gam_webass <- gam(value ~ year + s(elev.mean), 
+                data = filter(netmet, metric == "web.asymmetry"),
+                method = "REML")
+
+gam_webass.vis <- getViz(gam_webass)
+check.gamViz(gam_webass.vis)
+summary(gam_webass.vis)
+
+plot(gam_webass.vis, allTerms = TRUE) +
+  xlab("Elevation")
 ```

### 2020-06-04 data_proc.R

@@ -5,6 +5,7 @@ library(lubridate)
 ### Floral survey data
 fl_survey <- read_delim("./raw_data/data_plantcover_forDoug_2020.01.17.csv", delim = "\t", 
                            locale = locale(decimal_mark = ",", grouping_mark = ".")) %>%
+  na.omit() %>%
   separate(col = species_plant, into = c("genus", "species", "w", "x", "y", "z"), sep = " ") %>%
   unite(species_plant, genus, species, sep = " ") %>%
   select(-c(w,x,y,z)) %>%

### 2020-06-05 analysis.Rmd
updating lab computer

@@
 ```{r setup, include=FALSE}
+knitr::opts_chunk$set(include = FALSE)
 library(bipartite)
@@
 #### Colony phase breakpoint functions
-```{r}
+```{r include=FALSE}
 # Find breakpoints on GDD scale
@@
 #### Rarefied network analysis
-```{r}
+```{r, include=FALSE}
 # Rare_net function: rarefies networks by iteratively subsampling more frequently sampled networks to match the transect count of the least frequently sampled one
@@
 ### Load site data
-```{r}
+```{r, include=FALSE}
 site_data <- read_csv("./processed_data/site_data.csv") %>%
@@
 ### Climate data
-```{r}
+```{r, include=FALSE}
 gdd <- read_csv("./processed_data/climate.csv") %>%
@@
 ### Colony phase breakpoint analysis
-```{r}
+```{r, include=FALSE}
 castes <- read_csv("./processed_data/network.csv") %>%
@@
 ### Plot phase breakpoints
-```{r}
+```{r, include=FALSE}
 breaks_summary <- full_join(breaks_gdd_summary, breaks_yday_summary) %>%
@@
 ### Load network and survey data
-```{r}
+```{r, include=FALSE}
 bb_traits <- read_csv("./processed_data/bb_traits.csv") %>%
@@
 ### BB and floral turnover ~ elevation
-```{r}
+```{r, include=FALSE}
 bb_range <- net %>%
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+### Plot only the plants that drop out at 1750 m
+ggplot(filter(fl_net_elev, sites > 0 & elev.ceiling < 1750), 
+       aes(reorder(plant.sp, elev.med), elev.mean)) +
+  geom_line(size = 1, color = "gray60") +
+  geom_point(alpha = 0.5, aes(size = abund)) +
+  theme_light(8) +
+  xlab("Plant species") +
+  ylab("Elevation") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+ggplot(filter(fl_survey_elev, sites > 0 & elev.ceiling < 1750), 
+       aes(reorder(plant.sp, elev.med), elev.mean)) +
+  geom_line(size = 1, color = "gray60") +
+  geom_point(alpha = 0.5, aes(size = abund)) +
+  theme_light(8) +
+  xlab("Plant species") +
+  ylab("Elevation") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 ```
@@
 ### BB abundance, diversity, and per-capita resource availability
-```{r eval=FALSE, echo=FALSE}
+```{r eval=FALSE, echo=FALSE, include=FALSE}
 ### Per-transect Bombus density
@@
   summarize(value = mean(value, na.rm = TRUE))
-
+```
+
+### Plotting abundance and diversity patterns
+```{r}
 ### BB abundance ~ floral abundance
@@
 
+ggplot(abund_div, aes(elev.mean, fl.rich)) +
+  geom_point(alpha = 0.5, aes(color = year)) +
+  geom_smooth(color = "black", linetype = "dashed", size = 0.75) +
+  theme_light(16) +
+  xlab("Elevation") +
+  ylab("Floral abundance")
+
 ### Floral richness ~ GDD
@@
   theme_light()
+
+ggplot(abund_div, aes(elev.mean, log(fl.abund), color = year)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth() +
+  theme_light()
+
+ggplot(abund_div, aes(elev.mean, fl.abund)) +
+  geom_point(alpha = 0.5, aes(color = year)) +
+  geom_smooth(color = "black", linetype = "dashed", size = 0.75) +
+  theme_light(16) +
+  xlab("Elevation") +
+  ylab("Floral abundance")
+
+ggplot(abund_div, aes(elev.mean, log(fl.abund))) +
+  geom_point(alpha = 0.5, aes(color = year)) +
+  geom_smooth(color = "black", linetype = "dashed", size = 0.75) +
+  theme_light(16) +
+  xlab("Elevation") +
+  ylab("log(Floral abundance)")
 
@@
 ##### Floral abundance
-```{r}
+```{r eval=FALSE, include=FALSE}
 fl_abund_yday <- gam(fl.abund ~ year +
@@
 ##### Floral richness
-```{r}
+```{r eval=FALSE, include=FALSE}
 fl_rich_yday <- gam(fl.rich ~ year +
@@
 ##### BB abundance
-```{r}
+```{r eval=FALSE, include=FALSE}
 bb_abund_yday <- gam(bb.abund ~ year +
@@
 ##### BB richness
-```{r}
+```{r eval=FALSE, include=FALSE}
 bb_rich_yday <- gam(bb.rich ~ year +
@@
 ### Rarefied network analysis (not broken down by phase)
-```{r}
+```{r, include = FALSE}
 ### Determine the minimum transect count per site*year to set rarefaction sampling
@@
   geom_point(aes(color = year)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
   xlab("Elevation") +
@@
 ### Let's look at how abundance varies on a per-species basis
-```{r}
+```{r eval=FALSE, include=FALSE}
 species <- net %>%
@@
 ### Plot some dang networks
-```{r}
+```{r eval=FALSE, include=FALSE}
 web <- net %>%
@@
 #### Network level
-```{r}
+```{r eval=FALSE, include=FALSE}
 per.transect_netmet <- net %>%
@@
 #### Species level
-```{r}
+```{r eval=FALSE, include=FALSE}
 per.transect_spmet <- net %>%
@@
 ### Phase x elev network analysis
-```{r}
+```{r eval=FALSE, include=FALSE}
 ### Prepare pxe abundance and diversity table

### 2020-06-05 analysis.Rmd
added floral taxonomy up to family level, tweaked some things in data_proc

@@
 #library(lmerTest)
+install.packages("taxize")
+library(taxize)
 library(tidyverse)
@@
 
+fl_traits <- read.csv("./processed_data/floral_k_type.csv") %>%
+  select(plant.sp, k.type)
+
+fl_tax <- read_csv("./processed_data/floral_tax.csv")
+
 net <- read_csv("./processed_data/network.csv") %>%
@@
   left_join(breaks, by = c("site", "year")) %>%
+  left_join(fl_traits) %>%
+  left_join(fl_tax, by = "plant.genus") %>%
   ungroup() %>%
@@
   mutate(year = factor(year)) %>% # convert year to factor
-  #semi_join(net, by = c("site", "plant.sp")) %>% # Consider only plants visited at a given site-date
+  semi_join(net, by = "plant.sp") %>% # Consider only plants visited at a given site-date
   left_join(gdd, by = c("site", "date")) %>%
@@
   left_join(site_data, by = "site") %>%
+  left_join(fl_traits) %>%
+  left_join(fl_tax) %>%
   ungroup() %>%
@@
   arrange(site, date)
+
 ```
@@
   left_join(site_data) %>%
-  semi_join(net, by = "plant.sp") %>% # constrain to plant species that were actually visited
   group_by(plant.sp) %>%
+  mutate(elev.floor = min(elev.mean),
+         elev.ceiling = max(elev.mean),
+         elev.range = elev.ceiling - elev.floor,
+         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2),
+         sites = length(unique(site)))
+
+fl_survey_elev_genus <- survey %>%
+  filter(flower.cover > 0) %>%
+  group_by(site, genus) %>%
+  summarize(abund = sum(flower.cover)) %>%
+  left_join(site_data) %>%
+  group_by(genus) %>%
   mutate(elev.floor = min(elev.mean),
@@
 
+ggplot(filter(fl_survey_elev_genus, sites > 0), 
+       aes(reorder(genus, elev.med), elev.mean)) +
+  geom_line(size = 1, color = "gray60") +
+  geom_point(alpha = 0.5) +
+  theme_light(10) +
+  #theme(axis.text.x = element_blank()) +
+  xlab("Plant genus") +
+  ylab("Elevation") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
 ggplot(filter(fl_net_elev, sites > 0), 
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
+### FL traits by elevation -- do we see a sharp change at the tree line?
+```{r}
+survey_trait_elev <- survey %>%
+  group_by(site, year, elev.mean,  k.type) %>%
+  summarize(abund = sum(flower.cover)) %>%
+  group_by(site, elev.mean, year) %>%
+  mutate(div = length(unique(k.type)))
+
+net_trait_elev <- net %>%
+  group_by(site, year, elev.mean,  k.type) %>%
+  summarize(abund = n()) %>%
+  group_by(site, elev.mean, year) %>%
+  mutate(div = length(unique(k.type)))
+
+survey_genus_elev <- survey %>%
+  group_by(site, year, elev.mean, plant.genus) %>%
+  summarize(abund = sum(flower.cover)) %>%
+  group_by(site, elev.mean, year) %>%
+  mutate(div = length(unique(plant.genus)))
+
+survey_family_elev <- survey %>%
+  group_by(site, year, elev.mean,  plant.family) %>%
+  summarize(abund = sum(flower.cover)) %>%
+  group_by(site, elev.mean, year) %>%
+  mutate(div = length(unique(plant.family)))
+
+net_genus_elev <- survey %>%
+  group_by(site, year, elev.mean,  plant.genus) %>%
+  summarize(abund = n()) %>%
+  group_by(site, elev.mean, year) %>%
+  mutate(div = length(unique(plant.genus)))
+
+net_family_elev <- survey %>%
+  group_by(site, year, elev.mean,  plant.family) %>%
+  summarize(abund = n()) %>%
+  group_by(site, elev.mean, year) %>%
+  mutate(div = length(unique(plant.family)))
+
+ggplot(survey_trait_elev, aes(elev.mean, log(abund))) +
+  geom_point() +
+  stat_smooth() +
+  facet_wrap(~k.type, scales = "free_y")
+
+ggplot(net_trait_elev, aes(elev.mean, log(abund))) +
+  geom_point() +
+  stat_smooth() +
+  facet_wrap(~k.type, scales = "free_y")
+
+ggplot(survey_trait_elev, aes(elev.mean, div)) +
+  geom_point() +
+  stat_smooth()
+
+ggplot(net_trait_elev, aes(elev.mean, div)) +
+  geom_point() +
+  stat_smooth()
+
+ggplot(survey_genus_elev, aes(elev.mean, log(abund))) +
+  geom_point() +
+  stat_smooth() +
+  facet_wrap(~plant.genus, scales = "free_y")
+
+ggplot(survey_family_elev, aes(elev.mean, log(abund))) +
+  geom_point() +
+  stat_smooth() +
+  facet_wrap(~plant.family, scales = "free_y")
+
+ggplot(survey_genus_elev, aes(elev.mean, div)) +
+  geom_point() +
+  stat_smooth()
+
+ggplot(survey_family_elev, aes(elev.mean, div)) +
+  geom_point() +
+  stat_smooth()
+
+ggplot(net_genus_elev, aes(elev.mean, log(abund))) +
+  geom_point() +
+  stat_smooth() +
+  facet_wrap(~plant.genus, scales = "free_y")
+
+ggplot(net_family_elev, aes(elev.mean, log(abund))) +
+  geom_point() +
+  stat_smooth() +
+  facet_wrap(~plant.family, scales = "free_y")
+
+ggplot(net_genus_elev, aes(elev.mean, log(abund))) +
+  geom_point() +
+  stat_smooth(se = FALSE) +
+  facet_wrap(~plant.genus)
+
+ggplot(net_genus_elev, aes(elev.mean, log(abund))) +
+  geom_point() +
+  stat_smooth() +
+  facet_wrap(~plant.genus, scales = "free_y")
+ggplot(net_genus_elev, aes(elev.mean, abund)) +
+  geom_point() +
+  stat_smooth(se = FALSE) +
+  facet_wrap(~plant.genus)
+
+ggplot(net_genus_elev, aes(elev.mean, div)) +
+  geom_point() +
+  stat_smooth()
 ```

### 2020-06-05 data_proc.R

@@ -17,7 +17,7 @@ fl_survey <- read_delim("./raw_data/data_plantcover_forDoug_2020.01.17.csv", del
                                            "Gymnadenia bifolia" = "Platanthera bifolia"))) %>%
   separate(species_plant, c("genus", "species"), remove = FALSE) %>%
   dplyr::select(year, dayofyear, site = site_name, snowcover, 
-                plant.sp = species_plant, flower.cover = cover_all_flowers) %>%
+                plant.sp = species_plant, plant.genus = genus, flower.cover = cover_all_flowers) %>%
   mutate(temp.date = as_date(dayofyear),
          day = day(temp.date),
          month = month(temp.date),
@@ -34,7 +34,7 @@ network <- read_delim("./raw_data/data_bumblebee_forDoug_2020.04.17.csv", delim 
   separate(plant_sp_latin, c("genus", "species"), remove = FALSE) %>%
   dplyr::select(year, dayofyear, site = site_name, trap.time = trapping_time, caste, pollen, 
                 bb.sp = bumb_species, bb.sp.lat = bumb_sp_latin, plant.sp.abb = forage_plant, 
-                plant.sp = plant_sp_latin) %>%
+                plant.sp = plant_sp_latin, plant.genus = genus) %>%
   mutate(temp.date = as_date(dayofyear),
          day = day(temp.date),
          month = month(temp.date),
@@ -53,6 +53,34 @@ network <- read_delim("./raw_data/data_bumblebee_forDoug_2020.04.17.csv", delim 
                                         "Bombus flavidus" = "Psithyrus")))
 
 write_csv(network, "./processed_data/network.csv")
+
+### Floral taxonomy
+Sys.setenv(ENTREZ_KEY = "ab9a55ec842df6f86a750929aefc69143608")
+
+fl_sp <- survey %>%
+  select(plant.sp) %>%
+  unique() %>%
+  bind_rows(select(network, plant.sp)) %>%
+  unique()
+
+fl_tax <- tax_name(fl_sp$plant.sp, get = "family", db = "ncbi") %>%
+  select(plant.sp = query, plant.family = family) %>%
+  separate(plant.sp, c("plant.genus", "plant.sp"), remove = FALSE) %>%
+  select(plant.genus, plant.family) %>%
+  unique()
+
[…]

### 2020-06-05 analysis.Rmd
did something

@@
 #library(lmerTest)
-install.packages("taxize")
 library(taxize)
@@
 
-fl_survey_elev_genus <- survey %>%
+fl_survey_elev_family <- survey %>%
   filter(flower.cover > 0) %>%
-  group_by(site, genus) %>%
+  group_by(site, plant.family) %>%
   summarize(abund = sum(flower.cover)) %>%
   left_join(site_data) %>%
-  group_by(genus) %>%
+  group_by(plant.family) %>%
   mutate(elev.floor = min(elev.mean),
@@
 ggplot(filter(fl_survey_elev, sites > 0), 
-       aes(reorder(plant.sp, elev.med), elev.mean)) +
+       aes(reorder(plant.sp, elev.ceiling), elev.mean)) +
   geom_line(size = 1, color = "gray60") +
@@
 
-ggplot(filter(fl_survey_elev_genus, sites > 0), 
-       aes(reorder(genus, elev.med), elev.mean)) +
+ggplot(filter(fl_survey_elev_family, sites > 0), 
+       aes(reorder(plant.family, elev.med), elev.mean)) +
   geom_line(size = 1, color = "gray60") +
@@
   #theme(axis.text.x = element_blank()) +
-  xlab("Plant genus") +
+  xlab("Plant family") +
   ylab("Elevation") +
@@
   xlab("Elevation") +
-  ylab("Floral abundance")
+  ylab("Floral richness")
 
@@
 summary(fl_abund_yday.vis)
-plot(sm(fl_abund_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("FL abundance 2010 (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-plot(sm(fl_abund_yday.vis, 2)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("FL abundance 2011 (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-plot(sm(fl_abund_yday.vis, 3)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("FL abundance 2012 (yday)") +
+
+plot(fl_abund_yday.vis, allTerms = TRUE) +
+  ggtitle("FL abundance (yday)") +
   xlab("Day of year") +
@@
   ggtitle("FL abundance (gdd)") +
-  xlab("Day of year") +
+  xlab("GDD") +
   ylab("Elevation")
@@
 
-##### Floral richness
+##### Floral richness -- since year does not seem to change the overall pattern (though perhaps the magnitude), I could use a simpler model, perhaps with year as a random intercept effect
 ```{r eval=FALSE, include=FALSE}
@@
 summary(fl_rich_yday.vis)
-plot(sm(fl_rich_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("FL richness 2010 (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-plot(sm(fl_rich_yday.vis, 2)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("FL richness 2011 (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-plot(sm(fl_rich_yday.vis, 3)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("FL richness 2012 (yday)") +
+
+plot(fl_rich_yday.vis, allTerms = TRUE) +
+  ggtitle("FL richness (yday)") +
   xlab("Day of year") +
@@
 summary(bb_abund_yday.vis)
-plot(sm(bb_abund_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("bb abundance 2010 (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-plot(sm(bb_abund_yday.vis, 2)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("bb abundance 2011 (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-plot(sm(bb_abund_yday.vis, 3)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("bb abundance 2012 (yday)") +
+
+plot(bb_abund_yday.vis, allTerms = TRUE) +
+  ggtitle("BB abundance (yday)") +
   xlab("Day of year") +
@@
   ggtitle("BB abundance (gdd)") +
-  xlab("Day of year") +
+  xlab("GDD") +
   ylab("Elevation")

### 2020-06-10 analysis.Rmd
update

@@
 
-# network level PxE
-rare_net_nt_pxe <- function(net, n, iter) {
+```
+
+### Rarefied network analysis by floral k.type instead of floral species
+```{r}
+# Rare_net function: rarefies networks by iteratively subsampling more frequently sampled networks to match the transect count of the least frequently sampled one
+# species level
+rare_net_sp_k.type.s <- function(net, n, iter) {
 
   out <- data.frame()
-
-  t_filter <- net %>%
-    select(site, date, phase, year) %>%
-    unique() %>%
-    group_by(site, phase, year) %>%
-    mutate(transects = n()) %>%
-    filter(transects >= n)
 
@@
     temp <- net %>%
-      semi_join(t_filter) %>%
-      select(site, date, phase, year) %>%
+      select(site, date) %>%
       unique() %>%
-      group_by(site, phase, year) %>%
+      group_by(site) %>%
       sample_n(n) %>%
       left_join(net) %>%
-      group_by(site, phase, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
+      group_by(site, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
       summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      group_by(site, phase, year) %>%
-      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
+      na.omit() %>%
+      group_by(site) %>%
+      filter(length(unique(bb.sp)) > 1 & length(unique(k.type.s)) > 1) %>%
       ungroup() %>%
-      unite(webID, c(site, phase, year), sep = "_") %>%
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
+      filter(freq > 1) %>% # remove singletons
+      dplyr::select(higher = bb.sp, lower = k.type.s, webID = site, freq) %>% # rename columns to match bipartite's expectations
       data.frame() %>% # convert to data frame
       frame2webs() %>% # convert to to bipartite web
-      map(networklevel, weighted = TRUE, level = "higher",
-          index = c("web asymmetry", "weighted NODF", "H2", "C score", "niche overlap", "generality")) %>% 
+      map(specieslevel, level = "higher", index = c("d", "species strength", "proportional generality")) %>%
+      map(rownames_to_column) %>%
+      map(as_tibble) %>%
+      bind_rows(.id = "site") %>% # Collapse list into big data frame
+      rename(sp = rowname) %>%
+      mutate(iteration = rep(i, n()))
+
+    out <- bind_rows(out, temp)
+  } 
+  out <- out %>%
+    gather(metric, value, -c(sp, site, iteration)) %>% # get mean value across iterations for each metric
+    group_by(sp, site, metric) %>%
+    summarize(value = mean(value)) %>%
+    left_join(site_data) %>%
+    left_join(bb_traits, by = c("sp" = "bb.sp"))
+}
+
+# network level
+rare_net_nt_k.type.s <- function(net, n, iter) {
+
+  out <- data.frame()
+
+  for(i in 1:iter) {
+    temp <- net %>%
+      select(site, date) %>%
+      unique() %>%
+      group_by(site) %>%
+      sample_n(n) %>%
+      left_join(net) %>%
+      group_by(site, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
+      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
+      na.omit() %>%
+      group_by(site) %>%
+      filter(length(unique(bb.sp)) > 1 & length(unique(k.type.s)) > 1) %>%
+      ungroup() %>%
+      dplyr::select(higher = bb.sp, lower = k.type.s, webID = site, freq) %>% # rename columns to match bipartite's expectations
+      data.frame() %>% # convert to data frame
+      frame2webs() %>% # convert to to bipartite web
+      map(networklevel, weighted = TRUE, level = "higher", index = c("H2", "weighted NODF", "niche overlap", "C score", "generality", "web asymmetry", "weighted connectance")) %>% # We probably care mainly about the BB level of the network
       data.frame() %>%
@@
       rownames_to_column() %>%
-      dplyr::select(id = rowname, everything()) %>%
+      dplyr::select(site = rowname, everything()) %>%
       as_tibble() %>%
@@
   out <- out  %>%
-      gather(metric, value, -c(id, iteration)) %>% # get mean value across iterations for each metric
-      group_by(id, metric) %>%
+      gather(metric, value, -c(site, iteration)) %>% # get mean value across iterations for each metric
+      group_by(site, metric) %>%
       summarize(value = mean(value)) %>%
-      separate(id, sep = "_", into = c("site", "phase", "year")) %>%
-      mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
       left_join(site_data)
-}
-
-# species level version 2
-rare_net_sp2 <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  t_filter <- net %>%
-    select(site, date, phase, year) %>%
-    unique() %>%
-    group_by(site, phase, year) %>%
-    mutate(transects = n()) %>%
-    filter(transects >= n)
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      semi_join(t_filter) %>%
-      select(site, date, phase, year) %>%
-      unique() %>%
-      group_by(site, phase, year) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, phase, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      group_by(site, phase, year) %>%
-      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
-      ungroup() %>%
-      unite(webID, c(site, phase, year), sep = "_") %>%
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(specieslevel, level = "higher", index = c("d", "species strength")) %>%
-      map(rownames_to_column) %>%
-      map(as_tibble) %>%
-      bind_rows(.id = "id") %>% # Collapse list into big data frame
-      rename(sp = rowname) %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out %>%
-    gather(metric, value, -c(sp, id, iteration)) %>% # get mean value across iterations for each metric
-    group_by(sp, id, metric) %>%
-    summarize(value = mean(value)) %>%
-    separate(id, sep = "_", into = c("site", "phase", "year")) %>%
-    mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
-    left_join(site_data) %>%
-    left_join(bb_traits, by = c("sp" = "bb.sp"))
 }
@@
 
-fl_traits <- read.csv("./processed_data/floral_k_type.csv") %>%
-  select(plant.sp, k.type)
+fl_traits <- read.csv("./processed_data/floral_k_type_nom_fix.csv") %>%
+  select(plant.sp, k.type) %>%
+  mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"),
+         k.type.ss = str_extract(k.type, "[-+]?[0-9]*"))
 
@@
 survey_trait_elev <- survey %>%
-  group_by(site, year, elev.mean,  k.type) %>%
-  summarize(abund = sum(flower.cover)) %>%
-  group_by(site, elev.mean, year) %>%
-  mutate(div = length(unique(k.type)))
+  group_by(site, year, elev.mean,  k.type.s) %>%
+  summarize(abund = sum(flower.cover))
+
+survey_trait_elev2 <- survey %>%
+  group_by(site, year, elev.mean,  k.type.ss) %>%
+  summarize(abund = sum(flower.cover))
 
 net_trait_elev <- net %>%
-  group_by(site, year, elev.mean,  k.type) %>%
-  summarize(abund = n()) %>%
-  group_by(site, elev.mean, year) %>%
-  mutate(div = length(unique(k.type)))
+  group_by(site, year, elev.mean,  k.type.s) %>%
+  summarize(abund = n())
+
+net_trait_elev2 <- net %>%
+  group_by(site, year, elev.mean,  k.type.ss) %>%
+  summarize(abund = n())
+
+ggplot(survey_trait_elev, aes(elev.mean, sqrt(abund))) +
+  geom_point() +
+  geom_smooth() +
+  geom_vline(xintercept = 1750, color = "red") +
+  facet_wrap(~k.type.s, scales = "free_y")
+
+ggplot(survey_trait_elev2, aes(elev.mean, sqrt(abund))) +
+  geom_point() +
+  geom_smooth() +
+  geom_vline(xintercept = 1750, color = "red") +
+  facet_wrap(~k.type.ss, scales = "free_y")
+
+ggplot(net_trait_elev, aes(elev.mean, sqrt(abund))) +
+  geom_point() +
+  geom_smooth() +
+  geom_vline(xintercept = 1750, color = "red") +
+  facet_wrap(~k.type.s, scales = "free_y")
+
+ggplot(net_trait_elev2, aes(elev.mean, sqrt(abund))) +
+  geom_point() +
+  geom_smooth() +
+  geom_vline(xintercept = 1750, color = "red") +
+  facet_wrap(~k.type.ss, scales = "free_y")
+
+net_trait_genus_elev <- net %>%
+  group_by(site, year, elev.mean, plant.genus, k.type.s) %>%
+  summarize(abund = n())
+
+survey_trait_genus_elev <- survey %>%
+  group_by(site, year, elev.mean, plant.genus, k.type.s) %>%
+  summarize(abund = sum(flower.cover))
+
+ggplot(filter(net_trait_genus_elev, k.type.s == 5.1), 
+       aes(elev.mean, sqrt(abund))) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~plant.genus)
+
+ggplot(filter(survey_trait_genus_elev, k.type.s %in% c(5.1)), 
+       aes(elev.mean, sqrt(abund))) +
+  geom_point() +
+  geom_smooth() +
+  geom_vline(xintercept = 1750, color = "red") +
+  facet_wrap(~plant.genus, ncol = 5) +
+  ylab("sqrt(Abundance)") +
+  xlab("Elevation")
+
+ggplot(filter(net_trait_genus_elev, k.type.s %in% c(5.1)), 
+       aes(elev.mean, sqrt(abund))) +
+  geom_point() +
+  geom_smooth() +
+  geom_vline(xintercept = 1750, color = "red") +
+  facet_wrap(~plant.genus, ncol = 5) +
+  ylab("sqrt(Visitation frequency)") +
+  xlab("Elevation")
+```
+
+### Genera and families by elevation
+```{r}
 
@@
 
-ggplot(survey_trait_elev, aes(elev.mean, log(abund))) +
-  geom_point() +
-  stat_smooth() +
-  facet_wrap(~k.type, scales = "free_y")
-
-ggplot(net_trait_elev, aes(elev.mean, log(abund))) +
-  geom_point() +
-  stat_smooth() +
-  facet_wrap(~k.type, scales = "free_y")
-
-ggplot(survey_trait_elev, aes(elev.mean, div)) +
-  geom_point() +
-  stat_smooth()
-
-ggplot(net_trait_elev, aes(elev.mean, div)) +
-  geom_point() +
-  stat_smooth()
-
+#### Plot
 ggplot(survey_genus_elev, aes(elev.mean, log(abund))) +
@@
 
+ggplot(abund_div, aes(log(fl.abund), bb.abund, color = year)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth() +
+  ylab("BB abundance") +
+  xlab("log(floral cover)") +
+  theme_bw()
+
 ### BB richness ~ floral abundance
@@
   facet_grid(year~phase, scales = "free") +
+  ylab("log(fl-per-bb)") +
+  xlab("Elevation") +
+  theme_light()
+
+ggplot(abund_div, aes(elev.mean, log(fl.per.bb))) +
+  geom_point(alpha = 0.5) +
+  #stat_summary(aes(y = log(fl.per.bb), group=1), fun.y=mean, colour="black", geom="point",group=1) +
+  geom_smooth() +
+  facet_grid(.~year, scales = "free") +
   ylab("log(fl-per-bb)") +
@@
   theme_light()
+
+ggplot(filter(abund_div_mean, metric == "fl.per.bb"), aes(elev.mean, log(value), color = year)) +
+  geom_point(alpha = 1) +
+  #geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
+  geom_smooth() +
+  ylab("FL per BB") +
+  xlab("Elevation") +
+  theme_light()
+
+ggplot(filter(abund_div_mean, metric == "fl.per.bb"), aes(elev.mean, log(value))) +
+  geom_point(alpha = 1, aes(color = year)) +
+  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
+  #geom_smooth() +
+  ylab("FL per BB") +
+  xlab("Elevation") +
+  theme_light(16)
 
@@
   geom_smooth() +
-  facet_grid(year ~ phase)
+  facet_grid(.~year)
 
@@
   facet_grid(year ~ phase, scales = "free")
+
+ggplot(abund_div, aes(elev.mean, bb.abund)) +
+  geom_point() +
+  geom_smooth() +
+  facet_grid(.~year, scales = "free")
+
+ggplot(abund_div, aes(elev.mean, bb.abund, color = year)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth() +
+  xlab("Elevation") +
+  ylab("BB abundance")  +
+  theme_light(16)
+
+ggplot(abund_div, aes(elev.mean, sqrt(bb.abund), color = year)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth() +
+  xlab("Elevation") +
+  ylab("sqrt(BB abundance)")  +
+  theme_light(16)
 
@@
   ylab("log(Floral abundance)")
+
+ggplot(abund_div, aes(elev.mean, fl.abund, color = year)) +
+  geom_point(alpha = 0.5) +
+  theme_light(16) +
+  xlab("Elevation") +
+  ylab("Floral abundance")
+
+ggplot(abund_div, aes(elev.mean, log(fl.abund), color = year)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth() +
+  theme_light(16) +
+  xlab("Elevation") +
+  ylab("log(Floral abundance)")
+
+ggplot(abund_div, aes(elev.mean, sqrt(bb.abund), color = year)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth() +
+  xlab("Elevation") +
+  ylab("sqrt(BB abundance)")  +
+  theme_light(16)
 
@@
 
-##### Floral richness -- since year does not seem to change the overall pattern (though perhaps the magnitude), I could use a simpler model, perhaps with year as a random intercept effect
+##### Floral richness -- since year does not seem to change the overall pattern (though perhaps the magnitude), I could use a simpler model, perhaps with year as a random intercept effect; some explanat
 ```{r eval=FALSE, include=FALSE}
+### By year
 fl_rich_yday <- gam(fl.rich ~ year +
@@
 plot(fl_rich_yday.vis, allTerms = TRUE) +
+  ggtitle("FL abundance (gdd)") +
+  xlab("Day of year") +
+  ylab("Elevation")
+
+### With year as parametric term
+fl_rich_yday2 <- gam(fl.rich ~ year +
+                    te(yday, elev.mean, bs= "cc", k= 10, m=1),
+                    data = filter(abund_div),
+                    method = "REML")
+
+fl_rich_yday2.vis <- getViz(fl_rich_yday2)
+check.gamViz(fl_rich_yday2.vis)
+summary(fl_rich_yday2.vis)
+
+plot(sm(fl_rich_yday2.vis, 1)) + 
+  l_fitRaster() + 
+  l_fitContour() + 
+  l_points(size = 1) +
   ggtitle("FL richness (yday)") +
@@
 
+plot(pterm(fl_rich_yday2.vis, 1))
+
+### By year
 fl_rich_gdd <- gam(fl.rich ~ year +
@@
   ggtitle("FL richness (gdd)") +
+  xlab("GDD") +
+  ylab("Elevation")
+
+### With year as parametric term
+fl_rich_gdd2 <- gam(fl.rich ~ year +
+                    te(gdd.cum, elev.mean, bs= "cc", k= 10, m=1),
+                    data = filter(abund_div),
+                    method = "REML")
+
+fl_rich_gdd2.vis <- getViz(fl_rich_gdd2)
+check.gamViz(fl_rich_gdd2.vis)
+summary(fl_rich_gdd2.vis)
+
+plot(sm(fl_rich_gdd2.vis, 1)) + 
+  l_fitRaster() + 
+  l_fitContour() + 
+  l_points(size = 1) +
+  ggtitle("FL richness (gdd)") +
   xlab("Day of year") +
   ylab("Elevation")
+
+plot(pterm(fl_rich_gdd2.vis, 1))
 ```
@@
   xlab("Elevation")
+```
+
+### Rarefied network analysis by k.type instead of by plant sp
+```{r, include = FALSE}
+### Create yearly subsets to feed to rare_net
+net_2012 <- filter(net, year == "2012")
+net_2011 <- filter(net, year == "2011")
+net_2010 <- filter(net, year == "2010")
+
+### Species level
+spmet_2012 <- rare_net_sp_k.type(net_2012, 9, 20) %>%
+  mutate(year = factor(rep("2012", n())))
+spmet_2011 <- rare_net_sp(net_2011, 4, 20) %>%
+  mutate(year = factor(rep("2011", n())))  
+spmet_2010 <- rare_net_sp(net_2010, 3, 20) %>%
+  mutate(year = factor(rep("2010", n())))
+
+spmet <- bind_rows(spmet_2012, spmet_2011, spmet_2010)
+
+spmet_mean <- spmet %>%
+  group_by(site, year, elev.mean, metric) %>%
+  summarize(value = mean(value))
+
+### Network- and group-level
+netmet_rare_2012 <- rare_net_nt_k.type(net_2012, n = 9, iter = 20) %>%
+  mutate(year = factor(rep("2012", n())))
+netmet_rare_2011 <- rare_net_nt_k.type(net_2011, n = 4, iter = 20)  %>%
+  mutate(year = factor(rep("2011", n())))
+netmet_rare_2010 <- rare_net_nt_k.type(net_2010, n = 3, iter = 20) %>%
+  mutate(year = factor(rep("2010", n())))
+
+netmet <- bind_rows(netmet_rare_2012, 
+                    netmet_rare_2011, 
+                    netmet_rare_2010) %>%
+  mutate(metric = factor(metric, levels = c("H2",
+                                            "weighted.NODF",
+                                            "niche.overlap.HL", 
+                                            "C.score.HL",
+                                            "generality.HL",
+                                            "web.asymmetry",
+                                            "weighted.connectance")))
+
+### Plot
+#### Per species values
+ggplot(filter(spmet_2012,
+              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
+       aes(elev.mean, value)) +
+  geom_point() +
+  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  geom_smooth() +
+  facet_grid(metric~sp, scales = "free")
+
+ggplot(filter(spmet_2011,
+              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
+       aes(elev.mean, value)) +
+  geom_point() +
+  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  geom_smooth() +
+  facet_grid(metric~sp, scales = "free")
+
+ggplot(filter(spmet_2010,
+              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
+       aes(elev.mean, value)) +
+  geom_point() +
+  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  geom_smooth() +
+  facet_grid(metric~sp, scales = "free")
+
+#### Community mean d
+ggplot(filter(spmet_mean, metric == "d"), aes(elev.mean, value)) +
+  geom_point(aes(color = year)) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
+  xlab("Elevation") +
+  ylab("Community mean d'") +
+  theme_light(16)
+
+gam_d <- gam(value ~ year + s(elev.mean, k = 10), 
+             data = filter(spmet, metric == "d"),
+             method = "REML")
+
+gam_d.vis <- getViz(gam_d)
+check.gamViz(gam_d.vis)
+summary(gam_d.vis)
+
+plot(gam_d.vis, allTerms = TRUE) +
+  xlab("Elevation")
+
+#### Network level
+ggplot(netmet, aes(elev.mean, value)) +
+  geom_point() +
+  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  geom_smooth() +
+  facet_grid(metric~year, scales = "free")
+
+##### H2
+ggplot(filter(netmet, metric == "H2"), aes(elev.mean, value)) +
+  geom_point(aes(color = year)) +
+  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
+  xlab("Elevation") +
+  ylab("H2'") +
+  theme_light(16)
+
+gam_H2 <- gam(value ~ year + s(elev.mean), 
+              data = filter(netmet, metric == "H2"),
+              method = "REML")
+
+gam_H2.vis <- getViz(gam_H2)
+check.gamViz(gam_H2.vis)
+summary(gam_H2.vis)
+
+plot(gam_H2.vis, allTerms = TRUE) +
+  xlab("Elevation")
+
+##### niche overlap
+ggplot(filter(netmet, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
+  geom_point(aes(color = year)) +
+  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
+  #geom_smooth() +
+  xlab("Elevation") +
+  ylab("Niche overlap") +
+  theme_light(16)
+
+gam_niche.overlap <- gam(value ~ year + s(elev.mean), 
+                         data = filter(netmet, metric == "niche.overlap.HL"),
+                         method = "REML")
+
+gam_niche.overlap.vis <- getViz(gam_niche.overlap)
+check.gamViz(gam_niche.overlap.vis)
+summary(gam_niche.overlap.vis)
+
+plot(gam_niche.overlap.vis, allTerms = TRUE) +
+  xlab("Elevation")
+
+##### C.score
+ggplot(filter(netmet, metric == "C.score.HL"), aes(elev.mean, value)) +
+  geom_point(aes(color = year)) +
+  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
+  #geom_smooth() +
+  xlab("Elevation") +
+  ylab("C score") +
+  theme_light(16)
+
+gam_cscore <- gam(value ~ year + s(elev.mean), 
+                  data = filter(netmet, metric == "C.score.HL"),
+                  method = "REML")
+
+gam_cscore.vis <- getViz(gam_cscore)
+check.gamViz(gam_cscore.vis)
+summary(gam_cscore.vis)
+
+plot(gam_cscore.vis, allTerms = TRUE) +
+  xlab("Elevation")
+
+##### Connectance
+ggplot(filter(netmet, metric == "weighted.connectance"), aes(elev.mean, value)) +
+  geom_point(aes(color = year)) +
+  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
+  #geom_smooth() +
+  xlab("Elevation") +
+  ylab("Connectance (weighted)") +
+  theme_light(16)
+
+gam_connectance <- gam(value ~ year + s(elev.mean), 
+                  data = filter(netmet, metric == "weighted.connectance"),
+                  method = "REML")
+
+gam_connectance.vis <- getViz(gam_connectance)
+check.gamViz(gam_connectance.vis)
+summary(gam_connectance.vis)
+
+plot(gam_connectance.vis, allTerms = TRUE) +
+  xlab("Elevation")
+
+##### NODF
+ggplot(filter(netmet, metric == "weighted.NODF"), aes(elev.mean, value)) +
+  geom_point(aes(color = year)) +
+  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
+  #geom_smooth() +
+  xlab("Elevation") +
+  ylab("NODF (weighted)") +
+  theme_light(16)
+
+gam_NODF <- gam(value ~ year + s(elev.mean), 
+                data = filter(netmet, metric == "weighted.NODF"),
+                method = "REML")
+
+gam_NODF.vis <- getViz(gam_NODF)
+check.gamViz(gam_NODF.vis)
+summary(gam_NODF.vis)
+
+plot(gam_NODF.vis, allTerms = TRUE) +
+  xlab("Elevation")
+
+##### Web asymmetry
+ggplot(filter(netmet, metric == "web.asymmetry"), aes(elev.mean, value)) +
+  geom_point(aes(color = year)) +
+  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
+  #geom_smooth() +
+  xlab("Elevation") +
+  ylab("Web asymmetry") +
+  theme_light(16)
+
+gam_webass <- gam(value ~ year + s(elev.mean), 
+                data = filter(netmet, metric == "web.asymmetry"),
+                method = "REML")
+
+gam_webass.vis <- getViz(gam_webass)
+check.gamViz(gam_webass.vis)
+summary(gam_webass.vis)
+
+plot(gam_webass.vis, allTerms = TRUE) +
+  xlab("Elevation")
+```
+
+### Do BB species differ in their k.type associations?
+```{r}
+web_by_site <- net %>%
+  group_by(site, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
+  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
+  na.omit() %>%
+  group_by(site) %>%
+  filter(length(unique(bb.sp)) > 1 & length(unique(k.type.s)) > 1) %>%
+  ungroup() %>%
+  dplyr::select(higher = bb.sp, lower = k.type.s, webID = site, freq) %>% # rename columns to match bipartite's expectations
+  data.frame() %>% # convert to data frame
+  frame2webs()
+
+visweb(web[[1]])
+plotweb(web[[1]])
+
+web <- net %>%
+  group_by(year, elev.class2, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
+  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
+  na.omit() %>%
+  unite(webID, c(year, elev.class2), sep = "_") %>%
+  dplyr::select(higher = bb.sp, lower = k.type.s, webID, freq) %>% # rename columns to match bipartite's expectations
+  data.frame() %>% # convert to data frame
+  frame2webs()
+
+# 2010 high
+visweb(web[[1]])
+plotweb(web[[1]])
+
+# 2010 low
+visweb(web[[2]])
+plotweb(web[[2]])
+
+# 2010 med
+visweb(web[[3]])
+plotweb(web[[3]])
+
+# 2011 high
+visweb(web[[4]])
+plotweb(web[[4]])
+
+# 2011 low
+visweb(web[[5]])
+plotweb(web[[5]])
+
+# 2011 med
+visweb(web[[6]])
+plotweb(web[[6]])
+
+# 2012 high
+visweb(web[[7]])
+plotweb(web[[7]])
+
+# 2012 low
+visweb(web[[8]])
+plotweb(web[[8]])
+
+# 2012 med
+visweb(web[[9]])
+plotweb(web[[9]])
+
+```
+
+### Does tongue length class predict k.type association? Remember that B. wurflenii is primarily a nectar robber.
+```{r}
+library(infer)
+
+kugler_desc <- read_delim("./kugler.tsv", delim = "\t") %>%
+  select(1,2)
+
+ggplot(bb_traits, aes(pbl.w)) +
+  geom_histogram()
+
+net_cont <- net %>%
+  group_by(year, elev.class2, pbl.w.class, k.type.ss) %>%
+  summarize(freq = n()) %>%
+  group_by(elev.class2, pbl.w.class) %>%
+  mutate(prop = freq/sum(freq))
+
+ggplot(net_cont, aes(pbl.w.class, freq, fill = k.type.ss)) +
+  geom_col() +
+  facet_grid(.~elev.class2)
+
+ggplot(net_cont, aes(pbl.w.class, prop, fill = k.type.ss)) +
+  geom_col() +
+  facet_grid(.~elev.class2)
+
+chisq_mod <- chisq.test(net$pbl.w.class, net$k.type.ss)
+chisq_mod
+
+  c <- net %>%
+  select(k.type.s, pbl.w.class) %>%
+  mutate(k.type.s = factor(k.type.s),
+         pbl.w.class = factor(pbl.w.class)) %>%
+  chisq_test(k.type.s ~ pbl.w.class)
+
 ```
@@
 ```
+
+```{r}
+
+# network level PxE
+rare_net_nt_pxe <- function(net, n, iter) {
+
+  out <- data.frame()
+
+  t_filter <- net %>%
+    select(site, date, phase, year) %>%
+    unique() %>%
+    group_by(site, phase, year) %>%
+    mutate(transects = n()) %>%
+    filter(transects >= n)
+
+  for(i in 1:iter) {
+    temp <- net %>%
+      semi_join(t_filter) %>%
+      select(site, date, phase, year) %>%
+      unique() %>%
+      group_by(site, phase, year) %>%
+      sample_n(n) %>%
+      left_join(net) %>%
+      group_by(site, phase, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
+      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
+      group_by(site, phase, year) %>%
+      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
+      ungroup() %>%
+      unite(webID, c(site, phase, year), sep = "_") %>%
+      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
+      data.frame() %>% # convert to data frame
+      frame2webs() %>% # convert to to bipartite web
+      map(networklevel, weighted = TRUE, level = "higher",
+          index = c("web asymmetry", "weighted NODF", "H2", "C score", "niche overlap", "generality")) %>% 
+      data.frame() %>%
+      t() %>%
+      data.frame() %>%
+      rownames_to_column() %>%
+      dplyr::select(id = rowname, everything()) %>%
+      as_tibble() %>%
+      mutate(iteration = rep(i, n()))
+
+    out <- bind_rows(out, temp)
+  } 
+  out <- out  %>%
+      gather(metric, value, -c(id, iteration)) %>% # get mean value across iterations for each metric
+      group_by(id, metric) %>%
+      summarize(value = mean(value)) %>%
+      separate(id, sep = "_", into = c("site", "phase", "year")) %>%
+      mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
+      left_join(site_data)
+}
+
+# species level version 2
+rare_net_sp2 <- function(net, n, iter) {
+
+  out <- data.frame()
+
+  t_filter <- net %>%
+    select(site, date, phase, year) %>%
+    unique() %>%
+    group_by(site, phase, year) %>%
+    mutate(transects = n()) %>%
+    filter(transects >= n)
+
+  for(i in 1:iter) {
+    temp <- net %>%
+      semi_join(t_filter) %>%
+      select(site, date, phase, year) %>%
+      unique() %>%
+      group_by(site, phase, year) %>%
+      sample_n(n) %>%
+      left_join(net) %>%
+      group_by(site, phase, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
+      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
+      group_by(site, phase, year) %>%
+      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
+      ungroup() %>%
+      unite(webID, c(site, phase, year), sep = "_") %>%
+      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
+      data.frame() %>% # convert to data frame
+      frame2webs() %>% # convert to to bipartite web
+      map(specieslevel, level = "higher", index = c("d", "species strength")) %>%
+      map(rownames_to_column) %>%
+      map(as_tibble) %>%
+      bind_rows(.id = "id") %>% # Collapse list into big data frame
+      rename(sp = rowname) %>%
+      mutate(iteration = rep(i, n()))
+
+    out <- bind_rows(out, temp)
+  } 
+  out <- out %>%
+    gather(metric, value, -c(sp, id, iteration)) %>% # get mean value across iterations for each metric
+    group_by(sp, id, metric) %>%
+    summarize(value = mean(value)) %>%
+    separate(id, sep = "_", into = c("site", "phase", "year")) %>%
+    mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
+    left_join(site_data) %>%
+    left_join(bb_traits, by = c("sp" = "bb.sp"))
+}
+
+```

### 2020-06-10 data_proc.R

@@ -28,6 +28,12 @@ write_csv(fl_survey, "./processed_data/floral_survey.csv")
 ### Visitation data
 network <- read_delim("./raw_data/data_bumblebee_forDoug_2020.04.17.csv", delim = ";", 
                            locale = locale(decimal_mark = ",", grouping_mark = ".")) %>%
+  mutate(plant_sp_latin = str_replace_all(plant_sp_latin, 
+                                         c("Arabis caerula" = "Arabis caerulea", # correct nomenclature typos and synonyms
+                                           "Arctostaphylos alpina" = "Arctous alpina",
+                                           "Calluna vulagris" = "Calluna vulgaris",
+                                           "Deschampsia caespitosa" = "Deschampsia cespitosa",
+                                           "Gymnadenia bifolia" = "Platanthera bifolia")))%>%
   separate(col = plant_sp_latin, into = c("genus", "species", "w", "x", "y", "z"), sep = " ") %>%
   unite(plant_sp_latin, genus, species, sep = " ") %>%
   dplyr::select(-c(w,x,y,z)) %>%
@@ -92,7 +98,7 @@ site_data <- read_tsv("./raw_data/SiteLocationFabriceUpdate2020.tsv",
   mutate(elev.class2 = case_when(
     elev.mean < 1000 ~ "low",
     elev.mean >= 1000 & elev.mean < 1500 ~ "mid",
-    elev.mean >= 1500 ~ "high"
+    elev.mean >= 1750 ~ "high"
   ))
 
 write_csv(site_data, "./processed_data/site_data.csv")

### 2020-06-12 analysis.Rmd
Polished the GAM analyses

@@
 # species level
-rare_net_sp <- function(net, n, iter) {
+rare_net_sp_BB <- function(net, n, iter) {
 
@@
 
+rare_net_sp_FL <- function(net, n, iter) {
+
+  out <- data.frame()
+
+  for(i in 1:iter) {
+    temp <- net %>%
+      select(site, date) %>%
+      unique() %>%
+      group_by(site) %>%
+      sample_n(n) %>%
+      left_join(net) %>%
+      group_by(site, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
+      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
+      group_by(site) %>%
+      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
+      ungroup() %>%
+      filter(freq > 1) %>% # remove singletons
+      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
+      data.frame() %>% # convert to data frame
+      frame2webs() %>% # convert to to bipartite web
+      map(specieslevel, level = "lower", index = c("d", "species strength", "proportional generality")) %>%
+      map(rownames_to_column) %>%
+      map(as_tibble) %>%
+      bind_rows(.id = "site") %>% # Collapse list into big data frame
+      rename(sp = rowname) %>%
+      mutate(iteration = rep(i, n()))
+
+    out <- bind_rows(out, temp)
+  } 
+  out <- out %>%
+    gather(metric, value, -c(sp, site, iteration)) %>% # get mean value across iterations for each metric
+    group_by(sp, site, metric) %>%
+    summarize(value = mean(value)) %>%
+    left_join(site_data) %>%
+    left_join(fl_traits, by = c("sp" = "plant.sp"))
+}
+
 # network level
@@
       frame2webs() %>% # convert to to bipartite web
-      map(networklevel, weighted = TRUE, level = "higher", index = c("H2", "weighted NODF", "niche overlap", "C score", "generality", "web asymmetry", "weighted connectance")) %>% # We probably care mainly about the BB level of the network
+      map(networklevel, weighted = TRUE, index = c("H2", "weighted NODF", "niche overlap", "C score", "generality", "web asymmetry", "weighted connectance")) %>% # We probably care mainly about the BB level of the network
       data.frame() %>%
@@
 
+```
+
+#### Rarefied network analysis by colony phase
+```{r, include=FALSE}
+# network level version 2
+rare_net_nt2 <- function(net, n, iter) {
+
+  out <- data.frame()
+
+  t_filter <- net %>%
+    select(site, date, phase, year) %>%
+    unique() %>%
+    group_by(site, phase, year) %>%
+    mutate(transects = n()) %>%
+    filter(transects >= n)
+
+  for(i in 1:iter) {
+    temp <- net %>%
+      semi_join(t_filter) %>%
+      select(site, date, phase, year) %>%
+      unique() %>%
+      group_by(site, phase, year) %>%
+      sample_n(n) %>%
+      left_join(net) %>%
+      group_by(site, phase, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
+      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
+      #filter(freq > 1) %>%
+      unite(webID, c(site, phase, year), sep = "_") %>%
+      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
+      data.frame() %>% # convert to data frame
+      frame2webs() %>% # convert to to bipartite web
+      map(networklevel, weighted = TRUE, level = "higher",
+          index = c("web asymmetry", "weighted NODF", "H2", "C score", "niche overlap", "generality")) %>% 
+      data.frame() %>%
+      t() %>%
+      data.frame() %>%
+      rownames_to_column() %>%
+      dplyr::select(id = rowname, everything()) %>%
+      as_tibble() %>%
+      mutate(iteration = rep(i, n()))
+
+    out <- bind_rows(out, temp)
+  } 
+  out <- out  %>%
+      gather(metric, value, -c(id, iteration)) %>% # get mean value across iterations for each metric
+      group_by(id, metric) %>%
+      summarize(value = mean(value)) %>%
+      separate(id, sep = "_", into = c("site", "phase", "year")) %>%
+      mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
+      left_join(site_data)
+}
+# species level version 2
+rare_net_sp2 <- function(net, n, iter) {
+
+  out <- data.frame()
+
+  t_filter <- net %>%
+    select(site, date, phase, year) %>%
+    unique() %>%
+    group_by(site, phase, year) %>%
+    mutate(transects = n()) %>%
+    filter(transects >= n)
+
+  for(i in 1:iter) {
+    temp <- net %>%
+      semi_join(t_filter) %>%
+      select(site, date, phase, year) %>%
+      unique() %>%
+      group_by(site, phase, year) %>%
+      sample_n(n) %>%
+      left_join(net) %>%
+      group_by(site, phase, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
+      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
+      #filter(freq > 1) %>%
+      unite(webID, c(site, phase, year), sep = "_") %>%
+      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
+      data.frame() %>% # convert to data frame
+      frame2webs() %>% # convert to to bipartite web
+      map(specieslevel, level = "higher", index = c("d", "species strength")) %>%
+      map(rownames_to_column) %>%
+      map(as_tibble) %>%
+      bind_rows(.id = "id") %>% # Collapse list into big data frame
+      rename(sp = rowname) %>%
+      mutate(iteration = rep(i, n()))
+
+    out <- bind_rows(out, temp)
+  } 
+  out <- out %>%
+    gather(metric, value, -c(sp, id, iteration)) %>% # get mean value across iterations for each metric
+    group_by(sp, id, metric) %>%
+    summarize(value = mean(value)) %>%
+    separate(id, sep = "_", into = c("site", "phase", "year")) %>%
+    mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
+    left_join(site_data) %>%
+    left_join(bb_traits, by = c("sp" = "bb.sp"))
+}
 ```
@@
 ```{r eval=FALSE, include=FALSE}
-fl_abund_yday <- gam(fl.abund ~ year +
-                    te(yday, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
+fl_abund_yday <- gamV(fl.abund ~ year +
+                    te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
+                    data = abund_div,
                     family = "tw",
@@
 
-fl_abund_yday.vis <- getViz(fl_abund_yday)
-check.gamViz(fl_abund_yday.vis)
-summary(fl_abund_yday.vis)
-
-plot(fl_abund_yday.vis, allTerms = TRUE) +
-  ggtitle("FL abundance (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
+fl_abund_yday2 <- gamV(log(fl.abund + 0.001) ~ year +
+                    te(yday, elev.mean, by=year, bs= c("cr", "cr"), k=c(10, 10), m=1),
+                    data = abund_div,
+                    #family = "scat",
+                    method = "REML")
+
+fl_abund_yday3 <- gamV(log(fl.abund + 0.001) ~ year +
+                    te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
+                    data = abund_div,
+                    #family = "scat",
+                    method = "REML")
+
+AIC(fl_abund_yday, fl_abund_yday2, fl_abund_yday3)
+
+check.gamViz(fl_abund_yday)
+summary(fl_abund_yday)
+
+print(plot(fl_abund_yday, allTerms = TRUE), pages = 1)
+
+check.gamViz(fl_abund_yday2)
+summary(fl_abund_yday2)
+
+print(plot(fl_abund_yday2, allTerms = TRUE), pages = 1)
+
+check.gamViz(fl_abund_yday3)
+summary(fl_abund_yday3)
+
+print(plot(fl_abund_yday3, allTerms = TRUE), pages = 1)
 
 fl_abund_gdd <- gam(fl.abund ~ year +
-                    te(gdd.cum, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
+                    te(gdd.cum, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
                     data = filter(abund_div),
@@
 fl_rich_yday <- gam(fl.rich ~ year +
-                    te(yday, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
+                    te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
                     data = filter(abund_div),
@@
 plot(fl_rich_yday.vis, allTerms = TRUE) +
-  ggtitle("FL abundance (gdd)") +
+  ggtitle("FL richness (gdd)") +
   xlab("Day of year") +
@@
 fl_rich_yday2 <- gam(fl.rich ~ year +
-                    te(yday, elev.mean, bs= "cc", k= 10, m=1),
+                    te(yday, elev.mean, bs= "tp", k= 10, m=1),
                     data = filter(abund_div),
@@
 fl_rich_gdd <- gam(fl.rich ~ year +
-                    te(gdd.cum, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
+                    te(gdd.cum, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
                     data = filter(abund_div),
@@
 fl_rich_gdd2 <- gam(fl.rich ~ year +
-                    te(gdd.cum, elev.mean, bs= "cc", k= 10, m=1),
+                    te(gdd.cum, elev.mean, bs= "tp", k= 10, m=1),
                     data = filter(abund_div),
@@
 bb_abund_yday <- gam(bb.abund ~ year +
-                    te(yday, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
+                    te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
                     data = filter(abund_div),
@@
 bb_abund_gdd <- gam(bb.abund ~ year +
-                    te(gdd.cum, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
+                    te(gdd.cum, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
                     data = filter(abund_div),
@@
 bb_rich_yday <- gam(bb.rich ~ year +
-                    te(yday, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
+                    te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
                     data = filter(abund_div),
@@
 bb_rich_gdd <- gam(bb.rich ~ year +
-                    te(gdd.cum, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
+                    te(gdd.cum, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
                     data = filter(abund_div),
@@
 ### Rarefied network analysis (not broken down by phase)
+#### Inspect transect counts to set subsampling parameter
 ```{r, include = FALSE}
@@
   facet_wrap(~year)
-
+```
+
+#### Generate network metric tables
+```{r}
 ### Species level
-spmet_2012 <- rare_net_sp(net_2012, 9, 20) %>%
+#### BBs
+spmet_2012_BB <- rare_net_sp_BB(net_2012, 9, 100) %>%
   mutate(year = factor(rep("2012", n())))
-spmet_2011 <- rare_net_sp(net_2011, 4, 20) %>%
+spmet_2011_BB <- rare_net_sp_BB(net_2011, 4, 100) %>%
   mutate(year = factor(rep("2011", n())))  
-spmet_2010 <- rare_net_sp(net_2010, 3, 20) %>%
+spmet_2010_BB <- rare_net_sp_BB(net_2010, 3, 100) %>%
   mutate(year = factor(rep("2010", n())))
 
-spmet <- bind_rows(spmet_2012, spmet_2011, spmet_2010)
-
-spmet_mean <- spmet %>%
+spmet_BB <- bind_rows(spmet_2012_BB, spmet_2011_BB, spmet_2010_BB)
+
+spmet_mean_BB <- spmet_BB %>%
+  group_by(site, year, elev.mean, metric) %>%
+  summarize(value = mean(value, na.rm = TRUE))
+
+#### plants
+spmet_2012_FL <- rare_net_sp_FL(net_2012, 9, 100) %>%
+  mutate(year = factor(rep("2012", n())))
+spmet_2011_FL <- rare_net_sp_FL(net_2011, 4, 100) %>%
+  mutate(year = factor(rep("2011", n())))  
+spmet_2010_FL <- rare_net_sp_FL(net_2010, 3, 100) %>%
+  mutate(year = factor(rep("2010", n())))
+
+spmet_FL <- bind_rows(spmet_2012_FL, spmet_2011_FL, spmet_2010_FL)
+
+spmet_mean_FL <- spmet_FL %>%
   group_by(site, year, elev.mean, metric) %>%
@@
 ### Network- and group-level
-netmet_rare_2012 <- rare_net_nt(net_2012, n = 9, iter = 20) %>%
+netmet_rare_2012 <- rare_net_nt(net_2012, n = 9, iter = 25) %>%
   mutate(year = factor(rep("2012", n())))
-netmet_rare_2011 <- rare_net_nt(net_2011, n = 4, iter = 20)  %>%
+netmet_rare_2011 <- rare_net_nt(net_2011, n = 4, iter = 25)  %>%
   mutate(year = factor(rep("2011", n())))
-netmet_rare_2010 <- rare_net_nt(net_2010, n = 3, iter = 20) %>%
+netmet_rare_2010 <- rare_net_nt(net_2010, n = 3, iter = 25) %>%
   mutate(year = factor(rep("2010", n())))
@@
 netmet <- bind_rows(netmet_rare_2012, 
+                    netmet_rare_2011, 
+                    netmet_rare_2010) %>%
+  mutate(metric = factor(metric, levels = c("H2",
+                                            "weighted.NODF",
+                                            "niche.overlap.HL", 
+                                            "C.score.HL",
+                                            "generality.HL",
+                                            "niche.overlap.LL", 
+                                            "C.score.LL",
+                                            "generality.LL",
+                                            "web.asymmetry",
+                                            "weighted.connectance")))
+```
+
+### Plot network analysis
+```{r}
+#### Per species values
+ggplot(filter(spmet_2012_BB,
+              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf") & metric == "d"), 
+       aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
+  facet_wrap(~sp, scales = "free")
+
+ggplot(filter(spmet_2011_BB,
+              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf") & metric == "d"), 
+       aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
+  facet_wrap(~sp, scales = "free")
+
+ggplot(filter(spmet_2010_BB,
+              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf") & metric == "d"), 
+       aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
+  facet_wrap(~sp, scales = "free")
+
+#### Community mean d
+##### BB
+ggplot(filter(spmet_mean_BB, metric == "d"), aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 10)) +
+  xlab("Elevation") +
+  ylab("Community mean d'") +
+  facet_wrap(~year) +
+  theme_light(16)
+
+gam_d <- gamV(value ~ year + s(elev.mean, k = 15, bs = "tp", by = factor(year)), 
+             data = filter(spmet_BB, metric == "d"),
+             method = "REML")
+
+check.gamViz(gam_d)
+summary(gam_d)
+
+plot(gam_d, allTerms = TRUE)
+
+##### FL
+ggplot(filter(spmet_mean_FL, metric == "d"), aes(elev.mean, value)) +
+  geom_point(aes(color = year)) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
+  xlab("Elevation") +
+  ylab("Community mean d'") +
+  facet_wrap(~year) +
+  theme_light(16)
+
+#### Network level
+
+##### niche overlap
+ggplot(filter(netmet, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x)) +
+  #geom_smooth() +
+  xlab("Elevation") +
+  ylab("Niche overlap") +
+  facet_wrap(~year) +
+  theme_light(16)
+
+gam_niche.overlap <- gamV(value ~ year + s(elev.mean, k = 10, bs = "tp", by = factor(year)), 
+                         data = filter(netmet, metric == "niche.overlap.HL"),
+                         method = "REML")
+
+check.gamViz(gam_niche.overlap)
+summary(gam_niche.overlap)
+
+plot(gam_niche.overlap, allTerms = TRUE)
+
+ggplot(filter(netmet, metric == "niche.overlap.LL"), aes(elev.mean, value)) +
+  geom_point(aes(color = year)) +
+  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
+  #geom_smooth() +
+  xlab("Elevation") +
+  ylab("Niche overlap") +
+  facet_wrap(~year, scales = "free") +
+  theme_light(16)
+
+##### C.score
+ggplot(filter(netmet, metric == "C.score.HL"), aes(elev.mean, value)) +
+  geom_point(aes(color = year)) +
+  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
+  xlab("Elevation") +
+  ylab("C score") +
+  theme_light(16)
+
+ggplot(filter(netmet, metric == "C.score.HL"), aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x)) +
+  xlab("Elevation") +
+  ylab("C score") +
+  facet_wrap(~year) +
+  theme_light(16)
+
+gam_cscore <- gamV(value ~ year + s(elev.mean, bs = "tp", k = 10, by = factor(year)), 
+                  data = filter(netmet, metric == "C.score.HL"),
+                  #family = "scat",
+                  method = "REML")
+
+check.gamViz(gam_cscore)
+summary(gam_cscore)
+
+plot(gam_cscore.vis, allTerms = TRUE)
+
+ggplot(filter(netmet, metric == "C.score.LL"), aes(elev.mean, value)) +
+  geom_point(aes(color = year)) +
+  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
+  #geom_smooth() +
+  xlab("Elevation") +
+  ylab("C score") +
+  facet_wrap(~year, scales = "free") +
+  theme_light(16)
+```
+
+### Rarefied network analysis by k.type instead of by plant sp
+```{r, include = FALSE}
+### Create yearly subsets to feed to rare_net
+net_2012 <- filter(net, year == "2012")
+net_2011 <- filter(net, year == "2011")
+net_2010 <- filter(net, year == "2010")
+
+### Species level
+spmet_2012_k <- rare_net_sp_k.type.s(net_2012, 9, 20) %>%
+  mutate(year = factor(rep("2012", n())))
+spmet_2011_k <- rare_net_sp_k.type.s(net_2011, 4, 20) %>%
+  mutate(year = factor(rep("2011", n())))  
+spmet_2010_k <- rare_net_sp_k.type.s(net_2010, 3, 20) %>%
+  mutate(year = factor(rep("2010", n())))
+
+spmet_k <- bind_rows(spmet_2012_k, spmet_2011_k, spmet_2010_k)
+
+spmet_mean_k <- spmet_k %>%
+  group_by(site, year, elev.mean, metric) %>%
+  summarize(value = mean(value))
+
+### Network- and group-level
+netmet_rare_2012_k <- rare_net_nt_k.type.s(net_2012, n = 9, iter = 20) %>%
+  mutate(year = factor(rep("2012", n())))
+netmet_rare_2011_k <- rare_net_nt_k.type.s(net_2011, n = 4, iter = 20)  %>%
+  mutate(year = factor(rep("2011", n())))
+netmet_rare_2010_k <- rare_net_nt_k.type.s(net_2010, n = 3, iter = 20) %>%
+  mutate(year = factor(rep("2010", n())))
+
+netmet_k <- bind_rows(netmet_rare_2012, 
                     netmet_rare_2011, 
@@
 #### Community mean d
-ggplot(filter(spmet_mean, metric == "d"), aes(elev.mean, value)) +
+ggplot(filter(spmet_mean_k, metric == "d"), aes(elev.mean, value)) +
   geom_point(aes(color = year)) +
@@
   ylab("Community mean d'") +
+  theme_light(16)
+
+ggplot(filter(spmet_mean_k, metric == "d"), aes(elev.mean, value)) +
+  geom_point(aes(color = year)) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
+  xlab("Elevation") +
+  ylab("Community mean d'") +
+  facet_wrap(~year) +
   theme_light(16)
@@
 
+ggplot(filter(netmet, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
+  geom_point(aes(color = year)) +
+  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
+  #geom_smooth() +
+  xlab("Elevation") +
+  ylab("Niche overlap") +
+  facet_wrap(~year) +
+  theme_light(16)
+
 gam_niche.overlap <- gam(value ~ year + s(elev.mean), 
@@
 
-gam_cscore <- gam(value ~ year + s(elev.mean), 
-                  data = filter(netmet, metric == "C.score.HL"),
-                  method = "REML")
-
-gam_cscore.vis <- getViz(gam_cscore)
-check.gamViz(gam_cscore.vis)
-summary(gam_cscore.vis)
-
-plot(gam_cscore.vis, allTerms = TRUE) +
-  xlab("Elevation")
-
-##### Connectance
-ggplot(filter(netmet, metric == "weighted.connectance"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
-  #geom_smooth() +
-  xlab("Elevation") +
-  ylab("Connectance (weighted)") +
-  theme_light(16)
-
-gam_connectance <- gam(value ~ year + s(elev.mean), 
-                  data = filter(netmet, metric == "weighted.connectance"),
-                  method = "REML")
-
-gam_connectance.vis <- getViz(gam_connectance)
-check.gamViz(gam_connectance.vis)
-summary(gam_connectance.vis)
-
-plot(gam_connectance.vis, allTerms = TRUE) +
-  xlab("Elevation")
-
-##### NODF
-ggplot(filter(netmet, metric == "weighted.NODF"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
-  #geom_smooth() +
-  xlab("Elevation") +
-  ylab("NODF (weighted)") +
-  theme_light(16)
-
-gam_NODF <- gam(value ~ year + s(elev.mean), 
-                data = filter(netmet, metric == "weighted.NODF"),
-                method = "REML")
-
-gam_NODF.vis <- getViz(gam_NODF)
-check.gamViz(gam_NODF.vis)
-summary(gam_NODF.vis)
-
-plot(gam_NODF.vis, allTerms = TRUE) +
-  xlab("Elevation")
-
-##### Web asymmetry
-ggplot(filter(netmet, metric == "web.asymmetry"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
-  #geom_smooth() +
-  xlab("Elevation") +
-  ylab("Web asymmetry") +
-  theme_light(16)
-
-gam_webass <- gam(value ~ year + s(elev.mean), 
-                data = filter(netmet, metric == "web.asymmetry"),
-                method = "REML")
-
-gam_webass.vis <- getViz(gam_webass)
-check.gamViz(gam_webass.vis)
-summary(gam_webass.vis)
-
-plot(gam_webass.vis, allTerms = TRUE) +
-  xlab("Elevation")
-```
-
-### Rarefied network analysis by k.type instead of by plant sp
-```{r, include = FALSE}
-### Create yearly subsets to feed to rare_net
-net_2012 <- filter(net, year == "2012")
-net_2011 <- filter(net, year == "2011")
-net_2010 <- filter(net, year == "2010")
-
-### Species level
-spmet_2012 <- rare_net_sp_k.type(net_2012, 9, 20) %>%
-  mutate(year = factor(rep("2012", n())))
-spmet_2011 <- rare_net_sp(net_2011, 4, 20) %>%
-  mutate(year = factor(rep("2011", n())))  
-spmet_2010 <- rare_net_sp(net_2010, 3, 20) %>%
-  mutate(year = factor(rep("2010", n())))
-
-spmet <- bind_rows(spmet_2012, spmet_2011, spmet_2010)
-
-spmet_mean <- spmet %>%
-  group_by(site, year, elev.mean, metric) %>%
-  summarize(value = mean(value))
-
-### Network- and group-level
-netmet_rare_2012 <- rare_net_nt_k.type(net_2012, n = 9, iter = 20) %>%
-  mutate(year = factor(rep("2012", n())))
-netmet_rare_2011 <- rare_net_nt_k.type(net_2011, n = 4, iter = 20)  %>%
-  mutate(year = factor(rep("2011", n())))
-netmet_rare_2010 <- rare_net_nt_k.type(net_2010, n = 3, iter = 20) %>%
-  mutate(year = factor(rep("2010", n())))
-
-netmet <- bind_rows(netmet_rare_2012, 
-                    netmet_rare_2011, 
-                    netmet_rare_2010) %>%
-  mutate(metric = factor(metric, levels = c("H2",
-                                            "weighted.NODF",
-                                            "niche.overlap.HL", 
-                                            "C.score.HL",
-                                            "generality.HL",
-                                            "web.asymmetry",
-                                            "weighted.connectance")))
-
-### Plot
-#### Per species values
-ggplot(filter(spmet_2012,
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
-       aes(elev.mean, value)) +
-  geom_point() +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth() +
-  facet_grid(metric~sp, scales = "free")
-
-ggplot(filter(spmet_2011,
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
-       aes(elev.mean, value)) +
-  geom_point() +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth() +
-  facet_grid(metric~sp, scales = "free")
-
-ggplot(filter(spmet_2010,
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
-       aes(elev.mean, value)) +
-  geom_point() +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth() +
-  facet_grid(metric~sp, scales = "free")
-
-#### Community mean d
-ggplot(filter(spmet_mean, metric == "d"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
-  xlab("Elevation") +
-  ylab("Community mean d'") +
-  theme_light(16)
-
-gam_d <- gam(value ~ year + s(elev.mean, k = 10), 
-             data = filter(spmet, metric == "d"),
-             method = "REML")
-
-gam_d.vis <- getViz(gam_d)
-check.gamViz(gam_d.vis)
-summary(gam_d.vis)
-
-plot(gam_d.vis, allTerms = TRUE) +
-  xlab("Elevation")
-
-#### Network level
-ggplot(netmet, aes(elev.mean, value)) +
-  geom_point() +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth() +
-  facet_grid(metric~year, scales = "free")
-
-##### H2
-ggplot(filter(netmet, metric == "H2"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
-  xlab("Elevation") +
-  ylab("H2'") +
-  theme_light(16)
-
-gam_H2 <- gam(value ~ year + s(elev.mean), 
-              data = filter(netmet, metric == "H2"),
-              method = "REML")
-
-gam_H2.vis <- getViz(gam_H2)
-check.gamViz(gam_H2.vis)
-summary(gam_H2.vis)
-
-plot(gam_H2.vis, allTerms = TRUE) +
-  xlab("Elevation")
-
-##### niche overlap
-ggplot(filter(netmet, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
-  #geom_smooth() +
-  xlab("Elevation") +
-  ylab("Niche overlap") +
-  theme_light(16)
-
-gam_niche.overlap <- gam(value ~ year + s(elev.mean), 
-                         data = filter(netmet, metric == "niche.overlap.HL"),
-                         method = "REML")
-
-gam_niche.overlap.vis <- getViz(gam_niche.overlap)
-check.gamViz(gam_niche.overlap.vis)
-summary(gam_niche.overlap.vis)
-
-plot(gam_niche.overlap.vis, allTerms = TRUE) +
-  xlab("Elevation")
-
-##### C.score
 ggplot(filter(netmet, metric == "C.score.HL"), aes(elev.mean, value)) +
@@
   ylab("C score") +
+  facet_wrap(~year) +
   theme_light(16)
@@
 plot(gam_cscore.vis, allTerms = TRUE) +
-  xlab("Elevation")
-
-##### Connectance
-ggplot(filter(netmet, metric == "weighted.connectance"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
-  #geom_smooth() +
-  xlab("Elevation") +
-  ylab("Connectance (weighted)") +
-  theme_light(16)
-
-gam_connectance <- gam(value ~ year + s(elev.mean), 
-                  data = filter(netmet, metric == "weighted.connectance"),
-                  method = "REML")
-
-gam_connectance.vis <- getViz(gam_connectance)
-check.gamViz(gam_connectance.vis)
-summary(gam_connectance.vis)
-
-plot(gam_connectance.vis, allTerms = TRUE) +
-  xlab("Elevation")
-
-##### NODF
-ggplot(filter(netmet, metric == "weighted.NODF"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
-  #geom_smooth() +
-  xlab("Elevation") +
-  ylab("NODF (weighted)") +
-  theme_light(16)
-
-gam_NODF <- gam(value ~ year + s(elev.mean), 
-                data = filter(netmet, metric == "weighted.NODF"),
-                method = "REML")
-
-gam_NODF.vis <- getViz(gam_NODF)
-check.gamViz(gam_NODF.vis)
-summary(gam_NODF.vis)
-
-plot(gam_NODF.vis, allTerms = TRUE) +
-  xlab("Elevation")
-
-##### Web asymmetry
-ggplot(filter(netmet, metric == "web.asymmetry"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
-  #geom_smooth() +
-  xlab("Elevation") +
-  ylab("Web asymmetry") +
-  theme_light(16)
-
-gam_webass <- gam(value ~ year + s(elev.mean), 
-                data = filter(netmet, metric == "web.asymmetry"),
-                method = "REML")
-
-gam_webass.vis <- getViz(gam_webass)
-check.gamViz(gam_webass.vis)
-summary(gam_webass.vis)
-
-plot(gam_webass.vis, allTerms = TRUE) +
   xlab("Elevation")
@@
 
-### Plot some dang networks
-```{r eval=FALSE, include=FALSE}
-web <- net %>%
-  group_by(year, elev.class2, phase, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-  summarize(freq = n()) %>%
-  unite(webID, c(year, elev.class2, phase), sep = "_") %>% # calculate interaction frequency per species pair within each site
-  dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs()  # convert to to bipartite web
-
-visweb(web[[1]])
-visweb(web[[2]])
-```
-
 ######## Old analyses that I probably don't care to use but don't want to delete yet #########
@@
          total.rich = bb.rich + fl.rich) %>%
-  filter(bb.rich > 1 & fl.rich > 1) %>%
+  filter(bb.rich > 3 & fl.rich > 3) %>%
   unite(webID, c(site, dayofyear, gdd.cum, year, size), sep = "_") %>%
@@
   frame2webs() %>% # convert to to bipartite web
-  map(networklevel, weighted = TRUE, level = "higher", index = c("web asymmetry", "weighted NODF", "niche overlap", "C score", "H2")) %>% # We probably care mainly about the BB level of the network
+  map(networklevel, weighted = TRUE, level = "higher", index = c("niche overlap", "C score", "H2")) %>% # We probably care mainly about the BB level of the network
   data.frame() %>%
@@
          total.rich = bb.rich + fl.rich) %>%
-  filter(bb.rich > 1 & fl.rich > 1) %>%
+  filter(bb.rich > 5 & fl.rich > 5) %>%
   unite(webID, c(site, dayofyear, gdd.cum, year, size), sep = "_") %>%
@@
 
-### Phase x elev network analysis
+### Phase x elev network analysis -- this only makes sense in 2012, where we have sufficient sampling across the whole colony lifecycle
 ```{r eval=FALSE, include=FALSE}
@@
 
-pxe_net <- rare_net_nt2(net, 4, 20) %>%
+pxe_net <- rare_net_nt2(filter(net, year == 2012), 3, 20) %>%
   full_join((pxe_abund_div))
 
-ggplot(filter(pxe_net, 
-              metric == "C.score.HL" & ad.metric == "fl.abund"),
-       aes(log(ad.value), value)) +
-  geom_point() +
-  geom_smooth(method = "lm")  +
+ggplot(filter(pxe_net, metric == "H2"), aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth() +
   facet_grid(year ~ phase, scales = "free") +
-  xlab("log(floral abundance)") +
+  xlab("Elevation") +
+  ylab("H2")
+
+ggplot(filter(pxe_net, metric == "C.score.HL"), aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth() +
+  facet_grid(year ~ phase, scales = "free") +
+  xlab("Elevation") +
   ylab("C.score")
 
-ggplot(filter(pxe_net, 
-              metric == "C.score.HL" & ad.metric == "fl.per.bb"),
-       aes(log(ad.value), value)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
+ggplot(filter(pxe_net, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth() +
   facet_grid(year ~ phase, scales = "free") +
-  xlab("log(per-capita floral abundance)") +
-  ylab("C.score")
-
-ggplot(filter(pxe_net, 
-              metric == "H2" & ad.metric == "fl.rich"),
-       aes(ad.value, value)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_grid(year ~ phase, scales = "free") +
-  xlab("Floral richness") +
-  ylab("H2")
-
-ggplot(filter(pxe_net, 
-              metric == "H2" & ad.metric == "bb.abund"),
-       aes(log(ad.value), value)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_grid(year ~ phase, scales = "free") +
-  xlab("BB abund") +
-  ylab("H2")
-
-ggplot(filter(pxe_net, 
-              metric == "H2" & ad.metric == "bb.rich"),
-       aes(ad.value, value)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_grid(year ~ phase, scales = "free") +
-  xlab("BB richness") +
-  ylab("H2")
-
-ggplot(filter(pxe_net, metric == "weighted.NODF"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(year ~ phase, scales = "free") +
-  xlab("Elevation") +
-  ylab("Weighted NODF")
-
-ggplot(filter(pxe_net, metric == "H2"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(year ~ phase, scales = "free") +
-  xlab("Elevation") +
-  ylab("H2")
-
-ggplot(filter(pxe_net, metric == "C.score.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(year ~ phase, scales = "free") +
-  xlab("Elevation") +
-  ylab("C.score")
-
-ggplot(filter(pxe_net, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(year ~ phase, scales = "free") +
   xlab("Elevation") +
   ylab("Niche overlap")
-
-ggplot(filter(pxe_net, metric == "generality.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(year ~ phase, scales = "free") +
-  xlab("Elevation") +
-  ylab("Generality")
-
-ggplot(filter(pxe_net, metric == "web.asymmetry"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(year ~ phase, scales = "free") +
-  xlab("Elevation") +
-  ylab("Web asymmetry")
 

### 2020-06-12 data_proc.R

@@ -131,3 +131,139 @@ write_csv(climate, "./processed_data/climate.csv")
 
 ### TRY data
 try <- read_delim("./raw_data/TRY_13_05_2020.txt", delim = "\t")
+
+### Parasite data
+#### 2010
+parasites2010 <- read_delim("./raw_data/lab-data_parasites_2010_raw-data.csv", delim = ";") %>%
+  select(-c(year, ovary_development, fatbody_look, comment, dissector)) %>%
+  gather(var, value, -sample_number) %>%
+  mutate(value = as.numeric(value),
+         var = str_replace_all(var, "-", "_")) %>%
+  na.omit() %>%
+  spread(var, value) %>%
+  filter(!is.na(size)) %>% #remove the few cases where size was not measured 
+  replace(is.na(.), 0) %>%
+  mutate(ecto.mites = ecto_mites_l + ecto_mites_m + ecto_mites_s + ecto_mites_xs,
+         conopids = parasitoid_conopidae_dead_larvae +
+           parasitoid_conopidae_dead_larvae +
+           parasitoid_conopidae_dorsal_egg +
+           parasitoid_conopidae_dorsal_larvae +
+           parasitoid_conopidae_ventral_egg +
+           parasitoid_conopidae_ventral_larvae) %>%
+  select(sample_number, size, ecto.mites, conopids, 
+         nosema = parasite_nosema, apicystis = parasite_apicystis,
+         crithidia = parasite_crithidia, braconids = parasitoid_braconidae_larvae,
+         tracheal.mites = tracheal_mites, nematodes = nematode) %>%
+  gather(var, value, -sample_number, -size) %>%
+  mutate(value = as.numeric(value),
+         bin.value = case_when(
+           value > 0 ~ TRUE,
+           value == 0 ~ FALSE
+         )) %>%
+  na.omit() 
+  
+sampling2010 <- read_delim("./raw_data/field-data_bumblebees_2010_raw-data.csv", delim = ";") %>%
+  select(site = site_name, elev.mean = site_elevation, year, date, yday = dayofyear, bb.sp = bumb_species, 
+         caste, plant.sp = forage_plant, sample_number) %>%
+  mutate(date = dmy(date),
+         caste = str_trim(caste))
+
+psite_2010 <- parasites2010 %>%
+  left_join(sampling2010, by = c("sample_number")) %>%
+  select(site, date, year, yday, elev.mean, bb.sp, size, caste, 
[…]

### 2020-06-12 analysis.Rmd
further refined GAM analysis, exploring the use of year as a random effect smooth instead of parametric term

@@
   group_by(site, year, elev.mean, metric) %>%
-  summarize(value = mean(value, na.rm = TRUE))
+  summarize(value = mean(value, na.rm = TRUE)) %>%
+  ungroup() %>%
+  mutate(year = factor(year))
 
@@
   group_by(site, year, elev.mean, metric) %>%
-  summarize(value = mean(value))
+  summarize(value = mean(value, na.rm = TRUE)) %>%
+  ungroup() %>%
+  mutate(year = factor(year))
 
@@
                                             "web.asymmetry",
-                                            "weighted.connectance")))
+                                            "weighted.connectance"))) %>%
+  ungroup() %>%
+  mutate(year = factor(year))
 ```
@@
   geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10)) +
+  #geom_smooth(method = "gam", formula = y ~ s(x, k = 10)) +
   xlab("Elevation") +
@@
 
-gam_d <- gamV(value ~ year + s(elev.mean, k = 15, bs = "tp", by = factor(year)), 
-             data = filter(spmet_BB, metric == "d"),
+gam_d1 <- gamV(value ~ s(year, bs = "re") + s(elev.mean, k = 5, bs = "tp", by = year, m = 1), 
+             data = filter(spmet_mean_BB, metric == "d"),
              method = "REML")
 
-check.gamViz(gam_d)
-summary(gam_d)
-
-plot(gam_d, allTerms = TRUE)
+check.gamViz(gam_d1)
+summary(gam_d1)
+plot(gam_d1, allTerms = TRUE)
+
+summary(lm(value ~ elev.mean, data =  filter(spmet_mean_BB, metric == "d" & year == 2012)))
+plot(lm(value ~ elev.mean, data =  filter(spmet_mean_BB, metric == "d" & year == 2012)))
 
@@
 
-gam_niche.overlap <- gamV(value ~ year + s(elev.mean, k = 10, bs = "tp", by = factor(year)), 
+gam_niche.overlap <- gamV(value ~ s(year, bs = "re") + s(elev.mean, by = year, k = 10, bs = "tp", m = 2), 
                          data = filter(netmet, metric == "niche.overlap.HL"),
@@
 
-ggplot(filter(netmet, metric == "niche.overlap.LL"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
-  #geom_smooth() +
-  xlab("Elevation") +
-  ylab("Niche overlap") +
-  facet_wrap(~year, scales = "free") +
-  theme_light(16)
-
 ##### C.score
-ggplot(filter(netmet, metric == "C.score.HL"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
-  xlab("Elevation") +
-  ylab("C score") +
-  theme_light(16)
-
 ggplot(filter(netmet, metric == "C.score.HL"), aes(elev.mean, value)) +
@@
 
-gam_cscore <- gamV(value ~ year + s(elev.mean, bs = "tp", k = 10, by = factor(year)), 
+gam_cscore <- gamV(value ~ s(year, bs = "re") + s(elev.mean, by = factor(year), bs = "tp", k = 6, m = 1), 
                   data = filter(netmet, metric == "C.score.HL"),
@@
 
-plot(gam_cscore.vis, allTerms = TRUE)
-
-ggplot(filter(netmet, metric == "C.score.LL"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
-  #geom_smooth() +
-  xlab("Elevation") +
-  ylab("C score") +
-  facet_wrap(~year, scales = "free") +
-  theme_light(16)
+plot(gam_cscore, allTerms = TRUE)
 ```

### 2020-07-02 analyses/analysis.Rmd
broke down analyses into separate scripts; did some other things

--- 2020-06-12 analysis.Rmd
@@
 
-breaks_2012_yday <- breakpoints_yday(castes, 2012, t1 = 0.75, t2 = 0.15)
-breaks_2011_yday <- breakpoints_yday(castes, 2011, t1 = 0.75, t2 = 0.15)
-breaks_2010_yday <- breakpoints_yday(castes, 2010, t1 = 0.75, t2 = 0.15)
+breaks_2012_yday <- breakpoints_yday(castes, 2012, t1 = 0.5, t2 = 0.15)
+breaks_2011_yday <- breakpoints_yday(castes, 2011, t1 = 0.5, t2 = 0.15)
+breaks_2010_yday <- breakpoints_yday(castes, 2010, t1 = 0.5, t2 = 0.15)
 
@@
 ```
+
 ### BB and floral turnover ~ elevation
@@
            elev.mean, break1.gdd, break2.gdd) %>%
-  summarize(fl.abund = sum(flower.cover))
+  summarize(fl.abund = sum(flower.cover)) %>%
+  rename("yday" = "dayofyear")
 
@@
            elev.mean, break1.gdd, break2.gdd) %>%
-  summarize(fl.rich = n())
+  summarize(fl.rich = n()) %>%
+  rename("yday" = "dayofyear")
 
@@
 ```{r eval=FALSE, include=FALSE}
-fl_abund_yday <- gamV(fl.abund ~ year +
+fl_abund_yday <- gamV(log(fl.abund + 0.001) ~ year +
+                    te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
+                    data = fl_abund,
+                    method = "REML")
+
+check.gamViz(fl_abund_yday)
+summary(fl_abund_yday)
+print(plot(fl_abund_yday, allTerms = TRUE), pages = 1)
+
+fl_abund_gdd <- gamV(log(fl.abund + 0.001) ~ year +
+                    te(gdd.cum, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
+                    data = fl_abund,
+                    method = "REML")
+
+check.gamViz(fl_abund_gdd)
+summary(fl_abund_gdd)
+print(plot(fl_abund_gdd, allTerms = TRUE), pages = 1)
+```
+
+##### Floral richness -- since year does not seem to change the overall pattern (though perhaps the magnitude), I could use a simpler model, perhaps with year as a random intercept effect; some explanat
+```{r eval=FALSE, include=FALSE}
+fl_rich_yday <- gamV(fl.rich ~ year +
+                    te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
+                    data = fl_rich,
+                    method = "REML")
+
+check.gamViz(fl_rich_yday)
+summary(fl_rich_yday)
+print(plot(fl_rich_yday, allTerms = TRUE), pages = 1)
+
+fl_rich_gdd <- gamV(fl.rich ~ year +
+                    te(gdd.cum, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
+                    data = fl_rich,
+                    method = "REML")
+
+check.gamViz(fl_rich_gdd)
+summary(fl_rich_gdd)
+print(plot(fl_rich_gdd, allTerms = TRUE), pages = 1)
+
+```
+
+##### BB abundance
+```{r eval=FALSE, include=FALSE}
+bb_abund_yday <- gamV(sqrt(bb.abund) ~ year +
                     te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
                     data = abund_div,
-                    family = "tw",
                     method = "REML")
 
-fl_abund_yday2 <- gamV(log(fl.abund + 0.001) ~ year +
-                    te(yday, elev.mean, by=year, bs= c("cr", "cr"), k=c(10, 10), m=1),
+check.gamViz(bb_abund_yday)
+summary(bb_abund_yday))
+print(plot(bb_abund_yday, allTerms = TRUE), pages = 1)
+
+bb_abund_gdd <- gamV(sqrt(bb.abund) ~ year +
+                    te(gdd.cum, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
                     data = abund_div,
-                    #family = "scat",
                     method = "REML")
 
-fl_abund_yday3 <- gamV(log(fl.abund + 0.001) ~ year +
+check.gamViz(bb_abund_gdd)
+summary(bb_abund_gdd)
+print(plot(bb_abund_gdd, allTerms = TRUE), pages = 1)
+```
+
+##### BB abundance with species as a "by" factor smooth
+##### I am using family = "poisson"; this results in somewhat overdispersed models, but with a higher variance explained than when I use quasipoisson or negative binomial distributions, and the latter two distributions seem to introduce skew.
+I am more concerned with maximizing the explanatory power of my models than I am with having accurate p-values.
+We'll see what reviewers think, I guess.
+```{r eval=FALSE, include=FALSE}
+### Species-level abundance
+bb_abund_sp <- net %>%
+  group_by(site, date, dayofyear, gdd.cum, year, elev.class2, 
+           elev.mean, break1.gdd, break2.gdd, bb.sp) %>%
+  summarize(bb.abund = n()) %>%
+  rename(yday = dayofyear) %>%
+  ungroup() %>%
+  mutate(bb.sp = factor(bb.sp)) %>%
+  filter(bb.sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")) # we only want the common species I think
+
+### 2010
+bb_abund_sp_yday_2010_nb <- gamV(bb.abund ~ te(yday, elev.mean, 
+                                             by= bb.sp, 
+                                             bs= c("tp", "tp"), 
+                                             k=c(8, 8), m=1),
+                    data = filter(bb_abund_sp, year == "2010"),
+                    family = negbin(10),
+                    method = "REML")
+
+bb_abund_sp_yday_2010 <- gamV(bb.abund ~ te(yday, elev.mean, 
+                                             by= bb.sp, 
+                                             bs= c("tp", "tp"), 
+                                             k=c(8, 8), m=1),
+                    data = filter(bb_abund_sp, year == "2010"),
+                    family = "poisson",
+                    method = "REML")
+
+check.gamViz(bb_abund_sp_yday_2010)
+summary(bb_abund_sp_yday_2010)
+print(plot(bb_abund_sp_yday_2010, allTerms = TRUE), pages = 1)
+
+### 2011
+bb_abund_sp_yday_2011 <- gamV(bb.abund ~ te(yday, elev.mean, 
+                                             by= bb.sp, 
+                                             bs= c("tp", "tp"), 
+                                             k=c(8, 8), m=1),
+                    data = filter(bb_abund_sp, year == "2011"),
+                    family = "poisson",
+                    method = "REML")
+
+check.gamViz(bb_abund_sp_yday_2011)
+summary(bb_abund_sp_yday_2011)
+print(plot(bb_abund_sp_yday_2011, allTerms = TRUE), pages = 1)
+
+### 2012
+bb_abund_sp_yday_2012 <- gamV(bb.abund ~ te(yday, elev.mean, 
+                                             by= bb.sp, 
+                                             bs= c("tp", "tp"), 
+                                             k=c(8, 8), m=1),
+                    data = filter(bb_abund_sp, year == "2012"),
+                    family = "poisson",
+                    method = "REML")
+
+check.gamViz(bb_abund_sp_yday_2012)
+summary(bb_abund_sp_yday_2012)
+print(plot(bb_abund_sp_yday_2012, allTerms = TRUE), pages = 1)
+
+### This time, we'll try modeling the interaction of year and bb.sp in one model. bam() + discrete=TRUE + fREML might speed things up. I hope.
+
+bb_abund_sp_yday_pool <- bam(bb.abund ~ year + te(yday, elev.mean, 
+                                          by= interaction(bb.sp, year), 
+                                          bs= c("tp", "tp"), 
+                                          k=c(10, 10), m=1),
+                    data = bb_abund_sp,
+                    discrete = TRUE,
+                    family = "poisson",
+                    method = "fREML")
+
+bb_abund_sp_yday_pool_vis <- getViz(bb_abund_sp_yday_pool)
+check.gamViz(bb_abund_sp_yday_pool_vis)
+summary(bb_abund_sp_yday_pool_vis)
+print(plot(bb_abund_sp_yday_pool_vis, allTerms = TRUE), pages = 19)
+```
+
+##### FL abundance by k.type
+##### Need to go back through to simplify, include only ktypes that are used heavily
+```{r}
+fl_abund_ktype <- survey %>%
+  group_by(site, date, dayofyear, gdd.cum, year, elev.class2, 
+           elev.mean, break1.gdd, break2.gdd, k.type.ss) %>%
+  summarize(fl.abund = sum(flower.cover)) %>%
+  rename("yday" = "dayofyear")
+
+fl_abund_ktype_yday_pool <- bam(log(fl.abund + 0.001) ~ year + te(yday, elev.mean, 
+                                          by= interaction(k.type.ss, year), 
+                                          bs= c("tp", "tp"), 
+                                          k=c(10, 10), m=1),
+                    data = fl_abund_ktype,
+                    discrete = TRUE,
+                    #family = "poisson",
+                    method = "fREML")
+
+fl_abund_ktype_yday_pool_vis <- getViz(fl_abund_ktype_yday_pool)
+check.gamViz(fl_abund_ktype_yday_pool_vis)
+summary(fl_abund_ktype_yday_pool_vis)
+print(plot(fl_abund_ktype_yday_pool_vis, allTerms = TRUE), pages = 10)
+```
+
+##### BB richness
+```{r eval=FALSE, include=FALSE}
+bb_rich_yday <- gamV(bb.rich ~ year +
                     te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
                     data = abund_div,
-                    #family = "scat",
+                    family = "poisson",
                     method = "REML")
 
-AIC(fl_abund_yday, fl_abund_yday2, fl_abund_yday3)
-
-check.gamViz(fl_abund_yday)
-summary(fl_abund_yday)
-
-print(plot(fl_abund_yday, allTerms = TRUE), pages = 1)
-
-check.gamViz(fl_abund_yday2)
-summary(fl_abund_yday2)
-
-print(plot(fl_abund_yday2, allTerms = TRUE), pages = 1)
-
-check.gamViz(fl_abund_yday3)
-summary(fl_abund_yday3)
-
-print(plot(fl_abund_yday3, allTerms = TRUE), pages = 1)
-
-fl_abund_gdd <- gam(fl.abund ~ year +
+check.gamViz(bb_rich_yday)
+summary(bb_rich_yday)
+print(plot(bb_rich_yday, allTerms = TRUE), pages = 1)
+
+ggplot(abund_div, aes(elev.mean, bb.rich)) +
+  geom_point() +
+  facet_wrap(~year)
+
+ggplot(abund_div, aes(gdd.cum, bb.rich)) +
+  geom_point() +
+  facet_wrap(~year)
+
+ggplot(abund_div, aes(year, bb.rich)) +
+  geom_boxplot() +
+  facet_wrap(~year)
+
+bb_rich_gdd <- gamV(bb.rich ~ year +
                     te(gdd.cum, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    family = "tw",
+                    data = abund_div,
+                    family = "quasipoisson",
                     method = "REML")
 
-fl_abund_gdd.vis <- getViz(fl_abund_gdd)
-check.gamViz(fl_abund_gdd.vis)
-summary(fl_abund_gdd.vis)
-
-plot(fl_abund_gdd.vis, allTerms = TRUE) +
-  ggtitle("FL abundance (gdd)") +
-  xlab("GDD") +
-  ylab("Elevation")
-```
-
-##### Floral richness -- since year does not seem to change the overall pattern (though perhaps the magnitude), I could use a simpler model, perhaps with year as a random intercept effect; some explanat
-```{r eval=FALSE, include=FALSE}
-### By year
-fl_rich_yday <- gam(fl.rich ~ year +
+check.gamViz(bb_rich_gdd)
+summary(bb_rich_gdd)
+print(plot(bb_rich_gdd, allTerms = TRUE), pages = 1)
+```
+
+##### FL.per.BB
+```{r}
+fl.per.bb_yday <- gamV(log(fl.per.bb) ~ year +
                     te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
+                    data = abund_div,
                     method = "REML")
 
-fl_rich_yday.vis <- getViz(fl_rich_yday)
-check.gamViz(fl_rich_yday.vis)
-summary(fl_rich_yday.vis)
-
-plot(fl_rich_yday.vis, allTerms = TRUE) +
-  ggtitle("FL richness (gdd)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### With year as parametric term
-fl_rich_yday2 <- gam(fl.rich ~ year +
-                    te(yday, elev.mean, bs= "tp", k= 10, m=1),
-                    data = filter(abund_div),
+check.gamViz(fl.per.bb_yday)
+summary(fl.per.bb_yday)
+print(plot(fl.per.bb_yday, allTerms = TRUE), pages = 1)
+
+fl.per.bb_gdd <- gamV(log(fl.per.bb) ~ year +
+                    te(gdd.cum, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
+                    data = abund_div,
                     method = "REML")
 
-fl_rich_yday2.vis <- getViz(fl_rich_yday2)
-check.gamViz(fl_rich_yday2.vis)
-summary(fl_rich_yday2.vis)
-
-plot(sm(fl_rich_yday2.vis, 1)) + 
-  l_fitRaster() + 
-  l_fitContour() + 
-  l_points(size = 1) +
-  ggtitle("FL richness (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-plot(pterm(fl_rich_yday2.vis, 1))
-
-### By year
-fl_rich_gdd <- gam(fl.rich ~ year +
-                    te(gdd.cum, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-fl_rich_gdd.vis <- getViz(fl_rich_gdd)
-check.gamViz(fl_rich_gdd.vis)
-summary(fl_rich_gdd.vis)
-
-plot(fl_rich_gdd.vis, allTerms = TRUE) +
-  ggtitle("FL richness (gdd)") +
-  xlab("GDD") +
-  ylab("Elevation")
-
-### With year as parametric term
-fl_rich_gdd2 <- gam(fl.rich ~ year +
-                    te(gdd.cum, elev.mean, bs= "tp", k= 10, m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-fl_rich_gdd2.vis <- getViz(fl_rich_gdd2)
-check.gamViz(fl_rich_gdd2.vis)
-summary(fl_rich_gdd2.vis)
-
-plot(sm(fl_rich_gdd2.vis, 1)) + 
-  l_fitRaster() + 
-  l_fitContour() + 
-  l_points(size = 1) +
-  ggtitle("FL richness (gdd)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-plot(pterm(fl_rich_gdd2.vis, 1))
-```
-
-##### BB abundance
-```{r eval=FALSE, include=FALSE}
-bb_abund_yday <- gam(bb.abund ~ year +
-                    te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-bb_abund_yday.vis <- getViz(bb_abund_yday)
-check.gamViz(bb_abund_yday.vis)
-summary(bb_abund_yday.vis)
-
-plot(bb_abund_yday.vis, allTerms = TRUE) +
-  ggtitle("BB abundance (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-bb_abund_gdd <- gam(bb.abund ~ year +
-                    te(gdd.cum, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-bb_abund_gdd.vis <- getViz(bb_abund_gdd)
-check.gamViz(bb_abund_gdd.vis)
-summary(bb_abund_gdd.vis)
-
-plot(bb_abund_gdd.vis, allTerms = TRUE) +
-  ggtitle("BB abundance (gdd)") +
-  xlab("GDD") +
-  ylab("Elevation")
-```
-
-##### BB richness
-```{r eval=FALSE, include=FALSE}
-bb_rich_yday <- gam(bb.rich ~ year +
-                    te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-bb_rich_yday.vis <- getViz(bb_rich_yday)
-check.gamViz(bb_rich_yday.vis)
-summary(bb_rich_yday.vis)
-plot(sm(bb_rich_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("bb richness 2010 (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-plot(sm(bb_rich_yday.vis, 2)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("bb richness 2011 (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-plot(sm(bb_rich_yday.vis, 3)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("bb richness 2012 (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-bb_rich_gdd <- gam(bb.rich ~ year +
-                    te(gdd.cum, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-bb_rich_gdd.vis <- getViz(bb_rich_gdd)
-check.gamViz(bb_rich_gdd.vis)
-summary(bb_rich_gdd.vis)
-
-plot(bb_rich_gdd.vis, allTerms = TRUE) +
-  ggtitle("BB richness (gdd)") +
-  xlab("Day of year") +
-  ylab("Elevation")
+check.gamViz(fl.per.bb_gdd)
+summary(fl.per.bb_gdd)
+print(plot(fl.per.bb_gdd, allTerms = TRUE), pages = 1)
 ```
@@
   frame2webs()
-
-visweb(web[[1]])
-plotweb(web[[1]])
 

### 2020-07-02 analyses/beta_diversity.Rmd
broke down analyses into separate scripts; did some other things

---
title: "Beta Diversity through Time and Space"
output: html_document
---

### Prepare environment
```{r message=FALSE, warning=FALSE}
library(vegan)
library(lubridate)
library(gdm)
library(bipartite)
library(sf)
library(mgcv)
library(mgcViz)
library(tidyverse)
```

### Climate data
```{r, include=FALSE, echo=FALSE}
gdd <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year),
         yday = yday(date)) %>%
  dplyr::select(site, year, date, yday, gdd.cum) %>%
  semi_join(read_csv("../processed_data/network.csv"), by = "site")

gdd_total <- gdd %>%
  group_by(site, year) %>%
  summarize(gdd.total = max(gdd.cum))
```

### Load site data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) %>% # turn elav.class into an ordered factor %>%
  semi_join(read_csv("../processed_data/network.csv"), by = "site")  %>%
  left_join(gdd_total, by = "site")

### A quick little analysis
ggplot(site_data, aes(elev.mean, gdd.total, color = year)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  ylab("Total GDD")

# intercept only model
gdd_year.gam1 <- gam(gdd.total ~ s(elev.mean) + year,
                    data = site_data) %>% getViz()
summary(gdd_year.gam1)
check.gamViz(gdd_year.gam1)
plot(gdd_year.gam1)

# factor smooth model
gdd_year.gam2 <- gam(gdd.total ~ s(elev.mean, year, bs = "fs", k = 5),
                    data = site_data) %>% getViz()
summary(gdd_year.gam2)
check.gamViz(gdd_year.gam2)
plot(gdd_year.gam2)

# AIC
AIC(gdd_year.gam1, gdd_year.gam2)

site_pairwise_distance_t <- site_data %>%
  ungroup() %>%
  dplyr::select(site, lat, lon) %>%
  unique() %>%
  st_as_sf(coords = c("lon", "lat")) %>%
  st_set_crs(4326) %>%
  st_distance() %>%
  data.frame()

colnames(site_pairwise_distance_t) <- unique(site_data$site)
rownames(site_pairwise_distance_t) <- unique(site_data$site)

site_pairwise_distance <- site_pairwise_distance_t %>%
  rownames_to_column("site1") %>%
  pivot_longer(names_to = "site2", values_to = "distance", -site1)
```

### Load network and survey data and generate site.year and site.month.year webs
```{r}
bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
  dplyr::select(subgenus, bb.sp, pbl.w, pbl.w.class)

fl_traits <- read.csv("../processed_data/floral_k_type_nom_fix.csv") %>%
  dplyr::select(plant.sp, k.type) %>%
  mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"),
         k.type.ss = str_extract(k.type, "[-+]?[0-9]*"))

fl_tax <- read_csv("../processed_data/floral_tax.csv")

breaks <- read_csv("./output/breaks.csv") %>%
  mutate(year = factor(year))

net <- read_csv("../processed_data/network.csv") %>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  unite(site.date, site, date, sep = "_", remove = FALSE) %>%
  mutate(year = factor(year)) %>% # convert year to factor
  full_join(bb_traits, by = c("bb.sp" = "bb.sp")) %>%
  left_join(site_data, by = "site") %>%
  left_join(gdd, by = c("site", "date")) %>%
  left_join(breaks, by = c("site", "year")) %>%
  left_join(fl_traits) %>%
  left_join(fl_tax, by = "plant.genus") %>%
  ungroup() %>%
  mutate(phase = case_when(
    gdd.cum < break1.gdd ~ "founding",
    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
    gdd.cum > break2.gdd ~ "reproductive"
  )) %>%
  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
  arrange(site, date)

web_site.year <- net %>%
  group_by(site.year, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>% 
  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() # convert to to bipartite web

web_site <- net %>%
  group_by(site, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>% 
  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() # convert to to bipartite web

survey <- read_csv("../processed_data/floral_survey.csv") %>%
  mutate(year = factor(year)) %>% # convert year to factor
  semi_join(net, by = "plant.sp") %>% # Consider only plants visited at a given site-date
  left_join(gdd, by = c("site", "date", "year")) %>%
  left_join(breaks, by = c("site", "year")) %>%
  left_join(site_data, by = "site") %>%
  left_join(fl_traits) %>%
  left_join(fl_tax) %>%
  ungroup() %>%
  mutate(phase = case_when(
    gdd.cum < break1.gdd ~ "founding",
    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
    gdd.cum > break2.gdd ~ "reproductive"
  )) %>%
  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
  arrange(site, date)
```

### Tabulate webs
S	= beta_S, the dissimilarity in species composition
OS = beta_OS, the dissimilarity (component) explained by "rewiring" among shared species
WN = beta_WN, the dissimilarity between the two networks
ST = beta_ST, the dissimilarity (component) explained by difference in species community composition 
```{r}
array_site <- webs2array(web_site)
array_site.year <- webs2array(web_site.year)

beta_site <- betalinkr_multi(array_site, partition.st = TRUE, partition.rr = TRUE) %>%
  as_tibble() %>%
  rename(site1 = i) %>%
  rename(site2 = j) %>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, site2, elev1, elev2, elev.diff, 
                           man1, man2, man.diff, mean.elev, lon1, lat1, lon2, lat2)) %>%
  left_join(site_pairwise_distance) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
  ))

beta_site.year <- betalinkr_multi(array_site.year, partition.st = TRUE, partition.rr = TRUE) %>%
  as_tibble() %>%
  separate(i, into = c("site1", "year1"), sep = "_") %>%
  separate(j, into = c("site2", "year2"), sep = "_") %>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, year1, site2, year2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, year1, site2, year2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, year1, site2, year2, elev1, elev2, elev.diff, 
                           man1, man2, man.diff, mean.elev, lon1, lat1, lon2, lat2))
```

# Beta diversity by elevation difference
```{r}
# No apparent differences across years, so we will use the pooled data
ggplot(filter(beta_site.year, year1 == year2), 
       aes(metric, value, color = year1)) +
  geom_boxplot() +
  facet_wrap(~metric, scales = "free")

# And not much of a pattern in response to geographical distance
ggplot(beta_site, 
       aes(as.numeric(distance), value, color = metric)) +
  geom_smooth()

# Interaction beta diversity by elevation difference
ggplot(beta_site, 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  facet_wrap(~metric.class)
```

### Let's look at floral survey diversity patterns
```{r}
survey_bioData <- survey %>%
  filter(flower.cover > 0) %>%
  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
  group_by(site, plant.sp) %>%
  summarize(cover = mean(flower.cover)) %>%
  spread(plant.sp, cover) %>%
  replace(is.na(.), 0)

predData <- site_data %>%
  ungroup() %>%
  dplyr::select(site, elev.mean, lat, lon) %>%
  unique()

survey_sitepair <- formatsitepair(survey_bioData, bioFormat = 1, siteColumn = "site", dist = "jaccard",
                              XColumn = "lon", YColumn = "lat", predData = predData)

### Call gdm()
gdm_survey <- gdm(survey_sitepair, geo = TRUE)
gdm.varImp(survey_sitepair, geo = TRUE)

summary(gdm_survey)
plot(gdm_survey)
plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
```

### And now Bombus diversity patterns
```{r}
BB_bioData <- net %>%
  group_by(site, date, bb.sp) %>%
  summarize(count = n()) %>%
  group_by(site, date) %>%
  mutate(prop = count/sum(count)) %>%
  group_by(site, bb.sp) %>%
  summarize(abund = mean(prop)) %>%
  spread(bb.sp, abund) %>%
  replace(is.na(.), 0)

predData <- site_data %>%
  ungroup() %>%
  dplyr::select(site, elev.mean, lat, lon) %>%
  unique()

bb_sitepair <- formatsitepair(BB_bioData, bioFormat = 1, siteColumn = "site", dist = "jaccard",
                              XColumn = "lon", YColumn = "lat", predData = predData)

### Call gdm()
gdm_bb <- gdm(bb_sitepair, geo = TRUE)
gdm.varImp(bb_sitepair, geo = TRUE)

summary(gdm_bb)
plot(gdm_bb)
plotUncertainty(bb_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
```

### Now let's look at interaction beta diversity
```{r}
net_sitepair <- beta_site %>%
  filter(metric == "WN") %>%
  mutate(weights = rep(1, n())) %>%
  dplyr::select(distance = value, weights,
         s1.xCoord = lon1, s1.yCoord = lat1, 
         s2.xCoord = lon2, s2.yCoord = lat2, 
         s1.elev.mean = elev1, s2.elev.mean = elev2) %>%
  data.frame()

### Call gdm()
gdm_net <- gdm(net_sitepair, geo = TRUE)
gdm.varImp(net_sitepair, geo = TRUE)

summary(gdm_net)
plot(gdm_net)
plotUncertainty(net_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
```

### 2020-07-02 analyses/climate.Rmd
broke down analyses into separate scripts; did some other things

---
title: "climate"
output: html_document
---

### Prepare environment
```{r}
library(tidyverse)
```

### Climate data
```{r, include=FALSE}
samples <- read_csv("../processed_data/network.csv")

climate <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year),
         yday = yday(date)) %>%
  dplyr::select(site, date, yday, year, gdd.cum)
```

### Plot
```{r}
ggplot(climate, aes(yday, gdd.cum, color = year)) +
  geom_point(alpha = 0.01) +
  geom_smooth()
  #facet_wrap(~site) +
  xlim(c(83, 262))

```

### 2020-07-02 analyses/colony_phase.Rmd
broke down analyses into separate scripts; did some other things

--- 2020-06-12 analysis.Rmd
@@
 ---
-title: "analysis"
+title: "colony_phase.Rmd"
 output: html_document
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(include = FALSE)
-library(bipartite)
-library(mgcv)
-library(mgcViz)
-#library(lme4)
-#library(lmerTest)
-library(taxize)
 library(tidyverse)
 library(lubridate)
-library(GGally)
 ```
@@
 ### Define functions
-#### Colony phase breakpoint functions
 ```{r include=FALSE}
@@
 
-#### Rarefied network analysis
-```{r, include=FALSE}
-# Rare_net function: rarefies networks by iteratively subsampling more frequently sampled networks to match the transect count of the least frequently sampled one
-# species level
-rare_net_sp_BB <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      select(site, date) %>%
-      unique() %>%
-      group_by(site) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      group_by(site) %>%
-      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
-      ungroup() %>%
-      filter(freq > 1) %>% # remove singletons
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(specieslevel, level = "higher", index = c("d", "species strength", "proportional generality")) %>%
-      map(rownames_to_column) %>%
-      map(as_tibble) %>%
-      bind_rows(.id = "site") %>% # Collapse list into big data frame
-      rename(sp = rowname) %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out %>%
-    gather(metric, value, -c(sp, site, iteration)) %>% # get mean value across iterations for each metric
-    group_by(sp, site, metric) %>%
-    summarize(value = mean(value)) %>%
-    left_join(site_data) %>%
-    left_join(bb_traits, by = c("sp" = "bb.sp"))
-}
-
-rare_net_sp_FL <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      select(site, date) %>%
-      unique() %>%
-      group_by(site) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      group_by(site) %>%
-      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
-      ungroup() %>%
-      filter(freq > 1) %>% # remove singletons
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(specieslevel, level = "lower", index = c("d", "species strength", "proportional generality")) %>%
-      map(rownames_to_column) %>%
-      map(as_tibble) %>%
-      bind_rows(.id = "site") %>% # Collapse list into big data frame
-      rename(sp = rowname) %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out %>%
-    gather(metric, value, -c(sp, site, iteration)) %>% # get mean value across iterations for each metric
-    group_by(sp, site, metric) %>%
-    summarize(value = mean(value)) %>%
-    left_join(site_data) %>%
-    left_join(fl_traits, by = c("sp" = "plant.sp"))
-}
-
-# network level
-rare_net_nt <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      select(site, date) %>%
-      unique() %>%
-      group_by(site) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      group_by(site) %>%
-      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
-      ungroup() %>%
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(networklevel, weighted = TRUE, index = c("H2", "weighted NODF", "niche overlap", "C score", "generality", "web asymmetry", "weighted connectance")) %>% # We probably care mainly about the BB level of the network
-      data.frame() %>%
-      t() %>%
-      data.frame() %>%
-      rownames_to_column() %>%
-      dplyr::select(site = rowname, everything()) %>%
-      as_tibble() %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out  %>%
-      gather(metric, value, -c(site, iteration)) %>% # get mean value across iterations for each metric
-      group_by(site, metric) %>%
-      summarize(value = mean(value)) %>%
-      left_join(site_data)
-}
-
-```
-
-#### Rarefied network analysis by colony phase
-```{r, include=FALSE}
-# network level version 2
-rare_net_nt2 <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  t_filter <- net %>%
-    select(site, date, phase, year) %>%
-    unique() %>%
-    group_by(site, phase, year) %>%
-    mutate(transects = n()) %>%
-    filter(transects >= n)
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      semi_join(t_filter) %>%
-      select(site, date, phase, year) %>%
-      unique() %>%
-      group_by(site, phase, year) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, phase, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      #filter(freq > 1) %>%
-      unite(webID, c(site, phase, year), sep = "_") %>%
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(networklevel, weighted = TRUE, level = "higher",
-          index = c("web asymmetry", "weighted NODF", "H2", "C score", "niche overlap", "generality")) %>% 
-      data.frame() %>%
-      t() %>%
-      data.frame() %>%
-      rownames_to_column() %>%
-      dplyr::select(id = rowname, everything()) %>%
-      as_tibble() %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out  %>%
-      gather(metric, value, -c(id, iteration)) %>% # get mean value across iterations for each metric
-      group_by(id, metric) %>%
-      summarize(value = mean(value)) %>%
-      separate(id, sep = "_", into = c("site", "phase", "year")) %>%
-      mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
-      left_join(site_data)
-}
-# species level version 2
-rare_net_sp2 <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  t_filter <- net %>%
-    select(site, date, phase, year) %>%
-    unique() %>%
-    group_by(site, phase, year) %>%
-    mutate(transects = n()) %>%
-    filter(transects >= n)
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      semi_join(t_filter) %>%
-      select(site, date, phase, year) %>%
-      unique() %>%
-      group_by(site, phase, year) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, phase, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      #filter(freq > 1) %>%
-      unite(webID, c(site, phase, year), sep = "_") %>%
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(specieslevel, level = "higher", index = c("d", "species strength")) %>%
-      map(rownames_to_column) %>%
-      map(as_tibble) %>%
-      bind_rows(.id = "id") %>% # Collapse list into big data frame
-      rename(sp = rowname) %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out %>%
-    gather(metric, value, -c(sp, id, iteration)) %>% # get mean value across iterations for each metric
-    group_by(sp, id, metric) %>%
-    summarize(value = mean(value)) %>%
-    separate(id, sep = "_", into = c("site", "phase", "year")) %>%
-    mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
-    left_join(site_data) %>%
-    left_join(bb_traits, by = c("sp" = "bb.sp"))
-}
-```
-
-### Rarefied network analysis by floral k.type instead of floral species
-```{r}
-# Rare_net function: rarefies networks by iteratively subsampling more frequently sampled networks to match the transect count of the least frequently sampled one
-# species level
-rare_net_sp_k.type.s <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      select(site, date) %>%
-      unique() %>%
-      group_by(site) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      na.omit() %>%
-      group_by(site) %>%
-      filter(length(unique(bb.sp)) > 1 & length(unique(k.type.s)) > 1) %>%
-      ungroup() %>%
-      filter(freq > 1) %>% # remove singletons
-      dplyr::select(higher = bb.sp, lower = k.type.s, webID = site, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(specieslevel, level = "higher", index = c("d", "species strength", "proportional generality")) %>%
-      map(rownames_to_column) %>%
-      map(as_tibble) %>%
-      bind_rows(.id = "site") %>% # Collapse list into big data frame
-      rename(sp = rowname) %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out %>%
-    gather(metric, value, -c(sp, site, iteration)) %>% # get mean value across iterations for each metric
-    group_by(sp, site, metric) %>%
-    summarize(value = mean(value)) %>%
-    left_join(site_data) %>%
-    left_join(bb_traits, by = c("sp" = "bb.sp"))
-}
-
-# network level
-rare_net_nt_k.type.s <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      select(site, date) %>%
-      unique() %>%
-      group_by(site) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      na.omit() %>%
-      group_by(site) %>%
-      filter(length(unique(bb.sp)) > 1 & length(unique(k.type.s)) > 1) %>%
-      ungroup() %>%
-      dplyr::select(higher = bb.sp, lower = k.type.s, webID = site, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(networklevel, weighted = TRUE, level = "higher", index = c("H2", "weighted NODF", "niche overlap", "C score", "generality", "web asymmetry", "weighted connectance")) %>% # We probably care mainly about the BB level of the network
-      data.frame() %>%
-      t() %>%
-      data.frame() %>%
-      rownames_to_column() %>%
-      dplyr::select(site = rowname, everything()) %>%
-      as_tibble() %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out  %>%
-      gather(metric, value, -c(site, iteration)) %>% # get mean value across iterations for each metric
-      group_by(site, metric) %>%
-      summarize(value = mean(value)) %>%
-      left_join(site_data)
-}
-```
-
 ### Load site data
 ```{r, include=FALSE}
-site_data <- read_csv("./processed_data/site_data.csv") %>%
+site_data <- read_csv("../processed_data/site_data.csv") %>%
   dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
@@
 
-### Climate data
+### Load climate data
 ```{r, include=FALSE}
-gdd <- read_csv("./processed_data/climate.csv") %>%
+gdd <- read_csv("../processed_data/climate.csv") %>%
   mutate(year = factor(year)) %>%
@@
 ```{r, include=FALSE}
-castes <- read_csv("./processed_data/network.csv") %>%
+castes <- read_csv("../processed_data/network.csv") %>%
   filter(bb.sp != "psit") %>% # We do not want to consider Psithyrus, since these parasitic species do not have a worker caste
@@
   mutate(year = factor(year))
+
+write_csv(breaks, "./output/breaks.csv")
 ```
@@
   facet_grid(elev.class2 ~ year) +
-  xlab("Julian day") +
+  xlab("Calendar day") +
   ylab("Proportional abundance")
@@
   ylab("Proportional abundance")
+
+ggplot(breaks_summary, aes(value, prop, color = caste)) +
+  geom_point(alpha = 0.25) +
+  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
+  facet_grid(elev.class2~ time.metric, scales = "free") +
+  xlab("Julian day") +
+  ylab("Proportional abundance")
+
+ggplot(filter(breaks_summary, time.metric == "yday"), 
+       aes(value, prop, color = caste)) +
+  geom_point(alpha = 0.25) +
+  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
+  facet_grid(elev.class2 ~ factor(year), scales = "free") +
+  xlab("Day of year") +
+  ylab("Proportional abundance") 
 ```
-
-### Load network and survey data
-```{r, include=FALSE}
-bb_traits <- read_csv("./processed_data/bb_traits.csv") %>%
-  select(subgenus, bb.sp, pbl.w, pbl.w.class)
-
-fl_traits <- read.csv("./processed_data/floral_k_type_nom_fix.csv") %>%
-  select(plant.sp, k.type) %>%
-  mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"),
-         k.type.ss = str_extract(k.type, "[-+]?[0-9]*"))
-
-fl_tax <- read_csv("./processed_data/floral_tax.csv")
-
-net <- read_csv("./processed_data/network.csv") %>%
-  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
-  unite(site.date, site, date, sep = "_", remove = FALSE) %>%
-  mutate(year = factor(year)) %>% # convert year to factor
-  full_join(bb_traits, by = c("bb.sp" = "bb.sp")) %>%
-  left_join(site_data, by = "site") %>%
-  left_join(gdd, by = c("site", "date")) %>%
-  left_join(breaks, by = c("site", "year")) %>%
-  left_join(fl_traits) %>%
-  left_join(fl_tax, by = "plant.genus") %>%
-  ungroup() %>%
-  mutate(phase = case_when(
-    gdd.cum < break1.gdd ~ "founding",
-    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
-    gdd.cum > break2.gdd ~ "reproductive"
-  )) %>%
-  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
-  arrange(site, date)
-
-survey <- read_csv("./processed_data/floral_survey.csv") %>%
-  mutate(year = factor(year)) %>% # convert year to factor
-  semi_join(net, by = "plant.sp") %>% # Consider only plants visited at a given site-date
-  left_join(gdd, by = c("site", "date")) %>%
-  left_join(breaks, by = c("site", "year")) %>%
-  left_join(site_data, by = "site") %>%
-  left_join(fl_traits) %>%
-  left_join(fl_tax) %>%
-  ungroup() %>%
-  mutate(phase = case_when(
-    gdd.cum < break1.gdd ~ "founding",
-    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
-    gdd.cum > break2.gdd ~ "reproductive"
-  )) %>%
-  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
-  arrange(site, date)
-
-```
-### BB and floral turnover ~ elevation
-```{r, include=FALSE}
-bb_range <- net %>%
-  group_by(site, elev.mean, bb.sp) %>%
-  summarize(abund = n()) %>%
-  group_by(bb.sp) %>%
-  mutate(elev.floor = min(elev.mean),
-         elev.ceiling = max(elev.mean),
-         elev.range = elev.ceiling - elev.floor,
-         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))
-
-fl_survey_elev <- survey %>%
-  filter(flower.cover > 0) %>%
-  group_by(site, plant.sp) %>%
-  summarize(abund = sum(flower.cover)) %>%
-  left_join(site_data) %>%
-  group_by(plant.sp) %>%
-  mutate(elev.floor = min(elev.mean),
-         elev.ceiling = max(elev.mean),
-         elev.range = elev.ceiling - elev.floor,
-         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2),
-         sites = length(unique(site)))
-
-fl_survey_elev_family <- survey %>%
-  filter(flower.cover > 0) %>%
-  group_by(site, plant.family) %>%
-  summarize(abund = sum(flower.cover)) %>%
-  left_join(site_data) %>%
-  group_by(plant.family) %>%
-  mutate(elev.floor = min(elev.mean),
-         elev.ceiling = max(elev.mean),
-         elev.range = elev.ceiling - elev.floor,
-         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2),
-         sites = length(unique(site)))
-
-fl_net_elev <- net %>%
-  group_by(site, elev.mean, plant.sp) %>%
-  summarize(abund = n()) %>%
-  group_by(plant.sp) %>%
-  mutate(elev.floor = min(elev.mean),
-         elev.ceiling = max(elev.mean),
-         elev.range = elev.ceiling - elev.floor,
-         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2),
-         sites = length(unique(site)))
-
-ggplot(filter(fl_survey_elev, sites > 0), 
-       aes(reorder(plant.sp, elev.ceiling), elev.mean)) +
-  geom_line(size = 1, color = "gray60") +
-  geom_point(alpha = 0.5) +
-  theme_light(14) +
-  theme(axis.text.x = element_blank()) +
-  xlab("Plant species") +
-  ylab("Elevation")
-
-ggplot(filter(fl_survey_elev_family, sites > 0), 
-       aes(reorder(plant.family, elev.med), elev.mean)) +
-  geom_line(size = 1, color = "gray60") +
-  geom_point(alpha = 0.5) +
-  theme_light(10) +
-  #theme(axis.text.x = element_blank()) +
-  xlab("Plant family") +
-  ylab("Elevation") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-ggplot(filter(fl_net_elev, sites > 0), 
-       aes(reorder(plant.sp, elev.med), elev.mean)) +
-  geom_line(size = 1, color = "gray60") +
-  geom_point(alpha = 0.5) +
-  theme_light(14) +
-  theme(axis.text.x = element_blank()) +
-  xlab("Plant species") +
-  ylab("Elevation")
-
-ggplot(bb_range, aes(reorder(bb.sp, elev.med), elev.mean)) +
-  geom_line(size = 3, color = "gray60") +
-  geom_point(aes(size = abund), alpha = 0.5) +
-  theme_light(14) +
-  xlab("BB species") +
-  ylab("Elevation") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-### Plot only the plants that drop out at 1750 m
-ggplot(filter(fl_net_elev, sites > 0 & elev.ceiling < 1750), 
-       aes(reorder(plant.sp, elev.med), elev.mean)) +
-  geom_line(size = 1, color = "gray60") +
-  geom_point(alpha = 0.5, aes(size = abund)) +
-  theme_light(8) +
-  xlab("Plant species") +
-  ylab("Elevation") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-ggplot(filter(fl_survey_elev, sites > 0 & elev.ceiling < 1750), 
-       aes(reorder(plant.sp, elev.med), elev.mean)) +
-  geom_line(size = 1, color = "gray60") +
-  geom_point(alpha = 0.5, aes(size = abund)) +
-  theme_light(8) +
-  xlab("Plant species") +
-  ylab("Elevation") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-```
-
-### FL traits by elevation -- do we see a sharp change at the tree line?
-```{r}
-survey_trait_elev <- survey %>%
-  group_by(site, year, elev.mean,  k.type.s) %>%
-  summarize(abund = sum(flower.cover))
-
-survey_trait_elev2 <- survey %>%
-  group_by(site, year, elev.mean,  k.type.ss) %>%
-  summarize(abund = sum(flower.cover))
-
-net_trait_elev <- net %>%
-  group_by(site, year, elev.mean,  k.type.s) %>%
-  summarize(abund = n())
-
-net_trait_elev2 <- net %>%
-  group_by(site, year, elev.mean,  k.type.ss) %>%
-  summarize(abund = n())
-
-ggplot(survey_trait_elev, aes(elev.mean, sqrt(abund))) +
-  geom_point() +
-  geom_smooth() +
-  geom_vline(xintercept = 1750, color = "red") +
-  facet_wrap(~k.type.s, scales = "free_y")
-
-ggplot(survey_trait_elev2, aes(elev.mean, sqrt(abund))) +
-  geom_point() +
-  geom_smooth() +
-  geom_vline(xintercept = 1750, color = "red") +
-  facet_wrap(~k.type.ss, scales = "free_y")
-
-ggplot(net_trait_elev, aes(elev.mean, sqrt(abund))) +
-  geom_point() +
-  geom_smooth() +
-  geom_vline(xintercept = 1750, color = "red") +
-  facet_wrap(~k.type.s, scales = "free_y")
-
-ggplot(net_trait_elev2, aes(elev.mean, sqrt(abund))) +
-  geom_point() +
-  geom_smooth() +
-  geom_vline(xintercept = 1750, color = "red") +
-  facet_wrap(~k.type.ss, scales = "free_y")
-
-net_trait_genus_elev <- net %>%
-  group_by(site, year, elev.mean, plant.genus, k.type.s) %>%
-  summarize(abund = n())
-
-survey_trait_genus_elev <- survey %>%
-  group_by(site, year, elev.mean, plant.genus, k.type.s) %>%
-  summarize(abund = sum(flower.cover))
-
-ggplot(filter(net_trait_genus_elev, k.type.s == 5.1), 
-       aes(elev.mean, sqrt(abund))) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~plant.genus)
-
-ggplot(filter(survey_trait_genus_elev, k.type.s %in% c(5.1)), 
-       aes(elev.mean, sqrt(abund))) +
-  geom_point() +
-  geom_smooth() +
-  geom_vline(xintercept = 1750, color = "red") +
-  facet_wrap(~plant.genus, ncol = 5) +
-  ylab("sqrt(Abundance)") +
-  xlab("Elevation")
-
-ggplot(filter(net_trait_genus_elev, k.type.s %in% c(5.1)), 
-       aes(elev.mean, sqrt(abund))) +
-  geom_point() +
-  geom_smooth() +
-  geom_vline(xintercept = 1750, color = "red") +
-  facet_wrap(~plant.genus, ncol = 5) +
-  ylab("sqrt(Visitation frequency)") +
-  xlab("Elevation")
-```
-
-### Genera and families by elevation
-```{r}
-
-survey_genus_elev <- survey %>%
-  group_by(site, year, elev.mean, plant.genus) %>%
-  summarize(abund = sum(flower.cover)) %>%
-  group_by(site, elev.mean, year) %>%
-  mutate(div = length(unique(plant.genus)))
-
-survey_family_elev <- survey %>%
-  group_by(site, year, elev.mean,  plant.family) %>%
-  summarize(abund = sum(flower.cover)) %>%
-  group_by(site, elev.mean, year) %>%
-  mutate(div = length(unique(plant.family)))
-
-net_genus_elev <- survey %>%
-  group_by(site, year, elev.mean,  plant.genus) %>%
-  summarize(abund = n()) %>%
-  group_by(site, elev.mean, year) %>%
-  mutate(div = length(unique(plant.genus)))
-
-net_family_elev <- survey %>%
-  group_by(site, year, elev.mean,  plant.family) %>%
-  summarize(abund = n()) %>%
-  group_by(site, elev.mean, year) %>%
-  mutate(div = length(unique(plant.family)))
-
-#### Plot
-ggplot(survey_genus_elev, aes(elev.mean, log(abund))) +
-  geom_point() +
-  stat_smooth() +
-  facet_wrap(~plant.genus, scales = "free_y")
-
-ggplot(survey_family_elev, aes(elev.mean, log(abund))) +
-  geom_point() +
-  stat_smooth() +
-  facet_wrap(~plant.family, scales = "free_y")
-
-ggplot(survey_genus_elev, aes(elev.mean, div)) +
-  geom_point() +
-  stat_smooth()
-
-ggplot(survey_family_elev, aes(elev.mean, div)) +
-  geom_point() +
-  stat_smooth()
-
-ggplot(net_genus_elev, aes(elev.mean, log(abund))) +
-  geom_point() +
-  stat_smooth() +
-  facet_wrap(~plant.genus, scales = "free_y")
-
-ggplot(net_family_elev, aes(elev.mean, log(abund))) +
-  geom_point() +
-  stat_smooth() +
-  facet_wrap(~plant.family, scales = "free_y")
-
-ggplot(net_genus_elev, aes(elev.mean, log(abund))) +
-  geom_point() +
-  stat_smooth(se = FALSE) +
-  facet_wrap(~plant.genus)
-
-ggplot(net_genus_elev, aes(elev.mean, log(abund))) +
-  geom_point() +
-  stat_smooth() +
-  facet_wrap(~plant.genus, scales = "free_y")
-ggplot(net_genus_elev, aes(elev.mean, abund)) +
-  geom_point() +
-  stat_smooth(se = FALSE) +
-  facet_wrap(~plant.genus)
-
-ggplot(net_genus_elev, aes(elev.mean, div)) +
-  geom_point() +
-  stat_smooth()
-```
-
-### BB abundance, diversity, and per-capita resource availability
-```{r eval=FALSE, echo=FALSE, include=FALSE}
-### Per-transect Bombus density
-bb_abund <- net %>%
-  group_by(site, date, dayofyear, gdd.cum, year, elev.class2, 
-           elev.mean, break1.gdd, break2.gdd) %>%
-  summarize(bb.abund = n())
-
-### Per-transect Bombus richness
-bb_rich <- net %>%
-  group_by(site, date, dayofyear, gdd.cum, year, elev.class2,
-           elev.mean, break1.gdd, break2.gdd) %>%
-  summarize(bb.rich = length(unique(bb.sp)))
-
-### Per-transect floral abund
-fl_abund <- survey %>%
-  group_by(site, date, dayofyear, gdd.cum, year, elev.class2, 
-           elev.mean, break1.gdd, break2.gdd) %>%
-  summarize(fl.abund = sum(flower.cover))
-
-### Per-transect floral richness
-fl_rich <- survey %>%
-  filter(flower.cover > 0) %>%
-  group_by(site, date, dayofyear, gdd.cum, year, elev.class2, 
-           elev.mean, break1.gdd, break2.gdd) %>%
-  summarize(fl.rich = n())
-
-### Join BB density and floral cover
-abund_div <- bb_abund %>%
-  full_join(bb_rich) %>%
-  full_join(fl_abund) %>% # Note that there is not perfect alignment in site-dates between the network observations and the floral surveying, so this join will cause some NAs that will need to be removed, causing some transects to drop out of the dataset
-  full_join(fl_rich) %>%
-  #na.omit() %>%
-  ungroup %>%
-  mutate(fl.per.bb = fl.abund/bb.abund,
-         site = factor(site),
-         yday = dayofyear,
-         week = week(date),
-         month = month(date)) %>%
-  ungroup() %>%
-  mutate(phase = case_when(
-    gdd.cum < break1.gdd ~ "founding",
-    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
-    gdd.cum > break2.gdd ~ "reproductive"
-  )) %>%
-  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive")),
-         total.rich = bb.rich + fl.rich,
-         web.asymmetry = (bb.rich - fl.rich) / (bb.rich + fl.rich))
-
-abund_div_mean <- abund_div %>%
-  select(-dayofyear, -week, -month, -break1.gdd, -break2.gdd) %>%
-  gather(metric, value, -c(site, date, gdd.cum, year, elev.class2, elev.mean, phase)) %>%
-  group_by(site, year, elev.class2, elev.mean, metric) %>%
-  summarize(value = mean(value, na.rm = TRUE))
-```
-
-### Plotting abundance and diversity patterns
-```{r}
-### BB abundance ~ floral abundance
-ggplot(abund_div, aes(log(fl.abund), bb.abund, color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "lm", se = TRUE) +
-  facet_grid(elev.class2 ~ phase, scales = "free") +
-  ylab("BB abundance") +
-  xlab("log(floral cover)") +
-  theme_bw()
-
-ggplot(abund_div, aes(log(fl.abund), bb.abund, color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  ylab("BB abundance") +
-  xlab("log(floral cover)") +
-  theme_bw()
-
-### BB richness ~ floral abundance
-ggplot(abund_div, aes(log(fl.abund), bb.rich, color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "lm", se = TRUE) +
-  facet_grid(elev.class2 ~ phase, scales = "free") +
-  ylab("BB richness") +
-  xlab("log(floral cover)") +
-  theme_bw()
-
-### BB abundance ~ floral richness
-ggplot(abund_div, aes(fl.rich, bb.abund, color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "lm", se = TRUE) +
-  facet_grid(elev.class2 ~ phase, scales = "free") +
-  ylab("BB abundance") +
-  xlab("Floral richness")
-
-### BB richness ~ floral richness
-ggplot(abund_div, aes(fl.rich, bb.rich, color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "lm", se = TRUE) +
-  facet_grid(elev.class2 ~ phase, scales = "free") +
-  ylab("BB richness") +
-  xlab("Floral richness")
-
-### Floral richness ~ floral abundance
-ggplot(abund_div, aes(log(fl.abund), fl.rich, color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "lm", se = TRUE) +
-  facet_grid(elev.class2 ~ phase, scales = "free") +
-  ylab("Floral richness") +
-  xlab("log(floral cover)")
-
-### Per-capita floral resource abundance ~ elevation
-ggplot(abund_div, aes(elev.mean, log(fl.per.bb))) +
-  geom_point(alpha = 0.5) +
-  #stat_summary(aes(y = log(fl.per.bb), group=1), fun.y=mean, colour="black", geom="point",group=1) +
-  geom_smooth() +
-  facet_grid(year~phase, scales = "free") +
-  ylab("log(fl-per-bb)") +
-  xlab("Elevation") +
-  theme_light()
-
-ggplot(abund_div, aes(elev.mean, log(fl.per.bb))) +
-  geom_point(alpha = 0.5) +
-  #stat_summary(aes(y = log(fl.per.bb), group=1), fun.y=mean, colour="black", geom="point",group=1) +
-  geom_smooth() +
-  facet_grid(.~year, scales = "free") +
-  ylab("log(fl-per-bb)") +
-  xlab("Elevation") +
-  theme_light()
-
-ggplot(abund_div, aes(elev.mean, log(fl.per.bb))) +
-  geom_point(alpha = 0.5, aes(color = phase)) +
-  #stat_summary(aes(y = log(fl.per.bb), group=1), fun.y=mean, colour="black", geom="point",group=1) +
-  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
-  facet_grid(year~., scales = "free") +
-  ylab("log(fl-per-bb)") +
-  xlab("Elevation") +
-  theme_light()
-
-ggplot(abund_div, aes(elev.mean, log(fl.per.bb))) +
-  geom_point(alpha = 0.5, aes(color = phase)) +
-  #stat_summary(aes(y = log(fl.per.bb), group=1), fun.y=mean, colour="black", geom="point",group=1) +
-  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
-  ylab("log(fl-per-bb)") +
-  xlab("Elevation") +
-  theme_light()
-
-ggplot(abund_div, aes(elev.mean, fl.per.bb)) +
-  geom_point(alpha = 0.5, aes(color = phase)) +
-  #stat_summary(aes(y = log(fl.per.bb), group=1), fun.y=mean, colour="black", geom="point",group=1) +
-  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
-  ylab("log(fl-per-bb)") +
-  xlab("Elevation") +
-  theme_light()
-
-ggplot(filter(abund_div_mean, metric == "fl.per.bb"), aes(elev.mean, log(value))) +
-  geom_point(alpha = 1) +
-  #geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) +
-  ylab("log(fl-per-bb)") +
-  xlab("Elevation") +
-  facet_wrap(~year) +
-  theme_light()
-
-lm1 <- lm(value ~ elev.mean * year, data = filter(abund_div_mean, metric == "fl.per.bb"))
-summary(lm1)
-
-ggplot(filter(abund_div_mean, metric == "fl.per.bb"), aes(elev.mean, value)) +
-  geom_point(alpha = 1) +
-  #geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
-  geom_smooth() +
-  ylab("fl-per-bb") +
-  xlab("Elevation") +
-  theme_light()
-
-ggplot(filter(abund_div_mean, metric == "fl.per.bb"), aes(elev.mean, log(value), color = year)) +
-  geom_point(alpha = 1) +
-  #geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
-  geom_smooth() +
-  ylab("FL per BB") +
-  xlab("Elevation") +
-  theme_light()
-
-ggplot(filter(abund_div_mean, metric == "fl.per.bb"), aes(elev.mean, log(value))) +
-  geom_point(alpha = 1, aes(color = year)) +
-  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
-  #geom_smooth() +
-  ylab("FL per BB") +
-  xlab("Elevation") +
-  theme_light(16)
-
-### Per-capita floral resource abundance ~ GDD
-ggplot(abund_div, aes(gdd.cum, log(fl.per.bb))) +
-  geom_point(alpha = 0.5, aes(color = phase)) +
-  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
-  facet_grid(elev.class2 ~ year, scales = "free") +
-  ylab("log(fl-per-bb)") +
-  xlab("GDD") +
-  theme_light()
-
-ggplot(abund_div, aes(gdd.cum, log(fl.per.bb))) +
-  geom_point(alpha = 0.5, aes(color = phase)) +
-  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
-  ylab("log(fl-per-bb)") +
-  xlab("GDD") +
-  theme_light()
-
-### Per-capita floral resource abundance ~ Julian day
-ggplot(abund_div, aes(yday, log(fl.per.bb))) +
-  geom_point(alpha = 0.5, aes(color = phase)) +
-  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
-  facet_grid(elev.class2 ~ year, scales = "free") +
-  ylab("log(fl-per-bb)") +
-  xlab("Julian day") +
-  theme_light()
-
-### BB richness ~ elevation
-ggplot(abund_div, aes(elev.mean, bb.rich)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(.~year)
-
-### BB richness ~ GDD
-ggplot(abund_div, aes(gdd.cum, bb.rich)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(elev.class2 ~ year)
-
-### BB richness ~ Julian day
-ggplot(abund_div, aes(yday, bb.rich)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(elev.class2 ~ year)
-
-### BB abundance ~ elevation
-ggplot(abund_div, aes(elev.mean, bb.abund)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(year ~ phase, scales = "free")
-
-ggplot(abund_div, aes(elev.mean, bb.abund)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(.~year, scales = "free")
-
-ggplot(abund_div, aes(elev.mean, bb.abund, color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  xlab("Elevation") +
-  ylab("BB abundance")  +
-  theme_light(16)
-
-ggplot(abund_div, aes(elev.mean, sqrt(bb.abund), color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  xlab("Elevation") +
-  ylab("sqrt(BB abundance)")  +
-  theme_light(16)
-
-### BB abundance ~ GDD
-ggplot(abund_div, aes(gdd.cum, bb.abund)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(elev.class2 ~ year)
-
-### BB abundance ~ Julian day
-ggplot(abund_div, aes(yday, bb.abund)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(elev.class2 ~ year)
-
-### Floral richness ~ elevation
-ggplot(abund_div, aes(elev.mean, fl.rich)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(year ~ phase)
-
-ggplot(abund_div, aes(elev.mean, fl.rich)) +
-  geom_point(alpha = 0.5, aes(color = year)) +
-  geom_smooth(color = "black", linetype = "dashed", size = 0.75) +
-  theme_light(16) +
-  xlab("Elevation") +
-  ylab("Floral richness")
-
-### Floral richness ~ GDD
-ggplot(abund_div, aes(gdd.cum, fl.rich)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(elev.class2 ~ year)
-
-### Floral richness ~ Julian day
-ggplot(abund_div, aes(yday, fl.rich)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(elev.class2 ~ year)
-
-### Floral abundance ~ elevation
-ggplot(abund_div, aes(elev.mean, log(fl.abund))) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  facet_grid(year ~ phase, scales = "free") +
-  theme_light()
-
-ggplot(abund_div, aes(elev.mean, log(fl.abund), color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  theme_light()
-
-ggplot(abund_div, aes(elev.mean, fl.abund)) +
-  geom_point(alpha = 0.5, aes(color = year)) +
-  geom_smooth(color = "black", linetype = "dashed", size = 0.75) +
-  theme_light(16) +
-  xlab("Elevation") +
-  ylab("Floral abundance")
-
-ggplot(abund_div, aes(elev.mean, log(fl.abund))) +
-  geom_point(alpha = 0.5, aes(color = year)) +
-  geom_smooth(color = "black", linetype = "dashed", size = 0.75) +
-  theme_light(16) +
-  xlab("Elevation") +
-  ylab("log(Floral abundance)")
-
-ggplot(abund_div, aes(elev.mean, fl.abund, color = year)) +
-  geom_point(alpha = 0.5) +
-  theme_light(16) +
-  xlab("Elevation") +
-  ylab("Floral abundance")
-
-ggplot(abund_div, aes(elev.mean, log(fl.abund), color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  theme_light(16) +
-  xlab("Elevation") +
-  ylab("log(Floral abundance)")
-
-ggplot(abund_div, aes(elev.mean, sqrt(bb.abund), color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  xlab("Elevation") +
-  ylab("sqrt(BB abundance)")  +
-  theme_light(16)
-
-### Floral abundance ~ GDD
-ggplot(abund_div, aes(gdd.cum, log(fl.abund))) +
-  geom_point(aes(color = phase), alpha =  0.5) +
-  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
-  facet_grid(elev.class2 ~ year, scales = "free") +
-  theme_light()
-
-### Floral abundance ~ Julian day
-ggplot(abund_div, aes(yday, log(fl.abund))) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(elev.class2 ~ year)
-
-### Total (BB = FL) richness ~ elevation
-ggplot(abund_div, aes(elev.mean, total.rich)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(year ~ phase, scales = "free")
-
-### Web asymmetry ~ elevation
-ggplot(abund_div, aes(elev.mean, web.asymmetry)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  facet_grid(year ~ phase, scales = "free")
-
-ggplot(filter(abund_div_mean, metric == "web.asymmetry"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(year ~ ., scales = "free")
-
-### Web asymmetry ~ gdd
-ggplot(abund_div, aes(gdd.cum, web.asymmetry)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  facet_grid(year ~.)
-
-### Web asymmetry ~ yday
-ggplot(abund_div, aes(yday, web.asymmetry)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  facet_grid(year ~.)
-
-### Web asymmetry ~ floral abundance
-ggplot(abund_div, aes(log(fl.abund), web.asymmetry)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  facet_grid(year ~ phase, scales = "free")
-
-### Web asymmetry ~ BB abundance
-ggplot(abund_div, aes(bb.abund, web.asymmetry)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  facet_grid(year ~ phase, scales = "free")
-
-### Web asymmetry ~ Per-capita floral resource abundance
-ggplot(abund_div, aes(log(fl.per.bb), web.asymmetry)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  facet_grid(year ~ phase, scales = "free")
-
-ggplot(abund_div, aes(log(fl.per.bb), web.asymmetry)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  facet_grid(year ~ .)
-```
-
-### GAM surface plots of abundance and diversity metrics
-#### HGAM elevation x time x year (I would really like someone -- maybe my future self -- to check that I'm doing this right)
-##### Floral abundance
-```{r eval=FALSE, include=FALSE}
-fl_abund_yday <- gamV(fl.abund ~ year +
-                    te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = abund_div,
-                    family = "tw",
-                    method = "REML")
-
-fl_abund_yday2 <- gamV(log(fl.abund + 0.001) ~ year +
-                    te(yday, elev.mean, by=year, bs= c("cr", "cr"), k=c(10, 10), m=1),
-                    data = abund_div,
-                    #family = "scat",
-                    method = "REML")
-
-fl_abund_yday3 <- gamV(log(fl.abund + 0.001) ~ year +
-                    te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = abund_div,
-                    #family = "scat",
-                    method = "REML")
-
-AIC(fl_abund_yday, fl_abund_yday2, fl_abund_yday3)
-
-check.gamViz(fl_abund_yday)
-summary(fl_abund_yday)
-
-print(plot(fl_abund_yday, allTerms = TRUE), pages = 1)
-
-check.gamViz(fl_abund_yday2)
-summary(fl_abund_yday2)
-
-print(plot(fl_abund_yday2, allTerms = TRUE), pages = 1)
-
-check.gamViz(fl_abund_yday3)
-summary(fl_abund_yday3)
-
-print(plot(fl_abund_yday3, allTerms = TRUE), pages = 1)
-
-fl_abund_gdd <- gam(fl.abund ~ year +
-                    te(gdd.cum, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    family = "tw",
-                    method = "REML")
-
-fl_abund_gdd.vis <- getViz(fl_abund_gdd)
-check.gamViz(fl_abund_gdd.vis)
-summary(fl_abund_gdd.vis)
-
-plot(fl_abund_gdd.vis, allTerms = TRUE) +
-  ggtitle("FL abundance (gdd)") +
-  xlab("GDD") +
-  ylab("Elevation")
-```
-
-##### Floral richness -- since year does not seem to change the overall pattern (though perhaps the magnitude), I could use a simpler model, perhaps with year as a random intercept effect; some explanat
-```{r eval=FALSE, include=FALSE}
-### By year
-fl_rich_yday <- gam(fl.rich ~ year +
-                    te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-fl_rich_yday.vis <- getViz(fl_rich_yday)
-check.gamViz(fl_rich_yday.vis)
-summary(fl_rich_yday.vis)
-
-plot(fl_rich_yday.vis, allTerms = TRUE) +
-  ggtitle("FL richness (gdd)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### With year as parametric term
-fl_rich_yday2 <- gam(fl.rich ~ year +
-                    te(yday, elev.mean, bs= "tp", k= 10, m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-fl_rich_yday2.vis <- getViz(fl_rich_yday2)
-check.gamViz(fl_rich_yday2.vis)
-summary(fl_rich_yday2.vis)
-
-plot(sm(fl_rich_yday2.vis, 1)) + 
-  l_fitRaster() + 
-  l_fitContour() + 
-  l_points(size = 1) +
-  ggtitle("FL richness (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-plot(pterm(fl_rich_yday2.vis, 1))
-
-### By year
-fl_rich_gdd <- gam(fl.rich ~ year +
-                    te(gdd.cum, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-fl_rich_gdd.vis <- getViz(fl_rich_gdd)
-check.gamViz(fl_rich_gdd.vis)
-summary(fl_rich_gdd.vis)
-
-plot(fl_rich_gdd.vis, allTerms = TRUE) +
-  ggtitle("FL richness (gdd)") +
-  xlab("GDD") +
-  ylab("Elevation")
-
-### With year as parametric term
-fl_rich_gdd2 <- gam(fl.rich ~ year +
-                    te(gdd.cum, elev.mean, bs= "tp", k= 10, m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-fl_rich_gdd2.vis <- getViz(fl_rich_gdd2)
-check.gamViz(fl_rich_gdd2.vis)
-summary(fl_rich_gdd2.vis)
-
-plot(sm(fl_rich_gdd2.vis, 1)) + 
-  l_fitRaster() + 
-  l_fitContour() + 
-  l_points(size = 1) +
-  ggtitle("FL richness (gdd)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-plot(pterm(fl_rich_gdd2.vis, 1))
-```
-
-##### BB abundance
-```{r eval=FALSE, include=FALSE}
-bb_abund_yday <- gam(bb.abund ~ year +
-                    te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-bb_abund_yday.vis <- getViz(bb_abund_yday)
-check.gamViz(bb_abund_yday.vis)
-summary(bb_abund_yday.vis)
-
-plot(bb_abund_yday.vis, allTerms = TRUE) +
-  ggtitle("BB abundance (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-bb_abund_gdd <- gam(bb.abund ~ year +
-                    te(gdd.cum, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-bb_abund_gdd.vis <- getViz(bb_abund_gdd)
-check.gamViz(bb_abund_gdd.vis)
-summary(bb_abund_gdd.vis)
-
-plot(bb_abund_gdd.vis, allTerms = TRUE) +
-  ggtitle("BB abundance (gdd)") +
-  xlab("GDD") +
-  ylab("Elevation")
-```
-
-##### BB richness
-```{r eval=FALSE, include=FALSE}
-bb_rich_yday <- gam(bb.rich ~ year +
-                    te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-bb_rich_yday.vis <- getViz(bb_rich_yday)
-check.gamViz(bb_rich_yday.vis)
-summary(bb_rich_yday.vis)
-plot(sm(bb_rich_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("bb richness 2010 (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-plot(sm(bb_rich_yday.vis, 2)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("bb richness 2011 (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-plot(sm(bb_rich_yday.vis, 3)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("bb richness 2012 (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-bb_rich_gdd <- gam(bb.rich ~ year +
-                    te(gdd.cum, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-bb_rich_gdd.vis <- getViz(bb_rich_gdd)
-check.gamViz(bb_rich_gdd.vis)
-summary(bb_rich_gdd.vis)
-
-plot(bb_rich_gdd.vis, allTerms = TRUE) +
-  ggtitle("BB richness (gdd)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-```
-
-### Rarefied network analysis (not broken down by phase)
-#### Inspect transect counts to set subsampling parameter
-```{r, include = FALSE}
-### Determine the minimum transect count per site*year to set rarefaction sampling
-transects <- net %>%
-  select(site, year, date) %>%
-  unique()
-
-transects_per_siteyear <- transects %>%
-  group_by(site, year) %>%
-  summarize(transects = n())
-
-min_transects_per_year <- transects_per_siteyear %>%
-  group_by(year) %>%
-  summarize(min_transects = min(transects))
-
-### Create yearly subsets to feed to rare_net
-net_2012 <- filter(net, year == "2012")
-net_2011 <- filter(net, year == "2011")
-net_2010 <- filter(net, year == "2010")
-
-### Demonstrate that the distribution of interaction frequencies does not differ across elevation classes; in other words, no systematic bias is introduced to the network analysis due to uneven frequency of species with few observations
-net_freq <- net %>%
-  group_by(site, dayofyear, year, bb.sp, elev.mean, elev.class2) %>%
-  summarize(freq = n())
-
-ggplot(net_freq, aes(elev.class2, log(freq))) +
-  geom_boxplot() +
-  facet_wrap(~year)
-```
-
-#### Generate network metric tables
-```{r}
-### Species level
-#### BBs
-spmet_2012_BB <- rare_net_sp_BB(net_2012, 9, 100) %>%
-  mutate(year = factor(rep("2012", n())))
-spmet_2011_BB <- rare_net_sp_BB(net_2011, 4, 100) %>%
-  mutate(year = factor(rep("2011", n())))  
-spmet_2010_BB <- rare_net_sp_BB(net_2010, 3, 100) %>%
-  mutate(year = factor(rep("2010", n())))
-
-spmet_BB <- bind_rows(spmet_2012_BB, spmet_2011_BB, spmet_2010_BB)
-
-spmet_mean_BB <- spmet_BB %>%
-  group_by(site, year, elev.mean, metric) %>%
-  summarize(value = mean(value, na.rm = TRUE)) %>%
-  ungroup() %>%
-  mutate(year = factor(year))
-
-#### plants
-spmet_2012_FL <- rare_net_sp_FL(net_2012, 9, 100) %>%
-  mutate(year = factor(rep("2012", n())))
-spmet_2011_FL <- rare_net_sp_FL(net_2011, 4, 100) %>%
-  mutate(year = factor(rep("2011", n())))  
-spmet_2010_FL <- rare_net_sp_FL(net_2010, 3, 100) %>%
-  mutate(year = factor(rep("2010", n())))
-
-spmet_FL <- bind_rows(spmet_2012_FL, spmet_2011_FL, spmet_2010_FL)
-
-spmet_mean_FL <- spmet_FL %>%
-  group_by(site, year, elev.mean, metric) %>%
-  summarize(value = mean(value, na.rm = TRUE)) %>%
-  ungroup() %>%
-  mutate(year = factor(year))
-
-### Network- and group-level
-netmet_rare_2012 <- rare_net_nt(net_2012, n = 9, iter = 25) %>%
-  mutate(year = factor(rep("2012", n())))
-netmet_rare_2011 <- rare_net_nt(net_2011, n = 4, iter = 25)  %>%
-  mutate(year = factor(rep("2011", n())))
-netmet_rare_2010 <- rare_net_nt(net_2010, n = 3, iter = 25) %>%
-  mutate(year = factor(rep("2010", n())))
-
-netmet <- bind_rows(netmet_rare_2012, 
-                    netmet_rare_2011, 
-                    netmet_rare_2010) %>%
-  mutate(metric = factor(metric, levels = c("H2",
-                                            "weighted.NODF",
-                                            "niche.overlap.HL", 
-                                            "C.score.HL",
-                                            "generality.HL",
-                                            "niche.overlap.LL", 
-                                            "C.score.LL",
-                                            "generality.LL",
-                                            "web.asymmetry",
-                                            "weighted.connectance"))) %>%
-  ungroup() %>%
-  mutate(year = factor(year))
-```
-
-### Plot network analysis
-```{r}
-#### Per species values
-ggplot(filter(spmet_2012_BB,
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf") & metric == "d"), 
-       aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
-  facet_wrap(~sp, scales = "free")
-
-ggplot(filter(spmet_2011_BB,
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf") & metric == "d"), 
-       aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
-  facet_wrap(~sp, scales = "free")
-
-ggplot(filter(spmet_2010_BB,
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf") & metric == "d"), 
-       aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
-  facet_wrap(~sp, scales = "free")
-
-#### Community mean d
-##### BB
-ggplot(filter(spmet_mean_BB, metric == "d"), aes(elev.mean, value)) +
-  geom_point() +
-  #geom_smooth(method = "gam", formula = y ~ s(x, k = 10)) +
-  xlab("Elevation") +
-  ylab("Community mean d'") +
-  facet_wrap(~year) +
-  theme_light(16)
-
-gam_d1 <- gamV(value ~ s(year, bs = "re") + s(elev.mean, k = 5, bs = "tp", by = year, m = 1), 
-             data = filter(spmet_mean_BB, metric == "d"),
-             method = "REML")
-
-check.gamViz(gam_d1)
-summary(gam_d1)
-plot(gam_d1, allTerms = TRUE)
-
-summary(lm(value ~ elev.mean, data =  filter(spmet_mean_BB, metric == "d" & year == 2012)))
-plot(lm(value ~ elev.mean, data =  filter(spmet_mean_BB, metric == "d" & year == 2012)))
-
-##### FL
-ggplot(filter(spmet_mean_FL, metric == "d"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
-  xlab("Elevation") +
-  ylab("Community mean d'") +
-  facet_wrap(~year) +
-  theme_light(16)
-
-#### Network level
-
-##### niche overlap
-ggplot(filter(netmet, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x)) +
-  #geom_smooth() +
-  xlab("Elevation") +
-  ylab("Niche overlap") +
-  facet_wrap(~year) +
-  theme_light(16)
-
-gam_niche.overlap <- gamV(value ~ s(year, bs = "re") + s(elev.mean, by = year, k = 10, bs = "tp", m = 2), 
-                         data = filter(netmet, metric == "niche.overlap.HL"),
-                         method = "REML")
-
-check.gamViz(gam_niche.overlap)
-summary(gam_niche.overlap)
-
-plot(gam_niche.overlap, allTerms = TRUE)
-
-##### C.score
-ggplot(filter(netmet, metric == "C.score.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x)) +
-  xlab("Elevation") +
-  ylab("C score") +
-  facet_wrap(~year) +
-  theme_light(16)
-
-gam_cscore <- gamV(value ~ s(year, bs = "re") + s(elev.mean, by = factor(year), bs = "tp", k = 6, m = 1), 
-                  data = filter(netmet, metric == "C.score.HL"),
-                  #family = "scat",
-                  method = "REML")
-
-check.gamViz(gam_cscore)
-summary(gam_cscore)
-
-plot(gam_cscore, allTerms = TRUE)
-```
-
-### Rarefied network analysis by k.type instead of by plant sp
-```{r, include = FALSE}
-### Create yearly subsets to feed to rare_net
-net_2012 <- filter(net, year == "2012")
-net_2011 <- filter(net, year == "2011")
-net_2010 <- filter(net, year == "2010")
-
-### Species level
-spmet_2012_k <- rare_net_sp_k.type.s(net_2012, 9, 20) %>%
-  mutate(year = factor(rep("2012", n())))
-spmet_2011_k <- rare_net_sp_k.type.s(net_2011, 4, 20) %>%
-  mutate(year = factor(rep("2011", n())))  
-spmet_2010_k <- rare_net_sp_k.type.s(net_2010, 3, 20) %>%
-  mutate(year = factor(rep("2010", n())))
-
-spmet_k <- bind_rows(spmet_2012_k, spmet_2011_k, spmet_2010_k)
-
-spmet_mean_k <- spmet_k %>%
-  group_by(site, year, elev.mean, metric) %>%
-  summarize(value = mean(value))
-
-### Network- and group-level
-netmet_rare_2012_k <- rare_net_nt_k.type.s(net_2012, n = 9, iter = 20) %>%
-  mutate(year = factor(rep("2012", n())))
-netmet_rare_2011_k <- rare_net_nt_k.type.s(net_2011, n = 4, iter = 20)  %>%
-  mutate(year = factor(rep("2011", n())))
-netmet_rare_2010_k <- rare_net_nt_k.type.s(net_2010, n = 3, iter = 20) %>%
-  mutate(year = factor(rep("2010", n())))
-
-netmet_k <- bind_rows(netmet_rare_2012, 
-                    netmet_rare_2011, 
-                    netmet_rare_2010) %>%
-  mutate(metric = factor(metric, levels = c("H2",
-                                            "weighted.NODF",
-                                            "niche.overlap.HL", 
-                                            "C.score.HL",
-                                            "generality.HL",
-                                            "web.asymmetry",
-                                            "weighted.connectance")))
-
-### Plot
-#### Per species values
-ggplot(filter(spmet_2012,
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
-       aes(elev.mean, value)) +
-  geom_point() +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth() +
-  facet_grid(metric~sp, scales = "free")
-
-ggplot(filter(spmet_2011,
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
-       aes(elev.mean, value)) +
-  geom_point() +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth() +
-  facet_grid(metric~sp, scales = "free")
-
-ggplot(filter(spmet_2010,
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
-       aes(elev.mean, value)) +
-  geom_point() +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth() +
-  facet_grid(metric~sp, scales = "free")
-
-#### Community mean d
-ggplot(filter(spmet_mean_k, metric == "d"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
-  xlab("Elevation") +
-  ylab("Community mean d'") +
-  theme_light(16)
-
-ggplot(filter(spmet_mean_k, metric == "d"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
-  xlab("Elevation") +
-  ylab("Community mean d'") +
-  facet_wrap(~year) +
-  theme_light(16)
-
-gam_d <- gam(value ~ year + s(elev.mean, k = 10), 
-             data = filter(spmet, metric == "d"),
-             method = "REML")
-
-gam_d.vis <- getViz(gam_d)
-check.gamViz(gam_d.vis)
-summary(gam_d.vis)
-
-plot(gam_d.vis, allTerms = TRUE) +
-  xlab("Elevation")
-
-#### Network level
-ggplot(netmet, aes(elev.mean, value)) +
-  geom_point() +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth() +
-  facet_grid(metric~year, scales = "free")
-
-##### H2
-ggplot(filter(netmet, metric == "H2"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
-  xlab("Elevation") +
-  ylab("H2'") +
-  theme_light(16)
-
-gam_H2 <- gam(value ~ year + s(elev.mean), 
-              data = filter(netmet, metric == "H2"),
-              method = "REML")
-
-gam_H2.vis <- getViz(gam_H2)
-check.gamViz(gam_H2.vis)
-summary(gam_H2.vis)
-
-plot(gam_H2.vis, allTerms = TRUE) +
-  xlab("Elevation")
-
-##### niche overlap
-ggplot(filter(netmet, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
-  #geom_smooth() +
-  xlab("Elevation") +
-  ylab("Niche overlap") +
-  theme_light(16)
-
-ggplot(filter(netmet, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
-  #geom_smooth() +
-  xlab("Elevation") +
-  ylab("Niche overlap") +
-  facet_wrap(~year) +
-  theme_light(16)
-
-gam_niche.overlap <- gam(value ~ year + s(elev.mean), 
-                         data = filter(netmet, metric == "niche.overlap.HL"),
-                         method = "REML")
-
-gam_niche.overlap.vis <- getViz(gam_niche.overlap)
-check.gamViz(gam_niche.overlap.vis)
-summary(gam_niche.overlap.vis)
-
-plot(gam_niche.overlap.vis, allTerms = TRUE) +
-  xlab("Elevation")
-
-##### C.score
-ggplot(filter(netmet, metric == "C.score.HL"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
-  #geom_smooth() +
-  xlab("Elevation") +
-  ylab("C score") +
-  theme_light(16)
-
-ggplot(filter(netmet, metric == "C.score.HL"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
-  #geom_smooth() +
-  xlab("Elevation") +
-  ylab("C score") +
-  facet_wrap(~year) +
-  theme_light(16)
-
-gam_cscore <- gam(value ~ year + s(elev.mean), 
-                  data = filter(netmet, metric == "C.score.HL"),
-                  method = "REML")
-
-gam_cscore.vis <- getViz(gam_cscore)
-check.gamViz(gam_cscore.vis)
-summary(gam_cscore.vis)
-
-plot(gam_cscore.vis, allTerms = TRUE) +
-  xlab("Elevation")
-```
-
-### Do BB species differ in their k.type associations?
-```{r}
-web_by_site <- net %>%
-  group_by(site, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
-  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-  na.omit() %>%
-  group_by(site) %>%
-  filter(length(unique(bb.sp)) > 1 & length(unique(k.type.s)) > 1) %>%
-  ungroup() %>%
-  dplyr::select(higher = bb.sp, lower = k.type.s, webID = site, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs()
-
-visweb(web[[1]])
-plotweb(web[[1]])
-
-web <- net %>%
-  group_by(year, elev.class2, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
-  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-  na.omit() %>%
-  unite(webID, c(year, elev.class2), sep = "_") %>%
-  dplyr::select(higher = bb.sp, lower = k.type.s, webID, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs()
-
-# 2010 high
-visweb(web[[1]])
-plotweb(web[[1]])
-
-# 2010 low
-visweb(web[[2]])
-plotweb(web[[2]])
-
-# 2010 med
-visweb(web[[3]])
-plotweb(web[[3]])
-
-# 2011 high
-visweb(web[[4]])
-plotweb(web[[4]])
-
-# 2011 low
-visweb(web[[5]])
-plotweb(web[[5]])
-
-# 2011 med
-visweb(web[[6]])
-plotweb(web[[6]])
-
-# 2012 high
-visweb(web[[7]])
-plotweb(web[[7]])
-
-# 2012 low
-visweb(web[[8]])
-plotweb(web[[8]])
-
-# 2012 med
-visweb(web[[9]])
-plotweb(web[[9]])
-
-```
-
-### Does tongue length class predict k.type association? Remember that B. wurflenii is primarily a nectar robber.
-```{r}
-library(infer)
-
-kugler_desc <- read_delim("./kugler.tsv", delim = "\t") %>%
-  select(1,2)
-
-ggplot(bb_traits, aes(pbl.w)) +
-  geom_histogram()
-
-net_cont <- net %>%
-  group_by(year, elev.class2, pbl.w.class, k.type.ss) %>%
-  summarize(freq = n()) %>%
-  group_by(elev.class2, pbl.w.class) %>%
-  mutate(prop = freq/sum(freq))
-
-ggplot(net_cont, aes(pbl.w.class, freq, fill = k.type.ss)) +
-  geom_col() +
-  facet_grid(.~elev.class2)
-
-ggplot(net_cont, aes(pbl.w.class, prop, fill = k.type.ss)) +
-  geom_col() +
-  facet_grid(.~elev.class2)
-
-chisq_mod <- chisq.test(net$pbl.w.class, net$k.type.ss)
-chisq_mod
-
-  c <- net %>%
-  select(k.type.s, pbl.w.class) %>%
-  mutate(k.type.s = factor(k.type.s),
-         pbl.w.class = factor(pbl.w.class)) %>%
-  chisq_test(k.type.s ~ pbl.w.class)
-
-```
-
-### Let's look at how abundance varies on a per-species basis
-```{r eval=FALSE, include=FALSE}
-species <- net %>%
-  group_by(site, date, year, dayofyear, gdd.cum, phase, elev.mean, elev.class2, bb.sp, subgenus, pbl.w, pbl.w.class) %>%
-  summarize(abund = n()) %>%
-  group_by(site, date) %>%
-  mutate(prop = abund/sum(abund))
-
-subgenera <- net %>%
-  group_by(site, date, year, dayofyear, gdd.cum, phase, elev.mean, elev.class2, subgenus, pbl.w, pbl.w.class) %>%
-  summarize(abund = n()) %>%
-  group_by(site, date) %>%
-  mutate(prop = abund/sum(abund))
-
-ggplot(species, aes(dayofyear, abund)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth() +
-  facet_grid(year ~ bb.sp)
-
-ggplot(species, aes(dayofyear, prop)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth() +
-  facet_grid(year ~ bb.sp)
-
-ggplot(species, aes(elev.mean, abund)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth(se = FALSE, span = 1) +
-  facet_grid(year ~ bb.sp, scales = "free")
-
-ggplot(species, aes(elev.mean, prop)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth(se = FALSE, span = 1) +
-  facet_grid(year ~ bb.sp, scales = "free")
-
-ggplot(subgenera, aes(dayofyear, abund)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth() +
-  facet_grid(year ~ subgenus)
-
-ggplot(subgenera, aes(dayofyear, prop)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth() +
-  facet_grid(year ~ subgenus)
-
-ggplot(subgenera, aes(elev.mean, abund)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth() +
-  facet_grid(year ~ subgenus)
-
-ggplot(subgenera, aes(elev.mean, prop)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth(se = FALSE, span = 1) +
-  facet_grid(year ~ subgenus)
-
-### Pairs analysis
-species_pairs <- species %>%
-  select(site, year, date, bb.sp, abund) %>%
-  ungroup() %>%
-  spread(bb.sp, abund) %>%
-  replace(is.na(.), 0)
-
-ggplot(species_pairs, aes(pasc, prat)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_wrap(~year)
-
-ggpairs(species_pairs, columns = c("bss", "camp", "flav", 
-                                   "gers", "hort", "humi",
-                                   "hypn", "jone", "lapi",
-                                   "mend", "mont", "muci",
-                                   "pasc", "prat", "psyt",
-                                   "pyre", "soro", "wurf"))
-
-ggpairs(species, columns = c("bb.sp", "abund"),
-        cardinality_threshold = 20)
-
-```
-
-######## Old analyses that I probably don't care to use but don't want to delete yet #########
-
-### Per-transect network analysis -- is it worth it? I would have to filter out the transects for which the networks are too small.
-#### Network level
-```{r eval=FALSE, include=FALSE}
-per.transect_netmet <- net %>%
-  group_by(site, dayofyear, gdd.cum, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-  group_by(site, dayofyear, gdd.cum, year) %>%
-  mutate(size = sum(freq),
-         bb.rich = length(unique(bb.sp)),
-         fl.rich = length(unique(plant.sp)),
-         total.rich = bb.rich + fl.rich) %>%
-  filter(bb.rich > 3 & fl.rich > 3) %>%
-  unite(webID, c(site, dayofyear, gdd.cum, year, size), sep = "_") %>%
-  dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs() %>% # convert to to bipartite web
-  map(networklevel, weighted = TRUE, level = "higher", index = c("niche overlap", "C score", "H2")) %>% # We probably care mainly about the BB level of the network
-  data.frame() %>%
-  t() %>%
-  data.frame() %>%
-  rownames_to_column() %>%
-  dplyr::select(id = rowname, everything()) %>%
-  as_tibble() %>%
-  gather(metric, value, -id) %>%
-  separate(id, into = c("site", "yday", "gdd.cum", "year", "size"), sep = "_") %>%
-  mutate(size = as.numeric(size),
-         year = factor(year),
-         gdd.cum = as.numeric(gdd.cum),
-         yday = as.numeric(yday),
-         site = as.factor(site)) %>%
-  left_join(site_data) %>%
-  left_join(abund_div)
-
-### Can we explain network metrics with abundance and diversity data?
-# BB rich
-ggplot(per.transect_netmet, aes(bb.rich, value)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_grid(metric ~ year, scales = "free")
-# FL rich
-ggplot(per.transect_netmet, aes(fl.rich, value)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_grid(metric ~ year, scales = "free")
-# BB abund
-ggplot(per.transect_netmet, aes(log(bb.abund), value)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_grid(metric ~ year, scales = "free")
-# FL abund
-ggplot(per.transect_netmet, aes(log(fl.abund), value)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_grid(metric ~ year, scales = "free")
-# Per-capita FL abund
-ggplot(per.transect_netmet, aes(log(fl.per.bb), value)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_grid(metric ~ year, scales = "free")
-
-### Can we explain network metrics with elevation and/or time?
-ggplot(per.transect_netmet, aes(elev.mean, value, color = log(size))) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(metric ~ year, scales = "free")
-
-ggplot(per.transect_netmet, aes(log(size), value, color = log(size))) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(metric ~ year, scales = "free")
-
-ggplot(filter(per.transect_netmet, size > 1), 
-       aes(yday, value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(metric ~ year, scales = "free")
-
-ggplot(filter(per.transect_netmet, size > 10), 
-       aes(gdd.cum, value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(metric ~ year, scales = "free")
-
-ggplot(filter(per.transect_netmet, size > 10), 
-       aes(yday, elev.mean)) +
-  geom_point(aes(color = value)) +
-  geom_smooth() +
-  facet_grid(metric ~ year, scales = "free")
-```
-
-#### Species level
-```{r eval=FALSE, include=FALSE}
-per.transect_spmet <- net %>%
-  group_by(site, dayofyear, gdd.cum, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-  group_by(site, dayofyear, gdd.cum, year) %>%
-  mutate(size = sum(freq),
-         bb.rich = length(unique(bb.sp)),
-         fl.rich = length(unique(plant.sp)),
-         total.rich = bb.rich + fl.rich) %>%
-  filter(bb.rich > 5 & fl.rich > 5) %>%
-  unite(webID, c(site, dayofyear, gdd.cum, year, size), sep = "_") %>%
-  dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs() %>% # convert to to bipartite web
-  map(specieslevel, level = "higher", index = "d") %>% # We probably care mainly about the BB level of the network
-  map(rownames_to_column) %>%
-  map(as_tibble) %>%
-  bind_rows(.id = "id") %>% # Collapse list into big data frame
-  rename(sp = rowname) %>%
-  gather(metric, value, -c(sp, id)) %>% # get mean value across iterations for each metric
-  separate(id, sep = "_", into = c("site", "yday", "gdd.cum", "year", "size")) %>%
-  mutate(size = as.numeric(size),
-         year = factor(year),
-         gdd.cum = as.numeric(gdd.cum),
-         yday = as.numeric(yday)) %>%
-  left_join(site_data) %>%
-  left_join(bb_traits, by = c("sp" = "bb.sp")) %>%
-  left_join(abund_div)
-
-per.transect_spmet.mean <- per.transect_spmet %>%
-  group_by(site, yday, gdd.cum, year) %>%
-  summarize(value = mean(value)) %>%
-  left_join(site_data)
-
-ggplot(filter(per.transect_spmet), 
-       aes(log(fl.abund), value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(. ~ year, scales = "free")
-
-ggplot(filter(per.transect_spmet), 
-       aes(elev.mean, size)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(. ~ year, scales = "free")
-
-ggplot(filter(per.transect_spmet.mean), 
-       aes(yday, value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(. ~ year, scales = "free")
-
-ggplot(filter(per.transect_spmet), 
-       aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(. ~ year, scales = "free")
-
-ggplot(filter(per.transect_spmet, size >= 10), 
-       aes(gdd.cum, value, color = log(size))) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(metric ~ year, scales = "free")
-
-ggplot(filter(per.transect_spmet, size >= 10), 
-       aes(yday, value, color = log(size))) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(metric ~ year, scales = "free")
-```
-
-### Phase x elev network analysis -- this only makes sense in 2012, where we have sufficient sampling across the whole colony lifecycle
-```{r eval=FALSE, include=FALSE}
-### Prepare pxe abundance and diversity table
-pxe_abund_div <- abund_div %>%
-  select(-gdd.cum, -yday, -week, -month, -date, -elev.class2) %>%
-  gather(ad.metric, ad.value, -c(site, year, phase, elev.mean)) %>%
-  group_by(site, phase, year, ad.metric) %>%
-  summarize(ad.value = mean(ad.value, na.rm = TRUE))
-
-pxe_net <- rare_net_nt2(filter(net, year == 2012), 3, 20) %>%
-  full_join((pxe_abund_div))
-
-ggplot(filter(pxe_net, metric == "H2"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(year ~ phase, scales = "free") +
-  xlab("Elevation") +
-  ylab("H2")
-
-ggplot(filter(pxe_net, metric == "C.score.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(year ~ phase, scales = "free") +
-  xlab("Elevation") +
-  ylab("C.score")
-
-ggplot(filter(pxe_net, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(year ~ phase, scales = "free") +
-  xlab("Elevation") +
-  ylab("Niche overlap")
-
-pxe_sp <- rare_net_sp2(net, 4, 20)  %>%
-  full_join((pxe_abund_div))
-
-ggplot(filter(pxe_sp, 
-              metric == "d" &
-                sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit") &
-                ad.metric == "fl.rich"), 
-       aes(ad.value, value, color = year)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +
-  facet_grid(sp ~ phase, scales = "free") +
-  ylim(c(0,1)) +
-  xlab("Floral richness") +
-  ylab("d'")
-
-pxe_sp_mean <- pxe_sp %>%
-  filter(sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")) %>%
-  group_by(site, year, phase, elev.mean, metric) %>%
-  summarize(value = mean(value, na.rm = TRUE))
-
-ggplot(filter(pxe_sp_mean, metric == "d"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(year ~ phase, scales = "free") +
-  xlab("Elevation") +
-  ylab("Mean d'")
-
-ggplot(filter(pxe_sp_mean, metric == "species.strength"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(year ~ phase, scales = "free") +
-  xlab("Elevation") +
-  ylab("Mean species strength'")
-
-ggplot(filter(pxe_sp, 
-              metric == "species.strength",
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
-       aes(elev.mean, value, color = year)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(sp ~ phase, scales = "free") +
-  xlab("Elevation") +
-  ylab("Species strength")
-
-ggplot(filter(pxe_sp, 
-              metric == "d",
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
-       aes(elev.mean, value, color = year)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(sp ~ phase, scales = "free") +
-  ylim(c(0,1)) +
-  xlab("Elevation") +
-  ylab("d'")
-
-ggplot(filter(pxe_sp, 
-              metric == "d",
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
-       aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(sp ~ phase, scales = "free") +
-  ylim(c(0,1)) +
-  xlab("Elevation") +
-  ylab("d'")
-
-ggplot(filter(pxe_sp, 
-              metric == "d",
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
-       aes(elev.mean, value, color = phase)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(year ~ sp, scales = "free") +
-  theme_light() +
-  xlab("Elevation") +
-  ylab("d'")
-```
-
-```{r}
-
-# network level PxE
-rare_net_nt_pxe <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  t_filter <- net %>%
-    select(site, date, phase, year) %>%
-    unique() %>%
-    group_by(site, phase, year) %>%
-    mutate(transects = n()) %>%
-    filter(transects >= n)
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      semi_join(t_filter) %>%
-      select(site, date, phase, year) %>%
-      unique() %>%
-      group_by(site, phase, year) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, phase, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      group_by(site, phase, year) %>%
-      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
-      ungroup() %>%
-      unite(webID, c(site, phase, year), sep = "_") %>%
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(networklevel, weighted = TRUE, level = "higher",
-          index = c("web asymmetry", "weighted NODF", "H2", "C score", "niche overlap", "generality")) %>% 
-      data.frame() %>%
-      t() %>%
-      data.frame() %>%
-      rownames_to_column() %>%
-      dplyr::select(id = rowname, everything()) %>%
-      as_tibble() %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out  %>%
-      gather(metric, value, -c(id, iteration)) %>% # get mean value across iterations for each metric
-      group_by(id, metric) %>%
-      summarize(value = mean(value)) %>%
-      separate(id, sep = "_", into = c("site", "phase", "year")) %>%
-      mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
-      left_join(site_data)
-}
-
-# species level version 2
-rare_net_sp2 <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  t_filter <- net %>%
-    select(site, date, phase, year) %>%
-    unique() %>%
-    group_by(site, phase, year) %>%
-    mutate(transects = n()) %>%
-    filter(transects >= n)
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      semi_join(t_filter) %>%
-      select(site, date, phase, year) %>%
-      unique() %>%
-      group_by(site, phase, year) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, phase, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      group_by(site, phase, year) %>%
-      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
-      ungroup() %>%
-      unite(webID, c(site, phase, year), sep = "_") %>%
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(specieslevel, level = "higher", index = c("d", "species strength")) %>%
-      map(rownames_to_column) %>%
-      map(as_tibble) %>%
-      bind_rows(.id = "id") %>% # Collapse list into big data frame
-      rename(sp = rowname) %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out %>%
-    gather(metric, value, -c(sp, id, iteration)) %>% # get mean value across iterations for each metric
-    group_by(sp, id, metric) %>%
-    summarize(value = mean(value)) %>%
-    separate(id, sep = "_", into = c("site", "phase", "year")) %>%
-    mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
-    left_join(site_data) %>%
-    left_join(bb_traits, by = c("sp" = "bb.sp"))
-}
-
-```

### 2020-07-02 analyses/community_GAM_analysis.Rmd
broke down analyses into separate scripts; did some other things

---
title: "community_GAM_analysis"
output:
  pdf_document: default
  html_document: default
---

### Prepare environment
```{r setup, include=FALSE, echo=FALSE}
library(mgcv)
library(mgcViz)
library(tidymv)
#library(lme4)
#library(lmerTest)
library(tidyverse)
library(lubridate)
```

### Climate and weather data
```{r, include=FALSE, echo=FALSE}
gdd <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year)) %>%
  dplyr::select(site, year, date, gdd.cum) %>%
  semi_join(read_csv("../processed_data/network.csv"), by = "site")
# 
# temperature <- read_csv("../processed_data/temperature.csv")
# 
# gdd_total <- gdd %>%
#   group_by(site, year) %>%
#   summarize(gdd.total = max(gdd.cum))
```

### Load site data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))

# %>% # turn elav.class into an ordered factor
#   left_join(gdd_total, by = "site")
```

### Load network and survey data
```{r, echo=FALSE, include=FALSE}
bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
  dplyr::select(subgenus, bb.sp, pbl.w, pbl.w.class)

fl_traits <- read.csv("../processed_data/floral_k_type_nom_fix.csv") %>%
  dplyr::select(plant.sp, k.type) %>%
  mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"),
         k.type.ss = str_extract(k.type, "[-+]?[0-9]*"))

fl_corolla_depth <- read.csv("../processed_data/flower_morphology_GitaBenadi.csv") %>%
  group_by(plant.sp = Species) %>%
  summarize(corolla.depth = mean(LN))

fl_tax <- read_csv("../processed_data/floral_tax.csv")

breaks <- read_csv("./output/breaks.csv") %>%
  mutate(year = factor(year))

net <- read_csv("../processed_data/network.csv") %>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  unite(site.date, site, date, sep = "_", remove = FALSE) %>%
  mutate(year = factor(year)) %>% # convert year to factor
  full_join(bb_traits, by = c("bb.sp" = "bb.sp")) %>%
  left_join(site_data, by = c("site")) %>%
  left_join(gdd, by = c("site", "date", "year")) %>%
  left_join(breaks, by = c("site", "year")) %>%
  left_join(fl_traits) %>%
  left_join(fl_corolla_depth) %>%
  left_join(fl_tax, by = "plant.genus") %>%
  #left_join(temperature, by = c("site", "date")) %>%
  ungroup() %>%
  mutate(phase = case_when(
    gdd.cum < break1.gdd ~ "founding",
    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
    gdd.cum > break2.gdd ~ "reproductive"
  )) %>%
  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
  mutate(caste = case_when(
    caste == "worker" ~ "worker",
    caste == "male" ~ "reproductive",
    caste == "youngqueen" ~ "reproductive",
    caste == "queen" ~ "foundress",
    caste == "oldqueen" ~ "foundress"
  )) %>%
  rename(yday = dayofyear) %>%
  arrange(site, date)

survey <- read_csv("../processed_data/floral_survey.csv") %>%
  mutate(year = factor(year)) %>% # convert year to factor
  semi_join(net, by = "plant.sp") %>% # Consider only plants visited at a given site-date
  left_join(gdd, by = c("site", "date", "year")) %>%
  left_join(breaks, by = c("site", "year")) %>%
  left_join(site_data, by = c("site")) %>%
  left_join(fl_traits) %>%
  left_join(fl_traits) %>%
  left_join(fl_tax) %>%
  #left_join(temperature, by = c("site", "date")) %>%
  ungroup() %>%
  mutate(phase = case_when(
    gdd.cum < break1.gdd ~ "founding",
    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
    gdd.cum > break2.gdd ~ "reproductive"
  )) %>%
  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
  rename(yday = dayofyear) %>%
  arrange(site, date)
```

### FL and BB abundance / diversity
```{r echo=FALSE, include=FALSE}
### Per-transect Bombus abundance
bb_abund <- net %>%
  group_by(site, date, yday = dayofyear, gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd) %>%
  summarize(bb.abund = n())

ggplot(bb_abund, aes(log(top.fl.abund + 0.01), bb.abund, color = year)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(elev.class2~phase, scales = "free")

ggplot(bb_abund, aes(log(top.fl.abund + 0.01), bb.abund, color = year)) +
  geom_point(alpha = 0.5) +
  #geom_smooth(method = "lm") +
  geom_smooth() +
  facet_wrap(~bb.sp, scales = "free")

#### Species-level 
bb_abund_sp <- net %>%
  group_by(site, date, yday, gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd, bb.sp) %>%
  summarize(bb.abund = n()) %>%
  ungroup() %>%
  mutate(bb.sp = factor(bb.sp))

bb_abund_sp_workers <- net %>%
  filter(caste == "worker") %>%
  group_by(site, date, yday = dayofyear, gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd, bb.sp) %>%
  summarize(bb.abund = n()) %>%
  ungroup() %>%
  mutate(bb.sp = factor(bb.sp))

bb_abund_caste <- net %>%
  group_by(site, date, yday = dayofyear, gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd, caste) %>%
  summarize(bb.abund = n()) %>%
  ungroup() %>%
  mutate(caste = factor(caste))

bb_abund_caste_sp <- net %>%
  group_by(site, date, yday = dayofyear, gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd, caste, bb.sp) %>%
  summarize(bb.abund = n()) %>%
  ungroup() %>%
  mutate(caste = factor(caste),
         bb.sp = factor(bb.sp))

### Per-transect Bombus richness
bb_rich <- net %>%
  group_by(site, date, yday = dayofyear, gdd.cum, year, elev.class2,
           elev.mean, break1.gdd, break2.gdd) %>%
  summarize(bb.rich = length(unique(bb.sp))) %>%
  mutate(year2010 = ifelse(year == 2010, 1, 0),
         year2011 = ifelse(year == 2011, 1, 0),
         year2012 = ifelse(year == 2012, 1, 0))

ggplot(bb_rich, aes(yday, bb.rich)) +
  geom_point() +
  facet_wrap(~year)

### Per-transect floral abund
fl_abund <- survey %>%
  group_by(site, date, yday = dayofyear, gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd) %>%
  summarize(fl.abund = sum(flower.cover)) %>%
  mutate(year2010 = ifelse(year == 2010, 1, 0),
         year2011 = ifelse(year == 2011, 1, 0),
         year2012 = ifelse(year == 2012, 1, 0))

### Per-transect floral richness
fl_rich <- survey %>%
  filter(flower.cover > 0) %>%
  group_by(site, date, yday = dayofyear, gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd) %>%
  summarize(fl.rich = n())  %>%
  mutate(year2010 = ifelse(year == 2010, 1, 0),
         year2011 = ifelse(year == 2011, 1, 0),
         year2012 = ifelse(year == 2012, 1, 0))
```

### Get most important plants for each BB species
```{r}
# Per transect visitation rates of flower species per BB species
visit_rates <- net %>%
  group_by(bb.sp, plant.sp, year, site, date, yday, gdd.cum, elev.class2, 
           elev.mean, break1.gdd, break2.gdd, phase) %>%
  summarize(visits = n()) %>%
  group_by(year, site, date, bb.sp, yday, gdd.cum, elev.class2, 
           elev.mean, break1.gdd, break2.gdd) %>%
  mutate(prop.visits = visits/sum(visits)) %>%
  arrange(site, year, date, bb.sp, -prop.visits) %>%
  mutate(cum.prop.visits = cumsum(prop.visits))

### Per-transect floral abund
fl_abund_sp <- survey %>%
  group_by(site, date, year, plant.sp, yday, gdd.cum, elev.class2, 
           elev.mean, break1.gdd, break2.gdd, phase) %>%
  summarize(fl.abund = sum(flower.cover))

### Preference per transect per species
prop_fl_abund <- visit_rates %>%
  left_join(fl_abund_sp) %>%
  na.omit() %>% # drop the samples without floral survey complement
  group_by(site, date, year, bb.sp, yday, gdd.cum, elev.class2, 
           elev.mean, break1.gdd, break2.gdd, phase) %>%
  mutate(prop.fl.abund = fl.abund/sum(fl.abund))

### Median preference per bb.sp per plant.sp
prop_fl_abund.med <- prop_fl_abund %>%
  group_by(bb.sp, plant.sp, year) %>%
  summarize(visits.median = median(visits),
            prop.visits.median = median(prop.visits),
            prop.fl.abund.median = median(prop.fl.abund)) %>%
  filter(prop.fl.abund.median < 1)

prop_fl_abund_median.lm <- lm(prop.visits.median ~ poly(prop.fl.abund.median, 2) + year, 
                       data = prop_fl_abund.med)
summary(prop_fl_abund_median.lm)
plot(prop_fl_abund_median.lm)

### Extract residuals
prop_fl_abund_mean.resid <- residuals(prop_fl_abund_median.lm) %>%
  data.frame() %>%
  as_tibble() %>%
  rename(preference = 1)

### Join residuals back into data
preference <- prop_fl_abund.mean %>%
  filter(prop.fl.abund.mean != 1) %>%
  bind_cols(prop_fl_abund_mean.resid)

### Join preference data to floral abundance data
fl_abund_sp_pref <- fl_abund_sp %>%
  inner_join(preference) %>%
  mutate(fl.abund.weighted = fl.abund^(1 + preference)) # Is this the right way to model preference?

ggplot(fl_abund_sp_pref, aes(preference)) +
  geom_histogram()

### Summarize weighted fl.abund by site*date (i.e. transect)
fl_abund_pref <- fl_abund_sp_pref %>%
  group_by(site, date, year, bb.sp, yday, gdd.cum, elev.class2, 
           elev.mean, break1.gdd, break2.gdd, phase) %>%
  summarize(fl.abund = sum(fl.abund),
            fl.abund.weighted = sum(fl.abund.weighted)) %>%
  left_join(bb_abund_sp) %>%
  replace_na(list(bb.abund = 0))

ggplot(fl_abund_pref, aes(log(fl.abund.weighted + 0.01), bb.abund, color = year)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp, scales = "free")

ggplot(fl_abund_pref, aes(log(fl.abund + 0.01), bb.abund, color = year)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp, scales = "free")

fl_abund_pref.sum <- fl_abund_pref %>%
  group_by(site, year, bb.sp) %>%
  summarize(tot.fl.abund.weighted = sum(fl.abund.weighted),
            tot.fl.abund = sum(fl.abund),
            tot.bb.abund = sum(bb.abund))

ggplot(fl_abund_pref.sum, aes(log(tot.fl.abund.weighted + 0.01), tot.bb.abund, color = year)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp, scales = "free")

ggplot(fl_abund_pref.sum, aes(log(tot.fl.abund + 0.01), tot.bb.abund, color = year)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp, scales = "free")

ggplot(filter(preference, plant.sp == "Trifolium pratense"), 
       aes(bb.sp, preference, color = year)) +
  geom_point()

ggplot(filter(preference, plant.sp == "Acinos alpinus"), 
       aes(bb.sp, preference, color = year)) +
  geom_point()

ggplot(filter(prop_fl_abund, prop.fl.abund != 1), aes(prop.fl.abund, prop.visits)) +
  geom_point(alpha = 0.05) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) 
  #geom_smooth()

ggplot(filter(prop_fl_abund.mean, prop.fl.abund.mean != 1), 
       aes(prop.fl.abund.mean, prop.visits.mean)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "lm")

```

**The shape of floral resource variation in time, space, and space-time** \

If we consider floral resource abundance with respect to time alone, we basically get a monotonic and unimodal response
```{r}
fl_abund_time <- bam(log(fl.abund + 0.01) ~
                       year +
                       s(yday, 
                         bs = "tp", 
                         k = 10, m = 2),
                     data = fl_abund,
                     discrete = TRUE,
                     method = "fREML") %>% getViz()

check.gamViz(fl_abund_time)
summary(fl_abund_time)
print(plot(fl_abund_time , allTerms = TRUE), pages = 1)
plot(sm(fl_abund_time, 1)) + 
  l_points() + 
  l_fitLine() + 
  l_ciLine()

fl_abund_time_2010 <- bam(log(fl.abund + 0.01) ~
                            s(yday, bs = "tp", k = 10, m = 2),
                          data = filter(fl_abund, year == 2010),
                          discrete = TRUE,
                          method = "fREML") %>% getViz()

check.gamViz(fl_abund_time_2010)
summary(fl_abund_time_2010)
print(plot(fl_abund_time_2010 , allTerms = TRUE), pages = 1)

fl_abund_time_2011 <- bam(log(fl.abund + 0.01) ~
                            s(yday, bs = "tp", k = 10, m = 2),
                          data = filter(fl_abund, year == 2011),
                          discrete = TRUE,
                          method = "fREML") %>% getViz()

check.gamViz(fl_abund_time_2011)
summary(fl_abund_time_2011)
print(plot(fl_abund_time_2011 , allTerms = TRUE), pages = 1)

fl_abund_time_2012 <- bam(log(fl.abund + 0.01) ~
                            s(yday, bs = "tp", k = 10, m = 2),
                          data = filter(fl_abund, year == 2012),
                          discrete = TRUE,
                          method = "fREML") %>% getViz()

check.gamViz(fl_abund_time_2012)
summary(fl_abund_time_2012)
print(plot(fl_abund_time_2012 , allTerms = TRUE), pages = 1)

ggplot(fl_abund, aes(yday, fl.abund, color = year)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~site, scales = "free")
```

If we look at floral abundance solely in terms of elevation, we get a more wiggly picture, with peaks ~1000 m and ~1600 m.
```{r}
fl_abund_elev <- bam(log(fl.abund + 0.01) ~
                       year +
                       s(elev.mean, 
                         bs = "tp", 
                         k = 10, m = 2),
                     data = fl_abund,
                     discrete = TRUE,
                     method = "fREML") %>% getViz()

check.gamViz(fl_abund_elev)
summary(fl_abund_elev)
print(plot(fl_abund_elev, allTerms = TRUE), pages = 1)
plot(sm(fl_abund_elev, 1)) + 
  l_points() + 
  l_fitLine() + 
  l_ciLine()

fl_abund_elev_2010 <- bam(log(fl.abund + 0.01) ~
                            s(elev.mean, bs = "tp", k = 10, m = 2),
                          data = filter(fl_abund, year == 2010),
                          discrete = TRUE,
                          method = "fREML") %>% getViz()

check.gamViz(fl_abund_elev_2010)
summary(fl_abund_elev_2010)
print(plot(fl_abund_elev_2010 , allTerms = TRUE), pages = 1)

fl_abund_elev_2011 <- bam(log(fl.abund + 0.01) ~
                            s(elev.mean, bs = "tp", k = 10, m = 2),
                          data = filter(fl_abund, year == 2011),
                          discrete = TRUE,
                          method = "fREML") %>% getViz()

check.gamViz(fl_abund_elev_2011)
summary(fl_abund_elev_2011)
print(plot(fl_abund_elev_2011 , allTerms = TRUE), pages = 1)

fl_abund_elev_2012 <- bam(log(fl.abund + 0.01) ~
                            s(elev.mean, bs = "tp", k = 10, m = 2),
                          data = filter(fl_abund, year == 2012),
                          discrete = TRUE,
                          method = "fREML") %>% getViz()

check.gamViz(fl_abund_elev_2012)
summary(fl_abund_elev_2012)
print(plot(fl_abund_elev_2012 , allTerms = TRUE), pages = 1)

```

But now let's look at space and time together in a 2D GAM.
```{r}
# Call models
fl_abund_2D <- bam(log(fl.abund + 0.01) ~ year +
                            te(yday, elev.mean, 
                               bs= c("tp", "tp"), 
                               k=c(10, 10), m=2),
                    data = fl_abund,
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(fl_abund_2D)
summary(fl_abund_2D)
plot(sm(fl_abund_2D, 1), too.far = 0.05) + l_fitRaster() + l_fitContour() + l_points()
plot(pterm(fl_abund_2D, 1))

fl_abund_2D_2010 <- bam(log(fl.abund + 0.01) ~ te(yday, elev.mean, 
                                                  bs= c("tp", "tp"), 
                                                  k=c(10, 10), m=2),
                        data = filter(fl_abund, year == 2010),
                        discrete = TRUE,
                        method = "fREML") %>% getViz()

check.gamViz(fl_abund_2D_2010)
summary(fl_abund_2D_2010)
plot(sm(fl_abund_2D_2010, 1)) + 
  l_fitRaster() + 
  l_fitContour() + 
  l_points()

fl_abund_2D_2011 <- bam(log(fl.abund + 0.01) ~ te(yday, elev.mean, 
                                                  bs= c("tp", "tp"), 
                                                  k=c(10, 10), m=2),
                        data = filter(fl_abund, year == 2011),
                        discrete = TRUE,
                        method = "fREML") %>% getViz()

check.gamViz(fl_abund_2D_2011)
summary(fl_abund_2D_2011)
plot(sm(fl_abund_2D_2011, 1)) + 
  l_fitRaster() + 
  l_fitContour() + 
  l_points()

fl_abund_2D_2012 <- bam(log(fl.abund + 0.01) ~ te(yday, elev.mean, 
                                                  bs= c("tp", "tp"), 
                                                  k=c(10, 10), m=2),
                        data = filter(fl_abund, year == 2012),
                        discrete = TRUE,
                        method = "fREML") %>% getViz()

check.gamViz(fl_abund_2D_2012)
summary(fl_abund_2D_2012)
plot(sm(fl_abund_2D_2012, 1)) + 
  l_fitRaster() + 
  l_fitContour() + 
  l_points()
```

How does the pooled abundance of each bumble bee species' top 5 floral hosts vary in time and elevation?
```{r}
fl_abund_2D_2010_top <- bam(log(top.fl.abund + 0.01) ~ te(yday, elev.mean, 
                                                  by = factor(bb.sp),
                                                  bs= c("tp", "tp"), 
                                                  k=c(10, 10), m=1),
                        data = filter(fl_abund_top_plants_pooled, year == 2010),
                        discrete = TRUE,
                        method = "fREML") %>% getViz()

check.gamViz(fl_abund_2D_2010_top)
summary(fl_abund_2D_2010_top)
print(plot(fl_abund_2D_2010_top , allTerms = TRUE), pages = 1)

fl_abund_2D_2011_top <- bam(log(top.fl.abund + 0.01) ~ te(yday, elev.mean, 
                                                  by = factor(bb.sp),
                                                  bs= c("tp", "tp"), 
                                                  k=c(10, 10), m=1),
                        data = filter(fl_abund_top_plants_pooled, year == 2011),
                        discrete = TRUE,
                        method = "fREML") %>% getViz()

check.gamViz(fl_abund_2D_2011_top)
summary(fl_abund_2D_2011_top)
print(plot(fl_abund_2D_2011_top , allTerms = TRUE), pages = 1)

fl_abund_2D_2012_top <- bam(log(top.fl.abund + 0.01) ~ te(yday, elev.mean, 
                                                  by = factor(bb.sp),
                                                  bs= c("tp", "tp"), 
                                                  k=c(10, 10), m=1),
                        data = filter(fl_abund_top_plants_pooled, year == 2012),
                        discrete = TRUE,
                        method = "fREML") %>% getViz()

check.gamViz(fl_abund_2D_2012_top)
summary(fl_abund_2D_2012_top)
print(plot(fl_abund_2D_2012_top , allTerms = TRUE), pages = 1)

```

Now let's do the same with BB abundance. First by time.
```{r}
bb_abund_time_2010 <- bam(bb.abund ~ s(yday, bs = "tp", k = 20, m = 2),
                          data = filter(bb_abund, year == 2010),
                          discrete = TRUE,
                          family = "poisson",
                          method = "fREML") %>% getViz()

bb_abund_time_2011 <- bam(bb.abund ~ s(yday, bs = "tp", k = 20, m = 2),
                          data = filter(bb_abund, year == 2011),
                          discrete = TRUE,
                          family = "poisson",
                          method = "fREML") %>% getViz()

bb_abund_time_2012 <- bam(bb.abund ~ s(yday, bs = "tp", k = 20, m = 2),
                          data = filter(bb_abund, year == 2012),
                          discrete = TRUE,
                          family = "poisson",
                          method = "fREML") %>% getViz()

check.gamViz(bb_abund_time_2010)
summary(bb_abund_time_2010)
print(plot(bb_abund_time_2010 , allTerms = TRUE), pages = 1)

check.gamViz(bb_abund_time_2011)
summary(bb_abund_time_2011)
print(plot(bb_abund_time_2011 , allTerms = TRUE), pages = 1)

check.gamViz(bb_abund_time_2012)
summary(bb_abund_time_2012)
print(plot(bb_abund_time_2012 , allTerms = TRUE), pages = 1)

ggplot(bb_abund_sp, aes(yday, bb.abund, color = bb.sp)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_grid(year~site, scales = "free")

ggplot(filter(bb_abund_sp, bb.sp == "prat"), aes(yday, bb.abund)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_grid(year~site, scales = "free")
```

Then by elevation...this kinda sucks
```{r}
bb_abund_elev_2010 <- bam(bb.abund ~ s(elev.mean, bs = "tp", k = 10, m = 2),
                          data = filter(bb_abund, year == 2010),
                          discrete = TRUE,
                          family = "poisson",
                          method = "fREML") %>% getViz()

bb_abund_elev_2011 <- bam(bb.abund ~ s(elev.mean, bs = "tp", k = 10, m = 2),
                          data = filter(bb_abund, year == 2011),
                          discrete = TRUE,
                          family = "poisson",
                          method = "fREML") %>% getViz()

bb_abund_elev_2012 <- bam(bb.abund ~ s(elev.mean, bs = "tp", k = 10, m = 2),
                          data = filter(bb_abund, year == 2012),
                          discrete = TRUE,
                          family = "poisson",
                          method = "fREML") %>% getViz()

check.gamViz(bb_abund_elev_2010)
summary(bb_abund_elev_2010)
print(plot(bb_abund_elev_2010 , allTerms = TRUE), pages = 1)

check.gamViz(bb_abund_elev_2011)
summary(bb_abund_elev_2011)
print(plot(bb_abund_elev_2011 , allTerms = TRUE), pages = 1)

check.gamViz(bb_abund_elev_2012)
summary(bb_abund_elev_2012)
print(plot(bb_abund_elev_2012 , allTerms = TRUE), pages = 1)
```

But now let's look at space and time together in a 2D GAM.
```{r}
### Pooled across species
bb_abund_2010 <- bam(bb.abund ~ te(yday, elev.mean, 
                            bs = c("tp", "tp"), k = c(10, 10)),
                       data = filter(bb_abund, year == 2010),
                       discrete = TRUE,
                       method = "fREML",
                       family = "poisson",
                       select = TRUE) %>% getViz()

check.gamViz(bb_abund_2010)
summary(bb_abund_2010)
print(plot(bb_abund_2010, allTerms = TRUE), pages = 1)

bb_abund_2011 <- bam(bb.abund ~ te(yday, elev.mean, 
                            bs = c("tp", "tp"), k = c(10, 10)),
                       data = filter(bb_abund, year == 2011),
                       discrete = TRUE,
                       method = "fREML",
                       family = "poisson",
                       select = TRUE) %>% getViz()

check.gamViz(bb_abund_2011)
summary(bb_abund_2011)
print(plot(bb_abund_2011, allTerms = TRUE), pages = 1)

bb_abund_2012 <- bam(bb.abund ~ te(yday, elev.mean, 
                            bs = c("tp", "tp"), k = c(10, 10)),
                       data = filter(bb_abund, year == 2012),
                       discrete = TRUE,
                       method = "fREML",
                       family = "poisson",
                       select = TRUE) %>% getViz()

check.gamViz(bb_abund_2012)
summary(bb_abund_2012)
print(plot(bb_abund_2012, allTerms = TRUE), pages = 1)
# 
# 
# bb_abund_2010.caste <- bam(bb.abund ~ caste + 
#                              te(yday, elev.mean, by = caste,
#                                 bs = c("tp", "tp"), k = c(10, 10)),
#                            data = filter(bb_abund_caste, year == 2010),
#                            discrete = TRUE,
#                            method = "fREML",
#                            family = "poisson",
#                            select = TRUE) %>% getViz()
#                          
# check.gamViz(bb_abund_2010.caste)
# summary(bb_abund_2010.caste)
# print(plot(bb_abund_2010.caste, allTerms = TRUE), pages = 1)
# 
# bb_abund_2011.caste <- bam(bb.abund ~ caste + 
#                              te(yday, elev.mean, by = caste,
#                                 bs = c("tp", "tp"), k = c(10, 10)),
#                            data = filter(bb_abund_caste, year == 2011),
#                            discrete = TRUE,
#                            method = "fREML",
#                            family = "poisson",
#                            select = TRUE) %>% getViz()
#                          
# check.gamViz(bb_abund_2011.caste)
# summary(bb_abund_2011.caste)
# print(plot(bb_abund_2011.caste, allTerms = TRUE), pages = 1)
# 
# 
# bb_abund_2012.caste <- bam(bb.abund ~ caste + 
#                              te(yday, elev.mean, by = caste,
#                                 bs = c("tp", "tp"), k = c(10, 10)),
#                            data = filter(bb_abund_caste, year == 2012),
#                            discrete = TRUE,
#                            method = "fREML",
#                            family = "poisson",
#                            select = TRUE) %>% getViz()
#                          
# check.gamViz(bb_abund_2012.caste)
# summary(bb_abund_2012.caste)
# print(plot(bb_abund_2012.caste, allTerms = TRUE), pages = 1)
#

### Species level
bb_abund_sp <- net %>%
  group_by(site, date, yday = dayofyear,  gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd, bb.sp) %>%
  summarize(bb.abund = n()) %>%
  ungroup() %>%
  mutate(bb.sp = factor(bb.sp)) %>%
  spread(bb.sp, bb.abund) %>%
  gather(bb.sp, bb.abund, -c(site, date, yday,  
                             gdd.cum, year, elev.class2, elev.mean, 
                             break1.gdd, break2.gdd)) %>%
  replace_na(list(bb.abund = 0)) %>%
  mutate(bb.sp = factor(bb.sp)) %>%
  group_by(bb.sp, year) %>%
  filter(sum(bb.abund) > 5) # remove rare species (on a per-year basis)

# bb_fl_abund_sp <- left_join(bb_abund_sp, fl_abund)
# bb_fl_abund_sp_2010 <- left_join(bb_abund_sp, fl_abund) %>% filter(year == 2010)
# bb_fl_abund_sp_2011 <- left_join(bb_abund_sp, fl_abund) %>% filter(year == 2011)
# bb_fl_abund_sp_2012 <- left_join(bb_abund_sp, fl_abund) %>% filter(year == 2012)

bb_abund_sp_2010 <- bam(bb.abund ~ bb.sp +
                                 te(yday, elev.mean, by = bb.sp, 
                                    bs = c("tp", "tp"), k = c(10, 10), m = 1),
                               data = filter(bb_abund_sp, year == 2010),
                               discrete = TRUE,
                               method = "fREML",
                               family = "poisson") %>% getViz()

check.gamViz(bb_abund_sp_2010)
summary(bb_abund_sp_2010)
print(plot(bb_abund_sp_2010, allTerms = TRUE), pages =4)

bb_abund_sp_2011 <- bam(bb.abund ~ bb.sp +
                                 te(yday, elev.mean, by = bb.sp, 
                                    bs = c("tp", "tp"), k = c(10, 10), m = 1),
                               data = filter(bb_abund_sp, year == 2011),
                               discrete = TRUE,
                               method = "fREML",
                               family = "poisson") %>% getViz()

check.gamViz(bb_abund_sp_2011)
summary(bb_abund_sp_2011)
print(plot(bb_abund_sp_2011, allTerms = TRUE), pages = 4)

bb_abund_sp_2012 <- bam(bb.abund ~ bb.sp +
                                 te(yday, elev.mean, by = bb.sp, 
                                    bs = c("tp", "tp"), k = c(10, 10), m = 1),
                               data = filter(bb_abund_sp, year == 2012),
                               discrete = TRUE,
                               method = "fREML",
                               family = "poisson") %>% getViz()

check.gamViz(bb_abund_sp_2012)
summary(bb_abund_sp_2012)
print(plot(bb_abund_sp_2012, allTerms = TRUE), pages = 4)
```

And BB richness
```{r}
bb_rich_2010 <- bam(bb.rich ~ te(yday, elev.mean, 
                            bs = c("tp", "tp"), k = c(10, 10), m = 1),
                       data = filter(bb_rich, year == 2010),
                       discrete = TRUE,
                       method = "fREML",
                       family = "poisson",
                       select = TRUE) %>% getViz()

check.gamViz(bb_rich_2010)
summary(bb_rich_2010)
print(plot(bb_rich_2010, allTerms = TRUE), pages = 1)

bb_rich_2011 <- bam(bb.rich ~ te(yday, elev.mean, 
                            bs = c("tp", "tp"), k = c(10, 10), m = 1),
                       data = filter(bb_rich, year == 2011),
                       discrete = TRUE,
                       method = "fREML",
                       family = "poisson",
                       select = TRUE) %>% getViz()

check.gamViz(bb_rich_2011)
summary(bb_rich_2011)
print(plot(bb_rich_2011, allTerms = TRUE), pages = 1)

bb_rich_2012 <- bam(bb.rich ~ te(yday, elev.mean, 
                            bs = c("tp", "tp"), k = c(10, 10), m = 1),
                       data = filter(bb_rich, year == 2012),
                       discrete = TRUE,
                       method = "fREML",
                       family = "poisson",
                       select = TRUE) %>% getViz()

check.gamViz(bb_rich_2012)
summary(bb_rich_2012)
print(plot(bb_rich_2012, allTerms = TRUE), pages = 1)

```

To what extent is the relationship between BB abundance and elev*time reflect a relationship between BB abundance and floral abundance?
```{r}
bb_fl_abund <- left_join(bb_abund, fl_abund)
bb_fl_abund_sp <- left_join(bb_abund_sp, fl_abund)
bb_fl_abund.ktype <- left_join(bb_abund, fl_abund_ktype)
bb_fl_abund_sp.ktype <- left_join(bb_abund_sp, fl_abund_ktype)
bb_fl_abund.family <- left_join(bb_abund, fl_abund_family)
bb_fl_abund_sp.family <- left_join(bb_abund_sp, fl_abund_family)
bb_fl_abund_sp.genus <- left_join(bb_abund_sp, fl_abund_genus)

ggplot(bb_fl_abund, aes(log(fl.abund), bb.abund)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

ggplot(bb_fl_abund_sp, aes(log(fl.abund), bb.abund)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp)

ggplot(bb_fl_abund.ktype, aes(log(fl.abund), bb.abund)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~k.type.ss)

ggplot(bb_fl_abund_sp.ktype, aes(log(fl.abund), bb.abund)) +
  geom_point() +
  geom_smooth() +
  facet_grid(k.type.ss~bb.sp)

ggplot(bb_fl_abund.family, aes(log(fl.abund), bb.abund, color = year)) +
  #geom_point() +
  geom_smooth(se = FALSE, method = "lm") +
  facet_wrap(~plant.family)

ggplot(filter(bb_fl_abund_sp.family,
              bb.sp %in% c("pasc", "prat", "bss", "mont", "hort", "mont", "soro", "wurf", "psit") &
                plant.family %in% c("Rosaceae", "Lamiaceae", "Asteraceae")), 
       aes(log(fl.abund), bb.abund, color = year)) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  facet_grid(bb.sp~plant.family, scales = "free") +
  theme_gray(8)

### Let's see how tightly BB abundance is correlated with FL abundance within each site*year
ggplot(filter(bb_fl_abund, year == 2012), aes(fl.abund, bb.abund)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~site, scales = "free")

ggplot(filter(bb_fl_abund, year == 2011), aes(fl.abund, bb.abund)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~site, scales = "free")
```
Hmm...so, bumble bee abundance is correlated with floral abundance, but with a ton of noise. Perhaps that is due to the temporal patterns of colony growth.
```{r}
bb_fl_abund <- left_join(bb_abund, fl_abund)

bb_abund_2010_2 <- gam(bb.abund ~ te(yday, log(fl.abund + 0.01), bs = c("tp", "tp"), k = c(10, 10)),
                     data = filter(bb_fl_abund, year == 2010),
                     method = "REML",
                     family = "gaussian") %>% getViz()

check.gamViz(bb_abund_2010_2)
summary(bb_abund_2010_2)
print(plot(bb_abund_2010_2, allTerms = TRUE), pages = 1)

bb_abund_2012_2 <- gam(bb.abund ~ te(yday, log(fl.abund + 0.01), bs = c("tp", "tp"), k = c(10, 10)),
                     data = filter(bb_fl_abund, year == 2012),
                     method = "REML",
                     family = "gaussian") %>% getViz()

check.gamViz(bb_abund_2012_2)
summary(bb_abund_2012_2)
  plot <- print(plot(bb_abund_2012_2, allTerms = TRUE), pages = 1)

```

And dare we attempt to model the ratio of FL to BB abundance?
```{r}
fl_per_bb <- bb_fl_abund %>%
  ungroup() %>%
  mutate(fl.per.bb = fl.abund/bb.abund)

fl_per_bb_2010 <- bam(log(fl.abund/bb.abund + 0.01) ~ te(yday, elev.mean, 
                                             bs = c("tp", "tp"), 
                                             k = c(10, 10)),
                      data = filter(fl_per_bb, year == 2010),
                      discrete = FALSE,
                      method = "REML",
                      #family = "poisson",
                      select = TRUE) %>% getViz()

check.gamViz(fl_per_bb_2010)
summary(fl_per_bb_2010)
print(plot(fl_per_bb_2010, allTerms = TRUE), pages = 1)

fl_per_bb_2011 <- bam(log(fl.abund/bb.abund + 0.01) ~ te(yday, elev.mean, 
                                             bs = c("tp", "tp"), 
                                             k = c(10, 10)),
                      data = filter(fl_per_bb, year == 2011),
                      discrete = FALSE,
                      method = "REML",
                      #family = "poisson",
                      select = TRUE) %>% getViz()

check.gamViz(fl_per_bb_2011)
summary(fl_per_bb_2011)
print(plot(fl_per_bb_2011, allTerms = TRUE), pages = 1)

fl_per_bb_2012 <- bam(log(fl.abund/bb.abund + 0.01) ~ te(yday, elev.mean, 
                                             bs = c("tp", "tp"), 
                                             k = c(10, 10)),
                      data = filter(fl_per_bb, year == 2012),
                      discrete = FALSE,
                      method = "REML",
                      #family = "poisson",
                      select = TRUE) %>% getViz()

check.gamViz(fl_per_bb_2012)
summary(fl_per_bb_2012)
print(plot(fl_per_bb_2012, allTerms = TRUE), pages = 1)

```

What if we split up flora by...stuff?
```{r}
# fl_abund_family <- survey %>%
#   group_by(site, date, yday = dayofyear,  gdd.cum, year, elev.class2, 
#            elev.mean, break1.gdd, break2.gdd, gdd.total, plant.family) %>%
#   summarize(fl.abund = sum(flower.cover)) %>%
#   mutate(year2010 = ifelse(year == 2010, 1, 0),
#          year2011 = ifelse(year == 2011, 1, 0),
#          year2012 = ifelse(year == 2012, 1, 0))
# 
# fl_abund_fam_2010 <- bam(log(fl.abund + 0.01) ~ plant.family +
#                            te(yday, elev.mean, by = factor(plant.family), bs = c("tp", "tp"), k = c(10, 10)),
#                          data = filter(fl_abund_family, year == 2010),
#                          discrete = TRUE,
#                          method = "fREML") %>% getViz()
#                          
# check.gamViz(fl_abund_fam_2010)
# summary(fl_abund_fam_2010)
# print(plot(fl_abund_fam_2010, allTerms = TRUE), pages = 1)

### Prepare data
# fl_abund_ktype.ss <- survey %>%
#   group_by(site, date, yday = dayofyear,  gdd.cum, year, elev.mean, k.type.ss) %>%
#   summarize(fl.abund = sum(flower.cover)) %>%
#   ungroup() %>%
#   spread(k.type.ss, fl.abund) %>%
#   gather(k.type.ss, fl.abund, -c(site, date, yday, gdd.cum, year, elev.mean))
# %>%
#   gather()
#   mutate(year2010 = ifelse(year == 2010, 1, 0),
#          year2011 = ifelse(year == 2011, 1, 0),
#          year2012 = ifelse(year == 2012, 1, 0),
#          k.type.ss = factor(k.type.ss))
# 
# fl_abund_ktype.s <- survey %>%
#   group_by(site, date, yday = dayofyear,  gdd.cum, year, elev.class2, 
#            elev.mean, break1.gdd, break2.gdd, k.type.s) %>%
#   summarize(fl.abund = sum(flower.cover)) %>%
#   ungroup() %>%
#   mutate(year2010 = ifelse(year == 2010, 1, 0),
#          year2011 = ifelse(year == 2011, 1, 0),
#          year2012 = ifelse(year == 2012, 1, 0),
#          k.type.s = factor(k.type.s))

top_plants <- net %>%
  group_by(year, bb.sp, plant.sp) %>%
  summarize(visits = n()) %>%
  group_by(year, bb.sp) %>%
  top_n(3, visits) %>%
  arrange(bb.sp, year, -visits)

ggplot(top_plants, aes(year, visits, fill = plant.sp)) +
  geom_col(position = "fill") +
  facet_wrap(~bb.sp)

fl_abund_sp <- survey %>%
  group_by(site, date, yday = dayofyear,  gdd.cum, year, 
           elev.mean, plant.sp) %>%
  summarize(fl.abund = sum(flower.cover)) %>%
  ungroup() %>%
  spread(plant.sp, fl.abund) %>%
  gather(plant.sp, fl.abund, -c(site, date, yday, gdd.cum, year, elev.mean)) %>%
  replace_na(list(fl.abund = 0)) %>%
  semi_join(top_plants, by = c("year", "plant.sp")) %>%
  mutate(year2010 = ifelse(year == 2010, 1, 0),
         year2011 = ifelse(year == 2011, 1, 0),
         year2012 = ifelse(year == 2012, 1, 0),
         plant.sp = factor(plant.sp))

### As an example, let's see how some BB species are predicted by their #1 plant
Stachys_alopecuros_2012 <- fl_abund_sp %>%
  filter(plant.sp == "Stachys alopecuros", year == 2012) %>%
  filter(fl.abund > 0)

prat_2012 <- bb_abund_sp %>% 
  filter(bb.sp == "prat", year == 2012) %>%
  left_join(Stachys_alopecuros_2012)

ggplot(prat_2012, aes(log(fl.abund), bb.abund)) +
  geom_point()

Rhododendron_hirsutum_2011 <- fl_abund_sp %>%
  filter(plant.sp == "Rhododendron hirsutum", year == 2011) %>%
  filter(fl.abund > 0)

prat_2011 <- bb_abund_sp %>% 
  filter(bb.sp == "prat", year == 2011) %>%
  left_join(Rhododendron_hirsutum_2011)

ggplot(prat_2011, aes(log(fl.abund), bb.abund)) +
  geom_point()

Campanula_scheuchzeri_2012 <-  fl_abund_sp %>%
  filter(plant.sp == "Campanula scheuchzeri", year == 2012) %>%
  filter(fl.abund > 0)

soro_2012 <- bb_abund_sp %>% 
  filter(bb.sp == "soro", year == 2012) %>%
  left_join(Campanula_scheuchzeri_2012)

ggplot(soro_2012, aes(log(fl.abund), bb.abund)) +
  geom_point()

# ### Call ktype.ss models
# fl_abund_ktype.ss_2010 <- bam(log(fl.abund + 0.01) ~ k.type.ss +
#                            te(yday, elev.mean, by = k.type.ss, bs = c("tp", "tp"), k = c(10, 10), m = 1),
#                          data = filter(fl_abund_ktype.ss, year == 2010),
#                          discrete = TRUE,
#                          method = "fREML") %>% getViz()
#                          
# check.gamViz(fl_abund_ktype.ss_2010)
# summary(fl_abund_ktype.ss_2010)
# print(plot(fl_abund_ktype.ss_2010, allTerms = TRUE), pages = 1)
# 
# fl_abund_ktype.ss_2011 <- bam(log(fl.abund + 0.01) ~ k.type.ss +
#                            te(yday, elev.mean, by = k.type.ss, bs = c("tp", "tp"), k = c(10, 10), m = 1),
#                          data = filter(fl_abund_ktype.ss, year == 2011),
#                          discrete = TRUE,
#                          method = "fREML") %>% getViz()
#                          
# check.gamViz(fl_abund_ktype.ss_2011)
# summary(fl_abund_ktype.ss_2011)
# print(plot(fl_abund_ktype.ss_2011, allTerms = TRUE), pages = 1)
# 
# fl_abund_ktype.ss_2012 <- bam(log(fl.abund + 0.01) ~ k.type.ss +
#                            te(yday, elev.mean, by = k.type.ss, bs = c("tp", "tp"), k = c(15, 15)),
#                          data = filter(fl_abund_ktype.ss, year == 2012),
#                          discrete = TRUE,
#                          method = "fREML") %>% getViz()
#                          
# check.gamViz(fl_abund_ktype.ss_2012)
# summary(fl_abund_ktype.ss_2012)
# print(plot(fl_abund_ktype.ss_2012, allTerms = TRUE), pages = 1)
# 
# ### Call ktype.s models
# fl_abund_ktype.s_2010 <- bam(log(fl.abund + 0.01) ~ k.type.s +
#                            te(yday, elev.mean, by = k.type.s, bs = c("tp", "tp"), k = c(10, 10), m = 1),
#                          data = filter(fl_abund_ktype.s, year == 2010),
#                          discrete = TRUE,
#                          method = "fREML") %>% getViz()
#                          
# check.gamViz(fl_abund_ktype.s_2010)
# summary(fl_abund_ktype.s_2010)
# print(plot(fl_abund_ktype.s_2010, allTerms = TRUE), pages = 1)
# 
# fl_abund_ktype.s_2011 <- bam(log(fl.abund + 0.01) ~ k.type.s +
#                            te(yday, elev.mean, by = k.type.s, bs = c("tp", "tp"), k = c(10, 10), m = 1),
#                          data = filter(fl_abund_ktype.s, year == 2011),
#                          discrete = TRUE,
#                          method = "fREML") %>% getViz()
#                          
# check.gamViz(fl_abund_ktype.s_2011)
# summary(fl_abund_ktype.s_2011)
# print(plot(fl_abund_ktype.s_2011, allTerms = TRUE), pages = 1)
# 
# fl_abund_ktype.s_2012 <- bam(log(fl.abund + 0.01) ~ k.type.s +
#                            te(yday, elev.mean, by = k.type.s, bs = c("tp", "tp"), k = c(15, 15)),
#                          data = filter(fl_abund_ktype.s, year == 2012),
#                          discrete = TRUE,
#                          method = "fREML") %>% getViz()
#                          
# check.gamViz(fl_abund_ktype.s_2012)
# summary(fl_abund_ktype.s_2012)
# print(plot(fl_abund_ktype.s_2012, allTerms = TRUE), pages = 1)

### Call species models
### What are the most visited species?
fl_abund_sp_2010 <- bam(log(fl.abund + 0.01) ~ plant.sp +
                           te(yday, elev.mean, by = plant.sp, bs = c("tp", "tp"), k = c(10, 10), m = 1),
                         data = filter(fl_abund_sp, year == 2010),
                         discrete = TRUE,
                         method = "fREML") %>% getViz()

check.gamViz(fl_abund_sp_2010)
summary(fl_abund_sp_2010)
print(plot(fl_abund_sp_2010, allTerms = TRUE), pages = 1)

fl_abund_sp_2011 <- bam(log(fl.abund + 0.01) ~ plant.sp +
                           te(yday, elev.mean, by = plant.sp, bs = c("tp", "tp"), k = c(10, 10), m = 1),
                         data = filter(fl_abund_sp, year == 2011),
                         discrete = TRUE,
                         method = "fREML") %>% getViz()

check.gamViz(fl_abund_sp_2011)
summary(fl_abund_sp_2011)
print(plot(fl_abund_sp_2011, allTerms = TRUE), pages = 1)

fl_abund_sp_2012 <- bam(log(fl.abund + 0.01) ~ plant.sp +
                           te(yday, elev.mean, by = plant.sp, bs = c("tp", "tp"), k = c(10, 10), m = 1),
                         data = filter(fl_abund_sp, year == 2012),
                         discrete = TRUE,
                         method = "fREML") %>% getViz()

check.gamViz(fl_abund_sp_2012)
summary(fl_abund_sp_2012)
print(plot(fl_abund_sp_2012, allTerms = TRUE), pages = 1)
```

Alright -- how does the visitation rate of each floral species differ (or not) from its surveyed abundance?
```{r}
fl_abund_net_sp <- net %>%
  group_by(site, date, yday = dayofyear, gdd.cum, year, 
           elev.mean, plant.sp) %>%
  summarize(fl.abund = n()) %>%
  ungroup() %>%
  spread(plant.sp, fl.abund) %>%
  gather(plant.sp, fl.abund, -c(site, date, yday, gdd.cum, year, elev.mean)) %>%
  replace_na(list(fl.abund = 0)) %>%
  semi_join(top_plants, by = c("year", "plant.sp")) %>%
  mutate(year2010 = ifelse(year == 2010, 1, 0),
         year2011 = ifelse(year == 2011, 1, 0),
         year2012 = ifelse(year == 2012, 1, 0),
         plant.sp = factor(plant.sp))

fl_abund_net_sp_2012 <- bam(fl.abund ~ plant.sp +
                           te(yday, elev.mean, by = plant.sp, bs = c("tp", "tp"), k = c(10, 10), m = 1),
                         data = filter(fl_abund_net_sp, year == 2012 & plant.sp %in% c("Trifolium pratense", "Origanum vulgare", "Prunella vulgaris", "Stachys alopecuros")),
                         discrete = TRUE,
                         family = "poisson",
                         method = "fREML") %>% getViz()

check.gamViz(fl_abund_net_sp_2012)
summary(fl_abund_net_sp_2012)
print(plot(fl_abund_net_sp_2012, allTerms = TRUE), pages = 1)
```

Now let's get really granular
```{r}
sp_by_sp <- net %>%
  group_by(site, date, yday = dayofyear, gdd.cum, year, 
           elev.mean, bb.sp, plant.sp) %>%
  summarize(fl.abund = n()) %>%
  ungroup() %>%
  spread(plant.sp, fl.abund) %>%
  gather(plant.sp, fl.abund, -c(site, date, yday, gdd.cum, year, elev.mean, bb.sp)) %>%
  replace_na(list(fl.abund = 0)) %>%
  semi_join(top_plants, by = c("year", "plant.sp")) %>%
  mutate(year2010 = ifelse(year == 2010, 1, 0),
         year2011 = ifelse(year == 2011, 1, 0),
         year2012 = ifelse(year == 2012, 1, 0),
         plant.sp = factor(plant.sp))

Stachys_alopecuros_2012 <-  bam(fl.abund ~ bb.sp +
                           te(yday, elev.mean, by = factor(bb.sp), bs = c("tp", "tp"), k = c(10, 10), m = 1),
                         data = filter(sp_by_sp, 
                                       year == 2012 & 
                                         bb.sp %in% c("prat", "pasc", "bss", "soro", "mont", "wurf") &
                                         plant.sp == "Stachys alopecuros"),
                         discrete = TRUE,
                         family = "poisson",
                         method = "fREML") %>% getViz()

check.gamViz(Stachys_alopecuros_2012)
summary(Stachys_alopecuros_2012)
print(plot(Stachys_alopecuros_2012, allTerms = TRUE), pages = 1)

Trifolium_pratense_2011 <-  bam(fl.abund ~ bb.sp +
                           te(yday, elev.mean, by = factor(bb.sp), bs = c("tp", "tp"), k = c(10, 10), m = 1),
                         data = filter(sp_by_sp, 
                                       year == 2011 & 
                                         bb.sp %in% c("prat", "pasc", "bss", "soro", "wurf") &
                                         plant.sp == "Trifolium pratense"),
                         discrete = TRUE,
                         family = "poisson",
                         method = "fREML") %>% getViz()

check.gamViz(Trifolium_pratense_2011)
summary(Trifolium_pratense_2011)
print(plot(Trifolium_pratense_2011, allTerms = TRUE), pages = 1)
```

### GAM surface plots of abundance and diversity metrics
#### After puzzling over whether to model the three years together, using year as a "by" factor, or to set up an independent model for each year, I've settled (for now) on the latter. Since the yday range of each year is determined largely by when the snow melted, I do not want to extrapolate beyond my data.
##### Floral abundance
```{r eval=FALSE, include=FALSE}
### yday ranges
tapply(fl_abund$yday, fl_abund$year, summary)
tapply(fl_abund$gdd.cum, fl_abund$year, summary)

fl_abund_yday_2010 <- bam(log(fl.abund + 0.001) ~ te(yday, elev.mean, bs= c("tp", "tp"), k=c(12, 12), m=1),
                    data = filter(fl_abund, year == "2010"),
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(fl_abund_yday_2010)
summary(fl_abund_yday_2010)
print(plot(fl_abund_yday_2010, allTerms = TRUE), pages = 1)

fl_abund_yday_2011 <- bam(log(fl.abund + 0.001) ~ te(yday, elev.mean, bs= c("tp", "tp"), k=c(12, 12), m=1),
                    data = filter(fl_abund, year == "2011"),
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(fl_abund_yday_2011)
summary(fl_abund_yday_2011)
print(plot(fl_abund_yday_2011, allTerms = TRUE), pages = 1)

fl_abund_yday_2012 <- bam(log(fl.abund + 0.001) ~ te(yday, elev.mean, bs= c("tp", "tp"), k=c(12, 12), m=1),
                    data = filter(fl_abund, year == "2012"),
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(fl_abund_yday_2012)
summary(fl_abund_yday_2012)
print(plot(fl_abund_yday_2012, allTerms = TRUE), pages = 1)
plot(predict(fl_abund_yday_2012))

### HGAM as per Gavin's recommendation, but based on GDD instead of yday
fl_abund_gdd <- bam(log(fl.abund + 0.001) ~ year +
                      te(gdd.cum, elev.mean, k=c(10, 10)) +
                      te(gdd.cum, elev.mean, by = year2010, bs= c("tp", "tp"), k=c(12, 12), m=1) +
                      te(gdd.cum, elev.mean, by = year2011, bs= c("tp", "tp"), k=c(12, 12), m=1) +
                      te(gdd.cum, elev.mean, by = year2012, bs= c("tp", "tp"), k=c(12, 12), m=1),
                    data = fl_abund,
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

fl_abund_yday <- bam(log(fl.abund + 0.001) ~ year +
                      te(yday, elev.mean, by = year, bs= c("tp", "tp"), k=c(15, 15), m=1),
                    data = fl_abund,
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

fl_abund_yday <- bam(log(fl.abund + 0.001) ~ year +
                      te(yday, elev.mean, by = year2010, bs= c("tp", "tp"), k=c(15, 15), m=1) +
                       te(yday, elev.mean, by = year2011, bs= c("tp", "tp"), k=c(15, 15), m=1) +
                       te(yday, elev.mean, by = year2012, bs= c("tp", "tp"), k=c(15, 15), m=1),
                    data = fl_abund,
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(fl_abund_yday)
summary(fl_abund_yday)
print(plot(fl_abund_yday, allTerms = TRUE), pages = 1)
```

### Let's refine our plotting with tidymv
```{r, eval=FALSE, include=FALSE}
p2010 <- predict_gam(fl_abund_yday_2010)
p2011 <- predict_gam(fl_abund_yday_2011)
p2012 <- predict_gam(fl_abund_yday_2012)
p <- predict_gam(fl_abund_gdd)

ggplot(p2010, aes(yday, elev.mean, fit)) +
  geom_raster(aes(fill = exp(fit))) +
  #geom_contour(aes(z = fit), color = "black") +
  geom_point(data = filter(fl_abund, year == 2010), 
             aes(yday, elev.mean, size = fl.abund),
             color = "white", alpha = 0.25) +
  scale_fill_viridis_c()

ggplot(p2011, aes(yday, elev.mean, fit)) +
  geom_raster(aes(fill = exp(fit))) +
  #geom_contour(aes(z = fit), color = "black") +
  geom_point(data = filter(fl_abund, year == 2011), 
             aes(yday, elev.mean, size = fl.abund),
             color = "white", alpha = 0.25) +
  scale_fill_viridis_c()

ggplot(p2012, aes(yday, elev.mean, fit)) +
  geom_raster(aes(fill = exp(fit))) +
  #geom_contour(aes(z = fit), color = "black") +
  geom_point(data = filter(fl_abund, year == 2012), 
             aes(yday, elev.mean, size = fl.abund),
             color = "white", alpha = 0.25) +
  scale_fill_viridis_c()

ggplot(p, aes(yday, elev.mean, fit)) +
  geom_raster(aes(fill = exp(fit))) +
  geom_contour(aes(z = fit), color = "black") +
  geom_point(data = fl_abund,
  aes(yday, elev.mean, size = fl.abund),
  color = "white", alpha = 0.25) +
  scale_fill_viridis_c() +
  facet_wrap(~year)

```

##### Floral richness 
```{r eval=FALSE, include=FALSE}
fl_rich_yday_2010 <- bam(fl.rich ~ te(yday, elev.mean, bs= c("tp", "tp"), k=c(10, 10), m=1),
                    data = filter(fl_rich, year == "2010"),
                    discrete = TRUE,
                    family = "poisson",
                    method = "fREML") %>% getViz()

check.gamViz(fl_rich_yday_2010)
summary(fl_rich_yday_2010)
print(plot(fl_rich_yday_2010, allTerms = TRUE), pages = 1)

fl_rich_yday_2011 <- bam(fl.rich ~ te(yday, elev.mean, bs= c("tp", "tp"), k=c(10, 10), m=1),
                    data = filter(fl_rich, year == "2011"),
                    discrete = TRUE,
                    family = "poisson",
                    method = "fREML") %>% getViz()

check.gamViz(fl_rich_yday_2011)
summary(fl_rich_yday_2011)
print(plot(fl_rich_yday_2011, allTerms = TRUE), pages = 1)

fl_rich_yday_2012 <- bam(fl.rich ~ te(yday, elev.mean, bs= c("tp", "tp"), k=c(10, 10), m=1),
                    data = filter(fl_rich, year == "2012"),
                    discrete = TRUE,
                    family = "poisson",
                    method = "fREML") %>% getViz()

check.gamViz(fl_rich_yday_2012)
summary(fl_rich_yday_2012)
print(plot(fl_rich_yday_2012, allTerms = TRUE), pages = 1)
```

##### BB abundance
```{r eval=FALSE, include=FALSE}
bb_abund_yday_2012 <- bam(bb.abund ~ te(yday, elev.mean, bs= c("tp", "tp"), k=c(10, 10), m=1),
                    data = filter(bb_abund, year == "2012"),
                    family = "poisson",
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(bb_abund_yday_2012)
summary(bb_abund_yday_2012)
print(plot(bb_abund_yday_2012, allTerms = TRUE), pages = 1)

bb_abund_yday_2011 <- bam(bb.abund ~ te(yday, elev.mean, bs= c("tp", "tp"), k=c(10, 10), m=1),
                    data = filter(bb_abund, year == "2011"),
                    family = "poisson",
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(bb_abund_yday_2011)
summary(bb_abund_yday_2011)
print(plot(bb_abund_yday_2011, allTerms = TRUE), pages = 1)

bb_abund_yday_2010 <- bam(bb.abund ~ te(yday, elev.mean, bs= c("tp", "tp"), k=c(10, 10), m=1),
                    data = filter(bb_abund, year == "2010"),
                    family = "poisson",
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(bb_abund_yday_2010)
summary(bb_abund_yday_2010)
print(plot(bb_abund_yday_2010, allTerms = TRUE), pages = 1)
```

##### BB abundance with species as a "by" factor smooth
##### I am using family = "poisson"; this results in somewhat overdispersed models, but with a higher variance explained than when I use quasipoisson or negative binomial distributions, and the latter two distributions seem to introduce skew. I am more concerned with maximizing the explanatory power of my models than I am with having accurate p-values. We'll see what reviewers think, I guess. 
```{r eval=FALSE, include=FALSE}
### Species-level abundance
bb_abund_sp <- net %>%
  group_by(site, date, dayofyear, gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd, bb.sp) %>%
  summarize(bb.abund = n()) %>%
  rename(yday = dayofyear) %>%
  ungroup() %>%
  filter(bb.sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")) %>% # we only want the common species I think
  mutate(bb.sp = factor(bb.sp))

### 2010
bb_abund_sp_yday_2010 <- bam(bb.abund ~ te(yday, elev.mean, by= bb.sp, bs= c("tp", "tp"), k=c(12, 12), m=2),
                             data = filter(bb_abund_sp, year == "2010"),
                             family = "poisson",
                             discrete = TRUE,
                             method = "fREML") %>% getViz()

check.gamViz(bb_abund_sp_yday_2010)
summary(bb_abund_sp_yday_2010)

### 2011
bb_abund_sp_yday_2011 <- bam(bb.abund ~ te(yday, elev.mean, by= bb.sp, bs= c("tp", "tp"), k=c(12, 12), m=2),
                             data = filter(bb_abund_sp, year == "2011"),
                             family = "poisson",
                             discrete = TRUE,
                             method = "fREML") %>% getViz()

check.gamViz(bb_abund_sp_yday_2011)
summary(bb_abund_sp_yday_2011)

### 2012
bb_abund_sp_yday_2012 <- bam(bb.abund ~ te(yday, elev.mean, by= bb.sp, bs= c("tp", "tp"), k=c(15, 15), m=2),
                             data = filter(bb_abund_sp, year == "2012"),
                             family = "poisson",
                             discrete = TRUE,
                             method = "fREML") %>% getViz()

check.gamViz(bb_abund_sp_yday_2012)
summary(bb_abund_sp_yday_2012)

### Global HGAM
bb_abund_sp_yday <- bam(bb.abund ~ year +
                          te(elev.mean, yday, bs= c("tp", "tp"), k=c(10, 10), m=2) +
                          te(yday, elev.mean, by= bb.sp, bs= c("tp", "tp"), k=c(15, 15), m=1),
                             data = bb_abund_sp,
                             family = "nb",
                             discrete = TRUE,
                             method = "fREML") %>% getViz()

check.gamViz(bb_abund_sp_yday)
summary(bb_abund_sp_yday)

### Extract fits
bb2010 <- predict_gam(bb_abund_sp_yday_2010)
bb2011 <- predict_gam(bb_abund_sp_yday_2011)
bb2012 <- predict_gam(bb_abund_sp_yday_2012)
bb <- predict_gam(bb_abund_sp_yday)

### Plot
ggplot(bb2010, aes(yday, elev.mean, fit)) +
  geom_raster(aes(fill = fit)) +
  geom_contour(aes(z = fit), color = "black") +
  geom_point(data = filter(bb_abund_sp, year == 2010), 
             aes(yday, elev.mean, size = bb.abund),
             color = "white", alpha = 0.5) +
  scale_fill_viridis_c() +
  facet_wrap(~bb.sp)

ggplot(bb2011, aes(yday, elev.mean, fit)) +
  geom_raster(aes(fill = fit)) +
  geom_contour(aes(z = fit), color = "black") +
  geom_point(data = filter(bb_abund_sp, year == 2011), 
             aes(yday, elev.mean, size = bb.abund),
             color = "white", alpha = 0.5) +
  scale_fill_viridis_c() +
  facet_wrap(~bb.sp)

ggplot(bb2012, aes(yday, elev.mean, fit)) +
  geom_raster(aes(fill = fit)) +
  geom_contour(aes(z = fit), color = "black") +
  geom_point(data = filter(bb_abund_sp, year == 2012), 
             aes(yday, elev.mean, size = bb.abund),
             color = "white", alpha = 0.5) +
  scale_fill_viridis_c() +
  facet_wrap(~bb.sp)
```

##### FL abundance by k.type
##### Need to go back through to simplify, include only ktypes that are used heavily
```{r, eval=FALSE, include=FALSE}
k.type_key <- read_delim("../processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE)

fl_abund_ktype_yday_2010 <- bam(log(fl.abund + 0.001) ~ te(yday, elev.mean, 
                                                           by= k.type.ss, 
                                                           bs= c("tp", "tp"), 
                                                           k=c(12, 12), m=1),
                    data = filter(fl_abund_ktype, year == "2010"),
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(fl_abund_ktype_yday_2010)
summary(fl_abund_ktype_yday_2010)
print(plot(fl_abund_ktype_yday_2010, allTerms = TRUE), pages = 10)

fl_abund_ktype_yday_2011 <- bam(log(fl.abund + 0.001) ~ te(yday, elev.mean, 
                                                           by= k.type.ss, 
                                                           bs= c("tp", "tp"), 
                                                           k=c(12, 12), m=1),
                    data = filter(fl_abund_ktype, year == "2011"),
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(fl_abund_ktype_yday_2011)
summary(fl_abund_ktype_yday_2011)
print(plot(fl_abund_ktype_yday_2011, allTerms = TRUE), pages = 10)

fl_abund_ktype_yday_2012 <- bam(log(fl.abund + 0.001) ~ te(yday, elev.mean, 
                                                           by= k.type.ss, 
                                                           bs= c("tp", "tp"), 
                                                           k=c(12, 12), m=1),
                    data = filter(fl_abund_ktype, year == "2012"),
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(fl_abund_ktype_yday_2012)
summary(fl_abund_ktype_yday_2012)
print(plot(fl_abund_ktype_yday_2012, allTerms = TRUE), pages = 10)
```

##### FL abundance by genus
##### Need to go back through to simplify, include only ktypes that are used heavily
```{r, eval=FALSE, include=FALSE}
fl_abund_genus <- survey %>%
  group_by(site, date, yday = dayofyear, gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd, plant.genus) %>%
  summarize(fl.abund = sum(flower.cover)) %>%
  ungroup() %>%
  mutate(plant.genus = factor(plant.genus))

fl_abund_genus_yday_2010 <- bam(log(fl.abund + 0.001) ~ te(yday, elev.mean, 
                                                           by= plant.genus, 
                                                           bs= c("tp", "tp"), 
                                                           k=c(8, 8), m=1),
                    data = filter(fl_abund_genus, year == "2010"),
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(fl_abund_genus_yday_2010)
summary(fl_abund_genus_yday_2010)
print(plot(fl_abund_genus_yday_2010, allTerms = TRUE), pages = 10)

```

##### BB richness
```{r eval=FALSE, include=FALSE}
bb_rich_yday_2010 <- gamV(bb.rich ~ te(yday, elev.mean, bs= c("tp", "tp"), k=c(10, 10), m=1),
                    data = filter(bb_rich, year == "2010"),
                    family = "poisson",
                    method = "REML")

check.gamViz(bb_rich_yday_2010)
summary(bb_rich_yday_2010)
print(plot(bb_rich_yday_2010, allTerms = TRUE), pages = 1)

bb_rich_yday_2011 <- gamV(bb.rich ~ te(yday, elev.mean, bs= c("tp", "tp"), k=c(10, 10), m=1),
                    data = filter(bb_rich, year == "2011"),
                    family = "poisson",
                    method = "REML")

check.gamViz(bb_rich_yday_2011)
summary(bb_rich_yday_2011)
print(plot(bb_rich_yday_2011, allTerms = TRUE), pages = 1)

bb_rich_yday_2012 <- gamV(bb.rich ~ te(yday, elev.mean, bs= c("tp", "tp"), k=c(10, 10), m=1),
                    data = filter(bb_rich, year == "2012"),
                    family = "poisson",
                    method = "REML")

check.gamViz(bb_rich_yday_2012)
summary(bb_rich_yday_2012)
print(plot(bb_rich_yday_2012, allTerms = TRUE), pages = 1)

```

##### FL.per.BB
```{r, eval=FALSE, include=FALSE}
fl.per.bb_yday <- gamV(log(fl.per.bb) ~ year +
                    te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
                    data = abund_div,
                    method = "REML")

check.gamViz(fl.per.bb_yday)
summary(fl.per.bb_yday)
print(plot(fl.per.bb_yday, allTerms = TRUE), pages = 1)

fl.per.bb_gdd <- gamV(log(fl.per.bb) ~ year +
                    te(gdd.cum, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
                    data = abund_div,
                    method = "REML")

check.gamViz(fl.per.bb_gdd)
summary(fl.per.bb_gdd)
print(plot(fl.per.bb_gdd, allTerms = TRUE), pages = 1)
```

### 2020-07-02 analyses/community_basic_analysis.Rmd
broke down analyses into separate scripts; did some other things

--- 2020-06-12 analysis.Rmd
@@
 ---
-title: "analysis"
+title: "community_basic_analysis"
 output: html_document
@@
 knitr::opts_chunk$set(include = FALSE)
-library(bipartite)
-library(mgcv)
-library(mgcViz)
+# library(bipartite)
+# library(mgcv)
+# library(mgcViz)
 #library(lme4)
 #library(lmerTest)
-library(taxize)
 library(tidyverse)
@@
 
-### Define functions
-#### Colony phase breakpoint functions
-```{r include=FALSE}
-# Find breakpoints on GDD scale
-breakpoints_gdd <- function(x, year, t1, t2) {
-
-  data <- filter(x, year == !!year)
-
-  # Logistic regression for first break point (founding to buildup)
-  glm1.low <- glm(q.prop ~ gdd.cum,
-                  family = quasibinomial,
-                  data = filter(data, elev.class2 == "low"))
-
-  glm1.mid <- glm(q.prop ~ gdd.cum,
-                  family = quasibinomial,
-                  data = filter(data, elev.class2 == "mid"))
-
-  glm1.high <- glm(q.prop ~ gdd.cum,
-                   family = quasibinomial,
-                   data = filter(data, elev.class2 == "high"))
-
-  # Logistic regression for second break point (buildup to reproductive)
-  glm2.low <- glm(r.prop ~ gdd.cum,
-                  family = quasibinomial,
-                  data = filter(data, elev.class2 == "low"))
-
-  glm2.mid <- glm(r.prop ~ gdd.cum,
-                  family = quasibinomial,
-                  data = filter(data, elev.class2 == "mid"))
-
-  glm2.high <- glm(r.prop ~ gdd.cum,
-                   family = quasibinomial,
-                   data = filter(data, elev.class2 == "high"))
-
-  ### Extract gdd corresponding to desired proportion; based on https://stackoverflow.com/questions/32040504/regression-logistic-in-r-finding-x-value-predictor-for-a-particular-y-value
-
-  getX <- function(x, y) {
-    (log(y/(1-y)) - coef(x)[1])/coef(x)[2]
-    }
-
-  break1.low <- getX(glm1.low, t1) %>%
-    data.frame() %>%
-    select(gdd = 1) %>%
-    mutate(elev.class2 = "low",
-           breakID = "break1")
-
-  break1.mid <- getX(glm1.mid, t1) %>%
-    data.frame() %>%
-    select(gdd = 1) %>%
-    mutate(elev.class2 = "mid",
-           breakID = "break1")
-
-  break1.high <- getX(glm1.high, t1) %>%
-    data.frame() %>%
-    select(gdd = 1) %>%
-    mutate(elev.class2 = "high",
-           breakID = "break1")
-
-  break2.low <- getX(glm2.low, t2) %>%
-    data.frame() %>%
-    select(gdd = 1) %>%
-    mutate(elev.class2 = "low",
-           breakID = "break2")
-
-  break2.mid <- getX(glm2.mid, t2) %>%
-    data.frame() %>%
-    select(gdd = 1) %>%
-    mutate(elev.class2 = "mid",
-           breakID = "break2")
-
-  break2.high <- getX(glm2.high, t2) %>%
-    data.frame() %>%
-    select(gdd = 1) %>%
-    mutate(elev.class2 = "high",
-           breakID = "break2")
-
-  # Join into data frame, convert elev.class2 to ordered factor, spread
-  breaks <- full_join(break1.low, break1.mid) %>%
-    full_join(break1.high) %>%
-    full_join(break2.low) %>%
-    full_join(break2.mid) %>%
-    full_join(break2.high) %>%
-    mutate(elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) %>%
-    spread(breakID, gdd)
-
-  #
-  out <- data %>%
-    full_join(breaks) %>%
-    gather(caste, count, -c(site, date, year, q.prop, r.prop,
-                          gdd.cum, yday, elev.class, elev.class2, 
-                          management, transect, elev.mean, 
-                          lat, lon, break1, break2)) %>%
-  group_by(site, date) %>%
-  mutate(prop = count/sum(count))
-}
-
-# Find breakpoints on Julian scale
-breakpoints_yday <- function(x, year, t1, t2) {
-
-  data <- filter(x, year == !!year)
-
-  # Logistic regression for first break point (founding to buildup)
-  glm1.low <- glm(q.prop ~ yday,
-                  family = quasibinomial,
-                  data = filter(data, elev.class2 == "low"))
-
-  glm1.mid <- glm(q.prop ~ yday,
-                  family = quasibinomial,
-                  data = filter(data, elev.class2 == "mid"))
-
-  glm1.high <- glm(q.prop ~ yday,
-                   family = quasibinomial,
-                   data = filter(data, elev.class2 == "high"))
-
-  # Logistic regression for second break point (buildup to reproductive)
-  glm2.low <- glm(r.prop ~ yday,
-                  family = quasibinomial,
-                  data = filter(data, elev.class2 == "low"))
-
-  glm2.mid <- glm(r.prop ~ yday,
-                  family = quasibinomial,
-                  data = filter(data, elev.class2 == "mid"))
-
-  glm2.high <- glm(r.prop ~ yday,
-                   family = quasibinomial,
-                   data = filter(data, elev.class2 == "high"))
-
-  ### Extract gdd corresponding to desired proportion; based on https://stackoverflow.com/questions/32040504/regression-logistic-in-r-finding-x-value-predictor-for-a-particular-y-value
-
-  getX <- function(x, y) {
-    (log(y/(1-y)) - coef(x)[1])/coef(x)[2]
-    }
-
-  break1.low <- getX(glm1.low, t1) %>%
-    data.frame() %>%
-    select(day = 1) %>%
-    mutate(elev.class2 = "low",
-           breakID = "break1")
-
-  break1.mid <- getX(glm1.mid, t1) %>%
-    data.frame() %>%
-    select(day = 1) %>%
-    mutate(elev.class2 = "mid",
-           breakID = "break1")
-
-  break1.high <- getX(glm1.high, t1) %>%
-    data.frame() %>%
-    select(day = 1) %>%
-    mutate(elev.class2 = "high",
-           breakID = "break1")
-
-  break2.low <- getX(glm2.low, t2) %>%
-    data.frame() %>%
-    select(day = 1) %>%
-    mutate(elev.class2 = "low",
-           breakID = "break2")
-
-  break2.mid <- getX(glm2.mid, t2) %>%
-    data.frame() %>%
-    select(day = 1) %>%
-    mutate(elev.class2 = "mid",
-           breakID = "break2")
-
-  break2.high <- getX(glm2.high, t2) %>%
-    data.frame() %>%
-    select(day = 1) %>%
-    mutate(elev.class2 = "high",
-           breakID = "break2")
-
-  # Join into data frame, convert elev.class2 to ordered factor, spread
-  breaks <- full_join(break1.low, break1.mid) %>%
-    full_join(break1.high) %>%
-    full_join(break2.low) %>%
-    full_join(break2.mid) %>%
-    full_join(break2.high) %>%
-    mutate(elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) %>%
-    spread(breakID, day)
-
-  #
-  out <- data %>%
-    full_join(breaks) %>%
-    gather(caste, count, -c(site, date, year, q.prop, r.prop,
-                          gdd.cum, yday, elev.class, elev.class2, 
-                          management, transect, elev.mean, 
-                          lat, lon, break1, break2)) %>%
-  group_by(site, date) %>%
-  mutate(prop = count/sum(count))
-}
-
-# Find Juian breakpoints across all elevations
-breakpoints_yday_all <- function(x, year, t1, t2) {
-
-  data <- filter(x, year == !!year)
-
-  # Logistic regression for first break point (founding to buildup)
-  glm1 <- glm(q.prop ~ yday,
-                  family = quasibinomial,
-                  data = data)
-
-  # Logistic regression for second break point (buildup to reproductive)
-  glm2 <- glm(r.prop ~ yday,
-                  family = quasibinomial,
-                  data = data)
-
-  ### Extract day corresponding to desired proportion; based on https://stackoverflow.com/questions/32040504/regression-logistic-in-r-finding-x-value-predictor-for-a-particular-y-value
-
-  getX <- function(x, y) {
-    (log(y/(1-y)) - coef(x)[1])/coef(x)[2]
-    }
-
-  break1 <- getX(glm1, t1) %>%
-    data.frame() %>%
-    select(day = 1) %>%
-    mutate(elev.class2 = "low",
-           breakID = "break1")
-
-  break2 <- getX(glm2, t2) %>%
-    data.frame() %>%
-    select(day = 1) %>%
-    mutate(elev.class2 = "low",
-           breakID = "break2")
-
-  # Join into data frame, convert elev.class2 to ordered factor, spread
-  breaks <- full_join(break1, break2) %>%
-    spread(breakID, day)
-
-  #
-  out <- data %>%
-    full_join(breaks) %>%
-    gather(caste, count, -c(site, date, year, q.prop, r.prop,
-                          gdd.cum, yday, elev.class, elev.class2, 
-                          management, transect, elev.mean, 
-                          lat, lon, break1, break2)) %>%
-  group_by(site, date) %>%
-  mutate(prop = count/sum(count))
-}
-
-# Binomial ggplot smoother
-binomial_smooth <- function(...) {
-  geom_smooth(method = "glm", method.args = list(family = "quasibinomial"), se = FALSE, ...)
-}
-```
-
-#### Rarefied network analysis
-```{r, include=FALSE}
-# Rare_net function: rarefies networks by iteratively subsampling more frequently sampled networks to match the transect count of the least frequently sampled one
-# species level
-rare_net_sp_BB <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      select(site, date) %>%
-      unique() %>%
-      group_by(site) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      group_by(site) %>%
-      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
-      ungroup() %>%
-      filter(freq > 1) %>% # remove singletons
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(specieslevel, level = "higher", index = c("d", "species strength", "proportional generality")) %>%
-      map(rownames_to_column) %>%
-      map(as_tibble) %>%
-      bind_rows(.id = "site") %>% # Collapse list into big data frame
-      rename(sp = rowname) %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out %>%
-    gather(metric, value, -c(sp, site, iteration)) %>% # get mean value across iterations for each metric
-    group_by(sp, site, metric) %>%
-    summarize(value = mean(value)) %>%
-    left_join(site_data) %>%
-    left_join(bb_traits, by = c("sp" = "bb.sp"))
-}
-
-rare_net_sp_FL <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      select(site, date) %>%
-      unique() %>%
-      group_by(site) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      group_by(site) %>%
-      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
-      ungroup() %>%
-      filter(freq > 1) %>% # remove singletons
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(specieslevel, level = "lower", index = c("d", "species strength", "proportional generality")) %>%
-      map(rownames_to_column) %>%
-      map(as_tibble) %>%
-      bind_rows(.id = "site") %>% # Collapse list into big data frame
-      rename(sp = rowname) %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out %>%
-    gather(metric, value, -c(sp, site, iteration)) %>% # get mean value across iterations for each metric
-    group_by(sp, site, metric) %>%
-    summarize(value = mean(value)) %>%
-    left_join(site_data) %>%
-    left_join(fl_traits, by = c("sp" = "plant.sp"))
-}
-
-# network level
-rare_net_nt <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      select(site, date) %>%
-      unique() %>%
-      group_by(site) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      group_by(site) %>%
-      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
-      ungroup() %>%
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(networklevel, weighted = TRUE, index = c("H2", "weighted NODF", "niche overlap", "C score", "generality", "web asymmetry", "weighted connectance")) %>% # We probably care mainly about the BB level of the network
-      data.frame() %>%
-      t() %>%
-      data.frame() %>%
-      rownames_to_column() %>%
-      dplyr::select(site = rowname, everything()) %>%
-      as_tibble() %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out  %>%
-      gather(metric, value, -c(site, iteration)) %>% # get mean value across iterations for each metric
-      group_by(site, metric) %>%
-      summarize(value = mean(value)) %>%
-      left_join(site_data)
-}
-
-```
-
-#### Rarefied network analysis by colony phase
-```{r, include=FALSE}
-# network level version 2
-rare_net_nt2 <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  t_filter <- net %>%
-    select(site, date, phase, year) %>%
-    unique() %>%
-    group_by(site, phase, year) %>%
-    mutate(transects = n()) %>%
-    filter(transects >= n)
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      semi_join(t_filter) %>%
-      select(site, date, phase, year) %>%
-      unique() %>%
-      group_by(site, phase, year) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, phase, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      #filter(freq > 1) %>%
-      unite(webID, c(site, phase, year), sep = "_") %>%
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(networklevel, weighted = TRUE, level = "higher",
-          index = c("web asymmetry", "weighted NODF", "H2", "C score", "niche overlap", "generality")) %>% 
-      data.frame() %>%
-      t() %>%
-      data.frame() %>%
-      rownames_to_column() %>%
-      dplyr::select(id = rowname, everything()) %>%
-      as_tibble() %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out  %>%
-      gather(metric, value, -c(id, iteration)) %>% # get mean value across iterations for each metric
-      group_by(id, metric) %>%
-      summarize(value = mean(value)) %>%
-      separate(id, sep = "_", into = c("site", "phase", "year")) %>%
-      mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
-      left_join(site_data)
-}
-# species level version 2
-rare_net_sp2 <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  t_filter <- net %>%
-    select(site, date, phase, year) %>%
-    unique() %>%
-    group_by(site, phase, year) %>%
-    mutate(transects = n()) %>%
-    filter(transects >= n)
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      semi_join(t_filter) %>%
-      select(site, date, phase, year) %>%
-      unique() %>%
-      group_by(site, phase, year) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, phase, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      #filter(freq > 1) %>%
-      unite(webID, c(site, phase, year), sep = "_") %>%
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(specieslevel, level = "higher", index = c("d", "species strength")) %>%
-      map(rownames_to_column) %>%
-      map(as_tibble) %>%
-      bind_rows(.id = "id") %>% # Collapse list into big data frame
-      rename(sp = rowname) %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out %>%
-    gather(metric, value, -c(sp, id, iteration)) %>% # get mean value across iterations for each metric
-    group_by(sp, id, metric) %>%
-    summarize(value = mean(value)) %>%
-    separate(id, sep = "_", into = c("site", "phase", "year")) %>%
-    mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
-    left_join(site_data) %>%
-    left_join(bb_traits, by = c("sp" = "bb.sp"))
-}
-```
-
-### Rarefied network analysis by floral k.type instead of floral species
-```{r}
-# Rare_net function: rarefies networks by iteratively subsampling more frequently sampled networks to match the transect count of the least frequently sampled one
-# species level
-rare_net_sp_k.type.s <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      select(site, date) %>%
-      unique() %>%
-      group_by(site) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      na.omit() %>%
-      group_by(site) %>%
-      filter(length(unique(bb.sp)) > 1 & length(unique(k.type.s)) > 1) %>%
-      ungroup() %>%
-      filter(freq > 1) %>% # remove singletons
-      dplyr::select(higher = bb.sp, lower = k.type.s, webID = site, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(specieslevel, level = "higher", index = c("d", "species strength", "proportional generality")) %>%
-      map(rownames_to_column) %>%
-      map(as_tibble) %>%
-      bind_rows(.id = "site") %>% # Collapse list into big data frame
-      rename(sp = rowname) %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out %>%
-    gather(metric, value, -c(sp, site, iteration)) %>% # get mean value across iterations for each metric
-    group_by(sp, site, metric) %>%
-    summarize(value = mean(value)) %>%
-    left_join(site_data) %>%
-    left_join(bb_traits, by = c("sp" = "bb.sp"))
-}
-
-# network level
-rare_net_nt_k.type.s <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      select(site, date) %>%
-      unique() %>%
-      group_by(site) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      na.omit() %>%
-      group_by(site) %>%
-      filter(length(unique(bb.sp)) > 1 & length(unique(k.type.s)) > 1) %>%
-      ungroup() %>%
-      dplyr::select(higher = bb.sp, lower = k.type.s, webID = site, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(networklevel, weighted = TRUE, level = "higher", index = c("H2", "weighted NODF", "niche overlap", "C score", "generality", "web asymmetry", "weighted connectance")) %>% # We probably care mainly about the BB level of the network
-      data.frame() %>%
-      t() %>%
-      data.frame() %>%
-      rownames_to_column() %>%
-      dplyr::select(site = rowname, everything()) %>%
-      as_tibble() %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out  %>%
-      gather(metric, value, -c(site, iteration)) %>% # get mean value across iterations for each metric
-      group_by(site, metric) %>%
-      summarize(value = mean(value)) %>%
-      left_join(site_data)
-}
-```
-
 ### Load site data
 ```{r, include=FALSE}
-site_data <- read_csv("./processed_data/site_data.csv") %>%
+site_data <- read_csv("../processed_data/site_data.csv") %>%
   dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
@@
 ```{r, include=FALSE}
-gdd <- read_csv("./processed_data/climate.csv") %>%
+gdd <- read_csv("../processed_data/climate.csv") %>%
   mutate(year = factor(year)) %>%
@@
 
-### Colony phase breakpoint analysis
-```{r, include=FALSE}
-castes <- read_csv("./processed_data/network.csv") %>%
-  filter(bb.sp != "psit") %>% # We do not want to consider Psithyrus, since these parasitic species do not have a worker caste
-  group_by(site, date, year, caste) %>%
-  summarize(caste.count = n()) %>%
-  spread(caste, caste.count) %>%
-  replace(is.na(.), 0) %>%
-  mutate(queen = oldqueen + queen,
-         reproductive = male + youngqueen,
-         q.prop = queen/(queen+worker+reproductive),
-         r.prop = reproductive/(queen+worker+reproductive)) %>%
-  select(-oldqueen, -youngqueen, -male) %>%
-  left_join(gdd, by = c("site", "date")) %>%
-  left_join(site_data, by = c("site")) %>%
-  arrange(gdd.cum) %>%
-  mutate(yday = yday(date))
-
-breaks_2012_gdd <- breakpoints_gdd(castes, 2012, t1 = 0.5, t2 = 0.15)
-breaks_2011_gdd <- breakpoints_gdd(castes, 2011, t1 = 0.5, t2 = 0.15)
-breaks_2010_gdd <- breakpoints_gdd(castes, 2010, t1 = 0.5, t2 = 0.15)
-
-breaks_2012_yday <- breakpoints_yday(castes, 2012, t1 = 0.75, t2 = 0.15)
-breaks_2011_yday <- breakpoints_yday(castes, 2011, t1 = 0.75, t2 = 0.15)
-breaks_2010_yday <- breakpoints_yday(castes, 2010, t1 = 0.75, t2 = 0.15)
-
-breaks_gdd_summary <- full_join(breaks_2012_gdd, 
-                                breaks_2011_gdd) %>%
-  full_join(breaks_2010_gdd)
-
-write_csv(breaks_gdd_summary, "./output/breaks_gdd.csv")
-
-breaks_yday_summary <- full_join(breaks_2012_yday, 
-                                 breaks_2011_yday) %>%
-  full_join(breaks_2010_yday)
-
-write_csv(breaks_yday_summary, "./output/breaks_yday.csv")
-
-breaks_yday <- bind_rows(breaks_2012_yday, breaks_2011_yday, breaks_2010_yday) %>%
-  ungroup() %>%
-  select(site, year, break1.yday = break1, break2.yday = break2) %>%
-  unique()
-
-breaks_gdd <- bind_rows(breaks_2012_gdd, breaks_2011_gdd, breaks_2010_gdd) %>%
-  ungroup() %>%
-  select(site, year, break1.gdd = break1, break2.gdd = break2) %>%
-  unique()
-
-breaks <- full_join(breaks_yday, breaks_gdd, by = c("site", "year")) %>%
-  mutate(year = factor(year))
-```
-
-### Plot phase breakpoints
-```{r, include=FALSE}
-breaks_summary <- full_join(breaks_gdd_summary, breaks_yday_summary) %>%
-  select(site, date, year, elev.class2, yday, gdd.cum, break1, break2, caste, prop) %>%
-  gather(time.metric, value, -c(site, date, year, elev.class2, break1, break2, caste, prop))
-
-ggplot(breaks_yday_summary, aes(yday, prop, color = caste)) +
-  geom_point(alpha = 0.5) +
-  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
-  geom_vline(aes(xintercept = break1), linetype = "dashed") +
-  geom_vline(aes(xintercept = break2), linetype = "dashed") +
-  facet_grid(elev.class2 ~ year) +
-  xlab("Julian day") +
-  ylab("Proportional abundance")
-
-ggplot(breaks_gdd_summary, aes(gdd.cum, prop, color = caste)) +
-  geom_point(alpha = 0.5) +
-  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
-  geom_vline(aes(xintercept = break1), linetype = "dashed") +
-  geom_vline(aes(xintercept = break2), linetype = "dashed") +
-  facet_grid(elev.class2 ~ year) +
-  xlab("GDD") +
-  ylab("Proportional abundance")
-
-ggplot(breaks_summary, aes(value, prop, color = caste)) +
-  geom_point(alpha = 0.25) +
-  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
-  facet_grid(elev.class2~ time.metric, scales = "free") +
-  xlab("Julian day") +
-  ylab("Proportional abundance") 
-```
-
 ### Load network and survey data
 ```{r, include=FALSE}
-bb_traits <- read_csv("./processed_data/bb_traits.csv") %>%
+bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
   select(subgenus, bb.sp, pbl.w, pbl.w.class)
 
-fl_traits <- read.csv("./processed_data/floral_k_type_nom_fix.csv") %>%
+fl_traits <- read.csv("../processed_data/floral_k_type_nom_fix.csv") %>%
   select(plant.sp, k.type) %>%
@@
 
-fl_tax <- read_csv("./processed_data/floral_tax.csv")
-
-net <- read_csv("./processed_data/network.csv") %>%
+fl_tax <- read_csv("../processed_data/floral_tax.csv")
+
+breaks <- read_csv("./output/breaks.csv") %>%
+  mutate(year = factor(year))
+
+net <- read_csv("../processed_data/network.csv") %>%
   unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
@@
 
-survey <- read_csv("./processed_data/floral_survey.csv") %>%
+survey <- read_csv("../processed_data/floral_survey.csv") %>%
   mutate(year = factor(year)) %>% # convert year to factor
@@
 
-```
+ggplot(net, aes(site, elev.mean, 
+                color = elev.class2, 
+                shape = transect)) +
+  geom_point() +
+  geom_hline(yintercept = 1750, linetype = "dashed")
+```
+
 ### BB and floral turnover ~ elevation
@@
 bb_range <- net %>%
-  group_by(site, elev.mean, bb.sp) %>%
+  group_by(site, year, elev.mean, bb.sp) %>%
   summarize(abund = n()) %>%
-  group_by(bb.sp) %>%
+  group_by(year, bb.sp) %>%
   mutate(elev.floor = min(elev.mean),
@@
   ylab("Elevation") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  facet_wrap(~year)
 
@@
 ```{r}
+### Floral survey based
 survey_trait_elev <- survey %>%
@@
 
+### Network based
 net_trait_elev <- net %>%
@@
 
+### Plot
 ggplot(survey_trait_elev, aes(elev.mean, sqrt(abund))) +
@@
 
+### See which genera comprise each k.type
 net_trait_genus_elev <- net %>%
@@
 
+### Plot
 ggplot(filter(net_trait_genus_elev, k.type.s == 5.1), 
@@
   xlab("Elevation")
+
+ggplot(filter(survey_trait_genus_elev, k.type.s %in% c(6.1)), 
+       aes(elev.mean, sqrt(abund))) +
+  geom_point() +
+  geom_smooth() +
+  geom_vline(xintercept = 1750, color = "red") +
+  facet_wrap(~plant.genus, ncol = 5) +
+  ylab("sqrt(Abundance)") +
+  xlab("Elevation")
+
+ggplot(filter(net_trait_genus_elev, k.type.s %in% c(6.1)), 
+       aes(elev.mean, sqrt(abund))) +
+  geom_point() +
+  geom_smooth() +
+  geom_vline(xintercept = 1750, color = "red") +
+  facet_wrap(~plant.genus, ncol = 5) +
+  ylab("sqrt(Visitation frequency)") +
+  xlab("Elevation")
 ```
@@
 ```{r}
-
 survey_genus_elev <- survey %>%
@@
 
-### GAM surface plots of abundance and diversity metrics
-#### HGAM elevation x time x year (I would really like someone -- maybe my future self -- to check that I'm doing this right)
-##### Floral abundance
-```{r eval=FALSE, include=FALSE}
-fl_abund_yday <- gamV(fl.abund ~ year +
-                    te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = abund_div,
-                    family = "tw",
-                    method = "REML")
-
-fl_abund_yday2 <- gamV(log(fl.abund + 0.001) ~ year +
-                    te(yday, elev.mean, by=year, bs= c("cr", "cr"), k=c(10, 10), m=1),
-                    data = abund_div,
-                    #family = "scat",
-                    method = "REML")
-
-fl_abund_yday3 <- gamV(log(fl.abund + 0.001) ~ year +
-                    te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = abund_div,
-                    #family = "scat",
-                    method = "REML")
-
-AIC(fl_abund_yday, fl_abund_yday2, fl_abund_yday3)
-
-check.gamViz(fl_abund_yday)
-summary(fl_abund_yday)
-
-print(plot(fl_abund_yday, allTerms = TRUE), pages = 1)
-
-check.gamViz(fl_abund_yday2)
-summary(fl_abund_yday2)
-
-print(plot(fl_abund_yday2, allTerms = TRUE), pages = 1)
-
-check.gamViz(fl_abund_yday3)
-summary(fl_abund_yday3)
-
-print(plot(fl_abund_yday3, allTerms = TRUE), pages = 1)
-
-fl_abund_gdd <- gam(fl.abund ~ year +
-                    te(gdd.cum, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    family = "tw",
-                    method = "REML")
-
-fl_abund_gdd.vis <- getViz(fl_abund_gdd)
-check.gamViz(fl_abund_gdd.vis)
-summary(fl_abund_gdd.vis)
-
-plot(fl_abund_gdd.vis, allTerms = TRUE) +
-  ggtitle("FL abundance (gdd)") +
-  xlab("GDD") +
-  ylab("Elevation")
-```
-
-##### Floral richness -- since year does not seem to change the overall pattern (though perhaps the magnitude), I could use a simpler model, perhaps with year as a random intercept effect; some explanat
-```{r eval=FALSE, include=FALSE}
-### By year
-fl_rich_yday <- gam(fl.rich ~ year +
-                    te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-fl_rich_yday.vis <- getViz(fl_rich_yday)
-check.gamViz(fl_rich_yday.vis)
-summary(fl_rich_yday.vis)
-
-plot(fl_rich_yday.vis, allTerms = TRUE) +
-  ggtitle("FL richness (gdd)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### With year as parametric term
-fl_rich_yday2 <- gam(fl.rich ~ year +
-                    te(yday, elev.mean, bs= "tp", k= 10, m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-fl_rich_yday2.vis <- getViz(fl_rich_yday2)
-check.gamViz(fl_rich_yday2.vis)
-summary(fl_rich_yday2.vis)
-
-plot(sm(fl_rich_yday2.vis, 1)) + 
-  l_fitRaster() + 
-  l_fitContour() + 
-  l_points(size = 1) +
-  ggtitle("FL richness (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-plot(pterm(fl_rich_yday2.vis, 1))
-
-### By year
-fl_rich_gdd <- gam(fl.rich ~ year +
-                    te(gdd.cum, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-fl_rich_gdd.vis <- getViz(fl_rich_gdd)
-check.gamViz(fl_rich_gdd.vis)
-summary(fl_rich_gdd.vis)
-
-plot(fl_rich_gdd.vis, allTerms = TRUE) +
-  ggtitle("FL richness (gdd)") +
-  xlab("GDD") +
-  ylab("Elevation")
-
-### With year as parametric term
-fl_rich_gdd2 <- gam(fl.rich ~ year +
-                    te(gdd.cum, elev.mean, bs= "tp", k= 10, m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-fl_rich_gdd2.vis <- getViz(fl_rich_gdd2)
-check.gamViz(fl_rich_gdd2.vis)
-summary(fl_rich_gdd2.vis)
-
-plot(sm(fl_rich_gdd2.vis, 1)) + 
-  l_fitRaster() + 
-  l_fitContour() + 
-  l_points(size = 1) +
-  ggtitle("FL richness (gdd)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-plot(pterm(fl_rich_gdd2.vis, 1))
-```
-
-##### BB abundance
-```{r eval=FALSE, include=FALSE}
-bb_abund_yday <- gam(bb.abund ~ year +
-                    te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-bb_abund_yday.vis <- getViz(bb_abund_yday)
-check.gamViz(bb_abund_yday.vis)
-summary(bb_abund_yday.vis)
-
-plot(bb_abund_yday.vis, allTerms = TRUE) +
-  ggtitle("BB abundance (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-bb_abund_gdd <- gam(bb.abund ~ year +
-                    te(gdd.cum, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-bb_abund_gdd.vis <- getViz(bb_abund_gdd)
-check.gamViz(bb_abund_gdd.vis)
-summary(bb_abund_gdd.vis)
-
-plot(bb_abund_gdd.vis, allTerms = TRUE) +
-  ggtitle("BB abundance (gdd)") +
-  xlab("GDD") +
-  ylab("Elevation")
-```
-
-##### BB richness
-```{r eval=FALSE, include=FALSE}
-bb_rich_yday <- gam(bb.rich ~ year +
-                    te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-bb_rich_yday.vis <- getViz(bb_rich_yday)
-check.gamViz(bb_rich_yday.vis)
-summary(bb_rich_yday.vis)
-plot(sm(bb_rich_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("bb richness 2010 (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-plot(sm(bb_rich_yday.vis, 2)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("bb richness 2011 (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-plot(sm(bb_rich_yday.vis, 3)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("bb richness 2012 (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-bb_rich_gdd <- gam(bb.rich ~ year +
-                    te(gdd.cum, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-bb_rich_gdd.vis <- getViz(bb_rich_gdd)
-check.gamViz(bb_rich_gdd.vis)
-summary(bb_rich_gdd.vis)
-
-plot(bb_rich_gdd.vis, allTerms = TRUE) +
-  ggtitle("BB richness (gdd)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-```
-
-### Rarefied network analysis (not broken down by phase)
-#### Inspect transect counts to set subsampling parameter
-```{r, include = FALSE}
-### Determine the minimum transect count per site*year to set rarefaction sampling
-transects <- net %>%
-  select(site, year, date) %>%
-  unique()
-
-transects_per_siteyear <- transects %>%
-  group_by(site, year) %>%
-  summarize(transects = n())
-
-min_transects_per_year <- transects_per_siteyear %>%
-  group_by(year) %>%
-  summarize(min_transects = min(transects))
-
-### Create yearly subsets to feed to rare_net
-net_2012 <- filter(net, year == "2012")
-net_2011 <- filter(net, year == "2011")
-net_2010 <- filter(net, year == "2010")
-
-### Demonstrate that the distribution of interaction frequencies does not differ across elevation classes; in other words, no systematic bias is introduced to the network analysis due to uneven frequency of species with few observations
-net_freq <- net %>%
-  group_by(site, dayofyear, year, bb.sp, elev.mean, elev.class2) %>%
-  summarize(freq = n())
-
-ggplot(net_freq, aes(elev.class2, log(freq))) +
-  geom_boxplot() +
-  facet_wrap(~year)
-```
-
-#### Generate network metric tables
-```{r}
-### Species level
-#### BBs
-spmet_2012_BB <- rare_net_sp_BB(net_2012, 9, 100) %>%
-  mutate(year = factor(rep("2012", n())))
-spmet_2011_BB <- rare_net_sp_BB(net_2011, 4, 100) %>%
-  mutate(year = factor(rep("2011", n())))  
-spmet_2010_BB <- rare_net_sp_BB(net_2010, 3, 100) %>%
-  mutate(year = factor(rep("2010", n())))
-
-spmet_BB <- bind_rows(spmet_2012_BB, spmet_2011_BB, spmet_2010_BB)
-
-spmet_mean_BB <- spmet_BB %>%
-  group_by(site, year, elev.mean, metric) %>%
-  summarize(value = mean(value, na.rm = TRUE)) %>%
-  ungroup() %>%
-  mutate(year = factor(year))
-
-#### plants
-spmet_2012_FL <- rare_net_sp_FL(net_2012, 9, 100) %>%
-  mutate(year = factor(rep("2012", n())))
-spmet_2011_FL <- rare_net_sp_FL(net_2011, 4, 100) %>%
-  mutate(year = factor(rep("2011", n())))  
-spmet_2010_FL <- rare_net_sp_FL(net_2010, 3, 100) %>%
-  mutate(year = factor(rep("2010", n())))
-
-spmet_FL <- bind_rows(spmet_2012_FL, spmet_2011_FL, spmet_2010_FL)
-
-spmet_mean_FL <- spmet_FL %>%
-  group_by(site, year, elev.mean, metric) %>%
-  summarize(value = mean(value, na.rm = TRUE)) %>%
-  ungroup() %>%
-  mutate(year = factor(year))
-
-### Network- and group-level
-netmet_rare_2012 <- rare_net_nt(net_2012, n = 9, iter = 25) %>%
-  mutate(year = factor(rep("2012", n())))
-netmet_rare_2011 <- rare_net_nt(net_2011, n = 4, iter = 25)  %>%
-  mutate(year = factor(rep("2011", n())))
-netmet_rare_2010 <- rare_net_nt(net_2010, n = 3, iter = 25) %>%
-  mutate(year = factor(rep("2010", n())))
-
-netmet <- bind_rows(netmet_rare_2012, 
-                    netmet_rare_2011, 
-                    netmet_rare_2010) %>%
-  mutate(metric = factor(metric, levels = c("H2",
-                                            "weighted.NODF",
-                                            "niche.overlap.HL", 
-                                            "C.score.HL",
-                                            "generality.HL",
-                                            "niche.overlap.LL", 
-                                            "C.score.LL",
-                                            "generality.LL",
-                                            "web.asymmetry",
-                                            "weighted.connectance"))) %>%
-  ungroup() %>%
-  mutate(year = factor(year))
-```
-
-### Plot network analysis
-```{r}
-#### Per species values
-ggplot(filter(spmet_2012_BB,
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf") & metric == "d"), 
-       aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
-  facet_wrap(~sp, scales = "free")
-
-ggplot(filter(spmet_2011_BB,
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf") & metric == "d"), 
-       aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
-  facet_wrap(~sp, scales = "free")
-
-ggplot(filter(spmet_2010_BB,
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf") & metric == "d"), 
-       aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
-  facet_wrap(~sp, scales = "free")
-
-#### Community mean d
-##### BB
-ggplot(filter(spmet_mean_BB, metric == "d"), aes(elev.mean, value)) +
-  geom_point() +
-  #geom_smooth(method = "gam", formula = y ~ s(x, k = 10)) +
-  xlab("Elevation") +
-  ylab("Community mean d'") +
-  facet_wrap(~year) +
-  theme_light(16)
-
-gam_d1 <- gamV(value ~ s(year, bs = "re") + s(elev.mean, k = 5, bs = "tp", by = year, m = 1), 
-             data = filter(spmet_mean_BB, metric == "d"),
-             method = "REML")
-
-check.gamViz(gam_d1)
-summary(gam_d1)
-plot(gam_d1, allTerms = TRUE)
-
-summary(lm(value ~ elev.mean, data =  filter(spmet_mean_BB, metric == "d" & year == 2012)))
-plot(lm(value ~ elev.mean, data =  filter(spmet_mean_BB, metric == "d" & year == 2012)))
-
-##### FL
-ggplot(filter(spmet_mean_FL, metric == "d"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
-  xlab("Elevation") +
-  ylab("Community mean d'") +
-  facet_wrap(~year) +
-  theme_light(16)
-
-#### Network level
-
-##### niche overlap
-ggplot(filter(netmet, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x)) +
-  #geom_smooth() +
-  xlab("Elevation") +
-  ylab("Niche overlap") +
-  facet_wrap(~year) +
-  theme_light(16)
-
-gam_niche.overlap <- gamV(value ~ s(year, bs = "re") + s(elev.mean, by = year, k = 10, bs = "tp", m = 2), 
-                         data = filter(netmet, metric == "niche.overlap.HL"),
-                         method = "REML")
-
-check.gamViz(gam_niche.overlap)
-summary(gam_niche.overlap)
-
-plot(gam_niche.overlap, allTerms = TRUE)
-
-##### C.score
-ggplot(filter(netmet, metric == "C.score.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x)) +
-  xlab("Elevation") +
-  ylab("C score") +
-  facet_wrap(~year) +
-  theme_light(16)
-
-gam_cscore <- gamV(value ~ s(year, bs = "re") + s(elev.mean, by = factor(year), bs = "tp", k = 6, m = 1), 
-                  data = filter(netmet, metric == "C.score.HL"),
-                  #family = "scat",
-                  method = "REML")
-
-check.gamViz(gam_cscore)
-summary(gam_cscore)
-
-plot(gam_cscore, allTerms = TRUE)
-```
-
-### Rarefied network analysis by k.type instead of by plant sp
-```{r, include = FALSE}
-### Create yearly subsets to feed to rare_net
-net_2012 <- filter(net, year == "2012")
-net_2011 <- filter(net, year == "2011")
-net_2010 <- filter(net, year == "2010")
-
-### Species level
-spmet_2012_k <- rare_net_sp_k.type.s(net_2012, 9, 20) %>%
-  mutate(year = factor(rep("2012", n())))
-spmet_2011_k <- rare_net_sp_k.type.s(net_2011, 4, 20) %>%
-  mutate(year = factor(rep("2011", n())))  
-spmet_2010_k <- rare_net_sp_k.type.s(net_2010, 3, 20) %>%
-  mutate(year = factor(rep("2010", n())))
-
-spmet_k <- bind_rows(spmet_2012_k, spmet_2011_k, spmet_2010_k)
-
-spmet_mean_k <- spmet_k %>%
-  group_by(site, year, elev.mean, metric) %>%
-  summarize(value = mean(value))
-
-### Network- and group-level
-netmet_rare_2012_k <- rare_net_nt_k.type.s(net_2012, n = 9, iter = 20) %>%
-  mutate(year = factor(rep("2012", n())))
-netmet_rare_2011_k <- rare_net_nt_k.type.s(net_2011, n = 4, iter = 20)  %>%
-  mutate(year = factor(rep("2011", n())))
-netmet_rare_2010_k <- rare_net_nt_k.type.s(net_2010, n = 3, iter = 20) %>%
-  mutate(year = factor(rep("2010", n())))
-
-netmet_k <- bind_rows(netmet_rare_2012, 
-                    netmet_rare_2011, 
-                    netmet_rare_2010) %>%
-  mutate(metric = factor(metric, levels = c("H2",
-                                            "weighted.NODF",
-                                            "niche.overlap.HL", 
-                                            "C.score.HL",
-                                            "generality.HL",
-                                            "web.asymmetry",
-                                            "weighted.connectance")))
-
-### Plot
-#### Per species values
-ggplot(filter(spmet_2012,
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
-       aes(elev.mean, value)) +
-  geom_point() +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth() +
-  facet_grid(metric~sp, scales = "free")
-
-ggplot(filter(spmet_2011,
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
-       aes(elev.mean, value)) +
-  geom_point() +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth() +
-  facet_grid(metric~sp, scales = "free")
-
-ggplot(filter(spmet_2010,
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
-       aes(elev.mean, value)) +
-  geom_point() +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth() +
-  facet_grid(metric~sp, scales = "free")
-
-#### Community mean d
-ggplot(filter(spmet_mean_k, metric == "d"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
-  xlab("Elevation") +
-  ylab("Community mean d'") +
-  theme_light(16)
-
-ggplot(filter(spmet_mean_k, metric == "d"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
-  xlab("Elevation") +
-  ylab("Community mean d'") +
-  facet_wrap(~year) +
-  theme_light(16)
-
-gam_d <- gam(value ~ year + s(elev.mean, k = 10), 
-             data = filter(spmet, metric == "d"),
-             method = "REML")
-
-gam_d.vis <- getViz(gam_d)
-check.gamViz(gam_d.vis)
-summary(gam_d.vis)
-
-plot(gam_d.vis, allTerms = TRUE) +
-  xlab("Elevation")
-
-#### Network level
-ggplot(netmet, aes(elev.mean, value)) +
-  geom_point() +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth() +
-  facet_grid(metric~year, scales = "free")
-
-##### H2
-ggplot(filter(netmet, metric == "H2"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
-  xlab("Elevation") +
-  ylab("H2'") +
-  theme_light(16)
-
-gam_H2 <- gam(value ~ year + s(elev.mean), 
-              data = filter(netmet, metric == "H2"),
-              method = "REML")
-
-gam_H2.vis <- getViz(gam_H2)
-check.gamViz(gam_H2.vis)
-summary(gam_H2.vis)
-
-plot(gam_H2.vis, allTerms = TRUE) +
-  xlab("Elevation")
-
-##### niche overlap
-ggplot(filter(netmet, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
-  #geom_smooth() +
-  xlab("Elevation") +
-  ylab("Niche overlap") +
-  theme_light(16)
-
-ggplot(filter(netmet, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
-  #geom_smooth() +
-  xlab("Elevation") +
-  ylab("Niche overlap") +
-  facet_wrap(~year) +
-  theme_light(16)
-
-gam_niche.overlap <- gam(value ~ year + s(elev.mean), 
-                         data = filter(netmet, metric == "niche.overlap.HL"),
-                         method = "REML")
-
-gam_niche.overlap.vis <- getViz(gam_niche.overlap)
-check.gamViz(gam_niche.overlap.vis)
-summary(gam_niche.overlap.vis)
-
-plot(gam_niche.overlap.vis, allTerms = TRUE) +
-  xlab("Elevation")
-
-##### C.score
-ggplot(filter(netmet, metric == "C.score.HL"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
-  #geom_smooth() +
-  xlab("Elevation") +
-  ylab("C score") +
-  theme_light(16)
-
-ggplot(filter(netmet, metric == "C.score.HL"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
-  #geom_smooth() +
-  xlab("Elevation") +
-  ylab("C score") +
-  facet_wrap(~year) +
-  theme_light(16)
-
-gam_cscore <- gam(value ~ year + s(elev.mean), 
-                  data = filter(netmet, metric == "C.score.HL"),
-                  method = "REML")
-
-gam_cscore.vis <- getViz(gam_cscore)
-check.gamViz(gam_cscore.vis)
-summary(gam_cscore.vis)
-
-plot(gam_cscore.vis, allTerms = TRUE) +
-  xlab("Elevation")
-```
-
-### Do BB species differ in their k.type associations?
-```{r}
-web_by_site <- net %>%
-  group_by(site, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
-  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-  na.omit() %>%
-  group_by(site) %>%
-  filter(length(unique(bb.sp)) > 1 & length(unique(k.type.s)) > 1) %>%
-  ungroup() %>%
-  dplyr::select(higher = bb.sp, lower = k.type.s, webID = site, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs()
-
-visweb(web[[1]])
-plotweb(web[[1]])
-
-web <- net %>%
-  group_by(year, elev.class2, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
-  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-  na.omit() %>%
-  unite(webID, c(year, elev.class2), sep = "_") %>%
-  dplyr::select(higher = bb.sp, lower = k.type.s, webID, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs()
-
-# 2010 high
-visweb(web[[1]])
-plotweb(web[[1]])
-
-# 2010 low
-visweb(web[[2]])
-plotweb(web[[2]])
-
-# 2010 med
-visweb(web[[3]])
-plotweb(web[[3]])
-
-# 2011 high
-visweb(web[[4]])
-plotweb(web[[4]])
-
-# 2011 low
-visweb(web[[5]])
-plotweb(web[[5]])
-
-# 2011 med
-visweb(web[[6]])
-plotweb(web[[6]])
-
-# 2012 high
-visweb(web[[7]])
-plotweb(web[[7]])
-
-# 2012 low
-visweb(web[[8]])
-plotweb(web[[8]])
-
-# 2012 med
-visweb(web[[9]])
-plotweb(web[[9]])
-
-```
-
-### Does tongue length class predict k.type association? Remember that B. wurflenii is primarily a nectar robber.
-```{r}
-library(infer)
-
-kugler_desc <- read_delim("./kugler.tsv", delim = "\t") %>%
-  select(1,2)
-
-ggplot(bb_traits, aes(pbl.w)) +
-  geom_histogram()
-
-net_cont <- net %>%
-  group_by(year, elev.class2, pbl.w.class, k.type.ss) %>%
-  summarize(freq = n()) %>%
-  group_by(elev.class2, pbl.w.class) %>%
-  mutate(prop = freq/sum(freq))
-
-ggplot(net_cont, aes(pbl.w.class, freq, fill = k.type.ss)) +
-  geom_col() +
-  facet_grid(.~elev.class2)
-
-ggplot(net_cont, aes(pbl.w.class, prop, fill = k.type.ss)) +
-  geom_col() +
-  facet_grid(.~elev.class2)
-
-chisq_mod <- chisq.test(net$pbl.w.class, net$k.type.ss)
-chisq_mod
-
-  c <- net %>%
-  select(k.type.s, pbl.w.class) %>%
-  mutate(k.type.s = factor(k.type.s),
-         pbl.w.class = factor(pbl.w.class)) %>%
-  chisq_test(k.type.s ~ pbl.w.class)
-
-```
-
 ### Let's look at how abundance varies on a per-species basis
@@
 ```
-
-######## Old analyses that I probably don't care to use but don't want to delete yet #########
-
-### Per-transect network analysis -- is it worth it? I would have to filter out the transects for which the networks are too small.
-#### Network level
-```{r eval=FALSE, include=FALSE}
-per.transect_netmet <- net %>%
-  group_by(site, dayofyear, gdd.cum, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-  group_by(site, dayofyear, gdd.cum, year) %>%
-  mutate(size = sum(freq),
-         bb.rich = length(unique(bb.sp)),
-         fl.rich = length(unique(plant.sp)),
-         total.rich = bb.rich + fl.rich) %>%
-  filter(bb.rich > 3 & fl.rich > 3) %>%
-  unite(webID, c(site, dayofyear, gdd.cum, year, size), sep = "_") %>%
-  dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs() %>% # convert to to bipartite web
-  map(networklevel, weighted = TRUE, level = "higher", index = c("niche overlap", "C score", "H2")) %>% # We probably care mainly about the BB level of the network
-  data.frame() %>%
-  t() %>%
-  data.frame() %>%
-  rownames_to_column() %>%
-  dplyr::select(id = rowname, everything()) %>%
-  as_tibble() %>%
-  gather(metric, value, -id) %>%
-  separate(id, into = c("site", "yday", "gdd.cum", "year", "size"), sep = "_") %>%
-  mutate(size = as.numeric(size),
-         year = factor(year),
-         gdd.cum = as.numeric(gdd.cum),
-         yday = as.numeric(yday),
-         site = as.factor(site)) %>%
-  left_join(site_data) %>%
-  left_join(abund_div)
-
-### Can we explain network metrics with abundance and diversity data?
-# BB rich
-ggplot(per.transect_netmet, aes(bb.rich, value)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_grid(metric ~ year, scales = "free")
-# FL rich
-ggplot(per.transect_netmet, aes(fl.rich, value)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_grid(metric ~ year, scales = "free")
-# BB abund
-ggplot(per.transect_netmet, aes(log(bb.abund), value)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_grid(metric ~ year, scales = "free")
-# FL abund
-ggplot(per.transect_netmet, aes(log(fl.abund), value)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_grid(metric ~ year, scales = "free")
-# Per-capita FL abund
-ggplot(per.transect_netmet, aes(log(fl.per.bb), value)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_grid(metric ~ year, scales = "free")
-
-### Can we explain network metrics with elevation and/or time?
-ggplot(per.transect_netmet, aes(elev.mean, value, color = log(size))) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(metric ~ year, scales = "free")
-
-ggplot(per.transect_netmet, aes(log(size), value, color = log(size))) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(metric ~ year, scales = "free")
-
-ggplot(filter(per.transect_netmet, size > 1), 
-       aes(yday, value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(metric ~ year, scales = "free")
-
-ggplot(filter(per.transect_netmet, size > 10), 
-       aes(gdd.cum, value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(metric ~ year, scales = "free")
-
-ggplot(filter(per.transect_netmet, size > 10), 
-       aes(yday, elev.mean)) +
-  geom_point(aes(color = value)) +
-  geom_smooth() +
-  facet_grid(metric ~ year, scales = "free")
-```
-
-#### Species level
-```{r eval=FALSE, include=FALSE}
-per.transect_spmet <- net %>%
-  group_by(site, dayofyear, gdd.cum, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-  group_by(site, dayofyear, gdd.cum, year) %>%
-  mutate(size = sum(freq),
-         bb.rich = length(unique(bb.sp)),
-         fl.rich = length(unique(plant.sp)),
-         total.rich = bb.rich + fl.rich) %>%
-  filter(bb.rich > 5 & fl.rich > 5) %>%
-  unite(webID, c(site, dayofyear, gdd.cum, year, size), sep = "_") %>%
-  dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs() %>% # convert to to bipartite web
-  map(specieslevel, level = "higher", index = "d") %>% # We probably care mainly about the BB level of the network
-  map(rownames_to_column) %>%
-  map(as_tibble) %>%
-  bind_rows(.id = "id") %>% # Collapse list into big data frame
-  rename(sp = rowname) %>%
-  gather(metric, value, -c(sp, id)) %>% # get mean value across iterations for each metric
-  separate(id, sep = "_", into = c("site", "yday", "gdd.cum", "year", "size")) %>%
-  mutate(size = as.numeric(size),
-         year = factor(year),
-         gdd.cum = as.numeric(gdd.cum),
-         yday = as.numeric(yday)) %>%
-  left_join(site_data) %>%
-  left_join(bb_traits, by = c("sp" = "bb.sp")) %>%
-  left_join(abund_div)
-
-per.transect_spmet.mean <- per.transect_spmet %>%
-  group_by(site, yday, gdd.cum, year) %>%
-  summarize(value = mean(value)) %>%
-  left_join(site_data)
-
-ggplot(filter(per.transect_spmet), 
-       aes(log(fl.abund), value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(. ~ year, scales = "free")
-
-ggplot(filter(per.transect_spmet), 
-       aes(elev.mean, size)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(. ~ year, scales = "free")
-
-ggplot(filter(per.transect_spmet.mean), 
-       aes(yday, value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(. ~ year, scales = "free")
-
-ggplot(filter(per.transect_spmet), 
-       aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(. ~ year, scales = "free")
-
-ggplot(filter(per.transect_spmet, size >= 10), 
-       aes(gdd.cum, value, color = log(size))) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(metric ~ year, scales = "free")
-
-ggplot(filter(per.transect_spmet, size >= 10), 
-       aes(yday, value, color = log(size))) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(metric ~ year, scales = "free")
-```
-
-### Phase x elev network analysis -- this only makes sense in 2012, where we have sufficient sampling across the whole colony lifecycle
-```{r eval=FALSE, include=FALSE}
-### Prepare pxe abundance and diversity table
-pxe_abund_div <- abund_div %>%
-  select(-gdd.cum, -yday, -week, -month, -date, -elev.class2) %>%
-  gather(ad.metric, ad.value, -c(site, year, phase, elev.mean)) %>%
-  group_by(site, phase, year, ad.metric) %>%
-  summarize(ad.value = mean(ad.value, na.rm = TRUE))
-
-pxe_net <- rare_net_nt2(filter(net, year == 2012), 3, 20) %>%
-  full_join((pxe_abund_div))
-
-ggplot(filter(pxe_net, metric == "H2"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(year ~ phase, scales = "free") +
-  xlab("Elevation") +
-  ylab("H2")
-
-ggplot(filter(pxe_net, metric == "C.score.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(year ~ phase, scales = "free") +
-  xlab("Elevation") +
-  ylab("C.score")
-
-ggplot(filter(pxe_net, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(year ~ phase, scales = "free") +
-  xlab("Elevation") +
-  ylab("Niche overlap")
-
-pxe_sp <- rare_net_sp2(net, 4, 20)  %>%
-  full_join((pxe_abund_div))
-
-ggplot(filter(pxe_sp, 
-              metric == "d" &
-                sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit") &
-                ad.metric == "fl.rich"), 
-       aes(ad.value, value, color = year)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +
-  facet_grid(sp ~ phase, scales = "free") +
-  ylim(c(0,1)) +
-  xlab("Floral richness") +
-  ylab("d'")
-
-pxe_sp_mean <- pxe_sp %>%
-  filter(sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")) %>%
-  group_by(site, year, phase, elev.mean, metric) %>%
-  summarize(value = mean(value, na.rm = TRUE))
-
-ggplot(filter(pxe_sp_mean, metric == "d"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(year ~ phase, scales = "free") +
-  xlab("Elevation") +
-  ylab("Mean d'")
-
-ggplot(filter(pxe_sp_mean, metric == "species.strength"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(year ~ phase, scales = "free") +
-  xlab("Elevation") +
-  ylab("Mean species strength'")
-
-ggplot(filter(pxe_sp, 
-              metric == "species.strength",
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
-       aes(elev.mean, value, color = year)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(sp ~ phase, scales = "free") +
-  xlab("Elevation") +
-  ylab("Species strength")
-
-ggplot(filter(pxe_sp, 
-              metric == "d",
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
-       aes(elev.mean, value, color = year)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(sp ~ phase, scales = "free") +
-  ylim(c(0,1)) +
-  xlab("Elevation") +
-  ylab("d'")
-
-ggplot(filter(pxe_sp, 
-              metric == "d",
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
-       aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(sp ~ phase, scales = "free") +
-  ylim(c(0,1)) +
-  xlab("Elevation") +
-  ylab("d'")
-
-ggplot(filter(pxe_sp, 
-              metric == "d",
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
-       aes(elev.mean, value, color = phase)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(year ~ sp, scales = "free") +
-  theme_light() +
-  xlab("Elevation") +
-  ylab("d'")
-```
-
-```{r}
-
-# network level PxE
-rare_net_nt_pxe <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  t_filter <- net %>%
-    select(site, date, phase, year) %>%
-    unique() %>%
-    group_by(site, phase, year) %>%
-    mutate(transects = n()) %>%
-    filter(transects >= n)
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      semi_join(t_filter) %>%
-      select(site, date, phase, year) %>%
-      unique() %>%
-      group_by(site, phase, year) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, phase, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      group_by(site, phase, year) %>%
-      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
-      ungroup() %>%
-      unite(webID, c(site, phase, year), sep = "_") %>%
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(networklevel, weighted = TRUE, level = "higher",
-          index = c("web asymmetry", "weighted NODF", "H2", "C score", "niche overlap", "generality")) %>% 
-      data.frame() %>%
-      t() %>%
-      data.frame() %>%
-      rownames_to_column() %>%
-      dplyr::select(id = rowname, everything()) %>%
-      as_tibble() %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out  %>%
-      gather(metric, value, -c(id, iteration)) %>% # get mean value across iterations for each metric
-      group_by(id, metric) %>%
-      summarize(value = mean(value)) %>%
-      separate(id, sep = "_", into = c("site", "phase", "year")) %>%
-      mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
-      left_join(site_data)
-}
-
-# species level version 2
-rare_net_sp2 <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  t_filter <- net %>%
-    select(site, date, phase, year) %>%
-    unique() %>%
-    group_by(site, phase, year) %>%
-    mutate(transects = n()) %>%
-    filter(transects >= n)
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      semi_join(t_filter) %>%
-      select(site, date, phase, year) %>%
-      unique() %>%
-      group_by(site, phase, year) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, phase, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      group_by(site, phase, year) %>%
-      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
-      ungroup() %>%
-      unite(webID, c(site, phase, year), sep = "_") %>%
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(specieslevel, level = "higher", index = c("d", "species strength")) %>%
-      map(rownames_to_column) %>%
-      map(as_tibble) %>%
-      bind_rows(.id = "id") %>% # Collapse list into big data frame
-      rename(sp = rowname) %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out %>%
-    gather(metric, value, -c(sp, id, iteration)) %>% # get mean value across iterations for each metric
-    group_by(sp, id, metric) %>%
-    summarize(value = mean(value)) %>%
-    separate(id, sep = "_", into = c("site", "phase", "year")) %>%
-    mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
-    left_join(site_data) %>%
-    left_join(bb_traits, by = c("sp" = "bb.sp"))
-}
-
-```

### 2020-07-02 analyses/mapping.Rmd
broke down analyses into separate scripts; did some other things

---
title: "mapping"
output: html_document
---

```{r}
library(tmap)
library(sf)
library(tidyverse)
```

### Explore the spatial patterns of the study sites
#### Interactive map
```{r}
network <- read_csv("./processed_data/network.csv")

sites <- read_csv("./processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) %>% # turn elav.class into an ordered factor 
  semi_join(network, by = "site") %>%
  st_as_sf(coords = c("lon", "lat")) %>%
  st_set_crs(4326)

tmap_mode("view")
  tm_basemap("Esri.WorldTopoMap") +
    tm_shape(sites) +
    tm_symbols(col = "black") +
    #tm_text("site", auto.placement = TRUE, ymod = 2) +
    tm_scale_bar()

tmap_mode("view")
tm_basemap("Esri.WorldImagery") +
  tm_shape(sites) +
  tm_symbols(col = "yellow") +
  #tm_text("site", auto.placement = TRUE, ymod = 2, col = "yellow") +
  tm_scale_bar()
```

### 2020-07-02 analyses/network_analysis.Rmd
broke down analyses into separate scripts; did some other things

--- 2020-06-12 analysis.Rmd
@@
 ---
-title: "analysis"
+title: "network_analysis"
 output: html_document
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(include = FALSE)
 library(bipartite)
@@
 library(mgcViz)
-#library(lme4)
-#library(lmerTest)
-library(taxize)
 library(tidyverse)
 library(lubridate)
-library(GGally)
 ```
@@
 ### Define functions
-#### Colony phase breakpoint functions
-```{r include=FALSE}
-# Find breakpoints on GDD scale
-breakpoints_gdd <- function(x, year, t1, t2) {
-
-  data <- filter(x, year == !!year)
-
-  # Logistic regression for first break point (founding to buildup)
-  glm1.low <- glm(q.prop ~ gdd.cum,
-                  family = quasibinomial,
-                  data = filter(data, elev.class2 == "low"))
-
-  glm1.mid <- glm(q.prop ~ gdd.cum,
-                  family = quasibinomial,
-                  data = filter(data, elev.class2 == "mid"))
-
-  glm1.high <- glm(q.prop ~ gdd.cum,
-                   family = quasibinomial,
-                   data = filter(data, elev.class2 == "high"))
-
-  # Logistic regression for second break point (buildup to reproductive)
-  glm2.low <- glm(r.prop ~ gdd.cum,
-                  family = quasibinomial,
-                  data = filter(data, elev.class2 == "low"))
-
-  glm2.mid <- glm(r.prop ~ gdd.cum,
-                  family = quasibinomial,
-                  data = filter(data, elev.class2 == "mid"))
-
-  glm2.high <- glm(r.prop ~ gdd.cum,
-                   family = quasibinomial,
-                   data = filter(data, elev.class2 == "high"))
-
-  ### Extract gdd corresponding to desired proportion; based on https://stackoverflow.com/questions/32040504/regression-logistic-in-r-finding-x-value-predictor-for-a-particular-y-value
-
-  getX <- function(x, y) {
-    (log(y/(1-y)) - coef(x)[1])/coef(x)[2]
-    }
-
-  break1.low <- getX(glm1.low, t1) %>%
-    data.frame() %>%
-    select(gdd = 1) %>%
-    mutate(elev.class2 = "low",
-           breakID = "break1")
-
-  break1.mid <- getX(glm1.mid, t1) %>%
-    data.frame() %>%
-    select(gdd = 1) %>%
-    mutate(elev.class2 = "mid",
-           breakID = "break1")
-
-  break1.high <- getX(glm1.high, t1) %>%
-    data.frame() %>%
-    select(gdd = 1) %>%
-    mutate(elev.class2 = "high",
-           breakID = "break1")
-
-  break2.low <- getX(glm2.low, t2) %>%
-    data.frame() %>%
-    select(gdd = 1) %>%
-    mutate(elev.class2 = "low",
-           breakID = "break2")
-
-  break2.mid <- getX(glm2.mid, t2) %>%
-    data.frame() %>%
-    select(gdd = 1) %>%
-    mutate(elev.class2 = "mid",
-           breakID = "break2")
-
-  break2.high <- getX(glm2.high, t2) %>%
-    data.frame() %>%
-    select(gdd = 1) %>%
-    mutate(elev.class2 = "high",
-           breakID = "break2")
-
-  # Join into data frame, convert elev.class2 to ordered factor, spread
-  breaks <- full_join(break1.low, break1.mid) %>%
-    full_join(break1.high) %>%
-    full_join(break2.low) %>%
-    full_join(break2.mid) %>%
-    full_join(break2.high) %>%
-    mutate(elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) %>%
-    spread(breakID, gdd)
-
-  #
-  out <- data %>%
-    full_join(breaks) %>%
-    gather(caste, count, -c(site, date, year, q.prop, r.prop,
-                          gdd.cum, yday, elev.class, elev.class2, 
-                          management, transect, elev.mean, 
-                          lat, lon, break1, break2)) %>%
-  group_by(site, date) %>%
-  mutate(prop = count/sum(count))
-}
-
-# Find breakpoints on Julian scale
-breakpoints_yday <- function(x, year, t1, t2) {
-
-  data <- filter(x, year == !!year)
-
-  # Logistic regression for first break point (founding to buildup)
-  glm1.low <- glm(q.prop ~ yday,
-                  family = quasibinomial,
-                  data = filter(data, elev.class2 == "low"))
-
-  glm1.mid <- glm(q.prop ~ yday,
-                  family = quasibinomial,
-                  data = filter(data, elev.class2 == "mid"))
-
-  glm1.high <- glm(q.prop ~ yday,
-                   family = quasibinomial,
-                   data = filter(data, elev.class2 == "high"))
-
-  # Logistic regression for second break point (buildup to reproductive)
-  glm2.low <- glm(r.prop ~ yday,
-                  family = quasibinomial,
-                  data = filter(data, elev.class2 == "low"))
-
-  glm2.mid <- glm(r.prop ~ yday,
-                  family = quasibinomial,
-                  data = filter(data, elev.class2 == "mid"))
-
-  glm2.high <- glm(r.prop ~ yday,
-                   family = quasibinomial,
-                   data = filter(data, elev.class2 == "high"))
-
-  ### Extract gdd corresponding to desired proportion; based on https://stackoverflow.com/questions/32040504/regression-logistic-in-r-finding-x-value-predictor-for-a-particular-y-value
-
-  getX <- function(x, y) {
-    (log(y/(1-y)) - coef(x)[1])/coef(x)[2]
-    }
-
-  break1.low <- getX(glm1.low, t1) %>%
-    data.frame() %>%
-    select(day = 1) %>%
-    mutate(elev.class2 = "low",
-           breakID = "break1")
-
-  break1.mid <- getX(glm1.mid, t1) %>%
-    data.frame() %>%
-    select(day = 1) %>%
-    mutate(elev.class2 = "mid",
-           breakID = "break1")
-
-  break1.high <- getX(glm1.high, t1) %>%
-    data.frame() %>%
-    select(day = 1) %>%
-    mutate(elev.class2 = "high",
-           breakID = "break1")
-
-  break2.low <- getX(glm2.low, t2) %>%
-    data.frame() %>%
-    select(day = 1) %>%
-    mutate(elev.class2 = "low",
-           breakID = "break2")
-
-  break2.mid <- getX(glm2.mid, t2) %>%
-    data.frame() %>%
-    select(day = 1) %>%
-    mutate(elev.class2 = "mid",
-           breakID = "break2")
-
-  break2.high <- getX(glm2.high, t2) %>%
-    data.frame() %>%
-    select(day = 1) %>%
-    mutate(elev.class2 = "high",
-           breakID = "break2")
-
-  # Join into data frame, convert elev.class2 to ordered factor, spread
-  breaks <- full_join(break1.low, break1.mid) %>%
-    full_join(break1.high) %>%
-    full_join(break2.low) %>%
-    full_join(break2.mid) %>%
-    full_join(break2.high) %>%
-    mutate(elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) %>%
-    spread(breakID, day)
-
-  #
-  out <- data %>%
-    full_join(breaks) %>%
-    gather(caste, count, -c(site, date, year, q.prop, r.prop,
-                          gdd.cum, yday, elev.class, elev.class2, 
-                          management, transect, elev.mean, 
-                          lat, lon, break1, break2)) %>%
-  group_by(site, date) %>%
-  mutate(prop = count/sum(count))
-}
-
-# Find Juian breakpoints across all elevations
-breakpoints_yday_all <- function(x, year, t1, t2) {
-
-  data <- filter(x, year == !!year)
-
-  # Logistic regression for first break point (founding to buildup)
-  glm1 <- glm(q.prop ~ yday,
-                  family = quasibinomial,
-                  data = data)
-
-  # Logistic regression for second break point (buildup to reproductive)
-  glm2 <- glm(r.prop ~ yday,
-                  family = quasibinomial,
-                  data = data)
-
-  ### Extract day corresponding to desired proportion; based on https://stackoverflow.com/questions/32040504/regression-logistic-in-r-finding-x-value-predictor-for-a-particular-y-value
-
-  getX <- function(x, y) {
-    (log(y/(1-y)) - coef(x)[1])/coef(x)[2]
-    }
-
-  break1 <- getX(glm1, t1) %>%
-    data.frame() %>%
-    select(day = 1) %>%
-    mutate(elev.class2 = "low",
-           breakID = "break1")
-
-  break2 <- getX(glm2, t2) %>%
-    data.frame() %>%
-    select(day = 1) %>%
-    mutate(elev.class2 = "low",
-           breakID = "break2")
-
-  # Join into data frame, convert elev.class2 to ordered factor, spread
-  breaks <- full_join(break1, break2) %>%
-    spread(breakID, day)
-
-  #
-  out <- data %>%
-    full_join(breaks) %>%
-    gather(caste, count, -c(site, date, year, q.prop, r.prop,
-                          gdd.cum, yday, elev.class, elev.class2, 
-                          management, transect, elev.mean, 
-                          lat, lon, break1, break2)) %>%
-  group_by(site, date) %>%
-  mutate(prop = count/sum(count))
-}
-
-# Binomial ggplot smoother
-binomial_smooth <- function(...) {
-  geom_smooth(method = "glm", method.args = list(family = "quasibinomial"), se = FALSE, ...)
-}
-```
-
 #### Rarefied network analysis
@@
 
+  t_filter <- net %>%
+    dplyr::select(site, date, year) %>%
+    unique() %>%
+    group_by(site, year) %>%
+    mutate(transects = n()) %>%
+    filter(transects >= n)
+
   for(i in 1:iter) {
     temp <- net %>%
-      select(site, date) %>%
-      unique() %>%
-      group_by(site) %>%
+      semi_join(t_filter) %>%
+      dplyr::select(site, date, year) %>%
+      unique() %>%
+      group_by(site, year) %>%
       sample_n(n) %>%
       left_join(net) %>%
-      group_by(site, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
+      group_by(site, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
       summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      group_by(site) %>%
-      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
-      ungroup() %>%
-      filter(freq > 1) %>% # remove singletons
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
+      unite(webID, c(site, year), sep = "_") %>%
+      #group_by(site_year) %>%
+      #filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
+      #ungroup() %>%
+      #filter(freq > 1) %>% # remove singletons
+      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
       data.frame() %>% # convert to data frame
       frame2webs() %>% # convert to to bipartite web
-      map(specieslevel, level = "higher", index = c("d", "species strength", "proportional generality")) %>%
+      map(specieslevel, level = "higher", index = "d") %>%
       map(rownames_to_column) %>%
       map(as_tibble) %>%
-      bind_rows(.id = "site") %>% # Collapse list into big data frame
+      bind_rows(.id = "site_year") %>% # Collapse list into big data frame
       rename(sp = rowname) %>%
@@
   out <- out %>%
-    gather(metric, value, -c(sp, site, iteration)) %>% # get mean value across iterations for each metric
-    group_by(sp, site, metric) %>%
+    gather(metric, value, -c(sp, site_year, iteration)) %>% # get mean value across iterations for each metric
+    group_by(sp, site_year, metric) %>%
     summarize(value = mean(value)) %>%
+    separate(site_year, c("site", "year"), sep = "_") %>%
     left_join(site_data) %>%
@@
     temp <- net %>%
-      select(site, date) %>%
+      dplyr::select(site, date) %>%
       unique() %>%
@@
   t_filter <- net %>%
-    select(site, date, phase, year) %>%
+    dplyr::select(site, date, phase, year) %>%
     unique() %>%
@@
       semi_join(t_filter) %>%
-      select(site, date, phase, year) %>%
+      dplyr::select(site, date, phase, year) %>%
       unique() %>%
@@
 ```{r, include=FALSE}
-site_data <- read_csv("./processed_data/site_data.csv") %>%
+site_data <- read_csv("../processed_data/site_data.csv") %>%
   dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
@@
 ```{r, include=FALSE}
-gdd <- read_csv("./processed_data/climate.csv") %>%
+gdd <- read_csv("../processed_data/climate.csv") %>%
   mutate(year = factor(year)) %>%
-  select(site, date, gdd.cum)
-```
-
-### Colony phase breakpoint analysis
-```{r, include=FALSE}
-castes <- read_csv("./processed_data/network.csv") %>%
-  filter(bb.sp != "psit") %>% # We do not want to consider Psithyrus, since these parasitic species do not have a worker caste
-  group_by(site, date, year, caste) %>%
-  summarize(caste.count = n()) %>%
-  spread(caste, caste.count) %>%
-  replace(is.na(.), 0) %>%
-  mutate(queen = oldqueen + queen,
-         reproductive = male + youngqueen,
-         q.prop = queen/(queen+worker+reproductive),
-         r.prop = reproductive/(queen+worker+reproductive)) %>%
-  select(-oldqueen, -youngqueen, -male) %>%
-  left_join(gdd, by = c("site", "date")) %>%
-  left_join(site_data, by = c("site")) %>%
-  arrange(gdd.cum) %>%
-  mutate(yday = yday(date))
-
-breaks_2012_gdd <- breakpoints_gdd(castes, 2012, t1 = 0.5, t2 = 0.15)
-breaks_2011_gdd <- breakpoints_gdd(castes, 2011, t1 = 0.5, t2 = 0.15)
-breaks_2010_gdd <- breakpoints_gdd(castes, 2010, t1 = 0.5, t2 = 0.15)
-
-breaks_2012_yday <- breakpoints_yday(castes, 2012, t1 = 0.75, t2 = 0.15)
-breaks_2011_yday <- breakpoints_yday(castes, 2011, t1 = 0.75, t2 = 0.15)
-breaks_2010_yday <- breakpoints_yday(castes, 2010, t1 = 0.75, t2 = 0.15)
-
-breaks_gdd_summary <- full_join(breaks_2012_gdd, 
-                                breaks_2011_gdd) %>%
-  full_join(breaks_2010_gdd)
-
-write_csv(breaks_gdd_summary, "./output/breaks_gdd.csv")
-
-breaks_yday_summary <- full_join(breaks_2012_yday, 
-                                 breaks_2011_yday) %>%
-  full_join(breaks_2010_yday)
-
-write_csv(breaks_yday_summary, "./output/breaks_yday.csv")
-
-breaks_yday <- bind_rows(breaks_2012_yday, breaks_2011_yday, breaks_2010_yday) %>%
-  ungroup() %>%
-  select(site, year, break1.yday = break1, break2.yday = break2) %>%
-  unique()
-
-breaks_gdd <- bind_rows(breaks_2012_gdd, breaks_2011_gdd, breaks_2010_gdd) %>%
-  ungroup() %>%
-  select(site, year, break1.gdd = break1, break2.gdd = break2) %>%
-  unique()
-
-breaks <- full_join(breaks_yday, breaks_gdd, by = c("site", "year")) %>%
-  mutate(year = factor(year))
-```
-
-### Plot phase breakpoints
-```{r, include=FALSE}
-breaks_summary <- full_join(breaks_gdd_summary, breaks_yday_summary) %>%
-  select(site, date, year, elev.class2, yday, gdd.cum, break1, break2, caste, prop) %>%
-  gather(time.metric, value, -c(site, date, year, elev.class2, break1, break2, caste, prop))
-
-ggplot(breaks_yday_summary, aes(yday, prop, color = caste)) +
-  geom_point(alpha = 0.5) +
-  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
-  geom_vline(aes(xintercept = break1), linetype = "dashed") +
-  geom_vline(aes(xintercept = break2), linetype = "dashed") +
-  facet_grid(elev.class2 ~ year) +
-  xlab("Julian day") +
-  ylab("Proportional abundance")
-
-ggplot(breaks_gdd_summary, aes(gdd.cum, prop, color = caste)) +
-  geom_point(alpha = 0.5) +
-  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
-  geom_vline(aes(xintercept = break1), linetype = "dashed") +
-  geom_vline(aes(xintercept = break2), linetype = "dashed") +
-  facet_grid(elev.class2 ~ year) +
-  xlab("GDD") +
-  ylab("Proportional abundance")
-
-ggplot(breaks_summary, aes(value, prop, color = caste)) +
-  geom_point(alpha = 0.25) +
-  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
-  facet_grid(elev.class2~ time.metric, scales = "free") +
-  xlab("Julian day") +
-  ylab("Proportional abundance") 
+  dplyr::select(site, date, gdd.cum)
 ```
@@
 ```{r, include=FALSE}
-bb_traits <- read_csv("./processed_data/bb_traits.csv") %>%
-  select(subgenus, bb.sp, pbl.w, pbl.w.class)
-
-fl_traits <- read.csv("./processed_data/floral_k_type_nom_fix.csv") %>%
-  select(plant.sp, k.type) %>%
+bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
+  dplyr::select(subgenus, bb.sp, pbl.w, pbl.w.class)
+
+fl_traits <- read.csv("../processed_data/floral_k_type_nom_fix.csv") %>%
+  dplyr::select(plant.sp, k.type) %>%
   mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"),
@@
 
-survey <- read_csv("./processed_data/floral_survey.csv") %>%
+net_sum <- net %>%
+  group_by(site, year, elev.mean, date) %>%
+  summarize(total.obs = n()) %>%
+  ungroup() %>%
+  mutate(elev.fac = factor(elev.mean))
+
+ggplot(net_sum, aes(total.obs)) +
+  geom_histogram() +
+  facet_grid(elev.fac ~ year)
+
+ggplot(net_sum, aes(total.obs)) +
+  geom_histogram() +
+  facet_wrap(~year)
+
+survey <- read_csv("../processed_data/floral_survey.csv") %>%
   mutate(year = factor(year)) %>% # convert year to factor
@@
   arrange(site, date)
-
-```
-### BB and floral turnover ~ elevation
-```{r, include=FALSE}
-bb_range <- net %>%
-  group_by(site, elev.mean, bb.sp) %>%
-  summarize(abund = n()) %>%
-  group_by(bb.sp) %>%
-  mutate(elev.floor = min(elev.mean),
-         elev.ceiling = max(elev.mean),
-         elev.range = elev.ceiling - elev.floor,
-         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))
-
-fl_survey_elev <- survey %>%
-  filter(flower.cover > 0) %>%
-  group_by(site, plant.sp) %>%
-  summarize(abund = sum(flower.cover)) %>%
-  left_join(site_data) %>%
-  group_by(plant.sp) %>%
-  mutate(elev.floor = min(elev.mean),
-         elev.ceiling = max(elev.mean),
-         elev.range = elev.ceiling - elev.floor,
-         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2),
-         sites = length(unique(site)))
-
-fl_survey_elev_family <- survey %>%
-  filter(flower.cover > 0) %>%
-  group_by(site, plant.family) %>%
-  summarize(abund = sum(flower.cover)) %>%
-  left_join(site_data) %>%
-  group_by(plant.family) %>%
-  mutate(elev.floor = min(elev.mean),
-         elev.ceiling = max(elev.mean),
-         elev.range = elev.ceiling - elev.floor,
-         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2),
-         sites = length(unique(site)))
-
-fl_net_elev <- net %>%
-  group_by(site, elev.mean, plant.sp) %>%
-  summarize(abund = n()) %>%
-  group_by(plant.sp) %>%
-  mutate(elev.floor = min(elev.mean),
-         elev.ceiling = max(elev.mean),
-         elev.range = elev.ceiling - elev.floor,
-         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2),
-         sites = length(unique(site)))
-
-ggplot(filter(fl_survey_elev, sites > 0), 
-       aes(reorder(plant.sp, elev.ceiling), elev.mean)) +
-  geom_line(size = 1, color = "gray60") +
-  geom_point(alpha = 0.5) +
-  theme_light(14) +
-  theme(axis.text.x = element_blank()) +
-  xlab("Plant species") +
-  ylab("Elevation")
-
-ggplot(filter(fl_survey_elev_family, sites > 0), 
-       aes(reorder(plant.family, elev.med), elev.mean)) +
-  geom_line(size = 1, color = "gray60") +
-  geom_point(alpha = 0.5) +
-  theme_light(10) +
-  #theme(axis.text.x = element_blank()) +
-  xlab("Plant family") +
-  ylab("Elevation") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-ggplot(filter(fl_net_elev, sites > 0), 
-       aes(reorder(plant.sp, elev.med), elev.mean)) +
-  geom_line(size = 1, color = "gray60") +
-  geom_point(alpha = 0.5) +
-  theme_light(14) +
-  theme(axis.text.x = element_blank()) +
-  xlab("Plant species") +
-  ylab("Elevation")
-
-ggplot(bb_range, aes(reorder(bb.sp, elev.med), elev.mean)) +
-  geom_line(size = 3, color = "gray60") +
-  geom_point(aes(size = abund), alpha = 0.5) +
-  theme_light(14) +
-  xlab("BB species") +
-  ylab("Elevation") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-### Plot only the plants that drop out at 1750 m
-ggplot(filter(fl_net_elev, sites > 0 & elev.ceiling < 1750), 
-       aes(reorder(plant.sp, elev.med), elev.mean)) +
-  geom_line(size = 1, color = "gray60") +
-  geom_point(alpha = 0.5, aes(size = abund)) +
-  theme_light(8) +
-  xlab("Plant species") +
-  ylab("Elevation") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-ggplot(filter(fl_survey_elev, sites > 0 & elev.ceiling < 1750), 
-       aes(reorder(plant.sp, elev.med), elev.mean)) +
-  geom_line(size = 1, color = "gray60") +
-  geom_point(alpha = 0.5, aes(size = abund)) +
-  theme_light(8) +
-  xlab("Plant species") +
-  ylab("Elevation") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-```
-
-### FL traits by elevation -- do we see a sharp change at the tree line?
-```{r}
-survey_trait_elev <- survey %>%
-  group_by(site, year, elev.mean,  k.type.s) %>%
-  summarize(abund = sum(flower.cover))
-
-survey_trait_elev2 <- survey %>%
-  group_by(site, year, elev.mean,  k.type.ss) %>%
-  summarize(abund = sum(flower.cover))
-
-net_trait_elev <- net %>%
-  group_by(site, year, elev.mean,  k.type.s) %>%
-  summarize(abund = n())
-
-net_trait_elev2 <- net %>%
-  group_by(site, year, elev.mean,  k.type.ss) %>%
-  summarize(abund = n())
-
-ggplot(survey_trait_elev, aes(elev.mean, sqrt(abund))) +
-  geom_point() +
-  geom_smooth() +
-  geom_vline(xintercept = 1750, color = "red") +
-  facet_wrap(~k.type.s, scales = "free_y")
-
-ggplot(survey_trait_elev2, aes(elev.mean, sqrt(abund))) +
-  geom_point() +
-  geom_smooth() +
-  geom_vline(xintercept = 1750, color = "red") +
-  facet_wrap(~k.type.ss, scales = "free_y")
-
-ggplot(net_trait_elev, aes(elev.mean, sqrt(abund))) +
-  geom_point() +
-  geom_smooth() +
-  geom_vline(xintercept = 1750, color = "red") +
-  facet_wrap(~k.type.s, scales = "free_y")
-
-ggplot(net_trait_elev2, aes(elev.mean, sqrt(abund))) +
-  geom_point() +
-  geom_smooth() +
-  geom_vline(xintercept = 1750, color = "red") +
-  facet_wrap(~k.type.ss, scales = "free_y")
-
-net_trait_genus_elev <- net %>%
-  group_by(site, year, elev.mean, plant.genus, k.type.s) %>%
-  summarize(abund = n())
-
-survey_trait_genus_elev <- survey %>%
-  group_by(site, year, elev.mean, plant.genus, k.type.s) %>%
-  summarize(abund = sum(flower.cover))
-
-ggplot(filter(net_trait_genus_elev, k.type.s == 5.1), 
-       aes(elev.mean, sqrt(abund))) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~plant.genus)
-
-ggplot(filter(survey_trait_genus_elev, k.type.s %in% c(5.1)), 
-       aes(elev.mean, sqrt(abund))) +
-  geom_point() +
-  geom_smooth() +
-  geom_vline(xintercept = 1750, color = "red") +
-  facet_wrap(~plant.genus, ncol = 5) +
-  ylab("sqrt(Abundance)") +
-  xlab("Elevation")
-
-ggplot(filter(net_trait_genus_elev, k.type.s %in% c(5.1)), 
-       aes(elev.mean, sqrt(abund))) +
-  geom_point() +
-  geom_smooth() +
-  geom_vline(xintercept = 1750, color = "red") +
-  facet_wrap(~plant.genus, ncol = 5) +
-  ylab("sqrt(Visitation frequency)") +
-  xlab("Elevation")
-```
-
-### Genera and families by elevation
-```{r}
-
-survey_genus_elev <- survey %>%
-  group_by(site, year, elev.mean, plant.genus) %>%
-  summarize(abund = sum(flower.cover)) %>%
-  group_by(site, elev.mean, year) %>%
-  mutate(div = length(unique(plant.genus)))
-
-survey_family_elev <- survey %>%
-  group_by(site, year, elev.mean,  plant.family) %>%
-  summarize(abund = sum(flower.cover)) %>%
-  group_by(site, elev.mean, year) %>%
-  mutate(div = length(unique(plant.family)))
-
-net_genus_elev <- survey %>%
-  group_by(site, year, elev.mean,  plant.genus) %>%
-  summarize(abund = n()) %>%
-  group_by(site, elev.mean, year) %>%
-  mutate(div = length(unique(plant.genus)))
-
-net_family_elev <- survey %>%
-  group_by(site, year, elev.mean,  plant.family) %>%
-  summarize(abund = n()) %>%
-  group_by(site, elev.mean, year) %>%
-  mutate(div = length(unique(plant.family)))
-
-#### Plot
-ggplot(survey_genus_elev, aes(elev.mean, log(abund))) +
-  geom_point() +
-  stat_smooth() +
-  facet_wrap(~plant.genus, scales = "free_y")
-
-ggplot(survey_family_elev, aes(elev.mean, log(abund))) +
-  geom_point() +
-  stat_smooth() +
-  facet_wrap(~plant.family, scales = "free_y")
-
-ggplot(survey_genus_elev, aes(elev.mean, div)) +
-  geom_point() +
-  stat_smooth()
-
-ggplot(survey_family_elev, aes(elev.mean, div)) +
-  geom_point() +
-  stat_smooth()
-
-ggplot(net_genus_elev, aes(elev.mean, log(abund))) +
-  geom_point() +
-  stat_smooth() +
-  facet_wrap(~plant.genus, scales = "free_y")
-
-ggplot(net_family_elev, aes(elev.mean, log(abund))) +
-  geom_point() +
-  stat_smooth() +
-  facet_wrap(~plant.family, scales = "free_y")
-
-ggplot(net_genus_elev, aes(elev.mean, log(abund))) +
-  geom_point() +
-  stat_smooth(se = FALSE) +
-  facet_wrap(~plant.genus)
-
-ggplot(net_genus_elev, aes(elev.mean, log(abund))) +
-  geom_point() +
-  stat_smooth() +
-  facet_wrap(~plant.genus, scales = "free_y")
-ggplot(net_genus_elev, aes(elev.mean, abund)) +
-  geom_point() +
-  stat_smooth(se = FALSE) +
-  facet_wrap(~plant.genus)
-
-ggplot(net_genus_elev, aes(elev.mean, div)) +
-  geom_point() +
-  stat_smooth()
-```
-
-### BB abundance, diversity, and per-capita resource availability
-```{r eval=FALSE, echo=FALSE, include=FALSE}
-### Per-transect Bombus density
-bb_abund <- net %>%
-  group_by(site, date, dayofyear, gdd.cum, year, elev.class2, 
-           elev.mean, break1.gdd, break2.gdd) %>%
-  summarize(bb.abund = n())
-
-### Per-transect Bombus richness
-bb_rich <- net %>%
-  group_by(site, date, dayofyear, gdd.cum, year, elev.class2,
-           elev.mean, break1.gdd, break2.gdd) %>%
-  summarize(bb.rich = length(unique(bb.sp)))
-
-### Per-transect floral abund
-fl_abund <- survey %>%
-  group_by(site, date, dayofyear, gdd.cum, year, elev.class2, 
-           elev.mean, break1.gdd, break2.gdd) %>%
-  summarize(fl.abund = sum(flower.cover))
-
-### Per-transect floral richness
-fl_rich <- survey %>%
-  filter(flower.cover > 0) %>%
-  group_by(site, date, dayofyear, gdd.cum, year, elev.class2, 
-           elev.mean, break1.gdd, break2.gdd) %>%
-  summarize(fl.rich = n())
-
-### Join BB density and floral cover
-abund_div <- bb_abund %>%
-  full_join(bb_rich) %>%
-  full_join(fl_abund) %>% # Note that there is not perfect alignment in site-dates between the network observations and the floral surveying, so this join will cause some NAs that will need to be removed, causing some transects to drop out of the dataset
-  full_join(fl_rich) %>%
-  #na.omit() %>%
-  ungroup %>%
-  mutate(fl.per.bb = fl.abund/bb.abund,
-         site = factor(site),
-         yday = dayofyear,
-         week = week(date),
-         month = month(date)) %>%
-  ungroup() %>%
-  mutate(phase = case_when(
-    gdd.cum < break1.gdd ~ "founding",
-    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
-    gdd.cum > break2.gdd ~ "reproductive"
-  )) %>%
-  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive")),
-         total.rich = bb.rich + fl.rich,
-         web.asymmetry = (bb.rich - fl.rich) / (bb.rich + fl.rich))
-
-abund_div_mean <- abund_div %>%
-  select(-dayofyear, -week, -month, -break1.gdd, -break2.gdd) %>%
-  gather(metric, value, -c(site, date, gdd.cum, year, elev.class2, elev.mean, phase)) %>%
-  group_by(site, year, elev.class2, elev.mean, metric) %>%
-  summarize(value = mean(value, na.rm = TRUE))
-```
-
-### Plotting abundance and diversity patterns
-```{r}
-### BB abundance ~ floral abundance
-ggplot(abund_div, aes(log(fl.abund), bb.abund, color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "lm", se = TRUE) +
-  facet_grid(elev.class2 ~ phase, scales = "free") +
-  ylab("BB abundance") +
-  xlab("log(floral cover)") +
-  theme_bw()
-
-ggplot(abund_div, aes(log(fl.abund), bb.abund, color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  ylab("BB abundance") +
-  xlab("log(floral cover)") +
-  theme_bw()
-
-### BB richness ~ floral abundance
-ggplot(abund_div, aes(log(fl.abund), bb.rich, color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "lm", se = TRUE) +
-  facet_grid(elev.class2 ~ phase, scales = "free") +
-  ylab("BB richness") +
-  xlab("log(floral cover)") +
-  theme_bw()
-
-### BB abundance ~ floral richness
-ggplot(abund_div, aes(fl.rich, bb.abund, color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "lm", se = TRUE) +
-  facet_grid(elev.class2 ~ phase, scales = "free") +
-  ylab("BB abundance") +
-  xlab("Floral richness")
-
-### BB richness ~ floral richness
-ggplot(abund_div, aes(fl.rich, bb.rich, color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "lm", se = TRUE) +
-  facet_grid(elev.class2 ~ phase, scales = "free") +
-  ylab("BB richness") +
-  xlab("Floral richness")
-
-### Floral richness ~ floral abundance
-ggplot(abund_div, aes(log(fl.abund), fl.rich, color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "lm", se = TRUE) +
-  facet_grid(elev.class2 ~ phase, scales = "free") +
-  ylab("Floral richness") +
-  xlab("log(floral cover)")
-
-### Per-capita floral resource abundance ~ elevation
-ggplot(abund_div, aes(elev.mean, log(fl.per.bb))) +
-  geom_point(alpha = 0.5) +
-  #stat_summary(aes(y = log(fl.per.bb), group=1), fun.y=mean, colour="black", geom="point",group=1) +
-  geom_smooth() +
-  facet_grid(year~phase, scales = "free") +
-  ylab("log(fl-per-bb)") +
-  xlab("Elevation") +
-  theme_light()
-
-ggplot(abund_div, aes(elev.mean, log(fl.per.bb))) +
-  geom_point(alpha = 0.5) +
-  #stat_summary(aes(y = log(fl.per.bb), group=1), fun.y=mean, colour="black", geom="point",group=1) +
-  geom_smooth() +
-  facet_grid(.~year, scales = "free") +
-  ylab("log(fl-per-bb)") +
-  xlab("Elevation") +
-  theme_light()
-
-ggplot(abund_div, aes(elev.mean, log(fl.per.bb))) +
-  geom_point(alpha = 0.5, aes(color = phase)) +
-  #stat_summary(aes(y = log(fl.per.bb), group=1), fun.y=mean, colour="black", geom="point",group=1) +
-  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
-  facet_grid(year~., scales = "free") +
-  ylab("log(fl-per-bb)") +
-  xlab("Elevation") +
-  theme_light()
-
-ggplot(abund_div, aes(elev.mean, log(fl.per.bb))) +
-  geom_point(alpha = 0.5, aes(color = phase)) +
-  #stat_summary(aes(y = log(fl.per.bb), group=1), fun.y=mean, colour="black", geom="point",group=1) +
-  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
-  ylab("log(fl-per-bb)") +
-  xlab("Elevation") +
-  theme_light()
-
-ggplot(abund_div, aes(elev.mean, fl.per.bb)) +
-  geom_point(alpha = 0.5, aes(color = phase)) +
-  #stat_summary(aes(y = log(fl.per.bb), group=1), fun.y=mean, colour="black", geom="point",group=1) +
-  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
-  ylab("log(fl-per-bb)") +
-  xlab("Elevation") +
-  theme_light()
-
-ggplot(filter(abund_div_mean, metric == "fl.per.bb"), aes(elev.mean, log(value))) +
-  geom_point(alpha = 1) +
-  #geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) +
-  ylab("log(fl-per-bb)") +
-  xlab("Elevation") +
-  facet_wrap(~year) +
-  theme_light()
-
-lm1 <- lm(value ~ elev.mean * year, data = filter(abund_div_mean, metric == "fl.per.bb"))
-summary(lm1)
-
-ggplot(filter(abund_div_mean, metric == "fl.per.bb"), aes(elev.mean, value)) +
-  geom_point(alpha = 1) +
-  #geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
-  geom_smooth() +
-  ylab("fl-per-bb") +
-  xlab("Elevation") +
-  theme_light()
-
-ggplot(filter(abund_div_mean, metric == "fl.per.bb"), aes(elev.mean, log(value), color = year)) +
-  geom_point(alpha = 1) +
-  #geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
-  geom_smooth() +
-  ylab("FL per BB") +
-  xlab("Elevation") +
-  theme_light()
-
-ggplot(filter(abund_div_mean, metric == "fl.per.bb"), aes(elev.mean, log(value))) +
-  geom_point(alpha = 1, aes(color = year)) +
-  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
-  #geom_smooth() +
-  ylab("FL per BB") +
-  xlab("Elevation") +
-  theme_light(16)
-
-### Per-capita floral resource abundance ~ GDD
-ggplot(abund_div, aes(gdd.cum, log(fl.per.bb))) +
-  geom_point(alpha = 0.5, aes(color = phase)) +
-  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
-  facet_grid(elev.class2 ~ year, scales = "free") +
-  ylab("log(fl-per-bb)") +
-  xlab("GDD") +
-  theme_light()
-
-ggplot(abund_div, aes(gdd.cum, log(fl.per.bb))) +
-  geom_point(alpha = 0.5, aes(color = phase)) +
-  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
-  ylab("log(fl-per-bb)") +
-  xlab("GDD") +
-  theme_light()
-
-### Per-capita floral resource abundance ~ Julian day
-ggplot(abund_div, aes(yday, log(fl.per.bb))) +
-  geom_point(alpha = 0.5, aes(color = phase)) +
-  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
-  facet_grid(elev.class2 ~ year, scales = "free") +
-  ylab("log(fl-per-bb)") +
-  xlab("Julian day") +
-  theme_light()
-
-### BB richness ~ elevation
-ggplot(abund_div, aes(elev.mean, bb.rich)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(.~year)
-
-### BB richness ~ GDD
-ggplot(abund_div, aes(gdd.cum, bb.rich)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(elev.class2 ~ year)
-
-### BB richness ~ Julian day
-ggplot(abund_div, aes(yday, bb.rich)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(elev.class2 ~ year)
-
-### BB abundance ~ elevation
-ggplot(abund_div, aes(elev.mean, bb.abund)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(year ~ phase, scales = "free")
-
-ggplot(abund_div, aes(elev.mean, bb.abund)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(.~year, scales = "free")
-
-ggplot(abund_div, aes(elev.mean, bb.abund, color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  xlab("Elevation") +
-  ylab("BB abundance")  +
-  theme_light(16)
-
-ggplot(abund_div, aes(elev.mean, sqrt(bb.abund), color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  xlab("Elevation") +
-  ylab("sqrt(BB abundance)")  +
-  theme_light(16)
-
-### BB abundance ~ GDD
-ggplot(abund_div, aes(gdd.cum, bb.abund)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(elev.class2 ~ year)
-
-### BB abundance ~ Julian day
-ggplot(abund_div, aes(yday, bb.abund)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(elev.class2 ~ year)
-
-### Floral richness ~ elevation
-ggplot(abund_div, aes(elev.mean, fl.rich)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(year ~ phase)
-
-ggplot(abund_div, aes(elev.mean, fl.rich)) +
-  geom_point(alpha = 0.5, aes(color = year)) +
-  geom_smooth(color = "black", linetype = "dashed", size = 0.75) +
-  theme_light(16) +
-  xlab("Elevation") +
-  ylab("Floral richness")
-
-### Floral richness ~ GDD
-ggplot(abund_div, aes(gdd.cum, fl.rich)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(elev.class2 ~ year)
-
-### Floral richness ~ Julian day
-ggplot(abund_div, aes(yday, fl.rich)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(elev.class2 ~ year)
-
-### Floral abundance ~ elevation
-ggplot(abund_div, aes(elev.mean, log(fl.abund))) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  facet_grid(year ~ phase, scales = "free") +
-  theme_light()
-
-ggplot(abund_div, aes(elev.mean, log(fl.abund), color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  theme_light()
-
-ggplot(abund_div, aes(elev.mean, fl.abund)) +
-  geom_point(alpha = 0.5, aes(color = year)) +
-  geom_smooth(color = "black", linetype = "dashed", size = 0.75) +
-  theme_light(16) +
-  xlab("Elevation") +
-  ylab("Floral abundance")
-
-ggplot(abund_div, aes(elev.mean, log(fl.abund))) +
-  geom_point(alpha = 0.5, aes(color = year)) +
-  geom_smooth(color = "black", linetype = "dashed", size = 0.75) +
-  theme_light(16) +
-  xlab("Elevation") +
-  ylab("log(Floral abundance)")
-
-ggplot(abund_div, aes(elev.mean, fl.abund, color = year)) +
-  geom_point(alpha = 0.5) +
-  theme_light(16) +
-  xlab("Elevation") +
-  ylab("Floral abundance")
-
-ggplot(abund_div, aes(elev.mean, log(fl.abund), color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  theme_light(16) +
-  xlab("Elevation") +
-  ylab("log(Floral abundance)")
-
-ggplot(abund_div, aes(elev.mean, sqrt(bb.abund), color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  xlab("Elevation") +
-  ylab("sqrt(BB abundance)")  +
-  theme_light(16)
-
-### Floral abundance ~ GDD
-ggplot(abund_div, aes(gdd.cum, log(fl.abund))) +
-  geom_point(aes(color = phase), alpha =  0.5) +
-  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
-  facet_grid(elev.class2 ~ year, scales = "free") +
-  theme_light()
-
-### Floral abundance ~ Julian day
-ggplot(abund_div, aes(yday, log(fl.abund))) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(elev.class2 ~ year)
-
-### Total (BB = FL) richness ~ elevation
-ggplot(abund_div, aes(elev.mean, total.rich)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(year ~ phase, scales = "free")
-
-### Web asymmetry ~ elevation
-ggplot(abund_div, aes(elev.mean, web.asymmetry)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  facet_grid(year ~ phase, scales = "free")
-
-ggplot(filter(abund_div_mean, metric == "web.asymmetry"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(year ~ ., scales = "free")
-
-### Web asymmetry ~ gdd
-ggplot(abund_div, aes(gdd.cum, web.asymmetry)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  facet_grid(year ~.)
-
-### Web asymmetry ~ yday
-ggplot(abund_div, aes(yday, web.asymmetry)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  facet_grid(year ~.)
-
-### Web asymmetry ~ floral abundance
-ggplot(abund_div, aes(log(fl.abund), web.asymmetry)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  facet_grid(year ~ phase, scales = "free")
-
-### Web asymmetry ~ BB abundance
-ggplot(abund_div, aes(bb.abund, web.asymmetry)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  facet_grid(year ~ phase, scales = "free")
-
-### Web asymmetry ~ Per-capita floral resource abundance
-ggplot(abund_div, aes(log(fl.per.bb), web.asymmetry)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  facet_grid(year ~ phase, scales = "free")
-
-ggplot(abund_div, aes(log(fl.per.bb), web.asymmetry)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth() +
-  facet_grid(year ~ .)
-```
-
-### GAM surface plots of abundance and diversity metrics
-#### HGAM elevation x time x year (I would really like someone -- maybe my future self -- to check that I'm doing this right)
-##### Floral abundance
-```{r eval=FALSE, include=FALSE}
-fl_abund_yday <- gamV(fl.abund ~ year +
-                    te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = abund_div,
-                    family = "tw",
-                    method = "REML")
-
-fl_abund_yday2 <- gamV(log(fl.abund + 0.001) ~ year +
-                    te(yday, elev.mean, by=year, bs= c("cr", "cr"), k=c(10, 10), m=1),
-                    data = abund_div,
-                    #family = "scat",
-                    method = "REML")
-
-fl_abund_yday3 <- gamV(log(fl.abund + 0.001) ~ year +
-                    te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = abund_div,
-                    #family = "scat",
-                    method = "REML")
-
-AIC(fl_abund_yday, fl_abund_yday2, fl_abund_yday3)
-
-check.gamViz(fl_abund_yday)
-summary(fl_abund_yday)
-
-print(plot(fl_abund_yday, allTerms = TRUE), pages = 1)
-
-check.gamViz(fl_abund_yday2)
-summary(fl_abund_yday2)
-
-print(plot(fl_abund_yday2, allTerms = TRUE), pages = 1)
-
-check.gamViz(fl_abund_yday3)
-summary(fl_abund_yday3)
-
-print(plot(fl_abund_yday3, allTerms = TRUE), pages = 1)
-
-fl_abund_gdd <- gam(fl.abund ~ year +
-                    te(gdd.cum, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    family = "tw",
-                    method = "REML")
-
-fl_abund_gdd.vis <- getViz(fl_abund_gdd)
-check.gamViz(fl_abund_gdd.vis)
-summary(fl_abund_gdd.vis)
-
-plot(fl_abund_gdd.vis, allTerms = TRUE) +
-  ggtitle("FL abundance (gdd)") +
-  xlab("GDD") +
-  ylab("Elevation")
-```
-
-##### Floral richness -- since year does not seem to change the overall pattern (though perhaps the magnitude), I could use a simpler model, perhaps with year as a random intercept effect; some explanat
-```{r eval=FALSE, include=FALSE}
-### By year
-fl_rich_yday <- gam(fl.rich ~ year +
-                    te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-fl_rich_yday.vis <- getViz(fl_rich_yday)
-check.gamViz(fl_rich_yday.vis)
-summary(fl_rich_yday.vis)
-
-plot(fl_rich_yday.vis, allTerms = TRUE) +
-  ggtitle("FL richness (gdd)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-### With year as parametric term
-fl_rich_yday2 <- gam(fl.rich ~ year +
-                    te(yday, elev.mean, bs= "tp", k= 10, m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-fl_rich_yday2.vis <- getViz(fl_rich_yday2)
-check.gamViz(fl_rich_yday2.vis)
-summary(fl_rich_yday2.vis)
-
-plot(sm(fl_rich_yday2.vis, 1)) + 
-  l_fitRaster() + 
-  l_fitContour() + 
-  l_points(size = 1) +
-  ggtitle("FL richness (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-plot(pterm(fl_rich_yday2.vis, 1))
-
-### By year
-fl_rich_gdd <- gam(fl.rich ~ year +
-                    te(gdd.cum, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-fl_rich_gdd.vis <- getViz(fl_rich_gdd)
-check.gamViz(fl_rich_gdd.vis)
-summary(fl_rich_gdd.vis)
-
-plot(fl_rich_gdd.vis, allTerms = TRUE) +
-  ggtitle("FL richness (gdd)") +
-  xlab("GDD") +
-  ylab("Elevation")
-
-### With year as parametric term
-fl_rich_gdd2 <- gam(fl.rich ~ year +
-                    te(gdd.cum, elev.mean, bs= "tp", k= 10, m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-fl_rich_gdd2.vis <- getViz(fl_rich_gdd2)
-check.gamViz(fl_rich_gdd2.vis)
-summary(fl_rich_gdd2.vis)
-
-plot(sm(fl_rich_gdd2.vis, 1)) + 
-  l_fitRaster() + 
-  l_fitContour() + 
-  l_points(size = 1) +
-  ggtitle("FL richness (gdd)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-plot(pterm(fl_rich_gdd2.vis, 1))
-```
-
-##### BB abundance
-```{r eval=FALSE, include=FALSE}
-bb_abund_yday <- gam(bb.abund ~ year +
-                    te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-bb_abund_yday.vis <- getViz(bb_abund_yday)
-check.gamViz(bb_abund_yday.vis)
-summary(bb_abund_yday.vis)
-
-plot(bb_abund_yday.vis, allTerms = TRUE) +
-  ggtitle("BB abundance (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-bb_abund_gdd <- gam(bb.abund ~ year +
-                    te(gdd.cum, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-bb_abund_gdd.vis <- getViz(bb_abund_gdd)
-check.gamViz(bb_abund_gdd.vis)
-summary(bb_abund_gdd.vis)
-
-plot(bb_abund_gdd.vis, allTerms = TRUE) +
-  ggtitle("BB abundance (gdd)") +
-  xlab("GDD") +
-  ylab("Elevation")
-```
-
-##### BB richness
-```{r eval=FALSE, include=FALSE}
-bb_rich_yday <- gam(bb.rich ~ year +
-                    te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-bb_rich_yday.vis <- getViz(bb_rich_yday)
-check.gamViz(bb_rich_yday.vis)
-summary(bb_rich_yday.vis)
-plot(sm(bb_rich_yday.vis, 1)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("bb richness 2010 (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-plot(sm(bb_rich_yday.vis, 2)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("bb richness 2011 (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-plot(sm(bb_rich_yday.vis, 3)) +
-  l_fitRaster() +
-  l_fitContour() +
-  l_points() +
-  ggtitle("bb richness 2012 (yday)") +
-  xlab("Day of year") +
-  ylab("Elevation")
-
-bb_rich_gdd <- gam(bb.rich ~ year +
-                    te(gdd.cum, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
-                    data = filter(abund_div),
-                    method = "REML")
-
-bb_rich_gdd.vis <- getViz(bb_rich_gdd)
-check.gamViz(bb_rich_gdd.vis)
-summary(bb_rich_gdd.vis)
-
-plot(bb_rich_gdd.vis, allTerms = TRUE) +
-  ggtitle("BB richness (gdd)") +
-  xlab("Day of year") +
-  ylab("Elevation")
 ```
@@
 transects <- net %>%
-  select(site, year, date) %>%
+  dplyr::select(site, year, date) %>%
   unique()
@@
 net_2012 <- filter(net, year == "2012")
-net_2011 <- filter(net, year == "2011")
-net_2010 <- filter(net, year == "2010")
+# net_2011 <- filter(net, year == "2011")
+# net_2010 <- filter(net, year == "2010")
 
@@
 #### Generate network metric tables
-```{r}
+```{r, message=FALSE}
 ### Species level
 #### BBs
-spmet_2012_BB <- rare_net_sp_BB(net_2012, 9, 100) %>%
+spmet_2012_BB <- rare_net_sp_BB(net_2012, 9, 40) %>%
   mutate(year = factor(rep("2012", n())))
-spmet_2011_BB <- rare_net_sp_BB(net_2011, 4, 100) %>%
+spmet_2011_BB <- rare_net_sp_BB(net_2011, 4, 40) %>%
   mutate(year = factor(rep("2011", n())))  
-spmet_2010_BB <- rare_net_sp_BB(net_2010, 3, 100) %>%
+spmet_2010_BB <- rare_net_sp_BB(net_2010, 3, 40) %>%
   mutate(year = factor(rep("2010", n())))
 
-spmet_BB <- bind_rows(spmet_2012_BB, spmet_2011_BB, spmet_2010_BB)
+spmet_BB <- rare_net_sp_BB(net, 5, 50) %>%
+  ungroup() %>%
+  mutate(year = factor(year))
 
@@
 
+spmet_mean_BB_2012 <- spmet_2012_BB %>%
+  group_by(site, elev.mean, metric) %>%
+  summarize(value = mean(value, na.rm = TRUE))
+
+ggplot(spmet_mean_BB, aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~year)
+
+ggplot(spmet_2012_BB, aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~year)
+
+ggplot(spmet_mean_BB_2012, aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth()
+
+spmet_BB_gam <- bam(log(value + 0.01) ~ year + 
+                      s(elev.mean, bs = "tp", k = 10, m = 2) +
+                      s(elev.mean, by = factor(sp), k = 20, m = 1),
+                    data = spmet_BB,
+                    method = "fREML",
+                    discrete = TRUE) %>% getViz()
+
+check.gamViz(spmet_BB_gam)
+summary(spmet_BB_gam)
+print(plot(spmet_BB_gam, allTerms = TRUE), pages = 1)
+
+spmet_BB <- bind_rows(spmet_2012_BB, spmet_2011_BB, spmet_2010_BB)
+
 #### plants
-spmet_2012_FL <- rare_net_sp_FL(net_2012, 9, 100) %>%
+spmet_2012_FL <- rare_net_sp_FL(net_2012, 9, 40) %>%
   mutate(year = factor(rep("2012", n())))
-spmet_2011_FL <- rare_net_sp_FL(net_2011, 4, 100) %>%
+spmet_2011_FL <- rare_net_sp_FL(net_2011, 4, 40) %>%
   mutate(year = factor(rep("2011", n())))  
-spmet_2010_FL <- rare_net_sp_FL(net_2010, 3, 100) %>%
+spmet_2010_FL <- rare_net_sp_FL(net_2010, 3, 40) %>%
   mutate(year = factor(rep("2010", n())))
@@
 ### Network- and group-level
-netmet_rare_2012 <- rare_net_nt(net_2012, n = 9, iter = 25) %>%
+netmet_rare_2012 <- rare_net_nt(net_2012, n = 9, iter = 40) %>%
   mutate(year = factor(rep("2012", n())))
-netmet_rare_2011 <- rare_net_nt(net_2011, n = 4, iter = 25)  %>%
+netmet_rare_2011 <- rare_net_nt(net_2011, n = 4, iter = 40)  %>%
   mutate(year = factor(rep("2011", n())))
-netmet_rare_2010 <- rare_net_nt(net_2010, n = 3, iter = 25) %>%
+netmet_rare_2010 <- rare_net_nt(net_2010, n = 3, iter = 40) %>%
   mutate(year = factor(rep("2010", n())))
@@
                     netmet_rare_2010) %>%
+  ungroup() %>%
   mutate(metric = factor(metric, levels = c("H2",
@@
                                             "weighted.connectance"))) %>%
-  ungroup() %>%
   mutate(year = factor(year))
@@
   geom_point() +
-  #geom_smooth(method = "gam", formula = y ~ s(x, k = 10)) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 10)) +
   xlab("Elevation") +
@@
 
-gam_d1 <- gamV(value ~ s(year, bs = "re") + s(elev.mean, k = 5, bs = "tp", by = year, m = 1), 
+gam_d1 <- gamV(value ~ year + 
+                 s(elev.mean, k = 8, bs = "tp", by = year, m = 1), 
              data = filter(spmet_mean_BB, metric == "d"),
@@
 ### Rarefied network analysis by k.type instead of by plant sp
-```{r, include = FALSE}
+```{r, message = FALSE}
 ### Create yearly subsets to feed to rare_net
@@
 ### Species level
-spmet_2012_k <- rare_net_sp_k.type.s(net_2012, 9, 20) %>%
+spmet_2012_k <- rare_net_sp_k.type.s(net_2012, 9, 40) %>%
   mutate(year = factor(rep("2012", n())))
-spmet_2011_k <- rare_net_sp_k.type.s(net_2011, 4, 20) %>%
+spmet_2011_k <- rare_net_sp_k.type.s(net_2011, 4, 40) %>%
   mutate(year = factor(rep("2011", n())))  
-spmet_2010_k <- rare_net_sp_k.type.s(net_2010, 3, 20) %>%
+spmet_2010_k <- rare_net_sp_k.type.s(net_2010, 3, 40) %>%
   mutate(year = factor(rep("2010", n())))
@@
 ### Network- and group-level
-netmet_rare_2012_k <- rare_net_nt_k.type.s(net_2012, n = 9, iter = 20) %>%
+netmet_rare_2012_k <- rare_net_nt_k.type.s(net_2012, n = 9, iter = 40) %>%
   mutate(year = factor(rep("2012", n())))
-netmet_rare_2011_k <- rare_net_nt_k.type.s(net_2011, n = 4, iter = 20)  %>%
+netmet_rare_2011_k <- rare_net_nt_k.type.s(net_2011, n = 4, iter = 40)  %>%
   mutate(year = factor(rep("2011", n())))
-netmet_rare_2010_k <- rare_net_nt_k.type.s(net_2010, n = 3, iter = 20) %>%
+netmet_rare_2010_k <- rare_net_nt_k.type.s(net_2010, n = 3, iter = 40) %>%
   mutate(year = factor(rep("2010", n())))
@@
                                             "weighted.connectance")))
-
-### Plot
+```
+
+# Plot k.type network analysis
+```{r}
 #### Per species values
-ggplot(filter(spmet_2012,
+ggplot(filter(spmet_2012_k,
               sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
@@
 
-ggplot(filter(spmet_2011,
+ggplot(filter(spmet_2011_k,
               sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
@@
 
-ggplot(filter(spmet_2010,
+ggplot(filter(spmet_2010_k,
               sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
@@
 gam_d <- gam(value ~ year + s(elev.mean, k = 10), 
-             data = filter(spmet, metric == "d"),
+             data = filter(spmet_k, metric == "d"),
              method = "REML")
@@
 #### Network level
-ggplot(netmet, aes(elev.mean, value)) +
+ggplot(netmet_k, aes(elev.mean, value)) +
   geom_point() +
@@
 ##### H2
-ggplot(filter(netmet, metric == "H2"), aes(elev.mean, value)) +
+ggplot(filter(netmet_k, metric == "H2"), aes(elev.mean, value)) +
   geom_point(aes(color = year)) +
@@
 gam_H2 <- gam(value ~ year + s(elev.mean), 
-              data = filter(netmet, metric == "H2"),
+              data = filter(netmet_k, metric == "H2"),
               method = "REML")
@@
 ##### niche overlap
-ggplot(filter(netmet, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
+ggplot(filter(netmet_k, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
   geom_point(aes(color = year)) +
@@
 
-ggplot(filter(netmet, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
+ggplot(filter(netmet_k, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
   geom_point(aes(color = year)) +
@@
 gam_niche.overlap <- gam(value ~ year + s(elev.mean), 
-                         data = filter(netmet, metric == "niche.overlap.HL"),
+                         data = filter(netmet_k, metric == "niche.overlap.HL"),
                          method = "REML")
@@
 ##### C.score
-ggplot(filter(netmet, metric == "C.score.HL"), aes(elev.mean, value)) +
+ggplot(filter(netmet_k, metric == "C.score.HL"), aes(elev.mean, value)) +
   geom_point(aes(color = year)) +
@@
 
-ggplot(filter(netmet, metric == "C.score.HL"), aes(elev.mean, value)) +
+ggplot(filter(netmet_k, metric == "C.score.HL"), aes(elev.mean, value)) +
   geom_point(aes(color = year)) +
@@
 gam_cscore <- gam(value ~ year + s(elev.mean), 
-                  data = filter(netmet, metric == "C.score.HL"),
+                  data = filter(netmet_k, metric == "C.score.HL"),
                   method = "REML")
@@
 
-visweb(web[[1]])
-plotweb(web[[1]])
-
 web <- net %>%
@@
 ```{r}
-library(infer)
-
-kugler_desc <- read_delim("./kugler.tsv", delim = "\t") %>%
-  select(1,2)
+kugler_key <- read_delim("../processed_data/kugler_key.tsv", delim = "\t") %>%
+  dplyr::select(1,2)
 
@@
   summarize(freq = n()) %>%
-  group_by(elev.class2, pbl.w.class) %>%
+  group_by(year, elev.class2, pbl.w.class) %>%
+  mutate(prop = freq/sum(freq))
+
+net_cont2 <- net %>%
+  group_by(year, elev.class2, bb.sp, k.type.ss) %>%
+  summarize(freq = n()) %>%
+  group_by(year, elev.class2, bb.sp) %>%
   mutate(prop = freq/sum(freq))
@@
   geom_col() +
-  facet_grid(.~elev.class2)
+  facet_grid(elev.class2 ~ year)
 
@@
   geom_col() +
-  facet_grid(.~elev.class2)
+  facet_grid(elev.class2 ~ year)
+
+ggplot(net_cont2, aes(bb.sp, prop, fill = k.type.ss)) +
+  geom_col() +
+  facet_grid(elev.class2 ~ year)
 
@@
 
-  c <- net %>%
-  select(k.type.s, pbl.w.class) %>%
-  mutate(k.type.s = factor(k.type.s),
-         pbl.w.class = factor(pbl.w.class)) %>%
-  chisq_test(k.type.s ~ pbl.w.class)
-
-```
-
-### Let's look at how abundance varies on a per-species basis
-```{r eval=FALSE, include=FALSE}
-species <- net %>%
-  group_by(site, date, year, dayofyear, gdd.cum, phase, elev.mean, elev.class2, bb.sp, subgenus, pbl.w, pbl.w.class) %>%
-  summarize(abund = n()) %>%
-  group_by(site, date) %>%
-  mutate(prop = abund/sum(abund))
-
-subgenera <- net %>%
-  group_by(site, date, year, dayofyear, gdd.cum, phase, elev.mean, elev.class2, subgenus, pbl.w, pbl.w.class) %>%
-  summarize(abund = n()) %>%
-  group_by(site, date) %>%
-  mutate(prop = abund/sum(abund))
-
-ggplot(species, aes(dayofyear, abund)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth() +
-  facet_grid(year ~ bb.sp)
-
-ggplot(species, aes(dayofyear, prop)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth() +
-  facet_grid(year ~ bb.sp)
-
-ggplot(species, aes(elev.mean, abund)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth(se = FALSE, span = 1) +
-  facet_grid(year ~ bb.sp, scales = "free")
-
-ggplot(species, aes(elev.mean, prop)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth(se = FALSE, span = 1) +
-  facet_grid(year ~ bb.sp, scales = "free")
-
-ggplot(subgenera, aes(dayofyear, abund)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth() +
-  facet_grid(year ~ subgenus)
-
-ggplot(subgenera, aes(dayofyear, prop)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth() +
-  facet_grid(year ~ subgenus)
-
-ggplot(subgenera, aes(elev.mean, abund)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth() +
-  facet_grid(year ~ subgenus)
-
-ggplot(subgenera, aes(elev.mean, prop)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth(se = FALSE, span = 1) +
-  facet_grid(year ~ subgenus)
-
-### Pairs analysis
-species_pairs <- species %>%
-  select(site, year, date, bb.sp, abund) %>%
-  ungroup() %>%
-  spread(bb.sp, abund) %>%
-  replace(is.na(.), 0)
-
-ggplot(species_pairs, aes(pasc, prat)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
+ggplot(net, aes(log(pbl.w), corolla.depth)) +
+  geom_point() +
+  geom_smooth() +
   facet_wrap(~year)
 
-ggpairs(species_pairs, columns = c("bss", "camp", "flav", 
-                                   "gers", "hort", "humi",
-                                   "hypn", "jone", "lapi",
-                                   "mend", "mont", "muci",
-                                   "pasc", "prat", "psyt",
-                                   "pyre", "soro", "wurf"))
-
-ggpairs(species, columns = c("bb.sp", "abund"),
-        cardinality_threshold = 20)
-
+ggplot(net, aes(dayofyear, corolla.depth)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~year)
+
+ggplot(net, aes(elev.mean, corolla.depth)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~year)
+
+ggplot(net, aes(factor(pbl.w.class), corolla.depth)) +
+  geom_boxplot() 
 ```
@@
 pxe_abund_div <- abund_div %>%
-  select(-gdd.cum, -yday, -week, -month, -date, -elev.class2) %>%
+  dplyr::select(-gdd.cum, -yday, -week, -month, -date, -elev.class2) %>%
   gather(ad.metric, ad.value, -c(site, year, phase, elev.mean)) %>%

### 2020-07-02 analyses/parasites.Rmd
broke down analyses into separate scripts; did some other things

---
title: "parasites"
output: html_document
---

```{r setup, include=FALSE}
library(mgcv)
library(lubridate)
library(tidyverse)
```

```{r}
parasites <- read_csv("./processed_data/parasites.csv")
```

```{r}
ggplot(filter(parasites, caste == "worker"), aes(factor(year), size)) +
  geom_boxplot() +
  facet_wrap(~bb.sp, scales = "free")

TukeyHSD(aov(size ~ factor(year), data = filter(parasites, caste == "worker" & bb.sp == "prat")))
TukeyHSD(aov(size ~ factor(year), data = filter(parasites, caste == "worker" & bb.sp == "hort")))
TukeyHSD(aov(size ~ factor(year), data = filter(parasites, caste == "worker" & bb.sp == "mend")))
TukeyHSD(aov(size ~ factor(year), data = filter(parasites, caste == "worker" & bb.sp == "mont")))
TukeyHSD(aov(size ~ factor(year), data = filter(parasites, caste == "worker" & bb.sp == "muci")))
TukeyHSD(aov(size ~ factor(year), data = filter(parasites, caste == "worker" & bb.sp == "pasc")))
TukeyHSD(aov(size ~ factor(year), data = filter(parasites, caste == "worker" & bb.sp == "pyre")))
TukeyHSD(aov(size ~ factor(year), data = filter(parasites, caste == "worker" & bb.sp == "soro")))
TukeyHSD(aov(size ~ factor(year), data = filter(parasites, caste == "worker" & bb.sp == "telu")))
TukeyHSD(aov(size ~ factor(year), data = filter(parasites, caste == "worker" & bb.sp == "wurf")))

ggplot(filter(parasites, caste == "worker"), aes(elev.mean, size, color = factor(year))) +
  stat_summary(fun.y = mean, geom = "point") +
  #geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~bb.sp, scales = "free")

ggplot(parasites, aes(elev.mean, value)) +
  #stat_summary(fun.y = mean, geom = "point") +
  geom_point() +
  geom_smooth() +
  facet_wrap(~var, scales = "free")
```

```{r}

ggplot(filter(parasites, var %in% c("conopids", "ecto.mites", "tracheal.mites")), 
       aes(bin.value, fill = bin.value)) +
  geom_bar() +
  facet_grid(year~var, scales = "free")

ggplot(filter(parasites, var %in% c("conopids", "ecto.mites", "tracheal.mites")), 
       aes(elev.mean, value)) +
  geom_point() +
  facet_grid(year~var, scales = "free")

ggplot(filter(parasites, var %in% c("conopids", "ecto.mites", "tracheal.mites")), 
       aes(size, value)) +
  geom_point() +
  facet_grid(year~var, scales = "free")

ggplot(filter(parasites, var %in% c("conopids", "ecto.mites", "tracheal.mites")), 
       aes(yday, value)) +
  geom_point() +
  facet_grid(year~var, scales = "free")

ggplot(filter(parasites, var %in% c("conopids", "ecto.mites", "tracheal.mites")), 
       aes(bb.sp, value)) +
  geom_boxplot() +
  facet_grid(year~var, scales = "free")

ggplot(filter(parasites, var %in% c("conopids", "ecto.mites", "tracheal.mites")), 
       aes(plant.sp, value)) +
  geom_boxplot() +
  facet_grid(year~var, scales = "free")

```

### 2020-07-02 analyses/preference.Rmd
broke down analyses into separate scripts; did some other things

---
title: "preference"
output: html_document
---
### Prepare environment
```{r setup, include=FALSE, echo=FALSE}
library(mgcv)
library(mgcViz)
library(tidyverse)
library(lubridate)
```

### Climate data
```{r, include=FALSE, echo=FALSE}
gdd <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year)) %>%
  dplyr::select(site, year, date, gdd.cum) %>%
  semi_join(read_csv("../processed_data/network.csv"), by = "site")
```

### Load site data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))
```

### Load network and survey data
```{r, echo=FALSE, include=FALSE}
### BB traits
bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
  dplyr::select(subgenus, bb.sp, pbl.w, pbl.w.class)

### FL Kugler classification
fl_traits <- read.csv("../processed_data/floral_k_type_nom_fix.csv") %>%
  dplyr::select(plant.sp, k.type) %>%
  mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"),
         k.type.ss = str_extract(k.type, "[-+]?[0-9]*"))

### FL corolla depth (from Benadi)
fl_corolla_depth <- read.csv("../processed_data/flower_morphology_GitaBenadi.csv") %>%
  group_by(plant.sp = Species) %>%
  summarize(corolla.depth = mean(LN))

### FL higher  taxonomy
fl_tax <- read_csv("../processed_data/floral_tax.csv")

### Colony life cycle break points
breaks <- read_csv("./output/breaks.csv") %>%
  mutate(year = factor(year))

### BB-FL visitation data
net <- read_csv("../processed_data/network.csv") %>%
  mutate(year = factor(year)) %>% # convert year to factor
  full_join(bb_traits, by = c("bb.sp" = "bb.sp")) %>%
  left_join(site_data, by = c("site")) %>%
  left_join(gdd, by = c("site", "date", "year")) %>%
  left_join(breaks, by = c("site", "year")) %>%
  left_join(fl_traits) %>%
  left_join(fl_corolla_depth) %>%
  left_join(fl_tax, by = "plant.genus") %>%
  #left_join(temperature, by = c("site", "date")) %>%
  ungroup() %>%
  mutate(phase = case_when(
    gdd.cum < break1.gdd ~ "founding",
    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
    gdd.cum > break2.gdd ~ "reproductive"
  )) %>%
  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
  mutate(caste = case_when(
    caste == "worker" ~ "worker",
    caste == "male" ~ "reproductive",
    caste == "youngqueen" ~ "reproductive",
    caste == "queen" ~ "foundress",
    caste == "oldqueen" ~ "foundress"
  )) %>%
  rename(yday = dayofyear) %>%
  arrange(site, date)

### Floral survey data
survey <- read_csv("../processed_data/floral_survey.csv") %>%
  mutate(year = factor(year)) %>% # convert year to factor
  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime, not specifically for a given site*date)
  left_join(gdd, by = c("site", "date", "year")) %>%
  left_join(breaks, by = c("site", "year")) %>%
  left_join(site_data, by = c("site")) %>%
  left_join(fl_traits) %>%
  left_join(fl_traits) %>%
  left_join(fl_tax) %>%
  #left_join(temperature, by = c("site", "date")) %>%
  ungroup() %>%
  mutate(phase = case_when(
    gdd.cum < break1.gdd ~ "founding",
    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
    gdd.cum > break2.gdd ~ "reproductive"
  )) %>%
  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
  rename(yday = dayofyear) %>%
  arrange(site, date)
```

### FL and BB abundance / diversity
```{r echo=FALSE, include=FALSE}
### Per-transect BB abundance: species-level
bb_abund_sp <- net %>%
  group_by(site, date, yday, gdd.cum, year, elev.class2, 
           elev.mean, phase, break1.gdd, break2.gdd, bb.sp) %>%
  summarize(bb.abund = n()) %>%
  ungroup() %>%
  spread(bb.sp, bb.abund) %>%
  gather(bb.sp, bb.abund, -c(site, date, year, yday, gdd.cum, elev.class2, 
                             elev.mean, phase, break1.gdd, break2.gdd)) %>%
  replace_na(list(bb.abund = 0)) %>%
  mutate(bb.sp = factor(bb.sp))

### Per-transect BB abundance: species-level, by caste
bb_abund_sp_caste <- net %>%
  group_by(site, date, yday, gdd.cum, year, elev.class2, 
           elev.mean, phase, break1.gdd, break2.gdd, bb.sp, caste) %>%
  summarize(bb.abund = n()) %>%
  group_by(caste) %>%  # The next several lines of crazy nonsense are all about filling in zeros where a given caste*species was not observed.
  spread(bb.sp, bb.abund) %>%
  gather(bb.sp, bb.abund, -c(site, date, year, yday, gdd.cum, elev.class2, 
                             elev.mean, phase, break1.gdd, break2.gdd, caste)) %>%
  group_by(bb.sp) %>%
  spread(caste, bb.abund) %>%
  gather(caste, bb.abund, -c(site, date, year, yday, gdd.cum, elev.class2, 
                             elev.mean, phase, break1.gdd, break2.gdd, bb.sp)) %>%
  replace_na(list(bb.abund = 0)) %>%
  ungroup() %>%
  mutate(bb.sp = factor(bb.sp))

### Per-transect floral abund: by species
fl_abund_sp <- survey %>%
  group_by(site, date, year, plant.sp, yday) %>%
  summarize(fl.abund = sum(flower.cover)) %>%
  ungroup() %>%
  spread(plant.sp, fl.abund) %>%
  gather(plant.sp, fl.abund, -c(site, date, year, yday)) %>%
  replace_na(list(fl.abund = 0)) %>%
  mutate(plant.sp = factor(plant.sp)) %>%
  dplyr::select(site, year, yday, plant.sp, fl.abund) %>%  # this is an attempt to manage the memory crashes I'm getting with a later join step
  group_by(site, year, yday) %>%
  mutate(prop.fl.abund = fl.abund/sum(fl.abund))
```

### Get most important plants for each BB species
```{r}
# Per transect visitation rates of flower species per BB species
visit_rates <- net %>%
  group_by(bb.sp, plant.sp, year, site, yday) %>%
  summarize(visits = n()) %>%
  group_by(bb.sp) %>%
  spread(plant.sp, visits) %>%
  gather(plant.sp, visits, -c(bb.sp, year, site, yday)) %>%
  group_by(plant.sp) %>%
  spread(bb.sp, visits) %>%
  gather(bb.sp, visits, -c(plant.sp, year, site, yday)) %>%
  group_by(year, site, bb.sp, yday) %>%
  replace_na(list(visits = 0)) %>%
  group_by(site, year, yday, bb.sp) %>%
  mutate(prop.visits = visits/sum(visits)) %>% # This will create NAs wherever a given bb=fl combo was all zeros, so we need to set the NAs to 0 
  replace_na(list(prop.visits = 0)) %>%
  arrange(site, year, yday, bb.sp, -prop.visits) %>%
  ungroup() %>%
  dplyr::select(year, site, yday, bb.sp, plant.sp, visits, prop.visits) # this is an attempt to manage the memory crashes I'm getting with the subsequent join step

### Join FL abundance data to FL visitation data
fl_abund_x_visitation <- visit_rates %>%
  inner_join(fl_abund_sp) %>%
  filter(fl.abund > 0 & prop.fl.abund < 1) %>% # Can't calculate preference where the plant.sp is absent or where it is the only plant present
  inner_join(bb_abund_sp, by = c("site", "year", "yday", "bb.sp")) %>%
  filter(bb.abund > 0) # Can't calculate preference where the bb.sp is absent

fl_abund_x_visitation.wm <- fl_abund_x_visitation %>%
  group_by(year, bb.sp, plant.sp) %>%
  summarize(prop.fl.abund.m = mean(prop.fl.abund),
            prop.visits.wm = weighted.mean(prop.visits, bb.abund))
```

### Okay, here's what we're going to do. Preference has a strong bimodal pattern, with lots of plant species totally ignored while the ones that are foraged upon are visited in proportion to their abundance. We can't model that latter pattern well if we retain all the totally ignored plant species. So, let's model the visited plants separately from the unvisited ones. 
```{r}
### Call linear model on the subset of the floral species that were visited at least once
pref.lm <- lm(log(prop.visits.wm) ~ log(prop.fl.abund.m), # log on log regression is common for proportional data
              data = filter(fl_abund_x_visitation.wm, prop.visits.wm > 0))

summary(pref.lm)
plot(pref.lm)

ggplot(filter(fl_abund_x_visitation.wm, prop.visits.wm > 0),
       aes(log(prop.fl.abund.m), log(prop.visits.wm))) +
  geom_point() +
  geom_smooth(method = "lm")

### Extract residuals
pref.lm.resid <- residuals(pref.lm) %>%
  data.frame() %>%
  as_tibble() %>%
  rename(preference = 1)
```

Join residuals back into data and summarize by site-date-bb.sp (this deserves its own code chunk)
```{r}
#################################
### Build preference data set ###
#################################
# Start with the mean prop.fl.abund and weighted mean prop.visits
pref <- fl_abund_x_visitation.wm %>%  
  ungroup() %>%

# Remove bb*fl*year combos that had 0 visits
  filter(prop.visits.wm > 0) %>%

# With the zeroes removed, what's left can be joined 1:1 with the 
# residuals of the model above, i.e. the raw preference data
  bind_cols(pref.lm.resid) %>%

# Exponentiate the residuals to back-transform the data and ensure 
# that preference is > 0 for all visited plant species.
  mutate(preference = exp(preference)) %>%

# Now we need to add back all the bb*fl*year combos that had 0 visits
# They we have NAs for preference and for floral abundance columns
  full_join(unique(dplyr::select(visit_rates, year, bb.sp, plant.sp))) %>%

# Replace the NAs of the preference column with 0s; 
  replace_na(list(preference = 0)) %>%

# Drop the columns used to calculate preference so that they
# don't confuse us when we add site-date level abundance data
  dplyr::select(year, bb.sp, plant.sp, preference)

#################################
### Join with FL and BB abund ###
#################################

pref_wt_abund <- fl_abund_sp %>%
  full_join(pref, by = c("year", "plant.sp")) %>%  # This step introduces a small handful of NA because there were certain cases where a plant species was not observed at all in a given year. We will just omit those zeros from our data since they won't affect the analysis significantly
  mutate(fl.abund.weighted = fl.abund*preference) %>%
  group_by(site, year, yday, bb.sp) %>%
  summarize(fl.abund.weighted = sum(fl.abund.weighted, na.rm = TRUE)) %>%
  left_join(site_data, by = "site") %>%
  filter(!is.na(site)) %>%
  left_join(bb_abund_sp)
```

Preference-weighted floral abundance by year
```{r}
pref_sum <- pref_wt_abund %>%
  group_by(site, year, bb.sp) %>%
  summarize(fl.abund.weighted = mean(fl.abund.weighted))

ggplot(pref_sum, aes(year, fl.abund.weighted, color = bb.sp)) +
  geom_point()

ggplot(pref_sum, aes(year, log(fl.abund.weighted))) +
  geom_boxplot() +
  facet_wrap(~bb.sp, scales = "free")
```

GAM modeling of preference-weighted abundance per bb species
## Against any potential objections to my use of an added constant to avoid the log(0) problem, I would argue that when what is being measured is a resource, the presence of a resource in vanishingly small amounts is biologically equivalent to its absence. I suppose the strength of this argument hinges on how much we trust that our samples represent our sites. But if they don't, what are we even doing here? note that this argument does not hold when one is concerned with diversity and species distributions, since then a species rarity is a very different thing that its absence.
```{r}
### Set constants
k <- min(
  filter(pref_wt_abund, fl.abund.weighted > 0)$fl.abund.weighted
)

### Call models
weighted_fl_abundance.2010 <- bam(log(fl.abund.weighted + k) ~ te(yday, elev.mean, 
                                                  by = factor(bb.sp),
                                                  bs= c("tp", "tp"), 
                                                  k=c(10, 10), m=1),
                        data = filter(pref_wt_abund, year == 2010),
                        discrete = TRUE,
                        method = "fREML") %>% getViz()

check.gamViz(weighted_fl_abundance.2010)
summary(weighted_fl_abundance.2010)
print(plot(weighted_fl_abundance.2010 , allTerms = TRUE), pages = 4)

weighted_fl_abundance.2011 <- bam(log(fl.abund.weighted + k) ~ te(yday, elev.mean, 
                                                  by = factor(bb.sp),
                                                  bs= c("tp", "tp"), 
                                                  k=c(10, 10), m=1),
                        data = filter(pref_wt_abund, year == 2011),
                        discrete = TRUE,
                        method = "fREML") %>% getViz()

check.gamViz(weighted_fl_abundance.2011)
summary(weighted_fl_abundance.2011)
print(plot(weighted_fl_abundance.2011 , allTerms = TRUE), pages = 4)

weighted_fl_abundance.2012 <- bam(log(fl.abund.weighted + k) ~ te(yday, elev.mean, 
                                                  by = factor(bb.sp),
                                                  bs= c("tp", "tp"), 
                                                  k=c(10, 10), m=1),
                        data = filter(pref_wt_abund, year == 2012),
                        discrete = TRUE,
                        method = "fREML") %>% getViz()

check.gamViz(weighted_fl_abundance.2012)
summary(weighted_fl_abundance.2012)
print(plot(weighted_fl_abundance.2012 , allTerms = TRUE), pages = 4)
```

### Plotting bb.abund against fl.abund.weighted
```{r}
ggplot(pref_wt_abund, aes(log(fl.abund.weighted + 0.01), log(bb.abund), color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm") +
  facet_wrap(~bb.sp, scales = "free")

ggplot(pref_wt_abund, aes(log(fl.abund.weighted + 0.01), log(bb.abund), color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm") 
```

### Plot fl.abund.weighted against phase and elev.class
```{r}
ggplot(pref_wt_abund, aes(phase, log(fl.abund.weighted + k), color = year)) +
  geom_boxplot() +
  facet_wrap(~bb.sp)

ggplot(pref_wt_abund, aes(elev.class2, log(fl.abund.weighted + k), color = year)) +
  geom_boxplot() +
  facet_wrap(~bb.sp)
```

### 2020-07-02 writing/docx_template_generator.Rmd
broke down analyses into separate scripts; did some other things

---
title: "docx_template"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2020-07-02 writing/manuscript_v3.Rmd
broke down analyses into separate scripts; did some other things

---
title: Bumble bees and wildflowers along an elevational climate gradient exhibit niche
  asymmetry and non-linear abundance patterns
author:
- Douglas B. Sponsler
- Fabrice Requier
- Katharina Kallnik
- Alice Claßen
- A. Fabienne Maihoff
- Ingolf Steffan-Dewenter 
output:
  # pdf_document: default
  # word_document:
  #   reference_docx: docx_template.docx
bibliography: refs.bib
output: html_document
---

**Introduction**

The nascent reality of climate change has already revealed the fragile contingency of ecological paradigms that rely on the assumption of climate stability [@Hannah2002-nv; @Walther2002-mq]. Thus, one of the central questions facing ecologists today is how climate change will affect the future fitness and distribution of species, together with the composition and function of ecological communities. While it is intuitive that climate change should affect species and communities through environmental filtering, it is now recognized that climate change can also restructure biotic interactions [@Tylianakis2008-fi], altering the fitness landscape that governs community assembly and species coexistence [@Chesson2000-aw].

*[paragraph about Bombus being cold-adapted and generally distributed in high latitude or high elevation habitats]*
Bumble bees (Hymenoptera: *Bombus* spp.) are cold-adapted species that likely originated in cooling climate of the Eocene-Oligocene boundary (34 mya) and now distributed mainly at higher latitudes and elevations [@Hines2008-er].

Steep elevation gradients capture a broad range of climatic conditions within a small geographical area, and can thus serve

**Methods**
*Study sites*
*Sampling*
*Growing degree day analysis*
*Diversity modeling*
- GDM
*Abundance modeling*
- Floral preference modeling
- Preference-weighted floral abundance
- GAM
*Network analysis*
- Bootstrapping
- Network metrics

**Results**
*Diversity: Turnover asymmetry between bumble bees and floral hosts*
Beta diversity analysis

*Abundance: Heterogeneous distribution of bumble bees and floral hosts over elevation x time surface*
- Contrast the broadly heterogeneous distribution of habitat generalist BBs to the restricted distribution of high-elevation specialists
- Heterogeneous distribution of generalists indicates either or both movement along the elevation gradient or heterogeneous colonization along the elevation gradient.
- Concentration of preferred resources near the tree line
- Habitat specialists are often separated from their preferred host plants, probably due to thermal constraints, possibly due to competition, maybe both
- Habitat generalists vary in the degree to which their distribution matches that of their preferred hosts, suggesting that they are constrained by colonization, competition, or both.

*Network structure*
- Marked changes in network structure associated with the tree line

*Annual differences*
- 2010 was much cooler than 2011 and 2012 (2012 was slightly warmer than 2011)
- Overall BB abundance was highest in 2011
- Overall FL abundance was lowest in 2012
- The bss species group (B. terrestris + B. lucorum + B. cryptarum; mostly the first 2 since B. cryptarum is rare) seems to have largely failed to establish in 2012, with an early season low elevation peak followed by near absence the rest of the year across the whole elevation gradient

**Discussion**

### 2020-07-02 data_proc.R

@@ -8,7 +8,7 @@ fl_survey <- read_delim("./raw_data/data_plantcover_forDoug_2020.01.17.csv", del
   na.omit() %>%
   separate(col = species_plant, into = c("genus", "species", "w", "x", "y", "z"), sep = " ") %>%
   unite(species_plant, genus, species, sep = " ") %>%
-  select(-c(w,x,y,z)) %>%
+  dplyr::select(-c(w,x,y,z)) %>%
   mutate(species_plant = str_replace_all(species_plant, 
                                          c("Arabis caerula" = "Arabis caerulea", # correct nomenclature typos and synonyms
                                            "Arctostaphylos alpina" = "Arctous alpina",
@@ -21,7 +21,8 @@ fl_survey <- read_delim("./raw_data/data_plantcover_forDoug_2020.01.17.csv", del
   mutate(temp.date = as_date(dayofyear),
          day = day(temp.date),
          month = month(temp.date),
-         date = make_date(year, month, day))
+         date = make_date(year, month, day)) %>%
+  dplyr::select(-temp.date)
   
 write_csv(fl_survey, "./processed_data/floral_survey.csv")
 
@@ -56,7 +57,8 @@ network <- read_delim("./raw_data/data_bumblebee_forDoug_2020.04.17.csv", delim 
                                         "Bombus sylvestris" = "Psithyrus",
                                         "Bombus quadricolor" = "Psithyrus",
                                         "Bombus campestris" = "Psithyrus",
-                                        "Bombus flavidus" = "Psithyrus")))
+                                        "Bombus flavidus" = "Psithyrus"))) %>%
+  dplyr::select(-temp.date)
 
 write_csv(network, "./processed_data/network.csv")
 
@@ -97,7 +99,7 @@ site_data <- read_tsv("./raw_data/SiteLocationFabriceUpdate2020.tsv",
   mutate(elev.class = factor(elev.class, ordered = TRUE, levels = c("unten", "mitte", "oben"))) %>%
   mutate(elev.class2 = case_when(
     elev.mean < 1000 ~ "low",
-    elev.mean >= 1000 & elev.mean < 1500 ~ "mid",
+    elev.mean >= 1000 & elev.mean < 1705 ~ "mid",
     elev.mean >= 1750 ~ "high"
   ))
 
@@ -267,3 +269,29 @@ psite_2012 <- parasites2012 %>%
[…]

### 2020-07-03 analyses/preference.Rmd
fixed some things, broke some things

@@
 ### Load network and survey data
-```{r, echo=FALSE, include=FALSE}
-### BB traits
-bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
-  dplyr::select(subgenus, bb.sp, pbl.w, pbl.w.class)
-
-### FL Kugler classification
-fl_traits <- read.csv("../processed_data/floral_k_type_nom_fix.csv") %>%
-  dplyr::select(plant.sp, k.type) %>%
-  mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"),
-         k.type.ss = str_extract(k.type, "[-+]?[0-9]*"))
-
-### FL corolla depth (from Benadi)
-fl_corolla_depth <- read.csv("../processed_data/flower_morphology_GitaBenadi.csv") %>%
-  group_by(plant.sp = Species) %>%
-  summarize(corolla.depth = mean(LN))
-
-### FL higher  taxonomy
-fl_tax <- read_csv("../processed_data/floral_tax.csv")
-
-### Colony life cycle break points
-breaks <- read_csv("./output/breaks.csv") %>%
-  mutate(year = factor(year))
-
+```{r}
 ### BB-FL visitation data
 net <- read_csv("../processed_data/network.csv") %>%
-  mutate(year = factor(year)) %>% # convert year to factor
-  full_join(bb_traits, by = c("bb.sp" = "bb.sp")) %>%
-  left_join(site_data, by = c("site")) %>%
-  left_join(gdd, by = c("site", "date", "year")) %>%
-  left_join(breaks, by = c("site", "year")) %>%
-  left_join(fl_traits) %>%
-  left_join(fl_corolla_depth) %>%
-  left_join(fl_tax, by = "plant.genus") %>%
-  #left_join(temperature, by = c("site", "date")) %>%
-  ungroup() %>%
-  mutate(phase = case_when(
-    gdd.cum < break1.gdd ~ "founding",
-    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
-    gdd.cum > break2.gdd ~ "reproductive"
-  )) %>%
-  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
-  mutate(caste = case_when(
-    caste == "worker" ~ "worker",
-    caste == "male" ~ "reproductive",
-    caste == "youngqueen" ~ "reproductive",
-    caste == "queen" ~ "foundress",
-    caste == "oldqueen" ~ "foundress"
-  )) %>%
-  rename(yday = dayofyear) %>%
-  arrange(site, date)
+  dplyr::select(site, date, bb.sp, plant.sp)
 
@@
 survey <- read_csv("../processed_data/floral_survey.csv") %>%
-  mutate(year = factor(year)) %>% # convert year to factor
-  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime, not specifically for a given site*date)
-  left_join(gdd, by = c("site", "date", "year")) %>%
-  left_join(breaks, by = c("site", "year")) %>%
-  left_join(site_data, by = c("site")) %>%
-  left_join(fl_traits) %>%
-  left_join(fl_traits) %>%
-  left_join(fl_tax) %>%
-  #left_join(temperature, by = c("site", "date")) %>%
-  ungroup() %>%
-  mutate(phase = case_when(
-    gdd.cum < break1.gdd ~ "founding",
-    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
-    gdd.cum > break2.gdd ~ "reproductive"
-  )) %>%
-  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
-  rename(yday = dayofyear) %>%
-  arrange(site, date)
+  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
+  dplyr::select(site, date, plant.sp, flower.cover)
+
+shared_transects <- inner_join(net, survey, by = c("site", "date")) %>%
+  dplyr::select(site, date) %>%
+  unique()
 ```
@@
 bb_abund_sp <- net %>%
-  group_by(site, date, yday, gdd.cum, year, elev.class2, 
-           elev.mean, phase, break1.gdd, break2.gdd, bb.sp) %>%
-  summarize(bb.abund = n()) %>%
-  ungroup() %>%
-  spread(bb.sp, bb.abund) %>%
-  gather(bb.sp, bb.abund, -c(site, date, year, yday, gdd.cum, elev.class2, 
-                             elev.mean, phase, break1.gdd, break2.gdd)) %>%
-  replace_na(list(bb.abund = 0)) %>%
-  mutate(bb.sp = factor(bb.sp))
-
-### Per-transect BB abundance: species-level, by caste
-bb_abund_sp_caste <- net %>%
-  group_by(site, date, yday, gdd.cum, year, elev.class2, 
-           elev.mean, phase, break1.gdd, break2.gdd, bb.sp, caste) %>%
-  summarize(bb.abund = n()) %>%
-  group_by(caste) %>%  # The next several lines of crazy nonsense are all about filling in zeros where a given caste*species was not observed.
-  spread(bb.sp, bb.abund) %>%
-  gather(bb.sp, bb.abund, -c(site, date, year, yday, gdd.cum, elev.class2, 
-                             elev.mean, phase, break1.gdd, break2.gdd, caste)) %>%
-  group_by(bb.sp) %>%
-  spread(caste, bb.abund) %>%
-  gather(caste, bb.abund, -c(site, date, year, yday, gdd.cum, elev.class2, 
-                             elev.mean, phase, break1.gdd, break2.gdd, bb.sp)) %>%
-  replace_na(list(bb.abund = 0)) %>%
-  ungroup() %>%
-  mutate(bb.sp = factor(bb.sp))
+  dplyr::select(site, date, bb.sp) %>%
+  semi_join(shared_transects) %>% # I only want BB data with corresponding floral data
+  group_by(site, date, bb.sp) %>% # group by bb.sp*transect (and other variables we want to retain)
+  summarize(bb.abund = n()) %>% # get bb.sp*transect abundance
+  ungroup %>%
+  complete(site, date, bb.sp, fill = list(bb.abund = 0)) # complete implicit absences
 
@@
 fl_abund_sp <- survey %>%
-  group_by(site, date, year, plant.sp, yday) %>%
-  summarize(fl.abund = sum(flower.cover)) %>%
+  semi_join(shared_transects) %>%
+  group_by(site, date, plant.sp) %>% # group by plant.sp*transect
+  summarize(fl.abund = sum(flower.cover)) %>% # get plant.sp*transect total flower cover; this step really shouldn't be required, but in a few cases, there is more than one abundance entry for the same species on a given transect; not sure how that happened, but it's in the original data that Fabrice sent me; alternatively, I could drop the dups or average them.
+We'll see what Kalli and Fabrice have to say.
+  group_by(site, date) %>%
+  mutate(prop.fl.abund = fl.abund/sum(fl.abund)) %>% # get the proportional flower cover for each species.
   ungroup() %>%
-  spread(plant.sp, fl.abund) %>%
-  gather(plant.sp, fl.abund, -c(site, date, year, yday)) %>%
-  replace_na(list(fl.abund = 0)) %>%
-  mutate(plant.sp = factor(plant.sp)) %>%
-  dplyr::select(site, year, yday, plant.sp, fl.abund) %>%  # this is an attempt to manage the memory crashes I'm getting with a later join step
-  group_by(site, year, yday) %>%
-  mutate(prop.fl.abund = fl.abund/sum(fl.abund))
+  full_join(dplyr::select(net, site, date, plant.sp), by = c("site", "date", "plant.sp")) %>% # add the handful of species occurring in the visitation network but missing from the survey data
+  complete(site, date, plant.sp, fill = list(fl.abund = 0, prop.fl.abund = 0)) %>%
+  unique()
 ```
@@
 ```{r}
-# Per transect visitation rates of flower species per BB species
-visit_rates <- net %>%
-  group_by(bb.sp, plant.sp, year, site, yday) %>%
-  summarize(visits = n()) %>%
-  group_by(bb.sp) %>%
-  spread(plant.sp, visits) %>%
-  gather(plant.sp, visits, -c(bb.sp, year, site, yday)) %>%
-  group_by(plant.sp) %>%
-  spread(bb.sp, visits) %>%
-  gather(bb.sp, visits, -c(plant.sp, year, site, yday)) %>%
-  group_by(year, site, bb.sp, yday) %>%
-  replace_na(list(visits = 0)) %>%
-  group_by(site, year, yday, bb.sp) %>%
-  mutate(prop.visits = visits/sum(visits)) %>% # This will create NAs wherever a given bb=fl combo was all zeros, so we need to set the NAs to 0 
-  replace_na(list(prop.visits = 0)) %>%
-  arrange(site, year, yday, bb.sp, -prop.visits) %>%
-  ungroup() %>%
-  dplyr::select(year, site, yday, bb.sp, plant.sp, visits, prop.visits) # this is an attempt to manage the memory crashes I'm getting with the subsequent join step
+# All possible plant-bb combos
+fl_bb_pair_filler <- full_join(bb_abund_sp, fl_abund_sp) %>%
+  dplyr::select(site, date, bb.sp, plant.sp) %>%
+  complete(site, date, bb.sp, plant.sp)
 
-### Join FL abundance data to FL visitation data
-fl_abund_x_visitation <- visit_rates %>%
-  inner_join(fl_abund_sp) %>%
-  filter(fl.abund > 0 & prop.fl.abund < 1) %>% # Can't calculate preference where the plant.sp is absent or where it is the only plant present
-  inner_join(bb_abund_sp, by = c("site", "year", "yday", "bb.sp")) %>%
-  filter(bb.abund > 0) # Can't calculate preference where the bb.sp is absent
+fl_bb_pair_filter_sub <- fl_bb_pair_filler %>%
+  right_join(filter())
 
-fl_abund_x_visitation.wm <- fl_abund_x_visitation %>%
-  group_by(year, bb.sp, plant.sp) %>%
+filter(fl_bb_all, is.na(site))
+filter(fl_bb_all, is.na(date))
+filter(fl_bb_all, is.na(bb.sp))
+filter(fl_bb_all, is.na(plant.sp))
+filter(fl_bb_all, plant.sp == "Rubus idaeus")
+
+fl_abund_sp_unique <- unique(fl_abund_sp)
+
+# This creates a data set where each bumble bee species present at a given site*date is paired with each floral species present at a given site date, even if there was never a visitation between them.
+observed_visits <- net %>%
+  semi_join(shared_transects) %>% # include only site*date combos where floral surveying was done
+  group_by(site, date, bb.sp, plant.sp) %>% # group by bb*plant*transect
+  summarize(visits = n()) %>% # total visits per bb-fl pair
+  group_by(site, date, bb.sp) %>%
+  mutate(prop.visits = visits/sum(visits))
+
+# All possible plants * site * date
+possible_plants <- bind_rows(dplyr::select(net, c(site, date, plant.sp)),
+                             dplyr::select(survey, c(site, date, plant.sp))) %>%
+  unique() %>%
+  complete(site, date, plant.sp)
+
+# All possible bumbles * site * date
+possible_bb <- net %>%
+  dplyr::select(site, date, bb.sp) %>%
+  unique() %>%
+  complete(site, date, bb.sp)
+
+# All possible bumble-plant pairings *site * date, filtered to include only those when the given bb.sp was present and the given plant.sp was *either* present *or* visited (see note below)
+visitation_by_abundance <- full_join(possible_plants, possible_bb) %>%
+  left_join(bb_abund_sp) %>%
+  left_join(fl_abund_sp) %>%
+  #filter(fl.abund > 0) %>%
+  filter(bb.abund > 0) %>%
+  left_join(observed_visits) %>%
+  filter(fl.abund > 0 | visits > 0) %>% # In about 10% of the bb.sp*plant.sp*site*year combos, the visited plant was not recorded in the floral survey data.
+There are a few different ways we could handle this, but I think an argument could be make that treating those plants as if they had zero abundance (even though they were evidently present gets close to the truth, since they must have been rather rare to have been missed in the surveying.)
+  replace_na(list(visits = 0, prop.visits = 0))
+
+write_csv(visitation_by_abundance, "../analyses/output/visitation_by_abundance.csv")
+
+ggplot(visitation_by_abundance, aes(log(prop.fl.abund + 0.01), log(prop.visits + 0.01))) +
+  geom_point() +
+  geom_smooth(method = "lm")
+
+### Get mean abundance/visitation for each plant*bb combo
+visitation_by_abundance_mean <- visitation_by_abundance %>%
+  mutate(year = factor(year(date))) %>%
+  #group_by(year, plant.sp, bb.sp) %>%
+  group_by(plant.sp, bb.sp) %>%
   summarize(prop.fl.abund.m = mean(prop.fl.abund),
-            prop.visits.wm = weighted.mean(prop.visits, bb.abund))
+            prop.visits.wm = weighted.mean(prop.visits, bb.abund)) %>%
+  mutate(selection.ratio = prop.visits.wm/prop.fl.abund.m)
+
+write_csv(visitation_by_abundance_mean, "../analyses/output/visitation_by_abundance_mean.csv")
+
+ggplot(visitation_by_abundance_mean, aes(selection.ratio)) +
+  geom_histogram()
+
+filter(visitation_by_abundance_mean, plant.sp == "Trifolium pratense")
+
+ggplot(visitation_by_abundance_mean, 
+       aes(log(prop.fl.abund.m + 0.01), log(prop.visits.wm + 0.01))) +
+  geom_point() +
+  geom_smooth(method = "lm")
+
+ggplot(filter(visitation_by_abundance_mean, prop.visits.wm > 0), 
+       aes(log(prop.fl.abund.m + 0.01), log(prop.visits.wm + 0.01))) +
+  geom_point() +
+  geom_smooth(method = "lm")
 ```
@@
 ### Call linear model on the subset of the floral species that were visited at least once
-pref.lm <- lm(log(prop.visits.wm) ~ log(prop.fl.abund.m), # log on log regression is common for proportional data
-              data = filter(fl_abund_x_visitation.wm, prop.visits.wm > 0))
+pref.lm <- lm(log(prop.visits.wm) ~ log(prop.fl.abund.m + 0.01), # log on log regression is common for proportional data
+              data = filter(visitation_by_abundance_mean, prop.visits.wm > 0))
 
@@
 plot(pref.lm)
-
-ggplot(filter(fl_abund_x_visitation.wm, prop.visits.wm > 0),
-       aes(log(prop.fl.abund.m), log(prop.visits.wm))) +
-  geom_point() +
-  geom_smooth(method = "lm")
 
@@
 # Start with the mean prop.fl.abund and weighted mean prop.visits
-pref <- fl_abund_x_visitation.wm %>%  
+preference <- visitation_by_abundance_mean %>%  
   ungroup() %>%
@@
 # Now we need to add back all the bb*fl*year combos that had 0 visits
-# They we have NAs for preference and for floral abundance columns
-  full_join(unique(dplyr::select(visit_rates, year, bb.sp, plant.sp))) %>%
+# Then we have NAs for preference and floral abundance columns
+  full_join(unique(dplyr::select(visitation_by_abundance_mean, bb.sp, plant.sp))) %>%
 
@@
 # don't confuse us when we add site-date level abundance data
-  dplyr::select(year, bb.sp, plant.sp, preference)
+  dplyr::select(bb.sp, plant.sp, preference)
 
@@
 
-pref_wt_abund <- fl_abund_sp %>%
-  full_join(pref, by = c("year", "plant.sp")) %>%  # This step introduces a small handful of NA because there were certain cases where a plant species was not observed at all in a given year.
-We will just omit those zeros from our data since they won't affect the analysis significantly
-  mutate(fl.abund.weighted = fl.abund*preference) %>%
-  group_by(site, year, yday, bb.sp) %>%
-  summarize(fl.abund.weighted = sum(fl.abund.weighted, na.rm = TRUE)) %>%
-  left_join(site_data, by = "site") %>%
-  filter(!is.na(site)) %>%
-  left_join(bb_abund_sp)
+preference_weighted_abund <- fl_abund_sp %>%
+  right_join(preference, by = c("plant.sp")) %>% 
+  mutate(fl.abund.pref_wt = fl.abund*preference) %>%
+  mutate(year = factor(year(date)),
+         yday = yday(date)) %>%
+  group_by(site, year, date, yday, bb.sp) %>%
+  summarize(net.fl.abund.pref_wt = sum(fl.abund.pref_wt, na.rm = TRUE)) %>%
+  ungroup() %>%
+  left_join(site_data, by = "site")
+
+ggplot(preference_weighted_abund, aes(net.fl.abund.pref_wt)) +
+  geom_histogram()
+
+ggplot(preference_weighted_abund, aes(log(net.fl.abund.pref_wt + 0.01))) +
+  geom_histogram()
 ```
@@
 ```{r}
-pref_sum <- pref_wt_abund %>%
+pref_sum <- preference_weighted_abund %>%
   group_by(site, year, bb.sp) %>%
-  summarize(fl.abund.weighted = mean(fl.abund.weighted))
+  summarize(mean.net.fl.abund.pref_wt = mean(net.fl.abund.pref_wt))
 
-ggplot(pref_sum, aes(year, fl.abund.weighted, color = bb.sp)) +
+ggplot(pref_sum, aes(year, mean.net.fl.abund.pref_wt, color = bb.sp)) +
   geom_point()
 
-ggplot(pref_sum, aes(year, log(fl.abund.weighted))) +
+ggplot(pref_sum, aes(year, log(mean.net.fl.abund.pref_wt + 0.01))) +
   geom_boxplot() +
@@
 k <- min(
-  filter(pref_wt_abund, fl.abund.weighted > 0)$fl.abund.weighted
+  filter(preference_weighted_abund, net.fl.abund.pref_wt > 0)$net.fl.abund.pref_wt
 )
@@
 ### Call models
-weighted_fl_abundance.2010 <- bam(log(fl.abund.weighted + k) ~ te(yday, elev.mean, 
+weighted_fl_abundance.2010 <- bam(net.fl.abund.pref_wt ~ te(yday, elev.mean, 
                                                             by = factor(bb.sp),
                                                             bs= c("tp", "tp"), 
-                                                  k=c(10, 10), m=1),
-                        data = filter(pref_wt_abund, year == 2010),
+                                                            k=c(12, 12), m=1),
+                        data = filter(preference_weighted_abund, year == 2010),
                         discrete = TRUE,
@@
 
-weighted_fl_abundance.2011 <- bam(log(fl.abund.weighted + k) ~ te(yday, elev.mean, 
+weighted_fl_abundance.2011 <- bam(log(net.fl.abund.pref_wt + k) ~ te(yday, elev.mean, 
                                                   by = factor(bb.sp),
@@
                                                   k=c(10, 10), m=1),
-                        data = filter(pref_wt_abund, year == 2011),
+                        data = filter(preference_weighted_abund, year == 2011),
                         discrete = TRUE,
@@
 
-weighted_fl_abundance.2012 <- bam(log(fl.abund.weighted + k) ~ te(yday, elev.mean, 
+weighted_fl_abundance.2012 <- bam(log(net.fl.abund.pref_wt + k) ~ te(yday, elev.mean, 
                                                   by = factor(bb.sp),
@@
                                                   k=c(10, 10), m=1),
-                        data = filter(pref_wt_abund, year == 2012),
+                        data = filter(preference_weighted_abund, year == 2012),
                         discrete = TRUE,

### 2020-07-03 writing/manuscript_v3.Rmd
fixed some things, broke some things

@@
 title: Bumble bees and wildflowers along an elevational climate gradient exhibit niche
-  asymmetry and non-linear abundance patterns
+  asymmetry and heterogeneous abundance patterns
 author:
@@
 - Floral preference modeling
+
+Floral preference was estimated for all pairwise combinations of bumble bee and plant species.
+
 - Preference-weighted floral abundance

### 2020-07-08 analyses/preference.Rmd
update

@@
 site_data <- read_csv("../processed_data/site_data.csv") %>%
-  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
+  #dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
   mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
@@
 
+survey_dups <- survey %>%
+  group_by(site, date, plant.sp) %>%
+  mutate(dups = n()) %>%
+  filter(dups > 1)
+
+dup_species <- survey_dups %>%
+  group_by(plant.sp) %>%
+  mutate(dups = n()/2) %>%
+  group_by(site, date, plant.sp, dups) %>%
+  mutate(different = length(unique(flower.cover)) -1) %>%
+  group_by(site, date, plant.sp, dups) %>%
+  summarize(different = mean(different)) %>%
+  group_by(plant.sp, dups) %>%
+  summarize(different.dups = sum(different)) %>% 
+  arrange(-dups)
+
+### Shared transects
 shared_transects <- inner_join(net, survey, by = c("site", "date")) %>%
@@
   ungroup %>%
-  complete(site, date, bb.sp, fill = list(bb.abund = 0)) # complete implicit absences
+  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
+  mutate(year = factor(year(date)))
 
@@
   complete(site, date, plant.sp, fill = list(fl.abund = 0, prop.fl.abund = 0)) %>%
-  unique()
+  unique() %>%
+  mutate(year = factor(year(date)))
 ```
@@
 ```{r}
-# All possible plant-bb combos
-fl_bb_pair_filler <- full_join(bb_abund_sp, fl_abund_sp) %>%
-  dplyr::select(site, date, bb.sp, plant.sp) %>%
-  complete(site, date, bb.sp, plant.sp)
-
-fl_bb_pair_filter_sub <- fl_bb_pair_filler %>%
-  right_join(filter())
-
-filter(fl_bb_all, is.na(site))
-filter(fl_bb_all, is.na(date))
-filter(fl_bb_all, is.na(bb.sp))
-filter(fl_bb_all, is.na(plant.sp))
-filter(fl_bb_all, plant.sp == "Rubus idaeus")
-
-fl_abund_sp_unique <- unique(fl_abund_sp)
-
 # This creates a data set where each bumble bee species present at a given site*date is paired with each floral species present at a given site date, even if there was never a visitation between them.
@@
   left_join(fl_abund_sp) %>%
-  #filter(fl.abund > 0) %>%
   filter(bb.abund > 0) %>%
@@
 
-write_csv(visitation_by_abundance, "../analyses/output/visitation_by_abundance.csv")
-
-ggplot(visitation_by_abundance, aes(log(prop.fl.abund + 0.01), log(prop.visits + 0.01))) +
-  geom_point() +
-  geom_smooth(method = "lm")
+#write_csv(visitation_by_abundance, "../analyses/output/visitation_by_abundance.csv")
 
@@
   mutate(year = factor(year(date))) %>%
-  #group_by(year, plant.sp, bb.sp) %>%
   group_by(plant.sp, bb.sp) %>%
@@
 
-ggplot(visitation_by_abundance_mean, aes(selection.ratio)) +
-  geom_histogram()
-
-filter(visitation_by_abundance_mean, plant.sp == "Trifolium pratense")
-
-ggplot(visitation_by_abundance_mean, 
-       aes(log(prop.fl.abund.m + 0.01), log(prop.visits.wm + 0.01))) +
-  geom_point() +
-  geom_smooth(method = "lm")
-
 ggplot(filter(visitation_by_abundance_mean, prop.visits.wm > 0), 
-       aes(log(prop.fl.abund.m + 0.01), log(prop.visits.wm + 0.01))) +
-  geom_point() +
+       aes(log(prop.fl.abund.m + 0.01), log(prop.visits.wm))) +
+  geom_point(alpha = 0.5, aes(color = bb.sp)) +
   geom_smooth(method = "lm")
@@
 ### Call linear model on the subset of the floral species that were visited at least once
-pref.lm <- lm(log(prop.visits.wm) ~ log(prop.fl.abund.m + 0.01), # log on log regression is common for proportional data
+# log on log regression is common for proportional datam (log ratio regression)
+pref.lm <- lm(log(prop.visits.wm) ~ log(prop.fl.abund.m + 0.01), 
               data = filter(visitation_by_abundance_mean, prop.visits.wm > 0))
@@
 #################################
+bb_abund_sp_sum <- bb_abund_sp %>% # Get annual totals for BB species so that we can filter out really rare ones
+  group_by(year, bb.sp) %>%
+  summarize(bb.total.abund = sum(bb.abund))
 
-preference_weighted_abund <- fl_abund_sp %>%
-  right_join(preference, by = c("plant.sp")) %>% 
-  mutate(fl.abund.pref_wt = fl.abund*preference) %>%
-  mutate(year = factor(year(date)),
-         yday = yday(date)) %>%
-  group_by(site, year, date, yday, bb.sp) %>%
-  summarize(net.fl.abund.pref_wt = sum(fl.abund.pref_wt, na.rm = TRUE)) %>%
+fl_abund_wt <- fl_abund_sp %>%
+  right_join(preference) %>% 
+  mutate(fl.abund.wt = fl.abund*preference) %>%
+  mutate(yday = yday(date)) %>%
+  group_by(site, date, yday, bb.sp) %>%
+  summarize(fl.abund.wt = sum(fl.abund.wt, na.rm = TRUE)) %>%
   ungroup() %>%
-  left_join(site_data, by = "site")
+  left_join(bb_abund_sp) %>%
+  left_join(bb_abund_sp_sum) %>%
+  filter(bb.total.abund >= 5) %>%
+  left_join(site_data, by = "site") %>%
+  semi_join(survey, by = c("site", "date")) %>% # This gets rid of site*date combos that don't exist
+  mutate(year = factor(year(date)), # add in the years missing from the bb_abund join
+         bb.sp = factor(bb.sp), # convert bb.sp to a factor
+         management = factor(management)) # convert management to a factor
 
-ggplot(preference_weighted_abund, aes(net.fl.abund.pref_wt)) +
-  geom_histogram()
-
-ggplot(preference_weighted_abund, aes(log(net.fl.abund.pref_wt + 0.01))) +
-  geom_histogram()
-```
-
-Preference-weighted floral abundance by year
-```{r}
-pref_sum <- preference_weighted_abund %>%
-  group_by(site, year, bb.sp) %>%
-  summarize(mean.net.fl.abund.pref_wt = mean(net.fl.abund.pref_wt))
-
-ggplot(pref_sum, aes(year, mean.net.fl.abund.pref_wt, color = bb.sp)) +
-  geom_point()
-
-ggplot(pref_sum, aes(year, log(mean.net.fl.abund.pref_wt + 0.01))) +
-  geom_boxplot() +
-  facet_wrap(~bb.sp, scales = "free")
+write_csv(fl_abund_wt, "../analyses/output/fl_abund_wt.csv")
 ```
@@
 k <- min(
-  filter(preference_weighted_abund, net.fl.abund.pref_wt > 0)$net.fl.abund.pref_wt
+  filter(fl_abund_wt, fl.abund.wt > 0)$fl.abund.wt
 )
@@
 ### Call models
-weighted_fl_abundance.2010 <- bam(net.fl.abund.pref_wt ~ te(yday, elev.mean, 
-                                                            by = factor(bb.sp),
+weighted_fl_abundance.2010.qp <- bam(fl.abund.wt ~ bb.sp + te(yday, elev.mean, 
+                                                            by = bb.sp,
                                                             bs= c("tp", "tp"), 
-                                                            k=c(12, 12), m=1),
-                        data = filter(preference_weighted_abund, year == 2010),
+                                                            k=c(10, 10), m=1),
+                        data = filter(fl_abund_wt, year == 2010),
                         discrete = TRUE,
+                        family = "quasipoisson",
                         method = "fREML") %>% getViz()
 
-check.gamViz(weighted_fl_abundance.2010)
-summary(weighted_fl_abundance.2010)
-print(plot(weighted_fl_abundance.2010 , allTerms = TRUE), pages = 4)
+check.gamViz(weighted_fl_abundance.2010.qp)
+summary(weighted_fl_abundance.2010.qp)
+print(plot(weighted_fl_abundance.2010.qp , allTerms = TRUE), pages = 4)
 
-weighted_fl_abundance.2011 <- bam(log(net.fl.abund.pref_wt + k) ~ te(yday, elev.mean, 
-                                                  by = factor(bb.sp),
+weighted_fl_abundance.2010.qp2 <- bam(sqrt(fl.abund.wt) ~ bb.sp + te(yday, elev.mean, 
+                                                            by = bb.sp,
                                                             bs= c("tp", "tp"), 
                                                             k=c(10, 10), m=1),
-                        data = filter(preference_weighted_abund, year == 2011),
+                        data = filter(fl_abund_wt, year == 2010),
                         discrete = TRUE,
+                        family = "quasipoisson",
                         method = "fREML") %>% getViz()
 
-check.gamViz(weighted_fl_abundance.2011)
-summary(weighted_fl_abundance.2011)
-print(plot(weighted_fl_abundance.2011 , allTerms = TRUE), pages = 4)
+check.gamViz(weighted_fl_abundance.2010.qp2)
+summary(weighted_fl_abundance.2010.qp2)
+print(plot(weighted_fl_abundance.2010.qp2, allTerms = TRUE), pages = 3)
 
-weighted_fl_abundance.2012 <- bam(log(net.fl.abund.pref_wt + k) ~ te(yday, elev.mean, 
-                                                  by = factor(bb.sp),
+weighted_fl_abundance.2011.qp <- bam(fl.abund.wt ~ bb.sp + te(yday, elev.mean, 
+                                                            by = bb.sp,
                                                             bs= c("tp", "tp"), 
                                                             k=c(10, 10), m=1),
-                        data = filter(preference_weighted_abund, year == 2012),
+                        data = filter(fl_abund_wt, year == 2011),
                         discrete = TRUE,
+                        family = "quasipoisson",
                         method = "fREML") %>% getViz()
 
-check.gamViz(weighted_fl_abundance.2012)
-summary(weighted_fl_abundance.2012)
-print(plot(weighted_fl_abundance.2012 , allTerms = TRUE), pages = 4)
+check.gamViz(weighted_fl_abundance.2011.qp)
+summary(weighted_fl_abundance.2011.qp)
+print(plot(weighted_fl_abundance.2011.qp , allTerms = TRUE), pages = 4)
+
+weighted_fl_abundance.2011.qp2 <- bam(sqrt(fl.abund.wt) ~ bb.sp + te(yday, elev.mean, 
+                                                            by = bb.sp,
+                                                            bs= c("tp", "tp"), 
+                                                            k=c(10, 10), m=1),
+                        data = filter(fl_abund_wt, year == 2011),
+                        discrete = TRUE,
+                        family = "quasipoisson",
+                        method = "fREML") %>% getViz()
+
+check.gamViz(weighted_fl_abundance.2011.qp2)
+summary(weighted_fl_abundance.2011.qp2)
+print(plot(weighted_fl_abundance.2011.qp2, allTerms = TRUE), pages = 4)
+
+weighted_fl_abundance.2012.qp <- bam(fl.abund.wt ~ bb.sp + te(yday, elev.mean, 
+                                                            by = bb.sp,
+                                                            bs= c("tp", "tp"), 
+                                                            k=c(10, 10), m=1),
+                        data = filter(fl_abund_wt, year == 2012),
+                        discrete = TRUE,
+                        family = "quasipoisson",
+                        method = "fREML") %>% getViz()
+
+check.gamViz(weighted_fl_abundance.2012.qp)
+summary(weighted_fl_abundance.2012.qp)
+print(plot(weighted_fl_abundance.2012.qp , allTerms = TRUE), pages = 4)
+
+weighted_fl_abundance.2012.qp2_re <- bam(sqrt(fl.abund.wt) ~ s(bb.sp, bs = "re") + 
+                                        management + 
+                                        te(yday, elev.mean, 
+                                           by = bb.sp,
+                                           bs= c("tp", "tp"), 
+                                           k=c(10, 10), m=1),
+                        data = filter(fl_abund_wt, year == 2012),
+                        discrete = TRUE,
+                        select = TRUE,
+                        family = "quasipoisson",
+                        method = "fREML") %>% getViz()
+
+check.gamViz(weighted_fl_abundance.2012.qp2_re)
+summary(weighted_fl_abundance.2012.qp2_re)
+print(plot(weighted_fl_abundance.2012.qp2_re, allTerms = TRUE), pages = 4)
+
+weighted_fl_abundance.2012.qp2_p <- bam(sqrt(fl.abund.wt) ~ bb.sp + management + 
+                                        te(yday, elev.mean, 
+                                           by = bb.sp,
+                                           bs= c("tp", "tp"), 
+                                           k=c(10, 10), m=1),
+                        data = filter(fl_abund_wt, year == 2012),
+                        discrete = TRUE,
+                        family = "quasipoisson",
+                        method = "fREML") %>% getViz()
+
+check.gamViz(weighted_fl_abundance.2012.qp2_p)
+summary(weighted_fl_abundance.2012.qp2_p)
+print(plot(weighted_fl_abundance.2012.qp2_p, allTerms = TRUE), pages = 4)
+
+weighted_fl_abundance.qp2 <- bam(sqrt(fl.abund.wt) ~ bb.sp + year + te(yday, elev.mean, 
+                                                            by = bb.sp,
+                                                            bs= c("tp", "tp"), 
+                                                            k=c(10, 10), m=1),
+                        data = fl_abund_wt,
+                        discrete = TRUE,
+                        family = "quasipoisson",
+                        method = "fREML") %>% getViz()
+
+check.gamViz(weighted_fl_abundance.qp2)
+summary(weighted_fl_abundance.qp2)
+print(plot(weighted_fl_abundance.qp2 , allTerms = TRUE), pages = 4)
+
 ```
 
-### Plotting bb.abund against fl.abund.weighted
+### Abundance mismatch
 ```{r}
-ggplot(pref_wt_abund, aes(log(fl.abund.weighted + 0.01), log(bb.abund), color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "lm") +
+mismatch <- fl_abund_wt %>%
+  filter(!is.na(bb.abund)) %>%
+  group_by(year, bb.sp) %>%
+  mutate(bb.abund.rank = rank(-bb.abund, ties.method = "average"),
+            fl.abund.wt.rank = rank(-fl.abund.wt, ties.method = "average"))
+
+v <- c(10,9, 5, 8,7)
+v.rank = rank(v)
+
+ggplot(mismatch, aes(fl.abund.wt.rank, bb.abund.rank)) +
+  geom_point() +
+  facet_wrap(~bb.sp)
+
+lm_mismatch <- lmer(bb.abund ~ log(fl.abund.wt + 0.01) + (1 +fl.abund.wt | bb.sp),
+                    data = mismatch)
+summary(lm_mismatch)
+plot(lm_mismatch)
+
+ggplot(mismatch, aes(log(fl.abund.wt + 0.01), sqrt(bb.abund))) +
+  geom_point(aes(color = year)) +
   facet_wrap(~bb.sp, scales = "free")
 
-ggplot(pref_wt_abund, aes(log(fl.abund.weighted + 0.01), log(bb.abund), color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "lm") 
+mismatch_gam <- bam(bb.abund ~ bb.sp + year + s(log(fl.abund.wt + 0.01), by = bb.sp) + s(yday, k = 10),
+                    family = "quasipoisson",
+                    method = "fREML",
+                    discrete = TRUE,
+                    data = mismatch) %>% getViz()
+
+check.gamViz(mismatch_gam)
+summary(mismatch_gam)
+print(plot(mismatch_gam , allTerms = TRUE), pages = 4)
 ```
 
-### Plot fl.abund.weighted against phase and elev.class
+### Now let's model floral richness, but considering only species with preference > 0
 ```{r}
-ggplot(pref_wt_abund, aes(phase, log(fl.abund.weighted + k), color = year)) +
-  geom_boxplot() +
-  facet_wrap(~bb.sp)
+fl_rich_wt <- fl_abund_sp %>%
+  right_join(preference) %>% 
+  filter(preference > 0 & fl.abund > 0) %>%
+  mutate(yday = yday(date)) %>%
+  group_by(site, date, yday, bb.sp) %>%
+  summarize(fl.rich.wt = length(unique(plant.sp))) %>%
+  ungroup() %>%
+  left_join(bb_abund_sp) %>%
+  left_join(site_data, by = "site") %>%
+  semi_join(survey, by = c("site", "date")) %>% # This gets rid of site*date combos that don't exist
+  mutate(year = factor(year(date)), # add in the years missing from the bb_abund join
+         bb.sp = factor(bb.sp), # convert bb.sp to a factor
+         management = factor(management)) # convert management to a factor
 
-ggplot(pref_wt_abund, aes(elev.class2, log(fl.abund.weighted + k), color = year)) +
-  geom_boxplot() +
-  facet_wrap(~bb.sp)
+weighted_fl_rich <- bam(fl.rich.wt ~ bb.sp + year + te(yday, elev.mean, 
+                                                by = bb.sp,
+                                                bs= c("tp", "tp"), 
+                                                k=c(10, 10), m=1),
+                        data = fl_rich_wt,
+                        discrete = TRUE,
+                        family = "quasipoisson",
+                        method = "fREML") %>% getViz()
+
+check.gamViz(weighted_fl_rich)
+summary(weighted_fl_rich)
+print(plot(weighted_fl_rich , allTerms = TRUE), pages = 4)
 ```

### 2020-07-08 writing/manuscript_v3.Rmd
update

@@
 *Sampling*
+Because the frequency, duration, and seasonal interval of sampling varied across years, we model each year separately in all analyses.
+
 *Growing degree day analysis*
@@
 *Abundance modeling*
-- Floral preference modeling
+To represent floral abundance from the perspective of foraging bumble bees, we multiplied raw floral abundance by species-specific "preference" coefficients.
+The latter were calculated by regressing log-transformed mean proportional visitation against log-transformed mean proportional flower cover for all possible pairs of plant and bumble bee species, then extracting and exponentiating the residuals for each bee-plant pair (Supplemental Material).
+Using the R package mgcv, preference-weighted floral abundance was modeled using hierarchical generalized additive models consisting of a 2D tensor-product smooth of elevation and time (day-of-year), with bumble bee species as both a "by" smoothing factor and a random intercept term (Supplemental Material).
 
-Floral preference was estimated for all pairwise combinations of bumble bee and plant species.
-
-- Preference-weighted floral abundance
-- GAM
 *Network analysis*
@@
 **Discussion**
+
+While diversity (i.e.
+richness), of both bumble bees and flora, exhibits a more-or-less unimodal response to elevation, as predicted by previous studies, the abundances of bumble bees and flora are more complex.
+
+At around 1600 m, just below the tree line where the forest starts to thin, there seems to be a uniquely important elevation band.
+For many bumble bee species in our system, this is where preference-weighted floral resource abundance is highest.
+For a few species, it is the only peak in preference-weighted floral abundance.
+
+There also seems to be a consistent peak in preference-weighted floral abundance around 1000 m.

### 2020-07-08 analyses/BB_abund_GAM.Rmd
update

---
title: "preference"
output: html_document
---
### Prepare environment
```{r setup, include=FALSE, echo=FALSE}
library(mgcv)
library(mgcViz)
library(tidyverse)
library(lubridate)
```

### Climate data
```{r, include=FALSE, echo=FALSE}
gdd <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year)) %>%
  dplyr::select(site, year, date, gdd.cum) %>%
  semi_join(read_csv("../processed_data/network.csv"), by = "site")
```

### Load site data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../processed_data/site_data.csv") %>%
  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))
```

### Load network data
```{r}
### BB-FL visitation data
net <- read_csv("../processed_data/network.csv") %>%
  dplyr::select(site, date, bb.sp, plant.sp)
```

### FL and BB abundance / diversity
```{r echo=FALSE, include=FALSE}
### Per-transect BB abundance: species-level
bb_abund_sp <- net %>%
  dplyr::select(site, date, bb.sp) %>%
  group_by(site, date, bb.sp) %>% # group by bb.sp*transect (and other variables we want to retain)
  summarize(bb.abund = n()) %>% # get bb.sp*transect abundance
  ungroup %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  mutate(year = factor(year(date))) %>%
  left_join(site_data) %>%
  mutate(yday = yday(date),
         bb.sp = factor(bb.sp),
         management = factor(management)) %>%
  semi_join(net, by = c("site", "date")) %>%
  group_by(year, bb.sp) %>%
  mutate(bb.total.abund = sum(bb.abund))
```

GAM modeling of abundance per bb species
```{r}
### Call models
bb_abund.2010 <- bam(bb.abund ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
                                                            by = bb.sp,
                                                            bs= c("tp", "tp"), 
                                                            k=c(10, 10), m=1),
                        data = filter(bb_abund_sp, year == 2010 & bb.total.abund >= 5),
                        discrete = TRUE,
                        family = "poisson",
                        select = TRUE,
                        method = "fREML") %>% getViz()

check.gamViz(bb_abund.2010)
bb_abund.2010.sum <- summary(bb_abund.2010)
print(plot(bb_abund.2010, allTerms = TRUE), pages = 4)

bb_abund.2011 <- bam(bb.abund ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
                                                            by = bb.sp,
                                                            bs= c("tp", "tp"), 
                                                            k=c(10, 10), m=1),
                        data = filter(bb_abund_sp, year == 2011 & bb.total.abund >= 5),
                        discrete = TRUE,
                        family = "poisson",
                        select = TRUE,
                        method = "fREML") %>% getViz()

check.gamViz(bb_abund.2011)
bb_abund.2011.sum <- summary(bb_abund.2011)
print(plot(bb_abund.2011 , allTerms = TRUE), pages = 4)

bb_abund.2012 <- bam(bb.abund ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
                                                            by = bb.sp,
                                                            bs= c("tp", "tp"), 
                                                            k=c(10, 10), m=1),
                        data = filter(bb_abund_sp, year == 2012 & bb.total.abund >= 5),
                        discrete = TRUE,
                        family = "poisson",
                        select = TRUE,
                        method = "fREML") %>% getViz()

check.gamViz(bb_abund.2012)
bb_abund.2012.sum <- summary(bb_abund.2012)
print(plot(bb_abund.2012 , allTerms = TRUE), pages = 4)
plot(bb_abund.2012)

```

```{r}
bb_abund.2012.pred <- predict(bb_abund.2012)

ggplot(bb_abund.2012.pred, aes(yday, elev.mean, fit)) +
  geom_raster(aes(fill = exp(fit))) +
  #geom_contour(aes(z = fit), color = "black") +
  geom_point(data = filter(fl_abund, year == 2012), 
             aes(yday, elev.mean, size = fl.abund),
             color = "white", alpha = 0.25) +
  scale_fill_viridis_c()
```

### 2020-07-08 analyses/FL_abund_GAM.Rmd
update

--- 2020-07-08 analyses/preference.Rmd
@@
 ### Call models
-weighted_fl_abundance.2010.qp <- bam(fl.abund.wt ~ bb.sp + te(yday, elev.mean, 
+weighted_fl_abundance.2010.qp <- bam(fl.abund.wt ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
                                                             by = bb.sp,
@@
                         family = "quasipoisson",
+                        select = TRUE,
                         method = "fREML") %>% getViz()
@@
 
-weighted_fl_abundance.2010.qp2 <- bam(sqrt(fl.abund.wt) ~ bb.sp + te(yday, elev.mean, 
+weighted_fl_abundance.2010.qp2 <- bam(sqrt(fl.abund.wt) ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
                                                             by = bb.sp,
@@
 
-weighted_fl_abundance.2011.qp <- bam(fl.abund.wt ~ bb.sp + te(yday, elev.mean, 
+weighted_fl_abundance.2011.qp <- bam(fl.abund.wt ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
                                                             by = bb.sp,
@@
 
-weighted_fl_abundance.2011.qp2 <- bam(sqrt(fl.abund.wt) ~ bb.sp + te(yday, elev.mean, 
+weighted_fl_abundance.2011.qp2 <- bam(sqrt(fl.abund.wt) ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
                                                             by = bb.sp,
@@
 
-weighted_fl_abundance.2012.qp <- bam(fl.abund.wt ~ bb.sp + te(yday, elev.mean, 
+weighted_fl_abundance.2012.qp <- bam(fl.abund.wt ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
                                                             by = bb.sp,
@@
 print(plot(weighted_fl_abundance.2012.qp , allTerms = TRUE), pages = 4)
-
-weighted_fl_abundance.2012.qp2_re <- bam(sqrt(fl.abund.wt) ~ s(bb.sp, bs = "re") + 
-                                        management + 
-                                        te(yday, elev.mean, 
-                                           by = bb.sp,
-                                           bs= c("tp", "tp"), 
-                                           k=c(10, 10), m=1),
-                        data = filter(fl_abund_wt, year == 2012),
-                        discrete = TRUE,
-                        select = TRUE,
-                        family = "quasipoisson",
-                        method = "fREML") %>% getViz()
-
-check.gamViz(weighted_fl_abundance.2012.qp2_re)
-summary(weighted_fl_abundance.2012.qp2_re)
-print(plot(weighted_fl_abundance.2012.qp2_re, allTerms = TRUE), pages = 4)
-
-weighted_fl_abundance.2012.qp2_p <- bam(sqrt(fl.abund.wt) ~ bb.sp + management + 
-                                        te(yday, elev.mean, 
-                                           by = bb.sp,
-                                           bs= c("tp", "tp"), 
-                                           k=c(10, 10), m=1),
-                        data = filter(fl_abund_wt, year == 2012),
-                        discrete = TRUE,
-                        family = "quasipoisson",
-                        method = "fREML") %>% getViz()
-
-check.gamViz(weighted_fl_abundance.2012.qp2_p)
-summary(weighted_fl_abundance.2012.qp2_p)
-print(plot(weighted_fl_abundance.2012.qp2_p, allTerms = TRUE), pages = 4)
-
-weighted_fl_abundance.qp2 <- bam(sqrt(fl.abund.wt) ~ bb.sp + year + te(yday, elev.mean, 
-                                                            by = bb.sp,
-                                                            bs= c("tp", "tp"), 
-                                                            k=c(10, 10), m=1),
-                        data = fl_abund_wt,
-                        discrete = TRUE,
-                        family = "quasipoisson",
-                        method = "fREML") %>% getViz()
-
-check.gamViz(weighted_fl_abundance.qp2)
-summary(weighted_fl_abundance.qp2)
-print(plot(weighted_fl_abundance.qp2 , allTerms = TRUE), pages = 4)
 
@@
 
-v <- c(10,9, 5, 8,7)
-v.rank = rank(v)
-
-ggplot(mismatch, aes(fl.abund.wt.rank, bb.abund.rank)) +
+ggplot(mismatch, aes(fl.abund.wt, bb.abund)) +
   geom_point() +

### 2020-07-08 analyses/beta_diversity.Rmd
update

@@
   mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
-         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) %>% # turn elav.class into an ordered factor %>%
-  semi_join(read_csv("../processed_data/network.csv"), by = "site")  %>%
-  left_join(gdd_total, by = "site")
+         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) %>% # turn elav.class into an ordered factor
+  semi_join(read_csv("../processed_data/network.csv"), by = "site")
 
-### A quick little analysis
+# %>%
+#   left_join(gdd_total, by = "site")
+
+### Plot total GDD
 ggplot(site_data, aes(elev.mean, gdd.total, color = year)) +
@@
   ylab("Total GDD")
-
-# intercept only model
-gdd_year.gam1 <- gam(gdd.total ~ s(elev.mean) + year,
-                    data = site_data) %>% getViz()
-summary(gdd_year.gam1)
-check.gamViz(gdd_year.gam1)
-plot(gdd_year.gam1)
-
-# factor smooth model
-gdd_year.gam2 <- gam(gdd.total ~ s(elev.mean, year, bs = "fs", k = 5),
-                    data = site_data) %>% getViz()
-summary(gdd_year.gam2)
-check.gamViz(gdd_year.gam2)
-plot(gdd_year.gam2)
-
-# AIC
-AIC(gdd_year.gam1, gdd_year.gam2)
-
-site_pairwise_distance_t <- site_data %>%
-  ungroup() %>%
-  dplyr::select(site, lat, lon) %>%
-  unique() %>%
-  st_as_sf(coords = c("lon", "lat")) %>%
-  st_set_crs(4326) %>%
-  st_distance() %>%
-  data.frame()
-
-colnames(site_pairwise_distance_t) <- unique(site_data$site)
-rownames(site_pairwise_distance_t) <- unique(site_data$site)
-
-site_pairwise_distance <- site_pairwise_distance_t %>%
-  rownames_to_column("site1") %>%
-  pivot_longer(names_to = "site2", values_to = "distance", -site1)
 ```
@@
 ```{r}
-bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
-  dplyr::select(subgenus, bb.sp, pbl.w, pbl.w.class)
+# bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
+#   dplyr::select(subgenus, bb.sp, pbl.w, pbl.w.class)
+# 
+# fl_traits <- read.csv("../processed_data/floral_k_type_nom_fix.csv") %>%
+#   dplyr::select(plant.sp, k.type) %>%
+#   mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"),
+#          k.type.ss = str_extract(k.type, "[-+]?[0-9]*"))
 
-fl_traits <- read.csv("../processed_data/floral_k_type_nom_fix.csv") %>%
-  dplyr::select(plant.sp, k.type) %>%
-  mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"),
-         k.type.ss = str_extract(k.type, "[-+]?[0-9]*"))
-
-fl_tax <- read_csv("../processed_data/floral_tax.csv")
-
-breaks <- read_csv("./output/breaks.csv") %>%
-  mutate(year = factor(year))
+# fl_tax <- read_csv("../processed_data/floral_tax.csv") 
+# 
+# breaks <- read_csv("./output/breaks.csv") %>%
+#   mutate(year = factor(year))
 
@@
   mutate(year = factor(year)) %>% # convert year to factor
-  full_join(bb_traits, by = c("bb.sp" = "bb.sp")) %>%
+  #full_join(bb_traits, by = c("bb.sp" = "bb.sp")) %>%
   left_join(site_data, by = "site") %>%
-  left_join(gdd, by = c("site", "date")) %>%
-  left_join(breaks, by = c("site", "year")) %>%
-  left_join(fl_traits) %>%
-  left_join(fl_tax, by = "plant.genus") %>%
+  #left_join(gdd, by = c("site", "date")) %>%
+  #left_join(breaks, by = c("site", "year")) %>%
+  #left_join(fl_traits) %>%
+  #left_join(fl_tax, by = "plant.genus") %>%
   ungroup() %>%
-  mutate(phase = case_when(
-    gdd.cum < break1.gdd ~ "founding",
-    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
-    gdd.cum > break2.gdd ~ "reproductive"
-  )) %>%
-  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
+  #mutate(phase = case_when(
+    # gdd.cum < break1.gdd ~ "founding",
+    # gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
+    # gdd.cum > break2.gdd ~ "reproductive"
+  #)) %>%
+  #mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
   arrange(site, date)
+
+survey <- read_csv("../processed_data/floral_survey.csv") %>%
+  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
+  dplyr::select(site, date, plant.sp, flower.cover)
 
@@
   frame2webs() # convert to to bipartite web
-
-survey <- read_csv("../processed_data/floral_survey.csv") %>%
-  mutate(year = factor(year)) %>% # convert year to factor
-  semi_join(net, by = "plant.sp") %>% # Consider only plants visited at a given site-date
-  left_join(gdd, by = c("site", "date", "year")) %>%
-  left_join(breaks, by = c("site", "year")) %>%
-  left_join(site_data, by = "site") %>%
-  left_join(fl_traits) %>%
-  left_join(fl_tax) %>%
-  ungroup() %>%
-  mutate(phase = case_when(
-    gdd.cum < break1.gdd ~ "founding",
-    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
-    gdd.cum > break2.gdd ~ "reproductive"
-  )) %>%
-  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
-  arrange(site, date)
 ```
@@
                            man1, man2, man.diff, mean.elev, lon1, lat1, lon2, lat2)) %>%
-  left_join(site_pairwise_distance) %>%
+  #left_join(site_pairwise_distance) %>%
   mutate(metric.class = case_when(
@@
   ))
+
+write_csv(beta_site, "../analyses/output/beta_site.csv")
 
@@
 
-# And not much of a pattern in response to geographical distance
-ggplot(beta_site, 
-       aes(as.numeric(distance), value, color = metric)) +
-  geom_smooth()
+# # And not much of a pattern in response to geographical distance
+# ggplot(beta_site, 
+#        aes(as.numeric(distance), value, color = metric)) +
+#   geom_smooth()
 
@@
   group_by(site, plant.sp) %>%
-  summarize(cover = mean(flower.cover)) %>%
+  summarize(cover = sum(flower.cover)) %>%
   spread(plant.sp, cover) %>%
@@
   group_by(site, bb.sp) %>%
-  summarize(abund = mean(prop)) %>%
+  summarize(abund = sum(prop)) %>%
   spread(bb.sp, abund) %>%

### 2020-07-08 writing/manuscript_v3.Rmd
update

@@
 
+In mountain habitats, bumble bees (Hymenoptera: *Bombus* spp.) and their floral hosts provide an ideal study system in which to explore the effects of climate at both the basic level of species' distribution and abundance and the higher level of interaction network structure and dynamics.
+
 *[paragraph about Bombus being cold-adapted and generally distributed in high latitude or high elevation habitats]*
-Bumble bees (Hymenoptera: *Bombus* spp.) are cold-adapted species that likely originated in cooling climate of the Eocene-Oligocene boundary (34 mya) and now distributed mainly at higher latitudes and elevations [@Hines2008-er].
+Bumble bees (Hymenoptera: *Bombus* spp.) are cold-adapted species that likely originated in cooling climate of the Eocene-Oligocene boundary (34 mya) and are  now distributed mainly at higher latitudes and elevations [@Hines2008-er].
 
@@
 **Methods**
-*Study sites*
-*Sampling*
-Because the frequency, duration, and seasonal interval of sampling varied across years, we model each year separately in all analyses.
+*Study sites and sampling protocol*
 
-*Growing degree day analysis*
-*Diversity modeling*
-- GDM
-*Abundance modeling*
+*Diversity and abundance modeling*
 To represent floral abundance from the perspective of foraging bumble bees, we multiplied raw floral abundance by species-specific "preference" coefficients.
@@
 *Network analysis*
-- Bootstrapping
-- Network metrics
+- Interaction beta-diversity
+- Network metrics (with bootstrapping)
 
@@
 *Diversity: Turnover asymmetry between bumble bees and floral hosts*
-Beta diversity analysis
+Interaction beta diversity between pairs of sites was generally high, and
 
-*Abundance: Heterogeneous distribution of bumble bees and floral hosts over elevation x time surface*
+```{r}
+beta_site <- read_csv("../analyses/output/beta_site.csv") %>%
+  mutate(metric.class = factor(metric.class))
+
+ggplot(beta_site, 
+       aes(elev.diff, value, color = metric)) +
+  geom_point(alpha = 0.05) +
+  geom_smooth() +
+  facet_wrap(~metric.class)
+
+ggplot(filter(beta_site, metric.class == "aggregate" & metric != "S"), 
+       aes(elev.diff, value, color = metric)) +
+  geom_point(alpha = 0.05) +
+  geom_smooth(method = "lm")
+
+ggplot(filter(beta_site, metric.class == "ST.partition" & metric != "S"), 
+       aes(elev.diff, value, color = metric)) +
+  geom_point(alpha = 0.05) +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 2))
+```
+
+*Abundance and diversity: Heterogeneous distribution of bumble bees and floral hosts over elevation x time surface*
 - Contrast the broadly heterogeneous distribution of habitat generalist BBs to the restricted distribution of high-elevation specialists
@@
 For a few species, it is the only peak in preference-weighted floral abundance.
+"This 'range over-lap effect' is particularly pronounced in the zone of gradual or patch-wise transition from montane forests to alpine vegetation." [@Becker2007-cg]
 
-There also seems to be a consistent peak in preference-weighted floral abundance around 1000 m.
+There also seems to be a consistent peak in preference-weighted floral abundance around 1000 m. The slope breaks from the valley floor around 800-900 m, so 1000 m is low-mid slope.

### 2020-07-13 analyses/BB_abund_GAM.Rmd
update repo

@@
   complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
+  semi_join(net, by = c("site", "date")) %>%
   mutate(year = factor(year(date))) %>%
@@
          bb.sp = factor(bb.sp),
-         management = factor(management)) %>%
-  semi_join(net, by = c("site", "date")) %>%
+         management = factor(management),
+         site = factor(site)) %>%
   group_by(year, bb.sp) %>%
@@
 ```{r}
-### Call models
-bb_abund.2010 <- bam(bb.abund ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
+### Pooled years
+bb_abund.tw <- bam(bb.abund ~ s(bb.sp, bs = "re") + s(management, bs = "re") + year +
+                     s(site, bs = "re") +
+                     te(yday, elev.mean, 
+                        by = bb.sp,
+                        bs= c("tp", "tp"), 
+                        k=c(10, 10), m=1),
+                   data = bb_abund_sp,
+                   discrete = TRUE,
+                   family = "tw",
+                   select = TRUE,
+                   method = "fREML") %>% getViz()
+
+check.gamViz(bb_abund.tw)
+bb_abund.tw.sum <- summary(bb_abund.tw)
+print(plot(bb_abund.tw, allTerms = TRUE), pages = 1)
+
+### 2010
+bb_abund.2010.tw <- bam(bb.abund ~ s(bb.sp, bs = "re") + s(management, bs = "re") +
+                       s(site, bs = "re") +
+                       te(yday, elev.mean, 
                           by = bb.sp,
@@
                         discrete = TRUE,
-                        family = "poisson",
+                        family = "tw",
                         select = TRUE,
@@
 
-check.gamViz(bb_abund.2010)
-bb_abund.2010.sum <- summary(bb_abund.2010)
-print(plot(bb_abund.2010, allTerms = TRUE), pages = 4)
+check.gamViz(bb_abund.2010.tw)
+bb_abund.2010.sum.tw <- summary(bb_abund.2010.tw)
+print(plot(bb_abund.2010.tw, allTerms = TRUE), pages = 1)
 
-bb_abund.2011 <- bam(bb.abund ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
+### 2011
+bb_abund.2011.tw <- bam(bb.abund ~ s(bb.sp, bs = "re") + s(management, bs = "re") +
+                       s(site, bs = "re") +
+                       te(yday, elev.mean, 
                           by = bb.sp,
@@
 
-check.gamViz(bb_abund.2011)
-bb_abund.2011.sum <- summary(bb_abund.2011)
-print(plot(bb_abund.2011 , allTerms = TRUE), pages = 4)
+check.gamViz(bb_abund.2011.tw)
+bb_abund.2011.sum.tw <- summary(bb_abund.2011.tw)
+print(plot(bb_abund.2011.tw, allTerms = TRUE), pages = 1)
 
-bb_abund.2012 <- bam(bb.abund ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
+### 2012
+bb_abund.2012.tw <- bam(bb.abund ~ s(bb.sp, bs = "re") + s(management, bs = "re") +
+                       s(site, bs = "re") +
+                       te(yday, elev.mean, 
                           by = bb.sp,
@@
 
-check.gamViz(bb_abund.2012)
-bb_abund.2012.sum <- summary(bb_abund.2012)
-print(plot(bb_abund.2012 , allTerms = TRUE), pages = 4)
-plot(bb_abund.2012)
+check.gamViz(bb_abund.2012.tw)
+bb_abund.2012.sum.tw <- summary(bb_abund.2012.tw)
+print(plot(bb_abund.2012.tw, allTerms = TRUE), pages = 1)
 
 ```
-
-```{r}
-bb_abund.2012.pred <- predict(bb_abund.2012)
-
-ggplot(bb_abund.2012.pred, aes(yday, elev.mean, fit)) +
-  geom_raster(aes(fill = exp(fit))) +
-  #geom_contour(aes(z = fit), color = "black") +
-  geom_point(data = filter(fl_abund, year == 2012), 
-             aes(yday, elev.mean, size = fl.abund),
-             color = "white", alpha = 0.25) +
-  scale_fill_viridis_c()
-```

### 2020-07-13 analyses/FL_abund_GAM.Rmd
update repo

@@
   complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
-  mutate(year = factor(year(date)))
+  mutate(year = factor(year(date))) %>%
+  semi_join(net, by = c("site", "date")) # only include site-dates that actually occurred
 
@@
   complete(site, date, plant.sp, fill = list(fl.abund = 0, prop.fl.abund = 0)) %>%
+  semi_join(survey, by = c("site", "date")) %>%  # only include site-dates that actually occurred
   unique() %>%
@@
   unique() %>%
-  complete(site, date, plant.sp)
+  complete(site, date, plant.sp) %>%
+  semi_join(survey, by = c("site", "date"))
 
@@
   unique() %>%
-  complete(site, date, bb.sp)
+  complete(site, date, bb.sp) %>%
+  semi_join(net, by = c("site", "date"))
 
@@
   left_join(bb_abund_sp) %>%
-  left_join(bb_abund_sp_sum) %>%
-  filter(bb.total.abund >= 5) %>%
+  left_join(bb_abund_sp_sum)%>%
+  filter(bb.total.abund >= 5) %>% # It's not worth trying to model really rare species
+  semi_join(survey, by = c("site", "date")) %>% # This gets rid of site*date combos that don't exist
   left_join(site_data, by = "site") %>%
-  semi_join(survey, by = c("site", "date")) %>% # This gets rid of site*date combos that don't exist
   mutate(year = factor(year(date)), # add in the years missing from the bb_abund join
          bb.sp = factor(bb.sp), # convert bb.sp to a factor
-         management = factor(management)) # convert management to a factor
+         management = factor(management),
+         site = factor(site)) %>%
+  group_by(year) %>%
+  mutate(total.transects.year = n()) %>%
+  ungroup()
 
@@
 ```{r}
-### Set constants
-k <- min(
-  filter(fl_abund_wt, fl.abund.wt > 0)$fl.abund.wt
-)
+### Pooled years
+weighted_fl_abundance.tw <- bam(fl.abund.wt ~ s(bb.sp, bs = "re") + s(management, bs = "re") + year + 
+                                  s(site, bs = "re") +
+                                  te(yday, elev.mean, 
+                                     by = bb.sp,
+                                     bs= c("tp", "tp"), 
+                                     k=c(10, 10), m=1),
+                                data = fl_abund_wt,
+                                discrete = TRUE,
+                                family = "tw",
+                                select = TRUE,
+                                method = "fREML") %>% getViz()
 
-### Call models
-weighted_fl_abundance.2010.qp <- bam(fl.abund.wt ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
+check.gamViz(weighted_fl_abundance.tw)
+weighted_fl_abundance.tw.sum <- summary(weighted_fl_abundance.tw)
+print(plot(weighted_fl_abundance.tw, allTerms = TRUE), pages = 1)
+saveRDS(weighted_fl_abundance.tw, "../analyses/output/weighted_fl_abundance.rds")
+
+### 2010
+weighted_fl_abundance.2010.tw <- bam(fl.abund.wt ~ s(bb.sp, bs = "re") + s(management, bs = "re") + 
+                                       s(site, bs = "re") +
+                                       te(yday, elev.mean, 
                                           by = bb.sp,
@@
                                      discrete = TRUE,
-                        family = "quasipoisson",
+                                     family = "tw",
                                      select = TRUE,
@@
 
-check.gamViz(weighted_fl_abundance.2010.qp)
-summary(weighted_fl_abundance.2010.qp)
-print(plot(weighted_fl_abundance.2010.qp , allTerms = TRUE), pages = 4)
+check.gamViz(weighted_fl_abundance.2010.tw)
+weighted_fl_abundance.2010.tw.sum <- summary(weighted_fl_abundance.2010.tw)
+print(plot(weighted_fl_abundance.2010.tw , allTerms = TRUE), pages = 1)
 
-weighted_fl_abundance.2010.qp2 <- bam(sqrt(fl.abund.wt) ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
-                                                            by = bb.sp,
-                                                            bs= c("tp", "tp"), 
-                                                            k=c(10, 10), m=1),
-                        data = filter(fl_abund_wt, year == 2010),
-                        discrete = TRUE,
-                        family = "quasipoisson",
-                        select = TRUE,
-                        method = "fREML") %>% getViz()
-
-check.gamViz(weighted_fl_abundance.2010.qp2)
-summary(weighted_fl_abundance.2010.qp2)
-print(plot(weighted_fl_abundance.2010.qp2, allTerms = TRUE), pages = 3)
-
-weighted_fl_abundance.2011.qp <- bam(fl.abund.wt ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
+### 2011
+weighted_fl_abundance.2011.qp <- bam(fl.abund.wt ~ s(bb.sp, bs = "re") + s(management, bs = "re") + 
+                                       s(site, bs = "re") +
+                                       te(yday, elev.mean, 
                                           by = bb.sp,
@@
 check.gamViz(weighted_fl_abundance.2011.qp)
-summary(weighted_fl_abundance.2011.qp)
-print(plot(weighted_fl_abundance.2011.qp , allTerms = TRUE), pages = 4)
+weighted_fl_abundance.2011.qp.sum <- summary(weighted_fl_abundance.2011.qp)
+print(plot(weighted_fl_abundance.2011.qp , allTerms = TRUE), pages = 1)
 
-weighted_fl_abundance.2011.qp2 <- bam(sqrt(fl.abund.wt) ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
-                                                            by = bb.sp,
-                                                            bs= c("tp", "tp"), 
-                                                            k=c(10, 10), m=1),
-                        data = filter(fl_abund_wt, year == 2011),
-                        discrete = TRUE,
-                        family = "quasipoisson",
-                        select = TRUE,
-                        method = "fREML") %>% getViz()
-
-check.gamViz(weighted_fl_abundance.2011.qp2)
-summary(weighted_fl_abundance.2011.qp2)
-print(plot(weighted_fl_abundance.2011.qp2, allTerms = TRUE), pages = 4)
-
-weighted_fl_abundance.2012.qp <- bam(fl.abund.wt ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
+### 2012
+weighted_fl_abundance.2012.qp <- bam(fl.abund.wt ~ s(bb.sp, bs = "re") + s(management, bs = "re") + 
+                                       s(site, bs = "re") +
+                                       #slope.est + 
+                                       te(yday, elev.mean, 
                                           by = bb.sp,
@@
 check.gamViz(weighted_fl_abundance.2012.qp)
-summary(weighted_fl_abundance.2012.qp)
-print(plot(weighted_fl_abundance.2012.qp , allTerms = TRUE), pages = 4)
-
+weighted_fl_abundance.2012.qp.sum <- summary(weighted_fl_abundance.2012.qp)
+print(plot(weighted_fl_abundance.2012.qp , allTerms = TRUE), pages = 1)
 ```

### 2020-07-13 analyses/beta_diversity.Rmd
update repo

@@
   semi_join(read_csv("../processed_data/network.csv"), by = "site")
-
-gdd_total <- gdd %>%
-  group_by(site, year) %>%
-  summarize(gdd.total = max(gdd.cum))
 ```
@@
   semi_join(read_csv("../processed_data/network.csv"), by = "site")
-
-# %>%
-#   left_join(gdd_total, by = "site")
-
-### Plot total GDD
-ggplot(site_data, aes(elev.mean, gdd.total, color = year)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
-  ylab("Total GDD")
 ```
@@
 ```{r}
-# bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
-#   dplyr::select(subgenus, bb.sp, pbl.w, pbl.w.class)
-# 
-# fl_traits <- read.csv("../processed_data/floral_k_type_nom_fix.csv") %>%
-#   dplyr::select(plant.sp, k.type) %>%
-#   mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"),
-#          k.type.ss = str_extract(k.type, "[-+]?[0-9]*"))
-
-# fl_tax <- read_csv("../processed_data/floral_tax.csv") 
-# 
-# breaks <- read_csv("./output/breaks.csv") %>%
-#   mutate(year = factor(year))
-
 net <- read_csv("../processed_data/network.csv") %>%
@@
   mutate(year = factor(year)) %>% # convert year to factor
-  #full_join(bb_traits, by = c("bb.sp" = "bb.sp")) %>%
-  left_join(site_data, by = "site") %>%
-  #left_join(gdd, by = c("site", "date")) %>%
-  #left_join(breaks, by = c("site", "year")) %>%
-  #left_join(fl_traits) %>%
-  #left_join(fl_tax, by = "plant.genus") %>%
-  ungroup() %>%
-  #mutate(phase = case_when(
-    # gdd.cum < break1.gdd ~ "founding",
-    # gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
-    # gdd.cum > break2.gdd ~ "reproductive"
-  #)) %>%
-  #mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
+  group_by(site, date) %>%
+  #mutate(network.size = n()) %>%
   arrange(site, date)
@@
   dplyr::select(site, date, plant.sp, flower.cover)
+
+site.year_transects <- net %>%
+  group_by(site, year) %>%
+  summarize(transects = length(unique(date)))
+
+site_transects <- net %>%
+  group_by(site) %>%
+  summarize(transects = length(unique(date)))
 
@@
   rename(site1 = i) %>%
-  rename(site2 = j) %>%
+  rename(site2 = j)%>%
   left_join(site_data, by = c("site1" = "site")) %>%
@@
          elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
+  left_join(site_transects, by = c("site1" = "site")) %>%
+  dplyr::select(site1, site2, S, OS, WN, ST, 
+                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
+                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1 = transects) %>%
+  left_join(site_transects, by = c("site2" = "site")) %>%
+  dplyr::select(site1, site2, S, OS, WN, ST, 
+                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
+                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1, transects2 = transects) %>%
   mutate(elev.diff = abs(elev1 - elev2),
          mean.elev = (elev1 + elev2)/2,
+         transects.diff = abs(transects1 - transects2),
          man.diff = case_when(
@@
   gather(metric, value, -c(site1, site2, elev1, elev2, elev.diff, 
-                           man1, man2, man.diff, mean.elev, lon1, lat1, lon2, lat2)) %>%
-  #left_join(site_pairwise_distance) %>%
+                           man1, man2, man.diff, mean.elev, lon1, 
+                           lat1, lon2, lat2, transects1, transects2, transects.diff)) %>%
   mutate(metric.class = case_when(
@@
 
-write_csv(beta_site, "../analyses/output/beta_site.csv")
+#write_csv(beta_site, "../analyses/output/beta_site.csv")
 
-beta_site.year <- betalinkr_multi(array_site.year, partition.st = TRUE, partition.rr = TRUE) %>%
+beta_site.year <- betalinkr_multi(array_site.year, partition.st = TRUE, partition.rr = TRUE, binary = TRUE) %>%
   as_tibble() %>%
@@
          elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
+  left_join(site.year_transects, by = c("site1" = "site", "year1" = "year")) %>%
+  dplyr::select(site1, site2, year1, year2, S, OS, WN, ST, 
+                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
+                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1 = transects) %>%
+  left_join(site.year_transects, by = c("site2" = "site", "year2" = "year")) %>%
+  dplyr::select(site1, site2, year1, year2, S, OS, WN, ST, 
+                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
+                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1, transects2 = transects) %>%
   mutate(elev.diff = abs(elev1 - elev2),
          mean.elev = (elev1 + elev2)/2,
+         transects.diff = abs(transects1 - transects2),
          man.diff = case_when(
@@
   gather(metric, value, -c(site1, year1, site2, year2, elev1, elev2, elev.diff, 
-                           man1, man2, man.diff, mean.elev, lon1, lat1, lon2, lat2))
+                           man1, man2, man.diff, mean.elev, lon1, lat1, lon2, lat2,
+                           transects1, transects2, transects.diff)) %>%
+  mutate(metric.class = case_when(
+    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
+    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
+    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
+    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
+  ))
 ```
@@
 ```{r}
-# No apparent differences across years, so we will use the pooled data
-ggplot(filter(beta_site.year, year1 == year2), 
-       aes(metric, value, color = year1)) +
-  geom_boxplot() +
-  facet_wrap(~metric, scales = "free")
+# Interaction beta diversity by elevation difference (by year)
+## This shows that the patterns are extremely consistent across years, which is not surprising, since they are driven overwhelmingly by floral turnover, and plants tend to be in the same place from year to year.
+ggplot(filter(beta_site.year, year1 == year2 & 
+                metric != "S" & 
+                metric.class != "OS.partition"), 
+       aes(elev.diff, value, color = metric)) +
+  geom_point(alpha = 0.25) +
+  geom_smooth() +
+  facet_grid(metric.class ~ year1)
 
-# # And not much of a pattern in response to geographical distance
-# ggplot(beta_site, 
-#        aes(as.numeric(distance), value, color = metric)) +
-#   geom_smooth()
+ggplot(filter(beta_site.year, year1 == year2 & 
+                metric != "S" & 
+                metric.class != "OS.partition"), 
+       aes(transects.diff, value, color = metric)) +
+  geom_point(alpha = 0.25) +
+  geom_smooth() +
+  facet_grid(metric.class ~ year1, scales = "free_x")
 
-# Interaction beta diversity by elevation difference
-ggplot(beta_site, 
+# Interaction beta diversity by elevation difference, this time pooling across years
+ggplot(filter(beta_site, metric != "S" & 
+                metric.class != "OS.partition"), 
        aes(elev.diff, value, color = metric)) +
@@
   facet_wrap(~metric.class)
+
+# How important is the difference in transects? Not as important as the difference in elevation.
+ggplot(filter(beta_site, metric != "S" & 
+                metric.class != "OS.partition"), 
+       aes(transects.diff, value, color = metric)) +
+  geom_point(alpha = 0.05) +
+  geom_smooth() +
+  facet_wrap(~metric.class)
+
+ggplot(beta_site, 
+       aes(elev.diff, transects.diff)) +
+  geom_point(alpha = 0.05) +
+  geom_smooth()
+
+### Okay, we'll just do partial mantel tests to partial out the effect of transects.diff; and we'll only do that for the aggregate measures because it seems a bit silly to doi it for the partitions  
+WN_diff <- beta_site %>%
+  filter(metric == "WN") %>%
+  dplyr::select(site1, site2, value) %>%
+  spread(site2, value) %>%
+  dist()
+
+WN_diff <- beta_site %>%
+  filter(metric == "WN") %>%
+  dplyr::select(site1, site2, value) %>%
+  spread(site2, value) %>%
+  dist()
+
+ST_diff <- beta_site %>%
+  filter(metric == "ST") %>%
+  dplyr::select(site1, site2, value) %>%
+  spread(site2, value) %>%
+  dist()
+
+OS_diff <- beta_site %>%
+  filter(metric == "OS") %>%
+  dplyr::select(site1, site2, value) %>%
+  spread(site2, value) %>%
+  dist()
+
+elev_diff <- beta_site %>%
+  dplyr::select(site1, site2, elev.diff) %>%
+  unique() %>%
+  spread(site2, elev.diff) %>%
+  dist()
+
+transects_diff <- beta_site %>%
+  dplyr::select(site1, site2, transects.diff) %>%
+  unique() %>%
+  spread(site2, transects.diff) %>%
+  dist()
+
+WN_mantel <- mantel.partial(elev_diff, WN_diff, transects_diff)
+ST_mantel <- mantel.partial(elev_diff, ST_diff, transects_diff)
+OS_mantel <- mantel.partial(elev_diff, OS_diff, transects_diff)
 ```
@@
 ```{r}
+survey_transects <- survey %>%
+  group_by(site) %>%
+  summarize(transects = length(unique(date)))
+
 survey_bioData <- survey %>%
@@
   unique()
+%>%
+  left_join(survey_transects)
 

### 2020-07-13 analyses/network_analysis.Rmd
update repo

@@
 ### Define functions
-#### Rarefied network analysis
+#### Bootstrapped network analysis: *boot_net* function bootstraps networks by iteratively subsampling more frequently sampled networks to match the transect count of the least frequently sampled one
 ```{r, include=FALSE}
-# Rare_net function: rarefies networks by iteratively subsampling more frequently sampled networks to match the transect count of the least frequently sampled one
-# species level
-rare_net_sp_BB <- function(net, n, iter) {
+#####################################
+######## species level ##############
+#####################################
+boot_net_sp <- function(net, n, iter) {
 
@@
 
-rare_net_sp_FL <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      dplyr::select(site, date) %>%
-      unique() %>%
-      group_by(site) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      group_by(site) %>%
-      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
-      ungroup() %>%
-      filter(freq > 1) %>% # remove singletons
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(specieslevel, level = "lower", index = c("d", "species strength", "proportional generality")) %>%
-      map(rownames_to_column) %>%
-      map(as_tibble) %>%
-      bind_rows(.id = "site") %>% # Collapse list into big data frame
-      rename(sp = rowname) %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out %>%
-    gather(metric, value, -c(sp, site, iteration)) %>% # get mean value across iterations for each metric
-    group_by(sp, site, metric) %>%
-    summarize(value = mean(value)) %>%
-    left_join(site_data) %>%
-    left_join(fl_traits, by = c("sp" = "plant.sp"))
-}
-
-# network level
-rare_net_nt <- function(net, n, iter) {
+#####################################
+######## network level ##############
+#####################################
+boot_net <- function(net, n, iter) {
 
@@
       frame2webs() %>% # convert to to bipartite web
-      map(networklevel, weighted = TRUE, index = c("H2", "weighted NODF", "niche overlap", "C score", "generality", "web asymmetry", "weighted connectance")) %>% # We probably care mainly about the BB level of the network
-      data.frame() %>%
-      t() %>%
-      data.frame() %>%
-      rownames_to_column() %>%
-      dplyr::select(site = rowname, everything()) %>%
-      as_tibble() %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out  %>%
-      gather(metric, value, -c(site, iteration)) %>% # get mean value across iterations for each metric
-      group_by(site, metric) %>%
-      summarize(value = mean(value)) %>%
-      left_join(site_data)
-}
-
-```
-
-#### Rarefied network analysis by colony phase
-```{r, include=FALSE}
-# network level version 2
-rare_net_nt2 <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  t_filter <- net %>%
-    dplyr::select(site, date, phase, year) %>%
-    unique() %>%
-    group_by(site, phase, year) %>%
-    mutate(transects = n()) %>%
-    filter(transects >= n)
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      semi_join(t_filter) %>%
-      dplyr::select(site, date, phase, year) %>%
-      unique() %>%
-      group_by(site, phase, year) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, phase, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      #filter(freq > 1) %>%
-      unite(webID, c(site, phase, year), sep = "_") %>%
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(networklevel, weighted = TRUE, level = "higher",
-          index = c("web asymmetry", "weighted NODF", "H2", "C score", "niche overlap", "generality")) %>% 
-      data.frame() %>%
-      t() %>%
-      data.frame() %>%
-      rownames_to_column() %>%
-      dplyr::select(id = rowname, everything()) %>%
-      as_tibble() %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out  %>%
-      gather(metric, value, -c(id, iteration)) %>% # get mean value across iterations for each metric
-      group_by(id, metric) %>%
-      summarize(value = mean(value)) %>%
-      separate(id, sep = "_", into = c("site", "phase", "year")) %>%
-      mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
-      left_join(site_data)
-}
-# species level version 2
-rare_net_sp2 <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  t_filter <- net %>%
-    dplyr::select(site, date, phase, year) %>%
-    unique() %>%
-    group_by(site, phase, year) %>%
-    mutate(transects = n()) %>%
-    filter(transects >= n)
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      semi_join(t_filter) %>%
-      dplyr::select(site, date, phase, year) %>%
-      unique() %>%
-      group_by(site, phase, year) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, phase, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      #filter(freq > 1) %>%
-      unite(webID, c(site, phase, year), sep = "_") %>%
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(specieslevel, level = "higher", index = c("d", "species strength")) %>%
-      map(rownames_to_column) %>%
-      map(as_tibble) %>%
-      bind_rows(.id = "id") %>% # Collapse list into big data frame
-      rename(sp = rowname) %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out %>%
-    gather(metric, value, -c(sp, id, iteration)) %>% # get mean value across iterations for each metric
-    group_by(sp, id, metric) %>%
-    summarize(value = mean(value)) %>%
-    separate(id, sep = "_", into = c("site", "phase", "year")) %>%
-    mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
-    left_join(site_data) %>%
-    left_join(bb_traits, by = c("sp" = "bb.sp"))
-}
-```
-
-### Rarefied network analysis by floral k.type instead of floral species
-```{r}
-# Rare_net function: rarefies networks by iteratively subsampling more frequently sampled networks to match the transect count of the least frequently sampled one
-# species level
-rare_net_sp_k.type.s <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      dplyr::select(site, date) %>%
-      unique() %>%
-      group_by(site) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      na.omit() %>%
-      group_by(site) %>%
-      filter(length(unique(bb.sp)) > 1 & length(unique(k.type.s)) > 1) %>%
-      ungroup() %>%
-      filter(freq > 1) %>% # remove singletons
-      dplyr::select(higher = bb.sp, lower = k.type.s, webID = site, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(specieslevel, level = "higher", index = c("d", "species strength", "proportional generality")) %>%
-      map(rownames_to_column) %>%
-      map(as_tibble) %>%
-      bind_rows(.id = "site") %>% # Collapse list into big data frame
-      rename(sp = rowname) %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out %>%
-    gather(metric, value, -c(sp, site, iteration)) %>% # get mean value across iterations for each metric
-    group_by(sp, site, metric) %>%
-    summarize(value = mean(value)) %>%
-    left_join(site_data) %>%
-    left_join(bb_traits, by = c("sp" = "bb.sp"))
-}
-
-# network level
-rare_net_nt_k.type.s <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      dplyr::select(site, date) %>%
-      unique() %>%
-      group_by(site) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      na.omit() %>%
-      group_by(site) %>%
-      filter(length(unique(bb.sp)) > 1 & length(unique(k.type.s)) > 1) %>%
-      ungroup() %>%
-      dplyr::select(higher = bb.sp, lower = k.type.s, webID = site, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(networklevel, weighted = TRUE, level = "higher", index = c("H2", "weighted NODF", "niche overlap", "C score", "generality", "web asymmetry", "weighted connectance")) %>% # We probably care mainly about the BB level of the network
+      map(networklevel, level = "higher", weighted = TRUE, 
+          index = c("weighted NODF", "niche overlap", 
+                    "C score", "generality", "web asymmetry", 
+                    "weighted connectance")) %>% # We probably care mainly about the BB level of the network
       data.frame() %>%
@@
 ```{r, include=FALSE}
-bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
-  dplyr::select(subgenus, bb.sp, pbl.w, pbl.w.class)
-
-fl_traits <- read.csv("../processed_data/floral_k_type_nom_fix.csv") %>%
-  dplyr::select(plant.sp, k.type) %>%
-  mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"),
-         k.type.ss = str_extract(k.type, "[-+]?[0-9]*"))
-
-fl_tax <- read_csv("../processed_data/floral_tax.csv")
-
-breaks <- read_csv("./output/breaks.csv") %>%
-  mutate(year = factor(year))
-
 net <- read_csv("../processed_data/network.csv") %>%
   unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
-  unite(site.date, site, date, sep = "_", remove = FALSE) %>%
   mutate(year = factor(year)) %>% # convert year to factor
-  full_join(bb_traits, by = c("bb.sp" = "bb.sp")) %>%
   left_join(site_data, by = "site") %>%
   left_join(gdd, by = c("site", "date")) %>%
-  left_join(breaks, by = c("site", "year")) %>%
-  left_join(fl_traits) %>%
-  left_join(fl_tax, by = "plant.genus") %>%
   ungroup() %>%
-  mutate(phase = case_when(
-    gdd.cum < break1.gdd ~ "founding",
-    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
-    gdd.cum > break2.gdd ~ "reproductive"
-  )) %>%
-  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
   arrange(site, date)
 
-net_sum <- net %>%
-  group_by(site, year, elev.mean, date) %>%
-  summarize(total.obs = n()) %>%
-  ungroup() %>%
-  mutate(elev.fac = factor(elev.mean))
-
-ggplot(net_sum, aes(total.obs)) +
-  geom_histogram() +
-  facet_grid(elev.fac ~ year)
-
-ggplot(net_sum, aes(total.obs)) +
-  geom_histogram() +
-  facet_wrap(~year)
-
-survey <- read_csv("../processed_data/floral_survey.csv") %>%
-  mutate(year = factor(year)) %>% # convert year to factor
-  semi_join(net, by = "plant.sp") %>% # Consider only plants visited at a given site-date
-  left_join(gdd, by = c("site", "date")) %>%
-  left_join(breaks, by = c("site", "year")) %>%
-  left_join(site_data, by = "site") %>%
-  left_join(fl_traits) %>%
-  left_join(fl_tax) %>%
-  ungroup() %>%
-  mutate(phase = case_when(
-    gdd.cum < break1.gdd ~ "founding",
-    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
-    gdd.cum > break2.gdd ~ "reproductive"
-  )) %>%
-  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
-  arrange(site, date)
+# Yearly subsets
+net_2012 <- filter(net, year == "2012")
+net_2011 <- filter(net, year == "2011")
+net_2010 <- filter(net, year == "2010")
 ```
@@
   dplyr::select(site, year, date) %>%
-  unique()
-
-transects_per_siteyear <- transects %>%
+  unique() %>%
   group_by(site, year) %>%
@@
 
-min_transects_per_year <- transects_per_siteyear %>%
+min_transects_per_year <- transects %>%
   group_by(year) %>%
   summarize(min_transects = min(transects))
-
-### Create yearly subsets to feed to rare_net
-net_2012 <- filter(net, year == "2012")
-# net_2011 <- filter(net, year == "2011")
-# net_2010 <- filter(net, year == "2010")
-
-### Demonstrate that the distribution of interaction frequencies does not differ across elevation classes; in other words, no systematic bias is introduced to the network analysis due to uneven frequency of species with few observations
-net_freq <- net %>%
-  group_by(site, dayofyear, year, bb.sp, elev.mean, elev.class2) %>%
-  summarize(freq = n())
-
-ggplot(net_freq, aes(elev.class2, log(freq))) +
-  geom_boxplot() +
-  facet_wrap(~year)
 ```
 
-#### Generate network metric tables
+### Network- and group-level metric tables
+```{r}
+netmet_boot_2012 <- boot_net(net_2012, n = 9, iter = 40) %>%
+  mutate(year = factor(rep("2012", n())))
+netmet_boot_2011 <- boot_net(net_2011, n = 4, iter = 40)  %>%
+  mutate(year = factor(rep("2011", n())))
+netmet_boot_2010 <- boot_net(net_2010, n = 3, iter = 40) %>%
+  mutate(year = factor(rep("2010", n())))
+
+netmet <- bind_rows(netmet_boot_2012, 
+                    netmet_boot_2011, 
+                    netmet_boot_2010) %>%
+  ungroup() %>%
+  mutate(metric = factor(metric, levels = c("weighted.NODF",
+                                            "niche.overlap.HL", 
+                                            "C.score.HL",
+                                            "generality.HL",
+                                            "web.asymmetry",
+                                            "weighted.connectance"))) %>%
+  mutate(year = factor(year),
+         site = factor(site))
+```
+
+#### Species-level metric tables
 ```{r, message=FALSE}
@@
 #### BBs
-spmet_2012_BB <- rare_net_sp_BB(net_2012, 9, 40) %>%
+spmet_2012_BB <- boot_net_sp_BB(net_2012, 9, 40) %>%
   mutate(year = factor(rep("2012", n())))
-spmet_2011_BB <- rare_net_sp_BB(net_2011, 4, 40) %>%
+spmet_2011_BB <- boot_net_sp_BB(net_2011, 4, 40) %>%
   mutate(year = factor(rep("2011", n())))  
-spmet_2010_BB <- rare_net_sp_BB(net_2010, 3, 40) %>%
+spmet_2010_BB <- boot_net_sp_BB(net_2010, 3, 40) %>%
   mutate(year = factor(rep("2010", n())))
 
-spmet_BB <- rare_net_sp_BB(net, 5, 50) %>%
+spmet_BB <- boot_net_sp_BB(net, 5, 50) %>%
   ungroup() %>%
@@
 spmet_BB <- bind_rows(spmet_2012_BB, spmet_2011_BB, spmet_2010_BB)
-
-#### plants
-spmet_2012_FL <- rare_net_sp_FL(net_2012, 9, 40) %>%
-  mutate(year = factor(rep("2012", n())))
-spmet_2011_FL <- rare_net_sp_FL(net_2011, 4, 40) %>%
-  mutate(year = factor(rep("2011", n())))  
-spmet_2010_FL <- rare_net_sp_FL(net_2010, 3, 40) %>%
-  mutate(year = factor(rep("2010", n())))
-
-spmet_FL <- bind_rows(spmet_2012_FL, spmet_2011_FL, spmet_2010_FL)
-
-spmet_mean_FL <- spmet_FL %>%
-  group_by(site, year, elev.mean, metric) %>%
-  summarize(value = mean(value, na.rm = TRUE)) %>%
-  ungroup() %>%
-  mutate(year = factor(year))
-
-### Network- and group-level
-netmet_rare_2012 <- rare_net_nt(net_2012, n = 9, iter = 40) %>%
-  mutate(year = factor(rep("2012", n())))
-netmet_rare_2011 <- rare_net_nt(net_2011, n = 4, iter = 40)  %>%
-  mutate(year = factor(rep("2011", n())))
-netmet_rare_2010 <- rare_net_nt(net_2010, n = 3, iter = 40) %>%
-  mutate(year = factor(rep("2010", n())))
-
-netmet <- bind_rows(netmet_rare_2012, 
-                    netmet_rare_2011, 
-                    netmet_rare_2010) %>%
-  ungroup() %>%
-  mutate(metric = factor(metric, levels = c("H2",
-                                            "weighted.NODF",
-                                            "niche.overlap.HL", 
-                                            "C.score.HL",
-                                            "generality.HL",
-                                            "niche.overlap.LL", 
-                                            "C.score.LL",
-                                            "generality.LL",
-                                            "web.asymmetry",
-                                            "weighted.connectance"))) %>%
-  mutate(year = factor(year))
 ```
@@
 ```{r}
+#### Network level
+##### niche overlap
+ggplot(filter(netmet, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x)) +
+  #geom_smooth() +
+  xlab("Elevation") +
+  ylab("Niche overlap") +
+  facet_wrap(~year) +
+  theme_light(16)
+
+gam_niche.overlap <- gamV(value ~ year + s(site, bs = "re") +
+                            s(elev.mean, by = year, k = 10, bs = "tp", m = 2), 
+                         data = filter(netmet, metric == "niche.overlap.HL"),
+                         method = "REML")
+
+gam_niche.overlap2 <- gamV(value ~ year + s(site, bs = "re") +
+                            s(elev.mean, k = 10, bs = "tp", m = 2), 
+                         data = filter(netmet, metric == "niche.overlap.HL"),
+                         method = "REML")
+
+AIC(gam_niche.overlap, gam_niche.overlap2)
+
+check.gamViz(gam_niche.overlap)
+summary(gam_niche.overlap)
+print(plot(gam_niche.overlap, allTerms = TRUE), pages = 1)
+
+##### C.score
+ggplot(filter(netmet, metric == "C.score.HL"), aes(elev.mean, value)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x)) +
+  xlab("Elevation") +
+  ylab("C score") +
+  facet_wrap(~year) +
+  theme_light(16)
+
+gam_cscore <- gamV(value ~ year + s(site, bs = "re") + 
+                     s(elev.mean, bs = "tp", k = 10, m = 2), 
+                  data = filter(netmet, metric == "C.score.HL"),
+                  #family = "scat",
+                  method = "REML")
+
+check.gamViz(gam_cscore)
+summary(gam_cscore)
+print(plot(gam_cscore, allTerms = TRUE), pages = 1)
+
+#####################################################################################3
+
 #### Per species values
@@
 plot(lm(value ~ elev.mean, data =  filter(spmet_mean_BB, metric == "d" & year == 2012)))
-
-##### FL
-ggplot(filter(spmet_mean_FL, metric == "d"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
-  xlab("Elevation") +
-  ylab("Community mean d'") +
-  facet_wrap(~year) +
-  theme_light(16)
-
-#### Network level
-
-##### niche overlap
-ggplot(filter(netmet, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x)) +
-  #geom_smooth() +
-  xlab("Elevation") +
-  ylab("Niche overlap") +
-  facet_wrap(~year) +
-  theme_light(16)
-
-gam_niche.overlap <- gamV(value ~ s(year, bs = "re") + s(elev.mean, by = year, k = 10, bs = "tp", m = 2), 
-                         data = filter(netmet, metric == "niche.overlap.HL"),
-                         method = "REML")
-
-check.gamViz(gam_niche.overlap)
-summary(gam_niche.overlap)
-
-plot(gam_niche.overlap, allTerms = TRUE)
-
-##### C.score
-ggplot(filter(netmet, metric == "C.score.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x)) +
-  xlab("Elevation") +
-  ylab("C score") +
-  facet_wrap(~year) +
-  theme_light(16)
-
-gam_cscore <- gamV(value ~ s(year, bs = "re") + s(elev.mean, by = factor(year), bs = "tp", k = 6, m = 1), 
-                  data = filter(netmet, metric == "C.score.HL"),
-                  #family = "scat",
-                  method = "REML")
-
-check.gamViz(gam_cscore)
-summary(gam_cscore)
-
-plot(gam_cscore, allTerms = TRUE)
 ```
-
-### Rarefied network analysis by k.type instead of by plant sp
-```{r, message = FALSE}
-### Create yearly subsets to feed to rare_net
-net_2012 <- filter(net, year == "2012")
-net_2011 <- filter(net, year == "2011")
-net_2010 <- filter(net, year == "2010")
-
-### Species level
-spmet_2012_k <- rare_net_sp_k.type.s(net_2012, 9, 40) %>%
-  mutate(year = factor(rep("2012", n())))
-spmet_2011_k <- rare_net_sp_k.type.s(net_2011, 4, 40) %>%
-  mutate(year = factor(rep("2011", n())))  
-spmet_2010_k <- rare_net_sp_k.type.s(net_2010, 3, 40) %>%
-  mutate(year = factor(rep("2010", n())))
-
-spmet_k <- bind_rows(spmet_2012_k, spmet_2011_k, spmet_2010_k)
-
-spmet_mean_k <- spmet_k %>%
-  group_by(site, year, elev.mean, metric) %>%
-  summarize(value = mean(value))
-
-### Network- and group-level
-netmet_rare_2012_k <- rare_net_nt_k.type.s(net_2012, n = 9, iter = 40) %>%
-  mutate(year = factor(rep("2012", n())))
-netmet_rare_2011_k <- rare_net_nt_k.type.s(net_2011, n = 4, iter = 40)  %>%
-  mutate(year = factor(rep("2011", n())))
-netmet_rare_2010_k <- rare_net_nt_k.type.s(net_2010, n = 3, iter = 40) %>%
-  mutate(year = factor(rep("2010", n())))
-
-netmet_k <- bind_rows(netmet_rare_2012, 
-                    netmet_rare_2011, 
-                    netmet_rare_2010) %>%
-  mutate(metric = factor(metric, levels = c("H2",
-                                            "weighted.NODF",
-                                            "niche.overlap.HL", 
-                                            "C.score.HL",
-                                            "generality.HL",
-                                            "web.asymmetry",
-                                            "weighted.connectance")))
-```
-
-# Plot k.type network analysis
-```{r}
-#### Per species values
-ggplot(filter(spmet_2012_k,
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
-       aes(elev.mean, value)) +
-  geom_point() +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth() +
-  facet_grid(metric~sp, scales = "free")
-
-ggplot(filter(spmet_2011_k,
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
-       aes(elev.mean, value)) +
-  geom_point() +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth() +
-  facet_grid(metric~sp, scales = "free")
-
-ggplot(filter(spmet_2010_k,
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
-       aes(elev.mean, value)) +
-  geom_point() +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth() +
-  facet_grid(metric~sp, scales = "free")
-
-#### Community mean d
-ggplot(filter(spmet_mean_k, metric == "d"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
-  xlab("Elevation") +
-  ylab("Community mean d'") +
-  theme_light(16)
-
-ggplot(filter(spmet_mean_k, metric == "d"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
-  xlab("Elevation") +
-  ylab("Community mean d'") +
-  facet_wrap(~year) +
-  theme_light(16)
-
-gam_d <- gam(value ~ year + s(elev.mean, k = 10), 
-             data = filter(spmet_k, metric == "d"),
-             method = "REML")
-
-gam_d.vis <- getViz(gam_d)
-check.gamViz(gam_d.vis)
-summary(gam_d.vis)
-
-plot(gam_d.vis, allTerms = TRUE) +
-  xlab("Elevation")
-
-#### Network level
-ggplot(netmet_k, aes(elev.mean, value)) +
-  geom_point() +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth() +
-  facet_grid(metric~year, scales = "free")
-
-##### H2
-ggplot(filter(netmet_k, metric == "H2"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
-  xlab("Elevation") +
-  ylab("H2'") +
-  theme_light(16)
-
-gam_H2 <- gam(value ~ year + s(elev.mean), 
-              data = filter(netmet_k, metric == "H2"),
-              method = "REML")
-
-gam_H2.vis <- getViz(gam_H2)
-check.gamViz(gam_H2.vis)
-summary(gam_H2.vis)
-
-plot(gam_H2.vis, allTerms = TRUE) +
-  xlab("Elevation")
-
-##### niche overlap
-ggplot(filter(netmet_k, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
-  #geom_smooth() +
-  xlab("Elevation") +
-  ylab("Niche overlap") +
-  theme_light(16)
-
-ggplot(filter(netmet_k, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
-  #geom_smooth() +
-  xlab("Elevation") +
-  ylab("Niche overlap") +
-  facet_wrap(~year) +
-  theme_light(16)
-
-gam_niche.overlap <- gam(value ~ year + s(elev.mean), 
-                         data = filter(netmet_k, metric == "niche.overlap.HL"),
-                         method = "REML")
-
-gam_niche.overlap.vis <- getViz(gam_niche.overlap)
-check.gamViz(gam_niche.overlap.vis)
-summary(gam_niche.overlap.vis)
-
-plot(gam_niche.overlap.vis, allTerms = TRUE) +
-  xlab("Elevation")
-
-##### C.score
-ggplot(filter(netmet_k, metric == "C.score.HL"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
-  #geom_smooth() +
-  xlab("Elevation") +
-  ylab("C score") +
-  theme_light(16)
-
-ggplot(filter(netmet_k, metric == "C.score.HL"), aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
-  #geom_smooth() +
-  xlab("Elevation") +
-  ylab("C score") +
-  facet_wrap(~year) +
-  theme_light(16)
-
-gam_cscore <- gam(value ~ year + s(elev.mean), 
-                  data = filter(netmet_k, metric == "C.score.HL"),
-                  method = "REML")
-
-gam_cscore.vis <- getViz(gam_cscore)
-check.gamViz(gam_cscore.vis)
-summary(gam_cscore.vis)
-
-plot(gam_cscore.vis, allTerms = TRUE) +
-  xlab("Elevation")
-```
-
-### Do BB species differ in their k.type associations?
-```{r}
-web_by_site <- net %>%
-  group_by(site, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
-  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-  na.omit() %>%
-  group_by(site) %>%
-  filter(length(unique(bb.sp)) > 1 & length(unique(k.type.s)) > 1) %>%
-  ungroup() %>%
-  dplyr::select(higher = bb.sp, lower = k.type.s, webID = site, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs()
-
-web <- net %>%
-  group_by(year, elev.class2, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
-  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-  na.omit() %>%
-  unite(webID, c(year, elev.class2), sep = "_") %>%
-  dplyr::select(higher = bb.sp, lower = k.type.s, webID, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs()
-
-# 2010 high
-visweb(web[[1]])
-plotweb(web[[1]])
-
-# 2010 low
-visweb(web[[2]])
-plotweb(web[[2]])
-
-# 2010 med
-visweb(web[[3]])
-plotweb(web[[3]])
-
-# 2011 high
-visweb(web[[4]])
-plotweb(web[[4]])
-
-# 2011 low
-visweb(web[[5]])
-plotweb(web[[5]])
-
-# 2011 med
-visweb(web[[6]])
-plotweb(web[[6]])
-
-# 2012 high
-visweb(web[[7]])
-plotweb(web[[7]])
-
-# 2012 low
-visweb(web[[8]])
-plotweb(web[[8]])
-
-# 2012 med
-visweb(web[[9]])
-plotweb(web[[9]])
-
-```
-
-### Does tongue length class predict k.type association? Remember that B. wurflenii is primarily a nectar robber.
-```{r}
-kugler_key <- read_delim("../processed_data/kugler_key.tsv", delim = "\t") %>%
-  dplyr::select(1,2)
-
-ggplot(bb_traits, aes(pbl.w)) +
-  geom_histogram()
-
-net_cont <- net %>%
-  group_by(year, elev.class2, pbl.w.class, k.type.ss) %>%
-  summarize(freq = n()) %>%
-  group_by(year, elev.class2, pbl.w.class) %>%
-  mutate(prop = freq/sum(freq))
-
-net_cont2 <- net %>%
-  group_by(year, elev.class2, bb.sp, k.type.ss) %>%
-  summarize(freq = n()) %>%
-  group_by(year, elev.class2, bb.sp) %>%
-  mutate(prop = freq/sum(freq))
-
-ggplot(net_cont, aes(pbl.w.class, freq, fill = k.type.ss)) +
-  geom_col() +
-  facet_grid(elev.class2 ~ year)
-
-ggplot(net_cont, aes(pbl.w.class, prop, fill = k.type.ss)) +
-  geom_col() +
-  facet_grid(elev.class2 ~ year)
-
-ggplot(net_cont2, aes(bb.sp, prop, fill = k.type.ss)) +
-  geom_col() +
-  facet_grid(elev.class2 ~ year)
-
-chisq_mod <- chisq.test(net$pbl.w.class, net$k.type.ss)
-chisq_mod
-
-ggplot(net, aes(log(pbl.w), corolla.depth)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~year)
-
-ggplot(net, aes(dayofyear, corolla.depth)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~year)
-
-ggplot(net, aes(elev.mean, corolla.depth)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~year)
-
-ggplot(net, aes(factor(pbl.w.class), corolla.depth)) +
-  geom_boxplot() 
-```
-
-######## Old analyses that I probably don't care to use but don't want to delete yet #########
-
-### Per-transect network analysis -- is it worth it? I would have to filter out the transects for which the networks are too small.
-#### Network level
-```{r eval=FALSE, include=FALSE}
-per.transect_netmet <- net %>%
-  group_by(site, dayofyear, gdd.cum, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-  group_by(site, dayofyear, gdd.cum, year) %>%
-  mutate(size = sum(freq),
-         bb.rich = length(unique(bb.sp)),
-         fl.rich = length(unique(plant.sp)),
-         total.rich = bb.rich + fl.rich) %>%
-  filter(bb.rich > 3 & fl.rich > 3) %>%
-  unite(webID, c(site, dayofyear, gdd.cum, year, size), sep = "_") %>%
-  dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs() %>% # convert to to bipartite web
-  map(networklevel, weighted = TRUE, level = "higher", index = c("niche overlap", "C score", "H2")) %>% # We probably care mainly about the BB level of the network
-  data.frame() %>%
-  t() %>%
-  data.frame() %>%
-  rownames_to_column() %>%
-  dplyr::select(id = rowname, everything()) %>%
-  as_tibble() %>%
-  gather(metric, value, -id) %>%
-  separate(id, into = c("site", "yday", "gdd.cum", "year", "size"), sep = "_") %>%
-  mutate(size = as.numeric(size),
-         year = factor(year),
-         gdd.cum = as.numeric(gdd.cum),
-         yday = as.numeric(yday),
-         site = as.factor(site)) %>%
-  left_join(site_data) %>%
-  left_join(abund_div)
-
-### Can we explain network metrics with abundance and diversity data?
-# BB rich
-ggplot(per.transect_netmet, aes(bb.rich, value)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_grid(metric ~ year, scales = "free")
-# FL rich
-ggplot(per.transect_netmet, aes(fl.rich, value)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_grid(metric ~ year, scales = "free")
-# BB abund
-ggplot(per.transect_netmet, aes(log(bb.abund), value)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_grid(metric ~ year, scales = "free")
-# FL abund
-ggplot(per.transect_netmet, aes(log(fl.abund), value)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_grid(metric ~ year, scales = "free")
-# Per-capita FL abund
-ggplot(per.transect_netmet, aes(log(fl.per.bb), value)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_grid(metric ~ year, scales = "free")
-
-### Can we explain network metrics with elevation and/or time?
-ggplot(per.transect_netmet, aes(elev.mean, value, color = log(size))) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(metric ~ year, scales = "free")
-
-ggplot(per.transect_netmet, aes(log(size), value, color = log(size))) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(metric ~ year, scales = "free")
-
-ggplot(filter(per.transect_netmet, size > 1), 
-       aes(yday, value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(metric ~ year, scales = "free")
-
-ggplot(filter(per.transect_netmet, size > 10), 
-       aes(gdd.cum, value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(metric ~ year, scales = "free")
-
-ggplot(filter(per.transect_netmet, size > 10), 
-       aes(yday, elev.mean)) +
-  geom_point(aes(color = value)) +
-  geom_smooth() +
-  facet_grid(metric ~ year, scales = "free")
-```
-
-#### Species level
-```{r eval=FALSE, include=FALSE}
-per.transect_spmet <- net %>%
-  group_by(site, dayofyear, gdd.cum, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-  group_by(site, dayofyear, gdd.cum, year) %>%
-  mutate(size = sum(freq),
-         bb.rich = length(unique(bb.sp)),
-         fl.rich = length(unique(plant.sp)),
-         total.rich = bb.rich + fl.rich) %>%
-  filter(bb.rich > 5 & fl.rich > 5) %>%
-  unite(webID, c(site, dayofyear, gdd.cum, year, size), sep = "_") %>%
-  dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs() %>% # convert to to bipartite web
-  map(specieslevel, level = "higher", index = "d") %>% # We probably care mainly about the BB level of the network
-  map(rownames_to_column) %>%
-  map(as_tibble) %>%
-  bind_rows(.id = "id") %>% # Collapse list into big data frame
-  rename(sp = rowname) %>%
-  gather(metric, value, -c(sp, id)) %>% # get mean value across iterations for each metric
-  separate(id, sep = "_", into = c("site", "yday", "gdd.cum", "year", "size")) %>%
-  mutate(size = as.numeric(size),
-         year = factor(year),
-         gdd.cum = as.numeric(gdd.cum),
-         yday = as.numeric(yday)) %>%
-  left_join(site_data) %>%
-  left_join(bb_traits, by = c("sp" = "bb.sp")) %>%
-  left_join(abund_div)
-
-per.transect_spmet.mean <- per.transect_spmet %>%
-  group_by(site, yday, gdd.cum, year) %>%
-  summarize(value = mean(value)) %>%
-  left_join(site_data)
-
-ggplot(filter(per.transect_spmet), 
-       aes(log(fl.abund), value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(. ~ year, scales = "free")
-
-ggplot(filter(per.transect_spmet), 
-       aes(elev.mean, size)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(. ~ year, scales = "free")
-
-ggplot(filter(per.transect_spmet.mean), 
-       aes(yday, value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(. ~ year, scales = "free")
-
-ggplot(filter(per.transect_spmet), 
-       aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(. ~ year, scales = "free")
-
-ggplot(filter(per.transect_spmet, size >= 10), 
-       aes(gdd.cum, value, color = log(size))) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(metric ~ year, scales = "free")
-
-ggplot(filter(per.transect_spmet, size >= 10), 
-       aes(yday, value, color = log(size))) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(metric ~ year, scales = "free")
-```
-
-### Phase x elev network analysis -- this only makes sense in 2012, where we have sufficient sampling across the whole colony lifecycle
-```{r eval=FALSE, include=FALSE}
-### Prepare pxe abundance and diversity table
-pxe_abund_div <- abund_div %>%
-  dplyr::select(-gdd.cum, -yday, -week, -month, -date, -elev.class2) %>%
-  gather(ad.metric, ad.value, -c(site, year, phase, elev.mean)) %>%
-  group_by(site, phase, year, ad.metric) %>%
-  summarize(ad.value = mean(ad.value, na.rm = TRUE))
-
-pxe_net <- rare_net_nt2(filter(net, year == 2012), 3, 20) %>%
-  full_join((pxe_abund_div))
-
-ggplot(filter(pxe_net, metric == "H2"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(year ~ phase, scales = "free") +
-  xlab("Elevation") +
-  ylab("H2")
-
-ggplot(filter(pxe_net, metric == "C.score.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(year ~ phase, scales = "free") +
-  xlab("Elevation") +
-  ylab("C.score")
-
-ggplot(filter(pxe_net, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(year ~ phase, scales = "free") +
-  xlab("Elevation") +
-  ylab("Niche overlap")
-
-pxe_sp <- rare_net_sp2(net, 4, 20)  %>%
-  full_join((pxe_abund_div))
-
-ggplot(filter(pxe_sp, 
-              metric == "d" &
-                sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit") &
-                ad.metric == "fl.rich"), 
-       aes(ad.value, value, color = year)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +
-  facet_grid(sp ~ phase, scales = "free") +
-  ylim(c(0,1)) +
-  xlab("Floral richness") +
-  ylab("d'")
-
-pxe_sp_mean <- pxe_sp %>%
-  filter(sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")) %>%
-  group_by(site, year, phase, elev.mean, metric) %>%
-  summarize(value = mean(value, na.rm = TRUE))
-
-ggplot(filter(pxe_sp_mean, metric == "d"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(year ~ phase, scales = "free") +
-  xlab("Elevation") +
-  ylab("Mean d'")
-
-ggplot(filter(pxe_sp_mean, metric == "species.strength"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(year ~ phase, scales = "free") +
-  xlab("Elevation") +
-  ylab("Mean species strength'")
-
-ggplot(filter(pxe_sp, 
-              metric == "species.strength",
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
-       aes(elev.mean, value, color = year)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(sp ~ phase, scales = "free") +
-  xlab("Elevation") +
-  ylab("Species strength")
-
-ggplot(filter(pxe_sp, 
-              metric == "d",
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
-       aes(elev.mean, value, color = year)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(sp ~ phase, scales = "free") +
-  ylim(c(0,1)) +
-  xlab("Elevation") +
-  ylab("d'")
-
-ggplot(filter(pxe_sp, 
-              metric == "d",
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
-       aes(elev.mean, value)) +
-  geom_point(aes(color = year)) +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(sp ~ phase, scales = "free") +
-  ylim(c(0,1)) +
-  xlab("Elevation") +
-  ylab("d'")
-
-ggplot(filter(pxe_sp, 
-              metric == "d",
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
-       aes(elev.mean, value, color = phase)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
-  facet_grid(year ~ sp, scales = "free") +
-  theme_light() +
-  xlab("Elevation") +
-  ylab("d'")
-```
-
-```{r}
-
-# network level PxE
-rare_net_nt_pxe <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  t_filter <- net %>%
-    dplyr::select(site, date, phase, year) %>%
-    unique() %>%
-    group_by(site, phase, year) %>%
-    mutate(transects = n()) %>%
-    filter(transects >= n)
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      semi_join(t_filter) %>%
-      dplyr::select(site, date, phase, year) %>%
-      unique() %>%
-      group_by(site, phase, year) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, phase, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      group_by(site, phase, year) %>%
-      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
-      ungroup() %>%
-      unite(webID, c(site, phase, year), sep = "_") %>%
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(networklevel, weighted = TRUE, level = "higher",
-          index = c("web asymmetry", "weighted NODF", "H2", "C score", "niche overlap", "generality")) %>% 
-      data.frame() %>%
-      t() %>%
-      data.frame() %>%
-      rownames_to_column() %>%
-      dplyr::select(id = rowname, everything()) %>%
-      as_tibble() %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out  %>%
-      gather(metric, value, -c(id, iteration)) %>% # get mean value across iterations for each metric
-      group_by(id, metric) %>%
-      summarize(value = mean(value)) %>%
-      separate(id, sep = "_", into = c("site", "phase", "year")) %>%
-      mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
-      left_join(site_data)
-}
-
-# species level version 2
-rare_net_sp2 <- function(net, n, iter) {
-
-  out <- data.frame()
-
-  t_filter <- net %>%
-    dplyr::select(site, date, phase, year) %>%
-    unique() %>%
-    group_by(site, phase, year) %>%
-    mutate(transects = n()) %>%
-    filter(transects >= n)
-
-  for(i in 1:iter) {
-    temp <- net %>%
-      semi_join(t_filter) %>%
-      dplyr::select(site, date, phase, year) %>%
-      unique() %>%
-      group_by(site, phase, year) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, phase, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      group_by(site, phase, year) %>%
-      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
-      ungroup() %>%
-      unite(webID, c(site, phase, year), sep = "_") %>%
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(specieslevel, level = "higher", index = c("d", "species strength")) %>%
-      map(rownames_to_column) %>%
-      map(as_tibble) %>%
-      bind_rows(.id = "id") %>% # Collapse list into big data frame
-      rename(sp = rowname) %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out %>%
-    gather(metric, value, -c(sp, id, iteration)) %>% # get mean value across iterations for each metric
-    group_by(sp, id, metric) %>%
-    summarize(value = mean(value)) %>%
-    separate(id, sep = "_", into = c("site", "phase", "year")) %>%
-    mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
-    left_join(site_data) %>%
-    left_join(bb_traits, by = c("sp" = "bb.sp"))
-}
-
-```

### 2020-07-13 analyses/trash_can.R

@@ -1,0 +1,983 @@
+
+boot_net_sp_FL <- function(net, n, iter) {
+  
+  out <- data.frame()
+  
+  for(i in 1:iter) {
+    temp <- net %>%
+      dplyr::select(site, date) %>%
+      unique() %>%
+      group_by(site) %>%
+      sample_n(n) %>%
+      left_join(net) %>%
+      group_by(site, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
+      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
+      group_by(site) %>%
+      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
+      ungroup() %>%
+      filter(freq > 1) %>% # remove singletons
+      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
+      data.frame() %>% # convert to data frame
+      frame2webs() %>% # convert to to bipartite web
+      map(specieslevel, level = "lower", index = c("d", "species strength", "proportional generality")) %>%
+      map(rownames_to_column) %>%
+      map(as_tibble) %>%
+      bind_rows(.id = "site") %>% # Collapse list into big data frame
+      rename(sp = rowname) %>%
+      mutate(iteration = rep(i, n()))
+    
+    out <- bind_rows(out, temp)
+  } 
+  out <- out %>%
+    gather(metric, value, -c(sp, site, iteration)) %>% # get mean value across iterations for each metric
+    group_by(sp, site, metric) %>%
+    summarize(value = mean(value)) %>%
+    left_join(site_data) %>%
+    left_join(fl_traits, by = c("sp" = "plant.sp"))
+}
+
+
+#### Rarefied network analysis by colony phase
+```{r, include=FALSE}
+# network level version 2
+boot_net_nt2 <- function(net, n, iter) {
+  
+  out <- data.frame()
+  
+  t_filter <- net %>%
+    dplyr::select(site, date, phase, year) %>%
+    unique() %>%
+    group_by(site, phase, year) %>%
+    mutate(transects = n()) %>%
+    filter(transects >= n)
+  
+  for(i in 1:iter) {
+    temp <- net %>%
+      semi_join(t_filter) %>%
+      dplyr::select(site, date, phase, year) %>%
+      unique() %>%
+      group_by(site, phase, year) %>%
+      sample_n(n) %>%
[…]

### 2020-07-13 analyses/beta_diversity.Rmd
update

@@
 ```{r}
-net <- read_csv("../processed_data/network.csv") %>%
+net <- read_csv("../processed_data/network.csv")%>%
   unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
-  unite(site.date, site, date, sep = "_", remove = FALSE) %>%
+  unite(site.year.day, site, year, dayofyear, sep = "_", remove = FALSE) %>%
   mutate(year = factor(year)) %>% # convert year to factor
   group_by(site, date) %>%
-  #mutate(network.size = n()) %>%
   arrange(site, date)
@@
   frame2webs() # convert to to bipartite web
+
+web_site.date.2010 <- net %>%
+  filter(year == 2010) %>%
+  group_by(site.year.day, bb.sp, plant.sp) %>%
+  summarize(freq = n()) %>%
+  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year.day, freq) %>% # rename columns to match bipartite's expectations
+  data.frame() %>% # convert to data frame
+  frame2webs() # convert to to bipartite web
+
+web_site.date.2011 <- net %>%
+  filter(year == 2011) %>%
+  group_by(site.year.day, bb.sp, plant.sp) %>%
+  summarize(freq = n()) %>%
+  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year.day, freq) %>% # rename columns to match bipartite's expectations
+  data.frame() %>% # convert to data frame
+  frame2webs() # convert to to bipartite web
+
+web_site.date.2012 <- net %>%
+  filter(year == 2012) %>%
+  group_by(site.year.day, bb.sp, plant.sp) %>%
+  summarize(freq = n()) %>%
+  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year.day, freq) %>% # rename columns to match bipartite's expectations
+  data.frame() %>% # convert to data frame
+  frame2webs() # convert to to bipartite web
+
+array_site <- webs2array(web_site)
+array_site.year <- webs2array(web_site.year)
+array_site.date.2010 <- webs2array(web_site.date.2010)
+array_site.date.2011 <- webs2array(web_site.date.2011)
+array_site.date.2012 <- webs2array(web_site.date.2012)
 ```
@@
 ```{r}
-array_site <- webs2array(web_site)
-array_site.year <- webs2array(web_site.year)
-
 beta_site <- betalinkr_multi(array_site, partition.st = TRUE, partition.rr = TRUE) %>%
@@
   ))
+
+### By individual transect
+
+beta_site.date.2010 <- betalinkr_multi(array_site.date.2010, 
+                                       partition.st = TRUE, 
+                                       partition.rr = TRUE) %>%
+  as_tibble() %>%
+  separate(i, into = c("site1", "date1"), sep = "_") %>%
+  separate(j, into = c("site2", "date2"), sep = "_") %>%
+  left_join(site_data, by = c("site1" = "site"))%>%
+  dplyr::select(site1, date1, site2, date2, S, OS, WN, ST, 
+         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
+         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
+  left_join(site_data, by = c("site2" = "site")) %>%
+  dplyr::select(site1, date1, site2, date2, S, OS, WN, ST, 
+         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
+         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
+  mutate(elev.diff = abs(elev1 - elev2),
+         mean.elev = (elev1 + elev2)/2,
+         man.diff = case_when(
+           man1 == man2 ~ FALSE,
+           man1 != man2 ~ TRUE
+         )) %>%
+  gather(metric, value, -c(site1, date1, site2, date2, elev1, elev2, elev.diff, 
+                           man1, man2, man.diff, mean.elev, lon1, lat1, lon2, lat2)) %>%
+  mutate(metric.class = case_when(
+    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
+    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
+    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
+    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
+  ))
+
+beta_site.date.2011 <- betalinkr_multi(array_site.date.2011, 
+                                       partition.st = TRUE, 
+                                       partition.rr = TRUE)
+
+beta_site.date.2012 <- betalinkr_multi(array_site.date.2012, 
+                                       partition.st = TRUE, 
+                                       partition.rr = TRUE)
 ```

### 2020-07-13 analyses/network_analysis.Rmd
update

@@
     separate(site_year, c("site", "year"), sep = "_") %>%
-    left_join(site_data) %>%
-    left_join(bb_traits, by = c("sp" = "bb.sp"))
+    left_join(site_data) 
+  # %>%
+  #   left_join(bb_traits, by = c("sp" = "bb.sp"))
 }
@@
 net_2010 <- filter(net, year == "2010")
+
+### Per-transect BB abundance: species-level
+bb_abund_sp <- net %>%
+  dplyr::select(site, date, bb.sp) %>%
+  group_by(site, date, bb.sp) %>% # group by bb.sp*transect (and other variables we want to retain)
+  summarize(bb.abund = n()) %>% # get bb.sp*transect abundance
+  ungroup %>%
+  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
+  semi_join(net, by = c("site", "date")) %>%
+  mutate(year = factor(year(date))) %>%
+  mutate(yday = yday(date),
+         sp = factor(bb.sp),
+         site = factor(site)) %>%
+  group_by(site, year, sp) %>%
+  summarize(bb.total.abund = sum(bb.abund))
 ```
@@
 #### BBs
-spmet_2012_BB <- boot_net_sp_BB(net_2012, 9, 40) %>%
+spmet_2012 <- boot_net_sp(net_2012, 9, 40) %>%
   mutate(year = factor(rep("2012", n())))
-spmet_2011_BB <- boot_net_sp_BB(net_2011, 4, 40) %>%
+spmet_2011 <- boot_net_sp(net_2011, 4, 40) %>%
   mutate(year = factor(rep("2011", n())))  
-spmet_2010_BB <- boot_net_sp_BB(net_2010, 3, 40) %>%
+spmet_2010 <- boot_net_sp(net_2010, 3, 40) %>%
   mutate(year = factor(rep("2010", n())))
 
-spmet_BB <- boot_net_sp_BB(net, 5, 50) %>%
-  ungroup() %>%
-  mutate(year = factor(year))
+spmet <- bind_rows(spmet_2012, spmet_2011, spmet_2010) %>%
+  left_join(bb_abund_sp) %>%
+  mutate(sp = factor(sp),
+         management = factor(management),
+         site = factor(site))
 
-spmet_mean_BB <- spmet_BB %>%
-  group_by(site, year, elev.mean, metric) %>%
-  summarize(value = mean(value, na.rm = TRUE)) %>%
-  ungroup() %>%
-  mutate(year = factor(year))
+ggplot(filter(spmet, bb.total.abund >= 20 & sp %in% c("bss", "pasc", "prat", "wurf", "soro")), aes(elev.mean, value, color = year)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = TRUE) +
+  facet_wrap(~sp, scales = "free")
 
-spmet_mean_BB_2012 <- spmet_2012_BB %>%
-  group_by(site, elev.mean, metric) %>%
-  summarize(value = mean(value, na.rm = TRUE))
-
-ggplot(spmet_mean_BB, aes(elev.mean, value)) +
+ggplot(filter(spmet, bb.total.abund >= 20), aes(elev.mean, value)) +
   geom_point() +
-  geom_smooth() +
+  geom_smooth(method = "lm", se = TRUE) +
   facet_wrap(~year)
 
-ggplot(spmet_2012_BB, aes(elev.mean, value)) +
+specialization_GAM.2010 <- bam(value ~ 
+                            s(elev.mean, by = sp) +
+                            #s(bb.total.abund, by = sp) +
+                            s(site, bs = "re") +
+                            s(sp, bs = "re"),
+                   data = filter(spmet, year == 2010),
+                   discrete = TRUE,
+                   select = TRUE,
+                   method = "fREML") %>% getViz()
+
+check.gamViz(specialization_GAM.2010)
+specialization_GAM.2010.sum <- summary(specialization_GAM.2010)
+print(plot(specialization_GAM.2010, allTerms = TRUE), pages = 1)
+
+spmet_mean <- spmet %>%
+  group_by(site, year, elev.mean, metric) %>%
+  summarize(value = weighted.mean(value, bb.total.abund)) %>%
+  ungroup()
+
+ggplot(spmet_mean, aes(elev.mean, value)) +
   geom_point() +
-  geom_smooth() +
+  geom_smooth(method = "lm") +
   facet_wrap(~year)
 
-ggplot(spmet_mean_BB_2012, aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth()
-
-spmet_BB_gam <- bam(log(value + 0.01) ~ year + 
-                      s(elev.mean, bs = "tp", k = 10, m = 2) +
-                      s(elev.mean, by = factor(sp), k = 20, m = 1),
-                    data = spmet_BB,
-                    method = "fREML",
-                    discrete = TRUE) %>% getViz()
-
-check.gamViz(spmet_BB_gam)
-summary(spmet_BB_gam)
-print(plot(spmet_BB_gam, allTerms = TRUE), pages = 1)
-
-spmet_BB <- bind_rows(spmet_2012_BB, spmet_2011_BB, spmet_2010_BB)
 ```

### 2020-07-16 analyses/FL_abund_GAM.Rmd
preparing to try SEM

@@
 
+write_csv(preference, "../analyses/output/preference.csv")
 #################################
@@
 ```
+
+Does preference relate to BB-FL traits
+```{r}
+bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
+  dplyr::select(subgenus, bb.sp, pbl.w, pbl.w.class)
+
+fl_ktype <- read.csv("../processed_data/floral_k_type_nom_fix.csv") %>%
+  dplyr::select(plant.sp, k.type) %>%
+  mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"),
+         k.type.ss = str_extract(k.type, "[-+]?[0-9]*"))
+
+fl_morph <- read.csv("../processed_data/flower_morphology_GitaBenadi.csv") %>%
+  rename(plant.sp = Species) %>%
+  group_by(plant.sp) %>%
+  summarize(corolla.depth = mean(LN))
+
+fl_tax <- read_csv("../processed_data/floral_tax.csv")
+
+pref_traits <- preference %>%
+  left_join(bb_traits) %>%
+  left_join(fl_ktype) %>%
+  left_join(fl_morph) %>%
+  separate(plant.sp, c("plant.genus", "plant.species"), sep = " ", remove = FALSE) %>%
+  left_join(fl_tax) %>%
+  mutate(preference.bin = case_when(
+    preference > 0 ~ 1,
+    preference == 0 ~ 0
+  )) %>%
+  mutate(tongue.mismatch = abs(pbl.w - corolla.depth),
+         bb.sp = factor(bb.sp),
+         plant.sp = factor(plant.sp),
+         plant.genus = factor(plant.genus),
+         subgenus = factor(subgenus),
+         k.type.s = factor(k.type.s),
+         plant.family = factor(plant.family))
+
+ggplot(pref_traits, aes(k.type.s, preference)) +
+  geom_boxplot()
+
+ggplot(filter(pref_traits, preference > 0), aes(k.type.s, log(preference))) +
+  geom_boxplot()
+
+ggplot(pref_traits, aes(tongue.mismatch, preference)) + 
+  geom_point() +
+  geom_smooth()
+
+ggplot(pref_traits, aes(abs(pbl.w - corolla.depth), preference)) + 
+  geom_point() +
+  geom_smooth(method = "lm")
+
+### To deal with zero inflation, binarize and do binomial regression
+binomial_smooth <- function(...) {
+  geom_smooth(method = "glm", method.args = list(family = "binomial"), ...)
+}
+
+ggplot(pref_traits, aes(tongue.mismatch, preference.bin)) + 
+  geom_point() +
+  binomial_smooth()
+
+### Then, to focus only on non-zero preference...
+ggplot(filter(pref_traits, preference > 0), aes(tongue.mismatch, log(preference))) + 
+  geom_point() +
+  geom_smooth(method = "lm")
+
+### Model
+library(DHARMa)
+lmer_nonzero <- lmer(log(preference) ~ tongue.mismatch + 
+                       k.type.s +
+                       (1 | plant.sp) +
+                       (1 | bb.sp), 
+                     data = filter(pref_traits, preference > 0 & !is.na(corolla.depth)))
+
+summary(lmer_nonzero)
+plot(lmer_nonzero)
+simulationOutput <- simulateResiduals(fittedModel = lmer_nonzero, plot = T)
+testResiduals(simulationOutput)
+plot(simulationOutput)
+
+lmer_withzeros <- glmer(preference ~ tongue.mismatch + 
+                          k.type.s +
+                          (1 | plant.sp) +
+                          (1 | bb.sp), 
+                       family = "poisson",
+                       data = filter(pref_traits, !is.na(corolla.depth)))
+
+summary(lmer_withzeros)
+plot(lmer_withzeros)
+simulationOutput_withzeros <- simulateResiduals(fittedModel = lmer_withzeros, plot = T)
+testResiduals(simulationOutput_withzeros)
+plot(simulationOutput_withzeros)
+
+```

### 2020-07-16 analyses/FL_abund_GAM_sp.Rmd
preparing to try SEM

---
title: "FL_abund_GAM_sp"
output: html_document
---

### Prepare environment
```{r setup, include=FALSE, echo=FALSE}
library(mgcv)
library(mgcViz)
library(tidyverse)
library(lubridate)
```

### Climate data
```{r, include=FALSE, echo=FALSE}
gdd <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year)) %>%
  dplyr::select(site, year, date, gdd.cum) %>%
  semi_join(read_csv("../processed_data/network.csv"), by = "site")
```

### Load site data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../processed_data/site_data.csv") %>%
  #dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))
```

### Floral survey data
```{r}
survey <- read_csv("../processed_data/floral_survey.csv") %>%
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  group_by(site, date, plant.sp) %>%
  summarize(flower.cover = sum(flower.cover)) %>%
  mutate(yday = yday(date),
       year = year(date)) %>%
  left_join(site_data) %>%
  ungroup() %>%
  mutate(site = factor(site),
         plant.sp = factor(plant.sp),
         year = factor(year),
         management = factor(management))

preference <- read_csv("../analyses/output/preference.csv")

top_plants <- net %>%
  mutate(year = year(date)) %>%
  group_by(bb.sp, year, plant.sp) %>%
  summarize(visits = n()) %>%
  mutate(prop.visits = visits/sum(visits)) %>%
  group_by(bb.sp, year) %>%
  arrange(bb.sp, year, -prop.visits) %>%
  top_n(3)

top_3 <- read_csv("../analyses/output/visitation_by_abundance_mean.csv") %>%
  group_by(bb.sp) %>%
  arrange(-prop.visits.wm) %>%
  top_n(3)

```

```{r}
ggplot(filter(survey, plant.sp == "Trifolium pratense"), aes(yday, elev.mean, size = flower.cover)) +
  geom_point(alpha = 0.25) +
  facet_wrap(~year, scales = "free")

ggplot(filter(survey, plant.sp == "Erica carnea"), aes(yday, elev.mean, size = flower.cover)) +
  geom_point(alpha = 0.25) +
  facet_wrap(~year, scales = "free")

ggplot(filter(survey, plant.sp == "Stachys alopecuros"), aes(yday, elev.mean, size = flower.cover)) +
  geom_point(alpha = 0.25) +
  facet_wrap(~year, scales = "free")

t.pratense_GAM <- bam(flower.cover ~ s(management, bs = "re") + year + 
                                  s(site, bs = "re") +
                                  te(yday, elev.mean, 
                                     by = year,
                                     bs= c("tp", "tp"), 
                                     k=c(10, 10), m=1),
                                data = filter(survey, plant.sp == "Trifolium pratense"),
                                discrete = TRUE,
                                family = "tw",
                                select = TRUE,
                                method = "fREML") %>% getViz()

check.gamViz(t.pratense_GAM)
t.pratense_GAM.sum <- summary(t.pratense_GAM)
print(plot(t.pratense_GAM, allTerms = TRUE), pages = 1)

e.carnea_GAM <- bam(flower.cover ~ s(management, bs = "re") + year + 
                                  s(site, bs = "re") +
                                  te(yday, elev.mean, 
                                     by = year,
                                     bs= c("tp", "tp"), 
                                     k=c(3, 3), m=1),
                                data = filter(survey, plant.sp == "Erica carnea"),
                                discrete = TRUE,
                                family = "tw",
                                select = TRUE,
                                method = "fREML") %>% getViz()

check.gamViz(e.carnea_GAM)
e.carnea_GAM.sum <- summary(e.carnea_GAM)
print(plot(e.carnea_GAM, allTerms = TRUE), pages = 1)
```

### 2020-07-16 analyses/aggregate_network_analysis.Rmd
preparing to try SEM

---
title: "aggregate_network_analysis"
output: html_document
---

### Prepare environment
```{r setup, include=FALSE}
library(bipartite)
library(tidyverse)
library(lubridate)
```

### Load network and survey data
```{r, include=FALSE}
net <- read_csv("../processed_data/network.csv") %>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  mutate(year = factor(year)) %>% # convert year to factor
  ungroup()

web <- net %>%
  group_by(bb.sp, plant.sp, year) %>%
  summarize(freq = n()) %>%
  select(higher = bb.sp, lower = plant.sp, freq, webID = year) %>%
  as.data.frame() %>%
  frame2webs()

web2010 <- net %>%
  filter(year == 2010) %>%
  group_by(bb.sp, plant.sp, year) %>%
  summarize(freq = n()) %>%
  select(higher = bb.sp, lower = plant.sp, freq, webID = year) %>%
  as.data.frame() %>%
  frame2webs()

network_metric <- web %>%
  map(networklevel, level = "higher", weighted = TRUE) %>% 
  data.frame() %>%
  t() %>%
  data.frame() %>%
  rownames_to_column() %>%
  dplyr::select(site_year = rowname, everything()) %>%
  as_tibble()

modules2010 <- web2010 %>%
  map(computeModules)

web2010 <- web[[1]]
web2011 <- web[[2]]
web2012 <- web[[3]]

plotweb(web2010)
plotweb(web2011)
plotweb(web2012)

visweb(web2010)
visweb(web2011)
visweb(web2012)
```

### 2020-07-16 analyses/beta_diversity.Rmd
preparing to try SEM

@@
   as_tibble() %>%
-  separate(i, into = c("site1", "date1"), sep = "_") %>%
-  separate(j, into = c("site2", "date2"), sep = "_") %>%
+  separate(i, into = c("site1", "year1", "yday1"), sep = "_") %>%
+  separate(j, into = c("site2", "year2", "yday2"), sep = "_") %>%
   left_join(site_data, by = c("site1" = "site"))%>%
-  dplyr::select(site1, date1, site2, date2, S, OS, WN, ST, 
+  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
          ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
@@
   left_join(site_data, by = c("site2" = "site")) %>%
-  dplyr::select(site1, date1, site2, date2, S, OS, WN, ST, 
+  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
          ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
          elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
-  mutate(elev.diff = abs(elev1 - elev2),
+  mutate(site1 = factor(site1),
+         site2 = factor(site2),
+         year1 = factor(year1),
+         year2 = factor(year2),
+         yday1 = as.numeric(yday1),
+         yday2 = as.numeric(yday2),
+         elev.diff = abs(elev1 - elev2),
          mean.elev = (elev1 + elev2)/2,
+         yday.diff = abs(yday1 - yday2),
          man.diff = case_when(
@@
          )) %>%
-  gather(metric, value, -c(site1, date1, site2, date2, elev1, elev2, elev.diff, 
-                           man1, man2, man.diff, mean.elev, lon1, lat1, lon2, lat2)) %>%
+  gather(metric, value, -c(site1, year1, yday1, site2, year2, yday2,
+                           elev1, elev2, elev.diff, yday.diff, 
+                           man1, man2, man.diff, mean.elev, 
+                           lon1, lat1, lon2, lat2)) %>%
   mutate(metric.class = case_when(
@@
                                        partition.st = TRUE, 
-                                       partition.rr = TRUE)
+                                       partition.rr = TRUE) %>%
+  as_tibble() %>%
+  separate(i, into = c("site1", "year1", "yday1"), sep = "_") %>%
+  separate(j, into = c("site2", "year2", "yday2"), sep = "_") %>%
+  left_join(site_data, by = c("site1" = "site"))%>%
+  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
+         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
+         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
+  left_join(site_data, by = c("site2" = "site")) %>%
+  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
+         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
+         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
+  mutate(site1 = factor(site1),
+         site2 = factor(site2),
+         year1 = factor(year1),
+         year2 = factor(year2),
+         yday1 = as.numeric(yday1),
+         yday2 = as.numeric(yday2),
+         elev.diff = abs(elev1 - elev2),
+         mean.elev = (elev1 + elev2)/2,
+         yday.diff = abs(yday1 - yday2),
+         man.diff = case_when(
+           man1 == man2 ~ FALSE,
+           man1 != man2 ~ TRUE
+         )) %>%
+  gather(metric, value, -c(site1, year1, yday1, site2, year2, yday2,
+                           elev1, elev2, elev.diff, yday.diff, 
+                           man1, man2, man.diff, mean.elev, 
+                           lon1, lat1, lon2, lat2)) %>%
+  mutate(metric.class = case_when(
+    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
+    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
+    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
+    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
+  ))
 
@@
                                        partition.st = TRUE, 
-                                       partition.rr = TRUE)
+                                       partition.rr = TRUE) %>%
+  as_tibble() %>%
+  separate(i, into = c("site1", "year1", "yday1"), sep = "_") %>%
+  separate(j, into = c("site2", "year2", "yday2"), sep = "_") %>%
+  left_join(site_data, by = c("site1" = "site"))%>%
+  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
+         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
+         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
+  left_join(site_data, by = c("site2" = "site")) %>%
+  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
+         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
+         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
+  mutate(site1 = factor(site1),
+         site2 = factor(site2),
+         year1 = factor(year1),
+         year2 = factor(year2),
+         yday1 = as.numeric(yday1),
+         yday2 = as.numeric(yday2),
+         elev.diff = abs(elev1 - elev2),
+         mean.elev = (elev1 + elev2)/2,
+         yday.diff = abs(yday1 - yday2),
+         man.diff = case_when(
+           man1 == man2 ~ FALSE,
+           man1 != man2 ~ TRUE
+         )) %>%
+  gather(metric, value, -c(site1, year1, yday1, site2, year2, yday2,
+                           elev1, elev2, elev.diff, yday.diff, 
+                           man1, man2, man.diff, mean.elev, 
+                           lon1, lat1, lon2, lat2)) %>%
+  mutate(metric.class = case_when(
+    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
+    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
+    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
+    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
+  ))
+
+beta_site.date <- bind_rows(beta_site.date.2010, beta_site.date.2011, beta_site.date.2012)
+beta_site.date_gdm.prep <- beta_site.date %>%
+  filter(metric == "WN")
+
+write_csv(beta_site.date, "../analyses/output/beta_site.date.csv")
+```
+
+### Visualize transect-level beta-diversity
+```{r}
+ggplot(filter(beta_site.date.2010, metric != "S" & 
+                metric.class != "OS.partition"), 
+       aes(elev.diff, value, color = metric)) +
+  geom_point(alpha = 0.05) +
+  geom_smooth() +
+  facet_wrap(~metric.class)
+
+ggplot(filter(beta_site.date.2010, yday.diff <= 7 &
+                metric != "S" & 
+                metric.class != "OS.partition"), 
+       aes(elev.diff, value, color = metric)) +
+  geom_point(alpha = 0.05) +
+  geom_smooth() +
+  facet_wrap(~metric.class)
+
+ggplot(filter(beta_site.date.2011, yday.diff <= 7 &
+                metric != "S" & 
+                metric.class != "OS.partition"), 
+       aes(elev.diff, value, color = metric)) +
+  geom_point(alpha = 0.05) +
+  geom_smooth() +
+  facet_wrap(~metric.class)
+
+ggplot(filter(beta_site.date.2012, yday.diff <= 7 &
+                metric != "S" & 
+                metric.class != "OS.partition"), 
+       aes(elev.diff, value, color = metric)) +
+  geom_point(alpha = 0.05) +
+  geom_smooth() +
+  facet_wrap(~metric.class)
+
+ggplot(filter(beta_site.date.2012, yday.diff == 0 &
+                metric != "S" & 
+                metric.class != "OS.partition"),
+       aes(value)) +
+  geom_histogram()  +
+  facet_wrap(~metric)
+
+ggplot(filter(beta_site.date.2012, yday.diff <= 10 &
+                metric != "S" & 
+                metric.class != "OS.partition"),
+       aes(value)) +
+  geom_histogram()  +
+  facet_wrap(~metric)
+
+ggplot(filter(beta_site.date.2012, yday.diff <= 7 &
+                metric != "S" & 
+                metric.class != "OS.partition" &
+                site1 == site2),
+       aes(value)) +
+  geom_histogram()  +
+  facet_wrap(~metric)
+
 ```
@@
 ```{r}
-survey_transects <- survey %>%
-  group_by(site) %>%
-  summarize(transects = length(unique(date)))
-
-survey_bioData <- survey %>%
-  filter(flower.cover > 0) %>%
+### 2012
+survey_bioData.2012 <- survey %>%
+  mutate(year = year(date),
+         yday = yday(date)) %>%
+  filter(flower.cover > 0 & year == "2012") %>%
   mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
-  group_by(site, plant.sp) %>%
+  group_by(site, yday, plant.sp) %>%
   summarize(cover = sum(flower.cover)) %>%
   spread(plant.sp, cover) %>%
-  replace(is.na(.), 0)
+  replace(is.na(.), 0) %>%
+  arrange(site, yday) %>%
+  unite(site.day, site, yday, sep = "_")
 
-predData <- site_data %>%
+predData.2012 <- survey %>%
+  mutate(year = year(date),
+         yday = yday(date)) %>%
+  filter(flower.cover > 0 & year == "2012") %>%
+  select(site, yday) %>%
+  left_join(site_data) %>%
   ungroup() %>%
-  dplyr::select(site, elev.mean, lat, lon) %>%
-  unique()
-%>%
-  left_join(survey_transects)
+  dplyr::select(site, elev.mean, lat, lon, yday) %>%
+  unique() %>%
+  arrange(site, yday) %>%
+  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
+  select(-site)
 
-survey_sitepair <- formatsitepair(survey_bioData, bioFormat = 1, siteColumn = "site", dist = "jaccard",
-                              XColumn = "lon", YColumn = "lat", predData = predData)
+survey_sitepair.2012 <- formatsitepair(survey_bioData.2012, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
+                              XColumn = "lon", YColumn = "lat", predData = predData.2012)
 
 ### Call gdm()
-gdm_survey <- gdm(survey_sitepair, geo = TRUE)
-gdm.varImp(survey_sitepair, geo = TRUE)
+gdm_survey.2012 <- gdm(survey_sitepair.2012, geo = TRUE)
+#gdm.varImp(survey_sitepair, geo = TRUE)
 
-summary(gdm_survey)
-plot(gdm_survey)
-plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
+summary(gdm_survey.2012)
+plot(gdm_survey.2012)
+#plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
+
+### 2011
+survey_bioData.2011 <- survey %>%
+  mutate(year = year(date),
+         yday = yday(date)) %>%
+  filter(flower.cover > 0 & year == "2011") %>%
+  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
+  group_by(site, yday, plant.sp) %>%
+  summarize(cover = sum(flower.cover)) %>%
+  spread(plant.sp, cover) %>%
+  replace(is.na(.), 0) %>%
+  arrange(site, yday) %>%
+  unite(site.day, site, yday, sep = "_")
+
+predData.2011 <- survey %>%
+  mutate(year = year(date),
+         yday = yday(date)) %>%
+  filter(flower.cover > 0 & year == "2011") %>%
+  select(site, yday) %>%
+  left_join(site_data) %>%
+  ungroup() %>%
+  dplyr::select(site, elev.mean, lat, lon, yday) %>%
+  unique() %>%
+  arrange(site, yday) %>%
+  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
+  select(-site)
+
+survey_sitepair.2011 <- formatsitepair(survey_bioData.2011, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
+                              XColumn = "lon", YColumn = "lat", predData = predData.2011)
+
+### Call gdm()
+gdm_survey.2011 <- gdm(survey_sitepair.2011, geo = TRUE)
+#gdm.varImp(survey_sitepair, geo = TRUE)
+
+summary(gdm_survey.2011)
+plot(gdm_survey.2011)
+#plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
+
+### 2010
+survey_bioData.2010 <- survey %>%
+  mutate(year = year(date),
+         yday = yday(date)) %>%
+  filter(flower.cover > 0 & year == "2010") %>%
+  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
+  group_by(site, yday, plant.sp) %>%
+  summarize(cover = sum(flower.cover)) %>%
+  spread(plant.sp, cover) %>%
+  replace(is.na(.), 0) %>%
+  arrange(site, yday) %>%
+  unite(site.day, site, yday, sep = "_")
+
+predData.2010 <- survey %>%
+  mutate(year = year(date),
+         yday = yday(date)) %>%
+  filter(flower.cover > 0 & year == "2010") %>%
+  select(site, yday) %>%
+  left_join(site_data) %>%
+  ungroup() %>%
+  dplyr::select(site, elev.mean, lat, lon, yday) %>%
+  unique() %>%
+  arrange(site, yday) %>%
+  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
+  select(-site)
+
+survey_sitepair.2010 <- formatsitepair(survey_bioData.2010, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
+                              XColumn = "lon", YColumn = "lat", predData = predData.2010)
+
+### Call gdm()
+gdm_survey.2010 <- gdm(survey_sitepair.2010, geo = TRUE)
+#gdm.varImp(survey_sitepair, geo = TRUE)
+
+summary(gdm_survey.2010)
+plot(gdm_survey.2010)
+#plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
 ```
@@
 ```
+
+### At the transect level
+```{r}
+beta_site.date_gdm.prep
+
+transects_2010 <- beta_site.date.2010 %>%
+  filter(metric == "WN") %>%
+  mutate(weights = rep(1, n())) %>%
+  dplyr::select(distance = value, weights,
+         s1.xCoord = lon1, s1.yCoord = lat1, 
+         s2.xCoord = lon2, s2.yCoord = lat2, 
+         s1.elev = elev1, s2.elev = elev2,
+         s1.yday = yday1, s2.yday = yday2) %>%
+  data.frame()
+
+### Call gdm()
+gdm_net <- gdm(transects_2010, geo = TRUE)
+gdm.varImp(transects_2010, geo = TRUE)
+
+summary(gdm_net)
+plot(gdm_net)
+plotUncertainty(net_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
+```

### 2020-07-16 analyses/network_analysis.Rmd
preparing to try SEM

--- 2020-07-13 analyses/network_analysis.Rmd
@@
       unite(webID, c(site, year), sep = "_") %>%
-      #group_by(site_year) %>%
-      #filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
-      #ungroup() %>%
-      #filter(freq > 1) %>% # remove singletons
+      ungroup() %>%
       dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
@@
       bind_rows(.id = "site_year") %>% # Collapse list into big data frame
-      rename(sp = rowname) %>%
+      rename(bb.sp = rowname) %>%
       mutate(iteration = rep(i, n()))
@@
   out <- out %>%
-    gather(metric, value, -c(sp, site_year, iteration)) %>% # get mean value across iterations for each metric
-    group_by(sp, site_year, metric) %>%
+    gather(metric, value, -c(bb.sp, site_year, iteration)) %>% # get mean value across iterations for each metric
+    group_by(bb.sp, site_year, metric) %>%
     summarize(value = mean(value)) %>%
-    separate(site_year, c("site", "year"), sep = "_") %>%
-    left_join(site_data) %>%
-    left_join(bb_traits, by = c("sp" = "bb.sp"))
+    separate(site_year, c("site", "year"), sep = "_") 
 }
@@
 
+  t_filter <- net %>%
+    dplyr::select(site, date, year) %>%
+    unique() %>%
+    group_by(site, year) %>%
+    mutate(transects = n()) %>%
+    filter(transects >= n)
+
   for(i in 1:iter) {
     temp <- net %>%
-      dplyr::select(site, date) %>%
+      semi_join(t_filter) %>%
+      dplyr::select(site, date, year) %>%
       unique() %>%
-      group_by(site) %>%
+      group_by(site, year) %>%
       sample_n(n) %>%
       left_join(net) %>%
-      group_by(site, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
+      group_by(site, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
       summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      group_by(site) %>%
-      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
+      group_by(site, year) %>%
+      unite(webID, c(site, year), sep = "_") %>%
+      #filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
       ungroup() %>%
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
+      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
+      data.frame() %>% # convert to data frame
+      frame2webs() %>% # convert to to bipartite web
+      map(networklevel, level = "higher", weighted = TRUE) %>% # We probably care mainly about the BB level of the network
+      data.frame() %>%
+      t() %>%
+      data.frame() %>%
+      rownames_to_column() %>%
+      dplyr::select(site_year = rowname, everything()) %>%
+      as_tibble() %>%
+      mutate(iteration = rep(i, n()))
+
+    out <- bind_rows(out, temp)
+  } 
+  out <- out  %>%
+      gather(metric, value, -c(site_year, iteration)) %>% # get mean value across iterations for each metric
+      group_by(site_year, metric) %>%
+      summarize(value = mean(value)) %>%
+      separate(site_year, c("site", "year"), sep = "_") 
+}
+
+##################################################
+######## network level by site-date ##############
+##################################################
+site_date_net <- function(net, min_higher, min_lower) {
+
+  size_filter <- net %>%
+    group_by(site, date) %>%
+    summarize(higher.species = length(unique(bb.sp)),
+              lower.species = length(unique(plant.sp))) %>%
+    filter(higher.species >= min_higher & lower.species >= min_lower)
+
+  out <- net %>%
+    semi_join(size_filter, by = c("site", "date")) %>%
+    group_by(site, date, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
+    summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
+    unite(webID, c(site, date), sep = "_") %>%
+    ungroup() %>%
+    dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
     data.frame() %>% # convert to data frame
@@
     rownames_to_column() %>%
-      dplyr::select(site = rowname, everything()) %>%
+    dplyr::select(site_date = rowname, everything()) %>%
     as_tibble() %>%
-      mutate(iteration = rep(i, n()))
-
-    out <- bind_rows(out, temp)
-  } 
-  out <- out  %>%
-      gather(metric, value, -c(site, iteration)) %>% # get mean value across iterations for each metric
-      group_by(site, metric) %>%
-      summarize(value = mean(value)) %>%
-      left_join(site_data)
+    separate(site_date, c("site", "date"), sep = "_")
 }
@@
 net_2010 <- filter(net, year == "2010")
+
+# Floral survey
+survey <- read_csv("../processed_data/floral_survey.csv") %>%
+  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
+  dplyr::select(site, date, plant.sp, flower.cover)
 ```
@@
 
-### Network- and group-level metric tables
-```{r}
+### Tabulate metrics
+```{r message=TRUE, warning=TRUE, include=FALSE}
+### Auxiliary data
+site_BB_aux <- net %>%
+  group_by(site, year, date) %>%
+  summarize(bb.rich = length(unique(bb.sp)),
+            bb.abund = n()) %>%
+  group_by(site, year) %>%
+  summarize(bb.rich = mean(bb.rich),
+            log.bb.abund = log(mean(bb.abund)),
+            bb.transects = length(unique(date)))
+
+site_BB_aux.sp <- net %>%
+  group_by(site, date, bb.sp) %>% # group by bb.sp*transect (and other variables we want to retain)
+  summarize(bb.abund = n()) %>% # get bb.sp*transect abundance
+  ungroup %>%
+  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
+  semi_join(net, by = c("site", "date")) %>%
+  mutate(year = factor(year(date))) %>%
+  group_by(site, year, bb.sp) %>%
+  summarize(bb.sp.abund = mean(bb.abund))
+
+site_FL_aux <- survey %>%
+  mutate(year = factor(year(date))) %>%
+  filter(flower.cover > 0) %>%
+  group_by(site, year, date) %>%
+  summarize(fl.rich = length(unique(plant.sp)),
+            fl.abund = sum(flower.cover)) %>%
+  group_by(site, year) %>%
+  summarize(fl.rich = mean(fl.rich),
+            log.fl.abund = log(mean(fl.abund)),
+            fl.transects = length(unique(date)))
+
+# Network- and group-level metrics
 netmet_boot_2012 <- boot_net(net_2012, n = 9, iter = 40) %>%
@@
 
+### Species-level metrics 
+spmet_2012 <- boot_net_sp(net_2012, 9, 40) %>%
+  mutate(year = factor(rep("2012", n())))
+spmet_2011 <- boot_net_sp(net_2011, 4, 40) %>%
+  mutate(year = factor(rep("2011", n())))  
+spmet_2010 <- boot_net_sp(net_2010, 3, 40) %>%
+  mutate(year = factor(rep("2010", n())))
+
+spmet <- bind_rows(spmet_2012, spmet_2011, spmet_2010) %>%
+  left_join(site_BB_aux.sp) %>%
+  inner_join(site_FL_aux) %>%
+  spread(metric, value) %>%
+  left_join(site_data) %>%
+  ungroup() %>%
+  mutate(bb.sp = factor(bb.sp),
+         year = factor(year),
+         site = factor(site),
+         management = factor(management))
+
+spmet_mean <- spmet %>%
+  group_by(site, year) %>%
+  summarize(d = weighted.mean(d, bb.sp.abund, na.rm = TRUE)) %>%
+  ungroup()
+
+### Merge network metrics and auxiliary data
 netmet <- bind_rows(netmet_boot_2012, 
@@
   ungroup() %>%
-  mutate(metric = factor(metric, levels = c("weighted.NODF",
-                                            "niche.overlap.HL", 
-                                            "C.score.HL",
-                                            "generality.HL",
-                                            "web.asymmetry",
-                                            "weighted.connectance"))) %>%
+  spread(metric, value) %>%
+  left_join(spmet_mean, by = c("site", "year")) %>%
+  left_join(site_BB_aux, by = c("site", "year")) %>%
+  inner_join(site_FL_aux, by = c("site", "year")) %>% # one site-year is missing from floral survey data
+  left_join(site_data, by = "site") %>%
   mutate(year = factor(year),
-         site = factor(site))
+         site = factor(site),
+         management = factor(management))
 ```
 
-#### Species-level metric tables
-```{r, message=FALSE}
-### Species level
-#### BBs
-spmet_2012_BB <- boot_net_sp_BB(net_2012, 9, 40) %>%
-  mutate(year = factor(rep("2012", n())))
-spmet_2011_BB <- boot_net_sp_BB(net_2011, 4, 40) %>%
-  mutate(year = factor(rep("2011", n())))  
-spmet_2010_BB <- boot_net_sp_BB(net_2010, 3, 40) %>%
-  mutate(year = factor(rep("2010", n())))
+# Let's inspect correlations within our predictors
+```{r}
+library(GGally)
 
-spmet_BB <- boot_net_sp_BB(net, 5, 50) %>%
-  ungroup() %>%
-  mutate(year = factor(year))
+predictors <- netmet %>%
+  select(elev.mean, bb.rich, log.bb.abund, fl.rich, log.fl.abund)
 
-spmet_mean_BB <- spmet_BB %>%
-  group_by(site, year, elev.mean, metric) %>%
-  summarize(value = mean(value, na.rm = TRUE)) %>%
-  ungroup() %>%
-  mutate(year = factor(year))
-
-spmet_mean_BB_2012 <- spmet_2012_BB %>%
-  group_by(site, elev.mean, metric) %>%
-  summarize(value = mean(value, na.rm = TRUE))
-
-ggplot(spmet_mean_BB, aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~year)
-
-ggplot(spmet_2012_BB, aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~year)
-
-ggplot(spmet_mean_BB_2012, aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth()
-
-spmet_BB_gam <- bam(log(value + 0.01) ~ year + 
-                      s(elev.mean, bs = "tp", k = 10, m = 2) +
-                      s(elev.mean, by = factor(sp), k = 20, m = 1),
-                    data = spmet_BB,
-                    method = "fREML",
-                    discrete = TRUE) %>% getViz()
-
-check.gamViz(spmet_BB_gam)
-summary(spmet_BB_gam)
-print(plot(spmet_BB_gam, allTerms = TRUE), pages = 1)
-
-spmet_BB <- bind_rows(spmet_2012_BB, spmet_2011_BB, spmet_2010_BB)
+ggpairs(predictors)
 ```
 
-### Plot network analysis
+### Maybe the answer is not to abandon network analysis, but to include all the variables that confound my target relationship with elevation in a big model, then use variable selection to see what *does* predict network structure.
+I have massive concurvity, of course, so if the select() feature doesn't deal with that, I might need to opt for a model selection approach.
+
+A source of embuggerance that I have not yet dealt with is that when I bootstrap to equalize transects count, I do not do anything to equalize the temporal (either yday or GDD) interval spanned by the randomly drawn transects.
+This will, of course, average out as I increase iterations, so I guess the real problem is that the time interval spanned by the full set of transects for each site varies, so no matter how much bootstrapping I do, there will still be potential bias.
+This seems a fairly insurmountable limitation of the sampling structure of my data, so I will probably resort to just acknowledging it as a caveat of my analysis.
 ```{r}
@@
 ##### niche overlap
-ggplot(filter(netmet, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
+gam_niche.overlap <- gam(niche.overlap.HL ~ year + 
+                           s(site, bs = "re") +
+                           #management + # no effect
+                           s(elev.mean, k = 5) +
+                           #s(bb.rich, k = 5) + # too collinear with bb.abund
+                           s(fl.rich, k = 5) +
+                           #s(log.fl.abund, k = 5) + # no effect
+                           s(log.bb.abund, k = 5),
+                          select = TRUE,
+                          data = netmet,
+                          method = "REML") %>% getViz()
+
+ggplot(netmet, aes(elev.mean, niche.overlap.HL)) +
   geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x)) +
-  #geom_smooth() +
-  xlab("Elevation") +
-  ylab("Niche overlap") +
-  facet_wrap(~year) +
-  theme_light(16)
-
-gam_niche.overlap <- gamV(value ~ year + s(site, bs = "re") +
-                            s(elev.mean, by = year, k = 10, bs = "tp", m = 2), 
-                         data = filter(netmet, metric == "niche.overlap.HL"),
-                         method = "REML")
-
-gam_niche.overlap2 <- gamV(value ~ year + s(site, bs = "re") +
-                            s(elev.mean, k = 10, bs = "tp", m = 2), 
-                         data = filter(netmet, metric == "niche.overlap.HL"),
-                         method = "REML")
-
-AIC(gam_niche.overlap, gam_niche.overlap2)
+  theme_modern()
 
 check.gamViz(gam_niche.overlap)
+concurvity(gam_niche.overlap, full = TRUE)
+concurvity(gam_niche.overlap, full = FALSE)
 summary(gam_niche.overlap)
 print(plot(gam_niche.overlap, allTerms = TRUE), pages = 1)
+ggpredict(gam_niche.overlap)
+check_model(gam_niche.overlap)
 
 ##### C.score
-ggplot(filter(netmet, metric == "C.score.HL"), aes(elev.mean, value)) +
+gam_c.score <- gam(C.score.HL ~ year +
+                     s(site, bs = "re") +
+                     management +
+                     s(elev.mean, k = 5) +
+                     s(bb.rich, k = 5) +
+                     s(fl.rich, k = 5) +
+                     s(log.fl.abund, k = 5) +
+                     s(log.bb.abund, k = 5),
+                   select = TRUE,
+                   data = netmet,
+                   method = "REML") %>% getViz()
+
+check.gamViz(gam_c.score)
+concurvity(gam_c.score, full = TRUE)
+concurvity(gam_c.score, full = FALSE)
+summary(gam_c.score)
+print(plot(gam_c.score, allTerms = TRUE), pages = 1)
+
+##### weighted NODF: the main takeaway here is that higher BB abundance = higher nestedness, and that could easily be a sampling effect more than an ecologicla effect.
+Nevertheless, there remains a significant linear partial effect of elevation.
+It is worth noting that where elevation has an effect on network structure, it is linear after accounting for the influence of other variables.
+gam_weighted.NODF <- gam(weighted.NODF ~ year + 
+                           s(site, bs = "re") +
+                           #management +
+                           s(elev.mean, k = 5) +
+                           #s(bb.rich, k = 5) +
+                           s(fl.rich, k = 5) +
+                           #s(log.fl.abund, k = 5) +
+                           s(log.bb.abund, k = 5),
+                          select = TRUE,
+                          data = netmet,
+                          method = "REML") %>% getViz()
+
+check.gamViz(gam_weighted.NODF)
+concurvity(gam_weighted.NODF, full = TRUE)
+concurvity(gam_weighted.NODF, full = FALSE)
+summary(gam_weighted.NODF)
+print(plot(gam_weighted.NODF, allTerms = TRUE), pages = 1)
+ggpredict(gam_weighted.NODF)
+check_model(gam_weighted.NODF)
+
+##### togetherness
+gam_togetherness <- gam(togetherness.HL ~ year + 
+                           s(site, bs = "re") +
+                           #management +
+                           s(elev.mean, k = 5) +
+                           s(bb.rich, k = 5) +
+                           #s(log.fl.abund, k = 5) +
+                           #s(log.bb.abund, k = 5) +
+                           s(fl.rich, k = 5),
+                          select = TRUE,
+                          data = netmet,
+                          method = "REML") %>% getViz()
+
+check.gamViz(gam_togetherness)
+concurvity(gam_togetherness, full = TRUE)
+concurvity(gam_togetherness, full = FALSE)
+summary(gam_togetherness)
+print(plot(gam_togetherness, allTerms = TRUE), pages = 1)
+ggpredict(gam_togetherness)
+check_model(gam_togetherness)
+
+##### generality; this could be an interesting alternative to something like d', but I'm not sure whether it deals with the confunding effects of abundance -- singletons, in extreme cases.
+gam_generality <- gam(generality.HL ~ year + 
+                           s(site, bs = "re") +
+                           management +
+                           s(elev.mean, k = 5) +
+                           s(bb.rich, k = 5) +
+                           s(fl.rich, k = 5) +
+                           s(log.fl.abund, k = 5) +
+                           s(log.bb.abund, k = 5),
+                          select = TRUE,
+                          data = netmet,
+                          method = "REML") %>% getViz()
+
+check.gamViz(gam_generality)
+concurvity(gam_generality, full = TRUE)
+concurvity(gam_generality, full = FALSE)
+summary(gam_generality)
+print(plot(gam_generality, allTerms = TRUE), pages = 1)
+ggpredict(gam_generality)
+check_model(gam_generality)
+
+##### community mean d'
+gam_mean_d <- gam(d ~ year + 
+                           s(site, bs = "re") +
+                           #management +
+                           #s(elev.mean, k = 5) +
+                           #s(bb.rich, k = 5) +
+                           #s(log.fl.abund, k = 5) +
+                           #s(fl.rich, k = 5) +
+                           s(log.bb.abund, k = 5),
+                          select = TRUE,
+                          data = netmet,
+                          method = "REML") %>% getViz()
+
+check.gamViz(gam_mean_d)
+concurvity(gam_mean_d, full = TRUE)
+concurvity(gam_mean_d, full = FALSE)
+summary(gam_mean_d)
+print(plot(gam_mean_d, allTerms = TRUE), pages = 1)
+ggpredict(gam_mean_d)
+check_model(gam_mean_d)
+
+ggplot(netmet, aes(log.bb.abund, d, color = year)) +
   geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x)) +
-  xlab("Elevation") +
-  ylab("C score") +
-  facet_wrap(~year) +
-  theme_light(16)
+  geom_smooth(method = "lm") +
+  theme_abyss()
 
-gam_cscore <- gamV(value ~ year + s(site, bs = "re") + 
-                     s(elev.mean, bs = "tp", k = 10, m = 2), 
-                  data = filter(netmet, metric == "C.score.HL"),
-                  #family = "scat",
-                  method = "REML")
+ggplot(netmet, aes(elev.mean, d, color = year)) +
+  geom_point() +
+  geom_smooth() +
+  theme_abyss()
+```
 
-check.gamViz(gam_cscore)
-summary(gam_cscore)
-print(plot(gam_cscore, allTerms = TRUE), pages = 1)
+### But let's zoom in a bit closer on d' and see if we can model it per-species
+```{r}
+gam_d.sp_GI <- gam(d ~ year + 
+                     s(site, bs = "re") +
+                     management +
+                     s(bb.sp, bs = "re") +
+                     s(elev.mean, bs = "tp", k = 8) +
+                     s(elev.mean, by = bb.sp, bs = "tp", k = 8) +
+                     s(fl.rich, k = 8) +
+                     s(log.fl.abund, k = 8) +
+                     s(log.bb.sp.abund, k = 5),
+                   select = TRUE,
+                   data = spmet,
+                   method = "REML") %>% getViz()
 
-#####################################################################################3
+check.gamViz(gam_d.sp_GI)
+concurvity(gam_d.sp_GI, full = TRUE)
+concurvity(gam_d.sp_GI, full = FALSE)
+summary(gam_d.sp_GI)
+print(plot(gam_d.sp_GI, allTerms = TRUE), pages = 4)
 
-#### Per species values
-ggplot(filter(spmet_2012_BB,
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf") & metric == "d"), 
-       aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
-  facet_wrap(~sp, scales = "free")
+gam_d.sp_G <- gam(d ~ year + 
+                    s(site, bs = "re") +
+                    management +
+                    s(bb.sp, bs = "re") +
+                    s(elev.mean, bs = "tp", k = 8) +
+                    s(fl.rich, k = 8) +
+                    s(log.fl.abund, k = 8) +
+                    s(log.bb.sp.abund, k = 5),
+                  select = TRUE,
+                  data = spmet,
+                  method = "REML") %>% getViz()
 
-ggplot(filter(spmet_2011_BB,
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf") & metric == "d"), 
-       aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
-  facet_wrap(~sp, scales = "free")
+check.gamViz(gam_d.sp_G)
+concurvity(gam_d.sp_G, full = TRUE)
+concurvity(gam_d.sp_G, full = FALSE)
+summary(gam_d.sp_G)
+print(plot(gam_d.sp_G, allTerms = TRUE), pages = 1)
+check_model(gam_d.sp_G)
+modelG_elev.mean <- ggpredict(gam_d.sp_G, terms = "elev.mean")
+plot(modelG_elev.mean)
 
-ggplot(filter(spmet_2010_BB,
-              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf") & metric == "d"), 
-       aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
-  facet_wrap(~sp, scales = "free")
+# This simpler varying-smooth model explains just as much variance as the more complex GI model, so it should be preferred. 
+gam_d.sp_GS <- gam(d ~ year + 
+                     management +
+                     s(site, bs = "re") +
+                     s(elev.mean, bs = "tp", k = 8) +
+                     s(elev.mean, bb.sp, bs = "fs", k = 8) +
+                     s(log.fl.abund, k = 8) +
+                     s(fl.rich, k = 8) +
+                     s(log.bb.sp.abund, k = 5),
+                   select = TRUE,
+                   data = spmet,
+                   method = "REML") %>% getViz()
 
-#### Community mean d
-##### BB
-ggplot(filter(spmet_mean_BB, metric == "d"), aes(elev.mean, value)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10)) +
-  xlab("Elevation") +
-  ylab("Community mean d'") +
-  facet_wrap(~year) +
-  theme_light(16)
+check.gamViz(gam_d.sp_GS)
+concurvity(gam_d.sp_GS, full = TRUE)
+concurvity(gam_d.sp_GS, full = FALSE)
+summary(gam_d.sp_GS)
+print(plot(gam_d.sp_GS, allTerms = TRUE), pages = 1)
+check_model(gam_d.sp_GS)
+plot(ggpredict(gam_d.sp_GS))
 
-gam_d1 <- gamV(value ~ year + 
-                 s(elev.mean, k = 8, bs = "tp", by = year, m = 1), 
-             data = filter(spmet_mean_BB, metric == "d"),
-             method = "REML")
+# What about I? Does it do any better than S?
+gam_d.sp_I <- gam(d ~ year + 
+                     s(site, bs = "re") +
+                     management +
+                     s(bb.sp, bs = "re") +
+                     s(elev.mean, by = bb.sp, bs = "tp", k = 8) +
+                     s(fl.rich, k = 8) +
+                     s(log.fl.abund, k = 8) +
+                     s(log.bb.sp.abund, k = 5),
+                   select = TRUE,
+                   data = spmet,
+                   method = "REML") %>% getViz()
 
-check.gamViz(gam_d1)
-summary(gam_d1)
-plot(gam_d1, allTerms = TRUE)
+check.gamViz(gam_d.sp_I)
+concurvity(gam_d.sp_I, full = TRUE)
+concurvity(gam_d.sp_I, full = FALSE)
+summary(gam_d.sp_I)
+print(plot(gam_d.sp_I, allTerms = TRUE), pages = 4)
 
-summary(lm(value ~ elev.mean, data =  filter(spmet_mean_BB, metric == "d" & year == 2012)))
-plot(lm(value ~ elev.mean, data =  filter(spmet_mean_BB, metric == "d" & year == 2012)))
+AIC(gam_d.sp_I, gam_d.sp_S, gam_d.sp_G, gam_d.sp_GI, gam_d.sp_GS)
+
+ggplot(spmet, aes(log.bb.sp.abund)) +
+  geom_histogram() +
+  theme_modern()
+
+ggplot(spmet, aes(log.bb.sp.abund)) +
+  geom_histogram() +
+  theme_blackboard()
+
+ggplot(spmet, aes(elev.mean, d)) +
+  geom_point(aes(color = bb.sp)) +
+  geom_smooth() +
+  theme_blackboard() +
+  facet_wrap(~year)
+
+# We lose virtually no explanatory power in going from GI, to GS, to S. So, S is the preferred model right now.
+gam_d.sp_S <- gam(d ~ year + 
+                    s(site, bs = "re") +
+                    management +
+                    s(elev.mean, bb.sp, bs = "fs", k = 8) +
+                    s(fl.rich, k = 8) +
+                    s(log.fl.abund, k = 8) +
+                    s(log.bb.sp.abund, k = 5),
+                  select = TRUE,
+                  data = spmet,
+                  method = "REML") %>% getViz()
+
+check.gamViz(gam_d.sp_S)
+concurvity(gam_d.sp_S, full = TRUE)
+concurvity(gam_d.sp_S, full = FALSE)
+summary(gam_d.sp_S)
+print(plot(gam_d.sp_S, allTerms = TRUE), pages = 1)
+
+###  There remains a lot of noise in my multi-year models above, and a possible explanation for that is that the influence of year extends beyond just variation in the intercept.
+What if year affects the shape of Y ~ X?
+In these year subset models, I can drop both 'year' and 's(site, bs = "re")'.
+
+gam_d.sp_GI.2010 <- gam(d ~ management +
+                          s(bb.sp, bs = "re") +
+                          s(elev.mean, bs = "tp", k = 8) +
+                          s(elev.mean, by = bb.sp, bs = "tp", k = 8) +
+                          s(fl.rich, k = 8) +
+                          s(log.fl.abund, k = 8) +
+                          s(log.bb.sp.abund, k = 5),
+                        select = TRUE,
+                        data = filter(spmet, year == 2010),
+                        method = "REML") %>% getViz()
+
+check.gamViz(gam_d.sp_GI.2010)
+concurvity(gam_d.sp_GI.2010, full = TRUE)
+concurvity(gam_d.sp_GI.2010, full = FALSE)
+summary(gam_d.sp_GI.2010)
+print(plot(gam_d.sp_GI.2010, allTerms = TRUE), pages = 4)
+
+gam_d.sp_GI.2011 <- gam(d ~ management +
+                          s(bb.sp, bs = "re") +
+                          s(elev.mean, bs = "tp", k = 8) +
+                          s(elev.mean, by = bb.sp, bs = "tp", k = 8) +
+                          s(fl.rich, k = 8) +
+                          s(log.fl.abund, k = 8) +
+                          s(log.bb.sp.abund, k = 5),
+                        select = TRUE,
+                        data = filter(spmet, year == 2011),
+                        method = "REML") %>% getViz()
+
+check.gamViz(gam_d.sp_GI.2011)
+concurvity(gam_d.sp_GI.2011, full = TRUE)
+concurvity(gam_d.sp_GI.2011, full = FALSE)
+summary(gam_d.sp_GI.2011)
+print(plot(gam_d.sp_GI.2011, allTerms = TRUE), pages = 4)
+
+gam_d.sp_GI.2012 <- gam(d ~ management +
+                          s(bb.sp, bs = "re") +
+                          s(elev.mean, bs = "tp", k = 8) +
+                          s(elev.mean, by = bb.sp, bs = "tp", k = 8) +
+                          s(fl.rich, k = 8) +
+                          s(log.fl.abund, k = 8) +
+                          s(log.bb.sp.abund, k = 5),
+                        select = TRUE,
+                        data = filter(spmet, year == 2012),
+                        method = "REML") %>% getViz()
+
+check.gamViz(gam_d.sp_GI.2012)
+concurvity(gam_d.sp_GI.2012, full = TRUE)
+concurvity(gam_d.sp_GI.2012, full = FALSE)
+summary(gam_d.sp_GI.2012)
+print(plot(gam_d.sp_GI.2012, allTerms = TRUE), pages = 4)
+
+### I-form models
+gam_d.sp_I.2010 <- gam(d ~ management +
+                          s(bb.sp, bs = "re") +
+                          s(elev.mean, by = bb.sp, bs = "tp", k = 8) +
+                          s(fl.rich, k = 8) +
+                          s(log.fl.abund, k = 8) +
+                          s(log.bb.sp.abund, k = 5),
+                        select = TRUE,
+                        data = filter(spmet, year == 2010),
+                        method = "REML") %>% getViz()
+
+check.gamViz(gam_d.sp_I.2010)
+concurvity(gam_d.sp_I.2010, full = TRUE)
+concurvity(gam_d.sp_I.2010, full = FALSE)
+summary(gam_d.sp_I.2010)
+print(plot(gam_d.sp_I.2010, allTerms = TRUE), pages = 4)
+
+gam_d.sp_I.2011 <- gam(d ~ management +
+                          s(bb.sp, bs = "re") +
+                          s(elev.mean, by = bb.sp, bs = "tp", k = 8) +
+                          s(fl.rich, k = 8) +
+                          s(log.fl.abund, k = 8) +
+                          s(log.bb.sp.abund, k = 5),
+                        select = TRUE,
+                        data = filter(spmet, year == 2011),
+                        method = "REML") %>% getViz()
+
+check.gamViz(gam_d.sp_I.2011)
+concurvity(gam_d.sp_I.2011, full = TRUE)
+concurvity(gam_d.sp_I.2011, full = FALSE)
+summary(gam_d.sp_I.2011)
+print(plot(gam_d.sp_I.2011, allTerms = TRUE), pages = 4)
+
+gam_d.sp_I.2012 <- gam(d ~ management +
+                          s(bb.sp, bs = "re") +
+                          s(elev.mean, by = bb.sp, bs = "tp", k = 8) +
+                          s(fl.rich, k = 8) +
+                          s(log.fl.abund, k = 8) +
+                          s(log.bb.sp.abund, k = 5),
+                        select = TRUE,
+                        data = filter(spmet, year == 2012),
+                        method = "REML") %>% getViz()
+
+check.gamViz(gam_d.sp_I.2012)
+concurvity(gam_d.sp_I.2012, full = TRUE)
+concurvity(gam_d.sp_I.2012, full = FALSE)
+summary(gam_d.sp_I.2012)
+print(plot(gam_d.sp_I.2012, allTerms = TRUE), pages = 4)
 ```
+
+### Now let's try all this with GLMMs
+```{r}
+##### niche overlap
+lmer_niche.overlap <- lmer(niche.overlap.HL ~ elev.mean + 
+                            I(elev.mean^2) +
+                            I(elev.mean^3) +
+                            bb.rich +
+                            fl.rich +
+                            log.fl.abund +
+                            bb.abund +
+                            (1 | year),
+                          data = netmet)
+
+step(lmer_niche.overlap,reduce.random=FALSE) # stepwise variable selection; commented-out variables were removed based on the results of this step
+summary(glm_niche.overlap)
+
+#####  c-score
+lmer_c.score <- lmer(C.score.HL ~ elev.mean + 
+                            I(elev.mean^2) +
+                            I(elev.mean^3) +
+                            bb.rich +
+                            fl.rich +
+                            log.fl.abund +
+                            bb.abund +
+                            (1 | year),
+                          data = netmet)
+
+step(lmer_c.score,reduce.random=FALSE) # stepwise variable selection; commented-out variables were removed based on the results of this step
+summary(lmer_c.score)
+
+```
+
+### Does tongue-length / corolla depth matching vary with elevation?
+```{r}
+bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
+  dplyr::select(subgenus, bb.sp, pbl.w, pbl.w.class)
+
+fl_ktype <- read.csv("../processed_data/floral_k_type_nom_fix.csv") %>%
+  dplyr::select(plant.sp, k.type) %>%
+  mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"),
+         k.type.ss = str_extract(k.type, "[-+]?[0-9]*"))
+
+fl_morph <- read.csv("../processed_data/flower_morphology_GitaBenadi.csv") %>%
+  rename(plant.sp = Species) %>%
+  group_by(plant.sp) %>%
+  summarize(corolla.depth = mean(LN))
+
+fl_tax <- read_csv("../processed_data/floral_tax.csv")
+
+net_traits <- net %>%
+  # group_by(site, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
+  # summarize(freq = n()) %>%
+  # ungroup() %>%
+  left_join(bb_traits) %>%
+  left_join(fl_ktype) %>%
+  left_join(fl_morph) %>%
+  separate(plant.sp, c("plant.genus", "plant.species"), sep = " ", remove = FALSE) %>%
+  left_join(fl_tax) %>%
+  mutate(tongue.mismatch = abs(pbl.w - corolla.depth),
+         bb.sp = factor(bb.sp),
+         plant.sp = factor(plant.sp),
+         plant.genus = factor(plant.genus),
+         subgenus = factor(subgenus),
+         k.type.s = factor(k.type.s),
+         plant.family = factor(plant.family))
+
+ggplot(net_traits, aes(elev.mean, tongue.mismatch)) +
+  geom_point(color = "yellow") +
+  geom_smooth(method = "lm") +
+  facet_wrap(~bb.sp) +
+  theme_abyss()
+```
+
+### Since we clearly have structured collinearity in our preditors, i.e.
+elevation shapes abundance and diversity which shape network structure, perhaps SEM could be appropriate?
+The main question is whether it can handle the potentially complex (nonlinear) relationships between elevation and diversity/abundance, and whether it can handle the hierarchical structure of my data.
+```{r}
+library(lavaan)
+
+```

### 2020-07-16 analyses/trash_can.R

@@ -981,3 +981,53 @@ betalinkr_boot <- function(net, n, iter) {
 ```{r}
 beta1 <- betalinkr_boot(net, 3, 10)
 ```
+
+netmet_sitedate <- site_date_net(net, min_higher = 2, min_lower = 2) %>%
+  mutate(site = factor(site),
+         date = ymd(date),
+         yday = yday(date),
+         year = factor(year(date)))
+
+ggplot(netmet_sitedate, aes(elev.mean, niche.overlap.HL)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~year)
+
+ggplot(netmet_sitedate, aes(elev.mean, niche.overlap.HL)) +
+  geom_point() +
+  geom_smooth()
+
+ggplot(netmet_sitedate, aes(yday, niche.overlap.HL, color = year)) +
+  geom_point() +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE) +
+  facet_wrap(~site, scales = "free")
+
+bigbam_2010 <- bamV(niche.overlap.HL ~ s(site, bs = "re") +
+                      te(yday, elev.mean, 
+                         bs= c("tp", "tp"), 
+                         k=c(8, 8), m=1),
+                    data = filter(netmet_sitedate, year == 2010))
+
+check.gamViz(bigbam_2010)
+bigbam_2010.sum <- summary(bigbam_2010)
+print(plot(bigbam_2010, allTerms = TRUE), pages = 1)
+
+bigbam_2011 <- bamV(niche.overlap.HL ~ s(site, bs = "re") +
+                      te(yday, elev.mean, 
+                         bs= c("tp", "tp"), 
+                         k=c(8, 8), m=1),
+                    data = filter(netmet_sitedate, year == 2011))
+
+check.gamViz(bigbam_2011)
+bigbam_2011.sum <- summary(bigbam_2011)
+print(plot(bigbam_2011, allTerms = TRUE), pages = 1)
+
+bigbam_2012 <- bamV(niche.overlap.HL ~ s(site, bs = "re") +
+                      te(yday, elev.mean, 
+                         bs= c("tp", "tp"), 
+                         k=c(8, 8), m=1),
+                    data = filter(netmet_sitedate, year == 2012))
+
+check.gamViz(bigbam_2012)
+bigbam_2012.sum <- summary(bigbam_2012)
+print(plot(bigbam_2012, allTerms = TRUE), pages = 1)

### 2020-07-16 analyses/network_analysis.Rmd
woohoo SEM!

@@
 library(mgcViz)
+library(lavaan)
 library(tidyverse)
@@
 
+### Since we clearly have structured collinearity in our preditors, i.e.
+elevation shapes abundance and diversity which shape network structure, perhaps SEM could be appropriate?
+The main question is whether it can handle the potentially complex (nonlinear) relationships between elevation and diversity/abundance, and whether it can handle the hierarchical structure of my data.
+```{r}
+library(piecewiseSEM)
+
+netmet.scaled <- netmet %>%
+  select(site, year, niche.overlap.HL, elev.mean, fl.rich, bb.rich, log.fl.abund, log.bb.abund) %>%
+  mutate(niche.overlap.HL = scale(niche.overlap.HL),
+         elev.mean = scale(elev.mean),
+         fl.rich = scale(fl.rich),
+         log.fl.abund = scale(log.fl.abund),
+         log.bb.abund = scale(log.bb.abund)) %>%
+  as.matrix() %>%
+  as_tibble() %>%
+  mutate(niche.overlap.HL = as.numeric(niche.overlap.HL),
+         elev.mean = as.numeric(elev.mean),
+         fl.rich = as.numeric(fl.rich),
+         log.fl.abund = as.numeric(log.fl.abund),
+         log.bb.abund = as.numeric(log.bb.abund),
+         site = factor(site),
+         year = factor(year)) %>%
+  na.omit()
+
+mod_niche.overlap <- psem(lmer(niche.overlap.HL ~ elev.mean + 
+                                fl.rich + log.bb.abund + 
+                                 (1 | year) + (1 | site), 
+                              netmet.scaled),
+                          lm(fl.rich ~ elev.mean, 
+                             netmet.scaled),
+                          lm(log.bb.abund ~ elev.mean, 
+                             netmet.scaled))
+summary(mod_niche.overlap)
+plot(mod_niche.overlap)
+```
+
 ### But let's zoom in a bit closer on d' and see if we can model it per-species
@@
 
-### Now let's try all this with GLMMs
-```{r}
-##### niche overlap
-lmer_niche.overlap <- lmer(niche.overlap.HL ~ elev.mean + 
-                            I(elev.mean^2) +
-                            I(elev.mean^3) +
-                            bb.rich +
-                            fl.rich +
-                            log.fl.abund +
-                            bb.abund +
-                            (1 | year),
-                          data = netmet)
-
-step(lmer_niche.overlap,reduce.random=FALSE) # stepwise variable selection; commented-out variables were removed based on the results of this step
-summary(glm_niche.overlap)
-
-#####  c-score
-lmer_c.score <- lmer(C.score.HL ~ elev.mean + 
-                            I(elev.mean^2) +
-                            I(elev.mean^3) +
-                            bb.rich +
-                            fl.rich +
-                            log.fl.abund +
-                            bb.abund +
-                            (1 | year),
-                          data = netmet)
-
-step(lmer_c.score,reduce.random=FALSE) # stepwise variable selection; commented-out variables were removed based on the results of this step
-summary(lmer_c.score)
-
-```
-
 ### Does tongue-length / corolla depth matching vary with elevation?
@@
 ```
-
-### Since we clearly have structured collinearity in our preditors, i.e.
-elevation shapes abundance and diversity which shape network structure, perhaps SEM could be appropriate?
-The main question is whether it can handle the potentially complex (nonlinear) relationships between elevation and diversity/abundance, and whether it can handle the hierarchical structure of my data.
-```{r}
-library(lavaan)
-
-```

### 2020-07-27 analyses/BB_abund_GAM.Rmd
Refining GAMs, exploring global hierarchical specification

--- 2020-07-08 analyses/BB_abund_GAM.Rmd
@@
 library(lubridate)
+```
+
+### Reprex for Strenge Jacke
+```{r}
+library(reprex)
+
+library(mgcv)
+library(dplyr)
+library(ggeffects)
+
+data(iris)
+
+data <- iris %>%
+  data.frame()
+
+mod <- gam(Petal.Width ~ Species +
+             te(Sepal.Length, Sepal.Width, by = Species),
+           data = data)
+
+effects <- ggeffect(mod)
+means <- ggemmeans(mod)
+pred <- ggpredict(mod)
 ```
@@
 
-### FL and BB abundance / diversity
+### BB abundance / diversity
 ```{r echo=FALSE, include=FALSE}
@@
          bb.sp = factor(bb.sp),
-         management = factor(management)) %>%
-  semi_join(net, by = c("site", "date")) %>%
+         management = factor(management),
+         site = factor(site),
+         transect = factor(transect)) %>%
   group_by(year, bb.sp) %>%
@@
 GAM modeling of abundance per bb species
+
+I think I need to include site as a random effect because if I don't the pseudoreplication of elevation is pretty extreme, especially in 2012 where I have lots of repeated measures.
+I am not, for example, measuring floral abundance at 1000 m 20 independent times; I'm measuring it 20 times at the same spot.
+This does, of  course, create identifiability issues insofar as site and elevation are concurved.
+But I think I need to accept these problems.
+
 ```{r}
-### Call models
-bb_abund.2010 <- bam(bb.abund ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
+### Global model
+bb_abund.global <- bam(bb.abund ~
+                s(bb.sp, bs = "re") +
+                s(year, bs = "re") +
+                s(site, bs = "re") +
+                te(yday, elev.mean, 
                           by = bb.sp,
-                                                            bs= c("tp", "tp"), 
-                                                            k=c(10, 10), m=1),
+                          #id = 0,
+                          bs= c("gp", "tp"), 
+                          k=c(8, 8), 
+                          m=1) +
+                te(yday, elev.mean, 
+                          by = year,
+                          #id = 0,
+                          bs= c("gp", "tp"), 
+                          k=c(4, 4), 
+                          m=2),
+              data = filter(bb_abund_sp, bb.total.abund >= 5),
+              discrete = TRUE,
+              family = "tw",
+              select = TRUE,
+              method = "fREML") %>% getViz(nsim = 500)
+
+bb_abund.global2 <- gam(bb.abund ~
+                s(bb.sp, bs = "re") +
+                s(year, bs = "re") +
+                s(site, bs = "re") +
+                te(yday, elev.mean, 
+                          by = bb.sp,
+                          #id = 0,
+                          bs= c("gp", "tp"), 
+                          k=c(6, 6), 
+                          m=1) +
+                te(yday, elev.mean, 
+                          by = year,
+                          #id = 0,
+                          bs= c("gp", "tp"), 
+                          k=c(4, 4), 
+                          m=2),
+              data = filter(bb_abund_sp, bb.total.abund >= 5 & bb.sp %in% c("bss", "pasc", "prat", "pyre")),
+              discrete = FALSE,
+              family = ziP(),
+              select = TRUE,
+              method = "REML") %>% getViz(nsim = 500)
+
+ziptest <- gam(bb.abund ~
+                t2(yday, elev.mean, 
+                          by = bb.sp,
+                          #id = 0,
+                          bs= c("gp", "tp"), 
+                          k=c(3, 3), 
+                          m=1),
+              data = filter(bb_abund_sp, bb.total.abund >= 5),
+              #discrete = FALSE,
+              family = ziP(),
+              select = TRUE,
+              method = "REML") %>% getViz(nsim = 50)
+
+check.gamViz(bb_abund.global)
+check.gamViz(bb_abund.global2)
+check.gamViz(ziptest)
+
+bb_abund.global.sum <- summary(bb_abund.global)
+bb_abund.global2.sum <- summary(bb_abund.global2)
+
+AIC(bb_abund.global, bb_abund.global2)
+
+print(plot(bb_abund.global, allTerms = TRUE), pages = 3)
+print(plot(bb_abund.global2, allTerms = TRUE), pages = 3)
+
+bb_abund.global.ck <- check2D(bb_abund.global, x1 = "yday", x2 = "elev.mean")
+bb_abund.global.ck + l_gridCheck2D(gridFun = mean)
+
+### 2010
+bb_abund.2010 <- bam(bb.abund ~ 
+                       s(bb.sp, bs = "re") + 
+                       #bb.sp +
+                       #s(transect, bs = "re") +
+                       management + 
+                       s(site, bs = "re") +
+                       # te(yday, elev.mean,
+                       #    bs= c("gp", "cr"),
+                       #    k=c(10, 10),
+                       #    m=2) +
+                       te(yday, elev.mean, 
+                          by = bb.sp,
+                          #id = 0,
+                          bs= c("gp", "cr"), 
+                          k=c(5, 5), 
+                          m=2),
                      data = filter(bb_abund_sp, year == 2010 & bb.total.abund >= 5),
                      discrete = TRUE,
-                        family = "poisson",
+                     family = "tw",
                      select = TRUE,
-                        method = "fREML") %>% getViz()
+                     method = "fREML") %>% getViz(nsim = 500)
+
+bb_abund.2010.2 <- gam(bb.abund ~ 
+                       s(bb.sp, bs = "re") + 
+                       #bb.sp +
+                       #s(transect, bs = "re") +
+                       management + 
+                       s(site, bs = "re") +
+                       # te(yday, elev.mean,
+                       #    bs= c("gp", "cr"),
+                       #    k=c(10, 10),
+                       #    m=2) +
+                       te(yday, elev.mean, 
+                          by = bb.sp,
+                          #id = 0,
+                          bs= c("gp", "cr"), 
+                          k=c(5, 5), 
+                          m=2),
+                     data = filter(bb_abund_sp, year == 2010 & bb.total.abund >= 5),
+                     #discrete = TRUE,
+                     family = ziP(),
+                     select = TRUE,
+                     method = "REML") %>% getViz(nsim = 500)
 
 check.gamViz(bb_abund.2010)
+check.gamViz(bb_abund.2010.2)
 bb_abund.2010.sum <- summary(bb_abund.2010)
-print(plot(bb_abund.2010, allTerms = TRUE), pages = 4)
+print(plot(bb_abund.2010, allTerms = TRUE), pages = 1)
+bb_abund.2010.ck <- check2D(bb_abund.2010, x1 = "yday", x2 = "elev.mean")
+bb_abund.2010.ck + l_gridCheck2D(gridFun = mean)
 
-bb_abund.2011 <- bam(bb.abund ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
+### 2011
+bb_abund.2011 <- bam(bb.abund ~ 
+                       s(bb.sp, bs = "re") + 
+                       #bb.sp +
+                       #s(transect, bs = "re") +
+                       #management + 
+                       s(site, bs = "re") +
+                       # te(yday, elev.mean,
+                       #    bs= c("gp", "cr"),
+                       #    k=c(10, 10),
+                       #    m=2) +
+                       te(yday, elev.mean, 
                           by = bb.sp,
-                                                            bs= c("tp", "tp"), 
-                                                            k=c(10, 10), m=1),
+                          #id = 0,
+                          bs= c("gp", "cr"), 
+                          k=c(10, 10), 
+                          m=1),
                      data = filter(bb_abund_sp, year == 2011 & bb.total.abund >= 5),
                      discrete = TRUE,
-                        family = "poisson",
+                     family = "tw",
                      select = TRUE,
-                        method = "fREML") %>% getViz()
+                     method = "fREML") %>% getViz(nsim = 500)
 
   check.gamViz(bb_abund.2011)
+concurvity(bb_abund.2011)
 bb_abund.2011.sum <- summary(bb_abund.2011)
-print(plot(bb_abund.2011 , allTerms = TRUE), pages = 4)
+print(plot(bb_abund.2011, allTerms = TRUE), pages = 1)
+bb_abund.2011.ck <- check2D(bb_abund.2011, x1 = "yday", x2 = "elev.mean")
+bb_abund.2011.ck + l_gridCheck2D(gridFun = mean)
 
-bb_abund.2012 <- bam(bb.abund ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
+### 2012
+bb_abund.2012 <- bam(bb.abund ~ 
+                       s(bb.sp, bs = "re") + 
+                       #bb.sp +
+                       #s(transect, bs = "re") +
+                       management + 
+                       s(site, bs = "re") +
+                       # te(yday, elev.mean,
+                       #    bs= c("gp", "cr"),
+                       #    k=c(10, 10),
+                       #    m=2) +
+                       te(yday, elev.mean, 
                           by = bb.sp,
-                                                            bs= c("tp", "tp"), 
-                                                            k=c(10, 10), m=1),
+                          #id = 0,
+                          bs= c("gp", "cr"), 
+                          k=c(10, 10), 
+                          m=2),
                      data = filter(bb_abund_sp, year == 2012 & bb.total.abund >= 5),
                      discrete = TRUE,
-                        family = "poisson",
+                     family = "tw",
                      select = TRUE,
-                        method = "fREML") %>% getViz()
+                     method = "fREML") %>% getViz(nsim = 500)
 
@@
 bb_abund.2012.sum <- summary(bb_abund.2012)
-print(plot(bb_abund.2012 , allTerms = TRUE), pages = 4)
-plot(bb_abund.2012)
-
+print(plot(bb_abund.2012, allTerms = TRUE), pages = 1)
+bb_abund.2012.ck <- check2D(bb_abund.2012, x1 = "yday", x2 = "elev.mean")
+bb_abund.2012.ck + l_gridCheck2D(gridFun = mean)
 ```
 
+### Tabulate plots
 ```{r}
-bb_abund.2012.pred <- predict(bb_abund.2012)
+t1 <- ggpredict(bb_abund.2010, terms = c("elev.mean"))
+plot(t1)
 
-ggplot(bb_abund.2012.pred, aes(yday, elev.mean, fit)) +
-  geom_raster(aes(fill = exp(fit))) +
-  #geom_contour(aes(z = fit), color = "black") +
-  geom_point(data = filter(fl_abund, year == 2012), 
-             aes(yday, elev.mean, size = fl.abund),
-             color = "white", alpha = 0.25) +
-  scale_fill_viridis_c()
+bss2010 <- plot(sm(bb_abund.2010, 4)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. terrestris/lucorum")), x = NULL, y = NULL)
+
+bss2011 <- plot(sm(bb_abund.2011, 4)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. terrestris/lucorum")), x = NULL, y = NULL)
+bss2012 <- plot(sm(bb_abund.2012, 4)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. terrestris/lucorum")), x = NULL, y = NULL)
+
+gers2010 <- plot(sm(bb_abund.2010, 5)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. gerstaeckeri")), x = NULL, y = NULL)
+gers2011 <- ggplot() #too rare
+gers2012 <- plot(sm(bb_abund.2012, 5)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. gerstaeckeri")), x = NULL, y = NULL)
+
+hort2010 <- plot(sm(bb_abund.2010, 6)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. hortorum")), x = NULL, y = NULL)
+hort2011 <- plot(sm(bb_abund.2011, 5)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. hortorum")), x = NULL, y = NULL)
+hort2012 <- plot(sm(bb_abund.2012, 6)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. hortorum")), x = NULL, y = NULL)
+
+hypn2010 <- plot(sm(bb_abund.2010, 7)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. hypnorum")), x = NULL, y = NULL)
+hypn2011 <- plot(sm(bb_abund.2011, 6)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. hypnorum")), x = NULL, y = NULL)
+hypn2012 <- ggplot()  #too rare
+
+jone2010 <- plot(sm(bb_abund.2010, 8)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. jonellus")), x = NULL, y = NULL)
+jone2011 <- plot(sm(bb_abund.2011, 7)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. jonellus")), x = NULL, y = NULL)
+jone2012 <- plot(sm(bb_abund.2012, 7)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. jonellus")), x = NULL, y = NULL)
+
+lapi2010 <- ggplot() # too rare 
+lapi2011 <- plot(sm(bb_abund.2011, 8)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. lapidarius")), x = NULL, y = NULL)
+lapi2012 <- plot(sm(bb_abund.2012, 8)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. lapidarius")), x = NULL, y = NULL)
+
+mend2010 <- plot(sm(bb_abund.2010, 9)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. mendax")), x = NULL, y = NULL)
+mend2011 <- plot(sm(bb_abund.2011, 9)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. mendax")), x = NULL, y = NULL)
+mend2012 <- plot(sm(bb_abund.2012, 9)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. mendax")), x = NULL, y = NULL)
+
+mont2010 <- plot(sm(bb_abund.2010, 10)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. monticola")), x = NULL, y = NULL)
+mont2011 <- plot(sm(bb_abund.2011, 10)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. monticola")), x = NULL, y = NULL)
+mont2012 <- plot(sm(bb_abund.2012, 10)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. monticola")), x = NULL, y = NULL)
+
+muci2010 <- plot(sm(bb_abund.2010, 11)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. mucidis")), x = NULL, y = NULL)
+muci2011 <- plot(sm(bb_abund.2011, 11)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. mucidis")), x = NULL, y = NULL)
+muci2012 <- plot(sm(bb_abund.2012, 11)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. mucidis")), x = NULL, y = NULL)
+
+pasc2010 <- plot(sm(bb_abund.2010, 12)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. pascuorum")), x = NULL, y = NULL)
+pasc2011 <- plot(sm(bb_abund.2011, 12)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. pascuorum")), x = NULL, y = NULL)
+pasc2012 <- plot(sm(bb_abund.2012, 12)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. pascuorum")), x = NULL, y = NULL)
+
+prat2010 <- plot(sm(bb_abund.2010, 13)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. pratorum")), x = NULL, y = NULL)
+prat2011 <- plot(sm(bb_abund.2011, 13)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. pratorum")), x = NULL, y = NULL)
+prat2012 <- plot(sm(bb_abund.2012, 13)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. pratorum")), x = NULL, y = NULL)
+
+psit2010 <- plot(sm(bb_abund.2010, 14)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. psithyrus")), x = NULL, y = NULL)
+psit2011 <- plot(sm(bb_abund.2011, 14)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. psithyrus")), x = NULL, y = NULL)
+psit2012 <- plot(sm(bb_abund.2012, 14)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. psithyrus")), x = NULL, y = NULL)
+
+pyre2010 <- plot(sm(bb_abund.2010, 15)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. pyrenaeus")), x = NULL, y = NULL)
+pyre2011 <- plot(sm(bb_abund.2011, 15)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. pyrenaeus")), x = NULL, y = NULL)
+pyre2012 <- plot(sm(bb_abund.2012, 15)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. pyrenaeus")), x = NULL, y = NULL)
+
+soro2010 <- plot(sm(bb_abund.2010, 16)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. soroensis")), x = NULL, y = NULL)
+soro2011 <- plot(sm(bb_abund.2011, 16)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. soroensis")), x = NULL, y = NULL)
+soro2012 <- plot(sm(bb_abund.2012, 16)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. soroensis")), x = NULL, y = NULL)
+
+wurf2010 <- plot(sm(bb_abund.2010, 17)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. wurflenii")), x = NULL, y = NULL)
+wurf2011 <- plot(sm(bb_abund.2011, 17)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. wurflenii")), x = NULL, y = NULL)
+wurf2012 <- plot(sm(bb_abund.2012, 17)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  #l_points() +
+  guides(fill = FALSE) +
+  labs(title = expression(italic("B. wurflenii")), x = NULL, y = NULL)
+
+pdf("./output/BB_abund_major.pdf", height = 11, width = 8.5)
+major_sp <- gridPrint(bss2010, bss2011, bss2012, 
+                      pasc2010, pasc2011, pasc2012,
+                      prat2010, prat2011, prat2012,
+                      psit2010, psit2011, psit2012,
+                      soro2010, soro2011, soro2012,
+                      wurf2010, wurf2011, wurf2012, ncol = 3)
+dev.off()
+
+pdf("./output/BB_abund_highelev.pdf", height = 5.5, width = 8.5)
+high_elev_sp <- gridPrint(pyre2010, pyre2011, pyre2012, 
+                          mend2010, mend2011, mend2012,
+                          mont2010, mont2011, mont2012, ncol = 3)
+dev.off()
+
+pdf("./output/BB_abund_minor.pdf", height = 11, width = 8.5)
+minor_sp <- gridPrint(gers2010, gers2011, gers2012,
+                      hort2010, hort2011, hort2012,
+                      hypn2010, hypn2011, hypn2012,
+                      jone2010, jone2011, jone2012,
+                      lapi2010, lapi2011, lapi2012,
+                      muci2010, muci2011, muci2012, ncol = 3)
+dev.off()
+
+plot(major_sp)
+ggsave(major_sp, "./output/BB_abund_major.pdf")
+plot(high_elev_sp)
+plot(minor_sp)
 ```
+
+### Bumble bee diversity
+```{r}
+bb_rich <- bb_abund_sp %>%
+  group_by(site, date, yday, year, elev.mean, management) %>%
+  summarize(bb.rich = length(unique(bb.sp)))
+```

### 2020-07-27 analyses/FL_abund_GAM.Rmd
Refining GAMs, exploring global hierarchical specification

--- 2020-07-13 analyses/FL_abund_GAM.Rmd
@@
 fl_abund_sp <- survey %>%
-  semi_join(shared_transects) %>%
+  #semi_join(shared_transects) %>%
   group_by(site, date, plant.sp) %>% # group by plant.sp*transect
@@
   mutate(year = factor(year(date)))
+
+test <- fl_abund_sp %>%
+  group_by(site, date) %>%
+  summarize(n = n())
 ```
@@
 ```{r}
-# This creates a data set where each bumble bee species present at a given site*date is paired with each floral species present at a given site date, even if there was never a visitation between them.
 observed_visits <- net %>%
@@
 
-# All possible plants * site * date
-possible_plants <- bind_rows(dplyr::select(net, c(site, date, plant.sp)),
-                             dplyr::select(survey, c(site, date, plant.sp))) %>%
-  unique() %>%
-  complete(site, date, plant.sp) %>%
-  semi_join(survey, by = c("site", "date"))
-
-# All possible bumbles * site * date
-possible_bb <- net %>%
-  dplyr::select(site, date, bb.sp) %>%
-  unique() %>%
-  complete(site, date, bb.sp) %>%
-  semi_join(net, by = c("site", "date"))
-
 # All possible bumble-plant pairings *site * date, filtered to include only those when the given bb.sp was present and the given plant.sp was *either* present *or* visited (see note below)
-visitation_by_abundance <- full_join(possible_plants, possible_bb) %>%
-  left_join(bb_abund_sp) %>%
-  left_join(fl_abund_sp) %>%
+visitation_by_abundance <- fl_abund_sp %>%
+  inner_join(bb_abund_sp) %>%
   filter(bb.abund > 0) %>%
+  filter(fl.abund > 0) %>% 
   left_join(observed_visits) %>%
-  filter(fl.abund > 0 | visits > 0) %>% # In about 10% of the bb.sp*plant.sp*site*year combos, the visited plant was not recorded in the floral survey data.
+  #filter(fl.abund > 0 | visits > 0) %>% # In about 10% of the bb.sp*plant.sp*site*year combos, the visited plant was not recorded in the floral survey data.
 There are a few different ways we could handle this, but I think an argument could be make that treating those plants as if they had zero abundance (even though they were evidently present gets close to the truth, since they must have been rather rare to have been missed in the surveying.)
@@
   summarize(prop.fl.abund.m = mean(prop.fl.abund),
-            prop.visits.wm = weighted.mean(prop.visits, bb.abund)) %>%
-  mutate(selection.ratio = prop.visits.wm/prop.fl.abund.m)
+            prop.visits.wm = weighted.mean(prop.visits, bb.abund))
 
-write_csv(visitation_by_abundance_mean, "../analyses/output/visitation_by_abundance_mean.csv")
+#write_csv(visitation_by_abundance_mean, "../analyses/output/visitation_by_abundance_mean.csv")
 
@@
        aes(log(prop.fl.abund.m + 0.01), log(prop.visits.wm))) +
+  geom_point(alpha = 0.5, aes(color = bb.sp)) +
+  geom_smooth(method = "lm")
+
+ggplot(filter(visitation_by_abundance_mean), 
+       aes(prop.fl.abund.m, prop.visits.wm)) +
   geom_point(alpha = 0.5, aes(color = bb.sp)) +
@@
 
+write_csv(preference, "../analyses/output/preference.csv")
+
 #################################
@@
 #################################
-bb_abund_sp_sum <- bb_abund_sp %>% # Get annual totals for BB species so that we can filter out really rare ones
-  group_by(year, bb.sp) %>%
-  summarize(bb.total.abund = sum(bb.abund))
+
+# For bb-fl combos that never co-occurred, set preference = 1
+pref_forbidden_link_fill <- inner_join(fl_abund_sp, bb_abund_sp) %>%
+  mutate(co = case_when(
+    fl.abund == 0 | bb.abund == 0 ~ 0,
+    fl.abund != 0 & bb.abund != 0 ~ 1)
+  ) %>%
+  group_by(plant.sp, bb.sp) %>%
+  summarize(co.max = max(co)) %>%
+  filter(co.max == 0) %>%
+  mutate(preference.update = rep(1, n())) %>%
+  dplyr::select(-co.max) %>%
+  inner_join(fl_abund_sp)
 
 fl_abund_wt <- fl_abund_sp %>%
-  right_join(preference) %>% 
+  select(-year) %>%
+  left_join(preference) %>%
+
+  # Add the forbidden links with preference set to 1 (neutral)
+  full_join(pref_forbidden_link_fill) %>%
+  mutate(preference = case_when(
+    is.na(preference) ~ preference.update,
+    !is.na(preference) ~ preference)
+    ) %>%
+  dplyr::select(-preference.update) %>%
+
   mutate(fl.abund.wt = fl.abund*preference) %>%
@@
   ungroup() %>%
-  left_join(bb_abund_sp) %>%
-  left_join(bb_abund_sp_sum)%>%
-  filter(bb.total.abund >= 5) %>% # It's not worth trying to model really rare species
   semi_join(survey, by = c("site", "date")) %>% # This gets rid of site*date combos that don't exist
@@
          management = factor(management),
-         site = factor(site)) %>%
+         site = factor(site),
+         transect = factor(transect)) %>%
   group_by(year) %>%
@@
 GAM modeling of preference-weighted abundance per bb species
-## Against any potential objections to my use of an added constant to avoid the log(0) problem, I would argue that when what is being measured is a resource, the presence of a resource in vanishingly small amounts is biologically equivalent to its absence.
-I suppose the strength of this argument hinges on how much we trust that our samples represent our sites.
-But if they don't, what are we even doing here?
-note that this argument does not hold when one is concerned with diversity and species distributions, since then a species rarity is a very different thing that its absence.
+
+** One thing I need to go back and reconsider is that right now, I think forbidden links (i.e.
+links between species that never co-occurred) are given a zero for preference.
+A better approach would be to set their preference to a neutral value, which would be 1 in my case, since exp(0) = 1 and my preference values are calculated as exp(residual) of my linear model of visitation ~ abundance.
+**
+
+I think I need to include site as a random effect because if I don't the pseudoreplication of elevation is pretty extreme, especially in 2012 where I have lots of repeated measures.
+I am not, for example, measuring floral abundance at 1000 m 20 independent times; I'm measuring it 20 times at the same spot.
+This does, of  course, create identifiability issues insofar as site and elevation are concurved.
+But I think I need to accept these problems.
+The seriousness of the identifiability problems depends in part, I think, on how mgcv handles random effect smooths.
+Does it first account for all the variation it can with the main effects, then model the residuals with the random effect?
+In this case, it's probably not a big problem, since the "partial" influence of site would orthogonal (right?) with that of te(elev.mean, yday).
+
 ```{r}
-### Pooled years
-weighted_fl_abundance.tw <- bam(fl.abund.wt ~ s(bb.sp, bs = "re") + s(management, bs = "re") + year + 
+### Global model
+weighted_fl_abundance.global <- bam(log(fl.abund.wt + 0.001) ~
+                                      s(bb.sp, bs = "re") +
+                                      #s(management, bs = "re") +
+                                      s(year, bs = "re") +
+                                      s(site, bs = "re") +
+                                        te(yday, elev.mean, 
+                                         by = year,
+                                                id = 0,
+                                                bs= c("gp", "tp"), 
+                                                k=c(8, 8), 
+                                                m=2) +
+                                      te(yday, elev.mean, 
+                                                by = bb.sp,
+                                                #id = 0,
+                                                bs= c("gp", "tp"), 
+                                                k=c(8, 8), 
+                                                m=1),
+                                    data = fl_abund_wt,
+                                    discrete = TRUE,
+                                    family = "gaussian",
+                                    select = TRUE,
+                                    method = "fREML") %>% getViz(nsim = 500)
+
+weighted_fl_abundance.global2 <- bam(log(fl.abund.wt + 0.001) ~
+                                      s(bb.sp, bs = "re") +
+                                      #s(management, bs = "re") +
+                                      s(year, bs = "re") +
                                       s(site, bs = "re") +
@@
                                                 by = bb.sp,
-                                     bs= c("tp", "tp"), 
-                                     k=c(10, 10), m=1),
+                                                #id = 0,
+                                                bs= c("gp", "tp"), 
+                                                k=c(10, 10), 
+                                                m=1),
                                     data = fl_abund_wt,
                                     discrete = TRUE,
-                                family = "tw",
+                                    family = "gaussian",
                                     select = TRUE,
-                                method = "fREML") %>% getViz()
+                                    method = "fREML") %>% getViz(nsim = 500)
 
-check.gamViz(weighted_fl_abundance.tw)
-weighted_fl_abundance.tw.sum <- summary(weighted_fl_abundance.tw)
-print(plot(weighted_fl_abundance.tw, allTerms = TRUE), pages = 1)
-saveRDS(weighted_fl_abundance.tw, "../analyses/output/weighted_fl_abundance.rds")
+weighted_fl_abundance.global3 <- bam(log(fl.abund.wt + 0.001) ~
+                                      s(bb.sp, bs = "re") +
+                                      #s(management, bs = "re") +
+                                      s(year, bs = "re") +
+                                      s(site, bs = "re") +
+                                      te(yday, elev.mean, 
+                                              id = 0,
+                                              bs= c("gp", "tp"), 
+                                              k=c(10, 10), 
+                                              m=2) +
+                                      te(yday, elev.mean, 
+                                                by = bb.sp,
+                                                #id = 0,
+                                                bs= c("gp", "tp"), 
+                                                k=c(10, 10), 
+                                                m=1),
+                                    data = fl_abund_wt,
+                                    discrete = TRUE,
+                                    family = "gaussian",
+                                    select = TRUE,
+                                    method = "fREML") %>% getViz(nsim = 500)
+
+weighted_fl_abundance.global4 <- bam(log(fl.abund.wt + 0.001) ~
+                                      s(bb.sp, bs = "re") +
+                                      #s(management, bs = "re") +
+                                      s(year, bs = "re") +
+                                      s(site, bs = "re") +
+                                      te(yday, elev.mean, year,
+                                                by = bb.sp,
+                                                #id = 0,
+                                                bs= c("gp", "tp", "re"), 
+                                                k=c(5, 5), 
+                                                m=1),
+                                    data = fl_abund_wt,
+                                    discrete = TRUE,
+                                    family = "gaussian",
+                                    select = TRUE,
+                                    method = "fREML") %>% getViz(nsim = 500)
+
+AIC(weighted_fl_abundance.global, weighted_fl_abundance.global2, weighted_fl_abundance.global3, weighted_fl_abundance.global4)
+
+check.gamViz(weighted_fl_abundance.global)
+check.gamViz(weighted_fl_abundance.global2)
+check.gamViz(weighted_fl_abundance.global3)
+check.gamViz(weighted_fl_abundance.global4)
+
+weighted_fl_abundance.global.sum <- summary(weighted_fl_abundance.global)
+weighted_fl_abundance.global.sum2 <- summary(weighted_fl_abundance.global2)
+weighted_fl_abundance.global.sum3 <- summary(weighted_fl_abundance.global3)
+weighted_fl_abundance.global.sum3 <- summary(weighted_fl_abundance.global4)
+
+print(plot(weighted_fl_abundance.global, allTerms = TRUE), pages = 4)
+print(plot(weighted_fl_abundance.global2, allTerms = TRUE), pages = 4)
+print(plot(weighted_fl_abundance.global3, allTerms = TRUE), pages = 4)
+print(plot(weighted_fl_abundance.global4, allTerms = TRUE), pages = 4)
+
+weighted_fl_abundance.global.ck <- check2D(weighted_fl_abundance.global, 
+                                         x1 = "yday", x2 = "elev.mean")
+weighted_fl_abundance.global.ck + l_gridCheck2D(gridFun = mean)
 
 ### 2010
-weighted_fl_abundance.2010.tw <- bam(fl.abund.wt ~ s(bb.sp, bs = "re") + s(management, bs = "re") + 
-                                       s(site, bs = "re") +
+weighted_fl_abundance.2010 <- bam(fl.abund.wt ~ s(bb.sp, bs = "re") + 
+                                    management + 
+                                    #slope.est +
+                                    #s(site, bs = "re") +
                                     te(yday, elev.mean, 
                                        by = bb.sp,
-                                          bs= c("tp", "tp"), 
-                                          k=c(10, 10), m=1),
+                                       bs= c("gp", "tp"), 
+                                       k=c(10, 10), m=2),
                                   data = filter(fl_abund_wt, year == 2010),
@@
                                   select = TRUE,
-                                     method = "fREML") %>% getViz()
+                                  method = "fREML") %>% getViz(nsim = 500)
 
-check.gamViz(weighted_fl_abundance.2010.tw)
-weighted_fl_abundance.2010.tw.sum <- summary(weighted_fl_abundance.2010.tw)
-print(plot(weighted_fl_abundance.2010.tw , allTerms = TRUE), pages = 1)
+check.gamViz(weighted_fl_abundance.2010)
+weighted_fl_abundance.2010.sum <- summary(weighted_fl_abundance.2010)
+print(plot(weighted_fl_abundance.2010, allTerms = TRUE), pages = 1)
+weighted_fl_abundance.2010.ck <- check2D(weighted_fl_abundance.2010, 
+                                         x1 = "yday", x2 = "elev.mean")
+weighted_fl_abundance.2010.ck + l_gridCheck2D(gridFun = mean)
 
 ### 2011
-weighted_fl_abundance.2011.qp <- bam(fl.abund.wt ~ s(bb.sp, bs = "re") + s(management, bs = "re") + 
+weighted_fl_abundance.2011 <- bam(fl.abund.wt ~ s(bb.sp, bs = "re") + 
+                                    #s(transect, bs = "re") +
+                                    management + 
+                                    #slope.est +
+                                    #s(site, bs = "re") +
+                                    te(yday, elev.mean, 
+                                       by = bb.sp,
+                                       bs= c("gp", "tp"), 
+                                       k=c(10, 10), m=2),
+                                  data = filter(fl_abund_wt, year == 2011),
+                                  discrete = TRUE,
+                                  family = "tw",
+                                  select = TRUE,
+                                  method = "fREML") %>% getViz(nsim = 500)
+
+check.gamViz(weighted_fl_abundance.2011)
+weighted_fl_abundance.2011.sum <- summary(weighted_fl_abundance.2011)
+print(plot(weighted_fl_abundance.2011, allTerms = TRUE), pages = 1)
+weighted_fl_abundance.2011.ck <- check2D(weighted_fl_abundance.2011, 
+                                         x1 = "yday", x2 = "elev.mean")
+weighted_fl_abundance.2011.ck + l_gridCheck2D(gridFun = mean)
+
+### 2012
+weighted_fl_abundance.2012 <- bam(fl.abund.wt ~ s(bb.sp, bs = "re") + 
+                                    #s(transect, bs = "re") +
+                                    management + 
+                                    #slope.est +
                                     s(site, bs = "re") +
@@
                                        by = bb.sp,
-                                          bs= c("tp", "tp"), 
-                                          k=c(10, 10), m=1),
-                                     data = filter(fl_abund_wt, year == 2011),
-                                     discrete = TRUE,
-                                     family = "quasipoisson",
-                                     select = TRUE,
-                                     method = "fREML") %>% getViz()
-
-check.gamViz(weighted_fl_abundance.2011.qp)
-weighted_fl_abundance.2011.qp.sum <- summary(weighted_fl_abundance.2011.qp)
-print(plot(weighted_fl_abundance.2011.qp , allTerms = TRUE), pages = 1)
-
-### 2012
-weighted_fl_abundance.2012.qp <- bam(fl.abund.wt ~ s(bb.sp, bs = "re") + s(management, bs = "re") + 
-                                       s(site, bs = "re") +
-                                       #slope.est + 
-                                       te(yday, elev.mean, 
-                                          by = bb.sp,
-                                          bs= c("tp", "tp"), 
+                                       bs= c("gp", "tp"), 
                                        k=c(10, 10), m=1),
@@
                                   discrete = TRUE,
-                                     family = "quasipoisson",
+                                  family = "tw",
                                   select = TRUE,
-                                     method = "fREML") %>% getViz()
+                                  method = "fREML") %>% getViz(nsim = 500)
 
-check.gamViz(weighted_fl_abundance.2012.qp)
-weighted_fl_abundance.2012.qp.sum <- summary(weighted_fl_abundance.2012.qp)
-print(plot(weighted_fl_abundance.2012.qp , allTerms = TRUE), pages = 1)
+check.gamViz(weighted_fl_abundance.2012)
+weighted_fl_abundance.2012.sum <- summary(weighted_fl_abundance.2012)
+print(plot(weighted_fl_abundance.2012, allTerms = TRUE), pages = 1)
+weighted_fl_abundance.2012.ck <- check2D(weighted_fl_abundance.2012, 
+                                         x1 = "yday", x2 = "elev.mean")
+weighted_fl_abundance.2012.ck + l_gridCheck2D(gridFun = mean)
+```
+
+### Tabulate plots
+```{r}
+bssFL_2010 <- plot(sm(weighted_fl_abundance.2010, 4)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. terrestris/lucorum")), x = NULL, y = NULL)
+bssFL_2011 <- plot(sm(weighted_fl_abundance.2011, 4)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. terrestris/lucorum")), x = NULL, y = NULL)
+bssFL_2012 <- plot(sm(weighted_fl_abundance.2012, 4)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. terrestris/lucorum")), x = NULL, y = NULL)
+
+gersFL_2010 <- plot(sm(weighted_fl_abundance.2010, 5)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. gerstaeckeri")), x = NULL, y = NULL)
+gersFL_2011 <- plot(sm(weighted_fl_abundance.2011, 5)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. gerstaeckeri")), x = NULL, y = NULL) 
+gersFL_2012 <- plot(sm(weighted_fl_abundance.2012, 5)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. gerstaeckeri")), x = NULL, y = NULL)
+
+hortFL_2010 <- plot(sm(weighted_fl_abundance.2010, 6)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. hortorum")), x = NULL, y = NULL)
+hortFL_2011 <- plot(sm(weighted_fl_abundance.2011, 6)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. hortorum")), x = NULL, y = NULL)
+hortFL_2012 <- plot(sm(weighted_fl_abundance.2012, 6)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. hortorum")), x = NULL, y = NULL)
+
+humiFL_2010 <- plot(sm(weighted_fl_abundance.2010, 7)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. humilis")), x = NULL, y = NULL)
+humiFL_2011 <- plot(sm(weighted_fl_abundance.2011, 7)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. humilis")), x = NULL, y = NULL)
+humiFL_2012 <- plot(sm(weighted_fl_abundance.2012, 7)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. humilis")), x = NULL, y = NULL)
+
+hypnFL_2010 <- plot(sm(weighted_fl_abundance.2010, 8)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. hypnorum")), x = NULL, y = NULL)
+hypnFL_2011 <- plot(sm(weighted_fl_abundance.2011, 8)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. hypnorum")), x = NULL, y = NULL)
+hypnFL_2012 <- plot(sm(weighted_fl_abundance.2012, 8)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. hypnorum")), x = NULL, y = NULL)
+
+joneFL_2010 <- plot(sm(weighted_fl_abundance.2010, 9)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. jonellus")), x = NULL, y = NULL)
+joneFL_2011 <- plot(sm(weighted_fl_abundance.2011, 9)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. jonellus")), x = NULL, y = NULL)
+joneFL_2012 <- plot(sm(weighted_fl_abundance.2012, 9)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. jonellus")), x = NULL, y = NULL)
+
+lapiFL_2010 <- plot(sm(weighted_fl_abundance.2010, 10)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. lapidarius")), x = NULL, y = NULL)
+lapiFL_2011 <- plot(sm(weighted_fl_abundance.2011, 10)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. lapidarius")), x = NULL, y = NULL)
+lapiFL_2012 <- plot(sm(weighted_fl_abundance.2012, 10)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. lapidarius")), x = NULL, y = NULL)
+
+mendFL_2010 <- plot(sm(weighted_fl_abundance.2010, 11)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. mendax")), x = NULL, y = NULL)
+mendFL_2011 <- plot(sm(weighted_fl_abundance.2011, 11)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. mendax")), x = NULL, y = NULL)
+mendFL_2012 <- plot(sm(weighted_fl_abundance.2012, 11)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. mendax")), x = NULL, y = NULL)
+
+montFL_2010 <- plot(sm(weighted_fl_abundance.2010, 12)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. monticola")), x = NULL, y = NULL)
+montFL_2011 <- plot(sm(weighted_fl_abundance.2011, 12)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. monticola")), x = NULL, y = NULL)
+montFL_2012 <- plot(sm(weighted_fl_abundance.2012, 12)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. monticola")), x = NULL, y = NULL)
+
+muciFL_2010 <- plot(sm(weighted_fl_abundance.2010, 13)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. mucidis")), x = NULL, y = NULL)
+muciFL_2011 <- plot(sm(weighted_fl_abundance.2011, 13)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. mucidis")), x = NULL, y = NULL)
+muciFL_2012 <- plot(sm(weighted_fl_abundance.2012, 13)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. mucidis")), x = NULL, y = NULL)
+
+pascFL_2010 <- plot(sm(weighted_fl_abundance.2010, 14)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. pascuorum")), x = NULL, y = NULL)
+pascFL_2011 <- plot(sm(weighted_fl_abundance.2011, 14)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. pascuorum")), x = NULL, y = NULL)
+pascFL_2012 <- plot(sm(weighted_fl_abundance.2012, 14)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. pascuorum")), x = NULL, y = NULL)
+
+pratFL_2010 <- plot(sm(weighted_fl_abundance.2010, 15)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. pratorum")), x = NULL, y = NULL)
+pratFL_2011 <- plot(sm(weighted_fl_abundance.2011, 15)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. pratorum")), x = NULL, y = NULL)
+pratFL_2012 <- plot(sm(weighted_fl_abundance.2012, 15)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. pratorum")), x = NULL, y = NULL)
+
+psitFL_2010 <- plot(sm(weighted_fl_abundance.2010, 16)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. psithyrus")), x = NULL, y = NULL)
+psitFL_2011 <- plot(sm(weighted_fl_abundance.2011, 16)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. psithyrus")), x = NULL, y = NULL)
+psitFL_2012 <- plot(sm(weighted_fl_abundance.2012, 16)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. psithyrus")), x = NULL, y = NULL)
+
+pyreFL_2010 <- plot(sm(weighted_fl_abundance.2010, 17)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. pyrenaeus")), x = NULL, y = NULL)
+pyreFL_2011 <- plot(sm(weighted_fl_abundance.2011, 17)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. pyrenaeus")), x = NULL, y = NULL)
+pyreFL_2012 <- plot(sm(weighted_fl_abundance.2012, 17)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. pyrenaeus")), x = NULL, y = NULL)
+
+soroFL_2010 <- plot(sm(weighted_fl_abundance.2010, 18)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. soroensis")), x = NULL, y = NULL)
+soroFL_2011 <- plot(sm(weighted_fl_abundance.2011, 18)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. soroensis")), x = NULL, y = NULL)
+soroFL_2012 <- plot(sm(weighted_fl_abundance.2012, 18)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. soroensis")), x = NULL, y = NULL)
+
+wurfFL_2010 <- plot(sm(weighted_fl_abundance.2010, 19)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. wurflenii")), x = NULL, y = NULL)
+wurfFL_2011 <- plot(sm(weighted_fl_abundance.2011, 19)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. wurflenii")), x = NULL, y = NULL)
+wurfFL_2012 <- plot(sm(weighted_fl_abundance.2012, 19)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  #l_points() +
+  labs(title = expression(italic("B. wurflenii")), x = NULL, y = NULL)
+
+major_sp <- gridPrint(bssFL_2010, bssFL_2011, bssFL_2012, 
+                      pascFL_2010, pascFL_2011, pascFL_2012,
+                      pratFL_2010, pratFL_2011, pratFL_2012,
+                      soroFL_2010, soroFL_2011, soroFL_2012,
+                      wurfFL_2010, wurfFL_2011, wurfFL_2012, ncol = 3)
+
+high_elev_sp <- gridPrint(pyreFL_2010, pyreFL_2011, pyreFL_2012, 
+                          mendFL_2010, mendFL_2011, mendFL_2012,
+                          montFL_2010, montFL_2011, montFL_2012, ncol = 3)
+
+minor_sp <- gridPrint(gersFL_2010, gersFL_2011, gersFL_2012,
+                      hortFL_2010, hortFL_2011, hortFL_2012,
+                      #humiFL_2010, humiFL_2011, humiFL_2012, # too rarely seen, I think only 3 times
+                      hypnFL_2010, hypnFL_2011, hypnFL_2012,
+                      joneFL_2010, joneFL_2011, joneFL_2012,
+                      lapiFL_2010, lapiFL_2011, lapiFL_2012,
+                      #psitFL_2010, psitFL_2011, psitFL_2012, # Let's ignore 
+                      muciFL_2010, muciFL_2011, muciFL_2012, ncol = 3)
+plot(major_sp)
+plot(high_elev_sp)
+plot(minor_sp)
+
 ```
@@
 
-### Now let's model floral richness, but considering only species with preference > 0
+### Now let's model floral richness, but considering only species with preference > 0.
+This is very consistent year-to-year, so we can get away with a single model.
+In fact, AIC favors the global-only model, so we can make a case for not considering species differences (indeed, they appear to be minimal, with the exception of a couple of rare species (gers, humi) that we need to take with a grain of salt anyway).
 ```{r}
@@
          bb.sp = factor(bb.sp), # convert bb.sp to a factor
-         management = factor(management)) # convert management to a factor
+         management = factor(management),
+         site = factor(site)) # convert management to a factor
 
-weighted_fl_rich <- bam(fl.rich.wt ~ bb.sp + year + te(yday, elev.mean, 
+fl.rich_I <- bam(fl.rich.wt ~ s(bb.sp, bs = "re") + 
+                 year +
+                 s(management, bs = "re") + 
+                 s(site, bs = "re") +
+                 te(yday, elev.mean, 
                     by = bb.sp,
                     bs= c("tp", "tp"), 
-                                                k=c(10, 10), m=1),
+                    k=c(6, 6), m=2),
                data = fl_rich_wt,
@@
                family = "quasipoisson",
+               select = TRUE,
                method = "fREML") %>% getViz()
 
-check.gamViz(weighted_fl_rich)
-summary(weighted_fl_rich)
-print(plot(weighted_fl_rich , allTerms = TRUE), pages = 4)
+check.gamViz(fl.rich)
+fl.rich.sum <- summary(fl.rich)
+print(plot(fl.rich, allTerms = TRUE), pages = 1)
+
+fl.rich_GI <- bam(fl.rich.wt ~ s(bb.sp, bs = "re") + 
+                 year +
+                 s(management, bs = "re") + 
+                 s(site, bs = "re") +
+                 te(yday, elev.mean, 
+                    bs= c("tp", "tp"), 
+                    k=c(6, 6), m=2) +
+                 te(yday, elev.mean, 
+                    by = bb.sp,
+                    bs= c("tp", "tp"), 
+                    k=c(6, 6), m=2),
+               data = fl_rich_wt,
+               discrete = TRUE,
+               family = "quasipoisson",
+               select = TRUE,
+               method = "fREML") %>% getViz()
+
+check.gamViz(fl.rich)
+fl.rich_GI.sum <- summary(fl.rich_GI)
+print(plot(fl.rich_GI, allTerms = TRUE), pages = 1)
+
+fl.rich_G <- bam(fl.rich.wt ~ s(bb.sp, bs = "re") + 
+                 year +
+                 s(management, bs = "re") + 
+                 s(site, bs = "re") +
+                 te(yday, elev.mean, 
+                    bs= c("tp", "tp"), 
+                    k=c(6, 6), m=2),
+               data = fl_rich_wt,
+               discrete = TRUE,
+               family = "quasipoisson",
+               select = TRUE,
+               method = "fREML") %>% getViz()
+
+check.gamViz(fl.rich_G)
+fl.rich_G.sum <- summary(fl.rich_G)
+print(plot(fl.rich_G, allTerms = TRUE), pages = 1)
+
+AIC(fl.rich_GI, fl.rich_I, fl.rich_G)
 ```
+
+### Does preference relate to BB-FL traits? This serves mainly as a sanity check for my floral preference assessment.
+```{r}
+bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
+  dplyr::select(subgenus, bb.sp, pbl.w, pbl.w.class)
+
+fl_ktype <- read.csv("../processed_data/floral_k_type_nom_fix.csv") %>%
+  dplyr::select(plant.sp, k.type) %>%
+  mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"),
+         k.type.ss = str_extract(k.type, "[-+]?[0-9]*"))
+
+fl_morph <- read.csv("../processed_data/flower_morphology_GitaBenadi.csv") %>%
+  rename(plant.sp = Species) %>%
+  group_by(plant.sp) %>%
+  summarize(corolla.depth = mean(LN))
+
+fl_tax <- read_csv("../processed_data/floral_tax.csv")
+
+pref_traits <- preference %>%
+  left_join(bb_traits) %>%
+  left_join(fl_ktype) %>%
+  left_join(fl_morph) %>%
+  separate(plant.sp, c("plant.genus", "plant.species"), sep = " ", remove = FALSE) %>%
+  left_join(fl_tax) %>%
+  mutate(preference.bin = case_when(
+    preference > 0 ~ 1,
+    preference == 0 ~ 0
+  )) %>%
+  mutate(tongue.mismatch = abs(pbl.w - corolla.depth),
+         bb.sp = factor(bb.sp),
+         plant.sp = factor(plant.sp),
+         plant.genus = factor(plant.genus),
+         subgenus = factor(subgenus),
+         k.type.s = factor(k.type.s),
+         plant.family = factor(plant.family))
+
+### To deal with zero inflation, binarize and do binomial regression
+binomial_smooth <- function(...) {
+  geom_smooth(method = "glm", method.args = list(family = "binomial"), ...)
+}
+
+ggplot(pref_traits, aes(tongue.mismatch, preference.bin)) + 
+  geom_point() +
+  binomial_smooth()
+
+### Model: the binomial model might be the best bet
+library(DHARMa)
+library(lmerTest)
+lmer_bin <- glmer(preference.bin ~ tongue.mismatch + 
+                       (1 | plant.sp) +
+                       (1 | bb.sp), 
+                 family = "binomial",
+                 data = pref_traits)
+
+summary(lmer_bin)
+plot(lmer_bin)
+simulationOutput_bin <- simulateResiduals(fittedModel = lmer_bin, plot = T)
+testResiduals(simulationOutput_bin)
+plot(simulationOutput_bin)
+```

### 2020-07-27 analyses/network_analysis.Rmd
Refining GAMs, exploring global hierarchical specification

--- 2020-07-16 analyses/network_analysis.Rmd
@@
 library(mgcViz)
+library(piecewiseSEM)
+library(GGally)
 library(tidyverse)
@@
       frame2webs() %>% # convert to to bipartite web
-      map(networklevel, level = "higher", weighted = TRUE) %>% # We probably care mainly about the BB level of the network
+      map(networklevel, level = "higher", weighted = TRUE, index = c("niche overlap")) %>% # We probably care mainly about the BB level of the network
       data.frame() %>%
@@
 
+### Analyze sampling patterns
+```{r}
+samples <- net %>%
+  mutate(yday = yday(date)) %>%
+  group_by(site, year, yday, gdd.cum, elev.mean) %>%
+  summarize(obs = n())
+
+ggplot(samples, aes(yday, elev.mean, size = obs)) +
+  geom_point(alpha = 0.25) + 
+  facet_wrap(~year)
+
+ggplot(samples, aes(gdd.cum, elev.mean, size = obs)) +
+  geom_point(alpha = 0.25) + 
+  facet_wrap(~year)
+
+### Can we make an argument for salvaging individual years, or parts of year?
+#### 2011
+ggplot(filter(samples, year == 2011), aes(elev.mean, obs)) +
+  geom_point() +
+  geom_smooth()
+```
+
 ### Rarefied network analysis (not broken down by phase)
@@
   summarize(min_transects = min(transects))
+```
+
+### The problem with calculating network metrics on data pooled in time, particularly when you know that there is strong diversity and abundance structure in time, is that a metric like niche overlap becomes virtually uninterpretable, at least if competition is the interpretation in view.
+Bee 1 and Bee 2 could visit the same flower in different weeks and different flowers in the same week.
+In the pooled data, they might appear to have high niche overlap, but in fact they were never visiting the same flower at the same time, so there is no competition.
+Perhaps other metrics, like nestedness, or other interpretations -- not competition -- fare better with temporally aggregated data.
+So, I can consider those options.
+But, if I want to target an interpretation like competition through a metric like niche overlap, I have to control for time.
+The problem then becomes that so many of my site-dates have only a few observations, and network structure becomes uninformative (as far as I can reason) when interaction density becomes very small.
+You're just measuring so little of the structure that you intend to interpret.
+So, I've played around with the idea of setting some lower bound to drop the site-dates with too few observations, but this is inevitably arbitrary, and so many of my site-dates have very few observations that even a modest threshold culls a large number of site dates.
+What if I took the opposite approach, and targeted only the date(s) for each site with maximal observation density, then modeled those by elevation?
+The result would probably be very similar to the pooled analysis I have been doing, since pooled data becomes dominated by site-dates with high observation density.
+Really, all this would do is make transparent the fact that I am not controlling for time when I pool the data, I'm just masking a structure in which a few time points dominate for each site, and which time points dominate varies by site.
+So, this is an honest form of what I've already been doing, with no need to bootstrap.
+```{r}
+net.obs <- net %>%
+  group_by(site, date) %>%
+  summarize(total.obs = n()) %>%
+  ungroup() %>%
+  mutate(year = factor(year(date)))
+
+ggplot(net.obs, aes(total.obs)) +
+  geom_histogram() +
+  facet_grid(site ~ year)
+
+net.obs.top <- net.obs %>%
+  group_by(site, year) %>%
+  top_n(1, total.obs)
+
+web.top <- net %>%
+  semi_join(net.obs.top) %>%
+  unite(site.year.date, c(site, year, date), sep = "_") %>%
+  group_by(site.year.date, bb.sp, plant.sp) %>%
+  summarize(freq = n()) %>%
+  group_by(site.year.date) %>%
+  mutate(bb.rich = length(unique(bb.sp)),
+         plant.rich = length(unique(plant.sp))) %>%
+  filter(bb.rich >= 2 & plant.rich >= 2) %>%
+  select(higher = bb.sp, lower = plant.sp, webID = site.year.date, freq) %>%
+  data.frame() %>%
+  frame2webs() %>%
+  map(networklevel, level = "higher", weighted = TRUE, index = c("niche overlap", "weighted NODF")) %>% # We probably care mainly about the BB level of the network
+  data.frame() %>%
+  t() %>%
+  data.frame() %>%
+  rownames_to_column() %>%
+  dplyr::select(site.year.date = rowname, everything()) %>%
+  separate(site.year.date, c("site", "year", "date"), sep = "_") %>%
+  as_tibble() %>%
+  left_join(site_data)
+
+ggplot(web.top, aes(elev.mean, niche.overlap.HL, color = year)) +
+  geom_point() +
+  geom_smooth() +
+  theme_abyss() +
+  facet_wrap(~year)
+
+ggplot(web.top, aes(elev.mean, niche.overlap.HL)) +
+  geom_point(color = "goldenrod") +
+  geom_smooth() +
+  theme_abyss() +
+  facet_wrap(~year)
 ```
@@
   mutate(year = factor(rep("2012", n())))
-netmet_boot_2011 <- boot_net(net_2011, n = 4, iter = 40)  %>%
+netmet_boot_2011 <- boot_net(net_2011, n = 7, iter = 40)  %>% # Note that this results in site t5 being dropped, since it had only 4 transects.
+I think it worth dropping problematic t5 to improve overall quality
   mutate(year = factor(rep("2011", n())))
-netmet_boot_2010 <- boot_net(net_2010, n = 3, iter = 40) %>%
-  mutate(year = factor(rep("2010", n())))
+# Due to the unevenness of temporal interval of sampling in 2010, I don't think my bootstrapping approach handles  the embuggerance sufficiently.
+# netmet_boot_2010 <- boot_net(net_2010, n = 3, iter = 40) %>%
+#   mutate(year = factor(rep("2010", n())))
 
@@
   mutate(year = factor(rep("2012", n())))
-spmet_2011 <- boot_net_sp(net_2011, 4, 40) %>%
+spmet_2011 <- boot_net_sp(net_2011, 7, 40) %>%
   mutate(year = factor(rep("2011", n())))  
-spmet_2010 <- boot_net_sp(net_2010, 3, 40) %>%
-  mutate(year = factor(rep("2010", n())))
+# spmet_2010 <- boot_net_sp(net_2010, 3, 40) %>%
+#   mutate(year = factor(rep("2010", n())))
 
-spmet <- bind_rows(spmet_2012, spmet_2011, spmet_2010) %>%
+spmet <- bind_rows(spmet_2012, spmet_2011) %>%
   left_join(site_BB_aux.sp) %>%
@@
 netmet <- bind_rows(netmet_boot_2012, 
-                    netmet_boot_2011, 
-                    netmet_boot_2010) %>%
+                    netmet_boot_2011) %>%
+  ungroup() %>%
+  spread(metric, value) %>%
+  left_join(spmet_mean, by = c("site", "year")) %>%
+  left_join(site_BB_aux, by = c("site", "year")) %>%
+  inner_join(site_FL_aux, by = c("site", "year")) %>% # one site-year is missing from floral survey data
+  left_join(site_data, by = "site") %>%
+  mutate(year = factor(year),
+         site = factor(site),
+         management = factor(management))
+
+netmet_2011 <- netmet_boot_2011 %>%
   ungroup() %>%
@@
 ```{r}
-library(GGally)
-
 predictors <- netmet %>%
@@
 
-ggplot(netmet, aes(elev.mean, niche.overlap.HL)) +
+ggplot(filter(netmet, year == 2011), aes(elev.mean, niche.overlap.HL)) +
   geom_point() +
+  geom_smooth() +
   theme_modern()
@@
                      s(site, bs = "re") +
-                     management +
+                     #management +
                      s(elev.mean, k = 5) +
                      s(bb.rich, k = 5) +
-                     s(fl.rich, k = 5) +
-                     s(log.fl.abund, k = 5) +
-                     s(log.bb.abund, k = 5),
+                     #s(log.bb.abund, k = 5) +
+                     #s(log.fl.abund, k = 5) +
+                     s(fl.rich, k = 5),
                    select = TRUE,
@@
 
-##### weighted NODF: the main takeaway here is that higher BB abundance = higher nestedness, and that could easily be a sampling effect more than an ecologicla effect.
+##### weighted NODF: the main takeaway here is that higher BB abundance = higher nestedness, and that could easily be a sampling effect more than an ecological effect.
 Nevertheless, there remains a significant linear partial effect of elevation.
@@
                            #management +
-                           s(elev.mean, k = 5) +
+                           s(elev.mean, by = year, k = 3) +
                            #s(bb.rich, k = 5) +
-                           s(fl.rich, k = 5) +
-                           #s(log.fl.abund, k = 5) +
-                           s(log.bb.abund, k = 5),
+                           s(fl.rich, by = year, k = 3) +
+                          # s(log.fl.abund, k = 5) +
+                           s(log.bb.abund, by = year, k = 3),
                           select = TRUE,
@@
                            #management +
-                           s(elev.mean, k = 5) +
-                           s(bb.rich, k = 5) +
+                           #s(elev.mean, k = 5) +
+                           #s(bb.rich, k = 5) +
                            #s(log.fl.abund, k = 5) +
-                           #s(log.bb.abund, k = 5) +
+                           s(log.bb.abund, k = 5) +
                            s(fl.rich, k = 5),
@@
                            #management +
-                           #s(elev.mean, k = 5) +
+                           s(elev.mean, k = 5) +
                            #s(bb.rich, k = 5) +
                            #s(log.fl.abund, k = 5) +
-                           #s(fl.rich, k = 5) +
+                           s(fl.rich, k = 5) +
                            s(log.bb.abund, k = 5),
@@
 
-ggplot(netmet, aes(elev.mean, d, color = year)) +
+ggplot(filter(netmet, year == 2011), aes(elev.mean, d)) +
   geom_point() +
   geom_smooth() +
-  theme_abyss()
+  theme_modern()
+```
+
+### Since we clearly have structured collinearity in our preditors, i.e.
+elevation shapes abundance and diversity which shape network structure, perhaps SEM could be appropriate?
+The main question is whether it can handle the potentially complex (nonlinear) relationships between elevation and diversity/abundance, and whether it can handle the hierarchical structure of my data.
+```{r}
+# Create scales variables; this is important since SEM is based on covariance (right?)
+# For starters, let's just try to model niche.overlap
+netmet.scaled <- netmet %>%
+  select(site, year, niche.overlap.HL, elev.mean, fl.rich, bb.rich, log.fl.abund, log.bb.abund) %>%
+  mutate(elev.mean.2 = elev.mean^2,
+         elev.mean.3 = elev.mean^3) %>%
+  mutate(niche.overlap.HL = scale(niche.overlap.HL),
+         elev.mean = scale(elev.mean),
+         elev.mean.2 = scale(elev.mean),
+         elev.mean.3 = scale(elev.mean^3),
+         fl.rich = scale(fl.rich),
+         log.fl.abund = scale(log.fl.abund),
+         log.bb.abund = scale(log.bb.abund)) %>%
+  as.matrix() %>%
+  as_tibble() %>%
+  mutate(niche.overlap.HL = as.numeric(niche.overlap.HL),
+         elev.mean = as.numeric(elev.mean),
+         elev.mean.2 = as.numeric(elev.mean^2),
+         elev.mean.3 = as.numeric(elev.mean^3),
+         fl.rich = as.numeric(fl.rich),
+         bb.rich = as.numeric(bb.rich),
+         log.fl.abund = as.numeric(log.fl.abund),
+         log.bb.abund = as.numeric(log.bb.abund),
+         site = factor(site),
+         year = factor(year)) %>%
+  na.omit()
+
+# Pairs plot to determine best model form
+predictors.scaled <- netmet.scaled %>%
+  select(elev.mean, elev.mean.2, elev.mean.3, bb.rich, log.bb.abund, fl.rich, log.fl.abund, niche.overlap.HL)
+
+ggpairs(predictors.scaled)
+
+# Diversity and abundance metrics clearly have a quadratic relationship with elevation
+
+# Call models
+# To avoid figuring out how to handle year and site as random effects, let's start with separate year models
+netmet.scaled.2011 <- filter(netmet.scaled, year == 2011)
+
+mod_niche.overlap.2011 <- psem(
+  lm(niche.overlap.HL ~ elev.mean + elev.mean.2 + elev.mean.3 + fl.rich + log.bb.abund, 
+     netmet.scaled.2011),
+  lm(fl.rich ~ elev.mean + elev.mean.2, 
+     netmet.scaled.2011),
+  lm(log.bb.abund ~ elev.mean + elev.mean.2 + fl.rich, 
+     netmet.scaled.2011)
+  )
+
+summary(mod_niche.overlap.2011)
+plot(mod_niche.overlap.2011)
 ```
@@
 
-gam_d.sp_GI.2011 <- gam(d ~ management +
+gam_d.sp_GI.2011 <- gam(d ~ 
+                          #management +
                           s(bb.sp, bs = "re") +
@@
                           s(log.fl.abund, k = 8) +
-                          s(log.bb.sp.abund, k = 5),
+                          s(bb.sp.abund, k = 5),
                         select = TRUE,
@@
                         method = "REML") %>% getViz()
+
+ggplot(filter(spmet, year == 2011), aes(elev.mean, d)) +
+  geom_point() +
+  geom_smooth() +
+  theme_modern() +
+  facet_wrap(~bb.sp)
 

### 2020-09-09 analyses/FL_abund_GAM.Rmd
update

@@
 
-** One thing I need to go back and reconsider is that right now, I think forbidden links (i.e.
-links between species that never co-occurred) are given a zero for preference.
-A better approach would be to set their preference to a neutral value, which would be 1 in my case, since exp(0) = 1 and my preference values are calculated as exp(residual) of my linear model of visitation ~ abundance.
-**
+### Would these models be better specified using gamm4? Would that let me handle the crossed and nested random effects better?
 
@@
 ```{r}
+### A gamm4 version
+weighted_fl_abundance.gamm4 <- gamm4(log(fl.abund.wt + 0.001) ~
+                                       t2(yday, elev.mean, 
+                                          by = bb.sp,
+                                          #id = 0,
+                                          bs= c("gp", "tp"), 
+                                          k=c(8, 8), 
+                                          m=1),
+                                     random = ~(1 | site:year),
+                                     data = fl_abund_wt,
+                                     discrete = TRUE,
+                                     family = "gaussian",
+                                     select = TRUE,
+                                     method = "fREML") %>% getViz(nsim = 500)
+
 ### Global model
+#### My concern with using a global model is that, because I have so much higher sampling density in 2012, the global model will basically be just a 2012 model with 2010 and 2011 in the residuals, so to speak.
+I think this is at least partly mitigated by included a global by = year smooth, but I still don't quite understand how hierarchical effects relate to one another in a GAM.
+I'm also concerned that if I include the by = year smooth, it will be too concurve with my by = bb.sp smooths, and it will be hard to interpret the pattern of resource availability per species.
 weighted_fl_abundance.global <- bam(log(fl.abund.wt + 0.001) ~
@@
                                                 bs= c("gp", "tp"), 
-                                                k=c(8, 8), 
+                                                k=c(5, 5), 
                                                 m=2) +
@@
                                                 bs= c("gp", "tp"), 
-                                                k=c(8, 8), 
+                                                k=c(10, 10), 
                                                 m=1),
@@
 
-weighted_fl_abundance.global3 <- bam(log(fl.abund.wt + 0.001) ~
-                                      s(bb.sp, bs = "re") +
-                                      #s(management, bs = "re") +
-                                      s(year, bs = "re") +
-                                      s(site, bs = "re") +
-                                      te(yday, elev.mean, 
-                                              id = 0,
-                                              bs= c("gp", "tp"), 
-                                              k=c(10, 10), 
-                                              m=2) +
-                                      te(yday, elev.mean, 
-                                                by = bb.sp,
-                                                #id = 0,
-                                                bs= c("gp", "tp"), 
-                                                k=c(10, 10), 
-                                                m=1),
-                                    data = fl_abund_wt,
-                                    discrete = TRUE,
-                                    family = "gaussian",
-                                    select = TRUE,
-                                    method = "fREML") %>% getViz(nsim = 500)
-
-weighted_fl_abundance.global4 <- bam(log(fl.abund.wt + 0.001) ~
-                                      s(bb.sp, bs = "re") +
-                                      #s(management, bs = "re") +
-                                      s(year, bs = "re") +
-                                      s(site, bs = "re") +
-                                      te(yday, elev.mean, year,
-                                                by = bb.sp,
-                                                #id = 0,
-                                                bs= c("gp", "tp", "re"), 
-                                                k=c(5, 5), 
-                                                m=1),
-                                    data = fl_abund_wt,
-                                    discrete = TRUE,
-                                    family = "gaussian",
-                                    select = TRUE,
-                                    method = "fREML") %>% getViz(nsim = 500)
-
-AIC(weighted_fl_abundance.global, weighted_fl_abundance.global2, weighted_fl_abundance.global3, weighted_fl_abundance.global4)
+AIC(weighted_fl_abundance.global, weighted_fl_abundance.global2)
 
@@
 weighted_fl_abundance.global.sum2 <- summary(weighted_fl_abundance.global2)
-weighted_fl_abundance.global.sum3 <- summary(weighted_fl_abundance.global3)
-weighted_fl_abundance.global.sum3 <- summary(weighted_fl_abundance.global4)
 
@@
 print(plot(weighted_fl_abundance.global2, allTerms = TRUE), pages = 4)
-print(plot(weighted_fl_abundance.global3, allTerms = TRUE), pages = 4)
-print(plot(weighted_fl_abundance.global4, allTerms = TRUE), pages = 4)
 
@@
 
+weighted_fl_abundance.global2.ck <- check2D(weighted_fl_abundance.global2, 
+                                         x1 = "yday", x2 = "elev.mean")
+weighted_fl_abundance.global2.ck + l_gridCheck2D(gridFun = mean)
+
 ### 2010
-weighted_fl_abundance.2010 <- bam(fl.abund.wt ~ s(bb.sp, bs = "re") + 
-                                    management + 
-                                    #slope.est +
-                                    #s(site, bs = "re") +
+weighted_fl_abundance.2010 <- bam(log(fl.abund.wt + 0.001) ~
+                                      s(bb.sp, bs = "re") +
+                                      #s(management, bs = "re") +
+                                      s(site, bs = "re") +
                                       te(yday, elev.mean, 
                                                 by = bb.sp,
+                                                #id = 0,
                                                 bs= c("gp", "tp"), 
-                                       k=c(10, 10), m=2),
+                                                k=c(10, 10), 
+                                                m=2),
                                     data = filter(fl_abund_wt, year == 2010),
                                     discrete = TRUE,
-                                  family = "tw",
+                                    family = "gaussian",
                                     select = TRUE,
@@
 weighted_fl_abundance.2010.sum <- summary(weighted_fl_abundance.2010)
-print(plot(weighted_fl_abundance.2010, allTerms = TRUE), pages = 1)
+print(plot(weighted_fl_abundance.2010, allTerms = TRUE), pages = 4)
 weighted_fl_abundance.2010.ck <- check2D(weighted_fl_abundance.2010, 
@@
 ### 2011
-weighted_fl_abundance.2011 <- bam(fl.abund.wt ~ s(bb.sp, bs = "re") + 
-                                    #s(transect, bs = "re") +
-                                    management + 
-                                    #slope.est +
-                                    #s(site, bs = "re") +
+weighted_fl_abundance.2011 <- bam(log(fl.abund.wt + 0.001) ~
+                                      s(bb.sp, bs = "re") +
+                                      #s(management, bs = "re") +
+                                      s(site, bs = "re") +
                                       te(yday, elev.mean, 
                                                 by = bb.sp,
+                                                #id = 0,
                                                 bs= c("gp", "tp"), 
-                                       k=c(10, 10), m=2),
+                                                k=c(10, 10), 
+                                                m=2),
                                     data = filter(fl_abund_wt, year == 2011),
                                     discrete = TRUE,
-                                  family = "tw",
+                                    family = "gaussian",
                                     select = TRUE,
@@
 weighted_fl_abundance.2011.sum <- summary(weighted_fl_abundance.2011)
-print(plot(weighted_fl_abundance.2011, allTerms = TRUE), pages = 1)
+print(plot(weighted_fl_abundance.2011, allTerms = TRUE), pages = 4)
 weighted_fl_abundance.2011.ck <- check2D(weighted_fl_abundance.2011, 
@@
 ### 2012
-weighted_fl_abundance.2012 <- bam(fl.abund.wt ~ s(bb.sp, bs = "re") + 
-                                    #s(transect, bs = "re") +
-                                    management + 
-                                    #slope.est +
+weighted_fl_abundance.2012 <- bam(log(fl.abund.wt + 0.001) ~
+                                      s(bb.sp, bs = "re") +
+                                      #s(management, bs = "re") +
                                       s(site, bs = "re") +
@@
                                                 by = bb.sp,
+                                                #id = 0,
                                                 bs= c("gp", "tp"), 
-                                       k=c(10, 10), m=1),
+                                                k=c(10, 10), 
+                                                m=2),
                                     data = filter(fl_abund_wt, year == 2012),
                                     discrete = TRUE,
-                                  family = "tw",
+                                    family = "gaussian",
                                     select = TRUE,
@@
 weighted_fl_abundance.2012.sum <- summary(weighted_fl_abundance.2012)
-print(plot(weighted_fl_abundance.2012, allTerms = TRUE), pages = 1)
+print(plot(weighted_fl_abundance.2012, allTerms = TRUE), pages = 4)
 weighted_fl_abundance.2012.ck <- check2D(weighted_fl_abundance.2012, 

### 2020-09-09 analyses/beta_diversity.Rmd
update

@@
 ```{r message=FALSE, warning=FALSE}
-library(vegan)
 library(lubridate)
+library(sf)
 library(gdm)
 library(bipartite)
-library(sf)
+library(vegan)
 library(mgcv)
 library(mgcViz)
+library(see)
 library(tidyverse)
@@
 ### Climate data
-```{r, include=FALSE, echo=FALSE}
+```{r echo=FALSE, message=FALSE, include=FALSE}
 gdd <- read_csv("../processed_data/climate.csv") %>%
@@
 
-### Load network and survey data and generate site.year and site.month.year webs
+### Load network and survey data
 ```{r}
@@
   dplyr::select(site, date, plant.sp, flower.cover)
+```
 
-site.year_transects <- net %>%
+### betaboot function: This bootstraps the partitioning of interaction beta-diversity to account for differences in the number of sampling dates for each site in the same way the my other bootstrapping function ("bootnet" or something like that) bootstraps the calculation of network metrics.
+Note that due to some annoying behavior by webs2array(), I have to define an object within the for loop  of my function using the <<- global assignment operator.
+This is bad coding, but I can't figure out another way to get webs2array() to find the object it needs to find.
+```{r message=FALSE}
+betaboot <- function(x, n, iter) {
+
+  out <- data.frame()
+
+  index <- 0
+
+  t_filter <- x %>%
+    dplyr::select(site, date, year) %>%
+    unique() %>%
     group_by(site, year) %>%
-  summarize(transects = length(unique(date)))
+    mutate(transects = n()) %>%
+    filter(transects >= n)
 
-site_transects <- net %>%
+  for(i in 1:iter){
+
+    index <- index + 1
+
+    net.sub <- x %>%
+      semi_join(t_filter) %>%
+      ungroup() %>%
+      dplyr::select(site, dayofyear) %>%
+      unique() %>%
       group_by(site) %>%
-  summarize(transects = length(unique(date)))
+      sample_n(n) %>%
+      left_join(x)
 
-web_site.year <- net %>%
-  group_by(site.year, bb.sp, plant.sp) %>%
-  summarize(freq = n()) %>%
-  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs() # convert to to bipartite web
-
-web_site <- net %>%
+    web <<- net.sub %>%
       group_by(site, bb.sp, plant.sp) %>%
@@
 
-web_site.date.2010 <- net %>%
-  filter(year == 2010) %>%
-  group_by(site.year.day, bb.sp, plant.sp) %>%
-  summarize(freq = n()) %>%
-  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year.day, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs() # convert to to bipartite web
+    array <- webs2array(web)
 
-web_site.date.2011 <- net %>%
-  filter(year == 2011) %>%
-  group_by(site.year.day, bb.sp, plant.sp) %>%
-  summarize(freq = n()) %>%
-  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year.day, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs() # convert to to bipartite web
+    beta <- betalinkr_multi(array, partition.st = TRUE, partition.rr = TRUE, binary = TRUE) %>%
+      as_tibble() %>%
+      mutate(iteration = rep(index, n()))
 
-web_site.date.2012 <- net %>%
-  filter(year == 2012) %>%
-  group_by(site.year.day, bb.sp, plant.sp) %>%
-  summarize(freq = n()) %>%
-  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year.day, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs() # convert to to bipartite web
+    out <- bind_rows(out, beta)
+  }
 
-array_site <- webs2array(web_site)
-array_site.year <- webs2array(web_site.year)
-array_site.date.2010 <- webs2array(web_site.date.2010)
-array_site.date.2011 <- webs2array(web_site.date.2011)
-array_site.date.2012 <- webs2array(web_site.date.2012)
+  out <- out %>%
+    as_tibble() %>%
+    dplyr::select(site1 = i, site2 = j, everything()) %>%
+    gather(metric, value, -c(site1, site2, iteration)) %>%
+    mutate(metric.class = case_when(
+        metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
+        metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
+        metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
+        metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
+      ),
+      iteration = factor(iteration))
+}
 ```
@@
 ST = beta_ST, the dissimilarity (component) explained by difference in species community composition 
-```{r}
-beta_site <- betalinkr_multi(array_site, partition.st = TRUE, partition.rr = TRUE) %>%
-  as_tibble() %>%
-  rename(site1 = i) %>%
-  rename(site2 = j)%>%
+```{r message=FALSE}
+net2010 <- filter(net, year == 2010)
+net2011 <- filter(net, year == 2011)
+net2012 <- filter(net, year == 2012)
+
+beta_2010 <- betaboot(net2010, 5, 50) %>%
+  group_by(site1, site2, metric, metric.class) %>%
+  summarize(value = mean(value)) %>%
   left_join(site_data, by = c("site1" = "site")) %>%
-  dplyr::select(site1, site2, S, OS, WN, ST, 
-         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
+  dplyr::select(site1, site2, metric, metric.class, value, 
                 elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
   left_join(site_data, by = c("site2" = "site")) %>%
-  dplyr::select(site1, site2, S, OS, WN, ST, 
-         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
+  dplyr::select(site1, site2, metric, metric.class, value, 
          elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
-  left_join(site_transects, by = c("site1" = "site")) %>%
-  dplyr::select(site1, site2, S, OS, WN, ST, 
-                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
-                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1 = transects) %>%
-  left_join(site_transects, by = c("site2" = "site")) %>%
-  dplyr::select(site1, site2, S, OS, WN, ST, 
-                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
-                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1, transects2 = transects) %>%
   mutate(elev.diff = abs(elev1 - elev2),
          mean.elev = (elev1 + elev2)/2,
-         transects.diff = abs(transects1 - transects2),
          man.diff = case_when(
@@
            man1 != man2 ~ TRUE
-         )) %>%
-  gather(metric, value, -c(site1, site2, elev1, elev2, elev.diff, 
-                           man1, man2, man.diff, mean.elev, lon1, 
-                           lat1, lon2, lat2, transects1, transects2, transects.diff)) %>%
-  mutate(metric.class = case_when(
-    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
-    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
-    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
-    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
-  ))
+         ),
+         year = factor(rep("2010", n())))
 
-#write_csv(beta_site, "../analyses/output/beta_site.csv")
-
-beta_site.year <- betalinkr_multi(array_site.year, partition.st = TRUE, partition.rr = TRUE, binary = TRUE) %>%
-  as_tibble() %>%
-  separate(i, into = c("site1", "year1"), sep = "_") %>%
-  separate(j, into = c("site2", "year2"), sep = "_") %>%
+beta_2011 <- betaboot(net2011, 5, 50) %>%
+  group_by(site1, site2, metric, metric.class) %>%
+  summarize(value = mean(value)) %>%
   left_join(site_data, by = c("site1" = "site")) %>%
-  dplyr::select(site1, year1, site2, year2, S, OS, WN, ST, 
-         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
+  dplyr::select(site1, site2, metric, metric.class, value, 
                 elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
   left_join(site_data, by = c("site2" = "site")) %>%
-  dplyr::select(site1, year1, site2, year2, S, OS, WN, ST, 
-         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
+  dplyr::select(site1, site2, metric, metric.class, value, 
          elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
-  left_join(site.year_transects, by = c("site1" = "site", "year1" = "year")) %>%
-  dplyr::select(site1, site2, year1, year2, S, OS, WN, ST, 
-                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
-                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1 = transects) %>%
-  left_join(site.year_transects, by = c("site2" = "site", "year2" = "year")) %>%
-  dplyr::select(site1, site2, year1, year2, S, OS, WN, ST, 
-                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
-                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1, transects2 = transects) %>%
   mutate(elev.diff = abs(elev1 - elev2),
          mean.elev = (elev1 + elev2)/2,
-         transects.diff = abs(transects1 - transects2),
          man.diff = case_when(
@@
            man1 != man2 ~ TRUE
-         )) %>%
-  gather(metric, value, -c(site1, year1, site2, year2, elev1, elev2, elev.diff, 
-                           man1, man2, man.diff, mean.elev, lon1, lat1, lon2, lat2,
-                           transects1, transects2, transects.diff)) %>%
-  mutate(metric.class = case_when(
-    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
-    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
-    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
-    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
-  ))
+         ),
+         year = factor(rep("2011", n())))
 
-### By individual transect
+beta_2012 <- betaboot(net2012, 9, 50) %>%
+  group_by(site1, site2, metric, metric.class) %>%
+  summarize(value = mean(value)) %>%
+  left_join(site_data, by = c("site1" = "site")) %>%
+  dplyr::select(site1, site2, metric, metric.class, value, 
+                elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
+  left_join(site_data, by = c("site2" = "site")) %>%
+  dplyr::select(site1, site2, metric, metric.class, value, 
+         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
+  mutate(elev.diff = abs(elev1 - elev2),
+         mean.elev = (elev1 + elev2)/2,
+         man.diff = case_when(
+           man1 == man2 ~ FALSE,
+           man1 != man2 ~ TRUE
+         ),
+         year = factor(rep("2012", n())))
 
+beta_agg <- bind_rows(beta_2010, beta_2011, beta_2012)
+
+#write_csv(beta_agg, "./output/aggregated_beta_diversity.csv")
+#beta_agg <- read_csv("./output/aggregated_beta_diversity.csv")
+```
+
+### Plot results. There's nothing a Mantel test can tell me that I don't already know from the combination of plotting and GDM analysis.
+```{r}
+### Elevation
+ggplot(filter(beta, year == "2010" & 
+                metric != "S" & 
+                metric.class != "OS.partition"), 
+       aes(elev.diff, value, color = metric)) +
+  geom_point(alpha = 0.45) +
+  geom_smooth(method = "lm", formula =  y ~ poly(x, 3)) +
+  facet_wrap(~metric.class) +
+  theme_abyss()
+
+ggplot(filter(beta, year == "2011" & 
+                metric != "S" & 
+                metric.class != "OS.partition"), 
+       aes(elev.diff, value, color = metric)) +
+  geom_point(alpha = 0.45) +
+  geom_smooth(method = "lm", formula =  y ~ poly(x, 3)) +
+  facet_wrap(~metric.class) +
+  theme_abyss()
+
+ggplot(filter(beta, year == "2012" & 
+                metric != "S" & 
+                metric.class != "OS.partition"), 
+       aes(elev.diff, value, color = metric)) +
+  geom_point(alpha = 0.45) +
+  geom_smooth(method = "lm", formula =  y ~ poly(x, 3)) +
+  facet_wrap(~metric.class) +
+  theme_abyss()
+
+ggplot(filter(beta, 
+                metric != "S" & 
+                metric.class != "OS.partition"), 
+       aes(elev.diff, value, color = metric)) +
+  geom_point(alpha = 0.45) +
+  geom_smooth(method = "lm", formula =  y ~ poly(x, 3)) +
+  facet_wrap(~metric.class) +
+  theme_gray()
+
+### Management
+ggplot(filter(beta, year == "2010" & 
+                metric != "S" & 
+                metric.class != "OS.partition"), 
+       aes(man.diff, value)) +
+  geom_boxplot() +
+  facet_wrap(~metric, scales = "free") +
+  theme_gray()
+
+ggplot(filter(beta, year == "2011" & 
+                metric != "S" & 
+                metric.class != "OS.partition"), 
+       aes(man.diff, value)) +
+  geom_boxplot() +
+  facet_wrap(~metric, scales = "free") +
+  theme_gray()
+
+ggplot(filter(beta, year == "2012" & 
+                metric != "S" & 
+                metric.class != "OS.partition"), 
+       aes(man.diff, value)) +
+  geom_boxplot() +
+  facet_wrap(~metric, scales = "free") +
+  theme_gray()
+```
+
+### Let's look at per-transect floral survey beta-diversity
+```{r}
+### 2012
+survey_bioData.2012 <- survey %>%
+  mutate(year = year(date),
+         yday = yday(date)) %>%
+  filter(flower.cover > 0 & year == "2012") %>%
+  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
+  group_by(site, yday, plant.sp) %>%
+  summarize(cover = sum(flower.cover)) %>%
+  spread(plant.sp, cover) %>%
+  replace(is.na(.), 0) %>%
+  arrange(site, yday) %>%
+  unite(site.day, site, yday, sep = "_")
+
+survey_predData.2012 <- survey %>%
+  mutate(year = year(date),
+         yday = yday(date)) %>%
+  filter(flower.cover > 0 & year == "2012") %>%
+  dplyr::select(site, yday) %>%
+  left_join(site_data) %>%
+  ungroup() %>%
+  dplyr::select(site, elev.mean, lat, lon, yday) %>%
+  unique() %>%
+  arrange(site, yday) %>%
+  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
+  dplyr::select(-site)
+
+survey_sitepair.2012 <- formatsitepair(survey_bioData.2012, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
+                              XColumn = "lon", YColumn = "lat", predData = survey_predData.2012)
+
+### Call gdm()
+gdm_survey.2012 <- gdm(survey_sitepair.2012, geo = TRUE)
+gdm.varImp(survey_sitepair.2012, geo = TRUE)
+
+summary(gdm_survey.2012)
+plot(gdm_survey.2012)
+plotUncertainty(survey_sitepair.2012, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
+
+### 2011
+survey_bioData.2011 <- survey %>%
+  mutate(year = year(date),
+         yday = yday(date)) %>%
+  filter(flower.cover > 0 & year == "2011") %>%
+  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
+  group_by(site, yday, plant.sp) %>%
+  summarize(cover = sum(flower.cover)) %>%
+  spread(plant.sp, cover) %>%
+  replace(is.na(.), 0) %>%
+  arrange(site, yday) %>%
+  unite(site.day, site, yday, sep = "_")
+
+predData.2011 <- survey %>%
+  mutate(year = year(date),
+         yday = yday(date)) %>%
+  filter(flower.cover > 0 & year == "2011") %>%
+  select(site, yday) %>%
+  left_join(site_data) %>%
+  ungroup() %>%
+  dplyr::select(site, elev.mean, lat, lon, yday) %>%
+  unique() %>%
+  arrange(site, yday) %>%
+  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
+  select(-site)
+
+survey_sitepair.2011 <- formatsitepair(survey_bioData.2011, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
+                              XColumn = "lon", YColumn = "lat", predData = predData.2011)
+
+### Call gdm()
+gdm_survey.2011 <- gdm(survey_sitepair.2011, geo = TRUE)
+#gdm.varImp(survey_sitepair, geo = TRUE)
+
+summary(gdm_survey.2011)
+plot(gdm_survey.2011)
+#plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
+
+### 2010
+survey_bioData.2010 <- survey %>%
+  mutate(year = year(date),
+         yday = yday(date)) %>%
+  filter(flower.cover > 0 & year == "2010") %>%
+  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
+  group_by(site, yday, plant.sp) %>%
+  summarize(cover = sum(flower.cover)) %>%
+  spread(plant.sp, cover) %>%
+  replace(is.na(.), 0) %>%
+  arrange(site, yday) %>%
+  unite(site.day, site, yday, sep = "_")
+
+predData.2010 <- survey %>%
+  mutate(year = year(date),
+         yday = yday(date)) %>%
+  filter(flower.cover > 0 & year == "2010") %>%
+  select(site, yday) %>%
+  left_join(site_data) %>%
+  ungroup() %>%
+  dplyr::select(site, elev.mean, lat, lon, yday) %>%
+  unique() %>%
+  arrange(site, yday) %>%
+  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
+  select(-site)
+
+survey_sitepair.2010 <- formatsitepair(survey_bioData.2010, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
+                              XColumn = "lon", YColumn = "lat", predData = predData.2010)
+
+### Call gdm()
+gdm_survey.2010 <- gdm(survey_sitepair.2010, geo = TRUE)
+#gdm.varImp(survey_sitepair, geo = TRUE)
+
+summary(gdm_survey.2010)
+plot(gdm_survey.2010)
+#plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
+```
+
+### And now Bombus diversity patterns
+```{r}
+### 2010
+BB_bioData.2010 <- net %>%
+  filter(year == 2010) %>%
+  group_by(site, date, bb.sp) %>%
+  summarize(count = n()) %>%
+  group_by(site, date) %>%
+  #mutate(prop = count/sum(count)) %>%
+  #group_by(site, bb.sp) %>%
+  # summarize(abund = sum(prop)) %>%
+  spread(bb.sp, count) %>%
+  replace(is.na(.), 0) %>%
+  mutate(yday = yday(date)) %>%
+  unite(site.day, site, yday, sep = "_", remove = "TRUE") %>%
+  ungroup() %>%
+  select(-date)
+
+BB_predData.2010 <- net %>%
+  filter(year == 2010) %>%
+  mutate(yday = yday(date)) %>%
+  dplyr::select(site, yday) %>%
+  left_join(site_data) %>%
+  ungroup() %>%
+  dplyr::select(site, elev.mean, lat, lon, yday) %>%
+  unique() %>%
+  arrange(site, yday) %>%
+  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
+  dplyr::select(-site)
+
+bb_sitepair.2010 <- formatsitepair(BB_bioData.2010, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
+                              XColumn = "lon", YColumn = "lat", predData = BB_predData.2010)
+
+### Call gdm()
+gdm_bb.2010 <- gdm(bb_sitepair.2010, geo = TRUE)
+gdm.varImp(bb_sitepair.2010, geo = TRUE)
+
+summary(gdm_bb.2010)
+plot(gdm_bb.2010)
+plotUncertainty(bb_sitepair.2010, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
+
+### 2011
+BB_bioData.2011 <- net %>%
+  filter(year == 2011) %>%
+  group_by(site, date, bb.sp) %>%
+  summarize(count = n()) %>%
+  group_by(site, date) %>%
+  #mutate(prop = count/sum(count)) %>%
+  #group_by(site, bb.sp) %>%
+  # summarize(abund = sum(prop)) %>%
+  spread(bb.sp, count) %>%
+  replace(is.na(.), 0) %>%
+  mutate(yday = yday(date)) %>%
+  unite(site.day, site, yday, sep = "_", remove = "TRUE") %>%
+  ungroup() %>%
+  select(-date)
+
+BB_predData.2011 <- net %>%
+  filter(year == 2011) %>%
+  mutate(yday = yday(date)) %>%
+  dplyr::select(site, yday) %>%
+  left_join(site_data) %>%
+  ungroup() %>%
+  dplyr::select(site, elev.mean, lat, lon, yday) %>%
+  unique() %>%
+  arrange(site, yday) %>%
+  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
+  dplyr::select(-site)
+
+bb_sitepair.2011 <- formatsitepair(BB_bioData.2011, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
+                              XColumn = "lon", YColumn = "lat", predData = BB_predData.2011)
+
+### Call gdm()
+gdm_bb.2011 <- gdm(bb_sitepair.2011, geo = TRUE)
+gdm.varImp(bb_sitepair.2011, geo = TRUE)
+
+summary(gdm_bb.2011)
+plot(gdm_bb.2011)
+plotUncertainty(bb_sitepair.2011, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
+
+### 2012
+BB_bioData.2012 <- net %>%
+  filter(year == 2012) %>%
+  group_by(site, date, bb.sp) %>%
+  summarize(count = n()) %>%
+  group_by(site, date) %>%
+  #mutate(prop = count/sum(count)) %>%
+  #group_by(site, bb.sp) %>%
+  # summarize(abund = sum(prop)) %>%
+  spread(bb.sp, count) %>%
+  replace(is.na(.), 0) %>%
+  mutate(yday = yday(date)) %>%
+  unite(site.day, site, yday, sep = "_", remove = "TRUE") %>%
+  ungroup() %>%
+  select(-date)
+
+BB_predData.2012 <- net %>%
+  filter(year == 2012) %>%
+  mutate(yday = yday(date)) %>%
+  dplyr::select(site, yday) %>%
+  left_join(site_data) %>%
+  ungroup() %>%
+  dplyr::select(site, elev.mean, lat, lon, yday) %>%
+  unique() %>%
+  arrange(site, yday) %>%
+  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
+  dplyr::select(-site)
+
+bb_sitepair.2012 <- formatsitepair(BB_bioData.2012, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
+                              XColumn = "lon", YColumn = "lat", predData = BB_predData.2012)
+
+### Call gdm()
+gdm_bb.2012 <- gdm(bb_sitepair.2012, geo = TRUE)
+gdm.varImp(bb_sitepair.2012, geo = TRUE)
+
+summary(gdm_bb.2012)
+plot(gdm_bb.2012)
+plotUncertainty(bb_sitepair.2012, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
+```
+
+### Now let's look at interaction beta diversity.
+I don't think p-values would add anything to the interpretation I get from the plotted uncertainty, so it's okay that I can't get gdm.varImp() to work.
+```{r}
+### 2010
+sitepair_2010 <- beta_2010 %>%
+  ungroup() %>%
+  filter(metric == "WN") %>%
+  mutate(weights = rep(1, n())) %>%
+  dplyr::select(distance = value, weights,
+         s1.xCoord = lon1, s1.yCoord = lat1, 
+         s2.xCoord = lon2, s2.yCoord = lat2, 
+         s1.elev.mean = elev1, s2.elev.mean = elev2) %>%
+  data.frame()
+
+gdm_2010 <- gdm(sitepair_2010, geo = TRUE)
+gdm.varImp(sitepair_2010, geo = TRUE)
+
+summary(gdm_2010)
+plot(gdm_2010)
+plotUncertainty(sitepair_2010, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
+
+### 2011
+sitepair_2011 <- beta_2011 %>%
+  ungroup() %>%
+  filter(metric == "WN") %>%
+  mutate(weights = rep(1, n())) %>%
+  dplyr::select(distance = value, weights,
+         s1.xCoord = lon1, s1.yCoord = lat1, 
+         s2.xCoord = lon2, s2.yCoord = lat2, 
+         s1.elev.mean = elev1, s2.elev.mean = elev2) %>%
+  data.frame()
+
+gdm_2011 <- gdm(sitepair_2011, geo = TRUE)
+gdm.varImp(sitepair_2011, geo = TRUE)
+
+summary(gdm_2011)
+plot(gdm_2011)
+plotUncertainty(sitepair_2011, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
+
+### 2012
+sitepair_2012 <- beta_2012 %>%
+  ungroup() %>%
+  filter(metric == "WN") %>%
+  mutate(weights = rep(1, n())) %>%
+  dplyr::select(distance = value, weights,
+         s1.xCoord = lon1, s1.yCoord = lat1, 
+         s2.xCoord = lon2, s2.yCoord = lat2, 
+         s1.elev.mean = elev1, s2.elev.mean = elev2) %>%
+  data.frame()
+
+gdm_2012 <- gdm(sitepair_2012, geo = TRUE)
+gdm.varImp(sitepair_2012, geo = TRUE)
+
+summary(gdm_2012)
+plot(gdm_2012)
+plotUncertainty(sitepair_2012, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
+```
+
+### Transect-level analysis
+```{r}
 beta_site.date.2010 <- betalinkr_multi(array_site.date.2010, 
@@
 write_csv(beta_site.date, "../analyses/output/beta_site.date.csv")
+beta_site.date <- read_csv("../analyses/output/beta_site.date.csv") %>%
+  mutate(year1 = factor(year1),
+         year2 = factor(year2),
+         metric = factor(metric))
 ```
 
-### Visualize transect-level beta-diversity
+### Plot transect-level analysis
 ```{r}
-ggplot(filter(beta_site.date.2010, metric != "S" & 
-                metric.class != "OS.partition"), 
-       aes(elev.diff, value, color = metric)) +
-  geom_point(alpha = 0.05) +
-  geom_smooth() +
-  facet_wrap(~metric.class)
-
-ggplot(filter(beta_site.date.2010, yday.diff <= 7 &
+### Elevation diff
+ggplot(filter(beta_site.date, year1 == year2,
                 metric != "S" & 
@@
        aes(elev.diff, value, color = metric)) +
-  geom_point(alpha = 0.05) +
+  geom_point(alpha = 0.15, size = 0.5) +
   geom_smooth() +
-  facet_wrap(~metric.class)
+  facet_grid(metric.class ~ year1) +
+  theme_abyss()
 
-ggplot(filter(beta_site.date.2011, yday.diff <= 7 &
+### Time diff
+ggplot(filter(beta_site.date, year1 == year2,
                 metric != "S" & 
                 metric.class != "OS.partition"), 
-       aes(elev.diff, value, color = metric)) +
-  geom_point(alpha = 0.05) +
+       aes(yday.diff, value, color = metric)) +
+  geom_point(alpha = 0.15, size = 0.5) +
   geom_smooth() +
-  facet_wrap(~metric.class)
+  facet_grid(metric.class ~ year1) +
+  theme_abyss()
 
-ggplot(filter(beta_site.date.2012, yday.diff <= 7 &
-                metric != "S" & 
-                metric.class != "OS.partition"), 
-       aes(elev.diff, value, color = metric)) +
-  geom_point(alpha = 0.05) +
-  geom_smooth() +
-  facet_wrap(~metric.class)
-
-ggplot(filter(beta_site.date.2012, yday.diff == 0 &
-                metric != "S" & 
-                metric.class != "OS.partition"),
-       aes(value)) +
-  geom_histogram()  +
-  facet_wrap(~metric)
-
-ggplot(filter(beta_site.date.2012, yday.diff <= 10 &
-                metric != "S" & 
-                metric.class != "OS.partition"),
-       aes(value)) +
-  geom_histogram()  +
-  facet_wrap(~metric)
-
-ggplot(filter(beta_site.date.2012, yday.diff <= 7 &
+### Elevation diff within 1 wk time diff
+ggplot(filter(beta_site.date, year1 == year2,
                 metric != "S" & 
                 metric.class != "OS.partition" &
-                site1 == site2),
+                yday.diff <= 7), 
+       aes(elev.diff, value, color = metric)) +
+  geom_point(alpha = 0.15, size = 0.5) +
+  geom_smooth() +
+  facet_wrap(~metric.class) +
+  theme_abyss()
+
+ggplot(filter(beta_site.date, year1 == year2,
+                metric != "S" & 
+                metric.class != "OS.partition" &
+                yday.diff <= 7), 
+       aes(elev.diff, value, color = metric)) +
+  geom_point(alpha = 0.15, size = 0.5) +
+  geom_smooth() +
+  facet_grid(metric.class ~ year1) +
+  theme_abyss()
+
+### Time diff within 250 m elev diff
+ggplot(filter(beta_site.date, year1 == year2,
+                metric != "S" & 
+                metric.class != "OS.partition" &
+                elev.diff <= 7), 
+       aes(yday.diff, value, color = metric)) +
+  geom_point(alpha = 0.15, size = 0.5) +
+  geom_smooth() +
+  facet_wrap(~metric.class) +
+  theme_light()
+
+ggplot(filter(beta_site.date, year1 == year2,
+                metric != "S" & 
+                metric.class != "OS.partition" &
+                elev.diff <= 7), 
+       aes(yday.diff, value, color = metric)) +
+  geom_point(alpha = 0.15, size = 0.5) +
+  geom_smooth() +
+  facet_grid(metric.class ~ year1) +
+  theme_abyss()
+
+### 2D ElevationDiff*TimeDiff
+gam_betadiv <- bamV(value ~ te(elev.diff, yday.diff, 
+                               bs = c("te", "te"),
+                               k = c(10, 10)),
+                    family = "quasibinomial",
+                    data = filter(beta_site.date, year1 == year2 &
+                                    metric == "WN"),
+                    method = "REML")
+
+gam_betadiv <- bam(value ~ 
+                     s(year1, bs = "re") +
+                     te(yday.diff, elev.diff, k = c(10, 10)),
+                   data = filter(beta_site.date, year1 == year2 &
+                                   metric == "WN.rich"),
+                   family = "quasibinomial",
+                   discrete = TRUE,
+                   method = "fREML") %>% getViz()
+
+check.gamViz(gam_betadiv)
+plot(gam_betadiv)
+
+ggplot(filter(beta_site.date, year1 == year2 & metric == "WN"),
        aes(value)) +
-  geom_histogram()  +
-  facet_wrap(~metric)
-
+  geom_histogram()
 ```
-
-# Beta diversity by elevation difference
-```{r}
-# Interaction beta diversity by elevation difference (by year)
-## This shows that the patterns are extremely consistent across years, which is not surprising, since they are driven overwhelmingly by floral turnover, and plants tend to be in the same place from year to year.
-ggplot(filter(beta_site.year, year1 == year2 & 
-                metric != "S" & 
-                metric.class != "OS.partition"), 
-       aes(elev.diff, value, color = metric)) +
-  geom_point(alpha = 0.25) +
-  geom_smooth() +
-  facet_grid(metric.class ~ year1)
-
-ggplot(filter(beta_site.year, year1 == year2 & 
-                metric != "S" & 
-                metric.class != "OS.partition"), 
-       aes(transects.diff, value, color = metric)) +
-  geom_point(alpha = 0.25) +
-  geom_smooth() +
-  facet_grid(metric.class ~ year1, scales = "free_x")
-
-# Interaction beta diversity by elevation difference, this time pooling across years
-ggplot(filter(beta_site, metric != "S" & 
-                metric.class != "OS.partition"), 
-       aes(elev.diff, value, color = metric)) +
-  geom_point(alpha = 0.05) +
-  geom_smooth() +
-  facet_wrap(~metric.class)
-
-# How important is the difference in transects? Not as important as the difference in elevation.
-ggplot(filter(beta_site, metric != "S" & 
-                metric.class != "OS.partition"), 
-       aes(transects.diff, value, color = metric)) +
-  geom_point(alpha = 0.05) +
-  geom_smooth() +
-  facet_wrap(~metric.class)
-
-ggplot(beta_site, 
-       aes(elev.diff, transects.diff)) +
-  geom_point(alpha = 0.05) +
-  geom_smooth()
-
-### Okay, we'll just do partial mantel tests to partial out the effect of transects.diff; and we'll only do that for the aggregate measures because it seems a bit silly to doi it for the partitions  
-WN_diff <- beta_site %>%
-  filter(metric == "WN") %>%
-  dplyr::select(site1, site2, value) %>%
-  spread(site2, value) %>%
-  dist()
-
-WN_diff <- beta_site %>%
-  filter(metric == "WN") %>%
-  dplyr::select(site1, site2, value) %>%
-  spread(site2, value) %>%
-  dist()
-
-ST_diff <- beta_site %>%
-  filter(metric == "ST") %>%
-  dplyr::select(site1, site2, value) %>%
-  spread(site2, value) %>%
-  dist()
-
-OS_diff <- beta_site %>%
-  filter(metric == "OS") %>%
-  dplyr::select(site1, site2, value) %>%
-  spread(site2, value) %>%
-  dist()
-
-elev_diff <- beta_site %>%
-  dplyr::select(site1, site2, elev.diff) %>%
-  unique() %>%
-  spread(site2, elev.diff) %>%
-  dist()
-
-transects_diff <- beta_site %>%
-  dplyr::select(site1, site2, transects.diff) %>%
-  unique() %>%
-  spread(site2, transects.diff) %>%
-  dist()
-
-WN_mantel <- mantel.partial(elev_diff, WN_diff, transects_diff)
-ST_mantel <- mantel.partial(elev_diff, ST_diff, transects_diff)
-OS_mantel <- mantel.partial(elev_diff, OS_diff, transects_diff)
-```
-
-### Let's look at floral survey diversity patterns
-```{r}
-### 2012
-survey_bioData.2012 <- survey %>%
-  mutate(year = year(date),
-         yday = yday(date)) %>%
-  filter(flower.cover > 0 & year == "2012") %>%
-  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
-  group_by(site, yday, plant.sp) %>%
-  summarize(cover = sum(flower.cover)) %>%
-  spread(plant.sp, cover) %>%
-  replace(is.na(.), 0) %>%
-  arrange(site, yday) %>%
-  unite(site.day, site, yday, sep = "_")
-
-predData.2012 <- survey %>%
-  mutate(year = year(date),
-         yday = yday(date)) %>%
-  filter(flower.cover > 0 & year == "2012") %>%
-  select(site, yday) %>%
-  left_join(site_data) %>%
-  ungroup() %>%
-  dplyr::select(site, elev.mean, lat, lon, yday) %>%
-  unique() %>%
-  arrange(site, yday) %>%
-  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
-  select(-site)
-
-survey_sitepair.2012 <- formatsitepair(survey_bioData.2012, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
-                              XColumn = "lon", YColumn = "lat", predData = predData.2012)
-
-### Call gdm()
-gdm_survey.2012 <- gdm(survey_sitepair.2012, geo = TRUE)
-#gdm.varImp(survey_sitepair, geo = TRUE)
-
-summary(gdm_survey.2012)
-plot(gdm_survey.2012)
-#plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
-
-### 2011
-survey_bioData.2011 <- survey %>%
-  mutate(year = year(date),
-         yday = yday(date)) %>%
-  filter(flower.cover > 0 & year == "2011") %>%
-  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
-  group_by(site, yday, plant.sp) %>%
-  summarize(cover = sum(flower.cover)) %>%
-  spread(plant.sp, cover) %>%
-  replace(is.na(.), 0) %>%
-  arrange(site, yday) %>%
-  unite(site.day, site, yday, sep = "_")
-
-predData.2011 <- survey %>%
-  mutate(year = year(date),
-         yday = yday(date)) %>%
-  filter(flower.cover > 0 & year == "2011") %>%
-  select(site, yday) %>%
-  left_join(site_data) %>%
-  ungroup() %>%
-  dplyr::select(site, elev.mean, lat, lon, yday) %>%
-  unique() %>%
-  arrange(site, yday) %>%
-  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
-  select(-site)
-
-survey_sitepair.2011 <- formatsitepair(survey_bioData.2011, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
-                              XColumn = "lon", YColumn = "lat", predData = predData.2011)
-
-### Call gdm()
-gdm_survey.2011 <- gdm(survey_sitepair.2011, geo = TRUE)
-#gdm.varImp(survey_sitepair, geo = TRUE)
-
-summary(gdm_survey.2011)
-plot(gdm_survey.2011)
-#plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
-
-### 2010
-survey_bioData.2010 <- survey %>%
-  mutate(year = year(date),
-         yday = yday(date)) %>%
-  filter(flower.cover > 0 & year == "2010") %>%
-  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
-  group_by(site, yday, plant.sp) %>%
-  summarize(cover = sum(flower.cover)) %>%
-  spread(plant.sp, cover) %>%
-  replace(is.na(.), 0) %>%
-  arrange(site, yday) %>%
-  unite(site.day, site, yday, sep = "_")
-
-predData.2010 <- survey %>%
-  mutate(year = year(date),
-         yday = yday(date)) %>%
-  filter(flower.cover > 0 & year == "2010") %>%
-  select(site, yday) %>%
-  left_join(site_data) %>%
-  ungroup() %>%
-  dplyr::select(site, elev.mean, lat, lon, yday) %>%
-  unique() %>%
-  arrange(site, yday) %>%
-  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
-  select(-site)
-
-survey_sitepair.2010 <- formatsitepair(survey_bioData.2010, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
-                              XColumn = "lon", YColumn = "lat", predData = predData.2010)
-
-### Call gdm()
-gdm_survey.2010 <- gdm(survey_sitepair.2010, geo = TRUE)
-#gdm.varImp(survey_sitepair, geo = TRUE)
-
-summary(gdm_survey.2010)
-plot(gdm_survey.2010)
-#plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
-```
-
-### And now Bombus diversity patterns
-```{r}
-BB_bioData <- net %>%
-  group_by(site, date, bb.sp) %>%
-  summarize(count = n()) %>%
-  group_by(site, date) %>%
-  mutate(prop = count/sum(count)) %>%
-  group_by(site, bb.sp) %>%
-  summarize(abund = sum(prop)) %>%
-  spread(bb.sp, abund) %>%
-  replace(is.na(.), 0)
-
-predData <- site_data %>%
-  ungroup() %>%
-  dplyr::select(site, elev.mean, lat, lon) %>%
-  unique()
-
-bb_sitepair <- formatsitepair(BB_bioData, bioFormat = 1, siteColumn = "site", dist = "jaccard",
-                              XColumn = "lon", YColumn = "lat", predData = predData)
-
-### Call gdm()
-gdm_bb <- gdm(bb_sitepair, geo = TRUE)
-gdm.varImp(bb_sitepair, geo = TRUE)
-
-summary(gdm_bb)
-plot(gdm_bb)
-plotUncertainty(bb_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
-```
-
-### Now let's look at interaction beta diversity
-```{r}
-net_sitepair <- beta_site %>%
-  filter(metric == "WN") %>%
-  mutate(weights = rep(1, n())) %>%
-  dplyr::select(distance = value, weights,
-         s1.xCoord = lon1, s1.yCoord = lat1, 
-         s2.xCoord = lon2, s2.yCoord = lat2, 
-         s1.elev.mean = elev1, s2.elev.mean = elev2) %>%
-  data.frame()
-
-### Call gdm()
-gdm_net <- gdm(net_sitepair, geo = TRUE)
-gdm.varImp(net_sitepair, geo = TRUE)
-
-summary(gdm_net)
-plot(gdm_net)
-plotUncertainty(net_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
-```
-
-### At the transect level
-```{r}
-beta_site.date_gdm.prep
-
-transects_2010 <- beta_site.date.2010 %>%
-  filter(metric == "WN") %>%
-  mutate(weights = rep(1, n())) %>%
-  dplyr::select(distance = value, weights,
-         s1.xCoord = lon1, s1.yCoord = lat1, 
-         s2.xCoord = lon2, s2.yCoord = lat2, 
-         s1.elev = elev1, s2.elev = elev2,
-         s1.yday = yday1, s2.yday = yday2) %>%
-  data.frame()
-
-### Call gdm()
-gdm_net <- gdm(transects_2010, geo = TRUE)
-gdm.varImp(transects_2010, geo = TRUE)
-
-summary(gdm_net)
-plot(gdm_net)
-plotUncertainty(net_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
-```

### 2020-09-09 analyses/beta_diversity_v2.Rmd
update

--- 2020-09-09 analyses/beta_diversity.Rmd
@@
 
-### betaboot function: This bootstraps the partitioning of interaction beta-diversity to account for differences in the number of sampling dates for each site in the same way the my other bootstrapping function ("bootnet" or something like that) bootstraps the calculation of network metrics.
-Note that due to some annoying behavior by webs2array(), I have to define an object within the for loop  of my function using the <<- global assignment operator.
-This is bad coding, but I can't figure out another way to get webs2array() to find the object it needs to find.
-```{r message=FALSE}
-betaboot <- function(x, n, iter) {
-
-  out <- data.frame()
-
-  index <- 0
-
-  t_filter <- x %>%
-    dplyr::select(site, date, year) %>%
-    unique() %>%
-    group_by(site, year) %>%
-    mutate(transects = n()) %>%
-    filter(transects >= n)
-
-  for(i in 1:iter){
-
-    index <- index + 1
-
-    net.sub <- x %>%
-      semi_join(t_filter) %>%
-      ungroup() %>%
-      dplyr::select(site, dayofyear) %>%
-      unique() %>%
-      group_by(site) %>%
-      sample_n(n) %>%
-      left_join(x)
-
-    web <<- net.sub %>%
-      group_by(site, bb.sp, plant.sp) %>%
-      summarize(freq = n()) %>%
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() # convert to to bipartite web
-
-    array <- webs2array(web)
-
-    beta <- betalinkr_multi(array, partition.st = TRUE, partition.rr = TRUE, binary = TRUE) %>%
-      as_tibble() %>%
-      mutate(iteration = rep(index, n()))
-
-    out <- bind_rows(out, beta)
-  }
-
-  out <- out %>%
-    as_tibble() %>%
-    dplyr::select(site1 = i, site2 = j, everything()) %>%
-    gather(metric, value, -c(site1, site2, iteration)) %>%
-    mutate(metric.class = case_when(
-        metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
-        metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
-        metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
-        metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
-      ),
-      iteration = factor(iteration))
-}
-```
-
 ### Tabulate webs
@@
 ```{r message=FALSE}
-net2010 <- filter(net, year == 2010)
-net2011 <- filter(net, year == 2011)
-net2012 <- filter(net, year == 2012)
+net2010 <- filter(net, year == 2010) %>%
+  unite(site.date, c(site, date), sep = "_") %>%
+  group_by(site.date, bb.sp, plant.sp) %>%
+  summarize(freq = n()) %>% 
+  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.date, freq) %>% # rename columns to match bipartite's expectations
+  data.frame() %>% # convert to data frame
+  frame2webs()
 
-beta_2010 <- betaboot(net2010, 5, 50) %>%
-  group_by(site1, site2, metric, metric.class) %>%
-  summarize(value = mean(value)) %>%
-  left_join(site_data, by = c("site1" = "site")) %>%
-  dplyr::select(site1, site2, metric, metric.class, value, 
-                elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
-  left_join(site_data, by = c("site2" = "site")) %>%
-  dplyr::select(site1, site2, metric, metric.class, value, 
-         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
-  mutate(elev.diff = abs(elev1 - elev2),
-         mean.elev = (elev1 + elev2)/2,
-         man.diff = case_when(
-           man1 == man2 ~ FALSE,
-           man1 != man2 ~ TRUE
-         ),
-         year = factor(rep("2010", n())))
+net2011 <- filter(net, year == 2011) %>%
+  unite(site.date, c(site, date), sep = "_") %>%
+  group_by(site.date, bb.sp, plant.sp) %>%
+  summarize(freq = n()) %>% 
+  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.date, freq) %>% # rename columns to match bipartite's expectations
+  data.frame() %>% # convert to data frame
+  frame2webs()
 
-beta_2011 <- betaboot(net2011, 5, 50) %>%
-  group_by(site1, site2, metric, metric.class) %>%
-  summarize(value = mean(value)) %>%
-  left_join(site_data, by = c("site1" = "site")) %>%
-  dplyr::select(site1, site2, metric, metric.class, value, 
-                elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
-  left_join(site_data, by = c("site2" = "site")) %>%
-  dplyr::select(site1, site2, metric, metric.class, value, 
-         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
-  mutate(elev.diff = abs(elev1 - elev2),
-         mean.elev = (elev1 + elev2)/2,
-         man.diff = case_when(
-           man1 == man2 ~ FALSE,
-           man1 != man2 ~ TRUE
-         ),
-         year = factor(rep("2011", n())))
+net2012 <- filter(net, year == 2012) %>%
+  unite(site.date, c(site, date), sep = "_") %>%
+  group_by(site.date, bb.sp, plant.sp) %>%
+  summarize(freq = n()) %>% 
+  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.date, freq) %>% # rename columns to match bipartite's expectations
+  data.frame() %>% # convert to data frame
+  frame2webs()
 
-beta_2012 <- betaboot(net2012, 9, 50) %>%
-  group_by(site1, site2, metric, metric.class) %>%
-  summarize(value = mean(value)) %>%
-  left_join(site_data, by = c("site1" = "site")) %>%
-  dplyr::select(site1, site2, metric, metric.class, value, 
-                elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
-  left_join(site_data, by = c("site2" = "site")) %>%
-  dplyr::select(site1, site2, metric, metric.class, value, 
-         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
-  mutate(elev.diff = abs(elev1 - elev2),
-         mean.elev = (elev1 + elev2)/2,
-         man.diff = case_when(
-           man1 == man2 ~ FALSE,
-           man1 != man2 ~ TRUE
-         ),
-         year = factor(rep("2012", n())))
+array_site.date.2010 <- webs2array(net2010) # convert to array of webs
+array_site.date.2011 <- webs2array(net2011) # convert to array of webs
+array_site.date.2012 <- webs2array(net2012) # convert to array of webs
 
-beta_agg <- bind_rows(beta_2010, beta_2011, beta_2012)
-
-#write_csv(beta_agg, "./output/aggregated_beta_diversity.csv")
-#beta_agg <- read_csv("./output/aggregated_beta_diversity.csv")
-```
-
-### Plot results. There's nothing a Mantel test can tell me that I don't already know from the combination of plotting and GDM analysis.
-```{r}
-### Elevation
-ggplot(filter(beta, year == "2010" & 
-                metric != "S" & 
-                metric.class != "OS.partition"), 
-       aes(elev.diff, value, color = metric)) +
-  geom_point(alpha = 0.45) +
-  geom_smooth(method = "lm", formula =  y ~ poly(x, 3)) +
-  facet_wrap(~metric.class) +
-  theme_abyss()
-
-ggplot(filter(beta, year == "2011" & 
-                metric != "S" & 
-                metric.class != "OS.partition"), 
-       aes(elev.diff, value, color = metric)) +
-  geom_point(alpha = 0.45) +
-  geom_smooth(method = "lm", formula =  y ~ poly(x, 3)) +
-  facet_wrap(~metric.class) +
-  theme_abyss()
-
-ggplot(filter(beta, year == "2012" & 
-                metric != "S" & 
-                metric.class != "OS.partition"), 
-       aes(elev.diff, value, color = metric)) +
-  geom_point(alpha = 0.45) +
-  geom_smooth(method = "lm", formula =  y ~ poly(x, 3)) +
-  facet_wrap(~metric.class) +
-  theme_abyss()
-
-ggplot(filter(beta, 
-                metric != "S" & 
-                metric.class != "OS.partition"), 
-       aes(elev.diff, value, color = metric)) +
-  geom_point(alpha = 0.45) +
-  geom_smooth(method = "lm", formula =  y ~ poly(x, 3)) +
-  facet_wrap(~metric.class) +
-  theme_gray()
-
-### Management
-ggplot(filter(beta, year == "2010" & 
-                metric != "S" & 
-                metric.class != "OS.partition"), 
-       aes(man.diff, value)) +
-  geom_boxplot() +
-  facet_wrap(~metric, scales = "free") +
-  theme_gray()
-
-ggplot(filter(beta, year == "2011" & 
-                metric != "S" & 
-                metric.class != "OS.partition"), 
-       aes(man.diff, value)) +
-  geom_boxplot() +
-  facet_wrap(~metric, scales = "free") +
-  theme_gray()
-
-ggplot(filter(beta, year == "2012" & 
-                metric != "S" & 
-                metric.class != "OS.partition"), 
-       aes(man.diff, value)) +
-  geom_boxplot() +
-  facet_wrap(~metric, scales = "free") +
-  theme_gray()
-```
-
-### Let's look at per-transect floral survey beta-diversity
-```{r}
-### 2012
-survey_bioData.2012 <- survey %>%
-  mutate(year = year(date),
-         yday = yday(date)) %>%
-  filter(flower.cover > 0 & year == "2012") %>%
-  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
-  group_by(site, yday, plant.sp) %>%
-  summarize(cover = sum(flower.cover)) %>%
-  spread(plant.sp, cover) %>%
-  replace(is.na(.), 0) %>%
-  arrange(site, yday) %>%
-  unite(site.day, site, yday, sep = "_")
-
-survey_predData.2012 <- survey %>%
-  mutate(year = year(date),
-         yday = yday(date)) %>%
-  filter(flower.cover > 0 & year == "2012") %>%
-  dplyr::select(site, yday) %>%
-  left_join(site_data) %>%
-  ungroup() %>%
-  dplyr::select(site, elev.mean, lat, lon, yday) %>%
-  unique() %>%
-  arrange(site, yday) %>%
-  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
-  dplyr::select(-site)
-
-survey_sitepair.2012 <- formatsitepair(survey_bioData.2012, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
-                              XColumn = "lon", YColumn = "lat", predData = survey_predData.2012)
-
-### Call gdm()
-gdm_survey.2012 <- gdm(survey_sitepair.2012, geo = TRUE)
-gdm.varImp(survey_sitepair.2012, geo = TRUE)
-
-summary(gdm_survey.2012)
-plot(gdm_survey.2012)
-plotUncertainty(survey_sitepair.2012, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
-
-### 2011
-survey_bioData.2011 <- survey %>%
-  mutate(year = year(date),
-         yday = yday(date)) %>%
-  filter(flower.cover > 0 & year == "2011") %>%
-  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
-  group_by(site, yday, plant.sp) %>%
-  summarize(cover = sum(flower.cover)) %>%
-  spread(plant.sp, cover) %>%
-  replace(is.na(.), 0) %>%
-  arrange(site, yday) %>%
-  unite(site.day, site, yday, sep = "_")
-
-predData.2011 <- survey %>%
-  mutate(year = year(date),
-         yday = yday(date)) %>%
-  filter(flower.cover > 0 & year == "2011") %>%
-  select(site, yday) %>%
-  left_join(site_data) %>%
-  ungroup() %>%
-  dplyr::select(site, elev.mean, lat, lon, yday) %>%
-  unique() %>%
-  arrange(site, yday) %>%
-  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
-  select(-site)
-
-survey_sitepair.2011 <- formatsitepair(survey_bioData.2011, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
-                              XColumn = "lon", YColumn = "lat", predData = predData.2011)
-
-### Call gdm()
-gdm_survey.2011 <- gdm(survey_sitepair.2011, geo = TRUE)
-#gdm.varImp(survey_sitepair, geo = TRUE)
-
-summary(gdm_survey.2011)
-plot(gdm_survey.2011)
-#plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
-
-### 2010
-survey_bioData.2010 <- survey %>%
-  mutate(year = year(date),
-         yday = yday(date)) %>%
-  filter(flower.cover > 0 & year == "2010") %>%
-  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
-  group_by(site, yday, plant.sp) %>%
-  summarize(cover = sum(flower.cover)) %>%
-  spread(plant.sp, cover) %>%
-  replace(is.na(.), 0) %>%
-  arrange(site, yday) %>%
-  unite(site.day, site, yday, sep = "_")
-
-predData.2010 <- survey %>%
-  mutate(year = year(date),
-         yday = yday(date)) %>%
-  filter(flower.cover > 0 & year == "2010") %>%
-  select(site, yday) %>%
-  left_join(site_data) %>%
-  ungroup() %>%
-  dplyr::select(site, elev.mean, lat, lon, yday) %>%
-  unique() %>%
-  arrange(site, yday) %>%
-  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
-  select(-site)
-
-survey_sitepair.2010 <- formatsitepair(survey_bioData.2010, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
-                              XColumn = "lon", YColumn = "lat", predData = predData.2010)
-
-### Call gdm()
-gdm_survey.2010 <- gdm(survey_sitepair.2010, geo = TRUE)
-#gdm.varImp(survey_sitepair, geo = TRUE)
-
-summary(gdm_survey.2010)
-plot(gdm_survey.2010)
-#plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
-```
-
-### And now Bombus diversity patterns
-```{r}
-### 2010
-BB_bioData.2010 <- net %>%
-  filter(year == 2010) %>%
-  group_by(site, date, bb.sp) %>%
-  summarize(count = n()) %>%
-  group_by(site, date) %>%
-  #mutate(prop = count/sum(count)) %>%
-  #group_by(site, bb.sp) %>%
-  # summarize(abund = sum(prop)) %>%
-  spread(bb.sp, count) %>%
-  replace(is.na(.), 0) %>%
-  mutate(yday = yday(date)) %>%
-  unite(site.day, site, yday, sep = "_", remove = "TRUE") %>%
-  ungroup() %>%
-  select(-date)
-
-BB_predData.2010 <- net %>%
-  filter(year == 2010) %>%
-  mutate(yday = yday(date)) %>%
-  dplyr::select(site, yday) %>%
-  left_join(site_data) %>%
-  ungroup() %>%
-  dplyr::select(site, elev.mean, lat, lon, yday) %>%
-  unique() %>%
-  arrange(site, yday) %>%
-  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
-  dplyr::select(-site)
-
-bb_sitepair.2010 <- formatsitepair(BB_bioData.2010, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
-                              XColumn = "lon", YColumn = "lat", predData = BB_predData.2010)
-
-### Call gdm()
-gdm_bb.2010 <- gdm(bb_sitepair.2010, geo = TRUE)
-gdm.varImp(bb_sitepair.2010, geo = TRUE)
-
-summary(gdm_bb.2010)
-plot(gdm_bb.2010)
-plotUncertainty(bb_sitepair.2010, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
-
-### 2011
-BB_bioData.2011 <- net %>%
-  filter(year == 2011) %>%
-  group_by(site, date, bb.sp) %>%
-  summarize(count = n()) %>%
-  group_by(site, date) %>%
-  #mutate(prop = count/sum(count)) %>%
-  #group_by(site, bb.sp) %>%
-  # summarize(abund = sum(prop)) %>%
-  spread(bb.sp, count) %>%
-  replace(is.na(.), 0) %>%
-  mutate(yday = yday(date)) %>%
-  unite(site.day, site, yday, sep = "_", remove = "TRUE") %>%
-  ungroup() %>%
-  select(-date)
-
-BB_predData.2011 <- net %>%
-  filter(year == 2011) %>%
-  mutate(yday = yday(date)) %>%
-  dplyr::select(site, yday) %>%
-  left_join(site_data) %>%
-  ungroup() %>%
-  dplyr::select(site, elev.mean, lat, lon, yday) %>%
-  unique() %>%
-  arrange(site, yday) %>%
-  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
-  dplyr::select(-site)
-
-bb_sitepair.2011 <- formatsitepair(BB_bioData.2011, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
-                              XColumn = "lon", YColumn = "lat", predData = BB_predData.2011)
-
-### Call gdm()
-gdm_bb.2011 <- gdm(bb_sitepair.2011, geo = TRUE)
-gdm.varImp(bb_sitepair.2011, geo = TRUE)
-
-summary(gdm_bb.2011)
-plot(gdm_bb.2011)
-plotUncertainty(bb_sitepair.2011, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
-
-### 2012
-BB_bioData.2012 <- net %>%
-  filter(year == 2012) %>%
-  group_by(site, date, bb.sp) %>%
-  summarize(count = n()) %>%
-  group_by(site, date) %>%
-  #mutate(prop = count/sum(count)) %>%
-  #group_by(site, bb.sp) %>%
-  # summarize(abund = sum(prop)) %>%
-  spread(bb.sp, count) %>%
-  replace(is.na(.), 0) %>%
-  mutate(yday = yday(date)) %>%
-  unite(site.day, site, yday, sep = "_", remove = "TRUE") %>%
-  ungroup() %>%
-  select(-date)
-
-BB_predData.2012 <- net %>%
-  filter(year == 2012) %>%
-  mutate(yday = yday(date)) %>%
-  dplyr::select(site, yday) %>%
-  left_join(site_data) %>%
-  ungroup() %>%
-  dplyr::select(site, elev.mean, lat, lon, yday) %>%
-  unique() %>%
-  arrange(site, yday) %>%
-  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
-  dplyr::select(-site)
-
-bb_sitepair.2012 <- formatsitepair(BB_bioData.2012, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
-                              XColumn = "lon", YColumn = "lat", predData = BB_predData.2012)
-
-### Call gdm()
-gdm_bb.2012 <- gdm(bb_sitepair.2012, geo = TRUE)
-gdm.varImp(bb_sitepair.2012, geo = TRUE)
-
-summary(gdm_bb.2012)
-plot(gdm_bb.2012)
-plotUncertainty(bb_sitepair.2012, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
-```
-
-### Now let's look at interaction beta diversity.
-I don't think p-values would add anything to the interpretation I get from the plotted uncertainty, so it's okay that I can't get gdm.varImp() to work.
-```{r}
-### 2010
-sitepair_2010 <- beta_2010 %>%
-  ungroup() %>%
-  filter(metric == "WN") %>%
-  mutate(weights = rep(1, n())) %>%
-  dplyr::select(distance = value, weights,
-         s1.xCoord = lon1, s1.yCoord = lat1, 
-         s2.xCoord = lon2, s2.yCoord = lat2, 
-         s1.elev.mean = elev1, s2.elev.mean = elev2) %>%
-  data.frame()
-
-gdm_2010 <- gdm(sitepair_2010, geo = TRUE)
-gdm.varImp(sitepair_2010, geo = TRUE)
-
-summary(gdm_2010)
-plot(gdm_2010)
-plotUncertainty(sitepair_2010, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
-
-### 2011
-sitepair_2011 <- beta_2011 %>%
-  ungroup() %>%
-  filter(metric == "WN") %>%
-  mutate(weights = rep(1, n())) %>%
-  dplyr::select(distance = value, weights,
-         s1.xCoord = lon1, s1.yCoord = lat1, 
-         s2.xCoord = lon2, s2.yCoord = lat2, 
-         s1.elev.mean = elev1, s2.elev.mean = elev2) %>%
-  data.frame()
-
-gdm_2011 <- gdm(sitepair_2011, geo = TRUE)
-gdm.varImp(sitepair_2011, geo = TRUE)
-
-summary(gdm_2011)
-plot(gdm_2011)
-plotUncertainty(sitepair_2011, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
-
-### 2012
-sitepair_2012 <- beta_2012 %>%
-  ungroup() %>%
-  filter(metric == "WN") %>%
-  mutate(weights = rep(1, n())) %>%
-  dplyr::select(distance = value, weights,
-         s1.xCoord = lon1, s1.yCoord = lat1, 
-         s2.xCoord = lon2, s2.yCoord = lat2, 
-         s1.elev.mean = elev1, s2.elev.mean = elev2) %>%
-  data.frame()
-
-gdm_2012 <- gdm(sitepair_2012, geo = TRUE)
-gdm.varImp(sitepair_2012, geo = TRUE)
-
-summary(gdm_2012)
-plot(gdm_2012)
-plotUncertainty(sitepair_2012, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
-```
-
-### Transect-level analysis
-```{r}
 beta_site.date.2010 <- betalinkr_multi(array_site.date.2010, 
@@
 
-### Plot transect-level analysis
+### Plot results. There's nothing a Mantel test can tell me that I don't already know from the combination of plotting and GDM analysis.
 ```{r}
@@
   facet_wrap(~metric.class) +
-  theme_abyss()
+  theme_light()
 
@@
 ```
+
+### GDM analysis. I don't think p-values would add anything to the interpretation I get from the plotted uncertainty, so it's okay that I can't get gdm.varImp() to work.
+```{r}
+### 2010
+sitepair_2010.WN <- beta_site.date %>%
+  ungroup() %>%
+  filter(year1 == year2 & year1 == 2010) %>%
+  filter(metric == "WN") %>%
+  mutate(weights = rep(1, n())) %>%
+  dplyr::select(distance = value, weights,
+         s1.xCoord = lon1, s1.yCoord = lat1, 
+         s2.xCoord = lon2, s2.yCoord = lat2, 
+         s1.elev = elev1, s1.yday = yday1,
+         s2.elev = elev2, s2.yday = yday2) %>%
+  data.frame()
+
+gdm_WN.2010 <- gdm(sitepair_2010.WN, geo = TRUE)
+gdm_WN.2010.imp <- gdm.varImp(sitepair_2010.WN, geo = TRUE)
+
+summary(gdm_WN.2010)
+plot(gdm_WN.2010)
+plotUncertainty(sitepair_2010.WN, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
+
+### 2011
+sitepair_2011.WN <- beta_site.date %>%
+  ungroup() %>%
+  filter(year1 == year2 & year1 == 2011) %>%
+  filter(metric == "WN") %>%
+  mutate(weights = rep(1, n())) %>%
+  dplyr::select(distance = value, weights,
+         s1.xCoord = lon1, s1.yCoord = lat1, 
+         s2.xCoord = lon2, s2.yCoord = lat2, 
+         s1.elev = elev1, s1.yday = yday1,
+         s2.elev = elev2, s2.yday = yday2) %>%
+  data.frame()
+
+gdm_WN.2011 <- gdm(sitepair_2011.WN, geo = TRUE)
+gdm_WN.2011.imp <- gdm.varImp(sitepair_2011.WN, geo = TRUE)
+
+summary(gdm_WN.2011)
+plot(gdm_WN.2011)
+plotUncertainty(sitepair_2011.WN, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
+
+### 2012
+sitepair_2012.WN <- beta_site.date %>%
+  ungroup() %>%
+  filter(year1 == year2 & year1 == 2012) %>%
+  filter(metric == "WN") %>%
+  mutate(weights = rep(1, n())) %>%
+  dplyr::select(distance = value, weights,
+         s1.xCoord = lon1, s1.yCoord = lat1, 
+         s2.xCoord = lon2, s2.yCoord = lat2, 
+         s1.elev = elev1, s1.yday = yday1,
+         s2.elev = elev2, s2.yday = yday2) %>%
+  data.frame()
+
+gdm_WN.2012 <- gdm(sitepair_2012.WN, geo = TRUE)
+gdm_WN.2012.imp <- gdm.varImp(sitepair_2012.WN, geo = TRUE)
+
+summary(gdm_WN.2012)
+plot(gdm_WN.2012)
+plotUncertainty(sitepair_2012.WN, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
+```
+
+### Let's look at per-transect floral survey beta-diversity
+```{r}
+### 2012
+survey_bioData.2012 <- survey %>%
+  mutate(year = year(date),
+         yday = yday(date)) %>%
+  filter(flower.cover > 0 & year == "2012") %>%
+  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
+  group_by(site, yday, plant.sp) %>%
+  summarize(cover = sum(flower.cover)) %>%
+  spread(plant.sp, cover) %>%
+  replace(is.na(.), 0) %>%
+  arrange(site, yday) %>%
+  unite(site.day, site, yday, sep = "_")
+
+survey_predData.2012 <- survey %>%
+  mutate(year = year(date),
+         yday = yday(date)) %>%
+  filter(flower.cover > 0 & year == "2012") %>%
+  dplyr::select(site, yday) %>%
+  left_join(site_data) %>%
+  ungroup() %>%
+  dplyr::select(site, elev.mean, lat, lon, yday) %>%
+  unique() %>%
+  arrange(site, yday) %>%
+  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
+  dplyr::select(-site)
+
+survey_sitepair.2012 <- formatsitepair(survey_bioData.2012, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
+                              XColumn = "lon", YColumn = "lat", predData = survey_predData.2012)
+
+### Call gdm()
+gdm_survey.2012 <- gdm(survey_sitepair.2012, geo = TRUE)
+gdm_survey.2012.imp <- gdm.varImp(survey_sitepair.2012, geo = TRUE)
+
+summary(gdm_survey.2012)
+plot(gdm_survey.2012)
+plotUncertainty(survey_sitepair.2012, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
+
+### 2011
+survey_bioData.2011 <- survey %>%
+  mutate(year = year(date),
+         yday = yday(date)) %>%
+  filter(flower.cover > 0 & year == "2011") %>%
+  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
+  group_by(site, yday, plant.sp) %>%
+  summarize(cover = sum(flower.cover)) %>%
+  spread(plant.sp, cover) %>%
+  replace(is.na(.), 0) %>%
+  arrange(site, yday) %>%
+  unite(site.day, site, yday, sep = "_")
+
+predData.2011 <- survey %>%
+  mutate(year = year(date),
+         yday = yday(date)) %>%
+  filter(flower.cover > 0 & year == "2011") %>%
+  select(site, yday) %>%
+  left_join(site_data) %>%
+  ungroup() %>%
+  dplyr::select(site, elev.mean, lat, lon, yday) %>%
+  unique() %>%
+  arrange(site, yday) %>%
+  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
+  select(-site)
+
+survey_sitepair.2011 <- formatsitepair(survey_bioData.2011, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
+                              XColumn = "lon", YColumn = "lat", predData = predData.2011)
+
+### Call gdm()
+gdm_survey.2011 <- gdm(survey_sitepair.2011, geo = TRUE)
+gdm_survey.2011.imp <- gdm.varImp(survey_sitepair.2011, geo = TRUE)
+
+summary(gdm_survey.2011)
+plot(gdm_survey.2011)
+plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
+
+### 2010
+survey_bioData.2010 <- survey %>%
+  mutate(year = year(date),
+         yday = yday(date)) %>%
+  filter(flower.cover > 0 & year == "2010") %>%
+  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
+  group_by(site, yday, plant.sp) %>%
+  summarize(cover = sum(flower.cover)) %>%
+  spread(plant.sp, cover) %>%
+  replace(is.na(.), 0) %>%
+  arrange(site, yday) %>%
+  unite(site.day, site, yday, sep = "_")
+
+predData.2010 <- survey %>%
+  mutate(year = year(date),
+         yday = yday(date)) %>%
+  filter(flower.cover > 0 & year == "2010") %>%
+  select(site, yday) %>%
+  left_join(site_data) %>%
+  ungroup() %>%
+  dplyr::select(site, elev.mean, lat, lon, yday) %>%
+  unique() %>%
+  arrange(site, yday) %>%
+  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
+  select(-site)
+
+survey_sitepair.2010 <- formatsitepair(survey_bioData.2010, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
+                              XColumn = "lon", YColumn = "lat", predData = predData.2010)
+
+### Call gdm()
+gdm_survey.2010 <- gdm(survey_sitepair.2010, geo = TRUE)
+gdm_survey.2010.imp <- gdm.varImp(survey_sitepair.2010, geo = TRUE)
+
+summary(gdm_survey.2010)
+plot(gdm_survey.2010)
+plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
+```
+
+### And now Bombus diversity patterns
+```{r}
+### 2010
+BB_bioData.2010 <- net %>%
+  filter(year == 2010) %>%
+  group_by(site, date, bb.sp) %>%
+  summarize(count = n()) %>%
+  group_by(site, date) %>%
+  #mutate(prop = count/sum(count)) %>%
+  #group_by(site, bb.sp) %>%
+  # summarize(abund = sum(prop)) %>%
+  spread(bb.sp, count) %>%
+  replace(is.na(.), 0) %>%
+  mutate(yday = yday(date)) %>%
+  unite(site.day, site, yday, sep = "_", remove = "TRUE") %>%
+  ungroup() %>%
+  select(-date)
+
+BB_predData.2010 <- net %>%
+  filter(year == 2010) %>%
+  mutate(yday = yday(date)) %>%
+  dplyr::select(site, yday) %>%
+  left_join(site_data) %>%
+  ungroup() %>%
+  dplyr::select(site, elev.mean, lat, lon, yday) %>%
+  unique() %>%
+  arrange(site, yday) %>%
+  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
+  dplyr::select(-site)
+
+bb_sitepair.2010 <- formatsitepair(BB_bioData.2010, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
+                              XColumn = "lon", YColumn = "lat", predData = BB_predData.2010)
+
+### Call gdm()
+gdm_bb.2010 <- gdm(bb_sitepair.2010, geo = TRUE)
+gdm_bb.2010.imp <- gdm.varImp(bb_sitepair.2010, geo = TRUE)
+
+summary(gdm_bb.2010)
+plot(gdm_bb.2010)
+plotUncertainty(bb_sitepair.2010, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
+
+### 2011
+BB_bioData.2011 <- net %>%
+  filter(year == 2011) %>%
+  group_by(site, date, bb.sp) %>%
+  summarize(count = n()) %>%
+  group_by(site, date) %>%
+  #mutate(prop = count/sum(count)) %>%
+  #group_by(site, bb.sp) %>%
+  # summarize(abund = sum(prop)) %>%
+  spread(bb.sp, count) %>%
+  replace(is.na(.), 0) %>%
+  mutate(yday = yday(date)) %>%
+  unite(site.day, site, yday, sep = "_", remove = "TRUE") %>%
+  ungroup() %>%
+  select(-date)
+
+BB_predData.2011 <- net %>%
+  filter(year == 2011) %>%
+  mutate(yday = yday(date)) %>%
+  dplyr::select(site, yday) %>%
+  left_join(site_data) %>%
+  ungroup() %>%
+  dplyr::select(site, elev.mean, lat, lon, yday) %>%
+  unique() %>%
+  arrange(site, yday) %>%
+  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
+  dplyr::select(-site)
+
+bb_sitepair.2011 <- formatsitepair(BB_bioData.2011, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
+                              XColumn = "lon", YColumn = "lat", predData = BB_predData.2011)
+
+### Call gdm()
+gdm_bb.2011 <- gdm(bb_sitepair.2011, geo = TRUE)
+gdm.varImp(bb_sitepair.2011, geo = TRUE)
+
+summary(gdm_bb.2011)
+plot(gdm_bb.2011)
+plotUncertainty(bb_sitepair.2011, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
+
+### 2012
+BB_bioData.2012 <- net %>%
+  filter(year == 2012) %>%
+  group_by(site, date, bb.sp) %>%
+  summarize(count = n()) %>%
+  group_by(site, date) %>%
+  #mutate(prop = count/sum(count)) %>%
+  #group_by(site, bb.sp) %>%
+  # summarize(abund = sum(prop)) %>%
+  spread(bb.sp, count) %>%
+  replace(is.na(.), 0) %>%
+  mutate(yday = yday(date)) %>%
+  unite(site.day, site, yday, sep = "_", remove = "TRUE") %>%
+  ungroup() %>%
+  select(-date)
+
+BB_predData.2012 <- net %>%
+  filter(year == 2012) %>%
+  mutate(yday = yday(date)) %>%
+  dplyr::select(site, yday) %>%
+  left_join(site_data) %>%
+  ungroup() %>%
+  dplyr::select(site, elev.mean, lat, lon, yday) %>%
+  unique() %>%
+  arrange(site, yday) %>%
+  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
+  dplyr::select(-site)
+
+bb_sitepair.2012 <- formatsitepair(BB_bioData.2012, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
+                              XColumn = "lon", YColumn = "lat", predData = BB_predData.2012)
+
+### Call gdm()
+gdm_bb.2012 <- gdm(bb_sitepair.2012, geo = TRUE)
+gdm.varImp(bb_sitepair.2012, geo = TRUE)
+
+summary(gdm_bb.2012)
+plot(gdm_bb.2012)
+plotUncertainty(bb_sitepair.2012, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
+```

### 2020-09-09 analyses/network_analysis.Rmd
update

@@
   geom_smooth()
+
+### How evenly is the presence of species distributed across samples
+bb_abund_sp <- net %>%
+  group_by(site, dayofyear, elev.mean, year, bb.sp) %>%
+  summarize(abund = n())
+
+ggplot(filter(bb_abund_sp, year == 2011), aes(dayofyear, elev.mean)) +
+  geom_point(aes(alpha = abund)) +
+  facet_wrap(~bb.sp)
 ```
@@
 gam_d.sp_I <- gam(d ~ year + 
-                     s(site, bs = "re") +
+                     #s(site, bs = "re") +
                      management +
@@
                      s(log.fl.abund, k = 8) +
-                     s(log.bb.sp.abund, k = 5),
+                     s(bb.sp.abund, k = 5),
                    select = TRUE,

### 2020-09-09 analyses/trash_can.R

@@ -1031,3 +1031,173 @@ bigbam_2012 <- bamV(niche.overlap.HL ~ s(site, bs = "re") +
 check.gamViz(bigbam_2012)
 bigbam_2012.sum <- summary(bigbam_2012)
 print(plot(bigbam_2012, allTerms = TRUE), pages = 1)
+
+### Transect-level analysis
+```{r}
+beta_site.date.2010 <- betalinkr_multi(array_site.date.2010, 
+                                       partition.st = TRUE, 
+                                       partition.rr = TRUE) %>%
+  as_tibble() %>%
+  separate(i, into = c("site1", "year1", "yday1"), sep = "_") %>%
+  separate(j, into = c("site2", "year2", "yday2"), sep = "_") %>%
+  left_join(site_data, by = c("site1" = "site"))%>%
+  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
+                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
+                elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
+  left_join(site_data, by = c("site2" = "site")) %>%
+  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
+                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
+                elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
+  mutate(site1 = factor(site1),
+         site2 = factor(site2),
+         year1 = factor(year1),
+         year2 = factor(year2),
+         yday1 = as.numeric(yday1),
+         yday2 = as.numeric(yday2),
+         elev.diff = abs(elev1 - elev2),
+         mean.elev = (elev1 + elev2)/2,
+         yday.diff = abs(yday1 - yday2),
+         man.diff = case_when(
+           man1 == man2 ~ FALSE,
+           man1 != man2 ~ TRUE
+         )) %>%
+  gather(metric, value, -c(site1, year1, yday1, site2, year2, yday2,
+                           elev1, elev2, elev.diff, yday.diff, 
+                           man1, man2, man.diff, mean.elev, 
+                           lon1, lat1, lon2, lat2)) %>%
+  mutate(metric.class = case_when(
[…]

### 2020-09-09 analyses/beta_diversity_v2.Rmd
make beta-div analysis more solid

@@
 ```{r message=FALSE}
+# Create webs
 net2010 <- filter(net, year == 2010) %>%
@@
   summarize(freq = n()) %>%
+  group_by(site.date) %>%
+  mutate(bb.sp.count = length(unique(bb.sp)),
+         plant.sp.count = length(unique(plant.sp))) %>%
+  ungroup() %>%
+  filter(bb.sp.count > 1 & plant.sp.count > 1) %>% # the betalinkr function fails if you feed it webs with only 1 species in either (or perhaps both) of the levels; this is probably for good reason.
   dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.date, freq) %>% # rename columns to match bipartite's expectations
   data.frame() %>% # convert to data frame
-  frame2webs()
+  frame2webs(type.out = "array")
 
@@
   summarize(freq = n()) %>%
+  group_by(site.date) %>%
+  mutate(bb.sp.count = length(unique(bb.sp)),
+         plant.sp.count = length(unique(plant.sp))) %>%
+  ungroup() %>%
+  filter(bb.sp.count > 1 & plant.sp.count > 1) %>% 
   dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.date, freq) %>% # rename columns to match bipartite's expectations
   data.frame() %>% # convert to data frame
-  frame2webs()
+  frame2webs(type.out = "array")
 
@@
   summarize(freq = n()) %>%
+  group_by(site.date) %>%
+  mutate(bb.sp.count = length(unique(bb.sp)),
+         plant.sp.count = length(unique(plant.sp))) %>%
+  ungroup() %>%
+  filter(bb.sp.count > 1 & plant.sp.count > 1) %>% 
   dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.date, freq) %>% # rename columns to match bipartite's expectations
   data.frame() %>% # convert to data frame
-  frame2webs()
+  frame2webs(type.out = "array")
 
-array_site.date.2010 <- webs2array(net2010) # convert to array of webs
-array_site.date.2011 <- webs2array(net2011) # convert to array of webs
-array_site.date.2012 <- webs2array(net2012) # convert to array of webs
+# Calculate beta-diversity
+beta_site.date.2010 <- betalinkr_multi(net2010, 
+                                       partition.st = TRUE, 
+                                       partition.rr = TRUE,
+                                       binary = TRUE,
+                                       index = "jaccard") %>%
+  as_tibble()
 
-beta_site.date.2010 <- betalinkr_multi(array_site.date.2010, 
+beta_site.date.2011 <- betalinkr_multi(net2011, 
                                        partition.st = TRUE, 
-                                       partition.rr = TRUE) %>%
-  as_tibble() %>%
-  separate(i, into = c("site1", "year1", "yday1"), sep = "_") %>%
-  separate(j, into = c("site2", "year2", "yday2"), sep = "_") %>%
-  left_join(site_data, by = c("site1" = "site"))%>%
+                                       partition.rr = TRUE,
+                                       binary = TRUE,
+                                       index = "jaccard") %>%
+  as_tibble()
+
+beta_site.date.2012 <- betalinkr_multi(net2012, 
+                                       partition.st = TRUE, 
+                                       partition.rr = TRUE,
+                                       binary = TRUE,
+                                       index = "jaccard") %>%
+  as_tibble()
+
+# Collate results into a single tibble
+beta_site.date <- bind_rows(beta_site.date.2010, beta_site.date.2011, beta_site.date.2012)
+
+# Save to file to avoid redoing that long step
+write_csv(beta_site.date, "./output/beta_site.date.csv")
+#beta_site.date <- read_csv("./output/beta_site.date.csv")
+
+# Add site data
+beta_mit_data <- beta_site.date %>%
+  separate(i, into = c("site1", "date1"), sep = "_") %>%
+  separate(j, into = c("site2", "date2"), sep = "_") %>%
+  mutate(date1 = ymd(date1),
+         date2 = ymd(date2),
+         year1 = factor(year(date1)),
+         year2 = factor(year(date2)),
+         yday1 = yday(date1),
+         yday2 = yday(date2)) %>%
+  left_join(site_data, by = c("site1" = "site")) %>%
   dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
@@
   ))
-
-beta_site.date.2011 <- betalinkr_multi(array_site.date.2011, 
-                                       partition.st = TRUE, 
-                                       partition.rr = TRUE) %>%
-  as_tibble() %>%
-  separate(i, into = c("site1", "year1", "yday1"), sep = "_") %>%
-  separate(j, into = c("site2", "year2", "yday2"), sep = "_") %>%
-  left_join(site_data, by = c("site1" = "site"))%>%
-  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
-                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
-                elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
-  left_join(site_data, by = c("site2" = "site")) %>%
-  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
-                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
-                elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
-  mutate(site1 = factor(site1),
-         site2 = factor(site2),
-         year1 = factor(year1),
-         year2 = factor(year2),
-         yday1 = as.numeric(yday1),
-         yday2 = as.numeric(yday2),
-         elev.diff = abs(elev1 - elev2),
-         mean.elev = (elev1 + elev2)/2,
-         yday.diff = abs(yday1 - yday2),
-         man.diff = case_when(
-           man1 == man2 ~ FALSE,
-           man1 != man2 ~ TRUE
-         )) %>%
-  gather(metric, value, -c(site1, year1, yday1, site2, year2, yday2,
-                           elev1, elev2, elev.diff, yday.diff, 
-                           man1, man2, man.diff, mean.elev, 
-                           lon1, lat1, lon2, lat2)) %>%
-  mutate(metric.class = case_when(
-    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
-    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
-    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
-    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
-  ))
-
-beta_site.date.2012 <- betalinkr_multi(array_site.date.2012, 
-                                       partition.st = TRUE, 
-                                       partition.rr = TRUE) %>%
-  as_tibble() %>%
-  separate(i, into = c("site1", "year1", "yday1"), sep = "_") %>%
-  separate(j, into = c("site2", "year2", "yday2"), sep = "_") %>%
-  left_join(site_data, by = c("site1" = "site"))%>%
-  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
-                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
-                elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
-  left_join(site_data, by = c("site2" = "site")) %>%
-  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
-                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
-                elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
-  mutate(site1 = factor(site1),
-         site2 = factor(site2),
-         year1 = factor(year1),
-         year2 = factor(year2),
-         yday1 = as.numeric(yday1),
-         yday2 = as.numeric(yday2),
-         elev.diff = abs(elev1 - elev2),
-         mean.elev = (elev1 + elev2)/2,
-         yday.diff = abs(yday1 - yday2),
-         man.diff = case_when(
-           man1 == man2 ~ FALSE,
-           man1 != man2 ~ TRUE
-         )) %>%
-  gather(metric, value, -c(site1, year1, yday1, site2, year2, yday2,
-                           elev1, elev2, elev.diff, yday.diff, 
-                           man1, man2, man.diff, mean.elev, 
-                           lon1, lat1, lon2, lat2)) %>%
-  mutate(metric.class = case_when(
-    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
-    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
-    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
-    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
-  ))
-
-beta_site.date <- bind_rows(beta_site.date.2010, beta_site.date.2011, beta_site.date.2012)
-beta_site.date_gdm.prep <- beta_site.date %>%
-  filter(metric == "WN")
-
-write_csv(beta_site.date, "../analyses/output/beta_site.date.csv")
-beta_site.date <- read_csv("../analyses/output/beta_site.date.csv") %>%
-  mutate(year1 = factor(year1),
-         year2 = factor(year2),
-         metric = factor(metric))
 ```
@@
 ### Elevation diff
-ggplot(filter(beta_site.date, year1 == year2,
+ggplot(filter(beta_mit_data, year1 == year2 &
+                site1 != site2 &
                 metric != "S" & 
@@
        aes(elev.diff, value, color = metric)) +
-  geom_point(alpha = 0.15, size = 0.5) +
+  geom_point(alpha = 0.25, pch = 16) +
   geom_smooth() +
@@
 ### Time diff
-ggplot(filter(beta_site.date, year1 == year2,
+ggplot(filter(beta_mit_data, year1 == year2 &
+                site1 != site2 &
                 metric != "S" & 
@@
        aes(yday.diff, value, color = metric)) +
-  geom_point(alpha = 0.15, size = 0.5) +
+  geom_point(alpha = 0.25, pch = 16) +
   geom_smooth() +
@@
 ### Elevation diff within 1 wk time diff
-ggplot(filter(beta_site.date, year1 == year2,
+ggplot(filter(beta_mit_data, year1 == year2 &
+                site1 != site2 &
                 metric != "S" & 
@@
        aes(elev.diff, value, color = metric)) +
-  geom_point(alpha = 0.15, size = 0.5) +
+  geom_point(alpha = 0.15, pch = 16, size = 1) +
   geom_smooth() +
@@
 
-ggplot(filter(beta_site.date, year1 == year2,
+### Time diff within 250 m elev diff
+ggplot(filter(beta_mit_data, year1 == year2 &
+                site1 != site2 &
                 metric != "S" & 
                 metric.class != "OS.partition" &
-                yday.diff <= 7), 
-       aes(elev.diff, value, color = metric)) +
-  geom_point(alpha = 0.15, size = 0.5) +
-  geom_smooth() +
-  facet_grid(metric.class ~ year1) +
-  theme_abyss()
-
-### Time diff within 250 m elev diff
-ggplot(filter(beta_site.date, year1 == year2,
-                metric != "S" & 
-                metric.class != "OS.partition" &
-                elev.diff <= 7), 
+                elev.diff <= 250), 
        aes(yday.diff, value, color = metric)) +
-  geom_point(alpha = 0.15, size = 0.5) +
+  geom_point(alpha = 0.15, pch = 16, size = 1) +
   geom_smooth() +
@@
   theme_light()
-
-ggplot(filter(beta_site.date, year1 == year2,
-                metric != "S" & 
-                metric.class != "OS.partition" &
-                elev.diff <= 7), 
-       aes(yday.diff, value, color = metric)) +
-  geom_point(alpha = 0.15, size = 0.5) +
-  geom_smooth() +
-  facet_grid(metric.class ~ year1) +
-  theme_abyss()
 
@@
                     family = "quasibinomial",
-                    data = filter(beta_site.date, year1 == year2 &
+                    data = filter(beta_mit_data, year1 == year2 &
                                     metric == "WN"),
@@
                      te(yday.diff, elev.diff, k = c(10, 10)),
-                   data = filter(beta_site.date, year1 == year2 &
+                   data = filter(beta_mit_data, year1 == year2 &
                                    metric == "WN.rich"),
@@
 
-### GDM analysis. I don't think p-values would add anything to the interpretation I get from the plotted uncertainty, so it's okay that I can't get gdm.varImp() to work.
+### GDM analysis.
 ```{r}
 ### 2010
-sitepair_2010.WN <- beta_site.date %>%
+sitepair_2010.WN <- beta_mit_data %>%
   ungroup() %>%
-  filter(year1 == year2 & year1 == 2010) %>%
+  filter(year1 == year2 & year1 == 2010 & site1 != site2) %>% # I don't think we want to compare samples through within sites, since this introduces a new form of non-independence.
   filter(metric == "WN") %>%
@@
 ### 2011
-sitepair_2011.WN <- beta_site.date %>%
+sitepair_2011.WN <- beta_mit_data %>%
   ungroup() %>%
-  filter(year1 == year2 & year1 == 2011) %>%
+  filter(year1 == year2 & year1 == 2011 & site1 != site2) %>%
   filter(metric == "WN") %>%
@@
 ### 2012
-sitepair_2012.WN <- beta_site.date %>%
+sitepair_2012.WN <- beta_mit_data %>%
   ungroup() %>%
-  filter(year1 == year2 & year1 == 2012) %>%
+  filter(year1 == year2 & year1 == 2012 & site1 != site2) %>%
   filter(metric == "WN") %>%
@@
   summarize(cover = sum(flower.cover)) %>%
+  mutate(cover = rep(1, n())) %>% # this makes the data binary, which I think is what I want here; otherwise, just comment out
   spread(plant.sp, cover) %>%
@@
 survey_sitepair.2012 <- formatsitepair(survey_bioData.2012, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
-                              XColumn = "lon", YColumn = "lat", predData = survey_predData.2012)
+                              XColumn = "lon", YColumn = "lat", predData = survey_predData.2012, abundance = FALSE)
 
@@
 
-predData.2011 <- survey %>%
+survey_predData.2011 <- survey %>%
   mutate(year = year(date),
@@
 survey_sitepair.2011 <- formatsitepair(survey_bioData.2011, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
-                              XColumn = "lon", YColumn = "lat", predData = predData.2011)
+                              XColumn = "lon", YColumn = "lat", predData = survey_predData.2011)
 
@@
 plotUncertainty(bb_sitepair.2012, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
+
+                                                 Drosophila?
 ```

### 2020-09-10 analyses/beta_diversity_v2.Rmd
haaaands

@@
                 metric != "S" & 
-                metric.class != "OS.partition"), 
+                !metric.class %in% c("OS.partition", "WN.partition")), 
        aes(elev.diff, value, color = metric)) +
@@
   facet_grid(metric.class ~ year1) +
+  scale_color_brewer(type = "qual", palette = 6) +
   theme_abyss()
@@
                 metric != "S" & 
-                metric.class != "OS.partition"), 
+                !metric.class %in% c("OS.partition", "WN.partition")), 
        aes(yday.diff, value, color = metric)) +
@@
                 metric != "S" & 
-                metric.class != "OS.partition" &
+                !metric.class %in% c("OS.partition", "WN.partition") &
                 yday.diff <= 7), 
        aes(elev.diff, value, color = metric)) +
-  geom_point(alpha = 0.15, pch = 16, size = 1) +
+  geom_point(alpha = 0.1, pch = 16, size = 1) +
   geom_smooth() +
   facet_wrap(~metric.class) +
+  facet_grid(metric.class ~ year1) +
+  scale_color_brewer(type = "qual", palette = 6) +
   theme_light()
@@
                 metric != "S" & 
-                metric.class != "OS.partition" &
+                !metric.class %in% c("OS.partition", "WN.partition") &
                 elev.diff <= 250), 
        aes(yday.diff, value, color = metric)) +
-  geom_point(alpha = 0.15, pch = 16, size = 1) +
+  geom_point(alpha = 0.1, pch = 16, size = 1) +
   geom_smooth() +
   facet_wrap(~metric.class) +
+  scale_color_brewer(type = "qual", palette = 6) +
   theme_light()
-
-### 2D ElevationDiff*TimeDiff
-gam_betadiv <- bamV(value ~ te(elev.diff, yday.diff, 
-                               bs = c("te", "te"),
-                               k = c(10, 10)),
-                    family = "quasibinomial",
-                    data = filter(beta_mit_data, year1 == year2 &
-                                    metric == "WN"),
-                    method = "REML")
-
-gam_betadiv <- bam(value ~ 
-                     s(year1, bs = "re") +
-                     te(yday.diff, elev.diff, k = c(10, 10)),
-                   data = filter(beta_mit_data, year1 == year2 &
-                                   metric == "WN.rich"),
-                   family = "quasibinomial",
-                   discrete = TRUE,
-                   method = "fREML") %>% getViz()
-
-check.gamViz(gam_betadiv)
-plot(gam_betadiv)
-
-ggplot(filter(beta_site.date, year1 == year2 & metric == "WN"),
-       aes(value)) +
-  geom_histogram()
 ```

### 2020-09-10 analyses/BB_abund_GAM.Rmd
rearranged directories

--- 2020-07-08 analyses/BB_abund_GAM.Rmd
@@
 ---
-title: "preference"
+title: "BB_abund_GAM"
 output: html_document
@@
 gdd <- read_csv("../processed_data/climate.csv") %>%
-  mutate(year = factor(year)) %>%
-  dplyr::select(site, year, date, gdd.cum) %>%
-  semi_join(read_csv("../processed_data/network.csv"), by = "site")
+  mutate(year = factor(year)) %>% # make year a factor
+  dplyr::select(site, year, date, gdd.cum) %>% # drop unused columns
+  semi_join(read_csv("../processed_data/network.csv"), by = "site") # drop sites that were not samples
 ```
@@
 site_data <- read_csv("../processed_data/site_data.csv") %>%
-  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
+  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")), # rename in English
          elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))
@@
 net <- read_csv("../processed_data/network.csv") %>%
-  dplyr::select(site, date, bb.sp, plant.sp)
+  dplyr::select(site, date, bb.sp, plant.sp) # drop unused columns
 ```
 
-### FL and BB abundance / diversity
+### BB abundance / diversity
 ```{r echo=FALSE, include=FALSE}
@@
   dplyr::select(site, date, bb.sp) %>%
-  group_by(site, date, bb.sp) %>% # group by bb.sp*transect (and other variables we want to retain)
-  summarize(bb.abund = n()) %>% # get bb.sp*transect abundance
-  ungroup %>%
+  group_by(site, date, bb.sp) %>% # 
+  summarize(bb.abund = n()) %>% # get per transect bb abundance
+  ungroup() %>%
   complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
-  mutate(year = factor(year(date))) %>%
-  left_join(site_data) %>%
-  mutate(yday = yday(date),
+  semi_join(net, by = c("site", "date")) %>% # drop site-dates that did not exist in the data set
+  left_join(site_data) %>% # add site data
+  mutate(year = factor(year(date)),
+         yday = yday(date),
          bb.sp = factor(bb.sp),
-         management = factor(management)) %>%
-  semi_join(net, by = c("site", "date")) %>%
+         management = factor(management),
+         site = factor(site),
+         transect = factor(transect)) %>%
   group_by(year, bb.sp) %>%
@@
 
-GAM modeling of abundance per bb species
+## Global GAM: this is more elegant than breaking it down by year, but my concern is that the model will be totally dominated by the 2012 data, since I have so many more samples that year.
 ```{r}
-### Call models
-bb_abund.2010 <- bam(bb.abund ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
+bb_abund.global <- bam(bb.abund ~
+                         s(bb.sp, bs = "re") +
+                         s(year, bs = "re") +
+                         s(site, bs = "re") + ### I think I need to include site as a random effect because if I don't the pseudoreplication of elevation is pretty extreme, especially in 2012 where I have lots of repeated measures.
+I am not, for example, measuring floral abundance at 1000 m 20 independent times; I'm measuring it 20 times at the same spot.
+This does, of  course, create identifiability issues insofar as site and elevation are concurved.
+But I think I need to accept these problems.
+                         te(yday, elev.mean, 
                                    by = bb.sp,
-                                                            bs= c("tp", "tp"), 
-                                                            k=c(10, 10), m=1),
+                                   #id = 0,
+                                   bs= c("gp", "tp"), 
+                                   k=c(8, 8), 
+                                   m=1) +
+                         te(yday, elev.mean, 
+                                   by = year,
+                                   #id = 0,
+                                   bs= c("gp", "tp"), 
+                                   k=c(4, 4), 
+                                   m=2),
+                       data = filter(bb_abund_sp, bb.total.abund >= 5),
+                       discrete = TRUE,
+                       family = "tw",
+                       select = TRUE,
+                       method = "fREML") %>% getViz(nsim = 500)
+
+bb_abund.global2 <- gam(bb.abund ~
+                s(bb.sp, bs = "re") +
+                s(year, bs = "re") +
+                s(site, bs = "re") +
+                te(yday, elev.mean, 
+                          by = bb.sp,
+                          #id = 0,
+                          bs= c("gp", "tp"), 
+                          k=c(6, 6), 
+                          m=1) +
+                te(yday, elev.mean, 
+                          by = year,
+                          #id = 0,
+                          bs= c("gp", "tp"), 
+                          k=c(4, 4), 
+                          m=2),
+              data = filter(bb_abund_sp, bb.total.abund >= 5 & bb.sp %in% c("bss", "pasc", "prat", "pyre")),
+              discrete = FALSE,
+              family = ziP(),
+              select = TRUE,
+              method = "REML") %>% getViz(nsim = 500)
+
+ziptest <- gam(bb.abund ~
+                t2(yday, elev.mean, 
+                          by = bb.sp,
+                          #id = 0,
+                          bs= c("gp", "tp"), 
+                          k=c(3, 3), 
+                          m=1),
+              data = filter(bb_abund_sp, bb.total.abund >= 5),
+              #discrete = FALSE,
+              family = ziP(),
+              select = TRUE,
+              method = "REML") %>% getViz(nsim = 50)
+
+check.gamViz(bb_abund.global)
+check.gamViz(bb_abund.global2)
+check.gamViz(ziptest)
+
+bb_abund.global.sum <- summary(bb_abund.global)
+bb_abund.global2.sum <- summary(bb_abund.global2)
+
+AIC(bb_abund.global, bb_abund.global2)
+
+print(plot(bb_abund.global, allTerms = TRUE), pages = 3)
+print(plot(bb_abund.global2, allTerms = TRUE), pages = 3)
+
+bb_abund.global.ck <- check2D(bb_abund.global, x1 = "yday", x2 = "elev.mean")
+bb_abund.global.ck + l_gridCheck2D(gridFun = mean)
+```
+
+## Annual GAMs
+### My argument for not including site as a random effect is that it would create identifiability issues with the elevation variable that I am actally interested in.
+This will probably cause the p-values to be underestimated, maybe to the point of being uninformative.
+But I don't think I care.
+My goal is not to establish inferentially that time and elevation affect BB abundance; I want to say how they do.
+And I don't want to obscure the "how" for the sake of inferential scruples.
+```{r}
+### 2010
+bb_abund.2010 <- bam(bb.abund ~ 
+                       s(bb.sp, bs = "re") + 
+                       te(yday, elev.mean, 
+                          by = bb.sp,
+                          bs= c("gp", "cr"), 
+                          k=c(8, 8), 
+                          m=2),
                      data = filter(bb_abund_sp, year == 2010 & bb.total.abund >= 5),
                      discrete = TRUE,
-                        family = "poisson",
+                     family = "tw",
                      select = TRUE,
-                        method = "fREML") %>% getViz()
+                     method = "fREML") %>% getViz(nsim = 500)
 
@@
 bb_abund.2010.sum <- summary(bb_abund.2010)
-print(plot(bb_abund.2010, allTerms = TRUE), pages = 4)
+print(plot(bb_abund.2010, allTerms = TRUE), pages = 1)
+bb_abund.2010.ck <- check2D(bb_abund.2010, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals
 
-bb_abund.2011 <- bam(bb.abund ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
+### 2011
+bb_abund.2011 <- bam(bb.abund ~ 
+                       s(bb.sp, bs = "re") + 
+                       te(yday, elev.mean, 
                           by = bb.sp,
-                                                            bs= c("tp", "tp"), 
-                                                            k=c(10, 10), m=1),
+                          bs= c("gp", "cr"), 
+                          k=c(8, 8), 
+                          m=2),
                      data = filter(bb_abund_sp, year == 2011 & bb.total.abund >= 5),
                      discrete = TRUE,
-                        family = "poisson",
+                     family = "tw",
                      select = TRUE,
-                        method = "fREML") %>% getViz()
+                     method = "fREML") %>% getViz(nsim = 500)
 
@@
 bb_abund.2011.sum <- summary(bb_abund.2011)
-print(plot(bb_abund.2011 , allTerms = TRUE), pages = 4)
+print(plot(bb_abund.2011, allTerms = TRUE), pages = 1)
+bb_abund.2011.ck <- check2D(bb_abund.2011, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals
 
-bb_abund.2012 <- bam(bb.abund ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
+### 2012
+bb_abund.2012 <- bam(bb.abund ~ 
+                       s(bb.sp, bs = "re") + 
+                       te(yday, elev.mean, 
                           by = bb.sp,
-                                                            bs= c("tp", "tp"), 
-                                                            k=c(10, 10), m=1),
+                          bs= c("gp", "cr"), 
+                          k=c(8, 8), 
+                          m=2),
                      data = filter(bb_abund_sp, year == 2012 & bb.total.abund >= 5),
                      discrete = TRUE,
-                        family = "poisson",
+                     family = "tw",
                      select = TRUE,
-                        method = "fREML") %>% getViz()
+                     method = "fREML") %>% getViz(nsim = 500)
 
@@
 bb_abund.2012.sum <- summary(bb_abund.2012)
-print(plot(bb_abund.2012 , allTerms = TRUE), pages = 4)
-plot(bb_abund.2012)
-
+print(plot(bb_abund.2012, allTerms = TRUE), pages = 1)
+bb_abund.2012.ck <- check2D(bb_abund.2012, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals
 ```
 
+### Tabulate plots
 ```{r}
-bb_abund.2012.pred <- predict(bb_abund.2012)
+bss2010 <- plot(sm(bb_abund.2010, 2)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. terrestris/lucorum")), x = NULL, y = NULL)
+bss2011 <- plot(sm(bb_abund.2011, 2)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. terrestris/lucorum")), x = NULL, y = NULL)
+bss2012 <- plot(sm(bb_abund.2012, 2)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. terrestris/lucorum")), x = NULL, y = NULL)
 
-ggplot(bb_abund.2012.pred, aes(yday, elev.mean, fit)) +
-  geom_raster(aes(fill = exp(fit))) +
-  #geom_contour(aes(z = fit), color = "black") +
-  geom_point(data = filter(fl_abund, year == 2012), 
-             aes(yday, elev.mean, size = fl.abund),
-             color = "white", alpha = 0.25) +
-  scale_fill_viridis_c()
+gers2010 <- plot(sm(bb_abund.2010, 3)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. gerstaeckeri")), x = NULL, y = NULL)
+gers2011 <- ggplot() #too rare
+gers2012 <- plot(sm(bb_abund.2012, 3)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. gerstaeckeri")), x = NULL, y = NULL)
+
+hort2010 <- plot(sm(bb_abund.2010, 4)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. hortorum")), x = NULL, y = NULL)
+hort2011 <- plot(sm(bb_abund.2011, 3)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. hortorum")), x = NULL, y = NULL)
+hort2012 <- plot(sm(bb_abund.2012, 4)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. hortorum")), x = NULL, y = NULL)
+
+hypn2010 <- plot(sm(bb_abund.2010, 5)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. hypnorum")), x = NULL, y = NULL)
+hypn2011 <- plot(sm(bb_abund.2011, 4)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. hypnorum")), x = NULL, y = NULL)
+hypn2012 <- ggplot()  #too rare
+
+jone2010 <- plot(sm(bb_abund.2010, 6)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. jonellus")), x = NULL, y = NULL)
+jone2011 <- plot(sm(bb_abund.2011, 5)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. jonellus")), x = NULL, y = NULL)
+jone2012 <- plot(sm(bb_abund.2012, 5)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. jonellus")), x = NULL, y = NULL)
+
+lapi2010 <- ggplot() # too rare 
+lapi2011 <- plot(sm(bb_abund.2011, 6)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. lapidarius")), x = NULL, y = NULL)
+lapi2012 <- plot(sm(bb_abund.2012, 6)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. lapidarius")), x = NULL, y = NULL)
+
+mend2010 <- plot(sm(bb_abund.2010, 7)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. mendax")), x = NULL, y = NULL)
+mend2011 <- plot(sm(bb_abund.2011, 7)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. mendax")), x = NULL, y = NULL)
+mend2012 <- plot(sm(bb_abund.2012, 7)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. mendax")), x = NULL, y = NULL)
+
+mont2010 <- plot(sm(bb_abund.2010, 8)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. monticola")), x = NULL, y = NULL)
+mont2011 <- plot(sm(bb_abund.2011, 8)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE)+
+  l_points() +
+  labs(title = expression(italic("B. monticola")), x = NULL, y = NULL)
+mont2012 <- plot(sm(bb_abund.2012, 8)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. monticola")), x = NULL, y = NULL)
+
+muci2010 <- plot(sm(bb_abund.2010, 9)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. mucidis")), x = NULL, y = NULL)
+muci2011 <- plot(sm(bb_abund.2011, 9)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. mucidis")), x = NULL, y = NULL)
+muci2012 <- plot(sm(bb_abund.2012, 9)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. mucidis")), x = NULL, y = NULL)
+
+pasc2010 <- plot(sm(bb_abund.2010, 10)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. pascuorum")), x = NULL, y = NULL)
+pasc2011 <- plot(sm(bb_abund.2011, 10)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. pascuorum")), x = NULL, y = NULL)
+pasc2012 <- plot(sm(bb_abund.2012, 10)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. pascuorum")), x = NULL, y = NULL)
+
+prat2010 <- plot(sm(bb_abund.2010, 11)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. pratorum")), x = NULL, y = NULL)
+prat2011 <- plot(sm(bb_abund.2011, 11)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. pratorum")), x = NULL, y = NULL)
+prat2012 <- plot(sm(bb_abund.2012, 11)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. pratorum")), x = NULL, y = NULL)
+
+psit2010 <- plot(sm(bb_abund.2010, 12)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. psithyrus")), x = NULL, y = NULL)
+psit2011 <- plot(sm(bb_abund.2011, 12)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. psithyrus")), x = NULL, y = NULL)
+psit2012 <- plot(sm(bb_abund.2012, 12)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. psithyrus")), x = NULL, y = NULL)
+
+pyre2010 <- plot(sm(bb_abund.2010, 13)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. pyrenaeus")), x = NULL, y = NULL)
+pyre2011 <- plot(sm(bb_abund.2011, 13)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. pyrenaeus")), x = NULL, y = NULL)
+pyre2012 <- plot(sm(bb_abund.2012, 13)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. pyrenaeus")), x = NULL, y = NULL)
+
+soro2010 <- plot(sm(bb_abund.2010, 14)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. soroensis")), x = NULL, y = NULL)
+soro2011 <- plot(sm(bb_abund.2011, 14)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. soroensis")), x = NULL, y = NULL)
+soro2012 <- plot(sm(bb_abund.2012, 14)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. soroensis")), x = NULL, y = NULL)
+
+wurf2010 <- plot(sm(bb_abund.2010, 15)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. wurflenii")), x = NULL, y = NULL)
+wurf2011 <- plot(sm(bb_abund.2011, 15)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  guides(fill = FALSE) +
+  l_points() +
+  labs(title = expression(italic("B. wurflenii")), x = NULL, y = NULL)
+wurf2012 <- plot(sm(bb_abund.2012, 15)) +
+  l_fitRaster() +
+  l_fitContour(size = 0.1) +
+  l_points() +
+  guides(fill = FALSE) +
+  labs(title = expression(italic("B. wurflenii")), x = NULL, y = NULL)
+
+pdf("./output/BB_abund_major.pdf", height = 11, width = 8.5)
+major_sp <- gridPrint(bss2010, bss2011, bss2012, 
+                      pasc2010, pasc2011, pasc2012,
+                      prat2010, prat2011, prat2012,
+                      psit2010, psit2011, psit2012,
+                      soro2010, soro2011, soro2012,
+                      wurf2010, wurf2011, wurf2012, ncol = 3)
+dev.off()
+
+pdf("./output/BB_abund_highelev.pdf", height = 5.5, width = 8.5)
+high_elev_sp <- gridPrint(pyre2010, pyre2011, pyre2012, 
+                          mend2010, mend2011, mend2012,
+                          mont2010, mont2011, mont2012, ncol = 3)
+dev.off()
+
+pdf("./output/BB_abund_minor.pdf", height = 11, width = 8.5)
+minor_sp <- gridPrint(gers2010, gers2011, gers2012,
+                      hort2010, hort2011, hort2012,
+                      hypn2010, hypn2011, hypn2012,
+                      jone2010, jone2011, jone2012,
+                      lapi2010, lapi2011, lapi2012,
+                      muci2010, muci2011, muci2012, ncol = 3)
+dev.off()
 ```
+
+## Bumble bee species richness
+```{r}
+bb_rich <- bb_abund_sp %>%
+  filter(bb.abund > 0) %>%
+  group_by(site, date, yday, year, elev.mean, management) %>%
+  summarize(bb.rich = length(unique(bb.sp)))
+```
+
+## Global species richness GAM (no need to separate the years, meiner Meinung nach)
+```{r}
+### 2010
+bb_rich.global <- bam(bb.rich ~ 
+                        te(yday, elev.mean, 
+                           by = year,
+                           bs= c("gp", "cr"), 
+                           k=c(6, 6), 
+                           m=2),
+                      data = filter(bb_rich),
+                      discrete = TRUE,
+                      family = "gaussian",
+                      select = TRUE,
+                      method = "fREML") %>% getViz(nsim = 500)
+
+check.gamViz(bb_rich.global)
+bb_rich.global.sum <- summary(bb_rich.global)
+print(plot(bb_rich.global, allTerms = TRUE), pages = 1)
+bb_rich.global.ck <- check2D(bb_rich.global, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals
+```

### 2020-09-10 analyses/FL_abund_GAM.Rmd
rearranged directories

@@
 ---
-title: "preference"
+title: "FL_abund_GAM"
 output: html_document
@@
 site_data <- read_csv("../processed_data/site_data.csv") %>%
-  #dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
-  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
+  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")), # rename in English
          elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))
@@
 net <- read_csv("../processed_data/network.csv") %>%
-  dplyr::select(site, date, bb.sp, plant.sp)
+  dplyr::select(site, date, bb.sp, plant.sp) # drop unused columns
 
@@
 survey <- read_csv("../processed_data/floral_survey.csv") %>%
-  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
+  semi_join(net, by = "plant.sp") %>% # Consider only visited plant species (anytime anywhere, not specifically for a given site*date)
   dplyr::select(site, date, plant.sp, flower.cover)
 
+### Identify duplicate observations; these may be errors or intentionally duplicated (see correspondence with Kalli)
 survey_dups <- survey %>%
@@
   filter(dups > 1)
-
-dup_species <- survey_dups %>%
-  group_by(plant.sp) %>%
-  mutate(dups = n()/2) %>%
-  group_by(site, date, plant.sp, dups) %>%
-  mutate(different = length(unique(flower.cover)) -1) %>%
-  group_by(site, date, plant.sp, dups) %>%
-  summarize(different = mean(different)) %>%
-  group_by(plant.sp, dups) %>%
-  summarize(different.dups = sum(different)) %>% 
-  arrange(-dups)
 
@@
   mutate(year = factor(year(date))) %>%
-  semi_join(net, by = c("site", "date")) # only include site-dates that actually occurred
+  semi_join(shared_transects, by = c("site", "date")) # only include site-dates that actually occurred
 
@@
                                       #s(management, bs = "re") +
-                                      s(site, bs = "re") +
+                                      #s(site, bs = "re") +
                                       te(yday, elev.mean, 
@@
 
+weighted_fl_abundance.2010b <- bam(fl.abund.wt + 0.001 ~
+                                      s(bb.sp, bs = "re") +
+                                      #s(management, bs = "re") +
+                                      #s(site, bs = "re") +
+                                      te(yday, elev.mean, 
+                                                by = bb.sp,
+                                                #id = 0,
+                                                bs= c("gp", "tp"), 
+                                                k=c(15, 15), 
+                                                m=2),
+                                    data = filter(fl_abund_wt, year == 2010),
+                                    discrete = TRUE,
+                                    family = "tw",
+                                    select = TRUE,
+                                    method = "fREML") %>% getViz(nsim = 500)
+
 check.gamViz(weighted_fl_abundance.2010)
+check.gamViz(weighted_fl_abundance.2010b)
+
 weighted_fl_abundance.2010.sum <- summary(weighted_fl_abundance.2010)
 print(plot(weighted_fl_abundance.2010, allTerms = TRUE), pages = 4)
+print(plot(weighted_fl_abundance.2010b, allTerms = TRUE), pages = 4)
+
+plot(sm(weighted_fl_abundance.2010, 2))
+plot(sm(weighted_fl_abundance.2010b, 2))
+
 weighted_fl_abundance.2010.ck <- check2D(weighted_fl_abundance.2010, 
                                          x1 = "yday", x2 = "elev.mean")
+weighted_fl_abundance.2010b.ck <- check2D(weighted_fl_abundance.2010b, 
+                                         x1 = "yday", x2 = "elev.mean")
+
 weighted_fl_abundance.2010.ck + l_gridCheck2D(gridFun = mean)

### 2020-09-10 analyses/beta_diversity_v2.Rmd
rearranged directories

@@
          elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) %>% # turn elav.class into an ordered factor
-  semi_join(read_csv("../processed_data/network.csv"), by = "site")
+  semi_join(read_csv("../processed_data/network.csv"), by = "site") # drop sites that were not sampled
 ```
@@
 ```{r}
-net <- read_csv("../processed_data/network.csv")%>%
-  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
-  unite(site.year.day, site, year, dayofyear, sep = "_", remove = FALSE) %>%
-  mutate(year = factor(year)) %>% # convert year to factor
-  group_by(site, date) %>%
-  arrange(site, date)
+net <- read_csv("../processed_data/network.csv") %>%
+  unite(site.year.day, site, year, dayofyear, sep = "_", remove = FALSE) %>% # create site.date field that can be split up again later
+  mutate(year = factor(year)) # convert year to factor
 
 survey <- read_csv("../processed_data/floral_survey.csv") %>%
-  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
+  semi_join(net, by = "plant.sp") %>% # Consider only visited plant spp. (anytime, anywhere, not specifically for a given site*date)
   dplyr::select(site, date, plant.sp, flower.cover)
@@
 ```{r message=FALSE}
-# Create webs
-net2010 <- filter(net, year == 2010) %>%
+# Function to convert networks to webs arrays
+web_prep <- function(net, y) {
+  out <- net %>%
+    filter(year == y) %>%
     unite(site.date, c(site, date), sep = "_") %>%
@@
   frame2webs(type.out = "array")
+}
 
-net2011 <- filter(net, year == 2011) %>%
-  unite(site.date, c(site, date), sep = "_") %>%
-  group_by(site.date, bb.sp, plant.sp) %>%
-  summarize(freq = n()) %>%
-  group_by(site.date) %>%
-  mutate(bb.sp.count = length(unique(bb.sp)),
-         plant.sp.count = length(unique(plant.sp))) %>%
-  ungroup() %>%
-  filter(bb.sp.count > 1 & plant.sp.count > 1) %>% 
-  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.date, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs(type.out = "array")
+# Create webs
+net2010 <- web_prep(net, 2010)
+net2011 <- web_prep(net, 2011)
+net2012 <- web_prep(net, 2012)
 
-net2012 <- filter(net, year == 2012) %>%
-  unite(site.date, c(site, date), sep = "_") %>%
-  group_by(site.date, bb.sp, plant.sp) %>%
-  summarize(freq = n()) %>%
-  group_by(site.date) %>%
-  mutate(bb.sp.count = length(unique(bb.sp)),
-         plant.sp.count = length(unique(plant.sp))) %>%
-  ungroup() %>%
-  filter(bb.sp.count > 1 & plant.sp.count > 1) %>% 
-  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.date, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs(type.out = "array")
+# Function with betalinkr settings and conversion to tibble()
+betalinkr_set1 <- function(x) {
+  betalinkr_multi(x, partition.st = TRUE, partition.rr = TRUE, # partition both species turnover and richness/replacement
+                  binary = TRUE, # treating beta-diversity as binary is an appropriate complement to our abundance analysis
+                  index = "jaccard") %>% # the Jaccard index is best, I think, for binary analysis
+  as_tibble() # for convenience
+}
 
 # Calculate beta-diversity
-beta_site.date.2010 <- betalinkr_multi(net2010, 
-                                       partition.st = TRUE, 
-                                       partition.rr = TRUE,
-                                       binary = TRUE,
-                                       index = "jaccard") %>%
-  as_tibble()
-
-beta_site.date.2011 <- betalinkr_multi(net2011, 
-                                       partition.st = TRUE, 
-                                       partition.rr = TRUE,
-                                       binary = TRUE,
-                                       index = "jaccard") %>%
-  as_tibble()
-
-beta_site.date.2012 <- betalinkr_multi(net2012, 
-                                       partition.st = TRUE, 
-                                       partition.rr = TRUE,
-                                       binary = TRUE,
-                                       index = "jaccard") %>%
-  as_tibble()
+beta_site.date.2010 <- betalinkr_set1(net2010)
+beta_site.date.2011 <- betalinkr_set1(net2011) 
+beta_site.date.2012 <- betalinkr_set1(net2012)
 
@@
 
-### Plot results. There's nothing a Mantel test can tell me that I don't already know from the combination of plotting and GDM analysis.
+### Plot interaction beta-diversity against elevation difference and time difference
 ```{r}
-### Elevation diff
+# Set palette
+p <- c("OS" = "deepskyblue", "ST" = "green3", 
+       "ST.h" = "firebrick1", "ST.l" = "darkorchid", 
+       "ST.lh" = "gold", "WN" = "black")
+
+# Elevation diff
 ggplot(filter(beta_mit_data, year1 == year2 &
@@
   facet_grid(metric.class ~ year1) +
-  scale_color_brewer(type = "qual", palette = 6) +
+  scale_color_manual(values = p) +
   theme_abyss()
 
-### Time diff
+# Time diff
 ggplot(filter(beta_mit_data, year1 == year2 &
@@
   facet_grid(metric.class ~ year1) +
-  scale_color_brewer(type = "qual", palette = 6) +
+  scale_color_manual(values = p) +
   theme_abyss()
 
-### Elevation diff within 1 wk time diff
+# Elevation diff within 1 wk time diff
 ggplot(filter(beta_mit_data, year1 == year2 &
@@
   geom_smooth() +
-  facet_wrap(~metric.class) +
   facet_grid(metric.class ~ year1) +
-  scale_color_brewer(type = "qual", palette = 6) +
+  scale_color_manual(values = p) +
   theme_light()
 
-### Time diff within 250 m elev diff
+# Time diff within 250 m elev diff
 ggplot(filter(beta_mit_data, year1 == year2 &
@@
   geom_smooth() +
-  facet_wrap(~metric.class) +
-  scale_color_brewer(type = "qual", palette = 6) +
+  facet_grid(metric.class ~ year1) +
+  scale_color_manual(values = p) +
   theme_light()
@@
 
-### GDM analysis.
-```{r}
-### 2010
-sitepair_2010.WN <- beta_mit_data %>%
+### GDM: Interactions
+```{r message=FALSE, warning=FALSE}
+# Site_pair prep function
+site_pair_prep_net <- function(dat, y) {
+  beta_mit_data %>%
     ungroup() %>%
-  filter(year1 == year2 & year1 == 2010 & site1 != site2) %>% # I don't think we want to compare samples through within sites, since this introduces a new form of non-independence.
-  filter(metric == "WN") %>%
-  mutate(weights = rep(1, n())) %>%
+    filter(year1 == year2 & 
+             year1 == y & 
+             site1 != site2) %>% # I don't think we want to compare samples through within sites, introduces a new form of non-independence.
+    filter(metric == "WN") %>% # we'll focus for now on WN, but this could be done for partitions, too
+    mutate(weights = rep(1, n())) %>% # right now this analysis is unweighted; could weight by abundance, etc, if we wanted to
     dplyr::select(distance = value, weights,
@@
     data.frame()
+}
 
+# Prep site-pairs for GDM analysis
+### 2010
+sitepair_2010.WN <- site_pair_prep_net(beta_mit_data, 2010)
+sitepair_2011.WN <- site_pair_prep_net(beta_mit_data, 2011)
+sitepair_2012.WN <- site_pair_prep_net(beta_mit_data, 2012)
+
+# Run GDM analysis
 gdm_WN.2010 <- gdm(sitepair_2010.WN, geo = TRUE)
-gdm_WN.2010.imp <- gdm.varImp(sitepair_2010.WN, geo = TRUE)
+gdm_WN.2011 <- gdm(sitepair_2011.WN, geo = TRUE)
+gdm_WN.2012 <- gdm(sitepair_2012.WN, geo = TRUE)
 
 summary(gdm_WN.2010)
-plot(gdm_WN.2010)
-plotUncertainty(sitepair_2010.WN, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
+summary(gdm_WN.2011)
+summary(gdm_WN.2012)
 
-### 2011
-sitepair_2011.WN <- beta_mit_data %>%
-  ungroup() %>%
-  filter(year1 == year2 & year1 == 2011 & site1 != site2) %>%
-  filter(metric == "WN") %>%
-  mutate(weights = rep(1, n())) %>%
-  dplyr::select(distance = value, weights,
-         s1.xCoord = lon1, s1.yCoord = lat1, 
-         s2.xCoord = lon2, s2.yCoord = lat2, 
-         s1.elev = elev1, s1.yday = yday1,
-         s2.elev = elev2, s2.yday = yday2) %>%
-  data.frame()
-
-gdm_WN.2011 <- gdm(sitepair_2011.WN, geo = TRUE)
+# Permutation tests (slow)
+gdm_WN.2010.imp <- gdm.varImp(sitepair_2010.WN, geo = TRUE)
 gdm_WN.2011.imp <- gdm.varImp(sitepair_2011.WN, geo = TRUE)
-
-summary(gdm_WN.2011)
-plot(gdm_WN.2011)
-plotUncertainty(sitepair_2011.WN, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
-
-### 2012
-sitepair_2012.WN <- beta_mit_data %>%
-  ungroup() %>%
-  filter(year1 == year2 & year1 == 2012 & site1 != site2) %>%
-  filter(metric == "WN") %>%
-  mutate(weights = rep(1, n())) %>%
-  dplyr::select(distance = value, weights,
-         s1.xCoord = lon1, s1.yCoord = lat1, 
-         s2.xCoord = lon2, s2.yCoord = lat2, 
-         s1.elev = elev1, s1.yday = yday1,
-         s2.elev = elev2, s2.yday = yday2) %>%
-  data.frame()
-
-gdm_WN.2012 <- gdm(sitepair_2012.WN, geo = TRUE)
 gdm_WN.2012.imp <- gdm.varImp(sitepair_2012.WN, geo = TRUE)
 
-summary(gdm_WN.2012)
+# Plot
+plot(gdm_WN.2010)
+plot(gdm_WN.2011)
 plot(gdm_WN.2012)
+
+# Plot with confidence intervals (slow; also, I'm not sure about the current settings)
+plotUncertainty(sitepair_2010.WN, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
+plotUncertainty(sitepair_2011.WN, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
 plotUncertainty(sitepair_2012.WN, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
@@
 
-### Let's look at per-transect floral survey beta-diversity
+### GDM: Floral survey
 ```{r}
-### 2012
-survey_bioData.2012 <- survey %>%
+# Create functions for prepping data
+biodata_prep_survey <- function(surv, y) {
+  surv %>%
     mutate(year = year(date),
            yday = yday(date)) %>%
-  filter(flower.cover > 0 & year == "2012") %>%
+    filter(flower.cover > 0 & year == y) %>%
     mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
@@
     unite(site.day, site, yday, sep = "_")
+}
 
-survey_predData.2012 <- survey %>%
+predData_prep_survey <- function(surv, y) {
+  surv %>%
     mutate(year = year(date),
            yday = yday(date)) %>%
-  filter(flower.cover > 0 & year == "2012") %>%
+    filter(flower.cover > 0 & year == y) %>%
     dplyr::select(site, yday) %>%
@@
     dplyr::select(-site)
+}
 
-survey_sitepair.2012 <- formatsitepair(survey_bioData.2012, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
-                              XColumn = "lon", YColumn = "lat", predData = survey_predData.2012, abundance = FALSE)
+format_set <- function(bio, pred) {
+  formatsitepair(bio, bioFormat = 1, siteColumn = "site.day", 
+                 dist = "jaccard", XColumn = "lon", YColumn = "lat", 
+                 predData = pred, abundance = FALSE)
+}
 
-### Call gdm()
+# Prep input for GDM
+survey_bioData.2010 <- biodata_prep_survey(survey, 2010)
+survey_predData.2010 <- predData_prep_survey(survey, 2010)
+survey_sitepair.2010 <- format_set(survey_bioData.2010, survey_predData.2010)
+
+survey_bioData.2011 <- biodata_prep_survey(survey, 2011)
+survey_predData.2011 <- predData_prep_survey(survey, 2011)
+survey_sitepair.2011 <- format_set(survey_bioData.2011, survey_predData.2011)
+
+survey_bioData.2012 <- biodata_prep_survey(survey, 2012)
+survey_predData.2012 <- predData_prep_survey(survey, 2012)
+survey_sitepair.2012 <- format_set(survey_bioData.2012, survey_predData.2012)
+
+# Call GDMs
+gdm_survey.2010 <- gdm(survey_sitepair.2010, geo = TRUE)
+gdm_survey.2011 <- gdm(survey_sitepair.2011, geo = TRUE)
 gdm_survey.2012 <- gdm(survey_sitepair.2012, geo = TRUE)
+
+summary(gdm_survey.2010)
+summary(gdm_survey.2011)
+summary(gdm_survey.2012)
+
+# Permutations tests (slow)
+gdm_survey.2010.imp <- gdm.varImp(survey_sitepair.2010, geo = TRUE)
+gdm_survey.2011.imp <- gdm.varImp(survey_sitepair.2011, geo = TRUE)
 gdm_survey.2012.imp <- gdm.varImp(survey_sitepair.2012, geo = TRUE)
 
-summary(gdm_survey.2012)
+# Plot
+plot(gdm_survey.2010)
+plot(gdm_survey.2011)
 plot(gdm_survey.2012)
+
+# Plot with confidence intervals (slow)
+plotUncertainty(survey_sitepair.2010, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
+plotUncertainty(survey_sitepair.2011, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
 plotUncertainty(survey_sitepair.2012, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
-
-### 2011
-survey_bioData.2011 <- survey %>%
-  mutate(year = year(date),
-         yday = yday(date)) %>%
-  filter(flower.cover > 0 & year == "2011") %>%
-  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
-  group_by(site, yday, plant.sp) %>%
-  summarize(cover = sum(flower.cover)) %>%
-  mutate(cover = rep(1, n())) %>% # this makes the data binary, which I think is what I want here; otherwise, just comment out
-  spread(plant.sp, cover) %>%
-  replace(is.na(.), 0) %>%
-  arrange(site, yday) %>%
-  unite(site.day, site, yday, sep = "_")
-
-survey_predData.2011 <- survey %>%
-  mutate(year = year(date),
-         yday = yday(date)) %>%
-  filter(flower.cover > 0 & year == "2011") %>%
-  select(site, yday) %>%
-  left_join(site_data) %>%
-  ungroup() %>%
-  dplyr::select(site, elev.mean, lat, lon, yday) %>%
-  unique() %>%
-  arrange(site, yday) %>%
-  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
-  select(-site)
-
-survey_sitepair.2011 <- formatsitepair(survey_bioData.2011, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
-                              XColumn = "lon", YColumn = "lat", predData = survey_predData.2011)
-
-### Call gdm()
-gdm_survey.2011 <- gdm(survey_sitepair.2011, geo = TRUE)
-gdm_survey.2011.imp <- gdm.varImp(survey_sitepair.2011, geo = TRUE)
-
-summary(gdm_survey.2011)
-plot(gdm_survey.2011)
-plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
-
-### 2010
-survey_bioData.2010 <- survey %>%
-  mutate(year = year(date),
-         yday = yday(date)) %>%
-  filter(flower.cover > 0 & year == "2010") %>%
-  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
-  group_by(site, yday, plant.sp) %>%
-  summarize(cover = sum(flower.cover)) %>%
-  mutate(cover = rep(1, n())) %>% # this makes the data binary, which I think is what I want here; otherwise, just comment out
-  spread(plant.sp, cover) %>%
-  replace(is.na(.), 0) %>%
-  arrange(site, yday) %>%
-  unite(site.day, site, yday, sep = "_")
-
-predData.2010 <- survey %>%
-  mutate(year = year(date),
-         yday = yday(date)) %>%
-  filter(flower.cover > 0 & year == "2010") %>%
-  select(site, yday) %>%
-  left_join(site_data) %>%
-  ungroup() %>%
-  dplyr::select(site, elev.mean, lat, lon, yday) %>%
-  unique() %>%
-  arrange(site, yday) %>%
-  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
-  select(-site)
-
-survey_sitepair.2010 <- formatsitepair(survey_bioData.2010, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
-                              XColumn = "lon", YColumn = "lat", predData = predData.2010)
-
-### Call gdm()
-gdm_survey.2010 <- gdm(survey_sitepair.2010, geo = TRUE)
-gdm_survey.2010.imp <- gdm.varImp(survey_sitepair.2010, geo = TRUE)
-
-summary(gdm_survey.2010)
-plot(gdm_survey.2010)
-plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
 ```
 
-### And now Bombus diversity patterns
+### GDM: BB observations
 ```{r}
-### 2010
-BB_bioData.2010 <- net %>%
-  filter(year == 2010) %>%
+# Create functions for prepping data
+biodata_prep_BB <- function(net, y) {
+  net %>%
+    filter(year == y) %>%
     group_by(site, date, bb.sp) %>%
     summarize(count = n()) %>%
-  group_by(site, date) %>%
-  #mutate(prop = count/sum(count)) %>%
-  #group_by(site, bb.sp) %>%
-  # summarize(abund = sum(prop)) %>%
+    mutate(count = rep(1, n())) %>% # this makes the data binary, which I think is what I want here; otherwise, just comment out
     spread(bb.sp, count) %>%
@@
     select(-date)
+}
 
-BB_predData.2010 <- net %>%
-  filter(year == 2010) %>%
+predData_prep_BB <- function(net, y) {
+  net %>%
+    filter(year == y) %>%
     mutate(yday = yday(date)) %>%
@@
     dplyr::select(-site)
+}
 
-bb_sitepair.2010 <- formatsitepair(BB_bioData.2010, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
-                              XColumn = "lon", YColumn = "lat", predData = BB_predData.2010)
+# Prep input for GDM
+BB_bioData.2010 <- biodata_prep_BB(net, 2010)
+BB_predData.2010 <- predData_prep_BB(net, 2010)
+BB_sitepair.2010 <- format_set(BB_bioData.2010, BB_predData.2010)
 
-### Call gdm()
-gdm_bb.2010 <- gdm(bb_sitepair.2010, geo = TRUE)
-gdm_bb.2010.imp <- gdm.varImp(bb_sitepair.2010, geo = TRUE)
+BB_bioData.2011 <- biodata_prep_BB(net, 2011)
+BB_predData.2011 <- predData_prep_BB(net, 2011)
+BB_sitepair.2011 <- format_set(BB_bioData.2011, BB_predData.2011)
 
-summary(gdm_bb.2010)
-plot(gdm_bb.2010)
-plotUncertainty(bb_sitepair.2010, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
+BB_bioData.2012 <- biodata_prep_BB(net, 2012)
+BB_predData.2012 <- predData_prep_BB(net, 2012)
+BB_sitepair.2012 <- format_set(BB_bioData.2012, BB_predData.2012)
 
-### 2011
-BB_bioData.2011 <- net %>%
-  filter(year == 2011) %>%
-  group_by(site, date, bb.sp) %>%
-  summarize(count = n()) %>%
-  group_by(site, date) %>%
-  #mutate(prop = count/sum(count)) %>%
-  #group_by(site, bb.sp) %>%
-  # summarize(abund = sum(prop)) %>%
-  spread(bb.sp, count) %>%
-  replace(is.na(.), 0) %>%
-  mutate(yday = yday(date)) %>%
-  unite(site.day, site, yday, sep = "_", remove = "TRUE") %>%
-  ungroup() %>%
-  select(-date)
+# Call GDMs
+gdm_BB.2010 <- gdm(BB_sitepair.2010, geo = TRUE)
+gdm_BB.2011 <- gdm(BB_sitepair.2011, geo = TRUE)
+gdm_BB.2012 <- gdm(BB_sitepair.2012, geo = TRUE)
 
-BB_predData.2011 <- net %>%
-  filter(year == 2011) %>%
-  mutate(yday = yday(date)) %>%
-  dplyr::select(site, yday) %>%
-  left_join(site_data) %>%
-  ungroup() %>%
-  dplyr::select(site, elev.mean, lat, lon, yday) %>%
-  unique() %>%
-  arrange(site, yday) %>%
-  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
-  dplyr::select(-site)
+summary(gdm_BB.2010)
+summary(gdm_BB.2011)
+summary(gdm_BB.2012)
 
-bb_sitepair.2011 <- formatsitepair(BB_bioData.2011, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
-                              XColumn = "lon", YColumn = "lat", predData = BB_predData.2011)
+# Permutations tests (slow)
+gdm_BB.2010.imp <- gdm.varImp(BB_sitepair.2010, geo = TRUE)
+gdm_BB.2011.imp <- gdm.varImp(BB_sitepair.2011, geo = TRUE)
+gdm_BB.2012.imp <- gdm.varImp(BB_sitepair.2012, geo = TRUE)
 
-### Call gdm()
-gdm_bb.2011 <- gdm(bb_sitepair.2011, geo = TRUE)
-gdm.varImp(bb_sitepair.2011, geo = TRUE)
+# Plot
+plot(gdm_BB.2010)
+plot(gdm_BB.2011)
+plot(gdm_BB.2012)
 
-summary(gdm_bb.2011)
-plot(gdm_bb.2011)
-plotUncertainty(bb_sitepair.2011, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
-
-### 2012
-BB_bioData.2012 <- net %>%
-  filter(year == 2012) %>%
-  group_by(site, date, bb.sp) %>%
-  summarize(count = n()) %>%
-  group_by(site, date) %>%
-  #mutate(prop = count/sum(count)) %>%
-  #group_by(site, bb.sp) %>%
-  # summarize(abund = sum(prop)) %>%
-  spread(bb.sp, count) %>%
-  replace(is.na(.), 0) %>%
-  mutate(yday = yday(date)) %>%
-  unite(site.day, site, yday, sep = "_", remove = "TRUE") %>%
-  ungroup() %>%
-  select(-date)
-
-BB_predData.2012 <- net %>%
-  filter(year == 2012) %>%
-  mutate(yday = yday(date)) %>%
-  dplyr::select(site, yday) %>%
-  left_join(site_data) %>%
-  ungroup() %>%
-  dplyr::select(site, elev.mean, lat, lon, yday) %>%
-  unique() %>%
-  arrange(site, yday) %>%
-  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
-  dplyr::select(-site)
-
-bb_sitepair.2012 <- formatsitepair(BB_bioData.2012, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
-                              XColumn = "lon", YColumn = "lat", predData = BB_predData.2012)
-
-### Call gdm()
-gdm_bb.2012 <- gdm(bb_sitepair.2012, geo = TRUE)
-gdm.varImp(bb_sitepair.2012, geo = TRUE)
-
-summary(gdm_bb.2012)
-plot(gdm_bb.2012)
-plotUncertainty(bb_sitepair.2012, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
-
-                                                 Drosophila?
+# Plot with confidence intervals (slow)
+plotUncertainty(BB_sitepair.2010, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
+plotUncertainty(BB_sitepair.2011, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
+plotUncertainty(BB_sitepair.2012, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
 ```

### 2020-09-10 analyses/old_scripts/floral_ordination.Rmd
rearranged directories

---
title: "floral ordination"
output: html_document
---
### Prepare environment
```{r setup, include=FALSE, echo=FALSE}
library(vegan)
library(tidyverse)
library(lubridate)
```

### Load site data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) 
```

### Load network and survey data
```{r, echo=FALSE, include=FALSE}
bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
  dplyr::select(subgenus, bb.sp, pbl.w, pbl.w.class)

fl_traits <- read.csv("../processed_data/floral_k_type_nom_fix.csv") %>%
  dplyr::select(plant.sp, k.type) %>%
  mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"),
         k.type.ss = str_extract(k.type, "[-+]?[0-9]*"))

fl_corolla_depth <- read.csv("../processed_data/flower_morphology_GitaBenadi.csv") %>%
  group_by(plant.sp = Species) %>%
  summarize(corolla.depth = mean(LN))

fl_tax <- read_csv("../processed_data/floral_tax.csv")

breaks <- read_csv("./output/breaks.csv") %>%
  mutate(year = factor(year))

net <- read_csv("../processed_data/network.csv") %>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  unite(site.date, site, date, sep = "_", remove = FALSE) %>%
  mutate(year = factor(year)) %>% # convert year to factor
  full_join(bb_traits, by = c("bb.sp" = "bb.sp")) %>%
  left_join(site_data, by = c("site")) %>%
  left_join(breaks, by = c("site", "year")) %>%
  left_join(fl_traits) %>%
  left_join(fl_corolla_depth) %>%
  left_join(fl_tax, by = "plant.genus") %>%
  left_join(temperature, by = c("site", "date")) %>%
  ungroup() %>%
  arrange(site, date)

survey <- read_csv("../processed_data/floral_survey.csv") %>%
  mutate(year = factor(year)) %>% # convert year to factor
  semi_join(net, by = "plant.sp") %>% # Consider only plants visited at a given site-date
  left_join(breaks, by = c("site", "year")) %>%
  left_join(site_data, by = c("site")) %>%
  left_join(fl_traits) %>%
  left_join(fl_traits) %>%
  left_join(fl_tax) %>%
  left_join(temperature, by = c("site", "date")) %>%
  ungroup() %>%
  arrange(site, date)

```

### Floral ordination
```{r}
library(ggvegan)

survey_ord <- survey %>%
  filter(flower.cover > 0) %>%
  select(site, date, year, elev.class2, plant.sp, flower.cover) %>%
  group_by(site, date, elev.class2, year, plant.sp) %>%
  summarize(abund = sum(flower.cover)) %>%
  ungroup() %>%
  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
  spread(plant.sp, abund) %>%
  replace(is.na(.), 0) %>%
  unite(col = id, c(site, date, elev.class2, year)) %>%
  column_to_rownames("id") %>%
  decostand(method = "hellinger")

survey_ord <- survey %>%
  mutate(elev.class2 = case_when(
    elev.mean < 1000 ~ "low",
    elev.mean >= 1000 & elev.mean < 1750 ~ "mid",
    elev.mean >= 1750 ~ "high"
  )) %>%
  filter(flower.cover > 0) %>%
  select(site, year, elev.class2, plant.sp, flower.cover) %>%
  group_by(site, year, elev.class2, plant.sp) %>%
  summarize(abund = sum(flower.cover)) %>%
  ungroup() %>%
  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
  spread(plant.sp, abund) %>%
  replace(is.na(.), 0) %>%
  unite(col = id, c(site, elev.class2, year)) %>%
  column_to_rownames("id") %>%
  decostand(method = "hellinger")

pcoa1 <- capscale(survey_ord ~ 1, distance = "jaccard") %>%
  fortify() %>%
  filter(Score == "sites") %>%
  separate(Label, into = c("site", "elev.class2", "year"), sep = "_")

hulls <- pcoa1 %>%
  group_by(year, elev.class2) %>%
  slice(chull(MDS1, MDS2))

ggplot(pcoa1, aes(MDS1, MDS2, fill = elev.class2)) +
  geom_point() +
  geom_polygon(data = hulls, alpha = 0.2, lwd = 0.1) +
  theme_minimal(10) +
  xlab("MDS1") +
  ylab("MDS2") +
  facet_wrap(~year)
```

### 2020-09-11 analyses/FL_abund_GAM.Rmd
Refined and combined FL and BB GAMs

@@
 library(lubridate)
-```
-
-### Climate data
-```{r, include=FALSE, echo=FALSE}
-gdd <- read_csv("../processed_data/climate.csv") %>%
-  mutate(year = factor(year)) %>% # make year a factor
-  dplyr::select(site, year, date, gdd.cum) %>% # drop unused columns
-  semi_join(read_csv("../processed_data/network.csv"), by = "site") # drop sites that were not samples
 ```
@@
 ```{r echo=FALSE, include=FALSE}
-### Per-transect BB abundance: species-level
+### Per-transect BB abundance, filtered to include only samples with corresponding floral survey data
 bb_abund_sp <- net %>%
@@
   dplyr::select(-co.max) %>%
-  inner_join(fl_abund_sp)
+  inner_join(fl_abund_sp) %>%
+  select(-year)
 
@@
   dplyr::select(-preference.update) %>%
-
   mutate(fl.abund.wt = fl.abund*preference) %>%
@@
 
-GAM modeling of preference-weighted abundance per bb species
-
-### Would these models be better specified using gamm4? Would that let me handle the crossed and nested random effects better?
-
-I think I need to include site as a random effect because if I don't the pseudoreplication of elevation is pretty extreme, especially in 2012 where I have lots of repeated measures.
-I am not, for example, measuring floral abundance at 1000 m 20 independent times; I'm measuring it 20 times at the same spot.
-This does, of  course, create identifiability issues insofar as site and elevation are concurved.
-But I think I need to accept these problems.
-The seriousness of the identifiability problems depends in part, I think, on how mgcv handles random effect smooths.
-Does it first account for all the variation it can with the main effects, then model the residuals with the random effect?
-In this case, it's probably not a big problem, since the "partial" influence of site would orthogonal (right?) with that of te(elev.mean, yday).
-
+# GAM modeling of preference-weighted abundance per bb species
+## Global model
 ```{r}
-### A gamm4 version
-weighted_fl_abundance.gamm4 <- gamm4(log(fl.abund.wt + 0.001) ~
-                                       t2(yday, elev.mean, 
-                                          by = bb.sp,
-                                          #id = 0,
-                                          bs= c("gp", "tp"), 
-                                          k=c(8, 8), 
-                                          m=1),
-                                     random = ~(1 | site:year),
-                                     data = fl_abund_wt,
-                                     discrete = TRUE,
-                                     family = "gaussian",
-                                     select = TRUE,
-                                     method = "fREML") %>% getViz(nsim = 500)
-
-### Global model
 #### My concern with using a global model is that, because I have so much higher sampling density in 2012, the global model will basically be just a 2012 model with 2010 and 2011 in the residuals, so to speak.
@@
                                       s(year, bs = "re") +
-                                      s(site, bs = "re") +
+                                      #s(site, bs = "re") +
                                         te(yday, elev.mean, 
@@
 weighted_fl_abundance.global2.ck + l_gridCheck2D(gridFun = mean)
+```
 
+## Annual models
+```{r}
 ### 2010
@@
                                                 bs= c("gp", "tp"), 
-                                                k=c(10, 10), 
+                                                k=c(8, 8), 
                                                 m=2),
@@
 
-weighted_fl_abundance.2010b <- bam(fl.abund.wt + 0.001 ~
+check.gamViz(weighted_fl_abundance.2010)
+weighted_fl_abundance.2010.sum <- summary(weighted_fl_abundance.2010)
+print(plot(weighted_fl_abundance.2010, allTerms = TRUE), pages = 4)
+weighted_fl_abundance.2010.ck <- check2D(weighted_fl_abundance.2010, 
+                                         x1 = "yday", x2 = "elev.mean")
+
+### 2011
+weighted_fl_abundance.2011 <- bam(log(fl.abund.wt + 0.001) ~
                                       s(bb.sp, bs = "re") +
@@
                                                 bs= c("gp", "tp"), 
-                                                k=c(15, 15), 
-                                                m=2),
-                                    data = filter(fl_abund_wt, year == 2010),
-                                    discrete = TRUE,
-                                    family = "tw",
-                                    select = TRUE,
-                                    method = "fREML") %>% getViz(nsim = 500)
-
-check.gamViz(weighted_fl_abundance.2010)
-check.gamViz(weighted_fl_abundance.2010b)
-
-weighted_fl_abundance.2010.sum <- summary(weighted_fl_abundance.2010)
-print(plot(weighted_fl_abundance.2010, allTerms = TRUE), pages = 4)
-print(plot(weighted_fl_abundance.2010b, allTerms = TRUE), pages = 4)
-
-plot(sm(weighted_fl_abundance.2010, 2))
-plot(sm(weighted_fl_abundance.2010b, 2))
-
-weighted_fl_abundance.2010.ck <- check2D(weighted_fl_abundance.2010, 
-                                         x1 = "yday", x2 = "elev.mean")
-weighted_fl_abundance.2010b.ck <- check2D(weighted_fl_abundance.2010b, 
-                                         x1 = "yday", x2 = "elev.mean")
-
-weighted_fl_abundance.2010.ck + l_gridCheck2D(gridFun = mean)
-
-### 2011
-weighted_fl_abundance.2011 <- bam(log(fl.abund.wt + 0.001) ~
-                                      s(bb.sp, bs = "re") +
-                                      #s(management, bs = "re") +
-                                      s(site, bs = "re") +
-                                      te(yday, elev.mean, 
-                                                by = bb.sp,
-                                                #id = 0,
-                                                bs= c("gp", "tp"), 
-                                                k=c(10, 10), 
+                                                k=c(8, 8), 
                                                 m=2),
@@
                                          x1 = "yday", x2 = "elev.mean")
-weighted_fl_abundance.2011.ck + l_gridCheck2D(gridFun = mean)
 
@@
                                          x1 = "yday", x2 = "elev.mean")
-weighted_fl_abundance.2012.ck + l_gridCheck2D(gridFun = mean)
 ```
 
-### Tabulate plots
+# Extract predictions
 ```{r}
-bssFL_2010 <- plot(sm(weighted_fl_abundance.2010, 4)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. terrestris/lucorum")), x = NULL, y = NULL)
-bssFL_2011 <- plot(sm(weighted_fl_abundance.2011, 4)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. terrestris/lucorum")), x = NULL, y = NULL)
-bssFL_2012 <- plot(sm(weighted_fl_abundance.2012, 4)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. terrestris/lucorum")), x = NULL, y = NULL)
+# Function to build data set over which to generate predictions
+newdata_build <- function(dat, y) {
 
-gersFL_2010 <- plot(sm(weighted_fl_abundance.2010, 5)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. gerstaeckeri")), x = NULL, y = NULL)
-gersFL_2011 <- plot(sm(weighted_fl_abundance.2011, 5)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. gerstaeckeri")), x = NULL, y = NULL) 
-gersFL_2012 <- plot(sm(weighted_fl_abundance.2012, 5)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. gerstaeckeri")), x = NULL, y = NULL)
+  ## yday min/max
+  min_yday <- min(filter(dat, year == y)$yday)
+  max_yday <- max(filter(dat, year == y)$yday)
 
-hortFL_2010 <- plot(sm(weighted_fl_abundance.2010, 6)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. hortorum")), x = NULL, y = NULL)
-hortFL_2011 <- plot(sm(weighted_fl_abundance.2011, 6)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. hortorum")), x = NULL, y = NULL)
-hortFL_2012 <- plot(sm(weighted_fl_abundance.2012, 6)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. hortorum")), x = NULL, y = NULL)
+  ## elev min.max
+  min_elev <- min(filter(dat, year == y)$elev.mean)
+  max_elev <- max(filter(dat, year == y)$elev.mean)
 
-humiFL_2010 <- plot(sm(weighted_fl_abundance.2010, 7)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. humilis")), x = NULL, y = NULL)
-humiFL_2011 <- plot(sm(weighted_fl_abundance.2011, 7)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. humilis")), x = NULL, y = NULL)
-humiFL_2012 <- plot(sm(weighted_fl_abundance.2012, 7)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. humilis")), x = NULL, y = NULL)
+  ## species list
+  spp <- dat %>%
+    filter(year == y) %>%
+    select(bb.sp) %>%
+    unique() %>%
+    na.omit() ##### NEED TO FIGURE OUT WHERE THE NA COMES FROM!
 
-hypnFL_2010 <- plot(sm(weighted_fl_abundance.2010, 8)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. hypnorum")), x = NULL, y = NULL)
-hypnFL_2011 <- plot(sm(weighted_fl_abundance.2011, 8)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. hypnorum")), x = NULL, y = NULL)
-hypnFL_2012 <- plot(sm(weighted_fl_abundance.2012, 8)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. hypnorum")), x = NULL, y = NULL)
+  # sequence of 50 points along the yday and elev ranges, respectively
+  seq_yday <- seq(min_yday, max_yday, length.out = 50)
+  seq_elev <- seq(min_elev, max_elev, length.out = 50)
 
-joneFL_2010 <- plot(sm(weighted_fl_abundance.2010, 9)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. jonellus")), x = NULL, y = NULL)
-joneFL_2011 <- plot(sm(weighted_fl_abundance.2011, 9)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. jonellus")), x = NULL, y = NULL)
-joneFL_2012 <- plot(sm(weighted_fl_abundance.2012, 9)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. jonellus")), x = NULL, y = NULL)
+  newdata <- tibble(seq_yday, seq_elev) %>% # combine seqs into data frame
+    complete(seq_yday, seq_elev) %>% # get all possible values
+    crossing(spp) %>% # get all possible combos with species
+    select(yday = seq_yday, elev.mean = seq_elev, bb.sp) # rename columns to match the original model inputs
+}
 
-lapiFL_2010 <- plot(sm(weighted_fl_abundance.2010, 10)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. lapidarius")), x = NULL, y = NULL)
-lapiFL_2011 <- plot(sm(weighted_fl_abundance.2011, 10)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. lapidarius")), x = NULL, y = NULL)
-lapiFL_2012 <- plot(sm(weighted_fl_abundance.2012, 10)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. lapidarius")), x = NULL, y = NULL)
+newdata2010 <- newdata_build(fl_abund_wt, 2010)
+newdata2011 <- newdata_build(fl_abund_wt, 2011)
+newdata2012 <- newdata_build(fl_abund_wt, 2012)
 
-mendFL_2010 <- plot(sm(weighted_fl_abundance.2010, 11)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. mendax")), x = NULL, y = NULL)
-mendFL_2011 <- plot(sm(weighted_fl_abundance.2011, 11)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. mendax")), x = NULL, y = NULL)
-mendFL_2012 <- plot(sm(weighted_fl_abundance.2012, 11)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. mendax")), x = NULL, y = NULL)
+# A little wrapper to make predictions and return a tibble with desired columns 
+predict_func <- function(mod, new, y) {
+  predict.bam(mod, newdata = new, type = "response") %>%
+    data.frame() %>%
+    tibble() %>%
+    bind_cols(new) %>%
+    select(yday, elev.mean, bb.sp, pred = 1) %>%
+    mutate(pred.backtrans = exp(pred),
+           year = rep(y, n()))
+}
 
-montFL_2010 <- plot(sm(weighted_fl_abundance.2010, 12)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. monticola")), x = NULL, y = NULL)
-montFL_2011 <- plot(sm(weighted_fl_abundance.2011, 12)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. monticola")), x = NULL, y = NULL)
-montFL_2012 <- plot(sm(weighted_fl_abundance.2012, 12)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. monticola")), x = NULL, y = NULL)
+# Generate predictions
+predict2010 <- predict_func(weighted_fl_abundance.2010, newdata2010, 2010)
+predict2011 <- predict_func(weighted_fl_abundance.2011, newdata2011, 2011)
+predict2012 <- predict_func(weighted_fl_abundance.2012, newdata2012, 2012)
+predict_pool <- bind_rows(predict2010, predict2011, predict2012)
 
-muciFL_2010 <- plot(sm(weighted_fl_abundance.2010, 13)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. mucidis")), x = NULL, y = NULL)
-muciFL_2011 <- plot(sm(weighted_fl_abundance.2011, 13)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. mucidis")), x = NULL, y = NULL)
-muciFL_2012 <- plot(sm(weighted_fl_abundance.2012, 13)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. mucidis")), x = NULL, y = NULL)
+# Subset predictions
+predict_major <- filter(predict_pool, bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf"))
+predict_highelev <- filter(predict_pool, bb.sp %in% c("pyre", "mend", "mont"))
+predict_minor <- filter(predict_pool, bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci"))
 
-pascFL_2010 <- plot(sm(weighted_fl_abundance.2010, 14)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. pascuorum")), x = NULL, y = NULL)
-pascFL_2011 <- plot(sm(weighted_fl_abundance.2011, 14)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. pascuorum")), x = NULL, y = NULL)
-pascFL_2012 <- plot(sm(weighted_fl_abundance.2012, 14)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. pascuorum")), x = NULL, y = NULL)
+ggplot(predict_major, aes(yday, elev.mean, fill = pred)) +
+  geom_raster() +
+  geom_contour(aes(z = pred), color = "black", size = 0.25, binwidth = 1) +
+  scale_fill_viridis_c() +
+  facet_grid(year~bb.sp)
 
-pratFL_2010 <- plot(sm(weighted_fl_abundance.2010, 15)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. pratorum")), x = NULL, y = NULL)
-pratFL_2011 <- plot(sm(weighted_fl_abundance.2011, 15)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. pratorum")), x = NULL, y = NULL)
-pratFL_2012 <- plot(sm(weighted_fl_abundance.2012, 15)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. pratorum")), x = NULL, y = NULL)
+ggplot(predict_highelev, aes(yday, elev.mean, fill = pred)) +
+  geom_raster() +
+  geom_contour(aes(z = pred), color = "black", size = 0.25, binwidth = 1) +
+  scale_fill_viridis_c() +
+  facet_grid(year~bb.sp)
 
-psitFL_2010 <- plot(sm(weighted_fl_abundance.2010, 16)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. psithyrus")), x = NULL, y = NULL)
-psitFL_2011 <- plot(sm(weighted_fl_abundance.2011, 16)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. psithyrus")), x = NULL, y = NULL)
-psitFL_2012 <- plot(sm(weighted_fl_abundance.2012, 16)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. psithyrus")), x = NULL, y = NULL)
-
-pyreFL_2010 <- plot(sm(weighted_fl_abundance.2010, 17)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. pyrenaeus")), x = NULL, y = NULL)
-pyreFL_2011 <- plot(sm(weighted_fl_abundance.2011, 17)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. pyrenaeus")), x = NULL, y = NULL)
-pyreFL_2012 <- plot(sm(weighted_fl_abundance.2012, 17)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. pyrenaeus")), x = NULL, y = NULL)
-
-soroFL_2010 <- plot(sm(weighted_fl_abundance.2010, 18)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. soroensis")), x = NULL, y = NULL)
-soroFL_2011 <- plot(sm(weighted_fl_abundance.2011, 18)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. soroensis")), x = NULL, y = NULL)
-soroFL_2012 <- plot(sm(weighted_fl_abundance.2012, 18)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. soroensis")), x = NULL, y = NULL)
-
-wurfFL_2010 <- plot(sm(weighted_fl_abundance.2010, 19)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. wurflenii")), x = NULL, y = NULL)
-wurfFL_2011 <- plot(sm(weighted_fl_abundance.2011, 19)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. wurflenii")), x = NULL, y = NULL)
-wurfFL_2012 <- plot(sm(weighted_fl_abundance.2012, 19)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  #l_points() +
-  labs(title = expression(italic("B. wurflenii")), x = NULL, y = NULL)
-
-major_sp <- gridPrint(bssFL_2010, bssFL_2011, bssFL_2012, 
-                      pascFL_2010, pascFL_2011, pascFL_2012,
-                      pratFL_2010, pratFL_2011, pratFL_2012,
-                      soroFL_2010, soroFL_2011, soroFL_2012,
-                      wurfFL_2010, wurfFL_2011, wurfFL_2012, ncol = 3)
-
-high_elev_sp <- gridPrint(pyreFL_2010, pyreFL_2011, pyreFL_2012, 
-                          mendFL_2010, mendFL_2011, mendFL_2012,
-                          montFL_2010, montFL_2011, montFL_2012, ncol = 3)
-
-minor_sp <- gridPrint(gersFL_2010, gersFL_2011, gersFL_2012,
-                      hortFL_2010, hortFL_2011, hortFL_2012,
-                      #humiFL_2010, humiFL_2011, humiFL_2012, # too rarely seen, I think only 3 times
-                      hypnFL_2010, hypnFL_2011, hypnFL_2012,
-                      joneFL_2010, joneFL_2011, joneFL_2012,
-                      lapiFL_2010, lapiFL_2011, lapiFL_2012,
-                      #psitFL_2010, psitFL_2011, psitFL_2012, # Let's ignore 
-                      muciFL_2010, muciFL_2011, muciFL_2012, ncol = 3)
-plot(major_sp)
-plot(high_elev_sp)
-plot(minor_sp)
-
+ggplot(predict_minor, aes(yday, elev.mean, fill = pred)) +
+  geom_raster() +
+  geom_contour(aes(z = pred), color = "black", size = 0.25, binwidth = 1) +
+  scale_fill_viridis_c() +
+  facet_grid(year~bb.sp)
 ```

### 2020-09-11 analyses/abund_div_GAM.Rmd
Refined and combined FL and BB GAMs

---
title: "Abundance/diversity GAM"
output: html_document
---

### Prepare environment
```{r setup, include=FALSE, echo=FALSE}
library(mgcv)
library(mgcViz)
library(tidyverse)
library(lubridate)
```

### Load site data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../processed_data/site_data.csv") %>%
  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")), # rename in English
         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))
```

### Load network and survey data
```{r}
### BB-FL visitation data
net <- read_csv("../processed_data/network.csv") %>%
  dplyr::select(site, date, bb.sp, plant.sp) # drop unused columns

### Floral survey data
survey <- read_csv("../processed_data/floral_survey.csv") %>%
  semi_join(net, by = "plant.sp") %>% # Consider only visited plant species (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover)

### Identify duplicate observations; these may be errors or intentionally duplicated (see correspondence with Kalli)
survey_dups <- survey %>%
  group_by(site, date, plant.sp) %>%
  mutate(dups = n()) %>%
  filter(dups > 1)

### Shared transects
shared_transects <- inner_join(net, survey, by = c("site", "date")) %>%
  dplyr::select(site, date) %>%
  unique() 
```

### FL and BB abundance / diversity
```{r echo=FALSE, include=TRUE}
# Per-transect BB abundance
## full dataset
bb_abund_sp <- net %>%
  dplyr::select(site, date, bb.sp) %>%
  group_by(site, date, bb.sp) %>% # 
  summarize(bb.abund = n()) %>% # get per transect bb abundance
  ungroup() %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  semi_join(net, by = c("site", "date")) %>% # drop site-dates that did not exist in the data set
  left_join(site_data) %>% # add site data
  mutate(year = factor(year(date)),
         yday = yday(date),
         bb.sp = factor(bb.sp),
         management = factor(management),
         site = factor(site),
         transect = factor(transect)) %>%
  group_by(year, bb.sp) %>%
  mutate(bb.total.abund = sum(bb.abund))

## filtered to include only samples with corresponding floral survey data
bb_abund_sp.sub <- bb_abund_sp %>%
  semi_join(shared_transects)

### Per-transect floral abund: by species
fl_abund_sp <- survey %>%
  group_by(site, date, plant.sp) %>% # group by plant.sp*transect
  summarize(fl.abund = sum(flower.cover)) %>% # get plant.sp*transect total flower cover; this step really shouldn't be required, but in a few cases, there is more than one abundance entry for the same species on a given transect; not sure how that happened, but it's in the original data that Fabrice sent me; alternatively, I could drop the dups or average them. We'll see what Kalli and Fabrice have to say.
  group_by(site, date) %>%
  mutate(prop.fl.abund = fl.abund/sum(fl.abund)) %>% # get the proportional flower cover for each species.
  ungroup() %>%
  full_join(dplyr::select(net, site, date, plant.sp), by = c("site", "date", "plant.sp")) %>% # add the handful of species occurring in the visitation network but missing from the survey data
  complete(site, date, plant.sp, fill = list(fl.abund = 0, prop.fl.abund = 0)) %>%
  semi_join(survey, by = c("site", "date")) %>%  # only include site-dates that actually occurred
  unique() %>%
  mutate(year = factor(year(date)))
```

### Preference analysis
```{r}
observed_visits <- net %>%
  semi_join(shared_transects) %>% # include only site*date combos where floral surveying was done
  group_by(site, date, bb.sp, plant.sp) %>% # group by bb*plant*transect
  summarize(visits = n()) %>% # total visits per bb-fl pair
  group_by(site, date, bb.sp) %>%
  mutate(prop.visits = visits/sum(visits))

# All possible bumble-plant pairings *site * date, filtered to include only those when the given bb.sp was present and the given plant.sp was *either* present *or* visited (see note below)
visitation_by_abundance <- fl_abund_sp %>%
  inner_join(bb_abund_sp.sub) %>%
  filter(bb.abund > 0) %>%
  filter(fl.abund > 0) %>% 
  left_join(observed_visits) %>%
  #filter(fl.abund > 0 | visits > 0) %>% # In about 10% of the bb.sp*plant.sp*site*year combos, the visited plant was not recorded in the floral survey data. There are a few different ways we could handle this, but I think an argument could be make that treating those plants as if they had zero abundance (even though they were evidently present gets close to the truth, since they must have been rather rare to have been missed in the surveying.)
  replace_na(list(visits = 0, prop.visits = 0))

### Get mean abundance/visitation for each plant*bb combo
visitation_by_abundance_mean <- visitation_by_abundance %>%
  mutate(year = factor(year(date))) %>%
  group_by(plant.sp, bb.sp) %>%
  summarize(prop.fl.abund.m = mean(prop.fl.abund),
            prop.visits.wm = weighted.mean(prop.visits, bb.abund))

# Visualize
ggplot(filter(visitation_by_abundance_mean, prop.visits.wm > 0), 
       aes(log(prop.fl.abund.m + 0.01), log(prop.visits.wm))) +
  geom_point(alpha = 0.5, aes(color = bb.sp)) +
  geom_smooth(method = "lm")

ggplot(filter(visitation_by_abundance_mean), 
       aes(prop.fl.abund.m, prop.visits.wm)) +
  geom_point(alpha = 0.5, aes(color = bb.sp)) +
  geom_smooth(method = "lm")

### Okay, here's what we're going to do. Preference has a strong bimodal pattern, with lots of plant species totally ignored while the ones that are foraged upon are visited in proportion to their abundance. We can't model that latter pattern well if we retain all the totally ignored plant species. So, let's model the visited plants separately from the unvisited ones.

### Call linear model on the subset of the floral species that were visited at least once
# log on log regression is common for proportional datam (log ratio regression)
pref.lm <- lm(log(prop.visits.wm) ~ log(prop.fl.abund.m + 0.01), 
              data = filter(visitation_by_abundance_mean, prop.visits.wm > 0))
summary(pref.lm)
plot(pref.lm)

### Extract residuals
pref.lm.resid <- residuals(pref.lm) %>%
  data.frame() %>%
  as_tibble() %>%
  rename(preference = 1)

#################################
### Build preference data set ###
#################################

# Start with the mean prop.fl.abund and weighted mean prop.visits
preference <- visitation_by_abundance_mean %>%  
  ungroup() %>%

# Remove bb*fl*year combos that had 0 visits
  filter(prop.visits.wm > 0) %>%

# With the zeroes removed, what's left can be joined 1:1 with the 
# residuals of the model above, i.e. the raw preference data
  bind_cols(pref.lm.resid) %>%

# Exponentiate the residuals to back-transform the data and ensure 
# that preference is > 0 for all visited plant species.
  mutate(preference = exp(preference)) %>%

# Now we need to add back all the bb*fl*year combos that had 0 visits
# Then we have NAs for preference and floral abundance columns
  full_join(unique(dplyr::select(visitation_by_abundance_mean, bb.sp, plant.sp))) %>%

# Replace the NAs of the preference column with 0s; 
  replace_na(list(preference = 0)) %>%

# Drop the columns used to calculate preference so that they don't confuse us when we add site-date level abundance data
  dplyr::select(bb.sp, plant.sp, preference)

write_csv(preference, "../analyses/output/preference.csv")

#################################
### Join with FL and BB abund ###
#################################

# For bb-fl combos that never co-occurred, set preference = 1
pref_forbidden_link_fill <- inner_join(fl_abund_sp, bb_abund_sp.sub) %>%
  mutate(co = case_when(
    fl.abund == 0 | bb.abund == 0 ~ 0,
    fl.abund != 0 & bb.abund != 0 ~ 1)
  ) %>%
  group_by(plant.sp, bb.sp) %>%
  summarize(co.max = max(co)) %>%
  filter(co.max == 0) %>%
  mutate(preference.update = rep(1, n())) %>%
  dplyr::select(-co.max) %>%
  inner_join(fl_abund_sp) %>%
  select(-year)

fl_abund_wt <- fl_abund_sp %>%
  select(-year) %>%
  left_join(preference) %>%

# Add the forbidden links with preference set to 1 (neutral)
  full_join(pref_forbidden_link_fill) %>%
  mutate(preference = case_when(
    is.na(preference) ~ preference.update,
    !is.na(preference) ~ preference)
    ) %>%
  dplyr::select(-preference.update) %>%
  mutate(fl.abund.wt = fl.abund*preference) %>%
  mutate(yday = yday(date)) %>%
  group_by(site, date, yday, bb.sp) %>%
  summarize(fl.abund.wt = sum(fl.abund.wt, na.rm = TRUE)) %>%
  ungroup() %>%
  semi_join(survey, by = c("site", "date")) %>% # This gets rid of site*date combos that don't exist
  left_join(site_data, by = "site") %>%
  mutate(year = factor(year(date)), # add in the years missing from the bb_abund join
         bb.sp = factor(bb.sp), # convert bb.sp to a factor
         management = factor(management),
         site = factor(site),
         transect = factor(transect)) %>%
  group_by(year) %>%
  mutate(total.transects.year = n()) %>%
  ungroup()

write_csv(fl_abund_wt, "../analyses/output/fl_abund_wt.csv")
```

### Call GAMs
##### My argument for not including site as a random effect is that it would create identifiability issues with the elevation variable that I am actally interested in. This will probably cause the p-values to be underestimated, maybe to the point of being uninformative. But I don't think I care. My goal is not to establish inferentially that time and elevation affect BB abundance; I want to say how they do. And I don't want to obscure the "how" for the sake of inferential scruples.
#### Floral resources
```{r}
### 2010
weighted_fl_abundance.2010 <- bam(log(fl.abund.wt + 0.001) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      #s(site, bs = "re") +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(8, 8), 
                                                m=2),
                                    data = filter(fl_abund_wt, year == 2010),
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

check.gamViz(weighted_fl_abundance.2010)
#weighted_fl_abundance.2010.sum <- summary(weighted_fl_abundance.2010)
#print(plot(weighted_fl_abundance.2010, allTerms = TRUE), pages = 4)
weighted_fl_abundance.2010.ck <- check2D(weighted_fl_abundance.2010, 
                                         x1 = "yday", x2 = "elev.mean")

### 2011
weighted_fl_abundance.2011 <- bam(log(fl.abund.wt + 0.001) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      #s(site, bs = "re") +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(8, 8), 
                                                m=2),
                                    data = filter(fl_abund_wt, year == 2011),
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

check.gamViz(weighted_fl_abundance.2011)
#weighted_fl_abundance.2011.sum <- summary(weighted_fl_abundance.2011)
#print(plot(weighted_fl_abundance.2011, allTerms = TRUE), pages = 4)
weighted_fl_abundance.2011.ck <- check2D(weighted_fl_abundance.2011, 
                                         x1 = "yday", x2 = "elev.mean")

### 2012
weighted_fl_abundance.2012 <- bam(log(fl.abund.wt + 0.001) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      #s(site, bs = "re") +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(8, 8), 
                                                m=2),
                                    data = filter(fl_abund_wt, year == 2012),
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

check.gamViz(weighted_fl_abundance.2012)
#weighted_fl_abundance.2012.sum <- summary(weighted_fl_abundance.2012)
#print(plot(weighted_fl_abundance.2012, allTerms = TRUE), pages = 4)
weighted_fl_abundance.2012.ck <- check2D(weighted_fl_abundance.2012, 
                                         x1 = "yday", x2 = "elev.mean")
```

#### Bumbles
```{r}
### 2010
bb_abund.2010 <- bam(bb.abund ~ 
                       s(bb.sp, bs = "re") + 
                       te(yday, elev.mean, 
                          by = bb.sp,
                          bs= c("gp", "cr"), 
                          k=c(8, 8), 
                          m=2),
                     data = filter(bb_abund_sp, year == 2010 & bb.total.abund >= 5),
                     discrete = TRUE,
                     family = "tw",
                     select = TRUE,
                     method = "fREML") %>% getViz(nsim = 500)

check.gamViz(bb_abund.2010)
#bb_abund.2010.sum <- summary(bb_abund.2010)
#print(plot(bb_abund.2010, allTerms = TRUE), pages = 1)
bb_abund.2010.ck <- check2D(bb_abund.2010, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals

### 2011
bb_abund.2011 <- bam(bb.abund ~ 
                       s(bb.sp, bs = "re") + 
                       te(yday, elev.mean, 
                          by = bb.sp,
                          bs= c("gp", "cr"), 
                          k=c(8, 8), 
                          m=2),
                     data = filter(bb_abund_sp, year == 2011 & bb.total.abund >= 5),
                     discrete = TRUE,
                     family = "tw",
                     select = TRUE,
                     method = "fREML") %>% getViz(nsim = 500)

check.gamViz(bb_abund.2011)
#bb_abund.2011.sum <- summary(bb_abund.2011)
#print(plot(bb_abund.2011, allTerms = TRUE), pages = 1)
bb_abund.2011.ck <- check2D(bb_abund.2011, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals

### 2012
bb_abund.2012 <- bam(bb.abund ~ 
                       s(bb.sp, bs = "re") + 
                       te(yday, elev.mean, 
                          by = bb.sp,
                          bs= c("gp", "cr"), 
                          k=c(8, 8), 
                          m=2),
                     data = filter(bb_abund_sp, year == 2012 & bb.total.abund >= 5),
                     discrete = TRUE,
                     family = "tw",
                     select = TRUE,
                     method = "fREML") %>% getViz(nsim = 500)

check.gamViz(bb_abund.2012)
#bb_abund.2012.sum <- summary(bb_abund.2012)
#print(plot(bb_abund.2012, allTerms = TRUE), pages = 1)
bb_abund.2012.ck <- check2D(bb_abund.2012, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals
```

### Extract predictions
```{r}
# Function to build data set over which to generate predictions
newdata_build <- function(dat, y) {

  ## yday min/max
  min_yday <- min(filter(dat, year == y)$yday)
  max_yday <- max(filter(dat, year == y)$yday)

  ## elev min.max
  min_elev <- min(filter(dat, year == y)$elev.mean)
  max_elev <- max(filter(dat, year == y)$elev.mean)

  ## species list
  spp <- dat %>%
    filter(year == y) %>%
    select(bb.sp) %>%
    unique() %>%
    na.omit() ##### NEED TO FIGURE OUT WHERE THE NA COMES FROM!

  # sequence of 50 points along the yday and elev ranges, respectively
  seq_yday <- seq(min_yday, max_yday, length.out = 50)
  seq_elev <- seq(min_elev, max_elev, length.out = 50)

  newdata <- tibble(seq_yday, seq_elev) %>% # combine seqs into data frame
    complete(seq_yday, seq_elev) %>% # get all possible values
    crossing(spp) %>% # get all possible combos with species
    select(yday = seq_yday, elev.mean = seq_elev, bb.sp) # rename columns to match the original model inputs
}

newdata2010.FL <- newdata_build(fl_abund_wt, 2010)
newdata2011.FL <- newdata_build(fl_abund_wt, 2011)
newdata2012.FL <- newdata_build(fl_abund_wt, 2012)

newdata2010.BB <- newdata_build(filter(bb_abund_sp, bb.total.abund >= 5), 2010)
newdata2011.BB <- newdata_build(filter(bb_abund_sp, bb.total.abund >= 5), 2011)
newdata2012.BB <- newdata_build(filter(bb_abund_sp, bb.total.abund >= 5), 2012)

# A little wrapper to make predictions and return a tibble with desired columns 
predict_func <- function(mod, new, y) {
  predict.bam(mod, newdata = new, type = "response") %>%
    data.frame() %>%
    tibble() %>%
    bind_cols(new) %>%
    select(yday, elev.mean, bb.sp, pred = 1) %>%
    mutate(year = rep(y, n()))
}

# Generate predictions
predict2010.FL <- predict_func(weighted_fl_abundance.2010, newdata2010.FL, 2010)
predict2011.FL <- predict_func(weighted_fl_abundance.2011, newdata2011.FL, 2011)
predict2012.FL <- predict_func(weighted_fl_abundance.2012, newdata2012.FL, 2012)
predict_pool.FL <- bind_rows(predict2010.FL, predict2011.FL, predict2012.FL)

predict2010.BB <- predict_func(bb_abund.2010, newdata2010.BB, 2010)
predict2011.BB <- predict_func(bb_abund.2011, newdata2011.BB, 2011)
predict2012.BB <- predict_func(bb_abund.2012, newdata2012.BB, 2012)
predict_pool.BB <- bind_rows(predict2010.BB, predict2011.BB, predict2012.BB)

# Subset predictions
predict_major.FL <- filter(predict_pool.FL, bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf"))
predict_highelev.FL <- filter(predict_pool.FL, bb.sp %in% c("pyre", "mend", "mont"))
predict_minor.FL <- filter(predict_pool.FL, bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci"))

predict_major.BB <- filter(predict_pool.BB, bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf"))
predict_highelev.BB <- filter(predict_pool.BB, bb.sp %in% c("pyre", "mend", "mont"))
predict_minor.BB <- filter(predict_pool.BB, bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci"))
```

### Visualize
```{r}
ggplot(predict_major.FL, aes(yday, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25, binwidth = 1) +
  scale_fill_viridis_c() +
  facet_grid(year~bb.sp)

ggplot(predict_highelev.FL, aes(yday, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25, binwidth = 1) +
  scale_fill_viridis_c() +
  facet_grid(year~bb.sp)

ggplot(predict_minor.FL, aes(yday, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25, binwidth = 1) +
  scale_fill_viridis_c() +
  facet_grid(year~bb.sp)

# Plot
ggplot(predict_major.BB, aes(yday, elev.mean, fill = log(pred))) +
  geom_raster() +
  geom_contour(aes(z = log(pred)), color = "black", size = 0.25, binwidth = 1) +
  scale_fill_viridis_c() +
  facet_grid(year~bb.sp)

ggplot(predict_highelev.BB, aes(yday, elev.mean, fill = log(pred))) +
  geom_raster() +
  geom_contour(aes(z = log(pred)), color = "black", size = 0.25, binwidth = 1) +
  scale_fill_viridis_c() +
  facet_grid(year~bb.sp)

ggplot(predict_minor.BB, aes(yday, elev.mean, fill = log(pred))) +
  geom_raster() +
  geom_contour(aes(z = log(pred)), color = "black", size = 0.25, binwidth = 1) +
  scale_fill_viridis_c() +
  facet_grid(year~bb.sp)
```

### 2020-09-14 analyses/abund_div_GAM.Rmd
polished GAM script

@@
 ### Load network and survey data
-```{r}
+```{r message=FALSE}
 ### BB-FL visitation data
@@
   semi_join(net, by = "plant.sp") %>% # Consider only visited plant species (anytime anywhere, not specifically for a given site*date)
-  dplyr::select(site, date, plant.sp, flower.cover)
+  dplyr::select(site, date, plant.sp, flower.cover) %>%
+  group_by(plant.sp) %>%
+  mutate(present = case_when(
+    max(flower.cover) > 0 ~ TRUE,
+    max(flower.cover) == 0 ~ FALSE
+  )) %>%
+  filter(present == TRUE) # Ignore plants that never had any flower cover
 
@@
 ### FL and BB abundance / diversity
-```{r echo=FALSE, include=TRUE}
+```{r echo=FALSE, include=TRUE, message=FALSE}
 # Per-transect BB abundance
@@
   group_by(year, bb.sp) %>%
-  mutate(bb.total.abund = sum(bb.abund))
+  mutate(bb.annual.abund = sum(bb.abund))
 
@@
 bb_abund_sp.sub <- bb_abund_sp %>%
-  semi_join(shared_transects)
+  ungroup() %>%
+  semi_join(shared_transects) %>%
+  select(site, date, yday, bb.sp, bb.abund, bb.annual.abund, elev.mean)
 
@@
   ungroup() %>%
-  full_join(dplyr::select(net, site, date, plant.sp), by = c("site", "date", "plant.sp")) %>% # add the handful of species occurring in the visitation network but missing from the survey data
+  #full_join(dplyr::select(net, site, date, plant.sp), by = c("site", "date", "plant.sp")) %>% # add the handful of species occurring in the visitation network but missing from the survey data
   complete(site, date, plant.sp, fill = list(fl.abund = 0, prop.fl.abund = 0)) %>%
@@
 ### Preference analysis
-```{r}
+```{r message=FALSE}
 observed_visits <- net %>%
@@
 
-# All possible bumble-plant pairings *site * date, filtered to include only those when the given bb.sp was present and the given plant.sp was *either* present *or* visited (see note below)
+# All possible bumble-plant pairings *site * date, filtered to include only those when the given bb.sp was present and the given plant.sp.
+Note that in about 10% of the bb.sp*plant.sp*site*year combos, the visited plant was not recorded in the floral survey data.
+There are a few different ways we could handle this, but for now I am simply ignoring plant spp.
+for which we have no abundance data.
 visitation_by_abundance <- fl_abund_sp %>%
   inner_join(bb_abund_sp.sub) %>%
-  filter(bb.abund > 0) %>%
-  filter(fl.abund > 0) %>% 
+  filter(bb.abund > 0 & fl.abund > 0) %>%
   left_join(observed_visits) %>%
-  #filter(fl.abund > 0 | visits > 0) %>% # In about 10% of the bb.sp*plant.sp*site*year combos, the visited plant was not recorded in the floral survey data.
-There are a few different ways we could handle this, but I think an argument could be make that treating those plants as if they had zero abundance (even though they were evidently present gets close to the truth, since they must have been rather rare to have been missed in the surveying.)
   replace_na(list(visits = 0, prop.visits = 0))
@@
 visitation_by_abundance_mean <- visitation_by_abundance %>%
-  mutate(year = factor(year(date))) %>%
   group_by(plant.sp, bb.sp) %>%
@@
 
-# Remove bb*fl*year combos that had 0 visits
+# Remove bb*fl combos that had 0 visits
   filter(prop.visits.wm > 0) %>%
@@
 # Drop the columns used to calculate preference so that they don't confuse us when we add site-date level abundance data
-  dplyr::select(bb.sp, plant.sp, preference)
+  dplyr::select(bb.sp, plant.sp, preference) %>%
+  mutate(sqrt.preference = sqrt(preference),
+         bin.preference = case_when(
+           preference > 0 ~ 1,
+           preference == 0 ~ 0
+         ))
 
 write_csv(preference, "../analyses/output/preference.csv")
+
+ggplot(preference, aes(preference)) +
+  geom_histogram() +
+  facet_wrap(~bb.sp)
+
+ggplot(preference, aes(sqrt.preference)) +
+  geom_histogram() +
+  facet_wrap(~bb.sp)
 
@@
 
-# For bb-fl combos that never co-occurred, set preference = 1
-pref_forbidden_link_fill <- inner_join(fl_abund_sp, bb_abund_sp.sub) %>%
-  mutate(co = case_when(
-    fl.abund == 0 | bb.abund == 0 ~ 0,
-    fl.abund != 0 & bb.abund != 0 ~ 1)
-  ) %>%
-  group_by(plant.sp, bb.sp) %>%
-  summarize(co.max = max(co)) %>%
-  filter(co.max == 0) %>%
-  mutate(preference.update = rep(1, n())) %>%
-  dplyr::select(-co.max) %>%
-  inner_join(fl_abund_sp) %>%
-  select(-year)
+#After reconsidering this approach, I think it is better to assume preference = 0  for non-co-occuring spp.
+
+# # For bb-fl combos that never co-occurred, set preference = 0
+# pref_forbidden_link_fill <- inner_join(fl_abund_sp, bb_abund_sp.sub) %>%
+#   mutate(co = case_when(
+#     fl.abund == 0 | bb.abund == 0 ~ 0, # bb.sp and fl.sp that did not co-occur on given site date
+#     fl.abund != 0 & bb.abund != 0 ~ 1) # bb.sp and fl.sp that did co-occur on a given site date
+#   ) %>%
+#   group_by(plant.sp, bb.sp) %>%
+#   summarize(co.max = max(co)) %>% # if max(co) = 0, a given bb.sp*fl.sp combo *never* co-occurred, and therefore will be assigned aneutral preference
+#   filter(co.max == 0) %>%
+#   mutate(preference.update = rep(1, n())) %>%
+#   dplyr::select(-co.max) %>%
+#   inner_join(fl_abund_sp) %>%
+#   select(-year)
 
@@
 
-# Add the forbidden links with preference set to 1 (neutral)
-  full_join(pref_forbidden_link_fill) %>%
-  mutate(preference = case_when(
-    is.na(preference) ~ preference.update,
-    !is.na(preference) ~ preference)
-    ) %>%
-  dplyr::select(-preference.update) %>%
-  mutate(fl.abund.wt = fl.abund*preference) %>%
+# Add the forbidden links with preference set to 1 (neutral) (not doing this anymore)
+  # full_join(pref_forbidden_link_fill) %>%
+  # mutate(preference = case_when(
+  #   is.na(preference) ~ preference.update,
+  #   !is.na(preference) ~ preference)
+  #   ) %>%
+  # dplyr::select(-preference.update) %>%
+  mutate(fl.abund.wt.bin = fl.abund*bin.preference,
+         fl.abund.wt.sqrt = fl.abund*sqrt.preference,
+         fl.abund.wt = fl.abund*preference) %>% # using sqrt-transformed preference allows me to avoid log-transforming when trying to fit GAMs
   mutate(yday = yday(date)) %>%
   group_by(site, date, yday, bb.sp) %>%
-  summarize(fl.abund.wt = sum(fl.abund.wt, na.rm = TRUE)) %>%
+  summarize(fl.abund.wt.bin = sum(fl.abund.wt.bin, na.rm = TRUE),
+            fl.abund.wt.sqrt = sum(fl.abund.wt.sqrt, na.rm = TRUE),
+            fl.abund.wt = sum(fl.abund.wt, na.rm = TRUE)) %>%
   ungroup() %>%
@@
 #### Floral resources
-```{r}
+##### After trying a lot of things, I'm pretty sure the only option I have for fitting a good GAM is using log-transformation.
+The question that remains is how to treat the concept of preference.
+My "raw" preference-weighted abundance gives our measure of preference a lot of influence.
+If we trust it, this may be most informative, and it allows the greatest variation in floral resources between BB species.
+But if there is anything misleading in our preference estimates, using the "raw" measures will do nothing to mitigate this.
+An approach that gives us a "softer" notion of preference is to take the sqrt of the raw prference before multiplying floral abundance by it.
+An argument for this approach is that it retains most of the meaning of the raw preference data while dampening its effect, which may help avoid spurious hotspots based on overestimated preference.
+Finally, an approach that says "we don't really trust our data very much" is to treat preference as binary: visited species get a 1, non-visited species get a 0, and all the "weighting" of floral resource abundance means is dropping floral species that a given BB species did not visit.
+In all these cases, though, the distribution of weighted floral abundance is strongly logarithmic (or exponential or something), so a log transformation is necessary.
+I tried softer methods, e.g.
+sqrt-transformation, Tweedie family, etc., but the problem with all these is that the variance fans out like crazy as you move up to the higher end of the distribution.
+Happily, it makes little difference which of these options we choose, as the resulting plots lead to the same interpretation.
+```{r message=FALSE}
+# These are the distributions of my response variable options 
+ggplot(fl_abund_wt, aes(fl.abund.wt.bin)) +
+  geom_histogram()
+
+ggplot(fl_abund_wt, aes(fl.abund.wt.sqrt)) +
+  geom_histogram()
+
+ggplot(fl_abund_wt, aes(fl.abund.wt)) +
+  geom_histogram()
+
 ### 2010
-weighted_fl_abundance.2010 <- bam(log(fl.abund.wt + 0.001) ~
+weighted_fl_abundance.2010 <- bam(log(fl.abund.wt + 0.01) ~
                                       s(bb.sp, bs = "re") +
@@
 #print(plot(weighted_fl_abundance.2010, allTerms = TRUE), pages = 4)
-weighted_fl_abundance.2010.ck <- check2D(weighted_fl_abundance.2010, 
-                                         x1 = "yday", x2 = "elev.mean")
+#weighted_fl_abundance.2010.ck <- check2D(weighted_fl_abundance.2010, x1 = "yday", x2 = "elev.mean")
+
+weighted_fl_abundance.2010.sqrt <- bam(log(fl.abund.wt.sqrt + 0.01) ~
+                                      s(bb.sp, bs = "re") +
+                                      #s(management, bs = "re") +
+                                      #s(site, bs = "re") +
+                                      te(yday, elev.mean, 
+                                                by = bb.sp,
+                                                #id = 0,
+                                                bs= c("gp", "tp"), 
+                                                k=c(8, 8), 
+                                                m=2),
+                                    data = filter(fl_abund_wt, year == 2010),
+                                    discrete = TRUE,
+                                    family = "gaussian",
+                                    select = TRUE,
+                                    method = "fREML") %>% getViz(nsim = 500)
+
+check.gamViz(weighted_fl_abundance.2010.sqrt)
+#weighted_fl_abundance.2010.sum.sqrt <- summary(weighted_fl_abundance.2010.sqrt)
+#print(plot(weighted_fl_abundance.2010.sqrt, allTerms = TRUE), pages = 4)
+#weighted_fl_abundance.2010.ck.sqrt <- check2D(weighted_fl_abundance.2010.sqrt, x1 = "yday", x2 = "elev.mean")
+
+weighted_fl_abundance.2010.bin <- bam(log(fl.abund.wt.bin + 0.01) ~
+                                      s(bb.sp, bs = "re") +
+                                      #s(management, bs = "re") +
+                                      #s(site, bs = "re") +
+                                      te(yday, elev.mean, 
+                                                by = bb.sp,
+                                                #id = 0,
+                                                bs= c("gp", "tp"), 
+                                                k=c(8, 8), 
+                                                m=2),
+                                    data = filter(fl_abund_wt, year == 2010),
+                                    discrete = TRUE,
+                                    family = "gaussian",
+                                    select = TRUE,
+                                    method = "fREML") %>% getViz(nsim = 500)
+
+check.gamViz(weighted_fl_abundance.2010.bin)
+#weighted_fl_abundance.2010.sum.bin <- summary(weighted_fl_abundance.2010.bin)
+#print(plot(weighted_fl_abundance.2010.bin, allTerms = TRUE), pages = 4)
+#weighted_fl_abundance.2010.ck.bin <- check2D(weighted_fl_abundance.2010.bin, x1 = "yday", x2 = "elev.mean")
 
 ### 2011
-weighted_fl_abundance.2011 <- bam(log(fl.abund.wt + 0.001) ~
+weighted_fl_abundance.2011 <- bam(log(fl.abund.wt + 0.01) ~
                                       s(bb.sp, bs = "re") +
@@
 #print(plot(weighted_fl_abundance.2011, allTerms = TRUE), pages = 4)
-weighted_fl_abundance.2011.ck <- check2D(weighted_fl_abundance.2011, 
-                                         x1 = "yday", x2 = "elev.mean")
+#weighted_fl_abundance.2011.ck <- check2D(weighted_fl_abundance.2011, x1 = "yday", x2 = "elev.mean")
+
+weighted_fl_abundance.2011.sqrt <- bam(log(fl.abund.wt.sqrt + 0.01) ~
+                                      s(bb.sp, bs = "re") +
+                                      #s(management, bs = "re") +
+                                      #s(site, bs = "re") +
+                                      te(yday, elev.mean, 
+                                                by = bb.sp,
+                                                #id = 0,
+                                                bs= c("gp", "tp"), 
+                                                k=c(8, 8), 
+                                                m=2),
+                                    data = filter(fl_abund_wt, year == 2011),
+                                    discrete = TRUE,
+                                    family = "gaussian",
+                                    select = TRUE,
+                                    method = "fREML") %>% getViz(nsim = 500)
+
+check.gamViz(weighted_fl_abundance.2011.sqrt)
+#weighted_fl_abundance.2011.sum.sqrt <- summary(weighted_fl_abundance.2011.sqrt)
+#print(plot(weighted_fl_abundance.2011.sqrt, allTerms = TRUE), pages = 4)
+#weighted_fl_abundance.2011.ck.sqrt <- check2D(weighted_fl_abundance.2011.sqrt, x1 = "yday", x2 = "elev.mean")
+
+weighted_fl_abundance.2011.bin <- bam(log(fl.abund.wt.bin + 0.01) ~
+                                      s(bb.sp, bs = "re") +
+                                      #s(management, bs = "re") +
+                                      #s(site, bs = "re") +
+                                      te(yday, elev.mean, 
+                                                by = bb.sp,
+                                                #id = 0,
+                                                bs= c("gp", "tp"), 
+                                                k=c(8, 8), 
+                                                m=2),
+                                    data = filter(fl_abund_wt, year == 2011),
+                                    discrete = TRUE,
+                                    family = "gaussian",
+                                    select = TRUE,
+                                    method = "fREML") %>% getViz(nsim = 500)
+
+check.gamViz(weighted_fl_abundance.2011.bin)
+#weighted_fl_abundance.2011.sum.bin <- summary(weighted_fl_abundance.2011.bin)
+#print(plot(weighted_fl_abundance.2011.bin, allTerms = TRUE), pages = 4)
+#weighted_fl_abundance.2011.ck.bin <- check2D(weighted_fl_abundance.2011.bin, x1 = "yday", x2 = "elev.mean")
 
 ### 2012
-weighted_fl_abundance.2012 <- bam(log(fl.abund.wt + 0.001) ~
+weighted_fl_abundance.2012 <- bam(log(fl.abund.wt + 0.01) ~
                                       s(bb.sp, bs = "re") +
@@
 #print(plot(weighted_fl_abundance.2012, allTerms = TRUE), pages = 4)
-weighted_fl_abundance.2012.ck <- check2D(weighted_fl_abundance.2012, 
-                                         x1 = "yday", x2 = "elev.mean")
+#weighted_fl_abundance.2012.ck <- check2D(weighted_fl_abundance.2012, x1 = "yday", x2 = "elev.mean")
+
+weighted_fl_abundance.2012.sqrt <- bam(log(fl.abund.wt.sqrt + 0.01) ~
+                                      s(bb.sp, bs = "re") +
+                                      #s(management, bs = "re") +
+                                      #s(site, bs = "re") +
+                                      te(yday, elev.mean, 
+                                                by = bb.sp,
+                                                #id = 0,
+                                                bs= c("gp", "tp"), 
+                                                k=c(8, 8), 
+                                                m=2),
+                                    data = filter(fl_abund_wt, year == 2012),
+                                    discrete = TRUE,
+                                    family = "gaussian",
+                                    select = TRUE,
+                                    method = "fREML") %>% getViz(nsim = 500)
+
+check.gamViz(weighted_fl_abundance.2012.sqrt)
+#weighted_fl_abundance.2012.sum.sqrt <- summary(weighted_fl_abundance.2012.sqrt)
+#print(plot(weighted_fl_abundance.2012.sqrt, allTerms = TRUE), pages = 4)
+#weighted_fl_abundance.2012.ck.sqrt <- check2D(weighted_fl_abundance.2012.sqrt, x1 = "yday", x2 = "elev.mean")
+
+weighted_fl_abundance.2012.bin <- bam(log(fl.abund.wt.bin + 0.01) ~
+                                      s(bb.sp, bs = "re") +
+                                      #s(management, bs = "re") +
+                                      #s(site, bs = "re") +
+                                      te(yday, elev.mean, 
+                                                by = bb.sp,
+                                                #id = 0,
+                                                bs= c("gp", "tp"), 
+                                                k=c(8, 8), 
+                                                m=2),
+                                    data = filter(fl_abund_wt, year == 2012),
+                                    discrete = TRUE,
+                                    family = "gaussian",
+                                    select = TRUE,
+                                    method = "fREML") %>% getViz(nsim = 500)
+
+check.gamViz(weighted_fl_abundance.2012.bin)
+#weighted_fl_abundance.2012.sum.bin <- summary(weighted_fl_abundance.2012.bin)
+#print(plot(weighted_fl_abundance.2012.bin, allTerms = TRUE), pages = 4)
+#weighted_fl_abundance.2012.ck.bin <- check2D(weighted_fl_abundance.2012.bin, x1 = "yday", x2 = "elev.mean")
 ```
@@
 #### Bumbles
-```{r}
+```{r message=FALSE}
 ### 2010
@@
                           m=2),
-                     data = filter(bb_abund_sp, year == 2010 & bb.total.abund >= 5),
+                     data = filter(bb_abund_sp, year == 2010 & bb.annual.abund >= 5),
                      discrete = TRUE,
@@
 #print(plot(bb_abund.2010, allTerms = TRUE), pages = 1)
-bb_abund.2010.ck <- check2D(bb_abund.2010, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals
+#bb_abund.2010.ck <- check2D(bb_abund.2010, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals
 
@@
                           m=2),
-                     data = filter(bb_abund_sp, year == 2011 & bb.total.abund >= 5),
+                     data = filter(bb_abund_sp, year == 2011 & bb.annual.abund >= 5),
                      discrete = TRUE,
@@
 #print(plot(bb_abund.2011, allTerms = TRUE), pages = 1)
-bb_abund.2011.ck <- check2D(bb_abund.2011, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals
+#bb_abund.2011.ck <- check2D(bb_abund.2011, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals
 
@@
                           m=2),
-                     data = filter(bb_abund_sp, year == 2012 & bb.total.abund >= 5),
+                     data = filter(bb_abund_sp, year == 2012 & bb.annual.abund >= 5),
                      discrete = TRUE,
@@
 #print(plot(bb_abund.2012, allTerms = TRUE), pages = 1)
-bb_abund.2012.ck <- check2D(bb_abund.2012, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals
+#bb_abund.2012.ck <- check2D(bb_abund.2012, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals
 ```
@@
 ### Extract predictions
-```{r}
+```{r message=FALSE}
 # Function to build data set over which to generate predictions
@@
     select(bb.sp) %>%
-    unique() %>%
-    na.omit() ##### NEED TO FIGURE OUT WHERE THE NA COMES FROM!
+    unique()
 
@@
 
-newdata2010.BB <- newdata_build(filter(bb_abund_sp, bb.total.abund >= 5), 2010)
-newdata2011.BB <- newdata_build(filter(bb_abund_sp, bb.total.abund >= 5), 2011)
-newdata2012.BB <- newdata_build(filter(bb_abund_sp, bb.total.abund >= 5), 2012)
+newdata2010.BB <- newdata_build(filter(bb_abund_sp, bb.annual.abund >= 5), 2010)
+newdata2011.BB <- newdata_build(filter(bb_abund_sp, bb.annual.abund >= 5), 2011)
+newdata2012.BB <- newdata_build(filter(bb_abund_sp, bb.annual.abund >= 5), 2012)
 
@@
 predict2010.FL <- predict_func(weighted_fl_abundance.2010, newdata2010.FL, 2010)
+predict2010.FL.sqrt <- predict_func(weighted_fl_abundance.2010.sqrt, newdata2010.FL, 2010)
+predict2010.FL.bin <- predict_func(weighted_fl_abundance.2010.bin, newdata2010.FL, 2010)
+
 predict2011.FL <- predict_func(weighted_fl_abundance.2011, newdata2011.FL, 2011)
+predict2011.FL.sqrt <- predict_func(weighted_fl_abundance.2011.sqrt, newdata2011.FL, 2011)
+predict2011.FL.bin <- predict_func(weighted_fl_abundance.2011.bin, newdata2011.FL, 2011)
+
 predict2012.FL <- predict_func(weighted_fl_abundance.2012, newdata2012.FL, 2012)
-predict_pool.FL <- bind_rows(predict2010.FL, predict2011.FL, predict2012.FL)
+predict2012.FL.sqrt <- predict_func(weighted_fl_abundance.2012.sqrt, newdata2012.FL, 2012)
+predict2012.FL.bin <- predict_func(weighted_fl_abundance.2012.bin, newdata2012.FL, 2012)
+
+predict_pool.FL <- bind_rows(predict2010.FL, predict2011.FL, predict2012.FL) %>%
+  mutate(date = as_date(yday))
+predict_pool.FL.sqrt <- bind_rows(predict2010.FL.sqrt, predict2011.FL.sqrt, predict2012.FL.sqrt) %>%
+  mutate(date = as_date(yday))
+predict_pool.FL.bin <- bind_rows(predict2010.FL.bin, predict2011.FL.bin, predict2012.FL.bin) %>%
+  mutate(date = as_date(yday))
 
@@
 predict2012.BB <- predict_func(bb_abund.2012, newdata2012.BB, 2012)
-predict_pool.BB <- bind_rows(predict2010.BB, predict2011.BB, predict2012.BB)
+predict_pool.BB <- bind_rows(predict2010.BB, predict2011.BB, predict2012.BB) %>%
+  mutate(date = as_date(yday))
 
 # Subset predictions
-predict_major.FL <- filter(predict_pool.FL, bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf"))
-predict_highelev.FL <- filter(predict_pool.FL, bb.sp %in% c("pyre", "mend", "mont"))
-predict_minor.FL <- filter(predict_pool.FL, bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci"))
+predict_major.FL <- filter(predict_pool.FL, 
+                           bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf")) %>%
+  mutate(bb.sp = case_when(
+    bb.sp == "bss" ~ "B. sensu-strictu",
+    bb.sp == "pasc" ~ "B. pascuorum",
+    bb.sp == "prat" ~ "B. pratorum",
+    bb.sp == "soro" ~ "B. soroensis",
+    bb.sp == "wurf" ~ "B. wurflenii"
+  ))
+predict_major.FL.sqrt <- filter(predict_pool.FL.sqrt, 
+                                bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf")) %>%
+  mutate(bb.sp = case_when(
+    bb.sp == "bss" ~ "B. sensu-strictu",
+    bb.sp == "pasc" ~ "B. pascuorum",
+    bb.sp == "prat" ~ "B. pratorum",
+    bb.sp == "soro" ~ "B. soroensis",
+    bb.sp == "wurf" ~ "B. wurflenii"
+  ))
+predict_major.FL.bin <- filter(predict_pool.FL.bin, 
+                               bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf")) %>%
+  mutate(bb.sp = case_when(
+    bb.sp == "bss" ~ "B. sensu-strictu",
+    bb.sp == "pasc" ~ "B. pascuorum",
+    bb.sp == "prat" ~ "B. pratorum",
+    bb.sp == "soro" ~ "B. soroensis",
+    bb.sp == "wurf" ~ "B. wurflenii"
+  ))
 
-predict_major.BB <- filter(predict_pool.BB, bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf"))
-predict_highelev.BB <- filter(predict_pool.BB, bb.sp %in% c("pyre", "mend", "mont"))
-predict_minor.BB <- filter(predict_pool.BB, bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci"))
+predict_highelev.FL <- filter(predict_pool.FL, 
+                              bb.sp %in% c("pyre", "mend", "mont")) %>%
+  mutate(bb.sp = case_when(
+    bb.sp == "pyre" ~ "B. pyrenaeus",
+    bb.sp == "mend" ~ "B. mendax",
+    bb.sp == "mont" ~ "B. monticola"
+  ))
+predict_highelev.FL.sqrt <- filter(predict_pool.FL.sqrt, 
+                                   bb.sp %in% c("pyre", "mend", "mont")) %>%
+  mutate(bb.sp = case_when(
+    bb.sp == "pyre" ~ "B. pyrenaeus",
+    bb.sp == "mend" ~ "B. mendax",
+    bb.sp == "mont" ~ "B. monticola"
+  ))
+predict_highelev.FL.bin <- filter(predict_pool.FL.bin, 
+                                  bb.sp %in% c("pyre", "mend", "mont")) %>%
+  mutate(bb.sp = case_when(
+    bb.sp == "pyre" ~ "B. pyrenaeus",
+    bb.sp == "mend" ~ "B. mendax",
+    bb.sp == "mont" ~ "B. monticola"
+  ))
+
+predict_minor.FL <- filter(predict_pool.FL, 
+                           bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci")) %>%
+  mutate(bb.sp = case_when(
+    bb.sp == "gers" ~ "B. gerstaeckeri",
+    bb.sp == "hort" ~ "B. hortorum",
+    bb.sp == "hypn" ~ "B. hypnorum",
+    bb.sp == "jone" ~ "B. jonellus",
+    bb.sp == "lapi" ~ "B. lapidarius",
+    bb.sp == "muci" ~ "B. mucidus"
+  ))
+predict_minor.FL.sqrt <- filter(predict_pool.FL.sqrt, 
+                                bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci")) %>% 
+  mutate(bb.sp = case_when(
+    bb.sp == "gers" ~ "B. gerstaeckeri",
+    bb.sp == "hort" ~ "B. hortorum",
+    bb.sp == "hypn" ~ "B. hypnorum",
+    bb.sp == "jone" ~ "B. jonellus",
+    bb.sp == "lapi" ~ "B. lapidarius",
+    bb.sp == "muci" ~ "B. mucidus"
+  ))
+predict_minor.FL.bin <- filter(predict_pool.FL.bin, 
+                               bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci")) %>% 
+  mutate(bb.sp = case_when(
+    bb.sp == "gers" ~ "B. gerstaeckeri",
+    bb.sp == "hort" ~ "B. hortorum",
+    bb.sp == "hypn" ~ "B. hypnorum",
+    bb.sp == "jone" ~ "B. jonellus",
+    bb.sp == "lapi" ~ "B. lapidarius",
+    bb.sp == "muci" ~ "B. mucidus"
+  ))
+
+predict_major.BB <- filter(predict_pool.BB, 
+                           bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf")) %>%
+  mutate(bb.sp = case_when(
+    bb.sp == "bss" ~ "B. sensu-strictu",
+    bb.sp == "pasc" ~ "B. pascuorum",
+    bb.sp == "prat" ~ "B. pratorum",
+    bb.sp == "soro" ~ "B. soroensis",
+    bb.sp == "wurf" ~ "B. wurflenii"
+  ))
+predict_highelev.BB <- filter(predict_pool.BB, 
+                              bb.sp %in% c("pyre", "mend", "mont")) %>%
+  mutate(bb.sp = case_when(
+    bb.sp == "pyre" ~ "B. pyrenaeus",
+    bb.sp == "mend" ~ "B. mendax",
+    bb.sp == "mont" ~ "B. monticola"
+  ))
+predict_minor.BB <- filter(predict_pool.BB, 
+                           bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci")) %>%
+  mutate(bb.sp = case_when(
+    bb.sp == "gers" ~ "B. gerstaeckeri",
+    bb.sp == "hort" ~ "B. hortorum",
+    bb.sp == "hypn" ~ "B. hypnorum",
+    bb.sp == "jone" ~ "B. jonellus",
+    bb.sp == "lapi" ~ "B. lapidarius",
+    bb.sp == "muci" ~ "B. mucidus"
+  ))
 ```
@@
 ```{r}
-ggplot(predict_major.FL, aes(yday, elev.mean, fill = pred)) +
+### FL GAMS
+#### Major
+ggplot(predict_major.FL, aes(date, elev.mean, fill = pred)) +
   geom_raster() +
-  geom_contour(aes(z = pred), color = "black", size = 0.25, binwidth = 1) +
-  scale_fill_viridis_c() +
-  facet_grid(year~bb.sp)
+  geom_contour(aes(z = pred), color = "black", size = 0.25) +
+  scale_fill_viridis_c(name = "log(wt. abund.)") +
+  facet_grid(year~bb.sp, ) +
+  xlab("Date") +
+  ylab("Elevation (m)") +
+  theme(strip.text.x = element_text(face = "italic"),
+        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
+ggsave("./output/GAM_major_FL.png")
 
-ggplot(predict_highelev.FL, aes(yday, elev.mean, fill = pred)) +
+ggplot(predict_major.FL.sqrt, aes(date, elev.mean, fill = pred)) +
   geom_raster() +
-  geom_contour(aes(z = pred), color = "black", size = 0.25, binwidth = 1) +
-  scale_fill_viridis_c() +
-  facet_grid(year~bb.sp)
+  geom_contour(aes(z = pred), color = "black", size = 0.25) +
+  scale_fill_viridis_c(name = "log(wt. abund.)") +
+  facet_grid(year~bb.sp) +
+  xlab("Date") +
+  ylab("Elevation (m)") +
+  theme(strip.text.x = element_text(face = "italic"),
+        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
+ggsave("./output/GAM_major_FL_sqrt.png")
 
-ggplot(predict_minor.FL, aes(yday, elev.mean, fill = pred)) +
+ggplot(predict_major.FL.bin, aes(date, elev.mean, fill = pred)) +
   geom_raster() +
-  geom_contour(aes(z = pred), color = "black", size = 0.25, binwidth = 1) +
-  scale_fill_viridis_c() +
-  facet_grid(year~bb.sp)
+  geom_contour(aes(z = pred), color = "black", size = 0.25) +
+  scale_fill_viridis_c(name = "log(wt. abund.)") +
+  facet_grid(year~bb.sp) +
+  xlab("Date") +
+  ylab("Elevation (m)") +
+  theme(strip.text.x = element_text(face = "italic"),
+        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
+ggsave("./output/GAM_major_FL_bin.png")
 
-# Plot
-ggplot(predict_major.BB, aes(yday, elev.mean, fill = log(pred))) +
+#### High Elev
+ggplot(predict_highelev.FL, aes(date, elev.mean, fill = pred)) +
   geom_raster() +
-  geom_contour(aes(z = log(pred)), color = "black", size = 0.25, binwidth = 1) +
-  scale_fill_viridis_c() +
-  facet_grid(year~bb.sp)
+  geom_contour(aes(z = pred), color = "black", size = 0.25) +
+  scale_fill_viridis_c(name = "log(wt. abund.)") +
+  facet_grid(year~bb.sp) +
+  xlab("Date") +
+  ylab("Elevation (m)") +
+  theme(strip.text.x = element_text(face = "italic"),
+        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
+ggsave("./output/GAM_highelev_FL.png")
 
-ggplot(predict_highelev.BB, aes(yday, elev.mean, fill = log(pred))) +
+ggplot(predict_highelev.FL.sqrt, aes(date, elev.mean, fill = pred)) +
   geom_raster() +
-  geom_contour(aes(z = log(pred)), color = "black", size = 0.25, binwidth = 1) +
-  scale_fill_viridis_c() +
-  facet_grid(year~bb.sp)
+  scale_fill_viridis_c(name = "log(wt. abund.)") +
+  facet_grid(year~bb.sp) +
+  xlab("Date") +
+  ylab("Elevation (m)") +
+  theme(strip.text.x = element_text(face = "italic"),
+        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
+ggsave("./output/GAM_highelev_FL_sqrt.png")
 
-ggplot(predict_minor.BB, aes(yday, elev.mean, fill = log(pred))) +
+ggplot(predict_highelev.FL.bin, aes(date, elev.mean, fill = pred)) +
   geom_raster() +
-  geom_contour(aes(z = log(pred)), color = "black", size = 0.25, binwidth = 1) +
-  scale_fill_viridis_c() +
-  facet_grid(year~bb.sp)
+  geom_contour(aes(z = pred), color = "black", size = 0.25) +
+  scale_fill_viridis_c(name = "log(wt. abund.)") +
+  facet_grid(year~bb.sp) +
+  xlab("Date") +
+  ylab("Elevation (m)") +
+  theme(strip.text.x = element_text(face = "italic"),
+        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
+ggsave("./output/GAM_highelev_FL_bin.png")
+
+#### Minor
+ggplot(predict_minor.FL, aes(date, elev.mean, fill = pred)) +
+  geom_raster() +
+  geom_contour(aes(z = pred), color = "black", size = 0.25) +
+  scale_fill_viridis_c(name = "log(wt. abund.)") +
+  facet_grid(year~bb.sp) +
+  xlab("Date") +
+  ylab("Elevation (m)") +
+  theme(strip.text.x = element_text(face = "italic"),
+        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
+ggsave("./output/GAM_minor_FL.png")
+
+ggplot(predict_minor.FL.sqrt, aes(date, elev.mean, fill = pred)) +
+  geom_raster() +
+  geom_contour(aes(z = pred), color = "black", size = 0.25) +
+  scale_fill_viridis_c(name = "log(wt. abund.)") +
+  facet_grid(year~bb.sp) +
+  xlab("Date") +
+  ylab("Elevation (m)") +
+  theme(strip.text.x = element_text(face = "italic"),
+        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
+ggsave("./output/GAM_minor_FL_sqrt.png")
+
+ggplot(predict_minor.FL.bin, aes(date, elev.mean, fill = pred)) +
+  geom_raster() +
+  geom_contour(aes(z = pred), color = "black", size = 0.25) +
+  scale_fill_viridis_c(name = "log(wt. abund.)") +
+  facet_grid(year~bb.sp) +
+  xlab("Date") +
+  ylab("Elevation (m)") +
+  theme(strip.text.x = element_text(face = "italic"),
+        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
+ggsave("./output/GAM_minor_FL_bin.png")
+
+### BB GAMs
+ggplot(predict_major.BB, aes(date, elev.mean, fill = pred)) +
+  geom_raster() +
+  geom_contour(aes(z = log(pred)), color = "black", size = 0.25) +
+  scale_fill_viridis_c(name = "wt. abund.") +
+  facet_grid(year~bb.sp) +
+  xlab("Date") +
+  ylab("Elevation (m)") +
+  theme(strip.text.x = element_text(face = "italic"),
+        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
+ggsave("./output/GAM_major_BB.png")
+
+ggplot(predict_highelev.BB, aes(date, elev.mean, fill = pred)) +
+  geom_raster() +
+  geom_contour(aes(z = log(pred)), color = "black", size = 0.25) +
+  scale_fill_viridis_c(name = "wt. abund.") +
+  facet_grid(year~bb.sp) +
+  xlab("Date") +
+  ylab("Elevation (m)") +
+  theme(strip.text.x = element_text(face = "italic"),
+        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
+ggsave("./output/GAM_highelev_BB.png")
+
+ggplot(predict_minor.BB, aes(date, elev.mean, fill = pred)) +
+  geom_raster() +
+  geom_contour(aes(z = log(pred)), color = "black", size = 0.25) +
+  scale_fill_viridis_c(name = "wt. abund") +
+  facet_grid(year~bb.sp) +
+  xlab("Date") +
+  ylab("Elevation (m)") +
+  theme(strip.text.x = element_text(face = "italic"),
+        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
+ggsave("./output/GAM_minor_BB.png") 
 ```

### 2020-09-14 analyses/trash_can.R

@@ -1200,4 +1200,146 @@ transects_diff <- beta_site %>%
 WN_mantel <- mantel.partial(elev_diff, WN_diff, transects_diff)
 ST_mantel <- mantel.partial(elev_diff, ST_diff, transects_diff)
 OS_mantel <- mantel.partial(elev_diff, OS_diff, transects_diff)
+
+### 2011
+#### Let's explore whether the different transformation options I'm weighing have any effect on the overall pattern that we end up visualizing
+
+#### Option A: log(n + 0.001) with gaussian
+weighted_fl_abundance.2011.1 <- bam(log(fl.abund.wt + 0.01) ~
+                                      s(bb.sp, bs = "re") +
+                                      #s(management, bs = "re") +
+                                      #s(site, bs = "re") +
+                                      te(yday, elev.mean, 
+                                         by = bb.sp,
+                                         #id = 0,
+                                         bs= c("gp", "tp"), 
+                                         k=c(8, 8), 
+                                         m=2),
+                                    data = filter(fl_abund_wt, year == 2011),
+                                    discrete = TRUE,
+                                    family = "gaussian",
+                                    select = TRUE,
+                                    method = "fREML") %>% getViz(nsim = 500)
+
+check.gamViz(weighted_fl_abundance.2011.1)
+#weighted_fl_abundance.2011.sum.1 <- summary(weighted_fl_abundance.2011.1)
+#print(plot(weighted_fl_abundance.2011.1, allTerms = TRUE), pages = 4)
+weighted_fl_abundance.2011.ck.1 <- check2D(weighted_fl_abundance.2011.1, 
+                                           x1 = "yday", x2 = "elev.mean")
+
+#### Option B: sqrt(n) with tweedie
+weighted_fl_abundance.2011.2 <- bam(sqrt(fl.abund.wt) ~
+                                      s(bb.sp, bs = "re") +
+                                      #s(management, bs = "re") +
+                                      #s(site, bs = "re") +
+                                      te(yday, elev.mean, 
+                                         by = bb.sp,
+                                         #id = 0,
+                                         bs= c("gp", "tp"), 
+                                         k=c(8, 8), 
+                                         m=2),
+                                    data = filter(fl_abund_wt, year == 2011),
+                                    discrete = TRUE,
+                                    family = "tw",
+                                    select = TRUE,
[…]

### 2020-09-15 analyses/abund_div_GAM.Rmd
refined GDM analysis

@@
 Happily, it makes little difference which of these options we choose, as the resulting plots lead to the same interpretation.
-```{r message=FALSE}
+```{r message=FALSE, warning=FALSE}
 # These are the distributions of my response variable options 
@@
 
-check.gamViz(weighted_fl_abundance.2010)
+#check.gamViz(weighted_fl_abundance.2010)
 #weighted_fl_abundance.2010.sum <- summary(weighted_fl_abundance.2010)
@@
 
-check.gamViz(weighted_fl_abundance.2010.sqrt)
+#check.gamViz(weighted_fl_abundance.2010.sqrt)
 #weighted_fl_abundance.2010.sum.sqrt <- summary(weighted_fl_abundance.2010.sqrt)
@@
 
-check.gamViz(weighted_fl_abundance.2010.bin)
+#check.gamViz(weighted_fl_abundance.2010.bin)
 #weighted_fl_abundance.2010.sum.bin <- summary(weighted_fl_abundance.2010.bin)
@@
 
-check.gamViz(weighted_fl_abundance.2011)
+#check.gamViz(weighted_fl_abundance.2011)
 #weighted_fl_abundance.2011.sum <- summary(weighted_fl_abundance.2011)
@@
 
-check.gamViz(weighted_fl_abundance.2011.sqrt)
+#check.gamViz(weighted_fl_abundance.2011.sqrt)
 #weighted_fl_abundance.2011.sum.sqrt <- summary(weighted_fl_abundance.2011.sqrt)
@@
 
-check.gamViz(weighted_fl_abundance.2011.bin)
+#check.gamViz(weighted_fl_abundance.2011.bin)
 #weighted_fl_abundance.2011.sum.bin <- summary(weighted_fl_abundance.2011.bin)
@@
 
-check.gamViz(weighted_fl_abundance.2012)
+#check.gamViz(weighted_fl_abundance.2012)
 #weighted_fl_abundance.2012.sum <- summary(weighted_fl_abundance.2012)
@@
 
-check.gamViz(weighted_fl_abundance.2012.sqrt)
+#check.gamViz(weighted_fl_abundance.2012.sqrt)
 #weighted_fl_abundance.2012.sum.sqrt <- summary(weighted_fl_abundance.2012.sqrt)
@@
 
-check.gamViz(weighted_fl_abundance.2012.bin)
+#check.gamViz(weighted_fl_abundance.2012.bin)
 #weighted_fl_abundance.2012.sum.bin <- summary(weighted_fl_abundance.2012.bin)
@@
 
-check.gamViz(bb_abund.2010)
+#check.gamViz(bb_abund.2010)
 #bb_abund.2010.sum <- summary(bb_abund.2010)
@@
 
-check.gamViz(bb_abund.2011)
+#check.gamViz(bb_abund.2011)
 #bb_abund.2011.sum <- summary(bb_abund.2011)
@@
 
-check.gamViz(bb_abund.2012)
+#check.gamViz(bb_abund.2012)
 #bb_abund.2012.sum <- summary(bb_abund.2012)
@@
 predict_pool.FL <- bind_rows(predict2010.FL, predict2011.FL, predict2012.FL) %>%
-  mutate(date = as_date(yday))
+  mutate(date = as_date(yday),
+         subject = factor(rep("FL", n())))
 predict_pool.FL.sqrt <- bind_rows(predict2010.FL.sqrt, predict2011.FL.sqrt, predict2012.FL.sqrt) %>%
-  mutate(date = as_date(yday))
+  mutate(date = as_date(yday),
+         subject = factor(rep("FL", n())))
 predict_pool.FL.bin <- bind_rows(predict2010.FL.bin, predict2011.FL.bin, predict2012.FL.bin) %>%
-  mutate(date = as_date(yday))
+  mutate(date = as_date(yday),
+         subject = factor(rep("FL", n())))
 
@@
 predict_pool.BB <- bind_rows(predict2010.BB, predict2011.BB, predict2012.BB) %>%
-  mutate(date = as_date(yday))
+  mutate(date = as_date(yday),
+         subject = factor(rep("BB", n())),
+         pred = log(pred))
+
+predict_pool <- bind_rows(predict_pool.FL, predict_pool.BB)
+predict_pool.sqrt <- bind_rows(predict_pool.FL.sqrt, predict_pool.BB)
+predict_pool.bin <- bind_rows(predict_pool.FL.bin, predict_pool.BB)
 
 # Subset predictions
-predict_major.FL <- filter(predict_pool.FL, 
+predict_major <- filter(predict_pool, 
                            bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf")) %>%
   mutate(bb.sp = case_when(
-    bb.sp == "bss" ~ "B. sensu-strictu",
-    bb.sp == "pasc" ~ "B. pascuorum",
-    bb.sp == "prat" ~ "B. pratorum",
-    bb.sp == "soro" ~ "B. soroensis",
-    bb.sp == "wurf" ~ "B. wurflenii"
+    bb.sp == "bss" ~ "B. ss.",
+    bb.sp == "pasc" ~ "B. pasc",
+    bb.sp == "prat" ~ "B. prat",
+    bb.sp == "soro" ~ "B. soro",
+    bb.sp == "wurf" ~ "B. wurf"
   ))
-predict_major.FL.sqrt <- filter(predict_pool.FL.sqrt, 
+
+predict_major.sqrt <- filter(predict_pool.sqrt, 
                            bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf")) %>%
   mutate(bb.sp = case_when(
-    bb.sp == "bss" ~ "B. sensu-strictu",
-    bb.sp == "pasc" ~ "B. pascuorum",
-    bb.sp == "prat" ~ "B. pratorum",
-    bb.sp == "soro" ~ "B. soroensis",
-    bb.sp == "wurf" ~ "B. wurflenii"
+    bb.sp == "bss" ~ "B. ss.",
+    bb.sp == "pasc" ~ "B. pasc",
+    bb.sp == "prat" ~ "B. prat",
+    bb.sp == "soro" ~ "B. soro",
+    bb.sp == "wurf" ~ "B. wurf"
   ))
-predict_major.FL.bin <- filter(predict_pool.FL.bin, 
+
+predict_major.bin <- filter(predict_pool.bin, 
                            bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf")) %>%
   mutate(bb.sp = case_when(
-    bb.sp == "bss" ~ "B. sensu-strictu",
-    bb.sp == "pasc" ~ "B. pascuorum",
-    bb.sp == "prat" ~ "B. pratorum",
-    bb.sp == "soro" ~ "B. soroensis",
-    bb.sp == "wurf" ~ "B. wurflenii"
+    bb.sp == "bss" ~ "B. ss.",
+    bb.sp == "pasc" ~ "B. pasc",
+    bb.sp == "prat" ~ "B. prat",
+    bb.sp == "soro" ~ "B. soro",
+    bb.sp == "wurf" ~ "B. wurf"
   ))
 
-predict_highelev.FL <- filter(predict_pool.FL, 
+predict_highelev <- filter(predict_pool, 
                               bb.sp %in% c("pyre", "mend", "mont")) %>%
   mutate(bb.sp = case_when(
-    bb.sp == "pyre" ~ "B. pyrenaeus",
-    bb.sp == "mend" ~ "B. mendax",
-    bb.sp == "mont" ~ "B. monticola"
+    bb.sp == "pyre" ~ "B. pyre",
+    bb.sp == "mend" ~ "B. mend",
+    bb.sp == "mont" ~ "B. mont"
   ))
-predict_highelev.FL.sqrt <- filter(predict_pool.FL.sqrt, 
+
+predict_highelev.sqrt <- filter(predict_pool.sqrt, 
                                    bb.sp %in% c("pyre", "mend", "mont")) %>%
   mutate(bb.sp = case_when(
-    bb.sp == "pyre" ~ "B. pyrenaeus",
-    bb.sp == "mend" ~ "B. mendax",
-    bb.sp == "mont" ~ "B. monticola"
+    bb.sp == "pyre" ~ "B. pyre",
+    bb.sp == "mend" ~ "B. mend",
+    bb.sp == "mont" ~ "B. mont"
   ))
-predict_highelev.FL.bin <- filter(predict_pool.FL.bin, 
+
+predict_highelev.bin <- filter(predict_pool.bin, 
                                   bb.sp %in% c("pyre", "mend", "mont")) %>%
   mutate(bb.sp = case_when(
-    bb.sp == "pyre" ~ "B. pyrenaeus",
-    bb.sp == "mend" ~ "B. mendax",
-    bb.sp == "mont" ~ "B. monticola"
+    bb.sp == "pyre" ~ "B. pyre",
+    bb.sp == "mend" ~ "B. mend",
+    bb.sp == "mont" ~ "B. mont"
   ))
 
-predict_minor.FL <- filter(predict_pool.FL, 
+predict_minor <- filter(predict_pool, 
                            bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci")) %>%
   mutate(bb.sp = case_when(
-    bb.sp == "gers" ~ "B. gerstaeckeri",
-    bb.sp == "hort" ~ "B. hortorum",
-    bb.sp == "hypn" ~ "B. hypnorum",
-    bb.sp == "jone" ~ "B. jonellus",
-    bb.sp == "lapi" ~ "B. lapidarius",
-    bb.sp == "muci" ~ "B. mucidus"
+    bb.sp == "gers" ~ "B. gers",
+    bb.sp == "hort" ~ "B. hort",
+    bb.sp == "hypn" ~ "B. hypn",
+    bb.sp == "jone" ~ "B. jone",
+    bb.sp == "lapi" ~ "B. lapi",
+    bb.sp == "muci" ~ "B. muci"
   ))
-predict_minor.FL.sqrt <- filter(predict_pool.FL.sqrt, 
+
+predict_minor.sqrt <- filter(predict_pool.sqrt, 
                                 bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci")) %>% 
   mutate(bb.sp = case_when(
-    bb.sp == "gers" ~ "B. gerstaeckeri",
-    bb.sp == "hort" ~ "B. hortorum",
-    bb.sp == "hypn" ~ "B. hypnorum",
-    bb.sp == "jone" ~ "B. jonellus",
-    bb.sp == "lapi" ~ "B. lapidarius",
-    bb.sp == "muci" ~ "B. mucidus"
+    bb.sp == "gers" ~ "B. gers",
+    bb.sp == "hort" ~ "B. hort",
+    bb.sp == "hypn" ~ "B. hypn",
+    bb.sp == "jone" ~ "B. jone",
+    bb.sp == "lapi" ~ "B. lapi",
+    bb.sp == "muci" ~ "B. muci"
   ))
-predict_minor.FL.bin <- filter(predict_pool.FL.bin, 
+
+predict_minor.bin <- filter(predict_pool.bin, 
                                bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci")) %>% 
   mutate(bb.sp = case_when(
-    bb.sp == "gers" ~ "B. gerstaeckeri",
-    bb.sp == "hort" ~ "B. hortorum",
-    bb.sp == "hypn" ~ "B. hypnorum",
-    bb.sp == "jone" ~ "B. jonellus",
-    bb.sp == "lapi" ~ "B. lapidarius",
-    bb.sp == "muci" ~ "B. mucidus"
-  ))
-
-predict_major.BB <- filter(predict_pool.BB, 
-                           bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf")) %>%
-  mutate(bb.sp = case_when(
-    bb.sp == "bss" ~ "B. sensu-strictu",
-    bb.sp == "pasc" ~ "B. pascuorum",
-    bb.sp == "prat" ~ "B. pratorum",
-    bb.sp == "soro" ~ "B. soroensis",
-    bb.sp == "wurf" ~ "B. wurflenii"
-  ))
-predict_highelev.BB <- filter(predict_pool.BB, 
-                              bb.sp %in% c("pyre", "mend", "mont")) %>%
-  mutate(bb.sp = case_when(
-    bb.sp == "pyre" ~ "B. pyrenaeus",
-    bb.sp == "mend" ~ "B. mendax",
-    bb.sp == "mont" ~ "B. monticola"
-  ))
-predict_minor.BB <- filter(predict_pool.BB, 
-                           bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci")) %>%
-  mutate(bb.sp = case_when(
-    bb.sp == "gers" ~ "B. gerstaeckeri",
-    bb.sp == "hort" ~ "B. hortorum",
-    bb.sp == "hypn" ~ "B. hypnorum",
-    bb.sp == "jone" ~ "B. jonellus",
-    bb.sp == "lapi" ~ "B. lapidarius",
-    bb.sp == "muci" ~ "B. mucidus"
+    bb.sp == "gers" ~ "B. gers",
+    bb.sp == "hort" ~ "B. hort",
+    bb.sp == "hypn" ~ "B. hypn",
+    bb.sp == "jone" ~ "B. jone",
+    bb.sp == "lapi" ~ "B. lapi",
+    bb.sp == "muci" ~ "B. muci"
   ))
@@
 ```{r}
-### FL GAMS
 #### Major
-ggplot(predict_major.FL, aes(date, elev.mean, fill = pred)) +
+ggplot(predict_major, aes(date, elev.mean, fill = pred)) +
   geom_raster() +
   geom_contour(aes(z = pred), color = "black", size = 0.25) +
-  scale_fill_viridis_c(name = "log(wt. abund.)") +
-  facet_grid(year~bb.sp, ) +
+  scale_fill_viridis_c(name = "log(abund.)") +
+  #facet_grid(bb.sp + subject ~ year) +
+  facet_grid(year ~ bb.sp + subject) +
   xlab("Date") +
   ylab("Elevation (m)") +
-  theme(strip.text.x = element_text(face = "italic"),
-        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
-ggsave("./output/GAM_major_FL.png")
+  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
+ggsave("./output/GAM_major.png", height = 5, width = 11, units = "in")
 
-ggplot(predict_major.FL.sqrt, aes(date, elev.mean, fill = pred)) +
+ggplot(predict_major.sqrt, aes(date, elev.mean, fill = pred)) +
   geom_raster() +
   geom_contour(aes(z = pred), color = "black", size = 0.25) +
-  scale_fill_viridis_c(name = "log(wt. abund.)") +
-  facet_grid(year~bb.sp) +
+  scale_fill_viridis_c(name = "log(abund.)") +
+  #facet_grid(bb.sp + subject ~ year) +
+  facet_grid(year ~ bb.sp + subject) +
   xlab("Date") +
   ylab("Elevation (m)") +
-  theme(strip.text.x = element_text(face = "italic"),
-        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
-ggsave("./output/GAM_major_FL_sqrt.png")
+  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
+ggsave("./output/GAM_major_sqrt.png")
 
-ggplot(predict_major.FL.bin, aes(date, elev.mean, fill = pred)) +
+ggplot(predict_major.bin, aes(date, elev.mean, fill = pred)) +
   geom_raster() +
   geom_contour(aes(z = pred), color = "black", size = 0.25) +
-  scale_fill_viridis_c(name = "log(wt. abund.)") +
-  facet_grid(year~bb.sp) +
+  scale_fill_viridis_c(name = "log(abund.)") +
+  #facet_grid(bb.sp + subject ~ year) +
+  facet_grid(year ~ bb.sp + subject) +
   xlab("Date") +
   ylab("Elevation (m)") +
-  theme(strip.text.x = element_text(face = "italic"),
-        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
-ggsave("./output/GAM_major_FL_bin.png")
+  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
+ggsave("./output/GAM_major_bin.png")
 
 #### High Elev
-ggplot(predict_highelev.FL, aes(date, elev.mean, fill = pred)) +
+ggplot(predict_highelev, aes(date, elev.mean, fill = pred)) +
   geom_raster() +
   geom_contour(aes(z = pred), color = "black", size = 0.25) +
-  scale_fill_viridis_c(name = "log(wt. abund.)") +
-  facet_grid(year~bb.sp) +
+  scale_fill_viridis_c(name = "log(abund.)") +
+  #facet_grid(bb.sp + subject ~ year) +
+  facet_grid(year ~ bb.sp + subject) +
   xlab("Date") +
   ylab("Elevation (m)") +
-  theme(strip.text.x = element_text(face = "italic"),
-        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
-ggsave("./output/GAM_highelev_FL.png")
+  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
+ggsave("./output/GAM_highelev.png", height = 5, width = 11, units = "in")
 
-ggplot(predict_highelev.FL.sqrt, aes(date, elev.mean, fill = pred)) +
+ggplot(predict_highelev.sqrt, aes(date, elev.mean, fill = pred)) +
   geom_raster() +
-  scale_fill_viridis_c(name = "log(wt. abund.)") +
-  facet_grid(year~bb.sp) +
+  geom_contour(aes(z = pred), color = "black", size = 0.25) +
+  scale_fill_viridis_c(name = "log(abund.)") +
+  #facet_grid(bb.sp + subject ~ year) +
+  facet_grid(year ~ bb.sp + subject) +
   xlab("Date") +
   ylab("Elevation (m)") +
-  theme(strip.text.x = element_text(face = "italic"),
-        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
-ggsave("./output/GAM_highelev_FL_sqrt.png")
+  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
+ggsave("./output/GAM_highelev_sqrt.png", height = 5, width = 11, units = "in")
 
-ggplot(predict_highelev.FL.bin, aes(date, elev.mean, fill = pred)) +
+ggplot(predict_highelev.bin, aes(date, elev.mean, fill = pred)) +
   geom_raster() +
   geom_contour(aes(z = pred), color = "black", size = 0.25) +
-  scale_fill_viridis_c(name = "log(wt. abund.)") +
-  facet_grid(year~bb.sp) +
+  scale_fill_viridis_c(name = "log(abund.)") +
+  #facet_grid(bb.sp + subject ~ year) +
+  facet_grid(year ~ bb.sp + subject) +
   xlab("Date") +
   ylab("Elevation (m)") +
-  theme(strip.text.x = element_text(face = "italic"),
-        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
-ggsave("./output/GAM_highelev_FL_bin.png")
+  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
+ggsave("./output/GAM_highelev_bin.png", height = 5, width = 11, units = "in")
 
 #### Minor
-ggplot(predict_minor.FL, aes(date, elev.mean, fill = pred)) +
+ggplot(predict_minor, aes(date, elev.mean, fill = pred)) +
   geom_raster() +
   geom_contour(aes(z = pred), color = "black", size = 0.25) +
-  scale_fill_viridis_c(name = "log(wt. abund.)") +
-  facet_grid(year~bb.sp) +
+  scale_fill_viridis_c(name = "log(abund.)") +
+  #facet_grid(bb.sp + subject ~ year) +
+  facet_grid(year ~ bb.sp + subject) +
   xlab("Date") +
   ylab("Elevation (m)") +
-  theme(strip.text.x = element_text(face = "italic"),
-        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
-ggsave("./output/GAM_minor_FL.png")
+  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
+ggsave("./output/GAM_minor.png", height = 5, width = 11, units = "in")
 
-ggplot(predict_minor.FL.sqrt, aes(date, elev.mean, fill = pred)) +
+ggplot(predict_minor.sqrt, aes(date, elev.mean, fill = pred)) +
   geom_raster() +
   geom_contour(aes(z = pred), color = "black", size = 0.25) +
-  scale_fill_viridis_c(name = "log(wt. abund.)") +
-  facet_grid(year~bb.sp) +
+  scale_fill_viridis_c(name = "log(abund.)") +
+  #facet_grid(bb.sp + subject ~ year) +
+  facet_grid(year ~ bb.sp + subject) +
   xlab("Date") +
   ylab("Elevation (m)") +
-  theme(strip.text.x = element_text(face = "italic"),
-        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
-ggsave("./output/GAM_minor_FL_sqrt.png")
+  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
+ggsave("./output/GAM_minor_sqrt.png", height = 5, width = 11, units = "in")
 
-ggplot(predict_minor.FL.bin, aes(date, elev.mean, fill = pred)) +
+ggplot(predict_minor.bin, aes(date, elev.mean, fill = pred)) +
   geom_raster() +
   geom_contour(aes(z = pred), color = "black", size = 0.25) +
-  scale_fill_viridis_c(name = "log(wt. abund.)") +
-  facet_grid(year~bb.sp) +
+  scale_fill_viridis_c(name = "log(abund.)") +
+  #facet_grid(bb.sp + subject ~ year) +
+  facet_grid(year ~ bb.sp + subject) +
   xlab("Date") +
   ylab("Elevation (m)") +
-  theme(strip.text.x = element_text(face = "italic"),
-        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
-ggsave("./output/GAM_minor_FL_bin.png")
-
-### BB GAMs
-ggplot(predict_major.BB, aes(date, elev.mean, fill = pred)) +
-  geom_raster() +
-  geom_contour(aes(z = log(pred)), color = "black", size = 0.25) +
-  scale_fill_viridis_c(name = "wt. abund.") +
-  facet_grid(year~bb.sp) +
-  xlab("Date") +
-  ylab("Elevation (m)") +
-  theme(strip.text.x = element_text(face = "italic"),
-        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
-ggsave("./output/GAM_major_BB.png")
-
-ggplot(predict_highelev.BB, aes(date, elev.mean, fill = pred)) +
-  geom_raster() +
-  geom_contour(aes(z = log(pred)), color = "black", size = 0.25) +
-  scale_fill_viridis_c(name = "wt. abund.") +
-  facet_grid(year~bb.sp) +
-  xlab("Date") +
-  ylab("Elevation (m)") +
-  theme(strip.text.x = element_text(face = "italic"),
-        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
-ggsave("./output/GAM_highelev_BB.png")
-
-ggplot(predict_minor.BB, aes(date, elev.mean, fill = pred)) +
-  geom_raster() +
-  geom_contour(aes(z = log(pred)), color = "black", size = 0.25) +
-  scale_fill_viridis_c(name = "wt. abund") +
-  facet_grid(year~bb.sp) +
-  xlab("Date") +
-  ylab("Elevation (m)") +
-  theme(strip.text.x = element_text(face = "italic"),
-        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
-ggsave("./output/GAM_minor_BB.png") 
+  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
+ggsave("./output/GAM_minor_bin.png", height = 5, width = 11, units = "in")
 ```

### 2020-09-15 analyses/beta_diversity.Rmd
refined GDM analysis

--- 2020-07-13 analyses/beta_diversity.Rmd
@@
 ```{r, include=FALSE, echo=FALSE}
+# A function for converting a geometry column back to x and y columns
+# https://github.com/r-spatial/sf/issues/231
+sfc_as_cols <- function(x, names = c("x","y")) {
+  stopifnot(inherits(x,"sf") && inherits(sf::st_geometry(x),"sfc_POINT"))
+  ret <- sf::st_coordinates(x)
+  ret <- tibble::as_tibble(ret)
+  stopifnot(length(names) == ncol(ret))
+  x <- x[ , !names(x) %in% names]
+  ret <- setNames(ret,names)
+  dplyr::bind_cols(x,ret)
+}
+
 site_data <- read_csv("../processed_data/site_data.csv") %>%
@@
          elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) %>% # turn elav.class into an ordered factor
-  semi_join(read_csv("../processed_data/network.csv"), by = "site")
+  semi_join(read_csv("../processed_data/network.csv"), by = "site") %>% # drop sites that were not sampled
+  st_as_sf(coords = c("lon", "lat")) %>%
+  st_set_crs(4326) %>%
+  st_transform(5684) %>%
+  sfc_as_cols() %>%
+  as_tibble() %>%
+  dplyr::select(-geometry)
 ```
 
-### Load network and survey data and generate site.year and site.month.year webs
+### Load network and survey data
 ```{r}
-net <- read_csv("../processed_data/network.csv")%>%
-  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
-  unite(site.year.day, site, year, dayofyear, sep = "_", remove = FALSE) %>%
-  mutate(year = factor(year)) %>% # convert year to factor
-  group_by(site, date) %>%
-  arrange(site, date)
+net <- read_csv("../processed_data/network.csv") %>%
+  unite(site.year.day, site, year, dayofyear, sep = "_", remove = FALSE) %>% # create site.date field that can be split up again later
+  mutate(year = factor(year)) # convert year to factor
 
 survey <- read_csv("../processed_data/floral_survey.csv") %>%
-  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
+  semi_join(net, by = "plant.sp") %>% # Consider only visited plant spp. (anytime, anywhere, not specifically for a given site*date)
   dplyr::select(site, date, plant.sp, flower.cover)
-
-site.year_transects <- net %>%
-  group_by(site, year) %>%
-  summarize(transects = length(unique(date)))
-
-site_transects <- net %>%
-  group_by(site) %>%
-  summarize(transects = length(unique(date)))
-
-web_site.year <- net %>%
-  group_by(site.year, bb.sp, plant.sp) %>%
-  summarize(freq = n()) %>%
-  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs() # convert to to bipartite web
-
-web_site <- net %>%
-  group_by(site, bb.sp, plant.sp) %>%
-  summarize(freq = n()) %>% 
-  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs() # convert to to bipartite web
-
-web_site.date.2010 <- net %>%
-  filter(year == 2010) %>%
-  group_by(site.year.day, bb.sp, plant.sp) %>%
-  summarize(freq = n()) %>%
-  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year.day, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs() # convert to to bipartite web
-
-web_site.date.2011 <- net %>%
-  filter(year == 2011) %>%
-  group_by(site.year.day, bb.sp, plant.sp) %>%
-  summarize(freq = n()) %>%
-  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year.day, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs() # convert to to bipartite web
-
-web_site.date.2012 <- net %>%
-  filter(year == 2012) %>%
-  group_by(site.year.day, bb.sp, plant.sp) %>%
-  summarize(freq = n()) %>%
-  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year.day, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs() # convert to to bipartite web
-
-array_site <- webs2array(web_site)
-array_site.year <- webs2array(web_site.year)
-array_site.date.2010 <- webs2array(web_site.date.2010)
-array_site.date.2011 <- webs2array(web_site.date.2011)
-array_site.date.2012 <- webs2array(web_site.date.2012)
 ```
@@
 ST = beta_ST, the dissimilarity (component) explained by difference in species community composition 
-```{r}
-beta_site <- betalinkr_multi(array_site, partition.st = TRUE, partition.rr = TRUE) %>%
-  as_tibble() %>%
-  rename(site1 = i) %>%
-  rename(site2 = j)%>%
+```{r message=FALSE}
+# Function to convert networks to webs arrays
+web_prep <- function(net, y) {
+  out <- net %>%
+    filter(year == y) %>%
+    unite(site.date, c(site, date), sep = "_") %>%
+  group_by(site.date, bb.sp, plant.sp) %>%
+  summarize(freq = n()) %>%
+  group_by(site.date) %>%
+  mutate(bb.sp.count = length(unique(bb.sp)),
+         plant.sp.count = length(unique(plant.sp))) %>%
+  ungroup() %>%
+  filter(bb.sp.count > 1 & plant.sp.count > 1) %>% # the betalinkr function fails if you feed it webs with only 1 species in either (or perhaps both) of the levels; this is probably for good reason.
+  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.date, freq) %>% # rename columns to match bipartite's expectations
+  data.frame() %>% # convert to data frame
+  frame2webs(type.out = "array")
+}
+
+# Create webs
+net2010 <- web_prep(net, 2010)
+net2011 <- web_prep(net, 2011)
+net2012 <- web_prep(net, 2012)
+
+# Function with betalinkr settings and conversion to tibble()
+betalinkr_set1 <- function(x) {
+  betalinkr_multi(x, partition.st = TRUE, partition.rr = TRUE, # partition both species turnover and richness/replacement
+                  binary = TRUE, # treating beta-diversity as binary is an appropriate complement to our abundance analysis
+                  index = "jaccard") %>% # the Jaccard index is best, I think, for binary analysis
+  as_tibble() # for convenience
+}
+
+# Calculate beta-diversity
+beta_site.date.2010 <- betalinkr_set1(net2010)
+beta_site.date.2011 <- betalinkr_set1(net2011) 
+beta_site.date.2012 <- betalinkr_set1(net2012)
+
+# Collate results into a single tibble
+beta_site.date <- bind_rows(beta_site.date.2010, beta_site.date.2011, beta_site.date.2012)
+
+# Save to file to avoid redoing that long step
+write_csv(beta_site.date, "./output/beta_site.date.csv")
+beta_site.date <- read_csv("./output/beta_site.date.csv")
+
+# Add site data
+beta_mit_data <- beta_site.date %>%
+  separate(i, into = c("site1", "date1"), sep = "_") %>%
+  separate(j, into = c("site2", "date2"), sep = "_") %>%
+  mutate(date1 = ymd(date1),
+         date2 = ymd(date2),
+         year1 = factor(year(date1)),
+         year2 = factor(year(date2)),
+         yday1 = yday(date1),
+         yday2 = yday(date2)) %>%
   left_join(site_data, by = c("site1" = "site")) %>%
-  dplyr::select(site1, site2, S, OS, WN, ST, 
+  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
                 ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
-         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
+                elev1 = elev.mean, man1 = management, x1 = x, y1 = y) %>%
   left_join(site_data, by = c("site2" = "site")) %>%
-  dplyr::select(site1, site2, S, OS, WN, ST, 
+  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
                 ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
-         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
-  left_join(site_transects, by = c("site1" = "site")) %>%
-  dplyr::select(site1, site2, S, OS, WN, ST, 
-                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
-                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1 = transects) %>%
-  left_join(site_transects, by = c("site2" = "site")) %>%
-  dplyr::select(site1, site2, S, OS, WN, ST, 
-                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
-                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1, transects2 = transects) %>%
-  mutate(elev.diff = abs(elev1 - elev2),
+                elev1, elev2 = elev.mean, man1, man2 = management, x1, y1, x2 = x, y2 = y) %>%
+  mutate(site1 = factor(site1),
+         site2 = factor(site2),
+         year1 = factor(year1),
+         year2 = factor(year2),
+         yday1 = as.numeric(yday1),
+         yday2 = as.numeric(yday2),
+         elev.diff = abs(elev1 - elev2),
          mean.elev = (elev1 + elev2)/2,
-         transects.diff = abs(transects1 - transects2),
+         yday.diff = abs(yday1 - yday2),
          man.diff = case_when(
@@
          )) %>%
-  gather(metric, value, -c(site1, site2, elev1, elev2, elev.diff, 
-                           man1, man2, man.diff, mean.elev, lon1, 
-                           lat1, lon2, lat2, transects1, transects2, transects.diff)) %>%
+  gather(metric, value, -c(site1, year1, yday1, site2, year2, yday2,
+                           elev1, elev2, elev.diff, yday.diff, 
+                           man1, man2, man.diff, mean.elev, 
+                           x1, y1, x2, y2)) %>%
   mutate(metric.class = case_when(
@@
     metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
-    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
-  ))
-
-#write_csv(beta_site, "../analyses/output/beta_site.csv")
-
-beta_site.year <- betalinkr_multi(array_site.year, partition.st = TRUE, partition.rr = TRUE, binary = TRUE) %>%
-  as_tibble() %>%
-  separate(i, into = c("site1", "year1"), sep = "_") %>%
-  separate(j, into = c("site2", "year2"), sep = "_") %>%
-  left_join(site_data, by = c("site1" = "site")) %>%
-  dplyr::select(site1, year1, site2, year2, S, OS, WN, ST, 
-         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
-         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
-  left_join(site_data, by = c("site2" = "site")) %>%
-  dplyr::select(site1, year1, site2, year2, S, OS, WN, ST, 
-         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
-         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
-  left_join(site.year_transects, by = c("site1" = "site", "year1" = "year")) %>%
-  dplyr::select(site1, site2, year1, year2, S, OS, WN, ST, 
-                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
-                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1 = transects) %>%
-  left_join(site.year_transects, by = c("site2" = "site", "year2" = "year")) %>%
-  dplyr::select(site1, site2, year1, year2, S, OS, WN, ST, 
-                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
-                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1, transects2 = transects) %>%
-  mutate(elev.diff = abs(elev1 - elev2),
-         mean.elev = (elev1 + elev2)/2,
-         transects.diff = abs(transects1 - transects2),
-         man.diff = case_when(
-           man1 == man2 ~ FALSE,
-           man1 != man2 ~ TRUE
-         )) %>%
-  gather(metric, value, -c(site1, year1, site2, year2, elev1, elev2, elev.diff, 
-                           man1, man2, man.diff, mean.elev, lon1, lat1, lon2, lat2,
-                           transects1, transects2, transects.diff)) %>%
-  mutate(metric.class = case_when(
-    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
-    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
-    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
-    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
-  ))
-
-### By individual transect
-
-beta_site.date.2010 <- betalinkr_multi(array_site.date.2010, 
-                                       partition.st = TRUE, 
-                                       partition.rr = TRUE) %>%
-  as_tibble() %>%
-  separate(i, into = c("site1", "date1"), sep = "_") %>%
-  separate(j, into = c("site2", "date2"), sep = "_") %>%
-  left_join(site_data, by = c("site1" = "site"))%>%
-  dplyr::select(site1, date1, site2, date2, S, OS, WN, ST, 
-         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
-         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
-  left_join(site_data, by = c("site2" = "site")) %>%
-  dplyr::select(site1, date1, site2, date2, S, OS, WN, ST, 
-         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
-         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
-  mutate(elev.diff = abs(elev1 - elev2),
-         mean.elev = (elev1 + elev2)/2,
-         man.diff = case_when(
-           man1 == man2 ~ FALSE,
-           man1 != man2 ~ TRUE
-         )) %>%
-  gather(metric, value, -c(site1, date1, site2, date2, elev1, elev2, elev.diff, 
-                           man1, man2, man.diff, mean.elev, lon1, lat1, lon2, lat2)) %>%
-  mutate(metric.class = case_when(
-    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
-    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
-    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
-    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
-  ))
-
-beta_site.date.2011 <- betalinkr_multi(array_site.date.2011, 
-                                       partition.st = TRUE, 
-                                       partition.rr = TRUE)
-
-beta_site.date.2012 <- betalinkr_multi(array_site.date.2012, 
-                                       partition.st = TRUE, 
-                                       partition.rr = TRUE)
+    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"),
+    metric = factor(metric, levels = c("WN", "ST", "OS", "ST.h", "ST.l", 
+                                       "ST.lh", "WN.repl", "WN.rich", 
+                                       "OS.repl", "OS.rich"))
+    )
 ```
 
-# Beta diversity by elevation difference
+### Plot interaction beta-diversity against elevation difference and time difference
 ```{r}
-# Interaction beta diversity by elevation difference (by year)
-## This shows that the patterns are extremely consistent across years, which is not surprising, since they are driven overwhelmingly by floral turnover, and plants tend to be in the same place from year to year.
-ggplot(filter(beta_site.year, year1 == year2 & 
+# Set palette
+p <- c("OS" = "darkorange1", "ST" = "green3", 
+       "ST.h" = "firebrick1", "ST.l" = "darkorchid", 
+       "ST.lh" = "deepskyblue", "WN" = "black")
+
+# Elevation diff
+ggplot(filter(beta_mit_data, year1 == year2 &
+                site1 != site2 &
                 metric != "S" & 
-                metric.class != "OS.partition"), 
+                !metric.class %in% c("OS.partition", "WN.partition")), 
        aes(elev.diff, value, color = metric)) +
-  geom_point(alpha = 0.25) +
+  geom_point(alpha = 0.25, pch = 16) +
   geom_smooth() +
-  facet_grid(metric.class ~ year1)
+  facet_grid(metric.class ~ year1) +
+  scale_color_manual(values = p) +
+  theme_abyss()
 
-ggplot(filter(beta_site.year, year1 == year2 & 
+# Time diff
+ggplot(filter(beta_mit_data, year1 == year2 &
+                site1 != site2 &
                 metric != "S" & 
-                metric.class != "OS.partition"), 
-       aes(transects.diff, value, color = metric)) +
-  geom_point(alpha = 0.25) +
+                !metric.class %in% c("OS.partition", "WN.partition")), 
+       aes(yday.diff, value, color = metric)) +
+  geom_point(alpha = 0.25, pch = 16) +
   geom_smooth() +
-  facet_grid(metric.class ~ year1, scales = "free_x")
+  facet_grid(metric.class ~ year1) +
+  scale_color_manual(values = p) +
+  theme_abyss()
 
-# Interaction beta diversity by elevation difference, this time pooling across years
-ggplot(filter(beta_site, metric != "S" & 
-                metric.class != "OS.partition"), 
+# Elevation diff within 1 wk time diff
+ggplot(filter(beta_mit_data, year1 == year2 &
+                site1 != site2 &
+                metric != "S" & 
+                !metric.class %in% c("OS.partition", "WN.partition") &
+                yday.diff <= 7), 
        aes(elev.diff, value, color = metric)) +
-  geom_point(alpha = 0.05) +
+  geom_point(alpha = 0.1, pch = 16, size = 1) +
   geom_smooth() +
-  facet_wrap(~metric.class)
+  facet_grid(metric.class ~ year1) +
+  scale_color_manual(values = p, name = "partition") +
+  xlab("Elevation difference") +
+  ylab("β-diversity")
+ggsave("./output/beta_elev.png", height = 4.5, width = 6.5, units = "in")
 
-# How important is the difference in transects? Not as important as the difference in elevation.
-ggplot(filter(beta_site, metric != "S" & 
-                metric.class != "OS.partition"), 
-       aes(transects.diff, value, color = metric)) +
-  geom_point(alpha = 0.05) +
+# Time diff within 250 m elev diff
+ggplot(filter(beta_mit_data, year1 == year2 &
+                site1 != site2 &
+                metric != "S" & 
+                !metric.class %in% c("OS.partition", "WN.partition") &
+                elev.diff <= 250), 
+       aes(yday.diff, value, color = metric)) +
+  geom_point(alpha = 0.1, pch = 16, size = 1) +
   geom_smooth() +
-  facet_wrap(~metric.class)
-
-ggplot(beta_site, 
-       aes(elev.diff, transects.diff)) +
-  geom_point(alpha = 0.05) +
-  geom_smooth()
-
-### Okay, we'll just do partial mantel tests to partial out the effect of transects.diff; and we'll only do that for the aggregate measures because it seems a bit silly to doi it for the partitions  
-WN_diff <- beta_site %>%
-  filter(metric == "WN") %>%
-  dplyr::select(site1, site2, value) %>%
-  spread(site2, value) %>%
-  dist()
-
-WN_diff <- beta_site %>%
-  filter(metric == "WN") %>%
-  dplyr::select(site1, site2, value) %>%
-  spread(site2, value) %>%
-  dist()
-
-ST_diff <- beta_site %>%
-  filter(metric == "ST") %>%
-  dplyr::select(site1, site2, value) %>%
-  spread(site2, value) %>%
-  dist()
-
-OS_diff <- beta_site %>%
-  filter(metric == "OS") %>%
-  dplyr::select(site1, site2, value) %>%
-  spread(site2, value) %>%
-  dist()
-
-elev_diff <- beta_site %>%
-  dplyr::select(site1, site2, elev.diff) %>%
-  unique() %>%
-  spread(site2, elev.diff) %>%
-  dist()
-
-transects_diff <- beta_site %>%
-  dplyr::select(site1, site2, transects.diff) %>%
-  unique() %>%
-  spread(site2, transects.diff) %>%
-  dist()
-
-WN_mantel <- mantel.partial(elev_diff, WN_diff, transects_diff)
-ST_mantel <- mantel.partial(elev_diff, ST_diff, transects_diff)
-OS_mantel <- mantel.partial(elev_diff, OS_diff, transects_diff)
+  facet_grid(metric.class ~ year1) +
+  scale_color_manual(values = p, name = "partition") +
+  xlab("Elevation difference") +
+  ylab("β-diversity")
+ggsave("./output/beta_time.png", height = 4.5, width = 6.5, units = "in")
 ```
 
-### Let's look at floral survey diversity patterns
-```{r}
-survey_transects <- survey %>%
-  group_by(site) %>%
-  summarize(transects = length(unique(date)))
+### GDM: Interactions
+```{r message=FALSE, warning=FALSE}
+# Site_pair prep function
+site_pair_prep_net <- function(dat, y) {
+  beta_mit_data %>%
+    ungroup() %>%
+    filter(year1 == year2 & 
+             year1 == y & 
+             site1 != site2) %>% # I don't think we want to compare samples through within sites, introduces a new form of non-independence.
+    filter(metric == "WN") %>% # we'll focus for now on WN, but this could be done for partitions, too
+    mutate(weights = rep(1, n())) %>% # right now this analysis is unweighted; could weight by abundance, etc, if we wanted to
+    dplyr::select(distance = value, weights,
+                  s1.xCoord = x1, s1.yCoord = y1, 
+                  s2.xCoord = x2, s2.yCoord = y2, 
+                  s1.elev.mean = elev1, s1.yday = yday1,
+                  s2.elev.mean = elev2, s2.yday = yday2) %>%
+    data.frame()
+}
 
-survey_bioData <- survey %>%
-  filter(flower.cover > 0) %>%
-  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
-  group_by(site, plant.sp) %>%
-  summarize(cover = sum(flower.cover)) %>%
-  spread(plant.sp, cover) %>%
-  replace(is.na(.), 0)
+# Prep site-pairs for GDM analysis
+sitepair_2010.WN <- site_pair_prep_net(beta_mit_data, 2010)
+sitepair_2011.WN <- site_pair_prep_net(beta_mit_data, 2011)
+sitepair_2012.WN <- site_pair_prep_net(beta_mit_data, 2012)
 
-predData <- site_data %>%
-  ungroup() %>%
-  dplyr::select(site, elev.mean, lat, lon) %>%
-  unique()
-%>%
-  left_join(survey_transects)
+# Run GDM analysis
+gdm_WN.2010 <- gdm(sitepair_2010.WN, geo = TRUE)
+gdm_WN.2011 <- gdm(sitepair_2011.WN, geo = TRUE)
+gdm_WN.2012 <- gdm(sitepair_2012.WN, geo = TRUE)
 
-survey_sitepair <- formatsitepair(survey_bioData, bioFormat = 1, siteColumn = "site", dist = "jaccard",
-                              XColumn = "lon", YColumn = "lat", predData = predData)
+summary(gdm_WN.2010)
+summary(gdm_WN.2011)
+summary(gdm_WN.2012)
 
-### Call gdm()
-gdm_survey <- gdm(survey_sitepair, geo = TRUE)
-gdm.varImp(survey_sitepair, geo = TRUE)
+# Permutation tests (slow)
+gdm_WN.2010.imp <- gdm.varImp(sitepair_2010.WN, geo = TRUE)
+gdm_WN.2011.imp <- gdm.varImp(sitepair_2011.WN, geo = TRUE)
+gdm_WN.2012.imp <- gdm.varImp(sitepair_2012.WN, geo = TRUE)
 
-summary(gdm_survey)
-plot(gdm_survey)
-plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
+# Plot
+plot(gdm_WN.2010, include.rug = TRUE, 
+     rug.sitepair = sitepair_2010.WN, 
+     plot.layout = c(2,3))
+plot(gdm_WN.2011, include.rug = TRUE, 
+     rug.sitepair = sitepair_2011.WN, 
+     plot.layout = c(2,3))
+plot(gdm_WN.2012, include.rug = TRUE, 
+     rug.sitepair = sitepair_2012.WN, 
+     plot.layout = c(2,3))
+
+# Plot with confidence intervals (slow; also, I'm not sure about the current settings)
+plotUncertainty(sitepair_2010.WN, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
+plotUncertainty(sitepair_2011.WN, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
+plotUncertainty(sitepair_2012.WN, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
+
+# Plot nicely
+## wrapper for extracting splines and prepping tibble for plotting
+spline_prep <- function(x, y) {
+  isplineExtract(x) %>%
+    as.data.frame() %>%
+    tibble() %>%
+    mutate(x.Geographic = x.Geographic/1000) %>%
+    pivot_longer(cols = c(x.Geographic, x.elev.mean, x.yday), 
+                 names_to = "x.var", values_to = "predictor", 
+                 names_prefix = "x.") %>%
+    pivot_longer(cols = c(y.Geographic, y.elev.mean, y.yday), 
+                 names_to = "y.var", values_to = "response", 
+                 names_prefix = "y.") %>%
+    filter(x.var == y.var) %>%
+    mutate() %>%
+    dplyr::select(-y.var) %>%
+    mutate(year = factor(rep(y, n())))
+}
+
+## extract splines
+interaction_splines2010 <- spline_prep(gdm_WN.2010, 2010)
+interaction_splines2011 <- spline_prep(gdm_WN.2011, 2011)
+interaction_splines2012 <- spline_prep(gdm_WN.2012, 2012)
+
+## collate extracted splines  
+interaction_splines <- bind_rows(interaction_splines2010, 
+                                 interaction_splines2011, 
+                                 interaction_splines2012) %>%
+  mutate(x.var = case_when(
+    x.var == "elev.mean" ~ "Elevation (m)",
+    x.var == "Geographic" ~ "Geographic (km)",
+    x.var == "yday" ~ "Day of Year"),
+    x.var = factor(x.var, levels = c("Geographic (km)", "Elevation (m)", "Day of Year"))
+    )
+
+## plot
+ggplot(interaction_splines, aes(predictor, response)) +
+  geom_line() +
+  facet_grid(year ~ x.var, scales = "free_x") +
+  xlab(NULL) +
+  ylab("f(x)") +
+  theme_gray(12)
+ggsave("./output/interaction_GDM.png")
 ```
 
-### And now Bombus diversity patterns
+### GDM: Floral survey
 ```{r}
-BB_bioData <- net %>%
+# Create functions for prepping data
+biodata_prep_survey <- function(surv, y) {
+  surv %>%
+    mutate(year = year(date),
+           yday = yday(date)) %>%
+    filter(flower.cover > 0 & year == y) %>%
+    mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
+    group_by(site, yday, plant.sp) %>%
+    summarize(cover = sum(flower.cover)) %>%
+    mutate(cover = rep(1, n())) %>% # this makes the data binary, which I think is what I want here; otherwise, just comment out
+    spread(plant.sp, cover) %>%
+    replace(is.na(.), 0) %>%
+    arrange(site, yday) %>%
+    unite(site.day, site, yday, sep = "_")
+}
+
+predData_prep_survey <- function(surv, y) {
+  surv %>%
+    mutate(year = year(date),
+           yday = yday(date)) %>%
+    filter(flower.cover > 0 & year == y) %>%
+    dplyr::select(site, yday) %>%
+    left_join(site_data) %>%
+    ungroup() %>%
+    dplyr::select(site, elev.mean, y, x, yday) %>%
+    unique() %>%
+    arrange(site, yday) %>%
+    unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
+    dplyr::select(-site)
+}
+
+format_set <- function(bio, pred) {
+  formatsitepair(bio, bioFormat = 1, siteColumn = "site.day", 
+                 dist = "jaccard", XColumn = "x", YColumn = "y", 
+                 predData = pred, abundance = FALSE)
+}
+
+# Prep input for GDM
+survey_bioData.2010 <- biodata_prep_survey(survey, 2010)
+survey_predData.2010 <- predData_prep_survey(survey, 2010)
+survey_sitepair.2010 <- format_set(survey_bioData.2010, survey_predData.2010)
+
+survey_bioData.2011 <- biodata_prep_survey(survey, 2011)
+survey_predData.2011 <- predData_prep_survey(survey, 2011)
+survey_sitepair.2011 <- format_set(survey_bioData.2011, survey_predData.2011)
+
+survey_bioData.2012 <- biodata_prep_survey(survey, 2012)
+survey_predData.2012 <- predData_prep_survey(survey, 2012)
+survey_sitepair.2012 <- format_set(survey_bioData.2012, survey_predData.2012)
+
+# Call GDMs
+gdm_survey.2010 <- gdm(survey_sitepair.2010, geo = TRUE)
+gdm_survey.2011 <- gdm(survey_sitepair.2011, geo = TRUE)
+gdm_survey.2012 <- gdm(survey_sitepair.2012, geo = TRUE)
+
+summary(gdm_survey.2010)
+summary(gdm_survey.2011)
+summary(gdm_survey.2012)
+
+# Permutations tests (slow)
+gdm_survey.2010.imp <- gdm.varImp(survey_sitepair.2010, geo = TRUE)
+gdm_survey.2011.imp <- gdm.varImp(survey_sitepair.2011, geo = TRUE)
+gdm_survey.2012.imp <- gdm.varImp(survey_sitepair.2012, geo = TRUE)
+
+# Plot
+plot(gdm_survey.2010)
+plot(gdm_survey.2011)
+plot(gdm_survey.2012)
+
+# Plot with confidence intervals (slow)
+plotUncertainty(survey_sitepair.2010, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
+plotUncertainty(survey_sitepair.2011, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
+plotUncertainty(survey_sitepair.2012, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
+
+# Plot nicely
+## extract splines
+survey_splines2010 <- spline_prep(gdm_survey.2010, 2010)
+survey_splines2011 <- spline_prep(gdm_survey.2011, 2011)
+survey_splines2012 <- spline_prep(gdm_survey.2012, 2012)
+
+## collate extracted splines  
+survey_splines <- bind_rows(survey_splines2010,
+                            survey_splines2011, 
+                            survey_splines2012) %>%
+  mutate(x.var = case_when(
+    x.var == "elev.mean" ~ "Elevation (m)",
+    x.var == "Geographic" ~ "Geographic (km)",
+    x.var == "yday" ~ "Day of Year"),
+    x.var = factor(x.var, levels = c("Geographic (km)", "Elevation (m)", "Day of Year"))
+    )
+
+## plot
+ggplot(survey_splines, aes(predictor, response)) +
+  geom_line() +
+  facet_grid(year ~ x.var, scales = "free_x") +
+  xlab(NULL) +
+  ylab("f(x)") +
+  theme_gray(12)
+ggsave("./output/survey_GDM.png")
+```
+
+### GDM: BB observations
+```{r}
+# Create functions for prepping data
+biodata_prep_BB <- function(net, y) {
+  net %>%
+    filter(year == y) %>%
     group_by(site, date, bb.sp) %>%
     summarize(count = n()) %>%
-  group_by(site, date) %>%
-  mutate(prop = count/sum(count)) %>%
-  group_by(site, bb.sp) %>%
-  summarize(abund = sum(prop)) %>%
-  spread(bb.sp, abund) %>%
-  replace(is.na(.), 0)
+    mutate(count = rep(1, n())) %>% # this makes the data binary, which I think is what I want here; otherwise, just comment out
+    spread(bb.sp, count) %>%
+    replace(is.na(.), 0) %>%
+    mutate(yday = yday(date)) %>%
+    unite(site.day, site, yday, sep = "_", remove = "TRUE") %>%
+    ungroup() %>%
+    dplyr::select(-date)
+}
 
-predData <- site_data %>%
+predData_prep_BB <- function(net, y) {
+  net %>%
+    filter(year == y) %>%
+    mutate(yday = yday(date)) %>%
+    dplyr::select(site, yday) %>%
+    left_join(site_data) %>%
     ungroup() %>%
-  dplyr::select(site, elev.mean, lat, lon) %>%
-  unique()
+    dplyr::select(site, elev.mean, y, x, yday) %>%
+    unique() %>%
+    arrange(site, yday) %>%
+    unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
+    dplyr::select(-site)
+}
 
-bb_sitepair <- formatsitepair(BB_bioData, bioFormat = 1, siteColumn = "site", dist = "jaccard",
-                              XColumn = "lon", YColumn = "lat", predData = predData)
+# Prep input for GDM
+BB_bioData.2010 <- biodata_prep_BB(net, 2010)
+BB_predData.2010 <- predData_prep_BB(net, 2010)
+BB_sitepair.2010 <- format_set(BB_bioData.2010, BB_predData.2010)
 
-### Call gdm()
-gdm_bb <- gdm(bb_sitepair, geo = TRUE)
-gdm.varImp(bb_sitepair, geo = TRUE)
+BB_bioData.2011 <- biodata_prep_BB(net, 2011)
+BB_predData.2011 <- predData_prep_BB(net, 2011)
+BB_sitepair.2011 <- format_set(BB_bioData.2011, BB_predData.2011)
 
-summary(gdm_bb)
-plot(gdm_bb)
-plotUncertainty(bb_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
+BB_bioData.2012 <- biodata_prep_BB(net, 2012)
+BB_predData.2012 <- predData_prep_BB(net, 2012)
+BB_sitepair.2012 <- format_set(BB_bioData.2012, BB_predData.2012)
+
+# Call GDMs
+gdm_BB.2010 <- gdm(BB_sitepair.2010, geo = TRUE)
+gdm_BB.2011 <- gdm(BB_sitepair.2011, geo = TRUE)
+gdm_BB.2012 <- gdm(BB_sitepair.2012, geo = TRUE)
+
+summary(gdm_BB.2010)
+summary(gdm_BB.2011)
+summary(gdm_BB.2012)
+
+# Permutations tests (slow)
+gdm_BB.2010.imp <- gdm.varImp(BB_sitepair.2010, geo = TRUE, 
+                              parallel = TRUE, cores = 4,
+                              outFile = "./output/bb_permGDM_2010")
+gdm_BB.2011.imp <- gdm.varImp(BB_sitepair.2011, geo = TRUE, 
+                              parallel = TRUE, cores = 4,
+                              outFile = "./output/bb_permGDM_2011")
+gdm_BB.2012.imp <- gdm.varImp(BB_sitepair.2012, geo = TRUE, 
+                              parallel = TRUE, cores = 4,
+                              outFile = "./output/bb_permGDM_2012")
+
+# Plot
+plot(gdm_BB.2010)
+plot(gdm_BB.2011)
+plot(gdm_BB.2012)
+
+# Plot with confidence intervals (slow)
+plotUncertainty(BB_sitepair.2010, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
+plotUncertainty(BB_sitepair.2011, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
+plotUncertainty(BB_sitepair.2012, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
+
+# Plot nicely
+## extract splines
+bb_splines2010 <- spline_prep(gdm_BB.2010, 2010)
+bb_splines2011 <- spline_prep(gdm_BB.2011, 2011)
+bb_splines2012 <- spline_prep(gdm_BB.2012, 2012)
+
+## collate extracted splines  
+bb_splines <- bind_rows(bb_splines2010,
+                            bb_splines2011, 
+                            bb_splines2012) %>%
+  mutate(x.var = case_when(
+    x.var == "elev.mean" ~ "Elevation (m)",
+    x.var == "Geographic" ~ "Geographic (km)",
+    x.var == "yday" ~ "Day of Year"),
+    x.var = factor(x.var, levels = c("Geographic (km)", "Elevation (m)", "Day of Year"))
+    )
+
+## plot
+ggplot(bb_splines, aes(predictor, response)) +
+  geom_line() +
+  facet_grid(year ~ x.var, scales = "free_x") +
+  xlab(NULL) +
+  ylab("f(x)") +
+  theme_gray(12)
+ggsave("./output/bb_GDM.png")
 ```
-
-### Now let's look at interaction beta diversity
-```{r}
-net_sitepair <- beta_site %>%
-  filter(metric == "WN") %>%
-  mutate(weights = rep(1, n())) %>%
-  dplyr::select(distance = value, weights,
-         s1.xCoord = lon1, s1.yCoord = lat1, 
-         s2.xCoord = lon2, s2.yCoord = lat2, 
-         s1.elev.mean = elev1, s2.elev.mean = elev2) %>%
-  data.frame()
-
-### Call gdm()
-gdm_net <- gdm(net_sitepair, geo = TRUE)
-gdm.varImp(net_sitepair, geo = TRUE)
-
-summary(gdm_net)
-plot(gdm_net)
-plotUncertainty(net_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
-```

### 2020-09-17 analyses/beta_diversity.Rmd
refined network analysis

@@
 library(tidyverse)
-```
-
-### Climate data
-```{r echo=FALSE, message=FALSE, include=FALSE}
-gdd <- read_csv("../processed_data/climate.csv") %>%
-  mutate(year = factor(year),
-         yday = yday(date)) %>%
-  dplyr::select(site, year, date, yday, gdd.cum) %>%
-  semi_join(read_csv("../processed_data/network.csv"), by = "site")
 ```
@@
 # Permutation tests (slow)
-gdm_WN.2010.imp <- gdm.varImp(sitepair_2010.WN, geo = TRUE)
-gdm_WN.2011.imp <- gdm.varImp(sitepair_2011.WN, geo = TRUE)
-gdm_WN.2012.imp <- gdm.varImp(sitepair_2012.WN, geo = TRUE)
+gdm_WN.2010.imp <- gdm.varImp(sitepair_2010.WN, geo = TRUE, 
+                              parallel = TRUE, cores = 4,
+                              outFile = "interaction_permGDM_2010")
+gdm_WN.2011.imp <- gdm.varImp(sitepair_2011.WN, geo = TRUE, 
+                              parallel = TRUE, cores = 4,
+                              outFile = "interaction_permGDM_2011")
+gdm_WN.2012.imp <- gdm.varImp(sitepair_2012.WN, geo = TRUE, 
+                              parallel = TRUE, cores = 4,
+                              outFile = "interaction_permGDM_2012")
 
@@
                  dist = "jaccard", XColumn = "x", YColumn = "y", 
-                 predData = pred, abundance = FALSE)
+                 predData = pred, abundance = FALSE) %>%
+  filter(s1.elev.mean != s2.elev.mean) # this avoids comparing repeated measures of the same site (taking advantage of the fact that each site has a unique elevation in our dataset)
 }
@@
 # Permutations tests (slow)
-gdm_survey.2010.imp <- gdm.varImp(survey_sitepair.2010, geo = TRUE)
-gdm_survey.2011.imp <- gdm.varImp(survey_sitepair.2011, geo = TRUE)
-gdm_survey.2012.imp <- gdm.varImp(survey_sitepair.2012, geo = TRUE)
+gdm_survey.2010.imp <- gdm.varImp(survey_sitepair.2010, geo = TRUE, 
+                              parallel = TRUE, cores = 3,
+                              outFile = "survey_permGDM_2010")
+gdm_survey.2011.imp <- gdm.varImp(survey_sitepair.2011, geo = TRUE, 
+                              parallel = TRUE, cores = 3,
+                              outFile = "survey_permGDM_2011")
+gdm_survey.2012.imp <- gdm.varImp(survey_sitepair.2012, geo = TRUE, 
+                              parallel = TRUE, cores = 3,
+                              outFile = "survey_permGDM_2012")
 
@@
 
-### GDM: BB observations
+### GDM: BB observations; this won't work in it's present state because I cannot 
 ```{r}
@@
                               parallel = TRUE, cores = 4,
-                              outFile = "./output/bb_permGDM_2010")
+                              outFile = "bb_permGDM_2010")
 gdm_BB.2011.imp <- gdm.varImp(BB_sitepair.2011, geo = TRUE, 
                               parallel = TRUE, cores = 4,
-                              outFile = "./output/bb_permGDM_2011")
+                              outFile = "bb_permGDM_2011")
 gdm_BB.2012.imp <- gdm.varImp(BB_sitepair.2012, geo = TRUE, 
                               parallel = TRUE, cores = 4,
-                              outFile = "./output/bb_permGDM_2012")
+                              outFile = "bb_permGDM_2012")
 
@@
 # Plot with confidence intervals (slow)
-plotUncertainty(BB_sitepair.2010, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
+p <- plotUncertainty(BB_sitepair.2010, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
 plotUncertainty(BB_sitepair.2011, sampleSites = 0.7, bsIters = 10, geo  = TRUE)

### 2020-09-17 analyses/network_analysis.Rmd
refined network analysis

--- 2020-07-16 analyses/network_analysis.Rmd
@@
 library(mgcViz)
-library(lavaan)
+library(piecewiseSEM)
+library(GGally)
 library(tidyverse)
@@
       frame2webs() %>% # convert to to bipartite web
-      map(specieslevel, level = "higher", index = "d") %>%
+      #map(specieslevel, level = "higher", index = "d") %>%
+      map(specieslevel, level = "higher", index = c("d", "resource range")) %>%
       map(rownames_to_column) %>%
@@
 
+### Analyze sampling patterns
+```{r}
+samples <- net %>%
+  mutate(yday = yday(date)) %>%
+  group_by(site, year, yday, gdd.cum, elev.mean) %>%
+  summarize(obs = n())
+
+ggplot(samples, aes(yday, elev.mean, size = obs)) +
+  geom_point(alpha = 0.25) + 
+  facet_wrap(~year)
+
+ggplot(samples, aes(gdd.cum, elev.mean, size = obs)) +
+  geom_point(alpha = 0.25) + 
+  facet_wrap(~year)
+
+### Can we make an argument for salvaging individual years, or parts of year?
+#### 2011
+ggplot(filter(samples, year == 2011), aes(elev.mean, obs)) +
+  geom_point() +
+  geom_smooth()
+
+### How evenly is the presence of species distributed across samples
+bb_abund_sp <- net %>%
+  group_by(site, dayofyear, elev.mean, year, bb.sp) %>%
+  summarize(abund = n())
+
+ggplot(filter(bb_abund_sp, year == 2011), aes(dayofyear, elev.mean)) +
+  geom_point(aes(alpha = abund)) +
+  facet_wrap(~bb.sp)
+```
+
 ### Rarefied network analysis (not broken down by phase)
@@
   mutate(year = factor(rep("2012", n())))
-netmet_boot_2011 <- boot_net(net_2011, n = 4, iter = 40)  %>%
+netmet_boot_2011 <- boot_net(net_2011, n = 7, iter = 40)  %>% # Note that this results in site t5 being dropped, since it had only 4 transects.
+I think it worth dropping problematic t5 to improve overall quality
   mutate(year = factor(rep("2011", n())))
+# Due to the unevenness of temporal interval of sampling in 2010, I don't think my bootstrapping approach handles  the embuggerance sufficiently.
 netmet_boot_2010 <- boot_net(net_2010, n = 3, iter = 40) %>%
@@
   mutate(year = factor(rep("2012", n())))
-spmet_2011 <- boot_net_sp(net_2011, 4, 40) %>%
+spmet_2011 <- boot_net_sp(net_2011, 7, 40) %>%
   mutate(year = factor(rep("2011", n())))  
@@
   mutate(year = factor(rep("2010", n())))
+
+wma1_2010 <- filter(net, site == "wma1", year == "2010")
 
@@
   group_by(site, year) %>%
-  summarize(d = weighted.mean(d, bb.sp.abund, na.rm = TRUE)) %>%
+  #summarize(d = weighted.mean(d, bb.sp.abund, na.rm = TRUE)) %>%
+  summarize(resource.range = weighted.mean(resource.range, bb.sp.abund, na.rm = TRUE)) %>%
   ungroup()
@@
   left_join(site_data, by = "site") %>%
-  mutate(year = factor(year),
+  mutate(year = factor(year, levels = c("2010", "2011", "2012")),
          site = factor(site),
          management = factor(management))
-```
-
-# Let's inspect correlations within our predictors
-```{r}
-library(GGally)
-
-predictors <- netmet %>%
-  select(elev.mean, bb.rich, log.bb.abund, fl.rich, log.fl.abund)
-
-ggpairs(predictors)
 ```
@@
 
-A source of embuggerance that I have not yet dealt with is that when I bootstrap to equalize transects count, I do not do anything to equalize the temporal (either yday or GDD) interval spanned by the randomly drawn transects.
+#### A source of embuggerance that I have not yet dealt with is that when I bootstrap to equalize transects count, I do not do anything to equalize the temporal (either yday or GDD) interval spanned by the randomly drawn transects.
 This will, of course, average out as I increase iterations, so I guess the real problem is that the time interval spanned by the full set of transects for each site varies, so no matter how much bootstrapping I do, there will still be potential bias.
 This seems a fairly insurmountable limitation of the sampling structure of my data, so I will probably resort to just acknowledging it as a caveat of my analysis.
+
+#### The way I see it, there are two schools of thought that could guide this analysis.
+(1) If I am only interested in modeling elevational patterns in my response variable, then I should include only elevation and bb.transects in the model.
+(2) If my goal is to explain the response variable as best I can, then I should throw all my variables into the model, knowing that they may covary with elevation, or that the effects of elevation may be mediated by these other variables.
+The latter strategy would shed more light on what DOES influence the response variables, but it would obscure pruely elevational patterns.
+The former would reveal any elevation patterns that might exist, but it would probably not do as good a job of explaining the response variable.
+Let's try both and see what we learn.
+Maybe we'll end up turning to SEM in the end if I can figure it out.
+
+### Elevation-focused modeling
 ```{r}
-#### Network level
-##### niche overlap
-gam_niche.overlap <- gam(niche.overlap.HL ~ year + 
-                           s(site, bs = "re") +
-                           #management + # no effect
-                           s(elev.mean, k = 5) +
-                           #s(bb.rich, k = 5) + # too collinear with bb.abund
-                           s(fl.rich, k = 5) +
-                           #s(log.fl.abund, k = 5) + # no effect
-                           s(log.bb.abund, k = 5),
+#### Niche overlap
+gam_niche.overlap.1 <- gam(niche.overlap.HL ~ 
+                             s(year, bs = "re") + 
+                             s(bb.transects) + 
+                             s(elev.mean, by = year, k = 10),
                            select = TRUE,
@@
 
+check.gamViz(gam_niche.overlap.1)
+concurvity(gam_niche.overlap.1, full = TRUE)
+concurvity(gam_niche.overlap.1, full = FALSE)
+summary(gam_niche.overlap.1)
+print(plot(gam_niche.overlap.1, allTerms = TRUE), pages = 1)
+
+# bb.transects was zeroed out by select = TRUE, so this plot should reflect the model above perfectly
 ggplot(netmet, aes(elev.mean, niche.overlap.HL)) +
   geom_point() +
-  theme_modern()
+  geom_smooth(method = "gam", formula = y ~ s(x, bs = "tp", k = 10)) +
+  facet_wrap(~year)
 
-check.gamViz(gam_niche.overlap)
-concurvity(gam_niche.overlap, full = TRUE)
-concurvity(gam_niche.overlap, full = FALSE)
-summary(gam_niche.overlap)
-print(plot(gam_niche.overlap, allTerms = TRUE), pages = 1)
-ggpredict(gam_niche.overlap)
-check_model(gam_niche.overlap)
-
-##### C.score
-gam_c.score <- gam(C.score.HL ~ year +
-                     s(site, bs = "re") +
-                     management +
-                     s(elev.mean, k = 5) +
-                     s(bb.rich, k = 5) +
-                     s(fl.rich, k = 5) +
-                     s(log.fl.abund, k = 5) +
-                     s(log.bb.abund, k = 5),
+##### Community mean d'
+gam_mean_d.1 <- gam(d ~ 
+                      s(year, bs = "re") + 
+                      s(bb.transects) +
+                      s(elev.mean, by = year, k = 10),
                     select = TRUE,
@@
 
-check.gamViz(gam_c.score)
-concurvity(gam_c.score, full = TRUE)
-concurvity(gam_c.score, full = FALSE)
-summary(gam_c.score)
-print(plot(gam_c.score, allTerms = TRUE), pages = 1)
+check.gamViz(gam_mean_d.1)
+concurvity(gam_mean_d.1, full = TRUE)
+concurvity(gam_mean_d.1, full = FALSE)
+summary(gam_mean_d.1)
+print(plot(gam_mean_d.1, allTerms = TRUE), pages = 1)
 
-##### weighted NODF: the main takeaway here is that higher BB abundance = higher nestedness, and that could easily be a sampling effect more than an ecologicla effect.
-Nevertheless, there remains a significant linear partial effect of elevation.
-It is worth noting that where elevation has an effect on network structure, it is linear after accounting for the influence of other variables.
-gam_weighted.NODF <- gam(weighted.NODF ~ year + 
-                           s(site, bs = "re") +
-                           #management +
-                           s(elev.mean, k = 5) +
-                           #s(bb.rich, k = 5) +
-                           s(fl.rich, k = 5) +
-                           #s(log.fl.abund, k = 5) +
-                           s(log.bb.abund, k = 5),
+ggplot(netmet, aes(elev.mean, d)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, bs = "tp", k = 10)) +
+  facet_wrap(~year)
+
+ggplot(netmet, aes(bb.transects, d)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~year, scales = "free_x")
+
+# I would be reluctant to interpret the year effect biologically; more likely an artifact of sampling intensity
+ggplot(netmet, aes(year, d)) +
+  geom_boxplot()
+
+# It is interesting that, while elevation explains a lot of niche overlap and virtually none of specialization, niche overlap and specialization are strongly correlated with each other.
+ggplot(netmet, aes(niche.overlap.HL, d)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~year)
+
+##### Resource range
+gam_mean_rr.1 <- gam(resource.range ~ 
+                      s(year, bs = "re") + 
+                      s(bb.transects) +
+                      s(elev.mean, by = year, k = 10),
                     select = TRUE,
@@
 
-check.gamViz(gam_weighted.NODF)
-concurvity(gam_weighted.NODF, full = TRUE)
-concurvity(gam_weighted.NODF, full = FALSE)
-summary(gam_weighted.NODF)
-print(plot(gam_weighted.NODF, allTerms = TRUE), pages = 1)
-ggpredict(gam_weighted.NODF)
-check_model(gam_weighted.NODF)
+check.gamViz(gam_mean_rr.1)
+concurvity(gam_mean_rr.1, full = TRUE)
+concurvity(gam_mean_rr.1, full = FALSE)
+summary(gam_mean_rr.1)
+print(plot(gam_mean_rr.1, allTerms = TRUE), pages = 1)
 
-##### togetherness
-gam_togetherness <- gam(togetherness.HL ~ year + 
-                           s(site, bs = "re") +
+ggplot(netmet, aes(elev.mean, resource.range)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, bs = "tp", k = 10)) +
+  facet_wrap(~year)
+
+ggplot(spmet, aes(elev.mean, resource.range, color = year)) +
+  geom_point() +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
+  facet_wrap(~bb.sp, scales = "free")
+
+## Species-level resource range
+gam_rr.sp_I <- gam(resource.range ~ 
+                    s(year, bs = "re") + 
+                    #s(site, bs = "re") +
                     #management +
-                           s(elev.mean, k = 5) +
+                    s(bb.transects) +
+                    s(bb.sp, bs = "re") +
+                    s(elev.mean, by = bb.sp, bs = "tp", k = 5),
+                  #s(fl.rich, k = 8) +
+                  #s(log.fl.abund, k = 8) +
+                  #s(bb.sp.abund, k = 5),
+                  select = TRUE,
+                  data = spmet,
+                  method = "REML") %>% getViz()
+
+check.gamViz(gam_rr.sp_I)
+concurvity(gam_rr.sp_I, full = TRUE)
+concurvity(gam_rr.sp_I, full = FALSE)
+summary(gam_rr.sp_I)
+print(plot(gam_rr.sp_I, allTerms = TRUE), pages = 4)
+
+```
+
+### Let's inspect correlations within our predictors
+#### Collinearity among my predictors will be unavoidable, but I can try to mitigate the situation.
+Bumble bee richness and abundance are almost perfectly correlated, so I definitely don't want to include them both in the same model.
+Floral abundance and richness are not significantly correlated (interestingly), but each is correlated with both bumble bee richness and bumble bee abundance.
+So, I think the goal is to pick the combination that minimizes collinearity.
+The weakest correlation between a bumble bee variable and a floral variable is between bumble bee richness and floral abundance, though the correlation is still pretty strong.
+Thus, I think an argument can be made for including only bumble bee richness and floral abundance in my models.
+```{r message=FALSE}
+predictors <- netmet %>%
+  select(elev.mean, bb.rich, log.bb.abund, 
+         fl.rich, log.fl.abund, bb.transects, year)
+
+ggpairs(predictors)
+
+ggplot(predictors, aes(log.fl.abund, bb.rich)) +
+  geom_point() +
+  facet_wrap(~year)
+
+ggplot(predictors, aes(log.fl.abund, log.bb.abund)) +
+  geom_point() +
+  facet_wrap(~year)
+
+ggplot(predictors, aes(fl.rich, log.bb.abund)) +
+  geom_point() +
+  facet_wrap(~year)
+
+ggplot(predictors, aes(fl.rich, bb.rich)) +
+  geom_point() +
+  facet_wrap(~year)
+```
+
+### Response-focused modeling
+```{r}
+#### Niche overlap
+gam_niche.overlap.2 <- gam(niche.overlap.HL ~ 
+                             s(year, bs = "re") + 
+                             #s(management, bs = "re") +
+                             s(bb.transects) + 
+                             s(elev.mean, by = year, k = 10) +
                              s(bb.rich, k = 5) + 
-                           #s(log.fl.abund, k = 5) +
-                           #s(log.bb.abund, k = 5) +
-                           s(fl.rich, k = 5),
+                             s(log.fl.abund, k = 5),  # no effect
                            select = TRUE,
@@
 
-check.gamViz(gam_togetherness)
-concurvity(gam_togetherness, full = TRUE)
-concurvity(gam_togetherness, full = FALSE)
-summary(gam_togetherness)
-print(plot(gam_togetherness, allTerms = TRUE), pages = 1)
-ggpredict(gam_togetherness)
-check_model(gam_togetherness)
+check.gamViz(gam_niche.overlap.2)
+concurvity(gam_niche.overlap.2, full = TRUE)
+concurvity(gam_niche.overlap.2, full = FALSE)
+summary(gam_niche.overlap.2)
+print(plot(gam_niche.overlap.2, allTerms = TRUE), pages = 1)
 
-##### generality; this could be an interesting alternative to something like d', but I'm not sure whether it deals with the confunding effects of abundance -- singletons, in extreme cases.
-gam_generality <- gam(generality.HL ~ year + 
-                           s(site, bs = "re") +
-                           management +
-                           s(elev.mean, k = 5) +
+ggplot(netmet, aes(elev.mean, niche.overlap.HL)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~year)
+
+ggplot(netmet, aes(bb.transects, niche.overlap.HL)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~year, scales = "free_x")
+
+ggplot(netmet, aes(elev.mean, bb.transects)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~year)
+
+#### Community mean d'
+gam_mean_d.2 <- gam(d ~ 
+                      s(year, bs = "re") + 
+                      s(management, bs = "re") +
+                      s(bb.transects) +
+                      s(elev.mean, by = year, k = 10) +
                       s(bb.rich, k = 5) +
-                           s(fl.rich, k = 5) +
-                           s(log.fl.abund, k = 5) +
-                           s(log.bb.abund, k = 5),
+                      s(log.fl.abund, k = 5),
                     select = TRUE,
@@
 
-check.gamViz(gam_generality)
-concurvity(gam_generality, full = TRUE)
-concurvity(gam_generality, full = FALSE)
-summary(gam_generality)
-print(plot(gam_generality, allTerms = TRUE), pages = 1)
-ggpredict(gam_generality)
-check_model(gam_generality)
+check.gamViz(gam_mean_d.2)
+concurvity(gam_mean_d.2, full = TRUE)
+concurvity(gam_mean_d.2, full = FALSE)
+summary(gam_mean_d.2)
+print(plot(gam_mean_d.2, allTerms = TRUE), pages = 1)
 
-##### community mean d'
-gam_mean_d <- gam(d ~ year + 
-                           s(site, bs = "re") +
-                           #management +
-                           #s(elev.mean, k = 5) +
-                           #s(bb.rich, k = 5) +
-                           #s(log.fl.abund, k = 5) +
-                           #s(fl.rich, k = 5) +
-                           s(log.bb.abund, k = 5),
-                          select = TRUE,
-                          data = netmet,
-                          method = "REML") %>% getViz()
-
-check.gamViz(gam_mean_d)
-concurvity(gam_mean_d, full = TRUE)
-concurvity(gam_mean_d, full = FALSE)
-summary(gam_mean_d)
-print(plot(gam_mean_d, allTerms = TRUE), pages = 1)
-ggpredict(gam_mean_d)
-check_model(gam_mean_d)
-
-ggplot(netmet, aes(log.bb.abund, d, color = year)) +
+ggplot(netmet, aes(bb.rich, d)) +
   geom_point() +
   geom_smooth(method = "lm") +
-  theme_abyss()
+  facet_wrap(~year)
 
-ggplot(netmet, aes(elev.mean, d, color = year)) +
+ggplot(netmet, aes(log.bb.abund, d)) +
   geom_point() +
-  geom_smooth() +
-  theme_abyss()
+  geom_smooth(method = "lm") +
+  facet_wrap(~year)
+
+ggplot(netmet, aes(log.fl.abund, d)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~year)
+
+ggplot(netmet, aes(fl.rich, d)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~year)
+```
+
+### But shouldn't I be using the abundance data from the floral survey instead of the marginal totals?
+Yes.
+But no.
+In many cases, the species quantified during floral surveying do not include all of the species observed in the network.
+In those cases, I would have to either make up abundance data or drop floral species from the network, and both of those seem like bad ideas.
+Perhaps this is an argument for not using d'?
+If marginal totals are bad measures of abundance -- and I believe they are -- then maybe we need to either abandon the idea of analyzing specialization or opt for a simpler specialization metric, like PDI
+```{r}
+test <- net %>%
+      # semi_join(t_filter) %>%
+      # dplyr::select(site, date, year) %>%
+      # unique() %>%
+      # group_by(site, year) %>%
+      # sample_n(n) %>%
+      # left_join(net) %>%
+      group_by(site, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
+      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
+      unite(webID, c(site, year), sep = "_") %>%
+      ungroup() %>%
+      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
+      data.frame()
+
+survey_sum <- survey %>%
+  group_by(site, date, plant.sp) %>%
+  summarize(raw.abund = sum(flower.cover)) %>%
+  group_by(site, date) %>%
+  mutate(prop.abund = raw.abund/sum(raw.abund),
+         year = year(date)) %>%
+  group_by(site, year, plant.sp) %>%
+  summarize(mean.prop.abund = mean(prop.abund)) %>%
+  unite(webID, c(site, year), sep = "_") %>%
+  rename(lower = plant.sp) %>%
+  semi_join(test)
+
+webs <- test %>% # convert to data frame
+      frame2webs()
+
+h1_2010 <- specieslevel(webs[[1]], level = "higher", index = "d", low.abun = filter(survey_sum, webID == "h1_2010")$mean.prop.abund)
+
+h1_2011 <- specieslevel(webs[[2]], level = "higher", index = "d", low.abun = filter(survey_sum, webID == "h1_2011")$mean.prop.abund)
+
+h1_2012 <- specieslevel(webs[[2]], level = "higher", index = "d", low.abun = filter(survey_sum, webID == "h1_2012")$mean.prop.abund)
+
+%>% # convert to to bipartite web
+      map(specieslevel, level = "higher", index = "d") %>%
+      map(rownames_to_column) %>%
+      map(as_tibble) %>%
+      bind_rows(.id = "site_year") %>% # Collapse list into big data frame
+      rename(bb.sp = rowname)
+
+ggplot(test, aes(bb.sp, d)) +
+  geom_boxplot()
 ```
@@
 ```{r}
-library(piecewiseSEM)
-
+# Create scales variables; this is important since SEM is based on covariance (right?)
+# For starters, let's just try to model niche.overlap
 netmet.scaled <- netmet %>%
-  select(site, year, niche.overlap.HL, elev.mean, fl.rich, bb.rich, log.fl.abund, log.bb.abund) %>%
-  mutate(niche.overlap.HL = scale(niche.overlap.HL),
-         elev.mean = scale(elev.mean),
-         fl.rich = scale(fl.rich),
-         log.fl.abund = scale(log.fl.abund),
-         log.bb.abund = scale(log.bb.abund)) %>%
-  as.matrix() %>%
-  as_tibble() %>%
-  mutate(niche.overlap.HL = as.numeric(niche.overlap.HL),
-         elev.mean = as.numeric(elev.mean),
-         fl.rich = as.numeric(fl.rich),
-         log.fl.abund = as.numeric(log.fl.abund),
-         log.bb.abund = as.numeric(log.bb.abund),
-         site = factor(site),
-         year = factor(year)) %>%
+  select(site, year, niche.overlap.HL, elev.mean, 
+         fl.rich, bb.rich, log.fl.abund, log.bb.abund) %>%
+  mutate(elev.mean.2 = elev.mean^2,
+         elev.mean.3 = elev.mean^3) %>%
+  # mutate_at(c("niche.overlap.HL", "elev.mean", "elev.mean.2", "elev.mean.3", 
+  #        "fl.rich", "bb.rich", "log.fl.abund", "log.bb.abund"), ~(scale(.) %>% as.vector)) %>% # I did not know you could put pipes inside a mutate call -- good to know: https://stackoverflow.com/questions/15215457/standardize-data-columns-in-r
   na.omit()
 
-mod_niche.overlap <- psem(lmer(niche.overlap.HL ~ elev.mean + 
-                                fl.rich + log.bb.abund + 
-                                 (1 | year) + (1 | site), 
-                              netmet.scaled),
-                          lm(fl.rich ~ elev.mean, 
-                             netmet.scaled),
-                          lm(log.bb.abund ~ elev.mean, 
-                             netmet.scaled))
-summary(mod_niche.overlap)
-plot(mod_niche.overlap)
+# Pairs plot to determine best model form
+predictors.scaled <- netmet.scaled %>%
+  select(elev.mean, elev.mean.2, elev.mean.3, bb.rich, 
+         log.bb.abund, fl.rich, log.fl.abund, niche.overlap.HL)
+
+ggpairs(predictors.scaled)
+
+# Diversity and abundance metrics clearly have a quadratic relationship with elevation
+
+# Call models
+# To avoid figuring out how to handle year and site as random effects, let's start with separate year models
+netmet.scaled.2010 <- filter(netmet.scaled, year == 2010)
+netmet.scaled.2011 <- filter(netmet.scaled, year == 2011)
+netmet.scaled.2012 <- filter(netmet.scaled, year == 2012)
+
+mod_niche.overlap.2010 <- psem(
+  lm(niche.overlap.HL ~ elev.mean + log.fl.abund + bb.rich, 
+     netmet.scaled.2010),
+  lm(log.fl.abund ~ elev.mean, 
+     netmet.scaled.2010),
+  lm(bb.rich ~ elev.mean, 
+     netmet.scaled.2010)
+)
+
+summary(mod_niche.overlap.2010)
+plot(mod_niche.overlap.2010)
+
+mod_niche.overlap.2011 <- psem(
+  lm(niche.overlap.HL ~ elev.mean + log.fl.abund + bb.rich, 
+     netmet.scaled.2011),
+  lm(log.fl.abund ~ elev.mean, 
+     netmet.scaled.2011),
+  lm(bb.rich ~ elev.mean, 
+     netmet.scaled.2011)
+)
+
+summary(mod_niche.overlap.2011)
+plot(mod_niche.overlap.2011)
+
+mod_niche.overlap.2012 <- psem(
+  lm(niche.overlap.HL ~ elev.mean + log.fl.abund + bb.rich, 
+     netmet.scaled.2012),
+  lm(log.fl.abund ~ elev.mean, 
+     netmet.scaled.2012))
+)
+
+summary(mod_niche.overlap.2012)
+plot(mod_niche.overlap.2012)
 ```
-
-### But let's zoom in a bit closer on d' and see if we can model it per-species
-```{r}
-gam_d.sp_GI <- gam(d ~ year + 
-                     s(site, bs = "re") +
-                     management +
-                     s(bb.sp, bs = "re") +
-                     s(elev.mean, bs = "tp", k = 8) +
-                     s(elev.mean, by = bb.sp, bs = "tp", k = 8) +
-                     s(fl.rich, k = 8) +
-                     s(log.fl.abund, k = 8) +
-                     s(log.bb.sp.abund, k = 5),
-                   select = TRUE,
-                   data = spmet,
-                   method = "REML") %>% getViz()
-
-check.gamViz(gam_d.sp_GI)
-concurvity(gam_d.sp_GI, full = TRUE)
-concurvity(gam_d.sp_GI, full = FALSE)
-summary(gam_d.sp_GI)
-print(plot(gam_d.sp_GI, allTerms = TRUE), pages = 4)
-
-gam_d.sp_G <- gam(d ~ year + 
-                    s(site, bs = "re") +
-                    management +
-                    s(bb.sp, bs = "re") +
-                    s(elev.mean, bs = "tp", k = 8) +
-                    s(fl.rich, k = 8) +
-                    s(log.fl.abund, k = 8) +
-                    s(log.bb.sp.abund, k = 5),
-                  select = TRUE,
-                  data = spmet,
-                  method = "REML") %>% getViz()
-
-check.gamViz(gam_d.sp_G)
-concurvity(gam_d.sp_G, full = TRUE)
-concurvity(gam_d.sp_G, full = FALSE)
-summary(gam_d.sp_G)
-print(plot(gam_d.sp_G, allTerms = TRUE), pages = 1)
-check_model(gam_d.sp_G)
-modelG_elev.mean <- ggpredict(gam_d.sp_G, terms = "elev.mean")
-plot(modelG_elev.mean)
-
-# This simpler varying-smooth model explains just as much variance as the more complex GI model, so it should be preferred. 
-gam_d.sp_GS <- gam(d ~ year + 
-                     management +
-                     s(site, bs = "re") +
-                     s(elev.mean, bs = "tp", k = 8) +
-                     s(elev.mean, bb.sp, bs = "fs", k = 8) +
-                     s(log.fl.abund, k = 8) +
-                     s(fl.rich, k = 8) +
-                     s(log.bb.sp.abund, k = 5),
-                   select = TRUE,
-                   data = spmet,
-                   method = "REML") %>% getViz()
-
-check.gamViz(gam_d.sp_GS)
-concurvity(gam_d.sp_GS, full = TRUE)
-concurvity(gam_d.sp_GS, full = FALSE)
-summary(gam_d.sp_GS)
-print(plot(gam_d.sp_GS, allTerms = TRUE), pages = 1)
-check_model(gam_d.sp_GS)
-plot(ggpredict(gam_d.sp_GS))
-
-# What about I? Does it do any better than S?
-gam_d.sp_I <- gam(d ~ year + 
-                     s(site, bs = "re") +
-                     management +
-                     s(bb.sp, bs = "re") +
-                     s(elev.mean, by = bb.sp, bs = "tp", k = 8) +
-                     s(fl.rich, k = 8) +
-                     s(log.fl.abund, k = 8) +
-                     s(log.bb.sp.abund, k = 5),
-                   select = TRUE,
-                   data = spmet,
-                   method = "REML") %>% getViz()
-
-check.gamViz(gam_d.sp_I)
-concurvity(gam_d.sp_I, full = TRUE)
-concurvity(gam_d.sp_I, full = FALSE)
-summary(gam_d.sp_I)
-print(plot(gam_d.sp_I, allTerms = TRUE), pages = 4)
-
-AIC(gam_d.sp_I, gam_d.sp_S, gam_d.sp_G, gam_d.sp_GI, gam_d.sp_GS)
-
-ggplot(spmet, aes(log.bb.sp.abund)) +
-  geom_histogram() +
-  theme_modern()
-
-ggplot(spmet, aes(log.bb.sp.abund)) +
-  geom_histogram() +
-  theme_blackboard()
-
-ggplot(spmet, aes(elev.mean, d)) +
-  geom_point(aes(color = bb.sp)) +
-  geom_smooth() +
-  theme_blackboard() +
-  facet_wrap(~year)
-
-# We lose virtually no explanatory power in going from GI, to GS, to S. So, S is the preferred model right now.
-gam_d.sp_S <- gam(d ~ year + 
-                    s(site, bs = "re") +
-                    management +
-                    s(elev.mean, bb.sp, bs = "fs", k = 8) +
-                    s(fl.rich, k = 8) +
-                    s(log.fl.abund, k = 8) +
-                    s(log.bb.sp.abund, k = 5),
-                  select = TRUE,
-                  data = spmet,
-                  method = "REML") %>% getViz()
-
-check.gamViz(gam_d.sp_S)
-concurvity(gam_d.sp_S, full = TRUE)
-concurvity(gam_d.sp_S, full = FALSE)
-summary(gam_d.sp_S)
-print(plot(gam_d.sp_S, allTerms = TRUE), pages = 1)
-
-###  There remains a lot of noise in my multi-year models above, and a possible explanation for that is that the influence of year extends beyond just variation in the intercept.
-What if year affects the shape of Y ~ X?
-In these year subset models, I can drop both 'year' and 's(site, bs = "re")'.
-
-gam_d.sp_GI.2010 <- gam(d ~ management +
-                          s(bb.sp, bs = "re") +
-                          s(elev.mean, bs = "tp", k = 8) +
-                          s(elev.mean, by = bb.sp, bs = "tp", k = 8) +
-                          s(fl.rich, k = 8) +
-                          s(log.fl.abund, k = 8) +
-                          s(log.bb.sp.abund, k = 5),
-                        select = TRUE,
-                        data = filter(spmet, year == 2010),
-                        method = "REML") %>% getViz()
-
-check.gamViz(gam_d.sp_GI.2010)
-concurvity(gam_d.sp_GI.2010, full = TRUE)
-concurvity(gam_d.sp_GI.2010, full = FALSE)
-summary(gam_d.sp_GI.2010)
-print(plot(gam_d.sp_GI.2010, allTerms = TRUE), pages = 4)
-
-gam_d.sp_GI.2011 <- gam(d ~ management +
-                          s(bb.sp, bs = "re") +
-                          s(elev.mean, bs = "tp", k = 8) +
-                          s(elev.mean, by = bb.sp, bs = "tp", k = 8) +
-                          s(fl.rich, k = 8) +
-                          s(log.fl.abund, k = 8) +
-                          s(log.bb.sp.abund, k = 5),
-                        select = TRUE,
-                        data = filter(spmet, year == 2011),
-                        method = "REML") %>% getViz()
-
-check.gamViz(gam_d.sp_GI.2011)
-concurvity(gam_d.sp_GI.2011, full = TRUE)
-concurvity(gam_d.sp_GI.2011, full = FALSE)
-summary(gam_d.sp_GI.2011)
-print(plot(gam_d.sp_GI.2011, allTerms = TRUE), pages = 4)
-
-gam_d.sp_GI.2012 <- gam(d ~ management +
-                          s(bb.sp, bs = "re") +
-                          s(elev.mean, bs = "tp", k = 8) +
-                          s(elev.mean, by = bb.sp, bs = "tp", k = 8) +
-                          s(fl.rich, k = 8) +
-                          s(log.fl.abund, k = 8) +
-                          s(log.bb.sp.abund, k = 5),
-                        select = TRUE,
-                        data = filter(spmet, year == 2012),
-                        method = "REML") %>% getViz()
-
-check.gamViz(gam_d.sp_GI.2012)
-concurvity(gam_d.sp_GI.2012, full = TRUE)
-concurvity(gam_d.sp_GI.2012, full = FALSE)
-summary(gam_d.sp_GI.2012)
-print(plot(gam_d.sp_GI.2012, allTerms = TRUE), pages = 4)
-
-### I-form models
-gam_d.sp_I.2010 <- gam(d ~ management +
-                          s(bb.sp, bs = "re") +
-                          s(elev.mean, by = bb.sp, bs = "tp", k = 8) +
-                          s(fl.rich, k = 8) +
-                          s(log.fl.abund, k = 8) +
-                          s(log.bb.sp.abund, k = 5),
-                        select = TRUE,
-                        data = filter(spmet, year == 2010),
-                        method = "REML") %>% getViz()
-
-check.gamViz(gam_d.sp_I.2010)
-concurvity(gam_d.sp_I.2010, full = TRUE)
-concurvity(gam_d.sp_I.2010, full = FALSE)
-summary(gam_d.sp_I.2010)
-print(plot(gam_d.sp_I.2010, allTerms = TRUE), pages = 4)
-
-gam_d.sp_I.2011 <- gam(d ~ management +
-                          s(bb.sp, bs = "re") +
-                          s(elev.mean, by = bb.sp, bs = "tp", k = 8) +
-                          s(fl.rich, k = 8) +
-                          s(log.fl.abund, k = 8) +
-                          s(log.bb.sp.abund, k = 5),
-                        select = TRUE,
-                        data = filter(spmet, year == 2011),
-                        method = "REML") %>% getViz()
-
-check.gamViz(gam_d.sp_I.2011)
-concurvity(gam_d.sp_I.2011, full = TRUE)
-concurvity(gam_d.sp_I.2011, full = FALSE)
-summary(gam_d.sp_I.2011)
-print(plot(gam_d.sp_I.2011, allTerms = TRUE), pages = 4)
-
-gam_d.sp_I.2012 <- gam(d ~ management +
-                          s(bb.sp, bs = "re") +
-                          s(elev.mean, by = bb.sp, bs = "tp", k = 8) +
-                          s(fl.rich, k = 8) +
-                          s(log.fl.abund, k = 8) +
-                          s(log.bb.sp.abund, k = 5),
-                        select = TRUE,
-                        data = filter(spmet, year == 2012),
-                        method = "REML") %>% getViz()
-
-check.gamViz(gam_d.sp_I.2012)
-concurvity(gam_d.sp_I.2012, full = TRUE)
-concurvity(gam_d.sp_I.2012, full = FALSE)
-summary(gam_d.sp_I.2012)
-print(plot(gam_d.sp_I.2012, allTerms = TRUE), pages = 4)
-```
-
-### Does tongue-length / corolla depth matching vary with elevation?
-```{r}
-bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
-  dplyr::select(subgenus, bb.sp, pbl.w, pbl.w.class)
-
-fl_ktype <- read.csv("../processed_data/floral_k_type_nom_fix.csv") %>%
-  dplyr::select(plant.sp, k.type) %>%
-  mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"),
-         k.type.ss = str_extract(k.type, "[-+]?[0-9]*"))
-
-fl_morph <- read.csv("../processed_data/flower_morphology_GitaBenadi.csv") %>%
-  rename(plant.sp = Species) %>%
-  group_by(plant.sp) %>%
-  summarize(corolla.depth = mean(LN))
-
-fl_tax <- read_csv("../processed_data/floral_tax.csv")
-
-net_traits <- net %>%
-  # group_by(site, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-  # summarize(freq = n()) %>%
-  # ungroup() %>%
-  left_join(bb_traits) %>%
-  left_join(fl_ktype) %>%
-  left_join(fl_morph) %>%
-  separate(plant.sp, c("plant.genus", "plant.species"), sep = " ", remove = FALSE) %>%
-  left_join(fl_tax) %>%
-  mutate(tongue.mismatch = abs(pbl.w - corolla.depth),
-         bb.sp = factor(bb.sp),
-         plant.sp = factor(plant.sp),
-         plant.genus = factor(plant.genus),
-         subgenus = factor(subgenus),
-         k.type.s = factor(k.type.s),
-         plant.family = factor(plant.family))
-
-ggplot(net_traits, aes(elev.mean, tongue.mismatch)) +
-  geom_point(color = "yellow") +
-  geom_smooth(method = "lm") +
-  facet_wrap(~bb.sp) +
-  theme_abyss()
-```

### 2020-09-17 analyses/trash_can.R

@@ -1343,3 +1343,404 @@ check.gamViz(weighted_fl_abundance.2011.6)
 weighted_fl_abundance.2011.ck.6 <- check2D(weighted_fl_abundance.2011.6, 
                                            x1 = "yday", x2 = "elev.mean")
 ```
+
+### Since we clearly have structured collinearity in our preditors, i.e. elevation shapes abundance and diversity which shape network structure, perhaps SEM could be appropriate? The main question is whether it can handle the potentially complex (nonlinear) relationships between elevation and diversity/abundance, and whether it can handle the hierarchical structure of my data.
+```{r}
+# Create scales variables; this is important since SEM is based on covariance (right?)
+# For starters, let's just try to model niche.overlap
+netmet.scaled <- netmet %>%
+  select(site, year, niche.overlap.HL, elev.mean, fl.rich, bb.rich, log.fl.abund, log.bb.abund) %>%
+  mutate(elev.mean.2 = elev.mean^2,
+         elev.mean.3 = elev.mean^3) %>%
+  mutate(niche.overlap.HL = scale(niche.overlap.HL),
+         elev.mean = scale(elev.mean),
+         elev.mean.2 = scale(elev.mean),
+         elev.mean.3 = scale(elev.mean^3),
+         fl.rich = scale(fl.rich),
+         log.fl.abund = scale(log.fl.abund),
+         log.bb.abund = scale(log.bb.abund)) %>%
+  as.matrix() %>%
+  as_tibble() %>%
+  mutate(niche.overlap.HL = as.numeric(niche.overlap.HL),
+         elev.mean = as.numeric(elev.mean),
+         elev.mean.2 = as.numeric(elev.mean^2),
+         elev.mean.3 = as.numeric(elev.mean^3),
+         fl.rich = as.numeric(fl.rich),
+         bb.rich = as.numeric(bb.rich),
+         log.fl.abund = as.numeric(log.fl.abund),
+         log.bb.abund = as.numeric(log.bb.abund),
+         site = factor(site),
+         year = factor(year)) %>%
+  na.omit()
+
+# Pairs plot to determine best model form
+predictors.scaled <- netmet.scaled %>%
+  select(elev.mean, elev.mean.2, elev.mean.3, bb.rich, log.bb.abund, fl.rich, log.fl.abund, niche.overlap.HL)
+
+ggpairs(predictors.scaled)
+
[…]

### 2020-09-21 analyses/abund_div_GAM.Rmd
update prior to trying to recover from a broken projecy

@@
   group_by(year, bb.sp) %>%
-  mutate(bb.annual.abund = sum(bb.abund))
+  mutate(bb.annual.abund = sum(bb.abund),
+         abund.class = case_when(
+           bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf") ~ "dominant",
+           bb.sp %in% c("jone", "psit", "hort") ~ "common",
+           bb.sp %in% c("pyre", "mend", "mont", "hypn", "lapi") ~ "high elev.",
+           bb.sp %in% c("gers", "humi", "muci") ~ "minor"
+         ),
+         abund.class = factor(abund.class, levels = c("dominant", "common", "high elev.", "minor")),
+         bb.sp.ord = factor(bb.sp, levels = c("bss", "pasc", "prat", "soro", "wurf", 
+                                              "hort", "jone", "psit",
+                                              "pyre", "mend", "mont",
+                                              "gers", "humi", "muci", "hypn", "lapi")))
 
@@
   mutate(year = factor(year(date)))
+
+### Visualize raw abundance/occurrence
+ggplot(filter(bb_abund_sp, bb.abund > 0), aes(yday, elev.mean, color = abund.class)) +
+  geom_point(aes(alpha = bb.abund)) +
+  #geom_point() +
+  #facet_grid(year~ bb.sp) 
+  facet_wrap(~bb.sp.ord)
+
+ggplot(bb_abund_sp, aes(bb.sp, log(bb.abund + 1))) +
+  geom_boxplot()
+
+ggplot(bb_abund_sp, aes(log(bb.abund + 1))) +
+  geom_histogram() +
+  facet_wrap(~bb.sp)
 ```
@@
 #check.gamViz(weighted_fl_abundance.2010)
-#weighted_fl_abundance.2010.sum <- summary(weighted_fl_abundance.2010)
+weighted_fl_abundance.2010.sum <- summary(weighted_fl_abundance.2010)
+saveRDS(weighted_fl_abundance.2010.sum, file = "./output/weighted_fl_abundance_2010.rds")
 #print(plot(weighted_fl_abundance.2010, allTerms = TRUE), pages = 4)
@@
 
+### Make a data set that sets the boundaries for plotting predictions extrapolated (or interpolated) beyond samples
+```{r}
+survey_samples <- survey %>%
+  ungroup() %>%
+  left_join(site_data) %>%
+  mutate(yday = yday(date),
+         year = year(date)) %>%
+  select(year, yday, elev.mean) %>%
+  unique() %>%
+  mutate(sample = factor(c(1:length(year)))) %>%
+  group_by(year, sample) %>%
+  arrange(yday)
+%>%
+  nest()
+
+dat <- survey_samples$data[[1]]
+
+func <- function(x, days, meters) {
+  for(i in 1:dim(x)[1]) {
+
+    yday = x[i, 2]
+    print(yday)
+   # x = seq(x$)
+  }
+}
+
+func(survey_samples),
+
+dat <- survey_samples$data[[1]]
+yday <- dat$yday # set reference yday
+ydays <- seq((yday - 7), (yday + 7), 0.5) # add all days within range
+elev <- dat$elev.mean # set reference elev
+elevs <- seq((elev - 100), (elev + 100), 1) # add all elevs within range
+out.temp <- expand.grid(ydays, elevs) # expand grid to all possible combos
+
+bb_samples <- net %>%
+  ungroup() %>%
+  left_join(site_data) %>%
+  mutate(yday = yday(date),
+         year = year(date)) %>%
+  select(year, yday, elev.mean) %>%
+  unique() %>%
+  mutate(sample = factor(c(1:length(year)))) %>%
+  group_by(sample, year) %>%
+  nest()
+
+### Fill func
+#### Creates a set of yday*elevation combinations that can be used to filter surface plots to exclude predictions too far from the data
+#### This is very hard-wired right now, and it would break if any of the column names changed.
+plotting_boundaries <- function(x,
+                                year_sub, # a sloppy way to let me filter by year
+                                days, # how many days of outside our samples will we plot?
+                                meters) { # how many meters of outside our samples will we plot?
+
+  #out <- tibble() # create empty tibble to fill
+  out <- c()
+
+  data <- x %>%
+    filter(year == year_sub)
+
+  for(i in 1:dim(data)[1]) {
+
+    dat <- x$data[[i]] # this pull out the ith data frame of the data column in the nested data frame
+
+    yday <- dat$yday # set reference yday
+    # ydays <- seq((yday - days), (yday + days), 0.5) # add all days within range
+    # 
+    # elev <- dat$elev.mean # set reference elev
+    # elevs <- seq((elev - meters), (elev + meters), 1) # add all elevs within range
+    # 
+    # out.temp <- expand.grid(ydays, elevs) # expand grid to all possible combos
+    # out <- bind_rows(out, out.temp)
+    out[i] <- yday
+  }
+  # out <- unique(out) %>% # drop duplicates created by overlapping ranges
+  #   select(yday = 1, elev.mean = 2) # name columns
+  out
+}
+
+FL_plotting_2010 <- plotting_boundaries(survey_samples, 2010, 7, 100)
+FL_plotting_2011 <- plotting_boundaries(survey_samples, 2011, 7, 100)
+FL_plotting_2012 <- plotting_boundaries(survey_samples, 2012, 7, 100)
+View(unique(FL_plotting_2012$yday))
+
+BB_plotting_2010 <- plotting_boundaries(bb_samples, 2010, 7, 100)
+BB_plotting_2011 <- plotting_boundaries(bb_samples, 2011, 7, 100)
+BB_plotting_2012 <- plotting_boundaries(bb_samples, 2012, 7, 100)
+```
+
 ### Extract predictions
 ```{r message=FALSE}
+### Species names and such for plotting
+species_names <- cbind(c("bss", "pasc", "prat", "soro", "wurf", 
+                         "hort", "pyre", "mend", "mont", "gers", 
+                         "hypn","jone", "lapi", "muci", "psit", "himi"),
+                       c("B. sensu-strictu", "B. pascuorum", "B. pratorum", 
+                         "B. soroensis", "B. wurflenii", "B. hortorum",
+                         "B. pyrenaeus", "B. mendax", "B. monticola",
+                         "B. gerstaeckeri", "B. hypnorum", "B. jonellus", 
+                         "B. lapidarius", "B. mucidus", "B. psithyrus", "B. humilis"),
+                       c("B. ss.", "B. pasc.", "B. prat.", 
+                         "B. soro.", "B. wurf.", "B. hort.",
+                         "B. pyre.", "B. mend.", "B. mont.",
+                         "B. gers.", "B. hypn.", "B. jone.", 
+                         "B. lapi.", "B. muci.", "B. psit.", "B. humi"),
+                       c("major", "major", "major", "major", "major",
+                         "major", "high elev.", "high elev.", "high elev.",
+                         "minor", "minor", "minor", "minor", "minor", "minor", "minor")
+                       ) %>%
+  as.data.frame() %>%
+  as_tibble() %>%
+  dplyr::select(bb.sp = 1, species.name = 2, species.abb = 3, abund.class = 4)
 # Function to build data set over which to generate predictions
@@
 # Generate predictions
-predict2010.FL <- predict_func(weighted_fl_abundance.2010, newdata2010.FL, 2010)
-predict2010.FL.sqrt <- predict_func(weighted_fl_abundance.2010.sqrt, newdata2010.FL, 2010)
-predict2010.FL.bin <- predict_func(weighted_fl_abundance.2010.bin, newdata2010.FL, 2010)
+predict2010.FL <- predict_func(weighted_fl_abundance.2010, newdata2010.FL, 2010) %>%
+  mutate(yday = ceiling(yday*2)/2) %>%
+  semi_join(FL_plotting_2010, by =  c("yday", "elev.mean"))
+predict2010.FL.sqrt <- predict_func(weighted_fl_abundance.2010.sqrt, newdata2010.FL, 2010) %>%
+  mutate(yday = ceiling(yday*2)/2) %>%
+  semi_join(FL_plotting_2010, by =  c("yday", "elev.mean"))
+predict2010.FL.bin <- predict_func(weighted_fl_abundance.2010.bin, newdata2010.FL, 2010) %>%
+  mutate(yday = ceiling(yday*2)/2) %>%
+  semi_join(FL_plotting_2010, by =  c("yday", "elev.mean"))
 
-predict2011.FL <- predict_func(weighted_fl_abundance.2011, newdata2011.FL, 2011)
-predict2011.FL.sqrt <- predict_func(weighted_fl_abundance.2011.sqrt, newdata2011.FL, 2011)
-predict2011.FL.bin <- predict_func(weighted_fl_abundance.2011.bin, newdata2011.FL, 2011)
+predict2011.FL <- predict_func(weighted_fl_abundance.2011, newdata2011.FL, 2011) %>%
+  mutate(yday = ceiling(yday*2)/2) %>%
+  semi_join(FL_plotting_2011, by =  c("yday", "elev.mean"))
+predict2011.FL.sqrt <- predict_func(weighted_fl_abundance.2011.sqrt, newdata2011.FL, 2011) %>%
+  mutate(yday = ceiling(yday*2)/2) %>%
+  semi_join(FL_plotting_2011, by =  c("yday", "elev.mean"))
+predict2011.FL.bin <- predict_func(weighted_fl_abundance.2011.bin, newdata2011.FL, 2011) %>%
+  mutate(yday = ceiling(yday*2)/2) %>%
+  semi_join(FL_plotting_2011, by =  c("yday", "elev.mean"))
 
 predict2012.FL <- predict_func(weighted_fl_abundance.2012, newdata2012.FL, 2012)
-predict2012.FL.sqrt <- predict_func(weighted_fl_abundance.2012.sqrt, newdata2012.FL, 2012)
-predict2012.FL.bin <- predict_func(weighted_fl_abundance.2012.bin, newdata2012.FL, 2012)
+predict2012.FL2 <- predict2012.FL %>%
+  mutate(yday = ceiling(yday*2)/2) %>%
+  semi_join(FL_plotting_2012, by =  c("yday", "elev.mean"))
+
+ggplot(predict2012.FL2, aes(yday, elev.mean, fill = pred)) +
+  geom_raster() +
+  geom_contour(aes(z = pred), color = "black", size = 0.25) +
+  scale_fill_viridis_c(name = "log(abund.)") +
+  facet_wrap(~ bb.sp)
+
+predict2012.FL.sqrt <- predict_func(weighted_fl_abundance.2012.sqrt, newdata2012.FL, 2012) %>%
+  mutate(yday = ceiling(yday*2)/2) %>%
+  semi_join(FL_plotting_2012, by =  c("yday", "elev.mean"))
+predict2012.FL.bin <- predict_func(weighted_fl_abundance.2012.bin, newdata2012.FL, 2012) %>%
+  mutate(yday = ceiling(yday*2)/2) %>%
+  semi_join(FL_plotting_2012, by =  c("yday", "elev.mean"))
 
@@
 
-predict_pool <- bind_rows(predict_pool.FL, predict_pool.BB)
-predict_pool.sqrt <- bind_rows(predict_pool.FL.sqrt, predict_pool.BB)
-predict_pool.bin <- bind_rows(predict_pool.FL.bin, predict_pool.BB)
-
-# Subset predictions
-predict_major <- filter(predict_pool, 
-                           bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf")) %>%
-  mutate(bb.sp = case_when(
-    bb.sp == "bss" ~ "B. ss.",
-    bb.sp == "pasc" ~ "B. pasc",
-    bb.sp == "prat" ~ "B. prat",
-    bb.sp == "soro" ~ "B. soro",
-    bb.sp == "wurf" ~ "B. wurf"
-  ))
-
-predict_major.sqrt <- filter(predict_pool.sqrt, 
-                           bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf")) %>%
-  mutate(bb.sp = case_when(
-    bb.sp == "bss" ~ "B. ss.",
-    bb.sp == "pasc" ~ "B. pasc",
-    bb.sp == "prat" ~ "B. prat",
-    bb.sp == "soro" ~ "B. soro",
-    bb.sp == "wurf" ~ "B. wurf"
-  ))
-
-predict_major.bin <- filter(predict_pool.bin, 
-                           bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf")) %>%
-  mutate(bb.sp = case_when(
-    bb.sp == "bss" ~ "B. ss.",
-    bb.sp == "pasc" ~ "B. pasc",
-    bb.sp == "prat" ~ "B. prat",
-    bb.sp == "soro" ~ "B. soro",
-    bb.sp == "wurf" ~ "B. wurf"
-  ))
-
-predict_highelev <- filter(predict_pool, 
-                              bb.sp %in% c("pyre", "mend", "mont")) %>%
-  mutate(bb.sp = case_when(
-    bb.sp == "pyre" ~ "B. pyre",
-    bb.sp == "mend" ~ "B. mend",
-    bb.sp == "mont" ~ "B. mont"
-  ))
-
-predict_highelev.sqrt <- filter(predict_pool.sqrt, 
-                                   bb.sp %in% c("pyre", "mend", "mont")) %>%
-  mutate(bb.sp = case_when(
-    bb.sp == "pyre" ~ "B. pyre",
-    bb.sp == "mend" ~ "B. mend",
-    bb.sp == "mont" ~ "B. mont"
-  ))
-
-predict_highelev.bin <- filter(predict_pool.bin, 
-                                  bb.sp %in% c("pyre", "mend", "mont")) %>%
-  mutate(bb.sp = case_when(
-    bb.sp == "pyre" ~ "B. pyre",
-    bb.sp == "mend" ~ "B. mend",
-    bb.sp == "mont" ~ "B. mont"
-  ))
-
-predict_minor <- filter(predict_pool, 
-                           bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci")) %>%
-  mutate(bb.sp = case_when(
-    bb.sp == "gers" ~ "B. gers",
-    bb.sp == "hort" ~ "B. hort",
-    bb.sp == "hypn" ~ "B. hypn",
-    bb.sp == "jone" ~ "B. jone",
-    bb.sp == "lapi" ~ "B. lapi",
-    bb.sp == "muci" ~ "B. muci"
-  ))
-
-predict_minor.sqrt <- filter(predict_pool.sqrt, 
-                                bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci")) %>% 
-  mutate(bb.sp = case_when(
-    bb.sp == "gers" ~ "B. gers",
-    bb.sp == "hort" ~ "B. hort",
-    bb.sp == "hypn" ~ "B. hypn",
-    bb.sp == "jone" ~ "B. jone",
-    bb.sp == "lapi" ~ "B. lapi",
-    bb.sp == "muci" ~ "B. muci"
-  ))
-
-predict_minor.bin <- filter(predict_pool.bin, 
-                               bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci")) %>% 
-  mutate(bb.sp = case_when(
-    bb.sp == "gers" ~ "B. gers",
-    bb.sp == "hort" ~ "B. hort",
-    bb.sp == "hypn" ~ "B. hypn",
-    bb.sp == "jone" ~ "B. jone",
-    bb.sp == "lapi" ~ "B. lapi",
-    bb.sp == "muci" ~ "B. muci"
-  ))
+predict_pool <- bind_rows(predict_pool.FL, predict_pool.BB) %>%
+  left_join(species_names)
+predict_pool.sqrt <- bind_rows(predict_pool.FL.sqrt, predict_pool.BB) %>%
+  left_join(species_names)
+predict_pool.bin <- bind_rows(predict_pool.FL.bin, predict_pool.BB) %>%
+  left_join(species_names)
 ```
@@
 #### Major
-ggplot(predict_major, aes(date, elev.mean, fill = pred)) +
+ggplot(filter(predict_pool, abund.class == "major"), aes(date, elev.mean, fill = pred)) +
   geom_raster() +
@@
   theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
-ggsave("./output/GAM_major.png", height = 5, width = 11, units = "in")
+ggsave("./output/GAM_major.png", height = 3, width = 6.5, units = "in")
 

### 2020-09-21 analyses/network_analysis_v2.Rmd
update prior to trying to recover from a broken projecy

--- 2020-09-17 analyses/network_analysis.Rmd
@@
       #map(specieslevel, level = "higher", index = "d") %>%
-      map(specieslevel, level = "higher", index = c("d", "resource range")) %>%
+      map(specieslevel, level = "higher", index = c("d", "resource range", "degree", "partner diversity")) %>%
       map(rownames_to_column) %>%
@@
       frame2webs() %>% # convert to to bipartite web
-      map(networklevel, level = "higher", weighted = TRUE, index = c("niche overlap")) %>% # We probably care mainly about the BB level of the network
+      map(networklevel, level = "higher", weighted = TRUE, index = c("niche overlap", "web asymmetry", "weighted NODF", "partner diversity", "C score")) %>% # We probably care mainly about the BB level of the network
       data.frame() %>%
@@
   group_by(site, year, bb.sp) %>%
-  summarize(bb.sp.abund = mean(bb.abund))
+  summarize(bb.sp.abund = mean(bb.abund),
+            log.bb.abund = log(mean(bb.abund)),
+            bb.transects = length(unique(date)))
 
@@
 
-wma1_2010 <- filter(net, site == "wma1", year == "2010")
-
 spmet <- bind_rows(spmet_2012, spmet_2011, spmet_2010) %>%
@@
   group_by(site, year) %>%
-  #summarize(d = weighted.mean(d, bb.sp.abund, na.rm = TRUE)) %>%
-  summarize(resource.range = weighted.mean(resource.range, bb.sp.abund, na.rm = TRUE)) %>%
+  summarize(resource.range = weighted.mean(resource.range, bb.sp.abund, na.rm = TRUE),
+            d = weighted.mean(d, bb.sp.abund, na.rm = TRUE),
+            degree = weighted.mean(degree, bb.sp.abund, na.rm = TRUE),
+            partner.diversity = weighted.mean(partner.diversity, bb.sp.abund, na.rm = TRUE)) %>%
+  #summarize(resource.range = mean(resource.range, na.rm = TRUE)) %>%
   ungroup()
@@
          site = factor(site),
-         management = factor(management))
+         management = factor(management),
+         threshold = factor(case_when(
+           elev.mean > 1300 ~ "above",
+           elev.mean <= 1300 ~ "below"
+           ))
+         )
 ```
@@
 ```{r}
+ggplot(netmet, aes(elev.mean, C.score.HL,)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~year)
+
 #### Niche overlap
@@
                              s(bb.transects) + 
+                             s(log.bb.abund) +
+                             #s(bb.rich) +
                              s(elev.mean, by = year, k = 10),
@@
 
-# bb.transects was zeroed out by select = TRUE, so this plot should reflect the model above perfectly
 ggplot(netmet, aes(elev.mean, niche.overlap.HL)) +
@@
 
-##### Community mean d'
-gam_mean_d.1 <- gam(d ~ 
+ggplot(netmet, aes(log.bb.abund, niche.overlap.HL)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, bs = "tp", k = 10)) +
+  facet_wrap(~year)
+
+ggplot(netmet, aes(bb.rich, niche.overlap.HL)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, bs = "tp", k = 10)) +
+  facet_wrap(~year)
+
+#### Niche overlap
+gam_C.score.1 <- gam(C.score.HL ~ 
                              s(year, bs = "re") + 
                              s(bb.transects) + 
+                             s(log.bb.abund) +
+                             #s(bb.rich) +
                              s(elev.mean, by = year, k = 10),
@@
 
-check.gamViz(gam_mean_d.1)
-concurvity(gam_mean_d.1, full = TRUE)
-concurvity(gam_mean_d.1, full = FALSE)
-summary(gam_mean_d.1)
-print(plot(gam_mean_d.1, allTerms = TRUE), pages = 1)
+check.gamViz(gam_C.score.1)
+concurvity(gam_C.score.1, full = TRUE)
+concurvity(gam_C.score.1, full = FALSE)
+summary(gam_C.score.1)
+print(plot(gam_C.score.1, allTerms = TRUE), pages = 1)
 
-ggplot(netmet, aes(elev.mean, d)) +
+ggplot(netmet, aes(elev.mean, C.score.HL)) +
   geom_point() +
@@
 
-ggplot(netmet, aes(bb.transects, d)) +
+ggplot(netmet, aes(log.bb.abund, C.score.HL)) +
   geom_point() +
-  geom_smooth(method = "lm") +
-  facet_wrap(~year, scales = "free_x")
+  geom_smooth(method = "gam", formula = y ~ s(x, bs = "tp", k = 10)) +
+  facet_wrap(~year)
 
-# I would be reluctant to interpret the year effect biologically; more likely an artifact of sampling intensity
-ggplot(netmet, aes(year, d)) +
-  geom_boxplot()
+ggplot(netmet, aes(bb.rich, C.score.HL)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, bs = "tp", k = 10)) +
+  facet_wrap(~year)
 
-# It is interesting that, while elevation explains a lot of niche overlap and virtually none of specialization, niche overlap and specialization are strongly correlated with each other.
-ggplot(netmet, aes(niche.overlap.HL, d)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_wrap(~year)
+### NODF
+gam_NODF.1 <- gam(weighted.NODF ~ 
+                             s(year, bs = "re") + 
+                             s(bb.transects) + 
+                             s(log.bb.abund) +
+                             #s(bb.rich) +
+                             s(elev.mean, by = year, k = 10),
+                           select = TRUE,
+                           data = netmet,
+                           method = "REML") %>% getViz()
+
+check.gamViz(gam_NODF.1)
+concurvity(gam_NODF.1, full = TRUE)
+concurvity(gam_NODF.1, full = FALSE)
+summary(gam_NODF.1)
+print(plot(gam_NODF.1, allTerms = TRUE), pages = 1)
 
@@
                       s(bb.transects) +
+                       s(log.bb.abund) +
                       s(elev.mean, by = year, k = 10),
@@
                     s(bb.sp, bs = "re") +
+                     s(log.bb.abund, k = 5) +
                     s(elev.mean, by = bb.sp, bs = "tp", k = 5),
@@
 print(plot(gam_rr.sp_I, allTerms = TRUE), pages = 4)
+
+ggplot(spmet, aes(log.bb.abund, resource.range)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~year, scales =  "free")
 
@@
 
-#### Community mean d'
-gam_mean_d.2 <- gam(d ~ 
+#### Resource range
+gam_mean_rr.2 <- gam(resource.range ~ 
                       s(year, bs = "re") + 
-                      s(management, bs = "re") +
+                      #s(management, bs = "re") +
                       s(bb.transects) +
                       s(elev.mean, by = year, k = 10) +
-                      s(bb.rich, k = 5) +
-                      s(log.fl.abund, k = 5),
+                      #s(bb.rich, k = 5) +
+                       s(fl.rich, k = 5) +
+                       s(log.bb.abund, k = 5),
+                     # s(log.fl.abund, k = 5),
                     select = TRUE,
@@
 
-check.gamViz(gam_mean_d.2)
-concurvity(gam_mean_d.2, full = TRUE)
-concurvity(gam_mean_d.2, full = FALSE)
-summary(gam_mean_d.2)
-print(plot(gam_mean_d.2, allTerms = TRUE), pages = 1)
+check.gamViz(gam_mean_rr.2)
+concurvity(gam_mean_rr.2, full = TRUE)
+concurvity(gam_mean_rr.2, full = FALSE)
+summary(gam_mean_rr.2)
+print(plot(gam_mean_rr.2, allTerms = TRUE), pages = 1)
 
-ggplot(netmet, aes(bb.rich, d)) +
+ggplot(netmet, aes(bb.rich, resource.range)) +
   geom_point() +
@@
 
-ggplot(netmet, aes(log.bb.abund, d)) +
+ggplot(netmet, aes(log.bb.abund, resource.range)) +
   geom_point() +
@@
 
-ggplot(netmet, aes(log.fl.abund, d)) +
+ggplot(netmet, aes(log.fl.abund, resource.range)) +
   geom_point() +
@@
 
-ggplot(netmet, aes(fl.rich, d)) +
+ggplot(netmet, aes(fl.rich, resource.range)) +
   geom_point() +
@@
 ```
+
+## Threshold-based network analysis: do network properties differ meaningfully above and below the threshold (~1300m) evident in my abundance and beta-diversity analyses?
+```{r}
+
+```

### 2020-09-21 analyses/trash_can.R

@@ -1744,3 +1744,23 @@ ggplot(web.top_splevel, aes(elev.mean, d)) +
   theme_abyss() +
   facet_wrap(~year)
 ```
+
+
+### Play with null models
+```{r}
+webs <- net %>%
+  group_by(site, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
+  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
+  ungroup() %>%
+  unite(webID, c(site, year), sep = "_") %>%
+  dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
+  data.frame() %>% # convert to data frame
+  frame2webs()
+
+null.t.test(webs[[2]], level = "higher", index = "niche overlap")
+
+n1 <- nullmodel(webs[[1]], N = 1000, method = 1)
+met1 <- map(n1, specieslevel, level = "higher", index = "resource range") %>%
+  bind_rows(.id = "id") %>%
+  
+  ```

### 2020-09-25 analyses/abund_div_GAM.Rmd
refined plotting

--- 2020-09-15 analyses/abund_div_GAM.Rmd
@@
 #weighted_fl_abundance.2010.sum <- summary(weighted_fl_abundance.2010)
+#saveRDS(weighted_fl_abundance.2010.sum, file = "./output/weighted_fl_abundance_2010.rds")
 #print(plot(weighted_fl_abundance.2010, allTerms = TRUE), pages = 4)
@@
 
+# We might be interested in whether and where (in time*elev) a given pair of smooths differs from each other.
+It might only make sense to do this, though, if we first normalized abundance for each species; otherwise, the differences may be driven by differences in overall abundance.
+I'm not sure the species random effect takes care of that.
+```{r}
+# 2010 pasc vs prat
+plotDiff.mgcv.smooth.2D(s1 = sm(bb_abund.2010, 10), s2 = sm(bb_abund.2010, 11)) +
+  l_pvRaster() + 
+  l_pvContour()
+
+plotDiff.mgcv.smooth.2D(s1 = sm(weighted_fl_abundance.2010, 12), 
+                        s2 = sm(weighted_fl_abundance.2010, 13)) +
+  l_pvRaster() + 
+  l_pvContour()
+
+# 2010 pasc vs prat
+plotDiff.mgcv.smooth.2D(s1 = sm(bb_abund.2011, 10), s2 = sm(bb_abund.2011, 11)) +
+  l_pvRaster() + 
+  l_pvContour()
+
+# 2010 pasc vs prat
+plotDiff.mgcv.smooth.2D(s1 = sm(bb_abund.2012, 10), s2 = sm(bb_abund.2012, 11)) +
+  l_pvRaster() + 
+  l_pvContour()
+```
+
 ### Extract predictions
 ```{r message=FALSE}
+### Species names and such for plotting
+species_names <- cbind(c("bss", "pasc", "prat", 
+                         "soro", "wurf", "hort", 
+                         "pyre", "mend", "mont", 
+                         "gers", "hypn","jone", 
+                         "lapi", "muci", "psit", 
+                         "humi"),
+                       c("B. sensu-strictu", "B. pascuorum", "B. pratorum", 
+                         "B. soroensis", "B. wurflenii", "B. hortorum",
+                         "B. pyrenaeus", "B. mendax", "B. monticola",
+                         "B. gerstaeckeri", "B. hypnorum", "B. jonellus", 
+                         "B. lapidarius", "B. mucidus", "B. psithyrus", 
+                         "B. humilis"),
+                       c("B. ss.", "B. pasc.", "B. prat.", 
+                         "B. soro.", "B. wurf.", "B. hort.",
+                         "B. pyre.", "B. mend.", "B. mont.",
+                         "B. gers.", "B. hypn.", "B. jone.", 
+                         "B. lapi.", "B. muci.", "B. psit.", 
+                         "B. humi"),
+                       c("major", "major", "major", 
+                         "major", "major", "common", 
+                         "high elev.", "high elev.", "high elev.",
+                         "minor", "minor", "common", 
+                         "minor", "minor", "common", 
+                         "minor")
+                       ) %>%
+  as.data.frame() %>%
+  as_tibble() %>%
+  dplyr::select(bb.sp = 1, species.name = 2, species.abb = 3, abund.class = 4)
+
+BB_samples <- net %>%
+  mutate(year = year(date),
+         date = as_date(yday(date))) %>%
+  left_join(site_data) %>%
+  select(year, date, elev.mean) %>%
+  mutate(subject = rep("BB", n()),
+         sampled = rep(TRUE, n())) %>%
+  unique()
+
+FL_samples <- survey %>%
+  ungroup() %>%
+  mutate(year = year(date),
+         date = as_date(yday(date))) %>%
+  unique() %>%
+  left_join(site_data) %>%
+  select(year, date, elev.mean) %>%
+  mutate(subject = rep("FL", n()),
+         sampled = rep(TRUE, n())) %>%
+  unique()
+
+samples <- bind_rows(BB_samples, FL_samples)
+
 # Function to build data set over which to generate predictions
@@
     filter(year == y) %>%
-    select(bb.sp) %>%
+    dplyr::select(bb.sp) %>%
     unique()
@@
     complete(seq_yday, seq_elev) %>% # get all possible values
-    crossing(spp) %>% # get all possible combos with species
-    select(yday = seq_yday, elev.mean = seq_elev, bb.sp) # rename columns to match the original model inputs
+    tidyr::crossing(spp) %>% # get all possible combos with species
+    dplyr::select(yday = seq_yday, elev.mean = seq_elev, bb.sp) # rename columns to match the original model inputs
 }
@@
 predict_pool.FL <- bind_rows(predict2010.FL, predict2011.FL, predict2012.FL) %>%
-  mutate(date = as_date(yday),
-         subject = factor(rep("FL", n())))
+  mutate(subject = factor(rep("FL", n())))
+
 predict_pool.FL.sqrt <- bind_rows(predict2010.FL.sqrt, predict2011.FL.sqrt, predict2012.FL.sqrt) %>%
-  mutate(date = as_date(yday),
-         subject = factor(rep("FL", n())))
+  mutate(subject = factor(rep("FL", n())))
+
 predict_pool.FL.bin <- bind_rows(predict2010.FL.bin, predict2011.FL.bin, predict2012.FL.bin) %>%
-  mutate(date = as_date(yday),
-         subject = factor(rep("FL", n())))
+  mutate(subject = factor(rep("FL", n())))
 
@@
 predict_pool.BB <- bind_rows(predict2010.BB, predict2011.BB, predict2012.BB) %>%
-  mutate(date = as_date(yday),
-         subject = factor(rep("BB", n())),
-         pred = log(pred))
+  mutate(subject = factor(rep("BB", n()))) %>%
+  mutate(pred = log(pred))
 
-predict_pool <- bind_rows(predict_pool.FL, predict_pool.BB)
-predict_pool.sqrt <- bind_rows(predict_pool.FL.sqrt, predict_pool.BB)
-predict_pool.bin <- bind_rows(predict_pool.FL.bin, predict_pool.BB)
-
-# Subset predictions
-predict_major <- filter(predict_pool, 
-                           bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf")) %>%
-  mutate(bb.sp = case_when(
-    bb.sp == "bss" ~ "B. ss.",
-    bb.sp == "pasc" ~ "B. pasc",
-    bb.sp == "prat" ~ "B. prat",
-    bb.sp == "soro" ~ "B. soro",
-    bb.sp == "wurf" ~ "B. wurf"
-  ))
-
-predict_major.sqrt <- filter(predict_pool.sqrt, 
-                           bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf")) %>%
-  mutate(bb.sp = case_when(
-    bb.sp == "bss" ~ "B. ss.",
-    bb.sp == "pasc" ~ "B. pasc",
-    bb.sp == "prat" ~ "B. prat",
-    bb.sp == "soro" ~ "B. soro",
-    bb.sp == "wurf" ~ "B. wurf"
-  ))
-
-predict_major.bin <- filter(predict_pool.bin, 
-                           bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf")) %>%
-  mutate(bb.sp = case_when(
-    bb.sp == "bss" ~ "B. ss.",
-    bb.sp == "pasc" ~ "B. pasc",
-    bb.sp == "prat" ~ "B. prat",
-    bb.sp == "soro" ~ "B. soro",
-    bb.sp == "wurf" ~ "B. wurf"
-  ))
-
-predict_highelev <- filter(predict_pool, 
-                              bb.sp %in% c("pyre", "mend", "mont")) %>%
-  mutate(bb.sp = case_when(
-    bb.sp == "pyre" ~ "B. pyre",
-    bb.sp == "mend" ~ "B. mend",
-    bb.sp == "mont" ~ "B. mont"
-  ))
-
-predict_highelev.sqrt <- filter(predict_pool.sqrt, 
-                                   bb.sp %in% c("pyre", "mend", "mont")) %>%
-  mutate(bb.sp = case_when(
-    bb.sp == "pyre" ~ "B. pyre",
-    bb.sp == "mend" ~ "B. mend",
-    bb.sp == "mont" ~ "B. mont"
-  ))
-
-predict_highelev.bin <- filter(predict_pool.bin, 
-                                  bb.sp %in% c("pyre", "mend", "mont")) %>%
-  mutate(bb.sp = case_when(
-    bb.sp == "pyre" ~ "B. pyre",
-    bb.sp == "mend" ~ "B. mend",
-    bb.sp == "mont" ~ "B. mont"
-  ))
-
-predict_minor <- filter(predict_pool, 
-                           bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci")) %>%
-  mutate(bb.sp = case_when(
-    bb.sp == "gers" ~ "B. gers",
-    bb.sp == "hort" ~ "B. hort",
-    bb.sp == "hypn" ~ "B. hypn",
-    bb.sp == "jone" ~ "B. jone",
-    bb.sp == "lapi" ~ "B. lapi",
-    bb.sp == "muci" ~ "B. muci"
-  ))
-
-predict_minor.sqrt <- filter(predict_pool.sqrt, 
-                                bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci")) %>% 
-  mutate(bb.sp = case_when(
-    bb.sp == "gers" ~ "B. gers",
-    bb.sp == "hort" ~ "B. hort",
-    bb.sp == "hypn" ~ "B. hypn",
-    bb.sp == "jone" ~ "B. jone",
-    bb.sp == "lapi" ~ "B. lapi",
-    bb.sp == "muci" ~ "B. muci"
-  ))
-
-predict_minor.bin <- filter(predict_pool.bin, 
-                               bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci")) %>% 
-  mutate(bb.sp = case_when(
-    bb.sp == "gers" ~ "B. gers",
-    bb.sp == "hort" ~ "B. hort",
-    bb.sp == "hypn" ~ "B. hypn",
-    bb.sp == "jone" ~ "B. jone",
-    bb.sp == "lapi" ~ "B. lapi",
-    bb.sp == "muci" ~ "B. muci"
-  ))
+predict_pool <- bind_rows(predict_pool.FL, predict_pool.BB) %>%
+  left_join(species_names) %>%
+  mutate(date = as_date(yday))
+predict_pool.sqrt <- bind_rows(predict_pool.FL.sqrt, predict_pool.BB) %>%
+  left_join(species_names) %>%
+  mutate(date = as_date(yday))
+predict_pool.bin <- bind_rows(predict_pool.FL.bin, predict_pool.BB) %>%
+  left_join(species_names) %>%
+  mutate(date = as_date(yday))
 ```
@@
 ```{r}
+# Convex hull function: https://cran.r-project.org/web/packages/ggplot2/vignettes/extending-ggplot2.html
+
+StatChull <- ggproto("StatChull", Stat,
+  compute_group = function(data, scales) {
+    data[chull(data$x, data$y), , drop = FALSE]
+  },
+
+  required_aes = c("x", "y")
+)
+
+stat_chull <- function(mapping = NULL, data = NULL, geom = "polygon",
+                       position = "identity", na.rm = FALSE, show.legend = NA, 
+                       inherit.aes = TRUE, ...) {
+  layer(
+    stat = StatChull, data = data, mapping = mapping, geom = geom, 
+    position = position, show.legend = show.legend, inherit.aes = inherit.aes,
+    params = list(na.rm = na.rm, ...)
+  )
+}
+##
+
 #### Major
-ggplot(predict_major, aes(date, elev.mean, fill = pred)) +
-  geom_raster() +
-  geom_contour(aes(z = pred), color = "black", size = 0.25) +
+ggplot() +
+  geom_raster(data = filter(predict_pool, abund.class == "major"), 
+              aes(date, elev.mean, fill = pred), 
+              interpolate = TRUE) +
+  geom_contour(data = filter(predict_pool, abund.class == "major"),
+               aes(date, elev.mean, z = pred), color = "black", size = 0.25) +
   scale_fill_viridis_c(name = "log(abund.)") +
-  #facet_grid(bb.sp + subject ~ year) +
-  facet_grid(year ~ bb.sp + subject) +
+  stat_chull(data = samples, aes(date, elev.mean), fill = NA, 
+             color = "deeppink", size = 0.25) +
+  facet_grid(year ~ species.abb + subject, scales = "free_y") +
   xlab("Date") +
   ylab("Elevation (m)") +
-  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
-ggsave("./output/GAM_major.png", height = 5, width = 11, units = "in")
+  theme_gray(6) +
+  theme(axis.text.x = element_text(angle = 90, size = 5, vjust = 0.5)) 
+ggsave("./output/GAM_major.png", height = 2.5, width = 6.5, units = "in")
+ggsave("./output/GAM_major.pdf", height = 2.5, width = 6.5)
 
-ggplot(predict_major.sqrt, aes(date, elev.mean, fill = pred)) +
-  geom_raster() +
-  geom_contour(aes(z = pred), color = "black", size = 0.25) +
+#### High Elev
+ggplot() +
+  geom_raster(data = filter(predict_pool, abund.class == "high elev."), 
+              aes(date, elev.mean, fill = pred), 
+              interpolate = TRUE) +
+  geom_contour(data = filter(predict_pool, abund.class == "high elev."),
+               aes(date, elev.mean, z = pred), color = "black", size = 0.25) +
   scale_fill_viridis_c(name = "log(abund.)") +
-  #facet_grid(bb.sp + subject ~ year) +
-  facet_grid(year ~ bb.sp + subject) +
+  stat_chull(data = samples, aes(date, elev.mean), fill = NA, 
+             color = "deeppink", size = 0.25) +
+  facet_grid(year ~ species.abb + subject, scales = "free_y") +
   xlab("Date") +
   ylab("Elevation (m)") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
-ggsave("./output/GAM_major_sqrt.png")
-
-ggplot(predict_major.bin, aes(date, elev.mean, fill = pred)) +
-  geom_raster() +
-  geom_contour(aes(z = pred), color = "black", size = 0.25) +
-  scale_fill_viridis_c(name = "log(abund.)") +
-  #facet_grid(bb.sp + subject ~ year) +
-  facet_grid(year ~ bb.sp + subject) +
-  xlab("Date") +
-  ylab("Elevation (m)") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
-ggsave("./output/GAM_major_bin.png")
-
-#### High Elev
-ggplot(predict_highelev, aes(date, elev.mean, fill = pred)) +
-  geom_raster() +
-  geom_contour(aes(z = pred), color = "black", size = 0.25) +
-  scale_fill_viridis_c(name = "log(abund.)") +
-  #facet_grid(bb.sp + subject ~ year) +
-  facet_grid(year ~ bb.sp + subject) +
-  xlab("Date") +
-  ylab("Elevation (m)") +
-  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
-ggsave("./output/GAM_highelev.png", height = 5, width = 11, units = "in")
-
-ggplot(predict_highelev.sqrt, aes(date, elev.mean, fill = pred)) +
-  geom_raster() +
-  geom_contour(aes(z = pred), color = "black", size = 0.25) +
-  scale_fill_viridis_c(name = "log(abund.)") +
-  #facet_grid(bb.sp + subject ~ year) +
-  facet_grid(year ~ bb.sp + subject) +
-  xlab("Date") +
-  ylab("Elevation (m)") +
-  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
-ggsave("./output/GAM_highelev_sqrt.png", height = 5, width = 11, units = "in")
-
-ggplot(predict_highelev.bin, aes(date, elev.mean, fill = pred)) +
-  geom_raster() +
-  geom_contour(aes(z = pred), color = "black", size = 0.25) +
-  scale_fill_viridis_c(name = "log(abund.)") +
-  #facet_grid(bb.sp + subject ~ year) +
-  facet_grid(year ~ bb.sp + subject) +
-  xlab("Date") +
-  ylab("Elevation (m)") +
-  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
-ggsave("./output/GAM_highelev_bin.png", height = 5, width = 11, units = "in")
+  theme_gray(6) +
+  theme(axis.text.x = element_text(angle = 90, size = 5, vjust = 0.5)) 
+ggsave("./output/GAM_highelev.png", height = 2.5, width = 6.5, units = "in")
+ggsave("./output/GAM_highelev.pdf", height = 2.5, width = 6.5)
 
 #### Minor
-ggplot(predict_minor, aes(date, elev.mean, fill = pred)) +
-  geom_raster() +
-  geom_contour(aes(z = pred), color = "black", size = 0.25) +
+ggplot() +
+  geom_raster(data = filter(predict_pool, abund.class == "minor"), 
+              aes(date, elev.mean, fill = pred), 
+              interpolate = TRUE) +
+  geom_contour(data = filter(predict_pool, abund.class == "minor"),
+               aes(date, elev.mean, z = pred), color = "black", size = 0.25) +
   scale_fill_viridis_c(name = "log(abund.)") +
-  #facet_grid(bb.sp + subject ~ year) +
+  stat_chull(data = samples, aes(date, elev.mean), fill = NA, 
+             color = "white", size = 0.5, linetype = "dashed") +
   facet_grid(year ~ bb.sp + subject) +
@@
 
-ggplot(predict_minor.sqrt, aes(date, elev.mean, fill = pred)) +
-  geom_raster() +
-  geom_contour(aes(z = pred), color = "black", size = 0.25) +
+### Just for kicks and chortles, let's see what happens if we scale the prediction for each species, then average across species to create a single plot.
+predict_pool.mean <- predict_pool %>%
+  group_by(bb.sp, species.name, species.abb, year, subject) %>%
+  mutate(pred.norm = scale(pred)) %>%
+  group_by(date, elev.mean, year, subject, abund.class) %>%
+  summarize(pred.mean = mean(pred.norm))
+
+ggplot() +
+  geom_raster(data = predict_pool.mean, 
+              aes(date, elev.mean, fill = pred.mean), 
+              interpolate = TRUE) +
+  geom_contour(data = predict_pool.mean,
+               aes(date, elev.mean, z = pred.mean), 
+               color = "black", size = 0.25) +
   scale_fill_viridis_c(name = "log(abund.)") +
-  #facet_grid(bb.sp + subject ~ year) +
-  facet_grid(year ~ bb.sp + subject) +
+  stat_chull(data = samples, aes(date, elev.mean), fill = NA, 
+             color = "white", size = 0.5, linetype = "dashed") +
+  facet_grid(year ~ abund.class + subject) +
   xlab("Date") +
@@
   theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
-ggsave("./output/GAM_minor_sqrt.png", height = 5, width = 11, units = "in")
-
-ggplot(predict_minor.bin, aes(date, elev.mean, fill = pred)) +
-  geom_raster() +
-  geom_contour(aes(z = pred), color = "black", size = 0.25) +
-  scale_fill_viridis_c(name = "log(abund.)") +
-  #facet_grid(bb.sp + subject ~ year) +
-  facet_grid(year ~ bb.sp + subject) +
-  xlab("Date") +
-  ylab("Elevation (m)") +
-  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
-ggsave("./output/GAM_minor_bin.png", height = 5, width = 11, units = "in")
+ggsave("./output/GAM_mean.png", height = 5, width = 11, units = "in")
 ```

### 2020-09-25 analyses/beta_diversity.Rmd
refined plotting

@@
 library(gdm)
+library(patchwork)
 library(bipartite)
@@
 # Elevation diff within 1 wk time diff
-ggplot(filter(beta_mit_data, year1 == year2 &
+p1 <- ggplot(filter(beta_mit_data, year1 == year2 &
                 site1 != site2 &
                 metric != "S" & 
-                !metric.class %in% c("OS.partition", "WN.partition") &
-                yday.diff <= 7), 
+                !metric.class %in% c("OS.partition", "WN.partition") 
+                #& yday.diff <= 7
+                ), 
        aes(elev.diff, value, color = metric)) +
-  geom_point(alpha = 0.1, pch = 16, size = 1) +
-  geom_smooth() +
+  geom_point(alpha = 0.03, pch = 16, size = 0.25) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 6)) +
   facet_grid(metric.class ~ year1) +
   scale_color_manual(values = p, name = "partition") +
-  xlab("Elevation difference") +
+  xlab("Elevation difference (m)") +
   ylab("β-diversity")
@@
 # Time diff within 250 m elev diff
-ggplot(filter(beta_mit_data, year1 == year2 &
+p2 <- ggplot(filter(beta_mit_data, year1 == year2 &
                 site1 != site2 &
                 metric != "S" & 
-                !metric.class %in% c("OS.partition", "WN.partition") &
-                elev.diff <= 250), 
+                !metric.class %in% c("OS.partition", "WN.partition")
+               # & elev.diff <= 250
+                ), 
        aes(yday.diff, value, color = metric)) +
-  geom_point(alpha = 0.1, pch = 16, size = 1) +
-  geom_smooth() +
+  geom_point(alpha = 0.03, pch = 16, size = 0.25) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 6)) +
   facet_grid(metric.class ~ year1) +
   scale_color_manual(values = p, name = "partition") +
-  xlab("Elevation difference") +
-  ylab("β-diversity")
+  xlab("Time difference (days)") +
+  ylab("β-diversity") + 
+  theme(legend.position="none")
 ggsave("./output/beta_time.png", height = 4.5, width = 6.5, units = "in")
+
+png("./output/beta_stack.png", width = 6.5, height = 6.5, units = "in", res = 300)
+p1 / p2 + plot_annotation(tag_levels = c("A", "B", "C")) + plot_layout(guides = "collect")
+dev.off()
+
+WN.hist <- ggplot(filter(beta_mit_data, year1 == year2 &
+                           site1 != site2 &
+                           metric == "WN"), 
+                aes(value)) +
+  geom_histogram() +
+  scale_y_continuous(trans="pseudo_log", breaks = c(0, 10, 100, 1000, 10000)) +
+  geom_vline(aes(xintercept = mean(value))) +
+  facet_wrap(~year1) +
+  xlab("Total β-diversity") +
+  ylab("Site-date pairs")
+ggsave("./output/WN_hist.png", width = 6.5, height = 3, units = "in")
 ```

### 2020-09-25 analyses/overview.Rmd
refined plotting

---
title: "overview"
output: html_document
---

### Set up
```{r}
library(bipartite)
library(ggbipart)
library(GGally)
library(tidyverse)
library(lubridate)
library(patchwork)
```

### Load site data
```{r, include=FALSE}
site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) # turn elav.class into an ordered factor
```

### Load network and survey data
```{r, include=FALSE}
net <- read_csv("../processed_data/network.csv") %>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  mutate(year = factor(year)) %>% # convert year to factor
  left_join(site_data, by = "site") %>%
  ungroup() %>%
  arrange(site, date)

# Yearly subsets
net_2012 <- filter(net, year == "2012")
net_2011 <- filter(net, year == "2011")
net_2010 <- filter(net, year == "2010")

# Floral survey
survey <- read_csv("../processed_data/floral_survey.csv") %>%
  dplyr::select(site, date, plant.sp, flower.cover)

### Species names and such for plotting
species_names <- cbind(c("bss", "pasc", "prat", 
                         "soro", "wurf", "hort", 
                         "pyre", "mend", "mont", 
                         "gers", "hypn","jone", 
                         "lapi", "muci", "psit", 
                         "humi"),
                       c("B. sensu-strictu", "B. pascuorum", "B. pratorum", 
                         "B. soroensis", "B. wurflenii", "B. hortorum",
                         "B. pyrenaeus", "B. mendax", "B. monticola",
                         "B. gerstaeckeri", "B. hypnorum", "B. jonellus", 
                         "B. lapidarius", "B. mucidus", "B. psithyrus", 
                         "B. humilis"),
                       c("B. ss.", "B. pasc.", "B. prat.", 
                         "B. soro.", "B. wurf.", "B. hort.",
                         "B. pyre.", "B. mend.", "B. mont.",
                         "B. gers.", "B. hypn.", "B. jone.", 
                         "B. lapi.", "B. muci.", "B. psit.", 
                         "B. humi"),
                       c("major", "major", "major", 
                         "major", "major", "common", 
                         "high elev.", "high elev.", "high elev.",
                         "minor", "minor", "common", 
                         "minor", "minor", "common", 
                         "minor")
                       ) %>%
  as.data.frame() %>%
  as_tibble() %>%
  dplyr::select(bb.sp = 1, species.name = 2, species.abb = 3, abund.class = 4)
```

### Visualize metaweb
```{r}
taxa <- read_csv("../processed_data/floral_tax.csv")

bb_range <- net %>%
  left_join(site_data) %>%
  group_by(site, date, elev.mean, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, bb.sp) %>%
  summarize(abund = mean(abund)) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2)) %>%
  left_join(species_names) %>%
  mutate(abund.class = factor(abund.class, levels = c("major", "common", "high elev.", "minor")))

bb_range.year <- net %>%
  left_join(site_data) %>%
  group_by(site, date, elev.mean, bb.sp, year) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, bb.sp, year) %>%
  summarize(abund = mean(abund)) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2)) %>%
  left_join(species_names) %>%
  mutate(abund.class = factor(abund.class, levels = c("major", "common", "high elev.", "minor")))

metaweb <- net %>%
  separate(plant.sp, c("plant.genus", "species"), sep = " ") %>%
  left_join(taxa) %>%
  group_by(bb.sp, plant.family) %>%
  summarize(freq = n()) %>%
  pivot_wider(names_from = bb.sp, values_from = freq) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("plant.family")

metaweb.rev <- net %>%
  separate(plant.sp, c("plant.genus", "species"), sep = " ") %>%
  left_join(taxa) %>%
  group_by(bb.sp, plant.family) %>%
  summarize(freq = n()) %>%
  pivot_wider(names_from = plant.family, values_from = freq) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("bb.sp")

bip_ggnet(bip_init_network(as.matrix(metaweb.rev)), 
                           as.matrix(metaweb.rev),
          edge.label = NULL,
          label= TRUE)

visweb(metaweb, type = "diagonal", labsize = 2, pred.lablength = 5, prey.lablength = 5)
visweb(metaweb.rev, type = "nested", labsize = 2, pred.lablength = 5, prey.lablength = 5)

pdf("./output/metaweb.pdf", width = 8.5, height = 4.5)
plotweb(metaweb, text.rot = 90, labsize = 1, low.lablength = 5)
dev.off()

ggplot(bb_range, aes(reorder(bb.sp, elev.med), elev.mean, color = abund.class)) +
  geom_line(size = 2, alpha = 0.25) +
  geom_point(aes(size = abund), alpha = 0.5) +
  #facet_wrap(~year, ncol = 1) +
  theme_light(12) +
  scale_size_continuous(name = "Abundance") +
  scale_color_discrete(name = "Abund. class") +
  xlab("BB species") +
  ylab("Elevation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("./output/bb_range.png", width = 6, height = 4, units = "in")
ggsave("./output/bb_range.pdf", width = 6, height = 4)

```

### Observations per sample per year
```{r}
bb_obs_per_year <- net %>%
  mutate(year = year(date)) %>%
  group_by(year) %>%
  summarize(obs = n())

bb_obs_per_sample <- net %>%
  mutate(year = factor(year(date))) %>%
  group_by(year, site, date) %>%
  summarize(obs = n()) %>%
  left_join(site_data) %>%
  mutate(date =as_date(yday(date)))

bb_samples_per_year <- net %>%
  mutate(year = year(date)) %>%
  select(year, site, date) %>%
  unique() %>%
  group_by(year) %>%
  summarize(samples = n())

bb_summary <- full_join(bb_obs_per_year, bb_samples_per_year) %>%
  mutate(obs.per.sample = obs/samples)
```

### Network observations through elevation and time
```{r}
obs_count_plot <- ggplot(net, aes(x = "x")) +
  geom_bar(fill = "gray60") +
  ylab("Observations") +
  facet_wrap(~year(date)) +
  theme_light(12) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

sample_count_plot <- ggplot(bb_obs_per_sample, aes(x = "x")) +
  geom_bar(fill = "gray60") +
  ylab("Samples") +
  facet_wrap(~year) +
  theme_light(12) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank())

obs_plot <- ggplot(bb_obs_per_sample, aes(x = "x", y = obs)) +
  geom_violin(draw_quantiles = 0.5, fill = "gray60", color = NA) +
  geom_boxplot(fill = NA) +
  #geom_density(outline.type = "full", position = "fill") +
  ylab("Observations/sample") +
  #xlab(NULL) +
  facet_wrap(~year) +
  theme_light(12) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank())

samples_plot <- ggplot(bb_obs_per_sample, aes(as_date(yday(date)), elev.mean/10)) +
  geom_point(size = 1, aes(alpha = obs)) +
  facet_wrap(~year) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  #scale_color_gradient(low = "gray80", high = "black") +
  scale_alpha_continuous(name = "Observations") +
  #scale_x_date(date_labels = "%m", date_breaks = "1 month", limit) +
  scale_x_date(date_labels = "%b", breaks = as_date(c("1970-04-01", "1970-05-01", "1970-06-01",
                                                    "1970-07-01", "1970-08-01", "1970-09-01"))) +
  theme_light(12) +
  labs(x = NULL,
       y = "Elevation (m x 10)") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank())

png("./output/sampling.png", width = 8, height = 8, units = "in", res = 300)
obs_count_plot / sample_count_plot / obs_plot / samples_plot + 
  plot_annotation(tag_levels = c("A", "B", "C", "D"))
dev.off()
```

### Survey data through elevation and time
```{r}
survey <- survey %>%
  #semi_join(net, by = c("plant.sp")) %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
  separate(plant.sp, c("plant.genus", "plant.sp"), sep = " ") %>%
  left_join(taxa)

survey_sub <- survey %>%
  semi_join(net, by = c("plant.sp")) %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
  separate(plant.sp, c("plant.genus", "plant.sp"), sep = " ") %>%
  left_join(taxa)

length(unique(survey_sub$plant.genus))

fl_obs_per_sample <- survey %>%
  mutate(year = factor(year(date))) %>%
  group_by(year, site, date) %>%
  summarize(obs = n()) %>%
  left_join(site_data) %>%
  mutate(date =as_date(yday(date)))

obs_count_plot <- ggplot(survey, aes(x = "x")) +
  geom_bar(fill = "gray60") +
  ylab("Observations") +
  facet_wrap(~year(date)) +
  theme_light(12) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

sample_count_plot <- ggplot(fl_obs_per_sample, aes(x = "x")) +
  geom_bar(fill = "gray60") +
  ylab("Samples") +
  facet_wrap(~year) +
  theme_light(12) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank())

obs_plot <- ggplot(fl_obs_per_sample, aes(x = "x", y = obs)) +
  geom_violin(draw_quantiles = 0.5, fill = "gray60", color = NA) +
  geom_boxplot(fill = NA) +
  #geom_density(outline.type = "full", position = "fill") +
  ylab("Observations/sample") +
  #xlab(NULL) +
  facet_wrap(~year) +
  theme_light(12) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank())

samples_plot <- ggplot(fl_obs_per_sample, aes(as_date(yday(date)), elev.mean/10)) +
  geom_point(size = 1, aes(alpha = obs)) +
  facet_wrap(~year) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  #scale_color_gradient(low = "gray80", high = "black") +
  scale_alpha_continuous(name = "Observations") +
  #scale_x_date(date_labels = "%m", date_breaks = "1 month", limit) +
  scale_x_date(date_labels = "%b", breaks = as_date(c("1970-04-01", "1970-05-01", "1970-06-01",
                                                    "1970-07-01", "1970-08-01", "1970-09-01"))) +
  theme_light(12) +
  labs(x = NULL,
       y = "Elevation (m x 10)") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank())

png("./output/sampling.png", width = 8, height = 8, units = "in", res = 300)
obs_count_plot / sample_count_plot / obs_plot / samples_plot + 
  plot_annotation(tag_levels = c("A", "B", "C", "D"))
dev.off()

```

### 2020-09-25 analyses/plotting_filter.R

@@ -1,0 +1,157 @@
+### Need to run outside main script to avoid memory crashes
+
+library(tidyverse)
+library(lubridate)
+
+site_data <- read_csv("./processed_data/site_data.csv") %>%
+  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")), # rename in English
+         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))
+
+net <- read_csv("./processed_data/network.csv") %>%
+  dplyr::select(site, date, bb.sp, plant.sp) # drop unused columns
+
+survey <- read_csv("./processed_data/floral_survey.csv") %>%
+  semi_join(net, by = "plant.sp") %>% # Consider only visited plant species (anytime anywhere, not specifically for a given site*date)
+  dplyr::select(site, date, plant.sp, flower.cover) %>%
+  group_by(plant.sp) %>%
+  mutate(present = case_when(
+    max(flower.cover) > 0 ~ TRUE,
+    max(flower.cover) == 0 ~ FALSE
+  )) %>%
+  filter(present == TRUE) 
+
+### This could be so much more beautiful, but I don't have it in me right now. This works.
+seq_yday <- function(x){
+  seq((x - 10), (x + 10), 0.5)
+}
+
+seq_elev <- function(x){
+  seq((x - 150), (x + 150), 1)
+}
+
+df <- function(x) {
+  as.data.frame(x)
+}
+
+survey_range.2010 <- survey %>%
+  ungroup() %>%
+  left_join(site_data) %>%
+  mutate(yday = yday(date),
+         year = year(date)) %>%
+  filter(year == 2010) %>%
+  select(yday, elev.mean) %>%
+  unique() %>%
+  arrange(yday) %>%
+  mutate(ydays = map(yday, seq_yday),
+         elevs = map(elev.mean, seq_elev)) %>%
+  mutate(ydays = map(ydays, df),
+         elevs = map(elevs, df)) %>%
+  unnest(cols = ydays) %>%
+  select(yday = x, elevs) %>%
+  unnest(cols = elevs) %>%
+  unique() %>%
+  select(yday, elev.mean = x) %>%
+  mutate(year = rep(2010, n()))
+
+survey_range.2011 <- survey %>%
+  ungroup() %>%
+  left_join(site_data) %>%
[…]

### 2020-09-25 analyses/trash_can.R

@@ -532,6 +532,157 @@ ggplot(net, aes(factor(pbl.w.class), corolla.depth)) +
   geom_boxplot() 
 ```
 
+### Make a data set that sets the boundaries for plotting predictions extrapolated (or interpolated) beyond samples
+```{r}
+### This could be so much more beautiful, but I don't have it in me right now. This works.
+seq_yday <- function(x){
+  seq((x - 14), (x + 14), 1) # This says I'm willing to go 10 days beyond my data...
+}
+
+seq_elev <- function(x){
+  seq((x - 250), (x + 250), 5) # ...and 150 meters beyond my data
+}
+
+df <- function(x) {
+  as.data.frame(x)
+}
+
+survey_range.2010 <- survey %>%
+  ungroup() %>%
+  left_join(site_data) %>%
+  mutate(yday = yday(date),
+         year = year(date)) %>%
+  filter(year == 2010) %>%
+  select(yday, elev.mean) %>%
+  unique() %>%
+  arrange(yday) %>%
+  mutate(ydays = map(yday, seq_yday),
+         elevs = map(elev.mean, seq_elev)) %>%
+  mutate(ydays = map(ydays, df),
+         elevs = map(elevs, df)) %>%
+  unnest(cols = ydays) %>%
+  select(yday = x, elevs) %>%
+  unnest(cols = elevs) %>%
+  unique() %>%
+  select(yday, elev.mean = x) %>%
+  mutate(year = rep(2010, n()))
+
+survey_range.2011 <- survey %>%
+  ungroup() %>%
+  left_join(site_data) %>%
+  mutate(yday = yday(date),
+         year = year(date)) %>%
+  filter(year == 2011) %>%
+  select(yday, elev.mean) %>%
+  unique() %>%
+  arrange(yday) %>%
+  mutate(ydays = map(yday, seq_yday),
+         elevs = map(elev.mean, seq_elev)) %>%
+  mutate(ydays = map(ydays, df),
+         elevs = map(elevs, df)) %>%
+  unnest(cols = ydays) %>%
+  select(yday = x, elevs) %>%
+  unnest(cols = elevs) %>%
+  unique() %>%
+  select(yday, elev.mean = x) %>%
+  mutate(year = rep(2011, n()))
+
+survey_range.2012 <- survey %>%
+  ungroup() %>%
+  left_join(site_data) %>%
+  mutate(yday = yday(date),
+         year = year(date)) %>%
[…]

### 2020-09-28 analyses/beta_diversity.Rmd
Updated beta div analysis

@@
   dplyr::select(site, date, plant.sp, flower.cover)
+```
+
+### Get beta-diversity for BBs and plants (not their interactions)
+```{r}
+bb_beta <- net %>%
+  mutate(yday = yday(date)) %>%
+  group_by(site, yday, year, bb.sp) %>%
+  summarize(count = n()) %>%
+  unite(id, c(site, yday, year), sep = "_") %>%
+  pivot_wider(names_from = bb.sp, values_from = count) %>%
+  replace(is.na(.), 0) %>%
+  column_to_rownames("id") %>%
+  vegdist(method = "bray", binary = TRUE) %>%
+  as.matrix() %>%
+  as.data.frame() %>%
+  rownames_to_column(var = "id1") %>%
+  pivot_longer(cols = -id1, names_to = "id2", values_to = "BB") %>%
+  separate(id1, into = c("site1", "yday1", "year1"), sep = "_") %>%
+  separate(id2, into = c("site2", "yday2", "year2"), sep = "_")
+
+fl_beta <- survey %>%
+  filter(flower.cover > 0) %>%
+  mutate(yday = yday(date),
+         year= year(date)) %>%
+  group_by(site, yday, year, plant.sp) %>%
+  summarize(count = sum(flower.cover)) %>%
+  unite(id, c(site, yday, year), sep = "_") %>%
+  pivot_wider(names_from = plant.sp, values_from = count) %>%
+  replace(is.na(.), 0) %>%
+  column_to_rownames("id") %>%
+  vegdist(method = "bray", binary = TRUE, upper = TRUE) %>%
+  as.matrix() %>%
+  as.data.frame() %>%
+  rownames_to_column(var = "id1") %>%
+  pivot_longer(cols = -id1, names_to = "id2", values_to = "FL") %>%
+  separate(id1, into = c("site1", "yday1", "year1"), sep = "_") %>%
+  separate(id2, into = c("site2", "yday2", "year2"), sep = "_")
+
+species_beta <- full_join(bb_beta, fl_beta) %>%
+  left_join(site_data, by = c("site1" = "site")) %>%
+  dplyr::select(site1, year1, yday1, site2, year2, yday2, FL, BB,
+                elev1 = elev.mean, x1 = x, y1 = y) %>%
+  left_join(site_data, by = c("site2" = "site")) %>%
+  dplyr::select(site1, year1, yday1, site2, year2, yday2, FL, BB,
+                elev1, elev2 = elev.mean, x1, y1, x2 = x, y2 = y) %>%
+  mutate(site1 = factor(site1),
+         site2 = factor(site2),
+         year1 = factor(year1),
+         year2 = factor(year2),
+         yday1 = as.numeric(yday1),
+         yday2 = as.numeric(yday2),
+         elev.diff = elev1 - elev2,
+         mean.elev = (elev1 + elev2)/2,
+         yday.diff = yday1 - yday2,
+         yday.diff.abs = abs(yday1 - yday2)) %>%
+  pivot_longer(cols = c(BB, FL), names_to = "group", values_to = "beta") %>%
+  filter(year1 == year2 &
+           elev.diff >= 0) # this simply deduplicates the data so that site1 vs site2 and site2 vs. site 1 are not both included; only works because elev.diff is unique for each site pair
 ```
@@
          mean.elev = (elev1 + elev2)/2,
-         yday.diff = abs(yday1 - yday2),
+         yday.diff.abs = abs(yday1 - yday2),
          man.diff = case_when(
@@
   gather(metric, value, -c(site1, year1, yday1, site2, year2, yday2,
-                           elev1, elev2, elev.diff, yday.diff, 
+                           elev1, elev2, elev.diff, yday.diff.abs, 
                            man1, man2, man.diff, mean.elev, 
@@
 # Set palette
-p <- c("OS" = "darkorange1", "ST" = "green3", 
+p <- c("OS" = "deepskyblue", "ST" = "green3", 
        "ST.h" = "firebrick1", "ST.l" = "darkorchid", 
-       "ST.lh" = "deepskyblue", "WN" = "black")
+       "ST.lh" = "darkorange1", "WN" = "black")
 
-# Elevation diff
-ggplot(filter(beta_mit_data, year1 == year2 &
+# BB + FL elevation diff
+g1 <- ggplot(filter(species_beta, site1 != site2 & yday.diff.abs <= 5), aes(elev.diff, beta)) +
+  geom_point(alpha = 0.05, pch = 16) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 6)) +
+  facet_grid(group ~ year1) +
+  xlab("Elevation difference (m)") +
+  ylab("β-diversity")
+
+# BB + FL time diff
+g2 <- ggplot(filter(species_beta, site1 == site2 & yday.diff > 0), aes(yday.diff.abs, beta)) +
+  geom_point(alpha = 0.05, pch = 16) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 6)) +
+  facet_grid(group ~ year1) +
+  xlab("Time difference (days)") +
+  ylab("β-diversity")
+
+# BB * FL elevation diff
+g3 <- ggplot(filter(beta_mit_data, year1 == year2 &
                       site1 != site2 &
                       metric != "S" & 
-                !metric.class %in% c("OS.partition", "WN.partition")), 
+                      !metric.class %in% c("OS.partition", "WN.partition") & 
+                      yday.diff.abs <= 5), 
              aes(elev.diff, value, color = metric)) +
-  geom_point(alpha = 0.25, pch = 16) +
-  geom_smooth() +
-  facet_grid(metric.class ~ year1) +
-  scale_color_manual(values = p) +
-  theme_abyss()
-
-# Time diff
-ggplot(filter(beta_mit_data, year1 == year2 &
-                site1 != site2 &
-                metric != "S" & 
-                !metric.class %in% c("OS.partition", "WN.partition")), 
-       aes(yday.diff, value, color = metric)) +
-  geom_point(alpha = 0.25, pch = 16) +
-  geom_smooth() +
-  facet_grid(metric.class ~ year1) +
-  scale_color_manual(values = p) +
-  theme_abyss()
-
-# Elevation diff within 1 wk time diff
-p1 <- ggplot(filter(beta_mit_data, year1 == year2 &
-                site1 != site2 &
-                metric != "S" & 
-                !metric.class %in% c("OS.partition", "WN.partition") 
-                #& yday.diff <= 7
-                ), 
-       aes(elev.diff, value, color = metric)) +
-  geom_point(alpha = 0.03, pch = 16, size = 0.25) +
+  #geom_point(alpha = 0.03, pch = 16, size = 0.25) +
+  geom_point(alpha = 0.05, pch = 16, size = 1) +
   geom_smooth(method = "gam", formula = y ~ s(x, k = 6)) +
-  facet_grid(metric.class ~ year1) +
+  #facet_grid(metric.class ~ year1) +
+  facet_wrap(~year1) +
   scale_color_manual(values = p, name = "partition") +
@@
   ylab("β-diversity")
-ggsave("./output/beta_elev.png", height = 4.5, width = 6.5, units = "in")
 
-# Time diff within 250 m elev diff
-p2 <- ggplot(filter(beta_mit_data, year1 == year2 &
-                site1 != site2 &
+# BB * FL time diff
+g4 <- ggplot(filter(beta_mit_data, year1 == year2 &
+                      site1 == site2 &
                       metric != "S" & 
-                !metric.class %in% c("OS.partition", "WN.partition")
-               # & elev.diff <= 250
-                ), 
-       aes(yday.diff, value, color = metric)) +
-  geom_point(alpha = 0.03, pch = 16, size = 0.25) +
+                      !metric.class %in% c("OS.partition", "WN.partition")), 
+             aes(yday.diff.abs, value, color = metric)) +
+  #geom_point(alpha = 0.03, pch = 16, size = 0.25) +
+  geom_point(alpha = 0.05, pch = 16, size = 1) +
   geom_smooth(method = "gam", formula = y ~ s(x, k = 6)) +
-  facet_grid(metric.class ~ year1) +
+  #facet_grid(metric.class ~ year1) +
+  facet_wrap(~year1) +
   scale_color_manual(values = p, name = "partition") +
@@
   theme(legend.position="none")
-ggsave("./output/beta_time.png", height = 4.5, width = 6.5, units = "in")
 
-png("./output/beta_stack.png", width = 6.5, height = 6.5, units = "in", res = 300)
-p1 / p2 + plot_annotation(tag_levels = c("A", "B", "C")) + plot_layout(guides = "collect")
+png("./output/beta_stack.png", width = 9, height = 6.5, units = "in", res = 300)
+(g1 + g3) / (g2 + g4) + 
+  plot_annotation(tag_levels = c("A", "B", "C", "D")) + 
+  plot_layout(guides = "collect") & theme_light()
 dev.off()
@@
     x.var == "yday" ~ "Day of Year"),
-    x.var = factor(x.var, levels = c("Geographic (km)", "Elevation (m)", "Day of Year"))
+    x.var = factor(x.var, levels = c("Geographic (km)", "Elevation (m)", "Day of Year")),
+    subject = rep("BB*FL", n())
     )
@@
     x.var == "yday" ~ "Day of Year"),
-    x.var = factor(x.var, levels = c("Geographic (km)", "Elevation (m)", "Day of Year"))
+    x.var = factor(x.var, levels = c("Geographic (km)", "Elevation (m)", "Day of Year")),
+    subject = rep("FL", n())
     )
@@
 
-### GDM: BB observations; this won't work in it's present state because I cannot 
+### GDM: BB observations
 ```{r}
@@
     x.var == "yday" ~ "Day of Year"),
-    x.var = factor(x.var, levels = c("Geographic (km)", "Elevation (m)", "Day of Year"))
+    x.var = factor(x.var, levels = c("Geographic (km)", "Elevation (m)", "Day of Year")),
+    subject = rep("BB", n())
     )
@@
 ```
+
+### GDM all together now
+```{r}
+splines <- bind_rows(interaction_splines, bb_splines, survey_splines) %>%
+  mutate(subject = factor(subject))
+
+ggplot(splines, aes(predictor, response, color = subject)) +
+  geom_line() +
+  facet_grid(year ~ x.var, scales = "free_x") +
+  xlab(NULL) +
+  ylab("f(x)") +
+  theme_light(12)
+ggsave("./output/GDM.png")
+```

### 2020-09-29 analyses/beta_diversity.Rmd
tabulatin' stuff

--- 2020-09-17 analyses/beta_diversity.Rmd
@@
 ```{r message=FALSE, warning=FALSE}
-library(lubridate)
 library(sf)
 library(gdm)
+library(patchwork)
 library(bipartite)
 library(vegan)
-library(mgcv)
-library(mgcViz)
 library(see)
 library(tidyverse)
+library(lubridate)
 ```
 
-### Load site data
-```{r, include=FALSE, echo=FALSE}
+### Create functions for later
+```{r}
 # A function for converting a geometry column back to x and y columns
@@
 
+# Function to convert networks to webs arrays
+web_prep <- function(net, y) {
+  out <- net %>%
+    filter(year == y) %>%
+    unite(site.date, c(site, date), sep = "_") %>%
+  group_by(site.date, bb.sp, plant.sp) %>%
+  summarize(freq = n()) %>%
+  group_by(site.date) %>%
+  mutate(bb.sp.count = length(unique(bb.sp)),
+         plant.sp.count = length(unique(plant.sp))) %>%
+  ungroup() %>%
+  filter(bb.sp.count > 1 & plant.sp.count > 1) %>% # the betalinkr function fails if you feed it webs with only 1 species in either (or perhaps both) of the levels; this is probably for good reason.
+  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.date, freq) %>% # rename columns to match bipartite's expectations
+  data.frame() %>% # convert to data frame
+  frame2webs(type.out = "array")
+}
+
+# Function with betalinkr settings and conversion to tibble()
+betalinkr_set1 <- function(x) {
+  betalinkr_multi(x, partition.st = TRUE, partition.rr = TRUE, # partition both species turnover and richness/replacement
+                  binary = TRUE, # treating beta-diversity as binary is an appropriate complement to our abundance analysis
+                  index = "jaccard") %>% # the Jaccard index is best, I think, for binary analysis
+  as_tibble() # for convenience
+}
+
+# Binomial smooth function (thanks, Hadley): https://ggplot2.tidyverse.org/reference/geom_smooth.html
+binomial_smooth <- function(...) {
+  geom_smooth(method = "glm", method.args = list(family = "binomial"), ...)
+}
+
+# Site_pair prep function
+# site_pair_prep_net <- function(dat, y) {
+#   dat %>%
+#     ungroup() %>%
+#     filter(year1 == year2 & 
+#              year1 == y & 
+#              site1 != site2) %>% # I don't think we want to compare samples through within sites, introduces a new form of non-independence.
+#     filter(metric == "WN") %>% # we'll focus for now on WN, but this could be done for partitions, too
+#     mutate(weights = rep(1, dim(.)[1])) %>% # right now this analysis is unweighted; could weight by abundance, etc, if we wanted to
+#     dplyr::select(distance = value, weights,
+#                   s1.xCoord = x1, s1.yCoord = y1, 
+#                   s2.xCoord = x2, s2.yCoord = y2, 
+#                   s1.elev.mean = elev1, s1.yday = yday1,
+#                   s2.elev.mean = elev2, s2.yday = yday2) %>%
+#     data.frame()
+# }
+
+interaction_gdm_prep <- function(x, y) {
+
+  spdat <- net %>%
+    filter(year == y) %>% # filter to selected year
+    select(bb.sp, plant.sp.abb, site, date) %>% 
+    distinct() %>%# each unique bb-fl-site-date combo
+    mutate(present = rep(1, length(.[,1]))) %>% # we will treat abundance as binary
+    unite(interaction, c(bb.sp, plant.sp.abb), sep = "_") %>% # unite bb and fl into interaction "species", i.e. a unique bb-fl combo
+    unite(site.date, c("site", "date"), sep = "_") %>% # unite site and date into a sample id
+    pivot_wider(names_from = interaction, values_from = present) %>% # spread so that columns are interaction "species"
+    replace(is.na(.), 0) %>% # replace NAs with absences
+    as_tibble()
+
+  envdat <- net %>%
+    filter(year == y) %>% # filter to selected year
+    select(bb.sp, plant.sp.abb, site, date) %>% #
+    mutate(yday = yday(date)) %>%
+    left_join(site_data) %>% # add site data
+    unite(site.date, c("site", "date"), sep = "_") %>% # unite site and date
+    select(site.date, elev.mean, yday, x, y) %>%
+    distinct() # return unique site-dates
+
+  # Prep site-pairs for GDM analysis
+  sitepair <- formatsitepair(spdat, 
+                             bioFormat = 1, # i.e. samples as rows, species as columns, site-dates as rows
+                             siteColumn = "site.date", 
+                             dist = "jaccard", # I think Jaccard is the right choice
+                             XColumn = "x", YColumn = "y", # spatial columns
+                             predData = envdat, 
+                             abundance = FALSE) %>% # abundance = FALSE means binary
+    filter(s1.elev.mean != s2.elev.mean) # this drops within-site comparisons
+}
+
+## wrapper for extracting splines and prepping tibble for plotting
+spline_prep <- function(x, y) {
+  isplineExtract(x) %>%
+    as.data.frame() %>%
+    tibble() %>%
+    mutate(x.Geographic = x.Geographic/1000) %>%
+    pivot_longer(cols = c(x.Geographic, x.elev.mean, x.yday), 
+                 names_to = "x.var", values_to = "predictor", 
+                 names_prefix = "x.") %>%
+    pivot_longer(cols = c(y.Geographic, y.elev.mean, y.yday), 
+                 names_to = "y.var", values_to = "response", 
+                 names_prefix = "y.") %>%
+    filter(x.var == y.var) %>%
+    mutate() %>%
+    dplyr::select(-y.var) %>%
+    mutate(year = factor(rep(y, dim(.)[1])))
+}
+
+# Create functions for prepping data
+biodata_prep_survey <- function(surv, y) {
+  surv %>%
+    mutate(year = year(date),
+           yday = yday(date)) %>%
+    filter(flower.cover > 0 & year == y) %>%
+    mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
+    group_by(site, yday, plant.sp) %>%
+    summarize(cover = sum(flower.cover)) %>%
+    mutate(cover = rep(1, n())) %>% # this makes the data binary, which I think is what I want here; otherwise, just comment out
+    spread(plant.sp, cover) %>%
+    replace(is.na(.), 0) %>%
+    arrange(site, yday) %>%
+    unite(site.day, site, yday, sep = "_")
+}
+
+predData_prep_survey <- function(surv, y) {
+  surv %>%
+    mutate(year = year(date),
+           yday = yday(date)) %>%
+    filter(flower.cover > 0 & year == y) %>%
+    dplyr::select(site, yday) %>%
+    left_join(site_data) %>%
+    ungroup() %>%
+    dplyr::select(site, elev.mean, y, x, yday) %>%
+    unique() %>%
+    arrange(site, yday) %>%
+    unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
+    dplyr::select(-site)
+}
+
+format_set <- function(bio, pred) {
+  formatsitepair(bio, bioFormat = 1, siteColumn = "site.day", 
+                 dist = "jaccard", XColumn = "x", YColumn = "y", 
+                 predData = pred, abundance = FALSE) %>%
+  filter(s1.elev.mean != s2.elev.mean) # this avoids comparing repeated measures of the same site (taking advantage of the fact that each site has a unique elevation in our dataset)
+}
+
+biodata_prep_BB <- function(net, y) {
+  net %>%
+    filter(year == y) %>%
+    group_by(site, date, bb.sp) %>%
+    summarize(count = n()) %>%
+    mutate(count = rep(1, n())) %>% # this makes the data binary, which I think is what I want here; otherwise, just comment out
+    spread(bb.sp, count) %>%
+    replace(is.na(.), 0) %>%
+    mutate(yday = yday(date)) %>%
+    unite(site.day, site, yday, sep = "_", remove = "TRUE") %>%
+    ungroup() %>%
+    dplyr::select(-date)
+}
+
+predData_prep_BB <- function(net, y) {
+  net %>%
+    filter(year == y) %>%
+    mutate(yday = yday(date)) %>%
+    dplyr::select(site, yday) %>%
+    left_join(site_data) %>%
+    ungroup() %>%
+    dplyr::select(site, elev.mean, y, x, yday) %>%
+    unique() %>%
+    arrange(site, yday) %>%
+    unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
+    dplyr::select(-site)
+}
+```
+
+### Load site data
+```{r, include=FALSE, echo=FALSE}
 site_data <- read_csv("../processed_data/site_data.csv") %>%
@@
 survey <- read_csv("../processed_data/floral_survey.csv") %>%
+  filter(flower.cover > 0) %>%
   semi_join(net, by = "plant.sp") %>% # Consider only visited plant spp. (anytime, anywhere, not specifically for a given site*date)
   dplyr::select(site, date, plant.sp, flower.cover)
+```
+
+### Get beta-diversity for BBs and plants (not their interactions)
+```{r}
+# Bumble bee beta-diversity based on interaction sampling
+bb_beta <- net %>%
+  mutate(yday = yday(date)) %>% 
+  group_by(site, yday, year, bb.sp) %>%
+  summarize(count = n()) %>% 
+  unite(id, c(site, yday, year), sep = "_") %>%
+  pivot_wider(names_from = bb.sp, values_from = count) %>%
+  replace(is.na(.), 0) %>%
+  column_to_rownames("id") %>%
+  vegdist(method = "jaccard", binary = TRUE) %>% # Jaccard and binary settings make this comparable to the way I'm running betalinkr
+  as.matrix() %>%
+  as.data.frame() %>%
+  rownames_to_column(var = "id1") %>%
+  pivot_longer(cols = -id1, names_to = "id2", values_to = "BB") %>%
+  separate(id1, into = c("site1", "yday1", "year1"), sep = "_") %>%
+  separate(id2, into = c("site2", "yday2", "year2"), sep = "_")
+
+# Flora beta-diversity based on floral survey data (filtered to include only species visited by BBs)
+fl_beta <- survey %>%
+  mutate(yday = yday(date),
+         year= year(date)) %>%
+  group_by(site, yday, year, plant.sp) %>%
+  summarize(count = sum(flower.cover)) %>%
+  unite(id, c(site, yday, year), sep = "_") %>%
+  pivot_wider(names_from = plant.sp, values_from = count) %>%
+  replace(is.na(.), 0) %>%
+  column_to_rownames("id") %>%
+  vegdist(method = "jaccard", binary = TRUE, upper = TRUE) %>%
+  as.matrix() %>%
+  as.data.frame() %>%
+  rownames_to_column(var = "id1") %>%
+  pivot_longer(cols = -id1, names_to = "id2", values_to = "FL") %>%
+  separate(id1, into = c("site1", "yday1", "year1"), sep = "_") %>%
+  separate(id2, into = c("site2", "yday2", "year2"), sep = "_")
+
+# Join into single data frame for visualization
+species_beta <- full_join(bb_beta, fl_beta) %>%
+  left_join(site_data, by = c("site1" = "site")) %>% # add site data
+  dplyr::select(site1, year1, yday1, site2, year2, yday2, FL, BB, # first for site 1
+                elev1 = elev.mean, x1 = x, y1 = y) %>%
+  left_join(site_data, by = c("site2" = "site")) %>%
+  dplyr::select(site1, year1, yday1, site2, year2, yday2, FL, BB, # then for site 2
+                elev1, elev2 = elev.mean, x1, y1, x2 = x, y2 = y) %>%
+  mutate(site1 = factor(site1), 
+         site2 = factor(site2),
+         year1 = factor(year1),
+         year2 = factor(year2),
+         yday1 = as.numeric(yday1),
+         yday2 = as.numeric(yday2),
+         elev.diff = elev1 - elev2,
+         mean.elev = (elev1 + elev2)/2,
+         yday.diff = yday1 - yday2, 
+         yday.diff.abs = abs(yday1 - yday2)) %>%
+  pivot_longer(cols = c(BB, FL), names_to = "group", values_to = "beta") %>%
+  filter(year1 == year2 & # we don;t want to compare across years
+           elev.diff >= 0) # For each site pair, one will be higher and the other lower; thus, when their order is reversed only the sign of the difference changes.
+Since no two sites have exactly the same elevation, this deduplicates the data by dropping the ordering of a given pair of sites (one of the two orderings) that yields a negative elevation difference while retaining the alternative ordering that yields a positive elevation difference.
+By >= 0, I retain comparisons of a site to itself.
+I drop (in the ggplot call) within-site comparisons when I plot against elevation difference, and I isolate within-site comparison when I plot against time.
+In the latter case, it is also necessary to deduplicate, since each pair of site dates occurs in A-B and B-A.
+For that, I filter (again, in the ggplot call) so that yday.diff > 0, similarly dropping the orderings that yield negative differences.
+Note that all this deduplication nonsense is not necessary for the betalinkr call because it deduplicates internally.
 ```
@@
 ```{r message=FALSE}
-# Function to convert networks to webs arrays
-web_prep <- function(net, y) {
-  out <- net %>%
-    filter(year == y) %>%
-    unite(site.date, c(site, date), sep = "_") %>%
-  group_by(site.date, bb.sp, plant.sp) %>%
-  summarize(freq = n()) %>%
-  group_by(site.date) %>%
-  mutate(bb.sp.count = length(unique(bb.sp)),
-         plant.sp.count = length(unique(plant.sp))) %>%
-  ungroup() %>%
-  filter(bb.sp.count > 1 & plant.sp.count > 1) %>% # the betalinkr function fails if you feed it webs with only 1 species in either (or perhaps both) of the levels; this is probably for good reason.
-  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.date, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs(type.out = "array")
-}
-
 # Create webs
@@
 net2012 <- web_prep(net, 2012)
-
-# Function with betalinkr settings and conversion to tibble()
-betalinkr_set1 <- function(x) {
-  betalinkr_multi(x, partition.st = TRUE, partition.rr = TRUE, # partition both species turnover and richness/replacement
-                  binary = TRUE, # treating beta-diversity as binary is an appropriate complement to our abundance analysis
-                  index = "jaccard") %>% # the Jaccard index is best, I think, for binary analysis
-  as_tibble() # for convenience
-}
 
@@
                                        "ST.lh", "WN.repl", "WN.rich", 
-                                       "OS.repl", "OS.rich"))
-    )
+                                       "OS.repl", "OS.rich"))) %>%
+  filter(year1 == year2) # I am only interested in within-year comparisons at the moment
 ```
@@
 ### Plot interaction beta-diversity against elevation difference and time difference
+#### For elevation difference, we will "control" for time by only considering pairs separated by <= 5 days
+#### For time difference, we will control for elevation by only comparing within site.
 ```{r}
 # Set palette
-p <- c("OS" = "darkorange1", "ST" = "green3", 
+p <- c("OS" = "deepskyblue", "ST" = "green3", 
        "ST.h" = "firebrick1", "ST.l" = "darkorchid", 
-       "ST.lh" = "deepskyblue", "WN" = "black")
+       "ST.lh" = "darkorange1", "WN" = "black")
 
-# Elevation diff
-ggplot(filter(beta_mit_data, year1 == year2 &
-                site1 != site2 &
-                metric != "S" & 
-                !metric.class %in% c("OS.partition", "WN.partition")), 
+# BB + FL elevation diff
+g1 <- ggplot(filter(species_beta, 
+                    site1 != site2 & # no within-site comparisons
+                      yday.diff.abs <= 5), # only site-dates within 5 days
+             aes(elev.diff, beta)) +
+  geom_point(alpha = 0.05, pch = 16) +
+  binomial_smooth(formula = y ~ splines::ns(x, 3)) +
+  facet_wrap(~group, ncol = 1) +
+  xlab("Elevation difference (m)") +
+  ylab("β-diversity")
+
+# BB + FL time diff
+g2 <- ggplot(filter(species_beta, 
+                    site1 == site2 & # ONLY within-site comparisons
+                      yday.diff > 0), # deduplicate by date
+             aes(yday.diff.abs, beta)) +
+  geom_point(alpha = 0.05, pch = 16) +
+  binomial_smooth(formula = y ~ splines::ns(x, 3)) +
+  facet_wrap(~group, ncol = 1) +
+  xlab("Time difference (days)") +
+  ylab("β-diversity")
+
+# BB * FL elevation diff
+g3 <- ggplot(filter(beta_mit_data, 
+                      site1 != site2 & # no within-site comparisons
+                      metric != "S" & # drop this metric
+                      !metric.class %in% c("OS.partition", "WN.partition") & # drop this whole metric class
+                      yday.diff.abs <= 5), # only site-dates within 5 days 
              aes(elev.diff, value, color = metric)) +
-  geom_point(alpha = 0.25, pch = 16) +
-  geom_smooth() +
-  facet_grid(metric.class ~ year1) +
-  scale_color_manual(values = p) +
-  theme_abyss()
+  binomial_smooth(formula = y ~ splines::ns(x, 3)) +
+  #facet_wrap(~year1) +
+  scale_color_manual(values = , name = "partition") +
+  #scale_color_brewer(type = "qual", name = "partition") +
+  xlab("Elevation difference (m)") +
+  ylab("β-diversity")
 
-# Time diff
-ggplot(filter(beta_mit_data, year1 == year2 &
-                site1 != site2 &
-                metric != "S" & 
-                !metric.class %in% c("OS.partition", "WN.partition")), 
-       aes(yday.diff, value, color = metric)) +
-  geom_point(alpha = 0.25, pch = 16) +
-  geom_smooth() +
-  facet_grid(metric.class ~ year1) +
-  scale_color_manual(values = p) +
-  theme_abyss()
+# BB * FL time diff
+g4 <- ggplot(filter(beta_mit_data, 
+                      site1 == site2 & # ONLY sithin-site comparisons
+                      metric != "S" & # drop this metric
+                      !metric.class %in% c("OS.partition", "WN.partition")),  # drop this whole metric class
+             aes(yday.diff.abs, value, color = metric)) +
+  #geom_point(alpha = 0.03, pch = 16, size = 0.25) +
+  #geom_point(alpha = 0.05, pch = 16, size = 1) +
+  #geom_smooth(method = "gam", formula = y ~ s(x, k = 6)) +
+  binomial_smooth(formula = y ~ splines::ns(x, 3)) +
+  #facet_grid(metric.class ~ year1) +
+  #facet_wrap(~year1) +
+  scale_color_manual(values = p, name = "partition") +
+  xlab("Time difference (days)") +
+  ylab("β-diversity") + 
+  theme(legend.position="none")
 
-# Elevation diff within 1 wk time diff
-ggplot(filter(beta_mit_data, year1 == year2 &
-                site1 != site2 &
-                metric != "S" & 
-                !metric.class %in% c("OS.partition", "WN.partition") &
-                yday.diff <= 7), 
-       aes(elev.diff, value, color = metric)) +
-  geom_point(alpha = 0.1, pch = 16, size = 1) +
-  geom_smooth() +
-  facet_grid(metric.class ~ year1) +
-  scale_color_manual(values = p, name = "partition") +
-  xlab("Elevation difference") +
-  ylab("β-diversity")
-ggsave("./output/beta_elev.png", height = 4.5, width = 6.5, units = "in")
+# Patchwork plot for comparing species beta-diversity with interaction beta-diversity
+png("./output/beta_stack.png", width = 9, height = 6.5, units = "in", res = 300)
+(g1 | g2) / (g3 | g4) + 
+  plot_annotation(tag_levels = c("A", "B", "C", "D")) + 
+  plot_layout(guides = "collect") & 
+  theme_light() 
+dev.off()
 
-# Time diff within 250 m elev diff
-ggplot(filter(beta_mit_data, year1 == year2 &
-                site1 != site2 &
-                metric != "S" & 
-                !metric.class %in% c("OS.partition", "WN.partition") &
-                elev.diff <= 250), 
-       aes(yday.diff, value, color = metric)) +
-  geom_point(alpha = 0.1, pch = 16, size = 1) +
-  geom_smooth() +
-  facet_grid(metric.class ~ year1) +
-  scale_color_manual(values = p, name = "partition") +
-  xlab("Elevation difference") +
-  ylab("β-diversity")
-ggsave("./output/beta_time.png", height = 4.5, width = 6.5, units = "in")
+# Histogram of interaction beta-diversity comparable to the elevation comparison (i.e. excluding within-site comparisons)
+WN.hist <- ggplot(filter(beta_mit_data,
+                           #site1 != site2 &
+                           metric == "WN"), 
+                aes(value)) +
+  geom_histogram() +
+  #scale_y_continuous(trans="pseudo_log", breaks = c(0, 10, 100, 1000, 10000)) +
+  #facet_wrap(~year1) +
+  xlab("Total β-diversity") +
+  ylab("Site-date pairs") +
+  theme_light()
+
+BB.hist <- ggplot(filter(species_beta,
+                           #site1 != site2 &
+                           group == "BB"), 
+                aes(beta)) +
+  geom_histogram() +
+  #facet_wrap(~year1) +
+  xlab("Bumble bee β-diversity") +
+  ylab("Site-date pairs") +
+  theme_light()
+
+FL.hist <- ggplot(filter(species_beta,
+                           #site1 != site2 &
+                           group == "FL"), 
+                aes(beta)) +
+  geom_histogram() +
+  #facet_wrap(~year1) +
+  xlab("Floral β-diversity") +
+  ylab("Site-date pairs") +
+  theme_light()
+
+# Patchwork plot
+png("./output/beta_histograms.png", width = 6.5, height = 9, units = "in", res = 300)
+BB.hist / FL.hist / WN.hist & theme_light(14)
+dev.off()
 ```
@@
 ```{r message=FALSE, warning=FALSE}
-# Site_pair prep function
-site_pair_prep_net <- function(dat, y) {
-  beta_mit_data %>%
-    ungroup() %>%
-    filter(year1 == year2 & 
-             year1 == y & 
-             site1 != site2) %>% # I don't think we want to compare samples through within sites, introduces a new form of non-independence.
-    filter(metric == "WN") %>% # we'll focus for now on WN, but this could be done for partitions, too
-    mutate(weights = rep(1, n())) %>% # right now this analysis is unweighted; could weight by abundance, etc, if we wanted to
-    dplyr::select(distance = value, weights,
-                  s1.xCoord = x1, s1.yCoord = y1, 
-                  s2.xCoord = x2, s2.yCoord = y2, 
-                  s1.elev.mean = elev1, s1.yday = yday1,
-                  s2.elev.mean = elev2, s2.yday = yday2) %>%
-    data.frame()
-}
-
-# Prep site-pairs for GDM analysis
-sitepair_2010.WN <- site_pair_prep_net(beta_mit_data, 2010)
-sitepair_2011.WN <- site_pair_prep_net(beta_mit_data, 2011)
-sitepair_2012.WN <- site_pair_prep_net(beta_mit_data, 2012)
+# Prep sitepairs
+sitepair_2010.WN <- interaction_gdm_prep(net, 2010)
+sitepair_2011.WN <- interaction_gdm_prep(net, 2011)
+sitepair_2012.WN <- interaction_gdm_prep(net, 2012)
 
@@
 gdm_WN.2010.imp <- gdm.varImp(sitepair_2010.WN, geo = TRUE, 
-                              parallel = TRUE, cores = 4,
-                              outFile = "interaction_permGDM_2010")
+                              parallel = TRUE, cores = 2, # I think parallelization is a good idea (?)
+                              sampleSitePairs = 0.5, # this randomly subsamples 50% of the site-pairs; a necessary sacrifice to get this to run smoothly, particularly for 2012
+                              outFile = "./output/interaction_permGDM_2010") # save the output
 gdm_WN.2011.imp <- gdm.varImp(sitepair_2011.WN, geo = TRUE, 
-                              parallel = TRUE, cores = 4,
-                              outFile = "interaction_permGDM_2011")
+                              parallel = TRUE, cores = 2,
+                              sampleSitePairs = 0.5,
+                              outFile = "./output/interaction_permGDM_2011")
 gdm_WN.2012.imp <- gdm.varImp(sitepair_2012.WN, geo = TRUE, 
-                              parallel = TRUE, cores = 4,
-                              outFile = "interaction_permGDM_2012")
+                              parallel = TRUE, cores = 2,
+                              sampleSitePairs = 0.5, # it keep crashing...
+                              outFile = "./output/winteraction_permGDM_2012")
 
@@
 # Plot with confidence intervals (slow; also, I'm not sure about the current settings)
-plotUncertainty(sitepair_2010.WN, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
-plotUncertainty(sitepair_2011.WN, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
-plotUncertainty(sitepair_2012.WN, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
+plotUncertainty(sitepair_2010.WN, sampleSites = 0.5, bsIters = 10, geo  = TRUE)
+plotUncertainty(sitepair_2011.WN, sampleSites = 0.5, bsIters = 10, geo  = TRUE)
+plotUncertainty(sitepair_2012.WN, sampleSites = 0.5, bsIters = 10, geo  = TRUE)
 
 # Plot nicely
-## wrapper for extracting splines and prepping tibble for plotting
-spline_prep <- function(x, y) {
-  isplineExtract(x) %>%
-    as.data.frame() %>%
-    tibble() %>%
-    mutate(x.Geographic = x.Geographic/1000) %>%
-    pivot_longer(cols = c(x.Geographic, x.elev.mean, x.yday), 
-                 names_to = "x.var", values_to = "predictor", 
-                 names_prefix = "x.") %>%
-    pivot_longer(cols = c(y.Geographic, y.elev.mean, y.yday), 
-                 names_to = "y.var", values_to = "response", 
-                 names_prefix = "y.") %>%
-    filter(x.var == y.var) %>%
-    mutate() %>%
-    dplyr::select(-y.var) %>%
-    mutate(year = factor(rep(y, n())))
-}
-
 ## extract splines
@@
     x.var == "Geographic" ~ "Geographic (km)",
-    x.var == "yday" ~ "Day of Year"),
-    x.var = factor(x.var, levels = c("Geographic (km)", "Elevation (m)", "Day of Year"))
-    )
-
-## plot
-ggplot(interaction_splines, aes(predictor, response)) +
-  geom_line() +
-  facet_grid(year ~ x.var, scales = "free_x") +
-  xlab(NULL) +
-  ylab("f(x)") +
-  theme_gray(12)
-ggsave("./output/interaction_GDM.png")
+    x.var == "yday" ~ "Day of Year")) %>%
+  mutate(x.var = factor(x.var, levels = c("Geographic (km)", "Elevation (m)", "Day of Year")),
+    subject = rep("BB*FL", dim(.)[1]))
 ```
@@
 ```{r}
-# Create functions for prepping data
-biodata_prep_survey <- function(surv, y) {
-  surv %>%
-    mutate(year = year(date),
-           yday = yday(date)) %>%
-    filter(flower.cover > 0 & year == y) %>%
-    mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
-    group_by(site, yday, plant.sp) %>%
-    summarize(cover = sum(flower.cover)) %>%
-    mutate(cover = rep(1, n())) %>% # this makes the data binary, which I think is what I want here; otherwise, just comment out
-    spread(plant.sp, cover) %>%
-    replace(is.na(.), 0) %>%
-    arrange(site, yday) %>%
-    unite(site.day, site, yday, sep = "_")
-}
-
-predData_prep_survey <- function(surv, y) {
-  surv %>%
-    mutate(year = year(date),
-           yday = yday(date)) %>%
-    filter(flower.cover > 0 & year == y) %>%
-    dplyr::select(site, yday) %>%
-    left_join(site_data) %>%
-    ungroup() %>%
-    dplyr::select(site, elev.mean, y, x, yday) %>%
-    unique() %>%
-    arrange(site, yday) %>%
-    unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
-    dplyr::select(-site)
-}
-
-format_set <- function(bio, pred) {
-  formatsitepair(bio, bioFormat = 1, siteColumn = "site.day", 
-                 dist = "jaccard", XColumn = "x", YColumn = "y", 
-                 predData = pred, abundance = FALSE) %>%
-  filter(s1.elev.mean != s2.elev.mean) # this avoids comparing repeated measures of the same site (taking advantage of the fact that each site has a unique elevation in our dataset)
-}
-
 # Prep input for GDM
@@
 gdm_survey.2010.imp <- gdm.varImp(survey_sitepair.2010, geo = TRUE, 
-                              parallel = TRUE, cores = 3,
+                              parallel = TRUE, cores = 2,
+                              sampleSitePairs = 0.5,
                               outFile = "survey_permGDM_2010")
 gdm_survey.2011.imp <- gdm.varImp(survey_sitepair.2011, geo = TRUE, 
-                              parallel = TRUE, cores = 3,
+                              parallel = TRUE, cores = 2,
+                              sampleSitePairs = 0.5,
                               outFile = "survey_permGDM_2011")
 gdm_survey.2012.imp <- gdm.varImp(survey_sitepair.2012, geo = TRUE, 
-                              parallel = TRUE, cores = 3,
+                              parallel = TRUE, cores = 2,
+                              sampleSitePairs = 0.5,
                               outFile = "survey_permGDM_2012")
@@
 
-### GDM: BB observations; this won't work in it's present state because I cannot 
+### GDM: BB observations
 ```{r}
-# Create functions for prepping data
-biodata_prep_BB <- function(net, y) {
-  net %>%
-    filter(year == y) %>%
-    group_by(site, date, bb.sp) %>%
-    summarize(count = n()) %>%
-    mutate(count = rep(1, n())) %>% # this makes the data binary, which I think is what I want here; otherwise, just comment out
-    spread(bb.sp, count) %>%
-    replace(is.na(.), 0) %>%
-    mutate(yday = yday(date)) %>%
-    unite(site.day, site, yday, sep = "_", remove = "TRUE") %>%
-    ungroup() %>%
-    dplyr::select(-date)
-}
-
-predData_prep_BB <- function(net, y) {
-  net %>%
-    filter(year == y) %>%
-    mutate(yday = yday(date)) %>%
-    dplyr::select(site, yday) %>%
-    left_join(site_data) %>%
-    ungroup() %>%
-    dplyr::select(site, elev.mean, y, x, yday) %>%
-    unique() %>%
-    arrange(site, yday) %>%
-    unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
-    dplyr::select(-site)
-}
-
 # Prep input for GDM
@@
 gdm_BB.2010.imp <- gdm.varImp(BB_sitepair.2010, geo = TRUE, 
-                              parallel = TRUE, cores = 4,
+                              parallel = TRUE, cores = 2,
+                              sampleSitePairs = 0.5,
                               outFile = "bb_permGDM_2010")
 gdm_BB.2011.imp <- gdm.varImp(BB_sitepair.2011, geo = TRUE, 
-                              parallel = TRUE, cores = 4,
+                              parallel = TRUE, cores = 2,
+                              sampleSitePairs = 0.5,
                               outFile = "bb_permGDM_2011")
 gdm_BB.2012.imp <- gdm.varImp(BB_sitepair.2012, geo = TRUE, 
-                              parallel = TRUE, cores = 4,
+                              parallel = TRUE, cores = 2,
+                              sampleSitePairs = 0.5,
                               outFile = "bb_permGDM_2012")
@@
 # Plot with confidence intervals (slow)
-p <- plotUncertainty(BB_sitepair.2010, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
-plotUncertainty(BB_sitepair.2011, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
-plotUncertainty(BB_sitepair.2012, sampleSites = 0.7, bsIters = 10, geo  = TRUE)
+plotUncertainty(BB_sitepair.2010, sampleSites = 0.5, bsIters = 10, geo  = TRUE)
+plotUncertainty(BB_sitepair.2011, sampleSites = 0.5, bsIters = 10, geo  = TRUE)
+plotUncertainty(BB_sitepair.2012, sampleSites = 0.25, bsIters = 10, geo  = TRUE)
 
@@
 ```
+
+### Tabulate model output
+```{r}
+####### survey
+gdm_survey.2010.tab1 <- gdm_survey.2010.imp[1] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble() %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2) %>%
+  mutate(subject = "survey",
+         year = 2010) %>%
+  select(subject, year, 
+         mod.dev = `Model deviance`, 
+         dev.exp = `Percent deviance explained`, 
+         mod.pval = `Model p-value`)
+
+gdm_survey.2010.tab2 <- gdm_survey.2010.imp[2] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble() %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2) %>%
+  select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)
+
+gdm_survey.2010.tab3 <- gdm_survey.2010.imp[3] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble()  %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2)  %>%
+  select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)
+
+gdm_survey.2011.tab1 <- gdm_survey.2011.imp[1] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble() %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2) %>%
+  mutate(subject = "survey",
+         year = 2011) %>%
+  select(subject, year, 
+         mod.dev = `Model deviance`, 
+         dev.exp = `Percent deviance explained`, 
+         mod.pval = `Model p-value`)
+
+gdm_survey.2011.tab2 <- gdm_survey.2011.imp[2] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble() %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2) %>%
+  select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)
+
+gdm_survey.2011.tab3 <- gdm_survey.2011.imp[3] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble()  %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2)  %>%
+  select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)
+
+gdm_survey.2012.tab1 <- gdm_survey.2012.imp[1] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble() %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2) %>%
+  mutate(subject = "survey",
+         year = 2012) %>%
+  select(subject, year, 
+         mod.dev = `Model deviance`, 
+         dev.exp = `Percent deviance explained`, 
+         mod.pval = `Model p-value`)
+
+gdm_survey.2012.tab2 <- gdm_survey.2012.imp[2] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble() %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2) %>%
+  select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)
+
+gdm_survey.2012.tab3 <- gdm_survey.2012.imp[3] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble()  %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2)  %>%
+  select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)
+
+gdm_survey.2010.tab_join <- bind_cols(gdm_survey.2010.tab1,
+                                  gdm_survey.2010.tab2, 
+                                  gdm_survey.2010.tab3)
+
+gdm_survey.2011.tab_join <- bind_cols(gdm_survey.2011.tab1,
+                                  gdm_survey.2011.tab2, 
+                                  gdm_survey.2011.tab3)
+
+gdm_survey.2012.tab_join <- bind_cols(gdm_survey.2012.tab1,
+                                  gdm_survey.2012.tab2, 
+                                  gdm_survey.2012.tab3)
+
+######## BB
+gdm_BB.2010.tab1 <- gdm_BB.2010.imp[1] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble() %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2) %>%
+  mutate(subject = "BB",
+         year = 2010) %>%
+  select(subject, year, 
+         mod.dev = `Model deviance`, 
+         dev.exp = `Percent deviance explained`, 
+         mod.pval = `Model p-value`)
+
+gdm_BB.2010.tab2 <- gdm_BB.2010.imp[2] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble() %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2) %>%
+  select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)
+
+gdm_BB.2010.tab3 <- gdm_BB.2010.imp[3] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble()  %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2)  %>%
+  select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)
+
+gdm_BB.2011.tab1 <- gdm_BB.2011.imp[1] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble() %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2) %>%
+  mutate(subject = "BB",
+         year = 2011) %>%
+  select(subject, year, 
+         mod.dev = `Model deviance`, 
+         dev.exp = `Percent deviance explained`, 
+         mod.pval = `Model p-value`)
+
+gdm_BB.2011.tab2 <- gdm_BB.2011.imp[2] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble() %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2) %>%
+  select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)
+
+gdm_BB.2011.tab3 <- gdm_BB.2011.imp[3] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble()  %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2)  %>%
+  select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)
+
+gdm_BB.2012.tab1 <- gdm_BB.2012.imp[1] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble() %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2) %>%
+  mutate(subject = "BB",
+         year = 2012) %>%
+  select(subject, year, 
+         mod.dev = `Model deviance`, 
+         dev.exp = `Percent deviance explained`, 
+         mod.pval = `Model p-value`)
+
+gdm_BB.2012.tab2 <- gdm_BB.2012.imp[2] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble() %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2) %>%
+  select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)
+
+gdm_BB.2012.tab3 <- gdm_BB.2012.imp[3] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble()  %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2)  %>%
+  select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)
+
+gdm_BB.2010.tab_join <- bind_cols(gdm_BB.2010.tab1,
+                                  gdm_BB.2010.tab2, 
+                                  gdm_BB.2010.tab3)
+
+gdm_BB.2011.tab_join <- bind_cols(gdm_BB.2011.tab1,
+                                  gdm_BB.2011.tab2, 
+                                  gdm_BB.2011.tab3)
+
+gdm_BB.2012.tab_join <- bind_cols(gdm_BB.2012.tab1,
+                                  gdm_BB.2012.tab2, 
+                                  gdm_BB.2012.tab3)
+
+#######################
+
+gdm_BB.tab_join <- bind_rows(gdm_BB.2010.tab_join,
+                             gdm_BB.2011.tab_join,
+                             gdm_BB.2012.tab_join) %>%
+ # pivot_longer(cols = -c(subject, year), names_to = "var", values_to = "value") %>%
+  gt() %>%
+  tab_header(title = "GDM Model Output") %>%
+  tab_spanner(
+    label = "Model",
+    columns = vars(mod.dev, dev.exp, mod.pval)
+  ) %>%
+  tab_spanner(
+    label = "Variable importance",
+    columns = vars(geo.imp, elev.imp, yday.imp)
+  ) %>%
+  tab_spanner(
+    label = "Variable p-value",
+    columns = vars(geo.pval, elev.pval, yday.pval)
+  ) %>% 
+  tab_row_group(
+    group = "Bumble bee",
+    rows = 1:3
+  ) %>%
+  fmt_number(
+    columns = c("mod.dev", "dev.exp",
+                "geo.imp", "elev.imp", "yday.imp"),
+    decimals = 1
+  )
+```
+
+### GDM all together now
+```{r}
+splines <- bind_rows(interaction_splines, bb_splines, survey_splines) %>%
+  mutate(subject = factor(subject))
+
+ggplot(splines, aes(predictor, response, color = subject)) +
+  geom_line() +
+  facet_grid(year ~ x.var, scales = "free_x") +
+  xlab(NULL) +
+  ylab("f(x)") +
+  theme_light(12)
+ggsave("./output/GDM.png", width = 6.5, height = 5, units = "in")
+```
+# If I could get this to work, it would be cool to transform and plot my GDMs. But the script crashes every time. 
+```{r}
+
+# Function to build data set for gdm.transform()
+transform_data <- function(dat) {
+
+  ## yday min/max
+  min_yday <- min(dat$s1.yday)
+  max_yday <- max(dat$s1.yday)
+
+  ## elev min/max
+  min_elev <- min(dat$s1.elev.mean)
+  max_elev <- max(dat$s1.elev.mean)
+
+  ## lat/lon min/max
+  min_x1 <- min(dat$s1.xCoord)
+  max_x1 <- max(dat$s1.xCoord)
+
+  min_y1 <- min(dat$s1.yCoord)
+  max_y1 <- max(dat$s1.yCoord)
+
+  min_x2 <- min(dat$s2.xCoord)
+  max_x2 <- max(dat$s2.xCoord)
+
+  min_y2 <- max(dat$s2.yCoord)
+  max_y2 <- max(dat$s2.yCoord)
+
+  # sequence of 10 points along the yday and elev ranges, respectively
+  s1.xCoord <- seq(min_x1, max_x1, length.out = 10)
+  s1.yCoord <- seq(min_y1, max_y1, length.out = 10)
+  s2.xCoord <- seq(min_x2, max_x2, length.out = 10)
+  s2.yCoord <- seq(min_y2, max_y2, length.out = 10)
+  s1.elev.mean <- seq(min_elev, max_elev, length.out = 10)
+  s1.yday <- seq(min_yday, max_yday, length.out = 10)
+  s2.elev.mean <- seq(min_elev, max_elev, length.out = 10)
+  s2.yday <- seq(min_yday, max_yday, length.out = 10)
+
+  newdata <- tibble(s1.xCoord, s1.yCoord, s2.xCoord, s2.yCoord, 
+                    s1.elev.mean, s1.yday, s2.elev.mean, s2.yday) %>% # combine seqs into data frame
+    complete(s1.xCoord, s1.yCoord, s2.xCoord, s2.yCoord, 
+             s1.elev.mean, s1.yday, s2.elev.mean, s2.yday) %>% # get all possible values
+    mutate(distance = rep(1, n()),
+           weights = rep(1, n())) %>%
+    as.data.frame()
+
+  # coordinates(newdata) <- ~ X + Y
+  # # coerce to SpatialPixelsDataFrame
+  # gridded(newdata) <- TRUE
+  # # coerce to raster
+  # out <- raster(newdata)
+}
+
+transdata_gdm_WN.2010 <- transform_data(sitepair_2010.WN)
+
+gdm_WN.2010.trans <- gdm.transform(gdm_WN.2010, transdata_gdm_WN.2010)
+gdm_WN.2010.pred <- predict(gdm_WN.2010, gdm_WN.2010.trans)
+
+```
+
+## GDM with BBGDM
+```{r}
+library(bbgdm)
+
+spdat2010 <- net %>%
+  filter(year == 2010) %>%
+  select(bb.sp, plant.sp.abb, site, date) %>%
+  distinct() %>%
+  mutate(present = rep(1, length(.[,1]))) %>%
+  unite(interaction, c(bb.sp, plant.sp.abb), sep = "_") %>%
+  unite(site.date, c("site", "date"), sep = "_") %>%
+  pivot_wider(names_from = interaction, values_from = present) %>%
+  replace(is.na(.), 0) %>%
+  column_to_rownames("site.date") %>%
+  as.matrix()
+
+envdat2010 <- net %>%
+  filter(year == 2010) %>%
+  select(bb.sp, plant.sp.abb, site, date) %>%
+  mutate(yday = yday(date)) %>%
+  left_join(site_data) %>%
+  unite(site.date, c("site", "date"), sep = "_") %>%
+  select(site.date, elev.mean, yday, x, y) %>%
+  distinct() %>%
+  column_to_rownames("site.date") %>%
+  as.matrix()
+
+form <- ~ 1 + elev.mean + yday
+
+bbgdm2010 <- bbgdm(form, sp.dat = spdat2010, env.dat = envdat2010,
+                   family = "binomial", link = "logit", dism_metric = "bray_curtis",
+                   nboot = 10, spline_type = "ispline", spline_df = 5, 
+                   spline_knots = 3, geo = FALSE, 
+                   #geo.type = "euclidean", coord.names = c("x", "y"), 
+                   optim.meth = "nlmnib", 
+                   est.var = FALSE, trace = FALSE, prior = FALSE)
+
+plot(as.response(bbgdm2010))
+bbgdm.wald.test(bbgdm2010)
+plot(diagnostics(bbgdm2010))
+```

### 2020-11-26 analyses/beta_diversity.Rmd
did so many things

@@
                               sampleSitePairs = 0.5, # it keep crashing...
-                              outFile = "./output/winteraction_permGDM_2012")
+                              outFile = "./output/interaction_permGDM_2012")
 
@@
   mutate(x.var = case_when(
-    x.var == "elev.mean" ~ "Elevation (m)",
-    x.var == "Geographic" ~ "Geographic (km)",
-    x.var == "yday" ~ "Day of Year")) %>%
-  mutate(x.var = factor(x.var, levels = c("Geographic (km)", "Elevation (m)", "Day of Year")),
-    subject = rep("BB*FL", dim(.)[1]))
+    x.var == "elev.mean" ~ "Elevation",
+    x.var == "Geographic" ~ "Geographic dist.",
+    x.var == "yday" ~ "Time")) %>%
+  mutate(x.var = factor(x.var, levels = c("Geographic dist.", "Elevation", "Time")),
+    subject = rep("interaction", dim(.)[1]))
 ```
@@
   mutate(x.var = case_when(
-    x.var == "elev.mean" ~ "Elevation (m)",
-    x.var == "Geographic" ~ "Geographic (km)",
-    x.var == "yday" ~ "Day of Year"),
-    x.var = factor(x.var, levels = c("Geographic (km)", "Elevation (m)", "Day of Year")),
+    x.var == "elev.mean" ~ "Elevation",
+    x.var == "Geographic" ~ "Geographic dist.",
+    x.var == "yday" ~ "Time"),
+    x.var = factor(x.var, levels = c("Geographic dist.", "Elevation", "Time")),
     subject = rep("FL", n())
@@
   mutate(x.var = case_when(
-    x.var == "elev.mean" ~ "Elevation (m)",
-    x.var == "Geographic" ~ "Geographic (km)",
-    x.var == "yday" ~ "Day of Year"),
-    x.var = factor(x.var, levels = c("Geographic (km)", "Elevation (m)", "Day of Year")),
+    x.var == "elev.mean" ~ "Elevation",
+    x.var == "Geographic" ~ "Geographic dist.",
+    x.var == "yday" ~ "Time"),
+    x.var = factor(x.var, levels = c("Geographic dist.", "Elevation", "Time")),
     subject = rep("BB", n())
@@
 ```{r}
+# assign("gdm_WN.2010.imp", get(load("~/Documents/Research/bb_berchtesgaden/analyses/output/interaction_permGDM_2010.RData")))
+# 
+# assign("gdm_WN.2011.imp", get(load("~/Documents/Research/bb_berchtesgaden/analyses/output/interaction_permGDM_2011.RData")))
+# 
+# assign("gdm_WN.2012.imp", get(load("~/Documents/Research/bb_berchtesgaden/analyses/output/interaction_permGDM_2012.RData")))
+
+####### WN
+gdm_WN.2010.tab1 <- gdm_WN.2010.imp[1] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble() %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2) %>%
+  mutate(subject = "WN",
+         year = 2010) %>%
+  select(subject, year, 
+         mod.dev = `Model deviance`, 
+         dev.exp = `Percent deviance explained`, 
+         mod.pval = `Model p-value`)
+
+gdm_WN.2010.tab2 <- gdm_WN.2010.imp[2] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble() %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2) %>%
+  select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)
+
+gdm_WN.2010.tab3 <- gdm_WN.2010.imp[3] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble()  %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2)  %>%
+  select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)
+
+gdm_WN.2011.tab1 <- gdm_WN.2011.imp[1] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble() %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2) %>%
+  mutate(subject = "WN",
+         year = 2011) %>%
+  select(subject, year, 
+         mod.dev = `Model deviance`, 
+         dev.exp = `Percent deviance explained`, 
+         mod.pval = `Model p-value`)
+
+gdm_WN.2011.tab2 <- gdm_WN.2011.imp[2] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble() %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2) %>%
+  select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)
+
+gdm_WN.2011.tab3 <- gdm_WN.2011.imp[3] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble()  %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2)  %>%
+  select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)
+
+gdm_WN.2012.tab1 <- gdm_WN.2012.imp[1] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble() %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2) %>%
+  mutate(subject = "WN",
+         year = 2012) %>%
+  select(subject, year, 
+         mod.dev = `Model deviance`, 
+         dev.exp = `Percent deviance explained`, 
+         mod.pval = `Model p-value`)
+
+gdm_WN.2012.tab2 <- gdm_WN.2012.imp[2] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble() %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2) %>%
+  select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)
+
+gdm_WN.2012.tab3 <- gdm_WN.2012.imp[3] %>% 
+  as.data.frame() %>%
+  rownames_to_column() %>%
+  as_tibble()  %>%
+  select(1, 2) %>%
+  pivot_wider(names_from = rowname, values_from = 2)  %>%
+  select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)
+
+gdm_WN.2010.tab_join <- bind_cols(gdm_WN.2010.tab1,
+                                  gdm_WN.2010.tab2, 
+                                  gdm_WN.2010.tab3)
+
+gdm_WN.2011.tab_join <- bind_cols(gdm_WN.2011.tab1,
+                                  gdm_WN.2011.tab2, 
+                                  gdm_WN.2011.tab3)
+
+gdm_WN.2012.tab_join <- bind_cols(gdm_WN.2012.tab1,
+                                  gdm_WN.2012.tab2, 
+                                  gdm_WN.2012.tab3)
+
 ####### survey
@@
                              gdm_BB.2011.tab_join,
-                             gdm_BB.2012.tab_join) %>%
- # pivot_longer(cols = -c(subject, year), names_to = "var", values_to = "value") %>%
-  gt() %>%
-  tab_header(title = "GDM Model Output") %>%
+                             gdm_BB.2012.tab_join,
+                             gdm_survey.2010.tab_join,
+                             gdm_survey.2011.tab_join,
+                             gdm_survey.2012.tab_join,
+                             gdm_WN.2010.tab_join,
+                             gdm_WN.2011.tab_join,
+                             gdm_WN.2012.tab_join) %>%
+  select(-subject)
+
+gdm_tab <- gdm_BB.tab_join %>%
+  gt(rowname_col = "year") %>%
   tab_spanner(
@@
   tab_spanner(
-    label = "Variable importance",
-    columns = vars(geo.imp, elev.imp, yday.imp)
+    label = "Geographic distance",
+    columns = vars(geo.imp, geo.pval)
   ) %>%
   tab_spanner(
-    label = "Variable p-value",
-    columns = vars(geo.pval, elev.pval, yday.pval)
+    label = "Elevational distance",
+    columns = vars(elev.imp, elev.pval)
+  ) %>%
+  tab_spanner(
+    label = "Temporal distance",
+    columns = vars(yday.imp, yday.pval)
   ) %>%
   tab_row_group(
-    group = "Bumble bee",
+    group = "Bumble bees",
     rows = 1:3
+  ) %>% 
+  tab_row_group(
+    group = "Flora",
+    rows = 4:6
+  ) %>%
+  tab_row_group(
+    group = "Interactions",
+    rows = 7:9
   ) %>%
@@
     decimals = 1
+  ) %>%
+  fmt_number(
+    columns = c("mod.pval", "geo.pval", 
+                "elev.pval", "yday.pval"),
+    decimals = 3
+  ) %>%
+  cols_label(mod.dev = "tot. dev.",
+             dev.exp = "dev. expl. (%)",
+             mod.pval = "p",
+             geo.imp = "imp.",
+             elev.imp = "imp.",
+             yday.imp = "imp.",
+             geo.pval = "p-val",
+             elev.pval = "p-val",
+             yday.pval = "p-val")
+
+gdm_tab.nopval <- gdm_BB.tab_join %>%
+  dplyr::select(-c(geo.pval, elev.pval, yday.pval, mod.pval)) %>%
+  gt(rowname_col = "year") %>%
+  tab_spanner(
+    label = "Model",
+    columns = vars(mod.dev, dev.exp)
+  ) %>%
+  tab_spanner(
+    label = "Variable importance (%)",
+    columns = vars(geo.imp, elev.imp, yday.imp)
+  ) %>%
+  tab_row_group(
+    group = "Bumble bees",
+    rows = 1:3
+  ) %>% 
+  tab_row_group(
+    group = "Flora",
+    rows = 4:6
+  ) %>%
+  tab_row_group(
+    group = "Interactions",
+    rows = 7:9
+  ) %>%
+  fmt_number(
+    columns = c("mod.dev", "dev.exp",
+                "geo.imp", "elev.imp", "yday.imp"),
+    decimals = 1
+  ) %>%
+  cols_label(mod.dev = "tot. dev.",
+             dev.exp = "dev. expl. (%)",
+             geo.imp = "geo. dist.",
+             elev.imp = "elev. dist.",
+             yday.imp = "temp. dist.") %>%
+  data_color(
+    columns = vars(geo.imp, elev.imp, yday.imp),
+    colors = scales::col_numeric(
+      palette = c("white", "red"), 
+      domain = c(-5, 100))
+    ) %>%
+  data_color(
+    columns = vars(dev.exp),
+    colors = scales::col_numeric(
+      palette = c("white", "red"), 
+      domain = c(-5, 100))
   )
+
+gtsave(gdm_tab.nopval, "gdm_tab.png","./output/gdm_tab.png")
+
 ```
@@
 
-ggplot(splines, aes(predictor, response, color = subject)) +
+# ggplot(splines, aes(predictor, response, color = subject)) +
+#   geom_line() +
+#   facet_grid(year ~ x.var, scales = "free_x") +
+#   xlab(NULL) +
+#   ylab("f(x)") +
+#   theme_light(12)
+# ggsave("./output/GDM.png", width = 6.5, height = 5, units = "in")
+
+splot1 <- ggplot(filter(splines, x.var == "Geographic dist."), 
+                 aes(predictor, response, color = subject)) +
   geom_line() +
+  #facet_wrap(~year, ncol = 1, scales = "free_x") +
   facet_grid(year ~ x.var, scales = "free_x") +
-  xlab(NULL) +
+  xlab("km") +
   ylab("f(x)") +
-  theme_light(12)
-ggsave("./output/GDM.png", width = 6.5, height = 5, units = "in")
-```
+  ylim(c(0, 7)) +
+  theme_light(12) +
+  theme(legend.position = "none")
+
+splot2 <- ggplot(filter(splines, x.var == "Elevation"), 
+       aes(predictor, response, color = subject)) +
+  geom_line() +
+  #facet_wrap(~year, ncol = 1, scales = "free_x") +
+  facet_grid(year ~ x.var, scales = "free_x") +
+  xlab("m") +
+  ylab(NULL) +
+  ylim(c(0, 7)) +
+  theme_light(12) +
+  theme(legend.position = "none")
+
+splot3 <- ggplot(filter(splines, x.var == "Time"), 
+       aes(as_date(predictor), response, color = subject)) +
+  geom_line() +
+  #facet_wrap(~year, ncol = 1, scales = "free_x") +
+  facet_grid(year ~ x.var, scales = "free_x") +
+  xlab("date") +
+  ylab(NULL) +
+  ylim(c(0, 7)) +
+  theme_light(12) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
+
+png("./output/GDM_splines.png", width = 8, height = 5, units = "in", res = 300)
+splot1 | splot2 | splot3
+dev.off()
+```
+
 # If I could get this to work, it would be cool to transform and plot my GDMs. But the script crashes every time. 

### 2020-11-26 analyses/niche_relations_v4.Rmd
did so many things

---
title: "Dietary niche relationships between core bumble bee species in the Berchtesgadener Alps"
output: html_document
---

# Prepare workspace
## Load packages 
```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
library(bipartite)
library(mgcv)
library(mgcViz)
library(vegan)
library(ggvegan)
library(ggrepel)
library(treemapify)
library(RColorBrewer)
library(tidyverse)
library(lubridate)
library(patchwork)
library(tidymodels)
```

## Load data
```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
# Site data
site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, levels = c("unten", "mitte", "oben"))) # turn elav.class into an ordered factor

# BB traits
bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
  mutate(pbl.fac = factor(if_else(
    bb.sp %in% c("hort", "pasc", "wurf", "gers", "mend", "hypn"), "long", "short"))
    )

# FL traits
## family-level taxonomy
fl_tax <- read_csv("../processed_data/floral_tax.csv")

## Kugler morphotypes (accessed through Bioflor)
fl_ktype <- read_csv("../processed_data/floral_k_type.csv")

## diameter and corolla depth data from Gita Benadi
fl_morph <- read_csv("../processed_data/flower_morphology_GitaBenadi.csv") %>% 
  dplyr::select(SampleID, plant.family = Family, plant.sp = Species, 
                Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam), LN = mean(LN))

## verbal description of Kugler morphotypes
fl_ktype_key <- read_delim("../processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  dplyr::select(k.type = 1, description = 2)

## join all trait tables
fl_traits <- fl_tax %>%
  full_join(fl_ktype) %>%
  full_join(fl_morph) %>%
  dplyr::select(plant.sp, plant.genus, plant.family, k.type, k.type.s, k.type.ss, diam, LN)

# Network
net <- read_csv("../processed_data/network.csv") %>%
  mutate(year = factor(year)) %>% # convert year to factor
  left_join(site_data, by = "site") %>% # add site data
  ungroup() %>%
  arrange(site, date) %>%
  left_join(fl_traits) %>% # add floral traits
  group_by(elev.class) %>%
  mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) # divide each elev.class into 3 equal time slices

# Floral survey
survey <- read_csv("../processed_data/floral_survey.csv") %>%
  filter(flower.cover > 0) %>% # only count plants in bloom
  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  left_join(fl_traits) # add floral traits
```

We will focus on six species that are abundant and widespread:  
  1. *B. sensu-strictu*
  2. *B. pascuorum*
  3. *B. pratorum*
  4. *B. hortorum*
  5. *B. soroensis*
  6. *B. wurflenii*

```{r fig.cap="**Figure 1: Focal species**", message=FALSE, warning=FALSE, echo=FALSE}
bb_range.year <- net %>%
  left_join(site_data) %>%
  group_by(site, elev.mean, bb.sp, year) %>%
  summarize(abund = n()) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2)) %>%
  mutate(focal = if_else(
    bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf"), TRUE, FALSE)
    )

ggplot(bb_range.year, aes(reorder(bb.sp, elev.med), elev.mean, color = focal)) +
  geom_line(size = 2, alpha = 0.25) +
  geom_point(aes(size = abund), alpha = 0.5) +
  facet_wrap(~year, ncol = 1) +
  theme_light(12) +
  scale_size_continuous(name = "Abundance") +
  scale_color_discrete(name = "Focal species") +
  xlab("BB species") +
  ylab("Elevation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("./output/bb_range_year.pdf", height = 6, width = 6)
```

## Define functions
```{r, echo = FALSE}
# Convert PCoA to tidy array ("fortify")
pcoa_fort <- function(x){
  fortify(x, axes = 1:2) %>% 
  mutate(Score = case_when( # rename default scores to match my data
    Score == "species" ~ "FL.sp",
    Score == "sites" ~ "BB.sp"))
}

# Plot PCoA; to avoid overplotting, scores < min_score are not plotted
pcoa_plot <- function(x, min_score = 0.25) {

  # Add T.class variable for the BB.sp subset of the PCoA
  t.class <- x %>%
    filter(Score == "BB.sp") %>%
    mutate(T.class = factor(if_else(
      Label %in% c("hort", "pasc", "wurf", "gers", "mend", "hypn"), "long", "short"))
      )

  ggplot(filter(x, (abs(MDS1) > min_score | abs(MDS2) > min_score) | (Score == "BB.sp"))) + # no threshold for BB spp
    geom_polygon(data = t.class, aes(MDS1, MDS2, linetype = T.class), # draw polygons around T classes
                 color = "gray70", fill = NA, size = 1.25, key_glyph = "abline") +
    geom_point(aes(MDS1, MDS2, color = Score), size = 3, alpha = 0.75) + # add points
    ggrepel::geom_text_repel(aes(MDS1, MDS2, color = Score, label = Label), key_glyph = "rect") + # add labels
    coord_fixed() + # Gavin Simpson says this is **really** important
    theme_classic() 
}

# Generates PCoAs, ADONIS tests, and ordination plots (fully aggregated)
## The ... represents the variables we will use to parse our analysis. For aggregate analysis, just leave the ...
## unpopulated, e.g. niche_analysis(net, bb_traits, plant.sp). 
## To parse by elev.class: niche_analysis(net, bb_traits, plant.sp)
## To parse by  elev.class, period, and pollen: niche_analysis(net, bb_traits, plant.sp, period, pollen)
## You get the idea. You should anyway -- you wrote it. https://tidyeval.tidyverse.org/multiple.html
niche_analysis <- function(x, y, criterion, ...) {

  criterion <- enquo(criterion) # this argument toggles taxonomic (plant.sp) vs. morphological (k.type.s) analysis

  x %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
    group_by(..., bb.sp, !!criterion) %>% # pooling all sites and years
    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
    group_by(..., bb.sp) %>%
    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
    dplyr::select(-freq) %>% # drop freq variable
    group_by(...) %>%
    nest() %>% # nest into list-column
    mutate(comm1 = map(data, # spread into the kind of array vegan likes
                       function(x) pivot_wider(x, names_from = !!criterion, values_from = prop) %>%
                         replace(is.na(.), 0)
                       ), 
           comm2 = map(comm1, # column to rowname
                       function(x) column_to_rownames(x, var = "bb.sp")
                       ),
           pcoa = map(comm2, # call PCoA
                      function(x) capscale(x ~ 1, distance = "horn")
                      ),
           traits = map(comm1, # join BB traits
                        function(x) x %>% 
                          left_join(y) %>% # y is the trait table, an argument in the main function call
                          column_to_rownames(var = "bb.sp") %>% 
                          select(pbl.fac) # pbl.fac is the variable I'm interested in
                        ),
           adonis = map2(comm2, traits, # run adonis permuatation test for distance ~ tongue class
                         function(x, y) adonis(x ~ pbl.fac, data = y, method = "horn")
                         ),
           betadis = map2(comm2, traits, # betadisper to test for difference in dispersion between tongue classes
                         function(x, y) x %>%
                           vegdist(method = "horn") %>% # betadisper runs a PCoA under the hood; match method to my PCoA
                           betadisper(group = y$pbl.fac) 
                         ),
           betadis.permutest = map(betadis, permutest), # perumutation test
           betadis.p = map(betadis.permutest, # extract betadisper p-value
                           function(x) x$tab$'Pr(>F)'[[1]]
                           ),
           betadis.boxplot = map(betadis, boxplot), # boxplot of betadisper
           pcoa.fort = map(pcoa, pcoa_fort), 
           pcoa.plot = map(pcoa.fort, pcoa_plot),
           adonis.r2 = map(adonis, # extract adonis r2
                           function(x) x$aov.tab$R2[[1]]
                           ),
           adonis.p = map(adonis, # extract adonis p-value
                          function(x) x$aov.tab$'Pr(>F)'[[1]]
                          ),
           MDS1 = map(pcoa, # extract variance explained by MDS1
                      function(x) x$CA$eig[[1]] / sum(x$CA$eig)
                      ), 
           MDS2 = map(pcoa, # extract variance explained by MDS2
                      function(x) x$CA$eig[[2]] / sum(x$CA$eig)
                      )) %>%
    unnest(cols = c(adonis.r2, adonis.p, betadis.p, MDS1, MDS2)) %>% # these are lists of length = 1, so unnesting them simply displays the value
    select(..., comm2, pcoa, betadis, betadis.permutest, betadis.p, adonis, 
           adonis.r2, adonis.p, betadis.boxplot, pcoa.plot, MDS1, MDS2) 
}

beta_func <- function(x, traits, sitedata, level, criterion, binary, ...) {

  criterion <- enquo(criterion) # plant.sp or k.type.s
  level <- enquo(level) # bb.sp or pbl.fac (i.e. tongue length  guild)

  x %>% 
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    left_join(traits) %>% # add BB traits
    group_by(site, !!level, !!criterion, ...) %>% # in addition to grouping by level (bb.sp/pbl.fac) and criterion (plant.sp/k.type.s), the ... argument will also group by any variables supplied in the function call, e.g. elev.class, period, year.
    summarize(freq = n()) %>% # get visitation frequency
    group_by(site, !!level, ...) %>%
    mutate(prop = freq/sum(freq)) %>% # get visitation proportions
    group_by(!!level, ...) %>%
    select(-freq) %>%
    nest() %>% # nest into list-column
    mutate(comm = map(data, function(x) pivot_wider(x, names_from = !!criterion, values_from = prop, values_fill = 0)), # convert to community matrix
           comm2 = map(comm, function(x) column_to_rownames(x, var = "site")), # column to rowname
           dist = map(comm2, function(x) vegdist(x, method = "horn", binary = binary)), # conver to Horn's distance matrix
           tidy.dist = map(dist, tidy), # convert matrix to tidy array
           elevation = map(comm, function(x) left_join(x, sitedata, by = "site") %>% # add elevation data
                             select(site, elev.mean) %>% 
                             column_to_rownames(var = "site")),
           adonis = map2(dist, elevation, function(x, y) adonis(x ~ elev.mean, data = y)), # run adonis test
           adonis.r2 = map(adonis, function(x) x$aov.tab$R2[[1]]), # extract adonis r2
           adonis.p = map(adonis, function(x) x$aov.tab$'Pr(>F)'[[1]])) %>% # extract adonis p-val
    unnest(c(tidy.dist, adonis.r2, adonis.p)) %>% 
    select(..., !!level, site1 = item1, site2 = item2, distance, adonis.r2, adonis.p) %>%
    left_join(select(site_data, site, elev.mean), by = c("site1" = "site")) %>%
    rename(elev1 = elev.mean) %>%
    left_join(select(site_data, site, elev.mean), by = c("site2" = "site")) %>%
    rename(elev2 = elev.mean) %>%
    mutate(elev.diff = abs(elev1 - elev2)) 
}

# For the floral survey data
beta_func_fl <- function(x, sitedata, criterion, binary, ...) {

  criterion <- enquo(criterion) # plant.sp or k.type.s

  x %>% 
    group_by(site, !!criterion, ...) %>% # 
    summarize(abund = sum(flower.cover)) %>% # get visitation frequency
    group_by(site, ...) %>%
    mutate(prop = abund/sum(abund)) %>% # get visitation proportions
    select(-abund) %>%
    pivot_wider(names_from = !!criterion, values_from = prop, values_fill = 0) %>%
    column_to_rownames(var = "site") %>%
    vegdist(method = "horn", binary = binary) %>%
    tidy() %>%
    select(..., site1 = item1, site2 = item2, distance) %>%
    left_join(select(site_data, site, elev.mean), by = c("site1" = "site")) %>%
    rename(elev1 = elev.mean) %>%
    left_join(select(site_data, site, elev.mean), by = c("site2" = "site")) %>%
    rename(elev2 = elev.mean) %>%
    mutate(elev.diff = abs(elev1 - elev2)) 
}

# Niche overlap by sp.pair * site * year
niche_overlapper <- function(x, criterion) {

  criterion <- enquo(criterion) # criterion is set to either plant.sp (for taxonomic) or k.type.s (for morphological)

  x %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    group_by(bb.sp, !!criterion, site, year) %>% # group by site-year and BB-FL pair
    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-year
    group_by(bb.sp, site, year) %>% # group by site-year and BB species
    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
    dplyr::select(-freq) %>% # drop freq variable
    pivot_wider(names_from = !!criterion, values_from = prop, values_fill = 0) %>% # spread into the kind of array vegan likes
    group_by(site, year) %>%
    nest() %>% # nest into list-column
    mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
           dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
           tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
    dplyr::select(site, year, tidy_dist) %>% # drop intermediate list-columns
    unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
    left_join(site_data) %>% # add site data
    dplyr::select(site, year, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
    left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
    rename(subgenus1 = subgenus, pbl.w1 = pbl.w, pbl.w.class1 = pbl.w.class, pbl.fac1 = pbl.fac) %>%
    left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
    rename(subgenus2 = subgenus, pbl.w2 = pbl.w, pbl.w.class2 = pbl.w.class, pbl.fac2 = pbl.fac) %>%
    dplyr::select(site, year, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
           pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, pbl.fac1, pbl.fac2, elev.mean) %>%
    mutate(pbl.diff = abs(pbl.w1 - pbl.w2), # tongue length differences
           pbl.fac.diff = if_else(pbl.fac1 == pbl.fac2, "same", "different"),
           subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different"), # binary: same or different subgenus

           # Logical columns indicating whether a given species belongs to a given species pair
           # This is handy for extracting data for individual species across all their counterparts
           bss = if_else(bb.sp1 == "bss" | bb.sp2 == "bss", TRUE, FALSE),
           pasc = if_else(bb.sp1 == "pasc" | bb.sp2 == "pasc", TRUE, FALSE),
           wurf = if_else(bb.sp1 == "wurf" | bb.sp2 == "wurf", TRUE, FALSE),
           prat = if_else(bb.sp1 == "prat" | bb.sp2 == "prat", TRUE, FALSE),
           soro = if_else(bb.sp1 == "soro" | bb.sp2 == "soro", TRUE, FALSE),
           hort = if_else(bb.sp1 == "hort" | bb.sp2 == "hort", TRUE, FALSE)) %>%

    unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
    mutate(sp.pair = factor(sp.pair)) %>% # ...and make it a factor so that mgcv can handle it
    mutate(niche.overlap = 1-distance) # express niche overlap so that 1 = perfect overlap and 0 = no overlap
}

# Diet breadth analysis
diet_breit <- function(x, y, z, criterion, indices = c("normalised degree")) {

  criterion <- enquo(criterion) # plant.sp or k.type.s

  a <- x %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    left_join(y) %>%
    group_by(elev.class) %>%
    mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
    group_by(site, period, pbl.fac, !!criterion) %>% # group by site, date, bee species, plant species
    summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
    unite(webID, c(site, period), sep = "_") %>%
    ungroup() %>%
    dplyr::select(higher = pbl.fac, lower = !!criterion, webID, freq) %>% # rename columns to match bipartite's expectations
    data.frame() %>% # convert to data frame
    frame2webs() %>% # convert to to bipartite web
    map(specieslevel, level = "higher", index = indices) %>%
    map(rownames_to_column) %>%
    map(as_tibble) %>%
    bind_rows(.id = "webID") %>% # Collapse list into big data frame
    rename(pbl.fac = rowname) %>%
    separate(webID, c("site", "period"), sep =  "_") %>%
    left_join(z) %>%
    filter(pbl.fac %in% c("long", "short")) %>% # for some reason, a few rows have a "1" for pbl.fac, and I don't know where this comes from; this gets rid of it, but I would still like to know why
    mutate(site = factor(site),
           period = factor(period, levels = c("early", "mid", "late")))

  b <- x %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    group_by(elev.class) %>%
    mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
    left_join(bb_traits) %>%
    group_by(site, period, pbl.fac) %>% # group by site, date, bee species, plant species
    summarize(abund = n()) %>%
    group_by(site, period) %>%
    mutate(prop = abund/sum(abund))

  out <- left_join(a, b)
}

diet_breit2 <- function(x, y, z, criterion, indices = c("normalised degree")) {

  criterion <- enquo(criterion) # plant.sp or k.type.s

  a <- x %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    left_join(y) %>%
    # group_by(elev.class) %>%
    # mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
    #group_by(site, period, pbl.fac, !!criterion) %>% # group by site, date, bee species, plant species
    group_by(site, pbl.fac, !!criterion) %>% # group by site, date, bee species, plant species
    summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
    #unite(webID, c(site, period), sep = "_") %>%
    ungroup() %>%
    #dplyr::select(higher = pbl.fac, lower = !!criterion, webID, freq) %>% # rename columns to match bipartite's expectations
    dplyr::select(higher = pbl.fac, lower = !!criterion, webID = site, freq) %>% # rename columns to match bipartite's expectations
    data.frame() %>% # convert to data frame
    frame2webs() %>% # convert to to bipartite web
    map(specieslevel, level = "higher", index = indices) %>%
    map(rownames_to_column) %>%
    map(as_tibble) %>%
    bind_rows(.id = "webID") %>% # Collapse list into big data frame
    #rename(pbl.fac = rowname) %>%
    rename(pbl.fac = rowname,
           site = webID) %>%
    #separate(webID, c("site", "period"), sep =  "_") %>%
    left_join(z) %>%
    filter(pbl.fac %in% c("long", "short")) %>% # for some reason, a few rows have a "1" for pbl.fac, and I don't know where this comes from; this gets rid of it, but I would still like to know why
    # mutate(site = factor(site),
    #        period = factor(period, levels = c("early", "mid", "late")))
    mutate(site = factor(site))

  b <- x %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    # group_by(elev.class) %>%
    # mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
    left_join(bb_traits) %>%
   # group_by(site, period, pbl.fac) %>% # group by site, date, bee species, plant species
    group_by(site, pbl.fac) %>% # group by site, date, bee species, plant species
    summarize(abund = n()) %>%
    #group_by(site, period) %>%
    group_by(site) %>%
    mutate(prop = abund/sum(abund))

  out <- left_join(a, b)
}

# Null model generator
nullify1 <- function(x, n, criterion){

  criterion <- enquo(criterion) # plant.sp or k.type.s

  x %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    select(site, bb.sp, !!criterion) %>% 
    group_by(site, bb.sp, !!criterion) %>%
    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
    group_by(site) %>%
    nest() %>% # nest data into list-column grouped by pollen, site, date
    mutate(data = map(data, function(x) pivot_wider(x, 
                                                    names_from = !!criterion, 
                                                    values_from = freq, 
                                                    values_fill = 0) %>% column_to_rownames("bb.sp")), # turn table into community matrix
           data = map(data, function(x) vegan::permatfull(x, 
                                                          fixedmar = "rows", 
                                                          shuffle = "ind", 
                                                          times = n)), # permute community matrices
           data = map(data, function(x) return(x$perm))) %>% # extract permutations from permat objects
    unnest(data) %>% # unnest
    mutate(data = map(data, function(x) vegdist(x, method = "horn")), # convert each permutation into distance matrix
           data = map(data, tidy)) %>% # convert distance matrices to tidy arrays
    group_by(site) %>%
    mutate(iter = row_number()) %>% # add index to denote iterations
    unnest(cols = data) %>% # unnest again
    rename(bb.sp1 = item1, bb.sp2 = item2) %>%
    unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
    mutate(sp.pair = factor(sp.pair), # ...and make it a factor so that mgcv can handle it
           niche.overlap = 1-distance) %>% # express niche overlap so that 1 = perfect overlap and 0 = no overlap
    select(site, sp.pair, iter, niche.overlap.sim = niche.overlap)
}

# Null model generator
nullify2 <- function(x, n, criterion){

  criterion <- enquo(criterion) # plant.sp or k.type.s

  x %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    select(site, bb.sp, !!criterion) %>% 
    group_by(site, bb.sp, !!criterion) %>%
    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
    group_by(site) %>%
    nest() %>% # nest data into list-column grouped by pollen, site, date
    mutate(data = map(data, function(x) pivot_wider(x, 
                                                    names_from = !!criterion, 
                                                    values_from = freq, 
                                                    values_fill = 0) %>% column_to_rownames("bb.sp")), # turn table into community matrix
           data = map(data, function(x) vegan::permatfull(x, 
                                                          fixedmar = "both", 
                                                          shuffle = "ind", 
                                                          times = n)), # permute community matrices
           data = map(data, function(x) return(x$perm))) %>% # extract permutations from permat objects
    unnest(data) %>% # unnest
    mutate(data = map(data, function(x) vegdist(x, method = "horn")), # convert each permutation into distance matrix
           data = map(data, tidy)) %>% # convert distance matrices to tidy arrays
    group_by(site) %>%
    mutate(iter = row_number()) %>% # add index to denote iterations
    unnest(cols = data) %>% # unnest again
    rename(bb.sp1 = item1, bb.sp2 = item2) %>%
    unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
    mutate(sp.pair = factor(sp.pair), # ...and make it a factor so that mgcv can handle it
           niche.overlap = 1-distance) %>% # express niche overlap so that 1 = perfect overlap and 0 = no overlap
    select(site, sp.pair, iter, niche.overlap.sim = niche.overlap)
}
```

# 1. What is the logic of niche partitioning between core species, and does it change through elevation and time?

For ordination-based analyses, I have opted to pool across years to emphasize species relationships rather than interannual differences. To aid interpretation, I have classified BB species by tongue length. The two "true" long-tongued species are *B. pascuorum* and *B. hortorum*, with the latter having a considerably longer tongue than the former, but I have opted to consider *B. wurflenii* as "behaviorally long-tongued" because it characteristically forages as a robber, and evidently converges in flower choice with *B. pascuorum* and *B. hortorum*. This decision was made *post-hoc* after noting the clustering patterns.

I have also opted to not distinguish between putative pollen and putative nectar foragers. First, the classification of nectar vs. pollen foragers based on whether they are carrying pollen at the time of observation is imperfect (though probably still informative). Second, and more importantly in my opinion, the more I break the data up, the more difficult it becomes to visualize and interpret the data. Finally, when I ran the analysis parsing by pollen vs. nectar foraging, the patterns tended to be similar between the two.

Note that in this version of the analysis, I am not excluding observations that were unclassified with respect to pollen carrying. A closer inspection of the data revealed that this is simply how Kalli handled males and Psithyrus queens, since they never collect pollen. I am omitting Psithyrus from my analysis, but I am not including the males of the focal specis. I suppose an argument could be made that parsing by caste would be informative, but again, there are so many ways these data could be parsed, and I can't use all of them without whittling the data down to something intractable.

Kugler classes are coded in the form of number.decimal.letter.letter. I have chosen to simplify to the level of number.decimal.

To avoid overplotting plant species / Kugler classes, I have only plotted plant species scoring above a given threshold on either the first or second axis. Thus, the plotted species are those that contribute most to the separation, not necessarily those that contribute most to diet. A species that is visited heavily by all BB species will fall close to (0,0) and will therefore not be plotted. This is because they are added based on weighted averaging. i.e. they get "pulled" toward BB species in proportion to their (relative) abundance (i.e. visitation rate) for each species.

## 1.1 Ordination by species with post-hoc fitting of tongue length  
### Script
```{r message=FALSE, warning=FALSE, echo = FALSE}
# By plant.sp
niche.tax <- niche_analysis(net, bb_traits, criterion = plant.sp.abb)
niche.tax.plot <- niche.tax$pcoa.plot[[1]]

# By ktype
niche.morph <- niche_analysis(net, bb_traits, criterion = k.type.s)
niche.morph.plot <- niche.morph$pcoa.plot[[1]]

## Taxonomic and morphological together
pdf("../analyses/output/niche_pcoa.pdf", height = 8.5, width = 11)
(niche.tax.plot + ggtitle("Taxonomic") | niche.morph.plot + ggtitle("Morphological")) + plot_layout(guides = 'collect') 
dev.off()
```

### Results

In parentheses following the initial mention of each MDS axis, the variance explained is shown for the taxonomic (T:) and morphological (M:) ordinations.

1. MDS1 (T:48%, M:86%) separates long-tongued from short--tongued species. This mirrors the findings of Goulson et al. (2008). MDS2 (T:27%, M:12%) mainly accounts for the divergent foraging behavior of *B. sensu-stricto*. I wonder if this has been described before? A caveat to the interpretation of this pattern, though, is that *B. sensu-stricto* exhibited some weird temporal abundance patterns, particularly in 2012. It's possible that this biased its foraging in a way that could account for some of its apparent deviance.

2. LT species cluster around the lip flowers *Salvia glutinosa* and *Rhinanthus glacialis* (Kugler 5.1) and the flag flower *Anthyllis vulneraria* (Kugler 6.1). ST species *B. soroensis* and *B. pratorum* cluster around the bell flower *Campanula scheuchzeri* (Kugler 3.2) and flower heads *Carduus defloratus* and *Centaurea jacea* (Kugler 7.2). The short-tongued *B. sensu-stricto* seems to be especially associated with the bell flower *Erica carnea* (Kugler 3.1), which scores near 0 on MDS1 (i.e. not accounting for the separation of tongue-legnth guilds) but strongly negative on MDS2, mirroring the scoring *B. sensu-stricto*. The lip flower *Stachys alopecuros* (Kugler 5.1) also scores near 0 on MDS1 but high on MDS2, and together with *Erica carnea* seems to contribute to the separation of *B. sensu-strictu* from the other species of both tongue classes.

## 1.2 Ordination by species with post-hoc fitting of tongue length, broken down by elevation class
### Script
```{r message=FALSE, warning=FALSE, echo = FALSE}
# Taxonomic
niche_x_elev.tax <- niche_analysis(net, bb_traits, criterion = plant.sp.abb, elev.class)

## Extract and name plots
niche_oben.tax.plot <- filter(niche_x_elev.tax, elev.class == "oben")$pcoa.plot[[1]] + ggtitle("Oben/Tax")
niche_mitte.tax.plot <- filter(niche_x_elev.tax, elev.class == "mitte")$pcoa.plot[[1]] + ggtitle("Mitte/Tax")
niche_unten.tax.plot <- filter(niche_x_elev.tax, elev.class == "unten")$pcoa.plot[[1]] + ggtitle("Unten/Tax")

# Morphological
niche_x_elev.morph <- niche_analysis(net, bb_traits, criterion = k.type.s, elev.class)

## Extract and name plots
niche_oben.morph.plot <- filter(niche_x_elev.morph, elev.class == "oben")$pcoa.plot[[1]] + ggtitle("Oben/morph")
niche_mitte.morph.plot <- filter(niche_x_elev.morph, elev.class == "mitte")$pcoa.plot[[1]] + ggtitle("Mitte/morph")
niche_unten.morph.plot <- filter(niche_x_elev.morph, elev.class == "unten")$pcoa.plot[[1]] + ggtitle("Unten/morph")

## Display plots on grid
pdf("../analyses/output/niche_x_elev_pcoa.pdf", height = 8.5, width = 11)
(niche_oben.tax.plot | niche_oben.morph.plot) /
(niche_mitte.tax.plot | niche_mitte.morph.plot) /
(niche_unten.tax.plot | niche_unten.morph.plot) + plot_layout(guides = 'collect') 
dev.off()
```

### Results  
1. At low elevation, MDS1 (T:63%, M:64%) does *not* clearly separate LT from ST species; instead, it seems to be driven by the strong association between *B. soroensis* and the flower head *Centaurea jacea* (Kugler 7.2), which score high on MDS1. MDS2 (T:34%. M:35%) seems to be driven by the associated of *B. pratorum* and *B. hortorum* with the stalk-disc flower *Primula elatior* (Kugler 4.1). In general, though, the BB species at low elevation do not seem to cluster into clear groups.

2. At mid elevation, MDS1 (T:48%, M:79%) seems to be strongly driven by tongue-length. LT species are associated with the lip flowers *Salvia glutinosa* and *Rhinanthus glacialis* (Kugler 5.1) and the flag flower *Anthyllis vulneraria* (Kugler 6.1). ST species are associated with the flower head *Centaurea jacea* (Kugler 7.2). MDS2 (T:25%, M:16%) seems to be driven by the divergent foraging of *B. sensu-stricto*, particularly its association with the bell flower *Erica carnea* (Kugler 3.1), though *E. carnea* plots centrally because it was also visited by other BB species.

3. At high elevation, MDS1(T:52%, M:79%) again separates LT and ST species (though not as strongly as at mid elevation) while MDS2 (T:38%, M:18%) separates *B. sensu-stricto*. In this case, though, *B. sensu-stricto* seems to be especially associated with the disk flower *Helianthemum nummularium* (Kugler 1.1). LT species are associated with the lip flower *Stachys alopecuros* (Kugler 5.1) and the flag flower *Anthyllis vulneraria* (Kugler 6.1), while the ST species *B. pratorum* and *B. soroensis* are associated with the bell flower *Campanula scheuchzeri* (Kugler 3.2) and flower head *Carduus defloratus* (Kugler 7.2).

## 1.3 Ordination by BB species * elev.class * period, pooling across years
### Script
```{r message=FALSE, warning=FALSE, echo = FALSE}
# Taxonomic
niche_x_elevperiod.tax <- niche_analysis(net, bb_traits, criterion = plant.sp.abb, elev.class, period)

## Extract and name plots
niche_oben_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "oben" & period == "early")$pcoa.plot[[1]] +
  ggtitle("Oben/Early/Tax")
niche_oben_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "oben" & period == "mid")$pcoa.plot[[1]] +
  ggtitle("Oben/Mid/Tax")
niche_oben_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "oben" & period == "late")$pcoa.plot[[1]] +
  ggtitle("Oben/Late/Tax")

niche_mitte_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "mitte" & period == "early")$pcoa.plot[[1]] +
  ggtitle("Mitte/Early/Tax")
niche_mitte_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "mitte" & period == "mid")$pcoa.plot[[1]] +
  ggtitle("Mitte/Mid/Tax")
niche_mitte_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "mitte" & period == "late")$pcoa.plot[[1]] +
  ggtitle("Mitte/Late/Tax")

niche_unten_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "unten" & period == "early")$pcoa.plot[[1]] +
  ggtitle("Unten/Early/Tax")
niche_unten_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "unten" & period == "mid")$pcoa.plot[[1]] +
  ggtitle("Unten/Mid/Tax")
niche_unten_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "unten" & period == "late")$pcoa.plot[[1]] +
  ggtitle("Unten/Late/Tax")

# Morphological
niche_x_elevperiod.morph <- niche_analysis(net, bb_traits, criterion = k.type.s, elev.class, period)

## Extract and name plots
niche_oben_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "oben" & period == "early")$pcoa.plot[[1]] +
  ggtitle("Oben/Early/Morph")
niche_oben_mid.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "oben" & period == "mid")$pcoa.plot[[1]] +
  ggtitle("Oben/Mid/Morph")
niche_oben_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "oben" & period == "late")$pcoa.plot[[1]] +
  ggtitle("Oben/Late/Morph")

niche_mitte_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "mitte" & period == "early")$pcoa.plot[[1]] +
  ggtitle("Mitte/Early/Morph")
niche_mitte_mid.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "mitte" & period == "mid")$pcoa.plot[[1]] +
  ggtitle("Mitte/Mid/Morph")
niche_mitte_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "mitte" & period == "late")$pcoa.plot[[1]] +
  ggtitle("Mitte/Late/Morph")

niche_unten_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "unten" & period == "early")$pcoa.plot[[1]] +
  ggtitle("Unten/Early/Morph")
niche_unten_mid.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "unten" & period == "mid")$pcoa.plot[[1]] +
  ggtitle("Unten/Mid/Morph")
niche_unten_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "unten" & period == "late")$pcoa.plot[[1]] +
  ggtitle("Unten/Late/Morph")

## Display plots on grid
#pdf("../analyses/output/niche_x_elevperiod_pcoa_tax.pdf", height = 8.5, width = 11)
(niche_oben_early.tax.plot | niche_oben_mid.tax.plot  | niche_oben_late.tax.plot) /
(niche_mitte_early.tax.plot | niche_mitte_mid.tax.plot | niche_mitte_late.tax.plot) /
(niche_unten_early.tax.plot | niche_unten_mid.tax.plot  | niche_unten_late.tax.plot)  + plot_layout(guides = 'collect') 
#dev.off()

#pdf("../analyses/output/niche_x_elevperiod_pcoa_morph.pdf", height = 8.5, width = 11)
(niche_oben_early.morph.plot | niche_oben_mid.morph.plot  | niche_oben_late.morph.plot) /
(niche_mitte_early.morph.plot | niche_mitte_mid.morph.plot | niche_mitte_late.morph.plot) /
(niche_unten_early.morph.plot | niche_unten_mid.morph.plot  | niche_unten_late.morph.plot)  + plot_layout(guides = 'collect') 
#dev.off()
```

### Results

This is probably the most interpretable analysis. One thing I like about it is how it reveals important floral species that didn't score highly in the aggregate analysis. But there is also a lot to say.

1. *Low elevation*: The most salient pattern is a lack of one -- tongue length does not seem to be important at low elevation *until late in the season*, when all the short ST bees agree that *Centaurea jacea* is the only game in town.

  - *Early*: MDS1 (T:98%, M:97%) dominates the ordination and mainly separates *B. soroensis*, associated with the bell flower *Geum rivale* (Kugler 3.2) from the other BB species, associated with the stalk-disc flower *Primula elatior* (Kugler 4.1). MDS2 (T:2%, M:2%) accounted  for almost no variation.

  - *Mid*: MDS1 (T:90%, M:94%) again dominates the ordination, apparently representing a gradient of association ranging from the lip flower *Rhinanthus alectorolophus* (Kugler 5.1) and the flag blossom *Trifolium pratense* (Kugler 6.1) on the  low end of the axis to the flower head *Knautia arvensis* (Kugler 7.1). BB species are spread fairly evenly along this gradient, with *B. sensu-stricto* and *B. wurflenii* leaning hardest in the *Rhinanthus/Trifolium* direction and *B. pratorum* and *B. pascuorum* leaning hardest in the *Knautia* direction. Interestingly, the little variance MDS2 (T:9%, M:6%) explains seems to be associated with tongue length, though because MDS2 accounts for so little variance, we must conclude that tongue length is of little importance in structuring the visitation network in this elevation-time slice.

  - *Late*: MDS1 (T:51%, M:70%) mainly separates *B. wurflenii* from the rest of the BB species, apparently due to its strong association with the flag blossom *Lotus corniculatus* (Kugler 6.1) and the funnel flower *Gentiana aspera* (Kugler 2.1). MDS2 (T:41%, M:27%), which matters again, seems to separate LT ans ST species, though the pattern is not so much one of separation as of dispersion. All the ST species are huddled tightly around the flower head *Centaurea jacea* (Kugler 7.2), while the LT species are rather spread out, with *B. wurflenii* as described, *B. hortorum* expressing a distinct taste for the lip flower *Stachys officinalis* (Kugler 5.1), and *B. pascuorum* middling as usual.

2. *Mid elevation*:

  - *Early*: MDS1 (T:52%, M:68%) is clearly the tongue-length axis. ST species are all clustered around the bell flower *Erica carnea* (Kugler 3.1), while the LT species are more spread out, with *B. wurflenii* committed to the flag blossom *Polygala chamaebuxus*, *B. hortorum* more interested in the stalk-disk flower *Gentiana verna* (Kugler 4.1), and *B. pascuorum* falls in between (though a lot closer to *B. wurflenii*), evidentally pulled a bit toward the the stalk-disc flower *Primula elatior* (Kugler 4.1). MDS2 (T:32%, M:29%) seems to be driven strongly by *B. hortorum* divergence from the other two LT species.

  - *Mid*: MDS1 (T:49%, M:68%) separates *B. soroensis* and *B. pratorum* from the other BB species. The former are associated with the flower head *Phyteuma orbiculare* (Kugler 7.1) and the bell flower *Geum rivale* (Kugler 3.2). The latter consist of a fairly tight cluster of *B. hortorum*, *B. pascuorum*, and *B. wurflenii* around the flag flowers *Anthyllis vulneraria* and *Lotus corniculatus* (Kugler 6.1) and the lip flower *Ajuga reptans* (Kugler 5.1). MDS2 (T:26%, M:17%), together with MDS1, accounts for the uniqueness of *B. sensu-strictu* foraging, with that species plotting both very low on MDS1 and very high on MDS2. Interestingly, none of the plant species plot close to it, so it must exceptional not because it forages on something unique but because it forages a little bit on everything. That's the  only explanation I can come up with, anyway, for why the weighted-average plotting pulls plants away from *B. sensu-stricto*. The apparent exceptionality of *B. sensu-stricto* is a theme in this analysis.

  - *Late*: MDS1 (T:67%, M:98%) sort of separates LT and ST species, except that *B. pratorum* and *B. pascuorum* fall close to each other on this axis (but are separated widely on MDS2). The LT species are fairly well-clustered around the funnel flower *Gentiana aspera* (Kugler 2.1), the lip flowers *Salvia glutinosa* and *Rhinanthus glacialis*. The ST species are spreadwidely along MDS2(T:18%, M:1%), with *B. sensu-stricto* and *B. pratorum* associated with the lip flower *Origanum vulgare* and *B. soroensis* (Kugler 5.1) favoring the flower head *Centaurea jacea* (Kugler 7.2).

3. *High elevation*:
  - *Early*: MDS1 (T:54%, M:74%) is sort of tongue-length axis, but not perfectly, again because *B. sensu-stricto* is so abberant. In general, the species are fairly widespread with distincntfloral associations -- consistent with the observed decline in overall niche overlap at high elevation -- with the exception of *B. soroensis* and *B. pratorum*, which cluster tightly around the bell flower *Rhododendron hirsutum* (Kugler 3.1). *B. hortorum* likes the flag flower *Anthyllis vulneraria* (Kugler 6.1), *B. wurflenii* like the flag flower *Lotus corniculatus* (Kugler 6.1), *B. pascuorum* (middling as usual) seems to be associated with both the flag flower *Trifolium pratense* (Kugler 6.1) and the bell flower *Rhododendron hirsutum* (Kugler 3.1), and *B. sensu-stricto* seems slightly partial to disk flower *Trollius europaeus* (Kugler 1.2), though again its position suggests that it is exceptional in its generalism. MDS2 (T:28%, M:23%) is mostly about accounting for the deviance of *B. sensu-stricto*.

  - *Mid*: By mid-season at high elevation, the LT species are tightly clustered with each other and associated with the flag blossom *Trifolium pratense* (Kugler 6.1) and the lip flower *Stachys alopecuros* (Kugler 5.1). The ST species are widely spread, with *B. soroensis* and *B. pratorum* suggetsing some affinity for the flower head *Carduus defloratus* (Kugler 7.2) and the bell flower *Campanula scheuchzeri* (Kugler 3.2). *B. pratorum* plots similarly close to *T. pratense*. Again, *B. sensu-stricto* is off on its own with no defining floral association. MDS1 (T:64%, M:91%), MDS2(T:35%, M:9%).

  - *Late*: Finally, late in the season at high elevation, LT and ST species are separated diagonally with respect to MDS1 (T:75%, M:79%) and MDS2(T:20%, M:18%), with LT species falling high on one or both axes and ST species falling low on one or both axes. LT species are loosely associated with the flag blossom *Trifolium pratense* (Kugler 6.1), the funnel flowers *Gentiana asclepiadea* and congener *Gentiana aspera* (Kugler 2.1) and the flower head **(unusual for LT to be associated with flower head!)** *Carlina acaulus* (Kugler 7.2). *B. soroensis* and *B. pratorum* are associated with the flower heads *Carduus defloratus* (Kugler 7.2) and *Scabiosa lucida* (Kugler 7.1) and the bell flower *Campanula scheuchzeri* (Kugler 3.2). *B. sensu-stricto*, ever off on its own, was most closely associated with the lip flower *Thymus pulegioides* (Kugler 5.1).

## 1.4 A follow up to the question of diet breadth in long- vs. short-tongued BB species
The question is which guild -- long-tongued or short-tongued species -- uses greater proportion of available floral species. We will keep this really, really simple, and calculate the normalized degree for each guild. That is just the number of floral species visited divided by the total number of floral species in the network. It ranges between 0 and 1, and 0 and 1 mean exactly what you would expect them to mean. I guess we'll go ahead and calculate the raw degree, too, just so that we can see the absolute diet breadth.

For starters, I'm going to use diet_breit2, which does not parse by period.

**This dovetails with Goulson's hypothesis that long-tongued bees are more specialized than short-tongued bees.** I can test this.

### Script
```{r}
# Calculate die breadth metrics
diet_breadth.tax <- diet_breit2(net, bb_traits, site_data, criterion = plant.sp,
                               indices = c("degree", "normalised degree")) %>%
  mutate(normalised.degree.abundnorm = normalised.degree/abund,
         normalised.degree.propnorm = normalised.degree/prop,
         degree.abundnorm = degree/abund,
         degree.propnorm = degree/prop)

diet_breadth.morph <- diet_breit2(net, bb_traits, site_data, criterion = k.type.s,
                               indices = c("degree", "normalised degree")) %>%
  mutate(normalised.degree.abundnorm = normalised.degree/abund,
         normalised.degree.propnorm = normalised.degree/prop,
         degree.abundnorm = degree/abund,
         degree.propnorm = degree/prop)

# Relative diet breadth ~ elevation
ggplot(diet_breadth.tax, aes(elev.mean, normalised.degree, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

ggplot(diet_breadth.morph, aes(elev.mean, normalised.degree, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

# Absolute diet breadth ~ elevation
ggplot(diet_breadth.tax, aes(elev.mean, degree, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

ggplot(diet_breadth.morph, aes(elev.mean, degree, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

# Relative diet breadth / abundance
ggplot(diet_breadth.tax, aes(elev.mean, normalised.degree.abundnorm, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

ggplot(diet_breadth.morph, aes(elev.mean, normalised.degree.abundnorm, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

# Relative diet breadth / proportional abundance
ggplot(diet_breadth.tax, aes(elev.mean, normalised.degree.propnorm, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

ggplot(diet_breadth.morph, aes(elev.mean, normalised.degree.propnorm, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

# Absolute diet breadth / abundance
ggplot(diet_breadth.tax, aes(elev.mean, degree.abundnorm, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

ggplot(diet_breadth.morph, aes(elev.mean, degree.abundnorm, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

# Absolute diet breadth / proportional abundance
ggplot(diet_breadth.tax, aes(elev.mean, degree.propnorm, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

ggplot(diet_breadth.morph, aes(elev.mean, degree.propnorm, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

# BB abundance ~ elevation
ggplot(diet_breadth.tax, aes(elev.mean, abund, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

# BB proportional abundance ~ elevation
ggplot(diet_breadth.tax, aes(elev.mean, prop, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

# Absolute diet breadth ~ abundance
ggplot(diet_breadth.tax, aes(abund, degree, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "lm")

# Relative diet breadth ~ abundance
ggplot(diet_breadth.tax, aes(abund, normalised.degree, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "lm")

# Relative diet breadth ~ proportional abundance
ggplot(diet_breadth.tax, aes(prop, normalised.degree, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "lm")
```
### Results
In a nutshell, proportional diet breadth is extremely tightly correlated with proportional anbundance. The story here is that LT species are more abundant than ST species at lower elevations, and ST species are more abundant than LT species at higher elevations. This relationship is strong and basically linear, with the relative abundances "crossing" at

## 1.5 Another follow-up. How does the morphological composition of the plant community change with elevation? 
To make the graphs readable, we'll simplify Kugler classes to integers.

### Script
```{r}
# Define periods based on the network data
periods <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
  group_by(elev.class) %>%
  mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
  ungroup() %>%
  select(site, date, period) %>%
  distinct()

# Calculate the absolute and relative abundance of each Kugler class ~ site * period

## Floral survey (constrained to visited species)
survey_morph_abund <- survey %>%
  left_join(periods) %>%
  fill(period, .direction = "downup") %>%
  mutate(k.type.ss = factor(k.type.ss)) %>%
  group_by(site, period, k.type.ss) %>%
  summarize(abund = sum(flower.cover)) %>%
  group_by(site, period) %>%
  mutate(prop = abund/sum(abund)) %>%
  left_join(site_data)

## Visitation network
### Aggregate
visitation_morph_abund <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
  group_by(elev.class) %>%
  mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
  mutate(k.type.ss = factor(k.type.ss)) %>%
  group_by(site, period, k.type.ss) %>%
  summarize(abund = n()) %>%
  group_by(site, period) %>%
  mutate(prop = abund/sum(abund)) %>%
  left_join(site_data)

### By BB species
visitation_morph_abund.sp <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
  group_by(elev.class) %>%
  mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
  mutate(k.type.ss = factor(k.type.ss)) %>%
  group_by(site, period, k.type.ss, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(site, period, bb.sp) %>%
  mutate(prop = abund/sum(abund)) %>%
  left_join(site_data)

### Taxonomic by BB species
visitation_tax_abund.sp <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
  group_by(elev.class) %>%
  mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
  mutate(k.type.ss = factor(k.type.ss),
         bb.sp = factor(bb.sp, levels = c("hort", "pasc", "wurf", "bss", "prat", "soro"))) %>%
  group_by(period, elev.class, plant.genus, k.type.ss, plant.family, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(elev.class, period, bb.sp) %>%
  mutate(prop = abund/sum(abund)) %>%
  arrange(elev.class, period, bb.sp, -prop)

# Morphological composition ~ elevation (survey)
## Relative abundance
ggplot(survey_morph_abund, aes(factor(elev.mean), prop, fill = k.type.ss)) +
  geom_col() +
  facet_wrap(~period)

## Absolute abundance
ggplot(survey_morph_abund, aes(factor(elev.mean), abund, fill = k.type.ss)) +
  geom_col() +
  facet_wrap(~period)

# Morphological composition ~ elevation (visitation)
## Relative abundance 
ggplot(visitation_morph_abund, aes(factor(elev.mean), prop, fill = k.type.ss)) +
  geom_col() +
  facet_wrap(~period)

## Relative abundance ~ species
ggplot(visitation_morph_abund.sp, aes(factor(elev.mean), prop, fill = k.type.ss)) +
  geom_col() +
  facet_grid(bb.sp ~ period)

## Absolute abundance ~ species
ggplot(visitation_morph_abund.sp, aes(factor(elev.mean), abund, fill = k.type.ss)) +
  geom_col() +
  facet_grid(bb.sp ~ period)

## Relative abundance ~ species
ggplot(filter(visitation_tax_abund.sp, period == "early"), aes(factor(elev.mean), prop, fill = names)) +
  geom_col() +
  facet_wrap(~bb.sp)

ggplot(filter(visitation_tax_abund.sp, period == "mid"), aes(factor(elev.mean), prop, fill = names)) +
  geom_col() +
  facet_wrap(~bb.sp)

ggplot(filter(visitation_tax_abund.sp, period == "late"), aes(factor(elev.mean), prop, fill = names)) +
  geom_col() +
  facet_wrap(~bb.sp)

## Absolute abundance ~ species
ggplot(visitation_tax_abund.sp, aes(factor(elev.mean), abund, fill = names)) +
  geom_col() +
  facet_grid(bb.sp ~ period)
```

### Results

## 1.6 Tree plots
This visualization deserves to be pulled out as a stand-alone chunk. It's so good.
```{r}
### Visitation by BB species, parsed by elev.class and period
visitation_abund.sp <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
  group_by(elev.class) %>%
  mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
  mutate(k.type.ss = factor(k.type.ss),
         bb.sp = factor(bb.sp, levels = c("hort", "pasc", "wurf", "bss", "prat", "soro"))) %>%
  group_by(period, elev.class, plant.genus, k.type.ss, plant.family, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(elev.class, period, bb.sp) %>%
  mutate(prop = abund/sum(abund)) %>%
  arrange(elev.class, period, bb.sp, -prop)

# Color palette
cols = c("0" = "gray60", "1" = "#8DD3C7", "2" = "#FFFFB3", "3" = "#BEBADA",
         "4" = "#FB8072", "5" = "#80B1D3", "6" = "#FDB462", "7" = "#B3DE69",
         "8" = "#FCCDE5", "9" = "#D9D9D9")

# Early
ggplot(filter(visitation_abund.sp, period == "early"), 
       aes(area = prop, fill = k.type.ss, label = plant.genus)) +
  geom_treemap(color = "black") +
  geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
  scale_fill_manual(values = cols) + 
  facet_grid(fct_rev(elev.class)~bb.sp)

# Mid
ggplot(filter(visitation_abund.sp, period == "mid"), 
       aes(area = prop, fill = k.type.ss, label = plant.genus)) +
  geom_treemap(color = "black") +
  geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
  scale_fill_manual(values = cols) + 
  facet_grid(fct_rev(elev.class)~bb.sp)

# Late
ggplot(filter(visitation_abund.sp, period == "late"), 
       aes(area = prop, fill = k.type.ss, label = plant.genus)) +
  geom_treemap(color = "black") +
  geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
  scale_fill_manual(values = cols) + 
  facet_grid(fct_rev(elev.class)~bb.sp)

```

## 1.7 How do abundant generalist bumble bees deal with a dynamic floral community? Based on the ordination analysis, it looks like they maintain fairly steady morphological preferences, but they are taxonomically flexible. Let's take a closer look at that.

A question to consider is whether the shallower slope of morphological turnover is caused by the "reluctance" of BBs to switch morphotypes or lag in the floristic turnover of morphotypes relative to species. Morphotypes might be less "sensitive" than species to elevational change, so you first get species turnover within morphotypes, then with enough elevational chnage you start to see the morphotypes turning over. **I can freaking test this with my floral survey data!**

### Aggregate analysis
```{r}
## Binary
### Taxonomic
#### BB species
beta.tax.sp.bin <- beta_func(net, traits = bb_traits, sitedata = site_data, level = bb.sp, 
                             criterion = plant.sp, binary = TRUE) %>%
  mutate(criterion = rep("tax", n()))

#### Tongnue length guild
beta.tax.guild.bin <- beta_func(net, traits = bb_traits, sitedata = site_data, level = pbl.fac,
                                criterion = plant.sp, binary = TRUE) %>%
  mutate(criterion = rep("tax", n()))

#### Plants
betaFL_tax.bin <- beta_func_fl(survey, sitedata = site_data, criterion = plant.sp, binary = TRUE) %>%
  mutate(criterion = rep("tax", n()))

### Morphological
#### BB species
beta.morph.sp.bin <- beta_func(net, traits = bb_traits, sitedata = site_data, level = bb.sp,
                               criterion = k.type.s, binary = TRUE) %>%
  mutate(criterion = rep("morph", n()))

#### Tongue length guild
beta.morph.guild.bin <- beta_func(net, traits = bb_traits, sitedata = site_data, level = pbl.fac,
                                  criterion = k.type.s, binary = TRUE) %>%
  mutate(criterion = rep("morph", n()))

#### Plants
betaFL_morph.bin <- beta_func_fl(survey, sitedata = site_data, criterion = k.type.s, binary = TRUE) %>%
  mutate(criterion = rep("morph", n()))

#### Collated
beta.guild.bin <- bind_rows(beta.tax.guild.bin, beta.morph.guild.bin)
beta.sp.bin <- bind_rows(beta.tax.sp.bin, beta.morph.sp.bin)
beta.fl.bin <- bind_rows(betaFL_tax.bin, betaFL_morph.bin)

## Not binary (i.e. abundance-weighted)
### Taxonomic
#### BB species
beta.tax.sp.wt <- beta_func(net,  traits = bb_traits, sitedata = site_data, level = bb.sp,
                         criterion = plant.sp, binary = FALSE) %>%
  mutate(criterion = rep("tax", n()))

beta.tax.guild.wt <- beta_func(net,  traits = bb_traits, sitedata = site_data, level = pbl.fac,
                            criterion = plant.sp, binary = FALSE) %>%
  mutate(criterion = rep("tax", n()))

beta.morph.sp.wt <- beta_func(net,  traits = bb_traits, sitedata = site_data, level = bb.sp,
                           criterion = k.type.s, binary = FALSE) %>%
  mutate(criterion = rep("morph", n()))

beta.morph.guild.wt <- beta_func(net,  traits = bb_traits, sitedata = site_data, level = pbl.fac, 
                              criterion = k.type.s, binary = FALSE) %>%
  mutate(criterion = rep("morph", n()))

beta.sp.wt <- bind_rows(beta.tax.sp.wt, beta.morph.sp.wt)
beta.guild.wt <- bind_rows(beta.tax.guild.wt, beta.morph.guild.wt)

#### Plants
betaFL_tax.wt <- beta_func_fl(survey, sitedata = site_data, criterion = plant.sp, binary = FALSE) %>%
  mutate(criterion = rep("tax", n()))
betaFL_morph.wt <- beta_func_fl(survey, sitedata = site_data, criterion = k.type.s, binary = FALSE) %>%
  mutate(criterion = rep("morph", n()))
beta.fl.wt <- bind_rows(betaFL_tax.wt, betaFL_morph.wt)

# Pull out ADONIS results
## Binary
beta.tax.sp.adonis.bin <- beta.tax.sp.bin %>%
  select(bb.sp, adonis.r2, adonis.p) %>%
  distinct()

beta.morph.sp.adonis.bin <- beta.morph.sp.bin %>%
  select(bb.sp, adonis.r2, adonis.p) %>%
  distinct()

beta.tax.guild.adonis.bin <- beta.tax.guild.bin %>%
  select(pbl.fac, adonis.r2, adonis.p) %>%
  distinct()

beta.morph.guild.adonis.bin <- beta.morph.guild.bin %>%
  select(pbl.fac, adonis.r2, adonis.p) %>%
  distinct()

## Not binary
beta.tax.sp.adonis <- beta.tax.sp.wt %>%
  select(bb.sp, adonis.r2, adonis.p) %>%
  distinct()

beta.morph.sp.adonis <- beta.morph.sp.wt %>%
  select(bb.sp, adonis.r2, adonis.p) %>%
  distinct()

beta.tax.guild.adonis <- beta.tax.guild.wt %>%
  select(pbl.fac, adonis.r2, adonis.p) %>%
  distinct()

beta.morph.guild.adonis <- beta.morph.guild.wt %>%
  select(pbl.fac, adonis.r2, adonis.p) %>%
  distinct()

# Plot
## Binary
beta.sp.plot.bin <- ggplot(beta.sp.bin, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.05) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
  facet_wrap(~bb.sp) + 
  ggtitle("Species")

beta.guild.plot.bin <- ggplot(beta.guild.bin, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.05) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
  facet_wrap(~pbl.fac) + 
  ggtitle("Tongue guild")

(beta.sp.plot.bin | beta.guild.plot.bin) + plot_layout(guides = 'collect')

## Not binary
beta.sp.plot.wt <- ggplot(beta.sp.wt, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.05) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
  facet_wrap(~bb.sp) + 
  ggtitle("Species")

beta.guild.plot.wt <- ggplot(beta.guild.wt, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.125) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
  facet_wrap(~pbl.fac) + 
  ggtitle("Tongue guild")

(beta.sp.plot.wt | beta.guild.plot.wt) + plot_layout(guides = 'collect')

# Plants
ggplot(beta.fl.bin, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.125) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE)

ggplot(beta.fl.wt, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.125) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE)

# Plotting visitation beta-diversity with floral survey beta-diversity

beta.sp.plot.bin2 <- ggplot(beta.sp.bin, aes(elev.diff, distance, color = criterion)) +
  #geom_point(alpha = 0.05) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = TRUE) +
  geom_smooth(inherit.aes = FALSE, data = beta.fl.wt, 
              aes(elev.diff, distance, color = criterion), 
              linetype = "dashed", se = TRUE,
              method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~bb.sp) + 
  ggtitle("Species") +
  theme_light()

beta.guild.plot.bin2 <- ggplot(beta.guild.bin, aes(elev.diff, distance, color = criterion)) +
  #geom_point(alpha = 0.125) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = TRUE) +
  geom_smooth(inherit.aes = FALSE, data = beta.fl.wt, 
              aes(elev.diff, distance, color = criterion), 
              linetype = "dashed", se = TRUE,
              method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~pbl.fac) + 
  ggtitle("Tongue guild") +
  theme_light()

(beta.sp.plot.bin2 | beta.guild.plot.bin2) + plot_layout(guides = 'collect')

beta.sp.plot.wt2 <- ggplot(beta.sp.wt, aes(elev.diff, distance, color = criterion)) +
  #geom_point(alpha = 0.05) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = TRUE) +
  geom_smooth(inherit.aes = FALSE, data = beta.fl.wt, 
              aes(elev.diff, distance, color = criterion), 
              linetype = "dashed", se = TRUE,
              method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~bb.sp) + 
  ggtitle("Species") +
  theme_light()

beta.guild.plot.wt2 <- ggplot(beta.guild.wt, aes(elev.diff, distance, color = criterion)) +
  #geom_point(alpha = 0.125) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = TRUE) +
  geom_smooth(inherit.aes = FALSE, data = beta.fl.wt, 
              aes(elev.diff, distance, color = criterion), 
              linetype = "dashed", se = TRUE,
              method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~pbl.fac) + 
  ggtitle("Tongue guild") +
  theme_light()

(beta.sp.plot.wt2 | beta.guild.plot.wt2) + plot_layout(guides = 'collect')

# Plotting visitation beta-diversity against floral survey beta-diversity
beta.sp.bin2 <- beta.sp.bin %>%
  full_join(beta.fl.bin, by = c("site1", "site2", "elev1", "elev2", "elev.diff")) %>%
  rename(distance.vis = distance.x, distance.surv = distance.y, 
         criterion.vis = criterion.x, criterion.surv = criterion.y)

beta.guild.bin2 <- beta.guild.bin %>%
  full_join(beta.fl.bin, by = c("site1", "site2", "elev1", "elev2", "elev.diff")) %>%
  rename(distance.vis = distance.x, distance.surv = distance.y, 
         criterion.vis = criterion.x, criterion.surv = criterion.y)

beta.sp.wt2 <- beta.sp.wt %>%
  full_join(beta.fl.wt, by = c("site1", "site2", "elev1", "elev2", "elev.diff")) %>%
  rename(distance.vis = distance.x, distance.surv = distance.y, 
         criterion.vis = criterion.x, criterion.surv = criterion.y)

beta.guild.wt2 <- beta.guild.wt %>%
  full_join(beta.fl.wt, by = c("site1", "site2", "elev1", "elev2", "elev.diff")) %>%
  rename(distance.vis = distance.x, distance.surv = distance.y, 
         criterion.vis = criterion.x, criterion.surv = criterion.y)

ggplot(beta.sp.bin2, aes(distance.surv, distance.vis, color = criterion.vis)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~bb.sp)

ggplot(beta.sp.wt2, aes(distance.surv, distance.vis, color = criterion.vis)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~bb.sp)

ggplot(beta.guild.bin2, aes(distance.surv, distance.vis, color = criterion.vis)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~pbl.fac)

ggplot(beta.guild.wt2, aes(distance.surv, distance.vis, color = criterion.vis)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~bb.guild)

```

### Parsed by period
```{r}
# Binary
## Taxonomic
### BB species
beta_x_period.tax.sp.bin <- beta_func(net, traits = bb_traits, sitedata = site_data, level = bb.sp, criterion = plant.sp, 
                                    binary = TRUE, period) %>%
  mutate(criterion = rep("tax", n()))

### Tongue length guild
beta_x_period.tax.guild.bin <- beta_func(net, traits = bb_traits, sitedata = site_data, level = pbl.fac, criterion = plant.sp, 
                                       binary = TRUE, period) %>%
  mutate(criterion = rep("tax", n()))

## Morphological
### BB species
beta_x_period.morph.sp.bin <- beta_func(net, traits = bb_traits, sitedata = site_data, level = bb.sp, criterion = k.type.s, 
                                    binary = TRUE, period) %>%
  mutate(criterion = rep("morph", n()))

### Tongue length guild
beta_x_period.morph.guild.bin <- beta_func(net, traits = bb_traits, sitedata = site_data, level = pbl.fac, criterion = k.type.s, 
                                       binary = TRUE, period) %>%
  mutate(criterion = rep("morph", n()))

### Collated
beta_x_period.sp.bin <- bind_rows(beta_x_period.tax.sp.bin, beta_x_period.morph.sp.bin)
beta_x_period.guild.bin <- bind_rows(beta_x_period.tax.guild.bin, beta_x_period.morph.guild.bin)

# Not binary
## Taxonomic
### BB species
beta_x_period.tax.sp.wt <- beta_func(net, traits = bb_traits, sitedata = site_data, level = bb.sp, criterion = plant.sp, 
                                    binary = FALSE, period) %>%
  mutate(criterion = rep("tax", n()))

### Tongue length guild
beta_x_period.tax.guild.wt <- beta_func(net, traits = bb_traits, sitedata = site_data, level = pbl.fac, criterion = plant.sp, 
                                       binary = FALSE, period) %>%
  mutate(criterion = rep("tax", n()))

## Morphological
### BB species
beta_x_period.morph.sp.wt <- beta_func(net, traits = bb_traits, sitedata = site_data, level = bb.sp, criterion = k.type.s, 
                                    binary = FALSE, period) %>%
  mutate(criterion = rep("morph", n()))

### Tongue length guild
beta_x_period.morph.guild.wt <- beta_func(net, traits = bb_traits, sitedata = site_data, level = pbl.fac, criterion = k.type.s, 
                                       binary = FALSE, period) %>%
  mutate(criterion = rep("morph", n()))

### Collated
beta_x_period.sp.wt <- bind_rows(beta_x_period.tax.sp.wt, beta_x_period.morph.sp.wt)
beta_x_period.guild.wt <- bind_rows(beta_x_period.tax.guild.wt, beta_x_period.morph.guild.wt)

# Pull out ADONIS results
## Binary
beta_x_period.tax.sp.adonis.bin <- beta_x_period.tax.sp.bin %>%
  select(bb.sp, period, adonis.r2, adonis.p) %>%
  distinct()

beta_x_period.morph.sp.adonis.bin <- beta_x_period.morph.sp.bin %>%
  select(bb.sp, period, adonis.r2, adonis.p) %>%
  distinct()

beta_x_period.tax.guild.adonis.bin <- beta_x_period.tax.guild.bin %>%
  select(pbl.fac, period, adonis.r2, adonis.p) %>%
  distinct()

beta_x_period.morph.guild.adonis.bin <- beta_x_period.morph.guild.bin %>%
  select(pbl.fac, period, adonis.r2, adonis.p) %>%
  distinct()

## Not binary
beta_x_period.tax.sp.adonis <- beta_x_period.tax.sp.wt %>%
  select(bb.sp, adonis.r2, adonis.p) %>%
  distinct()

beta_x_period.morph.sp.adonis <- beta_x_period.morph.sp.wt %>%
  select(bb.sp, adonis.r2, adonis.p) %>%
  distinct()

beta_x_period.tax.guild.adonis <- beta_x_period.tax.guild.wt %>%
  select(pbl.fac, adonis.r2, adonis.p) %>%
  distinct()

beta_x_period.morph.guild.adonis <- beta_x_period.morph.guild.wt %>%
  select(pbl.fac, adonis.r2, adonis.p) %>%
  distinct()

# Plot
## Binary
beta_x_period.sp.plot.bin <- ggplot(beta_x_period.sp.bin, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
  facet_grid(bb.sp ~ period) + 
  ggtitle("Species")

beta_x_period.guild.plot.bin <- ggplot(beta_x_period.guild.bin, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
  facet_grid(pbl.fac ~ period) + 
  ggtitle("Tongue guild")

(beta_x_period.sp.plot.bin | beta_x_period.guild.plot.bin) + plot_layout(guides = 'collect')

## Not binary
beta_x_period.sp.plot.wt <- ggplot(beta_x_period.sp.wt, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
  facet_grid(bb.sp ~ period) + 
  ggtitle("Species")

beta_x_period.guild.plot.wt <- ggplot(beta_x_period.guild.wt, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
  facet_grid(pbl.fac ~ period) + 
  ggtitle("Tongue guild")

(beta_x_period.sp.plot.wt | beta_x_period.guild.plot.wt) + plot_layout(guides = 'collect')

########## A different way to look at this
ggplot(beta_x_period.guild.wt, aes(pbl.fac, distance, color = criterion)) +
  geom_boxplot() +
  facet_wrap( ~ period)

ggplot(beta_x_period.sp.wt, aes(bb.sp, distance, color = criterion)) +
  geom_boxplot() +
  facet_wrap( ~ period) 
```

## 1.8 Does the influence of tongue length on niche partitioning vary with elevation or time?
```{r, echo = FALSE}
# By plant.sp
ggplot(niche_x_elevperiod.tax, aes(elev.class, adonis.r2)) +
  geom_boxplot()

ggplot(niche_x_elevperiod.tax, aes(period, adonis.r2)) +
  geom_boxplot()

# By ktype
ggplot(niche_x_elevperiod.morph, aes(elev.class, adonis.r2)) +
  geom_boxplot()

ggplot(niche_x_elevperiod.morph, aes(period, adonis.r2)) +
  geom_boxplot()
```

# 2. Testing the stress-gradient hypothesis and the Ranta and Vepsäläinen heterogeneity hypothesis

**Stress-gradient hypothesis:** Is there any evidence that niche partitioning amongst generalist core bumble bee species is governed by an inverse relationship between elevation and competition, broadly consistent with the stress-gradient hypothesis (Bertness and Callaway 1994). This mechanism has been inferred from trait patterns in the whole bee community of the Berchtesgadener Alps (Hoiss et al. 2012) and from interaction specialization patterns in the plant-pollinator community of Mt. Kilimanjaro (Classen et al. 2020). It predicts that niche overlap will increase with elevation, reflecting competitive pressures at low elevations and reduced competition but increased need for opportunism at high elevations.

**Ranta/Vepsäläinen heterogeneity hypothesis :** Ranta and Vepsäläinen (1981) proposed that bumble bees communities are governed by a spatiotemporally heterogeneous mosaic of competitive conditions, such that competitive inequalities between species are sufficiently variable to allow more species to coexist than predicted by trait-based competition models.

## 2.1: Testing the stress-gradient hypothesis
### 2.1.1 Face-value analysis

*Results:*

```{r message=FALSE, warning=FALSE, echo = FALSE}
# Pairwise (bb.sp:bb.sp) niche overlap by site, pooled across years

niche_overlap.tax <- niche_overlapper(net, criterion = plant.sp.abb)
niche_overlap.morph <- niche_overlapper(net, criterion = k.type.s)

# All spp pairs
sg1 <- ggplot(niche_overlap.tax, aes(elev.mean, niche.overlap)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

sg2 <- ggplot(niche_overlap.morph, aes(elev.mean, niche.overlap)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

(sg1 + ggtitle("Taxonomic") | sg2 + ggtitle("Morphological")) + plot_layout(guides = 'collect')

## Within LT spp
sg3 <- ggplot(filter(niche_overlap.tax, pbl.fac1 == "long" & pbl.fac2 == "long"), 
       aes(elev.mean, niche.overlap)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

sg4 <- ggplot(filter(niche_overlap.morph, pbl.fac1 == "long" & pbl.fac2 == "long"), 
       aes(elev.mean, niche.overlap)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

(sg3 + ggtitle("Taxonomic") | sg4 + ggtitle("Morphological")) + plot_layout(guides = 'collect')

## Within ST spp
sg5 <- ggplot(filter(niche_overlap.tax, pbl.fac1 == "short" & pbl.fac2 == "short"), 
       aes(elev.mean, niche.overlap)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

sg6 <- ggplot(filter(niche_overlap.morph, pbl.fac1 == "short" & pbl.fac2 == "short"), 
       aes(elev.mean, niche.overlap)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

(sg5 + ggtitle("Taxonomic") | sg6 + ggtitle("Morphological")) + plot_layout(guides = 'collect')

## Between ST and LT
sg7 <- ggplot(filter(niche_overlap.tax, pbl.fac.diff == "different"), 
       aes(elev.mean, niche.overlap)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

sg8 <- ggplot(filter(niche_overlap.morph, pbl.fac.diff == "different"), 
       aes(elev.mean, niche.overlap)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8))

(sg7 + ggtitle("Taxonomic") | sg8 + ggtitle("Morphological")) + plot_layout(guides = 'collect')

# GAMs
glob_GAM.tax <- gam(niche.overlap ~ 
                      s(year, bs  = "re") + # random intercept effect for year
                      s(elev.mean, k = 8, m = 2) + # global smooth ~ elevation
                      s(elev.mean, sp.pair, bs = "fs", k = 8, m = 1), # group-level smooth ~ elevation*sp.pair
                    method = "REML",
                    family = "gaussian",
                    data = niche_overlap.tax) %>% getViz()

check.gamViz(glob_GAM.tax)
summary(glob_GAM.tax)
print(plot(glob_GAM.tax, allTerms = TRUE), pages = 1)

glob_GAM.morph <- gam(niche.overlap ~ 
                      s(year, bs  = "re") + # random intercept effect for year
                      s(elev.mean, k = 8, m = 2) + # global smooth ~ elevation
                      s(elev.mean, sp.pair, bs = "fs", k = 8, m = 1), # group-level smooth ~ elevation*sp.pair
                    method = "REML",
                    family = "gaussian",
                    data = niche_overlap.morph) %>% getViz()

check.gamViz(glob_GAM.morph)
summary(glob_GAM.morph)
print(plot(glob_GAM.morph, allTerms = TRUE), pages = 1)

# Within LT class
LT_GAM.tax <- gam(niche.overlap ~ 
                      s(year, bs  = "re") + # random intercept effect for year
                      s(elev.mean, k = 8, m = 2) + # global smooth ~ elevation
                      s(elev.mean, sp.pair, bs = "fs", k = 8, m = 1), # group-level smooth ~ elevation*sp.pair
                  method = "REML",
                  family = "gaussian",
                  data = filter(niche_overlap.tax, pbl.fac1 == "long" & pbl.fac2 == "long")) %>% getViz()

check.gamViz(LT_GAM.tax)
summary(LT_GAM.tax)
print(plot(LT_GAM.tax, allTerms = TRUE), pages = 1)

LT_GAM.morph <- gam(niche.overlap ~ 
                      s(year, bs  = "re") + # random intercept effect for year
                      s(elev.mean, k = 8, m = 2) + # global smooth ~ elevation
                      s(elev.mean, sp.pair, bs = "fs", k = 8, m = 1), # group-level smooth ~ elevation*sp.pair
                  method = "REML",
                  family = "gaussian",
                  data = filter(niche_overlap.morph, pbl.fac1 == "long" & pbl.fac2 == "long")) %>% getViz()

check.gamViz(LT_GAM.morph)
summary(LT_GAM.morph)
print(plot(LT_GAM.morph, allTerms = TRUE), pages = 1)

# Within ST class
ST_GAM.tax <- gam(niche.overlap ~ 
                      s(year, bs  = "re") + # random intercept effect for year
                      s(elev.mean, k = 8, m = 2) + # global smooth ~ elevation
                      s(elev.mean, sp.pair, bs = "fs", k = 8, m = 1), # group-level smooth ~ elevation*sp.pair
                  method = "REML",
                  family = "gaussian",
                  data = filter(niche_overlap.tax, pbl.fac1 == "short" & pbl.fac2 == "short")) %>% getViz()

check.gamViz(ST_GAM.tax)
summary(ST_GAM.tax)
print(plot(ST_GAM.tax, allTerms = TRUE), pages = 1)

ST_GAM.morph <- gam(niche.overlap ~ 
                      s(year, bs  = "re") + # random intercept effect for year
                      s(elev.mean, k = 8, m = 2) + # global smooth ~ elevation
                      s(elev.mean, sp.pair, bs = "fs", k = 8, m = 1), # group-level smooth ~ elevation*sp.pair
                  method = "REML",
                  family = "gaussian",
                  data = filter(niche_overlap.morph, pbl.fac1 == "short" & pbl.fac2 == "short")) %>% getViz()

check.gamViz(ST_GAM.morph)
summary(ST_GAM.morph)
print(plot(ST_GAM.morph, allTerms = TRUE), pages = 1)

# Between ST and LT
STxLT_GAM.tax <- gam(niche.overlap ~ 
                      s(year, bs  = "re") + # random intercept effect for year
                      s(elev.mean, k = 8, m = 2) + # global smooth ~ elevation
                      s(elev.mean, sp.pair, bs = "fs", k = 8, m = 1), # group-level smooth ~ elevation*sp.pair
                  method = "REML",
                  family = "gaussian",
                  data = filter(niche_overlap.tax, pbl.fac.diff == "different")) %>% getViz()

check.gamViz(STxLT_GAM.tax)
summary(STxLT_GAM.tax)
print(plot(STxLT_GAM.tax, allTerms = TRUE), pages = 1)

STxLT_GAM.morph <- gam(niche.overlap ~ 
                      s(year, bs  = "re") + # random intercept effect for year
                      s(elev.mean, k = 8, m = 2) + # global smooth ~ elevation
                      s(elev.mean, sp.pair, bs = "fs", k = 8, m = 1), # group-level smooth ~ elevation*sp.pair
                  method = "REML",
                  family = "gaussian",
                  data = filter(niche_overlap.morph, pbl.fac.diff == "different")) %>% getViz()

check.gamViz(STxLT_GAM.morph)
summary(STxLT_GAM.morph)
print(plot(STxLT_GAM.morph, allTerms = TRUE), pages = 1)
```

### 2.1.2 Null-model residual analysis

Arguably the most rigorous way to tease out the effect of elevation while controlling for community composition (abundance and richness patterns) is to compare the observed niche overlap at each site to that predicted by a random null model, then see how the residuals vary with elevation.

I'll be the first to admit that I'm new to null models, so I want to be extremely clear about the logic of how I'm setting this up. Step-by-step:  
1. Generate community matrices by site and pollen/nectar
2. Permute each matrix *n* times by randomly redistributing the visits of each bumble bee species while holding the total visits for each bumble bee species constant. This is, therefore, a random selection model constrained to the set of flora visited for a given site-pollen/nectar matrix. Importantly, I am opting not to hold the floral marginal totals constant, i.e. the total number of visits that each plant receives will vary with each permutation. My gut feeling is that this is the right place to start, but this is a decision that requires scrutiny.
3. Calculate niche overlap for each permuted matrix and store as tidy arrays; this mirrors the workflow for calculating empirical niche overlap, the only difference being that I had to deal with some tricky nested lists; I'm pretty sure I did this correctly.
4. Take the mean niche overlap for each permuted community.
5. Join null model mean niche overlap table to empirical niche overlap table, joining by site and pollen/nectar. Thus, for each site-pollen/nectar combo in the empirical niche overlap table, we have *n* iterations of random niche overlap.
6. Calculate the difference between empirical niche overlap and null niche overlap for each iteration; we'll call this "residual niche overlap".
7. Calculate mean residual niche overlap for each site-pollen/nectar combo. This can then be regressed against elevation to see how the deviation from null expectations varies along the elevation gradient, with the interpretation that this represents the "pure" influence of elevation on niche overlap whle controlling for community differences. We could also use the raw differences without collapsing them to a mean. But that's not what I did this first time through.

There are surely a lot of other ways a null model could be constructed and analyzed, but this seems to me a good place to start.

*Results*

```{r message=FALSE, warning=TRUE, echo = FALSE}
# Holding BB marginal totals constant
random_niche.tax <- nullify1(net, 50, plant.sp)
random_niche.morph <- nullify1(net, 50, k.type.s)

# Holding both BB and plant marginal totals constant
# random_niche.tax <- nullify2(net, 50, plant.sp)
# random_niche.morph <- nullify2(net, 50, k.type.s)

## Join null model to empirical niche overlap
# Taxonomic niche overlap
comparison.tax <- niche_overlap.tax %>%
  left_join(random_niche.tax, by = c("site", "sp.pair")) %>%
  mutate(niche.overlap.resid = niche.overlap - niche.overlap.sim) %>%
  group_by(site, elev.mean, sp.pair, pbl.fac1, pbl.fac2, pbl.fac.diff) %>%
  summarize(niche.overlap.resid = mean(niche.overlap.resid))

# Morphological niche overlap
comparison.morph <- niche_overlap.morph %>%
  left_join(random_niche.morph, by = c("site", "sp.pair")) %>%
  mutate(niche.overlap.resid = niche.overlap - niche.overlap.sim) %>%
  group_by(site, elev.mean, sp.pair, pbl.fac1, pbl.fac2, pbl.fac.diff) %>%
  summarize(niche.overlap.resid = mean(niche.overlap.resid))

# Plot
# All spp pairs
sgn1 <- ggplot(comparison.tax, aes(elev.mean, niche.overlap.resid)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

sgn2 <- ggplot(comparison.morph, aes(elev.mean, niche.overlap.resid)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

(sgn1 + ggtitle("Taxonomic") | sgn2 + ggtitle("Morphological")) + plot_layout(guides = 'collect')

## Within LT spp
sgn3 <- ggplot(filter(comparison.tax, pbl.fac1 == "long" & pbl.fac2 == "long"), 
       aes(elev.mean, niche.overlap.resid)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

sgn4 <- ggplot(filter(comparison.morph, pbl.fac1 == "long" & pbl.fac2 == "long"), 
       aes(elev.mean, niche.overlap.resid)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

(sgn3 + ggtitle("Taxonomic") | sgn4 + ggtitle("Morphological")) + plot_layout(guides = 'collect')

## Within ST spp
sgn5 <- ggplot(filter(comparison.tax, pbl.fac1 == "short" & pbl.fac2 == "short"), 
       aes(elev.mean, niche.overlap.resid)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

sgn6 <- ggplot(filter(comparison.morph, pbl.fac1 == "short" & pbl.fac2 == "short"), 
       aes(elev.mean, niche.overlap.resid)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

(sgn5 + ggtitle("Taxonomic") | sgn6 + ggtitle("Morphological")) + plot_layout(guides = 'collect')

## Between ST and LT
sgn7 <- ggplot(filter(comparison.tax, pbl.fac.diff == "different"), 
       aes(elev.mean, niche.overlap.resid)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

sgn8 <- ggplot(filter(comparison.morph, pbl.fac.diff == "different"), 
       aes(elev.mean, niche.overlap.resid)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8))

(sgn7 + ggtitle("Taxonomic") | sgn8 + ggtitle("Morphological")) + plot_layout(guides = 'collect')

# GAMs
glob_nullmodel_GAM.tax <- gam(niche.overlap.resid ~ 
                      s(elev.mean, k = 10, m = 2),
                    method = "REML",
                    family = "gaussian",
                    data = comparison.tax) %>% getViz()

check.gamViz(glob_nullmodel_GAM.tax)
summary(glob_nullmodel_GAM.tax)
print(plot(glob_nullmodel_GAM.tax, allTerms = TRUE), pages = 1)

glob_nullmodel_GAM.morph <- gam(niche.overlap.resid ~ 
                      s(elev.mean, k = 10, m = 2),
                    method = "REML",
                    family = "gaussian",
                    data = comparison.morph) %>% getViz()

check.gamViz(glob_nullmodel_GAM.morph)
summary(glob_nullmodel_GAM.morph)
print(plot(glob_nullmodel_GAM.morph, allTerms = TRUE), pages = 1)

# Within LT class
LT_nullmodel_GAM.tax <- gam(niche.overlap.resid ~ 
                    s(elev.mean, k = 10, m = 2),
                  method = "REML",
                  family = "gaussian",
                  data = filter(comparison.tax, pbl.fac1 == "long" & pbl.fac2 == "long")) %>% getViz()

check.gamViz(LT_nullmodel_GAM.tax)
summary(LT_nullmodel_GAM.tax)
print(plot(LT_nullmodel_GAM.tax, allTerms = TRUE), pages = 1)

LT_nullmodel_GAM.morph <- gam(niche.overlap.resid ~ 
                    s(elev.mean, k = 10, m = 2),
                  method = "REML",
                  family = "gaussian",
                  data = filter(comparison.morph, pbl.fac1 == "long" & pbl.fac2 == "long")) %>% getViz()

check.gamViz(LT_nullmodel_GAM.morph)
summary(LT_nullmodel_GAM.morph)
print(plot(LT_nullmodel_GAM.morph, allTerms = TRUE), pages = 1)

# Within ST class
ST_nullmodel_GAM.tax <- gam(niche.overlap.resid ~ 
                    s(elev.mean, k = 10, m = 2),
                  method = "REML",
                  family = "gaussian",
                  data = filter(comparison.tax, pbl.fac1 == "short" & pbl.fac2 == "short")) %>% getViz()

check.gamViz(ST_nullmodel_GAM.tax)
summary(ST_nullmodel_GAM.tax)
print(plot(ST_nullmodel_GAM.tax, allTerms = TRUE), pages = 1)

ST_nullmodel_GAM.morph <- gam(niche.overlap.resid ~ 
                    s(elev.mean, k = 10, m = 2),
                  method = "REML",
                  family = "gaussian",
                  data = filter(comparison.morph, pbl.fac1 == "short" & pbl.fac2 == "short")) %>% getViz()

check.gamViz(ST_nullmodel_GAM.morph)
summary(ST_nullmodel_GAM.morph)
print(plot(ST_nullmodel_GAM.morph, allTerms = TRUE), pages = 1)

# Between ST and LT
STxLT_nullmodel_GAM.tax <- gam(niche.overlap.resid ~ 
                    s(elev.mean, k = 10, m = 2),
                  method = "REML",
                  family = "gaussian",
                  data = filter(comparison.tax, pbl.fac.diff == "different")) %>% getViz()

check.gamViz(STxLT_nullmodel_GAM.tax)
summary(STxLT_nullmodel_GAM.tax)
print(plot(STxLT_nullmodel_GAM.tax, allTerms = TRUE), pages = 1)

STxLT_nullmodel_GAM.morph <- gam(niche.overlap.resid ~ 
                    s(elev.mean, k = 10, m = 2),
                  method = "REML",
                  family = "gaussian",
                  data = filter(comparison.morph, pbl.fac.diff == "different")) %>% getViz()

check.gamViz(STxLT_nullmodel_GAM.morph)
summary(STxLT_nullmodel_GAM.morph)
print(plot(STxLT_nullmodel_GAM.morph, allTerms = TRUE), pages = 1)
```

### 2.1.3. Parsing niche contraction and niche separation

Contrary to the predictions of H1, niche overlap decreases sharply at high elevation. This is true for both taxonomic and morphological niche overlap, and it holds in both the face-value and null-residual approaches. This decrease in niche overlap might be explained by two modest, proximate hypotheses:

1.  **H1:** The decrease of niche overlap at high elevations is due to *niche contraction*, i.e. BB spp. visiting a smaller proportion of available (and "eligible", i.e. visited at least once by a bumble bee) flora. 
  *Prediction:* "resource range" (Poisot et al. 2012) will decrease with elevation (as Dormann notes, resource range is a "backward" metric insofar as a large resource range means a small proportional niche breadth; I will calculate the inverse of resource range so that it makes more sense)
2.  **H2:** The decrease of niche overlap at high elevations is due to *niche separation*, i.e. BB spp. avoiding one another's niches but not contracting their own.
  *Predictions:* C score will increase with elevation

*Results:* I haven't run formal stats on this analysis yet, but from the visualization, it looks like the overall story is that niche overlap is driven mainly by niche separation, not so much but niche contraction. My first impression is that this suggests a behavior response to competition.

```{r, echo = FALSE}
resource_range.sp <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
  filter(!is.na(pollen)) %>%
  group_by(site, year, pollen, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
  unite(webID, c(site, year, pollen), sep = "_") %>%
  ungroup() %>%
  dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() %>% # convert to to bipartite web
  map(specieslevel, level = "higher", index = "resource range") %>%
  map(rownames_to_column) %>%
  map(as_tibble) %>%
  bind_rows(.id = "webID") %>% # Collapse list into big data frame
  rename(bb.sp = rowname) %>%
  separate(webID, c("site", "year", "pollen"), sep =  "_") %>%
  left_join(site_data) %>%
  mutate(site = factor(site),
         year = factor(year),
         pollen = factor(pollen)) %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # for some reason, one row has a "1" for bb.sp, and I don't know where this comes from; this gets rid of it, but I would still like to know why
  mutate(resource.range = 1-resource.range) %>%
  left_join(BB_covars.splevel) %>%
  filter(bb.sp.abund > 1)

resource_range.sp.mean <- resource_range.sp %>%
  group_by(site, year, elev.mean, pollen) %>%
  summarize(mean.resource.range = mean(resource.range))

ggplot(resource_range.sp, aes(elev.mean, resource.range, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_grid(bb.sp ~ year) +
  ylim(c(0,1))

ggplot(resource_range.sp.mean, aes(elev.mean, mean.resource.range, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year) +
  ylim(c(0,1))

resource_range.ktype <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
  filter(!is.na(pollen)) %>%
  group_by(site, year, pollen, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
  unite(webID, c(site, year, pollen), sep = "_") %>%
  ungroup() %>%
  dplyr::select(higher = bb.sp, lower = k.type.s, webID, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() %>% # convert to to bipartite web
  map(specieslevel, level = "higher", index = "resource range") %>%
  map(rownames_to_column) %>%
  map(as_tibble) %>%
  bind_rows(.id = "webID") %>% # Collapse list into big data frame
  rename(bb.sp = rowname) %>%
  separate(webID, c("site", "year", "pollen"), sep =  "_") %>%
  left_join(site_data) %>%
  mutate(site = factor(site),
         year = factor(year),
         pollen = factor(pollen)) %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # for some reason, one row has a "1" for bb.sp, and I don't know where this comes from; this gets rid of it, but I would still like to know why
  mutate(resource.range = 1-resource.range) %>%
  left_join(BB_covars.splevel) %>%
  filter(bb.sp.abund > 1)

resource_range.ktype.mean <- resource_range.ktype %>%
  group_by(site, year, elev.mean, pollen) %>%
  summarize(mean.resource.range = mean(resource.range))

ggplot(resource_range.ktype, aes(elev.mean, resource.range, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_grid(bb.sp ~ year) +
  ylim(c(0,1))

ggplot(resource_range.ktype.mean, aes(elev.mean, mean.resource.range, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year) +
  ylim(c(0,1))

cscore.sp <- net %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    filter(!is.na(pollen)) %>%
    group_by(site, year, pollen, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
    summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
    group_by(site, year, pollen) %>%
    unite(webID, c(site, year, pollen), sep = "_") %>%
    ungroup() %>%
    dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
    data.frame() %>% # convert to data frame
    frame2webs() %>% # convert to to bipartite web
    map(grouplevel, level = "higher", weighted = TRUE, index = c("C score")) %>% 
    data.frame() %>%
    t() %>%
    data.frame() %>%
    rownames_to_column() %>%
    dplyr::select(webID = rowname, everything()) %>%
    as_tibble() %>%
    separate(webID, c("site", "year", "pollen"), sep =  "_") %>%
    left_join(site_data) %>%
    mutate(site = factor(site),
           year = factor(year),
           pollen = factor(pollen))

ggplot(cscore.sp, aes(elev.mean, C.score.HL, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

cscore.ktype <- net %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    filter(!is.na(pollen)) %>%
    group_by(site, year, pollen, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
    summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
    group_by(site, year, pollen) %>%
    unite(webID, c(site, year, pollen), sep = "_") %>%
    ungroup() %>%
    dplyr::select(higher = bb.sp, lower = k.type.s, webID, freq) %>% # rename columns to match bipartite's expectations
    data.frame() %>% # convert to data frame
    frame2webs() %>% # convert to to bipartite web
    map(grouplevel, level = "higher", weighted = TRUE, index = c("C score")) %>% 
    data.frame() %>%
    t() %>%
    data.frame() %>%
    rownames_to_column() %>%
    dplyr::select(webID = rowname, everything()) %>%
    as_tibble() %>%
    separate(webID, c("site", "year", "pollen"), sep =  "_") %>%
    left_join(site_data) %>%
    mutate(site = factor(site),
           year = factor(year),
           pollen = factor(pollen))

ggplot(cscore.ktype, aes(elev.mean, C.score.HL, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

### 2020-11-26 analyses/old_scripts/BB_abund_GAM.Rmd
did so many things

--- 2020-09-10 analyses/BB_abund_GAM.Rmd
@@
 
-### Tabulate plots
+# Extract predictions
 ```{r}
-bss2010 <- plot(sm(bb_abund.2010, 2)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. terrestris/lucorum")), x = NULL, y = NULL)
-bss2011 <- plot(sm(bb_abund.2011, 2)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. terrestris/lucorum")), x = NULL, y = NULL)
-bss2012 <- plot(sm(bb_abund.2012, 2)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. terrestris/lucorum")), x = NULL, y = NULL)
+# Function to build data set over which to generate predictions
+newdata_build <- function(dat, y) {
 
-gers2010 <- plot(sm(bb_abund.2010, 3)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. gerstaeckeri")), x = NULL, y = NULL)
-gers2011 <- ggplot() #too rare
-gers2012 <- plot(sm(bb_abund.2012, 3)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. gerstaeckeri")), x = NULL, y = NULL)
+  ## yday min/max
+  min_yday <- min(filter(dat, year == y)$yday)
+  max_yday <- max(filter(dat, year == y)$yday)
 
-hort2010 <- plot(sm(bb_abund.2010, 4)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. hortorum")), x = NULL, y = NULL)
-hort2011 <- plot(sm(bb_abund.2011, 3)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. hortorum")), x = NULL, y = NULL)
-hort2012 <- plot(sm(bb_abund.2012, 4)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. hortorum")), x = NULL, y = NULL)
+  ## elev min.max
+  min_elev <- min(filter(dat, year == y)$elev.mean)
+  max_elev <- max(filter(dat, year == y)$elev.mean)
 
-hypn2010 <- plot(sm(bb_abund.2010, 5)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. hypnorum")), x = NULL, y = NULL)
-hypn2011 <- plot(sm(bb_abund.2011, 4)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. hypnorum")), x = NULL, y = NULL)
-hypn2012 <- ggplot()  #too rare
+  ## species list
+  spp <- dat %>%
+    filter(year == y & bb.total.abund >= 5) %>%
+    select(bb.sp) %>%
+    unique() %>%
+    na.omit() ##### NEED TO FIGURE OUT WHERE THE NA COMES FROM!
 
-jone2010 <- plot(sm(bb_abund.2010, 6)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. jonellus")), x = NULL, y = NULL)
-jone2011 <- plot(sm(bb_abund.2011, 5)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. jonellus")), x = NULL, y = NULL)
-jone2012 <- plot(sm(bb_abund.2012, 5)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. jonellus")), x = NULL, y = NULL)
+  # sequence of 50 points along the yday and elev ranges, respectively
+  seq_yday <- seq(min_yday, max_yday, length.out = 50)
+  seq_elev <- seq(min_elev, max_elev, length.out = 50)
 
-lapi2010 <- ggplot() # too rare 
-lapi2011 <- plot(sm(bb_abund.2011, 6)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. lapidarius")), x = NULL, y = NULL)
-lapi2012 <- plot(sm(bb_abund.2012, 6)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. lapidarius")), x = NULL, y = NULL)
+  newdata <- tibble(seq_yday, seq_elev) %>% # combine seqs into data frame
+    complete(seq_yday, seq_elev) %>% # get all possible values
+    crossing(spp) %>% # get all possible combos with species
+    select(yday = seq_yday, elev.mean = seq_elev, bb.sp) # rename columns to match the original model inputs
+}
 
-mend2010 <- plot(sm(bb_abund.2010, 7)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. mendax")), x = NULL, y = NULL)
-mend2011 <- plot(sm(bb_abund.2011, 7)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. mendax")), x = NULL, y = NULL)
-mend2012 <- plot(sm(bb_abund.2012, 7)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. mendax")), x = NULL, y = NULL)
+newdata2010 <- newdata_build(bb_abund_sp, 2010)
+newdata2011 <- newdata_build(bb_abund_sp, 2011)
+newdata2012 <- newdata_build(bb_abund_sp, 2012)
 
-mont2010 <- plot(sm(bb_abund.2010, 8)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. monticola")), x = NULL, y = NULL)
-mont2011 <- plot(sm(bb_abund.2011, 8)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE)+
-  l_points() +
-  labs(title = expression(italic("B. monticola")), x = NULL, y = NULL)
-mont2012 <- plot(sm(bb_abund.2012, 8)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. monticola")), x = NULL, y = NULL)
+# A little wrapper to make predictions and return a tibble with desired columns 
+predict_func <- function(mod, new, y) {
+  predict.bam(mod, newdata = new, type = "response") %>%
+    data.frame() %>%
+    tibble() %>%
+    bind_cols(new) %>%
+    select(yday, elev.mean, bb.sp, pred = 1) %>%
+    mutate(year = rep(y, n()))
+}
 
-muci2010 <- plot(sm(bb_abund.2010, 9)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. mucidis")), x = NULL, y = NULL)
-muci2011 <- plot(sm(bb_abund.2011, 9)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. mucidis")), x = NULL, y = NULL)
-muci2012 <- plot(sm(bb_abund.2012, 9)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. mucidis")), x = NULL, y = NULL)
+# Generate predictions
+predict2010 <- predict_func(bb_abund.2010, newdata2010, 2010)
+predict2011 <- predict_func(bb_abund.2011, newdata2011, 2011)
+predict2012 <- predict_func(bb_abund.2012, newdata2012, 2012)
+predict_pool <- bind_rows(predict2010, predict2011, predict2012)
 
-pasc2010 <- plot(sm(bb_abund.2010, 10)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. pascuorum")), x = NULL, y = NULL)
-pasc2011 <- plot(sm(bb_abund.2011, 10)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. pascuorum")), x = NULL, y = NULL)
-pasc2012 <- plot(sm(bb_abund.2012, 10)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. pascuorum")), x = NULL, y = NULL)
+# Subset predictions
+predict_major <- filter(predict_pool, bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf"))
+predict_highelev <- filter(predict_pool, bb.sp %in% c("pyre", "mend", "mont"))
+predict_minor <- filter(predict_pool, bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci"))
 
-prat2010 <- plot(sm(bb_abund.2010, 11)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. pratorum")), x = NULL, y = NULL)
-prat2011 <- plot(sm(bb_abund.2011, 11)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. pratorum")), x = NULL, y = NULL)
-prat2012 <- plot(sm(bb_abund.2012, 11)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. pratorum")), x = NULL, y = NULL)
+ggplot(predict_major, aes(yday, elev.mean, fill = log(pred))) +
+  geom_raster() +
+  geom_contour(aes(z = log(pred)), color = "black", size = 0.25) +
+  scale_fill_viridis_c() +
+  facet_grid(year~bb.sp)
 
-psit2010 <- plot(sm(bb_abund.2010, 12)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. psithyrus")), x = NULL, y = NULL)
-psit2011 <- plot(sm(bb_abund.2011, 12)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. psithyrus")), x = NULL, y = NULL)
-psit2012 <- plot(sm(bb_abund.2012, 12)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. psithyrus")), x = NULL, y = NULL)
+ggplot(predict_highelev, aes(yday, elev.mean, fill = log(pred))) +
+  geom_raster() +
+  geom_contour(aes(z = log(pred)), color = "black", size = 0.25) +
+  scale_fill_viridis_c() +
+  facet_grid(year~bb.sp)
 
-pyre2010 <- plot(sm(bb_abund.2010, 13)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. pyrenaeus")), x = NULL, y = NULL)
-pyre2011 <- plot(sm(bb_abund.2011, 13)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. pyrenaeus")), x = NULL, y = NULL)
-pyre2012 <- plot(sm(bb_abund.2012, 13)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. pyrenaeus")), x = NULL, y = NULL)
-
-soro2010 <- plot(sm(bb_abund.2010, 14)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. soroensis")), x = NULL, y = NULL)
-soro2011 <- plot(sm(bb_abund.2011, 14)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. soroensis")), x = NULL, y = NULL)
-soro2012 <- plot(sm(bb_abund.2012, 14)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. soroensis")), x = NULL, y = NULL)
-
-wurf2010 <- plot(sm(bb_abund.2010, 15)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. wurflenii")), x = NULL, y = NULL)
-wurf2011 <- plot(sm(bb_abund.2011, 15)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  guides(fill = FALSE) +
-  l_points() +
-  labs(title = expression(italic("B. wurflenii")), x = NULL, y = NULL)
-wurf2012 <- plot(sm(bb_abund.2012, 15)) +
-  l_fitRaster() +
-  l_fitContour(size = 0.1) +
-  l_points() +
-  guides(fill = FALSE) +
-  labs(title = expression(italic("B. wurflenii")), x = NULL, y = NULL)
-
-pdf("./output/BB_abund_major.pdf", height = 11, width = 8.5)
-major_sp <- gridPrint(bss2010, bss2011, bss2012, 
-                      pasc2010, pasc2011, pasc2012,
-                      prat2010, prat2011, prat2012,
-                      psit2010, psit2011, psit2012,
-                      soro2010, soro2011, soro2012,
-                      wurf2010, wurf2011, wurf2012, ncol = 3)
-dev.off()
-
-pdf("./output/BB_abund_highelev.pdf", height = 5.5, width = 8.5)
-high_elev_sp <- gridPrint(pyre2010, pyre2011, pyre2012, 
-                          mend2010, mend2011, mend2012,
-                          mont2010, mont2011, mont2012, ncol = 3)
-dev.off()
-
-pdf("./output/BB_abund_minor.pdf", height = 11, width = 8.5)
-minor_sp <- gridPrint(gers2010, gers2011, gers2012,
-                      hort2010, hort2011, hort2012,
-                      hypn2010, hypn2011, hypn2012,
-                      jone2010, jone2011, jone2012,
-                      lapi2010, lapi2011, lapi2012,
-                      muci2010, muci2011, muci2012, ncol = 3)
-dev.off()
+ggplot(predict_minor, aes(yday, elev.mean, fill = log(pred))) +
+  geom_raster() +
+  geom_contour(aes(z = log(pred)), color = "black", size = 0.25) +
+  scale_fill_viridis_c() +
+  facet_grid(year~bb.sp)
 ```

### 2020-11-26 analyses/old_scripts/flower_choice.Rmd
did so many things

---
title: "flower_choice"
output: html_document
---

# Packages 
```{r}
library(tidyverse)
```

# Load site data and temperature data
```{r, include=FALSE}
site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) # turn elav.class into an ordered factor

gdd <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year)) %>%
  dplyr::select(site, date, gdd.cum)
```

# Load floral taxonomy and trait data
```{r}
fl_tax <- read_csv("../processed_data/floral_tax.csv")

fl_ktype <- read_csv("../processed_data/floral_k_type.csv") %>%
  select(plant.sp, plant.genus = genus, k.type) %>%
  mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"),
         k.type.ss = str_extract(k.type, "[-+]?[0-9]*"))

fl_morph <- read_csv("../processed_data/flower_morphology_GitaBenadi.csv") %>%
  select(SampleID, plant.family = Family, plant.sp = Species, 
         Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam),
            LN = mean(LN))

fl_ktype_key <- read_delim("../processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  select(k.type = 1, description = 2)

fl_traits <- fl_tax %>%
  full_join(fl_ktype) %>%
  full_join(fl_morph) %>%
  select(plant.sp, plant.genus, plant.family, 
         k.type, k.type.s, k.type.ss, diam, LN)
```

# Load BB trait data 
```{r}
bb_traits <- read_csv("../processed_data/bb_traits.csv")
```

# Load network and survey data
```{r, include=FALSE}
# Visitation network
net <- read_csv("../processed_data/network.csv") %>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  mutate(year = factor(year)) %>% # convert year to factor
  left_join(site_data, by = "site") %>%
  left_join(gdd, by = c("site", "date")) %>%
  left_join(fl_traits) %>%
  left_join(bb_traits) %>%
  ungroup() %>%
  arrange(site, date)

visitation <- net %>%
  group_by(site, pollen, date, year, bb.sp, plant.sp) %>%
  summarize(visits = n()) %>%
  left_join(site_data, by = "site") %>%
  left_join(gdd, by = c("site", "date")) %>%
  left_join(fl_traits)

# Floral survey
survey <- read_csv("../processed_data/floral_survey.csv") %>%
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  group_by(site, date) %>%
  left_join(fl_traits) %>%
  left_join(site_data)

survey_sub <- read_csv("../processed_data/floral_survey.csv") %>%
  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  left_join(fl_traits) %>%
  left_join(site_data)
```

# How does the absolute abundance of floral trait classes compare to the visitation rates of floral trait classes?
```{r}
# Whole plant community
## By elevation
ggplot(survey, aes(cut_interval(elev.mean, n = 8), flower.cover, fill = k.type.s)) +
  geom_col(position = "fill") +
  facet_wrap(~year(date), ncol = 1)

ggplot(survey, aes(cut_interval(elev.mean, n = 8), flower.cover, fill = k.type.ss)) +
  geom_col(position = "fill") +
  facet_wrap(~year(date), ncol = 1)

## By week
ggplot(survey, aes(cut_interval(as_date(yday(date)), 10), flower.cover, fill = k.type.s)) +
  geom_col(position = "fill") +
  facet_wrap(~year(date))

ggplot(survey, aes(cut_interval(as_date(yday(date)), 10), flower.cover, fill = k.type.ss)) +
  geom_col(position = "fill") +
  facet_wrap(~year(date))

# Visited plant community
## By elevation
ggplot(survey_sub, aes(cut_interval(elev.mean, n = 8), flower.cover, fill = k.type.s)) +
  geom_col(position = "fill") +
  facet_wrap(~year(date), ncol = 1)

ggplot(survey_sub, aes(cut_interval(elev.mean, n = 8), flower.cover, fill = k.type.ss)) +
  geom_col(position = "fill") +
  facet_wrap(~year(date), ncol = 1)

## By week
ggplot(survey_sub, aes(cut_interval(as_date(yday(date)), 10), flower.cover, fill = k.type.s)) +
  geom_col(position = "fill") +
  facet_wrap(~year(date))

ggplot(survey_sub, aes(cut_interval(as_date(yday(date)), 10), flower.cover, fill = k.type.ss)) +
  geom_col(position = "fill") +
  facet_wrap(~year(date))

# Visitation rates
## By elevation
ggplot(visitation, aes(cut_interval(elev.mean, n = 8), visits, fill = k.type.s)) +
  geom_col(position = "fill") +
  facet_wrap(~year(date), ncol = 1)

ggplot(visitation, aes(cut_interval(elev.mean, n = 8), visits, fill = k.type.s)) +
  geom_col(position = "fill") +
  facet_grid(pollen~bb.sp)

ggplot(visitation, aes(cut_interval(elev.mean, n = 8), visits, fill = k.type.ss)) +
  geom_col(position = "fill") +
  facet_wrap(~year(date), ncol = 1)

## By week
ggplot(filter(visitation, !is.na(pollen)), aes(cut_interval(as_date(yday(date)), 10), visits, fill = k.type.s)) +
  geom_col(position = "fill") +
  facet_grid(pollen~bb.sp)

ggplot(visitation, aes(cut_interval(as_date(yday(date)), 6), visits, fill = k.type.ss)) +
  geom_col(position = "fill") +
  facet_grid(cut_interval(elev.mean, n = 5) ~ year(date))
```

# Tongue length and corolla depth
```{r}
ggplot(net, aes(pbl.w, LN)) +
  geom_point() +
  facet_grid(cut_interval(elev.mean, n = 5) ~ .)

ggplot(net, aes(pbl.w, LN)) +
  geom_point() +
  facet_grid(cut_interval(as_date(yday(date)), 6) ~ .)

```

# Tongue length and elevation
```{r}
ggplot(net, aes(cut_interval(elev.mean, n = 8), fill = factor(pbl.w.class))) +
  geom_bar(position = "fill") +
  facet_wrap(~year)

ggplot(net, aes(cut_interval(elev.mean, n = 8), fill = factor(bb.sp))) +
  geom_bar(position = "fill") +
  facet_wrap(~year)
```

# Do bumble bee species differ in the abundance of the floral resources they select?
```{r}
fl_abund_sp <- survey %>%
  group_by(site, date, plant.sp) %>% # group by plant.sp*transect
  summarize(fl.abund = sum(flower.cover)) %>% # get plant.sp*transect total flower cover; this step really shouldn't be required, but in a few cases, there is more than one abundance entry for the same species on a given transect; not sure how that happened, but it's in the original data that Fabrice sent me; alternatively, I could drop the dups or average them. We'll see what Kalli and Fabrice have to say.
  group_by(site, date) %>%
  mutate(prop.fl.abund = fl.abund/sum(fl.abund)) %>% # get the proportional flower cover for each species.
  ungroup()

abund_vis <- net %>%
  #group_by(site, date, bb.sp, plant.sp) %>%
  #summarize(visits = n()) %>%
  #group_by(site, date, bb.sp) %>%
  #mutate(prop_visits = visits/sum(visits)) %>%
  left_join(fl_abund_sp)

ggplot(abund_vis, aes(bb.sp, prop.fl.abund)) +
  geom_boxplot()

ggplot(abund_vis, aes(elev.mean, log(prop.fl.abund))) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~ bb.sp)

ggplot(abund_vis, aes(yday(date), log(prop.fl.abund))) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~ bb.sp)
```

#
```{r}
ggplot(net, aes(bb.sp, fill = plant.genus)) +
  geom_bar(position = "fill")

ggplot(net, aes(bb.sp, fill = plant.family)) +
  geom_bar(position = "fill")

  ggplot(net, aes(bb.sp, fill = k.type.s)) +
    geom_bar(position = "fill")
```

### 2020-11-26 analyses/old_scripts/network_analysis_bootstrapped.Rmd
did so many things

--- 2020-09-17 analyses/network_analysis.Rmd
@@
   summarize(resource.range = weighted.mean(resource.range, bb.sp.abund, na.rm = TRUE)) %>%
+  #summarize(resource.range = mean(resource.range, na.rm = TRUE)) %>%
   ungroup()

### 2020-11-26 analyses/old_scripts/network_analysis_raw.Rmd
did so many things

---
title: "network analysis raw"
output: html_document
---

### Introduction
Our philosophy in this iteration of the analysis is that we can simply treat the number of transects as a covariate to be controlled statistically. This avoids the need to bootstrap and throw away a lot of data, but we pay a price in terms of model identifiability due to the collinearity between elevation and transect number.

### Prepare environment
```{r setup, message=FALSE, warning=FALSE, include=FALSE}
library(bipartite)
library(mgcv)
library(mgcViz)
library(ggforce)
library(vegan)
library(tidyverse)
library(lubridate)
```

### Load data
```{r message=FALSE, warning=FALSE, include=FALSE}
# Site data
site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, 
                   elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, 
                             levels = c("oben", "mitte", "unten")),
         elev.class2 = factor(elev.class2, 
                             levels = c("high", "mid", "low"))) # turn elav.class into an ordered factor

# Climate data
gdd <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year)) %>%
  dplyr::select(site, date, gdd.cum)

total_gdd <- gdd %>%
  mutate(year = factor(year(date))) %>%
  group_by(site, year) %>%
  summarize(total.gdd = max(gdd.cum))

# Network
net <- read_csv("../processed_data/network.csv") %>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  mutate(year = factor(year)) %>% # convert year to factor
  left_join(site_data, by = "site") %>%
  left_join(gdd, by = c("site", "date")) %>%
  ungroup() %>%
  arrange(site, date)

# Floral survey
survey <- read_csv("../processed_data/floral_survey.csv") %>%
  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover)

# BB traits
bb_traits <- read_csv("../processed_data/bb_traits.csv")

### Auxiliary data, i.e. all the covariates we need to control for
BB_aux <- net %>%
  group_by(site, year, date) %>%
  summarize(bb.rich = length(unique(bb.sp)),
            bb.abund = n()) %>%
  group_by(site, year) %>%
  summarize(bb.rich = mean(bb.rich),
            log.bb.abund = log(mean(bb.abund)),
            bb.transects = length(unique(date)))

FL_aux <- survey %>%
  mutate(year = factor(year(date))) %>%
  filter(flower.cover > 0) %>%
  group_by(site, year, date) %>%
  summarize(fl.rich = length(unique(plant.sp)),
            fl.abund = sum(flower.cover)) %>%
  group_by(site, year) %>%
  summarize(fl.rich = mean(fl.rich),
            log.fl.abund = log(mean(fl.abund)),
            fl.transects = length(unique(date)))
```

### Calculate network/group-level metrics
This is straightforward because none of these metrics involve independent abundance data (i.e. our survey data). We will focus on the network-level metrics of connectance and NODF (nestedness) and the group-level metrics of niche overlap and C score (both are measures of dietary overlap / partitioning)
```{r echo=FALSE, message=FALSE, warning=FALSE}
netmet <- net %>%
  group_by(site, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
  group_by(site, year) %>%
  unite(webID, c(site, year), sep = "_") %>%
  ungroup() %>%
  dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() %>% # convert to to bipartite web
  map(networklevel, level = "higher", weighted = TRUE,
      index = c("connectance", "web asymmetry", "weighted NODF", "C score", "niche overlap")) %>% 
  data.frame() %>%
  t() %>%
  data.frame() %>%
  rownames_to_column() %>%
  dplyr::select(site_year = rowname, everything()) %>%
  as_tibble() %>%
  separate(site_year, c("site", "year"), sep =  "_") %>%
  left_join(site_data) %>%
  left_join(BB_aux) %>%
  left_join(FL_aux) %>%
  left_join(total_gdd) %>%
  mutate(site = factor(site),
         year = factor(year))
```

### Temperature ~ elevation 
The question has been raised of whether we might learn more by using temperature instead of elevation as a predictor variable. We see here, though, that total degree-day accumulation and elevation are almost perfectly correlated, so I think there is little to be gained by using temperature as an explanatory variable. We do, however, see that 2011 and 2012 were substantially warmer than 2010.
```{r  message=FALSE, warning=FALSE, echo = FALSE}
ggplot(netmet, aes(elev.mean, total.gdd, color = year)) +
  geom_point() +
  xlab("Elevation") +
  ylab("Total GDD") +
  theme_gray(14)
```

### Extract species-level metrics
```{r echo = FALSE, message=FALSE, warning=FALSE}
# This is ugly code. The reason is that to calculate d' the right way, we need to use our floral survey data as an independent measure of floral species abundance. I can't figure out a way to wrap this up in a nice neat map() call like I did for the network-level metrics, so instead I am analyzing each site-year one at a time, then collating them. First, the prep work.

# Calculate the mean proportional flower cover of each floral species per site-year
# Complete() expands the array to include all possible site/date/sp  combos, 
# filling in zeros for combos that did not occur.
survey_abundances <- survey %>%
  mutate(year = factor(year(date))) %>%
  filter(flower.cover > 0) %>%

  # get plant.sp*transect total flower cover; this step really shouldn't be required, 
  # but in a few cases, there is more than one abundance entry for the same species on a given transect; 
  # not sure how that happened, but it's in the original data that Fabrice sent me; alternatively, 
  # I could drop the dups or average them.

  group_by(site, date, year, plant.sp) %>%
  summarize(flower.cover = sum(flower.cover)) %>%

  mutate(site = factor(site), # Need to make site and plant.sp factors for complete() to work
         plant.sp = factor(plant.sp)) %>%
  complete(site, date, plant.sp, fill = list(flower.cover = 0)) %>% # get all combos, fill in zeros
  group_by(site, date, year) %>%
  mutate(prop = if_else( # get proportional abundance per site date
    flower.cover > 0, flower.cover/sum(flower.cover), 0)
    ) %>%
  group_by(site, year, plant.sp) %>% # get mean proportional abundance per site-year
  summarize(mean.cover.prop = mean(prop))

# Now, the marginal totals. This enables comparison with the survey abundances, and it
# also enables a convenient join to include only surveyed plants that occur 
# in the visitation network
network_marginal_totals <- net %>%
  group_by(site, date, year, plant.sp) %>%
  summarize(marg.tot = n()) %>% # marginal totals are just the total visits to each plant across all BB spp.
  mutate(site = factor(site),
         plant.sp = factor(plant.sp)) %>%
  complete(site, year, plant.sp, fill = list(marg.tot = 0)) %>% # get all combos and fill in zeros
  group_by(site, date, year)  %>%
  mutate(marg.tot.prop = if_else( # get proportional abundance per site date
    marg.tot > 0, marg.tot/sum(marg.tot), 0)
    ) %>%
  group_by(site, year, plant.sp) %>% 
  summarize(mean.marg.tot.prop = mean(marg.tot.prop))

# Left join survey abundances to network marginal totals
abundance_join <- left_join(network_marginal_totals,
                            survey_abundances) %>%
  arrange(as.character(plant.sp)) %>%

  # For species that we missed in the survey but visited in the network for a given site-year,
  # we will use the proportional marginal totals in place of the survey data rather than
  # having treating visited plants as having zero-abundance 
  replace_na(replace = list(mean.cover.prop = 0)) %>% 
  mutate(abundance = case_when(
    mean.cover.prop > 0 ~ mean.cover.prop,
    mean.cover.prop == 0 ~ mean.marg.tot.prop)
    ) %>%

  filter(mean.marg.tot.prop > 0) # for each site date, drop unvisited species

#################### Extract species metrics ########################
# Turn network into webs
sp_webs <- net %>%
  group_by(site, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
  group_by(site, year) %>%
  unite(webID, c(site, year), sep = "_") %>%
  ungroup() %>%
  dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs(emptylist = TRUE)

# Check names (don't screw this up!)
web_names <- names(sp_webs)

# Extract metrics one freaking web at a time
## Site H1
spmet_h1_2010 <- specieslevel(sp_webs[[1]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "h1", 
                                                year == 2010)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("h1", n()),
         year = rep(2010, n()))

spmet_h1_2011 <- specieslevel(sp_webs[[2]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "h1", 
                                                year == 2011)$mean.cover.prop)  %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("h1", n()),
         year = rep(2011, n()))

spmet_h1_2012 <- specieslevel(sp_webs[[3]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "h1", 
                                                year == 2012)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("h1", n()),
         year = rep(2012, n()))

## Site H2
spmet_h2_2010 <- specieslevel(sp_webs[[4]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "h2", 
                                                year == 2010)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("h2", n()),
         year = rep(2010, n()))

spmet_h2_2011 <- specieslevel(sp_webs[[5]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "h2", 
                                                year == 2011)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("h2", n()),
         year = rep(2011, n()))

spmet_h2_2012 <- specieslevel(sp_webs[[6]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "h2", 
                                                year == 2012)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("h2", n()),
         year = rep(2012, n()))

## Site H3
spmet_h3_2010 <- specieslevel(sp_webs[[7]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "h3", 
                                                year == 2010)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("h3", n()),
         year = rep(2010, n()))

spmet_h3_2011 <- specieslevel(sp_webs[[8]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "h3", 
                                                year == 2011)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("h3", n()),
         year = rep(2011, n()))

spmet_h3_2012 <- specieslevel(sp_webs[[9]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "h3", 
                                                year == 2012)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("h3", n()),
         year = rep(2012, n()))

## Site H4
spmet_h4_2010 <- specieslevel(sp_webs[[10]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "h4", 
                                                year == 2010)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("h4", n()),
         year = rep(2010, n()))

spmet_h4_2011 <- specieslevel(sp_webs[[11]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "h4", 
                                                year == 2011)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("h4", n()),
         year = rep(2011, n()))

spmet_h4_2012 <- specieslevel(sp_webs[[12]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "h4", 
                                                year == 2012)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("h4", n()),
         year = rep(2012, n()))

## Site H5
spmet_h5_2010 <- specieslevel(sp_webs[[13]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "h5", 
                                                year == 2010)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("h5", n()),
         year = rep(2010, n()))

spmet_h5_2011 <- specieslevel(sp_webs[[14]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "h5", 
                                                year == 2011)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("h5", n()),
         year = rep(2011, n()))

spmet_h5_2012 <- specieslevel(sp_webs[[15]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "h5", 
                                                year == 2012)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("h5", n()),
         year = rep(2012, n()))

## Site H6
spmet_h6_2010 <- specieslevel(sp_webs[[16]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "h6", 
                                                year == 2010)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("h6", n()),
         year = rep(2010, n()))

spmet_h6_2011 <- specieslevel(sp_webs[[17]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "h6", 
                                                year == 2011)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("h6", n()),
         year = rep(2011, n()))

spmet_h6_2012 <- specieslevel(sp_webs[[18]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "h6", 
                                                year == 2012)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("h6", n()),
         year = rep(2012, n()))

## Site H7
spmet_h7_2010 <- specieslevel(sp_webs[[19]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "h7", 
                                                year == 2010)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("h7", n()),
         year = rep(2010, n()))

spmet_h7_2011 <- specieslevel(sp_webs[[20]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "h7", 
                                                year == 2011)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("h7", n()),
         year = rep(2011, n()))

spmet_h7_2012 <- specieslevel(sp_webs[[21]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "h7", 
                                                year == 2012)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("h7", n()),
         year = rep(2012, n()))

## Site HO1
spmet_ho1_2010 <- specieslevel(sp_webs[[22]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "ho1", 
                                                year == 2010)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("ho1", n()),
         year = rep(2010, n()))

spmet_ho1_2011 <- specieslevel(sp_webs[[23]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "ho1", 
                                                year == 2011)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("ho1", n()),
         year = rep(2011, n()))

spmet_ho1_2012 <- specieslevel(sp_webs[[24]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "ho1", 
                                                year == 2012)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("ho1", n()),
         year = rep(2012, n()))

## Site ho2
spmet_ho2_2010 <- specieslevel(sp_webs[[25]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "ho2", 
                                                year == 2010)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("ho2", n()),
         year = rep(2010, n()))

spmet_ho2_2011 <- specieslevel(sp_webs[[26]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "ho2", 
                                                year == 2011)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("ho2", n()),
         year = rep(2011, n()))

spmet_ho2_2012 <- specieslevel(sp_webs[[27]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "ho2", 
                                                year == 2012)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("ho2", n()),
         year = rep(2012, n()))

## Site ho3
spmet_ho3_2010 <- specieslevel(sp_webs[[28]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "ho3", 
                                                year == 2010)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("ho3", n()),
         year = rep(2010, n()))

spmet_ho3_2011 <- specieslevel(sp_webs[[29]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "ho3", 
                                                year == 2011)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("ho3", n()),
         year = rep(2011, n()))

spmet_ho3_2012 <- specieslevel(sp_webs[[30]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "ho3", 
                                                year == 2012)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("ho3", n()),
         year = rep(2012, n()))

## Site ho4
spmet_ho4_2010 <- specieslevel(sp_webs[[31]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "ho4", 
                                                year == 2010)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("ho4", n()),
         year = rep(2010, n()))

spmet_ho4_2011 <- specieslevel(sp_webs[[32]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "ho4", 
                                                year == 2011)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("ho4", n()),
         year = rep(2011, n()))

spmet_ho4_2012 <- specieslevel(sp_webs[[33]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "ho4", 
                                                year == 2012)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("ho4", n()),
         year = rep(2012, n()))

## Site ho5
spmet_ho5_2010 <- specieslevel(sp_webs[[34]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "ho5", 
                                                year == 2010)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("ho5", n()),
         year = rep(2010, n()))

spmet_ho5_2011 <- specieslevel(sp_webs[[35]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "ho5", 
                                                year == 2011)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("ho5", n()),
         year = rep(2011, n()))

spmet_ho5_2012 <- specieslevel(sp_webs[[36]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "ho5", 
                                                year == 2012)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("ho5", n()),
         year = rep(2012, n()))

## Site t1
spmet_t1_2010 <- specieslevel(sp_webs[[37]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "t1", 
                                                year == 2010)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("t1", n()),
         year = rep(2010, n()))

spmet_t1_2011 <- specieslevel(sp_webs[[38]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "t1", 
                                                year == 2011)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("t1", n()),
         year = rep(2011, n()))

spmet_t1_2012 <- specieslevel(sp_webs[[39]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "t1", 
                                                year == 2012)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("t1", n()),
         year = rep(2012, n()))

## Site t4
spmet_t4_2010 <- specieslevel(sp_webs[[40]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "t4", 
                                                year == 2010)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("t4", n()),
         year = rep(2010, n()))

spmet_t4_2012 <- specieslevel(sp_webs[[41]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "t4", 
                                                year == 2012)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("t4", n()),
         year = rep(2012, n()))

## Site t5
spmet_t5_2010 <- specieslevel(sp_webs[[42]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "t5", 
                                                year == 2010)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("t5", n()),
         year = rep(2010, n()))

spmet_t5_2011 <- specieslevel(sp_webs[[43]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "t5", 
                                                year == 2011)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("t5", n()),
         year = rep(2011, n()))

spmet_t5_2012 <- specieslevel(sp_webs[[44]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "t5", 
                                                year == 2012)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("t5", n()),
         year = rep(2012, n()))

## Site wma1
spmet_wma1_2010 <- specieslevel(sp_webs[[45]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wma1", 
                                                year == 2010)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wma1", n()),
         year = rep(2010, n()))

spmet_wma1_2011 <- specieslevel(sp_webs[[46]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wma1", 
                                                year == 2011)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wma1", n()),
         year = rep(2011, n()))

spmet_wma1_2012 <- specieslevel(sp_webs[[47]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wma1", 
                                                year == 2012)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wma1", n()),
         year = rep(2012, n()))

## Site wma2
spmet_wma2_2010 <- specieslevel(sp_webs[[48]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wma2", 
                                                year == 2010)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wma2", n()),
         year = rep(2010, n()))

spmet_wma2_2011 <- specieslevel(sp_webs[[49]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wma2", 
                                                year == 2011)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wma2", n()),
         year = rep(2011, n()))

spmet_wma2_2012 <- specieslevel(sp_webs[[50]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wma2", 
                                                year == 2012)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wma2", n()),
         year = rep(2012, n()))

## Site wma3
spmet_wma3_2010 <- specieslevel(sp_webs[[51]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wma3", 
                                                year == 2010)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wma3", n()),
         year = rep(2010, n()))

spmet_wma3_2011 <- specieslevel(sp_webs[[52]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wma3", 
                                                year == 2011)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wma3", n()),
         year = rep(2011, n()))

spmet_wma3_2012 <- specieslevel(sp_webs[[53]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wma3", 
                                                year == 2012)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wma3", n()),
         year = rep(2012, n()))

## Site wma4
spmet_wma4_2010 <- specieslevel(sp_webs[[51]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wma4", 
                                                year == 2010)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wma4", n()),
         year = rep(2010, n()))

spmet_wma4_2011 <- specieslevel(sp_webs[[52]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wma4", 
                                                year == 2011)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wma4", n()),
         year = rep(2011, n()))

spmet_wma4_2012 <- specieslevel(sp_webs[[53]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wma4", 
                                                year == 2012)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wma4", n()),
         year = rep(2012, n()))

## Site wma5
spmet_wma5_2010 <- specieslevel(sp_webs[[54]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wma5", 
                                                year == 2010)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wma5", n()),
         year = rep(2010, n()))

spmet_wma5_2011 <- specieslevel(sp_webs[[55]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wma5", 
                                                year == 2011)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wma5", n()),
         year = rep(2011, n()))

spmet_wma5_2012 <- specieslevel(sp_webs[[56]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wma5", 
                                                year == 2012)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wma5", n()),
         year = rep(2012, n()))

## Site wmb1
spmet_wmb1_2010 <- specieslevel(sp_webs[[57]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wmb1", 
                                                year == 2010)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wmb1", n()),
         year = rep(2010, n()))

spmet_wmb1_2011 <- specieslevel(sp_webs[[58]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wmb1", 
                                                year == 2011)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wmb1", n()),
         year = rep(2011, n()))

spmet_wmb1_2012 <- specieslevel(sp_webs[[59]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wmb1", 
                                                year == 2012)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wmb1", n()),
         year = rep(2012, n()))

## Site wmb2
spmet_wmb2_2010 <- specieslevel(sp_webs[[60]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wmb2", 
                                                year == 2010)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wmb2", n()),
         year = rep(2010, n()))

spmet_wmb2_2011 <- specieslevel(sp_webs[[61]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wmb2", 
                                                year == 2011)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wmb2", n()),
         year = rep(2011, n()))

spmet_wmb2_2012 <- specieslevel(sp_webs[[62]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wmb2", 
                                                year == 2012)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wmb2", n()),
         year = rep(2012, n()))

## Site wmb3
spmet_wmb3_2010 <- specieslevel(sp_webs[[63]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wmb3", 
                                                year == 2010)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wmb3", n()),
         year = rep(2010, n()))

spmet_wmb3_2011 <- specieslevel(sp_webs[[64]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wmb3", 
                                                year == 2011)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wmb3", n()),
         year = rep(2011, n()))

spmet_wmb3_2012 <- specieslevel(sp_webs[[65]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wmb3", 
                                                year == 2012)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wmb3", n()),
         year = rep(2012, n()))

## Site wmb4
spmet_wmb4_2010 <- specieslevel(sp_webs[[66]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wmb4", 
                                                year == 2010)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wmb4", n()),
         year = rep(2010, n()))

spmet_wmb4_2011 <- specieslevel(sp_webs[[67]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wmb4", 
                                                year == 2011)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wmb4", n()),
         year = rep(2011, n()))

spmet_wmb4_2012 <- specieslevel(sp_webs[[68]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wmb4", 
                                                year == 2012)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wmb4", n()),
         year = rep(2012, n()))

## Site wmb5
spmet_wmb5_2010 <- specieslevel(sp_webs[[69]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wmb5", 
                                                year == 2010)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wmb5", n()),
         year = rep(2010, n()))

spmet_wmb5_2011 <- specieslevel(sp_webs[[70]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wmb5", 
                                                year == 2011)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wmb5", n()),
         year = rep(2011, n()))

spmet_wmb5_2012 <- specieslevel(sp_webs[[71]],
                              level = "higher",
                              high.abun = filter(abundance_join, 
                                                site == "wmb5", 
                                                year == 2012)$mean.cover.prop) %>%
  rownames_to_column(var = "bb.sp") %>%
  mutate(site = rep("wmb5", n()),
         year = rep(2012, n()))

# Collate metrics
spmet <- bind_rows(spmet_h1_2010, spmet_h1_2011, spmet_h1_2012,
                   spmet_h2_2010, spmet_h2_2011, spmet_h2_2012,
                   spmet_h3_2010, spmet_h3_2011, spmet_h3_2012,
                   spmet_h4_2010, spmet_h4_2011, spmet_h4_2012,
                   spmet_h5_2010, spmet_h5_2011, spmet_h5_2012,
                   spmet_h6_2010, spmet_h6_2011, spmet_h6_2012,
                   spmet_h7_2010, spmet_h7_2011, spmet_h7_2012,
                   spmet_ho1_2010, spmet_ho1_2011, spmet_ho1_2012,
                   spmet_ho2_2010, spmet_ho2_2011, spmet_ho2_2012,
                   spmet_ho3_2010, spmet_ho3_2011, spmet_ho3_2012,
                   spmet_ho4_2010, spmet_ho4_2011, spmet_ho4_2012,
                   spmet_ho5_2010, spmet_ho5_2011, spmet_ho5_2012,
                   spmet_t1_2010, spmet_t1_2011, spmet_t1_2012,
                   spmet_t4_2010, spmet_t4_2012,
                   spmet_t5_2010, spmet_t5_2011, spmet_t5_2012,
                   spmet_wma1_2010, spmet_wma1_2011, spmet_wma1_2012,
                   spmet_wma2_2010, spmet_wma2_2011, spmet_wma2_2012,
                   spmet_wma3_2010, spmet_wma3_2011, spmet_wma3_2012,
                   spmet_wma4_2010, spmet_wma4_2011, spmet_wma4_2012,
                   spmet_wma5_2010, spmet_wma5_2011, spmet_wma5_2012,
                   spmet_wmb1_2010, spmet_wmb1_2011, spmet_wmb1_2012,
                   spmet_wmb2_2010, spmet_wmb2_2011, spmet_wmb2_2012,
                   spmet_wmb3_2010, spmet_wmb3_2011, spmet_wmb3_2012,
                   spmet_wmb4_2010, spmet_wmb4_2011, spmet_wmb4_2012,
                   spmet_wmb5_2010, spmet_wmb5_2011, spmet_wmb5_2012) %>%
  mutate(site = factor(site),
         year = factor(year)) %>%
  select(site, year, everything()) %>%
  left_join(site_data) %>%
  left_join(bb_traits) %>%
  left_join(FL_aux) %>%
  left_join(BB_aux) %>%
  as_tibble() %>%
  mutate(bb.sp = factor(bb.sp))
```

#### How well does mean proportional marginal total predict mean proportional flower cover?

It's actually not so bad.

```{r message=FALSE, warning=FALSE, echo = FALSE}
ggplot(abundance_join, aes(log(mean.marg.tot.prop), log(mean.cover.prop))) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~year)
```

### Since we have curvy relationships and nested variables, we will analyze our data using hierarchical generalized additive models (HGAMs)
I think there are two ways we could approach this.

1. If our goal is simply to explain network metrics, we can include all covariates (floral abundance, floral richness, BB abundance, BB richness, transect number) in the models, understanding that we have collinearity problems, and see which ones explain the most. Because of the (curvilinear) collinearity between elevation and the other covariates, though, this approach would make it harder to detect and interpret the infleunce of elevation, which is the main question we're interested in.

2. The second approach is to say that while network metrics may, indeed, be influenced by abundance and richness covariates, these are just mediated effects of elevation. We may not care exactly how (i.e. via which mediating covariate) elevation influences network metrics; we might just want to know the shape of the response. But I think we **must** include transects as a covariate even in this approach; otherwise, we would need to go down the road of bootstrapping/rarefaction, and I think we all agree that we don't want to do that.

**My recommendation would be to go with option 2, and that is what will be presented below.**

#### Network/group-level metrics
For these models, we will use the bs = "fs" factor smooth method. This preserves degrees of freedom by forcing all smooths to have the same wiggliness.

**Results**

  * Niche overlap (15% explained)
    + Linear negative effect of elevation
    + No significant effect of transect number
  * C score (30% explained)
    + Positive quadratic/cubic effect of elevation
    + No significant effect of transect number

**Discussion** 
These models have a lot of unexplained variation, but they suggest that the diets of co-existing bumble bee species become increasingly dissimilar (niche overlap) or disaggregated (C score) with increasing elevation. This pattern does not appear to be influenced significantly by transect count.

```{r message=FALSE, warning=FALSE}
# Web asymmetry
gam_web.asym <- gam(web.asymmetry ~ 
                            s(elev.mean, year, bs = "fs", k = 5) +
                            s(bb.transects, year, bs = "fs", k = 4),
                    data = netmet,
                    family = "gaussian",
                    method = "REML",
                    select = FALSE) %>% getViz()

check.gamViz(gam_web.asym) 
summary(gam_web.asym)
plot(sm(gam_web.asym, 1)) +
  l_fitLine(alpha = 0.6)
plot(sm(gam_web.asym, 2)) +
  l_fitLine(alpha = 0.6)

# Niche overlap
gam_niche.overlap <- gam(niche.overlap.HL ~ 
                            s(elev.mean, year, bs = "fs", k = 5) +
                            s(bb.transects, year, bs = "fs", k = 4),
                         data = netmet,
                         family = "gaussian",
                         method = "REML",
                         select = FALSE) %>% getViz()

check.gamViz(gam_niche.overlap) 
summary(gam_niche.overlap)
plot(sm(gam_niche.overlap, 1)) +
  l_fitLine(alpha = 0.6)
plot(sm(gam_niche.overlap, 2)) +
  l_fitLine(alpha = 0.6)

# C score
gam_C.score <- gam(C.score.HL ~ 
                            s(elev.mean, year, bs = "fs", k = 5) +
                            s(bb.transects, year, bs = "fs", k = 4),
                   data = netmet,
                   family = "gaussian",
                   method = "REML",
                   select = FALSE) %>% getViz()

check.gamViz(gam_C.score) 
summary(gam_C.score)
plot(sm(gam_C.score, 1)) +
  l_fitLine(alpha = 0.6)
plot(sm(gam_C.score, 2)) +
  l_fitLine(alpha = 0.6)
```

#### Species-level d'
First let's look at the abundance and distribution of each species in each year. It's probably not worth trying to fit the model to species that are extremely rare or limited in their distirbution. We will drop B. humilis and B. lapidarius from 2010, B. humilis, B. gerstaeckeri, and B. mucidus from 2011, and B. humilis, B. hypnorum, and B. mucidus from 2012.
```{r message=FALSE, warning=FALSE}
bb_range.year <- net %>%
  left_join(site_data) %>%
  group_by(site, date, elev.mean, bb.sp, year) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, bb.sp, year) %>%
  summarize(abund = mean(abund)) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))

ggplot(bb_range.year, aes(reorder(bb.sp, elev.med), elev.mean)) +
  geom_line(size = 2, alpha = 0.25) +
  geom_point(aes(size = abund), alpha = 0.5) +
  facet_wrap(~year, ncol = 1) +
  theme_light(12) +
  scale_size_continuous(name = "Abundance") +
  scale_color_discrete(name = "Abund. class") +
  xlab("BB species") +
  ylab("Elevation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

For species-level d', we will use the "by = factor" factor smooth method. This allows different wiggliness and independent significance tests for each species, which I think is the right way to handle species-level differences in the response of d' to elevation. We will also create separate models for each year so that we can drop the rare species from each year and avoid a complicated interaction term (bb.sp * year) in the by=factor call.

**Results (only significant terms shown)**

* 2010 (33% explained)
  + B. gerstaeckeri (positive, linear)
  + B. pratorum (positive, quadratic)
  + **B. pyrenaeus (negative, quadratic)**
  + B. soroensis (positive, quadratic)
* 2011 (41% explained)
  + B. hortorum (positive, linear)
  + B. psithyrus (positive, linear)
  + Transects (positive, quadratic)
* 2012 (37% explained)
  + **B. sensu-strictu (positive, cubic)**
  + B. hortorum (negative, linear)
  + B. mendax (negative, leveling off above 1500)
  + B. pascuorum (negative, linear)
  + B. pyrenaeus (negative, linear)

**Discussion**
The majority of species do not exhibit a significant relationship between elevation and discrimination, and which species do varies markedly by year. Only B. hortorum and B. pyrenaeus show significant relationships in more than a single year. Transect number was significant only in 2011. **The most interesting finding is that the relationship between elevation and discrimination is generally positive in 2010 and 2011, but it changes to negative in 2012.**

```{r}
### Species-specific models
gam_d.2010 <- gam(d ~ 
                    bb.sp +
                    s(elev.mean, by = bb.sp, bs = "tp", k = 5) +
                    s(bb.transects, k = 4),
                  data = filter(spmet, year == 2010 & 
                                  !bb.sp %in% c("humi", "lapi")),
                  family = "gaussian",
                  method = "REML",
                  select = FALSE) %>% getViz()

check.gamViz(gam_d.2010) 
summary(gam_d.2010)
print(plot(gam_d.2010, select = c(2, 10, 12, 13)), pages = 1)

gam_d.2011 <- gam(d ~ 
                     bb.sp +
                     s(elev.mean, by = bb.sp, bs = "tp", k = 5) +
                     s(bb.transects, k = 4),
                   data = filter(spmet, year == 2011 &
                                   !bb.sp %in% c("humi", "gers", "muci")),
                   family = "gaussian",
                   method = "REML",
                   select = FALSE) %>% getViz()

check.gamViz(gam_d.2011) 
summary(gam_d.2011)
print(plot(gam_d.2011, select = c(2, 10, 14)), pages = 1)

gam_d.2012 <- gam(d ~ 
                     bb.sp +
                     s(elev.mean, by = bb.sp, bs = "tp", k = 5) +
                     s(bb.transects, k = 4),
                   data = filter(spmet, year == 2012 &
                                   !bb.sp %in% c("humi", "hypn", "muci")),
                   family = "gaussian",
                   method = "REML",
                   select = FALSE) %>% getViz()

check.gamViz(gam_d.2012) 
summary(gam_d.2012)
print(plot(gam_d.2012, select = c(1, 3, 6, 8, 11)), pages = 1)
```

### Let's revisit option 1. 
If we're not committed to a story about elevation, can we explain our network metrics by including a larger suite of covariates, some of which may be collinear with elevation?

First, let's inspect collinearity amongst the covariates. A PCA won't be much help here since some of these relationships are nonlinear. So, we'll make a pairs plot and interpret it manually. There's a lot that could be said about these graphs, but here are some highlights.

* Elevation x covariates 
  + Roughly quadratic relationship with BB richness, BB abundance (log-transformed), and floral richness
  + Roughly linear (negative) with floral abundance (log-transformed)
  + More complex relationship with BB and floral transects
* Collinearity among covariates
  + BB richness and BB abundance are very strongly correlated
  + There is also a strong correlation between BB richness and floral richness
  + In 2011, there is a very strong correlation between BB abundance and floral abundance; interestingly, this relationship does not hold up in the other two years

For a start, we will use all our covariates except for floral transects, since that one has to most tangential relationship to d' (it's only effect would be via the estimation of floral species' relative abundances). We will use the select = TRUE option to perform double-penalty variable selection; this potentially penalizes some terms out of the model.
```{r  message=FALSE, warning=FALSE, echo = FALSE}
# 2010
ggplot(filter(netmet, year == 2010) %>% na.omit(), 
       aes(x = .panel_x, y = .panel_y)) + 
  geom_point() +
  geom_autodensity(alpha = 0.5, colour = NA, position = 'identity') + 
  geom_smooth(aes(colour = NULL, fill = NULL)) + 
  facet_matrix(vars(elev.mean, bb.rich, log.bb.abund, fl.rich, 
                    log.fl.abund, bb.transects, fl.transects), 
               layer.diag = 2, 
               grid.y.diag = FALSE) +
  theme_light()

# 2011
ggplot(filter(netmet, year == 2011) %>% na.omit(), 
       aes(x = .panel_x, y = .panel_y)) + 
  geom_point() +
  geom_autodensity(alpha = 0.5, colour = NA, position = 'identity') + 
  geom_smooth(aes(colour = NULL, fill = NULL)) + 
  facet_matrix(vars(elev.mean, bb.rich, log.bb.abund, fl.rich, 
                    log.fl.abund, bb.transects, fl.transects), 
               layer.diag = 2, 
               grid.y.diag = FALSE) +
  theme_light()

# 2012
ggplot(filter(netmet, year == 2010) %>% na.omit(), 
       aes(x = .panel_x, y = .panel_y)) + 
  geom_point() +
  geom_autodensity(alpha = 0.5, colour = NA, position = 'identity') + 
  geom_smooth(aes(colour = NULL, fill = NULL)) + 
  facet_matrix(vars(elev.mean, bb.rich, log.bb.abund, fl.rich, 
                    log.fl.abund, bb.transects, fl.transects), 
               layer.diag = 2, 
               grid.y.diag = FALSE) +
  theme_light()
```

#### Network/group-level metrics
For these models, we will use the bs = "fs" factor smooth method. This preserves degrees of freedom by forcing all smooths to have the same wiggliness.

**Results**

  * Niche overlap (37% explained)
    + Weakly quadratic negative effect of floral richness
    + Linear positive effect of (log) BB abundance
    + No significant effect of elevation
    + Linear or weakly quadratic positive effect of transect count in 2011 and 2012, but negative (weakly quadratic) effect in 2010.
  * C score (57% explained)
    + Cubic negative effect of BB richness.
    + Positive linear effect of elevation
    + Negative linear effect of transect number

**Discussion** 
These models explain a lot more variance than the elevation-only models did. Interestingly, the effect of elevation drops out entirely in the niche overlap model with the addition of the other covariates. This suggests that the effects of elevation may, indeed, be mediated by effects on abundance and diversity, but moreover that abundance and diversity affect network metrics apart from the influence of elevation.

The story for niche overlap seems to be that diets become more similar as floral richness goes down and bumble bee abundance goes up. The relationship with floral richness seems easy to interpret: when there are fewer options, bumble bees are forced to overlap more in their choices. The positive relationship with BB abundance might be a samping effect; when a lot of bumble bees are seen, it is more likely that their observed diets will overlap.

C score increased lineary with elevation, indicating that, as one moves upslope, BB diets tend to become more disaggregated (in a presence/absence sense -- C score is basically an inverse and binary version of niche overlap). It also decreased linearly with transect number, which is probably a sampling effect: with higher sampling intensity, more links are observed, resulting in more observed overlap and less disaggregation. The negative cubic relationship with BB richness -- and remember that BB richness and BB abundance are almost perfectly correlated -- also suggests a sampling effect: when there are more and more kinds of BB, observed overlap is higher. 
```{r message=FALSE, warning=FALSE}
# Niche overlap
gam_niche.overlap.global <- gam(niche.overlap.HL ~ 
                                  s(bb.rich, k = 4) +
                                  s(fl.rich, k = 4) +
                                  s(log.fl.abund, k = 4) +
                                  s(log.bb.abund, k= 4) +
                                  s(elev.mean, year, bs = "fs", k = 5) +
                                  s(bb.transects, year, bs = "fs", k = 4),
                                data = netmet,
                                family = "gaussian",
                                method = "REML",
                                select = TRUE) %>% getViz()

check.gamViz(gam_niche.overlap.global) 
summary(gam_niche.overlap.global)
print(plot(gam_niche.overlap.global, allTerms = T), pages = 1)

plot(sm(gam_niche.overlap.global, 6)) +
  l_fitLine(alpha = 0.6)

# C score
gam_cscore.global <- gam(C.score.HL ~ 
                           s(bb.rich, k = 4) +
                           s(fl.rich, k = 4) +
                           s(log.fl.abund, k = 4) +
                           s(log.bb.abund, k= 4) +
                           s(elev.mean, year, bs = "fs", k = 5) +
                           s(bb.transects, year, bs = "fs", k = 4),
                         data = netmet,
                         family = "gaussian",
                         method = "REML",
                         select = FALSE) %>% getViz()

check.gamViz(gam_cscore.global) 
summary(gam_cscore.global)
print(plot(gam_cscore.global, allTerms = T), pages = 1)

plot(sm(gam_cscore.global, 4)) +
  l_fitLine(alpha = 0.6)

plot(sm(gam_cscore.global, 5)) +
  l_fitLine(alpha = 0.6) 
```

### Species-level d'

**Results (only significant terms shown)**

* 2010 (41% explained)
  + B. gerstaeckeri (positive, linear)
  + B. hortorum (positive, linear)
  + B. mendax (positive, linear)
  + B. pratorum (positive, quadratic)
  + B. psithyrus (positive, linear)
  + **B. pyrenaeus (negative, quadratic)**
  + B. soroensis (positive, quadratic)
* 2011 (58% explained)
  + Negative effect of BB richness (weakly cubic)
  + Positive effect of FL richness (cubic)
  + B. sensu-strictu (positive, linear)
  + B. hortorum (positive, linear)
  + B. hypnorum (positive, linear)
  + B. lapidarius (positive, linear)
  + B. monticola (positive, logistic)
  + B. pascuorum (positive, linear)
  + B. pratorum (positive, logistic)
  + B. psithyrus (positive, linear)
  + B. pyrenaeus (positive, linear)
  + B. soroensis (positive, linear)
  + B. wurflenii (positive, linear)
  + Transects (positive, quadratic)
* 2012 (43% explained)
  + Positive effect of BB richness (linear)
  + Negative effect of FL abundance (cubic)
  + Negative effect of BB abundance (linear)
  + B. hortorum (negative, linear)
  + B. mendax (negative, linear)
  + B. pascuorum (negative, linear)
  + B/ pyrenaeus (negative, linear)

**Discussion**
This is interesting. I do not fully understand why, but the addition of abundance and diversity covariates actually strengthens the apparent effects of elevation. It could also have something to do with setting select = TRUE. The change is most evident in 2011, where most BB species now show a significant relationship with elevation. The overall patterns are the same, though. Elevation has a positive effect on discrimination in 2010 and 2011, but this switches to a negative relationship in 2012.

**Conclusion**

```{r}
### Species-specific models
gam_d.2010.global <- gam(d ~ 
                           s(bb.rich, k = 4) +
                           s(fl.rich, k = 4) +
                           s(log.fl.abund, k = 4) +
                           s(log.bb.abund, k= 4) +
                           bb.sp +
                           s(elev.mean, by = bb.sp, bs = "tp", k = 5) +
                           s(bb.transects, k = 4),
                         data = filter(spmet, year == 2010 & 
                                         !bb.sp %in% c("humi", "lapi")),
                         family = "gaussian",
                         method = "REML",
                         select = FALSE) %>% getViz()

gam_d.2010.global <- gam(d ~ 
                           s(bb.rich, k = 4) +
                           s(fl.rich, k = 4) +
                           s(log.fl.abund, k = 4) +
                           s(log.bb.abund, k= 4) +
                           s(elev.mean, bb.sp, bs = "fs", k = 5) +
                           s(bb.transects, k = 4),
                         data = filter(spmet, year == 2010 & 
                                         !bb.sp %in% c("humi", "lapi")),
                         family = "gaussian",
                         method = "REML",
                         select = FALSE) %>% getViz()

check.gamViz(gam_d.2010.global) 
summary(gam_d.2010.global)
print(plot(gam_d.2010.global, allTerms = TRUE), pages = 1)
print(plot(gam_d.2010.global, select = c(6, 7, 10, 14, 15, 16, 17)), pages = 1)

gam_d.2011.global <- gam(d ~ 
                           s(bb.rich, k = 4) +
                           s(fl.rich, k = 4) +
                           s(log.fl.abund, k = 4) +
                           s(log.bb.abund, k= 4) +
                           bb.sp +
                           s(elev.mean, by = bb.sp, bs = "tp", k = 5) +
                           s(bb.transects, k = 4),
                         data = filter(spmet, year == 2011 &
                                         !bb.sp %in% c("humi", "gers", "muci")),
                         family = "gaussian",
                         method = "REML",
                         select = FALSE) %>% getViz()

gam_d.2011.global <- gam(d ~ 
                           s(bb.rich, k = 4) +
                           s(fl.rich, k = 4) +
                           s(log.fl.abund, k = 4) +
                           s(log.bb.abund, k= 4) +
                           s(elev.mean, bb.sp, bs = "fs", k = 5) +
                           s(bb.transects, k = 4),
                         data = filter(spmet, year == 2011 &
                                         !bb.sp %in% c("humi", "gers", "muci")),
                         family = "gaussian",
                         method = "REML",
                         select = FALSE) %>% getViz()

check.gamViz(gam_d.2011.global) 
summary(gam_d.2011.global)
print(plot(gam_d.2011.global, allTerms = TRUE), pages = 1)
print(plot(gam_d.2011.global, select = c(1, 2, 5, 6, 7, 9, 11, 12, 13, 14, 15, 16, 17, 18)), pages = 1)

gam_d.2012.global <- gam(d ~ 
                           s(bb.rich, k = 4) +
                           s(fl.rich, k = 4) +
                           s(log.fl.abund, k = 4) +
                           s(log.bb.abund, k= 4) +
                           bb.sp +
                           s(elev.mean, by = bb.sp, bs = "tp", k = 5) +
                           s(bb.transects, k = 4),
                         data = filter(spmet, year == 2012 &
                                         !bb.sp %in% c("humi", "hypn", "muci")),
                         family = "gaussian",
                         method = "REML",
                         select = FALSE) %>% getViz()

gam_d.2012.global <- gam(d ~ 
                           s(bb.rich, k = 4) +
                           s(fl.rich, k = 4) +
                           s(log.fl.abund, k = 4) +
                           s(log.bb.abund, k= 4) +
                           s(elev.mean, bb.sp, bs = "fs", k = 5) +
                           s(bb.transects, k = 4),
                         data = filter(spmet, year == 2012 &
                                         !bb.sp %in% c("humi", "hypn", "muci")),
                         family = "gaussian",
                         method = "REML",
                         select = FALSE) %>% getViz()

gam_d.2012.global <- gam(d ~ 
                           s(bb.rich, k = 4) +
                           s(fl.rich, k = 4) +
                           s(log.fl.abund, k = 4) +
                           s(log.bb.abund, k= 4) +
                           s(elev.mean, bb.sp, bs = "fs", k = 5) +
                           s(bb.transects, k = 4),
                         data = spmet,
                         family = "gaussian",
                         method = "REML",
                         select = FALSE) %>% getViz()

check.gamViz(gam_d.2012.global) 
summary(gam_d.2012.global)
print(plot(gam_d.2012.global, select = c(1, 3, 4, 6, 10, 12, 15)), pages = 1)
plot(gam_d.2012.global, select = c(5))
plot(gam_d.2011.global, select = c(5))
plot(gam_d.2010.global, select = c(5))
```

### Conclusion ###
When niche overlap and C score are modeled only as functions of elevation and transect count, we can find interpretable, if weak, elevational patterns. When we include the other covariates, though, the influence of elevation is largely eclipsed by that of BB abundance/richness, and I am not confident that these relationships can be interpreted as anything more than sampling effects. This casts doubt on the legitimacy of the elevation patterns found when abundance and richness covariates are excluded; it's hard to make the case that they are not just sampling effects, too.

### Preliminary visualizations
#### Network metrics ~ elevation
```{r message=FALSE, warning=FALSE, echo = FALSE, include = FALSE}
ggplot(netmet, aes(elev.mean, connectance)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

ggplot(netmet, aes(elev.mean, weighted.NODF)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

#### Group-level metrics ~ elevation
```{r message=FALSE, warning=FALSE, echo = FALSE, include = FALSE}
ggplot(netmet, aes(elev.mean, niche.overlap.HL)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

ggplot(netmet, aes(elev.mean, C.score.HL)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

#### Species-level d' ~ elevation
```{r  message=FALSE, warning=FALSE, echo = FALSE, include = FALSE}
# All together
ggplot(spmet, aes(elev.mean, d)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

ggplot(spmet, aes(elev.mean, d)) +
  geom_point() +
  geom_smooth()

# Broken down by species
ggplot(spmet, aes(elev.mean, d, color = year)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE) +
  facet_wrap(~bb.sp)

ggplot(spmet, aes(log.bb.abund, d, color = year)) +
  geom_point() +
  #geom_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE) +
  facet_wrap(~bb.sp)

ggplot(spmet, aes(log.fl.abund, d, color = year)) +
  geom_point() +
  #geom_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE) +
  facet_wrap(~bb.sp)

```

#### Species-level d' ~ bb.sp, year, and tongue length
```{r  message=FALSE, warning=FALSE, echo = FALSE, include = FALSE}
ggplot(spmet, aes(bb.sp, d)) +
  geom_boxplot()

ggplot(spmet, aes(bb.sp, d, fill = year)) +
  geom_boxplot()

ggplot(spmet, aes(year, d)) +
  geom_boxplot()

ggplot(spmet, aes(pbl.w, d)) +
  geom_point() +
  geom_smooth()
```

### 2020-11-26 analyses/old_scripts/niche.Rmd
did so many things

---
title: "niche"
output: html_document
---

### Prepare environment
```{r setup, message=FALSE, warning=FALSE, include=FALSE}
library(bipartite)
library(mgcv)
library(mgcViz)
library(ggforce)
library(vegan)
library(tidyverse)
library(lubridate)
library(tidymodels)
```

### Load data
```{r message=FALSE, warning=FALSE, include=FALSE}
# Site data
site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, 
                   elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, 
                             levels = c("oben", "mitte", "unten")),
         elev.class2 = factor(elev.class2, 
                             levels = c("high", "mid", "low"))) # turn elav.class into an ordered factor

# Climate data
gdd <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year)) %>%
  dplyr::select(site, date, gdd.cum)

total_gdd <- gdd %>%
  mutate(year = factor(year(date))) %>%
  group_by(site, year) %>%
  summarize(total.gdd = max(gdd.cum))

# Network
net <- read_csv("../processed_data/network.csv") %>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  mutate(year = factor(year)) %>% # convert year to factor
  left_join(site_data, by = "site") %>%
  left_join(gdd, by = c("site", "date")) %>%
  ungroup() %>%
  arrange(site, date)

ggplot(net, aes(bb.sp, fill = pollen)) +
  geom_bar(position = "fill")

# Floral survey
survey <- read_csv("../processed_data/floral_survey.csv") %>%
  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover)

# BB traits
bb_traits <- read_csv("../processed_data/bb_traits.csv")

# FL traits
fl_tax <- read_csv("../processed_data/floral_tax.csv")

fl_ktype <- read_csv("../processed_data/floral_k_type.csv") %>%
  select(plant.sp, plant.genus = genus, k.type) %>%
  mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"),
         k.type.ss = str_extract(k.type, "[-+]?[0-9]*"))

fl_morph <- read_csv("../processed_data/flower_morphology_GitaBenadi.csv") %>%
  select(SampleID, plant.family = Family, plant.sp = Species, 
         Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam),
            LN = mean(LN))

fl_ktype_key <- read_delim("../processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  select(k.type = 1, description = 2)

fl_traits <- fl_tax %>%
  full_join(fl_ktype) %>%
  full_join(fl_morph) %>%
  select(plant.sp, plant.genus, plant.family, 
         k.type, k.type.s, k.type.ss, diam, LN)

### Auxiliary data, i.e. all the covariates we need to control for
BB_covars <- net %>%
  group_by(site, year, date) %>%
  summarize(bb.rich = length(unique(bb.sp)),
            sqrt.bb.abund = sqrt(n()))

FL_covars <- survey %>%
  mutate(year = factor(year(date))) %>%
  filter(flower.cover > 0) %>%
  group_by(site, year, date) %>%
  summarize(fl.rich = length(unique(plant.sp)),
            log.fl.abund = log(sum(flower.cover)))

BB_covars.mean <- net %>%
  group_by(site, year, date) %>%
  summarize(bb.rich = length(unique(bb.sp)),
            bb.abund = n()) %>%
  group_by(site, year) %>%
  summarize(mean.bb.rich = mean(bb.rich),
            mean.sqrt.bb.abund = sqrt(mean(bb.abund)))

FL_covars.mean <- survey %>%
  mutate(year = factor(year(date))) %>%
  filter(flower.cover > 0) %>%
  group_by(site, year, date) %>%
  summarize(fl.rich = length(unique(plant.sp)),
            fl.abund = sum(flower.cover))  %>%
  group_by(site, year) %>%
  summarize(mean.fl.rich = mean(fl.rich),
            mean.log.fl.abund = log(mean(fl.abund)))
```

### Get pariwise niche overlap
```{r eval=FALSE, include=FALSE}
### Species-level by transect
niche_overlap.sp <- net %>%
  group_by(bb.sp, plant.sp, site, date, year) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
  group_by(bb.sp, site, date, year) %>%
  mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
  select(-freq) %>% # drop freq variable
  pivot_wider(names_from = plant.sp, values_from = prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  group_by(site, date, year) %>%
  nest() %>% # nest into list-column
  mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
         dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
         tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
  select(site, date, year, tidy_dist) %>% # drop intermediate list-columns
  unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
  left_join(site_data) %>% # add site data
  select(site, date, year, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # select and rename vars
  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
  rename(subgenus1 = subgenus, pbl.w1 = pbl.w, pbl.w.class1 = pbl.w.class) %>%
  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
  rename(subgenus2 = subgenus, pbl.w2 = pbl.w, pbl.w.class2 = pbl.w.class) %>%
  select(site, date, year, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
         pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
  mutate(pbl.diff = abs(pbl.w1 - pbl.w2), # tongue length differences
         yday = yday(date), # yday variable will come in handy
         subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different"), # binary: same or different subgenus

         # Logical columns indicating whether a given species belongs to a given species pair
         # This is handy for extracting data for individual species across all their counterparts
         bss = if_else(bb.sp1 == "bss" | bb.sp2 == "bss", TRUE, FALSE),
         pasc = if_else(bb.sp1 == "pasc" | bb.sp2 == "pasc", TRUE, FALSE),
         wurf = if_else(bb.sp1 == "wurf" | bb.sp2 == "wurf", TRUE, FALSE),
         prat = if_else(bb.sp1 == "prat" | bb.sp2 == "prat", TRUE, FALSE),
         soro = if_else(bb.sp1 == "soro" | bb.sp2 == "soro", TRUE, FALSE),
         hort = if_else(bb.sp1 == "hort" | bb.sp2 == "hort", TRUE, FALSE),
         jone = if_else(bb.sp1 == "jone" | bb.sp2 == "jone", TRUE, FALSE),
         psit = if_else(bb.sp1 == "psit" | bb.sp2 == "psit", TRUE, FALSE),
         mont = if_else(bb.sp1 == "mont" | bb.sp2 == "mont", TRUE, FALSE),
         hypn = if_else(bb.sp1 == "hypn" | bb.sp2 == "hypn", TRUE, FALSE),
         lapi = if_else(bb.sp1 == "lapi" | bb.sp2 == "lapi", TRUE, FALSE),
         muci = if_else(bb.sp1 == "muci" | bb.sp2 == "muci", TRUE, FALSE),
         gers = if_else(bb.sp1 == "gers" | bb.sp2 == "gers", TRUE, FALSE),
         mend = if_else(bb.sp1 == "mend" | bb.sp2 == "mend", TRUE, FALSE),
         pyre = if_else(bb.sp1 == "pyre" | bb.sp2 == "pyre", TRUE, FALSE),
         humi = if_else(bb.sp1 == "humi" | bb.sp2 == "humi", TRUE, FALSE)) %>%

  unite(sp.pair, c(bb.sp1, bb.sp2), sep = "_") %>% # create a single species pair column...
  mutate(sp.pair = factor(sp.pair)) %>% # ...and make it a factor so that mgcv can handle it
  left_join(BB_covars) %>% # add BB covars
  left_join(FL_covars) %>% # add floral covars
  mutate(niche.overlap = 1-distance) # express niche overlap so that 1 = perfect overlap and 0 = no overlap

niche_overlap.sp.mean <- niche_overlap.sp %>%
  group_by(sp.pair, site, year, elev.mean, pbl.diff, subgenus.diff, 
           bss, pasc, wurf, prat, soro, hort, jone, psit, mont, hypn, 
           lapi, muci, gers, mend, pyre, humi) %>%
  summarize(mean.niche.overlap = mean(niche.overlap)) %>%
  left_join(BB_covars.mean) %>%
  left_join(FL_covars.mean)

niche_overlap.sp.mean2x <- niche_overlap.sp %>%
  group_by(site, year, date, elev.mean) %>% # Average across species pairs within transects
  summarize(mean.niche.overlap = mean(niche.overlap)) %>% 
  group_by(site, year, elev.mean) %>% # Average across transects
  summarize(mean.niche.overlap = mean(mean.niche.overlap)) %>% 
  left_join(BB_covars.mean) %>%
  left_join(FL_covars.mean)

### Genus-level by transect
niche_overlap.gen <- net %>%
  group_by(bb.sp, plant.genus, site, date, year) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
  group_by(bb.sp, site, date, year) %>%
  mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
  select(-freq) %>% # drop freq variable
  pivot_wider(names_from = plant.genus, values_from = prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  group_by(site, date, year) %>%
  nest() %>% # nest into list-column
  mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
         dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
         tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
  select(site, date, year, tidy_dist) %>% # drop intermediate list-columns
  unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
  left_join(site_data) %>% # add site data
  select(site, date, year, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # select and rename vars
  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
  rename(subgenus1 = subgenus, pbl.w1 = pbl.w, pbl.w.class1 = pbl.w.class) %>%
  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
  rename(subgenus2 = subgenus, pbl.w2 = pbl.w, pbl.w.class2 = pbl.w.class) %>%
  select(site, date, year, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
         pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
  mutate(pbl.diff = abs(pbl.w1 - pbl.w2), # tongue length differences
         yday = yday(date), # yday variable will come in handy
         subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different"), # binary: same or different subgenus

         # Logical columns indicating whether a given species belongs to a given species pair
         # This is handy for extracting data for individual species across all their counterparts
         bss = if_else(bb.sp1 == "bss" | bb.sp2 == "bss", TRUE, FALSE),
         pasc = if_else(bb.sp1 == "pasc" | bb.sp2 == "pasc", TRUE, FALSE),
         wurf = if_else(bb.sp1 == "wurf" | bb.sp2 == "wurf", TRUE, FALSE),
         prat = if_else(bb.sp1 == "prat" | bb.sp2 == "prat", TRUE, FALSE),
         soro = if_else(bb.sp1 == "soro" | bb.sp2 == "soro", TRUE, FALSE),
         hort = if_else(bb.sp1 == "hort" | bb.sp2 == "hort", TRUE, FALSE),
         jone = if_else(bb.sp1 == "jone" | bb.sp2 == "jone", TRUE, FALSE),
         psit = if_else(bb.sp1 == "psit" | bb.sp2 == "psit", TRUE, FALSE),
         mont = if_else(bb.sp1 == "mont" | bb.sp2 == "mont", TRUE, FALSE),
         hypn = if_else(bb.sp1 == "hypn" | bb.sp2 == "hypn", TRUE, FALSE),
         lapi = if_else(bb.sp1 == "lapi" | bb.sp2 == "lapi", TRUE, FALSE),
         muci = if_else(bb.sp1 == "muci" | bb.sp2 == "muci", TRUE, FALSE),
         gers = if_else(bb.sp1 == "gers" | bb.sp2 == "gers", TRUE, FALSE),
         mend = if_else(bb.sp1 == "mend" | bb.sp2 == "mend", TRUE, FALSE),
         pyre = if_else(bb.sp1 == "pyre" | bb.sp2 == "pyre", TRUE, FALSE),
         humi = if_else(bb.sp1 == "humi" | bb.sp2 == "humi", TRUE, FALSE)) %>%

  unite(sp.pair, c(bb.sp1, bb.sp2), sep = "_") %>% # create a single species pair column...
  mutate(sp.pair = factor(sp.pair)) %>% # ...and make it a factor so that mgcv can handle it
  left_join(BB_covars) %>% # add BB covars
  left_join(FL_covars) %>% # add floral covars
  mutate(niche.overlap = 1-distance) # express niche overlap so that 1 = perfect overlap and 0 = no overlap

niche_overlap.gen.mean <- niche_overlap.gen %>%
  group_by(sp.pair, site, year, elev.mean, pbl.diff, subgenus.diff, 
           bss, pasc, wurf, prat, soro, hort, jone, psit, mont, hypn, 
           lapi, muci, gers, mend, pyre, humi) %>%
  summarize(mean.niche.overlap = mean(niche.overlap)) %>%
  left_join(BB_covars.mean) %>%
  left_join(FL_covars.mean)

### K-type by transect
niche_overlap.ktype <- net %>%
  left_join(fl_traits) %>%
  group_by(bb.sp, k.type.s, site, date, year) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
  group_by(bb.sp, site, date, year) %>%
  mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
  select(-freq) %>% # drop freq variable
  pivot_wider(names_from = k.type.s, values_from = prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  group_by(site, date, year) %>%
  nest() %>% # nest into list-column
  mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
         dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
         tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
  select(site, date, year, tidy_dist) %>% # drop intermediate list-columns
  unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
  left_join(site_data) %>% # add site data
  select(site, date, year, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # select and rename vars
  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
  rename(subgenus1 = subgenus, pbl.w1 = pbl.w, pbl.w.class1 = pbl.w.class) %>%
  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
  rename(subgenus2 = subgenus, pbl.w2 = pbl.w, pbl.w.class2 = pbl.w.class) %>%
  select(site, date, year, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
         pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
  mutate(pbl.diff = abs(pbl.w1 - pbl.w2), # tongue length differences
         yday = yday(date), # yday variable will come in handy
         subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different"), # binary: same or different subgenus

         # Logical columns indicating whether a given species belongs to a given species pair
         # This is handy for extracting data for individual species across all their counterparts
         bss = if_else(bb.sp1 == "bss" | bb.sp2 == "bss", TRUE, FALSE),
         pasc = if_else(bb.sp1 == "pasc" | bb.sp2 == "pasc", TRUE, FALSE),
         wurf = if_else(bb.sp1 == "wurf" | bb.sp2 == "wurf", TRUE, FALSE),
         prat = if_else(bb.sp1 == "prat" | bb.sp2 == "prat", TRUE, FALSE),
         soro = if_else(bb.sp1 == "soro" | bb.sp2 == "soro", TRUE, FALSE),
         hort = if_else(bb.sp1 == "hort" | bb.sp2 == "hort", TRUE, FALSE),
         jone = if_else(bb.sp1 == "jone" | bb.sp2 == "jone", TRUE, FALSE),
         psit = if_else(bb.sp1 == "psit" | bb.sp2 == "psit", TRUE, FALSE),
         mont = if_else(bb.sp1 == "mont" | bb.sp2 == "mont", TRUE, FALSE),
         hypn = if_else(bb.sp1 == "hypn" | bb.sp2 == "hypn", TRUE, FALSE),
         lapi = if_else(bb.sp1 == "lapi" | bb.sp2 == "lapi", TRUE, FALSE),
         muci = if_else(bb.sp1 == "muci" | bb.sp2 == "muci", TRUE, FALSE),
         gers = if_else(bb.sp1 == "gers" | bb.sp2 == "gers", TRUE, FALSE),
         mend = if_else(bb.sp1 == "mend" | bb.sp2 == "mend", TRUE, FALSE),
         pyre = if_else(bb.sp1 == "pyre" | bb.sp2 == "pyre", TRUE, FALSE),
         humi = if_else(bb.sp1 == "humi" | bb.sp2 == "humi", TRUE, FALSE)) %>%

  unite(sp.pair, c(bb.sp1, bb.sp2), sep = "_") %>% # create a single species pair column...
  mutate(sp.pair = factor(sp.pair)) %>% # ...and make it a factor so that mgcv can handle it
  left_join(BB_covars) %>% # add BB covars
  left_join(FL_covars) %>% # add floral covars
  mutate(niche.overlap = 1-distance) # express niche overlap so that 1 = perfect overlap and 0 = no overlap

niche_overlap.ktype.mean <- niche_overlap.ktype %>%
  group_by(sp.pair, site, year, elev.mean, pbl.diff, subgenus.diff, 
           bss, pasc, wurf, prat, soro, hort, jone, psit, mont, hypn, 
           lapi, muci, gers, mend, pyre, humi) %>%
  summarize(mean.niche.overlap = mean(niche.overlap)) %>%
  left_join(BB_covars.mean) %>%
  left_join(FL_covars.mean)
```

### Niche-overlap by year  
```{r}
ggplot(niche_overlap.sp, aes(year, niche.overlap)) +
  geom_boxplot()

ggplot(niche_overlap.gen, aes(year, niche.overlap)) +
  geom_boxplot()

ggplot(niche_overlap.ktype, aes(year, niche.overlap)) +
  geom_boxplot()

ggplot(niche_overlap.sp.mean, aes(year, mean.niche.overlap)) +
  geom_boxplot()

ggplot(niche_overlap.gen.mean, aes(year, mean.niche.overlap)) +
  geom_boxplot()

ggplot(niche_overlap.ktype.mean, aes(year, mean.niche.overlap)) +
  geom_boxplot()

ggplot(niche_overlap.sp.mean2x, aes(year, mean.niche.overlap)) +
  geom_boxplot()

ggplot(niche_overlap.gen.mean2x, aes(year, mean.niche.overlap)) +
  geom_boxplot()

ggplot(niche_overlap.ktype.mean2x, aes(year, mean.niche.overlap)) +
  geom_boxplot()
```

### Niche-overlap by subgenus
```{r}
ggplot(niche_overlap.sp, aes(subgenus.diff, niche.overlap)) +
  geom_boxplot()

ggplot(niche_overlap.gen, aes(subgenus.diff, niche.overlap)) +
  geom_boxplot()

ggplot(niche_overlap.ktype, aes(subgenus.diff, niche.overlap)) +
  geom_boxplot()

ggplot(niche_overlap.sp.mean, aes(subgenus.diff, mean.niche.overlap)) +
  geom_boxplot()

ggplot(niche_overlap.gen.mean, aes(subgenus.diff, mean.niche.overlap)) +
  geom_boxplot()

ggplot(niche_overlap.ktype.mean, aes(subgenus.diff, mean.niche.overlap)) +
  geom_boxplot()
```

### Niche-overlap by tongue length differences
```{r}
ggplot(niche_overlap.sp, aes(sqrt(pbl.diff), niche.overlap)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4))

ggplot(niche_overlap.gen, aes(sqrt(pbl.diff), niche.overlap)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4))

ggplot(niche_overlap.ktype, aes(sqrt(pbl.diff), niche.overlap)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4))

ggplot(niche_overlap.sp.mean, aes(sqrt(pbl.diff), mean.niche.overlap)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4))

ggplot(niche_overlap.gen.mean, aes(sqrt(pbl.diff), mean.niche.overlap)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4))

ggplot(niche_overlap.ktype.mean, aes(sqrt(pbl.diff), mean.niche.overlap)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) 
```

### Niche-overlap by elevation  
```{r}
ggplot(niche_overlap.sp, aes(elev.mean, niche.overlap)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~sp.pair) +
  #facet_wrap(~year) +
  ylim(c(0,1))

ggplot(niche_overlap.gen, aes(elev.mean, niche.overlap)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~sp.pair) +
  #facet_wrap(~year) +
  ylim(c(0,1))

ggplot(niche_overlap.ktype, aes(elev.mean, niche.overlap)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~sp.pair) +
  #facet_wrap(~year) +
  ylim(c(0,1))

ggplot(niche_overlap.sp.mean, aes(elev.mean, mean.niche.overlap)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~sp.pair) +
  #facet_wrap(~year) +
  ylim(c(0,1))

ggplot(niche_overlap.sp.mean2x, aes(elev.mean, mean.niche.overlap)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~year)

```

### Niche overlap BB/FL covars
```{r}
# Species level
ggplot(niche_overlap.sp, aes(fl.rich, distance)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  #facet_wrap(~sp.pair) +
  #facet_wrap(~year) +
  ylim(c(0,1))

ggplot(niche_overlap.sp, aes(log.fl.abund, distance)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  #facet_wrap(~sp.pair) +
  #facet_wrap(~year) +
  ylim(c(0,1))

ggplot(niche_overlap.sp, aes(bb.rich, distance)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  #facet_wrap(~sp.pair) +
  #facet_wrap(~year) +
  ylim(c(0,1))

ggplot(niche_overlap.sp, aes(sqrt.bb.abund, distance)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  #facet_wrap(~sp.pair) +
  #facet_wrap(~year) +
  ylim(c(0,1))

ggplot(niche_overlap.sp.mean, aes(mean.fl.rich, mean.dist)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  #facet_wrap(~sp.pair) +
  #facet_wrap(~year) +
  ylim(c(0,1))

ggplot(niche_overlap.sp.mean, aes(mean.bb.rich, mean.dist)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  #facet_wrap(~sp.pair) +
  #facet_wrap(~year) +
  ylim(c(0,1))

ggplot(niche_overlap.sp.mean, aes(mean.sqrt.bb.abund, mean.dist)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  #facet_wrap(~sp.pair) +
  #facet_wrap(~year) +
  ylim(c(0,1))

ggplot(niche_overlap.sp.mean, aes(mean.log.fl.abund, mean.dist)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  #facet_wrap(~sp.pair) +
  #facet_wrap(~year) +
  ylim(c(0,1))
```

### A first try at some GAMs
```{r}
no.sp <- bam(distance ~
               s(year, bs = "re") +
               s(sp.pair, bs = "re") +
               s(pbl.diff, k = 4) +
               #s(bb.rich, k = 4) +
               s(fl.rich, k = 4) +
               #s(sqrt.bb.abund, k = 4) +
               s(log.fl.abund, k = 4) +
               te(yday, elev.mean, bs= c("gp", "tp"), k = 10),
             data = niche_overlap.sp,
             discrete = TRUE,
             family = "gaussian",
             select = TRUE,
             method = "fREML") %>% getViz()

no.sp <- bam(distance ~
               s(year, bs = "re") +
               s(sp.pair, bs = "re") +
               s(pbl.diff, k = 4) +
               #s(bb.rich, k = 4) +
               #s(fl.rich, k = 4) +
               #s(sqrt.bb.abund, k = 4) +
               #s(log.fl.abund, k = 4) +
               s(yday, by = year, bs = "gp", k = 12) +
               s(elev.mean, by = year, k = 10),
             data = niche_overlap.sp,
             discrete = TRUE,
             family = "gaussian",
             select = TRUE,
             method = "fREML") %>% getViz()

check.gamViz(no.sp)
summary(no.sp)
print(plot(no.sp, allTerms = TRUE), pages = 1)
concurvity(no.sp, full = FALSE)

no.ktype <- bam(distance ~
               s(year, bs = "re") +
               s(sp.pair, bs = "re") +
               s(pbl.diff, k = 4) +
               s(bb.rich, k = 4) +
               s(fl.rich, k = 4) +
               s(sqrt.bb.abund, k = 4) +
               s(log.fl.abund, k = 4) +
               te(yday, elev.mean, bs= c("gp", "tp"), k = 10),
             data = niche_overlap.ktype,
             discrete = TRUE,
             family = "gaussian",
             select = TRUE,
             method = "fREML") %>% getViz()

check.gamViz(no.ktype)
summary(no.ktype)
print(plot(no.ktype, allTerms = TRUE), pages = 1)
concurvity(no.ktype, full = TRUE)

no.sp.mean <- bam(mean.dist ~
            s(year, bs = "re") +
            s(sp.pair, bs= "re") +
            s(pbl.diff, k = 4) +
            s(mean.bb.rich, k = 4) +
            s(mean.fl.rich, k = 4) +
            #s(mean.sqrt.bb.abund, k = 4) +
            s(mean.log.fl.abund, k = 4) +
            s(elev.mean, k = 10),
         data = niche_overlap.sp.mean,
         family = "gaussian",
         select = TRUE,
         discrete = TRUE,
         method = "fREML") %>% getViz()

no.sp.mean <- bam(mean.dist ~
            s(sp.pair, bs= "re") +
            s(pbl.diff, k = 4) +
            #s(mean.bb.rich, year, bs = "fs", k = 4) +
            s(mean.fl.rich, year, bs = "fs", k = 4) +
            #s(mean.sqrt.bb.abund, year, bs = "fs", k = 4) +
            s(mean.log.fl.abund, year, bs = "fs", k = 4) +
            s(elev.mean, year, bs= "fs", k = 10),
         data = niche_overlap.sp.mean,
         family = "gaussian",
         select = TRUE,
         discrete = TRUE,
         method = "fREML") %>% getViz()

no.sp.mean <- bam(mean.dist ~
            s(year, bs = "re") +
            s(sp.pair, bs= "re") +
            s(pbl.diff, k = 4) +
            #s(mean.bb.rich, year, bs = "fs", k = 4) +
            s(mean.fl.rich, by = year, k = 4) +
            #s(mean.sqrt.bb.abund, year, bs = "fs", k = 4) +
            s(mean.log.fl.abund, by = year, k = 4) +
            s(elev.mean, by = year, k = 10),
         data = niche_overlap.sp.mean,
         family = "gaussian",
         select = TRUE,
         discrete = TRUE,
         method = "fREML") %>% getViz()

no.sp.mean <- bam(mean.dist ~
            s(year, bs = "re") +
            s(sp.pair, bs= "re") +
            s(pbl.diff, k = 4) +
            #s(mean.bb.rich, year, bs = "fs", k = 4) +
            #s(mean.fl.rich, by = year, k = 4) +
            #s(mean.sqrt.bb.abund, year, bs = "fs", k = 4) +
            #s(mean.log.fl.abund, by = year, k = 4) +
            s(elev.mean, by = year, k = 8),
         data = niche_overlap.sp.mean,
         family = "gaussian",
         select = TRUE,
         discrete = TRUE,
         method = "fREML") %>% getViz()

check.gamViz(no.sp.mean)
summary(no.sp.mean)
print(plot(no.sp.mean, allTerms = TRUE), pages = 1)

ggplot(niche_overlap.sp.mean, aes(elev.mean, mean.dist)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

ggplot(niche_overlap.sp.mean, aes(mean.bb.rich, mean.sqrt.bb.abund)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

### Covariates
```{r}
ggplot(niche_overlap.sp %>% na.omit(), 
       aes(x = .panel_x, y = .panel_y)) + 
  geom_point(alpha = 0.025) +
  geom_autodensity(alpha = 0.5, colour = NA, position = 'identity') + 
  geom_smooth(aes(colour = NULL, fill = NULL)) + 
  facet_matrix(vars(elev.mean, yday, bb.rich, sqrt.bb.abund, fl.rich, 
                    log.fl.abund, distance), 
               layer.diag = 2, 
               grid.y.diag = FALSE) +
  theme_light()

ggplot(niche_overlap.sp.mean %>% na.omit(), 
       aes(x = .panel_x, y = .panel_y)) + 
  geom_point(alpha = 0.025) +
  geom_autodensity(alpha = 0.5, colour = NA, position = 'identity') + 
  geom_smooth(aes(colour = NULL, fill = NULL)) + 
  facet_matrix(vars(elev.mean, mean.bb.rich, mean.sqrt.bb.abund, mean.fl.rich, 
                    mean.log.fl.abund, mean.dist), 
               layer.diag = 2, 
               grid.y.diag = FALSE) +
  theme_light()

covars.mean <- niche_overlap.sp.mean %>%
  ungroup() %>%
  select(mean.bb.rich, mean.fl.rich, mean.log.fl.abund, mean.sqrt.bb.abund) %>%
  distinct() %>%
  na.omit()
pca_covars.mean <- rda(covars.mean)
biplot(pca_covars.mean)
```

### Niche overlap by time
```{r}
ggplot(niche_overlap.sp, aes(yday(date), distance)) +
  geom_point() +
  geom_smooth() +
  #facet_wrap(~sp.pair) +
  ylim(c(0,1))

ggplot(niche_overlap.gen, aes(yday(date), distance)) +
  geom_point() +
  geom_smooth() +
  ylim(c(0,1))

ggplot(niche_overlap.ktype, aes(yday(date), distance)) +
  geom_point() +
  geom_smooth() +
  ylim(c(0,1))
```

### Dare we try a GAM or two?
```{r}
gam_bss <- bam(distance ~ 
                 te(yday, elev.mean, 
                    bs= c("gp", "cr"), 
                    k=c(8, 8), 
                    m=2),
               data = filter(niche_overlap.sp, bss == TRUE),
               discrete = TRUE,
               family = "quasibinomial",
               select = TRUE,
               method = "fREML") %>% getViz(nsim = 500)

check.gamViz(gam_bss)
summary(gam_bss)
print(plot(gam_bss, allTerms = TRUE), pages = 1)
gam_bss.ck <- check2D(gam_bss, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals

gam_prat <- bam(distance ~ 
                 te(yday, elev.mean, 
                    bs= c("gp", "cr"), 
                    k=c(8, 8), 
                    m=2),
               data = filter(niche_overlap.sp.sitedate, prat == TRUE),
               discrete = TRUE,
               family = "gaussian",
               select = TRUE,
               method = "fREML") %>% getViz(nsim = 500)

check.gamViz(gam_prat)
summary(gam_prat)
print(plot(gam_prat, allTerms = TRUE), pages = 1)
gam_prat.ck <- check2D(gam_prat, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals

gam_pyre <- bam(distance ~ 
                 te(yday, elev.mean, 
                    bs= c("gp", "cr"), 
                    k=c(12, 12), 
                    m=2),
               data = filter(niche_overlap.sp.sitedate, pyre == TRUE),
               discrete = TRUE,
               family = "gaussian",
               select = TRUE,
               method = "fREML") %>% getViz(nsim = 500)

check.gamViz(gam_pyre)
summary(gam_pyre)
print(plot(gam_pyre, allTerms = TRUE), pages = 1)
gam_pyre.ck <- check2D(gam_pyre, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals

gam_pasc <- bam(distance ~ 
                 te(yday, elev.mean, 
                    bs= c("gp", "cr"), 
                    k=c(10, 10), 
                    m=2),
               data = filter(niche_overlap.sp.sitedate, pasc == TRUE),
               discrete = TRUE,
               family = "gaussian",
               select = TRUE,
               method = "fREML") %>% getViz(nsim = 500)

check.gamViz(gam_pasc)
summary(gam_pasc)
print(plot(gam_pasc, allTerms = TRUE), pages = 1)
gam_pasc.ck <- check2D(gam_pasc, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals

gam_wurf <- bam(distance ~ 
                 te(yday, elev.mean, 
                    bs= c("gp", "cr"), 
                    k=c(10, 10), 
                    m=2),
               data = filter(niche_overlap.sp.sitedate, wurf == TRUE),
               discrete = TRUE,
               family = "quasibinomial",
               select = TRUE,
               method = "fREML") %>% getViz(nsim = 500)

check.gamViz(gam_wurf)
summary(gam_wurf)
print(plot(gam_wurf, allTerms = TRUE), pages = 1)
gam_wurf.ck <- check2D(gam_wurf, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals
```

### Other stuff
```{r}

net_ord_genus_dist <- vegdist(net_ord_genus, method = "horn")
net_ord_genus_clust <- hclust(net_ord_genus_dist, "ward.D")
plot(net_ord_genus_clust)
groups <- cutree(net_ord_genus_clust, 4)

nmds_genus <- metaMDS(net_ord_genus, distance = "horn")
ordiplot(nmds_genus, type = "t")

ordiplot(nmds_genus, type = 'n')
points(nmds_genus, pch = groups, col = groups)

# Species level

net_ord_species_dist <- vegdist(net_ord_species, method = "horn")
net_ord_species_clust <- hclust(net_ord_species_dist, "ward.D")
plot(net_ord_species_clust)
groups <- cutree(net_ord_species_clust, 4)

nmds_species <- metaMDS(net_ord_species, distance = "horn")
ordiplot(nmds_species, type = "t")

ordiplot(nmds_species, type = 'n')
points(nmds_species, pch = groups, col = groups)
```

### Module analysis
First for the metaweb
```{r eval=FALSE, include=FALSE}
metaweb <- net %>%
  filter(year == 2012) %>%
  separate(plant.sp, c("plant.genus", "species"), sep = " ") %>%
  group_by(bb.sp, plant.genus) %>%
  summarize(freq = n()) %>%
  pivot_wider(names_from = bb.sp, values_from = freq) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("plant.genus")

metaweb_modules <- computeModules(metaweb)

pdf("./output/modules.pdf", height = 11, width = 8.5)
plotModuleWeb(metaweb_modules, plotModules = TRUE, weighted =  TRUE,
              labsize = 0.5)
dev.off()
```

### 2020-11-26 analyses/old_scripts/niche_partitioning.Rmd
did so many things

---
title: "Flexibility and constraints in resource selection within a generalist pollinator guild"
output: html_document
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(bipartite)
library(mgcv)
library(mgcViz)
library(ggbipart)
library(GGally)
library(vegan)
library(ggvegan)
library(tidyverse)
library(lubridate)
library(patchwork)
library(tidymodels)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
### Load data

# Site data
site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, 
                   elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, 
                             levels = c("oben", "mitte", "unten")),
         elev.class2 = factor(elev.class2, 
                             levels = c("high", "mid", "low"))) # turn elav.class into an ordered factor

# Climate data
gdd <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year)) %>%
  dplyr::select(site, date, gdd.cum)

total_gdd <- gdd %>%
  mutate(year = factor(year(date))) %>%
  group_by(site, year) %>%
  summarize(total.gdd = max(gdd.cum))

# Network
net <- read_csv("../processed_data/network.csv") %>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  mutate(year = factor(year)) %>% # convert year to factor
  left_join(site_data, by = "site") %>%
  left_join(gdd, by = c("site", "date")) %>%
  ungroup() %>%
  arrange(site, date)

ggplot(net, aes(bb.sp, fill = pollen)) +
  geom_bar(position = "fill")

# Floral survey
survey <- read_csv("../processed_data/floral_survey.csv") %>%
  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover)

# BB traits
bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf"))

# FL traits
fl_tax <- read_csv("../processed_data/floral_tax.csv")

fl_ktype <- read_csv("../processed_data/floral_k_type.csv") %>%
  select(plant.sp, plant.genus = genus, k.type) %>%
  mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"),
         k.type.ss = str_extract(k.type, "[-+]?[0-9]*"))

fl_morph <- read_csv("../processed_data/flower_morphology_GitaBenadi.csv") %>%
  select(SampleID, plant.family = Family, plant.sp = Species, 
         Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam),
            LN = mean(LN))

fl_ktype_key <- read_delim("../processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  select(k.type = 1, description = 2)

fl_traits <- fl_tax %>%
  full_join(fl_ktype) %>%
  full_join(fl_morph) %>%
  select(plant.sp, plant.genus, plant.family, 
         k.type, k.type.s, k.type.ss, diam, LN)

### Auxiliary data, i.e. all the covariates we need to control for
BB_covars <- net %>%
  group_by(site, year, date) %>%
  summarize(bb.rich = length(unique(bb.sp)),
            sqrt.bb.abund = sqrt(n()))

FL_covars <- survey %>%
  mutate(year = factor(year(date))) %>%
  filter(flower.cover > 0) %>%
  group_by(site, year, date) %>%
  summarize(fl.rich = length(unique(plant.sp)),
            log.fl.abund = log(sum(flower.cover)))

BB_covars.site <- net %>%
  group_by(site, year) %>%
  summarize(bb.rich = length(unique(bb.sp)),
            sqrt.bb.abund = sqrt(n()))

FL_covars.site <- survey %>%
  mutate(year = factor(year(date))) %>%
  filter(flower.cover > 0) %>%
  group_by(site, year) %>%
  summarize(fl.rich = length(unique(plant.sp)),
            log.fl.abund = log(sum(flower.cover)))

BB_covars.mean <- net %>%
  group_by(site, year, date) %>%
  summarize(bb.rich = length(unique(bb.sp)),
            bb.abund = n()) %>%
  group_by(site, year) %>%
  summarize(mean.bb.rich = mean(bb.rich),
            mean.sqrt.bb.abund = sqrt(mean(bb.abund)))

FL_covars.mean <- survey %>%
  mutate(year = factor(year(date))) %>%
  filter(flower.cover > 0) %>%
  group_by(site, year, date) %>%
  summarize(fl.rich = length(unique(plant.sp)),
            fl.abund = sum(flower.cover))  %>%
  group_by(site, year) %>%
  summarize(mean.fl.rich = mean(fl.rich),
            mean.log.fl.abund = log(mean(fl.abund)))

### Species names and such for plotting
species_names <- cbind(c("bss", "pasc", "prat", 
                         "soro", "wurf", "hort", 
                         "pyre", "mend", "mont", 
                         "gers", "hypn","jone", 
                         "lapi", "muci", "psit", 
                         "humi"),
                       c("B. sensu-strictu", "B. pascuorum", "B. pratorum", 
                         "B. soroensis", "B. wurflenii", "B. hortorum",
                         "B. pyrenaeus", "B. mendax", "B. monticola",
                         "B. gerstaeckeri", "B. hypnorum", "B. jonellus", 
                         "B. lapidarius", "B. mucidus", "B. psithyrus", 
                         "B. humilis"),
                       c("B. ss.", "B. pasc.", "B. prat.", 
                         "B. soro.", "B. wurf.", "B. hort.",
                         "B. pyre.", "B. mend.", "B. mont.",
                         "B. gers.", "B. hypn.", "B. jone.", 
                         "B. lapi.", "B. muci.", "B. psit.", 
                         "B. humi"),
                       c("major", "major", "major", 
                         "major", "major", "common", 
                         "high elev.", "high elev.", "high elev.",
                         "minor", "minor", "common", 
                         "minor", "minor", "common", 
                         "minor")
                       ) %>%
  as.data.frame() %>%
  as_tibble() %>%
  dplyr::select(bb.sp = 1, species.name = 2, species.abb = 3, abund.class = 4) %>%
  mutate(focal = if_else(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf"), TRUE, FALSE))
```

```{r message=FALSE, warning=FALSE, include=FALSE}
### Species-level by transect
bb_abund.transect <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  group_by(bb.sp, site, date, year) %>%
  summarize(bb.sp.abund = n())

bb_abund.site <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  group_by(bb.sp, site, year) %>%
  summarize(bb.sp.abund = n())

niche_overlap.sp <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
  filter(!is.na(pollen)) %>% # drop observations with no pollen/nectar classification
  group_by(bb.sp, pollen, plant.sp, site, date, year) %>% # group by transect and BB-FL pair
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per transect
  group_by(bb.sp, pollen, site, date, year) %>% # group by transect and BB species
  mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
  select(-freq) %>% # drop freq variable
  pivot_wider(names_from = plant.sp, values_from = prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  group_by(pollen, site, date, year) %>%
  nest() %>% # nest into list-column
  mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
         dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
         tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
  select(pollen, site, date, year, tidy_dist) %>% # drop intermediate list-columns
  unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
  left_join(site_data) %>% # add site data
  select(site, date, year, pollen, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # select and rename vars
  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
  rename(subgenus1 = subgenus, pbl.w1 = pbl.w, pbl.w.class1 = pbl.w.class) %>%
  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
  rename(subgenus2 = subgenus, pbl.w2 = pbl.w, pbl.w.class2 = pbl.w.class) %>%
  select(site, date, year, pollen, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
         pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
  left_join(bb_abund.transect, by = c("bb.sp1" = "bb.sp", "site", "date", "year")) %>%
  rename(bb.sp1.abund = bb.sp.abund) %>%
  select(site, date, year, pollen, bb.sp1, bb.sp2, bb.sp1.abund, subgenus1, subgenus2, distance, 
         pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
  left_join(bb_abund.transect, by = c("bb.sp2" = "bb.sp", "site", "date", "year")) %>%
  rename(bb.sp2.abund = bb.sp.abund) %>%
  select(site, date, year, pollen, bb.sp1, bb.sp2, bb.sp1.abund, bb.sp2.abund, subgenus1, subgenus2, distance, 
         pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
  mutate(pbl.diff = abs(pbl.w1 - pbl.w2), # tongue length differences
         yday = yday(date), # yday variable will come in handy
         abund.diff = abs(bb.sp1.abund - bb.sp2.abund),
         subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different"), # binary: same or different subgenus

         # Logical columns indicating whether a given species belongs to a given species pair
         # This is handy for extracting data for individual species across all their counterparts
         bss = if_else(bb.sp1 == "bss" | bb.sp2 == "bss", TRUE, FALSE),
         pasc = if_else(bb.sp1 == "pasc" | bb.sp2 == "pasc", TRUE, FALSE),
         wurf = if_else(bb.sp1 == "wurf" | bb.sp2 == "wurf", TRUE, FALSE),
         prat = if_else(bb.sp1 == "prat" | bb.sp2 == "prat", TRUE, FALSE),
         soro = if_else(bb.sp1 == "soro" | bb.sp2 == "soro", TRUE, FALSE),
         hort = if_else(bb.sp1 == "hort" | bb.sp2 == "hort", TRUE, FALSE)) %>%

  unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
  mutate(sp.pair = factor(sp.pair)) %>% # ...and make it a factor so that mgcv can handle it
  left_join(BB_covars) %>% # add BB covars
  left_join(FL_covars) %>% # add floral covars
  mutate(niche.overlap = 1-distance) # express niche overlap so that 1 = perfect overlap and 0 = no overlap

niche_overlap.sp_site <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
  filter(!is.na(pollen)) %>% # drop observations with no pollen/nectar classification
  group_by(bb.sp, pollen, plant.sp, site, year) %>% # group by site-year and BB-FL pair
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-year
  group_by(bb.sp, pollen, site, year) %>% # group by site-year and BB species
  mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
  select(-freq) %>% # drop freq variable
  pivot_wider(names_from = plant.sp, values_from = prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  group_by(pollen, site, year) %>%
  nest() %>% # nest into list-column
  mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
         dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
         tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
  select(pollen, site, year, tidy_dist) %>% # drop intermediate list-columns
  unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
  left_join(site_data) %>% # add site data
  select(site, year, pollen, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # select and rename vars
  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
  rename(subgenus1 = subgenus, pbl.w1 = pbl.w, pbl.w.class1 = pbl.w.class) %>%
  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
  rename(subgenus2 = subgenus, pbl.w2 = pbl.w, pbl.w.class2 = pbl.w.class) %>%
  select(site, year, pollen, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
         pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
  left_join(bb_abund.site, by = c("bb.sp1" = "bb.sp", "site", "year")) %>%
  rename(bb.sp1.abund = bb.sp.abund) %>%
  select(site, year, pollen, bb.sp1, bb.sp2, bb.sp1.abund, subgenus1, subgenus2, distance, 
         pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
  left_join(bb_abund.site, by = c("bb.sp2" = "bb.sp", "site", "year")) %>%
  rename(bb.sp2.abund = bb.sp.abund) %>%
  select(site, year, pollen, bb.sp1, bb.sp2, bb.sp1.abund, bb.sp2.abund, subgenus1, subgenus2, distance, 
         pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
  mutate(pbl.diff = abs(pbl.w1 - pbl.w2), # tongue length differences
         abund.diff = abs(bb.sp1.abund - bb.sp2.abund),
         subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different"), # binary: same or different subgenus

         # Logical columns indicating whether a given species belongs to a given species pair
         # This is handy for extracting data for individual species across all their counterparts
         bss = if_else(bb.sp1 == "bss" | bb.sp2 == "bss", TRUE, FALSE),
         pasc = if_else(bb.sp1 == "pasc" | bb.sp2 == "pasc", TRUE, FALSE),
         wurf = if_else(bb.sp1 == "wurf" | bb.sp2 == "wurf", TRUE, FALSE),
         prat = if_else(bb.sp1 == "prat" | bb.sp2 == "prat", TRUE, FALSE),
         soro = if_else(bb.sp1 == "soro" | bb.sp2 == "soro", TRUE, FALSE),
         hort = if_else(bb.sp1 == "hort" | bb.sp2 == "hort", TRUE, FALSE)) %>%

  unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
  mutate(sp.pair = factor(sp.pair)) %>% # ...and make it a factor so that mgcv can handle it
  left_join(BB_covars.site) %>% # add BB covars
  left_join(FL_covars.site) %>% # add floral covars
  mutate(niche.overlap = 1-distance) # express niche overlap so that 1 = perfect overlap and 0 = no overlap

niche_overlap.sp.strongpairs <- niche_overlap.sp %>%
  filter(bb.sp1.abund >= 5 & bb.sp2.abund >= 5)

niche_overlap.sp.mean <- niche_overlap.sp %>%
  group_by(pollen, sp.pair, site, year, elev.mean, pbl.diff, subgenus.diff, 
           bss, pasc, wurf, prat, soro, hort) %>%
  summarize(mean.niche.overlap = mean(niche.overlap)) %>%
  left_join(BB_covars.mean) %>%
  left_join(FL_covars.mean)

niche_overlap.sp.mean2x <- niche_overlap.sp %>%
  group_by(pollen, site, year, date, elev.mean) %>% # Average across species pairs within transects
  summarize(mean.niche.overlap = mean(niche.overlap)) %>% 
  group_by(pollen, site, year, elev.mean) %>% # Average across transects
  summarize(mean.niche.overlap = mean(mean.niche.overlap)) %>% 
  left_join(BB_covars.mean) %>%
  left_join(FL_covars.mean)

### K-type by transect
niche_overlap.ktype <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(!is.na(pollen)) %>%
  left_join(fl_traits) %>%
  group_by(bb.sp, pollen, k.type.s, site, date, year) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
  group_by(bb.sp, pollen, site, date, year) %>%
  mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
  select(-freq) %>% # drop freq variable
  #unite(ID, c(bb.sp, pollen), sep = "_"),
  pivot_wider(names_from = k.type.s, values_from = prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  group_by(pollen, site, date, year) %>%
  nest() %>% # nest into list-column
  mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
         dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
         tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
  select(pollen, site, date, year, tidy_dist) %>% # drop intermediate list-columns
  unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
  left_join(site_data) %>% # add site data
  select(site, date, year, pollen, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # select and rename vars
  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
  rename(subgenus1 = subgenus, pbl.w1 = pbl.w, pbl.w.class1 = pbl.w.class) %>%
  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
  rename(subgenus2 = subgenus, pbl.w2 = pbl.w, pbl.w.class2 = pbl.w.class) %>%
  select(site, date, year, pollen, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
         pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
  mutate(pbl.diff = abs(pbl.w1 - pbl.w2), # tongue length differences
         yday = yday(date), # yday variable will come in handy
         subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different"), # binary: same or different subgenus

         # Logical columns indicating whether a given species belongs to a given species pair
         # This is handy for extracting data for individual species across all their counterparts
         bss = if_else(bb.sp1 == "bss" | bb.sp2 == "bss", TRUE, FALSE),
         pasc = if_else(bb.sp1 == "pasc" | bb.sp2 == "pasc", TRUE, FALSE),
         wurf = if_else(bb.sp1 == "wurf" | bb.sp2 == "wurf", TRUE, FALSE),
         prat = if_else(bb.sp1 == "prat" | bb.sp2 == "prat", TRUE, FALSE),
         soro = if_else(bb.sp1 == "soro" | bb.sp2 == "soro", TRUE, FALSE),
         hort = if_else(bb.sp1 == "hort" | bb.sp2 == "hort", TRUE, FALSE)) %>%

  unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
  mutate(sp.pair = factor(sp.pair)) %>% # ...and make it a factor so that mgcv can handle it
  left_join(BB_covars) %>% # add BB covars
  left_join(FL_covars) %>% # add floral covars
  mutate(niche.overlap = 1-distance) # express niche overlap so that 1 = perfect overlap and 0 = no overlap

niche_overlap.ktype.mean <- niche_overlap.ktype %>%
  group_by(pollen, sp.pair, site, year, elev.mean, pbl.diff, subgenus.diff, 
           bss, pasc, wurf, prat, soro, hort) %>%
  summarize(mean.niche.overlap = mean(niche.overlap)) %>%
  left_join(BB_covars.mean) %>%
  left_join(FL_covars.mean)

niche_overlap.ktype.mean2x <- niche_overlap.ktype %>%
  group_by(pollen, site, year, date, elev.mean) %>% # Average across species pairs within transects
  summarize(mean.niche.overlap = mean(niche.overlap)) %>% 
  group_by(pollen, site, year, elev.mean) %>% # Average across transects
  summarize(mean.niche.overlap = mean(mean.niche.overlap)) %>% 
  left_join(BB_covars.mean) %>%
  left_join(FL_covars.mean)
```

### This story starts with three observations arising from my first set of analyses.

  1. The bumble bee species in the Berchtesgaden system -- even the abundant and widespread ones -- exhibit substantial non-overlap in resource use (Figure 1).
  2. The bumble bee community exhibits little turnover in space/elevation and time.
  3. The floral community exhibits dramatic turnover in space/elevation and time.

### We will focus on six species that are abundant and widespread:  
  1. *B. sensu-strictu*
  2. *B. pascuorum*
  3. *B. pratorum*
  4. *B. hortorum*
  5. *B. soroensis*
  6. *B. wurflenii*

```{r fig.cap="**Figure 1: Focal species**", message=FALSE, warning=FALSE, echo=FALSE}
bb_range.year <- net %>%
  left_join(site_data) %>%
  group_by(site, date, elev.mean, bb.sp, year) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, bb.sp, year) %>%
  summarize(abund = mean(abund)) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2)) %>%
  left_join(species_names)

ggplot(bb_range.year, aes(reorder(bb.sp, elev.med), elev.mean, color = focal)) +
  geom_line(size = 2, alpha = 0.25) +
  geom_point(aes(size = abund), alpha = 0.5) +
  facet_wrap(~year, ncol = 1) +
  theme_light(12) +
  scale_size_continuous(name = "Abundance") +
  scale_color_discrete(name = "Focal species") +
  xlab("BB species") +
  ylab("Elevation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### To improve readability, we will bin our floral taxa to family level. 
```{r fig.cap="**Figure 2: Metaweb diagram**", message=FALSE, warning=FALSE, echo=FALSE}
# Species level
## All
metaweb.sp <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  left_join(fl_traits) %>%
  group_by(bb.sp, plant.sp) %>%
  summarize(freq = n()) %>%
  pivot_wider(names_from = bb.sp, values_from = freq) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("plant.sp")

## Nectar
metaweb.sp.n <- net %>%
  filter(pollen == "nopollen") %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  left_join(fl_traits) %>%
  group_by(bb.sp, plant.sp) %>%
  summarize(freq = n()) %>%
  pivot_wider(names_from = bb.sp, values_from = freq) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("plant.sp")

## Pollen
metaweb.sp.p <- net %>%
  filter(pollen == "pollen") %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  left_join(fl_traits) %>%
  group_by(bb.sp, plant.sp) %>%
  summarize(freq = n()) %>%
  pivot_wider(names_from = bb.sp, values_from = freq) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("plant.sp")

# Family level
## All
metaweb.fam <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  separate(plant.sp, c("plant.genus", "species"), sep = " ") %>%
  left_join(fl_traits) %>%
  group_by(bb.sp, plant.family) %>%
  summarize(freq = n()) %>%
  pivot_wider(names_from = bb.sp, values_from = freq) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("plant.family")

## Nectar
metaweb.fam.n <- net %>%
  filter(pollen == "nopollen") %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  separate(plant.sp, c("plant.genus", "species"), sep = " ") %>%
  left_join(fl_traits) %>%
  group_by(bb.sp, plant.family) %>%
  summarize(freq = n()) %>%
  pivot_wider(names_from = bb.sp, values_from = freq) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("plant.family")

## Pollen
metaweb.fam.p <- net %>%
  filter(pollen == "pollen") %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  separate(plant.sp, c("plant.genus", "species"), sep = " ") %>%
  left_join(fl_traits) %>%
  group_by(bb.sp, plant.family) %>%
  summarize(freq = n()) %>%
  pivot_wider(names_from = bb.sp, values_from = freq) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("plant.family")

# K-type
## All
metaweb.ktype <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  separate(plant.sp, c("plant.genus", "species"), sep = " ") %>%
  left_join(fl_traits) %>%
  group_by(bb.sp, k.type.s) %>%
  summarize(freq = n()) %>%
  pivot_wider(names_from = bb.sp, values_from = freq) %>%
  replace_na(replace = list(bss = 0, hort = 0, pasc = 0, prat = 0, soro = 0, wurf = 0)) %>%
  replace_na(replace = list(k.type.s = "undet")) %>%
  column_to_rownames("k.type.s")

## Nectar
metaweb.ktype.n <- net %>%
  filter(pollen == "nopollen") %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  separate(plant.sp, c("plant.genus", "species"), sep = " ") %>%
  left_join(fl_traits) %>%
  group_by(bb.sp, k.type.s) %>%
  summarize(freq = n()) %>%
  pivot_wider(names_from = bb.sp, values_from = freq) %>%
  replace_na(replace = list(bss = 0, hort = 0, pasc = 0, prat = 0, soro = 0, wurf = 0)) %>%
  replace_na(replace = list(k.type.s = "undet")) %>%
  column_to_rownames("k.type.s")

## Pollen
metaweb.ktype.p <- net %>%
  filter(pollen == "pollen") %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  separate(plant.sp, c("plant.genus", "species"), sep = " ") %>%
  left_join(fl_traits) %>%
  group_by(bb.sp, k.type.s) %>%
  summarize(freq = n()) %>%
  pivot_wider(names_from = bb.sp, values_from = freq) %>%
  replace_na(replace = list(bss = 0, hort = 0, pasc = 0, prat = 0, soro = 0, wurf = 0)) %>%
  replace_na(replace = list(k.type.s = "undet")) %>%
  column_to_rownames("k.type.s")

#visweb(metaweb.sp, type = "diagonal", labsize = 2, pred.lablength = 5, prey.lablength = 5)
#visweb(metaweb.gen, type = "diagonal", labsize = 2, pred.lablength = 5, prey.lablength = 5)
visweb(metaweb.fam, type = "diagonal", labsize = 2, pred.lablength = 5, prey.lablength = 5)
visweb(metaweb.fam.n, type = "diagonal", labsize = 2, pred.lablength = 5, prey.lablength = 5)
visweb(metaweb.fam.p, type = "diagonal", labsize = 2, pred.lablength = 5, prey.lablength = 5)
# visweb(metaweb.ktype.s, type = "diagonal", labsize = 1, pred.lablength = 5, prey.lablength = 5)
# visweb(metaweb.ktype.ss, type = "diagonal", labsize = 1, pred.lablength = 5, prey.lablength = 5)
# 
# plotweb(metaweb.sp, text.rot = 90, labsize = 1, low.lablength = 5)
# plotweb(metaweb.gen, text.rot = 90, labsize = 1, low.lablength = 5)
plotweb(metaweb.fam, text.rot = 90, labsize = 1, low.lablength = 5)
plotweb(metaweb.fam.n, text.rot = 90, labsize = 1, low.lablength = 5)
plotweb(metaweb.fam.p, text.rot = 90, labsize = 1, low.lablength = 5)
# plotweb(metaweb.ktype.s, text.rot = 90, labsize = 1, low.lablength = 5)
# plotweb(metaweb.ktype.ss, text.rot = 90, labsize = 1, low.lablength = 5)
```

### H0: Does niche overlap differ from that predicted by random foraging?
#### Random null model
```{r message=FALSE, warning=FALSE}
### Define function that will be used inside map() calls
df_to_cmat <- function(x) {
  pivot_wider(x, names_from = plant.sp, values_from = freq) %>%
    replace(is.na(.), 0)
}

permat_set <- function(x){
  x %>%
    column_to_rownames("bb.sp") %>%
    vegan::permatfull(fixedmar = "rows", shuffle = "ind", times = 99)
}

ret <- function(x){
  return(x$perm)
}

tidy_dist <- function(x) {
  x %>%
    tidy() %>%
    mutate(index = seq_along())
}

### Generate random niche overlap data
#### Need to look into th the missing value warnings I'm getting from the vegdist() call; I think it's because some plants end up with zero visits in some permutations.
random_niche <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
  filter(!is.na(pollen)) %>% # drop observations with no pollen/nectar classification
  select(site, date, pollen, year, bb.sp, plant.sp) %>% 
  group_by(bb.sp, pollen, plant.sp, site, date) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
  group_by(pollen, site, date) %>%
  nest() %>% # nest data into list-column grouped by pollen, site, date
  mutate(data = map(data, df_to_cmat), # turn table into community matrix
         data = map(data, permat_set), # permute community matrices
         data = map(data, ret)) %>% # extract permutations from permat objects
  unnest(data) %>% # unnest
  mutate(data = map(data, function(x) vegdist(x, method = "horn")), # convert each permutation into distance matrix
         data = map(data, tidy)) %>% # convert distance matrices to tidy arrays
  group_by(pollen, site, date) %>%
  mutate(iter = row_number()) %>% # add index to denote iterations
  unnest(cols = data) %>% # unnest again
  rename(bb.sp1 = item1, bb.sp2 = item2) %>%
  unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
  mutate(sp.pair = factor(sp.pair), # ...and make it a factor so that mgcv can handle it
         niche.overlap = 1-distance) %>% # express niche overlap so that 1 = perfect overlap and 0 = no overlap
  select(pollen, site, date, sp.pair, iter, niche.overlap)

```

#### Niche overlap by species pair
```{r}
ggplot(niche_overlap.sp, aes(sp.pair, niche.overlap)) +
  geom_boxplot(fill = "gray60") +
  facet_wrap(~pollen, ncol = 1) +
  theme_light()

ggplot(niche_overlap.sp.strongpairs, aes(sp.pair, niche.overlap)) +
  geom_boxplot(fill = "gray60") +
  facet_wrap(~pollen, ncol = 1) +
  theme_light()

ggplot(niche_overlap.sp_site, aes(sp.pair, niche.overlap)) +
  geom_boxplot(fill = "gray60") +
  facet_wrap(~pollen, ncol = 1) +
  theme_light()
```

#### Niche overlap: pollen vs. nectar
```{r}
ggplot(niche_overlap.sp, aes(pollen, niche.overlap)) +
  geom_violin(fill = "gray60") +
  geom_boxplot(fill = "NA") +
  theme_light()

ggplot(niche_overlap.sp.strongpairs, aes(pollen, niche.overlap)) +
  geom_violin(fill = "gray60") +
  geom_boxplot(fill = "NA") +
  theme_light()

ggplot(niche_overlap.sp_site, aes(pollen, niche.overlap)) +
  geom_violin(fill = "gray60") +
  geom_boxplot(fill = "NA") +
  theme_light()

lm_p.v.n <- lm(niche.overlap ~ pollen, data = niche_overlap.sp) 
summary(lm_p.v.n)
plot(lm_p.v.n)
```

### Question 1: Why do these bumble bee species partition the floral community as they do?
#### **H0: Random null model.** Shuffle network and see if observed niche partitioning differs significantly from random niche partitioning.
#### **H1: Tongue-length / corolla depth matching.** This is the classic explanation for resource partitioning in BB communities. 
#### **H2: Competition.** It is also possble that competition could --- perhaps transiently --- drive BB species to partition floral resources according to some criterion other than tongue length matching. For example, several BB species may have very similar tongue lengths and yet show marked resource partitioning. 
#### Others? Phylogenetic patterns could represent unmeasured floral trait variation, such as nutritional parameters / secondary chemistry / visual or olfactory cues and their corresponding traits in the BBs (physiology, perception). Differences in spatial foraging strategies? Differences in specialization, e.g. maybe some BB species can manage fewer floral species at a time.

### We can do a quick test of H1: I see no evidence that tongue length differences can explain dietary differences
```{r, echo = FALSE}
ggplot(niche_overlap.sp, aes(pbl.diff, niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth() +
  labs(title = "transect-wise")

ggplot(filter(niche_overlap.sp.strongpairs, wurf == FALSE), aes(pbl.diff, niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "lm") +
  labs(title = "transect-wise")

ggplot(filter(niche_overlap.sp_site, wurf == FALSE), aes(pbl.diff, niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "lm") +
  labs(title = "transect-wise")

ggplot(niche_overlap.sp, aes(pbl.diff, niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth() +
  facet_wrap(~cut_interval(yday(date), 5)) +
  labs(title = "transect-wise by date interval")

ggplot(niche_overlap.sp.mean, aes(pbl.diff, mean.niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth() +
  labs(title = "site-wise")

ggplot(niche_overlap.sp.mean, aes(pbl.diff, mean.niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth() +
  facet_wrap(~cut_interval(elev.mean, 5)) +
  labs(title = "site-wise by elevation interval")
```

```{r}
ggplot(niche_overlap.sp, aes(sqrt(abund.diff), niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth() +
  labs(title = "transect-wise")
```
### Niche overlap by elevation/time
```{r, echo=FALSE}
# Elevation
ggplot(niche_overlap.sp, aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_grid(year~sp.pair) +
  labs(title = "transect-wise pair-wise niche overlap by elevation") +
  ylim(c(0,1))

ggplot(niche_overlap.sp.strongpairs, aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(year~sp.pair) +
  labs(title = "transect-wise pair-wise niche overlap by elevation") +
  ylim(c(0,1))

ggplot(niche_overlap.sp_site, aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  facet_grid(year~sp.pair) +
  labs(title = "transect-wise pair-wise niche overlap by elevation") +
  ylim(c(0,1))

ggplot(niche_overlap.sp.mean, aes(elev.mean, mean.niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_grid(year~sp.pair) +
  labs(title = "site-wise pair-wise niche overlap by elevation") +
  ylim(c(0,1))

# Time
ggplot(niche_overlap.sp, aes(yday, niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_grid(year~sp.pair) +
  labs(title = "transect-wise pair-wise niche overlap by elevation") +
  ylim(c(0,1))
```

### Cluster and ordination analysis
```{r}
# Combined
net.combined <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  group_by(bb.sp, plant.sp, site, date) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, site, date) %>%
  mutate(prop = freq/sum(freq))%>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, plant.sp) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = plant.sp, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.combined_dist <- vegdist(net.combined, method = "horn")
net.combined_clust <- hclust(net.combined_dist, "ward.D")
plot(net.combined_clust)
groups.combined <- cutree(net.combined_clust, 3)

pcoa.combined_dist <- capscale(net.combined ~ 1, distance = "horn", scale = TRUE)
ordiplot(pcoa.combined_dist, type = "t")

ordiplot(pcoa.combined_dist, type = 'n')
points(pcoa.combined_dist, pch = groups, col = groups)

# Pollen
net.pollen <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(pollen == "pollen") %>%
   group_by(bb.sp, plant.sp, site, date) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, site, date) %>%
  mutate(prop = freq/sum(freq))%>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, plant.sp) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = plant.sp, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.pollen.sitepool <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(pollen == "pollen") %>%
   group_by(bb.sp, plant.sp, site) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, site) %>%
  mutate(prop = freq/sum(freq))%>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, plant.sp) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = plant.sp, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.pollen_dist <- vegdist(net.pollen, method = "horn")
net.pollen_clust <- hclust(net.pollen_dist, "ward.D")
plot(net.pollen_clust)
groups.pollen <- cutree(net.pollen_clust, 3)

net.pollen_dist.sitepool <- vegdist(net.pollen.sitepool, method = "horn")
net.pollen_clust.sitepool <- hclust(net.pollen_dist.sitepool, "ward.D")
plot(net.pollen_clust.sitepool)
groups.pollen.sitepool <- cutree(net.pollen_clust.sitepool, 3)

pcoa.pollen_dist.sitepool <- capscale(net.pollen.sitepool ~ 1, distance = "horn", scale = TRUE)
ordiplot(pcoa.pollen_dist.sitepool, type = "t")

ordiplot(pcoa.pollen_dist.sitepool, type = 'n')
points(pcoa.pollen_dist.sitepool, pch = groups.pollen.sitepool, col = groups.pollen.sitepool)

# Nectar
net.nectar <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(pollen == "nopollen") %>%
   group_by(bb.sp, plant.sp, site, date) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, site, date) %>%
  mutate(prop = freq/sum(freq))%>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, plant.sp) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = plant.sp, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.nectar.sitepool <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(pollen == "nopollen") %>%
   group_by(bb.sp, plant.sp, site) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, site) %>%
  mutate(prop = freq/sum(freq))%>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, plant.sp) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = plant.sp, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.nectar_dist <- vegdist(net.nectar, method = "horn")
net.nectar_clust <- hclust(net.nectar_dist, "ward.D")
plot(net.nectar_clust)
groups.nectar <- cutree(net.nectar_clust, 3)

pcoa.nectar_dist <- capscale(net.nectar ~ 1, distance = "horn")
ordiplot(pcoa.nectar_dist, type = "t")

ordiplot(pcoa.nectar_dist, type = 'n')
points(pcoa.nectar_dist, pch = groups, col = groups)

net.nectar_dist.sitepool <- vegdist(net.nectar.sitepool, method = "horn")
net.nectar_clust.sitepool <- hclust(net.nectar_dist.sitepool, "ward.D")
plot(net.nectar_clust.sitepool)
groups.nectar.sitepool <- cutree(net.nectar_clust.sitepool, 2)

pcoa.nectar_dist.sitepool <- capscale(net.nectar.sitepool ~ 1, distance = "horn")
ordiplot(pcoa.nectar_dist.sitepool, type = "t")

ordiplot(pcoa.nectar_dist.sitepool, type = 'n')
points(pcoa.nectar_dist.sitepool, pch = groups.nectar.sitepool, col = groups.nectar.sitepool)

net.nectar.allpool <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(pollen == "nopollen") %>%
   group_by(bb.sp, plant.sp) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp) %>%
  mutate(prop = freq/sum(freq))%>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, plant.sp) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = plant.sp, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.nectar_dist.allpool <- vegdist(net.nectar.allpool, method = "horn")
net.nectar_clust.allpool <- hclust(net.nectar_dist.allpool, "ward.D")
plot(net.nectar_clust.allpool)
groups.nectar.allpool <- cutree(net.nectar_clust.allpool, 2)

pcoa.nectar_dist.allpool <- capscale(net.nectar.allpool ~ 1, distance = "horn")
ordiplot(pcoa.nectar_dist.allpool, type = "t")

ordiplot(pcoa.nectar_dist.allpool, type = 'n')
points(pcoa.nectar_dist.allpool, pch = groups.nectar.allpool, col = groups.nectar.allpool)

net.nectar.highpool <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(elev.mean >= 1500) %>%
  filter(pollen == "nopollen") %>%
   group_by(bb.sp, plant.sp) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp) %>%
  mutate(prop = freq/sum(freq))%>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, plant.sp) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = plant.sp, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.nectar_dist.highpool <- vegdist(net.nectar.highpool, method = "horn")
net.nectar_clust.highpool <- hclust(net.nectar_dist.highpool, "ward.D")
plot(net.nectar_clust.highpool)
groups.nectar.highpool <- cutree(net.nectar_clust.highpool, 2)

pcoa.nectar_dist.highpool <- capscale(net.nectar.highpool ~ 1, distance = "horn")
ordiplot(pcoa.nectar_dist.highpool, type = "t")

ordiplot(pcoa.nectar_dist.highpool, type = 'n')
points(pcoa.nectar_dist.highpool, pch = groups.nectar.highpool, col = groups.nectar.highpool)

net.nectar.lowpool <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(elev.mean <= 1000) %>%
  filter(pollen == "nopollen") %>%
   group_by(bb.sp, plant.sp) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp) %>%
  mutate(prop = freq/sum(freq))%>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, plant.sp) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = plant.sp, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.nectar_dist.lowpool <- vegdist(net.nectar.lowpool, method = "horn")
net.nectar_clust.lowpool <- hclust(net.nectar_dist.lowpool, "ward.D")
plot(net.nectar_clust.lowpool)
groups.nectar.lowpool <- cutree(net.nectar_clust.lowpool, 3)

pcoa.nectar_dist.lowpool <- capscale(net.nectar.lowpool ~ 1, distance = "horn")
ordiplot(pcoa.nectar_dist.lowpool, type = "t")

ordiplot(pcoa.nectar_dist.lowpool, type = 'n')
points(pcoa.nectar_dist.lowpool, pch = groups.nectar.lowpool, col = groups.nectar.lowpool)

net.nectar.midpool <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(elev.mean > 1000 & elev.mean < 1500) %>%
  filter(pollen == "nopollen") %>%
   group_by(bb.sp, plant.sp) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp) %>%
  mutate(prop = freq/sum(freq))%>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, plant.sp) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = plant.sp, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.nectar_dist.midpool <- vegdist(net.nectar.midpool, method = "horn")
net.nectar_clust.midpool <- hclust(net.nectar_dist.midpool, "ward.D")
plot(net.nectar_clust.midpool)
groups.nectar.midpool <- cutree(net.nectar_clust.midpool, 3)

pcoa.nectar_dist.midpool <- capscale(net.nectar.midpool ~ 1, distance = "horn")
ordiplot(pcoa.nectar_dist.midpool, type = "t")

ordiplot(pcoa.nectar_dist.midpool, type = 'n')
points(pcoa.nectar_dist.midpool, pch = groups.nectar.midpool, col = groups.nectar.midpool)

```

```
library(ggordiplots)

global_ordiplot <- gg_ordiplot(pcoa.nectar_dist, groups = groups, 
                               spiders = FALSE, hull = TRUE, 
                               ellipse = FALSE, label = FALSE)

ordiplot(pcoa.nectar_dist, type = "t")
plot(envfit(pcoa.nectar_dist, env = c(6.9, 12.7, 8.4, 7.3, 6.2, 7.5)))

str(pcoa.nectar_dist)

library(ggvegan)
fortify(pcoa.nectar_dist)

autoplot(pcoa.nectar_dist)
```

### GAMS
```{r, echo = FALSE}
pollen.gam.2010 <- bam(niche.overlap ~ 
                         s(sp.pair, bs = "re") +
                         te(yday, elev.mean, by = sp.pair, bs = c("gp", "tp"), k = 5),
                       data = filter(niche_overlap.sp, year == 2010),
                       method = "fREML",
                       discrete= TRUE,
                       family = "scat") %>% getViz()

check.gamViz(pollen.gam.2010)
pollen.gam.2010.sum <- summary(pollen.gam.2010)
print(plot(pollen.gam.2010, allTerms = TRUE), pages = 1)

pollen.gam.2011 <- bam(niche.overlap ~ 
                         s(sp.pair, bs = "re") +
                         te(yday, elev.mean, by = sp.pair, bs = c("gp", "tp"), k = 5),
                       data = filter(niche_overlap.sp, year == 2011),
                       method = "fREML",
                       discrete= TRUE,
                       family = "scat") %>% getViz()

check.gamViz(pollen.gam.2011)
pollen.gam.2011.sum <- summary(pollen.gam.2011)
print(plot(pollen.gam.2011, allTerms = TRUE), pages = 1)

pollen.gam.2012 <- bam(niche.overlap ~ 
                         s(sp.pair, bs = "re") +
                         te(yday, elev.mean, by = sp.pair, bs = c("gp", "tp"), k = 5),
                       data = filter(niche_overlap.sp, year == 2012),
                       method = "fREML",
                       discrete= TRUE,
                       family = "scat") %>% getViz()

check.gamViz(pollen.gam.2012)
pollen.gam.2012.sum <- summary(pollen.gam.2012)
print(plot(pollen.gam.2012, allTerms = TRUE), pages = 1)
```

### 2020-11-26 analyses/old_scripts/niche_partitioning_v2.Rmd
did so many things

---
title: "Flexibility and constraints in resource selection within a generalist pollinator guild"
output: html_document
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(bipartite)
library(mgcv)
library(mgcViz)
library(ggbipart)
library(GGally)
library(vegan)
library(ggvegan)
library(tidyverse)
library(lubridate)
library(patchwork)
library(tidymodels)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
### Load data

# Site data
site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, 
                   elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, 
                             levels = c("oben", "mitte", "unten")),
         elev.class2 = factor(elev.class2, 
                             levels = c("high", "mid", "low"))) # turn elav.class into an ordered factor

# Climate data
gdd <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year)) %>%
  dplyr::select(site, date, gdd.cum)

total_gdd <- gdd %>%
  mutate(year = factor(year(date))) %>%
  group_by(site, year) %>%
  summarize(total.gdd = max(gdd.cum))

# BB traits
bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf"))

# FL traits
fl_tax <- read_csv("../processed_data/floral_tax.csv")

fl_ktype <- read_csv("../processed_data/floral_k_type.csv")

fl_morph <- read_csv("../processed_data/flower_morphology_GitaBenadi.csv") %>%
  select(SampleID, plant.family = Family, plant.sp = Species, 
         Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam),
            LN = mean(LN))

fl_ktype_key <- read_delim("../processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  select(k.type = 1, description = 2)

fl_traits <- fl_tax %>%
  full_join(fl_ktype) %>%
  full_join(fl_morph) %>%
  select(plant.sp, plant.genus, plant.family, 
         k.type, k.type.s, k.type.ss, diam, LN)

# Network
net <- read_csv("../processed_data/network.csv") %>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  mutate(year = factor(year)) %>% # convert year to factor
  left_join(site_data, by = "site") %>%
  left_join(gdd, by = c("site", "date")) %>%
  ungroup() %>%
  arrange(site, date) %>%
  left_join(fl_traits) %>%
  mutate(k.type.s = if_else(plant.genus == "Phyteuma", 7.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Phyteuma", 7, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Aconitum", 5.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Aconitum", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Cardamine", 1.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Cardamine", 1, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Euphrasia", 5.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Euphrasia", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Lamium", 5.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Lamium", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Mentha", 2.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Mentha", 2, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Rumex", 0.0, k.type.s),
         k.type.ss = if_else(plant.genus == "Rumex", 0, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Salix", 9.0, k.type.s),
         k.type.ss = if_else(plant.genus == "Salix", 9, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Silene", 4.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Silene", 4, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Stachys", 5.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Stachys", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Taraxacum", 7.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Taraxacum", 7, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.sp == "Gentiana aspera", 2.1, k.type.s),
         k.type.ss = if_else(plant.sp == "Gentiana aspera", 2, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.sp == "Gentiana ciliata", 4.1, k.type.s),
         k.type.ss = if_else(plant.sp == "Gentiana ciliata", 4, k.type.ss))

ggplot(net, aes(bb.sp, fill = pollen)) +
  geom_bar(position = "fill")

# Floral survey
survey <- read_csv("../processed_data/floral_survey.csv") %>%
  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  left_join(fl_traits) %>%
  mutate(k.type.s = if_else(plant.genus == "Phyteuma", 7.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Phyteuma", 7, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Aconitum", 5.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Aconitum", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Cardamine", 1.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Cardamine", 1, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Euphrasia", 5.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Euphrasia", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Lamium", 5.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Lamium", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Mentha", 2.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Mentha", 2, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Rumex", 0.0, k.type.s),
         k.type.ss = if_else(plant.genus == "Rumex", 0, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Salix", 9.0, k.type.s),
         k.type.ss = if_else(plant.genus == "Salix", 9, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Silene", 4.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Silene", 4, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Stachys", 5.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Stachys", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Taraxacum", 7.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Taraxacum", 7, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.sp == "Gentiana aspera", 2.1, k.type.s),
         k.type.ss = if_else(plant.sp == "Gentiana aspera", 2, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.sp == "Gentiana ciliata", 4.1, k.type.s),
         k.type.ss = if_else(plant.sp == "Gentiana ciliata", 4, k.type.ss))

# Figure out which major species are scored NA for ktype
missing_ktype <- net %>%
  filter(is.na(k.type.s)) %>%
  group_by(plant.sp, plant.genus) %>%
  summarize(freq = n())

### Auxiliary data, i.e. all the covariates we need to control for
BB_covars <- net %>%
  group_by(site, year) %>%
  summarize(bb.rich = length(unique(bb.sp)),
            sqrt.bb.abund = sqrt(n()))

FL_covars <- survey %>%
  mutate(year = factor(year(date))) %>%
  filter(flower.cover > 0) %>%
  group_by(site, year) %>%
  summarize(fl.rich = length(unique(plant.sp)),
            log.fl.abund = log(sum(flower.cover)))

### Species names and such for plotting
species_names <- cbind(c("bss", "pasc", "prat", 
                         "soro", "wurf", "hort", 
                         "pyre", "mend", "mont", 
                         "gers", "hypn","jone", 
                         "lapi", "muci", "psit", 
                         "humi"),
                       c("B. sensu-strictu", "B. pascuorum", "B. pratorum", 
                         "B. soroensis", "B. wurflenii", "B. hortorum",
                         "B. pyrenaeus", "B. mendax", "B. monticola",
                         "B. gerstaeckeri", "B. hypnorum", "B. jonellus", 
                         "B. lapidarius", "B. mucidus", "B. psithyrus", 
                         "B. humilis"),
                       c("B. ss.", "B. pasc.", "B. prat.", 
                         "B. soro.", "B. wurf.", "B. hort.",
                         "B. pyre.", "B. mend.", "B. mont.",
                         "B. gers.", "B. hypn.", "B. jone.", 
                         "B. lapi.", "B. muci.", "B. psit.", 
                         "B. humi"),
                       c("major", "major", "major", 
                         "major", "major", "common", 
                         "high elev.", "high elev.", "high elev.",
                         "minor", "minor", "common", 
                         "minor", "minor", "common", 
                         "minor")
                       ) %>%
  as.data.frame() %>%
  as_tibble() %>%
  dplyr::select(bb.sp = 1, species.name = 2, species.abb = 3, abund.class = 4) %>%
  mutate(focal = if_else(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf"), TRUE, FALSE))
```

```{r message=FALSE, warning=FALSE, include=FALSE}
### Species-level by site
bb_abund.site <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  group_by(bb.sp, site, year) %>%
  summarize(bb.sp.abund = n())

# Pairwise niche overlap by site
niche_overlap.sp <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
  filter(!is.na(pollen)) %>% # drop observations with no pollen/nectar classification
  group_by(bb.sp, pollen, plant.sp, site, year) %>% # group by site-year and BB-FL pair
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-year
  group_by(bb.sp, pollen, site, year) %>% # group by site-year and BB species
  mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
  select(-freq) %>% # drop freq variable
  pivot_wider(names_from = plant.sp, values_from = prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  group_by(pollen, site, year) %>%
  nest() %>% # nest into list-column
  mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
         dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
         tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
  select(pollen, site, year, tidy_dist) %>% # drop intermediate list-columns
  unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
  left_join(site_data) %>% # add site data
  select(site, year, pollen, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # select and rename vars
  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
  rename(subgenus1 = subgenus, pbl.w1 = pbl.w, pbl.w.class1 = pbl.w.class) %>%
  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
  rename(subgenus2 = subgenus, pbl.w2 = pbl.w, pbl.w.class2 = pbl.w.class) %>%
  select(site, year, pollen, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
         pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
  left_join(bb_abund.site, by = c("bb.sp1" = "bb.sp", "site", "year")) %>%
  rename(bb.sp1.abund = bb.sp.abund) %>%
  select(site, year, pollen, bb.sp1, bb.sp2, bb.sp1.abund, subgenus1, subgenus2, distance, 
         pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
  left_join(bb_abund.site, by = c("bb.sp2" = "bb.sp", "site", "year")) %>%
  rename(bb.sp2.abund = bb.sp.abund) %>%
  select(site, year, pollen, bb.sp1, bb.sp2, bb.sp1.abund, bb.sp2.abund, subgenus1, subgenus2, distance, 
         pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
  mutate(pbl.diff = abs(pbl.w1 - pbl.w2), # tongue length differences
         abund.diff = abs(bb.sp1.abund - bb.sp2.abund),
         subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different"), # binary: same or different subgenus

         # Logical columns indicating whether a given species belongs to a given species pair
         # This is handy for extracting data for individual species across all their counterparts
         bss = if_else(bb.sp1 == "bss" | bb.sp2 == "bss", TRUE, FALSE),
         pasc = if_else(bb.sp1 == "pasc" | bb.sp2 == "pasc", TRUE, FALSE),
         wurf = if_else(bb.sp1 == "wurf" | bb.sp2 == "wurf", TRUE, FALSE),
         prat = if_else(bb.sp1 == "prat" | bb.sp2 == "prat", TRUE, FALSE),
         soro = if_else(bb.sp1 == "soro" | bb.sp2 == "soro", TRUE, FALSE),
         hort = if_else(bb.sp1 == "hort" | bb.sp2 == "hort", TRUE, FALSE)) %>%

  unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
  mutate(sp.pair = factor(sp.pair)) %>% # ...and make it a factor so that mgcv can handle it
  left_join(BB_covars.site) %>% # add BB covars
  left_join(FL_covars.site) %>% # add floral covars
  mutate(niche.overlap = 1-distance) # express niche overlap so that 1 = perfect overlap and 0 = no overlap

niche_overlap.ktype <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
  filter(!is.na(pollen)) %>% # drop observations with no pollen/nectar classification
  group_by(bb.sp, pollen, k.type.s, site, year) %>% # group by site-year and BB-FL pair
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-year
  group_by(bb.sp, pollen, site, year) %>% # group by site-year and BB species
  mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
  select(-freq) %>% # drop freq variable
  pivot_wider(names_from = k.type.s, values_from = prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  group_by(pollen, site, year) %>%
  nest() %>% # nest into list-column
  mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
         dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
         tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
  select(pollen, site, year, tidy_dist) %>% # drop intermediate list-columns
  unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
  left_join(site_data) %>% # add site data
  select(site, year, pollen, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # select and rename vars
  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
  rename(subgenus1 = subgenus, pbl.w1 = pbl.w, pbl.w.class1 = pbl.w.class) %>%
  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
  rename(subgenus2 = subgenus, pbl.w2 = pbl.w, pbl.w.class2 = pbl.w.class) %>%
  select(site, year, pollen, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
         pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
  left_join(bb_abund.site, by = c("bb.sp1" = "bb.sp", "site", "year")) %>%
  rename(bb.sp1.abund = bb.sp.abund) %>%
  select(site, year, pollen, bb.sp1, bb.sp2, bb.sp1.abund, subgenus1, subgenus2, distance, 
         pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
  left_join(bb_abund.site, by = c("bb.sp2" = "bb.sp", "site", "year")) %>%
  rename(bb.sp2.abund = bb.sp.abund) %>%
  select(site, year, pollen, bb.sp1, bb.sp2, bb.sp1.abund, bb.sp2.abund, subgenus1, subgenus2, distance, 
         pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
  mutate(pbl.diff = abs(pbl.w1 - pbl.w2), # tongue length differences
         abund.diff = abs(bb.sp1.abund - bb.sp2.abund),
         subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different"), # binary: same or different subgenus

         # Logical columns indicating whether a given species belongs to a given species pair
         # This is handy for extracting data for individual species across all their counterparts
         bss = if_else(bb.sp1 == "bss" | bb.sp2 == "bss", TRUE, FALSE),
         pasc = if_else(bb.sp1 == "pasc" | bb.sp2 == "pasc", TRUE, FALSE),
         wurf = if_else(bb.sp1 == "wurf" | bb.sp2 == "wurf", TRUE, FALSE),
         prat = if_else(bb.sp1 == "prat" | bb.sp2 == "prat", TRUE, FALSE),
         soro = if_else(bb.sp1 == "soro" | bb.sp2 == "soro", TRUE, FALSE),
         hort = if_else(bb.sp1 == "hort" | bb.sp2 == "hort", TRUE, FALSE)) %>%

  unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
  mutate(sp.pair = factor(sp.pair)) %>% # ...and make it a factor so that mgcv can handle it
  left_join(BB_covars.site) %>% # add BB covars
  left_join(FL_covars.site) %>% # add floral covars
  mutate(niche.overlap = 1-distance) # express niche overlap so that 1 = perfect overlap and 0 = no overlap
```

### This story starts with three observations arising from my first set of analyses.

  1. The bumble bee species in the Berchtesgaden system -- even the abundant and widespread ones -- exhibit substantial non-overlap in resource use (Figure 1).
  2. The bumble bee community exhibits little turnover in space/elevation and time.
  3. The floral community exhibits dramatic turnover in space/elevation and time.

### We will focus on six species that are abundant and widespread:  
  1. *B. sensu-strictu*
  2. *B. pascuorum*
  3. *B. pratorum*
  4. *B. hortorum*
  5. *B. soroensis*
  6. *B. wurflenii*

```{r fig.cap="**Figure 1: Focal species**", message=FALSE, warning=FALSE, echo=FALSE}
bb_range.year <- net %>%
  left_join(site_data) %>%
  group_by(site, date, elev.mean, bb.sp, year) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, bb.sp, year) %>%
  summarize(abund = mean(abund)) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2)) %>%
  left_join(species_names)

ggplot(bb_range.year, aes(reorder(bb.sp, elev.med), elev.mean, color = focal)) +
  geom_line(size = 2, alpha = 0.25) +
  geom_point(aes(size = abund), alpha = 0.5) +
  facet_wrap(~year, ncol = 1) +
  theme_light(12) +
  scale_size_continuous(name = "Abundance") +
  scale_color_discrete(name = "Focal species") +
  xlab("BB species") +
  ylab("Elevation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### To improve readability, we will bin our floral taxa to family level. 
```{r fig.cap="**Figure 2: Metaweb diagram**", message=FALSE, warning=FALSE, echo=FALSE}
# Species level
## All
metaweb.sp <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  group_by(bb.sp, plant.sp) %>%
  summarize(freq = n()) %>%
  pivot_wider(names_from = bb.sp, values_from = freq) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("plant.sp")

## Nectar
metaweb.sp.n <- net %>%
  filter(pollen == "nopollen") %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  group_by(bb.sp, plant.sp) %>%
  summarize(freq = n()) %>%
  pivot_wider(names_from = bb.sp, values_from = freq) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("plant.sp")

## Pollen
metaweb.sp.p <- net %>%
  filter(pollen == "pollen") %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  group_by(bb.sp, plant.sp) %>%
  summarize(freq = n()) %>%
  pivot_wider(names_from = bb.sp, values_from = freq) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("plant.sp")

# Family level
## All
metaweb.fam <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  separate(plant.sp, c("plant.genus", "species"), sep = " ") %>%
  group_by(bb.sp, plant.family) %>%
  summarize(freq = n()) %>%
  pivot_wider(names_from = bb.sp, values_from = freq) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("plant.family")

## Nectar
metaweb.fam.n <- net %>%
  filter(pollen == "nopollen") %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  separate(plant.sp, c("plant.genus", "species"), sep = " ") %>%
  group_by(bb.sp, plant.family) %>%
  summarize(freq = n()) %>%
  pivot_wider(names_from = bb.sp, values_from = freq) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("plant.family")

## Pollen
metaweb.fam.p <- net %>%
  filter(pollen == "pollen") %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  separate(plant.sp, c("plant.genus", "species"), sep = " ") %>%
  group_by(bb.sp, plant.family) %>%
  summarize(freq = n()) %>%
  pivot_wider(names_from = bb.sp, values_from = freq) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("plant.family")

# K-type
## All
metaweb.ktype <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  separate(plant.sp, c("plant.genus", "species"), sep = " ") %>%
  group_by(bb.sp, k.type.s) %>%
  summarize(freq = n()) %>%
  pivot_wider(names_from = bb.sp, values_from = freq) %>%
  replace_na(replace = list(bss = 0, hort = 0, pasc = 0, prat = 0, soro = 0, wurf = 0)) %>%
  replace_na(replace = list(k.type.s = "undet")) %>%
  column_to_rownames("k.type.s")

## Nectar
metaweb.ktype.n <- net %>%
  filter(pollen == "nopollen") %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  separate(plant.sp, c("plant.genus", "species"), sep = " ") %>%
  group_by(bb.sp, k.type.s) %>%
  summarize(freq = n()) %>%
  pivot_wider(names_from = bb.sp, values_from = freq) %>%
  replace_na(replace = list(bss = 0, hort = 0, pasc = 0, prat = 0, soro = 0, wurf = 0)) %>%
  replace_na(replace = list(k.type.s = "undet")) %>%
  column_to_rownames("k.type.s")

## Pollen
metaweb.ktype.p <- net %>%
  filter(pollen == "pollen") %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  separate(plant.sp, c("plant.genus", "species"), sep = " ") %>%
  group_by(bb.sp, k.type.s) %>%
  summarize(freq = n()) %>%
  pivot_wider(names_from = bb.sp, values_from = freq) %>%
  replace_na(replace = list(bss = 0, hort = 0, pasc = 0, prat = 0, soro = 0, wurf = 0)) %>%
  replace_na(replace = list(k.type.s = "undet")) %>%
  column_to_rownames("k.type.s")

#visweb(metaweb.sp, type = "diagonal", labsize = 2, pred.lablength = 5, prey.lablength = 5)
#visweb(metaweb.gen, type = "diagonal", labsize = 2, pred.lablength = 5, prey.lablength = 5)
visweb(metaweb.fam, type = "diagonal", labsize = 2, pred.lablength = 5, prey.lablength = 5)
visweb(metaweb.fam.n, type = "diagonal", labsize = 2, pred.lablength = 5, prey.lablength = 5)
visweb(metaweb.fam.p, type = "diagonal", labsize = 2, pred.lablength = 5, prey.lablength = 5)
# visweb(metaweb.ktype.n, type = "diagonal", labsize = 1, pred.lablength = 5, prey.lablength = 5)
# visweb(metaweb.ktype.p, type = "diagonal", labsize = 1, pred.lablength = 5, prey.lablength = 5)
# 
# plotweb(metaweb.sp, text.rot = 90, labsize = 1, low.lablength = 5)
# plotweb(metaweb.gen, text.rot = 90, labsize = 1, low.lablength = 5)
plotweb(metaweb.fam, text.rot = 90, labsize = 1, low.lablength = 5)
plotweb(metaweb.fam.n, text.rot = 90, labsize = 1, low.lablength = 5)
plotweb(metaweb.fam.p, text.rot = 90, labsize = 1, low.lablength = 5)
# plotweb(metaweb.ktype.n, text.rot = 90, labsize = 1, low.lablength = 5)
# plotweb(metaweb.ktype.p, text.rot = 90, labsize = 1, low.lablength = 5)
```

### H0: Does niche overlap differ from that predicted by random foraging?
#### Random null model
```{r message=FALSE, warning=FALSE}
### Define function that will be used inside map() calls
df_to_cmat <- function(x) {
  pivot_wider(x, names_from = plant.sp, values_from = freq) %>%
    replace(is.na(.), 0)
}

permat_set <- function(x){
  x %>%
    column_to_rownames("bb.sp") %>%
    vegan::permatfull(fixedmar = "rows", shuffle = "ind", times = 10)
}

ret <- function(x){
  return(x$perm)
}

tidy_dist <- function(x) {
  x %>%
    tidy() %>%
    mutate(index = seq_along())
}

### Generate random niche overlap data
#### Need to look into th the missing value warnings I'm getting from the vegdist() call; I think it's because some plants end up with zero visits in some permutations.
random_niche <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
  filter(!is.na(pollen)) %>% # drop observations with no pollen/nectar classification
  select(site, pollen, year, bb.sp, plant.sp) %>% 
  group_by(site, pollen, year, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
  group_by(site, pollen, year) %>%
  nest() %>% # nest data into list-column grouped by pollen, site, date
  mutate(data = map(data, df_to_cmat), # turn table into community matrix
         data = map(data, permat_set), # permute community matrices
         data = map(data, ret)) %>% # extract permutations from permat objects
  unnest(data) %>% # unnest
  mutate(data = map(data, function(x) vegdist(x, method = "horn")), # convert each permutation into distance matrix
         data = map(data, tidy)) %>% # convert distance matrices to tidy arrays
  group_by(site, pollen, year) %>%
  mutate(iter = row_number()) %>% # add index to denote iterations
  unnest(cols = data) %>% # unnest again
  rename(bb.sp1 = item1, bb.sp2 = item2) %>%
  unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
  mutate(sp.pair = factor(sp.pair), # ...and make it a factor so that mgcv can handle it
         niche.overlap = 1-distance) %>% # express niche overlap so that 1 = perfect overlap and 0 = no overlap
  select(site, pollen, year, sp.pair, iter, niche.overlap)
```

#### Niche overlap by species pair
```{r}
# By plant.sp
ggplot(niche_overlap.sp, aes(sp.pair, niche.overlap)) +
  geom_boxplot(fill = "gray60") +
  facet_grid(year~pollen) +
  theme_light()

# By plant ktype
ggplot(niche_overlap.ktype, aes(sp.pair, niche.overlap)) +
  geom_boxplot(fill = "gray60") +
  facet_grid(year~pollen) +
  theme_light()
```

#### Niche overlap: pollen vs. nectar
```{r}
# By plant.sp
ggplot(niche_overlap.sp, aes(pollen, niche.overlap)) +
  geom_violin(fill = "gray60") +
  geom_boxplot(fill = "NA") +
  facet_wrap(~year) +
  theme_light()

ggplot(niche_overlap.sp, aes(pollen, niche.overlap)) +
  geom_violin(fill = "gray60") +
  geom_boxplot(fill = "NA") +
  facet_grid(year~sp.pair) +
  theme_light()

# By plant ktype
ggplot(niche_overlap.ktype, aes(pollen, niche.overlap)) +
  geom_violin(fill = "gray60") +
  geom_boxplot(fill = "NA") +
  facet_wrap(~year) +
  theme_light()

ggplot(niche_overlap.ktype, aes(pollen, niche.overlap)) +
  geom_violin(fill = "gray60") +
  geom_boxplot(fill = "NA") +
  facet_grid(year~sp.pair) +
  theme_light()
```

### Question 1: Why do these bumble bee species partition the floral community as they do?
#### **H0: Random null model.** Shuffle network and see if observed niche partitioning differs significantly from random niche partitioning.
#### **H1: Tongue-length / corolla depth matching.** This is the classic explanation for resource partitioning in BB communities. 
#### **H2: Competition.** It is also possble that competition could --- perhaps transiently --- drive BB species to partition floral resources according to some criterion other than tongue length matching. For example, several BB species may have very similar tongue lengths and yet show marked resource partitioning. 
#### Others? Phylogenetic patterns could represent unmeasured floral trait variation, such as nutritional parameters / secondary chemistry / visual or olfactory cues and their corresponding traits in the BBs (physiology, perception). Differences in spatial foraging strategies? Differences in specialization, e.g. maybe some BB species can manage fewer floral species at a time.

### We can do a quick test of H1
```{r, echo = FALSE}
# By plant.sp
## All focal species
ggplot(niche_overlap.sp, aes(pbl.diff, niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth() +
  facet_wrap(~year) +
  labs(title = "transect-wise")

ggplot(niche_overlap.sp, aes(pbl.diff, niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "lm") +
  facet_wrap(~year) +
  labs(title = "transect-wise")

## Drop B. wurflenii since  it is a robber
ggplot(filter(niche_overlap.sp, wurf == FALSE), aes(pbl.diff, niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth() +
  facet_wrap(~year) +
  labs(title = "transect-wise")

ggplot(filter(niche_overlap.sp, wurf == FALSE), aes(pbl.diff, niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method  = "lm") +
  facet_wrap(~year) +
  labs(title = "transect-wise")

# By plant ktype
## All focal species
ggplot(niche_overlap.ktype, aes(pbl.diff, niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth() +
  facet_wrap(~year) +
  labs(title = "transect-wise")

ggplot(niche_overlap.ktype, aes(pbl.diff, niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "lm") +
  facet_wrap(~year) +
  labs(title = "transect-wise")

## Drop B. wurflenii since  it is a robber
ggplot(filter(niche_overlap.ktype, wurf == FALSE), aes(pbl.diff, niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth() +
  facet_wrap(~year) +
  labs(title = "transect-wise")

ggplot(filter(niche_overlap.ktype, wurf == FALSE), aes(pbl.diff, niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method  = "lm") +
  facet_wrap(~year) +
  labs(title = "transect-wise")
```

### Niche overlap by elevation
```{r, echo=FALSE}
# By plant.sp
## All sp pairs
ggplot(niche_overlap.sp, aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year) +
  ylim(c(0,1))

## By sp pair
ggplot(niche_overlap.sp, aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  facet_grid(year~sp.pair) +
  ylim(c(0,1))

# By plant ktype
## All sp pairs
ggplot(niche_overlap.ktype, aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year) +
  ylim(c(0,1))

## By sp pair
ggplot(niche_overlap.ktype, aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  facet_grid(year~sp.pair) +
  ylim(c(0,1))
```

### Cluster and ordination analysis by plant.sp
```{r}
# Combined
net.combined <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  group_by(bb.sp, plant.sp, site) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, site) %>%
  mutate(prop = freq/sum(freq))%>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, plant.sp) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = plant.sp, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.combined_dist <- vegdist(net.combined, method = "horn")
net.combined_clust <- hclust(net.combined_dist, "ward.D")
plot(net.combined_clust)
groups.combined <- cutree(net.combined_clust, 2)

pcoa.combined_dist <- capscale(net.combined ~ 1, distance = "horn", scale = TRUE)
ordiplot(pcoa.combined_dist, type = "t")

ordiplot(pcoa.combined_dist, type = 'n')
points(pcoa.combined_dist, pch = groups.combined, col = groups.combined)

# Pollen
net.pollen <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(pollen == "pollen") %>%
   group_by(bb.sp, plant.sp, site) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, site) %>%
  mutate(prop = freq/sum(freq))%>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, plant.sp) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = plant.sp, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.pollen_dist <- vegdist(net.pollen, method = "horn")
net.pollen_clust <- hclust(net.pollen_dist, "ward.D")
plot(net.pollen_clust)
groups.pollen <- cutree(net.pollen_clust, 3)

pcoa.pollen_dist <- capscale(net.pollen ~ 1, distance = "horn", scale = TRUE)
ordiplot(pcoa.pollen_dist, type = "t")

ordiplot(pcoa.pollen_dist, type = 'n')
points(pcoa.pollen_dist, pch = groups.pollen, col = groups.pollen)

# Nectar
net.nectar <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(pollen == "nopollen") %>%
   group_by(bb.sp, plant.sp, site) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, site) %>%
  mutate(prop = freq/sum(freq))%>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, plant.sp) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = plant.sp, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.nectar_dist <- vegdist(net.nectar, method = "horn")
net.nectar_clust <- hclust(net.nectar_dist, "ward.D")
plot(net.nectar_clust)
groups.nectar <- cutree(net.nectar_clust, 3)

pcoa.nectar_dist <- capscale(net.nectar ~ 1, distance = "horn")
ordiplot(pcoa.nectar_dist, type = "t")

ordiplot(pcoa.nectar_dist, type = 'n')
points(pcoa.nectar_dist, pch = groups.nectar, col = groups.nectar)

# Broken down by elevation class
## High
net.nectar.highpool <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(elev.class == "oben") %>%
  filter(pollen == "nopollen") %>%
   group_by(bb.sp, plant.sp) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp) %>%
  mutate(prop = freq/sum(freq))%>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, plant.sp) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = plant.sp, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.nectar_dist.highpool <- vegdist(net.nectar.highpool, method = "horn")
net.nectar_clust.highpool <- hclust(net.nectar_dist.highpool, "ward.D")
plot(net.nectar_clust.highpool)
groups.nectar.highpool <- cutree(net.nectar_clust.highpool, 2)

pcoa.nectar_dist.highpool <- capscale(net.nectar.highpool ~ 1, distance = "horn")
ordiplot(pcoa.nectar_dist.highpool, type = "t")

ordiplot(pcoa.nectar_dist.highpool, type = 'n')
points(pcoa.nectar_dist.highpool, pch = groups.nectar.highpool, col = groups.nectar.highpool)

net.pollen.highpool <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(elev.class == "oben") %>%
  filter(pollen == "pollen") %>%
   group_by(bb.sp, plant.sp) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp) %>%
  mutate(prop = freq/sum(freq))%>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, plant.sp) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = plant.sp, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.pollen_dist.highpool <- vegdist(net.pollen.highpool, method = "horn")
net.pollen_clust.highpool <- hclust(net.pollen_dist.highpool, "ward.D")
plot(net.pollen_clust.highpool)
groups.pollen.highpool <- cutree(net.pollen_clust.highpool, 3)

pcoa.pollen_dist.highpool <- capscale(net.pollen.highpool ~ 1, distance = "horn")
ordiplot(pcoa.pollen_dist.highpool, type = "t")

ordiplot(pcoa.pollen_dist.highpool, type = 'n')
points(pcoa.pollen_dist.highpool, pch = groups.pollen.highpool, col = groups.pollen.highpool)

## Low
net.nectar.lowpool <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(elev.class == "unten") %>%
  filter(pollen == "nopollen") %>%
   group_by(bb.sp, plant.sp) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp) %>%
  mutate(prop = freq/sum(freq))%>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, plant.sp) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = plant.sp, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.nectar_dist.lowpool <- vegdist(net.nectar.lowpool, method = "horn")
net.nectar_clust.lowpool <- hclust(net.nectar_dist.lowpool, "ward.D")
plot(net.nectar_clust.lowpool)
groups.nectar.lowpool <- cutree(net.nectar_clust.lowpool, 3)

pcoa.nectar_dist.lowpool <- capscale(net.nectar.lowpool ~ 1, distance = "horn")
ordiplot(pcoa.nectar_dist.lowpool, type = "t")

ordiplot(pcoa.nectar_dist.lowpool, type = 'n')
points(pcoa.nectar_dist.lowpool, pch = groups.nectar.lowpool, col = groups.nectar.lowpool)

net.pollen.lowpool <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(elev.class == "unten") %>%
  filter(pollen == "pollen") %>%
   group_by(bb.sp, plant.sp) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp) %>%
  mutate(prop = freq/sum(freq))%>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, plant.sp) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = plant.sp, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.pollen_dist.lowpool <- vegdist(net.pollen.lowpool, method = "horn")
net.pollen_clust.lowpool <- hclust(net.pollen_dist.lowpool, "ward.D")
plot(net.pollen_clust.lowpool)
groups.pollen.lowpool <- cutree(net.pollen_clust.lowpool, 3)

pcoa.pollen_dist.lowpool <- capscale(net.pollen.lowpool ~ 1, distance = "horn")
ordiplot(pcoa.pollen_dist.lowpool, type = "t")

ordiplot(pcoa.pollen_dist.lowpool, type = 'n')
points(pcoa.pollen_dist.lowpool, pch = groups.pollen.lowpool, col = groups.pollen.lowpool)

## Middle
net.nectar.midpool <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(elev.class == "mitte") %>%
  filter(pollen == "nopollen") %>%
   group_by(bb.sp, plant.sp) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp) %>%
  mutate(prop = freq/sum(freq))%>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, plant.sp) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = plant.sp, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.nectar_dist.midpool <- vegdist(net.nectar.midpool, method = "horn")
net.nectar_clust.midpool <- hclust(net.nectar_dist.midpool, "ward.D")
plot(net.nectar_clust.midpool)
groups.nectar.midpool <- cutree(net.nectar_clust.midpool, 4)

pcoa.nectar_dist.midpool <- capscale(net.nectar.midpool ~ 1, distance = "horn")
ordiplot(pcoa.nectar_dist.midpool, type = "t")

ordiplot(pcoa.nectar_dist.midpool, type = 'n')
points(pcoa.nectar_dist.midpool, pch = groups.nectar.midpool, col = groups.nectar.midpool)

net.pollen.midpool <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(elev.class == "mitte") %>%
  filter(pollen == "pollen") %>%
   group_by(bb.sp, plant.sp) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp) %>%
  mutate(prop = freq/sum(freq))%>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, plant.sp) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = plant.sp, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.pollen_dist.midpool <- vegdist(net.pollen.midpool, method = "horn")
net.pollen_clust.midpool <- hclust(net.pollen_dist.midpool, "ward.D")
plot(net.pollen_clust.midpool)
groups.pollen.midpool <- cutree(net.pollen_clust.midpool, 4)

pcoa.pollen_dist.midpool <- capscale(net.pollen.midpool ~ 1, distance = "horn")
ordiplot(pcoa.pollen_dist.midpool, type = "t")

ordiplot(pcoa.pollen_dist.midpool, type = 'n')
points(pcoa.pollen_dist.midpool, pch = groups.pollen.midpool, col = groups.pollen.midpool)

```

### Cluster and ordination analysis by k.type.s
```{r}
# Combined
net.combined.ktype <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  group_by(bb.sp, k.type.s, site) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, site) %>%
  mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, k.type.s) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = k.type.s, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.combined_dist.ktype <- vegdist(net.combined.ktype, method = "horn")
net.combined_clust.ktype <- hclust(net.combined_dist.ktype, "ward.D")
plot(net.combined_clust.ktype)
groups.combined.ktype <- cutree(net.combined_clust.ktype, 3)

pcoa.combined_dist.ktype <- capscale(net.combined.ktype ~ 1, distance = "horn", scale = TRUE)
ordiplot(pcoa.combined_dist.ktype, type = "t")

ordiplot(pcoa.combined_dist, type = 'n')
points(pcoa.combined_dist, pch = groups.combined, col = groups.combined)

# Pollen
net.pollen.ktype <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(pollen == "pollen") %>%
  group_by(bb.sp, k.type.s, site) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, site) %>%
  mutate(prop = freq/sum(freq))%>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, k.type.s) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = k.type.s, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.pollen_dist.ktype <- vegdist(net.pollen.ktype, method = "horn")
net.pollen_clust.ktype <- hclust(net.pollen_dist.ktype, "ward.D")
plot(net.pollen_clust.ktype)
groups.pollen.ktype <- cutree(net.pollen_clust.ktype, 3)

pcoa.pollen_dist.ktype <- capscale(net.pollen.ktype ~ 1, distance = "horn", scale = TRUE)
ordiplot(pcoa.pollen_dist.ktype, type = "t")

ordiplot(pcoa.pollen_dist, type = 'n')
points(pcoa.pollen_dist, pch = groups.pollen, col = groups.pollen)

# Nectar
net.nectar.ktype <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(pollen == "nopollen") %>%
  group_by(bb.sp, k.type.s, site) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, site) %>%
  mutate(prop = freq/sum(freq))%>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, k.type.s) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = k.type.s, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.nectar_dist.ktype <- vegdist(net.nectar.ktype, method = "horn")
net.nectar_clust.ktype <- hclust(net.nectar_dist.ktype, "ward.D")
plot(net.nectar_clust.ktype)
groups.nectar.ktype <- cutree(net.nectar_clust.ktype, 3)

pcoa.nectar_dist.ktype <- capscale(net.nectar.ktype ~ 1, distance = "horn", scale = TRUE)
ordiplot(pcoa.nectar_dist.ktype, type = "t")

ordiplot(pcoa.nectar_dist, type = 'n')
points(pcoa.nectar_dist, pch = groups.nectar, col = groups.nectar)

# Broken down by elevation class
## High
net.nectar.highpool.ktype <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(pollen == "nopollen") %>%
  filter(elev.class == "oben") %>%
  group_by(bb.sp, k.type.s, site) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, site) %>%
  mutate(prop = freq/sum(freq))%>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, k.type.s) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = k.type.s, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.nectar_dist.highpool.ktype <- vegdist(net.nectar.highpool.ktype, method = "horn")
net.nectar_clust.highpool.ktype <- hclust(net.nectar_dist.highpool.ktype, "ward.D")
plot(net.nectar_clust.highpool.ktype)
groups.nectar.highpool.ktype <- cutree(net.nectar_clust.highpool.ktype, 3)

pcoa.nectar_dist.highpool.ktype <- capscale(net.nectar.highpool.ktype ~ 1, distance = "horn", scale = TRUE)
ordiplot(pcoa.nectar_dist.highpool.ktype, type = "t")

ordiplot(pcoa.nectar_dist, type = 'n')
points(pcoa.nectar_dist, pch = groups.nectar, col = groups.nectar)

net.pollen.highpool.ktype <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(pollen == "pollen") %>%
  filter(elev.class == "oben") %>%
  group_by(bb.sp, k.type.s, site) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, site) %>%
  mutate(prop = freq/sum(freq))%>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, k.type.s) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = k.type.s, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.pollen_dist.highpool.ktype <- vegdist(net.pollen.highpool.ktype, method = "horn")
net.pollen_clust.highpool.ktype <- hclust(net.pollen_dist.highpool.ktype, "ward.D")
plot(net.pollen_clust.highpool.ktype)
groups.pollen.highpool.ktype <- cutree(net.pollen_clust.highpool.ktype, 2)

pcoa.pollen_dist.highpool.ktype <- capscale(net.pollen.highpool.ktype ~ 1, distance = "horn", scale = TRUE)
ordiplot(pcoa.pollen_dist.highpool.ktype, type = "t")

ordiplot(pcoa.pollen_dist, type = 'n')
points(pcoa.pollen_dist, pch = groups.pollen.highpool.ktype, col = groups.pollen.highpool.ktype)

## Low
net.nectar.lowpool.ktype <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(pollen == "nopollen") %>%
  filter(elev.class == "unten") %>%
  group_by(bb.sp, k.type.s, site) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, site) %>%
  mutate(prop = freq/sum(freq))%>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, k.type.s) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = k.type.s, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.nectar_dist.lowpool.ktype <- vegdist(net.nectar.lowpool.ktype, method = "horn")
net.nectar_clust.lowpool.ktype <- hclust(net.nectar_dist.lowpool.ktype, "ward.D")
plot(net.nectar_clust.lowpool.ktype)
groups.nectar.lowpool.ktype <- cutree(net.nectar_clust.lowpool.ktype, 3)

pcoa.nectar_dist.lowpool.ktype <- capscale(net.nectar.lowpool.ktype ~ 1, distance = "horn", scale = TRUE)
ordiplot(pcoa.nectar_dist.lowpool.ktype, type = "t")

ordiplot(pcoa.nectar_dist, type = 'n')
points(pcoa.nectar_dist, pch = groups.nectar, col = groups.nectar)

net.pollen.lowpool.ktype <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(pollen == "pollen") %>%
  filter(elev.class == "unten") %>%
  group_by(bb.sp, k.type.s, site) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, site) %>%
  mutate(prop = freq/sum(freq))%>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, k.type.s) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = k.type.s, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.pollen_dist.lowpool.ktype <- vegdist(net.pollen.lowpool.ktype, method = "horn")
net.pollen_clust.lowpool.ktype <- hclust(net.pollen_dist.lowpool.ktype, "ward.D")
plot(net.pollen_clust.lowpool.ktype)
groups.pollen.lowpool.ktype <- cutree(net.pollen_clust.lowpool.ktype, 2)

pcoa.pollen_dist.lowpool.ktype <- capscale(net.pollen.lowpool.ktype ~ 1, distance = "horn", scale = TRUE)
ordiplot(pcoa.pollen_dist.lowpool.ktype, type = "t")

ordiplot(pcoa.pollen_dist, type = 'n')
points(pcoa.pollen_dist, pch = groups.pollen.lowpool.ktype, col = groups.pollen.lowpool.ktype)

## Middle
net.nectar.midpool.ktype <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(pollen == "nopollen") %>%
  filter(elev.class == "mitte") %>%
  group_by(bb.sp, k.type.s, site) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, site) %>%
  mutate(prop = freq/sum(freq))%>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, k.type.s) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = k.type.s, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.nectar_dist.midpool.ktype <- vegdist(net.nectar.midpool.ktype, method = "horn")
net.nectar_clust.midpool.ktype <- hclust(net.nectar_dist.midpool.ktype, "ward.D")
plot(net.nectar_clust.midpool.ktype)
groups.nectar.midpool.ktype <- cutree(net.nectar_clust.midpool.ktype, 3)

pcoa.nectar_dist.midpool.ktype <- capscale(net.nectar.midpool.ktype ~ 1, distance = "horn", scale = TRUE)
ordiplot(pcoa.nectar_dist.midpool.ktype, type = "t")

ordiplot(pcoa.nectar_dist, type = 'n')
points(pcoa.nectar_dist, pch = groups.nectar, col = groups.nectar)

net.pollen.midpool.ktype <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(pollen == "pollen") %>%
  filter(elev.class == "mitte") %>%
  group_by(bb.sp, k.type.s, site) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, site) %>%
  mutate(prop = freq/sum(freq))%>% # get proportional interaction frequency for each BB-FL pair per site-date
  group_by(bb.sp, k.type.s) %>%
  summarize(mean.prop = mean(prop)) %>% # get mean proportional interaction frequency for each BB-FL pair across site-dates
  pivot_wider(names_from = k.type.s, values_from = mean.prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  column_to_rownames("bb.sp")

net.pollen_dist.midpool.ktype <- vegdist(net.pollen.midpool.ktype, method = "horn")
net.pollen_clust.midpool.ktype <- hclust(net.pollen_dist.midpool.ktype, "ward.D")
plot(net.pollen_clust.midpool.ktype)
groups.pollen.midpool.ktype <- cutree(net.pollen_clust.midpool.ktype, 2)

pcoa.pollen_dist.midpool.ktype <- capscale(net.pollen.midpool.ktype ~ 1, distance = "horn", scale = TRUE)
ordiplot(pcoa.pollen_dist.midpool.ktype, type = "t")

ordiplot(pcoa.pollen_dist, type = 'n')
points(pcoa.pollen_dist, pch = groups.pollen.midpool.ktype, col = groups.pollen.midpool.ktype)
```

# Taxonomic vs. morphological beta-diversity
```{r}
library(gdm)
library(sf)

# A function for converting a geometry column back to x and y columns
# https://github.com/r-spatial/sf/issues/231
sfc_as_cols <- function(x, names = c("x","y")) {
  stopifnot(inherits(x,"sf") && inherits(sf::st_geometry(x),"sfc_POINT"))
  ret <- sf::st_coordinates(x)
  ret <- tibble::as_tibble(ret)
  stopifnot(length(names) == ncol(ret))
  x <- x[ , !names(x) %in% names]
  ret <- setNames(ret,names)
  dplyr::bind_cols(x,ret)
}

site_data_xy <- site_data %>% # turn elav.class into an ordered factor
  semi_join(read_csv("../processed_data/network.csv"), by = "site") %>% # drop sites that were not sampled
  st_as_sf(coords = c("lon", "lat")) %>%
  st_set_crs(4326) %>%
  st_transform(5684) %>%
  sfc_as_cols() %>%
  as_tibble() %>%
  dplyr::select(-geometry)

# Create functions for prepping data
biodata_prep_ktype <- function(surv, y) {
  surv %>%
    mutate(year = year(date)) %>%
    filter(flower.cover > 0 & year == y) %>%
    group_by(site, k.type.s) %>%
    summarize(cover = sum(flower.cover)) %>%
    #mutate(cover = rep(1, n())) %>% # this makes the data binary, which I think is what I want here; otherwise, just comment out
    spread(k.type.s, cover) %>%
    replace(is.na(.), 0) %>%
    arrange(site)
}

biodata_prep_sp <- function(surv, y) {
  surv %>%
    mutate(year = year(date)) %>%
    filter(flower.cover > 0 & year == y) %>%
    group_by(site, plant.sp) %>%
    summarize(cover = sum(flower.cover)) %>%
    #mutate(cover = rep(1, n())) %>% # this makes the data binary, which I think is what I want here; otherwise, just comment out
    spread(plant.sp, cover) %>%
    replace(is.na(.), 0) %>%
    arrange(site)
}

predData_prep_survey <- function(surv, y) {
  surv %>%
    mutate(year = year(date)) %>%
    filter(flower.cover > 0 & year == y) %>%
    dplyr::select(site) %>%
    left_join(site_data_xy) %>%
    ungroup() %>%
    dplyr::select(site, elev.mean, x, y) %>%
    unique() %>%
    arrange(site) 
}

format_set <- function(bio, pred) {
  formatsitepair(bio, bioFormat = 1, siteColumn = "site", 
                 dist = "jaccard", predData = pred, XColumn = "x", YColumn = "y",
                 abundance = TRUE) %>%
  filter(s1.elev.mean != s2.elev.mean) # this avoids comparing repeated measures of the same site (taking advantage of the fact that each site has a unique elevation in our dataset)
}

bio2012.ktype <- biodata_prep_ktype(survey, 2012)
bio2012.sp <- biodata_prep_sp(survey, 2012)
pred2012 <- predData_prep_survey(survey, 2012)
set2012.ktype <- format_set(bio2012.ktype, pred2012)
set2012.sp <- format_set(bio2012.sp, pred2012)

gdm2012.ktype <- gdm(set2012.ktype, geo = TRUE)
plot(gdm2012.ktype)
summary(gdm2012.ktype)

gdm2012.sp <- gdm(set2012.sp, geo = TRUE)
plot(gdm2012.sp)
```

### Taxonomic vs. morphological diversity
```{r}
diversity <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(!is.na(pollen)) %>%
  group_by(site, year, bb.sp, elev.mean, pollen) %>%
  summarize(taxon.rich = length(unique(plant.sp)),
            morpho.rich = length(unique(k.type.s)))

diversity_long <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  filter(!is.na(pollen)) %>%
  group_by(site, year, bb.sp, elev.mean, pollen) %>%
  summarize(taxon.rich = length(unique(plant.sp)),
            morpho.rich = length(unique(k.type.s))) %>%
  pivot_longer(cols = c(taxon.rich, morpho.rich))

ggplot(diversity, aes(taxon.rich, morpho.rich, color = bb.sp)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(year~pollen)

ggplot(diversity_long, aes(elev.mean, value, color = name)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_grid(bb.sp~pollen + year, scales = "free")

ggplot(diversity_long, aes(bb.sp, value, fill = name, color = name)) +
  geom_violin(alpha = 0.5, draw_quantiles = c(0.25, 0.5, 0.75)) +
  facet_wrap(~year)
```

```
library(ggordiplots)

global_ordiplot <- gg_ordiplot(pcoa.nectar_dist, groups = groups, 
                               spiders = FALSE, hull = TRUE, 
                               ellipse = FALSE, label = FALSE)

ordiplot(pcoa.nectar_dist, type = "t")
plot(envfit(pcoa.nectar_dist, env = c(6.9, 12.7, 8.4, 7.3, 6.2, 7.5)))

str(pcoa.nectar_dist)

library(ggvegan)
fortify(pcoa.nectar_dist)

autoplot(pcoa.nectar_dist)
```

### GAMS
```{r, echo = FALSE}
pollen.gam.2010 <- bam(niche.overlap ~ 
                         s(sp.pair, bs = "re") +
                         te(yday, elev.mean, by = sp.pair, bs = c("gp", "tp"), k = 5),
                       data = filter(niche_overlap.sp, year == 2010),
                       method = "fREML",
                       discrete= TRUE,
                       family = "scat") %>% getViz()

check.gamViz(pollen.gam.2010)
pollen.gam.2010.sum <- summary(pollen.gam.2010)
print(plot(pollen.gam.2010, allTerms = TRUE), pages = 1)

pollen.gam.2011 <- bam(niche.overlap ~ 
                         s(sp.pair, bs = "re") +
                         te(yday, elev.mean, by = sp.pair, bs = c("gp", "tp"), k = 5),
                       data = filter(niche_overlap.sp, year == 2011),
                       method = "fREML",
                       discrete= TRUE,
                       family = "scat") %>% getViz()

check.gamViz(pollen.gam.2011)
pollen.gam.2011.sum <- summary(pollen.gam.2011)
print(plot(pollen.gam.2011, allTerms = TRUE), pages = 1)

pollen.gam.2012 <- bam(niche.overlap ~ 
                         s(sp.pair, bs = "re") +
                         te(yday, elev.mean, by = sp.pair, bs = c("gp", "tp"), k = 5),
                       data = filter(niche_overlap.sp, year == 2012),
                       method = "fREML",
                       discrete= TRUE,
                       family = "scat") %>% getViz()

check.gamViz(pollen.gam.2012)
pollen.gam.2012.sum <- summary(pollen.gam.2012)
print(plot(pollen.gam.2012, allTerms = TRUE), pages = 1)
```

### 2020-11-26 analyses/old_scripts/niche_relations.Rmd
did so many things

---
title: "site-wise niche-overlap by elevation"
output: html_document
---

Load packages 
```{r message=FALSE, warning=FALSE, include=FALSE}
library(bipartite)
library(mgcv)
library(mgcViz)
library(ggbipart)
library(GGally)
library(vegan)
library(ggvegan)
library(ggrepel)
library(gridExtra)
library(tidyverse)
library(lubridate)
library(patchwork)
library(tidymodels)
```

Load data
```{r message=FALSE, warning=FALSE, include=FALSE}
# Site data
site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, 
                   elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, 
                             levels = c("unten", "mitte", "oben")),
         elev.class2 = factor(elev.class2, 
                             levels = c("low", "mid", "high"))) # turn elav.class into an ordered factor

# BB traits
bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  mutate(pbl.fac = factor(if_else(bb.sp %in% c("hort", "pasc", "wurf"), "long", "short")))

# FL traits
fl_tax <- read_csv("../processed_data/floral_tax.csv")

fl_ktype <- read_csv("../processed_data/floral_k_type.csv")

fl_morph <- read_csv("../processed_data/flower_morphology_GitaBenadi.csv") %>%
  dplyr::select(SampleID, plant.family = Family, plant.sp = Species, 
         Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam),
            LN = mean(LN))

fl_ktype_key <- read_delim("../processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  dplyr::select(k.type = 1, description = 2)

fl_traits <- fl_tax %>%
  full_join(fl_ktype) %>%
  full_join(fl_morph) %>%
  dplyr::select(plant.sp, plant.genus, plant.family, 
         k.type, k.type.s, k.type.ss, diam, LN)

# Network
net <- read_csv("../processed_data/network.csv") %>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  mutate(year = factor(year)) %>% # convert year to factor
  left_join(site_data, by = "site") %>%
  #left_join(gdd, by = c("site", "date")) %>%
  ungroup() %>%
  arrange(site, date) %>%
  left_join(fl_traits) %>%

  # Fill in missing k.type scores
  mutate(k.type.s = if_else(plant.genus == "Phyteuma", 7.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Phyteuma", 7, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Aconitum", 5.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Aconitum", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Cardamine", 1.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Cardamine", 1, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Euphrasia", 5.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Euphrasia", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Lamium", 5.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Lamium", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Mentha", 2.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Mentha", 2, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Rumex", 0.0, k.type.s),
         k.type.ss = if_else(plant.genus == "Rumex", 0, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Salix", 9.0, k.type.s),
         k.type.ss = if_else(plant.genus == "Salix", 9, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Silene", 4.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Silene", 4, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Stachys", 5.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Stachys", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Taraxacum", 7.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Taraxacum", 7, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.sp == "Gentiana aspera", 2.1, k.type.s),
         k.type.ss = if_else(plant.sp == "Gentiana aspera", 2, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.sp == "Gentiana ciliata", 4.1, k.type.s),
         k.type.ss = if_else(plant.sp == "Gentiana ciliata", 4, k.type.ss))

# Floral survey
survey <- read_csv("../processed_data/floral_survey.csv") %>%
  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  left_join(fl_traits) %>%

  # Fill in missing k.type scores
  mutate(k.type.s = if_else(plant.genus == "Phyteuma", 7.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Phyteuma", 7, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Aconitum", 5.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Aconitum", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Cardamine", 1.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Cardamine", 1, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Euphrasia", 5.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Euphrasia", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Lamium", 5.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Lamium", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Mentha", 2.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Mentha", 2, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Rumex", 0.0, k.type.s),
         k.type.ss = if_else(plant.genus == "Rumex", 0, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Salix", 9.0, k.type.s),
         k.type.ss = if_else(plant.genus == "Salix", 9, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Silene", 4.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Silene", 4, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Stachys", 5.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Stachys", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Taraxacum", 7.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Taraxacum", 7, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.sp == "Gentiana aspera", 2.1, k.type.s),
         k.type.ss = if_else(plant.sp == "Gentiana aspera", 2, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.sp == "Gentiana ciliata", 4.1, k.type.s),
         k.type.ss = if_else(plant.sp == "Gentiana ciliata", 4, k.type.ss))

### Auxiliary data, i.e. all the covariates we need to control for
BB_covars <- net %>%
  group_by(site, year) %>%
  summarize(bb.rich = length(unique(bb.sp)),
            sqrt.bb.abund = sqrt(n()))

FL_covars <- survey %>%
  mutate(year = factor(year(date))) %>%
  filter(flower.cover > 0) %>%
  group_by(site, year) %>%
  summarize(fl.rich = length(unique(plant.sp)),
            log.fl.abund = log(sum(flower.cover)))

### Species-level by site (we will also separate pollen and nectar foragers)
BB_covars.splevel <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  group_by(bb.sp, site, year, pollen) %>%
  summarize(bb.sp.abund = n())

### Species names and such for plotting
species_names <- cbind(c("bss", "pasc", "prat", 
                         "soro", "wurf", "hort", 
                         "pyre", "mend", "mont", 
                         "gers", "hypn","jone", 
                         "lapi", "muci", "psit", 
                         "humi"),
                       c("B. sensu-strictu", "B. pascuorum", "B. pratorum", 
                         "B. soroensis", "B. wurflenii", "B. hortorum",
                         "B. pyrenaeus", "B. mendax", "B. monticola",
                         "B. gerstaeckeri", "B. hypnorum", "B. jonellus", 
                         "B. lapidarius", "B. mucidus", "B. psithyrus", 
                         "B. humilis"),
                       c("B. ss.", "B. pasc.", "B. prat.", 
                         "B. soro.", "B. wurf.", "B. hort.",
                         "B. pyre.", "B. mend.", "B. mont.",
                         "B. gers.", "B. hypn.", "B. jone.", 
                         "B. lapi.", "B. muci.", "B. psit.", 
                         "B. humi"),
                       c("major", "major", "major", 
                         "major", "major", "common", 
                         "high elev.", "high elev.", "high elev.",
                         "minor", "minor", "common", 
                         "minor", "minor", "common", 
                         "minor")
                       ) %>%
  as.data.frame() %>%
  as_tibble() %>%
  dplyr::select(bb.sp = 1, species.name = 2, species.abb = 3, abund.class = 4) %>%
  mutate(focal = if_else(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf"), TRUE, FALSE))
```

We will focus on six species that are abundant and widespread:  
  1. *B. sensu-strictu*
  2. *B. pascuorum*
  3. *B. pratorum*
  4. *B. hortorum*
  5. *B. soroensis*
  6. *B. wurflenii*

```{r fig.cap="**Figure 1: Focal species**", message=FALSE, warning=FALSE, echo=FALSE}
bb_range.year <- net %>%
  left_join(site_data) %>%
  group_by(site, date, elev.mean, bb.sp, year) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, bb.sp, year) %>%
  summarize(abund = mean(abund)) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2)) %>%
  left_join(species_names)

ggplot(bb_range.year, aes(reorder(bb.sp, elev.med), elev.mean, color = focal)) +
  geom_line(size = 2, alpha = 0.25) +
  geom_point(aes(size = abund), alpha = 0.5) +
  facet_wrap(~year, ncol = 1) +
  theme_light(12) +
  scale_size_continuous(name = "Abundance") +
  scale_color_discrete(name = "Focal species") +
  xlab("BB species") +
  ylab("Elevation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Functions
```{r}
# Niche overlap by sp.pair * site * year
niche_overlapper <- function(x, criterion) {

  criterion <- enquo(criterion) # criterion is set to either plant.sp (for taxonomic) or k.type.s (for morphological)

  x %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    filter(!is.na(pollen)) %>% # drop observations with no pollen/nectar classification
    group_by(bb.sp, pollen, !!criterion, site, year) %>% # group by site-year and BB-FL pair
    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-year
    group_by(bb.sp, pollen, site, year) %>% # group by site-year and BB species
    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
    dplyr::select(-freq) %>% # drop freq variable
    pivot_wider(names_from = !!criterion, values_from = prop) %>% # spread into the kind of array vegan likes
    replace(is.na(.), 0) %>% # replace NAs with zeros
    group_by(pollen, site, year) %>%
    nest() %>% # nest into list-column
    mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
           dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
           tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
    dplyr::select(pollen, site, year, tidy_dist) %>% # drop intermediate list-columns
    unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
    left_join(site_data) %>% # add site data
    dplyr::select(site, year, pollen, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
    left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
    rename(subgenus1 = subgenus, pbl.w1 = pbl.w, pbl.w.class1 = pbl.w.class) %>%
    left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
    rename(subgenus2 = subgenus, pbl.w2 = pbl.w, pbl.w.class2 = pbl.w.class) %>%
    dplyr::select(site, year, pollen, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
           pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
    left_join(BB_covars.splevel, by = c("bb.sp1" = "bb.sp", "site", "year", "pollen")) %>%
    rename(bb.sp1.abund = bb.sp.abund) %>%
    dplyr::select(site, year, pollen, bb.sp1, bb.sp2, bb.sp1.abund, subgenus1, subgenus2, distance, 
           pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
    left_join(BB_covars.splevel, by = c("bb.sp2" = "bb.sp", "site", "year", "pollen")) %>%
    rename(bb.sp2.abund = bb.sp.abund) %>%
    dplyr::select(site, year, pollen, bb.sp1, bb.sp2, bb.sp1.abund, bb.sp2.abund, subgenus1, subgenus2, distance, 
           pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
    mutate(pbl.diff = abs(pbl.w1 - pbl.w2), # tongue length differences
           abund.diff = abs(bb.sp1.abund - bb.sp2.abund),
           subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different"), # binary: same or different subgenus

           # Logical columns indicating whether a given species belongs to a given species pair
           # This is handy for extracting data for individual species across all their counterparts
           bss = if_else(bb.sp1 == "bss" | bb.sp2 == "bss", TRUE, FALSE),
           pasc = if_else(bb.sp1 == "pasc" | bb.sp2 == "pasc", TRUE, FALSE),
           wurf = if_else(bb.sp1 == "wurf" | bb.sp2 == "wurf", TRUE, FALSE),
           prat = if_else(bb.sp1 == "prat" | bb.sp2 == "prat", TRUE, FALSE),
           soro = if_else(bb.sp1 == "soro" | bb.sp2 == "soro", TRUE, FALSE),
           hort = if_else(bb.sp1 == "hort" | bb.sp2 == "hort", TRUE, FALSE)) %>%

    unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
    mutate(sp.pair = factor(sp.pair),
           pollen = factor(pollen)) %>% # ...and make it a factor so that mgcv can handle it
    left_join(BB_covars) %>% # add BB covars
    left_join(FL_covars) %>% # add floral covars
    mutate(niche.overlap = 1-distance) # express niche overlap so that 1 = perfect overlap and 0 = no overlap
}

niche_overlapper2 <- function(x, criterion) {

  criterion <- enquo(criterion) # criterion is set to either plant.sp (for taxonomic) or k.type.s (for morphological)

  x %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    filter(!is.na(pollen)) %>% # drop observations with no pollen/nectar classification
    group_by(bb.sp, pollen, !!criterion, site) %>% # group by site-year and BB-FL pair
    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-year
    group_by(bb.sp, pollen, site) %>% # group by site-year and BB species
    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
    dplyr::select(-freq) %>% # drop freq variable
    pivot_wider(names_from = !!criterion, values_from = prop) %>% # spread into the kind of array vegan likes
    replace(is.na(.), 0) %>% # replace NAs with zeros
    group_by(pollen, site) %>%
    nest() %>% # nest into list-column
    mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
           dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
           tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
    dplyr::select(pollen, site, tidy_dist) %>% # drop intermediate list-columns
    unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
    left_join(site_data) %>% # add site data
    dplyr::select(site, pollen, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
    left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
    rename(subgenus1 = subgenus, pbl.w1 = pbl.w, pbl.w.class1 = pbl.w.class) %>%
    left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
    rename(subgenus2 = subgenus, pbl.w2 = pbl.w, pbl.w.class2 = pbl.w.class) %>%
    dplyr::select(site, pollen, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
           pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
    mutate(pbl.diff = abs(pbl.w1 - pbl.w2), # tongue length differences
           subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different"), # binary: same or different subgenus

           # Logical columns indicating whether a given species belongs to a given species pair
           # This is handy for extracting data for individual species across all their counterparts
           bss = if_else(bb.sp1 == "bss" | bb.sp2 == "bss", TRUE, FALSE),
           pasc = if_else(bb.sp1 == "pasc" | bb.sp2 == "pasc", TRUE, FALSE),
           wurf = if_else(bb.sp1 == "wurf" | bb.sp2 == "wurf", TRUE, FALSE),
           prat = if_else(bb.sp1 == "prat" | bb.sp2 == "prat", TRUE, FALSE),
           soro = if_else(bb.sp1 == "soro" | bb.sp2 == "soro", TRUE, FALSE),
           hort = if_else(bb.sp1 == "hort" | bb.sp2 == "hort", TRUE, FALSE)) %>%

    unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
    mutate(sp.pair = factor(sp.pair),
           pollen = factor(pollen)) %>% # ...and make it a factor so that mgcv can handle it
    mutate(niche.overlap = 1-distance) # express niche overlap so that 1 = perfect overlap and 0 = no overlap
}

# Add BB trait data for envfit() call inside map2()
trait_join <- function(x, y = bb_traits) {
  x %>%
    select(bb.sp) %>%
    left_join(y) %>%
    column_to_rownames(var = "bb.sp") %>%
    select(pbl.w)
}

# Fortify envfit() object fitted to BB tongue length data
envfit_fort <- function(x){
  fortify(x) %>%
  select(Score = Type, Label, MDS1, MDS2) %>%
  mutate(Label = case_when(Label == "pbl.w" ~ "T.len"),
         Score = case_when(Score == "Vector" ~ "T.len"))
}

# Fortify PCOA and add the results of an envfit() to BB tongue length
pcoa_fort <- function(x, y){
  fortify(x, axes = 1:2) %>%
  bind_rows(y) %>%
  mutate(Score = case_when(
    Score == "species" ~ "FL.sp",
    Score == "sites" ~ "BB.sp",
    Score == "T.len" ~ "T.len"))
}

# Plot PCOA with repelled text and arrow for tongue length; to avoid overplotting, loadings < min_load are not plotted
pcoa_plot <- function(x, y, min_load = 0.25) {
  ggplot(filter(x, (abs(MDS1) > min_load | abs(MDS2) > min_load) | (Score %in% c("BB.sp", "T.len")))) + # to avoid overplotting, we can filter to include only floral species > loading threshold
    geom_point(aes(MDS1, MDS2, color = Score)) +
    geom_text_repel(aes(MDS1, MDS2, color = Score, label = Label), key_glyph = "rect") +
    geom_segment(data = y, 
               aes(x = 0, y = 0, xend = MDS1, yend = MDS2, colour = Label),
               arrow = arrow(), size = 2, alpha = 0.5, key_glyph = "rect")
}

# Niche analysis function; generates PCOA, envfit, ordination plot, and tongue length R2 coming out of envfit
niche_analysis1.year <- function(x, criterion) {

  criterion <- enquo(criterion)

  x %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
    filter(!is.na(pollen)) %>%
    group_by(year, bb.sp, !!criterion, pollen) %>%
    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site
    group_by(year, bb.sp, pollen) %>%
    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair per site-date
    dplyr::select(-freq) %>%
    group_by(pollen) %>%
    nest() %>%
    mutate(data2 = map(data, function(x) pivot_wider(x, names_from = !!criterion, values_from = prop)), # spread into the kind of array vegan likes
           data3 = map(data2, function(x) replace(x, is.na(x), 0)),
           data4 = map(data3, function(x) unite(x, id, c("year", "bb.sp"))),
           data5 = map(data4, function(x) column_to_rownames(x, var = "id")),
           pcoa = map(data5,  function(x) capscale(x ~ 1, distance = "horn", scale = TRUE)),
           traits = map(data3, function(x) x %>% 
                          left_join(bb_traits) %>% 
                          unite(id, c(year, bb.sp)) %>% 
                          column_to_rownames(var = "id") %>% 
                          select(pbl.w)),
           envfit = map2(pcoa, traits, function(x, y) envfit(x, y)),
           envfit.fort = map(envfit, envfit_fort),
           pcoa.fort = map2(pcoa, envfit.fort, pcoa_fort),
           pcoa.plot = map2(pcoa.fort, envfit.fort, pcoa_plot), 
           tongue.r2 = map(envfit, function(x) x$vectors$r),
           tongue.p = map(envfit, function(x) x$vectors$pvals),
           tongue.MDS1 = map(envfit, function(x) x$vectors$arrows[[1]]),
           tongue.MDS2 = map(envfit, function(x) x$vectors$arrows[[2]])) %>%
    unnest(cols = c(tongue.r2, tongue.p, tongue.MDS1, tongue.MDS2)) %>%
    select(pollen, pcoa, envfit, envfit.fort, pcoa.fort, pcoa.plot, tongue.r2, tongue.p, tongue.MDS1, tongue.MDS2) 
}

niche_analysis1 <- function(x, criterion) {

  criterion <- enquo(criterion)

  x %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
    filter(!is.na(pollen)) %>%
    group_by(bb.sp, !!criterion, pollen) %>%
    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site
    group_by(bb.sp, pollen) %>%
    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair per site-date
    dplyr::select(-freq) %>%
    group_by(pollen) %>%
    nest() %>%
    mutate(data2 = map(data, function(x) pivot_wider(x, names_from = !!criterion, values_from = prop)), # spread into the kind of array vegan likes
           data3 = map(data2, function(x) replace(x, is.na(x), 0)),
           data4 = map(data3, function(x) column_to_rownames(x, var = "bb.sp")),
           pcoa = map(data4,  function(x) capscale(x ~ 1, distance = "horn")),
           traits = map(data3, function(x) x %>% 
                          left_join(bb_traits) %>% 
                          column_to_rownames(var = "bb.sp") %>% 
                          select(pbl.w)),
           envfit = map2(pcoa, traits, function(x, y) envfit(x, y)),
           envfit.fort = map(envfit, envfit_fort),
           pcoa.fort = map2(pcoa, envfit.fort, pcoa_fort),
           pcoa.plot = map2(pcoa.fort, envfit.fort, pcoa_plot), 
           tongue.r2 = map(envfit, function(x) x$vectors$r),
           tongue.p = map(envfit, function(x) x$vectors$pvals),
           MDS1.varexp = map(pcoa, function(x) x$CA$eig[[1]]),
           MDS2.varexp = map(pcoa, function(x) x$CA$eig[[2]])) %>%
    unnest(cols = c(tongue.r2, tongue.p, MDS1.varexp, MDS2.varexp)) %>%
    select(pollen, pcoa, envfit, envfit.fort, pcoa.fort, pcoa.plot, tongue.r2, tongue.p, MDS1.varexp, MDS2.varexp) 
}

# This version break it down by elev.class
niche_analysis2.year <- function(x, criterion) {

  criterion <- enquo(criterion)

  x %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
    filter(!is.na(pollen)) %>%
    group_by(year, elev.class, bb.sp, !!criterion, pollen) %>%
    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site
    group_by(year, elev.class, bb.sp, pollen) %>%
    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair per site-date
    dplyr::select(-freq) %>%
    group_by(pollen, elev.class) %>%
    nest() %>%
    mutate(data2 = map(data, function(x) pivot_wider(x, names_from = !!criterion, values_from = prop)), # spread into the kind of array vegan likes
           data3 = map(data2, function(x) replace(x, is.na(x), 0)),
           data4 = map(data3, function(x) unite(x, id, c("year", "bb.sp"))),
           data5 = map(data4, function(x) column_to_rownames(x, var = "id")),
           pcoa = map(data5,  function(x) capscale(x ~ 1, distance = "horn", scale = TRUE)),
           traits = map(data3, function(x) x %>% 
                          left_join(bb_traits) %>% 
                          unite(id, c(year, bb.sp)) %>% 
                          column_to_rownames(var = "id") %>% 
                          select(pbl.w)),
           envfit = map2(pcoa, traits, function(x, y) envfit(x, y)),
           envfit.fort = map(envfit, envfit_fort),
           pcoa.fort = map2(pcoa, envfit.fort, pcoa_fort),
           pcoa.plot = map2(pcoa.fort, envfit.fort, pcoa_plot),
           tongue.r2 = map(envfit, function(x) x$vectors$r),
           tongue.p = map(envfit, function(x) x$vectors$pvals),
           tongue.MDS1 = map(envfit, function(x) x$vectors$arrows[[1]]),
           tongue.MDS2 = map(envfit, function(x) x$vectors$arrows[[2]])) %>%
    unnest(cols = c(tongue.r2, tongue.p, tongue.MDS1, tongue.MDS2)) %>%
    select(pollen, elev.class, pcoa, envfit, envfit.fort, pcoa.fort, 
           pcoa.plot, tongue.r2, tongue.p, tongue.MDS1, tongue.MDS2) 
}

niche_analysis2 <- function(x, criterion) {

  criterion <- enquo(criterion)

  x %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
    filter(!is.na(pollen)) %>%
    group_by(elev.class, bb.sp, !!criterion, pollen) %>%
    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site
    group_by(elev.class, bb.sp, pollen) %>%
    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair per site-date
    dplyr::select(-freq) %>%
    group_by(pollen, elev.class) %>%
    nest() %>%
    mutate(data2 = map(data, function(x) pivot_wider(x, names_from = !!criterion, values_from = prop)), # spread into the kind of array vegan likes
           data3 = map(data2, function(x) replace(x, is.na(x), 0)),
           data4 = map(data3, function(x) column_to_rownames(x, var = "bb.sp")),
           pcoa = map(data4,  function(x) capscale(x ~ 1, distance = "horn", scale = TRUE)),
           traits = map(data3, function(x) x %>% 
                          left_join(bb_traits) %>% 
                          column_to_rownames(var = "bb.sp") %>% 
                          select(pbl.w)),
           envfit = map2(pcoa, traits, function(x, y) envfit(x, y)),
           envfit.fort = map(envfit, envfit_fort),
           pcoa.fort = map2(pcoa, envfit.fort, pcoa_fort),
           pcoa.plot = map2(pcoa.fort, envfit.fort, pcoa_plot),
           tongue.r2 = map(envfit, function(x) x$vectors$r),
           tongue.p = map(envfit, function(x) x$vectors$pvals),
           MDS1.varexp = map(pcoa, function(x) x$CA$eig[[1]]),
           MDS2.varexp = map(pcoa, function(x) x$CA$eig[[2]])) %>%
    unnest(cols = c(tongue.r2, tongue.p, MDS1.varexp, MDS2.varexp)) %>%
    select(pollen, elev.class, pcoa, envfit, envfit.fort, pcoa.fort, 
           pcoa.plot, tongue.r2, tongue.p, MDS1.varexp, MDS2.varexp)
  }

# ...and this version breaks it down by elev.class and period
niche_analysis3 <- function(x, criterion) {

  criterion <- enquo(criterion)

  x %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
    filter(!is.na(pollen)) %>%
    group_by(elev.class) %>%
    mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
    group_by(elev.class, period, pollen, bb.sp, !!criterion) %>%
    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site
    group_by(elev.class, period, bb.sp, pollen) %>%
    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair per site-date
    dplyr::select(-freq) %>%
    group_by(pollen, elev.class, period) %>%
    nest() %>%
    mutate(data2 = map(data, function(x) pivot_wider(x, names_from = !!criterion, values_from = prop)), # spread into the kind of array vegan likes
           data3 = map(data2, function(x) replace(x, is.na(x), 0)),
           data4 = map(data3, function(x) column_to_rownames(x, var = "bb.sp")),
           pcoa = map(data4,  function(x) capscale(x ~ 1, distance = "horn", scale = TRUE)),
           traits = map(data3, function(x) x %>% 
                          left_join(bb_traits) %>% 
                          column_to_rownames(var = "bb.sp") %>% 
                          select(pbl.w)),
           envfit = map2(pcoa, traits, function(x, y) envfit(x, y)),
           envfit.fort = map(envfit, envfit_fort),
           pcoa.fort = map2(pcoa, envfit.fort, pcoa_fort),
           pcoa.plot = map2(pcoa.fort, envfit.fort, pcoa_plot), 
           tongue.r2 = map(envfit, function(x) x$vectors$r),
           tongue.p = map(envfit, function(x) x$vectors$pvals),
           MDS1.varexp = map(pcoa, function(x) x$CA$eig[[1]]),
           MDS2.varexp = map(pcoa, function(x) x$CA$eig[[2]])) %>%
    unnest(cols = c(tongue.r2, tongue.p, MDS1.varexp, MDS2.varexp)) %>%
    select(pollen, elev.class, period, pcoa, envfit, envfit.fort,
           pcoa.fort, pcoa.plot, tongue.r2, tongue.p, MDS1.varexp, MDS2.varexp)
  }
```

Pairwise (bb.sp:bb.sp) niche overlap, pooled by site-year
```{r}
# By plant.sp

## Resolved to sp.pair
niche_overlap.tax1 <- niche_overlapper(net, criterion = plant.sp.abb)
niche_overlap.tax2 <- niche_overlapper2(net, criterion = plant.sp.abb)

## Collapsed to comunity mean
niche_overlap.tax.mean1 <- niche_overlap.tax1 %>%
  group_by(site, year, pollen, elev.mean) %>%
  summarize(niche.overlap.mean = mean(niche.overlap)) %>%
  left_join(BB_covars) %>% # add BB covars
  left_join(FL_covars)

niche_overlap.tax.mean2 <- niche_overlap.tax2 %>%
  group_by(site, pollen, elev.mean) %>%
  summarize(niche.overlap.mean = mean(niche.overlap))

# By floral k.type (simplified)

## Resolved to sp.pair
niche_overlap.morph1 <- niche_overlapper(net, criterion = plant.sp.abb)
niche_overlap.morph2 <- niche_overlapper2(net, criterion = plant.sp.abb)

## Collapsed to comunity mean
niche_overlap.morph.mean1 <- niche_overlap.morph1 %>%
  group_by(site, year, pollen, elev.mean) %>%
  summarize(niche.overlap.mean = mean(niche.overlap)) %>%
  left_join(BB_covars) %>% # add BB covars
  left_join(FL_covars)

niche_overlap.morph.mean2 <- niche_overlap.morph %>%
  group_by(site, pollen, elev.mean) %>%
  summarize(niche.overlap.mean = mean(niche.overlap))
```

# 1. What is the logic of niche partitioning between core species, and does it change through elevation and time?
## 1.1 Ordination by species-year with post-hoc fitting of tongue length
```{r}
# By plant.sp
niche.tax <- niche_analysis1(net, plant.sp.abb)
niche.tax <- niche_analysis1.year(net, plant.sp.abb)

## Extract and name plots
pollen.tax <- filter(niche.tax, pollen == "pollen")$pcoa.plot[[1]]
nectar.tax <- filter(niche.tax, pollen == "nopollen")$pcoa.plot[[1]]

## Display plots on grid
pdf("../analyses/output/tax_niche.pdf", height = 8.5, width = 11)
(pollen.tax | nectar.tax) + plot_layout(guides = 'collect') 
dev.off()

# By ktype
niche.morph <- niche_analysis1(net, k.type.s)

## Extract and name plots
pollen.morph <- filter(niche.morph, pollen == "pollen")$pcoa.plot[[1]]
nectar.morph <- filter(niche.morph, pollen == "nopollen")$pcoa.plot[[1]]

## Display plots on grid
pdf("../analyses/output/morph_niche.pdf", height = 8.5, width = 11)
(pollen.morph | nectar.morph) + plot_layout(guides = 'collect') 
dev.off()
```

## 1.2 Ordination by species-year with post-hoc fitting of tongue length, broken down by elevation class
```{r}
# By plant.sp
niche_x_elev.tax <- niche_analysis2(net, plant.sp.abb)

## Extract and name plots
pollen_oben.tax <- filter(niche_x_elev.tax, pollen == "pollen" & elev.class == "oben")$pcoa.plot[[1]]
pollen_mitte.tax <- filter(niche_x_elev.tax, pollen == "pollen" & elev.class == "mitte")$pcoa.plot[[1]]
pollen_unten.tax <- filter(niche_x_elev.tax, pollen == "pollen" & elev.class == "unten")$pcoa.plot[[1]]

nectar_oben.tax <- filter(niche_x_elev.tax, pollen == "nopollen" & elev.class == "oben")$pcoa.plot[[1]]
nectar_mitte.tax <- filter(niche_x_elev.tax, pollen == "nopollen" & elev.class == "mitte")$pcoa.plot[[1]]
nectar_unten.tax <- filter(niche_x_elev.tax, pollen == "nopollen" & elev.class == "unten")$pcoa.plot[[1]]

## Display plots on grid
pdf("../analyses/output/tax_niche_x_elev.pdf", height = 8.5, width = 11)
(pollen_oben.tax | nectar_oben.tax) /
(pollen_mitte.tax | nectar_mitte.tax) /
(pollen_unten.tax | nectar_unten.tax) + plot_layout(guides = 'collect') 
dev.off()

# By ktype
niche_x_elev.morph <- niche_analysis2(net, k.type.s)

## Extract and name plots
pollen_oben.morph <- filter(niche_x_elev.morph, pollen == "pollen" & elev.class == "oben")$pcoa.plot[[1]]
pollen_mitte.morph <- filter(niche_x_elev.morph, pollen == "pollen" & elev.class == "mitte")$pcoa.plot[[1]]
pollen_unten.morph <- filter(niche_x_elev.morph, pollen == "pollen" & elev.class == "unten")$pcoa.plot[[1]]

nectar_oben.morph <- filter(niche_x_elev.morph, pollen == "nopollen" & elev.class == "oben")$pcoa.plot[[1]]
nectar_mitte.morph <- filter(niche_x_elev.morph, pollen == "nopollen" & elev.class == "mitte")$pcoa.plot[[1]]
nectar_unten.morph <- filter(niche_x_elev.morph, pollen == "nopollen" & elev.class == "unten")$pcoa.plot[[1]]

## Display plots on grid
pdf("../analyses/output/morp_niche_x_elev.pdf", height = 8.5, width = 11)
(pollen_oben.morph | nectar_oben.morph) /
(pollen_mitte.morph | nectar_mitte.morph) /
(pollen_unten.morph | nectar_unten.morph) + plot_layout(guides = 'collect') 
dev.off()
```

## 1.3 Ordination by BB species * elev.class * period, pooling across years
```{r}
# By plant species
niche_x_elevperiod.tax <- niche_analysis3(net, plant.sp.abb)

## Extract and name plots
pollen_oben_early.tax <- filter(niche_x_elevperiod.tax, pollen == "pollen" & elev.class == "oben" & period == "early")$pcoa.plot[[1]]
pollen_oben_mid.tax <- filter(niche_x_elevperiod.tax, pollen == "pollen" & elev.class == "oben" & period == "mid")$pcoa.plot[[1]]
pollen_oben_late.tax <- filter(niche_x_elevperiod.tax, pollen == "pollen" & elev.class == "oben" & period == "late")$pcoa.plot[[1]]

pollen_mitte_early.tax <- filter(niche_x_elevperiod.tax, pollen == "pollen" & elev.class == "mitte" & period == "early")$pcoa.plot[[1]]
pollen_mitte_mid.tax <- filter(niche_x_elevperiod.tax, pollen == "pollen" & elev.class == "mitte" & period == "mid")$pcoa.plot[[1]]
pollen_mitte_late.tax <- filter(niche_x_elevperiod.tax, pollen == "pollen" & elev.class == "mitte" & period == "late")$pcoa.plot[[1]]

#pollen_unten_early.tax <- filter(niche_x_elevperiod.tax, pollen == "pollen" & elev.class == "unten" & period == "early")$pcoa.plot[[1]]
pollen_unten_mid.tax <- filter(niche_x_elevperiod.tax, pollen == "pollen" & elev.class == "unten" & period == "mid")$pcoa.plot[[1]]
pollen_unten_late.tax <- filter(niche_x_elevperiod.tax, pollen == "pollen" & elev.class == "unten" & period == "late")$pcoa.plot[[1]]

nectar_oben_early.tax <- filter(niche_x_elevperiod.tax, pollen == "nopollen" & elev.class == "oben" & period == "early")$pcoa.plot[[1]]
nectar_oben_mid.tax <- filter(niche_x_elevperiod.tax, pollen == "nopollen" & elev.class == "oben" & period == "mid")$pcoa.plot[[1]]
nectar_oben_late.tax <- filter(niche_x_elevperiod.tax, pollen == "nopollen" & elev.class == "oben" & period == "late")$pcoa.plot[[1]]

nectar_mitte_early.tax <- filter(niche_x_elevperiod.tax, pollen == "nopollen" & elev.class == "mitte" & period == "early")$pcoa.plot[[1]]
nectar_mitte_mid.tax <- filter(niche_x_elevperiod.tax, pollen == "nopollen" & elev.class == "mitte" & period == "mid")$pcoa.plot[[1]]
nectar_mitte_late.tax <- filter(niche_x_elevperiod.tax, pollen == "nopollen" & elev.class == "mitte" & period == "late")$pcoa.plot[[1]]

nectar_unten_early.tax <- filter(niche_x_elevperiod.tax, pollen == "nopollen" & elev.class == "unten" & period == "early")$pcoa.plot[[1]]
nectar_unten_mid.tax <- filter(niche_x_elevperiod.tax, pollen == "nopollen" & elev.class == "unten" & period == "mid")$pcoa.plot[[1]]
nectar_unten_late.tax <- filter(niche_x_elevperiod.tax, pollen == "nopollen" & elev.class == "unten" & period == "late")$pcoa.plot[[1]]

## Display plots on grid
pdf("../analyses/output/tax_niche_x_elevperiod_pollen.pdf", height = 8.5, width = 11)
(pollen_oben_early.tax | pollen_oben_mid.tax | pollen_oben_late.tax) /
(pollen_mitte_early.tax | pollen_mitte_mid.tax | pollen_mitte_late.tax) /
(plot_spacer() | pollen_unten_mid.tax  | pollen_unten_late.tax)  + plot_layout(guides = 'collect') 
dev.off()

pdf("../analyses/output/tax_niche_x_elevperiod_nectar.pdf", height = 8.5, width = 11)
(nectar_oben_early.tax | nectar_oben_mid.tax | nectar_oben_late.tax) /
(nectar_mitte_early.tax | nectar_mitte_mid.tax | nectar_mitte_late.tax) /
(nectar_unten_early.tax | nectar_unten_mid.tax  | nectar_unten_late.tax)  + plot_layout(guides = 'collect') 
dev.off()

# By ktype
niche_x_elevperiod.morph <- niche_analysis3(net, k.type.s)

## Extract and name plots
pollen_oben_early.morph <- filter(niche_x_elevperiod.morph, pollen == "pollen" & elev.class == "oben" & period == "early")$pcoa.plot[[1]]
pollen_oben_mid.morph <- filter(niche_x_elevperiod.morph, pollen == "pollen" & elev.class == "oben" & period == "mid")$pcoa.plot[[1]]
pollen_oben_late.morph <- filter(niche_x_elevperiod.morph, pollen == "pollen" & elev.class == "oben" & period == "late")$pcoa.plot[[1]]

pollen_mitte_early.morph <- filter(niche_x_elevperiod.morph, pollen == "pollen" & elev.class == "mitte" & period == "early")$pcoa.plot[[1]]
pollen_mitte_mid.morph <- filter(niche_x_elevperiod.morph, pollen == "pollen" & elev.class == "mitte" & period == "mid")$pcoa.plot[[1]]
pollen_mitte_late.morph <- filter(niche_x_elevperiod.morph, pollen == "pollen" & elev.class == "mitte" & period == "late")$pcoa.plot[[1]]

#pollen_unten_early.morph <- filter(niche_x_elevperiod.morph, pollen == "pollen" & elev.class == "unten" & period == "early")$pcoa.plot[[1]]
pollen_unten_mid.morph <- filter(niche_x_elevperiod.morph, pollen == "pollen" & elev.class == "unten" & period == "mid")$pcoa.plot[[1]]
pollen_unten_late.morph <- filter(niche_x_elevperiod.morph, pollen == "pollen" & elev.class == "unten" & period == "late")$pcoa.plot[[1]]

nectar_oben_early.morph <- filter(niche_x_elevperiod.morph, pollen == "nopollen" & elev.class == "oben" & period == "early")$pcoa.plot[[1]]
nectar_oben_mid.morph <- filter(niche_x_elevperiod.morph, pollen == "nopollen" & elev.class == "oben" & period == "mid")$pcoa.plot[[1]]
nectar_oben_late.morph <- filter(niche_x_elevperiod.morph, pollen == "nopollen" & elev.class == "oben" & period == "late")$pcoa.plot[[1]]

nectar_mitte_early.morph <- filter(niche_x_elevperiod.morph, pollen == "nopollen" & elev.class == "mitte" & period == "early")$pcoa.plot[[1]]
nectar_mitte_mid.morph <- filter(niche_x_elevperiod.morph, pollen == "nopollen" & elev.class == "mitte" & period == "mid")$pcoa.plot[[1]]
nectar_mitte_late.morph <- filter(niche_x_elevperiod.morph, pollen == "nopollen" & elev.class == "mitte" & period == "late")$pcoa.plot[[1]]

nectar_unten_early.morph <- filter(niche_x_elevperiod.morph, pollen == "nopollen" & elev.class == "unten" & period == "early")$pcoa.plot[[1]]
nectar_unten_mid.morph <- filter(niche_x_elevperiod.morph, pollen == "nopollen" & elev.class == "unten" & period == "mid")$pcoa.plot[[1]]
nectar_unten_late.morph <- filter(niche_x_elevperiod.morph, pollen == "nopollen" & elev.class == "unten" & period == "late")$pcoa.plot[[1]]

## Display plots on grid
pdf("../analyses/output/morph_niche_x_elevperiod_pollen.pdf", height = 8.5, width = 11)
(pollen_oben_early.morph | pollen_oben_mid.morph | pollen_oben_late.morph) /
(pollen_mitte_early.morph | pollen_mitte_mid.morph | pollen_mitte_late.morph) /
(plot_spacer() | pollen_unten_mid.morph  | pollen_unten_late.morph)  + plot_layout(guides = 'collect') 
dev.off()

pdf("../analyses/output/morph_niche_x_elevperiod_nectar.pdf", height = 8.5, width = 11)
(nectar_oben_early.morph | nectar_oben_mid.morph | nectar_oben_late.morph) /
(nectar_mitte_early.morph | nectar_mitte_mid.morph | nectar_mitte_late.morph) /
(nectar_unten_early.morph | nectar_unten_mid.morph  | nectar_unten_late.morph)  + plot_layout(guides = 'collect') 
dev.off()

```

## 1.4 Does the influence of tongue length on niche partitioning vary with elevation or time?
```{r}
# By plant.sp
ggplot(niche_x_elevperiod.tax, aes(elev.class, tongue.r2)) +
  geom_boxplot() +
  facet_wrap(~pollen)

ggplot(niche_x_elevperiod.tax, aes(period, tongue.r2)) +
  geom_boxplot() +
  facet_wrap(~pollen)

# By ktype
ggplot(niche_x_elevperiod.morph, aes(elev.class, tongue.r2)) +
  geom_boxplot() +
  facet_wrap(~pollen)

ggplot(niche_x_elevperiod.morph, aes(period, tongue.r2)) +
  geom_boxplot() +
  facet_wrap(~pollen)
```

# 2. Testing the stress-gradient hypothesis and the Ranta and Vepsäläinen heterogeneity hypothesis

**Stress-gradient hypothesis:** Is there any evidence that niche partitioning amongst generalist core bumble bee species is governed by an inverse relationship between elevation and competition, broadly consistent with the stress-gradient hypothesis (Bertness and Callaway 1994). This mechanism has been inferred from trait patterns in the whole bee community of the Berchtesgadener Alps (Hoiss et al. 2012) and from interaction specialization patterns in the plant-pollinator community of Mt. Kilimanjaro (Classen et al. 2020). It predicts that niche overlap will increase with elevation, reflecting competitive pressures at low elevations and reduced competition but increased need for opportunism at high elevations.

**Ranta/Vepsäläinen heterogeneity hypothesis :** Ranta and Vepsäläinen (1981) proposed that bumble bees communities are governed by a spatiotemporally heterogeneous mosaic of competitive conditions, such that competitive inequalities between species are sufficiently variable to allow more species to coexist than predicted by trait-based competition models.

## 2.1: Testing the stress-gradient hypothesis
### 2.1.1 Face-value analysis

*Results:*  
In the taxonomic model (28% variance explained), niche overlap exhibited a unimodal relationship with elevation in 2010 and 2012 (both significant at p < 0.01), with peak overlap at ~1600 m and ~1250 m, respectively. In 2011, peak niche overlap occurred at the lowest elevations, but after an initial descent, it exhibit a unimodal relationship rather like that of 2010; **this relationship was not significant**, though (p = 0.23). Niche overlap of pollen visits was significantly (p = 0.01) lower than that of nectar visits. There was no significant effect of year.

In the morphological model (31% variance explained), niche overlap exhibited a significant relationship with elevation in all three years. In each year, the relationship was approximately quadratic, peaking at low elevation and descending with increasing elevation. In 2010 and 2012, the curve drops off steepy above 1500 m. In 2011, the curve is nearly linear.  Niche overlap of pollen visits was significantly (p = 0.01) lower than that of nectar visits, and there was no significant effect of year.

These results, of course, do not support H1. But there is a caveat. These models try to explain face-value niche overlap, but they do not try to distinguish observed niche overlap from the predictions of a null model. It is possible that the patterns we see would be predicted simply by the abundance relationships in our community. Niche overlap is, in fact, strongly predicted by bumble bee abundance, which is itself strongly correlated with bumble bee richness. One response is, well, so what? These are the observed patterns of niche overlap, and we may not care to disentangle the effects of abundance from those of behavioral changes.

```{r}
## Taxonomic niche partitioning

# Whole community (pairwise)
ggplot(niche_overlap.tax1, aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10)) +
  facet_wrap(~year)

ggplot(niche_overlap.tax2, aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

# Whole community (collapsed to mean)
ggplot(niche_overlap.tax.mean, aes(elev.mean, niche.overlap.mean, color = pollen)) +
  geom_point(alpha = 0.25) +
  facet_wrap(~year) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

ggplot(niche_overlap.tax.mean2, aes(elev.mean, niche.overlap.mean, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

# GAM
niche_comm_tax_GAM <- gam(niche.overlap.mean ~ 
                           year +
                           pollen +
                          # s(sqrt.bb.abund, k = 4) +
                           s(elev.mean, by = year, k = 10, m = 2),
                         method = "REML",
                         family = "gaussian",
                         data = niche_overlap.tax.mean) %>% getViz()

check.gamViz(niche_comm_tax_GAM)
summary(niche_comm_tax_GAM)
print(plot(niche_comm_tax_GAM, allTerms = TRUE), pages = 1)

niche_comm_tax_GAM2 <- gam(niche.overlap.mean ~ 
                           pollen +
                           s(elev.mean, by = pollen, k = 10),
                         method = "REML",
                         family = "gaussian",
                         data = niche_overlap.tax.mean2) %>% getViz()

check.gamViz(niche_comm_tax_GAM2)
summary(niche_comm_tax_GAM2)
print(plot(niche_comm_tax_GAM2, allTerms = TRUE), pages = 1)

## Morphological niche partitioning

# Whole community
ggplot(niche_overlap.morph, aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10)) +
  facet_wrap(~year)

ggplot(niche_overlap.morph2, aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

# Whole community (collapsed to mean)
ggplot(niche_overlap.morph.mean, aes(elev.mean, niche.overlap.mean, color = pollen)) +
  geom_point(alpha = 0.25) +
  facet_wrap(~year) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

ggplot(niche_overlap.morph.mean2, aes(elev.mean, niche.overlap.mean, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

# GAM
niche_comm_morph_GAM <- gam(niche.overlap.mean ~ 
                              year +
                              pollen +
                              #s(sqrt.bb.abund, k = 4) +
                              s(elev.mean, by = year, k = 10),
                            method = "REML",
                            family = "gaussian",
                            data = niche_overlap.morph.mean) %>% getViz()

check.gamViz(niche_comm_morph_GAM)
summary(niche_comm_morph_GAM)
print(plot(niche_comm_morph_GAM, allTerms = TRUE), pages = 1)

niche_comm_morph_GAM2 <- gam(niche.overlap.mean ~ 
                              pollen +
                              s(elev.mean, by = pollen, k = 10),
                            method = "REML",
                            family = "gaussian",
                            data = niche_overlap.morph.mean2) %>% getViz()

check.gamViz(niche_comm_morph_GAM2)
summary(niche_comm_morph_GAM2)
print(plot(niche_comm_morph_GAM2, allTerms = TRUE), pages = 1)
```

We find no support for H1 at the community-mean level. But perhaps this conceals a diversity of responses to elevation in **individual species pairs**. Let's talk about that.

*Results*
These models are hard to fit nicely due to the sparcity of data and high noise at the level of individual species pairs, but what we see here does not appear to add much to what we learned from the community mean model. Individual species pairs tend to exhibit either unimodal or linear decreasing responses, some of which are statistically significant, some of which are.

```{r}
## Taxonomic niche partitioning

# Whole community (pairwise)
ggplot(niche_overlap.tax, aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10)) +
  facet_grid(year~sp.pair) +
  ylim(c(0,1))

ggplot(niche_overlap.tax2, aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10)) +
  facet_wrap(~sp.pair) +
  ylim(c(0,1))

# GAM
niche_pairwise_tax_GAM <- gam(niche.overlap ~ 
                               s(year, bs = "re") +
                               s(pollen, bs = "re") +
                               s(elev.mean, by = sp.pair, k = 10),
                             method = "REML",
                             family = "gaussian",
                             data = niche_overlap.tax) %>% getViz()

check.gamViz(niche_pairwise_tax_GAM)
summary(niche_pairwise_tax_GAM)
print(plot(niche_pairwise_tax_GAM, allTerms = TRUE), pages = 1)

niche_pairwise_tax_GAM2 <- gam(niche.overlap ~ 
                               s(pollen, bs = "re") +
                               s(elev.mean, by = sp.pair, k = 10),
                             method = "REML",
                             family = "gaussian",
                             data = niche_overlap.tax2) %>% getViz()

check.gamViz(niche_pairwise_tax_GAM2)
summary(niche_pairwise_tax_GAM2)
print(plot(niche_pairwise_tax_GAM2, allTerms = TRUE), pages = 1)

## Morphological niche partitioning

# Whole community (pairwise)
ggplot(niche_overlap.morph, aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10)) +
  facet_grid(year~sp.pair) +
  ylim(c(0,1))

ggplot(niche_overlap.morph2, aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10)) +
  facet_wrap(~sp.pair) +
  ylim(c(0,1))

# GAM
niche_pairwise_morph_GAM <- gam(niche.overlap ~ 
                               s(year, bs = "re") +
                               s(pollen, bs = "re") +
                               s(elev.mean, by = sp.pair, k = 10),
                             method = "REML",
                             family = "gaussian",
                             data = niche_overlap.morph) %>% getViz()

check.gamViz(niche_pairwise_morph_GAM)
summary(niche_pairwise_morph_GAM)
print(plot(niche_pairwise_morph_GAM, allTerms = TRUE), pages = 1)

niche_pairwise_morph_GAM2 <- gam(niche.overlap ~ 
                               s(pollen, bs = "re") +
                               s(elev.mean, by = sp.pair, k = 10),
                             method = "REML",
                             family = "gaussian",
                             data = niche_overlap.morph2) %>% getViz()

check.gamViz(niche_pairwise_morph_GAM2)
summary(niche_pairwise_morph_GAM2)
print(plot(niche_pairwise_morph_GAM2, allTerms = TRUE), pages = 1)
```

### 2.1.2 Null-model residual analysis

Arguably the most rigorous way to tease out the effect of elevation while controlling for community composition (abundance and richness patterns) is to compare the observed niche overlap at each site to that predicted by a random null model, then see how the residuals vary with elevation.

I'll be the first to admit that I'm new to null models, so I want to be extremely clear about the logic of how I'm setting this up. Step-by-step:  
1. Generate community matrices by site, pollen/nectar, and year
2. Permute each matrix *n* times by randomly redistributing the visits of each bumble bee species while holding the total visits for each bumble bee species constant. This is, therefore, a random selection model constrained to the set of flora visited for a given site-pollen-year matrix. Importantly, I am opting not to hold the floral marginal totals constant, i.e. the total number of visits that each plant receives will vary with each permutation. My gut feeling is that this is the right place to start, but this is a decision that requires scrutiny.
3. Calculate niche overlap for each permuted matrix and store as tidy arrays; this mirrors the workflow for calculating empirical niche overlap, the only difference being that I had to deal with some tricky nested lists; I'm pretty sure I did this correctly.
4. Take the mean niche overlap for each permuted community.
5. Join null model mean niche overlap table to empirical niche overlap table, joining by site, year, and pollen. Thus, for each site-year-pollen combo in the empirical niche overlap table, we have *n* iterations of random niche overlap.
6. Calculate the difference between empirical niche overlap and null niche overlap for each iteration; we'll call this "residual niche overlap".
7. Calculate mean residual niche overlap for each site-year-pollen combo. This can then be regressed against elevation to see how the deviation from null expectations varies along the elevation gradient, with the understanding that this represents the "pure" influence of elevation on niche overlap whle controlling for community differences. We could also use the raw differences without collapsing them to a mean. But that's not what I did this first time through.

There are surely a lot of other ways a null model could be constructed and analyzed, but this seems to me a good place to start.

*Results*
Encouragingly, the results of the null model look very similar to those of the "face-value" analysis. The only difference that stands out is that there is no longer a significant difference between pollen and nectar visits. Perhaps that was an artifact of abundance? In any case, our conclusions regarding H1 are the same.

```{r message=FALSE, warning=FALSE}
### Define function that will be used inside map() calls
df_to_cmat <- function(x, criterion) {
  pivot_wider(x, names_from = criterion, values_from = freq) %>%
    replace(is.na(.), 0)
}

permat_set <- function(x){
  x %>%
    column_to_rownames("bb.sp") %>%
    vegan::permatfull(fixedmar = "rows", shuffle = "ind", times = 50)
}

ret <- function(x){
  return(x$perm)
}

tidy_dist <- function(x) {
  x %>%
    tidy() %>%
    mutate(index = seq_along())
}

### Generate random niche overlap data
#### Need to look into th the missing value warnings I'm getting from the vegdist() call; I think it's because some plants end up with zero visits in some permutations.
random_niche.tax <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
  filter(!is.na(pollen)) %>% # drop observations with no pollen/nectar classification
  select(site, pollen, year, bb.sp, plant.sp) %>% 
  group_by(site, pollen, year, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
  group_by(site, pollen, year) %>%
  nest() %>% # nest data into list-column grouped by pollen, site, date
  mutate(data = map(data, 
                    function(x) pivot_wider(x, 
                                            names_from = plant.sp, 
                                            values_from = freq) %>% replace(is.na(.), 0)), # turn table into community matrix
         data = map(data, permat_set), # permute community matrices
         data = map(data, ret)) %>% # extract permutations from permat objects
  unnest(data) %>% # unnest
  mutate(data = map(data, function(x) vegdist(x, method = "horn")), # convert each permutation into distance matrix
         data = map(data, tidy)) %>% # convert distance matrices to tidy arrays
  group_by(site, pollen, year) %>%
  mutate(iter = row_number()) %>% # add index to denote iterations
  unnest(cols = data) %>% # unnest again
  rename(bb.sp1 = item1, bb.sp2 = item2) %>%
  unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
  mutate(sp.pair = factor(sp.pair), # ...and make it a factor so that mgcv can handle it
         niche.overlap = 1-distance) %>% # express niche overlap so that 1 = perfect overlap and 0 = no overlap
  select(site, pollen, year, sp.pair, iter, niche.overlap)

random_niche.tax.mean <- random_niche.tax %>%
  group_by(site, pollen, year, iter) %>%
  summarize(random.niche.overlap.mean = mean(niche.overlap))

random_niche.tax2 <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
  filter(!is.na(pollen)) %>% # drop observations with no pollen/nectar classification
  select(site, pollen, bb.sp, plant.sp) %>% 
  group_by(site, pollen, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
  group_by(site, pollen) %>%
  nest() %>% # nest data into list-column grouped by pollen, site, date
  mutate(data = map(data, df_to_cmat.sp), # turn table into community matrix
         data = map(data, permat_set), # permute community matrices
         data = map(data, ret)) %>% # extract permutations from permat objects
  unnest(data) %>% # unnest
  mutate(data = map(data, function(x) vegdist(x, method = "horn")), # convert each permutation into distance matrix
         data = map(data, tidy)) %>% # convert distance matrices to tidy arrays
  group_by(site, pollen) %>%
  mutate(iter = row_number()) %>% # add index to denote iterations
  unnest(cols = data) %>% # unnest again
  rename(bb.sp1 = item1, bb.sp2 = item2) %>%
  unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
  mutate(sp.pair = factor(sp.pair), # ...and make it a factor so that mgcv can handle it
         niche.overlap = 1-distance) %>% # express niche overlap so that 1 = perfect overlap and 0 = no overlap
  select(site, pollen, sp.pair, iter, niche.overlap)

random_niche.sp.mean <- random_niche %>%
  group_by(site, pollen, year, iter) %>%
  summarize(random.niche.overlap.mean = mean(niche.overlap))

random_niche.sp.mean2 <- random_niche2 %>%
  group_by(site, pollen, iter) %>%
  summarize(random.niche.overlap.mean = mean(niche.overlap))

random_niche.ktype <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
  filter(!is.na(pollen)) %>% # drop observations with no pollen/nectar classification
  select(site, pollen, year, bb.sp, k.type.s) %>% 
  group_by(site, pollen, year, bb.sp, k.type.s) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
  group_by(site, pollen, year) %>%
  nest() %>% # nest data into list-column grouped by pollen, site, date
  mutate(data = map(data, df_to_cmat.ktype), # turn table into community matrix
         data = map(data, permat_set), # permute community matrices
         data = map(data, ret)) %>% # extract permutations from permat objects
  unnest(data) %>% # unnest
  mutate(data = map(data, function(x) vegdist(x, method = "horn")), # convert each permutation into distance matrix
         data = map(data, tidy)) %>% # convert distance matrices to tidy arrays
  group_by(site, pollen, year) %>%
  mutate(iter = row_number()) %>% # add index to denote iterations
  unnest(cols = data) %>% # unnest again
  rename(bb.sp1 = item1, bb.sp2 = item2) %>%
  unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
  mutate(sp.pair = factor(sp.pair), # ...and make it a factor so that mgcv can handle it
         niche.overlap = 1-distance) %>% # express niche overlap so that 1 = perfect overlap and 0 = no overlap
  select(site, pollen, year, sp.pair, iter, niche.overlap)

random_niche.ktype2 <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
  filter(!is.na(pollen)) %>% # drop observations with no pollen/nectar classification
  select(site, pollen, bb.sp, k.type.s) %>% 
  group_by(site, pollen, bb.sp, k.type.s) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
  group_by(site, pollen) %>%
  nest() %>% # nest data into list-column grouped by pollen, site, date
  mutate(data = map(data, df_to_cmat.ktype), # turn table into community matrix
         data = map(data, permat_set), # permute community matrices
         data = map(data, ret)) %>% # extract permutations from permat objects
  unnest(data) %>% # unnest
  mutate(data = map(data, function(x) vegdist(x, method = "horn")), # convert each permutation into distance matrix
         data = map(data, tidy)) %>% # convert distance matrices to tidy arrays
  group_by(site, pollen) %>%
  mutate(iter = row_number()) %>% # add index to denote iterations
  unnest(cols = data) %>% # unnest again
  rename(bb.sp1 = item1, bb.sp2 = item2) %>%
  unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
  mutate(sp.pair = factor(sp.pair), # ...and make it a factor so that mgcv can handle it
         niche.overlap = 1-distance) %>% # express niche overlap so that 1 = perfect overlap and 0 = no overlap
  select(site, pollen, sp.pair, iter, niche.overlap)

random_niche.ktype.mean <- random_niche %>%
  group_by(site, pollen, year, iter) %>%
  summarize(random.niche.overlap.mean = mean(niche.overlap))

random_niche.ktype.mean2 <- random_niche2 %>%
  group_by(site, pollen, iter) %>%
  summarize(random.niche.overlap.mean = mean(niche.overlap))

## Join null model to empirical niche overlap
# Taxonomic niche overlap
comparison.sp <- niche_overlap.sp.mean %>%
  left_join(random_niche.sp.mean) %>%
  mutate(niche.overlap.resid = niche.overlap.mean - random.niche.overlap.mean) %>%
  group_by(site, year, pollen, elev.mean, niche.overlap.mean) %>%
  summarize(mean.resid = mean(niche.overlap.resid))

comparison.sp2 <- niche_overlap.sp.mean2 %>%
  left_join(random_niche.sp.mean2) %>%
  mutate(niche.overlap.resid = niche.overlap.mean - random.niche.overlap.mean) %>%
  group_by(site, year, pollen, elev.mean, niche.overlap.mean) %>%
  summarize(mean.resid = mean(niche.overlap.resid))

ggplot(comparison.sp, aes(elev.mean, mean.resid, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~ year)

ggplot(comparison.sp2, aes(elev.mean, mean.resid, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~ year)

## GAM
null_resid_sp_GAM <- gam(mean.resid ~ 
                           year +
                           pollen +
                           s(elev.mean, by = year, k = 10),
                         method = "REML",
                         family = "gaussian",
                         data = comparison.sp) %>% getViz()

check.gamViz(null_resid_sp_GAM)
summary(null_resid_sp_GAM)
print(plot(null_resid_sp_GAM, allTerms = TRUE), pages = 1)

null_resid_sp_GAM2 <- gam(mean.resid ~ 
                           pollen +
                           s(elev.mean, by = pollen, k = 10),
                         method = "REML",
                         family = "gaussian",
                         data = comparison.sp2) %>% getViz()

check.gamViz(null_resid_sp_GAM2)
summary(null_resid_sp_GAM2)
print(plot(null_resid_sp_GAM2, allTerms = TRUE), pages = 1)

# Morphological niche overlap
comparison.ktype <- niche_overlap.ktype.mean %>%
  left_join(random_niche.ktype.mean) %>%
  mutate(niche.overlap.resid = niche.overlap.mean - random.niche.overlap.mean) %>%
  group_by(site, year, pollen, elev.mean, niche.overlap.mean) %>%
  summarize(mean.resid = mean(niche.overlap.resid))

comparison.ktype2 <- niche_overlap.ktype.mean2 %>%
  left_join(random_niche.ktype.mean2) %>%
  mutate(niche.overlap.resid = niche.overlap.mean - random.niche.overlap.mean) %>%
  group_by(site, year, pollen, elev.mean, niche.overlap.mean) %>%
  summarize(mean.resid = mean(niche.overlap.resid))

ggplot(comparison.ktype, aes(elev.mean, mean.resid, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~ year)

ggplot(comparison.ktype2, aes(elev.mean, mean.resid, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~ year)

## GAM
null_resid_ktype_GAM <- gam(mean.resid ~ 
                              year +
                              pollen +
                              s(elev.mean, by = year, k = 10),
                            method = "REML",
                            family = "gaussian",
                            data = comparison.ktype) %>% getViz()

check.gamViz(null_resid_ktype_GAM)
summary(null_resid_ktype_GAM)
print(plot(null_resid_ktype_GAM, allTerms = TRUE), pages = 1)

null_resid_ktype_GAM2 <- gam(mean.resid ~ 
                              pollen +
                              s(elev.mean, by = pollen, k = 10),
                            method = "REML",
                            family = "gaussian",
                            data = comparison.ktype2) %>% getViz()

check.gamViz(null_resid_ktype_GAM2)
summary(null_resid_ktype_GAM2)
print(plot(null_resid_ktype_GAM2, allTerms = TRUE), pages = 1)
```

### 2.1.3. Parsing niche contraction and niche separation

Contrary to the predictions of H1, niche overlap decreases sharply at high elevation. This is true for both taxonomic and morphological niche overlap, and it holds in both the face-value and null-residual approaches. This decrease in niche overlap might be explained by two modest, proximate hypotheses:

1.  **H1:** The decrease of niche overlap at high elevations is due to *niche contraction*, i.e. BB spp. visiting a smaller proportion of available (and "eligible", i.e. visited at least once by a bumble bee) flora. 
  *Prediction:* "resource range" (Poisot et al. 2012) will decrease with elevation (as Dormann notes, resource range is a "backward" metric insofar as a large resource range means a small proportional niche breadth; I will calculate the inverse of resource range so that it makes more sense)
2.  **H2:** The decrease of niche overlap at high elevations is due to *niche separation*, i.e. BB spp. avoiding one another's niches but not contracting their own.
  *Predictions:* C score will increase with elevation

*Results:* I haven't run formal stats on this analysis yet, but from the visualization, it looks like the overall story is that niche overlap is driven mainly by niche separation, not so much but niche contraction. My first impression is that this suggests a behavior response to competition.

```{r}
resource_range.sp <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
  filter(!is.na(pollen)) %>%
  group_by(site, year, pollen, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
  unite(webID, c(site, year, pollen), sep = "_") %>%
  ungroup() %>%
  dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() %>% # convert to to bipartite web
  map(specieslevel, level = "higher", index = "resource range") %>%
  map(rownames_to_column) %>%
  map(as_tibble) %>%
  bind_rows(.id = "webID") %>% # Collapse list into big data frame
  rename(bb.sp = rowname) %>%
  separate(webID, c("site", "year", "pollen"), sep =  "_") %>%
  left_join(site_data) %>%
  mutate(site = factor(site),
         year = factor(year),
         pollen = factor(pollen)) %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # for some reason, one row has a "1" for bb.sp, and I don't know where this comes from; this gets rid of it, but I would still like to know why
  mutate(resource.range = 1-resource.range) %>%
  left_join(BB_covars.splevel) %>%
  filter(bb.sp.abund > 1)

resource_range.sp.mean <- resource_range.sp %>%
  group_by(site, year, elev.mean, pollen) %>%
  summarize(mean.resource.range = mean(resource.range))

ggplot(resource_range.sp, aes(elev.mean, resource.range, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_grid(bb.sp ~ year) +
  ylim(c(0,1))

ggplot(resource_range.sp.mean, aes(elev.mean, mean.resource.range, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year) +
  ylim(c(0,1))

resource_range.ktype <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
  filter(!is.na(pollen)) %>%
  group_by(site, year, pollen, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
  unite(webID, c(site, year, pollen), sep = "_") %>%
  ungroup() %>%
  dplyr::select(higher = bb.sp, lower = k.type.s, webID, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() %>% # convert to to bipartite web
  map(specieslevel, level = "higher", index = "resource range") %>%
  map(rownames_to_column) %>%
  map(as_tibble) %>%
  bind_rows(.id = "webID") %>% # Collapse list into big data frame
  rename(bb.sp = rowname) %>%
  separate(webID, c("site", "year", "pollen"), sep =  "_") %>%
  left_join(site_data) %>%
  mutate(site = factor(site),
         year = factor(year),
         pollen = factor(pollen)) %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # for some reason, one row has a "1" for bb.sp, and I don't know where this comes from; this gets rid of it, but I would still like to know why
  mutate(resource.range = 1-resource.range) %>%
  left_join(BB_covars.splevel) %>%
  filter(bb.sp.abund > 1)

resource_range.ktype.mean <- resource_range.ktype %>%
  group_by(site, year, elev.mean, pollen) %>%
  summarize(mean.resource.range = mean(resource.range))

ggplot(resource_range.ktype, aes(elev.mean, resource.range, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_grid(bb.sp ~ year) +
  ylim(c(0,1))

ggplot(resource_range.ktype.mean, aes(elev.mean, mean.resource.range, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year) +
  ylim(c(0,1))

cscore.sp <- net %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    filter(!is.na(pollen)) %>%
    group_by(site, year, pollen, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
    summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
    group_by(site, year, pollen) %>%
    unite(webID, c(site, year, pollen), sep = "_") %>%
    ungroup() %>%
    dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
    data.frame() %>% # convert to data frame
    frame2webs() %>% # convert to to bipartite web
    map(grouplevel, level = "higher", weighted = TRUE, index = c("C score")) %>% 
    data.frame() %>%
    t() %>%
    data.frame() %>%
    rownames_to_column() %>%
    dplyr::select(webID = rowname, everything()) %>%
    as_tibble() %>%
    separate(webID, c("site", "year", "pollen"), sep =  "_") %>%
    left_join(site_data) %>%
    mutate(site = factor(site),
           year = factor(year),
           pollen = factor(pollen))

ggplot(cscore.sp, aes(elev.mean, C.score.HL, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

cscore.ktype <- net %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    filter(!is.na(pollen)) %>%
    group_by(site, year, pollen, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
    summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
    group_by(site, year, pollen) %>%
    unite(webID, c(site, year, pollen), sep = "_") %>%
    ungroup() %>%
    dplyr::select(higher = bb.sp, lower = k.type.s, webID, freq) %>% # rename columns to match bipartite's expectations
    data.frame() %>% # convert to data frame
    frame2webs() %>% # convert to to bipartite web
    map(grouplevel, level = "higher", weighted = TRUE, index = c("C score")) %>% 
    data.frame() %>%
    t() %>%
    data.frame() %>%
    rownames_to_column() %>%
    dplyr::select(webID = rowname, everything()) %>%
    as_tibble() %>%
    separate(webID, c("site", "year", "pollen"), sep =  "_") %>%
    left_join(site_data) %>%
    mutate(site = factor(site),
           year = factor(year),
           pollen = factor(pollen))

ggplot(cscore.ktype, aes(elev.mean, C.score.HL, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

### 2020-11-26 analyses/old_scripts/niche_relations_v2.Rmd
did so many things

---
title: "site-wise niche-overlap by elevation"
output: html_document
---

Load packages 
```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
library(bipartite)
library(mgcv)
library(mgcViz)
library(ggbipart)
library(GGally)
library(vegan)
library(ggvegan)
library(ggrepel)
library(gridExtra)
library(tidyverse)
library(lubridate)
library(patchwork)
library(tidymodels)
```

Load data
```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
# Site data
site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, 
                   elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, 
                             levels = c("unten", "mitte", "oben")),
         elev.class2 = factor(elev.class2, 
                             levels = c("low", "mid", "high"))) # turn elav.class into an ordered factor

# BB traits
bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  mutate(pbl.fac = factor(if_else(bb.sp %in% c("hort", "pasc", "wurf"), "long", "short")))

# FL traits
fl_tax <- read_csv("../processed_data/floral_tax.csv")

fl_ktype <- read_csv("../processed_data/floral_k_type.csv")

fl_morph <- read_csv("../processed_data/flower_morphology_GitaBenadi.csv") %>%
  dplyr::select(SampleID, plant.family = Family, plant.sp = Species, 
         Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam),
            LN = mean(LN))

fl_ktype_key <- read_delim("../processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  dplyr::select(k.type = 1, description = 2)

fl_traits <- fl_tax %>%
  full_join(fl_ktype) %>%
  full_join(fl_morph) %>%
  dplyr::select(plant.sp, plant.genus, plant.family, 
         k.type, k.type.s, k.type.ss, diam, LN)

# Network
net <- read_csv("../processed_data/network.csv") %>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  mutate(year = factor(year)) %>% # convert year to factor
  left_join(site_data, by = "site") %>%
  #left_join(gdd, by = c("site", "date")) %>%
  ungroup() %>%
  arrange(site, date) %>%
  left_join(fl_traits) %>%

  # Fill in missing k.type scores
  mutate(k.type.s = if_else(plant.genus == "Phyteuma", 7.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Phyteuma", 7, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Aconitum", 5.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Aconitum", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Cardamine", 1.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Cardamine", 1, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Euphrasia", 5.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Euphrasia", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Lamium", 5.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Lamium", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Mentha", 2.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Mentha", 2, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Rumex", 0.0, k.type.s),
         k.type.ss = if_else(plant.genus == "Rumex", 0, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Salix", 9.0, k.type.s),
         k.type.ss = if_else(plant.genus == "Salix", 9, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Silene", 4.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Silene", 4, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Stachys", 5.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Stachys", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Taraxacum", 7.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Taraxacum", 7, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.sp == "Gentiana aspera", 2.1, k.type.s),
         k.type.ss = if_else(plant.sp == "Gentiana aspera", 2, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.sp == "Gentiana ciliata", 4.1, k.type.s),
         k.type.ss = if_else(plant.sp == "Gentiana ciliata", 4, k.type.ss))

# Floral survey
survey <- read_csv("../processed_data/floral_survey.csv") %>%
  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  left_join(fl_traits) %>%

  # Fill in missing k.type scores
  mutate(k.type.s = if_else(plant.genus == "Phyteuma", 7.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Phyteuma", 7, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Aconitum", 5.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Aconitum", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Cardamine", 1.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Cardamine", 1, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Euphrasia", 5.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Euphrasia", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Lamium", 5.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Lamium", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Mentha", 2.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Mentha", 2, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Rumex", 0.0, k.type.s),
         k.type.ss = if_else(plant.genus == "Rumex", 0, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Salix", 9.0, k.type.s),
         k.type.ss = if_else(plant.genus == "Salix", 9, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Silene", 4.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Silene", 4, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Stachys", 5.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Stachys", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Taraxacum", 7.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Taraxacum", 7, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.sp == "Gentiana aspera", 2.1, k.type.s),
         k.type.ss = if_else(plant.sp == "Gentiana aspera", 2, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.sp == "Gentiana ciliata", 4.1, k.type.s),
         k.type.ss = if_else(plant.sp == "Gentiana ciliata", 4, k.type.ss))

### Auxiliary data, i.e. all the covariates we need to control for
BB_covars <- net %>%
  group_by(site, year) %>%
  summarize(bb.rich = length(unique(bb.sp)),
            sqrt.bb.abund = sqrt(n()))

FL_covars <- survey %>%
  mutate(year = factor(year(date))) %>%
  filter(flower.cover > 0) %>%
  group_by(site, year) %>%
  summarize(fl.rich = length(unique(plant.sp)),
            log.fl.abund = log(sum(flower.cover)))

### Species-level by site (we will also separate pollen and nectar foragers)
BB_covars.splevel <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  group_by(bb.sp, site, year, pollen) %>%
  summarize(bb.sp.abund = n())

### Species names and such for plotting
species_names <- cbind(c("bss", "pasc", "prat", 
                         "soro", "wurf", "hort", 
                         "pyre", "mend", "mont", 
                         "gers", "hypn","jone", 
                         "lapi", "muci", "psit", 
                         "humi"),
                       c("B. sensu-strictu", "B. pascuorum", "B. pratorum", 
                         "B. soroensis", "B. wurflenii", "B. hortorum",
                         "B. pyrenaeus", "B. mendax", "B. monticola",
                         "B. gerstaeckeri", "B. hypnorum", "B. jonellus", 
                         "B. lapidarius", "B. mucidus", "B. psithyrus", 
                         "B. humilis"),
                       c("B. ss.", "B. pasc.", "B. prat.", 
                         "B. soro.", "B. wurf.", "B. hort.",
                         "B. pyre.", "B. mend.", "B. mont.",
                         "B. gers.", "B. hypn.", "B. jone.", 
                         "B. lapi.", "B. muci.", "B. psit.", 
                         "B. humi"),
                       c("major", "major", "major", 
                         "major", "major", "common", 
                         "high elev.", "high elev.", "high elev.",
                         "minor", "minor", "common", 
                         "minor", "minor", "common", 
                         "minor")
                       ) %>%
  as.data.frame() %>%
  as_tibble() %>%
  dplyr::select(bb.sp = 1, species.name = 2, species.abb = 3, abund.class = 4) %>%
  mutate(focal = if_else(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf"), TRUE, FALSE))
```

We will focus on six species that are abundant and widespread:  
  1. *B. sensu-strictu*
  2. *B. pascuorum*
  3. *B. pratorum*
  4. *B. hortorum*
  5. *B. soroensis*
  6. *B. wurflenii*

```{r fig.cap="**Figure 1: Focal species**", message=FALSE, warning=FALSE, echo=FALSE}
bb_range.year <- net %>%
  left_join(site_data) %>%
  #group_by(site, date, elev.mean, bb.sp, year) %>%
  group_by(site, elev.mean, bb.sp, year) %>%
  summarize(abund = n()) %>%
  #group_by(site, elev.mean, bb.sp, year) %>%
  #summarize(abund = mean(abund)) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2)) %>%
  left_join(species_names)

ggplot(bb_range.year, aes(reorder(bb.sp, elev.med), elev.mean, color = focal)) +
  geom_line(size = 2, alpha = 0.25) +
  geom_point(aes(size = abund), alpha = 0.5) +
  facet_wrap(~year, ncol = 1) +
  theme_light(12) +
  scale_size_continuous(name = "Abundance") +
  scale_color_discrete(name = "Focal species") +
  xlab("BB species") +
  ylab("Elevation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("./output/bb_range.pdf", height = 6, width = 6)

bb_range <- net %>%
  left_join(site_data) %>%
  group_by(site, elev.mean, bb.sp, year) %>%
  summarize(abund = n()) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2)) %>%
  left_join(species_names)

ggplot(bb_range, aes(reorder(bb.sp, elev.med), elev.mean, color = focal)) +
  geom_line(size = 2, alpha = 0.25) +
  geom_point(aes(size = abund), alpha = 0.5) +
  theme_light(12) +
  scale_size_continuous(name = "Abundance") +
  scale_color_discrete(name = "Focal species") +
  xlab("BB species") +
  ylab("Elevation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Functions
```{r, echo = FALSE}
# Add BB trait data for envfit() call inside map2()
trait_join <- function(x, y = bb_traits) {
  x %>%
    select(bb.sp) %>%
    left_join(y) %>%
    column_to_rownames(var = "bb.sp") %>%
    select(pbl.w)
}

# Fortify envfit() object fitted to BB tongue length data
envfit_fort <- function(x){
  fortify(x) %>%
  select(Score = Type, Label, MDS1, MDS2) %>%
  mutate(Label = case_when(Label == "pbl.cat" ~ "T.class"),
         Score = case_when(Score == "Vector" ~ "T.class"))
}

# Fortify PCOA 
pcoa_fort <- function(x){
  fortify(x, axes = 1:2) %>%
  mutate(Score = case_when(
    Score == "species" ~ "FL.sp",
    Score == "sites" ~ "BB.sp"))
}

# Plot PCOA; to avoid overplotting, loadings < min_load are not plotted
pcoa_plot <- function(x, min_load = 0.25) {

  t.class <- x %>%
    filter(Score == "BB.sp") %>%
    mutate(T.class = factor(if_else(Label %in% c("hort", "pasc", "wurf"), "long", "short")))

  ggplot(filter(x, (abs(MDS1) > min_load | abs(MDS2) > min_load) | (Score %in% c("BB.sp", "T.len")))) + # to avoid overplotting, we can filter to include only floral species > loading threshold
    geom_polygon(data = t.class, aes(MDS1, MDS2, linetype = T.class), 
                 color = "gray70", fill = NA, size = 1.25, key_glyph = "abline") +
    geom_point(aes(MDS1, MDS2, color = Score), size = 3, alpha = 0.75) +
    geom_text_repel(aes(MDS1, MDS2, color = Score, label = Label), key_glyph = "rect") +
    theme_classic()
}

# Niche analysis function; generates PCOA, envfit, ordination plot, and tongue length R2 coming out of envfit
niche_analysis1 <- function(x, y, criterion) {

  criterion <- enquo(criterion)

  x %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
    filter(!is.na(pollen)) %>%
    group_by(bb.sp, !!criterion, pollen) %>%
    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site
    group_by(bb.sp, pollen) %>%
    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair per site-date
    dplyr::select(-freq) %>%
    group_by(pollen) %>%
    nest() %>%
    mutate(comm1 = map(data, function(x) pivot_wider(x, names_from = !!criterion, values_from = prop)), # spread into the kind of array vegan likes
           comm2 = map(comm1, function(x) replace(x, is.na(x), 0)),
           comm3 = map(comm2, function(x) column_to_rownames(x, var = "bb.sp")),
           pcoa = map(comm3,  function(x) capscale(x ~ 1, distance = "horn")),
           traits = map(comm2, function(x) x %>% 
                          left_join(y) %>% 
                          column_to_rownames(var = "bb.sp") %>% 
                          select(pbl.fac)),
           adonis = map2(comm3, traits, function(x, y) adonis(x ~ pbl.fac, data = y, method = "horn")),
           pcoa.fort = map(pcoa, pcoa_fort),
           pcoa.plot = map(pcoa.fort, pcoa_plot),
           tongue.r2 = map(adonis, function(x) x$aov.tab$R2[[1]]),
           tongue.p = map(adonis, function(x) x$aov.tab$'Pr(>F)'[[1]]),
           MDS1.varexp = map(pcoa, function(x) x$CA$eig[[1]]),
           MDS2.varexp = map(pcoa, function(x) x$CA$eig[[2]])) %>%
    unnest(cols = c(tongue.r2, tongue.p, MDS1.varexp, MDS2.varexp)) %>%
    select(pollen, comm2, comm3, traits, pcoa, pcoa.fort, adonis,
           pcoa.plot, tongue.r2, tongue.p, MDS1.varexp, MDS2.varexp) 
}

# This version break it down by elev.class
niche_analysis2 <- function(x, y, criterion) {

  criterion <- enquo(criterion)

  x %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
    filter(!is.na(pollen)) %>%
    group_by(elev.class, bb.sp, !!criterion, pollen) %>%
    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site
    group_by(elev.class, bb.sp, pollen) %>%
    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair per site-date
    dplyr::select(-freq) %>%
    group_by(pollen, elev.class) %>%
    nest() %>%
    mutate(comm1 = map(data, function(x) pivot_wider(x, names_from = !!criterion, values_from = prop)), # spread into the kind of array vegan likes
           comm2 = map(comm1, function(x) replace(x, is.na(x), 0)),
           comm3 = map(comm2, function(x) column_to_rownames(x, var = "bb.sp")),
           pcoa = map(comm3,  function(x) capscale(x ~ 1, distance = "horn")),
           traits = map(comm2, function(x) x %>% 
                          left_join(y) %>% 
                          column_to_rownames(var = "bb.sp") %>% 
                          select(pbl.fac)),
           adonis = map2(comm3, traits, function(x, y) adonis(x ~ pbl.fac, data = y, method = "horn")),
           pcoa.fort = map(pcoa, pcoa_fort),
           pcoa.plot = map(pcoa.fort, pcoa_plot),
           tongue.r2 = map(adonis, function(x) x$aov.tab$R2[[1]]),
           tongue.p = map(adonis, function(x) x$aov.tab$'Pr(>F)'[[1]]),
           MDS1.varexp = map(pcoa, function(x) x$CA$eig[[1]]),
           MDS2.varexp = map(pcoa, function(x) x$CA$eig[[2]])) %>%
    unnest(cols = c(tongue.r2, tongue.p, MDS1.varexp, MDS2.varexp)) %>%
    select(pollen, elev.class, comm2, comm3, traits, pcoa, pcoa.fort, adonis,
           pcoa.plot, tongue.r2, tongue.p, MDS1.varexp, MDS2.varexp) 
}

# ...and this version breaks it down by elev.class and period
niche_analysis3 <- function(x, y, criterion) {

  criterion <- enquo(criterion)

  x %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
    filter(!is.na(pollen)) %>%
    group_by(elev.class) %>%
    mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
    group_by(elev.class, period, pollen, bb.sp, !!criterion) %>%
    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site
    group_by(elev.class, period, bb.sp, pollen) %>%
    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair per site-date
    dplyr::select(-freq) %>%
    group_by(pollen, elev.class, period) %>%
    nest() %>%
    mutate(comm1 = map(data, function(x) pivot_wider(x, names_from = !!criterion, values_from = prop)), # spread into the kind of array vegan likes
           comm2 = map(comm1, function(x) replace(x, is.na(x), 0)),
           comm3 = map(comm2, function(x) column_to_rownames(x, var = "bb.sp")),
           pcoa = map(comm3,  function(x) capscale(x ~ 1, distance = "horn")),
           traits = map(comm2, function(x) x %>% 
                          left_join(y) %>% 
                          column_to_rownames(var = "bb.sp") %>% 
                          select(pbl.fac)),
           adonis = map2(comm3, traits, function(x, y) adonis(x ~ pbl.fac, data = y, method = "horn")),
           pcoa.fort = map(pcoa, pcoa_fort),
           pcoa.plot = map(pcoa.fort, pcoa_plot),
           tongue.r2 = map(adonis, function(x) x$aov.tab$R2[[1]]),
           tongue.p = map(adonis, function(x) x$aov.tab$'Pr(>F)'[[1]]),
           MDS1.varexp = map(pcoa, function(x) x$CA$eig[[1]]),
           MDS2.varexp = map(pcoa, function(x) x$CA$eig[[2]])) %>%
    unnest(cols = c(tongue.r2, tongue.p, MDS1.varexp, MDS2.varexp)) %>%
    select(pollen, elev.class, period, comm2, comm3, traits, pcoa, pcoa.fort, adonis,
           pcoa.plot, tongue.r2, tongue.p, MDS1.varexp, MDS2.varexp) 
}

beta_func <- function(x, y, level, criterion, binary) {

  criterion <- enquo(criterion)
  level <- enquo(level)

  x %>% 
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    left_join(bb_traits) %>%
    #filter(!is.na(pollen)) %>%
    #group_by(pollen, site, !!level, !!criterion) %>%
    group_by(site, !!level, !!criterion) %>%
    summarize(freq = n()) %>%
    #group_by(pollen, site, !!level) %>%
    group_by(site, !!level) %>%
    mutate(prop = freq/sum(freq)) %>%
    #group_by(pollen, !!level) %>%
    group_by(!!level) %>%
    select(-freq) %>%
    nest() %>%
    mutate(comm = map(data, function(x) pivot_wider(x, names_from = !!criterion, values_from = prop, values_fill = 0)),
           comm2 = map(comm, function(x) column_to_rownames(x, var = "site")),
           dist = map(comm2, function(x) vegdist(x, method = "horn", binary = binary)),
           tidy.dist = map(dist, tidy),
           elevation = map(comm, function(x) left_join(x, y, by = "site") %>% 
                             select(site, elev.mean) %>% 
                             column_to_rownames(var = "site")),
           adonis = map2(comm2, elevation, function(x, y) adonis(x ~ elev.mean, data = y)),
           elev.r2 = map(adonis, function(x) x$aov.tab$R2[[1]]),
           elev.p = map(adonis, function(x) x$aov.tab$'Pr(>F)'[[1]])) %>%
    unnest(c(tidy.dist, elev.r2, elev.p)) %>%
    #select(!!level, pollen, site1 = item1, site2 = item2, distance, elev.r2, elev.p) %>%
    select(!!level, site1 = item1, site2 = item2, distance, elev.r2, elev.p) %>%
    left_join(select(site_data, site, elev.mean), by = c("site1" = "site")) %>%
    rename(elev1 = elev.mean) %>%
    left_join(select(site_data, site, elev.mean), by = c("site2" = "site")) %>%
    rename(elev2 = elev.mean) %>%
    mutate(elev.diff = abs(elev1 - elev2)) 
}

# Niche overlap by sp.pair * site * year
niche_overlapper <- function(x, criterion) {

  criterion <- enquo(criterion) # criterion is set to either plant.sp (for taxonomic) or k.type.s (for morphological)

  x %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    filter(!is.na(pollen)) %>% # drop observations with no pollen/nectar classification
    group_by(bb.sp, pollen, !!criterion, site, year) %>% # group by site-year and BB-FL pair
    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-year
    group_by(bb.sp, pollen, site, year) %>% # group by site-year and BB species
    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
    dplyr::select(-freq) %>% # drop freq variable
    pivot_wider(names_from = !!criterion, values_from = prop) %>% # spread into the kind of array vegan likes
    replace(is.na(.), 0) %>% # replace NAs with zeros
    group_by(pollen, site, year) %>%
    nest() %>% # nest into list-column
    mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
           dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
           tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
    dplyr::select(pollen, site, year, tidy_dist) %>% # drop intermediate list-columns
    unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
    left_join(site_data) %>% # add site data
    dplyr::select(site, year, pollen, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
    left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
    rename(subgenus1 = subgenus, pbl.w1 = pbl.w, pbl.w.class1 = pbl.w.class, pbl.fac1 = pbl.fac) %>%
    left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
    rename(subgenus2 = subgenus, pbl.w2 = pbl.w, pbl.w.class2 = pbl.w.class, pbl.fac2 = pbl.fac) %>%
    dplyr::select(site, year, pollen, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
           pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, pbl.fac1, pbl.fac2, elev.mean) %>%
    left_join(BB_covars.splevel, by = c("bb.sp1" = "bb.sp", "site", "year", "pollen")) %>%
    rename(bb.sp1.abund = bb.sp.abund) %>%
    dplyr::select(site, year, pollen, bb.sp1, bb.sp2, bb.sp1.abund, subgenus1, subgenus2, distance, 
           pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, pbl.fac1, pbl.fac2, elev.mean) %>%
    left_join(BB_covars.splevel, by = c("bb.sp2" = "bb.sp", "site", "year", "pollen")) %>%
    rename(bb.sp2.abund = bb.sp.abund) %>%
    dplyr::select(site, year, pollen, bb.sp1, bb.sp2, bb.sp1.abund, bb.sp2.abund, subgenus1, subgenus2, distance, 
           pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, pbl.fac1, pbl.fac2, elev.mean) %>%
    mutate(pbl.diff = abs(pbl.w1 - pbl.w2), # tongue length differences
           pbl.fac.diff = if_else(pbl.fac1 == pbl.fac2, "same", "different"),
           abund.diff = abs(bb.sp1.abund - bb.sp2.abund),
           subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different"), # binary: same or different subgenus

           # Logical columns indicating whether a given species belongs to a given species pair
           # This is handy for extracting data for individual species across all their counterparts
           bss = if_else(bb.sp1 == "bss" | bb.sp2 == "bss", TRUE, FALSE),
           pasc = if_else(bb.sp1 == "pasc" | bb.sp2 == "pasc", TRUE, FALSE),
           wurf = if_else(bb.sp1 == "wurf" | bb.sp2 == "wurf", TRUE, FALSE),
           prat = if_else(bb.sp1 == "prat" | bb.sp2 == "prat", TRUE, FALSE),
           soro = if_else(bb.sp1 == "soro" | bb.sp2 == "soro", TRUE, FALSE),
           hort = if_else(bb.sp1 == "hort" | bb.sp2 == "hort", TRUE, FALSE)) %>%

    unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
    mutate(sp.pair = factor(sp.pair),
           pollen = factor(pollen)) %>% # ...and make it a factor so that mgcv can handle it
    left_join(BB_covars) %>% # add BB covars
    left_join(FL_covars) %>% # add floral covars
    mutate(niche.overlap = 1-distance) # express niche overlap so that 1 = perfect overlap and 0 = no overlap
}

niche_overlapper2 <- function(x, criterion) {

  criterion <- enquo(criterion) # criterion is set to either plant.sp (for taxonomic) or k.type.s (for morphological)

  x %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    filter(!is.na(pollen)) %>% # drop observations with no pollen/nectar classification
    group_by(bb.sp, pollen, !!criterion, site) %>% # group by site-year and BB-FL pair
    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-year
    group_by(bb.sp, pollen, site) %>% # group by site-year and BB species
    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
    dplyr::select(-freq) %>% # drop freq variable
    pivot_wider(names_from = !!criterion, values_from = prop) %>% # spread into the kind of array vegan likes
    replace(is.na(.), 0) %>% # replace NAs with zeros
    group_by(pollen, site) %>%
    nest() %>% # nest into list-column
    mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
           dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
           tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
    dplyr::select(pollen, site, tidy_dist) %>% # drop intermediate list-columns
    unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
    left_join(site_data) %>% # add site data
    dplyr::select(site, pollen, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
    left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
    rename(subgenus1 = subgenus, pbl.w1 = pbl.w, pbl.w.class1 = pbl.w.class, pbl.fac1 = pbl.fac) %>%
    left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
    rename(subgenus2 = subgenus, pbl.w2 = pbl.w, pbl.w.class2 = pbl.w.class, pbl.fac2 = pbl.fac) %>%
    dplyr::select(site, pollen, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
           pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, pbl.fac1, pbl.fac2, elev.mean) %>%
    mutate(pbl.diff = abs(pbl.w1 - pbl.w2), # tongue length differences
           pbl.fac.diff = if_else(pbl.fac1 == pbl.fac2, "same", "different"),
           subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different"), # binary: same or different subgenus

           # Logical columns indicating whether a given species belongs to a given species pair
           # This is handy for extracting data for individual species across all their counterparts
           bss = if_else(bb.sp1 == "bss" | bb.sp2 == "bss", TRUE, FALSE),
           pasc = if_else(bb.sp1 == "pasc" | bb.sp2 == "pasc", TRUE, FALSE),
           wurf = if_else(bb.sp1 == "wurf" | bb.sp2 == "wurf", TRUE, FALSE),
           prat = if_else(bb.sp1 == "prat" | bb.sp2 == "prat", TRUE, FALSE),
           soro = if_else(bb.sp1 == "soro" | bb.sp2 == "soro", TRUE, FALSE),
           hort = if_else(bb.sp1 == "hort" | bb.sp2 == "hort", TRUE, FALSE)) %>%

    unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
    mutate(sp.pair = factor(sp.pair),
           pollen = factor(pollen)) %>% # ...and make it a factor so that mgcv can handle it
    left_join(BB_covars) %>% # add BB covars
    left_join(FL_covars) %>% # add floral covars
    mutate(niche.overlap = 1-distance) # express niche overlap so that 1 = perfect overlap and 0 = no overlap
}
```

# 1. What is the logic of niche partitioning between core species, and does it change through elevation and time?

For ordination-based analyses, I have opted to pool across years to emphasize species relationships rather than interannual differences. To aid interpretation, I have classified BB species by tongue length. The two "true" long-tongued species are *B. pascuorum* and *B. hortorum*, with the latter having a considerably longer tongue than the former, but I have opted to consider *B. wurflenii* as "behaviorally long-tongued" because it characteristically forages as a robber, and evidently converges in flower choice with *B. pascuorum* and *B. hortorum*. This decision was made *post-hoc* after noting the clustering patterns.

To avoid overplotting plant species / Kugler classes, I have only plotted species loading above a given threshold on either the first or second axis. Thus, the plotted species are those that contribute most to the separation, not necessarily those that contribute most to diet. A species that is visited heavily by all BB species will fall close to (0,0) and will therefore not be plotted.

## 1.1 Ordination by species with post-hoc fitting of tongue length

**Results**  
1. The most salient pattern is that the long-tongued species cluster tightly with one another, especially in pollen foraging, while the short-tongued species are widely spread out. In other words, with respect to dietary niche space, long-tongued species converge while short-tongued species diverge.

2. Broadly speaking, MDS1 explains the difference between short- and long-tongued species while MDS2 explains the separation within those groups.

3. *Pollen foraging*: Long-tongued species were strongly associated with flag flowers (Kugler 6.1; *Trifolium pratense*, *Lotus corniculatus*, *Anthyllis vulneraria*) and the true lip flower (Kugler 5.1) *Rhinanthus glacialis*. Short-tongued species were more diffusely associated with disk/bowl flowers (Kugler 1.1-2; *Helianthemum nummularium*), flower heads (Kugler 7.1-2; *Phyteuma orbiculare*), and bell-shaped flowers of the sticky pollen persuasion (Kugler 3.2; *Campanula scheuchzeri*).

4. *Nectar foraging*: Long-tongued species were strongly associated with true lip flowers (Kugler 5.1; *Rhinanthus glacialis*, *Salvia glutinosa*). Short-tongued species are more diffusely associated with Asteraceae flower heads (Kugler 7.2; *Carduus defloratus*, *Centaurea jacea*) and bell flowers with powdery pollen (Kugler 3.1; *Erica carnea*). *B. soroensis* was especially associated with the Asteraceae.

**Discussion**
1. No surprises here with respect to the short-tongued/long-tongued divide; long-tonguers like deep/closed flowers, short-tonguers like shallow/open flowers. It is interesting, though, to see just how strongly B. wurflenii behaves as a long-tongued species. This is consistent, I suppose, with the generalization that deep/closed flowers off higher rewards. If you can rob, I guess you choose to rob the flowers that give you the highest reward rather than forage legitimately from open flowers. One thing that I find a bit puzzling is that these patterns hold for pollen as well as nectar; in fact, they are stronger for pollen than for nectar. Pollen collecting has nothing to do with tongue length, right? It's all in the legs. But I can guess at two explanations for this. First, just because a bee was noted as carrying pollen doesn't mean it was not nectar foraging (but it is probably safer to assume that a bee *not* carrying pollen was not foraging for pollen), so my pollen category probably represents a combination of pollen and nectar foraging. Second, and I think more convincingly, pollen and nectar foraging are not independent activities. A bee that learns to associate a given flower type with reward might be expected to return to that flower type to seek a different reward. And this "spillover" could end up being evolutionarily imprinted on the innate floral preferences BB spp. Right?

2. Note how far-flung *B. sensu-strictu* tends to be. With respect to nectar foraging, this might have something to do with *B. ss*'s affinity for *Erica carnea*. But it seems to be a bit of a wildcard among our dominant *Bombus* spp.

```{r message=FALSE, warning=FALSE, echo = FALSE}
# By plant.sp
niche.tax <- niche_analysis1(net, bb_traits, criterion = plant.sp.abb)

## Extract and name plots
pollen.tax <- filter(niche.tax, pollen == "pollen")$pcoa.plot[[1]]
nectar.tax <- filter(niche.tax, pollen == "nopollen")$pcoa.plot[[1]]

## Display plots on grid
# pdf("../analyses/output/tax_niche.pdf", height = 6, width = 11)
# (pollen.tax | nectar.tax) + plot_layout(guides = 'collect')
# dev.off()

# By ktype
niche.morph <- niche_analysis1(net, bb_traits, criterion = k.type.s)

## Extract and name plots
pollen.morph <- filter(niche.morph, pollen == "pollen")$pcoa.plot[[1]]
nectar.morph <- filter(niche.morph, pollen == "nopollen")$pcoa.plot[[1]]

## Display plots on grid
# pdf("../analyses/output/morph_niche.pdf", height = 6, width = 11)
# (pollen.morph | nectar.morph) + plot_layout(guides = 'collect') 
# dev.off()

## Taxonomic and morphological together
#pdf("../analyses/output/stack_niche.pdf", height = 8.5, width = 11)
(pollen.tax + ggtitle("Pollen/Taxonomic") | nectar.tax + ggtitle("Nectar/Taxonomic")) /
(pollen.morph + ggtitle("Pollen/Morphological") | nectar.morph + ggtitle("Nectar/Morphological")) + plot_layout(guides = 'collect') 
#dev.off()
```

## 1.2 Ordination by species with post-hoc fitting of tongue length, broken down by elevation class

**Results**  
1. The degree of separation between short- and long-tongued species varies markedly, as does the tightness of the clustering. At mid and high elevations, the patterns generally echo those described in the aggregate analysis above. At low elevations, though, the separation breaks down and the clustering relaxes. This is especially marked in nectar visitation.

2. *Pollen foraging*

2.1  At low elevation, pollen foraging is a story of the true lip flower (Kugler 5.1) *Rhinanthus alectorolophus* for long-tonguers and flower heads (Kugler 7.1-2; *Centaurea jacea*, *Knautia arvensis*) for short-tonguers. Note that *Rhinanthus alectorolophus* is a different species of *Rhinanthus* than the one that was prominent in the aggregate analysis (*Rhinanthus glacialis*).

2.2 At mid elevation, a bunch of morphological traits come into play. As in the aggregate analysis, long-tonguers are associated with flag flowers (Kugler 6.1; *Trifolium pratense*, *Anthyllis vulneraria*) and the true lip flower (Kugler 5.1) *Rhinanthus glacialis*. Short-tongue species, however, are spread along a wide and almost linear gradient with *B. sensu-strictu* on one end, *B. soroensis* on the other, and *B. pratorum* in the middle. Toward the *bss* end of the gradient, we have the disk/bowl flower (Kugler 1.2) *Hypericum maculatum* (actually K 1.1, but I think 1.1 and 1.2 loaded very close to each other, but 1.1 was dropped by the overplotting threshold) and the unclassified flower (Kugler 0) *Plantago lanceolata*; I think unclassified flowers are those that Kugler reckoned to be wind-pollinated. On the *B. soroensis* end of the gradient, we have the flower head (Kugler 7.1) *Phyteuma orbiculare*. In the middle, near *B. pratorum*, we have the bell flower (Kugler 3.2) *Geum rivale*.

2.3. At high elevation, long-tonguers are still associate with flag flowers and true lip flowers, and indeed by *T. pratense* and *Anthyllis vulneraria*, but *Rhianthus* has been replaced by *Lotus corniculatus* (Kugler 6.1) and *Stachys alopecuros* (Kugler 5.1). Short-tonguers are widely spread and associated with the disk/bowl flower (Kugler 1.1-2) *Helianthemum nummularium*, the flower head (Kugler 7.1) *Phyteuma orbiculare*, and the bell flower (Kugler 3.2) *Campanula scheuchzeri*..

3. *Nectar foraging*

3.1 At low elevation, there is little separation between long-tongued and short-tongued species. The most salient separation is that of *B. soroensis* from other species, driven by its apparent affinity for flower heads (Kugler 7.2; *Centaurea jacea*). *B. hortorum* and *B. pratorum*, despite their huge difference in tongue length, fall close to each other based on their shared affinity for the stalk-disc flower (Kugler 4.1) *Primula elatior*.

3.2 At mid elevation, short- and long-tongued species are separated more strongly. As we saw for pollen foraging, long-tongued species are associated with flag flowers (Kugler 6.1-2; *Polygala chamaebuxus*, *Anthyllis vulneraria*) and true lip flowers (Kugler 5.1; *Rhinanthus glacialis*, *Salvia glutinosa*). Short-tongued species were associated with the bell flower (Kugler 5.1) *Erica carnea*, the flower head (Kugler 7.2) *Centaurea jacea*, and the lip flower (Kugler 5.1) *Origanum vulgare*.

3.3 At high elevation, only two flower species stand out: *Carduus defloratus* (Kugler 7.2) and *Stachys alopecuros* (Kugler 5.1). Not surprisingly, the short-tongued species --- and especially *B. soroensis* --- leaned toward *Carduus* and the long-tongued species --- especially *B. hortorum* and *B. wurflenii* --- leaned toward *Stachys*. Interestingly, *B. pascuorum* and *B. pratorum* landed right next to each other in the middle of the spread.

**Discussion**  
1. It is interesting, though not surprising, that morphological associations are relatively constant even though taxonimic associations (by necessity) turn over strongly.

```{r message=FALSE, warning=FALSE, echo = FALSE}
# By plant.sp
niche_x_elev.tax <- niche_analysis2(net, bb_traits, plant.sp.abb)

## Extract and name plots
pollen_oben.tax <- filter(niche_x_elev.tax, pollen == "pollen" & elev.class == "oben")$pcoa.plot[[1]] + ggtitle("Pollen/Oben/Tax")
pollen_mitte.tax <- filter(niche_x_elev.tax, pollen == "pollen" & elev.class == "mitte")$pcoa.plot[[1]] + ggtitle("Pollen/Mitte/Tax")
pollen_unten.tax <- filter(niche_x_elev.tax, pollen == "pollen" & elev.class == "unten")$pcoa.plot[[1]] + ggtitle("Pollen/Unten/Tax")

nectar_oben.tax <- filter(niche_x_elev.tax, pollen == "nopollen" & elev.class == "oben")$pcoa.plot[[1]] + ggtitle("Nectar/Oben/Tax")
nectar_mitte.tax <- filter(niche_x_elev.tax, pollen == "nopollen" & elev.class == "mitte")$pcoa.plot[[1]] + ggtitle("Nectar/Mitte/Tax")
nectar_unten.tax <- filter(niche_x_elev.tax, pollen == "nopollen" & elev.class == "unten")$pcoa.plot[[1]] + ggtitle("Nectar/Unten/Tax")

# By ktype
niche_x_elev.morph <- niche_analysis2(net, bb_traits, k.type.s)

## Extract and name plots
pollen_oben.morph <- filter(niche_x_elev.morph, pollen == "pollen" & elev.class == "oben")$pcoa.plot[[1]] + ggtitle("Pollen/Oben/Morph")
pollen_mitte.morph <- filter(niche_x_elev.morph, pollen == "pollen" & elev.class == "mitte")$pcoa.plot[[1]] + ggtitle("Pollen/Mitte/Morph")
pollen_unten.morph <- filter(niche_x_elev.morph, pollen == "pollen" & elev.class == "unten")$pcoa.plot[[1]] + ggtitle("Pollen/Unten/Morph")

nectar_oben.morph <- filter(niche_x_elev.morph, pollen == "nopollen" & elev.class == "oben")$pcoa.plot[[1]] + ggtitle("Nectar/Oben/Morph")
nectar_mitte.morph <- filter(niche_x_elev.morph, pollen == "nopollen" & elev.class == "mitte")$pcoa.plot[[1]] + ggtitle("Nectar/Mitte/Morph")
nectar_unten.morph <- filter(niche_x_elev.morph, pollen == "nopollen" & elev.class == "unten")$pcoa.plot[[1]] + ggtitle("Nectar/Unten/Morph")

## Display plots on grid

## Pollen
pdf("../analyses/output/pollen_niche_x_elev.pdf", height = 8.5, width = 11)
(pollen_oben.morph | pollen_oben.tax) /
(pollen_mitte.morph | pollen_mitte.tax) /
(pollen_unten.morph | pollen_unten.tax) + plot_layout(guides = 'collect') 
dev.off()

## Nectar
pdf("../analyses/output/nectar_niche_x_elev.pdf", height = 8.5, width = 11)
(nectar_oben.morph | nectar_oben.tax) /
(nectar_mitte.morph | nectar_mitte.tax) /
(nectar_unten.morph | nectar_unten.tax) + plot_layout(guides = 'collect') 
dev.off()
```

## 1.3 Ordination by BB species * elev.class * period, pooling across years

**Results/Discussion**  
When we subdivide the visitation data by both time and elevation, we see that in many (if not most) elevation-time compartments, only a few floral species account for the niche patterns among BB species, and BB species of a given tongue length are often clustered around a single compatible floral species (though, remember, there could be many shared species that load near the origin and are therefore not plotted). For example, short-tongued bees foraging for nectar early in the season at mid elevation converged tightly around the open bell flower *Erica carnea* while the long-tongued species were able to spread out in the niche space provided by the deep-flowers *Polygala chamaebuxus*, *Primula elatior*, and *Gentiana verna*. Similarly, short-tongued bees foraging for nectar late in the season at low elevation were evidently forced to content themselves with *Centaurea jacea* while their long-tongued counterparts spread themselves broadly in the niche space provided by *Stachys officinalis*, *Lotus corniculatus*, and *Gentiana aspera*. This presents an interesting paradox worth unpacking. When aggregated through elevation and time, long-tongued species appear to be more tightly clustered in niche space, while short-tongued species are broadly spread. This suggests that the long-tongued species are either unable (due to incompatibility with open flowers or competition from short-tongued species) or unwilling (on account of the superior rewards offered by deep flowers) to diverge in niche space, while short-tongued species are either forced to diverge to relieve competition or else vary intrinsically in their preference for and compatibility with different open flower species. When we zoom into to smaller chunks of time and elevation, however, it is more often the short-tongued species that are pidgeon-holed in dietary niche space, presumably because most of the options are simply inaccessible to them on account of their short tongues, while long-tongued species can diverge among available deep flowers. This might suggest that potential competition is stronger among the long-tongued species, since their diet is regionally narrower, but realized competition is higher among short-tongued species, since their diet is locally narrower. This conclusion can only be drawn tentatively, though, since this ordination analysis shows how BB species separate, but it does not show the abundance or diversity of their floral hosts. 
```{r message=FALSE, warning=FALSE, echo = FALSE}
# By plant species
niche_x_elevperiod.tax <- niche_analysis3(net, bb_traits, plant.sp.abb)

## Extract and name plots
pollen_oben_early.tax <- filter(niche_x_elevperiod.tax, pollen == "pollen" & elev.class == "oben" & period == "early")$pcoa.plot[[1]] +
  ggtitle("Pollen/Oben/Early/Tax")
pollen_oben_mid.tax <- filter(niche_x_elevperiod.tax, pollen == "pollen" & elev.class == "oben" & period == "mid")$pcoa.plot[[1]] +
  ggtitle("Pollen/Oben/Mid/Tax")
pollen_oben_late.tax <- filter(niche_x_elevperiod.tax, pollen == "pollen" & elev.class == "oben" & period == "late")$pcoa.plot[[1]] +
  ggtitle("Pollen/Oben/Late/Tax")

pollen_mitte_early.tax <- filter(niche_x_elevperiod.tax, pollen == "pollen" & elev.class == "mitte" & period == "early")$pcoa.plot[[1]] +
  ggtitle("Pollen/Mitte/Early/Tax")
pollen_mitte_mid.tax <- filter(niche_x_elevperiod.tax, pollen == "pollen" & elev.class == "mitte" & period == "mid")$pcoa.plot[[1]] +
  ggtitle("Pollen/Mitte/Mid/Tax")
pollen_mitte_late.tax <- filter(niche_x_elevperiod.tax, pollen == "pollen" & elev.class == "mitte" & period == "late")$pcoa.plot[[1]] +
  ggtitle("Pollen/Mitte/Late/Tax")

#pollen_unten_early.tax <- filter(niche_x_elevperiod.tax, pollen == "pollen" & elev.class == "unten" & period == "early")$pcoa.plot[[1]] +
  #ggtitle("Pollen/Unten/Early/Tax")
pollen_unten_mid.tax <- filter(niche_x_elevperiod.tax, pollen == "pollen" & elev.class == "unten" & period == "mid")$pcoa.plot[[1]] +
  ggtitle("Pollen/Unten/Mid/Tax")
pollen_unten_late.tax <- filter(niche_x_elevperiod.tax, pollen == "pollen" & elev.class == "unten" & period == "late")$pcoa.plot[[1]] +
  ggtitle("Pollen/Unten/Late/Tax")

nectar_oben_early.tax <- filter(niche_x_elevperiod.tax, pollen == "nopollen" & elev.class == "oben" & period == "early")$pcoa.plot[[1]] +
  ggtitle("Nectar/Oben/Early/Tax")
nectar_oben_mid.tax <- filter(niche_x_elevperiod.tax, pollen == "nopollen" & elev.class == "oben" & period == "mid")$pcoa.plot[[1]] +
  ggtitle("Nectar/Oben/Mid/Tax")
nectar_oben_late.tax <- filter(niche_x_elevperiod.tax, pollen == "nopollen" & elev.class == "oben" & period == "late")$pcoa.plot[[1]] +
  ggtitle("Nectar/Oben/Late/Tax")

nectar_mitte_early.tax <- filter(niche_x_elevperiod.tax, pollen == "nopollen" & elev.class == "mitte" & period == "early")$pcoa.plot[[1]] +
  ggtitle("Nectar/Mitte/Early/Tax")
nectar_mitte_mid.tax <- filter(niche_x_elevperiod.tax, pollen == "nopollen" & elev.class == "mitte" & period == "mid")$pcoa.plot[[1]] +
  ggtitle("Nectar/Mitte/Mid/Tax")
nectar_mitte_late.tax <- filter(niche_x_elevperiod.tax, pollen == "nopollen" & elev.class == "mitte" & period == "late")$pcoa.plot[[1]] +
  ggtitle("Nectar/Mitte/Late/Tax")

nectar_unten_early.tax <- filter(niche_x_elevperiod.tax, pollen == "nopollen" & elev.class == "unten" & period == "early")$pcoa.plot[[1]] +
  ggtitle("Nectar/Unten/Early/Tax")
nectar_unten_mid.tax <- filter(niche_x_elevperiod.tax, pollen == "nopollen" & elev.class == "unten" & period == "mid")$pcoa.plot[[1]] +
  ggtitle("Nectar/Unten/Mid/Tax")
nectar_unten_late.tax <- filter(niche_x_elevperiod.tax, pollen == "nopollen" & elev.class == "unten" & period == "late")$pcoa.plot[[1]] +
  ggtitle("Nectar/Unten/Late/Tax")

# By ktype
niche_x_elevperiod.morph <- niche_analysis3(net, bb_traits, k.type.s)

## Extract and name plots
pollen_oben_early.morph <- filter(niche_x_elevperiod.morph, pollen == "pollen" & elev.class == "oben" & period == "early")$pcoa.plot[[1]] +
  ggtitle("Pollen/Oben/Early/Tax")
pollen_oben_mid.morph <- filter(niche_x_elevperiod.morph, pollen == "pollen" & elev.class == "oben" & period == "mid")$pcoa.plot[[1]] +
  ggtitle("Pollen/Oben/Mid/Tax")
pollen_oben_late.morph <- filter(niche_x_elevperiod.morph, pollen == "pollen" & elev.class == "oben" & period == "late")$pcoa.plot[[1]] +
  ggtitle("Pollen/Oben/Late/Tax")

pollen_mitte_early.morph <- filter(niche_x_elevperiod.morph, pollen == "pollen" & elev.class == "mitte" & period == "early")$pcoa.plot[[1]] +
  ggtitle("Pollen/Mitte/Early/Tax")
pollen_mitte_mid.morph <- filter(niche_x_elevperiod.morph, pollen == "pollen" & elev.class == "mitte" & period == "mid")$pcoa.plot[[1]] +
  ggtitle("Pollen/Mitte/Mid/Tax")
pollen_mitte_late.morph <- filter(niche_x_elevperiod.morph, pollen == "pollen" & elev.class == "mitte" & period == "late")$pcoa.plot[[1]] +
  ggtitle("Pollen/Mitte/Late/Tax")

#pollen_unten_early.morph <- filter(niche_x_elevperiod.morph, pollen == "pollen" & elev.class == "unten" & period == "early")$pcoa.plot[[1]] +
  #ggtitle("Pollen/Unten/Early/Tax")
pollen_unten_mid.morph <- filter(niche_x_elevperiod.morph, pollen == "pollen" & elev.class == "unten" & period == "mid")$pcoa.plot[[1]] +
  ggtitle("Pollen/Unten/Mid/Tax")
pollen_unten_late.morph <- filter(niche_x_elevperiod.morph, pollen == "pollen" & elev.class == "unten" & period == "late")$pcoa.plot[[1]] +
  ggtitle("Pollen/Unten/Late/Tax")

nectar_oben_early.morph <- filter(niche_x_elevperiod.morph, pollen == "nopollen" & elev.class == "oben" & period == "early")$pcoa.plot[[1]] +
  ggtitle("Nectar/Oben/Early/Tax")
nectar_oben_mid.morph <- filter(niche_x_elevperiod.morph, pollen == "nopollen" & elev.class == "oben" & period == "mid")$pcoa.plot[[1]] +
  ggtitle("Nectar/Oben/Mid/Tax")
nectar_oben_late.morph <- filter(niche_x_elevperiod.morph, pollen == "nopollen" & elev.class == "oben" & period == "late")$pcoa.plot[[1]] +
  ggtitle("Nectar/Oben/Late/Tax")

nectar_mitte_early.morph <- filter(niche_x_elevperiod.morph, pollen == "nopollen" & elev.class == "mitte" & period == "early")$pcoa.plot[[1]] +
  ggtitle("Nectar/Mitte/Early/Tax")
nectar_mitte_mid.morph <- filter(niche_x_elevperiod.morph, pollen == "nopollen" & elev.class == "mitte" & period == "mid")$pcoa.plot[[1]] +
  ggtitle("Nectar/Mitte/Mid/Tax")
nectar_mitte_late.morph <- filter(niche_x_elevperiod.morph, pollen == "nopollen" & elev.class == "mitte" & period == "late")$pcoa.plot[[1]] +
  ggtitle("Nectar/Mitte/Late/Tax")

nectar_unten_early.morph <- filter(niche_x_elevperiod.morph, pollen == "nopollen" & elev.class == "unten" & period == "early")$pcoa.plot[[1]] +
  ggtitle("Nectar/Unten/Early/Tax")
nectar_unten_mid.morph <- filter(niche_x_elevperiod.morph, pollen == "nopollen" & elev.class == "unten" & period == "mid")$pcoa.plot[[1]] +
  ggtitle("Nectar/Unten/Mid/Tax")
nectar_unten_late.morph <- filter(niche_x_elevperiod.morph, pollen == "nopollen" & elev.class == "unten" & period == "late")$pcoa.plot[[1]] +
  ggtitle("Nectar/Unten/Late/Tax")

## Display plots on grid
pdf("../analyses/output/tax_niche_x_elevperiod_pollen.pdf", height = 8.5, width = 11)
(pollen_oben_early.tax | pollen_oben_mid.tax | pollen_oben_late.tax) /
(pollen_mitte_early.tax | pollen_mitte_mid.tax | pollen_mitte_late.tax) /
(plot_spacer() | pollen_unten_mid.tax  | pollen_unten_late.tax)  + plot_layout(guides = 'collect') 
dev.off()

pdf("../analyses/output/tax_niche_x_elevperiod_nectar.pdf", height = 8.5, width = 11)
(nectar_oben_early.tax | nectar_oben_mid.tax | nectar_oben_late.tax) /
(nectar_mitte_early.tax | nectar_mitte_mid.tax | nectar_mitte_late.tax) /
(nectar_unten_early.tax | nectar_unten_mid.tax  | nectar_unten_late.tax)  + plot_layout(guides = 'collect') 
dev.off()

pdf("../analyses/output/morph_niche_x_elevperiod_pollen.pdf", height = 8.5, width = 11)
(pollen_oben_early.morph | pollen_oben_mid.morph | pollen_oben_late.morph) /
(pollen_mitte_early.morph | pollen_mitte_mid.morph | pollen_mitte_late.morph) /
(plot_spacer() | pollen_unten_mid.morph  | pollen_unten_late.morph)  + plot_layout(guides = 'collect') 
dev.off()

pdf("../analyses/output/morph_niche_x_elevperiod_nectar.pdf", height = 8.5, width = 11)
(nectar_oben_early.morph | nectar_oben_mid.morph | nectar_oben_late.morph) /
(nectar_mitte_early.morph | nectar_mitte_mid.morph | nectar_mitte_late.morph) /
(nectar_unten_early.morph | nectar_unten_mid.morph  | nectar_unten_late.morph)  + plot_layout(guides = 'collect') 
dev.off()
```

## 1.4 A follow up to the question of resource range in long- vs. short-tongued BB species

```{r}
res.rang_x_tongue_x_siteperiod <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
  left_join(bb_traits) %>%
  filter(!is.na(pollen)) %>%
  group_by(elev.class) %>%
  mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
  group_by(site, pollen, period, pbl.fac, plant.sp) %>% # group by site, date, bee species, plant species
  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
  unite(webID, c(site, pollen, period), sep = "_") %>%
  ungroup() %>%
  dplyr::select(higher = pbl.fac, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() %>% # convert to to bipartite web
  map(specieslevel, level = "higher", index = c("resource range", "partner diversity", 
                                                "species specificity", "proportional generality")) %>%
  map(rownames_to_column) %>%
  map(as_tibble) %>%
  bind_rows(.id = "webID") %>% # Collapse list into big data frame
  rename(pbl.fac = rowname) %>%
  separate(webID, c("site", "pollen", "period"), sep =  "_") %>%
  left_join(site_data) %>%
  filter(pbl.fac %in% c("long", "short")) %>% # for some reason, a few rows have a "1" for pbl.fac, and I don't know where this comes from; this gets rid of it, but I would still like to know why
  mutate(site = factor(site),
         pollen = factor(pollen)) %>%
  mutate(period = factor(period, levels = c("early", "mid", "late"))) %>%
  mutate(resource.range = 1-resource.range)

ggplot(res.rang_x_tongue_x_siteperiod, aes(elev.mean, resource.range, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_grid(pollen ~ period)
ggsave("./output/resource_range_x_tongue.pdf", width = 6, height = 4)

ggplot(res.rang_x_tongue_x_siteperiod, aes(elev.mean, partner.diversity, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_grid(pollen ~ period)

ggplot(res.rang_x_tongue_x_siteperiod, aes(elev.mean, species.specificity.index, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_grid(pollen ~ period)

ggplot(res.rang_x_tongue_x_siteperiod, aes(elev.mean, proportional.generality, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_grid(pollen ~ period)

abundance_x_tongue_x_siteperiod <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
  group_by(elev.class) %>%
  mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
  left_join(bb_traits) %>%
  filter(!is.na(pollen)) %>%
  group_by(site, period, pbl.fac) %>% # group by site, date, bee species, plant species
  summarize(abund = n()) %>%
  left_join(site_data)

ggplot(abundance_x_tongue_x_siteperiod, aes(elev.mean, abund, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~period)
```

## 1.5 How do abundant generalist bumble bees deal with a dynamic floral community? Based on the ordination analysis, it looks like they maintain fairly steady morphological preferences, but they are taxonomically flexible. Let's take a closer look at that.

Right now, I am calculating beta-diversity on the basis of presence-absence; if I calculate based on relative abundance, it is extremely high. 
```{r}
# Beta-diversity tables
## Binary
beta.tax.sp.bin <- beta_func(net, site_data, level = bb.sp, 
                             criterion = plant.sp, binary = TRUE) %>%
  mutate(criterion = rep("tax", n()))
beta.tax.guild.bin <- beta_func(net, site_data, level = pbl.fac,
                                criterion = plant.sp, binary = TRUE) %>%
  mutate(criterion = rep("tax", n()))

beta.morph.sp.bin <- beta_func(net, site_data, level = bb.sp,
                               criterion = k.type.s, binary = TRUE) %>%
  mutate(criterion = rep("morph", n()))
beta.morph.guild.bin <- beta_func(net, site_data, level = pbl.fac,
                                  criterion = k.type.s, binary = TRUE) %>%
  mutate(criterion = rep("morph", n()))

beta.sp.bin <- bind_rows(beta.tax.sp.bin, beta.morph.sp.bin)
beta.guild.bin <- bind_rows(beta.tax.guild.bin, beta.morph.guild.bin)

## Not binary
beta.tax.sp <- beta_func(net, site_data, level = bb.sp,
                         criterion = plant.sp, binary = FALSE) %>%
  mutate(criterion = rep("tax", n()))
beta.tax.guild <- beta_func(net, site_data, level = pbl.fac,
                            criterion = plant.sp, binary = FALSE) %>%
  mutate(criterion = rep("tax", n()))

beta.morph.sp <- beta_func(net, site_data, level = bb.sp,
                           criterion = k.type.s, binary = FALSE) %>%
  mutate(criterion = rep("morph", n()))
beta.morph.guild <- beta_func(net, site_data, level = pbl.fac, 
                              criterion = k.type.s, binary = FALSE) %>%
  mutate(criterion = rep("morph", n()))

beta.sp <- bind_rows(beta.tax.sp, beta.morph.sp)
beta.guild <- bind_rows(beta.tax.guild, beta.morph.guild)

# Pull out ADONIS results
## Binary
beta.tax.sp.adonis.bin <- beta.tax.sp %>%
  #select(bb.sp, pollen, elev.r2, elev.p) %>%
  select(bb.sp, elev.r2, elev.p) %>%
  distinct()

beta.morph.sp.adonis.bin <- beta.morph.sp.bin %>%
  #select(bb.sp, pollen, elev.r2, elev.p) %>%
  select(bb.sp, elev.r2, elev.p) %>%
  distinct()

beta.tax.guild.adonis.bin <- beta.tax.guild.bin %>%
  #select(pbl.fac, pollen, elev.r2, elev.p) %>%
  select(pbl.fac, elev.r2, elev.p) %>%
  distinct()

beta.morph.guild.adonis.bin <- beta.morph.guild.bin %>%
  #select(pbl.fac, pollen, elev.r2, elev.p) %>%
  select(pbl.fac, elev.r2, elev.p) %>%
  distinct()

## Not binary
beta.tax.sp.adonis <- beta.tax.sp %>%
  #select(bb.sp, pollen, elev.r2, elev.p) %>%
  select(bb.sp, elev.r2, elev.p) %>%
  distinct()

beta.morph.sp.adonis <- beta.morph.sp %>%
  #select(bb.sp, pollen, elev.r2, elev.p) %>%
  select(bb.sp, elev.r2, elev.p) %>%
  distinct()

beta.tax.guild.adonis <- beta.tax.guild %>%
  #select(pbl.fac, pollen, elev.r2, elev.p) %>%
  select(pbl.fac, elev.r2, elev.p) %>%
  distinct()

beta.morph.guild.adonis <- beta.morph.guild %>%
  #select(pbl.fac, pollen, elev.r2, elev.p) %>%
  select(pbl.fac, elev.r2, elev.p) %>%
  distinct()

# Plot
## Binary
beta.sp.plot.bin <- ggplot(beta.sp.bin, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.125) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
  facet_wrap(~bb.sp) + 
  ggtitle("Species")

beta.guild.plot.bin <- ggplot(beta.guild.bin, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
  facet_wrap(~pbl.fac) + 
  ggtitle("Tongue guild")

(beta.sp.plot.bin | beta.guild.plot.bin) + plot_layout(guides = 'collect')

## Not binary
beta.sp.plot <- ggplot(beta.sp, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.125) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
  facet_wrap(~bb.sp) + 
  ggtitle("Species")

beta.guild.plot <- ggplot(beta.guild, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
  facet_wrap(~pbl.fac) + 
  ggtitle("Tongue guild")

(beta.sp.plot | beta.guild.plot) + plot_layout(guides = 'collect')
```

## 1.6 Does the influence of tongue length on niche partitioning vary with elevation or time?
```{r, echo = FALSE}
# By plant.sp
ggplot(niche_x_elevperiod.tax, aes(elev.class, tongue.r2)) +
  geom_boxplot() +
  facet_wrap(~pollen)

ggplot(niche_x_elevperiod.tax, aes(period, tongue.r2)) +
  geom_boxplot() +
  facet_wrap(~pollen)

# By ktype
ggplot(niche_x_elevperiod.morph, aes(elev.class, tongue.r2)) +
  geom_boxplot() +
  facet_wrap(~pollen)

ggplot(niche_x_elevperiod.morph, aes(period, tongue.r2)) +
  geom_boxplot() +
  facet_wrap(~pollen)
```

# 2. Testing the stress-gradient hypothesis and the Ranta and Vepsäläinen heterogeneity hypothesis

**Stress-gradient hypothesis:** Is there any evidence that niche partitioning amongst generalist core bumble bee species is governed by an inverse relationship between elevation and competition, broadly consistent with the stress-gradient hypothesis (Bertness and Callaway 1994). This mechanism has been inferred from trait patterns in the whole bee community of the Berchtesgadener Alps (Hoiss et al. 2012) and from interaction specialization patterns in the plant-pollinator community of Mt. Kilimanjaro (Classen et al. 2020). It predicts that niche overlap will increase with elevation, reflecting competitive pressures at low elevations and reduced competition but increased need for opportunism at high elevations.

**Ranta/Vepsäläinen heterogeneity hypothesis :** Ranta and Vepsäläinen (1981) proposed that bumble bees communities are governed by a spatiotemporally heterogeneous mosaic of competitive conditions, such that competitive inequalities between species are sufficiently variable to allow more species to coexist than predicted by trait-based competition models.

## 2.1: Testing the stress-gradient hypothesis
### 2.1.1 Face-value analysis

*Results:*  
In the taxonomic model (28% variance explained), niche overlap exhibited a unimodal relationship with elevation in 2010 and 2012 (both significant at p < 0.01), with peak overlap at ~1600 m and ~1250 m, respectively. In 2011, peak niche overlap occurred at the lowest elevations, but after an initial descent, it exhibit a unimodal relationship rather like that of 2010; **this relationship was not significant**, though (p = 0.23). Niche overlap of pollen visits was significantly (p = 0.01) lower than that of nectar visits. There was no significant effect of year.

In the morphological model (31% variance explained), niche overlap exhibited a significant relationship with elevation in all three years. In each year, the relationship was approximately quadratic, peaking at low elevation and descending with increasing elevation. In 2010 and 2012, the curve drops off steepy above 1500 m. In 2011, the curve is nearly linear.  Niche overlap of pollen visits was significantly (p = 0.01) lower than that of nectar visits, and there was no significant effect of year.

These results, of course, do not support H1. But there is a caveat. These models try to explain face-value niche overlap, but they do not try to distinguish observed niche overlap from the predictions of a null model. It is possible that the patterns we see would be predicted simply by the abundance relationships in our community. Niche overlap is, in fact, strongly predicted by bumble bee abundance, which is itself strongly correlated with bumble bee richness. One response is, well, so what? These are the observed patterns of niche overlap, and we may not care to disentangle the effects of abundance from those of behavioral changes.

```{r message=FALSE, warning=FALSE, echo = FALSE}
# Pairwise (bb.sp:bb.sp) niche overlap by site, pooled across years

niche_overlap.tax <- niche_overlapper2(net, criterion = plant.sp.abb)
niche_overlap.morph <- niche_overlapper2(net, criterion = k.type.s)

# All spp pairs
ggplot(niche_overlap.tax, aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth()

ggplot(niche_overlap.morph, aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth()

## Within LT spp
ggplot(filter(niche_overlap.tax, pbl.fac1 == "long" & pbl.fac2 == "long"), 
       aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8))

ggplot(filter(niche_overlap.morph, pbl.fac1 == "long" & pbl.fac2 == "long"), 
       aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8))

## Within ST spp
ggplot(filter(niche_overlap.tax, pbl.fac1 == "short" & pbl.fac2 == "short"), 
       aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8))

ggplot(filter(niche_overlap.morph, pbl.fac1 == "short" & pbl.fac2 == "short"), 
       aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8))

## Between ST and LT
ggplot(filter(niche_overlap.tax, pbl.fac.diff == "different"), 
       aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8))

ggplot(filter(niche_overlap.morph, pbl.fac.diff == "different"), 
       aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8))

# GAMs
glob_GAM.tax <- gam(niche.overlap ~ 
                      pollen +
                      s(elev.mean, by = pollen, k = 10, m = 2),
                    method = "REML",
                    family = "gaussian",
                    data = niche_overlap.tax) %>% getViz()

check.gamViz(glob_GAM.tax)
summary(glob_GAM.tax)
print(plot(glob_GAM.tax, allTerms = TRUE), pages = 1)

glob_GAM.morph <- gam(niche.overlap ~ 
                      pollen +
                      s(elev.mean, by = pollen, k = 10, m = 2),
                    method = "REML",
                    family = "gaussian",
                    data = niche_overlap.morph) %>% getViz()

check.gamViz(glob_GAM.morph)
summary(glob_GAM.morph)
print(plot(glob_GAM.morph, allTerms = TRUE), pages = 1)

## Morphological niche partitioning

# Whole community
ggplot(niche_overlap.morph, aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10)) +
  facet_wrap(~year)

ggplot(niche_overlap.morph2, aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

# Whole community (collapsed to mean)
ggplot(niche_overlap.morph.mean, aes(elev.mean, niche.overlap.mean, color = pollen)) +
  geom_point(alpha = 0.25) +
  facet_wrap(~year) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

ggplot(niche_overlap.morph.mean2, aes(elev.mean, niche.overlap.mean, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

# GAM
niche_comm_morph_GAM <- gam(niche.overlap.mean ~ 
                              year +
                              pollen +
                              #s(sqrt.bb.abund, k = 4) +
                              s(elev.mean, by = year, k = 10),
                            method = "REML",
                            family = "gaussian",
                            data = niche_overlap.morph.mean) %>% getViz()

check.gamViz(niche_comm_morph_GAM)
summary(niche_comm_morph_GAM)
print(plot(niche_comm_morph_GAM, allTerms = TRUE), pages = 1)

niche_comm_morph_GAM2 <- gam(niche.overlap.mean ~ 
                              pollen +
                              s(elev.mean, by = pollen, k = 10),
                            method = "REML",
                            family = "gaussian",
                            data = niche_overlap.morph.mean2) %>% getViz()

check.gamViz(niche_comm_morph_GAM2)
summary(niche_comm_morph_GAM2)
print(plot(niche_comm_morph_GAM2, allTerms = TRUE), pages = 1)
```

We find no support for H1 at the community-mean level. But perhaps this conceals a diversity of responses to elevation in **individual species pairs**. Let's talk about that.

*Results*
These models are hard to fit nicely due to the sparcity of data and high noise at the level of individual species pairs, but what we see here does not appear to add much to what we learned from the community mean model. Individual species pairs tend to exhibit either unimodal or linear decreasing responses, a few of which are statistically significant, most of which are not.

```{r, echo = FALSE}
## Taxonomic niche partitioning

# Whole community (pairwise)
ggplot(niche_overlap.tax, aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10)) +
  facet_grid(year~sp.pair) +
  ylim(c(0,1))

ggplot(niche_overlap.tax2, aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10)) +
  facet_wrap(~sp.pair) +
  ylim(c(0,1))

# GAM
niche_pairwise_tax_GAM <- gam(niche.overlap ~ 
                               s(year, bs = "re") +
                               s(pollen, bs = "re") +
                               s(elev.mean, by = sp.pair, k = 10),
                             method = "REML",
                             family = "gaussian",
                             data = niche_overlap.tax) %>% getViz()

check.gamViz(niche_pairwise_tax_GAM)
summary(niche_pairwise_tax_GAM)
print(plot(niche_pairwise_tax_GAM, allTerms = TRUE), pages = 1)

niche_pairwise_tax_GAM2 <- gam(niche.overlap ~ 
                               s(pollen, bs = "re") +
                               s(elev.mean, by = sp.pair, k = 10),
                             method = "REML",
                             family = "gaussian",
                             data = niche_overlap.tax2) %>% getViz()

check.gamViz(niche_pairwise_tax_GAM2)
summary(niche_pairwise_tax_GAM2)
print(plot(niche_pairwise_tax_GAM2, allTerms = TRUE), pages = 1)

## Morphological niche partitioning

# Whole community (pairwise)
ggplot(niche_overlap.morph, aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10)) +
  facet_grid(year~sp.pair) +
  ylim(c(0,1))

ggplot(niche_overlap.morph2, aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10)) +
  facet_wrap(~sp.pair) +
  ylim(c(0,1))

# GAM
niche_pairwise_morph_GAM <- gam(niche.overlap ~ 
                               s(year, bs = "re") +
                               s(pollen, bs = "re") +
                               s(elev.mean, by = sp.pair, k = 10),
                             method = "REML",
                             family = "gaussian",
                             data = niche_overlap.morph) %>% getViz()

check.gamViz(niche_pairwise_morph_GAM)
summary(niche_pairwise_morph_GAM)
print(plot(niche_pairwise_morph_GAM, allTerms = TRUE), pages = 1)

niche_pairwise_morph_GAM2 <- gam(niche.overlap ~ 
                               s(pollen, bs = "re") +
                               s(elev.mean, by = sp.pair, k = 10),
                             method = "REML",
                             family = "gaussian",
                             data = niche_overlap.morph2) %>% getViz()

check.gamViz(niche_pairwise_morph_GAM2)
summary(niche_pairwise_morph_GAM2)
print(plot(niche_pairwise_morph_GAM2, allTerms = TRUE), pages = 1)
```

### 2.1.2 Null-model residual analysis

Arguably the most rigorous way to tease out the effect of elevation while controlling for community composition (abundance and richness patterns) is to compare the observed niche overlap at each site to that predicted by a random null model, then see how the residuals vary with elevation.

I'll be the first to admit that I'm new to null models, so I want to be extremely clear about the logic of how I'm setting this up. Step-by-step:  
1. Generate community matrices by site, pollen/nectar, and year
2. Permute each matrix *n* times by randomly redistributing the visits of each bumble bee species while holding the total visits for each bumble bee species constant. This is, therefore, a random selection model constrained to the set of flora visited for a given site-pollen-year matrix. Importantly, I am opting not to hold the floral marginal totals constant, i.e. the total number of visits that each plant receives will vary with each permutation. My gut feeling is that this is the right place to start, but this is a decision that requires scrutiny.
3. Calculate niche overlap for each permuted matrix and store as tidy arrays; this mirrors the workflow for calculating empirical niche overlap, the only difference being that I had to deal with some tricky nested lists; I'm pretty sure I did this correctly.
4. Take the mean niche overlap for each permuted community.
5. Join null model mean niche overlap table to empirical niche overlap table, joining by site, year, and pollen. Thus, for each site-year-pollen combo in the empirical niche overlap table, we have *n* iterations of random niche overlap.
6. Calculate the difference between empirical niche overlap and null niche overlap for each iteration; we'll call this "residual niche overlap".
7. Calculate mean residual niche overlap for each site-year-pollen combo. This can then be regressed against elevation to see how the deviation from null expectations varies along the elevation gradient, with the interpretation that this represents the "pure" influence of elevation on niche overlap whle controlling for community differences. We could also use the raw differences without collapsing them to a mean. But that's not what I did this first time through.

There are surely a lot of other ways a null model could be constructed and analyzed, but this seems to me a good place to start.

*Results*
Encouragingly, the results of the null model look very similar to those of the "face-value" analysis. The only difference that stands out is that there is no longer a significant difference between pollen and nectar visits. Perhaps that was an artifact of abundance? In any case, our conclusions regarding H1 are the same.

```{r message=FALSE, warning=TRUE, echo = FALSE}
### Define function that will be used inside map() calls
permat_set <- function(x){
  x %>%
    column_to_rownames("bb.sp") %>%
    vegan::permatfull(fixedmar = "rows", shuffle = "ind", times = 50)
}

ret <- function(x){
  return(x$perm)
}

tidy_dist <- function(x) {
  x %>%
    tidy() %>%
    mutate(index = seq_along())
}

### Generate random niche overlap data
#### Need to look into th the missing value warnings I'm getting from the vegdist() call; I think it's because some plants end up with zero visits in some permutations.
random_niche.tax <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
  filter(!is.na(pollen)) %>% # drop observations with no pollen/nectar classification
  select(site, pollen, year, bb.sp, plant.sp) %>% 
  group_by(site, pollen, year, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
  group_by(site, pollen, year) %>%
  nest() %>% # nest data into list-column grouped by pollen, site, date
  mutate(data = map(data, 
                    function(x) pivot_wider(x, 
                                            names_from = plant.sp, 
                                            values_from = freq) %>% replace(is.na(.), 0)), # turn table into community matrix
         data = map(data, permat_set), # permute community matrices
         data = map(data, ret)) %>% # extract permutations from permat objects
  unnest(data) %>% # unnest
  mutate(data = map(data, function(x) vegdist(x, method = "horn")), # convert each permutation into distance matrix
         data = map(data, tidy)) %>% # convert distance matrices to tidy arrays
  group_by(site, pollen, year) %>%
  mutate(iter = row_number()) %>% # add index to denote iterations
  unnest(cols = data) %>% # unnest again
  rename(bb.sp1 = item1, bb.sp2 = item2) %>%
  unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
  mutate(sp.pair = factor(sp.pair), # ...and make it a factor so that mgcv can handle it
         niche.overlap = 1-distance) %>% # express niche overlap so that 1 = perfect overlap and 0 = no overlap
  select(site, pollen, year, sp.pair, iter, niche.overlap)

random_niche.tax.mean <- random_niche.tax %>%
  group_by(site, pollen, year, iter) %>%
  summarize(random.niche.overlap.mean = mean(niche.overlap))

random_niche.morph <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
  filter(!is.na(pollen)) %>% # drop observations with no pollen/nectar classification
  left_join(fl_traits) %>%
  select(site, pollen, year, bb.sp, k.type.s) %>% 
  group_by(site, pollen, year, bb.sp, k.type.s) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
  group_by(site, pollen, year) %>%
  nest() %>% # nest data into list-column grouped by pollen, site, date
  mutate(data = map(data, function(x) pivot_wider(x, 
                                            names_from = k.type.s, 
                                            values_from = freq) %>% replace(is.na(.), 0)), # turn table into community matrix
         data = map(data, permat_set), # permute community matrices
         data = map(data, ret)) %>% # extract permutations from permat objects
  unnest(data) %>% # unnest
  mutate(data = map(data, function(x) vegdist(x, method = "horn")), # convert each permutation into distance matrix
         data = map(data, tidy)) %>% # convert distance matrices to tidy arrays
  group_by(site, pollen, year) %>%
  mutate(iter = row_number()) %>% # add index to denote iterations
  unnest(cols = data) %>% # unnest again
  rename(bb.sp1 = item1, bb.sp2 = item2) %>%
  unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
  mutate(sp.pair = factor(sp.pair), # ...and make it a factor so that mgcv can handle it
         niche.overlap = 1-distance) %>% # express niche overlap so that 1 = perfect overlap and 0 = no overlap
  select(site, pollen, year, sp.pair, iter, niche.overlap)

random_niche.ktype.mean <- random_niche %>%
  group_by(site, pollen, year, iter) %>%
  summarize(random.niche.overlap.mean = mean(niche.overlap))

## Join null model to empirical niche overlap
# Taxonomic niche overlap
comparison.sp <- niche_overlap.sp.mean %>%
  left_join(random_niche.sp.mean) %>%
  mutate(niche.overlap.resid = niche.overlap.mean - random.niche.overlap.mean) %>%
  group_by(site, year, pollen, elev.mean, niche.overlap.mean) %>%
  summarize(mean.resid = mean(niche.overlap.resid))

comparison.sp2 <- niche_overlap.sp.mean2 %>%
  left_join(random_niche.sp.mean2) %>%
  mutate(niche.overlap.resid = niche.overlap.mean - random.niche.overlap.mean) %>%
  group_by(site, year, pollen, elev.mean, niche.overlap.mean) %>%
  summarize(mean.resid = mean(niche.overlap.resid))

ggplot(comparison.sp, aes(elev.mean, mean.resid, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~ year)

ggplot(comparison.sp2, aes(elev.mean, mean.resid, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~ year)

## GAM
null_resid_sp_GAM <- gam(mean.resid ~ 
                           year +
                           pollen +
                           s(elev.mean, by = year, k = 10),
                         method = "REML",
                         family = "gaussian",
                         data = comparison.sp) %>% getViz()

check.gamViz(null_resid_sp_GAM)
summary(null_resid_sp_GAM)
print(plot(null_resid_sp_GAM, allTerms = TRUE), pages = 1)

null_resid_sp_GAM2 <- gam(mean.resid ~ 
                           pollen +
                           s(elev.mean, by = pollen, k = 10),
                         method = "REML",
                         family = "gaussian",
                         data = comparison.sp2) %>% getViz()

check.gamViz(null_resid_sp_GAM2)
summary(null_resid_sp_GAM2)
print(plot(null_resid_sp_GAM2, allTerms = TRUE), pages = 1)

# Morphological niche overlap
comparison.ktype <- niche_overlap.ktype.mean %>%
  left_join(random_niche.ktype.mean) %>%
  mutate(niche.overlap.resid = niche.overlap.mean - random.niche.overlap.mean) %>%
  group_by(site, year, pollen, elev.mean, niche.overlap.mean) %>%
  summarize(mean.resid = mean(niche.overlap.resid))

comparison.ktype2 <- niche_overlap.ktype.mean2 %>%
  left_join(random_niche.ktype.mean2) %>%
  mutate(niche.overlap.resid = niche.overlap.mean - random.niche.overlap.mean) %>%
  group_by(site, year, pollen, elev.mean, niche.overlap.mean) %>%
  summarize(mean.resid = mean(niche.overlap.resid))

ggplot(comparison.ktype, aes(elev.mean, mean.resid, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~ year)

ggplot(comparison.ktype2, aes(elev.mean, mean.resid, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~ year)

## GAM
null_resid_ktype_GAM <- gam(mean.resid ~ 
                              year +
                              pollen +
                              s(elev.mean, by = year, k = 10),
                            method = "REML",
                            family = "gaussian",
                            data = comparison.ktype) %>% getViz()

check.gamViz(null_resid_ktype_GAM)
summary(null_resid_ktype_GAM)
print(plot(null_resid_ktype_GAM, allTerms = TRUE), pages = 1)

null_resid_ktype_GAM2 <- gam(mean.resid ~ 
                              pollen +
                              s(elev.mean, by = pollen, k = 10),
                            method = "REML",
                            family = "gaussian",
                            data = comparison.ktype2) %>% getViz()

check.gamViz(null_resid_ktype_GAM2)
summary(null_resid_ktype_GAM2)
print(plot(null_resid_ktype_GAM2, allTerms = TRUE), pages = 1)
```

### 2.1.3. Parsing niche contraction and niche separation

Contrary to the predictions of H1, niche overlap decreases sharply at high elevation. This is true for both taxonomic and morphological niche overlap, and it holds in both the face-value and null-residual approaches. This decrease in niche overlap might be explained by two modest, proximate hypotheses:

1.  **H1:** The decrease of niche overlap at high elevations is due to *niche contraction*, i.e. BB spp. visiting a smaller proportion of available (and "eligible", i.e. visited at least once by a bumble bee) flora. 
  *Prediction:* "resource range" (Poisot et al. 2012) will decrease with elevation (as Dormann notes, resource range is a "backward" metric insofar as a large resource range means a small proportional niche breadth; I will calculate the inverse of resource range so that it makes more sense)
2.  **H2:** The decrease of niche overlap at high elevations is due to *niche separation*, i.e. BB spp. avoiding one another's niches but not contracting their own.
  *Predictions:* C score will increase with elevation

*Results:* I haven't run formal stats on this analysis yet, but from the visualization, it looks like the overall story is that niche overlap is driven mainly by niche separation, not so much but niche contraction. My first impression is that this suggests a behavior response to competition.

```{r, echo = FALSE}
resource_range.sp <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
  filter(!is.na(pollen)) %>%
  group_by(site, year, pollen, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
  unite(webID, c(site, year, pollen), sep = "_") %>%
  ungroup() %>%
  dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() %>% # convert to to bipartite web
  map(specieslevel, level = "higher", index = "resource range") %>%
  map(rownames_to_column) %>%
  map(as_tibble) %>%
  bind_rows(.id = "webID") %>% # Collapse list into big data frame
  rename(bb.sp = rowname) %>%
  separate(webID, c("site", "year", "pollen"), sep =  "_") %>%
  left_join(site_data) %>%
  mutate(site = factor(site),
         year = factor(year),
         pollen = factor(pollen)) %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # for some reason, one row has a "1" for bb.sp, and I don't know where this comes from; this gets rid of it, but I would still like to know why
  mutate(resource.range = 1-resource.range) %>%
  left_join(BB_covars.splevel) %>%
  filter(bb.sp.abund > 1)

resource_range.sp.mean <- resource_range.sp %>%
  group_by(site, year, elev.mean, pollen) %>%
  summarize(mean.resource.range = mean(resource.range))

ggplot(resource_range.sp, aes(elev.mean, resource.range, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_grid(bb.sp ~ year) +
  ylim(c(0,1))

ggplot(resource_range.sp.mean, aes(elev.mean, mean.resource.range, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year) +
  ylim(c(0,1))

resource_range.ktype <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
  filter(!is.na(pollen)) %>%
  group_by(site, year, pollen, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
  unite(webID, c(site, year, pollen), sep = "_") %>%
  ungroup() %>%
  dplyr::select(higher = bb.sp, lower = k.type.s, webID, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() %>% # convert to to bipartite web
  map(specieslevel, level = "higher", index = "resource range") %>%
  map(rownames_to_column) %>%
  map(as_tibble) %>%
  bind_rows(.id = "webID") %>% # Collapse list into big data frame
  rename(bb.sp = rowname) %>%
  separate(webID, c("site", "year", "pollen"), sep =  "_") %>%
  left_join(site_data) %>%
  mutate(site = factor(site),
         year = factor(year),
         pollen = factor(pollen)) %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # for some reason, one row has a "1" for bb.sp, and I don't know where this comes from; this gets rid of it, but I would still like to know why
  mutate(resource.range = 1-resource.range) %>%
  left_join(BB_covars.splevel) %>%
  filter(bb.sp.abund > 1)

resource_range.ktype.mean <- resource_range.ktype %>%
  group_by(site, year, elev.mean, pollen) %>%
  summarize(mean.resource.range = mean(resource.range))

ggplot(resource_range.ktype, aes(elev.mean, resource.range, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_grid(bb.sp ~ year) +
  ylim(c(0,1))

ggplot(resource_range.ktype.mean, aes(elev.mean, mean.resource.range, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year) +
  ylim(c(0,1))

cscore.sp <- net %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    filter(!is.na(pollen)) %>%
    group_by(site, year, pollen, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
    summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
    group_by(site, year, pollen) %>%
    unite(webID, c(site, year, pollen), sep = "_") %>%
    ungroup() %>%
    dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
    data.frame() %>% # convert to data frame
    frame2webs() %>% # convert to to bipartite web
    map(grouplevel, level = "higher", weighted = TRUE, index = c("C score")) %>% 
    data.frame() %>%
    t() %>%
    data.frame() %>%
    rownames_to_column() %>%
    dplyr::select(webID = rowname, everything()) %>%
    as_tibble() %>%
    separate(webID, c("site", "year", "pollen"), sep =  "_") %>%
    left_join(site_data) %>%
    mutate(site = factor(site),
           year = factor(year),
           pollen = factor(pollen))

ggplot(cscore.sp, aes(elev.mean, C.score.HL, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

cscore.ktype <- net %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    filter(!is.na(pollen)) %>%
    group_by(site, year, pollen, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
    summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
    group_by(site, year, pollen) %>%
    unite(webID, c(site, year, pollen), sep = "_") %>%
    ungroup() %>%
    dplyr::select(higher = bb.sp, lower = k.type.s, webID, freq) %>% # rename columns to match bipartite's expectations
    data.frame() %>% # convert to data frame
    frame2webs() %>% # convert to to bipartite web
    map(grouplevel, level = "higher", weighted = TRUE, index = c("C score")) %>% 
    data.frame() %>%
    t() %>%
    data.frame() %>%
    rownames_to_column() %>%
    dplyr::select(webID = rowname, everything()) %>%
    as_tibble() %>%
    separate(webID, c("site", "year", "pollen"), sep =  "_") %>%
    left_join(site_data) %>%
    mutate(site = factor(site),
           year = factor(year),
           pollen = factor(pollen))

ggplot(cscore.ktype, aes(elev.mean, C.score.HL, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

### 2020-11-26 analyses/old_scripts/niche_relations_v3.Rmd
did so many things

--- 2020-11-26 analyses/old_scripts/niche_relations_v2.Rmd
@@
 library(mgcViz)
-library(ggbipart)
-library(GGally)
+#library(GGally)
 library(vegan)
@@
 library(ggrepel)
-library(gridExtra)
+#library(gridExtra)
 library(tidyverse)
@@
 site_data <- read_csv("../processed_data/site_data.csv") %>%
-  dplyr::select(-c(slope.calc, slope.est, elev.min, 
-                   elev.max, temp.mean)) %>% # drop these variables
-  mutate(elev.class = factor(elev.class, 
-                             levels = c("unten", "mitte", "oben")),
-         elev.class2 = factor(elev.class2, 
-                             levels = c("low", "mid", "high"))) # turn elav.class into an ordered factor
+  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
+  mutate(elev.class = factor(elev.class, levels = c("unten", "mitte", "oben"))) # turn elav.class into an ordered factor
 
@@
 bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
-  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
-  mutate(pbl.fac = factor(if_else(bb.sp %in% c("hort", "pasc", "wurf"), "long", "short")))
+  mutate(pbl.fac = factor(if_else(
+    bb.sp %in% c("hort", "pasc", "wurf", "gers", "mend", "hypn"), "long", "short"))
+    )
 
 # FL traits
+## family-level taxonomy
 fl_tax <- read_csv("../processed_data/floral_tax.csv")
 
+## Kugler morphotypes (accessed through Bioflor)
 fl_ktype <- read_csv("../processed_data/floral_k_type.csv")
 
+## diameter and corolla depth data from Gita Benadi
 fl_morph <- read_csv("../processed_data/flower_morphology_GitaBenadi.csv") %>% 
@@
   group_by(plant.sp) %>%
-  summarize(diam = mean(diam),
-            LN = mean(LN))
-
+  summarize(diam = mean(diam), LN = mean(LN))
+
+## verbal description of Kugler morphotypes
 fl_ktype_key <- read_delim("../processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
@@
 
+## join all trait tables
 fl_traits <- fl_tax %>%
@@
   full_join(fl_morph) %>%
-  dplyr::select(plant.sp, plant.genus, plant.family, 
-         k.type, k.type.s, k.type.ss, diam, LN)
+  dplyr::select(plant.sp, plant.genus, plant.family, k.type, k.type.s, k.type.ss, diam, LN)
 
@@
 net <- read_csv("../processed_data/network.csv") %>%
-  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
   mutate(year = factor(year)) %>% # convert year to factor
-  left_join(site_data, by = "site") %>%
-  #left_join(gdd, by = c("site", "date")) %>%
+  left_join(site_data, by = "site") %>% # add site data
   ungroup() %>%
   arrange(site, date) %>%
-  left_join(fl_traits) %>%
-
-  # Fill in missing k.type scores
-  mutate(k.type.s = if_else(plant.genus == "Phyteuma", 7.1, k.type.s),
-         k.type.ss = if_else(plant.genus == "Phyteuma", 7, k.type.ss)) %>%
-  mutate(k.type.s = if_else(plant.genus == "Aconitum", 5.2, k.type.s),
-         k.type.ss = if_else(plant.genus == "Aconitum", 5, k.type.ss)) %>%
-  mutate(k.type.s = if_else(plant.genus == "Cardamine", 1.2, k.type.s),
-         k.type.ss = if_else(plant.genus == "Cardamine", 1, k.type.ss)) %>%
-  mutate(k.type.s = if_else(plant.genus == "Euphrasia", 5.1, k.type.s),
-         k.type.ss = if_else(plant.genus == "Euphrasia", 5, k.type.ss)) %>%
-  mutate(k.type.s = if_else(plant.genus == "Lamium", 5.1, k.type.s),
-         k.type.ss = if_else(plant.genus == "Lamium", 5, k.type.ss)) %>%
-  mutate(k.type.s = if_else(plant.genus == "Mentha", 2.2, k.type.s),
-         k.type.ss = if_else(plant.genus == "Mentha", 2, k.type.ss)) %>%
-  mutate(k.type.s = if_else(plant.genus == "Rumex", 0.0, k.type.s),
-         k.type.ss = if_else(plant.genus == "Rumex", 0, k.type.ss)) %>%
-  mutate(k.type.s = if_else(plant.genus == "Salix", 9.0, k.type.s),
-         k.type.ss = if_else(plant.genus == "Salix", 9, k.type.ss)) %>%
-  mutate(k.type.s = if_else(plant.genus == "Silene", 4.2, k.type.s),
-         k.type.ss = if_else(plant.genus == "Silene", 4, k.type.ss)) %>%
-  mutate(k.type.s = if_else(plant.genus == "Stachys", 5.1, k.type.s),
-         k.type.ss = if_else(plant.genus == "Stachys", 5, k.type.ss)) %>%
-  mutate(k.type.s = if_else(plant.genus == "Taraxacum", 7.2, k.type.s),
-         k.type.ss = if_else(plant.genus == "Taraxacum", 7, k.type.ss)) %>%
-  mutate(k.type.s = if_else(plant.sp == "Gentiana aspera", 2.1, k.type.s),
-         k.type.ss = if_else(plant.sp == "Gentiana aspera", 2, k.type.ss)) %>%
-  mutate(k.type.s = if_else(plant.sp == "Gentiana ciliata", 4.1, k.type.s),
-         k.type.ss = if_else(plant.sp == "Gentiana ciliata", 4, k.type.ss))
+  left_join(fl_traits) %>% # add floral traits
+  group_by(elev.class) %>%
+  mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) # divide each elev.class into 3 equal time slices
 
@@
   dplyr::select(site, date, plant.sp, flower.cover) %>%
-  left_join(fl_traits) %>%
-
-  # Fill in missing k.type scores
-  mutate(k.type.s = if_else(plant.genus == "Phyteuma", 7.1, k.type.s),
-         k.type.ss = if_else(plant.genus == "Phyteuma", 7, k.type.ss)) %>%
-  mutate(k.type.s = if_else(plant.genus == "Aconitum", 5.2, k.type.s),
-         k.type.ss = if_else(plant.genus == "Aconitum", 5, k.type.ss)) %>%
-  mutate(k.type.s = if_else(plant.genus == "Cardamine", 1.2, k.type.s),
-         k.type.ss = if_else(plant.genus == "Cardamine", 1, k.type.ss)) %>%
-  mutate(k.type.s = if_else(plant.genus == "Euphrasia", 5.1, k.type.s),
-         k.type.ss = if_else(plant.genus == "Euphrasia", 5, k.type.ss)) %>%
-  mutate(k.type.s = if_else(plant.genus == "Lamium", 5.1, k.type.s),
-         k.type.ss = if_else(plant.genus == "Lamium", 5, k.type.ss)) %>%
-  mutate(k.type.s = if_else(plant.genus == "Mentha", 2.2, k.type.s),
-         k.type.ss = if_else(plant.genus == "Mentha", 2, k.type.ss)) %>%
-  mutate(k.type.s = if_else(plant.genus == "Rumex", 0.0, k.type.s),
-         k.type.ss = if_else(plant.genus == "Rumex", 0, k.type.ss)) %>%
-  mutate(k.type.s = if_else(plant.genus == "Salix", 9.0, k.type.s),
-         k.type.ss = if_else(plant.genus == "Salix", 9, k.type.ss)) %>%
-  mutate(k.type.s = if_else(plant.genus == "Silene", 4.2, k.type.s),
-         k.type.ss = if_else(plant.genus == "Silene", 4, k.type.ss)) %>%
-  mutate(k.type.s = if_else(plant.genus == "Stachys", 5.1, k.type.s),
-         k.type.ss = if_else(plant.genus == "Stachys", 5, k.type.ss)) %>%
-  mutate(k.type.s = if_else(plant.genus == "Taraxacum", 7.2, k.type.s),
-         k.type.ss = if_else(plant.genus == "Taraxacum", 7, k.type.ss)) %>%
-  mutate(k.type.s = if_else(plant.sp == "Gentiana aspera", 2.1, k.type.s),
-         k.type.ss = if_else(plant.sp == "Gentiana aspera", 2, k.type.ss)) %>%
-  mutate(k.type.s = if_else(plant.sp == "Gentiana ciliata", 4.1, k.type.s),
-         k.type.ss = if_else(plant.sp == "Gentiana ciliata", 4, k.type.ss))
-
-### Auxiliary data, i.e. all the covariates we need to control for
-BB_covars <- net %>%
-  group_by(site, year) %>%
-  summarize(bb.rich = length(unique(bb.sp)),
-            sqrt.bb.abund = sqrt(n()))
-
-FL_covars <- survey %>%
-  mutate(year = factor(year(date))) %>%
-  filter(flower.cover > 0) %>%
-  group_by(site, year) %>%
-  summarize(fl.rich = length(unique(plant.sp)),
-            log.fl.abund = log(sum(flower.cover)))
-
-### Species-level by site (we will also separate pollen and nectar foragers)
-BB_covars.splevel <- net %>%
-  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
-  group_by(bb.sp, site, year, pollen) %>%
-  summarize(bb.sp.abund = n())
-
-### Species names and such for plotting
-species_names <- cbind(c("bss", "pasc", "prat", 
-                         "soro", "wurf", "hort", 
-                         "pyre", "mend", "mont", 
-                         "gers", "hypn","jone", 
-                         "lapi", "muci", "psit", 
-                         "humi"),
-                       c("B. sensu-strictu", "B. pascuorum", "B. pratorum", 
-                         "B. soroensis", "B. wurflenii", "B. hortorum",
-                         "B. pyrenaeus", "B. mendax", "B. monticola",
-                         "B. gerstaeckeri", "B. hypnorum", "B. jonellus", 
-                         "B. lapidarius", "B. mucidus", "B. psithyrus", 
-                         "B. humilis"),
-                       c("B. ss.", "B. pasc.", "B. prat.", 
-                         "B. soro.", "B. wurf.", "B. hort.",
-                         "B. pyre.", "B. mend.", "B. mont.",
-                         "B. gers.", "B. hypn.", "B. jone.", 
-                         "B. lapi.", "B. muci.", "B. psit.", 
-                         "B. humi"),
-                       c("major", "major", "major", 
-                         "major", "major", "common", 
-                         "high elev.", "high elev.", "high elev.",
-                         "minor", "minor", "common", 
-                         "minor", "minor", "common", 
-                         "minor")
-                       ) %>%
-  as.data.frame() %>%
-  as_tibble() %>%
-  dplyr::select(bb.sp = 1, species.name = 2, species.abb = 3, abund.class = 4) %>%
-  mutate(focal = if_else(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf"), TRUE, FALSE))
+  left_join(fl_traits) # add floral traits
 ```
@@
   left_join(site_data) %>%
-  #group_by(site, date, elev.mean, bb.sp, year) %>%
   group_by(site, elev.mean, bb.sp, year) %>%
   summarize(abund = n()) %>%
-  #group_by(site, elev.mean, bb.sp, year) %>%
-  #summarize(abund = mean(abund)) %>%
   group_by(bb.sp) %>%
@@
          elev.med = elev.floor + ((elev.ceiling - elev.floor)/2)) %>%
-  left_join(species_names)
+  mutate(focal = if_else(
+    bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf"), TRUE, FALSE)
+    )
 
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
-ggsave("./output/bb_range.pdf", height = 6, width = 6)
-
-bb_range <- net %>%
-  left_join(site_data) %>%
-  group_by(site, elev.mean, bb.sp, year) %>%
-  summarize(abund = n()) %>%
-  group_by(bb.sp) %>%
-  mutate(elev.floor = min(elev.mean),
-         elev.ceiling = max(elev.mean),
-         elev.range = elev.ceiling - elev.floor,
-         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2)) %>%
-  left_join(species_names)
-
-ggplot(bb_range, aes(reorder(bb.sp, elev.med), elev.mean, color = focal)) +
-  geom_line(size = 2, alpha = 0.25) +
-  geom_point(aes(size = abund), alpha = 0.5) +
-  theme_light(12) +
-  scale_size_continuous(name = "Abundance") +
-  scale_color_discrete(name = "Focal species") +
-  xlab("BB species") +
-  ylab("Elevation") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+ggsave("./output/bb_range_year.pdf", height = 6, width = 6)
 ```
@@
 ```{r, echo = FALSE}
-# Add BB trait data for envfit() call inside map2()
-trait_join <- function(x, y = bb_traits) {
-  x %>%
-    select(bb.sp) %>%
-    left_join(y) %>%
-    column_to_rownames(var = "bb.sp") %>%
-    select(pbl.w)
-}
-
-# Fortify envfit() object fitted to BB tongue length data
-envfit_fort <- function(x){
-  fortify(x) %>%
-  select(Score = Type, Label, MDS1, MDS2) %>%
-  mutate(Label = case_when(Label == "pbl.cat" ~ "T.class"),
-         Score = case_when(Score == "Vector" ~ "T.class"))
-}
-
-# Fortify PCOA 
+# Convert PCoA to tidy array ("fortify")
 pcoa_fort <- function(x){
   fortify(x, axes = 1:2) %>% 
-  mutate(Score = case_when(
+  mutate(Score = case_when( # rename default scores to match my data
     Score == "species" ~ "FL.sp",
@@
 
-# Plot PCOA; to avoid overplotting, loadings < min_load are not plotted
-pcoa_plot <- function(x, min_load = 0.25) {
-
+# Plot PCoA; to avoid overplotting, scores < min_score are not plotted
+pcoa_plot <- function(x, min_score = 0.25) {
+
+  # Add T.class variable for the BB.sp subset of the PCoA
   t.class <- x %>%
     filter(Score == "BB.sp") %>%
-    mutate(T.class = factor(if_else(Label %in% c("hort", "pasc", "wurf"), "long", "short")))
-
-  ggplot(filter(x, (abs(MDS1) > min_load | abs(MDS2) > min_load) | (Score %in% c("BB.sp", "T.len")))) + # to avoid overplotting, we can filter to include only floral species > loading threshold
-    geom_polygon(data = t.class, aes(MDS1, MDS2, linetype = T.class), 
+    mutate(T.class = factor(if_else(
+      Label %in% c("hort", "pasc", "wurf", "gers", "mend", "hypn"), "long", "short"))
+      )
+
+  ggplot(filter(x, (abs(MDS1) > min_score | abs(MDS2) > min_score) | (Score == "BB.sp"))) + # no threshold for BB spp
+    geom_polygon(data = t.class, aes(MDS1, MDS2, linetype = T.class), # draw polygons around T classes
                  color = "gray70", fill = NA, size = 1.25, key_glyph = "abline") +
-    geom_point(aes(MDS1, MDS2, color = Score), size = 3, alpha = 0.75) +
-    geom_text_repel(aes(MDS1, MDS2, color = Score, label = Label), key_glyph = "rect") +
+    geom_point(aes(MDS1, MDS2, color = Score), size = 3, alpha = 0.75) + # add points
+    ggrepel::geom_text_repel(aes(MDS1, MDS2, color = Score, label = Label), key_glyph = "rect") + # add labels
     theme_classic()
@@
 
-# Niche analysis function; generates PCOA, envfit, ordination plot, and tongue length R2 coming out of envfit
-niche_analysis1 <- function(x, y, criterion) {
-
-  criterion <- enquo(criterion)
+# Generates PCoAs, ADONIS tests, and ordination plots (fully aggregated)
+## The ... represents the variables we will use to parse our analysis. For aggregate analysis, just leave the ...
+## unpopulated, e.g. niche_analysis(net, bb_traits, plant.sp). 
+## To parse by elev.class: niche_analysis(net, bb_traits, plant.sp)
+## To parse by  elev.class, period, and pollen: niche_analysis(net, bb_traits, plant.sp, period, pollen)
+## You get the idea. You should anyway -- you wrote it. https://tidyeval.tidyverse.org/multiple.html
+niche_analysis <- function(x, y, criterion, ...) {
+
+  criterion <- enquo(criterion) # this argument toggles taxonomic (plant.sp) vs. morphological (k.type.s) analysis
 
   x %>%
-    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
-    filter(!is.na(pollen)) %>%
-    group_by(bb.sp, !!criterion, pollen) %>%
-    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site
-    group_by(bb.sp, pollen) %>%
-    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair per site-date
-    dplyr::select(-freq) %>%
-    group_by(pollen) %>%
-    nest() %>%
-    mutate(comm1 = map(data, function(x) pivot_wider(x, names_from = !!criterion, values_from = prop)), # spread into the kind of array vegan likes
-           comm2 = map(comm1, function(x) replace(x, is.na(x), 0)),
-           comm3 = map(comm2, function(x) column_to_rownames(x, var = "bb.sp")),
-           pcoa = map(comm3,  function(x) capscale(x ~ 1, distance = "horn")),
-           traits = map(comm2, function(x) x %>% 
-                          left_join(y) %>% 
+    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
+    group_by(..., bb.sp, !!criterion) %>% # pooling all sites and years
+    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
+    group_by(..., bb.sp) %>%
+    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
+    dplyr::select(-freq) %>% # drop freq variable
+    group_by(...) %>%
+    nest() %>% # nest into list-column
+    mutate(comm1 = map(data, # spread into the kind of array vegan likes
+                       function(x) pivot_wider(x, names_from = !!criterion, values_from = prop) %>%
+                         replace(is.na(.), 0)
+                       ), 
+           comm2 = map(comm1, # column to rowname
+                       function(x) column_to_rownames(x, var = "bb.sp")
+                       ),
+           pcoa = map(comm2, # call PCoA
+                      function(x) capscale(x ~ 1, distance = "horn")
+                      ),
+           traits = map(comm1, # join BB traits
+                        function(x) x %>% 
+                          left_join(y) %>% # y is the trait table, an argument in the main function call
                           column_to_rownames(var = "bb.sp") %>% 
-                          select(pbl.fac)),
-           adonis = map2(comm3, traits, function(x, y) adonis(x ~ pbl.fac, data = y, method = "horn")),
+                          select(pbl.fac) # pbl.fac is the variable I'm interested in
+                        ),
+           adonis = map2(comm2, traits, # run adonis permuatation test for distance ~ tongue class
+                         function(x, y) adonis(x ~ pbl.fac, data = y, method = "horn")
+                         ),
+           betadis = map2(comm2, traits, # betadisper to test for difference in dispersion between tongue classes
+                         function(x, y) x %>%
+                           vegdist(method = "horn") %>% # betadisper runs a PCoA under the hood; match method to my PCoA
+                           betadisper(group = y$pbl.fac) 
+                         ),
+           betadis.permutest = map(betadis, permutest), # perumutation test
+           betadis.p = map(betadis.permutest, # extract betadisper p-value
+                           function(x) x$tab$'Pr(>F)'[[1]]
+                           ),
+           betadis.boxplot = map(betadis, boxplot), # boxplot of betadisper
            pcoa.fort = map(pcoa, pcoa_fort), 
            pcoa.plot = map(pcoa.fort, pcoa_plot),
-           tongue.r2 = map(adonis, function(x) x$aov.tab$R2[[1]]),
-           tongue.p = map(adonis, function(x) x$aov.tab$'Pr(>F)'[[1]]),
-           MDS1.varexp = map(pcoa, function(x) x$CA$eig[[1]]),
-           MDS2.varexp = map(pcoa, function(x) x$CA$eig[[2]])) %>%
-    unnest(cols = c(tongue.r2, tongue.p, MDS1.varexp, MDS2.varexp)) %>%
-    select(pollen, comm2, comm3, traits, pcoa, pcoa.fort, adonis,
-           pcoa.plot, tongue.r2, tongue.p, MDS1.varexp, MDS2.varexp) 
-}
-
-# This version break it down by elev.class
-niche_analysis2 <- function(x, y, criterion) {
-
-  criterion <- enquo(criterion)
-
-  x %>%
-    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
-    filter(!is.na(pollen)) %>%
-    group_by(elev.class, bb.sp, !!criterion, pollen) %>%
-    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site
-    group_by(elev.class, bb.sp, pollen) %>%
-    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair per site-date
-    dplyr::select(-freq) %>%
-    group_by(pollen, elev.class) %>%
-    nest() %>%
-    mutate(comm1 = map(data, function(x) pivot_wider(x, names_from = !!criterion, values_from = prop)), # spread into the kind of array vegan likes
-           comm2 = map(comm1, function(x) replace(x, is.na(x), 0)),
-           comm3 = map(comm2, function(x) column_to_rownames(x, var = "bb.sp")),
-           pcoa = map(comm3,  function(x) capscale(x ~ 1, distance = "horn")),
-           traits = map(comm2, function(x) x %>% 
-                          left_join(y) %>% 
-                          column_to_rownames(var = "bb.sp") %>% 
-                          select(pbl.fac)),
-           adonis = map2(comm3, traits, function(x, y) adonis(x ~ pbl.fac, data = y, method = "horn")),
-           pcoa.fort = map(pcoa, pcoa_fort),
-           pcoa.plot = map(pcoa.fort, pcoa_plot),
-           tongue.r2 = map(adonis, function(x) x$aov.tab$R2[[1]]),
-           tongue.p = map(adonis, function(x) x$aov.tab$'Pr(>F)'[[1]]),
-           MDS1.varexp = map(pcoa, function(x) x$CA$eig[[1]]),
-           MDS2.varexp = map(pcoa, function(x) x$CA$eig[[2]])) %>%
-    unnest(cols = c(tongue.r2, tongue.p, MDS1.varexp, MDS2.varexp)) %>%
-    select(pollen, elev.class, comm2, comm3, traits, pcoa, pcoa.fort, adonis,
-           pcoa.plot, tongue.r2, tongue.p, MDS1.varexp, MDS2.varexp) 
-}
-
-# ...and this version breaks it down by elev.class and period
-niche_analysis3 <- function(x, y, criterion) {
-
-  criterion <- enquo(criterion)
-
-  x %>%
-    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
-    filter(!is.na(pollen)) %>%
-    group_by(elev.class) %>%
-    mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
-    group_by(elev.class, period, pollen, bb.sp, !!criterion) %>%
-    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site
-    group_by(elev.class, period, bb.sp, pollen) %>%
-    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair per site-date
-    dplyr::select(-freq) %>%
-    group_by(pollen, elev.class, period) %>%
-    nest() %>%
-    mutate(comm1 = map(data, function(x) pivot_wider(x, names_from = !!criterion, values_from = prop)), # spread into the kind of array vegan likes
-           comm2 = map(comm1, function(x) replace(x, is.na(x), 0)),
-           comm3 = map(comm2, function(x) column_to_rownames(x, var = "bb.sp")),
-           pcoa = map(comm3,  function(x) capscale(x ~ 1, distance = "horn")),
-           traits = map(comm2, function(x) x %>% 
-                          left_join(y) %>% 
-                          column_to_rownames(var = "bb.sp") %>% 
-                          select(pbl.fac)),
-           adonis = map2(comm3, traits, function(x, y) adonis(x ~ pbl.fac, data = y, method = "horn")),
-           pcoa.fort = map(pcoa, pcoa_fort),
-           pcoa.plot = map(pcoa.fort, pcoa_plot),
-           tongue.r2 = map(adonis, function(x) x$aov.tab$R2[[1]]),
-           tongue.p = map(adonis, function(x) x$aov.tab$'Pr(>F)'[[1]]),
-           MDS1.varexp = map(pcoa, function(x) x$CA$eig[[1]]),
-           MDS2.varexp = map(pcoa, function(x) x$CA$eig[[2]])) %>%
-    unnest(cols = c(tongue.r2, tongue.p, MDS1.varexp, MDS2.varexp)) %>%
-    select(pollen, elev.class, period, comm2, comm3, traits, pcoa, pcoa.fort, adonis,
-           pcoa.plot, tongue.r2, tongue.p, MDS1.varexp, MDS2.varexp) 
+           adonis.r2 = map(adonis, # extract adonis r2
+                           function(x) x$aov.tab$R2[[1]]
+                           ),
+           adonis.p = map(adonis, # extract adonis p-value
+                          function(x) x$aov.tab$'Pr(>F)'[[1]]
+                          ),
+           MDS1 = map(pcoa, # extract eigenvalue for MDS1
+                      function(x) x$CA$eig[[1]]
+                      ), 
+           MDS2 = map(pcoa, # extract eigenvalue for MDS2
+                      function(x) x$CA$eig[[2]]
+                      )) %>%
+    unnest(cols = c(adonis.r2, adonis.p, betadis.p, MDS1, MDS2)) %>% # these are lists of length = 1, so unnesting them simply displays the value
+    select(..., pcoa, betadis, betadis.permutest, betadis.p, adonis, 
+           adonis.r2, adonis.p, betadis.boxplot, pcoa.plot, MDS1, MDS2) 
 }
@@
            pollen = factor(pollen)) %>% # ...and make it a factor so that mgcv can handle it
-    left_join(BB_covars) %>% # add BB covars
-    left_join(FL_covars) %>% # add floral covars
     mutate(niche.overlap = 1-distance) # express niche overlap so that 1 = perfect overlap and 0 = no overlap
@@
 
-To avoid overplotting plant species / Kugler classes, I have only plotted species loading above a given threshold on either the first or second axis.
+To avoid overplotting plant species / Kugler classes, I have only plotted species scoring above a given threshold on either the first or second axis.
 Thus, the plotted species are those that contribute most to the separation, not necessarily those that contribute most to diet.
@@
 # By plant.sp
-niche.tax <- niche_analysis1(net, bb_traits, criterion = plant.sp.abb)
+niche.tax <- niche_analysis(net, bb_traits, criterion = plant.sp.abb)
 
@@
 # By ktype
-niche.morph <- niche_analysis1(net, bb_traits, criterion = k.type.s)
+niche.morph <- niche_analysis(net, bb_traits, criterion = k.type.s)
 
@@
 pratorum* in the middle.
-Toward the *bss* end of the gradient, we have the disk/bowl flower (Kugler 1.2) *Hypericum maculatum* (actually K 1.1, but I think 1.1 and 1.2 loaded very close to each other, but 1.1 was dropped by the overplotting threshold) and the unclassified flower (Kugler 0) *Plantago lanceolata*; I think unclassified flowers are those that Kugler reckoned to be wind-pollinated.
+Toward the *bss* end of the gradient, we have the disk/bowl flower (Kugler 1.2) *Hypericum maculatum* (actually K 1.1, but I think 1.1 and 1.2 scored very close to each other, but 1.1 was dropped by the overplotting threshold) and the unclassified flower (Kugler 0) *Plantago lanceolata*; I think unclassified flowers are those that Kugler reckoned to be wind-pollinated.
 On the *B.
@@
 # By plant.sp
-niche_x_elev.tax <- niche_analysis2(net, bb_traits, plant.sp.abb)
+niche_x_elev.tax <- niche_analysis(net, bb_traits, criterion = plant.sp.abb, elev.class)
 
@@
 # By ktype
-niche_x_elev.morph <- niche_analysis2(net, bb_traits, k.type.s)
+niche_x_elev.morph <- niche_analysis(net, bb_traits, criterion = k.type.s, elev.class)
 
@@
 **Results/Discussion**  
-When we subdivide the visitation data by both time and elevation, we see that in many (if not most) elevation-time compartments, only a few floral species account for the niche patterns among BB species, and BB species of a given tongue length are often clustered around a single compatible floral species (though, remember, there could be many shared species that load near the origin and are therefore not plotted).
+When we subdivide the visitation data by both time and elevation, we see that in many (if not most) elevation-time compartments, only a few floral species account for the niche patterns among BB species, and BB species of a given tongue length are often clustered around a single compatible floral species (though, remember, there could be many shared species that score near the origin and are therefore not plotted).
 For example, short-tongued bees foraging for nectar early in the season at mid elevation converged tightly around the open bell flower *Erica carnea* while the long-tongued species were able to spread out in the niche space provided by the deep-flowers *Polygala chamaebuxus*, *Primula elatior*, and *Gentiana verna*.
@@
 # By plant species
-niche_x_elevperiod.tax <- niche_analysis3(net, bb_traits, plant.sp.abb)
+niche_x_elevperiod.tax <- niche_analysis(net, bb_traits, criterion = plant.sp.abb, elev.class, period)
 
@@
 # By ktype
-niche_x_elevperiod.morph <- niche_analysis3(net, bb_traits, k.type.s)
+niche_x_elevperiod.morph <- niche_analysis(net, bb_traits, criterion = k.type.s, elev.class, period)
 
@@
 beta.sp.plot.bin <- ggplot(beta.sp.bin, aes(elev.diff, distance, color = criterion)) +
-  geom_point(alpha = 0.125) +
+  geom_point(alpha = 0.05) +
   geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
@@
 beta.guild.plot.bin <- ggplot(beta.guild.bin, aes(elev.diff, distance, color = criterion)) +
-  geom_point(alpha = 0.25) +
+  geom_point(alpha = 0.125) +
   geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
@@
 beta.sp.plot <- ggplot(beta.sp, aes(elev.diff, distance, color = criterion)) +
-  geom_point(alpha = 0.125) +
+  geom_point(alpha = 0.05) +
   geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
@@
 beta.guild.plot <- ggplot(beta.guild, aes(elev.diff, distance, color = criterion)) +
-  geom_point(alpha = 0.25) +
+  geom_point(alpha = 0.125) +
   geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
@@
 # All spp pairs
-ggplot(niche_overlap.tax, aes(elev.mean, niche.overlap, color = pollen)) +
-  geom_point() +
-  geom_smooth()
-
-ggplot(niche_overlap.morph, aes(elev.mean, niche.overlap, color = pollen)) +
-  geom_point() +
-  geom_smooth()
+sg1 <- ggplot(niche_overlap.tax, aes(elev.mean, niche.overlap, color = pollen)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
+
+sg2 <- ggplot(niche_overlap.morph, aes(elev.mean, niche.overlap, color = pollen)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
+
+(sg1 + ggtitle("Taxonomic") | sg2 + ggtitle("Morphological")) + plot_layout(guides = 'collect')
 
 ## Within LT spp
-ggplot(filter(niche_overlap.tax, pbl.fac1 == "long" & pbl.fac2 == "long"), 
+sg3 <- ggplot(filter(niche_overlap.tax, pbl.fac1 == "long" & pbl.fac2 == "long"), 
        aes(elev.mean, niche.overlap, color = pollen)) +
-  geom_point() +
+  geom_point(alpha = 0.5) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
+
+sg4 <- ggplot(filter(niche_overlap.morph, pbl.fac1 == "long" & pbl.fac2 == "long"), 
+       aes(elev.mean, niche.overlap, color = pollen)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
+
+(sg3 + ggtitle("Taxonomic") | sg4 + ggtitle("Morphological")) + plot_layout(guides = 'collect')
+
+## Within ST spp
+sg5 <- ggplot(filter(niche_overlap.tax, pbl.fac1 == "short" & pbl.fac2 == "short"), 
+       aes(elev.mean, niche.overlap, color = pollen)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
+
+sg6 <- ggplot(filter(niche_overlap.morph, pbl.fac1 == "short" & pbl.fac2 == "short"), 
+       aes(elev.mean, niche.overlap, color = pollen)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
+
+(sg5 + ggtitle("Taxonomic") | sg6 + ggtitle("Morphological")) + plot_layout(guides = 'collect')
+
+## Between ST and LT
+sg7 <- ggplot(filter(niche_overlap.tax, pbl.fac.diff == "different"), 
+       aes(elev.mean, niche.overlap, color = pollen)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
+
+sg8 <- ggplot(filter(niche_overlap.morph, pbl.fac.diff == "different"), 
+       aes(elev.mean, niche.overlap, color = pollen)) +
+  geom_point(alpha = 0.5) +
   geom_smooth(method = "gam", formula = y ~ s(x, k = 8))
 
-ggplot(filter(niche_overlap.morph, pbl.fac1 == "long" & pbl.fac2 == "long"), 
-       aes(elev.mean, niche.overlap, color = pollen)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 8))
-
-## Within ST spp
-ggplot(filter(niche_overlap.tax, pbl.fac1 == "short" & pbl.fac2 == "short"), 
-       aes(elev.mean, niche.overlap, color = pollen)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 8))
-
-ggplot(filter(niche_overlap.morph, pbl.fac1 == "short" & pbl.fac2 == "short"), 
-       aes(elev.mean, niche.overlap, color = pollen)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 8))
-
-## Between ST and LT
-ggplot(filter(niche_overlap.tax, pbl.fac.diff == "different"), 
-       aes(elev.mean, niche.overlap, color = pollen)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 8))
-
-ggplot(filter(niche_overlap.morph, pbl.fac.diff == "different"), 
-       aes(elev.mean, niche.overlap, color = pollen)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 8))
+(sg7 + ggtitle("Taxonomic") | sg8 + ggtitle("Morphological")) + plot_layout(guides = 'collect')
 
@@
 
-## Morphological niche partitioning
-
-# Whole community
-ggplot(niche_overlap.morph, aes(elev.mean, niche.overlap, color = pollen)) +
-  geom_point(alpha = 0.25) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10)) +
-  facet_wrap(~year)
-
-ggplot(niche_overlap.morph2, aes(elev.mean, niche.overlap, color = pollen)) +
-  geom_point(alpha = 0.25) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
-
-# Whole community (collapsed to mean)
-ggplot(niche_overlap.morph.mean, aes(elev.mean, niche.overlap.mean, color = pollen)) +
-  geom_point(alpha = 0.25) +
-  facet_wrap(~year) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
-
-ggplot(niche_overlap.morph.mean2, aes(elev.mean, niche.overlap.mean, color = pollen)) +
-  geom_point(alpha = 0.25) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
-
-# GAM
-niche_comm_morph_GAM <- gam(niche.overlap.mean ~ 
-                              year +
+# Within LT class
+LT_GAM.tax <- gam(niche.overlap ~ 
                     pollen +
-                              #s(sqrt.bb.abund, k = 4) +
-                              s(elev.mean, by = year, k = 10),
-                            method = "REML",
-                            family = "gaussian",
-                            data = niche_overlap.morph.mean) %>% getViz()
-
-check.gamViz(niche_comm_morph_GAM)
-summary(niche_comm_morph_GAM)
-print(plot(niche_comm_morph_GAM, allTerms = TRUE), pages = 1)
-
-niche_comm_morph_GAM2 <- gam(niche.overlap.mean ~ 
+                    s(elev.mean, by = pollen, k = 10, m = 2),
+                  method = "REML",
+                  family = "gaussian",
+                  data = filter(niche_overlap.tax, pbl.fac1 == "long" & pbl.fac2 == "long")) %>% getViz()
+
+check.gamViz(LT_GAM.tax)
+summary(LT_GAM.tax)
+print(plot(LT_GAM.tax, allTerms = TRUE), pages = 1)
+
+LT_GAM.morph <- gam(niche.overlap ~ 
                     pollen +
-                              s(elev.mean, by = pollen, k = 10),
-                            method = "REML",
-                            family = "gaussian",
-                            data = niche_overlap.morph.mean2) %>% getViz()
-
-check.gamViz(niche_comm_morph_GAM2)
-summary(niche_comm_morph_GAM2)
-print(plot(niche_comm_morph_GAM2, allTerms = TRUE), pages = 1)
+                    s(elev.mean, by = pollen, k = 10, m = 2),
+                  method = "REML",
+                  family = "gaussian",
+                  data = filter(niche_overlap.morph, pbl.fac1 == "long" & pbl.fac2 == "long")) %>% getViz()
+
+check.gamViz(LT_GAM.morph)
+summary(LT_GAM.morph)
+print(plot(LT_GAM.morph, allTerms = TRUE), pages = 1)
+
+# Within ST class
+ST_GAM.tax <- gam(niche.overlap ~ 
+                    pollen +
+                    s(elev.mean, by = pollen, k = 10, m = 2),
+                  method = "REML",
+                  family = "gaussian",
+                  data = filter(niche_overlap.tax, pbl.fac1 == "short" & pbl.fac2 == "short")) %>% getViz()
+
+check.gamViz(ST_GAM.tax)
+summary(ST_GAM.tax)
+print(plot(ST_GAM.tax, allTerms = TRUE), pages = 1)
+
+ST_GAM.morph <- gam(niche.overlap ~ 
+                    pollen +
+                    s(elev.mean, by = pollen, k = 10, m = 2),
+                  method = "REML",
+                  family = "gaussian",
+                  data = filter(niche_overlap.morph, pbl.fac1 == "short" & pbl.fac2 == "short")) %>% getViz()
+
+check.gamViz(ST_GAM.morph)
+summary(ST_GAM.morph)
+print(plot(ST_GAM.morph, allTerms = TRUE), pages = 1)
+
+# Between ST and LT
+STxLT_GAM.tax <- gam(niche.overlap ~ 
+                    pollen +
+                    s(elev.mean, by = pollen, k = 10, m = 2),
+                  method = "REML",
+                  family = "gaussian",
+                  data = filter(niche_overlap.tax, pbl.fac.diff == "different")) %>% getViz()
+
+check.gamViz(STxLT_GAM.tax)
+summary(STxLT_GAM.tax)
+print(plot(STxLT_GAM.tax, allTerms = TRUE), pages = 1)
+
+STxLT_GAM.morph <- gam(niche.overlap ~ 
+                    pollen +
+                    s(elev.mean, by = pollen, k = 10, m = 2),
+                  method = "REML",
+                  family = "gaussian",
+                  data = filter(niche_overlap.morph, pbl.fac.diff == "different")) %>% getViz()
+
+check.gamViz(STxLT_GAM.morph)
+summary(STxLT_GAM.morph)
+print(plot(STxLT_GAM.morph, allTerms = TRUE), pages = 1)
 ```
@@
   geom_smooth(method = "gam", formula = y ~ s(x, k = 10)) +
-  facet_grid(year~sp.pair) +
+  facet_wrap(~sp.pair) +
   ylim(c(0,1))
@@
 I'll be the first to admit that I'm new to null models, so I want to be extremely clear about the logic of how I'm setting this up. Step-by-step:  
-1. Generate community matrices by site, pollen/nectar, and year
+1. Generate community matrices by site and pollen/nectar
 2.
 Permute each matrix *n* times by randomly redistributing the visits of each bumble bee species while holding the total visits for each bumble bee species constant.
-This is, therefore, a random selection model constrained to the set of flora visited for a given site-pollen-year matrix.
+This is, therefore, a random selection model constrained to the set of flora visited for a given site-pollen/nectar matrix.
 Importantly, I am opting not to hold the floral marginal totals constant, i.e.
@@
 5.
-Join null model mean niche overlap table to empirical niche overlap table, joining by site, year, and pollen.
-Thus, for each site-year-pollen combo in the empirical niche overlap table, we have *n* iterations of random niche overlap.
+Join null model mean niche overlap table to empirical niche overlap table, joining by site and pollen/nectar.
+Thus, for each site-pollen/nectar combo in the empirical niche overlap table, we have *n* iterations of random niche overlap.
 6. Calculate the difference between empirical niche overlap and null niche overlap for each iteration; we'll call this "residual niche overlap".
 7.
-Calculate mean residual niche overlap for each site-year-pollen combo.
+Calculate mean residual niche overlap for each site-pollen/nectar combo.
 This can then be regressed against elevation to see how the deviation from null expectations varies along the elevation gradient, with the interpretation that this represents the "pure" influence of elevation on niche overlap whle controlling for community differences.
@@
 *Results*
-Encouragingly, the results of the null model look very similar to those of the "face-value" analysis.
-The only difference that stands out is that there is no longer a significant difference between pollen and nectar visits.
-Perhaps that was an artifact of abundance?
-In any case, our conclusions regarding H1 are the same.
 
@@
   filter(!is.na(pollen)) %>% # drop observations with no pollen/nectar classification
-  select(site, pollen, year, bb.sp, plant.sp) %>% 
-  group_by(site, pollen, year, bb.sp, plant.sp) %>%
+  select(site, pollen, bb.sp, plant.sp) %>% 
+  group_by(site, pollen, bb.sp, plant.sp) %>%
   summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
-  group_by(site, pollen, year) %>%
+  group_by(site, pollen) %>%
   nest() %>% # nest data into list-column grouped by pollen, site, date
@@
          data = map(data, tidy)) %>% # convert distance matrices to tidy arrays
-  group_by(site, pollen, year) %>%
+  group_by(site, pollen) %>%
   mutate(iter = row_number()) %>% # add index to denote iterations
@@
          niche.overlap = 1-distance) %>% # express niche overlap so that 1 = perfect overlap and 0 = no overlap
-  select(site, pollen, year, sp.pair, iter, niche.overlap)
-
-random_niche.tax.mean <- random_niche.tax %>%
-  group_by(site, pollen, year, iter) %>%
-  summarize(random.niche.overlap.mean = mean(niche.overlap))
+  select(site, pollen, sp.pair, iter, niche.overlap.sim = niche.overlap)
 
@@
   filter(!is.na(pollen)) %>% # drop observations with no pollen/nectar classification
-  left_join(fl_traits) %>%
-  select(site, pollen, year, bb.sp, k.type.s) %>% 
-  group_by(site, pollen, year, bb.sp, k.type.s) %>%
+  select(site, pollen, bb.sp, k.type.s) %>% 
+  group_by(site, pollen, bb.sp, k.type.s) %>%
   summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
-  group_by(site, pollen, year) %>%
+  group_by(site, pollen) %>%
   nest() %>% # nest data into list-column grouped by pollen, site, date
-  mutate(data = map(data, function(x) pivot_wider(x, 
+  mutate(data = map(data, 
+                    function(x) pivot_wider(x, 
                                             names_from = k.type.s, 
@@
          niche.overlap = 1-distance) %>% # express niche overlap so that 1 = perfect overlap and 0 = no overlap
-  select(site, pollen, year, sp.pair, iter, niche.overlap)
-
-random_niche.ktype.mean <- random_niche %>%
-  group_by(site, pollen, year, iter) %>%
-  summarize(random.niche.overlap.mean = mean(niche.overlap))
+  select(site, pollen, sp.pair, iter, niche.overlap.sim = niche.overlap)
 
@@
 # Taxonomic niche overlap
-comparison.sp <- niche_overlap.sp.mean %>%
-  left_join(random_niche.sp.mean) %>%
-  mutate(niche.overlap.resid = niche.overlap.mean - random.niche.overlap.mean) %>%
-  group_by(site, year, pollen, elev.mean, niche.overlap.mean) %>%
-  summarize(mean.resid = mean(niche.overlap.resid))
-
-comparison.sp2 <- niche_overlap.sp.mean2 %>%
-  left_join(random_niche.sp.mean2) %>%
-  mutate(niche.overlap.resid = niche.overlap.mean - random.niche.overlap.mean) %>%
-  group_by(site, year, pollen, elev.mean, niche.overlap.mean) %>%
-  summarize(mean.resid = mean(niche.overlap.resid))
-
-ggplot(comparison.sp, aes(elev.mean, mean.resid, color = pollen)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~ year)
-
-ggplot(comparison.sp2, aes(elev.mean, mean.resid, color = pollen)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~ year)
-
-## GAM
-null_resid_sp_GAM <- gam(mean.resid ~ 
-                           year +
+comparison.tax <- niche_overlap.tax %>%
+  left_join(random_niche.tax, by = c("site", "pollen", "sp.pair")) %>%
+  mutate(niche.overlap.resid = niche.overlap - niche.overlap.sim) %>%
+  group_by(site, elev.mean, sp.pair, pollen, pbl.fac1, pbl.fac2, pbl.fac.diff) %>%
+  summarize(niche.overlap.resid = mean(niche.overlap.resid)) %>%
+  mutate(pollen = factor(pollen))
+
+# Morphological niche overlap
+comparison.morph <- niche_overlap.morph %>%
+  left_join(random_niche.morph, by = c("site", "pollen", "sp.pair")) %>%
+  mutate(niche.overlap.resid = niche.overlap - niche.overlap.sim) %>%
+  group_by(site, elev.mean, sp.pair, pollen, pbl.fac1, pbl.fac2, pbl.fac.diff) %>%
+  summarize(niche.overlap.resid = mean(niche.overlap.resid) )%>%
+  mutate(pollen = factor(pollen))
+
+# Plot
+# All spp pairs
+sgn1 <- ggplot(comparison.tax, aes(elev.mean, niche.overlap.resid, color = pollen)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
+
+sgn2 <- ggplot(comparison.morph, aes(elev.mean, niche.overlap.resid, color = pollen)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
+
+(sgn1 + ggtitle("Taxonomic") | sgn2 + ggtitle("Morphological")) + plot_layout(guides = 'collect')
+
+## Within LT spp
+sgn3 <- ggplot(filter(comparison.tax, pbl.fac1 == "long" & pbl.fac2 == "long"), 
+       aes(elev.mean, niche.overlap.resid, color = pollen)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
+
+sgn4 <- ggplot(filter(comparison.morph, pbl.fac1 == "long" & pbl.fac2 == "long"), 
+       aes(elev.mean, niche.overlap.resid, color = pollen)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
+
+(sgn3 + ggtitle("Taxonomic") | sgn4 + ggtitle("Morphological")) + plot_layout(guides = 'collect')
+
+## Within ST spp
+sgn5 <- ggplot(filter(comparison.tax, pbl.fac1 == "short" & pbl.fac2 == "short"), 
+       aes(elev.mean, niche.overlap.resid, color = pollen)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
+
+sgn6 <- ggplot(filter(comparison.morph, pbl.fac1 == "short" & pbl.fac2 == "short"), 
+       aes(elev.mean, niche.overlap.resid, color = pollen)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
+
+(sgn5 + ggtitle("Taxonomic") | sgn6 + ggtitle("Morphological")) + plot_layout(guides = 'collect')
+
+## Between ST and LT
+sgn7 <- ggplot(filter(comparison.tax, pbl.fac.diff == "different"), 
+       aes(elev.mean, niche.overlap.resid, color = pollen)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
+
+sgn8 <- ggplot(filter(comparison.morph, pbl.fac.diff == "different"), 
+       aes(elev.mean, niche.overlap.resid, color = pollen)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 8))
+
+(sgn7 + ggtitle("Taxonomic") | sgn8 + ggtitle("Morphological")) + plot_layout(guides = 'collect')
+
+# GAMs
+glob_nullmodel_GAM.tax <- gam(niche.overlap.resid ~ 
                       pollen +
-                           s(elev.mean, by = year, k = 10),
-                         method = "REML",
-                         family = "gaussian",
-                         data = comparison.sp) %>% getViz()
-
-check.gamViz(null_resid_sp_GAM)
-summary(null_resid_sp_GAM)
-print(plot(null_resid_sp_GAM, allTerms = TRUE), pages = 1)
-
-null_resid_sp_GAM2 <- gam(mean.resid ~ 
+                      s(elev.mean, by = pollen, k = 10, m = 2),
+                    method = "REML",
+                    family = "gaussian",
+                    data = comparison.tax) %>% getViz()
+
+check.gamViz(glob_nullmodel_GAM.tax)
+summary(glob_nullmodel_GAM.tax)
+print(plot(glob_nullmodel_GAM.tax, allTerms = TRUE), pages = 1)
+
+glob_nullmodel_GAM.morph <- gam(niche.overlap.resid ~ 
                       pollen +
-                           s(elev.mean, by = pollen, k = 10),
-                         method = "REML",
-                         family = "gaussian",
-                         data = comparison.sp2) %>% getViz()
-
-check.gamViz(null_resid_sp_GAM2)
-summary(null_resid_sp_GAM2)
-print(plot(null_resid_sp_GAM2, allTerms = TRUE), pages = 1)
-
-# Morphological niche overlap
-comparison.ktype <- niche_overlap.ktype.mean %>%
-  left_join(random_niche.ktype.mean) %>%
-  mutate(niche.overlap.resid = niche.overlap.mean - random.niche.overlap.mean) %>%
-  group_by(site, year, pollen, elev.mean, niche.overlap.mean) %>%
-  summarize(mean.resid = mean(niche.overlap.resid))
-
-comparison.ktype2 <- niche_overlap.ktype.mean2 %>%
-  left_join(random_niche.ktype.mean2) %>%
-  mutate(niche.overlap.resid = niche.overlap.mean - random.niche.overlap.mean) %>%
-  group_by(site, year, pollen, elev.mean, niche.overlap.mean) %>%
-  summarize(mean.resid = mean(niche.overlap.resid))
-
-ggplot(comparison.ktype, aes(elev.mean, mean.resid, color = pollen)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~ year)
-
-ggplot(comparison.ktype2, aes(elev.mean, mean.resid, color = pollen)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~ year)
-
-## GAM
-null_resid_ktype_GAM <- gam(mean.resid ~ 
-                              year +
+                      s(elev.mean, by = pollen, k = 10, m = 2),
+                    method = "REML",
+                    family = "gaussian",
+                    data = comparison.morph) %>% getViz()
+
+check.gamViz(glob_nullmodel_GAM.morph)
+summary(glob_nullmodel_GAM.morph)
+print(plot(glob_nullmodel_GAM.morph, allTerms = TRUE), pages = 1)
+
+# Within LT class
+LT_nullmodel_GAM.tax <- gam(niche.overlap.resid ~ 
                     pollen +
-                              s(elev.mean, by = year, k = 10),
-                            method = "REML",
-                            family = "gaussian",
-                            data = comparison.ktype) %>% getViz()
-
-check.gamViz(null_resid_ktype_GAM)
-summary(null_resid_ktype_GAM)
-print(plot(null_resid_ktype_GAM, allTerms = TRUE), pages = 1)
-
-null_resid_ktype_GAM2 <- gam(mean.resid ~ 
+                    s(elev.mean, by = pollen, k = 10, m = 2),
+                  method = "REML",
+                  family = "gaussian",
+                  data = filter(comparison.tax, pbl.fac1 == "long" & pbl.fac2 == "long")) %>% getViz()
+
+check.gamViz(LT_nullmodel_GAM.tax)
+summary(LT_nullmodel_GAM.tax)
+print(plot(LT_nullmodel_GAM.tax, allTerms = TRUE), pages = 1)
+
+LT_nullmodel_GAM.morph <- gam(niche.overlap.resid ~ 
                     pollen +
-                              s(elev.mean, by = pollen, k = 10),
-                            method = "REML",
-                            family = "gaussian",
-                            data = comparison.ktype2) %>% getViz()
-
-check.gamViz(null_resid_ktype_GAM2)
-summary(null_resid_ktype_GAM2)
-print(plot(null_resid_ktype_GAM2, allTerms = TRUE), pages = 1)
+                    s(elev.mean, by = pollen, k = 10, m = 2),
+                  method = "REML",
+                  family = "gaussian",
+                  data = filter(comparison.morph, pbl.fac1 == "long" & pbl.fac2 == "long")) %>% getViz()
+
+check.gamViz(LT_nullmodel_GAM.morph)
+summary(LT_nullmodel_GAM.morph)
+print(plot(LT_nullmodel_GAM.morph, allTerms = TRUE), pages = 1)
+
+# Within ST class
+ST_nullmodel_GAM.tax <- gam(niche.overlap.resid ~ 
+                    pollen +
+                    s(elev.mean, by = pollen, k = 10, m = 2),
+                  method = "REML",
+                  family = "gaussian",
+                  data = filter(comparison.tax, pbl.fac1 == "short" & pbl.fac2 == "short")) %>% getViz()
+
+check.gamViz(ST_nullmodel_GAM.tax)
+summary(ST_nullmodel_GAM.tax)
+print(plot(ST_nullmodel_GAM.tax, allTerms = TRUE), pages = 1)
+
+ST_nullmodel_GAM.morph <- gam(niche.overlap.resid ~ 
+                    pollen +
+                    s(elev.mean, by = pollen, k = 10, m = 2),
+                  method = "REML",
+                  family = "gaussian",
+                  data = filter(comparison.morph, pbl.fac1 == "short" & pbl.fac2 == "short")) %>% getViz()
+
+check.gamViz(ST_nullmodel_GAM.morph)
+summary(ST_nullmodel_GAM.morph)
+print(plot(ST_nullmodel_GAM.morph, allTerms = TRUE), pages = 1)
+
+# Between ST and LT
+STxLT_nullmodel_GAM.tax <- gam(niche.overlap.resid ~ 
+                    pollen +
+                    s(elev.mean, by = pollen, k = 10, m = 2),
+                  method = "REML",
+                  family = "gaussian",
+                  data = filter(comparison.tax, pbl.fac.diff == "different")) %>% getViz()
+
+check.gamViz(STxLT_nullmodel_GAM.tax)
+summary(STxLT_nullmodel_GAM.tax)
+print(plot(STxLT_nullmodel_GAM.tax, allTerms = TRUE), pages = 1)
+
+STxLT_nullmodel_GAM.morph <- gam(niche.overlap.resid ~ 
+                    pollen +
+                    s(elev.mean, by = pollen, k = 10, m = 2),
+                  method = "REML",
+                  family = "gaussian",
+                  data = filter(comparison.morph, pbl.fac.diff == "different")) %>% getViz()
+
+check.gamViz(STxLT_nullmodel_GAM.morph)
+summary(STxLT_nullmodel_GAM.morph)
+print(plot(STxLT_nullmodel_GAM.morph, allTerms = TRUE), pages = 1)
 ```

### 2020-11-26 analyses/old_scripts/niche_v2.Rmd
did so many things

--- 2020-11-26 analyses/old_scripts/niche.Rmd
@@
 niche_overlap.sp <- net %>%
-  group_by(bb.sp, plant.sp, site, date, year) %>%
+  filter(!is.na(pollen)) %>%
+  group_by(bb.sp, pollen, plant.sp, site, date, year) %>%
   summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
-  group_by(bb.sp, site, date, year) %>%
+  group_by(bb.sp, pollen, site, date, year) %>%
   mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
   select(-freq) %>% # drop freq variable
+  #unite(ID, c(bb.sp, pollen), sep = "_"),
   pivot_wider(names_from = plant.sp, values_from = prop) %>% # spread into the kind of array vegan likes
   replace(is.na(.), 0) %>% # replace NAs with zeros
-  group_by(site, date, year) %>%
+  group_by(pollen, site, date, year) %>%
   nest() %>% # nest into list-column
@@
          tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
-  select(site, date, year, tidy_dist) %>% # drop intermediate list-columns
+  select(pollen, site, date, year, tidy_dist) %>% # drop intermediate list-columns
   unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
   left_join(site_data) %>% # add site data
-  select(site, date, year, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # select and rename vars
+  select(site, date, year, pollen, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # select and rename vars
   left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
@@
   rename(subgenus2 = subgenus, pbl.w2 = pbl.w, pbl.w.class2 = pbl.w.class) %>%
-  select(site, date, year, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
+  select(site, date, year, pollen, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
          pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
@@
 niche_overlap.sp.mean <- niche_overlap.sp %>%
-  group_by(sp.pair, site, year, elev.mean, pbl.diff, subgenus.diff, 
+  group_by(pollen, sp.pair, site, year, elev.mean, pbl.diff, subgenus.diff, 
            bss, pasc, wurf, prat, soro, hort, jone, psit, mont, hypn, 
@@
 niche_overlap.sp.mean2x <- niche_overlap.sp %>%
-  group_by(site, year, date, elev.mean) %>% # Average across species pairs within transects
+  group_by(pollen, site, year, date, elev.mean) %>% # Average across species pairs within transects
   summarize(mean.niche.overlap = mean(niche.overlap)) %>% 
-  group_by(site, year, elev.mean) %>% # Average across transects
+  group_by(pollen, site, year, elev.mean) %>% # Average across transects
   summarize(mean.niche.overlap = mean(mean.niche.overlap)) %>% 
-  left_join(BB_covars.mean) %>%
-  left_join(FL_covars.mean)
-
-### Genus-level by transect
-niche_overlap.gen <- net %>%
-  group_by(bb.sp, plant.genus, site, date, year) %>%
-  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
-  group_by(bb.sp, site, date, year) %>%
-  mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
-  select(-freq) %>% # drop freq variable
-  pivot_wider(names_from = plant.genus, values_from = prop) %>% # spread into the kind of array vegan likes
-  replace(is.na(.), 0) %>% # replace NAs with zeros
-  group_by(site, date, year) %>%
-  nest() %>% # nest into list-column
-  mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
-         dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
-         tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
-  select(site, date, year, tidy_dist) %>% # drop intermediate list-columns
-  unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
-  left_join(site_data) %>% # add site data
-  select(site, date, year, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # select and rename vars
-  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
-  rename(subgenus1 = subgenus, pbl.w1 = pbl.w, pbl.w.class1 = pbl.w.class) %>%
-  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
-  rename(subgenus2 = subgenus, pbl.w2 = pbl.w, pbl.w.class2 = pbl.w.class) %>%
-  select(site, date, year, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
-         pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
-  mutate(pbl.diff = abs(pbl.w1 - pbl.w2), # tongue length differences
-         yday = yday(date), # yday variable will come in handy
-         subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different"), # binary: same or different subgenus
-
-         # Logical columns indicating whether a given species belongs to a given species pair
-         # This is handy for extracting data for individual species across all their counterparts
-         bss = if_else(bb.sp1 == "bss" | bb.sp2 == "bss", TRUE, FALSE),
-         pasc = if_else(bb.sp1 == "pasc" | bb.sp2 == "pasc", TRUE, FALSE),
-         wurf = if_else(bb.sp1 == "wurf" | bb.sp2 == "wurf", TRUE, FALSE),
-         prat = if_else(bb.sp1 == "prat" | bb.sp2 == "prat", TRUE, FALSE),
-         soro = if_else(bb.sp1 == "soro" | bb.sp2 == "soro", TRUE, FALSE),
-         hort = if_else(bb.sp1 == "hort" | bb.sp2 == "hort", TRUE, FALSE),
-         jone = if_else(bb.sp1 == "jone" | bb.sp2 == "jone", TRUE, FALSE),
-         psit = if_else(bb.sp1 == "psit" | bb.sp2 == "psit", TRUE, FALSE),
-         mont = if_else(bb.sp1 == "mont" | bb.sp2 == "mont", TRUE, FALSE),
-         hypn = if_else(bb.sp1 == "hypn" | bb.sp2 == "hypn", TRUE, FALSE),
-         lapi = if_else(bb.sp1 == "lapi" | bb.sp2 == "lapi", TRUE, FALSE),
-         muci = if_else(bb.sp1 == "muci" | bb.sp2 == "muci", TRUE, FALSE),
-         gers = if_else(bb.sp1 == "gers" | bb.sp2 == "gers", TRUE, FALSE),
-         mend = if_else(bb.sp1 == "mend" | bb.sp2 == "mend", TRUE, FALSE),
-         pyre = if_else(bb.sp1 == "pyre" | bb.sp2 == "pyre", TRUE, FALSE),
-         humi = if_else(bb.sp1 == "humi" | bb.sp2 == "humi", TRUE, FALSE)) %>%
-
-  unite(sp.pair, c(bb.sp1, bb.sp2), sep = "_") %>% # create a single species pair column...
-  mutate(sp.pair = factor(sp.pair)) %>% # ...and make it a factor so that mgcv can handle it
-  left_join(BB_covars) %>% # add BB covars
-  left_join(FL_covars) %>% # add floral covars
-  mutate(niche.overlap = 1-distance) # express niche overlap so that 1 = perfect overlap and 0 = no overlap
-
-niche_overlap.gen.mean <- niche_overlap.gen %>%
-  group_by(sp.pair, site, year, elev.mean, pbl.diff, subgenus.diff, 
-           bss, pasc, wurf, prat, soro, hort, jone, psit, mont, hypn, 
-           lapi, muci, gers, mend, pyre, humi) %>%
-  summarize(mean.niche.overlap = mean(niche.overlap)) %>%
   left_join(BB_covars.mean) %>%
@@
 niche_overlap.ktype <- net %>%
+  filter(!is.na(pollen)) %>%
   left_join(fl_traits) %>%
-  group_by(bb.sp, k.type.s, site, date, year) %>%
+  group_by(bb.sp, pollen, k.type.s, site, date, year) %>%
   summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
-  group_by(bb.sp, site, date, year) %>%
+  group_by(bb.sp, pollen, site, date, year) %>%
   mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
   select(-freq) %>% # drop freq variable
+  #unite(ID, c(bb.sp, pollen), sep = "_"),
   pivot_wider(names_from = k.type.s, values_from = prop) %>% # spread into the kind of array vegan likes
   replace(is.na(.), 0) %>% # replace NAs with zeros
-  group_by(site, date, year) %>%
+  group_by(pollen, site, date, year) %>%
   nest() %>% # nest into list-column
@@
 niche_overlap.ktype.mean <- niche_overlap.ktype %>%
-  group_by(sp.pair, site, year, elev.mean, pbl.diff, subgenus.diff, 
+  group_by(pollen, sp.pair, site, year, elev.mean, pbl.diff, subgenus.diff, 
            bss, pasc, wurf, prat, soro, hort, jone, psit, mont, hypn, 
@@
   summarize(mean.niche.overlap = mean(niche.overlap)) %>%
+  left_join(BB_covars.mean) %>%
+  left_join(FL_covars.mean)
+
+niche_overlap.ktype.mean2x <- niche_overlap.ktype %>%
+  group_by(pollen, site, year, date, elev.mean) %>% # Average across species pairs within transects
+  summarize(mean.niche.overlap = mean(niche.overlap)) %>% 
+  group_by(pollen, site, year, elev.mean) %>% # Average across transects
+  summarize(mean.niche.overlap = mean(mean.niche.overlap)) %>% 
   left_join(BB_covars.mean) %>%
@@
 ```{r}
-ggplot(niche_overlap.sp, aes(year, niche.overlap)) +
-  geom_boxplot()
-
-ggplot(niche_overlap.gen, aes(year, niche.overlap)) +
+ggplot(niche_overlap.sp, aes(year, niche.overlap, color = pollen)) +
   geom_boxplot()
@@
 
-ggplot(niche_overlap.sp.mean, aes(year, mean.niche.overlap)) +
-  geom_boxplot()
-
-ggplot(niche_overlap.gen.mean, aes(year, mean.niche.overlap)) +
+ggplot(niche_overlap.sp.mean, aes(year, mean.niche.overlap, color = pollen)) +
   geom_boxplot()
@@
 
-ggplot(niche_overlap.sp.mean2x, aes(year, mean.niche.overlap)) +
-  geom_boxplot()
-
-ggplot(niche_overlap.gen.mean2x, aes(year, mean.niche.overlap)) +
+ggplot(niche_overlap.sp.mean2x, aes(year, mean.niche.overlap, color = pollen)) +
   geom_boxplot()
@@
 ```{r}
-ggplot(niche_overlap.sp, aes(subgenus.diff, niche.overlap)) +
+ggplot(niche_overlap.sp, aes(subgenus.diff, niche.overlap, color = pollen)) +
   geom_boxplot()
@@
 
-ggplot(niche_overlap.sp.mean, aes(subgenus.diff, mean.niche.overlap)) +
+ggplot(niche_overlap.sp.mean, aes(subgenus.diff, mean.niche.overlap, color = pollen)) +
   geom_boxplot()
@@
 ```{r}
-ggplot(niche_overlap.sp, aes(sqrt(pbl.diff), niche.overlap)) +
+ggplot(niche_overlap.sp, aes(sqrt(pbl.diff), niche.overlap, color = pollen)) +
   geom_point() +
@@
 
-ggplot(niche_overlap.sp.mean, aes(sqrt(pbl.diff), mean.niche.overlap)) +
+ggplot(niche_overlap.sp.mean, aes(sqrt(pbl.diff), mean.niche.overlap, color = pollen)) +
   geom_point() +
@@
 ```{r}
-ggplot(niche_overlap.sp, aes(elev.mean, niche.overlap)) +
+ggplot(niche_overlap.sp, aes(elev.mean, niche.overlap, color = pollen)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
+  #facet_wrap(~sp.pair) +
+  facet_wrap(~year) +
+  ylim(c(0,1))
+
+ggplot(niche_overlap.ktype, aes(elev.mean, niche.overlap, color = pollen)) +
   geom_point() +
@@
 
-ggplot(niche_overlap.gen, aes(elev.mean, niche.overlap)) +
+ggplot(niche_overlap.sp.mean, aes(elev.mean, mean.niche.overlap, color = pollen)) +
   geom_point() +
@@
 
-ggplot(niche_overlap.ktype, aes(elev.mean, niche.overlap)) +
+ggplot(niche_overlap.ktype.mean, aes(elev.mean, mean.niche.overlap, color = pollen)) +
   geom_point() +
@@
 
-ggplot(niche_overlap.sp.mean, aes(elev.mean, mean.niche.overlap)) +
+ggplot(niche_overlap.sp.mean2x, aes(elev.mean, mean.niche.overlap, color = pollen)) +
   geom_point() +
   geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
-  facet_wrap(~sp.pair) +
-  #facet_wrap(~year) +
-  ylim(c(0,1))
+  facet_wrap(~year)
 
-ggplot(niche_overlap.sp.mean2x, aes(elev.mean, mean.niche.overlap)) +
+ggplot(niche_overlap.ktype.mean2x, aes(elev.mean, mean.niche.overlap, color = pollen)) +
   geom_point() +
@@
 # Species level
-ggplot(niche_overlap.sp, aes(fl.rich, distance)) +
+ggplot(niche_overlap.sp, aes(fl.rich, niche.overlap, color = pollen)) +
   geom_point() +
@@
 ```{r}
-ggplot(niche_overlap.sp, aes(yday(date), distance)) +
+ggplot(niche_overlap.sp, aes(yday(date), niche.overlap, color = pollen)) +
   geom_point() +
@@
 ```{r}
-
 net_ord_genus_dist <- vegdist(net_ord_genus, method = "horn")

### 2020-11-26 analyses/old_scripts/resource_phenology.Rmd
did so many things

---
title: "floral_resources"
output: html_document
---

### Prepare environment
```{r setup, message=FALSE, warning=FALSE, include=FALSE}
library(ggforce)
library(vegan)
library(tidyverse)
library(lubridate)
library(tidymodels)
```

```{r}
# Network
net <- read_csv("../processed_data/network.csv") %>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  mutate(year = factor(year)) %>% # convert year to factor
  left_join(site_data, by = "site") %>%
  left_join(gdd, by = c("site", "date")) %>%
  ungroup() %>%
  arrange(site, date)

# Floral survey
# FL traits
fl_tax <- read_csv("../processed_data/floral_tax.csv")

fl_ktype <- read_csv("../processed_data/floral_k_type.csv") %>%
  select(plant.sp, plant.genus = genus, k.type) %>%
  mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"),
         k.type.ss = str_extract(k.type, "[-+]?[0-9]*"))

fl_morph <- read_csv("../processed_data/flower_morphology_GitaBenadi.csv") %>%
  select(SampleID, plant.family = Family, plant.sp = Species, 
         Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam),
            LN = mean(LN))

fl_ktype_key <- read_delim("../processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  select(k.type = 1, description = 2)

fl_traits <- fl_tax %>%
  full_join(fl_ktype) %>%
  full_join(fl_morph) %>%
  select(plant.sp, plant.genus, plant.family, 
         k.type, k.type.s, k.type.ss, diam, LN)

survey <- read_csv("../processed_data/floral_survey.csv") %>%
  group_by(site, date, plant.sp) %>%
  summarize(flower.cover = sum(flower.cover)) %>%
  group_by(site, date) %>%
  mutate(prop = flower.cover/sum(flower.cover)) %>%
  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover, prop) %>%
  mutate(yday = yday(date),
         year = year(date)) %>%
  left_join(fl_traits)

top_plants <- survey %>%
  select(site, date, plant.sp, prop) %>%
  complete(fill = list(prop = 0)) %>%
  group_by(plant.sp) %>%
  summarize(mean.prop = mean(prop))

```

```{r}
ggplot(survey, aes(month(date), flower.cover, fill = k.type.ss)) +
  geom_col(position = "fill") +
  facet_grid(year ~ site)
+
  theme(legend.position = "none")
```

### 2020-11-26 analyses/old_scripts/treemap.Rmd
did so many things

---
title: "trees"
output: html_document
---

```{r}
library(treemapify)
library(tidyverse)
```

# Survey data
```{r}
taxa <- read_csv("../processed_data/floral_tax.csv")

survey <- read_csv("../processed_data/floral_survey.csv") %>%
  filter(flower.cover > 0) %>%
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  separate(plant.sp, c("plant.genus", "plant.sp"), sep = " ") %>%
  left_join(taxa) %>%
  select(site, date, plant.family, plant.genus, plant.sp, flower.cover) %>%
  group_by(site, plant.family, plant.genus, plant.sp) %>%
  summarize(flower.cover = sum(flower.cover))
```

# Tree plot
```{r}
ggplot(survey, aes(area = flower.cover, fill = plant.family)) +
  geom_treemap() +
  theme(legend.position = "none") +
  facet_wrap(~site)

ggplot(survey, aes(area = flower.cover, fill = plant.family, label = plant.family,
                subgroup = plant.family)) +
  geom_treemap()
```

### 2020-11-26 analyses/plotting_filter.R

@@ -1,157 +1,0 @@
-### Need to run outside main script to avoid memory crashes
-
-library(tidyverse)
-library(lubridate)
-
-site_data <- read_csv("./processed_data/site_data.csv") %>%
-  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")), # rename in English
-         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))
-
-net <- read_csv("./processed_data/network.csv") %>%
-  dplyr::select(site, date, bb.sp, plant.sp) # drop unused columns
-
-survey <- read_csv("./processed_data/floral_survey.csv") %>%
-  semi_join(net, by = "plant.sp") %>% # Consider only visited plant species (anytime anywhere, not specifically for a given site*date)
-  dplyr::select(site, date, plant.sp, flower.cover) %>%
-  group_by(plant.sp) %>%
-  mutate(present = case_when(
-    max(flower.cover) > 0 ~ TRUE,
-    max(flower.cover) == 0 ~ FALSE
-  )) %>%
-  filter(present == TRUE) 
-
-### This could be so much more beautiful, but I don't have it in me right now. This works.
-seq_yday <- function(x){
-  seq((x - 10), (x + 10), 0.5)
-}
-
-seq_elev <- function(x){
-  seq((x - 150), (x + 150), 1)
-}
-
-df <- function(x) {
-  as.data.frame(x)
-}
-
-survey_range.2010 <- survey %>%
-  ungroup() %>%
-  left_join(site_data) %>%
-  mutate(yday = yday(date),
-         year = year(date)) %>%
-  filter(year == 2010) %>%
-  select(yday, elev.mean) %>%
-  unique() %>%
-  arrange(yday) %>%
-  mutate(ydays = map(yday, seq_yday),
-         elevs = map(elev.mean, seq_elev)) %>%
-  mutate(ydays = map(ydays, df),
-         elevs = map(elevs, df)) %>%
-  unnest(cols = ydays) %>%
-  select(yday = x, elevs) %>%
-  unnest(cols = elevs) %>%
-  unique() %>%
-  select(yday, elev.mean = x) %>%
-  mutate(year = rep(2010, n()))
-
-survey_range.2011 <- survey %>%
-  ungroup() %>%
-  left_join(site_data) %>%
[…]

### 2020-11-26 analyses/trash_can.R

@@ -1915,3 +1915,216 @@ met1 <- map(n1, specieslevel, level = "higher", index = "resource range") %>%
   bind_rows(.id = "id") %>%
   
   ```
+
+### Could we pool across years? That introduces a bit of pseudo replication, but I'm not sure it matters much. It would certainly simplify the workflow and presentation to pool across years, and it would correspond to the pooling I did in the simple visualizations of beta-diversity.
+```{r}
+###################
+### Interaction ###
+###################
+# Site_pair prep function
+site_pair_prep_net_pool <- function(dat) {
+  beta_mit_data %>%
+    ungroup() %>%
+    filter(year1 == year2 & 
+             site1 != site2) %>% # I don't think we want to compare samples through within sites, introduces a new form of non-independence.
+    filter(metric == "WN") %>% # we'll focus for now on WN, but this could be done for partitions, too
+    mutate(weights = rep(1, n())) %>% # right now this analysis is unweighted; could weight by abundance, etc, if we wanted to
+    dplyr::select(distance = value, weights,
+                  s1.xCoord = x1, s1.yCoord = y1, 
+                  s2.xCoord = x2, s2.yCoord = y2, 
+                  s1.elev.mean = elev1, s1.yday = yday1,
+                  s2.elev.mean = elev2, s2.yday = yday2) %>%
+    data.frame()
+}
+
+# Prep site-pairs for GDM analysis
+sitepair.WN <- site_pair_prep_net_pool(beta_mit_data)
+
+# Run GDM analysis
+gdm.WN <- gdm(sitepair.WN, geo = TRUE)
+summary(gdm.WN)
+
+# Permutation tests (slow)
+gdm.WN.imp <- gdm.varImp(sitepair.WN, geo = TRUE, 
+                         parallel = TRUE, cores = 2,
+                         outFile = "interaction_permGDM")
+
+# Plot nicely
+## wrapper for extracting splines and prepping tibble for plotting
+spline_prep_pool <- function(x) {
+  isplineExtract(x) %>%
+    as.data.frame() %>%
+    tibble() %>%
+    mutate(x.Geographic = x.Geographic/1000) %>%
+    pivot_longer(cols = c(x.Geographic, x.elev.mean, x.yday), 
[…]

### 2020-11-26 data_proc.R

@@ -98,9 +98,9 @@ site_data <- read_tsv("./raw_data/SiteLocationFabriceUpdate2020.tsv",
                 elev.min = Min_altitude, elev.max  = Max_altitude, lat = Lat, lon = Long) %>%
   mutate(elev.class = factor(elev.class, ordered = TRUE, levels = c("unten", "mitte", "oben"))) %>%
   mutate(elev.class2 = case_when(
-    elev.mean < 1000 ~ "low",
-    elev.mean >= 1000 & elev.mean < 1705 ~ "mid",
-    elev.mean >= 1750 ~ "high"
+    elev.class == "unten" ~ "low",
+    elev.class == "mitte" ~ "mid",
+    elev.class == "oben" ~ "high"
   ))
 
 write_csv(site_data, "./processed_data/site_data.csv")
@@ -111,9 +111,64 @@ floral_k_type <-  read_delim("./BioFlor_traits/BioFlor_Kugler_classification.csv
   na.omit() %>%
   mutate(taxon = str_replace_all(taxon, c(" x " = " x"))) %>%
   separate(taxon, c("genus", "species"), sep = " ") %>%
-  unite(plant.sp, genus, species, sep = " ", remove = FALSE)
+  unite(plant.sp, genus, species, sep = " ", remove = FALSE) %>%
+  select(plant.sp, plant.genus = genus, k.type) %>%
+  mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"), # make simplified (and more simplified) k types
+         k.type.ss = str_extract(k.type, "[-+]?[0-9]*")) %>%
   
+  # Fill in ktype data for plants present in my samples but missing from the Bioflor database
+  # In most cases
+  add_row(plant.sp = "Phyteuma orbiculare", plant.genus = "Phyteuma", 
+          k.type = "NA", k.type.s = "7.1", k.type.ss = "7") %>% # Following Neumayer and Paulus (1999)
+  add_row(plant.sp = "Phyteuma spicatum", plant.genus = "Phyteuma",
+          k.type = "NA", k.type.s = "7.1", k.type.ss = "7") %>% # Following Neumayer and Paulus (1999)
+  add_row(plant.sp = "Aconitum vulparia", plant.genus = "Aconitum",
+          k.type = "NA", k.type.s = "5.2", k.type.ss = "5") %>% # All scored Aconitum were classified as 5.2
[…]

### 2020-11-26 analyses/niche_relations_v4.Rmd
polished tree plots

@@
   mutate(Score = case_when( # rename default scores to match my data
-    Score == "species" ~ "FL.sp",
-    Score == "sites" ~ "BB.sp"))
+    Score == "species" ~ "FL sp.",
+    Score == "sites" ~ "BB sp."))
 }
@@
     geom_point(aes(MDS1, MDS2, color = Score), size = 3, alpha = 0.75) + # add points
-    ggrepel::geom_text_repel(aes(MDS1, MDS2, color = Score, label = Label), key_glyph = "rect") + # add labels
+    ggrepel::geom_text_repel(aes(MDS1, MDS2, color = Score, label = Label), key_glyph = "rect", size = 3) + # add labels
     coord_fixed() + # Gavin Simpson says this is **really** important
-    theme_classic() 
+    theme_classic() +
+    theme(legend.title=element_blank())
 }
@@
 ## Extract and name plots
-niche_oben_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "oben" & period == "early")$pcoa.plot[[1]] +
-  ggtitle("Oben/Early/Tax")
-niche_oben_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "oben" & period == "mid")$pcoa.plot[[1]] +
-  ggtitle("Oben/Mid/Tax")
-niche_oben_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "oben" & period == "late")$pcoa.plot[[1]] +
-  ggtitle("Oben/Late/Tax")
-
-niche_mitte_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "mitte" & period == "early")$pcoa.plot[[1]] +
-  ggtitle("Mitte/Early/Tax")
-niche_mitte_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "mitte" & period == "mid")$pcoa.plot[[1]] +
-  ggtitle("Mitte/Mid/Tax")
-niche_mitte_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "mitte" & period == "late")$pcoa.plot[[1]] +
-  ggtitle("Mitte/Late/Tax")
-
-niche_unten_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "unten" & period == "early")$pcoa.plot[[1]] +
-  ggtitle("Unten/Early/Tax")
-niche_unten_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "unten" & period == "mid")$pcoa.plot[[1]] +
-  ggtitle("Unten/Mid/Tax")
-niche_unten_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "unten" & period == "late")$pcoa.plot[[1]] +
-  ggtitle("Unten/Late/Tax")
+niche_oben_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "oben" & period == "early")$pcoa.plot[[1]] 
+niche_oben_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "oben" & period == "mid")$pcoa.plot[[1]] 
+niche_oben_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "oben" & period == "late")$pcoa.plot[[1]]
+
+niche_mitte_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "mitte" & period == "early")$pcoa.plot[[1]]
+niche_mitte_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "mitte" & period == "mid")$pcoa.plot[[1]] 
+niche_mitte_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "mitte" & period == "late")$pcoa.plot[[1]]
+
+niche_unten_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "unten" & period == "early")$pcoa.plot[[1]]
+niche_unten_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "unten" & period == "mid")$pcoa.plot[[1]]
+niche_unten_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "unten" & period == "late")$pcoa.plot[[1]]
 
@@
 The story here is that LT species are more abundant than ST species at lower elevations, and ST species are more abundant than LT species at higher elevations.
-This relationship is strong and basically linear, with the relative abundances "crossing" at
+This relationship is strong and basically linear, with the relative abundances crossing at mid-elevation.
 
@@
          bb.sp = factor(bb.sp, levels = c("hort", "pasc", "wurf", "bss", "prat", "soro"))) %>%
-  group_by(period, elev.class, plant.genus, k.type.ss, plant.family, bb.sp) %>%
+  group_by(period, elev.class, plant.sp.abb, plant.genus, k.type.ss, plant.family, bb.sp) %>%
   summarize(abund = n()) %>%
@@
 # Early
-ggplot(filter(visitation_abund.sp, period == "early"), 
-       aes(area = prop, fill = k.type.ss, label = plant.genus)) +
-  geom_treemap(color = "black") +
+tp.early <- ggplot(filter(visitation_abund.sp, period == "early"), 
+                aes(area = prop, fill = k.type.ss, label = plant.sp.abb)) +
+  geom_treemap(color = "white") +
   geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
   scale_fill_manual(values = cols) + 
+  labs(fill = "Morphotype") +
   facet_grid(fct_rev(elev.class)~bb.sp)
@@
 # Mid
-ggplot(filter(visitation_abund.sp, period == "mid"), 
-       aes(area = prop, fill = k.type.ss, label = plant.genus)) +
-  geom_treemap(color = "black") +
+tp.mid <- ggplot(filter(visitation_abund.sp, period == "mid"), 
+                 aes(area = prop, fill = k.type.ss, label = plant.sp.abb)) +
+  geom_treemap(color = "white") +
   geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
   scale_fill_manual(values = cols) + 
+  labs(fill = "Morphotype") +
   facet_grid(fct_rev(elev.class)~bb.sp)
@@
 # Late
-ggplot(filter(visitation_abund.sp, period == "late"), 
-       aes(area = prop, fill = k.type.ss, label = plant.genus)) +
-  geom_treemap(color = "black") +
+tp.late <- ggplot(filter(visitation_abund.sp, period == "late"), 
+                  aes(area = prop, fill = k.type.ss, label = plant.sp.abb)) +
+  geom_treemap(color = "white") +
   geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
   scale_fill_manual(values = cols) + 
+  labs(fill = "Morphotype") +
   facet_grid(fct_rev(elev.class)~bb.sp)
+
+pdf("./output/tree_ord_early.pdf", width = 14, height = 7)
+(tp.early | (niche_oben_early.tax.plot / niche_mitte_early.tax.plot / niche_unten_early.tax.plot) + plot_layout(guides = 'collect')) + plot_layout(widths = c(3, 1))
+dev.off()
+
+pdf("./output/tree_ord_mid.pdf", width = 14, height = 7)
+(tp.mid | (niche_oben_mid.tax.plot / niche_mitte_mid.tax.plot / niche_unten_mid.tax.plot) + plot_layout(guides = 'collect')) + plot_layout(widths = c(3, 1))
+dev.off()
+
+pdf("./output/tree_ord_late.pdf", width = 14, height = 7)
+(tp.late | (niche_oben_late.tax.plot / niche_mitte_late.tax.plot / niche_unten_late.tax.plot) + plot_layout(guides = 'collect')) + plot_layout(widths = c(3, 1))
+dev.off()
 

### 2020-11-30 analyses/niche_relations_v5.Rmd
update after conversation with Kalli

--- 2020-11-26 analyses/niche_relations_v4.Rmd
@@
   dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
-  mutate(elev.class = factor(elev.class, levels = c("unten", "mitte", "oben"))) # turn elav.class into an ordered factor
+  mutate(elev.class2 = factor(elev.class2, levels = c("low", "mid", "high"))) # turn elav.class into an ordered factor
 
@@
 fl_ktype <- read_csv("../processed_data/floral_k_type.csv")
+
+## Flower colors
+fl_color <- read_csv("../processed_data/floral_color.csv")
 
@@
   full_join(fl_ktype) %>%
+  full_join(fl_color) %>%
   full_join(fl_morph) %>%
-  dplyr::select(plant.sp, plant.genus, plant.family, k.type, k.type.s, k.type.ss, diam, LN)
+  dplyr::select(plant.sp, plant.genus, plant.family, k.type, k.type.s, k.type.ss, color, diam, LN)
 
@@
   left_join(fl_traits) %>% # add floral traits
-  group_by(elev.class) %>%
-  mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) # divide each elev.class into 3 equal time slices
+  #group_by(elev.class2) %>%
+  group_by(site) %>%
+  mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) # divide each elev.class2 into 3 equal time slices
 
@@
   left_join(fl_traits) # add floral traits
+
+missing_colors <- net %>%
+  bind_rows(survey) %>%
+  ungroup() %>%
+  select(plant.sp, color) %>%
+  filter(is.na(color)) %>%
+  distinct()
 ```
@@
 ggsave("./output/bb_range_year.pdf", height = 6, width = 6)
+
+# Do we see the bimodal abundance pattern predicted by the core-satellite hypothesis? Well, sort of.
+bb_abund <- net %>%
+  group_by(bb.sp) %>%
+  summarize(abund = n())
+
+ggplot(bb_abund, aes(abund)) +
+  geom_histogram(binwidth = 250)
+```
+
+And we will parse the samples by elevation class and season, dividing the samples from each site into equal-interval thirds
+```{r}
+samples <- net %>%
+  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
+  #group_by(elev.class2) %>%
+  group_by(site) %>%
+  mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
+  group_by(site, period, dayofyear) %>%
+  summarize(obs = n()) %>%
+  left_join(site_data) %>%
+  mutate(date = as_date(dayofyear))
+
+ggplot(samples, aes(as_date(yday(date)), elev.mean, color = period)) +
+  #geom_point(size = 1, aes(alpha = obs)) +
+  geom_point(size = 2, alpha = 0.75) +
+  #facet_wrap(~year) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_alpha_continuous(name = "Observations") +
+  scale_x_date(date_labels = "%b", breaks = as_date(c("1970-04-01", "1970-05-01", "1970-06-01",
+                                                    "1970-07-01", "1970-08-01", "1970-09-01"))) +
+  #geom_hline(yintercept = c(755, 1500), linetype = "dashed") +
+  theme_light(12) +
+  labs(x = "Time",
+       y = "Elevation (m x 10)") 
 ```
@@
 # Convert PCoA to tidy array ("fortify")
+# pcoa_fort <- function(x){
+#   fortify(x, axes = 1:2) %>% 
+#   mutate(Score = case_when( # rename default scores to match my data
+#     Score == "species" ~ "FL sp.",
+#     Score == "sites" ~ "BB sp."))
+# }
+
 pcoa_fort <- function(x){
@@
   mutate(Score = case_when( # rename default scores to match my data
-    Score == "species" ~ "FL.sp",
-    Score == "sites" ~ "BB.sp"))
+    Score == "species" ~ "flora",
+    #Label %in% c("hort", "pasc", "wurf") ~ "LT spp.",
+    Label %in% c("hort", "pasc") ~ "LT spp.",
+    Label == "wurf" ~ "robber",
+    Label %in% c("bss", "soro", "prat") ~ "ST spp."))
 }
@@
 
-  # Add T.class variable for the BB.sp subset of the PCoA
-  t.class <- x %>%
-    filter(Score == "BB.sp") %>%
-    mutate(T.class = factor(if_else(
-      Label %in% c("hort", "pasc", "wurf", "gers", "mend", "hypn"), "long", "short"))
-      )
-
-  ggplot(filter(x, (abs(MDS1) > min_score | abs(MDS2) > min_score) | (Score == "BB.sp"))) + # no threshold for BB spp
-    geom_polygon(data = t.class, aes(MDS1, MDS2, linetype = T.class), # draw polygons around T classes
-                 color = "gray70", fill = NA, size = 1.25, key_glyph = "abline") +
-    geom_point(aes(MDS1, MDS2, color = Score), size = 3, alpha = 0.75) + # add points
-    ggrepel::geom_text_repel(aes(MDS1, MDS2, color = Score, label = Label), key_glyph = "rect") + # add labels
+  # # Add T.class variable for the BB.sp subset of the PCoA
+  # t.class <- x %>%
+  #   filter(Score == "BB sp.") %>%
+  #   mutate(T.class = factor(if_else(
+  #     Label %in% c("hort", "pasc", "wurf", "gers", "mend", "hypn"), "long", "short"))
+  #     )
+
+  cols <- c("flora" = "gray40", "LT spp." = "#F8766D", "ST spp." = "#00BA38", "robber" = "#619CFF")
+
+  ggplot(filter(x, (abs(MDS1) > min_score | abs(MDS2) > min_score) | (Score %in% c("LT spp.", "ST spp.", "robber"))), 
+         aes(MDS1, MDS2, color = Score)) + # no threshold for BB spp
+    #geom_polygon(data = t.class, aes(MDS1, MDS2, linetype = T.class), # draw polygons around T classes
+     #            color = "gray70", fill = NA, size = 1.25, key_glyph = "abline") +
+    geom_point(size = 2, alpha = 0.75) + # add points
+    ggrepel::geom_text_repel(aes(label = Label), key_glyph = "rect", size = 3) + # add labels
     coord_fixed() + # Gavin Simpson says this is **really** important
-    theme_classic() 
+    theme_classic() +
+    theme(legend.title=element_blank()) +
+    scale_color_manual(values = cols)
 }
@@
 ## unpopulated, e.g. niche_analysis(net, bb_traits, plant.sp). 
-## To parse by elev.class: niche_analysis(net, bb_traits, plant.sp)
-## To parse by  elev.class, period, and pollen: niche_analysis(net, bb_traits, plant.sp, period, pollen)
+## To parse by elev.class2: niche_analysis(net, bb_traits, plant.sp)
+## To parse by  elev.class2, period, and pollen: niche_analysis(net, bb_traits, plant.sp, period, pollen)
 ## You get the idea. You should anyway -- you wrote it. https://tidyeval.tidyverse.org/multiple.html
@@
                         ),
-           adonis = map2(comm2, traits, # run adonis permuatation test for distance ~ tongue class
-                         function(x, y) adonis(x ~ pbl.fac, data = y, method = "horn")
-                         ),
-           betadis = map2(comm2, traits, # betadisper to test for difference in dispersion between tongue classes
-                         function(x, y) x %>%
-                           vegdist(method = "horn") %>% # betadisper runs a PCoA under the hood; match method to my PCoA
-                           betadisper(group = y$pbl.fac) 
-                         ),
-           betadis.permutest = map(betadis, permutest), # perumutation test
-           betadis.p = map(betadis.permutest, # extract betadisper p-value
-                           function(x) x$tab$'Pr(>F)'[[1]]
-                           ),
-           betadis.boxplot = map(betadis, boxplot), # boxplot of betadisper
+           # adonis = map2(comm2, traits, # run adonis permuatation test for distance ~ tongue class
+           #               function(x, y) adonis(x ~ pbl.fac, data = y, method = "horn")
+           #               ),
+           # betadis = map2(comm2, traits, # betadisper to test for difference in dispersion between tongue classes
+           #               function(x, y) x %>%
+           #                 vegdist(method = "horn") %>% # betadisper runs a PCoA under the hood; match method to my PCoA
+           #                 betadisper(group = y$pbl.fac) 
+           #               ),
+           # betadis.permutest = map(betadis, permutest), # perumutation test
+           # betadis.p = map(betadis.permutest, # extract betadisper p-value
+           #                 function(x) x$tab$'Pr(>F)'[[1]]
+           #                 ),
+           # betadis.boxplot = map(betadis, boxplot), # boxplot of betadisper
            pcoa.fort = map(pcoa, pcoa_fort), 
            pcoa.plot = map(pcoa.fort, pcoa_plot),
-           adonis.r2 = map(adonis, # extract adonis r2
-                           function(x) x$aov.tab$R2[[1]]
-                           ),
-           adonis.p = map(adonis, # extract adonis p-value
-                          function(x) x$aov.tab$'Pr(>F)'[[1]]
-                          ),
+           # adonis.r2 = map(adonis, # extract adonis r2
+           #                 function(x) x$aov.tab$R2[[1]]
+           #                 ),
+           # adonis.p = map(adonis, # extract adonis p-value
+           #                function(x) x$aov.tab$'Pr(>F)'[[1]]
+           #                ),
            MDS1 = map(pcoa, # extract variance explained by MDS1
@@
                       )) %>%
-    unnest(cols = c(adonis.r2, adonis.p, betadis.p, MDS1, MDS2)) %>% # these are lists of length = 1, so unnesting them simply displays the value
-    select(..., comm2, pcoa, betadis, betadis.permutest, betadis.p, adonis, 
-           adonis.r2, adonis.p, betadis.boxplot, pcoa.plot, MDS1, MDS2) 
+    #unnest(cols = c(adonis.r2, adonis.p, betadis.p, MDS1, MDS2)) %>% # these are lists of length = 1, so unnesting them simply displays the value
+    unnest(cols = c(MDS1, MDS2)) %>% # these are lists of length = 1, so unnesting them simply displays the value
+    # select(..., comm2, pcoa, betadis, betadis.permutest, betadis.p, adonis, 
+    #        adonis.r2, adonis.p, betadis.boxplot, pcoa.plot, MDS1, MDS2) 
+    select(..., comm2, pcoa, pcoa.plot, MDS1, MDS2)
 }
@@
 argument will also group by any variables supplied in the function call, e.g.
-elev.class, period, year.
+elev.class2, period, year.
     summarize(freq = n()) %>% # get visitation frequency
@@
     left_join(y) %>%
-    group_by(elev.class) %>%
-    mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
+    #group_by(elev.class2) %>%
+    group_by(site) %>%
+    mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
     group_by(site, period, pbl.fac, !!criterion) %>% # group by site, date, bee species, plant species
@@
     filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
-    group_by(elev.class) %>%
-    mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
+    #group_by(elev.class2) %>%
+    group_by(site) %>%
+    mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
     left_join(bb_traits) %>%
@@
     left_join(y) %>%
-    # group_by(elev.class) %>%
-    # mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
+    # group_by(elev.class2) %>%
+    # mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
     #group_by(site, period, pbl.fac, !!criterion) %>% # group by site, date, bee species, plant species
@@
     filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
-    # group_by(elev.class) %>%
-    # mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
+    # group_by(elev.class2) %>%
+    # mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
     left_join(bb_traits) %>%
@@
 # Taxonomic
-niche_x_elev.tax <- niche_analysis(net, bb_traits, criterion = plant.sp.abb, elev.class)
+niche_x_elev.tax <- niche_analysis(net, bb_traits, criterion = plant.sp.abb, elev.class2)
 
 ## Extract and name plots
-niche_oben.tax.plot <- filter(niche_x_elev.tax, elev.class == "oben")$pcoa.plot[[1]] + ggtitle("Oben/Tax")
-niche_mitte.tax.plot <- filter(niche_x_elev.tax, elev.class == "mitte")$pcoa.plot[[1]] + ggtitle("Mitte/Tax")
-niche_unten.tax.plot <- filter(niche_x_elev.tax, elev.class == "unten")$pcoa.plot[[1]] + ggtitle("Unten/Tax")
+niche_high.tax.plot <- filter(niche_x_elev.tax, elev.class2 == "high")$pcoa.plot[[1]] + ggtitle("high/Tax")
+niche_mid.tax.plot <- filter(niche_x_elev.tax, elev.class2 == "mid")$pcoa.plot[[1]] + ggtitle("mid/Tax")
+niche_low.tax.plot <- filter(niche_x_elev.tax, elev.class2 == "low")$pcoa.plot[[1]] + ggtitle("low/Tax")
 
 # Morphological
-niche_x_elev.morph <- niche_analysis(net, bb_traits, criterion = k.type.s, elev.class)
+niche_x_elev.morph <- niche_analysis(net, bb_traits, criterion = k.type.s, elev.class2)
 
 ## Extract and name plots
-niche_oben.morph.plot <- filter(niche_x_elev.morph, elev.class == "oben")$pcoa.plot[[1]] + ggtitle("Oben/morph")
-niche_mitte.morph.plot <- filter(niche_x_elev.morph, elev.class == "mitte")$pcoa.plot[[1]] + ggtitle("Mitte/morph")
-niche_unten.morph.plot <- filter(niche_x_elev.morph, elev.class == "unten")$pcoa.plot[[1]] + ggtitle("Unten/morph")
+niche_high.morph.plot <- filter(niche_x_elev.morph, elev.class2 == "high")$pcoa.plot[[1]] + ggtitle("high/morph")
+niche_mid.morph.plot <- filter(niche_x_elev.morph, elev.class2 == "mid")$pcoa.plot[[1]] + ggtitle("mid/morph")
+niche_low.morph.plot <- filter(niche_x_elev.morph, elev.class2 == "low")$pcoa.plot[[1]] + ggtitle("low/morph")
 
@@
 pdf("../analyses/output/niche_x_elev_pcoa.pdf", height = 8.5, width = 11)
-(niche_oben.tax.plot | niche_oben.morph.plot) /
-(niche_mitte.tax.plot | niche_mitte.morph.plot) /
-(niche_unten.tax.plot | niche_unten.morph.plot) + plot_layout(guides = 'collect') 
+(niche_high.tax.plot | niche_high.morph.plot) /
+(niche_mid.tax.plot | niche_mid.morph.plot) /
+(niche_low.tax.plot | niche_low.morph.plot) + plot_layout(guides = 'collect') 
 dev.off()
@@
 
-## 1.3 Ordination by BB species * elev.class * period, pooling across years
+## 1.3 Ordination by BB species * elev.class2 * period, pooling across years
 ### Script
@@
 # Taxonomic
-niche_x_elevperiod.tax <- niche_analysis(net, bb_traits, criterion = plant.sp.abb, elev.class, period)
+niche_x_elevperiod.tax <- niche_analysis(net, bb_traits, criterion = plant.sp.abb, elev.class2, period)
 
 ## Extract and name plots
-niche_oben_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "oben" & period == "early")$pcoa.plot[[1]] +
-  ggtitle("Oben/Early/Tax")
-niche_oben_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "oben" & period == "mid")$pcoa.plot[[1]] +
-  ggtitle("Oben/Mid/Tax")
-niche_oben_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "oben" & period == "late")$pcoa.plot[[1]] +
-  ggtitle("Oben/Late/Tax")
-
-niche_mitte_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "mitte" & period == "early")$pcoa.plot[[1]] +
-  ggtitle("Mitte/Early/Tax")
-niche_mitte_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "mitte" & period == "mid")$pcoa.plot[[1]] +
-  ggtitle("Mitte/Mid/Tax")
-niche_mitte_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "mitte" & period == "late")$pcoa.plot[[1]] +
-  ggtitle("Mitte/Late/Tax")
-
-niche_unten_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "unten" & period == "early")$pcoa.plot[[1]] +
-  ggtitle("Unten/Early/Tax")
-niche_unten_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "unten" & period == "mid")$pcoa.plot[[1]] +
-  ggtitle("Unten/Mid/Tax")
-niche_unten_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "unten" & period == "late")$pcoa.plot[[1]] +
-  ggtitle("Unten/Late/Tax")
+niche_high_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "high" & period == "early")$pcoa.plot[[1]] 
+niche_high_mid1.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "high" & period == "mid1")$pcoa.plot[[1]] 
+niche_high_mid2.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "high" & period == "mid2")$pcoa.plot[[1]] 
+niche_high_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "high" & period == "late")$pcoa.plot[[1]]
+
+niche_mid_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "mid" & period == "early")$pcoa.plot[[1]]
+niche_mid_mid1.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "mid" & period == "mid1")$pcoa.plot[[1]] 
+niche_mid_mid2.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "mid" & period == "mid2")$pcoa.plot[[1]]
+niche_mid_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "mid" & period == "late")$pcoa.plot[[1]]
+
+niche_low_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "low" & period == "early")$pcoa.plot[[1]]
+niche_low_mid1.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "low" & period == "mid1")$pcoa.plot[[1]]
+niche_low_mid2.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "low" & period == "mid2")$pcoa.plot[[1]]
+niche_low_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "low" & period == "late")$pcoa.plot[[1]]
 
 # Morphological
-niche_x_elevperiod.morph <- niche_analysis(net, bb_traits, criterion = k.type.s, elev.class, period)
+niche_x_elevperiod.morph <- niche_analysis(net, bb_traits, criterion = k.type.s, elev.class2, period)
 
 ## Extract and name plots
-niche_oben_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "oben" & period == "early")$pcoa.plot[[1]] +
-  ggtitle("Oben/Early/Morph")
-niche_oben_mid.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "oben" & period == "mid")$pcoa.plot[[1]] +
-  ggtitle("Oben/Mid/Morph")
-niche_oben_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "oben" & period == "late")$pcoa.plot[[1]] +
-  ggtitle("Oben/Late/Morph")
-
-niche_mitte_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "mitte" & period == "early")$pcoa.plot[[1]] +
-  ggtitle("Mitte/Early/Morph")
-niche_mitte_mid.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "mitte" & period == "mid")$pcoa.plot[[1]] +
-  ggtitle("Mitte/Mid/Morph")
-niche_mitte_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "mitte" & period == "late")$pcoa.plot[[1]] +
-  ggtitle("Mitte/Late/Morph")
-
-niche_unten_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "unten" & period == "early")$pcoa.plot[[1]] +
-  ggtitle("Unten/Early/Morph")
-niche_unten_mid.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "unten" & period == "mid")$pcoa.plot[[1]] +
-  ggtitle("Unten/Mid/Morph")
-niche_unten_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "unten" & period == "late")$pcoa.plot[[1]] +
-  ggtitle("Unten/Late/Morph")
+niche_high_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "high" & period == "early")$pcoa.plot[[1]]
+niche_high_mid1.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "high" & period == "mid1")$pcoa.plot[[1]] 
+niche_high_mid2.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "high" & period == "mid2")$pcoa.plot[[1]] 
+niche_high_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "high" & period == "late")$pcoa.plot[[1]]
+
+niche_mid_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "mid" & period == "early")$pcoa.plot[[1]]
+niche_mid_mid1.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "mid" & period == "mid1")$pcoa.plot[[1]]
+niche_mid_mid2.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "mid" & period == "mid2")$pcoa.plot[[1]] 
+niche_mid_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "mid" & period == "late")$pcoa.plot[[1]]
+
+niche_low_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "low" & period == "early")$pcoa.plot[[1]]
+niche_low_mid1.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "low" & period == "mid1")$pcoa.plot[[1]] 
+niche_low_mid2.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "low" & period == "mid2")$pcoa.plot[[1]] 
+niche_low_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "low" & period == "late")$pcoa.plot[[1]]
 
@@
 #pdf("../analyses/output/niche_x_elevperiod_pcoa_tax.pdf", height = 8.5, width = 11)
-(niche_oben_early.tax.plot | niche_oben_mid.tax.plot  | niche_oben_late.tax.plot) /
-(niche_mitte_early.tax.plot | niche_mitte_mid.tax.plot | niche_mitte_late.tax.plot) /
-(niche_unten_early.tax.plot | niche_unten_mid.tax.plot  | niche_unten_late.tax.plot)  + plot_layout(guides = 'collect') 
+(niche_high_early.tax.plot | niche_high_mid1.tax.plot  | niche_high_mid2.tax.plot | niche_high_late.tax.plot) /
+(niche_mid_early.tax.plot | niche_mid_mid1.tax.plot | niche_mid_mid2.tax.plot | niche_mid_late.tax.plot) /
+(niche_low_early.tax.plot | niche_low_mid1.tax.plot  | niche_low_mid2.tax.plot  | niche_low_late.tax.plot)  + 
+  plot_layout(guides = 'collect') 
 #dev.off()
@@
 #pdf("../analyses/output/niche_x_elevperiod_pcoa_morph.pdf", height = 8.5, width = 11)
-(niche_oben_early.morph.plot | niche_oben_mid.morph.plot  | niche_oben_late.morph.plot) /
-(niche_mitte_early.morph.plot | niche_mitte_mid.morph.plot | niche_mitte_late.morph.plot) /
-(niche_unten_early.morph.plot | niche_unten_mid.morph.plot  | niche_unten_late.morph.plot)  + plot_layout(guides = 'collect') 
+(niche_high_early.morph.plot | niche_high_mid1.morph.plot  | niche_high_mid2.morph.plot | niche_high_late.morph.plot) /
+(niche_mid_early.morph.plot | niche_mid_mid1.morph.plot | niche_mid_mid2.morph.plot | niche_mid_late.morph.plot) /
+(niche_low_early.morph.plot | niche_low_mid1.morph.plot  | niche_low_mid2.morph.plot  | niche_low_late.morph.plot)  + 
+  plot_layout(guides = 'collect') 
 #dev.off()
@@
 ST species are all clustered around the bell flower *Erica carnea* (Kugler 3.1), while the LT species are more spread out, with *B.
-wurflenii* committed to the flag blossom *Polygala chamaebuxus*, *B.
+wurflenii* commidd to the flag blossom *Polygala chamaebuxus*, *B.
 hortorum* more interested in the stalk-disk flower *Gentiana verna* (Kugler 4.1), and *B.
@@
   filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
-  group_by(elev.class) %>%
-  mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
+  #group_by(elev.class2) %>%
+  group_by(site) %>%
+  mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
   ungroup() %>%
@@
   filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
-  group_by(elev.class) %>%
-  mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
+  #group_by(elev.class2) %>%
+  group_by(site) %>%
+  mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
   mutate(k.type.ss = factor(k.type.ss)) %>%
@@
   filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
-  group_by(elev.class) %>%
-  mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
+  #group_by(elev.class2) %>%
+  group_by(site) %>%
+  mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
   mutate(k.type.ss = factor(k.type.ss),
          bb.sp = factor(bb.sp, levels = c("hort", "pasc", "wurf", "bss", "prat", "soro"))) %>%
-  group_by(period, elev.class, plant.genus, k.type.ss, plant.family, bb.sp) %>%
+  group_by(period, elev.class2, plant.genus, k.type.ss, plant.family, bb.sp) %>%
   summarize(abund = n()) %>%
-  group_by(elev.class, period, bb.sp) %>%
+  group_by(elev.class2, period, bb.sp) %>%
   mutate(prop = abund/sum(abund)) %>%
-  arrange(elev.class, period, bb.sp, -prop)
+  arrange(elev.class2, period, bb.sp, -prop)
 
@@
 ```{r}
-### Visitation by BB species, parsed by elev.class and period
+### Visitation by BB species, parsed by elev.class2 and period
 visitation_abund.sp <- net %>%
   filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
-  group_by(elev.class) %>%
-  mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
+  #group_by(elev.class2) %>%
+  group_by(site) %>%
+  mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
   mutate(k.type.ss = factor(k.type.ss),
          bb.sp = factor(bb.sp, levels = c("hort", "pasc", "wurf", "bss", "prat", "soro"))) %>%
-  group_by(period, elev.class, plant.genus, k.type.ss, plant.family, bb.sp) %>%
+  group_by(period, elev.class2, plant.sp.abb, plant.genus, k.type.ss, plant.family, bb.sp) %>%
   summarize(abund = n()) %>%
-  group_by(elev.class, period, bb.sp) %>%
+  group_by(elev.class2, period, bb.sp) %>%
   mutate(prop = abund/sum(abund)) %>%
-  arrange(elev.class, period, bb.sp, -prop)
+  arrange(elev.class2, period, bb.sp, -prop)
 
@@
 # Early
-ggplot(filter(visitation_abund.sp, period == "early"), 
-       aes(area = prop, fill = k.type.ss, label = plant.genus)) +
-  geom_treemap(color = "black") +
+tp.early <- ggplot(filter(visitation_abund.sp, period == "early"), 
+                aes(area = prop, fill = k.type.ss, label = plant.sp.abb)) +
+  geom_treemap(color = "white") +
   geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
   scale_fill_manual(values = cols) + 
-  facet_grid(fct_rev(elev.class)~bb.sp)
+  labs(fill = "Morphotype") +
+  facet_grid(fct_rev(elev.class2)~bb.sp) +
+  theme(strip.text = element_text(size = 14))
 
 # Mid
-ggplot(filter(visitation_abund.sp, period == "mid"), 
-       aes(area = prop, fill = k.type.ss, label = plant.genus)) +
-  geom_treemap(color = "black") +
+tp.mid1 <- ggplot(filter(visitation_abund.sp, period == "mid1"), 
+                 aes(area = prop, fill = k.type.ss, label = plant.sp.abb)) +
+  geom_treemap(color = "white") +
   geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
   scale_fill_manual(values = cols) + 
-  facet_grid(fct_rev(elev.class)~bb.sp)
-
-# Late
-ggplot(filter(visitation_abund.sp, period == "late"), 
-       aes(area = prop, fill = k.type.ss, label = plant.genus)) +
-  geom_treemap(color = "black") +
+  labs(fill = "Morphotype") +
+  facet_grid(fct_rev(elev.class2)~bb.sp) +
+  theme(strip.text = element_text(size = 14))
+
+tp.mid2 <- ggplot(filter(visitation_abund.sp, period == "mid2"), 
+                 aes(area = prop, fill = k.type.ss, label = plant.sp.abb)) +
+  geom_treemap(color = "white") +
   geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
   scale_fill_manual(values = cols) + 
-  facet_grid(fct_rev(elev.class)~bb.sp)
-
+  labs(fill = "Morphotype") +
+  facet_grid(fct_rev(elev.class2)~bb.sp) +
+  theme(strip.text = element_text(size = 14))
+
+# Late
+tp.late <- ggplot(filter(visitation_abund.sp, period == "late"), 
+                  aes(area = prop, fill = k.type.ss, label = plant.sp.abb)) +
+  geom_treemap(color = "white") +
+  geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
+  scale_fill_manual(values = cols) + 
+  labs(fill = "Morphotype") +
+  facet_grid(fct_rev(elev.class2)~bb.sp) +
+  theme(strip.text = element_text(size = 14))
+
+# Pair with ordination plots
+tp_ord1 <- (
+  tp.early | (
+    niche_high_early.tax.plot / 
+    niche_mid_early.tax.plot / 
+    niche_low_early.tax.plot
+    ) + 
+    plot_layout(guides = 'collect')
+  ) + 
+  plot_layout(widths = c(5, 1))
+ggsave("./output/tree_ord_early.png", tp_ord1, width = 14, height = 8)
+
+tp_ord2 <- (
+  tp.mid1 | (
+    niche_high_mid1.tax.plot / 
+    niche_mid_mid1.tax.plot / 
+    niche_low_mid1.tax.plot
+    ) + 
+    plot_layout(guides = 'collect')
+  ) + 
+  plot_layout(widths = c(5, 1))
+ggsave("./output/tree_ord_mid1.png", tp_ord2, width = 14, height = 8)
+
+tp_ord3 <- (
+  tp.mid2 | (
+    niche_high_mid2.tax.plot / 
+    niche_mid_mid2.tax.plot / 
+    niche_low_mid2.tax.plot
+    ) + 
+    plot_layout(guides = 'collect')
+  ) + 
+  plot_layout(widths = c(5, 1))
+ggsave("./output/tree_ord_mid2.png", tp_ord3, width = 14, height = 8)
+
+tp_ord4 <- (
+  tp.late | (
+    niche_high_late.tax.plot / 
+    niche_mid_late.tax.plot / 
+    niche_low_late.tax.plot
+    ) + 
+    plot_layout(guides = 'collect')
+  ) + 
+  plot_layout(widths = c(5, 1))
+ggsave("./output/tree_ord_late.png", tp_ord4, width = 14, height = 8)
 ```
@@
 # By plant.sp
-ggplot(niche_x_elevperiod.tax, aes(elev.class, adonis.r2)) +
+ggplot(niche_x_elevperiod.tax, aes(elev.class2, adonis.r2)) +
   geom_boxplot()
@@
 # By ktype
-ggplot(niche_x_elevperiod.morph, aes(elev.class, adonis.r2)) +
+ggplot(niche_x_elevperiod.morph, aes(elev.class2, adonis.r2)) +
   geom_boxplot()

### 2020-11-30 data_proc.R

@@ -98,9 +98,9 @@ site_data <- read_tsv("./raw_data/SiteLocationFabriceUpdate2020.tsv",
                 elev.min = Min_altitude, elev.max  = Max_altitude, lat = Lat, lon = Long) %>%
   mutate(elev.class = factor(elev.class, ordered = TRUE, levels = c("unten", "mitte", "oben"))) %>%
   mutate(elev.class2 = case_when(
-    elev.class == "unten" ~ "low",
-    elev.class == "mitte" ~ "mid",
-    elev.class == "oben" ~ "high"
+    elev.mean <= 850 ~ "low",
+    elev.mean > 850 & elev.mean <= 1500 ~ "mid",
+    elev.mean > 1500 ~ "high"
   ))
 
 write_csv(site_data, "./processed_data/site_data.csv")
@@ -117,7 +117,6 @@ floral_k_type <-  read_delim("./BioFlor_traits/BioFlor_Kugler_classification.csv
          k.type.ss = str_extract(k.type, "[-+]?[0-9]*")) %>%
   
   # Fill in ktype data for plants present in my samples but missing from the Bioflor database
-  # In most cases
   add_row(plant.sp = "Phyteuma orbiculare", plant.genus = "Phyteuma", 
           k.type = "NA", k.type.s = "7.1", k.type.ss = "7") %>% # Following Neumayer and Paulus (1999)
   add_row(plant.sp = "Phyteuma spicatum", plant.genus = "Phyteuma",
@@ -168,7 +167,37 @@ floral_m_type <-  read_delim("./BioFlor_traits/BioFlor_Mueller_classification.cs
   unite(plant.sp, genus, species, sep = " ", remove = FALSE) %>%
   select(plant.sp, plant.genus = genus, m.type) 
 
-#write_csv(floral_k_type, "./processed_data/floral_m_type.csv")
+floral_color <-  read_delim("./BioFlor_traits/BioFlor_flower_color.csv", delim = ";") %>%
+  slice(-1) %>%
+  select(taxon = 1, color = 2) %>% 
+  na.omit() %>%
+  head(-1) %>%
+  mutate(taxon = str_replace_all(taxon, c(" x " = " x"))) %>%
+  separate(taxon, c("genus", "species"), sep = " ") %>%
+  unite(plant.sp, genus, species, sep = " ", remove = FALSE) %>%
+  select(plant.sp, plant.genus = genus, color) %>%
+  
+  # Fill in ktype data for plants present in my samples but missing from the Bioflor database
[…]

### 2020-11-30 analyses/niche_relations_v5.Rmd
getting ready to break everything

@@
   x %>%
+    filter(caste != "male") %>%
     filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
@@
   x %>% 
+    filter(caste != "male") %>%
     filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
@@
   x %>%
+    filter(caste != "male") %>%
     filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
@@
   a <- x %>%
+    filter(caste != "male") %>%
     filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
@@
   b <- x %>%
+    filter(caste != "male") %>%
     filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
@@
 
-  out <- left_join(a, b)
+  filter(caste != "male") %>%out <- left_join(a, b)
 }
@@
 visitation_abund.sp <- net %>%
-  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
+  filter(caste != "male") %>%
+  #filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
   #group_by(elev.class2) %>%
@@
   mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
+  # mutate(k.type.ss = factor(k.type.ss),
+  #        color = factor(color),
+  #        bb.sp = factor(bb.sp, levels = c("hort", "pasc", "wurf", "bss", "prat", "soro"))) %>%
   mutate(k.type.ss = factor(k.type.ss),
-         bb.sp = factor(bb.sp, levels = c("hort", "pasc", "wurf", "bss", "prat", "soro"))) %>%
-  group_by(period, elev.class2, plant.sp.abb, plant.genus, k.type.ss, plant.family, bb.sp) %>%
+         color = factor(color),
+         bb.sp = factor(bb.sp)) %>%
+  group_by(period, elev.class2, plant.sp.abb, plant.genus, k.type.ss, color, plant.family, bb.sp) %>%
   summarize(abund = n()) %>%
@@
 # Color palette
+
 cols = c("0" = "gray60", "1" = "#8DD3C7", "2" = "#FFFFB3", "3" = "#BEBADA",
@@
          "8" = "#FCCDE5", "9" = "#D9D9D9")
+
+cols2 = c("blue" = "cornflowerblue", 
+          "yellow" = "gold1", 
+          "pink" = "orchid2", 
+          "red" = "firebrick1",
+         "violet" = "mediumpurple2", 
+         "purple" = "purple", 
+         "white" = "white", 
+         "various colors" = "black",
+         "brown" = "sienna",
+         "undet" = "gray40", 
+         "green" = "olivedrab1", 
+         "orange" = "orange", 
+         "lilac" = "mediumorchid")
 
@@
 
+tp.early.col <- ggplot(filter(visitation_abund.sp, period == "early"), 
+                aes(area = prop, fill = color, label = plant.sp.abb)) +
+  geom_treemap(color = "white") +
+  geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
+  scale_fill_manual(values = cols2) + 
+  labs(fill = "Morphotype") +
+  facet_grid(fct_rev(elev.class2)~bb.sp) +
+  theme(strip.text = element_text(size = 14))
+
 # Mid
@@
 
+tp.mid1.col <- ggplot(filter(visitation_abund.sp, period == "mid1"), 
+                 aes(area = prop, fill = color, label = plant.sp.abb)) +
+  geom_treemap(color = "white") +
+  geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
+  scale_fill_manual(values = cols2) + 
+  labs(fill = "Morphotype") +
+  facet_grid(fct_rev(elev.class2)~bb.sp) +
+  theme(strip.text = element_text(size = 14))
+
 tp.mid2 <- ggplot(filter(visitation_abund.sp, period == "mid2"), 
@@
 
+tp.mid2.col <- ggplot(filter(visitation_abund.sp, period == "mid2"), 
+                 aes(area = prop, fill = color, label = plant.sp.abb)) +
+  geom_treemap(color = "white") +
+  geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
+  scale_fill_manual(values = cols2) + 
+  labs(fill = "Morphotype") +
+  facet_grid(fct_rev(elev.class2)~bb.sp) +
+  theme(strip.text = element_text(size = 14))
+
 # Late
@@
   scale_fill_manual(values = cols) + 
+  labs(fill = "Morphotype") +
+  facet_grid(fct_rev(elev.class2)~bb.sp) +
+  theme(strip.text = element_text(size = 14))
+
+tp.late.col <- ggplot(filter(visitation_abund.sp, period == "late"), 
+                  aes(area = prop, fill = color, label = plant.sp.abb)) +
+  geom_treemap(color = "white") +
+  geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
+  scale_fill_manual(values = cols2) + 
   labs(fill = "Morphotype") +

### 2020-11-30 data_proc.R

@@ -179,22 +179,32 @@ floral_color <-  read_delim("./BioFlor_traits/BioFlor_flower_color.csv", delim =
   
   # Fill in ktype data for plants present in my samples but missing from the Bioflor database
   # Colors based on review of iNaturalist "research-grade" observations. Where color is too variable to classify, "various_colors". 
-  # Where essentially absent as in conifers, "undet".
+  # Where essentially absent as in conifers, "NA".
   add_row(plant.sp = "Euphrasia rostkoviana", plant.genus = "Euphrasia", color = "various colors") %>%
   add_row(plant.sp = "Euphrasia officinalis", plant.genus = "Euphrasia", color = "various colors") %>%
   add_row(plant.sp = "Taraxacum officinale", plant.genus = "Taraxacum", color = "yellow") %>%
-  add_row(plant.sp = "Juniperus communis", plant.genus = "Juniperus", color = "undet") %>%
+  add_row(plant.sp = "Juniperus communis", plant.genus = "Juniperus", color = "NA") %>%
   add_row(plant.sp = "Gentiana aspera", plant.genus = "Gentiana", color = "purple") %>%
   add_row(plant.sp = "Gentianella aspera", plant.genus = "Gentianella", color = "purple") %>%
-  add_row(plant.sp = "Carex ericetorum", plant.genus = "Carex", color = "undet") %>%
+  add_row(plant.sp = "Carex ericetorum", plant.genus = "Carex", color = "NA") %>%
   add_row(plant.sp = "Euphrasia picta", plant.genus = "Euphrasia", color = "various colors") %>%
   add_row(plant.sp = "Gentiana ciliata", plant.genus = "Gentiana", color = "violet") %>%
   add_row(plant.sp = "Gentianopsis ciliata", plant.genus = "Gentiana", color = "violet") %>%
   add_row(plant.sp = "Stachys alopecuros", plant.genus = "Stachys", color = "yellow") %>%
   add_row(plant.sp = "Betonica alopecuros", plant.genus = "Betonica", color = "yellow") %>%
-  
-  
-  
+  add_row(plant.sp = "Cardamine enneaphyllos", plant.genus = "Cardamine", color = "yellow") %>%
+  add_row(plant.sp = "Lamium galeobdolon", plant.genus = "Lamium", color = "yellow") %>%
[…]

### 2020-11-30 analyses/niche_relations_v6.Rmd
went ordi surfing

--- 2020-11-30 analyses/niche_relations_v5.Rmd
@@
 ---
-title: "Dietary niche relationships between core bumble bee species in the Berchtesgadener Alps"
+title: "Floral resource choice by bumble bee species in the Berchtesgadener Alps"
 output: html_document
@@
 library(treemapify)
-library(RColorBrewer)
+#library(RColorBrewer)
 library(tidyverse)
@@
 # BB traits
-bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
-  mutate(pbl.fac = factor(if_else(
-    bb.sp %in% c("hort", "pasc", "wurf", "gers", "mend", "hypn"), "long", "short"))
-    )
+bb_traits <- read_csv("../processed_data/bb_traits.csv")
 
@@
 
-## Flower colors
+## Flower colors (accessed through BioFlor)
 fl_color <- read_csv("../processed_data/floral_color.csv")
@@
 
-## join all trait tables
+## join all floral trait tables
 fl_traits <- fl_tax %>%
@@
   left_join(fl_traits) %>% # add floral traits
-  #group_by(elev.class2) %>%
   group_by(site) %>%
-  mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) # divide each elev.class2 into 3 equal time slices
+  mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) # divide each elev.class2 into 4 equal time slices
 
@@
   left_join(fl_traits) # add floral traits
-
-missing_colors <- net %>%
-  bind_rows(survey) %>%
-  ungroup() %>%
-  select(plant.sp, color) %>%
-  filter(is.na(color)) %>%
-  distinct()
 ```
 
-We will focus on six species that are abundant and widespread:  
+Core species:  
   1. *B. sensu-strictu*
@@
 
-And we will parse the samples by elevation class and season, dividing the samples from each site into equal-interval thirds
+We will parse the samples by elevation class and season, dividing the samples from each site into equal-interval quarters
 ```{r}
@@
                                                     "1970-07-01", "1970-08-01", "1970-09-01"))) +
-  #geom_hline(yintercept = c(755, 1500), linetype = "dashed") +
+  geom_hline(yintercept = c(900, 1500), linetype = "dashed") +
   theme_light(12) +
@@
 ```{r, echo = FALSE}
-# Convert PCoA to tidy array ("fortify")
-# pcoa_fort <- function(x){
-#   fortify(x, axes = 1:2) %>% 
-#   mutate(Score = case_when( # rename default scores to match my data
-#     Score == "species" ~ "FL sp.",
-#     Score == "sites" ~ "BB sp."))
-# }
-
-pcoa_fort <- function(x){
+# Fortify envfit() object fitted to BB tongue length data
+envfit_fort <- function(x){
+  fortify(x) %>%
+  select(Score = Type, Label, MDS1, MDS2) %>%
+  mutate(Label = case_when(Label == "pbl.w" ~ "T.len"),
+         Score = case_when(Score == "Vector" ~ "T.len"))
+}
+
+# Fortify PCOA and add the results of an envfit() to BB tongue length
+pcoa_fort <- function(x, y){
   fortify(x, axes = 1:2) %>%
-  mutate(Score = case_when( # rename default scores to match my data
-    Score == "species" ~ "flora",
-    #Label %in% c("hort", "pasc", "wurf") ~ "LT spp.",
-    Label %in% c("hort", "pasc") ~ "LT spp.",
-    Label == "wurf" ~ "robber",
-    Label %in% c("bss", "soro", "prat") ~ "ST spp."))
+  bind_rows(y) %>%
+  mutate(Score = case_when(
+    Score == "species" ~ "FL.sp",
+    Score == "sites" ~ "BB.sp",
+    Score == "T.len" ~ "T.len"))
 }
 
-# Plot PCoA; to avoid overplotting, scores < min_score are not plotted
-pcoa_plot <- function(x, min_score = 0.25) {
-
-  # # Add T.class variable for the BB.sp subset of the PCoA
-  # t.class <- x %>%
-  #   filter(Score == "BB sp.") %>%
-  #   mutate(T.class = factor(if_else(
-  #     Label %in% c("hort", "pasc", "wurf", "gers", "mend", "hypn"), "long", "short"))
-  #     )
-
-  cols <- c("flora" = "gray40", "LT spp." = "#F8766D", "ST spp." = "#00BA38", "robber" = "#619CFF")
-
-  ggplot(filter(x, (abs(MDS1) > min_score | abs(MDS2) > min_score) | (Score %in% c("LT spp.", "ST spp.", "robber"))), 
-         aes(MDS1, MDS2, color = Score)) + # no threshold for BB spp
-    #geom_polygon(data = t.class, aes(MDS1, MDS2, linetype = T.class), # draw polygons around T classes
-     #            color = "gray70", fill = NA, size = 1.25, key_glyph = "abline") +
-    geom_point(size = 2, alpha = 0.75) + # add points
-    ggrepel::geom_text_repel(aes(label = Label), key_glyph = "rect", size = 3) + # add labels
-    coord_fixed() + # Gavin Simpson says this is **really** important
-    theme_classic() +
-    theme(legend.title=element_blank()) +
-    scale_color_manual(values = cols)
+# Plot PCoA ordination with arrow for tongue length envfit()
+pcoa_plot <- function(x, y, min_load = 0.25) {
+  ggplot(filter(x, (abs(MDS1) > min_load | abs(MDS2) > min_load) | (Score %in% c("BB.sp", "T.len")))) + # to avoid overplotting, we can filter to include only floral species > loading threshold
+    geom_point(aes(MDS1, MDS2, color = Score)) +
+    geom_text_repel(aes(MDS1, MDS2, color = Score, label = Label), key_glyph = "rect") +
+    geom_segment(data = y, 
+               aes(x = 0, y = 0, xend = MDS1, yend = MDS2, colour = Label),
+               arrow = arrow(), size = 2, alpha = 0.5, key_glyph = "rect") +
+    theme_light() +
+    coord_fixed() # Gavin Simpson says this is **really** important
 }
@@
     filter(caste != "male") %>%
-    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
+    filter(!bb.sp %in% c("psit", "gers", "humi")) %>% # omit parasitic species, gers because it's a specialist, humi because it's so rare
     group_by(..., bb.sp, !!criterion) %>% # pooling all sites and years
@@
                           column_to_rownames(var = "bb.sp") %>% 
-                          select(pbl.fac) # pbl.fac is the variable I'm interested in
+                          select(pbl.w) # pbl.w is the variable I'm interested in
                         ),
-           # adonis = map2(comm2, traits, # run adonis permuatation test for distance ~ tongue class
-           #               function(x, y) adonis(x ~ pbl.fac, data = y, method = "horn")
-           #               ),
-           # betadis = map2(comm2, traits, # betadisper to test for difference in dispersion between tongue classes
-           #               function(x, y) x %>%
-           #                 vegdist(method = "horn") %>% # betadisper runs a PCoA under the hood; match method to my PCoA
-           #                 betadisper(group = y$pbl.fac) 
-           #               ),
-           # betadis.permutest = map(betadis, permutest), # perumutation test
-           # betadis.p = map(betadis.permutest, # extract betadisper p-value
-           #                 function(x) x$tab$'Pr(>F)'[[1]]
-           #                 ),
-           # betadis.boxplot = map(betadis, boxplot), # boxplot of betadisper
-           pcoa.fort = map(pcoa, pcoa_fort), 
-           pcoa.plot = map(pcoa.fort, pcoa_plot),
-           # adonis.r2 = map(adonis, # extract adonis r2
-           #                 function(x) x$aov.tab$R2[[1]]
-           #                 ),
-           # adonis.p = map(adonis, # extract adonis p-value
-           #                function(x) x$aov.tab$'Pr(>F)'[[1]]
-           #                ),
+           adonis = map2(comm2, traits, # run adonis permuatation test for distance ~ tongue class
+                         function(x, y) adonis(x ~ pbl.w, data = y, method = "horn")
+                         ),
+           envfit = map2(pcoa, traits, function(x, y) envfit(x, y)),
+           envfit.fort = map(envfit, envfit_fort),
+           pcoa.fort = map2(pcoa, envfit.fort, pcoa_fort),
+           pcoa.plot = map2(pcoa.fort, envfit.fort, pcoa_plot), 
+           envfit.r2 = map(envfit, function(x) x$vectors$r),
+           envfit.p = map(envfit, function(x) x$vectors$pvals),
+           adonis.r2 = map(adonis, # extract adonis r2
+                           function(x) x$aov.tab$R2[[1]]
+                           ),
+           adonis.p = map(adonis, # extract adonis p-value
+                          function(x) x$aov.tab$'Pr(>F)'[[1]]
+                          ),
+           ordisurf = map2(pcoa, traits, function(x, y) ordisurf(x ~ pbl.w, data = y, k = 8)),
+           ordisurf.dev_exp = map(ordisurf, function(x) summary(x)$dev.expl),
+           ordisurf.p = map(ordisurf, function(x) summary(x)$s.pv),
            MDS1 = map(pcoa, # extract variance explained by MDS1
@@
                       )) %>%
-    #unnest(cols = c(adonis.r2, adonis.p, betadis.p, MDS1, MDS2)) %>% # these are lists of length = 1, so unnesting them simply displays the value
-    unnest(cols = c(MDS1, MDS2)) %>% # these are lists of length = 1, so unnesting them simply displays the value
-    # select(..., comm2, pcoa, betadis, betadis.permutest, betadis.p, adonis, 
-    #        adonis.r2, adonis.p, betadis.boxplot, pcoa.plot, MDS1, MDS2) 
-    select(..., comm2, pcoa, pcoa.plot, MDS1, MDS2)
+    unnest(cols = c(ordisurf.dev_exp, ordisurf.p, envfit.r2, envfit.p, adonis.r2, adonis.p, MDS1, MDS2)) %>% # these are lists of length = 1, so unnesting them simply displays the value
+    select(..., comm2, traits, pcoa, ordisurf, ordisurf.dev_exp, ordisurf.p, envfit, envfit.fort, pcoa.fort, pcoa.plot, envfit.r2, envfit.p, adonis.r2, adonis.p, MDS1, MDS2)
 }
 
+ordi1 <- ordisurf(niche.tax$pcoa[[1]] ~ pbl.w, data = niche.tax$traits[[1]])
+
+# Calculate site-wise beta diversity from visitation data
 beta_func <- function(x, traits, sitedata, level, criterion, binary, ...) {
@@
 
-# For the floral survey data
+# Calculate site-wise beta diversity from survey data
 beta_func_fl <- function(x, sitedata, criterion, binary, ...) {
@@
 dev.off()
+
+# Playing with ordisurf as alternative to envfit()
+library(ggordiplots)
+
+## Based on https://oliviarata.wordpress.com/2014/07/17/ordinations-in-ggplot2-v2-ordisurf/
+# Get ordination object
+ordi1 <- niche.tax$ordisurf[[1]] 
+ordi1.grid <- ordi1$grid #extracts the ordisurf object
+ordi1.mite <- expand.grid(x = ordi1.grid$x, y = ordi1.grid$y) #get x and ys
+ordi1.mite$z <- as.vector(ordi1.grid$z) #unravel the matrix for the z scores
+ordi1.mite.na <- data.frame(na.omit(ordi1.mite)) %>% #gets rid of the nas
+  as_tibble() %>%
+  rename(MDS1 = x, MDS2 = y)
+
+# Get scores
+scores <- fortify(niche.tax$pcoa[[1]], axes = 1:2) %>%
+  mutate(Score = case_when(
+    Score == "species" ~ "FL.sp",
+    Score == "sites" ~ "BB.sp")
+    ) %>%
+  as_tibble()
+
+# Plot
+dat <- filter(scores, (abs(MDS1) > 0.25 | abs(MDS2) > 0.25) | Score == "BB.sp")
+
+ggplot(dat, aes(MDS1, MDS2, pch = Score)) +
+  geom_contour(data = ordi1.mite.na, aes(MDS1, MDS2, z = z, color = stat(level)), size = 1.5, alpha = 0.75, inherit.aes = FALSE) +
+  geom_point(size = 2) +
+  geom_text_repel(aes(MDS1, MDS2, label = Label), key_glyph = "rect") +
+  theme_light() +
+  scale_color_viridis_c(name = "Tongue length") +
+  scale_shape_manual(values = c(16, 1), name = "Level") +
+  coord_fixed() # Gavin Simpson says this is **really** important
 ```
@@
   filter(caste != "male") %>%
-  #filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
+  filter(!bb.sp %in% c("psit", "gers", "humi")) %>% # only focal species
   #group_by(elev.class2) %>%
@@
   mutate(prop = abund/sum(abund)) %>% 
+  left_join(bb_traits) %>%
   arrange(elev.class2, period, bb.sp, -prop)
@@
   labs(fill = "Morphotype") +
-  facet_grid(fct_rev(elev.class2)~bb.sp) +
+  facet_grid(fct_rev(elev.class2)~reorder(bb.sp, pbl.w)) +
   theme(strip.text = element_text(size = 14))
@@
 tp_ord1 <- (
-  tp.early | (
-    niche_high_early.tax.plot / 
-    niche_mid_early.tax.plot / 
-    niche_low_early.tax.plot
+  tp.early / (
+    niche_low_early.tax.plot + ggtitle("low") | 
+    niche_mid_early.tax.plot + ggtitle("mid") | 
+    niche_high_early.tax.plot + ggtitle("high")
     ) + 
@@
   ) + 
-  plot_layout(widths = c(5, 1))
-ggsave("./output/tree_ord_early.png", tp_ord1, width = 14, height = 8)
+  plot_layout(heights = c(2, 1))
+ggsave("./output/tree_ord_early.png", tp_ord1, width = 16, height = 10)
 

### 2020-12-02 analyses/niche_relations_v6.Rmd
update lab deskttop

@@
 ```{r, echo = FALSE}
+# Plot ordisurf
+## Based on https://oliviarata.wordpress.com/2014/07/17/ordinations-in-ggplot2-v2-ordisurf/
+surfplot <- function(x, y, min_score = 0.25) {
+
+  grid <- x$grid # extracts the ordisurf object
+  grid_exp <- expand.grid(x = grid$x, y = grid$y) # get x and ys
+  grid_exp$z <- as.vector(grid$z) # unravel the matrix for the z scores
+
+  contours <- data.frame(na.omit(grid_exp)) %>% # gets rid of the nas
+    as_tibble() %>%
+    rename(MDS1 = x, MDS2 = y)
+
+  scores <- y %>% 
+    filter(Score != "T.len") %>%
+    filter((abs(MDS1) > 0.25 | abs(MDS2) > 0.25) | Score == "BB.sp")
+
+  ggplot(scores, aes(MDS1, MDS2, pch = Score)) +
+    geom_contour(data = contours, aes(MDS1, MDS2, z = z, color = stat(level)), size = 1.5, alpha = 0.75, inherit.aes = FALSE) +
+    geom_point(size = 2) +
+    geom_text_repel(aes(MDS1, MDS2, label = Label), key_glyph = "rect") +
+    theme_light() +
+    scale_color_viridis_c(name = "Tongue length") +
+    scale_shape_manual(values = c(16, 1), name = "Level") +
+    coord_fixed() # Gavin Simpson says this is **really** important
+}
+
 # Fortify envfit() object fitted to BB tongue length data
@@
 # Plot PCoA ordination with arrow for tongue length envfit()
-pcoa_plot <- function(x, y, min_load = 0.25) {
-  ggplot(filter(x, (abs(MDS1) > min_load | abs(MDS2) > min_load) | (Score %in% c("BB.sp", "T.len")))) + # to avoid overplotting, we can filter to include only floral species > loading threshold
+pcoa_plot <- function(x, y, min_score = 0.25) {
+  ggplot(filter(x, (abs(MDS1) > min_score | abs(MDS2) > min_score) | (Score %in% c("BB.sp", "T.len")))) + # to avoid overplotting, we can filter to include only floral species > loading threshold
     geom_point(aes(MDS1, MDS2, color = Score)) +
@@
            ordisurf.p = map(ordisurf, function(x) summary(x)$s.pv),
+           ordisurf.plot = map2(ordisurf, pcoa.fort, surfplot),
            MDS1 = map(pcoa, # extract variance explained by MDS1
@@
     unnest(cols = c(ordisurf.dev_exp, ordisurf.p, envfit.r2, envfit.p, adonis.r2, adonis.p, MDS1, MDS2)) %>% # these are lists of length = 1, so unnesting them simply displays the value
-    select(..., comm2, traits, pcoa, ordisurf, ordisurf.dev_exp, ordisurf.p, envfit, envfit.fort, pcoa.fort, pcoa.plot, envfit.r2, envfit.p, adonis.r2, adonis.p, MDS1, MDS2)
+    select(..., comm2, traits, pcoa, ordisurf, ordisurf.dev_exp, ordisurf.p, ordisurf.plot, envfit, envfit.fort, pcoa.fort, pcoa.plot, envfit.r2, envfit.p, adonis.r2, adonis.p, MDS1, MDS2)
 }
-
-ordi1 <- ordisurf(niche.tax$pcoa[[1]] ~ pbl.w, data = niche.tax$traits[[1]])
 
@@
 niche.tax.plot <- niche.tax$pcoa.plot[[1]]
+niche.tax.surfplot <- niche.tax$ordisurf.plot
+nich.tax.baseplot <- plot(niche.tax$ordisurf[[1]])
+niche.tax$ordisurf.plot
+plot(niche.tax$ordisurf[[1]])
 
@@
 
-# Playing with ordisurf as alternative to envfit()
-library(ggordiplots)
-
-## Based on https://oliviarata.wordpress.com/2014/07/17/ordinations-in-ggplot2-v2-ordisurf/
-# Get ordination object
-ordi1 <- niche.tax$ordisurf[[1]] 
-ordi1.grid <- ordi1$grid #extracts the ordisurf object
-ordi1.mite <- expand.grid(x = ordi1.grid$x, y = ordi1.grid$y) #get x and ys
-ordi1.mite$z <- as.vector(ordi1.grid$z) #unravel the matrix for the z scores
-ordi1.mite.na <- data.frame(na.omit(ordi1.mite)) %>% #gets rid of the nas
-  as_tibble() %>%
-  rename(MDS1 = x, MDS2 = y)
-
-# Get scores
-scores <- fortify(niche.tax$pcoa[[1]], axes = 1:2) %>%
-  mutate(Score = case_when(
-    Score == "species" ~ "FL.sp",
-    Score == "sites" ~ "BB.sp")
-    ) %>%
-  as_tibble()
-
-# Plot
-dat <- filter(scores, (abs(MDS1) > 0.25 | abs(MDS2) > 0.25) | Score == "BB.sp")
-
-ggplot(dat, aes(MDS1, MDS2, pch = Score)) +
-  geom_contour(data = ordi1.mite.na, aes(MDS1, MDS2, z = z, color = stat(level)), size = 1.5, alpha = 0.75, inherit.aes = FALSE) +
-  geom_point(size = 2) +
-  geom_text_repel(aes(MDS1, MDS2, label = Label), key_glyph = "rect") +
-  theme_light() +
-  scale_color_viridis_c(name = "Tongue length") +
-  scale_shape_manual(values = c(16, 1), name = "Level") +
-  coord_fixed() # Gavin Simpson says this is **really** important
 ```

### 2020-12-03 analyses/niche_relations_v6B.Rmd
realized mistake about not standardizing by site-date

--- 2020-12-02 analyses/niche_relations_v6.Rmd
@@
 library(treemapify)
-#library(RColorBrewer)
 library(tidyverse)
@@
   group_by(site) %>%
-  mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) # divide each elev.class2 into 4 equal time slices
+  mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) # divide each elev.class2 into 4 equal time slices
 
@@
   group_by(site) %>%
-  mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
+  mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) %>%
   group_by(site, period, dayofyear) %>%
@@
   scores <- y %>% 
-    filter(Score != "T.len") %>%
-    filter((abs(MDS1) > 0.25 | abs(MDS2) > 0.25) | Score == "BB.sp")
+    #filter((abs(MDS1) > 0.25 | abs(MDS2) > 0.25) | Score == "BB.sp")
+    filter((abs(MDS1) > 0.1 | abs(MDS2) > 0.1) | Score == "BB.sp")
 
@@
 
-# Fortify envfit() object fitted to BB tongue length data
-envfit_fort <- function(x){
-  fortify(x) %>%
-  select(Score = Type, Label, MDS1, MDS2) %>%
-  mutate(Label = case_when(Label == "pbl.w" ~ "T.len"),
-         Score = case_when(Score == "Vector" ~ "T.len"))
-}
-
 # Fortify PCOA and add the results of an envfit() to BB tongue length
-pcoa_fort <- function(x, y){
+pcoa_fort <- function(x){
   fortify(x, axes = 1:2) %>%
-  bind_rows(y) %>%
   mutate(Score = case_when(
     Score == "species" ~ "FL.sp",
-    Score == "sites" ~ "BB.sp",
-    Score == "T.len" ~ "T.len"))
+    Score == "sites" ~ "BB.sp"))
 }
 
-# Plot PCoA ordination with arrow for tongue length envfit()
-pcoa_plot <- function(x, y, min_score = 0.25) {
-  ggplot(filter(x, (abs(MDS1) > min_score | abs(MDS2) > min_score) | (Score %in% c("BB.sp", "T.len")))) + # to avoid overplotting, we can filter to include only floral species > loading threshold
-    geom_point(aes(MDS1, MDS2, color = Score)) +
-    geom_text_repel(aes(MDS1, MDS2, color = Score, label = Label), key_glyph = "rect") +
-    geom_segment(data = y, 
-               aes(x = 0, y = 0, xend = MDS1, yend = MDS2, colour = Label),
-               arrow = arrow(), size = 2, alpha = 0.5, key_glyph = "rect") +
-    theme_light() +
-    coord_fixed() # Gavin Simpson says this is **really** important
+# Plot PCoA; to avoid overplotting, scores < min_score are not plotted
+pcoa_plot <- function(x, min_score = 0.25) {
+
+  # # Add T.class variable for the BB.sp subset of the PCoA
+  # t.class <- x %>%
+  #   filter(Score == "BB sp.") %>%
+  #   mutate(T.class = factor(if_else(
+  #     Label %in% c("hort", "pasc", "wurf", "gers", "mend", "hypn"), "long", "short"))
+  #     )
+
+  # ggplot(filter(x, (abs(MDS1) > min_score | abs(MDS2) > min_score) | (Score == "BB.sp")), 
+  #        aes(MDS1, MDS2, color = Score)) + # no threshold for BB spp
+  ggplot(x, aes(MDS1, MDS2, color = Score)) +
+    #geom_polygon(data = t.class, aes(MDS1, MDS2, linetype = T.class), # draw polygons around T classes
+     #            color = "gray70", fill = NA, size = 1.25, key_glyph = "abline") +
+    geom_point(size = 2, alpha = 0.75) + # add points
+    ggrepel::geom_text_repel(aes(label = Label), key_glyph = "rect", size = 3) + # add labels
+    coord_fixed() + # Gavin Simpson says this is **really** important
+    theme_classic() +
+    theme(legend.title=element_blank())
 }
 
-# Generates PCoAs, ADONIS tests, and ordination plots (fully aggregated)
+# Generates PCoAs, ordisurfs, ADONIS tests, and ordination plots 
 ## The ... represents the variables we will use to parse our analysis. For aggregate analysis, just leave the ...
@@
     filter(caste != "male") %>%
-    filter(!bb.sp %in% c("psit", "gers", "humi")) %>% # omit parasitic species, gers because it's a specialist, humi because it's so rare
-    group_by(..., bb.sp, !!criterion) %>% # pooling all sites and years
+    #filter(!bb.sp %in% c("psit", "gers", "humi")) %>% # omit parasitic species, gers because it's a specialist, humi because it's so rare
+    filter(!bb.sp %in% c("gers", "humi")) %>% # gers because it's a specialist, humi because it's so rare
+    #filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
+    group_by(site, date, bb.sp, !!criterion, ...) %>% # pooling all sites and years
     summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
-    group_by(..., bb.sp) %>%
-    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
-    dplyr::select(-freq) %>% # drop freq variable
+    group_by(site, date, bb.sp) %>%
+    mutate(prop = freq/sum(freq)) %>%
+    group_by(site, date, ...) %>%
+    complete(bb.sp, !!criterion, fill = list(prop = 0)) %>% # fill in zeros for unvisited plants
+    group_by(..., bb.sp, !!criterion) %>%
+    summarize(prop = mean(prop)) %>%
+    #dplyr::select(-freq) %>% # drop freq variable
     group_by(...) %>%
@@
                          ),
-           envfit = map2(pcoa, traits, function(x, y) envfit(x, y)),
-           envfit.fort = map(envfit, envfit_fort),
-           pcoa.fort = map2(pcoa, envfit.fort, pcoa_fort),
-           pcoa.plot = map2(pcoa.fort, envfit.fort, pcoa_plot), 
-           envfit.r2 = map(envfit, function(x) x$vectors$r),
-           envfit.p = map(envfit, function(x) x$vectors$pvals),
+           pcoa.fort = map(pcoa, pcoa_fort),
+           pcoa.plot = map(pcoa.fort, pcoa_plot),
            adonis.r2 = map(adonis, # extract adonis r2
@@
            ordisurf = map2(pcoa, traits, function(x, y) ordisurf(x ~ pbl.w, data = y, k = 8)),
+           #ordisurf = map2(pcoa, traits, function(x, y) ordisurf(x ~ pbl.w, data = y, k = 5)),
            ordisurf.dev_exp = map(ordisurf, function(x) summary(x)$dev.expl),
@@
            ordisurf.plot = map2(ordisurf, pcoa.fort, surfplot),
-           MDS1 = map(pcoa, # extract variance explained by MDS1
+           var.exp.MDS1 = map(pcoa, # extract variance explained by MDS1
                       function(x) x$CA$eig[[1]] / sum(x$CA$eig)
                       ), 
-           MDS2 = map(pcoa, # extract variance explained by MDS2
+           var.exp.MDS2 = map(pcoa, # extract variance explained by MDS2
                       function(x) x$CA$eig[[2]] / sum(x$CA$eig)
                       )) %>%
-    unnest(cols = c(ordisurf.dev_exp, ordisurf.p, envfit.r2, envfit.p, adonis.r2, adonis.p, MDS1, MDS2)) %>% # these are lists of length = 1, so unnesting them simply displays the value
-    select(..., comm2, traits, pcoa, ordisurf, ordisurf.dev_exp, ordisurf.p, ordisurf.plot, envfit, envfit.fort, pcoa.fort, pcoa.plot, envfit.r2, envfit.p, adonis.r2, adonis.p, MDS1, MDS2)
+    unnest(cols = c(ordisurf.dev_exp, ordisurf.p, adonis.r2, 
+                    adonis.p, var.exp.MDS1, var.exp.MDS2)) %>% # these are lists of length = 1, so unnesting them simply displays the value
+    mutate(var.exp.MDS1_MDS2 = var.exp.MDS1 + var.exp.MDS2) %>%
+    select(..., pcoa, pcoa.plot, var.exp.MDS1_MDS2, var.exp.MDS1, var.exp.MDS2, 
+            ordisurf.dev_exp, ordisurf.p, adonis.r2, adonis.p,
+            ordisurf, ordisurf.plot, adonis, comm2, traits, pcoa.fort)
 }
@@
     group_by(site) %>%
-    mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
+    mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) %>%
     group_by(site, period, pbl.fac, !!criterion) %>% # group by site, date, bee species, plant species
@@
     group_by(site) %>%
-    mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
+    mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) %>%
     left_join(bb_traits) %>%
@@
     # group_by(elev.class2) %>%
-    # mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
+    # mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) %>%
     #group_by(site, period, pbl.fac, !!criterion) %>% # group by site, date, bee species, plant species
@@
     # group_by(elev.class2) %>%
-    # mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
+    # mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) %>%
     left_join(bb_traits) %>%
@@
 niche.tax <- niche_analysis(net, bb_traits, criterion = plant.sp.abb)
-niche.tax.plot <- niche.tax$pcoa.plot[[1]]
-niche.tax.surfplot <- niche.tax$ordisurf.plot
-nich.tax.baseplot <- plot(niche.tax$ordisurf[[1]])
-niche.tax$ordisurf.plot
-plot(niche.tax$ordisurf[[1]])
+niche.tax.plot <- niche.tax$ordisurf.plot[[1]]
 
@@
 niche.morph <- niche_analysis(net, bb_traits, criterion = k.type.s)
-niche.morph.plot <- niche.morph$pcoa.plot[[1]]
+niche.morph.plot <- niche.morph$ordisurf.plot[[1]]
+
+# By color
+niche.color <- niche_analysis(net, bb_traits, criterion = color)
+niche.color.plot <- niche.color$pcoa.plot[[1]]
 
@@
 pdf("../analyses/output/niche_pcoa.pdf", height = 8.5, width = 11)
-(niche.tax.plot + ggtitle("Taxonomic") | niche.morph.plot + ggtitle("Morphological")) + plot_layout(guides = 'collect') 
+(niche.tax.plot + ggtitle("Taxonomic") | niche.morph.plot + ggtitle("Morphological")) + plot_layout(guides = 'keep') 
 dev.off()
@@
 ## Extract and name plots
-niche_high.tax.plot <- filter(niche_x_elev.tax, elev.class2 == "high")$pcoa.plot[[1]] + ggtitle("high/Tax")
-niche_mid.tax.plot <- filter(niche_x_elev.tax, elev.class2 == "mid")$pcoa.plot[[1]] + ggtitle("mid/Tax")
-niche_low.tax.plot <- filter(niche_x_elev.tax, elev.class2 == "low")$pcoa.plot[[1]] + ggtitle("low/Tax")
+niche_high.tax.plot <- filter(niche_x_elev.tax, elev.class2 == "high")$ordisurf.plot[[1]] + ggtitle("high/Tax")
+niche_mid.tax.plot <- filter(niche_x_elev.tax, elev.class2 == "mid")$ordisurf.plot[[1]] + ggtitle("mid/Tax")
+niche_low.tax.plot <- filter(niche_x_elev.tax, elev.class2 == "low")$ordisurf.plot[[1]] + ggtitle("low/Tax")
 
 # Morphological
-niche_x_elev.morph <- niche_analysis(net, bb_traits, criterion = k.type.s, elev.class2)
+niche_x_elev.morph <- niche_analysis(net, bb_traits, criterion = k.type.ss, elev.class2)
 
 ## Extract and name plots
-niche_high.morph.plot <- filter(niche_x_elev.morph, elev.class2 == "high")$pcoa.plot[[1]] + ggtitle("high/morph")
-niche_mid.morph.plot <- filter(niche_x_elev.morph, elev.class2 == "mid")$pcoa.plot[[1]] + ggtitle("mid/morph")
-niche_low.morph.plot <- filter(niche_x_elev.morph, elev.class2 == "low")$pcoa.plot[[1]] + ggtitle("low/morph")
+niche_high.morph.plot <- filter(niche_x_elev.morph, elev.class2 == "high")$ordisurf.plot[[1]] + ggtitle("high/morph")
+niche_mid.morph.plot <- filter(niche_x_elev.morph, elev.class2 == "mid")$ordisurf.plot[[1]] + ggtitle("mid/morph")
+niche_low.morph.plot <- filter(niche_x_elev.morph, elev.class2 == "low")$ordisurf.plot[[1]] + ggtitle("low/morph")
+
+# Color
+niche_x_elev.color <- niche_analysis(net, bb_traits, criterion = color, elev.class2)
+
+## Extract and name plots
+niche_high.color.plot <- filter(niche_x_elev.color, elev.class2 == "high")$pcoa.plot[[1]] + ggtitle("high/color")
+niche_mid.color.plot <- filter(niche_x_elev.color, elev.class2 == "mid")$pcoa.plot[[1]] + ggtitle("mid/color")
+niche_low.color.plot <- filter(niche_x_elev.color, elev.class2 == "low")$pcoa.plot[[1]] + ggtitle("low/color")
 
@@
 (niche_mid.tax.plot | niche_mid.morph.plot) /
-(niche_low.tax.plot | niche_low.morph.plot) + plot_layout(guides = 'collect') 
+(niche_low.tax.plot | niche_low.morph.plot) + plot_layout(guides = 'keep') 
 dev.off()
@@
 ## Extract and name plots
-niche_high_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "high" & period == "early")$pcoa.plot[[1]] 
-niche_high_mid1.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "high" & period == "mid1")$pcoa.plot[[1]] 
-niche_high_mid2.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "high" & period == "mid2")$pcoa.plot[[1]] 
-niche_high_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "high" & period == "late")$pcoa.plot[[1]]
-
-niche_mid_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "mid" & period == "early")$pcoa.plot[[1]]
-niche_mid_mid1.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "mid" & period == "mid1")$pcoa.plot[[1]] 
-niche_mid_mid2.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "mid" & period == "mid2")$pcoa.plot[[1]]
-niche_mid_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "mid" & period == "late")$pcoa.plot[[1]]
-
-niche_low_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "low" & period == "early")$pcoa.plot[[1]]
-niche_low_mid1.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "low" & period == "mid1")$pcoa.plot[[1]]
-niche_low_mid2.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "low" & period == "mid2")$pcoa.plot[[1]]
-niche_low_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "low" & period == "late")$pcoa.plot[[1]]
+niche_high_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "high" & period == "early")$ordisurf.plot[[1]] 
+# niche_high_mid1.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "high" & period == "mid1")$ordisurf.plot[[1]] 
+# niche_high_mid2.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "high" & period == "mid2")$ordisurf.plot[[1]]
+niche_high_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "high" & period == "mid")$ordisurf.plot[[1]]
+niche_high_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "high" & period == "late")$ordisurf.plot[[1]]
+
+niche_mid_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "mid" & period == "early")$ordisurf.plot[[1]]
+# niche_mid_mid1.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "mid" & period == "mid1")$ordisurf.plot[[1]] 
+# niche_mid_mid2.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "mid" & period == "mid2")$ordisurf.plot[[1]]
+niche_mid_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "mid" & period == "mid")$ordisurf.plot[[1]]
+niche_mid_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "mid" & period == "late")$ordisurf.plot[[1]]
+
+niche_low_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "low" & period == "early")$ordisurf.plot[[1]]
+# niche_low_mid1.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "low" & period == "mid1")$ordisurf.plot[[1]]
+# niche_low_mid2.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "low" & period == "mid2")$ordisurf.plot[[1]]
+niche_low_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "low" & period == "mid")$ordisurf.plot[[1]]
+niche_low_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "low" & period == "late")$ordisurf.plot[[1]]
 
@@
 ## Extract and name plots
-niche_high_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "high" & period == "early")$pcoa.plot[[1]]
-niche_high_mid1.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "high" & period == "mid1")$pcoa.plot[[1]] 
-niche_high_mid2.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "high" & period == "mid2")$pcoa.plot[[1]] 
-niche_high_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "high" & period == "late")$pcoa.plot[[1]]
-
-niche_mid_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "mid" & period == "early")$pcoa.plot[[1]]
-niche_mid_mid1.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "mid" & period == "mid1")$pcoa.plot[[1]]
-niche_mid_mid2.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "mid" & period == "mid2")$pcoa.plot[[1]] 
-niche_mid_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "mid" & period == "late")$pcoa.plot[[1]]
-
-niche_low_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "low" & period == "early")$pcoa.plot[[1]]
-niche_low_mid1.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "low" & period == "mid1")$pcoa.plot[[1]] 
-niche_low_mid2.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "low" & period == "mid2")$pcoa.plot[[1]] 
-niche_low_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "low" & period == "late")$pcoa.plot[[1]]
+niche_high_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "high" & period == "early")$ordisurf.plot[[1]]
+niche_high_mid1.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "high" & period == "mid1")$ordisurf.plot[[1]] 
+niche_high_mid2.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "high" & period == "mid2")$ordisurf.plot[[1]] 
+niche_high_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "high" & period == "late")$ordisurf.plot[[1]]
+
+niche_mid_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "mid" & period == "early")$ordisurf.plot[[1]]
+niche_mid_mid1.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "mid" & period == "mid1")$ordisurf.plot[[1]]
+niche_mid_mid2.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "mid" & period == "mid2")$ordisurf.plot[[1]] 
+niche_mid_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "mid" & period == "late")$ordisurf.plot[[1]]
+
+niche_low_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "low" & period == "early")$ordisurf.plot[[1]]
+niche_low_mid1.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "low" & period == "mid1")$ordisurf.plot[[1]] 
+niche_low_mid2.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "low" & period == "mid2")$ordisurf.plot[[1]] 
+niche_low_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "low" & period == "late")$ordisurf.plot[[1]]
+
+# Color
+niche_x_elevperiod.color <- niche_analysis(net, bb_traits, criterion = k.type.s, elev.class2, period)
+
+## Extract and name plots
+niche_high_early.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "high" & period == "early")$pcoa.plot[[1]]
+niche_high_mid1.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "high" & period == "mid1")$pcoa.plot[[1]] 
+niche_high_mid2.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "high" & period == "mid2")$pcoa.plot[[1]] 
+niche_high_late.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "high" & period == "late")$pcoa.plot[[1]]
+
+niche_mid_early.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "mid" & period == "early")$pcoa.plot[[1]]
+niche_mid_mid1.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "mid" & period == "mid1")$pcoa.plot[[1]]
+niche_mid_mid2.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "mid" & period == "mid2")$pcoa.plot[[1]] 
+niche_mid_late.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "mid" & period == "late")$pcoa.plot[[1]]
+
+niche_low_early.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "low" & period == "early")$pcoa.plot[[1]]
+niche_low_mid1.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "low" & period == "mid1")$pcoa.plot[[1]] 
+niche_low_mid2.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "low" & period == "mid2")$pcoa.plot[[1]] 
+niche_low_late.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "low" & period == "late")$pcoa.plot[[1]]
 
@@
   group_by(site) %>%
-  mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
+  mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) %>%
   ungroup() %>%
@@
   group_by(site) %>%
-  mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
+  mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) %>%
   mutate(k.type.ss = factor(k.type.ss)) %>%
@@
   group_by(site) %>%
-  mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
+  mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) %>%
   mutate(k.type.ss = factor(k.type.ss),
@@
   group_by(site) %>%
-  mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
+  mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) %>%
   # mutate(k.type.ss = factor(k.type.ss),
@@
 ggsave("./output/tree_ord_late.png", tp_ord4, width = 14, height = 8)
+```
+
+# 
+```{r}
+vis_sum <- net %>%
+  group_by(bb.sp, plant.sp, k.type.ss, color) %>%
+  summarize()
+
 ```

### 2020-12-07 analyses/diet_ordination.Rmd
update lab desktop

--- 2020-12-03 analyses/niche_relations_v6B.Rmd
@@
 ---
-title: "Floral resource choice by bumble bee species in the Berchtesgadener Alps"
+title: "ordination"
+author: "Doug Sponsler"
+date: "12/4/2020"
 output: html_document
@@
 
-# Prepare workspace
-## Load packages 
-```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
-library(bipartite)
-library(mgcv)
-library(mgcViz)
+# Load packages
+```{r}
 library(vegan)
@@
   left_join(fl_traits) %>% # add floral traits
-  group_by(site) %>%
-  mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) # divide each elev.class2 into 4 equal time slices
+  #group_by(site) %>%
+  group_by(elev.class2) %>%
+  mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) %>% # divide each elev.class2 into 4 equal time slices
+  mutate(k.type.s = as.character(k.type.s), # these can't be factors or else complete() will try to preserve all their levels for each group, even the levels that do not occur
+         k.type.ss = as.character(k.type.ss))
 
@@
 
-Core species:  
-  1. *B. sensu-strictu*
-  2. *B. pascuorum*
-  3. *B. pratorum*
-  4. *B. hortorum*
-  5. *B. soroensis*
-  6. *B. wurflenii*
-
-```{r fig.cap="**Figure 1: Focal species**", message=FALSE, warning=FALSE, echo=FALSE}
-bb_range.year <- net %>%
-  left_join(site_data) %>%
-  group_by(site, elev.mean, bb.sp, year) %>%
-  summarize(abund = n()) %>%
-  group_by(bb.sp) %>%
-  mutate(elev.floor = min(elev.mean),
-         elev.ceiling = max(elev.mean),
-         elev.range = elev.ceiling - elev.floor,
-         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2)) %>%
-  mutate(focal = if_else(
-    bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf"), TRUE, FALSE)
-    )
-
-ggplot(bb_range.year, aes(reorder(bb.sp, elev.med), elev.mean, color = focal)) +
-  geom_line(size = 2, alpha = 0.25) +
-  geom_point(aes(size = abund), alpha = 0.5) +
-  facet_wrap(~year, ncol = 1) +
-  theme_light(12) +
-  scale_size_continuous(name = "Abundance") +
-  scale_color_discrete(name = "Focal species") +
-  xlab("BB species") +
-  ylab("Elevation") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-ggsave("./output/bb_range_year.pdf", height = 6, width = 6)
-
-# Do we see the bimodal abundance pattern predicted by the core-satellite hypothesis? Well, sort of.
-bb_abund <- net %>%
-  group_by(bb.sp) %>%
-  summarize(abund = n())
-
-ggplot(bb_abund, aes(abund)) +
-  geom_histogram(binwidth = 250)
-```
-
-We will parse the samples by elevation class and season, dividing the samples from each site into equal-interval quarters
+# Define sample parsing
+We will parse the samples by elevation class and season, dividing the samples from each elevation class into equal-interval thirds
 ```{r}
@@
   filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
-  #group_by(elev.class2) %>%
-  group_by(site) %>%
+  group_by(elev.class2) %>%
+  #group_by(site) %>%
   mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) %>%
@@
 ggplot(samples, aes(as_date(yday(date)), elev.mean, color = period)) +
-  #geom_point(size = 1, aes(alpha = obs)) +
-  geom_point(size = 2, alpha = 0.75) +
-  #facet_wrap(~year) +
+  geom_point(size = 2, aes(alpha = obs)) +
+  #geom_point(size = 2, alpha = 0.75) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
@@
 
-## Define functions
-```{r, echo = FALSE}
+# Define functions
+```{r}
 # Plot ordisurf
 ## Based on https://oliviarata.wordpress.com/2014/07/17/ordinations-in-ggplot2-v2-ordisurf/
-surfplot <- function(x, y, min_score = 0.25) {
+surfplot <- function(x, y, min_score = 0.2) {
 
@@
   scores <- y %>% 
-    #filter((abs(MDS1) > 0.25 | abs(MDS2) > 0.25) | Score == "BB.sp")
-    filter((abs(MDS1) > 0.1 | abs(MDS2) > 0.1) | Score == "BB.sp")
+    filter((abs(MDS1) > min_score | abs(MDS2) > min_score) | Score == "BB.sp") %>%
+    mutate(max.score = max(MDS1, MDS2))
 
   ggplot(scores, aes(MDS1, MDS2, pch = Score)) +
-    geom_contour(data = contours, aes(MDS1, MDS2, z = z, color = stat(level)), size = 1.5, alpha = 0.75, inherit.aes = FALSE) +
+    geom_contour(data = contours, 
+                 aes(MDS1, MDS2, z = z, color = stat(level)),
+                 size = 1.5, alpha = 0.75, inherit.aes = FALSE) +
     geom_point(size = 2) +
     geom_text_repel(aes(MDS1, MDS2, label = Label), key_glyph = "rect") +
+    #geom_text(aes(MDS1, MDS2, label = Label, alpha = max.score), key_glyph = "rect") +
     theme_light() +
@@
 ## You get the idea. You should anyway -- you wrote it. https://tidyeval.tidyverse.org/multiple.html
-niche_analysis <- function(x, y, criterion, ...) {
+niche_analysis <- function(x, y, criterion, k, ...) {
 
@@
   x %>%
-    filter(caste != "male") %>%
-    #filter(!bb.sp %in% c("psit", "gers", "humi")) %>% # omit parasitic species, gers because it's a specialist, humi because it's so rare
-    filter(!bb.sp %in% c("gers", "humi")) %>% # gers because it's a specialist, humi because it's so rare
-    #filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
-    group_by(site, date, bb.sp, !!criterion, ...) %>% # pooling all sites and years
-    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
+    filter(caste != "male") %>% # exclude males
+    filter(!bb.sp %in% c("gers", "humi")) %>% # exclude gers because it's a specialist, humi because it's so rare
+    group_by(site, date, bb.sp, !!criterion, ...) %>% 
+    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
+    group_by(site, date, ...) %>%
+    complete(bb.sp, !!criterion, fill = list(freq = 0)) %>% # fill in zeros for unvisited plants
     group_by(site, date, bb.sp) %>%
-    mutate(prop = freq/sum(freq)) %>%
-    group_by(site, date, ...) %>%
-    complete(bb.sp, !!criterion, fill = list(prop = 0)) %>% # fill in zeros for unvisited plants
+    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair per site-date 
     group_by(..., bb.sp, !!criterion) %>%
-    summarize(prop = mean(prop)) %>%
-    #dplyr::select(-freq) %>% # drop freq variable
+    summarize(prop = mean(prop)) %>% # get mean proportional intneraction frequency aggregated by ...
     group_by(...) %>%
@@
                           ),
-           ordisurf = map2(pcoa, traits, function(x, y) ordisurf(x ~ pbl.w, data = y, k = 8)),
-           #ordisurf = map2(pcoa, traits, function(x, y) ordisurf(x ~ pbl.w, data = y, k = 5)),
+           ordisurf = map2(pcoa, traits, function(x, y) ordisurf(x ~ pbl.w, data = y, k = k)),
            ordisurf.dev_exp = map(ordisurf, function(x) summary(x)$dev.expl),
@@
 }
-
-# Calculate site-wise beta diversity from visitation data
-beta_func <- function(x, traits, sitedata, level, criterion, binary, ...) {
-
-  criterion <- enquo(criterion) # plant.sp or k.type.s
-  level <- enquo(level) # bb.sp or pbl.fac (i.e. tongue length  guild)
-
-  x %>% 
-    filter(caste != "male") %>%
-    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
-    left_join(traits) %>% # add BB traits
-    group_by(site, !!level, !!criterion, ...) %>% # in addition to grouping by level (bb.sp/pbl.fac) and criterion (plant.sp/k.type.s), the ...
-argument will also group by any variables supplied in the function call, e.g.
-elev.class2, period, year.
-    summarize(freq = n()) %>% # get visitation frequency
-    group_by(site, !!level, ...) %>%
-    mutate(prop = freq/sum(freq)) %>% # get visitation proportions
-    group_by(!!level, ...) %>%
-    select(-freq) %>%
-    nest() %>% # nest into list-column
-    mutate(comm = map(data, function(x) pivot_wider(x, names_from = !!criterion, values_from = prop, values_fill = 0)), # convert to community matrix
-           comm2 = map(comm, function(x) column_to_rownames(x, var = "site")), # column to rowname
-           dist = map(comm2, function(x) vegdist(x, method = "horn", binary = binary)), # conver to Horn's distance matrix
-           tidy.dist = map(dist, tidy), # convert matrix to tidy array
-           elevation = map(comm, function(x) left_join(x, sitedata, by = "site") %>% # add elevation data
-                             select(site, elev.mean) %>% 
-                             column_to_rownames(var = "site")),
-           adonis = map2(dist, elevation, function(x, y) adonis(x ~ elev.mean, data = y)), # run adonis test
-           adonis.r2 = map(adonis, function(x) x$aov.tab$R2[[1]]), # extract adonis r2
-           adonis.p = map(adonis, function(x) x$aov.tab$'Pr(>F)'[[1]])) %>% # extract adonis p-val
-    unnest(c(tidy.dist, adonis.r2, adonis.p)) %>% 
-    select(..., !!level, site1 = item1, site2 = item2, distance, adonis.r2, adonis.p) %>%
-    left_join(select(site_data, site, elev.mean), by = c("site1" = "site")) %>%
-    rename(elev1 = elev.mean) %>%
-    left_join(select(site_data, site, elev.mean), by = c("site2" = "site")) %>%
-    rename(elev2 = elev.mean) %>%
-    mutate(elev.diff = abs(elev1 - elev2)) 
-}
-
-# Calculate site-wise beta diversity from survey data
-beta_func_fl <- function(x, sitedata, criterion, binary, ...) {
-
-  criterion <- enquo(criterion) # plant.sp or k.type.s
-
-  x %>% 
-    group_by(site, !!criterion, ...) %>% # 
-    summarize(abund = sum(flower.cover)) %>% # get visitation frequency
-    group_by(site, ...) %>%
-    mutate(prop = abund/sum(abund)) %>% # get visitation proportions
-    select(-abund) %>%
-    pivot_wider(names_from = !!criterion, values_from = prop, values_fill = 0) %>%
-    column_to_rownames(var = "site") %>%
-    vegdist(method = "horn", binary = binary) %>%
-    tidy() %>%
-    select(..., site1 = item1, site2 = item2, distance) %>%
-    left_join(select(site_data, site, elev.mean), by = c("site1" = "site")) %>%
-    rename(elev1 = elev.mean) %>%
-    left_join(select(site_data, site, elev.mean), by = c("site2" = "site")) %>%
-    rename(elev2 = elev.mean) %>%
-    mutate(elev.diff = abs(elev1 - elev2)) 
-}
-
-# Niche overlap by sp.pair * site * year
-niche_overlapper <- function(x, criterion) {
-
-  criterion <- enquo(criterion) # criterion is set to either plant.sp (for taxonomic) or k.type.s (for morphological)
-
-  x %>%
-    filter(caste != "male") %>%
-    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
-    group_by(bb.sp, !!criterion, site, year) %>% # group by site-year and BB-FL pair
-    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-year
-    group_by(bb.sp, site, year) %>% # group by site-year and BB species
-    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
-    dplyr::select(-freq) %>% # drop freq variable
-    pivot_wider(names_from = !!criterion, values_from = prop, values_fill = 0) %>% # spread into the kind of array vegan likes
-    group_by(site, year) %>%
-    nest() %>% # nest into list-column
-    mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
-           dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
-           tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
-    dplyr::select(site, year, tidy_dist) %>% # drop intermediate list-columns
-    unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
-    left_join(site_data) %>% # add site data
-    dplyr::select(site, year, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
-    left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
-    rename(subgenus1 = subgenus, pbl.w1 = pbl.w, pbl.w.class1 = pbl.w.class, pbl.fac1 = pbl.fac) %>%
-    left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
-    rename(subgenus2 = subgenus, pbl.w2 = pbl.w, pbl.w.class2 = pbl.w.class, pbl.fac2 = pbl.fac) %>%
-    dplyr::select(site, year, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
-           pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, pbl.fac1, pbl.fac2, elev.mean) %>%
-    mutate(pbl.diff = abs(pbl.w1 - pbl.w2), # tongue length differences
-           pbl.fac.diff = if_else(pbl.fac1 == pbl.fac2, "same", "different"),
-           subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different"), # binary: same or different subgenus
-
-           # Logical columns indicating whether a given species belongs to a given species pair
-           # This is handy for extracting data for individual species across all their counterparts
-           bss = if_else(bb.sp1 == "bss" | bb.sp2 == "bss", TRUE, FALSE),
-           pasc = if_else(bb.sp1 == "pasc" | bb.sp2 == "pasc", TRUE, FALSE),
-           wurf = if_else(bb.sp1 == "wurf" | bb.sp2 == "wurf", TRUE, FALSE),
-           prat = if_else(bb.sp1 == "prat" | bb.sp2 == "prat", TRUE, FALSE),
-           soro = if_else(bb.sp1 == "soro" | bb.sp2 == "soro", TRUE, FALSE),
-           hort = if_else(bb.sp1 == "hort" | bb.sp2 == "hort", TRUE, FALSE)) %>%
-
-    unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
-    mutate(sp.pair = factor(sp.pair)) %>% # ...and make it a factor so that mgcv can handle it
-    mutate(niche.overlap = 1-distance) # express niche overlap so that 1 = perfect overlap and 0 = no overlap
-}
-
-# Diet breadth analysis
-diet_breit <- function(x, y, z, criterion, indices = c("normalised degree")) {
-
-  criterion <- enquo(criterion) # plant.sp or k.type.s
-
-  a <- x %>%
-    filter(caste != "male") %>%
-    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
-    left_join(y) %>%
-    #group_by(elev.class2) %>%
-    group_by(site) %>%
-    mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) %>%
-    group_by(site, period, pbl.fac, !!criterion) %>% # group by site, date, bee species, plant species
-    summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-    unite(webID, c(site, period), sep = "_") %>%
-    ungroup() %>%
-    dplyr::select(higher = pbl.fac, lower = !!criterion, webID, freq) %>% # rename columns to match bipartite's expectations
-    data.frame() %>% # convert to data frame
-    frame2webs() %>% # convert to to bipartite web
-    map(specieslevel, level = "higher", index = indices) %>%
-    map(rownames_to_column) %>%
-    map(as_tibble) %>%
-    bind_rows(.id = "webID") %>% # Collapse list into big data frame
-    rename(pbl.fac = rowname) %>%
-    separate(webID, c("site", "period"), sep =  "_") %>%
-    left_join(z) %>%
-    filter(pbl.fac %in% c("long", "short")) %>% # for some reason, a few rows have a "1" for pbl.fac, and I don't know where this comes from; this gets rid of it, but I would still like to know why
-    mutate(site = factor(site),
-           period = factor(period, levels = c("early", "mid", "late")))
-
-  b <- x %>%
-    filter(caste != "male") %>%
-    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
-    #group_by(elev.class2) %>%
-    group_by(site) %>%
-    mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) %>%
-    left_join(bb_traits) %>%
-    group_by(site, period, pbl.fac) %>% # group by site, date, bee species, plant species
-    summarize(abund = n()) %>%
-    group_by(site, period) %>%
-    mutate(prop = abund/sum(abund))
-
-  out <- left_join(a, b)
-}
-
-diet_breit2 <- function(x, y, z, criterion, indices = c("normalised degree")) {
-
-  criterion <- enquo(criterion) # plant.sp or k.type.s
-
-  a <- x %>%
-    filter(caste != "male") %>%
-    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
-    left_join(y) %>%
-    # group_by(elev.class2) %>%
-    # mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) %>%
-    #group_by(site, period, pbl.fac, !!criterion) %>% # group by site, date, bee species, plant species
-    group_by(site, pbl.fac, !!criterion) %>% # group by site, date, bee species, plant species
-    summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-    #unite(webID, c(site, period), sep = "_") %>%
-    ungroup() %>%
-    #dplyr::select(higher = pbl.fac, lower = !!criterion, webID, freq) %>% # rename columns to match bipartite's expectations
-    dplyr::select(higher = pbl.fac, lower = !!criterion, webID = site, freq) %>% # rename columns to match bipartite's expectations
-    data.frame() %>% # convert to data frame
-    frame2webs() %>% # convert to to bipartite web
-    map(specieslevel, level = "higher", index = indices) %>%
-    map(rownames_to_column) %>%
-    map(as_tibble) %>%
-    bind_rows(.id = "webID") %>% # Collapse list into big data frame
-    #rename(pbl.fac = rowname) %>%
-    rename(pbl.fac = rowname,
-           site = webID) %>%
-    #separate(webID, c("site", "period"), sep =  "_") %>%
-    left_join(z) %>%
-    filter(pbl.fac %in% c("long", "short")) %>% # for some reason, a few rows have a "1" for pbl.fac, and I don't know where this comes from; this gets rid of it, but I would still like to know why
-    # mutate(site = factor(site),
-    #        period = factor(period, levels = c("early", "mid", "late")))
-    mutate(site = factor(site))
-
-  b <- x %>%
-    filter(caste != "male") %>%
-    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
-    # group_by(elev.class2) %>%
-    # mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) %>%
-    left_join(bb_traits) %>%
-   # group_by(site, period, pbl.fac) %>% # group by site, date, bee species, plant species
-    group_by(site, pbl.fac) %>% # group by site, date, bee species, plant species
-    summarize(abund = n()) %>%
-    #group_by(site, period) %>%
-    group_by(site) %>%
-    mutate(prop = abund/sum(abund))
-
-  filter(caste != "male") %>%out <- left_join(a, b)
-}
-
-# Null model generator
-nullify1 <- function(x, n, criterion){
-
-  criterion <- enquo(criterion) # plant.sp or k.type.s
-
-  x %>%
-    filter(caste != "male") %>%
-    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
-    select(site, bb.sp, !!criterion) %>% 
-    group_by(site, bb.sp, !!criterion) %>%
-    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
-    group_by(site) %>%
-    nest() %>% # nest data into list-column grouped by pollen, site, date
-    mutate(data = map(data, function(x) pivot_wider(x, 
-                                                    names_from = !!criterion, 
-                                                    values_from = freq, 
-                                                    values_fill = 0) %>% column_to_rownames("bb.sp")), # turn table into community matrix
-           data = map(data, function(x) vegan::permatfull(x, 
-                                                          fixedmar = "rows", 
-                                                          shuffle = "ind", 
-                                                          times = n)), # permute community matrices
-           data = map(data, function(x) return(x$perm))) %>% # extract permutations from permat objects
-    unnest(data) %>% # unnest
-    mutate(data = map(data, function(x) vegdist(x, method = "horn")), # convert each permutation into distance matrix
-           data = map(data, tidy)) %>% # convert distance matrices to tidy arrays
-    group_by(site) %>%
-    mutate(iter = row_number()) %>% # add index to denote iterations
-    unnest(cols = data) %>% # unnest again
-    rename(bb.sp1 = item1, bb.sp2 = item2) %>%
-    unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
-    mutate(sp.pair = factor(sp.pair), # ...and make it a factor so that mgcv can handle it
-           niche.overlap = 1-distance) %>% # express niche overlap so that 1 = perfect overlap and 0 = no overlap
-    select(site, sp.pair, iter, niche.overlap.sim = niche.overlap)
-}
-
-# Null model generator
-nullify2 <- function(x, n, criterion){
-
-  criterion <- enquo(criterion) # plant.sp or k.type.s
-
-  x %>%
-    filter(caste != "male") %>%
-    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
-    select(site, bb.sp, !!criterion) %>% 
-    group_by(site, bb.sp, !!criterion) %>%
-    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
-    group_by(site) %>%
-    nest() %>% # nest data into list-column grouped by pollen, site, date
-    mutate(data = map(data, function(x) pivot_wider(x, 
-                                                    names_from = !!criterion, 
-                                                    values_from = freq, 
-                                                    values_fill = 0) %>% column_to_rownames("bb.sp")), # turn table into community matrix
-           data = map(data, function(x) vegan::permatfull(x, 
-                                                          fixedmar = "both", 
-                                                          shuffle = "ind", 
-                                                          times = n)), # permute community matrices
-           data = map(data, function(x) return(x$perm))) %>% # extract permutations from permat objects
-    unnest(data) %>% # unnest
-    mutate(data = map(data, function(x) vegdist(x, method = "horn")), # convert each permutation into distance matrix
-           data = map(data, tidy)) %>% # convert distance matrices to tidy arrays
-    group_by(site) %>%
-    mutate(iter = row_number()) %>% # add index to denote iterations
-    unnest(cols = data) %>% # unnest again
-    rename(bb.sp1 = item1, bb.sp2 = item2) %>%
-    unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
-    mutate(sp.pair = factor(sp.pair), # ...and make it a factor so that mgcv can handle it
-           niche.overlap = 1-distance) %>% # express niche overlap so that 1 = perfect overlap and 0 = no overlap
-    select(site, sp.pair, iter, niche.overlap.sim = niche.overlap)
-}
-```
-
-# 1. What is the logic of niche partitioning between core species, and does it change through elevation and time?
-
-For ordination-based analyses, I have opted to pool across years to emphasize species relationships rather than interannual differences.
-To aid interpretation, I have classified BB species by tongue length.
-The two "true" long-tongued species are *B.
-pascuorum* and *B.
-hortorum*, with the latter having a considerably longer tongue than the former, but I have opted to consider *B.
-wurflenii* as "behaviorally long-tongued" because it characteristically forages as a robber, and evidently converges in flower choice with *B.
-pascuorum* and *B.
-hortorum*.
-This decision was made *post-hoc* after noting the clustering patterns.
-
-I have also opted to not distinguish between putative pollen and putative nectar foragers.
-First, the classification of nectar vs.
-pollen foragers based on whether they are carrying pollen at the time of observation is imperfect (though probably still informative).
-Second, and more importantly in my opinion, the more I break the data up, the more difficult it becomes to visualize and interpret the data.
-Finally, when I ran the analysis parsing by pollen vs.
-nectar foraging, the patterns tended to be similar between the two.
-
-Note that in this version of the analysis, I am not excluding observations that were unclassified with respect to pollen carrying.
-A closer inspection of the data revealed that this is simply how Kalli handled males and Psithyrus queens, since they never collect pollen.
-I am omitting Psithyrus from my analysis, but I am not including the males of the focal specis.
-I suppose an argument could be made that parsing by caste would be informative, but again, there are so many ways these data could be parsed, and I can't use all of them without whittling the data down to something intractable.
-
-Kugler classes are coded in the form of number.decimal.letter.letter. I have chosen to simplify to the level of number.decimal.
-
-To avoid overplotting plant species / Kugler classes, I have only plotted plant species scoring above a given threshold on either the first or second axis.
-Thus, the plotted species are those that contribute most to the separation, not necessarily those that contribute most to diet.
-A species that is visited heavily by all BB species will fall close to (0,0) and will therefore not be plotted.
-This is because they are added based on weighted averaging.
-i.e.
-they get "pulled" toward BB species in proportion to their (relative) abundance (i.e.
-visitation rate) for each species.
-
-## 1.1 Ordination by species with post-hoc fitting of tongue length  
-### Script
+```
+
+# 1.1 Ordination by species with post-hoc fitting of tongue length  
 ```{r message=FALSE, warning=FALSE, echo = FALSE}
 # By plant.sp
-niche.tax <- niche_analysis(net, bb_traits, criterion = plant.sp.abb)
+niche.tax <- niche_analysis(net, bb_traits, criterion = plant.sp.abb, k = 8)
 niche.tax.plot <- niche.tax$ordisurf.plot[[1]]
@@
 # By ktype
-niche.morph <- niche_analysis(net, bb_traits, criterion = k.type.s)
+niche.morph <- niche_analysis(net, bb_traits, criterion = k.type.ss, k = 8)
 niche.morph.plot <- niche.morph$ordisurf.plot[[1]]
@@
 # By color
-niche.color <- niche_analysis(net, bb_traits, criterion = color)
+niche.color <- niche_analysis(net, bb_traits, criterion = color, k = 8)
 niche.color.plot <- niche.color$pcoa.plot[[1]]
@@
 dev.off()
-
-```
-
-### Results
-
-In parentheses following the initial mention of each MDS axis, the variance explained is shown for the taxonomic (T:) and morphological (M:) ordinations.
-
-1.
-MDS1 (T:48%, M:86%) separates long-tongued from short--tongued species.
-This mirrors the findings of Goulson et al.
-(2008).
-MDS2 (T:27%, M:12%) mainly accounts for the divergent foraging behavior of *B.
-sensu-stricto*.
-I wonder if this has been described before?
-A caveat to the interpretation of this pattern, though, is that *B.
-sensu-stricto* exhibited some weird temporal abundance patterns, particularly in 2012.
-It's possible that this biased its foraging in a way that could account for some of its apparent deviance.
-
-2.
-LT species cluster around the lip flowers *Salvia glutinosa* and *Rhinanthus glacialis* (Kugler 5.1) and the flag flower *Anthyllis vulneraria* (Kugler 6.1).
-ST species *B.
-soroensis* and *B.
-pratorum* cluster around the bell flower *Campanula scheuchzeri* (Kugler 3.2) and flower heads *Carduus defloratus* and *Centaurea jacea* (Kugler 7.2).
-The short-tongued *B.
-sensu-stricto* seems to be especially associated with the bell flower *Erica carnea* (Kugler 3.1), which scores near 0 on MDS1 (i.e.
-not accounting for the separation of tongue-legnth guilds) but strongly negative on MDS2, mirroring the scoring *B.
-sensu-stricto*.
-The lip flower *Stachys alopecuros* (Kugler 5.1) also scores near 0 on MDS1 but high on MDS2, and together with *Erica carnea* seems to contribute to the separation of *B.
-sensu-strictu* from the other species of both tongue classes.
-
-## 1.2 Ordination by species with post-hoc fitting of tongue length, broken down by elevation class
-### Script
+```
+
+# 1.2 Ordination by species with post-hoc fitting of tongue length, broken down by elevation class
 ```{r message=FALSE, warning=FALSE, echo = FALSE}
 # Taxonomic
-niche_x_elev.tax <- niche_analysis(net, bb_traits, criterion = plant.sp.abb, elev.class2)
+niche_x_elev.tax <- niche_analysis(net, bb_traits, criterion = plant.sp.abb, k = 8, elev.class2)
 
@@
 # Morphological
-niche_x_elev.morph <- niche_analysis(net, bb_traits, criterion = k.type.ss, elev.class2)
+niche_x_elev.morph <- niche_analysis(net, bb_traits, criterion = k.type.ss, k = 8, elev.class2)
 
@@
 # Color
-niche_x_elev.color <- niche_analysis(net, bb_traits, criterion = color, elev.class2)
+niche_x_elev.color <- niche_analysis(net, bb_traits, criterion = color, k = 8, elev.class2)
 
@@
 
-### Results  
-1.
-At low elevation, MDS1 (T:63%, M:64%) does *not* clearly separate LT from ST species; instead, it seems to be driven by the strong association between *B.
-soroensis* and the flower head *Centaurea jacea* (Kugler 7.2), which score high on MDS1.
-MDS2 (T:34%.
-M:35%) seems to be driven by the associated of *B.
-pratorum* and *B.
-hortorum* with the stalk-disc flower *Primula elatior* (Kugler 4.1).
-In general, though, the BB species at low elevation do not seem to cluster into clear groups.
-
-2.
-At mid elevation, MDS1 (T:48%, M:79%) seems to be strongly driven by tongue-length.
-LT species are associated with the lip flowers *Salvia glutinosa* and *Rhinanthus glacialis* (Kugler 5.1) and the flag flower *Anthyllis vulneraria* (Kugler 6.1).
-ST species are associated with the flower head *Centaurea jacea* (Kugler 7.2).
-MDS2 (T:25%, M:16%) seems to be driven by the divergent foraging of *B.
-sensu-stricto*, particularly its association with the bell flower *Erica carnea* (Kugler 3.1), though *E.
-carnea* plots centrally because it was also visited by other BB species.
-
-3.
-At high elevation, MDS1(T:52%, M:79%) again separates LT and ST species (though not as strongly as at mid elevation) while MDS2 (T:38%, M:18%) separates *B.
-sensu-stricto*.
-In this case, though, *B.
-sensu-stricto* seems to be especially associated with the disk flower *Helianthemum nummularium* (Kugler 1.1).
-LT species are associated with the lip flower *Stachys alopecuros* (Kugler 5.1) and the flag flower *Anthyllis vulneraria* (Kugler 6.1), while the ST species *B.
-pratorum* and *B.
-soroensis* are associated with the bell flower *Campanula scheuchzeri* (Kugler 3.2) and flower head *Carduus defloratus* (Kugler 7.2).
-
-## 1.3 Ordination by BB species * elev.class2 * period, pooling across years
-### Script
+# 1.3 Ordination by BB species * elev.class2 * period, pooling across years
 ```{r message=FALSE, warning=FALSE, echo = FALSE}
 # Taxonomic
-niche_x_elevperiod.tax <- niche_analysis(net, bb_traits, criterion = plant.sp.abb, elev.class2, period)
+niche_x_elevperiod.tax <- niche_analysis(net, bb_traits, criterion = plant.sp.abb, k = 7, elev.class2, period)
 
@@
 niche_high_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "high" & period == "early")$ordisurf.plot[[1]] 
-# niche_high_mid1.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "high" & period == "mid1")$ordisurf.plot[[1]] 
-# niche_high_mid2.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "high" & period == "mid2")$ordisurf.plot[[1]]
 niche_high_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "high" & period == "mid")$ordisurf.plot[[1]]
@@
 niche_mid_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "mid" & period == "early")$ordisurf.plot[[1]]
-# niche_mid_mid1.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "mid" & period == "mid1")$ordisurf.plot[[1]] 
-# niche_mid_mid2.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "mid" & period == "mid2")$ordisurf.plot[[1]]
 niche_mid_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "mid" & period == "mid")$ordisurf.plot[[1]]
@@
 niche_low_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "low" & period == "early")$ordisurf.plot[[1]]
-# niche_low_mid1.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "low" & period == "mid1")$ordisurf.plot[[1]]
-# niche_low_mid2.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "low" & period == "mid2")$ordisurf.plot[[1]]
 niche_low_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "low" & period == "mid")$ordisurf.plot[[1]]
@@
 # Morphological
-niche_x_elevperiod.morph <- niche_analysis(net, bb_traits, criterion = k.type.s, elev.class2, period)
+niche_x_elevperiod.morph <- niche_analysis(net, bb_traits, criterion = k.type.ss, k = 7, elev.class2, period)
 
@@
 niche_high_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "high" & period == "early")$ordisurf.plot[[1]]
-niche_high_mid1.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "high" & period == "mid1")$ordisurf.plot[[1]] 
-niche_high_mid2.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "high" & period == "mid2")$ordisurf.plot[[1]] 
+niche_high_mid.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "high" & period == "mid")$ordisurf.plot[[1]] 
 niche_high_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "high" & period == "late")$ordisurf.plot[[1]]
@@
 niche_mid_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "mid" & period == "early")$ordisurf.plot[[1]]
-niche_mid_mid1.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "mid" & period == "mid1")$ordisurf.plot[[1]]
-niche_mid_mid2.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "mid" & period == "mid2")$ordisurf.plot[[1]] 
+niche_mid_mid.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "mid" & period == "mid")$ordisurf.plot[[1]] 
 niche_mid_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "mid" & period == "late")$ordisurf.plot[[1]]
@@
 niche_low_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "low" & period == "early")$ordisurf.plot[[1]]
-niche_low_mid1.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "low" & period == "mid1")$ordisurf.plot[[1]] 
-niche_low_mid2.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "low" & period == "mid2")$ordisurf.plot[[1]] 
+niche_low_mid.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "low" & period == "mid")$ordisurf.plot[[1]] 
 niche_low_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "low" & period == "late")$ordisurf.plot[[1]]
@@
 # Color
-niche_x_elevperiod.color <- niche_analysis(net, bb_traits, criterion = k.type.s, elev.class2, period)
+niche_x_elevperiod.color <- niche_analysis(net, bb_traits, criterion = color, k = 7, elev.class2, period)
 
@@
 niche_high_early.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "high" & period == "early")$pcoa.plot[[1]]
-niche_high_mid1.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "high" & period == "mid1")$pcoa.plot[[1]] 
-niche_high_mid2.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "high" & period == "mid2")$pcoa.plot[[1]] 
+niche_high_mid.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "high" & period == "mid")$pcoa.plot[[1]] 
 niche_high_late.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "high" & period == "late")$pcoa.plot[[1]]
@@
 niche_mid_early.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "mid" & period == "early")$pcoa.plot[[1]]
-niche_mid_mid1.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "mid" & period == "mid1")$pcoa.plot[[1]]
-niche_mid_mid2.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "mid" & period == "mid2")$pcoa.plot[[1]] 
+niche_mid_mid.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "mid" & period == "mid")$pcoa.plot[[1]]
 niche_mid_late.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "mid" & period == "late")$pcoa.plot[[1]]
@@
 niche_low_early.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "low" & period == "early")$pcoa.plot[[1]]
-niche_low_mid1.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "low" & period == "mid1")$pcoa.plot[[1]] 
-niche_low_mid2.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "low" & period == "mid2")$pcoa.plot[[1]] 
+niche_low_mid.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "low" & period == "mid")$pcoa.plot[[1]] 
 niche_low_late.color.plot <- filter(niche_x_elevperiod.color, elev.class2 == "low" & period == "late")$pcoa.plot[[1]]
@@
 ```
-
-### Results
-
-This is probably the most interpretable analysis.
-One thing I like about it is how it reveals important floral species that didn't score highly in the aggregate analysis.
-But there is also a lot to say.
-
-1.
-*Low elevation*: The most salient pattern is a lack of one -- tongue length does not seem to be important at low elevation *until late in the season*, when all the short ST bees agree that *Centaurea jacea* is the only game in town.
-
-  - *Early*: MDS1 (T:98%, M:97%) dominates the ordination and mainly separates *B.
-soroensis*, associated with the bell flower *Geum rivale* (Kugler 3.2) from the other BB species, associated with the stalk-disc flower *Primula elatior* (Kugler 4.1).
-MDS2 (T:2%, M:2%) accounted  for almost no variation.
-
-  - *Mid*: MDS1 (T:90%, M:94%) again dominates the ordination, apparently representing a gradient of association ranging from the lip flower *Rhinanthus alectorolophus* (Kugler 5.1) and the flag blossom *Trifolium pratense* (Kugler 6.1) on the  low end of the axis to the flower head *Knautia arvensis* (Kugler 7.1).
-BB species are spread fairly evenly along this gradient, with *B.
-sensu-stricto* and *B.
-wurflenii* leaning hardest in the *Rhinanthus/Trifolium* direction and *B.
-pratorum* and *B.
-pascuorum* leaning hardest in the *Knautia* direction.
-Interestingly, the little variance MDS2 (T:9%, M:6%) explains seems to be associated with tongue length, though because MDS2 accounts for so little variance, we must conclude that tongue length is of little importance in structuring the visitation network in this elevation-time slice.
-
-  - *Late*: MDS1 (T:51%, M:70%) mainly separates *B.
-wurflenii* from the rest of the BB species, apparently due to its strong association with the flag blossom *Lotus corniculatus* (Kugler 6.1) and the funnel flower *Gentiana aspera* (Kugler 2.1).
-MDS2 (T:41%, M:27%), which matters again, seems to separate LT ans ST species, though the pattern is not so much one of separation as of dispersion.
-All the ST species are huddled tightly around the flower head *Centaurea jacea* (Kugler 7.2), while the LT species are rather spread out, with *B.
-wurflenii* as described, *B.
-hortorum* expressing a distinct taste for the lip flower *Stachys officinalis* (Kugler 5.1), and *B.
-pascuorum* middling as usual.
-
-2. *Mid elevation*:
-
-  - *Early*: MDS1 (T:52%, M:68%) is clearly the tongue-length axis.
-ST species are all clustered around the bell flower *Erica carnea* (Kugler 3.1), while the LT species are more spread out, with *B.
-wurflenii* commidd to the flag blossom *Polygala chamaebuxus*, *B.
-hortorum* more interested in the stalk-disk flower *Gentiana verna* (Kugler 4.1), and *B.
-pascuorum* falls in between (though a lot closer to *B.
-wurflenii*), evidentally pulled a bit toward the the stalk-disc flower *Primula elatior* (Kugler 4.1).
-MDS2 (T:32%, M:29%) seems to be driven strongly by *B.
-hortorum* divergence from the other two LT species.
-
-  - *Mid*: MDS1 (T:49%, M:68%) separates *B.
-soroensis* and *B.
-pratorum* from the other BB species.
-The former are associated with the flower head *Phyteuma orbiculare* (Kugler 7.1) and the bell flower *Geum rivale* (Kugler 3.2).
-The latter consist of a fairly tight cluster of *B.
-hortorum*, *B.
-pascuorum*, and *B.
-wurflenii* around the flag flowers *Anthyllis vulneraria* and *Lotus corniculatus* (Kugler 6.1) and the lip flower *Ajuga reptans* (Kugler 5.1).
-MDS2 (T:26%, M:17%), together with MDS1, accounts for the uniqueness of *B.
-sensu-strictu* foraging, with that species plotting both very low on MDS1 and very high on MDS2.
-Interestingly, none of the plant species plot close to it, so it must exceptional not because it forages on something unique but because it forages a little bit on everything.
-That's the  only explanation I can come up with, anyway, for why the weighted-average plotting pulls plants away from *B.
-sensu-stricto*.
-The apparent exceptionality of *B.
-sensu-stricto* is a theme in this analysis.
-
-  - *Late*: MDS1 (T:67%, M:98%) sort of separates LT and ST species, except that *B.
-pratorum* and *B.
-pascuorum* fall close to each other on this axis (but are separated widely on MDS2).
-The LT species are fairly well-clustered around the funnel flower *Gentiana aspera* (Kugler 2.1), the lip flowers *Salvia glutinosa* and *Rhinanthus glacialis*.
-The ST species are spreadwidely along MDS2(T:18%, M:1%), with *B.
-sensu-stricto* and *B.
-pratorum* associated with the lip flower *Origanum vulgare* and *B.
-soroensis* (Kugler 5.1) favoring the flower head *Centaurea jacea* (Kugler 7.2).
-
-3. *High elevation*:
-  - *Early*: MDS1 (T:54%, M:74%) is sort of tongue-length axis, but not perfectly, again because *B.
-sensu-stricto* is so abberant.
-In general, the species are fairly widespread with distincntfloral associations -- consistent with the observed decline in overall niche overlap at high elevation -- with the exception of *B.
-soroensis* and *B.
-pratorum*, which cluster tightly around the bell flower *Rhododendron hirsutum* (Kugler 3.1).
-*B.
-hortorum* likes the flag flower *Anthyllis vulneraria* (Kugler 6.1), *B.
-wurflenii* like the flag flower *Lotus corniculatus* (Kugler 6.1), *B.
-pascuorum* (middling as usual) seems to be associated with both the flag flower *Trifolium pratense* (Kugler 6.1) and the bell flower *Rhododendron hirsutum* (Kugler 3.1), and *B.
-sensu-stricto* seems slightly partial to disk flower *Trollius europaeus* (Kugler 1.2), though again its position suggests that it is exceptional in its generalism.
-MDS2 (T:28%, M:23%) is mostly about accounting for the deviance of *B.
-sensu-stricto*.
-
-  - *Mid*: By mid-season at high elevation, the LT species are tightly clustered with each other and associated with the flag blossom *Trifolium pratense* (Kugler 6.1) and the lip flower *Stachys alopecuros* (Kugler 5.1).
-The ST species are widely spread, with *B.
-soroensis* and *B.
-pratorum* suggetsing some affinity for the flower head *Carduus defloratus* (Kugler 7.2) and the bell flower *Campanula scheuchzeri* (Kugler 3.2).
-*B.
-pratorum* plots similarly close to *T.
-pratense*.
-Again, *B.
-sensu-stricto* is off on its own with no defining floral association.
-MDS1 (T:64%, M:91%), MDS2(T:35%, M:9%).
-
-  - *Late*: Finally, late in the season at high elevation, LT and ST species are separated diagonally with respect to MDS1 (T:75%, M:79%) and MDS2(T:20%, M:18%), with LT species falling high on one or both axes and ST species falling low on one or both axes.
-LT species are loosely associated with the flag blossom *Trifolium pratense* (Kugler 6.1), the funnel flowers *Gentiana asclepiadea* and congener *Gentiana aspera* (Kugler 2.1) and the flower head **(unusual for LT to be associated with flower head!)** *Carlina acaulus* (Kugler 7.2).
-*B.
-soroensis* and *B.
-pratorum* are associated with the flower heads *Carduus defloratus* (Kugler 7.2) and *Scabiosa lucida* (Kugler 7.1) and the bell flower *Campanula scheuchzeri* (Kugler 3.2).
-*B.
-sensu-stricto*, ever off on its own, was most closely associated with the lip flower *Thymus pulegioides* (Kugler 5.1).
-
-## 1.4 A follow up to the question of diet breadth in long- vs. short-tongued BB species
-The question is which guild -- long-tongued or short-tongued species -- uses greater proportion of available floral species.
-We will keep this really, really simple, and calculate the normalized degree for each guild.
-That is just the number of floral species visited divided by the total number of floral species in the network.
-It ranges between 0 and 1, and 0 and 1 mean exactly what you would expect them to mean.
-I guess we'll go ahead and calculate the raw degree, too, just so that we can see the absolute diet breadth.
-
-For starters, I'm going to use diet_breit2, which does not parse by period.
-
-**This dovetails with Goulson's hypothesis that long-tongued bees are more specialized than short-tongued bees.** I can test this.
-
-### Script
-```{r}
-# Calculate die breadth metrics
-diet_breadth.tax <- diet_breit2(net, bb_traits, site_data, criterion = plant.sp,
-                               indices = c("degree", "normalised degree")) %>%
-  mutate(normalised.degree.abundnorm = normalised.degree/abund,
-         normalised.degree.propnorm = normalised.degree/prop,
-         degree.abundnorm = degree/abund,
-         degree.propnorm = degree/prop)
-
-diet_breadth.morph <- diet_breit2(net, bb_traits, site_data, criterion = k.type.s,
-                               indices = c("degree", "normalised degree")) %>%
-  mutate(normalised.degree.abundnorm = normalised.degree/abund,
-         normalised.degree.propnorm = normalised.degree/prop,
-         degree.abundnorm = degree/abund,
-         degree.propnorm = degree/prop)
-
-# Relative diet breadth ~ elevation
-ggplot(diet_breadth.tax, aes(elev.mean, normalised.degree, color = pbl.fac)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))
-
-ggplot(diet_breadth.morph, aes(elev.mean, normalised.degree, color = pbl.fac)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))
-
-# Absolute diet breadth ~ elevation
-ggplot(diet_breadth.tax, aes(elev.mean, degree, color = pbl.fac)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))
-
-ggplot(diet_breadth.morph, aes(elev.mean, degree, color = pbl.fac)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))
-
-# Relative diet breadth / abundance
-ggplot(diet_breadth.tax, aes(elev.mean, normalised.degree.abundnorm, color = pbl.fac)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))
-
-ggplot(diet_breadth.morph, aes(elev.mean, normalised.degree.abundnorm, color = pbl.fac)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))
-
-# Relative diet breadth / proportional abundance
-ggplot(diet_breadth.tax, aes(elev.mean, normalised.degree.propnorm, color = pbl.fac)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))
-
-ggplot(diet_breadth.morph, aes(elev.mean, normalised.degree.propnorm, color = pbl.fac)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))
-
-# Absolute diet breadth / abundance
-ggplot(diet_breadth.tax, aes(elev.mean, degree.abundnorm, color = pbl.fac)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))
-
-ggplot(diet_breadth.morph, aes(elev.mean, degree.abundnorm, color = pbl.fac)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))
-
-# Absolute diet breadth / proportional abundance
-ggplot(diet_breadth.tax, aes(elev.mean, degree.propnorm, color = pbl.fac)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))
-
-ggplot(diet_breadth.morph, aes(elev.mean, degree.propnorm, color = pbl.fac)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))
-
-# BB abundance ~ elevation
-ggplot(diet_breadth.tax, aes(elev.mean, abund, color = pbl.fac)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))
-
-# BB proportional abundance ~ elevation
-ggplot(diet_breadth.tax, aes(elev.mean, prop, color = pbl.fac)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))
-
-# Absolute diet breadth ~ abundance
-ggplot(diet_breadth.tax, aes(abund, degree, color = pbl.fac)) +
-  geom_point() +
-  geom_smooth(method = "lm")
-
-# Relative diet breadth ~ abundance
-ggplot(diet_breadth.tax, aes(abund, normalised.degree, color = pbl.fac)) +
-  geom_point() +
-  geom_smooth(method = "lm")
-
-# Relative diet breadth ~ proportional abundance
-ggplot(diet_breadth.tax, aes(prop, normalised.degree, color = pbl.fac)) +
-  geom_point() +
-  geom_smooth(method = "lm")
-```
-### Results
-In a nutshell, proportional diet breadth is extremely tightly correlated with proportional anbundance.
-The story here is that LT species are more abundant than ST species at lower elevations, and ST species are more abundant than LT species at higher elevations.
-This relationship is strong and basically linear, with the relative abundances crossing at mid-elevation.
-
-## 1.5 Another follow-up. How does the morphological composition of the plant community change with elevation? 
-To make the graphs readable, we'll simplify Kugler classes to integers.
-
-### Script
-```{r}
-# Define periods based on the network data
-periods <- net %>%
-  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
-  #group_by(elev.class2) %>%
-  group_by(site) %>%
-  mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) %>%
-  ungroup() %>%
-  select(site, date, period) %>%
-  distinct()
-
-# Calculate the absolute and relative abundance of each Kugler class ~ site * period
-
-## Floral survey (constrained to visited species)
-survey_morph_abund <- survey %>%
-  left_join(periods) %>%
-  fill(period, .direction = "downup") %>%
-  mutate(k.type.ss = factor(k.type.ss)) %>%
-  group_by(site, period, k.type.ss) %>%
-  summarize(abund = sum(flower.cover)) %>%
-  group_by(site, period) %>%
-  mutate(prop = abund/sum(abund)) %>%
-  left_join(site_data)
-
-## Visitation network
-### Aggregate
-visitation_morph_abund <- net %>%
-  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
-  #group_by(elev.class2) %>%
-  group_by(site) %>%
-  mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) %>%
-  mutate(k.type.ss = factor(k.type.ss)) %>%
-  group_by(site, period, k.type.ss) %>%
-  summarize(abund = n()) %>%
-  group_by(site, period) %>%
-  mutate(prop = abund/sum(abund)) %>%
-  left_join(site_data)
-
-### By BB species
-visitation_morph_abund.sp <- net %>%
-  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
-  #group_by(elev.class2) %>%
-  group_by(site) %>%
-  mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) %>%
-  mutate(k.type.ss = factor(k.type.ss)) %>%
-  group_by(site, period, k.type.ss, bb.sp) %>%
-  summarize(abund = n()) %>%
-  group_by(site, period, bb.sp) %>%
-  mutate(prop = abund/sum(abund)) %>%
-  left_join(site_data)
-
-### Taxonomic by BB species
-visitation_tax_abund.sp <- net %>%
-  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
-  #group_by(elev.class2) %>%
-  group_by(site) %>%
-  mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) %>%
-  mutate(k.type.ss = factor(k.type.ss),
-         bb.sp = factor(bb.sp, levels = c("hort", "pasc", "wurf", "bss", "prat", "soro"))) %>%
-  group_by(period, elev.class2, plant.genus, k.type.ss, plant.family, bb.sp) %>%
-  summarize(abund = n()) %>%
-  group_by(elev.class2, period, bb.sp) %>%
-  mutate(prop = abund/sum(abund)) %>%
-  arrange(elev.class2, period, bb.sp, -prop)
-
-# Morphological composition ~ elevation (survey)
-## Relative abundance
-ggplot(survey_morph_abund, aes(factor(elev.mean), prop, fill = k.type.ss)) +
-  geom_col() +
-  facet_wrap(~period)
-
-## Absolute abundance
-ggplot(survey_morph_abund, aes(factor(elev.mean), abund, fill = k.type.ss)) +
-  geom_col() +
-  facet_wrap(~period)
-
-# Morphological composition ~ elevation (visitation)
-## Relative abundance 
-ggplot(visitation_morph_abund, aes(factor(elev.mean), prop, fill = k.type.ss)) +
-  geom_col() +
-  facet_wrap(~period)
-
-## Relative abundance ~ species
-ggplot(visitation_morph_abund.sp, aes(factor(elev.mean), prop, fill = k.type.ss)) +
-  geom_col() +
-  facet_grid(bb.sp ~ period)
-
-## Absolute abundance ~ species
-ggplot(visitation_morph_abund.sp, aes(factor(elev.mean), abund, fill = k.type.ss)) +
-  geom_col() +
-  facet_grid(bb.sp ~ period)
-
-## Relative abundance ~ species
-ggplot(filter(visitation_tax_abund.sp, period == "early"), aes(factor(elev.mean), prop, fill = names)) +
-  geom_col() +
-  facet_wrap(~bb.sp)
-
-ggplot(filter(visitation_tax_abund.sp, period == "mid"), aes(factor(elev.mean), prop, fill = names)) +
-  geom_col() +
-  facet_wrap(~bb.sp)
-
-ggplot(filter(visitation_tax_abund.sp, period == "late"), aes(factor(elev.mean), prop, fill = names)) +
-  geom_col() +
-  facet_wrap(~bb.sp)
-
-## Absolute abundance ~ species
-ggplot(visitation_tax_abund.sp, aes(factor(elev.mean), abund, fill = names)) +
-  geom_col() +
-  facet_grid(bb.sp ~ period)
-```
-
-### Results
 
@@
   filter(caste != "male") %>%
-  filter(!bb.sp %in% c("psit", "gers", "humi")) %>% # only focal species
-  #group_by(elev.class2) %>%
-  group_by(site) %>%
+  filter(!bb.sp %in% c("gers", "humi")) %>% 
+  #group_by(site) %>%
+  group_by(elev.class2) %>%
   mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) %>%
-  # mutate(k.type.ss = factor(k.type.ss),
-  #        color = factor(color),
-  #        bb.sp = factor(bb.sp, levels = c("hort", "pasc", "wurf", "bss", "prat", "soro"))) %>%
-  mutate(k.type.ss = factor(k.type.ss),
-         color = factor(color),
-         bb.sp = factor(bb.sp)) %>%
-  group_by(period, elev.class2, plant.sp.abb, plant.genus, k.type.ss, color, plant.family, bb.sp) %>%
-  summarize(abund = n()) %>%
-  group_by(elev.class2, period, bb.sp) %>%
-  mutate(prop = abund/sum(abund)) %>% 
+  group_by(site, date, bb.sp, period, elev.class2, plant.sp.abb, plant.genus, k.type.ss, color, plant.family) %>% 
+  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
+  group_by(site, date, period, elev.class2, plant.sp.abb, plant.genus, k.type.ss, color, plant.family) %>%
+  complete(bb.sp, plant.sp.abb, fill = list(freq = 0)) %>% # fill in zeros for unvisited plants
+  group_by(site, date, bb.sp) %>%
+  mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair per site-date 
+  group_by(bb.sp, period, elev.class2, plant.sp.abb, plant.genus, k.type.ss, color, plant.family) %>%
+  summarize(prop = mean(prop)) %>% # get mean proportional intneraction frequency aggregated by ...
   left_join(bb_traits) %>%
@@
 # Mid
-tp.mid1 <- ggplot(filter(visitation_abund.sp, period == "mid1"), 
+tp.mid <- ggplot(filter(visitation_abund.sp, period == "mid"), 
                  aes(area = prop, fill = k.type.ss, label = plant.sp.abb)) +
@@
 
-tp.mid1.col <- ggplot(filter(visitation_abund.sp, period == "mid1"), 
-                 aes(area = prop, fill = color, label = plant.sp.abb)) +
-  geom_treemap(color = "white") +
-  geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
-  scale_fill_manual(values = cols2) + 
-  labs(fill = "Morphotype") +
-  facet_grid(fct_rev(elev.class2)~reorder(bb.sp, pbl.w)) +
-  theme(strip.text = element_text(size = 14))
-
-tp.mid2 <- ggplot(filter(visitation_abund.sp, period == "mid2"), 
-                 aes(area = prop, fill = k.type.ss, label = plant.sp.abb)) +
-  geom_treemap(color = "white") +
-  geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
-  scale_fill_manual(values = cols) + 
-  labs(fill = "Morphotype") +
-  facet_grid(fct_rev(elev.class2)~reorder(bb.sp, pbl.w)) +
-  theme(strip.text = element_text(size = 14))
-
-tp.mid2.col <- ggplot(filter(visitation_abund.sp, period == "mid2"), 
+tp.mid.col <- ggplot(filter(visitation_abund.sp, period == "mid"), 
                  aes(area = prop, fill = color, label = plant.sp.abb)) +
@@
 ```
-
-# 
-```{r}
-vis_sum <- net %>%
-  group_by(bb.sp, plant.sp, k.type.ss, color) %>%
-  summarize()
-
-```
-
-## 1.7 How do abundant generalist bumble bees deal with a dynamic floral community?
-Based on the ordination analysis, it looks like they maintain fairly steady morphological preferences, but they are taxonomically flexible.
-Let's take a closer look at that.
-
-A question to consider is whether the shallower slope of morphological turnover is caused by the "reluctance" of BBs to switch morphotypes or lag in the floristic turnover of morphotypes relative to species.
-Morphotypes might be less "sensitive" than species to elevational change, so you first get species turnover within morphotypes, then with enough elevational chnage you start to see the morphotypes turning over.
-**I can freaking test this with my floral survey data!**
-
-### Aggregate analysis
-```{r}
-## Binary
-### Taxonomic
-#### BB species
-beta.tax.sp.bin <- beta_func(net, traits = bb_traits, sitedata = site_data, level = bb.sp, 
-                             criterion = plant.sp, binary = TRUE) %>%
-  mutate(criterion = rep("tax", n()))
-
-#### Tongnue length guild
-beta.tax.guild.bin <- beta_func(net, traits = bb_traits, sitedata = site_data, level = pbl.fac,
-                                criterion = plant.sp, binary = TRUE) %>%
-  mutate(criterion = rep("tax", n()))
-
-#### Plants
-betaFL_tax.bin <- beta_func_fl(survey, sitedata = site_data, criterion = plant.sp, binary = TRUE) %>%
-  mutate(criterion = rep("tax", n()))
-
-### Morphological
-#### BB species
-beta.morph.sp.bin <- beta_func(net, traits = bb_traits, sitedata = site_data, level = bb.sp,
-                               criterion = k.type.s, binary = TRUE) %>%
-  mutate(criterion = rep("morph", n()))
-
-#### Tongue length guild
-beta.morph.guild.bin <- beta_func(net, traits = bb_traits, sitedata = site_data, level = pbl.fac,
-                                  criterion = k.type.s, binary = TRUE) %>%
-  mutate(criterion = rep("morph", n()))
-
-#### Plants
-betaFL_morph.bin <- beta_func_fl(survey, sitedata = site_data, criterion = k.type.s, binary = TRUE) %>%
-  mutate(criterion = rep("morph", n()))
-
-#### Collated
-beta.guild.bin <- bind_rows(beta.tax.guild.bin, beta.morph.guild.bin)
-beta.sp.bin <- bind_rows(beta.tax.sp.bin, beta.morph.sp.bin)
-beta.fl.bin <- bind_rows(betaFL_tax.bin, betaFL_morph.bin)
-
-## Not binary (i.e. abundance-weighted)
-### Taxonomic
-#### BB species
-beta.tax.sp.wt <- beta_func(net,  traits = bb_traits, sitedata = site_data, level = bb.sp,
-                         criterion = plant.sp, binary = FALSE) %>%
-  mutate(criterion = rep("tax", n()))
-
-beta.tax.guild.wt <- beta_func(net,  traits = bb_traits, sitedata = site_data, level = pbl.fac,
-                            criterion = plant.sp, binary = FALSE) %>%
-  mutate(criterion = rep("tax", n()))
-
-beta.morph.sp.wt <- beta_func(net,  traits = bb_traits, sitedata = site_data, level = bb.sp,
-                           criterion = k.type.s, binary = FALSE) %>%
-  mutate(criterion = rep("morph", n()))
-
-beta.morph.guild.wt <- beta_func(net,  traits = bb_traits, sitedata = site_data, level = pbl.fac, 
-                              criterion = k.type.s, binary = FALSE) %>%
-  mutate(criterion = rep("morph", n()))
-
-beta.sp.wt <- bind_rows(beta.tax.sp.wt, beta.morph.sp.wt)
-beta.guild.wt <- bind_rows(beta.tax.guild.wt, beta.morph.guild.wt)
-
-#### Plants
-betaFL_tax.wt <- beta_func_fl(survey, sitedata = site_data, criterion = plant.sp, binary = FALSE) %>%
-  mutate(criterion = rep("tax", n()))
-betaFL_morph.wt <- beta_func_fl(survey, sitedata = site_data, criterion = k.type.s, binary = FALSE) %>%
-  mutate(criterion = rep("morph", n()))
-beta.fl.wt <- bind_rows(betaFL_tax.wt, betaFL_morph.wt)
-
-# Pull out ADONIS results
-## Binary
-beta.tax.sp.adonis.bin <- beta.tax.sp.bin %>%
-  select(bb.sp, adonis.r2, adonis.p) %>%
-  distinct()
-
-beta.morph.sp.adonis.bin <- beta.morph.sp.bin %>%
-  select(bb.sp, adonis.r2, adonis.p) %>%
-  distinct()
-
-beta.tax.guild.adonis.bin <- beta.tax.guild.bin %>%
-  select(pbl.fac, adonis.r2, adonis.p) %>%
-  distinct()
-
-beta.morph.guild.adonis.bin <- beta.morph.guild.bin %>%
-  select(pbl.fac, adonis.r2, adonis.p) %>%
-  distinct()
-
-## Not binary
-beta.tax.sp.adonis <- beta.tax.sp.wt %>%
-  select(bb.sp, adonis.r2, adonis.p) %>%
-  distinct()
-
-beta.morph.sp.adonis <- beta.morph.sp.wt %>%
-  select(bb.sp, adonis.r2, adonis.p) %>%
-  distinct()
-
-beta.tax.guild.adonis <- beta.tax.guild.wt %>%
-  select(pbl.fac, adonis.r2, adonis.p) %>%
-  distinct()
-
-beta.morph.guild.adonis <- beta.morph.guild.wt %>%
-  select(pbl.fac, adonis.r2, adonis.p) %>%
-  distinct()
-
-# Plot
-## Binary
-beta.sp.plot.bin <- ggplot(beta.sp.bin, aes(elev.diff, distance, color = criterion)) +
-  geom_point(alpha = 0.05) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
-  facet_wrap(~bb.sp) + 
-  ggtitle("Species")
-
-beta.guild.plot.bin <- ggplot(beta.guild.bin, aes(elev.diff, distance, color = criterion)) +
-  geom_point(alpha = 0.05) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
-  facet_wrap(~pbl.fac) + 
-  ggtitle("Tongue guild")
-
-(beta.sp.plot.bin | beta.guild.plot.bin) + plot_layout(guides = 'collect')
-
-## Not binary
-beta.sp.plot.wt <- ggplot(beta.sp.wt, aes(elev.diff, distance, color = criterion)) +
-  geom_point(alpha = 0.05) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
-  facet_wrap(~bb.sp) + 
-  ggtitle("Species")
-
-beta.guild.plot.wt <- ggplot(beta.guild.wt, aes(elev.diff, distance, color = criterion)) +
-  geom_point(alpha = 0.125) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
-  facet_wrap(~pbl.fac) + 
-  ggtitle("Tongue guild")
-
-(beta.sp.plot.wt | beta.guild.plot.wt) + plot_layout(guides = 'collect')
-
-# Plants
-ggplot(beta.fl.bin, aes(elev.diff, distance, color = criterion)) +
-  geom_point(alpha = 0.125) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE)
-
-ggplot(beta.fl.wt, aes(elev.diff, distance, color = criterion)) +
-  geom_point(alpha = 0.125) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE)
-
-# Plotting visitation beta-diversity with floral survey beta-diversity
-
-beta.sp.plot.bin2 <- ggplot(beta.sp.bin, aes(elev.diff, distance, color = criterion)) +
-  #geom_point(alpha = 0.05) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = TRUE) +
-  geom_smooth(inherit.aes = FALSE, data = beta.fl.wt, 
-              aes(elev.diff, distance, color = criterion), 
-              linetype = "dashed", se = TRUE,
-              method = "gam", formula = y ~ s(x, k = 4)) +
-  facet_wrap(~bb.sp) + 
-  ggtitle("Species") +
-  theme_light()
-
-beta.guild.plot.bin2 <- ggplot(beta.guild.bin, aes(elev.diff, distance, color = criterion)) +
-  #geom_point(alpha = 0.125) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = TRUE) +
-  geom_smooth(inherit.aes = FALSE, data = beta.fl.wt, 
-              aes(elev.diff, distance, color = criterion), 
-              linetype = "dashed", se = TRUE,
-              method = "gam", formula = y ~ s(x, k = 4)) +
-  facet_wrap(~pbl.fac) + 
-  ggtitle("Tongue guild") +
-  theme_light()
-
-(beta.sp.plot.bin2 | beta.guild.plot.bin2) + plot_layout(guides = 'collect')
-
-beta.sp.plot.wt2 <- ggplot(beta.sp.wt, aes(elev.diff, distance, color = criterion)) +
-  #geom_point(alpha = 0.05) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = TRUE) +
-  geom_smooth(inherit.aes = FALSE, data = beta.fl.wt, 
-              aes(elev.diff, distance, color = criterion), 
-              linetype = "dashed", se = TRUE,
-              method = "gam", formula = y ~ s(x, k = 4)) +
-  facet_wrap(~bb.sp) + 
-  ggtitle("Species") +
-  theme_light()
-
-beta.guild.plot.wt2 <- ggplot(beta.guild.wt, aes(elev.diff, distance, color = criterion)) +
-  #geom_point(alpha = 0.125) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = TRUE) +
-  geom_smooth(inherit.aes = FALSE, data = beta.fl.wt, 
-              aes(elev.diff, distance, color = criterion), 
-              linetype = "dashed", se = TRUE,
-              method = "gam", formula = y ~ s(x, k = 4)) +
-  facet_wrap(~pbl.fac) + 
-  ggtitle("Tongue guild") +
-  theme_light()
-
-(beta.sp.plot.wt2 | beta.guild.plot.wt2) + plot_layout(guides = 'collect')
-
-# Plotting visitation beta-diversity against floral survey beta-diversity
-beta.sp.bin2 <- beta.sp.bin %>%
-  full_join(beta.fl.bin, by = c("site1", "site2", "elev1", "elev2", "elev.diff")) %>%
-  rename(distance.vis = distance.x, distance.surv = distance.y, 
-         criterion.vis = criterion.x, criterion.surv = criterion.y)
-
-beta.guild.bin2 <- beta.guild.bin %>%
-  full_join(beta.fl.bin, by = c("site1", "site2", "elev1", "elev2", "elev.diff")) %>%
-  rename(distance.vis = distance.x, distance.surv = distance.y, 
-         criterion.vis = criterion.x, criterion.surv = criterion.y)
-
-beta.sp.wt2 <- beta.sp.wt %>%
-  full_join(beta.fl.wt, by = c("site1", "site2", "elev1", "elev2", "elev.diff")) %>%
-  rename(distance.vis = distance.x, distance.surv = distance.y, 
-         criterion.vis = criterion.x, criterion.surv = criterion.y)
-
-beta.guild.wt2 <- beta.guild.wt %>%
-  full_join(beta.fl.wt, by = c("site1", "site2", "elev1", "elev2", "elev.diff")) %>%
-  rename(distance.vis = distance.x, distance.surv = distance.y, 
-         criterion.vis = criterion.x, criterion.surv = criterion.y)
-
-ggplot(beta.sp.bin2, aes(distance.surv, distance.vis, color = criterion.vis)) +
-  geom_point(alpha = 0.25) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
-  facet_wrap(~bb.sp)
-
-ggplot(beta.sp.wt2, aes(distance.surv, distance.vis, color = criterion.vis)) +
-  geom_point(alpha = 0.25) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
-  facet_wrap(~bb.sp)
-
-ggplot(beta.guild.bin2, aes(distance.surv, distance.vis, color = criterion.vis)) +
-  geom_point(alpha = 0.25) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
-  facet_wrap(~pbl.fac)
-
-ggplot(beta.guild.wt2, aes(distance.surv, distance.vis, color = criterion.vis)) +
-  geom_point(alpha = 0.25) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
-  facet_wrap(~bb.guild)
-
-```
-
-### Parsed by period
-```{r}
-# Binary
-## Taxonomic
-### BB species
-beta_x_period.tax.sp.bin <- beta_func(net, traits = bb_traits, sitedata = site_data, level = bb.sp, criterion = plant.sp, 
-                                    binary = TRUE, period) %>%
-  mutate(criterion = rep("tax", n()))
-
-### Tongue length guild
-beta_x_period.tax.guild.bin <- beta_func(net, traits = bb_traits, sitedata = site_data, level = pbl.fac, criterion = plant.sp, 
-                                       binary = TRUE, period) %>%
-  mutate(criterion = rep("tax", n()))
-
-## Morphological
-### BB species
-beta_x_period.morph.sp.bin <- beta_func(net, traits = bb_traits, sitedata = site_data, level = bb.sp, criterion = k.type.s, 
-                                    binary = TRUE, period) %>%
-  mutate(criterion = rep("morph", n()))
-
-### Tongue length guild
-beta_x_period.morph.guild.bin <- beta_func(net, traits = bb_traits, sitedata = site_data, level = pbl.fac, criterion = k.type.s, 
-                                       binary = TRUE, period) %>%
-  mutate(criterion = rep("morph", n()))
-
-### Collated
-beta_x_period.sp.bin <- bind_rows(beta_x_period.tax.sp.bin, beta_x_period.morph.sp.bin)
-beta_x_period.guild.bin <- bind_rows(beta_x_period.tax.guild.bin, beta_x_period.morph.guild.bin)
-
-# Not binary
-## Taxonomic
-### BB species
-beta_x_period.tax.sp.wt <- beta_func(net, traits = bb_traits, sitedata = site_data, level = bb.sp, criterion = plant.sp, 
-                                    binary = FALSE, period) %>%
-  mutate(criterion = rep("tax", n()))
-
-### Tongue length guild
-beta_x_period.tax.guild.wt <- beta_func(net, traits = bb_traits, sitedata = site_data, level = pbl.fac, criterion = plant.sp, 
-                                       binary = FALSE, period) %>%
-  mutate(criterion = rep("tax", n()))
-
-## Morphological
-### BB species
-beta_x_period.morph.sp.wt <- beta_func(net, traits = bb_traits, sitedata = site_data, level = bb.sp, criterion = k.type.s, 
-                                    binary = FALSE, period) %>%
-  mutate(criterion = rep("morph", n()))
-
-### Tongue length guild
-beta_x_period.morph.guild.wt <- beta_func(net, traits = bb_traits, sitedata = site_data, level = pbl.fac, criterion = k.type.s, 
-                                       binary = FALSE, period) %>%
-  mutate(criterion = rep("morph", n()))
-
-### Collated
-beta_x_period.sp.wt <- bind_rows(beta_x_period.tax.sp.wt, beta_x_period.morph.sp.wt)
-beta_x_period.guild.wt <- bind_rows(beta_x_period.tax.guild.wt, beta_x_period.morph.guild.wt)
-
-# Pull out ADONIS results
-## Binary
-beta_x_period.tax.sp.adonis.bin <- beta_x_period.tax.sp.bin %>%
-  select(bb.sp, period, adonis.r2, adonis.p) %>%
-  distinct()
-
-beta_x_period.morph.sp.adonis.bin <- beta_x_period.morph.sp.bin %>%
-  select(bb.sp, period, adonis.r2, adonis.p) %>%
-  distinct()
-
-beta_x_period.tax.guild.adonis.bin <- beta_x_period.tax.guild.bin %>%
-  select(pbl.fac, period, adonis.r2, adonis.p) %>%
-  distinct()
-
-beta_x_period.morph.guild.adonis.bin <- beta_x_period.morph.guild.bin %>%
-  select(pbl.fac, period, adonis.r2, adonis.p) %>%
-  distinct()
-
-## Not binary
-beta_x_period.tax.sp.adonis <- beta_x_period.tax.sp.wt %>%
-  select(bb.sp, adonis.r2, adonis.p) %>%
-  distinct()
-
-beta_x_period.morph.sp.adonis <- beta_x_period.morph.sp.wt %>%
-  select(bb.sp, adonis.r2, adonis.p) %>%
-  distinct()
-
-beta_x_period.tax.guild.adonis <- beta_x_period.tax.guild.wt %>%
-  select(pbl.fac, adonis.r2, adonis.p) %>%
-  distinct()
-
-beta_x_period.morph.guild.adonis <- beta_x_period.morph.guild.wt %>%
-  select(pbl.fac, adonis.r2, adonis.p) %>%
-  distinct()
-
-# Plot
-## Binary
-beta_x_period.sp.plot.bin <- ggplot(beta_x_period.sp.bin, aes(elev.diff, distance, color = criterion)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
-  facet_grid(bb.sp ~ period) + 
-  ggtitle("Species")
-
-beta_x_period.guild.plot.bin <- ggplot(beta_x_period.guild.bin, aes(elev.diff, distance, color = criterion)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
-  facet_grid(pbl.fac ~ period) + 
-  ggtitle("Tongue guild")
-
-(beta_x_period.sp.plot.bin | beta_x_period.guild.plot.bin) + plot_layout(guides = 'collect')
-
-## Not binary
-beta_x_period.sp.plot.wt <- ggplot(beta_x_period.sp.wt, aes(elev.diff, distance, color = criterion)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
-  facet_grid(bb.sp ~ period) + 
-  ggtitle("Species")
-
-beta_x_period.guild.plot.wt <- ggplot(beta_x_period.guild.wt, aes(elev.diff, distance, color = criterion)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
-  facet_grid(pbl.fac ~ period) + 
-  ggtitle("Tongue guild")
-
-(beta_x_period.sp.plot.wt | beta_x_period.guild.plot.wt) + plot_layout(guides = 'collect')
-
-########## A different way to look at this
-ggplot(beta_x_period.guild.wt, aes(pbl.fac, distance, color = criterion)) +
-  geom_boxplot() +
-  facet_wrap( ~ period)
-
-ggplot(beta_x_period.sp.wt, aes(bb.sp, distance, color = criterion)) +
-  geom_boxplot() +
-  facet_wrap( ~ period) 
-```
-
-## 1.8 Does the influence of tongue length on niche partitioning vary with elevation or time?
-```{r, echo = FALSE}
-# By plant.sp
-ggplot(niche_x_elevperiod.tax, aes(elev.class2, adonis.r2)) +
-  geom_boxplot()
-
-ggplot(niche_x_elevperiod.tax, aes(period, adonis.r2)) +
-  geom_boxplot()
-
-# By ktype
-ggplot(niche_x_elevperiod.morph, aes(elev.class2, adonis.r2)) +
-  geom_boxplot()
-
-ggplot(niche_x_elevperiod.morph, aes(period, adonis.r2)) +
-  geom_boxplot()
-```
-
-# 2. Testing the stress-gradient hypothesis and the Ranta and Vepsäläinen heterogeneity hypothesis
-
-**Stress-gradient hypothesis:** Is there any evidence that niche partitioning amongst generalist core bumble bee species is governed by an inverse relationship between elevation and competition, broadly consistent with the stress-gradient hypothesis (Bertness and Callaway 1994).
-This mechanism has been inferred from trait patterns in the whole bee community of the Berchtesgadener Alps (Hoiss et al.
-2012) and from interaction specialization patterns in the plant-pollinator community of Mt.
-Kilimanjaro (Classen et al.
-2020).
-It predicts that niche overlap will increase with elevation, reflecting competitive pressures at low elevations and reduced competition but increased need for opportunism at high elevations.
-
-**Ranta/Vepsäläinen heterogeneity hypothesis :** Ranta and Vepsäläinen (1981) proposed that bumble bees communities are governed by a spatiotemporally heterogeneous mosaic of competitive conditions, such that competitive inequalities between species are sufficiently variable to allow more species to coexist than predicted by trait-based competition models.
-
-## 2.1: Testing the stress-gradient hypothesis
-### 2.1.1 Face-value analysis
-
-*Results:*
-
-```{r message=FALSE, warning=FALSE, echo = FALSE}
-# Pairwise (bb.sp:bb.sp) niche overlap by site, pooled across years
-
-niche_overlap.tax <- niche_overlapper(net, criterion = plant.sp.abb)
-niche_overlap.morph <- niche_overlapper(net, criterion = k.type.s)
-
-# All spp pairs
-sg1 <- ggplot(niche_overlap.tax, aes(elev.mean, niche.overlap)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
-
-sg2 <- ggplot(niche_overlap.morph, aes(elev.mean, niche.overlap)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
-
-(sg1 + ggtitle("Taxonomic") | sg2 + ggtitle("Morphological")) + plot_layout(guides = 'collect')
-
-## Within LT spp
-sg3 <- ggplot(filter(niche_overlap.tax, pbl.fac1 == "long" & pbl.fac2 == "long"), 
-       aes(elev.mean, niche.overlap)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
-
-sg4 <- ggplot(filter(niche_overlap.morph, pbl.fac1 == "long" & pbl.fac2 == "long"), 
-       aes(elev.mean, niche.overlap)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
-
-(sg3 + ggtitle("Taxonomic") | sg4 + ggtitle("Morphological")) + plot_layout(guides = 'collect')
-
-## Within ST spp
-sg5 <- ggplot(filter(niche_overlap.tax, pbl.fac1 == "short" & pbl.fac2 == "short"), 
-       aes(elev.mean, niche.overlap)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
-
-sg6 <- ggplot(filter(niche_overlap.morph, pbl.fac1 == "short" & pbl.fac2 == "short"), 
-       aes(elev.mean, niche.overlap)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
-
-(sg5 + ggtitle("Taxonomic") | sg6 + ggtitle("Morphological")) + plot_layout(guides = 'collect')
-
-## Between ST and LT
-sg7 <- ggplot(filter(niche_overlap.tax, pbl.fac.diff == "different"), 
-       aes(elev.mean, niche.overlap)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
-
-sg8 <- ggplot(filter(niche_overlap.morph, pbl.fac.diff == "different"), 
-       aes(elev.mean, niche.overlap)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 8))
-
-(sg7 + ggtitle("Taxonomic") | sg8 + ggtitle("Morphological")) + plot_layout(guides = 'collect')
-
-# GAMs
-glob_GAM.tax <- gam(niche.overlap ~ 
-                      s(year, bs  = "re") + # random intercept effect for year
-                      s(elev.mean, k = 8, m = 2) + # global smooth ~ elevation
-                      s(elev.mean, sp.pair, bs = "fs", k = 8, m = 1), # group-level smooth ~ elevation*sp.pair
-                    method = "REML",
-                    family = "gaussian",
-                    data = niche_overlap.tax) %>% getViz()
-
-check.gamViz(glob_GAM.tax)
-summary(glob_GAM.tax)
-print(plot(glob_GAM.tax, allTerms = TRUE), pages = 1)
-
-glob_GAM.morph <- gam(niche.overlap ~ 
-                      s(year, bs  = "re") + # random intercept effect for year
-                      s(elev.mean, k = 8, m = 2) + # global smooth ~ elevation
-                      s(elev.mean, sp.pair, bs = "fs", k = 8, m = 1), # group-level smooth ~ elevation*sp.pair
-                    method = "REML",
-                    family = "gaussian",
-                    data = niche_overlap.morph) %>% getViz()
-
-check.gamViz(glob_GAM.morph)
-summary(glob_GAM.morph)
-print(plot(glob_GAM.morph, allTerms = TRUE), pages = 1)
-
-# Within LT class
-LT_GAM.tax <- gam(niche.overlap ~ 
-                      s(year, bs  = "re") + # random intercept effect for year
-                      s(elev.mean, k = 8, m = 2) + # global smooth ~ elevation
-                      s(elev.mean, sp.pair, bs = "fs", k = 8, m = 1), # group-level smooth ~ elevation*sp.pair
-                  method = "REML",
-                  family = "gaussian",
-                  data = filter(niche_overlap.tax, pbl.fac1 == "long" & pbl.fac2 == "long")) %>% getViz()
-
-check.gamViz(LT_GAM.tax)
-summary(LT_GAM.tax)
-print(plot(LT_GAM.tax, allTerms = TRUE), pages = 1)
-
-LT_GAM.morph <- gam(niche.overlap ~ 
-                      s(year, bs  = "re") + # random intercept effect for year
-                      s(elev.mean, k = 8, m = 2) + # global smooth ~ elevation
-                      s(elev.mean, sp.pair, bs = "fs", k = 8, m = 1), # group-level smooth ~ elevation*sp.pair
-                  method = "REML",
-                  family = "gaussian",
-                  data = filter(niche_overlap.morph, pbl.fac1 == "long" & pbl.fac2 == "long")) %>% getViz()
-
-check.gamViz(LT_GAM.morph)
-summary(LT_GAM.morph)
-print(plot(LT_GAM.morph, allTerms = TRUE), pages = 1)
-
-# Within ST class
-ST_GAM.tax <- gam(niche.overlap ~ 
-                      s(year, bs  = "re") + # random intercept effect for year
-                      s(elev.mean, k = 8, m = 2) + # global smooth ~ elevation
-                      s(elev.mean, sp.pair, bs = "fs", k = 8, m = 1), # group-level smooth ~ elevation*sp.pair
-                  method = "REML",
-                  family = "gaussian",
-                  data = filter(niche_overlap.tax, pbl.fac1 == "short" & pbl.fac2 == "short")) %>% getViz()
-
-check.gamViz(ST_GAM.tax)
-summary(ST_GAM.tax)
-print(plot(ST_GAM.tax, allTerms = TRUE), pages = 1)
-
-ST_GAM.morph <- gam(niche.overlap ~ 
-                      s(year, bs  = "re") + # random intercept effect for year
-                      s(elev.mean, k = 8, m = 2) + # global smooth ~ elevation
-                      s(elev.mean, sp.pair, bs = "fs", k = 8, m = 1), # group-level smooth ~ elevation*sp.pair
-                  method = "REML",
-                  family = "gaussian",
-                  data = filter(niche_overlap.morph, pbl.fac1 == "short" & pbl.fac2 == "short")) %>% getViz()
-
-check.gamViz(ST_GAM.morph)
-summary(ST_GAM.morph)
-print(plot(ST_GAM.morph, allTerms = TRUE), pages = 1)
-
-# Between ST and LT
-STxLT_GAM.tax <- gam(niche.overlap ~ 
-                      s(year, bs  = "re") + # random intercept effect for year
-                      s(elev.mean, k = 8, m = 2) + # global smooth ~ elevation
-                      s(elev.mean, sp.pair, bs = "fs", k = 8, m = 1), # group-level smooth ~ elevation*sp.pair
-                  method = "REML",
-                  family = "gaussian",
-                  data = filter(niche_overlap.tax, pbl.fac.diff == "different")) %>% getViz()
-
-check.gamViz(STxLT_GAM.tax)
-summary(STxLT_GAM.tax)
-print(plot(STxLT_GAM.tax, allTerms = TRUE), pages = 1)
-
-STxLT_GAM.morph <- gam(niche.overlap ~ 
-                      s(year, bs  = "re") + # random intercept effect for year
-                      s(elev.mean, k = 8, m = 2) + # global smooth ~ elevation
-                      s(elev.mean, sp.pair, bs = "fs", k = 8, m = 1), # group-level smooth ~ elevation*sp.pair
-                  method = "REML",
-                  family = "gaussian",
-                  data = filter(niche_overlap.morph, pbl.fac.diff == "different")) %>% getViz()
-
-check.gamViz(STxLT_GAM.morph)
-summary(STxLT_GAM.morph)
-print(plot(STxLT_GAM.morph, allTerms = TRUE), pages = 1)
-```
-
-### 2.1.2 Null-model residual analysis
-
-Arguably the most rigorous way to tease out the effect of elevation while controlling for community composition (abundance and richness patterns) is to compare the observed niche overlap at each site to that predicted by a random null model, then see how the residuals vary with elevation.
-
-I'll be the first to admit that I'm new to null models, so I want to be extremely clear about the logic of how I'm setting this up. Step-by-step:  
-1. Generate community matrices by site and pollen/nectar
-2.
-Permute each matrix *n* times by randomly redistributing the visits of each bumble bee species while holding the total visits for each bumble bee species constant.
-This is, therefore, a random selection model constrained to the set of flora visited for a given site-pollen/nectar matrix.
-Importantly, I am opting not to hold the floral marginal totals constant, i.e.
-the total number of visits that each plant receives will vary with each permutation.
-My gut feeling is that this is the right place to start, but this is a decision that requires scrutiny.
-3.
-Calculate niche overlap for each permuted matrix and store as tidy arrays; this mirrors the workflow for calculating empirical niche overlap, the only difference being that I had to deal with some tricky nested lists; I'm pretty sure I did this correctly.
-4. Take the mean niche overlap for each permuted community.
-5.
-Join null model mean niche overlap table to empirical niche overlap table, joining by site and pollen/nectar.
-Thus, for each site-pollen/nectar combo in the empirical niche overlap table, we have *n* iterations of random niche overlap.
-6. Calculate the difference between empirical niche overlap and null niche overlap for each iteration; we'll call this "residual niche overlap".
-7.
-Calculate mean residual niche overlap for each site-pollen/nectar combo.
-This can then be regressed against elevation to see how the deviation from null expectations varies along the elevation gradient, with the interpretation that this represents the "pure" influence of elevation on niche overlap whle controlling for community differences.
-We could also use the raw differences without collapsing them to a mean.
-But that's not what I did this first time through.
-
-There are surely a lot of other ways a null model could be constructed and analyzed, but this seems to me a good place to start.
-
-*Results*
-
-```{r message=FALSE, warning=TRUE, echo = FALSE}
-# Holding BB marginal totals constant
-random_niche.tax <- nullify1(net, 50, plant.sp)
-random_niche.morph <- nullify1(net, 50, k.type.s)
-
-# Holding both BB and plant marginal totals constant
-# random_niche.tax <- nullify2(net, 50, plant.sp)
-# random_niche.morph <- nullify2(net, 50, k.type.s)
-
-## Join null model to empirical niche overlap
-# Taxonomic niche overlap
-comparison.tax <- niche_overlap.tax %>%
-  left_join(random_niche.tax, by = c("site", "sp.pair")) %>%
-  mutate(niche.overlap.resid = niche.overlap - niche.overlap.sim) %>%
-  group_by(site, elev.mean, sp.pair, pbl.fac1, pbl.fac2, pbl.fac.diff) %>%
-  summarize(niche.overlap.resid = mean(niche.overlap.resid))
-
-# Morphological niche overlap
-comparison.morph <- niche_overlap.morph %>%
-  left_join(random_niche.morph, by = c("site", "sp.pair")) %>%
-  mutate(niche.overlap.resid = niche.overlap - niche.overlap.sim) %>%
-  group_by(site, elev.mean, sp.pair, pbl.fac1, pbl.fac2, pbl.fac.diff) %>%
-  summarize(niche.overlap.resid = mean(niche.overlap.resid))
-
-# Plot
-# All spp pairs
-sgn1 <- ggplot(comparison.tax, aes(elev.mean, niche.overlap.resid)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
-
-sgn2 <- ggplot(comparison.morph, aes(elev.mean, niche.overlap.resid)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
-
-(sgn1 + ggtitle("Taxonomic") | sgn2 + ggtitle("Morphological")) + plot_layout(guides = 'collect')
-
-## Within LT spp
-sgn3 <- ggplot(filter(comparison.tax, pbl.fac1 == "long" & pbl.fac2 == "long"), 
-       aes(elev.mean, niche.overlap.resid)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
-
-sgn4 <- ggplot(filter(comparison.morph, pbl.fac1 == "long" & pbl.fac2 == "long"), 
-       aes(elev.mean, niche.overlap.resid)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
-
-(sgn3 + ggtitle("Taxonomic") | sgn4 + ggtitle("Morphological")) + plot_layout(guides = 'collect')
-
-## Within ST spp
-sgn5 <- ggplot(filter(comparison.tax, pbl.fac1 == "short" & pbl.fac2 == "short"), 
-       aes(elev.mean, niche.overlap.resid)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
-
-sgn6 <- ggplot(filter(comparison.morph, pbl.fac1 == "short" & pbl.fac2 == "short"), 
-       aes(elev.mean, niche.overlap.resid)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
-
-(sgn5 + ggtitle("Taxonomic") | sgn6 + ggtitle("Morphological")) + plot_layout(guides = 'collect')
-
-## Between ST and LT
-sgn7 <- ggplot(filter(comparison.tax, pbl.fac.diff == "different"), 
-       aes(elev.mean, niche.overlap.resid)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))
-
-sgn8 <- ggplot(filter(comparison.morph, pbl.fac.diff == "different"), 
-       aes(elev.mean, niche.overlap.resid)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 8))
-
-(sgn7 + ggtitle("Taxonomic") | sgn8 + ggtitle("Morphological")) + plot_layout(guides = 'collect')
-
-# GAMs
-glob_nullmodel_GAM.tax <- gam(niche.overlap.resid ~ 
-                      s(elev.mean, k = 10, m = 2),
-                    method = "REML",
-                    family = "gaussian",
-                    data = comparison.tax) %>% getViz()
-
-check.gamViz(glob_nullmodel_GAM.tax)
-summary(glob_nullmodel_GAM.tax)
-print(plot(glob_nullmodel_GAM.tax, allTerms = TRUE), pages = 1)
-
-glob_nullmodel_GAM.morph <- gam(niche.overlap.resid ~ 
-                      s(elev.mean, k = 10, m = 2),
-                    method = "REML",
-                    family = "gaussian",
-                    data = comparison.morph) %>% getViz()
-
-check.gamViz(glob_nullmodel_GAM.morph)
-summary(glob_nullmodel_GAM.morph)
-print(plot(glob_nullmodel_GAM.morph, allTerms = TRUE), pages = 1)
-
-# Within LT class
-LT_nullmodel_GAM.tax <- gam(niche.overlap.resid ~ 
-                    s(elev.mean, k = 10, m = 2),
-                  method = "REML",
-                  family = "gaussian",
-                  data = filter(comparison.tax, pbl.fac1 == "long" & pbl.fac2 == "long")) %>% getViz()
-
-check.gamViz(LT_nullmodel_GAM.tax)
-summary(LT_nullmodel_GAM.tax)
-print(plot(LT_nullmodel_GAM.tax, allTerms = TRUE), pages = 1)
-
-LT_nullmodel_GAM.morph <- gam(niche.overlap.resid ~ 
-                    s(elev.mean, k = 10, m = 2),
-                  method = "REML",
-                  family = "gaussian",
-                  data = filter(comparison.morph, pbl.fac1 == "long" & pbl.fac2 == "long")) %>% getViz()
-
-check.gamViz(LT_nullmodel_GAM.morph)
-summary(LT_nullmodel_GAM.morph)
-print(plot(LT_nullmodel_GAM.morph, allTerms = TRUE), pages = 1)
-
-# Within ST class
-ST_nullmodel_GAM.tax <- gam(niche.overlap.resid ~ 
-                    s(elev.mean, k = 10, m = 2),
-                  method = "REML",
-                  family = "gaussian",
-                  data = filter(comparison.tax, pbl.fac1 == "short" & pbl.fac2 == "short")) %>% getViz()
-
-check.gamViz(ST_nullmodel_GAM.tax)
-summary(ST_nullmodel_GAM.tax)
-print(plot(ST_nullmodel_GAM.tax, allTerms = TRUE), pages = 1)
-
-ST_nullmodel_GAM.morph <- gam(niche.overlap.resid ~ 
-                    s(elev.mean, k = 10, m = 2),
-                  method = "REML",
-                  family = "gaussian",
-                  data = filter(comparison.morph, pbl.fac1 == "short" & pbl.fac2 == "short")) %>% getViz()
-
-check.gamViz(ST_nullmodel_GAM.morph)
-summary(ST_nullmodel_GAM.morph)
-print(plot(ST_nullmodel_GAM.morph, allTerms = TRUE), pages = 1)
-
-# Between ST and LT
-STxLT_nullmodel_GAM.tax <- gam(niche.overlap.resid ~ 
-                    s(elev.mean, k = 10, m = 2),
-                  method = "REML",
-                  family = "gaussian",
-                  data = filter(comparison.tax, pbl.fac.diff == "different")) %>% getViz()
-
-check.gamViz(STxLT_nullmodel_GAM.tax)
-summary(STxLT_nullmodel_GAM.tax)
-print(plot(STxLT_nullmodel_GAM.tax, allTerms = TRUE), pages = 1)
-
-STxLT_nullmodel_GAM.morph <- gam(niche.overlap.resid ~ 
-                    s(elev.mean, k = 10, m = 2),
-                  method = "REML",
-                  family = "gaussian",
-                  data = filter(comparison.morph, pbl.fac.diff == "different")) %>% getViz()
-
-check.gamViz(STxLT_nullmodel_GAM.morph)
-summary(STxLT_nullmodel_GAM.morph)
-print(plot(STxLT_nullmodel_GAM.morph, allTerms = TRUE), pages = 1)
-```
-
-### 2.1.3. Parsing niche contraction and niche separation
-
-Contrary to the predictions of H1, niche overlap decreases sharply at high elevation.
-This is true for both taxonomic and morphological niche overlap, and it holds in both the face-value and null-residual approaches.
-This decrease in niche overlap might be explained by two modest, proximate hypotheses:
-
-1.
-**H1:** The decrease of niche overlap at high elevations is due to *niche contraction*, i.e.
-BB spp.
-visiting a smaller proportion of available (and "eligible", i.e.
-visited at least once by a bumble bee) flora.
-  *Prediction:* "resource range" (Poisot et al.
-2012) will decrease with elevation (as Dormann notes, resource range is a "backward" metric insofar as a large resource range means a small proportional niche breadth; I will calculate the inverse of resource range so that it makes more sense)
-2.  **H2:** The decrease of niche overlap at high elevations is due to *niche separation*, i.e. BB spp. avoiding one another's niches but not contracting their own.
-  *Predictions:* C score will increase with elevation
-
-*Results:* I haven't run formal stats on this analysis yet, but from the visualization, it looks like the overall story is that niche overlap is driven mainly by niche separation, not so much but niche contraction.
-My first impression is that this suggests a behavior response to competition.
-
-```{r, echo = FALSE}
-resource_range.sp <- net %>%
-  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
-  filter(!is.na(pollen)) %>%
-  group_by(site, year, pollen, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-  unite(webID, c(site, year, pollen), sep = "_") %>%
-  ungroup() %>%
-  dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs() %>% # convert to to bipartite web
-  map(specieslevel, level = "higher", index = "resource range") %>%
-  map(rownames_to_column) %>%
-  map(as_tibble) %>%
-  bind_rows(.id = "webID") %>% # Collapse list into big data frame
-  rename(bb.sp = rowname) %>%
-  separate(webID, c("site", "year", "pollen"), sep =  "_") %>%
-  left_join(site_data) %>%
-  mutate(site = factor(site),
-         year = factor(year),
-         pollen = factor(pollen)) %>%
-  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # for some reason, one row has a "1" for bb.sp, and I don't know where this comes from; this gets rid of it, but I would still like to know why
-  mutate(resource.range = 1-resource.range) %>%
-  left_join(BB_covars.splevel) %>%
-  filter(bb.sp.abund > 1)
-
-resource_range.sp.mean <- resource_range.sp %>%
-  group_by(site, year, elev.mean, pollen) %>%
-  summarize(mean.resource.range = mean(resource.range))
-
-ggplot(resource_range.sp, aes(elev.mean, resource.range, color = pollen)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(bb.sp ~ year) +
-  ylim(c(0,1))
-
-ggplot(resource_range.sp.mean, aes(elev.mean, mean.resource.range, color = pollen)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~year) +
-  ylim(c(0,1))
-
-resource_range.ktype <- net %>%
-  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
-  filter(!is.na(pollen)) %>%
-  group_by(site, year, pollen, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
-  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-  unite(webID, c(site, year, pollen), sep = "_") %>%
-  ungroup() %>%
-  dplyr::select(higher = bb.sp, lower = k.type.s, webID, freq) %>% # rename columns to match bipartite's expectations
-  data.frame() %>% # convert to data frame
-  frame2webs() %>% # convert to to bipartite web
-  map(specieslevel, level = "higher", index = "resource range") %>%
-  map(rownames_to_column) %>%
-  map(as_tibble) %>%
-  bind_rows(.id = "webID") %>% # Collapse list into big data frame
-  rename(bb.sp = rowname) %>%
-  separate(webID, c("site", "year", "pollen"), sep =  "_") %>%
-  left_join(site_data) %>%
-  mutate(site = factor(site),
-         year = factor(year),
-         pollen = factor(pollen)) %>%
-  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # for some reason, one row has a "1" for bb.sp, and I don't know where this comes from; this gets rid of it, but I would still like to know why
-  mutate(resource.range = 1-resource.range) %>%
-  left_join(BB_covars.splevel) %>%
-  filter(bb.sp.abund > 1)
-
-resource_range.ktype.mean <- resource_range.ktype %>%
-  group_by(site, year, elev.mean, pollen) %>%
-  summarize(mean.resource.range = mean(resource.range))
-
-ggplot(resource_range.ktype, aes(elev.mean, resource.range, color = pollen)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(bb.sp ~ year) +
-  ylim(c(0,1))
-
-ggplot(resource_range.ktype.mean, aes(elev.mean, mean.resource.range, color = pollen)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~year) +
-  ylim(c(0,1))
-
-cscore.sp <- net %>%
-    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
-    filter(!is.na(pollen)) %>%
-    group_by(site, year, pollen, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-    summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-    group_by(site, year, pollen) %>%
-    unite(webID, c(site, year, pollen), sep = "_") %>%
-    ungroup() %>%
-    dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
-    data.frame() %>% # convert to data frame
-    frame2webs() %>% # convert to to bipartite web
-    map(grouplevel, level = "higher", weighted = TRUE, index = c("C score")) %>% 
-    data.frame() %>%
-    t() %>%
-    data.frame() %>%
-    rownames_to_column() %>%
-    dplyr::select(webID = rowname, everything()) %>%
-    as_tibble() %>%
-    separate(webID, c("site", "year", "pollen"), sep =  "_") %>%
-    left_join(site_data) %>%
-    mutate(site = factor(site),
-           year = factor(year),
-           pollen = factor(pollen))
-
-ggplot(cscore.sp, aes(elev.mean, C.score.HL, color = pollen)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~year)
-
-cscore.ktype <- net %>%
-    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
-    filter(!is.na(pollen)) %>%
-    group_by(site, year, pollen, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
-    summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-    group_by(site, year, pollen) %>%
-    unite(webID, c(site, year, pollen), sep = "_") %>%
-    ungroup() %>%
-    dplyr::select(higher = bb.sp, lower = k.type.s, webID, freq) %>% # rename columns to match bipartite's expectations
-    data.frame() %>% # convert to data frame
-    frame2webs() %>% # convert to to bipartite web
-    map(grouplevel, level = "higher", weighted = TRUE, index = c("C score")) %>% 
-    data.frame() %>%
-    t() %>%
-    data.frame() %>%
-    rownames_to_column() %>%
-    dplyr::select(webID = rowname, everything()) %>%
-    as_tibble() %>%
-    separate(webID, c("site", "year", "pollen"), sep =  "_") %>%
-    left_join(site_data) %>%
-    mutate(site = factor(site),
-           year = factor(year),
-           pollen = factor(pollen))
-
-ggplot(cscore.ktype, aes(elev.mean, C.score.HL, color = pollen)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~year)
-```

### 2020-12-07 analyses/niche_relations_v6B.Rmd
update lab desktop

@@
     group_by(..., bb.sp, !!criterion) %>%
-    summarize(prop = mean(prop)) %>%
+    summarize(prop = median(prop)) %>%
     #dplyr::select(-freq) %>% # drop freq variable
@@
   filter(caste != "male") %>%
-  filter(!bb.sp %in% c("psit", "gers", "humi")) %>% # only focal species
-  #group_by(elev.class2) %>%
+  filter(!bb.sp %in% c("gers", "humi")) %>% # only focal species
   group_by(site) %>%

### 2020-12-07 analyses/old_scripts/niche_relations_v4.Rmd
update lab desktop

--- 2020-11-26 analyses/niche_relations_v4.Rmd
@@
   left_join(fl_traits) %>% # add floral traits
-  group_by(elev.class) %>%
-  mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) # divide each elev.class into 3 equal time slices
+  group_by(elev.class2) %>%
+  mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) # divide each elev.class2 into 3 equal time slices
 
@@
 
+And we will parse the samples by elevation class and season, dividing the samples from each elevation class into equal-interval thirds
+```{r}
+samples <- net %>%
+  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
+  #group_by(elev.class2) %>%
+  group_by(site) %>%
+  mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
+  group_by(site, period, dayofyear) %>%
+  summarize(obs = n()) %>%
+  left_join(site_data) %>%
+  mutate(date = as_date(dayofyear))
+
+ggplot(samples, aes(as_date(yday(date)), elev.mean, color = period, pch = elev.class2)) +
+  #geom_point(size = 1, aes(alpha = obs)) +
+  geom_point(size = 2, alpha = 0.75) +
+  #facet_wrap(~year) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_alpha_continuous(name = "Observations") +
+  scale_x_date(date_labels = "%b", breaks = as_date(c("1970-04-01", "1970-05-01", "1970-06-01",
+                                                    "1970-07-01", "1970-08-01", "1970-09-01"))) +
+  #geom_hline(yintercept = c(755, 1500), linetype = "dashed") +
+  theme_light(12) +
+  labs(x = "Time",
+       y = "Elevation (m x 10)") 
+```
+
 ## Define functions
@@
   mutate(Score = case_when( # rename default scores to match my data
-    Score == "species" ~ "FL.sp",
-    Score == "sites" ~ "BB.sp"))
+    Score == "species" ~ "flora",
+    Label %in% c("hort", "pasc", "wurf") ~ "LT spp.",
+    Label %in% c("bss", "soro", "prat") ~ "ST spp."))
 }
@@
 
-  # Add T.class variable for the BB.sp subset of the PCoA
-  t.class <- x %>%
-    filter(Score == "BB.sp") %>%
-    mutate(T.class = factor(if_else(
-      Label %in% c("hort", "pasc", "wurf", "gers", "mend", "hypn"), "long", "short"))
-      )
-
-  ggplot(filter(x, (abs(MDS1) > min_score | abs(MDS2) > min_score) | (Score == "BB.sp"))) + # no threshold for BB spp
-    geom_polygon(data = t.class, aes(MDS1, MDS2, linetype = T.class), # draw polygons around T classes
-                 color = "gray70", fill = NA, size = 1.25, key_glyph = "abline") +
-    geom_point(aes(MDS1, MDS2, color = Score), size = 3, alpha = 0.75) + # add points
-    ggrepel::geom_text_repel(aes(MDS1, MDS2, color = Score, label = Label), key_glyph = "rect") + # add labels
+  # # Add T.class variable for the BB.sp subset of the PCoA
+  # t.class <- x %>%
+  #   filter(Score == "BB sp.") %>%
+  #   mutate(T.class = factor(if_else(
+  #     Label %in% c("hort", "pasc", "wurf", "gers", "mend", "hypn"), "long", "short"))
+  #     )
+
+  cols <- c("flora" = "gray40", "LT spp." = "#F8766D", "ST spp." = "#00BFC4")
+
+  ggplot(filter(x, (abs(MDS1) > min_score | abs(MDS2) > min_score) | (Score %in% c("LT spp.", "ST spp."))), 
+         aes(MDS1, MDS2, color = Score)) + # no threshold for BB spp
+    #geom_polygon(data = t.class, aes(MDS1, MDS2, linetype = T.class), # draw polygons around T classes
+     #            color = "gray70", fill = NA, size = 1.25, key_glyph = "abline") +
+    geom_point(size = 2, alpha = 0.75) + # add points
+    ggrepel::geom_text_repel(aes(label = Label), key_glyph = "rect", size = 3) + # add labels
     coord_fixed() + # Gavin Simpson says this is **really** important
-    theme_classic() 
+    theme_classic() +
+    theme(legend.title=element_blank()) +
+    scale_color_manual(values = cols)
 }
@@
     left_join(y) %>%
-    group_by(elev.class) %>%
+    group_by(elev.class2) %>%
     mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
@@
     filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
-    group_by(elev.class) %>%
+    group_by(elev.class2) %>%
     mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
@@
     left_join(y) %>%
-    # group_by(elev.class) %>%
+    # group_by(elev.class2) %>%
     # mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
@@
     filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
-    # group_by(elev.class) %>%
+    # group_by(elev.class2) %>%
     # mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
@@
 # Taxonomic
-niche_x_elevperiod.tax <- niche_analysis(net, bb_traits, criterion = plant.sp.abb, elev.class, period)
+niche_x_elevperiod.tax <- niche_analysis(net, bb_traits, criterion = plant.sp.abb, elev.class2, period)
 
 ## Extract and name plots
-niche_oben_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "oben" & period == "early")$pcoa.plot[[1]] +
-  ggtitle("Oben/Early/Tax")
-niche_oben_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "oben" & period == "mid")$pcoa.plot[[1]] +
-  ggtitle("Oben/Mid/Tax")
-niche_oben_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "oben" & period == "late")$pcoa.plot[[1]] +
-  ggtitle("Oben/Late/Tax")
-
-niche_mitte_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "mitte" & period == "early")$pcoa.plot[[1]] +
-  ggtitle("Mitte/Early/Tax")
-niche_mitte_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "mitte" & period == "mid")$pcoa.plot[[1]] +
-  ggtitle("Mitte/Mid/Tax")
-niche_mitte_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "mitte" & period == "late")$pcoa.plot[[1]] +
-  ggtitle("Mitte/Late/Tax")
-
-niche_unten_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "unten" & period == "early")$pcoa.plot[[1]] +
-  ggtitle("Unten/Early/Tax")
-niche_unten_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "unten" & period == "mid")$pcoa.plot[[1]] +
-  ggtitle("Unten/Mid/Tax")
-niche_unten_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class == "unten" & period == "late")$pcoa.plot[[1]] +
-  ggtitle("Unten/Late/Tax")
+niche_high_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "high" & period == "early")$pcoa.plot[[1]] 
+niche_high_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "high" & period == "mid")$pcoa.plot[[1]] 
+niche_high_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "high" & period == "late")$pcoa.plot[[1]]
+
+niche_mid_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "mid" & period == "early")$pcoa.plot[[1]]
+niche_mid_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "mid" & period == "mid")$pcoa.plot[[1]] 
+niche_mid_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "mid" & period == "late")$pcoa.plot[[1]]
+
+niche_low_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "low" & period == "early")$pcoa.plot[[1]]
+niche_low_mid.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "low" & period == "mid")$pcoa.plot[[1]]
+niche_low_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "low" & period == "late")$pcoa.plot[[1]]
 
 # Morphological
-niche_x_elevperiod.morph <- niche_analysis(net, bb_traits, criterion = k.type.s, elev.class, period)
+niche_x_elevperiod.morph <- niche_analysis(net, bb_traits, criterion = k.type.s, elev.class2, period)
 
 ## Extract and name plots
-niche_oben_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "oben" & period == "early")$pcoa.plot[[1]] +
-  ggtitle("Oben/Early/Morph")
-niche_oben_mid.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "oben" & period == "mid")$pcoa.plot[[1]] +
-  ggtitle("Oben/Mid/Morph")
-niche_oben_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "oben" & period == "late")$pcoa.plot[[1]] +
-  ggtitle("Oben/Late/Morph")
-
-niche_mitte_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "mitte" & period == "early")$pcoa.plot[[1]] +
-  ggtitle("Mitte/Early/Morph")
-niche_mitte_mid.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "mitte" & period == "mid")$pcoa.plot[[1]] +
-  ggtitle("Mitte/Mid/Morph")
-niche_mitte_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "mitte" & period == "late")$pcoa.plot[[1]] +
-  ggtitle("Mitte/Late/Morph")
-
-niche_unten_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "unten" & period == "early")$pcoa.plot[[1]] +
-  ggtitle("Unten/Early/Morph")
-niche_unten_mid.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "unten" & period == "mid")$pcoa.plot[[1]] +
-  ggtitle("Unten/Mid/Morph")
-niche_unten_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class == "unten" & period == "late")$pcoa.plot[[1]] +
-  ggtitle("Unten/Late/Morph")
+niche_high_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "high" & period == "early")$pcoa.plot[[1]] +
+  ggtitle("high/Early/Morph")
+niche_high_mid.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "high" & period == "mid")$pcoa.plot[[1]] +
+  ggtitle("high/Mid/Morph")
+niche_high_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "high" & period == "late")$pcoa.plot[[1]] +
+  ggtitle("high/Late/Morph")
+
+niche_mid_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "mid" & period == "early")$pcoa.plot[[1]] +
+  ggtitle("mid/Early/Morph")
+niche_mid_mid.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "mid" & period == "mid")$pcoa.plot[[1]] +
+  ggtitle("mid/Mid/Morph")
+niche_mid_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "mid" & period == "late")$pcoa.plot[[1]] +
+  ggtitle("mid/Late/Morph")
+
+niche_low_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "low" & period == "early")$pcoa.plot[[1]] +
+  ggtitle("low/Early/Morph")
+niche_low_mid.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "low" & period == "mid")$pcoa.plot[[1]] +
+  ggtitle("low/Mid/Morph")
+niche_low_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "low" & period == "late")$pcoa.plot[[1]] +
+  ggtitle("low/Late/Morph")
 
@@
 #pdf("../analyses/output/niche_x_elevperiod_pcoa_tax.pdf", height = 8.5, width = 11)
-(niche_oben_early.tax.plot | niche_oben_mid.tax.plot  | niche_oben_late.tax.plot) /
-(niche_mitte_early.tax.plot | niche_mitte_mid.tax.plot | niche_mitte_late.tax.plot) /
-(niche_unten_early.tax.plot | niche_unten_mid.tax.plot  | niche_unten_late.tax.plot)  + plot_layout(guides = 'collect') 
+(niche_high_early.tax.plot | niche_high_mid.tax.plot  | niche_high_late.tax.plot) /
+(niche_mid_early.tax.plot | niche_mid_mid.tax.plot | niche_mid_late.tax.plot) /
+(niche_low_early.tax.plot | niche_low_mid.tax.plot  | niche_low_late.tax.plot)  + plot_layout(guides = 'collect') 
 #dev.off()
@@
 #pdf("../analyses/output/niche_x_elevperiod_pcoa_morph.pdf", height = 8.5, width = 11)
-(niche_oben_early.morph.plot | niche_oben_mid.morph.plot  | niche_oben_late.morph.plot) /
-(niche_mitte_early.morph.plot | niche_mitte_mid.morph.plot | niche_mitte_late.morph.plot) /
-(niche_unten_early.morph.plot | niche_unten_mid.morph.plot  | niche_unten_late.morph.plot)  + plot_layout(guides = 'collect') 
+(niche_high_early.morph.plot | niche_high_mid.morph.plot  | niche_high_late.morph.plot) /
+(niche_mid_early.morph.plot | niche_mid_mid.morph.plot | niche_mid_late.morph.plot) /
+(niche_low_early.morph.plot | niche_low_mid.morph.plot  | niche_low_late.morph.plot)  + plot_layout(guides = 'collect') 
 #dev.off()
@@
   filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
-  group_by(elev.class) %>%
+  group_by(elev.class2) %>%
   mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
@@
          bb.sp = factor(bb.sp, levels = c("hort", "pasc", "wurf", "bss", "prat", "soro"))) %>%
-  group_by(period, elev.class, plant.genus, k.type.ss, plant.family, bb.sp) %>%
+  group_by(period, elev.class2, plant.genus, k.type.ss, plant.family, bb.sp) %>%
   summarize(abund = n()) %>%
-  group_by(elev.class, period, bb.sp) %>%
+  group_by(elev.class2, period, bb.sp) %>%
   mutate(prop = abund/sum(abund)) %>%
-  arrange(elev.class, period, bb.sp, -prop)
+  arrange(elev.class2, period, bb.sp, -prop)
 
@@
 ```{r}
-### Visitation by BB species, parsed by elev.class and period
+### Visitation by BB species, parsed by elev.class2 and period
 visitation_abund.sp <- net %>%
   filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
-  group_by(elev.class) %>%
+  group_by(elev.class2) %>%
   mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
@@
          bb.sp = factor(bb.sp, levels = c("hort", "pasc", "wurf", "bss", "prat", "soro"))) %>%
-  group_by(period, elev.class, plant.genus, k.type.ss, plant.family, bb.sp) %>%
+  group_by(period, elev.class2, plant.sp.abb, plant.genus, k.type.ss, plant.family, bb.sp) %>%
   summarize(abund = n()) %>%
-  group_by(elev.class, period, bb.sp) %>%
+  group_by(elev.class2, period, bb.sp) %>%
   mutate(prop = abund/sum(abund)) %>%
-  arrange(elev.class, period, bb.sp, -prop)
+  arrange(elev.class2, period, bb.sp, -prop)
 
@@
 # Early
-ggplot(filter(visitation_abund.sp, period == "early"), 
-       aes(area = prop, fill = k.type.ss, label = plant.genus)) +
-  geom_treemap(color = "black") +
+tp.early <- ggplot(filter(visitation_abund.sp, period == "early"), 
+                aes(area = prop, fill = k.type.ss, label = plant.sp.abb)) +
+  geom_treemap(color = "white") +
   geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
   scale_fill_manual(values = cols) + 
-  facet_grid(fct_rev(elev.class)~bb.sp)
+  labs(fill = "Morphotype") +
+  facet_grid(fct_rev(elev.class2)~bb.sp) +
+  theme(strip.text = element_text(size = 14))
 
 # Mid
-ggplot(filter(visitation_abund.sp, period == "mid"), 
-       aes(area = prop, fill = k.type.ss, label = plant.genus)) +
-  geom_treemap(color = "black") +
+tp.mid <- ggplot(filter(visitation_abund.sp, period == "mid"), 
+                 aes(area = prop, fill = k.type.ss, label = plant.sp.abb)) +
+  geom_treemap(color = "white") +
   geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
   scale_fill_manual(values = cols) + 
-  facet_grid(fct_rev(elev.class)~bb.sp)
+  labs(fill = "Morphotype") +
+  facet_grid(fct_rev(elev.class2)~bb.sp) +
+  theme(strip.text = element_text(size = 14))
 
 # Late
-ggplot(filter(visitation_abund.sp, period == "late"), 
-       aes(area = prop, fill = k.type.ss, label = plant.genus)) +
-  geom_treemap(color = "black") +
+tp.late <- ggplot(filter(visitation_abund.sp, period == "late"), 
+                  aes(area = prop, fill = k.type.ss, label = plant.sp.abb)) +
+  geom_treemap(color = "white") +
   geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
   scale_fill_manual(values = cols) + 
-  facet_grid(fct_rev(elev.class)~bb.sp)
-
+  labs(fill = "Morphotype") +
+  facet_grid(fct_rev(elev.class2)~bb.sp) +
+  theme(strip.text = element_text(size = 14))
+
+tp_ord1 <- (
+  tp.early | (
+    niche_high_early.tax.plot / 
+    niche_mid_early.tax.plot / 
+    niche_low_early.tax.plot
+    ) + 
+    plot_layout(guides = 'collect')
+  ) + 
+  plot_layout(widths = c(5, 1))
+ggsave("./output/tree_ord_early.png", tp_ord1, width = 14, height = 8)
+
+tp_ord1 <- (
+  tp.mid | (
+    niche_high_mid.tax.plot / 
+    niche_mid_mid.tax.plot / 
+    niche_low_mid.tax.plot
+    ) + 
+    plot_layout(guides = 'collect')
+  ) + 
+  plot_layout(widths = c(5, 1))
+ggsave("./output/tree_ord_mid.png", tp_ord1, width = 14, height = 8)
+
+tp_ord1 <- (
+  tp.late | (
+    niche_high_late.tax.plot / 
+    niche_mid_late.tax.plot / 
+    niche_low_late.tax.plot
+    ) + 
+    plot_layout(guides = 'collect')
+  ) + 
+  plot_layout(widths = c(5, 1))
+ggsave("./output/tree_ord_late.png", tp_ord1, width = 14, height = 8)
 ```

### 2020-12-07 analyses/old_scripts/niche_relations_v6.Rmd
update lab desktop

--- 2020-12-03 analyses/niche_relations_v6B.Rmd
@@
   scores <- y %>% 
-    #filter((abs(MDS1) > 0.25 | abs(MDS2) > 0.25) | Score == "BB.sp")
-    filter((abs(MDS1) > 0.1 | abs(MDS2) > 0.1) | Score == "BB.sp")
+    filter((abs(MDS1) > 0.25 | abs(MDS2) > 0.25) | Score == "BB.sp")
 
@@
     #filter(!bb.sp %in% c("psit", "gers", "humi")) %>% # omit parasitic species, gers because it's a specialist, humi because it's so rare
-    filter(!bb.sp %in% c("gers", "humi")) %>% # gers because it's a specialist, humi because it's so rare
-    #filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
-    group_by(site, date, bb.sp, !!criterion, ...) %>% # pooling all sites and years
+    filter(!bb.sp %in% c("gers", "humi")) %>% # omit parasitic species, gers because it's a specialist, humi because it's so rare
+    group_by(..., bb.sp, !!criterion) %>% # pooling all sites and years
     summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
-    group_by(site, date, bb.sp) %>%
-    mutate(prop = freq/sum(freq)) %>%
-    group_by(site, date, ...) %>%
-    complete(bb.sp, !!criterion, fill = list(prop = 0)) %>% # fill in zeros for unvisited plants
-    group_by(..., bb.sp, !!criterion) %>%
-    summarize(prop = mean(prop)) %>%
-    #dplyr::select(-freq) %>% # drop freq variable
+    group_by(..., bb.sp) %>%
+    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
+    dplyr::select(-freq) %>% # drop freq variable
     group_by(...) %>%
@@
            ordisurf = map2(pcoa, traits, function(x, y) ordisurf(x ~ pbl.w, data = y, k = 8)),
-           #ordisurf = map2(pcoa, traits, function(x, y) ordisurf(x ~ pbl.w, data = y, k = 5)),
            ordisurf.dev_exp = map(ordisurf, function(x) summary(x)$dev.expl),
@@
 # Morphological
-niche_x_elev.morph <- niche_analysis(net, bb_traits, criterion = k.type.ss, elev.class2)
+niche_x_elev.morph <- niche_analysis(net, bb_traits, criterion = k.type.s, elev.class2)
 

### 2020-12-07 analyses/RLQ4th.Rmd
rlq

---
title: "RLQ and Fourth Corner"
output: html_document
---

## Load packages
```{r, message=FALSE}
library(ade4)
library(tidyverse)
```

## Load data
```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
# Site data
# Site data
site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class2 = factor(elev.class2, levels = c("low", "mid", "high"))) # turn elav.class into an ordered factor

# BB traits
bb_traits <- read_csv("../processed_data/bb_traits.csv")

# FL traits
## family-level taxonomy
fl_tax <- read_csv("../processed_data/floral_tax.csv")

## Kugler morphotypes (accessed through Bioflor)
fl_ktype <- read_csv("../processed_data/floral_k_type.csv")

## Flower colors (accessed through BioFlor)
fl_color <- read_csv("../processed_data/floral_color.csv")

## diameter and corolla depth data from Gita Benadi
fl_morph <- read_csv("../processed_data/flower_morphology_GitaBenadi.csv") %>% 
  dplyr::select(SampleID, plant.family = Family, plant.sp = Species, 
                Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam), LN = mean(LN))

## verbal description of Kugler morphotypes
fl_ktype_key <- read_delim("../processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  dplyr::select(k.type = 1, description = 2)

## join all floral trait tables
fl_traits <- fl_tax %>%
  full_join(fl_ktype) %>%
  full_join(fl_color) %>%
  full_join(fl_morph) %>%
  dplyr::select(plant.sp, plant.genus, plant.family, k.type, k.type.s, k.type.ss, color, diam, LN)

# Network
net <- read_csv("../processed_data/network.csv") %>%
  mutate(year = factor(year)) %>% # convert year to factor
  left_join(site_data, by = "site") %>% # add site data
  ungroup() %>%
  arrange(site, date) %>%
  left_join(fl_traits) %>% # add floral traits
  group_by(elev.class2) %>%
  mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) %>% # divide each elev.class2 into 4 equal time slices
  mutate(k.type.s = as.character(k.type.s), # these can't be factors or else complete() will try to preserve all their levels for each group, even the levels that do not occur
         k.type.ss = as.character(k.type.ss))

# Floral survey
survey <- read_csv("../processed_data/floral_survey.csv") %>%
  filter(flower.cover > 0) %>% # only count plants in bloom
  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  left_join(fl_traits) # add floral traits
```

## Aggregate RLQ + fourth corner
```{r}
library(factoextra)
data(aravo)

# RLQ tables
L1 <- net %>%
  filter(caste != "male") %>%
  filter(!bb.sp %in% c("gers", "humi")) %>%
  group_by(bb.sp, plant.sp) %>%
  summarize(abund = n()) %>%
  complete(bb.sp, plant.sp, fill = list(abund = 0)) %>%
  pivot_wider(values_from = abund, names_from = plant.sp, values_fill = 0) %>%
  arrange(bb.sp) %>%
  as_tibble() %>%
  column_to_rownames("bb.sp")

R1 <- bb_traits %>%
  filter(!bb.sp %in% c("gers", "humi")) %>%
  select(bb.sp, clade, pbl.w) %>%
  arrange(bb.sp) %>%
  as_tibble() %>%
  mutate(clade = factor(clade)) %>%
  column_to_rownames("bb.sp")

Q1 <- fl_traits %>% 
  distinct() %>%
  semi_join(as_tibble(colnames(L1)), by = c("plant.sp" = "value")) %>%
  arrange(plant.sp) %>%
  select(plant.sp, plant.family, k.type.ss, color) %>%
  mutate(k.type.ss = factor(k.type.ss),
         color = factor(color),
         plant.family = factor(plant.family)) %>%
  column_to_rownames("plant.sp")

# Ordinations
L1.ca <- dudi.coa(L1, scannf = FALSE)
fviz_ca_biplot(L1.ca)

L1.ca2 <- cca(L1)
plot(L1.ca2, type = "t")

R1.ca <- dudi.hillsmith(R1, row.w = L1.ca$lw, scannf = FALSE)
summary(R1)

Q1.ca <- dudi.acm(Q1, row.w = L1.ca$cw, scannf = FALSE)

RLQ1 <- rlq(R1.ca, L1.ca, Q1.ca, scannf = FALSE)

plot(RLQ1)
summary(RLQ1)
nrepet <- 999
randtest(RLQ1, modeltype = 6, nrepet = nrepet)

par(mfrow = c(1, 3))
s.arrow(RLQ1$l1)
s.arrow(RLQ1$c1)
s.label(RLQ1$lQ, boxes = FALSE)

fc1 <- fourthcorner(R1, L1, Q1, 
                    modeltype = 6, p.adjust.method.G = "none",
                    p.adjust.method.D = "none", nrepet = nrepet)
plot(fc1)

testQaxes.comb.1 <- fourthcorner.rlq(RLQ1, modeltype = 6,
                                     typetest = "Q.axes", nrepet = nrepet, 
                                     p.adjust.method.G = "none", p.adjust.method.D = "none")

testRaxes.comb.1 <- fourthcorner.rlq(RLQ1, modeltype = 6,
                                     typetest = "R.axes", nrepet = nrepet, 
                                     p.adjust.method.G = "none", p.adjust.method.D = "none")

print(testQaxes.comb.1, stat = "D")
```

# Sandbox
```{r}
L1 <- net %>%
  filter(caste != "male") %>%
  filter(!bb.sp %in% c("gers", "humi")) %>%
  group_by(bb.sp, k.type.ss) %>%
  summarize(abund = n()) %>%
  ungroup() %>%
  complete(bb.sp, k.type.ss, fill = list(abund = 0)) %>%
  pivot_wider(values_from = abund, names_from = k.type.ss) %>%
  column_to_rownames("bb.sp") %>%
  dudi.coa(scannf = FALSE)

fviz_ca_biplot(L1)

L1 <- net %>%
  filter(caste != "male") %>%
  filter(!bb.sp %in% c("gers", "humi")) %>%
  group_by(bb.sp, color) %>%
  summarize(abund = n()) %>%
  ungroup() %>%
  complete(bb.sp, color, fill = list(abund = 0)) %>%
  pivot_wider(values_from = abund, names_from = color) %>%
  column_to_rownames("bb.sp") %>%
  dudi.coa(scannf = FALSE)

fviz_ca_biplot(L1)
```

### 2020-12-07 analyses/diet_ordination_CA.Rmd
rlq

--- 2020-12-07 analyses/diet_ordination.Rmd
@@
     as_tibble() %>%
-    rename(MDS1 = x, MDS2 = y)
+    rename(CA1 = x, CA2 = y)
 
   scores <- y %>% 
-    filter((abs(MDS1) > min_score | abs(MDS2) > min_score) | Score == "BB.sp") %>%
-    mutate(max.score = max(MDS1, MDS2))
+    filter((abs(CA1) > min_score | abs(CA2) > min_score) | Score == "BB.sp") %>%
+    mutate(max.score = max(CA1, CA2))
 
-  ggplot(scores, aes(MDS1, MDS2, pch = Score)) +
+  ggplot(scores, aes(CA1, CA2, pch = Score)) +
     geom_contour(data = contours, 
-                 aes(MDS1, MDS2, z = z, color = stat(level)),
+                 aes(CA1, CA2, z = z, color = stat(level)),
                  size = 1.5, alpha = 0.75, inherit.aes = FALSE) +
     geom_point(size = 2) +
-    geom_text_repel(aes(MDS1, MDS2, label = Label), key_glyph = "rect") +
+    geom_text_repel(aes(CA1, CA2, label = Label), key_glyph = "rect") +
     #geom_text(aes(MDS1, MDS2, label = Label, alpha = max.score), key_glyph = "rect") +
@@
 # Fortify PCOA and add the results of an envfit() to BB tongue length
-pcoa_fort <- function(x){
+ca_fort <- function(x){
   fortify(x, axes = 1:2) %>%
@@
 # Plot PCoA; to avoid overplotting, scores < min_score are not plotted
-pcoa_plot <- function(x, min_score = 0.25) {
+ca_plot <- function(x) {
 
@@
   #        aes(MDS1, MDS2, color = Score)) + # no threshold for BB spp
-  ggplot(x, aes(MDS1, MDS2, color = Score)) +
+  ggplot(x, aes(CA1, CA2, color = Score)) +
     #geom_polygon(data = t.class, aes(MDS1, MDS2, linetype = T.class), # draw polygons around T classes
@@
            ),
-           pcoa = map(comm2, # call PCoA
-                      function(x) capscale(x ~ 1, distance = "horn")
+           ca = map(comm2, # call PCoA
+                      function(x) cca(x)
                       ),
@@
                          ),
-           pcoa.fort = map(pcoa, pcoa_fort),
-           pcoa.plot = map(pcoa.fort, pcoa_plot),
+           ca.fort = map(ca, ca_fort),
+           ca.plot = map(ca.fort, ca_plot),
            adonis.r2 = map(adonis, # extract adonis r2
@@
                           ),
-           ordisurf = map2(pcoa, traits, function(x, y) ordisurf(x ~ pbl.w, data = y, k = k)),
+           ordisurf = map2(ca, traits, function(x, y) ordisurf(x ~ pbl.w, data = y, k = k)),
            ordisurf.dev_exp = map(ordisurf, function(x) summary(x)$dev.expl),
            ordisurf.p = map(ordisurf, function(x) summary(x)$s.pv),
-           ordisurf.plot = map2(ordisurf, pcoa.fort, surfplot),
-           var.exp.MDS1 = map(pcoa, # extract variance explained by MDS1
+           ordisurf.plot = map2(ordisurf, ca.fort, surfplot),
+           var.exp.MDS1 = map(ca, # extract variance explained by MDS1
                       function(x) x$CA$eig[[1]] / sum(x$CA$eig)
                       ), 
-           var.exp.MDS2 = map(pcoa, # extract variance explained by MDS2
+           var.exp.MDS2 = map(ca, # extract variance explained by MDS2
                       function(x) x$CA$eig[[2]] / sum(x$CA$eig)
@@
     mutate(var.exp.MDS1_MDS2 = var.exp.MDS1 + var.exp.MDS2) %>%
-    select(..., pcoa, pcoa.plot, var.exp.MDS1_MDS2, var.exp.MDS1, var.exp.MDS2, 
+    select(..., ca, ca.plot, var.exp.MDS1_MDS2, var.exp.MDS1, var.exp.MDS2, 
             ordisurf.dev_exp, ordisurf.p, adonis.r2, adonis.p,
-            ordisurf, ordisurf.plot, adonis, comm2, traits, pcoa.fort)
+            ordisurf, ordisurf.plot, adonis, comm2, traits, ca.fort)
+}
+
+niche_analysis2 <- function(x, y, criterion, k) {
+
+  criterion <- enquo(criterion) # this argument toggles taxonomic (plant.sp) vs. morphological (k.type.s) analysis
+
+  x %>%
+    filter(caste != "male") %>% # exclude males
+    filter(!bb.sp %in% c("gers", "humi")) %>% # exclude gers because it's a specialist, humi because it's so rare
+    group_by(site, date, bb.sp, !!criterion) %>% 
+    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-date
+    group_by(site, date) %>%
+    complete(bb.sp, !!criterion, fill = list(freq = 0)) %>% # fill in zeros for unvisited plants
+    group_by(site, date, bb.sp) %>%
+    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair per site-date 
+    group_by(site, bb.sp, !!criterion) %>%
+    summarize(prop = mean(prop)) %>% # get mean proportional intneraction frequency aggregated by ...
+    group_by(site) %>%
+    nest() %>% # nest into list-column
+    mutate(comm1 = map(data, # spread into the kind of array vegan likes
+                       function(x) pivot_wider(x, names_from = !!criterion, values_from = prop) %>%
+                         replace(is.na(.), 0)
+                       ), 
+           comm2 = map(comm1, # column to rowname
+                       function(x) column_to_rownames(x, var = "bb.sp")
+           ),
+           ca = map(comm2, # call PCoA
+                      function(x) cca(x)
+                      ),
+           traits = map(comm1, # join BB traits
+                        function(x) x %>% 
+                          left_join(y) %>% # y is the trait table, an argument in the main function call
+                          column_to_rownames(var = "bb.sp") %>% 
+                          select(pbl.w) # pbl.w is the variable I'm interested in
+                        ),
+           adonis = map2(comm2, traits, # run adonis permuatation test for distance ~ tongue class
+                         function(x, y) adonis(x ~ pbl.w, data = y, method = "horn")
+                         ),
+           ca.fort = map(ca, ca_fort),
+           ca.plot = map(ca.fort, ca_plot),
+           adonis.r2 = map(adonis, # extract adonis r2
+                           function(x) x$aov.tab$R2[[1]]
+                           ),
+           adonis.p = map(adonis, # extract adonis p-value
+                          function(x) x$aov.tab$'Pr(>F)'[[1]]
+                          ),
+           ordisurf = map2(ca, traits, function(x, y) ordisurf(x ~ pbl.w, data = y, k = k)),
+           ordisurf.dev_exp = map(ordisurf, function(x) summary(x)$dev.expl),
+           ordisurf.p = map(ordisurf, function(x) summary(x)$s.pv),
+           ordisurf.plot = map2(ordisurf, ca.fort, surfplot),
+           var.exp.MDS1 = map(ca, # extract variance explained by MDS1
+                      function(x) x$CA$eig[[1]] / sum(x$CA$eig)
+                      ), 
+           var.exp.MDS2 = map(ca, # extract variance explained by MDS2
+                      function(x) x$CA$eig[[2]] / sum(x$CA$eig)
+                      )) %>%
+    unnest(cols = c(ordisurf.dev_exp, ordisurf.p, adonis.r2, 
+                    adonis.p, var.exp.MDS1, var.exp.MDS2)) %>% # these are lists of length = 1, so unnesting them simply displays the value
+    mutate(var.exp.MDS1_MDS2 = var.exp.MDS1 + var.exp.MDS2) %>%
+    select(site, ca, ca.plot, var.exp.MDS1_MDS2, var.exp.MDS1, var.exp.MDS2, 
+            ordisurf.dev_exp, ordisurf.p, adonis.r2, adonis.p,
+            ordisurf, ordisurf.plot, adonis, comm2, traits, ca.fort)
 }
@@
 niche_x_elevperiod.tax <- niche_analysis(net, bb_traits, criterion = plant.sp.abb, k = 7, elev.class2, period)
+niche_x_site.tax <- niche_analysis2(net, bb_traits, criterion = plant.sp.abb, k = 7)
 

### 2020-12-07 data_proc.R

@@ -178,15 +178,15 @@ floral_color <-  read_delim("./BioFlor_traits/BioFlor_flower_color.csv", delim =
   select(plant.sp, plant.genus = genus, color) %>%
   
   # Fill in ktype data for plants present in my samples but missing from the Bioflor database
-  # Colors based on review of iNaturalist "research-grade" observations. Where color is too variable to classify, "various_colors". 
+  # Colors based on review of iNaturalist "research-grade" observations or Wikimedia commons. Where color is too variable to classify, "various_colors". 
   # Where essentially absent as in conifers, "NA".
-  add_row(plant.sp = "Euphrasia rostkoviana", plant.genus = "Euphrasia", color = "various colors") %>%
-  add_row(plant.sp = "Euphrasia officinalis", plant.genus = "Euphrasia", color = "various colors") %>%
+  add_row(plant.sp = "Euphrasia rostkoviana", plant.genus = "Euphrasia", color = "var") %>%
+  add_row(plant.sp = "Euphrasia officinalis", plant.genus = "Euphrasia", color = "var") %>%
   add_row(plant.sp = "Taraxacum officinale", plant.genus = "Taraxacum", color = "yellow") %>%
-  add_row(plant.sp = "Juniperus communis", plant.genus = "Juniperus", color = "NA") %>%
+  add_row(plant.sp = "Juniperus communis", plant.genus = "Juniperus", color = "green") %>%
   add_row(plant.sp = "Gentiana aspera", plant.genus = "Gentiana", color = "purple") %>%
   add_row(plant.sp = "Gentianella aspera", plant.genus = "Gentianella", color = "purple") %>%
-  add_row(plant.sp = "Carex ericetorum", plant.genus = "Carex", color = "NA") %>%
+  add_row(plant.sp = "Carex ericetorum", plant.genus = "Carex", color = "yellow") %>%
   add_row(plant.sp = "Euphrasia picta", plant.genus = "Euphrasia", color = "various colors") %>%
   add_row(plant.sp = "Gentiana ciliata", plant.genus = "Gentiana", color = "violet") %>%
   add_row(plant.sp = "Gentianopsis ciliata", plant.genus = "Gentiana", color = "violet") %>%
@@ -194,13 +194,13 @@ floral_color <-  read_delim("./BioFlor_traits/BioFlor_flower_color.csv", delim =
[…]

### 2020-12-09 analyses/floral_choice.Rmd
second take on floral choice

---
title: "Floral choice"
output: html_document
---

# Load packages
```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
library(vegan)
library(ggvegan)
library(ggrepel)
library(treemapify)
library(mgcv)
library(mgcViz)
library(tidyverse)
library(lubridate)
library(patchwork)
library(tidymodels)
```

# Load data
```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
# Site data
site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class2 = factor(elev.class2, levels = c("low", "mid", "high"))) # turn elav.class into an ordered factor

# BB traits
bb_traits <- read_csv("../processed_data/bb_traits.csv")

# FL traits
## family-level taxonomy
fl_tax <- read_csv("../processed_data/floral_tax.csv")

## Kugler morphotypes (accessed through Bioflor)
fl_ktype <- read_csv("../processed_data/floral_k_type.csv")

## Flower colors (to the human eye) (accessed through BioFlor)
fl_color <- read_csv("../processed_data/floral_color.csv")

## diameter and corolla depth data from Gita Benadi
fl_morph <- read_csv("../processed_data/flower_morphology_GitaBenadi.csv") %>% 
  dplyr::select(SampleID, plant.family = Family, plant.sp = Species, 
                Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam), LN = mean(LN))

## verbal description of Kugler morphotypes
fl_ktype_key <- read_delim("../processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  dplyr::select(k.type = 1, description = 2)

## join all floral trait tables
fl_traits <- fl_tax %>%
  full_join(fl_ktype) %>%
  full_join(fl_color) %>%
  full_join(fl_morph) %>%
  dplyr::select(plant.sp, plant.genus, plant.family, k.type, k.type.s, k.type.ss, color, diam, LN)

# Network
net <- read_csv("../processed_data/network.csv") %>%
  mutate(year = factor(year)) %>% # convert year to factor
  left_join(site_data, by = "site") %>% # add site data
  ungroup() %>%
  arrange(site, date) %>%
  left_join(fl_traits) %>% # add floral traits
  group_by(elev.class2) %>%
  mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) %>% # divide each elev.class2 into 3 equal time slices
  mutate(k.type.s = as.character(k.type.s), # these can't be factors or else complete() will try to preserve all their levels for each group, even the levels that do not occur
         k.type.ss = as.character(k.type.ss))

# Define periods based on visitation data
periods <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
  #group_by(elev.class2) %>%
  group_by(site) %>%
  mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) %>%
  ungroup() %>%
  select(site, date, period) %>%
  distinct()

# Floral survey
survey <- read_csv("../processed_data/floral_survey.csv") %>%
  filter(plant.sp != "Huperzia selago") %>% # we can do without the fern
  left_join(periods) %>% # add periods
  fill(period, .direction = "downup") %>% # fill in periods
  filter(flower.cover > 0) %>% # only count plants in bloom
  dplyr::select(site, date, period, plant.sp, flower.cover) %>%
  mutate(year = factor(year(date))) %>% # convert year to factor
  left_join(fl_traits) # add floral traits

View(filter(survey, is.na(k.type.s)) %>% select(plant.sp) %>% unique())
```

# What proportion of the surveyed floral pool is visited, and how does this vary with elevation?
```{r}
surveyed.siteyear <- survey %>%
  group_by(site, plant.sp, year) %>%
  summarize(flower.cover = sum(flower.cover))

visited.siteyear <- net %>%
  group_by(site, plant.sp, year) %>%
  summarize(visits = n())

auxilliary_data.survey <- survey %>%
  group_by(site, year) %>%
  summarize(survey.dates = length(unique(date)))

auxilliary_data.visits <- net %>%
  group_by(site, year) %>%
  summarize(sample.dates = length(unique(date)),
            bb.abund = n())

join.siteyear <- full_join(surveyed.siteyear, visited.siteyear)

prop_visited.siteyear <- join.siteyear %>%
  replace_na(replace = list(visits = 0, flower.cover = 0)) %>%
  mutate(surveyed = if_else(flower.cover > 0, 1, 0),
         visited = if_else(visits > 0, 1, 0)) %>%
  group_by(site, year) %>%
  summarize(n.visited = sum(visited),
            n.surveyed = sum(surveyed),
            prop.visited = n.visited/n.surveyed) %>%
  left_join(auxilliary_data.survey) %>%
  left_join(auxilliary_data.visits) %>%
  left_join(site_data)

ggplot(prop_visited.siteyear, aes(year, prop.visited)) +
  geom_violin()

ggplot(prop_visited.siteyear, aes(elev.mean, prop.visited, color = year)) +
  geom_point() +
  geom_smooth()

prop_visited.siteyear.gam <- gam(prop.visited ~ 
                                   year +
                                   s(bb.abund, k = 5, bs = "tp") +
                                   s(elev.mean, k = 5, bs = "tp"),
                                 data = filter(prop_visited.siteyear, !is.infinite(prop.visited)),
                                 method = "REML",
                                 select = TRUE) %>% getViz()

summary(prop_visited.siteyear.gam)
check.gamViz(prop_visited.siteyear.gam)
print(plot(prop_visited.siteyear.gam, allTerms = TRUE), pages = 1)
```

# How is the surveyed pool biased -- phylogenetically, taxonomically, morphologically -- by BB foraging. that is, which taxa/clades/morphologies are over- or underrepresented in the visitation data? And how does this vary with elevation?
```{r}
# By plant family
families.survey <- survey %>% 
  group_by(plant.family) %>%
  summarize(flower.cover = sum(flower.cover)) %>%
  mutate(survey.prop = flower.cover/sum(flower.cover))

families.visits <- net %>% 
  group_by(plant.family) %>%
  summarize(visits = n()) %>%
  mutate(visit.prop = visits/sum(visits))

families <- full_join(families.survey, families.visits) %>%
  select(plant.family, survey.prop, visit.prop) %>%
  replace_na(replace = list(survey.prop = 0, visit.prop = 0)) %>%
  #mutate(representation = (visit.prop - survey.prop)/survey.prop) %>% # representation as fold-change
  mutate(representation = visit.prop - survey.prop) %>% # representation as difference
  pivot_longer(cols = c(survey.prop, visit.prop, representation))

ggplot(filter(families, name != "representation"), 
       aes(name, value, fill = plant.family)) +
  geom_col()

ggplot(filter(families, name == "representation"), 
       aes(plant.family, value)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# By Kugler morphotype
morphotypes.survey <- survey %>% 
  group_by(k.type.ss) %>%
  summarize(flower.cover = sum(flower.cover)) %>%
  mutate(survey.prop = flower.cover/sum(flower.cover),
         k.type.ss = factor(k.type.ss))

morphotypes.visits <- net %>% 
  group_by(k.type.ss) %>%
  summarize(visits = n()) %>%
  mutate(visit.prop = visits/sum(visits),
         k.type.ss = factor(k.type.ss))

morphotypes <- full_join(morphotypes.survey, morphotypes.visits) %>%
  select(k.type.ss, survey.prop, visit.prop) %>%
  replace_na(replace = list(survey.prop = 0, visit.prop = 0)) %>%
  #mutate(representation = (visit.prop - survey.prop)/survey.prop) %>% # representation as fold-change
  mutate(representation = visit.prop - survey.prop) %>% # representation as difference
  pivot_longer(cols = c(survey.prop, visit.prop, representation))

ggplot(filter(morphotypes, name != "representation"), 
       aes(name, value, fill = k.type.ss)) +
  geom_col()

ggplot(filter(morphotypes, name == "representation"), 
       aes(k.type.ss, value)) +
  geom_col()
```

### 2020-12-09 data_proc.R

@@ -65,28 +65,34 @@ write_csv(network, "./processed_data/network.csv")
 ### Floral taxonomy
 Sys.setenv(ENTREZ_KEY = "ab9a55ec842df6f86a750929aefc69143608")
 
-fl_sp <- survey %>%
+fl_sp <- fl_survey %>%
   select(plant.sp) %>%
   unique() %>%
   bind_rows(select(network, plant.sp)) %>%
   unique()
 
-fl_tax <- tax_name(fl_sp$plant.sp, get = "family", db = "ncbi") %>%
+fl_tax <- taxize::tax_name(fl_sp$plant.sp, get = "family", db = "ncbi") %>%
   select(plant.sp = query, plant.family = family) %>%
-  separate(plant.sp, c("plant.genus", "plant.sp"), remove = FALSE) %>%
-  select(plant.genus, plant.family) %>%
+  separate(plant.sp, c("plant.genus", "sp"), remove = FALSE) %>%
+  select(plant.sp, plant.genus, plant.family) %>%
   unique()
 
 fl_tax_gapfill <- fl_tax %>% # manually fill in some gaps in the NCBI database
-  mutate(plant.family = case_when(
-    plant.genus == "Crocus" ~ "Iridaceae",
-    plant.genus == "Rheum" ~ "Polygonaceae",
-    plant.genus == "Mentha" ~ "Lamiaceae",
-    plant.genus %in% c("Senecio", "Crepis") ~ "Asteraceae",
-    plant.genus == "Salix" ~ "Salicaceae",
-    !plant.genus %in% c("Senecio", "Crocus", "Crepis", "Mentha", "Salix", "Rheum") ~ plant.family
-  ))
-
+  mutate(plant.family = if_else(is.na(plant.family), 
+                                case_when(
+                                  plant.genus == "Agrostis" ~ "Poaceae",
+                                  plant.genus == "Calamintha" ~ "Lamiaceae",
+                                  plant.genus == "Crepis" ~ "Asteraceae",
+                                  plant.genus == "Crocus" ~ "Iridaceae",
+                                  plant.genus == "Mentha" ~ "Lamiaceae",
+                                  plant.genus == "Minuartia" ~ "Caryophyllaceae",
+                                  plant.genus == "Papaver" ~ "Papaveraceae",
+                                  plant.genus == "Rheum" ~ "Polygonaceae",
[…]

### 2020-12-10 analyses/floral_choice.Rmd
anova

@@
 library(mgcViz)
+library(rstatix)
 library(tidyverse)
@@
   left_join(fl_traits) # add floral traits
-
-View(filter(survey, is.na(k.type.s)) %>% select(plant.sp) %>% unique())
 ```
@@
 families.survey <- survey %>% 
-  group_by(plant.family) %>%
+  group_by(plant.family, site) %>%
   summarize(flower.cover = sum(flower.cover)) %>%
+  group_by(site) %>%
   mutate(survey.prop = flower.cover/sum(flower.cover))
@@
 families.visits <- net %>% 
-  group_by(plant.family) %>%
+  group_by(plant.family, site) %>%
   summarize(visits = n()) %>%
+  group_by(site) %>%
   mutate(visit.prop = visits/sum(visits))
@@
 families <- full_join(families.survey, families.visits) %>%
-  select(plant.family, survey.prop, visit.prop) %>%
+  select(plant.family, site,  survey.prop, visit.prop) %>%
   replace_na(replace = list(survey.prop = 0, visit.prop = 0)) %>%
-  #mutate(representation = (visit.prop - survey.prop)/survey.prop) %>% # representation as fold-change
-  mutate(representation = visit.prop - survey.prop) %>% # representation as difference
-  pivot_longer(cols = c(survey.prop, visit.prop, representation))
+  mutate(representation.fold = (visit.prop - survey.prop)/survey.prop, # representation as fold-change
+         representation = visit.prop - survey.prop) %>%
+  ungroup()
 
-ggplot(filter(families, name != "representation"), 
-       aes(name, value, fill = plant.family)) +
-  geom_col()
+ggplot(families, aes(plant.family, representation)) +
+  geom_boxplot()
 
-ggplot(filter(families, name == "representation"), 
-       aes(plant.family, value)) +
-  geom_col() +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+ggplot(families, aes(plant.family, representation.fold)) +
+  geom_boxplot()
+
+aov.fam <- aov(representation ~ plant.family, data = families)
+aov.fam.test <- anova_test(aov.fam)
+aov.fam.tuk <- tukey_hsd(aov.fam)
+
+aov.fam2 <- aov(representation.fold ~ plant.family, data = filter(families, !is.infinite(representation.fold)))
+aov.fam2.test <- anova_test(aov.fam2)
+aov.fam2.tuk <- tukey_hsd(aov.fam2)
 
@@
 morphotypes.survey <- survey %>% 
-  group_by(k.type.ss) %>%
+  group_by(k.type.ss, site) %>%
   summarize(flower.cover = sum(flower.cover)) %>%
+  group_by(site) %>%
   mutate(survey.prop = flower.cover/sum(flower.cover),
@@
 morphotypes.visits <- net %>% 
-  group_by(k.type.ss) %>%
+  group_by(k.type.ss, site) %>%
   summarize(visits = n()) %>%
+  group_by(site) %>%
   mutate(visit.prop = visits/sum(visits),
@@
 morphotypes <- full_join(morphotypes.survey, morphotypes.visits) %>%
-  select(k.type.ss, survey.prop, visit.prop) %>%
+  select(k.type.ss, site,  survey.prop, visit.prop) %>%
   replace_na(replace = list(survey.prop = 0, visit.prop = 0)) %>%
-  #mutate(representation = (visit.prop - survey.prop)/survey.prop) %>% # representation as fold-change
-  mutate(representation = visit.prop - survey.prop) %>% # representation as difference
-  pivot_longer(cols = c(survey.prop, visit.prop, representation))
+  mutate(representation.fold = (visit.prop - survey.prop)/survey.prop, # representation as fold-change
+         representation = visit.prop - survey.prop) %>%
+  ungroup()
 
-ggplot(filter(morphotypes, name != "representation"), 
-       aes(name, value, fill = k.type.ss)) +
-  geom_col()
+ggplot(morphotypes, aes(k.type.ss, representation)) +
+  geom_boxplot()
 
-ggplot(filter(morphotypes, name == "representation"), 
-       aes(k.type.ss, value)) +
-  geom_col()
+ggplot(morphotypes, aes(k.type.ss, representation.fold)) +
+  geom_boxplot()
+
+aov.morph <- aov(representation ~ k.type.ss, data = morphotypes)
+aov.morph.test <- anova_test(aov.morph)
+aov.morph.tuk <- tukey_hsd(aov.morph)
+
+aov.morph2 <- aov(representation.fold ~ k.type.ss, data = filter(morphotypes, !is.infinite(representation.fold)))
+aov.morph2.test <- anova_test(aov.morph2)
+aov.morph2.tuk <- tukey_hsd(aov.morph2)
+
+library(brms)
+brm.morph <- brm(representation ~ k.type.ss, data = morphotypes)
 ```

### 2020-12-18 analyses/floral_choice.Rmd
developed sweet visitation bias analysis

@@
 library(rstatix)
+library(brms)
 library(tidyverse)
@@
 visited.siteyear <- net %>%
+  filter(caste != "male") %>%
   group_by(site, plant.sp, year) %>%
@@
 auxilliary_data.visits <- net %>%
+  filter(caste != "male") %>%
   group_by(site, year) %>%
@@
 families.visits <- net %>% 
+  filter(caste != "male") %>%
   group_by(plant.family, site) %>%
@@
          representation = visit.prop - survey.prop) %>%
-  ungroup()
+  group_by(plant.family) %>%
+  mutate(survey.prop.norm = scale(survey.prop),
+         visit.prop.norm = scale(visit.prop)) %>%
+  ungroup() %>%
+  left_join(site_data)
 
@@
 
-aov.fam <- aov(representation ~ plant.family, data = families)
-aov.fam.test <- anova_test(aov.fam)
-aov.fam.tuk <- tukey_hsd(aov.fam)
-
-aov.fam2 <- aov(representation.fold ~ plant.family, data = filter(families, !is.infinite(representation.fold)))
-aov.fam2.test <- anova_test(aov.fam2)
-aov.fam2.tuk <- tukey_hsd(aov.fam2)
+brm.fam <- brm(representation ~ plant.family, data = families)
+plot(brm.fam)
+plot(marginal_effects(brm.fam))
+hist(resid(brm.fam))
+mcmc_plot(brm.fam, type = "intervals")
 
@@
 morphotypes.visits <- net %>% 
+  filter(caste != "male") %>%
   group_by(k.type.ss, site) %>%
@@
          representation = visit.prop - survey.prop) %>%
-  ungroup()
+  group_by(k.type.ss) %>%
+  mutate(survey.prop.norm = scale(survey.prop),
+         visit.prop.norm = scale(visit.prop)) %>%
+  ungroup() %>%
+  left_join(site_data)
 
@@
 
-aov.morph <- aov(representation ~ k.type.ss, data = morphotypes)
-aov.morph.test <- anova_test(aov.morph)
-aov.morph.tuk <- tukey_hsd(aov.morph)
+brm.morph <- brm(representation ~ k.type.ss, data = morphotypes)
+plot(brm.morph)
+plot(marginal_effects(brm.morph))
+hist(resid(brm.morph))
+mcmc_plot(brm.morph, type = "intervals", pars = c("b_k.type.ss1", "b_k.type.ss2", "b_k.type.ss3",
+                                                  "b_k.type.ss4", "b_k.type.ss5", "b_k.type.ss6",
+                                                  "b_k.type.ss7", "b_k.type.ss8", "b_k.type.ss9",
+                                                  "b_k.type.ss10"))
+mcmc_plot(brm.morph, type = "intervals")
 
-aov.morph2 <- aov(representation.fold ~ k.type.ss, data = filter(morphotypes, !is.infinite(representation.fold)))
-aov.morph2.test <- anova_test(aov.morph2)
-aov.morph2.tuk <- tukey_hsd(aov.morph2)
+### By elevation
+ggplot(families, aes(elev.mean, representation)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~plant.family)
 
-library(brms)
-brm.morph <- brm(representation ~ k.type.ss, data = morphotypes)
+ggplot(families, aes(elev.mean, visit.prop)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~plant.family)
+
+ggplot(families, aes(elev.mean, visit.prop.norm)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~plant.family)
+
+ggplot(families, aes(elev.mean, survey.prop)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~plant.family)
+
+ggplot(families, aes(elev.mean, survey.prop.norm)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~plant.family)
+
+ggplot(morphotypes, aes(elev.mean, representation)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~k.type.ss)
+
+ggplot(morphotypes, aes(elev.mean, visit.prop)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~k.type.ss)
+
+ggplot(morphotypes, aes(elev.mean, visit.prop.norm)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~k.type.ss)
+
+ggplot(morphotypes, aes(elev.mean, survey.prop)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~k.type.ss)
+
+ggplot(morphotypes, aes(elev.mean, survey.prop.norm)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~k.type.ss)
 ```
+
+# Do BB species differ in their taxonomic/phylogenetic/morphological biases? And how does this vary with elevation/time?
+```{r}
+morphotypes.visits_sp <- net %>% 
+  group_by(k.type.ss, site, bb.sp) %>%
+  summarize(visits = n()) %>%
+  group_by(site, bb.sp) %>%
+  mutate(visit.prop = visits/sum(visits))
+
+sp_by_site <- net %>%
+  ungroup() %>%
+  select(site, bb.sp) %>%
+  distinct()
+
+morphotypes_sp1 <- full_join(morphotypes.survey, morphotypes.visits_sp)
+
+morphotypes_never_visited <- filter(morphotypes_sp1, is.na(bb.sp)) %>%
+  select(-bb.sp) %>%
+  left_join(sp_by_site) %>%
+  replace_na(replace = list(visits = 0, visit.prop = 0))
+
+morphotypes_sp <- full_join(morphotypes.survey, morphotypes.visits_sp) %>%
+  filter(!is.na(bb.sp)) %>%
+  bind_rows(morphotypes_never_visited) %>%
+  select(k.type.ss, site, bb.sp, survey.prop, visit.prop) %>%
+  replace_na(replace = list(survey.prop = 0, visit.prop = 0)) %>%
+  mutate(representation.fold = (visit.prop - survey.prop)/survey.prop, # representation as fold-change
+         representation = visit.prop - survey.prop) %>%
+  group_by(k.type.ss, bb.sp) %>%
+  mutate(survey.prop.norm = scale(survey.prop),
+         visit.prop.norm = scale(visit.prop)) %>%
+  ungroup() %>%
+  left_join(site_data)
+
+ggplot(filter(morphotypes_sp, bb.sp %in% c("pasc", "prat", "bss", "wurf", "soro", "psit", "hort")), 
+       aes(elev.mean, representation, color = bb.sp)) +
+  geom_line() +
+  #geom_smooth() +
+  facet_wrap(~k.type.ss)
+```

### 2020-12-18 analyses/visitation_bias.Rmd
developed sweet visitation bias analysis

---
title: "visitation_bias"
author: "Doug Sponsler"
date: "12/14/2020"
output: html_document
---
# Preparations
## Packages 
```{r, warning=FALSE, include=FALSE, echo = FALSE}
library(bipartite)
library(mgcv)
library(mgcViz)
library(brms)
library(bayesplot)
#library(tidybayes)
library(tidyverse)
library(lubridate)
```

## Load data
```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
# Site data
site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class2 = factor(elev.class2, levels = c("low", "mid", "high")),
         site = factor(site)) # turn elav.class into an ordered factor and site into a regular factor

# BB traits
bb_traits <- read_csv("../processed_data/bb_traits.csv")

# FL traits
## family-level taxonomy
fl_tax <- read_csv("../processed_data/floral_tax.csv")

## Kugler morphotypes (accessed through Bioflor)
fl_ktype <- read_csv("../processed_data/floral_k_type.csv")

## Flower colors (accessed through BioFlor)
fl_color <- read_csv("../processed_data/floral_color.csv")

## diameter and corolla depth data from Gita Benadi
fl_morph <- read_csv("../processed_data/flower_morphology_GitaBenadi.csv") %>% 
  dplyr::select(SampleID, plant.family = Family, plant.sp = Species, 
                Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam), LN = mean(LN))

## verbal description of Kugler morphotypes
fl_ktype_key <- read_delim("../processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  dplyr::select(k.type = 1, description = 2)

## join all floral trait tables
fl_traits <- fl_tax %>%
  full_join(fl_ktype) %>%
  full_join(fl_color) %>%
  full_join(fl_morph) %>%
  dplyr::select(plant.sp, plant.genus, plant.family, k.type, k.type.s, k.type.ss, color, diam, LN)
```

## Align survey and visitation data with pseudodates
```{r}
# Site-dates where BB visitation was recorded (excluding dates where only males were recorded)
visitation_samples <- read_csv("../processed_data/network.csv") %>%
  filter(caste != "male") %>%
  select(site, date) %>%
  distinct()

# Site dates where floral surveying was conducted
survey_samples <- read_csv("../processed_data/floral_survey.csv") %>%
  select(site, date) %>%
  distinct()

# Site-dates where visitation sampling and floral sampling coincided
shared_dates <- inner_join(visitation_samples, survey_samples)

# Site-dates where visitation was recorded WITHOUT corresponding survey data
visitation_only <- anti_join(visitation_samples, survey_samples) %>%
  mutate(date.v = date) %>%
  select(-date)

# This creates a pseudo date column for unmatched visitation site-dates that will match each visitation date
# to its nearest survey data. For site t4, there is a 2-year mismatch; we will drop these site-dates later.
pseudodates <- full_join(visitation_only, survey_samples) %>%
  mutate(mismatch = abs(date.v - date)) %>%
  group_by(site, date.v) %>%
  filter(mismatch == min(mismatch)) %>%
  rename(date = date.v, pseudo.date = date)
```

## Generate a data frame that represents visitation and survey abundance
```{r}
# BB visitation data
visitation <- read_csv("../processed_data/network.csv") %>%
  filter(caste != "male") %>% # we will ignore males 
  group_by(site, date, bb.sp, plant.sp) %>%
  summarize(visits = n()) %>% # visits per BB-FL pair
  ungroup() %>%
  left_join(pseudodates) %>% # Now we will replace unmatched dates with their pseudo dates
  replace_na(replace = list(mismatch = 0)) %>% # set mismatch to zero for perfectly aligned dates
  filter(mismatch <= 7) %>% # this gets rid of irreconcilable dates (like the t4 dates mentioned above)
  mutate(date2 = if_else(is.na(pseudo.date), date, pseudo.date)) %>% # replace date with pseudodate for all site-dates lacking perfect alignments
  select(-c(date, pseudo.date, mismatch)) %>%
  rename(date = date2) %>%
  group_by(site, date) %>%
  complete(bb.sp, plant.sp, fill = list(visits = 0)) # generate all possible BB-FL combos for each site date and fill in zeros for visits

# Floral cover data
flower_cover <- read_csv("../processed_data/floral_survey.csv") %>%
  filter(flower.cover > 0) %>% # ignore plants that were not blooming
  group_by(site, date, plant.sp) %>%
  summarize(flower.cover = sum(flower.cover))

# Get shared dates after adding pseudodates. This will let us filter out survey dates that don't
# have corresponding visitation sampling dates
shared_dates_with_pseudo <- inner_join(visitation, flower_cover, by = c("site", "date")) %>%
  select(site, date) %>%
  distinct()

# Join visitation and flower cover data

## First we need a list of all BB spp. by site and date
sp_by_site <- visitation %>% 
  ungroup() %>%
  select(site, date, bb.sp) %>%
  distinct()

## Then we need a list of unvisited plant.sp by site and date
unvisited <- flower_cover %>% 
  semi_join(shared_dates_with_pseudo) %>% # filter flower cover data to include only site-dates with visitation complement (real or pseudodates)
  full_join(visitation) %>% # join to visitation data
  filter(is.na(bb.sp)) %>% # select only the site-date-plants that were totally unvisited (and therefore have no BB spp to join to)
  select(-bb.sp) %>% # drop bb.sp variable because I just want the plants
  left_join(sp_by_site) %>% # join in the species list by site and date, thus populating the unvisited plants with all the BB spp recorded for each site-date
  replace_na(replace = list(visits = 0)) # replace NAs (created in the line above) with zero

## Now we can combine everything into a single data frame
visit_cover <- flower_cover %>% 
  semi_join(shared_dates_with_pseudo) %>% # filter flower cover data to include only site-dates with visitation complement (real or pseudodates)
  full_join(visitation) %>% # join to visitation data
  replace_na(replace = list(flower.cover = 0.01)) %>% # If a plant was visited but not surveyed, we will assign it the minimum non-zero flower cover value of 0.01.
  filter(!is.na(bb.sp)) %>% # drop the univisited plants
  bind_rows(unvisited) %>% # add them back, with BB spp and zeros
  select(site, date, bb.sp, plant.sp, visits, flower.cover) %>%
  group_by(site, date, bb.sp, plant.sp, flower.cover) %>% # In a few cases, the use of pseudodates causes two sampling period to converge on the same date, which creates apparent duplicate visitation data. We will handle this by averaging the visits.
  summarize(visits = mean(visits)) %>%
  left_join(fl_traits) %>% # add floral traits
  ungroup()

## Get total visits by BB sp and site-date; this will be a measure of BB abundance that we can use after we've calculated d'. I could, of course, work this into the biaseval function, but I'll put that off for now.
bb_abundance <- visit_cover %>%
  group_by(site, date, bb.sp) %>%
  summarize(bb.sp.abund = sum(visits)) %>%
  group_by(site, date) %>%
  mutate(bb.sp.prop.abund = bb.sp.abund/sum(bb.sp.abund))
```

## A handy function for calculating Blüthgen's d'
```{r}
# A function for calculating Blüthgen's d', parsed in various ways
biaseval <- function(x, criterion, ...) { # criterion = plant grouping variable, e.g. plant.sp, k.type.ss, color

  criterion <- enquo(criterion) # tidy eval

  x %>%

# Select variables of interest; the ... represents the variables we will use to parse our analysis
    select(..., bb.sp, !!criterion, flower.cover, visits) %>% 
    group_by(..., bb.sp, !!criterion) %>%

# If criterion = plant.sp, this next step does nothing. But if criterion is some other grouping variable (e.g. k.type.ss), this sums flower cover and visits over that grouping variable.
    summarize(flower.cover = sum(flower.cover), 
              visits = sum(visits)) %>%

# Nest into list column
    group_by(...) %>%
    nest() %>%

# Pivot each nested data frame into the kind of web that bipartite likes
    mutate(webs = map(data, function(x) select(x, bb.sp, !!criterion, visits) %>%
                        pivot_wider(names_from = !!criterion, values_from = visits, values_fill = 0) %>%
                        column_to_rownames("bb.sp"))) %>%

# Pull out the flower.cover variable to use in the "abuns" argument of dfun below
    mutate(low.abund = map(data, function(x) select(x, !!criterion, flower.cover) %>%
                            unique() %>%
                            column_to_rownames(paste(criterion)[2]))) %>% # Tell me why I need to do this, Hadley.

# Call the dfun command from bipartite to calculate Blüthgen's d'
    mutate(d = map2(webs, low.abund, function(x, y) bipartite::dfun(x, abuns = y$flower.cover) %>%
                      data.frame() %>%
                      rownames_to_column(var = "bb.sp"))) %>%

    unnest(cols = d) %>%
    na.omit() %>% # in a few cases where the sampling is very sparse, d' cannot be calculated; these are safe to drop
    ungroup() %>%
    mutate(site = factor(site),
         bb.sp = factor(bb.sp),
         year = factor(lubridate::year(date)))
}
```

# 1. Is bumble bee foraging biased? How do taxonomic, phylogenetic, and morphological bias compare? Are some BB species more biased than others?
## 1.1 Calculate Blüthgen's d' (i.e. visitation bias)
```{r}
# No grouping -- just plant.sp
bias_plantsp.sitedate <- biaseval(visit_cover, plant.sp, site, date) %>%
  left_join(site_data) %>% # add site data
  left_join(bb_abundance) %>% # add BB abundance data
  mutate(criterion = factor(rep("plant.sp", n())))

# Grouped by k.type.ss
bias_ktype.sitedate <- biaseval(visit_cover, k.type.ss, site, date) %>%
  left_join(site_data) %>%
  left_join(bb_abundance) %>% # add BB abundance data
  mutate(criterion = factor(rep("ktype", n())))

# Grouped by plant.family
bias_family.sitedate <- biaseval(visit_cover, plant.family, site, date) %>%
  left_join(site_data) %>%
  left_join(bb_abundance) %>% # add BB abundance data
  mutate(criterion = factor(rep("plant.family", n())))

# Collated
bias_sitedate <- bind_rows(bias_plantsp.sitedate, 
                           bias_ktype.sitedate, 
                           bias_family.sitedate)
```

## 1.2 Visualize
```{r}
# Compare mean bias by criterion
ggplot(bias_sitedate, aes(criterion, dprime)) +
  geom_boxplot()

# Bias ~ elevation * criterion
ggplot(bias_sitedate, aes(elev.mean, dprime, color = criterion)) +
  geom_point(alpha = 0.25) +
  geom_smooth(se = TRUE)

ggplot(bias_sitedate, aes(elev.mean, dprime, color = criterion)) +
  geom_point(alpha = 0.25) +
  geom_smooth(se = FALSE, method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp)

# Bias ~ year; Blüthgen's d' seems very robust to sampling effects, since it does not vary across years
ggplot(bias_sitedate, aes(year, dprime, color = criterion)) +
  geom_boxplot()

ggplot(bias_sitedate, aes(year, dprime, color = criterion)) +
  geom_boxplot() +
  facet_wrap(~bb.sp)

# Does bias vary with abundance?
ggplot(bias_sitedate, aes(log(bb.sp.abund), dprime, color = criterion)) +
  #geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp)

ggplot(bias_plantsp.sitedate, aes(log(bb.sp.prop.abund), dprime)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp)
```

## 1.3 Are some BB species more biased than others? (brms models)
```{r}
# Do species vary in d' (simple intercept model)
brm.plantsp <- brm(dprime ~ (1 | bb.sp), data = filter(bias.sitedate, criterion ==  "plant.sp"))
summary(brm.plantsp)
mcmc_plot(brm.plantsp, pars = c("^r_"))
mcmc_plot(brm.plantsp, pars = c("^b_"))
parnames(brm.plantsp)

brm.ktype <- brm(dprime ~ (1 | bb.sp), data = filter(bias.sitedate, criterion ==  "ktype"))
summary(brm.ktype)
mcmc_plot(brm.ktype, pars = c("^r_"))
mcmc_plot(brm.ktype, pars = c("^b_"))
parnames(brm.ktype)

brm.family <- brm(dprime ~ (1 | bb.sp), data = filter(bias.sitedate, criterion ==  "plant.family"))
summary(brm.family)
mcmc_plot(brm.family, pars = c("^r_"))
mcmc_plot(brm.family, pars = c("^b_"))
parnames(brm.family)
```

## 1.4 Does bias vary with criterion? (brms models)
```{r}
brm.criterion <- brm(dprime ~ (1 | criterion), data = bias.sitedate)
summary(brm.criterion)
mcmc_plot(brm.criterion, pars = c("^r_"))
mcmc_plot(brm.criterion, pars = c("^b_"))
parnames(brm.criterion)
```

## 1.5 Does bias vary with elevation? (brms models)
```{r}
# Does d' vary with elevation
brm2.plantsp <- brm(dprime ~ elev.mean + (1 | bb.sp), 
                    control = list(max_treedepth = 12),
                    data = filter(bias.sitedate, criterion ==  "plant.sp"))
summary(brm2.plantsp)
mcmc_plot(brm2.plantsp, pars = c("^r_"))
mcmc_plot(brm2.plantsp, pars = c("^b_"))
conditional_effects(brm2.plantsp)
parnames(brm2.plantsp)

brm2.ktype <- brm(dprime ~ elev.mean + (1 | bb.sp), 
                  control = list(max_treedepth = 12),
                  data = filter(bias.sitedate, criterion ==  "ktype"))
summary(brm2.ktype)
mcmc_plot(brm2.ktype, pars = c("^r_"))
mcmc_plot(brm2.ktype, pars = c("^b_"))
conditional_effects(brm2.ktype)
parnames(brm2.ktype)

brm2.family <- brm(dprime ~ elev.mean + (1 + elev.mean | bb.sp), 
                  control = list(max_treedepth = 10),
                  data = filter(bias.sitedate, criterion ==  "plant.family"))
summary(brm2.family)
mcmc_plot(brm2.family, pars = c("^r_"))
mcmc_plot(brm2.family, pars = c("^b_"))
conditional_effects(brm2.family)
parnames(brm2.family)

brm2.family <- brm(dprime ~ elev.mean, 
                  control = list(max_treedepth = 10),
                  data = filter(bias.sitedate, criterion ==  "plant.family"))
summary(brm2.family)
mcmc_plot(brm2.family, pars = c("^r_"))
mcmc_plot(brm2.family, pars = c("^b_"))
conditional_effects(brm2.family)
parnames(brm2.family)
```

## 1.6 Define hypotheses
```{r}
# Is d' different from 0?
H1 <- paste("Intercept = 0")

# Are some species more biased in their foraging than others? 
H.bss_gers <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[gers,Intercept]")
H.bss_hort <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[hort,Intercept]")
H.bss_humi <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[humi,Intercept]")
H.bss_hypn <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[hypn,Intercept]")
H.bss_jone <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[jone,Intercept]")
H.bss_lapi <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[lapi,Intercept]")
H.bss_mend <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[mend,Intercept]")
H.bss_mont <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[mont,Intercept]")
H.bss_muci <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[muci,Intercept]")
H.bss_pasc <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[pasc,Intercept]")
H.bss_prat <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[prat,Intercept]")
H.bss_psit <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[psit,Intercept]")
H.bss_pyre <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[pyre,Intercept]")
H.bss_soro <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[soro,Intercept]")
H.bss_wurf <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[wurf,Intercept]")

H.gers_hort <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[hort,Intercept]")
H.gers_humi <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[humi,Intercept]")
H.gers_hypn <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[hypn,Intercept]")
H.gers_jone <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[jone,Intercept]")
H.gers_lapi <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[lapi,Intercept]")
H.gers_mend <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[mend,Intercept]")
H.gers_mont <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[mont,Intercept]")
H.gers_muci <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[muci,Intercept]")
H.gers_pasc <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[pasc,Intercept]")
H.gers_prat <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[prat,Intercept]")
H.gers_psit <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[psit,Intercept]")
H.gers_pyre <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[pyre,Intercept]")
H.gers_soro <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[soro,Intercept]")
H.gers_wurf <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[wurf,Intercept]")

H.hort_humi <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[humi,Intercept]")
H.hort_hypn <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[hypn,Intercept]")
H.hort_jone <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[jone,Intercept]")
H.hort_lapi <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[lapi,Intercept]")
H.hort_mend <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[mend,Intercept]")
H.hort_mont <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[mont,Intercept]")
H.hort_muci <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[muci,Intercept]")
H.hort_pasc <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[pasc,Intercept]")
H.hort_prat <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[prat,Intercept]")
H.hort_psit <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[psit,Intercept]")
H.hort_pyre <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[pyre,Intercept]")
H.hort_soro <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[soro,Intercept]")
H.hort_wurf <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[wurf,Intercept]")

H.humi_hypn <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[hypn,Intercept]")
H.humi_jone <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[jone,Intercept]")
H.humi_lapi <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[lapi,Intercept]")
H.humi_mend <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[mend,Intercept]")
H.humi_mont <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[mont,Intercept]")
H.humi_muci <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[muci,Intercept]")
H.humi_pasc <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[pasc,Intercept]")
H.humi_prat <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[prat,Intercept]")
H.humi_psit <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[psit,Intercept]")
H.humi_pyre <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[pyre,Intercept]")
H.humi_soro <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[soro,Intercept]")
H.humi_wurf <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[wurf,Intercept]")

H.hypn_jone <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[jone,Intercept]")
H.hypn_lapi <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[lapi,Intercept]")
H.hypn_mend <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[mend,Intercept]")
H.hypn_mont <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[mont,Intercept]")
H.hypn_muci <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[muci,Intercept]")
H.hypn_pasc <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[pasc,Intercept]")
H.hypn_prat <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[prat,Intercept]")
H.hypn_psit <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[psit,Intercept]")
H.hypn_pyre <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[pyre,Intercept]")
H.hypn_soro <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[soro,Intercept]")
H.hypn_wurf <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[wurf,Intercept]")

H.jone_lapi <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[lapi,Intercept]")
H.jone_mend <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[mend,Intercept]")
H.jone_mont <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[mont,Intercept]")
H.jone_muci <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[muci,Intercept]")
H.jone_pasc <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[pasc,Intercept]")
H.jone_prat <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[prat,Intercept]")
H.jone_psit <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[psit,Intercept]")
H.jone_pyre <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[pyre,Intercept]")
H.jone_soro <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[soro,Intercept]")
H.jone_wurf <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[wurf,Intercept]")

H.lapi_mend <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[mend,Intercept]")
H.lapi_mont <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[mont,Intercept]")
H.lapi_muci <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[muci,Intercept]")
H.lapi_pasc <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[pasc,Intercept]")
H.lapi_prat <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[prat,Intercept]")
H.lapi_psit <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[psit,Intercept]")
H.lapi_pyre <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[pyre,Intercept]")
H.lapi_soro <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[soro,Intercept]")
H.lapi_wurf <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[wurf,Intercept]")

H.mend_mont <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[mont,Intercept]")
H.mend_muci <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[muci,Intercept]")
H.mend_pasc <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[pasc,Intercept]")
H.mend_prat <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[prat,Intercept]")
H.mend_psit <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[psit,Intercept]")
H.mend_pyre <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[pyre,Intercept]")
H.mend_soro <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[soro,Intercept]")
H.mend_wurf <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[wurf,Intercept]")

H.mont_muci <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[muci,Intercept]")
H.mont_pasc <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[pasc,Intercept]")
H.mont_prat <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[prat,Intercept]")
H.mont_psit <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[psit,Intercept]")
H.mont_pyre <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[pyre,Intercept]")
H.mont_soro <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[soro,Intercept]")
H.mont_wurf <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[wurf,Intercept]")

H.muci_pasc <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[pasc,Intercept]")
H.muci_prat <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[prat,Intercept]")
H.muci_psit <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[psit,Intercept]")
H.muci_pyre <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[pyre,Intercept]")
H.muci_soro <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[soro,Intercept]")
H.muci_wurf <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[wurf,Intercept]")

H.pasc_prat <- paste("r_bb.sp[pasc,Intercept] = r_bb.sp[prat,Intercept]")
H.pasc_psit <- paste("r_bb.sp[pasc,Intercept] = r_bb.sp[psit,Intercept]")
H.pasc_pyre <- paste("r_bb.sp[pasc,Intercept] = r_bb.sp[pyre,Intercept]")
H.pasc_soro <- paste("r_bb.sp[pasc,Intercept] = r_bb.sp[soro,Intercept]")
H.pasc_wurf <- paste("r_bb.sp[pasc,Intercept] = r_bb.sp[wurf,Intercept]")

H.prat_psit <- paste("r_bb.sp[prat,Intercept] = r_bb.sp[psit,Intercept]")
H.prat_pyre <- paste("r_bb.sp[prat,Intercept] = r_bb.sp[pyre,Intercept]")
H.prat_soro <- paste("r_bb.sp[prat,Intercept] = r_bb.sp[soro,Intercept]")
H.prat_wurf <- paste("r_bb.sp[prat,Intercept] = r_bb.sp[wurf,Intercept]")

H.psit_pyre <- paste("r_bb.sp[psit,Intercept] = r_bb.sp[pyre,Intercept]")
H.psit_soro <- paste("r_bb.sp[psit,Intercept] = r_bb.sp[soro,Intercept]")
H.psit_wurf <- paste("r_bb.sp[psit,Intercept] = r_bb.sp[wurf,Intercept]")

H.pyre_soro <- paste("r_bb.sp[pyre,Intercept] = r_bb.sp[soro,Intercept]")
H.pyre_wurf <- paste("r_bb.sp[pyre,Intercept] = r_bb.sp[wurf,Intercept]")

H.soro_wurf <- paste("r_bb.sp[soro,Intercept] = r_bb.sp[wurf,Intercept]")
```

## 1.7 Test contrasts
```{r}
### Plant.sp
hypothesis(brm.plantsp, H1)

hypothesis(brm.plantsp, H.bss_gers, class = "")
hypothesis(brm.plantsp, H.bss_hort, class = "")
hypothesis(brm.plantsp, H.bss_humi, class = "")
hypothesis(brm.plantsp, H.bss_hypn, class = "")
hypothesis(brm.plantsp, H.bss_jone, class = "")
hypothesis(brm.plantsp, H.bss_lapi, class = "")
hypothesis(brm.plantsp, H.bss_mend, class = "")
hypothesis(brm.plantsp, H.bss_mont, class = "")
hypothesis(brm.plantsp, H.bss_muci, class = "")
hypothesis(brm.plantsp, H.bss_pasc, class = "")
hypothesis(brm.plantsp, H.bss_prat, class = "")
hypothesis(brm.plantsp, H.bss_psit, class = "")
hypothesis(brm.plantsp, H.bss_pyre, class = "")
hypothesis(brm.plantsp, H.bss_soro, class = "")
hypothesis(brm.plantsp, H.bss_wurf, class = "")

hypothesis(brm.plantsp, H.gers_hort, class = "")
hypothesis(brm.plantsp, H.gers_humi, class = "")
hypothesis(brm.plantsp, H.gers_hypn, class = "")
hypothesis(brm.plantsp, H.gers_jone, class = "")
hypothesis(brm.plantsp, H.gers_lapi, class = "")
hypothesis(brm.plantsp, H.gers_mend, class = "")
hypothesis(brm.plantsp, H.gers_mont, class = "")
hypothesis(brm.plantsp, H.gers_muci, class = "")
hypothesis(brm.plantsp, H.gers_pasc, class = "")
hypothesis(brm.plantsp, H.gers_prat, class = "")
hypothesis(brm.plantsp, H.gers_psit, class = "")
hypothesis(brm.plantsp, H.gers_pyre, class = "")
hypothesis(brm.plantsp, H.gers_soro, class = "")
hypothesis(brm.plantsp, H.gers_wurf, class = "")

hypothesis(brm.plantsp, H.hort_humi, class = "")
hypothesis(brm.plantsp, H.hort_hypn, class = "")
hypothesis(brm.plantsp, H.hort_jone, class = "")
hypothesis(brm.plantsp, H.hort_lapi, class = "")
hypothesis(brm.plantsp, H.hort_mend, class = "")
hypothesis(brm.plantsp, H.hort_mont, class = "")
hypothesis(brm.plantsp, H.hort_muci, class = "")
hypothesis(brm.plantsp, H.hort_pasc, class = "")
hypothesis(brm.plantsp, H.hort_prat, class = "")
hypothesis(brm.plantsp, H.hort_psit, class = "")
hypothesis(brm.plantsp, H.hort_pyre, class = "")
hypothesis(brm.plantsp, H.hort_soro, class = "")
hypothesis(brm.plantsp, H.hort_wurf, class = "")

hypothesis(brm.plantsp, H.humi_hypn, class = "")
hypothesis(brm.plantsp, H.humi_jone, class = "")
hypothesis(brm.plantsp, H.humi_lapi, class = "")
hypothesis(brm.plantsp, H.humi_mend, class = "")
hypothesis(brm.plantsp, H.humi_mont, class = "")
hypothesis(brm.plantsp, H.humi_muci, class = "")
hypothesis(brm.plantsp, H.humi_pasc, class = "")
hypothesis(brm.plantsp, H.humi_prat, class = "")
hypothesis(brm.plantsp, H.humi_psit, class = "")
hypothesis(brm.plantsp, H.humi_pyre, class = "")
hypothesis(brm.plantsp, H.humi_soro, class = "")
hypothesis(brm.plantsp, H.humi_wurf, class = "")

hypothesis(brm.plantsp, H.hypn_jone, class = "")
hypothesis(brm.plantsp, H.hypn_lapi, class = "")
hypothesis(brm.plantsp, H.hypn_mend, class = "")
hypothesis(brm.plantsp, H.hypn_mont, class = "")
hypothesis(brm.plantsp, H.hypn_muci, class = "")
hypothesis(brm.plantsp, H.hypn_pasc, class = "")
hypothesis(brm.plantsp, H.hypn_prat, class = "")
hypothesis(brm.plantsp, H.hypn_psit, class = "")
hypothesis(brm.plantsp, H.hypn_pyre, class = "")
hypothesis(brm.plantsp, H.hypn_soro, class = "")
hypothesis(brm.plantsp, H.hypn_wurf, class = "")

hypothesis(brm.plantsp, H.jone_lapi, class = "")
hypothesis(brm.plantsp, H.jone_mend, class = "")
hypothesis(brm.plantsp, H.jone_mont, class = "")
hypothesis(brm.plantsp, H.jone_muci, class = "")
hypothesis(brm.plantsp, H.jone_pasc, class = "")
hypothesis(brm.plantsp, H.jone_prat, class = "")
hypothesis(brm.plantsp, H.jone_psit, class = "")
hypothesis(brm.plantsp, H.jone_pyre, class = "")
hypothesis(brm.plantsp, H.jone_soro, class = "")
hypothesis(brm.plantsp, H.jone_wurf, class = "")

hypothesis(brm.plantsp, H.lapi_mend, class = "")
hypothesis(brm.plantsp, H.lapi_mont, class = "")
hypothesis(brm.plantsp, H.lapi_muci, class = "")
hypothesis(brm.plantsp, H.lapi_pasc, class = "")
hypothesis(brm.plantsp, H.lapi_prat, class = "")
hypothesis(brm.plantsp, H.lapi_psit, class = "")
hypothesis(brm.plantsp, H.lapi_pyre, class = "")
hypothesis(brm.plantsp, H.lapi_soro, class = "")
hypothesis(brm.plantsp, H.lapi_wurf, class = "")

hypothesis(brm.plantsp, H.mend_mont, class = "")
hypothesis(brm.plantsp, H.mend_muci, class = "")
hypothesis(brm.plantsp, H.mend_pasc, class = "")
hypothesis(brm.plantsp, H.mend_prat, class = "")
hypothesis(brm.plantsp, H.mend_psit, class = "")
hypothesis(brm.plantsp, H.mend_pyre, class = "")
hypothesis(brm.plantsp, H.mend_soro, class = "")
hypothesis(brm.plantsp, H.mend_wurf, class = "")

hypothesis(brm.plantsp, H.mont_muci, class = "")
hypothesis(brm.plantsp, H.mont_pasc, class = "")
hypothesis(brm.plantsp, H.mont_prat, class = "")
hypothesis(brm.plantsp, H.mont_psit, class = "")
hypothesis(brm.plantsp, H.mont_pyre, class = "")
hypothesis(brm.plantsp, H.mont_soro, class = "")
hypothesis(brm.plantsp, H.mont_wurf, class = "")

hypothesis(brm.plantsp, H.muci_pasc, class = "")
hypothesis(brm.plantsp, H.muci_prat, class = "")
hypothesis(brm.plantsp, H.muci_psit, class = "")
hypothesis(brm.plantsp, H.muci_pyre, class = "")
hypothesis(brm.plantsp, H.muci_soro, class = "")
hypothesis(brm.plantsp, H.muci_wurf, class = "")

hypothesis(brm.plantsp, H.pasc_prat, class = "")
hypothesis(brm.plantsp, H.pasc_psit, class = "")
hypothesis(brm.plantsp, H.pasc_pyre, class = "")
hypothesis(brm.plantsp, H.pasc_soro, class = "")
hypothesis(brm.plantsp, H.pasc_wurf, class = "")

hypothesis(brm.plantsp, H.prat_psit, class = "")
hypothesis(brm.plantsp, H.prat_pyre, class = "")
hypothesis(brm.plantsp, H.prat_soro, class = "")
hypothesis(brm.plantsp, H.prat_wurf, class = "")

hypothesis(brm.plantsp, H.psit_pyre, class = "")
hypothesis(brm.plantsp, H.psit_soro, class = "")
hypothesis(brm.plantsp, H.psit_wurf, class = "")

hypothesis(brm.plantsp, H.pyre_soro, class = "")
hypothesis(brm.plantsp, H.pyre_wurf, class = "")

hypothesis(brm.plantsp, H.soro_wurf, class = "")

### Kugler type
hypothesis(brm.ktype, H1)

hypothesis(brm.ktype, H.bss_gers, class = "")
hypothesis(brm.ktype, H.bss_hort, class = "")
hypothesis(brm.ktype, H.bss_humi, class = "")
hypothesis(brm.ktype, H.bss_hypn, class = "")
hypothesis(brm.ktype, H.bss_jone, class = "")
hypothesis(brm.ktype, H.bss_lapi, class = "")
hypothesis(brm.ktype, H.bss_mend, class = "")
hypothesis(brm.ktype, H.bss_mont, class = "")
hypothesis(brm.ktype, H.bss_muci, class = "")
hypothesis(brm.ktype, H.bss_pasc, class = "")
hypothesis(brm.ktype, H.bss_prat, class = "")
hypothesis(brm.ktype, H.bss_psit, class = "")
hypothesis(brm.ktype, H.bss_pyre, class = "")
hypothesis(brm.ktype, H.bss_soro, class = "")
hypothesis(brm.ktype, H.bss_wurf, class = "")

hypothesis(brm.ktype, H.gers_hort, class = "")
hypothesis(brm.ktype, H.gers_humi, class = "")
hypothesis(brm.ktype, H.gers_hypn, class = "")
hypothesis(brm.ktype, H.gers_jone, class = "")
hypothesis(brm.ktype, H.gers_lapi, class = "")
hypothesis(brm.ktype, H.gers_mend, class = "")
hypothesis(brm.ktype, H.gers_mont, class = "")
hypothesis(brm.ktype, H.gers_muci, class = "")
hypothesis(brm.ktype, H.gers_pasc, class = "")
hypothesis(brm.ktype, H.gers_prat, class = "")
hypothesis(brm.ktype, H.gers_psit, class = "")
hypothesis(brm.ktype, H.gers_pyre, class = "")
hypothesis(brm.ktype, H.gers_soro, class = "")
hypothesis(brm.ktype, H.gers_wurf, class = "")

hypothesis(brm.ktype, H.hort_humi, class = "")
hypothesis(brm.ktype, H.hort_hypn, class = "")
hypothesis(brm.ktype, H.hort_jone, class = "")
hypothesis(brm.ktype, H.hort_lapi, class = "")
hypothesis(brm.ktype, H.hort_mend, class = "")
hypothesis(brm.ktype, H.hort_mont, class = "")
hypothesis(brm.ktype, H.hort_muci, class = "")
hypothesis(brm.ktype, H.hort_pasc, class = "")
hypothesis(brm.ktype, H.hort_prat, class = "")
hypothesis(brm.ktype, H.hort_psit, class = "")
hypothesis(brm.ktype, H.hort_pyre, class = "")
hypothesis(brm.ktype, H.hort_soro, class = "")
hypothesis(brm.ktype, H.hort_wurf, class = "")

hypothesis(brm.ktype, H.humi_hypn, class = "")
hypothesis(brm.ktype, H.humi_jone, class = "")
hypothesis(brm.ktype, H.humi_lapi, class = "")
hypothesis(brm.ktype, H.humi_mend, class = "")
hypothesis(brm.ktype, H.humi_mont, class = "")
hypothesis(brm.ktype, H.humi_muci, class = "")
hypothesis(brm.ktype, H.humi_pasc, class = "")
hypothesis(brm.ktype, H.humi_prat, class = "")
hypothesis(brm.ktype, H.humi_psit, class = "")
hypothesis(brm.ktype, H.humi_pyre, class = "")
hypothesis(brm.ktype, H.humi_soro, class = "")
hypothesis(brm.ktype, H.humi_wurf, class = "")

hypothesis(brm.ktype, H.hypn_jone, class = "")
hypothesis(brm.ktype, H.hypn_lapi, class = "")
hypothesis(brm.ktype, H.hypn_mend, class = "")
hypothesis(brm.ktype, H.hypn_mont, class = "")
hypothesis(brm.ktype, H.hypn_muci, class = "")
hypothesis(brm.ktype, H.hypn_pasc, class = "")
hypothesis(brm.ktype, H.hypn_prat, class = "")
hypothesis(brm.ktype, H.hypn_psit, class = "")
hypothesis(brm.ktype, H.hypn_pyre, class = "")
hypothesis(brm.ktype, H.hypn_soro, class = "")
hypothesis(brm.ktype, H.hypn_wurf, class = "")

hypothesis(brm.ktype, H.jone_lapi, class = "")
hypothesis(brm.ktype, H.jone_mend, class = "")
hypothesis(brm.ktype, H.jone_mont, class = "")
hypothesis(brm.ktype, H.jone_muci, class = "")
hypothesis(brm.ktype, H.jone_pasc, class = "")
hypothesis(brm.ktype, H.jone_prat, class = "")
hypothesis(brm.ktype, H.jone_psit, class = "")
hypothesis(brm.ktype, H.jone_pyre, class = "")
hypothesis(brm.ktype, H.jone_soro, class = "")
hypothesis(brm.ktype, H.jone_wurf, class = "")

hypothesis(brm.ktype, H.lapi_mend, class = "")
hypothesis(brm.ktype, H.lapi_mont, class = "")
hypothesis(brm.ktype, H.lapi_muci, class = "")
hypothesis(brm.ktype, H.lapi_pasc, class = "")
hypothesis(brm.ktype, H.lapi_prat, class = "")
hypothesis(brm.ktype, H.lapi_psit, class = "")
hypothesis(brm.ktype, H.lapi_pyre, class = "")
hypothesis(brm.ktype, H.lapi_soro, class = "")
hypothesis(brm.ktype, H.lapi_wurf, class = "")

hypothesis(brm.ktype, H.mend_mont, class = "")
hypothesis(brm.ktype, H.mend_muci, class = "")
hypothesis(brm.ktype, H.mend_pasc, class = "")
hypothesis(brm.ktype, H.mend_prat, class = "")
hypothesis(brm.ktype, H.mend_psit, class = "")
hypothesis(brm.ktype, H.mend_pyre, class = "")
hypothesis(brm.ktype, H.mend_soro, class = "")
hypothesis(brm.ktype, H.mend_wurf, class = "")

hypothesis(brm.ktype, H.mont_muci, class = "")
hypothesis(brm.ktype, H.mont_pasc, class = "")
hypothesis(brm.ktype, H.mont_prat, class = "")
hypothesis(brm.ktype, H.mont_psit, class = "")
hypothesis(brm.ktype, H.mont_pyre, class = "")
hypothesis(brm.ktype, H.mont_soro, class = "")
hypothesis(brm.ktype, H.mont_wurf, class = "")

hypothesis(brm.ktype, H.muci_pasc, class = "")
hypothesis(brm.ktype, H.muci_prat, class = "")
hypothesis(brm.ktype, H.muci_psit, class = "")
hypothesis(brm.ktype, H.muci_pyre, class = "")
hypothesis(brm.ktype, H.muci_soro, class = "")
hypothesis(brm.ktype, H.muci_wurf, class = "")

hypothesis(brm.ktype, H.pasc_prat, class = "")
hypothesis(brm.ktype, H.pasc_psit, class = "")
hypothesis(brm.ktype, H.pasc_pyre, class = "")
hypothesis(brm.ktype, H.pasc_soro, class = "")
hypothesis(brm.ktype, H.pasc_wurf, class = "")

hypothesis(brm.ktype, H.prat_psit, class = "")
hypothesis(brm.ktype, H.prat_pyre, class = "")
hypothesis(brm.ktype, H.prat_soro, class = "")
hypothesis(brm.ktype, H.prat_wurf, class = "")

hypothesis(brm.ktype, H.psit_pyre, class = "")
hypothesis(brm.ktype, H.psit_soro, class = "")
hypothesis(brm.ktype, H.psit_wurf, class = "")

hypothesis(brm.ktype, H.pyre_soro, class = "")
hypothesis(brm.ktype, H.pyre_wurf, class = "")

hypothesis(brm.ktype, H.soro_wurf, class = "")

### Plant family
hypothesis(brm.family, H1)

hypothesis(brm.family, H.bss_gers, class = "")
hypothesis(brm.family, H.bss_hort, class = "")
hypothesis(brm.family, H.bss_humi, class = "")
hypothesis(brm.family, H.bss_hypn, class = "")
hypothesis(brm.family, H.bss_jone, class = "")
hypothesis(brm.family, H.bss_lapi, class = "")
hypothesis(brm.family, H.bss_mend, class = "")
hypothesis(brm.family, H.bss_mont, class = "")
hypothesis(brm.family, H.bss_muci, class = "")
hypothesis(brm.family, H.bss_pasc, class = "")
hypothesis(brm.family, H.bss_prat, class = "")
hypothesis(brm.family, H.bss_psit, class = "")
hypothesis(brm.family, H.bss_pyre, class = "")
hypothesis(brm.family, H.bss_soro, class = "")
hypothesis(brm.family, H.bss_wurf, class = "")

hypothesis(brm.family, H.gers_hort, class = "")
hypothesis(brm.family, H.gers_humi, class = "")
hypothesis(brm.family, H.gers_hypn, class = "")
hypothesis(brm.family, H.gers_jone, class = "")
hypothesis(brm.family, H.gers_lapi, class = "")
hypothesis(brm.family, H.gers_mend, class = "")
hypothesis(brm.family, H.gers_mont, class = "")
hypothesis(brm.family, H.gers_muci, class = "")
hypothesis(brm.family, H.gers_pasc, class = "")
hypothesis(brm.family, H.gers_prat, class = "")
hypothesis(brm.family, H.gers_psit, class = "")
hypothesis(brm.family, H.gers_pyre, class = "")
hypothesis(brm.family, H.gers_soro, class = "")
hypothesis(brm.family, H.gers_wurf, class = "")

hypothesis(brm.family, H.hort_humi, class = "")
hypothesis(brm.family, H.hort_hypn, class = "")
hypothesis(brm.family, H.hort_jone, class = "")
hypothesis(brm.family, H.hort_lapi, class = "")
hypothesis(brm.family, H.hort_mend, class = "")
hypothesis(brm.family, H.hort_mont, class = "")
hypothesis(brm.family, H.hort_muci, class = "")
hypothesis(brm.family, H.hort_pasc, class = "")
hypothesis(brm.family, H.hort_prat, class = "")
hypothesis(brm.family, H.hort_psit, class = "")
hypothesis(brm.family, H.hort_pyre, class = "")
hypothesis(brm.family, H.hort_soro, class = "")
hypothesis(brm.family, H.hort_wurf, class = "")

hypothesis(brm.family, H.humi_hypn, class = "")
hypothesis(brm.family, H.humi_jone, class = "")
hypothesis(brm.family, H.humi_lapi, class = "")
hypothesis(brm.family, H.humi_mend, class = "")
hypothesis(brm.family, H.humi_mont, class = "")
hypothesis(brm.family, H.humi_muci, class = "")
hypothesis(brm.family, H.humi_pasc, class = "")
hypothesis(brm.family, H.humi_prat, class = "")
hypothesis(brm.family, H.humi_psit, class = "")
hypothesis(brm.family, H.humi_pyre, class = "")
hypothesis(brm.family, H.humi_soro, class = "")
hypothesis(brm.family, H.humi_wurf, class = "")

hypothesis(brm.family, H.hypn_jone, class = "")
hypothesis(brm.family, H.hypn_lapi, class = "")
hypothesis(brm.family, H.hypn_mend, class = "")
hypothesis(brm.family, H.hypn_mont, class = "")
hypothesis(brm.family, H.hypn_muci, class = "")
hypothesis(brm.family, H.hypn_pasc, class = "")
hypothesis(brm.family, H.hypn_prat, class = "")
hypothesis(brm.family, H.hypn_psit, class = "")
hypothesis(brm.family, H.hypn_pyre, class = "")
hypothesis(brm.family, H.hypn_soro, class = "")
hypothesis(brm.family, H.hypn_wurf, class = "")

hypothesis(brm.family, H.jone_lapi, class = "")
hypothesis(brm.family, H.jone_mend, class = "")
hypothesis(brm.family, H.jone_mont, class = "")
hypothesis(brm.family, H.jone_muci, class = "")
hypothesis(brm.family, H.jone_pasc, class = "")
hypothesis(brm.family, H.jone_prat, class = "")
hypothesis(brm.family, H.jone_psit, class = "")
hypothesis(brm.family, H.jone_pyre, class = "")
hypothesis(brm.family, H.jone_soro, class = "")
hypothesis(brm.family, H.jone_wurf, class = "")

hypothesis(brm.family, H.lapi_mend, class = "")
hypothesis(brm.family, H.lapi_mont, class = "")
hypothesis(brm.family, H.lapi_muci, class = "")
hypothesis(brm.family, H.lapi_pasc, class = "")
hypothesis(brm.family, H.lapi_prat, class = "")
hypothesis(brm.family, H.lapi_psit, class = "")
hypothesis(brm.family, H.lapi_pyre, class = "")
hypothesis(brm.family, H.lapi_soro, class = "")
hypothesis(brm.family, H.lapi_wurf, class = "")

hypothesis(brm.family, H.mend_mont, class = "")
hypothesis(brm.family, H.mend_muci, class = "")
hypothesis(brm.family, H.mend_pasc, class = "")
hypothesis(brm.family, H.mend_prat, class = "")
hypothesis(brm.family, H.mend_psit, class = "")
hypothesis(brm.family, H.mend_pyre, class = "")
hypothesis(brm.family, H.mend_soro, class = "")
hypothesis(brm.family, H.mend_wurf, class = "")

hypothesis(brm.family, H.mont_muci, class = "")
hypothesis(brm.family, H.mont_pasc, class = "")
hypothesis(brm.family, H.mont_prat, class = "")
hypothesis(brm.family, H.mont_psit, class = "")
hypothesis(brm.family, H.mont_pyre, class = "")
hypothesis(brm.family, H.mont_soro, class = "")
hypothesis(brm.family, H.mont_wurf, class = "")

hypothesis(brm.family, H.muci_pasc, class = "")
hypothesis(brm.family, H.muci_prat, class = "")
hypothesis(brm.family, H.muci_psit, class = "")
hypothesis(brm.family, H.muci_pyre, class = "")
hypothesis(brm.family, H.muci_soro, class = "")
hypothesis(brm.family, H.muci_wurf, class = "")

hypothesis(brm.family, H.pasc_prat, class = "")
hypothesis(brm.family, H.pasc_psit, class = "")
hypothesis(brm.family, H.pasc_pyre, class = "")
hypothesis(brm.family, H.pasc_soro, class = "")
hypothesis(brm.family, H.pasc_wurf, class = "")

hypothesis(brm.family, H.prat_psit, class = "")
hypothesis(brm.family, H.prat_pyre, class = "")
hypothesis(brm.family, H.prat_soro, class = "")
hypothesis(brm.family, H.prat_wurf, class = "")

hypothesis(brm.family, H.psit_pyre, class = "")
hypothesis(brm.family, H.psit_soro, class = "")
hypothesis(brm.family, H.psit_wurf, class = "")

hypothesis(brm.family, H.pyre_soro, class = "")
hypothesis(brm.family, H.pyre_wurf, class = "")

hypothesis(brm.family, H.soro_wurf, class = "")
```

# 2. Having established the bias of BB foraging, let's zoom in on which morphotypes and families were over/under-represented
## 2.1 Calculate representation
```{r}
representation.ktype <- visit_cover %>%
  mutate(k.type.ss = factor(k.type.ss)) %>%
  group_by(site, date, bb.sp, k.type.ss) %>%
  summarize(flower.cover = sum(flower.cover),
            visits = sum(visits)) %>%
  group_by(site, date, bb.sp) %>%
  mutate(prop.flower.cover = flower.cover/sum(flower.cover),
         prop.visits = visits/sum(visits)) %>%
  mutate(representation = prop.visits - prop.flower.cover,
         representation.exp = if_else(visits == 0, 0, exp(prop.visits - prop.flower.cover)),
         representation.bin = if_else(representation > 0, 1, 0)) %>%
  left_join(site_data) %>%
  mutate(yday = yday(date)) %>%
  select(site, elev.mean, date, yday, bb.sp, k.type.ss, representation, representation.bin)

representation.family <- visit_cover %>%
  mutate(plant.family = factor(plant.family)) %>%
  group_by(site, date, bb.sp, plant.family) %>%
  summarize(flower.cover = sum(flower.cover),
            visits = sum(visits)) %>%
  group_by(site, date, bb.sp) %>%
  mutate(prop.flower.cover = flower.cover/sum(flower.cover),
         prop.visits = visits/sum(visits)) %>%
  mutate(representation = prop.visits - prop.flower.cover,
         representation.exp = if_else(visits == 0, 0, exp(prop.visits - prop.flower.cover)),
         representation.bin = if_else(representation > 0, 1, 0)) %>%
  left_join(site_data) %>%
  mutate(yday = yday(date)) %>%
  select(site, elev.mean, date, yday, bb.sp, plant.family, representation, representation.bin)
```

## 2.2 Visualize
```{r}
ggplot(representation.ktype, aes(representation)) +
  geom_histogram()

ggplot(representation.ktype, aes(representation.bin)) +
  geom_bar()

ggplot(representation.ktype, aes(k.type.ss, representation)) +
  geom_boxplot() +
  #geom_hline(yintercept = 1, linetype = "dashed")+
  facet_wrap(~bb.sp)

ggplot(representation.ktype, aes(k.type.ss, fill = representation.bin)) +
  geom_bar() +
  facet_wrap(~bb.sp)

ggplot(representation.ktype, aes(elev.mean, representation)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~k.type.ss)

ggplot(representation.family, aes(plant.family, representation)) +
  geom_boxplot() +
  facet_wrap(~bb.sp)

ggplot(representation.family, aes(elev.mean, representation)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~plant.family)

ggplot(representation.family, aes(yday, exp(representation))) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 3), se = FALSE) +
  facet_wrap(~plant.family)
```

## 2.3 Explaining representation by floral traits, BB spp, and elevation (and maybe time)
```{r}
repmod.ktype <- brm(representation ~ (1 | k.type.ss), 
                    data = representation.ktype)
mcmc_plot(repmod.ktype)
conditional_effects(repmod.ktype)

repmod.ktype2 <- update(repmod.ktype, formula. = representation ~ elev.mean + (1 + elev.mean | k.type.ss))
```

# 3. Can we explain bias by BB-Fl trait associations? (RLQ + fourth corner analysis)
## 3.1
## 3.2 Do BB-FL trait associations vary with elevation?

# 4. Can we explain BB-FL trait associations in terms of floral reward or handling time? We can use Neumayer and Paulus' (1999) data. 
## 4.1 
## 4.2 Do there appear to be suboptimalities in BB-FL trait associations? That is, are trait classes over/under-represented relative to their reward/efficiency? If so, how might explain such suboptimalities?  
  - Innate bias trumps learning?
  - Ecological modulators are involved, e.g. competition or distribution (see Sowig 1989)

### 2020-12-18 data_proc.R

@@ -22,6 +22,7 @@ fl_survey <- read_delim("./raw_data/data_plantcover_forDoug_2020.01.17.csv", del
          day = day(temp.date),
          month = month(temp.date),
          date = make_date(year, month, day)) %>%
+  filter(plant.genus != "Huperzia") %>% # because I don't need the clubmoss
   dplyr::select(-temp.date)
   
 write_csv(fl_survey, "./processed_data/floral_survey.csv")

### 2021-01-11 analyses/visitation_bias.Rmd
Started new manucript

@@
 library(brms)
+library(lme4)
+#library(inla)
 library(bayesplot)
+library(vegan)
 #library(tidybayes)
@@
 library(lubridate)
+library(tidymodels)
 ```
@@
   left_join(fl_traits) %>% # add floral traits
-  ungroup()
+  ungroup() %>%
+  mutate(k.type.ss = factor(k.type.ss))
 
@@
   mutate(bb.sp.prop.abund = bb.sp.abund/sum(bb.sp.abund))
+
+fl_abundance <- visit_cover %>%
+  group_by(site, date) %>%
+  summarize(total.flower.cover = sum(flower.cover))
 ```
@@
 
-# 1. Is bumble bee foraging biased? How do taxonomic, phylogenetic, and morphological bias compare? Are some BB species more biased than others?
+## And one for niche overlap
+```{r}
+# Niche overlap by sp.pair * site * year
+niche_overlapper <- function(x, criterion) {
+
+  criterion <- enquo(criterion) # criterion is set to either plant.sp (for taxonomic) or k.type.s (for morphological)
+
+  no1 <- x %>%
+    group_by(bb.sp, !!criterion, site, date) %>% # group by site-year and BB-FL pair
+    summarize(visits = sum(visits)) %>%
+    group_by(bb.sp, site, date) %>% # group by site-year and BB species
+    mutate(prop.visits = visits/sum(visits)) %>% # get proportional interaction frequency for each BB-FL pair
+    select(site, date, bb.sp, !!criterion, prop.visits) %>%
+    pivot_wider(names_from = !!criterion, values_from = prop.visits, values_fill = 0) %>% # spread into the kind of array vegan likes
+    group_by(site, date) %>%
+    nest() %>% # nest into list-column
+    mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
+           dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
+           tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
+    dplyr::select(site, date, tidy_dist) %>% # drop intermediate list-columns
+    unnest(cols = c(tidy_dist)) %>%
+    rename(bb.sp1 = item1, bb.sp2 = item2) %>%
+    mutate(niche.overlap = 1-distance,  # express niche overlap so that 1 = perfect overlap and 0 = no overlap
+           bss = if_else(bb.sp1 == "bss" | bb.sp2 == "bss", TRUE, FALSE),
+           pasc = if_else(bb.sp1 == "pasc" | bb.sp2 == "pasc", TRUE, FALSE),
+           wurf = if_else(bb.sp1 == "wurf" | bb.sp2 == "wurf", TRUE, FALSE),
+           prat = if_else(bb.sp1 == "prat" | bb.sp2 == "prat", TRUE, FALSE),
+           soro = if_else(bb.sp1 == "soro" | bb.sp2 == "soro", TRUE, FALSE),
+           humi = if_else(bb.sp1 == "humi" | bb.sp2 == "humi", TRUE, FALSE),
+           muci = if_else(bb.sp1 == "muci" | bb.sp2 == "muci", TRUE, FALSE),
+           pyre = if_else(bb.sp1 == "pyre" | bb.sp2 == "pyre", TRUE, FALSE),
+           mont = if_else(bb.sp1 == "mont" | bb.sp2 == "mont", TRUE, FALSE),
+           mend = if_else(bb.sp1 == "mend" | bb.sp2 == "mend", TRUE, FALSE),
+           psit = if_else(bb.sp1 == "psit" | bb.sp2 == "psit", TRUE, FALSE),
+           gers = if_else(bb.sp1 == "gers" | bb.sp2 == "gers", TRUE, FALSE),
+           jone = if_else(bb.sp1 == "jone" | bb.sp2 == "jone", TRUE, FALSE),
+           hypn = if_else(bb.sp1 == "hypn" | bb.sp2 == "hypn", TRUE, FALSE),
+           lapi = if_else(bb.sp1 == "lapi" | bb.sp2 == "lapi", TRUE, FALSE),
+           hort = if_else(bb.sp1 == "hort" | bb.sp2 == "hort", TRUE, FALSE))
+
+  no <- list(
+
+  no_bss <- no1 %>%
+    group_by(site, date, bss) %>% 
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%
+    mutate(bb.sp = rep("bss", n())) %>%
+    filter(bss == TRUE),
+
+  no_pasc <- no1 %>%
+    group_by(site, date, pasc) %>%
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%    
+    mutate(bb.sp = rep("pasc", n())) %>%
+    filter(pasc == TRUE),
+
+  no_wurf <- no1 %>%
+    group_by(site, date, wurf) %>%
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%    
+    mutate(bb.sp = rep("wurf", n())) %>%
+    filter(wurf == TRUE),
+
+  no_prat <- no1 %>%
+    group_by(site, date, prat) %>%
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%   
+    mutate(bb.sp = rep("prat", n())) %>%
+    filter(prat == TRUE),
+
+  no_soro <- no1 %>%
+    group_by(site, date, soro) %>% 
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%    
+    mutate(bb.sp = rep("soro", n())) %>%
+    filter(soro == TRUE),
+
+  no_humi <- no1 %>%
+    group_by(site, date, humi) %>%
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%    
+    mutate(bb.sp = rep("humi", n())) %>%
+    filter(humi == TRUE),
+
+  no_muci <- no1 %>%
+    group_by(site, date, muci) %>%
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%    
+    mutate(bb.sp = rep("muci", n())) %>%
+    filter(muci == TRUE),
+
+  no_pyre <- no1 %>%
+    group_by(site, date, pyre) %>%
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%   
+    mutate(bb.sp = rep("pyre", n())) %>%
+    filter(pyre == TRUE),
+
+  no_mont <- no1 %>%
+    group_by(site, date, mont) %>% 
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%    
+    mutate(bb.sp = rep("mont", n())) %>%
+    filter(mont == TRUE),
+
+  no_mend <- no1 %>%
+    group_by(site, date, mend) %>%
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%    
+    mutate(bb.sp = rep("mend", n())) %>%
+    filter(mend == TRUE),
+
+  no_psit <- no1 %>%
+    group_by(site, date, psit) %>%
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%    
+    mutate(bb.sp = rep("psit", n())) %>%
+    filter(psit == TRUE),
+
+  no_gers <- no1 %>%
+    group_by(site, date, gers) %>%
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%    
+    mutate(bb.sp = rep("gers", n())) %>%
+    filter(gers == TRUE),
+
+  no_jone <- no1 %>%
+    group_by(site, date, jone) %>% 
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%    
+    mutate(bb.sp = rep("jone", n())) %>%
+    filter(jone == TRUE),
+
+  no_hypn <- no1 %>%
+    group_by(site, date, hypn) %>%
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%    
+    mutate(bb.sp = rep("hypn", n())) %>%
+    filter(hypn == TRUE),
+
+  no_lapi <- no1 %>%
+    group_by(site, date, lapi) %>%
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%    
+    mutate(bb.sp = rep("lapi", n())) %>%
+    filter(lapi == TRUE),
+
+  no_hort <- no1 %>%
+    group_by(site, date, hort) %>%
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%    
+    mutate(bb.sp = rep("hort", n())) %>%
+    filter(hort == TRUE)
+  ) %>%
+
+    bind_rows() %>%
+    select(site, date, bb.sp, mean.niche.overlap, max.niche.overlap)
+  }
+```
+
+# 1.
+Is bumble bee foraging biased?
+How do taxonomic, phylogenetic, and morphological bias compare -- NO, can't analyze this, these should be considered separate repsonse variables and analyzed in separate models?
+Are some BB species more biased than others?
 ## 1.1 Calculate Blüthgen's d' (i.e. visitation bias)
@@
 # No grouping -- just plant.sp
+no_plant.sp <- niche_overlapper(visit_cover, plant.sp)
+
 bias_plantsp.sitedate <- biaseval(visit_cover, plant.sp, site, date) %>%
@@
   left_join(bb_abundance) %>% # add BB abundance data
-  mutate(criterion = factor(rep("plant.sp", n())))
+  left_join(fl_abundance) %>% # add fl abundance
+  mutate(criterion = factor(rep("plant.sp", n()))) %>%
+  left_join(no_plant.sp)
 
 # Grouped by k.type.ss
+no_ktype <- niche_overlapper(visit_cover, k.type.ss)
+
 bias_ktype.sitedate <- biaseval(visit_cover, k.type.ss, site, date) %>%
@@
   left_join(bb_abundance) %>% # add BB abundance data
-  mutate(criterion = factor(rep("ktype", n())))
+  left_join(fl_abundance) %>% # add fl abundance
+  mutate(criterion = factor(rep("ktype", n()))) %>%
+  left_join(no_ktype)
 
 # Grouped by plant.family
+no_plant.family <- niche_overlapper(visit_cover, plant.family)
+
 bias_family.sitedate <- biaseval(visit_cover, plant.family, site, date) %>%
@@
   left_join(bb_abundance) %>% # add BB abundance data
-  mutate(criterion = factor(rep("plant.family", n())))
+  left_join(fl_abundance) %>% # add fl abundance
+  mutate(criterion = factor(rep("plant.family", n()))) %>%
+  left_join(no_plant.family)
 
@@
                            bias_ktype.sitedate, 
-                           bias_family.sitedate)
+                           bias_family.sitedate) %>%
+  mutate(bb.sp = factor(bb.sp),
+         yday = yday(date),
+         elev = elev.mean/10) %>%
+  left_join(no)
+
+# When bias is high, it could mean that there are few good options on which all (or many) BB species converge.
+But apparently it doesn't.
+When bias is high, BB spp are biased toward DIFFERENT flora, and thus their niches diverge.
+ggplot(bias_sitedate, aes(dprime, mean.niche.overlap, color = criterion)) +
+  geom_point(alpha = 0.1) +
+  geom_smooth(method = "lm") +
+  #geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
+  facet_wrap(~bb.sp) +
+  ylim(c(0,1))
+
+ggplot(bias_sitedate, aes(dprime, max.niche.overlap, color = criterion)) +
+  geom_point(alpha = 0.1) +
+  geom_smooth(method = "lm") +
+  #geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
+  facet_wrap(~bb.sp) +
+  ylim(c(0,1))
+
+ggplot(bias_sitedate, aes(log(total.flower.cover), dprime, color = criterion)) +
+  geom_point(alpha = 0.1) +
+  geom_smooth(method = "lm") +
+  #geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
+  facet_wrap(~bb.sp) +
+  ylim(c(0,1))
+
+ggplot(bias_sitedate, aes(log(total.flower.cover), mean.niche.overlap, color = criterion)) +
+  geom_point(alpha = 0.1) +
+  geom_smooth(method = "lm") +
+  #geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
+  facet_wrap(~bb.sp) +
+  ylim(c(0,1))
+
+brm.bias_niche <- brm(mean.niche.overlap ~ 1 + dprime + (1 | bb.sp),
+                     cores = 3,
+                     data = filter(bias_sitedate, criterion == "ktype"))
+
+parnames(brm.bias_niche)
+conditional_effects(brm.bias_nich)
+hypothesis(brm.bias_niche, "dprime = 0")
+pp_check(brm.bias_niche)
 ```
@@
 
+ggplot(bias_sitedate, aes(yday, dprime, color = criterion)) +
+  geom_point(alpha = 0.25) +
+  geom_smooth(se = TRUE)
+
 ggplot(bias_sitedate, aes(elev.mean, dprime, color = criterion)) +
+  geom_point(alpha = 0.25) +
+  geom_smooth(se = FALSE, method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~bb.sp)
+
+ggplot(bias_sitedate, aes(yday, dprime, color = criterion)) +
   geom_point(alpha = 0.25) +
@@
 
+ggplot(bias_sitedate, aes(bb.sp.abund, dprime, color = criterion)) +
+  #geom_point() +
+  geom_smooth() +
+  facet_wrap(~bb.sp, )
+
+ggplot(bias_sitedate, aes(log(bb.sp.abund), dprime, color = criterion)) +
+  geom_point(alpha = 0.1) +
+  geom_smooth()
+
 ggplot(bias_plantsp.sitedate, aes(log(bb.sp.prop.abund), dprime)) +
@@
 
-## 1.3 Are some BB species more biased than others? (brms models)
+## 1.3 Multi-level model: bias ~ elevation + (1 + elevation | bb.sp)
 ```{r}
-# Do species vary in d' (simple intercept model)
-brm.plantsp <- brm(dprime ~ (1 | bb.sp), data = filter(bias.sitedate, criterion ==  "plant.sp"))
-summary(brm.plantsp)
-mcmc_plot(brm.plantsp, pars = c("^r_"))
-mcmc_plot(brm.plantsp, pars = c("^b_"))
-parnames(brm.plantsp)
-
-brm.ktype <- brm(dprime ~ (1 | bb.sp), data = filter(bias.sitedate, criterion ==  "ktype"))
-summary(brm.ktype)
-mcmc_plot(brm.ktype, pars = c("^r_"))
-mcmc_plot(brm.ktype, pars = c("^b_"))
-parnames(brm.ktype)
-
-brm.family <- brm(dprime ~ (1 | bb.sp), data = filter(bias.sitedate, criterion ==  "plant.family"))
-summary(brm.family)
-mcmc_plot(brm.family, pars = c("^r_"))
-mcmc_plot(brm.family, pars = c("^b_"))
-parnames(brm.family)
-```
-
-## 1.4 Does bias vary with criterion? (brms models)
-```{r}
-brm.criterion <- brm(dprime ~ (1 | criterion), data = bias.sitedate)
-summary(brm.criterion)
-mcmc_plot(brm.criterion, pars = c("^r_"))
-mcmc_plot(brm.criterion, pars = c("^b_"))
-parnames(brm.criterion)
-```
-
-## 1.5 Does bias vary with elevation? (brms models)
-```{r}
-# Does d' vary with elevation
-brm2.plantsp <- brm(dprime ~ elev.mean + (1 | bb.sp), 
-                    control = list(max_treedepth = 12),
-                    data = filter(bias.sitedate, criterion ==  "plant.sp"))
-summary(brm2.plantsp)
-mcmc_plot(brm2.plantsp, pars = c("^r_"))
-mcmc_plot(brm2.plantsp, pars = c("^b_"))
-conditional_effects(brm2.plantsp)
-parnames(brm2.plantsp)
-
-brm2.ktype <- brm(dprime ~ elev.mean + (1 | bb.sp), 
-                  control = list(max_treedepth = 12),
-                  data = filter(bias.sitedate, criterion ==  "ktype"))
-summary(brm2.ktype)
-mcmc_plot(brm2.ktype, pars = c("^r_"))
-mcmc_plot(brm2.ktype, pars = c("^b_"))
-conditional_effects(brm2.ktype)
-parnames(brm2.ktype)
-
-brm2.family <- brm(dprime ~ elev.mean + (1 + elev.mean | bb.sp), 
+# brms
+brm.ktype1 <- brm(dprime ~ yday + elev + (1 + elev | bb.sp), 
                   control = list(max_treedepth = 10),
-                  data = filter(bias.sitedate, criterion ==  "plant.family"))
-summary(brm2.family)
-mcmc_plot(brm2.family, pars = c("^r_"))
-mcmc_plot(brm2.family, pars = c("^b_"))
-conditional_effects(brm2.family)
-parnames(brm2.family)
-
-brm2.family <- brm(dprime ~ elev.mean, 
+                  cores = 3,
+                  iter = 2000,
+                  file = "./output/brm_ktype1",
+                  data = filter(bias_sitedate, criterion ==  "ktype"))
+
+summary(brm.ktype1)
+plot(brm.ktype1)
+parnames(brm.ktype1)
+mcmc_plot(brm.ktype1, pars = c("^r_"))
+mcmc_plot(brm.ktype1, pars = c("^b_"))
+mcmc_plot(brm.ktype1)
+conditional_effects(brm.ktype1)
+pp_check(brm.ktype1)
+
+brm.ktype2 <- brm(dprime ~ 1 + yday + elev + (1 | bb.sp), 
                   control = list(max_treedepth = 10),
-                  data = filter(bias.sitedate, criterion ==  "plant.family"))
-summary(brm2.family)
-mcmc_plot(brm2.family, pars = c("^r_"))
-mcmc_plot(brm2.family, pars = c("^b_"))
-conditional_effects(brm2.family)
-parnames(brm2.family)
+                  cores = 4,
+                  iter = 2000,
+                  file = "./output/brm_ktype2b",
+                  data = filter(bias_sitedate, criterion ==  "ktype"))
+
+summary(brm.ktype2)
+plot(brm.ktype2)
+parnames(brm.ktype2)
+mcmc_plot(brm.ktype2, pars = c("^r_"))
+mcmc_plot(brm.ktype2, pars = c("^b_"))
+mcmc_plot(brm.ktype2)
+conditional_effects(brm.ktype2)
+pp_check(brm.ktype2)
+
+brm.ktype3 <- brm(dprime ~ 1 + (1 | bb.sp), 
+                  control = list(max_treedepth = 10),
+                  cores = 4,
+                  iter = 2000,
+                  family = "poisson",
+                  file = "./output/brm_ktype3",
+                  data = filter(bias_sitedate, criterion ==  "ktype"))
+
+summary(brm.ktype3)
+plot(brm.ktype3)
+parnames(brm.ktype3)
+mcmc_plot(brm.ktype3, pars = c("^r_"))
+mcmc_plot(brm.ktype3, pars = c("^b_"))
+mcmc_plot(brm.ktype3)
+conditional_effects(brm.ktype3)
+pp_check(brm.ktype3)
+
+# lme4
+library(lme4)
+library(ggeffects)
+
+lmer.ktype <- lmer(dprime ~ 1 + elev:yday + (1 | bb.sp), 
+                  data = filter(bias_sitedate, criterion ==  "ktype"))
+summary(lmer.ktype)
+pred.ktype_yday <- ggpredict(lmer.ktype, "yday")
+pred.ktype_elev <- ggpredict(lmer.ktype, "elev")
+#pred.ktype_yday_ran <- ggpredict(lmer.ktype, "yday", type = "random")
+#pred.ktype_elev_ran <- ggpredict(lmer.ktype, "elev", type = "random")
+pred.ktype_bbsp <- ggpredict(lmer.ktype, "bb.sp", type = "random")
+
+plot(pred.ktype_yday)
+plot(pred.ktype_elev)
+#plot(pred.ktype_yday_ran)
+#plot(pred.ktype_elev_ran)
+plot(pred.ktype_bbsp)
+
+# mgcv
+gam.ktype <- bam(dprime ~ s(bb.sp, bs = "re") + 
+                   t2(yday, elev.mean, bs = c("tp", "gp")),
+                 method = "fREML",
+                 discrete = TRUE,
+                 data = filter(bias_sitedate, criterion ==  "ktype")) %>% getViz()
+
+summary(gam.ktype)
+plot(gam.ktype, allTerms = TRUE)
+
+# inla
+library(INLA)
+
+inla.ktype <- inla(dprime ~ elev.mean, 
+                   family = "gaussian",
+                   data = filter(bias_sitedate, criterion ==  "ktype"))
+
+inla.pen <- inla(dprime ~ 1 + elev.mean,
+  data = filter(bias_sitedate, criterion ==  "ktype"), control.predictor = list(compute = TRUE))
+summary(inla.pen)
 ```
@@
 # 2. Having established the bias of BB foraging, let's zoom in on which morphotypes and families were over/under-represented
-## 2.1 Calculate representation
+## 2.1 Calculate bias per plant speces/ktype/family
+
+Deciding exactly how to represent bias turns out to be rather tricky. I am trying out a few options right now.
+
+1.
+Bias could  be trated simply as the Euclidean difference between prop.flower.cover and prop.visits.
+This has the effect of weighting bias by flower.cover, particularly negative bias, since an abundant plant can have a greater negative bias (because it is farther from 0) than a rare plant.
+Euclidean bias is bounded between -1 and 1.
+Its symmetry is a nice property that sets it apart from proportional or fold-change bias.
+If, for whatever reason, I decided that a log-transformation of this measure were desirable, I think I could simply add a constant of 1 to all values, which would move the boundary up to (0, 2).
+I think the biggest problem with measuring bias in this way is that it makes no distinction between a plant that was visited a little and one that was visited not at all, and that could be an important biological difference.
+For example, consider two plants, one with a proportional abundance = 0.01 and proportional visitation = 0.01, and the second with proportional abundance = 0.01 and proportional visitation = 0.00.
+These will look identical in our analysis, but it is possible that the unvisited plant would not have been visited no matter how high its abundance was.
+**For starters, though, I am going to proceed with this metric because it is the simplest.
+And because DJ thought it was a good idea.**
+
+2.
+Alternatively, bias could be calculated as a simple proportion: prop.visits/prop.flower.cover.
+If anything, this might tend to reverse the weighting, since a small Euclidean change in a rare species would be a large proportional change.
+Proportional change has a lower boundary of 0 and no upper boundary.
+This approach highlights cases of positive bias, but it is even worse than euclidean bias in terms of obscuring the distinction between plants visited slightly and those visited not at all.
+Worse in a sense.
+In another sense, it is better than euclidean bias, because with proportional visitation a plant visited 0 times is at least numerical distinguishable from one visited 1 time (visited plants will always have non-zero proportional bias).
+
+3.
+Similarly to proportional bias, I could calculate bias as fold change.
+This converges to be virtually identical to proportional change as bias becomes more positive, but fold change has a lower bound of -1 instead of 0.
+
+4.
+If the distinction between plants that were visited and those that were not is what we're most interested in, we could simply treat bias as binary.
+This, of course, comes at the price of not being able to distinguish a plant that received 0.01 proportional visitation and one that recevied  0.99 proportional visitation, and this distinction is arguably important.
+
+5.
+An interesting option is to first separate visited from unvisited plants, set the bias of unvisited plants to 0, and then exponentiate the euclidean bias of all visited plant species.
+This has some desirable properties.
+First, it clearly distinguished visited plants from unvisited ones.
+Second, it assigns plants visited perfectly in proportion to their abundance a sensible bias of 1 (just like proportional bias).
+Thus, plants with a psotive bias are > 1 (bounded at 2.718282 -- okay, so that's a bit inelegant), those with a negative bias are < 1 (bounded at 0.367895), and those not visited at all = 0.
+
+6.
+The problem with the option above is that is adds weight to positive bias.
+Another approach is to separate the unvisited plants from the visited, set unvisited bias = 0, and visited bias = prop.visits - prop.flower.cover + 1.
+The +1 serves to make the distribution positive, with 1 represening perfect proportionality between visitation and floral abundance.
+This creates a symmetrically bounded distribution between 0 (not inclusive) and 2 (inclusive), plus the true zeros from the unvisited plants.
+It feels a bit...slapped together, though.
+
 ```{r}
-representation.ktype <- visit_cover %>%
-  mutate(k.type.ss = factor(k.type.ss)) %>%
-  group_by(site, date, bb.sp, k.type.ss) %>%
-  summarize(flower.cover = sum(flower.cover),
+# Function for calculating bias per group (= criterion)
+biaseval_group <- function(x, criterion) {
+
+  criterion <- enquo(criterion)
+
+  x %>%
+    mutate(criterion = factor(!!criterion)) %>% # having this as a factor will help with modeling later on
+    group_by(site, date, bb.sp, criterion) %>%
+    summarize(flower.cover = sum(flower.cover), # get raw floer cover and visits by group
               visits = sum(visits)) %>%
     group_by(site, date, bb.sp) %>%
-  mutate(prop.flower.cover = flower.cover/sum(flower.cover),
+    mutate(prop.flower.cover = flower.cover/sum(flower.cover), # get proportional flower cover and visits by group
            prop.visits = visits/sum(visits)) %>%
-  mutate(representation = prop.visits - prop.flower.cover,
-         representation.exp = if_else(visits == 0, 0, exp(prop.visits - prop.flower.cover)),
-         representation.bin = if_else(representation > 0, 1, 0)) %>%
+    mutate(bias = prop.visits - prop.flower.cover, # bias as difference
+           bias.prop = prop.visits/prop.flower.cover, # bias as proportion
+           bias.fold = bias/prop.flower.cover, # bias as fold change
+           bias.bin = if_else(bias > 0, 1, 0), # bias as binary
+           bias.exp = if_else(visits == 0, 0, exp(prop.visits - prop.flower.cover)),  # bias as exponentiated difference
+           bias.split = if_else(visits == 0, 0, (prop.visits - prop.flower.cover + 1))) %>%
+    mutate(yday = yday(date)) %>%
+    select(site, date, yday, bb.sp, criterion, 
+           bias, bias.prop, bias.fold, bias.bin, bias.exp, bias.split,
+           visits, flower.cover, prop.visits, prop.flower.cover)
+}
+
+bias_group_plantsp <- biaseval_group(visit_cover, plant.sp) %>%
   left_join(site_data) %>%
-  mutate(yday = yday(date)) %>%
-  select(site, elev.mean, date, yday, bb.sp, k.type.ss, representation, representation.bin)
-
-representation.family <- visit_cover %>%
-  mutate(plant.family = factor(plant.family)) %>%
-  group_by(site, date, bb.sp, plant.family) %>%
-  summarize(flower.cover = sum(flower.cover),
-            visits = sum(visits)) %>%
-  group_by(site, date, bb.sp) %>%
-  mutate(prop.flower.cover = flower.cover/sum(flower.cover),
-         prop.visits = visits/sum(visits)) %>%
-  mutate(representation = prop.visits - prop.flower.cover,
-         representation.exp = if_else(visits == 0, 0, exp(prop.visits - prop.flower.cover)),
-         representation.bin = if_else(representation > 0, 1, 0)) %>%
+  rename(group = criterion) %>%
+  mutate(criterion = rep("plant.sp", n()))
+
+bias_group_ktype <- biaseval_group(visit_cover, k.type.ss) %>%
   left_join(site_data) %>%
-  mutate(yday = yday(date)) %>%
-  select(site, elev.mean, date, yday, bb.sp, plant.family, representation, representation.bin)
+  rename(group = criterion) %>%
+  mutate(criterion = rep("ktype", n()))
+
+bias_group_family <- biaseval_group(visit_cover, plant.family) %>%
+  left_join(site_data) %>%
+  rename(group = criterion) %>%
+  mutate(criterion = rep("plant.family", n()))
+
+bias_group <- bind_rows(bias_group_plantsp, bias_group_ktype, bias_group_family)
+
+ggplot(bias_group, aes(prop.flower.cover, prop.visits)) +
+  geom_point(alpha = 0.1) +
+  #geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
+  geom_smooth(method = "lm") +
+  geom_abline(slope = 1, color = "red", linetype = "dashed") +
+  facet_wrap(~criterion)
+
+ggplot(bias_group, aes(log(prop.flower.cover), prop.visits)) +
+  geom_point(alpha = 0.1) +
+  geom_smooth() +
+  facet_wrap(~criterion)
+
+ggplot(filter(bias_group, criterion == "ktype"), aes(prop.flower.cover, prop.visits)) +
+  geom_point(alpha = 0.1) +
+  geom_smooth() +
+  geom_abline(slope = 1, color = "red", linetype = "dashed") +
+  facet_wrap(~group)
 ```
@@
 ```{r}
-ggplot(representation.ktype, aes(representation)) +
+# Check the distribution of bias
+ggplot(bias_group_ktype, aes(bias)) +
   geom_histogram()
 
-ggplot(representation.ktype, aes(representation.bin)) +
-  geom_bar()
-
-ggplot(representation.ktype, aes(k.type.ss, representation)) +
-  geom_boxplot() +
-  #geom_hline(yintercept = 1, linetype = "dashed")+
-  facet_wrap(~bb.sp)
-
-ggplot(representation.ktype, aes(k.type.ss, fill = representation.bin)) +
-  geom_bar() +
-  facet_wrap(~bb.sp)
-
-ggplot(representation.ktype, aes(elev.mean, representation)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
-  facet_wrap(~k.type.ss)
-
-ggplot(representation.family, aes(plant.family, representation)) +
+ggplot(bias_group_ktype, aes(bias)) +
+  geom_histogram() +
+  facet_wrap(~criterion)
+
+ggplot(bias_group_family, aes(bias)) +
+  geom_histogram()
+
+ggplot(bias_group_family, aes(bias)) +
+  geom_histogram() +
+  facet_wrap(~criterion)
+
+# Bias by Kugler morphotype
+ggplot(bias_ktype, aes(criterion, bias)) +
+  geom_boxplot()
+
+ggplot(bias_ktype, aes(criterion, bias)) +
   geom_boxplot() +
@@
 
-ggplot(representation.family, aes(elev.mean, representation)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
-  facet_wrap(~plant.family)
-
-ggplot(representation.family, aes(yday, exp(representation))) +
-  geom_point(alpha = 0.1) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 3), se = FALSE) +
-  facet_wrap(~plant.family)
+ggplot(bias_ktype, aes(bb.sp, bias)) +
+  geom_boxplot() +
+  facet_wrap(~criterion)
+
+ggplot(bias_ktype, aes(elev.mean, bias)) +
+  #geom_point() +
+  geom_smooth(method = gam, formula = y ~ s(x, k = 5)) +
+  facet_wrap(~criterion)
+
+# Bias by plant family
+ggplot(bias_family, aes(criterion, bias)) +
+  geom_boxplot()
+
+ggplot(bias_family, aes(criterion, bias)) +
+  geom_boxplot() +
+  facet_wrap(~bb.sp)
+
+ggplot(bias_family, aes(elev.mean, bias)) +
+  #geom_point() +
+  geom_smooth(method = gam, formula = y ~ s(x, k = 5)) +
+  facet_wrap(~criterion)
+
+ggplot(bias_family, aes(elev.mean, bias)) +
+  #geom_point() +
+  geom_smooth(method = gam, formula = y ~ s(x, k = 5)) +
+  facet_wrap(~criterion)
+
 ```
@@
 ```{r}
-repmod.ktype <- brm(representation ~ (1 | k.type.ss), 
-                    data = representation.ktype)
+repmod.ktype <- brm(bias ~ elev.mean + (1 | criterion), 
+                    data = bias_ktype)
+
+repmod.ktype <- brm(bias ~ criterion, 
+                    data = bias_ktype)
+parnames(repmod.ktype)
 mcmc_plot(repmod.ktype)
@@
 
-repmod.ktype2 <- update(repmod.ktype, formula. = representation ~ elev.mean + (1 + elev.mean | k.type.ss))
+H.K5_K6 <- paste("b_criterion5 = b_criterion6")
+hypothesis(repmod.ktype, H.K5_K6, class = "")
+```
+
+# How do species vary in their biases
+```{r}
+bias_matrix.plantsp <- bias_group_plantsp %>%
+  group_by(bb.sp, criterion) %>%
+  summarize(bias = mean(bias)) %>%
+  pivot_wider(values_from = bias, names_from = criterion, values_fill = 0) %>%
+  column_to_rownames("bb.sp")
+
+bias_matrix.ktype <- bias_group_ktype %>%
+  group_by(bb.sp, criterion) %>%
+  summarize(bias = mean(bias)) %>%
+  pivot_wider(values_from = bias, names_from = criterion, values_fill = 0) %>%
+  column_to_rownames("bb.sp")
+
+library(vegan)
+
+pca.plantsp <- rda(bias_matrix.plantsp)
+plot(pca.plantsp, type = "t")
+
+pca.ktype <- rda(bias_matrix.ktype)
+plot(pca.ktype)
+
 ```

### 2021-01-11 writing/manuscript_v4.Rmd
Started new manucript

---
title: Bumble bee foraging bias
author:
- Douglas B. Sponsler
- Fabrice Requier
- Katharina Kallnik
- Alice Claßen
- A. Fabienne Maihoff
- Ingolf Steffan-Dewenter 
bibliography: refs.bib
output: html_document
---

# Introduction

Plant-pollinator networks are characterized by unequal interaction strengths among partners. While the distribution of interaction strengths can be partially explained by the neutral effects of abundance, pollinators exhibit positive and negative biases toward available flora even when interaction strength is conditioned on floral abundance. This phenomenon is formalized by the discrimination (d’) metric of [@Bluthgen2006-if], but to avoid conflation with the metric *per se*, we will use the term bias in a more general sense to describe disproportionality between pollinator visitation and floral abundance.

Taken as a focal phenomenon, the realized floral biases of pollinators can be interposed between their underlying causal mechanisms and their higher-order significance **(Figure 1)**. From the causal perspective, realized biases can be partitioned into innate and learned components. Both these components depend on interactions between the functional traits of a given pollinator and those of its potential floral hosts, manifest over ecological time scales with respect to learned biases and evolutionary time scales with respect to innate biases. From the vantage of higher-order effects, the floral biases of pollinators contribute to the complex structures observed in plant-pollinator networks, which have consequences for both the ecological function of the network and the fitness of its constituent species. These effects can, in turn, feedback on the causes of bias via their modulation of community composition and biotic interactions.

Bumble bees (Hymenoptera: *Bombus* spp.) are an especially interesting group in which to study floral biases together with their causal mechanisms and downstream effects. On the one hand, bumble bees are exceptionally versatile foragers, and their ability to solve the puzzles posed by complex flower morphologies [@Plowright1984-lc], or indeed by human investigators [@Alem2016-rn; @Loukola2017-bt], renders almost any flower potentially accessible. Even when a gross mismatch between tongue length and corolla depth makes the nectar of a given flower inaccessible by ordinary means, some bumble bee species will resort readily to biting through the base of the flower to “rob” nectar, incidentally making the nectar available to other prospective foragers as well. Flower handling skills are, moreover, socially transmitted, amplifying the plasticity of individual behavior [@Leadbeater2008-rh; @Loukola2017-bt]. Nevertheless, the flexibility of bumble bee foraging is tuned to floral reward, which depends on both the intrinsic reward of a given flower and the searching/learning/handling costs associated with exploiting it. At an evolutionary level, the innate preferences of bumble bees for certain floral traits likely reflect historic associations between floral traits and floral rewards [@Raine2007-nq]. These innate preferences, however, can be modulated or even nullified by learned patterns of reward based on individual foraging experience [@Kugler1943-uy; Maharaj2019-wm], which depend not only intrinsic floral properties but also ecological modulators, such as the spatial distribution of floral patches or competition with other bumble bees [@Ishii2013-sm; @Plowright1984-lc; @Sowig1989-ld]. Thus, realized biases in bumble bee foraging are the labile product of bumble bee traits, floral traits, and ecological context. The influence of these biases on interaction network structure has consequences both for plant pollination and bumble bee species coexistence, and these consequences potentially feed back into the causes of bumble bee foraging bias via the modulation of competition between bumble bees and ultimately via the regulation of community assembly.  <!-- strengthen this sentence with some refs? -->

*mountain slopes give us floristic variation and abiotic variation while controlling for genetic variation within BB species, thus enabling us to study the lability of bias and partition it between innate and learned components*

The study bias in bumble bee foraging

Local elevation gradients along steep mountain slopes provide the opportunity to study broad environmental variation while controlling for potential confounding effects associated with geographic distance.

# Sandbox

constrained by the learning and handling costs associated with available flower morphologies, and these costs are determined principally by the morphological fittedness between bumble bee tongue length and floral corolla depth. Short-tongued bumble bee species exploit deep flowers less efficiently than their long-tongued counterparts (robbing notwithstanding), and long-tongued species are less efficient than short-tongued species when foraging on shallow flowers [@Inouye1980-mv].

# References

### 2021-01-11 analyses/visitation_bias.Rmd
syncing to office desktop, got rid of lots of old scripts and output

@@
 
-## And one for niche overlap
+# 1.
+Is bumble bee foraging biased?
+How do taxonomic, phylogenetic, and morphological bias compare -- NO, can't analyze this, these should be considered separate repsonse variables and analyzed in separate models?
+Are some BB species more biased than others?
+## 1.1 Calculate Blüthgen's d' (i.e. visitation bias)
+```{r}
+# No grouping -- just plant.sp
+bias_plantsp <- biaseval(visit_cover, plant.sp, site, date) %>%
+  left_join(site_data) %>% # add site data
+  left_join(bb_abundance) %>% # add BB abundance data
+  left_join(fl_abundance) %>% # add fl abundance
+  mutate(criterion = factor(rep("plant.sp", n()))) %>%
+  left_join(no_plant.sp)
+
+# Grouped by k.type.ss
+bias_ktype <- biaseval(visit_cover, k.type.ss, site, date) %>%
+  left_join(site_data) %>%
+  left_join(bb_abundance) %>% # add BB abundance data
+  left_join(fl_abundance) %>% # add fl abundance
+  mutate(criterion = factor(rep("ktype", n()))) %>%
+  left_join(no_ktype)
+
+# Grouped by plant.family
+bias_family <- biaseval(visit_cover, plant.family, site, date) %>%
+  left_join(site_data) %>%
+  left_join(bb_abundance) %>% # add BB abundance data
+  left_join(fl_abundance) %>% # add fl abundance
+  mutate(criterion = factor(rep("plant.family", n()))) %>%
+  left_join(no_plant.family)
+
+# Collated
+bias <- bind_rows(bias_plantsp.sitedate, 
+                  bias_ktype.sitedate, 
+                  bias_family.sitedate) %>%
+  mutate(bb.sp = factor(bb.sp),
+         yday = yday(date),
+         elev = elev.mean/10) 
+```
+
+## 1.2 Visualize
+```{r}
+# Compare mean bias by criterion
+ggplot(bias, aes(criterion, dprime)) +
+  geom_boxplot()
+
+# Bias ~ elevation * criterion
+ggplot(bias, aes(elev.mean, dprime, color = criterion)) +
+  geom_point(alpha = 0.25) +
+  geom_smooth(se = TRUE)
+
+ggplot(bias, aes(yday, dprime, color = criterion)) +
+  geom_point(alpha = 0.25) +
+  geom_smooth(se = TRUE)
+
+ggplot(bias, aes(elev.mean, dprime, color = criterion)) +
+  geom_point(alpha = 0.25) +
+  geom_smooth(se = FALSE, method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~bb.sp)
+
+ggplot(bias, aes(yday, dprime, color = criterion)) +
+  geom_point(alpha = 0.25) +
+  geom_smooth(se = FALSE, method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~bb.sp)
+
+# Bias ~ year; Blüthgen's d' seems very robust to sampling effects, since it does not vary across years
+ggplot(bias, aes(year, dprime, color = criterion)) +
+  geom_boxplot()
+
+ggplot(bias, aes(year, dprime, color = criterion)) +
+  geom_boxplot() +
+  facet_wrap(~bb.sp)
+
+# Does bias vary with abundance?
+ggplot(bias, aes(log(bb.sp.abund), dprime, color = criterion)) +
+  #geom_point() +
+  geom_smooth() +
+  facet_wrap(~bb.sp)
+
+ggplot(bias, aes(bb.sp.abund, dprime, color = criterion)) +
+  #geom_point() +
+  geom_smooth() +
+  facet_wrap(~bb.sp, )
+
+ggplot(bias, aes(log(bb.sp.abund), dprime, color = criterion)) +
+  geom_point(alpha = 0.1) +
+  geom_smooth()
+
+ggplot(bias_plantsp.sitedate, aes(log(bb.sp.prop.abund), dprime)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~bb.sp)
+```
+
+## 1.3 Multi-level model: bias ~ elevation + (1 + elevation | bb.sp)
+```{r}
+# brms
+brm.ktype1 <- brm(dprime ~ yday + elev + (1 + elev | bb.sp), 
+                  control = list(max_treedepth = 10),
+                  cores = 3,
+                  iter = 2000,
+                  file = "./output/brm_ktype1",
+                  data = filter(bias, criterion ==  "ktype"))
+
+summary(brm.ktype1)
+plot(brm.ktype1)
+parnames(brm.ktype1)
+mcmc_plot(brm.ktype1, pars = c("^r_"))
+mcmc_plot(brm.ktype1, pars = c("^b_"))
+mcmc_plot(brm.ktype1)
+conditional_effects(brm.ktype1)
+pp_check(brm.ktype1)
+
+brm.ktype2 <- brm(dprime ~ 1 + yday + elev + (1 | bb.sp), 
+                  control = list(max_treedepth = 10),
+                  cores = 4,
+                  iter = 2000,
+                  file = "./output/brm_ktype2b",
+                  data = filter(bias, criterion ==  "ktype"))
+
+summary(brm.ktype2)
+plot(brm.ktype2)
+parnames(brm.ktype2)
+mcmc_plot(brm.ktype2, pars = c("^r_"))
+mcmc_plot(brm.ktype2, pars = c("^b_"))
+mcmc_plot(brm.ktype2)
+conditional_effects(brm.ktype2)
+pp_check(brm.ktype2)
+
+brm.ktype3 <- brm(dprime ~ 1 + (1 | bb.sp), 
+                  control = list(max_treedepth = 10),
+                  cores = 4,
+                  iter = 2000,
+                  family = "poisson",
+                  file = "./output/brm_ktype3",
+                  data = filter(bias, criterion ==  "ktype"))
+
+summary(brm.ktype3)
+plot(brm.ktype3)
+parnames(brm.ktype3)
+mcmc_plot(brm.ktype3, pars = c("^r_"))
+mcmc_plot(brm.ktype3, pars = c("^b_"))
+mcmc_plot(brm.ktype3)
+conditional_effects(brm.ktype3)
+pp_check(brm.ktype3)
+
+# lme4
+library(lme4)
+library(ggeffects)
+
+lmer.ktype <- lmer(dprime ~ 1 + elev:yday + (1 | bb.sp), 
+                  data = filter(bias, criterion ==  "ktype"))
+summary(lmer.ktype)
+pred.ktype_yday <- ggpredict(lmer.ktype, "yday")
+pred.ktype_elev <- ggpredict(lmer.ktype, "elev")
+#pred.ktype_yday_ran <- ggpredict(lmer.ktype, "yday", type = "random")
+#pred.ktype_elev_ran <- ggpredict(lmer.ktype, "elev", type = "random")
+pred.ktype_bbsp <- ggpredict(lmer.ktype, "bb.sp", type = "random")
+
+plot(pred.ktype_yday)
+plot(pred.ktype_elev)
+#plot(pred.ktype_yday_ran)
+#plot(pred.ktype_elev_ran)
+plot(pred.ktype_bbsp)
+
+# mgcv
+gam.ktype <- bam(dprime ~ s(bb.sp, bs = "re") + 
+                   t2(yday, elev.mean, bs = c("tp", "gp")),
+                 method = "fREML",
+                 discrete = TRUE,
+                 data = filter(bias, criterion ==  "ktype")) %>% getViz()
+
+summary(gam.ktype)
+plot(gam.ktype, allTerms = TRUE)
+
+# inla
+library(INLA)
+
+inla.ktype <- inla(dprime ~ elev.mean, 
+                   family = "gaussian",
+                   data = filter(bias, criterion ==  "ktype"))
+
+inla.pen <- inla(dprime ~ 1 + elev.mean,
+  data = filter(bias, criterion ==  "ktype"), control.predictor = list(compute = TRUE))
+summary(inla.pen)
+```
+
+## 1.4 Define hypotheses
+```{r}
+# Is d' different from 0?
+H1 <- paste("Intercept = 0")
+
+# Are some species more biased in their foraging than others? 
+H.bss_gers <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[gers,Intercept]")
+H.bss_hort <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[hort,Intercept]")
+H.bss_humi <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[humi,Intercept]")
+H.bss_hypn <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[hypn,Intercept]")
+H.bss_jone <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[jone,Intercept]")
+H.bss_lapi <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[lapi,Intercept]")
+H.bss_mend <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[mend,Intercept]")
+H.bss_mont <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[mont,Intercept]")
+H.bss_muci <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[muci,Intercept]")
+H.bss_pasc <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[pasc,Intercept]")
+H.bss_prat <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[prat,Intercept]")
+H.bss_psit <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[psit,Intercept]")
+H.bss_pyre <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[pyre,Intercept]")
+H.bss_soro <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[soro,Intercept]")
+H.bss_wurf <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[wurf,Intercept]")
+
+H.gers_hort <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[hort,Intercept]")
+H.gers_humi <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[humi,Intercept]")
+H.gers_hypn <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[hypn,Intercept]")
+H.gers_jone <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[jone,Intercept]")
+H.gers_lapi <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[lapi,Intercept]")
+H.gers_mend <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[mend,Intercept]")
+H.gers_mont <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[mont,Intercept]")
+H.gers_muci <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[muci,Intercept]")
+H.gers_pasc <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[pasc,Intercept]")
+H.gers_prat <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[prat,Intercept]")
+H.gers_psit <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[psit,Intercept]")
+H.gers_pyre <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[pyre,Intercept]")
+H.gers_soro <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[soro,Intercept]")
+H.gers_wurf <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[wurf,Intercept]")
+
+H.hort_humi <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[humi,Intercept]")
+H.hort_hypn <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[hypn,Intercept]")
+H.hort_jone <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[jone,Intercept]")
+H.hort_lapi <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[lapi,Intercept]")
+H.hort_mend <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[mend,Intercept]")
+H.hort_mont <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[mont,Intercept]")
+H.hort_muci <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[muci,Intercept]")
+H.hort_pasc <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[pasc,Intercept]")
+H.hort_prat <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[prat,Intercept]")
+H.hort_psit <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[psit,Intercept]")
+H.hort_pyre <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[pyre,Intercept]")
+H.hort_soro <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[soro,Intercept]")
+H.hort_wurf <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[wurf,Intercept]")
+
+H.humi_hypn <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[hypn,Intercept]")
+H.humi_jone <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[jone,Intercept]")
+H.humi_lapi <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[lapi,Intercept]")
+H.humi_mend <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[mend,Intercept]")
+H.humi_mont <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[mont,Intercept]")
+H.humi_muci <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[muci,Intercept]")
+H.humi_pasc <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[pasc,Intercept]")
+H.humi_prat <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[prat,Intercept]")
+H.humi_psit <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[psit,Intercept]")
+H.humi_pyre <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[pyre,Intercept]")
+H.humi_soro <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[soro,Intercept]")
+H.humi_wurf <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[wurf,Intercept]")
+
+H.hypn_jone <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[jone,Intercept]")
+H.hypn_lapi <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[lapi,Intercept]")
+H.hypn_mend <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[mend,Intercept]")
+H.hypn_mont <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[mont,Intercept]")
+H.hypn_muci <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[muci,Intercept]")
+H.hypn_pasc <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[pasc,Intercept]")
+H.hypn_prat <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[prat,Intercept]")
+H.hypn_psit <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[psit,Intercept]")
+H.hypn_pyre <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[pyre,Intercept]")
+H.hypn_soro <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[soro,Intercept]")
+H.hypn_wurf <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[wurf,Intercept]")
+
+H.jone_lapi <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[lapi,Intercept]")
+H.jone_mend <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[mend,Intercept]")
+H.jone_mont <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[mont,Intercept]")
+H.jone_muci <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[muci,Intercept]")
+H.jone_pasc <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[pasc,Intercept]")
+H.jone_prat <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[prat,Intercept]")
+H.jone_psit <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[psit,Intercept]")
+H.jone_pyre <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[pyre,Intercept]")
+H.jone_soro <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[soro,Intercept]")
+H.jone_wurf <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[wurf,Intercept]")
+
+H.lapi_mend <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[mend,Intercept]")
+H.lapi_mont <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[mont,Intercept]")
+H.lapi_muci <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[muci,Intercept]")
+H.lapi_pasc <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[pasc,Intercept]")
+H.lapi_prat <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[prat,Intercept]")
+H.lapi_psit <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[psit,Intercept]")
+H.lapi_pyre <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[pyre,Intercept]")
+H.lapi_soro <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[soro,Intercept]")
+H.lapi_wurf <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[wurf,Intercept]")
+
+H.mend_mont <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[mont,Intercept]")
+H.mend_muci <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[muci,Intercept]")
+H.mend_pasc <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[pasc,Intercept]")
+H.mend_prat <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[prat,Intercept]")
+H.mend_psit <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[psit,Intercept]")
+H.mend_pyre <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[pyre,Intercept]")
+H.mend_soro <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[soro,Intercept]")
+H.mend_wurf <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[wurf,Intercept]")
+
+H.mont_muci <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[muci,Intercept]")
+H.mont_pasc <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[pasc,Intercept]")
+H.mont_prat <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[prat,Intercept]")
+H.mont_psit <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[psit,Intercept]")
+H.mont_pyre <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[pyre,Intercept]")
+H.mont_soro <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[soro,Intercept]")
+H.mont_wurf <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[wurf,Intercept]")
+
+H.muci_pasc <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[pasc,Intercept]")
+H.muci_prat <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[prat,Intercept]")
+H.muci_psit <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[psit,Intercept]")
+H.muci_pyre <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[pyre,Intercept]")
+H.muci_soro <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[soro,Intercept]")
+H.muci_wurf <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[wurf,Intercept]")
+
+H.pasc_prat <- paste("r_bb.sp[pasc,Intercept] = r_bb.sp[prat,Intercept]")
+H.pasc_psit <- paste("r_bb.sp[pasc,Intercept] = r_bb.sp[psit,Intercept]")
+H.pasc_pyre <- paste("r_bb.sp[pasc,Intercept] = r_bb.sp[pyre,Intercept]")
+H.pasc_soro <- paste("r_bb.sp[pasc,Intercept] = r_bb.sp[soro,Intercept]")
+H.pasc_wurf <- paste("r_bb.sp[pasc,Intercept] = r_bb.sp[wurf,Intercept]")
+
+H.prat_psit <- paste("r_bb.sp[prat,Intercept] = r_bb.sp[psit,Intercept]")
+H.prat_pyre <- paste("r_bb.sp[prat,Intercept] = r_bb.sp[pyre,Intercept]")
+H.prat_soro <- paste("r_bb.sp[prat,Intercept] = r_bb.sp[soro,Intercept]")
+H.prat_wurf <- paste("r_bb.sp[prat,Intercept] = r_bb.sp[wurf,Intercept]")
+
+H.psit_pyre <- paste("r_bb.sp[psit,Intercept] = r_bb.sp[pyre,Intercept]")
+H.psit_soro <- paste("r_bb.sp[psit,Intercept] = r_bb.sp[soro,Intercept]")
+H.psit_wurf <- paste("r_bb.sp[psit,Intercept] = r_bb.sp[wurf,Intercept]")
+
+H.pyre_soro <- paste("r_bb.sp[pyre,Intercept] = r_bb.sp[soro,Intercept]")
+H.pyre_wurf <- paste("r_bb.sp[pyre,Intercept] = r_bb.sp[wurf,Intercept]")
+
+H.soro_wurf <- paste("r_bb.sp[soro,Intercept] = r_bb.sp[wurf,Intercept]")
+```
+
+## 1.5 Test contrasts
+```{r}
+### Plant.sp
+hypothesis(brm.plantsp, H1)
+
+hypothesis(brm.plantsp, H.bss_gers, class = "")
+hypothesis(brm.plantsp, H.bss_hort, class = "")
+hypothesis(brm.plantsp, H.bss_humi, class = "")
+hypothesis(brm.plantsp, H.bss_hypn, class = "")
+hypothesis(brm.plantsp, H.bss_jone, class = "")
+hypothesis(brm.plantsp, H.bss_lapi, class = "")
+hypothesis(brm.plantsp, H.bss_mend, class = "")
+hypothesis(brm.plantsp, H.bss_mont, class = "")
+hypothesis(brm.plantsp, H.bss_muci, class = "")
+hypothesis(brm.plantsp, H.bss_pasc, class = "")
+hypothesis(brm.plantsp, H.bss_prat, class = "")
+hypothesis(brm.plantsp, H.bss_psit, class = "")
+hypothesis(brm.plantsp, H.bss_pyre, class = "")
+hypothesis(brm.plantsp, H.bss_soro, class = "")
+hypothesis(brm.plantsp, H.bss_wurf, class = "")
+
+hypothesis(brm.plantsp, H.gers_hort, class = "")
+hypothesis(brm.plantsp, H.gers_humi, class = "")
+hypothesis(brm.plantsp, H.gers_hypn, class = "")
+hypothesis(brm.plantsp, H.gers_jone, class = "")
+hypothesis(brm.plantsp, H.gers_lapi, class = "")
+hypothesis(brm.plantsp, H.gers_mend, class = "")
+hypothesis(brm.plantsp, H.gers_mont, class = "")
+hypothesis(brm.plantsp, H.gers_muci, class = "")
+hypothesis(brm.plantsp, H.gers_pasc, class = "")
+hypothesis(brm.plantsp, H.gers_prat, class = "")
+hypothesis(brm.plantsp, H.gers_psit, class = "")
+hypothesis(brm.plantsp, H.gers_pyre, class = "")
+hypothesis(brm.plantsp, H.gers_soro, class = "")
+hypothesis(brm.plantsp, H.gers_wurf, class = "")
+
+hypothesis(brm.plantsp, H.hort_humi, class = "")
+hypothesis(brm.plantsp, H.hort_hypn, class = "")
+hypothesis(brm.plantsp, H.hort_jone, class = "")
+hypothesis(brm.plantsp, H.hort_lapi, class = "")
+hypothesis(brm.plantsp, H.hort_mend, class = "")
+hypothesis(brm.plantsp, H.hort_mont, class = "")
+hypothesis(brm.plantsp, H.hort_muci, class = "")
+hypothesis(brm.plantsp, H.hort_pasc, class = "")
+hypothesis(brm.plantsp, H.hort_prat, class = "")
+hypothesis(brm.plantsp, H.hort_psit, class = "")
+hypothesis(brm.plantsp, H.hort_pyre, class = "")
+hypothesis(brm.plantsp, H.hort_soro, class = "")
+hypothesis(brm.plantsp, H.hort_wurf, class = "")
+
+hypothesis(brm.plantsp, H.humi_hypn, class = "")
+hypothesis(brm.plantsp, H.humi_jone, class = "")
+hypothesis(brm.plantsp, H.humi_lapi, class = "")
+hypothesis(brm.plantsp, H.humi_mend, class = "")
+hypothesis(brm.plantsp, H.humi_mont, class = "")
+hypothesis(brm.plantsp, H.humi_muci, class = "")
+hypothesis(brm.plantsp, H.humi_pasc, class = "")
+hypothesis(brm.plantsp, H.humi_prat, class = "")
+hypothesis(brm.plantsp, H.humi_psit, class = "")
+hypothesis(brm.plantsp, H.humi_pyre, class = "")
+hypothesis(brm.plantsp, H.humi_soro, class = "")
+hypothesis(brm.plantsp, H.humi_wurf, class = "")
+
+hypothesis(brm.plantsp, H.hypn_jone, class = "")
+hypothesis(brm.plantsp, H.hypn_lapi, class = "")
+hypothesis(brm.plantsp, H.hypn_mend, class = "")
+hypothesis(brm.plantsp, H.hypn_mont, class = "")
+hypothesis(brm.plantsp, H.hypn_muci, class = "")
+hypothesis(brm.plantsp, H.hypn_pasc, class = "")
+hypothesis(brm.plantsp, H.hypn_prat, class = "")
+hypothesis(brm.plantsp, H.hypn_psit, class = "")
+hypothesis(brm.plantsp, H.hypn_pyre, class = "")
+hypothesis(brm.plantsp, H.hypn_soro, class = "")
+hypothesis(brm.plantsp, H.hypn_wurf, class = "")
+
+hypothesis(brm.plantsp, H.jone_lapi, class = "")
+hypothesis(brm.plantsp, H.jone_mend, class = "")
+hypothesis(brm.plantsp, H.jone_mont, class = "")
+hypothesis(brm.plantsp, H.jone_muci, class = "")
+hypothesis(brm.plantsp, H.jone_pasc, class = "")
+hypothesis(brm.plantsp, H.jone_prat, class = "")
+hypothesis(brm.plantsp, H.jone_psit, class = "")
+hypothesis(brm.plantsp, H.jone_pyre, class = "")
+hypothesis(brm.plantsp, H.jone_soro, class = "")
+hypothesis(brm.plantsp, H.jone_wurf, class = "")
+
+hypothesis(brm.plantsp, H.lapi_mend, class = "")
+hypothesis(brm.plantsp, H.lapi_mont, class = "")
+hypothesis(brm.plantsp, H.lapi_muci, class = "")
+hypothesis(brm.plantsp, H.lapi_pasc, class = "")
+hypothesis(brm.plantsp, H.lapi_prat, class = "")
+hypothesis(brm.plantsp, H.lapi_psit, class = "")
+hypothesis(brm.plantsp, H.lapi_pyre, class = "")
+hypothesis(brm.plantsp, H.lapi_soro, class = "")
+hypothesis(brm.plantsp, H.lapi_wurf, class = "")
+
+hypothesis(brm.plantsp, H.mend_mont, class = "")
+hypothesis(brm.plantsp, H.mend_muci, class = "")
+hypothesis(brm.plantsp, H.mend_pasc, class = "")
+hypothesis(brm.plantsp, H.mend_prat, class = "")
+hypothesis(brm.plantsp, H.mend_psit, class = "")
+hypothesis(brm.plantsp, H.mend_pyre, class = "")
+hypothesis(brm.plantsp, H.mend_soro, class = "")
+hypothesis(brm.plantsp, H.mend_wurf, class = "")
+
+hypothesis(brm.plantsp, H.mont_muci, class = "")
+hypothesis(brm.plantsp, H.mont_pasc, class = "")
+hypothesis(brm.plantsp, H.mont_prat, class = "")
+hypothesis(brm.plantsp, H.mont_psit, class = "")
+hypothesis(brm.plantsp, H.mont_pyre, class = "")
+hypothesis(brm.plantsp, H.mont_soro, class = "")
+hypothesis(brm.plantsp, H.mont_wurf, class = "")
+
+hypothesis(brm.plantsp, H.muci_pasc, class = "")
+hypothesis(brm.plantsp, H.muci_prat, class = "")
+hypothesis(brm.plantsp, H.muci_psit, class = "")
+hypothesis(brm.plantsp, H.muci_pyre, class = "")
+hypothesis(brm.plantsp, H.muci_soro, class = "")
+hypothesis(brm.plantsp, H.muci_wurf, class = "")
+
+hypothesis(brm.plantsp, H.pasc_prat, class = "")
+hypothesis(brm.plantsp, H.pasc_psit, class = "")
+hypothesis(brm.plantsp, H.pasc_pyre, class = "")
+hypothesis(brm.plantsp, H.pasc_soro, class = "")
+hypothesis(brm.plantsp, H.pasc_wurf, class = "")
+
+hypothesis(brm.plantsp, H.prat_psit, class = "")
+hypothesis(brm.plantsp, H.prat_pyre, class = "")
+hypothesis(brm.plantsp, H.prat_soro, class = "")
+hypothesis(brm.plantsp, H.prat_wurf, class = "")
+
+hypothesis(brm.plantsp, H.psit_pyre, class = "")
+hypothesis(brm.plantsp, H.psit_soro, class = "")
+hypothesis(brm.plantsp, H.psit_wurf, class = "")
+
+hypothesis(brm.plantsp, H.pyre_soro, class = "")
+hypothesis(brm.plantsp, H.pyre_wurf, class = "")
+
+hypothesis(brm.plantsp, H.soro_wurf, class = "")
+
+### Kugler type
+hypothesis(brm.ktype, H1)
+
+hypothesis(brm.ktype, H.bss_gers, class = "")
+hypothesis(brm.ktype, H.bss_hort, class = "")
+hypothesis(brm.ktype, H.bss_humi, class = "")
+hypothesis(brm.ktype, H.bss_hypn, class = "")
+hypothesis(brm.ktype, H.bss_jone, class = "")
+hypothesis(brm.ktype, H.bss_lapi, class = "")
+hypothesis(brm.ktype, H.bss_mend, class = "")
+hypothesis(brm.ktype, H.bss_mont, class = "")
+hypothesis(brm.ktype, H.bss_muci, class = "")
+hypothesis(brm.ktype, H.bss_pasc, class = "")
+hypothesis(brm.ktype, H.bss_prat, class = "")
+hypothesis(brm.ktype, H.bss_psit, class = "")
+hypothesis(brm.ktype, H.bss_pyre, class = "")
+hypothesis(brm.ktype, H.bss_soro, class = "")
+hypothesis(brm.ktype, H.bss_wurf, class = "")
+
+hypothesis(brm.ktype, H.gers_hort, class = "")
+hypothesis(brm.ktype, H.gers_humi, class = "")
+hypothesis(brm.ktype, H.gers_hypn, class = "")
+hypothesis(brm.ktype, H.gers_jone, class = "")
+hypothesis(brm.ktype, H.gers_lapi, class = "")
+hypothesis(brm.ktype, H.gers_mend, class = "")
+hypothesis(brm.ktype, H.gers_mont, class = "")
+hypothesis(brm.ktype, H.gers_muci, class = "")
+hypothesis(brm.ktype, H.gers_pasc, class = "")
+hypothesis(brm.ktype, H.gers_prat, class = "")
+hypothesis(brm.ktype, H.gers_psit, class = "")
+hypothesis(brm.ktype, H.gers_pyre, class = "")
+hypothesis(brm.ktype, H.gers_soro, class = "")
+hypothesis(brm.ktype, H.gers_wurf, class = "")
+
+hypothesis(brm.ktype, H.hort_humi, class = "")
+hypothesis(brm.ktype, H.hort_hypn, class = "")
+hypothesis(brm.ktype, H.hort_jone, class = "")
+hypothesis(brm.ktype, H.hort_lapi, class = "")
+hypothesis(brm.ktype, H.hort_mend, class = "")
+hypothesis(brm.ktype, H.hort_mont, class = "")
+hypothesis(brm.ktype, H.hort_muci, class = "")
+hypothesis(brm.ktype, H.hort_pasc, class = "")
+hypothesis(brm.ktype, H.hort_prat, class = "")
+hypothesis(brm.ktype, H.hort_psit, class = "")
+hypothesis(brm.ktype, H.hort_pyre, class = "")
+hypothesis(brm.ktype, H.hort_soro, class = "")
+hypothesis(brm.ktype, H.hort_wurf, class = "")
+
+hypothesis(brm.ktype, H.humi_hypn, class = "")
+hypothesis(brm.ktype, H.humi_jone, class = "")
+hypothesis(brm.ktype, H.humi_lapi, class = "")
+hypothesis(brm.ktype, H.humi_mend, class = "")
+hypothesis(brm.ktype, H.humi_mont, class = "")
+hypothesis(brm.ktype, H.humi_muci, class = "")
+hypothesis(brm.ktype, H.humi_pasc, class = "")
+hypothesis(brm.ktype, H.humi_prat, class = "")
+hypothesis(brm.ktype, H.humi_psit, class = "")
+hypothesis(brm.ktype, H.humi_pyre, class = "")
+hypothesis(brm.ktype, H.humi_soro, class = "")
+hypothesis(brm.ktype, H.humi_wurf, class = "")
+
+hypothesis(brm.ktype, H.hypn_jone, class = "")
+hypothesis(brm.ktype, H.hypn_lapi, class = "")
+hypothesis(brm.ktype, H.hypn_mend, class = "")
+hypothesis(brm.ktype, H.hypn_mont, class = "")
+hypothesis(brm.ktype, H.hypn_muci, class = "")
+hypothesis(brm.ktype, H.hypn_pasc, class = "")
+hypothesis(brm.ktype, H.hypn_prat, class = "")
+hypothesis(brm.ktype, H.hypn_psit, class = "")
+hypothesis(brm.ktype, H.hypn_pyre, class = "")
+hypothesis(brm.ktype, H.hypn_soro, class = "")
+hypothesis(brm.ktype, H.hypn_wurf, class = "")
+
+hypothesis(brm.ktype, H.jone_lapi, class = "")
+hypothesis(brm.ktype, H.jone_mend, class = "")
+hypothesis(brm.ktype, H.jone_mont, class = "")
+hypothesis(brm.ktype, H.jone_muci, class = "")
+hypothesis(brm.ktype, H.jone_pasc, class = "")
+hypothesis(brm.ktype, H.jone_prat, class = "")
+hypothesis(brm.ktype, H.jone_psit, class = "")
+hypothesis(brm.ktype, H.jone_pyre, class = "")
+hypothesis(brm.ktype, H.jone_soro, class = "")
+hypothesis(brm.ktype, H.jone_wurf, class = "")
+
+hypothesis(brm.ktype, H.lapi_mend, class = "")
+hypothesis(brm.ktype, H.lapi_mont, class = "")
+hypothesis(brm.ktype, H.lapi_muci, class = "")
+hypothesis(brm.ktype, H.lapi_pasc, class = "")
+hypothesis(brm.ktype, H.lapi_prat, class = "")
+hypothesis(brm.ktype, H.lapi_psit, class = "")
+hypothesis(brm.ktype, H.lapi_pyre, class = "")
+hypothesis(brm.ktype, H.lapi_soro, class = "")
+hypothesis(brm.ktype, H.lapi_wurf, class = "")
+
+hypothesis(brm.ktype, H.mend_mont, class = "")
+hypothesis(brm.ktype, H.mend_muci, class = "")
+hypothesis(brm.ktype, H.mend_pasc, class = "")
+hypothesis(brm.ktype, H.mend_prat, class = "")
+hypothesis(brm.ktype, H.mend_psit, class = "")
+hypothesis(brm.ktype, H.mend_pyre, class = "")
+hypothesis(brm.ktype, H.mend_soro, class = "")
+hypothesis(brm.ktype, H.mend_wurf, class = "")
+
+hypothesis(brm.ktype, H.mont_muci, class = "")
+hypothesis(brm.ktype, H.mont_pasc, class = "")
+hypothesis(brm.ktype, H.mont_prat, class = "")
+hypothesis(brm.ktype, H.mont_psit, class = "")
+hypothesis(brm.ktype, H.mont_pyre, class = "")
+hypothesis(brm.ktype, H.mont_soro, class = "")
+hypothesis(brm.ktype, H.mont_wurf, class = "")
+
+hypothesis(brm.ktype, H.muci_pasc, class = "")
+hypothesis(brm.ktype, H.muci_prat, class = "")
+hypothesis(brm.ktype, H.muci_psit, class = "")
+hypothesis(brm.ktype, H.muci_pyre, class = "")
+hypothesis(brm.ktype, H.muci_soro, class = "")
+hypothesis(brm.ktype, H.muci_wurf, class = "")
+
+hypothesis(brm.ktype, H.pasc_prat, class = "")
+hypothesis(brm.ktype, H.pasc_psit, class = "")
+hypothesis(brm.ktype, H.pasc_pyre, class = "")
+hypothesis(brm.ktype, H.pasc_soro, class = "")
+hypothesis(brm.ktype, H.pasc_wurf, class = "")
+
+hypothesis(brm.ktype, H.prat_psit, class = "")
+hypothesis(brm.ktype, H.prat_pyre, class = "")
+hypothesis(brm.ktype, H.prat_soro, class = "")
+hypothesis(brm.ktype, H.prat_wurf, class = "")
+
+hypothesis(brm.ktype, H.psit_pyre, class = "")
+hypothesis(brm.ktype, H.psit_soro, class = "")
+hypothesis(brm.ktype, H.psit_wurf, class = "")
+
+hypothesis(brm.ktype, H.pyre_soro, class = "")
+hypothesis(brm.ktype, H.pyre_wurf, class = "")
+
+hypothesis(brm.ktype, H.soro_wurf, class = "")
+
+### Plant family
+hypothesis(brm.family, H1)
+
+hypothesis(brm.family, H.bss_gers, class = "")
+hypothesis(brm.family, H.bss_hort, class = "")
+hypothesis(brm.family, H.bss_humi, class = "")
+hypothesis(brm.family, H.bss_hypn, class = "")
+hypothesis(brm.family, H.bss_jone, class = "")
+hypothesis(brm.family, H.bss_lapi, class = "")
+hypothesis(brm.family, H.bss_mend, class = "")
+hypothesis(brm.family, H.bss_mont, class = "")
+hypothesis(brm.family, H.bss_muci, class = "")
+hypothesis(brm.family, H.bss_pasc, class = "")
+hypothesis(brm.family, H.bss_prat, class = "")
+hypothesis(brm.family, H.bss_psit, class = "")
+hypothesis(brm.family, H.bss_pyre, class = "")
+hypothesis(brm.family, H.bss_soro, class = "")
+hypothesis(brm.family, H.bss_wurf, class = "")
+
+hypothesis(brm.family, H.gers_hort, class = "")
+hypothesis(brm.family, H.gers_humi, class = "")
+hypothesis(brm.family, H.gers_hypn, class = "")
+hypothesis(brm.family, H.gers_jone, class = "")
+hypothesis(brm.family, H.gers_lapi, class = "")
+hypothesis(brm.family, H.gers_mend, class = "")
+hypothesis(brm.family, H.gers_mont, class = "")
+hypothesis(brm.family, H.gers_muci, class = "")
+hypothesis(brm.family, H.gers_pasc, class = "")
+hypothesis(brm.family, H.gers_prat, class = "")
+hypothesis(brm.family, H.gers_psit, class = "")
+hypothesis(brm.family, H.gers_pyre, class = "")
+hypothesis(brm.family, H.gers_soro, class = "")
+hypothesis(brm.family, H.gers_wurf, class = "")
+
+hypothesis(brm.family, H.hort_humi, class = "")
+hypothesis(brm.family, H.hort_hypn, class = "")
+hypothesis(brm.family, H.hort_jone, class = "")
+hypothesis(brm.family, H.hort_lapi, class = "")
+hypothesis(brm.family, H.hort_mend, class = "")
+hypothesis(brm.family, H.hort_mont, class = "")
+hypothesis(brm.family, H.hort_muci, class = "")
+hypothesis(brm.family, H.hort_pasc, class = "")
+hypothesis(brm.family, H.hort_prat, class = "")
+hypothesis(brm.family, H.hort_psit, class = "")
+hypothesis(brm.family, H.hort_pyre, class = "")
+hypothesis(brm.family, H.hort_soro, class = "")
+hypothesis(brm.family, H.hort_wurf, class = "")
+
+hypothesis(brm.family, H.humi_hypn, class = "")
+hypothesis(brm.family, H.humi_jone, class = "")
+hypothesis(brm.family, H.humi_lapi, class = "")
+hypothesis(brm.family, H.humi_mend, class = "")
+hypothesis(brm.family, H.humi_mont, class = "")
+hypothesis(brm.family, H.humi_muci, class = "")
+hypothesis(brm.family, H.humi_pasc, class = "")
+hypothesis(brm.family, H.humi_prat, class = "")
+hypothesis(brm.family, H.humi_psit, class = "")
+hypothesis(brm.family, H.humi_pyre, class = "")
+hypothesis(brm.family, H.humi_soro, class = "")
+hypothesis(brm.family, H.humi_wurf, class = "")
+
+hypothesis(brm.family, H.hypn_jone, class = "")
+hypothesis(brm.family, H.hypn_lapi, class = "")
+hypothesis(brm.family, H.hypn_mend, class = "")
+hypothesis(brm.family, H.hypn_mont, class = "")
+hypothesis(brm.family, H.hypn_muci, class = "")
+hypothesis(brm.family, H.hypn_pasc, class = "")
+hypothesis(brm.family, H.hypn_prat, class = "")
+hypothesis(brm.family, H.hypn_psit, class = "")
+hypothesis(brm.family, H.hypn_pyre, class = "")
+hypothesis(brm.family, H.hypn_soro, class = "")
+hypothesis(brm.family, H.hypn_wurf, class = "")
+
+hypothesis(brm.family, H.jone_lapi, class = "")
+hypothesis(brm.family, H.jone_mend, class = "")
+hypothesis(brm.family, H.jone_mont, class = "")
+hypothesis(brm.family, H.jone_muci, class = "")
+hypothesis(brm.family, H.jone_pasc, class = "")
+hypothesis(brm.family, H.jone_prat, class = "")
+hypothesis(brm.family, H.jone_psit, class = "")
+hypothesis(brm.family, H.jone_pyre, class = "")
+hypothesis(brm.family, H.jone_soro, class = "")
+hypothesis(brm.family, H.jone_wurf, class = "")
+
+hypothesis(brm.family, H.lapi_mend, class = "")
+hypothesis(brm.family, H.lapi_mont, class = "")
+hypothesis(brm.family, H.lapi_muci, class = "")
+hypothesis(brm.family, H.lapi_pasc, class = "")
+hypothesis(brm.family, H.lapi_prat, class = "")
+hypothesis(brm.family, H.lapi_psit, class = "")
+hypothesis(brm.family, H.lapi_pyre, class = "")
+hypothesis(brm.family, H.lapi_soro, class = "")
+hypothesis(brm.family, H.lapi_wurf, class = "")
+
+hypothesis(brm.family, H.mend_mont, class = "")
+hypothesis(brm.family, H.mend_muci, class = "")
+hypothesis(brm.family, H.mend_pasc, class = "")
+hypothesis(brm.family, H.mend_prat, class = "")
+hypothesis(brm.family, H.mend_psit, class = "")
+hypothesis(brm.family, H.mend_pyre, class = "")
+hypothesis(brm.family, H.mend_soro, class = "")
+hypothesis(brm.family, H.mend_wurf, class = "")
+
+hypothesis(brm.family, H.mont_muci, class = "")
+hypothesis(brm.family, H.mont_pasc, class = "")
+hypothesis(brm.family, H.mont_prat, class = "")
+hypothesis(brm.family, H.mont_psit, class = "")
+hypothesis(brm.family, H.mont_pyre, class = "")
+hypothesis(brm.family, H.mont_soro, class = "")
+hypothesis(brm.family, H.mont_wurf, class = "")
+
+hypothesis(brm.family, H.muci_pasc, class = "")
+hypothesis(brm.family, H.muci_prat, class = "")
+hypothesis(brm.family, H.muci_psit, class = "")
+hypothesis(brm.family, H.muci_pyre, class = "")
+hypothesis(brm.family, H.muci_soro, class = "")
+hypothesis(brm.family, H.muci_wurf, class = "")
+
+hypothesis(brm.family, H.pasc_prat, class = "")
+hypothesis(brm.family, H.pasc_psit, class = "")
+hypothesis(brm.family, H.pasc_pyre, class = "")
+hypothesis(brm.family, H.pasc_soro, class = "")
+hypothesis(brm.family, H.pasc_wurf, class = "")
+
+hypothesis(brm.family, H.prat_psit, class = "")
+hypothesis(brm.family, H.prat_pyre, class = "")
+hypothesis(brm.family, H.prat_soro, class = "")
+hypothesis(brm.family, H.prat_wurf, class = "")
+
+hypothesis(brm.family, H.psit_pyre, class = "")
+hypothesis(brm.family, H.psit_soro, class = "")
+hypothesis(brm.family, H.psit_wurf, class = "")
+
+hypothesis(brm.family, H.pyre_soro, class = "")
+hypothesis(brm.family, H.pyre_wurf, class = "")
+
+hypothesis(brm.family, H.soro_wurf, class = "")
+```
+
+# 2. Having established the bias of BB foraging, let's zoom in on which morphotypes and families were over/under-represented
+
+## 2.1 Defining bias
+
+Deciding exactly how to represent bias turns out to be rather tricky. I am trying out a few options right now.
+
+1.
+Bias could  be trated simply as the Euclidean difference between prop.flower.cover and prop.visits.
+This has the effect of weighting bias by flower.cover, particularly negative bias, since an abundant plant can have a greater negative bias (because it is farther from 0) than a rare plant.
+Euclidean bias is bounded between -1 and 1.
+Its symmetry is a nice property that sets it apart from proportional or fold-change bias.
+If, for whatever reason, I decided that a log-transformation of this measure were desirable, I think I could simply add a constant of 1 to all values, which would move the boundary up to (0, 2).
+I think the biggest problem with measuring bias in this way is that it makes no distinction between a plant that was visited a little and one that was visited not at all, and that could be an important biological difference.
+For example, consider two plants, one with a proportional abundance = 0.01 and proportional visitation = 0.01, and the second with proportional abundance = 0.01 and proportional visitation = 0.00.
+These will look identical in our analysis, but it is possible that the unvisited plant would not have been visited no matter how high its abundance was.
+**For starters, though, I am going to proceed with this metric because it is the simplest.
+And because DJ thought it was a good idea.**
+
+2.
+Alternatively, bias could be calculated as a simple proportion: prop.visits/prop.flower.cover.
+If anything, this might tend to reverse the weighting, since a small Euclidean change in a rare species would be a large proportional change.
+Proportional change has a lower boundary of 0 and no upper boundary.
+This approach highlights cases of positive bias, but it is even worse than euclidean bias in terms of obscuring the distinction between plants visited slightly and those visited not at all.
+Worse in a sense.
+In another sense, it is better than euclidean bias, because with proportional visitation a plant visited 0 times is at least numerical distinguishable from one visited 1 time (visited plants will always have non-zero proportional bias).
+
+3.
+Similarly to proportional bias, I could calculate bias as fold change.
+This converges to be virtually identical to proportional change as bias becomes more positive, but fold change has a lower bound of -1 instead of 0.
+
+4.
+If the distinction between plants that were visited and those that were not is what we're most interested in, we could simply treat bias as binary.
+This, of course, comes at the price of not being able to distinguish a plant that received 0.01 proportional visitation and one that recevied  0.99 proportional visitation, and this distinction is arguably important.
+
+5.
+An interesting option is to first separate visited from unvisited plants, set the bias of unvisited plants to 0, and then exponentiate the euclidean bias of all visited plant species.
+This has some desirable properties.
+First, it clearly distinguished visited plants from unvisited ones.
+Second, it assigns plants visited perfectly in proportion to their abundance a sensible bias of 1 (just like proportional bias).
+Thus, plants with a psotive bias are > 1 (bounded at 2.718282 -- okay, so that's a bit inelegant), those with a negative bias are < 1 (bounded at 0.367895), and those not visited at all = 0.
+
+6.
+The problem with the option above is that is adds weight to positive bias.
+Another approach is to separate the unvisited plants from the visited, set unvisited bias = 0, and visited bias = prop.visits - prop.flower.cover + 1.
+The +1 serves to make the distribution positive, with 1 represening perfect proportionality between visitation and floral abundance.
+This creates a symmetrically bounded distribution between 0 (not inclusive) and 2 (inclusive), plus the true zeros from the unvisited plants.
+It feels a bit...slapped together, though.
+
+## 2.2 Function for calculating bias per group (= criterion)
+```{r}
+biaseval_group <- function(x, criterion) {
+
+  criterion <- enquo(criterion)
+
+  x %>%
+    mutate(criterion = factor(!!criterion)) %>% # having this as a factor will help with modeling later on
+    group_by(site, date, bb.sp, criterion) %>%
+    summarize(flower.cover = sum(flower.cover), # get raw floer cover and visits by group
+              visits = sum(visits)) %>%
+    group_by(site, date, bb.sp) %>%
+    mutate(prop.flower.cover = flower.cover/sum(flower.cover), # get proportional flower cover and visits by group
+           prop.visits = visits/sum(visits)) %>%
+    mutate(bias = prop.visits - prop.flower.cover, # bias as difference
+           bias.prop = prop.visits/prop.flower.cover, # bias as proportion
+           bias.fold = bias/prop.flower.cover, # bias as fold change
+           bias.bin = if_else(bias > 0, 1, 0), # bias as binary
+           bias.exp = if_else(visits == 0, 0, exp(prop.visits - prop.flower.cover)),  # bias as exponentiated difference
+           bias.split = if_else(visits == 0, 0, (prop.visits - prop.flower.cover + 1))) %>%
+    mutate(yday = yday(date)) %>%
+    select(site, date, yday, bb.sp, criterion, 
+           bias, bias.prop, bias.fold, bias.bin, bias.exp, bias.split,
+           visits, flower.cover, prop.visits, prop.flower.cover)
+}
+```
+
+## 2.3 Calculate bias per plant speces/ktype/family
+```{r}
+bias_group_plantsp <- biaseval_group(visit_cover, plant.sp) %>%
+  left_join(site_data) %>%
+  rename(group = criterion) %>%
+  mutate(criterion = rep("plant.sp", n()))
+
+bias_group_ktype <- biaseval_group(visit_cover, k.type.ss) %>%
+  left_join(site_data) %>%
+  rename(group = criterion) %>%
+  mutate(criterion = rep("ktype", n()))
+
+bias_group_family <- biaseval_group(visit_cover, plant.family) %>%
+  left_join(site_data) %>%
+  rename(group = criterion) %>%
+  mutate(criterion = rep("plant.family", n()))
+
+bias_group <- bind_rows(bias_group_plantsp, bias_group_ktype, bias_group_family)
+
+ggplot(bias_group, aes(prop.flower.cover, prop.visits)) +
+  geom_point(alpha = 0.1) +
+  #geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
+  geom_smooth(method = "lm") +
+  geom_abline(slope = 1, color = "red", linetype = "dashed") +
+  facet_wrap(~criterion)
+
+ggplot(bias_group, aes(log(prop.flower.cover), prop.visits)) +
+  geom_point(alpha = 0.1) +
+  geom_smooth() +
+  facet_wrap(~criterion)
+
+ggplot(filter(bias_group, criterion == "ktype"), aes(prop.flower.cover, prop.visits)) +
+  geom_point(alpha = 0.1) +
+  geom_smooth() +
+  geom_abline(slope = 1, color = "red", linetype = "dashed") +
+  facet_wrap(~group)
+```
+
+## 2.4 Visualize
+```{r}
+# Check the distribution of bias
+ggplot(bias_group_ktype, aes(bias)) +
+  geom_histogram()
+
+ggplot(bias_group_ktype, aes(bias)) +
+  geom_histogram() +
+  facet_wrap(~criterion)
+
+ggplot(bias_group_family, aes(bias)) +
+  geom_histogram()
+
+ggplot(bias_group_family, aes(bias)) +
+  geom_histogram() +
+  facet_wrap(~criterion)
+
+# Bias by Kugler morphotype
+ggplot(bias_group_ktype, aes(group, bias)) +
+  geom_boxplot()
+
+ggplot(bias_group_ktype, aes(group, bias)) +
+  geom_boxplot() +
+  facet_wrap(~bb.sp)
+
+ggplot(bias_group_ktype, aes(bb.sp, bias)) +
+  geom_boxplot() +
+  facet_wrap(~group)
+
+ggplot(bias_group_ktype, aes(elev.mean, bias)) +
+  #geom_point() +
+  geom_smooth(method = gam, formula = y ~ s(x, k = 5)) +
+  facet_wrap(~group)
+
+# Bias by plant family
+ggplot(bias_group_family, aes(group, bias)) +
+  geom_boxplot()
+
+ggplot(bias_group_family, aes(group, bias)) +
+  geom_boxplot() +
+  facet_wrap(~bb.sp)
+
+ggplot(bias_group_family, aes(bb.sp, bias)) +
+  geom_boxplot() +
+  facet_wrap(~group)
+
+ggplot(bias_group_family, aes(elev.mean, bias)) +
+  #geom_point() +
+  geom_smooth(method = gam, formula = y ~ s(x, k = 5)) +
+  facet_wrap(~group)
+```
+
+## 2.5 Explaining bias by floral traits, BB spp, and elevation (and maybe time)
+```{r}
+repmod.ktype <- brm(bias ~ elev.mean + (1 | criterion), 
+                    data = bias_ktype)
+
+repmod.ktype <- brm(bias ~ criterion, 
+                    data = bias_ktype)
+parnames(repmod.ktype)
+mcmc_plot(repmod.ktype)
+conditional_effects(repmod.ktype)
+
+H.K5_K6 <- paste("b_criterion5 = b_criterion6")
+hypothesis(repmod.ktype, H.K5_K6, class = "")
+```
+
+# How do species vary in their biases
+```{r}
+bias_matrix.plantsp <- bias_group_plantsp %>%
+  group_by(bb.sp, criterion) %>%
+  summarize(bias = mean(bias)) %>%
+  pivot_wider(values_from = bias, names_from = criterion, values_fill = 0) %>%
+  column_to_rownames("bb.sp")
+
+bias_matrix.ktype <- bias_group_ktype %>%
+  group_by(bb.sp, criterion) %>%
+  summarize(bias = mean(bias)) %>%
+  pivot_wider(values_from = bias, names_from = criterion, values_fill = 0) %>%
+  column_to_rownames("bb.sp")
+
+library(vegan)
+
+pca.plantsp <- rda(bias_matrix.plantsp)
+plot(pca.plantsp, type = "t")
+
+pca.ktype <- rda(bias_matrix.ktype)
+plot(pca.ktype)
+
+```
+
+# 3. Can we explain bias by BB-Fl trait associations? (RLQ + fourth corner analysis)
+## 3.1
+## 3.2 Do BB-FL trait associations vary with elevation?
+
+# 4. How does bias relate to network structure?
+## 4.1 A function for calculating niche overlap
 ```{r}
@@
 
-# 1.
-Is bumble bee foraging biased?
-How do taxonomic, phylogenetic, and morphological bias compare -- NO, can't analyze this, these should be considered separate repsonse variables and analyzed in separate models?
-Are some BB species more biased than others?
-## 1.1 Calculate Blüthgen's d' (i.e. visitation bias)
-```{r}
-# No grouping -- just plant.sp
-no_plant.sp <- niche_overlapper(visit_cover, plant.sp)
-
-bias_plantsp.sitedate <- biaseval(visit_cover, plant.sp, site, date) %>%
-  left_join(site_data) %>% # add site data
-  left_join(bb_abundance) %>% # add BB abundance data
-  left_join(fl_abundance) %>% # add fl abundance
-  mutate(criterion = factor(rep("plant.sp", n()))) %>%
-  left_join(no_plant.sp)
-
-# Grouped by k.type.ss
-no_ktype <- niche_overlapper(visit_cover, k.type.ss)
-
-bias_ktype.sitedate <- biaseval(visit_cover, k.type.ss, site, date) %>%
-  left_join(site_data) %>%
-  left_join(bb_abundance) %>% # add BB abundance data
-  left_join(fl_abundance) %>% # add fl abundance
-  mutate(criterion = factor(rep("ktype", n()))) %>%
-  left_join(no_ktype)
-
-# Grouped by plant.family
-no_plant.family <- niche_overlapper(visit_cover, plant.family)
-
-bias_family.sitedate <- biaseval(visit_cover, plant.family, site, date) %>%
-  left_join(site_data) %>%
-  left_join(bb_abundance) %>% # add BB abundance data
-  left_join(fl_abundance) %>% # add fl abundance
-  mutate(criterion = factor(rep("plant.family", n()))) %>%
-  left_join(no_plant.family)
-
-# Collated
-bias_sitedate <- bind_rows(bias_plantsp.sitedate, 
-                           bias_ktype.sitedate, 
-                           bias_family.sitedate) %>%
-  mutate(bb.sp = factor(bb.sp),
-         yday = yday(date),
-         elev = elev.mean/10) %>%
-  left_join(no)
+## 4.2 Calculate niche overlap: NOT READY FOR INTERPRETATION YET
+```{r}
+niche_overlap_plant.sp <- niche_overlapper(visit_cover, plant.sp)
+
+niche_overlap_ktype <- niche_overlapper(visit_cover, k.type.ss)
+
+niche_overlap_plant.family <- niche_overlapper(visit_cover, plant.family)
+
+niche_overlap <- bind_rows(no_plant.sp, 
+                           no_ktype, 
+                           no_plant.family)
 
@@
 When bias is high, BB spp are biased toward DIFFERENT flora, and thus their niches diverge.
-ggplot(bias_sitedate, aes(dprime, mean.niche.overlap, color = criterion)) +
+ggplot(bias, aes(dprime, mean.niche.overlap, color = criterion)) +
   geom_point(alpha = 0.1) +
@@
 
-ggplot(bias_sitedate, aes(dprime, max.niche.overlap, color = criterion)) +
+ggplot(bias, aes(dprime, max.niche.overlap, color = criterion)) +
   geom_point(alpha = 0.1) +
@@
 
-ggplot(bias_sitedate, aes(log(total.flower.cover), dprime, color = criterion)) +
+ggplot(bias, aes(log(total.flower.cover), dprime, color = criterion)) +
   geom_point(alpha = 0.1) +
@@
 
-ggplot(bias_sitedate, aes(log(total.flower.cover), mean.niche.overlap, color = criterion)) +
+ggplot(bias, aes(log(total.flower.cover), mean.niche.overlap, color = criterion)) +
   geom_point(alpha = 0.1) +
@@
                      cores = 3,
-                     data = filter(bias_sitedate, criterion == "ktype"))
+                     data = filter(bias, criterion == "ktype"))
 
@@
 ```
-
-## 1.2 Visualize
-```{r}
-# Compare mean bias by criterion
-ggplot(bias_sitedate, aes(criterion, dprime)) +
-  geom_boxplot()
-
-# Bias ~ elevation * criterion
-ggplot(bias_sitedate, aes(elev.mean, dprime, color = criterion)) +
-  geom_point(alpha = 0.25) +
-  geom_smooth(se = TRUE)
-
-ggplot(bias_sitedate, aes(yday, dprime, color = criterion)) +
-  geom_point(alpha = 0.25) +
-  geom_smooth(se = TRUE)
-
-ggplot(bias_sitedate, aes(elev.mean, dprime, color = criterion)) +
-  geom_point(alpha = 0.25) +
-  geom_smooth(se = FALSE, method = "gam", formula = y ~ s(x, k = 5)) +
-  facet_wrap(~bb.sp)
-
-ggplot(bias_sitedate, aes(yday, dprime, color = criterion)) +
-  geom_point(alpha = 0.25) +
-  geom_smooth(se = FALSE, method = "gam", formula = y ~ s(x, k = 5)) +
-  facet_wrap(~bb.sp)
-
-# Bias ~ year; Blüthgen's d' seems very robust to sampling effects, since it does not vary across years
-ggplot(bias_sitedate, aes(year, dprime, color = criterion)) +
-  geom_boxplot()
-
-ggplot(bias_sitedate, aes(year, dprime, color = criterion)) +
-  geom_boxplot() +
-  facet_wrap(~bb.sp)
-
-# Does bias vary with abundance?
-ggplot(bias_sitedate, aes(log(bb.sp.abund), dprime, color = criterion)) +
-  #geom_point() +
-  geom_smooth() +
-  facet_wrap(~bb.sp)
-
-ggplot(bias_sitedate, aes(bb.sp.abund, dprime, color = criterion)) +
-  #geom_point() +
-  geom_smooth() +
-  facet_wrap(~bb.sp, )
-
-ggplot(bias_sitedate, aes(log(bb.sp.abund), dprime, color = criterion)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth()
-
-ggplot(bias_plantsp.sitedate, aes(log(bb.sp.prop.abund), dprime)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~bb.sp)
-```
-
-## 1.3 Multi-level model: bias ~ elevation + (1 + elevation | bb.sp)
-```{r}
-# brms
-brm.ktype1 <- brm(dprime ~ yday + elev + (1 + elev | bb.sp), 
-                  control = list(max_treedepth = 10),
-                  cores = 3,
-                  iter = 2000,
-                  file = "./output/brm_ktype1",
-                  data = filter(bias_sitedate, criterion ==  "ktype"))
-
-summary(brm.ktype1)
-plot(brm.ktype1)
-parnames(brm.ktype1)
-mcmc_plot(brm.ktype1, pars = c("^r_"))
-mcmc_plot(brm.ktype1, pars = c("^b_"))
-mcmc_plot(brm.ktype1)
-conditional_effects(brm.ktype1)
-pp_check(brm.ktype1)
-
-brm.ktype2 <- brm(dprime ~ 1 + yday + elev + (1 | bb.sp), 
-                  control = list(max_treedepth = 10),
-                  cores = 4,
-                  iter = 2000,
-                  file = "./output/brm_ktype2b",
-                  data = filter(bias_sitedate, criterion ==  "ktype"))
-
-summary(brm.ktype2)
-plot(brm.ktype2)
-parnames(brm.ktype2)
-mcmc_plot(brm.ktype2, pars = c("^r_"))
-mcmc_plot(brm.ktype2, pars = c("^b_"))
-mcmc_plot(brm.ktype2)
-conditional_effects(brm.ktype2)
-pp_check(brm.ktype2)
-
-brm.ktype3 <- brm(dprime ~ 1 + (1 | bb.sp), 
-                  control = list(max_treedepth = 10),
-                  cores = 4,
-                  iter = 2000,
-                  family = "poisson",
-                  file = "./output/brm_ktype3",
-                  data = filter(bias_sitedate, criterion ==  "ktype"))
-
-summary(brm.ktype3)
-plot(brm.ktype3)
-parnames(brm.ktype3)
-mcmc_plot(brm.ktype3, pars = c("^r_"))
-mcmc_plot(brm.ktype3, pars = c("^b_"))
-mcmc_plot(brm.ktype3)
-conditional_effects(brm.ktype3)
-pp_check(brm.ktype3)
-
-# lme4
-library(lme4)
-library(ggeffects)
-
-lmer.ktype <- lmer(dprime ~ 1 + elev:yday + (1 | bb.sp), 
-                  data = filter(bias_sitedate, criterion ==  "ktype"))
-summary(lmer.ktype)
-pred.ktype_yday <- ggpredict(lmer.ktype, "yday")
-pred.ktype_elev <- ggpredict(lmer.ktype, "elev")
-#pred.ktype_yday_ran <- ggpredict(lmer.ktype, "yday", type = "random")
-#pred.ktype_elev_ran <- ggpredict(lmer.ktype, "elev", type = "random")
-pred.ktype_bbsp <- ggpredict(lmer.ktype, "bb.sp", type = "random")
-
-plot(pred.ktype_yday)
-plot(pred.ktype_elev)
-#plot(pred.ktype_yday_ran)
-#plot(pred.ktype_elev_ran)
-plot(pred.ktype_bbsp)
-
-# mgcv
-gam.ktype <- bam(dprime ~ s(bb.sp, bs = "re") + 
-                   t2(yday, elev.mean, bs = c("tp", "gp")),
-                 method = "fREML",
-                 discrete = TRUE,
-                 data = filter(bias_sitedate, criterion ==  "ktype")) %>% getViz()
-
-summary(gam.ktype)
-plot(gam.ktype, allTerms = TRUE)
-
-# inla
-library(INLA)
-
-inla.ktype <- inla(dprime ~ elev.mean, 
-                   family = "gaussian",
-                   data = filter(bias_sitedate, criterion ==  "ktype"))
-
-inla.pen <- inla(dprime ~ 1 + elev.mean,
-  data = filter(bias_sitedate, criterion ==  "ktype"), control.predictor = list(compute = TRUE))
-summary(inla.pen)
-```
-
-## 1.6 Define hypotheses
-```{r}
-# Is d' different from 0?
-H1 <- paste("Intercept = 0")
-
-# Are some species more biased in their foraging than others? 
-H.bss_gers <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[gers,Intercept]")
-H.bss_hort <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[hort,Intercept]")
-H.bss_humi <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[humi,Intercept]")
-H.bss_hypn <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[hypn,Intercept]")
-H.bss_jone <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[jone,Intercept]")
-H.bss_lapi <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[lapi,Intercept]")
-H.bss_mend <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[mend,Intercept]")
-H.bss_mont <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[mont,Intercept]")
-H.bss_muci <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[muci,Intercept]")
-H.bss_pasc <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[pasc,Intercept]")
-H.bss_prat <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[prat,Intercept]")
-H.bss_psit <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[psit,Intercept]")
-H.bss_pyre <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[pyre,Intercept]")
-H.bss_soro <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[soro,Intercept]")
-H.bss_wurf <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.gers_hort <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[hort,Intercept]")
-H.gers_humi <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[humi,Intercept]")
-H.gers_hypn <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[hypn,Intercept]")
-H.gers_jone <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[jone,Intercept]")
-H.gers_lapi <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[lapi,Intercept]")
-H.gers_mend <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[mend,Intercept]")
-H.gers_mont <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[mont,Intercept]")
-H.gers_muci <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[muci,Intercept]")
-H.gers_pasc <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[pasc,Intercept]")
-H.gers_prat <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[prat,Intercept]")
-H.gers_psit <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[psit,Intercept]")
-H.gers_pyre <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[pyre,Intercept]")
-H.gers_soro <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[soro,Intercept]")
-H.gers_wurf <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.hort_humi <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[humi,Intercept]")
-H.hort_hypn <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[hypn,Intercept]")
-H.hort_jone <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[jone,Intercept]")
-H.hort_lapi <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[lapi,Intercept]")
-H.hort_mend <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[mend,Intercept]")
-H.hort_mont <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[mont,Intercept]")
-H.hort_muci <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[muci,Intercept]")
-H.hort_pasc <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[pasc,Intercept]")
-H.hort_prat <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[prat,Intercept]")
-H.hort_psit <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[psit,Intercept]")
-H.hort_pyre <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[pyre,Intercept]")
-H.hort_soro <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[soro,Intercept]")
-H.hort_wurf <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.humi_hypn <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[hypn,Intercept]")
-H.humi_jone <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[jone,Intercept]")
-H.humi_lapi <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[lapi,Intercept]")
-H.humi_mend <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[mend,Intercept]")
-H.humi_mont <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[mont,Intercept]")
-H.humi_muci <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[muci,Intercept]")
-H.humi_pasc <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[pasc,Intercept]")
-H.humi_prat <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[prat,Intercept]")
-H.humi_psit <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[psit,Intercept]")
-H.humi_pyre <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[pyre,Intercept]")
-H.humi_soro <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[soro,Intercept]")
-H.humi_wurf <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.hypn_jone <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[jone,Intercept]")
-H.hypn_lapi <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[lapi,Intercept]")
-H.hypn_mend <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[mend,Intercept]")
-H.hypn_mont <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[mont,Intercept]")
-H.hypn_muci <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[muci,Intercept]")
-H.hypn_pasc <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[pasc,Intercept]")
-H.hypn_prat <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[prat,Intercept]")
-H.hypn_psit <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[psit,Intercept]")
-H.hypn_pyre <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[pyre,Intercept]")
-H.hypn_soro <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[soro,Intercept]")
-H.hypn_wurf <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.jone_lapi <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[lapi,Intercept]")
-H.jone_mend <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[mend,Intercept]")
-H.jone_mont <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[mont,Intercept]")
-H.jone_muci <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[muci,Intercept]")
-H.jone_pasc <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[pasc,Intercept]")
-H.jone_prat <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[prat,Intercept]")
-H.jone_psit <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[psit,Intercept]")
-H.jone_pyre <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[pyre,Intercept]")
-H.jone_soro <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[soro,Intercept]")
-H.jone_wurf <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.lapi_mend <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[mend,Intercept]")
-H.lapi_mont <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[mont,Intercept]")
-H.lapi_muci <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[muci,Intercept]")
-H.lapi_pasc <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[pasc,Intercept]")
-H.lapi_prat <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[prat,Intercept]")
-H.lapi_psit <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[psit,Intercept]")
-H.lapi_pyre <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[pyre,Intercept]")
-H.lapi_soro <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[soro,Intercept]")
-H.lapi_wurf <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.mend_mont <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[mont,Intercept]")
-H.mend_muci <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[muci,Intercept]")
-H.mend_pasc <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[pasc,Intercept]")
-H.mend_prat <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[prat,Intercept]")
-H.mend_psit <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[psit,Intercept]")
-H.mend_pyre <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[pyre,Intercept]")
-H.mend_soro <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[soro,Intercept]")
-H.mend_wurf <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.mont_muci <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[muci,Intercept]")
-H.mont_pasc <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[pasc,Intercept]")
-H.mont_prat <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[prat,Intercept]")
-H.mont_psit <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[psit,Intercept]")
-H.mont_pyre <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[pyre,Intercept]")
-H.mont_soro <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[soro,Intercept]")
-H.mont_wurf <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.muci_pasc <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[pasc,Intercept]")
-H.muci_prat <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[prat,Intercept]")
-H.muci_psit <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[psit,Intercept]")
-H.muci_pyre <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[pyre,Intercept]")
-H.muci_soro <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[soro,Intercept]")
-H.muci_wurf <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.pasc_prat <- paste("r_bb.sp[pasc,Intercept] = r_bb.sp[prat,Intercept]")
-H.pasc_psit <- paste("r_bb.sp[pasc,Intercept] = r_bb.sp[psit,Intercept]")
-H.pasc_pyre <- paste("r_bb.sp[pasc,Intercept] = r_bb.sp[pyre,Intercept]")
-H.pasc_soro <- paste("r_bb.sp[pasc,Intercept] = r_bb.sp[soro,Intercept]")
-H.pasc_wurf <- paste("r_bb.sp[pasc,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.prat_psit <- paste("r_bb.sp[prat,Intercept] = r_bb.sp[psit,Intercept]")
-H.prat_pyre <- paste("r_bb.sp[prat,Intercept] = r_bb.sp[pyre,Intercept]")
-H.prat_soro <- paste("r_bb.sp[prat,Intercept] = r_bb.sp[soro,Intercept]")
-H.prat_wurf <- paste("r_bb.sp[prat,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.psit_pyre <- paste("r_bb.sp[psit,Intercept] = r_bb.sp[pyre,Intercept]")
-H.psit_soro <- paste("r_bb.sp[psit,Intercept] = r_bb.sp[soro,Intercept]")
-H.psit_wurf <- paste("r_bb.sp[psit,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.pyre_soro <- paste("r_bb.sp[pyre,Intercept] = r_bb.sp[soro,Intercept]")
-H.pyre_wurf <- paste("r_bb.sp[pyre,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.soro_wurf <- paste("r_bb.sp[soro,Intercept] = r_bb.sp[wurf,Intercept]")
-```
-
-## 1.7 Test contrasts
-```{r}
-### Plant.sp
-hypothesis(brm.plantsp, H1)
-
-hypothesis(brm.plantsp, H.bss_gers, class = "")
-hypothesis(brm.plantsp, H.bss_hort, class = "")
-hypothesis(brm.plantsp, H.bss_humi, class = "")
-hypothesis(brm.plantsp, H.bss_hypn, class = "")
-hypothesis(brm.plantsp, H.bss_jone, class = "")
-hypothesis(brm.plantsp, H.bss_lapi, class = "")
-hypothesis(brm.plantsp, H.bss_mend, class = "")
-hypothesis(brm.plantsp, H.bss_mont, class = "")
-hypothesis(brm.plantsp, H.bss_muci, class = "")
-hypothesis(brm.plantsp, H.bss_pasc, class = "")
-hypothesis(brm.plantsp, H.bss_prat, class = "")
-hypothesis(brm.plantsp, H.bss_psit, class = "")
-hypothesis(brm.plantsp, H.bss_pyre, class = "")
-hypothesis(brm.plantsp, H.bss_soro, class = "")
-hypothesis(brm.plantsp, H.bss_wurf, class = "")
-
-hypothesis(brm.plantsp, H.gers_hort, class = "")
-hypothesis(brm.plantsp, H.gers_humi, class = "")
-hypothesis(brm.plantsp, H.gers_hypn, class = "")
-hypothesis(brm.plantsp, H.gers_jone, class = "")
-hypothesis(brm.plantsp, H.gers_lapi, class = "")
-hypothesis(brm.plantsp, H.gers_mend, class = "")
-hypothesis(brm.plantsp, H.gers_mont, class = "")
-hypothesis(brm.plantsp, H.gers_muci, class = "")
-hypothesis(brm.plantsp, H.gers_pasc, class = "")
-hypothesis(brm.plantsp, H.gers_prat, class = "")
-hypothesis(brm.plantsp, H.gers_psit, class = "")
-hypothesis(brm.plantsp, H.gers_pyre, class = "")
-hypothesis(brm.plantsp, H.gers_soro, class = "")
-hypothesis(brm.plantsp, H.gers_wurf, class = "")
-
-hypothesis(brm.plantsp, H.hort_humi, class = "")
-hypothesis(brm.plantsp, H.hort_hypn, class = "")
-hypothesis(brm.plantsp, H.hort_jone, class = "")
-hypothesis(brm.plantsp, H.hort_lapi, class = "")
-hypothesis(brm.plantsp, H.hort_mend, class = "")
-hypothesis(brm.plantsp, H.hort_mont, class = "")
-hypothesis(brm.plantsp, H.hort_muci, class = "")
-hypothesis(brm.plantsp, H.hort_pasc, class = "")
-hypothesis(brm.plantsp, H.hort_prat, class = "")
-hypothesis(brm.plantsp, H.hort_psit, class = "")
-hypothesis(brm.plantsp, H.hort_pyre, class = "")
-hypothesis(brm.plantsp, H.hort_soro, class = "")
-hypothesis(brm.plantsp, H.hort_wurf, class = "")
-
-hypothesis(brm.plantsp, H.humi_hypn, class = "")
-hypothesis(brm.plantsp, H.humi_jone, class = "")
-hypothesis(brm.plantsp, H.humi_lapi, class = "")
-hypothesis(brm.plantsp, H.humi_mend, class = "")
-hypothesis(brm.plantsp, H.humi_mont, class = "")
-hypothesis(brm.plantsp, H.humi_muci, class = "")
-hypothesis(brm.plantsp, H.humi_pasc, class = "")
-hypothesis(brm.plantsp, H.humi_prat, class = "")
-hypothesis(brm.plantsp, H.humi_psit, class = "")
-hypothesis(brm.plantsp, H.humi_pyre, class = "")
-hypothesis(brm.plantsp, H.humi_soro, class = "")
-hypothesis(brm.plantsp, H.humi_wurf, class = "")
-
-hypothesis(brm.plantsp, H.hypn_jone, class = "")
-hypothesis(brm.plantsp, H.hypn_lapi, class = "")
-hypothesis(brm.plantsp, H.hypn_mend, class = "")
-hypothesis(brm.plantsp, H.hypn_mont, class = "")
-hypothesis(brm.plantsp, H.hypn_muci, class = "")
-hypothesis(brm.plantsp, H.hypn_pasc, class = "")
-hypothesis(brm.plantsp, H.hypn_prat, class = "")
-hypothesis(brm.plantsp, H.hypn_psit, class = "")
-hypothesis(brm.plantsp, H.hypn_pyre, class = "")
-hypothesis(brm.plantsp, H.hypn_soro, class = "")
-hypothesis(brm.plantsp, H.hypn_wurf, class = "")
-
-hypothesis(brm.plantsp, H.jone_lapi, class = "")
-hypothesis(brm.plantsp, H.jone_mend, class = "")
-hypothesis(brm.plantsp, H.jone_mont, class = "")
-hypothesis(brm.plantsp, H.jone_muci, class = "")
-hypothesis(brm.plantsp, H.jone_pasc, class = "")
-hypothesis(brm.plantsp, H.jone_prat, class = "")
-hypothesis(brm.plantsp, H.jone_psit, class = "")
-hypothesis(brm.plantsp, H.jone_pyre, class = "")
-hypothesis(brm.plantsp, H.jone_soro, class = "")
-hypothesis(brm.plantsp, H.jone_wurf, class = "")
-
-hypothesis(brm.plantsp, H.lapi_mend, class = "")
-hypothesis(brm.plantsp, H.lapi_mont, class = "")
-hypothesis(brm.plantsp, H.lapi_muci, class = "")
-hypothesis(brm.plantsp, H.lapi_pasc, class = "")
-hypothesis(brm.plantsp, H.lapi_prat, class = "")
-hypothesis(brm.plantsp, H.lapi_psit, class = "")
-hypothesis(brm.plantsp, H.lapi_pyre, class = "")
-hypothesis(brm.plantsp, H.lapi_soro, class = "")
-hypothesis(brm.plantsp, H.lapi_wurf, class = "")
-
-hypothesis(brm.plantsp, H.mend_mont, class = "")
-hypothesis(brm.plantsp, H.mend_muci, class = "")
-hypothesis(brm.plantsp, H.mend_pasc, class = "")
-hypothesis(brm.plantsp, H.mend_prat, class = "")
-hypothesis(brm.plantsp, H.mend_psit, class = "")
-hypothesis(brm.plantsp, H.mend_pyre, class = "")
-hypothesis(brm.plantsp, H.mend_soro, class = "")
-hypothesis(brm.plantsp, H.mend_wurf, class = "")
-
-hypothesis(brm.plantsp, H.mont_muci, class = "")
-hypothesis(brm.plantsp, H.mont_pasc, class = "")
-hypothesis(brm.plantsp, H.mont_prat, class = "")
-hypothesis(brm.plantsp, H.mont_psit, class = "")
-hypothesis(brm.plantsp, H.mont_pyre, class = "")
-hypothesis(brm.plantsp, H.mont_soro, class = "")
-hypothesis(brm.plantsp, H.mont_wurf, class = "")
-
-hypothesis(brm.plantsp, H.muci_pasc, class = "")
-hypothesis(brm.plantsp, H.muci_prat, class = "")
-hypothesis(brm.plantsp, H.muci_psit, class = "")
-hypothesis(brm.plantsp, H.muci_pyre, class = "")
-hypothesis(brm.plantsp, H.muci_soro, class = "")
-hypothesis(brm.plantsp, H.muci_wurf, class = "")
-
-hypothesis(brm.plantsp, H.pasc_prat, class = "")
-hypothesis(brm.plantsp, H.pasc_psit, class = "")
-hypothesis(brm.plantsp, H.pasc_pyre, class = "")
-hypothesis(brm.plantsp, H.pasc_soro, class = "")
-hypothesis(brm.plantsp, H.pasc_wurf, class = "")
-
-hypothesis(brm.plantsp, H.prat_psit, class = "")
-hypothesis(brm.plantsp, H.prat_pyre, class = "")
-hypothesis(brm.plantsp, H.prat_soro, class = "")
-hypothesis(brm.plantsp, H.prat_wurf, class = "")
-
-hypothesis(brm.plantsp, H.psit_pyre, class = "")
-hypothesis(brm.plantsp, H.psit_soro, class = "")
-hypothesis(brm.plantsp, H.psit_wurf, class = "")
-
-hypothesis(brm.plantsp, H.pyre_soro, class = "")
-hypothesis(brm.plantsp, H.pyre_wurf, class = "")
-
-hypothesis(brm.plantsp, H.soro_wurf, class = "")
-
-### Kugler type
-hypothesis(brm.ktype, H1)
-
-hypothesis(brm.ktype, H.bss_gers, class = "")
-hypothesis(brm.ktype, H.bss_hort, class = "")
-hypothesis(brm.ktype, H.bss_humi, class = "")
-hypothesis(brm.ktype, H.bss_hypn, class = "")
-hypothesis(brm.ktype, H.bss_jone, class = "")
-hypothesis(brm.ktype, H.bss_lapi, class = "")
-hypothesis(brm.ktype, H.bss_mend, class = "")
-hypothesis(brm.ktype, H.bss_mont, class = "")
-hypothesis(brm.ktype, H.bss_muci, class = "")
-hypothesis(brm.ktype, H.bss_pasc, class = "")
-hypothesis(brm.ktype, H.bss_prat, class = "")
-hypothesis(brm.ktype, H.bss_psit, class = "")
-hypothesis(brm.ktype, H.bss_pyre, class = "")
-hypothesis(brm.ktype, H.bss_soro, class = "")
-hypothesis(brm.ktype, H.bss_wurf, class = "")
-
-hypothesis(brm.ktype, H.gers_hort, class = "")
-hypothesis(brm.ktype, H.gers_humi, class = "")
-hypothesis(brm.ktype, H.gers_hypn, class = "")
-hypothesis(brm.ktype, H.gers_jone, class = "")
-hypothesis(brm.ktype, H.gers_lapi, class = "")
-hypothesis(brm.ktype, H.gers_mend, class = "")
-hypothesis(brm.ktype, H.gers_mont, class = "")
-hypothesis(brm.ktype, H.gers_muci, class = "")
-hypothesis(brm.ktype, H.gers_pasc, class = "")
-hypothesis(brm.ktype, H.gers_prat, class = "")
-hypothesis(brm.ktype, H.gers_psit, class = "")
-hypothesis(brm.ktype, H.gers_pyre, class = "")
-hypothesis(brm.ktype, H.gers_soro, class = "")
-hypothesis(brm.ktype, H.gers_wurf, class = "")
-
-hypothesis(brm.ktype, H.hort_humi, class = "")
-hypothesis(brm.ktype, H.hort_hypn, class = "")
-hypothesis(brm.ktype, H.hort_jone, class = "")
-hypothesis(brm.ktype, H.hort_lapi, class = "")
-hypothesis(brm.ktype, H.hort_mend, class = "")
-hypothesis(brm.ktype, H.hort_mont, class = "")
-hypothesis(brm.ktype, H.hort_muci, class = "")
-hypothesis(brm.ktype, H.hort_pasc, class = "")
-hypothesis(brm.ktype, H.hort_prat, class = "")
-hypothesis(brm.ktype, H.hort_psit, class = "")
-hypothesis(brm.ktype, H.hort_pyre, class = "")
-hypothesis(brm.ktype, H.hort_soro, class = "")
-hypothesis(brm.ktype, H.hort_wurf, class = "")
-
-hypothesis(brm.ktype, H.humi_hypn, class = "")
-hypothesis(brm.ktype, H.humi_jone, class = "")
-hypothesis(brm.ktype, H.humi_lapi, class = "")
-hypothesis(brm.ktype, H.humi_mend, class = "")
-hypothesis(brm.ktype, H.humi_mont, class = "")
-hypothesis(brm.ktype, H.humi_muci, class = "")
-hypothesis(brm.ktype, H.humi_pasc, class = "")
-hypothesis(brm.ktype, H.humi_prat, class = "")
-hypothesis(brm.ktype, H.humi_psit, class = "")
-hypothesis(brm.ktype, H.humi_pyre, class = "")
-hypothesis(brm.ktype, H.humi_soro, class = "")
-hypothesis(brm.ktype, H.humi_wurf, class = "")
-
-hypothesis(brm.ktype, H.hypn_jone, class = "")
-hypothesis(brm.ktype, H.hypn_lapi, class = "")
-hypothesis(brm.ktype, H.hypn_mend, class = "")
-hypothesis(brm.ktype, H.hypn_mont, class = "")
-hypothesis(brm.ktype, H.hypn_muci, class = "")
-hypothesis(brm.ktype, H.hypn_pasc, class = "")
-hypothesis(brm.ktype, H.hypn_prat, class = "")
-hypothesis(brm.ktype, H.hypn_psit, class = "")
-hypothesis(brm.ktype, H.hypn_pyre, class = "")
-hypothesis(brm.ktype, H.hypn_soro, class = "")
-hypothesis(brm.ktype, H.hypn_wurf, class = "")
-
-hypothesis(brm.ktype, H.jone_lapi, class = "")
-hypothesis(brm.ktype, H.jone_mend, class = "")
-hypothesis(brm.ktype, H.jone_mont, class = "")
-hypothesis(brm.ktype, H.jone_muci, class = "")
-hypothesis(brm.ktype, H.jone_pasc, class = "")
-hypothesis(brm.ktype, H.jone_prat, class = "")
-hypothesis(brm.ktype, H.jone_psit, class = "")
-hypothesis(brm.ktype, H.jone_pyre, class = "")
-hypothesis(brm.ktype, H.jone_soro, class = "")
-hypothesis(brm.ktype, H.jone_wurf, class = "")
-
-hypothesis(brm.ktype, H.lapi_mend, class = "")
-hypothesis(brm.ktype, H.lapi_mont, class = "")
-hypothesis(brm.ktype, H.lapi_muci, class = "")
-hypothesis(brm.ktype, H.lapi_pasc, class = "")
-hypothesis(brm.ktype, H.lapi_prat, class = "")
-hypothesis(brm.ktype, H.lapi_psit, class = "")
-hypothesis(brm.ktype, H.lapi_pyre, class = "")
-hypothesis(brm.ktype, H.lapi_soro, class = "")
-hypothesis(brm.ktype, H.lapi_wurf, class = "")
-
-hypothesis(brm.ktype, H.mend_mont, class = "")
-hypothesis(brm.ktype, H.mend_muci, class = "")
-hypothesis(brm.ktype, H.mend_pasc, class = "")
-hypothesis(brm.ktype, H.mend_prat, class = "")
-hypothesis(brm.ktype, H.mend_psit, class = "")
-hypothesis(brm.ktype, H.mend_pyre, class = "")
-hypothesis(brm.ktype, H.mend_soro, class = "")
-hypothesis(brm.ktype, H.mend_wurf, class = "")
-
-hypothesis(brm.ktype, H.mont_muci, class = "")
-hypothesis(brm.ktype, H.mont_pasc, class = "")
-hypothesis(brm.ktype, H.mont_prat, class = "")
-hypothesis(brm.ktype, H.mont_psit, class = "")
-hypothesis(brm.ktype, H.mont_pyre, class = "")
-hypothesis(brm.ktype, H.mont_soro, class = "")
-hypothesis(brm.ktype, H.mont_wurf, class = "")
-
-hypothesis(brm.ktype, H.muci_pasc, class = "")
-hypothesis(brm.ktype, H.muci_prat, class = "")
-hypothesis(brm.ktype, H.muci_psit, class = "")
-hypothesis(brm.ktype, H.muci_pyre, class = "")
-hypothesis(brm.ktype, H.muci_soro, class = "")
-hypothesis(brm.ktype, H.muci_wurf, class = "")
-
-hypothesis(brm.ktype, H.pasc_prat, class = "")
-hypothesis(brm.ktype, H.pasc_psit, class = "")
-hypothesis(brm.ktype, H.pasc_pyre, class = "")
-hypothesis(brm.ktype, H.pasc_soro, class = "")
-hypothesis(brm.ktype, H.pasc_wurf, class = "")
-
-hypothesis(brm.ktype, H.prat_psit, class = "")
-hypothesis(brm.ktype, H.prat_pyre, class = "")
-hypothesis(brm.ktype, H.prat_soro, class = "")
-hypothesis(brm.ktype, H.prat_wurf, class = "")
-
-hypothesis(brm.ktype, H.psit_pyre, class = "")
-hypothesis(brm.ktype, H.psit_soro, class = "")
-hypothesis(brm.ktype, H.psit_wurf, class = "")
-
-hypothesis(brm.ktype, H.pyre_soro, class = "")
-hypothesis(brm.ktype, H.pyre_wurf, class = "")
-
-hypothesis(brm.ktype, H.soro_wurf, class = "")
-
-### Plant family
-hypothesis(brm.family, H1)
-
-hypothesis(brm.family, H.bss_gers, class = "")
-hypothesis(brm.family, H.bss_hort, class = "")
-hypothesis(brm.family, H.bss_humi, class = "")
-hypothesis(brm.family, H.bss_hypn, class = "")
-hypothesis(brm.family, H.bss_jone, class = "")
-hypothesis(brm.family, H.bss_lapi, class = "")
-hypothesis(brm.family, H.bss_mend, class = "")
-hypothesis(brm.family, H.bss_mont, class = "")
-hypothesis(brm.family, H.bss_muci, class = "")
-hypothesis(brm.family, H.bss_pasc, class = "")
-hypothesis(brm.family, H.bss_prat, class = "")
-hypothesis(brm.family, H.bss_psit, class = "")
-hypothesis(brm.family, H.bss_pyre, class = "")
-hypothesis(brm.family, H.bss_soro, class = "")
-hypothesis(brm.family, H.bss_wurf, class = "")
-
-hypothesis(brm.family, H.gers_hort, class = "")
-hypothesis(brm.family, H.gers_humi, class = "")
-hypothesis(brm.family, H.gers_hypn, class = "")
-hypothesis(brm.family, H.gers_jone, class = "")
-hypothesis(brm.family, H.gers_lapi, class = "")
-hypothesis(brm.family, H.gers_mend, class = "")
-hypothesis(brm.family, H.gers_mont, class = "")
-hypothesis(brm.family, H.gers_muci, class = "")
-hypothesis(brm.family, H.gers_pasc, class = "")
-hypothesis(brm.family, H.gers_prat, class = "")
-hypothesis(brm.family, H.gers_psit, class = "")
-hypothesis(brm.family, H.gers_pyre, class = "")
-hypothesis(brm.family, H.gers_soro, class = "")
-hypothesis(brm.family, H.gers_wurf, class = "")
-
-hypothesis(brm.family, H.hort_humi, class = "")
-hypothesis(brm.family, H.hort_hypn, class = "")
-hypothesis(brm.family, H.hort_jone, class = "")
-hypothesis(brm.family, H.hort_lapi, class = "")
-hypothesis(brm.family, H.hort_mend, class = "")
-hypothesis(brm.family, H.hort_mont, class = "")
-hypothesis(brm.family, H.hort_muci, class = "")
-hypothesis(brm.family, H.hort_pasc, class = "")
-hypothesis(brm.family, H.hort_prat, class = "")
-hypothesis(brm.family, H.hort_psit, class = "")
-hypothesis(brm.family, H.hort_pyre, class = "")
-hypothesis(brm.family, H.hort_soro, class = "")
-hypothesis(brm.family, H.hort_wurf, class = "")
-
-hypothesis(brm.family, H.humi_hypn, class = "")
-hypothesis(brm.family, H.humi_jone, class = "")
-hypothesis(brm.family, H.humi_lapi, class = "")
-hypothesis(brm.family, H.humi_mend, class = "")
-hypothesis(brm.family, H.humi_mont, class = "")
-hypothesis(brm.family, H.humi_muci, class = "")
-hypothesis(brm.family, H.humi_pasc, class = "")
-hypothesis(brm.family, H.humi_prat, class = "")
-hypothesis(brm.family, H.humi_psit, class = "")
-hypothesis(brm.family, H.humi_pyre, class = "")
-hypothesis(brm.family, H.humi_soro, class = "")
-hypothesis(brm.family, H.humi_wurf, class = "")
-
-hypothesis(brm.family, H.hypn_jone, class = "")
-hypothesis(brm.family, H.hypn_lapi, class = "")
-hypothesis(brm.family, H.hypn_mend, class = "")
-hypothesis(brm.family, H.hypn_mont, class = "")
-hypothesis(brm.family, H.hypn_muci, class = "")
-hypothesis(brm.family, H.hypn_pasc, class = "")
-hypothesis(brm.family, H.hypn_prat, class = "")
-hypothesis(brm.family, H.hypn_psit, class = "")
-hypothesis(brm.family, H.hypn_pyre, class = "")
-hypothesis(brm.family, H.hypn_soro, class = "")
-hypothesis(brm.family, H.hypn_wurf, class = "")
-
-hypothesis(brm.family, H.jone_lapi, class = "")
-hypothesis(brm.family, H.jone_mend, class = "")
-hypothesis(brm.family, H.jone_mont, class = "")
-hypothesis(brm.family, H.jone_muci, class = "")
-hypothesis(brm.family, H.jone_pasc, class = "")
-hypothesis(brm.family, H.jone_prat, class = "")
-hypothesis(brm.family, H.jone_psit, class = "")
-hypothesis(brm.family, H.jone_pyre, class = "")
-hypothesis(brm.family, H.jone_soro, class = "")
-hypothesis(brm.family, H.jone_wurf, class = "")
-
-hypothesis(brm.family, H.lapi_mend, class = "")
-hypothesis(brm.family, H.lapi_mont, class = "")
-hypothesis(brm.family, H.lapi_muci, class = "")
-hypothesis(brm.family, H.lapi_pasc, class = "")
-hypothesis(brm.family, H.lapi_prat, class = "")
-hypothesis(brm.family, H.lapi_psit, class = "")
-hypothesis(brm.family, H.lapi_pyre, class = "")
-hypothesis(brm.family, H.lapi_soro, class = "")
-hypothesis(brm.family, H.lapi_wurf, class = "")
-
-hypothesis(brm.family, H.mend_mont, class = "")
-hypothesis(brm.family, H.mend_muci, class = "")
-hypothesis(brm.family, H.mend_pasc, class = "")
-hypothesis(brm.family, H.mend_prat, class = "")
-hypothesis(brm.family, H.mend_psit, class = "")
-hypothesis(brm.family, H.mend_pyre, class = "")
-hypothesis(brm.family, H.mend_soro, class = "")
-hypothesis(brm.family, H.mend_wurf, class = "")
-
-hypothesis(brm.family, H.mont_muci, class = "")
-hypothesis(brm.family, H.mont_pasc, class = "")
-hypothesis(brm.family, H.mont_prat, class = "")
-hypothesis(brm.family, H.mont_psit, class = "")
-hypothesis(brm.family, H.mont_pyre, class = "")
-hypothesis(brm.family, H.mont_soro, class = "")
-hypothesis(brm.family, H.mont_wurf, class = "")
-
-hypothesis(brm.family, H.muci_pasc, class = "")
-hypothesis(brm.family, H.muci_prat, class = "")
-hypothesis(brm.family, H.muci_psit, class = "")
-hypothesis(brm.family, H.muci_pyre, class = "")
-hypothesis(brm.family, H.muci_soro, class = "")
-hypothesis(brm.family, H.muci_wurf, class = "")
-
-hypothesis(brm.family, H.pasc_prat, class = "")
-hypothesis(brm.family, H.pasc_psit, class = "")
-hypothesis(brm.family, H.pasc_pyre, class = "")
-hypothesis(brm.family, H.pasc_soro, class = "")
-hypothesis(brm.family, H.pasc_wurf, class = "")
-
-hypothesis(brm.family, H.prat_psit, class = "")
-hypothesis(brm.family, H.prat_pyre, class = "")
-hypothesis(brm.family, H.prat_soro, class = "")
-hypothesis(brm.family, H.prat_wurf, class = "")
-
-hypothesis(brm.family, H.psit_pyre, class = "")
-hypothesis(brm.family, H.psit_soro, class = "")
-hypothesis(brm.family, H.psit_wurf, class = "")
-
-hypothesis(brm.family, H.pyre_soro, class = "")
-hypothesis(brm.family, H.pyre_wurf, class = "")
-
-hypothesis(brm.family, H.soro_wurf, class = "")
-```
-
-# 2. Having established the bias of BB foraging, let's zoom in on which morphotypes and families were over/under-represented
-## 2.1 Calculate bias per plant speces/ktype/family
-
-Deciding exactly how to represent bias turns out to be rather tricky. I am trying out a few options right now.
-
-1.
-Bias could  be trated simply as the Euclidean difference between prop.flower.cover and prop.visits.
-This has the effect of weighting bias by flower.cover, particularly negative bias, since an abundant plant can have a greater negative bias (because it is farther from 0) than a rare plant.
-Euclidean bias is bounded between -1 and 1.
-Its symmetry is a nice property that sets it apart from proportional or fold-change bias.
-If, for whatever reason, I decided that a log-transformation of this measure were desirable, I think I could simply add a constant of 1 to all values, which would move the boundary up to (0, 2).
-I think the biggest problem with measuring bias in this way is that it makes no distinction between a plant that was visited a little and one that was visited not at all, and that could be an important biological difference.
-For example, consider two plants, one with a proportional abundance = 0.01 and proportional visitation = 0.01, and the second with proportional abundance = 0.01 and proportional visitation = 0.00.
-These will look identical in our analysis, but it is possible that the unvisited plant would not have been visited no matter how high its abundance was.
-**For starters, though, I am going to proceed with this metric because it is the simplest.
-And because DJ thought it was a good idea.**
-
-2.
-Alternatively, bias could be calculated as a simple proportion: prop.visits/prop.flower.cover.
-If anything, this might tend to reverse the weighting, since a small Euclidean change in a rare species would be a large proportional change.
-Proportional change has a lower boundary of 0 and no upper boundary.
-This approach highlights cases of positive bias, but it is even worse than euclidean bias in terms of obscuring the distinction between plants visited slightly and those visited not at all.
-Worse in a sense.
-In another sense, it is better than euclidean bias, because with proportional visitation a plant visited 0 times is at least numerical distinguishable from one visited 1 time (visited plants will always have non-zero proportional bias).
-
-3.
-Similarly to proportional bias, I could calculate bias as fold change.
-This converges to be virtually identical to proportional change as bias becomes more positive, but fold change has a lower bound of -1 instead of 0.
-
-4.
-If the distinction between plants that were visited and those that were not is what we're most interested in, we could simply treat bias as binary.
-This, of course, comes at the price of not being able to distinguish a plant that received 0.01 proportional visitation and one that recevied  0.99 proportional visitation, and this distinction is arguably important.
-
-5.
-An interesting option is to first separate visited from unvisited plants, set the bias of unvisited plants to 0, and then exponentiate the euclidean bias of all visited plant species.
-This has some desirable properties.
-First, it clearly distinguished visited plants from unvisited ones.
-Second, it assigns plants visited perfectly in proportion to their abundance a sensible bias of 1 (just like proportional bias).
-Thus, plants with a psotive bias are > 1 (bounded at 2.718282 -- okay, so that's a bit inelegant), those with a negative bias are < 1 (bounded at 0.367895), and those not visited at all = 0.
-
-6.
-The problem with the option above is that is adds weight to positive bias.
-Another approach is to separate the unvisited plants from the visited, set unvisited bias = 0, and visited bias = prop.visits - prop.flower.cover + 1.
-The +1 serves to make the distribution positive, with 1 represening perfect proportionality between visitation and floral abundance.
-This creates a symmetrically bounded distribution between 0 (not inclusive) and 2 (inclusive), plus the true zeros from the unvisited plants.
-It feels a bit...slapped together, though.
-
-```{r}
-# Function for calculating bias per group (= criterion)
-biaseval_group <- function(x, criterion) {
-
-  criterion <- enquo(criterion)
-
-  x %>%
-    mutate(criterion = factor(!!criterion)) %>% # having this as a factor will help with modeling later on
-    group_by(site, date, bb.sp, criterion) %>%
-    summarize(flower.cover = sum(flower.cover), # get raw floer cover and visits by group
-              visits = sum(visits)) %>%
-    group_by(site, date, bb.sp) %>%
-    mutate(prop.flower.cover = flower.cover/sum(flower.cover), # get proportional flower cover and visits by group
-           prop.visits = visits/sum(visits)) %>%
-    mutate(bias = prop.visits - prop.flower.cover, # bias as difference
-           bias.prop = prop.visits/prop.flower.cover, # bias as proportion
-           bias.fold = bias/prop.flower.cover, # bias as fold change
-           bias.bin = if_else(bias > 0, 1, 0), # bias as binary
-           bias.exp = if_else(visits == 0, 0, exp(prop.visits - prop.flower.cover)),  # bias as exponentiated difference
-           bias.split = if_else(visits == 0, 0, (prop.visits - prop.flower.cover + 1))) %>%
-    mutate(yday = yday(date)) %>%
-    select(site, date, yday, bb.sp, criterion, 
-           bias, bias.prop, bias.fold, bias.bin, bias.exp, bias.split,
-           visits, flower.cover, prop.visits, prop.flower.cover)
-}
-
-bias_group_plantsp <- biaseval_group(visit_cover, plant.sp) %>%
-  left_join(site_data) %>%
-  rename(group = criterion) %>%
-  mutate(criterion = rep("plant.sp", n()))
-
-bias_group_ktype <- biaseval_group(visit_cover, k.type.ss) %>%
-  left_join(site_data) %>%
-  rename(group = criterion) %>%
-  mutate(criterion = rep("ktype", n()))
-
-bias_group_family <- biaseval_group(visit_cover, plant.family) %>%
-  left_join(site_data) %>%
-  rename(group = criterion) %>%
-  mutate(criterion = rep("plant.family", n()))
-
-bias_group <- bind_rows(bias_group_plantsp, bias_group_ktype, bias_group_family)
-
-ggplot(bias_group, aes(prop.flower.cover, prop.visits)) +
-  geom_point(alpha = 0.1) +
-  #geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
-  geom_smooth(method = "lm") +
-  geom_abline(slope = 1, color = "red", linetype = "dashed") +
-  facet_wrap(~criterion)
-
-ggplot(bias_group, aes(log(prop.flower.cover), prop.visits)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth() +
-  facet_wrap(~criterion)
-
-ggplot(filter(bias_group, criterion == "ktype"), aes(prop.flower.cover, prop.visits)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth() +
-  geom_abline(slope = 1, color = "red", linetype = "dashed") +
-  facet_wrap(~group)
-```
-
-## 2.2 Visualize
-```{r}
-# Check the distribution of bias
-ggplot(bias_group_ktype, aes(bias)) +
-  geom_histogram()
-
-ggplot(bias_group_ktype, aes(bias)) +
-  geom_histogram() +
-  facet_wrap(~criterion)
-
-ggplot(bias_group_family, aes(bias)) +
-  geom_histogram()
-
-ggplot(bias_group_family, aes(bias)) +
-  geom_histogram() +
-  facet_wrap(~criterion)
-
-# Bias by Kugler morphotype
-ggplot(bias_ktype, aes(criterion, bias)) +
-  geom_boxplot()
-
-ggplot(bias_ktype, aes(criterion, bias)) +
-  geom_boxplot() +
-  facet_wrap(~bb.sp)
-
-ggplot(bias_ktype, aes(bb.sp, bias)) +
-  geom_boxplot() +
-  facet_wrap(~criterion)
-
-ggplot(bias_ktype, aes(elev.mean, bias)) +
-  #geom_point() +
-  geom_smooth(method = gam, formula = y ~ s(x, k = 5)) +
-  facet_wrap(~criterion)
-
-# Bias by plant family
-ggplot(bias_family, aes(criterion, bias)) +
-  geom_boxplot()
-
-ggplot(bias_family, aes(criterion, bias)) +
-  geom_boxplot() +
-  facet_wrap(~bb.sp)
-
-ggplot(bias_family, aes(elev.mean, bias)) +
-  #geom_point() +
-  geom_smooth(method = gam, formula = y ~ s(x, k = 5)) +
-  facet_wrap(~criterion)
-
-ggplot(bias_family, aes(elev.mean, bias)) +
-  #geom_point() +
-  geom_smooth(method = gam, formula = y ~ s(x, k = 5)) +
-  facet_wrap(~criterion)
-
-```
-
-## 2.3 Explaining representation by floral traits, BB spp, and elevation (and maybe time)
-```{r}
-repmod.ktype <- brm(bias ~ elev.mean + (1 | criterion), 
-                    data = bias_ktype)
-
-repmod.ktype <- brm(bias ~ criterion, 
-                    data = bias_ktype)
-parnames(repmod.ktype)
-mcmc_plot(repmod.ktype)
-conditional_effects(repmod.ktype)
-
-H.K5_K6 <- paste("b_criterion5 = b_criterion6")
-hypothesis(repmod.ktype, H.K5_K6, class = "")
-```
-
-# How do species vary in their biases
-```{r}
-bias_matrix.plantsp <- bias_group_plantsp %>%
-  group_by(bb.sp, criterion) %>%
-  summarize(bias = mean(bias)) %>%
-  pivot_wider(values_from = bias, names_from = criterion, values_fill = 0) %>%
-  column_to_rownames("bb.sp")
-
-bias_matrix.ktype <- bias_group_ktype %>%
-  group_by(bb.sp, criterion) %>%
-  summarize(bias = mean(bias)) %>%
-  pivot_wider(values_from = bias, names_from = criterion, values_fill = 0) %>%
-  column_to_rownames("bb.sp")
-
-library(vegan)
-
-pca.plantsp <- rda(bias_matrix.plantsp)
-plot(pca.plantsp, type = "t")
-
-pca.ktype <- rda(bias_matrix.ktype)
-plot(pca.ktype)
-
-```
-
-# 3. Can we explain bias by BB-Fl trait associations? (RLQ + fourth corner analysis)
-## 3.1
-## 3.2 Do BB-FL trait associations vary with elevation?
-
-# 4. Can we explain BB-FL trait associations in terms of floral reward or handling time? We can use Neumayer and Paulus' (1999) data. 
-## 4.1 
-## 4.2 Do there appear to be suboptimalities in BB-FL trait associations?
-That is, are trait classes over/under-represented relative to their reward/efficiency?
-If so, how might explain such suboptimalities?
-  - Innate bias trumps learning?
-  - Ecological modulators are involved, e.g. competition or distribution (see Sowig 1989)

### 2021-01-11 writing/manuscript_v4.Rmd
syncing to office desktop, got rid of lots of old scripts and output

@@
 ---
-title: Bumble bee foraging bias
+title: "Bumble bee foraging bias"
 author:
@@
 bibliography: refs.bib
-output: html_document
+output:
+  pdf_document: default
+  html_document: default
 ---
@@
 
-*mountain slopes give us floristic variation and abiotic variation while controlling for genetic variation within BB species, thus enabling us to study the lability of bias and partition it between innate and learned components*
+The topic of floral preference in bumble bees has been approached via both laboratory experimentation [e.g.
+@Kugler1943-uy; @Raine2007-nq] and observational field study [e.g.
+Heinrich1976-wc; @Goulson2008-zy].
+The former approach enables more confident causal inference with respect to the mechanisms of floral choice, but it suffers all the usual limitations of biological realism associated with experimental control, and the higher-order effects of floral choices on network structure cannot be observed meaningfully under laboratory conditions.
+Observational field studies, however, have tended to be anecdotal in scope, typically describing bumble bee foraging patterns in particular localities, over short time frames, and with limited floristic data to enable a distinction between true foraging bias and the neutral effects of floral abundance [though see @Neumayer1999-il].
+The steep elevation gradients of mountain slopes provide an opportunity for more powerful observational study of bumble foraging because they generate broad gradients of floristic and abiotic conditions while controlling for confounding effects of geographic variation, such as genotypic differentiation within bumble bee species.
+When bumble bee visitation in such environments is sampled with sufficient spatial and temporal replication and complemented with quantitative floral surveying, inference can be made regarding both the trait interactions that underly floral biases and the network structures that result from them.
+Mountain habitats are also arguably the most relevant places to study bumble bee ecology, since bumble bees are adapted to cool climates, and their global distribution is concentrated in areas of high elevation or high latitude.
+In such environments, bumble bees tend to dominate the pollinator community, and local flora often depend primarily on bumble bee visitation for pollination <!-- This paragraph needs to be beefed up a bit, but it at least sketches out the points I want to make -->.
 
-The study bias in bumble bee foraging
+# Methods
 
-Local elevation gradients along steep mountain slopes provide the opportunity to study broad environmental variation while controlling for potential confounding effects associated with geographic distance.
+## Field sampling
+
+The study was conducted in Berchtesgaden National Park (47.55°N, 12.92°E), located in the Northern Limestone Alps of southeast Germany.
+The landscape is characterized by mountain pastures mainly surrounded by coniferous forests.
+We selected 25 study sites (60 x 60 m) on mountain pastures at elevations ranging from 641-2032 m above sea level.
+Fourteen of these pastures are extensively grazed by cattle or sheep, three are mowed for hay production, and eight have been abandoned throughout the last century and are no longer subject to any human management.
+The sites sampled in this study are the same as those used in a series of previous studies [@Hoiss2012-sn; @Hoiss2012-kr; Hoiss2015-gv], but the data reported here are otherwise independent.
+
+Sampling consisted of repeated visits to each study site at approximately weekly intervals.
+Samples were only collected on dry days when the air temperature was at least 6°C.
+During each visit, bumble bees and their activity were recorded during a 50-minute transect walk.
+Bumblebees observed on or a given flower were counted as floral visitors.
+Bumble bee queens were identified to species level in the field, whereas workers and males were stored in individually labeled tubes in the freezer for later identification in the laboratory after @Amiet1996-jz.
+
+## Data analysis
+
+Prior to analysis, we aligned each bumble bee visitation sample with corresponding floral survey data.
+For most samples, floral surveying was conducted jointly with bumble bee visitation observation.
+In cases where bumble bee visitation was sampled without a joint floral survey, the bumble bee visitation data was paired with the nearest available floral survey date within a maximum distance of 7 days.
+Bumble bee visitation samples lacking corresponding floral survey data within 7 days were omitted from further analysis.
+In a few cases where more than one bumble bee visitation date was paired with a given floral survey date, the bumble bee visitation data were pooled.
+
+To quantify cumulative bias per bumble bee species, we used the discrimination (d’) metric of @Bluthgen2006-if, which represents the net deviation of a species’ visitation from the predictions of a null model in which the probability of visitation is purely a function of floral abundance.
+
+# Results
+
+# Discussion
 

### 2021-01-11 analyses/trash_can.R

@@ -1,2130 +1,0 @@
-
-boot_net_sp_FL <- function(net, n, iter) {
-  
-  out <- data.frame()
-  
-  for(i in 1:iter) {
-    temp <- net %>%
-      dplyr::select(site, date) %>%
-      unique() %>%
-      group_by(site) %>%
-      sample_n(n) %>%
-      left_join(net) %>%
-      group_by(site, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
-      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
-      group_by(site) %>%
-      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
-      ungroup() %>%
-      filter(freq > 1) %>% # remove singletons
-      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
-      data.frame() %>% # convert to data frame
-      frame2webs() %>% # convert to to bipartite web
-      map(specieslevel, level = "lower", index = c("d", "species strength", "proportional generality")) %>%
-      map(rownames_to_column) %>%
-      map(as_tibble) %>%
-      bind_rows(.id = "site") %>% # Collapse list into big data frame
-      rename(sp = rowname) %>%
-      mutate(iteration = rep(i, n()))
-    
-    out <- bind_rows(out, temp)
-  } 
-  out <- out %>%
-    gather(metric, value, -c(sp, site, iteration)) %>% # get mean value across iterations for each metric
-    group_by(sp, site, metric) %>%
-    summarize(value = mean(value)) %>%
-    left_join(site_data) %>%
-    left_join(fl_traits, by = c("sp" = "plant.sp"))
-}
-
-
-#### Rarefied network analysis by colony phase
-```{r, include=FALSE}
-# network level version 2
-boot_net_nt2 <- function(net, n, iter) {
-  
-  out <- data.frame()
-  
-  t_filter <- net %>%
-    dplyr::select(site, date, phase, year) %>%
-    unique() %>%
-    group_by(site, phase, year) %>%
-    mutate(transects = n()) %>%
-    filter(transects >= n)
-  
-  for(i in 1:iter) {
-    temp <- net %>%
-      semi_join(t_filter) %>%
-      dplyr::select(site, date, phase, year) %>%
-      unique() %>%
-      group_by(site, phase, year) %>%
-      sample_n(n) %>%
[…]

### 2021-01-11 analyses/visitation_bias.Rmd
back to laptop because rstan won't run on Windows...

@@
   left_join(fl_abundance) %>% # add fl abundance
-  mutate(criterion = factor(rep("plant.sp", n()))) %>%
-  left_join(no_plant.sp)
+  mutate(criterion = factor(rep("plant.sp", n())))
 
@@
   left_join(fl_abundance) %>% # add fl abundance
-  mutate(criterion = factor(rep("ktype", n()))) %>%
-  left_join(no_ktype)
+  mutate(criterion = factor(rep("ktype", n())))
 
@@
   left_join(fl_abundance) %>% # add fl abundance
-  mutate(criterion = factor(rep("plant.family", n()))) %>%
-  left_join(no_plant.family)
+  mutate(criterion = factor(rep("plant.family", n())))
 
 # Collated
-bias <- bind_rows(bias_plantsp.sitedate, 
-                  bias_ktype.sitedate, 
-                  bias_family.sitedate) %>%
+bias <- bind_rows(bias_plantsp, 
+                  bias_ktype, 
+                  bias_family) %>%
   mutate(bb.sp = factor(bb.sp),

### 2021-01-13 code/visitation_bias.Rmd
restructured project directory; this will break some paths, but easy to fix, and clearer now

--- 2021-01-11 analyses/visitation_bias.Rmd
@@
 library(lme4)
+library(ggeffects)
+library(sjPlot)
 #library(inla)
@@
 # 1.
-Is bumble bee foraging biased?
-How do taxonomic, phylogenetic, and morphological bias compare -- NO, can't analyze this, these should be considered separate repsonse variables and analyzed in separate models?
-Are some BB species more biased than others?
+How does overall bias (Blüthgen's d') vary across species and through elev*time?
+For now, we will focus on bias at the level of plant species.
+We'll start considering traits later on when our goal is to explain patterns of bias.
 ## 1.1 Calculate Blüthgen's d' (i.e. visitation bias)
@@
 
-ggplot(bias_plantsp.sitedate, aes(log(bb.sp.prop.abund), dprime)) +
+ggplot(bias, aes(log(bb.sp.prop.abund), dprime)) +
   geom_point() +
@@
 
-## 1.3 Multi-level model: bias ~ elevation + (1 + elevation | bb.sp)
+## 1.3 Multi-level model: bias ~ yday * elevation + (1 + elevation | bb.sp)
 ```{r}
-# brms
-brm.ktype1 <- brm(dprime ~ yday + elev + (1 + elev | bb.sp), 
+# I am interested both in the variation of bias through elevation and time and in the intercept effect of each bumble bee species
+
+# First, let's look at the distribution of bias per BB sp
+ggplot(filter(bias, criterion == "plant.sp"), aes(dprime)) +
+  geom_histogram() +
+  facet_wrap(~bb.sp)
+
+# Now, let's try to get a sense for the shape of the response to elevation and time; we might expect to find nonlinearities here
+ggplot(filter(bias, criterion == "plant.sp"), aes(elev, dprime)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~bb.sp)
+
+ggplot(filter(bias, criterion == "plant.sp"), aes(yday, dprime)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~bb.sp)
+
+# After inspecting these plots, I see no reason to indulge in wiggly models.
+
+# This model includes a global interaction term for yday and elevation, with a varying intercept for bb.sp and a varying slow for elev | bb.sp.
+I think this should give me all the estimation I'm interested in.
+brm.plantsp11 <- brm(dprime ~ yday * elev + (1 + elev | bb.sp), 
                   control = list(max_treedepth = 10),
-                  cores = 3,
-                  iter = 2000,
-                  file = "./output/brm_ktype1",
-                  data = filter(bias, criterion ==  "ktype"))
-
-summary(brm.ktype1)
-plot(brm.ktype1)
-parnames(brm.ktype1)
-mcmc_plot(brm.ktype1, pars = c("^r_"))
-mcmc_plot(brm.ktype1, pars = c("^b_"))
-mcmc_plot(brm.ktype1)
-conditional_effects(brm.ktype1)
-pp_check(brm.ktype1)
-
-brm.ktype2 <- brm(dprime ~ 1 + yday + elev + (1 | bb.sp), 
+                  cores = 4,
+                  iter = 4000,
+                  #family = "skew_normal",
+                  file = "./output/brm_plantsp11",
+                  data = filter(bias, criterion ==  "plant.sp"))
+
+summary(brm.plantsp11)
+prior_summary(brm.plantsp11)
+plot(brm.plantsp11)
+parnames(brm.plantsp11)
+mcmc_plot(brm.plantsp11, pars = c("^r_"))
+mcmc_plot(brm.plantsp11, pars = c("elev"))
+mcmc_plot(brm.plantsp11, pars = c("^b_"))
+mcmc_plot(brm.plantsp11)
+conditional_effects(brm.plantsp11)
+pp_check(brm.plantsp11)
+bayes_R2(brm.plantsp11)
+hypothesis(brm.plantsp11, "b_elev = 0", class = NULL)
+
+brm.plantsp1 <- brm(dprime ~ yday * elev + (1 | bb.sp), 
                   control = list(max_treedepth = 10),
@@
                   iter = 2000,
-                  file = "./output/brm_ktype2b",
-                  data = filter(bias, criterion ==  "ktype"))
-
-summary(brm.ktype2)
-plot(brm.ktype2)
-parnames(brm.ktype2)
-mcmc_plot(brm.ktype2, pars = c("^r_"))
-mcmc_plot(brm.ktype2, pars = c("^b_"))
-mcmc_plot(brm.ktype2)
-conditional_effects(brm.ktype2)
-pp_check(brm.ktype2)
-
-brm.ktype3 <- brm(dprime ~ 1 + (1 | bb.sp), 
+                  file = "./output/brm_plantsp1",
+                  data = filter(bias, criterion ==  "plant.sp"))
+
+summary(brm.plantsp1)
+prior_summary(brm.plantsp1)
+plot(brm.plantsp1)
+parnames(brm.plantsp1)
+mcmc_plot(brm.plantsp1, pars = c("^r_"))
+mcmc_plot(brm.plantsp1, pars = c("elev"))
+mcmc_plot(brm.plantsp1, pars = c("^b_"))
+mcmc_plot(brm.plantsp1)
+conditional_effects(brm.plantsp1)
+pp_check(brm.plantsp1)
+bayes_R2(brm.plantsp1)
+hypothesis(brm.plantsp1, "b_elev = 0", class = NULL)
+
+brm.plantsp2 <- brm(dprime ~ yday + elev + (1 | bb.sp), 
                   control = list(max_treedepth = 10),
@@
                   iter = 2000,
-                  family = "poisson",
-                  file = "./output/brm_ktype3",
-                  data = filter(bias, criterion ==  "ktype"))
-
-summary(brm.ktype3)
-plot(brm.ktype3)
-parnames(brm.ktype3)
-mcmc_plot(brm.ktype3, pars = c("^r_"))
-mcmc_plot(brm.ktype3, pars = c("^b_"))
-mcmc_plot(brm.ktype3)
-conditional_effects(brm.ktype3)
-pp_check(brm.ktype3)
-
-# lme4
-library(lme4)
-library(ggeffects)
-
-lmer.ktype <- lmer(dprime ~ 1 + elev:yday + (1 | bb.sp), 
-                  data = filter(bias, criterion ==  "ktype"))
-summary(lmer.ktype)
-pred.ktype_yday <- ggpredict(lmer.ktype, "yday")
-pred.ktype_elev <- ggpredict(lmer.ktype, "elev")
-#pred.ktype_yday_ran <- ggpredict(lmer.ktype, "yday", type = "random")
-#pred.ktype_elev_ran <- ggpredict(lmer.ktype, "elev", type = "random")
-pred.ktype_bbsp <- ggpredict(lmer.ktype, "bb.sp", type = "random")
-
-plot(pred.ktype_yday)
-plot(pred.ktype_elev)
-#plot(pred.ktype_yday_ran)
-#plot(pred.ktype_elev_ran)
-plot(pred.ktype_bbsp)
-
-# mgcv
-gam.ktype <- bam(dprime ~ s(bb.sp, bs = "re") + 
-                   t2(yday, elev.mean, bs = c("tp", "gp")),
-                 method = "fREML",
-                 discrete = TRUE,
-                 data = filter(bias, criterion ==  "ktype")) %>% getViz()
-
-summary(gam.ktype)
-plot(gam.ktype, allTerms = TRUE)
-
-# inla
-library(INLA)
-
-inla.ktype <- inla(dprime ~ elev.mean, 
-                   family = "gaussian",
-                   data = filter(bias, criterion ==  "ktype"))
-
-inla.pen <- inla(dprime ~ 1 + elev.mean,
-  data = filter(bias, criterion ==  "ktype"), control.predictor = list(compute = TRUE))
-summary(inla.pen)
+                  file = "./output/brm_plantsp2",
+                  data = filter(bias, criterion ==  "plant.sp"))
+
+summary(brm.plantsp2)
+prior_summary(brm.plantsp2)
+plot(brm.plantsp2)
+parnames(brm.plantsp2)
+mcmc_plot(brm.plantsp2, pars = c("^r_"))
+mcmc_plot(brm.plantsp2, pars = c("elev"))
+mcmc_plot(brm.plantsp2, pars = c("^b_"))
+mcmc_plot(brm.plantsp2)
+conditional_effects(brm.plantsp2)
+pp_check(brm.plantsp2)
+bayes_R2(brm.plantsp2)
+hypothesis(brm.plantsp2, "b_elev = 0", class = NULL)
+
+# This, I think, would also be a legitimate way to specify the model, but it would not tell us anything about elevation or time. 
+brm.plantsp0 <- brm(dprime ~ 1 + (1 | bb.sp) + (1 | site),
+                  control = list(max_treedepth = 10),
+                  cores = 4,
+                  iter = 2000,
+                  file = "./output/brm_plantsp0",
+                  data = filter(bias, criterion ==  "plant.sp"))
+
+summary(brm.plantsp0)
+plot(brm.plantsp0)
+parnames(brm.plantsp0)
+mcmc_plot(brm.plantsp0, pars = c("^r_"))
+mcmc_plot(brm.plantsp0, pars = c("^b_"))
+pp_check(brm.plantsp0)
+bayes_R2(brm.plantsp0)
 ```
@@
 1.
-Bias could  be trated simply as the Euclidean difference between prop.flower.cover and prop.visits.
+Bias could  be treated simply as the Euclidean difference between prop.flower.cover and prop.visits.
 This has the effect of weighting bias by flower.cover, particularly negative bias, since an abundant plant can have a greater negative bias (because it is farther from 0) than a rare plant.
@@
 ```{r}
-bias_group_plantsp <- biaseval_group(visit_cover, plant.sp) %>%
+representation <- biaseval_group(visit_cover, plant.sp) %>%
+  rename(plant.sp = criterion) %>%
   left_join(site_data) %>%
-  rename(group = criterion) %>%
-  mutate(criterion = rep("plant.sp", n()))
-
-bias_group_ktype <- biaseval_group(visit_cover, k.type.ss) %>%
+  left_join(fl_traits) %>%
+  mutate(k.type.ss = factor(k.type.ss)) %>%
+  select(site, date, elev.mean, yday, bb.sp, plant.sp, k.type.ss, 
+         prop.visits, prop.flower.cover, bias, visits, 
+         flower.cover, plant.genus, plant.family)
+
+# 
+representation_ktype <- biaseval_group(visit_cover, k.type.ss) %>%
   left_join(site_data) %>%
@@
   mutate(criterion = rep("ktype", n()))
-
-bias_group_family <- biaseval_group(visit_cover, plant.family) %>%
-  left_join(site_data) %>%
-  rename(group = criterion) %>%
-  mutate(criterion = rep("plant.family", n()))
-
-bias_group <- bind_rows(bias_group_plantsp, bias_group_ktype, bias_group_family)
-
-ggplot(bias_group, aes(prop.flower.cover, prop.visits)) +
-  geom_point(alpha = 0.1) +
-  #geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
-  geom_smooth(method = "lm") +
-  geom_abline(slope = 1, color = "red", linetype = "dashed") +
-  facet_wrap(~criterion)
-
-ggplot(bias_group, aes(log(prop.flower.cover), prop.visits)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth() +
-  facet_wrap(~criterion)
-
-ggplot(filter(bias_group, criterion == "ktype"), aes(prop.flower.cover, prop.visits)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth() +
-  geom_abline(slope = 1, color = "red", linetype = "dashed") +
-  facet_wrap(~group)
+# 
+# bias_group_family <- biaseval_group(visit_cover, plant.family) %>%
+#   left_join(site_data) %>%
+#   rename(group = criterion) %>%
+#   mutate(criterion = rep("plant.family", n()))
+# 
+# bias_group <- bind_rows(bias_group_plantsp, bias_group_ktype, bias_group_family)
+# 
+# ggplot(bias_group, aes(prop.flower.cover, prop.visits)) +
+#   geom_point(alpha = 0.1) +
+#   #geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
+#   geom_smooth(method = "lm") +
+#   geom_abline(slope = 1, color = "red", linetype = "dashed") +
+#   facet_wrap(~criterion)
+# 
+# ggplot(bias_group, aes(log(prop.flower.cover), prop.visits)) +
+#   geom_point(alpha = 0.1) +
+#   geom_smooth() +
+#   facet_wrap(~criterion)
+# 
+# ggplot(filter(bias_group, criterion == "ktype"), aes(prop.flower.cover, prop.visits)) +
+#   geom_point(alpha = 0.1) +
+#   geom_smooth() +
+#   geom_abline(slope = 1, color = "red", linetype = "dashed") +
+#   facet_wrap(~group)
 ```
@@
 ```{r}
-repmod.ktype <- brm(bias ~ elev.mean + (1 | criterion), 
-                    data = bias_ktype)
-
-repmod.ktype <- brm(bias ~ criterion, 
-                    data = bias_ktype)
-parnames(repmod.ktype)
-mcmc_plot(repmod.ktype)
-conditional_effects(repmod.ktype)
-
-H.K5_K6 <- paste("b_criterion5 = b_criterion6")
-hypothesis(repmod.ktype, H.K5_K6, class = "")
-```
-
-# How do species vary in their biases
+ggplot(representation, aes(k.type.ss, bias)) +
+  geom_violin()
+
+ggplot(representation_ktype, aes(group, bias)) +
+  geom_boxplot()
+
+ggplot(representation, aes(bias)) +
+  geom_histogram() +
+  facet_wrap(~k.type.ss)
+
+ggplot(filter(representation, visits > 0), aes(bias)) +
+  geom_histogram() +
+  facet_wrap(~k.type.ss)
+
+ggplot(filter(representation, prop.flower.cover > 0.05), aes(k.type.ss, bias)) +
+  geom_boxplot()
+
+ggplot(representation, aes(prop.flower.cover, prop.visits)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~k.type.ss)
+
+summary(lm(representation ~ k.type.ss, data = bias))
+
+brm_representation0 <- brm(bias ~ 1 + (1 | group),
+                 control = list(max_treedepth = 10),
+                 cores = 4,
+                 iter = 2000,
+                 file = "./output/brm_representation1",
+                 data = representation_ktype)
+
+brm_representation1 <- brm(bias ~ 1 + (1 | k.type.ss),
+                 control = list(max_treedepth = 10),
+                 #cores = 4,
+                 cores = 4,
+                 chains = 4,
+                 #iter = 2000,
+                 iter = 2000,
+                 file = "./output/brm_representation1",
+                 data = representation)
+
+summary(brm_representation1)
+plot(brm_representation1)
+parnames(brm_representation1)
+mcmc_plot(brm_representation1, pars = c("^r_"))
+#mcmc_plot(representation.ktype1, pars = c("elev"))
+mcmc_plot(brm_representation1, pars = c("^b_"))
+mcmc_plot(brm_representation1)
+#conditional_effects(representation.ktype1)
+pp_check(brm_representation1)
+bayes_R2(brm_representation1)
+
+hypothesis(representation.ktype1, "r_group[5,Intercept] = r_group[6,Intercept]", class = "")
+hypothesis(representation.ktype1, "r_group[1,Intercept] = r_group[0,Intercept]", class = "")
+
+# One interesting way to think about bias is to focus on the slope of prop.visits ~ prop.flower.cover. 
+# We could analyze this with a varying slope model
+
+# Binomial ggplot smoother
+binomial_smooth <- function(...) {
+  geom_smooth(method = "glm", method.args = list(family = "quasibinomial"), se = FALSE, ...)
+}
+
+ggplot(representation, aes(sqrt(prop.flower.cover), sqrt(prop.visits))) +
+  geom_point() +
+  #geom_smooth(method = "lm") +
+  #binomial_smooth(formula = y ~ splines::ns(x, 3)) +
+  geom_smooth() +
+  #geom_smooth(method = "glm", method.args = list(family = "binomial")) +
+  facet_grid(k.type.ss~bb.sp)
+
+brm_representation3 <- brm(prop.visits ~ prop.flower.cover + (1 + prop.flower.cover | k.type.ss),
+                 control = list(max_treedepth = 10),
+                 #cores = 4,
+                 cores = 2,
+                 chains = 2,
+                 #iter = 2000,
+                 iter = 1000,
+                 file = "./output/brm_representation3",
+                 data = representation)
+
+lmer_representation3 <- lmer(prop.visits ~ log(prop.flower.cover) + (1 + log(prop.flower.cover) | k.type.ss),
+                             data = representation)
+
+summary(lmer_representation3)
+plot(lmer_representation3)
+```
+
+# 3. How do species vary in their biases?
 ```{r}
-bias_matrix.plantsp <- bias_group_plantsp %>%
-  group_by(bb.sp, criterion) %>%
+rep_matrix.plantsp <- representation %>%
+  group_by(bb.sp, group) %>%
   summarize(bias = mean(bias)) %>%
-  pivot_wider(values_from = bias, names_from = criterion, values_fill = 0) %>%
+  pivot_wider(values_from = bias, names_from = group, values_fill = 0) %>%
   column_to_rownames("bb.sp")
@@
 bias_matrix.ktype <- bias_group_ktype %>%
-  group_by(bb.sp, criterion) %>%
+  group_by(bb.sp, group) %>%
   summarize(bias = mean(bias)) %>%
-  pivot_wider(values_from = bias, names_from = criterion, values_fill = 0) %>%
+  pivot_wider(values_from = bias, names_from = group, values_fill = 0) %>%
   column_to_rownames("bb.sp")
 
-library(vegan)
+traits <- bb_traits %>%
+  arrange(bb.sp)
 
@@
 plot(pca.plantsp, type = "t")
+envfit(pca.plantsp ~ pbl.w.class, data = traits)
+summary(ordisurf(pca.plantsp ~ pbl.w.class, data = traits))
 
 pca.ktype <- rda(bias_matrix.ktype)
-plot(pca.ktype)
-
+plot(pca.ktype, type = "t")
+envfit(pca.ktype ~ pbl.w.class, data = traits)
+summary(ordisurf(pca.ktype ~ pbl.w.class, data = traits))
+pca.ktype_pca1_traits <- pca.ktype$CA$u[,1] %>%
+  as.data.frame() %>%
+  rownames_to_column(var = "bb.sp") %>%
+  rename(PC1 = 2) %>%
+  left_join(bb_traits)
+
+ggplot(pca.ktype_pca1_traits, aes(pbl.w, PC1)) +
+  geom_point() +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 2))
+
+ggplot(pca.ktype_pca1_traits, aes(pbl.w.class, PC1)) +
+  geom_point() +
+  geom_smooth(method = "lm")
 ```

### 2021-01-13 writing/manuscript_v4.Rmd
restructured project directory; this will break some paths, but easy to fix, and clearer now

@@
 While the distribution of interaction strengths can be partially explained by the neutral effects of abundance, pollinators exhibit positive and negative biases toward available flora even when interaction strength is conditioned on floral abundance.
-This phenomenon is formalized by the discrimination (d’) metric of [@Bluthgen2006-if], but to avoid conflation with the metric *per se*, we will use the term bias in a more general sense to describe disproportionality between pollinator visitation and floral abundance.
-
-Taken as a focal phenomenon, the realized floral biases of pollinators can be interposed between their underlying causal mechanisms and their higher-order significance **(Figure 1)**.
-From the causal perspective, realized biases can be partitioned into innate and learned components.
+Taken as a focal phenomenon, the floral biases of pollinators can be interposed between their underlying causal mechanisms and their higher-order significance **(Figure 1)**.
+From the causal perspective, floral biases can be partitioned into innate and learned components.
 Both these components depend on interactions between the functional traits of a given pollinator and those of its potential floral hosts, manifest over ecological time scales with respect to learned biases and evolutionary time scales with respect to innate biases.
@@
 At an evolutionary level, the innate preferences of bumble bees for certain floral traits likely reflect historic associations between floral traits and floral rewards [@Raine2007-nq].
-These innate preferences, however, can be modulated or even nullified by learned patterns of reward based on individual foraging experience [@Kugler1943-uy; Maharaj2019-wm], which depend not only intrinsic floral properties but also ecological modulators, such as the spatial distribution of floral patches or competition with other bumble bees [@Ishii2013-sm; @Plowright1984-lc; @Sowig1989-ld].
+These innate preferences, however, can be modulated or even nullified by patterns of reward learned during foraging [@Kugler1943-uy; Maharaj2019-wm], which depend not only intrinsic floral properties but also ecological modulators, such as the spatial distribution of floral patches or competition with other bumble bees [@Ishii2013-sm; @Plowright1984-lc; @Sowig1989-ld].
 Thus, realized biases in bumble bee foraging are the labile product of bumble bee traits, floral traits, and ecological context.
@@
 Observational field studies, however, have tended to be anecdotal in scope, typically describing bumble bee foraging patterns in particular localities, over short time frames, and with limited floristic data to enable a distinction between true foraging bias and the neutral effects of floral abundance [though see @Neumayer1999-il].
-The steep elevation gradients of mountain slopes provide an opportunity for more powerful observational study of bumble foraging because they generate broad gradients of floristic and abiotic conditions while controlling for confounding effects of geographic variation, such as genotypic differentiation within bumble bee species.
-When bumble bee visitation in such environments is sampled with sufficient spatial and temporal replication and complemented with quantitative floral surveying, inference can be made regarding both the trait interactions that underly floral biases and the network structures that result from them.
+The steep elevation gradients of mountain slopes provide an opportunity for more powerful observational study of bumble foraging because they generate broad gradients of floristic and abiotic variation while controlling for confounding effects of geographic variation, such as genotypic differentiation within bumble bee species.
+When bumble bee visitation in such environments is sampled with sufficient spatial and temporal replication and complemented with quantitative floral surveying, inference can be made regarding both the trait interactions that underpin floral biases and the network structures that result from them.
 Mountain habitats are also arguably the most relevant places to study bumble bee ecology, since bumble bees are adapted to cool climates, and their global distribution is concentrated in areas of high elevation or high latitude.
@@
 For most samples, floral surveying was conducted jointly with bumble bee visitation observation.
-In cases where bumble bee visitation was sampled without a joint floral survey, the bumble bee visitation data was paired with the nearest available floral survey date within a maximum distance of 7 days.
+In cases where bumble bee visitation was sampled without a joint floral survey, the bumble bee visitation data was paired with the nearest available floral survey data within a maximum distance of 7 days.
 Bumble bee visitation samples lacking corresponding floral survey data within 7 days were omitted from further analysis.
@@
 
+This phenomenon is formalized by the discrimination (d’) metric of @Bluthgen2006-if, but to avoid conflation with the metric *per se*, we use the term bias in a general sense to describe disproportionality between pollinator visitation and floral abundance.
+
 # References

### 2021-01-13 code/data_proc.R

@@ -1,0 +1,463 @@
+library(tidyverse)
+library(lubridate)
+
+# Get the data cleaned up and properly organized
+### Floral survey data
+fl_survey <- read_delim("./raw_data/data_plantcover_forDoug_2020.01.17.csv", delim = "\t", 
+                           locale = locale(decimal_mark = ",", grouping_mark = ".")) %>%
+  na.omit() %>%
+  separate(col = species_plant, into = c("genus", "species", "w", "x", "y", "z"), sep = " ") %>%
+  unite(species_plant, genus, species, sep = " ") %>%
+  dplyr::select(-c(w,x,y,z)) %>%
+  mutate(species_plant = str_replace_all(species_plant, 
+                                         c("Arabis caerula" = "Arabis caerulea", # correct nomenclature typos and synonyms
+                                           "Arctostaphylos alpina" = "Arctous alpina",
+                                           "Calluna vulagris" = "Calluna vulgaris",
+                                           "Deschampsia caespitosa" = "Deschampsia cespitosa",
+                                           "Gymnadenia bifolia" = "Platanthera bifolia"))) %>%
+  separate(species_plant, c("genus", "species"), remove = FALSE) %>%
+  dplyr::select(year, dayofyear, site = site_name, snowcover, 
+                plant.sp = species_plant, plant.genus = genus, flower.cover = cover_all_flowers) %>%
+  mutate(temp.date = as_date(dayofyear),
+         day = day(temp.date),
+         month = month(temp.date),
+         date = make_date(year, month, day)) %>%
+  filter(plant.genus != "Huperzia") %>% # because I don't need the clubmoss
+  dplyr::select(-temp.date)
+  
+write_csv(fl_survey, "./processed_data/floral_survey.csv")
+
+### Visitation data
+network <- read_delim("./raw_data/data_bumblebee_forDoug_2020.04.17.csv", delim = ";", 
+                           locale = locale(decimal_mark = ",", grouping_mark = ".")) %>%
+  mutate(plant_sp_latin = str_replace_all(plant_sp_latin, 
+                                         c("Arabis caerula" = "Arabis caerulea", # correct nomenclature typos and synonyms
[…]

### 2021-01-13 data_proc.R

@@ -1,463 +1,0 @@
-library(tidyverse)
-library(lubridate)
-
-# Get the data cleaned up and properly organized
-### Floral survey data
-fl_survey <- read_delim("./raw_data/data_plantcover_forDoug_2020.01.17.csv", delim = "\t", 
-                           locale = locale(decimal_mark = ",", grouping_mark = ".")) %>%
-  na.omit() %>%
-  separate(col = species_plant, into = c("genus", "species", "w", "x", "y", "z"), sep = " ") %>%
-  unite(species_plant, genus, species, sep = " ") %>%
-  dplyr::select(-c(w,x,y,z)) %>%
-  mutate(species_plant = str_replace_all(species_plant, 
-                                         c("Arabis caerula" = "Arabis caerulea", # correct nomenclature typos and synonyms
-                                           "Arctostaphylos alpina" = "Arctous alpina",
-                                           "Calluna vulagris" = "Calluna vulgaris",
-                                           "Deschampsia caespitosa" = "Deschampsia cespitosa",
-                                           "Gymnadenia bifolia" = "Platanthera bifolia"))) %>%
-  separate(species_plant, c("genus", "species"), remove = FALSE) %>%
-  dplyr::select(year, dayofyear, site = site_name, snowcover, 
-                plant.sp = species_plant, plant.genus = genus, flower.cover = cover_all_flowers) %>%
-  mutate(temp.date = as_date(dayofyear),
-         day = day(temp.date),
-         month = month(temp.date),
-         date = make_date(year, month, day)) %>%
-  filter(plant.genus != "Huperzia") %>% # because I don't need the clubmoss
-  dplyr::select(-temp.date)
-  
-write_csv(fl_survey, "./processed_data/floral_survey.csv")
-
-### Visitation data
-network <- read_delim("./raw_data/data_bumblebee_forDoug_2020.04.17.csv", delim = ";", 
-                           locale = locale(decimal_mark = ",", grouping_mark = ".")) %>%
-  mutate(plant_sp_latin = str_replace_all(plant_sp_latin, 
-                                         c("Arabis caerula" = "Arabis caerulea", # correct nomenclature typos and synonyms
[…]

### 2021-01-14 code/visitation_bias.Rmd
update desktop

@@
 # Site data
-site_data <- read_csv("../processed_data/site_data.csv") %>%
+site_data <- read_csv("../data/processed_data/site_data.csv") %>%
   dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
@@
 # BB traits
-bb_traits <- read_csv("../processed_data/bb_traits.csv")
+bb_traits <- read_csv("../data/processed_data/bb_traits.csv")
 
@@
 ## family-level taxonomy
-fl_tax <- read_csv("../processed_data/floral_tax.csv")
+fl_tax <- read_csv("../data/processed_data/floral_tax.csv")
 
 ## Kugler morphotypes (accessed through Bioflor)
-fl_ktype <- read_csv("../processed_data/floral_k_type.csv")
+fl_ktype <- read_csv("../data/processed_data/floral_k_type.csv")
 
 ## Flower colors (accessed through BioFlor)
-fl_color <- read_csv("../processed_data/floral_color.csv")
+fl_color <- read_csv("../data/processed_data/floral_color.csv")
 
 ## diameter and corolla depth data from Gita Benadi
-fl_morph <- read_csv("../processed_data/flower_morphology_GitaBenadi.csv") %>% 
+fl_morph <- read_csv("../data/processed_data/flower_morphology_GitaBenadi.csv") %>% 
   dplyr::select(SampleID, plant.family = Family, plant.sp = Species, 
@@
 ## verbal description of Kugler morphotypes
-fl_ktype_key <- read_delim("../processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
+fl_ktype_key <- read_delim("../data/processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
   dplyr::select(k.type = 1, description = 2)
@@
 # Site-dates where BB visitation was recorded (excluding dates where only males were recorded)
-visitation_samples <- read_csv("../processed_data/network.csv") %>%
+visitation_samples <- read_csv("../data/processed_data/network.csv") %>%
   filter(caste != "male") %>%
@@
 # Site dates where floral surveying was conducted
-survey_samples <- read_csv("../processed_data/floral_survey.csv") %>%
+survey_samples <- read_csv("../data/processed_data/floral_survey.csv") %>%
   select(site, date) %>%
@@
 # BB visitation data
-visitation <- read_csv("../processed_data/network.csv") %>%
+visitation <- read_csv("../data/processed_data/network.csv") %>%
   filter(caste != "male") %>% # we will ignore males 
@@
 # Floral cover data
-flower_cover <- read_csv("../processed_data/floral_survey.csv") %>%
+flower_cover <- read_csv("../data/processed_data/floral_survey.csv") %>%
   filter(flower.cover > 0) %>% # ignore plants that were not blooming
@@
 
-# 1.
+# 1. Quantifying bias
 How does overall bias (Blüthgen's d') vary across species and through elev*time?
@@
 I think this should give me all the estimation I'm interested in.
+It's tricky to make it converge, though.
 brm.plantsp11 <- brm(dprime ~ yday * elev + (1 + elev | bb.sp), 
@@
                   #family = "skew_normal",
-                  file = "./output/brm_plantsp11",
+                  file = "../output/brm_plantsp11",
                   data = filter(bias, criterion ==  "plant.sp"))
@@
 
+# Alternatively, we can drop the varying slope term to simplify the model a bit. Still doesn't quite converge.
 brm.plantsp1 <- brm(dprime ~ yday * elev + (1 | bb.sp), 
@@
                   iter = 2000,
-                  file = "./output/brm_plantsp1",
+                  file = "../output/brm_plantsp1",
                   data = filter(bias, criterion ==  "plant.sp"))
@@
 
+# We can simplify the model still more by dropping the yday*elev interaction term.
+Now it converges.
+But I think the takeaway is that neither elevation noy yday give us a clear signal with respect  to dprime.
 brm.plantsp2 <- brm(dprime ~ yday + elev + (1 | bb.sp), 
@@
                   iter = 2000,
-                  file = "./output/brm_plantsp2",
+                  file = "../output/brm_plantsp2",
                   data = filter(bias, criterion ==  "plant.sp"))
@@
 # This, I think, would also be a legitimate way to specify the model, but it would not tell us anything about elevation or time.
+It's the simplest, and that's nice.
+Well-behaved.
+Captures about twice as much variance as the previous model, and it wins the WAIC context easily.
 brm.plantsp0 <- brm(dprime ~ 1 + (1 | bb.sp) + (1 | site),
@@
                   iter = 2000,
-                  file = "./output/brm_plantsp0",
+                  file = "../output/brm_plantsp0",
                   data = filter(bias, criterion ==  "plant.sp"))
@@
 parnames(brm.plantsp0)
+
+# I guess the message here is that both bb.sp and site contribute to the variation of dprime, but site more so.
+Biologically, I think the meaning of prime variation by site is that some sites have greater relative abundance of *attractive* flora than others.
+A site with loads of something that is never visited and a little bit of clover will yield a high dprime, while a site with a high abundance of attractive floral would yield a low dprime.
 mcmc_plot(brm.plantsp0, pars = c("^r_"))
+mcmc_plot(brm.plantsp0, pars = c("^r_bb.sp")) 
+ggsave("../output/brm.plantsp0_plot1.png")
 mcmc_plot(brm.plantsp0, pars = c("^b_"))
@@
 bayes_R2(brm.plantsp0)
+
+# Another option would be to drop just elevation from the model. This runs clean. Doesn't explain much more variance than the simpler model, though.
+brm.plantsp00 <- brm(dprime ~ yday + (1 | bb.sp) + (1 | site),
+                  control = list(max_treedepth = 10),
+                  cores = 4,
+                  iter = 2000,
+                  file = "../output/brm_plantsp00",
+                  data = filter(bias, criterion ==  "plant.sp"))
+
+summary(brm.plantsp00)
+plot(brm.plantsp00)
+parnames(brm.plantsp00)
+mcmc_plot(brm.plantsp00, pars = c("^r_"))
+mcmc_plot(brm.plantsp00, pars = c("^b_"))
+conditional_effects(brm.plantsp00)
+pp_check(brm.plantsp00)
+bayes_R2(brm.plantsp00)
+
+# Let's do a formal comparison of the last three models
+waic(brm.plantsp2, brm.plantsp0, brm.plantsp00)
+
+# The one including yday is a *slight* winner, but the margin is so small as to be meaningless. I think the simplest model is the one we want (brm.plantsp0).
+
+# Let's try the same model now, but for ktype.
+brm.ktype0 <- brm(dprime ~ 1 + (1 | bb.sp) + (1 | site),
+                  control = list(max_treedepth = 10),
+                  cores = 4,
+                  iter = 2000,
+                  family = "skew_normal",
+                  file = "../output/brm_ktype0",
+                  data = filter(bias, criterion ==  "ktype"))
+
+summary(brm.ktype0)
+plot(brm.ktype0)
+parnames(brm.ktype0)
+mcmc_plot(brm.ktype0, pars = c("^r_"))
+mcmc_plot(brm.ktype0, pars = c("^b_"))
+conditional_effects(brm.ktype0)
+pp_check(brm.ktype0)
+bayes_R2(brm.ktype0)
 ```
@@
 
-# 2. Having established the bias of BB foraging, let's zoom in on which morphotypes and families were over/under-represented
+# 2. Explaining bias
 
@@
 It feels a bit...slapped together, though.
+
+*We will adopt the term representation to avoid confusion with the object defined above.*
 
@@
          prop.visits, prop.flower.cover, bias, visits, 
-         flower.cover, plant.genus, plant.family)
+         flower.cover, plant.genus, plant.family) %>%
+  left_join(bb_traits)
 
 # 
-representation_ktype <- biaseval_group(visit_cover, k.type.ss) %>%
-  left_join(site_data) %>%
-  rename(group = criterion) %>%
-  mutate(criterion = rep("ktype", n()))
+# representation_ktype <- biaseval_group(visit_cover, k.type.ss) %>%
+#   left_join(site_data) %>%
+#   rename(group = criterion) %>%
+#   mutate(criterion = rep("ktype", n()))
 # 
@@
 
-## 2.5 Explaining bias by floral traits, BB spp, and elevation (and maybe time)
+## 2.5 Multi-level models
 ```{r}
+# The tons of zeros are due to the lots of plant species that were present at very low abundance and received few or zero visits. 
 ggplot(representation, aes(k.type.ss, bias)) +
@@
 
-ggplot(representation_ktype, aes(group, bias)) +
+ggplot(representation, aes(k.type.ss, bias)) +
   geom_boxplot()
@@
 
-ggplot(filter(representation, visits > 0), aes(bias)) +
+ggplot(filter(representation, abs(bias) > 0.1), aes(bias)) +
   geom_histogram() +
@@
 
-summary(lm(representation ~ k.type.ss, data = bias))
-
-brm_representation0 <- brm(bias ~ 1 + (1 | group),
+brm_representation0 <- brm(bias ~ 1 + (1 | k.type.ss) + (1 | site),
                  control = list(max_treedepth = 10),
                  cores = 4,
+                 chains = 4,
                  iter = 2000,
-                 file = "./output/brm_representation1",
-                 data = representation_ktype)
-
-brm_representation1 <- brm(bias ~ 1 + (1 | k.type.ss),
+                 file = "../output/brm_representation0",
+                 data = representation)
+
+summary(brm_representation0)
+plot(brm_representation0)
+parnames(brm_representation0)
+mcmc_plot(brm_representation0, pars = c("^r_"))
+#mcmc_plot(representation.ktype0, pars = c("elev"))
+mcmc_plot(brm_representation0, pars = c("^b_"))
+mcmc_plot(brm_representation0)
+#conditional_effects(representation.ktype0)
+pp_check(brm_representation0)
+bayes_R2(brm_representation0)
+
+brm_representation1 <- brm(bias ~ pbl.w + (1 + pbl.w | k.type.ss) + (1 | bb.sp),
                  control = list(max_treedepth = 10),
-                 #cores = 4,
                  cores = 4,
                  chains = 4,
-                 #iter = 2000,
                  iter = 2000,
-                 file = "./output/brm_representation1",
+                 file = "../output/brm_representation1",
                  data = representation)
@@
 plot(lmer_representation3)
-```
-
-# 3. How do species vary in their biases?
+
+# Can we explain anything with tongue length
+
+ggplot(filter(representation, bb.sp != "gers" & abs(bias) >= 0.1), aes(pbl.w, bias)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~k.type.ss)
+
+lm_pbl <- lmer(bias ~ pbl.w + (1 + pbl.w | k.type.ss) + (1 | bb.sp),
+               data = representation)
+
+summary(lm_pbl)
+```
+
+## 2.6 Ordination of bias
 ```{r}
-rep_matrix.plantsp <- representation %>%
-  group_by(bb.sp, group) %>%
+rep_matrix.ktype <- representation %>%
+  group_by(bb.sp, k.type.ss) %>%
   summarize(bias = mean(bias)) %>%
-  pivot_wider(values_from = bias, names_from = group, values_fill = 0) %>%
-  column_to_rownames("bb.sp")
-
-bias_matrix.ktype <- bias_group_ktype %>%
-  group_by(bb.sp, group) %>%
-  summarize(bias = mean(bias)) %>%
-  pivot_wider(values_from = bias, names_from = group, values_fill = 0) %>%
+  pivot_wider(values_from = bias, names_from = k.type.ss, values_fill = 0) %>%
   column_to_rownames("bb.sp")
@@
 
-pca.plantsp <- rda(bias_matrix.plantsp)
-plot(pca.plantsp, type = "t")
-envfit(pca.plantsp ~ pbl.w.class, data = traits)
-summary(ordisurf(pca.plantsp ~ pbl.w.class, data = traits))
-
-pca.ktype <- rda(bias_matrix.ktype)
+pca.ktype <- rda(rep_matrix.ktype)
+#ef.ktype <- envfit(pca.ktype ~ pbl.w.class, data = traits)
+#os.ktype <- ordisurf(pca.ktype ~ pbl.w.class, data = traits)
+ef.ktype <- envfit(pca.ktype ~ pbl.w, data = traits)
+os.ktype <- ordisurf(pca.ktype ~ pbl.w, data = traits)
 plot(pca.ktype, type = "t")
-envfit(pca.ktype ~ pbl.w.class, data = traits)
-summary(ordisurf(pca.ktype ~ pbl.w.class, data = traits))
-pca.ktype_pca1_traits <- pca.ktype$CA$u[,1] %>%
-  as.data.frame() %>%
-  rownames_to_column(var = "bb.sp") %>%
-  rename(PC1 = 2) %>%
-  left_join(bb_traits)
-
-ggplot(pca.ktype_pca1_traits, aes(pbl.w, PC1)) +
-  geom_point() +
-  geom_smooth(method = "lm", formula = y ~ poly(x, 2))
-
-ggplot(pca.ktype_pca1_traits, aes(pbl.w.class, PC1)) +
-  geom_point() +
-  geom_smooth(method = "lm")
-```
-
-# 3. Can we explain bias by BB-Fl trait associations? (RLQ + fourth corner analysis)
-## 3.1
-## 3.2 Do BB-FL trait associations vary with elevation?
-
-# 4. How does bias relate to network structure?
-## 4.1 A function for calculating niche overlap
+plot(ef.ktype, add = TRUE)
+plot(os.ktype)
+```
+
+# 3. Effects of bias
+
+## 3.1 Relationship between bias(d') and niche overlap
+### 3.1.1 A function for calculating niche overlap
 ```{r}
@@
 
-## 4.2 Calculate niche overlap: NOT READY FOR INTERPRETATION YET
+### 3.1.2 Calculate niche overlap: NOT READY FOR INTERPRETATION YET
 ```{r}
@@
 ```
+
+## 3.2 Ordination of dietary niche
+```{r}
+rep_matrix.ktype <- representation %>%
+  group_by(bb.sp, k.type.ss) %>%
+  summarize(bias = mean(bias)) %>%
+  pivot_wider(values_from = bias, names_from = k.type.ss, values_fill = 0) %>%
+  column_to_rownames("bb.sp")
+
+traits <- bb_traits %>%
+  arrange(bb.sp)
+
+pca.ktype <- rda(rep_matrix.ktype)
+#ef.ktype <- envfit(pca.ktype ~ pbl.w.class, data = traits)
+#os.ktype <- ordisurf(pca.ktype ~ pbl.w.class, data = traits)
+ef.ktype <- envfit(pca.ktype ~ pbl.w, data = traits)
+os.ktype <- ordisurf(pca.ktype ~ pbl.w, data = traits)
+plot(pca.ktype, type = "t")
+plot(ef.ktype, add = TRUE)
+plot(os.ktype)
+```

### 2021-01-14 writing/manuscript_v4.Rmd
update desktop

@@
 At an evolutionary level, the innate preferences of bumble bees for certain floral traits likely reflect historic associations between floral traits and floral rewards [@Raine2007-nq].
-These innate preferences, however, can be modulated or even nullified by patterns of reward learned during foraging [@Kugler1943-uy; Maharaj2019-wm], which depend not only intrinsic floral properties but also ecological modulators, such as the spatial distribution of floral patches or competition with other bumble bees [@Ishii2013-sm; @Plowright1984-lc; @Sowig1989-ld].
+These innate preferences, however, can be modulated or even nullified by patterns of reward learned during foraging [@Kugler1943-uy; Maharaj2019-wm], which depend not only on intrinsic floral properties but also ecological modulators, such as the spatial distribution of floral patches or competition with other bumble bees [@Ishii2013-sm; @Plowright1984-lc; @Sowig1989-ld].
 Thus, realized biases in bumble bee foraging are the labile product of bumble bee traits, floral traits, and ecological context.
-The influence of these biases on interaction network structure has consequences both for plant pollination and bumble bee species coexistence, and these consequences potentially feed back into the causes of bumble bee foraging bias via the modulation of competition between bumble bees and ultimately via the regulation of community assembly.
+The influence of these biases on interaction network structure has consequences both for plant pollination and bumble bee species coexistence, and these consequences potentially feed back upon the causes of bumble bee foraging bias via the modulation of competition between bumble bees and ultimately via the regulation of community assembly.
 <!-- strengthen this sentence with some refs?
@@
 Heinrich1976-wc; @Goulson2008-zy].
-The former approach enables more confident causal inference with respect to the mechanisms of floral choice, but it suffers all the usual limitations of biological realism associated with experimental control, and the higher-order effects of floral choices on network structure cannot be observed meaningfully under laboratory conditions.
-Observational field studies, however, have tended to be anecdotal in scope, typically describing bumble bee foraging patterns in particular localities, over short time frames, and with limited floristic data to enable a distinction between true foraging bias and the neutral effects of floral abundance [though see @Neumayer1999-il].
+The former approach enables stronger causal inference with respect to the mechanisms of floral choice, but it suffers all the usual limitations of biological realism associated with experimental control, and the higher-order effects of floral choices on network structure cannot be observed meaningfully under laboratory conditions.
+Observational field studies, however, tend to be anecdotal in scope, typically describing bumble bee foraging patterns in particular localities, over short time frames, and with limited floristic data to enable a distinction between true foraging bias and the neutral effects of floral abundance [though see @Neumayer1999-il].
 The steep elevation gradients of mountain slopes provide an opportunity for more powerful observational study of bumble foraging because they generate broad gradients of floristic and abiotic variation while controlling for confounding effects of geographic variation, such as genotypic differentiation within bumble bee species.
@@
 Mountain habitats are also arguably the most relevant places to study bumble bee ecology, since bumble bees are adapted to cool climates, and their global distribution is concentrated in areas of high elevation or high latitude.
-In such environments, bumble bees tend to dominate the pollinator community, and local flora often depend primarily on bumble bee visitation for pollination <!-- This paragraph needs to be beefed up a bit, but it at least sketches out the points I want to make -->.
+In such environments, bumble bees tend to dominate the pollinator community, and local flora often depend primarily on bumble bee visitation for pollination.
+<!-- This paragraph needs to be beefed up a bit, but it at least sketches out the points I want to make.
+I remember a reference somewhere -- I think in one of the Nordic papers, and I think a recent one -- that suggested that the morphology of alpine flowers has been shaped by selection pressure to attract bumble bees -->
 
@@
 # Results
+## Cumulative bias across species and sites
+```{r}
+knitr::include_graphics("../output/brm.plantsp0_plot1.png")
+```
+
+## Parsing bias by floral morphotype [representation ~ 1 + (1 + ktype) + (1 + site)]
+
+## Linking morphological bias to species' traits [bias ordination w/ tongue length]
+
+##
 

### 2021-01-14 code/visitation_bias.Rmd
back to lappy

@@
 library(lme4)
-library(ggeffects)
-library(sjPlot)
-#library(inla)
 library(bayesplot)
 library(vegan)
-#library(tidybayes)
 library(tidyverse)

### 2021-01-15 code/visitation_bias.Rmd
oh yeah

@@
 library(bipartite)
-library(mgcv)
-library(mgcViz)
 library(brms)
 library(lme4)
+library(sjPlot)
+library(DHARMa)
 library(bayesplot)
@@
 
-# This model includes a global interaction term for yday and elevation, with a varying intercept for bb.sp and a varying slow for elev | bb.sp.
+# # This model includes a global interaction term for yday and elevation, with a varying intercept for bb.sp and a varying slow for elev | bb.sp.
 I think this should give me all the estimation I'm interested in.
 It's tricky to make it converge, though.
-brm.plantsp11 <- brm(dprime ~ yday * elev + (1 + elev | bb.sp), 
-                  control = list(max_treedepth = 10),
-                  cores = 4,
-                  iter = 4000,
-                  #family = "skew_normal",
-                  file = "../output/brm_plantsp11",
-                  data = filter(bias, criterion ==  "plant.sp"))
-
-summary(brm.plantsp11)
-prior_summary(brm.plantsp11)
-plot(brm.plantsp11)
-parnames(brm.plantsp11)
-mcmc_plot(brm.plantsp11, pars = c("^r_"))
-mcmc_plot(brm.plantsp11, pars = c("elev"))
-mcmc_plot(brm.plantsp11, pars = c("^b_"))
-mcmc_plot(brm.plantsp11)
-conditional_effects(brm.plantsp11)
-pp_check(brm.plantsp11)
-bayes_R2(brm.plantsp11)
-hypothesis(brm.plantsp11, "b_elev = 0", class = NULL)
-
-# Alternatively, we can drop the varying slope term to simplify the model a bit. Still doesn't quite converge.
-brm.plantsp1 <- brm(dprime ~ yday * elev + (1 | bb.sp), 
-                  control = list(max_treedepth = 10),
-                  cores = 4,
-                  iter = 2000,
-                  file = "../output/brm_plantsp1",
-                  data = filter(bias, criterion ==  "plant.sp"))
-
-summary(brm.plantsp1)
-prior_summary(brm.plantsp1)
-plot(brm.plantsp1)
-parnames(brm.plantsp1)
-mcmc_plot(brm.plantsp1, pars = c("^r_"))
-mcmc_plot(brm.plantsp1, pars = c("elev"))
-mcmc_plot(brm.plantsp1, pars = c("^b_"))
-mcmc_plot(brm.plantsp1)
-conditional_effects(brm.plantsp1)
-pp_check(brm.plantsp1)
-bayes_R2(brm.plantsp1)
-hypothesis(brm.plantsp1, "b_elev = 0", class = NULL)
-
-# We can simplify the model still more by dropping the yday*elev interaction term.
+# brm.plantsp11 <- brm(dprime ~ yday * elev + (1 + elev | bb.sp), 
+#                   control = list(max_treedepth = 10),
+#                   cores = 4,
+#                   iter = 4000,
+#                   #family = "skew_normal",
+#                   file = "../output/brm_plantsp11",
+#                   data = filter(bias, criterion ==  "plant.sp"))
+# 
+# summary(brm.plantsp11)
+# prior_summary(brm.plantsp11)
+# plot(brm.plantsp11)
+# parnames(brm.plantsp11)
+# mcmc_plot(brm.plantsp11, pars = c("^r_"))
+# mcmc_plot(brm.plantsp11, pars = c("elev"))
+# mcmc_plot(brm.plantsp11, pars = c("^b_"))
+# mcmc_plot(brm.plantsp11)
+# conditional_effects(brm.plantsp11)
+# pp_check(brm.plantsp11)
+# bayes_R2(brm.plantsp11)
+# hypothesis(brm.plantsp11, "b_elev = 0", class = NULL)
+# 
+# # Alternatively, we can drop the varying slope term to simplify the model a bit. Still doesn't quite converge.
+# brm.plantsp1 <- brm(dprime ~ yday * elev + (1 | bb.sp), 
+#                   control = list(max_treedepth = 10),
+#                   cores = 4,
+#                   iter = 2000,
+#                   file = "../output/brm_plantsp1",
+#                   data = filter(bias, criterion ==  "plant.sp"))
+# 
+# summary(brm.plantsp1)
+# prior_summary(brm.plantsp1)
+# plot(brm.plantsp1)
+# parnames(brm.plantsp1)
+# mcmc_plot(brm.plantsp1, pars = c("^r_"))
+# mcmc_plot(brm.plantsp1, pars = c("elev"))
+# mcmc_plot(brm.plantsp1, pars = c("^b_"))
+# mcmc_plot(brm.plantsp1)
+# conditional_effects(brm.plantsp1)
+# pp_check(brm.plantsp1)
+# bayes_R2(brm.plantsp1)
+# hypothesis(brm.plantsp1, "b_elev = 0", class = NULL)
+# 
+# # We can simplify the model still more by dropping the yday*elev interaction term.
 Now it converges.
 But I think the takeaway is that neither elevation noy yday give us a clear signal with respect  to dprime.
-brm.plantsp2 <- brm(dprime ~ yday + elev + (1 | bb.sp), 
-                  control = list(max_treedepth = 10),
-                  cores = 4,
-                  iter = 2000,
-                  file = "../output/brm_plantsp2",
-                  data = filter(bias, criterion ==  "plant.sp"))
-
-summary(brm.plantsp2)
-prior_summary(brm.plantsp2)
-plot(brm.plantsp2)
-parnames(brm.plantsp2)
-mcmc_plot(brm.plantsp2, pars = c("^r_"))
-mcmc_plot(brm.plantsp2, pars = c("elev"))
-mcmc_plot(brm.plantsp2, pars = c("^b_"))
-mcmc_plot(brm.plantsp2)
-conditional_effects(brm.plantsp2)
-pp_check(brm.plantsp2)
-bayes_R2(brm.plantsp2)
-hypothesis(brm.plantsp2, "b_elev = 0", class = NULL)
+# brm.plantsp2 <- brm(dprime ~ yday + elev + (1 | bb.sp), 
+#                   control = list(max_treedepth = 10),
+#                   cores = 4,
+#                   iter = 2000,
+#                   file = "../output/brm_plantsp2",
+#                   data = filter(bias, criterion ==  "plant.sp"))
+# 
+# summary(brm.plantsp2)
+# prior_summary(brm.plantsp2)
+# plot(brm.plantsp2)
+# parnames(brm.plantsp2)
+# mcmc_plot(brm.plantsp2, pars = c("^r_"))
+# mcmc_plot(brm.plantsp2, pars = c("elev"))
+# mcmc_plot(brm.plantsp2, pars = c("^b_"))
+# mcmc_plot(brm.plantsp2)
+# conditional_effects(brm.plantsp2)
+# pp_check(brm.plantsp2)
+# bayes_R2(brm.plantsp2)
+# hypothesis(brm.plantsp2, "b_elev = 0", class = NULL)
 
@@
 
-# I guess the message here is that both bb.sp and site contribute to the variation of dprime, but site more so.
+# Out of curiosity, let's see what we get from the lme4 approach to the same model; we can use this as a sanity check of sorts.
+If the results of lme4 differ dramatically from those of brms, we might wonder why.
+lme4.plantsp0 <- lmer(dprime ~ 1 + (1 | bb.sp) + (1 | site),
+                      data = filter(bias, criterion ==  "plant.sp"))
+
+summary(lme4.plantsp0)
+#plot_model(lme4.plantsp0, type = "pred")
+plot_model(lme4.plantsp0, type = "re")
+
+lme4.plantsp1 <- lmer(dprime ~ elev * yday + (1 | bb.sp) + (1 | site),
+                      data = filter(bias, criterion ==  "plant.sp"))
+
+summary(lme4.plantsp1)
+plot_model(lme4.plantsp1, type = "pred")
+plot_model(lme4.plantsp1, type = "pred", terms = c("yday", "elev"))
+plot_model(lme4.plantsp1, type = "re")
+
+AIC(lme4.plantsp1)
+AIC(lme4.plantsp0)
+
+# 
+# # I guess the message here is that both bb.sp and site contribute to the variation of dprime, but site more so.
 Biologically, I think the meaning of prime variation by site is that some sites have greater relative abundance of *attractive* flora than others.
 A site with loads of something that is never visited and a little bit of clover will yield a high dprime, while a site with a high abundance of attractive floral would yield a low dprime.
-mcmc_plot(brm.plantsp0, pars = c("^r_"))
-mcmc_plot(brm.plantsp0, pars = c("^r_bb.sp")) 
-ggsave("../output/brm.plantsp0_plot1.png")
-mcmc_plot(brm.plantsp0, pars = c("^b_"))
-pp_check(brm.plantsp0)
-bayes_R2(brm.plantsp0)
-
-# Another option would be to drop just elevation from the model. This runs clean. Doesn't explain much more variance than the simpler model, though.
-brm.plantsp00 <- brm(dprime ~ yday + (1 | bb.sp) + (1 | site),
-                  control = list(max_treedepth = 10),
-                  cores = 4,
-                  iter = 2000,
-                  file = "../output/brm_plantsp00",
-                  data = filter(bias, criterion ==  "plant.sp"))
-
-summary(brm.plantsp00)
-plot(brm.plantsp00)
-parnames(brm.plantsp00)
-mcmc_plot(brm.plantsp00, pars = c("^r_"))
-mcmc_plot(brm.plantsp00, pars = c("^b_"))
-conditional_effects(brm.plantsp00)
-pp_check(brm.plantsp00)
-bayes_R2(brm.plantsp00)
-
-# Let's do a formal comparison of the last three models
-waic(brm.plantsp2, brm.plantsp0, brm.plantsp00)
-
-# The one including yday is a *slight* winner, but the margin is so small as to be meaningless. I think the simplest model is the one we want (brm.plantsp0).
-
-# Let's try the same model now, but for ktype.
-brm.ktype0 <- brm(dprime ~ 1 + (1 | bb.sp) + (1 | site),
-                  control = list(max_treedepth = 10),
-                  cores = 4,
-                  iter = 2000,
-                  family = "skew_normal",
-                  file = "../output/brm_ktype0",
-                  data = filter(bias, criterion ==  "ktype"))
-
-summary(brm.ktype0)
-plot(brm.ktype0)
-parnames(brm.ktype0)
-mcmc_plot(brm.ktype0, pars = c("^r_"))
-mcmc_plot(brm.ktype0, pars = c("^b_"))
-conditional_effects(brm.ktype0)
-pp_check(brm.ktype0)
-bayes_R2(brm.ktype0)
+# mcmc_plot(brm.plantsp0, pars = c("^r_"))
+# mcmc_plot(brm.plantsp0, pars = c("^r_bb.sp")) 
+# ggsave("../output/brm.plantsp0_plot1.png")
+# mcmc_plot(brm.plantsp0, pars = c("^b_"))
+# pp_check(brm.plantsp0)
+# bayes_R2(brm.plantsp0)
+# 
+# # Another option would be to drop just elevation from the model. This runs clean. Doesn't explain much more variance than the simpler model, though.
+# brm.plantsp00 <- brm(dprime ~ yday + (1 | bb.sp) + (1 | site),
+#                   control = list(max_treedepth = 10),
+#                   cores = 4,
+#                   iter = 2000,
+#                   file = "../output/brm_plantsp00",
+#                   data = filter(bias, criterion ==  "plant.sp"))
+# 
+# summary(brm.plantsp00)
+# plot(brm.plantsp00)
+# parnames(brm.plantsp00)
+# mcmc_plot(brm.plantsp00, pars = c("^r_"))
+# mcmc_plot(brm.plantsp00, pars = c("^b_"))
+# conditional_effects(brm.plantsp00)
+# pp_check(brm.plantsp00)
+# bayes_R2(brm.plantsp00)
+# 
+# # Let's do a formal comparison of the last three models
+# waic(brm.plantsp2, brm.plantsp0, brm.plantsp00)
+# 
+# # The one including yday is a *slight* winner, but the margin is so small as to be meaningless. I think the simplest model is the one we want (brm.plantsp0).
+# 
+# # Let's try the same model now, but for ktype.
+# brm.ktype0 <- brm(dprime ~ 1 + (1 | bb.sp) + (1 | site),
+#                   control = list(max_treedepth = 10),
+#                   cores = 4,
+#                   iter = 2000,
+#                   family = "skew_normal",
+#                   file = "../output/brm_ktype0",
+#                   data = filter(bias, criterion ==  "ktype"))
+# 
+# summary(brm.ktype0)
+# plot(brm.ktype0)
+# parnames(brm.ktype0)
+# mcmc_plot(brm.ktype0, pars = c("^r_"))
+# mcmc_plot(brm.ktype0, pars = c("^b_"))
+# conditional_effects(brm.ktype0)
+# pp_check(brm.ktype0)
+# bayes_R2(brm.ktype0)
 ```
@@
 
-## 2.3 Calculate bias per plant speces/ktype/family
-```{r}
-representation <- biaseval_group(visit_cover, plant.sp) %>%
-  rename(plant.sp = criterion) %>%
-  left_join(site_data) %>%
+## 2.3 Calculate visitation probability per plant speces/ktype/family
+```{r}
+visit_prob <- visit_cover %>%
+  group_by(site, date, bb.sp, plant.sp) %>%
+  summarize(flower.cover = sum(flower.cover), # get raw flower cover and visits by group
+            visits = sum(visits)) %>%
+  group_by(site, date) %>%
+  mutate(prop.flower.cover = flower.cover/sum(flower.cover), # get proportional flower cover and visits by group
+         prop.visits = visits/sum(visits),
+         visits.bin = factor(if_else(visits > 0, 1, 0))) %>%
+  group_by(site, date, bb.sp) %>%
+  mutate(bb.sp.abund = sum(visits)) %>%
   left_join(fl_traits) %>%
-  mutate(k.type.ss = factor(k.type.ss)) %>%
-  select(site, date, elev.mean, yday, bb.sp, plant.sp, k.type.ss, 
-         prop.visits, prop.flower.cover, bias, visits, 
-         flower.cover, plant.genus, plant.family) %>%
-  left_join(bb_traits)
+  left_join(bb_traits) %>%
+  select(site, date, bb.sp, plant.sp, visits.bin, prop.flower.cover, k.type.ss, bb.sp.abund, pbl.w.class2) %>%
+  mutate(k.type.ss = factor(k.type.ss))
+
+# Pooled across BB spp
+visit_prob_pooled <- visit_cover %>%
+  group_by(site, date, plant.sp) %>%
+  summarize(flower.cover = sum(flower.cover), # get raw floer cover and visits by group
+            visits = sum(visits)) %>%
+  group_by(site, date) %>%
+  mutate(prop.flower.cover = flower.cover/sum(flower.cover), # get proportional flower cover and visits by group
+         prop.visits = visits/sum(visits),
+         bb.abund = sum(visits),
+         visits.bin = factor(if_else(visits > 0, 1, 0))) %>%
+  mutate(yday = yday(date)) %>%
+  left_join(fl_traits) %>%
+  mutate(k.type.ss = factor(k.type.ss))
 
@@
 ## 2.5 Multi-level models
-```{r}
-# The tons of zeros are due to the lots of plant species that were present at very low abundance and received few or zero visits. 
+My approach will be to fit models using both brms and lme4, partly as a learning exercise, partly to enable quicker model development, partly as a sanity check / back-up plan.
+
+### Preliminary visualization
+The tons of zeros are due to the lots of plant species that were present at very low abundance and received few or zero visits.
+This might not be ideal.
+We could opt for one of our alternative bias measure, like fold change, which all give something like a poisson distribution, but I don't know if this would be an improvement.
+We could also filter our data to include only plants that were above some threshold of relative abundance, say, 1%.
+```{r}
+# Bias
 ggplot(representation, aes(k.type.ss, bias)) +
@@
 
-ggplot(representation, aes(k.type.ss, bias)) +
-  geom_boxplot()
-
-ggplot(representation, aes(bias)) +
-  geom_histogram() +
-  facet_wrap(~k.type.ss)
-
-ggplot(filter(representation, abs(bias) > 0.1), aes(bias)) +
-  geom_histogram() +
-  facet_wrap(~k.type.ss)
-
-ggplot(filter(representation, prop.flower.cover > 0.05), aes(k.type.ss, bias)) +
-  geom_boxplot()
-
-ggplot(representation, aes(prop.flower.cover, prop.visits)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_wrap(~k.type.ss)
-
-brm_representation0 <- brm(bias ~ 1 + (1 | k.type.ss) + (1 | site),
-                 control = list(max_treedepth = 10),
+# Bias.fold
+ggplot(representation, aes(k.type.ss, bias.fold)) +
+  geom_violin()
+
+# Bias.exp
+ggplot(representation, aes(k.type.ss, bias.exp)) +
+  geom_violin()
+
+# Bias.bin
+ggplot(representation, aes(k.type.ss, fill = bias.bin)) +
+  geom_bar()
+
+# Visits.bin
+ggplot(representation, aes(k.type.ss, fill = visits.bin)) +
+  geom_bar()
+
+ggplot(visit_prob_pooled, aes(k.type.ss, fill = visits.bin)) +
+  geom_bar(position = "fill")
+
+# Bias filtered
+ggplot(filter(representation, prop.flower.cover >= 0.05), aes(k.type.ss, bias)) +
+  geom_violin()
+
+ggplot(filter(representation, visits > 0), aes(k.type.ss, bias)) +
+  geom_violin()
+
+# Visits.bin filtered
+ggplot(filter(representation, prop.flower.cover >= 0.05), aes(k.type.ss, fill = visits.bin)) +
+  geom_bar()
+```
+
+### lme4 models
+We will drop ktypes 9 and 10 from these models since both are vanishingly rare
+```{r}
+# This model does not parse by tongue length
+glmer_vp00 <- glmer(visits.bin ~ prop.flower.cover + bb.sp.abund + (1 | k.type.ss) + (1 | site),
+                    family  = "binomial",
+                    data = filter(visit_prob, !k.type.ss %in% c("9", "10")))
+
+plot_model(glmer_vp00, type = "pred")
+plot_model(glmer_vp00, type = "re")
+
+# This model does 
+glmer_vp01 <- glmer(visits.bin ~ prop.flower.cover + bb.sp.abund + (1 | k.type.ss:pbl.w.class2) + (1 | site),
+                    family  = "binomial",
+                    data = filter(visit_prob, !k.type.ss %in% c("9", "10")))
+
+plot_model(glmer_vp01, type = "pred")
+plot_model(glmer_vp01, type = "re")
+```
+
+### brms models
+```{r}
+# This model does not parse by tongue length
+brm_representation00 <- brm(visits.bin ~ prop.flower.cover + bb.sp.abund + (1 | k.type.ss) + (1 | site),
+                            family = "bernoulli",
+                            file = "../output/brm_vp00",
+                            chains = 4,
                             cores = 4,
+                            iter = 2000,
+                            data = filter(visit_prob, !k.type.ss %in% c("9", "10")))
+
+summary(brm_representation00)
+prior_summary(brm_representation00)
+parnames(brm_representation00)
+mcmc_plot(brm_representation00, pars = c("^r_k"))
+mcmc_plot(brm_representation00, pars = c("elev"))
+mcmc_plot(brm_representation00, pars = c("^b_"))
+mcmc_plot(brm_representation00)
+conditional_effects(brm_representation00)
+pp_check(brm_representation00)
+bayes_R2(brm_representation00)
+
+hypothesis(brm_representation00, "r_k.type.ss[0,Intercept] = r_k.type.ss[1,Intercept]", class = NULL)
+hypothesis(brm_representation00, "r_k.type.ss[1,Intercept] = r_k.type.ss[2,Intercept]", class = NULL)
+hypothesis(brm_representation00, "r_k.type.ss[6,Intercept] = r_k.type.ss[5,Intercept]", class = NULL)
+hypothesis(brm_representation00, "r_k.type.ss[6,Intercept] = r_k.type.ss[5,Intercept]", class = NULL)
+hypothesis(brm_representation00, "r_k.type.ss[6,Intercept] = r_k.type.ss[5,Intercept]", class = NULL)
+
+# This model does
+brm_representation01 <- brm(visits.bin ~ prop.flower.cover + bb.sp.abund + (1 | k.type.ss:pbl.w.class2) + (1 | site),
+                            family = "bernoulli",
+                            file = "../output/brm_vp01",
                             chains = 4,
+                            cores = 4,
                             iter = 2000,
-                 file = "../output/brm_representation0",
-                 data = representation)
-
-summary(brm_representation0)
-plot(brm_representation0)
-parnames(brm_representation0)
-mcmc_plot(brm_representation0, pars = c("^r_"))
-#mcmc_plot(representation.ktype0, pars = c("elev"))
-mcmc_plot(brm_representation0, pars = c("^b_"))
-mcmc_plot(brm_representation0)
-#conditional_effects(representation.ktype0)
-pp_check(brm_representation0)
-bayes_R2(brm_representation0)
-
-brm_representation1 <- brm(bias ~ pbl.w + (1 + pbl.w | k.type.ss) + (1 | bb.sp),
-                 control = list(max_treedepth = 10),
-                 cores = 4,
-                 chains = 4,
-                 iter = 2000,
-                 file = "../output/brm_representation1",
-                 data = representation)
-
-summary(brm_representation1)
-plot(brm_representation1)
-parnames(brm_representation1)
-mcmc_plot(brm_representation1, pars = c("^r_"))
-#mcmc_plot(representation.ktype1, pars = c("elev"))
-mcmc_plot(brm_representation1, pars = c("^b_"))
-mcmc_plot(brm_representation1)
-#conditional_effects(representation.ktype1)
-pp_check(brm_representation1)
-bayes_R2(brm_representation1)
-
-hypothesis(representation.ktype1, "r_group[5,Intercept] = r_group[6,Intercept]", class = "")
-hypothesis(representation.ktype1, "r_group[1,Intercept] = r_group[0,Intercept]", class = "")
-
-# One interesting way to think about bias is to focus on the slope of prop.visits ~ prop.flower.cover. 
-# We could analyze this with a varying slope model
-
-# Binomial ggplot smoother
-binomial_smooth <- function(...) {
-  geom_smooth(method = "glm", method.args = list(family = "quasibinomial"), se = FALSE, ...)
-}
-
-ggplot(representation, aes(sqrt(prop.flower.cover), sqrt(prop.visits))) +
-  geom_point() +
-  #geom_smooth(method = "lm") +
-  #binomial_smooth(formula = y ~ splines::ns(x, 3)) +
-  geom_smooth() +
-  #geom_smooth(method = "glm", method.args = list(family = "binomial")) +
-  facet_grid(k.type.ss~bb.sp)
-
-brm_representation3 <- brm(prop.visits ~ prop.flower.cover + (1 + prop.flower.cover | k.type.ss),
-                 control = list(max_treedepth = 10),
-                 #cores = 4,
-                 cores = 2,
-                 chains = 2,
-                 #iter = 2000,
-                 iter = 1000,
-                 file = "./output/brm_representation3",
-                 data = representation)
-
-lmer_representation3 <- lmer(prop.visits ~ log(prop.flower.cover) + (1 + log(prop.flower.cover) | k.type.ss),
-                             data = representation)
-
-summary(lmer_representation3)
-plot(lmer_representation3)
-
-# Can we explain anything with tongue length
-
-ggplot(filter(representation, bb.sp != "gers" & abs(bias) >= 0.1), aes(pbl.w, bias)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_wrap(~k.type.ss)
-
-lm_pbl <- lmer(bias ~ pbl.w + (1 + pbl.w | k.type.ss) + (1 | bb.sp),
-               data = representation)
-
-summary(lm_pbl)
+                            data = filter(visit_prob, !k.type.ss %in% c("9", "10")))
+
 ```
@@
 
+rep_matrix.ktype2 <- representation %>%
+  group_by(bb.sp, k.type.ss) %>%
+  summarize(visit.bin = mean(bias)) %>%
+  pivot_wider(values_from = visit.bin, names_from = k.type.ss, values_fill = 0) %>%
+  column_to_rownames("bb.sp")
+
 traits <- bb_traits %>%
@@
 
-pca.ktype <- rda(rep_matrix.ktype)
+pca.ktype <- rda(rep_matrix.ktype2)
 #ef.ktype <- envfit(pca.ktype ~ pbl.w.class, data = traits)

### 2021-01-15 writing/manuscript_v4.Rmd
oh yeah

@@
 
-# Introduction
+# 1. Introduction
 
@@
 
-# Methods
+# 2. Methods
 
-## Field sampling
+## 2.1 Field sampling
 
@@
 
-## Data analysis
+## 2.2. Data analysis
 
@@
 
+### 2.2.1 Modeling cumulative bias
+
 To quantify cumulative bias per bumble bee species, we used the discrimination (d’) metric of @Bluthgen2006-if, which represents the net deviation of a species’ visitation from the predictions of a null model in which the probability of visitation is purely a function of floral abundance.
 
-# Results
-## Cumulative bias across species and sites
+#### 2.2.2.1.1 Model 1: dprime ~
+
+### 2.2.2 Modeling visitation probability as a function of floral morphotype and bumble bee tongue length
+
+#### 2.2.2.1 Model 2: visits.bin ~ prop.flower.cover + bb.sp.abund + (1 | k.type.ss) + (1 | site)
+
+#### 2.2.2.2 Model 3: visits.bin ~ prop.flower.cover + bb.sp.abund + (1 | k.type.ss:pbl.w.class2) + (1 | site)
+
+#### 2.2.2.3 Model 4: Ordination with post-hoc fitting of tongue-length
+
+### 2.2.3 Modeling the relationship between bias and niche relationships 
+<!-- I might try to include a plant-centric analysis, too, to see how bias relates to potential competition between plants for BB pollinators.
+I think it is true to say that if floral resources are not limiting for bees, bee visitation is limiting for flora.
+-->
+
+#### Model 5: niche_overlap ~ dprime
+
+#### Model 6: Ordination of diet
+
+# 3. Results
+## 3.1 Model 1
 ```{r}
@@
 
-## Parsing bias by floral morphotype [representation ~ 1 + (1 + ktype) + (1 + site)]
+## 3.2 Model 2
 
-## Linking morphological bias to species' traits [bias ordination w/ tongue length]
+## 3.3. Model 3
 
-##
+## 3.4 Model 4
 
-# Discussion
+## 3.5 Model 5
+
+## 3.6 Model 6
+
+# 4. Discussion
+
+## 4.1 The phenomenon of bias: Cumulative bias and variation across species
+
+## 4.2 The mechanism of bias: Explaining bias by the interaction of flower and bumble bee morphology
+
+## 4.3 The outcome of bias: Linking bias to niche relationships
 
@@
 
-# References
+# 5. References

### 2021-01-22 code/visitation_bias.Rmd
backup

--- 2021-01-14 code/visitation_bias.Rmd
@@
 library(bipartite)
-library(mgcv)
-library(mgcViz)
 library(brms)
 library(lme4)
+library(sjPlot)
+library(DHARMa)
 library(bayesplot)
+library(tidybayes)
+library(modelr)
+library(ggdist)
 library(vegan)
+library(ggvegan)
+library(ggrepel)
+library(treemapify)
+library(patchwork)
+library(tidymodels)
 library(tidyverse)
 library(lubridate)
-library(tidymodels)
+
+rstan::rstan_options(auto_write = TRUE)
+options(mc.cores = parallel::detectCores())
 ```
@@
 bb_traits <- read_csv("../data/processed_data/bb_traits.csv")
+
+ggplot(bb_traits, aes(pbl.w, pbl.w.durieaux,
+                      color = bb.sp, pch = factor(pbl.w.class2))) +
+  geom_point()
 
@@
 
+ktype_families <- visit_cover %>%
+  select(plant.sp, plant.family, k.type.ss) %>%
+  distinct()
+
+ggplot(ktype_families, aes(plant.family, fill = k.type.ss)) +
+  geom_bar(position = "fill") +
+  theme(axis.text.x = element_text(angle = 90))
+
 ## Get total visits by BB sp and site-date; this will be a measure of BB abundance that we can use after we've calculated d'.
@@
 bias_plantsp <- biaseval(visit_cover, plant.sp, site, date) %>%
+  rename(dprime.sp = dprime) %>%
+  select(site, date, bb.sp, dprime.sp)
+
+# Grouped by k.type.ss
+bias_ktype <- biaseval(visit_cover, k.type.ss, site, date) %>%
+  rename(dprime.ktype = dprime) %>%
+  select(site, date, bb.sp, dprime.ktype)
+
+# Grouped by plant.family
+bias_family <- biaseval(visit_cover, plant.family, site, date) %>%
+  rename(dprime.fam = dprime) %>%
+  select(site, date, bb.sp, dprime.fam)
+
+# Collated
+bias <- full_join(bias_plantsp, bias_ktype) %>%
+  full_join(bias_family) %>%
   left_join(site_data) %>% # add site data
@@
   left_join(fl_abundance) %>% # add fl abundance
-  mutate(criterion = factor(rep("plant.sp", n())))
-
-# Grouped by k.type.ss
-bias_ktype <- biaseval(visit_cover, k.type.ss, site, date) %>%
-  left_join(site_data) %>%
-  left_join(bb_abundance) %>% # add BB abundance data
-  left_join(fl_abundance) %>% # add fl abundance
-  mutate(criterion = factor(rep("ktype", n())))
-
-# Grouped by plant.family
-bias_family <- biaseval(visit_cover, plant.family, site, date) %>%
-  left_join(site_data) %>%
-  left_join(bb_abundance) %>% # add BB abundance data
-  left_join(fl_abundance) %>% # add fl abundance
-  mutate(criterion = factor(rep("plant.family", n())))
-
-# Collated
-bias <- bind_rows(bias_plantsp, 
-                  bias_ktype, 
-                  bias_family) %>%
+  filter(bb.sp.abund >= 3) %>% # dprime tends to become binary when n is very low; this threshold removes more than half the data points, but almost all those were ones and zeros.
+These ones and zeros probably don't mean much, and they make the models hard to fit.
   mutate(bb.sp = factor(bb.sp), 
          yday = yday(date),
-         elev = elev.mean/10) 
-```
-
-## 1.2 Visualize
-```{r}
-# Compare mean bias by criterion
-ggplot(bias, aes(criterion, dprime)) +
-  geom_boxplot()
-
-# Bias ~ elevation * criterion
-ggplot(bias, aes(elev.mean, dprime, color = criterion)) +
-  geom_point(alpha = 0.25) +
-  geom_smooth(se = TRUE)
-
-ggplot(bias, aes(yday, dprime, color = criterion)) +
-  geom_point(alpha = 0.25) +
-  geom_smooth(se = TRUE)
-
-ggplot(bias, aes(elev.mean, dprime, color = criterion)) +
-  geom_point(alpha = 0.25) +
-  geom_smooth(se = FALSE, method = "gam", formula = y ~ s(x, k = 5)) +
-  facet_wrap(~bb.sp)
-
-ggplot(bias, aes(yday, dprime, color = criterion)) +
-  geom_point(alpha = 0.25) +
-  geom_smooth(se = FALSE, method = "gam", formula = y ~ s(x, k = 5)) +
-  facet_wrap(~bb.sp)
-
-# Bias ~ year; Blüthgen's d' seems very robust to sampling effects, since it does not vary across years
-ggplot(bias, aes(year, dprime, color = criterion)) +
-  geom_boxplot()
-
-ggplot(bias, aes(year, dprime, color = criterion)) +
-  geom_boxplot() +
-  facet_wrap(~bb.sp)
-
-# Does bias vary with abundance?
-ggplot(bias, aes(log(bb.sp.abund), dprime, color = criterion)) +
-  #geom_point() +
-  geom_smooth() +
-  facet_wrap(~bb.sp)
-
-ggplot(bias, aes(bb.sp.abund, dprime, color = criterion)) +
-  #geom_point() +
-  geom_smooth() +
-  facet_wrap(~bb.sp, )
-
-ggplot(bias, aes(log(bb.sp.abund), dprime, color = criterion)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth()
-
-ggplot(bias, aes(log(bb.sp.prop.abund), dprime)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~bb.sp)
+         elev = elev.mean/10) %>% #  scale this down for convenience
+  mutate(yday0 = scale(yday),  # normalized versions of variables
+         elev0 = scale(elev),
+         dprime.sp0 = scale(dprime.sp),
+         dprime.ktype0 = scale(dprime.ktype),
+         dprime.fam0 = scale(dprime.fam)) %>%
+  select(site, date, bb.sp, dprime.sp, dprime.sp0, dprime.ktype, 
+         dprime.ktype0, dprime.fam, dprime.fam0, elev, elev0, 
+         yday, yday0, bb.sp.abund, bb.sp.prop.abund, total.flower.cover)
 ```
@@
 ## 1.3 Multi-level model: bias ~ yday * elevation + (1 + elevation | bb.sp)
+### 1.3.1 Preliminary visualization
 ```{r}
@@
 
+ggplot(bias, aes(dprime.sp)) +
+  geom_histogram()
+
 # First, let's look at the distribution of bias per BB sp
-ggplot(filter(bias, criterion == "plant.sp"), aes(dprime)) +
+ggplot(bias, aes(dprime.sp)) +
   geom_histogram() +
@@
 
+ggplot(bias, aes(dprime.ktype)) +
+  geom_histogram() +
+  facet_wrap(~bb.sp)
+
+ggplot(bias, aes(dprime.fam)) +
+  geom_histogram() +
+  facet_wrap(~bb.sp)
+
 # Now, let's try to get a sense for the shape of the response to elevation and time; we might expect to find nonlinearities here
-ggplot(filter(bias, criterion == "plant.sp"), aes(elev, dprime)) +
+ggplot(bias, aes(elev, dprime.sp)) +
   geom_point() +
@@
 
-ggplot(filter(bias, criterion == "plant.sp"), aes(yday, dprime)) +
+ggplot(bias, aes(yday, dprime.sp)) +
   geom_point() +
@@
 
+ggplot(bias, aes(elev, dprime.ktype)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~bb.sp)
+
+ggplot(bias, aes(yday, dprime.ktype)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~bb.sp)
+
+ggplot(bias, aes(elev, dprime.fam)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~bb.sp)
+
+ggplot(bias, aes(yday, dprime.fam)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~bb.sp)
+
 # After inspecting these plots, I see no reason to indulge in wiggly models.
-
-# This model includes a global interaction term for yday and elevation, with a varying intercept for bb.sp and a varying slow for elev | bb.sp.
-I think this should give me all the estimation I'm interested in.
-It's tricky to make it converge, though.
-brm.plantsp11 <- brm(dprime ~ yday * elev + (1 + elev | bb.sp), 
-                  control = list(max_treedepth = 10),
-                  cores = 4,
-                  iter = 4000,
-                  #family = "skew_normal",
-                  file = "../output/brm_plantsp11",
-                  data = filter(bias, criterion ==  "plant.sp"))
-
-summary(brm.plantsp11)
-prior_summary(brm.plantsp11)
-plot(brm.plantsp11)
-parnames(brm.plantsp11)
-mcmc_plot(brm.plantsp11, pars = c("^r_"))
-mcmc_plot(brm.plantsp11, pars = c("elev"))
-mcmc_plot(brm.plantsp11, pars = c("^b_"))
-mcmc_plot(brm.plantsp11)
-conditional_effects(brm.plantsp11)
-pp_check(brm.plantsp11)
-bayes_R2(brm.plantsp11)
-hypothesis(brm.plantsp11, "b_elev = 0", class = NULL)
-
-# Alternatively, we can drop the varying slope term to simplify the model a bit. Still doesn't quite converge.
-brm.plantsp1 <- brm(dprime ~ yday * elev + (1 | bb.sp), 
-                  control = list(max_treedepth = 10),
+```
+
+### 1.3.3 brms models
+There's a lot I don't understand about the standardization of variables in the context of Bayesian MLM, but for now I have opted to scale and center the predictor variables but not the response.
+My reason for scaling and centering the predictors is that is massively speeds up estimation and enables the interaction model to converge (it would not converge with raw variables).
+My reason for not scaling the response is that it doesn't change the outcome of the model (I've checked) and I am interested in seeing the global intercept for d' on its native scale so that I can compare it with previously reported values.
+Right now, I am using all default priors.
+This means flat priors for global effects and student priors for varying effects (I think).
+My understanding is that brms internally scales priors to match the scale of the data.
+If I want to define my own priors, which I know I really should, particularly to replace the flat priors on the global effects, then I need to understand how the scaling of the variables affects the specification of the prior.
+```{r}
+# elev*yday interaction model
+brm_dprime.sp00 <- brm(dprime.sp ~ elev0 * yday0 + (1 | bb.sp),
+                       #+ (1 | site),
                        cores = 4,
                        iter = 2000,
-                  file = "../output/brm_plantsp1",
-                  data = filter(bias, criterion ==  "plant.sp"))
-
-summary(brm.plantsp1)
-prior_summary(brm.plantsp1)
-plot(brm.plantsp1)
-parnames(brm.plantsp1)
-mcmc_plot(brm.plantsp1, pars = c("^r_"))
-mcmc_plot(brm.plantsp1, pars = c("elev"))
-mcmc_plot(brm.plantsp1, pars = c("^b_"))
-mcmc_plot(brm.plantsp1)
-conditional_effects(brm.plantsp1)
-pp_check(brm.plantsp1)
-bayes_R2(brm.plantsp1)
-hypothesis(brm.plantsp1, "b_elev = 0", class = NULL)
-
-# We can simplify the model still more by dropping the yday*elev interaction term.
-Now it converges.
-But I think the takeaway is that neither elevation noy yday give us a clear signal with respect  to dprime.
-brm.plantsp2 <- brm(dprime ~ yday + elev + (1 | bb.sp), 
-                  control = list(max_treedepth = 10),
-                  cores = 4,
-                  iter = 2000,
-                  file = "../output/brm_plantsp2",
-                  data = filter(bias, criterion ==  "plant.sp"))
-
-summary(brm.plantsp2)
-prior_summary(brm.plantsp2)
-plot(brm.plantsp2)
-parnames(brm.plantsp2)
-mcmc_plot(brm.plantsp2, pars = c("^r_"))
-mcmc_plot(brm.plantsp2, pars = c("elev"))
-mcmc_plot(brm.plantsp2, pars = c("^b_"))
-mcmc_plot(brm.plantsp2)
-conditional_effects(brm.plantsp2)
-pp_check(brm.plantsp2)
-bayes_R2(brm.plantsp2)
-hypothesis(brm.plantsp2, "b_elev = 0", class = NULL)
-
-# This, I think, would also be a legitimate way to specify the model, but it would not tell us anything about elevation or time.
-It's the simplest, and that's nice.
-Well-behaved.
-Captures about twice as much variance as the previous model, and it wins the WAIC context easily.
-brm.plantsp0 <- brm(dprime ~ 1 + (1 | bb.sp) + (1 | site),
-                  control = list(max_treedepth = 10),
-                  cores = 4,
-                  iter = 2000,
-                  file = "../output/brm_plantsp0",
-                  data = filter(bias, criterion ==  "plant.sp"))
-
-summary(brm.plantsp0)
-plot(brm.plantsp0)
-parnames(brm.plantsp0)
-
-# I guess the message here is that both bb.sp and site contribute to the variation of dprime, but site more so.
-Biologically, I think the meaning of prime variation by site is that some sites have greater relative abundance of *attractive* flora than others.
-A site with loads of something that is never visited and a little bit of clover will yield a high dprime, while a site with a high abundance of attractive floral would yield a low dprime.
-mcmc_plot(brm.plantsp0, pars = c("^r_"))
-mcmc_plot(brm.plantsp0, pars = c("^r_bb.sp")) 
-ggsave("../output/brm.plantsp0_plot1.png")
-mcmc_plot(brm.plantsp0, pars = c("^b_"))
-pp_check(brm.plantsp0)
-bayes_R2(brm.plantsp0)
-
-# Another option would be to drop just elevation from the model. This runs clean. Doesn't explain much more variance than the simpler model, though.
-brm.plantsp00 <- brm(dprime ~ yday + (1 | bb.sp) + (1 | site),
-                  control = list(max_treedepth = 10),
-                  cores = 4,
-                  iter = 2000,
-                  file = "../output/brm_plantsp00",
-                  data = filter(bias, criterion ==  "plant.sp"))
-
-summary(brm.plantsp00)
-plot(brm.plantsp00)
-parnames(brm.plantsp00)
-mcmc_plot(brm.plantsp00, pars = c("^r_"))
-mcmc_plot(brm.plantsp00, pars = c("^b_"))
-conditional_effects(brm.plantsp00)
-pp_check(brm.plantsp00)
-bayes_R2(brm.plantsp00)
-
-# Let's do a formal comparison of the last three models
-waic(brm.plantsp2, brm.plantsp0, brm.plantsp00)
-
-# The one including yday is a *slight* winner, but the margin is so small as to be meaningless. I think the simplest model is the one we want (brm.plantsp0).
-
-# Let's try the same model now, but for ktype.
-brm.ktype0 <- brm(dprime ~ 1 + (1 | bb.sp) + (1 | site),
-                  control = list(max_treedepth = 10),
+                       family = "skew_normal", # this accounts for a slight left-skew in the posterior
+                       file = "../output/brm_dprime.sp00",
+                       data = bias)
+
+summary(brm_dprime.sp00)
+prior_summary(brm_dprime.sp00)
+parnames(brm_dprime.sp00)
+pp_check(brm_dprime.sp00)
+bayes_R2(brm_dprime.sp00)
+
+# Intercept-only model
+brm_dprime.sp01 <- brm(dprime.sp ~ 1 + (1 | bb.sp) + (1 | site),
                        cores = 4,
@@
                        family = "skew_normal",
-                  file = "../output/brm_ktype0",
-                  data = filter(bias, criterion ==  "ktype"))
-
-summary(brm.ktype0)
-plot(brm.ktype0)
-parnames(brm.ktype0)
-mcmc_plot(brm.ktype0, pars = c("^r_"))
-mcmc_plot(brm.ktype0, pars = c("^b_"))
-conditional_effects(brm.ktype0)
-pp_check(brm.ktype0)
-bayes_R2(brm.ktype0)
-```
-
-## 1.4 Define hypotheses
-```{r}
-# Is d' different from 0?
-H1 <- paste("Intercept = 0")
-
-# Are some species more biased in their foraging than others? 
-H.bss_gers <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[gers,Intercept]")
-H.bss_hort <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[hort,Intercept]")
-H.bss_humi <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[humi,Intercept]")
-H.bss_hypn <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[hypn,Intercept]")
-H.bss_jone <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[jone,Intercept]")
-H.bss_lapi <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[lapi,Intercept]")
-H.bss_mend <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[mend,Intercept]")
-H.bss_mont <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[mont,Intercept]")
-H.bss_muci <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[muci,Intercept]")
-H.bss_pasc <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[pasc,Intercept]")
-H.bss_prat <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[prat,Intercept]")
-H.bss_psit <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[psit,Intercept]")
-H.bss_pyre <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[pyre,Intercept]")
-H.bss_soro <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[soro,Intercept]")
-H.bss_wurf <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.gers_hort <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[hort,Intercept]")
-H.gers_humi <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[humi,Intercept]")
-H.gers_hypn <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[hypn,Intercept]")
-H.gers_jone <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[jone,Intercept]")
-H.gers_lapi <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[lapi,Intercept]")
-H.gers_mend <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[mend,Intercept]")
-H.gers_mont <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[mont,Intercept]")
-H.gers_muci <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[muci,Intercept]")
-H.gers_pasc <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[pasc,Intercept]")
-H.gers_prat <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[prat,Intercept]")
-H.gers_psit <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[psit,Intercept]")
-H.gers_pyre <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[pyre,Intercept]")
-H.gers_soro <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[soro,Intercept]")
-H.gers_wurf <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.hort_humi <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[humi,Intercept]")
-H.hort_hypn <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[hypn,Intercept]")
-H.hort_jone <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[jone,Intercept]")
-H.hort_lapi <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[lapi,Intercept]")
-H.hort_mend <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[mend,Intercept]")
-H.hort_mont <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[mont,Intercept]")
-H.hort_muci <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[muci,Intercept]")
-H.hort_pasc <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[pasc,Intercept]")
-H.hort_prat <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[prat,Intercept]")
-H.hort_psit <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[psit,Intercept]")
-H.hort_pyre <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[pyre,Intercept]")
-H.hort_soro <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[soro,Intercept]")
-H.hort_wurf <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.humi_hypn <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[hypn,Intercept]")
-H.humi_jone <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[jone,Intercept]")
-H.humi_lapi <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[lapi,Intercept]")
-H.humi_mend <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[mend,Intercept]")
-H.humi_mont <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[mont,Intercept]")
-H.humi_muci <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[muci,Intercept]")
-H.humi_pasc <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[pasc,Intercept]")
-H.humi_prat <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[prat,Intercept]")
-H.humi_psit <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[psit,Intercept]")
-H.humi_pyre <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[pyre,Intercept]")
-H.humi_soro <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[soro,Intercept]")
-H.humi_wurf <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.hypn_jone <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[jone,Intercept]")
-H.hypn_lapi <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[lapi,Intercept]")
-H.hypn_mend <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[mend,Intercept]")
-H.hypn_mont <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[mont,Intercept]")
-H.hypn_muci <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[muci,Intercept]")
-H.hypn_pasc <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[pasc,Intercept]")
-H.hypn_prat <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[prat,Intercept]")
-H.hypn_psit <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[psit,Intercept]")
-H.hypn_pyre <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[pyre,Intercept]")
-H.hypn_soro <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[soro,Intercept]")
-H.hypn_wurf <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.jone_lapi <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[lapi,Intercept]")
-H.jone_mend <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[mend,Intercept]")
-H.jone_mont <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[mont,Intercept]")
-H.jone_muci <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[muci,Intercept]")
-H.jone_pasc <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[pasc,Intercept]")
-H.jone_prat <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[prat,Intercept]")
-H.jone_psit <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[psit,Intercept]")
-H.jone_pyre <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[pyre,Intercept]")
-H.jone_soro <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[soro,Intercept]")
-H.jone_wurf <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.lapi_mend <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[mend,Intercept]")
-H.lapi_mont <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[mont,Intercept]")
-H.lapi_muci <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[muci,Intercept]")
-H.lapi_pasc <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[pasc,Intercept]")
-H.lapi_prat <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[prat,Intercept]")
-H.lapi_psit <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[psit,Intercept]")
-H.lapi_pyre <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[pyre,Intercept]")
-H.lapi_soro <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[soro,Intercept]")
-H.lapi_wurf <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.mend_mont <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[mont,Intercept]")
-H.mend_muci <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[muci,Intercept]")
-H.mend_pasc <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[pasc,Intercept]")
-H.mend_prat <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[prat,Intercept]")
-H.mend_psit <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[psit,Intercept]")
-H.mend_pyre <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[pyre,Intercept]")
-H.mend_soro <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[soro,Intercept]")
-H.mend_wurf <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.mont_muci <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[muci,Intercept]")
-H.mont_pasc <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[pasc,Intercept]")
-H.mont_prat <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[prat,Intercept]")
-H.mont_psit <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[psit,Intercept]")
-H.mont_pyre <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[pyre,Intercept]")
-H.mont_soro <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[soro,Intercept]")
-H.mont_wurf <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.muci_pasc <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[pasc,Intercept]")
-H.muci_prat <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[prat,Intercept]")
-H.muci_psit <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[psit,Intercept]")
-H.muci_pyre <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[pyre,Intercept]")
-H.muci_soro <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[soro,Intercept]")
-H.muci_wurf <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.pasc_prat <- paste("r_bb.sp[pasc,Intercept] = r_bb.sp[prat,Intercept]")
-H.pasc_psit <- paste("r_bb.sp[pasc,Intercept] = r_bb.sp[psit,Intercept]")
-H.pasc_pyre <- paste("r_bb.sp[pasc,Intercept] = r_bb.sp[pyre,Intercept]")
-H.pasc_soro <- paste("r_bb.sp[pasc,Intercept] = r_bb.sp[soro,Intercept]")
-H.pasc_wurf <- paste("r_bb.sp[pasc,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.prat_psit <- paste("r_bb.sp[prat,Intercept] = r_bb.sp[psit,Intercept]")
-H.prat_pyre <- paste("r_bb.sp[prat,Intercept] = r_bb.sp[pyre,Intercept]")
-H.prat_soro <- paste("r_bb.sp[prat,Intercept] = r_bb.sp[soro,Intercept]")
-H.prat_wurf <- paste("r_bb.sp[prat,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.psit_pyre <- paste("r_bb.sp[psit,Intercept] = r_bb.sp[pyre,Intercept]")
-H.psit_soro <- paste("r_bb.sp[psit,Intercept] = r_bb.sp[soro,Intercept]")
-H.psit_wurf <- paste("r_bb.sp[psit,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.pyre_soro <- paste("r_bb.sp[pyre,Intercept] = r_bb.sp[soro,Intercept]")
-H.pyre_wurf <- paste("r_bb.sp[pyre,Intercept] = r_bb.sp[wurf,Intercept]")
-
-H.soro_wurf <- paste("r_bb.sp[soro,Intercept] = r_bb.sp[wurf,Intercept]")
-```
-
-## 1.5 Test contrasts
-```{r}
-### Plant.sp
-hypothesis(brm.plantsp, H1)
-
-hypothesis(brm.plantsp, H.bss_gers, class = "")
-hypothesis(brm.plantsp, H.bss_hort, class = "")
-hypothesis(brm.plantsp, H.bss_humi, class = "")
-hypothesis(brm.plantsp, H.bss_hypn, class = "")
-hypothesis(brm.plantsp, H.bss_jone, class = "")
-hypothesis(brm.plantsp, H.bss_lapi, class = "")
-hypothesis(brm.plantsp, H.bss_mend, class = "")
-hypothesis(brm.plantsp, H.bss_mont, class = "")
-hypothesis(brm.plantsp, H.bss_muci, class = "")
-hypothesis(brm.plantsp, H.bss_pasc, class = "")
-hypothesis(brm.plantsp, H.bss_prat, class = "")
-hypothesis(brm.plantsp, H.bss_psit, class = "")
-hypothesis(brm.plantsp, H.bss_pyre, class = "")
-hypothesis(brm.plantsp, H.bss_soro, class = "")
-hypothesis(brm.plantsp, H.bss_wurf, class = "")
-
-hypothesis(brm.plantsp, H.gers_hort, class = "")
-hypothesis(brm.plantsp, H.gers_humi, class = "")
-hypothesis(brm.plantsp, H.gers_hypn, class = "")
-hypothesis(brm.plantsp, H.gers_jone, class = "")
-hypothesis(brm.plantsp, H.gers_lapi, class = "")
-hypothesis(brm.plantsp, H.gers_mend, class = "")
-hypothesis(brm.plantsp, H.gers_mont, class = "")
-hypothesis(brm.plantsp, H.gers_muci, class = "")
-hypothesis(brm.plantsp, H.gers_pasc, class = "")
-hypothesis(brm.plantsp, H.gers_prat, class = "")
-hypothesis(brm.plantsp, H.gers_psit, class = "")
-hypothesis(brm.plantsp, H.gers_pyre, class = "")
-hypothesis(brm.plantsp, H.gers_soro, class = "")
-hypothesis(brm.plantsp, H.gers_wurf, class = "")
-
-hypothesis(brm.plantsp, H.hort_humi, class = "")
-hypothesis(brm.plantsp, H.hort_hypn, class = "")
-hypothesis(brm.plantsp, H.hort_jone, class = "")
-hypothesis(brm.plantsp, H.hort_lapi, class = "")
-hypothesis(brm.plantsp, H.hort_mend, class = "")
-hypothesis(brm.plantsp, H.hort_mont, class = "")
-hypothesis(brm.plantsp, H.hort_muci, class = "")
-hypothesis(brm.plantsp, H.hort_pasc, class = "")
-hypothesis(brm.plantsp, H.hort_prat, class = "")
-hypothesis(brm.plantsp, H.hort_psit, class = "")
-hypothesis(brm.plantsp, H.hort_pyre, class = "")
-hypothesis(brm.plantsp, H.hort_soro, class = "")
-hypothesis(brm.plantsp, H.hort_wurf, class = "")
-
-hypothesis(brm.plantsp, H.humi_hypn, class = "")
-hypothesis(brm.plantsp, H.humi_jone, class = "")
-hypothesis(brm.plantsp, H.humi_lapi, class = "")
-hypothesis(brm.plantsp, H.humi_mend, class = "")
-hypothesis(brm.plantsp, H.humi_mont, class = "")
-hypothesis(brm.plantsp, H.humi_muci, class = "")
-hypothesis(brm.plantsp, H.humi_pasc, class = "")
-hypothesis(brm.plantsp, H.humi_prat, class = "")
-hypothesis(brm.plantsp, H.humi_psit, class = "")
-hypothesis(brm.plantsp, H.humi_pyre, class = "")
-hypothesis(brm.plantsp, H.humi_soro, class = "")
-hypothesis(brm.plantsp, H.humi_wurf, class = "")
-
-hypothesis(brm.plantsp, H.hypn_jone, class = "")
-hypothesis(brm.plantsp, H.hypn_lapi, class = "")
-hypothesis(brm.plantsp, H.hypn_mend, class = "")
-hypothesis(brm.plantsp, H.hypn_mont, class = "")
-hypothesis(brm.plantsp, H.hypn_muci, class = "")
-hypothesis(brm.plantsp, H.hypn_pasc, class = "")
-hypothesis(brm.plantsp, H.hypn_prat, class = "")
-hypothesis(brm.plantsp, H.hypn_psit, class = "")
-hypothesis(brm.plantsp, H.hypn_pyre, class = "")
-hypothesis(brm.plantsp, H.hypn_soro, class = "")
-hypothesis(brm.plantsp, H.hypn_wurf, class = "")
-
-hypothesis(brm.plantsp, H.jone_lapi, class = "")
-hypothesis(brm.plantsp, H.jone_mend, class = "")
-hypothesis(brm.plantsp, H.jone_mont, class = "")
-hypothesis(brm.plantsp, H.jone_muci, class = "")
-hypothesis(brm.plantsp, H.jone_pasc, class = "")
-hypothesis(brm.plantsp, H.jone_prat, class = "")
-hypothesis(brm.plantsp, H.jone_psit, class = "")
-hypothesis(brm.plantsp, H.jone_pyre, class = "")
-hypothesis(brm.plantsp, H.jone_soro, class = "")
-hypothesis(brm.plantsp, H.jone_wurf, class = "")
-
-hypothesis(brm.plantsp, H.lapi_mend, class = "")
-hypothesis(brm.plantsp, H.lapi_mont, class = "")
-hypothesis(brm.plantsp, H.lapi_muci, class = "")
-hypothesis(brm.plantsp, H.lapi_pasc, class = "")
-hypothesis(brm.plantsp, H.lapi_prat, class = "")
-hypothesis(brm.plantsp, H.lapi_psit, class = "")
-hypothesis(brm.plantsp, H.lapi_pyre, class = "")
-hypothesis(brm.plantsp, H.lapi_soro, class = "")
-hypothesis(brm.plantsp, H.lapi_wurf, class = "")
-
-hypothesis(brm.plantsp, H.mend_mont, class = "")
-hypothesis(brm.plantsp, H.mend_muci, class = "")
-hypothesis(brm.plantsp, H.mend_pasc, class = "")
-hypothesis(brm.plantsp, H.mend_prat, class = "")
-hypothesis(brm.plantsp, H.mend_psit, class = "")
-hypothesis(brm.plantsp, H.mend_pyre, class = "")
-hypothesis(brm.plantsp, H.mend_soro, class = "")
-hypothesis(brm.plantsp, H.mend_wurf, class = "")
-
-hypothesis(brm.plantsp, H.mont_muci, class = "")
-hypothesis(brm.plantsp, H.mont_pasc, class = "")
-hypothesis(brm.plantsp, H.mont_prat, class = "")
-hypothesis(brm.plantsp, H.mont_psit, class = "")
-hypothesis(brm.plantsp, H.mont_pyre, class = "")
-hypothesis(brm.plantsp, H.mont_soro, class = "")
-hypothesis(brm.plantsp, H.mont_wurf, class = "")
-
-hypothesis(brm.plantsp, H.muci_pasc, class = "")
-hypothesis(brm.plantsp, H.muci_prat, class = "")
-hypothesis(brm.plantsp, H.muci_psit, class = "")
-hypothesis(brm.plantsp, H.muci_pyre, class = "")
-hypothesis(brm.plantsp, H.muci_soro, class = "")
-hypothesis(brm.plantsp, H.muci_wurf, class = "")
-
-hypothesis(brm.plantsp, H.pasc_prat, class = "")
-hypothesis(brm.plantsp, H.pasc_psit, class = "")
-hypothesis(brm.plantsp, H.pasc_pyre, class = "")
-hypothesis(brm.plantsp, H.pasc_soro, class = "")
-hypothesis(brm.plantsp, H.pasc_wurf, class = "")
-
-hypothesis(brm.plantsp, H.prat_psit, class = "")
-hypothesis(brm.plantsp, H.prat_pyre, class = "")
-hypothesis(brm.plantsp, H.prat_soro, class = "")
-hypothesis(brm.plantsp, H.prat_wurf, class = "")
-
-hypothesis(brm.plantsp, H.psit_pyre, class = "")
-hypothesis(brm.plantsp, H.psit_soro, class = "")
-hypothesis(brm.plantsp, H.psit_wurf, class = "")
-
-hypothesis(brm.plantsp, H.pyre_soro, class = "")
-hypothesis(brm.plantsp, H.pyre_wurf, class = "")
-
-hypothesis(brm.plantsp, H.soro_wurf, class = "")
-
-### Kugler type
-hypothesis(brm.ktype, H1)
-
-hypothesis(brm.ktype, H.bss_gers, class = "")
-hypothesis(brm.ktype, H.bss_hort, class = "")
-hypothesis(brm.ktype, H.bss_humi, class = "")
-hypothesis(brm.ktype, H.bss_hypn, class = "")
-hypothesis(brm.ktype, H.bss_jone, class = "")
-hypothesis(brm.ktype, H.bss_lapi, class = "")
-hypothesis(brm.ktype, H.bss_mend, class = "")
-hypothesis(brm.ktype, H.bss_mont, class = "")
-hypothesis(brm.ktype, H.bss_muci, class = "")
-hypothesis(brm.ktype, H.bss_pasc, class = "")
-hypothesis(brm.ktype, H.bss_prat, class = "")
-hypothesis(brm.ktype, H.bss_psit, class = "")
-hypothesis(brm.ktype, H.bss_pyre, class = "")
-hypothesis(brm.ktype, H.bss_soro, class = "")
-hypothesis(brm.ktype, H.bss_wurf, class = "")
-
-hypothesis(brm.ktype, H.gers_hort, class = "")
-hypothesis(brm.ktype, H.gers_humi, class = "")
-hypothesis(brm.ktype, H.gers_hypn, class = "")
-hypothesis(brm.ktype, H.gers_jone, class = "")
-hypothesis(brm.ktype, H.gers_lapi, class = "")
-hypothesis(brm.ktype, H.gers_mend, class = "")
-hypothesis(brm.ktype, H.gers_mont, class = "")
-hypothesis(brm.ktype, H.gers_muci, class = "")
-hypothesis(brm.ktype, H.gers_pasc, class = "")
-hypothesis(brm.ktype, H.gers_prat, class = "")
-hypothesis(brm.ktype, H.gers_psit, class = "")
-hypothesis(brm.ktype, H.gers_pyre, class = "")
-hypothesis(brm.ktype, H.gers_soro, class = "")
-hypothesis(brm.ktype, H.gers_wurf, class = "")
-
-hypothesis(brm.ktype, H.hort_humi, class = "")
-hypothesis(brm.ktype, H.hort_hypn, class = "")
-hypothesis(brm.ktype, H.hort_jone, class = "")
-hypothesis(brm.ktype, H.hort_lapi, class = "")
-hypothesis(brm.ktype, H.hort_mend, class = "")
-hypothesis(brm.ktype, H.hort_mont, class = "")
-hypothesis(brm.ktype, H.hort_muci, class = "")
-hypothesis(brm.ktype, H.hort_pasc, class = "")
-hypothesis(brm.ktype, H.hort_prat, class = "")
-hypothesis(brm.ktype, H.hort_psit, class = "")
-hypothesis(brm.ktype, H.hort_pyre, class = "")
-hypothesis(brm.ktype, H.hort_soro, class = "")
-hypothesis(brm.ktype, H.hort_wurf, class = "")
-
-hypothesis(brm.ktype, H.humi_hypn, class = "")
-hypothesis(brm.ktype, H.humi_jone, class = "")
-hypothesis(brm.ktype, H.humi_lapi, class = "")
-hypothesis(brm.ktype, H.humi_mend, class = "")
-hypothesis(brm.ktype, H.humi_mont, class = "")
-hypothesis(brm.ktype, H.humi_muci, class = "")
-hypothesis(brm.ktype, H.humi_pasc, class = "")
-hypothesis(brm.ktype, H.humi_prat, class = "")
-hypothesis(brm.ktype, H.humi_psit, class = "")
-hypothesis(brm.ktype, H.humi_pyre, class = "")
-hypothesis(brm.ktype, H.humi_soro, class = "")
-hypothesis(brm.ktype, H.humi_wurf, class = "")
-
-hypothesis(brm.ktype, H.hypn_jone, class = "")
-hypothesis(brm.ktype, H.hypn_lapi, class = "")
-hypothesis(brm.ktype, H.hypn_mend, class = "")
-hypothesis(brm.ktype, H.hypn_mont, class = "")
-hypothesis(brm.ktype, H.hypn_muci, class = "")
-hypothesis(brm.ktype, H.hypn_pasc, class = "")
-hypothesis(brm.ktype, H.hypn_prat, class = "")
-hypothesis(brm.ktype, H.hypn_psit, class = "")
-hypothesis(brm.ktype, H.hypn_pyre, class = "")
-hypothesis(brm.ktype, H.hypn_soro, class = "")
-hypothesis(brm.ktype, H.hypn_wurf, class = "")
-
-hypothesis(brm.ktype, H.jone_lapi, class = "")
-hypothesis(brm.ktype, H.jone_mend, class = "")
-hypothesis(brm.ktype, H.jone_mont, class = "")
-hypothesis(brm.ktype, H.jone_muci, class = "")
-hypothesis(brm.ktype, H.jone_pasc, class = "")
-hypothesis(brm.ktype, H.jone_prat, class = "")
-hypothesis(brm.ktype, H.jone_psit, class = "")
-hypothesis(brm.ktype, H.jone_pyre, class = "")
-hypothesis(brm.ktype, H.jone_soro, class = "")
-hypothesis(brm.ktype, H.jone_wurf, class = "")
-
-hypothesis(brm.ktype, H.lapi_mend, class = "")
-hypothesis(brm.ktype, H.lapi_mont, class = "")
-hypothesis(brm.ktype, H.lapi_muci, class = "")
-hypothesis(brm.ktype, H.lapi_pasc, class = "")
-hypothesis(brm.ktype, H.lapi_prat, class = "")
-hypothesis(brm.ktype, H.lapi_psit, class = "")
-hypothesis(brm.ktype, H.lapi_pyre, class = "")
-hypothesis(brm.ktype, H.lapi_soro, class = "")
-hypothesis(brm.ktype, H.lapi_wurf, class = "")
-
-hypothesis(brm.ktype, H.mend_mont, class = "")
-hypothesis(brm.ktype, H.mend_muci, class = "")
-hypothesis(brm.ktype, H.mend_pasc, class = "")
-hypothesis(brm.ktype, H.mend_prat, class = "")
-hypothesis(brm.ktype, H.mend_psit, class = "")
-hypothesis(brm.ktype, H.mend_pyre, class = "")
-hypothesis(brm.ktype, H.mend_soro, class = "")
-hypothesis(brm.ktype, H.mend_wurf, class = "")
-
-hypothesis(brm.ktype, H.mont_muci, class = "")
-hypothesis(brm.ktype, H.mont_pasc, class = "")
-hypothesis(brm.ktype, H.mont_prat, class = "")
-hypothesis(brm.ktype, H.mont_psit, class = "")
-hypothesis(brm.ktype, H.mont_pyre, class = "")
-hypothesis(brm.ktype, H.mont_soro, class = "")
-hypothesis(brm.ktype, H.mont_wurf, class = "")
-
-hypothesis(brm.ktype, H.muci_pasc, class = "")
-hypothesis(brm.ktype, H.muci_prat, class = "")
-hypothesis(brm.ktype, H.muci_psit, class = "")
-hypothesis(brm.ktype, H.muci_pyre, class = "")
-hypothesis(brm.ktype, H.muci_soro, class = "")
-hypothesis(brm.ktype, H.muci_wurf, class = "")
-
-hypothesis(brm.ktype, H.pasc_prat, class = "")
-hypothesis(brm.ktype, H.pasc_psit, class = "")
-hypothesis(brm.ktype, H.pasc_pyre, class = "")
-hypothesis(brm.ktype, H.pasc_soro, class = "")
-hypothesis(brm.ktype, H.pasc_wurf, class = "")
-
-hypothesis(brm.ktype, H.prat_psit, class = "")
-hypothesis(brm.ktype, H.prat_pyre, class = "")
-hypothesis(brm.ktype, H.prat_soro, class = "")
-hypothesis(brm.ktype, H.prat_wurf, class = "")
-
-hypothesis(brm.ktype, H.psit_pyre, class = "")
-hypothesis(brm.ktype, H.psit_soro, class = "")
-hypothesis(brm.ktype, H.psit_wurf, class = "")
-
-hypothesis(brm.ktype, H.pyre_soro, class = "")
-hypothesis(brm.ktype, H.pyre_wurf, class = "")
-
-hypothesis(brm.ktype, H.soro_wurf, class = "")
-
-### Plant family
-hypothesis(brm.family, H1)
-
-hypothesis(brm.family, H.bss_gers, class = "")
-hypothesis(brm.family, H.bss_hort, class = "")
-hypothesis(brm.family, H.bss_humi, class = "")
-hypothesis(brm.family, H.bss_hypn, class = "")
-hypothesis(brm.family, H.bss_jone, class = "")
-hypothesis(brm.family, H.bss_lapi, class = "")
-hypothesis(brm.family, H.bss_mend, class = "")
-hypothesis(brm.family, H.bss_mont, class = "")
-hypothesis(brm.family, H.bss_muci, class = "")
-hypothesis(brm.family, H.bss_pasc, class = "")
-hypothesis(brm.family, H.bss_prat, class = "")
-hypothesis(brm.family, H.bss_psit, class = "")
-hypothesis(brm.family, H.bss_pyre, class = "")
-hypothesis(brm.family, H.bss_soro, class = "")
-hypothesis(brm.family, H.bss_wurf, class = "")
-
-hypothesis(brm.family, H.gers_hort, class = "")
-hypothesis(brm.family, H.gers_humi, class = "")
-hypothesis(brm.family, H.gers_hypn, class = "")
-hypothesis(brm.family, H.gers_jone, class = "")
-hypothesis(brm.family, H.gers_lapi, class = "")
-hypothesis(brm.family, H.gers_mend, class = "")
-hypothesis(brm.family, H.gers_mont, class = "")
-hypothesis(brm.family, H.gers_muci, class = "")
-hypothesis(brm.family, H.gers_pasc, class = "")
-hypothesis(brm.family, H.gers_prat, class = "")
-hypothesis(brm.family, H.gers_psit, class = "")
-hypothesis(brm.family, H.gers_pyre, class = "")
-hypothesis(brm.family, H.gers_soro, class = "")
-hypothesis(brm.family, H.gers_wurf, class = "")
-
-hypothesis(brm.family, H.hort_humi, class = "")
-hypothesis(brm.family, H.hort_hypn, class = "")
-hypothesis(brm.family, H.hort_jone, class = "")
-hypothesis(brm.family, H.hort_lapi, class = "")
-hypothesis(brm.family, H.hort_mend, class = "")
-hypothesis(brm.family, H.hort_mont, class = "")
-hypothesis(brm.family, H.hort_muci, class = "")
-hypothesis(brm.family, H.hort_pasc, class = "")
-hypothesis(brm.family, H.hort_prat, class = "")
-hypothesis(brm.family, H.hort_psit, class = "")
-hypothesis(brm.family, H.hort_pyre, class = "")
-hypothesis(brm.family, H.hort_soro, class = "")
-hypothesis(brm.family, H.hort_wurf, class = "")
-
-hypothesis(brm.family, H.humi_hypn, class = "")
-hypothesis(brm.family, H.humi_jone, class = "")
-hypothesis(brm.family, H.humi_lapi, class = "")
-hypothesis(brm.family, H.humi_mend, class = "")
-hypothesis(brm.family, H.humi_mont, class = "")
-hypothesis(brm.family, H.humi_muci, class = "")
-hypothesis(brm.family, H.humi_pasc, class = "")
-hypothesis(brm.family, H.humi_prat, class = "")
-hypothesis(brm.family, H.humi_psit, class = "")
-hypothesis(brm.family, H.humi_pyre, class = "")
-hypothesis(brm.family, H.humi_soro, class = "")
-hypothesis(brm.family, H.humi_wurf, class = "")
-
-hypothesis(brm.family, H.hypn_jone, class = "")
-hypothesis(brm.family, H.hypn_lapi, class = "")
-hypothesis(brm.family, H.hypn_mend, class = "")
-hypothesis(brm.family, H.hypn_mont, class = "")
-hypothesis(brm.family, H.hypn_muci, class = "")
-hypothesis(brm.family, H.hypn_pasc, class = "")
-hypothesis(brm.family, H.hypn_prat, class = "")
-hypothesis(brm.family, H.hypn_psit, class = "")
-hypothesis(brm.family, H.hypn_pyre, class = "")
-hypothesis(brm.family, H.hypn_soro, class = "")
-hypothesis(brm.family, H.hypn_wurf, class = "")
-
-hypothesis(brm.family, H.jone_lapi, class = "")
-hypothesis(brm.family, H.jone_mend, class = "")
-hypothesis(brm.family, H.jone_mont, class = "")
-hypothesis(brm.family, H.jone_muci, class = "")
-hypothesis(brm.family, H.jone_pasc, class = "")
-hypothesis(brm.family, H.jone_prat, class = "")
-hypothesis(brm.family, H.jone_psit, class = "")
-hypothesis(brm.family, H.jone_pyre, class = "")
-hypothesis(brm.family, H.jone_soro, class = "")
-hypothesis(brm.family, H.jone_wurf, class = "")
-
-hypothesis(brm.family, H.lapi_mend, class = "")
-hypothesis(brm.family, H.lapi_mont, class = "")
-hypothesis(brm.family, H.lapi_muci, class = "")
-hypothesis(brm.family, H.lapi_pasc, class = "")
-hypothesis(brm.family, H.lapi_prat, class = "")
-hypothesis(brm.family, H.lapi_psit, class = "")
-hypothesis(brm.family, H.lapi_pyre, class = "")
-hypothesis(brm.family, H.lapi_soro, class = "")
-hypothesis(brm.family, H.lapi_wurf, class = "")
-
-hypothesis(brm.family, H.mend_mont, class = "")
-hypothesis(brm.family, H.mend_muci, class = "")
-hypothesis(brm.family, H.mend_pasc, class = "")
-hypothesis(brm.family, H.mend_prat, class = "")
-hypothesis(brm.family, H.mend_psit, class = "")
-hypothesis(brm.family, H.mend_pyre, class = "")
-hypothesis(brm.family, H.mend_soro, class = "")
-hypothesis(brm.family, H.mend_wurf, class = "")
-
-hypothesis(brm.family, H.mont_muci, class = "")
-hypothesis(brm.family, H.mont_pasc, class = "")
-hypothesis(brm.family, H.mont_prat, class = "")
-hypothesis(brm.family, H.mont_psit, class = "")
-hypothesis(brm.family, H.mont_pyre, class = "")
-hypothesis(brm.family, H.mont_soro, class = "")
-hypothesis(brm.family, H.mont_wurf, class = "")
-
-hypothesis(brm.family, H.muci_pasc, class = "")
-hypothesis(brm.family, H.muci_prat, class = "")
-hypothesis(brm.family, H.muci_psit, class = "")
-hypothesis(brm.family, H.muci_pyre, class = "")
-hypothesis(brm.family, H.muci_soro, class = "")
-hypothesis(brm.family, H.muci_wurf, class = "")
-
-hypothesis(brm.family, H.pasc_prat, class = "")
-hypothesis(brm.family, H.pasc_psit, class = "")
-hypothesis(brm.family, H.pasc_pyre, class = "")
-hypothesis(brm.family, H.pasc_soro, class = "")
-hypothesis(brm.family, H.pasc_wurf, class = "")
-
-hypothesis(brm.family, H.prat_psit, class = "")
-hypothesis(brm.family, H.prat_pyre, class = "")
-hypothesis(brm.family, H.prat_soro, class = "")
-hypothesis(brm.family, H.prat_wurf, class = "")
-
-hypothesis(brm.family, H.psit_pyre, class = "")
-hypothesis(brm.family, H.psit_soro, class = "")
-hypothesis(brm.family, H.psit_wurf, class = "")
-
-hypothesis(brm.family, H.pyre_soro, class = "")
-hypothesis(brm.family, H.pyre_wurf, class = "")
-
-hypothesis(brm.family, H.soro_wurf, class = "")
+                       file = "../output/brm_dprime.sp01",
+                       data = bias)
+
+summary(brm_dprime.sp01)
+prior_summary(brm_dprime.sp01)
+parnames(brm_dprime.sp01)
+pp_check(brm_dprime.sp01)
+bayes_R2(brm_dprime.sp01)
+
+# Compare interaction model and intercept-only model. This favors the intercept-only model.
+loo(brm_dprime.sp00, brm_dprime.sp01)
+```
+
+#### 1.3.3.1 Plot
+```{r}
+get_variables(brm_dprime.sp01)
+
+brm_dprime.sp01_plotframe <- brm_dprime.sp01 %>%
+  spread_draws(b_Intercept, r_bb.sp[condition, term]) %>% 
+  mutate(condition_mean = b_Intercept + r_bb.sp)
+
+brm_dprime.sp01_plotframe2 <- brm_dprime.sp01 %>%
+  spread_draws(b_Intercept, r_site[condition, term]) %>% 
+  mutate(condition_mean = b_Intercept + r_site)
+
+brm_dprime.sp01_subplot1 <- ggplot(brm_dprime.sp01_plotframe, aes(condition_mean, condition)) +
+  stat_halfeye() +
+  ylab("Bumble bee species") +
+  xlab("Discrimination (d')") +
+  theme_light()
+
+brm_dprime.sp01_subplot2 <- ggplot(brm_dprime.sp01_plotframe2, aes(condition_mean, condition)) +
+  stat_halfeye() +
+  ylab("Site") +
+  xlab("Discrimination (d')") +
+  theme_light()
+
+brm_dprime.sp01_plot1 <- brm_dprime.sp01_subplot1 | brm_dprime.sp01_subplot2
+
+ggsave(brm_dprime.sp01_plot1, file = "../output/brm_dprime.sp01_plot1.png")
 ```
@@
 # 2. Explaining bias
+## 2.1 Calculate visitation probability per plant speces/ktype/family
+```{r}
+pvis_ktype <- visit_cover %>%
+  group_by(site, date, bb.sp, plant.sp) %>%
+  summarize(flower.cover = sum(flower.cover), # get raw flower cover and visits by site*date*bb.sp*plant.sp
+            visits = sum(visits)) %>%
+  group_by(site, date) %>%
+  mutate(prop.flower.cover = flower.cover/sum(flower.cover), # proportional flower cover
+         prop.visits = visits/sum(visits), # proportional visits
+         visits.bin = factor(if_else(visits > 0, 1, 0))) %>% # binary visits
+  group_by(site, date, bb.sp) %>%
+  mutate(bb.sp.abund = sum(visits)) %>% # abundance of each BB sp. per site-date
+  ungroup() %>%
+  left_join(fl_traits) %>% # add floral traits
+  left_join(bb_traits) %>% # add BB traits
+  mutate(k.type.ss = factor(k.type.ss), # recode variables
+         pbl.w.class2 = factor(pbl.w.class2),
+         visits = as.integer(visits)) %>%
+  filter(!k.type.ss %in% c("9", "10")) %>% # Drop ktypes 9 and 10 since both are extremely rare
+  left_join(site_data)  %>%
+  mutate(prop.flower.cover0 = scale(prop.flower.cover), # normalized predictors
+         bb.sp.abund0 = scale(bb.sp.abund),
+         elev.mean0 = scale(elev.mean),
+         yday = yday(date), # and add yday variable
+         yday0 = scale(yday)) %>%
+  select(site, elev.mean0, elev.mean, elev.class2, date, yday, yday0, bb.sp, plant.sp, visits, 
+         visits.bin, prop.visits, prop.flower.cover, prop.flower.cover0,
+         k.type.ss, bb.sp.abund, bb.sp.abund0, pbl.w.class2) 
+```
+
+## 2.2 Multi-level models
+
+### 2.2.1 Preliminary visualization
+The tons of zeros are due to the lots of plant species that were present at very low abundance and received few or zero visits.
+This might not be ideal.
+We could opt for one of our alternative bias measure, like fold change, which all give something like a poisson distribution, but I don't know if this would be an improvement.
+We could also filter our data to include only plants that were above some threshold of relative abundance, say, 1%.
+```{r}
+# Visit.prop
+ggplot(filter(pvis_ktype, prop.flower.cover >= 0.05), aes(prop.visits)) +
+  geom_histogram()
+
+# Visits.bin
+ggplot(pvis_ktype, aes(k.type.ss, fill = visits.bin)) +
+  geom_bar()
+
+ggplot(pvis_ktype, aes(elev.mean, prop.visits)) +
+  geom_point() +
+  geom_smooth() +
+  facet_grid(bb.sp~k.type.ss)
+
+# Visits.bin filtered
+ggplot(filter(pvis_ktype, prop.flower.cover >= 0.05), aes(k.type.ss, fill = visits.bin)) +
+  geom_bar()
+```
+
+### 2.2.2 brms models
+```{r}
+# Poisson models of raw visit counts
+brm_pvis00 <- brm(visits ~ prop.flower.cover0 + 
+                    bb.sp.abund0 + 
+                    (1 | k.type.ss) + 
+                    (1 | site) + 
+                    (1 | bb.sp),
+                  #control = list(adapt_delta = 0.99),
+                  family = "poisson",
+                  file = "./output/brm_pvis00b",
+                  chains = 4,
+                  #cores = 4,
+                  iter = 2000,
+                  data = pvis_ktype)
+
+summary(brm_pvis00)
+prior_summary(brm_pvis00)
+parnames(brm_pvis00)
+mcmc_plot(brm_pvis00, pars = c("^r_k"))
+mcmc_plot(brm_pvis00, pars = c("^r_s"))
+mcmc_plot(brm_pvis00, pars = c("^b_"))
+mcmc_plot(brm_pvis00)
+conditional_effects(brm_pvis00)
+pp_check(brm_pvis00)
+bayes_R2(brm_pvis00)
+
+# Perhaps I should add the bb.sp varying intercept effect; otherwise, I think I could be accused of pseudo replication.
+brm_pvis11 <- brm(visits ~ prop.flower.cover0 + 
+                    bb.sp.abund0 + 
+                    (1 | k.type.ss:pbl.w.class2) + 
+                    (1 | site) +
+                    (1 | bb.sp),
+                  family = "poisson",
+                  file = "../output/brm_pvis11",
+                  chains = 4,
+                  #cores = 4,
+                  iter = 2000,
+                  data = pvis_ktype)
+
+summary(brm_pvis11)
+prior_summary(brm_pvis11)
+parnames(brm_pvis11)
+mcmc_plot(brm_pvis11, pars = c("^r_k"))
+mcmc_plot(brm_pvis11, pars = c("^r_b"))
+mcmc_plot(brm_pvis11, pars = c("^b_"))
+mcmc_plot(brm_pvis11)
+conditional_effects(brm_pvis11)
+pp_check(brm_pvis11)
+bayes_R2(brm_pvis11)
+```
+
+#### 2.2.2.1 Plot
+```{r}
+# Note that the intercept is on the log scale, so if you want to transform it back to predicted # visits, you have to exponentiate it. I think.
+
+## Interaction model
+brm_pvis11_plotframe <- brm_pvis11 %>%
+  spread_draws(b_Intercept, `r_k.type.ss:pbl.w.class2`[condition, term]) %>% # the back ticks are necessary to escape the :
+  rename(r_ktype.pbl = `r_k.type.ss:pbl.w.class2`) %>%
+  mutate(condition_mean = b_Intercept + r_ktype.pbl) %>%
+  separate(condition, c("ktype", "pbl.class"), sep = "_", remove = FALSE) %>%
+  mutate(pbl.class = case_when(
+    pbl.class == "1" ~ "short",
+    pbl.class == "2" ~ "med",
+    pbl.class == "3" ~ "long")
+    ) %>%
+  mutate(ktype = case_when(
+    ktype == "0" ~ "NA (0)",
+    ktype == "1" ~ "Disc (1)",
+    ktype == "2" ~ "Funnel (2)",
+    ktype == "3" ~ "Bell (3)",
+    ktype == "4" ~ "Stalk-disc (4)",
+    ktype == "5" ~ "Lip (5)",
+    ktype == "6" ~ "Flag (6)",
+    ktype == "7" ~ "Head (7)")
+  ) %>%
+  mutate(ktype = factor(ktype,
+                        levels = c("Disc (1)", "Funnel (2)", "Bell (3)", "Stalk-disc (4)",
+                                   "Lip (5)", "Flag (6)", "Head (7)", "NA (0)")))
+
+# ggplot(brm_pvis11_plotframe, aes(condition_mean, pbl.class, color = pbl.class, fill = pbl.class)) +
+#   stat_pointinterval() +
+#   #stat_halfeye(alpha = 0.5, interval_alpha = 1, point_alpha = 1, size = 1, .width = 0.95) +
+#   facet_grid(ktype ~ .) +
+#   theme_light() +
+#   labs(color = "Tongue length", x = "Visits (log-scale)") +
+#   guides(fill = FALSE) +
+#   theme(axis.title.y=element_blank(),
+#         axis.text.y=element_blank(),
+#         axis.ticks.y=element_blank())
+
+ggplot(brm_pvis11_plotframe, aes(condition_mean, pbl.class, color = pbl.class, fill = pbl.class)) +
+  #stat_pointinterval() +
+  stat_halfeye(alpha = 0.4, interval_alpha = 1, point_alpha = 1) +
+  facet_wrap(~ktype, scales = "free") +
+  theme_light() +
+  labs(color = "Tongue length", x = "Visits (log-scale)") +
+  guides(fill = FALSE) +
+  theme(axis.title.y=element_blank(),
+        axis.text.y=element_blank(),
+        axis.ticks.y=element_blank())
+
+ggsave(file = "../output/brm_pvis11_plot1.png", width = 7, height = 5.5)
+
+# Just ktype model
+brm_pvis00_plotframe <- brm_pvis00 %>%
+  spread_draws(b_Intercept, r_k.type.ss[condition, term]) %>% # the back ticks are necessary to escape the :
+  mutate(condition_mean = b_Intercept + r_k.type.ss) %>%
+  mutate(ktype = case_when(
+    condition == "0" ~ "NA",
+    condition == "1" ~ "Disc",
+    condition == "2" ~ "Funnel",
+    condition == "3" ~ "Bell",
+    condition == "4" ~ "Stalk-disc",
+    condition == "5" ~ "Lip",
+    condition == "6" ~ "Flag",
+    condition == "7" ~ "Head")
+  ) %>%
+  mutate(ktype = factor(ktype,
+                        levels = c("Disc", "Funnel", "Bell", "Stalk-disc",
+                                   "Lip", "Flag", "Head", "NA")))
+
+ggplot(brm_pvis00_plotframe, aes(condition_mean, ktype)) +
+  stat_halfeye() +
+  theme_light() +
+  guides(fill = FALSE) +
+  xlab("Visits (log-scale)") +
+  ylab("Morphotype")
+
+ggsave(file = "../output/brm_pvis00_plot1.png", width = 7, height = 5.5)
+```
+
+### 2.2.3 The lability of bias: How does the foraging bias of each species with respect to each morphotype vary through elevation and time?
+```{r}
+ggplot(pvis_ktype, aes(elev.mean, prop.visits)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_grid(bb.sp ~ k.type.ss)
+
+ggplot(pvis_ktype, aes(elev.mean, prop.visits)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_grid(bb.sp ~ k.type.ss)
+
+ggplot(pvis_ktype, aes(yday0, prop.visits)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_grid(bb.sp ~ k.type.ss)
+
+brm_pviselev00 <- brm(visits ~ prop.flower.cover0 + 
+                        bb.sp.abund0 + 
+                        (1 + elev.mean0:yday0 | k.type.ss:bb.sp),
+                      family = "poisson",
+                      file = "../output/brm_pviselev00",
+                      chains = 4,
+                      cores = 4,
+                      iter = 2000,
+                      data = pvis_ktype)
+
+summary(brm_pviselev00)
+prior_summary(brm_pviselev00)
+parnames(brm_pviselev00)
+mcmc_plot(brm_pviselev00, pars = c("^r_k"))
+mcmc_plot(brm_pviselev00, pars = c("^r_s"))
+mcmc_plot(brm_pviselev00, pars = c("^b_"))
+mcmc_plot(brm_pviselev00)
+conditional_effects(brm_pviselev00)
+pp_check(brm_pviselev00)
+bayes_R2(brm_pviselev00)
+```
+
+# 3. Effects of bias
+
+## 3.1 Relationship between bias(d') and niche overlap
+### 3.1.1 A function for calculating niche overlap
+Niche overlap (Horn's distance) is first calculated for each species pair and each site-date.
+Naturally, site-dates in which only one BB species was observed return empty niche overlap matrices and so are excluded from further analysis.
+```{r}
+# Niche overlap by sp.pair * site * year
+niche_overlapper <- function(x, criterion) {
+
+  criterion <- enquo(criterion) # criterion is set to either plant.sp (for taxonomic) or k.type.s (for morphological)
+
+  no1 <- x %>%
+    group_by(bb.sp, !!criterion, site, date) %>% # group by site-year and BB-FL pair
+    summarize(visits = sum(visits)) %>%
+    group_by(bb.sp, site, date) %>% # group by site-year and BB species
+    mutate(prop.visits = visits/sum(visits)) %>% # get proportional interaction frequency for each BB-FL pair
+    select(site, date, bb.sp, !!criterion, prop.visits) %>%
+    pivot_wider(names_from = !!criterion, values_from = prop.visits, values_fill = 0) %>% # spread into the kind of array vegan likes
+    group_by(site, date) %>%
+    nest() %>% # nest into list-column
+    mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
+           richness = map(data, function(x) length(x$bb.sp)),
+           dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
+           tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
+    #dplyr::select(site, date, tidy_dist) %>% # drop intermediate list-columns
+    unnest(cols = c(tidy_dist)) %>%
+    rename(bb.sp1 = item1, bb.sp2 = item2) %>%
+    mutate(niche.overlap = 1-distance,  # express niche overlap so that 1 = perfect overlap and 0 = no overlap
+           telu = if_else(bb.sp1 == "telu" | bb.sp2 == "telu", TRUE, FALSE),
+           pasc = if_else(bb.sp1 == "pasc" | bb.sp2 == "pasc", TRUE, FALSE),
+           wurf = if_else(bb.sp1 == "wurf" | bb.sp2 == "wurf", TRUE, FALSE),
+           prat = if_else(bb.sp1 == "prat" | bb.sp2 == "prat", TRUE, FALSE),
+           soro = if_else(bb.sp1 == "soro" | bb.sp2 == "soro", TRUE, FALSE),
+           humi = if_else(bb.sp1 == "humi" | bb.sp2 == "humi", TRUE, FALSE),
+           muci = if_else(bb.sp1 == "muci" | bb.sp2 == "muci", TRUE, FALSE),
+           pyre = if_else(bb.sp1 == "pyre" | bb.sp2 == "pyre", TRUE, FALSE),
+           mont = if_else(bb.sp1 == "mont" | bb.sp2 == "mont", TRUE, FALSE),
+           mend = if_else(bb.sp1 == "mend" | bb.sp2 == "mend", TRUE, FALSE),
+           psit = if_else(bb.sp1 == "psit" | bb.sp2 == "psit", TRUE, FALSE),
+           gers = if_else(bb.sp1 == "gers" | bb.sp2 == "gers", TRUE, FALSE),
+           jone = if_else(bb.sp1 == "jone" | bb.sp2 == "jone", TRUE, FALSE),
+           hypn = if_else(bb.sp1 == "hypn" | bb.sp2 == "hypn", TRUE, FALSE),
+           lapi = if_else(bb.sp1 == "lapi" | bb.sp2 == "lapi", TRUE, FALSE),
+           hort = if_else(bb.sp1 == "hort" | bb.sp2 == "hort", TRUE, FALSE))
+}
+```
+
+### 3.1.2 My first question is how niche overlap compares across species pairs
+```{r}
+# Unaggregated niche overlap per species pair; not that this drops a lot of data because of the inner join with bias
+niche_overlap <- read_csv("../data/processed_data/network.csv") %>%
+  filter(caste != "male") %>% # we will ignore males 
+  group_by(site, date, bb.sp, plant.sp) %>%
+  summarize(visits = n()) %>%
+  niche_overlapper(plant.sp) %>%
+  mutate(yday = yday(date)) %>%
+  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>%
+  rename(pbl.w.class2_sp1 = pbl.w.class2) %>%
+  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>%
+  rename(pbl.w.class2_sp2 = pbl.w.class2) %>%
+  mutate(same.class = case_when(
+    pbl.w.class2_sp1 == "1" & pbl.w.class2_sp2 == "1" ~ "short:short",
+    pbl.w.class2_sp1 == "2" & pbl.w.class2_sp2 == "2" ~ "med:med",
+    pbl.w.class2_sp1 == "3" & pbl.w.class2_sp2 == "3" ~ "long:long",
+    pbl.w.class2_sp1 == "1" & pbl.w.class2_sp2 == "2" ~ "short:med",
+    pbl.w.class2_sp1 == "2" & pbl.w.class2_sp2 == "1" ~ "short:med",
+    pbl.w.class2_sp1 == "1" & pbl.w.class2_sp2 == "3" ~ "short:long",
+    pbl.w.class2_sp1 == "3" & pbl.w.class2_sp2 == "1" ~ "short:long",
+    pbl.w.class2_sp1 == "2" & pbl.w.class2_sp2 == "3" ~ "med:long",
+    pbl.w.class2_sp1 == "3" & pbl.w.class2_sp2 == "2" ~ "med:long")
+    ) %>%
+  left_join(bb_abundance, by = c("site", "date", "bb.sp1" = "bb.sp")) %>%
+  rename(bb.sp1.abund = bb.sp.abund) %>%
+  left_join(bb_abundance, by = c("site", "date", "bb.sp2" = "bb.sp")) %>%
+  rename(bb.sp2.abund = bb.sp.abund) %>%
+  unite(sp.pair, c(bb.sp1, bb.sp2), sep = "_", remove = FALSE) %>%
+  left_join(site_data) %>%
+  mutate(yday = yday(date)) %>%
+  select(site, date, yday, elev.mean, sp.pair, bb.sp1, bb.sp2, bb.sp1.abund, 
+         bb.sp2.abund, niche.overlap, same.class)
+
+# Niche overlap becomes rather binary when one or both species is rare. This is driven especially by singletons.
+ggplot(niche_overlap, aes(sp.pair, niche.overlap)) +
+  geom_violin() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+# Let's see if we can improve matters by dropping singletons
+ggplot(filter(niche_overlap, bb.sp1.abund >= 2 & bb.sp1.abund >= 2), 
+       aes(sp.pair, niche.overlap)) +
+  geom_violin() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+# Now a threshold of 3
+ggplot(filter(niche_overlap, bb.sp1.abund >= 3 & bb.sp1.abund >= 3), 
+       aes(sp.pair, niche.overlap)) +
+  geom_violin() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+ggplot(filter(niche_overlap, bb.sp1.abund >= 3 & bb.sp1.abund >= 3), 
+       aes(same.class, niche.overlap)) +
+  geom_violin() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+# Now a threshold of 5...bye, all the species pairs...
+ggplot(filter(niche_overlap, bb.sp1.abund >= 5 & bb.sp1.abund >= 5), 
+       aes(sp.pair, niche.overlap)) +
+  geom_violin() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+ggplot(filter(niche_overlap, bb.sp1.abund >= 10 & bb.sp1.abund >= 10), 
+       aes(sp.pair, niche.overlap)) +
+  geom_violin() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+# Let's see if there is any influence of site, elevation, yday
+ggplot(filter(niche_overlap, bb.sp1.abund >= 3 & bb.sp1.abund >= 3), 
+       aes(site, niche.overlap)) +
+  geom_boxplot() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+ggplot(filter(niche_overlap, bb.sp1.abund >= 3 & bb.sp1.abund >= 3), 
+       aes(elev.mean, niche.overlap)) +
+  geom_point() +
+  geom_smooth() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+ggplot(filter(niche_overlap, bb.sp1.abund >= 3 & bb.sp1.abund >= 3), 
+       aes(yday, niche.overlap)) +
+  geom_point() +
+  geom_smooth() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+# Abundance relatonships
+ggplot(niche_overlap, aes(log(bb.sp1.abund), log(bb.sp2.abund))) +
+  geom_point() +
+  geom_smooth(se = FALSE) +
+  facet_wrap(~sp.pair)
+
+#
+ggplot(niche_overlap, aes(same.class, niche.overlap)) +
+  geom_boxplot()
+
+# Now for what is probably a pretty naive model
+
+brm_niche00 <- brm(niche.overlap ~ bb.sp1.abund*bb.sp2.abund + 
+                     (1 | sp.pair) + 
+                     (1 | same.class) + 
+                     (1 | site),
+                   chains = 4,
+                   cores = 4,
+                   iter = 2000,
+                   file = "../output/brm_niche00",
+                   data = filter(niche_overlap, bb.sp1.abund >= 1 & bb.sp1.abund >= 1))
+
+parnames(brm_niche02)
+conditional_effects(brm_niche02)
+mcmc_plot(brm_niche02, pars = c("^r_sp"))
+mcmc_plot(brm_niche02, pars = c("^r_sa"))
+mcmc_plot(brm_niche02, pars = c("^b_"))
+hypothesis(brm_niche02, "dprime = 0")
+pp_check(brm_niche02)
+bayes_R2(brm_niche02)
+
+```
+
+### 3.1.3 Now I'd be interested to know how niche overlap relates to dprime
+```{r}
+niche_overlap_bias <- niche_overlap %>%
+  inner_join(bias, by = c("site", "date", c("bb.sp1" = "bb.sp"))) %>%
+  rename(dprime.sp1 = dprime.sp) %>%
+  inner_join(bias, by = c("site", "date", c("bb.sp2" = "bb.sp"))) %>%
+  rename(dprime.sp2 = dprime.sp) %>%
+  select(site, date, bb.sp1, bb.sp2, bb.sp1.abund, bb.sp2.abund, 
+         dprime.sp1, dprime.sp2, niche.overlap, same.class)
+
+# When bias is high, it could mean that there are few good options on which all (or many) BB species converge.
+But apparently it doesn't.
+When bias is high, BB spp are biased toward DIFFERENT flora, and thus their niches diverge.
+ggplot(niche_overlap_bias, aes(dprime.sp, mean.niche.overlap)) +
+  geom_point(alpha = 0.1) +
+  geom_smooth(method = "lm") +
+  #geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = TRUE) +
+  facet_wrap(~bb.sp) +
+  ylim(c(0,1))
+
+brm.bias_niche <- brm(mean.niche.overlap ~ dprime.sp + (1 + dprime.sp | bb.sp) + (1 + dprime.sp | site),
+                     cores = 4,
+                     chains = 4,
+                     iter = 2000,
+                     file = "../output/brm_bias_nicheoverlap",
+                     data = bias_nicheoverlap)
+
+parnames(brm.bias_niche)
+conditional_effects(brm.bias_niche)
+mcmc_plot(brm.bias_niche, pars = c("^r_"))
+hypothesis(brm.bias_niche, "dprime = 0")
+pp_check(brm.bias_niche)
+bayes_R2(brm.bias_niche)
+
+brm.dprime_niche00 <- brm(mean.niche.overlap ~ dprime.sp + (1 | bb.sp) + (1 | site),
+                     cores = 4,
+                     chains = 4,
+                     iter = 2000,
+                     file = "../output/brm_bias_niche00",
+                     data = bias_nicheoverlap)
+
+parnames(brm.dprime_niche00)
+conditional_effects(brm.dprime_niche00)
+mcmc_plot(brm.dprime_niche00, pars = c("^r_bb"))
+mcmc_plot(brm.dprime_niche00, pars = c("^b_"))
+hypothesis(brm.dprime_niche00, "dprime = 0")
+pp_check(brm.dprime_niche00)
+bayes_R2(brm.dprime_niche00)
+```
+
+## 3.2 Ordination of dietary niche
+```{r}
+# Aggregated by mean proportional visitation
+
+# We will use the whole dataset this time, since there is no need to align visitation to floral surveying
+visitation1 <- read_csv("../data/processed_data/network.csv") %>%
+  filter(caste != "male") %>% # we will ignore males 
+  group_by(site, date, bb.sp, plant.sp) %>%
+  summarize(visits = n()) %>%
+  group_by(site, date, bb.sp) %>%
+  mutate(prop.visits = visits/sum(visits)) %>%
+  group_by(bb.sp, plant.sp) %>%
+  summarize(mean.prop.visits = mean(prop.visits)) %>%
+  pivot_wider(names_from = plant.sp, values_from = mean.prop.visits, values_fill = 0) %>%
+  column_to_rownames("bb.sp")
+
+visitation2 <- read_csv("../data/processed_data/network.csv") %>%
+  filter(caste != "male") %>% # we will ignore males 
+  group_by(site, date, bb.sp, plant.sp) %>%
+  summarize(visits = n()) %>%
+  group_by(site, date, bb.sp) %>%
+  mutate(prop.visits = visits/sum(visits)) %>%
+  unite(id, c(site, date, bb.sp)) %>%
+  select(-visits) %>%
+  pivot_wider(names_from = plant.sp, values_from = prop.visits, values_fill = 0) %>%
+  column_to_rownames("id")
+
+# Aggregated
+pcoa_horn1 <- capscale(visitation1 ~ 1, distance = "horn")
+pcoa_horn1.fort1 <- fortify(pcoa_horn1, axes = 1:2) %>%
+  filter(Score == "species") %>%
+  rename(plant.sp = Label) %>%
+  left_join(fl_traits)
+pcoa_horn1.fort2 <- fortify(pcoa_horn1, axes = 1:2) %>%
+  filter(Score == "sites") %>%
+  rename(bb.sp = Label) %>%
+  left_join(bb_traits)
+
+ggplot(pcoa_horn1.fort2, aes(MDS1, MDS2)) +
+  geom_text(aes(label = bb.sp)) +
+  coord_fixed()
+
+# Disaggregated
+pcoa_horn <- capscale(visitation2 ~ 1, distance = "horn")
+pcoa_horn.fort <- fortify(pcoa_horn, axes = 1:2)
+pcoa_horn.fort1 <- pcoa_horn.fort %>%
+  filter(Score == "species") %>%
+  rename(plant.sp = Label) %>%
+  left_join(fl_traits)
+pcoa_horn.fort2 <- pcoa_horn.fort %>%
+  filter(Score == "sites") %>%
+  separate(Label, c("site", "date", "bb.sp"), sep = "_") %>%
+  left_join(bb_traits)
+
+ggplot(pcoa_horn.fort1, aes(MDS1, MDS2, color = factor(k.type.ss))) +
+  geom_point(alpha = 0.25) +
+  coord_fixed()
+
+ggplot(pcoa_horn.fort2, aes(MDS1, MDS2, color = factor(pbl.w.class2))) +
+  geom_point(alpha = 0.5) +
+  coord_fixed()
+
+ggplot(pcoa_horn.fort2, aes(MDS1, MDS2, color = factor(pbl.w.class2))) +
+  geom_density2d() +
+  coord_fixed()
+
+visitation3 <- read_csv("../data/processed_data/network.csv") %>%
+  filter(caste != "male") %>% # we will ignore males 
+  group_by(bb.sp, plant.sp) %>%
+  summarize(visits = n()) %>%
+  group_by(bb.sp) %>%
+  mutate(prop.visits = visits/sum(visits))
+
+visitation_aggregate2 <- visitation2 %>%
+  group_by(bb.sp, plant.sp) %>%
+  summarize(mean.prop.visits = mean(prop.visits)) %>%
+  pivot_wider(names_from = plant.sp, values_from = mean.prop.visits, values_fill = 0) %>%
+  column_to_rownames("bb.sp")
+
+visitation_aggregate3 <- visitation3 %>%
+  pivot_wider(names_from = plant.sp, values_from = prop.visits, values_fill = 0) %>%
+  column_to_rownames("bb.sp")
+
+bb_traits2 <- bb_traits %>%
+  select(bb.sp, pbl.w.class2) %>%
+  arrange(bb.sp)
+
+pcoa_horn <- capscale(visitation_aggregate ~ 1, distance = "horn")
+envfit_horn <- envfit(pcoa_horn ~ pbl.w.class2, data = bb_traits2)
+plot(pcoa_horn, type = "t")
+plot(envfit_horn, add = TRUE)
+pcoa_horn.fort <- fortify(pcoa_horn, axes = 1:2)
+
+# What if I did not aggregate the data?
+visit_cover_prop.plantsp2 <- visit_cover %>%
+  filter(bb.sp %in% c("pasc", "prat", "telu", "soro", "wurf", "psit")) %>%
+  select(site, date, bb.sp, plant.sp, k.type.ss, visits, flower.cover) %>%
+  group_by(site, date, bb.sp) %>%
+  mutate(prop.visits = visits/sum(visits)) %>%
+  group_by(site, date, plant.sp) %>%
+  mutate(prop.flower.cover = flower.cover/sum(flower.cover)) %>%
+  unite(id, c(site, date, bb.sp), remove = TRUE) %>%
+  select(id, plant.sp, prop.visits) %>%
+  pivot_wider(names_from = plant.sp, values_from = prop.visits, values_fill = 0) %>%
+  column_to_rownames("id")
+
+traits <- bb_traits %>%
+  filter(bb.sp %in% c("pasc", "prat", "telu", "soro", "wurf", "psit")) %>%
+  select(bb.sp, pbl.w.class2) %>%
+  arrange(bb.sp)
+
+pcoa_horn2 <- capscale(visit_cover_prop.plantsp2 ~ 1, distance = "horn")
+envfit_horn2 <- envfit(pcoa_horn2 ~ pbl.w.class2, data = traits)
+plot(pcoa_horn2, type = "t")
+plot(envfit_horn2, add = TRUE)
+
+pcoa_horn2_fort <- fortify(pcoa_horn2)
+
+pcoa_horn2_plot <- pcoa_horn2_fort %>%
+  filter(Score == "sites") %>%
+  separate(Label, c("site", "date", "bb.sp"), sep = "_") %>%
+  mutate(date = ymd(date)) %>%
+  mutate(yday = yday(date))
+
+ggplot(pcoa_horn2_plot, aes(MDS1, MDS2, color = bb.sp)) +
+  geom_point() +
+  coord_fixed()
+
+ggplot(pcoa_horn2_plot, aes(MDS1, MDS2, color = yday)) +
+  geom_point() +
+  coord_fixed() 
+```
+
+# Trash can
+
+```{r}
+niche_overlapper <- function(x, criterion) {
+
+  criterion <- enquo(criterion) # criterion is set to either plant.sp (for taxonomic) or k.type.s (for morphological)
+
+  no1 <- x %>%
+    group_by(bb.sp, !!criterion, site, date) %>% # group by site-year and BB-FL pair
+    summarize(visits = sum(visits)) %>%
+    group_by(bb.sp, site, date) %>% # group by site-year and BB species
+    mutate(prop.visits = visits/sum(visits)) %>% # get proportional interaction frequency for each BB-FL pair
+    select(site, date, bb.sp, !!criterion, prop.visits) %>%
+    pivot_wider(names_from = !!criterion, values_from = prop.visits, values_fill = 0) %>% # spread into the kind of array vegan likes
+    group_by(site, date) %>%
+    nest() %>% # nest into list-column
+    mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
+           richness = map(data, function(x) length(x$bb.sp)),
+           dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
+           tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
+    #dplyr::select(site, date, tidy_dist) %>% # drop intermediate list-columns
+    unnest(cols = c(tidy_dist)) %>%
+    rename(bb.sp1 = item1, bb.sp2 = item2) %>%
+    mutate(niche.overlap = 1-distance,  # express niche overlap so that 1 = perfect overlap and 0 = no overlap
+           telu = if_else(bb.sp1 == "telu" | bb.sp2 == "telu", TRUE, FALSE),
+           pasc = if_else(bb.sp1 == "pasc" | bb.sp2 == "pasc", TRUE, FALSE),
+           wurf = if_else(bb.sp1 == "wurf" | bb.sp2 == "wurf", TRUE, FALSE),
+           prat = if_else(bb.sp1 == "prat" | bb.sp2 == "prat", TRUE, FALSE),
+           soro = if_else(bb.sp1 == "soro" | bb.sp2 == "soro", TRUE, FALSE),
+           humi = if_else(bb.sp1 == "humi" | bb.sp2 == "humi", TRUE, FALSE),
+           muci = if_else(bb.sp1 == "muci" | bb.sp2 == "muci", TRUE, FALSE),
+           pyre = if_else(bb.sp1 == "pyre" | bb.sp2 == "pyre", TRUE, FALSE),
+           mont = if_else(bb.sp1 == "mont" | bb.sp2 == "mont", TRUE, FALSE),
+           mend = if_else(bb.sp1 == "mend" | bb.sp2 == "mend", TRUE, FALSE),
+           psit = if_else(bb.sp1 == "psit" | bb.sp2 == "psit", TRUE, FALSE),
+           gers = if_else(bb.sp1 == "gers" | bb.sp2 == "gers", TRUE, FALSE),
+           jone = if_else(bb.sp1 == "jone" | bb.sp2 == "jone", TRUE, FALSE),
+           hypn = if_else(bb.sp1 == "hypn" | bb.sp2 == "hypn", TRUE, FALSE),
+           lapi = if_else(bb.sp1 == "lapi" | bb.sp2 == "lapi", TRUE, FALSE),
+           hort = if_else(bb.sp1 == "hort" | bb.sp2 == "hort", TRUE, FALSE))
+
+  no <- list(
+
+  no_telu <- no1 %>%
+    group_by(site, date, telu) %>% 
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%
+    mutate(bb.sp = rep("telu", n())) %>%
+    filter(telu == TRUE),
+
+  no_pasc <- no1 %>%
+    group_by(site, date, pasc) %>%
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%    
+    mutate(bb.sp = rep("pasc", n())) %>%
+    filter(pasc == TRUE),
+
+  no_wurf <- no1 %>%
+    group_by(site, date, wurf) %>%
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%    
+    mutate(bb.sp = rep("wurf", n())) %>%
+    filter(wurf == TRUE),
+
+  no_prat <- no1 %>%
+    group_by(site, date, prat) %>%
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%   
+    mutate(bb.sp = rep("prat", n())) %>%
+    filter(prat == TRUE),
+
+  no_soro <- no1 %>%
+    group_by(site, date, soro) %>% 
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%    
+    mutate(bb.sp = rep("soro", n())) %>%
+    filter(soro == TRUE),
+
+  no_humi <- no1 %>%
+    group_by(site, date, humi) %>%
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%    
+    mutate(bb.sp = rep("humi", n())) %>%
+    filter(humi == TRUE),
+
+  no_muci <- no1 %>%
+    group_by(site, date, muci) %>%
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%    
+    mutate(bb.sp = rep("muci", n())) %>%
+    filter(muci == TRUE),
+
+  no_pyre <- no1 %>%
+    group_by(site, date, pyre) %>%
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%   
+    mutate(bb.sp = rep("pyre", n())) %>%
+    filter(pyre == TRUE),
+
+  no_mont <- no1 %>%
+    group_by(site, date, mont) %>% 
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%    
+    mutate(bb.sp = rep("mont", n())) %>%
+    filter(mont == TRUE),
+
+  no_mend <- no1 %>%
+    group_by(site, date, mend) %>%
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%    
+    mutate(bb.sp = rep("mend", n())) %>%
+    filter(mend == TRUE),
+
+  no_psit <- no1 %>%
+    group_by(site, date, psit) %>%
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%    
+    mutate(bb.sp = rep("psit", n())) %>%
+    filter(psit == TRUE),
+
+  no_gers <- no1 %>%
+    group_by(site, date, gers) %>%
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%    
+    mutate(bb.sp = rep("gers", n())) %>%
+    filter(gers == TRUE),
+
+  no_jone <- no1 %>%
+    group_by(site, date, jone) %>% 
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%    
+    mutate(bb.sp = rep("jone", n())) %>%
+    filter(jone == TRUE),
+
+  no_hypn <- no1 %>%
+    group_by(site, date, hypn) %>%
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%    
+    mutate(bb.sp = rep("hypn", n())) %>%
+    filter(hypn == TRUE),
+
+  no_lapi <- no1 %>%
+    group_by(site, date, lapi) %>%
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%    
+    mutate(bb.sp = rep("lapi", n())) %>%
+    filter(lapi == TRUE),
+
+  no_hort <- no1 %>%
+    group_by(site, date, hort) %>%
+    summarize(mean.niche.overlap = mean(niche.overlap),
+              max.niche.overlap = max(niche.overlap)) %>%    
+    mutate(bb.sp = rep("hort", n())) %>%
+    filter(hort == TRUE)
+  ) %>%
+
+    bind_rows() %>%
+    select(site, date, bb.sp, mean.niche.overlap, max.niche.overlap)
+}
+```
 
@@
 ```
-
-## 2.3 Calculate bias per plant speces/ktype/family
-```{r}
-representation <- biaseval_group(visit_cover, plant.sp) %>%
-  rename(plant.sp = criterion) %>%
-  left_join(site_data) %>%
-  left_join(fl_traits) %>%
-  mutate(k.type.ss = factor(k.type.ss)) %>%
-  select(site, date, elev.mean, yday, bb.sp, plant.sp, k.type.ss, 
-         prop.visits, prop.flower.cover, bias, visits, 
-         flower.cover, plant.genus, plant.family) %>%
-  left_join(bb_traits)
-
-# 
-# representation_ktype <- biaseval_group(visit_cover, k.type.ss) %>%
-#   left_join(site_data) %>%
-#   rename(group = criterion) %>%
-#   mutate(criterion = rep("ktype", n()))
-# 
-# bias_group_family <- biaseval_group(visit_cover, plant.family) %>%
-#   left_join(site_data) %>%
-#   rename(group = criterion) %>%
-#   mutate(criterion = rep("plant.family", n()))
-# 
-# bias_group <- bind_rows(bias_group_plantsp, bias_group_ktype, bias_group_family)
-# 
-# ggplot(bias_group, aes(prop.flower.cover, prop.visits)) +
-#   geom_point(alpha = 0.1) +
-#   #geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
-#   geom_smooth(method = "lm") +
-#   geom_abline(slope = 1, color = "red", linetype = "dashed") +
-#   facet_wrap(~criterion)
-# 
-# ggplot(bias_group, aes(log(prop.flower.cover), prop.visits)) +
-#   geom_point(alpha = 0.1) +
-#   geom_smooth() +
-#   facet_wrap(~criterion)
-# 
-# ggplot(filter(bias_group, criterion == "ktype"), aes(prop.flower.cover, prop.visits)) +
-#   geom_point(alpha = 0.1) +
-#   geom_smooth() +
-#   geom_abline(slope = 1, color = "red", linetype = "dashed") +
-#   facet_wrap(~group)
-```
-
-## 2.4 Visualize
-```{r}
-# Check the distribution of bias
-ggplot(bias_group_ktype, aes(bias)) +
-  geom_histogram()
-
-ggplot(bias_group_ktype, aes(bias)) +
-  geom_histogram() +
-  facet_wrap(~criterion)
-
-ggplot(bias_group_family, aes(bias)) +
-  geom_histogram()
-
-ggplot(bias_group_family, aes(bias)) +
-  geom_histogram() +
-  facet_wrap(~criterion)
-
-# Bias by Kugler morphotype
-ggplot(bias_group_ktype, aes(group, bias)) +
-  geom_boxplot()
-
-ggplot(bias_group_ktype, aes(group, bias)) +
-  geom_boxplot() +
-  facet_wrap(~bb.sp)
-
-ggplot(bias_group_ktype, aes(bb.sp, bias)) +
-  geom_boxplot() +
-  facet_wrap(~group)
-
-ggplot(bias_group_ktype, aes(elev.mean, bias)) +
-  #geom_point() +
-  geom_smooth(method = gam, formula = y ~ s(x, k = 5)) +
-  facet_wrap(~group)
-
-# Bias by plant family
-ggplot(bias_group_family, aes(group, bias)) +
-  geom_boxplot()
-
-ggplot(bias_group_family, aes(group, bias)) +
-  geom_boxplot() +
-  facet_wrap(~bb.sp)
-
-ggplot(bias_group_family, aes(bb.sp, bias)) +
-  geom_boxplot() +
-  facet_wrap(~group)
-
-ggplot(bias_group_family, aes(elev.mean, bias)) +
-  #geom_point() +
-  geom_smooth(method = gam, formula = y ~ s(x, k = 5)) +
-  facet_wrap(~group)
-```
-
-## 2.5 Multi-level models
-```{r}
-# The tons of zeros are due to the lots of plant species that were present at very low abundance and received few or zero visits. 
-ggplot(representation, aes(k.type.ss, bias)) +
-  geom_violin()
-
-ggplot(representation, aes(k.type.ss, bias)) +
-  geom_boxplot()
-
-ggplot(representation, aes(bias)) +
-  geom_histogram() +
-  facet_wrap(~k.type.ss)
-
-ggplot(filter(representation, abs(bias) > 0.1), aes(bias)) +
-  geom_histogram() +
-  facet_wrap(~k.type.ss)
-
-ggplot(filter(representation, prop.flower.cover > 0.05), aes(k.type.ss, bias)) +
-  geom_boxplot()
-
-ggplot(representation, aes(prop.flower.cover, prop.visits)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_wrap(~k.type.ss)
-
-brm_representation0 <- brm(bias ~ 1 + (1 | k.type.ss) + (1 | site),
-                 control = list(max_treedepth = 10),
-                 cores = 4,
-                 chains = 4,
-                 iter = 2000,
-                 file = "../output/brm_representation0",
-                 data = representation)
-
-summary(brm_representation0)
-plot(brm_representation0)
-parnames(brm_representation0)
-mcmc_plot(brm_representation0, pars = c("^r_"))
-#mcmc_plot(representation.ktype0, pars = c("elev"))
-mcmc_plot(brm_representation0, pars = c("^b_"))
-mcmc_plot(brm_representation0)
-#conditional_effects(representation.ktype0)
-pp_check(brm_representation0)
-bayes_R2(brm_representation0)
-
-brm_representation1 <- brm(bias ~ pbl.w + (1 + pbl.w | k.type.ss) + (1 | bb.sp),
-                 control = list(max_treedepth = 10),
-                 cores = 4,
-                 chains = 4,
-                 iter = 2000,
-                 file = "../output/brm_representation1",
-                 data = representation)
-
-summary(brm_representation1)
-plot(brm_representation1)
-parnames(brm_representation1)
-mcmc_plot(brm_representation1, pars = c("^r_"))
-#mcmc_plot(representation.ktype1, pars = c("elev"))
-mcmc_plot(brm_representation1, pars = c("^b_"))
-mcmc_plot(brm_representation1)
-#conditional_effects(representation.ktype1)
-pp_check(brm_representation1)
-bayes_R2(brm_representation1)
-
-hypothesis(representation.ktype1, "r_group[5,Intercept] = r_group[6,Intercept]", class = "")
-hypothesis(representation.ktype1, "r_group[1,Intercept] = r_group[0,Intercept]", class = "")
-
-# One interesting way to think about bias is to focus on the slope of prop.visits ~ prop.flower.cover. 
-# We could analyze this with a varying slope model
-
-# Binomial ggplot smoother
-binomial_smooth <- function(...) {
-  geom_smooth(method = "glm", method.args = list(family = "quasibinomial"), se = FALSE, ...)
-}
-
-ggplot(representation, aes(sqrt(prop.flower.cover), sqrt(prop.visits))) +
-  geom_point() +
-  #geom_smooth(method = "lm") +
-  #binomial_smooth(formula = y ~ splines::ns(x, 3)) +
-  geom_smooth() +
-  #geom_smooth(method = "glm", method.args = list(family = "binomial")) +
-  facet_grid(k.type.ss~bb.sp)
-
-brm_representation3 <- brm(prop.visits ~ prop.flower.cover + (1 + prop.flower.cover | k.type.ss),
-                 control = list(max_treedepth = 10),
-                 #cores = 4,
-                 cores = 2,
-                 chains = 2,
-                 #iter = 2000,
-                 iter = 1000,
-                 file = "./output/brm_representation3",
-                 data = representation)
-
-lmer_representation3 <- lmer(prop.visits ~ log(prop.flower.cover) + (1 + log(prop.flower.cover) | k.type.ss),
-                             data = representation)
-
-summary(lmer_representation3)
-plot(lmer_representation3)
-
-# Can we explain anything with tongue length
-
-ggplot(filter(representation, bb.sp != "gers" & abs(bias) >= 0.1), aes(pbl.w, bias)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  facet_wrap(~k.type.ss)
-
-lm_pbl <- lmer(bias ~ pbl.w + (1 + pbl.w | k.type.ss) + (1 | bb.sp),
-               data = representation)
-
-summary(lm_pbl)
-```
-
-## 2.6 Ordination of bias
-```{r}
-rep_matrix.ktype <- representation %>%
-  group_by(bb.sp, k.type.ss) %>%
-  summarize(bias = mean(bias)) %>%
-  pivot_wider(values_from = bias, names_from = k.type.ss, values_fill = 0) %>%
-  column_to_rownames("bb.sp")
-
-traits <- bb_traits %>%
-  arrange(bb.sp)
-
-pca.ktype <- rda(rep_matrix.ktype)
-#ef.ktype <- envfit(pca.ktype ~ pbl.w.class, data = traits)
-#os.ktype <- ordisurf(pca.ktype ~ pbl.w.class, data = traits)
-ef.ktype <- envfit(pca.ktype ~ pbl.w, data = traits)
-os.ktype <- ordisurf(pca.ktype ~ pbl.w, data = traits)
-plot(pca.ktype, type = "t")
-plot(ef.ktype, add = TRUE)
-plot(os.ktype)
-```
-
-# 3. Effects of bias
-
-## 3.1 Relationship between bias(d') and niche overlap
-### 3.1.1 A function for calculating niche overlap
-```{r}
-# Niche overlap by sp.pair * site * year
-niche_overlapper <- function(x, criterion) {
-
-  criterion <- enquo(criterion) # criterion is set to either plant.sp (for taxonomic) or k.type.s (for morphological)
-
-  no1 <- x %>%
-    group_by(bb.sp, !!criterion, site, date) %>% # group by site-year and BB-FL pair
-    summarize(visits = sum(visits)) %>%
-    group_by(bb.sp, site, date) %>% # group by site-year and BB species
-    mutate(prop.visits = visits/sum(visits)) %>% # get proportional interaction frequency for each BB-FL pair
-    select(site, date, bb.sp, !!criterion, prop.visits) %>%
-    pivot_wider(names_from = !!criterion, values_from = prop.visits, values_fill = 0) %>% # spread into the kind of array vegan likes
-    group_by(site, date) %>%
-    nest() %>% # nest into list-column
-    mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
-           dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
-           tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
-    dplyr::select(site, date, tidy_dist) %>% # drop intermediate list-columns
-    unnest(cols = c(tidy_dist)) %>%
-    rename(bb.sp1 = item1, bb.sp2 = item2) %>%
-    mutate(niche.overlap = 1-distance,  # express niche overlap so that 1 = perfect overlap and 0 = no overlap
-           bss = if_else(bb.sp1 == "bss" | bb.sp2 == "bss", TRUE, FALSE),
-           pasc = if_else(bb.sp1 == "pasc" | bb.sp2 == "pasc", TRUE, FALSE),
-           wurf = if_else(bb.sp1 == "wurf" | bb.sp2 == "wurf", TRUE, FALSE),
-           prat = if_else(bb.sp1 == "prat" | bb.sp2 == "prat", TRUE, FALSE),
-           soro = if_else(bb.sp1 == "soro" | bb.sp2 == "soro", TRUE, FALSE),
-           humi = if_else(bb.sp1 == "humi" | bb.sp2 == "humi", TRUE, FALSE),
-           muci = if_else(bb.sp1 == "muci" | bb.sp2 == "muci", TRUE, FALSE),
-           pyre = if_else(bb.sp1 == "pyre" | bb.sp2 == "pyre", TRUE, FALSE),
-           mont = if_else(bb.sp1 == "mont" | bb.sp2 == "mont", TRUE, FALSE),
-           mend = if_else(bb.sp1 == "mend" | bb.sp2 == "mend", TRUE, FALSE),
-           psit = if_else(bb.sp1 == "psit" | bb.sp2 == "psit", TRUE, FALSE),
-           gers = if_else(bb.sp1 == "gers" | bb.sp2 == "gers", TRUE, FALSE),
-           jone = if_else(bb.sp1 == "jone" | bb.sp2 == "jone", TRUE, FALSE),
-           hypn = if_else(bb.sp1 == "hypn" | bb.sp2 == "hypn", TRUE, FALSE),
-           lapi = if_else(bb.sp1 == "lapi" | bb.sp2 == "lapi", TRUE, FALSE),
-           hort = if_else(bb.sp1 == "hort" | bb.sp2 == "hort", TRUE, FALSE))
-
-  no <- list(
-
-  no_bss <- no1 %>%
-    group_by(site, date, bss) %>% 
-    summarize(mean.niche.overlap = mean(niche.overlap),
-              max.niche.overlap = max(niche.overlap)) %>%
-    mutate(bb.sp = rep("bss", n())) %>%
-    filter(bss == TRUE),
-
-  no_pasc <- no1 %>%
-    group_by(site, date, pasc) %>%
-    summarize(mean.niche.overlap = mean(niche.overlap),
-              max.niche.overlap = max(niche.overlap)) %>%    
-    mutate(bb.sp = rep("pasc", n())) %>%
-    filter(pasc == TRUE),
-
-  no_wurf <- no1 %>%
-    group_by(site, date, wurf) %>%
-    summarize(mean.niche.overlap = mean(niche.overlap),
-              max.niche.overlap = max(niche.overlap)) %>%    
-    mutate(bb.sp = rep("wurf", n())) %>%
-    filter(wurf == TRUE),
-
-  no_prat <- no1 %>%
-    group_by(site, date, prat) %>%
-    summarize(mean.niche.overlap = mean(niche.overlap),
-              max.niche.overlap = max(niche.overlap)) %>%   
-    mutate(bb.sp = rep("prat", n())) %>%
-    filter(prat == TRUE),
-
-  no_soro <- no1 %>%
-    group_by(site, date, soro) %>% 
-    summarize(mean.niche.overlap = mean(niche.overlap),
-              max.niche.overlap = max(niche.overlap)) %>%    
-    mutate(bb.sp = rep("soro", n())) %>%
-    filter(soro == TRUE),
-
-  no_humi <- no1 %>%
-    group_by(site, date, humi) %>%
-    summarize(mean.niche.overlap = mean(niche.overlap),
-              max.niche.overlap = max(niche.overlap)) %>%    
-    mutate(bb.sp = rep("humi", n())) %>%
-    filter(humi == TRUE),
-
-  no_muci <- no1 %>%
-    group_by(site, date, muci) %>%
-    summarize(mean.niche.overlap = mean(niche.overlap),
-              max.niche.overlap = max(niche.overlap)) %>%    
-    mutate(bb.sp = rep("muci", n())) %>%
-    filter(muci == TRUE),
-
-  no_pyre <- no1 %>%
-    group_by(site, date, pyre) %>%
-    summarize(mean.niche.overlap = mean(niche.overlap),
-              max.niche.overlap = max(niche.overlap)) %>%   
-    mutate(bb.sp = rep("pyre", n())) %>%
-    filter(pyre == TRUE),
-
-  no_mont <- no1 %>%
-    group_by(site, date, mont) %>% 
-    summarize(mean.niche.overlap = mean(niche.overlap),
-              max.niche.overlap = max(niche.overlap)) %>%    
-    mutate(bb.sp = rep("mont", n())) %>%
-    filter(mont == TRUE),
-
-  no_mend <- no1 %>%
-    group_by(site, date, mend) %>%
-    summarize(mean.niche.overlap = mean(niche.overlap),
-              max.niche.overlap = max(niche.overlap)) %>%    
-    mutate(bb.sp = rep("mend", n())) %>%
-    filter(mend == TRUE),
-
-  no_psit <- no1 %>%
-    group_by(site, date, psit) %>%
-    summarize(mean.niche.overlap = mean(niche.overlap),
-              max.niche.overlap = max(niche.overlap)) %>%    
-    mutate(bb.sp = rep("psit", n())) %>%
-    filter(psit == TRUE),
-
-  no_gers <- no1 %>%
-    group_by(site, date, gers) %>%
-    summarize(mean.niche.overlap = mean(niche.overlap),
-              max.niche.overlap = max(niche.overlap)) %>%    
-    mutate(bb.sp = rep("gers", n())) %>%
-    filter(gers == TRUE),
-
-  no_jone <- no1 %>%
-    group_by(site, date, jone) %>% 
-    summarize(mean.niche.overlap = mean(niche.overlap),
-              max.niche.overlap = max(niche.overlap)) %>%    
-    mutate(bb.sp = rep("jone", n())) %>%
-    filter(jone == TRUE),
-
-  no_hypn <- no1 %>%
-    group_by(site, date, hypn) %>%
-    summarize(mean.niche.overlap = mean(niche.overlap),
-              max.niche.overlap = max(niche.overlap)) %>%    
-    mutate(bb.sp = rep("hypn", n())) %>%
-    filter(hypn == TRUE),
-
-  no_lapi <- no1 %>%
-    group_by(site, date, lapi) %>%
-    summarize(mean.niche.overlap = mean(niche.overlap),
-              max.niche.overlap = max(niche.overlap)) %>%    
-    mutate(bb.sp = rep("lapi", n())) %>%
-    filter(lapi == TRUE),
-
-  no_hort <- no1 %>%
-    group_by(site, date, hort) %>%
-    summarize(mean.niche.overlap = mean(niche.overlap),
-              max.niche.overlap = max(niche.overlap)) %>%    
-    mutate(bb.sp = rep("hort", n())) %>%
-    filter(hort == TRUE)
-  ) %>%
-
-    bind_rows() %>%
-    select(site, date, bb.sp, mean.niche.overlap, max.niche.overlap)
-  }
-```
-
-### 3.1.2 Calculate niche overlap: NOT READY FOR INTERPRETATION YET
-```{r}
-niche_overlap_plant.sp <- niche_overlapper(visit_cover, plant.sp)
-
-niche_overlap_ktype <- niche_overlapper(visit_cover, k.type.ss)
-
-niche_overlap_plant.family <- niche_overlapper(visit_cover, plant.family)
-
-niche_overlap <- bind_rows(no_plant.sp, 
-                           no_ktype, 
-                           no_plant.family)
-
-# When bias is high, it could mean that there are few good options on which all (or many) BB species converge.
-But apparently it doesn't.
-When bias is high, BB spp are biased toward DIFFERENT flora, and thus their niches diverge.
-ggplot(bias, aes(dprime, mean.niche.overlap, color = criterion)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth(method = "lm") +
-  #geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
-  facet_wrap(~bb.sp) +
-  ylim(c(0,1))
-
-ggplot(bias, aes(dprime, max.niche.overlap, color = criterion)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth(method = "lm") +
-  #geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
-  facet_wrap(~bb.sp) +
-  ylim(c(0,1))
-
-ggplot(bias, aes(log(total.flower.cover), dprime, color = criterion)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth(method = "lm") +
-  #geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
-  facet_wrap(~bb.sp) +
-  ylim(c(0,1))
-
-ggplot(bias, aes(log(total.flower.cover), mean.niche.overlap, color = criterion)) +
-  geom_point(alpha = 0.1) +
-  geom_smooth(method = "lm") +
-  #geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
-  facet_wrap(~bb.sp) +
-  ylim(c(0,1))
-
-brm.bias_niche <- brm(mean.niche.overlap ~ 1 + dprime + (1 | bb.sp),
-                     cores = 3,
-                     data = filter(bias, criterion == "ktype"))
-
-parnames(brm.bias_niche)
-conditional_effects(brm.bias_nich)
-hypothesis(brm.bias_niche, "dprime = 0")
-pp_check(brm.bias_niche)
-```
-
-## 3.2 Ordination of dietary niche
-```{r}
-rep_matrix.ktype <- representation %>%
-  group_by(bb.sp, k.type.ss) %>%
-  summarize(bias = mean(bias)) %>%
-  pivot_wider(values_from = bias, names_from = k.type.ss, values_fill = 0) %>%
-  column_to_rownames("bb.sp")
-
-traits <- bb_traits %>%
-  arrange(bb.sp)
-
-pca.ktype <- rda(rep_matrix.ktype)
-#ef.ktype <- envfit(pca.ktype ~ pbl.w.class, data = traits)
-#os.ktype <- ordisurf(pca.ktype ~ pbl.w.class, data = traits)
-ef.ktype <- envfit(pca.ktype ~ pbl.w, data = traits)
-os.ktype <- ordisurf(pca.ktype ~ pbl.w, data = traits)
-plot(pca.ktype, type = "t")
-plot(ef.ktype, add = TRUE)
-plot(os.ktype)
-```
+```

### 2021-01-22 writing/manuscript_v4.Rmd
backup

@@
 output:
-  pdf_document: default
-  html_document: default
+  pdf_document:
+    fig_caption: yes
 ---
@@
 
+### Study sites
+
 The study was conducted in Berchtesgaden National Park (47.55°N, 12.92°E), located in the Northern Limestone Alps of southeast Germany.
@@
 The sites sampled in this study are the same as those used in a series of previous studies [@Hoiss2012-sn; @Hoiss2012-kr; Hoiss2015-gv], but the data reported here are otherwise independent.
+
+### Bumble bee visitation
 
@@
 
+visitation sample = site-date of bumble bee visitation observations
+
+While floral visitation by male bumble bees was recorded during visitation sampling, we chose to analyze only visitation by queens and workers.
+
+### Flora surveying
+
+floral survey sample = site-date of floral surveying
+
 ## 2.2. Data analysis
 
-Prior to analysis, we aligned each bumble bee visitation sample with corresponding floral survey data.
-For most samples, floral surveying was conducted jointly with bumble bee visitation observation.
-In cases where bumble bee visitation was sampled without a joint floral survey, the bumble bee visitation data was paired with the nearest available floral survey data within a maximum distance of 7 days.
-Bumble bee visitation samples lacking corresponding floral survey data within 7 days were omitted from further analysis.
-In a few cases where more than one bumble bee visitation date was paired with a given floral survey date, the bumble bee visitation data were pooled.
+Prior to analysis, we aligned each visitation sample with a corresponding floral survey sample.
+In most cases, floral surveying was conducted jointly with bumble bee observation for each site and date.
+Occasionally, though, floral surveying could not be completed on the same day as bumble bee observation, resulting in a small temporal offset between visitation samples and floral survey samples.
+In these cases, visitation samples were paired with the nearest available floral survey data within a maximum distance of 7 days.
+Visitation samples lacking corresponding floral survey data within 7 days were omitted from further analysis.
+In a few cases, two consecutive visitation samples at a given site shared the same nearest floral survey sample; in these cases, the visitation samples were first pooled, then aligned to their shared nearest floral survey sample.
+In rare cases, a given floral species was recorded during visitation sampling but missed during floral surveying.
+These species were retroactively assigned the minimum floral area value of 0.01 m^2.
 
@@
 
-To quantify cumulative bias per bumble bee species, we used the discrimination (d’) metric of @Bluthgen2006-if, which represents the net deviation of a species’ visitation from the predictions of a null model in which the probability of visitation is purely a function of floral abundance.
+To quantify cumulative bias per bumble bee species, we used the discrimination (*d’*) metric of @Bluthgen2006-if, which represents the net deviation of a species’ visitation patterns from the predictions of a null model in which the probability of visitation is purely a function of floral abundance.
+We calculated *d'* values separately for each visitation sample, with floral abundances supplied by its corresponding floral survey sample.
 
-#### 2.2.2.1.1 Model 1: dprime ~
+Preliminary visualization showed that *d'* values tended to become binary (0 or 1) when calculated for bumble bee species for which fewer than 3 individuals were observed in a given visitation sample.
+This generated a tri-modal distribution of *d'* values, with peaks at 0 and 1 flanking an otherwise approximately normal distribution.
+To ease model estimation and avoid the influence of the highly uncertain *d'* values calculated based on one or two observations, we chose to filter our data to include for each visitation sample only species that were observed at least 3 times.
 
-### 2.2.2 Modeling visitation probability as a function of floral morphotype and bumble bee tongue length
+The variation of *d'* was analyzed using Bayesian multi-level modeling.
+We modeled the response variable *d'* as a function of the interaction between elevation and day-of-year, with bumble bee species as a varying intercept term.
+The continuous predictors elevation and day-of-year were scaled and centered to optimize software performance.
+<!-- Explain choice of priors (or lack thereof).
+Maybe also explain the choice of modeling bb.sp as a varying intercept term rather than a fixed effect --> We then specified an alternative model in which the elevation and day of year terms were replaced with a varying intercept term for site.
+This simpler model was compared to the interaction model via leave-one-out (LOO) cross-validation using the *loo()* function from *brms*, which is a wrapper for a function of the same name from the R package *loo* [@Yao2017aa].
 
-#### 2.2.2.1 Model 2: visits.bin ~ prop.flower.cover + bb.sp.abund + (1 | k.type.ss) + (1 | site)
+### 2.2.2 Modeling visitation rate as a function of floral morphotype and bumble bee tongue length
 
-#### 2.2.2.2 Model 3: visits.bin ~ prop.flower.cover + bb.sp.abund + (1 | k.type.ss:pbl.w.class2) + (1 | site)
+The discrimination metric *d'* expresses the overall disproportionality between relative abundance and visitation, but it does not reveal the causes of bias, i.e.
+the functional relationships between bumble bees and flowers that lead to differential visitation.
+To explore these functional underpinnings of bumble foraging bias, we analyzed visitation rate as a function of the interaction between floral morphology and bumble bee tongue length, conditioned on the abundance of each flower and bumble bee species.
 
-#### 2.2.2.3 Model 4: Ordination with post-hoc fitting of tongue-length
+Bumble bee tongue-length data were obtained from the literature values compiled and summarized by @Arbetman2017-xj.
+Values for *B.
+gerstaeckeri* and *B.
+mendax*, which were not included in @Arbetman2017-xj, were obtained from @Obeso1992-yp and @Durieux2000-uk, respectively.
+We then binned bumble bee species into three tongue-length classes: short (< 7.5 mm), medium (7.5---10.0 mm), and long (>= 10.0 mm).
+Short-tongued species included *B.
+jonellus*, *B.
+lapidarius*, *B.
+pratorum*, *B.
+psithyrus-group*, *B.
+pyrenaeus*, and *B.
+terrestris/lucorum*.
+Medium-tongued species included *B.
+humilis*, *B.
+hypnorum*, *B.
+monticola*, *B.
+mucidus*, *B.
+pascuorum*, and *B.
+wurflenii*.
+Long-tongued species included *B.
+gerstaeckeri*, *B.
+hortorum*, and *B.
+mendax*.
+
+Floral morphology was summarized using floral morphotypes of @Kugler1970-gb, accessed via the BIOFLOR database [@Klotz2002-uc].
+Morphotypes were simplified to the basic categories numbered 0-10: 0 = not applicable (includes wind-pollinated flora such as grasses), 1 = disc- and bowl-shaped flowers, 2 = funnel flowers, 3 = bell-shaped flowers, 4 = stalk-disc flowers, 5 = lip flowers, 6 = flag flowers, 7 = flower heads, 8 = spike flowers, 9 = brush flowers, 10 = trap flowers **(Figure X)**.
+<!-- I think I would like to include a figure with photographs of flowers exemplifying each morphotype --> In a few cases, a plant species in our data set was not found in the BIOFLOR database.
+In most cases, morphotypes are constant at the genus level, and missing morphotypes were added based on the morphotypes found for congeners.
+The genus *Gentiana* consisted of both group 2 and group 4 representatives, so the missing species *Gentiana aspera* and *Gentiana ciliata* were classified based on visual comparison with their classified congeners.
+In the case of the genus *Phyteuma*, we followed @Neumayer1999-il and classified it as a group 7 flower head.
+
+We modeled visitation using multilevel Bayesian Poisson regression.
+First, we specified a model focused on comparing visitation rates across morphotypes.
+Visitation was modeled using varying intercept terms for morphotype, site, and bumble bee species, and fixed terms for the abundances of floral and bumble bee species.
+Then, we specified a more complex model in which the varying intercept term for morphotype was replaced with one for the interaction of morphotype and bumble bee tongue length class.
+In both models, the continuous predictors were scaled and centered to optimize software performance.
+<!-- Should I perhaps have included a fixed interaction of bb abundance and floral abundance?
+-->
 
@@
 
+### 2.2.4 Software and code
+
+All analyses were conducted in R v.
+3.6 .
+Discrimination (*d'*) was calculated using the function *dfun()* [@Dormann2011-aa] from the R package *bipartite* [@Dormann2008-ab].Bayesian multilevel models were implemented in *Stan* [@Stan2020aa] via the R package *brms* [@Buerkner2018aa], and visualizations were constructed with the R packages *tidybayes* [@Kay2020aa] and *ggdist* [Kay2020ab].
+Ordination analyses were performed with the R package *vegan* [@Oksanen2019aa] and visualized using the R packages *ggplot2* [@Wickham2016aa] and *ggvegan* [@Simpson2019aa].
+Data handling was performed the *tidyverse* suite of R packages [Wickham2019aa].
+Code and model diagnostics are available in Appendix X.
+
 # 3. Results
-## 3.1 Model 1
-```{r}
-knitr::include_graphics("../output/brm.plantsp0_plot1.png")
+## 3.1 Cumulative bias (d')
+
+LOO model comparison favored the simpler model that used a varying intercept effect for site in lieu of the elevation:day-of-year interaction term (elpd_diff = -104).
+The estimated global mean of *d'* was 0.38 (0.33---0.43 95% CI).
+Variation of bumble bee species from the global mean of *d'* was modest, with an estimated standard deviation of 0.05 (0.02---0.08 95% CI) and substantial overlap between species in their posterior distributions **(Figure XA)**.
+Spatial variation in *d'* was somewhat more pronounced, with an estimated standard deviation of 0.08 (0.06---0.11 95% CI) **(Figure XB)**.
+
+See supplemental material for the details of model specification and evaluation.
+
+```{r fig1, echo = FALSE,  fig.cap = paste("Posterior conditional mean discrimination (d') by bumble bee species and site. Half-eye intervals depict 66\\% and 95\\% CIs.")}
+knitr::include_graphics("../output/brm_dprime.sp01_plot1.png")
 ```
 
-## 3.2 Model 2
+## 3.2 Explaining bias via analysis of visitation rate by flower morphotype and bee tongue length
 
-## 3.3. Model 3
+Visitation rates varied markedly across morphotypes **(Figure X)**.
+Flag, lip, and bell flowers had the highest visitation rates, with broadly overlapping posterior distributions.
+Funnel flowers and bell flowers had the next highest visitation rates, followed by stalk-disc flowers and then disc flowers.
+Unclassified flowers, which are mainly wind-pollinated species, had the lowest visitation rates.
 
-## 3.4 Model 4
+```{r fig2, echo = FALSE,  fig.cap = paste("Posterior conditional mean visits (log-scale) by Kugler morphotype. Half-eye plots depict 66\\% and 95\\% CIs.")}
+knitr::include_graphics("../output/brm_pvis00_plot1.png")
+```
 
-## 3.5 Model 5
+Visitation also varied strongly within morphotypes across bumble bee tongue length classes **(Figure X)**.
+Disc flowers (group 1), bell flowers (group 3), and unclassified flowers (group 0) were visited chiefly by short-tongued bumble bee species; visitation of these morphotypes by medium- and long-tongued species was much lower, and there was little difference in visitation rates between medium- and long-tongued species.
+Funnel flowers (group 2) were visited mostly by long-tongued species, with medium- and short-tongued species following in rank order.
+The opposite pattern appears in head flowers (group 7), which were visited at the highest rate by short-tongued species, with medium- and long-tongued species following in reverse rank-order.
+Lip flowers (group 5) were visited at high rates by all tongue-length classes, with little differentiation between them.
+Flag flowers (group 6) were received similarly high rates of visitation from long- and medium-tongued species, with lower visitation from short-tongued species.
 
-## 3.6 Model 6
+```{r fig3, echo = FALSE,  fig.cap = paste("Posterior conditional mean visits (log-scale) by Kugler morphotype and bumble bee tongue length class. Half-eye plots depict 66\\% and 95\\% CIs.")}
+knitr::include_graphics("../output/brm_pvis11_plot1.png")
+```
+
+## 3.3. Consequences of bias for niche relationships
 

### 2021-01-22 code/data_proc.R

@@ -28,7 +28,7 @@ fl_survey <- read_delim("./raw_data/data_plantcover_forDoug_2020.01.17.csv", del
 write_csv(fl_survey, "./processed_data/floral_survey.csv")
 
 ### Visitation data
-network <- read_delim("./raw_data/data_bumblebee_forDoug_2020.04.17.csv", delim = ";", 
+network <- read_delim("./data/raw_data/data_bumblebee_forDoug_2020.04.17.csv", delim = ";", 
                            locale = locale(decimal_mark = ",", grouping_mark = ".")) %>%
   mutate(plant_sp_latin = str_replace_all(plant_sp_latin, 
                                          c("Arabis caerula" = "Arabis caerulea", # correct nomenclature typos and synonyms
@@ -47,13 +47,14 @@ network <- read_delim("./raw_data/data_bumblebee_forDoug_2020.04.17.csv", delim 
          day = day(temp.date),
          month = month(temp.date),
          date = make_date(year, month, day)) %>%
-  mutate(bb.sp = str_replace_all(bb.sp, c("luco" = "bss","terr" = "bss", "telu" = "bss", "cryp" = "bss", 
-                                          "sylv" = "psit", "quad" = "psit", "camp" = "psit", "flav" = "psit", "psyt" = "psit")),
+  filter(bb.sp != "cryp") %>% # this species was observed only oncne in three years
+  mutate(bb.sp = str_replace_all(bb.sp, c("luco" = "telu", "terr" = "telu", "sylv" = "psit", 
+                                          "quad" = "psit", "camp" = "psit", "flav" = "psit", 
+                                          "psyt" = "psit")),
          bb.sp.lat = str_replace_all(bb.sp.lat, c("Bombus terrestris/lucorum" = "Bombus ss",
-                                        "Bombus terrestris" = "Bombus ss",
-                                        "Bombus lucorum" = "Bombus ss",
-                                        "Bombus terrestris/lucorum/lucorum" = "Bombus ss",
-                                        "Bombus cryptarum" = "Bombus ss",
+                                        "Bombus terrestris" = "Bombus telu",
+                                        "Bombus lucorum" = "Bombus telu",
+                                        "Bombus terrestris/lucorum/lucorum" = "Bombus telu",
                                         "B. barbutellus, B. bohemicus, B. sylvestris" = "Psithyrus",
[…]

### 2021-01-26 code/visitation_bias.Rmd
developed ordination analysis

@@
 }
+
+# Aggregating over site
+niche_overlapper2 <- function(x, criterion) {
+
+  criterion <- enquo(criterion) # criterion is set to either plant.sp (for taxonomic) or k.type.s (for morphological)
+
+  no1 <- x %>%
+    group_by(bb.sp, !!criterion, site) %>% # group by site-year and BB-FL pair
+    summarize(visits = sum(visits)) %>%
+    group_by(bb.sp, site) %>% # group by site-year and BB species
+    mutate(prop.visits = visits/sum(visits)) %>% # get proportional interaction frequency for each BB-FL pair
+    select(site, bb.sp, !!criterion, prop.visits) %>%
+    pivot_wider(names_from = !!criterion, values_from = prop.visits, values_fill = 0) %>% # spread into the kind of array vegan likes
+    group_by(site) %>%
+    nest() %>% # nest into list-column
+    mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
+           richness = map(data, function(x) length(x$bb.sp)),
+           dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
+           tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
+    #dplyr::select(site, date, tidy_dist) %>% # drop intermediate list-columns
+    unnest(cols = c(tidy_dist)) %>%
+    rename(bb.sp1 = item1, bb.sp2 = item2) %>%
+    mutate(niche.overlap = 1-distance,  # express niche overlap so that 1 = perfect overlap and 0 = no overlap
+           telu = if_else(bb.sp1 == "telu" | bb.sp2 == "telu", TRUE, FALSE),
+           pasc = if_else(bb.sp1 == "pasc" | bb.sp2 == "pasc", TRUE, FALSE),
+           wurf = if_else(bb.sp1 == "wurf" | bb.sp2 == "wurf", TRUE, FALSE),
+           prat = if_else(bb.sp1 == "prat" | bb.sp2 == "prat", TRUE, FALSE),
+           soro = if_else(bb.sp1 == "soro" | bb.sp2 == "soro", TRUE, FALSE),
+           humi = if_else(bb.sp1 == "humi" | bb.sp2 == "humi", TRUE, FALSE),
+           muci = if_else(bb.sp1 == "muci" | bb.sp2 == "muci", TRUE, FALSE),
+           pyre = if_else(bb.sp1 == "pyre" | bb.sp2 == "pyre", TRUE, FALSE),
+           mont = if_else(bb.sp1 == "mont" | bb.sp2 == "mont", TRUE, FALSE),
+           mend = if_else(bb.sp1 == "mend" | bb.sp2 == "mend", TRUE, FALSE),
+           psit = if_else(bb.sp1 == "psit" | bb.sp2 == "psit", TRUE, FALSE),
+           gers = if_else(bb.sp1 == "gers" | bb.sp2 == "gers", TRUE, FALSE),
+           jone = if_else(bb.sp1 == "jone" | bb.sp2 == "jone", TRUE, FALSE),
+           hypn = if_else(bb.sp1 == "hypn" | bb.sp2 == "hypn", TRUE, FALSE),
+           lapi = if_else(bb.sp1 == "lapi" | bb.sp2 == "lapi", TRUE, FALSE),
+           hort = if_else(bb.sp1 == "hort" | bb.sp2 == "hort", TRUE, FALSE))
+}
+
+niche_overlapper3 <- function(x, criterion) {
+
+  criterion <- enquo(criterion) # criterion is set to either plant.sp (for taxonomic) or k.type.s (for morphological)
+
+  no1 <- x %>%
+    group_by(bb.sp, !!criterion, site) %>% # group by site-year and BB-FL pair
+    summarize(visits = sum(visits)) %>%
+    select(site, bb.sp, !!criterion, visits) %>%
+    pivot_wider(names_from = !!criterion, values_from = visits, values_fill = 0) %>% # spread into the kind of array vegan likes
+    group_by(site) %>%
+    nest() %>% # nest into list-column
+    mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
+           richness = map(data, function(x) length(x$bb.sp)),
+           dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
+           tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
+    #dplyr::select(site, date, tidy_dist) %>% # drop intermediate list-columns
+    unnest(cols = c(tidy_dist)) %>%
+    rename(bb.sp1 = item1, bb.sp2 = item2) %>%
+    mutate(niche.overlap = 1-distance,  # express niche overlap so that 1 = perfect overlap and 0 = no overlap
+           telu = if_else(bb.sp1 == "telu" | bb.sp2 == "telu", TRUE, FALSE),
+           pasc = if_else(bb.sp1 == "pasc" | bb.sp2 == "pasc", TRUE, FALSE),
+           wurf = if_else(bb.sp1 == "wurf" | bb.sp2 == "wurf", TRUE, FALSE),
+           prat = if_else(bb.sp1 == "prat" | bb.sp2 == "prat", TRUE, FALSE),
+           soro = if_else(bb.sp1 == "soro" | bb.sp2 == "soro", TRUE, FALSE),
+           humi = if_else(bb.sp1 == "humi" | bb.sp2 == "humi", TRUE, FALSE),
+           muci = if_else(bb.sp1 == "muci" | bb.sp2 == "muci", TRUE, FALSE),
+           pyre = if_else(bb.sp1 == "pyre" | bb.sp2 == "pyre", TRUE, FALSE),
+           mont = if_else(bb.sp1 == "mont" | bb.sp2 == "mont", TRUE, FALSE),
+           mend = if_else(bb.sp1 == "mend" | bb.sp2 == "mend", TRUE, FALSE),
+           psit = if_else(bb.sp1 == "psit" | bb.sp2 == "psit", TRUE, FALSE),
+           gers = if_else(bb.sp1 == "gers" | bb.sp2 == "gers", TRUE, FALSE),
+           jone = if_else(bb.sp1 == "jone" | bb.sp2 == "jone", TRUE, FALSE),
+           hypn = if_else(bb.sp1 == "hypn" | bb.sp2 == "hypn", TRUE, FALSE),
+           lapi = if_else(bb.sp1 == "lapi" | bb.sp2 == "lapi", TRUE, FALSE),
+           hort = if_else(bb.sp1 == "hort" | bb.sp2 == "hort", TRUE, FALSE))
+}
 ```
@@
 ### 3.1.2 My first question is how niche overlap compares across species pairs
-```{r}
-# Unaggregated niche overlap per species pair; not that this drops a lot of data because of the inner join with bias
+#### 3.1.2.1 Calculate pairwise niche overlap
+```{r}
+# Get bb.sp abundances; this time without filtering out the sites without matching floral data
+bb_abundance2 <-  read_csv("../data/processed_data/network.csv") %>%
+  filter(caste != "male") %>% # we will ignore males 
+  group_by(site, date, bb.sp, plant.sp) %>%
+  summarize(visits = n()) %>%
+  group_by(site, date, bb.sp) %>%
+  summarize(bb.sp.abund = sum(visits)) %>%
+  group_by(site, date) %>%
+  mutate(bb.sp.prop.abund = bb.sp.abund/sum(bb.sp.abund))
+
+bb_abundance2_agg <-  read_csv("../data/processed_data/network.csv") %>%
+  filter(caste != "male") %>% # we will ignore males 
+  group_by(site, bb.sp, plant.sp) %>%
+  summarize(visits = n()) %>%
+  group_by(site, bb.sp) %>%
+  summarize(bb.sp.abund = sum(visits)) %>%
+  group_by(site) %>%
+  mutate(bb.sp.prop.abund = bb.sp.abund/sum(bb.sp.abund))
+
+# Un-aggregated niche overlap per species pair per site-date
 niche_overlap <- read_csv("../data/processed_data/network.csv") %>%
@@
     ) %>%
-  left_join(bb_abundance, by = c("site", "date", "bb.sp1" = "bb.sp")) %>%
+  left_join(bb_abundance2, by = c("site", "date", "bb.sp1" = "bb.sp")) %>%
   rename(bb.sp1.abund = bb.sp.abund) %>%
-  left_join(bb_abundance, by = c("site", "date", "bb.sp2" = "bb.sp")) %>%
+  left_join(bb_abundance2, by = c("site", "date", "bb.sp2" = "bb.sp")) %>%
   rename(bb.sp2.abund = bb.sp.abund) %>%
@@
   select(site, date, yday, elev.mean, sp.pair, bb.sp1, bb.sp2, bb.sp1.abund, 
-         bb.sp2.abund, niche.overlap, same.class)
+         bb.sp2.abund, niche.overlap, same.class) %>%
+  mutate(log.bb.sp1.abund = log(bb.sp1.abund),
+         log.bb.sp2.abund = log(bb.sp2.abund))
+
+# Aggregated niche overlap per species pair per site
+niche_overlap2 <- read_csv("../data/processed_data/network.csv") %>%
+  filter(caste != "male") %>% # we will ignore males 
+  group_by(site, bb.sp, plant.sp) %>%
+  summarize(visits = n()) %>%
+  niche_overlapper2(plant.sp) %>%
+  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>%
+  rename(pbl.w.class2_sp1 = pbl.w.class2) %>%
+  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>%
+  rename(pbl.w.class2_sp2 = pbl.w.class2) %>%
+  mutate(same.class = case_when(
+    pbl.w.class2_sp1 == "1" & pbl.w.class2_sp2 == "1" ~ "short:short",
+    pbl.w.class2_sp1 == "2" & pbl.w.class2_sp2 == "2" ~ "med:med",
+    pbl.w.class2_sp1 == "3" & pbl.w.class2_sp2 == "3" ~ "long:long",
+    pbl.w.class2_sp1 == "1" & pbl.w.class2_sp2 == "2" ~ "short:med",
+    pbl.w.class2_sp1 == "2" & pbl.w.class2_sp2 == "1" ~ "short:med",
+    pbl.w.class2_sp1 == "1" & pbl.w.class2_sp2 == "3" ~ "short:long",
+    pbl.w.class2_sp1 == "3" & pbl.w.class2_sp2 == "1" ~ "short:long",
+    pbl.w.class2_sp1 == "2" & pbl.w.class2_sp2 == "3" ~ "med:long",
+    pbl.w.class2_sp1 == "3" & pbl.w.class2_sp2 == "2" ~ "med:long")
+    ) %>%
+  left_join(bb_abundance2_agg, by = c("site", "bb.sp1" = "bb.sp")) %>%
+  rename(bb.sp1.abund = bb.sp.abund) %>%
+  left_join(bb_abundance2_agg, by = c("site", "bb.sp2" = "bb.sp")) %>%
+  rename(bb.sp2.abund = bb.sp.abund) %>%
+  unite(sp.pair, c(bb.sp1, bb.sp2), sep = "_", remove = FALSE) %>%
+  left_join(site_data) %>%
+  select(site, elev.mean, sp.pair, bb.sp1, bb.sp2, bb.sp1.abund, 
+         bb.sp2.abund, niche.overlap, same.class) %>%
+  mutate(log.bb.sp1.abund = log(bb.sp1.abund),
+         log.bb.sp2.abund = log(bb.sp2.abund))
+
+write_csv(niche_overlap2, "../hpc_packets/niche_overlap2.csv")
+```
+
+#### 3.1.2.2 Preliminary visualization
+```{r}
+ggplot(niche_overlap, aes(bb.sp1.abund, bb.sp2.abund, color = niche.overlap)) +
+  geom_point()
+
+ggplot(niche_overlap, aes(bb.sp1.abund, niche.overlap)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 4))
+
+ggplot(niche_overlap, aes(bb.sp2.abund, niche.overlap)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 4))
+
+ggplot(niche_overlap, aes(log.bb.sp1.abund, niche.overlap)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 4))
+
+ggplot(niche_overlap, aes(log.bb.sp2.abund, niche.overlap)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 4))
+
+ggplot(niche_overlap, aes(log.bb.sp2.abund, niche.overlap, color = same.class)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 4))
+
+ggplot(niche_overlap, aes(sqrt(bb.sp2.abund), sqrt(niche.overlap), color = same.class)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 4))
+
+ggplot(niche_overlap, aes(bb.sp2.abund, niche.overlap, color = same.class)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 4))
+
+ggplot(niche_overlap, aes(log.bb.sp1.abund, log.bb.sp2.abund, color = niche.overlap)) +
+  geom_point()
 
@@
 ggplot(niche_overlap, aes(sp.pair, niche.overlap)) +
+  geom_violin() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+ggplot(niche_overlap, aes(same.class, niche.overlap)) +
   geom_violin() +
@@
 
+ggplot(filter(niche_overlap, bb.sp1.abund >= 2 & bb.sp1.abund >= 2), 
+       aes(same.class, niche.overlap)) +
+  geom_violin() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
 # Now a threshold of 3
@@
 ggplot(filter(niche_overlap, bb.sp1.abund >= 5 & bb.sp1.abund >= 5), 
-       aes(sp.pair, niche.overlap)) +
+       aes(same.class, niche.overlap)) +
   geom_violin() +
@@
 ggplot(filter(niche_overlap, bb.sp1.abund >= 10 & bb.sp1.abund >= 10), 
-       aes(sp.pair, niche.overlap)) +
+       aes(same.class, niche.overlap)) +
   geom_violin() +
@@
 
-# Now for what is probably a pretty naive model
-
+ggplot(niche_overlap, aes(sp.pair, niche.overlap)) +
+  geom_boxplot() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+# Aggregate by site
+ggplot(niche_overlap2, aes(bb.sp1.abund, niche.overlap)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))
+
+ggplot(niche_overlap2, aes(log.bb.sp1.abund, niche.overlap)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))
+
+ggplot(niche_overlap2, aes(log.bb.sp1.abund, niche.overlap)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~same.class)
+
+ggplot(niche_overlap2, aes(log.bb.sp1.abund, niche.overlap)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~sp.pair)
+
+ggplot(niche_overlap2, aes(sp.pair, niche.overlap)) +
+  geom_violin() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+ggplot(niche_overlap2, aes(sp.pair, niche.overlap)) +
+  geom_boxplot() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+ggplot(niche_overlap2, aes(same.class, niche.overlap)) +
+  geom_violin() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+ggplot(niche_overlap2, aes(elev.mean, niche.overlap)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE) +
+  facet_wrap(~sp.pair)
+
+ggplot(niche_overlap, aes(niche.overlap)) +
+  geom_histogram() +
+  facet_wrap(~sp.pair)
+
+## raw visits aggregate
+ggplot(niche_overlap3, aes(log.bb.sp1.abund, niche.overlap)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))
+
+ggplot(niche_overlap3, aes(same.class, niche.overlap)) +
+  geom_violin() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+ggplot(niche_overlap2, aes(niche.overlap)) +
+  geom_histogram() +
+  facet_wrap(~sp.pair)
+
+```
+
+#### 3.1.2.3 brms model ... still a mess
+```{r}
 brm_niche00 <- brm(niche.overlap ~ bb.sp1.abund*bb.sp2.abund + 
+                     (1 | sp.pair) + 
+                     (1 | same.class) + 
+                     (1 | site),
+                   control = list(adapt_delta = 0.90),
+                   chains = 4,
+                   cores = 4,
+                   iter = 2000,
+                   file = "../output/brm_niche00",
+                   data = niche_overlap)
+
+parnames(brm_niche00)
+conditional_effects(brm_niche00)
+mcmc_plot(brm_niche00, pars = c("^r_sp"))
+mcmc_plot(brm_niche00, pars = c("^r_sa"))
+mcmc_plot(brm_niche00, pars = c("^b_"))
+hypothesis(brm_niche00, "dprime = 0")
+pp_check(brm_niche00)
+bayes_R2(brm_niche00)
+
+# To stabilize niche overlap (avoid lots of  1s and 0s), we will filter the data to include only species pairs in which both species were represented by at least 3 observations.
+brm_niche01 <- brm(niche.overlap ~ bb.sp1.abund*bb.sp2.abund + 
                      (1 | sp.pair) + 
@@
                    iter = 2000,
-                   file = "../output/brm_niche00",
-                   data = filter(niche_overlap, bb.sp1.abund >= 1 & bb.sp1.abund >= 1))
+                   file = "../output/brm_niche01",
+                   data = filter(niche_overlap, bb.sp1.abund >= 2 & bb.sp1.abund >= 2))
+
+parnames(brm_niche01)
+conditional_effects(brm_niche01)
+mcmc_plot(brm_niche01, pars = c("^r_sp"))
+mcmc_plot(brm_niche01, pars = c("^r_sa"))
+mcmc_plot(brm_niche01, pars = c("^b_"))
+hypothesis(brm_niche01, "dprime = 0")
+pp_check(brm_niche01)
+bayes_R2(brm_niche01)
+
+# Would it help if we log-transformed the bb abundance data and stuck to unfiltered data?
+brm_niche02 <- brm(niche.overlap ~ log.bb.sp1.abund*log.bb.sp2.abund + 
+                     (1 | sp.pair) + 
+                     (1 | same.class) + 
+                     (1 | site),
+                   #control = list(adapt_delta = 0.90),
+                   chains = 4,
+                   cores = 4,
+                   iter = 2000,
+                   file = "../output/brm_niche02",
+                   data = niche_overlap)
 
@@
 
+brm_niche03 <- brm(niche.overlap ~ log.bb.sp1.abund*log.bb.sp2.abund + 
+                     (1 + log.bb.sp1.abund*log.bb.sp2.abund | sp.pair) + 
+                     (1 | same.class) + 
+                     (1 | site),
+                   #control = list(adapt_delta = 0.90),
+                   chains = 4,
+                   cores = 4,
+                   iter = 2000,
+                   file = "../output/brm_niche03",
+                   data = niche_overlap)
+
+parnames(brm_niche02)
+conditional_effects(brm_niche02)
+mcmc_plot(brm_niche02, pars = c("^r_sp"))
+mcmc_plot(brm_niche02, pars = c("^r_sa"))
+mcmc_plot(brm_niche02, pars = c("^b_"))
+hypothesis(brm_niche02, "dprime = 0")
+pp_check(brm_niche02)
+bayes_R2(brm_niche02)
+
+# A zero-one inflated model might be the key.
+brm_niche04 <- brm(niche.overlap ~ log.bb.sp1.abund*log.bb.sp2.abund + 
+                     (1 | sp.pair) + 
+                     (1 | same.class) + 
+                     (1 | site),
+                   #control = list(adapt_delta = 0.90),
+                   family = "zero_one_inflated_beta",
+                   chains = 4,
+                   cores = 4,
+                   iter = 2000,
+                   file = "../output/brm_niche04",
+                   data = niche_overlap)
+
+parnames(brm_niche04)
+conditional_effects(brm_niche04)
+mcmc_plot(brm_niche04, pars = c("^r_sp"))
+mcmc_plot(brm_niche04, pars = c("^r_sa"))
+mcmc_plot(brm_niche04, pars = c("^b_"))
+hypothesis(brm_niche04, "dprime = 0")
+pp_check(brm_niche04)
+bayes_R2(brm_niche04)
+
+# Aggregate models
+brm_niche60 <- brm(niche.overlap ~ log.bb.sp1.abund*log.bb.sp2.abund + 
+                     (1 | sp.pair) + 
+                     (1 | same.class) +
+                     (1 | site),
+                   family = "zero_one_inflated_beta",
+                   chains = 4,
+                   cores = 4,
+                   iter = 2000,
+                   file = "../output/brm_niche60",
+                   data = niche_overlap2)
+
+parnames(brm_niche60)
+conditional_effects(brm_niche60)
+mcmc_plot(brm_niche60, pars = c("^r_sp"))
+mcmc_plot(brm_niche60, pars = c("^r_sa"))
+mcmc_plot(brm_niche60, pars = c("^b_Int"))
+hypothesis(brm_niche60, "dprime = 0")
+pp_check(brm_niche60)
+bayes_R2(brm_niche60)
+
+brm_niche61 <- brm(niche.overlap ~ log.bb.sp1.abund*log.bb.sp2.abund + 
+                     (1 | sp.pair) + 
+                     (1 | same.class) +
+                     (1 | site),
+                   control = list(adapt_delta = 0.99),
+                   family = "zero_one_inflated_beta",
+                   chains = 4,
+                   cores = 4,
+                   iter = 2000,
+                   file = "../output/brm_niche61", # for local
+                   #file = "./output/brm_niche61", # for HPC
+                   data = niche_overlap2)
+
+parnames(brm_niche61)
+conditional_effects(brm_niche61)
+mcmc_plot(brm_niche61, pars = c("^r_sp"))
+mcmc_plot(brm_niche61, pars = c("^r_sa"))
+mcmc_plot(brm_niche61, pars = c("^b_Int"))
+pp_check(brm_niche61)
+bayes_R2(brm_niche61)
+
+brm_niche62 <- brm(niche.overlap ~ log.bb.sp1.abund*log.bb.sp2.abund + 
+                     (1 + log.bb.sp1.abund*log.bb.sp2.abund | sp.pair) + 
+                     (1 | same.class) +
+                     (1 | site),
+                   control = list(adapt_delta = 0.99),
+                   family = "zero_one_inflated_beta",
+                   chains = 4,
+                   cores = 4,
+                   iter = 2000,
+                   #file = "../output/brm_niche62", # for local
+                   file = "./output/brm_niche62", # for HPC
+                   data = niche_overlap2)
+```
+
+#### 3.2.2.4 Plot model
+```{r}
+## Sp.pair and same class, for convenience
+sp_pair_class <- niche_overlap2 %>%
+  ungroup() %>%
+  select(sp.pair, same.class) %>%
+  distinct()
+
+## Tongue length combos
+brm_niche61_plotframe1 <- brm_niche61 %>%
+  spread_draws(b_Intercept, r_same.class[condition, term]) %>% 
+  mutate(condition_mean = b_Intercept + r_same.class)
+
+ggplot(brm_niche61_plotframe1, aes(condition_mean, reorder(condition, condition_mean))) +
+  stat_halfeye() +
+  ylab("Tongue length combination") +
+  xlab("Niche overlap")
+
+ggsave(file = "../output/brm_niche61_plot1.png", width = 7, height = 5.5)
+
+## Species pairs
+brm_niche61_plotframe2 <- brm_niche61 %>%
+  spread_draws(b_Intercept, r_sp.pair[condition, term]) %>% 
+  mutate(condition_mean = b_Intercept + r_sp.pair) %>%
+  left_join(sp_pair_class, by = c("condition" = "sp.pair"))
+
+ggplot(brm_niche61_plotframe2, aes(condition_mean, 
+                                   reorder(condition, condition_mean), 
+                                   color = same.class)) +
+  stat_pointinterval() +
+  ylab("Species pair") +
+  xlab("Niche overlap")
+
+ggsave(file = "../output/brm_niche61_plot1.png", width = 7, height = 5.5)
 ```
@@
   rename(dprime.sp2 = dprime.sp) %>%
-  select(site, date, bb.sp1, bb.sp2, bb.sp1.abund, bb.sp2.abund, 
+  select(site, date, sp.pair, bb.sp1, bb.sp2, bb.sp1.abund, bb.sp2.abund, 
          dprime.sp1, dprime.sp2, niche.overlap, same.class)
@@
   #geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = TRUE) +
-  facet_wrap(~bb.sp) +
+  facet_wrap(~sp.pair) +
   ylim(c(0,1))
@@
 ## 3.2 Ordination of dietary niche
-```{r}
-# Aggregated by mean proportional visitation
-
-# We will use the whole dataset this time, since there is no need to align visitation to floral surveying
-visitation1 <- read_csv("../data/processed_data/network.csv") %>%
-  filter(caste != "male") %>% # we will ignore males 
-  group_by(site, date, bb.sp, plant.sp) %>%
-  summarize(visits = n()) %>%
-  group_by(site, date, bb.sp) %>%
-  mutate(prop.visits = visits/sum(visits)) %>%
-  group_by(bb.sp, plant.sp) %>%
-  summarize(mean.prop.visits = mean(prop.visits)) %>%
-  pivot_wider(names_from = plant.sp, values_from = mean.prop.visits, values_fill = 0) %>%
-  column_to_rownames("bb.sp")
-
+Aggregation changes meaning.
+Don't forget that.
+When I pool visitation across dates, I am introducing the effects of relative abundance and floral phenology.
+Species in bloom when BB abundance is high will be over-represented, and vice-versa.
+And species with a long bloom time will be over-represented, and vice-versa.
+This isn't necessarily a problem.
+I just need to make sure I interpret the results the right way.
+When I pool data like this, I am comparing the "net" diet of species, which includes the effects of abundance and phenology in addition to the effects of floral preferences.
+I think that is appropriate for this stage of the analysis, but I need to be careful not to assume that what I'm doing here is the same sort of analysis as what I did above without pooling samples.
+
+### 3.2.1 Fully pooled
+```{r}
+# Fl traits with abbreviated names
+fl_traits2 <- fl_traits %>%
+  left_join(read_csv("../data/processed_data/network.csv")) %>%
+  select(plant.sp, plant.sp.abb, plant.genus, plant.family, k.type.ss, color) %>%
+  mutate(k.type.ss = factor(k.type.ss)) %>%
+  distinct()
+
+# Totally aggregated; note I'm using the full data, not the aligned and trimmed data. No need for complementary floral survey data.
 visitation2 <- read_csv("../data/processed_data/network.csv") %>%
   filter(caste != "male") %>% # we will ignore males 
-  group_by(site, date, bb.sp, plant.sp) %>%
+  filter(!bb.sp %in% c("humi")) %>% # humi is too rare
+  group_by(bb.sp, plant.sp.abb) %>%
   summarize(visits = n()) %>%
-  group_by(site, date, bb.sp) %>%
+  group_by(bb.sp) %>%
   mutate(prop.visits = visits/sum(visits)) %>%
-  unite(id, c(site, date, bb.sp)) %>%
+  select(-visits) %>%
+  pivot_wider(names_from = plant.sp.abb, values_from = prop.visits, values_fill = 0) %>%
+  column_to_rownames("bb.sp")
+
+# Generate ordination
+pcoa_horn2 <- capscale(visitation2 ~ 1, distance = "horn")
+
+# Get variance explained
+pcoa_horn2_mds1 <- round(pcoa_horn2$CA$eig[[1]] / sum(pcoa_horn2$CA$eig)*100)
+pcoa_horn2_mds2 <- round(pcoa_horn2$CA$eig[[2]] / sum(pcoa_horn2$CA$eig)*100)
+
+# Fortify plants
+pcoa_horn2.fort1 <- fortify(pcoa_horn2, axes = 1:2) %>%
+  filter(Score == "species") %>%
+  rename(plant.sp.abb = Label) %>%
+  left_join(fl_traits2)
+
+# Fortify bees
+pcoa_horn2.fort2 <- fortify(pcoa_horn2, axes = 1:2) %>%
+  filter(Score == "sites") %>%
+  rename(bb.sp = Label) %>%
+  left_join(bb_traits) %>%
+  mutate(pbl.w.class2 = factor(pbl.w.class2)) %>%
+  mutate(pbl.w.class2 = case_when(
+    pbl.w.class2 == "1" ~ "short",
+    pbl.w.class2 == "2" ~ "med",
+    pbl.w.class2 == "3" ~ "long"
+    ))
+
+# Generate convex hulls around tongue length classes
+pcoa_horn2_hulls <- pcoa_horn2.fort2 %>%
+  group_by(pbl.w.class2) %>%
+  slice(chull(MDS1, MDS2))
+
+# Plot
+ggplot(pcoa_horn2.fort2, aes(MDS1, MDS2, color = pbl.w.class2)) +
+  geom_polygon(data = pcoa_horn2_hulls, 
+               aes(fill = pbl.w.class2, color = NULL),
+               alpha = 0.25) +
+  geom_text(aes(label = bb.sp), key_glyph = "rect") +
+  geom_text_repel(data = filter(pcoa_horn2.fort1, abs(MDS1) >= 0.2 | abs(MDS2) >= 0.2), 
+            aes(MDS1, MDS2, label = plant.sp.abb), inherit.aes = FALSE, color = "gray40") +
+  geom_point(data = filter(pcoa_horn2.fort1, abs(MDS1) >= 0.2 | abs(MDS2) >= 0.2),
+            aes(MDS1, MDS2, shape = k.type.ss), inherit.aes = FALSE, color = "gray40", size = 3) +
+  scale_shape_discrete() +
+  xlab(paste("MDS1 ", "(", pcoa_horn2_mds1, "%", ")", sep = "")) +
+  ylab(paste("MDS2 ", "(", pcoa_horn2_mds2, "%", ")", sep = "")) +
+  labs(fill = "Tongue length", 
+       shape = "Morphotype") +
+  guides(color = FALSE) +
+  coord_fixed() +
+  theme_light(10)
+
+ggsave("../output/pcoa_aggregate.png")
+```
+
+### 3.2.1 Pooled within elev.class2
+```{r}
+# Fl traits with abbreviated names
+fl_traits2 <- fl_traits %>%
+  left_join(read_csv("../data/processed_data/network.csv")) %>%
+  select(plant.sp, plant.sp.abb, plant.genus, plant.family, k.type.ss, color) %>%
+  mutate(k.type.ss = factor(k.type.ss)) %>%
+  distinct()
+
+# Get visitation
+visitation_low <- read_csv("../data/processed_data/network.csv") %>%
+  filter(caste != "male") %>% # we will ignore males 
+  filter(!bb.sp %in% c("humi")) %>% # humi is too rare
+  left_join(site_data) %>%
+  filter(elev.class2 == "low") %>%
+  group_by(bb.sp, plant.sp.abb) %>%
+  summarize(visits = n()) %>%
+  group_by(bb.sp) %>%
+  mutate(prop.visits = visits/sum(visits)) %>%
+  select(-visits) %>%
+  pivot_wider(names_from = plant.sp.abb, values_from = prop.visits, values_fill = 0) %>%
+  column_to_rownames("bb.sp")
+
+visitation_mid <- read_csv("../data/processed_data/network.csv") %>%
+  filter(caste != "male") %>% # we will ignore males 
+  filter(!bb.sp %in% c("humi")) %>% # humi is too rare
+  left_join(site_data) %>%
+  filter(elev.class2 == "mid") %>%
+  group_by(bb.sp, plant.sp.abb) %>%
+  summarize(visits = n()) %>%
+  group_by(bb.sp) %>%
+  mutate(prop.visits = visits/sum(visits)) %>%
+  select(-visits) %>%
+  pivot_wider(names_from = plant.sp.abb, values_from = prop.visits, values_fill = 0) %>%
+  column_to_rownames("bb.sp")
+
+visitation_high <- read_csv("../data/processed_data/network.csv") %>%
+  filter(caste != "male") %>% # we will ignore males 
+  filter(!bb.sp %in% c("humi")) %>% # humi is too rare
+  left_join(site_data) %>%
+  filter(elev.class2 == "high") %>%
+  group_by(bb.sp, plant.sp.abb) %>%
+  summarize(visits = n()) %>%
+  group_by(bb.sp) %>%
+  mutate(prop.visits = visits/sum(visits)) %>%
+  select(-visits) %>%
+  pivot_wider(names_from = plant.sp.abb, values_from = prop.visits, values_fill = 0) %>%
+  column_to_rownames("bb.sp")
+
+# Generate ordination
+pcoa_horn_low <- capscale(visitation_low ~ 1, distance = "horn")
+pcoa_horn_mid <- capscale(visitation_mid ~ 1, distance = "horn")
+pcoa_horn_high <- capscale(visitation_high ~ 1, distance = "horn")
+
+# Get variance explained
+pcoa_horn_low_mds1 <- round(pcoa_horn_low$CA$eig[[1]] / sum(pcoa_horn_low$CA$eig)*100)
+pcoa_horn_low_mds2 <- round(pcoa_horn_low$CA$eig[[2]] / sum(pcoa_horn_low$CA$eig)*100)
+
+pcoa_horn_mid_mds1 <- round(pcoa_horn_mid$CA$eig[[1]] / sum(pcoa_horn_mid$CA$eig)*100)
+pcoa_horn_mid_mds2 <- round(pcoa_horn_mid$CA$eig[[2]] / sum(pcoa_horn_mid$CA$eig)*100)
+
+pcoa_horn_high_mds1 <- round(pcoa_horn_high$CA$eig[[1]] / sum(pcoa_horn_high$CA$eig)*100)
+pcoa_horn_high_mds2 <- round(pcoa_horn_high$CA$eig[[2]] / sum(pcoa_horn_high$CA$eig)*100)
+
+# Fortify plants
+pcoa_horn.fort1_low <- fortify(pcoa_horn_low, axes = 1:2) %>%
+  filter(Score == "species") %>%
+  rename(plant.sp.abb = Label) %>%
+  left_join(fl_traits2)
+
+pcoa_horn.fort1_mid <- fortify(pcoa_horn_mid, axes = 1:2) %>%
+  filter(Score == "species") %>%
+  rename(plant.sp.abb = Label) %>%
+  left_join(fl_traits2)
+
+pcoa_horn.fort1_high <- fortify(pcoa_horn_high, axes = 1:2) %>%
+  filter(Score == "species") %>%
+  rename(plant.sp.abb = Label) %>%
+  left_join(fl_traits2)
+
+# Fortify bees
+pcoa_horn.fort2_low <- fortify(pcoa_horn_low, axes = 1:2) %>%
+  filter(Score == "sites") %>%
+  rename(bb.sp = Label) %>%
+  left_join(bb_traits) %>%
+  mutate(pbl.w.class2 = factor(pbl.w.class2)) %>%
+  mutate(pbl.w.class2 = case_when(
+    pbl.w.class2 == "1" ~ "short",
+    pbl.w.class2 == "2" ~ "med",
+    pbl.w.class2 == "3" ~ "long"
+    ))
+
+pcoa_horn.fort2_mid <- fortify(pcoa_horn_mid, axes = 1:2) %>%
+  filter(Score == "sites") %>%
+  rename(bb.sp = Label) %>%
+  left_join(bb_traits) %>%
+  mutate(pbl.w.class2 = factor(pbl.w.class2)) %>%
+  mutate(pbl.w.class2 = case_when(
+    pbl.w.class2 == "1" ~ "short",
+    pbl.w.class2 == "2" ~ "med",
+    pbl.w.class2 == "3" ~ "long"
+    ))
+
+pcoa_horn.fort2_high <- fortify(pcoa_horn_high, axes = 1:2) %>%
+  filter(Score == "sites") %>%
+  rename(bb.sp = Label) %>%
+  left_join(bb_traits) %>%
+  mutate(pbl.w.class2 = factor(pbl.w.class2)) %>%
+  mutate(pbl.w.class2 = case_when(
+    pbl.w.class2 == "1" ~ "short",
+    pbl.w.class2 == "2" ~ "med",
+    pbl.w.class2 == "3" ~ "long"
+    ))
+
+# Generate convex hulls around tongue length classes
+pcoa_horn_hulls_low <- pcoa_horn.fort2_low %>%
+  group_by(pbl.w.class2) %>%
+  slice(chull(MDS1, MDS2))
+
+pcoa_horn_hulls_mid <- pcoa_horn.fort2_mid %>%
+  group_by(pbl.w.class2) %>%
+  slice(chull(MDS1, MDS2))
+
+pcoa_horn_hulls_high <- pcoa_horn.fort2_high %>%
+  group_by(pbl.w.class2) %>%
+  slice(chull(MDS1, MDS2))
+
+# Plot
+ggplot(pcoa_horn.fort2_low, aes(MDS1, MDS2, color = pbl.w.class2)) +
+  geom_polygon(data = pcoa_horn_hulls_low, 
+               aes(fill = pbl.w.class2, color = NULL),
+               alpha = 0.25) +
+  geom_text(aes(label = bb.sp), key_glyph = "rect") +
+  geom_text_repel(data = filter(pcoa_horn.fort1_low, abs(MDS1) >= 0.2 | abs(MDS2) >= 0.2), 
+            aes(MDS1, MDS2, label = plant.sp.abb), inherit.aes = FALSE, color = "gray40") +
+  geom_point(data = filter(pcoa_horn.fort1_low, abs(MDS1) >= 0.2 | abs(MDS2) >= 0.2),
+            aes(MDS1, MDS2, shape = k.type.ss), inherit.aes = FALSE, color = "gray40", size = 3) +
+  scale_shape_discrete() +
+  xlab(paste("MDS1 ", "(", pcoa_horn_low_mds1, "%", ")", sep = "")) +
+  ylab(paste("MDS2 ", "(", pcoa_horn_low_mds2, "%", ")", sep = "")) +
+  labs(fill = "Tongue length", 
+       shape = "Morphotype") +
+  guides(color = FALSE) +
+  coord_fixed() +
+  theme_light(10)
+
+ggplot(pcoa_horn.fort2_mid, aes(MDS1, MDS2, color = pbl.w.class2)) +
+  geom_polygon(data = pcoa_horn_hulls_mid, 
+               aes(fill = pbl.w.class2, color = NULL),
+               alpha = 0.25) +
+  geom_text(aes(label = bb.sp), key_glyph = "rect") +
+  geom_text_repel(data = filter(pcoa_horn.fort1_mid, abs(MDS1) >= 0.2 | abs(MDS2) >= 0.2), 
+            aes(MDS1, MDS2, label = plant.sp.abb), inherit.aes = FALSE, color = "gray40") +
+  geom_point(data = filter(pcoa_horn.fort1_mid, abs(MDS1) >= 0.2 | abs(MDS2) >= 0.2),
+            aes(MDS1, MDS2, shape = k.type.ss), inherit.aes = FALSE, color = "gray40", size = 3) +
+  scale_shape_discrete() +
+  xlab(paste("MDS1 ", "(", pcoa_horn_mid_mds1, "%", ")", sep = "")) +
+  ylab(paste("MDS2 ", "(", pcoa_horn_mid_mds2, "%", ")", sep = "")) +
+  labs(fill = "Tongue length", 
+       shape = "Morphotype") +
+  guides(color = FALSE) +
+  coord_fixed() +
+  theme_light(10)
+
+ggplot(pcoa_horn.fort2_high, aes(MDS1, MDS2, color = pbl.w.class2)) +
+  geom_polygon(data = pcoa_horn_hulls_high, 
+               aes(fill = pbl.w.class2, color = NULL),
+               alpha = 0.25) +
+  geom_text(aes(label = bb.sp), key_glyph = "rect") +
+  geom_text_repel(data = filter(pcoa_horn.fort1_high, abs(MDS1) >= 0.2 | abs(MDS2) >= 0.2), 
+            aes(MDS1, MDS2, label = plant.sp.abb), inherit.aes = FALSE, color = "gray40") +
+  geom_point(data = filter(pcoa_horn.fort1_high, abs(MDS1) >= 0.2 | abs(MDS2) >= 0.2),
+            aes(MDS1, MDS2, shape = k.type.ss), inherit.aes = FALSE, color = "gray40", size = 3) +
+  scale_shape_discrete() +
+  xlab(paste("MDS1 ", "(", pcoa_horn_high_mds1, "%", ")", sep = "")) +
+  ylab(paste("MDS2 ", "(", pcoa_horn_high_mds2, "%", ")", sep = "")) +
+  labs(fill = "Tongue length", 
+       shape = "Morphotype") +
+  guides(color = FALSE) +
+  coord_fixed() +
+  theme_light(10)
+
+ggsave("../output/pcoa_aggregate.png")
+```
+
+```{r}
+# Totally aggregated
+visitation2 <- read_csv("../data/processed_data/network.csv") %>%
+  filter(caste != "male") %>% # we will ignore males 
+  filter(!bb.sp %in% c("humi")) %>% # humi is too rare
+  group_by(bb.sp, plant.sp) %>%
+  summarize(visits = n()) %>%
+  group_by(bb.sp) %>%
+  mutate(prop.visits = visits/sum(visits)) %>%
   select(-visits) %>%
   pivot_wider(names_from = plant.sp, values_from = prop.visits, values_fill = 0) %>%
-  column_to_rownames("id")
-
-# Aggregated
-pcoa_horn1 <- capscale(visitation1 ~ 1, distance = "horn")
-pcoa_horn1.fort1 <- fortify(pcoa_horn1, axes = 1:2) %>%
+  column_to_rownames("bb.sp")
+
+# Ordinated separately by site
+visitation3 <- read_csv("../data/processed_data/network.csv") %>%
+  filter(caste != "male") %>% # we will ignore males 
+  filter(!bb.sp %in% c("humi")) %>% # humi is too rare
+  group_by(site, bb.sp, plant.sp) %>%
+  summarize(visits = n()) %>%
+  group_by(site, bb.sp) %>%
+  mutate(prop.visits = visits/sum(visits)) %>%
+  select(-visits) %>%
+  group_by(site) %>%
+  nest() %>%
+  mutate(data2 = map(data, 
+                     function(x) pivot_wider(x, 
+                                             names_from = plant.sp, 
+                                             values_from = prop.visits, 
+                                             values_fill = 0)),
+         data3 = map(data2, function(x) column_to_rownames(x, "bb.sp")),
+         pcoa = map(data3, function(x) capscale(x ~ 1, distance = "horn")),
+         pcoa_quickplot = map(pcoa, function(x) plot(x, type = "t")))
+
+plot(visitation3$pcoa[[1]], type = "t")
+plot(visitation3$pcoa[[2]], type = "t")
+plot(visitation3$pcoa[[3]], type = "t")
+plot(visitation3$pcoa[[4]], type = "t")
+plot(visitation3$pcoa[[5]], type = "t")
+plot(visitation3$pcoa[[6]], type = "t")
+plot(visitation3$pcoa[[7]], type = "t")
+plot(visitation3$pcoa[[8]], type = "t")
+plot(visitation3$pcoa[[9]], type = "t")
+plot(visitation3$pcoa[[10]], type = "t")
+
+# Aggregation 1
+
+# Aggregation 2
+pcoa_horn2 <- capscale(visitation2 ~ 1, distance = "horn")
+plot(pcoa_horn2, type = "t")
+
+pcoa_horn2.fort1 <- fortify(pcoa_horn2, axes = 1:2) %>%
   filter(Score == "species") %>%
@@
   left_join(fl_traits)
-pcoa_horn1.fort2 <- fortify(pcoa_horn1, axes = 1:2) %>%
+pcoa_horn2.fort2 <- fortify(pcoa_horn2, axes = 1:2) %>%
   filter(Score == "sites") %>%
@@
 
-ggplot(pcoa_horn1.fort2, aes(MDS1, MDS2)) +
+ggplot(pcoa_horn2.fort2, aes(MDS1, MDS2)) +
   geom_text(aes(label = bb.sp)) +
+  coord_fixed()
+
+ggplot(pcoa_horn2.fort2, aes(MDS1, MDS2)) +
+  geom_text(aes(label = bb.sp, color = factor(pbl.w.class2))) +
+  geom_point(aes(color = factor(pbl.w.class2))) +
   coord_fixed()

### 2021-01-27 code/visitation_bias.Rmd
beta regression

@@
          elev0 = scale(elev),
+         bb.sp.abund0 = scale(bb.sp.abund),
          dprime.sp0 = scale(dprime.sp),
@@
          dprime.ktype0, dprime.fam, dprime.fam0, elev, elev0, 
-         yday, yday0, bb.sp.abund, bb.sp.prop.abund, total.flower.cover)
+         yday, yday0, bb.sp.abund, bb.sp.abund0, bb.sp.prop.abund, total.flower.cover)
+
+bias2 <- full_join(bias_plantsp, bias_ktype) %>%
+  full_join(bias_family) %>%
+  left_join(site_data) %>% # add site data
+  left_join(bb_abundance) %>% # add BB abundance data
+  left_join(fl_abundance) %>% # add fl abundance
+  #filter(bb.sp.abund >= 3) %>% # dprime tends to become binary when n is very low; this threshold removes more than half the data points, but almost all those were ones and zeros.
+These ones and zeros probably don't mean much, and they make the models hard to fit.
+  mutate(bb.sp = factor(bb.sp), 
+         yday = yday(date),
+         elev = elev.mean/10) %>% #  scale this down for convenience
+  mutate(yday0 = scale(yday),  # normalized versions of variables
+         elev0 = scale(elev),
+         bb.sp.abund0 = scale(bb.sp.abund),
+         dprime.sp0 = scale(dprime.sp),
+         dprime.ktype0 = scale(dprime.ktype),
+         dprime.fam0 = scale(dprime.fam)) %>%
+  select(site, date, bb.sp, dprime.sp, dprime.sp0, dprime.ktype, 
+         dprime.ktype0, dprime.fam, dprime.fam0, elev, elev0, 
+         yday, yday0, bb.sp.abund, bb.sp.abund0, bb.sp.prop.abund, total.flower.cover)
+
+ggplot(bias, aes(dprime.sp)) +
+  geom_histogram()
+
+ggplot(bias1, aes(dprime.sp)) +
+  geom_histogram()
+
+ggplot(bias2, aes(dprime.sp)) +
+  geom_histogram()
 ```
@@
   geom_histogram()
+
+ggplot(bias2, aes(bb.sp.abund, dprime.sp)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~bb.sp)
 
@@
 My reason for scaling and centering the predictors is that is massively speeds up estimation and enables the interaction model to converge (it would not converge with raw variables).
-My reason for not scaling the response is that it doesn't change the outcome of the model (I've checked) and I am interested in seeing the global intercept for d' on its native scale so that I can compare it with previously reported values.
+My reason for not scaling the response is that it doesn't change the outcome of the model (I've checked) and I am interested in seeing the global intercept for d' on its original scale so that I can compare it with previously reported values.
 Right now, I am using all default priors.
@@
 If I want to define my own priors, which I know I really should, particularly to replace the flat priors on the global effects, then I need to understand how the scaling of the variables affects the specification of the prior.
-```{r}
-# elev*yday interaction model
-brm_dprime.sp00 <- brm(dprime.sp ~ elev0 * yday0 + (1 | bb.sp),
-                       #+ (1 | site),
+
+I am also opting to use beta regression because it acknowledges the (0,1) boundedness of d'.
+There is a lovely paper on this topic (Douma and Weedon 2019, MEE), and in it the authors explicitly note that beta regression can be applied to derived metrics that are (0,1) bounded.
+
+"Although we present beta regression in this review as a method for analysing continuous‐based proportions, there are also many examples for using this techniques to analyse data from derived indices that are bound between 0 and 1 (e.g.
+an evenness index in Nogueira, González‐Troncoso, and Tolimieri (2016); or a straightness index in Shimada et al.
+(2016)).
+In addition, it can be applied to variables that are constrained to the interval a and b as they can be rescaled to [0, 1] through (y − a)/(b − a)."
+
+Importantly, the results of beta regression are on the logit scale, so they have to be back-transformed using the inv_logit_scaled() function.
+
+I've played around with every way I can think of to add elevation as a predictor, but there just isn't an elevation effect to be found, and loo model evaluation always favors a varying site intercept to a fixed elevation effect.
+I'm removing those models for the time being to clean up my script.
+I can always run them again if we decide it is important ot show the model selection process for the paper.
+```{r}
+# Full data
+brm_dprime.sp10 <- brm(dprime.sp ~ yday0 + bb.sp.abund0 + (1 | bb.sp) + (1 | site),
                        cores = 4,
                        iter = 2000,
-                       family = "skew_normal", # this accounts for a slight left-skew in the posterior
+                       family = "zero_one_inflated_beta", # 
+                       file = "../output/brm_dprime.sp10",
+                       data = bias2)
+
+summary(brm_dprime.sp10)
+conditional_effects(brm_dprime.sp10)
+prior_summary(brm_dprime.sp10)
+parnames(brm_dprime.sp10)
+pp_check(brm_dprime.sp10)
+bayes_R2(brm_dprime.sp10)
+mcmc_plot(brm_dprime.sp10, pars = "^r_bb.sp")
+mcmc_plot(brm_dprime.sp10, pars = "^r_site")
+
+# Abundance threshold >= 3
+brm_dprime.sp00 <- brm(dprime.sp ~ yday0 + bb.sp.abund0 + (1 | bb.sp) + (1 | site),
+                       cores = 4,
+                       iter = 2000,
+                       family = "zero_one_inflated_beta", # 
                        file = "../output/brm_dprime.sp00",
@@
 summary(brm_dprime.sp00)
+conditional_effects(brm_dprime.sp00)
 prior_summary(brm_dprime.sp00)
@@
 bayes_R2(brm_dprime.sp00)
-
-# Intercept-only model
-brm_dprime.sp01 <- brm(dprime.sp ~ 1 + (1 | bb.sp) + (1 | site),
-                       cores = 4,
-                       iter = 2000,
-                       family = "skew_normal",
-                       file = "../output/brm_dprime.sp01",
-                       data = bias)
-
-summary(brm_dprime.sp01)
-prior_summary(brm_dprime.sp01)
-parnames(brm_dprime.sp01)
-pp_check(brm_dprime.sp01)
-bayes_R2(brm_dprime.sp01)
-
-# Compare interaction model and intercept-only model. This favors the intercept-only model.
-loo(brm_dprime.sp00, brm_dprime.sp01)
+mcmc_plot(brm_dprime.sp00, pars = "^r_bb.sp")
+mcmc_plot(brm_dprime.sp00, pars = "^r_site")
 ```
@@
 ```{r}
-get_variables(brm_dprime.sp01)
-
-brm_dprime.sp01_plotframe <- brm_dprime.sp01 %>%
+# Full data with zero- one-inflation
+get_variables(brm_dprime.sp10)
+
+brm_dprime.sp10_plotframe <- brm_dprime.sp10 %>%
   spread_draws(b_Intercept, r_bb.sp[condition, term]) %>% 
-  mutate(condition_mean = b_Intercept + r_bb.sp)
-
-brm_dprime.sp01_plotframe2 <- brm_dprime.sp01 %>%
+  mutate(condition_mean = b_Intercept + r_bb.sp) %>%
+  mutate(condition_mean_invlogit = inv_logit_scaled(condition_mean))
+
+brm_dprime.sp10_plotframe2 <- brm_dprime.sp10 %>%
   spread_draws(b_Intercept, r_site[condition, term]) %>% 
-  mutate(condition_mean = b_Intercept + r_site)
-
-brm_dprime.sp01_subplot1 <- ggplot(brm_dprime.sp01_plotframe, aes(condition_mean, condition)) +
+  mutate(condition_mean = b_Intercept + r_site) %>%
+  mutate(condition_mean_invlogit = inv_logit_scaled(condition_mean))
+
+brm_dprime.sp10_subplot1 <- ggplot(brm_dprime.sp10_plotframe, aes(condition_mean_invlogit, condition)) +
   stat_halfeye() +
@@
 
-brm_dprime.sp01_subplot2 <- ggplot(brm_dprime.sp01_plotframe2, aes(condition_mean, condition)) +
+brm_dprime.sp10_subplot2 <- ggplot(brm_dprime.sp10_plotframe2, aes(condition_mean_invlogit, condition)) +
   stat_halfeye() +
@@
 
-brm_dprime.sp01_plot1 <- brm_dprime.sp01_subplot1 | brm_dprime.sp01_subplot2
-
-ggsave(brm_dprime.sp01_plot1, file = "../output/brm_dprime.sp01_plot1.png")
+brm_dprime.sp10_plot1 <- brm_dprime.sp10_subplot1 | brm_dprime.sp10_subplot2
+
+ggsave(brm_dprime.sp10_plot1, file = "../output/brm_dprime.sp10_plot1.png")
+
+# Subsetted data with skew normal 
+get_variables(brm_dprime.sp00)
+
+brm_dprime.sp00_plotframe <- brm_dprime.sp00 %>%
+  spread_draws(b_Intercept, r_bb.sp[condition, term]) %>% 
+  mutate(condition_mean = b_Intercept + r_bb.sp) %>%
+  mutate(condition_mean_invlogit = inv_logit_scaled(condition_mean))
+
+brm_dprime.sp00_plotframe2 <- brm_dprime.sp00 %>%
+  spread_draws(b_Intercept, r_site[condition, term]) %>% 
+  mutate(condition_mean = b_Intercept + r_site) %>%
+  mutate(condition_mean_invlogit = inv_logit_scaled(condition_mean))
+
+brm_dprime.sp00_subplot1 <- ggplot(brm_dprime.sp00_plotframe, aes(condition_mean_invlogit, condition)) +
+  stat_halfeye() +
+  ylab("Bumble bee species") +
+  xlab("Discrimination (d')") +
+  theme_light()
+
+brm_dprime.sp00_subplot2 <- ggplot(brm_dprime.sp00_plotframe2, aes(condition_mean_invlogit, condition)) +
+  stat_halfeye() +
+  ylab("Site") +
+  xlab("Discrimination (d')") +
+  theme_light()
+
+brm_dprime.sp00_plot1 <- brm_dprime.sp00_subplot1 | brm_dprime.sp00_subplot2
+
+ggsave(brm_dprime.sp00_plot1, file = "../output/brm_dprime.sp00_plot1.png")
 ```
@@
 ```{r}
+# Zero- one-inflated beta regression
+brm_pvis01 <- brm(prop.visits ~ prop.flower.cover0 + 
+                    (1 | k.type.ss) + 
+                    (1 | site) + 
+                    (1 | bb.sp),
+                  family = "zero_one_inflated_beta",
+                  file = "./output/brm_pvis01",
+                  chains = 4,
+                  cores = 4,
+                  iter = 2000,
+                  data = pvis_ktype)
+
 # Poisson models of raw visit counts
@@
                   family = "poisson",
-                  file = "./output/brm_pvis00b",
+                  file = "../output/brm_pvis00",
                   chains = 4,
@@
 
+niche_overlap_bias2 <- niche_overlap %>%
+  inner_join(bias2, by = c("site", "date", c("bb.sp1" = "bb.sp"))) %>%
+  rename(dprime.sp1 = dprime.sp) %>%
+  inner_join(bias, by = c("site", "date", c("bb.sp2" = "bb.sp"))) %>%
+  rename(dprime.sp2 = dprime.sp) %>%
+  select(site, date, sp.pair, bb.sp1, bb.sp2, bb.sp1.abund, bb.sp2.abund, 
+         dprime.sp1, dprime.sp2, niche.overlap, same.class)
+
+ggplot(niche_overlap_bias, aes(dprime.sp1, dprime.sp2, color = niche.overlap)) +
+  geom_point()
+
+ggplot(niche_overlap_bias, aes(log(bb.sp1.abund), log(bb.sp2.abund), color = niche.overlap)) +
+  geom_point()
+
 # When bias is high, it could mean that there are few good options on which all (or many) BB species converge.

### 2021-01-28 code/visitation_bias.Rmd
back to lappy

@@
 # 2. Explaining bias
-## 2.1 Calculate visitation probability per plant speces/ktype/family
+## 2.1 Calculate visitation probability per plant morphotype.
+Instead of keeping the plant species resolved, we will pool visitation by morphotype.
+The problem with not pooling by morphotype is that more species-rich morphotypes will be penalized simply because their visitation is diluted across more species.
 ```{r}
 pvis_ktype <- visit_cover %>%
-  group_by(site, date, bb.sp, plant.sp) %>%
+  group_by(site, date, bb.sp, k.type.ss) %>%
   summarize(flower.cover = sum(flower.cover), # get raw flower cover and visits by site*date*bb.sp*plant.sp
@@
   ungroup() %>%
-  left_join(fl_traits) %>% # add floral traits
   left_join(bb_traits) %>% # add BB traits
@@
   left_join(site_data)  %>%
+  ungroup() %>%
   mutate(prop.flower.cover0 = scale(prop.flower.cover), # normalized predictors
@@
          yday0 = scale(yday)) %>%
-  select(site, elev.mean0, elev.mean, elev.class2, date, yday, yday0, bb.sp, plant.sp, visits, 
-         visits.bin, prop.visits, prop.flower.cover, prop.flower.cover0,
-         k.type.ss, bb.sp.abund, bb.sp.abund0, pbl.w.class2) 
+  select(site, elev.mean0, elev.mean, elev.class2, date, yday, yday0, 
+         bb.sp, k.type.ss, visits, visits.bin, prop.visits, prop.flower.cover, 
+         prop.flower.cover0, bb.sp.abund, bb.sp.abund0, pbl.w.class2)
+
+write_csv(pvis_ktype, "../hpc_packets/pvis_ktype.csv")
+
+# We can also pool across BB spp. to focus just on how much visitation each ktype gets
+pvis_ktype_pool <- visit_cover %>%
+  group_by(site, date, k.type.ss) %>%
+  summarize(flower.cover = sum(flower.cover), # get raw flower cover and visits by site*date*bb.sp*plant.sp
+            visits = sum(visits)) %>%
+  group_by(site, date) %>%
+  mutate(prop.flower.cover = flower.cover/sum(flower.cover), # proportional flower cover
+         prop.visits = visits/sum(visits), # proportional visits
+         visits.bin = factor(if_else(visits > 0, 1, 0))) %>% # binary visits
+  group_by(site, date) %>%
+  mutate(bb.abund = sum(visits)) %>% # abundance of each BB sp. per site-date
+  ungroup() %>%
+  mutate(k.type.ss = factor(k.type.ss), # recode variables
+         visits = as.integer(visits)) %>%
+  filter(!k.type.ss %in% c("9", "10")) %>% # Drop ktypes 9 and 10 since both are extremely rare
+  left_join(site_data)  %>%
+  ungroup() %>%
+  mutate(prop.flower.cover0 = scale(prop.flower.cover), # normalized predictors
+         bb.abund0 = scale(bb.abund),
+         elev.mean0 = scale(elev.mean),
+         yday = yday(date), # and add yday variable
+         yday0 = scale(yday)) %>%
+  select(site, elev.mean0, elev.mean, elev.class2, date, yday, yday0, 
+         k.type.ss, visits, visits.bin, prop.visits, prop.flower.cover, 
+         prop.flower.cover0, bb.abund, bb.abund0)
+
+pvis_ktype_sitepool <- pvis_ktype %>%
+  group_by(site, elev.mean, bb.sp, pbl.w.class2, k.type.ss) %>%
+  summarize(mean.prop.visits = mean(prop.visits),
+            mean.prop.flower.cover = mean(prop.flower.cover))
+
+pvis_ktype_sitepool2 <- pvis_ktype %>%
+  group_by(site, elev.mean, bb.sp, pbl.w.class2, k.type.ss) %>%
+  summarize(mean.prop.visits = weighted.mean(prop.visits, bb.sp.abund),
+            mean.prop.flower.cover = mean(prop.flower.cover))
+
+ggplot(pvis_ktype_sitepool, aes(mean.prop.flower.cover, mean.prop.visits)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~k.type.ss)
+
+ggplot(pvis_ktype_sitepool, aes(mean.prop.visits)) +
+  geom_histogram() +
+  facet_grid(pbl.w.class2~k.type.ss)
+
+ggplot(pvis_ktype_sitepool2, aes(mean.prop.visits)) +
+  geom_histogram() +
+  facet_grid(pbl.w.class2~k.type.ss)
 ```
@@
 ## 2.2 Multi-level models
-
-### 2.2.1 Preliminary visualization
-The tons of zeros are due to the lots of plant species that were present at very low abundance and received few or zero visits.
-This might not be ideal.
-We could opt for one of our alternative bias measure, like fold change, which all give something like a poisson distribution, but I don't know if this would be an improvement.
-We could also filter our data to include only plants that were above some threshold of relative abundance, say, 1%.
-```{r}
-# Visit.prop
-ggplot(filter(pvis_ktype, prop.flower.cover >= 0.05), aes(prop.visits)) +
-  geom_histogram()
-
-# Visits.bin
-ggplot(pvis_ktype, aes(k.type.ss, fill = visits.bin)) +
-  geom_bar()
-
-ggplot(pvis_ktype, aes(elev.mean, prop.visits)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(bb.sp~k.type.ss)
-
-# Visits.bin filtered
-ggplot(filter(pvis_ktype, prop.flower.cover >= 0.05), aes(k.type.ss, fill = visits.bin)) +
-  geom_bar()
-```
-
-### 2.2.2 brms models
-```{r}
-# Zero- one-inflated beta regression
-brm_pvis01 <- brm(prop.visits ~ prop.flower.cover0 + 
-                    (1 | k.type.ss) + 
-                    (1 | site) + 
-                    (1 | bb.sp),
+### I think the structure of my data necessitates a complex model design.
+My observations of prop.visits ~ pbl.w.class:k.type.ss are repeated within BB species, and therefore not independent.
+Moreover, my observations are repeated within sites, adding another layer of non-independence.
+I think this means I need to include varying intercept terms for both BB sp and site.
+### I did some visualization, and it is also clear that the slope of prop.visits ~ prop.flower.cover varies strongly with pbl.w.class:ktype. This is very informative, and it makes good sense. 
+```{r}
+brm_pvis62 <- brm(mean.prop.visits ~ mean.prop.flower.cover + 
+                    (1 + mean.prop.flower.cover | k.type.ss:pbl.w.class2) + 
+                    (1 + mean.prop.flower.cover | k.type.ss:bb.sp),
                   family = "zero_one_inflated_beta",
-                  file = "./output/brm_pvis01",
+                  file = "../output/brm_pvis62",
+                  #file = "brm_pvis61",
+                  chains = 4,
+                  cores = 4,
+                  iter = 2000,
+                  data = pvis_ktype_sitepool2)
+bayes_R2(brm_pvis62)
+
+brm_pvis61 <- brm(mean.prop.visits ~ mean.prop.flower.cover + 
+                    (1 + mean.prop.flower.cover | k.type.ss:pbl.w.class2) + 
+                    (1 + mean.prop.flower.cover | k.type.ss:bb.sp),
+                  family = "zero_one_inflated_beta",
+                  file = "../output/brm_pvis61",
+                  #file = "brm_pvis61",
+                  chains = 4,
+                  cores = 4,
+                  iter = 2000,
+                  data = pvis_ktype_sitepool)
+
+summary(brm_pvis61)
+prior_summary(brm_pvis61)
+parnames(brm_pvis61)
+mcmc_plot(brm_pvis61, pars = c("^r_k"))
+mcmc_plot(brm_pvis61, pars = c("^r_s"))
+mcmc_plot(brm_pvis61)
+conditional_effects(brm_pvis61)
+pp_check(brm_pvis61)
+bayes_R2(brm_pvis61)
+mcmc_plot(brm_pvis50, pars = c("r_k.type.ss:pbl.w.class2[0_1,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[0_2,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[0_3,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[1_1,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[1_2,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[1_3,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[2_1,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[2_2,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[2_3,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[3_1,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[3_2,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[3_3,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[4_1,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[4_2,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[4_3,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[5_1,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[5_2,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[5_3,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[6_1,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[6_2,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[6_3,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[7_1,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[7_2,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[7_3,prop.flower.cover0]"),
+          fixed = TRUE)
+
+mcmc_plot(brm_pvis50, pars = c("r_k.type.ss:pbl.w.class2[0_1,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[0_2,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[0_3,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[1_1,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[1_2,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[1_3,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[2_1,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[2_2,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[2_3,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[3_1,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[3_2,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[3_3,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[4_1,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[4_2,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[4_3,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[5_1,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[5_2,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[5_3,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[6_1,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[6_2,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[6_3,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[7_1,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[7_2,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[7_3,Intercept]"),
+          fixed = TRUE)
+
+brm_pvis60 <- brm(prop.visits ~ prop.flower.cover0 + 
+                    (1 + prop.flower.cover0 | k.type.ss:pbl.w.class2) + 
+                    (1 + prop.flower.cover0 | k.type.ss:site) +
+                    (1 + prop.flower.cover0 | k.type.ss:bb.sp),
+                  family = "zero_one_inflated_beta",
+                  file = "../output/brm_pvis60",
+                  #file = "brm_pvis60",
                   chains = 4,
@@
                   data = pvis_ktype)
-
-# Poisson models of raw visit counts
-brm_pvis00 <- brm(visits ~ prop.flower.cover0 + 
-                    bb.sp.abund0 + 
-                    (1 | k.type.ss) + 
-                    (1 | site) + 
-                    (1 | bb.sp),
-                  #control = list(adapt_delta = 0.99),
-                  family = "poisson",
-                  file = "../output/brm_pvis00",
+bayes_R2(brm_pvis60)
+conditional_effects(brm_pvis60)
+mcmc_plot(brm_pvis60, pars = c("^b_"))
+mcmc_plot(brm_pvis60, pars = c("r_k.type.ss:pbl.w.class2[0_1,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[0_2,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[0_3,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[1_1,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[1_2,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[1_3,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[2_1,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[2_2,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[2_3,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[3_1,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[3_2,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[3_3,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[4_1,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[4_2,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[4_3,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[5_1,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[5_2,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[5_3,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[6_1,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[6_2,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[6_3,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[7_1,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[7_2,Intercept]",
+                               "r_k.type.ss:pbl.w.class2[7_3,Intercept]"),
+          fixed = TRUE)
+mcmc_plot(brm_pvis60, pars = c("r_k.type.ss:pbl.w.class2[0_1,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[0_2,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[0_3,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[1_1,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[1_2,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[1_3,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[2_1,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[2_2,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[2_3,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[3_1,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[3_2,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[3_3,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[4_1,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[4_2,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[4_3,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[5_1,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[5_2,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[5_3,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[6_1,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[6_2,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[6_3,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[7_1,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[7_2,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[7_3,prop.flower.cover0]"),
+          fixed = TRUE)
+
+brm_pvis50 <- brm(prop.visits ~ prop.flower.cover0 + 
+                    (1 + prop.flower.cover0 | k.type.ss:pbl.w.class2) + 
+                    (1 | k.type.ss:site) +
+                    (1 | k.type.ss:bb.sp),
+                  family = "zero_one_inflated_beta",
+                  file = "../output/brm_pvis50",
+                  #file = "brm_pvis50",
                   chains = 4,
-                  #cores = 4,
+                  cores = 4,
                   iter = 2000,
@@
 
-summary(brm_pvis00)
-prior_summary(brm_pvis00)
-parnames(brm_pvis00)
-mcmc_plot(brm_pvis00, pars = c("^r_k"))
-mcmc_plot(brm_pvis00, pars = c("^r_s"))
-mcmc_plot(brm_pvis00, pars = c("^b_"))
-mcmc_plot(brm_pvis00)
-conditional_effects(brm_pvis00)
-pp_check(brm_pvis00)
-bayes_R2(brm_pvis00)
-
-# Perhaps I should add the bb.sp varying intercept effect; otherwise, I think I could be accused of pseudo replication.
-brm_pvis11 <- brm(visits ~ prop.flower.cover0 + 
-                    bb.sp.abund0 + 
-                    (1 | k.type.ss:pbl.w.class2) + 
-                    (1 | site) +
-                    (1 | bb.sp),
-                  family = "poisson",
-                  file = "../output/brm_pvis11",
+summary(brm_pvis50)
+prior_summary(brm_pvis50)
+get_variables(brm_pvis50)
+mcmc_plot(brm_pvis50, pars = c("r_k.type.ss:pbl.w.class2[2_1,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[2_2,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[2_3,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[3_1,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[3_2,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[3_3,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[6_1,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[6_2,prop.flower.cover0]",
+                               "r_k.type.ss:pbl.w.class2[6_3,prop.flower.cover0]"),
+          fixed = TRUE)
+
+mcmc_plot(brm_pvis50, pars = c("r_k.type.ss:bb.sp[3_gers,Intercept]",
+                               "r_k.type.ss:bb.sp[3_hort,Intercept]",
+                               "r_k.type.ss:bb.sp[3_hypn,Intercept]",
+                               "r_k.type.ss:bb.sp[3_humi,Intercept]",
+                               "r_k.type.ss:bb.sp[3_jone,Intercept]",
+                               "r_k.type.ss:bb.sp[3_lapi,Intercept]",
+                               "r_k.type.ss:bb.sp[3_mend,Intercept]",
+                               "r_k.type.ss:bb.sp[3_mont,Intercept]",
+                               "r_k.type.ss:bb.sp[3_muci,Intercept]",
+                               "r_k.type.ss:bb.sp[3_pasc,Intercept]",
+                               "r_k.type.ss:bb.sp[3_prat,Intercept]",
+                               "r_k.type.ss:bb.sp[3_psit,Intercept]",
+                               "r_k.type.ss:bb.sp[3_pyre,Intercept]",
+                               "r_k.type.ss:bb.sp[3_soro,Intercept]",
+                               "r_k.type.ss:bb.sp[3_telu,Intercept]",
+                               "r_k.type.ss:bb.sp[3_wurf,Intercept]"),
+          fixed = TRUE)
+
+mcmc_plot(brm_pvis50, pars = c("r_k.type.ss:bb.sp[6_gers,Intercept]",
+                               "r_k.type.ss:bb.sp[6_hort,Intercept]",
+                               "r_k.type.ss:bb.sp[6_hypn,Intercept]",
+                               "r_k.type.ss:bb.sp[6_humi,Intercept]",
+                               "r_k.type.ss:bb.sp[6_jone,Intercept]",
+                               "r_k.type.ss:bb.sp[6_lapi,Intercept]",
+                               "r_k.type.ss:bb.sp[6_mend,Intercept]",
+                               "r_k.type.ss:bb.sp[6_mont,Intercept]",
+                               "r_k.type.ss:bb.sp[6_muci,Intercept]",
+                               "r_k.type.ss:bb.sp[6_pasc,Intercept]",
+                               "r_k.type.ss:bb.sp[6_prat,Intercept]",
+                               "r_k.type.ss:bb.sp[6_psit,Intercept]",
+                               "r_k.type.ss:bb.sp[6_pyre,Intercept]",
+                               "r_k.type.ss:bb.sp[6_soro,Intercept]",
+                               "r_k.type.ss:bb.sp[6_telu,Intercept]",
+                               "r_k.type.ss:bb.sp[6_wurf,Intercept]"),
+          fixed = TRUE)
+
+mcmc_plot(brm_pvis50, pars = c("^r_k.type.ss:site"))
+mcmc_plot(brm_pvis50)
+conditional_effects(brm_pvis50)
+pp_check(brm_pvis50)
+bayes_R2(brm_pvis50)
+
+## BB spp. pooled, only interested in comparing ktypes here
+brm_pvis12 <- brm(prop.visits ~ prop.flower.cover0 + 
+                    (1 | k.type.ss) + 
+                    (1 | site),
+                  family = "zero_one_inflated_beta",
+                  file = "../output/brm_pvis12",
+                  #file = "brm_pvis12",
                   chains = 4,
-                  #cores = 4,
+                  cores = 4,
                   iter = 2000,
-                  data = pvis_ktype)
-
-summary(brm_pvis11)
-prior_summary(brm_pvis11)
-parnames(brm_pvis11)
-mcmc_plot(brm_pvis11, pars = c("^r_k"))
-mcmc_plot(brm_pvis11, pars = c("^r_b"))
-mcmc_plot(brm_pvis11, pars = c("^b_"))
-mcmc_plot(brm_pvis11)
-conditional_effects(brm_pvis11)
-pp_check(brm_pvis11)
-bayes_R2(brm_pvis11)
+                  data = pvis_ktype_pool)
+
+summary(brm_pvis12)
+prior_summary(brm_pvis12)
+parnames(brm_pvis12)
+mcmc_plot(brm_pvis12, pars = c("^r_k"))
+mcmc_plot(brm_pvis12, pars = c("^r_s"))
+mcmc_plot(brm_pvis12)
+conditional_effects(brm_pvis12)
+pp_check(brm_pvis12)
+bayes_R2(brm_pvis12)
+
+```
+
+```{r}
+# mgcv time
+gam1 <- bam(prop.visits ~ prop.flower.cover0 +
+              te(yday, elev.mean, by = k.type.ss, bs = c("gp", "tp"), k = c(8, 8)),
+            family = "betar",
+            method = "fREML",
+            discrete = TRUE,
+            data = pvis_ktype_pool) %>% getViz()
+
+summary(gam1)
+check.gamViz(gam1)
+plot.gamViz(gam1)
+
+gam2 <- bam(visits ~ te(yday, elev.mean, by = k.type.ss, bs = c("gp", "tp"), k = c(8, 8)),
+            family = "quasipoisson",
+            method = "fREML",
+            discrete = TRUE,
+            data = pvis_ktype_pool) %>% getViz()
+
+summary(gam2)
+check.gamViz(gam2)
+plot.gamViz(gam2)
+
 ```
@@
 ## Interaction model
-brm_pvis11_plotframe <- brm_pvis11 %>%
+brm_pvi50_plotframe <- brm_pvis50 %>%
   spread_draws(b_Intercept, `r_k.type.ss:pbl.w.class2`[condition, term]) %>% # the back ticks are necessary to escape the :
@@
   mutate(condition_mean = b_Intercept + r_ktype.pbl) %>%
+  mutate(condition_mean = inv_logit_scaled(condition_mean)) %>%
   separate(condition, c("ktype", "pbl.class"), sep = "_", remove = FALSE) %>%
-  mutate(pbl.class = case_when(
-    pbl.class == "1" ~ "short",
-    pbl.class == "2" ~ "med",
-    pbl.class == "3" ~ "long")
-    ) %>%
   mutate(ktype = case_when(
@@
 
-ggplot(brm_pvis11_plotframe, aes(condition_mean, pbl.class, color = pbl.class, fill = pbl.class)) +
+ggplot(brm_pvi50_plotframe, aes(condition_mean, pbl.class, color = pbl.class, fill = pbl.class)) +
   #stat_pointinterval() +
@@
 ### 2.2.3 The lability of bias: How does the foraging bias of each species with respect to each morphotype vary through elevation and time?
+
+#### We need to do some kind of tensor product gam on this jawn!
+```{r}
+ggplot(pvis_ktype_pool, aes(yday, elev.mean, color = prop.visits)) +
+  geom_point() +
+  facet_wrap(~k.type.ss)
+```
+
 ```{r}

### 2021-02-05 code/preference_lability.Rmd
pretty much finalized analysis, almost done ms

---
title: "visitation_bias"
author: "Doug Sponsler"
date: "12/14/2020"
output: html_document
---
# Preparations
## Packages 
```{r, warning=FALSE, include=FALSE, echo = FALSE}
library(bipartite)
library(brms)
library(bayesplot)
library(tidybayes)
library(mgcv)
library(mgcViz)
library(modelr)
library(ggdist)
library(vegan)
library(ggvegan)
library(ggrepel)
library(treemapify)
library(patchwork)
library(tidymodels)
library(tidyverse)
library(lubridate)

rstan::rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

## Load data
```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
# Site data
site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class2 = factor(elev.class2, levels = c("low", "mid", "high")),
         site = factor(site)) # turn elav.class into an ordered factor and site into a regular factor

# BB traits
bb_traits <- read_csv("../data/processed_data/bb_traits.csv")

ggplot(bb_traits, aes(pbl.w, pbl.w.durieaux,
                      color = bb.sp, pch = factor(pbl.w.class2))) +
  geom_point()

# FL traits
## family-level taxonomy
fl_tax <- read_csv("../data/processed_data/floral_tax.csv")

## Kugler morphotypes (accessed through Bioflor)
fl_ktype <- read_csv("../data/processed_data/floral_k_type.csv")

## Flower colors (accessed through BioFlor)
fl_color <- read_csv("../data/processed_data/floral_color.csv")

## diameter and corolla depth data from Gita Benadi
fl_morph <- read_csv("../data/processed_data/flower_morphology_GitaBenadi.csv") %>% 
  dplyr::select(SampleID, plant.family = Family, plant.sp = Species, 
                Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam), LN = mean(LN))

## verbal description of Kugler morphotypes
fl_ktype_key <- read_delim("../data/processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  dplyr::select(k.type = 1, description = 2)

## join all floral trait tables
fl_traits <- fl_tax %>%
  full_join(fl_ktype) %>%
  full_join(fl_color) %>%
  full_join(fl_morph) %>%
  dplyr::select(plant.sp, plant.genus, plant.family, k.type, k.type.s, k.type.ss, color, diam, LN)
```

## Align survey and visitation data with pseudodates
```{r}
# Site-dates where BB visitation was recorded (excluding dates where only males were recorded)
visitation_samples <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>%
  select(site, date) %>%
  distinct()

# Site dates where floral surveying was conducted
survey_samples <- read_csv("../data/processed_data/floral_survey.csv") %>%
  select(site, date) %>%
  distinct()

# Site-dates where visitation sampling and floral sampling coincided
shared_dates <- inner_join(visitation_samples, survey_samples)

# Site-dates where visitation was recorded WITHOUT corresponding survey data
visitation_only <- anti_join(visitation_samples, survey_samples) %>%
  mutate(date.v = date) %>%
  select(-date)

# This creates a pseudo date column for unmatched visitation site-dates that will match each visitation date
# to its nearest survey data. For site t4, there is a 2-year mismatch; we will drop these site-dates later.
pseudodates <- full_join(visitation_only, survey_samples) %>%
  mutate(mismatch = abs(date.v - date)) %>%
  group_by(site, date.v) %>%
  filter(mismatch == min(mismatch)) %>%
  rename(date = date.v, pseudo.date = date)
```

## Generate a data frame that represents visitation and survey abundance
```{r}
# BB visitation data
visitation <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>% # we will ignore males 
  group_by(site, date, bb.sp, plant.sp) %>%
  summarize(visits = n()) %>% # visits per BB-FL pair
  ungroup() %>%
  left_join(pseudodates) %>% # Now we will replace unmatched dates with their pseudo dates
  replace_na(replace = list(mismatch = 0)) %>% # set mismatch to zero for perfectly aligned dates
  filter(mismatch <= 7) %>% # this gets rid of irreconcilable dates (like the t4 dates mentioned above)
  mutate(date2 = if_else(is.na(pseudo.date), date, pseudo.date)) %>% # replace date with pseudodate for all site-dates lacking perfect alignments
  select(-c(date, pseudo.date, mismatch)) %>%
  rename(date = date2) %>%
  group_by(site, date) %>%
  complete(bb.sp, plant.sp, fill = list(visits = 0)) # generate all possible BB-FL combos for each site date and fill in zeros for visits

# Floral cover data
flower_cover <- read_csv("../data/processed_data/floral_survey.csv") %>%
  filter(flower.cover > 0) %>% # ignore plants that were not blooming
  group_by(site, date, plant.sp) %>%
  summarize(flower.cover = sum(flower.cover))

# Get shared dates after adding pseudodates. This will let us filter out survey dates that don't
# have corresponding visitation sampling dates
shared_dates_with_pseudo <- inner_join(visitation, flower_cover, by = c("site", "date")) %>%
  select(site, date) %>%
  distinct()

# Join visitation and flower cover data

## First we need a list of all BB spp. by site and date
sp_by_site <- visitation %>% 
  ungroup() %>%
  select(site, date, bb.sp) %>%
  distinct()

## Then we need a list of unvisited plant.sp by site and date
unvisited <- flower_cover %>% 
  semi_join(shared_dates_with_pseudo) %>% # filter flower cover data to include only site-dates with visitation complement (real or pseudodates)
  full_join(visitation) %>% # join to visitation data
  filter(is.na(bb.sp)) %>% # select only the site-date-plants that were totally unvisited (and therefore have no BB spp to join to)
  select(-bb.sp) %>% # drop bb.sp variable because I just want the plants
  left_join(sp_by_site) %>% # join in the species list by site and date, thus populating the unvisited plants with all the BB spp recorded for each site-date
  replace_na(replace = list(visits = 0)) # replace NAs (created in the line above) with zero

## Now we can combine everything into a single data frame
visit_cover <- flower_cover %>% 
  semi_join(shared_dates_with_pseudo) %>% # filter flower cover data to include only site-dates with visitation complement (real or pseudodates)
  full_join(visitation) %>% # join to visitation data
  replace_na(replace = list(flower.cover = 0.01)) %>% # If a plant was visited but not surveyed, we will assign it the minimum non-zero flower cover value of 0.01.
  filter(!is.na(bb.sp)) %>% # drop the univisited plants
  bind_rows(unvisited) %>% # add them back, with BB spp and zeros
  select(site, date, bb.sp, plant.sp, visits, flower.cover) %>%
  group_by(site, date, bb.sp, plant.sp, flower.cover) %>% # In a few cases, the use of pseudodates causes two sampling period to converge on the same date, which creates apparent duplicate visitation data. We will handle this by averaging the visits.
  summarize(visits = mean(visits)) %>%
  left_join(fl_traits) %>% # add floral traits
  ungroup() %>%
  mutate(k.type.ss = factor(k.type.ss))

ktype_families <- visit_cover %>%
  select(plant.sp, plant.family, k.type.ss) %>%
  distinct()

ggplot(ktype_families, aes(plant.family, fill = k.type.ss)) +
  geom_bar(position = "fill") +
  theme(axis.text.x = element_text(angle = 90))

## Get total visits by BB sp and site-date; this will be a measure of BB abundance that we can use after we've calculated d'. I could, of course, work this into the biaseval function, but I'll put that off for now.
bb_abundance <- visit_cover %>%
  group_by(site, date, bb.sp) %>%
  summarize(bb.sp.abund = sum(visits)) %>%
  group_by(site, date) %>%
  mutate(bb.sp.prop.abund = bb.sp.abund/sum(bb.sp.abund))

fl_abundance <- visit_cover %>%
  group_by(site, date) %>%
  summarize(total.flower.cover = sum(flower.cover))
```

# 1. Overall bias
## 1.1 Calculate visitation probability per plant morphotype. Instead of keeping the plant species resolved, we will pool visitation by morphotype. The problem with not pooling by morphotype is that more species-rich morphotypes will be penalized simply because their visitation is diluted across more species.
```{r}
prop_visits_ktype <- visit_cover %>%
  group_by(site, date, bb.sp, k.type.ss) %>%
  summarize(flower.cover = sum(flower.cover), # get raw flower cover and visits by site*date*bb.sp*plant.sp
            visits = sum(visits)) %>%
  group_by(site, date) %>%
  mutate(prop.flower.cover = flower.cover/sum(flower.cover))%>% # binary visits
  group_by(site, date, bb.sp) %>%
  mutate(prop.visits = visits/sum(visits),
         bb.sp.abund = sum(visits)) %>% # abundance of each BB sp. per site-date
  ungroup() %>%
  left_join(bb_traits) %>% # add BB traits
  mutate(ktype = factor(k.type.ss), # recode variables
         tongue = factor(pbl.w.class2),
         visits = as.integer(visits)) %>% # so we can run a Poisson model
  left_join(site_data) %>% # add site data
  ungroup() %>%
  mutate(yday = yday(date),
         elev.mean0 = scale(elev.mean),
         yday0 = scale(yday),
         prop.flower.cover0 = scale(prop.flower.cover),
         perc.flower.cover = prop.flower.cover*100,
         bb.sp.abund0 = scale(bb.sp.abund)) %>%
  select(site, elev.mean, date, yday, bb.sp, ktype, visits, 
         prop.visits, perc.flower.cover, bb.sp.abund, tongue,
         elev.mean0, yday0, prop.flower.cover0, bb.sp.abund0) %>%
  mutate(visits.bin = if_else(visits > 0, 1, 0))

write_csv(prop_visits_ktype, "../hpc_packets/prop_visits_ktype.csv")

prop_visits_ktype_pooled <- visit_cover %>%
  group_by(site, date, k.type.ss) %>%
  summarize(flower.cover = sum(flower.cover), # get raw flower cover and visits by site*date*bb.sp*plant.sp
            visits = sum(visits)) %>%
  group_by(site, date) %>%
  mutate(prop.flower.cover = flower.cover/sum(flower.cover),
         prop.visits = visits/sum(visits),
         bb.abund = sum(visits),
         visits = as.integer(visits)) %>%
  mutate(ktype = factor(k.type.ss)) %>%
  left_join(site_data) %>% # add site data
  ungroup() %>%
  mutate(yday = yday(date),
         elev.mean0 = scale(elev.mean),
         yday0 = scale(yday),
         perc.flower.cover = prop.flower.cover*100,
         prop.flower.cover0 = scale(prop.flower.cover),
         bb.abund0 = scale(bb.abund)) %>%
  select(site, elev.mean, date, yday, ktype, visits, 
         prop.visits, bb.abund, perc.flower.cover,
         elev.mean0, yday0, prop.flower.cover0, bb.abund0) %>%
  mutate(visits.bin = if_else(visits > 0, 1, 0))

write_csv(prop_visits_ktype_pooled, "../hpc_packets/prop_visits_ktype_pool.csv")

```

## 1.2 Binomial multi-level regression models
```{r}
## Exploring the differential bias of bumble bee species and tongue length classes towards floral morphotypes
### Bernoulli with binarized visitation
brm_pvis00 <- brm(as.integer(visits.bin) ~
                    bb.sp.abund*perc.flower.cover + # fixed term: because more abundant flowers will tend to be more visited,and more abundance bees visit more
                    (1 + perc.flower.cover | ktype:bb.sp) + # a random intercept and random slope for ktype:species; I think I need to model this before I can make conclusions about ktype:pbl.class
                    (1 + perc.flower.cover | ktype:tongue) + # a random intercept and slope for ktype:tongue
                    (1 | site), # a random intercept and slope for ktype:site
                  family = "bernoulli", 
                  #file = "../output/brm_pvis_ktype_00",  # for running locally
                  file = "brm_pvis_ktype_00", # for running on the HPC
                  chains = 4,
                  cores = 12,
                  iter = 2000,
                  data = filter(prop_visits_ktype, !ktype %in% c("9", "10")))

brm_pvis01 <- brm(as.integer(visits.bin) ~
                    bb.abund*perc.flower.cover + # fixed term: because more abundant flowers will tend to be more visited,and more abundance bees visit more
                    (1 + perc.flower.cover | ktype) + # a random intercept and slope for ktype:site
                    (1 | site), 
                  family = "bernoulli", 
                  file = "../output/brm_pvis_ktype_01",  # for running locally
                  #file = "brm_pvis_ktype_01", # for running on the HPC
                  chains = 4,
                  cores = 12,
                  iter = 2000,
                  data = filter(prop_visits_ktype_pooled, !ktype %in% c("9", "10")))

# If I want to make inference about species visitation rates, I don't need the tongue-length class effect; I just need the site effect. But, when I want to make inference about the tongue length classes, I do need the species-level effect.
brm_pvis02 <- brm(as.integer(visits.bin) ~
                    bb.sp.abund*perc.flower.cover + # fixed term: because more abundant flowers will tend to be more visited,and more abundance bees visit more
                    (1 + perc.flower.cover | ktype:bb.sp) + # a random intercept and random slope for ktype:species; I think I need to model this before I can make conclusions about ktype:pbl.class
                    (1 | site), # a random intercept and slope for ktype:site
                  family = "bernoulli", 
                  #file = "../output/brm_pvis_ktype_02",  # for running locally
                  file = "brm_pvis_ktype_02", # for running on the HPC
                  chains = 4,
                  cores = 12,
                  iter = 2000,
                  data = filter(prop_visits_ktype, !ktype %in% c("9", "10")))
```
## 1.3 Evaluate models
```{r}
# Summary
summary(brm_pvis00)
summary(brm_pvis01)
summary(brm_pvis02)

# Posterior predictive checks
pp_check(brm_pvis00)
pp_check(brm_pvis01)
pp_check(brm_pvis02)

# Bayes R2
bayes_R2(brm_pvis00)
bayes_R2(brm_pvis01)
bayes_R2(brm_pvis02)

```

## 1.4 Preliminary visualization
```{r}
get_variables(brm_pvis00)
conditional_effects(brm_pvis00)
mcmc_plot(brm_pvis00, pars = c("^r_ktype:tongue.*Intercept]$"))
mcmc_plot(brm_pvis00, pars = c("^r_ktype:tongue.*perc.flower.cover]$"))
mcmc_plot(brm_pvis00, pars = c("^r_ktype:bb.sp.*Intercept]$"))
mcmc_plot(brm_pvis00, pars = c("^r_ktype:bb.sp.*perc.flower.cover]$"))
mcmc_plot(brm_pvis00, pars = c("^r_ktype:site.*Intercept]$"))
mcmc_plot(brm_pvis00, pars = c("^r_ktype:site.*perc.flower.cover]$"))

get_variables(brm_pvis01)
conditional_effects(brm_pvis01)
mcmc_plot(brm_pvis01, pars = c("^r_ktype"))
mcmc_plot(brm_pvis01, pars = c("^r_ktype:tongue.*perc.flower.cover]$"))
mcmc_plot(brm_pvis01, pars = c("^r_ktype:bb.sp.*Intercept]$"))
mcmc_plot(brm_pvis01, pars = c("^r_ktype:bb.sp.*perc.flower.cover]$"))
mcmc_plot(brm_pvis01, pars = c("^r_ktype:site.*Intercept]$"))
mcmc_plot(brm_pvis01, pars = c("^r_ktype:site.*perc.flower.cover]$"))

get_variables(brm_pvis02)
conditional_effects(brm_pvis02)
mcmc_plot(brm_pvis02, pars = c("^r_ktype"))
mcmc_plot(brm_pvis02, pars = c("^r_ktype:bb.sp.*Intercept]$"))
mcmc_plot(brm_pvis02, pars = c("^r_ktype:bb.sp.*perc.flower.cover]$"))
mcmc_plot(brm_pvis02, pars = c("^r_ktype:site.*Intercept]$"))
mcmc_plot(brm_pvis02, pars = c("^r_ktype:site.*perc.flower.cover]$"))
```

## 1.5 Final visualization
Note that the intercept is on the log scale, so if you want to transform it back to predicted # visits, you have to exponentiate it. I think. Here is a helplful conversation involving the esteemed Ben Bolker:
### 1.5.1 morphotype:tongue intercept
```{r}
brm_pvis00_plotframe1 <- brm_pvis00 %>%
  spread_draws(b_Intercept, `r_ktype:tongue`[condition, term]) %>% # the back ticks are necessary to escape the :
  rename(r_ktype.tongue = `r_ktype:tongue`) %>% # get rid of the problematic :
  filter(term == "Intercept") %>% # drop slope terms for now
  mutate(condition_mean = b_Intercept + r_ktype.tongue) %>% # get conditional mean
  group_by(condition) %>% # group by tongue:morpotype pair
  point_interval(condition_mean, .width = c(0.95, 0.66)) %>% # calculate credible intervals
  mutate(condition_mean_trans = exp(condition_mean), # exponentiating the estimates gets us back to the response scale (negbinomial uses log link function)
         .lower_trans = exp(.lower),
         .upper_trans = exp(.upper)) %>% 
separate(condition, c("ktype", "tongue"), sep = "_", remove = FALSE) %>% # separate condition term into ktype and tongue
  mutate(ktype = case_when( # rename
    ktype == "0" ~ "NA (0)",
    ktype == "1" ~ "Disc (1)",
    ktype == "2" ~ "Funnel (2)",
    ktype == "3" ~ "Bell (3)",
    ktype == "4" ~ "Stalk-disc (4)",
    ktype == "5" ~ "Lip (5)",
    ktype == "6" ~ "Flag (6)",
    ktype == "7" ~ "Head (7)")
  ) %>%
  mutate(ktype = factor(ktype, levels = c("Disc (1)", "Funnel (2)", "Bell (3)", # relevel
                                          "Stalk-disc (4)","Lip (5)", "Flag (6)", 
                                          "Head (7)", "NA (0)")))

ggplot(brm_pvis00_plotframe1, aes(condition_mean_trans, tongue, 
                                 xmin = .lower_trans, xmax = .upper_trans,
                                 color = tongue, fill = tongue)) +
  geom_pointinterval() +
  facet_wrap(~ktype, scales = "free") +
  theme_light() +
  labs(color = "Tongue length", x = "Visitation probability") +
  guides(fill = FALSE) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

ggsave(file = "../output/brm_pvis00_plot1.png")
ggsave(file = "../output/brm_pvis00_plot1.pdf")

```

### 1.5.2 morphotype:bb.sp intercept
```{r}
brm_pvis02_plotframe2 <- brm_pvis02 %>%
  spread_draws(b_Intercept, `r_ktype:bb.sp`[condition, term]) %>% # the back ticks are necessary to escape the :
  rename(r_ktype.bb.sp = `r_ktype:bb.sp`) %>% # get rid of the problematic :
  filter(term == "Intercept") %>% # drop slope terms for now
  mutate(condition_mean = b_Intercept + r_ktype.bb.sp) %>% # get conditional mean
  group_by(condition) %>% # group by bb.sp:morpotype pair
  point_interval(condition_mean, .width = c(0.95, 0.66)) %>% # calculate credible intervals
  mutate(condition_mean_trans = inv_logit_scaled(condition_mean), # exponentiating the estimates gets us back to the response scale (negbinomial uses log link function)
         .lower_trans = inv_logit_scaled(.lower),
         .upper_trans = inv_logit_scaled(.upper)) %>% 
separate(condition, c("ktype", "bb.sp"), sep = "_", remove = FALSE) %>% # separate condition term into ktype and bb.sp
  mutate(ktype = case_when( # rename
    ktype == "0" ~ "NA (0)",
    ktype == "1" ~ "Disc (1)",
    ktype == "2" ~ "Funnel (2)",
    ktype == "3" ~ "Bell (3)",
    ktype == "4" ~ "Stalk-disc (4)",
    ktype == "5" ~ "Lip (5)",
    ktype == "6" ~ "Flag (6)",
    ktype == "7" ~ "Head (7)")
  ) %>%
  mutate(ktype = factor(ktype, levels = c("Disc (1)", "Funnel (2)", "Bell (3)", # relevel
                                          "Stalk-disc (4)","Lip (5)", "Flag (6)", 
                                          "Head (7)", "NA (0)"))) %>%
  left_join(bb_traits)

ggplot(brm_pvis02_plotframe2, aes(condition_mean_trans, bb.sp, 
                                 xmin = .lower_trans, xmax = .upper_trans,
                                 color = pbl.w.class2, fill = pbl.w.class2)) +
  geom_pointinterval() +
  facet_wrap(~ktype, scales = "free") +
  theme_light() +
  labs(color = "Tongue length", x = "Visitation probability", y = "Morphotype") +
  guides(fill = FALSE)

ggsave(file = "../output/brm_pvis02_plot1.png", width = 6.5, height = 6.5)
```

### 1.5.3 morphotype:tongue slope
```{r}
brm_pvis00_plotframe3 <- brm_pvis00 %>%
  spread_draws(b_perc.flower.cover, `r_ktype:tongue`[condition, term]) %>% # the back ticks are necessary to escape the :
  rename(r_ktype.tongue = `r_ktype:tongue`) %>%
  filter(term == "perc.flower.cover") %>%
  mutate(condition_slope = b_perc.flower.cover + r_ktype.tongue) %>%
  separate(condition, c("ktype", "tongue"), sep = "_", remove = FALSE) %>%
  mutate(ktype = case_when(
    ktype == "0" ~ "NA (0)",
    ktype == "1" ~ "Disc (1)",
    ktype == "2" ~ "Funnel (2)",
    ktype == "3" ~ "Bell (3)",
    ktype == "4" ~ "Stalk-disc (4)",
    ktype == "5" ~ "Lip (5)",
    ktype == "6" ~ "Flag (6)",
    ktype == "7" ~ "Head (7)")
  ) %>%
  mutate(ktype = factor(ktype,
                        levels = c("Disc (1)", "Funnel (2)", "Bell (3)", "Stalk-disc (4)",
                                   "Lip (5)", "Flag (6)", "Head (7)", "NA (0)")))

ggplot(brm_pvis00_plotframe3, aes(condition_slope, tongue, color = tongue)) +
  stat_pointinterval() +
  facet_grid(ktype~.) +
  theme_light() +
  labs(color = "Tongue length", x = "Slope of logit-likelihood") +
  guides(fill = FALSE) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

ggsave(file = "../output/brm_pvis00_plot3.png", width = 7, height = 5.5)
```

### 1.5.3 morphotype intercept
```{r}
brm_pvis01_plotframe1 <- brm_pvis01 %>%
  spread_draws(b_Intercept, r_ktype[condition, term]) %>% # the back ticks are necessary to escape the :
  filter(term == "Intercept") %>% # drop slope terms for now
  mutate(condition_mean = b_Intercept + r_ktype) %>% # get conditional mean
  group_by(condition) %>% # group by tongue:morpotype pair
  point_interval(condition_mean, .width = c(0.95, 0.66)) %>% # calculate credible intervals
  mutate(condition_mean_trans = inv_logit_scaled(condition_mean), # exponentiating the estimates gets us back to the response scale (negbinomial uses log link function)
         .lower_trans = inv_logit_scaled(.lower),
         .upper_trans = inv_logit_scaled(.upper)) %>% 
  mutate(ktype = case_when( # rename
    condition == "0" ~ "NA (0)",
    condition == "1" ~ "Disc (1)",
    condition == "2" ~ "Funnel (2)",
    condition == "3" ~ "Bell (3)",
    condition == "4" ~ "Stalk-disc (4)",
    condition == "5" ~ "Lip (5)",
    condition == "6" ~ "Flag (6)",
    condition == "7" ~ "Head (7)")
  ) %>%
  mutate(ktype = factor(ktype, levels = c("Disc (1)", "Funnel (2)", "Bell (3)", # relevel
                                          "Stalk-disc (4)","Lip (5)", "Flag (6)", 
                                          "Head (7)", "NA (0)")))

ggplot(brm_pvis01_plotframe1, aes(condition_mean_trans, ktype, 
                                 xmin = .lower_trans, xmax = .upper_trans)) +
  geom_pointinterval() +
  theme_light() +
  labs(x = "Visitation probability", y= "Morphotype") +
  guides(fill = FALSE)

ggsave(file = "../output/brm_pvis01_plot1.png", width = 5, height = 7)
ggsave(file = "../output/brm_pvis01_plot1.pdf", width = 5, height = 7)
```

### 1.5.3 morphotype slope
```{r}
brm_pvis01_plotframe2 <- brm_pvis01 %>%
  spread_draws(b_perc.flower.cover, r_ktype[condition, term]) %>% # the back ticks are necessary to escape the :
  filter(term == "perc.flower.cover") %>%
  mutate(condition_slope = b_perc.flower.cover + r_ktype) %>%
  mutate(ktype = case_when(
    condition == "0" ~ "NA (0)",
    condition == "1" ~ "Disc (1)",
    condition == "2" ~ "Funnel (2)",
    condition == "3" ~ "Bell (3)",
    condition == "4" ~ "Stalk-disc (4)",
    condition == "5" ~ "Lip (5)",
    condition == "6" ~ "Flag (6)",
    condition == "7" ~ "Head (7)")
  ) %>%
  mutate(ktype = factor(ktype,
                        levels = c("Disc (1)", "Funnel (2)", "Bell (3)", "Stalk-disc (4)",
                                   "Lip (5)", "Flag (6)", "Head (7)", "NA (0)")))

ggplot(brm_pvis01_plotframe2, aes(condition_slope, ktype)) +
  stat_pointinterval() +
  theme_light() +
  labs(color = "Tongue length", x = "Slope of visitation probability ~ percent flower cover") +
  guides(fill = FALSE)

ggsave(file = "../output/brm_pvis01_plot2.png", width = 7, height = 5.5)
```

## 1.6 Model 1 hypothesis tests
I'm not doing every possible comparison. Just the ones where it is possible to have some doubt. It would be stupid to compare group whose 95% CIs are miles apart.
```{r}
get_variables(brm_pvis01)

hypothesis(brm_pvis01, "r_ktype[0,Intercept] = r_ktype[1,Intercept]", class = "") # this is lower than its closest neighbor; no need to make any other comparisons

hypothesis(brm_pvis01, "r_ktype[1,Intercept] = r_ktype[4,Intercept]", class = "") 
hypothesis(brm_pvis01, "r_ktype[2,Intercept] = r_ktype[4,Intercept]", class = "")

hypothesis(brm_pvis01, "r_ktype[2,Intercept] = r_ktype[1,Intercept]", class = "")
hypothesis(brm_pvis01, "r_ktype[1,Intercept] = r_ktype[3,Intercept]", class = "") 
hypothesis(brm_pvis01, "r_ktype[2,Intercept] = r_ktype[3,Intercept]", class = "")
hypothesis(brm_pvis01, "r_ktype[4,Intercept] = r_ktype[3,Intercept]", class = "")

hypothesis(brm_pvis01, "r_ktype[5,Intercept] = r_ktype[6,Intercept]", class = "") 
hypothesis(brm_pvis01, "r_ktype[5,Intercept] = r_ktype[7,Intercept]", class = "") 
```

## 1.7 Model 2 hypothesis tests
```{r}
get_variables(brm_pvis00)

# Disc (group  1)
hypothesis(brm_pvis00, "r_ktype:tongue[1_short,Intercept] = r_ktype:tongue[1_med,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[1_short,Intercept] = r_ktype:tongue[1_long,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[1_med,Intercept] = r_ktype:tongue[1_long,Intercept]", class = "")

# Funnel (group  2)
hypothesis(brm_pvis00, "r_ktype:tongue[2_short,Intercept] = r_ktype:tongue[2_med,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[2_short,Intercept] = r_ktype:tongue[2_long,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[2_med,Intercept] = r_ktype:tongue[2_long,Intercept]", class = "")

# Bell (group  3)
hypothesis(brm_pvis00, "r_ktype:tongue[3_short,Intercept] = r_ktype:tongue[3_med,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[3_short,Intercept] = r_ktype:tongue[3_long,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[3_med,Intercept] = r_ktype:tongue[3_long,Intercept]", class = "")

# Stalk-disc (group  4)
hypothesis(brm_pvis00, "r_ktype:tongue[4_short,Intercept] = r_ktype:tongue[4_med,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[4_short,Intercept] = r_ktype:tongue[4_long,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[4_med,Intercept] = r_ktype:tongue[4_long,Intercept]", class = "")

# Lip (group  5)
hypothesis(brm_pvis00, "r_ktype:tongue[5_short,Intercept] = r_ktype:tongue[5_med,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[5_short,Intercept] = r_ktype:tongue[5_long,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[5_med,Intercept] = r_ktype:tongue[5_long,Intercept]", class = "")

# Flag (group  6)
hypothesis(brm_pvis00, "r_ktype:tongue[6_short,Intercept] = r_ktype:tongue[6_med,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[6_short,Intercept] = r_ktype:tongue[6_long,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[6_med,Intercept] = r_ktype:tongue[6_long,Intercept]", class = "")

# Head (group  7)
hypothesis(brm_pvis00, "r_ktype:tongue[7_short,Intercept] = r_ktype:tongue[7_med,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[7_short,Intercept] = r_ktype:tongue[7_long,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[7_med,Intercept] = r_ktype:tongue[7_long,Intercept]", class = "")

# NA (group  0)
hypothesis(brm_pvis00, "r_ktype:tongue[0_short,Intercept] = r_ktype:tongue[0_med,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[0_short,Intercept] = r_ktype:tongue[0_long,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[0_med,Intercept] = r_ktype:tongue[0_long,Intercept]", class = "")
```

# 2 Ordination of dietary niche
Aggregation changes meaning. Don't forget that. When I pool visitation across dates, I am introducing the effects of relative abundance and floral phenology. Species in bloom when BB abundance is high will be over-represented, and vice-versa. And species with a long bloom time will be over-represented, and vice-versa. This isn't necessarily a problem. I just need to make sure I interpret the results the right way. When I pool data like this, I am comparing the "net" diet of species, which includes the effects of abundance and phenology in addition to the effects of floral preferences. I think that is appropriate for this stage of the analysis, but I need to be careful not to assume that what I'm doing here is the same sort of analysis as what I did above without pooling samples. 
```{r}
# Fl traits with abbreviated names
fl_traits2 <- fl_traits %>%
  left_join(read_csv("../data/processed_data/network.csv")) %>%
  select(plant.sp, plant.sp.abb, plant.genus, plant.family, k.type.ss, color) %>%
  mutate(k.type.ss = factor(k.type.ss)) %>%
  distinct()

# Totally aggregated; note I'm using the full data, not the aligned and trimmed data. No need for complementary floral survey data.
visitation2 <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>% # we will ignore males 
  filter(!bb.sp %in% c("humi")) %>% # humi is too rare
  group_by(bb.sp, plant.sp.abb) %>%
  summarize(visits = n()) %>%
  group_by(bb.sp) %>%
  mutate(prop.visits = visits/sum(visits)) %>%
  select(-visits) %>%
  pivot_wider(names_from = plant.sp.abb, values_from = prop.visits, values_fill = 0) %>%
  column_to_rownames("bb.sp")

bb_traits2 <- bb_traits %>%
  filter(bb.sp != "humi") %>%
  select(bb.sp, pbl.w.class2) %>%
  arrange(bb.sp) %>%
  mutate(pbl.w.class2.pool = case_when(
    pbl.w.class2 == "short" ~ "short",
    pbl.w.class2 == "med" ~ "med/long",
    pbl.w.class2 == "long" ~ "med/long"
  ))

# Generate ordination
pcoa_horn2 <- capscale(visitation2 ~ 1, distance = "horn")
pcoa_horn2_adonis <- adonis(visitation2 ~ pbl.w.class2, data = bb_traits2)
pcoa_horn2_betadisper <- betadisper(vegdist(visitation2, method = "horn"), 
                                    group = bb_traits2$pbl.w.class2,
                                    type = "median")
anova(pcoa_horn2_betadisper) # I think this is not significant because short tongued species are disperse on axis 2, but axis 2 does not explain nearly as much variation.
plot(pcoa_horn2_betadisper)
# I'm suprised that the betadisper test is not significant, but I think it might be driven by the fact that replication is so low for long and medium tongue species. Let's see what happens if we pool long and medium tongue species, since they overlap so much anyway.
pcoa_horn2_betadisper_pool <- betadisper(vegdist(visitation2, method = "horn"), 
                                    group = bb_traits2$pbl.w.class2.pool,
                                    type = "median")
anova(pcoa_horn2_betadisper_pool) # I think this is not significant because short tongued species are disperse on axis 2, but axis 2 does not explain nearly as much variation.
plot(pcoa_horn2_betadisper_pool)

# Get variance explained
pcoa_horn2_mds1 <- round(pcoa_horn2$CA$eig[[1]] / sum(pcoa_horn2$CA$eig)*100)
pcoa_horn2_mds2 <- round(pcoa_horn2$CA$eig[[2]] / sum(pcoa_horn2$CA$eig)*100)

# Fortify plants
pcoa_horn2.fort1 <- fortify(pcoa_horn2, axes = 1:2) %>%
  filter(Score == "species") %>%
  rename(plant.sp.abb = Label) %>%
  left_join(fl_traits2)

# Fortify bees
pcoa_horn2.fort2 <- fortify(pcoa_horn2, axes = 1:2) %>%
  filter(Score == "sites") %>%
  rename(bb.sp = Label) %>%
  left_join(bb_traits) %>%
  mutate(pbl.w.class2 = factor(pbl.w.class2))

# Generate convex hulls around tongue length classes
pcoa_horn2_hulls <- pcoa_horn2.fort2 %>%
  group_by(pbl.w.class2) %>%
  slice(chull(MDS1, MDS2))

# Plot
ggplot(pcoa_horn2.fort2, aes(MDS1, MDS2, color = pbl.w.class2)) +
  geom_polygon(data = pcoa_horn2_hulls, 
               aes(fill = pbl.w.class2, color = NULL),
               alpha = 0.25) +
  geom_text(aes(label = bb.sp), key_glyph = "rect") +
  geom_text_repel(data = filter(pcoa_horn2.fort1, abs(MDS1) >= 0.2 | abs(MDS2) >= 0.2), 
            aes(MDS1, MDS2, label = plant.sp.abb), inherit.aes = FALSE, color = "gray40") +
  geom_point(data = filter(pcoa_horn2.fort1, abs(MDS1) >= 0.2 | abs(MDS2) >= 0.2),
            aes(MDS1, MDS2, shape = k.type.ss), inherit.aes = FALSE, color = "gray40", size = 3) +
  scale_shape_discrete() +
  xlab(paste("MDS1 ", "(", pcoa_horn2_mds1, "%", ")", sep = "")) +
  ylab(paste("MDS2 ", "(", pcoa_horn2_mds2, "%", ")", sep = "")) +
  labs(fill = "Tongue length", 
       shape = "Morphotype") +
  guides(color = FALSE) +
  coord_fixed() +
  theme_light(10)

ggsave("../output/pcoa_aggregate.png")

# a quick peak at diet breadth...quick, I promise
taxon_diet_breadth <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>% # we will ignore males 
  filter(!bb.sp %in% c("humi")) %>% # humi is too rare
  group_by(site, date, bb.sp, plant.sp) %>%
  summarize(visits = n()) %>%
  group_by(site, date, bb.sp) %>%
  summarize(bb.sp.abund = sum(visits),
            tax.diet.breadth = length(plant.sp)) %>%
  mutate(log.bb.sp.abund = log(bb.sp.abund),
         prop.tax.diet.breadth = tax.diet.breadth/bb.sp.abund)

morph_diet_breadth <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>% # we will ignore males 
  filter(!bb.sp %in% c("humi")) %>% # humi is too rare
  left_join(fl_traits) %>%
  group_by(site, date, bb.sp, k.type.ss) %>%
  summarize(visits = n()) %>%
  group_by(site, date, bb.sp) %>%
  summarize(bb.sp.abund = sum(visits),
            morph.diet.breadth = length(k.type.ss)) %>%
  mutate(log.bb.sp.abund = log(bb.sp.abund),
         prop.morph.diet.breadth = morph.diet.breadth/bb.sp.abund)

diet_breadth <- full_join(taxon_diet_breadth, morph_diet_breadth)

ggplot(diet_breadth, aes(tax.diet.breadth, morph.diet.breadth, color = bb.sp)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(diet_breadth, aes(morph.diet.breadth, tax.diet.breadth, color = bb.sp)) +
  geom_point() +
  geom_smooth(method = "lm")
```

# 3. The lability of bias: How does the foraging bias of each species with respect to each morphotype vary through elevation and time?
```{r}
# Subset data to include only abundant BB spp and exclude rare morphotypes (9, 10)
gam_data <- prop_visits_ktype %>%
  filter(ktype %in% c("3", "5", "6", "7") & # focus on major ktypes
           bb.sp %in% c("telu", "prat", "pasc", "hort", "soro", "wurf")) %>% # focus on abundance bb.sp
  unite(pair, c(bb.sp, ktype), sep = "_", remove = FALSE) %>% # paste bb.sp and ktype into pair
  mutate(pair = factor(pair),
         bin.visits = if_else(visits > 0, 1, 0)) # binarize visitation

# Independent smooth model
gam_pvis_elev01 <- bam(bin.visits ~
                         s(pair, bs = "re") + # center by pair
                         bb.sp.abund*perc.flower.cover + # condition on a global bb/fl abundance term
                         perc.flower.cover*pair + # condition on a pair-specific fl abundance
                         #s(perc.flower.cover, by = pair, k = 3) +  term
                         te(yday, elev.mean, by = pair, bs = c("gp", "tp"), k = 5), # conditional effects of yday/elev by pair
                       family = "binomial", # binomial response
                       method = "fREML",
                       discrete = TRUE,
                       select =  TRUE,
                       data = filter(gam_data,
                                     bb.sp %in% c("soro", "wurf", "hort",
                                                  "telu", "pasc", "prat"))) %>%
  getViz()

#saveRDS(gam_pvis_elev01, "../output/gam_pvis_elev01.rds")
#gam_pvis_elev01 <- readRDS("../output/gam_pvis_elev01.rds") %>% getViz()

summary(gam_pvis_elev01)
check.gamViz(gam_pvis_elev01)
gam_pvis_elev01.plot <- plot(gam_pvis_elev01)
gam_pvis_elev01.plot[[1]]gam_pvis_elev01
str(gam_pvis_elev01)

hort3 <- plot(sm(gam_pvis_elev01, 2))
hort5 <- plot(sm(gam_pvis_elev01, 3))
hort6 <- plot(sm(gam_pvis_elev01, 4))
hort7 <- plot(sm(gam_pvis_elev01, 5))

pasc3 <- plot(sm(gam_pvis_elev01, 6))
pasc5 <- plot(sm(gam_pvis_elev01, 7))
pasc6 <- plot(sm(gam_pvis_elev01, 8))
pasc7 <- plot(sm(gam_pvis_elev01, 9))

prat3 <- plot(sm(gam_pvis_elev01, 10))
prat5 <- plot(sm(gam_pvis_elev01, 11))
prat6 <- plot(sm(gam_pvis_elev01, 12))
prat7 <- plot(sm(gam_pvis_elev01, 13))

soro3 <- plot(sm(gam_pvis_elev01, 14))
soro5 <- plot(sm(gam_pvis_elev01, 15))
soro6 <- plot(sm(gam_pvis_elev01, 16))
soro7 <- plot(sm(gam_pvis_elev01, 17))

telu3 <- plot(sm(gam_pvis_elev01, 18))
telu5 <- plot(sm(gam_pvis_elev01, 19))
telu6 <- plot(sm(gam_pvis_elev01, 20))
telu7 <- plot(sm(gam_pvis_elev01, 21))

wurf3 <- plot(sm(gam_pvis_elev01, 22))
wurf5 <- plot(sm(gam_pvis_elev01, 23))
wurf6 <- plot(sm(gam_pvis_elev01, 24))
wurf7 <- plot(sm(gam_pvis_elev01, 25))

pterm1 <- plot(pterm(gam_pvis_elev01, 1))
pterm2 <- plot(pterm(gam_pvis_elev01, 2))
pterm3 <- plot(pterm(gam_pvis_elev01, 3))

hort3
pasc3
prat3
soro3
telu3
wurf3

hort5
pasc5
prat5
soro5
telu5
wurf5

hort6
pasc6
prat6
soro6
telu6
wurf6

hort7
pasc7
prat7
soro7
telu7
wurf7

grid_plot <- gridPrint(hort3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
                        hort5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
                        hort6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
                        hort7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
                        pasc3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
                        pasc5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
                        pasc6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
                        pasc7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
                        prat3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
                        prat5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
                        prat6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
                        prat7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
                        soro3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
                        soro5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
                        soro6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
                        soro7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
                        telu3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
                        telu5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
                        telu6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
                        telu7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
                        wurf3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
                        wurf5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
                        wurf6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
                        wurf7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
                        ncol = 4)

ggsave(grid_plot, file = "../output/gam_plot1.pdf")

color_ramp <- gridPrint(hort3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL))
ggsave(color_ramp, file = "../output/color_ramp.pdf")

```

### 2021-02-05 code/visitation_bias.Rmd
pretty much finalized analysis, almost done ms

--- 2021-01-27 code/visitation_bias.Rmd
@@
 # Collated
+## Filtered t0 abund >= 3
 bias <- full_join(bias_plantsp, bias_ktype) %>%
@@
 
+## Unfiltered
 bias2 <- full_join(bias_plantsp, bias_ktype) %>%
@@
 
-ggplot(bias1, aes(dprime.sp)) +
-  geom_histogram()
-
 ggplot(bias2, aes(dprime.sp)) +
   geom_histogram()
+
+ggplot(filter(bias2, bb.sp.abund >= 5), aes(dprime.sp)) +
+  geom_histogram()
+
+ggplot(filter(bias2, bb.sp.abund >= 10), aes(dprime.sp)) +
+  geom_histogram()
+
+ggplot(filter(bias, bb.sp %in% c("pasc", "prat", "wurf", "soro", "psit", "hort", "telu")), aes(elev, dprime.sp)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~bb.sp)
 ```
@@
 I can always run them again if we decide it is important ot show the model selection process for the paper.
+
+#### Call models
 ```{r}
@@
 
-summary(brm_dprime.sp10)
-conditional_effects(brm_dprime.sp10)
-prior_summary(brm_dprime.sp10)
-parnames(brm_dprime.sp10)
-pp_check(brm_dprime.sp10)
-bayes_R2(brm_dprime.sp10)
-mcmc_plot(brm_dprime.sp10, pars = "^r_bb.sp")
-mcmc_plot(brm_dprime.sp10, pars = "^r_site")
-
 # Abundance threshold >= 3
@@
 
-summary(brm_dprime.sp00)
-conditional_effects(brm_dprime.sp00)
-prior_summary(brm_dprime.sp00)
-parnames(brm_dprime.sp00)
+# Abundance threshold >= 5
+brm_dprime.sp20 <- brm(dprime.sp ~ yday + bb.sp.abund + (1 | bb.sp) + (1 | site),
+                       cores = 4,
+                       iter = 2000,
+                       family = "zero_one_inflated_beta", # 
+                       file = "../output/brm_dprime.sp20",
+                       data = filter(bias, bb.sp.abund >= 5))
+
+# Abundance threshold >= 10
+brm_dprime.sp30 <- brm(dprime.sp ~ yday + bb.sp.abund + (1 | bb.sp) + (1 | site),
+                       cores = 4,
+                       iter = 2000,
+                       family = "zero_one_inflated_beta", # 
+                       file = "../output/brm_dprime.sp30",
+                       data = filter(bias, bb.sp.abund >= 10))
+```
+
+#### 1.3.3.1 Evaluate models
+They just explains so little variance... d' is too unstable at low abundance. I just don't think there is much that can be said here. 
+```{r}
+# Full data
+pp_check(brm_dprime.sp10)
+bayes_R2(brm_dprime.sp10)
+
+## QQ plot
+qres_dprime10 <- bias2 %>%
+  add_predicted_draws(brm_dprime.sp10, n = 20) %>%
+  summarise(
+    quant_residual = mean(.prediction < dprime.sp)
+  )
+
+ggplot(qres_dprime10,
+         aes(sample = quant_residual)) +
+  geom_qq(distribution = stats::qunif) +
+  geom_qq_line(distribution = stats::qunif)
+
+## Predicted vs. data
+pred_vs_data_dprime10 <- bias2 %>%
+  add_predicted_draws(brm_dprime.sp10, n = 20)
+
+ggplot(pred_vs_data_dprime10, aes(.prediction, dprime.sp)) +
+  geom_point(alpha = 0.25)
+
+# Abundance >= 3
 pp_check(brm_dprime.sp00)
 bayes_R2(brm_dprime.sp00)
+
+## QQ plot
+qres_dprime00 <- bias %>%
+  add_predicted_draws(brm_dprime.sp00, n = 20) %>%
+  summarise(
+    quant_residual = mean(.prediction < dprime.sp)
+  )
+
+ggplot(qres_dprime00,
+         aes(sample = quant_residual)) +
+  geom_qq(distribution = stats::qunif) +
+  geom_qq_line(distribution = stats::qunif)
+
+## Predicted vs. data
+pred_vs_data_dprime00 <- bias %>%
+  add_predicted_draws(brm_dprime.sp00, n = 20)
+
+ggplot(pred_vs_data_dprime00, aes(.prediction, dprime.sp)) +
+  geom_point(alpha = 0.25) +
+  geom_smooth(method = "lm")
+
+# Abundance >= 5
+pp_check(brm_dprime.sp20)
+bayes_R2(brm_dprime.sp20)
+
+## QQ plot
+qres_dprime20 <- filter(bias2, bb.sp.abund >= 5) %>%
+  add_predicted_draws(brm_dprime.sp20, n = 20) %>%
+  summarise(
+    quant_residual = mean(.prediction < dprime.sp)
+  )
+
+ggplot(qres_dprime20,
+         aes(sample = quant_residual)) +
+  geom_qq(distribution = stats::qunif) +
+  geom_qq_line(distribution = stats::qunif)
+
+## Predicted vs. data
+pred_vs_data_dprime20 <-  filter(bias2, bb.sp.abund >= 5) %>%
+  add_predicted_draws(brm_dprime.sp20, n = 20)
+
+ggplot(pred_vs_data_dprime20, aes(.prediction, dprime.sp)) +
+  geom_point(alpha = 0.25) +
+  geom_smooth(method = "lm")
+
+# Abundance >= 10
+pp_check(brm_dprime.sp30)
+bayes_R2(brm_dprime.sp30)
+
+## QQ plot
+qres_dprime30 <- filter(bias2, bb.sp.abund >= 10) %>%
+  add_predicted_draws(brm_dprime.sp30, n = 20) %>%
+  summarise(
+    quant_residual = mean(.prediction < dprime.sp)
+  )
+
+ggplot(qres_dprime30,
+         aes(sample = quant_residual)) +
+  geom_qq(distribution = stats::qunif) +
+  geom_qq_line(distribution = stats::qunif)
+
+## Predicted vs. data
+pred_vs_data_dprime30 <-  filter(bias2, bb.sp.abund >= 10) %>%
+  add_predicted_draws(brm_dprime.sp20, n = 20)
+
+ggplot(pred_vs_data_dprime30, aes(.prediction, dprime.sp)) +
+  geom_point(alpha = 0.210) +
+  geom_smooth(method = "lm")
+
+```
+
+#### 1.3.3.2 Preliminary visualization
+```{r}
+# Full data
+conditional_effects(brm_dprime.sp10)
+mcmc_plot(brm_dprime.sp10, pars = "^r_bb.sp")
+mcmc_plot(brm_dprime.sp10, pars = "^r_site")
+
+# Abundance >= 3
+conditional_effects(brm_dprime.sp00)
 mcmc_plot(brm_dprime.sp00, pars = "^r_bb.sp")
@@
 
-#### 1.3.3.1 Plot
+#### 1.3.3.4 Final visualization
 ```{r}
@@
 
-# Subsetted data with skew normal 
+# Subsetted data 
 get_variables(brm_dprime.sp00)
@@
 # 2. Explaining bias
-## 2.1 Calculate visitation probability per plant speces/ktype/family
-```{r}
-pvis_ktype <- visit_cover %>%
-  group_by(site, date, bb.sp, plant.sp) %>%
+## 2.1 Calculate visitation probability per plant morphotype.
+Instead of keeping the plant species resolved, we will pool visitation by morphotype.
+The problem with not pooling by morphotype is that more species-rich morphotypes will be penalized simply because their visitation is diluted across more species.
+```{r}
+prop_visits_ktype <- visit_cover %>%
+  group_by(site, date, bb.sp, k.type.ss) %>%
   summarize(flower.cover = sum(flower.cover), # get raw flower cover and visits by site*date*bb.sp*plant.sp
@@
   group_by(site, date) %>%
-  mutate(prop.flower.cover = flower.cover/sum(flower.cover), # proportional flower cover
-         prop.visits = visits/sum(visits), # proportional visits
-         visits.bin = factor(if_else(visits > 0, 1, 0))) %>% # binary visits
+  mutate(prop.flower.cover = flower.cover/sum(flower.cover))%>% # binary visits
   group_by(site, date, bb.sp) %>%
-  mutate(bb.sp.abund = sum(visits)) %>% # abundance of each BB sp. per site-date
+  mutate(prop.visits = visits/sum(visits),
+         bb.sp.abund = sum(visits)) %>% # abundance of each BB sp. per site-date
   ungroup() %>%
-  left_join(fl_traits) %>% # add floral traits
   left_join(bb_traits) %>% # add BB traits
-  mutate(k.type.ss = factor(k.type.ss), # recode variables
-         pbl.w.class2 = factor(pbl.w.class2),
+  mutate(ktype = factor(k.type.ss), # recode variables
+         tongue = factor(pbl.w.class2),
+         visits = as.integer(visits)) %>% # so we can run a Poisson model
+  left_join(site_data) %>% # add site data
+  ungroup() %>%
+  mutate(yday = yday(date),
+         elev.mean0 = scale(elev.mean),
+         yday0 = scale(yday),
+         prop.flower.cover0 = scale(prop.flower.cover),
+         perc.flower.cover = prop.flower.cover*100,
+         bb.sp.abund0 = scale(bb.sp.abund)) %>%
+  select(site, elev.mean, date, yday, bb.sp, ktype, visits, 
+         prop.visits, perc.flower.cover, bb.sp.abund, tongue,
+         elev.mean0, yday0, prop.flower.cover0, bb.sp.abund0) %>%
+  mutate(visits.bin = if_else(visits > 0, 1, 0))
+
+write_csv(prop_visits_ktype, "../hpc_packets/prop_visits_ktype.csv")
+
+prop_visits_ktype_pooled <- visit_cover %>%
+  group_by(site, date, k.type.ss) %>%
+  summarize(flower.cover = sum(flower.cover), # get raw flower cover and visits by site*date*bb.sp*plant.sp
+            visits = sum(visits)) %>%
+  group_by(site, date) %>%
+  mutate(prop.flower.cover = flower.cover/sum(flower.cover),
+         prop.visits = visits/sum(visits),
+         bb.abund = sum(visits),
          visits = as.integer(visits)) %>%
-  filter(!k.type.ss %in% c("9", "10")) %>% # Drop ktypes 9 and 10 since both are extremely rare
-  left_join(site_data)  %>%
-  mutate(prop.flower.cover0 = scale(prop.flower.cover), # normalized predictors
-         bb.sp.abund0 = scale(bb.sp.abund),
+  mutate(ktype = factor(k.type.ss)) %>%
+  left_join(site_data) %>% # add site data
+  ungroup() %>%
+  mutate(yday = yday(date),
          elev.mean0 = scale(elev.mean),
-         yday = yday(date), # and add yday variable
-         yday0 = scale(yday)) %>%
-  select(site, elev.mean0, elev.mean, elev.class2, date, yday, yday0, bb.sp, plant.sp, visits, 
-         visits.bin, prop.visits, prop.flower.cover, prop.flower.cover0,
-         k.type.ss, bb.sp.abund, bb.sp.abund0, pbl.w.class2) 
+         yday0 = scale(yday),
+         perc.flower.cover = prop.flower.cover*100,
+         prop.flower.cover0 = scale(prop.flower.cover),
+         bb.abund0 = scale(bb.abund)) %>%
+  select(site, elev.mean, date, yday, ktype, visits, 
+         prop.visits, bb.abund, perc.flower.cover,
+         elev.mean0, yday0, prop.flower.cover0, bb.abund0) %>%
+  mutate(visits.bin = if_else(visits > 0, 1, 0))
+
+write_csv(prop_visits_ktype_pooled, "../hpc_packets/prop_visits_ktype_pool.csv")
+
 ```
@@
 ## 2.2 Multi-level models
-
-### 2.2.1 Preliminary visualization
-The tons of zeros are due to the lots of plant species that were present at very low abundance and received few or zero visits.
-This might not be ideal.
-We could opt for one of our alternative bias measure, like fold change, which all give something like a poisson distribution, but I don't know if this would be an improvement.
-We could also filter our data to include only plants that were above some threshold of relative abundance, say, 1%.
-```{r}
-# Visit.prop
-ggplot(filter(pvis_ktype, prop.flower.cover >= 0.05), aes(prop.visits)) +
-  geom_histogram()
-
-# Visits.bin
-ggplot(pvis_ktype, aes(k.type.ss, fill = visits.bin)) +
-  geom_bar()
-
-ggplot(pvis_ktype, aes(elev.mean, prop.visits)) +
-  geom_point() +
-  geom_smooth() +
-  facet_grid(bb.sp~k.type.ss)
-
-# Visits.bin filtered
-ggplot(filter(pvis_ktype, prop.flower.cover >= 0.05), aes(k.type.ss, fill = visits.bin)) +
-  geom_bar()
-```
-
-### 2.2.2 brms models
-```{r}
-# Zero- one-inflated beta regression
-brm_pvis01 <- brm(prop.visits ~ prop.flower.cover0 + 
-                    (1 | k.type.ss) + 
-                    (1 | site) + 
-                    (1 | bb.sp),
-                  family = "zero_one_inflated_beta",
-                  file = "./output/brm_pvis01",
+### I tried fitting proportional visitation with beta regression models, but the models performed really badly, with R2 < 1%.
+I was depressed for some time.
+But then I realized that the problem was probably caused by low abundance species, which constitute most of my data.
+For low-abundance species, and especially singletons, proportional visitation tends to be 100% for a single ktype and 0% for all others.
+That's tough to model.
+But using the raw count data, that's not a problem, because we could accurately predict that the visitation of a rare species would be extremely low, and the difference between 1 visit and 0 visits is small.
+So, I am opting for negbinomial models (these are more robust than Poisson, allowing for overdispersion I think --- loo likes them better than Poisson anyway).
+The costs are (1) I now have to condition on the interaction of BB abundance and proportional floral abundance and (2) it's theoretically possible to predict more visits than bumble bees.
+As long as the models converge, I don't think that (1) is a problem.
+(2) is undesirable, but I don't think it's really that serious, since the point is get our predictions close to reality, even if the actual predictions are impossible.
+I would rather get an impossible prediction that is close to reality than a possible one that is far from it (say, a prediction of 0% when reality is 100% --- no wonder my beta models flopped!).
+With the negbinomial models, I'm explaining ~50% of the variance, and I call that a win.
+
+If these models are legit --- and I need to be patient in making sure of that --- then I have a jackpot of inference here.
+I have the morphotype:bb.sp and morphotype:tongue effects in spades.
+But a really cool and unexpected finding is that there is a decoupling of the intercept and slope terms.
+Some morphotypes have high intercepts but low slopes, indicating that they are visited heavily even when they are at relatively low abundance.
+Others have lower intercepts but response more steeply to prop flower cover.
+I think I can tell a story about that.
+Finally, there is substantial site variation, which could be interpreted as lability of bias.
+
+But these models need to be scrutinized.
+In particular, I don't really understand the pp.check output.
+The negbinomial models seem like they might be weirder than the Poisson models.
+I could always try zero-inflated models if that seems necessary.
+### 2.2.1 Call models
+```{r}
+## Exploring the differential bias of bumble bee species and tongue length classes towards floral morphotypes
+### Bernoulli with binarized visitation
+brm_pvis30 <- brm(as.integer(visits.bin) ~
+                    bb.sp.abund*perc.flower.cover + # fixed term: because more abundant flowers will tend to be more visited,and more abundance bees visit more
+                    (1 + perc.flower.cover | ktype:bb.sp) + # a random intercept and random slope for ktype:species; I think I need to model this before I can make conclusions about ktype:pbl.class
+                    (1 + perc.flower.cover | ktype:tongue) + # a random intercept and slope for ktype:tongue
+                    (1 + perc.flower.cover | ktype:site), # a random intercept and slope for ktype:site
+                  family = "bernoulli", 
+                  #file = "../output/brm_prop_visits_ktype_10",  # for running locally
+                  file = "brm_prop_visits_ktype_30", # for running on the HPC
                   chains = 4,
-                  cores = 4,
+                  cores = 12,
                   iter = 2000,
-                  data = pvis_ktype)
-
-# Poisson models of raw visit counts
-brm_pvis00 <- brm(visits ~ prop.flower.cover0 + 
-                    bb.sp.abund0 + 
-                    (1 | k.type.ss) + 
-                    (1 | site) + 
-                    (1 | bb.sp),
-                  #control = list(adapt_delta = 0.99),
-                  family = "poisson",
-                  file = "../output/brm_pvis00",
+                  data = filter(prop_visits_ktype, !ktype %in% c("9", "10")))
+
+### Negbinomial 
+brm_pvis10 <- brm(as.integer(visits) ~
+                    bb.sp.abund*perc.flower.cover + # fixed term: because more abundant flowers will tend to be more visited,and more abundance bees visit more
+                    (1 + perc.flower.cover | ktype:bb.sp) + # a random intercept and random slope for ktype:species; I think I need to model this before I can make conclusions about ktype:pbl.class
+                    (1 + perc.flower.cover | ktype:tongue) + # a random intercept and slope for ktype:tongue
+                    (1 + perc.flower.cover | ktype:site), # a random intercept and slope for ktype:site
+                  family = "negbinomial", 
+                  #file = "../output/brm_prop_visits_ktype_10",  # for running locally
+                  file = "brm_prop_visits_ktype_10", # for running on the HPC
                   chains = 4,
-                  #cores = 4,
+                  cores = 12,
                   iter = 2000,
-                  data = pvis_ktype)
-
-summary(brm_pvis00)
-prior_summary(brm_pvis00)
-parnames(brm_pvis00)
-mcmc_plot(brm_pvis00, pars = c("^r_k"))
-mcmc_plot(brm_pvis00, pars = c("^r_s"))
-mcmc_plot(brm_pvis00, pars = c("^b_"))
-mcmc_plot(brm_pvis00)
-conditional_effects(brm_pvis00)
-pp_check(brm_pvis00)
-bayes_R2(brm_pvis00)
-
-# Perhaps I should add the bb.sp varying intercept effect; otherwise, I think I could be accused of pseudo replication.
-brm_pvis11 <- brm(visits ~ prop.flower.cover0 + 
-                    bb.sp.abund0 + 
-                    (1 | k.type.ss:pbl.w.class2) + 
-                    (1 | site) +
-                    (1 | bb.sp),
-                  family = "poisson",
-                  file = "../output/brm_pvis11",
+                  data = filter(prop_visits_ktype, !ktype %in% c("9", "10")))
+
+### Zero-inflated negbinomial
+brm_pvis11 <- brm(as.integer(visits) ~
+                    bb.sp.abund*perc.flower.cover + # fixed term: because more abundant flowers will tend to be more visited,and more abundance bees visit more
+                    (1 + perc.flower.cover | ktype:bb.sp) + # a random intercept and random slope for ktype:species; I think I need to model this before I can make conclusions about ktype:pbl.class
+                    (1 + perc.flower.cover | ktype:tongue) + # a random intercept and slope for ktype:tongue
+                    (1 + perc.flower.cover | ktype:site), # a random intercept and slope for ktype:site
+                  family = "zero_inflated_negbinomial", 
+                  #file = "../output/brm_prop_visits_ktype_11",  # for running locally
+                  file = "brm_prop_visits_ktype_11", # for running on the HPC
                   chains = 4,
-                  #cores = 4,
+                  cores = 12,
                   iter = 2000,
-                  data = pvis_ktype)
-
-summary(brm_pvis11)
-prior_summary(brm_pvis11)
-parnames(brm_pvis11)
-mcmc_plot(brm_pvis11, pars = c("^r_k"))
-mcmc_plot(brm_pvis11, pars = c("^r_b"))
-mcmc_plot(brm_pvis11, pars = c("^b_"))
-mcmc_plot(brm_pvis11)
+                  data = filter(prop_visits_ktype, !ktype %in% c("9", "10")))
+
+### Poisson
+brm_pvis20 <- brm(as.integer(visits) ~
+                    bb.sp.abund*perc.flower.cover + # fixed term: because more abundant flowers will tend to be more visited,and more abundance bees visit more
+                    (1 + perc.flower.cover | ktype:bb.sp) + # a random intercept and random slope for ktype:species; I think I need to model this before I can make conclusions about ktype:pbl.class
+                    (1 + perc.flower.cover | ktype:tongue) + # a random intercept and slope for ktype:tongue
+                    (1 + perc.flower.cover | ktype:site), # a random intercept and slope for ktype:site
+                  family = "Poisson", # because my proportional data have zeros and ones
+                  #file = "../output/brm_prop_visits_ktype_20",  # for running locally
+                  file = "brm_prop_visits_ktype_20", # for running on the HPC
+                  chains = 4,
+                  cores = 12,
+                  iter = 2000,
+                  data = filter(prop_visits_ktype, !ktype %in% c("9", "10")))
+
+### Zero-inflated Poisson
+brm_pvis21 <- brm(as.integer(visits) ~
+                    bb.sp.abund*perc.flower.cover + # fixed term: because more abundant flowers will tend to be more visited,and more abundance bees visit more
+                    (1 + perc.flower.cover | ktype:bb.sp) + # a random intercept and random slope for ktype:species; I think I need to model this before I can make conclusions about ktype:pbl.class
+                    (1 + perc.flower.cover | ktype:tongue) + # a random intercept and slope for ktype:tongue
+                    (1 + perc.flower.cover | ktype:site), # a random intercept and slope for ktype:site
+                  family = "zero_inflated_poisson", # because my proportional data have zeros and ones
+                  #file = "../output/brm_prop_visits_ktype_21",  # for running locally
+                  file = "brm_prop_visits_ktype_21", # for running on the HPC
+                  chains = 4,
+                  cores = 12,
+                  iter = 2000,
+                  data = filter(prop_visits_ktype, !ktype %in% c("9", "10")))
+
+brm_pvis22 <- brm(as.integer(visits) ~
+                    bb.sp.abund*perc.flower.cover + # fixed term: because more abundant flowers will tend to be more visited,and more abundance bees visit more
+                    (1 | ktype:bb.sp) + # a random intercept and random slope for ktype:species; I think I need to model this before I can make conclusions about ktype:pbl.class
+                    (1 + perc.flower.cover | ktype:tongue) + # a random intercept and slope for ktype:tongue
+                    (1 | ktype:site), # a random intercept and slope for ktype:site
+                  family = "zero_inflated_poisson", # because my proportional data have zeros and ones
+                  #file = "../output/brm_prop_visits_ktype_22",  # for running locally
+                  file = "brm_prop_visits_ktype_22", # for running on the HPC
+                  chains = 4,
+                  cores = 12,
+                  iter = 2000,
+                  data = filter(prop_visits_ktype, !ktype %in% c("9", "10")))
+```
+### 2.2.2 Evaluate models
+```{r}
+# Posterior predictive checks
+pp_check(brm_pvis10)
+pp_check(brm_pvis11)
+pp_check(brm_pvis20)
+pp_check(brm_pvis21)
+
+# Predictions vs. data
+## Get predictions
+brm_pvis10_pred <- predict(brm_pvis10) %>%
+  as_tibble()
+
+brm_pvis11_pred <- predict(brm_pvis11) %>%
+  as_tibble()
+
+brm_pvis20_pred <- predict(brm_pvis20) %>%
+  as_tibble()
+
+brm_pvis21_pred <- predict(brm_pvis21) %>%
+  as_tibble()
+
+## Add draws
+qres10 <- filter(prop_visits_ktype, !ktype %in% c("9", "10")) %>%
+  add_predicted_draws(brm_pvis10, n = 10)
+
+qres11 <- filter(prop_visits_ktype, !ktype %in% c("9", "10")) %>%
+  add_predicted_draws(brm_pvis11, n = 10)
+
+qres20 <- filter(prop_visits_ktype, !ktype %in% c("9", "10")) %>%
+  add_predicted_draws(brm_pvis20, n = 10)
+
+qres21 <- filter(prop_visits_ktype, !ktype %in% c("9", "10")) %>%
+  add_predicted_draws(brm_pvis21, n = 10)
+
+## Plot predictions vs. data
+ggplot(qres10, aes(.prediction, visits)) +
+  geom_point() +
+  geom_smooth()
+
+ggplot(qres11, aes(.prediction, visits)) +
+  geom_point() +
+  geom_smooth()
+
+ggplot(qres20, aes(.prediction, visits)) +
+  geom_point() +
+  geom_smooth()
+
+ggplot(qres21, aes(.prediction, visits)) +
+  geom_point() +
+  geom_smooth()
+```
+
+### 2.2.3 Preliminary visualization
+```{r}
+# Negbinmial
+get_variables(brm_pvis10)
+conditional_effects(brm_pvis10)
+mcmc_plot(brm_pvis10, pars = c("^r_ktype:tongue.*Intercept]$"))
+mcmc_plot(brm_pvis10, pars = c("^r_ktype:tongue.*perc.flower.cover]$"))
+mcmc_plot(brm_pvis10, pars = c("^r_ktype:bb.sp.*Intercept]$"))
+mcmc_plot(brm_pvis10, pars = c("^r_ktype:bb.sp.*perc.flower.cover]$"))
+mcmc_plot(brm_pvis10, pars = c("^r_ktype:site.*Intercept]$"))
+mcmc_plot(brm_pvis10, pars = c("^r_ktype:site.*perc.flower.cover]$"))
+
+# Zero-inflated negbinomial
+get_variables(brm_pvis11)
 conditional_effects(brm_pvis11)
-pp_check(brm_pvis11)
-bayes_R2(brm_pvis11)
-```
-
-#### 2.2.2.1 Plot
-```{r}
-# Note that the intercept is on the log scale, so if you want to transform it back to predicted # visits, you have to exponentiate it. I think.
-
-## Interaction model
-brm_pvis11_plotframe <- brm_pvis11 %>%
-  spread_draws(b_Intercept, `r_k.type.ss:pbl.w.class2`[condition, term]) %>% # the back ticks are necessary to escape the :
-  rename(r_ktype.pbl = `r_k.type.ss:pbl.w.class2`) %>%
-  mutate(condition_mean = b_Intercept + r_ktype.pbl) %>%
-  separate(condition, c("ktype", "pbl.class"), sep = "_", remove = FALSE) %>%
-  mutate(pbl.class = case_when(
-    pbl.class == "1" ~ "short",
-    pbl.class == "2" ~ "med",
-    pbl.class == "3" ~ "long")
+mcmc_plot(brm_pvis11, pars = c("^r_ktype:tongue.*Intercept]$"))
+mcmc_plot(brm_pvis11, pars = c("^r_ktype:tongue.*perc.flower.cover]$"))
+mcmc_plot(brm_pvis11, pars = c("^r_ktype:bb.sp.*Intercept]$"))
+mcmc_plot(brm_pvis11, pars = c("^r_ktype:bb.sp.*perc.flower.cover]$"))
+mcmc_plot(brm_pvis11, pars = c("^r_ktype:site.*Intercept]$"))
+mcmc_plot(brm_pvis11, pars = c("^r_ktype:site.*perc.flower.cover]$"))
+
+# Poisson
+get_variables(brm_pvis20)
+conditional_effects(brm_pvis20)
+mcmc_plot(brm_pvis20, pars = c("^r_ktype:tongue.*Intercept]$"))
+mcmc_plot(brm_pvis20, pars = c("^r_ktype:tongue.*perc.flower.cover]$"))
+mcmc_plot(brm_pvis20, pars = c("^r_ktype:bb.sp.*Intercept]$"))
+mcmc_plot(brm_pvis20, pars = c("^r_ktype:bb.sp.*perc.flower.cover]$"))
+mcmc_plot(brm_pvis20, pars = c("^r_ktype:site.*Intercept]$"))
+mcmc_plot(brm_pvis20, pars = c("^r_ktype:site.*perc.flower.cover]$"))
+
+# zero-inflated Poisson
+get_variables(brm_pvis21)
+conditional_effects(brm_pvis21)
+mcmc_plot(brm_pvis21, pars = c("^r_ktype:tongue.*Intercept]$"))
+mcmc_plot(brm_pvis21, pars = c("^r_ktype:tongue.*perc.flower.cover]$"))
+mcmc_plot(brm_pvis21, pars = c("^r_ktype:bb.sp.*Intercept]$"))
+mcmc_plot(brm_pvis21, pars = c("^r_ktype:bb.sp.*perc.flower.cover]$"))
+mcmc_plot(brm_pvis21, pars = c("^r_ktype:site.*Intercept]$"))
+mcmc_plot(brm_pvis21, pars = c("^r_ktype:site.*perc.flower.cover]$"))
+```
+
+### 2.2.4 Final visualization
+# Note that the intercept is on the log scale, so if you want to transform it back to predicted # visits, you have to exponentiate it.
+I think.
+Here is a helplful conversation involving the esteemed Ben Bolker:
+
+---
+Dear all,
+
+I used the glm.nb with the default values from the MASS package to run a negative binomial regression.
+Here is a simple example...The question now is how do I report the results, say, for height?
+Do I simply take the anti logs.
+i.e.
+1.019613 = exp(0.019423)?
+
+Bolker:
+Yes, exactly.
+If you look at ?glm.nb you will see that it uses a log link function, and therefore you should exponentiate (anti-log) to back-transform.
+Natural, not base-10 logs, are used.
+Don't forget that back-transforming standard errors by themselves is meaningless, you have to back-transform lower and upper confidence limits ...
+
+---
+
+Another conversation, this time with Paul Buerkner
+
+---
+I ultimately want to discuss the results of the model fitting on the linear (not log) parameter scale.
+Is it fine to simply exp() the estimates (samples) when computing credible intervals etc?
+Obviously means would be stretched, but quantiles should be invariant.
+
+Buerkner: Yes, this is exactly, how it should be done :-)
+
+---
+
+In my workflow, I think this entails first calculating the conditional mean, then running pointinterval() to get the credible intervals **then** exponentiating the conditional mean and the credible intervals.
+
+##### 2.2.2.1.1 morphotype:tongue intercept
+```{r}
+brm_pvis12_plotframe <- brm_pvis12 %>%
+  spread_draws(b_Intercept, `r_ktype:tongue`[condition, term]) %>% # the back ticks are necessary to escape the :
+  rename(r_ktype.tongue = `r_ktype:tongue`) %>% # get rid of the problematic :
+  filter(term == "Intercept") %>% # drop slope terms for now
+  mutate(condition_mean = b_Intercept + r_ktype.tongue) %>% # get conditional mean
+  group_by(condition) %>% # group by tongue:morpotype pair
+  point_interval(condition_mean, .width = c(0.95, 0.66)) %>% # calculate credible intervals
+  mutate(condition_mean_trans = exp(condition_mean), # exponentiating the estimates gets us back to the response scale (negbinomial uses log link function)
+         .lower_trans = exp(.lower),
+         .upper_trans = exp(.upper)) %>% 
+separate(condition, c("ktype", "tongue"), sep = "_", remove = FALSE) %>% # separate condition term into ktype and tongue
+  mutate(ktype = case_when( # rename
+    ktype == "0" ~ "NA (0)",
+    ktype == "1" ~ "Disc (1)",
+    ktype == "2" ~ "Funnel (2)",
+    ktype == "3" ~ "Bell (3)",
+    ktype == "4" ~ "Stalk-disc (4)",
+    ktype == "5" ~ "Lip (5)",
+    ktype == "6" ~ "Flag (6)",
+    ktype == "7" ~ "Head (7)")
   ) %>%
+  mutate(ktype = factor(ktype, levels = c("Disc (1)", "Funnel (2)", "Bell (3)", # relevel
+                                          "Stalk-disc (4)","Lip (5)", "Flag (6)", 
+                                          "Head (7)", "NA (0)")))
+
+ggplot(brm_pvis12_plotframe, aes(condition_mean_trans, tongue, 
+                                 xmin = .lower_trans, xmax = .upper_trans,
+                                 color = tongue, fill = tongue)) +
+  geom_pointinterval() +
+  facet_grid(ktype~.) +
+  theme_light() +
+  labs(color = "Tongue length", x = "Mean visits") +
+  guides(fill = FALSE) +
+  theme(axis.title.y=element_blank(),
+        axis.text.y=element_blank(),
+        axis.ticks.y=element_blank())
+
+ggsave(file = "../output/brm_pvis12_plot1.png", width = 5, height = 7)
+```
+
+##### 2.2.2.1.2 morphotype:bb.sp intercept
+```{r}
+# ktype:bb.sp intercept
+brm_pvis12_plotframe2 <- brm_pvis12 %>%
+  spread_draws(b_Intercept, `r_ktype:bb.sp`[condition, term]) %>% # the back ticks are necessary to escape the :
+  rename(r_ktype.bb.sp = `r_ktype:bb.sp`) %>% # get rid of the problematic :
+  filter(term == "Intercept") %>% # drop slope terms for now
+  mutate(condition_mean = b_Intercept + r_ktype.bb.sp) %>% # get conditional mean
+  group_by(condition) %>% # group by bb.sp:morpotype pair
+  point_interval(condition_mean, .width = c(0.95, 0.66)) %>% # calculate credible intervals
+  mutate(condition_mean_trans = exp(condition_mean), # exponentiating the estimates gets us back to the response scale (negbinomial uses log link function)
+         .lower_trans = exp(.lower),
+         .upper_trans = exp(.upper)) %>% 
+  separate(condition, c("ktype", "bb.sp"), sep = "_", remove = FALSE) %>% # separate condition term into ktype and bb.sp
+  mutate(ktype = case_when( # rename
+    ktype == "0" ~ "NA (0)",
+    ktype == "1" ~ "Disc (1)",
+    ktype == "2" ~ "Funnel (2)",
+    ktype == "3" ~ "Bell (3)",
+    ktype == "4" ~ "Stalk-disc (4)",
+    ktype == "5" ~ "Lip (5)",
+    ktype == "6" ~ "Flag (6)",
+    ktype == "7" ~ "Head (7)")
+  ) %>%
+  mutate(ktype = factor(ktype, levels = c("Disc (1)", "Funnel (2)", "Bell (3)", # relevel
+                                          "Stalk-disc (4)","Lip (5)", "Flag (6)", 
+                                          "Head (7)", "NA (0)"))) %>%
+  left_join(bb_traits)
+
+ggplot(brm_pvis12_plotframe2, aes(condition_mean_trans, bb.sp, 
+                                 xmin = .lower_trans, xmax = .upper_trans,
+                                 color = pbl.w.class2, fill = pbl.w.class2)) +
+  geom_pointinterval() +
+  facet_wrap(~ktype, scales = "free") +
+  theme_light(10) +
+  labs(color = "Tongue length", x = "Mean visits") +
+  guides(fill = FALSE)
+
+ggsave(file = "../output/brm_pvis12_plot2.png", width = 9, height = 6.5)
+
+ggplot(brm_pvis12_plotframe2, aes(condition_mean_trans, ktype, 
+                                 xmin = .lower_trans, xmax = .upper_trans)) +
+  geom_pointinterval() +
+  facet_wrap(~bb.sp, scales = "free") +
+  #facet_grid(ktype~.) 
+  theme_light(10) +
+  labs(color = "Tongue length", x = "Mean visits") +
+  guides(fill = FALSE)
+```
+
+##### 2.2.2.1.3 morphotype:tongue slope
+```{r}
+brm_pvis12_plotframe3 <- brm_pvis12 %>%
+  spread_draws(b_prop.flower.cover0, `r_ktype:tongue`[condition, term]) %>% # the back ticks are necessary to escape the :
+  rename(r_ktype.tongue = `r_ktype:tongue`) %>%
+  filter(term == "prop.flower.cover0") %>%
+  mutate(condition_slope = b_prop.flower.cover0 + r_ktype.tongue) %>%
+  separate(condition, c("ktype", "tongue"), sep = "_", remove = FALSE) %>%
   mutate(ktype = case_when(
@@
 
-# ggplot(brm_pvis11_plotframe, aes(condition_mean, pbl.class, color = pbl.class, fill = pbl.class)) +
-#   stat_pointinterval() +
-#   #stat_halfeye(alpha = 0.5, interval_alpha = 1, point_alpha = 1, size = 1, .width = 0.95) +
-#   facet_grid(ktype ~ .) +
-#   theme_light() +
-#   labs(color = "Tongue length", x = "Visits (log-scale)") +
-#   guides(fill = FALSE) +
-#   theme(axis.title.y=element_blank(),
-#         axis.text.y=element_blank(),
-#         axis.ticks.y=element_blank())
-
-ggplot(brm_pvis11_plotframe, aes(condition_mean, pbl.class, color = pbl.class, fill = pbl.class)) +
-  #stat_pointinterval() +
-  stat_halfeye(alpha = 0.4, interval_alpha = 1, point_alpha = 1) +
-  facet_wrap(~ktype, scales = "free") +
+ggplot(brm_pvis12_plotframe3, aes(condition_slope, tongue, color = tongue, fill = tongue)) +
+  stat_pointinterval() +
+  facet_grid(ktype~.) +
   theme_light() +
@@
 
-ggsave(file = "../output/brm_pvis11_plot1.png", width = 7, height = 5.5)
-
+ggsave(file = "../output/brm_pvis12_plot2.png", width = 7, height = 5.5)
+```
+
+# ktype:bb.sp slope
+```{r}
+brm_pvis12_plotframe4 <- brm_pvis12 %>%
+  spread_draws(b_prop.flower.cover0, `r_ktype:bb.sp`[condition, term]) %>% # the back ticks are necessary to escape the :
+  rename(r_ktype.bb.sp = `r_ktype:bb.sp`) %>%
+  filter(term == "prop.flower.cover0") %>%
+  mutate(condition_slope = b_prop.flower.cover0 + r_ktype.bb.sp) %>%
+  separate(condition, c("ktype", "bb.sp"), sep = "_", remove = FALSE) %>%
+  mutate(ktype = case_when(
+    ktype == "0" ~ "NA (0)",
+    ktype == "1" ~ "Disc (1)",
+    ktype == "2" ~ "Funnel (2)",
+    ktype == "3" ~ "Bell (3)",
+    ktype == "4" ~ "Stalk-disc (4)",
+    ktype == "5" ~ "Lip (5)",
+    ktype == "6" ~ "Flag (6)",
+    ktype == "7" ~ "Head (7)")
+  ) %>%
+  mutate(ktype = factor(ktype,
+                        levels = c("Disc (1)", "Funnel (2)", "Bell (3)", "Stalk-disc (4)",
+                                   "Lip (5)", "Flag (6)", "Head (7)", "NA (0)"))) %>%
+  left_join(bb_traits)
+
+ggplot(brm_pvis12_plotframe4, aes(condition_slope, bb.sp, color = pbl.w.class2)) +
+  stat_pointinterval() +
+  facet_wrap(~ktype, scales = "free") +
+  theme_light() +
+  labs(color = "Tongue length", x = "Visits (log-scale)") +
+  guides(fill = FALSE)
+
+ggsave(file = "../output/brm_pvis12_plot1.png", width = 7, height = 5.5)
+```
+
+##### 2.2.2.1.2 Ktype-only model 
+```{r}
 # Just ktype model
@@
 ggsave(file = "../output/brm_pvis00_plot1.png", width = 7, height = 5.5)
+
+# What I could do here that I haven't done yet is to plot the predicted curves, incorporating all the slope and intercept information.
+ggplot(filter(prop_visits_ktype,!ktype %in% c("9", "10")), aes(bb.sp.abund, prop.flower.cover, size = visits, color = tongue)) +
+  geom_point(alpha = 0.15) +
+  facet_wrap(~ktype) +
+  theme_light()
 ```
@@
 ### 2.2.3 The lability of bias: How does the foraging bias of each species with respect to each morphotype vary through elevation and time?
+```{r}
+# mgcv time
+gam1 <- bam(prop.visits ~ prop.flower.cover0 +
+              te(yday, elev.mean, by = k.type.ss, bs = c("gp", "tp"), k = c(8, 8)),
+            family = "betar",
+            method = "fREML",
+            discrete = TRUE,
+            data = pvis_ktype_pool) %>% getViz()
+
+summary(gam1)
+check.gamViz(gam1)
+plot.gamViz(gam1)
+
+gam2 <- bam(visits ~ te(yday, elev.mean, by = k.type.ss, bs = c("gp", "tp"), k = c(8, 8)),
+            family = "quasipoisson",
+            method = "fREML",
+            discrete = TRUE,
+            data = pvis_ktype_pool) %>% getViz()
+
+summary(gam2)
+check.gamViz(gam2)
+plot.gamViz(gam2)
+
+```
+
+#### We need to do some kind of tensor product gam on this jawn!
+```{r}
+ggplot(pvis_ktype_pool, aes(yday, elev.mean, color = prop.visits)) +
+  geom_point() +
+  facet_wrap(~k.type.ss)
+```
+
 ```{r}
@@
 bayes_R2(brm_pviselev00)
+```
+
+## A related question: do bumble bee species adjust their morphotype preferences along an elevation gradient? Or maybe through time?
+```{r}
+# Conceptual plot
+ggplot(filter(prop_visits_ktype, bb.sp == "prat"), aes(elev.mean, (prop.visits - (perc.flower.cover/100)))) +
+  geom_point() +
+  facet_wrap(~ktype, scales = "free") +
+  geom_smooth()
+
+ggplot(filter(prop_visits_ktype, bb.sp == "pasc"), aes(elev.mean, (prop.visits - (perc.flower.cover/100)))) +
+  geom_point() +
+  facet_wrap(~ktype, scales = "free") +
+  geom_smooth()
+
+ggplot(filter(prop_visits_ktype, bb.sp == "wurf"), aes(elev.mean, (prop.visits - (perc.flower.cover/100)))) +
+  geom_point() +
+  facet_wrap(~ktype, scales = "free") +
+  geom_smooth()
+
+ggplot(filter(prop_visits_ktype, bb.sp == "soro"), aes(elev.mean, (prop.visits - (perc.flower.cover/100)))) +
+  geom_point() +
+  facet_wrap(~ktype, scales = "free") +
+  geom_smooth()
+
+ggplot(filter(prop_visits_ktype, bb.sp == "telu"), aes(elev.mean, (prop.visits - (perc.flower.cover/100)))) +
+  geom_point() +
+  facet_wrap(~ktype, scales = "free") +
+  geom_smooth()
+
+ggplot(filter(prop_visits_ktype, bb.sp == "hort"), aes(elev.mean, (prop.visits - (perc.flower.cover/100)))) +
+  geom_point() +
+  facet_wrap(~ktype, scales = "free") +
+  geom_smooth()
+
+# brms model
+brm_pvis_elev00 <- brm(as.integer(visits) ~
+                    bb.sp.abund*perc.flower.cover + # fixed term: because more abundant flowers will tend to be more visited,and more abundance bees visit more
+                    (1 + perc.flower.cover + elev.mean0 | ktype:bb.sp), # a random intercept and random slope for ktype:species; I think I need to model this before I can make conclusions about ktype:pbl.class
+                  family = "zero_inflated_poisson", # because my proportional data have zeros and ones
+                  #file = "../output/brm_prop_visits_elev00",  # for running locally
+                  file = "brm_prop_visits_elev00", # for running on the HPC
+                  chains = 4,
+                  cores = 12,
+                  iter = 2000,
+                  data = filter(prop_visits_ktype, !ktype %in% c("9", "10")))
+
+# an mgcv approach
+library(mgcv)
+library(mgcViz)
+gam_data <- prop_visits_ktype %>%
+  filter(!ktype %in% c("0", "1", "9", "10") &
+           bb.sp %in% c("telu", "prat", "pasc", "hort", "soro", "wurf")) %>%
+  unite(pair, c(bb.sp, ktype), sep = "_", remove = FALSE) %>%
+  mutate(pair = factor(pair),
+         bin.visits = if_else(visits > 0, 1, 0))
+
+ggplot(gam_data, aes(elev.mean, bin.visits)) +
+  geom_point() +
+  geom_smooth() +
+  facet_grid(ktype~bb.sp)
+
+gam_pvis_elev00 <- gam(as.integer(visits) ~ s(elev.mean0, pair, bs = "fs") +
+                         bb.sp.abund*perc.flower.cover,
+                       family = "ziP",
+                       method = "REML",
+                       #discrete = TRUE,
+                       data = gam_data) %>% getViz()
+
+summary(gam_pvis_elev00)
+check.gamViz(gam_pvis_elev00)
+plot.gamViz(gam_pvis_elev00)
+plot(gam_pvis_elev00)
+
+gam_pvis_elev01 <- gam(as.integer(visits) ~ s(elev.mean0, by = ktype, k = 4) +
+                         #s(elev.mean0, by = pair, k = 4) +
+                         bb.sp.abund*perc.flower.cover,
+                       family = "poisson",
+                       method = "REML",
+                      # discrete = TRUE,
+                       data = filter(gam_data, bb.sp %in% c("soro", "wurf", "hort", 
+                                                            "telu", "pasc", "prat"))) %>% getViz()
+
+summary(gam_pvis_elev01)
+check.gamViz(gam_pvis_elev01)
+print(plot.gamViz(gam_pvis_elev01, allTerms = TRUE), pages = 6)
+
+gam_pvis_elev02 <- bam(bin.visits ~ 
+                         s(elev.mean0, by = pair, k = 4) +
+                         s(bb.sp.abund, perc.flower.cover),
+                       family = "binomial",
+                       method = "fREML",
+                       discrete = TRUE,
+                       data = filter(gam_data, bb.sp %in% c("soro", "wurf", "hort", 
+                                                            "telu", "pasc", "prat"))) %>% getViz()
+
+summary(gam_pvis_elev02)
+check.gamViz(gam_pvis_elev02)
+plot.gamViz(gam_pvis_elev02, allTerms = TRUE)
 ```
@@
            richness = map(data, function(x) length(x$bb.sp)),
-           dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
+           dist = map(data2, function(x) vegdist(x, method = "jaccard")), # convert to Horn's distance matrix
            tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
@@
     summarize(visits = sum(visits)) %>%
-    select(site, bb.sp, !!criterion, visits) %>%
-    pivot_wider(names_from = !!criterion, values_from = visits, values_fill = 0) %>% # spread into the kind of array vegan likes
+    group_by(site, bb.sp) %>%
+    mutate(prop.visits = visits/sum(visits)) %>%
+    select(site, bb.sp, !!criterion, prop.visits) %>%
+    pivot_wider(names_from = !!criterion, values_from = prop.visits, values_fill = 0) %>% # spread into the kind of array vegan likes
     group_by(site) %>%
@@
   mutate(same.class = case_when(
-    pbl.w.class2_sp1 == "1" & pbl.w.class2_sp2 == "1" ~ "short:short",
-    pbl.w.class2_sp1 == "2" & pbl.w.class2_sp2 == "2" ~ "med:med",
-    pbl.w.class2_sp1 == "3" & pbl.w.class2_sp2 == "3" ~ "long:long",
-    pbl.w.class2_sp1 == "1" & pbl.w.class2_sp2 == "2" ~ "short:med",
-    pbl.w.class2_sp1 == "2" & pbl.w.class2_sp2 == "1" ~ "short:med",
-    pbl.w.class2_sp1 == "1" & pbl.w.class2_sp2 == "3" ~ "short:long",
-    pbl.w.class2_sp1 == "3" & pbl.w.class2_sp2 == "1" ~ "short:long",
-    pbl.w.class2_sp1 == "2" & pbl.w.class2_sp2 == "3" ~ "med:long",
-    pbl.w.class2_sp1 == "3" & pbl.w.class2_sp2 == "2" ~ "med:long")
+    pbl.w.class2_sp1 == "short" & pbl.w.class2_sp2 == "short" ~ "short:short",
+    pbl.w.class2_sp1 == "med" & pbl.w.class2_sp2 == "med" ~ "med:med",
+    pbl.w.class2_sp1 == "long" & pbl.w.class2_sp2 == "long" ~ "long:long",
+    pbl.w.class2_sp1 == "short" & pbl.w.class2_sp2 == "med" ~ "short:med",
+    pbl.w.class2_sp1 == "med" & pbl.w.class2_sp2 == "short" ~ "short:med",
+    pbl.w.class2_sp1 == "short" & pbl.w.class2_sp2 == "long" ~ "short:long",
+    pbl.w.class2_sp1 == "long" & pbl.w.class2_sp2 == "short" ~ "short:long",
+    pbl.w.class2_sp1 == "med" & pbl.w.class2_sp2 == "long" ~ "med:long",
+    pbl.w.class2_sp1 == "long" & pbl.w.class2_sp2 == "med" ~ "med:long")
     ) %>%
@@
 
-# Aggregated niche overlap per species pair per site
-niche_overlap2 <- read_csv("../data/processed_data/network.csv") %>%
-  filter(caste != "male") %>% # we will ignore males 
-  group_by(site, bb.sp, plant.sp) %>%
-  summarize(visits = n()) %>%
-  niche_overlapper2(plant.sp) %>%
-  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>%
-  rename(pbl.w.class2_sp1 = pbl.w.class2) %>%
-  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>%
-  rename(pbl.w.class2_sp2 = pbl.w.class2) %>%
-  mutate(same.class = case_when(
-    pbl.w.class2_sp1 == "1" & pbl.w.class2_sp2 == "1" ~ "short:short",
-    pbl.w.class2_sp1 == "2" & pbl.w.class2_sp2 == "2" ~ "med:med",
-    pbl.w.class2_sp1 == "3" & pbl.w.class2_sp2 == "3" ~ "long:long",
-    pbl.w.class2_sp1 == "1" & pbl.w.class2_sp2 == "2" ~ "short:med",
-    pbl.w.class2_sp1 == "2" & pbl.w.class2_sp2 == "1" ~ "short:med",
-    pbl.w.class2_sp1 == "1" & pbl.w.class2_sp2 == "3" ~ "short:long",
-    pbl.w.class2_sp1 == "3" & pbl.w.class2_sp2 == "1" ~ "short:long",
-    pbl.w.class2_sp1 == "2" & pbl.w.class2_sp2 == "3" ~ "med:long",
-    pbl.w.class2_sp1 == "3" & pbl.w.class2_sp2 == "2" ~ "med:long")
-    ) %>%
-  left_join(bb_abundance2_agg, by = c("site", "bb.sp1" = "bb.sp")) %>%
-  rename(bb.sp1.abund = bb.sp.abund) %>%
-  left_join(bb_abundance2_agg, by = c("site", "bb.sp2" = "bb.sp")) %>%
-  rename(bb.sp2.abund = bb.sp.abund) %>%
-  unite(sp.pair, c(bb.sp1, bb.sp2), sep = "_", remove = FALSE) %>%
-  left_join(site_data) %>%
-  select(site, elev.mean, sp.pair, bb.sp1, bb.sp2, bb.sp1.abund, 
-         bb.sp2.abund, niche.overlap, same.class) %>%
-  mutate(log.bb.sp1.abund = log(bb.sp1.abund),
-         log.bb.sp2.abund = log(bb.sp2.abund))
+write_csv(niche_overlap, "../hpc_packets/niche_overlap.csv")
+
+# # Aggregated niche overlap per species pair per site
+# niche_overlap2 <- read_csv("../data/processed_data/network.csv") %>%
+#   filter(caste != "male") %>% # we will ignore males 
+#   group_by(site, bb.sp, plant.sp) %>%
+#   summarize(visits = n()) %>%
+#   niche_overlapper2(plant.sp) %>%
+#   left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>%
+#   rename(pbl.w.class2_sp1 = pbl.w.class2) %>%
+#   left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>%
+#   rename(pbl.w.class2_sp2 = pbl.w.class2) %>%
+#   mutate(same.class = case_when(
+#     pbl.w.class2_sp1 == "short" & pbl.w.class2_sp2 == "short" ~ "short:short",
+#     pbl.w.class2_sp1 == "med" & pbl.w.class2_sp2 == "med" ~ "med:med",
+#     pbl.w.class2_sp1 == "long" & pbl.w.class2_sp2 == "long" ~ "long:long",
+#     pbl.w.class2_sp1 == "short" & pbl.w.class2_sp2 == "med" ~ "short:med",
+#     pbl.w.class2_sp1 == "med" & pbl.w.class2_sp2 == "short" ~ "short:med",
+#     pbl.w.class2_sp1 == "short" & pbl.w.class2_sp2 == "long" ~ "short:long",
+#     pbl.w.class2_sp1 == "long" & pbl.w.class2_sp2 == "short" ~ "short:long",
+#     pbl.w.class2_sp1 == "med" & pbl.w.class2_sp2 == "long" ~ "med:long",
+#     pbl.w.class2_sp1 == "long" & pbl.w.class2_sp2 == "med" ~ "med:long")
+#     ) %>%
+#   left_join(bb_abundance2_agg, by = c("site", "bb.sp1" = "bb.sp")) %>%
+#   rename(bb.sp1.abund = bb.sp.abund) %>%
+#   left_join(bb_abundance2_agg, by = c("site", "bb.sp2" = "bb.sp")) %>%
+#   rename(bb.sp2.abund = bb.sp.abund) %>%
+#   unite(sp.pair, c(bb.sp1, bb.sp2), sep = "_", remove = FALSE) %>%
+#   mutate(site = factor(site)) %>%
+#   select(site, bb.sp1, bb.sp2, sp.pair, niche.overlap, same.class, bb.sp1.abund, bb.sp2.abund) %>%
+#   left_join(site_data) %>%
+#   ungroup() %>%
+#   mutate(elev0 = scale(elev.mean)) %>%
+#   select(site, elev0, sp.pair, bb.sp1, bb.sp2, bb.sp1.abund, 
+#          bb.sp2.abund, niche.overlap, same.class) %>%
+#   mutate(log.bb.sp1.abund = log(bb.sp1.abund),
+#          log.bb.sp2.abund = log(bb.sp2.abund))
 
@@
 ```{r}
-ggplot(niche_overlap, aes(bb.sp1.abund, bb.sp2.abund, color = niche.overlap)) +
-  geom_point()
-
-ggplot(niche_overlap, aes(bb.sp1.abund, niche.overlap)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 4))
-
-ggplot(niche_overlap, aes(bb.sp2.abund, niche.overlap)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 4))
-
-ggplot(niche_overlap, aes(log.bb.sp1.abund, niche.overlap)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 4))
-
-ggplot(niche_overlap, aes(log.bb.sp2.abund, niche.overlap)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 4))
-
-ggplot(niche_overlap, aes(log.bb.sp2.abund, niche.overlap, color = same.class)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 4))
-
-ggplot(niche_overlap, aes(sqrt(bb.sp2.abund), sqrt(niche.overlap), color = same.class)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 4))
-
-ggplot(niche_overlap, aes(bb.sp2.abund, niche.overlap, color = same.class)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 4))
-
-ggplot(niche_overlap, aes(log.bb.sp1.abund, log.bb.sp2.abund, color = niche.overlap)) +
-  geom_point()
-
-# Niche overlap becomes rather binary when one or both species is rare. This is driven especially by singletons.
-ggplot(niche_overlap, aes(sp.pair, niche.overlap)) +
-  geom_violin() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-ggplot(niche_overlap, aes(same.class, niche.overlap)) +
-  geom_violin() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-# Let's see if we can improve matters by dropping singletons
-ggplot(filter(niche_overlap, bb.sp1.abund >= 2 & bb.sp1.abund >= 2), 
-       aes(sp.pair, niche.overlap)) +
-  geom_violin() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-ggplot(filter(niche_overlap, bb.sp1.abund >= 2 & bb.sp1.abund >= 2), 
-       aes(same.class, niche.overlap)) +
-  geom_violin() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-# Now a threshold of 3
-ggplot(filter(niche_overlap, bb.sp1.abund >= 3 & bb.sp1.abund >= 3), 
-       aes(sp.pair, niche.overlap)) +
-  geom_violin() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-ggplot(filter(niche_overlap, bb.sp1.abund >= 3 & bb.sp1.abund >= 3), 
-       aes(same.class, niche.overlap)) +
-  geom_violin() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-# Now a threshold of 5...bye, all the species pairs...
-ggplot(filter(niche_overlap, bb.sp1.abund >= 5 & bb.sp1.abund >= 5), 
-       aes(same.class, niche.overlap)) +
-  geom_violin() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-ggplot(filter(niche_overlap, bb.sp1.abund >= 10 & bb.sp1.abund >= 10), 
-       aes(same.class, niche.overlap)) +
-  geom_violin() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-# Let's see if there is any influence of site, elevation, yday
-ggplot(filter(niche_overlap, bb.sp1.abund >= 3 & bb.sp1.abund >= 3), 
-       aes(site, niche.overlap)) +
-  geom_boxplot() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-ggplot(filter(niche_overlap, bb.sp1.abund >= 3 & bb.sp1.abund >= 3), 
-       aes(elev.mean, niche.overlap)) +
-  geom_point() +
-  geom_smooth() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-ggplot(filter(niche_overlap, bb.sp1.abund >= 3 & bb.sp1.abund >= 3), 
-       aes(yday, niche.overlap)) +
-  geom_point() +
-  geom_smooth() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-# Abundance relatonships
-ggplot(niche_overlap, aes(log(bb.sp1.abund), log(bb.sp2.abund))) +
-  geom_point() +
-  geom_smooth(se = FALSE) +
-  facet_wrap(~sp.pair)
-
-#
-ggplot(niche_overlap, aes(same.class, niche.overlap)) +
-  geom_boxplot()
-
-ggplot(niche_overlap, aes(sp.pair, niche.overlap)) +
-  geom_boxplot() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-# Aggregate by site
-ggplot(niche_overlap2, aes(bb.sp1.abund, niche.overlap)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))
-
-ggplot(niche_overlap2, aes(log.bb.sp1.abund, niche.overlap)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))
-
-ggplot(niche_overlap2, aes(log.bb.sp1.abund, niche.overlap)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
-  facet_wrap(~same.class)
-
-ggplot(niche_overlap2, aes(log.bb.sp1.abund, niche.overlap)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
-  facet_wrap(~sp.pair)
-
-ggplot(niche_overlap2, aes(sp.pair, niche.overlap)) +
-  geom_violin() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-ggplot(niche_overlap2, aes(sp.pair, niche.overlap)) +
-  geom_boxplot() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-ggplot(niche_overlap2, aes(same.class, niche.overlap)) +
-  geom_violin() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-ggplot(niche_overlap2, aes(elev.mean, niche.overlap)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +
-  facet_wrap(~sp.pair)
+# Not aggregated
+ggplot(niche_overlap, aes(niche.overlap)) +
+  geom_histogram()
 
@@
 
-## raw visits aggregate
-ggplot(niche_overlap3, aes(log.bb.sp1.abund, niche.overlap)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))
-
-ggplot(niche_overlap3, aes(same.class, niche.overlap)) +
-  geom_violin() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+ggplot(niche_overlap, aes(sp.pair, niche.overlap)) +
+  geom_boxplot() +
+  facet_wrap(~same.class)
+
+ggplot(niche_overlap, aes(same.class, niche.overlap)) +
+  geom_boxplot()
+
+# Aggregate by site
+ggplot(niche_overlap2, aes(niche.overlap)) +
+  geom_histogram()
 
@@
   facet_wrap(~sp.pair)
-
 ```
@@
 ```{r}
-brm_niche00 <- brm(niche.overlap ~ bb.sp1.abund*bb.sp2.abund + 
-                     (1 | sp.pair) + 
-                     (1 | same.class) + 
-                     (1 | site),
-                   control = list(adapt_delta = 0.90),
-                   chains = 4,
-                   cores = 4,
-                   iter = 2000,
-                   file = "../output/brm_niche00",
-                   data = niche_overlap)
-
-parnames(brm_niche00)
-conditional_effects(brm_niche00)
-mcmc_plot(brm_niche00, pars = c("^r_sp"))
-mcmc_plot(brm_niche00, pars = c("^r_sa"))
-mcmc_plot(brm_niche00, pars = c("^b_"))
-hypothesis(brm_niche00, "dprime = 0")
-pp_check(brm_niche00)
-bayes_R2(brm_niche00)
-
-# To stabilize niche overlap (avoid lots of  1s and 0s), we will filter the data to include only species pairs in which both species were represented by at least 3 observations.
-brm_niche01 <- brm(niche.overlap ~ bb.sp1.abund*bb.sp2.abund + 
-                     (1 | sp.pair) + 
-                     (1 | same.class) + 
-                     (1 | site),
-                   chains = 4,
-                   cores = 4,
-                   iter = 2000,
-                   file = "../output/brm_niche01",
-                   data = filter(niche_overlap, bb.sp1.abund >= 2 & bb.sp1.abund >= 2))
-
-parnames(brm_niche01)
-conditional_effects(brm_niche01)
-mcmc_plot(brm_niche01, pars = c("^r_sp"))
-mcmc_plot(brm_niche01, pars = c("^r_sa"))
-mcmc_plot(brm_niche01, pars = c("^b_"))
-hypothesis(brm_niche01, "dprime = 0")
-pp_check(brm_niche01)
-bayes_R2(brm_niche01)
-
-# Would it help if we log-transformed the bb abundance data and stuck to unfiltered data?
-brm_niche02 <- brm(niche.overlap ~ log.bb.sp1.abund*log.bb.sp2.abund + 
-                     (1 | sp.pair) + 
-                     (1 | same.class) + 
-                     (1 | site),
-                   #control = list(adapt_delta = 0.90),
-                   chains = 4,
-                   cores = 4,
-                   iter = 2000,
-                   file = "../output/brm_niche02",
-                   data = niche_overlap)
-
-parnames(brm_niche02)
-conditional_effects(brm_niche02)
-mcmc_plot(brm_niche02, pars = c("^r_sp"))
-mcmc_plot(brm_niche02, pars = c("^r_sa"))
-mcmc_plot(brm_niche02, pars = c("^b_"))
-hypothesis(brm_niche02, "dprime = 0")
-pp_check(brm_niche02)
-bayes_R2(brm_niche02)
-
-brm_niche03 <- brm(niche.overlap ~ log.bb.sp1.abund*log.bb.sp2.abund + 
-                     (1 + log.bb.sp1.abund*log.bb.sp2.abund | sp.pair) + 
-                     (1 | same.class) + 
-                     (1 | site),
-                   #control = list(adapt_delta = 0.90),
-                   chains = 4,
-                   cores = 4,
-                   iter = 2000,
-                   file = "../output/brm_niche03",
-                   data = niche_overlap)
-
-parnames(brm_niche02)
-conditional_effects(brm_niche02)
-mcmc_plot(brm_niche02, pars = c("^r_sp"))
-mcmc_plot(brm_niche02, pars = c("^r_sa"))
-mcmc_plot(brm_niche02, pars = c("^b_"))
-hypothesis(brm_niche02, "dprime = 0")
-pp_check(brm_niche02)
-bayes_R2(brm_niche02)
-
-# A zero-one inflated model might be the key.
-brm_niche04 <- brm(niche.overlap ~ log.bb.sp1.abund*log.bb.sp2.abund + 
-                     (1 | sp.pair) + 
-                     (1 | same.class) + 
-                     (1 | site),
-                   #control = list(adapt_delta = 0.90),
-                   family = "zero_one_inflated_beta",
-                   chains = 4,
-                   cores = 4,
-                   iter = 2000,
-                   file = "../output/brm_niche04",
-                   data = niche_overlap)
-
-parnames(brm_niche04)
-conditional_effects(brm_niche04)
-mcmc_plot(brm_niche04, pars = c("^r_sp"))
-mcmc_plot(brm_niche04, pars = c("^r_sa"))
-mcmc_plot(brm_niche04, pars = c("^b_"))
-hypothesis(brm_niche04, "dprime = 0")
-pp_check(brm_niche04)
-bayes_R2(brm_niche04)
-
-# Aggregate models
-brm_niche60 <- brm(niche.overlap ~ log.bb.sp1.abund*log.bb.sp2.abund + 
+# Not aggregated
+brm_niche10 <- brm(niche.overlap ~ log.bb.sp1.abund*log.bb.sp2.abund + 
                      (1 | sp.pair) + 
@@
                    iter = 2000,
-                   file = "../output/brm_niche60",
-                   data = niche_overlap2)
-
-parnames(brm_niche60)
-conditional_effects(brm_niche60)
-mcmc_plot(brm_niche60, pars = c("^r_sp"))
-mcmc_plot(brm_niche60, pars = c("^r_sa"))
-mcmc_plot(brm_niche60, pars = c("^b_Int"))
-hypothesis(brm_niche60, "dprime = 0")
-pp_check(brm_niche60)
-bayes_R2(brm_niche60)
-
-brm_niche61 <- brm(niche.overlap ~ log.bb.sp1.abund*log.bb.sp2.abund + 
-                     (1 | sp.pair) + 
-                     (1 | same.class) +
-                     (1 | site),
-                   control = list(adapt_delta = 0.99),
-                   family = "zero_one_inflated_beta",
-                   chains = 4,
-                   cores = 4,
-                   iter = 2000,
-                   file = "../output/brm_niche61", # for local
-                   #file = "./output/brm_niche61", # for HPC
-                   data = niche_overlap2)
-
-parnames(brm_niche61)
-conditional_effects(brm_niche61)
-mcmc_plot(brm_niche61, pars = c("^r_sp"))
-mcmc_plot(brm_niche61, pars = c("^r_sa"))
-mcmc_plot(brm_niche61, pars = c("^b_Int"))
-pp_check(brm_niche61)
-bayes_R2(brm_niche61)
-
-brm_niche62 <- brm(niche.overlap ~ log.bb.sp1.abund*log.bb.sp2.abund + 
-                     (1 + log.bb.sp1.abund*log.bb.sp2.abund | sp.pair) + 
-                     (1 | same.class) +
-                     (1 | site),
-                   control = list(adapt_delta = 0.99),
-                   family = "zero_one_inflated_beta",
-                   chains = 4,
-                   cores = 4,
-                   iter = 2000,
-                   #file = "../output/brm_niche62", # for local
-                   file = "./output/brm_niche62", # for HPC
-                   data = niche_overlap2)
+                   #file = "../output/brm_niche10",
+                   file = "./brm_niche10",
+                   data = niche_overlap)
 ```
@@
   left_join(bb_traits) %>%
-  mutate(pbl.w.class2 = factor(pbl.w.class2)) %>%
-  mutate(pbl.w.class2 = case_when(
-    pbl.w.class2 == "1" ~ "short",
-    pbl.w.class2 == "2" ~ "med",
-    pbl.w.class2 == "3" ~ "long"
-    ))
+  mutate(pbl.w.class2 = factor(pbl.w.class2))
 
@@
 ```{r}
-# Fl traits with abbreviated names
-fl_traits2 <- fl_traits %>%
-  left_join(read_csv("../data/processed_data/network.csv")) %>%
-  select(plant.sp, plant.sp.abb, plant.genus, plant.family, k.type.ss, color) %>%
-  mutate(k.type.ss = factor(k.type.ss)) %>%
-  distinct()
-
 # Get visitation
@@
 ```
-```
+
+```{r}
+# # BB spp. pooled, only interested in comparing ktypes here
+# brm_pvis00 <- brm(visits ~ bb.abund*prop.flower.cover +
+#                     (1 + prop.flower.cover | ktype/site),
+#                   family = "negbinomial",
+#                   file = "../output/brm_pvis00",  # for running locally
+#                   #file = "brm_pvis00", # for running on the HPC
+#                   chains = 4,
+#                   cores = 4,
+#                   iter = 2000,
+#                   data = prop_visits_ktype_pooled)
+# 
+# get_variables(brm_pvis00)
+# pp_check(brm_pvis00, type='stat', stat='mean')
+# pp_check(brm_pvis00, type = "error_scatter_avg")
+# pp_check(brm_pvis00,  x = 'prop.flower.cover', type = "error_scatter_avg_vs_x")
+# pp_check(brm_pvis00,  x = 'bb.abund', type = "error_scatter_avg_vs_x")
+# 
+# 
+# bayes_R2(brm_pvis00)
+# conditional_effects(brm_pvis00)
+# 
+# mcmc_plot(brm_pvis00, pars = c("^r_ktype:site.*0_.*Intercept]$"))
+# mcmc_plot(brm_pvis00, pars = c("^r_ktype:site.*1_.*Intercept]$"))
+# mcmc_plot(brm_pvis00, pars = c("^r_ktype:site.*2_.*Intercept]$"))
+# mcmc_plot(brm_pvis00, pars = c("^r_ktype:site.*3_.*Intercept]$"))
+# mcmc_plot(brm_pvis00, pars = c("^r_ktype:site.*4_.*Intercept]$"))
+# mcmc_plot(brm_pvis00, pars = c("^r_ktype:site.*5_.*Intercept]$"))
+# mcmc_plot(brm_pvis00, pars = c("^r_ktype:site.*6_.*Intercept]$"))
+# mcmc_plot(brm_pvis00, pars = c("^r_ktype:site.*7_.*Intercept]$"))
+# 
+# mcmc_plot(brm_pvis00, pars = c("^r_ktype:site.*0_.*prop.flower.cover]$"))
+# mcmc_plot(brm_pvis00, pars = c("^r_ktype:site.*1_.*prop.flower.cover]$"))
+# mcmc_plot(brm_pvis00, pars = c("^r_ktype:site.*2_.*prop.flower.cover]$"))
+# mcmc_plot(brm_pvis00, pars = c("^r_ktype:site.*3_.*prop.flower.cover]$"))
+# mcmc_plot(brm_pvis00, pars = c("^r_ktype:site.*4_.*prop.flower.cover]$"))
+# mcmc_plot(brm_pvis00, pars = c("^r_ktype:site.*5_.*prop.flower.cover]$"))
+# mcmc_plot(brm_pvis00, pars = c("^r_ktype:site.*6_.*prop.flower.cover]$"))
+# mcmc_plot(brm_pvis00, pars = c("^r_ktype:site.*7_.*prop.flower.cover]$"))
+# 
+# mcmc_plot(brm_pvis00, pars = c("r_ktype[0,Intercept]",
+#                                "r_ktype[1,Intercept]",
+#                                "r_ktype[2,Intercept]",
+#                                "r_ktype[3,Intercept]",
+#                                "r_ktype[4,Intercept]",
+#                                "r_ktype[5,Intercept]",
+#                                "r_ktype[6,Intercept]",
+#                                "r_ktype[7,Intercept]"), fixed =  TRUE)
+# 
+# mcmc_plot(brm_pvis00, pars = c("r_ktype[0,prop.flower.cover]",
+#                                "r_ktype[1,prop.flower.cover]",
+#                                "r_ktype[2,prop.flower.cover]",
+#                                "r_ktype[3,prop.flower.cover]",
+#                                "r_ktype[4,prop.flower.cover]",
+#                                "r_ktype[5,prop.flower.cover]",
+#                                "r_ktype[6,prop.flower.cover]",
+#                                "r_ktype[7,prop.flower.cover]"), fixed =  TRUE)
+# 
+# 
+# brm_pvis03 <- brm(visits ~ bb.abund*prop.flower.cover +
+#                     (1 + prop.flower.cover | ktype),
+#                   family = "negbinomial",
+#                   file = "../output/brm_pvis03",  # for running locally
+#                   #file = "brm_pvis03", # for running on the HPC
+#                   chains = 4,
+#                   cores = 4,
+#                   iter = 2000,
+#                   data = prop_visits_ktype_pooled)
+# 
+# pp_check(brm_pvis03)
+# bayes_R2(brm_pvis03)
+# conditional_effects(brm_pvis03)
+# mcmc_plot(brm_pvis03, pars = c("^r_k"))
+#
+
+```

### 2021-02-05 writing/manuscript_v4.Rmd
developed ordination analysis
pretty much finalized analysis, almost done ms

@@
 ---
-title: "Bumble bee foraging bias"
+title: "Bias and lability in the association between bumble bee visitation floral morphology along a montane-to-alpine elevation gradient"
 author:
@@
 ## 3.3. Consequences of bias for niche relationships
+```{r fig4, echo = FALSE,  fig.cap = paste("Principal coordinates ordination of bumble bee visitation pooled across all samples.")}
+knitr::include_graphics("../output/pcoa_aggregate.png")
+```
 

### 2021-02-05 writing/manuscript_v5.Rmd
pretty much finalized analysis, almost done ms

---
title: "Bumble bees exhibit strong but flexible foraging biases through temporal and elevational gradients of floristic turnover"
author:
- Douglas B. Sponsler
- Fabrice Requier
- Katharina Kallnik
- Alice Claßen
- A. Fabienne Maihoff
- Ingolf Steffan-Dewenter
bibliography: refs.bib
output:
  html_document:
    df_print: paged
  pdf_document:
    fig_caption: yes
---

# 1. Introduction

Bumble bees are exceptionally versatile foragers whose ability to solve the puzzles posed by complex flower morphologies [@Plowright1984-lc], or indeed by human investigators [@Alem2016-rn; @Loukola2017-bt], renders almost any flower potentially accessible. Even when a gross mismatch between tongue length and corolla depth makes the nectar of a given flower inaccessible by ordinary means, some bumble bee species will resort readily to biting through the base of the flower to “rob” nectar, incidentally making the nectar available to other prospective foragers as well. Flower handling skills are, moreover, socially transmitted, amplifying the plasticity of individual behavior [@Leadbeater2008-rh; @Loukola2017-bt]. Nevertheless, the flexibility of bumble bee foraging is tuned to floral reward, which depends on both the intrinsic reward of a given flower and the searching/learning/handling costs associated with exploiting it. At an evolutionary level, the innate biases of bumble bees for certain floral traits likely reflect historic associations between floral traits and floral rewards [@Raine2007-nq]. These innate biases, however, can be modulated or even nullified by patterns of reward learned during foraging [@Kugler1943-uy; Maharaj2019-wm], which depend not only on intrinsic floral properties but also ecological modulators, such as the spatial distribution of floral patches or competition with other bumble bees [@Ishii2013-sm; @Plowright1984-lc; @Sowig1989-ld].

The topic of floral bias in bumble bees has been approached via both laboratory experimentation [e.g. @Kugler1943-uy; @Raine2007-nq] and observational field study [e.g. Heinrich1976-wc; @Goulson2008-zy]. The former approach enables stronger causal inference with respect to the mechanisms of floral choice, but it suffers from all the usual limitations of biological realism associated with laboratory conditions. Observational field studies, however, tend to be anecdotal in scope, typically describing bumble bee foraging patterns in particular localities, over short time frames, and with limited floristic data to enable a distinction between true foraging bias and the neutral effects of floral abundance. The steep elevation gradients of mountain slopes provide an opportunity for more powerful observational study of bumble foraging because they generate broad gradients of environmental variation while controlling for confounding effects of geographic variation, such as genotypic differentiation within bumble bee species. When bumble bee visitation in such environments is sampled with sufficient spatial and temporal replication and complemented with quantitative floral surveying, true bias can be distinguished from the neutral effects of abundance, and variation in bias can be explored along elevational and temporal dimensions of floristic turnover.

In the present study, we focus on the question of bumble bee foraging bias with respect to floral morphology, and the lability of morphological biases across gradients of floristic variation. At the outset, we emphasize the use of the term "bias" in favor of the term "preference", since the latter is laden with more assumptions and its inference under field conditions is dubious []. By "bias", we mean simply an empirical disproportionality between the relative abundance of a given floral taxon or group and its proportional visitation. The exact mechanism of this disproportionality --- whether an arbitrary "preference" on the part of the foraging bee, an unmeasured decoupling of relative abundance and rate of reward, an exogenous effect of competition, etc. --- is beyond the scope of our discussion, and probably best studied under laboratory conditions where strict factorial control is possible. We focus instead on the empirical phenomenon of bias and its morphological correlations.

The meaning of "morphology", however, is hardly more clear-cut than that of bias. Bumble bees and flowers vary in an array of interrelated morphological parameters, both within and between species, and the selection of relevant and measurable morphological metrics is nontrivial. For bumble bees, morphological variation consists principally (though not exclusively) of variation in tongue length. Tongue length in bumble bees corresponds functionally to corolla depth in flowers, and there is a well-established (though imperfect) association between tongue length and floral choice [e.g. @Inouye1980-mv]. Importantly, though, accessing floral reward involves more than just a match between tongue length and corolla depth. Such matching is entirely bypassed by nectar robbers, but even in the case of "legitimate" foraging, access to reward involves learning to manipulate floral structures that, in some floral morphotypes, are challenging irrespective of tongue length [@Laverty1994-hm]. For example, the characteristic Papilionaceous floral morphology of legumes requires a bee to separate tightly appressed keel petals to access anthers and nectaries [@Westerkamp1997-hs], and the narrowly campanulate flowers of many Ericaceous flora pose a stringent barrier to large-bodied visitors like bumble bees [@Wallace1977-kz]. Rather than considering various floral morphological traits independently, we will rely on the synthesis of @Kugler1970-gb, who developed a typology consisting of ten primary morphotypes and nested subdivision of each. Kugler's typology was informed by laboratory studies of bumble bee flower choice [@Kugler1943-uy], and there is a precedent for analyzing bumble bee visitation in terms of Kugler's morphotypes [@Neumayer1999-il; @Schneller2014-ry].

First, we compare the overall rate of visitation across floral morphotypes. Then, we parse visitation rates by bumble bee species and explore the functional relationship between floral morphotype and bumble bee tongue length. We then compare overall diet similarity across bumble bee species as an emergent consequence of their respective biases. Lastly, we analyze the flexibility of the morphological preferences of individual bumble bee species through elevation and time.

# 2. Methods

## 2.1 Field system

### Study sites

The study was conducted in Berchtesgaden National Park (47.55°N, 12.92°E), located in the Northern Limestone Alps of southeast Germany. The landscape is composed of mountain pastures mainly surrounded by coniferous forests. We selected 25 study sites (60 x 60 m) on mountain pastures at elevations ranging from 641-2032 m above sea level. Fourteen of these pastures are extensively grazed by cattle or sheep, three are mowed for hay production, and eight have been abandoned throughout the last century and are no longer subject to any human management. The sites sampled in this study are the same as those used in a series of previous studies [@Hoiss2012-sn; @Hoiss2012-kr; Hoiss2015-gv], but the data reported here are otherwise independent.

### Bumble bee visitation

Sampling consisted of repeated visits (hereafter, "visitation samples") to each study site at approximately weekly intervals. Samples were only collected on dry days when the air temperature was at least 6°C. During each visit, bumble bees and their activity were recorded during a 50-minute transect walk. Bumblebees observed on or a given flower were counted as floral visitors. Bumble bee queens were identified to species level in the field, whereas workers and males were stored in individually labeled tubes in the freezer for later identification in the laboratory after @Amiet1996-jz. While floral visitation by male bumble bees was recorded during visitation sampling, we chose to analyze only visitation by queens and workers. During observation, it was noted whether each bee was visibly carrying pollen at the time of its visit, but this provides at best an imperfect distinction between pollen and nectar foragers, and we opted not to differentiate between bees with and without pollen in our analyses.

### Floral surveying

floral survey sample = site-date of floral surveying

## 2.2. Data analysis

### 2.2.1 Bumble bee and floral morphology data

Bumble bee tongue-length data were obtained from the literature values compiled and summarized by @Arbetman2017-xj. Values for *B. gerstaeckeri* and *B. mendax*, which were not included in @Arbetman2017-xj, were obtained from @Obeso1992-yp and @Durieux2000-uk, respectively. We then binned bumble bee species into three tongue-length classes: short (< 7.5 mm), medium (7.5---10.0 mm), and long (>= 10.0 mm). Short-tongued species included *B. jonellus*, *B. lapidarius*, *B. pratorum*, *B. psithyrus-group*, *B. pyrenaeus*, and *B. terrestris/lucorum*. Medium-tongued species included *B. humilis*, *B. hypnorum*, *B. monticola*, *B. mucidus*, *B. pascuorum*, and *B. wurflenii*. Long-tongued species included *B. gerstaeckeri*, *B. hortorum*, and *B. mendax*.

Floral morphology was summarized using floral morphotypes of @Kugler1970-gb, accessed via the BIOFLOR database [@Klotz2002-uc]. Morphotypes were simplified to the basic categories numbered 0-10: 0 = not applicable (includes wind-pollinated flora such as grasses), 1 = disc- and bowl-shaped flowers, 2 = funnel flowers, 3 = bell-shaped flowers, 4 = stalk-disc flowers, 5 = lip flowers, 6 = flag flowers, 7 = flower heads, 8 = spike flowers, 9 = brush flowers, 10 = trap flowers **(Figure X)**. <!-- I think I would like to include a figure with photographs of flowers exemplifying each morphotype --> In a few cases, a plant species in our data set was not found in the BIOFLOR database. In most cases, morphotypes are constant at the genus level, and missing morphotypes were added based on the morphotypes found for congeners. The genus *Gentiana* consisted of both group 2 and group 4 representatives, so the missing species *Gentiana aspera* and *Gentiana ciliata* were classified based on visual comparison with their classified congeners. In the case of the genus *Phyteuma*, we followed @Neumayer1999-il and classified it as a group 7 flower head.

### 2.2.2 Processing visitation and survey data

Prior to analysis, we aligned each visitation sample with a corresponding floral survey sample. In most cases, floral surveying was conducted jointly with bumble bee observation for each site and date. Occasionally, though, floral surveying could not be completed on the same day as bumble bee observation, resulting in a small temporal offset between visitation samples and floral survey samples. In these cases, visitation samples were paired with the nearest available floral survey data within a maximum distance of 7 days. Visitation samples lacking corresponding floral survey data within 7 days were omitted from further analysis. In a few cases, two consecutive visitation samples at a given site shared the same nearest floral survey sample; in these cases, the visitation samples were first pooled, then aligned to their shared nearest floral survey sample. In rare cases, a given floral species was recorded during visitation sampling but missed during floral surveying. These species were retroactively assigned the minimum floral area value of 0.01 m^2.

Visitation data were extremely right-skewed because the majority of available flora received zero visits during a given sampling period. To facilitate model fitting, we created a binary version of visitation rate in which visited flora were assigned to 1 and unvisited flora to 0. We then modeled binary visitation using generalized models with binomial or Bernoulli distribution families. The conditional effects inferred from these models should, therefore, be interpreted as effects on *visitation probability* rather than visitation rate.

### 2.2.3 Model 1: Comparing overall visitation probability across floral morphotypes

We first compared overall visitation probability across floral morphotypes. In this model, visitation for each sample was pooled across bumble bee species and within floral morphotypes prior to binarization. Visitation probability was then modeled using multilevel Bayesian regression with a Bernoulli distribution family (logit link function). *Bumble bee abundance* and *percent flower cover* (per morphotype) were included as a global interaction term, thus conditioning visitation probability on abundance and enabling the inference of bias. *Floral morphotype* was specified as a varying effect term with both a varying intercept and a varying slope with respect to *percent flower cover*, and a varying intercept term was also included for *site*. See Appendix X for the details of model specification and evaluation.

### 2.2.4 Model 2: Visitation probability as a function of floral morphotype and bumble bee species

To explore the functional underpinnings of bumble foraging bias, we analyzed visitation probability as a function of the interaction between floral morphology and bumble bee tongue length class. For this analysis, visitation was pooled within floral morphotype but not across bumble bee species. We then specified a Bayesian model of the same form as Model 1, but this time included a varying slope and intercept term for *floral morphotype : bumble bee tongue length class* and *floral morphotype : bumble bee species*. In this model, the relationship between floral morphotype and bumble bee tongue length class was the focus of inference, and the bumble bee species effect was added to avoid pseudoreplication arising from the grouping of observation within species. See Appendix X for the details of model specification and evaluation.

### 2.2.4 Model 3: Visitation probability as a function of floral morphotype and bumble bee tongue length

To focus on species-level morphotype biases, we removed the tongue length effect from Model 2 and re-fit the model considering only bumble bee species. See Appendix X for the details of model specification and evaluation.

### 2.2.5 Model 4: Aggregate diet ordination

To visualize the net diet similarity between bumble bee species, we pooled visitation data across all samples within bumble bee species and plant species, then performed a principal coordinates analysis based on the Jaccard distance metric. To test the significance of tongue length class as a predictor of diet similarity, we performed a permutation test.

### 2.2.6 Model 5: The variation of morphological bias through elevation and time

To explore the plasticity of floral biases, we first subsetted our data to include only the dominant bumble bee species (*B. hortorum*, *B. pascuorum*, *B. pratorum*, *B. soroensis*, *B. terrestris/lucorum*, and *B. wurflenii*) and the most visited floral morphotypes (groups 3, 5, 6, and 7). We then constructed a generalized additive model in which binary visitation was modeled as a tensor-product smooth of *elevation* and *day-of-year*, grouped by floral morphotype. This response was conditioned on linear interaction terms for *bumble bee abundance* : *percent flower cover* and *percent flower cover* : *floral morphotype* and a random intercept term for *bumble bee species* : *floral morphotype*. See Appendix X for the details of model specification and evaluation.

### 2.2.7 Software

All analyses were conducted in R v. 3.6. Bayesian multilevel models were implemented in *Stan* [@Stan2020aa] via the R package *brms* [@Buerkner2018aa], and visualizations were constructed with the R packages *tidybayes* [@Kay2020aa] and *ggdist* [Kay2020ab]. Ordination analysis and permutation testing were performed with the R package *vegan* [@Oksanen2019aa] and visualized using the R packages *ggplot2* [@Wickham2016aa] and *ggvegan* [@Simpson2019aa]. GAMs were implemented with the R package *mgcv* [@Wood2017aa] and visualized with the R package *mgcViz* [@Fasiolo2018aa]. Data handling was performed the *tidyverse* suite of R packages [Wickham2019aa]. Input data and code are available in Appendix X.

# 3. Results
## 3.1 Model 1: Overall visitation probability by floral morphotype

Since we report the results of a Bernoulli regression model, the term "visitation rate" used here should be interpreted as conditional visitation *probability*, not conditional visitation frequency.

Lip (group 5) and flag (group 6) flowers had the highest overall visitation probability, with a conditional mean visitation probabilities of 0.36 (95% CI = 0.30--0.42) and 0.33 (95% CI = 0.27--0.39), respectively **(Figure X)**. In second place, head flowers (group 7) had a mean conditional visitation probability of 0.19 (95% CI = 0.16--0.24). Bell (group 3) flowers were a distant third with a conditional mean visitation probability of 0.092 (95% CI = 0.07--0.12), followed by stalk-disc (group 4), funnel (group 2), and disc flowers (group 1) with conditional visitation probabilities of 0.07 (95% CI = 0.05--0.09), 0.05 (95% CI = 0.04--0.07), and 0.05 (95% CI = 0.03--0.07), respectively. Unclassified (group 0) flowers, consisting mainly of wind-pollinated grasses, were the least-visited, with a conditional mean visitation probability of 0.3 (95% CI = 0.02--0.04). Our model explained approximately 35% of the total variance (97.5% CI = 34--36%).

```{r fig1, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability by floral morphotype. Point-interval plots depict 66\\% and 95\\% CIs.")}
knitr::include_graphics("../output/brm_pvis01_plot1.png")
```

## 3.2 Models 2 and 3: Visitation probability by flower morphotype, bumble bee species, and bumble bee tongue length class

Descriptions of tongue classes are based on Model 2 and refer to **(Figure X)**, and descriptions of species are based on Model 3 and refer to **(Figure X)**. Both models explained approximately 28% of total variance (97.5% CI = 27--29%). As in Model 1, visitation rates refer to conditional visitation probabilities.

Disc (group 1) flowers were visited at similar rates across tongue length classes. At the species level, disc flowers were visited at relatively high rates by the short-tongued *B. terrestris/lucorum* and the medium-tongued *Bombus monticola*. Visitation to funnel (group 2) flowers was dominated by long-tongued species, an effect driven entirely by the *Aconitum* (group 2) specialist *B. gerstaeckeri*. Bell (group 3) flowers were visited at a higher rate by short-tongued species than by long-tongued species, with medium-tongued species exhibiting intermediate visitation rates. At the species-level, bell flowers were visited at highest probability by the short-tongued species *B. jonellus*, *B. pyrenaeus*, and *B. soroensis*, and by the medium-tongued species *B. hypnorum* and *B. monticola*. Stalk-disc (group 4) flowers were visited at similar rates across all tongue length classes, and there was also broad overlap in visitation rates at the species level. Lip (group 5) flowers were also visited at similar rates across tongue length classes, but at the species level *B. pascuorum* exhibited an uniquely high preference for lip flowers. Flag (group 6) flowers were visited at a higher rate by medium-tongued species than by short-tongued species, with long-tongued species  exhibiting intermediate visitation rates. Flag flowers generated strong separation in visitation rate at the species level, receiving very high visitation from the medium-tongued *B. pascuorum* and *B. mucidus*, moderately high visitation from the short-tongued *B. terrestris/lucorum* and *B. lapidarius*, the medium-tongued *B. monticola*, and the long-tongued *B. mendax* and *B. hortorum*, and low visitation by the short-tongued *B. soroensis*,  *B. pyrenaeus*, *B. psithyrus*, *B. pratorum*, and *B. jonellus*, the medium-tongued *B. hypnorum*, and the long-tongued *B. gerstaeckeri*.

```{r fig2, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability by Kugler morphotype and bumble bee tongue length class. Point-interval plots depict 66\\% and 95\\% CIs. Letters indicate group differences based on evidence ratio at alpha = 0.05. Note that the range of the x-axis has been rescaled for each panel to emphasize the comparison of visitation probabilities across tongue length classes.")}
knitr::include_graphics("../output/brm_pvis00_plot1_mod.png")
```
```{r fig3, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability by Kugler morphotype and bumble bee species. The tongue length class of each species is color-coded. Point-interval plots depict 66\\% and 95\\% CIs. Note that the range of the x-axis has been rescaled for each panel to emphasize the comparison of visitation probabilities across bumble bee species.")}
knitr::include_graphics("../output/brm_pvis02_plot1.png")
```

## 3.3. Model 4: Diet ordination

Axis 1 accounted for 24% of total diet variance and primarily separated long- and medium-tongued species from short-tongued species. Long- and medium-tongued species scored low on Axis one and were associated with the lip flowers (group 5) *Ajuga reptans*, *Rhinanthus glacialis*, and *Stachys alopecuros*) and flag flowers (group 6) *Anthyllis vulneraria*, *Lotus corniculatus*, and *Trifolium pratense*. Short tongued species scored high on Axis one and were associated with the bell flowers (group 3) *Campanula scheuchzeri*, *Erica carnea*, and *Geum rivale* and the head flowers (group 7) *Carduus defloratus*, *Centaurea jacea*, and *Phyteuma orbiculare*. Axis 2 accounted for 19% of total variance and primarily explained variation within short-tongued species. Short-tongued species scoring high on Axis 2 were associated with *Campanula scheuchzeri* while those scoring low on Axis 2 were associated with *Erica carnea*. Short-tongued species scoring close to zero on Axis 2 also scored lower on Axis 1 (closer to medium- and long-tongued species) and were associated with head flowers. Tongue length class explained 22% of diet variance (Adonis permutation test, F = 1.68, p  = 0.006). While short-tongued bees appeared most broadly spread in ordination space, there were no significant differences in dispersion across tongue length classes (F = 0.62, p = 0.55).

```{r fig5, echo = FALSE,  fig.cap = paste("Principal coordinates ordination of bumble bee visitation pooled across all samples.")}
knitr::include_graphics("../output/pcoa_aggregate.png")
```

## 3.3 Model 5:

Model 5 explained 33.8% of total variance (adjusted R2 = 0.37). All tensor product smooths were significant at p < 0.001 except for the one representing *B. pratorum* and flag (group 6), which was more weakly significant at p = 0.02. As in Models 1-3, the results of Model 5 should be interpreted in terms of conditional visitation probability.

Bell (group 3) flowers were visited at highest rates early in the season and at low elevation by all species except *B. hortorum*, whose visitation rate to bell flowers peaked in mid-season and at mid-elevation. Notably, the medium-tongued *B. pascuorum* and *B. wurflenii* also exhibited relatively high visitation to bell flowers at mid-season and mid-elevation, while the visitation rates of the short-tongued *B. pratorum*, *B. soroensis*. and *B. terrestris lucorum* were more concentrated at the early-season and low-elevation peak.

Lip (group 5) flowers exhibited roughly the mirror image of the pattern of visitation rate seen in bell flowers, with peak visitation late in the season and at low elevation, but relatively high visitation extending toward mid-season and mid-elevation. This pattern was consistent for all bumble bee species.

The pattern of visitation rates to flag (group 6) flowers fell generally along the lines of tongue length class. The short-tongued *B. pratorum* and *B. soroensis* exhibited peak visitation early in the season and at low elevation. The long-tongued *B. hortorum* and the medium-tongued *B. pascuorum* and *B. wurflenii* exhibited peak visitation at mid-season, though *B. pascuorum* exhibited a bimodal pattern in elevation, with peaks around 1000 m and 1750 m, while the visitation rate of *B. hortorum* declined monotonically from high elevation to low. The short-tongued *B. terrestis/lucorum* broke the tongue-length pattern, exhibiting a diffuse pattern of visitation rate more similar to that of the long- and medium-tongued species than the other short-tongued species.

Visitation rates for head (group 7) flowers were more or less bimodal for all species, with a first peak in mid-season and low elevation and a second peak in late-season and high elevation.⎄

```{r fig6, echo = FALSE,  fig.cap = paste("GAM analysis of bias as a function of the tensor-product interaction of elevation and day of year, parsed by bumble bee species and floral morphotype and conditioned on the abundance of each bumble bee species and the relative abundance of each floral morphotype. The color ramp of each smooth has been scaled to equal color range, thereby obscuring the absolute magnitude of effect. To avoid cluttering the figure, we have opted not to include a unique color ramp key for each plot, since the focus of our inference is on relative rather than absolute patterns.")}
knitr::include_graphics("../output/gam_plot1_mod.png")
```

# 4. Discussion

## 4.1 Foraging bias and floral morphology

Our study is not the first to compare bumble bee visitation rates across Kugler's (1970) floral morphotypes [see @Neumayer1999-il; @Schneller2014-ry], but to our knowledge it is the first to enable the inference of bias across morphotypes by controlling statistically for the neutral effects of abundance.

When bumble bee visitation is pooled across all species, it is clear that lip, flag, and head flowers dominate the rest of the floral community in terms of conditional visitation probability. Patterns of pooled visitation are, of course, driven disproportionately by the behavior of the most abundant bumble bee species, which in our system were *B. hortorum*, *B. pascuorum*, *B. pratorum*, *B. psithyrus*, *B. soroensis*, *B. terrestris/lucorum*, and *B. wurflenii*. Visitation to lip and flag flowers was driven especially by the medium-tongued *B. pascuorum*, the most abundant bumble bee in our study system, and secondarily by the medium-tongued *B. wurflenii*. Visitation to head flowers was driven mainly by the short-tongued *B. soroensis* and *B. psithyrus*.

The influence of tongue length on visitation bias was pronounced for some morphotypes but not for others. In the cases of disc, and stalk-disc, and unclassified flowers, visitation rates were uniformly low across all tongue length classes. The opposite was true of lip flowers, which were visited at high rates by all tongue length classes. Bell flowers and head flowers were visited chiefly by short-tongued bees, consistent with the shallow corollas of these morphotypes. In contrast, funnel flowers, with their deep corollas were visited chiefly by long-tongued bees, though this effect was driven disproportionately by *B. gerstaeckeri*, which specializes on the funnel flowers of the genus *Aconitum*. Flag flowers were visited at the highest rate by medium-tongued species, and especially by *B. pascuorum* and *B. mucidus*.

While variation in floral bias across tongue length classes largely comports with the established relationship between tongue length and corolla depth, variation in bias *within* tongue length classes demonstrates the insufficiency of simple morphological matching to explain bumble bee foraging patterns. Nectar robbing is one mechanism by which the relationship between tongue length and corolla depth might be relaxed, but its influence in our data is unclear. Of the bumble bees in our study system, *B. terrestris/lucorum* and especially *B. wurflenii* are known to engage readily in nectar robbing. In the case of *B. wurflenii*, however, its morphotype affinities did not differ markedly from those of other medium-tongued species, and it may be that the condition of medium-tongue length alone confers enough to foraging flexibility in our study system to dampen the divergent effects of nectar robbing. Nectar robbing might, however, account for the deviation of *B. terrestris/lucorum* from most other short-tongued species with respect to flag flowers. With the exception of *B. lapidarius*, all short-tongued species besides *B. terrestris/lucorum* eschewed flag flowers, while *B. terrestris/lucorum* visited them at rates comparable to medium- and long-tongued species. Indeed, *B. terrestris/lucorum* was, on the whole, an idiosyncratic forager, demonstrating a relatively high affinity for disc flowers and unclassified flowers (particularly *Plantago lanceolata*) that were largely ignored by other species. Perhaps the most striking divergence of floral bias within tongue length class, though, was the divisive effect of bell flowers on short-tongued bumble bees. *B. pyrenaeus*, *B. soroensis*, *B. pratorum*, and *B. jonellus* visited bell flowers at high rates, while *B. terrestris/lucorum*, *B. psithyrus*, and *B. lapidarius* visited them relatively rarely. There is no obvious morphological distinction between these groups of short-tongued bees that would account for their divergent affinities for bell flowers; ecological drivers such as competition or differences in spatial or temporal activity patterns are more likely explanations.

Marked differences in the relative visitation rates of floral morphotypes compared to [@Schneller2014-ry], but also some consistencies. We reproduce their finding that *B. soroensis* is partial to bell flowers. There data collection was limited to July and August, which probably explains their finding that head (group 7) flowers were most visited, followed by lip (group 5) flowers. Bell flowers were strongly underrepresented in their data, but that makes sense given what I see in my smooth plots. They are most attractive early in the year.

## 4.2 The emergent phenomenon of diet similarity

Our finding that bumble bee diet ordination produces a first axis corresponding to tongue length and a second axis describing the dispersion of short-tongued species is strikingly consistent with past studies [@Goulson2008-zy; Durieux2000-uk].

## 4.3 The lability of bias through elevation and time

Species differed not just in their overall bias toward given morphotypes, but also in *when* and *where* they were biased toward a morphotype.

Ours is, to our knowledge, the first study to characterize formally the variation of bumble bee foraging bias through elevation an time.

# 5. References

### 2021-02-13 code/abund_div_GAM.Rmd
update

---
title: "Abundance/diversity GAM"
output: html_document
---

### Prepare environment
```{r setup, include=FALSE, echo=FALSE}
library(mgcv)
library(mgcViz)
library(tidyverse)
library(lubridate)
```

### Load site and trait data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class2 = factor(elev.class2, levels = c("low", "mid", "high")),
         site = factor(site)) # turn elav.class into an ordered factor and site into a regular factor

# BB traits
bb_traits <- read_csv("../data/processed_data/bb_traits.csv")

# FL traits
## family-level taxonomy
fl_tax <- read_csv("../data/processed_data/floral_tax.csv")

## Kugler morphotypes (accessed through Bioflor)
fl_ktype <- read_csv("../data/processed_data/floral_k_type.csv")

## Flower colors (accessed through BioFlor)
fl_color <- read_csv("../data/processed_data/floral_color.csv")

## diameter and corolla depth data from Gita Benadi
fl_morph <- read_csv("../data/processed_data/flower_morphology_GitaBenadi.csv") %>% 
  dplyr::select(SampleID, plant.family = Family, plant.sp = Species, 
                Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam), LN = mean(LN))

## verbal description of Kugler morphotypes
fl_ktype_key <- read_delim("../data/processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  dplyr::select(k.type = 1, description = 2)

## join all floral trait tables
fl_traits <- fl_tax %>%
  full_join(fl_ktype) %>%
  full_join(fl_color) %>%
  full_join(fl_morph) %>%
  dplyr::select(plant.sp, plant.genus, plant.family, k.type, k.type.s, k.type.ss, color, diam, LN)
```

### Load network and survey data
```{r message=FALSE}
### BB-FL visitation data
net <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>%
  dplyr::select(site, date, bb.sp, plant.sp) # drop unused columns

### Floral survey data
survey <- read_csv("../data/processed_data/floral_survey.csv") %>%
  semi_join(net, by = "plant.sp") %>% # Consider only visited plant species (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  group_by(plant.sp) %>%
  mutate(present = case_when(
    max(flower.cover) > 0 ~ TRUE,
    max(flower.cover) == 0 ~ FALSE
  )) %>%
  filter(present == TRUE) # Ignore plants that never had any flower cover

vis_sp <- net %>%
  left_join(fl_tax) %>%
  select(plant.sp, plant.genus, plant.family) %>%
  unique()

surv_sp <- survey %>%
  left_join(fl_tax) %>%
  select(plant.sp, plant.genus, plant.family) %>%
  unique()

missing <- anti_join(vis_sp, surv_sp)
```

### FL and BB abundance / diversity
```{r echo=FALSE, include=TRUE, message=FALSE}
# Per-transect BB abundance
## full dataset
bb_abund_sp <- net %>%
  dplyr::select(site, date, bb.sp) %>%
  group_by(site, date, bb.sp) %>% # 
  summarize(bb.abund = n()) %>% # get per transect bb abundance
  ungroup() %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  semi_join(net, by = c("site", "date")) %>% # drop site-dates that did not exist in the data set
  left_join(site_data) %>% # add site data
  mutate(year = factor(year(date)),
         yday = yday(date),
         bb.sp = factor(bb.sp),
         management = factor(management),
         site = factor(site),
         transect = factor(transect)) %>%
  group_by(year, bb.sp) %>%
  mutate(bb.annual.abund = sum(bb.abund),
         abund.class = case_when(
           bb.sp %in% c("telu", "pasc", "prat", "soro", "wurf") ~ "dominant",
           bb.sp %in% c("jone", "psit", "hort") ~ "common",
           bb.sp %in% c("pyre", "mend", "mont", "hypn", "lapi") ~ "high elev.",
           bb.sp %in% c("gers", "humi", "muci") ~ "minor"
         ),
         abund.class = factor(abund.class, levels = c("dominant", "common", "high elev.", "minor")),
         bb.sp.ord = factor(bb.sp, levels = c("telu", "pasc", "prat", "soro", "wurf", 
                                              "hort", "jone", "psit",
                                              "pyre", "mend", "mont",
                                              "gers", "humi", "muci", "hypn", "lapi"))) %>%
  left_join(bb_traits) %>%
  mutate(tongue = factor(pbl.w.class2))

# ## filtered to include only samples with corresponding floral survey data
# bb_abund_sp.sub <- bb_abund_sp %>%
#   ungroup() %>%
#   semi_join(shared_transects) %>%
#   select(site, date, yday, bb.sp, bb.abund, bb.annual.abund, elev.mean)

### Per-transect floral abund by species
fl_abund_sp <- survey %>%
  group_by(site, date, plant.sp) %>% # group by plant.sp*transect
  summarize(fl.abund = sum(flower.cover)) %>% # get plant.sp*transect total flower cover; this step really shouldn't be required, but in a few cases, there is more than one abundance entry for the same species on a given transect; not sure how that happened, but it's in the original data that Fabrice sent me; alternatively, I could drop the dups or average them. We'll see what Kalli and Fabrice have to say.
  group_by(site, date) %>%
  mutate(prop.fl.abund = fl.abund/sum(fl.abund)) %>% # get the proportional flower cover for each species.
  ungroup() %>%
  #full_join(dplyr::select(net, site, date, plant.sp), by = c("site", "date", "plant.sp")) %>% # add the handful of species occurring in the visitation network but missing from the survey data
  complete(site, date, plant.sp, fill = list(fl.abund = 0, prop.fl.abund = 0)) %>%
  semi_join(survey, by = c("site", "date")) %>%  # only include site-dates that actually occurred
  unique() %>%
  mutate(year = factor(year(date))) %>%
  left_join(fl_traits)

### Visualize raw abundance/occurrence
ggplot(filter(bb_abund_sp, bb.abund > 0), aes(yday, elev.mean, color = abund.class)) +
  geom_point(aes(alpha = bb.abund)) +
  #geom_point() +
  #facet_grid(year~ bb.sp) 
  facet_wrap(~bb.sp.ord)

ggplot(bb_abund_sp, aes(bb.sp, log(bb.abund + 1))) +
  geom_boxplot()

ggplot(bb_abund_sp, aes(log(bb.abund + 1))) +
  geom_histogram() +
  facet_wrap(~bb.sp)
```

### Binary "preference" for each plant genus by each bumble bee species; score species 0 or 1 based on whether they were visited
```{r message=FALSE}
weights <- fl_abund_sp %>%
  select(plant.sp, plant.genus) %>%
  unique() %>%
  left_join(select(net, bb.sp, plant.sp)) %>%
  select(plant.genus, bb.sp) %>%
  distinct() %>%
  mutate(visited = rep(1, n())) %>%
  pivot_wider(values_from = visited, names_from = bb.sp, values_fill = 0) %>%
  pivot_longer(cols = 2:17, values_to = "visited", names_to = "bb.sp")

weighted_fl_abund <- fl_abund_sp %>%
  group_by(site, date, plant.genus) %>%
  summarize(flower.cover = sum(fl.abund)) %>%
  left_join(weights) %>%
  mutate(weighted.flower.cover = flower.cover*visited) %>%
  group_by(site, date, bb.sp) %>%
  summarize(total.weighted.flower.cover = sum(weighted.flower.cover)) %>%
  left_join(site_data) %>%
  select(site, date, elev.mean, bb.sp, total.weighted.flower.cover) %>%
  mutate(yday = yday(date),
         bb.sp = factor(bb.sp),
         site = factor(site),
         year = factor(year(date)))

ggplot(weighted_fl_abund, aes(total.weighted.flower.cover)) +
  geom_histogram() +
  facet_wrap(~bb.sp)

ggplot(weighted_fl_abund, aes(bb.sp, total.weighted.flower.cover)) +
  geom_boxplot()

ggplot(weighted_fl_abund, aes(bb.sp, log(total.weighted.flower.cover))) +
  geom_boxplot()
```

### Call GAMs
#### Floral abundance
```{r message=FALSE, warning=FALSE}
# First, let's model overall floral abundance without parsing it out by which species were visited by which bumble bee species
fl_abund_pool <- fl_abund_sp %>%
  group_by(site, date, year) %>%
  summarize(fl.abund = sum(fl.abund)) %>%
  mutate(site = factor(site),
         yday = yday(date)) %>%
  left_join(site_data)

ggplot(fl_abund_pool, aes(fl.abund)) +
  geom_histogram()

fl_abund.qp <- bam(fl.abund ~ 
                      s(site, bs = "re") +
                      s(year, bs = "re") +
                      te(yday, elev.mean,
                         bs = c("gp", "tp"),
                         k = c(8, 8)),
                    data = fl_abund_pool,
                    family = "quasipoisson",
                    discrete = TRUE,
                    select = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(fl_abund.qp)
summary(fl_abund.qp)
plot(fl_abund.qp)

# Now, by bumble bee species
weighted_fl_abund.qp <- bam(total.weighted.flower.cover ~ 
                              s(bb.sp, bs = "re") +
                              s(site, bs = "re") +
                              s(year, bs = "re") +
                              te(yday, elev.mean,
                                 by = bb.sp,
                                 bs = c("gp", "tp"),
                                 k = c(5, 5)),
                            data = filter(weighted_fl_abund, bb.sp != "humi"),
                            family = "quasipoisson",
                            discrete = TRUE,
                            select = TRUE,
                            method = "fREML") %>% getViz(nsim = 500)

check.gamViz(weighted_fl_abund.qp)
sum_fl.qp <- summary(weighted_fl_abund.qp)
print(plot(weighted_fl_abund.qp, allTerms = TRUE), pages = 4)
check2D(weighted_fl_abund.qp, x1 = "yday", x2 = "elev.mean")

gers_fl_abund <- plot(sm(weighted_fl_abund.qp, 4))
hort_fl_abund <- plot(sm(weighted_fl_abund.qp, 5))
hypn_fl_abund <- plot(sm(weighted_fl_abund.qp, 6))
jone_fl_abund <- plot(sm(weighted_fl_abund.qp, 7))
lapi_fl_abund <- plot(sm(weighted_fl_abund.qp, 8))
mend_fl_abund <- plot(sm(weighted_fl_abund.qp, 9))
mont_fl_abund <- plot(sm(weighted_fl_abund.qp, 10))
muci_fl_abund <- plot(sm(weighted_fl_abund.qp, 11))
pasc_fl_abund <- plot(sm(weighted_fl_abund.qp, 12))
prat_fl_abund <- plot(sm(weighted_fl_abund.qp, 13))
psit_fl_abund <- plot(sm(weighted_fl_abund.qp, 14))
pyre_fl_abund <- plot(sm(weighted_fl_abund.qp, 15))
soro_fl_abund <- plot(sm(weighted_fl_abund.qp, 16))
telu_fl_abund <- plot(sm(weighted_fl_abund.qp, 17))
wurf_fl_abund <- plot(sm(weighted_fl_abund.qp, 18))

pdf("../output/gam_FL.pdf")
grid_plot_fl_abun <- gridPrint(gers_fl_abund + 
                                  l_fitRaster() + 
                                  l_fitContour() + 
                                  labs(title = "gers", x = NULL, y = NULL) +
                                  guides(fill=FALSE),
                               hort_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "hort", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               hypn_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "hypn", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               jone_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "jone", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               lapi_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "lapi", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               mend_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "mend", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               mont_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "mont", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               muci_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "muci", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               pasc_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "pasc", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               prat_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "prat", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               psit_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "psit", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               pyre_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "pyre", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               soro_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "soro", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               telu_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "telu", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               wurf_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "wurf", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               ncol = 4)
dev.off()
```

#### Bumbles
```{r message=FALSE}
# Zero-inflation may be an issue; a simple Poisson model is badly overdispersed
ggplot(bb_abund_sp, aes(bb.abund)) +
  geom_histogram() +
  facet_wrap(~bb.sp)

# Quasipoisson (winner)
bb_abund.qp <- bam(bb.abund ~ 
                     s(bb.sp.ord, bs = "re") + # center species
                     s(year, bs = "re") + # random intercept for year
                     s(site, bs = "re") + # random intercept to avoid pseudoreplicating elevation
                     te(yday, elev.mean, # tensor product smooth of elevation and time
                        by = bb.sp.ord, # by bb.sp
                        bs= c("gp", "tp"), # gaussian procees for time, thinplate spline for elevation
                        k=c(5, 5), 
                        m=2),
                   data = filter(bb_abund_sp, bb.sp != "humi"), # humi is just too rare
                   discrete = TRUE,
                   family = "quasipoisson",
                   select = TRUE,
                   method = "fREML") %>% getViz(nsim = 500)

check.gamViz(bb_abund.qp)
sum.qp <- summary(bb_abund.qp)
print(plot(bb_abund.qp, allTerms = TRUE), pages = 4)
check2D(bb_abund.qp, x1 = "yday", x2 = "elev.mean")

telu_abund <- plot(sm(bb_abund.qp, 4))
pasc_abund <- plot(sm(bb_abund.qp, 5))
prat_abund <- plot(sm(bb_abund.qp, 6))
soro_abund <- plot(sm(bb_abund.qp, 7))
wurf_abund <- plot(sm(bb_abund.qp, 8))
hort_abund <- plot(sm(bb_abund.qp, 9))
jone_abund <- plot(sm(bb_abund.qp, 10))
psit_abund <- plot(sm(bb_abund.qp, 11))
pyre_abund <- plot(sm(bb_abund.qp, 12))
mend_abund <- plot(sm(bb_abund.qp, 13))
mont_abund <- plot(sm(bb_abund.qp, 14))
gers_abund <- plot(sm(bb_abund.qp, 15))
muci_abund <- plot(sm(bb_abund.qp, 16))
hypn_abund <- plot(sm(bb_abund.qp, 17))
lapi_abund <- plot(sm(bb_abund.qp, 18))

pdf("../output/gam_BB.pdf")
grid_plot_fl_abun <- gridPrint(gers_abund + 
                                  l_fitRaster() + 
                                  l_fitContour() + 
                                  labs(title = "gers", x = NULL, y = NULL) +
                                  guides(fill=FALSE),
                               hort_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "hort", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               hypn_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "hypn", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               jone_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "jone", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               lapi_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "lapi", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               mend_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "mend", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               mont_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "mont", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               muci_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "muci", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               pasc_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "pasc", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               prat_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "prat", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               psit_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "psit", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               pyre_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "pyre", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               soro_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "soro", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               telu_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "telu", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               wurf_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "wurf", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               ncol = 4)
dev.off()
```

#### Functional bumbles
```{r}
bb_abund_sp_tongues <- bb_abund_sp %>%
  group_by(site, date, year, pbl.w.class2, elev.mean, yday) %>%
  summarize(bb.abund = sum(bb.abund)) %>%
  mutate(tongue = factor(pbl.w.class2)) %>%
  dplyr::select(-pbl.w.class2)

bb_abund_tongue.qp <- bam(bb.abund ~ 
                             s(bb.sp.ord, bs = "re") + # center species
                             s(tongue, bs = "re") + # center species
                             s(year, bs = "re") + # random intercept for year
                             s(site, bs = "re") + # random intercept to avoid pseudoreplicating elevation
                             te(yday, elev.mean, # tensor product smooth of elevation and time
                                by = tongue, # by bb.sp
                                bs= c("gp", "tp"), # gaussian procees for time, thinplate spline for elevation
                                k=c(8, 8), 
                                m=2),
                           data = bb_abund_sp, # humi is just too rare
                           discrete = TRUE,
                           family = "quasipoisson",
                           select = TRUE,
                           method = "fREML") %>% getViz()

check.gamViz(bb_abund_tongue.qp)
sum.tongue_qp <- summary(bb_abund_tongue.qp)
print(plot(bb_abund_tongue.qp, allTerms = TRUE), pages = 1)
plot(bb_abund_tongue.qp)
```

#### Occurrence clustering
```{r}
library(vegan)
library(ggvegan)
library(ggrepel)

# Unpooled data
bb_rel_abund <- bb_abund_sp %>%
  filter(bb.sp != "humi") %>%
  ungroup() %>%
  group_by(site, date, bb.sp) %>%
  summarize(bb.abund = sum(bb.abund)) %>%
  unite(id, c(site, date)) %>%
  pivot_wider(names_from = id, values_from = bb.abund) %>%
  column_to_rownames("bb.sp") %>%
  decostand(method = "hellinger")

dist <- vegdist(bb_rel_abund, method = "horn")
clust <- hclust(dist, method = "ward.D2")

pdf("../output/clust.pdf", height = 4.5, width = 9)
plot(clust)
dev.off()

pcoa <- capscale(bb_rel_abund ~ 1, method = "horn")
plot(pcoa, type = "t")

pcoa_mds1 <- round(pcoa$CA$eig[[1]] / sum(pcoa$CA$eig)*100)
pcoa_mds2 <- round(pcoa$CA$eig[[2]] / sum(pcoa$CA$eig)*100)

pcoa.fort <- fortify(pcoa)

ggplot(filter(pcoa.fort, Score == "sites"), aes(MDS1, MDS2)) +
  geom_point() +
  geom_text_repel(aes(label = Label), size = 6) +
  xlab(paste("MDS1 ", "(", pcoa_mds1, "%", ")", sep = "")) +
  ylab(paste("MDS2 ", "(", pcoa_mds2, "%", ")", sep = "")) +
  coord_fixed() +
  theme_gray(16)

ggsave("../output/pcoa_abund.pdf")

# Pooled data
bb_rel_abund2 <- bb_abund_sp %>%
  filter(bb.sp != "humi") %>%
  ungroup() %>%
  group_by(site, date, bb.sp) %>%
  summarize(bb.abund = sum(bb.abund)) %>%
  group_by(site, bb.sp) %>%
  summarize(bb.abund = mean(bb.abund)) %>%
  pivot_wider(names_from = site, values_from = bb.abund) %>%
  column_to_rownames("bb.sp") %>%
  decostand(method = "hellinger")

dist2 <- vegdist(bb_rel_abund2, method = "horn")
clust2 <- hclust(dist2, method = "ward.D2")

pdf("../output/clust2.pdf", height = 4.5, width = 9)
plot(clust2)
dev.off()

pcoa2 <- capscale(bb_rel_abund2 ~ 1, method = "horn")
plot(pcoa2, type = "t")

pcoa2_mds1 <- round(pcoa2$CA$eig[[1]] / sum(pcoa2$CA$eig)*100)
pcoa2_mds2 <- round(pcoa2$CA$eig[[2]] / sum(pcoa2$CA$eig)*100)

pcoa2.fort <- fortify(pcoa2)

ggplot(filter(pcoa2.fort, Score == "sites"), aes(MDS1, MDS2)) +
  geom_point() +
  geom_text_repel(aes(label = Label), size = 6) +
  xlab(paste("MDS1 ", "(", pcoa2_mds1, "%", ")", sep = "")) +
  ylab(paste("MDS2 ", "(", pcoa2_mds2, "%", ")", sep = "")) +
  coord_fixed() +
  theme_gray(16)

ggsave("../output/pcoa2_abund2.pdf")
```

### BB range and abundance
```{r}
species_names <- cbind(c("telu", "pasc", "prat", 
                         "soro", "wurf", "hort", 
                         "pyre", "mend", "mont", 
                         "gers", "hypn","jone", 
                         "lapi", "muci", "psit"),
                       c("B. terrestris/lucorum", "B. pascuorum", "B. pratorum", 
                         "B. soroensis", "B. wurflenii", "B. hortorum",
                         "B. pyrenaeus", "B. mendax", "B. monticola",
                         "B. gerstaeckeri", "B. hypnorum", "B. jonellus", 
                         "B. lapidarius", "B. mucidus", "B. psithyrus"),
                       c("B. telu.", "B. pasc.", "B. prat.", 
                         "B. soro.", "B. wurf.", "B. hort.",
                         "B. pyre.", "B. mend.", "B. mont.",
                         "B. gers.", "B. hypn.", "B. jone.", 
                         "B. lapi.", "B. muci.", "B. psit."),
                       c("major", "major", "major", 
                         "major", "major", "common", 
                         "high elev.", "high elev.", "high elev.",
                         "minor", "minor", "common", 
                         "minor", "high elev.", "common")
                       ) %>%
  as.data.frame() %>%
  as_tibble() %>%
  dplyr::select(bb.sp = 1, species.name = 2, species.abb = 3, abund.class = 4)

bb_range <- net %>%
  #filter(bb.sp != "humi") %>%
  left_join(site_data) %>%
  group_by(site, date, elev.mean, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, bb.sp) %>%
  summarize(abund = mean(abund)) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))
# %>%
#   left_join(species_names) %>%
#   mutate(abund.class = factor(abund.class, levels = c("major", "common", "high elev.", "minor")))

ggplot(bb_range, aes(reorder(bb.sp, elev.med), elev.mean)) +
  geom_line(size = 2, alpha = 0.25) +
  geom_point(aes(size = abund), alpha = 0.5) +
  #facet_wrap(~year, ncol = 1) +
  theme_light(12) +
  scale_size_continuous(name = "Abundance") +
  #scale_color_discrete(name = "Abund. class") +
  xlab("BB species") +
  ylab("Elevation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("../output/bb_range.png", width = 6, height = 4, units = "in")
ggsave("../output/bb_range.pdf", width = 6, height = 4)
```

### 2021-02-13 code/beta_diversity.Rmd
update

---
title: "Beta Diversity through Time and Space"
output: html_document
---

### Prepare environment
```{r message=FALSE, warning=FALSE}
library(sf)
library(gdm)
library(patchwork)
library(bipartite)
library(vegan)
library(see)
library(tidyverse)
library(lubridate)
```

### Create functions for later
```{r}
# A function for converting a geometry column back to x and y columns
# https://github.com/r-spatial/sf/issues/231
sfc_as_cols <- function(x, names = c("x","y")) {
  stopifnot(inherits(x,"sf") && inherits(sf::st_geometry(x),"sfc_POINT"))
  ret <- sf::st_coordinates(x)
  ret <- tibble::as_tibble(ret)
  stopifnot(length(names) == ncol(ret))
  x <- x[ , !names(x) %in% names]
  ret <- setNames(ret,names)
  dplyr::bind_cols(x,ret)
}

# Function to convert networks to webs arrays
web_prep <- function(net) {

  out <- net %>%
    filter(caste != "male") %>%
    group_by(site, bb.sp, plant.sp) %>%
    summarize(freq = n()) %>%
    group_by(site) %>%
    mutate(bb.sp.count = length(unique(bb.sp)),
           plant.sp.count = length(unique(plant.sp))) %>%
    ungroup() %>%
    dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
    data.frame() %>% # convert to data frame
    frame2webs(type.out = "array")
}

# Function with betalinkr settings and conversion to tibble()
betalinkr_set1 <- function(x) {
  betalinkr_multi(x, partition.st = TRUE, partition.rr = TRUE, # partition both species turnover and richness/replacement
                  binary = TRUE, # treating beta-diversity as binary is an appropriate complement to our abundance analysis
                  index = "jaccard") %>% # the Jaccard index is best, I think, for binary analysis
  as_tibble() # for convenience
}

# Binomial smooth function (thanks, Hadley): https://ggplot2.tidyverse.org/reference/geom_smooth.html
binomial_smooth <- function(...) {
  geom_smooth(method = "glm", method.args = list(family = "binomial"), ...)
}

# Site_pair prep function
# site_pair_prep_net <- function(dat, y) {
#   dat %>%
#     ungroup() %>%
#     filter(year1 == year2 & 
#              year1 == y & 
#              site1 != site2) %>% # I don't think we want to compare samples through within sites, introduces a new form of non-independence.
#     filter(metric == "WN") %>% # we'll focus for now on WN, but this could be done for partitions, too
#     mutate(weights = rep(1, dim(.)[1])) %>% # right now this analysis is unweighted; could weight by abundance, etc, if we wanted to
#     dplyr::select(distance = value, weights,
#                   s1.xCoord = x1, s1.yCoord = y1, 
#                   s2.xCoord = x2, s2.yCoord = y2, 
#                   s1.elev.mean = elev1, s1.yday = yday1,
#                   s2.elev.mean = elev2, s2.yday = yday2) %>%
#     data.frame()
# }

interaction_gdm_prep <- function(x) {

  spdat <- net %>%
    filter(caste != "male") %>%
    dplyr::select(bb.sp, plant.sp.abb, site) %>% 
    distinct() %>%# each unique bb-fl-site-date combo
    mutate(present = rep(1, length(.[,1]))) %>% # we will treat abundance as binary
    unite(interaction, c(bb.sp, plant.sp.abb), sep = "_") %>% # unite bb and fl into interaction "species", i.e. a unique bb-fl combo
    pivot_wider(names_from = interaction, values_from = present) %>% # spread so that columns are interaction "species"
    replace(is.na(.), 0) %>% # replace NAs with absences
    as_tibble()

  envdat <- net %>%
    filter(caste != "male") %>%
    group_by(site) %>%
    summarize(samples = length(unique(date))) %>%
    #dplyr::select(bb.sp, plant.sp.abb, site) %>% #
    left_join(site_data) %>% # add site data
    dplyr::select(site, elev.mean, x, y, samples) %>%
    distinct() # return unique site-dates

  # Prep site-pairs for GDM analysis
  sitepair <- formatsitepair(spdat, 
                             bioFormat = 1, # i.e. samples as rows, species as columns, site-dates as rows
                             siteColumn = "site", 
                             dist = "jaccard", # I think Jaccard is the right choice
                             XColumn = "x", YColumn = "y", # spatial columns
                             predData = envdat, 
                             abundance = FALSE) %>% # abundance = FALSE means binary
    filter(s1.elev.mean != s2.elev.mean) # this drops within-site comparisons
}

## wrapper for extracting splines and prepping tibble for plotting
spline_prep <- function(x, y) {
  isplineExtract(x) %>%
    as.data.frame() %>%
    tibble() %>%
    mutate(x.Geographic = x.Geographic/1000) %>%
    pivot_longer(cols = c(x.Geographic, x.elev.mean), 
                 names_to = "x.var", values_to = "predictor", 
                 names_prefix = "x.") %>%
    pivot_longer(cols = c(y.Geographic, y.elev.mean), 
                 names_to = "y.var", values_to = "response", 
                 names_prefix = "y.") %>%
    filter(x.var == y.var) %>%
    mutate() %>%
    dplyr::select(-y.var)
}

# Create functions for prepping data
biodata_prep_survey <- function(surv) {
  surv %>%
    filter(flower.cover > 0) %>%
    mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
    group_by(site, plant.sp) %>%
    summarize(cover = sum(flower.cover)) %>%
    mutate(cover = rep(1, n())) %>% # this makes the data binary, which I think is what I want here; otherwise, just comment out
    spread(plant.sp, cover) %>%
    replace(is.na(.), 0) %>%
    arrange(site)
}

predData_prep_survey <- function(surv) {
  surv %>%
    filter(flower.cover > 0) %>%
    group_by(site) %>%
    summarize(samples = length(unique(date))) %>%
    dplyr::select(site, samples) %>%
    left_join(site_data) %>%
    ungroup() %>%
    dplyr::select(site, elev.mean, y, x, samples) %>%
    unique() %>%
    arrange(site) 
}

format_set <- function(bio, pred) {
  formatsitepair(bio, bioFormat = 1, siteColumn = "site", 
                 dist = "jaccard", XColumn = "x", YColumn = "y", 
                 predData = pred, abundance = FALSE) %>%
  filter(s1.elev.mean != s2.elev.mean) # this avoids comparing repeated measures of the same site (taking advantage of the fact that each site has a unique elevation in our dataset)
}

biodata_prep_BB <- function(net) {
  net %>%
    filter(caste != "male") %>%
    group_by(site, bb.sp) %>%
    summarize(count = n()) %>%
    mutate(count = rep(1, n())) %>% # this makes the data binary, which I think is what I want here; otherwise, just comment out
    spread(bb.sp, count) %>%
    replace(is.na(.), 0) %>%
    ungroup()
}

predData_prep_BB <- function(net) {
  net %>%
    filter(caste != "male") %>%
    group_by(site) %>%
    summarize(samples = length(unique(date))) %>%
    dplyr::select(site, samples) %>%
    left_join(site_data) %>%
    ungroup() %>%
    dplyr::select(site, elev.mean, y, x, samples) %>%
    unique() %>%
    arrange(site) 
}
```

### Load site data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class2 = factor(elev.class2, levels = c("low", "mid", "high")),
         site = factor(site)) %>% # turn elav.class into an ordered factor and site into a regular factor
  semi_join(read_csv("../data/processed_data/network.csv"), by = "site") %>% # drop sites that were not sampled
  st_as_sf(coords = c("lon", "lat")) %>%
  st_set_crs(4326) %>%
  st_transform(5684) %>%
  sfc_as_cols() %>%
  as_tibble() %>%
  dplyr::select(-geometry)
```

### Load network and survey data
```{r}
net <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste  != "male") %>%
  unite(site.year.day, site, year, dayofyear, sep = "_", remove = FALSE) %>% # create site.date field that can be split up again later
  mutate(year = factor(year)) # convert year to factor

survey <- read_csv("../data/processed_data/floral_survey.csv") %>%
  filter(flower.cover > 0) %>%
  semi_join(net, by = "plant.sp") %>% # Consider only visited plant spp. (anytime, anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  left_join(fl_traits)
```

### Get beta-diversity for BBs and plants (not their interactions)
```{r}
# Bumble bee beta-diversity based on interaction sampling
bb_beta <- net %>%
  group_by(site, bb.sp) %>%
  summarize(count = n()) %>% 
  pivot_wider(names_from = bb.sp, values_from = count) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("site") %>%
  vegdist(method = "jaccard", binary = TRUE) %>% # Jaccard and binary settings make this comparable to the way I'm running betalinkr
  as.matrix() %>%
  as.data.frame() %>%
  rownames_to_column(var = "site1") %>%
  pivot_longer(cols = -site1, names_to = "site2", values_to = "BB")

# Flora beta-diversity based on floral survey data (filtered to include only species visited by BBs)
fl_beta <- survey %>%
  group_by(site, plant.sp) %>%
  summarize(count = sum(flower.cover)) %>%
  pivot_wider(names_from = plant.sp, values_from = count) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("site") %>%
  vegdist(method = "jaccard", binary = TRUE, upper = TRUE) %>%
  as.matrix() %>%
  as.data.frame() %>%
  rownames_to_column(var = "site1") %>%
  pivot_longer(cols = -site1, names_to = "site2", values_to = "FL")

# Join into single data frame for visualization
species_beta <- full_join(bb_beta, fl_beta) %>%
  left_join(site_data, by = c("site1" = "site")) %>% # add site data
  dplyr::select(site1, site2, FL, BB, # first for site 1
                elev1 = elev.mean, x1 = x, y1 = y) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, FL, BB, # then for site 2
                elev1, elev2 = elev.mean, x1, y1, x2 = x, y2 = y) %>%
  mutate(site1 = factor(site1), 
         site2 = factor(site2),
         elev.diff = elev1 - elev2,
         mean.elev = (elev1 + elev2)/2) %>%
  pivot_longer(cols = c(BB, FL), names_to = "group", values_to = "beta") %>%
  filter(elev.diff >= 0) # For each site pair, one will be higher and the other lower; thus, when their order is reversed only the sign of the difference changes. Since no two sites have exactly the same elevation, this deduplicates the data by dropping the ordering of a given pair of sites (one of the two orderings) that yields a negative elevation difference while retaining the alternative ordering that yields a positive elevation difference. By >= 0, I retain comparisons of a site to itself. I drop (in the ggplot call) within-site comparisons when I plot against elevation difference, and I isolate within-site comparison when I plot against time. In the latter case, it is also necessary to deduplicate, since each pair of site dates occurs in A-B and B-A. For that, I filter (again, in the ggplot call) so that yday.diff > 0, similarly dropping the orderings that yield negative differences. Note that all this deduplication nonsense is not necessary for the betalinkr call because it deduplicates internally.
```

### Tabulate webs
S	= beta_S, the dissimilarity in species composition
OS = beta_OS, the dissimilarity (component) explained by "rewiring" among shared species
WN = beta_WN, the dissimilarity between the two networks
ST = beta_ST, the dissimilarity (component) explained by difference in species community composition 
```{r message=FALSE}
# Create webs
net_web <- web_prep(net)

# Calculate beta-diversity
beta_site <- betalinkr_set1(net_web) %>%
  rename(site1 = i, site2 = j)

# Add site data
beta_mit_data <- beta_site %>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1 = elev.mean, x1 = x, y1 = y) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2 = elev.mean, x1, y1, x2 = x, y2 = y) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2) %>%
  gather(metric, value, -c(site1, site2, elev1, elev2, 
                           elev.diff, mean.elev, 
                           x1, y1, x2, y2)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"),
    metric = factor(metric, levels = c("WN", "ST", "OS", "S", "ST.h", "ST.l", 
                                       "ST.lh", "WN.repl", "WN.rich", 
                                       "OS.repl", "OS.rich")))
```

### Plot interaction beta-diversity against elevation difference and time difference
#### For elevation difference, we will "control" for time by only considering pairs separated by <= 5 days
#### For time difference, we will control for elevation by only comparing within site.
```{r}
# Set palette
# p <- c("OS" = "darkorange1", "ST" = "green3", 
#        "ST.h" = "firebrick1", "ST.l" = "darkorchid", 
#        "ST.lh" = "deepskyblue", "WN" = "black")
# 
# 
# p <- c("OS" = "#66c2a5", "ST" = "#fc8d62", 
#        "ST.h" = "#8da0cb", "ST.l" = "#e78ac3", 
#        "ST.lh" = "#a6d854", "WN" = "black")

p <- c("OS" = "#e7298a", "ST" = "#d95f02", 
       "ST.h" = "#7570b3", "ST.l" = "#1b9e77", 
       "ST.lh" = "#66a61e", "WN" = "black")

# BB + FL elevation diff
g1 <- ggplot(filter(species_beta, 
                    site1 != site2), # only site-dates within 5 days
             aes(elev.diff, beta, color = group)) +
  geom_point(alpha = 0.25, pch = 16) +
  #binomial_smooth(formula = y ~ splines::ns(x, 3)) +
  geom_smooth(method = "lm", se = FALSE) +
  #facet_wrap(~group, ncol = 1) +
  xlab("Elevation difference (m)") +
  ylab("β-diversity") +
  theme_light()

# BB * FL elevation diff
g2 <- ggplot(filter(beta_mit_data, 
                      site1 != site2 & # no within-site comparisons
                      metric != "S" & # drop this metric
                      !metric.class %in% c("OS.partition", "WN.partition")), 
             aes(elev.diff, value, color = metric))  +
  geom_point(alpha = 0.20, pch = 16) +
  binomial_smooth(formula = y ~ splines::ns(x, 3), se = FALSE) +
  scale_color_manual(values = p, name = "partition") +
  xlab("Elevation difference (m)") +
  ylab(NULL) +
  theme_light()

# Patchwork plot for comparing species beta-diversity with interaction beta-diversity
png("../output/beta_stack.png", width = 6.5, height = 4.5, units = "in", res = 300)
(g1 | g2) + 
  plot_annotation(tag_levels = c("A", "B")) + 
  #plot_layout(guides = "collect") & 
  theme_light() 
dev.off()
```

### GDM: Interactions
```{r message=FALSE, warning=FALSE}
# Prep sitepairs
sitepair.WN <- interaction_gdm_prep(net)

# Run GDM analysis
gdm_WN <- gdm(sitepair.WN, geo = TRUE)
summary(gdm_WN)

# Plot
plot(gdm_WN)

# Plot with confidence intervals 
pdf("../output/gdm_interactions.pdf", height = 2.5, width =  6.5)
plotUncertainty(sitepair.WN, sampleSites = 0.7, bsIters = 100, geo  = TRUE, 
                splineCol = "black", plot.layout = c(1,3))
dev.off()
```

### GDM: Floral survey
```{r}
# Prep input for GDM
survey_bioData <- biodata_prep_survey(survey)
survey_predData <- predData_prep_survey(survey)
survey_sitepair <- format_set(survey_bioData, survey_predData)

# Call GDMs
gdm_survey <- gdm(survey_sitepair, geo = TRUE)
summary(gdm_survey)

# Plot
plot(gdm_survey)

# Plot with confidence intervals (slow)
pdf("../output/gdm_flora.pdf", height = 2.5, width =  6.5)
plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 100, geo  = TRUE, 
                splineCol = "#00BFC4", plot.layout = c(1,3))
dev.off()
```

### GDM: BB observations
```{r}
# Prep input for GDM
BB_bioData <- biodata_prep_BB(net)
BB_predData <- predData_prep_BB(net)
BB_sitepair <- format_set(BB_bioData, BB_predData)

# Call GDMs
gdm_BB <- gdm(BB_sitepair, geo = TRUE)
summary(gdm_BB)

# Plot
plot(gdm_BB)

# Plot with confidence intervals (slow)
pdf("../output/gdm_BB.pdf", height = 2.5, width =  6.5)
plotUncertainty(BB_sitepair, sampleSites = 0.7, bsIters = 100, geo  = TRUE, 
                splineCol = "#F8766D", plot.layout = c(1,3))
dev.off()
```

### Tabulate model output
```{r}
# assign("gdm_WN.2010.imp", get(load("~/Documents/Research/bb_berchtesgaden/analyses/output/interaction_permGDM_2010.RData")))
# 
# assign("gdm_WN.2011.imp", get(load("~/Documents/Research/bb_berchtesgaden/analyses/output/interaction_permGDM_2011.RData")))
# 
# assign("gdm_WN.2012.imp", get(load("~/Documents/Research/bb_berchtesgaden/analyses/output/interaction_permGDM_2012.RData")))

####### WN
gdm_WN.2010.tab1 <- gdm_WN.2010.imp[1] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  mutate(subject = "WN",
         year = 2010) %>%
  dplyr::select(subject, year, 
         mod.dev = `Model deviance`, 
         dev.exp = `Percent deviance explained`, 
         mod.pval = `Model p-value`)

gdm_WN.2010.tab2 <- gdm_WN.2010.imp[2] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)

gdm_WN.2010.tab3 <- gdm_WN.2010.imp[3] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble()  %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2)  %>%
  dplyr::select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)

gdm_WN.2011.tab1 <- gdm_WN.2011.imp[1] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  mutate(subject = "WN",
         year = 2011) %>%
  dplyr::select(subject, year, 
         mod.dev = `Model deviance`, 
         dev.exp = `Percent deviance explained`, 
         mod.pval = `Model p-value`)

gdm_WN.2011.tab2 <- gdm_WN.2011.imp[2] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)

gdm_WN.2011.tab3 <- gdm_WN.2011.imp[3] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble()  %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2)  %>%
  dplyr::select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)

gdm_WN.2012.tab1 <- gdm_WN.2012.imp[1] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  mutate(subject = "WN",
         year = 2012) %>%
  dplyr::select(subject, year, 
         mod.dev = `Model deviance`, 
         dev.exp = `Percent deviance explained`, 
         mod.pval = `Model p-value`)

gdm_WN.2012.tab2 <- gdm_WN.2012.imp[2] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)

gdm_WN.2012.tab3 <- gdm_WN.2012.imp[3] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble()  %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2)  %>%
  dplyr::select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)

gdm_WN.2010.tab_join <- bind_cols(gdm_WN.2010.tab1,
                                  gdm_WN.2010.tab2, 
                                  gdm_WN.2010.tab3)

gdm_WN.2011.tab_join <- bind_cols(gdm_WN.2011.tab1,
                                  gdm_WN.2011.tab2, 
                                  gdm_WN.2011.tab3)

gdm_WN.2012.tab_join <- bind_cols(gdm_WN.2012.tab1,
                                  gdm_WN.2012.tab2, 
                                  gdm_WN.2012.tab3)

####### survey
gdm_survey.2010.tab1 <- gdm_survey.2010.imp[1] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  mutate(subject = "survey",
         year = 2010) %>%
  dplyr::select(subject, year, 
         mod.dev = `Model deviance`, 
         dev.exp = `Percent deviance explained`, 
         mod.pval = `Model p-value`)

gdm_survey.2010.tab2 <- gdm_survey.2010.imp[2] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)

gdm_survey.2010.tab3 <- gdm_survey.2010.imp[3] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble()  %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2)  %>%
  dplyr::select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)

gdm_survey.2011.tab1 <- gdm_survey.2011.imp[1] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  mutate(subject = "survey",
         year = 2011) %>%
  dplyr::select(subject, year, 
         mod.dev = `Model deviance`, 
         dev.exp = `Percent deviance explained`, 
         mod.pval = `Model p-value`)

gdm_survey.2011.tab2 <- gdm_survey.2011.imp[2] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)

gdm_survey.2011.tab3 <- gdm_survey.2011.imp[3] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble()  %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2)  %>%
  select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)

gdm_survey.2012.tab1 <- gdm_survey.2012.imp[1] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplylr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  mutate(subject = "survey",
         year = 2012) %>%
  dplyr::select(subject, year, 
         mod.dev = `Model deviance`, 
         dev.exp = `Percent deviance explained`, 
         mod.pval = `Model p-value`)

gdm_survey.2012.tab2 <- gdm_survey.2012.imp[2] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)

gdm_survey.2012.tab3 <- gdm_survey.2012.imp[3] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble()  %>%
  dplytr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2)  %>%
  dplyr::select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)

gdm_survey.2010.tab_join <- bind_cols(gdm_survey.2010.tab1,
                                  gdm_survey.2010.tab2, 
                                  gdm_survey.2010.tab3)

gdm_survey.2011.tab_join <- bind_cols(gdm_survey.2011.tab1,
                                  gdm_survey.2011.tab2, 
                                  gdm_survey.2011.tab3)

gdm_survey.2012.tab_join <- bind_cols(gdm_survey.2012.tab1,
                                  gdm_survey.2012.tab2, 
                                  gdm_survey.2012.tab3)

######## BB
gdm_BB.2010.tab1 <- gdm_BB.2010.imp[1] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  mutate(subject = "BB",
         year = 2010) %>%
  dplyr::select(subject, year, 
         mod.dev = `Model deviance`, 
         dev.exp = `Percent deviance explained`, 
         mod.pval = `Model p-value`)

gdm_BB.2010.tab2 <- gdm_BB.2010.imp[2] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)

gdm_BB.2010.tab3 <- gdm_BB.2010.imp[3] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble()  %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2)  %>%
  dplyr::select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)

gdm_BB.2011.tab1 <- gdm_BB.2011.imp[1] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  mutate(subject = "BB",
         year = 2011) %>%
  dplyr::select(subject, year, 
         mod.dev = `Model deviance`, 
         dev.exp = `Percent deviance explained`, 
         mod.pval = `Model p-value`)

gdm_BB.2011.tab2 <- gdm_BB.2011.imp[2] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)

gdm_BB.2011.tab3 <- gdm_BB.2011.imp[3] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble()  %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2)  %>%
  dplyr::select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)

gdm_BB.2012.tab1 <- gdm_BB.2012.imp[1] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  mutate(subject = "BB",
         year = 2012) %>%
  dplyr::select(subject, year, 
         mod.dev = `Model deviance`, 
         dev.exp = `Percent deviance explained`, 
         mod.pval = `Model p-value`)

gdm_BB.2012.tab2 <- gdm_BB.2012.imp[2] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)

gdm_BB.2012.tab3 <- gdm_BB.2012.imp[3] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble()  %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2)  %>%
  dplyr::select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)

gdm_BB.2010.tab_join <- bind_cols(gdm_BB.2010.tab1,
                                  gdm_BB.2010.tab2, 
                                  gdm_BB.2010.tab3)

gdm_BB.2011.tab_join <- bind_cols(gdm_BB.2011.tab1,
                                  gdm_BB.2011.tab2, 
                                  gdm_BB.2011.tab3)

gdm_BB.2012.tab_join <- bind_cols(gdm_BB.2012.tab1,
                                  gdm_BB.2012.tab2, 
                                  gdm_BB.2012.tab3)

#######################

gdm_BB.tab_join <- bind_rows(gdm_BB.2010.tab_join,
                             gdm_BB.2011.tab_join,
                             gdm_BB.2012.tab_join,
                             gdm_survey.2010.tab_join,
                             gdm_survey.2011.tab_join,
                             gdm_survey.2012.tab_join,
                             gdm_WN.2010.tab_join,
                             gdm_WN.2011.tab_join,
                             gdm_WN.2012.tab_join) %>%
  dplyr::select(-subject)

gdm_tab <- gdm_BB.tab_join %>%
  gt(rowname_col = "year") %>%
  tab_spanner(
    label = "Model",
    columns = vars(mod.dev, dev.exp, mod.pval)
  ) %>%
  tab_spanner(
    label = "Geographic distance",
    columns = vars(geo.imp, geo.pval)
  ) %>%
  tab_spanner(
    label = "Elevational distance",
    columns = vars(elev.imp, elev.pval)
  ) %>%
  tab_spanner(
    label = "Temporal distance",
    columns = vars(yday.imp, yday.pval)
  ) %>%
  tab_row_group(
    group = "Bumble bees",
    rows = 1:3
  ) %>% 
  tab_row_group(
    group = "Flora",
    rows = 4:6
  ) %>%
  tab_row_group(
    group = "Interactions",
    rows = 7:9
  ) %>%
  fmt_number(
    columns = c("mod.dev", "dev.exp",
                "geo.imp", "elev.imp", "yday.imp"),
    decimals = 1
  ) %>%
  fmt_number(
    columns = c("mod.pval", "geo.pval", 
                "elev.pval", "yday.pval"),
    decimals = 3
  ) %>%
  cols_label(mod.dev = "tot. dev.",
             dev.exp = "dev. expl. (%)",
             mod.pval = "p",
             geo.imp = "imp.",
             elev.imp = "imp.",
             yday.imp = "imp.",
             geo.pval = "p-val",
             elev.pval = "p-val",
             yday.pval = "p-val")

gdm_tab.nopval <- gdm_BB.tab_join %>%
  dplyr::select(-c(geo.pval, elev.pval, yday.pval, mod.pval)) %>%
  gt(rowname_col = "year") %>%
  tab_spanner(
    label = "Model",
    columns = vars(mod.dev, dev.exp)
  ) %>%
  tab_spanner(
    label = "Variable importance (%)",
    columns = vars(geo.imp, elev.imp, yday.imp)
  ) %>%
  tab_row_group(
    group = "Bumble bees",
    rows = 1:3
  ) %>% 
  tab_row_group(
    group = "Flora",
    rows = 4:6
  ) %>%
  tab_row_group(
    group = "Interactions",
    rows = 7:9
  ) %>%
  fmt_number(
    columns = c("mod.dev", "dev.exp",
                "geo.imp", "elev.imp", "yday.imp"),
    decimals = 1
  ) %>%
  cols_label(mod.dev = "tot. dev.",
             dev.exp = "dev. expl. (%)",
             geo.imp = "geo. dist.",
             elev.imp = "elev. dist.",
             yday.imp = "temp. dist.") %>%
  data_color(
    columns = vars(geo.imp, elev.imp, yday.imp),
    colors = scales::col_numeric(
      palette = c("white", "red"), 
      domain = c(-5, 100))
    ) %>%
  data_color(
    columns = vars(dev.exp),
    colors = scales::col_numeric(
      palette = c("white", "red"), 
      domain = c(-5, 100))
  )

gtsave(gdm_tab.nopval, "gdm_tab.png","./output/gdm_tab.png")

```

### GDM all together now
```{r}
splines <- bind_rows(interaction_splines, bb_splines, survey_splines) %>%
  mutate(subject = factor(subject))

# ggplot(splines, aes(predictor, response, color = subject)) +
#   geom_line() +
#   facet_grid(year ~ x.var, scales = "free_x") +
#   xlab(NULL) +
#   ylab("f(x)") +
#   theme_light(12)
# ggsave("./output/GDM.png", width = 6.5, height = 5, units = "in")

splot1 <- ggplot(filter(splines, x.var == "Geographic dist."), 
                 aes(predictor, response, color = subject)) +
  geom_line() +
  #facet_wrap(~year, ncol = 1, scales = "free_x") +
  facet_grid(year ~ x.var, scales = "free_x") +
  xlab("km") +
  ylab("f(x)") +
  ylim(c(0, 7)) +
  theme_light(12) +
  theme(legend.position = "none")

splot2 <- ggplot(filter(splines, x.var == "Elevation"), 
       aes(predictor, response, color = subject)) +
  geom_line() +
  #facet_wrap(~year, ncol = 1, scales = "free_x") +
  facet_grid(year ~ x.var, scales = "free_x") +
  xlab("m") +
  ylab(NULL) +
  ylim(c(0, 7)) +
  theme_light(12) +
  theme(legend.position = "none")

splot3 <- ggplot(filter(splines, x.var == "Time"), 
       aes(as_date(predictor), response, color = subject)) +
  geom_line() +
  #facet_wrap(~year, ncol = 1, scales = "free_x") +
  facet_grid(year ~ x.var, scales = "free_x") +
  xlab("date") +
  ylab(NULL) +
  ylim(c(0, 7)) +
  theme_light(12) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

png("./output/GDM_splines.png", width = 8, height = 5, units = "in", res = 300)
splot1 | splot2 | splot3
dev.off()
```

# If I could get this to work, it would be cool to transform and plot my GDMs. But the script crashes every time. 
```{r}

# Function to build data set for gdm.transform()
transform_data <- function(dat) {

  ## yday min/max
  min_yday <- min(dat$s1.yday)
  max_yday <- max(dat$s1.yday)

  ## elev min/max
  min_elev <- min(dat$s1.elev.mean)
  max_elev <- max(dat$s1.elev.mean)

  ## lat/lon min/max
  min_x1 <- min(dat$s1.xCoord)
  max_x1 <- max(dat$s1.xCoord)

  min_y1 <- min(dat$s1.yCoord)
  max_y1 <- max(dat$s1.yCoord)

  min_x2 <- min(dat$s2.xCoord)
  max_x2 <- max(dat$s2.xCoord)

  min_y2 <- max(dat$s2.yCoord)
  max_y2 <- max(dat$s2.yCoord)

  # sequence of 10 points along the yday and elev ranges, respectively
  s1.xCoord <- seq(min_x1, max_x1, length.out = 10)
  s1.yCoord <- seq(min_y1, max_y1, length.out = 10)
  s2.xCoord <- seq(min_x2, max_x2, length.out = 10)
  s2.yCoord <- seq(min_y2, max_y2, length.out = 10)
  s1.elev.mean <- seq(min_elev, max_elev, length.out = 10)
  s1.yday <- seq(min_yday, max_yday, length.out = 10)
  s2.elev.mean <- seq(min_elev, max_elev, length.out = 10)
  s2.yday <- seq(min_yday, max_yday, length.out = 10)

  newdata <- tibble(s1.xCoord, s1.yCoord, s2.xCoord, s2.yCoord, 
                    s1.elev.mean, s1.yday, s2.elev.mean, s2.yday) %>% # combine seqs into data frame
    complete(s1.xCoord, s1.yCoord, s2.xCoord, s2.yCoord, 
             s1.elev.mean, s1.yday, s2.elev.mean, s2.yday) %>% # get all possible values
    mutate(distance = rep(1, n()),
           weights = rep(1, n())) %>%
    as.data.frame()

  # coordinates(newdata) <- ~ X + Y
  # # coerce to SpatialPixelsDataFrame
  # gridded(newdata) <- TRUE
  # # coerce to raster
  # out <- raster(newdata)
}

transdata_gdm_WN.2010 <- transform_data(sitepair_2010.WN)

gdm_WN.2010.trans <- gdm.transform(gdm_WN.2010, transdata_gdm_WN.2010)
gdm_WN.2010.pred <- predict(gdm_WN.2010, gdm_WN.2010.trans)

```

## GDM with BBGDM
```{r}
library(bbgdm)

spdat2010 <- net %>%
  filter(year == 2010) %>%
  select(bb.sp, plant.sp.abb, site, date) %>%
  distinct() %>%
  mutate(present = rep(1, length(.[,1]))) %>%
  unite(interaction, c(bb.sp, plant.sp.abb), sep = "_") %>%
  unite(site.date, c("site", "date"), sep = "_") %>%
  pivot_wider(names_from = interaction, values_from = present) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("site.date") %>%
  as.matrix()

envdat2010 <- net %>%
  filter(year == 2010) %>%
  select(bb.sp, plant.sp.abb, site, date) %>%
  mutate(yday = yday(date)) %>%
  left_join(site_data) %>%
  unite(site.date, c("site", "date"), sep = "_") %>%
  select(site.date, elev.mean, yday, x, y) %>%
  distinct() %>%
  column_to_rownames("site.date") %>%
  as.matrix()

form <- ~ 1 + elev.mean + yday

bbgdm2010 <- bbgdm(form, sp.dat = spdat2010, env.dat = envdat2010,
                   family = "binomial", link = "logit", dism_metric = "bray_curtis",
                   nboot = 10, spline_type = "ispline", spline_df = 5, 
                   spline_knots = 3, geo = FALSE, 
                   #geo.type = "euclidean", coord.names = c("x", "y"), 
                   optim.meth = "nlmnib", 
                   est.var = FALSE, trace = FALSE, prior = FALSE)

plot(as.response(bbgdm2010))
bbgdm.wald.test(bbgdm2010)
plot(diagnostics(bbgdm2010))
```

### 2021-02-13 code/preference_lability.Rmd
update

@@
 bb_traits <- read_csv("../data/processed_data/bb_traits.csv")
-
-ggplot(bb_traits, aes(pbl.w, pbl.w.durieaux,
-                      color = bb.sp, pch = factor(pbl.w.class2))) +
-  geom_point()
 
@@
 ```
+
 ## 1.3 Evaluate models
@@
 ggsave(file = "../output/brm_pvis01_plot1.pdf", width = 5, height = 7)
+
+ggplot(brm_pvis01_plotframe1, aes(condition_mean_trans, ktype, 
+                                 xmin = .lower_trans, xmax = .upper_trans)) +
+  geom_pointinterval() +
+  theme_light(18) +
+  labs(x = "Visitation probability", y= "Morphotype") +
+  guides(fill = FALSE)
+
+ggsave(file = "../output/brm_pvis01_plot1_largeprint.png", width = 7, height = 5)
+
 ```
@@
 
-diet_breadth <- full_join(taxon_diet_breadth, morph_diet_breadth)
+diet_breadth <- full_join(taxon_diet_breadth, morph_diet_breadth) %>%
+  left_join(site_data)
 
@@
   geom_smooth(method = "lm")
+
+ggplot(diet_breadth, aes(elev.mean, tax.diet.breadth)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~bb.sp)
+
+ggplot(diet_breadth, aes(elev.mean, morph.diet.breadth)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~bb.sp)
 ```

### 2021-02-13 writing/manuscript_v5.Rmd
update

@@
 ---
-title: "Bumble bees exhibit strong but flexible foraging biases through temporal and elevational gradients of floristic turnover"
+title: Bumble bees exhibit strong but flexible foraging biases through temporal and
+  elevational gradients of floristic turnover
 author:
@@
 output:
+  word_document: default
   html_document:
@@
 At an evolutionary level, the innate biases of bumble bees for certain floral traits likely reflect historic associations between floral traits and floral rewards [@Raine2007-nq].
-These innate biases, however, can be modulated or even nullified by patterns of reward learned during foraging [@Kugler1943-uy; Maharaj2019-wm], which depend not only on intrinsic floral properties but also ecological modulators, such as the spatial distribution of floral patches or competition with other bumble bees [@Ishii2013-sm; @Plowright1984-lc; @Sowig1989-ld].
+These innate biases, however, can be modulated or even nullified by patterns of reward learned during foraging [@Kugler1943-uy; @Maharaj2019-wm], which depend not only on intrinsic floral properties but also ecological modulators, such as the spatial distribution of floral patches or competition with other bumble bees [@Ishii2013-sm; @Plowright1984-lc; @Sowig1989-ld].
 
@@
 @Kugler1943-uy; @Raine2007-nq] and observational field study [e.g.
-Heinrich1976-wc; @Goulson2008-zy].
+@Heinrich1976-wc; @Goulson2008-zy].
 The former approach enables stronger causal inference with respect to the mechanisms of floral choice, but it suffers from all the usual limitations of biological realism associated with laboratory conditions.
@@
 In the present study, we focus on the question of bumble bee foraging bias with respect to floral morphology, and the lability of morphological biases across gradients of floristic variation.
-At the outset, we emphasize the use of the term "bias" in favor of the term "preference", since the latter is laden with more assumptions and its inference under field conditions is dubious [].
+At the outset, we emphasize the use of the term "bias" in favor of the term "preference", since the latter is laden with more assumptions and its inference under field conditions is problematic [@Williams2005-yp].
 By "bias", we mean simply an empirical disproportionality between the relative abundance of a given floral taxon or group and its proportional visitation.
@@
 --- is beyond the scope of our discussion, and probably best studied under laboratory conditions where strict factorial control is possible.
-We focus instead on the empirical phenomenon of bias and its morphological correlations.
+We focus instead on the empirical phenomenon of bias, its variation, and its morphological correlations.
 
@@
 Such matching is entirely bypassed by nectar robbers, but even in the case of "legitimate" foraging, access to reward involves learning to manipulate floral structures that, in some floral morphotypes, are challenging irrespective of tongue length [@Laverty1994-hm].
-For example, the characteristic Papilionaceous floral morphology of legumes requires a bee to separate tightly appressed keel petals to access anthers and nectaries [@Westerkamp1997-hs], and the narrowly campanulate flowers of many Ericaceous flora pose a stringent barrier to large-bodied visitors like bumble bees [@Wallace1977-kz].
-Rather than considering various floral morphological traits independently, we will rely on the synthesis of @Kugler1970-gb, who developed a typology consisting of ten primary morphotypes and nested subdivision of each.
+For example, the characteristic Papilionaceous floral morphology of legumes requires a bee to separate tightly appressed keel petals to access anthers and nectaries [@Westerkamp1997-hs], and the narrowly campanulate flowers of many Ericaceous flora pose a significant barrier to large-bodied visitors like bumble bees [@Wallace1977-kz].
+Rather than considering various floral morphological traits independently, we rely on the synthesis of @Kugler1970-gb, who developed a typology consisting of ten primary morphotypes and nested subdivision of each.
 Kugler's typology was informed by laboratory studies of bumble bee flower choice [@Kugler1943-uy], and there is a precedent for analyzing bumble bee visitation in terms of Kugler's morphotypes [@Neumayer1999-il; @Schneller2014-ry].
 
-First, we compare the overall rate of visitation across floral morphotypes.
+First, we compare the overall rate of bumble bee visitation across floral morphotypes.
 Then, we parse visitation rates by bumble bee species and explore the functional relationship between floral morphotype and bumble bee tongue length.
-We then compare overall diet similarity across bumble bee species as an emergent consequence of their respective biases.
-Lastly, we analyze the flexibility of the morphological preferences of individual bumble bee species through elevation and time.
+These biases, together with the background effects of relative abundance, generate emergent patterns of diet (dis)similarity, which we present in the form of ordination analysis.
+Lastly, we analyze the lability of the morphological biases of individual bumble bee species through elevation and time, and compare these patterns of bias across bumble bee species and within floral morphtypes.
+The results of these analyses demonstrate both the strength of bumble bee floral bias, arising in part from morphological constraints, and the flexibility of bumble bee floral bias that enables them to forage adaptively in the dynamic floral communities of mountain landscapes.
 
@@
 Fourteen of these pastures are extensively grazed by cattle or sheep, three are mowed for hay production, and eight have been abandoned throughout the last century and are no longer subject to any human management.
-The sites sampled in this study are the same as those used in a series of previous studies [@Hoiss2012-sn; @Hoiss2012-kr; Hoiss2015-gv], but the data reported here are otherwise independent.
+The sites sampled in this study are the same as those used in a series of previous studies [@Hoiss2012-sn; @Hoiss2012-kr; @Hoiss2015-gv], but the data reported here are otherwise independent.
 
@@
 
-floral survey sample = site-date of floral surveying
+*I need Kalli's help to write this section.*
 
@@
 Floral morphology was summarized using floral morphotypes of @Kugler1970-gb, accessed via the BIOFLOR database [@Klotz2002-uc].
-Morphotypes were simplified to the basic categories numbered 0-10: 0 = not applicable (includes wind-pollinated flora such as grasses), 1 = disc- and bowl-shaped flowers, 2 = funnel flowers, 3 = bell-shaped flowers, 4 = stalk-disc flowers, 5 = lip flowers, 6 = flag flowers, 7 = flower heads, 8 = spike flowers, 9 = brush flowers, 10 = trap flowers **(Figure X)**.
-<!-- I think I would like to include a figure with photographs of flowers exemplifying each morphotype --> In a few cases, a plant species in our data set was not found in the BIOFLOR database.
+Morphotypes were simplified to the primary classes numbered 0-10: 0 = not applicable (includes wind-pollinated flora such as grasses), 1 = disc- and bowl-shaped flowers, 2 = funnel flowers, 3 = bell-shaped flowers, 4 = stalk-disc flowers, 5 = lip flowers, 6 = flag flowers, 7 = flower heads, 8 = spike flowers, 9 = brush flowers, 10 = trap flowers.
+In a few cases, a plant species in our data set was not found in the BIOFLOR database.
 In most cases, morphotypes are constant at the genus level, and missing morphotypes were added based on the morphotypes found for congeners.
@@
 We then modeled binary visitation using generalized models with binomial or Bernoulli distribution families.
-The conditional effects inferred from these models should, therefore, be interpreted as effects on *visitation probability* rather than visitation rate.
+The conditional effects inferred from these models should, therefore, be interpreted as effects on *visitation probability* rather than visitation frequency
 
@@
 
-To visualize the net diet similarity between bumble bee species, we pooled visitation data across all samples within bumble bee species and plant species, then performed a principal coordinates analysis based on the Jaccard distance metric.
+To visualize the net diet similarity between bumble bee species, we pooled visitation data across all samples within bumble bee species and plant species, then performed a principal coordinates analysis based on Horn-Morisita distance metric.
 To test the significance of tongue length class as a predictor of diet similarity, we performed a permutation test.
@@
 wurflenii*) and the most visited floral morphotypes (groups 3, 5, 6, and 7).
-We then constructed a generalized additive model in which binary visitation was modeled as a tensor-product smooth of *elevation* and *day-of-year*, grouped by floral morphotype.
+We then constructed a generalized additive model (GAM) in which binary visitation was modeled as a tensor-product smooth of *elevation* and *day-of-year*, grouped by floral morphotype.
 This response was conditioned on linear interaction terms for *bumble bee abundance* : *percent flower cover* and *percent flower cover* : *floral morphotype* and a random intercept term for *bumble bee species* : *floral morphotype*.
@@
 All analyses were conducted in R v.
-3.6.
-Bayesian multilevel models were implemented in *Stan* [@Stan2020aa] via the R package *brms* [@Buerkner2018aa], and visualizations were constructed with the R packages *tidybayes* [@Kay2020aa] and *ggdist* [Kay2020ab].
+3.6 [@Rcore2020].
+Bayesian multilevel models were implemented in *Stan* [@Stan2020aa] via the R package *brms* [@Buerkner2018aa], and visualizations were constructed with the R packages *tidybayes* [@Kay2020aa] and *ggdist* [@Kay2020ab].
 Ordination analysis and permutation testing were performed with the R package *vegan* [@Oksanen2019aa] and visualized using the R packages *ggplot2* [@Wickham2016aa] and *ggvegan* [@Simpson2019aa].
 GAMs were implemented with the R package *mgcv* [@Wood2017aa] and visualized with the R package *mgcViz* [@Fasiolo2018aa].
-Data handling was performed the *tidyverse* suite of R packages [Wickham2019aa].
+Data handling was performed the *tidyverse* suite of R packages [@Wickham2019aa].
 Input data and code are available in Appendix X.
@@
 
-Lip (group 5) and flag (group 6) flowers had the highest overall visitation probability, with a conditional mean visitation probabilities of 0.36 (95% CI = 0.30--0.42) and 0.33 (95% CI = 0.27--0.39), respectively **(Figure X)**.
+Lip (group 5) and flag (group 6) flowers had the highest overall visitation probability, with a conditional mean visitation probabilities of 0.36 (95% CI = 0.30--0.42) and 0.33 (95% CI = 0.27--0.39), respectively **(Figure 1)**.
 In second place, head flowers (group 7) had a mean conditional visitation probability of 0.19 (95% CI = 0.16--0.24).
@@
 
-Descriptions of tongue classes are based on Model 2 and refer to **(Figure X)**, and descriptions of species are based on Model 3 and refer to **(Figure X)**.
+Descriptions of tongue classes are based on Model 2 and refer to **Figure 2**, and descriptions of species are based on Model 3 and refer to **Figure 3**.
 Both models explained approximately 28% of total variance (97.5% CI = 27--29%).
@@
 
-Axis 1 accounted for 24% of total diet variance and primarily separated long- and medium-tongued species from short-tongued species.
+Axis 1 accounted for 24% of total diet variance and primarily separated long- and medium-tongued species from short-tongued species **(Figure 4)**.
 Long- and medium-tongued species scored low on Axis one and were associated with the lip flowers (group 5) *Ajuga reptans*, *Rhinanthus glacialis*, and *Stachys alopecuros*) and flag flowers (group 6) *Anthyllis vulneraria*, *Lotus corniculatus*, and *Trifolium pratense*.
@@
 Model 5 explained 33.8% of total variance (adjusted R2 = 0.37).
-All tensor product smooths were significant at p < 0.001 except for the one representing *B.
+All tensor product smooths **(Figure 5)** were significant at p < 0.001 except for the one representing *B.
 pratorum* and flag (group 6), which was more weakly significant at p = 0.02.
@@
 
-Visitation rates for head (group 7) flowers were more or less bimodal for all species, with a first peak in mid-season and low elevation and a second peak in late-season and high elevation.⎄
+Visitation rates for head (group 7) flowers were more or less bimodal for all species, with a first peak in mid-season and low elevation and a second peak in late-season and high elevation.
 
@@
 
-When bumble bee visitation is pooled across all species, it is clear that lip, flag, and head flowers dominate the rest of the floral community in terms of conditional visitation probability.
+When bumble bee visitation is pooled across all species, it is clear that lip, flag, and head flowers dominate the floral community in terms of conditional visitation probability.
 Patterns of pooled visitation are, of course, driven disproportionately by the behavior of the most abundant bumble bee species, which in our system were *B.
@@
 In the cases of disc, and stalk-disc, and unclassified flowers, visitation rates were uniformly low across all tongue length classes.
-The opposite was true of lip flowers, which were visited at high rates by all tongue length classes.
+The opposite was true of lip flowers, which were visited at similarly high rates by all tongue length classes.
 Bell flowers and head flowers were visited chiefly by short-tongued bees, consistent with the shallow corollas of these morphotypes.
@@
 
-While variation in floral bias across tongue length classes largely comports with the established relationship between tongue length and corolla depth, variation in bias *within* tongue length classes demonstrates the insufficiency of simple morphological matching to explain bumble bee foraging patterns.
+While the variation in floral bias across tongue length classes we observed largely comports with the established relationship between tongue length and corolla depth, variation in bias *within* tongue length classes demonstrates the insufficiency of simple morphological matching to explain bumble bee foraging patterns.
 Nectar robbing is one mechanism by which the relationship between tongue length and corolla depth might be relaxed, but its influence in our data is unclear.
@@
 terrestris/lucorum* and especially *B.
-wurflenii* are known to engage readily in nectar robbing.
+wurflenii* are known to engage frequently in nectar robbing [@Brian1954-cy].
 In the case of *B.
@@
 lapidarius* visited them relatively rarely.
-There is no obvious morphological distinction between these groups of short-tongued bees that would account for their divergent affinities for bell flowers; ecological drivers such as competition or differences in spatial or temporal activity patterns are more likely explanations.
+A similar split among the short-tongued species occurred with respect to head flowers, which were visited at high rates by *B.
+soroensis* and *B.
+psithyrus* but at relatively low rates by other short-tongued species.
+There are no obvious morphological distinctions within short-tongued bumble bees that would account for these divergent affinities; ecological drivers such as competition or differences in spatial or temporal activity patterns are more likely explanations, as will be revisited below in our discussion of overall diet similarity.
 
-Marked differences in the relative visitation rates of floral morphotypes compared to [@Schneller2014-ry], but also some consistencies.
-We reproduce their finding that *B.
-soroensis* is partial to bell flowers.
-There data collection was limited to July and August, which probably explains their finding that head (group 7) flowers were most visited, followed by lip (group 5) flowers.
-Bell flowers were strongly underrepresented in their data, but that makes sense given what I see in my smooth plots.
-They are most attractive early in the year.
+## 4.2 Diet similarity
 
-## 4.2 The emergent phenomenon of diet similarity
-
-Our finding that bumble bee diet ordination produces a first axis corresponding to tongue length and a second axis describing the dispersion of short-tongued species is strikingly consistent with past studies [@Goulson2008-zy; Durieux2000-uk].
+Our finding that bumble bee diet ordination produces a primary axis corresponding to tongue length --- and associated with affinity for legumes --- and a secondary axis describing the dispersion of short-tongued species is consistent with studies in Poland [@Goulson2008-zy], France [@Durieux2000-uk], and England [@Goulson2005-nz] that analyzed bumble bee diet by similar methods, suggesting that this may be a general pattern among bumble bee communities, at least within temperate Europe.
+While our test of beta-dispersion did not reveal significant differences between tongue length classes, the spread of short-tongued bees along the secondary axis of our ordination, together with their divergence with respect to visitation rates on bell and head flowers described above, suggests that resource partitioning may be more pronounced among short-tongued bumble bees than among their medium- and long-tongued counterparts.
+It is tempting to invoke interspecific competition as a cause of the apparent niche partitioning among short-tongued bumble bees, in agreement with previous studies that have demonstrated the potential for severe competition between bumble bees of similar tongue length [e.g.
+@Heinrich1976-wc; @Inouye1978-qi; @Pyke1982-gn].
+But the importance of competition in bumble bee communities is not universal [e.g.
+@Ranta1981-vx; @Williams1989-vr], and in the rich wildflower meadows of our (sub)alpine system, bumble bee populations are more likely limited by harsh abiotic conditions than by food availablility.
+Alternatively, it may be that short-tongued bees in our study system differed from one another in their affinities for specific habitats [@Schneller2014-ry] or spatial configurations of resource patches [@Sowig1989-ld], and their divergent morphological biases arose secondarily from these effects.
 
@@
 
-Species differed not just in their overall bias toward given morphotypes, but also in *when* and *where* they were biased toward a morphotype.
+While the intraspecific lability of bumble bee foraging biases has been demonstrated under laboratory conditions [@Ings2009-vc; @Maharaj2019-wm] and in a localized field study [@Raine2007-nq], ours is the first study, to our knowledge, to characterize the *in situ* lability of foraging biases within multiple species through elevational and temporal gradients of floristic turnover.
+In our analysis, it is evident that morphological biases in bumble bees, insofar as they are captured by the morphotypology used in our study, are not fixed but rather tuned to floristic (and possibly abiotic) context and therefore dynamic through elevation and time.
+Notably, bumble bee species differed from one another not just in their overall bias toward given morphotypes but also in *when* and *where* they were biased toward a given morphotype.
+This is especially evident in the case of flag flowers, for which positive bias was maximized in mid-season for *B.
+hortorum*, *B.
+pascuorum*, *B.
+terrestris/lucorum*, and *B.
+wurflenii* but in early-season for *B.
+pratorum* and *B.
+soroensis*.
+This pattern corresponds to the overall biases of these species toward flag flowers, which were primary resources for *B.
+hortorum*, *B.
+pascuorum*, *B.
+terrestris/lucorum*, and *B.
+wurflenii* but only marginal resources for *B.
+pratorum* and *B.
+soroensis*.
 
-Ours is, to our knowledge, the first study to characterize formally the variation of bumble bee foraging bias through elevation an time.
+## 4.4 Conclusion
+
+For bumble bees, floral bias is a labile product of innate preferences, the soft constraints of morphological compatibility, and the dynamic ecological context in which foraging occurs.
+By conditioning bumble bee visitation data upon concurrent floral abundance data across spatially and temporally replicated samples, we demonstrate the strength bumble bee floral bias, its functional underpinnings, its variation across species, and its lability within species in response to environmental variation.
+Such tuning of floral bias has obvious adaptive significance for bumble bees, which are characteristically --- and most likely primordially --- associated with mountain habitats [@Williams1985-ql; @Hines2008-er], where steep temperature gradients and short growing seasons generate extreme spatial and temporal floral turnover.
 

### 2021-02-18 code/beta_diversity.Rmd
revised ms1

@@
 
-# Binomial smooth function (thanks, Hadley): https://ggplot2.tidyverse.org/reference/geom_smooth.html
-binomial_smooth <- function(...) {
-  geom_smooth(method = "glm", method.args = list(family = "binomial"), ...)
-}
-
-# Site_pair prep function
-# site_pair_prep_net <- function(dat, y) {
-#   dat %>%
-#     ungroup() %>%
-#     filter(year1 == year2 & 
-#              year1 == y & 
-#              site1 != site2) %>% # I don't think we want to compare samples through within sites, introduces a new form of non-independence.
-#     filter(metric == "WN") %>% # we'll focus for now on WN, but this could be done for partitions, too
-#     mutate(weights = rep(1, dim(.)[1])) %>% # right now this analysis is unweighted; could weight by abundance, etc, if we wanted to
-#     dplyr::select(distance = value, weights,
-#                   s1.xCoord = x1, s1.yCoord = y1, 
-#                   s2.xCoord = x2, s2.yCoord = y2, 
-#                   s1.elev.mean = elev1, s1.yday = yday1,
-#                   s2.elev.mean = elev2, s2.yday = yday2) %>%
-#     data.frame()
+# # Binomial smooth function (thanks, Hadley): https://ggplot2.tidyverse.org/reference/geom_smooth.html
+# binomial_smooth <- function(...) {
+#   geom_smooth(method = "glm", method.args = list(family = "binomial"), ...)
 # }
@@
     filter(s1.elev.mean != s2.elev.mean) # this drops within-site comparisons
-}
-
-## wrapper for extracting splines and prepping tibble for plotting
-spline_prep <- function(x, y) {
-  isplineExtract(x) %>%
-    as.data.frame() %>%
-    tibble() %>%
-    mutate(x.Geographic = x.Geographic/1000) %>%
-    pivot_longer(cols = c(x.Geographic, x.elev.mean), 
-                 names_to = "x.var", values_to = "predictor", 
-                 names_prefix = "x.") %>%
-    pivot_longer(cols = c(y.Geographic, y.elev.mean), 
-                 names_to = "y.var", values_to = "response", 
-                 names_prefix = "y.") %>%
-    filter(x.var == y.var) %>%
-    mutate() %>%
-    dplyr::select(-y.var)
 }
@@
 Note that all this deduplication nonsense is not necessary for the betalinkr call because it deduplicates internally.
+
+# fl_beta_dist <- survey %>%
+#   group_by(site, plant.sp) %>%
+#   summarize(count = sum(flower.cover)) %>%
+#   pivot_wider(names_from = plant.sp, values_from = count) %>%
+#   replace(is.na(.), 0) %>%
+#   column_to_rownames("site") %>%
+#   vegdist(method = "jaccard", binary = TRUE, upper = TRUE) 
+# 
+# bb_beta_dist <- net %>%
+#   group_by(site, bb.sp) %>%
+#   summarize(count = n()) %>% 
+#   pivot_wider(names_from = bb.sp, values_from = count) %>%
+#   replace(is.na(.), 0) %>%
+#   column_to_rownames("site") %>%
+#   vegdist(method = "jaccard", binary = TRUE)
 ```
@@
 
-p <- c("OS" = "#e7298a", "ST" = "#d95f02", 
-       "ST.h" = "#7570b3", "ST.l" = "#1b9e77", 
-       "ST.lh" = "#66a61e", "WN" = "black")
+p1 <- c("OS" = "#e7298a", "ST" = "#66a61e", "WN" = "black")
+
+p2 <- c("ST.h" = "#7570b3", "ST.l" = "#1b9e77", "ST.lh" = "#d95f02")
 
@@
   #binomial_smooth(formula = y ~ splines::ns(x, 3)) +
-  geom_smooth(method = "lm", se = FALSE) +
+  #geom_smooth(method = "lm", se = FALSE) +
+  geom_smooth(method = "glm", 
+              method.args = list(family = "binomial"), 
+              se = FALSE) +
   #facet_wrap(~group, ncol = 1) +
@@
 
-# BB * FL elevation diff
+# BB * FL elevation diff (aggregate)
 g2 <- ggplot(filter(beta_mit_data, 
@@
                       metric != "S" & # drop this metric
-                      !metric.class %in% c("OS.partition", "WN.partition")), 
+                      metric.class == "aggregate"), 
              aes(elev.diff, value, color = metric))  +
   geom_point(alpha = 0.20, pch = 16) +
-  binomial_smooth(formula = y ~ splines::ns(x, 3), se = FALSE) +
-  scale_color_manual(values = p, name = "partition") +
+  #binomial_smooth(formula = y ~ splines::ns(x, 3), se = FALSE) +
+  #geom_smooth(method = "gam", formula = y ~ s(x, k  = 4), se = FALSE) +
+  geom_smooth(method = "glm", 
+              method.args = list(family = "binomial"), 
+              se = FALSE) +
+  scale_color_manual(values = p1, name = "β partition") +
+  xlab(NULL) +
+  ylab("Prop. β-diversity") +
+  theme_light()
+
+# BB * FL elevation diff (ST partition)
+g3 <- ggplot(filter(beta_mit_data, 
+                      site1 != site2 & # no within-site comparisons
+                      metric.class ==  "ST.partition"), 
+             aes(elev.diff, value, color = metric))  +
+  geom_point(alpha = 0.20, pch = 16) +
+  #binomial_smooth(formula = y ~ splines::ns(x, 3), se = FALSE) +
+  #geom_smooth(method = "gam", formula = y ~ s(x, k  = 4), se = FALSE) +
+  geom_smooth(method = "glm", 
+              method.args = list(family = "binomial"), 
+              se = FALSE) +
+  scale_color_manual(values = p2, name = "ST partition") +
   xlab("Elevation difference (m)") +
-  ylab(NULL) +
+  ylab("Prop. species turnover") +
   theme_light()
@@
 png("../output/beta_stack.png", width = 6.5, height = 4.5, units = "in", res = 300)
-(g1 | g2) + 
-  plot_annotation(tag_levels = c("A", "B")) + 
+(g1 | (g2 / g3)) + 
+  plot_annotation(tag_levels = c("A", "B", "C")) + 
   #plot_layout(guides = "collect") & 
@@
 dev.off()
+```
+
+### MRM analysis
+```{r}
+library(ecodist)
+
+survey_sample_diff <- survey %>%
+  group_by(site) %>%
+  summarize(samples = length(unique(date))) %>%
+  column_to_rownames("site") %>%
+  vegdist(method = "euclidean", upper = TRUE) %>%
+  as.matrix() %>%
+  as.data.frame() %>%
+  rownames_to_column(var = "site1") %>%
+  pivot_longer(cols = -site1, names_to = "site2", values_to = "sample.diff")
+
+net_sample_diff <- net %>%
+  group_by(site) %>%
+  summarize(samples = length(unique(date))) %>%
+  column_to_rownames("site") %>%
+  vegdist(method = "euclidean", upper = TRUE) %>%
+  as.matrix() %>%
+  as.data.frame() %>%
+  rownames_to_column(var = "site1") %>%
+  pivot_longer(cols = -site1, names_to = "site2", values_to = "sample.diff")
+
+mrm_survey_data <- species_beta %>%
+  filter(group == "FL" & site1 != site2) %>%
+  left_join(survey_sample_diff) %>%
+  dplyr::select(site1, site2, beta, elev.diff, sample.diff) %>%
+  mutate(elev.diff = scale(elev.diff), # gotta scale variables for MRM
+         sample.diff = scale(sample.diff))
+
+mrm_net_data <- species_beta %>%
+  filter(group == "BB" & site1 != site2) %>%
+  left_join(net_sample_diff)  %>%
+  dplyr::select(site1, site2, beta, elev.diff, sample.diff) %>%
+  mutate(elev.diff = scale(elev.diff),
+         sample.diff = scale(sample.diff))
+
+mrm_beta_data <- beta_mit_data %>%
+  filter(site1 != site2) %>%
+  left_join(net_sample_diff)  %>%
+  dplyr::select(site1, site2, value, metric, elev.diff, sample.diff) %>%
+  mutate(elev.diff = scale(elev.diff),
+         sample.diff = scale(sample.diff))
+
+mrm_fl <- MRM(beta ~ elev.diff + sample.diff, 
+              data = mrm_survey_data,
+              nperm = 1000,
+              method = "logistic")
+
+mrm_bb <- MRM(beta ~ elev.diff + sample.diff, 
+              data = mrm_net_data,
+              nperm = 1000,
+              method = "logistic")
+
+mrm_WN <- MRM(value ~ elev.diff + sample.diff, 
+              data = filter(mrm_beta_data, metric == "WN"),
+              nperm = 1000,
+              method = "logistic")
+
+mrm_ST <- MRM(value ~ elev.diff + sample.diff, 
+              data = filter(mrm_beta_data, metric == "ST"),
+              nperm = 1000,
+              method = "logistic")
+
+mrm_ST.l <- MRM(value ~ elev.diff + sample.diff, 
+              data = filter(mrm_beta_data, metric == "ST.l"),
+              nperm = 1000,
+              method = "logistic")
+
+mrm_ST.h <- MRM(value ~ elev.diff + sample.diff, 
+              data = filter(mrm_beta_data, metric == "ST.h"),
+              nperm = 1000,
+              method = "logistic")
+
+mrm_ST.lh <- MRM(value ~ elev.diff + sample.diff, 
+              data = filter(mrm_beta_data, metric == "ST.lh"),
+              nperm = 1000,
+              method = "logistic")
+
+mrm_OS <- MRM(value ~ elev.diff + sample.diff, 
+              data = filter(mrm_beta_data, metric == "OS"),
+              nperm = 1000,
+              method = "logistic")
+
+# Partial Mantel tests are another option, perhaps more routine to interpret
+# mantel_fl <- mantel(beta ~ elev.diff + sample.diff, 
+#               data = mrm_survey_data,
+#               mrank = TRUE)
+# 
+# mantel_bb <- mantel(beta ~ elev.diff + sample.diff, 
+#               data = mrm_net_data,
+#               mrank = TRUE)
+# 
+# mantel_WN <- mantel(value ~ elev.diff + sample.diff, 
+#                     data = filter(mrm_beta_data, metric == "WN"),
+#                     mrank = TRUE)
+# 
+# mantel_ST <- mantel(value ~ elev.diff + sample.diff, 
+#                     data = filter(mrm_beta_data, metric == "ST"),
+#                     mrank = TRUE)
+# 
+# mantel_ST.l <- mantel(value ~ elev.diff + sample.diff, 
+#                     data = filter(mrm_beta_data, metric == "ST.l"),
+#                     mrank = TRUE)
+# 
+# mantel_ST.h <- mantel(value ~ elev.diff + sample.diff, 
+#                     data = filter(mrm_beta_data, metric == "ST.h"),
+#                     mrank = TRUE)
+# 
+# mantel_ST.lh <- mantel(value ~ elev.diff + sample.diff, 
+#                     data = filter(mrm_beta_data, metric == "ST.lh"),
+#                     mrank = TRUE)
+# 
+# mantel_OS <- mantel(value ~ elev.diff + sample.diff, 
+#                     data = filter(mrm_beta_data, metric == "OS"),
+#                     mrank = TRUE)
 ```

### 2021-02-18 ms_1/ms_1.Rmd
revised ms1

---
title: "Bumble bees and wildflowers along a mountain slope exhibit nonlinear patterns of abundance and turnover, punctuated by the tree line ecotone"

author: Douglas B. Sponsler^1^*, Fabrice Requier^2^, Katharina Kallnik^1^, Alice Claßen^1^, A. Fabienne Maihoff^1^, Ingolf Steffan-Dewenter^1^

# date: `r Sys.Date()`

output:
  bookdown::word_document2:
    fig_caption: yes
    fig_width: 7
    fig_height: 7
    df_print: kable
    keep_md: yes
    number_sections: no
    toc: no
    reference_docx: Word_template.docx

bibliography: 
  - references.bib
#  - knitcitations.bib
csl: ecology.csl
# csl: https://www.zotero.org/styles/ecology

---

^1^ Department of Animal Ecology and Tropical Biology, Biocenter, University of Würzburg, Würzburg, Germany

^2^ Université Paris-Saclay, CNRS, IRD, UMR Évolution, Génomes, Comportement et Écologie, 91198, Gif-sur-Yvette, France

`*` Corresponding author: douglas.sponsler@uni-wuerzburg.de

# ABSTRACT

*Keywords*: climate, elevation gradient, pollinator, *Bombus*, alpine

```{r setup, include=FALSE, cache=FALSE, message = FALSE}

library("knitr")

### Chunk options: see http://yihui.name/knitr/options/ ###
## Text results
opts_chunk$set(echo = FALSE, warning = TRUE, message = TRUE)

## Code decoration
opts_chunk$set(tidy = TRUE, comment = NA, highlight = TRUE)

## Cache
# opts_chunk$set(cache = 2, cache.path = "output/cache/")

## Plots
opts_chunk$set(fig.path = "../output/")

```

```{r knitcitations, cache = FALSE}
library(knitcitations)
cleanbib()   
cite_options(citation_format = "pandoc")
```

# INTRODUCTION

The rate of modern climate change is forcing species to adapt ecologically within the constraints of gene pools shaped by cooler climates [@Visser2008-ct]. As temperatures rise, species are expected to track suitable conditions to higher latitudes and higher elevations, and such effects are already measurable across a broad range of taxa [@Lenoir2015-tj]. Species range shifts, however, are not merely abstract, Cartesian translations along axes of latitude, elevation, and temperature; they are movements *in situ*, involving the advent, extinction, and reorganization of biological interactions [@Tylianakis2008-fi; @Blois2013-dz] as well as the concomitant variation of abiotic conditions other than temperature [@Korner1995-mh; @Hodkinson2005-af]. Moreover, some of these factors, unlike temperature, cannot be assumed to vary linearly with latitude or elevation but instead exhibit marked nonlinearities and thresholds. A salient example is the tree line ecotone that marks the elevational transition between forest and grassland. At this relatively discrete band within a continuous elevational temperature gradient, plant communities exhibit rapid species turnover [@Descombes2017-id] and declining diversity [@Becker2007-cg] as abiotic conditions become more extreme away from the shelter of tree canopy [@Slatyer1992-ir]. Thus, species range shifts --- and, perhaps more importantly, the ecological consequences thereof --- are contingent upon patterns of species distributions and biotic interactions, and community ecology is as much prior as posterior to the ecology of global change.

Bumble bees (Hymenoptera: *Bombus* spp.) are both a classic model organism in community ecology and an emergent conservation priority due to recent declines [@Goulson2008-oe] and manifest vulnerability to climate change [@Soroye2020-zb]. The genus *Bombus* is believed to have arisen in the mountains of Asia during the cooling climate of the Eocene-Oligocene boundary (~34 mya), and its subsequent spread and diversification along mountain corridors and into lowland habitats was likely driven by alternating range expansion during cooling periods and retreat to higher elevations during warming periods [@Hines2008-er; @Stewart2010-jn; @Martinet2018-rb]. Today, bumble bees remain characteristically cold-adapted species [@Heinrich1994-cv], with peak abundance and diversity associated with mountain ranges and northern latitudes, where they are often the dominant guild of flower-visiting insects and the principal pollinators of entomophilous flora [@Williams1998-xi; @Goulson2010-lo]. Climate-induced range shifts are expected to be especially pronounced in the mountain habitats that host the core of bumble bee biodiversity, together with their floral mutualists, and upward movement of bumble bee species has already been documented [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc]. Populations of bumble bee species already restricted to the highest elevation zones can be expected to suffer geometric habitat shrinkage and increasing genetic isolation as they move toward mountain peaks; the upward advance of lowland species may induce novel competitive pressures and disrupt the pollination networks in which bumble bees participate [@Brosi2013-mu; @Ishii2013-sm]; and concurrent elevational [@Lenoir2008-yn] and/or phenological [@Pyke2016-ll] shifts in the floral community could generate mismatches between bumble bee species and their historic floral hosts.

The outcomes of range shifts in mountain bumble bee and floral communities have obvious significance from a conservation perspective, but they also provide a model system in which to study the general phenomenon of community assembly under climate change, a phenomenon of both historical interest and prospective urgency. The goal of anticipating, interpreting, and potentially mitigating the effects of ranges shifts, however, is predicated upon understanding the current elevational and temporal patterns of bumble bee communities and those of their floral hosts. In particular, it is important to detect nonlinearities in species-, community-, and interaction-level responses to elevational climate gradients, such as the floristic threshold effect that has been documented at the tree line ecotone [@Descombes2017-id].

In the present study, we approach this topic at three levels of biological organization. First, we ask how mountain bumble bee species are distributed in elevation and intra-annual time with respect to one another and with respect to their floral hosts. Then, we move from the level of individual species to that of community composition and explore the structure of species turnover (i.e. β-diversity) of bumble bees and flora in response to elevation. Finally, propagate the question to the level of species interactions, investigating the turnover of bee-flower interaction partners through elevation. In each of these analyses, we employ flexible modeling techniques capable of capturing nonlinear responses to elevation and time with the goal of identifying where, when, and for which species the impacts of climate-induced range shifts should be expected to be most acute.

# METHODS

## Field system

The study was conducted in Berchtesgaden National Park (47.55°N, 12.92°E), located in the Northern Limestone Alps of southeast Germany. The landscape is composed of mountain pastures mainly surrounded by coniferous forests. We selected 25 study sites (60 x 60 m) on mountain pastures at elevations ranging from 641-2032 m above sea level. Fourteen of these pastures are extensively grazed by cattle or sheep, three are mowed for hay production, and eight have been abandoned throughout the last century and are no longer subject to any human management. The sites sampled in this study are the same as those used in a series of previous studies [@Hoiss2012-sn; @Hoiss2012-kr; @Hoiss2015-gv], but the data reported here are otherwise independent.

Sampling consisted of repeated visits to each study site at approximately weekly intervals. Samples were only collected on dry days when the air temperature was at least 6°C. During each visit, bumble bees and their activity were recorded during a 50-minute transect walk. Bumblebees observed on or a given flower were counted as floral visitors. Bumble bee queens were identified to species level in the field, while workers and males were stored in individually labeled tubes in the freezer for later identification in the laboratory after [@Amiet1996-jz]. Floral visitation by male bumble bees was recorded during visitation sampling, but we chose to analyze only visitation by queens and workers.

In conjunction with visitation observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.1 m^2^. Species identification followed @Lauber2007-ha and @Oberndorfer2001-ro.

## Data analysis

Bumble bee abundance was quantified as the total number of recorded floral visits per bumble bee species per site-date (excluding males). To represent floral abundance from the perspective of each bumble bee species in our network, we first scored each plant genus in our data set as either visited or not visited by each bumble bee species. We then calculated the total floral resource availability for each bumble bee species for each site-date by summing the observed flower cover of visited genera. We opted to work at the genus level based on the reasoning that if a bumble bee species visits one member of a given genus, other members of the same genus should also be considered potential floral hosts. This approach was intended to dampen the effects of false non-detection on the estimation of floral resource abundance, particularly for rare bumble bee species whose diet breadth would tend to be underestimated simply due to sparsity of observations [@Williams2005-yp].

We analyzed bumble bee abundance and floral abundance using separate hierarchical generalized additive models (HGAM) [@Wood2011-pg; @Pedersen2019-kf], with abundance modeled as a response to the interaction (tensor product) of elevation and day-of-year. Bumble bee species was included in each model as both a smoothing factor and a random intercept term, resulting in centered 2-dimensional abundance smooths for each bumble bee species. We also included year and site as random intercept effects to account for annual differences in abundance patterns and repeated measurements within sites.

To complement GAM-based analysis of abundance patterns through elevation and time, we also analyzed the site occupancy pattern of each bumble bee species using cluster and ordination analyses. We first calculated the mean abundance across dates of each bumble bee species for each site in our study system. Then, we applied a Hellinger transformation to the matrix of mean abundances and calculated the Horn-Morisita distance metric for each pair of species. The resulting distance matrix was then used to generate a principal coordinates analysis (PCoA) and a cluster dendrogram, each depicting the (dis)similarity between species pairs in their distribution across sites. The primary purpose of this analysis was to organize the presentation of the GAM results by grouping species with similar site affinities, but because our sites were arranged along an elevation gradient, our cluster and ordination analyses also highlight differences between species in elevational distribution.

For our analyses of species and interaction β-diversity, we opted to focus on elevational patterns by pooling samples within sites, since the sparse observations of individual sampling events tended to inflate β-diversity and its variance to a degree that impaired model fitting and interpretation. After pooling, we calculated the species and interaction β-diversity between all pairs of sites and partitioned total interaction β-diversity into its components of species turnover (further partitioned into bumble bee and plant turnover, respectively) and interaction rewiring [@Novotny2009-my; @Poisot2012-fk].

As an initial analysis, we plotted the relationship between site-wise elevation difference and each metric of β-diversity and verified the significance of the relationship using logistic matrix regression [@Goslee2009-ln], with the difference in number of sampling dates between sites as a covariate to control for potential confounding effects of sampling frequency.

To investigate the relationship between elevation and β-diversity more deeply, we performed a second analysis using generalized dissimilarity modeling (GDM) [@Ferrier2007-ji] to analyze the elevational variation in the β-diversity of bumble bees, flora, and (unpartitioned) interactions. In addition to estimating the relationship between differences in the response variable and differences in the predictor variables, GDM captures the slope of this relationship over the range of each predictor variable, revealing potential variation in the amount of change in the response induced by a given change in a predictor. This enabled us to ask whether the relationship between β-diversity and elevation exhibits thresholds or other nonlinearities. To control for potential confounding effects of geographic proximity and number of sampling dates per site, these terms were added as covariates in the GDM.

All analyses were conducted in `R` `r citep(citation())`. GAM analyses were performed with packages `mgcv` [@Wood_2017] and `mgcViz` [@Fasiolo2018aa], and the models we fit correspond to the “type I” model form described in @Pedersen2019-kf. PCoA and species β-diversity calculation were performed with the package `vegan` `r citep(citation("vegan"))`. Calculation and partitioning of interaction β-diversity were performed with the package `bipartite` `r citep(citation("bipartite"))`. Matrix regression was performed with the package `ecodist` [@Goslee2007-aa]. GDM analysis of β-diversity was performed with the package `gdm` `r citep(citation("gdm"))`. Data handling and visualization was performed with the `tidyverse` suite of R packages `r citep(citation("gdm"))`. Annotated R code is available in the Supplementary Material.

# RESULTS

## Summary of visitation and floral survey data

We recorded a total of 12,918 bumble- bee-flower interactions (excluding males) over the three years of our study. The metaweb across all sites and dates consisted of 16 bumble bee species (with *Bombus terrestric/lucorum* and *Bombus psithyrus* species groups lumped into species groups), 163 plant species (110 genera, 37 families), and 736 unique bumble- bee-plant interaction pairs. Five species --- *B. pascuorum*, *B. pratorum*, *B. soroensis*, *B. terrestris-lucorum*, and *B. wurflenii* --- accounted for the bulk of overall bumble bee abundance and were present over the entire elevation range of our sites **(Figure 1)**. *B. hortorum*, *B. jonellus*, and *B. psithyrus* (species-group) were widespread but at lower abundance. *B. monticola* occurred at moderate abundance but was rare or absent at sites below 1000 m. The remaining species were relatively rare in our study system. *B. mendax*, *B. mucidus*, and *B. pyrenaeus* occurred primarily above 1500 m, consistent with their known affinity for high elevation habitats [@Rasmont2010-ci]. *B. gerstaeckeri*, *B. hypnorum*, *B. lapidarius* were sparsely distributed across mid-elevation sites. *B. humilis* was recorded only three times and only in one year, so we omitted it from all analyses.

```{r fig1, echo = FALSE,  fig.cap = paste("Mean abundances and elevational distributions of bumble bee species. Sampled points along the elevation gradient are depicted as circles, the diameter of which is scaled to the mean abundance (pooled across study years) of the given bumble bee species at that elevation. Abundances of zero are not plotted. Lines span between the highest and lowest occurrence of each species, highlighting its observed elevation range.")}
knitr::include_graphics("../output/bb_range.png")
```

Floral surveying yielded a total of 352 plant species, representing 191 genera and 52 families. Of these, 155 species from 103 genera and 35 families were observed to be visited by bumble bees. Eight species --- *Rubus idaeus*, *Rosa canina*, *Juniperus communis*, *Larix decidua*, *Salix* sp., *Caltha palustris*, *Pulmonaria officinalis*, and *Rheum barbarum* --- were recorded in visitation data but not recorded during floral surveying. Each accounted for no more than 4 visits in total over the three years of our study, and they were omitted from the analysis of floral abundance.

## Abundance over elevation and time

Cluster **(Figure 2A)** and ordination **(Figure 2B)** analyses highlighted the strong differences in site affinity between the high-elevation species *B. mucidus*, *B. monticola*, *B. mendax*, and *B. pyrenaeus* and the rest of the bumble bee community. PCoA captured 59% of total variation in site affinity.

Elevation-time GAM smooths **(Figure 2C)** of bumble bee abundance differed significantly from a flat surface (p < 0.00001) for all species except *B. hypnorum* (p = 0.8). Overall, the model explained 57.6% of total deviance (adjusted R^2^ = 0.346). The distinction between *B. mucidus*, *B. monticola*, *B. mendax*, and *B. pyrenaeus* and the rest of the community was again clear; all 4 high-elevation species exhibited peak abundance at the upper extreme of our elevation gradient, and only *B. monticola* was found at high abundance below 1500 m. *B. monticola* also peaked in abundance earlier in the season than the other high-elevation species.

*B. gerstaeckeri* and *B. pratorum* exhibited well-defined mid-elevation abundance peaks, with that of *B. gerstaeckeri* occurring very late in the season. The apparent mid-elevation abundance peak of *B. hypnorum* was not significant and should not be interpreted strongly.

*B. pascuorum* and *B. psithyrus* had well-defined low-elevation peaks, with that of *B. psithyrus* occurring early in the season, during the colony-founding --- or, in the case of *B. psithryus*, colony-usurping --- phase of the bumble bee life cycle.

The remaining species exhibited relatively diffuse patterns of abundance. Of these, *B. soroensis* was exceptional in that it exhibited peak abundance at the highest sites, but it was also found at relatively high abundance across the whole elevation gradient, peaking at mid-season. *B. hortorum* and *B. lapidarius* peaked at low elevation. *B. jonellus* exhibited little elevational variation in abundance until declining above 1500 m, but its abundance was strongly patterned in time, peaking early in the season. *B. terrestris/lucorum* was generally abundant across the whole elevation gradient and throughout the whole season, but it showed a weakly trimodal pattern in elevation, with abundance peaks at low, middle, and high sites. Its strongest abundance peak occurred at low elevation, early in the season during colony-founding. *B. wurflenii* was broadly abundant above 1000 m but was rarer at lower sites.

Elevation-time GAM smooths of floral abundance differed significantly from a flat surface for all species (p << 0.00001) and explained 64.8% of total deviance (adjusted R^2^ = 0.425). Floral resource availability exhibited similar patterns for most bumble bee species, with peak floral resources occurring below 1500 m and during mid-season **(Figure 2)**. There was a noticeable temporal lag in floral resource availability associated with increasing elevation, consistent with expected temperature constraints on plant phenology. The only species to exhibit marked anomalies in floral resource availability was *B. gerstaeckeri*, which specializes on the genus *Aconitum*[@Ponchau2006-af] but was also observed (presumably collecting nectar) on *Gentiana*, *Salvia*, *Rhinanthus*, *Cirsium*, and *Clinopodium.* Its floral resources were concentrated in two discrete peaks, the first at low elevation and mid-season, and the second at high elevation and late-season.

```{r fig2, echo = FALSE,  fig.cap = paste("Abundance of bumble bee species and their floral resources through elevation and time. For each bumble bee species, surface plots are arranged in two columns labeled BB and FL. Surface plots in the BB columns show the abundance of the bumble bee species, while plots in the FL columns show the abundance of floral resources for the corresponding bumble bee species. The color ramp of each smooth has been scaled to equal color range, thereby obscuring the absolute magnitude of effect. To avoid cluttering the figure, we have opted not to include a unique color ramp key for each plot, since the focus of our inference is on relative rather than absolute patterns. Surface plots with original color ramp scales are available in Appendix X.")}
knitr::include_graphics("../output/gam_BB_FL4.png")
```

## Species and interaction β-diversity

Floral β-diversity among our study sites was very high overall and responded steeply to elevation difference between sites **(Figure 3A)**. Even sites at similar elevation exhibited ~40% species turnover, and the most widely separated sites (elevation difference > 1250 m) differed by more than 85%.

Bumble bee β-diversity was, in comparison to floral β-diversity, both lower overall and less responsive to elevational difference between sites **(Figure 3A)**. Sites at similar elevation exhibited ~25%
species turnover, and species turnover between the most widely separated sites remained less than 50%.

Total interaction β-diversity was >75% between sites at similar elevation and approached perfect dissimilarity in the most widely separated sites **(Figure 3B)**. Partitioning revealed that total interaction β-diversity was driven primarily by species turnover, which accounted for ~50% of total β-diversity for sites at similar elevation and ~90% of total β-diversity for the most widely separated sites. Species turnover consisted mainly of turnover in the floral community **(Figure 3C)**, though the joint turnover of plants and bumble bees accounted for ~25% of total species turnover between the most widely separated sites. Interaction rewiring was most significant between sites at similar elevation, where it accounted for ~30% of total β-diversity, but its share of total β-diversity declined to ~10% in the most widely separated sites **(Figure 3B)**.

Logistic matrix regression confirmed the significance (p < 0.01) of the relationship between each β-diversity metric and elevation **(Table 1)**.

```{r fig3, echo = FALSE,  fig.cap = paste("Species-level (A) and interaction-level (B, C) plotted against elevational difference. Each point represents the β-diversity (or β-diversity partition) between a pair of sites, and lines represent the overall relationship between β-diversity and elevation difference using binomial regression smooths. Standard errors are not plotted because they would be misleading due to the non-independence inherent to distance matrix regression, but all regressions were significant (p < 0.01). Interaction β-diversity (B, C) is partitioned using Poisot et al.’s (2012) notation: WN = unpartitioned β-diversity, ST = β-diversity due to species turnover, OS =  β-diversity due to interaction rewiring, ST.h =  β-diversity due to species turnover in the higher trophic level (bumble bees), ST.l = β-diversity due to species turnover in the lower trophic level (plants), and ST.lh =  β-diversity due to joint species turnover in higher and lower trophic levels.")}
knitr::include_graphics("../output/beta_stack.png")
```

GDM analysis reproduced the finding that overall β-diversity was dominated by floral turnover, but it also revealed a strongly nonlinear response of floral and interaction β-diversity to elevation **(Figure 4)**. Turnover was high from ~600-1000 m, representing the transition from valley floor to lower slopes, and then leveled off from ~1000-1500 m. At around 1500 m, turnover accelerated sharply and remained steep for the remainder of the elevation gradient. The partial effects of geographic proximity and sampling intensity were negligible in all cases.

```{r fig4, echo = FALSE,  fig.cap = paste("GDM splines of species-level and interaction-level β-diversity in response to geographic, elevational, and temporal distance between site-dates. The maximum height of each spline represents a given variable’s partial effect on β-diversity (its effect when covariates are held constant), and the shape of each spline represents to the rate of species turnover as it varies along the corresponding gradient (i.e. geographic distance, elevation, or day of year), with steeper parts of the curve indicating regions of the gradient over which species/interaction turnover is more rapid. Within-site comparisons were omitted from GDM analysis to avoid pseudoreplication. Shaded bands depict the uncertainty of each partial effect curve (+/- one standard deviation) based on bootstrapping. Note the differences in scale between the y-axes of each variable.")}
knitr::include_graphics("../output/gdm_stack.png")
```

# DISCUSSION

The floristic importance of the tree line ecotone has been noted in previous work [@Pellissier2010-kg; @Descombes2017-id]. By studying the floral community through the lens of foraging bumble bees, we provide a functional extension of this floristic pattern, showing that the tree line ecotone both a boundary above which floral resource availability sharply declines and a critical distributional interface between high-elevation bumble bees and their habitat-generalist counterparts. Similar patterns with respect to the tree line ecotone have been described for bumble bees and their floral hosts in the Mediterranean system of Mount Olympus [@Minachilis2020-xd], suggesting that these findings are a general property of mountain ecosystems.

The fact that

As such, the tree line ecotone deserves special consideration in conservation management, and targeted studies of plant-pollinator relationships at the tree line ecotone would be warranted.

A salient question that could not be answered by our sampling approach is whether mountain bumble bees adaptively forage up- or downslope, as suggested by @Lundberg1980-qk. Bumble bees have large foraging ranges and have been shown both to track resources through space and time [@Devoto2014-ay] and to cross forest matrix to reach patches of foraging habitat [@Mola2020-yc]. Assuming an average slope of 20° and a foraging range of 1 km, a bumble bee could travel up- or downslope by more than 340 m, thus spanning an elevation belt nearly 700 m wide. Such 3-dimensional foraging would enable bumble bees both to exploit the elevational turnover of floral species and to track preferred species through their elevationally staggered phenology, the latter constituting a sort of physiological time travel [@Van_Straalen1983-df.

he rate of β-diversity --- whether that of bumble bees, flora, or interactions --- accelerates sharply near the tree line ecotone (~1500 m) and continues accelerating for the remainder of the elevation gradient (Figure 8). Thus, a linear elevation gradient generates a nonlinear biological response, punctuated by the tree line ecotone, above which further increase in elevation becomes an increasingly stringent filter of community assembly. Such patterns have been described previously for plants (Pellissier et al. 2010, Descombes et al. 2017); our study suggests they extend also to bumble bees, and that they are propagated from the level of community composition to that of species interactions between bumble bees and their floral hosts.

# CONCLUSIONS

The central finding of our study, evident in each of our analyses, is that a linear elevation gradient can generate nonlinear biological responses in terms of both abundance and β-diversity, and the latter at both at the level of community composition and at the level of species interactions. While elevational climate variation cannot be completely disentangled from other variables that covary with elevation [@Hodkinson2005-af], nonlinear responses of plants and pollinators along an elevational climate gradient suggest that we might expect similar nonlinearities in response to temporal climate change. Indeed, such nonlinear responses to climate have already been reported, including a recent study that documented nonlinear decline of insect abundance over a 37-year temporal climate gradient [@Lister2018-jp].

Independent of their utility as space-for-time climate proxies, mountains per se are unique ecological theaters in which extrinsic temporal climate change interacts with an intrinsic elevational climate gradation to alter the spatial and temporal distributions of species and interactions [@Telwala2013-sc; @CaraDonna2014-nq; @Miller-Struttmann2014-wz; @Rafferty2020-aj]. The nonlinearity of abundance observed in our study warns that temporal climate change on mountain slopes should be expected to involve the redistribution of discrete peaks of species abundance.

It should also be expected that climate change on mountain slopes will be punctuated by the tree line ecotone, which constitutes, at least with respect to bumble bees and wildflowers, a region of unique resource abundance and a transition to high and accelerating turnover of species and interactions [@@Descombes2017-id]. The observation that the upslope advancement of alpine tree lines often lags behind climate warming [@Dullinger2004-qu] suggests a danger that high elevation florivores could move upslope faster than their floral resources, or perhaps be forced to cope with increased competitive pressure from other florivore species formerly restricted to lower elevations [@Hoiss2012-sn].

though as a matter of applied conservation it is perhaps doubtful that local interventions could forestall for long the consequences of a process occurring at a global scale.

# ACKNOWLEDGEMENTS

We thank D.J. McNeil for helpful conversations during data analysis.

# REFERENCES

<!-- ```{r write_citations, cache=FALSE, include=FALSE} -->
<!-- write.bibtex(file = "knitcitations.bib") -->
<!-- ``` -->

<div id ="refs"></div>

### 2021-02-18 ms_1/ms_1.Rmd
ms

@@
 Climate-induced range shifts are expected to be especially pronounced in the mountain habitats that host the core of bumble bee biodiversity, together with their floral mutualists, and upward movement of bumble bee species has already been documented [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc].
-Populations of bumble bee species already restricted to the highest elevation zones can be expected to suffer geometric habitat shrinkage and increasing genetic isolation as they move toward mountain peaks; the upward advance of lowland species may induce novel competitive pressures and disrupt the pollination networks in which bumble bees participate [@Brosi2013-mu; @Ishii2013-sm]; and concurrent elevational [@Lenoir2008-yn] and/or phenological [@Pyke2016-ll] shifts in the floral community could generate mismatches between bumble bee species and their historic floral hosts.
+Populations of bumble bee species already restricted to the highest elevation zones can be expected to suffer geometric habitat shrinkage and increasing genetic isolation as they move toward mountain peaks; the upward advance of lowland species may introduce novel competitive pressures and disrupt the pollination networks in which bumble bees participate [@Brosi2013-mu; @Ishii2013-sm]; and concurrent elevational [@Lenoir2008-yn] and/or phenological [@Pyke2016-ll] shifts in the floral community could generate mismatches between bumble bee species and their historic floral hosts.
 
 The outcomes of range shifts in mountain bumble bee and floral communities have obvious significance from a conservation perspective, but they also provide a model system in which to study the general phenomenon of community assembly under climate change, a phenomenon of both historical interest and prospective urgency.
-The goal of anticipating, interpreting, and potentially mitigating the effects of ranges shifts, however, is predicated upon understanding the current elevational and temporal patterns of bumble bee communities and those of their floral hosts.
+Anticipating, interpreting, and potentially mitigating the effects of ranges shifts, however, necessitate an understanding the current elevational and temporal patterns of bumble bee communities and those of their floral hosts that will be acted upon by climate change.
 In particular, it is important to detect nonlinearities in species-, community-, and interaction-level responses to elevational climate gradients, such as the floristic threshold effect that has been documented at the tree line ecotone [@Descombes2017-id].
@@
 β-diversity) of bumble bees and flora in response to elevation.
-Finally, propagate the question to the level of species interactions, investigating the turnover of bee-flower interaction partners through elevation.
-In each of these analyses, we employ flexible modeling techniques capable of capturing nonlinear responses to elevation and time with the goal of identifying where, when, and for which species the impacts of climate-induced range shifts should be expected to be most acute.
+Finally, we move to the level of species interactions, investigating the turnover of bee-flower interaction partners through elevation.
+In each of these analyses, we employ flexible modeling techniques capable of capturing nonlinear responses with the goal of identifying where, when, and for which species the impacts of climate-induced range shifts should be expected to be most acute.
 
@@
 
+- Recap the point of the study
+
+The core finding of our study, evident in each of our analyses, is that a linear elevation gradient can generate nonlinear biological responses in terms of both abundance and β-diversity, and the latter both at the level of community composition and at the level of species interactions.
+
+The most salient pattern in our study is the unique importance of the elevation band falling at about 1500 m, which corresponds to the lower edge of the tree line ecotone in the Berchtesgadener Alps [@Kostler1970-ya; @Mayer1970-zn] and specifically that of the elevational transects used in our study.
 The floristic importance of the tree line ecotone has been noted in previous work [@Pellissier2010-kg; @Descombes2017-id].
-By studying the floral community through the lens of foraging bumble bees, we provide a functional extension of this floristic pattern, showing that the tree line ecotone both a boundary above which floral resource availability sharply declines and a critical distributional interface between high-elevation bumble bees and their habitat-generalist counterparts.
-Similar patterns with respect to the tree line ecotone have been described for bumble bees and their floral hosts in the Mediterranean system of Mount Olympus [@Minachilis2020-xd], suggesting that these findings are a general property of mountain ecosystems.
+By studying the floral community through the lens of foraging bumble bees, however, we provide a functional extension of this floristic pattern, demonstrating that the tree line ecotone comprises (1) a boundary above which floral resource availability sharply declines, (2) a distributional interface between high-elevation bumble bees and their low-elevation or habitat-generalist counterparts, and (3) a zone of rapid acceleration of species and interaction β-diversity.
+The significance of the tree line ecotone as a threshold of floral resource availability and as a zone of overlapping bumble bee distributions corroborates the recent findings of @Minachilis2020-xd in the Mediterranean system of Mount Olympus, indicating that these patterns are likely to generalize well across mountain ecosystems.
+Our finding that interaction β-diversity is also punctuated by the tree line ecotone follows naturally from these compositional patterns, and we expect that the β-diversity patterns found in our system will prove similarly generalizable.
+
+The observation that the upslope advancement of alpine tree lines often lags behind climate warming [@Dullinger2004-qu] suggests a danger that high elevation bumble bees could move upslope faster than their floral resources.
+Indeed, a somewhat unexpected finding of our study is that high-elevation bumble bee species (*B.
+mucidus*, *B.
+monticola*, *B.
+mendax*, and *B.
+pyrenaeus*) were associated with floral hosts that peaked in abundance at much lower elevations, largely outside the range of their bumble bee visitors.
+As the upper boundary of high floral resource abundance, the lower edge of the tree line ecotone likely functions as critical foraging habitat for high-elevation bumble bees.
+Future warming could force these species farther into resource-poor alpine grasslands while at the same time increasing resource competition at the tree line as dominant low-elevation species, like *B.
+pascuorum*, extend their ranges upslope.
+
+Our analyses of species and interaction β-diversity reveal a striking contrast between a relative stable bumble bee community and an extremely dynamic floral community.
+The dominance of total interaction β-diversity by floral species turnover is consistent with the findings of @Simanonok2014-ou in the Rocky Mountains of North America, and it is perhaps a predictable relationship given the exceptional cold-hardiness of bumble bees [@Heinrich1994-cv] and the sensitivity of plants to temperature gradients [].
+Nevertheless, it highlights the behavioral challenge that bumble bees have presumably faced for the entirety of their evolutionary history [] and that may explain the remarkable flexibility and intelligence of their flower handling []: how to forage efficiently in a diverse that turns over rapidly in both space and time.
+A fascinating question that could not be answered by our sampling approach is whether mountain bumble bees adaptively forage up- or downslope, as suggested by @Lundberg1980-qk.
+Bumble bees have large foraging ranges and have been shown both to track resources through space and time [@Devoto2014-ay] and to cross forest matrix to reach patches of foraging habitat [@Mola2020-yc].
+Assuming an average slope of 20° and a foraging range of 1 km, a bumble bee could travel up- or downslope by more than 340 m, thus spanning an elevation belt nearly 700 m wide.
+Such 3-dimensional foraging would enable bumble bees both to exploit the elevational turnover of floral species and to track preferred species through their elevationally staggered phenology, the latter constituting a sort of physiological time travel [@Van_Straalen1983-df.
+
+# CONCLUSIONS
+
+- What do these findings tell use about climate change and range shifts in mountain ecosystems more generally?
+- Do our results suggest any management strategies?
+
+As in the past, so now are mountains the primarily theaters of ecological change for bumble bees.
+
+Mountains unique ecological theaters in which extrinsic temporal climate change interacts with an intrinsic elevational climate gradation to alter the spatial and temporal distributions of species and interactions [@Telwala2013-sc; @CaraDonna2014-nq; @Miller-Struttmann2014-wz; @Rafferty2020-aj].
+
+It should also be expected that climate change on mountain slopes will be punctuated by the tree line ecotone, which constitutes, at least with respect to bumble bees and wildflowers, a region of unique resource abundance and a transition to high and accelerating turnover of species and interactions [@@Descombes2017-id].
+
+though as a matter of applied conservation it is perhaps doubtful that local interventions could forestall for long the consequences of a process occurring at a global scale.
+
+Retreating to higher elevation in response to climatic warming is not a new phenomenon in the evolutionary history of bumble bees
 
@@
 As such, the tree line ecotone deserves special consideration in conservation management, and targeted studies of plant-pollinator relationships at the tree line ecotone would be warranted.
-
-A salient question that could not be answered by our sampling approach is whether mountain bumble bees adaptively forage up- or downslope, as suggested by @Lundberg1980-qk.
-Bumble bees have large foraging ranges and have been shown both to track resources through space and time [@Devoto2014-ay] and to cross forest matrix to reach patches of foraging habitat [@Mola2020-yc].
-Assuming an average slope of 20° and a foraging range of 1 km, a bumble bee could travel up- or downslope by more than 340 m, thus spanning an elevation belt nearly 700 m wide.
-Such 3-dimensional foraging would enable bumble bees both to exploit the elevational turnover of floral species and to track preferred species through their elevationally staggered phenology, the latter constituting a sort of physiological time travel [@Van_Straalen1983-df.
-
-he rate of β-diversity --- whether that of bumble bees, flora, or interactions --- accelerates sharply near the tree line ecotone (~1500 m) and continues accelerating for the remainder of the elevation gradient (Figure 8).
-Thus, a linear elevation gradient generates a nonlinear biological response, punctuated by the tree line ecotone, above which further increase in elevation becomes an increasingly stringent filter of community assembly.
-Such patterns have been described previously for plants (Pellissier et al.
-2010, Descombes et al.
-2017); our study suggests they extend also to bumble bees, and that they are propagated from the level of community composition to that of species interactions between bumble bees and their floral hosts.
-
-# CONCLUSIONS
-
-The central finding of our study, evident in each of our analyses, is that a linear elevation gradient can generate nonlinear biological responses in terms of both abundance and β-diversity, and the latter at both at the level of community composition and at the level of species interactions.
-While elevational climate variation cannot be completely disentangled from other variables that covary with elevation [@Hodkinson2005-af], nonlinear responses of plants and pollinators along an elevational climate gradient suggest that we might expect similar nonlinearities in response to temporal climate change.
-Indeed, such nonlinear responses to climate have already been reported, including a recent study that documented nonlinear decline of insect abundance over a 37-year temporal climate gradient [@Lister2018-jp].
-
-Independent of their utility as space-for-time climate proxies, mountains per se are unique ecological theaters in which extrinsic temporal climate change interacts with an intrinsic elevational climate gradation to alter the spatial and temporal distributions of species and interactions [@Telwala2013-sc; @CaraDonna2014-nq; @Miller-Struttmann2014-wz; @Rafferty2020-aj].
-The nonlinearity of abundance observed in our study warns that temporal climate change on mountain slopes should be expected to involve the redistribution of discrete peaks of species abundance.
-
-It should also be expected that climate change on mountain slopes will be punctuated by the tree line ecotone, which constitutes, at least with respect to bumble bees and wildflowers, a region of unique resource abundance and a transition to high and accelerating turnover of species and interactions [@@Descombes2017-id].
-The observation that the upslope advancement of alpine tree lines often lags behind climate warming [@Dullinger2004-qu] suggests a danger that high elevation florivores could move upslope faster than their floral resources, or perhaps be forced to cope with increased competitive pressure from other florivore species formerly restricted to lower elevations [@Hoiss2012-sn].
-
-though as a matter of applied conservation it is perhaps doubtful that local interventions could forestall for long the consequences of a process occurring at a global scale.
 

### 2021-02-27 code/abund_div_GAM.Rmd
back up

@@
 ```
+
+# Just a quick peak at what would happen if I added a by = year smooth 
+```{r}
+weighted_fl_abund.qp2 <- bam(total.weighted.flower.cover ~ 
+                              s(bb.sp, bs = "re") +
+                              s(site, bs = "re") +
+                              s(year, bs = "re") +
+                              te(yday, elev.mean,
+                                 by = year,
+                                 bs = c("gp", "tp"),
+                                 k = c(5, 5)) +
+                              te(yday, elev.mean,
+                                 by = bb.sp,
+                                 bs = c("gp", "tp"),
+                                 k = c(5, 5)),
+                            data = filter(weighted_fl_abund, bb.sp != "humi"),
+                            family = "quasipoisson",
+                            discrete = TRUE,
+                            select = TRUE,
+                            method = "fREML") %>% getViz(nsim = 500)
+
+check.gamViz(weighted_fl_abund.qp2)
+sum_fl.qp2 <- summary(weighted_fl_abund.qp2)
+print(plot(weighted_fl_abund.qp2, allTerms = TRUE), pages = 4)
+check2D(weighted_fl_abund.qp2, x1 = "yday", x2 = "elev.mean")
+
+bb_abund.qp2 <- bam(bb.abund ~ 
+                     s(bb.sp.ord, bs = "re") + # center species
+                     s(year, bs = "re") + # random intercept for year
+                     s(site, bs = "re") + # random intercept to avoid pseudoreplicating elevation
+                     te(yday, elev.mean, # tensor product smooth of elevation and time
+                        by = year, # by bb.sp
+                        bs= c("gp", "tp"), # gaussian procees for time, thinplate spline for elevation
+                        k=c(5, 5), 
+                        m=2) +
+                     te(yday, elev.mean, # tensor product smooth of elevation and time
+                        by = bb.sp.ord, # by bb.sp
+                        bs= c("gp", "tp"), # gaussian procees for time, thinplate spline for elevation
+                        k=c(5, 5), 
+                        m=2),
+                   data = filter(bb_abund_sp, bb.sp != "humi"), # humi is just too rare
+                   discrete = TRUE,
+                   family = "quasipoisson",
+                   select = TRUE,
+                   method = "fREML") %>% getViz(nsim = 500)
+
+check.gamViz(bb_abund.qp2)
+sum.qp2 <- summary(bb_abund.qp2)
+print(plot(bb_abund.qp2, allTerms = TRUE), pages = 4)
+check2D(bb_abund.qp2, x1 = "yday", x2 = "elev.mean")
+```

### 2021-02-27 code/preference_lability.Rmd
back up

@@
 library(mgcViz)
+library(DHARMa)
 library(modelr)
@@
 
-ktype_families <- visit_cover %>%
-  select(plant.sp, plant.family, k.type.ss) %>%
-  distinct()
-
-ggplot(ktype_families, aes(plant.family, fill = k.type.ss)) +
-  geom_bar(position = "fill") +
-  theme(axis.text.x = element_text(angle = 90))
-
 ## Get total visits by BB sp and site-date; this will be a measure of BB abundance that we can use after we've calculated d'.
@@
 
-# 1. Overall bias
+# 1. Abundance of floral morphotypes through elevation and time
+
+## 1.1 Get absolute and percent flower cover by morphotype
+```{r}
+ktype_abund <- flower_cover %>%
+  left_join(fl_traits) %>%
+  group_by(site, date, k.type.ss) %>%
+  summarize(flower.cover = sum(flower.cover)) %>%
+  left_join(site_data) %>%
+  mutate(ktype = factor(k.type.ss),
+         year = factor(year(date)),
+         yday = yday(date),
+         site = factor(site)) %>%
+  group_by(site, date) %>%
+  mutate(prop.flower.cover = flower.cover/sum(flower.cover))
+```
+
+## 1.2 Fit elevation*time tensor-product GAMs 
+```{r}
+# Prop abundance; this works well with a quasibinomial distribution
+gam_ktype_propabund <- bam(prop.flower.cover ~
+                             s(ktype, bs = "re") +
+                             te(yday, elev.mean, by = ktype, bs = c("gp", "tp"), k = 8),
+                           family = "quasibinomial",
+                           method = "fREML",
+                           discrete = TRUE,
+                           select  = TRUE,
+                           data = filter(ktype_abund, !ktype %in% c("9", "10"))) %>% getViz()
+
+check.gamViz(gam_ktype_propabund)
+summary(gam_ktype_propabund)
+
+# Absolute abundance; I find that I really have to log-transform absolute flower cover to fit a good model
+gam_ktype_abund <- bam(log(flower.cover) ~
+                         s(ktype, bs = "re") +
+                         te(yday, elev.mean, by = ktype, bs = c("gp", "tp"), k = 8),
+                       family = "scat", # to account for heavy tails
+                       method = "fREML",
+                       discrete = TRUE,
+                       select  = TRUE,
+                       data = filter(ktype_abund, !ktype %in% c("9", "10"))) %>% getViz()
+
+check.gamViz(gam_ktype_abund)
+summary(gam_ktype_abund)
+```
+
+## 1.3 Plot morphotype abundances
+```{r}
+k1_abund <- plot(sm(gam_ktype_abund, 2))
+k2_abund <- plot(sm(gam_ktype_abund, 3))
+k3_abund <- plot(sm(gam_ktype_abund, 4))
+k4_abund <- plot(sm(gam_ktype_abund, 5))
+k5_abund <- plot(sm(gam_ktype_abund, 6))
+k6_abund <- plot(sm(gam_ktype_abund, 7))
+k7_abund <- plot(sm(gam_ktype_abund, 8))
+
+k1_propabund <- plot(sm(gam_ktype_propabund, 2))
+k2_propabund <- plot(sm(gam_ktype_propabund, 3))
+k3_propabund <- plot(sm(gam_ktype_propabund, 4))
+k4_propabund <- plot(sm(gam_ktype_propabund, 5))
+k5_propabund <- plot(sm(gam_ktype_propabund, 6))
+k6_propabund <- plot(sm(gam_ktype_propabund, 7))
+k7_propabund <- plot(sm(gam_ktype_propabund, 8))
+
+ktype_abund_grid <- gridPrint(
+
+                       k1_propabund + l_fitRaster() + l_fitContour() + 
+                         labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       k2_propabund + l_fitRaster() + l_fitContour() + 
+                         labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       k3_propabund + l_fitRaster() + l_fitContour() + 
+                         labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       k4_propabund + l_fitRaster() + l_fitContour() + 
+                         labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       k5_propabund + l_fitRaster() + l_fitContour() + 
+                         labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
+                       k6_propabund + l_fitRaster() + l_fitContour() + 
+                         labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       k7_propabund + l_fitRaster() + l_fitContour() + 
+                         labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+
+                       k1_abund + l_fitRaster() + l_fitContour() + 
+                         labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       k2_abund + l_fitRaster() + l_fitContour() + 
+                         labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       k3_abund + l_fitRaster() + l_fitContour() + 
+                         labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       k4_abund + l_fitRaster() + l_fitContour() + 
+                         labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       k5_abund + l_fitRaster() + l_fitContour() + 
+                         labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
+                       k6_abund + l_fitRaster() + l_fitContour() + 
+                         labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       k7_abund + l_fitRaster() + l_fitContour() + 
+                         labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+
+                       ncol = 7)
+
+ggsave(ktype_abund_grid, file = "../output/ktype_abund_gam.pdf", height = 3, width = 12)
+```
+
+# 1.
+Overall bias ** Need to refit these models without ktype 0 since it seems that these plants were ignored during floral surveying (always assigned cover = 0), and therefore I have no abundance data with which to compare visitation **
+
 ## 1.1 Calculate visitation probability per plant morphotype.
@@
   select(site, elev.mean, date, yday, bb.sp, ktype, visits, 
-         prop.visits, perc.flower.cover, bb.sp.abund, tongue,
+         prop.visits, flower.cover, perc.flower.cover, bb.sp.abund, tongue,
          elev.mean0, yday0, prop.flower.cover0, bb.sp.abund0) %>%
@@
          elev.mean0, yday0, prop.flower.cover0, bb.abund0) %>%
-  mutate(visits.bin = if_else(visits > 0, 1, 0))
+  mutate(visits.bin = if_else(visits > 0, 1, 0),
+         year = factor(year(date)))
 
@@
                     (1 + perc.flower.cover | ktype:tongue) + # a random intercept and slope for ktype:tongue
-                    (1 | site), # a random intercept and slope for ktype:site
+                    (1 | ktype:site), # a varying intercept and slope for ktype:site
                   family = "bernoulli", 
-                  #file = "../output/brm_pvis_ktype_00",  # for running locally
-                  file = "brm_pvis_ktype_00", # for running on the HPC
+                  file = "../output/brm_pvis_ktype_00",  # for running locally
+                  #file = "brm_pvis_ktype_00", # for running on the HPC
                   chains = 4,
@@
                   iter = 2000,
-                  data = filter(prop_visits_ktype, !ktype %in% c("9", "10")))
+                  data = filter(prop_visits_ktype, !ktype %in% c("0", "9", "10")))
 
@@
                     bb.abund*perc.flower.cover + # fixed term: because more abundant flowers will tend to be more visited,and more abundance bees visit more
-                    (1 + perc.flower.cover | ktype) + # a random intercept and slope for ktype:site
-                    (1 | site), 
+                    (1 + perc.flower.cover | ktype) + # a random intercept and slope for ktype
+                    (1 | ktype:site),  # a varying intercept term for ktype:site
                   family = "bernoulli", 
@@
                   iter = 2000,
-                  data = filter(prop_visits_ktype_pooled, !ktype %in% c("9", "10")))
+                  data = filter(prop_visits_ktype_pooled, !ktype %in% c("0", "9", "10")))
 
@@
                     (1 + perc.flower.cover | ktype:bb.sp) + # a random intercept and random slope for ktype:species; I think I need to model this before I can make conclusions about ktype:pbl.class
-                    (1 | site), # a random intercept and slope for ktype:site
+                    (1 | ktype:site), # a varying intercept and slope for ktype:site
                   family = "bernoulli", 
-                  #file = "../output/brm_pvis_ktype_02",  # for running locally
-                  file = "brm_pvis_ktype_02", # for running on the HPC
+                  file = "../output/brm_pvis_ktype_02",  # for running locally
+                  #file = "brm_pvis_ktype_02", # for running on the HPC
                   chains = 4,
                   cores = 12,
-                  iter = 2000,
-                  data = filter(prop_visits_ktype, !ktype %in% c("9", "10")))
+                  iter = 5000, # Bulk ESS was a bit low wit 2000 iter
+                  data = filter(prop_visits_ktype, !ktype %in% c("0", "9", "10")))
 ```
@@
 summary(brm_pvis02)
+summary(brm_pvis11)
+summary(brm_pvis21)
 
@@
 pp_check(brm_pvis02)
+pp_check(brm_pvis11)
+pp_check(brm_pvis21)
 
@@
 bayes_R2(brm_pvis02)
+bayes_R2(brm_pvis11)
+bayes_R2(brm_pvis21)
 
@@
 mcmc_plot(brm_pvis01, pars = c("^r_ktype:bb.sp.*Intercept]$"))
+mcmc_plot(brm_pvis01, pars = c("^r_ktype:site.*Intercept]$"))
 mcmc_plot(brm_pvis01, pars = c("^r_ktype:bb.sp.*perc.flower.cover]$"))
@@
 mcmc_plot(brm_pvis02, pars = c("^r_ktype:bb.sp.*Intercept]$"))
+mcmc_plot(brm_pvis01, pars = c("^r_ktype:site.*Intercept]$"))
 mcmc_plot(brm_pvis02, pars = c("^r_ktype:bb.sp.*perc.flower.cover]$"))
@@
 mcmc_plot(brm_pvis02, pars = c("^r_ktype:site.*perc.flower.cover]$"))
+
+get_variables(brm_pvis11)
+conditional_effects(brm_pvis11)
+mcmc_plot(brm_pvis11, pars = c("^r_ktype:year"))
+
+get_variables(brm_pvis21)
+conditional_effects(brm_pvis21)
+mcmc_plot(brm_pvis21, pars = c("^r_ktype"))
 ```
@@
   mutate(ktype = case_when( # rename
-    ktype == "0" ~ "NA (0)",
     ktype == "1" ~ "Disc (1)",
@@
   ) %>%
-  mutate(ktype = factor(ktype, levels = c("Disc (1)", "Funnel (2)", "Bell (3)", # relevel
-                                          "Stalk-disc (4)","Lip (5)", "Flag (6)", 
-                                          "Head (7)", "NA (0)")))
+  mutate(ktype = factor(ktype, levels = c("Disc (1)", "Funnel (2)", # relevel
+                                          "Bell (3)", "Stalk-disc (4)", "Lip (5)", 
+                                          "Flag (6)", "Head (7)")))
 
@@
 
-ggsave(file = "../output/brm_pvis00_plot1.png")
-ggsave(file = "../output/brm_pvis00_plot1.pdf")
-
+ggsave(file = "../output/brm_pvis00_plot1.pdf", width = 6, height = 5)
 ```
@@
   ) %>%
-  mutate(ktype = factor(ktype, levels = c("Disc (1)", "Funnel (2)", "Bell (3)", # relevel
-                                          "Stalk-disc (4)","Lip (5)", "Flag (6)", 
-                                          "Head (7)", "NA (0)"))) %>%
-  left_join(bb_traits)
+   mutate(ktype = factor(ktype, levels = c("Disc (1)", "Funnel (2)", # relevel
+                                          "Bell (3)", "Stalk-disc (4)", "Lip (5)", 
+                                          "Flag (6)", "Head (7)"))) %>%
+  left_join(bb_traits) %>%
+  mutate(bb.sp = factor(bb.sp, levels = c("telu", "soro", "pyre", "psit", "prat", "lapi", "jone",
+                                           "wurf", "pasc", "muci", "mont", "hypn",
+                                           "mend", "hort", "gers")))
 
@@
 
-ggsave(file = "../output/brm_pvis02_plot1.png", width = 6.5, height = 6.5)
+ggsave(file = "../output/brm_pvis02_plot1.pdf", width = 8, height = 8)
 ```
@@
   mutate(ktype = case_when(
-    ktype == "0" ~ "NA (0)",
     ktype == "1" ~ "Disc (1)",
@@
                         levels = c("Disc (1)", "Funnel (2)", "Bell (3)", "Stalk-disc (4)",
-                                   "Lip (5)", "Flag (6)", "Head (7)", "NA (0)")))
+                                   "Lip (5)", "Flag (6)", "Head (7)")))
 
@@
   mutate(ktype = case_when( # rename
-    condition == "0" ~ "NA (0)",
     condition == "1" ~ "Disc (1)",
@@
   ) %>%
-  mutate(ktype = factor(ktype, levels = c("Disc (1)", "Funnel (2)", "Bell (3)", # relevel
-                                          "Stalk-disc (4)","Lip (5)", "Flag (6)", 
-                                          "Head (7)", "NA (0)")))
+  mutate(ktype = factor(ktype, levels = c("Head (7)", "Flag (6)", # relevel
+                                          "Lip (5)", "Stalk-disc (4)", "Bell (3)", 
+                                          "Funnel (2)", "Disc (1)")))
 
@@
   geom_pointinterval() +
-  theme_light() +
+  theme_light(14) +
   labs(x = "Visitation probability", y= "Morphotype") +
@@
 
-ggsave(file = "../output/brm_pvis01_plot1.png", width = 5, height = 7)
-ggsave(file = "../output/brm_pvis01_plot1.pdf", width = 5, height = 7)
-
-ggplot(brm_pvis01_plotframe1, aes(condition_mean_trans, ktype, 
-                                 xmin = .lower_trans, xmax = .upper_trans)) +
-  geom_pointinterval() +
-  theme_light(18) +
-  labs(x = "Visitation probability", y= "Morphotype") +
-  guides(fill = FALSE)
-
-ggsave(file = "../output/brm_pvis01_plot1_largeprint.png", width = 7, height = 5)
+ggsave(file = "../output/brm_pvis01_plot1.pdf", width = 5, height = 5)
 
@@
 
-# 3. The lability of bias: How does the foraging bias of each species with respect to each morphotype vary through elevation and time?
-```{r}
-# Subset data to include only abundant BB spp and exclude rare morphotypes (9, 10)
+# 4. The lability of bias
+## 4.1 How does the foraging bias of each species with respect to each morphotype vary through elevation and time?
+```{r}
+# Subset data to include only abundant BB spp 
 gam_data <- prop_visits_ktype %>%
-  filter(ktype %in% c("3", "5", "6", "7") & # focus on major ktypes
+  filter(!ktype %in% c("0", "9", "10") & # focus on major ktypes
+           bb.sp %in% c("telu", "prat", "pasc", "hort", "soro", "wurf")) %>% # focus on abundance bb.sp
+  unite(pair, c(bb.sp, ktype), sep = "_", remove = FALSE) %>% # paste bb.sp and ktype into pair
+  mutate(pair = factor(pair),
+         bin.visits = if_else(visits > 0, 1, 0), # binarize visitation
+         year = factor(year(date)),
+         site = factor(site))
+
+# Independent smooth model
+gam_pvis_elev01 <- bam(bin.visits ~
+                         s(pair, bs = "re") + # center by pair
+                         bb.sp.abund:perc.flower.cover + # condition on a global bb/fl abundance term
+                         perc.flower.cover:pair + # condition on a pair-specific fl abundance
+                         te(yday, elev.mean, by = pair, bs = c("gp", "tp"), k = 5), # conditional effects of yday/elev by pair
+                       family = "binomial", # binomial response
+                       method = "fREML",
+                       discrete = TRUE,
+                       select =  TRUE,
+                       data = filter(gam_data,
+                                     bb.sp %in% c("soro", "wurf", "hort",
+                                                  "telu", "pasc", "prat"))) %>%
+  getViz()
+
+gam_pvis_elev01.sum <- summary(gam_pvis_elev01)
+check.gamViz(gam_pvis_elev01)
+concurvity(gam_pvis_elev01)
+```
+
+### 4.1.1 Plot 
+```{r}
+hort1 <- plot(sm(gam_pvis_elev01, 2))
+hort2 <- plot(sm(gam_pvis_elev01, 3))
+hort3 <- plot(sm(gam_pvis_elev01, 4))
+hort4 <- plot(sm(gam_pvis_elev01, 5))
+hort5 <- plot(sm(gam_pvis_elev01, 6))
+hort6 <- plot(sm(gam_pvis_elev01, 7))
+hort7 <- plot(sm(gam_pvis_elev01, 8))
+
+pasc1 <- plot(sm(gam_pvis_elev01, 9))
+pasc2 <- plot(sm(gam_pvis_elev01, 10))
+pasc3 <- plot(sm(gam_pvis_elev01, 11))
+pasc4 <- plot(sm(gam_pvis_elev01, 12))
+pasc5 <- plot(sm(gam_pvis_elev01, 13))
+pasc6 <- plot(sm(gam_pvis_elev01, 14))
+pasc7 <- plot(sm(gam_pvis_elev01, 15))
+
+prat1 <- plot(sm(gam_pvis_elev01, 16))
+prat2 <- plot(sm(gam_pvis_elev01, 17))
+prat3 <- plot(sm(gam_pvis_elev01, 18))
+prat4 <- plot(sm(gam_pvis_elev01, 19))
+prat5 <- plot(sm(gam_pvis_elev01, 20))
+prat6 <- plot(sm(gam_pvis_elev01, 21))
+prat7 <- plot(sm(gam_pvis_elev01, 22))
+
+soro1 <- plot(sm(gam_pvis_elev01, 23))
+soro2 <- plot(sm(gam_pvis_elev01, 24))
+soro3 <- plot(sm(gam_pvis_elev01, 25))
+soro4 <- plot(sm(gam_pvis_elev01, 26))
+soro5 <- plot(sm(gam_pvis_elev01, 27))
+soro6 <- plot(sm(gam_pvis_elev01, 28))
+soro7 <- plot(sm(gam_pvis_elev01, 29))
+
+telu1 <- plot(sm(gam_pvis_elev01, 30))
+telu2 <- plot(sm(gam_pvis_elev01, 31))
+telu3 <- plot(sm(gam_pvis_elev01, 32))
+telu4 <- plot(sm(gam_pvis_elev01, 33))
+telu5 <- plot(sm(gam_pvis_elev01, 34))
+telu6 <- plot(sm(gam_pvis_elev01, 35))
+telu7 <- plot(sm(gam_pvis_elev01, 36))
+
+wurf1 <- plot(sm(gam_pvis_elev01, 37))
+wurf2 <- plot(sm(gam_pvis_elev01, 38))
+wurf3 <- plot(sm(gam_pvis_elev01, 39))
+wurf4 <- plot(sm(gam_pvis_elev01, 40))
+wurf5 <- plot(sm(gam_pvis_elev01, 41))
+wurf6 <- plot(sm(gam_pvis_elev01, 42))
+wurf7 <- plot(sm(gam_pvis_elev01, 43))
+
+pterm1 <- plot(pterm(gam_pvis_elev01, 1))
+pterm2 <- plot(pterm(gam_pvis_elev01, 2))
+
+grid_plot <- gridPrint(hort1 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       hort2 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       hort3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       hort4 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       hort5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
+                       hort6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       hort7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       pasc1 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       pasc2 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       pasc3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       pasc4 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       pasc5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
+                       pasc6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       pasc7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       prat1 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       prat2 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       prat3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       prat4 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       prat5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
+                       prat6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       prat7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       soro1 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       soro2 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       soro3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       soro4 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       soro5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
+                       soro6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       soro7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       telu1 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       telu2 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       telu3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       telu4 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       telu5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
+                       telu6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       telu7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       wurf1 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       wurf2 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       wurf3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       wurf4 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       wurf5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
+                       wurf6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       wurf7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                       ncol = 7)
+
+ggsave(grid_plot, file = "../output/ktype_lability_gam.pdf", height = 9, width = 12)
+```
+
+### Another way to ask this question is to see how the visitation probability changes if I drop the floral abundance term
+```{r}
+gam_pvis_elev01_unconditioned <- bam(bin.visits ~
+                         s(pair, bs = "re") + # center by pair
+                         s(year, bs = "re") +
+                         bb.sp.abund +
+                         te(yday, elev.mean, by = pair, bs = c("gp", "tp"), k = 5), # conditional effects of yday/elev by pair
+                       family = "binomial", # binomial response
+                       method = "fREML",
+                       discrete = TRUE,
+                       select =  TRUE,
+                       data = filter(gam_data,
+                                     bb.sp %in% c("soro", "wurf", "hort",
+                                                  "telu", "pasc", "prat"))) %>%
+  getViz()
+
+check.gamViz(gam_pvis_elev01_unconditioned)
+
+print(plot(gam_pvis_elev01_unconditioned, allTerms = TRUE), pages = 1)
+
+u_hort3 <- plot(sm(gam_pvis_elev01_unconditioned, 3))
+u_hort5 <- plot(sm(gam_pvis_elev01_unconditioned, 4))
+u_hort6 <- plot(sm(gam_pvis_elev01_unconditioned, 5))
+u_hort7 <- plot(sm(gam_pvis_elev01_unconditioned, 6))
+
+u_pasc3 <- plot(sm(gam_pvis_elev01_unconditioned, 7))
+u_pasc5 <- plot(sm(gam_pvis_elev01_unconditioned, 8))
+u_pasc6 <- plot(sm(gam_pvis_elev01_unconditioned, 9))
+u_pasc7 <- plot(sm(gam_pvis_elev01_unconditioned, 10))
+
+u_prat3 <- plot(sm(gam_pvis_elev01_unconditioned, 11))
+u_prat5 <- plot(sm(gam_pvis_elev01_unconditioned, 12))
+u_prat6 <- plot(sm(gam_pvis_elev01_unconditioned, 13))
+u_prat7 <- plot(sm(gam_pvis_elev01_unconditioned, 14))
+
+u_soro3 <- plot(sm(gam_pvis_elev01_unconditioned, 15))
+u_soro5 <- plot(sm(gam_pvis_elev01_unconditioned, 16))
+u_soro6 <- plot(sm(gam_pvis_elev01_unconditioned, 17))
+u_soro7 <- plot(sm(gam_pvis_elev01_unconditioned, 18))
+
+u_telu3 <- plot(sm(gam_pvis_elev01_unconditioned, 19))
+u_telu5 <- plot(sm(gam_pvis_elev01_unconditioned, 20))
+u_telu6 <- plot(sm(gam_pvis_elev01_unconditioned, 21))
+u_telu7 <- plot(sm(gam_pvis_elev01_unconditioned, 22))
+
+u_wurf3 <- plot(sm(gam_pvis_elev01_unconditioned, 23))
+u_wurf5 <- plot(sm(gam_pvis_elev01_unconditioned, 24))
+u_wurf6 <- plot(sm(gam_pvis_elev01_unconditioned, 25))
+u_wurf7 <- plot(sm(gam_pvis_elev01_unconditioned, 26))
+
+u_hort3
+u_pasc3
+u_prat3
+u_soro3
+u_telu3
+u_wurf3
+
+u_hort5
+u_pasc5
+u_prat5
+u_soro5
+u_telu5
+u_wurf5
+
+u_hort6
+u_pasc6
+u_prat6
+u_soro6
+u_telu6
+u_wurf6
+
+u_hort7
+u_pasc7
+u_prat7
+u_soro7
+u_telu7
+u_wurf7
+
+grid_plot <- gridPrint(u_hort3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                        u_hort5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
+                        u_hort6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                        u_hort7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                        u_pasc3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                        u_pasc5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
+                        u_pasc6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                        u_pasc7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                        u_prat3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                        u_prat5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
+                        u_prat6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                        u_prat7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                        u_soro3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                        u_soro5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
+                        u_soro6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                        u_soro7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                        u_telu3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                        u_telu5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
+                        u_telu6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                        u_telu7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                        u_wurf3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                        u_wurf5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
+                        u_wurf6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                        u_wurf7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
+                        ncol = 4)
+```
+
+# Hmm...so it looks like floral abundance just isn't driving visitation patterns very strongly.
+Removing it from the model hardly changes anything, but modeling just floral abundance gives very different patterns than modeling visitation.
+I suppose we can look at this directly
+```{r}
+ggplot(gam_data, aes(perc.flower))
+```
+
+# But it looks like BB abundance is a very strong driver. What if we remove that from the model?
+```{r}
+
+```
+
+# Let's try some alternative formulations 
+```{r}
+# We could try using all the ktypes but still sticking to just the major BBs
+gam_data2 <- prop_visits_ktype %>%
+  filter(ktype %in% c("1", "2", "3", "4", "5", "6", "7") & # focus on major ktypes
            bb.sp %in% c("telu", "prat", "pasc", "hort", "soro", "wurf")) %>% # focus on abundance bb.sp
@@
 
-# Independent smooth model
-gam_pvis_elev01 <- bam(bin.visits ~
+gam_pvis_elev_alt00 <- bam(bin.visits ~
                              s(pair, bs = "re") + # center by pair
@@
                            select =  TRUE,
+                           data = filter(gam_data2,
+                                         bb.sp %in% c("soro", "wurf", "hort",
+                                                      "telu", "pasc", "prat"))) %>% getViz()
+
+check.gamViz(gam_pvis_elev_alt00)
+sum_alt00 <- summary(gam_pvis_elev_alt00)
+print(plot.gamViz(gam_pvis_elev_alt00, allTerms = TRUE), pages = 1)
+
+# We could also try a count model
+gam_pvis_elev_alt01 <- bam(visits ~
+                             s(pair, bs = "re") + # center by pair
+                             bb.sp.abund*perc.flower.cover + # condition on a global bb/fl abundance term
+                             perc.flower.cover*pair + # condition on a pair-specific fl abundance
+                             #s(perc.flower.cover, by = pair, k = 3) +  term
+                             te(yday, elev.mean, by = pair, bs = c("gp", "tp"), k = 5), # conditional effects of yday/elev by pair
+                           family = "quasipoisson", # Poisson response
+                           method = "fREML",
+                           discrete = TRUE,
+                           select =  TRUE,
                            data = filter(gam_data,
+                                         bb.sp %in% c("soro", "wurf", "hort",
+                                                      "telu", "pasc", "prat"))) %>% getViz()
+
+check.gamViz(gam_pvis_elev_alt01)
+sum_alt01 <- summary(gam_pvis_elev_alt01)
+print(plot.gamViz(gam_pvis_elev_alt01, allTerms = TRUE), pages = 1)
+plot.gamViz(pterm(gam_pvis_elev_alt01, 1))
+
+```
+
+# Can we say that certain morphotypes are preferred when overall resources are scarce?
+```{r}
+### Floral survey data
+survey <- read_csv("../data/processed_data/floral_survey.csv") %>%
+  semi_join(visitation, by = "plant.sp") %>% # Consider only visited plant species (anytime anywhere, not specifically for a given site*date)
+  dplyr::select(site, date, plant.sp, flower.cover) %>%
+  group_by(plant.sp) %>%
+  mutate(present = case_when(
+    max(flower.cover) > 0 ~ TRUE,
+    max(flower.cover) == 0 ~ FALSE
+  )) %>%
+  filter(present == TRUE) # Ignore plants that never had any flower cover
+
+fl_abund_sp <- survey %>%
+  group_by(site, date, plant.sp) %>% # group by plant.sp*transect
+  summarize(fl.abund = sum(flower.cover)) %>% # get plant.sp*transect total flower cover; this step really shouldn't be required, but in a few cases, there is more than one abundance entry for the same species on a given transect; not sure how that happened, but it's in the original data that Fabrice sent me; alternatively, I could drop the dups or average them.
+We'll see what Kalli and Fabrice have to say.
+  group_by(site, date) %>%
+  mutate(prop.fl.abund = fl.abund/sum(fl.abund)) %>% # get the proportional flower cover for each species.
+  ungroup() %>%
+  #full_join(dplyr::select(net, site, date, plant.sp), by = c("site", "date", "plant.sp")) %>% # add the handful of species occurring in the visitation network but missing from the survey data
+  complete(site, date, plant.sp, fill = list(fl.abund = 0, prop.fl.abund = 0)) %>%
+  semi_join(survey, by = c("site", "date")) %>%  # only include site-dates that actually occurred
+  unique() %>%
+  mutate(year = factor(year(date))) %>%
+  left_join(fl_traits)
+
+weights <- fl_abund_sp %>%
+  select(plant.sp, plant.genus) %>%
+  unique() %>%
+  left_join(select(visitation, bb.sp, plant.sp)) %>%
+  select(plant.genus, bb.sp) %>%
+  distinct() %>%
+  mutate(visited = rep(1, n())) %>%
+  pivot_wider(values_from = visited, names_from = bb.sp, values_fill = 0) %>%
+  pivot_longer(cols = 2:17, values_to = "visited", names_to = "bb.sp")
+
+weighted_fl_abund <- fl_abund_sp %>%
+  group_by(site, date, plant.genus) %>%
+  summarize(flower.cover = sum(fl.abund)) %>%
+  left_join(weights) %>%
+  mutate(weighted.flower.cover = flower.cover*visited) %>%
+  group_by(site, date, bb.sp) %>%
+  summarize(total.weighted.flower.cover = sum(weighted.flower.cover)) %>%
+  left_join(site_data) %>%
+  select(site, date, elev.mean, bb.sp, total.weighted.flower.cover) %>%
+  mutate(yday = yday(date),
+         bb.sp = factor(bb.sp),
+         site = factor(site),
+         year = factor(year(date))) %>%
+  filter(bb.sp %in% c("pasc", "prat", "soro", "wurf", "telu", "hort"))
+
+dat <- gam_data %>%
+  left_join(weighted_fl_abund) %>%
+  mutate(perc.vis = (visits/bb.sp.abund)*100,
+         pref = perc.vis - perc.flower.cover)
+
+ggplot(dat, aes(log(total.weighted.flower.cover), pref, color = ktype)) +
+  geom_point(alpha = 0.25) +
+  geom_smooth() +
+  facet_wrap(~bb.sp)
+
+gam0 <- bam(bin.visits ~
+                 s(pair, bs = "re") + # center by pair
+                 bb.sp.abund*perc.flower.cover + # condition on a global bb/fl abundance term
+                 perc.flower.cover*pair + # condition on a pair-specific fl abundance
+                 s(total.weighted.flower.cover, by = ktype),
+               family = "binomial", # binomial response
+               method = "fREML",
+               discrete = TRUE,
+               select =  TRUE,
+               data = filter(dat,
                              bb.sp %in% c("soro", "wurf", "hort",
@@
 
-#saveRDS(gam_pvis_elev01, "../output/gam_pvis_elev01.rds")
-#gam_pvis_elev01 <- readRDS("../output/gam_pvis_elev01.rds") %>% getViz()
-
-summary(gam_pvis_elev01)
+brm_lability00 <- brm(as.integer(visits.bin) ~
+                        bb.sp.abund*perc.flower.cover +
+                        total.weighted.flower.cover +
+                        (1 + perc.flower.cover + total.weighted.flower.cover | ktype:bb.sp) + 
+                        (1 | site), # a random intercept and for site
+                      family = "bernoulli", 
+                      #file = "../output/brm_lability_00",  # for running locally
+                      file = "brm_lability_00", # for running on the HPC
+                      chains = 4,
+                      cores = 12,
+                      iter = 2000,
+                      data = dat)
+
+library(glmmTMB)
+library(sjPlot)
+
+glm_lability00 <- glmmTMB(as.integer(bin.visits) ~
+                            perc.flower.cover*bb.sp.abund +
+                            total.weighted.flower.cover*pair +
+                            (1 + perc.flower.cover | pair) +
+                            (1 | site), # a random intercept and for site
+                          family = "binomial", 
+                          data = dat)
+
+summary(glm_lability00)
+
+plot_model(glm_lability00)
+
+```
+
+# Is preference labile at the level of individual floral species? We could pick out a widely distributed representative of each morphotype and see
+```{r}
+flower_distributions2012 <- flower_cover %>%
+  mutate(yday = yday(date),
+         year = year(date)) %>%
+  filter(year == 2012) %>%
+  left_join(site_data) %>%
+  left_join(fl_traits)
+
+# Choose the top three most frequently observed species of each morphotype
+top_sp_in_morph_2012 <- flower_distributions2012 %>%
+  filter(!k.type.ss %in% c("0", "9", "10")) %>%
+  group_by(plant.sp, k.type.ss) %>%
+  summarize(obs = n()) %>%
+  group_by(k.type.ss) %>%
+  top_n(3, obs) %>%
+  arrange(k.type.ss, -obs)
+
+# Disc (1)
+ggplot(filter(flower_distributions2012, plant.sp == "Potentilla erecta"), aes(yday, elev.mean)) +
+  geom_point(alpha = 0.25, aes(size = flower.cover))
+
+ggplot(filter(flower_distributions2012, plant.sp == "Alchemilla vulgaris"), aes(yday, elev.mean)) +
+  geom_point(alpha = 0.25, aes(size = flower.cover))
+
+ggplot(filter(flower_distributions2012, plant.sp == "Ranunculus acris"), aes(yday, elev.mean)) +
+  geom_point(alpha = 0.25, aes(size = flower.cover))
+
+# Funnel (2)
+ggplot(filter(flower_distributions2012, plant.sp == "Gentiana aspera"), aes(yday, elev.mean)) +
+  geom_point(alpha = 0.25, aes(size = flower.cover))
+
+ggplot(filter(flower_distributions2012, plant.sp == "Gentiana clusii"), aes(yday, elev.mean)) +
+  geom_point(alpha = 0.25, aes(size = flower.cover))
+
+ggplot(filter(flower_distributions2012, plant.sp == "Valeriana tripteris"), aes(yday, elev.mean)) +
+  geom_point(alpha = 0.25, aes(size = flower.cover))
+
+# Bell (3)
+ggplot(filter(flower_distributions2012, plant.sp == "Campanula scheuchzeri"), aes(yday, elev.mean)) +
+  geom_point(alpha = 0.25, aes(size = flower.cover))
+
+ggplot(filter(flower_distributions2012, plant.sp == "Soldanella alpina"), aes(yday, elev.mean)) +
+  geom_point(alpha = 0.25, aes(size = flower.cover))
+
+ggplot(filter(flower_distributions2012, plant.sp == "Geum rivale"), aes(yday, elev.mean)) +
+  geom_point(alpha = 0.25, aes(size = flower.cover))
+
+# Stalk-disc (4)
+ggplot(filter(flower_distributions2012, plant.sp == "Silene dioica"), aes(yday, elev.mean)) +
+  geom_point(alpha = 0.25, aes(size = flower.cover))
+
+ggplot(filter(flower_distributions2012, plant.sp == "Myosotis alpestris"), aes(yday, elev.mean)) +
+  geom_point(alpha = 0.25, aes(size = flower.cover))
+
+ggplot(filter(flower_distributions2012, plant.sp == "Gentiana verna"), aes(yday, elev.mean)) +
+  geom_point(alpha = 0.25, aes(size = flower.cover))
+
+# Lip (5)
+ggplot(filter(flower_distributions2012, plant.sp == "Prunella vulgaris"), aes(yday, elev.mean)) +
+  geom_point(alpha = 0.25, aes(size = flower.cover))
+
+ggplot(filter(flower_distributions2012, plant.sp == "Acinos alpinus"), aes(yday, elev.mean)) +
+  geom_point(alpha = 0.25, aes(size = flower.cover))
+
+ggplot(filter(flower_distributions2012, plant.sp == "Euphrasia rostkoviana"), aes(yday, elev.mean)) +
+  geom_point(alpha = 0.25, aes(size = flower.cover))
+
+# Flag (6)
+ggplot(filter(flower_distributions2012, plant.sp == "Lotus corniculatus"), aes(yday, elev.mean)) +
+  geom_point(alpha = 0.25, aes(size = flower.cover))
+
+ggplot(filter(flower_distributions2012, plant.sp == "Trifolium pratense"), aes(yday, elev.mean)) +
+  geom_point(alpha = 0.25, aes(size = flower.cover))
+
+ggplot(filter(flower_distributions2012, plant.sp == "Anthyllis vulneraria"), aes(yday, elev.mean)) +
+  geom_point(alpha = 0.25, aes(size = flower.cover))
+
+# Head (7)
+ggplot(filter(flower_distributions2012, plant.sp == "Leontodon hispidus"), aes(yday, elev.mean)) +
+  geom_point(alpha = 0.25, aes(size = flower.cover))
+
+ggplot(filter(flower_distributions2012, plant.sp == "Carduus defloratus"), aes(yday, elev.mean)) +
+  geom_point(alpha = 0.25, aes(size = flower.cover))
+
+ggplot(filter(flower_distributions2012, plant.sp == "Leucanthemum vulgare"), aes(yday, elev.mean)) +
+  geom_point(alpha = 0.25, aes(size = flower.cover))
+```
+
+```{r}
+prop_visits_focal <- visit_cover %>%
+  group_by(site, date, bb.sp, plant.sp) %>%
+  summarize(flower.cover = sum(flower.cover), # get raw flower cover and visits by site*date*bb.sp*plant.sp
+            visits = sum(visits)) %>%
+  group_by(site, date) %>%
+  mutate(prop.flower.cover = flower.cover/sum(flower.cover))%>% # binary visits
+  group_by(site, date, bb.sp) %>%
+  mutate(prop.visits = visits/sum(visits),
+         bb.sp.abund = sum(visits)) %>% # abundance of each BB sp. per site-date
+  ungroup() %>%
+  left_join(bb_traits) %>% # add BB traits
+  mutate(tongue = factor(pbl.w.class2),
+         visits = as.integer(visits)) %>% # so we can run a Poisson model
+  left_join(site_data) %>% # add site data
+  ungroup() %>%
+  mutate(yday = yday(date),
+         perc.flower.cover = prop.flower.cover*100) %>%
+  select(site, elev.mean, date, yday, bb.sp, plant.sp, visits, 
+         prop.visits, flower.cover, perc.flower.cover, bb.sp.abund, tongue) %>%
+  mutate(visits.bin = if_else(visits > 0, 1, 0),
+         year = factor(year(date))) %>%
+  filter(year ==  "2012") %>%
+  inner_join(top_sp_in_morph_2012)
+
+gam_data_focal <- prop_visits_focal %>%
+  unite(pair, c(bb.sp, plant.sp), sep = "_", remove = FALSE) %>% # paste bb.sp and ktype into pair
+  mutate(pair = factor(pair),
+         bin.visits = if_else(visits > 0, 1, 0), # binarize visitation
+         year = factor(year(date)),
+         site = factor(site))
+
+gam_pvis_elev01_focal <- bam(bin.visits ~
+                         s(pair, bs = "re") + # center by pair
+                         bb.sp.abund:perc.flower.cover + # condition on a global bb/fl abundance term
+                         perc.flower.cover:pair + # condition on a pair-specific fl abundance
+                         te(yday, elev.mean, by = pair, bs = c("gp", "tp"), k = 5), # conditional effects of yday/elev by pair
+                       family = "binomial", # binomial response
+                       method = "fREML",
+                       discrete = TRUE,
+                       select =  TRUE,
+                       data = filter(gam_data_focal,
+                                     bb.sp %in% c("soro", "wurf", "hort",
+                                                  "telu", "pasc", "prat"))) %>%
+  getViz()
+
+gam_pvis_elev01.sum <- summary(gam_pvis_elev01)
 check.gamViz(gam_pvis_elev01)
-gam_pvis_elev01.plot <- plot(gam_pvis_elev01)
-gam_pvis_elev01.plot[[1]]gam_pvis_elev01
-str(gam_pvis_elev01)
-
-hort3 <- plot(sm(gam_pvis_elev01, 2))
-hort5 <- plot(sm(gam_pvis_elev01, 3))
-hort6 <- plot(sm(gam_pvis_elev01, 4))
-hort7 <- plot(sm(gam_pvis_elev01, 5))
-
-pasc3 <- plot(sm(gam_pvis_elev01, 6))
-pasc5 <- plot(sm(gam_pvis_elev01, 7))
-pasc6 <- plot(sm(gam_pvis_elev01, 8))
-pasc7 <- plot(sm(gam_pvis_elev01, 9))
-
-prat3 <- plot(sm(gam_pvis_elev01, 10))
-prat5 <- plot(sm(gam_pvis_elev01, 11))
-prat6 <- plot(sm(gam_pvis_elev01, 12))
-prat7 <- plot(sm(gam_pvis_elev01, 13))
-
-soro3 <- plot(sm(gam_pvis_elev01, 14))
-soro5 <- plot(sm(gam_pvis_elev01, 15))
-soro6 <- plot(sm(gam_pvis_elev01, 16))
-soro7 <- plot(sm(gam_pvis_elev01, 17))
-
-telu3 <- plot(sm(gam_pvis_elev01, 18))
-telu5 <- plot(sm(gam_pvis_elev01, 19))
-telu6 <- plot(sm(gam_pvis_elev01, 20))
-telu7 <- plot(sm(gam_pvis_elev01, 21))
-
-wurf3 <- plot(sm(gam_pvis_elev01, 22))
-wurf5 <- plot(sm(gam_pvis_elev01, 23))
-wurf6 <- plot(sm(gam_pvis_elev01, 24))
-wurf7 <- plot(sm(gam_pvis_elev01, 25))
-
-pterm1 <- plot(pterm(gam_pvis_elev01, 1))
-pterm2 <- plot(pterm(gam_pvis_elev01, 2))
-pterm3 <- plot(pterm(gam_pvis_elev01, 3))
-
-hort3
-pasc3
-prat3
-soro3
-telu3
-wurf3
-
-hort5
-pasc5
-prat5
-soro5
-telu5
-wurf5
-
-hort6
-pasc6
-prat6
-soro6
-telu6
-wurf6
-
-hort7
-pasc7
-prat7
-soro7
-telu7
-wurf7
-
-grid_plot <- gridPrint(hort3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        hort5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
-                        hort6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        hort7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        pasc3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        pasc5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
-                        pasc6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        pasc7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        prat3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        prat5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
-                        prat6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        prat7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        soro3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        soro5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
-                        soro6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        soro7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        telu3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        telu5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
-                        telu6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        telu7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        wurf3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        wurf5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
-                        wurf6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        wurf7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        ncol = 4)
-
-ggsave(grid_plot, file = "../output/gam_plot1.pdf")
-
-color_ramp <- gridPrint(hort3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL))
-ggsave(color_ramp, file = "../output/color_ramp.pdf")
-
-```
+concurvity(gam_pvis_elev01)
+```
+
+# Niche overlap
+```{r}
+niche_data <- visitation %>%
+  left_join(fl_traits)
+
+niche_overlapper <- function(x, criterion) {
+
+  criterion <- enquo(criterion) # criterion is set to either plant.sp (for taxonomic) or k.type.s (for morphological)
+
+  x %>%
+    filter(bb.sp %in% c("telu", "pasc", "prat", "soro", "wurf")) %>% # focal species
+    group_by(bb.sp, !!criterion, site, date) %>% # group by site-year and BB-FL pair
+    summarize(freq = sum(visits)) %>% # get interaction frequency for each BB-FL pair per site-year
+    group_by(bb.sp, site, date) %>% # group by site-year and BB species
+    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
+    group_by(bb.sp, site, !!criterion) %>%
+    summarize(mean.prop = mean(prop)) %>%
+    pivot_wider(names_from = !!criterion, values_from = mean.prop, values_fill = 0) %>% # spread into the kind of array vegan likes
+    group_by(site) %>%
+    nest() %>% # nest into list-column
+    mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
+           dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
+           tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
+    dplyr::select(site, tidy_dist) %>% # drop intermediate list-columns
+    unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
+    left_join(site_data) %>% # add site data
+    dplyr::select(site, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
+    left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
+    rename(subgenus1 = subgenus, tongue1 = pbl.w.class2) %>%
+    left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
+    rename(subgenus2 = subgenus, tongue2 = pbl.w.class2) %>%
+    dplyr::select(site, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
+           tongue1, tongue2, elev.mean) %>%
+    mutate(tongue.diff = if_else(tongue1 == tongue2, "same", "different"),
+           subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different")) %>%
+
+    unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":", remove = FALSE) %>% # create a single species pair column...
+    mutate(sp.pair = factor(sp.pair)) %>% # ...and make it a factor so that mgcv can handle it
+    mutate(niche.overlap = 1-distance) # express niche overlap so that 1 = perfect overlap and 0 = no overlap
+}
+
+niche_overlap_gen <- niche_overlapper(niche_data, plant.genus) %>%
+  mutate(level = rep("genus", n()))
+
+niche_overlap_fam <- niche_overlapper(niche_data, plant.family) %>%
+  mutate(level = rep("family", n()))
+
+niche_overlap_ktype <- niche_overlapper(niche_data, k.type.ss) %>%
+  mutate(level = rep("morphotype", n()))
+
+niche_pool <- bind_rows(niche_overlap_gen,
+                        niche_overlap_fam, 
+                        niche_overlap_ktype,) %>%
+  mutate(level = factor(level, levels = c("morphotype", "family", "genus")),
+         tongue.diff = factor(tongue.diff))
+
+ggplot(niche_pool, aes(elev.mean, niche.overlap, color = level)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~sp.pair) +
+  theme_light()
+
+ggplot(niche_pool, aes(elev.mean, niche.overlap, color = level)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  theme_light()
+
+ggplot(niche_pool, aes(sp.pair, niche.overlap, color = level)) +
+  geom_boxplot()
+
+ggplot(niche_pool, aes(level, niche.overlap)) +
+  geom_boxplot()
+
+ggplot(niche_pool, aes(niche.overlap)) +
+  geom_histogram(bins = 10) +
+  facet_wrap(~level)
+
+morpho_gam_niche.I <- bam(niche.overlap ~
+                          sp.pair +
+                          #s(elev.mean, k = 5) +
+                          s(elev.mean, by = sp.pair, k = 5, m = 1),
+                        method = "fREML",
+                        family = "gaussian",
+                        discrete = TRUE,
+                        select = TRUE,
+                        data = filter(niche_pool, level == "morphotype")) %>% getViz()
+
+check.gamViz(morpho_gam_niche.I)
+summary(morpho_gam_niche.I)
+print(plot.gamViz(morpho_gam_niche.I, allTerms = TRUE), pages = 1)
+
+fam_gam_niche.I <- bam(niche.overlap ~
+                          sp.pair +
+                          s(elev.mean, by = sp.pair, k = 5),
+                        method = "fREML",
+                        family = "gaussian",
+                        discrete = TRUE,
+                        select = TRUE,
+                        data = filter(niche_pool, level == "family")) %>% getViz()
+
+check.gamViz(fam_gam_niche.I)
+summary(fam_gam_niche.I)
+print(plot.gamViz(fam_gam_niche.I, allTerms = TRUE), pages = 1)
+
+gen_gam_niche.I <- bam(niche.overlap ~
+                          sp.pair +
+                          s(elev.mean, by = sp.pair, k = 5),
+                        method = "fREML",
+                        family = "gaussian",
+                        discrete = TRUE,
+                        select = TRUE,
+                        data = filter(niche_pool, level == "genus")) %>% getViz()
+
+check.gamViz(gen_gam_niche.I)
+summary(gen_gam_niche.I)
+plot.gamViz(gen_gam_niche.I, allTerms = TRUE)
+
+```

### 2021-02-27 ms_1/ms_1.Rmd
back up

@@
 ---
-title: "Bumble bees and wildflowers along a mountain slope exhibit nonlinear patterns of abundance and turnover, punctuated by the tree line ecotone"
+title: "Elevational range shifts of mountain bumble bees and their floral hosts will be punctuated by nonlinear turnover of species and interactions"
 
@@
 
+Climate-driven range shifts are a fundamental evolutionary process, but the rapid pace of recent climate change has made biogeographic dynamics as much a matter of current events as of historic speculation.
+This provides a unique opportunity to study biogeographic processes in real time, but it also poses a grave threat to global biodiversity as species are forced to respond rapidly within the constraints of gene pools shaped by cooler climates.
+Climate-induced range shifts involve more than merely tracking suitable temperature envelopes; when species move, they do so in the context of biological communities and under the influence of more than just temperature, and the outcomes of climate-induced range shifts are thus contingent on potentially nonlinear processes that cannot be reduced to simple elevational or latitudinal temperature gradients.
+Bumble bees (Hymentopera: *Bombus* spp.) are cold-adapted species associated with mountain habitats, and they have already exhibited measurable upslope range shifts that are expected to result in habitat loss, novel competitive interactions, and disrupted pollination networks.
+Interpreting these shifts, however, requires an understanding of the current elevational patterns of bumble bees and the floral mutualists upon which climate change will act.
+In this study, we demonstrate that mountain bumble bees and flora are distributed in strongly nonlinear patterns in space and time, and that these patterns propagate from the level of species abundances to that of β-diversity, and from the level of community composition to that of biotic interactions.
+In particular, we show that the tree line ecotone is a critical threshold of floral resource availability, a distributional interface between high- and low-elevation bumble bees species, and a zone of rapid acceleration of species and interaction β-diversity.
+The implications of these findings extend beyond the particular case of bumble bees to demonstrate that linear elevational temperature gradients are ecologically punctuated, and the outcomes of climate-induced range shifts with depend on how species and species-interactions change as they cross --- or fail to cross --- thresholds like the tree line ecotone.
+
 *Keywords*: climate, elevation gradient, pollinator, *Bombus*, alpine
@@
 
-The rate of modern climate change is forcing species to adapt ecologically within the constraints of gene pools shaped by cooler climates [@Visser2008-ct].
+The rate of recent climate change is forcing species to adapt ecologically within the constraints of gene pools shaped by cooler climates [@Visser2008-ct].
 As temperatures rise, species are expected to track suitable conditions to higher latitudes and higher elevations, and such effects are already measurable across a broad range of taxa [@Lenoir2015-tj].
@@
 At this relatively discrete band within a continuous elevational temperature gradient, plant communities exhibit rapid species turnover [@Descombes2017-id] and declining diversity [@Becker2007-cg] as abiotic conditions become more extreme away from the shelter of tree canopy [@Slatyer1992-ir].
-Thus, species range shifts --- and, perhaps more importantly, the ecological consequences thereof --- are contingent upon patterns of species distributions and biotic interactions, and community ecology is as much prior as posterior to the ecology of global change.
+Thus, species range shifts --- and, perhaps more importantly, the ecological consequences thereof --- are contingent upon existing patterns of species distributions and biotic interactions, and community ecology is as much prior as posterior to the ecology of global change.
 
-Bumble bees (Hymenoptera: *Bombus* spp.) are both a classic model organism in community ecology and an emergent conservation priority due to recent declines [@Goulson2008-oe] and manifest vulnerability to climate change [@Soroye2020-zb].
+Bumble bees (Hymenoptera: *Bombus* spp.) are both classic model taxa in community ecology and an emergent conservation priority due to recent declines [@Goulson2008-oe] and manifest vulnerability to climate change [@Soroye2020-zb].
 The genus *Bombus* is believed to have arisen in the mountains of Asia during the cooling climate of the Eocene-Oligocene boundary (~34 mya), and its subsequent spread and diversification along mountain corridors and into lowland habitats was likely driven by alternating range expansion during cooling periods and retreat to higher elevations during warming periods [@Hines2008-er; @Stewart2010-jn; @Martinet2018-rb].
@@
 The outcomes of range shifts in mountain bumble bee and floral communities have obvious significance from a conservation perspective, but they also provide a model system in which to study the general phenomenon of community assembly under climate change, a phenomenon of both historical interest and prospective urgency.
-Anticipating, interpreting, and potentially mitigating the effects of ranges shifts, however, necessitate an understanding the current elevational and temporal patterns of bumble bee communities and those of their floral hosts that will be acted upon by climate change.
+Anticipating, interpreting, and potentially mitigating the effects of ranges shifts, however, hinge upon an understanding the current elevational and temporal patterns of the bumble bee and floral communities that will be acted upon by climate change.
 In particular, it is important to detect nonlinearities in species-, community-, and interaction-level responses to elevational climate gradients, such as the floristic threshold effect that has been documented at the tree line ecotone [@Descombes2017-id].
@@
 Finally, we move to the level of species interactions, investigating the turnover of bee-flower interaction partners through elevation.
-In each of these analyses, we employ flexible modeling techniques capable of capturing nonlinear responses with the goal of identifying where, when, and for which species the impacts of climate-induced range shifts should be expected to be most acute.
+In each of these analyses, we employ flexible modeling techniques capable of capturing nonlinear responses with the goal of identifying where, when, and for which species the impacts of climate-induced range shifts should be expected to be most severe
 
@@
 
-The study was conducted in Berchtesgaden National Park (47.55°N, 12.92°E), located in the Northern Limestone Alps of southeast Germany.
+The study was conducted 2010-2012 in Berchtesgaden National Park (47.55°N, 12.92°E), located in the Northern Limestone Alps of southeast Germany.
 The landscape is composed of mountain pastures mainly surrounded by coniferous forests.
@@
 
-Sampling consisted of repeated visits to each study site at approximately weekly intervals.
+Sampling in each year of the study consisted of repeated visits to each site at approximately weekly intervals.
 Samples were only collected on dry days when the air temperature was at least 6°C.
@@
 Floral visitation by male bumble bees was recorded during visitation sampling, but we chose to analyze only visitation by queens and workers.
-
 In conjunction with visitation observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.1 m^2^.
 Species identification followed @Lauber2007-ha and @Oberndorfer2001-ro.
+
+While per-sample effort was standardized across all years, the total number of samples (i.e.
+site-dates) varied markedly,  with 153 in 2010, 212 in 2011, and 539 in 2012.
+The increased total sampling effort in 2012 was driven in part by the overall scarcity of bumble bees that year.
+Since the start of field sampling each year was determined by snowmelt, lower sites were sampled as early as April while higher sites could not be sampled until June or July.
+This resulted in a temporally staggered pattern of sampling in 2010 and 2012, while in 2011 a combination of early snowmelt and a late start to field activities resulted in sampling that was approximately aligned in time across the whole elevation gradient.
+See Appendix X for details of sampling patterns.
 
@@
 This approach was intended to dampen the effects of false non-detection on the estimation of floral resource abundance, particularly for rare bumble bee species whose diet breadth would tend to be underestimated simply due to sparsity of observations [@Williams2005-yp].
+We also opted to treat the cryptic *B.
+terrestris/lucorum* species pair as a single morphospecies, since these two species cannot be distinguished reliably in the field.
+Similarly, we pooled representatives of the parasitic *Psithyrus* subgenus --- which included *B.
+barbutellus*, *B.
+bohemicus*, *B.
+campestris*, *B.
+flavidus*, *B.
+quadricolor*, and *B.
+sylvestris* --- into a single species group *B.
+psithryus*.
 
@@
 PCoA and species β-diversity calculation were performed with the package `vegan` `r citep(citation("vegan"))`.
-Calculation and partitioning of interaction β-diversity were performed with the package `bipartite` `r citep(citation("bipartite"))`.
+Calculation and partitioning of interaction β-diversity were performed with the package `bipartite` [@Dormann2008-aa; Dormann2009-aa].
 Matrix regression was performed with the package `ecodist` [@Goslee2007-aa].
@@
 We recorded a total of 12,918 bumble- bee-flower interactions (excluding males) over the three years of our study.
-The metaweb across all sites and dates consisted of 16 bumble bee species (with *Bombus terrestric/lucorum* and *Bombus psithyrus* species groups lumped into species groups), 163 plant species (110 genera, 37 families), and 736 unique bumble- bee-plant interaction pairs.
+The metaweb across all sites and dates consisted of 16 bumble bee species, 163 plant species (110 genera, 37 families), and 736 unique bumble- bee-plant interaction pairs.
 Five species --- *B.
@@
 pratorum*, *B.
-soroensis*, *B.
+soroeensis*, *B.
 terrestris-lucorum*, and *B.
@@
 Of these, *B.
-soroensis* was exceptional in that it exhibited peak abundance at the highest sites, but it was also found at relatively high abundance across the whole elevation gradient, peaking at mid-season.
+soroeensis* was exceptional in that it exhibited peak abundance at the highest sites, but it was also found at relatively high abundance across the whole elevation gradient, peaking at mid-season.
 *B.
@@
 The only species to exhibit marked anomalies in floral resource availability was *B.
-gerstaeckeri*, which specializes on the genus *Aconitum*[@Ponchau2006-af] but was also observed (presumably collecting nectar) on *Gentiana*, *Salvia*, *Rhinanthus*, *Cirsium*, and *Clinopodium.* Its floral resources were concentrated in two discrete peaks, the first at low elevation and mid-season, and the second at high elevation and late-season.
+gerstaeckeri*, which specializes on the genus *Aconitum* [@Ponchau2006-af] but was also observed (presumably collecting nectar) on *Gentiana*, *Salvia*, *Rhinanthus*, *Cirsium*, and *Clinopodium.* Its floral resources were concentrated in two discrete peaks, the first at low elevation and mid-season, and the second at high elevation and late-season.
 
@@
 Bumble bee β-diversity was, in comparison to floral β-diversity, both lower overall and less responsive to elevational difference between sites **(Figure 3A)**.
-Sites at similar elevation exhibited ~25%
-species turnover, and species turnover between the most widely separated sites remained less than 50%.
+Sites at similar elevation exhibited ~25% species turnover, and species turnover between the most widely separated sites remained less than 50%.
 
@@
 
-```{r fig4, echo = FALSE,  fig.cap = paste("GDM splines of species-level and interaction-level β-diversity in response to geographic, elevational, and temporal distance between site-dates.
-The maximum height of each spline represents a given variable’s partial effect on β-diversity (its effect when covariates are held constant), and the shape of each spline represents to the rate of species turnover as it varies along the corresponding gradient (i.e.
-geographic distance, elevation, or day of year), with steeper parts of the curve indicating regions of the gradient over which species/interaction turnover is more rapid.
+```{r fig4, echo = FALSE,  fig.cap = paste("GDM splines of species-level and interaction-level β-diversity in response to elevational, controlling for geographic proximity and number of sampling dates.
+The maximum height of each spline represents a given variable’s partial effect on β-diversity (its effect when covariates are held constant), and the shape of each spline represents to the rate of species turnover as it varies along the corresponding gradient, with steeper parts of the curve indicating regions of the gradient over which species/interaction turnover is more rapid.
 Within-site comparisons were omitted from GDM analysis to avoid pseudoreplication.
@@
 
-- Recap the point of the study
-
 The core finding of our study, evident in each of our analyses, is that a linear elevation gradient can generate nonlinear biological responses in terms of both abundance and β-diversity, and the latter both at the level of community composition and at the level of species interactions.
 
-The most salient pattern in our study is the unique importance of the elevation band falling at about 1500 m, which corresponds to the lower edge of the tree line ecotone in the Berchtesgadener Alps [@Kostler1970-ya; @Mayer1970-zn] and specifically that of the elevational transects used in our study.
+Arguably the most salient pattern in our data is the unique importance of the elevation band falling at about 1500 m, which corresponds to the lower edge of the tree line ecotone in the Berchtesgadener Alps [@Kostler1970-ya; @Mayer1970-zn] and specifically that of the elevational transects used in our study.
 The floristic importance of the tree line ecotone has been noted in previous work [@Pellissier2010-kg; @Descombes2017-id].
@@
 
-The observation that the upslope advancement of alpine tree lines often lags behind climate warming [@Dullinger2004-qu] suggests a danger that high elevation bumble bees could move upslope faster than their floral resources.
-Indeed, a somewhat unexpected finding of our study is that high-elevation bumble bee species (*B.
+The observation that the upslope advancement of alpine tree lines often lags behind climate warming [@Dullinger2004-qu] suggests that high elevation bumble bees could move upslope faster than their floral resources, though the opposite pattern could also occur if the climate-tracking of bumble bees is limited by other factors, such as nest site availability [@Marshall2020-pc].
+A somewhat unexpected finding of our study is that high-elevation bumble bee species (*B.
 mucidus*, *B.
@@
 Our analyses of species and interaction β-diversity reveal a striking contrast between a relative stable bumble bee community and an extremely dynamic floral community.
-The dominance of total interaction β-diversity by floral species turnover is consistent with the findings of @Simanonok2014-ou in the Rocky Mountains of North America, and it is perhaps a predictable relationship given the exceptional cold-hardiness of bumble bees [@Heinrich1994-cv] and the sensitivity of plants to temperature gradients [].
-Nevertheless, it highlights the behavioral challenge that bumble bees have presumably faced for the entirety of their evolutionary history [] and that may explain the remarkable flexibility and intelligence of their flower handling []: how to forage efficiently in a diverse that turns over rapidly in both space and time.
+The dominance of total interaction β-diversity by floral species turnover is consistent with the findings of @Simanonok2014-ou in the Rocky Mountains of North America, and it is perhaps a predictable relationship given the exceptional cold-hardiness of bumble bees and the sensitivity of plants to temperature gradients.
+Nevertheless, it highlights the behavioral challenge that bumble bees have presumably faced for the entirety of their evolutionary history as mountain florivores: how to forage efficiently in a diverse floral community that turns over rapidly in both space and time.
+Indeed, the remarkable flexibility and intelligence of bumble bee foraging [@Heinrich1979-ep; @Loukola2017-bt], which has been interpreted as a side-effect of their sociality [@Dukas1991-ig], could be interpreted alternatively as an adaptation to precisely this problem of foraging in the context of extreme β-diversity.
 A fascinating question that could not be answered by our sampling approach is whether mountain bumble bees adaptively forage up- or downslope, as suggested by @Lundberg1980-qk.
@@
 Assuming an average slope of 20° and a foraging range of 1 km, a bumble bee could travel up- or downslope by more than 340 m, thus spanning an elevation belt nearly 700 m wide.
-Such 3-dimensional foraging would enable bumble bees both to exploit the elevational turnover of floral species and to track preferred species through their elevationally staggered phenology, the latter constituting a sort of physiological time travel [@Van_Straalen1983-df.
+Such 3-dimensional foraging would enable bumble bees both to exploit the elevational turnover of floral species and to track preferred species through their elevationally staggered phenology, the latter constituting a sort of physiological time travel [@Van_Straalen1983-df].
 
@@
 
-- What do these findings tell use about climate change and range shifts in mountain ecosystems more generally?
-- Do our results suggest any management strategies?
+Mountains are unique ecological theaters in which extrinsic temporal climate change interacts with intrinsic elevational climate gradation to alter the spatial and temporal distributions of species and interactions [@Telwala2013-sc; @CaraDonna2014-nq; @Miller-Struttmann2014-wz; @Rafferty2020-aj].
+The central finding of our study, that elevation gradients are characterized by nonlinear responses of species, communities, and interaction networks --- punctuated dramatically by the tree line ecotone --- is almost certainly not an idiosyncrasy of bumble bees and their floral hosts but rather a pattern emerging from fundamental physical constraints that should be expected to obtain broadly across the biotic communities of mountains worldwide.
+Species interactions along tree line ecotones, therefore, deserve special consideration in research and conservation management, since climate-induced range shifts at the tree line ecotone should be expected to have disproportionate ecological consequences relative to range shifts at other elevation bands.
 
-As in the past, so now are mountains the primarily theaters of ecological change for bumble bees.
+<!-- Retreating to higher elevation in response to climatic warming is not a new phenomenon in the evolutionary history of bumble bees [@Hines2008-er; @Martinet2018-rb].
+Nevertheless, the rapid pace of recent climate change, together with the interacting effects of human land use, are predicted to [@Soroye2020-zb; @Sirois-Delisle2018-vi]  -->
 
-Mountains unique ecological theaters in which extrinsic temporal climate change interacts with an intrinsic elevational climate gradation to alter the spatial and temporal distributions of species and interactions [@Telwala2013-sc; @CaraDonna2014-nq; @Miller-Struttmann2014-wz; @Rafferty2020-aj].
+<!-- - What do these findings tell use about climate change and range shifts in mountain ecosystems more generally? -->
+<!-- - Do our results suggest any management strategies? -->
 
-It should also be expected that climate change on mountain slopes will be punctuated by the tree line ecotone, which constitutes, at least with respect to bumble bees and wildflowers, a region of unique resource abundance and a transition to high and accelerating turnover of species and interactions [@@Descombes2017-id].
+<!-- Retreating to higher elevation in response to climatic warming is not a new phenomenon in the evolutionary history of bumble bees, but the rate of modern climate change, together with the interacting effects of human land use, are predicted to [@Soroye2020-zb; @Sirois-Delisle2018-vi]  -->
 
-though as a matter of applied conservation it is perhaps doubtful that local interventions could forestall for long the consequences of a process occurring at a global scale.
-
-Retreating to higher elevation in response to climatic warming is not a new phenomenon in the evolutionary history of bumble bees
-
-The fact that
-
-As such, the tree line ecotone deserves special consideration in conservation management, and targeted studies of plant-pollinator relationships at the tree line ecotone would be warranted.
+<!-- though as a matter of applied conservation it is perhaps doubtful that local interventions could forestall for long the consequences of a process occurring at a global scale. -->
 

### 2021-02-27 ms_2/ms_2.Rmd
back up

---
title: "Bumble bees exhibit strong but flexible foraging biases through temporal and elevational gradients of floristic turnover"

author: Douglas B. Sponsler^1^*, Fabrice Requier^2^, Katharina Kallnik^1^, Alice Claßen^1^, A. Fabienne Maihoff^1^, Ingolf Steffan-Dewenter^1^

# date: `r Sys.Date()`

output:
  bookdown::word_document2:
    fig_caption: yes
    fig_width: 7
    fig_height: 7
    df_print: kable
    keep_md: yes
    number_sections: no
    toc: no
    reference_docx: Word_template.docx

bibliography: 
  - references.bib
#  - knitcitations.bib
csl: ecology.csl
# csl: https://www.zotero.org/styles/ecology

---

^1^ Department of Animal Ecology and Tropical Biology, Biocenter, University of Würzburg, Würzburg, Germany

^2^ Université Paris-Saclay, CNRS, IRD, UMR Évolution, Génomes, Comportement et Écologie, 91198, Gif-sur-Yvette, France

`*` Corresponding author: douglas.sponsler@uni-wuerzburg.de

# ABSTRACT

*Keywords*: optimal foraging, trait matching, pollinator, *Bombus*, alpine

```{r setup, include=FALSE, cache=FALSE, message = FALSE}

library("knitr")

### Chunk options: see http://yihui.name/knitr/options/ ###
## Text results
opts_chunk$set(echo = FALSE, warning = TRUE, message = TRUE)

## Code decoration
opts_chunk$set(tidy = TRUE, comment = NA, highlight = TRUE)

## Cache
# opts_chunk$set(cache = 2, cache.path = "output/cache/")

## Plots
opts_chunk$set(fig.path = "../output/")

```

```{r knitcitations, cache = FALSE}
library(knitcitations)
cleanbib()   
cite_options(citation_format = "pandoc")
```

# INTRODUCTION

<!-- Open with a discussion of the problem of foraging efficiently in a dynamic foraging environment, the balance of bias and flexibility, etc. -->

Bumble bees are exceptionally versatile foragers whose ability to solve the puzzles posed by complex flower morphologies (Plowright and Laverty 1984), or indeed by human investigators (Alem et al. 2016; Loukola et al. 2017), renders almost any flower potentially accessible. Even when a gross mismatch between tongue length and corolla depth makes the nectar of a given flower inaccessible by ordinary means, some bumble bee species will resort readily to biting through the base of the flower to “rob” nectar, incidentally making the nectar available to other prospective foragers as well. Flower handling skills are, moreover, socially transmitted, amplifying the plasticity of individual behavior (Leadbeater and Chittka 2008; Loukola et al. 2017). Nevertheless, the flexibility of bumble bee foraging is tuned to floral reward, which depends on both the intrinsic reward of a given flower and the searching/learning/handling costs associated with exploiting it. At an evolutionary level, the innate biases of bumble bees for certain floral traits likely reflect historic associations between floral traits and floral rewards (Raine and Chittka 2007). These innate biases, however, can be modulated or even nullified by patterns of reward learned during foraging (Kugler 1943; Maharaj et al. 2019), which depend not only on intrinsic floral properties but also ecological modulators, such as the spatial distribution of floral patches or competition with other bumble bees (Ishii 2013; Plowright and Laverty 1984; Sowig 1989).

The topic of floral bias in bumble bees has been approached via both laboratory experimentation (e.g. Kugler 1943; Raine and Chittka 2007) and observational field study (e.g. Heinrich 1976; Dave Goulson, Lye, and Darvill 2008). The former approach enables stronger causal inference with respect to the mechanisms of floral choice, but it suffers from all the usual limitations of biological realism associated with laboratory conditions. Observational field studies, however, tend to be anecdotal in scope, typically describing bumble bee foraging patterns in particular localities, over short time frames, and with limited floristic data to enable a distinction between true foraging bias and the neutral effects of floral abundance. The steep elevation gradients of mountain slopes provide an opportunity for more powerful observational study of bumble foraging because they generate broad gradients of environmental variation while controlling for confounding effects of geographic variation, such as genotypic differentiation within bumble bee species. When bumble bee visitation in such environments is sampled with sufficient spatial and temporal replication and complemented with quantitative floral surveying, true bias can be distinguished from the neutral effects of abundance, and variation in bias can be explored along elevational and temporal dimensions of floristic turnover.

In the present study, we focus on the question of bumble bee foraging bias with respect to floral morphology, and the lability of morphological biases across gradients of floristic variation. At the outset, we emphasize the use of the term “bias” in favor of the term “preference,” since the latter is laden with more assumptions and its inference under field conditions is problematic (Williams 2005). By “bias,” we mean simply an empirical disproportionality between the relative abundance of a given floral taxon or morphological class and its proportional visitation. The exact mechanism of this disproportionality --- whether an arbitrary “preference” on the part of the foraging bee, an unmeasured decoupling of relative abundance and rate of reward, an exogenous effect of competition, etc. --- is beyond the scope of our discussion and probably best studied under laboratory conditions where strict factorial control is possible. We focus instead on the empirical phenomenon of bias, its variation, and its morphological correlations.

The meaning of “morphology,” however, is hardly more clear-cut than that of bias. Bumble bees and flowers vary in an array of interrelated morphological parameters, both within and between species, and the selection of relevant and measurable morphological metrics is nontrivial. For bumble bees, morphological variation consists principally (though not exclusively) of variation in tongue length. Tongue length in bumble bees corresponds functionally to corolla depth in flowers, and there is a well-established (though imperfect) association between tongue length and floral choice (e.g. Inouye 1980). Importantly, though, accessing floral reward involves more than just a match between tongue length and corolla depth. Such matching is entirely bypassed by nectar robbers, but even in the case of “legitimate” foraging, access to reward involves learning to manipulate floral structures that, in some floral morphotypes, are challenging irrespective of tongue length (Laverty 1994). For example, the characteristic Papilionaceous floral morphology of legumes requires a bee to separate tightly appressed keel petals to access anthers and nectaries (Westerkamp 1997), and the narrowly campanulate flowers of many Ericaceous flora pose a significant barrier to large-bodied visitors like bumble bees (Wallace 1977). Rather than considering various floral morphological traits independently, we rely on the synthesis of Kugler (1970), who developed a typology consisting of ten primary morphotypes and nested subdivision of each. Kugler’s typology was informed by laboratory studies of bumble bee flower choice (Kugler 1943), and there is a precedent for analyzing bumble bee visitation in terms of Kugler’s morphotypes (Neumayer and Paulus 1999; Schneller et al. 2014).

<!-- Formulate this section more as a list of motivating questions; perhaps end with something that ties back into the more general introduction. -->

How do floral morphotypes compare in terms of total bumble bee visitation?

How does morphological bias vary across bumble bee species and tongue length classes?

Species' morphological biases give rise to patterns of diet (dis)similarity

How are the floral morphotypes distributed through elevation and time in our mountain ecosystem?

For ubiquitously distributed bumble bee species, how does morphological bias vary through elevation and time?

First, we compare the overall rate of bumble bee visitation across floral morphotypes. Then, we parse visitation rates by bumble bee species and explore the functional relationship between floral morphotype and bumble bee tongue length. These biases, together with the background effects of relative abundance, generate emergent patterns of diet (dis)similarity, which we present in the form of ordination analysis. Lastly, we analyze the lability of the morphological biases of individual bumble bee species through elevation and time, and compare these patterns of bias across bumble bee species and within floral morphotypes. The results of these analyses demonstrate both the strength of bumble bee floral bias, arising in part from morphological constraints, and the flexibility of bumble bee floral bias that enables them to forage adaptively in the dynamic floral communities of mountain landscapes.

# METHODS

## Field system

<!-- I guess I'll need to tweak the wording of this section so that it is not a verbatim copy of the other manuscript. But it's okay if it recycles a lot of language because it is, after all, the same thing being described. -->

<!-- Describe general floristic and temperature patterns. Describe sampling in more detail. -->

Field sampling was conducted 2010-2012 in Berchtesgaden National Park (47.55°N, 12.92°E), located in the Northern Limestone Alps of southeast Germany. The landscape is composed of mountain pastures in a coniferous forest matrix. We selected 25 study sites (60 x 60 m) on pastures at elevations ranging from 641-2032 m above sea level. Fourteen of these pastures are extensively grazed by cattle or sheep, three are mowed for hay production, and eight have been abandoned throughout the last century and are no longer subject to any human management. The sites sampled in this study are the same as those used in a series of previous studies [@Hoiss2012-sn; @Hoiss2012-kr; @Hoiss2015-gv], but the data reported here are otherwise independent.

Sampling consisted of repeated visits (hereafter, “visitation samples”) to each study site at approximately weekly intervals. Samples were only collected on dry days when the air temperature was at least 6°C. During each visit, bumble bees and their activity were recorded during a 50-minute transect walk. Bumblebees observed on or a given flower were counted as floral visitors. Bumble bee queens were identified to species level in the field, whereas workers and males were stored in individually labeled tubes in the freezer for later identification in the laboratory after Amiet (1996). While floral visitation by male bumble bees was recorded during visitation sampling, we chose to analyze only visitation by queens and workers. During observation, it was noted whether each bee was visibly carrying pollen at the time of its visit, but this provides at best an imperfect distinction between pollen and nectar foragers, and we opted not to differentiate between bees with and without pollen in our analyses. In conjunction with visitation observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.1 m^2^. Species identification followed @Lauber2007-ha and @Oberndorfer2001-ro.

While per-sample effort was standardized across all years, the total number of samples (i.e. site-dates) varied markedly,  with 153 in 2010, 212 in 2011, and 539 in 2012. The increased total sampling effort in 2012 was driven in part by the overall scarcity of bumble bees that year. Since the start of field sampling each year was determined by snowmelt, lower sites were sampled as early as April while higher sites could not be sampled until June or July. This resulted in a temporally staggered pattern of sampling in 2010 and 2012, while in 2011 a combination of early snowmelt and a late start to field activities resulted in sampling that was approximately aligned in time across the whole elevation gradient. See Appendix X for details of sampling patterns.

## Data analysis

### Data processing

Bumble bee tongue-length data were obtained from the literature values compiled and summarized by Arbetman et al. (2017). Values for *B. gerstaeckeri* and *B. mendax*, which were not included in Arbetman et al. (2017), were obtained from Obeso (1992) and Durieux (2000), respectively. We then binned bumble bee species into three tongue-length classes: short (< 7.5 mm), medium (7.5—10.0 mm), and long (>= 10.0 mm). Short-tongued species included *B. jonellus*, *B. lapidarius*, *B. pratorum*, *B. psithyrus*-group, *B. pyrenaeus*, and *B. terrestris/lucorum*. Medium-tongued species included *B. humilis*, *B. hypnorum*, *B. monticola*, *B. mucidus*, *B. pascuorum*, and *B. wurflenii*. Long-tongued species included *B. gerstaeckeri*, *B. hortorum*, and *B. mendax*.

Floral morphology was summarized using floral morphotypes of Kugler (1970), accessed via the BIOFLOR database (Klotz, Kühn, and Durka 2002). Morphotypes were simplified to the primary classes numbered 0-10: 0 = not applicable (includes wind-pollinated flora such as grasses), 1 = disc- and bowl-shaped flowers, 2 = funnel flowers, 3 = bell-shaped flowers, 4 = stalk-disc flowers, 5 = lip flowers, 6 = flag flowers, 7 = flower heads, 8 = spike flowers, 9 = brush flowers, 10 = trap flowers **(Figure 1)**. In a few cases, a plant species in our data set was not found in the BIOFLOR database. In most cases, morphotypes are constant at the genus level, and missing morphotypes were added based on the morphotypes found for congeners. The genus *Gentiana* consisted of both group 2 and group 4 representatives, so the missing species *Gentiana aspera* and *Gentiana ciliata* were classified based on visual comparison with their classified congeners. In the case of the genus *Phyteuma*, we followed Neumayer and Paulus (1999) and classified it as a group 7 flower head. Morphotypes 9 and 10 were extremely rare in our study system and were therefore omitted from all analyses. Plants belonging to morphotype 0 were typically not recorded during floral surveys due to their lack of conspicuous floral displays, so this morphotype was also omitted from all analyses.

Prior to analysis, we aligned each visitation sample with a corresponding floral survey sample. In most cases, floral surveying was conducted jointly with bumble bee observation for each site and date. Occasionally, though, floral surveying could not be completed on the same day as bumble bee observation, resulting in a small temporal offset between visitation samples and floral survey samples. In these cases, visitation samples were paired with the nearest available floral survey data within a maximum distance of 7 days. Visitation samples lacking corresponding floral survey data within 7 days were omitted from further analysis. In a few cases, two consecutive visitation samples at a given site shared the same nearest floral survey sample; in these cases, the visitation samples were first pooled, then aligned to their shared nearest floral survey sample. In rare cases, a given floral species was recorded during visitation sampling but missed during floral surveying. These species were retroactively assigned the minimum floral area value of 0.01 m^2.

Visitation data were extremely right-skewed because the majority of available flora received zero visits during a given sampling period. To facilitate model fitting, we created a binary version of visitation rate in which visited flora were assigned to 1 and unvisited flora to 0. We then modeled binary visitation using generalized models with binomial or Bernoulli distribution families. The conditional effects inferred from these models should, therefore, be interpreted as effects on visitation probability rather than visitation frequency.

While our study spanned three years, we chose not to model year effects. Differences in sampling intensity and duration across years, for the reasons described above, make comparisons across years difficult to interpret, and we deemed these comparisons not worth the cost of added model complexity.

<!-- Should I be including year as a varying effect in these models? -->

<!-- State more explicityl how the models should be interpreted withr expect to bias -->

<!-- Report the effects of abundance that I am conditioning on -->

<!-- Perhaps show what happens if we remove the abundance interaction terms -->

<!-- Mor edescriptive fgure captions -->

<!-- Add something about model diagnostics -->

<!-- Consider adding year as a varying effect -->

<!-- Figure 3 order by tongue length class -->

<!-- Describe the differences in models in the Figure2 and 3 captions -->

<!-- Explain filtering of ordination plot -->

<!-- Year should really be included in Model 5, or just use 2012. -->

<!-- Open each results paragraph with a summary sentence. -->

### Model 1: Comparing overall visitation probability across floral morphotypes

We first compared overall visitation probability across floral morphotypes. In this model, visitation for each sample was pooled across bumble bee species and within floral morphotypes prior to binarization. Visitation probability was then modeled using multilevel Bayesian regression with a Bernoulli distribution family (logit link function). *Bumble bee abundance* and *percent flower cover* (per morphotype) were included as a global interaction term, thus conditioning visitation probability on abundance and enabling the inference of bias. *Floral morphotype* was specified as a varying effect term with both a varying intercept and a varying slope with respect to *percent flower cover*, and a varying intercept term was also included for *site*. See Appendix X for the details of model specification and evaluation.

### Model 2: Visitation probability as a function of floral morphotype and bumble bee tongue length

To explore the functional underpinnings of bumble foraging bias, we analyzed visitation probability as a function of the interaction between floral morphology and bumble bee tongue length class. For this analysis, visitation was pooled within floral morphotype but not across bumble bee species. We then specified a Bayesian model of the same form as Model 1, but this time included a varying slope and intercept term for floral morphotype : bumble bee tongue length class and floral morphotype : bumble bee species. In this model, the relationship between floral morphotype and bumble bee tongue length class was the focus of inference, and the bumble bee species effect was added to avoid pseudoreplication arising from the grouping of observation within species. See Appendix X for the details of model specification and evaluation.

### Model 3: Visitation probability as a function of floral morphotype and bumble bee species

To focus on species-level morphotype biases, we removed the tongue length effect from Model 2 and re-fit the model considering only bumble bee species. See Appendix X for the details of model specification and evaluation.

### Model 4: Aggregate diet ordination

To visualize the net diet similarity between bumble bee species, we pooled visitation data across all samples within bumble bee species and plant species, then performed a principal coordinates analysis based on Horn-Morisita distance metric. To test the significance of tongue length class as a predictor of diet similarity, we performed a permutation test.

### Model 5: The variation of morphological bias through elevation and time

To explore the plasticity of floral biases, we first subsetted our data to include only the dominant bumble bee species (B. hortorum, B. pascuorum, B. pratorum, B. soroeensis, B. terrestris/lucorum, and B. wurflenii) and the most visited floral morphotypes (groups 3, 5, 6, and 7). We then constructed a generalized additive model (GAM) in which binary visitation was modeled as a tensor-product smooth of elevation and day-of-year, grouped by floral morphotype. This response was conditioned on linear interaction terms for bumble bee abundance : percent flower cover and percent flower cover : floral morphotype and a random intercept term for bumble bee species : floral morphotype. See Appendix X for the details of model specification and evaluation.

## Software

All analyses were conducted in `R` `r citep(citation())`. Bayesian multilevel models were implemented in Stan (Stan Development Team 2021) via the package `brms` `r citep(citation("brms"))`, and visualizations were constructed with the packages `tidybayes` `r citep(citation("tidybayes"))` and `ggdist` `r citep(citation("ggdist"))`. Ordination analysis and permutation testing were performed with the package `vegan` `r citep(citation("vegan"))` and visualized using the packages `ggplot2` `vegan` `r citep(citation("ggplot2"))` and `ggvegan` (Simpson 2019). GAMs were implemented with the package `mgcv` [@Wood_2017] and visualized with the package `mgcViz` [@Fasiolo2018aa]. Data handling and visualization was performed with the `tidyverse` suite of R packages `r citep(citation("gdm"))`. Annotated R code is available in the Supplementary Material.

# RESULTS

## Summary of visitation and floral survey data

<!-- Modify this section to describe how many observations/samples were dropped due to the non-alignment of visitation and survey data -->

We recorded a total of 12,918 bumble- bee-flower interactions (excluding males) over the three years of our study. The metaweb across all sites and dates consisted of 16 bumble bee species (with *Bombus terrestric/lucorum* and *Bombus psithyrus* species groups lumped into species groups), 163 plant species (110 genera, 37 families), and 736 unique bumble- bee-plant interaction pairs. Five species --- *B. pascuorum*, *B. pratorum*, *B. soroeensis*, *B. terrestris-lucorum*, and *B. wurflenii* --- accounted for the bulk of overall bumble bee abundance and were present over the entire elevation range of our sites **(Figure 1)**. *B. hortorum*, *B. jonellus*, and *B. psithyrus* (species-group) were widespread but at lower abundance. *B. monticola* occurred at moderate abundance but was rare or absent at sites below 1000 m. The remaining species were relatively rare in our study system. *B. mendax*, *B. mucidus*, and *B. pyrenaeus* occurred primarily above 1500 m, consistent with their known affinity for high elevation habitats [@Rasmont2010-ci]. *B. gerstaeckeri*, *B. hypnorum*, *B. lapidarius* were sparsely distributed across mid-elevation sites. *B. humilis* was recorded only three times and only in one year, so we omitted it from all analyses.

Floral surveying yielded a total of 352 plant species, representing 191 genera and 52 families. Of these, 155 species from 103 genera and 35 families were observed to be visited by bumble bees. Eight species --- *Rubus idaeus*, *Rosa canina*, *Juniperus communis*, *Larix decidua*, *Salix* sp., *Caltha palustris*, *Pulmonaria officinalis*, and *Rheum barbarum* --- were recorded in visitation data but not recorded during floral surveying. Each accounted for no more than 4 visits in total over the three years of our study, and they were omitted from the analysis of floral abundance.

```{r fig1, echo = FALSE,  fig.cap = paste("Each floral morphotype is exemplified by a representative species common in our study system. Illustrations are modifications of public domain works obtained from www.plantillustrations.org. For details of original sources, see Appendix X.")}
knitr::include_graphics("../output/ktype_abund_gam_mod.png")
```

## Overall visitation probability by floral morphotype

Lip (group 5) and flag (group 6) flowers had the highest overall visitation probability, with a conditional mean visitation probabilities of 0.36 (95% CI = 0.30–0.42) and 0.33 (95% CI = 0.27–0.39), respectively (Figure 1). In second place, head flowers (group 7) had a mean conditional visitation probability of 0.19 (95% CI = 0.16–0.24). Bell (group 3) flowers were a distant third with a conditional mean visitation probability of 0.092 (95% CI = 0.07–0.12), followed by stalk-disc (group 4), funnel (group 2), and disc flowers (group 1) with conditional visitation probabilities of 0.07 (95% CI = 0.05–0.09), 0.05 (95% CI = 0.04–0.07), and 0.05 (95% CI = 0.03–0.07), respectively. Unclassified (group 0) flowers, consisting mainly of wind-pollinated grasses, were the least-visited, with a conditional mean visitation probability of 0.03 (95% CI = 0.02–0.04). Our model explained approximately 35% of the total variance (97.5% CI = 34–36%).

```{r fig2, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability by floral morphotype. Point-interval plots depict 66\\% and 95\\% CIs.")}
knitr::include_graphics("../output/brm_pvis01_plot1.pdf")
```

## Visitation probability by flower morphotype, bumble bee species, and bumble bee tongue length class

Descriptions of tongue classes are based on Model 2 and refer to Figure 2, and descriptions of species are based on Model 3 and refer to Figure 3. Both models explained approximately 28% of total variance (97.5% CI = 27–29%). As in Model 1, visitation rates refer to conditional visitation probabilities.

Disc (group 1) flowers were visited at similar rates across tongue length classes. At the species level, disc flowers were visited at relatively high rates by the short-tongued B. terrestris/lucorum and the medium-tongued Bombus monticola. Visitation to funnel (group 2) flowers was dominated by long-tongued species, an effect driven entirely by the Aconitum (group 2) specialist B. gerstaeckeri. Bell (group 3) flowers were visited at a higher rate by short-tongued species than by long-tongued species, with medium-tongued species exhibiting intermediate visitation rates. At the species-level, bell flowers were visited at highest probability by the short-tongued species B. jonellus, B. pyrenaeus, and B. soroeensis, and by the medium-tongued species B. hypnorum and B. monticola. Stalk-disc (group 4) flowers were visited at similar rates across all tongue length classes, and there was also broad overlap in visitation rates at the species level. Lip (group 5) flowers were also visited at similar rates across tongue length classes, but at the species level B. pascuorum exhibited an uniquely high preference for lip flowers. Flag (group 6) flowers were visited at a higher rate by medium-tongued species than by short-tongued species, with long-tongued species exhibiting intermediate visitation rates. Flag flowers generated strong separation in visitation rate at the species level, receiving very high visitation from the medium-tongued B. pascuorum and B. mucidus, moderately high visitation from the short-tongued B. terrestris/lucorum and B. lapidarius, the medium-tongued B. monticola, and the long-tongued B. mendax and B. hortorum, and low visitation by the short-tongued B. soroeensis, B. pyrenaeus, B. psithyrus, B. pratorum, and B. jonellus, the medium-tongued B. hypnorum, and the long-tongued B. gerstaeckeri.

```{r fig3, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability by Kugler morphotype and bumble bee tongue length class. Point-interval plots depict 66\\% and 95\\% CIs. Letters indicate group differences based on evidence ratio at alpha = 0.05. Note that the range of the x-axis has been rescaled for each panel to emphasize the comparison of visitation probabilities across tongue length classes.")}
knitr::include_graphics("../output/brm_pvis00_plot1.pdf")
```

```{r fig4, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability by Kugler morphotype and bumble bee species. The tongue length class of each species is color-coded. Point-interval plots depict 66\\% and 95\\% CIs. Note that the range of the x-axis has been rescaled for each panel to emphasize the comparison of visitation probabilities across bumble bee species")}
knitr::include_graphics("../output/brm_pvis02_plot1.pdf")
```

## Diet ordination

Axis 1 accounted for 24% of total diet variance and primarily separated long- and medium-tongued species from short-tongued species (Figure 4). Long- and medium-tongued species scored low on Axis one and were associated with the lip flowers (group 5) Ajuga reptans, Rhinanthus glacialis, and Stachys alopecuros) and flag flowers (group 6) Anthyllis vulneraria, Lotus corniculatus, and Trifolium pratense. Short tongued species scored high on Axis one and were associated with the bell flowers (group 3) Campanula scheuchzeri, Erica carnea, and Geum rivale and the head flowers (group 7) Carduus defloratus, Centaurea jacea, and Phyteuma orbiculare. Axis 2 accounted for 19% of total variance and primarily explained variation within short-tongued species. Short-tongued species scoring high on Axis 2 were associated with Campanula scheuchzeri while those scoring low on Axis 2 were associated with Erica carnea. Short-tongued species scoring close to zero on Axis 2 also scored lower on Axis 1 (closer to medium- and long-tongued species) and were associated with head flowers. Tongue length class explained 22% of diet variance (Adonis permutation test, F = 1.68, p = 0.006). While short-tongued bees appeared most broadly spread in ordination space, there were no significant differences in dispersion across tongue length classes (F = 0.62, p = 0.55).

```{r fig5, echo = FALSE,  fig.cap = paste(" Principal coordinates ordination of bumble bee visitation pooled across all samples.")}
knitr::include_graphics("../output/pcoa_aggregate.png")
```

## The variation of morphological bias through elevation and time

Model 5 explained 33.8% of total variance (adjusted R2 = 0.37). All tensor product smooths (Figure 5) were significant at p < 0.001 except for the one representing B. pratorum and flag (group 6), which was more weakly significant at p = 0.02. As in Models 1-3, the results of Model 5 should be interpreted in terms of conditional visitation probability.

Bell (group 3) flowers were visited at highest rates early in the season and at low elevation by all species except B. hortorum, whose visitation rate to bell flowers peaked in mid-season and at mid-elevation. Notably, the medium-tongued B. pascuorum and B. wurflenii also exhibited relatively high visitation to bell flowers at mid-season and mid-elevation, while the visitation rates of the short-tongued B. pratorum, B. soroeensis. and B. terrestris lucorum were more concentrated at the early-season and low-elevation peak.

Lip (group 5) flowers exhibited roughly the mirror image of the pattern of visitation rate seen in bell flowers, with peak visitation late in the season and at low elevation, but relatively high visitation extending toward mid-season and mid-elevation. This pattern was consistent for all bumble bee species.

The pattern of visitation rates to flag (group 6) flowers fell generally along the lines of tongue length class. The short-tongued B. pratorum and B. soroeensis exhibited peak visitation early in the season and at low elevation. The long-tongued B. hortorum and the medium-tongued B. pascuorum and B. wurflenii exhibited peak visitation at mid-season, though B. pascuorum exhibited a bimodal pattern in elevation, with peaks around 1000 m and 1750 m, while the visitation rate of B. hortorum declined monotonically from high elevation to low. The short-tongued B. terrestis/lucorum broke the tongue-length pattern, exhibiting a diffuse pattern of visitation rate more similar to that of the long- and medium-tongued species than the other short-tongued species.

Visitation rates for head (group 7) flowers were more or less bimodal for all species, with a first peak in mid-season and low elevation and a second peak in late-season and high elevation.

```{r fig6, echo = FALSE,  fig.cap = paste("GAM analysis of bias as a function of the tensor-product interaction of elevation and day of year, parsed by bumble bee species and floral morphotype and conditioned on the abundance of each bumble bee species and the relative abundance of each floral morphotype. The color ramp of each smooth has been scaled to equal color range, thereby obscuring the absolute magnitude of effect. To avoid cluttering the figure, we have opted not to include a unique color ramp key for each plot, since the focus of our inference is on relative rather than absolute patterns.")}
knitr::include_graphics("../output/ktype_lability_gam_mod.png")
```

# DISCUSSION

## Foraging bias and floral morphology

Our study is not the first to compare bumble bee visitation rates across Kugler’s (1970) floral morphotypes (see Neumayer and Paulus 1999; Schneller et al. 2014), but to our knowledge it is the first to enable the inference of bias across morphotypes by controlling statistically for the neutral effects of abundance.

When bumble bee visitation is pooled across all species, it is clear that lip, flag, and head flowers dominate the floral community in terms of conditional visitation probability. Patterns of pooled visitation are, of course, driven disproportionately by the behavior of the most abundant bumble bee species, which in our system were B. hortorum, B. pascuorum, B. pratorum, B. psithyrus, B. soroeensis, B. terrestris/lucorum, and B. wurflenii. Visitation to lip and flag flowers was driven especially by the medium-tongued B. pascuorum, the most abundant bumble bee in our study system, and secondarily by the medium-tongued B. wurflenii. Visitation to head flowers was driven mainly by the short-tongued B. soroeensis and B. psithyrus.

The influence of tongue length on visitation bias was pronounced for some morphotypes but not for others. In the cases of disc, and stalk-disc, and unclassified flowers, visitation rates were uniformly low across all tongue length classes. The opposite was true of lip flowers, which were visited at similarly high rates by all tongue length classes. Bell flowers and head flowers were visited chiefly by short-tongued bees, consistent with the shallow corollas of these morphotypes. In contrast, funnel flowers, with their deep corollas were visited chiefly by long-tongued bees, though this effect was driven disproportionately by B. gerstaeckeri, which specializes on the funnel flowers of the genus Aconitum. Flag flowers were visited at the highest rate by medium-tongued species, and especially by B. pascuorum and B. mucidus.

While the variation in floral bias across tongue length classes we observed largely comports with the established relationship between tongue length and corolla depth, variation in bias within tongue length classes demonstrates the insufficiency of simple morphological matching to explain bumble bee foraging patterns. Nectar robbing is one mechanism by which the relationship between tongue length and corolla depth might be relaxed, but its influence in our data is unclear. Of the bumble bees in our study system, B. terrestris/lucorum and especially B. wurflenii are known to engage frequently in nectar robbing (Brian 1954). In the case of B. wurflenii, however, its morphotype affinities did not differ markedly from those of other medium-tongued species, and it may be that the condition of medium-tongue length alone confers enough to foraging flexibility in our study system to dampen the divergent effects of nectar robbing. Nectar robbing might, however, account for the deviation of B. terrestris/lucorum from most other short-tongued species with respect to flag flowers. With the exception of B. lapidarius, all short-tongued species besides B. terrestris/lucorum eschewed flag flowers, while B. terrestris/lucorum visited them at rates comparable to medium- and long-tongued species. Indeed, B. terrestris/lucorum was, on the whole, an idiosyncratic forager, demonstrating a relatively high affinity for disc flowers and unclassified flowers (particularly Plantago lanceolata) that were largely ignored by other species. Perhaps the most striking divergence of floral bias within tongue length class, though, was the divisive effect of bell flowers on short-tongued bumble bees. B. pyrenaeus, B. soroeensis, B. pratorum, and B. jonellus visited bell flowers at high rates, while B. terrestris/lucorum, B. psithyrus, and B. lapidarius visited them relatively rarely. A similar split among the short-tongued species occurred with respect to head flowers, which were visited at high rates by B. soroeensis and B. psithyrus but at relatively low rates by other short-tongued species. There are no obvious morphological distinctions within short-tongued bumble bees that would account for these divergent affinities; ecological drivers such as competition or differences in spatial or temporal activity patterns are more likely explanations, as will be revisited below in our discussion of overall diet similarity.

## Diet similarity

Our finding that bumble bee diet ordination produces a primary axis corresponding to tongue length — and associated with affinity for legumes — and a secondary axis describing the dispersion of short-tongued species is consistent with studies in Poland (Dave Goulson, Lye, and Darvill 2008), France (Durieux 2000), and England (D. Goulson et al. 2005) that analyzed bumble bee diet by similar methods, suggesting that this may be a general pattern among bumble bee communities, at least within temperate Europe. While our test of beta-dispersion did not reveal significant differences between tongue length classes, the spread of short-tongued bees along the secondary axis of our ordination, together with their divergence with respect to visitation rates on bell and head flowers described above, suggests that resource partitioning may be more pronounced among short-tongued bumble bees than among their medium- and long-tongued counterparts. It is tempting to invoke interspecific competition as a cause of the apparent niche partitioning among short-tongued bumble bees, in agreement with previous studies that have demonstrated the potential for severe competition between bumble bees of similar tongue length (e.g. Heinrich 1976; Inouye 1978; Pyke 1982). But the importance of competition in bumble bee communities is not universal (e.g. Ranta and Vepsäläinen 1981; Williams 1989), and in the rich wildflower meadows of our (sub)alpine system, bumble bee populations are more likely limited by harsh abiotic conditions than by food availablility. Alternatively, it may be that short-tongued bees in our study system differed from one another in their affinities for specific habitats (Schneller et al. 2014) or spatial configurations of resource patches (Sowig 1989), and their divergent morphological biases arose secondarily from these effects.

## The lability of bias through elevation and time

While the intraspecific lability of bumble bee foraging biases has been demonstrated under laboratory conditions (Ings, Raine, and Chittka 2009; Maharaj et al. 2019) and in a localized field study (Raine and Chittka 2007), ours is the first study, to our knowledge, to characterize the in situ lability of foraging biases within multiple species through elevational and temporal gradients of floristic turnover. In our analysis, it is evident that morphological biases in bumble bees, insofar as they are captured by the morphotypology used in our study, are not fixed but rather tuned to floristic (and possibly abiotic) context and therefore dynamic through elevation and time. Notably, bumble bee species differed from one another not just in their overall bias toward given morphotypes but also in when and where they were biased toward a given morphotype. This is especially evident in the case of flag flowers, for which positive bias was maximized in mid-season for B. hortorum, B. pascuorum, B. terrestris/lucorum, and B. wurflenii but in early-season for B. pratorum and B. soroeensis. This pattern corresponds to the overall biases of these species toward flag flowers, which were primary resources for B. hortorum, B. pascuorum, B. terrestris/lucorum, and B. wurflenii but only marginal resources for B. pratorum and B. soroeensis.

# CONCLUSIONS

For bumble bees, floral bias is a labile product of innate preferences, the soft constraints of morphological compatibility, and the dynamic ecological context in which foraging occurs. By conditioning bumble bee visitation data upon concurrent floral abundance data across spatially and temporally replicated samples, we demonstrate the strength bumble bee floral bias, its functional underpinnings, its variation across species, and its lability within species in response to environmental variation. Such tuning of floral bias has obvious adaptive significance for bumble bees, which are characteristically — and most likely primordially — associated with mountain habitats (Williams 1985; Hines 2008), where steep temperature gradients and short growing seasons generate extreme spatial and temporal floral turnover.

# ACKNOWLEDGEMENTS

We thank D.J. McNeil for helpful conversations during data analysis.

# REFERENCES

<!-- ```{r write_citations, cache=FALSE, include=FALSE} -->
<!-- write.bibtex(file = "knitcitations.bib") -->
<!-- ``` -->

<div id ="refs"></div>

### 2021-03-03 writing/ms_abundance_turnover.Rmd
update

--- 2021-02-13 writing/manuscript_v5.Rmd
@@
 ---
-title: Bumble bees exhibit strong but flexible foraging biases through temporal and
-  elevational gradients of floristic turnover
+title: Bumble bees and wildflowers along a mountain slope exhibit nonlinear patterns of abundance and turnover, punctuated by the tree line ecotone
 author:
@@
 
-Bumble bees are exceptionally versatile foragers whose ability to solve the puzzles posed by complex flower morphologies [@Plowright1984-lc], or indeed by human investigators [@Alem2016-rn; @Loukola2017-bt], renders almost any flower potentially accessible.
-Even when a gross mismatch between tongue length and corolla depth makes the nectar of a given flower inaccessible by ordinary means, some bumble bee species will resort readily to biting through the base of the flower to “rob” nectar, incidentally making the nectar available to other prospective foragers as well.
-Flower handling skills are, moreover, socially transmitted, amplifying the plasticity of individual behavior [@Leadbeater2008-rh; @Loukola2017-bt].
-Nevertheless, the flexibility of bumble bee foraging is tuned to floral reward, which depends on both the intrinsic reward of a given flower and the searching/learning/handling costs associated with exploiting it.
-At an evolutionary level, the innate biases of bumble bees for certain floral traits likely reflect historic associations between floral traits and floral rewards [@Raine2007-nq].
-These innate biases, however, can be modulated or even nullified by patterns of reward learned during foraging [@Kugler1943-uy; @Maharaj2019-wm], which depend not only on intrinsic floral properties but also ecological modulators, such as the spatial distribution of floral patches or competition with other bumble bees [@Ishii2013-sm; @Plowright1984-lc; @Sowig1989-ld].
-
-The topic of floral bias in bumble bees has been approached via both laboratory experimentation [e.g.
-@Kugler1943-uy; @Raine2007-nq] and observational field study [e.g.
-@Heinrich1976-wc; @Goulson2008-zy].
-The former approach enables stronger causal inference with respect to the mechanisms of floral choice, but it suffers from all the usual limitations of biological realism associated with laboratory conditions.
-Observational field studies, however, tend to be anecdotal in scope, typically describing bumble bee foraging patterns in particular localities, over short time frames, and with limited floristic data to enable a distinction between true foraging bias and the neutral effects of floral abundance.
-The steep elevation gradients of mountain slopes provide an opportunity for more powerful observational study of bumble foraging because they generate broad gradients of environmental variation while controlling for confounding effects of geographic variation, such as genotypic differentiation within bumble bee species.
-When bumble bee visitation in such environments is sampled with sufficient spatial and temporal replication and complemented with quantitative floral surveying, true bias can be distinguished from the neutral effects of abundance, and variation in bias can be explored along elevational and temporal dimensions of floristic turnover.
-
-In the present study, we focus on the question of bumble bee foraging bias with respect to floral morphology, and the lability of morphological biases across gradients of floristic variation.
-At the outset, we emphasize the use of the term "bias" in favor of the term "preference", since the latter is laden with more assumptions and its inference under field conditions is problematic [@Williams2005-yp].
-By "bias", we mean simply an empirical disproportionality between the relative abundance of a given floral taxon or group and its proportional visitation.
-The exact mechanism of this disproportionality --- whether an arbitrary "preference" on the part of the foraging bee, an unmeasured decoupling of relative abundance and rate of reward, an exogenous effect of competition, etc.
---- is beyond the scope of our discussion, and probably best studied under laboratory conditions where strict factorial control is possible.
-We focus instead on the empirical phenomenon of bias, its variation, and its morphological correlations.
-
-The meaning of "morphology", however, is hardly more clear-cut than that of bias.
-Bumble bees and flowers vary in an array of interrelated morphological parameters, both within and between species, and the selection of relevant and measurable morphological metrics is nontrivial.
-For bumble bees, morphological variation consists principally (though not exclusively) of variation in tongue length.
-Tongue length in bumble bees corresponds functionally to corolla depth in flowers, and there is a well-established (though imperfect) association between tongue length and floral choice [e.g.
-@Inouye1980-mv].
-Importantly, though, accessing floral reward involves more than just a match between tongue length and corolla depth.
-Such matching is entirely bypassed by nectar robbers, but even in the case of "legitimate" foraging, access to reward involves learning to manipulate floral structures that, in some floral morphotypes, are challenging irrespective of tongue length [@Laverty1994-hm].
-For example, the characteristic Papilionaceous floral morphology of legumes requires a bee to separate tightly appressed keel petals to access anthers and nectaries [@Westerkamp1997-hs], and the narrowly campanulate flowers of many Ericaceous flora pose a significant barrier to large-bodied visitors like bumble bees [@Wallace1977-kz].
-Rather than considering various floral morphological traits independently, we rely on the synthesis of @Kugler1970-gb, who developed a typology consisting of ten primary morphotypes and nested subdivision of each.
-Kugler's typology was informed by laboratory studies of bumble bee flower choice [@Kugler1943-uy], and there is a precedent for analyzing bumble bee visitation in terms of Kugler's morphotypes [@Neumayer1999-il; @Schneller2014-ry].
-
-First, we compare the overall rate of bumble bee visitation across floral morphotypes.
-Then, we parse visitation rates by bumble bee species and explore the functional relationship between floral morphotype and bumble bee tongue length.
-These biases, together with the background effects of relative abundance, generate emergent patterns of diet (dis)similarity, which we present in the form of ordination analysis.
-Lastly, we analyze the lability of the morphological biases of individual bumble bee species through elevation and time, and compare these patterns of bias across bumble bee species and within floral morphtypes.
-The results of these analyses demonstrate both the strength of bumble bee floral bias, arising in part from morphological constraints, and the flexibility of bumble bee floral bias that enables them to forage adaptively in the dynamic floral communities of mountain landscapes.
+Anthropogenic climate change is occurring at ecological rather than evolutionary time scales, forcing species to adapt within the constraints of gene pools shaped by cooler climates [@Visser2008-ct].
+As temperatures rise, species are expected to be pushed to higher latitudes and higher elevations, and such effects are already measurable across a broad range of taxa [@Lenoir2015-tj].
 

### 2021-03-03 code/bias_niche.Rmd
oops

--- 2021-02-13 code/preference_lability.Rmd
@@
 # 1. Overall bias
+
 ## 1.1 Calculate visitation probability per plant morphotype.
@@
 
-# 2 Ordination of dietary niche
+# 2. Ordination of dietary niche
 Aggregation changes meaning.
@@
 
-# 3. The lability of bias: How does the foraging bias of each species with respect to each morphotype vary through elevation and time?
+# 3. Dietary niche overlap (i.e. potential competition)
+
+## 3.1 Check BB distribution patterns to decide whom to retain in the analysis
 ```{r}
-# Subset data to include only abundant BB spp and exclude rare morphotypes (9, 10)
-gam_data <- prop_visits_ktype %>%
-  filter(ktype %in% c("3", "5", "6", "7") & # focus on major ktypes
-           bb.sp %in% c("telu", "prat", "pasc", "hort", "soro", "wurf")) %>% # focus on abundance bb.sp
-  unite(pair, c(bb.sp, ktype), sep = "_", remove = FALSE) %>% # paste bb.sp and ktype into pair
-  mutate(pair = factor(pair),
-         bin.visits = if_else(visits > 0, 1, 0)) # binarize visitation
+bb_distribution <- visitation %>%
+  mutate(year = year(date)) %>%
+  left_join(site_data) %>%
+  group_by(site, date, elev.mean, bb.sp, year) %>%
+  summarize(abund = n()) %>%
+  group_by(site, elev.mean, bb.sp, year) %>%
+  summarize(abund = mean(abund)) %>%
+  group_by(bb.sp) %>%
+  mutate(elev.floor = min(elev.mean),
+         elev.ceiling = max(elev.mean),
+         elev.range = elev.ceiling - elev.floor,
+         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))
 
-# Independent smooth model
-gam_pvis_elev01 <- bam(bin.visits ~
-                         s(pair, bs = "re") + # center by pair
-                         bb.sp.abund*perc.flower.cover + # condition on a global bb/fl abundance term
-                         perc.flower.cover*pair + # condition on a pair-specific fl abundance
-                         #s(perc.flower.cover, by = pair, k = 3) +  term
-                         te(yday, elev.mean, by = pair, bs = c("gp", "tp"), k = 5), # conditional effects of yday/elev by pair
-                       family = "binomial", # binomial response
+ggplot(bb_distribution, aes(reorder(bb.sp, elev.med), elev.mean)) +
+  geom_line(size = 2, alpha = 0.25) +
+  geom_point(aes(size = abund), alpha = 0.5) +
+  facet_wrap(~year, ncol = 1) +
+  theme_light(12) +
+  scale_size_continuous(name = "Abundance") +
+  xlab("BB species") +
+  ylab("Elevation") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
+## 3.2 Prep data for calculating niche overlap
+```{r}
+niche_data <- visitation %>%
+  #filter(!bb.sp %in% c("humi", "muci", "gers")) %>%
+  left_join(fl_traits) %>%
+  mutate(year = factor(year(date)))
+```
+
+## 3.3 Function for calculating niche overlap
+```{r}
+niche_overlapper <- function(x, criterion) {
+
+  criterion <- enquo(criterion) # criterion is set to either plant.sp (for taxonomic) or k.type.s (for morphological)
+
+  x %>%
+    #group_by(bb.sp, !!criterion, site, date, year) %>% # group by site-year and BB-FL pair
+    group_by(bb.sp, !!criterion, site, date) %>% # group by site-year and BB-FL pair
+    summarize(freq = sum(visits)) %>% # get interaction frequency for each BB-FL pair per site-year
+    group_by(bb.sp, site, date) %>% # group by site-year and BB species
+    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
+    #group_by(bb.sp, site, year, !!criterion) %>%
+    group_by(bb.sp, site, !!criterion) %>%
+    summarize(mean.prop = mean(prop)) %>%
+    pivot_wider(names_from = !!criterion, values_from = mean.prop, values_fill = 0) %>% # spread into the kind of array vegan likes
+    #group_by(site, year) %>%
+    group_by(site) %>%
+    nest() %>% # nest into list-column
+    mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
+           dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
+           tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
+    #dplyr::select(site, year, tidy_dist) %>% # drop intermediate list-columns
+    dplyr::select(site, tidy_dist) %>% # drop intermediate list-columns
+    unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
+    left_join(site_data) %>% # add site data
+    # dplyr::select(site, year, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
+    dplyr::select(site, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
+    left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
+    rename(subgenus1 = subgenus, tongue1 = pbl.w.class2) %>%
+    left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
+    rename(subgenus2 = subgenus, tongue2 = pbl.w.class2) %>%
+    # dplyr::select(site, year, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
+    #        tongue1, tongue2, elev.mean) %>%
+    dplyr::select(site, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
+       tongue1, tongue2, elev.mean) %>%
+    mutate(tongue.diff = as.logical(if_else(tongue1 == tongue2, FALSE, TRUE)),
+           tongue.pair = factor(case_when(
+             tongue1 == "med" & tongue2 == "med" ~ "med:med",
+             tongue1 == "med" & tongue2 == "long" ~ "med:long",
+             tongue1 == "med" & tongue2 == "short" ~ "short:med",
+             tongue1 == "long" & tongue2 == "med" ~ "med:long",
+             tongue1 == "short" & tongue2 == "med" ~ "short:med",
+             tongue1 == "long" & tongue2 == "short" ~ "short:long",
+             tongue1 == "short" & tongue2 == "long" ~ "short:long",
+             tongue1 == "long" & tongue2 == "long" ~ "long:long",
+             tongue1 == "short" & tongue2 == "short" ~ "short:short",
+           )),
+           subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different")) %>%
+    unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":", remove = FALSE) %>% # create a single species pair column...
+    mutate(sp.pair = factor(sp.pair)) %>% # ...and make it a factor so that mgcv can handle it
+    mutate(niche.overlap = 1-distance) %>% # express niche overlap so that 1 = perfect overlap and 0 = no overlap
+    select(-distance)
+}
+```
+
+## 3.4 Calculate niche overlap pooled within site
+```{r}
+# Pooled within site
+niche_overlap_gen <- niche_overlapper(niche_data, plant.genus) %>%
+  mutate(level = rep("genus", n()),
+         niche.overlap.genus = niche.overlap)
+
+niche_overlap_fam <- niche_overlapper(niche_data, plant.family) %>%
+  mutate(level = rep("family", n()),
+         niche.overlap.family = niche.overlap)
+
+niche_overlap_ktype <- niche_overlapper(niche_data, k.type.ss) %>%
+  mutate(level = rep("morphotype", n()),
+         niche.overlap.ktype = niche.overlap)
+
+niche_pool <- bind_rows(niche_overlap_gen,
+                        niche_overlap_fam, 
+                        niche_overlap_ktype,) %>%
+  mutate(level = factor(level, levels = c("morphotype", "family", "genus")),
+         tongue.diff = tongue.diff,
+         site = factor(site))
+
+write_csv(niche_pool, "../hpc_packets/niche_pool.csv")
+
+morph_genus_difference <- niche_overlap_gen %>%
+  select(-niche.overlap, -level) %>%
+  full_join(niche_overlap_fam) %>%
+  select(-niche.overlap, -level) %>%
+  full_join(niche_overlap_ktype) %>%
+  select(-niche.overlap, -level) %>%
+  mutate(tongue.diff = tongue.diff,
+         site = factor(site),
+         morph.genus.diff = niche.overlap.ktype - niche.overlap.genus)
+
+ggplot(morph_genus_difference, aes(elev.mean, morph.genus.diff)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~tongue.pair)
+
+ggplot(morph_genus_difference, aes(tongue.pair, morph.genus.diff)) +
+  geom_boxplot()
+```
+
+## 3.5 Niche overlap ~ level
+```{r}
+brm_niche00 <- brm(niche.overlap ~ 1 +
+                     level + # fixed effect of level 
+                     (1 | site) + # varying intercept for site
+                     (1 | sp.pair), # varying intercept for sp.pair
+                   family = "zero_one_inflated_beta", # for prop data with 0s and 1s
+                   file = "../output/brm_niche_00", # local
+                   #file = "./brm_niche_00", # HPC
+                   iter = 2000,
+                   chains = 4,
+                   data = niche_pool)
+
+pp_check(brm_niche00)
+bayes_R2(brm_niche00)
+
+get_variables(brm_niche00)
+conditional_effects(brm_niche00)
+mcmc_plot(brm_niche00, pars = c("^r_sp.pair\\[mend*"))
+mcmc_plot(brm_niche00, pars = c("^r_site*"))
+```
+
+## 3.6 Genus niche overlap ~ tongue
+```{r}
+brm_niche10 <- brm(niche.overlap ~ 1 +
+                     (1 | site) + # varying intercept for site
+                     (1 | tongue.pair) +
+                     (1 | sp.pair), # varying intercept for sp.pair
+                   family = "zero_one_inflated_beta", # for prop data with 0s and 1s
+                   file = "../output/brm_niche_10", # local
+                   #file = "./brm_niche_10", # HPC
+                   iter = 2000,
+                   chains = 4,
+                   data = filter(niche_pool, level == "genus"))
+
+pp_check(brm_niche10)
+bayes_R2(brm_niche10)
+
+get_variables(brm_niche10)
+conditional_effects(brm_niche10)
+mcmc_plot(brm_niche10, pars = c("^r_tongue*"))
+mcmc_plot(brm_niche10, pars = c("^r_site*"))
+```
+
+## 3.7 Genus niche overlap ~ sp.pair
+```{r}
+brm_niche20 <- brm(niche.overlap ~ 1 +
+                     (1 | site) + # varying intercept for site
+                     (1 | sp.pair), # varying intercept for sp.pair
+                   family = "zero_one_inflated_beta", # for prop data with 0s and 1s
+                   file = "../output/brm_niche_20", # local
+                   #file = "./brm_niche_20", # HPC
+                   iter = 2000,
+                   chains = 4,
+                   data = filter(niche_pool, level == "genus"))
+
+pp_check(brm_niche20)
+bayes_R2(brm_niche20)
+
+get_variables(brm_niche20)
+conditional_effects(brm_niche20)
+mcmc_plot(brm_niche20, pars = c("^r_sp.pair"))
+mcmc_plot(brm_niche20, pars = c("^r_site*"))
+```
+
+## 3.8 Niche overlap ~ elevation
+```{r}
+# Beta models
+gen_gam_niche.Ib <- bam(niche.overlap ~
+                            s(sp.pair, bs = "re") +
+                            s(elev.mean, by = sp.pair, k = 6, m = 2),
                           method = "fREML",
+                          family = betar(eps = 0.001),
                           discrete = TRUE,
                           select = TRUE,
-                       data = filter(gam_data,
-                                     bb.sp %in% c("soro", "wurf", "hort",
-                                                  "telu", "pasc", "prat"))) %>%
-  getViz()
+                          data = filter(niche_pool, level == "genus")) %>% getViz()
 
-#saveRDS(gam_pvis_elev01, "../output/gam_pvis_elev01.rds")
-#gam_pvis_elev01 <- readRDS("../output/gam_pvis_elev01.rds") %>% getViz()
+check.gamViz(gen_gam_niche.Ib)
+summary(gen_gam_niche.Ib)
+print(plot.gamViz(gen_gam_niche.Ib, allTerms = TRUE), pages = 1)
 
-summary(gam_pvis_elev01)
-check.gamViz(gam_pvis_elev01)
-gam_pvis_elev01.plot <- plot(gam_pvis_elev01)
-gam_pvis_elev01.plot[[1]]gam_pvis_elev01
-str(gam_pvis_elev01)
+gen_gam_niche.GI <- bam(niche.overlap ~
+                            s(sp.pair, bs = "re") +
+                            s(elev.mean, k = 6, m = 2) +
+                            s(elev.mean, by = sp.pair, k = 6, m = 1),
+                          method = "fREML",
+                          family = betar(eps = 0.001),
+                          discrete = TRUE,
+                          select = TRUE,
+                          data = filter(niche_pool, level == "genus")) %>% getViz()
 
-hort3 <- plot(sm(gam_pvis_elev01, 2))
-hort5 <- plot(sm(gam_pvis_elev01, 3))
-hort6 <- plot(sm(gam_pvis_elev01, 4))
-hort7 <- plot(sm(gam_pvis_elev01, 5))
+check.gamViz(gen_gam_niche.GI)
+summary(gen_gam_niche.GI)
+print(plot.gamViz(gen_gam_niche.GI, allTerms = TRUE), pages = 1)
 
-pasc3 <- plot(sm(gam_pvis_elev01, 6))
-pasc5 <- plot(sm(gam_pvis_elev01, 7))
-pasc6 <- plot(sm(gam_pvis_elev01, 8))
-pasc7 <- plot(sm(gam_pvis_elev01, 9))
-
-prat3 <- plot(sm(gam_pvis_elev01, 10))
-prat5 <- plot(sm(gam_pvis_elev01, 11))
-prat6 <- plot(sm(gam_pvis_elev01, 12))
-prat7 <- plot(sm(gam_pvis_elev01, 13))
-
-soro3 <- plot(sm(gam_pvis_elev01, 14))
-soro5 <- plot(sm(gam_pvis_elev01, 15))
-soro6 <- plot(sm(gam_pvis_elev01, 16))
-soro7 <- plot(sm(gam_pvis_elev01, 17))
-
-telu3 <- plot(sm(gam_pvis_elev01, 18))
-telu5 <- plot(sm(gam_pvis_elev01, 19))
-telu6 <- plot(sm(gam_pvis_elev01, 20))
-telu7 <- plot(sm(gam_pvis_elev01, 21))
-
-wurf3 <- plot(sm(gam_pvis_elev01, 22))
-wurf5 <- plot(sm(gam_pvis_elev01, 23))
-wurf6 <- plot(sm(gam_pvis_elev01, 24))
-wurf7 <- plot(sm(gam_pvis_elev01, 25))
-
-pterm1 <- plot(pterm(gam_pvis_elev01, 1))
-pterm2 <- plot(pterm(gam_pvis_elev01, 2))
-pterm3 <- plot(pterm(gam_pvis_elev01, 3))
-
-hort3
-pasc3
-prat3
-soro3
-telu3
-wurf3
-
-hort5
-pasc5
-prat5
-soro5
-telu5
-wurf5
-
-hort6
-pasc6
-prat6
-soro6
-telu6
-wurf6
-
-hort7
-pasc7
-prat7
-soro7
-telu7
-wurf7
-
-grid_plot <- gridPrint(hort3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        hort5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
-                        hort6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        hort7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        pasc3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        pasc5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
-                        pasc6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        pasc7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        prat3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        prat5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
-                        prat6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        prat7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        soro3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        soro5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
-                        soro6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        soro7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        telu3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        telu5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
-                        telu6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        telu7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        wurf3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        wurf5 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE), 
-                        wurf6 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        wurf7 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL) + guides(fill=FALSE),
-                        ncol = 4)
-
-ggsave(grid_plot, file = "../output/gam_plot1.pdf")
-
-color_ramp <- gridPrint(hort3 + l_fitRaster() + l_fitContour() + labs(title = NULL, x = NULL, y = NULL))
-ggsave(color_ramp, file = "../output/color_ramp.pdf")
-
+brm_niche01 <- brm(niche.overlap ~ 1 +
+                     s(elev.mean, by = sp.pair, k = 10) +
+                     (1 | sp.pair),
+                   family = "zero_one_inflated_beta",
+                   #file = "../output/brm_niche_01",
+                   file = "./brm_niche_01",
+                   iter = 2000,
+                   chains = 4,
+                   data = filter(niche_pool, level == "genus"))
 ```

### 2021-03-03 code/bias_niche.Rmd
update writing

@@
 
-morph_genus_difference <- niche_overlap_gen %>%
-  select(-niche.overlap, -level) %>%
-  full_join(niche_overlap_fam) %>%
-  select(-niche.overlap, -level) %>%
-  full_join(niche_overlap_ktype) %>%
-  select(-niche.overlap, -level) %>%
-  mutate(tongue.diff = tongue.diff,
-         site = factor(site),
-         morph.genus.diff = niche.overlap.ktype - niche.overlap.genus)
+niche_pool_sitemean <- niche_pool %>%
+  group_by(site, level, elev.mean) %>%
+  summarize(mean.niche.overlap = mean(niche.overlap))
 
-ggplot(morph_genus_difference, aes(elev.mean, morph.genus.diff)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~tongue.pair)
-
-ggplot(morph_genus_difference, aes(tongue.pair, morph.genus.diff)) +
-  geom_boxplot()
+# morph_genus_difference <- niche_overlap_gen %>%
+#   select(-niche.overlap, -level) %>%
+#   full_join(niche_overlap_fam) %>%
+#   select(-niche.overlap, -level) %>%
+#   full_join(niche_overlap_ktype) %>%
+#   select(-niche.overlap, -level) %>%
+#   mutate(tongue.diff = tongue.diff,
+#          site = factor(site),
+#          morph.genus.diff = niche.overlap.ktype - niche.overlap.genus) 
+# 
+# ggplot(morph_genus_difference, aes(elev.mean, morph.genus.diff)) +
+#   geom_point() +
+#   geom_smooth() +
+#   facet_wrap(~tongue.pair)
+# 
+# ggplot(morph_genus_difference, aes(tongue.pair, morph.genus.diff)) +
+#   geom_boxplot()
 ```
@@
 get_variables(brm_niche10)
-conditional_effects(brm_niche10)
 mcmc_plot(brm_niche10, pars = c("^r_tongue*"))
@@
 ```{r}
-# Beta models
-gen_gam_niche.Ib <- bam(niche.overlap ~
+# Pairwise
+gen_gam_niche.I <- bam(niche.overlap ~
                             s(sp.pair, bs = "re") +
-                            s(elev.mean, by = sp.pair, k = 6, m = 2),
+                            s(site, bs = "re") + # since I have measures repeated over species pairs within each elevation level
+                            s(elev.mean, by = sp.pair, k = 8, m = 2),
                           method = "fREML",
@@
                           select = TRUE,
-                          data = filter(niche_pool, level == "genus")) %>% getViz()
+                          data = filter(niche_pool, level == "genus" &
+                                          bb.sp1 %in% c("pasc", "prat", "soro", "telu", "wurf", # dominant species
+                                                        "mont", "muci", "pyre", "mend") & # high elevation specialists
+                                          bb.sp2 %in% c("pasc", "prat", "soro", "telu", "wurf",
+                                                        "mont", "muci", "pyre", "mend"))) %>% getViz()
 
-check.gamViz(gen_gam_niche.Ib)
-summary(gen_gam_niche.Ib)
-print(plot.gamViz(gen_gam_niche.Ib, allTerms = TRUE), pages = 1)
+check.gamViz(gen_gam_niche.I)
+summary(gen_gam_niche.I)
+plot(sm(gen_gam_niche.I, 9))
+plot(sm(gen_gam_niche.I, 12))
+plot(sm(gen_gam_niche.I, 14))
+plot(sm(gen_gam_niche.I, 15))
+plot(sm(gen_gam_niche.I, 16))
+plot(sm(gen_gam_niche.I, 32))
 
-gen_gam_niche.GI <- bam(niche.overlap ~
-                            s(sp.pair, bs = "re") +
-                            s(elev.mean, k = 6, m = 2) +
-                            s(elev.mean, by = sp.pair, k = 6, m = 1),
-                          method = "fREML",
+# Community mean
+ggplot(niche_pool_sitemean, aes(elev.mean, mean.niche.overlap, color = level)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 8))
+
+mean_niche_gam.I <- gam(mean.niche.overlap ~
+                            level + 
+                            s(elev.mean, by = level, k = 8, m = 2),
+                          method = "REML",
                           family = betar(eps = 0.001),
-                          discrete = TRUE,
                           select = TRUE,
-                          data = filter(niche_pool, level == "genus")) %>% getViz()
+                          data = niche_pool_sitemean) %>% getViz()
 
-check.gamViz(gen_gam_niche.GI)
-summary(gen_gam_niche.GI)
-print(plot.gamViz(gen_gam_niche.GI, allTerms = TRUE), pages = 1)
-
-brm_niche01 <- brm(niche.overlap ~ 1 +
-                     s(elev.mean, by = sp.pair, k = 10) +
-                     (1 | sp.pair),
-                   family = "zero_one_inflated_beta",
-                   #file = "../output/brm_niche_01",
-                   file = "./brm_niche_01",
-                   iter = 2000,
-                   chains = 4,
-                   data = filter(niche_pool, level == "genus"))
+check.gamViz(mean_niche_gam.I)
+summary(mean_niche_gam.I)
+print(plot.gamViz(mean_niche_gam.I, allTerms = TRUE), pages = 1)
 ```

### 2021-03-03 ms_2/ms_2_v2.Rmd
update
update writing

--- 2021-02-27 ms_2/ms_2.Rmd
@@
 ---
-title: "Bumble bees exhibit strong but flexible foraging biases through temporal and elevational gradients of floristic turnover"
+title: "Potential competition in alpine bumble bees is mitigated by imperfect trait-matching and subpartitioning within floral morphotypes"
 
@@
 
-*Keywords*: optimal foraging, trait matching, pollinator, *Bombus*, alpine
+Species coexistence is constrained in part by competition over shared food sources.
+Bumble bees are classic model organisms for the study of competition and coexistence, and this topic has taken on new urgency as climate-induced range shifts are expected to disrupt competitive relationships between bumble bee species, particularly along the climatic gradients of the mountain slopes that host the world's richest bumble bee communities.
+We surveyed floral abundance and recorded bumble bee floral visitation patterns along a ~1400 m elevation gradient in German Alps.
+Using these data, we first ask how floral biases compare across bumble bee species and whether these patterns can be explained by trait-matching between bumble bee tongue length and floral morphotype.
+We then investigate the patterns of diet overlap between bumble bee species that emerge from their respective floral biases, comparing diet overlap at the nested levels of floral morphotype, floral family, and floral genus.
+Finally, we ask how both community-mean and species-pairwise diet overlap vary with elevation.
+While tongue length was a strong predictor of bumble bees' floral morphotype biases, species with similar tongue lengths exhibited divergent biases with respect to certain flora morphotypes.
+This indicates that factors other than tongue length are involved in flower choice, which might explain the coexistence of bumble bee species with similar tongue lengths.
+Moreover, genus-level diet overlap was approximately one half that of morphotype-level diet overlap.
+Thus, even when bumble bees exhibit similar biases at the level of floral morphotype, partly due to the constraints of tongue length, floral diversity enables the mitigation of competition via the genus- or species-level subpartitioning of floral resources within morphotypes.
+This suggests that floral species richness may be as important as overall floral abundance in avoiding the competitive exclusion of bumble bee species.
+We find that community-mean diet overlap, indicative of the potential for diffuse competition, generally declines with elevation, while the diet overlap for certain species pairs exhibits mid- or high-elevation peaks.
+The long-term coexistence of mountain bumble bee species may depend on how climate induced range shifts in bumble bees and their floral hosts will alter the these elevational patterns of diet overlap.
+
+*Keywords*: foraging, trait matching, pollinator, *Bombus*, alpine
 
@@
 
-<!-- Open with a discussion of the problem of foraging efficiently in a dynamic foraging environment, the balance of bias and flexibility, etc. -->
+The role of competition in regulating community assembly and species coexistence has long been a central question in ecology.
+It has taken on new urgency, however, on account of recent climate change, which threatens to disrupt existing patterns of species interactions [@Tylianakis2008-fi], including competition [@Ettinger2017-pk], with consequences for community composition and function.
+These effects are likely to be most pronounced along the steep climate gradients of mountain slopes, where climate-induced range shifts are expected to introduce novel competitive interactions within trophic levels and cause spatial or phenological mismatches between organisms and their food resources.
 
-Bumble bees are exceptionally versatile foragers whose ability to solve the puzzles posed by complex flower morphologies (Plowright and Laverty 1984), or indeed by human investigators (Alem et al.
-2016; Loukola et al.
-2017), renders almost any flower potentially accessible.
-Even when a gross mismatch between tongue length and corolla depth makes the nectar of a given flower inaccessible by ordinary means, some bumble bee species will resort readily to biting through the base of the flower to “rob” nectar, incidentally making the nectar available to other prospective foragers as well.
-Flower handling skills are, moreover, socially transmitted, amplifying the plasticity of individual behavior (Leadbeater and Chittka 2008; Loukola et al.
-2017).
-Nevertheless, the flexibility of bumble bee foraging is tuned to floral reward, which depends on both the intrinsic reward of a given flower and the searching/learning/handling costs associated with exploiting it.
-At an evolutionary level, the innate biases of bumble bees for certain floral traits likely reflect historic associations between floral traits and floral rewards (Raine and Chittka 2007).
-These innate biases, however, can be modulated or even nullified by patterns of reward learned during foraging (Kugler 1943; Maharaj et al.
-2019), which depend not only on intrinsic floral properties but also ecological modulators, such as the spatial distribution of floral patches or competition with other bumble bees (Ishii 2013; Plowright and Laverty 1984; Sowig 1989).
+Bumble bees (Hymenoptera: *Bombus* spp.) have become classic model organisms for the study of competition and species coexistence due to their broad and overlapping distributions and their gross similarity in morphology, behavior, and dietary niche [@Plowright1984-lc; @Goulson2008-zy].
+In this role, however, bumble bees have proven somewhat enigmatic.
+On the one hand, there is evidence that resource limitation can be a significant constraint on local bumble communities [@Heinrich1976-wc; @Bowers1985-df], and certain bumble bee communities have been shown to be over-dispersed in terms of tongue length (a matching trait corresponding to floral corolla depth) or phylogeny (understood to encompass unmeasured trait variation), providing evidence for competition as a community assembly process [@Inouye1977-kl; @Pellissier2013-pk].
+Nevertheless, bumble bee communities with an apparent excess of species richness and deficit of trait (or phylogenetic) dispersion are commonly observed [@Ranta1981-vx; @Goulson2008-zy].
+Teasing apart the still-obscure mechanisms of bumble bee coexistence and niche partitioning remains a salient problem in basic ecology, but it has also become an urgent question in the context of conservation.
+Bumble bees are cold-adapted species [@Heinrich1994-cv] associated mainly with mountain habitats, where they serve as the principal pollinators of many alpine plants [@Williams1998-xi; @Goulson2010-lo].
+Warming climates have already driven measurable elevational range shifts in bumble bees and their floral hosts [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc], and these are predicted to cause declines or even extinction in many bumble bees species [@Soroye2020-zb].
+One mechanism by which climate change is expected to affect bumble bees is via the introduction of novel interspecific competition driven by the advance of lowland bumble bee species into the elevational distributions of their highland counterparts.
+Thus, the speculative question of how competition has acted in the historic assembly of bumble bee communities has become an unfolding process observable over ecological time scales.
 
-The topic of floral bias in bumble bees has been approached via both laboratory experimentation (e.g.
-Kugler 1943; Raine and Chittka 2007) and observational field study (e.g.
-Heinrich 1976; Dave Goulson, Lye, and Darvill 2008).
-The former approach enables stronger causal inference with respect to the mechanisms of floral choice, but it suffers from all the usual limitations of biological realism associated with laboratory conditions.
-Observational field studies, however, tend to be anecdotal in scope, typically describing bumble bee foraging patterns in particular localities, over short time frames, and with limited floristic data to enable a distinction between true foraging bias and the neutral effects of floral abundance.
-The steep elevation gradients of mountain slopes provide an opportunity for more powerful observational study of bumble foraging because they generate broad gradients of environmental variation while controlling for confounding effects of geographic variation, such as genotypic differentiation within bumble bee species.
-When bumble bee visitation in such environments is sampled with sufficient spatial and temporal replication and complemented with quantitative floral surveying, true bias can be distinguished from the neutral effects of abundance, and variation in bias can be explored along elevational and temporal dimensions of floristic turnover.
+All else held equal, species coexistence is favored when intraspecific competition exceeds interspecific competition, and this stabilizing inequality is favored to the extent that species differ in fundamental dietary niche [@Abrams1977-zp; @Chesson2000-aw].
+Bumble bees are exceptionally versatile foragers whose ability to solve the puzzles posed by complex flower morphologies [Plowright1984-lc], or indeed by human investigators [@Alem2016-rn; @Loukola2017-bt], renders almost any flower potentially accessible.
+Nevertheless, the flexibility of bumble bee foraging is tuned to floral reward, which depends on both the intrinsic reward of a given flower and the costs associated with exploiting it.
+These costs are closely related to the morphological fittedness between bumble bee species and their floral hosts.
+Bumble bees and flowers vary in an array of interrelated morphological parameters, both within and between species, but for bumble bees, morphological variation consists principally (though not exclusively) of variation in tongue length.
+Floral handling time depends on the degree of match between bumble bee tongue length and floral corolla depth [@Inouye1980-mv], and there is a well-established (though imperfect) association between tongue length and floral choice.
+Importantly, though, accessing floral reward involves more than just a match between tongue length and corolla depth, and in some flowers access to floral rewards involves learning to manipulate floral structures that are challenging irrespective of tongue length (Laverty 1994).
+For example, the characteristic Papilionaceous floral morphology of legumes requires a bee to separate tightly appressed keel petals to access anthers and nectaries (Westerkamp 1997), and the narrowly campanulate flowers of many Ericaceous flora pose a significant barrier to large-bodied visitors like bumble bees (Wallace 1977).
+Thus the dietary niche of a given bumble bee species in a given ecological context is an emergent product of innate floral biases, rewards experienced during foraging, and learned flower handling skills, and all of which contribute to the phenomenon of interspecific niche overlap upon which competition is predicated.
 
-In the present study, we focus on the question of bumble bee foraging bias with respect to floral morphology, and the lability of morphological biases across gradients of floristic variation.
-At the outset, we emphasize the use of the term “bias” in favor of the term “preference,” since the latter is laden with more assumptions and its inference under field conditions is problematic (Williams 2005).
-By “bias,” we mean simply an empirical disproportionality between the relative abundance of a given floral taxon or morphological class and its proportional visitation.
-The exact mechanism of this disproportionality --- whether an arbitrary “preference” on the part of the foraging bee, an unmeasured decoupling of relative abundance and rate of reward, an exogenous effect of competition, etc.
---- is beyond the scope of our discussion and probably best studied under laboratory conditions where strict factorial control is possible.
-We focus instead on the empirical phenomenon of bias, its variation, and its morphological correlations.
+Empirical measurements of niche overlap, however, are inherently equivocal in their interpretation with respect to competition [@Pianka1972-oi; @Pianka1974-qk; @Abrams1980-am].
+Low niche overlap between two species, for example, could be interpreted prospectively as an indication that the *potential* for interspecific competition is low, but the same pattern could also be interpreted retrospectively as a mitigating response to high interspecific competition.
+Which of these interpretations is most plausible hinges on how closely species' observed niches approximate their fundamental niches (within a given resource context).
+Where observed niches are close to absolute niches, low niche overlap is an indication of low potential for interspecific competition.
+Any difference between observed and fundamental niches, though, might be attributable to competition-driven niche partitioning, and niche overlap as a reflection of potential competition cannot be distinguished from niche overlap as a response to realized competition.
+In other words, niche overlap can only be intepreted as an index of *potential* competition if *realized* competition is assumed to be negligible, and this assumption can only be justified on the basis of additional knowledge of the study system under consideration.
 
-The meaning of “morphology,” however, is hardly more clear-cut than that of bias.
-Bumble bees and flowers vary in an array of interrelated morphological parameters, both within and between species, and the selection of relevant and measurable morphological metrics is nontrivial.
-For bumble bees, morphological variation consists principally (though not exclusively) of variation in tongue length.
-Tongue length in bumble bees corresponds functionally to corolla depth in flowers, and there is a well-established (though imperfect) association between tongue length and floral choice (e.g.
-Inouye 1980).
-Importantly, though, accessing floral reward involves more than just a match between tongue length and corolla depth.
-Such matching is entirely bypassed by nectar robbers, but even in the case of “legitimate” foraging, access to reward involves learning to manipulate floral structures that, in some floral morphotypes, are challenging irrespective of tongue length (Laverty 1994).
-For example, the characteristic Papilionaceous floral morphology of legumes requires a bee to separate tightly appressed keel petals to access anthers and nectaries (Westerkamp 1997), and the narrowly campanulate flowers of many Ericaceous flora pose a significant barrier to large-bodied visitors like bumble bees (Wallace 1977).
 Rather than considering various floral morphological traits independently, we rely on the synthesis of Kugler (1970), who developed a typology consisting of ten primary morphotypes and nested subdivision of each.
@@
 
-<!-- Formulate this section more as a list of motivating questions; perhaps end with something that ties back into the more general introduction. -->
+<!-- 
+That one paper [@Zhongming2020-xj] simply equated diet overlap with competition, and I think that is a mistake.
+But I think we can make a more nuanced case for its use as an index of potential competition.
+Indeed, Ingolf interpreted diet overlap in exactly this sense in one of his early papers [@Steffan-Dewenter2000-ll].
+In fact, there is a slew of papers, from across a wide range of study systems and organisms, that use the term "potential competition" to interpret diet overlap.
+I think I'm on solid ground.
+-->
 
-How do floral morphotypes compare in terms of total bumble bee visitation?
+Importantly, our study system is one of rich floral resources and harsh abiotic conditions, and it is likely that competition is low, or at least transient, and that bumble bees are limited instead by abiotic conditions during the critical period of spring colony founding.
+Thus, the patterns in our data likely represent floral choice unconstrained by competition and therefore approximate the absolute dietary niches of the species involved (within the context of the studied floral community).
+On this basis, we argue for the interpretation of niche overlap as an index of *potential* competition rather than a response to *realized* competition, with the caveat this interpretation may misrepresent certain temporal intervals (e.g.
+early spring colony founding) or elevational bands (e.g.
+the sparse alpine grasslands above the tree line ecotone, which in our system falls around 1500 m).
 
-How does morphological bias vary across bumble bee species and tongue length classes?
-
-Species' morphological biases give rise to patterns of diet (dis)similarity
-
-How are the floral morphotypes distributed through elevation and time in our mountain ecosystem?
-
-For ubiquitously distributed bumble bee species, how does morphological bias vary through elevation and time?
-
-First, we compare the overall rate of bumble bee visitation across floral morphotypes.
-Then, we parse visitation rates by bumble bee species and explore the functional relationship between floral morphotype and bumble bee tongue length.
-These biases, together with the background effects of relative abundance, generate emergent patterns of diet (dis)similarity, which we present in the form of ordination analysis.
-Lastly, we analyze the lability of the morphological biases of individual bumble bee species through elevation and time, and compare these patterns of bias across bumble bee species and within floral morphotypes.
-The results of these analyses demonstrate both the strength of bumble bee floral bias, arising in part from morphological constraints, and the flexibility of bumble bee floral bias that enables them to forage adaptively in the dynamic floral communities of mountain landscapes.
+We first ask how floral biases compare across bumble bee species whether these patterns can be explained by implicit trait-matching between bumble bee tongue length and floral morphotype.
+We then investigate the patterns of dietary niche overlap between bumble bee species that emerge from their respective floral biases, comparing niche overlap at the nested levels of floral morphotype, floral family, and floral genus.
+Finally, we ask how niche overlap varies with elevation.
+With respect to elevation, we analyze both pairwise niche overlap and community mean niche overlap, since low pairwise niche overlap could conceal a high potential for diffuse competition [@Pianka1974-qk; MacArthur1972-qz].
+We discuss the implications of these findings for bumble bee conservation in a changing climate
 
@@
 
-### Model 5: The variation of morphological bias through elevation and time
+### Model 5: Niche overlap by level
 
-To explore the plasticity of floral biases, we first subsetted our data to include only the dominant bumble bee species (B.
-hortorum, B.
-pascuorum, B.
-pratorum, B.
-soroeensis, B.
-terrestris/lucorum, and B.
-wurflenii) and the most visited floral morphotypes (groups 3, 5, 6, and 7).
-We then constructed a generalized additive model (GAM) in which binary visitation was modeled as a tensor-product smooth of elevation and day-of-year, grouped by floral morphotype.
-This response was conditioned on linear interaction terms for bumble bee abundance : percent flower cover and percent flower cover : floral morphotype and a random intercept term for bumble bee species : floral morphotype.
-See Appendix X for the details of model specification and evaluation.
+### Model 6: Genus-level overlap by species-pair
+
+### Model 7: Community-mean genus-level niche overlap by elevation
+
+### Model 8: Pairwise genus-level niche overlap by elevation
 
@@
 GAMs were implemented with the package `mgcv` [@Wood_2017] and visualized with the package `mgcViz` [@Fasiolo2018aa].
-Data handling and visualization was performed with the `tidyverse` suite of R packages `r citep(citation("gdm"))`.
+Data handling and visualization was performed with the `tidyverse` suite of R packages `r citep(citation("tidyverse"))`.
 Annotated R code is available in the Supplementary Material.
@@
 
-## Diet similarity
+## Diet overlap
 
@@
 
-## The lability of bias through elevation and time
+To account for this inconsistency, @Ranta1981-vx proposed that bumble bees communities are governed by a spatiotemporally heterogeneous mosaic of competitive conditions, such that competitive inequalities between species are sufficiently variable to allow more species to coexist than predicted by trait-based competition models.
+A complementary explanation, based on the core/satellite species hypothesis [@Hanski1982-hs], is that competitive forces are significant only among regionally abundant and widespread “core species”, and the effects of competition are more pronounced in the structure of regional species pools than in that of local assemblages [@Hanski1982-ie].
+In some cases, competition may simply be precluded by a saturating abundance of floral resources [@Williams1989-vr].
+Range shifts in bumble bees have already been documented [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc], and the introduction or extinction of species from bumble bee communities, with corresponding addition and release of competitive pressure, can alter both the floral visitation patterns of bumble bees and, consequently, the pollination biology of their host plants [@Brosi2013-mu; @Ishii2013-sm].
 
-While the intraspecific lability of bumble bee foraging biases has been demonstrated under laboratory conditions (Ings, Raine, and Chittka 2009; Maharaj et al.
-2019) and in a localized field study (Raine and Chittka 2007), ours is the first study, to our knowledge, to characterize the in situ lability of foraging biases within multiple species through elevational and temporal gradients of floristic turnover.
-In our analysis, it is evident that morphological biases in bumble bees, insofar as they are captured by the morphotypology used in our study, are not fixed but rather tuned to floristic (and possibly abiotic) context and therefore dynamic through elevation and time.
-Notably, bumble bee species differed from one another not just in their overall bias toward given morphotypes but also in when and where they were biased toward a given morphotype.
-This is especially evident in the case of flag flowers, for which positive bias was maximized in mid-season for B.
-hortorum, B.
-pascuorum, B.
-terrestris/lucorum, and B.
-wurflenii but in early-season for B.
-pratorum and B.
-soroeensis.
-This pattern corresponds to the overall biases of these species toward flag flowers, which were primary resources for B.
-hortorum, B.
-pascuorum, B.
-terrestris/lucorum, and B.
-wurflenii but only marginal resources for B.
-pratorum and B.
-soroeensis.
+We see that community-mean diet overlap, indicative of the potential for diffuse competition, generally declines with elevation [@Pianka1974-qk].
+
+"there  is  still  considerable  scope  for  niche  differentiation  among  species  with  tonguesthat  are  near-identical  in  length." @Goulson2008-zy
 

### 2021-03-05 code/bias_niche.Rmd
update

@@
 summary(brm_pvis02)
-summary(brm_pvis11)
-summary(brm_pvis21)
 
@@
 pp_check(brm_pvis02)
-pp_check(brm_pvis11)
-pp_check(brm_pvis21)
 
@@
 bayes_R2(brm_pvis02)
-bayes_R2(brm_pvis11)
-bayes_R2(brm_pvis21)
-
 ```
@@
   geom_pointinterval() +
-  theme_light(14) +
+  theme_light(16) +
   labs(x = "Visitation probability", y= "Morphotype") +
@@
 
-ggsave(file = "../output/brm_pvis01_plot1.pdf", width = 5, height = 5)
+ggsave(file = "../output/brm_pvis01_plot1.pdf", width = 6, height = 7)
 
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+bb_distribution_pool <- visitation %>%
+  left_join(site_data) %>%
+  group_by(site, date, elev.mean, bb.sp) %>%
+  summarize(abund = n()) %>%
+  ungroup() %>%
+  complete(site, date, elev.mean, bb.sp, fill = list(abund = 0)) %>%
+  group_by(site, elev.mean, bb.sp) %>%
+  summarize(abund = mean(abund)) %>%
+  filter(abund > 0) %>%
+  group_by(bb.sp) %>%
+  mutate(elev.floor = min(elev.mean),
+         elev.ceiling = max(elev.mean),
+         elev.range = elev.ceiling - elev.floor,
+         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))
+
+ggplot(bb_distribution_pool, aes(reorder(bb.sp, elev.med), elev.mean)) +
+  geom_line(size = 2, alpha = 0.25) +
+  geom_point(aes(size = abund), alpha = 0.5) +
+  theme_light(14) +
+  scale_size_continuous(name = "Abundance") +
+  xlab("BB species") +
+  ylab("Elevation") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+ggsave("../output/bb_range.pdf", width = 6, height = 5)
 ```
@@
 mcmc_plot(brm_niche00, pars = c("^r_site*"))
+
+brm_niche00_plotframe1 <- brm_niche00 %>%
+  spread_draws(b_Intercept, b_levelgenus, b_levelmorphotype) %>%
+  mutate(b_levelgenus = b_Intercept + b_levelgenus,
+         b_levelmorphotype = b_Intercept + b_levelmorphotype) %>%
+  pivot_longer(-c(.chain, .iteration, .draw), names_to = "level", values_to = "value") %>%
+  group_by(level) %>%
+  point_interval(value, .width = c(0.95, 0.66)) %>% # calculate credible intervals
+  mutate(value_trans = inv_logit_scaled(value), # exponentiating the estimates gets us back to the response scale (negbinomial uses log link function)
+         .lower_trans = inv_logit_scaled(.lower),
+         .upper_trans = inv_logit_scaled(.upper)) %>%
+  mutate(level = case_when(
+    level == "b_Intercept" ~ "family",
+    level == "b_levelmorphotype" ~ "morphotype",
+    level == "b_levelgenus" ~ "genus"
+  )) %>%
+  mutate(level = factor(level, levels = c("genus", "family", "morphotype")))
+
+ggplot(brm_niche00_plotframe1, aes(value_trans, level, 
+                                 xmin = .lower_trans, xmax = .upper_trans)) +
+  geom_pointinterval() +
+  theme_light(16) +
+  labs(x = "Niche overlap", y= "Level") +
+  guides(fill = FALSE)
+
+ggplot(brm_niche00_plotframe1, aes(level, value_trans, 
+                                 ymin = .lower_trans, ymax = .upper_trans)) +
+  geom_pointinterval() +
+  theme_light(16) +
+  labs(x = "Level", y= "Niche overlap") +
+  guides(fill = FALSE) +
+  ylim(c(0, 0.55))
+
+#ggsave(file = "../output/brm_pvis01_plot1.pdf", width = 6, height = 7)
 ```

### 2021-03-05 ms_2/ms_2_v2.Rmd
update

@@
 It has taken on new urgency, however, on account of recent climate change, which threatens to disrupt existing patterns of species interactions [@Tylianakis2008-fi], including competition [@Ettinger2017-pk], with consequences for community composition and function.
-These effects are likely to be most pronounced along the steep climate gradients of mountain slopes, where climate-induced range shifts are expected to introduce novel competitive interactions within trophic levels and cause spatial or phenological mismatches between organisms and their food resources.
+These effects are likely to be most pronounced along the steep climate gradients of mountain slopes, where rapid climate-induced range shifts are expected to introduce novel competitive interactions within trophic levels and cause spatial or phenological mismatches between organisms and their food resources.
 
 Bumble bees (Hymenoptera: *Bombus* spp.) have become classic model organisms for the study of competition and species coexistence due to their broad and overlapping distributions and their gross similarity in morphology, behavior, and dietary niche [@Plowright1984-lc; @Goulson2008-zy].
-In this role, however, bumble bees have proven somewhat enigmatic.
+In this role, however, they have proven somewhat enigmatic.
 On the one hand, there is evidence that resource limitation can be a significant constraint on local bumble communities [@Heinrich1976-wc; @Bowers1985-df], and certain bumble bee communities have been shown to be over-dispersed in terms of tongue length (a matching trait corresponding to floral corolla depth) or phylogeny (understood to encompass unmeasured trait variation), providing evidence for competition as a community assembly process [@Inouye1977-kl; @Pellissier2013-pk].
 Nevertheless, bumble bee communities with an apparent excess of species richness and deficit of trait (or phylogenetic) dispersion are commonly observed [@Ranta1981-vx; @Goulson2008-zy].
-Teasing apart the still-obscure mechanisms of bumble bee coexistence and niche partitioning remains a salient problem in basic ecology, but it has also become an urgent question in the context of conservation.
+Teasing apart the still-obscure mechanisms of bumble bee coexistence and niche partitioning remains a salient problem from a basic ecological perspective, but it has also become an urgent question in the context of applied conservation.
 Bumble bees are cold-adapted species [@Heinrich1994-cv] associated mainly with mountain habitats, where they serve as the principal pollinators of many alpine plants [@Williams1998-xi; @Goulson2010-lo].
-Warming climates have already driven measurable elevational range shifts in bumble bees and their floral hosts [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc], and these are predicted to cause declines or even extinction in many bumble bees species [@Soroye2020-zb].
-One mechanism by which climate change is expected to affect bumble bees is via the introduction of novel interspecific competition driven by the advance of lowland bumble bee species into the elevational distributions of their highland counterparts.
-Thus, the speculative question of how competition has acted in the historic assembly of bumble bee communities has become an unfolding process observable over ecological time scales.
+Warming climates have already caused measurable upslope shifts in bumble bees and their floral hosts [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc], and these are predicted to cause declines or even extinction in many bumble bees species [@Soroye2020-zb].
+One mechanism by which climate change is expected to affect bumble bees is the introduction of novel interspecific competition driven by the upslope advance of lowland bumble bee species into the elevational distributions of their highland counterparts.
+Thus, the speculative question of how competition has acted in the historic assembly of bumble bee communities has become an unfolding process observable over ecological time scales and instrumental in the future viability of bumble bee species and their floral mutualists.
 
-All else held equal, species coexistence is favored when intraspecific competition exceeds interspecific competition, and this stabilizing inequality is favored to the extent that species differ in fundamental dietary niche [@Abrams1977-zp; @Chesson2000-aw].
+All else held equal, species coexistence is favored when intraspecific competition exceeds interspecific competition, and this stabilizing inequality is favored to the extent that species differ in dietary niche [@Abrams1977-zp; @Chesson2000-aw].
 Bumble bees are exceptionally versatile foragers whose ability to solve the puzzles posed by complex flower morphologies [Plowright1984-lc], or indeed by human investigators [@Alem2016-rn; @Loukola2017-bt], renders almost any flower potentially accessible.
 Nevertheless, the flexibility of bumble bee foraging is tuned to floral reward, which depends on both the intrinsic reward of a given flower and the costs associated with exploiting it.
-These costs are closely related to the morphological fittedness between bumble bee species and their floral hosts.
-Bumble bees and flowers vary in an array of interrelated morphological parameters, both within and between species, but for bumble bees, morphological variation consists principally (though not exclusively) of variation in tongue length.
-Floral handling time depends on the degree of match between bumble bee tongue length and floral corolla depth [@Inouye1980-mv], and there is a well-established (though imperfect) association between tongue length and floral choice.
-Importantly, though, accessing floral reward involves more than just a match between tongue length and corolla depth, and in some flowers access to floral rewards involves learning to manipulate floral structures that are challenging irrespective of tongue length (Laverty 1994).
-For example, the characteristic Papilionaceous floral morphology of legumes requires a bee to separate tightly appressed keel petals to access anthers and nectaries (Westerkamp 1997), and the narrowly campanulate flowers of many Ericaceous flora pose a significant barrier to large-bodied visitors like bumble bees (Wallace 1977).
-Thus the dietary niche of a given bumble bee species in a given ecological context is an emergent product of innate floral biases, rewards experienced during foraging, and learned flower handling skills, and all of which contribute to the phenomenon of interspecific niche overlap upon which competition is predicated.
+These costs are closely related to the morphological fittedness between bumble bees and their floral hosts, which consists principally in the compatibility between bumble bee tongue length and floral corolla depth.
+Bumble bee species can differ markedly in tongue length, and floral handling time increases both for short-tongued bumble bees foraging on deep flowers and long-tongued bees foraging on shallow flowers [@Inouye1980-mv], which leads to an association between tongue length and floral niche in bumble bee communities [e.g.
+@Inouye1978-qi; @Inouye1980-mv].
+This association between tongue length and floral morphology, though, is imperfect [e.g.
+@Ranta1980-ok; @Goulson2008-oe], and access to floral rewards in some flowers involves learning to manipulate floral structures that are challenging irrespective of tongue length [@Laverty1994-hm].
+For example, the characteristic Papilionaceous floral morphology of legumes requires a bee to separate tightly appressed keel petals to access anthers and nectaries [@Westerkamp1997-hs], and the narrowly campanulate flowers of many Ericaceous flora pose a significant barrier to large-bodied visitors like bumble bees [@Wallace1977-kz].
+Thus the dietary niches of bumble bee species in a given ecological context are emergent products of innate floral biases, rewards experienced during foraging, and learned flower handling skills, all of which contribute to the phenomenon of interspecific niche overlap upon which interspecific competition is predicated.
 
-Empirical measurements of niche overlap, however, are inherently equivocal in their interpretation with respect to competition [@Pianka1972-oi; @Pianka1974-qk; @Abrams1980-am].
-Low niche overlap between two species, for example, could be interpreted prospectively as an indication that the *potential* for interspecific competition is low, but the same pattern could also be interpreted retrospectively as a mitigating response to high interspecific competition.
-Which of these interpretations is most plausible hinges on how closely species' observed niches approximate their fundamental niches (within a given resource context).
-Where observed niches are close to absolute niches, low niche overlap is an indication of low potential for interspecific competition.
-Any difference between observed and fundamental niches, though, might be attributable to competition-driven niche partitioning, and niche overlap as a reflection of potential competition cannot be distinguished from niche overlap as a response to realized competition.
-In other words, niche overlap can only be intepreted as an index of *potential* competition if *realized* competition is assumed to be negligible, and this assumption can only be justified on the basis of additional knowledge of the study system under consideration.
+<!-- I expect my co-authors to want the section below to get pushed to the discussion. I disagree, but I think we could make it work if they insist. -->
 
-Rather than considering various floral morphological traits independently, we rely on the synthesis of Kugler (1970), who developed a typology consisting of ten primary morphotypes and nested subdivision of each.
-Kugler’s typology was informed by laboratory studies of bumble bee flower choice (Kugler 1943), and there is a precedent for analyzing bumble bee visitation in terms of Kugler’s morphotypes (Neumayer and Paulus 1999; Schneller et al.
-2014).
+Empirical measurements of niche overlap are, however, inherently ambiguous with respect to competition.
+It has been well-demonstrated that niche overlap can never serve as a metric of competition *per se* [@Abrams1980-am], but it is often used as an index of *potential* competition, particularly in the context of dietary studies [e.g.
+@French1985-eo; @Steffan-Dewenter2000-ll; @Jung2015-cw], based on the reasoning that if and when food becomes limiting, competition can be expected to vary in proportion to the dietary overlap of co-occurring species [@Pianka1972-oi].
+Focusing on *potential* competition does not, however, escape the equivocity of niche overlap, since an observed pattern of niche overlap could just as plausibly be interpreted as adaptive niche partitioning in response to *realized* competition [@Pianka1974-qk].
+This creates a predicament in which niche overlap can only be interpreted as an index of potential competition when realized competition --- which cannot be inferred from niche overlap --- is assumed to be negligible, and this assumption can only be justified on the basis of additional knowledge of the study system under consideration.
+Our study is set in the Berchtesgadener Alps of southeast Germany, a landscape of flower-rich mountain meadows --- generated by past or ongoing grazing management --- embedded in a matrix of coniferous forest.
+This system is one of rich floral resources but harsh abiotic conditions, so it is likely that competition for floral resources is low, or at least transient, and that bumble bees are limited instead by abiotic conditions during the critical period of spring colony founding.
+On this basis, we argue for the interpretation of niche overlap as an index of *potential* competition rather than a response to *realized* competition, with the caveats this interpretation may misrepresent certain temporal intervals (e.g.
+early spring colony founding) or elevational bands (e.g.
+the sparse alpine grasslands above the tree line ecotone) [Sponsler et al., submitted] and that the residual effects of past competitive pressure on innate floral biases [@Connell1980-dn] are beyond the scope of our inference.
 
-<!-- 
-That one paper [@Zhongming2020-xj] simply equated diet overlap with competition, and I think that is a mistake.
-But I think we can make a more nuanced case for its use as an index of potential competition.
-Indeed, Ingolf interpreted diet overlap in exactly this sense in one of his early papers [@Steffan-Dewenter2000-ll].
-In fact, there is a slew of papers, from across a wide range of study systems and organisms, that use the term "potential competition" to interpret diet overlap.
-I think I'm on solid ground.
--->
-
-Importantly, our study system is one of rich floral resources and harsh abiotic conditions, and it is likely that competition is low, or at least transient, and that bumble bees are limited instead by abiotic conditions during the critical period of spring colony founding.
-Thus, the patterns in our data likely represent floral choice unconstrained by competition and therefore approximate the absolute dietary niches of the species involved (within the context of the studied floral community).
-On this basis, we argue for the interpretation of niche overlap as an index of *potential* competition rather than a response to *realized* competition, with the caveat this interpretation may misrepresent certain temporal intervals (e.g.
-early spring colony founding) or elevational bands (e.g.
-the sparse alpine grasslands above the tree line ecotone, which in our system falls around 1500 m).
-
-We first ask how floral biases compare across bumble bee species whether these patterns can be explained by implicit trait-matching between bumble bee tongue length and floral morphotype.
+The aim of our study is to synthesize the phenomena of floral bias, dietary niche overlap, and interspecific competition in a way that accounts for the coexistence of bumble bee species with similar tongue lengths and predicts the effects of climate-driven elevational range shifts on the future of potential competition between bumble bee species in mountain habitats.
+We first ask how floral biases compare across bumble bee species and to what extent these patterns can be explained by trait-matching between bumble bee tongue length and floral morphotype.
 We then investigate the patterns of dietary niche overlap between bumble bee species that emerge from their respective floral biases, comparing niche overlap at the nested levels of floral morphotype, floral family, and floral genus.
-Finally, we ask how niche overlap varies with elevation.
-With respect to elevation, we analyze both pairwise niche overlap and community mean niche overlap, since low pairwise niche overlap could conceal a high potential for diffuse competition [@Pianka1974-qk; MacArthur1972-qz].
-We discuss the implications of these findings for bumble bee conservation in a changing climate
+Finally, we ask how niche overlap varies with elevation and discuss the implications of this variation for bumble bee conservation in a changing climate.
 
@@
 Fourteen of these pastures are extensively grazed by cattle or sheep, three are mowed for hay production, and eight have been abandoned throughout the last century and are no longer subject to any human management.
-The sites sampled in this study are the same as those used in a series of previous studies [@Hoiss2012-sn; @Hoiss2012-kr; @Hoiss2015-gv], but the data reported here are otherwise independent.
 
@@
 
-## Data analysis
+## Data sources
 
-### Data processing
+### Floral morphology
+
+Rather than considering various floral morphological traits independently, we rely on the synthesis of @Kugler1970-gb, who developed a typology consisting of ten primary morphotypes and nested subdivision of each.
+Kugler’s typology was informed by laboratory studies of bumble bee flower choice [@Kugler1943-uy], and there is a precedent for analyzing bumble bee visitation in terms of Kugler’s morphotypes [@Neumayer1999-il; @Schneller2014-ry].
+
+Kugler morphotype classifications of the plants in our study system were accessed via the BIOFLOR database (Klotz, Kühn, and Durka 2002).
+Morphotypes were simplified to the primary classes numbered 0-10: 0 = not applicable (includes wind-pollinated flora such as grasses), 1 = disc- and bowl-shaped flowers, 2 = funnel flowers, 3 = bell-shaped flowers, 4 = stalk-disc flowers, 5 = lip flowers, 6 = flag flowers, 7 = flower heads, 8 = spike flowers, 9 = brush flowers, 10 = trap flowers **(Figure 1)**.
+In a few cases, a plant species in our data set was not found in the BIOFLOR database.
+In most cases, morphotypes are constant at the genus level, and missing morphotypes were added based on the morphotypes found for congeners.
+The genus *Gentiana* consisted of both group 2 and group 4 representatives, so the missing species *Gentiana aspera* and *Gentiana ciliata* were classified based on visual comparison with their classified congeners.
+In the case of the genus *Phyteuma*, we followed Neumayer and Paulus (1999) and classified it as a group 7 flower head.
+Morphotypes 9 and 10 were extremely rare in our study system and were therefore omitted from all analyses.
+Plants belonging to morphotype 0 were typically not recorded during floral surveys due to their lack of conspicuous floral displays, so this morphotype was also omitted from all analyses.
+
+### Bumble bee tongue length
 
@@
 
-Floral morphology was summarized using floral morphotypes of Kugler (1970), accessed via the BIOFLOR database (Klotz, Kühn, and Durka 2002).
-Morphotypes were simplified to the primary classes numbered 0-10: 0 = not applicable (includes wind-pollinated flora such as grasses), 1 = disc- and bowl-shaped flowers, 2 = funnel flowers, 3 = bell-shaped flowers, 4 = stalk-disc flowers, 5 = lip flowers, 6 = flag flowers, 7 = flower heads, 8 = spike flowers, 9 = brush flowers, 10 = trap flowers **(Figure 1)**.
-In a few cases, a plant species in our data set was not found in the BIOFLOR database.
-In most cases, morphotypes are constant at the genus level, and missing morphotypes were added based on the morphotypes found for congeners.
-The genus *Gentiana* consisted of both group 2 and group 4 representatives, so the missing species *Gentiana aspera* and *Gentiana ciliata* were classified based on visual comparison with their classified congeners.
-In the case of the genus *Phyteuma*, we followed Neumayer and Paulus (1999) and classified it as a group 7 flower head.
-Morphotypes 9 and 10 were extremely rare in our study system and were therefore omitted from all analyses.
-Plants belonging to morphotype 0 were typically not recorded during floral surveys due to their lack of conspicuous floral displays, so this morphotype was also omitted from all analyses.
+## Data analysis
+
+### Data processing
 
@@
 While our study spanned three years, we chose not to model year effects.
-Differences in sampling intensity and duration across years, for the reasons described above, make comparisons across years difficult to interpret, and we deemed these comparisons not worth the cost of added model complexity.
-
-<!-- Should I be including year as a varying effect in these models? -->
+Differences in sampling intensity and duration across years, for the reasons described above, would make comparisons across years difficult to interpret, and we deemed these comparisons not worth the cost of added model complexity.
+This decision entails the caveats that our analysis cannot detect year-to-year variation in floral biases and niche overlap, and main effects cannot be distinguished from annual anomalies.
+With only three years of data, though, such annual inferences would be of little value.
 
@@
 
-<!-- Consider adding year as a varying effect -->
-
 <!-- Figure 3 order by tongue length class -->
@@
 
-<!-- Year should really be included in Model 5, or just use 2012. -->
+<!-- Open each results paragraph with a summary sentence. -->
 
-<!-- Open each results paragraph with a summary sentence. -->
+### Defining bias
+
+We opt to use of the term “bias” in favor of the term “preference,” since the latter is laden with more assumptions and its inference under field conditions is problematic [@Williams2005-yp].
+By “bias,” we mean simply an empirical disproportionality between the relative abundance of a given floral taxon or group and its proportional visitation.
+The exact mechanism of this disproportionality — whether an arbitrary “preference” on the part of the foraging bee, an unmeasured decoupling of relative abundance and rate of reward, etc.
+— is beyond the scope of our discussion, and probably best studied under laboratory conditions where strict factorial control is possible.
+We focus instead on the empirical phenomenon of bias, its variation, and its correlation to morphology.
 
@@
 
+We analyze both pairwise niche overlap and community mean niche overlap, since low pairwise niche overlap could conceal a high potential for diffuse competition [@Pianka1974-qk; MacArthur1972-qz]
+
 ### Model 8: Pairwise genus-level niche overlap by elevation
@@
 terrestris-lucorum*, and *B.
-wurflenii* --- accounted for the bulk of overall bumble bee abundance and were present over the entire elevation range of our sites **(Figure 1)**.
+wurflenii* --- accounted for the bulk of overall bumble bee abundance and were present over the entire elevation range of our sites.
 *B.
@@
 
-```{r fig1, echo = FALSE,  fig.cap = paste("Each floral morphotype is exemplified by a representative species common in our study system.
-Illustrations are modifications of public domain works obtained from www.plantillustrations.org.
-For details of original sources, see Appendix X.")}
-knitr::include_graphics("../output/ktype_abund_gam_mod.png")
-```
-
 ## Overall visitation probability by floral morphotype
@@
 
-```{r fig2, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability by floral morphotype. Point-interval plots depict 66\\% and 95\\% CIs.")}
-knitr::include_graphics("../output/brm_pvis01_plot1.pdf")
+```{r fig1, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability by floral morphotype.
+Point-interval plots depict 66\\% and 95\\% CIs.
+Each floral morphotype is exemplified by a representative species common in our study system.
+Illustrations are modifications of public domain works obtained from www.plantillustrations.org.
+For details of original sources, see Appendix X.")}
+knitr::include_graphics("../output/brm_pvis01_plot1_mod.png")
 ```
@@
 Lip (group 5) flowers were also visited at similar rates across tongue length classes, but at the species level B.
-pascuorum exhibited an uniquely high preference for lip flowers.
+pascuorum exhibited an uniquely high bias toward lip flowers.
 Flag (group 6) flowers were visited at a higher rate by medium-tongued species than by short-tongued species, with long-tongued species exhibiting intermediate visitation rates.
@@
 knitr::include_graphics("../output/pcoa_aggregate.png")
-```
-
-## The variation of morphological bias through elevation and time
-
-Model 5 explained 33.8% of total variance (adjusted R2 = 0.37).
-All tensor product smooths (Figure 5) were significant at p < 0.001 except for the one representing B.
-pratorum and flag (group 6), which was more weakly significant at p = 0.02.
-As in Models 1-3, the results of Model 5 should be interpreted in terms of conditional visitation probability.
-
-Bell (group 3) flowers were visited at highest rates early in the season and at low elevation by all species except B.
-hortorum, whose visitation rate to bell flowers peaked in mid-season and at mid-elevation.
-Notably, the medium-tongued B.
-pascuorum and B.
-wurflenii also exhibited relatively high visitation to bell flowers at mid-season and mid-elevation, while the visitation rates of the short-tongued B.
-pratorum, B.
-soroeensis.
-and B.
-terrestris lucorum were more concentrated at the early-season and low-elevation peak.
-
-Lip (group 5) flowers exhibited roughly the mirror image of the pattern of visitation rate seen in bell flowers, with peak visitation late in the season and at low elevation, but relatively high visitation extending toward mid-season and mid-elevation.
-This pattern was consistent for all bumble bee species.
-
-The pattern of visitation rates to flag (group 6) flowers fell generally along the lines of tongue length class.
-The short-tongued B.
-pratorum and B.
-soroeensis exhibited peak visitation early in the season and at low elevation.
-The long-tongued B.
-hortorum and the medium-tongued B.
-pascuorum and B.
-wurflenii exhibited peak visitation at mid-season, though B.
-pascuorum exhibited a bimodal pattern in elevation, with peaks around 1000 m and 1750 m, while the visitation rate of B.
-hortorum declined monotonically from high elevation to low.
-The short-tongued B.
-terrestis/lucorum broke the tongue-length pattern, exhibiting a diffuse pattern of visitation rate more similar to that of the long- and medium-tongued species than the other short-tongued species.
-
-Visitation rates for head (group 7) flowers were more or less bimodal for all species, with a first peak in mid-season and low elevation and a second peak in late-season and high elevation.
-
-```{r fig6, echo = FALSE,  fig.cap = paste("GAM analysis of bias as a function of the tensor-product interaction of elevation and day of year, parsed by bumble bee species and floral morphotype and conditioned on the abundance of each bumble bee species and the relative abundance of each floral morphotype.
-The color ramp of each smooth has been scaled to equal color range, thereby obscuring the absolute magnitude of effect.
-To avoid cluttering the figure, we have opted not to include a unique color ramp key for each plot, since the focus of our inference is on relative rather than absolute patterns.")}
-knitr::include_graphics("../output/ktype_lability_gam_mod.png")
 ```
@@
 
-For bumble bees, floral bias is a labile product of innate preferences, the soft constraints of morphological compatibility, and the dynamic ecological context in which foraging occurs.
+For bumble bees, floral bias is a labile product of innate biases, the soft constraints of morphological compatibility, and the dynamic ecological context in which foraging occurs.
 By conditioning bumble bee visitation data upon concurrent floral abundance data across spatially and temporally replicated samples, we demonstrate the strength bumble bee floral bias, its functional underpinnings, its variation across species, and its lability within species in response to environmental variation.

### 2021-03-08 code/bias_niche.Rmd
improved niche overlap analysis

@@
 library(ggrepel)
-library(treemapify)
+library(tidymv)
 library(patchwork)
@@
 ## 1.3 Evaluate models
-```{r}
+```{r eval=FALSE, include=FALSE}
 # Summary
@@
 ## 1.4 Preliminary visualization
-```{r}
+```{r eval=FALSE, include=FALSE}
 get_variables(brm_pvis00)
@@
 ## 1.5 Final visualization
-Note that the intercept is on the log scale, so if you want to transform it back to predicted # visits, you have to exponentiate it.
-I think.
-Here is a helplful conversation involving the esteemed Ben Bolker:
 ### 1.5.1 morphotype:tongue intercept
@@
 ### 1.5.3 morphotype:tongue slope
-```{r}
+```{r eval=FALSE, include=FALSE}
 brm_pvis00_plotframe3 <- brm_pvis00 %>%
@@
 ### 1.5.3 morphotype slope
-```{r}
+```{r eval=FALSE, include=FALSE}
 brm_pvis01_plotframe2 <- brm_pvis01 %>%
@@
 ## 1.6 Model 1 hypothesis tests
-I'm not doing every possible comparison. Just the ones where it is possible to have some doubt. It would be stupid to compare group whose 95% CIs are miles apart.
 ```{r}
 get_variables(brm_pvis01)
-
-hypothesis(brm_pvis01, "r_ktype[0,Intercept] = r_ktype[1,Intercept]", class = "") # this is lower than its closest neighbor; no need to make any other comparisons
 
@@
 hypothesis(brm_pvis00, "r_ktype:tongue[7_med,Intercept] = r_ktype:tongue[7_long,Intercept]", class = "")
-
-# NA (group  0)
-hypothesis(brm_pvis00, "r_ktype:tongue[0_short,Intercept] = r_ktype:tongue[0_med,Intercept]", class = "")
-hypothesis(brm_pvis00, "r_ktype:tongue[0_short,Intercept] = r_ktype:tongue[0_long,Intercept]", class = "")
-hypothesis(brm_pvis00, "r_ktype:tongue[0_med,Intercept] = r_ktype:tongue[0_long,Intercept]", class = "")
 ```
@@
 plot(pcoa_horn2_betadisper)
-# I'm suprised that the betadisper test is not significant, but I think it might be driven by the fact that replication is so low for long and medium tongue species.
+# I'm surprised that the betadisper test is not significant, but I think it might be driven by the fact that replication is so low for long and medium tongue species.
 Let's see what happens if we pool long and medium tongue species, since they overlap so much anyway.
@@
                                     type = "median")
-anova(pcoa_horn2_betadisper_pool) # I think this is not significant because short tongued species are disperse on axis 2, but axis 2 does not explain nearly as much variation.
+anova(pcoa_horn2_betadisper_pool) # I think this is not significant because short tongued species are dispersed on axis 2, but axis 2 does not explain nearly as much variation.
 plot(pcoa_horn2_betadisper_pool)
@@
 ggsave("../output/pcoa_aggregate.png")
-
-# a quick peak at diet breadth...quick, I promise
-taxon_diet_breadth <- read_csv("../data/processed_data/network.csv") %>%
-  filter(caste != "male") %>% # we will ignore males 
-  filter(!bb.sp %in% c("humi")) %>% # humi is too rare
-  group_by(site, date, bb.sp, plant.sp) %>%
-  summarize(visits = n()) %>%
-  group_by(site, date, bb.sp) %>%
-  summarize(bb.sp.abund = sum(visits),
-            tax.diet.breadth = length(plant.sp)) %>%
-  mutate(log.bb.sp.abund = log(bb.sp.abund),
-         prop.tax.diet.breadth = tax.diet.breadth/bb.sp.abund)
-
-morph_diet_breadth <- read_csv("../data/processed_data/network.csv") %>%
-  filter(caste != "male") %>% # we will ignore males 
-  filter(!bb.sp %in% c("humi")) %>% # humi is too rare
-  left_join(fl_traits) %>%
-  group_by(site, date, bb.sp, k.type.ss) %>%
-  summarize(visits = n()) %>%
-  group_by(site, date, bb.sp) %>%
-  summarize(bb.sp.abund = sum(visits),
-            morph.diet.breadth = length(k.type.ss)) %>%
-  mutate(log.bb.sp.abund = log(bb.sp.abund),
-         prop.morph.diet.breadth = morph.diet.breadth/bb.sp.abund)
-
-diet_breadth <- full_join(taxon_diet_breadth, morph_diet_breadth) %>%
-  left_join(site_data)
-
-ggplot(diet_breadth, aes(tax.diet.breadth, morph.diet.breadth, color = bb.sp)) +
-  geom_point() +
-  geom_smooth(method = "lm")
-
-ggplot(diet_breadth, aes(morph.diet.breadth, tax.diet.breadth, color = bb.sp)) +
-  geom_point() +
-  geom_smooth(method = "lm")
-
-ggplot(diet_breadth, aes(elev.mean, tax.diet.breadth)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~bb.sp)
-
-ggplot(diet_breadth, aes(elev.mean, morph.diet.breadth)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~bb.sp)
 ```
@@
 ```{r}
+# This one averages visitation and then calculates niche overlaps
 niche_overlapper <- function(x, criterion) {
@@
        tongue1, tongue2, elev.mean) %>%
-    mutate(tongue.diff = as.logical(if_else(tongue1 == tongue2, FALSE, TRUE)),
+    mutate(tongue.diff = factor(if_else(tongue1 == tongue2, "same", "different")),
            tongue.pair = factor(case_when(
@@
 }
+
+# This one calculates niche overlap on fully resolved visitation data and then averages niche overlap across dates
+niche_overlapper2 <- function(x, criterion) {
+
+  criterion <- enquo(criterion) # criterion is set to either plant.sp (for taxonomic) or k.type.s (for morphological)
+
+  x %>%
+    #group_by(bb.sp, !!criterion, site, date, year) %>% # group by site-year and BB-FL pair
+    group_by(bb.sp, !!criterion, site, date) %>% # group by site-year and BB-FL pair
+    summarize(freq = sum(visits)) %>% # get interaction frequency for each BB-FL pair per site-year
+    group_by(bb.sp, site, date) %>% # group by site-year and BB species
+    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
+    #group_by(bb.sp, site, year, !!criterion) %>%
+    # group_by(bb.sp, site, !!criterion) %>%
+    # summarize(mean.prop = mean(prop)) %>%
+    unite(sitedate, c(site, date), sep = "_") %>%
+    #pivot_wider(names_from = !!criterion, values_from = mean.prop, values_fill = 0) %>% # spread into the kind of array vegan likes
+    #pivot_wider(names_from = !!criterion, values_from = prop, values_fill = 0) %>% # spread into the kind of array vegan likes
+    #group_by(site, year) %>%
+    #group_by(site) %>%
+    select(-freq) %>%
+    group_by(sitedate) %>%
+    nest() %>% # nest into list-column
+    mutate(data2 = map(data, function(x) pivot_wider(x, names_from = !!criterion, values_from = prop, values_fill = 0)),
+           data3 = map(data2, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
+           dist = map(data3, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
+           tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
+    #dplyr::select(site, year, tidy_dist) %>% # drop intermediate list-columns
+    #dplyr::select(site, tidy_dist) %>% # drop intermediate list-columns
+    dplyr::select(sitedate, tidy_dist) %>% # drop intermediate list-columns
+    unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
+    separate(sitedate, c("site", "date"), sep = "_") %>%
+    left_join(site_data) %>% # add site data
+    # dplyr::select(site, year, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
+    dplyr::select(site, date, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
+    left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
+    rename(subgenus1 = subgenus, tongue1 = pbl.w.class2) %>%
+    left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
+    rename(subgenus2 = subgenus, tongue2 = pbl.w.class2) %>%
+    # dplyr::select(site, year, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
+    #        tongue1, tongue2, elev.mean) %>%
+    dplyr::select(site, date, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
+       tongue1, tongue2, elev.mean) %>%
+    mutate(tongue.diff = factor(if_else(tongue1 == tongue2, "same", "different")),
+           tongue.pair = factor(case_when(
+             tongue1 == "med" & tongue2 == "med" ~ "med:med",
+             tongue1 == "med" & tongue2 == "long" ~ "med:long",
+             tongue1 == "med" & tongue2 == "short" ~ "short:med",
+             tongue1 == "long" & tongue2 == "med" ~ "med:long",
+             tongue1 == "short" & tongue2 == "med" ~ "short:med",
+             tongue1 == "long" & tongue2 == "short" ~ "short:long",
+             tongue1 == "short" & tongue2 == "long" ~ "short:long",
+             tongue1 == "long" & tongue2 == "long" ~ "long:long",
+             tongue1 == "short" & tongue2 == "short" ~ "short:short",
+           )),
+           subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different")) %>%
+    unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":", remove = FALSE) %>% # create a single species pair column...
+    mutate(sp.pair = factor(sp.pair)) %>% # ...and make it a factor so that mgcv can handle it
+    mutate(niche.overlap = 1-distance) %>% # express niche overlap so that 1 = perfect overlap and 0 = no overlap
+    select(-distance)
+}
 ```
@@
 ```{r}
-# Pooled within site
+# Pooled across dates within level, site, and species pair
 niche_overlap_gen <- niche_overlapper(niche_data, plant.genus) %>%
@@
 
-niche_pool <- bind_rows(niche_overlap_gen,
+# Collate across levels
+niche_overlap <- bind_rows(niche_overlap_gen,
                            niche_overlap_fam, 
                            niche_overlap_ktype,) %>%
-  mutate(level = factor(level, levels = c("morphotype", "family", "genus")),
-         tongue.diff = tongue.diff,
-         site = factor(site))
-
-write_csv(niche_pool, "../hpc_packets/niche_pool.csv")
-
-niche_pool_sitemean <- niche_pool %>%
+  mutate(level = factor(level, levels = c("family", "morphotype", "genus")),
+         site = factor(site)) %>%
+  separate(sp.pair, c("sp1", "sp2"), sep = ":", remove = FALSE)
+
+ggplot(niche_overlap, aes(niche.overlap)) +
+  geom_histogram()
+
+ggplot(filter(niche_overlap,
+              sp1 %in% c("pasc", "prat", "soro", "telu", "wurf") &
+                sp2 %in% c("pasc", "prat", "soro", "telu", "wurf")), aes(niche.overlap)) +
+  geom_histogram() +
+  facet_wrap(~level)
+
+# Pool across species pairs within level and site to get community mean niche overlap
+niche_overlap_levels <- niche_overlap %>%
   group_by(site, level, elev.mean) %>%
-  summarize(mean.niche.overlap = mean(niche.overlap))
-
-# morph_genus_difference <- niche_overlap_gen %>%
-#   select(-niche.overlap, -level) %>%
-#   full_join(niche_overlap_fam) %>%
-#   select(-niche.overlap, -level) %>%
-#   full_join(niche_overlap_ktype) %>%
-#   select(-niche.overlap, -level) %>%
-#   mutate(tongue.diff = tongue.diff,
-#          site = factor(site),
-#          morph.genus.diff = niche.overlap.ktype - niche.overlap.genus) 
-# 
-# ggplot(morph_genus_difference, aes(elev.mean, morph.genus.diff)) +
-#   geom_point() +
-#   geom_smooth() +
-#   facet_wrap(~tongue.pair)
-# 
-# ggplot(morph_genus_difference, aes(tongue.pair, morph.genus.diff)) +
-#   geom_boxplot()
-```
-
-## 3.5 Niche overlap ~ level
+  summarize(mean.niche.overlap = mean(niche.overlap)) %>%
+  mutate(level = factor(level, levels = c("family", "morphotype", "genus")))  # ordering them this way will make family the reference level, which is handy since I don't really care about it.
+
+# Isolate genus-level, then pool within dsp.pair and site
+niche_overlap_sppair <-  niche_overlap %>%
+  group_by(site, sp.pair, level, elev.mean) %>%
+  summarize(mean.niche.overlap = mean(niche.overlap)) %>%
+  mutate(level = factor(level, levels = c("family", "morphotype", "genus"))) %>%
+  filter(level == "genus") %>%
+  separate(sp.pair, c("sp1", "sp2"), sep = ":", remove = FALSE)
+
+# Isolate genus-level, then pool within tongue.pair and site
+niche_overlap_tongues <- niche_overlap %>%
+  group_by(site, level, tongue.pair, elev.mean) %>%
+  summarize(mean.niche.overlap = mean(niche.overlap)) %>%
+  filter(level == "genus")
+
+# Isolate genus-level, then pool within tongue.diff and site
+niche_overlap_tonguediff <- niche_overlap %>%
+  group_by(site, level, tongue.diff, elev.mean) %>%
+  summarize(mean.niche.overlap = mean(niche.overlap)) %>%
+  filter(level == "genus")
+
+#write_csv(niche_pool, "../hpc_packets/niche_pool.csv")
+
+```
+
+## 3.5 Niche overlap ~ level + elevation
+### 3.5.1 Model
+```{r}
+# Beta regression is probably the right way to do this; a gaussian model gives almost identical results, though, since our pooled data are approximately gaussian
+niche_gam00 <- gam(mean.niche.overlap ~
+                     level + 
+                     s(elev.mean, by = level, k = 8, m = 2),
+                   method = "REML",
+                   family = betar(eps = 0.001),
+                   select = TRUE,
+                   data = niche_overlap_levels) %>% getViz()
+
+check.gamViz(niche_gam00)
+summary(niche_gam00)
+print(plot.gamViz(niche_gam00, allTerms = TRUE), pages = 1)
+```
+
+### 3.5.2 Visualize
+```{r}
+# Adding raw data back in
+data_add <- niche_overlap_levels %>%
+  ungroup() %>%
+  arrange(elev.mean, level) %>%
+  select(site, mean.niche.overlap)
+
+# Get predictions using tidymv
+niche_gam00_pred <- get_gam_predictions(niche_gam00, elev.mean) %>%
+  as_tibble() %>%
+  mutate(mean.niche.overlap_trans = inv_logit_scaled(mean.niche.overlap),
+         CI_upper_trans = inv_logit_scaled(CI_upper),
+         CI_lower_trans = inv_logit_scaled(CI_lower)) %>%
+  select(-mean.niche.overlap) %>%
+  bind_cols(data_add)
+
+# Plot smooths
+ggplot(niche_gam00_pred, aes(elev.mean, mean.niche.overlap, color = level)) +
+  geom_point() +
+  geom_ribbon(aes(elev.mean, mean.niche.overlap_trans, 
+                  ymax = CI_upper_trans, ymin = CI_lower_trans,
+                  fill = level),
+              alpha = 0.3, inherit.aes = FALSE) +
+  geom_line(aes(elev.mean, mean.niche.overlap_trans, 
+                color = level), inherit.aes = FALSE) +
+  ylab("Niche overlap") +
+  xlab("Elevation (m)") +
+  guides(fill = FALSE) +
+  labs(color = "Level") +
+  theme_light()
+
+ggsave(file = "../output/niche_gam00_plot1.pdf", width = 5, height = 4)
+ggsave(file = "../output/niche_gam00_plot1.png", width = 5, height = 4)
+
+# Plot intercept effect of level
+plot(pterm(niche_gam00, 1))
+
+plotDiff(sm(niche_gam00, 1), sm(niche_gam00, 3)) +
+  l_ciPoly() + 
+  l_fitLine() + 
+  geom_hline(yintercept = 0, linetype = 2)
+```
+
+## 3.7 Genus-level niche overlap by tongue.diff
+
+### 3.7.1 Model
+```{r}
+niche_gam20 <- gam(mean.niche.overlap ~
+                     tongue.diff +
+                     s(elev.mean, by = tongue.diff, k = 8, m = 2),
+                   method = "REML",
+                   family = betar(eps = 0.001),
+                   select = TRUE,
+                   data = niche_overlap_tonguediff) %>% getViz()
+
+check.gamViz(niche_gam20)
+summary(niche_gam20)
+print(plot(niche_gam20, allTerms = TRUE), pages = 1)
+```
+
+### 3.7.2 Visualize
+```{r}
+# Get predictions
+niche_gam20_pred <- get_gam_predictions(niche_gam20, elev.mean) %>%
+  as_tibble() %>%
+  mutate(mean.niche.overlap_trans = inv_logit_scaled(mean.niche.overlap),
+         CI_upper_trans = inv_logit_scaled(CI_upper),
+         CI_lower_trans = inv_logit_scaled(CI_lower))
+
+# Plot smooths
+ggplot(niche_gam20_pred, aes(elev.mean, mean.niche.overlap_trans, color = tongue.diff)) +
+  geom_ribbon(aes(elev.mean, mean.niche.overlap_trans, 
+                  ymax = CI_upper_trans, ymin = CI_lower_trans,
+                  fill = tongue.diff),
+              alpha = 0.3, inherit.aes = FALSE) +
+  geom_line() +
+  ylab("Niche overlap") +
+  xlab("Elevation (m)") +
+  guides(fill = FALSE) +
+  labs(color = "Tongue class") +
+  theme_light()
+
+ggsave(file = "../output/niche_gam20_plot1.pdf", width = 5, height = 4)
+ggsave(file = "../output/niche_gam20_plot1.png", width = 5, height = 4)
+
+# Plot intercept effect of tongue.diff
+plot(pterm(niche_gam20, 1))
+```
+
+-------------------------------------
+# Old
+
+## 3.6 Genus-level niche overlap ~ tongue.pair + elevation
+### 3.6.1 Model
+```{r}
+niche_gam10 <- gam(mean.niche.overlap ~
+                     tongue.pair +
+                     s(elev.mean, by = tongue.pair, k = 8, m = 2),
+                   method = "REML",
+                   family = betar(eps = 0.001),
+                   select = TRUE,
+                   data = niche_overlap_tongues) %>% getViz()
+
+check.gamViz(niche_gam10)
+summary(niche_gam10)
+print(plot(niche_gam10, allTerms = TRUE), pages = 1)
+```
+
+### 3.6.2 Visualize
+```{r}
+# Get predictions using tidymv
+niche_gam10_pred <- get_gam_predictions(niche_gam10, elev.mean) %>%
+  as_tibble() %>%
+  mutate(mean.niche.overlap_trans = inv_logit_scaled(mean.niche.overlap),
+         CI_upper_trans = inv_logit_scaled(CI_upper),
+         CI_lower_trans = inv_logit_scaled(CI_lower))
+
+# Plot smooths
+ggplot(niche_gam10_pred, aes(elev.mean, mean.niche.overlap_trans, color = tongue.pair)) +
+  geom_ribbon(aes(elev.mean, mean.niche.overlap_trans,
+                  ymax = CI_upper_trans, ymin = CI_lower_trans,
+                  fill = tongue.pair),
+              alpha = 0.3, inherit.aes = FALSE) +
+  geom_line() +
+  ylab("Niche overlap") +
+  xlab("Elevation") +
+  theme_light()
+
+ggsave(file = "../output/niche_gam10_plot1.pdf", width = 5, height = 4)
+ggsave(file = "../output/niche_gam10_plot1.png", width = 5, height = 4)
+```
+
+### 3.5.1 Model
 ```{r}
@@
 mcmc_plot(brm_niche00, pars = c("^r_site*"))
-
+```
+
+### 3.5.2 Vizualize
+```{r}
 brm_niche00_plotframe1 <- brm_niche00 %>%
@@
   geom_pointinterval() +
-  theme_light(16) +
+  theme_light(12) +
   labs(x = "Niche overlap", y= "Level") +
-  guides(fill = FALSE)
-
-ggplot(brm_niche00_plotframe1, aes(level, value_trans, 
-                                 ymin = .lower_trans, ymax = .upper_trans)) +
-  geom_pointinterval() +
-  theme_light(16) +
-  labs(x = "Level", y= "Niche overlap") +
   guides(fill = FALSE) +
-  ylim(c(0, 0.55))
-
-#ggsave(file = "../output/brm_pvis01_plot1.pdf", width = 6, height = 7)
+  scale_x_continuous(breaks = seq(0.25, 0.5, by = 0.05))
+
+ggsave(file = "../output/brm_niche00_plot1.pdf", width = 4.5, height = 4)
+ggsave(file = "../output/brm_niche00_plot1.png", width = 4.5, height = 4)
 ```
@@
 ## 3.6 Genus niche overlap ~ tongue
+### 3.6.1 Model
 ```{r}
@@
 
+### 3.6.2 Vizualize
+```{r}
+brm_niche10_plotframe1 <- brm_niche10 %>%
+  spread_draws(b_Intercept, r_tongue.pair[condition, term]) %>%
+  mutate(condition_mean = b_Intercept + r_tongue.pair) %>%
+  group_by(condition) %>% # group by tongue:morpotype pair
+  point_interval(condition_mean, .width = c(0.95, 0.66)) %>% # calculate credible intervals
+  mutate(condition_mean_trans = inv_logit_scaled(condition_mean), # exponentiating the estimates gets us back to the response scale (negbinomial uses log link function)
+         .lower_trans = inv_logit_scaled(.lower),
+         .upper_trans = inv_logit_scaled(.upper)) %>%
+  mutate(condition = factor(condition, levels = c("short:short", "med:med", "long:long",
+                                          "short:med", "short:long", "med:long")))
+
+ggplot(brm_niche10_plotframe1, aes(condition_mean_trans, condition, 
+                                 xmin = .lower_trans, xmax = .upper_trans)) +
+  geom_pointinterval() +
+  theme_light(12) +
+  labs(x = "Niche overlap", y= "Tongue pair") +
+  guides(fill = FALSE)
+
+ggsave(file = "../output/brm_niche10_plot1.pdf", width = 4.5, height = 4)
+ggsave(file = "../output/brm_niche10_plot1.png", width = 4.5, height = 4)
+```
+
 ## 3.7 Genus niche overlap ~ sp.pair
-```{r}
+### 3.7.1 Model
+```{r eval=FALSE, include=FALSE}
 brm_niche20 <- brm(niche.overlap ~ 1 +
@@
                    chains = 4,
-                   data = filter(niche_pool, level == "genus"))
+                   data = filter(niche_overlap, level == "genus"))
 
@@
 
+### 3.7.2 Visualize
+```{r eval=FALSE, include=FALSE}
+brm_niche20_plotframe1 <- brm_niche20 %>%
+  spread_draws(b_Intercept, r_sp.pair[condition, term]) %>%
+  mutate(condition_mean = b_Intercept + r_sp.pair) %>%
+  group_by(condition) %>% # group by tongue:morpotype pair
+  point_interval(condition_mean, .width = c(0.95, 0.66)) %>% # calculate credible intervals
+  mutate(condition_mean_trans = inv_logit_scaled(condition_mean), # exponentiating the estimates gets us back to the response scale (negbinomial uses log link function)
+         .lower_trans = inv_logit_scaled(.lower),
+         .upper_trans = inv_logit_scaled(.upper)) %>%
+  mutate(condition = factor(condition)) %>%
+  left_join(niche_pool, by = c("condition" = "sp.pair")) %>%
+  select(condition, condition_mean_trans, .lower_trans, .upper_trans, tongue.pair) %>%
+  distinct() %>%
+  mutate(tongue.pair = factor(tongue.pair, levels = c("short:short", "med:med", "long:long",
+                                                      "short:med", "short:long", "med:long")))
+
+ggplot(brm_niche20_plotframe1, aes(condition_mean_trans, condition, 
+                                 xmin = .lower_trans, xmax = .upper_trans)) +
+  geom_pointinterval() +
+  theme_light(11) +
+  labs(x = "Niche overlap", y= "Tongue pair") +
+  guides(fill = FALSE) +
+ # facet_wrap(~tongue.pair, ncol = 1, scales = "free_y") +
+  facet_wrap(~tongue.pair, scales = "free_y")
+
+ggsave(file = "../output/brm_niche20_plot1.pdf", width = 4.5, height = 4)
+ggsave(file = "../output/brm_niche20_plot1.png", width = 7, height = 10)
+```
+
 ## 3.8 Niche overlap ~ elevation
-```{r}
+### 3.8.1 Models
+```{r}
+ggplot(filter(niche_overlap_sppair, !sp1 %in% c("humi") & !sp2 %in% c("humi")), 
+       aes(elev.mean, mean.niche.overlap, color = sp.pair)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
+  facet_wrap(~sp.pair) +
+  ylim(c(0,1))
+
+ggplot(filter(niche_overlap_sppair, 
+              sp1 %in% c("pasc", "prat", "soro", "telu", "wurf") & 
+                sp2 %in% c("pasc", "prat", "soro", "telu", "wurf")), 
+       aes(elev.mean, mean.niche.overlap, color = sp.pair)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
+  facet_wrap(~sp.pair) +
+  ylim(c(0,1))
+
+ggplot(filter(niche_overlap_sppair, 
+              sp1 %in% c("pasc", "prat", "soro", "telu", "wurf",
+                         "mont", "muci", "pyre", "mend", "gers") & 
+                sp2 %in% c("pasc", "prat", "soro", "telu", "wurf",
+                           "mont", "muci", "pyre", "mend", "gers")), 
+       aes(elev.mean, mean.niche.overlap, color = sp.pair)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
+  facet_wrap(~sp.pair) +
+  ylim(c(0,1))
+
+ggplot(filter(niche_overlap_sppair,
+              sp1 != "humi" & sp2 != "humi" &
+              sp1 %in% c("pasc", "prat", "soro", "telu", "wurf") | 
+                sp2 %in% c("pasc", "prat", "soro", "telu", "wurf")), 
+       aes(elev.mean, mean.niche.overlap, color = sp.pair)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
+  facet_wrap(~sp.pair) +
+  ylim(c(0,1))
+
 # Pairwise
-gen_gam_niche.I <- bam(niche.overlap ~
+gen_gam_niche.I <- bam(mean.niche.overlap ~
                             s(sp.pair, bs = "re") +
@@
                           select = TRUE,
-                          data = filter(niche_pool, level == "genus" &
-                                          bb.sp1 %in% c("pasc", "prat", "soro", "telu", "wurf", # dominant species
+                          data = filter(niche_overlap_sppair, level == "genus" &
+                                          sp1 %in% c("pasc", "prat", "soro", "telu", "wurf", # dominant species
                                                         "mont", "muci", "pyre", "mend") & # high elevation specialists
-                                          bb.sp2 %in% c("pasc", "prat", "soro", "telu", "wurf",
+                                          sp2 %in% c("pasc", "prat", "soro", "telu", "wurf",
                                                         "mont", "muci", "pyre", "mend"))) %>% getViz()
@@
 summary(gen_gam_niche.I)
-plot(sm(gen_gam_niche.I, 9))
-plot(sm(gen_gam_niche.I, 12))
-plot(sm(gen_gam_niche.I, 14))
-plot(sm(gen_gam_niche.I, 15))
-plot(sm(gen_gam_niche.I, 16))
-plot(sm(gen_gam_niche.I, 32))
+plot(sm(gen_gam_niche.I, 1))
+plot(sm(gen_gam_niche.I, 11))
+plot(sm(gen_gam_niche.I, 28))
+plot(sm(gen_gam_niche.I, 33))
+plot(sm(gen_gam_niche.I, 37))
 
@@
 
-mean_niche_gam.I <- gam(mean.niche.overlap ~
-                            level + 
-                            s(elev.mean, by = level, k = 8, m = 2),
-                          method = "REML",
-                          family = betar(eps = 0.001),
-                          select = TRUE,
-                          data = niche_pool_sitemean) %>% getViz()
-
-check.gamViz(mean_niche_gam.I)
-summary(mean_niche_gam.I)
-print(plot.gamViz(mean_niche_gam.I, allTerms = TRUE), pages = 1)
-```
+```

### 2021-03-08 ms_2/ms_2_v2.Rmd
improved niche overlap analysis

@@
 This association between tongue length and floral morphology, though, is imperfect [e.g.
-@Ranta1980-ok; @Goulson2008-oe], and access to floral rewards in some flowers involves learning to manipulate floral structures that are challenging irrespective of tongue length [@Laverty1994-hm].
+@Ranta1980-ok; @Goulson2008-oe], perhaps in part because access to floral rewards in some flowers involves learning to manipulate floral structures that are challenging irrespective of tongue length [@Laverty1994-hm].
 For example, the characteristic Papilionaceous floral morphology of legumes requires a bee to separate tightly appressed keel petals to access anthers and nectaries [@Westerkamp1997-hs], and the narrowly campanulate flowers of many Ericaceous flora pose a significant barrier to large-bodied visitors like bumble bees [@Wallace1977-kz].
@@
 
-Rather than considering various floral morphological traits independently, we rely on the synthesis of @Kugler1970-gb, who developed a typology consisting of ten primary morphotypes and nested subdivision of each.
+Rather than considering various floral morphological traits independently, we relied on the synthesis of @Kugler1970-gb, who developed a typology consisting of ten primary morphotypes and nested subdivision of each.
 Kugler’s typology was informed by laboratory studies of bumble bee flower choice [@Kugler1943-uy], and there is a precedent for analyzing bumble bee visitation in terms of Kugler’s morphotypes [@Neumayer1999-il; @Schneller2014-ry].
@@
 Morphotypes 9 and 10 were extremely rare in our study system and were therefore omitted from all analyses.
-Plants belonging to morphotype 0 were typically not recorded during floral surveys due to their lack of conspicuous floral displays, so this morphotype was also omitted from all analyses.
+Plants belonging to morphotype 0 were typically not recorded during floral surveys due to their lack of conspicuous floral displays, and they were rarely visited by bumble bees, so this morphotype was also omitted from all analyses.
 
@@
 
+### Defining bias
+
+We opt to use of the term “bias” in favor of the term “preference,” since the latter is laden with more assumptions and its inference under field conditions is problematic [@Williams2005-yp].
+By “bias,” we mean simply an empirical disproportionality between the relative abundance of a given floral taxon or group and its proportional visitation.
+The exact mechanism of this disproportionality — whether an arbitrary “preference” on the part of the foraging bee, an unmeasured decoupling of relative abundance and rate of reward, etc.
+— is beyond the scope of our discussion, and probably best studied under laboratory conditions where strict factorial control is possible.
+We focus instead on the empirical phenomenon of bias, its variation, and its correlation to morphology.
+
 ## Data analysis
@@
 In rare cases, a given floral species was recorded during visitation sampling but missed during floral surveying.
-These species were retroactively assigned the minimum floral area value of 0.01 m^2.
+These species were retroactively assigned the minimum floral area value of 0.01 m^2^.
 
-Visitation data were extremely right-skewed because the majority of available flora received zero visits during a given sampling period.
+Visitation data were extremely right-skewed because the majority of available flora received zero visits during any given sampling period.
 To facilitate model fitting, we created a binary version of visitation rate in which visited flora were assigned to 1 and unvisited flora to 0.
 We then modeled binary visitation using generalized models with binomial or Bernoulli distribution families.
-The conditional effects inferred from these models should, therefore, be interpreted as effects on visitation probability rather than visitation frequency.
+The conditional effects inferred from these models must, therefore, be interpreted as effects on visitation probability rather than visitation frequency.
+
+We calculated niche overlap with the goal of focusing on elevational patterns, sacrificing temporal resolution to pool observations within site.
+First, we calculated each bumble bee species' proportional visitation frequency at the nested levels of floral morphotype, plant family, and plant genus.
+Then, we calculated, for each bumble bee species, the mean proportional visitation rate to each morphotype/family/genus across all sampling dates within each site.
+Using these species-level visitation matrices, we calculated pairwise niche overlap as 1 - Horn-Morisita distance (calculated separately for morphotype-, family-, and genus-level matrices) for all species pairs within each site.
+This approach was chosen for its normalizing effect on the highly variable "snapshot" visitation data generated by our individual transect walks, but it has the disadvantage of conflating niche overlap due to similar floral preference and niche overlap due to similar temporal (i.e.
+seasonal) activity patterns.
+
+<!-- Maybe explain why we only went down to genus instead of species -->
 
 While our study spanned three years, we chose not to model year effects.
-Differences in sampling intensity and duration across years, for the reasons described above, would make comparisons across years difficult to interpret, and we deemed these comparisons not worth the cost of added model complexity.
+Differences in sampling intensity and duration across years, for the reasons described above, would make comparisons across years difficult to interpret.
 This decision entails the caveats that our analysis cannot detect year-to-year variation in floral biases and niche overlap, and main effects cannot be distinguished from annual anomalies.
-With only three years of data, though, such annual inferences would be of little value.
+With only three years of data, though, such annual inferences would be of little value, and we deemed these comparisons not worth the cost of added model complexity.
 
-<!-- State more explicityl how the models should be interpreted withr expect to bias -->
+<!-- State more explicityl how the models should be interpreted with respect to bias -->
 
@@
 
+<!-- Something about priors -->
+
 <!-- Figure 3 order by tongue length class -->
@@
 <!-- Open each results paragraph with a summary sentence. -->
-
-### Defining bias
-
-We opt to use of the term “bias” in favor of the term “preference,” since the latter is laden with more assumptions and its inference under field conditions is problematic [@Williams2005-yp].
-By “bias,” we mean simply an empirical disproportionality between the relative abundance of a given floral taxon or group and its proportional visitation.
-The exact mechanism of this disproportionality — whether an arbitrary “preference” on the part of the foraging bee, an unmeasured decoupling of relative abundance and rate of reward, etc.
-— is beyond the scope of our discussion, and probably best studied under laboratory conditions where strict factorial control is possible.
-We focus instead on the empirical phenomenon of bias, its variation, and its correlation to morphology.
 
@@
 
-### Model 5: Niche overlap by level
+### Model 5: Niche overlap by level and elevation
 
-### Model 6: Genus-level overlap by species-pair
+We modeled niche overlap as a function of level (morphotype, family, genus) and elevation using hierarchical generalized additive models (HGAMs).
+First, we calculated community mean niche overlap by taking the mean niche overlap (within level) across all species pairs for each site.
+Level was included as both a fixed intercept effect and a smoothing factor in conjunction with elevation, following the type "I" model form described by @Pedersen2019-kf.
+To account for the proportional nature of our data, we used a beta distribution family with a logit link function.
 
-### Model 7: Community-mean genus-level niche overlap by elevation
+### Model 6: Genus-level niche overlap by tongue length and elevation
 
-We analyze both pairwise niche overlap and community mean niche overlap, since low pairwise niche overlap could conceal a high potential for diffuse competition [@Pianka1974-qk; MacArthur1972-qz]
-
-### Model 8: Pairwise genus-level niche overlap by elevation
+To focus on genus-level niche overlap as a function of tongue length class, we generated a binary tongue-difference variable for each species pair: "same" for species belonging to the same tongue length class and "different" for species belonging to different tongue length classes.
+We then took the mean genus-level niche overlap across all species pairs within tongue-difference class and site.
+Then, we modeled genus-level niche overlap using tongue-difference as both a fixed intercept effect and a smoothing factor in conjunction with elevation, again following the type "I" model form @Pedersen2019-kf.
 
@@
 Ordination analysis and permutation testing were performed with the package `vegan` `r citep(citation("vegan"))` and visualized using the packages `ggplot2` `vegan` `r citep(citation("ggplot2"))` and `ggvegan` (Simpson 2019).
-GAMs were implemented with the package `mgcv` [@Wood_2017] and visualized with the package `mgcViz` [@Fasiolo2018aa].
+HGAMs were implemented with the package `mgcv` [@Wood_2017] and visualized with the package `mgcViz` [@Fasiolo2018aa].
 Data handling and visualization was performed with the `tidyverse` suite of R packages `r citep(citation("tidyverse"))`.
@@
 
-```{r fig5, echo = FALSE,  fig.cap = paste(" Principal coordinates ordination of bumble bee visitation pooled across all samples.")}
+```{r fig5, echo = FALSE,  fig.cap = paste("Principal coordinates ordination of bumble bee visitation pooled across all samples.")}
 knitr::include_graphics("../output/pcoa_aggregate.png")
+```
+
+## Niche overlap
+
+### Niche overlap through elevation by level
+
+Morphotype-level niche overlap was nearly double genus-level niche overlap throughout the whole elevation gradient, with family-level niche overlap taking intermediate values.
+The fixed intercepts of morphotype- and genus-level niche overlap each differed significantly (in opposite directions) from the reference level of family-level niche overlap (p << 0.001).
+The conditional smooths of each level were significantly different from flat lines (p < 0.001) and generally mirrored one another in shape **(Figure 8)**.
+Niche overlap was highest at the low end of the elevation gradient, decreased steeply from 600-1200 m, and leveled off or declined weakly for the remainder of the gradient, though morphotype-level niche overlap exhibited a second peak just above 1500 m that was not evident in genus- or family-level niche overlap.
+The model explained 78.4% of total deviance (adjusted R^2^ = 0.743).
+
+```{r fig8, echo = FALSE,  fig.cap = paste("Niche overlap by level through elevation.
+Points show empirical niche overlap values, lines show GAM smooths, and shaded bands show the standard errors of the mean")}
+knitr::include_graphics("../output/niche_gam00_plot1.png")
+```
+
+### Genus-level niche overlap by tongue length and elevation
+
+Whether a species pair belonged to the same or different tongue length classes had no effect on genus-level niche overlap, either in terms of intercept (p = 0.18) or response to elevation **(Figure 9)**.
+The model explained 36.5% of total deviance (adjusted R^2^ = 0.316).
+
+```{r fig9, echo = FALSE,  fig.cap = paste("Niche overlap by tongue length class difference through elevation. Lines show GAM smooths and shaded bands show standard errors of the mean")}
+knitr::include_graphics("../output/niche_gam20_plot1.png")
 ```
@@
 
+We analyze community mean niche overlap as an indication of diffuse competition [@Pianka1974-qk; MacArthur1972-qz]
+
 Our finding that bumble bee diet ordination produces a primary axis corresponding to tongue length — and associated with affinity for legumes — and a secondary axis describing the dispersion of short-tongued species is consistent with studies in Poland (Dave Goulson, Lye, and Darvill 2008), France (Durieux 2000), and England (D.

### 2021-03-09 code/bias_niche.Rmd
backup

@@
 ggsave(file = "../output/brm_pvis00_plot1.pdf", width = 6, height = 5)
+ggsave(file = "../output/brm_pvis00_plot1.png", width = 6, height = 5)
 ```
@@
 ggsave(file = "../output/brm_pvis02_plot1.pdf", width = 8, height = 8)
+ggsave(file = "../output/brm_pvis02_plot1.png", width = 8, height = 8)
 ```
@@
   left_join(read_csv("../data/processed_data/network.csv")) %>%
-  select(plant.sp, plant.sp.abb, plant.genus, plant.family, k.type.ss, color) %>%
+  #select(plant.sp, plant.sp.abb, plant.genus, plant.family, k.type.ss, color) %>%
+  select(plant.genus, k.type.ss) %>%
   mutate(k.type.ss = factor(k.type.ss)) %>%
-  distinct()
+  distinct() %>%
+  na.omit() # THIS IS JUST A PLACEHOLDER UNTIL I FIX THE GENTIANA/ELLA/OPSIS NOMENCALTURE
 
@@
   filter(!bb.sp %in% c("humi")) %>% # humi is too rare
-  group_by(bb.sp, plant.sp.abb) %>%
+  #group_by(bb.sp, plant.sp.abb) %>%
+  group_by(bb.sp, plant.genus) %>%
   summarize(visits = n()) %>%
@@
   select(-visits) %>%
-  pivot_wider(names_from = plant.sp.abb, values_from = prop.visits, values_fill = 0) %>%
+  #pivot_wider(names_from = plant.sp.abb, values_from = prop.visits, values_fill = 0) %>%
+  pivot_wider(names_from = plant.genus, values_from = prop.visits, values_fill = 0) %>%
   column_to_rownames("bb.sp")
@@
   filter(Score == "species") %>%
-  rename(plant.sp.abb = Label) %>%
+  #rename(plant.sp.abb = Label) %>%
+  rename(plant.genus = Label) %>%
   left_join(fl_traits2)
@@
   geom_text(aes(label = bb.sp), key_glyph = "rect") +
-  geom_text_repel(data = filter(pcoa_horn2.fort1, abs(MDS1) >= 0.2 | abs(MDS2) >= 0.2), 
-            aes(MDS1, MDS2, label = plant.sp.abb), inherit.aes = FALSE, color = "gray40") +
-  geom_point(data = filter(pcoa_horn2.fort1, abs(MDS1) >= 0.2 | abs(MDS2) >= 0.2),
+  geom_text_repel(data = filter(pcoa_horn2.fort1, abs(MDS1) >= 0.22 | abs(MDS2) >= 0.22), 
+            aes(MDS1, MDS2, label = plant.genus), inherit.aes = FALSE, color = "gray40") +
+  geom_point(data = filter(pcoa_horn2.fort1, abs(MDS1) >= 0.22 | abs(MDS2) >= 0.22),
             aes(MDS1, MDS2, shape = k.type.ss), inherit.aes = FALSE, color = "gray40", size = 3) +
@@
   coord_fixed() +
-  theme_light(10)
+  theme_light(9)
 

### 2021-03-09 ms_2/ms_2_v2.Rmd
backup

@@
 We surveyed floral abundance and recorded bumble bee floral visitation patterns along a ~1400 m elevation gradient in German Alps.
-Using these data, we first ask how floral biases compare across bumble bee species and whether these patterns can be explained by trait-matching between bumble bee tongue length and floral morphotype.
-We then investigate the patterns of diet overlap between bumble bee species that emerge from their respective floral biases, comparing diet overlap at the nested levels of floral morphotype, floral family, and floral genus.
-Finally, we ask how both community-mean and species-pairwise diet overlap vary with elevation.
-While tongue length was a strong predictor of bumble bees' floral morphotype biases, species with similar tongue lengths exhibited divergent biases with respect to certain flora morphotypes.
-This indicates that factors other than tongue length are involved in flower choice, which might explain the coexistence of bumble bee species with similar tongue lengths.
-Moreover, genus-level diet overlap was approximately one half that of morphotype-level diet overlap.
-Thus, even when bumble bees exhibit similar biases at the level of floral morphotype, partly due to the constraints of tongue length, floral diversity enables the mitigation of competition via the genus- or species-level subpartitioning of floral resources within morphotypes.
+Using these data, we first ask how floral preferences compare across bumble bee species and whether these patterns can be explained by trait-matching between bumble bee tongue length and floral morphotype.
+We then investigate the patterns of diet overlap between bumble bee species at the nested levels of floral morphotype, floral family, and floral genus, and model diet overlap as a function bumble bee tongue length and elevation.
+We find that while tongue length is a strong predictor of bumble bees' floral morphotype preferences, this pattern is far from perfect, and species with similar tongue lengths exhibit divergent preferences with respect to certain flora morphotypes.
+This indicates that factors other than tongue length are involved in flower choice, which might explain the puzzling coexistence of bumble bee species with similar tongue lengths.
+Moreover, genus-level diet overlap was approximately one half that of morphotype-level diet overlap, indicating that even when bumble bees exhibit similar biases at the level of floral morphotype, partly due to the constraints of tongue length, floral diversity enables the mitigation of competition via the genus- or species-level subpartitioning of floral resources within morphotypes.
 This suggests that floral species richness may be as important as overall floral abundance in avoiding the competitive exclusion of bumble bee species.
-We find that community-mean diet overlap, indicative of the potential for diffuse competition, generally declines with elevation, while the diet overlap for certain species pairs exhibits mid- or high-elevation peaks.
-The long-term coexistence of mountain bumble bee species may depend on how climate induced range shifts in bumble bees and their floral hosts will alter the these elevational patterns of diet overlap.
+Diet overlap was highest at the low end of our elevation gradient, declining sharply over the transition from valley floor to mid-slope, then leveling off or declining weakly for the remainder of the gradient.
+If diet overlap is interpreted as an index of potential competition, upslope shifts in bumble bees and flora could generate more competitive conditions at high elevations if and when floral resources become limiting.
 
@@
 In this role, however, they have proven somewhat enigmatic.
-On the one hand, there is evidence that resource limitation can be a significant constraint on local bumble communities [@Heinrich1976-wc; @Bowers1985-df], and certain bumble bee communities have been shown to be over-dispersed in terms of tongue length (a matching trait corresponding to floral corolla depth) or phylogeny (understood to encompass unmeasured trait variation), providing evidence for competition as a community assembly process [@Inouye1977-kl; @Pellissier2013-pk].
+On the one hand, there is evidence that resource limitation can be a significant constraint on local bumble communities [@Heinrich1976-wc; @Bowers1985-df], and certain bumble bee communities have been shown to be over-dispersed in terms of tongue length (a matching trait corresponding to floral corolla depth) or phylogeny (understood to encompass unmeasured trait variation), suggesting that competition functions as a community assembly process [@Inouye1977-kl; @Pellissier2013-pk].
 Nevertheless, bumble bee communities with an apparent excess of species richness and deficit of trait (or phylogenetic) dispersion are commonly observed [@Ranta1981-vx; @Goulson2008-zy].
@@
 Nevertheless, the flexibility of bumble bee foraging is tuned to floral reward, which depends on both the intrinsic reward of a given flower and the costs associated with exploiting it.
-These costs are closely related to the morphological fittedness between bumble bees and their floral hosts, which consists principally in the compatibility between bumble bee tongue length and floral corolla depth.
+These costs are closely related to the degree of morphological fittedness between bumble bees and their floral hosts, which consists principally in the matching of bumble bee tongue length to floral corolla depth.
 Bumble bee species can differ markedly in tongue length, and floral handling time increases both for short-tongued bumble bees foraging on deep flowers and long-tongued bees foraging on shallow flowers [@Inouye1980-mv], which leads to an association between tongue length and floral niche in bumble bee communities [e.g.
@@
 @Ranta1980-ok; @Goulson2008-oe], perhaps in part because access to floral rewards in some flowers involves learning to manipulate floral structures that are challenging irrespective of tongue length [@Laverty1994-hm].
-For example, the characteristic Papilionaceous floral morphology of legumes requires a bee to separate tightly appressed keel petals to access anthers and nectaries [@Westerkamp1997-hs], and the narrowly campanulate flowers of many Ericaceous flora pose a significant barrier to large-bodied visitors like bumble bees [@Wallace1977-kz].
-Thus the dietary niches of bumble bee species in a given ecological context are emergent products of innate floral biases, rewards experienced during foraging, and learned flower handling skills, all of which contribute to the phenomenon of interspecific niche overlap upon which interspecific competition is predicated.
-
-<!-- I expect my co-authors to want the section below to get pushed to the discussion. I disagree, but I think we could make it work if they insist. -->
 
@@
 This creates a predicament in which niche overlap can only be interpreted as an index of potential competition when realized competition --- which cannot be inferred from niche overlap --- is assumed to be negligible, and this assumption can only be justified on the basis of additional knowledge of the study system under consideration.
+
 Our study is set in the Berchtesgadener Alps of southeast Germany, a landscape of flower-rich mountain meadows --- generated by past or ongoing grazing management --- embedded in a matrix of coniferous forest.
@@
 
-The aim of our study is to synthesize the phenomena of floral bias, dietary niche overlap, and interspecific competition in a way that accounts for the coexistence of bumble bee species with similar tongue lengths and predicts the effects of climate-driven elevational range shifts on the future of potential competition between bumble bee species in mountain habitats.
-We first ask how floral biases compare across bumble bee species and to what extent these patterns can be explained by trait-matching between bumble bee tongue length and floral morphotype.
-We then investigate the patterns of dietary niche overlap between bumble bee species that emerge from their respective floral biases, comparing niche overlap at the nested levels of floral morphotype, floral family, and floral genus.
+The aim of our study is to synthesize the phenomena of floral preference, trait-matching, and dietary niche overlap in a way that accounts for the coexistence of bumble bee species with similar tongue lengths and predicts the effects of climate-driven elevational range shifts on the future of potential competition between bumble bee species in mountain habitats.
+We first ask how floral preferences compare across bumble bee species and to what extent these patterns can be explained by trait-matching between bumble bee tongue length and floral morphotype.
+We then investigate the patterns of dietary niche overlap between bumble bee species that emerge from their respective floral preferences, comparing niche overlap at the nested levels of floral morphotype, floral family, and floral genus.
 Finally, we ask how niche overlap varies with elevation and discuss the implications of this variation for bumble bee conservation in a changing climate.
@@
 
-### Defining bias
+## Data analysis
 
-We opt to use of the term “bias” in favor of the term “preference,” since the latter is laden with more assumptions and its inference under field conditions is problematic [@Williams2005-yp].
-By “bias,” we mean simply an empirical disproportionality between the relative abundance of a given floral taxon or group and its proportional visitation.
-The exact mechanism of this disproportionality — whether an arbitrary “preference” on the part of the foraging bee, an unmeasured decoupling of relative abundance and rate of reward, etc.
-— is beyond the scope of our discussion, and probably best studied under laboratory conditions where strict factorial control is possible.
-We focus instead on the empirical phenomenon of bias, its variation, and its correlation to morphology.
+### Software
 
-## Data analysis
+All analyses were conducted in `R` `r citep(citation())`.
+Data handling and visualization was performed with the `tidyverse` suite of R packages `r citep(citation("tidyverse"))`.
+Bayesian multilevel models were implemented in Stan (Stan Development Team 2021) via the package `brms` `r citep(citation("brms"))` using default priors, and visualizations were prepared with the packages `tidybayes` `r citep(citation("tidybayes"))` and `ggdist` `r citep(citation("ggdist"))`.
+Ordination analysis and permutation testing were performed with the package `vegan` `r citep(citation("vegan"))` and visualized using the packages `ggplot2` `vegan` `r citep(citation("ggplot2"))` and `ggvegan` (Simpson 2019).
+Hierarchical generalized additive models (HGAMs) were implemented with the package `mgcv` [@Wood_2017] and visualized with the packages `mgcViz` [@Fasiolo2018aa] and `tidymv` [@Coretta-aa2021].
+All R code used in this analysis, together with full model output and diagnostics, is available in Appendix X.
 
@@
 In these cases, visitation samples were paired with the nearest available floral survey data within a maximum distance of 7 days.
-Visitation samples lacking corresponding floral survey data within 7 days were omitted from further analysis.
+Visitation samples lacking corresponding floral survey data within 7 days were omitted from further analysis, which resulted in dropping 40 site-dates out of a total of 895, leaving 855 site-dates included in our analyses.
 In a few cases, two consecutive visitation samples at a given site shared the same nearest floral survey sample; in these cases, the visitation samples were first pooled, then aligned to their shared nearest floral survey sample.
@@
 
+We defined floral preference as the relative rate of visitation between floral morphotypes after conditioning on the neutral effects of abundance (see model descriptions below for details).
 Visitation data were extremely right-skewed because the majority of available flora received zero visits during any given sampling period.
@@
 seasonal) activity patterns.
-
-<!-- Maybe explain why we only went down to genus instead of species -->
+We opted to analyze niche overlap only down to the level of floral genus in order to dampen the possible effects of false non-detection of floral species due to undersampling of the diets of rare bumble bee species [@Williams2005-yp].
 
 While our study spanned three years, we chose not to model year effects.
-Differences in sampling intensity and duration across years, for the reasons described above, would make comparisons across years difficult to interpret.
+Differences in sampling intensity and duration across years (described above) would make comparisons across years difficult to interpret.
 This decision entails the caveats that our analysis cannot detect year-to-year variation in floral biases and niche overlap, and main effects cannot be distinguished from annual anomalies.
-With only three years of data, though, such annual inferences would be of little value, and we deemed these comparisons not worth the cost of added model complexity.
+With only three years of data, though, such annual inferences would be of little value, and we deemed them not worth the cost of added model complexity.
 
@@
 
-<!-- Report the effects of abundance that I am conditioning on -->
-
-<!-- Perhaps show what happens if we remove the abundance interaction terms -->
-
 <!-- Mor edescriptive fgure captions -->
 
-<!-- Add something about model diagnostics -->
-
-<!-- Something about priors -->
-
-<!-- Figure 3 order by tongue length class -->
-
 <!-- Describe the differences in models in the Figure2 and 3 captions -->
-
-<!-- Explain filtering of ordination plot -->
 
@@
 Visitation probability was then modeled using multilevel Bayesian regression with a Bernoulli distribution family (logit link function).
-*Bumble bee abundance* and *percent flower cover* (per morphotype) were included as a global interaction term, thus conditioning visitation probability on abundance and enabling the inference of bias.
-*Floral morphotype* was specified as a varying effect term with both a varying intercept and a varying slope with respect to *percent flower cover*, and a varying intercept term was also included for *site*.
-See Appendix X for the details of model specification and evaluation.
+*Bumble bee abundance* and *percent flower cover* (per morphotype) were included as a global interaction term, thus conditioning visitation probability on abundance and enabling the inference of preference independent of the neutral effects of abundance.
+*Floral morphotype* was specified as a both a varying intercept term and a varying slope term with respect to *percent flower cover*.
+*Site* was included as a varying intercept term to avoid pseudoreplication owing to the repeated observations within each site.
 
@@
 
-To explore the functional underpinnings of bumble foraging bias, we analyzed visitation probability as a function of the interaction between floral morphology and bumble bee tongue length class.
+To explore the functional underpinnings of bumble foraging preference, we analyzed visitation probability as a function of the interaction between floral morphology and bumble bee tongue length class.
 For this analysis, visitation was pooled within floral morphotype but not across bumble bee species.
 We then specified a Bayesian model of the same form as Model 1, but this time included a varying slope and intercept term for floral morphotype : bumble bee tongue length class and floral morphotype : bumble bee species.
-In this model, the relationship between floral morphotype and bumble bee tongue length class was the focus of inference, and the bumble bee species effect was added to avoid pseudoreplication arising from the grouping of observation within species.
-See Appendix X for the details of model specification and evaluation.
+In this model, the relationship between floral morphotype and bumble bee tongue length class was the focus of inference, and the bumble bee species effect was added to account for the grouping of observations within species.
 
@@
 
-To focus on species-level morphotype biases, we removed the tongue length effect from Model 2 and re-fit the model considering only bumble bee species.
-See Appendix X for the details of model specification and evaluation.
+To focus on species-level morphotype preferences, we removed the tongue length effect from Model 2 and re-fit the model considering only bumble bee species.
 
@@
 
-To visualize the net diet similarity between bumble bee species, we pooled visitation data across all samples within bumble bee species and plant species, then performed a principal coordinates analysis based on Horn-Morisita distance metric.
+To visualize the overall diet similarity between bumble bee species, we pooled visitation data across all samples within bumble bee species and plant genus, then performed a principal coordinates analysis based on Horn-Morisita distance metric.
 To test the significance of tongue length class as a predictor of diet similarity, we performed a permutation test.
+Importantly, this approach treats the whole study system in aggregate, thus conflating the causal effects of floral preference, seasonal activity patterns, and site occupancy patterns in the generation of overall diet similarity.
+In this sense, it can be understood as a model of potential diet overlap given the aggregate foraging patterns of the bumble bee species in our system, with the understanding that realized niche overlap at particular locations or times may vary.
 
@@
 
-We modeled niche overlap as a function of level (morphotype, family, genus) and elevation using hierarchical generalized additive models (HGAMs).
+To investigate elevational patterns of *within-site* niche overlap, we modeled niche overlap as a function of level (morphotype, family, genus) and elevation using HGAMs.
 First, we calculated community mean niche overlap by taking the mean niche overlap (within level) across all species pairs for each site.
@@
 
-## Software
-
-All analyses were conducted in `R` `r citep(citation())`.
-Bayesian multilevel models were implemented in Stan (Stan Development Team 2021) via the package `brms` `r citep(citation("brms"))`, and visualizations were constructed with the packages `tidybayes` `r citep(citation("tidybayes"))` and `ggdist` `r citep(citation("ggdist"))`.
-Ordination analysis and permutation testing were performed with the package `vegan` `r citep(citation("vegan"))` and visualized using the packages `ggplot2` `vegan` `r citep(citation("ggplot2"))` and `ggvegan` (Simpson 2019).
-HGAMs were implemented with the package `mgcv` [@Wood_2017] and visualized with the package `mgcViz` [@Fasiolo2018aa].
-Data handling and visualization was performed with the `tidyverse` suite of R packages `r citep(citation("tidyverse"))`.
-Annotated R code is available in the Supplementary Material.
-
 # RESULTS
@@
 ## Summary of visitation and floral survey data
-
-<!-- Modify this section to describe how many observations/samples were dropped due to the non-alignment of visitation and survey data -->
 
@@
 
-## Overall visitation probability by floral morphotype
+## Floral preference (Models 1-3)
 
-Lip (group 5) and flag (group 6) flowers had the highest overall visitation probability, with a conditional mean visitation probabilities of 0.36 (95% CI = 0.30–0.42) and 0.33 (95% CI = 0.27–0.39), respectively (Figure 1).
-In second place, head flowers (group 7) had a mean conditional visitation probability of 0.19 (95% CI = 0.16–0.24).
-Bell (group 3) flowers were a distant third with a conditional mean visitation probability of 0.092 (95% CI = 0.07–0.12), followed by stalk-disc (group 4), funnel (group 2), and disc flowers (group 1) with conditional visitation probabilities of 0.07 (95% CI = 0.05–0.09), 0.05 (95% CI = 0.04–0.07), and 0.05 (95% CI = 0.03–0.07), respectively.
-Unclassified (group 0) flowers, consisting mainly of wind-pollinated grasses, were the least-visited, with a conditional mean visitation probability of 0.03 (95% CI = 0.02–0.04).
-Our model explained approximately 35% of the total variance (97.5% CI = 34–36%).
+Our visitation models (Models 1-3) revealed strong morphotype preferences across the aggregate bumble bee community and marked differences in floral preference between bumble bee species.
+These differences in preference were partly explained by differences in tongue length, but certain floral morphotypes generated divergent preference patterns among bumble bee species with similar tongue lengths.
+
+Lip (group 5) and flag (group 6) flowers had the highest overall visitation probability, with a conditional mean visitation probabilities of 0.35 (95% CI = 0.27–0.4) and 0.31 (95% CI = 0.24–0.40), respectively (Figure 1).
+In second place, head flowers (group 7) had a mean conditional visitation probability of 0.18 (95% CI = 0.13–0.25).
+Bell (group 3) flowers were a distant third with a conditional mean visitation probability of 0.09 (95% CI = 0.06–0.13), followed by stalk-disc (group 4), funnel (group 2), and disc flowers (group 1) with conditional visitation probabilities of 0.06 (95% CI = 0.04–0.08), 0.04 (95% CI = 0.03–0.07), and 0.04 (95% CI = 0.02–0.06), respectively.
+Model 1 explained approximately 37% of the total variance (97.5% CI = 35–38%).
 
@@
 
-## Visitation probability by flower morphotype, bumble bee species, and bumble bee tongue length class
-
-Descriptions of tongue classes are based on Model 2 and refer to Figure 2, and descriptions of species are based on Model 3 and refer to Figure 3.
-Both models explained approximately 28% of total variance (97.5% CI = 27–29%).
-As in Model 1, visitation rates refer to conditional visitation probabilities.
-
-Disc (group 1) flowers were visited at similar rates across tongue length classes.
+Disc (group 1) flowers were visited at similar rates across tongue length classes **(Figure 2)**.
 At the species level, disc flowers were visited at relatively high rates by the short-tongued B.
-terrestris/lucorum and the medium-tongued Bombus monticola.
+terrestris/lucorum and the medium-tongued Bombus monticola **(Figure 3)**.
 Visitation to funnel (group 2) flowers was dominated by long-tongued species, an effect driven entirely by the Aconitum (group 2) specialist B.
@@
 gerstaeckeri.
+Models 2 and 3 both explained approximately 31% of total variance (97.5% CI = 30–32%).
 
@@
 Note that the range of the x-axis has been rescaled for each panel to emphasize the comparison of visitation probabilities across tongue length classes.")}
-knitr::include_graphics("../output/brm_pvis00_plot1.pdf")
+knitr::include_graphics("../output/brm_pvis00_plot1.png")
 ```
@@
 Note that the range of the x-axis has been rescaled for each panel to emphasize the comparison of visitation probabilities across bumble bee species")}
-knitr::include_graphics("../output/brm_pvis02_plot1.pdf")
+knitr::include_graphics("../output/brm_pvis02_plot1.png")
 ```
 
-## Diet ordination
+## Aggregate diet similarity
 
+Principal coordinates analysis (Model 4) suggested weak but significant effects of tongue length on the diet similarity while also highlighting the particular floral genera that accounted most strongly for the dietary separation of bumble bee species.
 Axis 1 accounted for 24% of total diet variance and primarily separated long- and medium-tongued species from short-tongued species (Figure 4).
-Long- and medium-tongued species scored low on Axis one and were associated with the lip flowers (group 5) Ajuga reptans, Rhinanthus glacialis, and Stachys alopecuros) and flag flowers (group 6) Anthyllis vulneraria, Lotus corniculatus, and Trifolium pratense.
-Short tongued species scored high on Axis one and were associated with the bell flowers (group 3) Campanula scheuchzeri, Erica carnea, and Geum rivale and the head flowers (group 7) Carduus defloratus, Centaurea jacea, and Phyteuma orbiculare.
+Long- and medium-tongued species scored low on Axis one and were associated with the lip flowers (group 5) *Ajuga*, *Rhinanthus*, and *Stachys* and flag flowers (group 6) *Anthyllis*, *Lotus*, and *Trifolium*.
+Short tongued species scored high on Axis one and were associated with the bell flowers (group 3) *Campanula*, *Erica*, and *Geum* and the head flowers (group 7) *Carduus*, *Centaurea*, and *Phyteuma*.
 Axis 2 accounted for 19% of total variance and primarily explained variation within short-tongued species.
-Short-tongued species scoring high on Axis 2 were associated with Campanula scheuchzeri while those scoring low on Axis 2 were associated with Erica carnea.
+Short-tongued species scoring high on Axis 2 were associated with *Campanula* while those scoring low on Axis 2 were associated with *Erica*.
 Short-tongued species scoring close to zero on Axis 2 also scored lower on Axis 1 (closer to medium- and long-tongued species) and were associated with head flowers.
-Tongue length class explained 22% of diet variance (Adonis permutation test, F = 1.68, p = 0.006).
-While short-tongued bees appeared most broadly spread in ordination space, there were no significant differences in dispersion across tongue length classes (F = 0.62, p = 0.55).
+Tongue length class explained 22% of diet variance (Adonis permutation test, F = 1.68, p = 0.005).
+While short-tongued bees appeared most broadly spread in ordination space, there were no significant differences in dispersion across tongue length classes (F = 0.26, p = 0.62).
 
-```{r fig5, echo = FALSE,  fig.cap = paste("Principal coordinates ordination of bumble bee visitation pooled across all samples.")}
+```{r fig5, echo = FALSE,  fig.cap = paste("Principal coordinates ordination of bumble bee visitation pooled across all samples.
+To avoid overplotting, only plant genera with an absolute score of at least 0.22 on one or both axes are plotted.")}
 knitr::include_graphics("../output/pcoa_aggregate.png")
@@
 
-## Niche overlap
+## Site-level diet overlap (Models 5 and 6)
 
-### Niche overlap through elevation by level
+Site-level analysis of diet overlap revealed strong differences in the magnitude of diet overlap at the nest levels of floral morphotype, floral family, and floral genus, as well as consistent patterns in response to elevation.
+At the site level, however, bumble bees belonging to the same tongue length class did not have significantly more similar diets than those belonging to different tongue length classes.
 
-Morphotype-level niche overlap was nearly double genus-level niche overlap throughout the whole elevation gradient, with family-level niche overlap taking intermediate values.
-The fixed intercepts of morphotype- and genus-level niche overlap each differed significantly (in opposite directions) from the reference level of family-level niche overlap (p << 0.001).
+Morphotype-level diet overlap was nearly double genus-level diet overlap throughout the whole elevation gradient, with family-level diet overlap taking intermediate values.
+The fixed intercepts of morphotype- and genus-level diet overlap each differed significantly (in opposite directions) from the reference level of family-level diet overlap (p << 0.001).
 The conditional smooths of each level were significantly different from flat lines (p < 0.001) and generally mirrored one another in shape **(Figure 8)**.
-Niche overlap was highest at the low end of the elevation gradient, decreased steeply from 600-1200 m, and leveled off or declined weakly for the remainder of the gradient, though morphotype-level niche overlap exhibited a second peak just above 1500 m that was not evident in genus- or family-level niche overlap.
-The model explained 78.4% of total deviance (adjusted R^2^ = 0.743).
+Diet overlap was highest at the low end of the elevation gradient, decreased steeply from 600-1200 m, and leveled off or declined weakly for the remainder of the gradient, though morphotype-level diet overlap exhibited a second peak just above 1500 m that was not evident in genus- or family-level diet overlap.
+The Model 5 explained 78.4% of total deviance (adjusted R^2^ = 0.743).
 
@@
 
-### Genus-level niche overlap by tongue length and elevation
-
-Whether a species pair belonged to the same or different tongue length classes had no effect on genus-level niche overlap, either in terms of intercept (p = 0.18) or response to elevation **(Figure 9)**.
-The model explained 36.5% of total deviance (adjusted R^2^ = 0.316).
+Whether a species pair belonged to the same or different tongue length classes had no effect on genus-level niche overlap, either in terms of intercept (p = 0.18) or shape of response to elevation **(Figure 9)**.
+Model 6 explained 36.5% of total deviance (adjusted R^2^ = 0.316).
 
@@
 
-## Foraging bias and floral morphology
+The mystery of bumble bee species coexistence hinges on the relationship between species-level floral preferences, trait-matching, and the emergent phenomenon of community-level patterns of diet overlap.
+In this study, we show that while tongue length is a relatively strong predictor of species-level floral preferences, it is at best a weak predictor of diet overlap.
+This is due in part to divergent preference of bumble bees with similar tongue lengths toward certain floral morphotypes.
+Perhaps more importantly, though, diet overlap at the level of floral genus tended to be about half as strong as diet overlap at the level of floral morphotype, indicating that bumble bees with similar morphological affinities, likely associated with tongue length, subdivide floral morphotypes at the genus- and species-level, reducing the potential for competition.
 
-Our study is not the first to compare bumble bee visitation rates across Kugler’s (1970) floral morphotypes (see Neumayer and Paulus 1999; Schneller et al.
-2014), but to our knowledge it is the first to enable the inference of bias across morphotypes by controlling statistically for the neutral effects of abundance.
+As discussed earlier, the interpretation of the patterns in our data hinge strongly on the extent to which interspecific competition actually occurs in our study system.
+We have argued that the assumption of minimal competition is reasonable in the context of our study, owing to the rich floral resources and harsh abiotic conditions of the Berchtesgadener Alps, together with the manifest co-occurrence of species with similar tongue length.
+This leads to the interpretation of abundance-conditioned visitation rates as *preferences* and diet overlap as *potential competition*.
+To the extent that interspecific competition does occur in our system, though, visitation rates would represent floral preference distorted by the pressure of competition toward niche partitioning, rendering diet overlap an underestimate of potential competition.
+While we maintain, based on our knowledge of our study system, that the first interpretation is more likely, both interpretations are plausible and must be kept in mind.
+
+and resulting patterns of dietary overlap.
+In this study, we demonstrate that while bumble bees are, indeed, remarkably generalized in their foraging, they are highly discriminating generalists, over-representing certain floral taxa and morphotypes and ignoring others.
+These patterns of floral preference are related to bumble bee tongue length, but there exists considerable variation in floral preference among species with similar tongue lengths, which may serve to relax competition between them.
+
+In the cases of disc and stalk-disc flowers visitation rates were uniformly low across all tongue length classes.
+The opposite was true of lip flowers, which were visited at similarly high rates by all tongue length classes.
+Bell flowers and head flowers were visited chiefly by short-tongued bees, consistent with the shallow corollas of these morphotypes.
+Conversely, funnel flowers, with their deep corollas were visited chiefly by long-tongued bees, though this effect was driven disproportionately by *B.
+gerstaeckeri*, which specializes on the funnel flowers of the genus *Aconitum* [@Ponchau2006-af].
+Flag flowers were visited at the highest rate by medium-tongued species, and especially by *B.
+pascuorum* and *B.
+mucidus*.
+
+If not tongue length, what else could explain the variation in floral morphotype preference witin tongue length classes?
 
 When bumble bee visitation is pooled across all species, it is clear that lip, flag, and head flowers dominate the floral community in terms of conditional visitation probability.
-Patterns of pooled visitation are, of course, driven disproportionately by the behavior of the most abundant bumble bee species, which in our system were B.
-hortorum, B.
-pascuorum, B.
-pratorum, B.
-psithyrus, B.
-soroeensis, B.
-terrestris/lucorum, and B.
-wurflenii.
-Visitation to lip and flag flowers was driven especially by the medium-tongued B.
-pascuorum, the most abundant bumble bee in our study system, and secondarily by the medium-tongued B.
-wurflenii.
-Visitation to head flowers was driven mainly by the short-tongued B.
-soroeensis and B.
-psithyrus.
+Patterns of pooled visitation are, of course, driven disproportionately by the behavior of the most abundant bumble bee species, which in our system were *B.
+pascuorum*, *B.
+pratorum*, *B.
+soroeensis*, *B.
+terrestris/lucorum*, and *B.
+wurflenii* (Sponsler et al., submitted).
+Visitation to lip and flag flowers was driven especially by the medium-tongued *B.
+pascuorum*, the most abundant bumble bee in our study system, and secondarily by the medium-tongued *B.
+wurflenii*.
+Visitation to head flowers was driven mainly by the short-tongued *B.
+soroeensis*.
 
 The influence of tongue length on visitation bias was pronounced for some morphotypes but not for others.
-In the cases of disc, and stalk-disc, and unclassified flowers, visitation rates were uniformly low across all tongue length classes.
+In the cases of disc and stalk-disc flowers visitation rates were uniformly low across all tongue length classes.
 The opposite was true of lip flowers, which were visited at similarly high rates by all tongue length classes.
 Bell flowers and head flowers were visited chiefly by short-tongued bees, consistent with the shallow corollas of these morphotypes.
-In contrast, funnel flowers, with their deep corollas were visited chiefly by long-tongued bees, though this effect was driven disproportionately by B.
-gerstaeckeri, which specializes on the funnel flowers of the genus Aconitum.
-Flag flowers were visited at the highest rate by medium-tongued species, and especially by B.
-pascuorum and B.
-mucidus.
+Conversely, funnel flowers, with their deep corollas were visited chiefly by long-tongued bees, though this effect was driven disproportionately by *B.
+gerstaeckeri*, which specializes on the funnel flowers of the genus *Aconitum* [@Ponchau2006-af].
+Flag flowers were visited at the highest rate by medium-tongued species, and especially by *B.
+pascuorum* and *B.
+mucidus*.
 
@@
 Nectar robbing is one mechanism by which the relationship between tongue length and corolla depth might be relaxed, but its influence in our data is unclear.
-Of the bumble bees in our study system, B.
-terrestris/lucorum and especially B.
-wurflenii are known to engage frequently in nectar robbing (Brian 1954).
-In the case of B.
-wurflenii, however, its morphotype affinities did not differ markedly from those of other medium-tongued species, and it may be that the condition of medium-tongue length alone confers enough to foraging flexibility in our study system to dampen the divergent effects of nectar robbing.
-Nectar robbing might, however, account for the deviation of B.
-terrestris/lucorum from most other short-tongued species with respect to flag flowers.
-With the exception of B.
-lapidarius, all short-tongued species besides B.
-terrestris/lucorum eschewed flag flowers, while B.
-terrestris/lucorum visited them at rates comparable to medium- and long-tongued species.
+Of the bumble bees in our study system, *B.
+terrestris/lucorum* and especially *B.
+wurflenii* are known to engage frequently in nectar robbing [@Brian1954-cy].
+In the case of *B.
+wurflenii*, however, its morphotype affinities did not differ markedly from those of other medium-tongued species, and it may be that the condition of medium-tongue length alone confers enough to foraging flexibility in our study system to dampen the divergent effects of nectar robbing.
+Nectar robbing might, however, account for the deviation of *B.
+terrestris/lucorum* from most other short-tongued species with respect to flag flowers.
+With the exception of *B.
+lapidarius*, all short-tongued species besides *B.
+terrestris/lucorum* eschewed flag flowers, while *B.
+terrestris/lucorum* visited them at rates comparable to medium- and long-tongued species.
 Indeed, B.
-terrestris/lucorum was, on the whole, an idiosyncratic forager, demonstrating a relatively high affinity for disc flowers and unclassified flowers (particularly Plantago lanceolata) that were largely ignored by other species.
+terrestris/lucorum was, on the whole, an idiosyncratic forager, also demonstrating a relatively high affinity for the disc flowers that were largely ignored by other species.
 Perhaps the most striking divergence of floral bias within tongue length class, though, was the divisive effect of bell flowers on short-tongued bumble bees.
-B.
-pyrenaeus, B.
-soroeensis, B.
-pratorum, and B.
-jonellus visited bell flowers at high rates, while B.
-terrestris/lucorum, B.
-psithyrus, and B.
-lapidarius visited them relatively rarely.
-A similar split among the short-tongued species occurred with respect to head flowers, which were visited at high rates by B.
-soroeensis and B.
-psithyrus but at relatively low rates by other short-tongued species.
-There are no obvious morphological distinctions within short-tongued bumble bees that would account for these divergent affinities; ecological drivers such as competition or differences in spatial or temporal activity patterns are more likely explanations, as will be revisited below in our discussion of overall diet similarity.
+*B.
+pyrenaeus*, *B.
+soroeensis*, *B.
+pratorum*, and *B.
+jonellus* visited bell flowers at high rates, while *B.
+terrestris/lucorum*, *B.
+psithyrus*, and *B.
+lapidarius* visited them relatively rarely.
+A similar split among the short-tongued species occurred with respect to head flowers, which were visited at high rates by *B.
+soroeensis* and *B.
+psithyrus* but at relatively low rates by other short-tongued species.
+There are no obvious morphological distinctions within short-tongued bumble bees that would account for these divergent affinities, and ecological drivers such as differences in spatial or temporal activity patterns are more likely explanations.
 
-## Diet overlap
+Our finding that bumble bee diet ordination produces a primary axis corresponding to tongue length (and associated with affinity for legumes) and a secondary axis describing the dispersion of short-tongued species is consistent with studies in Poland [@Goulson2008-oe], France [@Durieux2000-uk], and England [@Goulson2005-nz] that analyzed bumble bee diet by similar methods, suggesting that this may be a general pattern among bumble bee communities, at least within temperate Europe.
+While our test of beta-dispersion did not reveal significant differences between tongue length classes, the spread of short-tongued bees along the secondary axis of our ordination, together with their divergence with respect to visitation rates on bell and head flowers described above, suggests that resource partitioning may be more pronounced among short-tongued bumble bees than among their medium- and long-tongued counterparts.
 
-We analyze community mean niche overlap as an indication of diffuse competition [@Pianka1974-qk; MacArthur1972-qz]
-
-Our finding that bumble bee diet ordination produces a primary axis corresponding to tongue length — and associated with affinity for legumes — and a secondary axis describing the dispersion of short-tongued species is consistent with studies in Poland (Dave Goulson, Lye, and Darvill 2008), France (Durieux 2000), and England (D.
-Goulson et al.
-2005) that analyzed bumble bee diet by similar methods, suggesting that this may be a general pattern among bumble bee communities, at least within temperate Europe.
-While our test of beta-dispersion did not reveal significant differences between tongue length classes, the spread of short-tongued bees along the secondary axis of our ordination, together with their divergence with respect to visitation rates on bell and head flowers described above, suggests that resource partitioning may be more pronounced among short-tongued bumble bees than among their medium- and long-tongued counterparts.
 It is tempting to invoke interspecific competition as a cause of the apparent niche partitioning among short-tongued bumble bees, in agreement with previous studies that have demonstrated the potential for severe competition between bumble bees of similar tongue length (e.g.
@@
 2014) or spatial configurations of resource patches (Sowig 1989), and their divergent morphological biases arose secondarily from these effects.
+
+## Diet overlap
+
+We analyze community mean niche overlap as an indication of diffuse competition [@Pianka1974-qk; MacArthur1972-qz]
 
@@
 
+The aim of our study was to build an analysis of bumble bee diet overlap from the ground up, as it were, beginning with a description of floral preferences, modeling these preferences with respect to bumble bee and floral traits, and
+
+Our study begins at the level of bumble bee floral visitation patterns, which, when conditioned on the neutral effects of abundance, can be interpreted as floral preferences.
+These floral preferences, together with potential effects of spatial and temporal co-occurrence,
+
 # ACKNOWLEDGEMENTS

### 2021-03-09 ms_2/ms_2_v2.Rmd
did some writing and added some references

@@
 Nevertheless, bumble bee communities with an apparent excess of species richness and deficit of trait (or phylogenetic) dispersion are commonly observed [@Ranta1981-vx; @Goulson2008-zy].
+To account for this inconsistency, @Ranta1981-vx proposed that bumble bees communities are governed by a spatiotemporally heterogeneous mosaic of competitive conditions, such that competitive inequalities between species are sufficiently variable to allow more species to coexist than predicted by trait-based competition models.
+A complementary explanation, based on the core/satellite species hypothesis [@Hanski1982-hs], is that competitive forces are significant only among regionally abundant and widespread “core species”, and the effects of competition are more pronounced in the structure of regional species pools than in that of local assemblages [@Hanski1982-ie].
+In some cases, competition may simply be precluded by a saturating abundance of floral resources [@Williams1989-vr].
+
 Teasing apart the still-obscure mechanisms of bumble bee coexistence and niche partitioning remains a salient problem from a basic ecological perspective, but it has also become an urgent question in the context of applied conservation.
 Bumble bees are cold-adapted species [@Heinrich1994-cv] associated mainly with mountain habitats, where they serve as the principal pollinators of many alpine plants [@Williams1998-xi; @Goulson2010-lo].
-Warming climates have already caused measurable upslope shifts in bumble bees and their floral hosts [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc], and these are predicted to cause declines or even extinction in many bumble bees species [@Soroye2020-zb].
-One mechanism by which climate change is expected to affect bumble bees is the introduction of novel interspecific competition driven by the upslope advance of lowland bumble bee species into the elevational distributions of their highland counterparts.
+Warming climates have already caused measurable upslope shifts in bumble bees and their floral hosts [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc], and these are predicted to cause declines in many bumble bees species [@Soroye2020-zb].
+Moreover, the introduction or extinction of species from bumble bee communities, with corresponding addition and release of competitive pressure, can alter both the floral visitation patterns of bumble bees and, consequently, the pollination biology of their host plants [@Brosi2013-mu; @Ishii2013-sm].
 Thus, the speculative question of how competition has acted in the historic assembly of bumble bee communities has become an unfolding process observable over ecological time scales and instrumental in the future viability of bumble bee species and their floral mutualists.
@@
 
-Kugler morphotype classifications of the plants in our study system were accessed via the BIOFLOR database (Klotz, Kühn, and Durka 2002).
+Kugler morphotype classifications of the plants in our study system were accessed via the BIOFLOR database [@Klotz2002-uc].
 Morphotypes were simplified to the primary classes numbered 0-10: 0 = not applicable (includes wind-pollinated flora such as grasses), 1 = disc- and bowl-shaped flowers, 2 = funnel flowers, 3 = bell-shaped flowers, 4 = stalk-disc flowers, 5 = lip flowers, 6 = flag flowers, 7 = flower heads, 8 = spike flowers, 9 = brush flowers, 10 = trap flowers **(Figure 1)**.
@@
 The genus *Gentiana* consisted of both group 2 and group 4 representatives, so the missing species *Gentiana aspera* and *Gentiana ciliata* were classified based on visual comparison with their classified congeners.
-In the case of the genus *Phyteuma*, we followed Neumayer and Paulus (1999) and classified it as a group 7 flower head.
+In the case of the genus *Phyteuma*, we followed @Neumayer1999-il and classified it as a group 7 flower head.
 Morphotypes 9 and 10 were extremely rare in our study system and were therefore omitted from all analyses.
@@
 
-Bumble bee tongue-length data were obtained from the literature values compiled and summarized by Arbetman et al.
-(2017).
+Bumble bee tongue-length data were obtained from the literature values compiled and summarized by @Arbetman2017-xj.
 Values for *B.
 gerstaeckeri* and *B.
-mendax*, which were not included in Arbetman et al.
-(2017), were obtained from Obeso (1992) and Durieux (2000), respectively.
+mendax*, which were not included in @Arbetman2017-xj, were obtained from @Obeso1992-yp and @Durieux2000-uk, respectively.
 We then binned bumble bee species into three tongue-length classes: short (< 7.5 mm), medium (7.5—10.0 mm), and long (>= 10.0 mm).
@@
 All analyses were conducted in `R` `r citep(citation())`.
-Data handling and visualization was performed with the `tidyverse` suite of R packages `r citep(citation("tidyverse"))`.
-Bayesian multilevel models were implemented in Stan (Stan Development Team 2021) via the package `brms` `r citep(citation("brms"))` using default priors, and visualizations were prepared with the packages `tidybayes` `r citep(citation("tidybayes"))` and `ggdist` `r citep(citation("ggdist"))`.
-Ordination analysis and permutation testing were performed with the package `vegan` `r citep(citation("vegan"))` and visualized using the packages `ggplot2` `vegan` `r citep(citation("ggplot2"))` and `ggvegan` (Simpson 2019).
+Data handling and visualization was performed with the `tidyverse` suite of R packages [@Wickham2019-aa].
+Bayesian multilevel models were implemented in Stan [@Stan2020aa] via the package `brms` [@Burkner2017-aa] using default priors, and visualizations were prepared with the packages `tidybayes` `r citep(citation("tidybayes"))` and `ggdist` `r citep(citation("ggdist"))`.
+Ordination analysis and permutation testing were performed with the package `vegan` `r citep(citation("vegan"))` and visualized using the packages `ggplot2` [@Wickham2016-aa] and `ggvegan` [@Simpson2019aa].
 Hierarchical generalized additive models (HGAMs) were implemented with the package `mgcv` [@Wood_2017] and visualized with the packages `mgcViz` [@Fasiolo2018aa] and `tidymv` [@Coretta-aa2021].
@@
 
-To investigate elevational patterns of *within-site* niche overlap, we modeled niche overlap as a function of level (morphotype, family, genus) and elevation using HGAMs.
-First, we calculated community mean niche overlap by taking the mean niche overlap (within level) across all species pairs for each site.
+To investigate elevational patterns of *within-site* niche overlap, we modeled community-mean niche overlap as a function of level (morphotype, family, genus) and elevation using HGAMs.
+Community mean niche overlap was calculated by taking the mean niche overlap (within level) across all species pairs for each site.
 Level was included as both a fixed intercept effect and a smoothing factor in conjunction with elevation, following the type "I" model form described by @Pedersen2019-kf.
@@
 Disc (group 1) flowers were visited at similar rates across tongue length classes **(Figure 2)**.
-At the species level, disc flowers were visited at relatively high rates by the short-tongued B.
-terrestris/lucorum and the medium-tongued Bombus monticola **(Figure 3)**.
-Visitation to funnel (group 2) flowers was dominated by long-tongued species, an effect driven entirely by the Aconitum (group 2) specialist B.
-gerstaeckeri.
+At the species level, disc flowers were visited at relatively high rates by the short-tongued *B.
+terrestris/lucorum* and the medium-tongued *Bombus monticola* **(Figure 3)**.
+Visitation to funnel (group 2) flowers was dominated by long-tongued species, an effect driven entirely by the *Aconitum* (group 2) specialist *B.
+gerstaeckeri*.
 Bell (group 3) flowers were visited at a higher rate by short-tongued species than by long-tongued species, with medium-tongued species exhibiting intermediate visitation rates.
-At the species-level, bell flowers were visited at highest probability by the short-tongued species B.
-jonellus, B.
-pyrenaeus, and B.
-soroeensis, and by the medium-tongued species B.
-hypnorum and B.
-monticola.
+At the species-level, bell flowers were visited at highest probability by the short-tongued species *B.
+jonellus*, *B.
+pyrenaeus*, and *B.
+soroeensis*, and by the medium-tongued species *B.
+hypnorum* and *B.
+monticola*.
 Stalk-disc (group 4) flowers were visited at similar rates across all tongue length classes, and there was also broad overlap in visitation rates at the species level.
-Lip (group 5) flowers were also visited at similar rates across tongue length classes, but at the species level B.
-pascuorum exhibited an uniquely high bias toward lip flowers.
+Lip (group 5) flowers were also visited at similar rates across tongue length classes, but at the species level *B.
+pascuorum* exhibited an uniquely high bias toward lip flowers.
 Flag (group 6) flowers were visited at a higher rate by medium-tongued species than by short-tongued species, with long-tongued species exhibiting intermediate visitation rates.
-Flag flowers generated strong separation in visitation rate at the species level, receiving very high visitation from the medium-tongued B.
-pascuorum and B.
-mucidus, moderately high visitation from the short-tongued B.
-terrestris/lucorum and B.
-lapidarius, the medium-tongued B.
-monticola, and the long-tongued B.
-mendax and B.
-hortorum, and low visitation by the short-tongued B.
-soroeensis, B.
-pyrenaeus, B.
-psithyrus, B.
-pratorum, and B.
-jonellus, the medium-tongued B.
-hypnorum, and the long-tongued B.
-gerstaeckeri.
+Flag flowers generated strong separation in visitation rate at the species level, receiving very high visitation from the medium-tongued *B.
+pascuorum* and *B.
+mucidus*, moderately high visitation from the short-tongued *B.
+terrestris/lucorum* and *B.
+lapidarius*, the medium-tongued *B.
+monticola*, and the long-tongued *B.
+mendax* and *B.
+hortorum*, and low visitation by the short-tongued *B.
+soroeensis*, *B.
+pyrenaeus*, *B.
+psithyrus*, *B.
+pratorum*, and *B.
+jonellus*, the medium-tongued *B.
+hypnorum*, and the long-tongued *B.
+gerstaeckeri*.
 Models 2 and 3 both explained approximately 31% of total variance (97.5% CI = 30–32%).
@@
 
-The mystery of bumble bee species coexistence hinges on the relationship between species-level floral preferences, trait-matching, and the emergent phenomenon of community-level patterns of diet overlap.
-In this study, we show that while tongue length is a relatively strong predictor of species-level floral preferences, it is at best a weak predictor of diet overlap.
-This is due in part to divergent preference of bumble bees with similar tongue lengths toward certain floral morphotypes.
-Perhaps more importantly, though, diet overlap at the level of floral genus tended to be about half as strong as diet overlap at the level of floral morphotype, indicating that bumble bees with similar morphological affinities, likely associated with tongue length, subdivide floral morphotypes at the genus- and species-level, reducing the potential for competition.
-
-As discussed earlier, the interpretation of the patterns in our data hinge strongly on the extent to which interspecific competition actually occurs in our study system.
-We have argued that the assumption of minimal competition is reasonable in the context of our study, owing to the rich floral resources and harsh abiotic conditions of the Berchtesgadener Alps, together with the manifest co-occurrence of species with similar tongue length.
-This leads to the interpretation of abundance-conditioned visitation rates as *preferences* and diet overlap as *potential competition*.
-To the extent that interspecific competition does occur in our system, though, visitation rates would represent floral preference distorted by the pressure of competition toward niche partitioning, rendering diet overlap an underestimate of potential competition.
-While we maintain, based on our knowledge of our study system, that the first interpretation is more likely, both interpretations are plausible and must be kept in mind.
-
-and resulting patterns of dietary overlap.
-In this study, we demonstrate that while bumble bees are, indeed, remarkably generalized in their foraging, they are highly discriminating generalists, over-representing certain floral taxa and morphotypes and ignoring others.
-These patterns of floral preference are related to bumble bee tongue length, but there exists considerable variation in floral preference among species with similar tongue lengths, which may serve to relax competition between them.
-
-In the cases of disc and stalk-disc flowers visitation rates were uniformly low across all tongue length classes.
-The opposite was true of lip flowers, which were visited at similarly high rates by all tongue length classes.
-Bell flowers and head flowers were visited chiefly by short-tongued bees, consistent with the shallow corollas of these morphotypes.
-Conversely, funnel flowers, with their deep corollas were visited chiefly by long-tongued bees, though this effect was driven disproportionately by *B.
-gerstaeckeri*, which specializes on the funnel flowers of the genus *Aconitum* [@Ponchau2006-af].
-Flag flowers were visited at the highest rate by medium-tongued species, and especially by *B.
-pascuorum* and *B.
-mucidus*.
-
-If not tongue length, what else could explain the variation in floral morphotype preference witin tongue length classes?
-
-When bumble bee visitation is pooled across all species, it is clear that lip, flag, and head flowers dominate the floral community in terms of conditional visitation probability.
+While bumble bees are, indeed, remarkably generalized in their foraging, they are highly discriminating generalists, dramatically over-representing certain floral taxa and morphotypes and ignoring others.
+In our study system, lip, flag, and head flowers dominated overall visitation probability when visitation was pooled across all bumble bee species, while bell, disc, and stalk-disc flowers were underrepresented.
 Patterns of pooled visitation are, of course, driven disproportionately by the behavior of the most abundant bumble bee species, which in our system were *B.
@@
 
-The influence of tongue length on visitation bias was pronounced for some morphotypes but not for others.
-In the cases of disc and stalk-disc flowers visitation rates were uniformly low across all tongue length classes.
-The opposite was true of lip flowers, which were visited at similarly high rates by all tongue length classes.
-Bell flowers and head flowers were visited chiefly by short-tongued bees, consistent with the shallow corollas of these morphotypes.
-Conversely, funnel flowers, with their deep corollas were visited chiefly by long-tongued bees, though this effect was driven disproportionately by *B.
+Tongue length class was an important predictor of preference for some but not all floral morphotypes.
+For disc and stalk-disc flowers, visitation rates were uniformly low across all tongue length classes.
+The opposite was true of lip flowers, which were visited at high rates by all tongue length classes.
+Bell flowers and head flowers were preferred by short-tongued bees, consistent with the shallow corollas of these morphotypes.
+Conversely, funnel flowers, with their deep corollas were preferred by long-tongued bees, though this effect was driven disproportionately by *B.
 gerstaeckeri*, which specializes on the funnel flowers of the genus *Aconitum* [@Ponchau2006-af].
@@
 
-While the variation in floral bias across tongue length classes we observed largely comports with the established relationship between tongue length and corolla depth, variation in bias within tongue length classes demonstrates the insufficiency of simple morphological matching to explain bumble bee foraging patterns.
-Nectar robbing is one mechanism by which the relationship between tongue length and corolla depth might be relaxed, but its influence in our data is unclear.
+While the variation we observed in floral preference across tongue length classes largely comports with the established relationship between tongue length and corolla depth, variation in preference *within* tongue length classes echoes the findings of @Goulson2008-zy in demonstrating the insufficiency of simple morphological matching in explaining bumble bee foraging patterns.
+Perhaps the most striking divergence of floral preference within tongue length class was the divisive effect of bell flowers on short-tongued bumble bees.
+*B.
+pyrenaeus*, *B.
+soroeensis*, *B.
+pratorum*, and *B.
+jonellus* visited bell flowers at high rates, while *B.
+terrestris/lucorum*, *B.
+psithyrus*, and *B.
+lapidarius* visited them relatively rarely.
+A similar split among the short-tongued species occurred with respect to head flowers, which were visited at high rates by *B.
+soroeensis* and *B.
+psithyrus* but at relatively low rates by other short-tongued species.
+There are no obvious morphological distinctions within short-tongued bumble bees that would account for these divergent affinities, and ecological drivers such as differences in spatial or temporal activity patterns are more likely explanations [@Sowig1989-ld; @Schneller2014-ry].
+Nectar robbing is another mechanism by which the relationship between tongue length and corolla depth might be relaxed, but its influence in our data is unclear.
 Of the bumble bees in our study system, *B.
@@
 terrestris/lucorum was, on the whole, an idiosyncratic forager, also demonstrating a relatively high affinity for the disc flowers that were largely ignored by other species.
-Perhaps the most striking divergence of floral bias within tongue length class, though, was the divisive effect of bell flowers on short-tongued bumble bees.
-*B.
-pyrenaeus*, *B.
-soroeensis*, *B.
-pratorum*, and *B.
-jonellus* visited bell flowers at high rates, while *B.
-terrestris/lucorum*, *B.
-psithyrus*, and *B.
-lapidarius* visited them relatively rarely.
-A similar split among the short-tongued species occurred with respect to head flowers, which were visited at high rates by *B.
-soroeensis* and *B.
-psithyrus* but at relatively low rates by other short-tongued species.
-There are no obvious morphological distinctions within short-tongued bumble bees that would account for these divergent affinities, and ecological drivers such as differences in spatial or temporal activity patterns are more likely explanations.
+Together, these species-level idiosyncracies of preference within tongue length classes suggest a possible answer to the question of how bumble bees of similar tongue length can coexists, namely that factors other than tongue length are evidently at play in bumble bees' morphological affinities.
 
-Our finding that bumble bee diet ordination produces a primary axis corresponding to tongue length (and associated with affinity for legumes) and a secondary axis describing the dispersion of short-tongued species is consistent with studies in Poland [@Goulson2008-oe], France [@Durieux2000-uk], and England [@Goulson2005-nz] that analyzed bumble bee diet by similar methods, suggesting that this may be a general pattern among bumble bee communities, at least within temperate Europe.
-While our test of beta-dispersion did not reveal significant differences between tongue length classes, the spread of short-tongued bees along the secondary axis of our ordination, together with their divergence with respect to visitation rates on bell and head flowers described above, suggests that resource partitioning may be more pronounced among short-tongued bumble bees than among their medium- and long-tongued counterparts.
+With respect to diet overlap among bumble bees, our finding that bumble bee diet ordination produces a primary axis corresponding to tongue length (and associated with affinity for legumes) and a secondary axis describing the dispersion of short-tongued species is consistent with studies in Poland [@Goulson2008-oe], France [@Durieux2000-uk], and England [@Goulson2005-nz] that analyzed bumble bee diet by similar methods, suggesting that this may be a general pattern among bumble bee communities, at least within temperate Europe.
+Perhaps a more significant finding, though, is that community-mean diet overlap at the level of floral morphotype was consistently nearly twice as high as diet overlap at the level of floral genus.
+From a numerical perspective, this result is unsurprising simply because each morphotype comprises many genera.
+From an ecological perspective, though, this pattern suggests another answer to the question of how bumble bee species of similar tongue length can coexists.
+If tongue length primarily constrains bumble bee foraging at the coarse level of floral morphotype, taxonomic diversity within morphotypes creates space for bumble bees with similar tongue lengths to avoid actual dietary overlap, thus mitigating interspecific competition.
+Importantly, this suggests that taxonomic diversity *within* floral morphotypes may as important as morphotype diversity *per se* and overall floral abundance in facilitating the coexistence of bumble bee species, an insight which may be relevant to applied bumble bee conservation.
 
-It is tempting to invoke interspecific competition as a cause of the apparent niche partitioning among short-tongued bumble bees, in agreement with previous studies that have demonstrated the potential for severe competition between bumble bees of similar tongue length (e.g.
-Heinrich 1976; Inouye 1978; Pyke 1982).
-But the importance of competition in bumble bee communities is not universal (e.g.
-Ranta and Vepsäläinen 1981; Williams 1989), and in the rich wildflower meadows of our (sub)alpine system, bumble bee populations are more likely limited by harsh abiotic conditions than by food availablility.
-Alternatively, it may be that short-tongued bees in our study system differed from one another in their affinities for specific habitats (Schneller et al.
-2014) or spatial configurations of resource patches (Sowig 1989), and their divergent morphological biases arose secondarily from these effects.
+With respect to elevation, community mean niche overlap was highest at the lowest sites and declined sharply over the transition from valley floor to mid-slope.
+If this pattern of niche overlap moves upslope in response to climate change, it might result in increased competition potential at higher elevations, which could threaten high-elevation bumble bee species that currently exist under conditions of relatively low niche overlap.
+In our system, though, it must be noted that the lowest elevation sites are also the ones most influences by human land use, and it is possible that the elevation pattern evident in our data is driven by factors other than elevational climate variation.
 
-## Diet overlap
+<!-- If not tongue length, what else could explain the variation in floral morphotype preference witin tongue length classes? -->
 
-We analyze community mean niche overlap as an indication of diffuse competition [@Pianka1974-qk; MacArthur1972-qz]
+# CONCLUSION
 
-To account for this inconsistency, @Ranta1981-vx proposed that bumble bees communities are governed by a spatiotemporally heterogeneous mosaic of competitive conditions, such that competitive inequalities between species are sufficiently variable to allow more species to coexist than predicted by trait-based competition models.
-A complementary explanation, based on the core/satellite species hypothesis [@Hanski1982-hs], is that competitive forces are significant only among regionally abundant and widespread “core species”, and the effects of competition are more pronounced in the structure of regional species pools than in that of local assemblages [@Hanski1982-ie].
-In some cases, competition may simply be precluded by a saturating abundance of floral resources [@Williams1989-vr].
-Range shifts in bumble bees have already been documented [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc], and the introduction or extinction of species from bumble bee communities, with corresponding addition and release of competitive pressure, can alter both the floral visitation patterns of bumble bees and, consequently, the pollination biology of their host plants [@Brosi2013-mu; @Ishii2013-sm].
+The mystery of bumble bee species coexistence hinges on the relationship between species-level floral preferences, trait-matching, and the emergent phenomenon of community-level patterns of diet overlap.
+While tongue length is a relatively strong predictor of species-level floral preferences, it is at best a weak predictor of diet overlap.
+This is due in part to divergent preference of bumble bees with similar tongue lengths toward certain floral morphotypes.
+Perhaps more importantly, though, diet overlap at the level of floral genus tended to be about half as strong as diet overlap at the level of floral morphotype, indicating that bumble bees with similar morphological affinities, driven in part by tongue length compatibility, subdivide floral morphotypes at the genus- and species-level, reducing the potential for competition.
+This suggests that taxonomic floral diversity may be as important as overall floral abundance in enabling the coexistence of bumble bee species.
+Something about elevation...
 
-We see that community-mean diet overlap, indicative of the potential for diffuse competition, generally declines with elevation [@Pianka1974-qk].
+As discussed earlier, the interpretation of the patterns in our data hinge strongly on the extent to which interspecific competition actually occurs in our study system.
+We have argued that the assumption of minimal competition is reasonable in the context of our study, owing to the rich floral resources and harsh abiotic conditions of the Berchtesgadener Alps, together with the manifest co-occurrence of species with similar tongue length.
+This leads to the interpretation of abundance-conditioned visitation rates as *preferences* and diet overlap as *potential competition*.
+To the extent that interspecific competition does occur in our system, though, visitation rates would represent floral preference distorted by the pressure of competition toward niche partitioning, rendering diet overlap an underestimate of potential competition.
+While we maintain, based on our knowledge of our study system, that the first interpretation is more likely, both interpretations are plausible and must be kept in mind.
 
-"there  is  still  considerable  scope  for  niche  differentiation  among  species  with  tonguesthat  are  near-identical  in  length." @Goulson2008-zy
-
-# CONCLUSIONS
-
-For bumble bees, floral bias is a labile product of innate biases, the soft constraints of morphological compatibility, and the dynamic ecological context in which foraging occurs.
-By conditioning bumble bee visitation data upon concurrent floral abundance data across spatially and temporally replicated samples, we demonstrate the strength bumble bee floral bias, its functional underpinnings, its variation across species, and its lability within species in response to environmental variation.
-Such tuning of floral bias has obvious adaptive significance for bumble bees, which are characteristically — and most likely primordially — associated with mountain habitats (Williams 1985; Hines 2008), where steep temperature gradients and short growing seasons generate extreme spatial and temporal floral turnover.
-
-The aim of our study was to build an analysis of bumble bee diet overlap from the ground up, as it were, beginning with a description of floral preferences, modeling these preferences with respect to bumble bee and floral traits, and
-
-Our study begins at the level of bumble bee floral visitation patterns, which, when conditioned on the neutral effects of abundance, can be interpreted as floral preferences.
-These floral preferences, together with potential effects of spatial and temporal co-occurrence,
+<!-- We analyze community mean niche overlap as an indication of diffuse competition [@Pianka1974-qk; MacArthur1972-qz] -->
 

### 2021-03-10 ms_2/ms_2_v2.Rmd
back to lab computer

@@
 ```
+
+<!-- Don't forget: I need to fix the Gentiana/Gentianella/Gentianopsis nomenclature problem and rerun all the models before I finalize the paper.
+This won't change anything about the interpretation of the  analysis, but it needs to be done.
+It might make Gentiana show up in the PCoA.
+-->
 
@@
 All analyses were conducted in `R` `r citep(citation())`.
-Data handling and visualization was performed with the `tidyverse` suite of R packages [@Wickham2019-aa].
-Bayesian multilevel models were implemented in Stan [@Stan2020aa] via the package `brms` [@Burkner2017-aa] using default priors, and visualizations were prepared with the packages `tidybayes` `r citep(citation("tidybayes"))` and `ggdist` `r citep(citation("ggdist"))`.
-Ordination analysis and permutation testing were performed with the package `vegan` `r citep(citation("vegan"))` and visualized using the packages `ggplot2` [@Wickham2016-aa] and `ggvegan` [@Simpson2019aa].
+Data handling and visualization were performed with the `tidyverse` suite [@Wickham2019-aa].
+Bayesian multilevel models were implemented in Stan [@Stan2020aa] via the package `brms` [@Burkner2017-aa] using default priors, and visualizations were prepared with the packages `tidybayes` [@Kay2020-aa] and `ggdist` [@Kay2021-aa].
+Ordination analysis and permutation testing were performed with the package `vegan` [@Oksanen_2019] and visualized using the packages `ggplot2` [@Wickham2016-aa] and `ggvegan` [@Simpson2019aa].
 Hierarchical generalized additive models (HGAMs) were implemented with the package `mgcv` [@Wood_2017] and visualized with the packages `mgcViz` [@Fasiolo2018aa] and `tidymv` [@Coretta-aa2021].
@@
 With only three years of data, though, such annual inferences would be of little value, and we deemed them not worth the cost of added model complexity.
-
-<!-- State more explicityl how the models should be interpreted with respect to bias -->
 
@@
 To investigate elevational patterns of *within-site* niche overlap, we modeled community-mean niche overlap as a function of level (morphotype, family, genus) and elevation using HGAMs.
-Community mean niche overlap was calculated by taking the mean niche overlap (within level) across all species pairs for each site.
+Community mean niche overlap, which we interpret as an index of potential *diffuse* competition [@Pianka1974-qk; @MacArthur1972-qz], was calculated by taking the mean niche overlap (within level) across all species pairs for each site.
 Level was included as both a fixed intercept effect and a smoothing factor in conjunction with elevation, following the type "I" model form described by @Pedersen2019-kf.
@@
 
-```{r fig3, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability by Kugler morphotype and bumble bee tongue length class.
+```{r fig3, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability parsed by Kugler morphotype and bumble bee tongue length class.
 Point-interval plots depict 66\\% and 95\\% CIs.
 Letters indicate group differences based on evidence ratio at alpha = 0.05.
-Note that the range of the x-axis has been rescaled for each panel to emphasize the comparison of visitation probabilities across tongue length classes.")}
+Note that the range of the x-axis has been rescaled for each panel to emphasize the comparison of visitation probabilities across tongue length classes within floral morphotypes.")}
 knitr::include_graphics("../output/brm_pvis00_plot1.png")
@@
 Point-interval plots depict 66\\% and 95\\% CIs.
-Note that the range of the x-axis has been rescaled for each panel to emphasize the comparison of visitation probabilities across bumble bee species")}
+Note that the range of the x-axis has been rescaled for each panel to emphasize the comparison of visitation probabilities across bumble bee species within floral morphptypes.")}
 knitr::include_graphics("../output/brm_pvis02_plot1.png")
@@
 
-While bumble bees are, indeed, remarkably generalized in their foraging, they are highly discriminating generalists, dramatically over-representing certain floral taxa and morphotypes and ignoring others.
-In our study system, lip, flag, and head flowers dominated overall visitation probability when visitation was pooled across all bumble bee species, while bell, disc, and stalk-disc flowers were underrepresented.
+While bumble bees are, indeed, remarkably generalized in their foraging, they are highly discriminating generalists, dramatically over-representing certain floral taxa and morphotypes relative to their abundance while ignoring others.
+In our study system, lip, flag, and head flowers were by far the most strongly preferred when visitation was pooled across all bumble bee species, while bell, disc, and stalk-disc flowers were underrepresented.
 Patterns of pooled visitation are, of course, driven disproportionately by the behavior of the most abundant bumble bee species, which in our system were *B.
@@
 
-<!-- If not tongue length, what else could explain the variation in floral morphotype preference witin tongue length classes? -->
-
-# CONCLUSION
-
-The mystery of bumble bee species coexistence hinges on the relationship between species-level floral preferences, trait-matching, and the emergent phenomenon of community-level patterns of diet overlap.
-While tongue length is a relatively strong predictor of species-level floral preferences, it is at best a weak predictor of diet overlap.
-This is due in part to divergent preference of bumble bees with similar tongue lengths toward certain floral morphotypes.
-Perhaps more importantly, though, diet overlap at the level of floral genus tended to be about half as strong as diet overlap at the level of floral morphotype, indicating that bumble bees with similar morphological affinities, driven in part by tongue length compatibility, subdivide floral morphotypes at the genus- and species-level, reducing the potential for competition.
-This suggests that taxonomic floral diversity may be as important as overall floral abundance in enabling the coexistence of bumble bee species.
-Something about elevation...
-
-As discussed earlier, the interpretation of the patterns in our data hinge strongly on the extent to which interspecific competition actually occurs in our study system.
+As discussed earlier, the interpretation of floral visitation rates and diet overlap hinge strongly on the extent to which interspecific competition actually occurs in our study system.
 We have argued that the assumption of minimal competition is reasonable in the context of our study, owing to the rich floral resources and harsh abiotic conditions of the Berchtesgadener Alps, together with the manifest co-occurrence of species with similar tongue length.
@@
 
-<!-- We analyze community mean niche overlap as an indication of diffuse competition [@Pianka1974-qk; MacArthur1972-qz] -->
+# CONCLUSION
 
-# ACKNOWLEDGEMENTS
+The question of how bumble bee species with similar tongue lengths coexist hinges on the relationship between species-level floral preferences, trait-matching, and the emergent phenomenon of community-level patterns of diet overlap.
+While tongue length is a relatively strong predictor of species-level floral preferences, it is at best a weak predictor of diet overlap.
+This is due in part to divergent preference of bumble bees with similar tongue lengths toward certain floral morphotypes.
+Perhaps more importantly, though, diet overlap at the level of floral genus tended to be about half as strong as diet overlap at the level of floral morphotype, indicating that bumble bees with similar morphological affinities, driven in part by tongue length compatibility, subdivide floral morphotypes at the genus- and species-level, reducing the potential for competition.
+This suggests that taxonomic floral diversity may be as important as overall floral abundance in enabling the coexistence of bumble bee species.
+**Something about elevation...
+Somemthing that broadens back out to the general themes of species coexisitence and climate change...**
+
+<!-- Maybe add something somewhere about MacArthur's insight regarding the importance of transient and sporadic competition as a driver community assembly. -->
+
+# ACKNOWLEDGMENTS
 

### 2021-03-12 code/ms1_abundance.Rmd
a new take on ms1 analysis

---
title: "Abundance/diversity GAM"
output: html_document
---

# 1. Prep work
## 1.1 Prepare environment
```{r setup, include=FALSE, echo=FALSE}
library(mgcv)
library(mgcViz)
library(tidymv)
library(vegan)
library(ggvegan)
library(ggrepel)
library(tidyverse)
library(lubridate)
```

## 1.2 Load site and trait data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class2 = factor(elev.class2, levels = c("low", "mid", "high")),
         site = factor(site),
         tree_line = factor(tree_line)) # turn elev.class into an ordered factor and site into a regular factor

# BB traits
bb_traits <- read_csv("../data/processed_data/bb_traits.csv")

# FL traits
## family-level taxonomy
fl_tax <- read_csv("../data/processed_data/floral_tax.csv")

## Kugler morphotypes (accessed through Bioflor)
fl_ktype <- read_csv("../data/processed_data/floral_k_type.csv")

## Flower colors (accessed through BioFlor)
fl_color <- read_csv("../data/processed_data/floral_color.csv")

## diameter and corolla depth data from Gita Benadi
fl_morph <- read_csv("../data/processed_data/flower_morphology_GitaBenadi.csv") %>% 
  dplyr::select(SampleID, plant.family = Family, plant.sp = Species, 
                Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam), LN = mean(LN))

## verbal description of Kugler morphotypes
fl_ktype_key <- read_delim("../data/processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  dplyr::select(k.type = 1, description = 2)

## join all floral trait tables
fl_traits <- fl_tax %>%
  full_join(fl_ktype) %>%
  full_join(fl_color) %>%
  full_join(fl_morph) %>%
  dplyr::select(plant.sp, plant.genus, plant.family, k.type, k.type.s, k.type.ss, color, diam, LN)
```

## 1.3 Load network and survey data
```{r message=FALSE}
### BB-FL visitation data
net <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male",
         !bb.sp %in% c("humi", "hypn")) %>% # too rare
  dplyr::select(site, date, bb.sp, plant.sp) # drop unused columns

### Floral survey data
survey <- read_csv("../data/processed_data/floral_survey.csv") %>%
  semi_join(net, by = "plant.sp") %>% # Consider only visited plant species (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  group_by(plant.sp) %>%
  mutate(present = case_when(
    max(flower.cover) > 0 ~ TRUE,
    max(flower.cover) == 0 ~ FALSE
  )) %>%
  filter(present == TRUE) # Ignore plants that never had any flower cover

vis_sp <- net %>%
  left_join(fl_tax) %>%
  select(plant.sp, plant.genus, plant.family) %>%
  unique()

surv_sp <- survey %>%
  left_join(fl_tax) %>%
  select(plant.sp, plant.genus, plant.family) %>%
  unique()

missing <- anti_join(vis_sp, surv_sp)
```

## 1.4 Calculate FL and BB abundance / diversity
```{r echo=FALSE, include=TRUE, message=FALSE}
# Per-transect BB abundance
## full dataset
bb_abund_sp <- net %>%
  dplyr::select(site, date, bb.sp) %>%
  group_by(site, date, bb.sp) %>% # 
  summarize(bb.abund = n()) %>% # get per transect bb abundance
  ungroup() %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  semi_join(net, by = c("site", "date")) %>% # drop site-dates that did not exist in the data set
  left_join(site_data) %>% # add site data
  mutate(year = factor(year(date)),
         yday = yday(date),
         bb.sp = factor(bb.sp),
         management = factor(management),
         site = factor(site),
         transect = factor(transect)) %>%
  left_join(bb_traits) %>%
  mutate(tongue = factor(pbl.w.class2))

### Per-transect floral abund by species
fl_abund_sp <- survey %>%
  group_by(site, date, plant.sp) %>% # group by plant.sp*transect
  summarize(fl.abund = sum(flower.cover)) %>% # get plant.sp*transect total flower cover; this step really shouldn't be required, but in a few cases, there is more than one abundance entry for the same species on a given transect; not sure how that happened, but it's in the original data that Fabrice sent me; alternatively, I could drop the dups or average them. We'll see what Kalli and Fabrice have to say.
  group_by(site, date) %>%
  mutate(prop.fl.abund = fl.abund/sum(fl.abund)) %>% # get the proportional flower cover for each species.
  ungroup() %>%
  #full_join(dplyr::select(net, site, date, plant.sp), by = c("site", "date", "plant.sp")) %>% # add the handful of species occurring in the visitation network but missing from the survey data
  complete(site, date, plant.sp, fill = list(fl.abund = 0, prop.fl.abund = 0)) %>%
  semi_join(survey, by = c("site", "date")) %>%  # only include site-dates that actually occurred
  unique() %>%
  mutate(year = factor(year(date))) %>%
  left_join(fl_traits)
```

## 1.5 Calculate binary "preference" for each plant genus by each bumble bee species; score species 0 or 1 based on whether they were visited
```{r message=FALSE}
weights <- fl_abund_sp %>%
  select(plant.sp, plant.genus) %>%
  unique() %>%
  left_join(select(net, bb.sp, plant.sp)) %>%
  select(plant.genus, bb.sp) %>%
  distinct() %>%
  mutate(visited = rep(1, n())) %>%
  pivot_wider(values_from = visited, names_from = bb.sp, values_fill = 0) %>%
  pivot_longer(cols = 2:15, values_to = "visited", names_to = "bb.sp")

weighted_fl_abund <- fl_abund_sp %>%
  group_by(site, date, plant.genus) %>%
  summarize(flower.cover = sum(fl.abund)) %>%
  left_join(weights) %>%
  mutate(weighted.flower.cover = flower.cover*visited) %>%
  group_by(site, date, bb.sp) %>%
  summarize(total.weighted.flower.cover = sum(weighted.flower.cover)) %>%
  left_join(site_data) %>%
  select(site, date, elev.mean, bb.sp, total.weighted.flower.cover) %>%
  mutate(yday = yday(date),
         bb.sp = factor(bb.sp),
         site = factor(site),
         year = factor(year(date)))
```

# 2. Analyses

## 2.1 BB elevation distribution
```{r}
bb_range <- net %>%
  left_join(site_data) %>%
  group_by(site, date, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = max(abund)) %>%
  filter(abund > 0) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))

ggplot(bb_range, aes(reorder(bb.sp, elev.med), elev.mean)) +
  geom_line(size = 2, alpha = 0.25) +
  geom_point(aes(size = abund, color = tree_line), alpha = 0.5) +
  theme_light(14) +
  scale_size_continuous(name = "Peak abundance",
                        breaks = c(1, 10, 20, 40, 80)) +
  scale_color_discrete(name = "Tree line") +
  xlab("Bumble bee species") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave("../output/bb_range_ms1.pdf")
ggsave("../output/bb_range_ms1.png")

# What happens if I plot max abundance by elevation?

bb_range2 <- net %>%
  left_join(site_data) %>%
  group_by(site, date, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = max(abund))

ggplot(filter(bb_range2, bb.sp != "humi"), aes(elev.mean, abund)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 6)) +
  facet_wrap(~bb.sp, scales = "free_y", ncol = 4) +
  geom_vline(xintercept = 1500, color = "red", linetype = "dashed")
```

## 2.2. Aggregate analysis of max BB abundance and mean FL abundance  ~ elevation

### 2.2.1 Calculate max and mean abundances within site and year
```{r}
# Get mean total weighted floral abundance by BB sp
weighted_fl_abund_mean <- weighted_fl_abund %>%
  group_by(site, bb.sp, year) %>%
  summarize(mean.fl.abund = mean(total.weighted.flower.cover),
            max.fl.abund = max(total.weighted.flower.cover)) %>%
  left_join(site_data)

# Get max BB abundance by sp
bb_abund_sp_max <- bb_abund_sp %>%
  group_by(site, elev.mean, year, bb.sp) %>%
  summarize(max.bb.abund = max(bb.abund),
            mean.bb.abund = mean(bb.abund))

# Join
abund_sum <- full_join(weighted_fl_abund_mean, bb_abund_sp_max) %>%  
  unite(sp.year, c(bb.sp, year), remove = FALSE) %>%
  mutate(site = factor(site),
         bb.sp = factor(bb.sp),
         sp.year = factor(sp.year)) 
```

### 2.2.2 Preliminary visualizations
```{r}
## mean floral abundance
ggplot(abund_sum, aes(elev.mean, mean.fl.abund, color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8), se = FALSE) +
  facet_wrap(~bb.sp, scales = "free_y") +
  geom_vline(xintercept = 1500)

## log(mean floral abundance)
ggplot(abund_sum, aes(elev.mean, log(mean.fl.abund), color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8), se = FALSE) +
  facet_wrap(~bb.sp) +
  geom_vline(xintercept = 1500)

## Max floral abundance)
ggplot(abund_sum, aes(elev.mean, max.fl.abund, color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8), se = FALSE) +
  facet_wrap(~bb.sp) +
  geom_vline(xintercept = 1500)

## Max BB abundance
ggplot(abund_sum, aes(elev.mean, max.bb.abund, color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8),  se = FALSE) +
  facet_wrap(~bb.sp, scales = "free_y") +
  geom_vline(xintercept = 1500)

## Max  abundance ~ mean BB abundance, to see how interchangeable they are.
## They are almost perfectly correlated, and I think max abund is more informative
ggplot(abund_sum, aes(mean.bb.abund, max.bb.abund, color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8),  se = FALSE) +
  facet_wrap(~bb.sp, scales = "free_y")

## Are max FL and mean FL similarly interchangeable? Yes.
ggplot(abund_sum, aes(mean.fl.abund, max.fl.abund, color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8),  se = FALSE) +
  facet_wrap(~bb.sp, scales = "free_y")

## Max BB abundance ~ mean FL abundance
ggplot(abund_sum, aes(log(mean.fl.abund), max.bb.abund)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8),  se = FALSE) +
  facet_wrap(~bb.sp, scales = "free_y")
```

### 2.2.3 Abundance models
#### 2.2.3.1 Bumble bees
```{r}
gam_max_bb_abund_00 <- gam(max.bb.abund ~ 
                             year +
                             s(bb.sp, bs = "re") +
                             s(elev.mean, by = bb.sp, k = 8),
                           method = "REML",
                           family = "nb",
                           select = TRUE,
                           data = abund_sum) %>% getViz()

check.gamViz(gam_max_bb_abund_00)
summary(gam_max_bb_abund_00)
print(plot(gam_max_bb_abund_00, allTerms = TRUE), pages = 1)

gam_max_bb_abund_00_pred <- predict.gam(gam_max_bb_abund_00, type = "terms", se.fit = TRUE) %>%
  as.data.frame() %>%
  as_tibble() %>%
  select(sm.year = 1, sm.bb.sp = 2, gers = 3, hort = 4, jone = 5, lapi = 6, mend  = 7,
         mont = 8, muci = 9, pasc = 10, prat = 11, psit = 12, pyre = 13, soro = 14,
         telu = 15, wurf = 16,
         se.year = 17, se.bb.sp = 18, gers.se = 19, hort.se = 20, jone.se = 21, lapi.se = 22, mend.se  = 23,
         mont.se = 24, muci.se = 25, pasc.se = 26, prat.se = 27, psit.se = 28, pyre.se = 29, soro.se = 30,
         telu.se = 31, wurf.se = 32) %>%
  mutate(fit = sm.year + sm.bb.sp + gers + hort + jone + lapi + 
           mend + mont + muci + pasc + prat + psit + pyre + 
           soro + telu + wurf) %>%
  mutate(se = se.year + se.bb.sp + gers.se + hort.se + jone.se + lapi.se + mend.se + mont.se + 
           muci.se + pasc.se + prat.se + psit.se + pyre.se + soro.se + telu.se + wurf.se) %>%
  select(fit, se) %>%
  mutate(fit.plus = fit + se,
         fit.min = fit - se,
         fit.trans = exp(fit),
         fit.plus.trans = exp(fit.plus),
         fit.min.trans = exp(fit.min)) %>%
  bind_cols(abund_sum)

ggplot(gam_max_bb_abund_00_pred, aes(elev.mean, fit.trans, color = year)) +
  geom_ribbon(aes(elev.mean, fit.trans, 
                  ymax = fit.plus.trans, ymin = fit.min.trans,
                  fill = year), alpha = 0.2, inherit.aes = FALSE) +
  geom_line() +
  facet_wrap(~bb.sp, scales = "free_y") +
  geom_vline(xintercept = 1500, linetype = "dashed") +
  theme_light()
```

#### 2.2.3.2 Floral resources
```{r}
# There are lots of ways this could be specified. From looking at the data and the models, it seems like BB species-level variation in floral resources is very low, and there is a strong global effect of elevation. I think the global model has the strongest message, namely the collapse of overall floral resource availability above the tree line. I can all report the species-level predictions, though.
gam_max_fl_abund_00 <- gam(log(mean.fl.abund) ~ 
                             s(elev.mean, year, bs = "fs", k = 8) +
                             s(elev.mean, bb.sp, bs = "fs", k = 8),
                           method = "REML",
                           family = "gaussian",
                           select = TRUE,
                           data = abund_sum) %>% getViz()

check.gamViz(gam_max_fl_abund_00)
summary(gam_max_fl_abund_00)
concurvity(gam_max_fl_abund_00)
print(plot(gam_max_fl_abund_00, allTerms = TRUE), pages = 1)
plot(gam_max_fl_abund_00)

gam_max_fl_abund_00_pred <- predict.gam(gam_max_fl_abund_00, type = "terms", se.fit = TRUE) %>%
  as.data.frame() %>%
  as_tibble() %>%
  select(sm.year = 1, sm.bb.sp = 2,
         se.year = 3, se.bb.sp = 4) %>%
  mutate(fit = sm.year) %>%
  mutate(se = se.year) %>%
  select(fit, se) %>%
  mutate(fit.plus = fit + se,
         fit.min = fit - se,
         fit.trans = exp(fit),
         fit.plus.trans = exp(fit.plus),
         fit.min.trans = exp(fit.min)) %>%
  bind_cols(na.omit(abund_sum)) %>% # the na.omit() drops the rows where floral surveying data were missing
  select(fit.trans, fit.plus.trans, fit.min.trans, se, elev.mean, year) %>%
  distinct()

ggplot(gam_max_fl_abund_00_pred, aes(elev.mean, fit.trans, color = year)) +
  geom_ribbon(aes(elev.mean, fit.trans, 
                  ymax = fit.plus.trans, ymin = fit.min.trans,
                  fill = year), alpha = 0.2, inherit.aes = FALSE) +
  geom_line() +
  #geom_point(aes(elev.mean, mean.fl.abund, color = year), inherit.aes = FALSE, alpha = 0.25) +
  geom_vline(xintercept = 1500, linetype = "dashed") +
  theme_light()

# Alternatively, since it seems like species basically has just an intercept effect
gam_max_fl_abund_01 <- gam(log(mean.fl.abund) ~ 
                             s(bb.sp, bs = "re") +
                             s(elev.mean, year, bs = "fs", k = 8),
                           method = "REML",
                           family = "gaussian",
                           select = TRUE,
                           data = abund_sum) %>% getViz()

check.gamViz(gam_max_fl_abund_01)
summary(gam_max_fl_abund_01)
concurvity(gam_max_fl_abund_01)
print(plot(gam_max_fl_abund_01, allTerms = TRUE), pages = 1)
plot(gam_max_fl_abund_01)

gam_max_fl_abund_01_pred <- predict.gam(gam_max_fl_abund_00, type = "terms", se.fit = TRUE) %>%
  as.data.frame() %>%
  as_tibble() %>%
  select(sm.year = 1, sm.bb.sp = 2,
         se.year = 3, se.bb.sp = 4) %>%
  mutate(fit = sm.year) %>%
  mutate(se = se.year) %>%
  select(fit, se) %>%
  mutate(fit.plus = fit + se,
         fit.min = fit - se,
         fit.trans = exp(fit),
         fit.plus.trans = exp(fit.plus),
         fit.min.trans = exp(fit.min)) %>%
  bind_cols(na.omit(abund_sum)) %>% # the na.omit() drops the rows where floral surveying data were missing
  select(fit.trans, fit.plus.trans, fit.min.trans, se, elev.mean, year) %>%
  distinct()

ggplot(gam_max_fl_abund_00_pred, aes(elev.mean, fit.trans, color = year)) +
  geom_ribbon(aes(elev.mean, fit.trans, 
                  ymax = fit.plus.trans, ymin = fit.min.trans,
                  fill = year), alpha = 0.2, inherit.aes = FALSE) +
  geom_line() +
  #geom_point(aes(elev.mean, mean.fl.abund, color = year), inherit.aes = FALSE, alpha = 0.25) +
  geom_vline(xintercept = 1500, linetype = "dashed") +
  theme_light()
```

## 2.3 Elevation x time analysis

### 2.3.1 Floral abundance
```{r message=FALSE, warning=FALSE}
# First, let's model overall floral abundance without parsing it out by which species were visited by which bumble bee species
fl_abund_pool <- fl_abund_sp %>%
  group_by(site, date, year) %>%
  summarize(fl.abund = sum(fl.abund)) %>%
  mutate(site = factor(site),
         yday = yday(date)) %>%
  left_join(site_data)

ggplot(fl_abund_pool, aes(fl.abund)) +
  geom_histogram()

fl_abund.qp <- bam(fl.abund ~ 
                      s(site, bs = "re") +
                      s(year, bs = "re") +
                      te(yday, elev.mean,
                         bs = c("gp", "tp"),
                         k = c(8, 8)),
                    data = fl_abund_pool,
                    family = "quasipoisson",
                    discrete = TRUE,
                    select = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(fl_abund.qp)
summary(fl_abund.qp)
plot(fl_abund.qp)

# Now, by bumble bee species
weighted_fl_abund.qp <- bam(total.weighted.flower.cover ~ 
                              s(bb.sp, bs = "re") +
                              s(site, bs = "re") +
                              s(year, bs = "re") +
                              te(yday, elev.mean,
                                 by = bb.sp,
                                 bs = c("gp", "tp"),
                                 k = c(5, 5)),
                            data = filter(weighted_fl_abund, bb.sp != "humi"),
                            family = "quasipoisson",
                            discrete = TRUE,
                            select = TRUE,
                            method = "fREML") %>% getViz(nsim = 500)

check.gamViz(weighted_fl_abund.qp)
sum_fl.qp <- summary(weighted_fl_abund.qp)
print(plot(weighted_fl_abund.qp, allTerms = TRUE), pages = 4)
check2D(weighted_fl_abund.qp, x1 = "yday", x2 = "elev.mean")

gers_fl_abund <- plot(sm(weighted_fl_abund.qp, 4))
hort_fl_abund <- plot(sm(weighted_fl_abund.qp, 5))
hypn_fl_abund <- plot(sm(weighted_fl_abund.qp, 6))
jone_fl_abund <- plot(sm(weighted_fl_abund.qp, 7))
lapi_fl_abund <- plot(sm(weighted_fl_abund.qp, 8))
mend_fl_abund <- plot(sm(weighted_fl_abund.qp, 9))
mont_fl_abund <- plot(sm(weighted_fl_abund.qp, 10))
muci_fl_abund <- plot(sm(weighted_fl_abund.qp, 11))
pasc_fl_abund <- plot(sm(weighted_fl_abund.qp, 12))
prat_fl_abund <- plot(sm(weighted_fl_abund.qp, 13))
psit_fl_abund <- plot(sm(weighted_fl_abund.qp, 14))
pyre_fl_abund <- plot(sm(weighted_fl_abund.qp, 15))
soro_fl_abund <- plot(sm(weighted_fl_abund.qp, 16))
telu_fl_abund <- plot(sm(weighted_fl_abund.qp, 17))
wurf_fl_abund <- plot(sm(weighted_fl_abund.qp, 18))

pdf("../output/gam_FL.pdf")
grid_plot_fl_abun <- gridPrint(gers_fl_abund + 
                                  l_fitRaster() + 
                                  l_fitContour() + 
                                  labs(title = "gers", x = NULL, y = NULL) +
                                  guides(fill=FALSE),
                               hort_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "hort", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               hypn_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "hypn", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               jone_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "jone", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               lapi_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "lapi", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               mend_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "mend", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               mont_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "mont", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               muci_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "muci", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               pasc_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "pasc", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               prat_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "prat", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               psit_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "psit", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               pyre_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "pyre", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               soro_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "soro", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               telu_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "telu", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               wurf_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "wurf", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               ncol = 4)
dev.off()
```

### 2.3.2 Bumble bee abundance
```{r message=FALSE}
# Zero-inflation may be an issue; a simple Poisson model is badly overdispersed
ggplot(bb_abund_sp, aes(bb.abund)) +
  geom_histogram() +
  facet_wrap(~bb.sp)

# Quasipoisson (winner)
bb_abund.qp <- bam(bb.abund ~ 
                     s(bb.sp.ord, bs = "re") + # center species
                     s(year, bs = "re") + # random intercept for year
                     s(site, bs = "re") + # random intercept to avoid pseudoreplicating elevation
                     te(yday, elev.mean, # tensor product smooth of elevation and time
                        by = bb.sp.ord, # by bb.sp
                        bs= c("gp", "tp"), # gaussian procees for time, thinplate spline for elevation
                        k=c(5, 5), 
                        m=2),
                   data = filter(bb_abund_sp, bb.sp != "humi"), # humi is just too rare
                   discrete = TRUE,
                   family = "quasipoisson",
                   select = TRUE,
                   method = "fREML") %>% getViz(nsim = 500)

check.gamViz(bb_abund.qp)
sum.qp <- summary(bb_abund.qp)
print(plot(bb_abund.qp, allTerms = TRUE), pages = 4)
check2D(bb_abund.qp, x1 = "yday", x2 = "elev.mean")

telu_abund <- plot(sm(bb_abund.qp, 4))
pasc_abund <- plot(sm(bb_abund.qp, 5))
prat_abund <- plot(sm(bb_abund.qp, 6))
soro_abund <- plot(sm(bb_abund.qp, 7))
wurf_abund <- plot(sm(bb_abund.qp, 8))
hort_abund <- plot(sm(bb_abund.qp, 9))
jone_abund <- plot(sm(bb_abund.qp, 10))
psit_abund <- plot(sm(bb_abund.qp, 11))
pyre_abund <- plot(sm(bb_abund.qp, 12))
mend_abund <- plot(sm(bb_abund.qp, 13))
mont_abund <- plot(sm(bb_abund.qp, 14))
gers_abund <- plot(sm(bb_abund.qp, 15))
muci_abund <- plot(sm(bb_abund.qp, 16))
hypn_abund <- plot(sm(bb_abund.qp, 17))
lapi_abund <- plot(sm(bb_abund.qp, 18))

pdf("../output/gam_BB.pdf")
grid_plot_fl_abun <- gridPrint(gers_abund + 
                                  l_fitRaster() + 
                                  l_fitContour() + 
                                  labs(title = "gers", x = NULL, y = NULL) +
                                  guides(fill=FALSE),
                               hort_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "hort", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               hypn_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "hypn", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               jone_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "jone", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               lapi_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "lapi", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               mend_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "mend", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               mont_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "mont", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               muci_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "muci", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               pasc_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "pasc", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               prat_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "prat", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               psit_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "psit", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               pyre_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "pyre", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               soro_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "soro", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               telu_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "telu", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               wurf_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "wurf", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               ncol = 4)
dev.off()
```

### 2021-03-12 code/ms1_beta_diversity.Rmd
a new take on ms1 analysis

--- 2021-02-18 code/beta_diversity.Rmd
@@
   mutate(elev.class2 = factor(elev.class2, levels = c("low", "mid", "high")),
-         site = factor(site)) %>% # turn elav.class into an ordered factor and site into a regular factor
+         site = factor(site),
+         tree_line = factor(tree_line)) %>% # turn elav.class into an ordered factor and site into a regular factor
   semi_join(read_csv("../data/processed_data/network.csv"), by = "site") %>% # drop sites that were not sampled
@@
 png("../output/beta_stack.png", width = 6.5, height = 4.5, units = "in", res = 300)
+(g1 | (g2 / g3)) + 
+  plot_annotation(tag_levels = c("A", "B", "C")) + 
+  #plot_layout(guides = "collect") & 
+  theme_light() 
+dev.off()
+
+pdf("../output/beta_stack.pdf", width = 6.5, height = 4.5)
 (g1 | (g2 / g3)) + 

### 2021-03-12 ms_1/ms_1.Rmd
update
a new take on ms1 analysis

@@
 ---
-title: "Elevational range shifts of mountain bumble bees and their floral hosts will be punctuated by nonlinear turnover of species and interactions"
+title: "The tree line ecotone punctuates the elevational turnover of plants, pollinators, and their interactions "
 
@@
 
-Climate-driven range shifts are a fundamental evolutionary process, but the rapid pace of recent climate change has made biogeographic dynamics as much a matter of current events as of historic speculation.
-This provides a unique opportunity to study biogeographic processes in real time, but it also poses a grave threat to global biodiversity as species are forced to respond rapidly within the constraints of gene pools shaped by cooler climates.
-Climate-induced range shifts involve more than merely tracking suitable temperature envelopes; when species move, they do so in the context of biological communities and under the influence of more than just temperature, and the outcomes of climate-induced range shifts are thus contingent on potentially nonlinear processes that cannot be reduced to simple elevational or latitudinal temperature gradients.
+Climate-induced range shifts involve more than merely tracking suitable temperature envelopes; when species move, they do so in the context of biological communities and under the influence of more than just temperature, and the outcomes of climate-induced range shifts are thus contingent on processes that cannot be reduced to linear elevational or latitudinal temperature gradients.
 Bumble bees (Hymentopera: *Bombus* spp.) are cold-adapted species associated with mountain habitats, and they have already exhibited measurable upslope range shifts that are expected to result in habitat loss, novel competitive interactions, and disrupted pollination networks.
 Interpreting these shifts, however, requires an understanding of the current elevational patterns of bumble bees and the floral mutualists upon which climate change will act.
-In this study, we demonstrate that mountain bumble bees and flora are distributed in strongly nonlinear patterns in space and time, and that these patterns propagate from the level of species abundances to that of β-diversity, and from the level of community composition to that of biotic interactions.
+In this study, we analyze bumble bee, flowering plant, and interaction records sampled during 904 standardized transect walks conducted over three consecutive years along an elevational gradient in the German Alps ranging from 640 to 2030 meters above sea level.
+
+We demonstrate that mountain bumble bees and flora exhibit distinct peaks of abundance, are distributed in strongly nonlinear patterns in space and time, and that these patterns propagate from the level of species abundances to that of β-diversity, and from the level of community composition to that of biotic interactions.
 In particular, we show that the tree line ecotone is a critical threshold of floral resource availability, a distributional interface between high- and low-elevation bumble bees species, and a zone of rapid acceleration of species and interaction β-diversity.
@@
 As temperatures rise, species are expected to track suitable conditions to higher latitudes and higher elevations, and such effects are already measurable across a broad range of taxa [@Lenoir2015-tj].
-Species range shifts, however, are not merely abstract, Cartesian translations along axes of latitude, elevation, and temperature; they are movements *in situ*, involving the advent, extinction, and reorganization of biological interactions [@Tylianakis2008-fi; @Blois2013-dz] as well as the concomitant variation of abiotic conditions other than temperature [@Korner1995-mh; @Hodkinson2005-af].
-Moreover, some of these factors, unlike temperature, cannot be assumed to vary linearly with latitude or elevation but instead exhibit marked nonlinearities and thresholds.
+Species range shifts, however, are not merely abstract translations along axes of latitude and elevation; they are movements *in situ*, involving the advent, extinction, and reorganization of biological interactions [@Tylianakis2008-fi; @Blois2013-dz] as well as the concomitant variation of abiotic conditions other than temperature [@Korner1995-mh; @Hodkinson2005-af].
+Thus, when species move along continuous, linear climate gradients, they can experience punctuated and nonlinear ecological transitions.
 A salient example is the tree line ecotone that marks the elevational transition between forest and grassland.
 At this relatively discrete band within a continuous elevational temperature gradient, plant communities exhibit rapid species turnover [@Descombes2017-id] and declining diversity [@Becker2007-cg] as abiotic conditions become more extreme away from the shelter of tree canopy [@Slatyer1992-ir].
-Thus, species range shifts --- and, perhaps more importantly, the ecological consequences thereof --- are contingent upon existing patterns of species distributions and biotic interactions, and community ecology is as much prior as posterior to the ecology of global change.
+Thus, species range shifts and the consequences thereof are contingent upon existing patterns of species distributions and biotic interactions.
 
-Bumble bees (Hymenoptera: *Bombus* spp.) are both classic model taxa in community ecology and an emergent conservation priority due to recent declines [@Goulson2008-oe] and manifest vulnerability to climate change [@Soroye2020-zb].
-The genus *Bombus* is believed to have arisen in the mountains of Asia during the cooling climate of the Eocene-Oligocene boundary (~34 mya), and its subsequent spread and diversification along mountain corridors and into lowland habitats was likely driven by alternating range expansion during cooling periods and retreat to higher elevations during warming periods [@Hines2008-er; @Stewart2010-jn; @Martinet2018-rb].
-Today, bumble bees remain characteristically cold-adapted species [@Heinrich1994-cv], with peak abundance and diversity associated with mountain ranges and northern latitudes, where they are often the dominant guild of flower-visiting insects and the principal pollinators of entomophilous flora [@Williams1998-xi; @Goulson2010-lo].
+Bumble bees (Hymenoptera: *Bombus* spp.) are well-studied model organisms in the field of community ecology, but they have more recently become conservation priorities due to well-documented population declines and particular vulnerability to climate change [@Goulson2008-oe; @Soroye2020-zb].
+The genus *Bombus* is believed to have arisen in the mountains of Asia during the cooling climate of the Eocene-Oligocene boundary (~34 mya) [@Hines2008-er], and bumble bees remain characteristically cold-adapted species today [@Heinrich1994-cv], with peak abundance and diversity in mountain ranges and northern latitudes where they are often the dominant guild of flower-visiting insects and the principal pollinators of entomophilous flora [@Williams1998-xi; @Goulson2010-lo].
 Climate-induced range shifts are expected to be especially pronounced in the mountain habitats that host the core of bumble bee biodiversity, together with their floral mutualists, and upward movement of bumble bee species has already been documented [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc].
@@
 
-The outcomes of range shifts in mountain bumble bee and floral communities have obvious significance from a conservation perspective, but they also provide a model system in which to study the general phenomenon of community assembly under climate change, a phenomenon of both historical interest and prospective urgency.
+<!-- I either need to beef up the paragraph below or fold it into another one. But I need to get to the point of our study. Which means I need to know what the point is. -->
+
+The outcomes of range shifts in mountain bumble bee and floral communities have obvious significance from a conservation perspective [@Soroye2020-zb], but they also provide a model system in which to study the basic phenomenon of community assembly under climate change, a phenomenon of both historical interest and prospective urgency.
+As bumble bee species move up mountain slopes, they move in relation to plant species and other bumble bee species
+
 Anticipating, interpreting, and potentially mitigating the effects of ranges shifts, however, hinge upon an understanding the current elevational and temporal patterns of the bumble bee and floral communities that will be acted upon by climate change.
@@
 
+⎄
 In the present study, we approach this topic at three levels of biological organization.
@@
 β-diversity) of bumble bees and flora in response to elevation.
-Finally, we move to the level of species interactions, investigating the turnover of bee-flower interaction partners through elevation.
-In each of these analyses, we employ flexible modeling techniques capable of capturing nonlinear responses with the goal of identifying where, when, and for which species the impacts of climate-induced range shifts should be expected to be most severe
+Finally, we proceed to the level of species interactions, investigating the turnover of bee-flower interaction partners through elevation.
+In each of these analyses, we employ flexible modeling techniques capable of capturing nonlinear responses with the goal of identifying where, when, and for which species the impacts of climate-induced range shifts should be expected to be most severe.
 
@@
 
-The study was conducted 2010-2012 in Berchtesgaden National Park (47.55°N, 12.92°E), located in the Northern Limestone Alps of southeast Germany.
+The study was conducted 2010-2012 in Berchtesgaden National Park, located in the Northern Limestone Alps of southeast Germany (47.55°N, 12.92°E).
 The landscape is composed of mountain pastures mainly surrounded by coniferous forests.
-We selected 25 study sites (60 x 60 m) on mountain pastures at elevations ranging from 641-2032 m above sea level.
-Fourteen of these pastures are extensively grazed by cattle or sheep, three are mowed for hay production, and eight have been abandoned throughout the last century and are no longer subject to any human management.
-The sites sampled in this study are the same as those used in a series of previous studies [@Hoiss2012-sn; @Hoiss2012-kr; @Hoiss2015-gv], but the data reported here are otherwise independent.
+We selected 25 study sites (60 x 60 m) on mountain pastures at elevations ranging from 641-2032 meters above sea level (m.s.l.) **(Figure 1)**.
+Fourteen of these pastures are extensively grazed by cattle or sheep, 3 are mowed for hay production, and 8 have been abandoned throughout the last century and are no longer subject to any human management.
+Study sites were classified as falling either above or below the tree line based on an examination of Google Earth imagery and corroborated by the authors with field experience at our study sites (K.
+Kallnik, A.
+F.
+Maihoff, A.
+Claßen).
+Among our sites, the tree line generally fell at an elevation of around 1500 m.s.l., consistent with the previous descriptions of the tree line in the Berchtesgander Alps [@Kostler1970-ya; @Mayer1970-zn] **(Figure 2)**.
+Eight sites fell above the tree line and 17 fell below it **(Figures 1, 2)**.
+
+```{r fig1, echo = FALSE,  fig.cap = paste("Study sites in the Berchtesgadener Alps plotted over Google Earth imagery.
+Sites above the tree line are depecited in red and sites below the tree line in yellow.")}
+knitr::include_graphics("../sites2_2021-03-11_17-40-25.png")
+```
 
 Sampling in each year of the study consisted of repeated visits to each site at approximately weekly intervals.
-Samples were only collected on dry days when the air temperature was at least 6°C.
+Samples were only collected during periods without rain when the air temperature was at least 6°C.
 During each visit, bumble bees and their activity were recorded during a 50-minute transect walk.
@@
 Bumble bee queens were identified to species level in the field, while workers and males were stored in individually labeled tubes in the freezer for later identification in the laboratory after [@Amiet1996-jz].
-Floral visitation by male bumble bees was recorded during visitation sampling, but we chose to analyze only visitation by queens and workers.
-In conjunction with visitation observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.1 m^2^.
+Floral visitation by male bumble bees was recorded during visitation sampling, but we chose to analyze only visitation by queens and workers; males are often found resting on flowers, so their "visits" cannot necessarily be interpreted in the same way as those of queens and workers.
+In conjunction with visitation observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.01 m^2^.
 Species identification followed @Lauber2007-ha and @Oberndorfer2001-ro.
 
-While per-sample effort was standardized across all years, the total number of samples (i.e.
-site-dates) varied markedly,  with 153 in 2010, 212 in 2011, and 539 in 2012.
-The increased total sampling effort in 2012 was driven in part by the overall scarcity of bumble bees that year.
+Although per-sample effort was standardized across all years, the total number of samples (i.e.
+site-dates) varied: 153 in 2010, 212 in 2011, and 539 in 2012.
 Since the start of field sampling each year was determined by snowmelt, lower sites were sampled as early as April while higher sites could not be sampled until June or July.
@@
 
-Bumble bee abundance was quantified as the total number of recorded floral visits per bumble bee species per site-date (excluding males).
+### Data processing
+
+Bumble bee abundance was quantified as the total number of recorded floral visits per bumble bee species per site-date.
 To represent floral abundance from the perspective of each bumble bee species in our network, we first scored each plant genus in our data set as either visited or not visited by each bumble bee species.
@@
 We also opted to treat the cryptic *B.
-terrestris/lucorum* species pair as a single morphospecies, since these two species cannot be distinguished reliably in the field.
-Similarly, we pooled representatives of the parasitic *Psithyrus* subgenus --- which included *B.
+terrestris/lucorum* species pair as a single morphospecies, since these two species cannot be visually distinguished reliably.
+We also pooled representatives of the parasitic *Psithyrus* subgenus --- which included *B.
 barbutellus*, *B.
@@
 
+### Distribution of bumble bees and floral resources in elevation and time
+
 We analyzed bumble bee abundance and floral abundance using separate hierarchical generalized additive models (HGAM) [@Wood2011-pg; @Pedersen2019-kf], with abundance modeled as a response to the interaction (tensor product) of elevation and day-of-year.
 Bumble bee species was included in each model as both a smoothing factor and a random intercept term, resulting in centered 2-dimensional abundance smooths for each bumble bee species.
-We also included year and site as random intercept effects to account for annual differences in abundance patterns and repeated measurements within sites.
+We also included year and site as random intercept effects to account for annual differences in mean abundance and repeated measurements within sites.
 
-To complement GAM-based analysis of abundance patterns through elevation and time, we also analyzed the site occupancy pattern of each bumble bee species using cluster and ordination analyses.
-We first calculated the mean abundance across dates of each bumble bee species for each site in our study system.
-Then, we applied a Hellinger transformation to the matrix of mean abundances and calculated the Horn-Morisita distance metric for each pair of species.
-The resulting distance matrix was then used to generate a principal coordinates analysis (PCoA) and a cluster dendrogram, each depicting the (dis)similarity between species pairs in their distribution across sites.
-The primary purpose of this analysis was to organize the presentation of the GAM results by grouping species with similar site affinities, but because our sites were arranged along an elevation gradient, our cluster and ordination analyses also highlight differences between species in elevational distribution.
+### Species and interaction β-diversity
 
 For our analyses of species and interaction β-diversity, we opted to focus on elevational patterns by pooling samples within sites, since the sparse observations of individual sampling events tended to inflate β-diversity and its variance to a degree that impaired model fitting and interpretation.
-After pooling, we calculated the species and interaction β-diversity between all pairs of sites and partitioned total interaction β-diversity into its components of species turnover (further partitioned into bumble bee and plant turnover, respectively) and interaction rewiring [@Novotny2009-my; @Poisot2012-fk].
-
+After pooling, we calculated the species and interaction β-diversity between all pairs of sites and partitioned total interaction β-diversity into its components of species turnover (changes in species composition) and interaction rewiring (identical species interacting in different ways) [@Novotny2009-my; @Poisot2012-fk].
 As an initial analysis, we plotted the relationship between site-wise elevation difference and each metric of β-diversity and verified the significance of the relationship using logistic matrix regression [@Goslee2009-ln], with the difference in number of sampling dates between sites as a covariate to control for potential confounding effects of sampling frequency.
@@
 
-All analyses were conducted in `R` `r citep(citation())`.
+### Software
+
+All analyses were conducted in `R` [@RCoreTeam-2021aa].
+Data handling and visualization were performed with the `tidyverse` suite [@Wickham2019-aa].
 GAM analyses were performed with packages `mgcv` [@Wood_2017] and `mgcViz` [@Fasiolo2018aa], and the models we fit correspond to the “type I” model form described in @Pedersen2019-kf.
-PCoA and species β-diversity calculation were performed with the package `vegan` `r citep(citation("vegan"))`.
+PCoA and species β-diversity calculation were performed with the package `vegan` [@Oksanen_2019] and visualized using the packages `ggplot2` [@Wickham2016-aa] and `ggvegan` [@Simpson2019aa].
 Calculation and partitioning of interaction β-diversity were performed with the package `bipartite` [@Dormann2008-aa; Dormann2009-aa].
 Matrix regression was performed with the package `ecodist` [@Goslee2007-aa].
-GDM analysis of β-diversity was performed with the package `gdm` `r citep(citation("gdm"))`.
-Data handling and visualization was performed with the `tidyverse` suite of R packages `r citep(citation("gdm"))`.
-Annotated R code is available in the Supplementary Material.
+GDM analysis of β-diversity was performed with the package `gdm` [@Fitzpatrick-2021aa].
+Annotated `R` code is available in the Supplementary Material.
 
@@
 soroeensis*, *B.
-terrestris-lucorum*, and *B.
+terrestris/lucorum*, and *B.
 wurflenii* --- accounted for the bulk of overall bumble bee abundance and were present over the entire elevation range of our sites **(Figure 1)**.
@@
 *B.
-monticola* occurred at moderate abundance but was rare or absent at sites below 1000 m.
+monticola* occurred at moderate abundance but was rare or absent at sites below 1000 m.s.l.
 The remaining species were relatively rare in our study system.
@@
 mucidus*, and *B.
-pyrenaeus* occurred primarily above 1500 m, consistent with their known affinity for high elevation habitats [@Rasmont2010-ci].
+pyrenaeus* occurred primarily (peak abundance > 1) above the 1500 m.s.l., consistent with their known affinity for high elevation habitats [@Rasmont2010-ci].
 *B.
@@
 hypnorum*, *B.
-lapidarius* were sparsely distributed across mid-elevation sites.
+lapidarius* were sparsely distributed across low- and mid-elevation sites.
 *B.
@@
 
-```{r fig1, echo = FALSE,  fig.cap = paste("Mean abundances and elevational distributions of bumble bee species.
-Sampled points along the elevation gradient are depicted as circles, the diameter of which is scaled to the mean abundance (pooled across study years) of the given bumble bee species at that elevation.
+```{r fig2, echo = FALSE,  fig.cap = paste("Peak abundances and elevational distributions of bumble bee species.
+Sampled sites along the elevation gradient are depicted as circles, the diameter of which is scaled to the peak abundance (pooled across study years) of the given bumble bee species.
 Abundances of zero are not plotted.
-Lines span between the highest and lowest occurrence of each species, highlighting its observed elevation range.")}
-knitr::include_graphics("../output/bb_range.png")
+Lines span between the highest and lowest occurrence of each species, highlighting its observed elevation range.
+Species are plotted in order of increasing median elevation.")}
+knitr::include_graphics("../output/bb_range_ms1.png")
 ```
@@
 
-## Abundance over elevation and time
+## Distribution of bumble bees and floral resources in elevation and time
 
@@
 
-```{r fig2, echo = FALSE,  fig.cap = paste("Abundance of bumble bee species and their floral resources through elevation and time.
+```{r fig3, echo = FALSE,  fig.cap = paste("Abundance of bumble bee species and their floral resources through elevation and time.
 For each bumble bee species, surface plots are arranged in two columns labeled BB and FL.
@@
 
-```{r fig3, echo = FALSE,  fig.cap = paste("Species-level (A) and interaction-level (B, C) plotted against elevational difference.
+```{r fig4, echo = FALSE,  fig.cap = paste("Species-level (A) and interaction-level (B, C) plotted against elevational difference.
 Each point represents the β-diversity (or β-diversity partition) between a pair of sites, and lines represent the overall relationship between β-diversity and elevation difference using binomial regression smooths.
@@
 
-```{r fig4, echo = FALSE,  fig.cap = paste("GDM splines of species-level and interaction-level β-diversity in response to elevational, controlling for geographic proximity and number of sampling dates.
+```{r fig5, echo = FALSE,  fig.cap = paste("GDM splines of species-level and interaction-level β-diversity in response to elevational, controlling for geographic proximity and number of sampling dates.
 The maximum height of each spline represents a given variable’s partial effect on β-diversity (its effect when covariates are held constant), and the shape of each spline represents to the rate of species turnover as it varies along the corresponding gradient, with steeper parts of the curve indicating regions of the gradient over which species/interaction turnover is more rapid.
+Dotted vertical lines in the elevation panels indicates the location of the tree line.
 Within-site comparisons were omitted from GDM analysis to avoid pseudoreplication.

### 2021-03-12 ms_2/ms_2_v2.Rmd
a new take on ms1 analysis

@@
 Bumble bees are classic model organisms for the study of competition and coexistence, and this topic has taken on new urgency as climate-induced range shifts are expected to disrupt competitive relationships between bumble bee species, particularly along the climatic gradients of the mountain slopes that host the world's richest bumble bee communities.
-We surveyed floral abundance and recorded bumble bee floral visitation patterns along a ~1400 m elevation gradient in German Alps.
-Using these data, we first ask how floral preferences compare across bumble bee species and whether these patterns can be explained by trait-matching between bumble bee tongue length and floral morphotype.
+We surveyed floral abundance and recorded bumble bee floral visitation patterns along a ~1400 m elevation gradient in the German Alps.
+Using these data, we first ask how floral preferences compare across bumble bee species and to what extent these patterns can be explained by trait-matching between bumble bee tongue length and floral morphotype.
 We then investigate the patterns of diet overlap between bumble bee species at the nested levels of floral morphotype, floral family, and floral genus, and model diet overlap as a function bumble bee tongue length and elevation.
-We find that while tongue length is a strong predictor of bumble bees' floral morphotype preferences, this pattern is far from perfect, and species with similar tongue lengths exhibit divergent preferences with respect to certain flora morphotypes.
-This indicates that factors other than tongue length are involved in flower choice, which might explain the puzzling coexistence of bumble bee species with similar tongue lengths.
-Moreover, genus-level diet overlap was approximately one half that of morphotype-level diet overlap, indicating that even when bumble bees exhibit similar biases at the level of floral morphotype, partly due to the constraints of tongue length, floral diversity enables the mitigation of competition via the genus- or species-level subpartitioning of floral resources within morphotypes.
+We find that while tongue length is a significant predictor of bumble bees' floral morphotype preferences, species with similar tongue lengths exhibit divergent preferences with respect to certain flora morphotypes.
+This indicates that factors other than tongue length are involved in flower choice, which might partly explain the puzzling coexistence of bumble bee species with similar tongue lengths.
+Moreover, genus-level diet overlap was approximately one half that of morphotype-level diet overlap, indicating that even when bumble bees exhibit similar biases at the level of floral morphotype, partly due to the constraints of tongue length, floral diversity enables the mitigation of competition via the subpartitioning of floral taxa within morphotypes.
 This suggests that floral species richness may be as important as overall floral abundance in avoiding the competitive exclusion of bumble bee species.
 Diet overlap was highest at the low end of our elevation gradient, declining sharply over the transition from valley floor to mid-slope, then leveling off or declining weakly for the remainder of the gradient.
-If diet overlap is interpreted as an index of potential competition, upslope shifts in bumble bees and flora could generate more competitive conditions at high elevations if and when floral resources become limiting.
+If diet overlap is interpreted as an index of potential competition, upslope shifts in bumble bees and flora could generate more competitive conditions at high elevations in cases where floral resources are limiting.
 
@@
 To account for this inconsistency, @Ranta1981-vx proposed that bumble bees communities are governed by a spatiotemporally heterogeneous mosaic of competitive conditions, such that competitive inequalities between species are sufficiently variable to allow more species to coexist than predicted by trait-based competition models.
-A complementary explanation, based on the core/satellite species hypothesis [@Hanski1982-hs], is that competitive forces are significant only among regionally abundant and widespread “core species”, and the effects of competition are more pronounced in the structure of regional species pools than in that of local assemblages [@Hanski1982-ie].
-In some cases, competition may simply be precluded by a saturating abundance of floral resources [@Williams1989-vr].
+A complementary explanation, based on the core/satellite species hypothesis [@Hanski1982-hs], is that competitive forces are significant only among regionally abundant and widespread “core species”, and the effects of competition are more pronounced in the composition of regional species pools than in that of local assemblages [@Hanski1982-ie].
+In some cases, competition for food may simply be precluded by a saturating abundance of floral resources [@Williams1989-vr], implying that factors other than food availability --- such as abiotic stressors or predator/pathogen pressure --- limit populations.
 
@@
 These costs are closely related to the degree of morphological fittedness between bumble bees and their floral hosts, which consists principally in the matching of bumble bee tongue length to floral corolla depth.
-Bumble bee species can differ markedly in tongue length, and floral handling time increases both for short-tongued bumble bees foraging on deep flowers and long-tongued bees foraging on shallow flowers [@Inouye1980-mv], which leads to an association between tongue length and floral niche in bumble bee communities [e.g.
-@Inouye1978-qi; @Inouye1980-mv].
-This association between tongue length and floral morphology, though, is imperfect [e.g.
-@Ranta1980-ok; @Goulson2008-oe], perhaps in part because access to floral rewards in some flowers involves learning to manipulate floral structures that are challenging irrespective of tongue length [@Laverty1994-hm].
+Bumble bee species can differ markedly in tongue length, and floral handling time increases when tongue length and corolla depth as poorly matched [@Inouye1980-mv; @Harder1983-bk], which leads to an association between tongue length and floral selection in bumble bee communities [e.g.
+@Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn].
+This association between tongue length and floral niche, though, is imperfect, and there can be considerable overlap in floral resource use between bumble bees of different tongue lengths [e.g.
+@Ranta1980-ok; @Goulson2008-oe].
 
 Empirical measurements of niche overlap are, however, inherently ambiguous with respect to competition.
-It has been well-demonstrated that niche overlap can never serve as a metric of competition *per se* [@Abrams1980-am], but it is often used as an index of *potential* competition, particularly in the context of dietary studies [e.g.
+It has been well-demonstrated that niche overlap cannot serve as a metric of competition *per se* [@Abrams1980-am], but it is often used as an index of *potential* competition, particularly in the context of dietary studies [e.g.
 @French1985-eo; @Steffan-Dewenter2000-ll; @Jung2015-cw], based on the reasoning that if and when food becomes limiting, competition can be expected to vary in proportion to the dietary overlap of co-occurring species [@Pianka1972-oi].
@@
 
-Our study is set in the Berchtesgadener Alps of southeast Germany, a landscape of flower-rich mountain meadows --- generated by past or ongoing grazing management --- embedded in a matrix of coniferous forest.
-This system is one of rich floral resources but harsh abiotic conditions, so it is likely that competition for floral resources is low, or at least transient, and that bumble bees are limited instead by abiotic conditions during the critical period of spring colony founding.
-On this basis, we argue for the interpretation of niche overlap as an index of *potential* competition rather than a response to *realized* competition, with the caveats this interpretation may misrepresent certain temporal intervals (e.g.
-early spring colony founding) or elevational bands (e.g.
-the sparse alpine grasslands above the tree line ecotone) [Sponsler et al., submitted] and that the residual effects of past competitive pressure on innate floral biases [@Connell1980-dn] are beyond the scope of our inference.
+Our study is set in the Berchtesgadener Alps of southeast Germany, a landscape of flower-rich mountain meadows embedded in a matrix of coniferous forest.
+This system is one of rich floral resources but harsh abiotic conditions, so it is likely that competition for floral resources is low, or at least transient, and that bumble bees are limited instead by the shortness of the growing season and by fluctuating abiotic conditions during the critical period of spring colony founding.
+On this basis, we argue for the interpretation of niche overlap as an index of *potential* competition rather than a response to *realized* competition, with the caveat this interpretation may misrepresent certain temporal intervals (e.g.
+early spring) or elevational bands (e.g.
+the sparse alpine grasslands above the tree line ecotone) [Sponsler et al., submitted].
 
-The aim of our study is to synthesize the phenomena of floral preference, trait-matching, and dietary niche overlap in a way that accounts for the coexistence of bumble bee species with similar tongue lengths and predicts the effects of climate-driven elevational range shifts on the future of potential competition between bumble bee species in mountain habitats.
+The aim of our study is to synthesize the phenomena of floral preference, trait-matching, and dietary niche overlap in a way that accounts for the coexistence of bumble bee species with similar tongue lengths and sheds light on the possible effects of climate-driven elevational range shifts on the future of potential competition between bumble bee species in mountain habitats.
 We first ask how floral preferences compare across bumble bee species and to what extent these patterns can be explained by trait-matching between bumble bee tongue length and floral morphotype.
@@
 During each visit, bumble bees and their activity were recorded during a 50-minute transect walk.
-Bumblebees observed on or a given flower were counted as floral visitors.
-Bumble bee queens were identified to species level in the field, whereas workers and males were stored in individually labeled tubes in the freezer for later identification in the laboratory after Amiet (1996).
-While floral visitation by male bumble bees was recorded during visitation sampling, we chose to analyze only visitation by queens and workers.
+Bumblebees observed on a given flower were counted as floral visitors.
+Bumble bee queens were identified to species level in the field, whereas workers and males were stored in individually labeled tubes in the freezer for later identification in the laboratory after @Amiet1996-jz.
+While floral visitation by male bumble bees was recorded during visitation sampling, we chose to analyze only visitation by queens and workers, which we pooled.
 During observation, it was noted whether each bee was visibly carrying pollen at the time of its visit, but this provides at best an imperfect distinction between pollen and nectar foragers, and we opted not to differentiate between bees with and without pollen in our analyses.
 In conjunction with visitation observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.1 m^2^.
-Species identification followed @Lauber2007-ha and @Oberndorfer2001-ro.
+Plant species identification followed @Lauber2007-ha and @Oberndorfer2001-ro.
 
@@
 This resulted in a temporally staggered pattern of sampling in 2010 and 2012, while in 2011 a combination of early snowmelt and a late start to field activities resulted in sampling that was approximately aligned in time across the whole elevation gradient.
-See Appendix X for details of sampling patterns.
 
@@
 In most cases, morphotypes are constant at the genus level, and missing morphotypes were added based on the morphotypes found for congeners.
-The genus *Gentiana* consisted of both group 2 and group 4 representatives, so the missing species *Gentiana aspera* and *Gentiana ciliata* were classified based on visual comparison with their classified congeners.
 In the case of the genus *Phyteuma*, we followed @Neumayer1999-il and classified it as a group 7 flower head.
@@
 
-Bumble bee tongue-length data were obtained from the literature values compiled and summarized by @Arbetman2017-xj.
+Bumble bee tongue-length data (for the worker caste) were obtained from the literature values compiled and summarized by @Arbetman2017-xj.
 Values for *B.
@@
 
-## Data analysis
+## Software
 
-### Software
-
-All analyses were conducted in `R` `r citep(citation())`.
+All analyses were conducted in `R` [@RCoreTeam-2021aa].
 Data handling and visualization were performed with the `tidyverse` suite [@Wickham2019-aa].
@@
 Hierarchical generalized additive models (HGAMs) were implemented with the package `mgcv` [@Wood_2017] and visualized with the packages `mgcViz` [@Fasiolo2018aa] and `tidymv` [@Coretta-aa2021].
-All R code used in this analysis, together with full model output and diagnostics, is available in Appendix X.
+All `R` code used in this analysis, together with full model output and diagnostics, is available in Appendix X.
 
-### Data processing
+## Data processing
 
@@
 In a few cases, two consecutive visitation samples at a given site shared the same nearest floral survey sample; in these cases, the visitation samples were first pooled, then aligned to their shared nearest floral survey sample.
-In rare cases, a given floral species was recorded during visitation sampling but missed during floral surveying.
-These species were retroactively assigned the minimum floral area value of 0.01 m^2^.
+In rare cases, a given floral species was recorded during visitation sampling but missed during floral surveying, and these species were retroactively assigned the minimum floral area value of 0.01 m^2^.
 
@@
 To facilitate model fitting, we created a binary version of visitation rate in which visited flora were assigned to 1 and unvisited flora to 0.
-We then modeled binary visitation using generalized models with binomial or Bernoulli distribution families.
+We then modeled binary visitation using generalized models with Bernoulli distribution families.
 The conditional effects inferred from these models must, therefore, be interpreted as effects on visitation probability rather than visitation frequency.
@@
 Using these species-level visitation matrices, we calculated pairwise niche overlap as 1 - Horn-Morisita distance (calculated separately for morphotype-, family-, and genus-level matrices) for all species pairs within each site.
-This approach was chosen for its normalizing effect on the highly variable "snapshot" visitation data generated by our individual transect walks, but it has the disadvantage of conflating niche overlap due to similar floral preference and niche overlap due to similar temporal (i.e.
+This approach was chosen for its normalizing effect on the highly variable "snapshot" visitation data generated by our individual transect walks, but it entails the disadvantage of conflating niche overlap due to similar floral preference and niche overlap due to similar temporal (i.e.
 seasonal) activity patterns.
@@
 Differences in sampling intensity and duration across years (described above) would make comparisons across years difficult to interpret.
-This decision entails the caveats that our analysis cannot detect year-to-year variation in floral biases and niche overlap, and main effects cannot be distinguished from annual anomalies.
+This means that our analysis cannot detect year-to-year variation in floral biases and niche overlap, and main effects cannot be distinguished from annual anomalies.
 With only three years of data, though, such annual inferences would be of little value, and we deemed them not worth the cost of added model complexity.
 
-<!-- Mor edescriptive fgure captions -->
-
-<!-- Describe the differences in models in the Figure2 and 3 captions -->
-
-<!-- Open each results paragraph with a summary sentence. -->
+## Statistical models
 
@@
 Our visitation models (Models 1-3) revealed strong morphotype preferences across the aggregate bumble bee community and marked differences in floral preference between bumble bee species.
-These differences in preference were partly explained by differences in tongue length, but certain floral morphotypes generated divergent preference patterns among bumble bee species with similar tongue lengths.
+These differences in preference were partly explained by differences in tongue length, but certain floral morphotypes generated divergent preference patterns among bumble bee species with similar tongue lengths, and there was substantial overlap in floral preference between bumble bees of different tongue length.
 
@@
 terrestris/lucorum* and the medium-tongued *Bombus monticola* **(Figure 3)**.
-Visitation to funnel (group 2) flowers was dominated by long-tongued species, an effect driven entirely by the *Aconitum* (group 2) specialist *B.
+Funnel (group 2) flowers were preferred by long-tongued species, an effect driven by the *Aconitum* (group 2) specialist *B.
 gerstaeckeri*.
 Bell (group 3) flowers were visited at a higher rate by short-tongued species than by long-tongued species, with medium-tongued species exhibiting intermediate visitation rates.
-At the species-level, bell flowers were visited at highest probability by the short-tongued species *B.
+At the species-level, bell flowers were most preferred by the short-tongued species *B.
 jonellus*, *B.
@@
 Lip (group 5) flowers were also visited at similar rates across tongue length classes, but at the species level *B.
-pascuorum* exhibited an uniquely high bias toward lip flowers.
+pascuorum* exhibited an uniquely high preference for lip flowers.
 Flag (group 6) flowers were visited at a higher rate by medium-tongued species than by short-tongued species, with long-tongued species exhibiting intermediate visitation rates.
@@
 
-```{r fig3, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability parsed by Kugler morphotype and bumble bee tongue length class.
+```{r fig3, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability parsed by Kugler morphotype and bumble bee tongue length class (Model 2).
 Point-interval plots depict 66\\% and 95\\% CIs.
@@
 
-```{r fig4, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability by Kugler morphotype and bumble bee species.
+```{r fig4, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability by Kugler morphotype and bumble bee species (Model 3).
 The tongue length class of each species is color-coded.
@@
 Principal coordinates analysis (Model 4) suggested weak but significant effects of tongue length on the diet similarity while also highlighting the particular floral genera that accounted most strongly for the dietary separation of bumble bee species.
-Axis 1 accounted for 24% of total diet variance and primarily separated long- and medium-tongued species from short-tongued species (Figure 4).
-Long- and medium-tongued species scored low on Axis one and were associated with the lip flowers (group 5) *Ajuga*, *Rhinanthus*, and *Stachys* and flag flowers (group 6) *Anthyllis*, *Lotus*, and *Trifolium*.
-Short tongued species scored high on Axis one and were associated with the bell flowers (group 3) *Campanula*, *Erica*, and *Geum* and the head flowers (group 7) *Carduus*, *Centaurea*, and *Phyteuma*.
+Axis 1 accounted for 23% of total diet variance and primarily separated long- and medium-tongued species from short-tongued species (Figure 4).
+Long- and medium-tongued species scored low on Axis one and were associated with the lip flowers (group 5) *Rhinanthus*, and *Stachys* and flag flowers (group 6) *Anthyllis*, *Lotus*, and *Trifolium*.
+Short tongued species scored high on Axis one and were associated with the bell flowers (group 3) *Campanula* and *Erica*, and the head flower (group 7) *Centaurea*.
 Axis 2 accounted for 19% of total variance and primarily explained variation within short-tongued species.
@@
 Short-tongued species scoring close to zero on Axis 2 also scored lower on Axis 1 (closer to medium- and long-tongued species) and were associated with head flowers.
-Tongue length class explained 22% of diet variance (Adonis permutation test, F = 1.68, p = 0.005).
+Both the lip flower *Ajuga* and the head flower *Centaurea* scored near the middle of the ordination, reflecting their broad appeal across tongue length classes.
+Tongue length class explained 24% of diet variance (Adonis permutation test, F = 1.86, p = 0.003).
 While short-tongued bees appeared most broadly spread in ordination space, there were no significant differences in dispersion across tongue length classes (F = 0.26, p = 0.62).
@@
 
-Site-level analysis of diet overlap revealed strong differences in the magnitude of diet overlap at the nest levels of floral morphotype, floral family, and floral genus, as well as consistent patterns in response to elevation.
+Site-level analysis of diet overlap revealed strong differences in the magnitude of diet overlap at the nested levels of floral morphotype, floral family, and floral genus, as well as consistent patterns in response to elevation.
 At the site level, however, bumble bees belonging to the same tongue length class did not have significantly more similar diets than those belonging to different tongue length classes.
@@
 The Model 5 explained 78.4% of total deviance (adjusted R^2^ = 0.743).
+Whether a species pair belonged to the same or different tongue length classes had no effect on genus-level niche overlap, either in terms of intercept (p = 0.18) or shape of response to elevation **(Figure 9)**.
+Model 6 explained 36.5% of total deviance (adjusted R^2^ = 0.316).
 
@@
 ```
-
-Whether a species pair belonged to the same or different tongue length classes had no effect on genus-level niche overlap, either in terms of intercept (p = 0.18) or shape of response to elevation **(Figure 9)**.
-Model 6 explained 36.5% of total deviance (adjusted R^2^ = 0.316).
 
@@
 
+The question of how bumble bee species with similar tongue lengths coexist hinges on the relationships between species-level floral preferences, trait-matching, and the emergent phenomenon of diet similarity, and this set of relationships can be generalized to describe the niche relations of any group of organisms that share food resources.
+In this study, we clarify these relationships with respect to bumble bees and wild flowers interacting along a mountain slope elevation gradient.
+In addition to addressing the question of bumble bee species coexistence, our results suggest how elevational patterns of potential competition may change in response to climate-induced range shifts, an issue of immediate concern for the conservation of bumble bees and their floral mutualists [@Sirois-Delisle2018-vi; @Soroye2020-zb].
+
 While bumble bees are, indeed, remarkably generalized in their foraging, they are highly discriminating generalists, dramatically over-representing certain floral taxa and morphotypes relative to their abundance while ignoring others.
-In our study system, lip, flag, and head flowers were by far the most strongly preferred when visitation was pooled across all bumble bee species, while bell, disc, and stalk-disc flowers were underrepresented.
+In our study, lip, flag, and head flowers were by far the most strongly preferred when visitation was pooled across all bumble bee species, while bell, disc, and stalk-disc flowers were underrepresented.
 Patterns of pooled visitation are, of course, driven disproportionately by the behavior of the most abundant bumble bee species, which in our system were *B.
@@
 
-While the variation we observed in floral preference across tongue length classes largely comports with the established relationship between tongue length and corolla depth, variation in preference *within* tongue length classes echoes the findings of @Goulson2008-zy in demonstrating the insufficiency of simple morphological matching in explaining bumble bee foraging patterns.
+While the variation we observed in floral preference across tongue length classes largely comports with the established relationship between tongue length and corolla depth, variation in preference *within* tongue length classes agrees with studies demonstrating the insufficiency of simple morphological matching in explaining bumble bee foraging patterns [e.g.
+@Ranta1980-ok; @Goulson2008-oe].
 Perhaps the most striking divergence of floral preference within tongue length class was the divisive effect of bell flowers on short-tongued bumble bees.
@@
 terrestris/lucorum* visited them at rates comparable to medium- and long-tongued species.
-Indeed, B.
-terrestris/lucorum was, on the whole, an idiosyncratic forager, also demonstrating a relatively high affinity for the disc flowers that were largely ignored by other species.
-Together, these species-level idiosyncracies of preference within tongue length classes suggest a possible answer to the question of how bumble bees of similar tongue length can coexists, namely that factors other than tongue length are evidently at play in bumble bees' morphological affinities.
+Indeed, *B.
+terrestris/lucorum* was, on the whole, an eclectic forager, also demonstrating a relatively high affinity for the disc flowers that were largely ignored by other species.
+It should be noted, though, that an unusually high abundance of *B.
+terrestris/lucorum* queens was observed in 2012, but nest founding success was evidently low, leading to low *B.
+terrestris/lucorum* abundance later in the season, an anomaly that could account for the unusual foraging patterns of this species group in the pooled data.
+Together, these species-level idiosyncracies of preference within tongue length classes underscore the fact that factors other than the matching of tongue length and floral morphology --- such as affinities based on floral color, plant growth form, spatial distribution of floral patches, or habitat context [@Brian1957-dg; Harder1985-ls; @Sowig1989-ld] --- are at play in bumble bees' floral preferences, which may partly explain the coexistence of bumble bees of similar tongue lengths.
 
@@
 From an ecological perspective, though, this pattern suggests another answer to the question of how bumble bee species of similar tongue length can coexists.
-If tongue length primarily constrains bumble bee foraging at the coarse level of floral morphotype, taxonomic diversity within morphotypes creates space for bumble bees with similar tongue lengths to avoid actual dietary overlap, thus mitigating interspecific competition.
-Importantly, this suggests that taxonomic diversity *within* floral morphotypes may as important as morphotype diversity *per se* and overall floral abundance in facilitating the coexistence of bumble bee species, an insight which may be relevant to applied bumble bee conservation.
+If tongue length primarily constrains bumble bee foraging at the coarse level of floral morphotype, floral diversity *within* morphotypes creates space for bumble bees with similar tongue lengths to avoid actual dietary overlap, thus mitigating interspecific competition.
+Importantly, this suggests that diversity within floral morphotypes may as important as morphotype diversity *per se* and overall floral abundance in facilitating the coexistence of bumble bee species, an insight which may have implications for the design of floral enhancements aimed at bumble bee conservation.
 
-With respect to elevation, community mean niche overlap was highest at the lowest sites and declined sharply over the transition from valley floor to mid-slope.
+Community mean diet overlap was highest at the lowest sites and declined sharply over the transition from valley floor to mid-slope.
 If this pattern of niche overlap moves upslope in response to climate change, it might result in increased competition potential at higher elevations, which could threaten high-elevation bumble bee species that currently exist under conditions of relatively low niche overlap.
-In our system, though, it must be noted that the lowest elevation sites are also the ones most influences by human land use, and it is possible that the elevation pattern evident in our data is driven by factors other than elevational climate variation.
+In our system, though, it must be noted that the lowest elevation sites are also the ones most influenced by human land use, and it is possible that the elevation pattern evident in our data is driven by factors other than elevational climate variation.
 
 As discussed earlier, the interpretation of floral visitation rates and diet overlap hinge strongly on the extent to which interspecific competition actually occurs in our study system.
-We have argued that the assumption of minimal competition is reasonable in the context of our study, owing to the rich floral resources and harsh abiotic conditions of the Berchtesgadener Alps, together with the manifest co-occurrence of species with similar tongue length.
-This leads to the interpretation of abundance-conditioned visitation rates as *preferences* and diet overlap as *potential competition*.
-To the extent that interspecific competition does occur in our system, though, visitation rates would represent floral preference distorted by the pressure of competition toward niche partitioning, rendering diet overlap an underestimate of potential competition.
+We have argued that the assumption of minimal competition is reasonable in the context of the Berchtesgadener Alps, owing to their rich floral resources and restrictive abiotic conditions, together with the manifest co-occurrence of bumble bee species with similar tongue lengths.
+This leads to the interpretation of abundance-conditioned visitation rates as *preferences* and the interpretation of diet overlap as *potential competition*.
+To the extent that interspecific competition *does* occur in our system, though, visitation rates would represent floral preference distorted by the pressure of competition toward niche partitioning, rendering diet overlap an underestimate of potential competition.
 While we maintain, based on our knowledge of our study system, that the first interpretation is more likely, both interpretations are plausible and must be kept in mind.
-
-# CONCLUSION
-
-The question of how bumble bee species with similar tongue lengths coexist hinges on the relationship between species-level floral preferences, trait-matching, and the emergent phenomenon of community-level patterns of diet overlap.
-While tongue length is a relatively strong predictor of species-level floral preferences, it is at best a weak predictor of diet overlap.
-This is due in part to divergent preference of bumble bees with similar tongue lengths toward certain floral morphotypes.
-Perhaps more importantly, though, diet overlap at the level of floral genus tended to be about half as strong as diet overlap at the level of floral morphotype, indicating that bumble bees with similar morphological affinities, driven in part by tongue length compatibility, subdivide floral morphotypes at the genus- and species-level, reducing the potential for competition.
-This suggests that taxonomic floral diversity may be as important as overall floral abundance in enabling the coexistence of bumble bee species.
-**Something about elevation...
-Somemthing that broadens back out to the general themes of species coexisitence and climate change...**
-
-<!-- Maybe add something somewhere about MacArthur's insight regarding the importance of transient and sporadic competition as a driver community assembly. -->
+Given our assumption of negligible interspecific competition, however, it might be questioned whether the inference of *potential* competition is ecologically relevant, since only realized competition can cause competitive exclusion.
+In answer to this question, we point out that while the assumption of negligible interspecific competition is necessary for the inferences we make concerning floral preference and potential competition, the processes of (1) imperfect trait-matching bewteen bumble bee tongue length and floral morphotype and (2) taxonomic subpartitioning of floral morphotypes that we infer in our analysis still obtain under conditions of realized interspecific competition, even if their inference under those conditions would be obscured by the confounding effects of competitive niche partitioning.
+Indeed, conditions of realized competition likely occur in all bumble bee communities, including those of our study system, but these conditions may be spatially and/or temporally localized [@Ranta1981-vx], and perhaps associated with extreme climatic events that cause dramatic reductions in floral resource avaulability [@Rasmont2012-zq; @Thomson2016-gq; @Phillips2018-me].
+This calls to mind the insight of MacArthur1972-qz, that competition need not be frequent to have lasting ecological impacts.
+Potential competition in the form of dietary overlap is perhaps the best indication of the effects that future instances of realized competition will have, and it is likely that the "ghosts of competition past" [@Connell1980-dn] haunt the floral preferences exhibited by bumble bees today.
 

### 2021-03-15 code/ms1_abundance.Rmd
refined GAM analysis

@@
 library(ggrepel)
+library(ggforce)
 library(tidyverse)
@@
   filter(caste != "male",
-         !bb.sp %in% c("humi", "hypn")) %>% # too rare
+         !bb.sp %in% c("humi", "hypn", "lapi")) %>% # too rare
   dplyr::select(site, date, bb.sp, plant.sp) # drop unused columns
@@
   pivot_wider(values_from = visited, names_from = bb.sp, values_fill = 0) %>%
-  pivot_longer(cols = 2:15, values_to = "visited", names_to = "bb.sp")
+  pivot_longer(cols = 2:14, values_to = "visited", names_to = "bb.sp")
 
@@
   geom_vline(xintercept = 1500, color = "red", linetype = "dashed")
+
+# Overall bumble bee abundance
+bb_abund_sum <- bb_abund_sp %>%
+  group_by(site, bb.sp) %>%
+  summarize(cum.abund = sum(bb.abund),
+            mean.abund = mean(bb.abund)) %>%
+  group_by(site) %>%
+  mutate(prop.abund = cum.abund/sum(cum.abund)) %>%
+  group_by(bb.sp) %>%
+  mutate(mean.prop.abund = mean(prop.abund))
+
+ggplot(bb_abund_sum, aes(site, prop.abund, fill = bb.sp)) +
+  geom_col()
+
+ggplot(bb_abund_sum, aes(bb.sp, mean.abund)) +
+  geom_boxplot()
+
+ggplot(bb_abund_sum, aes(reorder(bb.sp, -mean.prop.abund), prop.abund)) +
+  geom_boxplot(fill = "gray85") +
+  theme_light() +
+  ylab("Proportional abundance") +
+  xlab("Bumble bee species")
+
+ggsave("../output/bb_abundsum.pdf")
+ggsave("../output/bb_abundsum.png")
 ```
@@
          bb.sp = factor(bb.sp),
-         sp.year = factor(sp.year)) 
+         sp.year = factor(sp.year)) %>%
+  unite(bb.sp.year, c(bb.sp, year), sep = "_", remove = FALSE) %>%
+  mutate(bb.sp.year = factor(bb.sp.year))
 ```
@@
   geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 8), se = FALSE) +
+  #geom_smooth(method = "gam", formula = y ~ s(x, k = 8), se = FALSE) +
   facet_wrap(~bb.sp, scales = "free_y") +
@@
 ### 2.2.3 Abundance models
+I'm choosing to use site as a random intercept effect in these models because I think that is the correct way to specify a model involving repeated observations (in this case, bumble bee species) within site.
+The effect of this decision is that the elevation smoother is less influenced by individual sites and converges more strongly toward the mean.
+This creates less exciting curves, but I think it is a more principled way to model the data and less prone to overfitting.
+I can have my cake and eat it too, though, by plotting the actual data over the GAM fit so that the strong patterns of certain sites are visible.
+Then I can talk about the conditional elevation effect and the importance of certain peaks that are likely to be idiosyncrasies of my system (i.e.
+it would be overfitting to call them elevation effects).
+
 #### 2.2.3.1 Bumble bees
 ```{r}
+# Call model
 gam_max_bb_abund_00 <- gam(max.bb.abund ~ 
-                             year +
                              s(bb.sp, bs = "re") +
-                             s(elev.mean, by = bb.sp, k = 8),
+                             s(bb.sp.year, bs = "re") + # fixed intercept for bb.sp:year
+                             s(site, bs = "re") + # random intercept for site
+                             s(elev.mean, by = bb.sp, k = 8), # smoother for elev by BB sp.
                            method = "REML",
-                           family = "nb",
+                           family = "tw", # Tweedie distribution performs well
                            select = TRUE,
@@
 
-check.gamViz(gam_max_bb_abund_00)
-summary(gam_max_bb_abund_00)
-print(plot(gam_max_bb_abund_00, allTerms = TRUE), pages = 1)
+gam_max_bb_abund_00_ilink <- family(gam_max_bb_abund_00)$linkinv # get inverse link function
 
-gam_max_bb_abund_00_pred <- predict.gam(gam_max_bb_abund_00, type = "terms", se.fit = TRUE) %>%
-  as.data.frame() %>%
-  as_tibble() %>%
-  select(sm.year = 1, sm.bb.sp = 2, gers = 3, hort = 4, jone = 5, lapi = 6, mend  = 7,
-         mont = 8, muci = 9, pasc = 10, prat = 11, psit = 12, pyre = 13, soro = 14,
-         telu = 15, wurf = 16,
-         se.year = 17, se.bb.sp = 18, gers.se = 19, hort.se = 20, jone.se = 21, lapi.se = 22, mend.se  = 23,
-         mont.se = 24, muci.se = 25, pasc.se = 26, prat.se = 27, psit.se = 28, pyre.se = 29, soro.se = 30,
-         telu.se = 31, wurf.se = 32) %>%
-  mutate(fit = sm.year + sm.bb.sp + gers + hort + jone + lapi + 
-           mend + mont + muci + pasc + prat + psit + pyre + 
-           soro + telu + wurf) %>%
-  mutate(se = se.year + se.bb.sp + gers.se + hort.se + jone.se + lapi.se + mend.se + mont.se + 
-           muci.se + pasc.se + prat.se + psit.se + pyre.se + soro.se + telu.se + wurf.se) %>%
-  select(fit, se) %>%
-  mutate(fit.plus = fit + se,
-         fit.min = fit - se,
-         fit.trans = exp(fit),
-         fit.plus.trans = exp(fit.plus),
-         fit.min.trans = exp(fit.min)) %>%
+check.gamViz(gam_max_bb_abund_00) # evaluate model
+gam_max_bb_abund_00_sum <- summary(gam_max_bb_abund_00) # summarize model
+print(plot(gam_max_bb_abund_00, allTerms = TRUE), pages = 1) # preliminary visualization
+
+# Plot smooths and year effect on the link scale
+## Individual plots
+gers_abund <- plot(sm(gam_max_bb_abund_00, 4)) + l_fitLine() + l_ciLine() + 
+  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "gers")
+hort_abund <- plot(sm(gam_max_bb_abund_00, 5)) + l_fitLine() + l_ciLine() + 
+  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "hort")
+jone_abund <- plot(sm(gam_max_bb_abund_00, 6)) + l_fitLine() + l_ciLine() + 
+  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "jone")
+mend_abund <- plot(sm(gam_max_bb_abund_00, 7)) + l_fitLine() + l_ciLine() + 
+  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "mend")
+mont_abund <- plot(sm(gam_max_bb_abund_00, 8)) + l_fitLine() + l_ciLine() + 
+  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "mont")
+muci_abund <- plot(sm(gam_max_bb_abund_00, 9)) + l_fitLine() + l_ciLine() + 
+  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "muci")
+pasc_abund <- plot(sm(gam_max_bb_abund_00, 10)) + l_fitLine() + l_ciLine() + 
+  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "pasc")
+prat_abund <- plot(sm(gam_max_bb_abund_00, 11)) + l_fitLine() + l_ciLine() + 
+  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "prat")
+psit_abund <- plot(sm(gam_max_bb_abund_00, 12)) + l_fitLine() + l_ciLine() + 
+  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "psit")
+pyre_abund <- plot(sm(gam_max_bb_abund_00, 13)) + l_fitLine() + l_ciLine() + 
+  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "pyre")
+soro_abund <- plot(sm(gam_max_bb_abund_00, 14)) + l_fitLine() + l_ciLine() + 
+  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "soro")
+telu_abund <- plot(sm(gam_max_bb_abund_00, 15)) + l_fitLine() + l_ciLine() + 
+  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "telu")
+wurf_abund <- plot(sm(gam_max_bb_abund_00, 16)) + l_fitLine() + l_ciLine() + 
+  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "wurf")
+
+## Arrange on grid plot
+grid_bb_abund <- gridPrint(gers_abund, hort_abund, jone_abund, 
+                           mend_abund, mont_abund, muci_abund, 
+                           pasc_abund, prat_abund, psit_abund, 
+                           pyre_abund, soro_abund, telu_abund, 
+                           wurf_abund, year_effect)
+
+pdf("../output/gam_max_bb_abund_grid.pdf")
+plot(grid_bb_abund)
+dev.off()
+
+# Plot model predictions on response scale and add data as points
+## Extract and prepare predictions
+gam_max_bb_abund_00_pred <- predict.gam(gam_max_bb_abund_00, # get model predictions
+                                        type = "terms") %>% 
+  as.data.frame() %>% # convert to data frame
+  as_tibble() %>% # convert to tibble
+  select(sm.bb.sp = 1, sm.bb.sp.year = 2, sm.site = 3, gers = 4, # rename
+         hort = 5, jone = 6, mend = 7, mont = 8, 
+         muci = 9, pasc = 10, prat = 11, psit = 12, 
+         pyre = 13, soro = 14, telu = 15, wurf = 16) %>%
+  mutate(fit = # get cumulative fit and se, ignoring site effect 
+           sm.bb.sp + sm.bb.sp.year + gers + 
+           hort + jone + mend + mont + 
+           muci + pasc + prat + psit + 
+           pyre + soro + telu + wurf 
+           ) %>%
+  select(fit) %>%
+  mutate(fit.trans = gam_max_bb_abund_00_ilink(fit)) %>%
   bind_cols(abund_sum)
 
+## Plot
 ggplot(gam_max_bb_abund_00_pred, aes(elev.mean, fit.trans, color = year)) +
-  geom_ribbon(aes(elev.mean, fit.trans, 
-                  ymax = fit.plus.trans, ymin = fit.min.trans,
-                  fill = year), alpha = 0.2, inherit.aes = FALSE) +
-  geom_line() +
+  # geom_ribbon(aes(elev.mean, fit.trans, 
+  #                 ymax = fit.plus.trans, ymin = fit.min.trans,
+  #                 fill = year), alpha = 0.2, inherit.aes = FALSE) +
+  geom_line(size = 1.5, alpha = 0.5) +
+  geom_point(aes(elev.mean, max.bb.abund, color = year), inherit.aes = FALSE, alpha = 0.25) +
   facet_wrap(~bb.sp, scales = "free_y") +
   geom_vline(xintercept = 1500, linetype = "dashed") +
-  theme_light()
+  theme_light() +
+  labs(x = "Elevation (m.s.l.)", 
+       y = bquote("Peak abundance"),
+       color = "Year",
+       fill = "Year")
+
+ggsave("../output/gam_max_bb_abund.pdf", width = 6.5, height = 4.5)
+ggsave("../output/gam_max_bb_abund.png", width = 6.5, height = 4.5)
 ```
@@
 ```{r}
-# There are lots of ways this could be specified.
-From looking at the data and the models, it seems like BB species-level variation in floral resources is very low, and there is a strong global effect of elevation.
-I think the global model has the strongest message, namely the collapse of overall floral resource availability above the tree line.
-I can all report the species-level predictions, though.
-gam_max_fl_abund_00 <- gam(log(mean.fl.abund) ~ 
+gam_mean_fl_abund_00 <- gam(mean.fl.abund ~ 
+                             s(site, bs = "re") + # random intercept for site
                              s(elev.mean, year, bs = "fs", k = 8) +
@@
                            method = "REML",
-                           family = "gaussian",
+                           family = "tw",
                            select = TRUE,
@@
 
-check.gamViz(gam_max_fl_abund_00)
-summary(gam_max_fl_abund_00)
-concurvity(gam_max_fl_abund_00)
-print(plot(gam_max_fl_abund_00, allTerms = TRUE), pages = 1)
-plot(gam_max_fl_abund_00)
+gam_mean_fl_abund_00_ilink <- family(gam_mean_fl_abund_00)$linkinv # get inverse link function
 
-gam_max_fl_abund_00_pred <- predict.gam(gam_max_fl_abund_00, type = "terms", se.fit = TRUE) %>%
-  as.data.frame() %>%
-  as_tibble() %>%
-  select(sm.year = 1, sm.bb.sp = 2,
-         se.year = 3, se.bb.sp = 4) %>%
-  mutate(fit = sm.year) %>%
-  mutate(se = se.year) %>%
+check.gamViz(gam_mean_fl_abund_00) # evaluate model
+gam_mean_fl_abund_00_sum <- summary(gam_mean_fl_abund_00) # summarize model
+print(plot(gam_mean_fl_abund_00, allTerms = TRUE), pages = 1) # preliminary visualization
+
+# Plot on the link scale
+elev_effect <- plot(sm(gam_mean_fl_abund_00, 2)) + l_fitLine(alpha = 0.6, size = 1)
+bb.sp_effect <- plot(sm(gam_mean_fl_abund_00, 3)) + l_fitLine(alpha = 0.6, size = 1)
+
+grid_fl_abund <- gridPrint(elev_effect, bb.sp_effect)
+
+pdf("../output/gam_mean_fl_abund_grid.pdf")
+plot(grid_fl_abund)
+dev.off()
+
+# Plot model predictions on response scale and add data as points
+## Extract and prepare predictions
+gam_mean_fl_abund_00_pred <- predict.gam(gam_mean_fl_abund_00, 
+                                         type = "terms", 
+                                         se.fit = TRUE) %>% # get model predictions and se
+  as.data.frame() %>% # convert to data frame
+  as_tibble() %>% # convert to tibble
+  select(sm.site = 1, sm.year = 2, sm.bb.sp = 3, # rename
+         se.site = 4, se.year = 5, se.bb.sp = 6) %>%
+  mutate(fit = sm.year, # get fit and se ignoring site and species
+         se = se.year) %>% 
   select(fit, se) %>%
-  mutate(fit.plus = fit + se,
-         fit.min = fit - se,
-         fit.trans = exp(fit),
-         fit.plus.trans = exp(fit.plus),
-         fit.min.trans = exp(fit.min)) %>%
+  mutate(fit.plus = fit + se, # upper CI bound
+         fit.min = fit - se, # lower CI bound
+         fit.trans = gam_mean_fl_abund_00_ilink(fit), # apply inverse link function
+         fit.plus.trans = gam_mean_fl_abund_00_ilink(fit.plus),
+         fit.min.trans = gam_mean_fl_abund_00_ilink(fit.min)) %>%
   bind_cols(na.omit(abund_sum)) %>% # the na.omit() drops the rows where floral surveying data were missing
-  select(fit.trans, fit.plus.trans, fit.min.trans, se, elev.mean, year) %>%
+  select(fit.trans, fit.plus.trans, fit.min.trans, 
+         se, elev.mean, year, mean.fl.abund) %>%
   distinct()
 
-ggplot(gam_max_fl_abund_00_pred, aes(elev.mean, fit.trans, color = year)) +
-  geom_ribbon(aes(elev.mean, fit.trans, 
-                  ymax = fit.plus.trans, ymin = fit.min.trans,
-                  fill = year), alpha = 0.2, inherit.aes = FALSE) +
-  geom_line() +
-  #geom_point(aes(elev.mean, mean.fl.abund, color = year), inherit.aes = FALSE, alpha = 0.25) +
+ggplot(gam_mean_fl_abund_00_pred, aes(elev.mean, fit.trans, color = year)) +
+  # geom_ribbon(aes(elev.mean, fit.trans, 
+  #                 ymax = fit.plus.trans, ymin = fit.min.trans,
+  #                 fill = year), alpha = 0.2, inherit.aes = FALSE) +
+  geom_line(size = 1.5, alpha = 0.5) +
+  geom_point(aes(elev.mean, mean.fl.abund, color = year), inherit.aes = FALSE, alpha = 0.15) +
   geom_vline(xintercept = 1500, linetype = "dashed") +
-  theme_light()
+  theme_light() +
+  labs(x = "Elevation (m.s.l.)", 
+       y = bquote("Mean flower cover" ~ (m^2)),
+       color = "Year",
+       fill = "Year")
 
-# Alternatively, since it seems like species basically has just an intercept effect
-gam_max_fl_abund_01 <- gam(log(mean.fl.abund) ~ 
-                             s(bb.sp, bs = "re") +
-                             s(elev.mean, year, bs = "fs", k = 8),
-                           method = "REML",
-                           family = "gaussian",
-                           select = TRUE,
-                           data = abund_sum) %>% getViz()
+ggsave("../output/gam_mean_fl_abund.pdf")
+ggsave("../output/gam_mean_fl_abund.png")
 
-check.gamViz(gam_max_fl_abund_01)
-summary(gam_max_fl_abund_01)
-concurvity(gam_max_fl_abund_01)
-print(plot(gam_max_fl_abund_01, allTerms = TRUE), pages = 1)
-plot(gam_max_fl_abund_01)
+ggplot(gam_mean_fl_abund_00_pred, aes(elev.mean, fit.trans, color = year)) +
+  # geom_ribbon(aes(elev.mean, fit.trans, 
+  #                 ymax = fit.plus.trans, ymin = fit.min.trans,
+  #                 fill = year), alpha = 0.2, inherit.aes = FALSE) +
+  geom_line(size = 1.5, alpha = 0.5) +
+  geom_point(aes(elev.mean, mean.fl.abund, color = year), inherit.aes = FALSE, alpha = 0.15) +
+  geom_vline(xintercept = 1500, linetype = "dashed") +
+  theme_light() +
+  labs(x = "Elevation (m.s.l.)", 
+       y = bquote("Mean flower cover" ~ (m^2)),
+       color = "Year",
+       fill = "Year") +
+  ylim(c(0, 10))
 
-gam_max_fl_abund_01_pred <- predict.gam(gam_max_fl_abund_00, type = "terms", se.fit = TRUE) %>%
-  as.data.frame() %>%
-  as_tibble() %>%
-  select(sm.year = 1, sm.bb.sp = 2,
-         se.year = 3, se.bb.sp = 4) %>%
-  mutate(fit = sm.year) %>%
-  mutate(se = se.year) %>%
-  select(fit, se) %>%
-  mutate(fit.plus = fit + se,
-         fit.min = fit - se,
-         fit.trans = exp(fit),
-         fit.plus.trans = exp(fit.plus),
-         fit.min.trans = exp(fit.min)) %>%
-  bind_cols(na.omit(abund_sum)) %>% # the na.omit() drops the rows where floral surveying data were missing
-  select(fit.trans, fit.plus.trans, fit.min.trans, se, elev.mean, year) %>%
-  distinct()
+ggsave("../output/gam_mean_fl_abund_sub.pdf")
+ggsave("../output/gam_mean_fl_abund_sub.png")
 
-ggplot(gam_max_fl_abund_00_pred, aes(elev.mean, fit.trans, color = year)) +
-  geom_ribbon(aes(elev.mean, fit.trans, 
-                  ymax = fit.plus.trans, ymin = fit.min.trans,
-                  fill = year), alpha = 0.2, inherit.aes = FALSE) +
-  geom_line() +
-  #geom_point(aes(elev.mean, mean.fl.abund, color = year), inherit.aes = FALSE, alpha = 0.25) +
+ggplot(gam_mean_fl_abund_00_pred, aes(elev.mean, fit.trans, color = year)) +
+  # geom_ribbon(aes(elev.mean, fit.trans, 
+  #                 ymax = fit.plus.trans, ymin = fit.min.trans,
+  #                 fill = year), alpha = 0.2, inherit.aes = FALSE) +
+  geom_line(size = 1.5, alpha = 0.5) +
+  geom_point(aes(elev.mean, mean.fl.abund, color = year), inherit.aes = FALSE, alpha = 0.15) +
   geom_vline(xintercept = 1500, linetype = "dashed") +
-  theme_light()
+  theme_light() +
+  labs(x = "Elevation (m.s.l.)", 
+       y = bquote("Mean flower cover" ~ (m^2)),
+       color = "Year",
+       fill = "Year") +
+  facet_zoom(ylim = c(0, 10), 
+             show.area = FALSE,
+             zoom.size = 1)
+
+ggsave("../output/gam_mean_fl_abund_zoom.pdf", width = 6.5, height = 4.5)
+ggsave("../output/gam_mean_fl_abund_zoom.png", width = 6.5, height = 4.5)
 ```

### 2021-03-15 ms_1/ms_1.Rmd
refined GAM analysis

@@
 ```{r fig1, echo = FALSE,  fig.cap = paste("Study sites in the Berchtesgadener Alps plotted over Google Earth imagery.
-Sites above the tree line are depecited in red and sites below the tree line in yellow.")}
-knitr::include_graphics("../sites2_2021-03-11_17-40-25.png")
+Sites above the tree line are depecited in red and sites below the tree line in yellow.
+The elevation of each site is given in meters above sea level.")}
+knitr::include_graphics("../output/sites_GE.png")
 ```
@@
 
-### Distribution of bumble bees and floral resources in elevation and time
+### Elevational distribution of bumble bees and floral resources
 
-We analyzed bumble bee abundance and floral abundance using separate hierarchical generalized additive models (HGAM) [@Wood2011-pg; @Pedersen2019-kf], with abundance modeled as a response to the interaction (tensor product) of elevation and day-of-year.
-Bumble bee species was included in each model as both a smoothing factor and a random intercept term, resulting in centered 2-dimensional abundance smooths for each bumble bee species.
-We also included year and site as random intercept effects to account for annual differences in mean abundance and repeated measurements within sites.
+We analyzed the abundance of bumble bees and floral resources using separate hierarchical generalized additive models (HGAM) [@Wood2011-pg; @Pedersen2019-kf].
+We opted in both models to the collapse temporal variation of abundance within sites and focus on elevational patterns.
+For bumble bees, we modeled *peak* abundance at each site as a smooth response to elevation, with bumble bee species as a "by" smoothing factor and a random intercept effect.
+We also included species:year as a random intercept effect to capture annual diffences in species abundance and site as a random intercept effect to avoid overfitting due to pseudoreplication of elevation arising from the measurement of multiple bumble bee species at each site.
+This construction is a modification of the "type I" model form described in @Pedersen2019-kf.
+For floral resources, we modeled *mean* flower cover, weighted per bumble be species, as a smooth response to elevation, with bumble bee species and year as "factor-smooths" and site as a random intercept effect (again to avoid overfitting due to pseudoreplication).
+This construction is a modification of the "type S" model form described in @Pedersen2019-kf.
 
@@
 
-For our analyses of species and interaction β-diversity, we opted to focus on elevational patterns by pooling samples within sites, since the sparse observations of individual sampling events tended to inflate β-diversity and its variance to a degree that impaired model fitting and interpretation.
+For our analyses of species and interaction β-diversity, we again opted to focus on elevational patterns by pooling samples within sites, since the sparse observations of individual sampling events tended to inflate β-diversity and its variance to a degree that impaired model fitting and interpretation.
 After pooling, we calculated the species and interaction β-diversity between all pairs of sites and partitioned total interaction β-diversity into its components of species turnover (changes in species composition) and interaction rewiring (identical species interacting in different ways) [@Novotny2009-my; @Poisot2012-fk].
@@
 Data handling and visualization were performed with the `tidyverse` suite [@Wickham2019-aa].
-GAM analyses were performed with packages `mgcv` [@Wood_2017] and `mgcViz` [@Fasiolo2018aa], and the models we fit correspond to the “type I” model form described in @Pedersen2019-kf.
+GAM analyses were performed with packages `mgcv` [@Wood_2017] and `mgcViz` [@Fasiolo2018aa].
 PCoA and species β-diversity calculation were performed with the package `vegan` [@Oksanen_2019] and visualized using the packages `ggplot2` [@Wickham2016-aa] and `ggvegan` [@Simpson2019aa].
@@
 The metaweb across all sites and dates consisted of 16 bumble bee species, 163 plant species (110 genera, 37 families), and 736 unique bumble- bee-plant interaction pairs.
-Five species --- *B.
-pascuorum*, *B.
-pratorum*, *B.
-soroeensis*, *B.
-terrestris/lucorum*, and *B.
-wurflenii* --- accounted for the bulk of overall bumble bee abundance and were present over the entire elevation range of our sites **(Figure 1)**.
 *B.
-hortorum*, *B.
-jonellus*, and *B.
-psithyrus* (species-group) were widespread but at lower abundance.
+pascuorum* ("pasc"), *B.
+pratorum* ("prat"), *B.
+soroeensis* ("soro"), *B.
+terrestris/lucorum* ("telu"), and *B.
+wurflenii* ("wurf") --- accounted for the bulk of overall bumble bee abundance, while *B.
+hortorum* ("hort"), *B.
+jonellus* ("jone"), *B.
+psithyrus* ("psit"), *B.
+monticola* ("mont"), *B.
+mendax* ("mend"), *B.
+mucidus* ("muci"), *B.
+pyrenaeus* ("pyre"), and *B.
+gerstaeckeri* ("gers") .
 *B.
-monticola* occurred at moderate abundance but was rare or absent at sites below 1000 m.s.l.
-The remaining species were relatively rare in our study system.
-*B.
-mendax*, *B.
-mucidus*, and *B.
-pyrenaeus* occurred primarily (peak abundance > 1) above the 1500 m.s.l., consistent with their known affinity for high elevation habitats [@Rasmont2010-ci].
-*B.
-gerstaeckeri*, *B.
 hypnorum*, *B.
-lapidarius* were sparsely distributed across low- and mid-elevation sites.
-*B.
-humilis* was recorded only three times and only in one year, so we omitted it from all analyses.
-
-```{r fig2, echo = FALSE,  fig.cap = paste("Peak abundances and elevational distributions of bumble bee species.
-Sampled sites along the elevation gradient are depicted as circles, the diameter of which is scaled to the peak abundance (pooled across study years) of the given bumble bee species.
-Abundances of zero are not plotted.
-Lines span between the highest and lowest occurrence of each species, highlighting its observed elevation range.
-Species are plotted in order of increasing median elevation.")}
-knitr::include_graphics("../output/bb_range_ms1.png")
-```
+humilis*, and *B.
+lapidarius* were recorded only sporadically and at low abundance, so we omitted them from all analyses.
 
@@
 
-## Distribution of bumble bees and floral resources in elevation and time
+```{r fig1, echo = FALSE,  fig.cap = paste("Proportional abundances (cumulative within site) of bumble bee species in our study system, ordered by mean proportional abundance.")}
+knitr::include_graphics("../output/bb_abundsum.png")
+```
 
-Cluster **(Figure 2A)** and ordination **(Figure 2B)** analyses highlighted the strong differences in site affinity between the high-elevation species *B.
-mucidus*, *B.
-monticola*, *B.
-mendax*, and *B.
-pyrenaeus* and the rest of the bumble bee community.
-PCoA captured 59% of total variation in site affinity.
+## Elevational distribution of bumble bees and floral resources
 
-Elevation-time GAM smooths **(Figure 2C)** of bumble bee abundance differed significantly from a flat surface (p < 0.00001) for all species except *B.
-hypnorum* (p = 0.8).
-Overall, the model explained 57.6% of total deviance (adjusted R^2^ = 0.346).
-The distinction between *B.
-mucidus*, *B.
-monticola*, *B.
-mendax*, and *B.
-pyrenaeus* and the rest of the community was again clear; all 4 high-elevation species exhibited peak abundance at the upper extreme of our elevation gradient, and only *B.
-monticola* was found at high abundance below 1500 m.
-*B.
-monticola* also peaked in abundance earlier in the season than the other high-elevation species.
+Our model explained 71.5% of total deviance (adjusted R^2^ = 0.511).
 
 *B.
-gerstaeckeri* and *B.
-pratorum* exhibited well-defined mid-elevation abundance peaks, with that of *B.
-gerstaeckeri* occurring very late in the season.
-The apparent mid-elevation abundance peak of *B.
-hypnorum* was not significant and should not be interpreted strongly.
+mendax*, *B.
+monticola*, *B.
+mucidus*, and *B.
+pyrenaeus* exhibited a positive response to elevation and peaked in abundance above the tree line (~ 1500 m.s.l.) (**Figure 3**), consistent with their known affinity for high elevation habitats [@Rasmont2010-ci].
+The abundance of *B.
+soroensis* also tended to increase with elevation and peaked above the tree line, but it was also abundant at lower elevations.
+In contrast, the abundance of *B.
+pascorum* --- the most abundant species in our study system --- declined steeply with elevation, and a similar pattern was seen for *B.
+psithyrus*.
 
-*B.
-pascuorum* and *B.
-psithyrus* had well-defined low-elevation peaks, with that of *B.
-psithyrus* occurring early in the season, during the colony-founding --- or, in the case of *B.
-psithryus*, colony-usurping --- phase of the bumble bee life cycle.
+while *B. gerstaeckeri* was sparsely distributed at mid-elevation sites.
 
-The remaining species exhibited relatively diffuse patterns of abundance.
-Of these, *B.
-soroeensis* was exceptional in that it exhibited peak abundance at the highest sites, but it was also found at relatively high abundance across the whole elevation gradient, peaking at mid-season.
-*B.
-hortorum* and *B.
-lapidarius* peaked at low elevation.
-*B.
-jonellus* exhibited little elevational variation in abundance until declining above 1500 m, but its abundance was strongly patterned in time, peaking early in the season.
-*B.
-terrestris/lucorum* was generally abundant across the whole elevation gradient and throughout the whole season, but it showed a weakly trimodal pattern in elevation, with abundance peaks at low, middle, and high sites.
-Its strongest abundance peak occurred at low elevation, early in the season during colony-founding.
-*B.
-wurflenii* was broadly abundant above 1000 m but was rarer at lower sites.
+```{r fig3, echo = FALSE,  fig.cap = paste("Peak bumble bee abundance by elevation and year.
+Curves shows the conditional GAM predictions excluding the site random effect, and points show original data.
+Dotted vertical lines indicate the approximate location of the tree line at our study sites.
+Note different scales of y-axes.")}
+knitr::include_graphics("../output/gam_max_bb_abund2.png")
+```
 
-Elevation-time GAM smooths of floral abundance differed significantly from a flat surface for all species (p << 0.00001) and explained 64.8% of total deviance (adjusted R^2^ = 0.425).
-Floral resource availability exhibited similar patterns for most bumble bee species, with peak floral resources occurring below 1500 m and during mid-season **(Figure 2)**.
-There was a noticeable temporal lag in floral resource availability associated with increasing elevation, consistent with expected temperature constraints on plant phenology.
-The only species to exhibit marked anomalies in floral resource availability was *B.
-gerstaeckeri*, which specializes on the genus *Aconitum* [@Ponchau2006-af] but was also observed (presumably collecting nectar) on *Gentiana*, *Salvia*, *Rhinanthus*, *Cirsium*, and *Clinopodium.* Its floral resources were concentrated in two discrete peaks, the first at low elevation and mid-season, and the second at high elevation and late-season.
-
-```{r fig3, echo = FALSE,  fig.cap = paste("Abundance of bumble bee species and their floral resources through elevation and time.
-For each bumble bee species, surface plots are arranged in two columns labeled BB and FL.
-Surface plots in the BB columns show the abundance of the bumble bee species, while plots in the FL columns show the abundance of floral resources for the corresponding bumble bee species.
-The color ramp of each smooth has been scaled to equal color range, thereby obscuring the absolute magnitude of effect.
-To avoid cluttering the figure, we have opted not to include a unique color ramp key for each plot, since the focus of our inference is on relative rather than absolute patterns.
-Surface plots with original color ramp scales are available in Appendix X.")}
-knitr::include_graphics("../output/gam_BB_FL4.png")
+```{r fig4, echo = FALSE,  fig.cap = paste("Mean weighted floral abundance (area of flower cover) by elevation and year.
+Lines shows the conditional GAM predictions excluding the random effects of site and species, and points show original data.
+The right panel displays the full range of the data while the left panel displays a subset of the range, omitting extreme values of flower cover to highlight the prediction curves.")}
+knitr::include_graphics("../output/gam_mean_fl_abund_zoom.png")
 ```
@@
 
-```{r fig4, echo = FALSE,  fig.cap = paste("Species-level (A) and interaction-level (B, C) plotted against elevational difference.
+```{r fig5, echo = FALSE,  fig.cap = paste("Species-level (A) and interaction-level (B, C) plotted against elevational difference.
 Each point represents the β-diversity (or β-diversity partition) between a pair of sites, and lines represent the overall relationship between β-diversity and elevation difference using binomial regression smooths.
@@
 
-```{r fig5, echo = FALSE,  fig.cap = paste("GDM splines of species-level and interaction-level β-diversity in response to elevational, controlling for geographic proximity and number of sampling dates.
+```{r fig6, echo = FALSE,  fig.cap = paste("GDM splines of species-level and interaction-level β-diversity in response to elevational, controlling for geographic proximity and number of sampling dates.
 The maximum height of each spline represents a given variable’s partial effect on β-diversity (its effect when covariates are held constant), and the shape of each spline represents to the rate of species turnover as it varies along the corresponding gradient, with steeper parts of the curve indicating regions of the gradient over which species/interaction turnover is more rapid.

### 2021-03-16 code/ms1_abundance.Rmd
simplified GAM visualization and opted to drop the site random effect

@@
 library(mgcViz)
+library(patchwork)
 library(tidymv)
@@
   summarize(mean.fl.abund = mean(total.weighted.flower.cover),
-            max.fl.abund = max(total.weighted.flower.cover)) %>%
-  left_join(site_data)
+            max.fl.abund = max(total.weighted.flower.cover))
 
@@
   unite(bb.sp.year, c(bb.sp, year), sep = "_", remove = FALSE) %>%
-  mutate(bb.sp.year = factor(bb.sp.year))
+  mutate(bb.sp.year = factor(bb.sp.year)) %>%
+  left_join(site_data)
 ```
@@
   geom_point(alpha = 0.5) +
-  #geom_smooth(method = "gam", formula = y ~ s(x, k = 8), se = FALSE) +
-  facet_wrap(~bb.sp, scales = "free_y") +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 8), se = FALSE) +
+  #facet_wrap(~bb.sp, scales = "free_y") +
   geom_vline(xintercept = 1500)
@@
                              s(bb.sp.year, bs = "re") + # fixed intercept for bb.sp:year
-                             s(site, bs = "re") + # random intercept for site
                              s(elev.mean, by = bb.sp, k = 8), # smoother for elev by BB sp.
@@
 ## Individual plots
-gers_abund <- plot(sm(gam_max_bb_abund_00, 4)) + l_fitLine() + l_ciLine() + 
-  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "gers")
-hort_abund <- plot(sm(gam_max_bb_abund_00, 5)) + l_fitLine() + l_ciLine() + 
-  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "hort")
-jone_abund <- plot(sm(gam_max_bb_abund_00, 6)) + l_fitLine() + l_ciLine() + 
-  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "jone")
-mend_abund <- plot(sm(gam_max_bb_abund_00, 7)) + l_fitLine() + l_ciLine() + 
-  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "mend")
-mont_abund <- plot(sm(gam_max_bb_abund_00, 8)) + l_fitLine() + l_ciLine() + 
-  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "mont")
-muci_abund <- plot(sm(gam_max_bb_abund_00, 9)) + l_fitLine() + l_ciLine() + 
-  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "muci")
-pasc_abund <- plot(sm(gam_max_bb_abund_00, 10)) + l_fitLine() + l_ciLine() + 
-  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "pasc")
-prat_abund <- plot(sm(gam_max_bb_abund_00, 11)) + l_fitLine() + l_ciLine() + 
-  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "prat")
-psit_abund <- plot(sm(gam_max_bb_abund_00, 12)) + l_fitLine() + l_ciLine() + 
-  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "psit")
-pyre_abund <- plot(sm(gam_max_bb_abund_00, 13)) + l_fitLine() + l_ciLine() + 
-  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "pyre")
-soro_abund <- plot(sm(gam_max_bb_abund_00, 14)) + l_fitLine() + l_ciLine() + 
-  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "soro")
-telu_abund <- plot(sm(gam_max_bb_abund_00, 15)) + l_fitLine() + l_ciLine() + 
-  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "telu")
-wurf_abund <- plot(sm(gam_max_bb_abund_00, 16)) + l_fitLine() + l_ciLine() + 
-  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "wurf")
+gers_abund <- plot(sm(gam_max_bb_abund_00, 3)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "gers")
+hort_abund <- plot(sm(gam_max_bb_abund_00, 4)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "hort")
+jone_abund <- plot(sm(gam_max_bb_abund_00, 5)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "jone")
+mend_abund <- plot(sm(gam_max_bb_abund_00, 6)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "mend")
+mont_abund <- plot(sm(gam_max_bb_abund_00, 7)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "mont")
+muci_abund <- plot(sm(gam_max_bb_abund_00, 8)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "muci")
+pasc_abund <- plot(sm(gam_max_bb_abund_00, 9)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "pasc")
+prat_abund <- plot(sm(gam_max_bb_abund_00, 10)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "prat")
+psit_abund <- plot(sm(gam_max_bb_abund_00, 11)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "psit")
+pyre_abund <- plot(sm(gam_max_bb_abund_00, 12)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "pyre")
+soro_abund <- plot(sm(gam_max_bb_abund_00, 13)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "soro")
+telu_abund <- plot(sm(gam_max_bb_abund_00, 14)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "telu")
+wurf_abund <- plot(sm(gam_max_bb_abund_00, 15)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "wurf")
 
-## Arrange on grid plot
-grid_bb_abund <- gridPrint(gers_abund, hort_abund, jone_abund, 
-                           mend_abund, mont_abund, muci_abund, 
-                           pasc_abund, prat_abund, psit_abund, 
-                           pyre_abund, soro_abund, telu_abund, 
-                           wurf_abund, year_effect)
+## Arrange significant smooths on grid plot
+grid_bb_abund <- gridPrint(gers_abund, mend_abund, mont_abund, muci_abund, 
+                           pasc_abund, prat_abund, psit_abund, pyre_abund, 
+                           soro_abund, telu_abund, wurf_abund, ncol = 4)
 
-pdf("../output/gam_max_bb_abund_grid.pdf")
+pdf("../output/gam_max_bb_abund_grid.pdf", width = 8, height = 6)
 plot(grid_bb_abund)
@@
 
-# Plot model predictions on response scale and add data as points
-## Extract and prepare predictions
-gam_max_bb_abund_00_pred <- predict.gam(gam_max_bb_abund_00, # get model predictions
-                                        type = "terms") %>% 
-  as.data.frame() %>% # convert to data frame
-  as_tibble() %>% # convert to tibble
-  select(sm.bb.sp = 1, sm.bb.sp.year = 2, sm.site = 3, gers = 4, # rename
-         hort = 5, jone = 6, mend = 7, mont = 8, 
-         muci = 9, pasc = 10, prat = 11, psit = 12, 
-         pyre = 13, soro = 14, telu = 15, wurf = 16) %>%
-  mutate(fit = # get cumulative fit and se, ignoring site effect 
-           sm.bb.sp + sm.bb.sp.year + gers + 
-           hort + jone + mend + mont + 
-           muci + pasc + prat + psit + 
-           pyre + soro + telu + wurf 
-           ) %>%
-  select(fit) %>%
-  mutate(fit.trans = gam_max_bb_abund_00_ilink(fit)) %>%
-  bind_cols(abund_sum)
-
-## Plot
-ggplot(gam_max_bb_abund_00_pred, aes(elev.mean, fit.trans, color = year)) +
-  # geom_ribbon(aes(elev.mean, fit.trans, 
-  #                 ymax = fit.plus.trans, ymin = fit.min.trans,
-  #                 fill = year), alpha = 0.2, inherit.aes = FALSE) +
-  geom_line(size = 1.5, alpha = 0.5) +
-  geom_point(aes(elev.mean, max.bb.abund, color = year), inherit.aes = FALSE, alpha = 0.25) +
-  facet_wrap(~bb.sp, scales = "free_y") +
-  geom_vline(xintercept = 1500, linetype = "dashed") +
-  theme_light() +
-  labs(x = "Elevation (m.s.l.)", 
-       y = bquote("Peak abundance"),
-       color = "Year",
-       fill = "Year")
-
-ggsave("../output/gam_max_bb_abund.pdf", width = 6.5, height = 4.5)
-ggsave("../output/gam_max_bb_abund.png", width = 6.5, height = 4.5)
+png("../output/gam_max_bb_abund_grid.png", res = 300, width = 8, height = 6, units = "in")
+plot(grid_bb_abund)
+dev.off()
 ```
@@
 gam_mean_fl_abund_00 <- gam(mean.fl.abund ~
-                             s(site, bs = "re") + # random intercept for site
-                             s(elev.mean, year, bs = "fs", k = 8) +
-                             s(elev.mean, bb.sp, bs = "fs", k = 8),
+                              s(elev.mean, bb.sp, bs = "fs", k = 8) +
+                              s(elev.mean, year, bs = "fs", k = 8),
                            method = "REML",
@@
                            data = abund_sum) %>% getViz()
-
-gam_mean_fl_abund_00_ilink <- family(gam_mean_fl_abund_00)$linkinv # get inverse link function
 
@@
 # Plot on the link scale
-elev_effect <- plot(sm(gam_mean_fl_abund_00, 2)) + l_fitLine(alpha = 0.6, size = 1)
-bb.sp_effect <- plot(sm(gam_mean_fl_abund_00, 3)) + l_fitLine(alpha = 0.6, size = 1)
+elev_effect <- plot(sm(gam_mean_fl_abund_00, 2)) + 
+  l_fitLine(alpha = 0.6, size = 1) + 
+  labs(x = "Elevation (m.s.l.)", y = "Conditional effect (link scale)", color = "Year")
 
-grid_fl_abund <- gridPrint(elev_effect, bb.sp_effect)
+bb.sp_effect <- plot(sm(gam_mean_fl_abund_00, 1)) +
+  l_fitLine(alpha = 0.6, size = 1) + 
+  labs(x = "Elevation (m.s.l.)", y = NULL, color = "Species")
 
-pdf("../output/gam_mean_fl_abund_grid.pdf")
+grid_fl_abund <- gridPrint(elev_effect, bb.sp_effect, ncol = 2)
+
+pdf("../output/gam_mean_fl_abund_grid.pdf", width = 7, height = 3.5)
 plot(grid_fl_abund)
@@
 
-# Plot model predictions on response scale and add data as points
-## Extract and prepare predictions
-gam_mean_fl_abund_00_pred <- predict.gam(gam_mean_fl_abund_00, 
-                                         type = "terms", 
-                                         se.fit = TRUE) %>% # get model predictions and se
-  as.data.frame() %>% # convert to data frame
-  as_tibble() %>% # convert to tibble
-  select(sm.site = 1, sm.year = 2, sm.bb.sp = 3, # rename
-         se.site = 4, se.year = 5, se.bb.sp = 6) %>%
-  mutate(fit = sm.year, # get fit and se ignoring site and species
-         se = se.year) %>% 
-  select(fit, se) %>%
-  mutate(fit.plus = fit + se, # upper CI bound
-         fit.min = fit - se, # lower CI bound
-         fit.trans = gam_mean_fl_abund_00_ilink(fit), # apply inverse link function
-         fit.plus.trans = gam_mean_fl_abund_00_ilink(fit.plus),
-         fit.min.trans = gam_mean_fl_abund_00_ilink(fit.min)) %>%
-  bind_cols(na.omit(abund_sum)) %>% # the na.omit() drops the rows where floral surveying data were missing
-  select(fit.trans, fit.plus.trans, fit.min.trans, 
-         se, elev.mean, year, mean.fl.abund) %>%
-  distinct()
-
-ggplot(gam_mean_fl_abund_00_pred, aes(elev.mean, fit.trans, color = year)) +
-  # geom_ribbon(aes(elev.mean, fit.trans, 
-  #                 ymax = fit.plus.trans, ymin = fit.min.trans,
-  #                 fill = year), alpha = 0.2, inherit.aes = FALSE) +
-  geom_line(size = 1.5, alpha = 0.5) +
-  geom_point(aes(elev.mean, mean.fl.abund, color = year), inherit.aes = FALSE, alpha = 0.15) +
-  geom_vline(xintercept = 1500, linetype = "dashed") +
-  theme_light() +
-  labs(x = "Elevation (m.s.l.)", 
-       y = bquote("Mean flower cover" ~ (m^2)),
-       color = "Year",
-       fill = "Year")
-
-ggsave("../output/gam_mean_fl_abund.pdf")
-ggsave("../output/gam_mean_fl_abund.png")
-
-ggplot(gam_mean_fl_abund_00_pred, aes(elev.mean, fit.trans, color = year)) +
-  # geom_ribbon(aes(elev.mean, fit.trans, 
-  #                 ymax = fit.plus.trans, ymin = fit.min.trans,
-  #                 fill = year), alpha = 0.2, inherit.aes = FALSE) +
-  geom_line(size = 1.5, alpha = 0.5) +
-  geom_point(aes(elev.mean, mean.fl.abund, color = year), inherit.aes = FALSE, alpha = 0.15) +
-  geom_vline(xintercept = 1500, linetype = "dashed") +
-  theme_light() +
-  labs(x = "Elevation (m.s.l.)", 
-       y = bquote("Mean flower cover" ~ (m^2)),
-       color = "Year",
-       fill = "Year") +
-  ylim(c(0, 10))
-
-ggsave("../output/gam_mean_fl_abund_sub.pdf")
-ggsave("../output/gam_mean_fl_abund_sub.png")
-
-ggplot(gam_mean_fl_abund_00_pred, aes(elev.mean, fit.trans, color = year)) +
-  # geom_ribbon(aes(elev.mean, fit.trans, 
-  #                 ymax = fit.plus.trans, ymin = fit.min.trans,
-  #                 fill = year), alpha = 0.2, inherit.aes = FALSE) +
-  geom_line(size = 1.5, alpha = 0.5) +
-  geom_point(aes(elev.mean, mean.fl.abund, color = year), inherit.aes = FALSE, alpha = 0.15) +
-  geom_vline(xintercept = 1500, linetype = "dashed") +
-  theme_light() +
-  labs(x = "Elevation (m.s.l.)", 
-       y = bquote("Mean flower cover" ~ (m^2)),
-       color = "Year",
-       fill = "Year") +
-  facet_zoom(ylim = c(0, 10), 
-             show.area = FALSE,
-             zoom.size = 1)
-
-ggsave("../output/gam_mean_fl_abund_zoom.pdf", width = 6.5, height = 4.5)
-ggsave("../output/gam_mean_fl_abund_zoom.png", width = 6.5, height = 4.5)
-```
-
-## 2.3 Elevation x time analysis
-
-### 2.3.1 Floral abundance
-```{r message=FALSE, warning=FALSE}
-# First, let's model overall floral abundance without parsing it out by which species were visited by which bumble bee species
-fl_abund_pool <- fl_abund_sp %>%
-  group_by(site, date, year) %>%
-  summarize(fl.abund = sum(fl.abund)) %>%
-  mutate(site = factor(site),
-         yday = yday(date)) %>%
-  left_join(site_data)
-
-ggplot(fl_abund_pool, aes(fl.abund)) +
-  geom_histogram()
-
-fl_abund.qp <- bam(fl.abund ~ 
-                      s(site, bs = "re") +
-                      s(year, bs = "re") +
-                      te(yday, elev.mean,
-                         bs = c("gp", "tp"),
-                         k = c(8, 8)),
-                    data = fl_abund_pool,
-                    family = "quasipoisson",
-                    discrete = TRUE,
-                    select = TRUE,
-                    method = "fREML") %>% getViz()
-
-check.gamViz(fl_abund.qp)
-summary(fl_abund.qp)
-plot(fl_abund.qp)
-
-# Now, by bumble bee species
-weighted_fl_abund.qp <- bam(total.weighted.flower.cover ~ 
-                              s(bb.sp, bs = "re") +
-                              s(site, bs = "re") +
-                              s(year, bs = "re") +
-                              te(yday, elev.mean,
-                                 by = bb.sp,
-                                 bs = c("gp", "tp"),
-                                 k = c(5, 5)),
-                            data = filter(weighted_fl_abund, bb.sp != "humi"),
-                            family = "quasipoisson",
-                            discrete = TRUE,
-                            select = TRUE,
-                            method = "fREML") %>% getViz(nsim = 500)
-
-check.gamViz(weighted_fl_abund.qp)
-sum_fl.qp <- summary(weighted_fl_abund.qp)
-print(plot(weighted_fl_abund.qp, allTerms = TRUE), pages = 4)
-check2D(weighted_fl_abund.qp, x1 = "yday", x2 = "elev.mean")
-
-gers_fl_abund <- plot(sm(weighted_fl_abund.qp, 4))
-hort_fl_abund <- plot(sm(weighted_fl_abund.qp, 5))
-hypn_fl_abund <- plot(sm(weighted_fl_abund.qp, 6))
-jone_fl_abund <- plot(sm(weighted_fl_abund.qp, 7))
-lapi_fl_abund <- plot(sm(weighted_fl_abund.qp, 8))
-mend_fl_abund <- plot(sm(weighted_fl_abund.qp, 9))
-mont_fl_abund <- plot(sm(weighted_fl_abund.qp, 10))
-muci_fl_abund <- plot(sm(weighted_fl_abund.qp, 11))
-pasc_fl_abund <- plot(sm(weighted_fl_abund.qp, 12))
-prat_fl_abund <- plot(sm(weighted_fl_abund.qp, 13))
-psit_fl_abund <- plot(sm(weighted_fl_abund.qp, 14))
-pyre_fl_abund <- plot(sm(weighted_fl_abund.qp, 15))
-soro_fl_abund <- plot(sm(weighted_fl_abund.qp, 16))
-telu_fl_abund <- plot(sm(weighted_fl_abund.qp, 17))
-wurf_fl_abund <- plot(sm(weighted_fl_abund.qp, 18))
-
-pdf("../output/gam_FL.pdf")
-grid_plot_fl_abun <- gridPrint(gers_fl_abund + 
-                                  l_fitRaster() + 
-                                  l_fitContour() + 
-                                  labs(title = "gers", x = NULL, y = NULL) +
-                                  guides(fill=FALSE),
-                               hort_fl_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "hort", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               hypn_fl_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "hypn", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               jone_fl_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "jone", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               lapi_fl_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "lapi", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               mend_fl_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "mend", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               mont_fl_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "mont", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               muci_fl_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "muci", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               pasc_fl_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "pasc", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               prat_fl_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "prat", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               psit_fl_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "psit", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               pyre_fl_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "pyre", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               soro_fl_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "soro", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               telu_fl_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "telu", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               wurf_fl_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "wurf", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               ncol = 4)
+png("../output/gam_mean_fl_abund_grid.png", res = 300, units = "in", width = 7, height = 3.5)
+plot(grid_fl_abund)
 dev.off()
 ```
-
-### 2.3.2 Bumble bee abundance
-```{r message=FALSE}
-# Zero-inflation may be an issue; a simple Poisson model is badly overdispersed
-ggplot(bb_abund_sp, aes(bb.abund)) +
-  geom_histogram() +
-  facet_wrap(~bb.sp)
-
-# Quasipoisson (winner)
-bb_abund.qp <- bam(bb.abund ~ 
-                     s(bb.sp.ord, bs = "re") + # center species
-                     s(year, bs = "re") + # random intercept for year
-                     s(site, bs = "re") + # random intercept to avoid pseudoreplicating elevation
-                     te(yday, elev.mean, # tensor product smooth of elevation and time
-                        by = bb.sp.ord, # by bb.sp
-                        bs= c("gp", "tp"), # gaussian procees for time, thinplate spline for elevation
-                        k=c(5, 5), 
-                        m=2),
-                   data = filter(bb_abund_sp, bb.sp != "humi"), # humi is just too rare
-                   discrete = TRUE,
-                   family = "quasipoisson",
-                   select = TRUE,
-                   method = "fREML") %>% getViz(nsim = 500)
-
-check.gamViz(bb_abund.qp)
-sum.qp <- summary(bb_abund.qp)
-print(plot(bb_abund.qp, allTerms = TRUE), pages = 4)
-check2D(bb_abund.qp, x1 = "yday", x2 = "elev.mean")
-
-telu_abund <- plot(sm(bb_abund.qp, 4))
-pasc_abund <- plot(sm(bb_abund.qp, 5))
-prat_abund <- plot(sm(bb_abund.qp, 6))
-soro_abund <- plot(sm(bb_abund.qp, 7))
-wurf_abund <- plot(sm(bb_abund.qp, 8))
-hort_abund <- plot(sm(bb_abund.qp, 9))
-jone_abund <- plot(sm(bb_abund.qp, 10))
-psit_abund <- plot(sm(bb_abund.qp, 11))
-pyre_abund <- plot(sm(bb_abund.qp, 12))
-mend_abund <- plot(sm(bb_abund.qp, 13))
-mont_abund <- plot(sm(bb_abund.qp, 14))
-gers_abund <- plot(sm(bb_abund.qp, 15))
-muci_abund <- plot(sm(bb_abund.qp, 16))
-hypn_abund <- plot(sm(bb_abund.qp, 17))
-lapi_abund <- plot(sm(bb_abund.qp, 18))
-
-pdf("../output/gam_BB.pdf")
-grid_plot_fl_abun <- gridPrint(gers_abund + 
-                                  l_fitRaster() + 
-                                  l_fitContour() + 
-                                  labs(title = "gers", x = NULL, y = NULL) +
-                                  guides(fill=FALSE),
-                               hort_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "hort", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               hypn_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "hypn", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               jone_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "jone", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               lapi_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "lapi", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               mend_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "mend", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               mont_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "mont", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               muci_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "muci", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               pasc_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "pasc", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               prat_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "prat", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               psit_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "psit", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               pyre_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "pyre", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               soro_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "soro", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               telu_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "telu", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               wurf_abund + 
-                                 l_fitRaster() + 
-                                 l_fitContour() + 
-                                 labs(title = "wurf", x = NULL, y = NULL) +
-                                 guides(fill=FALSE),
-                               ncol = 4)
-dev.off()
-```

### 2021-03-16 code/trash_can.Rmd
simplified GAM visualization and opted to drop the site random effect

---
title: "trash can"
output: html_document
---

# Old stuff

### Abundance models
I'm choosing to use site as a random intercept effect in these models because I think that is the correct way to specify a model involving repeated observations (in this case, bumble bee species) within site. The effect of this decision is that the elevation smoother is less influenced by individual sites and converges more strongly toward the mean. This creates less exciting curves, but I think it is a more principled way to model the data and less prone to overfitting. I can have my cake and eat it too, though, by plotting the actual data over the GAM fit so that the strong patterns of certain sites are visible. Then I can talk about the conditional elevation effect and the importance of certain peaks that are likely to be idiosyncrasies of my system (i.e. it would be overfitting to call them elevation effects).

#### Bumble bees
```{r}
# Call model
gam_max_bb_abund_00 <- gam(max.bb.abund ~ 
                             s(bb.sp, bs = "re") +
                             s(bb.sp.year, bs = "re") + # fixed intercept for bb.sp:year
                             s(site, bs = "re") + # random intercept for site
                             s(elev.mean, by = bb.sp, k = 8), # smoother for elev by BB sp.
                           method = "REML",
                           family = "tw", # Tweedie distribution performs well
                           select = TRUE,
                           data = abund_sum) %>% getViz()

gam_max_bb_abund_00_ilink <- family(gam_max_bb_abund_00)$linkinv # get inverse link function

check.gamViz(gam_max_bb_abund_00) # evaluate model
gam_max_bb_abund_00_sum <- summary(gam_max_bb_abund_00) # summarize model
print(plot(gam_max_bb_abund_00, allTerms = TRUE), pages = 1) # preliminary visualization

# Plot smooths and year effect on the link scale
## Individual plots
gers_abund <- plot(sm(gam_max_bb_abund_00, 4)) + l_fitLine() + l_ciLine() + 
  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "gers")
hort_abund <- plot(sm(gam_max_bb_abund_00, 5)) + l_fitLine() + l_ciLine() + 
  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "hort")
jone_abund <- plot(sm(gam_max_bb_abund_00, 6)) + l_fitLine() + l_ciLine() + 
  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "jone")
mend_abund <- plot(sm(gam_max_bb_abund_00, 7)) + l_fitLine() + l_ciLine() + 
  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "mend")
mont_abund <- plot(sm(gam_max_bb_abund_00, 8)) + l_fitLine() + l_ciLine() + 
  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "mont")
muci_abund <- plot(sm(gam_max_bb_abund_00, 9)) + l_fitLine() + l_ciLine() + 
  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "muci")
pasc_abund <- plot(sm(gam_max_bb_abund_00, 10)) + l_fitLine() + l_ciLine() + 
  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "pasc")
prat_abund <- plot(sm(gam_max_bb_abund_00, 11)) + l_fitLine() + l_ciLine() + 
  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "prat")
psit_abund <- plot(sm(gam_max_bb_abund_00, 12)) + l_fitLine() + l_ciLine() + 
  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "psit")
pyre_abund <- plot(sm(gam_max_bb_abund_00, 13)) + l_fitLine() + l_ciLine() + 
  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "pyre")
soro_abund <- plot(sm(gam_max_bb_abund_00, 14)) + l_fitLine() + l_ciLine() + 
  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "soro")
telu_abund <- plot(sm(gam_max_bb_abund_00, 15)) + l_fitLine() + l_ciLine() + 
  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "telu")
wurf_abund <- plot(sm(gam_max_bb_abund_00, 16)) + l_fitLine() + l_ciLine() + 
  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "wurf")

## Arrange on grid plot
grid_bb_abund <- gridPrint(gers_abund, hort_abund, jone_abund, 
                           mend_abund, mont_abund, muci_abund, 
                           pasc_abund, prat_abund, psit_abund, 
                           pyre_abund, soro_abund, telu_abund, 
                           wurf_abund, year_effect)

pdf("../output/gam_max_bb_abund_grid.pdf")
plot(grid_bb_abund)
dev.off()

# Plot model predictions on response scale and add data as points
## Extract and prepare predictions
gam_max_bb_abund_00_pred <- predict.gam(gam_max_bb_abund_00, # get model predictions
                                        type = "terms") %>% 
  as.data.frame() %>% # convert to data frame
  as_tibble() %>% # convert to tibble
  select(sm.bb.sp = 1, sm.bb.sp.year = 2, sm.site = 3, gers = 4, # rename
         hort = 5, jone = 6, mend = 7, mont = 8, 
         muci = 9, pasc = 10, prat = 11, psit = 12, 
         pyre = 13, soro = 14, telu = 15, wurf = 16) %>%
  mutate(fit = # get cumulative fit and se, ignoring site effect 
           sm.bb.sp + sm.bb.sp.year + gers + 
           hort + jone + mend + mont + 
           muci + pasc + prat + psit + 
           pyre + soro + telu + wurf 
           ) %>%
  select(fit) %>%
  mutate(fit.trans = gam_max_bb_abund_00_ilink(fit)) %>%
  bind_cols(abund_sum)

## Plot
ggplot(gam_max_bb_abund_00_pred, aes(elev.mean, fit.trans, linetype = year)) +
  geom_line(size = 0.5, alpha = 0.5) +
  geom_point(aes(elev.mean, max.bb.abund, color = tree_line), inherit.aes = FALSE, alpha = 0.5) +
  facet_wrap(~bb.sp, scales = "free_y") +
  theme_light() +
  labs(x = "Elevation (m.s.l.)", 
       y = bquote("Peak abundance"),
       color = "Tree line",
       shape = "Year",
       linetype = "Year")

# ggplot(gam_max_bb_abund_00_pred, aes(elev.mean, fit.trans, color = year)) +
#   geom_rect(aes(xmin = 1620, # add rectangles to indicate above/below tree line
#                 xmax = 2050,
#                 ymin = -Inf,
#                 ymax = Inf), fill = 'gray80', alpha = 0.01, color = NA) +
#   geom_rect(aes(xmin = 1530,
#                 xmax = 1570,
#                 ymin = -Inf,
#                 ymax = Inf), fill = 'gray80', alpha = 0.01, color = NA) +
#   geom_line(size = 1.5, alpha = 0.5) +
#   geom_point(aes(elev.mean, max.bb.abund, color = year), inherit.aes = FALSE, alpha = 0.5) +
#   facet_wrap(~bb.sp, scales = "free_y") +
#   theme_light() +
#   labs(x = "Elevation (m.s.l.)", 
#        y = bquote("Peak abundance"),
#        color = "Year",
#        fill = "Year")

ggsave("../output/gam_max_bb_abund.pdf", width = 6.5, height = 4.5)
ggsave("../output/gam_max_bb_abund.png", width = 6.5, height = 4.5)

############### Let's see what happens without the site random effect

gam_max_bb_abund_01 <- gam(max.bb.abund ~ 
                             s(bb.sp, bs = "re") +
                             s(bb.sp.year, bs = "re") + # fixed intercept for bb.sp:year
                             s(elev.mean, by = bb.sp, k = 8), # smoother for elev by BB sp.
                           method = "REML",
                           family = "tw", # Tweedie distribution performs well
                           select = TRUE,
                           data = abund_sum) %>% getViz()

gam_max_bb_abund_01_ilink <- family(gam_max_bb_abund_01)$linkinv # get inverse link function

check.gamViz(gam_max_bb_abund_01) # evaluate model
gam_max_bb_abund_01_sum <- summary(gam_max_bb_abund_01) # summarize model
print(plot(gam_max_bb_abund_01, allTerms = TRUE), pages = 1) # preliminary visualization

gam_max_bb_abund_01_pred <- predict.gam(gam_max_bb_abund_01, # get model predictions
                                        type = "terms") %>% 
  as.data.frame() %>% # convert to data frame
  as_tibble() %>% # convert to tibble
  select(sm.bb.sp = 1, sm.bb.sp.year = 2, gers = 3, # rename
         hort = 4, jone = 5, mend = 6, mont = 7, 
         muci = 8, pasc = 9, prat = 10, psit = 11, 
         pyre = 12, soro = 13, telu = 14, wurf = 15) %>%
  mutate(fit = # get cumulative fit
           sm.bb.sp + sm.bb.sp.year + gers + 
           hort + jone + mend + mont + 
           muci + pasc + prat + psit + 
           pyre + soro + telu + wurf 
           ) %>%
  select(fit) %>%
  mutate(fit.trans = gam_max_bb_abund_01_ilink(fit)) %>%
  bind_cols(abund_sum)

ggplot(gam_max_bb_abund_01_pred, aes(elev.mean, fit.trans, linetype = year)) +
  geom_line(size = 0.5, alpha = 0.5) +
  geom_point(aes(elev.mean, max.bb.abund, color = tree_line), inherit.aes = FALSE, alpha = 0.5) +
  facet_wrap(~bb.sp, scales = "free_y") +
  theme_light() +
  labs(x = "Elevation (m.s.l.)", 
       y = bquote("Peak abundance"),
       color = "Tree line",
       shape = "Year",
       linetype = "Year")
```

#### Floral resources
```{r}
gam_mean_fl_abund_00 <- gam(mean.fl.abund ~ 
                             s(site, bs = "re") + # random intercept for site
                             s(elev.mean, year, bs = "fs", k = 8) +
                             s(elev.mean, bb.sp, bs = "fs", k = 8),
                           method = "REML",
                           family = "tw",
                           select = TRUE,
                           data = abund_sum) %>% getViz()

gam_mean_fl_abund_00_ilink <- family(gam_mean_fl_abund_00)$linkinv # get inverse link function

check.gamViz(gam_mean_fl_abund_00) # evaluate model
gam_mean_fl_abund_00_sum <- summary(gam_mean_fl_abund_00) # summarize model
print(plot(gam_mean_fl_abund_00, allTerms = TRUE), pages = 1) # preliminary visualization

# Plot on the link scale
elev_effect <- plot(sm(gam_mean_fl_abund_00, 2)) + 
  l_fitLine(alpha = 0.6, size = 1) + 
  labs(x = "Elevation (m.s.l.)", y = "NULL", color = "Year") +
  ylim(c(-2.1, 1))

bb.sp_effect <- plot(sm(gam_mean_fl_abund_00, 3)) +
  l_fitLine(alpha = 0.6, size = 1) + 
  labs(x = "Elevation (m.s.l.)", y = "NULL", color = "Species") +
  ylim(c(-2.1, 1))

grid_fl_abund <- gridPrint(elev_effect, bb.sp_effect, ncol = 2)

pdf("../output/gam_mean_fl_abund_grid.pdf", width = 7, height = 3.5)
plot(grid_fl_abund)
dev.off()

png("../output/gam_mean_fl_abund_grid.png", res = 300, units = "in", width = 7, height = 3.5)
plot(grid_fl_abund)
dev.off()

# Plot model predictions on response scale and add data as points
## Extract and prepare predictions
gam_mean_fl_abund_00_pred <- predict.gam(gam_mean_fl_abund_00, 
                                         type = "terms", 
                                         se.fit = TRUE) %>% # get model predictions and se
  as.data.frame() %>% # convert to data frame
  as_tibble() %>% # convert to tibble
  select(sm.site = 1, sm.year = 2, sm.bb.sp = 3, # rename
         se.site = 4, se.year = 5, se.bb.sp = 6) %>%
  mutate(fit = sm.year, # get fit and se ignoring site and species
         se = se.year) %>% 
  select(fit, se) %>%
  mutate(fit.plus = fit + se, # upper CI bound
         fit.min = fit - se, # lower CI bound
         fit.trans = gam_mean_fl_abund_00_ilink(fit), # apply inverse link function
         fit.plus.trans = gam_mean_fl_abund_00_ilink(fit.plus),
         fit.min.trans = gam_mean_fl_abund_00_ilink(fit.min)) %>%
  bind_cols(na.omit(abund_sum)) %>% # the na.omit() drops the rows where floral surveying data were missing
  select(fit.trans, fit.plus.trans, fit.min.trans, 
         se, elev.mean, year, mean.fl.abund, tree_line) %>%
  distinct()

# Plot
## Full data
full <- ggplot(gam_mean_fl_abund_00_pred, aes(elev.mean, fit.trans, linetype = year)) +
  geom_point(aes(elev.mean, mean.fl.abund, color = tree_line), inherit.aes = FALSE, alpha = 0.25) +
  geom_line(size = 1, alpha = 0.5) +
  theme_light() +
  labs(x = "Elevation (m.s.l.)", 
       y = bquote("Mean flower cover" ~ (m^2)),
       color = "Tree line",
       linetype = "Year") +
  theme(legend.position = "none")

# full <- ggplot(gam_mean_fl_abund_00_pred, aes(elev.mean, fit.trans, color = year)) +
#   geom_rect(aes(xmin = 1620, # add rectangles to indicate above/below tree line
#                 xmax = 2050,
#                 ymin = -Inf,
#                 ymax = Inf), fill = 'gray80', alpha = 0.01, color = NA) +
#   geom_rect(aes(xmin = 1530,
#               xmax = 1570,
#               ymin = -Inf,
#               ymax = Inf), fill = 'gray80', alpha = 0.01, color = NA) +
#   geom_line(size = 1.5, alpha = 0.5) +
#   geom_point(aes(elev.mean, mean.fl.abund, color = year), inherit.aes = FALSE, alpha = 0.5) +
#   theme_light() +
#   labs(x = "Elevation (m.s.l.)", 
#        y = bquote("Mean flower cover" ~ (m^2)),
#        color = "Year",
#        #shape = "Tree line",
#        fill = "Year") +
#   theme(legend.position = "none")

## Subset to better show elevation smooths
sub <- ggplot(gam_mean_fl_abund_00_pred, aes(elev.mean, fit.trans, linetype = year)) +
  geom_point(aes(elev.mean, mean.fl.abund, color = tree_line), inherit.aes = FALSE, alpha = 0.25) +
  geom_line(size = 1, alpha = 0.5) +
  theme_light() +
  labs(x = "Elevation (m.s.l.)", 
       y = NULL,
       color = "Tree line",
       linetype = "Year") +
  ylim(c(0, 10))

# sub <- ggplot(gam_mean_fl_abund_00_pred, aes(elev.mean, fit.trans, color = year)) +
#   geom_rect(aes(xmin = 1620, # add rectangles to indicate above/below tree line
#                 xmax = 2050,
#                 ymin = -Inf,
#                 ymax = Inf), fill = 'gray80', alpha = 0.01, color = NA) +
#   geom_rect(aes(xmin = 1530,
#               xmax = 1570,
#               ymin = -Inf,
#               ymax = Inf), fill = 'gray80', alpha = 0.01, color = NA) +
#   geom_line(size = 1.5, alpha = 0.5) +
#   geom_point(aes(elev.mean, mean.fl.abund, color = year), inherit.aes = FALSE, alpha = 0.5) +
#   theme_light() +
#   labs(x = "Elevation (m.s.l.)", 
#        y = NULL,
#        color = "Year",
#        #shape = "Tree line",
#        fill = "Year") +
#   ylim(c(0, 10))

# Combine full and sub
fl_patchplot <- full | sub + plot_layout(guides = 'collect')

ggsave(plot = fl_patchplot, "../output/gam_mean_fl_abund_patchplot.pdf", width = 7, height = 4)
ggsave(plot = fl_patchplot, "../output/gam_mean_fl_abund_patchplot.png", width = 7, height = 4)

# A naive plot of floral abundance above vs. below the tree line
ggplot(abund_sum, aes(year, log(mean.fl.abund), color = tree_line)) +
  geom_boxplot() +
  theme_light() +
  labs(x = "Year", 
       y = bquote("Mean flower cover" ~ (m^2) ~ "(log-scale)"),
       color = "Tree line")

############### Let's see what happens without the site random effect
gam_mean_fl_abund_02 <- gam(mean.fl.abund ~
                              s(elev.mean, bb.sp, bs = "fs", k = 8) +
                              s(elev.mean, year, bs = "fs", k = 8),
                           method = "REML",
                           family = "tw",
                           select = TRUE,
                           data = abund_sum) %>% getViz()

check.gamViz(gam_mean_fl_abund_02) # evaluate model
gam_mean_fl_abund_02_sum <- summary(gam_mean_fl_abund_02) # summarize model
print(plot(gam_mean_fl_abund_02, allTerms = TRUE), pages = 1) # preliminary visualization
```

## Elevation x time analysis

### Floral abundance
```{r message=FALSE, warning=FALSE}
# First, let's model overall floral abundance without parsing it out by which species were visited by which bumble bee species
fl_abund_pool <- fl_abund_sp %>%
  group_by(site, date, year) %>%
  summarize(fl.abund = sum(fl.abund)) %>%
  mutate(site = factor(site),
         yday = yday(date)) %>%
  left_join(site_data)

ggplot(fl_abund_pool, aes(fl.abund)) +
  geom_histogram()

fl_abund.qp <- bam(fl.abund ~ 
                      s(site, bs = "re") +
                      s(year, bs = "re") +
                      te(yday, elev.mean,
                         bs = c("gp", "tp"),
                         k = c(8, 8)),
                    data = fl_abund_pool,
                    family = "quasipoisson",
                    discrete = TRUE,
                    select = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(fl_abund.qp)
summary(fl_abund.qp)
plot(fl_abund.qp)

# Now, by bumble bee species
weighted_fl_abund.qp <- bam(total.weighted.flower.cover ~ 
                              s(bb.sp, bs = "re") +
                              s(site, bs = "re") +
                              s(year, bs = "re") +
                              te(yday, elev.mean,
                                 by = bb.sp,
                                 bs = c("gp", "tp"),
                                 k = c(5, 5)),
                            data = filter(weighted_fl_abund, bb.sp != "humi"),
                            family = "quasipoisson",
                            discrete = TRUE,
                            select = TRUE,
                            method = "fREML") %>% getViz(nsim = 500)

check.gamViz(weighted_fl_abund.qp)
sum_fl.qp <- summary(weighted_fl_abund.qp)
print(plot(weighted_fl_abund.qp, allTerms = TRUE), pages = 4)
check2D(weighted_fl_abund.qp, x1 = "yday", x2 = "elev.mean")

gers_fl_abund <- plot(sm(weighted_fl_abund.qp, 4))
hort_fl_abund <- plot(sm(weighted_fl_abund.qp, 5))
hypn_fl_abund <- plot(sm(weighted_fl_abund.qp, 6))
jone_fl_abund <- plot(sm(weighted_fl_abund.qp, 7))
lapi_fl_abund <- plot(sm(weighted_fl_abund.qp, 8))
mend_fl_abund <- plot(sm(weighted_fl_abund.qp, 9))
mont_fl_abund <- plot(sm(weighted_fl_abund.qp, 10))
muci_fl_abund <- plot(sm(weighted_fl_abund.qp, 11))
pasc_fl_abund <- plot(sm(weighted_fl_abund.qp, 12))
prat_fl_abund <- plot(sm(weighted_fl_abund.qp, 13))
psit_fl_abund <- plot(sm(weighted_fl_abund.qp, 14))
pyre_fl_abund <- plot(sm(weighted_fl_abund.qp, 15))
soro_fl_abund <- plot(sm(weighted_fl_abund.qp, 16))
telu_fl_abund <- plot(sm(weighted_fl_abund.qp, 17))
wurf_fl_abund <- plot(sm(weighted_fl_abund.qp, 18))

pdf("../output/gam_FL.pdf")
grid_plot_fl_abun <- gridPrint(gers_fl_abund + 
                                  l_fitRaster() + 
                                  l_fitContour() + 
                                  labs(title = "gers", x = NULL, y = NULL) +
                                  guides(fill=FALSE),
                               hort_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "hort", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               hypn_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "hypn", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               jone_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "jone", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               lapi_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "lapi", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               mend_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "mend", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               mont_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "mont", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               muci_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "muci", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               pasc_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "pasc", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               prat_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "prat", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               psit_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "psit", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               pyre_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "pyre", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               soro_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "soro", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               telu_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "telu", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               wurf_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "wurf", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               ncol = 4)
dev.off()
```

### Bumble bee abundance
```{r message=FALSE}
# Zero-inflation may be an issue; a simple Poisson model is badly overdispersed
ggplot(bb_abund_sp, aes(bb.abund)) +
  geom_histogram() +
  facet_wrap(~bb.sp)

# Quasipoisson (winner)
bb_abund.qp <- bam(bb.abund ~ 
                     s(bb.sp.ord, bs = "re") + # center species
                     s(year, bs = "re") + # random intercept for year
                     s(site, bs = "re") + # random intercept to avoid pseudoreplicating elevation
                     te(yday, elev.mean, # tensor product smooth of elevation and time
                        by = bb.sp.ord, # by bb.sp
                        bs= c("gp", "tp"), # gaussian procees for time, thinplate spline for elevation
                        k=c(5, 5), 
                        m=2),
                   data = filter(bb_abund_sp, bb.sp != "humi"), # humi is just too rare
                   discrete = TRUE,
                   family = "quasipoisson",
                   select = TRUE,
                   method = "fREML") %>% getViz(nsim = 500)

check.gamViz(bb_abund.qp)
sum.qp <- summary(bb_abund.qp)
print(plot(bb_abund.qp, allTerms = TRUE), pages = 4)
check2D(bb_abund.qp, x1 = "yday", x2 = "elev.mean")

telu_abund <- plot(sm(bb_abund.qp, 4))
pasc_abund <- plot(sm(bb_abund.qp, 5))
prat_abund <- plot(sm(bb_abund.qp, 6))
soro_abund <- plot(sm(bb_abund.qp, 7))
wurf_abund <- plot(sm(bb_abund.qp, 8))
hort_abund <- plot(sm(bb_abund.qp, 9))
jone_abund <- plot(sm(bb_abund.qp, 10))
psit_abund <- plot(sm(bb_abund.qp, 11))
pyre_abund <- plot(sm(bb_abund.qp, 12))
mend_abund <- plot(sm(bb_abund.qp, 13))
mont_abund <- plot(sm(bb_abund.qp, 14))
gers_abund <- plot(sm(bb_abund.qp, 15))
muci_abund <- plot(sm(bb_abund.qp, 16))
hypn_abund <- plot(sm(bb_abund.qp, 17))
lapi_abund <- plot(sm(bb_abund.qp, 18))

pdf("../output/gam_BB.pdf")
grid_plot_fl_abun <- gridPrint(gers_abund + 
                                  l_fitRaster() + 
                                  l_fitContour() + 
                                  labs(title = "gers", x = NULL, y = NULL) +
                                  guides(fill=FALSE),
                               hort_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "hort", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               hypn_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "hypn", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               jone_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "jone", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               lapi_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "lapi", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               mend_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "mend", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               mont_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "mont", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               muci_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "muci", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               pasc_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "pasc", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               prat_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "prat", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               psit_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "psit", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               pyre_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "pyre", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               soro_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "soro", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               telu_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "telu", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               wurf_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "wurf", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               ncol = 4)
dev.off()
```

### 2021-03-16 ms_1/ms_1.Rmd
simplified GAM visualization and opted to drop the site random effect

@@
 Fourteen of these pastures are extensively grazed by cattle or sheep, 3 are mowed for hay production, and 8 have been abandoned throughout the last century and are no longer subject to any human management.
-Study sites were classified as falling either above or below the tree line based on an examination of Google Earth imagery and corroborated by the authors with field experience at our study sites (K.
+Study sites were classified as lying either above or below the tree line based on an examination of Google Earth imagery and corroborated by the authors with field experience at our study sites (K.
 Kallnik, A.
@@
 Claßen).
-Among our sites, the tree line generally fell at an elevation of around 1500 m.s.l., consistent with the previous descriptions of the tree line in the Berchtesgander Alps [@Kostler1970-ya; @Mayer1970-zn] **(Figure 2)**.
-Eight sites fell above the tree line and 17 fell below it **(Figures 1, 2)**.
+Eight sites were located above the tree line, which generally fell at an elevation of around 1500 m.s.l., consistent with the previous descriptions of the tree line in the Berchtesgadener Alps [@Kostler1970-ya; @Mayer1970-zn].
 
@@
 We also opted to treat the cryptic *B.
-terrestris/lucorum* species pair as a single morphospecies, since these two species cannot be visually distinguished reliably.
-We also pooled representatives of the parasitic *Psithyrus* subgenus --- which included *B.
+terrestris/lucorum* species pair as a single morphospecies, since these two species cannot be visually distinguished reliably, and we similarly pooled representatives of the parasitic *Psithyrus* subgenus --- *B.
 barbutellus*, *B.
@@
 quadricolor*, and *B.
-sylvestris* --- into a single species group *B.
+sylvestris* --- into the species group *B.
 psithryus*.
@@
 
-We analyzed the abundance of bumble bees and floral resources using separate hierarchical generalized additive models (HGAM) [@Wood2011-pg; @Pedersen2019-kf].
-We opted in both models to the collapse temporal variation of abundance within sites and focus on elevational patterns.
-For bumble bees, we modeled *peak* abundance at each site as a smooth response to elevation, with bumble bee species as a "by" smoothing factor and a random intercept effect.
-We also included species:year as a random intercept effect to capture annual diffences in species abundance and site as a random intercept effect to avoid overfitting due to pseudoreplication of elevation arising from the measurement of multiple bumble bee species at each site.
+We analyzed the abundance of bumble bees and floral resources using separate hierarchical generalized additive models (HGAMs) fit with the Tweedie distribution family (log link function) [@Wood2011-pg; @Pedersen2019-kf].
+We opted in both models to collapse the temporal variation of abundance within sites and focus on elevational patterns.
+
+For bumble bees, we modeled *peak* abundance at each site as a smooth response to elevation, with bumble bee *species* as a smoothing factor and a random intercept term.
+We also included *species:year* as a random intercept term because exploratory analysis indicated that species varied differently across years.
+Finally, we included site as a random intercept effect to avoid overfitting due to pseudoreplication of elevation arising from the measurement of multiple bumble bee species at each site.
 This construction is a modification of the "type I" model form described in @Pedersen2019-kf.
-For floral resources, we modeled *mean* flower cover, weighted per bumble be species, as a smooth response to elevation, with bumble bee species and year as "factor-smooths" and site as a random intercept effect (again to avoid overfitting due to pseudoreplication).
-This construction is a modification of the "type S" model form described in @Pedersen2019-kf.
+
+For floral resources, we modeled *mean* flower cover, weighted per bumble bee species, as a smooth response to elevation, with year as a smoothing factor.
+To account for differences in mean resource abundance --- and potentially differences in the elevational response curve of resource abundance --- across bumble bee species, we added a group-level smoothing term of flower cover by elevation with bumble bee species as a smoothing factor.
+Again, we included site as a random intercept effect to avoid overfitting due to pseudoreplication.
+This construction is a modification of the "type GS" model form described in @Pedersen2019-kf.
 
@@
 
-```{r fig1, echo = FALSE,  fig.cap = paste("Proportional abundances (cumulative within site) of bumble bee species in our study system, ordered by mean proportional abundance.")}
+```{r fig2, echo = FALSE,  fig.cap = paste("Proportional abundances (cumulative within site) of bumble bee species in our study system, ordered by mean proportional abundance.")}
 knitr::include_graphics("../output/bb_abundsum.png")
@@
 
-Our model explained 71.5% of total deviance (adjusted R^2^ = 0.511).
+Our bumble bee abundance model explained 72% of total deviance (adjusted R^2^ = 0.51).
 
@@
 ```{r fig3, echo = FALSE,  fig.cap = paste("Peak bumble bee abundance by elevation and year.
-Curves shows the conditional GAM predictions excluding the site random effect, and points show original data.
-Dotted vertical lines indicate the approximate location of the tree line at our study sites.
+Curves shows the conditional GAM predictions excluding the site random intercept effect, and points show original data and are color-coded to indicate sites falling above the tree line.
 Note different scales of y-axes.")}
-knitr::include_graphics("../output/gam_max_bb_abund2.png")
+knitr::include_graphics("../output/gam_max_bb_abund_grid.png")
 ```
 
-```{r fig4, echo = FALSE,  fig.cap = paste("Mean weighted floral abundance (area of flower cover) by elevation and year.
-Lines shows the conditional GAM predictions excluding the random effects of site and species, and points show original data.
-The right panel displays the full range of the data while the left panel displays a subset of the range, omitting extreme values of flower cover to highlight the prediction curves.")}
-knitr::include_graphics("../output/gam_mean_fl_abund_zoom.png")
+Floral resource abundance varied markedly across years **(Figure 4)**.
+In 2010, floral resource abundance peaked at the lowest sites and declined sharply with elevation.
+In 2011 and 2012, floral resource availability peaked at mid elevation.
+Overall floral resource abundance was highest in 2011 and lowest in 2012.
+
+Sites near the tree line tended to have relatively high floral resource abundance, but floral resource abundance declined sharply above the tree line, reaching a minimum at the highest sites.
+
+<!-- This is the message; not that it IS low above the tree line but that it turns steeply down above the tree line. -->
+
+Floral resource abundance differed across bumble bee species mainly as an intercept effect **(Figure 4C)**, with the most extreme case being *B.
+gerstaeckeri*, which had by far the lowest resource abundance.
+This is not surprising, since *B.
+gerstaeckeri* specializes on the genus *Aconitum* for pollen foraging (Ponchau2006-af), though in our study it was also observed (presumably collecting nectar) on *Gentiana*, *Salvia*, *Rhinanthus*, *Cirsium*, and *Clinopodium*.
+For most species, deviations from the shape of the global response curve were minimal, but the high-elevation species *B.
+mendax*, *B.
+monticola*, *B.
+mucidus*, and *B.
+pyrenaeus*, along with *B.
+jonellus*, experienced stable or increasing conditional floral resource availability with increasing elevation, while conditional floral resource availability declined for other species.
+Our floral resource abundance model explained 81% of total deviance (adjusted R^2^ = 0.56).
+
+When interpreting the response curves of our model, it is important to note that certain sites exhibited extremely high floral resource availability, and this variance was explained by the random effect intercept effect of *site* (not depicted in plotted smooths) rather than the fixed effect of *elevation*.
+We opted for this model specification because it is more resistant to overfitting and therefore more likely to express a generalizable effect of elevation, but we visualize the empirical floral abundance data in **Figure 4** so that these strong site effects, which account for most of the residual variation with respect to the fitted smooth, can be appreciated.
+Whether these are, in fact, consistent elevational effects rather than site-specific idiosyncrasies would require corroboration from other data sets.
+
+```{r fig4, echo = FALSE,  fig.cap = paste("Floral abundance HGAM analysis.
+Panels A and B depict mean weighted floral abundance (area of flower cover) by elevation and year, with lines showing the conditional GAM predictions (excluding the random effects of site and species) and points showing original data, color coded to show which sites fell above the tree line.
+Panel A displays the full data set while panel B displays a subset of the range to highlight the prediction curves.
+Panels C and D show the conditional smooth effects (on the link scale) of elevation-by-year and elevation-by-species")}
+knitr::include_graphics("../output/gam_mean_fl_abund_grid.png")
 ```

### 2021-03-17 code/ms1_abundance.Rmd
update

@@
 library(mgcViz)
+library(gratia)
 library(patchwork)
@@
 
-weighted_fl_abund <- fl_abund_sp %>%
+weighted_flfl_abund <- fl_abund_sp %>%
   group_by(site, date, plant.genus) %>%
@@
 # Get mean total weighted floral abundance by BB sp
-weighted_fl_abund_mean <- weighted_fl_abund %>%
+weighted_fl_abund_mean <- weighted_flfl_abund %>%
   group_by(site, bb.sp, year) %>%
@@
 
-### 2.2.2 Preliminary visualizations
+### 2.2.2 Call models
 ```{r}
-## mean floral abundance
-ggplot(abund_sum, aes(elev.mean, mean.fl.abund, color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 8), se = FALSE) +
-  #facet_wrap(~bb.sp, scales = "free_y") +
-  geom_vline(xintercept = 1500)
-
-## log(mean floral abundance)
-ggplot(abund_sum, aes(elev.mean, log(mean.fl.abund), color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 8), se = FALSE) +
-  facet_wrap(~bb.sp) +
-  geom_vline(xintercept = 1500)
-
-## Max floral abundance)
-ggplot(abund_sum, aes(elev.mean, max.fl.abund, color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 8), se = FALSE) +
-  facet_wrap(~bb.sp) +
-  geom_vline(xintercept = 1500)
-
-## Max BB abundance
-ggplot(abund_sum, aes(elev.mean, max.bb.abund, color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 8),  se = FALSE) +
-  facet_wrap(~bb.sp, scales = "free_y") +
-  geom_vline(xintercept = 1500)
-
-## Max  abundance ~ mean BB abundance, to see how interchangeable they are.
-## They are almost perfectly correlated, and I think max abund is more informative
-ggplot(abund_sum, aes(mean.bb.abund, max.bb.abund, color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 8),  se = FALSE) +
-  facet_wrap(~bb.sp, scales = "free_y")
-
-## Are max FL and mean FL similarly interchangeable? Yes.
-ggplot(abund_sum, aes(mean.fl.abund, max.fl.abund, color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 8),  se = FALSE) +
-  facet_wrap(~bb.sp, scales = "free_y")
-
-## Max BB abundance ~ mean FL abundance
-ggplot(abund_sum, aes(log(mean.fl.abund), max.bb.abund)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 8),  se = FALSE) +
-  facet_wrap(~bb.sp, scales = "free_y")
-```
-
-### 2.2.3 Abundance models
-I'm choosing to use site as a random intercept effect in these models because I think that is the correct way to specify a model involving repeated observations (in this case, bumble bee species) within site.
-The effect of this decision is that the elevation smoother is less influenced by individual sites and converges more strongly toward the mean.
-This creates less exciting curves, but I think it is a more principled way to model the data and less prone to overfitting.
-I can have my cake and eat it too, though, by plotting the actual data over the GAM fit so that the strong patterns of certain sites are visible.
-Then I can talk about the conditional elevation effect and the importance of certain peaks that are likely to be idiosyncrasies of my system (i.e.
-it would be overfitting to call them elevation effects).
-
-#### 2.2.3.1 Bumble bees
-```{r}
-# Call model
+# BB
 gam_max_bb_abund_00 <- gam(max.bb.abund ~ 
-                             s(bb.sp, bs = "re") +
-                             s(bb.sp.year, bs = "re") + # fixed intercept for bb.sp:year
-                             s(elev.mean, by = bb.sp, k = 8), # smoother for elev by BB sp.
+                             s(bb.sp, bs = "re") + # random intercept for bb.sp
+                             s(bb.sp.year, bs = "re") + # random intercept for bb.sp:year
+                             s(elev.mean, by = bb.sp, k = 8), # smoother for elev by bb.sp.
                            method = "REML",
@@
 
-gam_max_bb_abund_00_ilink <- family(gam_max_bb_abund_00)$linkinv # get inverse link function
+# FL
+gam_mean_fl_abund_00 <- gam(mean.fl.abund ~
+                              year +
+                              s(bb.sp, bs = "re") +
+                              s(elev.mean, by = bb.sp, k = 8),
+                           method = "REML",
+                           family = "tw",
+                           select = TRUE,
+                           data = abund_sum) %>% getViz()
+```
 
+### 2.2.3 Evaluate model
+```{r}
+# BB
 check.gamViz(gam_max_bb_abund_00) # evaluate model
@@
 
+# FL
+check.gamViz(gam_mean_fl_abund_00) # evaluate model
+gam_mean_fl_abund_00_sum <- summary(gam_mean_fl_abund_00) # summarize model
+print(plot(gam_mean_fl_abund_00, allTerms = TRUE), pages = 1) # preliminary visualization
+```
+
+### 2.2.4 Visualize
+
+#### 2.2.4.1 BB
+```{r}
 # Plot smooths and year effect on the link scale
+
 ## Individual plots
@@
 
-## Arrange significant smooths on grid plot
-grid_bb_abund <- gridPrint(gers_abund, mend_abund, mont_abund, muci_abund, 
+## Arrange smooths on grid plot
+
+grid_bb_abund <- gridPrint(gers_abund, hort_abund, jone_abund, mend_abund, mont_abund, muci_abund, 
                            pasc_abund, prat_abund, psit_abund, pyre_abund, 
@@
 
-pdf("../output/gam_max_bb_abund_grid.pdf", width = 8, height = 6)
+pdf("../output/gam_bb_abund_grid.pdf", width = 8, height = 6)
 plot(grid_bb_abund)
@@
 
-png("../output/gam_max_bb_abund_grid.png", res = 300, width = 8, height = 6, units = "in")
+png("../output/gam_bb_abund_grid.png", res = 300, width = 8, height = 6, units = "in")
 plot(grid_bb_abund)
 dev.off()
+
+# Extract predictions and plot on response scale
+
+gam_max_bb_abund_00_ilink <- family(gam_max_bb_abund_00)$linkinv # get inverse link function
+
+gam_max_bb_abund_00_pred <- predict.gam(gam_max_bb_abund_00, # get model predictions
+                                        type = "terms") %>% 
+  as.data.frame() %>% # convert to data frame
+  as_tibble() %>% # convert to tibble
+  select(bb.sp = 1, bb.sp.year = 2, gers = 3, # rename
+         hort = 4, jone = 5, mend = 6, mont = 7, 
+         muci = 8, pasc = 9, prat = 10, psit = 11, 
+         pyre = 12, soro = 13, telu = 14, wurf = 15) %>%
+  mutate(fit = # get cumulative fit
+           bb.sp + bb.sp.year + gers + 
+           hort + jone + mend + mont + 
+           muci + pasc + prat + psit + 
+           pyre + soro + telu + wurf 
+           ) %>%
+  select(fit) %>%
+  mutate(fit.trans = gam_max_bb_abund_00_ilink(fit)) %>% # back-transform to response scale
+  bind_cols(abund_sum)
+
+ggplot(gam_max_bb_abund_00_pred, aes(elev.mean, fit.trans, linetype = year)) + # map year to linetype
+  geom_point(aes(elev.mean, max.bb.abund, color = tree_line), inherit.aes = FALSE, alpha = 0.5) + # add original data and map tree line to color
+  geom_line(size = 0.5, alpha = 0.5) +
+  facet_wrap(~bb.sp, scales = "free_y") +
+  theme_light() +
+  labs(x = "Elevation (m.s.l.)", 
+       y = bquote("Peak abundance"),
+       color = "Tree line",
+       shape = "Year",
+       linetype = "Year")
+
+ggsave("../output/gam_bb_abund.pdf")
+ggsave("../output/gam_bb_abund.png")
 ```
 
-#### 2.2.3.2 Floral resources
+#### 2.2.4.2 FL
 ```{r}
-gam_mean_fl_abund_00 <- gam(mean.fl.abund ~
-                              s(elev.mean, bb.sp, bs = "fs", k = 8) +
-                              s(elev.mean, year, bs = "fs", k = 8),
-                           method = "REML",
-                           family = "tw",
-                           select = TRUE,
-                           data = abund_sum) %>% getViz()
+# Plot on the link scale
 
-check.gamViz(gam_mean_fl_abund_00) # evaluate model
-gam_mean_fl_abund_00_sum <- summary(gam_mean_fl_abund_00) # summarize model
-print(plot(gam_mean_fl_abund_00, allTerms = TRUE), pages = 1) # preliminary visualization
+## Individual plots
 
-# Plot on the link scale
-elev_effect <- plot(sm(gam_mean_fl_abund_00, 2)) + 
-  l_fitLine(alpha = 0.6, size = 1) + 
-  labs(x = "Elevation (m.s.l.)", y = "Conditional effect (link scale)", color = "Year")
+gers_fl_abund <- plot(sm(gam_mean_fl_abund_00, 2)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "gers")
+hort_fl_abund <- plot(sm(gam_mean_fl_abund_00, 3)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "hort")
+jone_fl_abund <- plot(sm(gam_mean_fl_abund_00, 4)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "jone")
+mend_fl_abund <- plot(sm(gam_mean_fl_abund_00, 5)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "mend")
+mont_fl_abund <- plot(sm(gam_mean_fl_abund_00, 6)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "mont")
+muci_fl_abund <- plot(sm(gam_mean_fl_abund_00, 7)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "muci")
+pasc_fl_abund <- plot(sm(gam_mean_fl_abund_00, 8)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "pasc")
+prat_fl_abund <- plot(sm(gam_mean_fl_abund_00, 9)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "prat")
+psit_fl_abund <- plot(sm(gam_mean_fl_abund_00, 10)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "psit")
+pyre_fl_abund <- plot(sm(gam_mean_fl_abund_00, 11)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "pyre")
+soro_fl_abund <- plot(sm(gam_mean_fl_abund_00, 12)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "soro")
+telu_fl_abund <- plot(sm(gam_mean_fl_abund_00, 13)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "telu")
+wurf_fl_abund <- plot(sm(gam_mean_fl_abund_00, 14)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "wurf")
+year_fl_effect <- plot(pterm(gam_mean_fl_abund_00, 1)) + l_fitPoints() + l_ciBar() + 
+  labs(x = NULL, y = NULL, title = "Year")
 
-bb.sp_effect <- plot(sm(gam_mean_fl_abund_00, 1)) +
-  l_fitLine(alpha = 0.6, size = 1) + 
-  labs(x = "Elevation (m.s.l.)", y = NULL, color = "Species")
+## Arrange significant smooths on grid plot
 
-grid_fl_abund <- gridPrint(elev_effect, bb.sp_effect, ncol = 2)
+grid_fl_abund <- gridPrint(gers_fl_abund, hort_fl_abund, jone_fl_abund, mend_fl_abund, mont_fl_abund, muci_fl_abund, 
+                           pasc_fl_abund, prat_fl_abund, psit_fl_abund, pyre_fl_abund, 
+                           soro_fl_abund, telu_fl_abund, wurf_fl_abund, year_fl_effect, ncol = 4)
 
-pdf("../output/gam_mean_fl_abund_grid.pdf", width = 7, height = 3.5)
+pdf("../output/gam_fl_abund_grid.pdf", width = 8, height = 6)
 plot(grid_fl_abund)
@@
 
-png("../output/gam_mean_fl_abund_grid.png", res = 300, units = "in", width = 7, height = 3.5)
+png("../output/gam_fl_abund_grid.png", res = 300, width = 8, height = 6, units = "in")
 plot(grid_fl_abund)
 dev.off()
+
+# Extract predictions and plot on response scale
+
+gam_mean_fl_abund_00_ilink <- family(gam_mean_fl_abund_00)$linkinv # get inverse link function to transform to the response scale
+
+gam_mean_fl_abund_00_pred <- predict.gam(gam_mean_fl_abund_00, # get model predictions
+                                        type = "terms") %>% 
+  as.data.frame() %>% # convert to data frame
+  as_tibble() %>% # convert to tibble
+  select(year = 1, bb.sp = 2, gers = 3, # rename
+         hort = 4, jone = 5, mend = 6, mont = 7, 
+         muci = 8, pasc = 9, prat = 10, psit = 11, 
+         pyre = 12, soro = 13, telu = 14, wurf = 15) %>%
+  mutate(fit = # get cumulative fit
+           year + bb.sp + gers + 
+           hort + jone + mend + mont + 
+           muci + pasc + prat + psit + 
+           pyre + soro + telu + wurf 
+           ) %>%
+  select(fit) %>%
+  mutate(fit.trans = gam_max_bb_abund_00_ilink(fit)) %>% # back-transform to response scale
+  bind_cols(na.omit(abund_sum)) # the na.omit() drops the rows where floral surveying data were missing
+
+ggplot(gam_mean_fl_abund_00_pred, aes(elev.mean, fit.trans, linetype = year)) + # map year to linetype
+  geom_point(aes(elev.mean, mean.fl.abund, color = tree_line), inherit.aes = FALSE, alpha = 0.5) + # add original data and map tree line to color
+  geom_line(size = 0.5, alpha = 0.5) +
+  facet_wrap(~bb.sp, scales = "free_y") +
+  theme_light() +
+  labs(x = "Elevation (m.s.l.)", 
+       y = bquote("Peak abundance"),
+       color = "Tree line",
+       shape = "Year",
+       linetype = "Year")
+
+ggsave("../output/gam_fl_abund.pdf")
+ggsave("../output/gam_fl_abund.png")
 ```

### 2021-03-17 ms_1/ms_1.Rmd
update

--- 2021-03-15 ms_1/ms_1.Rmd
@@
 
-Climate-induced range shifts involve more than merely tracking suitable temperature envelopes; when species move, they do so in the context of biological communities and under the influence of more than just temperature, and the outcomes of climate-induced range shifts are thus contingent on processes that cannot be reduced to linear elevational or latitudinal temperature gradients.
-Bumble bees (Hymentopera: *Bombus* spp.) are cold-adapted species associated with mountain habitats, and they have already exhibited measurable upslope range shifts that are expected to result in habitat loss, novel competitive interactions, and disrupted pollination networks.
-Interpreting these shifts, however, requires an understanding of the current elevational patterns of bumble bees and the floral mutualists upon which climate change will act.
-In this study, we analyze bumble bee, flowering plant, and interaction records sampled during 904 standardized transect walks conducted over three consecutive years along an elevational gradient in the German Alps ranging from 640 to 2030 meters above sea level.
-
-We demonstrate that mountain bumble bees and flora exhibit distinct peaks of abundance, are distributed in strongly nonlinear patterns in space and time, and that these patterns propagate from the level of species abundances to that of β-diversity, and from the level of community composition to that of biotic interactions.
-In particular, we show that the tree line ecotone is a critical threshold of floral resource availability, a distributional interface between high- and low-elevation bumble bees species, and a zone of rapid acceleration of species and interaction β-diversity.
+The structuring of biological communities along mountain slopes is complex, and the elevational range shifts that are now occurring in response to climate change involve more than merely tracking suitable temperature envelopes.
+When species move, they do so in the context of biological communities, and the outcomes of these movements depend on how and to what extent biotic interactions are reordered.
+Bumble bees (Hymentopera: *Bombus* spp.) are cold-adapted species associated with mountain habitats, and they have already exhibited measurable upslope range shifts that are expected to result in habitat loss, novel competitive interactions, and the rewiring of pollination networks.
+Predicting and interpreting these shifts, however, requires an understanding of the current elevational patterns of bumble bees and the floral mutualists that are being acted upon by climate change.
+In this study, we sampled bumble bees, flowering plants, and their interactions over three consecutive years along an 1400 m elevational gradient in the German Alps.
+Using nonlinear modeling techniques, we analyze the elevational patterns of this community at three levels of biological organization: species abundance, species β-diversity, and interaction β-diversity.
+We demonstrate complex, nonlinear responses to elevation at all three levels of organization.
+In particular, we identify the tree line ecotone as (1) a distributional interface between low/mid- and high-elevation bumble bee species, (2) a threshold above which floral resource availability sharply decreases, and (3) a zone of accelerated turnover of flowering plants and bumble- bee-flower interactions.
 The implications of these findings extend beyond the particular case of bumble bees to demonstrate that linear elevational temperature gradients are ecologically punctuated, and the outcomes of climate-induced range shifts with depend on how species and species-interactions change as they cross --- or fail to cross --- thresholds like the tree line ecotone.
 
-*Keywords*: climate, elevation gradient, pollinator, *Bombus*, alpine
+*Keywords*: climate, elevation gradient, pollinator, *Bombus*, mountain, tree line
 
@@
 
+<!-- This is a paper about mountains. -->
+
+Mountains are unique ecological theaters in which extrinsic climate dynamics interact with intrinsic climate gradation to shape the distributions of species and biotic interactions [@Telwala2013-sc; @CaraDonna2014-nq; @Miller-Struttmann2014-wz; @Rafferty2020-aj].
 The rate of recent climate change is forcing species to adapt ecologically within the constraints of gene pools shaped by cooler climates [@Visser2008-ct].
-As temperatures rise, species are expected to track suitable conditions to higher latitudes and higher elevations, and such effects are already measurable across a broad range of taxa [@Lenoir2015-tj].
-Species range shifts, however, are not merely abstract translations along axes of latitude and elevation; they are movements *in situ*, involving the advent, extinction, and reorganization of biological interactions [@Tylianakis2008-fi; @Blois2013-dz] as well as the concomitant variation of abiotic conditions other than temperature [@Korner1995-mh; @Hodkinson2005-af].
+As temperatures rise, species are expected to track suitable conditions to higher elevations, and such effects are already measurable across a broad range of taxa [@Lenoir2015-tj].
+Elevational range shifts, however, are not merely abstract translations along the linear lapse of temperature generated by mountain slopes; they are movements *in situ*, involving the advent, extinction, and reorganization of biological interactions [@Tylianakis2008-fi; @Blois2013-dz] as well as the concomitant variation of abiotic conditions other than temperature [@Korner1995-mh; @Hodkinson2005-af].
 Thus, when species move along continuous, linear climate gradients, they can experience punctuated and nonlinear ecological transitions.
@@
 At this relatively discrete band within a continuous elevational temperature gradient, plant communities exhibit rapid species turnover [@Descombes2017-id] and declining diversity [@Becker2007-cg] as abiotic conditions become more extreme away from the shelter of tree canopy [@Slatyer1992-ir].
-Thus, species range shifts and the consequences thereof are contingent upon existing patterns of species distributions and biotic interactions.
 
-Bumble bees (Hymenoptera: *Bombus* spp.) are well-studied model organisms in the field of community ecology, but they have more recently become conservation priorities due to well-documented population declines and particular vulnerability to climate change [@Goulson2008-oe; @Soroye2020-zb].
-The genus *Bombus* is believed to have arisen in the mountains of Asia during the cooling climate of the Eocene-Oligocene boundary (~34 mya) [@Hines2008-er], and bumble bees remain characteristically cold-adapted species today [@Heinrich1994-cv], with peak abundance and diversity in mountain ranges and northern latitudes where they are often the dominant guild of flower-visiting insects and the principal pollinators of entomophilous flora [@Williams1998-xi; @Goulson2010-lo].
-Climate-induced range shifts are expected to be especially pronounced in the mountain habitats that host the core of bumble bee biodiversity, together with their floral mutualists, and upward movement of bumble bee species has already been documented [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc].
-Populations of bumble bee species already restricted to the highest elevation zones can be expected to suffer geometric habitat shrinkage and increasing genetic isolation as they move toward mountain peaks; the upward advance of lowland species may introduce novel competitive pressures and disrupt the pollination networks in which bumble bees participate [@Brosi2013-mu; @Ishii2013-sm]; and concurrent elevational [@Lenoir2008-yn] and/or phenological [@Pyke2016-ll] shifts in the floral community could generate mismatches between bumble bee species and their historic floral hosts.
-
-<!-- I either need to beef up the paragraph below or fold it into another one. But I need to get to the point of our study. Which means I need to know what the point is. -->
+Bumble bees (Hymenoptera: *Bombus* spp.) are classic model organisms in the field of community ecology, but they have more recently become conservation priorities due to well-documented population declines and vulnerability to climate change [@Goulson2008-oe; @Soroye2020-zb].
+Bumble bees are characteristically cold-adapted species [@Heinrich1994-cv], with peak abundance and diversity in mountain ranges and northern latitudes, where they are often the dominant guild of flower-visiting insects and the principal pollinators of entomophilous flora [@Williams1998-xi; @Goulson2010-lo].
+Upslope movement of bumble bee species in response to climate change has already been documented, and this is expected to cause the declines or extinction of some bumble bee species [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc].
+From a purely geometric perspective, populations of bumble bee species already restricted to the highest elevation zones can be expected to suffer habitat shrinkage [@Dirnbock2011-jf], possibly accompanied by increasing genetic isolation [though see @Fijen2020-zg].
+But the outcomes of elevational range shifts on bumble bees also hinge upon the current elevational distributions of bumble bees and their floral resources that will be acted upon by climate change.
+The upward advance of lowland bumble bee species, for example, may introduce novel competitive pressures on highland species and rewire the pollination networks in which they participate [@Nagamitsu2010-wi; @Brosi2013-mu; @Ishii2013-sm], and elevational [@Lenoir2008-yn] and/or phenological [@Pyke2016-ll] shifts in the floral community could generate mismatches between bumble bee species and their floral hosts.
+When range shifts are considered in this complex biotic context, it is especially important to understand nonlinearities or thresholds in species-, community-, and interaction-level responses to elevational climate gradients, such as the floristic threshold effect that has been documented at the tree line ecotone [@Descombes2017-id], since these could produce dramatic ecological consequences that would not be predicted by geometric considerations alone.
 
 The outcomes of range shifts in mountain bumble bee and floral communities have obvious significance from a conservation perspective [@Soroye2020-zb], but they also provide a model system in which to study the basic phenomenon of community assembly under climate change, a phenomenon of both historical interest and prospective urgency.
-As bumble bee species move up mountain slopes, they move in relation to plant species and other bumble bee species
-
-Anticipating, interpreting, and potentially mitigating the effects of ranges shifts, however, hinge upon an understanding the current elevational and temporal patterns of the bumble bee and floral communities that will be acted upon by climate change.
-In particular, it is important to detect nonlinearities in species-, community-, and interaction-level responses to elevational climate gradients, such as the floristic threshold effect that has been documented at the tree line ecotone [@Descombes2017-id].
-
-⎄
-In the present study, we approach this topic at three levels of biological organization.
-First, we ask how mountain bumble bee species are distributed in elevation and intra-annual time with respect to one another and with respect to their floral hosts.
-Then, we move from the level of individual species to that of community composition and explore the structure of species turnover (i.e.
-β-diversity) of bumble bees and flora in response to elevation.
-Finally, we proceed to the level of species interactions, investigating the turnover of bee-flower interaction partners through elevation.
-In each of these analyses, we employ flexible modeling techniques capable of capturing nonlinear responses with the goal of identifying where, when, and for which species the impacts of climate-induced range shifts should be expected to be most severe.
+In the present study, set along ~1400 m elevation gradient in the German Alps, we investigate elevational patterns in a community of bumble bees and wildflowers at three levels of biological organization.
+First, we ask how elevation shapes the abundances of bumble bee species and their floral resources.
+Then, we move from the level of individual species to that of community composition and explore the elevational structure of species turnover (i.e.
+β-diversity) in bumble bees and flowering plants.
+Finally, we advance to the level of species interactions, investigating the turnover of bumble- bee-flower interaction partners through elevation.
+In each of these analyses, we employ flexible modeling techniques designed to capture nonlinear responses to elevation, with a particular emphasis on the tree line ecotone as a key threshold that structures the biological communities of mountains slopes.
 
@@
 
-We analyzed the abundance of bumble bees and floral resources using separate hierarchical generalized additive models (HGAM) [@Wood2011-pg; @Pedersen2019-kf].
-We opted in both models to the collapse temporal variation of abundance within sites and focus on elevational patterns.
-For bumble bees, we modeled *peak* abundance at each site as a smooth response to elevation, with bumble bee species as a "by" smoothing factor and a random intercept effect.
-We also included species:year as a random intercept effect to capture annual diffences in species abundance and site as a random intercept effect to avoid overfitting due to pseudoreplication of elevation arising from the measurement of multiple bumble bee species at each site.
-This construction is a modification of the "type I" model form described in @Pedersen2019-kf.
-For floral resources, we modeled *mean* flower cover, weighted per bumble be species, as a smooth response to elevation, with bumble bee species and year as "factor-smooths" and site as a random intercept effect (again to avoid overfitting due to pseudoreplication).
-This construction is a modification of the "type S" model form described in @Pedersen2019-kf.
+We analyzed the abundance of bumble bees and floral resources using separate hierarchical generalized additive models (HGAMs) fit with the Tweedie distribution family (log link function) [@Wood2011-pg; @Pedersen2019-kf].
+We opted in both models to collapse the temporal variation of abundance within sites and focus on elevational patterns.
+For bumble bees, we modeled *peak* abundance at each site as a smooth response to elevation, with bumble bee species as a smoothing factor and a random intercept term.
+We also included species:year as a random intercept term because exploratory analysis indicated that species varied differently across years.
+For floral resources, we modeled *mean* flower cover, weighted per bumble bee species, as a smooth response to elevation, with year and bumble bee species as random intercept terms.
+Both models are follow the "type I" model form described in @Pedersen2019-kf.
 
@@
 
-Our model explained 71.5% of total deviance (adjusted R^2^ = 0.511).
-
+The peak abundance of all bumble bee species responded significantly to elevation (p ≤ 0.04) **(Figures 3, 4)**.
 *B.
@@
 mucidus*, and *B.
-pyrenaeus* exhibited a positive response to elevation and peaked in abundance above the tree line (~ 1500 m.s.l.) (**Figure 3**), consistent with their known affinity for high elevation habitats [@Rasmont2010-ci].
+pyrenaeus* exhibited a positive response to elevation and peaked in abundance above the tree line (~ 1500 m.s.l.), consistent with their known affinity for high elevation habitats [@Rasmont2010-ci].
 The abundance of *B.
-soroensis* also tended to increase with elevation and peaked above the tree line, but it was also abundant at lower elevations.
+soroeensis* also tended to increase with elevation and peaked above the tree line, but it was also abundant at lower elevations.
 In contrast, the abundance of *B.
-pascorum* --- the most abundant species in our study system --- declined steeply with elevation, and a similar pattern was seen for *B.
-psithyrus*.
+hortorum*, *B.
+pascorum*, and *B.
+psithyrus* declined with elevation.
+This effect was especially pronounced in *B.
+pascuorum*, which was also the most abundant species in our study system overall.
+*B.
+gerstaeckeri*, *B.
+jonellus*, *B.
+terrestris-lucorum*, and *B.
+wurflenii* peaked at mid-elevation below the tree line, while *B.
+pratorum* was unique in exhibiting a sharp peak of abundance directly *at* the tree line.
 
-while *B. gerstaeckeri* was sparsely distributed at mid-elevation sites.
-
-```{r fig3, echo = FALSE,  fig.cap = paste("Peak bumble bee abundance by elevation and year.
-Curves shows the conditional GAM predictions excluding the site random effect, and points show original data.
-Dotted vertical lines indicate the approximate location of the tree line at our study sites.
-Note different scales of y-axes.")}
-knitr::include_graphics("../output/gam_max_bb_abund2.png")
+```{r fig3, echo = FALSE,  fig.cap = paste("Fitted GAM smooths of bumble bee abundance on elevation, faceted by bumble bee species and plotted over original data.
+Data points are color-coded to indicate whether they represent sites above (red) or below (blue) the tree line.
+Note the different scales of the y-axes.")}
+knitr::include_graphics("../output/gam_bb_abund.png")
 ```
 
-```{r fig4, echo = FALSE,  fig.cap = paste("Mean weighted floral abundance (area of flower cover) by elevation and year.
-Lines shows the conditional GAM predictions excluding the random effects of site and species, and points show original data.
-The right panel displays the full range of the data while the left panel displays a subset of the range, omitting extreme values of flower cover to highlight the prediction curves.")}
-knitr::include_graphics("../output/gam_mean_fl_abund_zoom.png")
+```{r fig4, echo = FALSE,  fig.cap = paste("Conditional effect of elevation on the peak abundance of each bumble bee species.
+Effects are plotted on the link scale, and dashed lines depict 95\\% confidence intervals.")}
+knitr::include_graphics("../output/gam_bb_abund_grid.png")
+```
+
+Mean floral resource abundance likewise responded significantly (p << 0.0001) to elevation for all bumble bee species **(Figures 5, 6)**.
+*B.
+mendax*, *B.
+monticola*, *B.
+mucidus*, *B.
+soroeensis*, and *B.
+wurflenii* exhibited a weakly to strongly bimodal pattern of flora resource abundance, with peaks around 1000 m.s.l.
+and 1600 m.s.l., the latter peak corresponding closely to the location of the tree line.
+*B.
+jonellus* and *B.
+pyrenaeus* exhibited a unimodal pattern peaking near 1600 m.s.l.
+The remaning species --- *B.
+gerstaeckeri*, *B.
+hortorum*, *B.
+pascuorum*, *B.
+pratorum*, *B.
+psithyrus*, and *B.
+terrestris-lucorum* experienced linear or quadratic decline in floral resource abundance with increasing elevation.
+Floral resource abundance was low for all species at elevations above ~1600 m.s.l.
+
+```{r fig5, echo = FALSE,  fig.cap = paste("Fitted GAM smooths of floral abundance on elevation, faceted by bumble bee species and plotted over original data.
+Data points are color-coded to indicate whether they represent sites above (red) or below (blue) the tree line.
+Note the different scales of the y-axes.")}
+knitr::include_graphics("../output/gam_fl_abund.png")
+```
+
+```{r fig6, echo = FALSE,  fig.cap = paste("Conditional effects of elevation on mean flower cover by bumble bee species and year.
+Effects are plotted on the link scale and dashed lines depict 95\\% confidence intervals.")}
+knitr::include_graphics("../output/gam_fl_abund_grid.png")
 ```
@@
 
-```{r fig5, echo = FALSE,  fig.cap = paste("Species-level (A) and interaction-level (B, C) plotted against elevational difference.
+```{r fig7, echo = FALSE,  fig.cap = paste("Species-level (A) and interaction-level (B, C) plotted against elevational difference.
 Each point represents the β-diversity (or β-diversity partition) between a pair of sites, and lines represent the overall relationship between β-diversity and elevation difference using binomial regression smooths.
@@
 
-```{r fig6, echo = FALSE,  fig.cap = paste("GDM splines of species-level and interaction-level β-diversity in response to elevational, controlling for geographic proximity and number of sampling dates.
+```{r fig8, echo = FALSE,  fig.cap = paste("GDM splines of species-level and interaction-level β-diversity in response to elevational, controlling for geographic proximity and number of sampling dates.
 The maximum height of each spline represents a given variable’s partial effect on β-diversity (its effect when covariates are held constant), and the shape of each spline represents to the rate of species turnover as it varies along the corresponding gradient, with steeper parts of the curve indicating regions of the gradient over which species/interaction turnover is more rapid.
@@
 
-The core finding of our study, evident in each of our analyses, is that a linear elevation gradient can generate nonlinear biological responses in terms of both abundance and β-diversity, and the latter both at the level of community composition and at the level of species interactions.
+The core finding of our study, evident in each of our analyses, is that a linear elevation gradient can generate complex and nonlinear biological responses in a mountain bumble- bee-wildflower community in terms of both abundance and β-diversity, and the latter both at the level of community composition and at the level of species interactions.
+As a rule, these responses are punctuated by the tree line ecotone, which constitutes a discrete ecological threshold along the linear elevation gradient of a mountain slope.
 
-Arguably the most salient pattern in our data is the unique importance of the elevation band falling at about 1500 m, which corresponds to the lower edge of the tree line ecotone in the Berchtesgadener Alps [@Kostler1970-ya; @Mayer1970-zn] and specifically that of the elevational transects used in our study.
-The floristic importance of the tree line ecotone has been noted in previous work [@Pellissier2010-kg; @Descombes2017-id].
-By studying the floral community through the lens of foraging bumble bees, however, we provide a functional extension of this floristic pattern, demonstrating that the tree line ecotone comprises (1) a boundary above which floral resource availability sharply declines, (2) a distributional interface between high-elevation bumble bees and their low-elevation or habitat-generalist counterparts, and (3) a zone of rapid acceleration of species and interaction β-diversity.
+We found a clear distinction between bumble bee species with low- or mid-elevation abundance peaks (*B.
+gerstaeckeri*, *B.
+hortorum*, *B.
+jonellus*, *B.
+pascorum*, *B.
+pratorum*, *B.
+psithyrus*, *B.
+terrestris-lucorum*, *B.
+wurflenii*) and those with high-elevation abundance peaks (*B.
+mendax*, *B.
+monticola*, *B.
+mucidus*, *B.
+pyrenaeus*, and *B.
+soroeensis*).
+The tree line forms an important zone of overlap at which the declining abundance of low- and mid-elevation species intersects with the rising abundance of high-elevation species.
+The tree line was also important in the elevational structure of floral resource abundance.
+For some (mostly high-elevation) bumble bee species, the tree line marked a distinct peak of floral resource availability, often mirrored by a second peak at lower elevation (~1000 m.s.l.).
+It is notable that, for some of the high-elevation species that exhibited this bimodal pattern of floral resource abundance --- *B.
+mendax*, *B.
+monticola*, and *B.
+mucidus* --- the lower peak of floral resource abundance is almost entirely outside their elevational range.
+The tree line, therefore, functions as critical foraging habitat for high-elevation bumble bees.
+Above the tree line, flora resource availability declined sharply as flower-rich grassland and shrubland transitioned to sparse alpine scree.
+The observation that the upslope advancement of alpine tree lines often lags behind climate warming [@Dullinger2004-qu] suggests that high elevation bumble bees could move upslope faster than their floral resources, increasingly confined to resource-poor habitat, though the opposite pattern could also occur if the climate-tracking of bumble bees is limited by other factors, such as nest site availability [@Marshall2020-pc].
+Moreover, several studies have found that the introduction or removal of bumble bee species from a community can have strong effects on the fitness of other species and on patterns of floral visitation [@Nagamitsu2010-wi; @Brosi2013-mu; @Ishii2013-sm]; if elevational range shifts introduce new competitive pressures via the encroachment of low-elevation species into the historic ranges of high-elevation species, this effect might be expected to be most pronounced at the tree line.
 The significance of the tree line ecotone as a threshold of floral resource availability and as a zone of overlapping bumble bee distributions corroborates the recent findings of @Minachilis2020-xd in the Mediterranean system of Mount Olympus, indicating that these patterns are likely to generalize well across mountain ecosystems.
-Our finding that interaction β-diversity is also punctuated by the tree line ecotone follows naturally from these compositional patterns, and we expect that the β-diversity patterns found in our system will prove similarly generalizable.
 
-The observation that the upslope advancement of alpine tree lines often lags behind climate warming [@Dullinger2004-qu] suggests that high elevation bumble bees could move upslope faster than their floral resources, though the opposite pattern could also occur if the climate-tracking of bumble bees is limited by other factors, such as nest site availability [@Marshall2020-pc].
-A somewhat unexpected finding of our study is that high-elevation bumble bee species (*B.
-mucidus*, *B.
-monticola*, *B.
-mendax*, and *B.
-pyrenaeus*) were associated with floral hosts that peaked in abundance at much lower elevations, largely outside the range of their bumble bee visitors.
-As the upper boundary of high floral resource abundance, the lower edge of the tree line ecotone likely functions as critical foraging habitat for high-elevation bumble bees.
-Future warming could force these species farther into resource-poor alpine grasslands while at the same time increasing resource competition at the tree line as dominant low-elevation species, like *B.
-pascuorum*, extend their ranges upslope.
-
-Our analyses of species and interaction β-diversity reveal a striking contrast between a relative stable bumble bee community and an extremely dynamic floral community.
-The dominance of total interaction β-diversity by floral species turnover is consistent with the findings of @Simanonok2014-ou in the Rocky Mountains of North America, and it is perhaps a predictable relationship given the exceptional cold-hardiness of bumble bees and the sensitivity of plants to temperature gradients.
-Nevertheless, it highlights the behavioral challenge that bumble bees have presumably faced for the entirety of their evolutionary history as mountain florivores: how to forage efficiently in a diverse floral community that turns over rapidly in both space and time.
-Indeed, the remarkable flexibility and intelligence of bumble bee foraging [@Heinrich1979-ep; @Loukola2017-bt], which has been interpreted as a side-effect of their sociality [@Dukas1991-ig], could be interpreted alternatively as an adaptation to precisely this problem of foraging in the context of extreme β-diversity.
+The tree line not only punctuated patterns of abundance but also those of species-level and interaction-level β-diversity.
+While the turnover rate of bumble bee species was low across the whole elevation gradient and showed no response to the tree line, the turnover rate of the floral community was an order of magnitude higher and exhibited a marked inflection point at the tree line, above which it increased steeply.
+This floristic importance of the tree line has been noted in previous work [@Pellissier2010-kg; @Descombes2017-id], but our study provides a functional extension to this pattern, demonstrating that the tree line also marks a zone of rapid acceleration in the turnover of bumble- bee-flower interactions that is evidently driven by the underlying turnover of the floral community.
+The picture that emerges is a striking contrast between a relative stable bumble bee community and an extremely dynamic floral community.
+The dominance that we observed of total interaction β-diversity by floral species turnover is consistent with the findings of @Simanonok2014-ou in the Rocky Mountains of North America, and it is perhaps unsurprising given the exceptional cold-hardiness of bumble bees and the sensitivity of plants to temperature gradients.
+Nevertheless, it highlights the behavioral challenge that bumble bees have presumably faced for the entirety of their evolutionary history as mountain florivores: how to forage efficiently in a diverse floral community that turns over so rapidly.
+Indeed, the remarkable flexibility and intelligence of bumble bee foraging [@Heinrich1979-ep; @Loukola2017-bt], which have been interpreted as side-effects of their sociality [@Dukas1991-ig], could be interpreted alternatively as an adaptation to precisely this problem of foraging in the context of extreme β-diversity.
 A fascinating question that could not be answered by our sampling approach is whether mountain bumble bees adaptively forage up- or downslope, as suggested by @Lundberg1980-qk.
@@
 
-# CONCLUSIONS
-
-Mountains are unique ecological theaters in which extrinsic temporal climate change interacts with intrinsic elevational climate gradation to alter the spatial and temporal distributions of species and interactions [@Telwala2013-sc; @CaraDonna2014-nq; @Miller-Struttmann2014-wz; @Rafferty2020-aj].
-The central finding of our study, that elevation gradients are characterized by nonlinear responses of species, communities, and interaction networks --- punctuated dramatically by the tree line ecotone --- is almost certainly not an idiosyncrasy of bumble bees and their floral hosts but rather a pattern emerging from fundamental physical constraints that should be expected to obtain broadly across the biotic communities of mountains worldwide.
-Species interactions along tree line ecotones, therefore, deserve special consideration in research and conservation management, since climate-induced range shifts at the tree line ecotone should be expected to have disproportionate ecological consequences relative to range shifts at other elevation bands.
+The central finding of our study, that elevation gradients are characterized by complex and nonlinear responses of species, communities, and biotic interactions --- punctuated by the tree line --- is almost certainly not an idiosyncrasy of bumble bees and their floral hosts but rather a pattern emerging from fundamental physical constraints that should be expected to obtain broadly across the biotic communities of mountains worldwide.
+Species interactions along tree line ecotones, therefore, deserve special consideration in research and conservation management, since climate-induced range shifts at tree lines should be expected to have disproportionate ecological consequences relative to range shifts at other elevation bands.
 
@@
 <!-- - What do these findings tell use about climate change and range shifts in mountain ecosystems more generally? -->
-<!-- - Do our results suggest any management strategies? -->
-
-<!-- Retreating to higher elevation in response to climatic warming is not a new phenomenon in the evolutionary history of bumble bees, but the rate of modern climate change, together with the interacting effects of human land use, are predicted to [@Soroye2020-zb; @Sirois-Delisle2018-vi]  -->
-
 <!-- though as a matter of applied conservation it is perhaps doubtful that local interventions could forestall for long the consequences of a process occurring at a global scale. -->

### 2021-03-23 code/ms2_bias_niche.Rmd
a new take on ms1 analysis
fixed nomenclature issue

--- 2021-03-09 code/bias_niche.Rmd
@@
   filter(caste != "male") %>% # we will ignore males 
+  filter(!bb.sp %in% c("humi", "hypn")) %>% # drop rare species
   group_by(site, date, bb.sp, plant.sp) %>%
@@
 ## 1.2 Binomial multi-level regression models
+### 1.2.1 Choose priors
+```{r}
+# For starters, I'm just going to replace the flat prior on the fixed effects we a weakly regularizing prior (on the logit scale)
+brm_pvis_prior <- prior(normal(0, 1.5), 
+                        class = b) 
+```
+
+### 1.2.2 Call models
+All these run cleanly.
 ```{r}
@@
 ### Bernoulli with binarized visitation
-brm_pvis00 <- brm(as.integer(visits.bin) ~
+brm_pvis_00 <- brm(as.integer(visits.bin) ~
                     bb.sp.abund*perc.flower.cover + # fixed term: because more abundant flowers will tend to be more visited,and more abundance bees visit more
@@
                   family = "bernoulli", 
+                  prior = brm_pvis_prior, # weakly regularizing
                   file = "../output/brm_pvis_ktype_00",  # for running locally
@@
 
-brm_pvis01 <- brm(as.integer(visits.bin) ~
+brm_pvis_01 <- brm(as.integer(visits.bin) ~
                     bb.abund*perc.flower.cover + # fixed term: because more abundant flowers will tend to be more visited,and more abundance bees visit more
@@
                   family = "bernoulli", 
+                  prior = brm_pvis_prior, # weakly regularizing
                   file = "../output/brm_pvis_ktype_01",  # for running locally
@@
 But, when I want to make inference about the tongue length classes, I do need the species-level effect.
-brm_pvis02 <- brm(as.integer(visits.bin) ~
+brm_pvis_02 <- brm(as.integer(visits.bin) ~
                     bb.sp.abund*perc.flower.cover + # fixed term: because more abundant flowers will tend to be more visited,and more abundance bees visit more
@@
                   family = "bernoulli", 
+                  prior = brm_pvis_prior, # weakly regularizing
                   file = "../output/brm_pvis_ktype_02",  # for running locally
@@
 # Summary
-summary(brm_pvis00)
-summary(brm_pvis01)
-summary(brm_pvis02)
+summary(brm_pvis_00)
+summary(brm_pvis_01)
+summary(brm_pvis_02)
 
 # Posterior predictive checks
-pp_check(brm_pvis00)
-pp_check(brm_pvis01)
-pp_check(brm_pvis02)
+pp_check(brm_pvis_00)
+pp_check(brm_pvis_01)
+pp_check(brm_pvis_02)
+
+# MCMC convergence
+plot(brm_pvis_00)
+plot(brm_pvis_01)
+plot(brm_pvis_02)
 
 # Bayes R2
-bayes_R2(brm_pvis00)
-bayes_R2(brm_pvis01)
-bayes_R2(brm_pvis02)
+bayes_R2(brm_pvis_00)
+bayes_R2(brm_pvis_01)
+bayes_R2(brm_pvis_02)
 ```
@@
 ```{r eval=FALSE, include=FALSE}
-get_variables(brm_pvis00)
-conditional_effects(brm_pvis00)
-mcmc_plot(brm_pvis00, pars = c("^r_ktype:tongue.*Intercept]$"))
-mcmc_plot(brm_pvis00, pars = c("^r_ktype:tongue.*perc.flower.cover]$"))
-mcmc_plot(brm_pvis00, pars = c("^r_ktype:bb.sp.*Intercept]$"))
-mcmc_plot(brm_pvis00, pars = c("^r_ktype:bb.sp.*perc.flower.cover]$"))
-mcmc_plot(brm_pvis00, pars = c("^r_ktype:site.*Intercept]$"))
-mcmc_plot(brm_pvis00, pars = c("^r_ktype:site.*perc.flower.cover]$"))
-
-get_variables(brm_pvis01)
-conditional_effects(brm_pvis01)
-mcmc_plot(brm_pvis01, pars = c("^r_ktype"))
-mcmc_plot(brm_pvis01, pars = c("^r_ktype:tongue.*perc.flower.cover]$"))
-mcmc_plot(brm_pvis01, pars = c("^r_ktype:bb.sp.*Intercept]$"))
-mcmc_plot(brm_pvis01, pars = c("^r_ktype:site.*Intercept]$"))
-mcmc_plot(brm_pvis01, pars = c("^r_ktype:bb.sp.*perc.flower.cover]$"))
-mcmc_plot(brm_pvis01, pars = c("^r_ktype:site.*Intercept]$"))
-mcmc_plot(brm_pvis01, pars = c("^r_ktype:site.*perc.flower.cover]$"))
-
-get_variables(brm_pvis02)
-conditional_effects(brm_pvis02)
-mcmc_plot(brm_pvis02, pars = c("^r_ktype"))
-mcmc_plot(brm_pvis02, pars = c("^r_ktype:bb.sp.*Intercept]$"))
-mcmc_plot(brm_pvis01, pars = c("^r_ktype:site.*Intercept]$"))
-mcmc_plot(brm_pvis02, pars = c("^r_ktype:bb.sp.*perc.flower.cover]$"))
-mcmc_plot(brm_pvis02, pars = c("^r_ktype:site.*Intercept]$"))
-mcmc_plot(brm_pvis02, pars = c("^r_ktype:site.*perc.flower.cover]$"))
-
-get_variables(brm_pvis11)
-conditional_effects(brm_pvis11)
-mcmc_plot(brm_pvis11, pars = c("^r_ktype:year"))
-
-get_variables(brm_pvis21)
-conditional_effects(brm_pvis21)
-mcmc_plot(brm_pvis21, pars = c("^r_ktype"))
+get_variables(brm_pvis_00)
+conditional_effects(brm_pvis_00)
+mcmc_plot(brm_pvis_00, pars = c("^r_ktype:tongue.*Intercept]$"))
+mcmc_plot(brm_pvis_00, pars = c("^r_ktype:tongue.*perc.flower.cover]$"))
+mcmc_plot(brm_pvis_00, pars = c("^r_ktype:bb.sp.*Intercept]$"))
+mcmc_plot(brm_pvis_00, pars = c("^r_ktype:bb.sp.*perc.flower.cover]$"))
+mcmc_plot(brm_pvis_00, pars = c("^r_ktype:site.*Intercept]$"))
+mcmc_plot(brm_pvis_00, pars = c("^r_ktype:site.*perc.flower.cover]$"))
+
+get_variables(brm_pvis_01)
+conditional_effects(brm_pvis_01)
+mcmc_plot(brm_pvis_01, pars = c("^r_ktype*"))
+mcmc_plot(brm_pvis_01, pars = c("^r_ktype:site.*Intercept]$"))
+mcmc_plot(brm_pvis_01, pars = c("^r_ktype:site.*Intercept]$"))
+mcmc_plot(brm_pvis_01, pars = c("^r_ktype:site.*perc.flower.cover]$"))
+
+get_variables(brm_pvis_02)
+conditional_effects(brm_pvis_02)
+mcmc_plot(brm_pvis_02, pars = c("^r_ktype"))
+mcmc_plot(brm_pvis_02, pars = c("^r_ktype:bb.sp.*Intercept]$"))
+mcmc_plot(brm_pvis_02, pars = c("^r_ktype:site.*Intercept]$"))
+mcmc_plot(brm_pvis_02, pars = c("^r_ktype:bb.sp.*perc.flower.cover]$"))
+mcmc_plot(brm_pvis_02, pars = c("^r_ktype:site.*Intercept]$"))
+mcmc_plot(brm_pvis_02, pars = c("^r_ktype:site.*perc.flower.cover]$"))
 ```
@@
 ```{r}
-brm_pvis00_plotframe1 <- brm_pvis00 %>%
+brm_pvis_00_plotframe1 <- brm_pvis_00 %>%
   spread_draws(b_Intercept, `r_ktype:tongue`[condition, term]) %>% # the back ticks are necessary to escape the :
@@
 
-ggplot(brm_pvis00_plotframe1, aes(condition_mean_trans, tongue, 
+ggplot(brm_pvis_00_plotframe1, aes(condition_mean_trans, tongue, 
                                  xmin = .lower_trans, xmax = .upper_trans,
@@
 
-ggsave(file = "../output/brm_pvis00_plot1.pdf", width = 6, height = 5)
-ggsave(file = "../output/brm_pvis00_plot1.png", width = 6, height = 5)
+ggsave(file = "../output/brm_pvis_00_plot1.pdf", width = 6, height = 5)
+ggsave(file = "../output/brm_pvis_00_plot1.png", width = 6, height = 5)
 ```
@@
 ```{r}
-brm_pvis02_plotframe2 <- brm_pvis02 %>%
+brm_pvis_02_plotframe2 <- brm_pvis_02 %>%
   spread_draws(b_Intercept, `r_ktype:bb.sp`[condition, term]) %>% # the back ticks are necessary to escape the :
@@
 
-ggplot(brm_pvis02_plotframe2, aes(condition_mean_trans, bb.sp, 
+ggplot(brm_pvis_02_plotframe2, aes(condition_mean_trans, bb.sp, 
                                  xmin = .lower_trans, xmax = .upper_trans,
@@
 
-ggsave(file = "../output/brm_pvis02_plot1.pdf", width = 8, height = 8)
-ggsave(file = "../output/brm_pvis02_plot1.png", width = 8, height = 8)
+ggsave(file = "../output/brm_pvis_02_plot1.pdf", width = 8, height = 8)
+ggsave(file = "../output/brm_pvis_02_plot1.png", width = 8, height = 8)
 ```
@@
 ```{r eval=FALSE, include=FALSE}
-brm_pvis00_plotframe3 <- brm_pvis00 %>%
+brm_pvis_00_plotframe3 <- brm_pvis_00 %>%
   spread_draws(b_perc.flower.cover, `r_ktype:tongue`[condition, term]) %>% # the back ticks are necessary to escape the :
@@
 
-ggplot(brm_pvis00_plotframe3, aes(condition_slope, tongue, color = tongue)) +
+ggplot(brm_pvis_00_plotframe3, aes(condition_slope, tongue, color = tongue)) +
   stat_pointinterval() +
@@
 
-ggsave(file = "../output/brm_pvis00_plot3.png", width = 7, height = 5.5)
+ggsave(file = "../output/brm_pvis_00_plot3.png", width = 7, height = 5.5)
 ```
@@
 ```{r}
-brm_pvis01_plotframe1 <- brm_pvis01 %>%
+brm_pvis_01_plotframe1 <- brm_pvis_01 %>%
   spread_draws(b_Intercept, r_ktype[condition, term]) %>% # the back ticks are necessary to escape the :
@@
 
-ggplot(brm_pvis01_plotframe1, aes(condition_mean_trans, ktype, 
+ggplot(brm_pvis_01_plotframe1, aes(condition_mean_trans, ktype, 
                                  xmin = .lower_trans, xmax = .upper_trans)) +
@@
 
-ggsave(file = "../output/brm_pvis01_plot1.pdf", width = 6, height = 7)
+ggsave(file = "../output/brm_pvis_01_plot1.pdf", width = 6, height = 7)
 
@@
 ```{r eval=FALSE, include=FALSE}
-brm_pvis01_plotframe2 <- brm_pvis01 %>%
+brm_pvis_01_plotframe2 <- brm_pvis_01 %>%
   spread_draws(b_perc.flower.cover, r_ktype[condition, term]) %>% # the back ticks are necessary to escape the :
@@
 
-ggplot(brm_pvis01_plotframe2, aes(condition_slope, ktype)) +
+ggplot(brm_pvis_01_plotframe2, aes(condition_slope, ktype)) +
   stat_pointinterval() +
@@
 
-ggsave(file = "../output/brm_pvis01_plot2.png", width = 7, height = 5.5)
+ggsave(file = "../output/brm_pvis_01_plot2.png", width = 7, height = 5.5)
 ```
@@
   filter(caste != "male") %>% # we will ignore males 
-  filter(!bb.sp %in% c("humi")) %>% # humi is too rare
+  filter(!bb.sp %in% c("hypn", "humi")) %>% # hypn and humi are too rare
   #group_by(bb.sp, plant.sp.abb) %>%
@@
 bb_traits2 <- bb_traits %>%
-  filter(bb.sp != "humi") %>%
+  filter(!bb.sp %in% c("humi", "hypn")) %>% # drop rare species
   select(bb.sp, pbl.w.class2) %>%
@@
   geom_text(aes(label = bb.sp), key_glyph = "rect") +
-  geom_text_repel(data = filter(pcoa_horn2.fort1, abs(MDS1) >= 0.22 | abs(MDS2) >= 0.22), 
+  geom_text_repel(data = filter(pcoa_horn2.fort1, abs(MDS1) >= 0.2 | abs(MDS2) >= 0.2), 
             aes(MDS1, MDS2, label = plant.genus), inherit.aes = FALSE, color = "gray40") +
-  geom_point(data = filter(pcoa_horn2.fort1, abs(MDS1) >= 0.22 | abs(MDS2) >= 0.22),
+  geom_point(data = filter(pcoa_horn2.fort1, abs(MDS1) >= 0.2 | abs(MDS2) >= 0.2),
             aes(MDS1, MDS2, shape = k.type.ss), inherit.aes = FALSE, color = "gray40", size = 3) +
@@
 
-ggsave("../output/bb_range.pdf", width = 6, height = 5)
+ggsave("../output/bb_range_ms2.pdf", width = 6, height = 5)
 ```
@@
 niche_data <- visitation %>%
-  #filter(!bb.sp %in% c("humi", "muci", "gers")) %>%
   left_join(fl_traits) %>%
@@
   filter(level == "genus")
-
-#write_csv(niche_pool, "../hpc_packets/niche_pool.csv")
-
 ```
@@
 # Beta regression is probably the right way to do this; a gaussian model gives almost identical results, though, since our pooled data are approximately gaussian
-niche_gam00 <- gam(mean.niche.overlap ~
+niche_gam_00 <- gam(mean.niche.overlap ~
                      level + 
@@
 
-check.gamViz(niche_gam00)
-summary(niche_gam00)
-print(plot.gamViz(niche_gam00, allTerms = TRUE), pages = 1)
+check.gamViz(niche_gam_00)
+summary(niche_gam_00)
+print(plot.gamViz(niche_gam_00, allTerms = TRUE), pages = 1)
 ```
@@
 # Get predictions using tidymv
-niche_gam00_pred <- get_gam_predictions(niche_gam00, elev.mean) %>%
+niche_gam_00_pred <- get_gam_predictions(niche_gam_00, elev.mean) %>%
   as_tibble() %>%
@@
 # Plot smooths
-ggplot(niche_gam00_pred, aes(elev.mean, mean.niche.overlap, color = level)) +
+ggplot(niche_gam_00_pred, aes(elev.mean, mean.niche.overlap, color = level)) +
   geom_point() +
@@
 
-ggsave(file = "../output/niche_gam00_plot1.pdf", width = 5, height = 4)
-ggsave(file = "../output/niche_gam00_plot1.png", width = 5, height = 4)
+ggsave(file = "../output/niche_gam_00_plot1.pdf", width = 5, height = 4)
+ggsave(file = "../output/niche_gam_00_plot1.png", width = 5, height = 4)
 
 # Plot intercept effect of level
-plot(pterm(niche_gam00, 1))
-
-plotDiff(sm(niche_gam00, 1), sm(niche_gam00, 3)) +
+plot(pterm(niche_gam_00, 1))
+
+plotDiff(sm(niche_gam_00, 1), sm(niche_gam_00, 3)) +
   l_ciPoly() + 
@@
 ```{r}
-niche_gam20 <- gam(mean.niche.overlap ~
+niche_gam_20 <- gam(mean.niche.overlap ~
                      tongue.diff +
@@
 
-check.gamViz(niche_gam20)
-summary(niche_gam20)
-print(plot(niche_gam20, allTerms = TRUE), pages = 1)
+check.gamViz(niche_gam_20)
+summary(niche_gam_20)
+print(plot(niche_gam_20, allTerms = TRUE), pages = 1)
 ```
@@
 # Get predictions
-niche_gam20_pred <- get_gam_predictions(niche_gam20, elev.mean) %>%
+niche_gam_20_pred <- get_gam_predictions(niche_gam_20, elev.mean) %>%
   as_tibble() %>%
@@
 # Plot smooths
-ggplot(niche_gam20_pred, aes(elev.mean, mean.niche.overlap_trans, color = tongue.diff)) +
+ggplot(niche_gam_20_pred, aes(elev.mean, mean.niche.overlap_trans, color = tongue.diff)) +
   geom_ribbon(aes(elev.mean, mean.niche.overlap_trans, 
@@
 
-ggsave(file = "../output/niche_gam20_plot1.pdf", width = 5, height = 4)
-ggsave(file = "../output/niche_gam20_plot1.png", width = 5, height = 4)
+ggsave(file = "../output/niche_gam_20_plot1.pdf", width = 5, height = 4)
+ggsave(file = "../output/niche_gam_20_plot1.png", width = 5, height = 4)
 
 # Plot intercept effect of tongue.diff
-plot(pterm(niche_gam20, 1))
-```
-
--------------------------------------
-# Old
-
-## 3.6 Genus-level niche overlap ~ tongue.pair + elevation
-### 3.6.1 Model
-```{r}
-niche_gam10 <- gam(mean.niche.overlap ~
-                     tongue.pair +
-                     s(elev.mean, by = tongue.pair, k = 8, m = 2),
-                   method = "REML",
-                   family = betar(eps = 0.001),
-                   select = TRUE,
-                   data = niche_overlap_tongues) %>% getViz()
-
-check.gamViz(niche_gam10)
-summary(niche_gam10)
-print(plot(niche_gam10, allTerms = TRUE), pages = 1)
-```
-
-### 3.6.2 Visualize
-```{r}
-# Get predictions using tidymv
-niche_gam10_pred <- get_gam_predictions(niche_gam10, elev.mean) %>%
-  as_tibble() %>%
-  mutate(mean.niche.overlap_trans = inv_logit_scaled(mean.niche.overlap),
-         CI_upper_trans = inv_logit_scaled(CI_upper),
-         CI_lower_trans = inv_logit_scaled(CI_lower))
-
-# Plot smooths
-ggplot(niche_gam10_pred, aes(elev.mean, mean.niche.overlap_trans, color = tongue.pair)) +
-  geom_ribbon(aes(elev.mean, mean.niche.overlap_trans,
-                  ymax = CI_upper_trans, ymin = CI_lower_trans,
-                  fill = tongue.pair),
-              alpha = 0.3, inherit.aes = FALSE) +
-  geom_line() +
-  ylab("Niche overlap") +
-  xlab("Elevation") +
-  theme_light()
-
-ggsave(file = "../output/niche_gam10_plot1.pdf", width = 5, height = 4)
-ggsave(file = "../output/niche_gam10_plot1.png", width = 5, height = 4)
-```
-
-### 3.5.1 Model
-```{r}
-brm_niche00 <- brm(niche.overlap ~ 1 +
-                     level + # fixed effect of level 
-                     (1 | site) + # varying intercept for site
-                     (1 | sp.pair), # varying intercept for sp.pair
-                   family = "zero_one_inflated_beta", # for prop data with 0s and 1s
-                   file = "../output/brm_niche_00", # local
-                   #file = "./brm_niche_00", # HPC
-                   iter = 2000,
-                   chains = 4,
-                   data = niche_pool)
-
-pp_check(brm_niche00)
-bayes_R2(brm_niche00)
-
-get_variables(brm_niche00)
-conditional_effects(brm_niche00)
-mcmc_plot(brm_niche00, pars = c("^r_sp.pair\\[mend*"))
-mcmc_plot(brm_niche00, pars = c("^r_site*"))
-```
-
-### 3.5.2 Vizualize
-```{r}
-brm_niche00_plotframe1 <- brm_niche00 %>%
-  spread_draws(b_Intercept, b_levelgenus, b_levelmorphotype) %>%
-  mutate(b_levelgenus = b_Intercept + b_levelgenus,
-         b_levelmorphotype = b_Intercept + b_levelmorphotype) %>%
-  pivot_longer(-c(.chain, .iteration, .draw), names_to = "level", values_to = "value") %>%
-  group_by(level) %>%
-  point_interval(value, .width = c(0.95, 0.66)) %>% # calculate credible intervals
-  mutate(value_trans = inv_logit_scaled(value), # exponentiating the estimates gets us back to the response scale (negbinomial uses log link function)
-         .lower_trans = inv_logit_scaled(.lower),
-         .upper_trans = inv_logit_scaled(.upper)) %>%
-  mutate(level = case_when(
-    level == "b_Intercept" ~ "family",
-    level == "b_levelmorphotype" ~ "morphotype",
-    level == "b_levelgenus" ~ "genus"
-  )) %>%
-  mutate(level = factor(level, levels = c("genus", "family", "morphotype")))
-
-ggplot(brm_niche00_plotframe1, aes(value_trans, level, 
-                                 xmin = .lower_trans, xmax = .upper_trans)) +
-  geom_pointinterval() +
-  theme_light(12) +
-  labs(x = "Niche overlap", y= "Level") +
-  guides(fill = FALSE) +
-  scale_x_continuous(breaks = seq(0.25, 0.5, by = 0.05))
-
-ggsave(file = "../output/brm_niche00_plot1.pdf", width = 4.5, height = 4)
-ggsave(file = "../output/brm_niche00_plot1.png", width = 4.5, height = 4)
-```
-
-## 3.6 Genus niche overlap ~ tongue
-### 3.6.1 Model
-```{r}
-brm_niche10 <- brm(niche.overlap ~ 1 +
-                     (1 | site) + # varying intercept for site
-                     (1 | tongue.pair) +
-                     (1 | sp.pair), # varying intercept for sp.pair
-                   family = "zero_one_inflated_beta", # for prop data with 0s and 1s
-                   file = "../output/brm_niche_10", # local
-                   #file = "./brm_niche_10", # HPC
-                   iter = 2000,
-                   chains = 4,
-                   data = filter(niche_pool, level == "genus"))
-
-pp_check(brm_niche10)
-bayes_R2(brm_niche10)
-
-get_variables(brm_niche10)
-mcmc_plot(brm_niche10, pars = c("^r_tongue*"))
-mcmc_plot(brm_niche10, pars = c("^r_site*"))
-```
-
-### 3.6.2 Vizualize
-```{r}
-brm_niche10_plotframe1 <- brm_niche10 %>%
-  spread_draws(b_Intercept, r_tongue.pair[condition, term]) %>%
-  mutate(condition_mean = b_Intercept + r_tongue.pair) %>%
-  group_by(condition) %>% # group by tongue:morpotype pair
-  point_interval(condition_mean, .width = c(0.95, 0.66)) %>% # calculate credible intervals
-  mutate(condition_mean_trans = inv_logit_scaled(condition_mean), # exponentiating the estimates gets us back to the response scale (negbinomial uses log link function)
-         .lower_trans = inv_logit_scaled(.lower),
-         .upper_trans = inv_logit_scaled(.upper)) %>%
-  mutate(condition = factor(condition, levels = c("short:short", "med:med", "long:long",
-                                          "short:med", "short:long", "med:long")))
-
-ggplot(brm_niche10_plotframe1, aes(condition_mean_trans, condition, 
-                                 xmin = .lower_trans, xmax = .upper_trans)) +
-  geom_pointinterval() +
-  theme_light(12) +
-  labs(x = "Niche overlap", y= "Tongue pair") +
-  guides(fill = FALSE)
-
-ggsave(file = "../output/brm_niche10_plot1.pdf", width = 4.5, height = 4)
-ggsave(file = "../output/brm_niche10_plot1.png", width = 4.5, height = 4)
-```
-
-## 3.7 Genus niche overlap ~ sp.pair
-### 3.7.1 Model
-```{r eval=FALSE, include=FALSE}
-brm_niche20 <- brm(niche.overlap ~ 1 +
-                     (1 | site) + # varying intercept for site
-                     (1 | sp.pair), # varying intercept for sp.pair
-                   family = "zero_one_inflated_beta", # for prop data with 0s and 1s
-                   file = "../output/brm_niche_20", # local
-                   #file = "./brm_niche_20", # HPC
-                   iter = 2000,
-                   chains = 4,
-                   data = filter(niche_overlap, level == "genus"))
-
-pp_check(brm_niche20)
-bayes_R2(brm_niche20)
-
-get_variables(brm_niche20)
-conditional_effects(brm_niche20)
-mcmc_plot(brm_niche20, pars = c("^r_sp.pair"))
-mcmc_plot(brm_niche20, pars = c("^r_site*"))
-```
-
-### 3.7.2 Visualize
-```{r eval=FALSE, include=FALSE}
-brm_niche20_plotframe1 <- brm_niche20 %>%
-  spread_draws(b_Intercept, r_sp.pair[condition, term]) %>%
-  mutate(condition_mean = b_Intercept + r_sp.pair) %>%
-  group_by(condition) %>% # group by tongue:morpotype pair
-  point_interval(condition_mean, .width = c(0.95, 0.66)) %>% # calculate credible intervals
-  mutate(condition_mean_trans = inv_logit_scaled(condition_mean), # exponentiating the estimates gets us back to the response scale (negbinomial uses log link function)
-         .lower_trans = inv_logit_scaled(.lower),
-         .upper_trans = inv_logit_scaled(.upper)) %>%
-  mutate(condition = factor(condition)) %>%
-  left_join(niche_pool, by = c("condition" = "sp.pair")) %>%
-  select(condition, condition_mean_trans, .lower_trans, .upper_trans, tongue.pair) %>%
-  distinct() %>%
-  mutate(tongue.pair = factor(tongue.pair, levels = c("short:short", "med:med", "long:long",
-                                                      "short:med", "short:long", "med:long")))
-
-ggplot(brm_niche20_plotframe1, aes(condition_mean_trans, condition, 
-                                 xmin = .lower_trans, xmax = .upper_trans)) +
-  geom_pointinterval() +
-  theme_light(11) +
-  labs(x = "Niche overlap", y= "Tongue pair") +
-  guides(fill = FALSE) +
- # facet_wrap(~tongue.pair, ncol = 1, scales = "free_y") +
-  facet_wrap(~tongue.pair, scales = "free_y")
-
-ggsave(file = "../output/brm_niche20_plot1.pdf", width = 4.5, height = 4)
-ggsave(file = "../output/brm_niche20_plot1.png", width = 7, height = 10)
-```
-
-## 3.8 Niche overlap ~ elevation
-### 3.8.1 Models
-```{r}
-ggplot(filter(niche_overlap_sppair, !sp1 %in% c("humi") & !sp2 %in% c("humi")), 
-       aes(elev.mean, mean.niche.overlap, color = sp.pair)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
-  facet_wrap(~sp.pair) +
-  ylim(c(0,1))
-
-ggplot(filter(niche_overlap_sppair, 
-              sp1 %in% c("pasc", "prat", "soro", "telu", "wurf") & 
-                sp2 %in% c("pasc", "prat", "soro", "telu", "wurf")), 
-       aes(elev.mean, mean.niche.overlap, color = sp.pair)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
-  facet_wrap(~sp.pair) +
-  ylim(c(0,1))
-
-ggplot(filter(niche_overlap_sppair, 
-              sp1 %in% c("pasc", "prat", "soro", "telu", "wurf",
-                         "mont", "muci", "pyre", "mend", "gers") & 
-                sp2 %in% c("pasc", "prat", "soro", "telu", "wurf",
-                           "mont", "muci", "pyre", "mend", "gers")), 
-       aes(elev.mean, mean.niche.overlap, color = sp.pair)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
-  facet_wrap(~sp.pair) +
-  ylim(c(0,1))
-
-ggplot(filter(niche_overlap_sppair,
-              sp1 != "humi" & sp2 != "humi" &
-              sp1 %in% c("pasc", "prat", "soro", "telu", "wurf") | 
-                sp2 %in% c("pasc", "prat", "soro", "telu", "wurf")), 
-       aes(elev.mean, mean.niche.overlap, color = sp.pair)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
-  facet_wrap(~sp.pair) +
-  ylim(c(0,1))
-
-# Pairwise
-gen_gam_niche.I <- bam(mean.niche.overlap ~
-                            s(sp.pair, bs = "re") +
-                            s(site, bs = "re") + # since I have measures repeated over species pairs within each elevation level
-                            s(elev.mean, by = sp.pair, k = 8, m = 2),
-                          method = "fREML",
-                          family = betar(eps = 0.001),
-                          discrete = TRUE,
-                          select = TRUE,
-                          data = filter(niche_overlap_sppair, level == "genus" &
-                                          sp1 %in% c("pasc", "prat", "soro", "telu", "wurf", # dominant species
-                                                        "mont", "muci", "pyre", "mend") & # high elevation specialists
-                                          sp2 %in% c("pasc", "prat", "soro", "telu", "wurf",
-                                                        "mont", "muci", "pyre", "mend"))) %>% getViz()
-
-check.gamViz(gen_gam_niche.I)
-summary(gen_gam_niche.I)
-plot(sm(gen_gam_niche.I, 1))
-plot(sm(gen_gam_niche.I, 11))
-plot(sm(gen_gam_niche.I, 28))
-plot(sm(gen_gam_niche.I, 33))
-plot(sm(gen_gam_niche.I, 37))
-
-# Community mean
-ggplot(niche_pool_sitemean, aes(elev.mean, mean.niche.overlap, color = level)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 8))
-
-```
+plot(pterm(niche_gam_20, 1))
+```

### 2021-03-23 code/trash_can.Rmd
fixed nomenclature issue

@@
 ```
+
+-------------------------------------
+# Old
+
+## 3.6 Genus-level niche overlap ~ tongue.pair + elevation
+### 3.6.1 Model
+```{r}
+niche_gam10 <- gam(mean.niche.overlap ~
+                     tongue.pair +
+                     s(elev.mean, by = tongue.pair, k = 8, m = 2),
+                   method = "REML",
+                   family = betar(eps = 0.001),
+                   select = TRUE,
+                   data = niche_overlap_tongues) %>% getViz()
+
+check.gamViz(niche_gam10)
+summary(niche_gam10)
+print(plot(niche_gam10, allTerms = TRUE), pages = 1)
+```
+
+### 3.6.2 Visualize
+```{r}
+# Get predictions using tidymv
+niche_gam10_pred <- get_gam_predictions(niche_gam10, elev.mean) %>%
+  as_tibble() %>%
+  mutate(mean.niche.overlap_trans = inv_logit_scaled(mean.niche.overlap),
+         CI_upper_trans = inv_logit_scaled(CI_upper),
+         CI_lower_trans = inv_logit_scaled(CI_lower))
+
+# Plot smooths
+ggplot(niche_gam10_pred, aes(elev.mean, mean.niche.overlap_trans, color = tongue.pair)) +
+  geom_ribbon(aes(elev.mean, mean.niche.overlap_trans,
+                  ymax = CI_upper_trans, ymin = CI_lower_trans,
+                  fill = tongue.pair),
+              alpha = 0.3, inherit.aes = FALSE) +
+  geom_line() +
+  ylab("Niche overlap") +
+  xlab("Elevation") +
+  theme_light()
+
+ggsave(file = "../output/niche_gam10_plot1.pdf", width = 5, height = 4)
+ggsave(file = "../output/niche_gam10_plot1.png", width = 5, height = 4)
+```
+
+### 3.5.1 Model
+```{r}
+brm_niche00 <- brm(niche.overlap ~ 1 +
+                     level + # fixed effect of level 
+                     (1 | site) + # varying intercept for site
+                     (1 | sp.pair), # varying intercept for sp.pair
+                   family = "zero_one_inflated_beta", # for prop data with 0s and 1s
+                   file = "../output/brm_niche_00", # local
+                   #file = "./brm_niche_00", # HPC
+                   iter = 2000,
+                   chains = 4,
+                   data = niche_pool)
+
+pp_check(brm_niche00)
+bayes_R2(brm_niche00)
+
+get_variables(brm_niche00)
+conditional_effects(brm_niche00)
+mcmc_plot(brm_niche00, pars = c("^r_sp.pair\\[mend*"))
+mcmc_plot(brm_niche00, pars = c("^r_site*"))
+```
+
+### 3.5.2 Vizualize
+```{r}
+brm_niche00_plotframe1 <- brm_niche00 %>%
+  spread_draws(b_Intercept, b_levelgenus, b_levelmorphotype) %>%
+  mutate(b_levelgenus = b_Intercept + b_levelgenus,
+         b_levelmorphotype = b_Intercept + b_levelmorphotype) %>%
+  pivot_longer(-c(.chain, .iteration, .draw), names_to = "level", values_to = "value") %>%
+  group_by(level) %>%
+  point_interval(value, .width = c(0.95, 0.66)) %>% # calculate credible intervals
+  mutate(value_trans = inv_logit_scaled(value), # exponentiating the estimates gets us back to the response scale (negbinomial uses log link function)
+         .lower_trans = inv_logit_scaled(.lower),
+         .upper_trans = inv_logit_scaled(.upper)) %>%
+  mutate(level = case_when(
+    level == "b_Intercept" ~ "family",
+    level == "b_levelmorphotype" ~ "morphotype",
+    level == "b_levelgenus" ~ "genus"
+  )) %>%
+  mutate(level = factor(level, levels = c("genus", "family", "morphotype")))
+
+ggplot(brm_niche00_plotframe1, aes(value_trans, level, 
+                                 xmin = .lower_trans, xmax = .upper_trans)) +
+  geom_pointinterval() +
+  theme_light(12) +
+  labs(x = "Niche overlap", y= "Level") +
+  guides(fill = FALSE) +
+  scale_x_continuous(breaks = seq(0.25, 0.5, by = 0.05))
+
+ggsave(file = "../output/brm_niche00_plot1.pdf", width = 4.5, height = 4)
+ggsave(file = "../output/brm_niche00_plot1.png", width = 4.5, height = 4)
+```
+
+## 3.6 Genus niche overlap ~ tongue
+### 3.6.1 Model
+```{r}
+brm_niche10 <- brm(niche.overlap ~ 1 +
+                     (1 | site) + # varying intercept for site
+                     (1 | tongue.pair) +
+                     (1 | sp.pair), # varying intercept for sp.pair
+                   family = "zero_one_inflated_beta", # for prop data with 0s and 1s
+                   file = "../output/brm_niche_10", # local
+                   #file = "./brm_niche_10", # HPC
+                   iter = 2000,
+                   chains = 4,
+                   data = filter(niche_pool, level == "genus"))
+
+pp_check(brm_niche10)
+bayes_R2(brm_niche10)
+
+get_variables(brm_niche10)
+mcmc_plot(brm_niche10, pars = c("^r_tongue*"))
+mcmc_plot(brm_niche10, pars = c("^r_site*"))
+```
+
+### 3.6.2 Vizualize
+```{r}
+brm_niche10_plotframe1 <- brm_niche10 %>%
+  spread_draws(b_Intercept, r_tongue.pair[condition, term]) %>%
+  mutate(condition_mean = b_Intercept + r_tongue.pair) %>%
+  group_by(condition) %>% # group by tongue:morpotype pair
+  point_interval(condition_mean, .width = c(0.95, 0.66)) %>% # calculate credible intervals
+  mutate(condition_mean_trans = inv_logit_scaled(condition_mean), # exponentiating the estimates gets us back to the response scale (negbinomial uses log link function)
+         .lower_trans = inv_logit_scaled(.lower),
+         .upper_trans = inv_logit_scaled(.upper)) %>%
+  mutate(condition = factor(condition, levels = c("short:short", "med:med", "long:long",
+                                          "short:med", "short:long", "med:long")))
+
+ggplot(brm_niche10_plotframe1, aes(condition_mean_trans, condition, 
+                                 xmin = .lower_trans, xmax = .upper_trans)) +
+  geom_pointinterval() +
+  theme_light(12) +
+  labs(x = "Niche overlap", y= "Tongue pair") +
+  guides(fill = FALSE)
+
+ggsave(file = "../output/brm_niche10_plot1.pdf", width = 4.5, height = 4)
+ggsave(file = "../output/brm_niche10_plot1.png", width = 4.5, height = 4)
+```
+
+## 3.7 Genus niche overlap ~ sp.pair
+### 3.7.1 Model
+```{r eval=FALSE, include=FALSE}
+brm_niche20 <- brm(niche.overlap ~ 1 +
+                     (1 | site) + # varying intercept for site
+                     (1 | sp.pair), # varying intercept for sp.pair
+                   family = "zero_one_inflated_beta", # for prop data with 0s and 1s
+                   file = "../output/brm_niche_20", # local
+                   #file = "./brm_niche_20", # HPC
+                   iter = 2000,
+                   chains = 4,
+                   data = filter(niche_overlap, level == "genus"))
+
+pp_check(brm_niche20)
+bayes_R2(brm_niche20)
+
+get_variables(brm_niche20)
+conditional_effects(brm_niche20)
+mcmc_plot(brm_niche20, pars = c("^r_sp.pair"))
+mcmc_plot(brm_niche20, pars = c("^r_site*"))
+```
+
+### 3.7.2 Visualize
+```{r eval=FALSE, include=FALSE}
+brm_niche20_plotframe1 <- brm_niche20 %>%
+  spread_draws(b_Intercept, r_sp.pair[condition, term]) %>%
+  mutate(condition_mean = b_Intercept + r_sp.pair) %>%
+  group_by(condition) %>% # group by tongue:morpotype pair
+  point_interval(condition_mean, .width = c(0.95, 0.66)) %>% # calculate credible intervals
+  mutate(condition_mean_trans = inv_logit_scaled(condition_mean), # exponentiating the estimates gets us back to the response scale (negbinomial uses log link function)
+         .lower_trans = inv_logit_scaled(.lower),
+         .upper_trans = inv_logit_scaled(.upper)) %>%
+  mutate(condition = factor(condition)) %>%
+  left_join(niche_pool, by = c("condition" = "sp.pair")) %>%
+  select(condition, condition_mean_trans, .lower_trans, .upper_trans, tongue.pair) %>%
+  distinct() %>%
+  mutate(tongue.pair = factor(tongue.pair, levels = c("short:short", "med:med", "long:long",
+                                                      "short:med", "short:long", "med:long")))
+
+ggplot(brm_niche20_plotframe1, aes(condition_mean_trans, condition, 
+                                 xmin = .lower_trans, xmax = .upper_trans)) +
+  geom_pointinterval() +
+  theme_light(11) +
+  labs(x = "Niche overlap", y= "Tongue pair") +
+  guides(fill = FALSE) +
+ # facet_wrap(~tongue.pair, ncol = 1, scales = "free_y") +
+  facet_wrap(~tongue.pair, scales = "free_y")
+
+ggsave(file = "../output/brm_niche20_plot1.pdf", width = 4.5, height = 4)
+ggsave(file = "../output/brm_niche20_plot1.png", width = 7, height = 10)
+```
+
+## 3.8 Niche overlap ~ elevation
+### 3.8.1 Models
+```{r}
+ggplot(niche_overlap_sppair, aes(elev.mean, mean.niche.overlap, color = sp.pair)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
+  facet_wrap(~sp.pair) +
+  ylim(c(0,1))
+
+ggplot(filter(niche_overlap_sppair, 
+              sp1 %in% c("pasc", "prat", "soro", "telu", "wurf") & 
+                sp2 %in% c("pasc", "prat", "soro", "telu", "wurf")), 
+       aes(elev.mean, mean.niche.overlap, color = sp.pair)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
+  facet_wrap(~sp.pair) +
+  ylim(c(0,1))
+
+ggplot(filter(niche_overlap_sppair, 
+              sp1 %in% c("pasc", "prat", "soro", "telu", "wurf",
+                         "mont", "muci", "pyre", "mend", "gers") & 
+                sp2 %in% c("pasc", "prat", "soro", "telu", "wurf",
+                           "mont", "muci", "pyre", "mend", "gers")), 
+       aes(elev.mean, mean.niche.overlap, color = sp.pair)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
+  facet_wrap(~sp.pair) +
+  ylim(c(0,1))
+
+ggplot(filter(niche_overlap_sppair,
+              sp1 %in% c("pasc", "prat", "soro", "telu", "wurf") | 
+                sp2 %in% c("pasc", "prat", "soro", "telu", "wurf")), 
+       aes(elev.mean, mean.niche.overlap, color = sp.pair)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
+  facet_wrap(~sp.pair) +
+  ylim(c(0,1))
+
+# Pairwise
+gen_gam_niche.I <- bam(mean.niche.overlap ~
+                            s(sp.pair, bs = "re") +
+                            s(site, bs = "re") + # since I have measures repeated over species pairs within each elevation level
+                            s(elev.mean, by = sp.pair, k = 8, m = 2),
+                          method = "fREML",
+                          family = betar(eps = 0.001),
+                          discrete = TRUE,
+                          select = TRUE,
+                          data = filter(niche_overlap_sppair, level == "genus" &
+                                          sp1 %in% c("pasc", "prat", "soro", "telu", "wurf", # dominant species
+                                                        "mont", "muci", "pyre", "mend") & # high elevation specialists
+                                          sp2 %in% c("pasc", "prat", "soro", "telu", "wurf",
+                                                        "mont", "muci", "pyre", "mend"))) %>% getViz()
+
+check.gamViz(gen_gam_niche.I)
+summary(gen_gam_niche.I)
+plot(sm(gen_gam_niche.I, 1))
+plot(sm(gen_gam_niche.I, 11))
+plot(sm(gen_gam_niche.I, 28))
+plot(sm(gen_gam_niche.I, 33))
+plot(sm(gen_gam_niche.I, 37))
+
+# Community mean
+ggplot(niche_pool_sitemean, aes(elev.mean, mean.niche.overlap, color = level)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 8))
+
+```

### 2021-03-23 ms_2/ms_2_v2.Rmd
fixed nomenclature issue

@@
 Species coexistence is constrained in part by competition over shared food sources.
-Bumble bees are classic model organisms for the study of competition and coexistence, and this topic has taken on new urgency as climate-induced range shifts are expected to disrupt competitive relationships between bumble bee species, particularly along the climatic gradients of the mountain slopes that host the world's richest bumble bee communities.
+Bumble bees are classic model organisms for the study of competition and coexistence, and this topic has taken on new urgency as climate-induced range shifts are expected to reshuffle competitive relationships between bumble bee species, particularly along the climatic gradients of the mountain slopes that host the world's richest bumble bee communities.
 We surveyed floral abundance and recorded bumble bee floral visitation patterns along a ~1400 m elevation gradient in the German Alps.
@@
 This indicates that factors other than tongue length are involved in flower choice, which might partly explain the puzzling coexistence of bumble bee species with similar tongue lengths.
-Moreover, genus-level diet overlap was approximately one half that of morphotype-level diet overlap, indicating that even when bumble bees exhibit similar biases at the level of floral morphotype, partly due to the constraints of tongue length, floral diversity enables the mitigation of competition via the subpartitioning of floral taxa within morphotypes.
+Moreover, genus-level diet overlap was approximately one half that of morphotype-level diet overlap, indicating that even when bumble bees exhibit similar biases at the level of floral morphotype, partly due to the constraints of tongue length, floral diversity enables the mitigation of competition via the subpartitioning within floral morphotypes.
 This suggests that floral species richness may be as important as overall floral abundance in avoiding the competitive exclusion of bumble bee species.
@@
 
-<!-- Don't forget: I need to fix the Gentiana/Gentianella/Gentianopsis nomenclature problem and rerun all the models before I finalize the paper.
-This won't change anything about the interpretation of the  analysis, but it needs to be done.
-It might make Gentiana show up in the PCoA.
--->
-
 # INTRODUCTION
@@
 In this role, however, they have proven somewhat enigmatic.
-On the one hand, there is evidence that resource limitation can be a significant constraint on local bumble communities [@Heinrich1976-wc; @Bowers1985-df], and certain bumble bee communities have been shown to be over-dispersed in terms of tongue length (a matching trait corresponding to floral corolla depth) or phylogeny (understood to encompass unmeasured trait variation), suggesting that competition functions as a community assembly process [@Inouye1977-kl; @Pellissier2013-pk].
-Nevertheless, bumble bee communities with an apparent excess of species richness and deficit of trait (or phylogenetic) dispersion are commonly observed [@Ranta1981-vx; @Goulson2008-zy].
+On the one hand, there is evidence that resource limitation can be a significant constraint on local bumble communities [@Heinrich1976-wc; @Bowers1985-df], and certain bumble bee communities have been shown to be over-dispersed in terms of tongue length or phylogeny, suggesting that competition functions as a community assembly process [@Inouye1977-kl; @Pellissier2013-pk].
+Nevertheless, bumble bee communities with an apparent deficit of trait (or phylogenetic) dispersion are commonly observed [@Ranta1981-vx; @Goulson2008-zy].
 To account for this inconsistency, @Ranta1981-vx proposed that bumble bees communities are governed by a spatiotemporally heterogeneous mosaic of competitive conditions, such that competitive inequalities between species are sufficiently variable to allow more species to coexist than predicted by trait-based competition models.
 A complementary explanation, based on the core/satellite species hypothesis [@Hanski1982-hs], is that competitive forces are significant only among regionally abundant and widespread “core species”, and the effects of competition are more pronounced in the composition of regional species pools than in that of local assemblages [@Hanski1982-ie].
-In some cases, competition for food may simply be precluded by a saturating abundance of floral resources [@Williams1989-vr], implying that factors other than food availability --- such as abiotic stressors or predator/pathogen pressure --- limit populations.
+In some cases, competition for food may simply be precluded by a saturating abundance of floral resources [@Williams1989-vr], implying that populations are limited by other factors, such as abiotic stressors, nest site availability, or predator/pathogen pressure.
 
@@
 These costs are closely related to the degree of morphological fittedness between bumble bees and their floral hosts, which consists principally in the matching of bumble bee tongue length to floral corolla depth.
-Bumble bee species can differ markedly in tongue length, and floral handling time increases when tongue length and corolla depth as poorly matched [@Inouye1980-mv; @Harder1983-bk], which leads to an association between tongue length and floral selection in bumble bee communities [e.g.
+Bumble bee species can differ markedly in tongue length, and floral handling time increases when tongue length and corolla depth are poorly matched [@Inouye1980-mv; @Harder1983-bk], which leads to an association between tongue length and floral selection in bumble bee communities [e.g.
 @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn].
@@
 Focusing on *potential* competition does not, however, escape the equivocity of niche overlap, since an observed pattern of niche overlap could just as plausibly be interpreted as adaptive niche partitioning in response to *realized* competition [@Pianka1974-qk].
-This creates a predicament in which niche overlap can only be interpreted as an index of potential competition when realized competition --- which cannot be inferred from niche overlap --- is assumed to be negligible, and this assumption can only be justified on the basis of additional knowledge of the study system under consideration.
+This creates a predicament in which niche overlap can only be interpreted as an index of potential competition when realized competition is assumed to be negligible, an assumption that can only be justified on the basis of additional knowledge of the study system under consideration.
 
@@
 While floral visitation by male bumble bees was recorded during visitation sampling, we chose to analyze only visitation by queens and workers, which we pooled.
-During observation, it was noted whether each bee was visibly carrying pollen at the time of its visit, but this provides at best an imperfect distinction between pollen and nectar foragers, and we opted not to differentiate between bees with and without pollen in our analyses.
 In conjunction with visitation observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.1 m^2^.
@@
 
-We recorded a total of 12,918 bumble- bee-flower interactions (excluding males) over the three years of our study.
-The metaweb across all sites and dates consisted of 16 bumble bee species (with *Bombus terrestric/lucorum* and *Bombus psithyrus* species groups lumped into species groups), 163 plant species (110 genera, 37 families), and 736 unique bumble- bee-plant interaction pairs.
+We recorded a total of 12,918 bumble- bee-flower interactions over the three years of our study.
+The metaweb across all sites and dates consisted of 16 bumble bee species (with *Bombus terrestric/lucorum* and *Bombus psithyrus* species lumped into species groups), 163 plant species (110 genera, 37 families), and 736 unique bumble- bee-plant interaction pairs.
 Five species --- *B.
@@
 *B.
-humilis* was recorded only three times and only in one year, so we omitted it from all analyses.
+hypnorum* and *B.
+humilis* were recorded only three times and only in one year, so we omitted it from all analyses.
 
@@
 ```{r fig8, echo = FALSE,  fig.cap = paste("Niche overlap by level through elevation.
-Points show empirical niche overlap values, lines show GAM smooths, and shaded bands show the standard errors of the mean")}
-knitr::include_graphics("../output/niche_gam00_plot1.png")
+Points show empirical niche overlap values, lines show GAM smooths, and shaded bands show 95\% confidence intervals.")}
+knitr::include_graphics("../output/niche_gam_00_plot1.png")
 ```
 
-```{r fig9, echo = FALSE,  fig.cap = paste("Niche overlap by tongue length class difference through elevation. Lines show GAM smooths and shaded bands show standard errors of the mean")}
-knitr::include_graphics("../output/niche_gam20_plot1.png")
+```{r fig9, echo = FALSE,  fig.cap = paste("Niche overlap by tongue length class difference through elevation. Lines show GAM smooths and shaded bands show 95\% confidence intervals.")}
+knitr::include_graphics("../output/niche_gam_20_plot1.png")
 ```
@@
 
-We thank D.J. McNeil for helpful conversations during data analysis.
+We thank D.J. McNeil for helpful conversations during data analysis. O. Mitesser kindly provided a high-performance computing environment for expediting Bayesian model fitting.
 

### 2021-03-23 code/data_proc.R

@@ -3,7 +3,7 @@ library(lubridate)
 
 # Get the data cleaned up and properly organized
 ### Floral survey data
-fl_survey <- read_delim("./raw_data/data_plantcover_forDoug_2020.01.17.csv", delim = "\t", 
+fl_survey <- read_delim("./data/raw_data/data_plantcover_forDoug_2020.01.17.csv", delim = "\t", 
                            locale = locale(decimal_mark = ",", grouping_mark = ".")) %>%
   na.omit() %>%
   separate(col = species_plant, into = c("genus", "species", "w", "x", "y", "z"), sep = " ") %>%
@@ -14,7 +14,9 @@ fl_survey <- read_delim("./raw_data/data_plantcover_forDoug_2020.01.17.csv", del
                                            "Arctostaphylos alpina" = "Arctous alpina",
                                            "Calluna vulagris" = "Calluna vulgaris",
                                            "Deschampsia caespitosa" = "Deschampsia cespitosa",
-                                           "Gymnadenia bifolia" = "Platanthera bifolia"))) %>%
+                                           "Gymnadenia bifolia" = "Platanthera bifolia",
+                                           "Gentiana aspera" = "Gentianella aspera",
+                                           "Gentiana ciliata" = "Gentianella ciliata"))) %>%
   separate(species_plant, c("genus", "species"), remove = FALSE) %>%
   dplyr::select(year, dayofyear, site = site_name, snowcover, 
                 plant.sp = species_plant, plant.genus = genus, flower.cover = cover_all_flowers) %>%
@@ -25,7 +27,13 @@ fl_survey <- read_delim("./raw_data/data_plantcover_forDoug_2020.01.17.csv", del
   filter(plant.genus != "Huperzia") %>% # because I don't need the clubmoss
   dplyr::select(-temp.date)
   
-write_csv(fl_survey, "./processed_data/floral_survey.csv")
+write_csv(fl_survey, "./data/processed_data/floral_survey.csv")
+
+# 
+gentian_problem <- fl_survey %>%
+  select(plant.genus, plant.sp) %>%
+  filter(plant.genus %in% c("Gentiana", "Gentianella", "Gentianopsis")) %>%
+  distinct()
 
 ### Visitation data
[…]

### 2021-03-25 code/ms2_bias_niche.Rmd
fixed nomenclature again

@@
 
-## Flower colors (accessed through BioFlor)
-fl_color <- read_csv("../data/processed_data/floral_color.csv")
-
-## diameter and corolla depth data from Gita Benadi
-fl_morph <- read_csv("../data/processed_data/flower_morphology_GitaBenadi.csv") %>% 
-  dplyr::select(SampleID, plant.family = Family, plant.sp = Species, 
-                Unit, Flowers_per_unit, diam = Diameter, LN) %>%
-  group_by(plant.sp) %>%
-  summarize(diam = mean(diam), LN = mean(LN))
-
 ## verbal description of Kugler morphotypes
@@
   full_join(fl_ktype) %>%
-  full_join(fl_color) %>%
-  full_join(fl_morph) %>%
-  dplyr::select(plant.sp, plant.genus, plant.family, k.type, k.type.s, k.type.ss, color, diam, LN)
+  dplyr::select(plant.sp, plant.genus, plant.family, k.type, k.type.s, k.type.ss) 
 ```
@@
   filter(caste != "male") %>% # we will ignore males 
-  filter(!bb.sp %in% c("humi", "hypn")) %>% # drop rare species
+  filter(!bb.sp %in% c("humi")) %>% # humi too rare
   group_by(site, date, bb.sp, plant.sp) %>%
@@
 All these run cleanly.
+Well, brm_pvis_01 throws maximum tree depth warnings, but these are just about efficiency, not validity.
+These warnings are evidently a consequence of increasing adapt_delta to 0.99, which succeeded in getting rid of the divergent transition warnings (those warnings are much more serious and pertain to the validity of the model).
 ```{r}
@@
                   chains = 4,
-                  cores = 12,
-                  iter = 2000,
+                  #cores = 12,
+                  iter = 5000,
                   data = filter(prop_visits_ktype, !ktype %in% c("0", "9", "10")))
@@
                   prior = brm_pvis_prior, # weakly regularizing
+                  control = list(adapt_delta = 0.99), # the default setting yielded divergent transitions
                   file = "../output/brm_pvis_ktype_01",  # for running locally
@@
                   chains = 4,
-                  cores = 12,
-                  iter = 2000,
+                  #cores = 12,
+                  iter = 5000,
                   data = filter(prop_visits_ktype_pooled, !ktype %in% c("0", "9", "10")))
@@
                   chains = 4,
-                  cores = 12,
+                  #cores = 12,
                   iter = 5000, # Bulk ESS was a bit low wit 2000 iter
@@
                                           "Bell (3)", "Stalk-disc (4)", "Lip (5)", 
-                                          "Flag (6)", "Head (7)")))
-
-ggplot(brm_pvis_00_plotframe1, aes(condition_mean_trans, tongue, 
+                                          "Flag (6)", "Head (7)"))) %>%
+  mutate(tongue = factor(tongue, levels = c("long", "med", "short")))
+
+ggplot(brm_pvis_00_plotframe1, aes(condition_mean_trans, rev(tongue), 
                                  xmin = .lower_trans, xmax = .upper_trans,
@@
 ggsave(file = "../output/brm_pvis_01_plot1.pdf", width = 6, height = 7)
-
+ggsave(file = "../output/brm_pvis_01_plot1.png", width = 6, height = 7)
 ```
@@
 ```{r}
-get_variables(brm_pvis01)
-
-hypothesis(brm_pvis01, "r_ktype[1,Intercept] = r_ktype[4,Intercept]", class = "") 
-hypothesis(brm_pvis01, "r_ktype[2,Intercept] = r_ktype[4,Intercept]", class = "")
-
-hypothesis(brm_pvis01, "r_ktype[2,Intercept] = r_ktype[1,Intercept]", class = "")
-hypothesis(brm_pvis01, "r_ktype[1,Intercept] = r_ktype[3,Intercept]", class = "") 
-hypothesis(brm_pvis01, "r_ktype[2,Intercept] = r_ktype[3,Intercept]", class = "")
-hypothesis(brm_pvis01, "r_ktype[4,Intercept] = r_ktype[3,Intercept]", class = "")
-
-hypothesis(brm_pvis01, "r_ktype[5,Intercept] = r_ktype[6,Intercept]", class = "") 
-hypothesis(brm_pvis01, "r_ktype[5,Intercept] = r_ktype[7,Intercept]", class = "") 
+get_variables(brm_pvis_01)
+
+hypothesis(brm_pvis_01, "r_ktype[1,Intercept] = r_ktype[4,Intercept]", class = "") 
+hypothesis(brm_pvis_01, "r_ktype[2,Intercept] = r_ktype[4,Intercept]", class = "")
+
+hypothesis(brm_pvis_01, "r_ktype[2,Intercept] = r_ktype[1,Intercept]", class = "")
+hypothesis(brm_pvis_01, "r_ktype[1,Intercept] = r_ktype[3,Intercept]", class = "") 
+hypothesis(brm_pvis_01, "r_ktype[2,Intercept] = r_ktype[3,Intercept]", class = "")
+hypothesis(brm_pvis_01, "r_ktype[4,Intercept] = r_ktype[3,Intercept]", class = "")
+
+hypothesis(brm_pvis_01, "r_ktype[5,Intercept] = r_ktype[6,Intercept]", class = "") 
+hypothesis(brm_pvis_01, "r_ktype[5,Intercept] = r_ktype[7,Intercept]", class = "") 
 ```
@@
 ```{r}
-get_variables(brm_pvis00)
+get_variables(brm_pvis_00)
 
 # Disc (group  1)
-hypothesis(brm_pvis00, "r_ktype:tongue[1_short,Intercept] = r_ktype:tongue[1_med,Intercept]", class = "")
-hypothesis(brm_pvis00, "r_ktype:tongue[1_short,Intercept] = r_ktype:tongue[1_long,Intercept]", class = "")
-hypothesis(brm_pvis00, "r_ktype:tongue[1_med,Intercept] = r_ktype:tongue[1_long,Intercept]", class = "")
+hypothesis(brm_pvis_00, "r_ktype:tongue[1_short,Intercept] = r_ktype:tongue[1_med,Intercept]", class = "")
+hypothesis(brm_pvis_00, "r_ktype:tongue[1_short,Intercept] = r_ktype:tongue[1_long,Intercept]", class = "")
+hypothesis(brm_pvis_00, "r_ktype:tongue[1_med,Intercept] = r_ktype:tongue[1_long,Intercept]", class = "")
 
 # Funnel (group  2)
-hypothesis(brm_pvis00, "r_ktype:tongue[2_short,Intercept] = r_ktype:tongue[2_med,Intercept]", class = "")
-hypothesis(brm_pvis00, "r_ktype:tongue[2_short,Intercept] = r_ktype:tongue[2_long,Intercept]", class = "")
-hypothesis(brm_pvis00, "r_ktype:tongue[2_med,Intercept] = r_ktype:tongue[2_long,Intercept]", class = "")
+hypothesis(brm_pvis_00, "r_ktype:tongue[2_short,Intercept] = r_ktype:tongue[2_med,Intercept]", class = "")
+hypothesis(brm_pvis_00, "r_ktype:tongue[2_short,Intercept] = r_ktype:tongue[2_long,Intercept]", class = "")
+hypothesis(brm_pvis_00, "r_ktype:tongue[2_med,Intercept] = r_ktype:tongue[2_long,Intercept]", class = "")
 
 # Bell (group  3)
-hypothesis(brm_pvis00, "r_ktype:tongue[3_short,Intercept] = r_ktype:tongue[3_med,Intercept]", class = "")
-hypothesis(brm_pvis00, "r_ktype:tongue[3_short,Intercept] = r_ktype:tongue[3_long,Intercept]", class = "")
-hypothesis(brm_pvis00, "r_ktype:tongue[3_med,Intercept] = r_ktype:tongue[3_long,Intercept]", class = "")
+hypothesis(brm_pvis_00, "r_ktype:tongue[3_short,Intercept] = r_ktype:tongue[3_med,Intercept]", class = "")
+hypothesis(brm_pvis_00, "r_ktype:tongue[3_short,Intercept] = r_ktype:tongue[3_long,Intercept]", class = "")
+hypothesis(brm_pvis_00, "r_ktype:tongue[3_med,Intercept] = r_ktype:tongue[3_long,Intercept]", class = "")
 
 # Stalk-disc (group  4)
-hypothesis(brm_pvis00, "r_ktype:tongue[4_short,Intercept] = r_ktype:tongue[4_med,Intercept]", class = "")
-hypothesis(brm_pvis00, "r_ktype:tongue[4_short,Intercept] = r_ktype:tongue[4_long,Intercept]", class = "")
-hypothesis(brm_pvis00, "r_ktype:tongue[4_med,Intercept] = r_ktype:tongue[4_long,Intercept]", class = "")
+hypothesis(brm_pvis_00, "r_ktype:tongue[4_short,Intercept] = r_ktype:tongue[4_med,Intercept]", class = "")
+hypothesis(brm_pvis_00, "r_ktype:tongue[4_short,Intercept] = r_ktype:tongue[4_long,Intercept]", class = "")
+hypothesis(brm_pvis_00, "r_ktype:tongue[4_med,Intercept] = r_ktype:tongue[4_long,Intercept]", class = "")
 
 # Lip (group  5)
-hypothesis(brm_pvis00, "r_ktype:tongue[5_short,Intercept] = r_ktype:tongue[5_med,Intercept]", class = "")
-hypothesis(brm_pvis00, "r_ktype:tongue[5_short,Intercept] = r_ktype:tongue[5_long,Intercept]", class = "")
-hypothesis(brm_pvis00, "r_ktype:tongue[5_med,Intercept] = r_ktype:tongue[5_long,Intercept]", class = "")
+hypothesis(brm_pvis_00, "r_ktype:tongue[5_short,Intercept] = r_ktype:tongue[5_med,Intercept]", class = "")
+hypothesis(brm_pvis_00, "r_ktype:tongue[5_short,Intercept] = r_ktype:tongue[5_long,Intercept]", class = "")
+hypothesis(brm_pvis_00, "r_ktype:tongue[5_med,Intercept] = r_ktype:tongue[5_long,Intercept]", class = "")
 
 # Flag (group  6)
-hypothesis(brm_pvis00, "r_ktype:tongue[6_short,Intercept] = r_ktype:tongue[6_med,Intercept]", class = "")
-hypothesis(brm_pvis00, "r_ktype:tongue[6_short,Intercept] = r_ktype:tongue[6_long,Intercept]", class = "")
-hypothesis(brm_pvis00, "r_ktype:tongue[6_med,Intercept] = r_ktype:tongue[6_long,Intercept]", class = "")
+hypothesis(brm_pvis_00, "r_ktype:tongue[6_short,Intercept] = r_ktype:tongue[6_med,Intercept]", class = "")
+hypothesis(brm_pvis_00, "r_ktype:tongue[6_short,Intercept] = r_ktype:tongue[6_long,Intercept]", class = "")
+hypothesis(brm_pvis_00, "r_ktype:tongue[6_med,Intercept] = r_ktype:tongue[6_long,Intercept]", class = "")
 
 # Head (group  7)
-hypothesis(brm_pvis00, "r_ktype:tongue[7_short,Intercept] = r_ktype:tongue[7_med,Intercept]", class = "")
-hypothesis(brm_pvis00, "r_ktype:tongue[7_short,Intercept] = r_ktype:tongue[7_long,Intercept]", class = "")
-hypothesis(brm_pvis00, "r_ktype:tongue[7_med,Intercept] = r_ktype:tongue[7_long,Intercept]", class = "")
+hypothesis(brm_pvis_00, "r_ktype:tongue[7_short,Intercept] = r_ktype:tongue[7_med,Intercept]", class = "")
+hypothesis(brm_pvis_00, "r_ktype:tongue[7_short,Intercept] = r_ktype:tongue[7_long,Intercept]", class = "")
+hypothesis(brm_pvis_00, "r_ktype:tongue[7_med,Intercept] = r_ktype:tongue[7_long,Intercept]", class = "")
 ```
@@
   mutate(k.type.ss = factor(k.type.ss)) %>%
-  distinct() %>%
-  na.omit() # THIS IS JUST A PLACEHOLDER UNTIL I FIX THE GENTIANA/ELLA/OPSIS NOMENCALTURE
+  distinct()
 
@@
   filter(caste != "male") %>% # we will ignore males 
-  filter(!bb.sp %in% c("hypn", "humi")) %>% # hypn and humi are too rare
+  filter(!bb.sp %in% c("humi")) %>% # humi is too rare
   #group_by(bb.sp, plant.sp.abb) %>%
@@
 bb_traits2 <- bb_traits %>%
-  filter(!bb.sp %in% c("humi", "hypn")) %>% # drop rare species
+  filter(!bb.sp %in% c("humi")) %>% # drop humi
   select(bb.sp, pbl.w.class2) %>%
@@
                alpha = 0.25) +
-  geom_text(aes(label = bb.sp), key_glyph = "rect") +
-  geom_text_repel(data = filter(pcoa_horn2.fort1, abs(MDS1) >= 0.2 | abs(MDS2) >= 0.2), 
-            aes(MDS1, MDS2, label = plant.genus), inherit.aes = FALSE, color = "gray40") +
-  geom_point(data = filter(pcoa_horn2.fort1, abs(MDS1) >= 0.2 | abs(MDS2) >= 0.2),
-            aes(MDS1, MDS2, shape = k.type.ss), inherit.aes = FALSE, color = "gray40", size = 3) +
+  geom_point(data = filter(pcoa_horn2.fort1, abs(MDS1) >= 0.20 | abs(MDS2) >= 0.20),
+            aes(MDS1, MDS2, shape = k.type.ss), inherit.aes = FALSE, color = "gray40", size = 2) +
+  geom_text(aes(label = bb.sp), key_glyph = "rect", size = 3) +
+  geom_text_repel(data = filter(pcoa_horn2.fort1, abs(MDS1) >= 0.20 | abs(MDS2) >= 0.20), 
+            aes(MDS1, MDS2, label = plant.genus), inherit.aes = FALSE, color = "gray40",
+            size = 3) +
   scale_shape_discrete() +
@@
     # dplyr::select(site, year, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
-    dplyr::select(site, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
+    dplyr::select(site, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean, management) %>% # dplyr::select and rename vars
     left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
@@
     rename(subgenus2 = subgenus, tongue2 = pbl.w.class2) %>%
-    # dplyr::select(site, year, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
-    #        tongue1, tongue2, elev.mean) %>%
     dplyr::select(site, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
-       tongue1, tongue2, elev.mean) %>%
+       tongue1, tongue2, elev.mean, management) %>%
     mutate(tongue.diff = factor(if_else(tongue1 == tongue2, "same", "different")),
@@
 # This one calculates niche overlap on fully resolved visitation data and then averages niche overlap across dates
-niche_overlapper2 <- function(x, criterion) {
-
-  criterion <- enquo(criterion) # criterion is set to either plant.sp (for taxonomic) or k.type.s (for morphological)
-
-  x %>%
-    #group_by(bb.sp, !!criterion, site, date, year) %>% # group by site-year and BB-FL pair
-    group_by(bb.sp, !!criterion, site, date) %>% # group by site-year and BB-FL pair
-    summarize(freq = sum(visits)) %>% # get interaction frequency for each BB-FL pair per site-year
-    group_by(bb.sp, site, date) %>% # group by site-year and BB species
-    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
-    #group_by(bb.sp, site, year, !!criterion) %>%
-    # group_by(bb.sp, site, !!criterion) %>%
-    # summarize(mean.prop = mean(prop)) %>%
-    unite(sitedate, c(site, date), sep = "_") %>%
-    #pivot_wider(names_from = !!criterion, values_from = mean.prop, values_fill = 0) %>% # spread into the kind of array vegan likes
-    #pivot_wider(names_from = !!criterion, values_from = prop, values_fill = 0) %>% # spread into the kind of array vegan likes
-    #group_by(site, year) %>%
-    #group_by(site) %>%
-    select(-freq) %>%
-    group_by(sitedate) %>%
-    nest() %>% # nest into list-column
-    mutate(data2 = map(data, function(x) pivot_wider(x, names_from = !!criterion, values_from = prop, values_fill = 0)),
-           data3 = map(data2, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
-           dist = map(data3, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
-           tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
-    #dplyr::select(site, year, tidy_dist) %>% # drop intermediate list-columns
-    #dplyr::select(site, tidy_dist) %>% # drop intermediate list-columns
-    dplyr::select(sitedate, tidy_dist) %>% # drop intermediate list-columns
-    unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
-    separate(sitedate, c("site", "date"), sep = "_") %>%
-    left_join(site_data) %>% # add site data
-    # dplyr::select(site, year, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
-    dplyr::select(site, date, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
-    left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
-    rename(subgenus1 = subgenus, tongue1 = pbl.w.class2) %>%
-    left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
-    rename(subgenus2 = subgenus, tongue2 = pbl.w.class2) %>%
-    # dplyr::select(site, year, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
+# niche_overlapper2 <- function(x, criterion) {
+#     
+#   criterion <- enquo(criterion) # criterion is set to either plant.sp (for taxonomic) or k.type.s (for morphological)
+#   
+#   x %>%
+#     #group_by(bb.sp, !!criterion, site, date, year) %>% # group by site-year and BB-FL pair
+#     group_by(bb.sp, !!criterion, site, date) %>% # group by site-year and BB-FL pair
+#     summarize(freq = sum(visits)) %>% # get interaction frequency for each BB-FL pair per site-year
+#     group_by(bb.sp, site, date) %>% # group by site-year and BB species
+#     mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
+#     #group_by(bb.sp, site, year, !!criterion) %>%
+#     # group_by(bb.sp, site, !!criterion) %>%
+#     # summarize(mean.prop = mean(prop)) %>%
+#     unite(sitedate, c(site, date), sep = "_") %>%
+#     #pivot_wider(names_from = !!criterion, values_from = mean.prop, values_fill = 0) %>% # spread into the kind of array vegan likes
+#     #pivot_wider(names_from = !!criterion, values_from = prop, values_fill = 0) %>% # spread into the kind of array vegan likes
+#     #group_by(site, year) %>%
+#     #group_by(site) %>%
+#     select(-freq) %>%
+#     group_by(sitedate) %>%
+#     nest() %>% # nest into list-column
+#     mutate(data2 = map(data, function(x) pivot_wider(x, names_from = !!criterion, values_from = prop, values_fill = 0)),
+#            data3 = map(data2, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
+#            dist = map(data3, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
+#            tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
+#     #dplyr::select(site, year, tidy_dist) %>% # drop intermediate list-columns
+#     #dplyr::select(site, tidy_dist) %>% # drop intermediate list-columns
+#     dplyr::select(sitedate, tidy_dist) %>% # drop intermediate list-columns
+#     unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
+#     separate(sitedate, c("site", "date"), sep = "_") %>%
+#     left_join(site_data) %>% # add site data
+#     # dplyr::select(site, year, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
+#     dplyr::select(site, date, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
+#     left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
+#     rename(subgenus1 = subgenus, tongue1 = pbl.w.class2) %>%
+#     left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
+#     rename(subgenus2 = subgenus, tongue2 = pbl.w.class2) %>%
+#     # dplyr::select(site, year, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
+#     #        tongue1, tongue2, elev.mean) %>%
+#     dplyr::select(site, date, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
 #        tongue1, tongue2, elev.mean) %>%
-    dplyr::select(site, date, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
-       tongue1, tongue2, elev.mean) %>%
-    mutate(tongue.diff = factor(if_else(tongue1 == tongue2, "same", "different")),
-           tongue.pair = factor(case_when(
-             tongue1 == "med" & tongue2 == "med" ~ "med:med",
-             tongue1 == "med" & tongue2 == "long" ~ "med:long",
-             tongue1 == "med" & tongue2 == "short" ~ "short:med",
-             tongue1 == "long" & tongue2 == "med" ~ "med:long",
-             tongue1 == "short" & tongue2 == "med" ~ "short:med",
-             tongue1 == "long" & tongue2 == "short" ~ "short:long",
-             tongue1 == "short" & tongue2 == "long" ~ "short:long",
-             tongue1 == "long" & tongue2 == "long" ~ "long:long",
-             tongue1 == "short" & tongue2 == "short" ~ "short:short",
-           )),
-           subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different")) %>%
-    unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":", remove = FALSE) %>% # create a single species pair column...
-    mutate(sp.pair = factor(sp.pair)) %>% # ...and make it a factor so that mgcv can handle it
-    mutate(niche.overlap = 1-distance) %>% # express niche overlap so that 1 = perfect overlap and 0 = no overlap
-    select(-distance)
-}
+#     mutate(tongue.diff = factor(if_else(tongue1 == tongue2, "same", "different")),
+#            tongue.pair = factor(case_when(
+#              tongue1 == "med" & tongue2 == "med" ~ "med:med",
+#              tongue1 == "med" & tongue2 == "long" ~ "med:long",
+#              tongue1 == "med" & tongue2 == "short" ~ "short:med",
+#              tongue1 == "long" & tongue2 == "med" ~ "med:long",
+#              tongue1 == "short" & tongue2 == "med" ~ "short:med",
+#              tongue1 == "long" & tongue2 == "short" ~ "short:long",
+#              tongue1 == "short" & tongue2 == "long" ~ "short:long",
+#              tongue1 == "long" & tongue2 == "long" ~ "long:long",
+#              tongue1 == "short" & tongue2 == "short" ~ "short:short",
+#            )),
+#            subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different")) %>%
+#     unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":", remove = FALSE) %>% # create a single species pair column...
+#     mutate(sp.pair = factor(sp.pair)) %>% # ...and make it a factor so that mgcv can handle it
+#     mutate(niche.overlap = 1-distance) %>% # express niche overlap so that 1 = perfect overlap and 0 = no overlap
+#     select(-distance)
+# }
 ```
@@
   mutate(level = factor(level, levels = c("family", "morphotype", "genus")),
-         site = factor(site)) %>%
+         site = factor(site),
+         management = factor(management, levels = c("none", "grasing", "mowing"))) %>%
   separate(sp.pair, c("sp1", "sp2"), sep = ":", remove = FALSE)
-
-ggplot(niche_overlap, aes(niche.overlap)) +
-  geom_histogram()
-
-ggplot(filter(niche_overlap,
-              sp1 %in% c("pasc", "prat", "soro", "telu", "wurf") &
-                sp2 %in% c("pasc", "prat", "soro", "telu", "wurf")), aes(niche.overlap)) +
-  geom_histogram() +
-  facet_wrap(~level)
 
@@
 niche_overlap_levels <- niche_overlap %>%
-  group_by(site, level, elev.mean) %>%
+  group_by(site, level, elev.mean, management) %>%
   summarize(mean.niche.overlap = mean(niche.overlap)) %>%
@@
 niche_overlap_sppair <-  niche_overlap %>%
-  group_by(site, sp.pair, level, elev.mean) %>%
+  group_by(site, sp.pair, level, elev.mean, management) %>%
   summarize(mean.niche.overlap = mean(niche.overlap)) %>%
@@
 niche_overlap_tongues <- niche_overlap %>%
-  group_by(site, level, tongue.pair, elev.mean) %>%
+  group_by(site, level, tongue.pair, elev.mean, management) %>%
   summarize(mean.niche.overlap = mean(niche.overlap)) %>%
@@
 niche_overlap_tonguediff <- niche_overlap %>%
-  group_by(site, level, tongue.diff, elev.mean) %>%
+  group_by(site, level, tongue.diff, elev.mean, management) %>%
   summarize(mean.niche.overlap = mean(niche.overlap)) %>%
   filter(level == "genus")
+
+#################
+
+ggplot(niche_overlap_sppair, aes(management, mean.niche.overlap)) +
+  geom_boxplot()
 ```
@@
                       level + 
+                      #management +
                       s(elev.mean, by = level, k = 8, m = 2),
@@
 print(plot(niche_gam_20, allTerms = TRUE), pages = 1)
+
+niche_gam_21 <- gam(mean.niche.overlap ~
+                     tongue.pair +
+                     s(elev.mean, by = tongue.pair, k = 8, m = 2),
+                   method = "REML",
+                   family = betar(eps = 0.001),
+                   select = TRUE,
+                   data = niche_overlap_tongues) %>% getViz()
+
+check.gamViz(niche_gam_21)
+summary(niche_gam_21)
+print(plot(niche_gam_21, allTerms = TRUE), pages = 1)
 ```
@@
 ```{r}
+# Tongue diff
 # Get predictions
@@
 plot(pterm(niche_gam_20, 1))
-```
+
+# Tongue pair
+# Get predictions
+niche_gam_21_pred <- get_gam_predictions(niche_gam_21, elev.mean) %>%
+  as_tibble() %>%
+  mutate(mean.niche.overlap_trans = inv_logit_scaled(mean.niche.overlap),
+         CI_upper_trans = inv_logit_scaled(CI_upper),
+         CI_lower_trans = inv_logit_scaled(CI_lower))
+
+# Plot smooths
+ggplot(niche_gam_21_pred, aes(elev.mean, mean.niche.overlap_trans, color = tongue.pair)) +
+  geom_ribbon(aes(elev.mean, mean.niche.overlap_trans, 
+                  ymax = CI_upper_trans, ymin = CI_lower_trans,
+                  fill = tongue.pair),
+              alpha = 0.3, inherit.aes = FALSE) +
+  geom_line() +
+  ylab("Niche overlap") +
+  xlab("Elevation (m)") +
+  guides(fill = FALSE) +
+  labs(color = "Tongue pair") +
+  theme_light()
+
+ggsave(file = "../output/niche_gam_21_plot1.pdf", width = 5, height = 4)
+ggsave(file = "../output/niche_gam_21_plot1.png", width = 5, height = 4)
+
+# Plot intercept effect of tongue.diff
+plot(pterm(niche_gam_20, 1))
+```

### 2021-03-25 ms_2/ms_2_v2.Rmd
fixed nomenclature again

@@
 
-The role of competition in regulating community assembly and species coexistence has long been a central question in ecology.
-It has taken on new urgency, however, on account of recent climate change, which threatens to disrupt existing patterns of species interactions [@Tylianakis2008-fi], including competition [@Ettinger2017-pk], with consequences for community composition and function.
-These effects are likely to be most pronounced along the steep climate gradients of mountain slopes, where rapid climate-induced range shifts are expected to introduce novel competitive interactions within trophic levels and cause spatial or phenological mismatches between organisms and their food resources.
+The role of competition and niche partitioning in regulating community assembly has long been a central question in ecology, but it has taken on new urgency on account of anthropogenic climate and land use change, which threaten to disrupt existing patterns of species interactions and niche relationships [@Tylianakis2008-fi].
+These effects are likely to be especially pronounced in mountain ecosystems, where steep elevational temperature gradients facilitate species range shifts in response to climate change, and socioeconomic factors are driving rapid changes in historical land use patterns [@Becker2007-cg; @Loffler2011-xo; @Klein2019-dq]
 
@@
 In this role, however, they have proven somewhat enigmatic.
-On the one hand, there is evidence that resource limitation can be a significant constraint on local bumble communities [@Heinrich1976-wc; @Bowers1985-df], and certain bumble bee communities have been shown to be over-dispersed in terms of tongue length or phylogeny, suggesting that competition functions as a community assembly process [@Inouye1977-kl; @Pellissier2013-pk].
+On the one hand, there is evidence that resource limitation can be a significant constraint on local bumble communities [@Heinrich1976-wc; @Bowers1985-df], and certain bumble bee communities have been shown to be overdispersed in terms of tongue length or phylogeny, suggesting that competition functions as a community assembly process [@Inouye1977-kl; @Pellissier2013-pk].
 Nevertheless, bumble bee communities with an apparent deficit of trait (or phylogenetic) dispersion are commonly observed [@Ranta1981-vx; @Goulson2008-zy].
@@
 A complementary explanation, based on the core/satellite species hypothesis [@Hanski1982-hs], is that competitive forces are significant only among regionally abundant and widespread “core species”, and the effects of competition are more pronounced in the composition of regional species pools than in that of local assemblages [@Hanski1982-ie].
-In some cases, competition for food may simply be precluded by a saturating abundance of floral resources [@Williams1989-vr], implying that populations are limited by other factors, such as abiotic stressors, nest site availability, or predator/pathogen pressure.
-
-Teasing apart the still-obscure mechanisms of bumble bee coexistence and niche partitioning remains a salient problem from a basic ecological perspective, but it has also become an urgent question in the context of applied conservation.
-Bumble bees are cold-adapted species [@Heinrich1994-cv] associated mainly with mountain habitats, where they serve as the principal pollinators of many alpine plants [@Williams1998-xi; @Goulson2010-lo].
-Warming climates have already caused measurable upslope shifts in bumble bees and their floral hosts [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc], and these are predicted to cause declines in many bumble bees species [@Soroye2020-zb].
-Moreover, the introduction or extinction of species from bumble bee communities, with corresponding addition and release of competitive pressure, can alter both the floral visitation patterns of bumble bees and, consequently, the pollination biology of their host plants [@Brosi2013-mu; @Ishii2013-sm].
-Thus, the speculative question of how competition has acted in the historic assembly of bumble bee communities has become an unfolding process observable over ecological time scales and instrumental in the future viability of bumble bee species and their floral mutualists.
+In some cases, competition for food may simply be precluded by a saturating abundance of floral resources, implying that populations are limited by other factors, such as abiotic stressors, nest site availability, or predator/pathogen pressure [@Ranta1981-vx; @Williams1989-vr].
 
@@
 These costs are closely related to the degree of morphological fittedness between bumble bees and their floral hosts, which consists principally in the matching of bumble bee tongue length to floral corolla depth.
-Bumble bee species can differ markedly in tongue length, and floral handling time increases when tongue length and corolla depth are poorly matched [@Inouye1980-mv; @Harder1983-bk], which leads to an association between tongue length and floral selection in bumble bee communities [e.g.
+Bumble bee species can differ markedly in tongue length, and floral handling time increases when tongue length and corolla depth are poorly matched [@Inouye1980-mv; @Harder1983-bk], which leads to an association between tongue length and floral choice in bumble bee communities [e.g.
 @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn].
-This association between tongue length and floral niche, though, is imperfect, and there can be considerable overlap in floral resource use between bumble bees of different tongue lengths [e.g.
+This association between tongue length and floral niche is imperfect, though, and there can be considerable overlap in floral resource use between bumble bees of different tongue lengths [e.g.
 @Ranta1980-ok; @Goulson2008-oe].
 
-Empirical measurements of niche overlap are, however, inherently ambiguous with respect to competition.
-It has been well-demonstrated that niche overlap cannot serve as a metric of competition *per se* [@Abrams1980-am], but it is often used as an index of *potential* competition, particularly in the context of dietary studies [e.g.
+While niche overlap cannot serve as a metric of competition *per se* [@Abrams1980-am], it is often used as an index of *potential* competition, particularly in the context of dietary studies [e.g.
 @French1985-eo; @Steffan-Dewenter2000-ll; @Jung2015-cw], based on the reasoning that if and when food becomes limiting, competition can be expected to vary in proportion to the dietary overlap of co-occurring species [@Pianka1972-oi].
-Focusing on *potential* competition does not, however, escape the equivocity of niche overlap, since an observed pattern of niche overlap could just as plausibly be interpreted as adaptive niche partitioning in response to *realized* competition [@Pianka1974-qk].
-This creates a predicament in which niche overlap can only be interpreted as an index of potential competition when realized competition is assumed to be negligible, an assumption that can only be justified on the basis of additional knowledge of the study system under consideration.
+Our study is set in the Berchtesgadener Alps of southeast Germany, a landscape of flower-rich mountain meadows embedded in a matrix of coniferous forest.
+This system is one of abundant floral resources but harsh abiotic conditions, so it is likely that competition for floral resources is low, or at least transient, and that bumble bee populations are limited instead by the shortness of the growing season.
+Yet while *realized* competition between bumble bees in our system may be low, the question of *potential* competition is highly relevant because mountain ecosystems are currently subject to at least two powerful drivers of directed change.
+First, warming climates have already caused measurable upslope shifts in bumble bees and their floral hosts [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc], which could introduce novel competitive interactions between bumble bee species that experience increased sympatry.
+Second, socioeconomic factors have led to a dramatic decline in traditional alpine grazing management, particularly in the European Alps, resulting in widespread pasture abandonment and reforestation of mountain meadows [@Tappeiner2006-jp; @Tasser2007-qo].
+The regression of meadows to forest entails a massive loss of floral resource availability, and even early stages of succession can cause declines in floral abundance and diversity, with corresponding effects on bumble bees [@Walcher2017-kw].
+At the same time, more accessible areas at the bases of mountains are undergoing agricultural intensification that displaces natural vegetation and favors grasses over forbs [@Tasser2002-om].
+Caught between adverse land use changes above and below and subject to climate-induced range shifts, mountain bumble bees are facing a future of reduced floral resource availability and altered sympatry, and potential competition is likely to become realized according to patterns of niche overlap between species.
 
-Our study is set in the Berchtesgadener Alps of southeast Germany, a landscape of flower-rich mountain meadows embedded in a matrix of coniferous forest.
-This system is one of rich floral resources but harsh abiotic conditions, so it is likely that competition for floral resources is low, or at least transient, and that bumble bees are limited instead by the shortness of the growing season and by fluctuating abiotic conditions during the critical period of spring colony founding.
-On this basis, we argue for the interpretation of niche overlap as an index of *potential* competition rather than a response to *realized* competition, with the caveat this interpretation may misrepresent certain temporal intervals (e.g.
-early spring) or elevational bands (e.g.
-the sparse alpine grasslands above the tree line ecotone) [Sponsler et al., submitted].
-
-The aim of our study is to synthesize the phenomena of floral preference, trait-matching, and dietary niche overlap in a way that accounts for the coexistence of bumble bee species with similar tongue lengths and sheds light on the possible effects of climate-driven elevational range shifts on the future of potential competition between bumble bee species in mountain habitats.
+Thus, the still-obscure mechanisms of bumble bee coexistence and niche partitioning remain a salient problem from a basic ecological perspective, but they have also become timely questions in the context of applied conservation.
+The aim of our study is to synthesize the phenomena of floral preference, trait-matching, and dietary niche overlap in a way that accounts for the coexistence of bumble bee species with similar tongue lengths and sheds light on the possible effects of habitat loss and climate-driven range shifts on the future of competition between bumble bee species in mountain habitats.
 We first ask how floral preferences compare across bumble bee species and to what extent these patterns can be explained by trait-matching between bumble bee tongue length and floral morphotype.
 We then investigate the patterns of dietary niche overlap between bumble bee species that emerge from their respective floral preferences, comparing niche overlap at the nested levels of floral morphotype, floral family, and floral genus.
-Finally, we ask how niche overlap varies with elevation and discuss the implications of this variation for bumble bee conservation in a changing climate.
+Finally, we ask how niche overlap varies with elevation and discuss the implications of this variation for bumble bee conservation in an era of global change.
 
@@
 ## Field system
-
-<!-- I guess I'll need to tweak the wording of this section so that it is not a verbatim copy of the other manuscript.
-But it's okay if it recycles a lot of language because it is, after all, the same thing being described.
--->
-
-<!-- Describe general floristic and temperature patterns. Describe sampling in more detail. -->
 
@@
 Sampling consisted of repeated visits (hereafter, “visitation samples”) to each study site at approximately weekly intervals.
-Samples were only collected on dry days when the air temperature was at least 6°C.
-During each visit, bumble bees and their activity were recorded during a 50-minute transect walk.
-Bumblebees observed on a given flower were counted as floral visitors.
-Bumble bee queens were identified to species level in the field, whereas workers and males were stored in individually labeled tubes in the freezer for later identification in the laboratory after @Amiet1996-jz.
-While floral visitation by male bumble bees was recorded during visitation sampling, we chose to analyze only visitation by queens and workers, which we pooled.
-In conjunction with visitation observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.1 m^2^.
-Plant species identification followed @Lauber2007-ha and @Oberndorfer2001-ro.
+Sampling was only conducted during periods without rain and with an air temperature of at least 6°C.
+Bumble bees were recorded during one 50-minute transect walk per visit, and bumblebees observed on a given flower were counted as floral visitors.
+Queens were identified to species level in the field to avoid destructively sampling colony foundresses, but workers and males were collected and frozen for later identification in the laboratory.
+Species were identified following @Amiet1996-jz.
+While male bumble bees was recorded, we chose to analyze only the pooled visitation of by queens and workers.
+In conjunction with bumble bee observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.01 m^2^ and identified the plant species according to @Lauber2007-ha and @Oberndorfer2001-ro.
 
-While per-sample effort was standardized across all years, the total number of samples (i.e.
-site-dates) varied markedly, with 153 in 2010, 212 in 2011, and 539 in 2012.
-The increased total sampling effort in 2012 was driven in part by the overall scarcity of bumble bees that year.
-Since the start of field sampling each year was determined by snowmelt, lower sites were sampled as early as April while higher sites could not be sampled until June or July.
-This resulted in a temporally staggered pattern of sampling in 2010 and 2012, while in 2011 a combination of early snowmelt and a late start to field activities resulted in sampling that was approximately aligned in time across the whole elevation gradient.
+While per-sample effort was standardized across all years, the total number of samples varied: 153 in 2010, 212 in 2011, and 539 in 2012.
+The start of field sampling each year was determined by snowmelt, so lower sites were sampled as early as April while higher sites could not be sampled until June or July.
+In 2010 and 2012, this resulted in a temporally staggered pattern of sampling, but a combination of early snowmelt and a late start to field activities in 2011 resulted in sampling that was approximately aligned in time across the whole elevation gradient.
+Due to these annual differences in sampling intensity and duration, we opted to pool data across years rather than include year effects in our models, which would multiply model complexity and be difficult to interpret.
+This means, however, that our analysis cannot detect year-to-year variation in floral preferences and niche overlap, and main effects cannot be distinguished from annual anomalies.
+With only three years of data, though, such annual inferences would be of little value, and we deemed them not worth the cost of added model complexity.
+
+The samples analyzed in this paper have been previously analyzed (Sponsler et al., submitted), but the analyses presented here are non-overlapping.
 
@@
 
-## Software
-
-All analyses were conducted in `R` [@RCoreTeam-2021aa].
-Data handling and visualization were performed with the `tidyverse` suite [@Wickham2019-aa].
-Bayesian multilevel models were implemented in Stan [@Stan2020aa] via the package `brms` [@Burkner2017-aa] using default priors, and visualizations were prepared with the packages `tidybayes` [@Kay2020-aa] and `ggdist` [@Kay2021-aa].
-Ordination analysis and permutation testing were performed with the package `vegan` [@Oksanen_2019] and visualized using the packages `ggplot2` [@Wickham2016-aa] and `ggvegan` [@Simpson2019aa].
-Hierarchical generalized additive models (HGAMs) were implemented with the package `mgcv` [@Wood_2017] and visualized with the packages `mgcViz` [@Fasiolo2018aa] and `tidymv` [@Coretta-aa2021].
-All `R` code used in this analysis, together with full model output and diagnostics, is available in Appendix X.
-
 ## Data processing
@@
 Using these species-level visitation matrices, we calculated pairwise niche overlap as 1 - Horn-Morisita distance (calculated separately for morphotype-, family-, and genus-level matrices) for all species pairs within each site.
-This approach was chosen for its normalizing effect on the highly variable "snapshot" visitation data generated by our individual transect walks, but it entails the disadvantage of conflating niche overlap due to similar floral preference and niche overlap due to similar temporal (i.e.
+This approach was chosen for its regularizing effect on the highly variable "snapshot" visitation data generated by our individual transect walks, but it entails the disadvantage of conflating niche overlap due to similar floral preference and niche overlap due to similar temporal (i.e.
 seasonal) activity patterns.
@@
 
-While our study spanned three years, we chose not to model year effects.
-Differences in sampling intensity and duration across years (described above) would make comparisons across years difficult to interpret.
-This means that our analysis cannot detect year-to-year variation in floral biases and niche overlap, and main effects cannot be distinguished from annual anomalies.
-With only three years of data, though, such annual inferences would be of little value, and we deemed them not worth the cost of added model complexity.
+## Statistical models
 
-## Statistical models
+All analyses were conducted in `R` [@RCoreTeam-2021aa].
+Data handling and visualization were performed with the `tidyverse` suite [@Wickham2019-aa].
+All `R` code used in this analysis, together with full model output and diagnostics, is available in Appendix X.
 
@@
 Visitation probability was then modeled using multilevel Bayesian regression with a Bernoulli distribution family (logit link function).
+Models were implemented in Stan [@Stan2020aa] via the package `brms` [@Burkner2017-aa], and visualizations were prepared with the packages `tidybayes` [@Kay2020-aa] and `ggdist` [@Kay2021-aa].
 *Bumble bee abundance* and *percent flower cover* (per morphotype) were included as a global interaction term, thus conditioning visitation probability on abundance and enabling the inference of preference independent of the neutral effects of abundance.
 *Floral morphotype* was specified as a both a varying intercept term and a varying slope term with respect to *percent flower cover*.
-*Site* was included as a varying intercept term to avoid pseudoreplication owing to the repeated observations within each site.
+*Site* was included as a varying intercept term to avoid overfitting due to pseudoreplication owing to the repeated observations within each site.
+We set a weakly informative (i.e.
+regularizing) Normal(0, 1.5) prior on the β-coefficients of the fixed effects [@Lemoine2019-xm] and used the default priors provided by `brms` version 2.14.4 for all other model parameters.
 
@@
 To explore the functional underpinnings of bumble foraging preference, we analyzed visitation probability as a function of the interaction between floral morphology and bumble bee tongue length class.
-For this analysis, visitation was pooled within floral morphotype but not across bumble bee species.
-We then specified a Bayesian model of the same form as Model 1, but this time included a varying slope and intercept term for floral morphotype : bumble bee tongue length class and floral morphotype : bumble bee species.
-In this model, the relationship between floral morphotype and bumble bee tongue length class was the focus of inference, and the bumble bee species effect was added to account for the grouping of observations within species.
+For this analysis, we pooled visitation within floral morphotype but not across bumble bee species.
+Then we specified a Bayesian model of the same form as Model 1, but this time included a varying slope and intercept term for *floral morphotype* : *bumble bee tongue length class* and *floral morphotype* : *bumble bee species*.
+In this model, the relationship between *floral morphotype* and *bumble bee tongue length class* was the focus of inference, and the *bumble bee species* effect was added to account for the grouping of observations within species.
 
@@
 
-To visualize the overall diet similarity between bumble bee species, we pooled visitation data across all samples within bumble bee species and plant genus, then performed a principal coordinates analysis based on Horn-Morisita distance metric.
+To visualize the overall diet similarity between bumble bee species, we pooled visitation data across all samples within bumble bee species and plant genus, then performed a principal coordinates analysis (PCoA) based on Horn-Morisita distance metric.
 To test the significance of tongue length class as a predictor of diet similarity, we performed a permutation test.
@@
 In this sense, it can be understood as a model of potential diet overlap given the aggregate foraging patterns of the bumble bee species in our system, with the understanding that realized niche overlap at particular locations or times may vary.
+PCoA and permutation tests were performed with the R package `vegan` [@Oksanen_2019], and visualizations were prepared with the packages `ggplot2` [@Wickham2016-aa] and `ggvegan` [@Simpson2019aa]
 
@@
 
-To investigate elevational patterns of *within-site* niche overlap, we modeled community-mean niche overlap as a function of level (morphotype, family, genus) and elevation using HGAMs.
-Community mean niche overlap, which we interpret as an index of potential *diffuse* competition [@Pianka1974-qk; @MacArthur1972-qz], was calculated by taking the mean niche overlap (within level) across all species pairs for each site.
+To investigate elevational patterns of *within-site* niche overlap, we modeled community-mean niche overlap as a function of level (morphotype, family, genus) and elevation using hierarchical generalized additive models (HGAMs).
+Community mean niche overlap, which can be interpreted as an index of potential *diffuse* competition [@Pianka1974-qk; @MacArthur1972-qz], was calculated by taking the mean niche overlap (within level) across all species pairs for each site.
 Level was included as both a fixed intercept effect and a smoothing factor in conjunction with elevation, following the type "I" model form described by @Pedersen2019-kf.
+Management was initially included as a fixed intercept effect, but its influence was insiginificant, so it was dropped from the final model.
 To account for the proportional nature of our data, we used a beta distribution family with a logit link function.
+Models were fit with the R package `mgcv` [@Wood_2017] and visualized with the packages `mgcViz` [@Fasiolo2018aa] and `tidymv` [@Coretta-aa2021].
 
@@
 We then took the mean genus-level niche overlap across all species pairs within tongue-difference class and site.
-Then, we modeled genus-level niche overlap using tongue-difference as both a fixed intercept effect and a smoothing factor in conjunction with elevation, again following the type "I" model form @Pedersen2019-kf.
+Then, we constructed an HGAM of genus-level niche overlap using tongue-difference as both a fixed intercept effect and a smoothing factor in conjunction with elevation, again following the type "I" model form @Pedersen2019-kf.
 
@@
 
-We recorded a total of 12,918 bumble- bee-flower interactions over the three years of our study.
-The metaweb across all sites and dates consisted of 16 bumble bee species (with *Bombus terrestric/lucorum* and *Bombus psithyrus* species lumped into species groups), 163 plant species (110 genera, 37 families), and 736 unique bumble- bee-plant interaction pairs.
-Five species --- *B.
+We recorded a total of 12,918 bumble- bee-flower interactions, including 16 bumble bee species (with *Bombus terrestric/lucorum* and *Bombus psithyrus* species lumped into species groups), 163 plant species (110 genera, 37 families), and 736 unique bumble- bee-plant interaction pairs.
+*B.
 pascuorum*, *B.
@@
 terrestris-lucorum*, and *B.
-wurflenii* --- accounted for the bulk of overall bumble bee abundance and were present over the entire elevation range of our sites.
+wurflenii* were the most abundant species, and they occurred at all sites.
 *B.
 hortorum*, *B.
-jonellus*, and *B.
-psithyrus* (species-group) were widespread but at lower abundance.
-*B.
-monticola* occurred at moderate abundance but was rare or absent at sites below 1000 m.
-The remaining species were relatively rare in our study system.
-*B.
+jonellus*,  *B.
+monticola*, and *B.
+psithyrus* occurred at moderate abundance.
+The remaining species --- *B.
+gerstaeckeri*, *B.
+hypnorum*, *B.
+lapidarius*, *B.
 mendax*, *B.
 mucidus*, and *B.
-pyrenaeus* occurred primarily above 1500 m, consistent with their known affinity for high elevation habitats [@Rasmont2010-ci].
+pyrenaeus* --- were relatively rare.
 *B.
-gerstaeckeri*, *B.
-hypnorum*, *B.
-lapidarius* were sparsely distributed across mid-elevation sites.
-*B.
-hypnorum* and *B.
-humilis* were recorded only three times and only in one year, so we omitted it from all analyses.
-
-Floral surveying yielded a total of 352 plant species, representing 191 genera and 52 families.
-Of these, 155 species from 103 genera and 35 families were observed to be visited by bumble bees.
-Eight species --- *Rubus idaeus*, *Rosa canina*, *Juniperus communis*, *Larix decidua*, *Salix* sp., *Caltha palustris*, *Pulmonaria officinalis*, and *Rheum barbarum* --- were recorded in visitation data but not recorded during floral surveying.
-Each accounted for no more than 4 visits in total over the three years of our study, and they were omitted from the analysis of floral abundance.
+humilis* was recorded only three times and only in one year, so we omitted it from all analyses.
+Floral surveying yielded a total of 352 plant species (191 genera, 52 families), of which 155 (103 genera, 35 families) observed to be visited by bumble bees.
+For a more full analysis of bumble bee and floral abundance patterns, see Sponsler at al.
+(submitted).
 
@@
 
-Lip (group 5) and flag (group 6) flowers had the highest overall visitation probability, with a conditional mean visitation probabilities of 0.35 (95% CI = 0.27–0.4) and 0.31 (95% CI = 0.24–0.40), respectively (Figure 1).
-In second place, head flowers (group 7) had a mean conditional visitation probability of 0.18 (95% CI = 0.13–0.25).
-Bell (group 3) flowers were a distant third with a conditional mean visitation probability of 0.09 (95% CI = 0.06–0.13), followed by stalk-disc (group 4), funnel (group 2), and disc flowers (group 1) with conditional visitation probabilities of 0.06 (95% CI = 0.04–0.08), 0.04 (95% CI = 0.03–0.07), and 0.04 (95% CI = 0.02–0.06), respectively.
+Lip and flag flowers were the most preferred morphotypes overall, with similarly high visitation probabilities of 0.35 (95% CI = 0.27–0.44) and 0.31 (95% CI = 0.24–0.40), respectively **(Figure 1)**.
+In second place, head flowers had a visitation probability of 0.18 (95% CI = 0.13–0.25).
+Bell flowers were a distant third with a visitation probability of 0.09 (95% CI = 0.06–0.13), followed by stalk-disc, funnel, and disc flowers with visitation probabilities of 0.06 (95% CI = 0.04–0.08), 0.04 (95% CI = 0.03–0.06), and 0.04 (95% CI = 0.02–0.06), respectively.
 Model 1 explained approximately 37% of the total variance (97.5% CI = 35–38%).
@@
 For details of original sources, see Appendix X.")}
-knitr::include_graphics("../output/brm_pvis01_plot1_mod.png")
+knitr::include_graphics("../output/brm_pvis_01_plot1_mod.png")
 ```
 
-Disc (group 1) flowers were visited at similar rates across tongue length classes **(Figure 2)**.
-At the species level, disc flowers were visited at relatively high rates by the short-tongued *B.
-terrestris/lucorum* and the medium-tongued *Bombus monticola* **(Figure 3)**.
-Funnel (group 2) flowers were preferred by long-tongued species, an effect driven by the *Aconitum* (group 2) specialist *B.
+Disc and stalk-disc flowers were visited at similarly low rates across tongue length classes, and lip flowers were visited at similarly high rates across tongue length classes **(Figure 2)**.
+There was considerable in visitation rates between individual species, though, including those belonging to the same tongue length class **(Figure 3)**.
+The short-tongued *B.
+terrestris/lucorum*, for example, stands out in its preference for disc flowers, and the medium-tongued *B.
+monticola* in its preference for stalk-disc flowers.
+While lip flowers were visited at relatively high rates by almost all species, the medium-tongued *B.
+pascuorum* exhibited an especially high preference for this morphotype, while the medium-tongued *B.
+mucidus* ignored lip flowers almost entirely.
+
+In contrast, visitation of funnel, bell, flag, and head flowers differed significantly both across tongue length classes **(Figure 2)** and between individual species **(Figure 3)**.
+Funnel flowers were preferred by long-tongued species, though effect was driven mainly by the *Aconitum* specialist *B.
 gerstaeckeri*.
-Bell (group 3) flowers were visited at a higher rate by short-tongued species than by long-tongued species, with medium-tongued species exhibiting intermediate visitation rates.
+Bell flowers were visited at a higher rate by short-tongued species than by long-tongued species, with medium-tongued species exhibiting intermediate visitation rates.
 At the species-level, bell flowers were most preferred by the short-tongued species *B.
@@
 monticola*.
-Stalk-disc (group 4) flowers were visited at similar rates across all tongue length classes, and there was also broad overlap in visitation rates at the species level.
-Lip (group 5) flowers were also visited at similar rates across tongue length classes, but at the species level *B.
-pascuorum* exhibited an uniquely high preference for lip flowers.
-Flag (group 6) flowers were visited at a higher rate by medium-tongued species than by short-tongued species, with long-tongued species exhibiting intermediate visitation rates.
+Flag flowers were visited at a higher rate by medium-tongued species than by short-tongued species, with long-tongued species exhibiting intermediate visitation rates.
 Flag flowers generated strong separation in visitation rate at the species level, receiving very high visitation from the medium-tongued *B.
@@
 
-```{r fig3, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability parsed by Kugler morphotype and bumble bee tongue length class (Model 2).
+```{r fig2, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability parsed by Kugler morphotype and bumble bee tongue length class (Model 2).
 Point-interval plots depict 66\\% and 95\\% CIs.
@@
 Note that the range of the x-axis has been rescaled for each panel to emphasize the comparison of visitation probabilities across tongue length classes within floral morphotypes.")}
-knitr::include_graphics("../output/brm_pvis00_plot1.png")
+knitr::include_graphics("../output/brm_pvis_00_plot1_mod.png")
 ```
 
-```{r fig4, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability by Kugler morphotype and bumble bee species (Model 3).
+```{r fig3, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability by Kugler morphotype and bumble bee species (Model 3).
 The tongue length class of each species is color-coded.
@@
 Note that the range of the x-axis has been rescaled for each panel to emphasize the comparison of visitation probabilities across bumble bee species within floral morphptypes.")}
-knitr::include_graphics("../output/brm_pvis02_plot1.png")
+knitr::include_graphics("../output/brm_pvis_02_plot1.png")
 ```
@@
 Principal coordinates analysis (Model 4) suggested weak but significant effects of tongue length on the diet similarity while also highlighting the particular floral genera that accounted most strongly for the dietary separation of bumble bee species.
-Axis 1 accounted for 23% of total diet variance and primarily separated long- and medium-tongued species from short-tongued species (Figure 4).
-Long- and medium-tongued species scored low on Axis one and were associated with the lip flowers (group 5) *Rhinanthus*, and *Stachys* and flag flowers (group 6) *Anthyllis*, *Lotus*, and *Trifolium*.
-Short tongued species scored high on Axis one and were associated with the bell flowers (group 3) *Campanula* and *Erica*, and the head flower (group 7) *Centaurea*.
+Axis 1 accounted for 23% of total diet variance and primarily separated long- and medium-tongued species from short-tongued species **(Figure 4)**.
+Long- and medium-tongued species scored low on Axis one and were associated with the lip flowers *Rhinanthus*, and *Stachys* and flag flowers *Anthyllis*, *Lotus*, and *Trifolium*.
+Short tongued species scored high on Axis one and were associated with the bell flowers *Campanula* and *Erica*, and the head flower *Centaurea*.
 Axis 2 accounted for 19% of total variance and primarily explained variation within short-tongued species.
@@
 Both the lip flower *Ajuga* and the head flower *Centaurea* scored near the middle of the ordination, reflecting their broad appeal across tongue length classes.
-Tongue length class explained 24% of diet variance (Adonis permutation test, F = 1.86, p = 0.003).
-While short-tongued bees appeared most broadly spread in ordination space, there were no significant differences in dispersion across tongue length classes (F = 0.26, p = 0.62).
+Tongue length class explained 24% of diet variance (Adonis permutation test, F = 1.79, p = 0.005).
+While short-tongued bees appeared most broadly spread in ordination space, there were no significant differences in dispersion across tongue length classes (F = 0.42, p = 0.66).
 
-```{r fig5, echo = FALSE,  fig.cap = paste("Principal coordinates ordination of bumble bee visitation pooled across all samples.
-To avoid overplotting, only plant genera with an absolute score of at least 0.22 on one or both axes are plotted.")}
+```{r fig4, echo = FALSE,  fig.cap = paste("Principal coordinates ordination of bumble bee visitation pooled across all samples.
+To avoid overplotting, only plant genera with an absolute score of at least 0.20 on one or both axes are plotted.")}
 knitr::include_graphics("../output/pcoa_aggregate.png")
@@
 Diet overlap was highest at the low end of the elevation gradient, decreased steeply from 600-1200 m, and leveled off or declined weakly for the remainder of the gradient, though morphotype-level diet overlap exhibited a second peak just above 1500 m that was not evident in genus- or family-level diet overlap.
-The Model 5 explained 78.4% of total deviance (adjusted R^2^ = 0.743).
-Whether a species pair belonged to the same or different tongue length classes had no effect on genus-level niche overlap, either in terms of intercept (p = 0.18) or shape of response to elevation **(Figure 9)**.
-Model 6 explained 36.5% of total deviance (adjusted R^2^ = 0.316).
+Model 5 explained 74.2% of total deviance (adjusted R^2^ = 0.70).
+Whether a species pair belonged to the same or different tongue length classes had no effect on genus-level niche overlap, either in terms of intercept (p = 0.12) or shape of response to elevation **(Figure 9)**.
+Model 6 explained 40.0% of total deviance (adjusted R^2^ = 0.35).
 
-```{r fig8, echo = FALSE,  fig.cap = paste("Niche overlap by level through elevation.
+```{r fig5, echo = FALSE,  fig.cap = paste("Niche overlap by level through elevation.
 Points show empirical niche overlap values, lines show GAM smooths, and shaded bands show 95\% confidence intervals.")}
@@
 
-```{r fig9, echo = FALSE,  fig.cap = paste("Niche overlap by tongue length class difference through elevation. Lines show GAM smooths and shaded bands show 95\% confidence intervals.")}
+```{r fig6, echo = FALSE,  fig.cap = paste("Niche overlap by tongue length class difference through elevation. Lines show GAM smooths and shaded bands show 95\% confidence intervals.")}
 knitr::include_graphics("../output/niche_gam_20_plot1.png")
@@
 
+<!-- Thus, the speculative question of how competition has acted in the historic assembly of bumble bee communities has become an unfolding process observable over ecological time scales and instrumental in the future viability of bumble bee species and their floral mutualists.
+-->
+
+<!-- Moreover, the introduction or extinction of species from bumble bee communities, with corresponding addition and release of competitive pressure, can alter the floral visitation patterns of bumble bees and the pollination rates of their host plants [@Brosi2013-mu; @Ishii2013-sm].
+-->
+
+<!-- Bumble bees are cold-adapted species [@Heinrich1994-cv] associated mainly with mountain habitats, where they serve as the principal pollinators of many alpine plants [@Williams1998-xi; @Goulson2010-lo].
+-->
+
 The question of how bumble bee species with similar tongue lengths coexist hinges on the relationships between species-level floral preferences, trait-matching, and the emergent phenomenon of diet similarity, and this set of relationships can be generalized to describe the niche relations of any group of organisms that share food resources.
 In this study, we clarify these relationships with respect to bumble bees and wild flowers interacting along a mountain slope elevation gradient.
-In addition to addressing the question of bumble bee species coexistence, our results suggest how elevational patterns of potential competition may change in response to climate-induced range shifts, an issue of immediate concern for the conservation of bumble bees and their floral mutualists [@Sirois-Delisle2018-vi; @Soroye2020-zb].
+In addition to addressing the question of bumble bee species coexistence, our results suggest how elevational patterns of potential competition may response to climate change, an issue of immediate concern for the conservation of bumble bees and their floral mutualists [@Sirois-Delisle2018-vi; @Soroye2020-zb].
 
@@
 
+Empirical measurements of niche overlap are, however, inherently ambiguous with respect to competition.
+
+Focusing on *potential* competition does not, however, escape the equivocity of niche overlap, since an observed pattern of niche overlap could just as plausibly be interpreted as adaptive niche partitioning in response to *realized* competition [@Pianka1974-qk].
+This creates a predicament in which niche overlap can only be interpreted as an index of potential competition when realized competition is assumed to be negligible, an assumption that can only be justified on the basis of additional knowledge of the study system under consideration.
+
+Our study is set in the Berchtesgadener Alps of southeast Germany, a landscape of flower-rich mountain meadows embedded in a matrix of coniferous forest.
+This system is one of rich floral resources but harsh abiotic conditions, so it is likely that competition for floral resources is low, or at least transient, and that bumble bees are limited instead by the shortness of the growing season and by weather fluctuations during the critical period of spring colony founding.
+On this basis, we argue for the interpretation of niche overlap as an index of *potential* competition rather than a response to *realized* competition, with the caveat this interpretation may misrepresent certain temporal intervals (e.g.
+early spring) or elevational bands (e.g.
+the sparse alpine grasslands above the tree line ecotone) [Sponsler et al., submitted].
+
+# Notes from conversation with Kalli
+
+- Explain overall preferences in figure 1 by noting that even short-tongued bumble bees represent the long-end of the complete pollinator spectrum; deeper flowers tend to have higher nectar rewards, and open flowers may be subject to competition from other visitors like Syrphids.
+
+- Tying competition back into the discussion.
+The premise of the paper is that we are modeling *potential competition* with the idea that it predicts realized competition if and when conditions change and competition ensues.
+So far, the main thrust has been climate change and how it could induce competition via changes in species distributions.
+But that is probably not the most important driver in Berchtesgaden.
+Prior to human influence, the landscape was mostly forested, and floral resources were presumably scarce, perhaps concentrated in forest gaps created by fire or tree falls or avalanches or whatever.
+When humans cleared the land for grazing, they created the flower-rich meadow habitats that now characterize the landscape.
+More recently, though, economic drivers have caused a withdraw of human management from mountain meadows and, consequently, reforestation [@Tasser2007-qo].
+At the same time, land use has intensified at lower elevations.
+So, mountain bumble bees may find themselves caught between the forest and the farm, so to speak, experience loss of habitat at high elevations due to the withdraw of human activity and loss of habitat at low elevations due to the intensification of human management.
+We can be confident that if these trends persist, competition will come into play.
+There is a wealth of literature on pasture abandonment and reforestation in the Alps, including at least one paper that deals explicitly with bumble bees (lots more that deal with floral diversity).
+
+- Explain that traits such as color (really especially color) probably explain much of the variance in preference not explained by morphology.
+
 # ACKNOWLEDGMENTS
 
-We thank D.J. McNeil for helpful conversations during data analysis. O. Mitesser kindly provided a high-performance computing environment for expediting Bayesian model fitting.
+We thank D.J. McNeil for helpful conversations during data analysis. O. Mitesser kindly provided a high-performance computing environment to expedite Bayesian model fitting.
 

### 2021-03-25 code/data_proc.R

@@ -29,11 +29,6 @@ fl_survey <- read_delim("./data/raw_data/data_plantcover_forDoug_2020.01.17.csv"
   
 write_csv(fl_survey, "./data/processed_data/floral_survey.csv")
 
-# 
-gentian_problem <- fl_survey %>%
-  select(plant.genus, plant.sp) %>%
-  filter(plant.genus %in% c("Gentiana", "Gentianella", "Gentianopsis")) %>%
-  distinct()
 
 ### Visitation data
 network <- read_delim("./data/raw_data/data_bumblebee_forDoug_2020.04.17.csv", delim = ";", 
@@ -45,7 +40,7 @@ network <- read_delim("./data/raw_data/data_bumblebee_forDoug_2020.04.17.csv", d
                                            "Deschampsia caespitosa" = "Deschampsia cespitosa",
                                            "Gymnadenia bifolia" = "Platanthera bifolia",
                                            "Gentiana aspera" = "Gentianella aspera",
-                                           "Gentiana ciliata" = "Gentianella ciliata")))%>%
+                                           "Gentiana ciliata" = "Gentianella ciliata"))) %>%
   separate(col = plant_sp_latin, into = c("genus", "species", "w", "x", "y", "z"), sep = " ") %>%
   unite(plant_sp_latin, genus, species, sep = " ") %>%
   dplyr::select(-c(w,x,y,z)) %>%
@@ -73,11 +68,6 @@ network <- read_delim("./data/raw_data/data_bumblebee_forDoug_2020.04.17.csv", d
   dplyr::select(-temp.date)
 
 write_csv(network, "./data/processed_data/network.csv")
-
-gentian_problem_visitation <- network %>%
-  select(plant.genus, plant.sp) %>%
-  filter(plant.genus %in% c("Gentiana", "Gentianella", "Gentianopsis")) %>%
-  distinct()
 
 ### Floral taxonomy
 Sys.setenv(ENTREZ_KEY = "ab9a55ec842df6f86a750929aefc69143608")
@@ -110,8 +100,8 @@ fl_tax_gapfill <- fl_tax %>% # manually fill in some gaps in the NCBI database
                                 plant.family)
   )
   
-write_csv(fl_tax_gapfill, "./processed_data/floral_tax.csv")
[…]

### 2021-03-26 ms_1/ms_1.Rmd
almost done ms2...

@@
 Bumble bees are characteristically cold-adapted species [@Heinrich1994-cv], with peak abundance and diversity in mountain ranges and northern latitudes, where they are often the dominant guild of flower-visiting insects and the principal pollinators of entomophilous flora [@Williams1998-xi; @Goulson2010-lo].
-Upslope movement of bumble bee species in response to climate change has already been documented, and this is expected to cause the declines or extinction of some bumble bee species [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc].
+Upslope movement of bumble bee species in response to climate change has already been documented, and this is expected to cause the declines or extinction of some bumble bee species [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; Biella2017-ni; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc].
 From a purely geometric perspective, populations of bumble bee species already restricted to the highest elevation zones can be expected to suffer habitat shrinkage [@Dirnbock2011-jf], possibly accompanied by increasing genetic isolation [though see @Fijen2020-zg].
@@
 This effect was especially pronounced in *B.
-pascuorum*, which was also the most abundant species in our study system overall.
+pascuorum*, likely because its exceptionally long nesting cycle is incompatible with the short growing season at high elevations [@Neumayer2009-gx].
 *B.
@@
 pratorum* was unique in exhibiting a sharp peak of abundance directly *at* the tree line.
+
+There’s a nice species-by-species discussion section.
+That would be a good place to look for descriptive anecdotes to draw upon in our writing.
+For example: “Bombus pascuorum (SCOPOLI 1763), the field bumblebee, is the most common bumblebee species of the alpine foothills and valley locations, but reaches its altitudinal limit in the area at about 1500 m.
+It can be assumed that the nest sites are even lower.
+Because of the extremely long nesting cycle of at least 5 months (REUTER & SCHWAMMBERGER 2002).
+there are limits to its altitudinal distribution.
+B.
+pascuorum was and is found in the ubiquitous distribution in submontane and montane regions and is one of the few bumblebee species which can still be found in cleared agricultural deserts.”...
+“Bombus pratorum (LINNAEUS 1761), the meadow bumblebee, is a species with an extremely short nest cycle, which can be found from the valley to the summits of the Gesäuse National Park.
+Its main distribution is in the montane zone, where it is often found in forested and near-forested areas.
+It is also well documented in the area by historical finds.
 

### 2021-03-26 ms_2/ms_2_v2.Rmd
almost done ms2...

@@
 On the one hand, there is evidence that resource limitation can be a significant constraint on local bumble communities [@Heinrich1976-wc; @Bowers1985-df], and certain bumble bee communities have been shown to be overdispersed in terms of tongue length or phylogeny, suggesting that competition functions as a community assembly process [@Inouye1977-kl; @Pellissier2013-pk].
-Nevertheless, bumble bee communities with an apparent deficit of trait (or phylogenetic) dispersion are commonly observed [@Ranta1981-vx; @Goulson2008-zy].
+Nevertheless, bumble bee communities with an apparent deficit of trait (or phylogenetic) dispersion are commonly observed [@Ranta1981-vx; @Goulson2008-zy; Iserbyt2008-jn].
 To account for this inconsistency, @Ranta1981-vx proposed that bumble bees communities are governed by a spatiotemporally heterogeneous mosaic of competitive conditions, such that competitive inequalities between species are sufficiently variable to allow more species to coexist than predicted by trait-based competition models.
 A complementary explanation, based on the core/satellite species hypothesis [@Hanski1982-hs], is that competitive forces are significant only among regionally abundant and widespread “core species”, and the effects of competition are more pronounced in the composition of regional species pools than in that of local assemblages [@Hanski1982-ie].
-In some cases, competition for food may simply be precluded by a saturating abundance of floral resources, implying that populations are limited by other factors, such as abiotic stressors, nest site availability, or predator/pathogen pressure [@Ranta1981-vx; @Williams1989-vr].
+In some cases, competition for food may simply be precluded by a saturating abundance of floral resources, implying that populations are limited by other factors, such as abiotic stressors, nest site availability, or predator/pathogen pressure [@Ranta1981-vx; @Williams1989-vr; Iserbyt2008-jn].
 
@@
 Our study is set in the Berchtesgadener Alps of southeast Germany, a landscape of flower-rich mountain meadows embedded in a matrix of coniferous forest.
-This system is one of abundant floral resources but harsh abiotic conditions, so it is likely that competition for floral resources is low, or at least transient, and that bumble bee populations are limited instead by the shortness of the growing season.
+This system is one of abundant floral resources but harsh abiotic conditions, so it is likely that competition for floral resources is low, or at least transient, and that bumble bee populations are limited instead by the shortness of the growing season, as has been suggested in the similar environment of the Eyne Valley of France [@Iserbyt2008-jn].
 Yet while *realized* competition between bumble bees in our system may be low, the question of *potential* competition is highly relevant because mountain ecosystems are currently subject to at least two powerful drivers of directed change.
@@
 Second, socioeconomic factors have led to a dramatic decline in traditional alpine grazing management, particularly in the European Alps, resulting in widespread pasture abandonment and reforestation of mountain meadows [@Tappeiner2006-jp; @Tasser2007-qo].
-The regression of meadows to forest entails a massive loss of floral resource availability, and even early stages of succession can cause declines in floral abundance and diversity, with corresponding effects on bumble bees [@Walcher2017-kw].
+The regression of meadows to forest entails a massive loss of floral resource availability, and even early stages of succession can cause declines in floral abundance and diversity, with corresponding effects on bumble bees [@Iserbyt2015-jc; @Walcher2017-kw; @Suzuki-Ohno2020-jv].
 At the same time, more accessible areas at the bases of mountains are undergoing agricultural intensification that displaces natural vegetation and favors grasses over forbs [@Tasser2002-om].
@@
 Rather than considering various floral morphological traits independently, we relied on the synthesis of @Kugler1970-gb, who developed a typology consisting of ten primary morphotypes and nested subdivision of each.
-Kugler’s typology was informed by laboratory studies of bumble bee flower choice [@Kugler1943-uy], and there is a precedent for analyzing bumble bee visitation in terms of Kugler’s morphotypes [@Neumayer1999-il; @Schneller2014-ry].
+Kugler’s typology was informed by laboratory studies of bumble bee flower choice [@Kugler1943-uy], and there is a precedent for analyzing bumble bee visitation in terms of Kugler’s morphotypes [@Farnberger1996-xo; @Neumayer1999-il; @Schneller2014-ry].
 
@@
 Funnel flowers were preferred by long-tongued species, though effect was driven mainly by the *Aconitum* specialist *B.
-gerstaeckeri*.
+gerstaeckeri* [@Ponchau2006-af; @Schneller2014-ry].
 Bell flowers were visited at a higher rate by short-tongued species than by long-tongued species, with medium-tongued species exhibiting intermediate visitation rates.
@@
 Note that the range of the x-axis has been rescaled for each panel to emphasize the comparison of visitation probabilities across tongue length classes within floral morphotypes.")}
-knitr::include_graphics("../output/brm_pvis_00_plot1_mod.png")
+knitr::include_graphics("../output/brm_pvis00_plot1_mod2.png")
 ```
@@
 Note that the range of the x-axis has been rescaled for each panel to emphasize the comparison of visitation probabilities across bumble bee species within floral morphptypes.")}
-knitr::include_graphics("../output/brm_pvis_02_plot1.png")
+knitr::include_graphics("../output/brm_pvis_02_plot1_mod.png")
 ```
@@
 To avoid overplotting, only plant genera with an absolute score of at least 0.20 on one or both axes are plotted.")}
-knitr::include_graphics("../output/pcoa_aggregate.png")
+knitr::include_graphics("../output/pcoa_aggregate_mod.png")
 ```
@@
 ```{r fig5, echo = FALSE,  fig.cap = paste("Niche overlap by level through elevation.
-Points show empirical niche overlap values, lines show GAM smooths, and shaded bands show 95\% confidence intervals.")}
+Points show empirical niche overlap values, lines show GAM smooths, and shaded bands show 95\\% confidence intervals.")}
 knitr::include_graphics("../output/niche_gam_00_plot1.png")
@@
 
-```{r fig6, echo = FALSE,  fig.cap = paste("Niche overlap by tongue length class difference through elevation. Lines show GAM smooths and shaded bands show 95\% confidence intervals.")}
+```{r fig6, echo = FALSE,  fig.cap = paste("Niche overlap by tongue length class difference through elevation. Lines show GAM smooths and shaded bands show 95\\% confidence intervals.")}
 knitr::include_graphics("../output/niche_gam_20_plot1.png")
@@
 
-<!-- Thus, the speculative question of how competition has acted in the historic assembly of bumble bee communities has become an unfolding process observable over ecological time scales and instrumental in the future viability of bumble bee species and their floral mutualists.
--->
+Any discussion of competition and niche partitioning among *Bombus* species must begin with the observation that bumble bees are unique among insect pollinators in the degree to which their foraging combines contrasting extremes of bias and flexibility.
+Coexistence between bumble bee species with similar tongue lengths hinges on the relationships between species-level floral preferences, trait-matching, and the emergent phenomenon of diet similarity.
+In this study, we clarify these relationships with respect to bumble bees and wild flowers interacting along a mountain slope elevation gradient.
+In addition to addressing the question of bumble bee species coexistence, our results suggest how patterns of competition may respond to climate and land use change along mountain slopes, an issue of immediate concern for the conservation of bumble bees and their floral mutualists [@Sirois-Delisle2018-vi; @Soroye2020-zb].
 
-<!-- Moreover, the introduction or extinction of species from bumble bee communities, with corresponding addition and release of competitive pressure, can alter the floral visitation patterns of bumble bees and the pollination rates of their host plants [@Brosi2013-mu; @Ishii2013-sm].
--->
+## Floral preference is only partly explained by the relationship between bumble bee tongue length and floral morphotype
 
-<!-- Bumble bees are cold-adapted species [@Heinrich1994-cv] associated mainly with mountain habitats, where they serve as the principal pollinators of many alpine plants [@Williams1998-xi; @Goulson2010-lo].
--->
+ While biases can be explained partly in terms of morphological matching between tongue length and floral morphotype
 
-The question of how bumble bee species with similar tongue lengths coexist hinges on the relationships between species-level floral preferences, trait-matching, and the emergent phenomenon of diet similarity, and this set of relationships can be generalized to describe the niche relations of any group of organisms that share food resources.
-In this study, we clarify these relationships with respect to bumble bees and wild flowers interacting along a mountain slope elevation gradient.
-In addition to addressing the question of bumble bee species coexistence, our results suggest how elevational patterns of potential competition may response to climate change, an issue of immediate concern for the conservation of bumble bees and their floral mutualists [@Sirois-Delisle2018-vi; @Soroye2020-zb].
+ The rule is flexibility, and floral morphotypes do not constitute hard boundaries to bumble bee flower choice.
+Nevertheless, we demonstrate that for certain floral morphotypes --- funnel flower, bell flowers, flag flowers, and head flower --- visitation rate is affected strongly by tongue length.
 
-While bumble bees are, indeed, remarkably generalized in their foraging, they are highly discriminating generalists, dramatically over-representing certain floral taxa and morphotypes relative to their abundance while ignoring others.
-In our study, lip, flag, and head flowers were by far the most strongly preferred when visitation was pooled across all bumble bee species, while bell, disc, and stalk-disc flowers were underrepresented.
-Patterns of pooled visitation are, of course, driven disproportionately by the behavior of the most abundant bumble bee species, which in our system were *B.
-pascuorum*, *B.
-pratorum*, *B.
-soroeensis*, *B.
-terrestris/lucorum*, and *B.
-wurflenii* (Sponsler et al., submitted).
-Visitation to lip and flag flowers was driven especially by the medium-tongued *B.
-pascuorum*, the most abundant bumble bee in our study system, and secondarily by the medium-tongued *B.
-wurflenii*.
-Visitation to head flowers was driven mainly by the short-tongued *B.
-soroeensis*.
-
-Tongue length class was an important predictor of preference for some but not all floral morphotypes.
-For disc and stalk-disc flowers, visitation rates were uniformly low across all tongue length classes.
-The opposite was true of lip flowers, which were visited at high rates by all tongue length classes.
-Bell flowers and head flowers were preferred by short-tongued bees, consistent with the shallow corollas of these morphotypes.
-Conversely, funnel flowers, with their deep corollas were preferred by long-tongued bees, though this effect was driven disproportionately by *B.
-gerstaeckeri*, which specializes on the funnel flowers of the genus *Aconitum* [@Ponchau2006-af].
-Flag flowers were visited at the highest rate by medium-tongued species, and especially by *B.
-pascuorum* and *B.
-mucidus*.
+In our study, lip, flag, and head flowers were by far the most strongly preferred when visitation was pooled across all bumble bee species, while disc, funnel, bell, and stalk-disc flowers were comparatively underrepresented.
+These findings are broadly consistent with the aggregate visitation patterns previously reported by @Schneller2014-ry in the central Alps at Vorarlberg and Neumayer1999-il ~50 km south of our study system, suggesting that our findings are at least regionally generalizeable.
+When considering these patterns from the perspective of pollination services to mountain flora, however, it is important to note that while bumble bees are the dominant Hymenopteran pollinators in our study system, Syrphid flies are also abundance flower visitors whose visitation is mainly restricted to the open flower morphotypes (e.g.
+disc flowers) that receive less attention from bumble bees [@Neumayer1999-il].
+In the context of our research question, we have emphasized the differences in tongue length between bumble bee species, but all bumble bees have relatively long tongues compared to the overall pollinator community, and the process of niche partitioning *among* bumble bee species is nested within the broader process of niche partitioning *between* bumble bees and other flower visitors.
 
@@
 
+<!-- - Explain that traits such as color (really especially color) probably explain much of the variance in preference not explained by morphology [@Neumayer1999-il].  -->
+
+## The influence of tongue length on floral preference does not carry on to genus-level niche overlap, and floral diversity within morphotypes relaxes potential competition between bumble bee species of similar tongue length
+
 With respect to diet overlap among bumble bees, our finding that bumble bee diet ordination produces a primary axis corresponding to tongue length (and associated with affinity for legumes) and a secondary axis describing the dispersion of short-tongued species is consistent with studies in Poland [@Goulson2008-oe], France [@Durieux2000-uk], and England [@Goulson2005-nz] that analyzed bumble bee diet by similar methods, suggesting that this may be a general pattern among bumble bee communities, at least within temperate Europe.
@@
 
+<!-- @Durieux2000-uk similarly ofund that while tongue length is correlated with floral preference, it is not correlated with diet similarity. But we did better stats! -->
+
+## Climate, land use, and the future of bumble bee niche relations
+
 Community mean diet overlap was highest at the lowest sites and declined sharply over the transition from valley floor to mid-slope.
@@
 
-As discussed earlier, the interpretation of floral visitation rates and diet overlap hinge strongly on the extent to which interspecific competition actually occurs in our study system.
+<!-- As discussed earlier, the interpretation of floral visitation rates and diet overlap hinge strongly on the extent to which interspecific competition actually occurs in our study system.
 We have argued that the assumption of minimal competition is reasonable in the context of the Berchtesgadener Alps, owing to their rich floral resources and restrictive abiotic conditions, together with the manifest co-occurrence of bumble bee species with similar tongue lengths.
@@
 Potential competition in the form of dietary overlap is perhaps the best indication of the effects that future instances of realized competition will have, and it is likely that the "ghosts of competition past" [@Connell1980-dn] haunt the floral preferences exhibited by bumble bees today.
+-->
 
-Empirical measurements of niche overlap are, however, inherently ambiguous with respect to competition.
+<!-- Empirical measurements of niche overlap are, however, inherently ambiguous with respect to competition.  -->
 
-Focusing on *potential* competition does not, however, escape the equivocity of niche overlap, since an observed pattern of niche overlap could just as plausibly be interpreted as adaptive niche partitioning in response to *realized* competition [@Pianka1974-qk].
+<!-- Focusing on *potential* competition does not, however, escape the equivocity of niche overlap, since an observed pattern of niche overlap could just as plausibly be interpreted as adaptive niche partitioning in response to *realized* competition [@Pianka1974-qk].
 This creates a predicament in which niche overlap can only be interpreted as an index of potential competition when realized competition is assumed to be negligible, an assumption that can only be justified on the basis of additional knowledge of the study system under consideration.
+-->
 
-Our study is set in the Berchtesgadener Alps of southeast Germany, a landscape of flower-rich mountain meadows embedded in a matrix of coniferous forest.
-This system is one of rich floral resources but harsh abiotic conditions, so it is likely that competition for floral resources is low, or at least transient, and that bumble bees are limited instead by the shortness of the growing season and by weather fluctuations during the critical period of spring colony founding.
-On this basis, we argue for the interpretation of niche overlap as an index of *potential* competition rather than a response to *realized* competition, with the caveat this interpretation may misrepresent certain temporal intervals (e.g.
-early spring) or elevational bands (e.g.
-the sparse alpine grasslands above the tree line ecotone) [Sponsler et al., submitted].
+<!-- - Explain that traits such as color (really especially color) probably explain much of the variance in preference not explained by morphology.  -->
 
-# Notes from conversation with Kalli
+<!-- Bumble bees are cold-adapted species [@Heinrich1994-cv] associated mainly with mountain habitats, where they serve as the principal pollinators of many alpine plants [@Williams1998-xi; @Goulson2010-lo].
+The introduction or extinction of species from bumble bee communities, with corresponding addition and release of competitive pressure, can alter the floral visitation patterns of bumble bees and the pollination rates of their host plants [@Brosi2013-mu; @Ishii2013-sm].
+-->
 
-- Explain overall preferences in figure 1 by noting that even short-tongued bumble bees represent the long-end of the complete pollinator spectrum; deeper flowers tend to have higher nectar rewards, and open flowers may be subject to competition from other visitors like Syrphids.
-
-- Tying competition back into the discussion.
-The premise of the paper is that we are modeling *potential competition* with the idea that it predicts realized competition if and when conditions change and competition ensues.
-So far, the main thrust has been climate change and how it could induce competition via changes in species distributions.
-But that is probably not the most important driver in Berchtesgaden.
-Prior to human influence, the landscape was mostly forested, and floral resources were presumably scarce, perhaps concentrated in forest gaps created by fire or tree falls or avalanches or whatever.
-When humans cleared the land for grazing, they created the flower-rich meadow habitats that now characterize the landscape.
-More recently, though, economic drivers have caused a withdraw of human management from mountain meadows and, consequently, reforestation [@Tasser2007-qo].
-At the same time, land use has intensified at lower elevations.
-So, mountain bumble bees may find themselves caught between the forest and the farm, so to speak, experience loss of habitat at high elevations due to the withdraw of human activity and loss of habitat at low elevations due to the intensification of human management.
-We can be confident that if these trends persist, competition will come into play.
-There is a wealth of literature on pasture abandonment and reforestation in the Alps, including at least one paper that deals explicitly with bumble bees (lots more that deal with floral diversity).
-
-- Explain that traits such as color (really especially color) probably explain much of the variance in preference not explained by morphology.
+<!-- Lip and head flowers have the highest nectar reward per inflorescence [@Neumayer1999-il]. -->
+<!-- Interestingly, @Farnberger1996-xo in a localized study near Wien found rather different results.
+Lip flowers were still found to be the favorite overall, and head flowers were popular, too, but flag flowers were of secondary importance, and bell flowers were the most visited morphotype for humi, pasc, and prat.
+Pasc almost totally ignored bell flowers in our study.
+-->
 

### 2021-03-29 code/ms2_bias_niche.Rmd
almost done ms2...

@@
 ggsave("../output/pcoa_aggregate.png")
+ggsave("../output/pcoa_aggregate.pdf")
+
 ```
@@
   summarize(mean.niche.overlap = mean(niche.overlap)) %>%
-  filter(level == "genus")
+  filter(level == "genus") %>%
+  mutate(tongue.pair = factor(tongue.pair, levels = c("short:long", "short:med", "short:short",
+                                                      "med:long", "med:med", "long:long")))
 
@@
 # Plot intercept effect of tongue.diff
-plot(pterm(niche_gam_20, 1))
-```
+plot(pterm(niche_gam_21, 1))
+```

### 2021-03-29 ms_2/ms_2_v2.Rmd
almost done ms2...

@@
 Bumble bee species can differ markedly in tongue length, and floral handling time increases when tongue length and corolla depth are poorly matched [@Inouye1980-mv; @Harder1983-bk], which leads to an association between tongue length and floral choice in bumble bee communities [e.g.
-@Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn].
+@Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn].
 This association between tongue length and floral niche is imperfect, though, and there can be considerable overlap in floral resource use between bumble bees of different tongue lengths [e.g.
@@
 
-To focus on genus-level niche overlap as a function of tongue length class, we generated a binary tongue-difference variable for each species pair: "same" for species belonging to the same tongue length class and "different" for species belonging to different tongue length classes.
+<!-- To focus on genus-level niche overlap as a function of tongue length class, we generated a binary tongue-difference variable for each species pair: "same" for species belonging to the same tongue length class and "different" for species belonging to different tongue length classes.
 We then took the mean genus-level niche overlap across all species pairs within tongue-difference class and site.
 Then, we constructed an HGAM of genus-level niche overlap using tongue-difference as both a fixed intercept effect and a smoothing factor in conjunction with elevation, again following the type "I" model form @Pedersen2019-kf.
+-->
+
+To focus on genus-level niche overlap as a function of tongue length class, we classified each species pair as a combination of "short:short", "short:medium", "short:long", "medium:medium", "medium:long", or "long:long".
+Within each site, we then took the mean genus-level niche overlap across all species pairs within each tongue-pair class.
+Then, we constructed an HGAM of genus-level niche overlap using tongue-pair as both a fixed intercept effect and a smoothing factor in conjunction with elevation, again following the type "I" model form @Pedersen2019-kf.
 
@@
 Model 5 explained 74.2% of total deviance (adjusted R^2^ = 0.70).
-Whether a species pair belonged to the same or different tongue length classes had no effect on genus-level niche overlap, either in terms of intercept (p = 0.12) or shape of response to elevation **(Figure 9)**.
+
+<!-- Whether a species pair belonged to the same or different tongue length classes had no effect on genus-level niche overlap, either in terms of intercept (p = 0.12) or shape of response to elevation **(Figure 9)**.
 Model 6 explained 40.0% of total deviance (adjusted R^2^ = 0.35).
+-->
+
+We recommend caution, however, in interpreting the p-values of this model, since the pairwise comparison of group differences entails a degree of non-independence.
 
@@
 ```{r fig6, echo = FALSE,  fig.cap = paste("Niche overlap by tongue length class difference through elevation. Lines show GAM smooths and shaded bands show 95\\% confidence intervals.")}
-knitr::include_graphics("../output/niche_gam_20_plot1.png")
+knitr::include_graphics("../output/niche_gam_21_plot1.png")
 ```
@@
 
-Any discussion of competition and niche partitioning among *Bombus* species must begin with the observation that bumble bees are unique among insect pollinators in the degree to which their foraging combines contrasting extremes of bias and flexibility.
-Coexistence between bumble bee species with similar tongue lengths hinges on the relationships between species-level floral preferences, trait-matching, and the emergent phenomenon of diet similarity.
-In this study, we clarify these relationships with respect to bumble bees and wild flowers interacting along a mountain slope elevation gradient.
-In addition to addressing the question of bumble bee species coexistence, our results suggest how patterns of competition may respond to climate and land use change along mountain slopes, an issue of immediate concern for the conservation of bumble bees and their floral mutualists [@Sirois-Delisle2018-vi; @Soroye2020-zb].
-
 ## Floral preference is only partly explained by the relationship between bumble bee tongue length and floral morphotype
 
- While biases can be explained partly in terms of morphological matching between tongue length and floral morphotype
+Any discussion of competition and niche partitioning among *Bombus* species must begin with the observation that bumble bees are unique among insect pollinators in the degree to which their foraging combines contrasting extremes of bias and flexibility.
+While the floral preferences of bumble bee species can be explained partly by morphological matching between tongue length and floral morphotype, floral morphotypes are not hard boundaries to flower choice [@Kugler1943-uy; @Brian1954-cy].
+Both terms of this paradox are evident in our study.
 
- The rule is flexibility, and floral morphotypes do not constitute hard boundaries to bumble bee flower choice.
-Nevertheless, we demonstrate that for certain floral morphotypes --- funnel flower, bell flowers, flag flowers, and head flower --- visitation rate is affected strongly by tongue length.
-
-In our study, lip, flag, and head flowers were by far the most strongly preferred when visitation was pooled across all bumble bee species, while disc, funnel, bell, and stalk-disc flowers were comparatively underrepresented.
-These findings are broadly consistent with the aggregate visitation patterns previously reported by @Schneller2014-ry in the central Alps at Vorarlberg and Neumayer1999-il ~50 km south of our study system, suggesting that our findings are at least regionally generalizeable.
-When considering these patterns from the perspective of pollination services to mountain flora, however, it is important to note that while bumble bees are the dominant Hymenopteran pollinators in our study system, Syrphid flies are also abundance flower visitors whose visitation is mainly restricted to the open flower morphotypes (e.g.
+Lip, flag, and head flowers were by far the most strongly preferred when visitation was pooled across all bumble bee species, while disc, funnel, bell, and stalk-disc flowers were comparatively underrepresented.
+These findings are broadly consistent with the aggregate visitation patterns previously reported by @Schneller2014-ry in the central Alps at Vorarlberg and Neumayer1999-il in the Austrian Alps ~55 km south of our study system, suggesting that our findings are at least regionally generalizeable.
+When considering these patterns from the perspective of pollination services to mountain flora, it is important to note that while bumble bees are the dominant Hymenopteran pollinators in our study system, flies (especially Syrphid hoverflies) are also abundant flower visitors in mountain ecosystems, and their visitation is mainly restricted to the open flower morphotypes (e.g.
 disc flowers) that receive less attention from bumble bees [@Neumayer1999-il].
@@
 
-While the variation we observed in floral preference across tongue length classes largely comports with the established relationship between tongue length and corolla depth, variation in preference *within* tongue length classes agrees with studies demonstrating the insufficiency of simple morphological matching in explaining bumble bee foraging patterns [e.g.
+The relationship between tongue length and floral preference was strong for funnel, bell, flag, and head flowers, which largely comports with the the established relationship between tongue length and corolla depth [e.g.
+@Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn].
+Nevertheless, tongue length was a poor predictor of preference for disc, stalk-disc, and lip flowers, which agrees with studies demonstrating the insufficiency of simple morphological matching as an explanation of bumble bee foraging patterns [e.g.
 @Ranta1980-ok; @Goulson2008-oe].
+Where tongue length fails as a predictor of floral preference, it can do so in two ways: the *convergence* of preference across species of *different* tongue length or the *divergence* of preference among species of *similar* tongue length.
+An example of the former is the near-universal affinity for lip flowers among the bumble bee species in our study.
+The only exception to this rule was the rare high-elevation species *B.
+mucidus*, which was effectively a flag flower specialist in our study.
+Lip flowers are known to be favored by bumble bees, particularly in the Alps [@Schneller2014-ry], and the apparent consistency of their attractiveness across tongue length differences might be due to their relatively permissive floral morphology (compared to flag or funnel flowers, for example) and/or their tendency to offer high per-inflorescence rewards compared to other floral morphotypes [@Neumayer1999-il].
 Perhaps the most striking divergence of floral preference within tongue length class was the divisive effect of bell flowers on short-tongued bumble bees.
@@
 There are no obvious morphological distinctions within short-tongued bumble bees that would account for these divergent affinities, and ecological drivers such as differences in spatial or temporal activity patterns are more likely explanations [@Sowig1989-ld; @Schneller2014-ry].
-Nectar robbing is another mechanism by which the relationship between tongue length and corolla depth might be relaxed, but its influence in our data is unclear.
+Nectar robbing is another mechanism by which the relationship between tongue length and corolla depth might be relaxed [@Brian1954-cy], but its influence in our data is unclear.
 Of the bumble bees in our study system, *B.
 terrestris/lucorum* and especially *B.
-wurflenii* are known to engage frequently in nectar robbing [@Brian1954-cy].
+wurflenii* are known to engage frequently in nectar robbing [@Brian1954-cy; @Neumayer1999-il].
 In the case of *B.
-wurflenii*, however, its morphotype affinities did not differ markedly from those of other medium-tongued species, and it may be that the condition of medium-tongue length alone confers enough to foraging flexibility in our study system to dampen the divergent effects of nectar robbing.
+wurflenii*, however, its morphotype affinities did not differ markedly from those of other medium-tongued species, and it may be that the condition of medium-tongue length alone confers enough foraging flexibility in our study system to dampen the divergent effects of nectar robbing.
 Nectar robbing might, however, account for the deviation of *B.
@@
 terrestris/lucorum* was, on the whole, an eclectic forager, also demonstrating a relatively high affinity for the disc flowers that were largely ignored by other species.
-It should be noted, though, that an unusually high abundance of *B.
-terrestris/lucorum* queens was observed in 2012, but nest founding success was evidently low, leading to low *B.
-terrestris/lucorum* abundance later in the season, an anomaly that could account for the unusual foraging patterns of this species group in the pooled data.
 Together, these species-level idiosyncracies of preference within tongue length classes underscore the fact that factors other than the matching of tongue length and floral morphology --- such as affinities based on floral color, plant growth form, spatial distribution of floral patches, or habitat context [@Brian1957-dg; Harder1985-ls; @Sowig1989-ld] --- are at play in bumble bees' floral preferences, which may partly explain the coexistence of bumble bees of similar tongue lengths.
-
-<!-- - Explain that traits such as color (really especially color) probably explain much of the variance in preference not explained by morphology [@Neumayer1999-il].  -->
 
@@
 
-With respect to diet overlap among bumble bees, our finding that bumble bee diet ordination produces a primary axis corresponding to tongue length (and associated with affinity for legumes) and a secondary axis describing the dispersion of short-tongued species is consistent with studies in Poland [@Goulson2008-oe], France [@Durieux2000-uk], and England [@Goulson2005-nz] that analyzed bumble bee diet by similar methods, suggesting that this may be a general pattern among bumble bee communities, at least within temperate Europe.
-Perhaps a more significant finding, though, is that community-mean diet overlap at the level of floral morphotype was consistently nearly twice as high as diet overlap at the level of floral genus.
+While tongue length was a relatively strong (though imperfect) predictor of bumble bee species' morphotype preferences, it was a weak predictor of diet similarity (at the level of floral genus) across bumble bees species, echoing the findnings of @Durieux2000-uk.
+Diet ordination produced a primary axis corresponding to tongue length (and associated with affinity for legumes) and a secondary axis describing the dispersion of short-tongued species, which is consistent with studies in Poland [@Goulson2008-oe], France [@Durieux2000-uk], and England [@Goulson2005-nz] that analyzed bumble bee diet by similar methods, suggesting that this may be a general pattern among bumble bee communities, at least within temperate Europe.
+Nevertheless, the ordination explained less than 50% of total variance, and while short-tongued species separated clearly from long- and medium-tongued species, the latter overlapped extensively with each other.
+These results were corroborated by our HGAM analysis of niche overlap, where the only significant in overall niche overlap was between short- and long-tongued species.
+A caveat to these analyses, though, is that our bumble bee community included only 3 long-tongued species (*B.
+gerstaeckeri*, *B.
+hortorum*, and *B.
+mendax*), two of which (*B.
+gerstaeckeri* and *B.
+mendax*) were relatively rare.
+It is possible that clearer patterns of tongue-length-based diet differentiation would be seen in a more balanced community.
+
+Community-mean diet overlap at the level of floral morphotype was consistently nearly twice as high as diet overlap at the level of floral genus.
 From a numerical perspective, this result is unsurprising simply because each morphotype comprises many genera.
-From an ecological perspective, though, this pattern suggests another answer to the question of how bumble bee species of similar tongue length can coexists.
-If tongue length primarily constrains bumble bee foraging at the coarse level of floral morphotype, floral diversity *within* morphotypes creates space for bumble bees with similar tongue lengths to avoid actual dietary overlap, thus mitigating interspecific competition.
-Importantly, this suggests that diversity within floral morphotypes may as important as morphotype diversity *per se* and overall floral abundance in facilitating the coexistence of bumble bee species, an insight which may have implications for the design of floral enhancements aimed at bumble bee conservation.
-
-<!-- @Durieux2000-uk similarly ofund that while tongue length is correlated with floral preference, it is not correlated with diet similarity. But we did better stats! -->
+From an ecological perspective, though, this pattern suggests another answer to the question of how bumble bee species of similar tongue length can coexists: if tongue length primarily constrains bumble bee foraging at the coarse level of floral morphotype, floral diversity *within* morphotypes creates space for bumble bees with similar tongue lengths to avoid actual dietary overlap, thus mitigating interspecific competition.
 
@@
 In our system, though, it must be noted that the lowest elevation sites are also the ones most influenced by human land use, and it is possible that the elevation pattern evident in our data is driven by factors other than elevational climate variation.
+
+## Conclusion
+Coexistence between bumble bee species with similar tongue lengths hinges on the relationships between species-level floral preferences, trait-matching, and the emergent phenomenon of diet similarity.
+In this study, we clarify these relationships with respect to bumble bees and wild flowers interacting along a mountain slope elevation gradient.
+In addition to addressing the question of bumble bee species coexistence, our results suggest how patterns of competition may respond to climate and land use change along mountain slopes, an issue of immediate concern for the conservation of bumble bees and their floral mutualists [@Sirois-Delisle2018-vi; @Soroye2020-zb].
+
+This suggests that diversity within floral morphotypes may as important as morphotype diversity *per se* and overall floral abundance in facilitating the coexistence of bumble bee species, an insight which may have implications for the design of floral enhancements aimed at bumble bee conservation.
 
@@
 
-<!-- Lip and head flowers have the highest nectar reward per inflorescence [@Neumayer1999-il]. -->
+<!-- Lip and head flowers have the highest nectar reward per inflorescence [@Neumayer1999-il] cited in . -->
 <!-- Interestingly, @Farnberger1996-xo in a localized study near Wien found rather different results.

### 2021-04-06 code/ms1_abundance.Rmd
added management covariate to models

--- 2021-03-16 code/ms1_abundance.Rmd
@@
 
-## 1.4 Calculate FL and BB abundance / diversity
+## 1.4 Calculate FL and BB abundance
 ```{r echo=FALSE, include=TRUE, message=FALSE}
@@
   mutate(bb.sp.year = factor(bb.sp.year)) %>%
-  left_join(site_data)
+  left_join(site_data) %>%
+  mutate(management = factor(management))
 ```
 
-### 2.2.2 Preliminary visualizations
+### 2.2.2 Call models
 ```{r}
-## mean floral abundance
-ggplot(abund_sum, aes(elev.mean, mean.fl.abund, color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 8), se = FALSE) +
-  #facet_wrap(~bb.sp, scales = "free_y") +
-  geom_vline(xintercept = 1500)
+# BB
+# gam_max_bb_abund_00 <- gam(max.bb.abund ~ 
+#                              s(bb.sp, bs = "re") + # random intercept for bb.sp
+#                              s(bb.sp.year, bs = "re") + # random intercept for bb.sp:year
+#                              s(elev.mean, by = bb.sp, k = 8), # smoother for elev by bb.sp.
+#                            method = "REML",
+#                            family = "tw", # Tweedie distribution performs well
+#                            select = TRUE,
+#                            data = abund_sum) %>% getViz()
+# 
+# # FL
+# gam_mean_fl_abund_00 <- gam(mean.fl.abund ~
+#                               year +
+#                               s(bb.sp, bs = "re") +
+#                               s(elev.mean, by = bb.sp, k = 8),
+#                            method = "REML",
+#                            family = "tw",
+#                            select = TRUE,
+#                            data = abund_sum) %>% getViz()
 
-## log(mean floral abundance)
-ggplot(abund_sum, aes(elev.mean, log(mean.fl.abund), color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 8), se = FALSE) +
-  facet_wrap(~bb.sp) +
-  geom_vline(xintercept = 1500)
-
-## Max floral abundance)
-ggplot(abund_sum, aes(elev.mean, max.fl.abund, color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 8), se = FALSE) +
-  facet_wrap(~bb.sp) +
-  geom_vline(xintercept = 1500)
-
-## Max BB abundance
-ggplot(abund_sum, aes(elev.mean, max.bb.abund, color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 8),  se = FALSE) +
-  facet_wrap(~bb.sp, scales = "free_y") +
-  geom_vline(xintercept = 1500)
-
-## Max  abundance ~ mean BB abundance, to see how interchangeable they are.
-## They are almost perfectly correlated, and I think max abund is more informative
-ggplot(abund_sum, aes(mean.bb.abund, max.bb.abund, color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 8),  se = FALSE) +
-  facet_wrap(~bb.sp, scales = "free_y")
-
-## Are max FL and mean FL similarly interchangeable? Yes.
-ggplot(abund_sum, aes(mean.fl.abund, max.fl.abund, color = year)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 8),  se = FALSE) +
-  facet_wrap(~bb.sp, scales = "free_y")
-
-## Max BB abundance ~ mean FL abundance
-ggplot(abund_sum, aes(log(mean.fl.abund), max.bb.abund)) +
-  geom_point(alpha = 0.5) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 8),  se = FALSE) +
-  facet_wrap(~bb.sp, scales = "free_y")
-```
-
-### 2.2.3 Abundance models
-I'm choosing to use site as a random intercept effect in these models because I think that is the correct way to specify a model involving repeated observations (in this case, bumble bee species) within site.
-The effect of this decision is that the elevation smoother is less influenced by individual sites and converges more strongly toward the mean.
-This creates less exciting curves, but I think it is a more principled way to model the data and less prone to overfitting.
-I can have my cake and eat it too, though, by plotting the actual data over the GAM fit so that the strong patterns of certain sites are visible.
-Then I can talk about the conditional elevation effect and the importance of certain peaks that are likely to be idiosyncrasies of my system (i.e.
-it would be overfitting to call them elevation effects).
-
-#### 2.2.3.1 Bumble bees
-```{r}
-# Call model
-gam_max_bb_abund_00 <- gam(max.bb.abund ~ 
-                             s(bb.sp, bs = "re") +
-                             s(bb.sp.year, bs = "re") + # fixed intercept for bb.sp:year
-                             s(elev.mean, by = bb.sp, k = 8), # smoother for elev by BB sp.
+# BB
+gam_max_bb_abund_01 <- gam(max.bb.abund ~ 
+                             management +
+                             s(bb.sp, bs = "re") + # random intercept for bb.sp
+                             s(bb.sp.year, bs = "re") + # random intercept for bb.sp:year
+                             s(elev.mean, by = bb.sp, k = 8), # smoother for elev by bb.sp.
                            method = "REML",
-                           family = "tw", # Tweedie distribution performs well
+                           family = "nb",
                            select = TRUE,
@@
 
-gam_max_bb_abund_00_ilink <- family(gam_max_bb_abund_00)$linkinv # get inverse link function
-
-check.gamViz(gam_max_bb_abund_00) # evaluate model
-gam_max_bb_abund_00_sum <- summary(gam_max_bb_abund_00) # summarize model
-print(plot(gam_max_bb_abund_00, allTerms = TRUE), pages = 1) # preliminary visualization
-
-# Plot smooths and year effect on the link scale
-## Individual plots
-gers_abund <- plot(sm(gam_max_bb_abund_00, 3)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "gers")
-hort_abund <- plot(sm(gam_max_bb_abund_00, 4)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "hort")
-jone_abund <- plot(sm(gam_max_bb_abund_00, 5)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "jone")
-mend_abund <- plot(sm(gam_max_bb_abund_00, 6)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "mend")
-mont_abund <- plot(sm(gam_max_bb_abund_00, 7)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "mont")
-muci_abund <- plot(sm(gam_max_bb_abund_00, 8)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "muci")
-pasc_abund <- plot(sm(gam_max_bb_abund_00, 9)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "pasc")
-prat_abund <- plot(sm(gam_max_bb_abund_00, 10)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "prat")
-psit_abund <- plot(sm(gam_max_bb_abund_00, 11)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "psit")
-pyre_abund <- plot(sm(gam_max_bb_abund_00, 12)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "pyre")
-soro_abund <- plot(sm(gam_max_bb_abund_00, 13)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "soro")
-telu_abund <- plot(sm(gam_max_bb_abund_00, 14)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "telu")
-wurf_abund <- plot(sm(gam_max_bb_abund_00, 15)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "wurf")
-
-## Arrange significant smooths on grid plot
-grid_bb_abund <- gridPrint(gers_abund, mend_abund, mont_abund, muci_abund, 
-                           pasc_abund, prat_abund, psit_abund, pyre_abund, 
-                           soro_abund, telu_abund, wurf_abund, ncol = 4)
-
-pdf("../output/gam_max_bb_abund_grid.pdf", width = 8, height = 6)
-plot(grid_bb_abund)
-dev.off()
-
-png("../output/gam_max_bb_abund_grid.png", res = 300, width = 8, height = 6, units = "in")
-plot(grid_bb_abund)
-dev.off()
-```
-
-#### 2.2.3.2 Floral resources
-```{r}
-gam_mean_fl_abund_00 <- gam(mean.fl.abund ~
-                              s(elev.mean, bb.sp, bs = "fs", k = 8) +
-                              s(elev.mean, year, bs = "fs", k = 8),
+# FL
+gam_mean_fl_abund_01 <- gam(mean.fl.abund ~
+                              management +
+                              year +
+                              s(bb.sp, bs = "re") +
+                              s(elev.mean, by = bb.sp, k = 8),
                            method = "REML",
@@
                            data = abund_sum) %>% getViz()
+```
 
-check.gamViz(gam_mean_fl_abund_00) # evaluate model
-gam_mean_fl_abund_00_sum <- summary(gam_mean_fl_abund_00) # summarize model
-print(plot(gam_mean_fl_abund_00, allTerms = TRUE), pages = 1) # preliminary visualization
+### 2.2.3 Evaluate model
+```{r}
+# BB
+check.gamViz(gam_max_bb_abund_01) # evaluate model
+gam_max_bb_abund_01_sum <- summary(gam_max_bb_abund_01) # summarize model
+print(plot(gam_max_bb_abund_01, allTerms = TRUE), pages = 1) # preliminary visualization
 
+# FL
+check.gamViz(gam_mean_fl_abund_01) # evaluate model
+gam_mean_fl_abund_01_sum <- summary(gam_mean_fl_abund_01) # summarize model
+print(plot(gam_mean_fl_abund_01, allTerms = TRUE), pages = 1) # preliminary visualization
+
+```
+
+### 2.2.4 Visualize
+
+#### 2.2.4.1 BB
+```{r}
+# Plot smooths and year effect on the link scale
+
+## Individual plots
+gers_abund <- plot(sm(gam_max_bb_abund_01, 3)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "gers")
+hort_abund <- plot(sm(gam_max_bb_abund_01, 4)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "hort")
+jone_abund <- plot(sm(gam_max_bb_abund_01, 5)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "jone")
+mend_abund <- plot(sm(gam_max_bb_abund_01, 6)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "mend")
+mont_abund <- plot(sm(gam_max_bb_abund_01, 7)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "mont")
+muci_abund <- plot(sm(gam_max_bb_abund_01, 8)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "muci")
+pasc_abund <- plot(sm(gam_max_bb_abund_01, 9)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "pasc")
+prat_abund <- plot(sm(gam_max_bb_abund_01, 10)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "prat")
+psit_abund <- plot(sm(gam_max_bb_abund_01, 11)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "psit")
+pyre_abund <- plot(sm(gam_max_bb_abund_01, 12)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "pyre")
+soro_abund <- plot(sm(gam_max_bb_abund_01, 13)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "soro")
+telu_abund <- plot(sm(gam_max_bb_abund_01, 14)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "telu")
+wurf_abund <- plot(sm(gam_max_bb_abund_01, 15)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "wurf")
+mgmt_bb_effect <- plot(pterm(gam_max_bb_abund_01, 1)) + l_fitPoints() + l_ciBar() + 
+  labs(x = NULL, y = NULL, title = "Management")
+
+## Arrange smooths on grid plot
+
+grid_bb_abund <- gridPrint(gers_abund, hort_abund, jone_abund, mend_abund, mont_abund, muci_abund, 
+                           pasc_abund, prat_abund, psit_abund, pyre_abund, 
+                           soro_abund, telu_abund, wurf_abund, mgmt_bb_effect, ncol = 4)
+
+pdf("../output/gam_bb_abund_grid.pdf", width = 8, height = 6)
+plot(grid_bb_abund)
+dev.off()
+
+png("../output/gam_bb_abund_grid.png", res = 300, width = 8, height = 6, units = "in")
+plot(grid_bb_abund)
+dev.off()
+
+# Extract predictions and plot on response scale
+
+gam_max_bb_abund_01_ilink <- family(gam_max_bb_abund_01)$linkinv # get inverse link function
+
+gam_max_bb_abund_01_pred <- predict.gam(gam_max_bb_abund_01, # get model predictions
+                                        type = "terms") %>% 
+  as.data.frame() %>% # convert to data frame
+  as_tibble() %>% # convert to tibble
+  dplyr::select(mgmt = 1, bb.sp = 2, bb.sp.year = 3, gers = 4, # rename
+         hort = 5, jone = 6, mend = 7, mont = 8, 
+         muci = 9, pasc = 10, prat = 11, psit = 12, 
+         pyre = 13, soro = 14, telu = 15, wurf = 16) %>%
+  mutate(fit = # get cumulative fit
+           mgmt + bb.sp + bb.sp.year + gers + 
+           hort + jone + mend + mont + 
+           muci + pasc + prat + psit + 
+           pyre + soro + telu + wurf 
+           ) %>%
+  dplyr::select(fit) %>%
+  mutate(fit.trans = gam_max_bb_abund_01_ilink(fit)) %>% # back-transform to response scale
+  bind_cols(abund_sum)
+
+ggplot(gam_max_bb_abund_01_pred, aes(elev.mean, fit.trans, linetype = year)) + # map year to linetype
+  geom_point(aes(elev.mean, max.bb.abund, shape = tree_line, color = management), 
+             inherit.aes = FALSE, alpha = 0.5, size = 2) + # add original data and map tree line to color
+  #geom_line(size = 0.5, alpha = 0.5) +
+  facet_wrap(~bb.sp, scales = "free_y") +
+  theme_light() +
+  labs(x = "Elevation (m.s.l.)", 
+       y = bquote("Peak bumble bee abundance"),
+       color = "Tree line",
+       shape = "Year",
+       linetype = "Year")
+
+ggsave("../output/gam_bb_abund.pdf")
+ggsave("../output/gam_bb_abund.png")
+```
+
+#### 2.2.4.2 FL
+```{r}
 # Plot on the link scale
-elev_effect <- plot(sm(gam_mean_fl_abund_00, 2)) + 
-  l_fitLine(alpha = 0.6, size = 1) + 
-  labs(x = "Elevation (m.s.l.)", y = "Conditional effect (link scale)", color = "Year")
 
-bb.sp_effect <- plot(sm(gam_mean_fl_abund_00, 1)) +
-  l_fitLine(alpha = 0.6, size = 1) + 
-  labs(x = "Elevation (m.s.l.)", y = NULL, color = "Species")
+## Individual plots
 
-grid_fl_abund <- gridPrint(elev_effect, bb.sp_effect, ncol = 2)
+gers_fl_abund <- plot(sm(gam_mean_fl_abund_01, 2)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "gers")
+hort_fl_abund <- plot(sm(gam_mean_fl_abund_01, 3)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "hort")
+jone_fl_abund <- plot(sm(gam_mean_fl_abund_01, 4)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "jone")
+mend_fl_abund <- plot(sm(gam_mean_fl_abund_01, 5)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "mend")
+mont_fl_abund <- plot(sm(gam_mean_fl_abund_01, 6)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "mont")
+muci_fl_abund <- plot(sm(gam_mean_fl_abund_01, 7)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "muci")
+pasc_fl_abund <- plot(sm(gam_mean_fl_abund_01, 8)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "pasc")
+prat_fl_abund <- plot(sm(gam_mean_fl_abund_01, 9)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "prat")
+psit_fl_abund <- plot(sm(gam_mean_fl_abund_01, 10)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "psit")
+pyre_fl_abund <- plot(sm(gam_mean_fl_abund_01, 11)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "pyre")
+soro_fl_abund <- plot(sm(gam_mean_fl_abund_01, 12)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "soro")
+telu_fl_abund <- plot(sm(gam_mean_fl_abund_01, 13)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "telu")
+wurf_fl_abund <- plot(sm(gam_mean_fl_abund_01, 14)) + l_fitLine() + l_ciLine() + 
+  labs(x = NULL, y = NULL, title = "wurf")
+mgmt_fl_effect <- plot(pterm(gam_mean_fl_abund_01, 1)) + l_fitPoints() + l_ciBar() + 
+  labs(x = NULL, y = NULL, title = "Management")
+year_fl_effect <- plot(pterm(gam_mean_fl_abund_01, 2)) + l_fitPoints() + l_ciBar() + 
+  labs(x = NULL, y = NULL, title = "Year")
 
-pdf("../output/gam_mean_fl_abund_grid.pdf", width = 7, height = 3.5)
+## Arrange significant smooths on grid plot
+
+grid_fl_abund <- gridPrint(gers_fl_abund, hort_fl_abund, jone_fl_abund, mend_fl_abund, mont_fl_abund, muci_fl_abund, 
+                           pasc_fl_abund, prat_fl_abund, psit_fl_abund, pyre_fl_abund, 
+                           soro_fl_abund, telu_fl_abund, wurf_fl_abund, year_fl_effect, mgmt_fl_effect, ncol = 4)
+
+pdf("../output/gam_fl_abund_grid.pdf", width = 8, height = 6)
 plot(grid_fl_abund)
@@
 
-png("../output/gam_mean_fl_abund_grid.png", res = 300, units = "in", width = 7, height = 3.5)
+png("../output/gam_fl_abund_grid.png", res = 300, width = 8, height = 6, units = "in")
 plot(grid_fl_abund)
 dev.off()
+
+# Extract predictions and plot on response scale
+
+gam_mean_fl_abund_01_ilink <- family(gam_mean_fl_abund_01)$linkinv # get inverse link function to transform to the response scale
+
+gam_mean_fl_abund_01_pred <- predict.gam(gam_mean_fl_abund_01, # get model predictions
+                                        type = "terms") %>% 
+  as.data.frame() %>% # convert to data frame
+  as_tibble() %>% # convert to tibble
+  dplyr::select(mgmt = 1, year = 2, bb.sp = 3, gers = 4, # rename
+         hort = 5, jone = 6, mend = 7, mont = 8, 
+         muci = 9, pasc = 10, prat = 11, psit = 12, 
+         pyre = 13, soro = 14, telu = 15, wurf = 16) %>%
+  mutate(fit = # get cumulative fit
+           mgmt + year + bb.sp + gers + 
+           hort + jone + mend + mont + 
+           muci + pasc + prat + psit + 
+           pyre + soro + telu + wurf 
+           ) %>%
+  dplyr::select(fit) %>%
+  mutate(fit.trans = gam_max_bb_abund_01_ilink(fit)) %>% # back-transform to response scale
+  bind_cols(na.omit(abund_sum)) # the na.omit() drops the rows where floral surveying data were missing
+
+ggplot(gam_mean_fl_abund_01_pred, aes(elev.mean, fit.trans, linetype = year)) + # map year to linetype
+  geom_point(aes(elev.mean, mean.fl.abund, color = management), inherit.aes = FALSE, alpha = 0.5) + # add original data and map tree line to color
+  #geom_line(size = 0.5, alpha = 0.5) +
+  facet_wrap(~bb.sp, scales = "free_y") +
+  theme_light() +
+  labs(x = "Elevation (m.s.l.)", 
+       y = bquote("Peak floral abundance"),
+       color = "Tree line",
+       shape = "Year",
+       linetype = "Year")
+
+ggsave("../output/gam_fl_abund.pdf")
+ggsave("../output/gam_fl_abund.png")
 ```

### 2021-04-06 code/ms1_beta_diversity.Rmd
added management covariate to models

@@
     summarize(samples = length(unique(date))) %>%
-    #dplyr::select(bb.sp, plant.sp.abb, site) %>% #
     left_join(site_data) %>% # add site data
-    dplyr::select(site, elev.mean, x, y, samples) %>%
+    ungroup() %>%
+    mutate(mowing = if_else(management == "mowing", 1, 0),
+           grasing = if_else(management == "grasing", 1, 0),
+           none = if_else(management == "none", 1, 0)) %>%
+    dplyr::select(site, elev.mean, x, y, samples, mowing, grasing, none) %>%
     distinct() # return unique site-dates
@@
   semi_join(net, by = "plant.sp") %>% # Consider only visited plant spp. (anytime, anywhere, not specifically for a given site*date)
-  dplyr::select(site, date, plant.sp, flower.cover) %>%
-  left_join(fl_traits)
+  dplyr::select(site, date, plant.sp, flower.cover) 
 ```
@@
   dplyr::select(site1, site2, FL, BB, # first for site 1
-                elev1 = elev.mean, x1 = x, y1 = y) %>%
+                elev1 = elev.mean, x1 = x, y1 = y, mgmt1 = management) %>%
   left_join(site_data, by = c("site2" = "site")) %>%
   dplyr::select(site1, site2, FL, BB, # then for site 2
-                elev1, elev2 = elev.mean, x1, y1, x2 = x, y2 = y) %>%
+                elev1, elev2 = elev.mean, x1, y1, x2 = x, y2 = y, mgmt1, mgmt2 = management) %>%
   mutate(site1 = factor(site1), 
@@
          elev.diff = elev1 - elev2,
-         mean.elev = (elev1 + elev2)/2) %>%
+         mean.elev = (elev1 + elev2)/2, mgmt.diff = if_else(mgmt1 == mgmt2, 0, 1)) %>%
   pivot_longer(cols = c(BB, FL), names_to = "group", values_to = "beta") %>%
@@
                 ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
-                elev1 = elev.mean, x1 = x, y1 = y) %>%
+                elev1 = elev.mean, x1 = x, y1 = y, mgmt1 = management) %>%
   left_join(site_data, by = c("site2" = "site")) %>%
@@
                 ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
-                elev1, elev2 = elev.mean, x1, y1, x2 = x, y2 = y) %>%
+                elev1, elev2 = elev.mean, x1, y1, x2 = x, y2 = y, mgmt1, mgmt2 = management) %>%
   mutate(elev.diff = abs(elev1 - elev2),
-         mean.elev = (elev1 + elev2)/2) %>%
+         mean.elev = (elev1 + elev2)/2,
+         mgmt.diff = if_else(mgmt1 == mgmt2, 0, 1)) %>%
   gather(metric, value, -c(site1, site2, elev1, elev2, 
                            elev.diff, mean.elev, 
-                           x1, y1, x2, y2)) %>%
+                           x1, y1, x2, y2, mgmt1, mgmt2, mgmt.diff)) %>%
   mutate(metric.class = case_when(
@@
 # Set palette
-# p <- c("OS" = "darkorange1", "ST" = "green3", 
-#        "ST.h" = "firebrick1", "ST.l" = "darkorchid", 
-#        "ST.lh" = "deepskyblue", "WN" = "black")
-# 
-# 
-# p <- c("OS" = "#66c2a5", "ST" = "#fc8d62", 
-#        "ST.h" = "#8da0cb", "ST.l" = "#e78ac3", 
-#        "ST.lh" = "#a6d854", "WN" = "black")
-
 p1 <- c("OS" = "#e7298a", "ST" = "#66a61e", "WN" = "black")
@@
   summarize(samples = length(unique(date))) %>%
+  remove_rownames() %>%
   column_to_rownames("site") %>%
@@
   left_join(survey_sample_diff) %>%
-  dplyr::select(site1, site2, beta, elev.diff, sample.diff) %>%
+  dplyr::select(site1, site2, beta, elev.diff, sample.diff, mgmt.diff) %>%
   mutate(elev.diff = scale(elev.diff), # gotta scale variables for MRM
@@
   left_join(net_sample_diff)  %>%
-  dplyr::select(site1, site2, beta, elev.diff, sample.diff) %>%
+  dplyr::select(site1, site2, beta, elev.diff, sample.diff, mgmt.diff) %>%
   mutate(elev.diff = scale(elev.diff),
@@
   left_join(net_sample_diff)  %>%
-  dplyr::select(site1, site2, value, metric, elev.diff, sample.diff) %>%
+  dplyr::select(site1, site2, value, metric, elev.diff, sample.diff, mgmt.diff) %>%
   mutate(elev.diff = scale(elev.diff),
@@
 
-mrm_fl <- MRM(beta ~ elev.diff + sample.diff, 
+mrm_fl <- MRM(beta ~ elev.diff + sample.diff + mgmt.diff, 
               data = mrm_survey_data,
@@
 
-mrm_bb <- MRM(beta ~ elev.diff + sample.diff, 
+mrm_bb <- MRM(beta ~ elev.diff + sample.diff + mgmt.diff, 
               data = mrm_net_data,
@@
 
-mrm_WN <- MRM(value ~ elev.diff + sample.diff, 
+mrm_WN <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
               data = filter(mrm_beta_data, metric == "WN"),
@@
 
-mrm_ST <- MRM(value ~ elev.diff + sample.diff, 
+mrm_ST <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
               data = filter(mrm_beta_data, metric == "ST"),
@@
 
-mrm_ST.l <- MRM(value ~ elev.diff + sample.diff, 
+mrm_ST.l <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
               data = filter(mrm_beta_data, metric == "ST.l"),
@@
 
-mrm_ST.h <- MRM(value ~ elev.diff + sample.diff, 
+mrm_ST.h <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
               data = filter(mrm_beta_data, metric == "ST.h"),
@@
 
-mrm_ST.lh <- MRM(value ~ elev.diff + sample.diff, 
+mrm_ST.lh <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
               data = filter(mrm_beta_data, metric == "ST.lh"),
@@
 
-mrm_OS <- MRM(value ~ elev.diff + sample.diff, 
+mrm_OS <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
               data = filter(mrm_beta_data, metric == "OS"),
@@
 # Run GDM analysis
-gdm_WN <- gdm(sitepair.WN, geo = TRUE)
+gdm_WN <- gdm(sitepair.WN, geo = FALSE)
 summary(gdm_WN)
+gdm.varImp_WN <- gdm.varImp(sitepair.WN, geo = FALSE)
 

### 2021-04-06 code/ms2_bias_niche.Rmd
added management covariate to models

@@
 
+ggplot(niche_gam_21_pred, aes(elev.mean, mean.niche.overlap_trans)) +
+  geom_ribbon(aes(elev.mean, mean.niche.overlap_trans, 
+                  ymax = CI_upper_trans, ymin = CI_lower_trans),
+              alpha = 0.3, inherit.aes = FALSE) +
+  geom_line() +
+  ylab("Niche overlap") +
+  xlab("Elevation (m)") +
+  guides(fill = FALSE) +
+  facet_wrap(~tongue.pair) +
+  theme_light()
+
 ggsave(file = "../output/niche_gam_21_plot1.pdf", width = 5, height = 4)
@@
 # Plot intercept effect of tongue.diff
-plot(pterm(niche_gam_21, 1))
-```
+pdf("../output/niche_gam_21_plot2.pdf", width = 5, height = 4)
+plot(pterm(niche_gam_21, 1)) + labs(x = "Tongue pair", y = "Niche overlap") + theme_light()
+dev.off()
+```

### 2021-04-06 ms_1/ms_1.Rmd
added management covariate to models

--- 2021-03-17 ms_1/ms_1.Rmd
@@
 ---
-title: "The tree line ecotone punctuates the elevational turnover of plants, pollinators, and their interactions "
+title: "The tree line ecotone punctuates the elevational turnover of plants, pollinators, and their interactions"
 
-author: Douglas B. Sponsler^1^*, Fabrice Requier^2^, Katharina Kallnik^1^, Alice Claßen^1^, A. Fabienne Maihoff^1^, Ingolf Steffan-Dewenter^1^
+author: Douglas B. Sponsler^1^*, Fabrice Requier^2^, Katharina Kallnik^1^, Alice Claßen^1^, A. Fabienne Maihoff^1^, Johanna Sieger^3^, Ingolf Steffan-Dewenter^1^
 
@@
 
+^3^ Naturpark Frankenhöhe, Ansbach, Germany
+
 `*` Corresponding author: douglas.sponsler@uni-wuerzburg.de
+
+# Open Research Statement
+
+Data are not yet provided but will be permanently archived on Dryad if the paper is accepted for publication.
+This submission uses novel code, which is provided as Supporting Information to be evaluated as part of the submission.
+If the paper is accepted, the code will be archived on Dryad alongside our raw data.
 
@@
 
-The structuring of biological communities along mountain slopes is complex, and the elevational range shifts that are now occurring in response to climate change involve more than merely tracking suitable temperature envelopes.
+The structuring of biological communities along mountain slopes is complex, and elevational range shifts in response to climate change involve more than merely tracking suitable temperature envelopes.
+Bumble bees (Hymentopera: *Bombus* spp.) are cold-adapted species associated with mountain habitats, and they are already exhibiting upslope range shifts that are expected to result in habitat loss, novel competitive interactions, and the rewiring of pollination networks.
+Interpreting these shifts, however, requires an understanding of the current elevational patterns of bumble bees and the floral mutualists that are being acted upon by climate change.
+We recorded bumble- bee-flower interactions over three years along an 1400 m elevational gradient in the German Alps.
+Using nonlinear modeling, we analyze the elevational patterns at the nested levels of species abundance, species β-diversity, and interaction β-diversity.
+We demonstrate that the tree line ecotone is (1) a distributional interface between low/mid- and high-elevation bumble bee species, (2) a threshold above which floral resource availability sharply decreases, and (3) a zone of accelerated turnover of floral composition and bumble- bee-flower interactions.
+Our results show that linear elevational temperature gradients are ecologically punctuated, and the outcomes of climate-induced range shifts may hinge upon dynamics at the tree line ecotone.
+
+<!-- The structuring of biological communities along mountain slopes is complex, and the elevational range shifts that are now occurring in response to climate change involve more than merely tracking suitable temperature envelopes.
 When species move, they do so in the context of biological communities, and the outcomes of these movements depend on how and to what extent biotic interactions are reordered.
@@
 The implications of these findings extend beyond the particular case of bumble bees to demonstrate that linear elevational temperature gradients are ecologically punctuated, and the outcomes of climate-induced range shifts with depend on how species and species-interactions change as they cross --- or fail to cross --- thresholds like the tree line ecotone.
+-->
 
-*Keywords*: climate, elevation gradient, pollinator, *Bombus*, mountain, tree line
+*Keywords*: climate, beta-diversity, elevation gradient, *Bombus*, bumble bee, mountain ecology, pollination network
 
@@
 
-<!-- This is a paper about mountains. -->
-
 Mountains are unique ecological theaters in which extrinsic climate dynamics interact with intrinsic climate gradation to shape the distributions of species and biotic interactions [@Telwala2013-sc; @CaraDonna2014-nq; @Miller-Struttmann2014-wz; @Rafferty2020-aj].
@@
 
-The outcomes of range shifts in mountain bumble bee and floral communities have obvious significance from a conservation perspective [@Soroye2020-zb], but they also provide a model system in which to study the basic phenomenon of community assembly under climate change, a phenomenon of both historical interest and prospective urgency.
+The outcomes of range shifts in mountain bumble bee and floral communities have obvious significance from a conservation perspective [@Soroye2020-zb], but they also provide a model system in which to study the basic process of community assembly under climate change, a phenomenon of both historical interest and prospective urgency.
 In the present study, set along ~1400 m elevation gradient in the German Alps, we investigate elevational patterns in a community of bumble bees and wildflowers at three levels of biological organization.
@@
 Fourteen of these pastures are extensively grazed by cattle or sheep, 3 are mowed for hay production, and 8 have been abandoned throughout the last century and are no longer subject to any human management.
-Study sites were classified as lying either above or below the tree line based on an examination of Google Earth imagery and corroborated by the authors with field experience at our study sites (K.
-Kallnik, A.
-F.
-Maihoff, A.
-Claßen).
+Study sites were classified as lying either above or below the tree line based on an examination of Google Earth imagery corroborated by field experience at our study sites.
 Eight sites were located above the tree line, which generally fell at an elevation of around 1500 m.s.l., consistent with the previous descriptions of the tree line in the Berchtesgadener Alps [@Kostler1970-ya; @Mayer1970-zn].
@@
 Bumble bee abundance was quantified as the total number of recorded floral visits per bumble bee species per site-date.
-To represent floral abundance from the perspective of each bumble bee species in our network, we first scored each plant genus in our data set as either visited or not visited by each bumble bee species.
-We then calculated the total floral resource availability for each bumble bee species for each site-date by summing the observed flower cover of visited genera.
+To weight floral abundance based on the foraging preferences of each bumble bee species in our network, we scored each plant genus in our data set as either visited or not visited by each bumble bee species.
+We then calculated the total floral resource availability for each bumble bee species for each site-date by summing the observed flower cover of *visited* genera.
 We opted to work at the genus level based on the reasoning that if a bumble bee species visits one member of a given genus, other members of the same genus should also be considered potential floral hosts.
@@
 We also opted to treat the cryptic *B.
-terrestris/lucorum* species pair as a single morphospecies, since these two species cannot be visually distinguished reliably, and we similarly pooled representatives of the parasitic *Psithyrus* subgenus --- *B.
+terrestris/lucorum* species pair as a single morphospecies, since these two species cannot be visually distinguished, and we similarly pooled representatives of the parasitic *Psithyrus* subgenus --- *B.
 barbutellus*, *B.
@@
 
-We analyzed the abundance of bumble bees and floral resources using separate hierarchical generalized additive models (HGAMs) fit with the Tweedie distribution family (log link function) [@Wood2011-pg; @Pedersen2019-kf].
-We opted in both models to collapse the temporal variation of abundance within sites and focus on elevational patterns.
-For bumble bees, we modeled *peak* abundance at each site as a smooth response to elevation, with bumble bee species as a smoothing factor and a random intercept term.
-We also included species:year as a random intercept term because exploratory analysis indicated that species varied differently across years.
-For floral resources, we modeled *mean* flower cover, weighted per bumble bee species, as a smooth response to elevation, with year and bumble bee species as random intercept terms.
-Both models are follow the "type I" model form described in @Pedersen2019-kf.
+We analyzed the abundance of bumble bees and floral resources using separate hierarchical generalized additive models (HGAMs) [@Wood2011-pg; @Pedersen2019-kf].
+In both models, we opted to collapse the temporal variation of abundance within sites and focus on elevational patterns.
+For bumble bees, we modeled *peak* abundance at each site as a smooth response to elevation, with bumble bee species as a smoothing factor, management (grazing, mowing, or unmanaged) as a fixed effect, and species:year as a random intercept term (exploratory analysis indicated that species varied differently across years).
+For floral resources, we modeled *mean* flower cover, weighted per bumble bee species (i.e.
+the sum of *visited* genera), as a smooth response to elevation, with management as a fixed effect, and year and bumble bee species as random intercept terms.
+For the bumble bee model, we used the negative binomial distribution family (log link function), and for the floral model we used a Tweedie family distribution (log link function).
+Both models follow the "type I" model form described in @Pedersen2019-kf.
 
@@
 For our analyses of species and interaction β-diversity, we again opted to focus on elevational patterns by pooling samples within sites, since the sparse observations of individual sampling events tended to inflate β-diversity and its variance to a degree that impaired model fitting and interpretation.
-After pooling, we calculated the species and interaction β-diversity between all pairs of sites and partitioned total interaction β-diversity into its components of species turnover (changes in species composition) and interaction rewiring (identical species interacting in different ways) [@Novotny2009-my; @Poisot2012-fk].
+After pooling, we calculated the species and interaction β-diversity between all pairs of sites and partitioned total interaction β-diversity into its components of species turnover (changes in species composition) and interaction rewiring (interaction turnover despite species constancy) [@Novotny2009-my; @Poisot2012-fk].
 As an initial analysis, we plotted the relationship between site-wise elevation difference and each metric of β-diversity and verified the significance of the relationship using logistic matrix regression [@Goslee2009-ln], with the difference in number of sampling dates between sites as a covariate to control for potential confounding effects of sampling frequency.
@@
 To control for potential confounding effects of geographic proximity and number of sampling dates per site, these terms were added as covariates in the GDM.
+GDM is designed to model continuous predictors, and while the inclusion of categorical variables is theoretically possible [@Ferrier2007-ji],
 
@@
 lapidarius* were recorded only sporadically and at low abundance, so we omitted them from all analyses.
+See Appendix X for a summary of species-wise abundances.
 
@@
 
-```{r fig2, echo = FALSE,  fig.cap = paste("Proportional abundances (cumulative within site) of bumble bee species in our study system, ordered by mean proportional abundance.")}
-knitr::include_graphics("../output/bb_abundsum.png")
-```
-
 ## Elevational distribution of bumble bees and floral resources
 
-The peak abundance of all bumble bee species responded significantly to elevation (p ≤ 0.04) **(Figures 3, 4)**.
+The peak abundance of all bumble bee species responded significantly to elevation (p ≤ 0.04) **(Figure 2)**.
 *B.
@@
 
-```{r fig3, echo = FALSE,  fig.cap = paste("Fitted GAM smooths of bumble bee abundance on elevation, faceted by bumble bee species and plotted over original data.
+<!-- ```{r fig3, echo = FALSE,  fig.cap = paste("Fitted GAM smooths of bumble bee abundance on elevation, faceted by bumble bee species and plotted over original data.
 Data points are color-coded to indicate whether they represent sites above (red) or below (blue) the tree line.
-Note the different scales of the y-axes.")}
-knitr::include_graphics("../output/gam_bb_abund.png")
-```
+Note the different scales of the y-axes.")} -->
+<!-- knitr::include_graphics("../output/gam_bb_abund.png") -->
+<!-- ``` -->
 
-```{r fig4, echo = FALSE,  fig.cap = paste("Conditional effect of elevation on the peak abundance of each bumble bee species.
+```{r fig2, echo = FALSE,  fig.cap = paste("Conditional effect of elevation on the peak abundance of each bumble bee species.
 Effects are plotted on the link scale, and dashed lines depict 95\\% confidence intervals.")}
@@
 
-Mean floral resource abundance likewise responded significantly (p << 0.0001) to elevation for all bumble bee species **(Figures 5, 6)**.
+Mean floral resource abundance likewise responded significantly (p << 0.0001) to elevation for all bumble bee species **(Figure 3)**.
 *B.
@@
 pyrenaeus* exhibited a unimodal pattern peaking near 1600 m.s.l.
-The remaning species --- *B.
+The remaining species --- *B.
 gerstaeckeri*, *B.
@@
 
-```{r fig5, echo = FALSE,  fig.cap = paste("Fitted GAM smooths of floral abundance on elevation, faceted by bumble bee species and plotted over original data.
+# ```{r fig5, echo = FALSE,  fig.cap = paste("Fitted GAM smooths of floral abundance on elevation, faceted by bumble bee species and plotted over original data.
 Data points are color-coded to indicate whether they represent sites above (red) or below (blue) the tree line.
 Note the different scales of the y-axes.")}
-knitr::include_graphics("../output/gam_fl_abund.png")
-```
+# knitr::include_graphics("../output/gam_fl_abund.png")
+# ```
 
@@
 
-Floral β-diversity among our study sites was very high overall and responded steeply to elevation difference between sites **(Figure 3A)**.
+Floral β-diversity among our study sites was very high overall and responded steeply to elevation difference between sites **(Figure 4A)**.
 Even sites at similar elevation exhibited ~40% species turnover, and the most widely separated sites (elevation difference > 1250 m) differed by more than 85%.
 
-Bumble bee β-diversity was, in comparison to floral β-diversity, both lower overall and less responsive to elevational difference between sites **(Figure 3A)**.
+Bumble bee β-diversity was, in comparison to floral β-diversity, both lower overall and less responsive to elevational difference between sites **(Figure 4A)**.
 Sites at similar elevation exhibited ~25% species turnover, and species turnover between the most widely separated sites remained less than 50%.
 
-Total interaction β-diversity was >75% between sites at similar elevation and approached perfect dissimilarity in the most widely separated sites **(Figure 3B)**.
+Total interaction β-diversity was >75% between sites at similar elevation and approached perfect dissimilarity in the most widely separated sites **(Figure 4B)**.
 Partitioning revealed that total interaction β-diversity was driven primarily by species turnover, which accounted for ~50% of total β-diversity for sites at similar elevation and ~90% of total β-diversity for the most widely separated sites.
-Species turnover consisted mainly of turnover in the floral community **(Figure 3C)**, though the joint turnover of plants and bumble bees accounted for ~25% of total species turnover between the most widely separated sites.
-Interaction rewiring was most significant between sites at similar elevation, where it accounted for ~30% of total β-diversity, but its share of total β-diversity declined to ~10% in the most widely separated sites **(Figure 3B)**.
+Species turnover consisted mainly of turnover in the floral community **(Figure 4C)**, though the joint turnover of plants and bumble bees accounted for ~25% of total species turnover between the most widely separated sites.
+Interaction rewiring was most significant between sites at similar elevation, where it accounted for ~30% of total β-diversity, but its share of total β-diversity declined to ~10% in the most widely separated sites **(Figure 4B)**.
 
-Logistic matrix regression confirmed the significance (p < 0.01) of the relationship between each β-diversity metric and elevation **(Table 1)**.
+Logistic matrix regression confirmed the significance (p < 0.01) of the relationship between each β-diversity metric and elevation.
 
-```{r fig7, echo = FALSE,  fig.cap = paste("Species-level (A) and interaction-level (B, C) plotted against elevational difference.
+```{r fig4, echo = FALSE,  fig.cap = paste("Species-level (A) and interaction-level (B, C) plotted against elevational difference.
 Each point represents the β-diversity (or β-diversity partition) between a pair of sites, and lines represent the overall relationship between β-diversity and elevation difference using binomial regression smooths.

### 2021-04-06 ms_2/ms_2_v2.Rmd
added management covariate to models

@@
 
-author: Douglas B. Sponsler^1^*, Fabrice Requier^2^, Katharina Kallnik^1^, Alice Claßen^1^, A. Fabienne Maihoff^1^, Ingolf Steffan-Dewenter^1^
+author: Douglas B. Sponsler^1^*, Katharina Kallnik^1^, Fabrice Requier^2^, Alice Claßen^1^, A. Fabienne Maihoff^1^, Johanna Sieger^3^, Ingolf Steffan-Dewenter^1^
 
@@
 
+^3^ Naturpark Frankenhöhe, Ansbach, Germany
+
 `*` Corresponding author: douglas.sponsler@uni-wuerzburg.de
@@
 
-Species coexistence is constrained in part by competition over shared food sources.
-Bumble bees are classic model organisms for the study of competition and coexistence, and this topic has taken on new urgency as climate-induced range shifts are expected to reshuffle competitive relationships between bumble bee species, particularly along the climatic gradients of the mountain slopes that host the world's richest bumble bee communities.
+Species coexistence is constrained by competition over shared food sources.
+Bumble bees are classic model organisms for the study of competition and coexistence, and this topic has taken on new urgency as climate and land use change threaten to reshuffle competitive relationships between bumble bee species along the mountain slopes that host the world's richest bumble bee communities.
 We surveyed floral abundance and recorded bumble bee floral visitation patterns along a ~1400 m elevation gradient in the German Alps.
@@
 We then investigate the patterns of diet overlap between bumble bee species at the nested levels of floral morphotype, floral family, and floral genus, and model diet overlap as a function bumble bee tongue length and elevation.
-We find that while tongue length is a significant predictor of bumble bees' floral morphotype preferences, species with similar tongue lengths exhibit divergent preferences with respect to certain flora morphotypes.
-This indicates that factors other than tongue length are involved in flower choice, which might partly explain the puzzling coexistence of bumble bee species with similar tongue lengths.
+We find that while tongue length is a significant predictor of bumble bees' floral morphotype preferences, there exists both convergence in floral preference across bumble bees with different tongue lengths and divergence of floral preference among bumble bees with similar tongue lengths.
+This indicates that factors other than tongue length are involved in flower choice, which might partly explain the coexistence of bumble bee species with similar tongue lengths.
 Moreover, genus-level diet overlap was approximately one half that of morphotype-level diet overlap, indicating that even when bumble bees exhibit similar biases at the level of floral morphotype, partly due to the constraints of tongue length, floral diversity enables the mitigation of competition via the subpartitioning within floral morphotypes.
-This suggests that floral species richness may be as important as overall floral abundance in avoiding the competitive exclusion of bumble bee species.
-Diet overlap was highest at the low end of our elevation gradient, declining sharply over the transition from valley floor to mid-slope, then leveling off or declining weakly for the remainder of the gradient.
-If diet overlap is interpreted as an index of potential competition, upslope shifts in bumble bees and flora could generate more competitive conditions at high elevations in cases where floral resources are limiting.
+Diet overlap was highest at the low end of our elevation gradient, declining sharply over the transition to mid-slope, then leveling off or declining weakly for the remainder of the gradient.
+Resource simplification at low elevations combined with warming climates could cause increasing spillover of low-elevation bumble bee species into high-elevation sites, introducing novel competition potential likely to be exacerbated by the reforestation of sub-alpine meadows.
 
@@
 
-# INTRODUCTION
+# 1. INTRODUCTION
 
@@
 
-# METHODS
+# 2. METHODS
 
-## Field system
+## 2.1 Field system
 
@@
 
-## Data sources
+## 2.2 Data sources
 
-### Floral morphology
+### 2.2.1 Floral morphology
 
@@
 
-### Bumble bee tongue length
+### 2.2.2 Bumble bee tongue length
 
@@
 
-## Data processing
+## 2.3 Data processing
 
@@
 
-## Statistical models
+## 2.4 Statistical models
 
@@
 
-### Model 1: Comparing overall visitation probability across floral morphotypes
+### 2.4.1 Model 1: Comparing overall visitation probability across floral morphotypes
 
@@
 
-### Model 2: Visitation probability as a function of floral morphotype and bumble bee tongue length
+### 2.4.2 Model 2: Visitation probability as a function of floral morphotype and bumble bee tongue length
 
@@
 
-### Model 3: Visitation probability as a function of floral morphotype and bumble bee species
+### 2.4.3 Model 3: Visitation probability as a function of floral morphotype and bumble bee species
 
@@
 
-### Model 4: Aggregate diet ordination
+### 2.4.4 Model 4: Aggregate diet ordination
 
@@
 
-### Model 5: Niche overlap by level and elevation
+### 2.4.5 Model 5: Niche overlap by level and elevation
 
@@
 
-### Model 6: Genus-level niche overlap by tongue length and elevation
+### 2.4.6 Model 6: Genus-level niche overlap by tongue length and elevation
 
@@
 
-# RESULTS
+# 3. RESULTS
 
-## Summary of visitation and floral survey data
+## 3.1 Summary of visitation and floral survey data
 
@@
 
-## Floral preference (Models 1-3)
+## 3.2 Floral preference (Models 1-3)
 
@@
 
-## Aggregate diet similarity
+## 3.3 Aggregate diet similarity
 
@@
 
-## Site-level diet overlap (Models 5 and 6)
+## 3.4 Site-level diet overlap (Models 5 and 6)
 
@@
 
-We recommend caution, however, in interpreting the p-values of this model, since the pairwise comparison of group differences entails a degree of non-independence.
+The response curve of niche overlap to elevation was significantly different (p < 0.01) from a flat line for *short:short*, *medium:long*, and *long:long*, but not for *short:long*, *short:medium*, or *medium:medium* **(Figure 6A)**.
+Niche overlap among short-tongued species (*short:short*) declined linearly with elevation, while niche overlap between medium- and long-tongued species had a u-shaped response to elevation, with maxima at the extreme of the elevation gradient.
+Niche overlap between long-tongued species exhibited a w-shaped response with peaks at low-, mid-, and high-elevation, though this erratic pattern is likely driven by small number of long-tongued species (*B.
+gerstaeckeri*.
+*B.
+hortorum*, and *B.
+mendax*) available for comparison.
+Compared to the reference intercept level of *short:long*, mean niche overlap was significantly higher for *medium:long* (p = 0.002) and *short:short* comparisons (p = 0.02) **(Figure 6B)**.
+Model 6 explained 43.5% of total deviance (adjusted R^2^ = 0.24).
+We recommend caution in interpreting the p-values of this model, since the pairwise comparison of differences entails a degree of non-independence not accounted for in the structure of the model.
 
@@
 
-```{r fig6, echo = FALSE,  fig.cap = paste("Niche overlap by tongue length class difference through elevation. Lines show GAM smooths and shaded bands show 95\\% confidence intervals.")}
-knitr::include_graphics("../output/niche_gam_21_plot1.png")
+```{r fig6, echo = FALSE,  fig.cap = paste("Niche overlap by tongue length pair and elevation.
+Lines show GAM smooths of niche overlap through elevation, and shaded bands show 95\\% confidence intervals (A).
+Dot-whisker plots show 95\\% confidence intervals of intercepts (i.e.
+mean niche overlap) for each tongue length pair (B).")}
+knitr::include_graphics("../output/niche_gam_21.png")
 ```
 
-# DISCUSSION
-
-## Floral preference is only partly explained by the relationship between bumble bee tongue length and floral morphotype
+# 4. DISCUSSION
 
@@
 
-## The influence of tongue length on floral preference does not carry on to genus-level niche overlap, and floral diversity within morphotypes relaxes potential competition between bumble bee species of similar tongue length
-
 While tongue length was a relatively strong (though imperfect) predictor of bumble bee species' morphotype preferences, it was a weak predictor of diet similarity (at the level of floral genus) across bumble bees species, echoing the findnings of @Durieux2000-uk.
@@
 Nevertheless, the ordination explained less than 50% of total variance, and while short-tongued species separated clearly from long- and medium-tongued species, the latter overlapped extensively with each other.
-These results were corroborated by our HGAM analysis of niche overlap, where the only significant in overall niche overlap was between short- and long-tongued species.
+These results were corroborated by our HGAM analysis of niche overlap, where overall niche overlap was lowest between short- and long-tongued species and significantly higher among short-tongued species and between medium- and long-tongued species.
 A caveat to these analyses, though, is that our bumble bee community included only 3 long-tongued species (*B.
@@
 
-Community-mean diet overlap at the level of floral morphotype was consistently nearly twice as high as diet overlap at the level of floral genus.
-From a numerical perspective, this result is unsurprising simply because each morphotype comprises many genera.
+From the perspective of interspecific competition in bumble bees, it is significant that community-mean diet overlap at the level of floral morphotype was consistently nearly twice as high as diet overlap at the level of floral genus.
+From a purely numerical perspective, this result is unsurprising simply because each morphotype comprises many genera.
 From an ecological perspective, though, this pattern suggests another answer to the question of how bumble bee species of similar tongue length can coexists: if tongue length primarily constrains bumble bee foraging at the coarse level of floral morphotype, floral diversity *within* morphotypes creates space for bumble bees with similar tongue lengths to avoid actual dietary overlap, thus mitigating interspecific competition.
 
-## Climate, land use, and the future of bumble bee niche relations
+Community mean diet overlap was highest at the lowest sites and declined sharply over the transition to mid-slope.
+If this pattern of niche overlap moves upslope in response to climate change, it might result in increased competition potential at higher elevations, which could threaten high-elevation bumble bee species that currently exist under conditions of relatively low niche overlap.
+In our system, though, it must be noted that the lowest elevation sites are also the ones most influenced by human land use.
+The elevated niche overlap we observed at the lowest sites could be interpreted as an indication that the deleterious effects of land use intensification at low elevation have simply outpaced those of agricultural abandonment at high-elevation, and mountain bumble bees populations will initially experience resource pressure from below as reforestation slowly closes high-elevation meadows.
+If the elevated niche overlap we observed at low elevation is to be interepreted as an indication of resource limitation, though, it must be noted that overall resource volume is stable in our study system up to the tree line ecotone at around 1500 m.s.l.
+(Sponsler et al., submitted); if bumble bee species are forced to converge on similar diets at low elevation sites, it is more likely because the floral communities at these sites have been *simplified* in terms of taxonomic/morphological diversity or spatiotemporal structure of availability.
 
-Community mean diet overlap was highest at the lowest sites and declined sharply over the transition from valley floor to mid-slope.
-If this pattern of niche overlap moves upslope in response to climate change, it might result in increased competition potential at higher elevations, which could threaten high-elevation bumble bee species that currently exist under conditions of relatively low niche overlap.
-In our system, though, it must be noted that the lowest elevation sites are also the ones most influenced by human land use, and it is possible that the elevation pattern evident in our data is driven by factors other than elevational climate variation.
+<!-- I could, of course, test this. But I don't want to right now. -->
 
 ## Conclusion
+
 Coexistence between bumble bee species with similar tongue lengths hinges on the relationships between species-level floral preferences, trait-matching, and the emergent phenomenon of diet similarity.
 In this study, we clarify these relationships with respect to bumble bees and wild flowers interacting along a mountain slope elevation gradient.
-In addition to addressing the question of bumble bee species coexistence, our results suggest how patterns of competition may respond to climate and land use change along mountain slopes, an issue of immediate concern for the conservation of bumble bees and their floral mutualists [@Sirois-Delisle2018-vi; @Soroye2020-zb].
+Our results indicate that bumble bee coexistence is favored by the fact that the morphological association between tongue length and floral morphotype is a soft constraint on floral preference, and there exists both convergence in floral preference across bumble bees with different tongue lengths and divergence of floral preference among bumble bees with similar tongue lengths.
+Moreover, we find that niche overlap at the level of floral genus is much lower than at the level of floral morphotype, suggesting that the partitioning of resources *within* floral morphotypes functions to mitigate potential competition.
+This means that diversity *within* floral morphotypes may be as important as overall floral abundance or morphotype-level diversity in facilitating the coexistence of bumble bee species, an insight with practical implications for floral resource management aimed at bumble bee conservation.
 
-This suggests that diversity within floral morphotypes may as important as morphotype diversity *per se* and overall floral abundance in facilitating the coexistence of bumble bee species, an insight which may have implications for the design of floral enhancements aimed at bumble bee conservation.
+The future of mountain bumble bees --- and, indeed, of mountain ecosystems in general --- involves the systematic pressure of climate warming as well as the twin land use pressures of agricultural intensification at low elevations and agricultural abandonment (with subsequent reforestation) at high elevations [@Iserbyt2015-jc; @Walcher2017-kw; @Sirois-Delisle2018-vi; @Soroye2020-zb; @Suzuki-Ohno2020-jv].
+The outcomes of these processes with respect to competition and coexistence between bumble bee species are not yet clear, but our finding that interspecific niche overlap (interpreted as an index of potential competiion) is highest at low elevations suggests that land use intensification may already be influencing bumble bee niche relationships.
+Resource simplification at low elevations combined with warming climates along mountain slopes could cause increasing spillover of low-elevation bumble bee species into the historical ranges of high elevation species, introducing novel competition potential likely to be exacerbated by the gradual reforestation of abandoned sub-alpine meadows.
 

### 2021-04-07 code/ms1_abundance.Rmd
updated analyses and figures ot include mamagement; dropped MRM analysis

@@
 # BB
-# gam_max_bb_abund_00 <- gam(max.bb.abund ~ 
-#                              s(bb.sp, bs = "re") + # random intercept for bb.sp
-#                              s(bb.sp.year, bs = "re") + # random intercept for bb.sp:year
-#                              s(elev.mean, by = bb.sp, k = 8), # smoother for elev by bb.sp.
-#                            method = "REML",
-#                            family = "tw", # Tweedie distribution performs well
-#                            select = TRUE,
-#                            data = abund_sum) %>% getViz()
-# 
-# # FL
-# gam_mean_fl_abund_00 <- gam(mean.fl.abund ~
-#                               year +
-#                               s(bb.sp, bs = "re") +
-#                               s(elev.mean, by = bb.sp, k = 8),
-#                            method = "REML",
-#                            family = "tw",
-#                            select = TRUE,
-#                            data = abund_sum) %>% getViz()
-
-# BB
 gam_max_bb_abund_01 <- gam(max.bb.abund ~ 
@@
 ```{r}
-# Plot smooths and year effect on the link scale
-
 ## Individual plots
 gers_abund <- plot(sm(gam_max_bb_abund_01, 3)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "gers")
+  labs(x = NULL, y = NULL, title = "gers") +
+  l_rug(length = unit(0.15, "cm"))
 hort_abund <- plot(sm(gam_max_bb_abund_01, 4)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "hort")
+  labs(x = NULL, y = NULL, title = "hort") +
+  l_rug(length = unit(0.15, "cm"))
 jone_abund <- plot(sm(gam_max_bb_abund_01, 5)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "jone")
+  labs(x = NULL, y = NULL, title = "jone") +
+  l_rug(length = unit(0.15, "cm"))
 mend_abund <- plot(sm(gam_max_bb_abund_01, 6)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "mend")
+  labs(x = NULL, y = NULL, title = "mend") +
+  l_rug(length = unit(0.15, "cm"))
 mont_abund <- plot(sm(gam_max_bb_abund_01, 7)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "mont")
+  labs(x = NULL, y = NULL, title = "mont") +
+  l_rug(length = unit(0.15, "cm"))
 muci_abund <- plot(sm(gam_max_bb_abund_01, 8)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "muci")
+  labs(x = NULL, y = NULL, title = "muci") +
+  l_rug(length = unit(0.15, "cm"))
 pasc_abund <- plot(sm(gam_max_bb_abund_01, 9)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "pasc")
+  labs(x = NULL, y = NULL, title = "pasc") +
+  l_rug(length = unit(0.15, "cm"))
 prat_abund <- plot(sm(gam_max_bb_abund_01, 10)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "prat")
+  labs(x = NULL, y = NULL, title = "prat") +
+  l_rug(length = unit(0.15, "cm"))
 psit_abund <- plot(sm(gam_max_bb_abund_01, 11)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "psit")
+  labs(x = NULL, y = NULL, title = "psit") +
+  l_rug(length = unit(0.15, "cm"))
 pyre_abund <- plot(sm(gam_max_bb_abund_01, 12)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "pyre")
+  labs(x = NULL, y = NULL, title = "pyre") +
+  l_rug(length = unit(0.15, "cm"))
 soro_abund <- plot(sm(gam_max_bb_abund_01, 13)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "soro")
+  labs(x = NULL, y = NULL, title = "soro") +
+  l_rug(length = unit(0.15, "cm"))
 telu_abund <- plot(sm(gam_max_bb_abund_01, 14)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "telu")
+  labs(x = NULL, y = NULL, title = "telu") +
+  l_rug(length = unit(0.15, "cm"))
 wurf_abund <- plot(sm(gam_max_bb_abund_01, 15)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "wurf")
+  labs(x = NULL, y = NULL, title = "wurf") +
+  l_rug(length = unit(0.15, "cm"))
 mgmt_bb_effect <- plot(pterm(gam_max_bb_abund_01, 1)) + l_fitPoints() + l_ciBar() + 
@@
 
-# Extract predictions and plot on response scale
-
-gam_max_bb_abund_01_ilink <- family(gam_max_bb_abund_01)$linkinv # get inverse link function
-
-gam_max_bb_abund_01_pred <- predict.gam(gam_max_bb_abund_01, # get model predictions
-                                        type = "terms") %>% 
-  as.data.frame() %>% # convert to data frame
-  as_tibble() %>% # convert to tibble
-  dplyr::select(mgmt = 1, bb.sp = 2, bb.sp.year = 3, gers = 4, # rename
-         hort = 5, jone = 6, mend = 7, mont = 8, 
-         muci = 9, pasc = 10, prat = 11, psit = 12, 
-         pyre = 13, soro = 14, telu = 15, wurf = 16) %>%
-  mutate(fit = # get cumulative fit
-           mgmt + bb.sp + bb.sp.year + gers + 
-           hort + jone + mend + mont + 
-           muci + pasc + prat + psit + 
-           pyre + soro + telu + wurf 
-           ) %>%
-  dplyr::select(fit) %>%
-  mutate(fit.trans = gam_max_bb_abund_01_ilink(fit)) %>% # back-transform to response scale
-  bind_cols(abund_sum)
-
-ggplot(gam_max_bb_abund_01_pred, aes(elev.mean, fit.trans, linetype = year)) + # map year to linetype
-  geom_point(aes(elev.mean, max.bb.abund, shape = tree_line, color = management), 
-             inherit.aes = FALSE, alpha = 0.5, size = 2) + # add original data and map tree line to color
-  #geom_line(size = 0.5, alpha = 0.5) +
-  facet_wrap(~bb.sp, scales = "free_y") +
-  theme_light() +
-  labs(x = "Elevation (m.s.l.)", 
-       y = bquote("Peak bumble bee abundance"),
-       color = "Tree line",
-       shape = "Year",
-       linetype = "Year")
-
-ggsave("../output/gam_bb_abund.pdf")
-ggsave("../output/gam_bb_abund.png")
+# 
+# 
+# # Extract predictions and plot on response scale
+# 
+# gam_max_bb_abund_01_ilink <- family(gam_max_bb_abund_01)$linkinv # get inverse link function
+# 
+# gam_max_bb_abund_01_pred <- predict.gam(gam_max_bb_abund_01, # get model predictions
+#                                         type = "terms") %>% 
+#   as.data.frame() %>% # convert to data frame
+#   as_tibble() %>% # convert to tibble
+#   dplyr::select(mgmt = 1, bb.sp = 2, bb.sp.year = 3, gers = 4, # rename
+#          hort = 5, jone = 6, mend = 7, mont = 8, 
+#          muci = 9, pasc = 10, prat = 11, psit = 12, 
+#          pyre = 13, soro = 14, telu = 15, wurf = 16) %>%
+#   mutate(fit = # get cumulative fit
+#            mgmt + bb.sp + bb.sp.year + gers + 
+#            hort + jone + mend + mont + 
+#            muci + pasc + prat + psit + 
+#            pyre + soro + telu + wurf 
+#            ) %>%
+#   dplyr::select(fit) %>%
+#   mutate(fit.trans = gam_max_bb_abund_01_ilink(fit)) %>% # back-transform to response scale
+#   bind_cols(abund_sum)
+# 
+# ggplot(gam_max_bb_abund_01_pred, aes(elev.mean, fit.trans, linetype = year)) + # map year to linetype
+#   geom_point(aes(elev.mean, max.bb.abund, shape = tree_line, color = management), 
+#              inherit.aes = FALSE, alpha = 0.5, size = 2) + # add original data and map tree line to color
+#   #geom_line(size = 0.5, alpha = 0.5) +
+#   facet_wrap(~bb.sp, scales = "free_y") +
+#   theme_light() +
+#   labs(x = "Elevation (m.s.l.)", 
+#        y = bquote("Peak bumble bee abundance"),
+#        color = "Tree line",
+#        shape = "Year",
+#        linetype = "Year")
+# 
+# ggsave("../output/gam_bb_abund.pdf")
+# ggsave("../output/gam_bb_abund.png")
 ```
@@
 gers_fl_abund <- plot(sm(gam_mean_fl_abund_01, 2)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "gers")
+  labs(x = NULL, y = NULL, title = "gers") +
+  l_rug(length = unit(0.15, "cm"))
 hort_fl_abund <- plot(sm(gam_mean_fl_abund_01, 3)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "hort")
+  labs(x = NULL, y = NULL, title = "hort") +
+  l_rug(length = unit(0.15, "cm"))
 jone_fl_abund <- plot(sm(gam_mean_fl_abund_01, 4)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "jone")
+  labs(x = NULL, y = NULL, title = "jone") +
+  l_rug(length = unit(0.15, "cm"))
 mend_fl_abund <- plot(sm(gam_mean_fl_abund_01, 5)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "mend")
+  labs(x = NULL, y = NULL, title = "mend") +
+  l_rug(length = unit(0.15, "cm"))
 mont_fl_abund <- plot(sm(gam_mean_fl_abund_01, 6)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "mont")
+  labs(x = NULL, y = NULL, title = "mont") +
+  l_rug(length = unit(0.15, "cm"))
 muci_fl_abund <- plot(sm(gam_mean_fl_abund_01, 7)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "muci")
+  labs(x = NULL, y = NULL, title = "muci") +
+  l_rug(length = unit(0.15, "cm"))
 pasc_fl_abund <- plot(sm(gam_mean_fl_abund_01, 8)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "pasc")
+  labs(x = NULL, y = NULL, title = "pasc") +
+  l_rug(length = unit(0.15, "cm"))
 prat_fl_abund <- plot(sm(gam_mean_fl_abund_01, 9)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "prat")
+  labs(x = NULL, y = NULL, title = "prat") +
+  l_rug(length = unit(0.15, "cm"))
 psit_fl_abund <- plot(sm(gam_mean_fl_abund_01, 10)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "psit")
+  labs(x = NULL, y = NULL, title = "psit") +
+  l_rug(length = unit(0.15, "cm"))
 pyre_fl_abund <- plot(sm(gam_mean_fl_abund_01, 11)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "pyre")
+  labs(x = NULL, y = NULL, title = "pyre") +
+  l_rug(length = unit(0.15, "cm"))
 soro_fl_abund <- plot(sm(gam_mean_fl_abund_01, 12)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "soro")
+  labs(x = NULL, y = NULL, title = "soro") +
+  l_rug(length = unit(0.15, "cm"))
 telu_fl_abund <- plot(sm(gam_mean_fl_abund_01, 13)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "telu")
+  labs(x = NULL, y = NULL, title = "telu") +
+  l_rug(length = unit(0.15, "cm"))
 wurf_fl_abund <- plot(sm(gam_mean_fl_abund_01, 14)) + l_fitLine() + l_ciLine() + 
-  labs(x = NULL, y = NULL, title = "wurf")
+  labs(x = NULL, y = NULL, title = "wurf") +
+  l_rug(length = unit(0.15, "cm"))
 mgmt_fl_effect <- plot(pterm(gam_mean_fl_abund_01, 1)) + l_fitPoints() + l_ciBar() + 
@@
 
-# Extract predictions and plot on response scale
-
-gam_mean_fl_abund_01_ilink <- family(gam_mean_fl_abund_01)$linkinv # get inverse link function to transform to the response scale
-
-gam_mean_fl_abund_01_pred <- predict.gam(gam_mean_fl_abund_01, # get model predictions
-                                        type = "terms") %>% 
-  as.data.frame() %>% # convert to data frame
-  as_tibble() %>% # convert to tibble
-  dplyr::select(mgmt = 1, year = 2, bb.sp = 3, gers = 4, # rename
-         hort = 5, jone = 6, mend = 7, mont = 8, 
-         muci = 9, pasc = 10, prat = 11, psit = 12, 
-         pyre = 13, soro = 14, telu = 15, wurf = 16) %>%
-  mutate(fit = # get cumulative fit
-           mgmt + year + bb.sp + gers + 
-           hort + jone + mend + mont + 
-           muci + pasc + prat + psit + 
-           pyre + soro + telu + wurf 
-           ) %>%
-  dplyr::select(fit) %>%
-  mutate(fit.trans = gam_max_bb_abund_01_ilink(fit)) %>% # back-transform to response scale
-  bind_cols(na.omit(abund_sum)) # the na.omit() drops the rows where floral surveying data were missing
-
-ggplot(gam_mean_fl_abund_01_pred, aes(elev.mean, fit.trans, linetype = year)) + # map year to linetype
-  geom_point(aes(elev.mean, mean.fl.abund, color = management), inherit.aes = FALSE, alpha = 0.5) + # add original data and map tree line to color
-  #geom_line(size = 0.5, alpha = 0.5) +
-  facet_wrap(~bb.sp, scales = "free_y") +
-  theme_light() +
-  labs(x = "Elevation (m.s.l.)", 
-       y = bquote("Peak floral abundance"),
-       color = "Tree line",
-       shape = "Year",
-       linetype = "Year")
-
-ggsave("../output/gam_fl_abund.pdf")
-ggsave("../output/gam_fl_abund.png")
+# 
+# 
+# # Extract predictions and plot on response scale
+# 
+# gam_mean_fl_abund_01_ilink <- family(gam_mean_fl_abund_01)$linkinv # get inverse link function to transform to the response scale
+# 
+# gam_mean_fl_abund_01_pred <- predict.gam(gam_mean_fl_abund_01, # get model predictions
+#                                         type = "terms") %>% 
+#   as.data.frame() %>% # convert to data frame
+#   as_tibble() %>% # convert to tibble
+#   dplyr::select(mgmt = 1, year = 2, bb.sp = 3, gers = 4, # rename
+#          hort = 5, jone = 6, mend = 7, mont = 8, 
+#          muci = 9, pasc = 10, prat = 11, psit = 12, 
+#          pyre = 13, soro = 14, telu = 15, wurf = 16) %>%
+#   mutate(fit = # get cumulative fit
+#            mgmt + year + bb.sp + gers + 
+#            hort + jone + mend + mont + 
+#            muci + pasc + prat + psit + 
+#            pyre + soro + telu + wurf 
+#            ) %>%
+#   dplyr::select(fit) %>%
+#   mutate(fit.trans = gam_max_bb_abund_01_ilink(fit)) %>% # back-transform to response scale
+#   bind_cols(na.omit(abund_sum)) # the na.omit() drops the rows where floral surveying data were missing
+# 
+# ggplot(gam_mean_fl_abund_01_pred, aes(elev.mean, fit.trans, linetype = year)) + # map year to linetype
+#   geom_point(aes(elev.mean, mean.fl.abund, color = management), inherit.aes = FALSE, alpha = 0.5) + # add original data and map tree line to color
+#   #geom_line(size = 0.5, alpha = 0.5) +
+#   facet_wrap(~bb.sp, scales = "free_y") +
+#   theme_light() +
+#   labs(x = "Elevation (m.s.l.)", 
+#        y = bquote("Peak floral abundance"),
+#        color = "Tree line",
+#        shape = "Year",
+#        linetype = "Year")
+# 
+# ggsave("../output/gam_fl_abund.pdf")
+# ggsave("../output/gam_fl_abund.png")
 ```

### 2021-04-07 code/ms1_beta_diversity.Rmd
updated analyses and figures ot include mamagement; dropped MRM analysis

@@
 library(vegan)
-library(see)
+library(ecodist)
+library(gt)
 library(tidyverse)
@@
     ungroup() %>%
-    dplyr::select(site, elev.mean, y, x, samples) %>%
-    unique() %>%
+    mutate(mowing = if_else(management == "mowing", 1, 0),
+           grasing = if_else(management == "grasing", 1, 0),
+           none = if_else(management == "none", 1, 0)) %>%
+    dplyr::select(site, elev.mean, x, y, samples, mowing, grasing, none) %>%
+    distinct() %>%
     arrange(site) 
@@
     ungroup() %>%
-    dplyr::select(site, elev.mean, y, x, samples) %>%
+    mutate(mowing = if_else(management == "mowing", 1, 0),
+           grasing = if_else(management == "grasing", 1, 0),
+           none = if_else(management == "none", 1, 0)) %>%
+    dplyr::select(site, elev.mean, x, y, samples, mowing, grasing, none) %>%
     unique() %>%
@@
 ### MRM analysis
-```{r}
-library(ecodist)
-
+```{r, warning=FALSE}
 survey_sample_diff <- survey %>%
@@
 
+# Function for tabulating modle coefficients
+mrm_tab <- function(x, response) {
+  x %>%
+    as.data.frame() %>%
+    rownames_to_column(var = "predictor") %>%
+    slice(-1) %>%
+    mutate(response = rep(response, n()),
+           Predictor = case_when(
+             predictor == "elev.diff" ~ "elevation",
+             predictor == "sample.diff" ~ "sampling freq.",
+             predictor == "mgmt.diff" ~ "management")) %>%
+    select(response, Predictor, b, pval) %>%
+    tibble()
+}
+
+# Call models
 mrm_fl <- MRM(beta ~ elev.diff + sample.diff + mgmt.diff, 
@@
               nperm = 1000,
-              method = "logistic")
+              method = "logistic")$coef %>%
+  mrm_tab("Flora")
 
@@
               nperm = 1000,
-              method = "logistic")
+              method = "logistic")$coef %>%
+  mrm_tab("Bumble bees")
 
@@
               nperm = 1000,
-              method = "logistic")
+              method = "logistic")$coef %>%
+  mrm_tab("Total interactions")
 
@@
               nperm = 1000,
-              method = "logistic")
+              method = "logistic")$coef %>%
+  mrm_tab("ST partition")
 
@@
               nperm = 1000,
-              method = "logistic")
+              method = "logistic")$coef %>%
+  mrm_tab("ST.l partition")
 
@@
               nperm = 1000,
-              method = "logistic")
+              method = "logistic")$coef %>%
+  mrm_tab("ST.h partition")
 
@@
               nperm = 1000,
-              method = "logistic")
+              method = "logistic")$coef %>%
+  mrm_tab("ST.lh partition")
 
@@
               nperm = 1000,
-              method = "logistic")
+              method = "logistic")$coef %>%
+  mrm_tab("Rewiring partition")
 
-# Partial Mantel tests are another option, perhaps more routine to interpret
-# mantel_fl <- mantel(beta ~ elev.diff + sample.diff, 
-#               data = mrm_survey_data,
-#               mrank = TRUE)
-# 
-# mantel_bb <- mantel(beta ~ elev.diff + sample.diff, 
-#               data = mrm_net_data,
-#               mrank = TRUE)
-# 
-# mantel_WN <- mantel(value ~ elev.diff + sample.diff, 
-#                     data = filter(mrm_beta_data, metric == "WN"),
-#                     mrank = TRUE)
-# 
-# mantel_ST <- mantel(value ~ elev.diff + sample.diff, 
-#                     data = filter(mrm_beta_data, metric == "ST"),
-#                     mrank = TRUE)
-# 
-# mantel_ST.l <- mantel(value ~ elev.diff + sample.diff, 
-#                     data = filter(mrm_beta_data, metric == "ST.l"),
-#                     mrank = TRUE)
-# 
-# mantel_ST.h <- mantel(value ~ elev.diff + sample.diff, 
-#                     data = filter(mrm_beta_data, metric == "ST.h"),
-#                     mrank = TRUE)
-# 
-# mantel_ST.lh <- mantel(value ~ elev.diff + sample.diff, 
-#                     data = filter(mrm_beta_data, metric == "ST.lh"),
-#                     mrank = TRUE)
-# 
-# mantel_OS <- mantel(value ~ elev.diff + sample.diff, 
-#                     data = filter(mrm_beta_data, metric == "OS"),
-#                     mrank = TRUE)
+# Tabulate results
+mrm_results <- bind_rows(mrm_fl, mrm_bb, mrm_WN, mrm_ST, 
+                         mrm_ST.h, mrm_ST.l, mrm_ST.lh, mrm_OS) %>%
+  gt(groupname_col = "response") %>%
+  tab_style(
+    style = list(cell_text(weight = "bold")),
+    locations = list(cells_column_labels(everything()))
+  ) %>%
+  tab_style(
+    style = list(cell_text(style = "italic")),
+    locations = list(cells_group(everything()))
+  ) %>%
+  # data_color(
+  #   columns = vars(b),
+  #   colors = scales::col_numeric(
+  #     palette = c("white", "red"),
+  #     domain = c(-0.01, 0.5)), alpha = 0.75
+  #   ) %>%
+  tab_style(
+    style = list(
+      #cell_fill(color = "lightcyan"),
+      cell_text(weight = "bold")
+      ),
+    locations = cells_body(
+      columns = vars(pval),
+      rows = pval < 0.05
+    ))
+
+gtsave(mrm_results,"../output/mrm_tab.pdf")
+
+mrm_results_sub <- bind_rows(mrm_fl, mrm_bb, mrm_WN) %>%
+  gt(groupname_col = "response") %>%
+  tab_style(
+    style = list(cell_text(weight = "bold")),
+    locations = list(cells_column_labels(everything()))
+  ) %>%
+  tab_style(
+    style = list(cell_text(style = "italic")),
+    locations = list(cells_group(everything()))
+  ) %>%
+  # data_color(
+  #   columns = vars(b),
+  #   colors = scales::col_numeric(
+  #     palette = c("white", "red"),
+  #     domain = c(-0.01, 0.5)), alpha = 0.75
+  #   ) %>%
+  tab_style(
+    style = list(
+      #cell_fill(color = "lightcyan"),
+      cell_text(weight = "bold")
+      ),
+    locations = cells_body(
+      columns = vars(pval),
+      rows = pval < 0.05
+    ))
 ```
@@
 # Prep sitepairs
-sitepair.WN <- interaction_gdm_prep(net)
+sitepair_WN <- interaction_gdm_prep(net)
 
 # Run GDM analysis
-gdm_WN <- gdm(sitepair.WN, geo = FALSE)
+gdm_WN <- gdm(sitepair_WN, geo = FALSE)
 summary(gdm_WN)
-gdm.varImp_WN <- gdm.varImp(sitepair.WN, geo = FALSE)
+gdm.varImp_WN <- gdm.varImp(sitepair_WN, geo = FALSE)
 
@@
 # Plot with confidence intervals 
-pdf("../output/gdm_interactions.pdf", height = 2.5, width =  6.5)
-plotUncertainty(sitepair.WN, sampleSites = 0.7, bsIters = 100, geo  = TRUE, 
-                splineCol = "black", plot.layout = c(1,3))
-dev.off()
+# pdf("../output/gdm_interactions.pdf", height = 2.5, width =  6.5)
+# plotUncertainty(sitepair.WN, sampleSites = 0.7, bsIters = 100, geo  = FALSE, 
+#                 splineCol = "black", plot.layout = c(1,3))
+# dev.off()
 ```
@@
 # Prep input for GDM
-survey_bioData <- biodata_prep_survey(survey)
-survey_predData <- predData_prep_survey(survey)
-survey_sitepair <- format_set(survey_bioData, survey_predData)
+bioData_survey <- biodata_prep_survey(survey)
+predData_survey <- predData_prep_survey(survey)
+sitepair_survey <- format_set(bioData_survey, predData_survey)
 
 # Call GDMs
-gdm_survey <- gdm(survey_sitepair, geo = TRUE)
+gdm_survey <- gdm(survey_sitepair, geo = FALSE)
 summary(gdm_survey)
+gdm.varImp_survey <- gdm.varImp(sitepair_survey, geo = FALSE)
 
@@
 # Plot with confidence intervals (slow)
-pdf("../output/gdm_flora.pdf", height = 2.5, width =  6.5)
-plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 100, geo  = TRUE, 
-                splineCol = "#00BFC4", plot.layout = c(1,3))
-dev.off()
+# pdf("../output/gdm_flora.pdf", height = 2.5, width =  6.5)
+# plotUncertainty(sitepair_survey, sampleSites = 0.7, bsIters = 100, geo  = FALSE, 
+#                 splineCol = "#00BFC4", plot.layout = c(1,3))
+# dev.off()
 ```
@@
 # Prep input for GDM
-BB_bioData <- biodata_prep_BB(net)
-BB_predData <- predData_prep_BB(net)
-BB_sitepair <- format_set(BB_bioData, BB_predData)
+bioData_BB <- biodata_prep_BB(net)
+predData_BB <- predData_prep_BB(net)
+sitepair_BB <- format_set(bioData_BB, predData_BB)
 
 # Call GDMs
-gdm_BB <- gdm(BB_sitepair, geo = TRUE)
+gdm_BB <- gdm(sitepair_BB, geo = FALSE)
 summary(gdm_BB)
+gdm.varImp_BB <- gdm.varImp(sitepair_BB, geo = FALSE)
 
@@
 # Plot with confidence intervals (slow)
-pdf("../output/gdm_BB.pdf", height = 2.5, width =  6.5)
-plotUncertainty(BB_sitepair, sampleSites = 0.7, bsIters = 100, geo  = TRUE, 
-                splineCol = "#F8766D", plot.layout = c(1,3))
-dev.off()
+# pdf("../output/gdm_BB.pdf", height = 2.5, width =  6.5)
+# plotUncertainty(sitepair_BB, sampleSites = 0.7, bsIters = 100, geo  = TRUE, 
+#                 splineCol = "#F8766D", plot.layout = c(1,3))
+# dev.off()
 ```
 
-### Tabulate model output
+### Tabulate gdm.varImp results
 ```{r}
-# assign("gdm_WN.2010.imp", get(load("~/Documents/Research/bb_berchtesgaden/analyses/output/interaction_permGDM_2010.RData")))
-# 
-# assign("gdm_WN.2011.imp", get(load("~/Documents/Research/bb_berchtesgaden/analyses/output/interaction_permGDM_2011.RData")))
-# 
-# assign("gdm_WN.2012.imp", get(load("~/Documents/Research/bb_berchtesgaden/analyses/output/interaction_permGDM_2012.RData")))
+# Function for tabulate gdm.varImp output 
+gdm_tab <- function(x, response) {
+  x %>%
+    as.data.frame() %>%
+    rownames_to_column(var = "predictor") %>%
+    mutate(response = rep(response, n()),
+           Predictor = case_when(
+             predictor == "elev.mean" ~ "elevation",
+             predictor == "samples" ~ "sampling freq.",
+             predictor == "mowing" ~ "mowing",
+             predictor == "grasing" ~ "grasing",
+             predictor == "none" ~ "none")) %>%
+    select(response, Predictor, Importance, pval) %>%
+    tibble()
+}
 
-####### WN
-gdm_WN.2010.tab1 <- gdm_WN.2010.imp[1] %>% 
+# Interaction model
+gdm.varImp_WN.tab2 <- gdm.varImp_WN[[2]] %>%
   as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble() %>%
-  dplyr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2) %>%
-  mutate(subject = "WN",
-         year = 2010) %>%
-  dplyr::select(subject, year, 
-         mod.dev = `Model deviance`, 
-         dev.exp = `Percent deviance explained`, 
-         mod.pval = `Model p-value`)
+  select(Importance = 1)
 
-gdm_WN.2010.tab2 <- gdm_WN.2010.imp[2] %>% 
+gdm.varImp_WN.tab3 <- gdm.varImp_WN[[3]] %>%
   as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble() %>%
-  dplyr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2) %>%
-  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)
+  select(pval = 1)
 
-gdm_WN.2010.tab3 <- gdm_WN.2010.imp[3] %>% 
+gdm.varImp_WN.tab <- bind_cols(gdm.varImp_WN.tab2, gdm.varImp_WN.tab3) %>%
+  gdm_tab("Interactions")
+
+# Floral model
+gdm.varImp_survey.tab2 <- gdm.varImp_survey[[2]] %>%
   as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble()  %>%
-  dplyr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2)  %>%
-  dplyr::select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)
+  select(Importance = 1)
 
-gdm_WN.2011.tab1 <- gdm_WN.2011.imp[1] %>% 
+gdm.varImp_survey.tab3 <- gdm.varImp_survey[[3]] %>%
   as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble() %>%
-  dplyr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2) %>%
-  mutate(subject = "WN",
-         year = 2011) %>%
-  dplyr::select(subject, year, 
-         mod.dev = `Model deviance`, 
-         dev.exp = `Percent deviance explained`, 
-         mod.pval = `Model p-value`)
+  select(pval = 1)
 
-gdm_WN.2011.tab2 <- gdm_WN.2011.imp[2] %>% 
+gdm.varImp_survey.tab <- bind_cols(gdm.varImp_survey.tab2, gdm.varImp_survey.tab3) %>%
+  gdm_tab("Flora")
+
+# BB model
+gdm.varImp_BB.tab2 <- gdm.varImp_BB[[2]] %>%
   as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble() %>%
-  dplyr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2) %>%
-  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)
+  select(Importance = 1)
 
-gdm_WN.2011.tab3 <- gdm_WN.2011.imp[3] %>% 
+gdm.varImp_BB.tab3 <- gdm.varImp_BB[[3]] %>%
   as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble()  %>%
-  dplyr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2)  %>%
-  dplyr::select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)
+  select(pval = 1)
 
-gdm_WN.2012.tab1 <- gdm_WN.2012.imp[1] %>% 
-  as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble() %>%
-  dplyr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2) %>%
-  mutate(subject = "WN",
-         year = 2012) %>%
-  dplyr::select(subject, year, 
-         mod.dev = `Model deviance`, 
-         dev.exp = `Percent deviance explained`, 
-         mod.pval = `Model p-value`)
+gdm.varImp_BB.tab <- bind_cols(gdm.varImp_BB.tab2, gdm.varImp_BB.tab3) %>%
+  gdm_tab("Bumble bees")
 
-gdm_WN.2012.tab2 <- gdm_WN.2012.imp[2] %>% 
-  as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble() %>%
-  dplyr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2) %>%
-  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)
-
-gdm_WN.2012.tab3 <- gdm_WN.2012.imp[3] %>% 
-  as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble()  %>%
-  dplyr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2)  %>%
-  dplyr::select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)
-
-gdm_WN.2010.tab_join <- bind_cols(gdm_WN.2010.tab1,
-                                  gdm_WN.2010.tab2, 
-                                  gdm_WN.2010.tab3)
-
-gdm_WN.2011.tab_join <- bind_cols(gdm_WN.2011.tab1,
-                                  gdm_WN.2011.tab2, 
-                                  gdm_WN.2011.tab3)
-
-gdm_WN.2012.tab_join <- bind_cols(gdm_WN.2012.tab1,
-                                  gdm_WN.2012.tab2, 
-                                  gdm_WN.2012.tab3)
-
-####### survey
-gdm_survey.2010.tab1 <- gdm_survey.2010.imp[1] %>% 
-  as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble() %>%
-  dplyr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2) %>%
-  mutate(subject = "survey",
-         year = 2010) %>%
-  dplyr::select(subject, year, 
-         mod.dev = `Model deviance`, 
-         dev.exp = `Percent deviance explained`, 
-         mod.pval = `Model p-value`)
-
-gdm_survey.2010.tab2 <- gdm_survey.2010.imp[2] %>% 
-  as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble() %>%
-  dplyr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2) %>%
-  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)
-
-gdm_survey.2010.tab3 <- gdm_survey.2010.imp[3] %>% 
-  as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble()  %>%
-  dplyr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2)  %>%
-  dplyr::select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)
-
-gdm_survey.2011.tab1 <- gdm_survey.2011.imp[1] %>% 
-  as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble() %>%
-  dplyr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2) %>%
-  mutate(subject = "survey",
-         year = 2011) %>%
-  dplyr::select(subject, year, 
-         mod.dev = `Model deviance`, 
-         dev.exp = `Percent deviance explained`, 
-         mod.pval = `Model p-value`)
-
-gdm_survey.2011.tab2 <- gdm_survey.2011.imp[2] %>% 
-  as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble() %>%
-  dplyr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2) %>%
-  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)
-
-gdm_survey.2011.tab3 <- gdm_survey.2011.imp[3] %>% 
-  as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble()  %>%
-  dplyr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2)  %>%
-  select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)
-
-gdm_survey.2012.tab1 <- gdm_survey.2012.imp[1] %>% 
-  as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble() %>%
-  dplylr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2) %>%
-  mutate(subject = "survey",
-         year = 2012) %>%
-  dplyr::select(subject, year, 
-         mod.dev = `Model deviance`, 
-         dev.exp = `Percent deviance explained`, 
-         mod.pval = `Model p-value`)
-
-gdm_survey.2012.tab2 <- gdm_survey.2012.imp[2] %>% 
-  as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble() %>%
-  dplyr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2) %>%
-  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)
-
-gdm_survey.2012.tab3 <- gdm_survey.2012.imp[3] %>% 
-  as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble()  %>%
-  dplytr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2)  %>%
-  dplyr::select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)
-
-gdm_survey.2010.tab_join <- bind_cols(gdm_survey.2010.tab1,
-                                  gdm_survey.2010.tab2, 
-                                  gdm_survey.2010.tab3)
-
-gdm_survey.2011.tab_join <- bind_cols(gdm_survey.2011.tab1,
-                                  gdm_survey.2011.tab2, 
-                                  gdm_survey.2011.tab3)
-
-gdm_survey.2012.tab_join <- bind_cols(gdm_survey.2012.tab1,
-                                  gdm_survey.2012.tab2, 
-                                  gdm_survey.2012.tab3)
-
-######## BB
-gdm_BB.2010.tab1 <- gdm_BB.2010.imp[1] %>% 
-  as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble() %>%
-  dplyr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2) %>%
-  mutate(subject = "BB",
-         year = 2010) %>%
-  dplyr::select(subject, year, 
-         mod.dev = `Model deviance`, 
-         dev.exp = `Percent deviance explained`, 
-         mod.pval = `Model p-value`)
-
-gdm_BB.2010.tab2 <- gdm_BB.2010.imp[2] %>% 
-  as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble() %>%
-  dplyr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2) %>%
-  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)
-
-gdm_BB.2010.tab3 <- gdm_BB.2010.imp[3] %>% 
-  as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble()  %>%
-  dplyr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2)  %>%
-  dplyr::select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)
-
-gdm_BB.2011.tab1 <- gdm_BB.2011.imp[1] %>% 
-  as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble() %>%
-  dplyr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2) %>%
-  mutate(subject = "BB",
-         year = 2011) %>%
-  dplyr::select(subject, year, 
-         mod.dev = `Model deviance`, 
-         dev.exp = `Percent deviance explained`, 
-         mod.pval = `Model p-value`)
-
-gdm_BB.2011.tab2 <- gdm_BB.2011.imp[2] %>% 
-  as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble() %>%
-  dplyr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2) %>%
-  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)
-
-gdm_BB.2011.tab3 <- gdm_BB.2011.imp[3] %>% 
-  as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble()  %>%
-  dplyr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2)  %>%
-  dplyr::select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)
-
-gdm_BB.2012.tab1 <- gdm_BB.2012.imp[1] %>% 
-  as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble() %>%
-  dplyr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2) %>%
-  mutate(subject = "BB",
-         year = 2012) %>%
-  dplyr::select(subject, year, 
-         mod.dev = `Model deviance`, 
-         dev.exp = `Percent deviance explained`, 
-         mod.pval = `Model p-value`)
-
-gdm_BB.2012.tab2 <- gdm_BB.2012.imp[2] %>% 
-  as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble() %>%
-  dplyr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2) %>%
-  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)
-
-gdm_BB.2012.tab3 <- gdm_BB.2012.imp[3] %>% 
-  as.data.frame() %>%
-  rownames_to_column() %>%
-  as_tibble()  %>%
-  dplyr::select(1, 2) %>%
-  pivot_wider(names_from = rowname, values_from = 2)  %>%
-  dplyr::select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)
-
-gdm_BB.2010.tab_join <- bind_cols(gdm_BB.2010.tab1,
-                                  gdm_BB.2010.tab2, 
-                                  gdm_BB.2010.tab3)
-
-gdm_BB.2011.tab_join <- bind_cols(gdm_BB.2011.tab1,
-                                  gdm_BB.2011.tab2, 
-                                  gdm_BB.2011.tab3)
-
-gdm_BB.2012.tab_join <- bind_cols(gdm_BB.2012.tab1,
-                                  gdm_BB.2012.tab2, 
-                                  gdm_BB.2012.tab3)
-
-#######################
-
-gdm_BB.tab_join <- bind_rows(gdm_BB.2010.tab_join,
-                             gdm_BB.2011.tab_join,
-                             gdm_BB.2012.tab_join,
-                             gdm_survey.2010.tab_join,
-                             gdm_survey.2011.tab_join,
-                             gdm_survey.2012.tab_join,
-                             gdm_WN.2010.tab_join,
-                             gdm_WN.2011.tab_join,
-                             gdm_WN.2012.tab_join) %>%
-  dplyr::select(-subject)
-
-gdm_tab <- gdm_BB.tab_join %>%
-  gt(rowname_col = "year") %>%
-  tab_spanner(
-    label = "Model",
-    columns = vars(mod.dev, dev.exp, mod.pval)
+# Merge and format
+gdm.varImp.tab <- bind_rows(gdm.varImp_WN.tab, gdm.varImp_survey.tab, gdm.varImp_BB.tab) %>%
+  mutate(Importance = round(Importance, 2)) %>%
+  gt(groupname_col = "response") %>%
+  tab_style(
+    style = list(cell_text(weight = "bold")),
+    locations = list(cells_column_labels(everything()))
   ) %>%
-  tab_spanner(
-    label = "Geographic distance",
-    columns = vars(geo.imp, geo.pval)
+  tab_style(
+    style = list(cell_text(style = "italic")),
+    locations = list(cells_row_groups(everything()))
   ) %>%
-  tab_spanner(
-    label = "Elevational distance",
-    columns = vars(elev.imp, elev.pval)
-  ) %>%
-  tab_spanner(
-    label = "Temporal distance",
-    columns = vars(yday.imp, yday.pval)
-  ) %>%
-  tab_row_group(
-    group = "Bumble bees",
-    rows = 1:3
-  ) %>% 
-  tab_row_group(
-    group = "Flora",
-    rows = 4:6
-  ) %>%
-  tab_row_group(
-    group = "Interactions",
-    rows = 7:9
-  ) %>%
-  fmt_number(
-    columns = c("mod.dev", "dev.exp",
-                "geo.imp", "elev.imp", "yday.imp"),
-    decimals = 1
-  ) %>%
-  fmt_number(
-    columns = c("mod.pval", "geo.pval", 
-                "elev.pval", "yday.pval"),
-    decimals = 3
-  ) %>%
-  cols_label(mod.dev = "tot. dev.",
-             dev.exp = "dev. expl. (%)",
-             mod.pval = "p",
-             geo.imp = "imp.",
-             elev.imp = "imp.",
-             yday.imp = "imp.",
-             geo.pval = "p-val",
-             elev.pval = "p-val",
-             yday.pval = "p-val")
-
-gdm_tab.nopval <- gdm_BB.tab_join %>%
-  dplyr::select(-c(geo.pval, elev.pval, yday.pval, mod.pval)) %>%
-  gt(rowname_col = "year") %>%
-  tab_spanner(
-    label = "Model",
-    columns = vars(mod.dev, dev.exp)
-  ) %>%
-  tab_spanner(
-    label = "Variable importance (%)",
-    columns = vars(geo.imp, elev.imp, yday.imp)
-  ) %>%
-  tab_row_group(
-    group = "Bumble bees",
-    rows = 1:3
-  ) %>% 
-  tab_row_group(
-    group = "Flora",
-    rows = 4:6
-  ) %>%
-  tab_row_group(
-    group = "Interactions",
-    rows = 7:9
-  ) %>%
-  fmt_number(
-    columns = c("mod.dev", "dev.exp",
-                "geo.imp", "elev.imp", "yday.imp"),
-    decimals = 1
-  ) %>%
-  cols_label(mod.dev = "tot. dev.",
-             dev.exp = "dev. expl. (%)",
-             geo.imp = "geo. dist.",
-             elev.imp = "elev. dist.",
-             yday.imp = "temp. dist.") %>%
   data_color(
-    columns = vars(geo.imp, elev.imp, yday.imp),
+    columns = vars(Importance),
     colors = scales::col_numeric(
       palette = c("white", "red"),
-      domain = c(-5, 100))
+      domain = c(-0, 56)), alpha = 0.75
     ) %>%
-  data_color(
-    columns = vars(dev.exp),
-    colors = scales::col_numeric(
-      palette = c("white", "red"), 
-      domain = c(-5, 100))
-  )
+  tab_style(
+    style = list(
+      cell_text(weight = "bold")
+      ),
+    locations = cells_body(
+      columns = vars(pval),
+      rows = pval < 0.05
+    ))
 
-gtsave(gdm_tab.nopval, "gdm_tab.png","./output/gdm_tab.png")
+gtsave(gdm.varImp.tab, "../output/gdm_tab.pdf")
+```
+
+### Plot GDM splines
+```{r}
+# Interaction model
+gdm_WN_splines1 <- isplineExtract(gdm_WN)$x %>%
+  as.data.frame() %>%
+  select(elevation = elev.mean)
+
+gdm_WN_splines2 <- isplineExtract(gdm_WN)$y %>%
+  as.data.frame() %>%
+  select(response = elev.mean)
+
+gdm_WN_splines <- bind_cols(gdm_WN_splines1, gdm_WN_splines2) %>%
+  mutate(level = rep("Interactions"))
+
+# Flora model
+gdm_survey_splines1 <- isplineExtract(gdm_survey)$x %>%
+  as.data.frame() %>%
+  select(elevation = elev.mean)
+
+gdm_survey_splines2 <- isplineExtract(gdm_survey)$y %>%
+  as.data.frame() %>%
+  select(response = elev.mean)
+
+gdm_survey_splines <- bind_cols(gdm_survey_splines1, gdm_survey_splines2) %>%
+  mutate(level = rep("Flora"))
+
+# BB model
+gdm_BB_splines1 <- isplineExtract(gdm_BB)$x %>%
+  as.data.frame() %>%
+  select(elevation = elev.mean)
+
+gdm_BB_splines2 <- isplineExtract(gdm_BB)$y %>%
+  as.data.frame() %>%
+  select(response = elev.mean)
+
+gdm_BB_splines <- bind_cols(gdm_BB_splines1, gdm_BB_splines2) %>%
+  mutate(level = rep("Bumble bees"))
+
+# Combine
+gdm_splines <- bind_rows(gdm_WN_splines, gdm_survey_splines, gdm_BB_splines) %>%
+  mutate(Level = factor(level, levels = c("Interactions", "Flora", "Bumble bees")))
+
+tree_line <- site_data %>%
+  select(elevation = elev.mean, tree_line)
+
+# Plot
+ggplot(gdm_splines, aes(elevation, response, linetype = Level)) +
+  geom_line() +
+  #geom_vline(xintercept = 1500) +
+  geom_rug(data = tree_line, aes(elevation, color = tree_line), inherit.aes = FALSE) +
+  labs(color = "Tree Line", x = "Elevation", y = "Response") +
+  theme_light(14)
+
+ggsave("../output/gdm_splines.pdf")
+ggsave("../output/gdm_splines.png")
 
 ```
-
-### GDM all together now
-```{r}
-splines <- bind_rows(interaction_splines, bb_splines, survey_splines) %>%
-  mutate(subject = factor(subject))
-
-# ggplot(splines, aes(predictor, response, color = subject)) +
-#   geom_line() +
-#   facet_grid(year ~ x.var, scales = "free_x") +
-#   xlab(NULL) +
-#   ylab("f(x)") +
-#   theme_light(12)
-# ggsave("./output/GDM.png", width = 6.5, height = 5, units = "in")
-
-splot1 <- ggplot(filter(splines, x.var == "Geographic dist."), 
-                 aes(predictor, response, color = subject)) +
-  geom_line() +
-  #facet_wrap(~year, ncol = 1, scales = "free_x") +
-  facet_grid(year ~ x.var, scales = "free_x") +
-  xlab("km") +
-  ylab("f(x)") +
-  ylim(c(0, 7)) +
-  theme_light(12) +
-  theme(legend.position = "none")
-
-splot2 <- ggplot(filter(splines, x.var == "Elevation"), 
-       aes(predictor, response, color = subject)) +
-  geom_line() +
-  #facet_wrap(~year, ncol = 1, scales = "free_x") +
-  facet_grid(year ~ x.var, scales = "free_x") +
-  xlab("m") +
-  ylab(NULL) +
-  ylim(c(0, 7)) +
-  theme_light(12) +
-  theme(legend.position = "none")
-
-splot3 <- ggplot(filter(splines, x.var == "Time"), 
-       aes(as_date(predictor), response, color = subject)) +
-  geom_line() +
-  #facet_wrap(~year, ncol = 1, scales = "free_x") +
-  facet_grid(year ~ x.var, scales = "free_x") +
-  xlab("date") +
-  ylab(NULL) +
-  ylim(c(0, 7)) +
-  theme_light(12) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
-
-png("./output/GDM_splines.png", width = 8, height = 5, units = "in", res = 300)
-splot1 | splot2 | splot3
-dev.off()
-```
-
-# If I could get this to work, it would be cool to transform and plot my GDMs. But the script crashes every time. 
-```{r}
-
-# Function to build data set for gdm.transform()
-transform_data <- function(dat) {
-
-  ## yday min/max
-  min_yday <- min(dat$s1.yday)
-  max_yday <- max(dat$s1.yday)
-
-  ## elev min/max
-  min_elev <- min(dat$s1.elev.mean)
-  max_elev <- max(dat$s1.elev.mean)
-
-  ## lat/lon min/max
-  min_x1 <- min(dat$s1.xCoord)
-  max_x1 <- max(dat$s1.xCoord)
-
-  min_y1 <- min(dat$s1.yCoord)
-  max_y1 <- max(dat$s1.yCoord)
-
-  min_x2 <- min(dat$s2.xCoord)
-  max_x2 <- max(dat$s2.xCoord)
-
-  min_y2 <- max(dat$s2.yCoord)
-  max_y2 <- max(dat$s2.yCoord)
-
-  # sequence of 10 points along the yday and elev ranges, respectively
-  s1.xCoord <- seq(min_x1, max_x1, length.out = 10)
-  s1.yCoord <- seq(min_y1, max_y1, length.out = 10)
-  s2.xCoord <- seq(min_x2, max_x2, length.out = 10)
-  s2.yCoord <- seq(min_y2, max_y2, length.out = 10)
-  s1.elev.mean <- seq(min_elev, max_elev, length.out = 10)
-  s1.yday <- seq(min_yday, max_yday, length.out = 10)
-  s2.elev.mean <- seq(min_elev, max_elev, length.out = 10)
-  s2.yday <- seq(min_yday, max_yday, length.out = 10)
-
-  newdata <- tibble(s1.xCoord, s1.yCoord, s2.xCoord, s2.yCoord, 
-                    s1.elev.mean, s1.yday, s2.elev.mean, s2.yday) %>% # combine seqs into data frame
-    complete(s1.xCoord, s1.yCoord, s2.xCoord, s2.yCoord, 
-             s1.elev.mean, s1.yday, s2.elev.mean, s2.yday) %>% # get all possible values
-    mutate(distance = rep(1, n()),
-           weights = rep(1, n())) %>%
-    as.data.frame()
-
-  # coordinates(newdata) <- ~ X + Y
-  # # coerce to SpatialPixelsDataFrame
-  # gridded(newdata) <- TRUE
-  # # coerce to raster
-  # out <- raster(newdata)
-}
-
-transdata_gdm_WN.2010 <- transform_data(sitepair_2010.WN)
-
-gdm_WN.2010.trans <- gdm.transform(gdm_WN.2010, transdata_gdm_WN.2010)
-gdm_WN.2010.pred <- predict(gdm_WN.2010, gdm_WN.2010.trans)
-
-```
-
-## GDM with BBGDM
-```{r}
-library(bbgdm)
-
-spdat2010 <- net %>%
-  filter(year == 2010) %>%
-  select(bb.sp, plant.sp.abb, site, date) %>%
-  distinct() %>%
-  mutate(present = rep(1, length(.[,1]))) %>%
-  unite(interaction, c(bb.sp, plant.sp.abb), sep = "_") %>%
-  unite(site.date, c("site", "date"), sep = "_") %>%
-  pivot_wider(names_from = interaction, values_from = present) %>%
-  replace(is.na(.), 0) %>%
-  column_to_rownames("site.date") %>%
-  as.matrix()
-
-envdat2010 <- net %>%
-  filter(year == 2010) %>%
-  select(bb.sp, plant.sp.abb, site, date) %>%
-  mutate(yday = yday(date)) %>%
-  left_join(site_data) %>%
-  unite(site.date, c("site", "date"), sep = "_") %>%
-  select(site.date, elev.mean, yday, x, y) %>%
-  distinct() %>%
-  column_to_rownames("site.date") %>%
-  as.matrix()
-
-form <- ~ 1 + elev.mean + yday
-
-bbgdm2010 <- bbgdm(form, sp.dat = spdat2010, env.dat = envdat2010,
-                   family = "binomial", link = "logit", dism_metric = "bray_curtis",
-                   nboot = 10, spline_type = "ispline", spline_df = 5, 
-                   spline_knots = 3, geo = FALSE, 
-                   #geo.type = "euclidean", coord.names = c("x", "y"), 
-                   optim.meth = "nlmnib", 
-                   est.var = FALSE, trace = FALSE, prior = FALSE)
-
-plot(as.response(bbgdm2010))
-bbgdm.wald.test(bbgdm2010)
-plot(diagnostics(bbgdm2010))
-```

### 2021-04-07 ms_1/ms_1.Rmd
updated analyses and figures ot include mamagement; dropped MRM analysis

@@
 We opted to work at the genus level based on the reasoning that if a bumble bee species visits one member of a given genus, other members of the same genus should also be considered potential floral hosts.
-This approach was intended to dampen the effects of false non-detection on the estimation of floral resource abundance, particularly for rare bumble bee species whose diet breadth would tend to be underestimated simply due to sparsity of observations [@Williams2005-yp].
+This approach was intended to dampen the effects of false non-detection, particularly for rare bumble bee species whose diet breadth would tend to be underestimated simply due to sparsity of observations [@Williams2005-yp].
 We also opted to treat the cryptic *B.
@@
 In both models, we opted to collapse the temporal variation of abundance within sites and focus on elevational patterns.
-For bumble bees, we modeled *peak* abundance at each site as a smooth response to elevation, with bumble bee species as a smoothing factor, management (grazing, mowing, or unmanaged) as a fixed effect, and species:year as a random intercept term (exploratory analysis indicated that species varied differently across years).
+For bumble bees, we modeled *peak* abundance at each site as a smooth response to elevation, with bumble bee species as a smoothing factor, site management (grazing, mowing, or unmanaged) as a fixed effect, and species:year as a random intercept term (exploratory analysis indicated that species varied differently across years).
 For floral resources, we modeled *mean* flower cover, weighted per bumble bee species (i.e.
@@
 After pooling, we calculated the species and interaction β-diversity between all pairs of sites and partitioned total interaction β-diversity into its components of species turnover (changes in species composition) and interaction rewiring (interaction turnover despite species constancy) [@Novotny2009-my; @Poisot2012-fk].
-As an initial analysis, we plotted the relationship between site-wise elevation difference and each metric of β-diversity and verified the significance of the relationship using logistic matrix regression [@Goslee2009-ln], with the difference in number of sampling dates between sites as a covariate to control for potential confounding effects of sampling frequency.
+As an initial analysis, we visualized the relationship between site-wise elevation difference and each metric of β-diversity, fitting each response with a binomial regression smooth.
 
-To investigate the relationship between elevation and β-diversity more deeply, we performed a second analysis using generalized dissimilarity modeling (GDM) [@Ferrier2007-ji] to analyze the elevational variation in the β-diversity of bumble bees, flora, and (unpartitioned) interactions.
-In addition to estimating the relationship between differences in the response variable and differences in the predictor variables, GDM captures the slope of this relationship over the range of each predictor variable, revealing potential variation in the amount of change in the response induced by a given change in a predictor.
+<!-- and modeled each relationship using logistic matrix regression [@Goslee2009-ln].
+Site management difference (i.e.
+same or different management category) was included as a covariate to account for potential management effects, and the difference in number of sampling dates between sites was included as covariate to control for the potential confounding effects of sampling frequency.
+-->
+
+To investigate the relationship between elevation and β-diversity more deeply, we performed a second analysis using generalized dissimilarity modeling (GDM) [@Ferrier2007-ji] to analyze the elevational variation in the β-diversity of bumble bees, flora, and (unpartitioned) interactions, respectively.
+In addition to estimating the relationship between differences in the response variable and differences in the predictor variables, GDM captures the slope of this relationship over the range of each predictor variable, revealing potential variation in the amount of change in the response induced by a given change in a predictor [@Ferrier2007-ji].
 This enabled us to ask whether the relationship between β-diversity and elevation exhibits thresholds or other nonlinearities.
-To control for potential confounding effects of geographic proximity and number of sampling dates per site, these terms were added as covariates in the GDM.
-GDM is designed to model continuous predictors, and while the inclusion of categorical variables is theoretically possible [@Ferrier2007-ji],
+We also included site management in each model by encoding management classes as a series of binary dummy variables, following @Ferrier2007-ji.
+To control for potential confounding effects of the number of sampling days per site, sampling days was included as a covariate.
+The importance of the variables in each model were assessed by a permutation test in which variable importance is defined as the percent loss of explained deviance when a given variable is permuted in isolation [@Fitzpatrick-2021aa].
 
@@
 GAM analyses were performed with packages `mgcv` [@Wood_2017] and `mgcViz` [@Fasiolo2018aa].
-PCoA and species β-diversity calculation were performed with the package `vegan` [@Oksanen_2019] and visualized using the packages `ggplot2` [@Wickham2016-aa] and `ggvegan` [@Simpson2019aa].
+β-diversity was calculated with the package `vegan` [@Oksanen_2019] and visualized using the packages `ggplot2` [@Wickham2016-aa] and `ggvegan` [@Simpson2019aa].
 Calculation and partitioning of interaction β-diversity were performed with the package `bipartite` [@Dormann2008-aa; Dormann2009-aa].
-Matrix regression was performed with the package `ecodist` [@Goslee2007-aa].
 GDM analysis of β-diversity was performed with the package `gdm` [@Fitzpatrick-2021aa].
 Annotated `R` code is available in the Supplementary Material.
+
+<!-- Matrix regression was performed with the package `ecodist` [@Goslee2007-aa]. -->
 
@@
 lapidarius* were recorded only sporadically and at low abundance, so we omitted them from all analyses.
-See Appendix X for a summary of species-wise abundances.
+See Supplementary Material for a summary of species-wise abundances.
 
@@
 
-The peak abundance of all bumble bee species responded significantly to elevation (p ≤ 0.04) **(Figure 2)**.
+For all bumble bee species, the response of peak abundance to elevation differed significantly (p ≤ 0.02) from a flat line **(Figure 2A)**.
 *B.
@@
 monticola*, *B.
-mucidus*, and *B.
-pyrenaeus* exhibited a positive response to elevation and peaked in abundance above the tree line (~ 1500 m.s.l.), consistent with their known affinity for high elevation habitats [@Rasmont2010-ci].
-The abundance of *B.
-soroeensis* also tended to increase with elevation and peaked above the tree line, but it was also abundant at lower elevations.
+mucidus*, *B.
+pyrenaeus*, and *B.
+soroeensis* exhibited a positive response to elevation and peaked in abundance above the tree line (~ 1500 m.s.l.), consistent with their known affinity for high elevation habitats [@Rasmont2010-ci].
 In contrast, the abundance of *B.
 hortorum*, *B.
-pascorum*, and *B.
-psithyrus* declined with elevation.
+pascorum*, *B.
+psithyrus*, and *B.
+terrestris/lucorum* declined with elevation.
 This effect was especially pronounced in *B.
@@
 gerstaeckeri*, *B.
-jonellus*, *B.
-terrestris-lucorum*, and *B.
+jonellus*, and *B.
 wurflenii* peaked at mid-elevation below the tree line, while *B.
 pratorum* was unique in exhibiting a sharp peak of abundance directly *at* the tree line.
+Our model explained 71% of total variance (adjusted R^2^ = 0.45).
+See Supplemental Material for tabulated model output and diagnostics.
 
-<!-- ```{r fig3, echo = FALSE,  fig.cap = paste("Fitted GAM smooths of bumble bee abundance on elevation, faceted by bumble bee species and plotted over original data.
-Data points are color-coded to indicate whether they represent sites above (red) or below (blue) the tree line.
-Note the different scales of the y-axes.")} -->
-<!-- knitr::include_graphics("../output/gam_bb_abund.png") -->
-<!-- ``` -->
-
-```{r fig2, echo = FALSE,  fig.cap = paste("Conditional effect of elevation on the peak abundance of each bumble bee species.
-Effects are plotted on the link scale, and dashed lines depict 95\\% confidence intervals.")}
-knitr::include_graphics("../output/gam_bb_abund_grid.png")
-```
-
-Mean floral resource abundance likewise responded significantly (p << 0.0001) to elevation for all bumble bee species **(Figure 3)**.
+Mean floral resource abundance likewise responded significantly (p << 0.0001) to elevation for all bumble bee species **(Figure 2B)**.
 *B.
+jonellus*, *B.
 mendax*, *B.
@@
 soroeensis*, and *B.
-wurflenii* exhibited a weakly to strongly bimodal pattern of flora resource abundance, with peaks around 1000 m.s.l.
-and 1600 m.s.l., the latter peak corresponding closely to the location of the tree line.
+wurflenii* exhibited bimodal patterns of flora resource abundance, with peaks around 1000 m.s.l.
+and 1600 m.s.l.
+The latter peak corresponds closely to the location of the tree line.
 *B.
-jonellus* and *B.
 pyrenaeus* exhibited a unimodal pattern peaking near 1600 m.s.l.
@@
 psithyrus*, and *B.
-terrestris-lucorum* experienced linear or quadratic decline in floral resource abundance with increasing elevation.
+terrestris/lucorum* experienced linear or quadratic decline in floral resource abundance with increasing elevation.
 Floral resource abundance was low for all species at elevations above ~1600 m.s.l.
+Our model explained 67% of total variance (adjusted R^2^ = 0.41).
+See Supplemental Material for tabulated model output and diagnostics.
 
-# ```{r fig5, echo = FALSE,  fig.cap = paste("Fitted GAM smooths of floral abundance on elevation, faceted by bumble bee species and plotted over original data.
-Data points are color-coded to indicate whether they represent sites above (red) or below (blue) the tree line.
-Note the different scales of the y-axes.")}
-# knitr::include_graphics("../output/gam_fl_abund.png")
-# ```
-
-```{r fig6, echo = FALSE,  fig.cap = paste("Conditional effects of elevation on mean flower cover by bumble bee species and year.
-Effects are plotted on the link scale and dashed lines depict 95\\% confidence intervals.")}
-knitr::include_graphics("../output/gam_fl_abund_grid.png")
+```{r fig2, echo = FALSE,  fig.cap = paste("Conditional effects of elevation an dmanagement on the maximum abundance of each bumble bee species (A) and conditional effects of elevation, management, and year on mean flower cover weighted for each bumble bee species (B).
+Effects are plotted on the link scale, and dashed lines depict 95\\% confidence intervals.
+Rug plots depict the locations of sites along the elevation axis and are color-coded to indicate whether each site fell above or below the tree line.")}
+knitr::include_graphics("../output/gam_abund_grid_mod.png")
 ```
@@
 
-Floral β-diversity among our study sites was very high overall and responded steeply to elevation difference between sites **(Figure 4A)**.
+Floral β-diversity among our study sites was very high overall and responded steeply to elevation difference between sites **(Figure 3A)**.
 Even sites at similar elevation exhibited ~40% species turnover, and the most widely separated sites (elevation difference > 1250 m) differed by more than 85%.
 
-Bumble bee β-diversity was, in comparison to floral β-diversity, both lower overall and less responsive to elevational difference between sites **(Figure 4A)**.
+Bumble bee β-diversity was, in comparison to floral β-diversity, both lower overall and less responsive to elevational difference between sites **(Figure 3A)**.
 Sites at similar elevation exhibited ~25% species turnover, and species turnover between the most widely separated sites remained less than 50%.
 
-Total interaction β-diversity was >75% between sites at similar elevation and approached perfect dissimilarity in the most widely separated sites **(Figure 4B)**.
+Total interaction β-diversity was >75% between sites at similar elevation and approached perfect dissimilarity in the most widely separated sites **(Figure 3B)**.
 Partitioning revealed that total interaction β-diversity was driven primarily by species turnover, which accounted for ~50% of total β-diversity for sites at similar elevation and ~90% of total β-diversity for the most widely separated sites.
-Species turnover consisted mainly of turnover in the floral community **(Figure 4C)**, though the joint turnover of plants and bumble bees accounted for ~25% of total species turnover between the most widely separated sites.
-Interaction rewiring was most significant between sites at similar elevation, where it accounted for ~30% of total β-diversity, but its share of total β-diversity declined to ~10% in the most widely separated sites **(Figure 4B)**.
+Species turnover consisted mainly of turnover in the floral community **(Figure 3C)**, though the joint turnover of plants and bumble bees accounted for ~25% of total species turnover between the most widely separated sites.
+Interaction rewiring was most significant between sites at similar elevation, where it accounted for ~30% of total β-diversity, but its share of total β-diversity declined to ~10% in the most widely separated sites **(Figure 3B)**.
 
-Logistic matrix regression confirmed the significance (p < 0.01) of the relationship between each β-diversity metric and elevation.
-
-```{r fig4, echo = FALSE,  fig.cap = paste("Species-level (A) and interaction-level (B, C) plotted against elevational difference.
+```{r fig3, echo = FALSE,  fig.cap = paste("Species-level (A) and interaction-level (B, C) plotted against elevational difference.
 Each point represents the β-diversity (or β-diversity partition) between a pair of sites, and lines represent the overall relationship between β-diversity and elevation difference using binomial regression smooths.
@@
 
-```{r fig8, echo = FALSE,  fig.cap = paste("GDM splines of species-level and interaction-level β-diversity in response to elevational, controlling for geographic proximity and number of sampling dates.
-The maximum height of each spline represents a given variable’s partial effect on β-diversity (its effect when covariates are held constant), and the shape of each spline represents to the rate of species turnover as it varies along the corresponding gradient, with steeper parts of the curve indicating regions of the gradient over which species/interaction turnover is more rapid.
-Dotted vertical lines in the elevation panels indicates the location of the tree line.
-Within-site comparisons were omitted from GDM analysis to avoid pseudoreplication.
-Shaded bands depict the uncertainty of each partial effect curve (+/- one standard deviation) based on bootstrapping.
-Note the differences in scale between the y-axes of each variable.")}
-knitr::include_graphics("../output/gdm_stack.png")
+```{r fig4, echo = FALSE,  fig.cap = paste("GDM splines of species-level and interaction-level β-diversity in response to elevation (A) and tabulated model output showing variable importance and significance for elevation, sampling frequency, and site management.
+The maximum height of each spline (A) represents the partial effect of elevation on β-diversity (its effect when covariates are held constant), and the shape of each spline represents to the rate of species turnover as it varies along the elevation gradient, with steeper parts of the curve indicating regions of the gradient over which species/interaction turnover is more rapid.
+Rug plots depict the locations of sites along the elevation axis and are color-coded to indicate whether each site fell above or below the tree line.
+Variable importance (B) is mapped to a color ramp, and p-values less than 0.05 are depicted in bold typeface.")}
+knitr::include_graphics("../output/gdm_combo.png")
 ```
@@
 psithyrus*, *B.
-terrestris-lucorum*, *B.
+terrestris/lucorum*, *B.
 wurflenii*) and those with high-elevation abundance peaks (*B.

### 2021-04-08 Dryad_ms1/README.Rmd
finalizing for submission

---
title: "Abundance and beta-diversity of bumble bees and wildflowers in the Berchtesgadener Alps"

author: Katharina Kallnik^1^, Johanna Sieger^3^, Fabrice Requier^2^, Douglas B. Sponsler^1^*,  Alice Claßen^1^, A. Fabienne Maihoff^1^,  Ingolf Steffan-Dewenter^1^

output: pdf_document
---

^1^ Department of Animal Ecology and Tropical Biology, Biocenter, University of Würzburg, Würzburg, Germany

^2^ Université Paris-Saclay, CNRS, IRD, UMR Évolution, Génomes, Comportement et Écologie, 91198, Gif-sur-Yvette, France

^3^ Naturpark Frankenhöhe, Ansbach, Germany

`*` Corresponding author: douglas.sponsler@uni-wuerzburg.de

<!-- # GENERAL INFORMATION -->

<!-- ## Timeframe of data collection -->

<!-- ## Location of data collection -->

<!-- ## Funding sources -->

# DATA FILES

## 1. `floral_tax.csv`

### Description: Supplies family-level taxonomy for the floral taxa in our study.

### Columns:
  1. `plant.sp` = plant species
  2. `plant.genus` = plant genus
  3. `plant.family` = plant family

## 2. `network.csv`

### Description: Bumble- bee-wildflower visitation data used in all analyses of (1) bumble bee abundance, (2) bumble bee beta-diversity, and (3) interaction beta-diversity.

### Columns:
  1. `year` = year
  2. `dayofyear` = day of year since January 1
  3. `site` = site name
  4. `snowcover` = whether site was snow-covered
  10. `plant.sp` = full plant species name
  11. `plant.genus` = plant genus name
  12. `flower cover` = flower cover in m^2^ to the nearest 0.01 m^2^
  12. `day` = day of month
  13. `month` = month
  14. `date` = date

## 3. `floral_survey.csv`

### Description: Floral survey data used in all analysis of (1) floral abundance and (2) floral beta-diversity.

### Columns:
  1. `year` = year
  2. `dayofyear` = day of year since January 1
  3. `site` = site name
  4. `trap.time` = time interaction was observed
  5. `caste` = bumble bee caste (queen, male, worker, NA = parasitic *Psithryus*)
  6. `pollen` = whether the bee had visible pollen loads
  7. `bb.sp` = abbreviated bumble bee species name
  8. `bb.sp.lat` = full bumble bee species name
  9. `plant.sp.abb` = abbreviated plant species name
  10. `plant.sp` = full plant species name
  11. `plant.genus` = plant genus name
  12. `day` = day of month
  13. `month` = month
  14. `date` = date

## 4. `site_data.csv`

### Description: Site data including elevation, management, and geographic coordinates.

### Columns:
 1. `site` = site name
 2. `elev.class` = elevation category (oben, mitte, unten)
 3. `management` = site management (mowing, grasing, none)
 4. `temp.mean` = mean temperature recorded with iButton devices
 5. `elev.mean` = mean elevation
 6. `transect` = transect name
 7. `slope.calc` = slope calculated from min and max elevation
 8. `slope.est` = estimated slope
 9. `elev.min` = minimum elevation within site
 10. `elev.max` = maximum elevation within site
 11. `lat` = latitude (decimal degrees)
 12. `lon` = longitude (decimal degrees)
 13. `elev.class2` = alternative elevation binning
 14. `tree_line` = whether a site was above or below the tree line; only for sites included in analysis (hence NAs)

# CODE FILES

## 1. `ms1_abundance.Rmd`

### Description: R script for running abundance analyses.

## 2. `ms1_beta_diversity.Rmd`

### Description: R script for running beta-diversity analyses.

### 2021-04-08 code/ms1_abundance.Rmd
finalizing for submission

@@
 library(mgcViz)
-library(gratia)
 library(patchwork)
@@
 library(vegan)
-library(ggvegan)
-library(ggrepel)
-library(ggforce)
 library(tidyverse)
@@
 
-## 1.2 Load site and trait data
+## 1.2 Load site data and floral taxonomy data
 ```{r, include=FALSE, echo=FALSE}
 site_data <- read_csv("../data/processed_data/site_data.csv") %>%
-  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
-  mutate(elev.class2 = factor(elev.class2, levels = c("low", "mid", "high")),
-         site = factor(site),
-         tree_line = factor(tree_line)) # turn elev.class into an ordered factor and site into a regular factor
-
-# BB traits
-bb_traits <- read_csv("../data/processed_data/bb_traits.csv")
+  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean, elev.class, elev.class2)) %>% 
+  mutate(site = factor(site),
+         tree_line = factor(tree_line))
 
@@
 fl_tax <- read_csv("../data/processed_data/floral_tax.csv") 
-
-## Kugler morphotypes (accessed through Bioflor)
-fl_ktype <- read_csv("../data/processed_data/floral_k_type.csv")
-
-## Flower colors (accessed through BioFlor)
-fl_color <- read_csv("../data/processed_data/floral_color.csv")
-
-## diameter and corolla depth data from Gita Benadi
-fl_morph <- read_csv("../data/processed_data/flower_morphology_GitaBenadi.csv") %>% 
-  dplyr::select(SampleID, plant.family = Family, plant.sp = Species, 
-                Unit, Flowers_per_unit, diam = Diameter, LN) %>%
-  group_by(plant.sp) %>%
-  summarize(diam = mean(diam), LN = mean(LN))
-
-## verbal description of Kugler morphotypes
-fl_ktype_key <- read_delim("../data/processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
-  dplyr::select(k.type = 1, description = 2)
-
-## join all floral trait tables
-fl_traits <- fl_tax %>%
-  full_join(fl_ktype) %>%
-  full_join(fl_color) %>%
-  full_join(fl_morph) %>%
-  dplyr::select(plant.sp, plant.genus, plant.family, k.type, k.type.s, k.type.ss, color, diam, LN)
 ```
@@
 
+### Identify species that were visited but not recorded in the survey
 vis_sp <- net %>%
@@
          site = factor(site),
-         transect = factor(transect)) %>%
-  left_join(bb_traits) %>%
-  mutate(tongue = factor(pbl.w.class2))
+         transect = factor(transect))
 
@@
   group_by(site, date, plant.sp) %>% # group by plant.sp*transect
-  summarize(fl.abund = sum(flower.cover)) %>% # get plant.sp*transect total flower cover; this step really shouldn't be required, but in a few cases, there is more than one abundance entry for the same species on a given transect; not sure how that happened, but it's in the original data that Fabrice sent me; alternatively, I could drop the dups or average them.
-We'll see what Kalli and Fabrice have to say.
+  summarize(fl.abund = sum(flower.cover)) %>% # get plant.sp*transect total flower cover; this step really shouldn't be required, but in a few cases, there is more than one abundance entry for the same species on a given transect; not sure how that happened, but it's in the original data; alternatively, I could drop the dups or average them.
   group_by(site, date) %>%
@@
   ungroup() %>%
-  #full_join(dplyr::select(net, site, date, plant.sp), by = c("site", "date", "plant.sp")) %>% # add the handful of species occurring in the visitation network but missing from the survey data
   complete(site, date, plant.sp, fill = list(fl.abund = 0, prop.fl.abund = 0)) %>%
@@
   mutate(year = factor(year(date))) %>%
-  left_join(fl_traits)
+  left_join(fl_tax)
 ```
@@
 
-# What happens if I plot max abundance by elevation?
-
-bb_range2 <- net %>%
-  left_join(site_data) %>%
-  group_by(site, date, elev.mean, tree_line, bb.sp) %>%
-  summarize(abund = n()) %>%
-  group_by(site, elev.mean, tree_line, bb.sp) %>%
-  summarize(abund = max(abund))
-
-ggplot(filter(bb_range2, bb.sp != "humi"), aes(elev.mean, abund)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 6)) +
-  facet_wrap(~bb.sp, scales = "free_y", ncol = 4) +
-  geom_vline(xintercept = 1500, color = "red", linetype = "dashed")
-
 # Overall bumble bee abundance
@@
   mutate(mean.prop.abund = mean(prop.abund))
-
-ggplot(bb_abund_sum, aes(site, prop.abund, fill = bb.sp)) +
-  geom_col()
-
-ggplot(bb_abund_sum, aes(bb.sp, mean.abund)) +
-  geom_boxplot()
 
@@
                              management +
-                             s(bb.sp, bs = "re") + # random intercept for bb.sp
-                             s(bb.sp.year, bs = "re") + # random intercept for bb.sp:year
-                             s(elev.mean, by = bb.sp, k = 8), # smoother for elev by bb.sp.
+                             s(bb.sp, bs = "re") + 
+                             s(bb.sp.year, bs = "re") + 
+                             s(elev.mean, by = bb.sp, k = 8), 
                            method = "REML",
@@
 
-### 2.2.3 Evaluate model
+### 2.2.3 Evaluate models
 ```{r}
@@
 ```{r}
+# Plot on the link scale
 ## Individual plots

### 2021-04-08 code/ms1_beta_diversity.Rmd
finalizing for submission

@@
 ---
-title: "Beta Diversity through Time and Space"
+title: "Beta Diversity through Elevation"
 output: html_document
@@
 
-### Create functions for later
+### Define functions
 ```{r}
@@
 
-# Function with betalinkr settings and conversion to tibble()
+# Function for betalinkr settings and conversion to tibble()
 betalinkr_set1 <- function(x) {
@@
 
-# # Binomial smooth function (thanks, Hadley): https://ggplot2.tidyverse.org/reference/geom_smooth.html
-# binomial_smooth <- function(...) {
-#   geom_smooth(method = "glm", method.args = list(family = "binomial"), ...)
-# }
+# Function for converting visitation data into site-pair table for GDM analysis
 
@@
     dplyr::select(bb.sp, plant.sp.abb, site) %>% 
-    distinct() %>%# each unique bb-fl-site-date combo
+    distinct() %>% # each unique bb-fl-site combo
     mutate(present = rep(1, length(.[,1]))) %>% # we will treat abundance as binary
@@
 
-# Create functions for prepping data
+# Function for preparing the floral survey biodata input to gdm's formatsitepair function
 biodata_prep_survey <- function(surv) {
@@
 
+# Function for preparing the floral survey predictor data input to gdm's formatsitepair function
 predData_prep_survey <- function(surv) {
@@
 
-format_set <- function(bio, pred) {
-  formatsitepair(bio, bioFormat = 1, siteColumn = "site", 
-                 dist = "jaccard", XColumn = "x", YColumn = "y", 
-                 predData = pred, abundance = FALSE) %>%
-  filter(s1.elev.mean != s2.elev.mean) # this avoids comparing repeated measures of the same site (taking advantage of the fact that each site has a unique elevation in our dataset)
-}
-
+# Function for preparing the visitation biodata input to gdm's formatsitepair function
 biodata_prep_BB <- function(net) {
@@
 
+# Function for preparing the visitation predictor data input to gdm's formatsitepair function
 predData_prep_BB <- function(net) {
@@
 }
+
+# Function with setting for gdm's formatsitepair function
+format_set <- function(bio, pred) {
+  formatsitepair(bio, bioFormat = 1, siteColumn = "site", 
+                 dist = "jaccard", XColumn = "x", YColumn = "y", 
+                 predData = pred, abundance = FALSE) %>%
+  filter(s1.elev.mean != s2.elev.mean) # this avoids comparing repeated measures of the same site (taking advantage of the fact that each site has a unique elevation in our dataset)
+}
+
+# Function for tabulating gdm.varImp output 
+gdm_tab <- function(x, response) {
+  x %>%
+    as.data.frame() %>%
+    rownames_to_column(var = "predictor") %>%
+    mutate(response = rep(response, n()),
+           Predictor = case_when(
+             predictor == "elev.mean" ~ "elevation",
+             predictor == "samples" ~ "sampling freq.",
+             predictor == "mowing" ~ "mowing",
+             predictor == "grasing" ~ "grasing",
+             predictor == "none" ~ "none")) %>%
+    select(response, Predictor, Importance, pval) %>%
+    tibble()
+}
 ```
@@
 site_data <- read_csv("../data/processed_data/site_data.csv") %>%
-  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
-  mutate(elev.class2 = factor(elev.class2, levels = c("low", "mid", "high")),
-         site = factor(site),
-         tree_line = factor(tree_line)) %>% # turn elav.class into an ordered factor and site into a regular factor
+  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean, elev.class, elev.class2)) %>% #
+  mutate(site = factor(site),
+         tree_line = factor(tree_line)) %>% 
   semi_join(read_csv("../data/processed_data/network.csv"), by = "site") %>% # drop sites that were not sampled
@@
   dplyr::select(-geometry)
+
+# Write to csv for archiving
+write_csv(site_data, "../Dryad_ms1/site_data.csv")
 ```
@@
   filter(caste  != "male") %>%
-  unite(site.year.day, site, year, dayofyear, sep = "_", remove = FALSE) %>% # create site.date field that can be split up again later
-  mutate(year = factor(year)) # convert year to factor
+  unite(site.year.day, site, year, dayofyear, sep = "_", remove = FALSE) %>% 
+  mutate(year = factor(year))
 
@@
   dplyr::select(site, date, plant.sp, flower.cover)
+
+# Write to csv for archiving
+write_csv(net, "../Dryad_ms1//visitation_data.csv")
+write_csv(survey, "../Dryad_ms1//floral_survey_data.csv")
 ```
@@
                                        "OS.repl", "OS.rich")))
+
+# Write to csv for archiving
+write_csv(beta_mit_data, "../Dryad_ms1/partitioned_beta_diversity")
 ```
 
-### Plot interaction beta-diversity against elevation difference and time difference
-#### For elevation difference, we will "control" for time by only considering pairs separated by <= 5 days
-#### For time difference, we will control for elevation by only comparing within site.
+### Plot interaction beta-diversity against elevation difference
 ```{r}
@@
 g1 <- ggplot(filter(species_beta, 
-                    site1 != site2), # only site-dates within 5 days
+                    site1 != site2), # no within-site comparisons
              aes(elev.diff, beta, color = group)) +
   geom_point(alpha = 0.25, pch = 16) +
-  #binomial_smooth(formula = y ~ splines::ns(x, 3)) +
-  #geom_smooth(method = "lm", se = FALSE) +
   geom_smooth(method = "glm", 
@@
               se = FALSE) +
-  #facet_wrap(~group, ncol = 1) +
   xlab("Elevation difference (m)") +
@@
                       site1 != site2 & # no within-site comparisons
-                      metric != "S" & # drop this metric
+                      metric != "S" &
                       metric.class == "aggregate"), 
@@
   geom_point(alpha = 0.20, pch = 16) +
-  #binomial_smooth(formula = y ~ splines::ns(x, 3), se = FALSE) +
-  #geom_smooth(method = "gam", formula = y ~ s(x, k  = 4), se = FALSE) +
   geom_smooth(method = "glm", 
@@
   plot_annotation(tag_levels = c("A", "B", "C")) + 
-  #plot_layout(guides = "collect") & 
   theme_light() 
@@
 
-### MRM analysis
+### MRM analysis (not included in paper)
 ```{r, warning=FALSE}
@@
   ) %>%
-  # data_color(
-  #   columns = vars(b),
-  #   colors = scales::col_numeric(
-  #     palette = c("white", "red"),
-  #     domain = c(-0.01, 0.5)), alpha = 0.75
-  #   ) %>%
   tab_style(
     style = list(
-      #cell_fill(color = "lightcyan"),
       cell_text(weight = "bold")
@@
 # Call GDMs
-gdm_survey <- gdm(survey_sitepair, geo = FALSE)
+gdm_survey <- gdm(sitepair_survey, geo = FALSE)
 summary(gdm_survey)
@@
 ```{r}
-# Function for tabulate gdm.varImp output 
-gdm_tab <- function(x, response) {
-  x %>%
-    as.data.frame() %>%
-    rownames_to_column(var = "predictor") %>%
-    mutate(response = rep(response, n()),
-           Predictor = case_when(
-             predictor == "elev.mean" ~ "elevation",
-             predictor == "samples" ~ "sampling freq.",
-             predictor == "mowing" ~ "mowing",
-             predictor == "grasing" ~ "grasing",
-             predictor == "none" ~ "none")) %>%
-    select(response, Predictor, Importance, pval) %>%
-    tibble()
-}
-
 # Interaction model

### 2021-04-08 ms_1/ms_1.Rmd
finalizing for submission

@@
 This submission uses novel code, which is provided as Supporting Information to be evaluated as part of the submission.
-If the paper is accepted, the code will be archived on Dryad alongside our raw data.
+If the paper is accepted, the code will be archived on Dryad alongside the data.
+
+\newpage
 
@@
 The structuring of biological communities along mountain slopes is complex, and elevational range shifts in response to climate change involve more than merely tracking suitable temperature envelopes.
+When species move, they do so in the context of biological communities, and the outcomes of these movements depend on how and to what extent biotic interactions are reordered.
 Bumble bees (Hymentopera: *Bombus* spp.) are cold-adapted species associated with mountain habitats, and they are already exhibiting upslope range shifts that are expected to result in habitat loss, novel competitive interactions, and the rewiring of pollination networks.
-Interpreting these shifts, however, requires an understanding of the current elevational patterns of bumble bees and the floral mutualists that are being acted upon by climate change.
+Predicting and interpreting these shifts, however, requires an understanding of the current elevational patterns of bumble bees and their floral mutualists that are being acted upon by climate change.
 We recorded bumble- bee-flower interactions over three years along an 1400 m elevational gradient in the German Alps.
@@
 We demonstrate that the tree line ecotone is (1) a distributional interface between low/mid- and high-elevation bumble bee species, (2) a threshold above which floral resource availability sharply decreases, and (3) a zone of accelerated turnover of floral composition and bumble- bee-flower interactions.
-Our results show that linear elevational temperature gradients are ecologically punctuated, and the outcomes of climate-induced range shifts may hinge upon dynamics at the tree line ecotone.
-
-<!-- The structuring of biological communities along mountain slopes is complex, and the elevational range shifts that are now occurring in response to climate change involve more than merely tracking suitable temperature envelopes.
-When species move, they do so in the context of biological communities, and the outcomes of these movements depend on how and to what extent biotic interactions are reordered.
-Bumble bees (Hymentopera: *Bombus* spp.) are cold-adapted species associated with mountain habitats, and they have already exhibited measurable upslope range shifts that are expected to result in habitat loss, novel competitive interactions, and the rewiring of pollination networks.
-Predicting and interpreting these shifts, however, requires an understanding of the current elevational patterns of bumble bees and the floral mutualists that are being acted upon by climate change.
-In this study, we sampled bumble bees, flowering plants, and their interactions over three consecutive years along an 1400 m elevational gradient in the German Alps.
-Using nonlinear modeling techniques, we analyze the elevational patterns of this community at three levels of biological organization: species abundance, species β-diversity, and interaction β-diversity.
-We demonstrate complex, nonlinear responses to elevation at all three levels of organization.
-In particular, we identify the tree line ecotone as (1) a distributional interface between low/mid- and high-elevation bumble bee species, (2) a threshold above which floral resource availability sharply decreases, and (3) a zone of accelerated turnover of flowering plants and bumble- bee-flower interactions.
-The implications of these findings extend beyond the particular case of bumble bees to demonstrate that linear elevational temperature gradients are ecologically punctuated, and the outcomes of climate-induced range shifts with depend on how species and species-interactions change as they cross --- or fail to cross --- thresholds like the tree line ecotone.
--->
+The implications of these findings extend beyond the particular case of bumble bees to demonstrate that linear elevational temperature gradients are ecologically punctuated, and the outcomes of climate-induced range shifts will depend on dynamics at the tree line ecotone.
 
@@
 Mountains are unique ecological theaters in which extrinsic climate dynamics interact with intrinsic climate gradation to shape the distributions of species and biotic interactions [@Telwala2013-sc; @CaraDonna2014-nq; @Miller-Struttmann2014-wz; @Rafferty2020-aj].
-The rate of recent climate change is forcing species to adapt ecologically within the constraints of gene pools shaped by cooler climates [@Visser2008-ct].
+The rate of recent climate change, however, is forcing species to adapt rapidly within the constraints of gene pools shaped by cooler climates [@Visser2008-ct].
 As temperatures rise, species are expected to track suitable conditions to higher elevations, and such effects are already measurable across a broad range of taxa [@Lenoir2015-tj].
@@
 A salient example is the tree line ecotone that marks the elevational transition between forest and grassland.
-At this relatively discrete band within a continuous elevational temperature gradient, plant communities exhibit rapid species turnover [@Descombes2017-id] and declining diversity [@Becker2007-cg] as abiotic conditions become more extreme away from the shelter of tree canopy [@Slatyer1992-ir].
+At this relatively discrete band within a linear elevational temperature gradient, plant communities exhibit rapid species turnover [@Descombes2017-id] and declining diversity [@Becker2007-cg] as abiotic conditions become more extreme away from the shelter of tree canopy [@Slatyer1992-ir].
 
-Bumble bees (Hymenoptera: *Bombus* spp.) are classic model organisms in the field of community ecology, but they have more recently become conservation priorities due to well-documented population declines and vulnerability to climate change [@Goulson2008-oe; @Soroye2020-zb].
+Bumble bees (Hymenoptera: *Bombus* spp.) are classic model organisms in the field of community ecology, and they have more recently become conservation priorities due to well-documented population declines and vulnerability to climate change [@Goulson2008-oe; @Soroye2020-zb].
 Bumble bees are characteristically cold-adapted species [@Heinrich1994-cv], with peak abundance and diversity in mountain ranges and northern latitudes, where they are often the dominant guild of flower-visiting insects and the principal pollinators of entomophilous flora [@Williams1998-xi; @Goulson2010-lo].
@@
 The upward advance of lowland bumble bee species, for example, may introduce novel competitive pressures on highland species and rewire the pollination networks in which they participate [@Nagamitsu2010-wi; @Brosi2013-mu; @Ishii2013-sm], and elevational [@Lenoir2008-yn] and/or phenological [@Pyke2016-ll] shifts in the floral community could generate mismatches between bumble bee species and their floral hosts.
-When range shifts are considered in this complex biotic context, it is especially important to understand nonlinearities or thresholds in species-, community-, and interaction-level responses to elevational climate gradients, such as the floristic threshold effect that has been documented at the tree line ecotone [@Descombes2017-id], since these could produce dramatic ecological consequences that would not be predicted by geometric considerations alone.
+When range shifts are considered in this complex biotic context, it is especially important to understand nonlinearities or thresholds in species-, community-, and interaction-level responses to elevational climate gradients, since these could produce dramatic ecological consequences that would not be predicted by geometric considerations alone.
 
-The outcomes of range shifts in mountain bumble bee and floral communities have obvious significance from a conservation perspective [@Soroye2020-zb], but they also provide a model system in which to study the basic process of community assembly under climate change, a phenomenon of both historical interest and prospective urgency.
-In the present study, set along ~1400 m elevation gradient in the German Alps, we investigate elevational patterns in a community of bumble bees and wildflowers at three levels of biological organization.
+<!-- The outcomes of range shifts in mountain bumble bee and floral communities have obvious significance from a conservation perspective [@Soroye2020-zb], but they also provide a model system in which to study the basic process of community assembly under climate change, a phenomenon of both historical interest and prospective urgency.
+-->
+In the present study, set along a ~1400 m elevation gradient in the German Alps, we investigate elevational patterns in a community of bumble bees and wildflowers at three levels of biological organization.
 First, we ask how elevation shapes the abundances of bumble bee species and their floral resources.
@@
 
-```{r fig1, echo = FALSE,  fig.cap = paste("Study sites in the Berchtesgadener Alps plotted over Google Earth imagery.
-Sites above the tree line are depecited in red and sites below the tree line in yellow.
-The elevation of each site is given in meters above sea level.")}
-knitr::include_graphics("../output/sites_GE.png")
-```
-
 Sampling in each year of the study consisted of repeated visits to each site at approximately weekly intervals.
@@
 During each visit, bumble bees and their activity were recorded during a 50-minute transect walk.
-Bumblebees observed on or a given flower were counted as floral visitors.
+Bumblebees observed on a given flower were counted as floral visitors.
 Bumble bee queens were identified to species level in the field, while workers and males were stored in individually labeled tubes in the freezer for later identification in the laboratory after [@Amiet1996-jz].
-Floral visitation by male bumble bees was recorded during visitation sampling, but we chose to analyze only visitation by queens and workers; males are often found resting on flowers, so their "visits" cannot necessarily be interpreted in the same way as those of queens and workers.
+Floral visitation by male bumble bees was recorded during visitation sampling, but we chose to analyze only visitation by queens and workers; males are often found resting on flowers, so their "visits" cannot always be interpreted in the same way as those of queens and workers.
 In conjunction with visitation observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.01 m^2^.
-Species identification followed @Lauber2007-ha and @Oberndorfer2001-ro.
+Plant species identification followed @Lauber2007-ha and @Oberndorfer2001-ro.
 
@@
 This resulted in a temporally staggered pattern of sampling in 2010 and 2012, while in 2011 a combination of early snowmelt and a late start to field activities resulted in sampling that was approximately aligned in time across the whole elevation gradient.
-See Appendix X for details of sampling patterns.
+See Supplemental Material for details of sampling patterns.
 
@@
 We also opted to treat the cryptic *B.
-terrestris/lucorum* species pair as a single morphospecies, since these two species cannot be visually distinguished, and we similarly pooled representatives of the parasitic *Psithyrus* subgenus --- *B.
+terrestris/lucorum* species pair as a single morphospecies, since these two species cannot be visually distinguished.
+We similarly pooled representatives of the parasitic *Psithyrus* subgenus --- *B.
 barbutellus*, *B.
@@
 
-We analyzed the abundance of bumble bees and floral resources using separate hierarchical generalized additive models (HGAMs) [@Wood2011-pg; @Pedersen2019-kf].
-In both models, we opted to collapse the temporal variation of abundance within sites and focus on elevational patterns.
-For bumble bees, we modeled *peak* abundance at each site as a smooth response to elevation, with bumble bee species as a smoothing factor, site management (grazing, mowing, or unmanaged) as a fixed effect, and species:year as a random intercept term (exploratory analysis indicated that species varied differently across years).
+We analyzed the abundance of bumble bees and floral resources, respectively, using hierarchical generalized additive models (HGAMs) [@Wood2011-pg; @Pedersen2019-kf].
+To focus on elevational patterns, we chose to collapse the temporal variation of abundance within sites into summary values.
+For bumble bees, we modeled *peak* abundance at each site as a smooth response to `elevation`, with bumble bee `species` as a smoothing factor, `management` (grazing, mowing, or unmanaged) as a fixed effect, and `species:year` as a random intercept term (exploratory analysis indicated that species varied differently across years).
 For floral resources, we modeled *mean* flower cover, weighted per bumble bee species (i.e.
-the sum of *visited* genera), as a smooth response to elevation, with management as a fixed effect, and year and bumble bee species as random intercept terms.
+the sum of *visited* genera), as a smooth response to `elevation`, with `management` as a fixed effect, and `year` and bumble bee `species` as random intercept terms.
 For the bumble bee model, we used the negative binomial distribution family (log link function), and for the floral model we used a Tweedie family distribution (log link function).
@@
 
-For our analyses of species and interaction β-diversity, we again opted to focus on elevational patterns by pooling samples within sites, since the sparse observations of individual sampling events tended to inflate β-diversity and its variance to a degree that impaired model fitting and interpretation.
-After pooling, we calculated the species and interaction β-diversity between all pairs of sites and partitioned total interaction β-diversity into its components of species turnover (changes in species composition) and interaction rewiring (interaction turnover despite species constancy) [@Novotny2009-my; @Poisot2012-fk].
+For our analyses of species and interaction β-diversity, we pooled all samples (including across years) within sites, since the sparse observations of individual sampling events tended to inflate β-diversity and its variance to a degree that impaired model fitting and interpretation.
+After pooling, we calculated the species and interaction β-diversity between all pairs of sites using the binary Jaccard index and partitioned total interaction β-diversity into its components of species turnover (changes in species composition) and interaction rewiring (interaction turnover despite species constancy) [@Novotny2009-my; @Poisot2012-fk].
 As an initial analysis, we visualized the relationship between site-wise elevation difference and each metric of β-diversity, fitting each response with a binomial regression smooth.
@@
 To investigate the relationship between elevation and β-diversity more deeply, we performed a second analysis using generalized dissimilarity modeling (GDM) [@Ferrier2007-ji] to analyze the elevational variation in the β-diversity of bumble bees, flora, and (unpartitioned) interactions, respectively.
-In addition to estimating the relationship between differences in the response variable and differences in the predictor variables, GDM captures the slope of this relationship over the range of each predictor variable, revealing potential variation in the amount of change in the response induced by a given change in a predictor [@Ferrier2007-ji].
+In addition to estimating the relationship between differences in a response variable and differences in a predictor variables, GDM captures the slope of this relationship over the range of the predictor variable, revealing potential variation in the amount of change in the response induced by a given change in the predictor [@Ferrier2007-ji].
 This enabled us to ask whether the relationship between β-diversity and elevation exhibits thresholds or other nonlinearities.
@@
 terrestris/lucorum* ("telu"), and *B.
-wurflenii* ("wurf") --- accounted for the bulk of overall bumble bee abundance, while *B.
+wurflenii* ("wurf") accounted for the bulk of overall bumble bee abundance, while *B.
 hortorum* ("hort"), *B.
@@
 pyrenaeus* ("pyre"), and *B.
-gerstaeckeri* ("gers") .
+gerstaeckeri* ("gers") were present at moderate to low abundance.
 *B.
@@
 humilis*, and *B.
-lapidarius* were recorded only sporadically and at low abundance, so we omitted them from all analyses.
+lapidarius* were recorded only sporadically, so we omitted them from abundance analyses but retained them for β-diversity analyses.
 See Supplementary Material for a summary of species-wise abundances.
 
-Floral surveying yielded a total of 352 plant species, representing 191 genera and 52 families.
-Of these, 155 species from 103 genera and 35 families were observed to be visited by bumble bees.
+Floral surveying yielded a total of 352 plant species (192 genera and 52 families).
+Of these, 163 species (111 genera and 37 families) were observed to be visited by bumble bees.
 Eight species --- *Rubus idaeus*, *Rosa canina*, *Juniperus communis*, *Larix decidua*, *Salix* sp., *Caltha palustris*, *Pulmonaria officinalis*, and *Rheum barbarum* --- were recorded in visitation data but not recorded during floral surveying.
@@
 
-For all bumble bee species, the response of peak abundance to elevation differed significantly (p ≤ 0.02) from a flat line **(Figure 2A)**.
+HGAM analysis revealed that the response of peak abundance to elevation differed significantly (p ≤ 0.02) from a flat line for all bumble bee species **(Figure 2A)**.
 *B.
@@
 pyrenaeus*, and *B.
-soroeensis* exhibited a positive response to elevation and peaked in abundance above the tree line (~ 1500 m.s.l.), consistent with their known affinity for high elevation habitats [@Rasmont2010-ci].
+soroeensis* exhibited a positive response to elevation and peaked in abundance above the tree line, consistent with their known affinity for high elevation habitats [@Rasmont2010-ci].
 In contrast, the abundance of *B.
@@
 wurflenii* peaked at mid-elevation below the tree line, while *B.
-pratorum* was unique in exhibiting a sharp peak of abundance directly *at* the tree line.
+pratorum* was unique in exhibiting a sharp peak of abundance *at* the tree line.
 Our model explained 71% of total variance (adjusted R^2^ = 0.45).
+Neither mowing (p = 0.08) nor non-management (p = 0.95) had a significant effect on bumble bee abundance relative to the reference level of grasing management.
+The random effect of species:year was also nonsignificant (p = 0.52).
 See Supplemental Material for tabulated model output and diagnostics.
@@
 soroeensis*, and *B.
-wurflenii* exhibited bimodal patterns of flora resource abundance, with peaks around 1000 m.s.l.
-and 1600 m.s.l.
-The latter peak corresponds closely to the location of the tree line.
+wurflenii* exhibited bimodal patterns of flora resource abundance, with peaks around 1000 and 1600 m.s.l, the latter peak corresponding closely to the location of the tree line.
 *B.
-pyrenaeus* exhibited a unimodal pattern peaking near 1600 m.s.l.
+pyrenaeus* exhibited a unimodal floral resource pattern peaking near 1600 m.s.l.
 The remaining species --- *B.
@@
 Our model explained 67% of total variance (adjusted R^2^ = 0.41).
+Both lack of management (p << 0.0001) and mowing (p << 0.0001) had a significant positive effect on mean flower cover relative to the reference level of grasing management.
+The effect of year was also significant, with 2010 (p << 0.0001) and 2011 (p << 0.0001) exhibiting higher mean flower cover than 2012.
 See Supplemental Material for tabulated model output and diagnostics.
-
-```{r fig2, echo = FALSE,  fig.cap = paste("Conditional effects of elevation an dmanagement on the maximum abundance of each bumble bee species (A) and conditional effects of elevation, management, and year on mean flower cover weighted for each bumble bee species (B).
-Effects are plotted on the link scale, and dashed lines depict 95\\% confidence intervals.
-Rug plots depict the locations of sites along the elevation axis and are color-coded to indicate whether each site fell above or below the tree line.")}
-knitr::include_graphics("../output/gam_abund_grid_mod.png")
-```
 
@@
 Even sites at similar elevation exhibited ~40% species turnover, and the most widely separated sites (elevation difference > 1250 m) differed by more than 85%.
-
 Bumble bee β-diversity was, in comparison to floral β-diversity, both lower overall and less responsive to elevational difference between sites **(Figure 3A)**.
@@
 
-```{r fig3, echo = FALSE,  fig.cap = paste("Species-level (A) and interaction-level (B, C) plotted against elevational difference.
-Each point represents the β-diversity (or β-diversity partition) between a pair of sites, and lines represent the overall relationship between β-diversity and elevation difference using binomial regression smooths.
-Standard errors are not plotted because they would be misleading due to the non-independence inherent to distance matrix regression, but all regressions were significant (p < 0.01).
-Interaction β-diversity (B, C) is partitioned using Poisot et al.’s (2012) notation: WN = unpartitioned β-diversity, ST = β-diversity due to species turnover, OS =  β-diversity due to interaction rewiring, ST.h =  β-diversity due to species turnover in the higher trophic level (bumble bees), ST.l = β-diversity due to species turnover in the lower trophic level (plants), and ST.lh =  β-diversity due to joint species turnover in higher and lower trophic levels.")}
-knitr::include_graphics("../output/beta_stack.png")
-```
-
-GDM analysis reproduced the finding that overall β-diversity was dominated by floral turnover, but it also revealed a strongly nonlinear response of floral and interaction β-diversity to elevation **(Figure 4)**.
-Turnover was high from ~600-1000 m, representing the transition from valley floor to lower slopes, and then leveled off from ~1000-1500 m.
-At around 1500 m, turnover accelerated sharply and remained steep for the remainder of the elevation gradient.
-The partial effects of geographic proximity and sampling intensity were negligible in all cases.
-
-```{r fig4, echo = FALSE,  fig.cap = paste("GDM splines of species-level and interaction-level β-diversity in response to elevation (A) and tabulated model output showing variable importance and significance for elevation, sampling frequency, and site management.
-The maximum height of each spline (A) represents the partial effect of elevation on β-diversity (its effect when covariates are held constant), and the shape of each spline represents to the rate of species turnover as it varies along the elevation gradient, with steeper parts of the curve indicating regions of the gradient over which species/interaction turnover is more rapid.
-Rug plots depict the locations of sites along the elevation axis and are color-coded to indicate whether each site fell above or below the tree line.
-Variable importance (B) is mapped to a color ramp, and p-values less than 0.05 are depicted in bold typeface.")}
-knitr::include_graphics("../output/gdm_combo.png")
-```
+GDM analysis reproduced the finding that overall β-diversity was dominated by floral turnover, but it also revealed a strongly nonlinear response of floral and interaction β-diversity to elevation **(Figure 4A)**.
+Turnover was high from ~600-1000 m, representing the transition from valley to lower slopes, and then leveled off from ~1000-1500 m.
+At around 1500 m --- the elevation of the tree line --- turnover accelerated sharply and remained steep for the remainder of the elevation gradient.
+In contrast, bumble bee turnover was modest and exhibited no marked thresholds.
+Our GDMs explained 65% of total deviance in interaction β-diversity, 74% of floral β-diversity, and 27% of bumble bee β-diversity.
+The conditional effect of elevation was significant (p ~ 0) in all models, and its importance was 1-2 orders of magnitude higher that of sampling frequency and management **(Figure 4B)**.
+Nevertheless, the conditional effect of sampling frequency was significant (p ~ 0) in the interaction β-diversity model, and mowing management was significant (p ~ 0) in the floral β-diversity model.
 
@@
 The core finding of our study, evident in each of our analyses, is that a linear elevation gradient can generate complex and nonlinear biological responses in a mountain bumble- bee-wildflower community in terms of both abundance and β-diversity, and the latter both at the level of community composition and at the level of species interactions.
-As a rule, these responses are punctuated by the tree line ecotone, which constitutes a discrete ecological threshold along the linear elevation gradient of a mountain slope.
+These responses are punctuated by the tree line ecotone, which constitutes a discrete ecological threshold along the linear elevation gradient of a mountain slope.
 
@@
 mucidus* --- the lower peak of floral resource abundance is almost entirely outside their elevational range.
-The tree line, therefore, functions as critical foraging habitat for high-elevation bumble bees.
-Above the tree line, flora resource availability declined sharply as flower-rich grassland and shrubland transitioned to sparse alpine scree.
-The observation that the upslope advancement of alpine tree lines often lags behind climate warming [@Dullinger2004-qu] suggests that high elevation bumble bees could move upslope faster than their floral resources, increasingly confined to resource-poor habitat, though the opposite pattern could also occur if the climate-tracking of bumble bees is limited by other factors, such as nest site availability [@Marshall2020-pc].
-Moreover, several studies have found that the introduction or removal of bumble bee species from a community can have strong effects on the fitness of other species and on patterns of floral visitation [@Nagamitsu2010-wi; @Brosi2013-mu; @Ishii2013-sm]; if elevational range shifts introduce new competitive pressures via the encroachment of low-elevation species into the historic ranges of high-elevation species, this effect might be expected to be most pronounced at the tree line.
+The tree line, therefore, probably functions as critical foraging habitat for high-elevation bumble bees.
+Above the tree line, we found that flora resource availability declines sharply as flower-rich grassland and shrubland transitions to sparse alpine scree.
+The observation that the upslope advancement of alpine tree lines often lags behind climate warming [@Dullinger2004-qu] suggests that high elevation bumble bees could move upslope faster than their floral resources and become increasingly confined to resource-poor habitat, though the opposite pattern could also occur if the climate-tracking of bumble bees is limited by other factors, such as nest site availability [@Marshall2020-pc].
+Several studies have found that the introduction or removal of bumble bee species from a community can have strong effects on the fitness of other species and on patterns of floral visitation [@Nagamitsu2010-wi; @Brosi2013-mu; @Ishii2013-sm].
+If elevational range shifts introduce new competitive pressures via the encroachment of low-elevation species into the historic ranges of high-elevation species, this effect can be expected to be most pronounced at the tree line.
 The significance of the tree line ecotone as a threshold of floral resource availability and as a zone of overlapping bumble bee distributions corroborates the recent findings of @Minachilis2020-xd in the Mediterranean system of Mount Olympus, indicating that these patterns are likely to generalize well across mountain ecosystems.
@@
 While the turnover rate of bumble bee species was low across the whole elevation gradient and showed no response to the tree line, the turnover rate of the floral community was an order of magnitude higher and exhibited a marked inflection point at the tree line, above which it increased steeply.
-This floristic importance of the tree line has been noted in previous work [@Pellissier2010-kg; @Descombes2017-id], but our study provides a functional extension to this pattern, demonstrating that the tree line also marks a zone of rapid acceleration in the turnover of bumble- bee-flower interactions that is evidently driven by the underlying turnover of the floral community.
-The picture that emerges is a striking contrast between a relative stable bumble bee community and an extremely dynamic floral community.
-The dominance that we observed of total interaction β-diversity by floral species turnover is consistent with the findings of @Simanonok2014-ou in the Rocky Mountains of North America, and it is perhaps unsurprising given the exceptional cold-hardiness of bumble bees and the sensitivity of plants to temperature gradients.
-Nevertheless, it highlights the behavioral challenge that bumble bees have presumably faced for the entirety of their evolutionary history as mountain florivores: how to forage efficiently in a diverse floral community that turns over so rapidly.
-Indeed, the remarkable flexibility and intelligence of bumble bee foraging [@Heinrich1979-ep; @Loukola2017-bt], which have been interpreted as side-effects of their sociality [@Dukas1991-ig], could be interpreted alternatively as an adaptation to precisely this problem of foraging in the context of extreme β-diversity.
+This floristic importance of the tree line has been noted in previous work [@Pellissier2010-kg; @Descombes2017-id], but our study provides a functional extension to this pattern, demonstrating that the tree line also marks a zone of rapid acceleration in the turnover of bumble- bee-flower *interactions*, evidently driven by the underlying turnover of the floral community.
+
+The contrast between the relative stability of the bumble bee community and the extreme dynamism of the floral community in our study is striking.
+The dominance of total interaction β-diversity by floral species turnover in our system is consistent with the findings of @Simanonok2014-ou in the Rocky Mountains of North America, and it is perhaps unsurprising given the exceptional cold-hardiness of bumble bees and the sensitivity of plants to temperature gradients.
+Nevertheless, it highlights the behavioral challenge that bumble bees have presumably faced for the entirety of their evolutionary history as mountain florivores: how to forage efficiently in a diverse floral community that exhibits rapid spatial and temporal turnover.
+Indeed, the remarkable flexibility and intelligence of bumble bee foraging [@Heinrich1979-ep; @Loukola2017-bt], which have been interpreted as side-effects of their sociality [@Dukas1991-ig], could alternatively be interpreted as adaptations to precisely this problem of foraging in the context of extreme β-diversity.
 A fascinating question that could not be answered by our sampling approach is whether mountain bumble bees adaptively forage up- or downslope, as suggested by @Lundberg1980-qk.
@@
 
+While elevational patterns were the focus of our study, our analyses also revealed a negative conditional effect of grasing management relative to mowing or non-management.
+This is surprising because other studies have linked agricultural abandonment in the Alps (typically a transition from grasing to non-management) with reduced abundance and diversity of both flora and bumble bees [@Iserbyt2015-jc; @Walcher2017-kw].
+We caution, however, against the over-interpretation of the negative effect of grasing found in our study, since our methods did not focus on the characterization of land use or land use history.
+
+The outcomes of range shifts in mountain bumble bee and floral communities have obvious significance from a conservation perspective [@Soroye2020-zb], but they also provide a model system in which to study the basic process of community assembly under climate change, a phenomenon of both historical interest and prospective urgency.
 The central finding of our study, that elevation gradients are characterized by complex and nonlinear responses of species, communities, and biotic interactions --- punctuated by the tree line --- is almost certainly not an idiosyncrasy of bumble bees and their floral hosts but rather a pattern emerging from fundamental physical constraints that should be expected to obtain broadly across the biotic communities of mountains worldwide.
@@
 
+<div id ="refs"></div>
+
+\newpage
+
+# FIGURE LEGENDS
+
+**Figure 1:** Study sites in the Berchtesgadener Alps plotted over Google Earth imagery.
+Sites above the tree line are depecited in red and sites below the tree line in yellow.
+The elevation of each site is given in meters above sea level.
+
+**Figure 2:** Conditional effects of elevation an dmanagement on the maximum abundance of each bumble bee species (A) and conditional effects of elevation, management, and year on mean flower cover weighted for each bumble bee species (B).
+Effects are plotted on the link scale, and dashed lines depict 95\\% confidence intervals.
+Rug plots depict the locations of sites along the elevation axis and are color-coded to indicate whether each site fell above or below the tree line.
+
+**Figure 3:** Species-level (A) and interaction-level (B, C) plotted against elevational difference.
+Each point represents the β-diversity (or β-diversity partition) between a pair of sites, and lines represent the overall relationship between β-diversity and elevation difference using binomial regression smooths.
+Standard errors are not plotted because they would be misleading due to the non-independence inherent to distance matrix regression, but all regressions were significant (p < 0.01).
+Interaction β-diversity (B, C) is partitioned using Poisot et al.’s (2012) notation: WN = unpartitioned β-diversity, ST = β-diversity due to species turnover, OS =  β-diversity due to interaction rewiring, ST.h =  β-diversity due to species turnover in the higher trophic level (bumble bees), ST.l = β-diversity due to species turnover in the lower trophic level (plants), and ST.lh =  β-diversity due to joint species turnover in higher and lower trophic levels.
+
+**Figure 4:** GDM splines of species-level and interaction-level β-diversity in response to elevation (A) and tabulated model output showing variable importance and significance for elevation, sampling frequency, and site management.
+The maximum height of each spline (A) represents the partial effect of elevation on β-diversity (its effect when covariates are held constant), and the shape of each spline represents to the rate of species turnover as it varies along the elevation gradient, with steeper parts of the curve indicating regions of the gradient over which species/interaction turnover is more rapid.
+Rug plots depict the locations of sites along the elevation axis and are color-coded to indicate whether each site fell above or below the tree line.
+Variable importance (B) is mapped to a color ramp, and p-values less than 0.05 are depicted in bold typeface.
+
+\newpage
+
+# FIGURES
+
+**Figure 1**
+
+```{r fig1, echo = FALSE}
+knitr::include_graphics("../output/sites_GE.png")
+```
+
+\newpage
+
+**Figure 2**
+
+```{r fig2, echo = FALSE}
+knitr::include_graphics("../output/gam_abund_grid_mod.png")
+```
+
+\newpage
+
+**Figure 3**
+
+```{r fig3, echo = FALSE}
+knitr::include_graphics("../output/beta_stack.png")
+```
+
+\newpage
+
+**Figure4**
+
+```{r fig4, echo = FALSE}
+knitr::include_graphics("../output/gdm_combo.png")
+```
+
+\newpage
+
 <!-- ```{r write_citations, cache=FALSE, include=FALSE} -->
@@
 <!-- ``` -->
-
-<div id ="refs"></div>

### 2021-04-14 code/ms1_abundance.Rmd
set up null model approach to preference analysis

@@
 
+saveRDS(gam_max_bb_abund_01, "../output/gam_max_bb_abund_01.rds")
+
 # FL
@@
                            data = abund_sum) %>% getViz()
+
+saveRDS(gam_mean_fl_abund_01, "../output/gam_mean_fl_abund_01.rds")
 ```
@@
 # BB
-check.gamViz(gam_max_bb_abund_01) # evaluate model
+gam_max_bb_abund_01_check <- check.gamViz(gam_max_bb_abund_01) # evaluate model
 gam_max_bb_abund_01_sum <- summary(gam_max_bb_abund_01) # summarize model
@@
 # FL
-check.gamViz(gam_mean_fl_abund_01) # evaluate model
+gam_mean_fl_abund_01_check <- check.gamViz(gam_mean_fl_abund_01) # evaluate model
 gam_mean_fl_abund_01_sum <- summary(gam_mean_fl_abund_01) # summarize model

### 2021-04-14 code/ms2_data.Rmd
set up null model approach to preference analysis

--- 2021-03-25 code/ms2_bias_niche.Rmd
@@
 ```{r, warning=FALSE, include=FALSE, echo = FALSE}
-library(bipartite)
-library(brms)
-library(bayesplot)
-library(tidybayes)
-library(mgcv)
-library(mgcViz)
-library(DHARMa)
-library(modelr)
-library(ggdist)
-library(vegan)
-library(ggvegan)
-library(ggrepel)
-library(tidymv)
-library(patchwork)
-library(tidymodels)
 library(tidyverse)
 library(lubridate)
-
-rstan::rstan_options(auto_write = TRUE)
-options(mc.cores = parallel::detectCores())
 ```
@@
 
-write_csv(prop_visits_ktype, "../hpc_packets/prop_visits_ktype.csv")
+write_csv(prop_visits_ktype, "../data/processed_data/prop_visits_ktype.csv")
 
@@
 
-write_csv(prop_visits_ktype_pooled, "../hpc_packets/prop_visits_ktype_pool.csv")
-
+write_csv(prop_visits_ktype_pooled, "../data/processed_data/prop_visits_ktype_pool.csv")
 ```
-
-## 1.2 Binomial multi-level regression models
-### 1.2.1 Choose priors
-```{r}
-# For starters, I'm just going to replace the flat prior on the fixed effects we a weakly regularizing prior (on the logit scale)
-brm_pvis_prior <- prior(normal(0, 1.5), 
-                        class = b) 
-```
-
-### 1.2.2 Call models
-All these run cleanly.
-Well, brm_pvis_01 throws maximum tree depth warnings, but these are just about efficiency, not validity.
-These warnings are evidently a consequence of increasing adapt_delta to 0.99, which succeeded in getting rid of the divergent transition warnings (those warnings are much more serious and pertain to the validity of the model).
-```{r}
-## Exploring the differential bias of bumble bee species and tongue length classes towards floral morphotypes
-### Bernoulli with binarized visitation
-brm_pvis_00 <- brm(as.integer(visits.bin) ~
-                    bb.sp.abund*perc.flower.cover + # fixed term: because more abundant flowers will tend to be more visited,and more abundance bees visit more
-                    (1 + perc.flower.cover | ktype:bb.sp) + # a random intercept and random slope for ktype:species; I think I need to model this before I can make conclusions about ktype:pbl.class
-                    (1 + perc.flower.cover | ktype:tongue) + # a random intercept and slope for ktype:tongue
-                    (1 | ktype:site), # a varying intercept and slope for ktype:site
-                  family = "bernoulli", 
-                  prior = brm_pvis_prior, # weakly regularizing
-                  file = "../output/brm_pvis_ktype_00",  # for running locally
-                  #file = "brm_pvis_ktype_00", # for running on the HPC
-                  chains = 4,
-                  #cores = 12,
-                  iter = 5000,
-                  data = filter(prop_visits_ktype, !ktype %in% c("0", "9", "10")))
-
-brm_pvis_01 <- brm(as.integer(visits.bin) ~
-                    bb.abund*perc.flower.cover + # fixed term: because more abundant flowers will tend to be more visited,and more abundance bees visit more
-                    (1 + perc.flower.cover | ktype) + # a random intercept and slope for ktype
-                    (1 | ktype:site),  # a varying intercept term for ktype:site
-                  family = "bernoulli", 
-                  prior = brm_pvis_prior, # weakly regularizing
-                  control = list(adapt_delta = 0.99), # the default setting yielded divergent transitions
-                  file = "../output/brm_pvis_ktype_01",  # for running locally
-                  #file = "brm_pvis_ktype_01", # for running on the HPC
-                  chains = 4,
-                  #cores = 12,
-                  iter = 5000,
-                  data = filter(prop_visits_ktype_pooled, !ktype %in% c("0", "9", "10")))
-
-# If I want to make inference about species visitation rates, I don't need the tongue-length class effect; I just need the site effect.
-But, when I want to make inference about the tongue length classes, I do need the species-level effect.
-brm_pvis_02 <- brm(as.integer(visits.bin) ~
-                    bb.sp.abund*perc.flower.cover + # fixed term: because more abundant flowers will tend to be more visited,and more abundance bees visit more
-                    (1 + perc.flower.cover | ktype:bb.sp) + # a random intercept and random slope for ktype:species; I think I need to model this before I can make conclusions about ktype:pbl.class
-                    (1 | ktype:site), # a varying intercept and slope for ktype:site
-                  family = "bernoulli", 
-                  prior = brm_pvis_prior, # weakly regularizing
-                  file = "../output/brm_pvis_ktype_02",  # for running locally
-                  #file = "brm_pvis_ktype_02", # for running on the HPC
-                  chains = 4,
-                  #cores = 12,
-                  iter = 5000, # Bulk ESS was a bit low wit 2000 iter
-                  data = filter(prop_visits_ktype, !ktype %in% c("0", "9", "10")))
-```
-
-## 1.3 Evaluate models
-```{r eval=FALSE, include=FALSE}
-# Summary
-summary(brm_pvis_00)
-summary(brm_pvis_01)
-summary(brm_pvis_02)
-
-# Posterior predictive checks
-pp_check(brm_pvis_00)
-pp_check(brm_pvis_01)
-pp_check(brm_pvis_02)
-
-# MCMC convergence
-plot(brm_pvis_00)
-plot(brm_pvis_01)
-plot(brm_pvis_02)
-
-# Bayes R2
-bayes_R2(brm_pvis_00)
-bayes_R2(brm_pvis_01)
-bayes_R2(brm_pvis_02)
-```
-
-## 1.4 Preliminary visualization
-```{r eval=FALSE, include=FALSE}
-get_variables(brm_pvis_00)
-conditional_effects(brm_pvis_00)
-mcmc_plot(brm_pvis_00, pars = c("^r_ktype:tongue.*Intercept]$"))
-mcmc_plot(brm_pvis_00, pars = c("^r_ktype:tongue.*perc.flower.cover]$"))
-mcmc_plot(brm_pvis_00, pars = c("^r_ktype:bb.sp.*Intercept]$"))
-mcmc_plot(brm_pvis_00, pars = c("^r_ktype:bb.sp.*perc.flower.cover]$"))
-mcmc_plot(brm_pvis_00, pars = c("^r_ktype:site.*Intercept]$"))
-mcmc_plot(brm_pvis_00, pars = c("^r_ktype:site.*perc.flower.cover]$"))
-
-get_variables(brm_pvis_01)
-conditional_effects(brm_pvis_01)
-mcmc_plot(brm_pvis_01, pars = c("^r_ktype*"))
-mcmc_plot(brm_pvis_01, pars = c("^r_ktype:site.*Intercept]$"))
-mcmc_plot(brm_pvis_01, pars = c("^r_ktype:site.*Intercept]$"))
-mcmc_plot(brm_pvis_01, pars = c("^r_ktype:site.*perc.flower.cover]$"))
-
-get_variables(brm_pvis_02)
-conditional_effects(brm_pvis_02)
-mcmc_plot(brm_pvis_02, pars = c("^r_ktype"))
-mcmc_plot(brm_pvis_02, pars = c("^r_ktype:bb.sp.*Intercept]$"))
-mcmc_plot(brm_pvis_02, pars = c("^r_ktype:site.*Intercept]$"))
-mcmc_plot(brm_pvis_02, pars = c("^r_ktype:bb.sp.*perc.flower.cover]$"))
-mcmc_plot(brm_pvis_02, pars = c("^r_ktype:site.*Intercept]$"))
-mcmc_plot(brm_pvis_02, pars = c("^r_ktype:site.*perc.flower.cover]$"))
-```
-
-## 1.5 Final visualization
-### 1.5.1 morphotype:tongue intercept
-```{r}
-brm_pvis_00_plotframe1 <- brm_pvis_00 %>%
-  spread_draws(b_Intercept, `r_ktype:tongue`[condition, term]) %>% # the back ticks are necessary to escape the :
-  rename(r_ktype.tongue = `r_ktype:tongue`) %>% # get rid of the problematic :
-  filter(term == "Intercept") %>% # drop slope terms for now
-  mutate(condition_mean = b_Intercept + r_ktype.tongue) %>% # get conditional mean
-  group_by(condition) %>% # group by tongue:morpotype pair
-  point_interval(condition_mean, .width = c(0.95, 0.66)) %>% # calculate credible intervals
-  mutate(condition_mean_trans = exp(condition_mean), # exponentiating the estimates gets us back to the response scale (negbinomial uses log link function)
-         .lower_trans = exp(.lower),
-         .upper_trans = exp(.upper)) %>% 
-separate(condition, c("ktype", "tongue"), sep = "_", remove = FALSE) %>% # separate condition term into ktype and tongue
-  mutate(ktype = case_when( # rename
-    ktype == "1" ~ "Disc (1)",
-    ktype == "2" ~ "Funnel (2)",
-    ktype == "3" ~ "Bell (3)",
-    ktype == "4" ~ "Stalk-disc (4)",
-    ktype == "5" ~ "Lip (5)",
-    ktype == "6" ~ "Flag (6)",
-    ktype == "7" ~ "Head (7)")
-  ) %>%
-  mutate(ktype = factor(ktype, levels = c("Disc (1)", "Funnel (2)", # relevel
-                                          "Bell (3)", "Stalk-disc (4)", "Lip (5)", 
-                                          "Flag (6)", "Head (7)"))) %>%
-  mutate(tongue = factor(tongue, levels = c("long", "med", "short")))
-
-ggplot(brm_pvis_00_plotframe1, aes(condition_mean_trans, rev(tongue), 
-                                 xmin = .lower_trans, xmax = .upper_trans,
-                                 color = tongue, fill = tongue)) +
-  geom_pointinterval() +
-  facet_wrap(~ktype, scales = "free") +
-  theme_light() +
-  labs(color = "Tongue length", x = "Visitation probability") +
-  guides(fill = FALSE) +
-  theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank())
-
-ggsave(file = "../output/brm_pvis_00_plot1.pdf", width = 6, height = 5)
-ggsave(file = "../output/brm_pvis_00_plot1.png", width = 6, height = 5)
-```
-
-### 1.5.2 morphotype:bb.sp intercept
-```{r}
-brm_pvis_02_plotframe2 <- brm_pvis_02 %>%
-  spread_draws(b_Intercept, `r_ktype:bb.sp`[condition, term]) %>% # the back ticks are necessary to escape the :
-  rename(r_ktype.bb.sp = `r_ktype:bb.sp`) %>% # get rid of the problematic :
-  filter(term == "Intercept") %>% # drop slope terms for now
-  mutate(condition_mean = b_Intercept + r_ktype.bb.sp) %>% # get conditional mean
-  group_by(condition) %>% # group by bb.sp:morpotype pair
-  point_interval(condition_mean, .width = c(0.95, 0.66)) %>% # calculate credible intervals
-  mutate(condition_mean_trans = inv_logit_scaled(condition_mean), # exponentiating the estimates gets us back to the response scale (negbinomial uses log link function)
-         .lower_trans = inv_logit_scaled(.lower),
-         .upper_trans = inv_logit_scaled(.upper)) %>% 
-separate(condition, c("ktype", "bb.sp"), sep = "_", remove = FALSE) %>% # separate condition term into ktype and bb.sp
-  mutate(ktype = case_when( # rename
-    ktype == "1" ~ "Disc (1)",
-    ktype == "2" ~ "Funnel (2)",
-    ktype == "3" ~ "Bell (3)",
-    ktype == "4" ~ "Stalk-disc (4)",
-    ktype == "5" ~ "Lip (5)",
-    ktype == "6" ~ "Flag (6)",
-    ktype == "7" ~ "Head (7)")
-  ) %>%
-   mutate(ktype = factor(ktype, levels = c("Disc (1)", "Funnel (2)", # relevel
-                                          "Bell (3)", "Stalk-disc (4)", "Lip (5)", 
-                                          "Flag (6)", "Head (7)"))) %>%
-  left_join(bb_traits) %>%
-  mutate(bb.sp = factor(bb.sp, levels = c("telu", "soro", "pyre", "psit", "prat", "lapi", "jone",
-                                           "wurf", "pasc", "muci", "mont", "hypn",
-                                           "mend", "hort", "gers")))
-
-ggplot(brm_pvis_02_plotframe2, aes(condition_mean_trans, bb.sp, 
-                                 xmin = .lower_trans, xmax = .upper_trans,
-                                 color = pbl.w.class2, fill = pbl.w.class2)) +
-  geom_pointinterval() +
-  facet_wrap(~ktype, scales = "free") +
-  theme_light() +
-  labs(color = "Tongue length", x = "Visitation probability", y = "Morphotype") +
-  guides(fill = FALSE)
-
-ggsave(file = "../output/brm_pvis_02_plot1.pdf", width = 8, height = 8)
-ggsave(file = "../output/brm_pvis_02_plot1.png", width = 8, height = 8)
-```
-
-### 1.5.3 morphotype:tongue slope
-```{r eval=FALSE, include=FALSE}
-brm_pvis_00_plotframe3 <- brm_pvis_00 %>%
-  spread_draws(b_perc.flower.cover, `r_ktype:tongue`[condition, term]) %>% # the back ticks are necessary to escape the :
-  rename(r_ktype.tongue = `r_ktype:tongue`) %>%
-  filter(term == "perc.flower.cover") %>%
-  mutate(condition_slope = b_perc.flower.cover + r_ktype.tongue) %>%
-  separate(condition, c("ktype", "tongue"), sep = "_", remove = FALSE) %>%
-  mutate(ktype = case_when(
-    ktype == "1" ~ "Disc (1)",
-    ktype == "2" ~ "Funnel (2)",
-    ktype == "3" ~ "Bell (3)",
-    ktype == "4" ~ "Stalk-disc (4)",
-    ktype == "5" ~ "Lip (5)",
-    ktype == "6" ~ "Flag (6)",
-    ktype == "7" ~ "Head (7)")
-  ) %>%
-  mutate(ktype = factor(ktype,
-                        levels = c("Disc (1)", "Funnel (2)", "Bell (3)", "Stalk-disc (4)",
-                                   "Lip (5)", "Flag (6)", "Head (7)")))
-
-ggplot(brm_pvis_00_plotframe3, aes(condition_slope, tongue, color = tongue)) +
-  stat_pointinterval() +
-  facet_grid(ktype~.) +
-  theme_light() +
-  labs(color = "Tongue length", x = "Slope of logit-likelihood") +
-  guides(fill = FALSE) +
-  theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank())
-
-ggsave(file = "../output/brm_pvis_00_plot3.png", width = 7, height = 5.5)
-```
-
-### 1.5.3 morphotype intercept
-```{r}
-brm_pvis_01_plotframe1 <- brm_pvis_01 %>%
-  spread_draws(b_Intercept, r_ktype[condition, term]) %>% # the back ticks are necessary to escape the :
-  filter(term == "Intercept") %>% # drop slope terms for now
-  mutate(condition_mean = b_Intercept + r_ktype) %>% # get conditional mean
-  group_by(condition) %>% # group by tongue:morpotype pair
-  point_interval(condition_mean, .width = c(0.95, 0.66)) %>% # calculate credible intervals
-  mutate(condition_mean_trans = inv_logit_scaled(condition_mean), # exponentiating the estimates gets us back to the response scale (negbinomial uses log link function)
-         .lower_trans = inv_logit_scaled(.lower),
-         .upper_trans = inv_logit_scaled(.upper)) %>% 
-  mutate(ktype = case_when( # rename
-    condition == "1" ~ "Disc (1)",
-    condition == "2" ~ "Funnel (2)",
-    condition == "3" ~ "Bell (3)",
-    condition == "4" ~ "Stalk-disc (4)",
-    condition == "5" ~ "Lip (5)",
-    condition == "6" ~ "Flag (6)",
-    condition == "7" ~ "Head (7)")
-  ) %>%
-  mutate(ktype = factor(ktype, levels = c("Head (7)", "Flag (6)", # relevel
-                                          "Lip (5)", "Stalk-disc (4)", "Bell (3)", 
-                                          "Funnel (2)", "Disc (1)")))
-
-ggplot(brm_pvis_01_plotframe1, aes(condition_mean_trans, ktype, 
-                                 xmin = .lower_trans, xmax = .upper_trans)) +
-  geom_pointinterval() +
-  theme_light(16) +
-  labs(x = "Visitation probability", y= "Morphotype") +
-  guides(fill = FALSE)
-
-ggsave(file = "../output/brm_pvis_01_plot1.pdf", width = 6, height = 7)
-ggsave(file = "../output/brm_pvis_01_plot1.png", width = 6, height = 7)
-```
-
-### 1.5.3 morphotype slope
-```{r eval=FALSE, include=FALSE}
-brm_pvis_01_plotframe2 <- brm_pvis_01 %>%
-  spread_draws(b_perc.flower.cover, r_ktype[condition, term]) %>% # the back ticks are necessary to escape the :
-  filter(term == "perc.flower.cover") %>%
-  mutate(condition_slope = b_perc.flower.cover + r_ktype) %>%
-  mutate(ktype = case_when(
-    condition == "0" ~ "NA (0)",
-    condition == "1" ~ "Disc (1)",
-    condition == "2" ~ "Funnel (2)",
-    condition == "3" ~ "Bell (3)",
-    condition == "4" ~ "Stalk-disc (4)",
-    condition == "5" ~ "Lip (5)",
-    condition == "6" ~ "Flag (6)",
-    condition == "7" ~ "Head (7)")
-  ) %>%
-  mutate(ktype = factor(ktype,
-                        levels = c("Disc (1)", "Funnel (2)", "Bell (3)", "Stalk-disc (4)",
-                                   "Lip (5)", "Flag (6)", "Head (7)", "NA (0)")))
-
-ggplot(brm_pvis_01_plotframe2, aes(condition_slope, ktype)) +
-  stat_pointinterval() +
-  theme_light() +
-  labs(color = "Tongue length", x = "Slope of visitation probability ~ percent flower cover") +
-  guides(fill = FALSE)
-
-ggsave(file = "../output/brm_pvis_01_plot2.png", width = 7, height = 5.5)
-```
-
-## 1.6 Model 1 hypothesis tests
-```{r}
-get_variables(brm_pvis_01)
-
-hypothesis(brm_pvis_01, "r_ktype[1,Intercept] = r_ktype[4,Intercept]", class = "") 
-hypothesis(brm_pvis_01, "r_ktype[2,Intercept] = r_ktype[4,Intercept]", class = "")
-
-hypothesis(brm_pvis_01, "r_ktype[2,Intercept] = r_ktype[1,Intercept]", class = "")
-hypothesis(brm_pvis_01, "r_ktype[1,Intercept] = r_ktype[3,Intercept]", class = "") 
-hypothesis(brm_pvis_01, "r_ktype[2,Intercept] = r_ktype[3,Intercept]", class = "")
-hypothesis(brm_pvis_01, "r_ktype[4,Intercept] = r_ktype[3,Intercept]", class = "")
-
-hypothesis(brm_pvis_01, "r_ktype[5,Intercept] = r_ktype[6,Intercept]", class = "") 
-hypothesis(brm_pvis_01, "r_ktype[5,Intercept] = r_ktype[7,Intercept]", class = "") 
-```
-
-## 1.7 Model 2 hypothesis tests
-```{r}
-get_variables(brm_pvis_00)
-
-# Disc (group  1)
-hypothesis(brm_pvis_00, "r_ktype:tongue[1_short,Intercept] = r_ktype:tongue[1_med,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[1_short,Intercept] = r_ktype:tongue[1_long,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[1_med,Intercept] = r_ktype:tongue[1_long,Intercept]", class = "")
-
-# Funnel (group  2)
-hypothesis(brm_pvis_00, "r_ktype:tongue[2_short,Intercept] = r_ktype:tongue[2_med,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[2_short,Intercept] = r_ktype:tongue[2_long,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[2_med,Intercept] = r_ktype:tongue[2_long,Intercept]", class = "")
-
-# Bell (group  3)
-hypothesis(brm_pvis_00, "r_ktype:tongue[3_short,Intercept] = r_ktype:tongue[3_med,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[3_short,Intercept] = r_ktype:tongue[3_long,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[3_med,Intercept] = r_ktype:tongue[3_long,Intercept]", class = "")
-
-# Stalk-disc (group  4)
-hypothesis(brm_pvis_00, "r_ktype:tongue[4_short,Intercept] = r_ktype:tongue[4_med,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[4_short,Intercept] = r_ktype:tongue[4_long,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[4_med,Intercept] = r_ktype:tongue[4_long,Intercept]", class = "")
-
-# Lip (group  5)
-hypothesis(brm_pvis_00, "r_ktype:tongue[5_short,Intercept] = r_ktype:tongue[5_med,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[5_short,Intercept] = r_ktype:tongue[5_long,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[5_med,Intercept] = r_ktype:tongue[5_long,Intercept]", class = "")
-
-# Flag (group  6)
-hypothesis(brm_pvis_00, "r_ktype:tongue[6_short,Intercept] = r_ktype:tongue[6_med,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[6_short,Intercept] = r_ktype:tongue[6_long,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[6_med,Intercept] = r_ktype:tongue[6_long,Intercept]", class = "")
-
-# Head (group  7)
-hypothesis(brm_pvis_00, "r_ktype:tongue[7_short,Intercept] = r_ktype:tongue[7_med,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[7_short,Intercept] = r_ktype:tongue[7_long,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[7_med,Intercept] = r_ktype:tongue[7_long,Intercept]", class = "")
-```
-
-# 2. Ordination of dietary niche
-Aggregation changes meaning.
-Don't forget that.
-When I pool visitation across dates, I am introducing the effects of relative abundance and floral phenology.
-Species in bloom when BB abundance is high will be over-represented, and vice-versa.
-And species with a long bloom time will be over-represented, and vice-versa.
-This isn't necessarily a problem.
-I just need to make sure I interpret the results the right way.
-When I pool data like this, I am comparing the "net" diet of species, which includes the effects of abundance and phenology in addition to the effects of floral preferences.
-I think that is appropriate for this stage of the analysis, but I need to be careful not to assume that what I'm doing here is the same sort of analysis as what I did above without pooling samples.
-```{r}
-# Fl traits with abbreviated names
-fl_traits2 <- fl_traits %>%
-  left_join(read_csv("../data/processed_data/network.csv")) %>%
-  #select(plant.sp, plant.sp.abb, plant.genus, plant.family, k.type.ss, color) %>%
-  select(plant.genus, k.type.ss) %>%
-  mutate(k.type.ss = factor(k.type.ss)) %>%
-  distinct()
-
-# Totally aggregated; note I'm using the full data, not the aligned and trimmed data. No need for complementary floral survey data.
-visitation2 <- read_csv("../data/processed_data/network.csv") %>%
-  filter(caste != "male") %>% # we will ignore males 
-  filter(!bb.sp %in% c("humi")) %>% # humi is too rare
-  #group_by(bb.sp, plant.sp.abb) %>%
-  group_by(bb.sp, plant.genus) %>%
-  summarize(visits = n()) %>%
-  group_by(bb.sp) %>%
-  mutate(prop.visits = visits/sum(visits)) %>%
-  select(-visits) %>%
-  #pivot_wider(names_from = plant.sp.abb, values_from = prop.visits, values_fill = 0) %>%
-  pivot_wider(names_from = plant.genus, values_from = prop.visits, values_fill = 0) %>%
-  column_to_rownames("bb.sp")
-
-bb_traits2 <- bb_traits %>%
-  filter(!bb.sp %in% c("humi")) %>% # drop humi
-  select(bb.sp, pbl.w.class2) %>%
-  arrange(bb.sp) %>%
-  mutate(pbl.w.class2.pool = case_when(
-    pbl.w.class2 == "short" ~ "short",
-    pbl.w.class2 == "med" ~ "med/long",
-    pbl.w.class2 == "long" ~ "med/long"
-  ))
-
-# Generate ordination
-pcoa_horn2 <- capscale(visitation2 ~ 1, distance = "horn")
-pcoa_horn2_adonis <- adonis(visitation2 ~ pbl.w.class2, data = bb_traits2)
-pcoa_horn2_betadisper <- betadisper(vegdist(visitation2, method = "horn"), 
-                                    group = bb_traits2$pbl.w.class2,
-                                    type = "median")
-anova(pcoa_horn2_betadisper) # I think this is not significant because short tongued species are disperse on axis 2, but axis 2 does not explain nearly as much variation.
-plot(pcoa_horn2_betadisper)
-# I'm surprised that the betadisper test is not significant, but I think it might be driven by the fact that replication is so low for long and medium tongue species.
-Let's see what happens if we pool long and medium tongue species, since they overlap so much anyway.
-pcoa_horn2_betadisper_pool <- betadisper(vegdist(visitation2, method = "horn"), 
-                                    group = bb_traits2$pbl.w.class2.pool,
-                                    type = "median")
-anova(pcoa_horn2_betadisper_pool) # I think this is not significant because short tongued species are dispersed on axis 2, but axis 2 does not explain nearly as much variation.
-plot(pcoa_horn2_betadisper_pool)
-
-# Get variance explained
-pcoa_horn2_mds1 <- round(pcoa_horn2$CA$eig[[1]] / sum(pcoa_horn2$CA$eig)*100)
-pcoa_horn2_mds2 <- round(pcoa_horn2$CA$eig[[2]] / sum(pcoa_horn2$CA$eig)*100)
-
-# Fortify plants
-pcoa_horn2.fort1 <- fortify(pcoa_horn2, axes = 1:2) %>%
-  filter(Score == "species") %>%
-  #rename(plant.sp.abb = Label) %>%
-  rename(plant.genus = Label) %>%
-  left_join(fl_traits2)
-
-# Fortify bees
-pcoa_horn2.fort2 <- fortify(pcoa_horn2, axes = 1:2) %>%
-  filter(Score == "sites") %>%
-  rename(bb.sp = Label) %>%
-  left_join(bb_traits) %>%
-  mutate(pbl.w.class2 = factor(pbl.w.class2))
-
-# Generate convex hulls around tongue length classes
-pcoa_horn2_hulls <- pcoa_horn2.fort2 %>%
-  group_by(pbl.w.class2) %>%
-  slice(chull(MDS1, MDS2))
-
-# Plot
-ggplot(pcoa_horn2.fort2, aes(MDS1, MDS2, color = pbl.w.class2)) +
-  geom_polygon(data = pcoa_horn2_hulls, 
-               aes(fill = pbl.w.class2, color = NULL),
-               alpha = 0.25) +
-  geom_point(data = filter(pcoa_horn2.fort1, abs(MDS1) >= 0.20 | abs(MDS2) >= 0.20),
-            aes(MDS1, MDS2, shape = k.type.ss), inherit.aes = FALSE, color = "gray40", size = 2) +
-  geom_text(aes(label = bb.sp), key_glyph = "rect", size = 3) +
-  geom_text_repel(data = filter(pcoa_horn2.fort1, abs(MDS1) >= 0.20 | abs(MDS2) >= 0.20), 
-            aes(MDS1, MDS2, label = plant.genus), inherit.aes = FALSE, color = "gray40",
-            size = 3) +
-  scale_shape_discrete() +
-  xlab(paste("MDS1 ", "(", pcoa_horn2_mds1, "%", ")", sep = "")) +
-  ylab(paste("MDS2 ", "(", pcoa_horn2_mds2, "%", ")", sep = "")) +
-  labs(fill = "Tongue length", 
-       shape = "Morphotype") +
-  guides(color = FALSE) +
-  coord_fixed() +
-  theme_light(9)
-
-ggsave("../output/pcoa_aggregate.png")
-```
-
-# 3. Dietary niche overlap (i.e. potential competition)
-
-## 3.1 Check BB distribution patterns to decide whom to retain in the analysis
-```{r}
-bb_distribution <- visitation %>%
-  mutate(year = year(date)) %>%
-  left_join(site_data) %>%
-  group_by(site, date, elev.mean, bb.sp, year) %>%
-  summarize(abund = n()) %>%
-  group_by(site, elev.mean, bb.sp, year) %>%
-  summarize(abund = mean(abund)) %>%
-  group_by(bb.sp) %>%
-  mutate(elev.floor = min(elev.mean),
-         elev.ceiling = max(elev.mean),
-         elev.range = elev.ceiling - elev.floor,
-         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))
-
-ggplot(bb_distribution, aes(reorder(bb.sp, elev.med), elev.mean)) +
-  geom_line(size = 2, alpha = 0.25) +
-  geom_point(aes(size = abund), alpha = 0.5) +
-  facet_wrap(~year, ncol = 1) +
-  theme_light(12) +
-  scale_size_continuous(name = "Abundance") +
-  xlab("BB species") +
-  ylab("Elevation") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-bb_distribution_pool <- visitation %>%
-  left_join(site_data) %>%
-  group_by(site, date, elev.mean, bb.sp) %>%
-  summarize(abund = n()) %>%
-  ungroup() %>%
-  complete(site, date, elev.mean, bb.sp, fill = list(abund = 0)) %>%
-  group_by(site, elev.mean, bb.sp) %>%
-  summarize(abund = mean(abund)) %>%
-  filter(abund > 0) %>%
-  group_by(bb.sp) %>%
-  mutate(elev.floor = min(elev.mean),
-         elev.ceiling = max(elev.mean),
-         elev.range = elev.ceiling - elev.floor,
-         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))
-
-ggplot(bb_distribution_pool, aes(reorder(bb.sp, elev.med), elev.mean)) +
-  geom_line(size = 2, alpha = 0.25) +
-  geom_point(aes(size = abund), alpha = 0.5) +
-  theme_light(14) +
-  scale_size_continuous(name = "Abundance") +
-  xlab("BB species") +
-  ylab("Elevation") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-ggsave("../output/bb_range_ms2.pdf", width = 6, height = 5)
-```
-
-## 3.2 Prep data for calculating niche overlap
-```{r}
-niche_data <- visitation %>%
-  left_join(fl_traits) %>%
-  mutate(year = factor(year(date)))
-```
-
-## 3.3 Function for calculating niche overlap
-```{r}
-# This one averages visitation and then calculates niche overlaps
-niche_overlapper <- function(x, criterion) {
-
-  criterion <- enquo(criterion) # criterion is set to either plant.sp (for taxonomic) or k.type.s (for morphological)
-
-  x %>%
-    #group_by(bb.sp, !!criterion, site, date, year) %>% # group by site-year and BB-FL pair
-    group_by(bb.sp, !!criterion, site, date) %>% # group by site-year and BB-FL pair
-    summarize(freq = sum(visits)) %>% # get interaction frequency for each BB-FL pair per site-year
-    group_by(bb.sp, site, date) %>% # group by site-year and BB species
-    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
-    #group_by(bb.sp, site, year, !!criterion) %>%
-    group_by(bb.sp, site, !!criterion) %>%
-    summarize(mean.prop = mean(prop)) %>%
-    pivot_wider(names_from = !!criterion, values_from = mean.prop, values_fill = 0) %>% # spread into the kind of array vegan likes
-    #group_by(site, year) %>%
-    group_by(site) %>%
-    nest() %>% # nest into list-column
-    mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
-           dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
-           tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
-    #dplyr::select(site, year, tidy_dist) %>% # drop intermediate list-columns
-    dplyr::select(site, tidy_dist) %>% # drop intermediate list-columns
-    unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
-    left_join(site_data) %>% # add site data
-    # dplyr::select(site, year, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
-    dplyr::select(site, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean, management) %>% # dplyr::select and rename vars
-    left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
-    rename(subgenus1 = subgenus, tongue1 = pbl.w.class2) %>%
-    left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
-    rename(subgenus2 = subgenus, tongue2 = pbl.w.class2) %>%
-    dplyr::select(site, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
-       tongue1, tongue2, elev.mean, management) %>%
-    mutate(tongue.diff = factor(if_else(tongue1 == tongue2, "same", "different")),
-           tongue.pair = factor(case_when(
-             tongue1 == "med" & tongue2 == "med" ~ "med:med",
-             tongue1 == "med" & tongue2 == "long" ~ "med:long",
-             tongue1 == "med" & tongue2 == "short" ~ "short:med",
-             tongue1 == "long" & tongue2 == "med" ~ "med:long",
-             tongue1 == "short" & tongue2 == "med" ~ "short:med",
-             tongue1 == "long" & tongue2 == "short" ~ "short:long",
-             tongue1 == "short" & tongue2 == "long" ~ "short:long",
-             tongue1 == "long" & tongue2 == "long" ~ "long:long",
-             tongue1 == "short" & tongue2 == "short" ~ "short:short",
-           )),
-           subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different")) %>%
-    unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":", remove = FALSE) %>% # create a single species pair column...
-    mutate(sp.pair = factor(sp.pair)) %>% # ...and make it a factor so that mgcv can handle it
-    mutate(niche.overlap = 1-distance) %>% # express niche overlap so that 1 = perfect overlap and 0 = no overlap
-    select(-distance)
-}
-
-# This one calculates niche overlap on fully resolved visitation data and then averages niche overlap across dates
-# niche_overlapper2 <- function(x, criterion) {
-#     
-#   criterion <- enquo(criterion) # criterion is set to either plant.sp (for taxonomic) or k.type.s (for morphological)
-#   
-#   x %>%
-#     #group_by(bb.sp, !!criterion, site, date, year) %>% # group by site-year and BB-FL pair
-#     group_by(bb.sp, !!criterion, site, date) %>% # group by site-year and BB-FL pair
-#     summarize(freq = sum(visits)) %>% # get interaction frequency for each BB-FL pair per site-year
-#     group_by(bb.sp, site, date) %>% # group by site-year and BB species
-#     mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
-#     #group_by(bb.sp, site, year, !!criterion) %>%
-#     # group_by(bb.sp, site, !!criterion) %>%
-#     # summarize(mean.prop = mean(prop)) %>%
-#     unite(sitedate, c(site, date), sep = "_") %>%
-#     #pivot_wider(names_from = !!criterion, values_from = mean.prop, values_fill = 0) %>% # spread into the kind of array vegan likes
-#     #pivot_wider(names_from = !!criterion, values_from = prop, values_fill = 0) %>% # spread into the kind of array vegan likes
-#     #group_by(site, year) %>%
-#     #group_by(site) %>%
-#     select(-freq) %>%
-#     group_by(sitedate) %>%
-#     nest() %>% # nest into list-column
-#     mutate(data2 = map(data, function(x) pivot_wider(x, names_from = !!criterion, values_from = prop, values_fill = 0)),
-#            data3 = map(data2, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
-#            dist = map(data3, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
-#            tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
-#     #dplyr::select(site, year, tidy_dist) %>% # drop intermediate list-columns
-#     #dplyr::select(site, tidy_dist) %>% # drop intermediate list-columns
-#     dplyr::select(sitedate, tidy_dist) %>% # drop intermediate list-columns
-#     unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
-#     separate(sitedate, c("site", "date"), sep = "_") %>%
-#     left_join(site_data) %>% # add site data
-#     # dplyr::select(site, year, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
-#     dplyr::select(site, date, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
-#     left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
-#     rename(subgenus1 = subgenus, tongue1 = pbl.w.class2) %>%
-#     left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
-#     rename(subgenus2 = subgenus, tongue2 = pbl.w.class2) %>%
-#     # dplyr::select(site, year, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
-#     #        tongue1, tongue2, elev.mean) %>%
-#     dplyr::select(site, date, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
-#        tongue1, tongue2, elev.mean) %>%
-#     mutate(tongue.diff = factor(if_else(tongue1 == tongue2, "same", "different")),
-#            tongue.pair = factor(case_when(
-#              tongue1 == "med" & tongue2 == "med" ~ "med:med",
-#              tongue1 == "med" & tongue2 == "long" ~ "med:long",
-#              tongue1 == "med" & tongue2 == "short" ~ "short:med",
-#              tongue1 == "long" & tongue2 == "med" ~ "med:long",
-#              tongue1 == "short" & tongue2 == "med" ~ "short:med",
-#              tongue1 == "long" & tongue2 == "short" ~ "short:long",
-#              tongue1 == "short" & tongue2 == "long" ~ "short:long",
-#              tongue1 == "long" & tongue2 == "long" ~ "long:long",
-#              tongue1 == "short" & tongue2 == "short" ~ "short:short",
-#            )),
-#            subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different")) %>%
-#     unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":", remove = FALSE) %>% # create a single species pair column...
-#     mutate(sp.pair = factor(sp.pair)) %>% # ...and make it a factor so that mgcv can handle it
-#     mutate(niche.overlap = 1-distance) %>% # express niche overlap so that 1 = perfect overlap and 0 = no overlap
-#     select(-distance)
-# }
-```
-
-## 3.4 Calculate niche overlap pooled within site
-```{r}
-# Pooled across dates within level, site, and species pair
-niche_overlap_gen <- niche_overlapper(niche_data, plant.genus) %>%
-  mutate(level = rep("genus", n()),
-         niche.overlap.genus = niche.overlap)
-
-niche_overlap_fam <- niche_overlapper(niche_data, plant.family) %>%
-  mutate(level = rep("family", n()),
-         niche.overlap.family = niche.overlap)
-
-niche_overlap_ktype <- niche_overlapper(niche_data, k.type.ss) %>%
-  mutate(level = rep("morphotype", n()),
-         niche.overlap.ktype = niche.overlap)
-
-# Collate across levels
-niche_overlap <- bind_rows(niche_overlap_gen,
-                           niche_overlap_fam, 
-                           niche_overlap_ktype,) %>%
-  mutate(level = factor(level, levels = c("family", "morphotype", "genus")),
-         site = factor(site),
-         management = factor(management, levels = c("none", "grasing", "mowing"))) %>%
-  separate(sp.pair, c("sp1", "sp2"), sep = ":", remove = FALSE)
-
-# Pool across species pairs within level and site to get community mean niche overlap
-niche_overlap_levels <- niche_overlap %>%
-  group_by(site, level, elev.mean, management) %>%
-  summarize(mean.niche.overlap = mean(niche.overlap)) %>%
-  mutate(level = factor(level, levels = c("family", "morphotype", "genus")))  # ordering them this way will make family the reference level, which is handy since I don't really care about it.
-
-# Isolate genus-level, then pool within dsp.pair and site
-niche_overlap_sppair <-  niche_overlap %>%
-  group_by(site, sp.pair, level, elev.mean, management) %>%
-  summarize(mean.niche.overlap = mean(niche.overlap)) %>%
-  mutate(level = factor(level, levels = c("family", "morphotype", "genus"))) %>%
-  filter(level == "genus") %>%
-  separate(sp.pair, c("sp1", "sp2"), sep = ":", remove = FALSE)
-
-# Isolate genus-level, then pool within tongue.pair and site
-niche_overlap_tongues <- niche_overlap %>%
-  group_by(site, level, tongue.pair, elev.mean, management) %>%
-  summarize(mean.niche.overlap = mean(niche.overlap)) %>%
-  filter(level == "genus")
-
-# Isolate genus-level, then pool within tongue.diff and site
-niche_overlap_tonguediff <- niche_overlap %>%
-  group_by(site, level, tongue.diff, elev.mean, management) %>%
-  summarize(mean.niche.overlap = mean(niche.overlap)) %>%
-  filter(level == "genus")
-
-#################
-
-ggplot(niche_overlap_sppair, aes(management, mean.niche.overlap)) +
-  geom_boxplot()
-```
-
-## 3.5 Niche overlap ~ level + elevation
-### 3.5.1 Model
-```{r}
-# Beta regression is probably the right way to do this; a gaussian model gives almost identical results, though, since our pooled data are approximately gaussian
-niche_gam_00 <- gam(mean.niche.overlap ~
-                      level + 
-                      #management +
-                      s(elev.mean, by = level, k = 8, m = 2),
-                    method = "REML",
-                    family = betar(eps = 0.001),
-                    select = TRUE,
-                    data = niche_overlap_levels) %>% getViz()
-
-check.gamViz(niche_gam_00)
-summary(niche_gam_00)
-print(plot.gamViz(niche_gam_00, allTerms = TRUE), pages = 1)
-```
-
-### 3.5.2 Visualize
-```{r}
-# Adding raw data back in
-data_add <- niche_overlap_levels %>%
-  ungroup() %>%
-  arrange(elev.mean, level) %>%
-  select(site, mean.niche.overlap)
-
-# Get predictions using tidymv
-niche_gam_00_pred <- get_gam_predictions(niche_gam_00, elev.mean) %>%
-  as_tibble() %>%
-  mutate(mean.niche.overlap_trans = inv_logit_scaled(mean.niche.overlap),
-         CI_upper_trans = inv_logit_scaled(CI_upper),
-         CI_lower_trans = inv_logit_scaled(CI_lower)) %>%
-  select(-mean.niche.overlap) %>%
-  bind_cols(data_add)
-
-# Plot smooths
-ggplot(niche_gam_00_pred, aes(elev.mean, mean.niche.overlap, color = level)) +
-  geom_point() +
-  geom_ribbon(aes(elev.mean, mean.niche.overlap_trans, 
-                  ymax = CI_upper_trans, ymin = CI_lower_trans,
-                  fill = level),
-              alpha = 0.3, inherit.aes = FALSE) +
-  geom_line(aes(elev.mean, mean.niche.overlap_trans, 
-                color = level), inherit.aes = FALSE) +
-  ylab("Niche overlap") +
-  xlab("Elevation (m)") +
-  guides(fill = FALSE) +
-  labs(color = "Level") +
-  theme_light()
-
-ggsave(file = "../output/niche_gam_00_plot1.pdf", width = 5, height = 4)
-ggsave(file = "../output/niche_gam_00_plot1.png", width = 5, height = 4)
-
-# Plot intercept effect of level
-plot(pterm(niche_gam_00, 1))
-
-plotDiff(sm(niche_gam_00, 1), sm(niche_gam_00, 3)) +
-  l_ciPoly() + 
-  l_fitLine() + 
-  geom_hline(yintercept = 0, linetype = 2)
-```
-
-## 3.7 Genus-level niche overlap by tongue.diff
-
-### 3.7.1 Model
-```{r}
-niche_gam_20 <- gam(mean.niche.overlap ~
-                     tongue.diff +
-                     s(elev.mean, by = tongue.diff, k = 8, m = 2),
-                   method = "REML",
-                   family = betar(eps = 0.001),
-                   select = TRUE,
-                   data = niche_overlap_tonguediff) %>% getViz()
-
-check.gamViz(niche_gam_20)
-summary(niche_gam_20)
-print(plot(niche_gam_20, allTerms = TRUE), pages = 1)
-
-niche_gam_21 <- gam(mean.niche.overlap ~
-                     tongue.pair +
-                     s(elev.mean, by = tongue.pair, k = 8, m = 2),
-                   method = "REML",
-                   family = betar(eps = 0.001),
-                   select = TRUE,
-                   data = niche_overlap_tongues) %>% getViz()
-
-check.gamViz(niche_gam_21)
-summary(niche_gam_21)
-print(plot(niche_gam_21, allTerms = TRUE), pages = 1)
-```
-
-### 3.7.2 Visualize
-```{r}
-# Tongue diff
-# Get predictions
-niche_gam_20_pred <- get_gam_predictions(niche_gam_20, elev.mean) %>%
-  as_tibble() %>%
-  mutate(mean.niche.overlap_trans = inv_logit_scaled(mean.niche.overlap),
-         CI_upper_trans = inv_logit_scaled(CI_upper),
-         CI_lower_trans = inv_logit_scaled(CI_lower))
-
-# Plot smooths
-ggplot(niche_gam_20_pred, aes(elev.mean, mean.niche.overlap_trans, color = tongue.diff)) +
-  geom_ribbon(aes(elev.mean, mean.niche.overlap_trans, 
-                  ymax = CI_upper_trans, ymin = CI_lower_trans,
-                  fill = tongue.diff),
-              alpha = 0.3, inherit.aes = FALSE) +
-  geom_line() +
-  ylab("Niche overlap") +
-  xlab("Elevation (m)") +
-  guides(fill = FALSE) +
-  labs(color = "Tongue class") +
-  theme_light()
-
-ggsave(file = "../output/niche_gam_20_plot1.pdf", width = 5, height = 4)
-ggsave(file = "../output/niche_gam_20_plot1.png", width = 5, height = 4)
-
-# Plot intercept effect of tongue.diff
-plot(pterm(niche_gam_20, 1))
-
-# Tongue pair
-# Get predictions
-niche_gam_21_pred <- get_gam_predictions(niche_gam_21, elev.mean) %>%
-  as_tibble() %>%
-  mutate(mean.niche.overlap_trans = inv_logit_scaled(mean.niche.overlap),
-         CI_upper_trans = inv_logit_scaled(CI_upper),
-         CI_lower_trans = inv_logit_scaled(CI_lower))
-
-# Plot smooths
-ggplot(niche_gam_21_pred, aes(elev.mean, mean.niche.overlap_trans, color = tongue.pair)) +
-  geom_ribbon(aes(elev.mean, mean.niche.overlap_trans, 
-                  ymax = CI_upper_trans, ymin = CI_lower_trans,
-                  fill = tongue.pair),
-              alpha = 0.3, inherit.aes = FALSE) +
-  geom_line() +
-  ylab("Niche overlap") +
-  xlab("Elevation (m)") +
-  guides(fill = FALSE) +
-  labs(color = "Tongue pair") +
-  theme_light()
-
-ggsave(file = "../output/niche_gam_21_plot1.pdf", width = 5, height = 4)
-ggsave(file = "../output/niche_gam_21_plot1.png", width = 5, height = 4)
-
-# Plot intercept effect of tongue.diff
-plot(pterm(niche_gam_20, 1))
-```

### 2021-04-14 code/ms2_ktype_preference.Rmd
set up null model approach to preference analysis

--- 2021-03-25 code/ms2_bias_niche.Rmd
@@
 ```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
-# Site data
-site_data <- read_csv("../data/processed_data/site_data.csv") %>%
-  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
-  mutate(elev.class2 = factor(elev.class2, levels = c("low", "mid", "high")),
-         site = factor(site)) # turn elav.class into an ordered factor and site into a regular factor
+prop_visits_ktype <- read_csv("../data/processed_data/prop_visits_ktype.csv") %>%
+  mutate(ktype = factor(ktype), # recode variables
+         tongue = factor(tongue),
+         visits = as.integer(visits))
 
-# BB traits
-bb_traits <- read_csv("../data/processed_data/bb_traits.csv")
-
-# FL traits
-## family-level taxonomy
-fl_tax <- read_csv("../data/processed_data/floral_tax.csv")
-
-## Kugler morphotypes (accessed through Bioflor)
-fl_ktype <- read_csv("../data/processed_data/floral_k_type.csv")
-
-## verbal description of Kugler morphotypes
-fl_ktype_key <- read_delim("../data/processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
-  dplyr::select(k.type = 1, description = 2)
-
-## join all floral trait tables
-fl_traits <- fl_tax %>%
-  full_join(fl_ktype) %>%
-  dplyr::select(plant.sp, plant.genus, plant.family, k.type, k.type.s, k.type.ss) 
+prop_visits_ktype_pooled <- read_csv("../data/processed_data/prop_visits_ktype_pool.csv") %>%
+  mutate(ktype = factor(ktype), # recode variables
+         visits = as.integer(visits))
 ```
 
-## Align survey and visitation data with pseudodates
-```{r}
-# Site-dates where BB visitation was recorded (excluding dates where only males were recorded)
-visitation_samples <- read_csv("../data/processed_data/network.csv") %>%
-  filter(caste != "male") %>%
-  select(site, date) %>%
-  distinct()
-
-# Site dates where floral surveying was conducted
-survey_samples <- read_csv("../data/processed_data/floral_survey.csv") %>%
-  select(site, date) %>%
-  distinct()
-
-# Site-dates where visitation sampling and floral sampling coincided
-shared_dates <- inner_join(visitation_samples, survey_samples)
-
-# Site-dates where visitation was recorded WITHOUT corresponding survey data
-visitation_only <- anti_join(visitation_samples, survey_samples) %>%
-  mutate(date.v = date) %>%
-  select(-date)
-
-# This creates a pseudo date column for unmatched visitation site-dates that will match each visitation date
-# to its nearest survey data. For site t4, there is a 2-year mismatch; we will drop these site-dates later.
-pseudodates <- full_join(visitation_only, survey_samples) %>%
-  mutate(mismatch = abs(date.v - date)) %>%
-  group_by(site, date.v) %>%
-  filter(mismatch == min(mismatch)) %>%
-  rename(date = date.v, pseudo.date = date)
-```
-
-## Generate a data frame that represents visitation and survey abundance
-```{r}
-# BB visitation data
-visitation <- read_csv("../data/processed_data/network.csv") %>%
-  filter(caste != "male") %>% # we will ignore males 
-  filter(!bb.sp %in% c("humi")) %>% # humi too rare
-  group_by(site, date, bb.sp, plant.sp) %>%
-  summarize(visits = n()) %>% # visits per BB-FL pair
-  ungroup() %>%
-  left_join(pseudodates) %>% # Now we will replace unmatched dates with their pseudo dates
-  replace_na(replace = list(mismatch = 0)) %>% # set mismatch to zero for perfectly aligned dates
-  filter(mismatch <= 7) %>% # this gets rid of irreconcilable dates (like the t4 dates mentioned above)
-  mutate(date2 = if_else(is.na(pseudo.date), date, pseudo.date)) %>% # replace date with pseudodate for all site-dates lacking perfect alignments
-  select(-c(date, pseudo.date, mismatch)) %>%
-  rename(date = date2) %>%
-  group_by(site, date) %>%
-  complete(bb.sp, plant.sp, fill = list(visits = 0)) # generate all possible BB-FL combos for each site date and fill in zeros for visits
-
-# Floral cover data
-flower_cover <- read_csv("../data/processed_data/floral_survey.csv") %>%
-  filter(flower.cover > 0) %>% # ignore plants that were not blooming
-  group_by(site, date, plant.sp) %>%
-  summarize(flower.cover = sum(flower.cover))
-
-# Get shared dates after adding pseudodates. This will let us filter out survey dates that don't
-# have corresponding visitation sampling dates
-shared_dates_with_pseudo <- inner_join(visitation, flower_cover, by = c("site", "date")) %>%
-  select(site, date) %>%
-  distinct()
-
-# Join visitation and flower cover data
-
-## First we need a list of all BB spp. by site and date
-sp_by_site <- visitation %>% 
-  ungroup() %>%
-  select(site, date, bb.sp) %>%
-  distinct()
-
-## Then we need a list of unvisited plant.sp by site and date
-unvisited <- flower_cover %>% 
-  semi_join(shared_dates_with_pseudo) %>% # filter flower cover data to include only site-dates with visitation complement (real or pseudodates)
-  full_join(visitation) %>% # join to visitation data
-  filter(is.na(bb.sp)) %>% # select only the site-date-plants that were totally unvisited (and therefore have no BB spp to join to)
-  select(-bb.sp) %>% # drop bb.sp variable because I just want the plants
-  left_join(sp_by_site) %>% # join in the species list by site and date, thus populating the unvisited plants with all the BB spp recorded for each site-date
-  replace_na(replace = list(visits = 0)) # replace NAs (created in the line above) with zero
-
-## Now we can combine everything into a single data frame
-visit_cover <- flower_cover %>% 
-  semi_join(shared_dates_with_pseudo) %>% # filter flower cover data to include only site-dates with visitation complement (real or pseudodates)
-  full_join(visitation) %>% # join to visitation data
-  replace_na(replace = list(flower.cover = 0.01)) %>% # If a plant was visited but not surveyed, we will assign it the minimum non-zero flower cover value of 0.01.
-  filter(!is.na(bb.sp)) %>% # drop the univisited plants
-  bind_rows(unvisited) %>% # add them back, with BB spp and zeros
-  select(site, date, bb.sp, plant.sp, visits, flower.cover) %>%
-  group_by(site, date, bb.sp, plant.sp, flower.cover) %>% # In a few cases, the use of pseudodates causes two sampling period to converge on the same date, which creates apparent duplicate visitation data.
-We will handle this by averaging the visits.
-  summarize(visits = mean(visits)) %>%
-  left_join(fl_traits) %>% # add floral traits
-  ungroup() %>%
-  mutate(k.type.ss = factor(k.type.ss))
-
-## Get total visits by BB sp and site-date; this will be a measure of BB abundance that we can use after we've calculated d'.
-I could, of course, work this into the biaseval function, but I'll put that off for now.
-bb_abundance <- visit_cover %>%
-  group_by(site, date, bb.sp) %>%
-  summarize(bb.sp.abund = sum(visits)) %>%
-  group_by(site, date) %>%
-  mutate(bb.sp.prop.abund = bb.sp.abund/sum(bb.sp.abund))
-
-fl_abundance <- visit_cover %>%
-  group_by(site, date) %>%
-  summarize(total.flower.cover = sum(flower.cover))
-```
-
-# 1. Overall bias
-
-## 1.1 Calculate visitation probability per plant morphotype.
-Instead of keeping the plant species resolved, we will pool visitation by morphotype.
-The problem with not pooling by morphotype is that more species-rich morphotypes will be penalized simply because their visitation is diluted across more species.
-```{r}
-prop_visits_ktype <- visit_cover %>%
-  group_by(site, date, bb.sp, k.type.ss) %>%
-  summarize(flower.cover = sum(flower.cover), # get raw flower cover and visits by site*date*bb.sp*plant.sp
-            visits = sum(visits)) %>%
-  group_by(site, date) %>%
-  mutate(prop.flower.cover = flower.cover/sum(flower.cover))%>% # binary visits
-  group_by(site, date, bb.sp) %>%
-  mutate(prop.visits = visits/sum(visits),
-         bb.sp.abund = sum(visits)) %>% # abundance of each BB sp. per site-date
-  ungroup() %>%
-  left_join(bb_traits) %>% # add BB traits
-  mutate(ktype = factor(k.type.ss), # recode variables
-         tongue = factor(pbl.w.class2),
-         visits = as.integer(visits)) %>% # so we can run a Poisson model
-  left_join(site_data) %>% # add site data
-  ungroup() %>%
-  mutate(yday = yday(date),
-         elev.mean0 = scale(elev.mean),
-         yday0 = scale(yday),
-         prop.flower.cover0 = scale(prop.flower.cover),
-         perc.flower.cover = prop.flower.cover*100,
-         bb.sp.abund0 = scale(bb.sp.abund)) %>%
-  select(site, elev.mean, date, yday, bb.sp, ktype, visits, 
-         prop.visits, flower.cover, perc.flower.cover, bb.sp.abund, tongue,
-         elev.mean0, yday0, prop.flower.cover0, bb.sp.abund0) %>%
-  mutate(visits.bin = if_else(visits > 0, 1, 0))
-
-write_csv(prop_visits_ktype, "../hpc_packets/prop_visits_ktype.csv")
-
-prop_visits_ktype_pooled <- visit_cover %>%
-  group_by(site, date, k.type.ss) %>%
-  summarize(flower.cover = sum(flower.cover), # get raw flower cover and visits by site*date*bb.sp*plant.sp
-            visits = sum(visits)) %>%
-  group_by(site, date) %>%
-  mutate(prop.flower.cover = flower.cover/sum(flower.cover),
-         prop.visits = visits/sum(visits),
-         bb.abund = sum(visits),
-         visits = as.integer(visits)) %>%
-  mutate(ktype = factor(k.type.ss)) %>%
-  left_join(site_data) %>% # add site data
-  ungroup() %>%
-  mutate(yday = yday(date),
-         elev.mean0 = scale(elev.mean),
-         yday0 = scale(yday),
-         perc.flower.cover = prop.flower.cover*100,
-         prop.flower.cover0 = scale(prop.flower.cover),
-         bb.abund0 = scale(bb.abund)) %>%
-  select(site, elev.mean, date, yday, ktype, visits, 
-         prop.visits, bb.abund, perc.flower.cover,
-         elev.mean0, yday0, prop.flower.cover0, bb.abund0) %>%
-  mutate(visits.bin = if_else(visits > 0, 1, 0),
-         year = factor(year(date)))
-
-write_csv(prop_visits_ktype_pooled, "../hpc_packets/prop_visits_ktype_pool.csv")
-
-```
-
-## 1.2 Binomial multi-level regression models
+1.2 Binomial multi-level regression models
 ### 1.2.1 Choose priors
@@
                   data = filter(prop_visits_ktype, !ktype %in% c("0", "9", "10")))
+```
+
+### Fixed effects versions of the models
+```{r}
+brm_pvis_10 <- brm(visits.bin ~
+                    bb.sp.abund*perc.flower.cover + # fixed term: because more abundant flowers will tend to be more visited,and more abundance bees visit more
+                    ktype:tongue +
+                     perc.flower.cover:ktype:tongue +
+                    (1 + perc.flower.cover | ktype:bb.sp) + # a random intercept and random slope for ktype:species; I think I need to model this before I can make conclusions about ktype:pbl.class
+                    (1 | ktype:site), # a varying intercept and slope for ktype:site
+                  family = "bernoulli", 
+                  prior = brm_pvis_prior, # weakly regularizing
+                  #file = "../output/brm_pvis_ktype_10",  # for running locally
+                  file = "brm_pvis_ktype_10", # for running on the HPC
+                  chains = 4,
+                  #cores = 12,
+                  iter = 2000,
+                  data = filter(prop_visits_ktype, !ktype %in% c("0", "9", "10")))
+
+conditional_effects(brm_pvis_10)
+
+brm_pvis_11 <- brm(as.integer(visits.bin) ~
+                    bb.abund*perc.flower.cover + # fixed term: because more abundant flowers will tend to be more visited,and more abundance bees visit more
+                    ktype +
+                    perc.flower.cover:ktype +
+                    (1 | ktype:site),  # a varying intercept term for ktype:site
+                  family = "bernoulli", 
+                  prior = brm_pvis_prior, # weakly regularizing
+                  #control = list(adapt_delta = 0.99), # the default setting yielded divergent transitions
+                  file = "../output/brm_pvis_ktype_11",  # for running locally
+                  #file = "brm_pvis_ktype_11", # for running on the HPC
+                  chains = 4,
+                  #cores = 12,
+                  iter = 2000,
+                  data = filter(prop_visits_ktype_pooled, !ktype %in% c("0", "9", "10")))
+
+conditional_effects(brm_pvis_11)
+
+# My intention is that ~ 0 will force the intercept of the fixed slope effects to be zero, since in fact visitation probability must be zero when there are either no  bees or  no flowers.
+brm_pvis_11b <- brm(as.integer(visits.bin) ~ 0 +
+                    bb.abund*perc.flower.cover + # fixed term: because more abundant flowers will tend to be more visited,and more abundance bees visit more
+                    ktype +
+                    perc.flower.cover:ktype +
+                    (1 | ktype:site),  # a varying intercept term for ktype:site
+                  family = "bernoulli", 
+                  prior = brm_pvis_prior, # weakly regularizing
+                  #control = list(adapt_delta = 0.99), # the default setting yielded divergent transitions
+                  file = "../output/brm_pvis_ktype_11b",  # for running locally
+                  #file = "brm_pvis_ktype_11b", # for running on the HPC
+                  chains = 4,
+                  #cores = 12,
+                  iter = 2000,
+                  data = filter(prop_visits_ktype_pooled, !ktype %in% c("0", "9", "10")))
+
+conditional_effects(brm_pvis_11b)
+
+brm_pvis_21 <- brm(prop.visits ~
+                    perc.flower.cover*ktype +
+                    (1 | ktype:site),  # a varying intercept term for ktype:site
+                  family = "zero_one_inflated_beta", 
+                  #prior = brm_pvis_prior, # weakly regularizing
+                  #control = list(adapt_delta = 0.99), # the default setting yielded divergent transitions
+                  #file = "../output/brm_pvis_ktype_21",  # for running locally
+                  file = "brm_pvis_ktype_21", # for running on the HPC
+                  chains = 4,
+                  #cores = 12,
+                  iter = 2000,
+                  data = filter(prop_visits_ktype_pooled, !ktype %in% c("0", "9", "10")))
+
+conditional_effects(brm_pvis_11)
 ```
@@
 ```
-
-# 2. Ordination of dietary niche
-Aggregation changes meaning.
-Don't forget that.
-When I pool visitation across dates, I am introducing the effects of relative abundance and floral phenology.
-Species in bloom when BB abundance is high will be over-represented, and vice-versa.
-And species with a long bloom time will be over-represented, and vice-versa.
-This isn't necessarily a problem.
-I just need to make sure I interpret the results the right way.
-When I pool data like this, I am comparing the "net" diet of species, which includes the effects of abundance and phenology in addition to the effects of floral preferences.
-I think that is appropriate for this stage of the analysis, but I need to be careful not to assume that what I'm doing here is the same sort of analysis as what I did above without pooling samples.
-```{r}
-# Fl traits with abbreviated names
-fl_traits2 <- fl_traits %>%
-  left_join(read_csv("../data/processed_data/network.csv")) %>%
-  #select(plant.sp, plant.sp.abb, plant.genus, plant.family, k.type.ss, color) %>%
-  select(plant.genus, k.type.ss) %>%
-  mutate(k.type.ss = factor(k.type.ss)) %>%
-  distinct()
-
-# Totally aggregated; note I'm using the full data, not the aligned and trimmed data. No need for complementary floral survey data.
-visitation2 <- read_csv("../data/processed_data/network.csv") %>%
-  filter(caste != "male") %>% # we will ignore males 
-  filter(!bb.sp %in% c("humi")) %>% # humi is too rare
-  #group_by(bb.sp, plant.sp.abb) %>%
-  group_by(bb.sp, plant.genus) %>%
-  summarize(visits = n()) %>%
-  group_by(bb.sp) %>%
-  mutate(prop.visits = visits/sum(visits)) %>%
-  select(-visits) %>%
-  #pivot_wider(names_from = plant.sp.abb, values_from = prop.visits, values_fill = 0) %>%
-  pivot_wider(names_from = plant.genus, values_from = prop.visits, values_fill = 0) %>%
-  column_to_rownames("bb.sp")
-
-bb_traits2 <- bb_traits %>%
-  filter(!bb.sp %in% c("humi")) %>% # drop humi
-  select(bb.sp, pbl.w.class2) %>%
-  arrange(bb.sp) %>%
-  mutate(pbl.w.class2.pool = case_when(
-    pbl.w.class2 == "short" ~ "short",
-    pbl.w.class2 == "med" ~ "med/long",
-    pbl.w.class2 == "long" ~ "med/long"
-  ))
-
-# Generate ordination
-pcoa_horn2 <- capscale(visitation2 ~ 1, distance = "horn")
-pcoa_horn2_adonis <- adonis(visitation2 ~ pbl.w.class2, data = bb_traits2)
-pcoa_horn2_betadisper <- betadisper(vegdist(visitation2, method = "horn"), 
-                                    group = bb_traits2$pbl.w.class2,
-                                    type = "median")
-anova(pcoa_horn2_betadisper) # I think this is not significant because short tongued species are disperse on axis 2, but axis 2 does not explain nearly as much variation.
-plot(pcoa_horn2_betadisper)
-# I'm surprised that the betadisper test is not significant, but I think it might be driven by the fact that replication is so low for long and medium tongue species.
-Let's see what happens if we pool long and medium tongue species, since they overlap so much anyway.
-pcoa_horn2_betadisper_pool <- betadisper(vegdist(visitation2, method = "horn"), 
-                                    group = bb_traits2$pbl.w.class2.pool,
-                                    type = "median")
-anova(pcoa_horn2_betadisper_pool) # I think this is not significant because short tongued species are dispersed on axis 2, but axis 2 does not explain nearly as much variation.
-plot(pcoa_horn2_betadisper_pool)
-
-# Get variance explained
-pcoa_horn2_mds1 <- round(pcoa_horn2$CA$eig[[1]] / sum(pcoa_horn2$CA$eig)*100)
-pcoa_horn2_mds2 <- round(pcoa_horn2$CA$eig[[2]] / sum(pcoa_horn2$CA$eig)*100)
-
-# Fortify plants
-pcoa_horn2.fort1 <- fortify(pcoa_horn2, axes = 1:2) %>%
-  filter(Score == "species") %>%
-  #rename(plant.sp.abb = Label) %>%
-  rename(plant.genus = Label) %>%
-  left_join(fl_traits2)
-
-# Fortify bees
-pcoa_horn2.fort2 <- fortify(pcoa_horn2, axes = 1:2) %>%
-  filter(Score == "sites") %>%
-  rename(bb.sp = Label) %>%
-  left_join(bb_traits) %>%
-  mutate(pbl.w.class2 = factor(pbl.w.class2))
-
-# Generate convex hulls around tongue length classes
-pcoa_horn2_hulls <- pcoa_horn2.fort2 %>%
-  group_by(pbl.w.class2) %>%
-  slice(chull(MDS1, MDS2))
-
-# Plot
-ggplot(pcoa_horn2.fort2, aes(MDS1, MDS2, color = pbl.w.class2)) +
-  geom_polygon(data = pcoa_horn2_hulls, 
-               aes(fill = pbl.w.class2, color = NULL),
-               alpha = 0.25) +
-  geom_point(data = filter(pcoa_horn2.fort1, abs(MDS1) >= 0.20 | abs(MDS2) >= 0.20),
-            aes(MDS1, MDS2, shape = k.type.ss), inherit.aes = FALSE, color = "gray40", size = 2) +
-  geom_text(aes(label = bb.sp), key_glyph = "rect", size = 3) +
-  geom_text_repel(data = filter(pcoa_horn2.fort1, abs(MDS1) >= 0.20 | abs(MDS2) >= 0.20), 
-            aes(MDS1, MDS2, label = plant.genus), inherit.aes = FALSE, color = "gray40",
-            size = 3) +
-  scale_shape_discrete() +
-  xlab(paste("MDS1 ", "(", pcoa_horn2_mds1, "%", ")", sep = "")) +
-  ylab(paste("MDS2 ", "(", pcoa_horn2_mds2, "%", ")", sep = "")) +
-  labs(fill = "Tongue length", 
-       shape = "Morphotype") +
-  guides(color = FALSE) +
-  coord_fixed() +
-  theme_light(9)
-
-ggsave("../output/pcoa_aggregate.png")
-```
-
-# 3. Dietary niche overlap (i.e. potential competition)
-
-## 3.1 Check BB distribution patterns to decide whom to retain in the analysis
-```{r}
-bb_distribution <- visitation %>%
-  mutate(year = year(date)) %>%
-  left_join(site_data) %>%
-  group_by(site, date, elev.mean, bb.sp, year) %>%
-  summarize(abund = n()) %>%
-  group_by(site, elev.mean, bb.sp, year) %>%
-  summarize(abund = mean(abund)) %>%
-  group_by(bb.sp) %>%
-  mutate(elev.floor = min(elev.mean),
-         elev.ceiling = max(elev.mean),
-         elev.range = elev.ceiling - elev.floor,
-         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))
-
-ggplot(bb_distribution, aes(reorder(bb.sp, elev.med), elev.mean)) +
-  geom_line(size = 2, alpha = 0.25) +
-  geom_point(aes(size = abund), alpha = 0.5) +
-  facet_wrap(~year, ncol = 1) +
-  theme_light(12) +
-  scale_size_continuous(name = "Abundance") +
-  xlab("BB species") +
-  ylab("Elevation") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-bb_distribution_pool <- visitation %>%
-  left_join(site_data) %>%
-  group_by(site, date, elev.mean, bb.sp) %>%
-  summarize(abund = n()) %>%
-  ungroup() %>%
-  complete(site, date, elev.mean, bb.sp, fill = list(abund = 0)) %>%
-  group_by(site, elev.mean, bb.sp) %>%
-  summarize(abund = mean(abund)) %>%
-  filter(abund > 0) %>%
-  group_by(bb.sp) %>%
-  mutate(elev.floor = min(elev.mean),
-         elev.ceiling = max(elev.mean),
-         elev.range = elev.ceiling - elev.floor,
-         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))
-
-ggplot(bb_distribution_pool, aes(reorder(bb.sp, elev.med), elev.mean)) +
-  geom_line(size = 2, alpha = 0.25) +
-  geom_point(aes(size = abund), alpha = 0.5) +
-  theme_light(14) +
-  scale_size_continuous(name = "Abundance") +
-  xlab("BB species") +
-  ylab("Elevation") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-ggsave("../output/bb_range_ms2.pdf", width = 6, height = 5)
-```
-
-## 3.2 Prep data for calculating niche overlap
-```{r}
-niche_data <- visitation %>%
-  left_join(fl_traits) %>%
-  mutate(year = factor(year(date)))
-```
-
-## 3.3 Function for calculating niche overlap
-```{r}
-# This one averages visitation and then calculates niche overlaps
-niche_overlapper <- function(x, criterion) {
-
-  criterion <- enquo(criterion) # criterion is set to either plant.sp (for taxonomic) or k.type.s (for morphological)
-
-  x %>%
-    #group_by(bb.sp, !!criterion, site, date, year) %>% # group by site-year and BB-FL pair
-    group_by(bb.sp, !!criterion, site, date) %>% # group by site-year and BB-FL pair
-    summarize(freq = sum(visits)) %>% # get interaction frequency for each BB-FL pair per site-year
-    group_by(bb.sp, site, date) %>% # group by site-year and BB species
-    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
-    #group_by(bb.sp, site, year, !!criterion) %>%
-    group_by(bb.sp, site, !!criterion) %>%
-    summarize(mean.prop = mean(prop)) %>%
-    pivot_wider(names_from = !!criterion, values_from = mean.prop, values_fill = 0) %>% # spread into the kind of array vegan likes
-    #group_by(site, year) %>%
-    group_by(site) %>%
-    nest() %>% # nest into list-column
-    mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
-           dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
-           tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
-    #dplyr::select(site, year, tidy_dist) %>% # drop intermediate list-columns
-    dplyr::select(site, tidy_dist) %>% # drop intermediate list-columns
-    unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
-    left_join(site_data) %>% # add site data
-    # dplyr::select(site, year, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
-    dplyr::select(site, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean, management) %>% # dplyr::select and rename vars
-    left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
-    rename(subgenus1 = subgenus, tongue1 = pbl.w.class2) %>%
-    left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
-    rename(subgenus2 = subgenus, tongue2 = pbl.w.class2) %>%
-    dplyr::select(site, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
-       tongue1, tongue2, elev.mean, management) %>%
-    mutate(tongue.diff = factor(if_else(tongue1 == tongue2, "same", "different")),
-           tongue.pair = factor(case_when(
-             tongue1 == "med" & tongue2 == "med" ~ "med:med",
-             tongue1 == "med" & tongue2 == "long" ~ "med:long",
-             tongue1 == "med" & tongue2 == "short" ~ "short:med",
-             tongue1 == "long" & tongue2 == "med" ~ "med:long",
-             tongue1 == "short" & tongue2 == "med" ~ "short:med",
-             tongue1 == "long" & tongue2 == "short" ~ "short:long",
-             tongue1 == "short" & tongue2 == "long" ~ "short:long",
-             tongue1 == "long" & tongue2 == "long" ~ "long:long",
-             tongue1 == "short" & tongue2 == "short" ~ "short:short",
-           )),
-           subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different")) %>%
-    unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":", remove = FALSE) %>% # create a single species pair column...
-    mutate(sp.pair = factor(sp.pair)) %>% # ...and make it a factor so that mgcv can handle it
-    mutate(niche.overlap = 1-distance) %>% # express niche overlap so that 1 = perfect overlap and 0 = no overlap
-    select(-distance)
-}
-
-# This one calculates niche overlap on fully resolved visitation data and then averages niche overlap across dates
-# niche_overlapper2 <- function(x, criterion) {
-#     
-#   criterion <- enquo(criterion) # criterion is set to either plant.sp (for taxonomic) or k.type.s (for morphological)
-#   
-#   x %>%
-#     #group_by(bb.sp, !!criterion, site, date, year) %>% # group by site-year and BB-FL pair
-#     group_by(bb.sp, !!criterion, site, date) %>% # group by site-year and BB-FL pair
-#     summarize(freq = sum(visits)) %>% # get interaction frequency for each BB-FL pair per site-year
-#     group_by(bb.sp, site, date) %>% # group by site-year and BB species
-#     mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
-#     #group_by(bb.sp, site, year, !!criterion) %>%
-#     # group_by(bb.sp, site, !!criterion) %>%
-#     # summarize(mean.prop = mean(prop)) %>%
-#     unite(sitedate, c(site, date), sep = "_") %>%
-#     #pivot_wider(names_from = !!criterion, values_from = mean.prop, values_fill = 0) %>% # spread into the kind of array vegan likes
-#     #pivot_wider(names_from = !!criterion, values_from = prop, values_fill = 0) %>% # spread into the kind of array vegan likes
-#     #group_by(site, year) %>%
-#     #group_by(site) %>%
-#     select(-freq) %>%
-#     group_by(sitedate) %>%
-#     nest() %>% # nest into list-column
-#     mutate(data2 = map(data, function(x) pivot_wider(x, names_from = !!criterion, values_from = prop, values_fill = 0)),
-#            data3 = map(data2, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
-#            dist = map(data3, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
-#            tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
-#     #dplyr::select(site, year, tidy_dist) %>% # drop intermediate list-columns
-#     #dplyr::select(site, tidy_dist) %>% # drop intermediate list-columns
-#     dplyr::select(sitedate, tidy_dist) %>% # drop intermediate list-columns
-#     unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
-#     separate(sitedate, c("site", "date"), sep = "_") %>%
-#     left_join(site_data) %>% # add site data
-#     # dplyr::select(site, year, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
-#     dplyr::select(site, date, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
-#     left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
-#     rename(subgenus1 = subgenus, tongue1 = pbl.w.class2) %>%
-#     left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
-#     rename(subgenus2 = subgenus, tongue2 = pbl.w.class2) %>%
-#     # dplyr::select(site, year, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
-#     #        tongue1, tongue2, elev.mean) %>%
-#     dplyr::select(site, date, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
-#        tongue1, tongue2, elev.mean) %>%
-#     mutate(tongue.diff = factor(if_else(tongue1 == tongue2, "same", "different")),
-#            tongue.pair = factor(case_when(
-#              tongue1 == "med" & tongue2 == "med" ~ "med:med",
-#              tongue1 == "med" & tongue2 == "long" ~ "med:long",
-#              tongue1 == "med" & tongue2 == "short" ~ "short:med",
-#              tongue1 == "long" & tongue2 == "med" ~ "med:long",
-#              tongue1 == "short" & tongue2 == "med" ~ "short:med",
-#              tongue1 == "long" & tongue2 == "short" ~ "short:long",
-#              tongue1 == "short" & tongue2 == "long" ~ "short:long",
-#              tongue1 == "long" & tongue2 == "long" ~ "long:long",
-#              tongue1 == "short" & tongue2 == "short" ~ "short:short",
-#            )),
-#            subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different")) %>%
-#     unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":", remove = FALSE) %>% # create a single species pair column...
-#     mutate(sp.pair = factor(sp.pair)) %>% # ...and make it a factor so that mgcv can handle it
-#     mutate(niche.overlap = 1-distance) %>% # express niche overlap so that 1 = perfect overlap and 0 = no overlap
-#     select(-distance)
-# }
-```
-
-## 3.4 Calculate niche overlap pooled within site
-```{r}
-# Pooled across dates within level, site, and species pair
-niche_overlap_gen <- niche_overlapper(niche_data, plant.genus) %>%
-  mutate(level = rep("genus", n()),
-         niche.overlap.genus = niche.overlap)
-
-niche_overlap_fam <- niche_overlapper(niche_data, plant.family) %>%
-  mutate(level = rep("family", n()),
-         niche.overlap.family = niche.overlap)
-
-niche_overlap_ktype <- niche_overlapper(niche_data, k.type.ss) %>%
-  mutate(level = rep("morphotype", n()),
-         niche.overlap.ktype = niche.overlap)
-
-# Collate across levels
-niche_overlap <- bind_rows(niche_overlap_gen,
-                           niche_overlap_fam, 
-                           niche_overlap_ktype,) %>%
-  mutate(level = factor(level, levels = c("family", "morphotype", "genus")),
-         site = factor(site),
-         management = factor(management, levels = c("none", "grasing", "mowing"))) %>%
-  separate(sp.pair, c("sp1", "sp2"), sep = ":", remove = FALSE)
-
-# Pool across species pairs within level and site to get community mean niche overlap
-niche_overlap_levels <- niche_overlap %>%
-  group_by(site, level, elev.mean, management) %>%
-  summarize(mean.niche.overlap = mean(niche.overlap)) %>%
-  mutate(level = factor(level, levels = c("family", "morphotype", "genus")))  # ordering them this way will make family the reference level, which is handy since I don't really care about it.
-
-# Isolate genus-level, then pool within dsp.pair and site
-niche_overlap_sppair <-  niche_overlap %>%
-  group_by(site, sp.pair, level, elev.mean, management) %>%
-  summarize(mean.niche.overlap = mean(niche.overlap)) %>%
-  mutate(level = factor(level, levels = c("family", "morphotype", "genus"))) %>%
-  filter(level == "genus") %>%
-  separate(sp.pair, c("sp1", "sp2"), sep = ":", remove = FALSE)
-
-# Isolate genus-level, then pool within tongue.pair and site
-niche_overlap_tongues <- niche_overlap %>%
-  group_by(site, level, tongue.pair, elev.mean, management) %>%
-  summarize(mean.niche.overlap = mean(niche.overlap)) %>%
-  filter(level == "genus")
-
-# Isolate genus-level, then pool within tongue.diff and site
-niche_overlap_tonguediff <- niche_overlap %>%
-  group_by(site, level, tongue.diff, elev.mean, management) %>%
-  summarize(mean.niche.overlap = mean(niche.overlap)) %>%
-  filter(level == "genus")
-
-#################
-
-ggplot(niche_overlap_sppair, aes(management, mean.niche.overlap)) +
-  geom_boxplot()
-```
-
-## 3.5 Niche overlap ~ level + elevation
-### 3.5.1 Model
-```{r}
-# Beta regression is probably the right way to do this; a gaussian model gives almost identical results, though, since our pooled data are approximately gaussian
-niche_gam_00 <- gam(mean.niche.overlap ~
-                      level + 
-                      #management +
-                      s(elev.mean, by = level, k = 8, m = 2),
-                    method = "REML",
-                    family = betar(eps = 0.001),
-                    select = TRUE,
-                    data = niche_overlap_levels) %>% getViz()
-
-check.gamViz(niche_gam_00)
-summary(niche_gam_00)
-print(plot.gamViz(niche_gam_00, allTerms = TRUE), pages = 1)
-```
-
-### 3.5.2 Visualize
-```{r}
-# Adding raw data back in
-data_add <- niche_overlap_levels %>%
-  ungroup() %>%
-  arrange(elev.mean, level) %>%
-  select(site, mean.niche.overlap)
-
-# Get predictions using tidymv
-niche_gam_00_pred <- get_gam_predictions(niche_gam_00, elev.mean) %>%
-  as_tibble() %>%
-  mutate(mean.niche.overlap_trans = inv_logit_scaled(mean.niche.overlap),
-         CI_upper_trans = inv_logit_scaled(CI_upper),
-         CI_lower_trans = inv_logit_scaled(CI_lower)) %>%
-  select(-mean.niche.overlap) %>%
-  bind_cols(data_add)
-
-# Plot smooths
-ggplot(niche_gam_00_pred, aes(elev.mean, mean.niche.overlap, color = level)) +
-  geom_point() +
-  geom_ribbon(aes(elev.mean, mean.niche.overlap_trans, 
-                  ymax = CI_upper_trans, ymin = CI_lower_trans,
-                  fill = level),
-              alpha = 0.3, inherit.aes = FALSE) +
-  geom_line(aes(elev.mean, mean.niche.overlap_trans, 
-                color = level), inherit.aes = FALSE) +
-  ylab("Niche overlap") +
-  xlab("Elevation (m)") +
-  guides(fill = FALSE) +
-  labs(color = "Level") +
-  theme_light()
-
-ggsave(file = "../output/niche_gam_00_plot1.pdf", width = 5, height = 4)
-ggsave(file = "../output/niche_gam_00_plot1.png", width = 5, height = 4)
-
-# Plot intercept effect of level
-plot(pterm(niche_gam_00, 1))
-
-plotDiff(sm(niche_gam_00, 1), sm(niche_gam_00, 3)) +
-  l_ciPoly() + 
-  l_fitLine() + 
-  geom_hline(yintercept = 0, linetype = 2)
-```
-
-## 3.7 Genus-level niche overlap by tongue.diff
-
-### 3.7.1 Model
-```{r}
-niche_gam_20 <- gam(mean.niche.overlap ~
-                     tongue.diff +
-                     s(elev.mean, by = tongue.diff, k = 8, m = 2),
-                   method = "REML",
-                   family = betar(eps = 0.001),
-                   select = TRUE,
-                   data = niche_overlap_tonguediff) %>% getViz()
-
-check.gamViz(niche_gam_20)
-summary(niche_gam_20)
-print(plot(niche_gam_20, allTerms = TRUE), pages = 1)
-
-niche_gam_21 <- gam(mean.niche.overlap ~
-                     tongue.pair +
-                     s(elev.mean, by = tongue.pair, k = 8, m = 2),
-                   method = "REML",
-                   family = betar(eps = 0.001),
-                   select = TRUE,
-                   data = niche_overlap_tongues) %>% getViz()
-
-check.gamViz(niche_gam_21)
-summary(niche_gam_21)
-print(plot(niche_gam_21, allTerms = TRUE), pages = 1)
-```
-
-### 3.7.2 Visualize
-```{r}
-# Tongue diff
-# Get predictions
-niche_gam_20_pred <- get_gam_predictions(niche_gam_20, elev.mean) %>%
-  as_tibble() %>%
-  mutate(mean.niche.overlap_trans = inv_logit_scaled(mean.niche.overlap),
-         CI_upper_trans = inv_logit_scaled(CI_upper),
-         CI_lower_trans = inv_logit_scaled(CI_lower))
-
-# Plot smooths
-ggplot(niche_gam_20_pred, aes(elev.mean, mean.niche.overlap_trans, color = tongue.diff)) +
-  geom_ribbon(aes(elev.mean, mean.niche.overlap_trans, 
-                  ymax = CI_upper_trans, ymin = CI_lower_trans,
-                  fill = tongue.diff),
-              alpha = 0.3, inherit.aes = FALSE) +
-  geom_line() +
-  ylab("Niche overlap") +
-  xlab("Elevation (m)") +
-  guides(fill = FALSE) +
-  labs(color = "Tongue class") +
-  theme_light()
-
-ggsave(file = "../output/niche_gam_20_plot1.pdf", width = 5, height = 4)
-ggsave(file = "../output/niche_gam_20_plot1.png", width = 5, height = 4)
-
-# Plot intercept effect of tongue.diff
-plot(pterm(niche_gam_20, 1))
-
-# Tongue pair
-# Get predictions
-niche_gam_21_pred <- get_gam_predictions(niche_gam_21, elev.mean) %>%
-  as_tibble() %>%
-  mutate(mean.niche.overlap_trans = inv_logit_scaled(mean.niche.overlap),
-         CI_upper_trans = inv_logit_scaled(CI_upper),
-         CI_lower_trans = inv_logit_scaled(CI_lower))
-
-# Plot smooths
-ggplot(niche_gam_21_pred, aes(elev.mean, mean.niche.overlap_trans, color = tongue.pair)) +
-  geom_ribbon(aes(elev.mean, mean.niche.overlap_trans, 
-                  ymax = CI_upper_trans, ymin = CI_lower_trans,
-                  fill = tongue.pair),
-              alpha = 0.3, inherit.aes = FALSE) +
-  geom_line() +
-  ylab("Niche overlap") +
-  xlab("Elevation (m)") +
-  guides(fill = FALSE) +
-  labs(color = "Tongue pair") +
-  theme_light()
-
-ggsave(file = "../output/niche_gam_21_plot1.pdf", width = 5, height = 4)
-ggsave(file = "../output/niche_gam_21_plot1.png", width = 5, height = 4)
-
-# Plot intercept effect of tongue.diff
-plot(pterm(niche_gam_20, 1))
-```

### 2021-04-14 code/ms2_ktype_preference_nullmodel_analysis.Rmd
set up null model approach to preference analysis

---
title: "visitation_bias"
author: "Doug Sponsler"
date: "12/14/2020"
output: html_document
---
# Preparations
## Packages 
```{r, warning=FALSE, include=FALSE, echo = FALSE}
library(bipartite)
library(econullnetr)
library(patchwork)
library(mgcv)
library(mgcViz)
library(tidymv)
library(tidyverse)
library(lubridate)
```

## Load data
```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
prop_visits_ktype <- read_csv("../data/processed_data/prop_visits_ktype.csv") %>%
  mutate(ktype = factor(ktype), # recode variables
         tongue = factor(tongue),
         visits = as.integer(visits))

prop_visits_ktype_pooled <- read_csv("../data/processed_data/prop_visits_ktype_pool.csv") %>%
  mutate(ktype = factor(ktype), # recode variables
         visits = as.integer(visits))

bb_traits <- read_csv("../data/processed_data/bb_traits.csv") %>%
  select(bb.sp, pbl.w, tongue = pbl.w.class2)

site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean, elev.class, elev.class2)) %>% 
  mutate(site = factor(site),
         tree_line = factor(tree_line))
```

# Null model by species
```{r}
# The "consumers" table input to generate_null_net()
con <- prop_visits_ktype %>%
  unite(id, c(site, date)) %>%
  select(id, bb.sp, ktype, visits) %>%
  pivot_wider(values_from = visits, names_from = ktype, values_fill = 0) %>%
  select(-c("0", "10", "9"))

# The "resources" table input to generate_null_net()
res <- prop_visits_ktype %>%
  unite(id, c(site, date)) %>%
  select(id, ktype, perc.flower.cover) %>%
  distinct() %>%
  pivot_wider(values_from = perc.flower.cover, names_from = ktype, values_fill = 0) %>%
  select(-c("0", "10", "9"))

# Generate null model
null1 <- generate_null_net(consumers = con[, 2:9],
                           resources = res[, 2:8],
                           sims = 5000, # raised from default of 1000 (takes a while to run)
                           data.type = "counts",
                           c.samples = con[, 1], # parsed by sample
                           r.samples = res[, 1])

#saveRDS(null1, file = "../output/nullmodel_ktype.rds")
null1 <- readRDS("../output/nullmodel_ktype.rds")

# Tabulate output and generate tests
null_tab1 <- test_interactions(null1, signif.level = 0.99) %>% # make it a bit more conservative to account for multiple comparisons
  as_tibble() %>%
  select(bb.sp = Consumer, ktype = Resource, everything()) %>%
  group_by(bb.sp) %>%
  mutate(total.abund = sum(Observed), # total abundance per species
         preference = Observed - Null, # a working definition of preference
         preference.upper99 = Observed - Upper.99.CL, 
         preference.lower99 = Observed - Lower.99.CL,

         # If preference equals the difference between # observed visits and the null model prediction, then dividing preference by the total abundance of each species yields a standardized preference. Right? 
         preference.std = preference/total.abund, 
         Observed.std = Observed/total.abund,
         Upper.99.CL.std = Upper.99.CL/total.abund,
         Lower.99.CL.std = Lower.99.CL/total.abund) %>%
  left_join(bb_traits) %>%
  ungroup() %>%
  mutate(bb.sp = fct_reorder(factor(bb.sp), pbl.w))
```

## Visualize network
```{r}
plot_bipartite(null1, signif.level = 0.99)
```

## Visualize species model
```{r}
# Standardized observed vs. null predictions
## Species-wise
ggplot(null_tab1, aes(Observed.std, ktype)) +
  geom_segment(aes(y = ktype, yend = ktype, 
                   x = Lower.99.CL.std, xend = Upper.99.CL.std), 
               color = "gray40", inherit.aes = FALSE) +
  geom_point(aes(color = Test)) +
  facet_wrap(~bb.sp) +
  scale_y_discrete(limits = rev) +
  scale_color_manual(values = c("gray40", "#00B6EB", "#F8766D")) +
  labs(x = "Standardized visits", y = "Floral morphotype", color = "Test (p < 0.01)") +
  theme_light()

ggsave("../output/null_model_spwise_preference.pdf", height = 6, width = 7)
ggsave("../output/null_model_spwise_preference.png", height = 6, width = 7)

## Morphotype-wise
ggplot(null_tab1, aes(Observed.std, bb.sp)) +
  geom_segment(aes(y = bb.sp, yend = bb.sp, 
                   x = Lower.99.CL.std, xend = Upper.99.CL.std), 
               color = "gray40", inherit.aes = FALSE) +
  geom_point(aes(color = Test)) +
  facet_wrap(~ktype) +
  scale_y_discrete(limits = rev) +
  scale_color_manual(values = c("gray40", "#00B6EB", "#F8766D")) +
  labs(x = "Standardized visits", y = "Bumble bee species", color = "Test (p < 0.01)") +
  theme_light()

ggsave("../output/null_model_ktypewise_preference.pdf", height = 6, width = 7)
ggsave("../output/null_model_ktypewise_preference.png", height = 6, width = 7)

# Standardized preference ~ morphotype
ggplot(null_tab1, aes(ktype, preference.std)) +
  #geom_violin(fill = "#F8766D", color = "#F8766D") +
  geom_violin(fill = "gray70", color = "gray70") +
  geom_boxplot(fill = NA, color = "black") +
  theme_light(14) +
  labs(x = "Floral morphotype", 
       y = "Standardized preference")

ggsave("../output/null_model_ktype_preference.pdf", height = 6, width = 7)
ggsave("../output/null_model_ktype_preference.png", height = 6, width = 7)

# Standardized preference ~ morphotype * tongue class
ggplot(null_tab1, aes(ktype, preference.std)) +
  #geom_violin(fill = "#F8766D", color = "#F8766D") +
  geom_violin(fill = "gray70", color = "gray70") +
  geom_boxplot(fill = NA, color = "black") +
  theme_light(14) +
  labs(x = "Floral morphotype", 
       y = "Standardized preference") +
  facet_wrap(~tongue)

ggsave("../output/null_model_ktypeXtongueclass_preference.pdf", height = 6, width = 7)
ggsave("../output/null_model_ktypeXtongueclass_preference.png", height = 6, width = 7)

# Standardized preference ~ tongue length
ggplot(null_tab1, aes(pbl.w, preference.std)) +
  geom_point(aes(color = Test)) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5), color = "gray20", size = 0.5) +
  facet_wrap(~ktype, scales = "free") +
  theme_light()
```

## Model standardized preference ~ tongue length
```{r}
gam1 <- gam(preference.std ~ 
              #s(ktype, bs = "re") + 
              ktype + 
              s(pbl.w, by = ktype, k = 5), 
            data = null_tab1) %>% getViz()

summary(gam1)
check.gamViz(gam1)
plot.gamViz(gam1, allTerms = TRUE)

# Extract predictions
gam1_pred <- predict_gam(gam1) %>%
  mutate(fit.plus.se = fit + (2*se.fit),
         fit.minus.se = fit - (2*se.fit))

# Plot
ggplot(gam1_pred, aes(pbl.w, fit)) +
  geom_line() + # fit line
  geom_ribbon(aes(ymin = fit.minus.se, ymax = fit.plus.se), # Uncertainty
              fill = "black", alpha = 0.2) +
  geom_point(data = null_tab1, # Original data
             aes(pbl.w, preference.std, color = Test), 
             inherit.aes = FALSE) +
  facet_wrap(~ktype) +
  labs(x = "Tongue length (mm)", y = "Standardized preference", 
       color = "Test (p < 0.01)") +
  theme_light()

ggsave("../output/null_model_ktypeXtonguelength_preference.pdf", height = 6, width = 7)
ggsave("../output/null_model_ktypeXtonguelength_preference.png", height = 6, width = 7)
```

## Model standardized preference by morphotype
```{r}
library(brms)

x <- seq(-2, 2, by = .1)
z <- seq(-20, 20, by = .1)

# This seems like a good place to start
y <- (dnorm(x, mean = 0, sd = 0.5))
plot(x, y, type = "l")

ktype_brm <- brm(preference.std ~ ktype,
                 prior = c(prior(normal(0, 0.5), class = b)),
                 iter = 2000,
                 chains = 4,
                 data = null_tab1)

conditional_effects(ktype_brm)

```

# Null model by site
```{r}
# Nest and join consumer and resource data frames
con_nest <- prop_visits_ktype %>%
  select(site, date, bb.sp, ktype, visits) %>%
  pivot_wider(values_from = visits, names_from = ktype, values_fill = 0) %>%
  select(-c("0", "10", "9")) %>%
  group_by(site) %>%
  nest() %>%
  rename(bees = data)

res_nest <- prop_visits_ktype %>%
  select(site, date, ktype, perc.flower.cover) %>%
  distinct() %>%
  pivot_wider(values_from = perc.flower.cover, names_from = ktype, values_fill = 0) %>%
  select(-c("0", "10", "9")) %>%
  group_by(site) %>%
  nest() %>%
  rename(plants = data)

conres <- full_join(con_nest, res_nest)

# Create a big beautiful table of data and models
sitewise_null_models <- conres %>%

  # Generate null models
  mutate(null.model = map2(bees, plants, function(x, y) 
    generate_null_net(consumers = x[, 2:9],
                      resources = y[, 2:8],
                      sims = 5000,
                      data.type = "counts",
                      c.samples = x[, 1],
                      r.samples = y[, 1]))) %>%

  # Tabulate results (with test)
  mutate(null.model.tab = map(null.model, function(x) 
    test_interactions(x, signif.level = 0.99) %>%
      as_tibble() %>%
      mutate(preference = Observed - Null,
             preference.upper99 = Observed - Upper.99.CL,
             preference.lower99 = Observed - Lower.99.CL) %>%
      select(bb.sp = Consumer, ktype = Resource, everything()))) %>%

  # Unnest and add site data and bumble bee trait data
  unnest(null.model.tab) %>%
  left_join(site_data) %>%
  left_join(bb_traits) %>%

  # Calculate preference/total abundance per species
  group_by(site, bb.sp) %>%
  mutate(total.abund = sum(Observed)) %>%
  ungroup() %>%
  mutate(preference.std = preference/total.abund)

write_rds(sitewise_null_models, "../output/nullmodel_ktype_sitewise.rds")
sitewise_null_models <- readRDS("../output/nullmodel_ktype_sitewise.rds")
```

## Visualize site model
```{r}
ggplot(sitewise_null_models, aes(elev.mean, preference, color = ktype)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp, scales = "free")

ggplot(sitewise_null_models, aes(elev.mean, SES, color = ktype)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp, scales = "free")

ggplot(sitewise_null_models, aes(elev.mean, preference.std, color = ktype)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp, scales = "free")
```

### 2021-04-14 ms_1/Appendix_S1.Rmd
set up null model approach to preference analysis

---
title: "Appendix S1"
output: pdf_document
---

# Packages
```{r message=FALSE, warning=FALSE}
library(mgcViz)
library(lubridate)
library(tidyverse)
```

# Sampling patterns
## Load sampling data
```{r message=FALSE, warning=FALSE}
### BB-FL visitation data
visitation_sampling <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>%
  dplyr::select(site, date, bb.sp, plant.sp) # drop unused columns

### Floral survey data
floral_surveying <- read_csv("../data/processed_data/floral_survey.csv") %>%
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  group_by(plant.sp) %>%
  mutate(present = case_when(
    max(flower.cover) > 0 ~ TRUE,
    max(flower.cover) == 0 ~ FALSE
  )) %>%
  filter(present == TRUE) # Ignore plants that never had any flower cover

site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean, elev.class, elev.class2)) %>% 
  mutate(site = factor(site),
         tree_line = factor(tree_line))
```

## Process sampling data
```{r message=FALSE, warning=FALSE}
bb_obs_per_year <- visitation_sampling %>%
  mutate(year = year(date)) %>%
  group_by(year) %>%
  summarize(obs = n())

bb_obs_per_sample <- visitation_sampling %>%
  mutate(year = factor(year(date))) %>%
  group_by(year, site, date) %>%
  summarize(obs = n()) %>%
  left_join(site_data) %>%
  mutate(date =as_date(yday(date)))

bb_samples_per_year <- visitation_sampling %>%
  mutate(year = year(date)) %>%
  select(year, site, date) %>%
  unique() %>%
  group_by(year) %>%
  summarize(samples = n())

bb_summary <- full_join(bb_obs_per_year, bb_samples_per_year) %>%
  mutate(obs.per.sample = obs/samples)
```

## Visualize
```{r message=FALSE, warning=FALSE}
ggplot(visitation_sampling, aes(x = "x")) +
  geom_bar(fill = "gray60") +
  ylab("Observations") +
  facet_wrap(~year(date)) +
  theme_light(12) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  ggtitle("Bumble bee observations per study year")

ggplot(bb_obs_per_sample, aes(x = "x")) +
  geom_bar(fill = "gray60") +
  ylab("Samples") +
  facet_wrap(~year) +
  theme_light(12) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  ggtitle("Sampling events (site/day) per year")

ggplot(bb_obs_per_sample, aes(x = "x", y = obs)) +
  geom_violin(draw_quantiles = 0.5, fill = "gray60", color = NA) +
  geom_boxplot(fill = NA) +
  #geom_density(outline.type = "full", position = "fill") +
  ylab("Observations/sample") +
  #xlab(NULL) +
  facet_wrap(~year) +
  theme_light(12) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  ggtitle("Bumble bee observations per sampling event")

ggplot(bb_obs_per_sample, aes(as_date(yday(date)), elev.mean/10)) +
  geom_point(size = 1, aes(alpha = obs)) +
  facet_wrap(~year) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_alpha_continuous(name = "Observations") +
  scale_x_date(date_labels = "%b", breaks = as_date(c("1970-04-01", "1970-05-01", "1970-06-01",
                                                    "1970-07-01", "1970-08-01", "1970-09-01"))) +
  theme_light(12) +
  labs(x = NULL,
       y = "Elevation (m x 10)") +
  ggtitle("Bumble bee observations through elevation and time")
```

# Bumble bee abundance
```{r message=FALSE, warning=FALSE}
bb_range <- visitation_sampling %>%
  left_join(site_data) %>%
  group_by(site, date, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = max(abund)) %>%
  filter(abund > 0) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))

ggplot(bb_range, aes(reorder(bb.sp, elev.med), elev.mean)) +
  geom_line(size = 2, alpha = 0.25) +
  geom_point(aes(size = abund, color = tree_line), alpha = 0.5) +
  theme_light(14) +
  scale_size_continuous(name = "Peak abundance",
                        breaks = c(1, 10, 20, 40, 80)) +
  scale_color_discrete(name = "Tree line") +
  xlab("Bumble bee species") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Range and peak abundance of each bumble bee sp.")

# Overall bumble bee abundance
bb_abund_sp <- visitation_sampling %>%
  dplyr::select(site, date, bb.sp) %>%
  group_by(site, date, bb.sp) %>% # 
  summarize(bb.abund = n()) %>% # get per transect bb abundance
  ungroup() %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  semi_join(visitation_sampling, by = c("site", "date")) %>% # drop site-dates that did not exist in the data set
  left_join(site_data) %>% # add site data
  mutate(year = factor(year(date)),
         yday = yday(date),
         bb.sp = factor(bb.sp),
         management = factor(management),
         site = factor(site),
         transect = factor(transect))

bb_abund_sum <- bb_abund_sp %>%
  group_by(site, bb.sp) %>%
  summarize(cum.abund = sum(bb.abund),
            mean.abund = mean(bb.abund)) %>%
  group_by(site) %>%
  mutate(prop.abund = cum.abund/sum(cum.abund)) %>%
  group_by(bb.sp) %>%
  mutate(mean.prop.abund = mean(prop.abund))

ggplot(bb_abund_sum, aes(reorder(bb.sp, -mean.prop.abund), prop.abund)) +
  geom_boxplot(fill = "gray85") +
  theme_light() +
  ylab("Proportional abundance") +
  xlab("Bumble bee species") +
  ggtitle("Relative abundances of bumble bee species")
```

# GAM diagnostics

## Bumble bee abundance GAM
```{r message=FALSE, warning=FALSE}
bb_abund_gam <- readRDS("../output/gam_max_bb_abund_01.rds")
summary(bb_abund_gam)
check.gamViz
```

## Floral abundance GAM
```{r message=FALSE, warning=FALSE}
fl_abund_gam <- readRDS("../output/gam_mean_fl_abund_01.rds")
summary(fl_abund_gam)
check.gamViz(fl_abund_gam)
```

### 2021-04-14 ms_1/ms_1.Rmd
set up null model approach to preference analysis

@@
 ---
-title: "The tree line ecotone punctuates the elevational turnover of plants, pollinators, and their interactions"
+title: "The tree line ecotone punctuates the abundance and turnover of plants, pollinators, and their interactions"
 
@@
 
-Data are not yet provided but will be permanently archived on Dryad if the paper is accepted for publication.
-This submission uses novel code, which is provided as Supporting Information to be evaluated as part of the submission.
-If the paper is accepted, the code will be archived on Dryad alongside the data.
+Data are provided as private-for-peer review.
+This study also uses novel code.
+All data and code required to reproduce our analyses are archived on Dryad: https://datadryad.org/stash/share/M_4Yt9VVXTRm36FSfwwGzUs2OTXP9Fv52jjETdSu6ys
 
@@
 
-Mountains are unique ecological theaters in which extrinsic climate dynamics interact with intrinsic climate gradation to shape the distributions of species and biotic interactions [@Telwala2013-sc; @CaraDonna2014-nq; @Miller-Struttmann2014-wz; @Rafferty2020-aj].
+Mountains are unique ecological theaters in which global climate dynamics interact with local climate gradation to shape the distributions of species and biotic interactions [@Telwala2013-sc; @CaraDonna2014-nq; @Miller-Struttmann2014-wz; Peters2019-km; @Rafferty2020-aj].
 The rate of recent climate change, however, is forcing species to adapt rapidly within the constraints of gene pools shaped by cooler climates [@Visser2008-ct].
@@
 This resulted in a temporally staggered pattern of sampling in 2010 and 2012, while in 2011 a combination of early snowmelt and a late start to field activities resulted in sampling that was approximately aligned in time across the whole elevation gradient.
-See Supplemental Material for details of sampling patterns.
+See Appendix S1 for details of sampling patterns.
 
@@
 This approach was intended to dampen the effects of false non-detection, particularly for rare bumble bee species whose diet breadth would tend to be underestimated simply due to sparsity of observations [@Williams2005-yp].
-We also opted to treat the cryptic *B.
-terrestris/lucorum* species pair as a single morphospecies, since these two species cannot be visually distinguished.
-We similarly pooled representatives of the parasitic *Psithyrus* subgenus --- *B.
+We treated the cryptic *B.
+terrestris/lucorum* species pair as a single morphospecies, since these two species cannot be visually distinguished, and we pooled representatives of the parasitic *Psithyrus* subgenus --- *B.
 barbutellus*, *B.
@@
 psithryus*.
+*B.
+hypnorum*, *B.
+humilis*, and *B.
+lapidarius* were extremely rare in our samples, so we omitted them from abundance analyses but retained them for β-diversity analyses.
+
+Eight plant species --- *Rubus idaeus*, *Rosa canina*, *Juniperus communis*, *Larix decidua*, *Salix* sp., *Caltha palustris*, *Pulmonaria officinalis*, and *Rheum barbarum* --- were recorded in visitation data but not recorded during floral surveying.
+Each accounted for no more than 4 visits in total over the three years of our study, and they were omitted from the analysis of floral abundance.
 
@@
 We analyzed the abundance of bumble bees and floral resources, respectively, using hierarchical generalized additive models (HGAMs) [@Wood2011-pg; @Pedersen2019-kf].
-To focus on elevational patterns, we chose to collapse the temporal variation of abundance within sites into summary values.
+To focus on elevational patterns, we collapsed the temporal variation of abundance within sites into summary values.
 For bumble bees, we modeled *peak* abundance at each site as a smooth response to `elevation`, with bumble bee `species` as a smoothing factor, `management` (grazing, mowing, or unmanaged) as a fixed effect, and `species:year` as a random intercept term (exploratory analysis indicated that species varied differently across years).
@@
 For our analyses of species and interaction β-diversity, we pooled all samples (including across years) within sites, since the sparse observations of individual sampling events tended to inflate β-diversity and its variance to a degree that impaired model fitting and interpretation.
-After pooling, we calculated the species and interaction β-diversity between all pairs of sites using the binary Jaccard index and partitioned total interaction β-diversity into its components of species turnover (changes in species composition) and interaction rewiring (interaction turnover despite species constancy) [@Novotny2009-my; @Poisot2012-fk].
-As an initial analysis, we visualized the relationship between site-wise elevation difference and each metric of β-diversity, fitting each response with a binomial regression smooth.
+After pooling, we calculated species (separately for bumble bees and plants) and interaction β-diversity between all pairs of sites using the binary Jaccard index and partitioned total interaction β-diversity into its components of species turnover (changes in species composition) and interaction rewiring (interaction turnover despite species constancy) [@Novotny2009-my; @Poisot2012-fk].
+Only plant species that were visited at least once by a bumble bee over the course of the study were included in the calculation of floral β-diversity.
 
@@
 
+As an initial analysis, we visualized the relationship between site-wise elevation difference and each metric of β-diversity, fitting each response with a binomial regression smooth.
 To investigate the relationship between elevation and β-diversity more deeply, we performed a second analysis using generalized dissimilarity modeling (GDM) [@Ferrier2007-ji] to analyze the elevational variation in the β-diversity of bumble bees, flora, and (unpartitioned) interactions, respectively.
@@
 GDM analysis of β-diversity was performed with the package `gdm` [@Fitzpatrick-2021aa].
-Annotated `R` code is available in the Supplementary Material.
+Data and code to reproduce these analyses are archived on Dryad: https://datadryad.org/stash/share/M_4Yt9VVXTRm36FSfwwGzUs2OTXP9Fv52jjETdSu6ys.
 
@@
 humilis*, and *B.
-lapidarius* were recorded only sporadically, so we omitted them from abundance analyses but retained them for β-diversity analyses.
-See Supplementary Material for a summary of species-wise abundances.
+lapidarius* were recorded only sporadically.
+See Appendix S1 for a summary of species-wise abundances.
 
-Floral surveying yielded a total of 352 plant species (192 genera and 52 families).
-Of these, 163 species (111 genera and 37 families) were observed to be visited by bumble bees.
-Eight species --- *Rubus idaeus*, *Rosa canina*, *Juniperus communis*, *Larix decidua*, *Salix* sp., *Caltha palustris*, *Pulmonaria officinalis*, and *Rheum barbarum* --- were recorded in visitation data but not recorded during floral surveying.
-Each accounted for no more than 4 visits in total over the three years of our study, and they were omitted from the analysis of floral abundance.
+Floral surveying yielded a total of 352 plant species (192 genera and 52 families). Of these, 163 species (111 genera and 37 families) were observed to be visited by bumble bees.
 
@@
 The random effect of species:year was also nonsignificant (p = 0.52).
-See Supplemental Material for tabulated model output and diagnostics.
+See Appendix S1 for tabulated model output and diagnostics.
 
@@
 The effect of year was also significant, with 2010 (p << 0.0001) and 2011 (p << 0.0001) exhibiting higher mean flower cover than 2012.
-See Supplemental Material for tabulated model output and diagnostics.
+See Appendix S1 for tabulated model output and diagnostics.
 
@@
 
-GDM analysis reproduced the finding that overall β-diversity was dominated by floral turnover, but it also revealed a strongly nonlinear response of floral and interaction β-diversity to elevation **(Figure 4A)**.
+GDM analysis revealed a strongly nonlinear response of floral and interaction β-diversity to elevation **(Figure 4A)**.
 Turnover was high from ~600-1000 m, representing the transition from valley to lower slopes, and then leveled off from ~1000-1500 m.
@@
 
-The outcomes of range shifts in mountain bumble bee and floral communities have obvious significance from a conservation perspective [@Soroye2020-zb], but they also provide a model system in which to study the basic process of community assembly under climate change, a phenomenon of both historical interest and prospective urgency.
+The outcomes of range shifts in mountain bumble bee and floral communities have obvious significance from a conservation perspective [@Soroye2020-zb], but they also provide a model system in which to study the basic process of community assembly under climate change [@Hoiss2012-sn; @Classen2015-jc].
 The central finding of our study, that elevation gradients are characterized by complex and nonlinear responses of species, communities, and biotic interactions --- punctuated by the tree line --- is almost certainly not an idiosyncrasy of bumble bees and their floral hosts but rather a pattern emerging from fundamental physical constraints that should be expected to obtain broadly across the biotic communities of mountains worldwide.
@@
 
-<!-- Retreating to higher elevation in response to climatic warming is not a new phenomenon in the evolutionary history of bumble bees [@Hines2008-er; @Martinet2018-rb].
-Nevertheless, the rapid pace of recent climate change, together with the interacting effects of human land use, are predicted to [@Soroye2020-zb; @Sirois-Delisle2018-vi]  -->
-
-<!-- - What do these findings tell use about climate change and range shifts in mountain ecosystems more generally? -->
-<!-- though as a matter of applied conservation it is perhaps doubtful that local interventions could forestall for long the consequences of a process occurring at a global scale. -->
-
 # ACKNOWLEDGEMENTS
 
-We thank D.J. McNeil for helpful conversations during data analysis.
+We thank the National Park Berchtesgaden for their support and the permission to work in the national park, and we thank the owners of the study sites outside the national park for access to their land.
+The field work of this study was performed within the framework of the joint research center FORKAST as part of the ‘Bavarian Climate Programme 2020'.
+D.J.
+McNeil provided helpful discussion of data analyses.
 
@@
 
-**Figure 1:** Study sites in the Berchtesgadener Alps plotted over Google Earth imagery.
+**Figure 1:** Study sites in the Berchtesgadener Alps plotted over Google Earth imagery (https://earth.google.com/).
 Sites above the tree line are depecited in red and sites below the tree line in yellow.
@@
 Variable importance (B) is mapped to a color ramp, and p-values less than 0.05 are depicted in bold typeface.
+P-values of 0.00 arise from the discreteness of the permutation test and should be interpreted as arbitrarily small.
 

### 2021-04-20 code/ms2_data.Rmd
null model analysis

@@
   dplyr::select(plant.sp, plant.genus, plant.family, k.type, k.type.s, k.type.ss)
+
+write_csv(fl_traits, "../data/processed_data/fl_traits.csv")
 ```
@@
 
+write_csv(visitation, "../data/processed_data/visitation.csv")
+
 # Floral cover data
@@
   mutate(k.type.ss = factor(k.type.ss))
+
+write_csv(visit_cover, "../data/processed_data/visit_cover.csv")
 

### 2021-04-20 code/ms2_ktype_preference_nullmodel_analysis.Rmd
null model analysis

@@
 library(lubridate)
+library(vegan)
+library(ggvegan)
+library(ggrepel)
 ```
@@
          tree_line = factor(tree_line))
+
+visit_cover <- read_csv("../data/processed_data/visit_cover.csv")
 ```
 
-# Null model by species
+# Null model 1: bb.sp x ktype, fully pooled
 ```{r}
@@
 
+## Similarity of preferences across species
+```{r}
+prefs <- null_tab1 %>%
+  select(bb.sp, ktype, preference.std) %>%
+  pivot_wider(names_from = ktype, values_from = preference.std) %>%
+  column_to_rownames("bb.sp")
+
+visits.ktype <- null_tab1 %>%
+  select(bb.sp, ktype, Observed.std) %>%
+  pivot_wider(names_from = ktype, values_from = Observed.std) %>%
+  column_to_rownames("bb.sp")
+
+visits_diffmat <- vegdist(visits, method = "horn") %>%
+  broom::tidy() %>%
+  select(bb.sp1 = 1, bb.sp2 = 2, distance) %>%
+  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>%
+  select(bb.sp1, bb.sp2, distance, pbl.w1 = pbl.w) %>%
+  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>%
+  select(bb.sp1, bb.sp2, distance, pbl.w1, pbl.w2 = pbl.w) %>%
+  mutate(tongue.diff = abs(pbl.w1 - pbl.w2),
+         diet.similarity = 1 - distance) %>%
+  select(-distance)
+
+prefs_diffmat <- vegdist(prefs, method = "euclidean") %>%
+  broom::tidy() %>%
+  select(bb.sp1 = 1, bb.sp2 = 2, distance) %>%
+  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>%
+  select(bb.sp1, bb.sp2, distance, pbl.w1 = pbl.w) %>%
+  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>%
+  select(bb.sp1, bb.sp2, distance, pbl.w1, pbl.w2 = pbl.w) %>%
+  mutate(tongue.diff = abs(pbl.w1 - pbl.w2),
+         preference.similarity = 1 - distance) %>%
+  select(-distance)
+
+visits_prefs <- full_join(visits_diffmat, prefs_diffmat)
+
+ggplot(visits_prefs, aes(preference.similarity, diet.similarity)) +
+  geom_point()
+
+ggplot(visits_prefs, aes(tongue.diff, diet.similarity)) +
+  geom_point()
+
+ggplot(visits_prefs, aes(tongue.diff, preference.similarity)) +
+  geom_point()
+
+pca_pref <- rda(prefs)
+plot(pca_pref)
+
+pca_visit <- rda(visits)
+plot(pca_visit)
+
+tongues <- bb_traits %>%
+  semi_join(null_tab1) %>%
+  arrange(bb.sp)
+
+ordisurf_pref <- ordisurf(pca_pref, tongues$pbl.w) 
+summary(ordisurf_pref)
+plot(ordisurf_pref)
+
+ordisurf_visit <- ordisurf(pca_visit, tongues$pbl.w) 
+summary(ordisurf_visit)
+plot(ordisurf_visit)
+
+fortify(ordisurf_pref)
+fortify(ordisurf_visit)
+
+# Get variance explained
+pcoa_pref_pc1 <- round(pca_pref$CA$eig[[1]] / sum(pca_pref$CA$eig)*100)
+pcoa_pref_pc2 <- round(pca_pref$CA$eig[[2]] / sum(pca_pref$CA$eig)*100)
+
+pcoa_visit_pc1 <- round(pca_visit$CA$eig[[1]] / sum(pca_visit$CA$eig)*100)
+pcoa_visit_pc2 <- round(pca_visit$CA$eig[[2]] / sum(pca_visit$CA$eig)*100)
+
+# Fortify whole model
+pca_pref.fort <- fortify(pca_pref, axes = 1:2) 
+pca_visit.fort <- fortify(pca_visit, axes = 1:2)
+
+# Fortify plants
+pca_pref.fort1 <- fortify(pca_pref, axes = 1:2) %>%
+  filter(Score == "species") %>%
+  rename(ktype = Label)
+
+pca_visit.fort1 <- fortify(pca_visit, axes = 1:2) %>%
+  filter(Score == "species") %>%
+  rename(ktype = Label)
+
+# Fortify bees
+pca_pref.fort2 <- fortify(pca_pref, axes = 1:2) %>%
+  filter(Score == "sites") %>%
+  rename(bb.sp = Label) %>%
+  left_join(bb_traits)
+
+pca_visit.fort2 <- fortify(pca_visit, axes = 1:2) %>%
+  filter(Score == "sites") %>%
+  rename(bb.sp = Label) %>%
+  left_join(bb_traits)
+
+# Generate convex hulls around tongue length classes
+# pca_pref_hulls <- pca_pref.fort2 %>%
+#   group_by(tongue) %>%
+#   slice(chull(PC1, PC2))
+
+# Plot
+ggplot(pca_pref.fort2, aes(PC1, PC2, color = tongue)) +
+  geom_polygon(data = pca_pref_hulls, 
+               aes(fill = tongue, color = NULL),
+               alpha = 0.25) +
+  geom_point(data = pca_pref.fort1, aes(PC1, PC2), inherit.aes = FALSE, color = "gray40", size = 2) +
+  geom_text(aes(label = bb.sp), key_glyph = "rect", size = 3) +
+  geom_text_repel(data = pca_pref.fort1, 
+            aes(PC1, PC2, label = ktype), inherit.aes = FALSE, color = "gray40",
+            size = 3) +
+  scale_shape_discrete() +
+  xlab(paste("PC1 ", "(", pcoa_pref_pc1, "%", ")", sep = "")) +
+  ylab(paste("PC2 ", "(", pcoa_pref_pc2, "%", ")", sep = "")) +
+  labs(fill = "Tongue length", 
+       shape = "Morphotype") +
+  guides(color = FALSE) +
+  coord_fixed() +
+  theme_light(9)
+
+# Plot ordisurf
+## Based on https://oliviarata.wordpress.com/2014/07/17/ordinations-in-ggplot2-v2-ordisurf/
+surfplot <- function(x, y, min_score = 0) {
+
+  grid <- x$grid # extracts the ordisurf object
+  grid_exp <- expand.grid(x = grid$x, y = grid$y) # get x and ys
+  grid_exp$z <- as.vector(grid$z) # unravel the matrix for the z scores
+
+  contours <- data.frame(na.omit(grid_exp)) %>% # gets rid of the nas
+    as_tibble() %>%
+    rename(PC1 = x, PC2 = y)
+
+  scores <- y %>% 
+    filter((abs(PC1) > min_score | abs(PC2) > min_score) | Score == "BB.sp") %>%
+    mutate(max.score = max(PC1, PC2))
+
+  ggplot(scores, aes(PC1, PC2, pch = Score)) +
+    geom_contour(data = contours, 
+                 aes(PC1, PC2, z = z, color = stat(level)),
+                 size = 1.5, alpha = 0.75, inherit.aes = FALSE) +
+    #geom_point(size = 2) +
+    geom_text_repel(aes(PC1, PC2, label = Label), key_glyph = "rect") +
+    #geom_text(aes(MDS1, MDS2, label = Label, alpha = max.score), key_glyph = "rect") +
+    theme_light() +
+    scale_color_viridis_c(name = "Tongue length") +
+    #scale_shape_manual(values = c(16, 1), name = "Level") +
+    coord_fixed() # Gavin Simpson says this is **really** important
+}
+
+surfplot(ordisurf_pref, pca_pref.fort)
+surfplot(ordisurf_visit, pca_visit.fort)
+```
+
 ## Visualize network
@@
   scale_y_discrete(limits = rev) +
-  scale_color_manual(values = c("gray40", "#00B6EB", "#F8766D")) +
+  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
   labs(x = "Standardized visits", y = "Floral morphotype", color = "Test (p < 0.01)") +
@@
   geom_point(aes(color = Test)) +
-  facet_wrap(~ktype) +
+  facet_wrap(~ktype, ncol = 4) +
   scale_y_discrete(limits = rev) +
-  scale_color_manual(values = c("gray40", "#00B6EB", "#F8766D")) +
+  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
   labs(x = "Standardized visits", y = "Bumble bee species", color = "Test (p < 0.01)") +
@@
 
-ggsave("../output/null_model_ktypewise_preference.pdf", height = 6, width = 7)
-ggsave("../output/null_model_ktypewise_preference.png", height = 6, width = 7)
+ggsave("../output/null_model_ktypewise_preference.pdf", height = 5, width = 7)
+ggsave("../output/null_model_ktypewise_preference.png", height = 5, width = 7)
 
@@
 ggplot(null_tab1, aes(ktype, preference.std)) +
-  #geom_violin(fill = "#F8766D", color = "#F8766D") +
-  geom_violin(fill = "gray70", color = "gray70") +
-  geom_boxplot(fill = NA, color = "black") +
+  #geom_violin(fill = "gray70", color = "white", draw_quantiles = c(0.25, 0.5, 0.75)) +
+  geom_boxplot(fill = NA, color = "black", outlier.shape = 4, size = 0.35) +
+  geom_jitter(aes(color = Test), width = 0.3, alpha = 0.6) +
   theme_light(14) +
+  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
   labs(x = "Floral morphotype", 
@@
 ggsave("../output/null_model_ktypeXtongueclass_preference.png", height = 6, width = 7)
-
-# Standardized preference ~ tongue length
-ggplot(null_tab1, aes(pbl.w, preference.std)) +
-  geom_point(aes(color = Test)) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5), color = "gray20", size = 0.5) +
-  facet_wrap(~ktype, scales = "free") +
-  theme_light()
 ```
@@
 summary(gam1)
-check.gamViz(gam1)
-plot.gamViz(gam1, allTerms = TRUE)
+#check.gamViz(gam1)
+#plot.gamViz(gam1, allTerms = TRUE)
 
@@
              inherit.aes = FALSE) +
+  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
   facet_wrap(~ktype) +
@@
 
-# Null model by site
+# Null model 2: bb.sp x ktype, by site
 ```{r}
@@
 # Create a big beautiful table of data and models
-sitewise_null_models <- conres %>%
 
-  # Generate null models
-  mutate(null.model = map2(bees, plants, function(x, y) 
-    generate_null_net(consumers = x[, 2:9],
-                      resources = y[, 2:8],
-                      sims = 5000,
-                      data.type = "counts",
-                      c.samples = x[, 1],
-                      r.samples = y[, 1]))) %>%
+## RUN ON HPC!
+# sitewise_null_models <- conres %>%
+#   
+#   # Generate null models
+#   mutate(null.model = map2(bees, plants, function(x, y) 
+#     generate_null_net(consumers = x[, 2:9],
+#                       resources = y[, 2:8],
+#                       sims = 2000,
+#                       data.type = "counts",
+#                       c.samples = x[, 1],
+#                       r.samples = y[, 1]))) %>%
+#   
+#   # Tabulate results (with test)
+#   mutate(null.model.tab = map(null.model, function(x) 
+#     test_interactions(x, signif.level = 0.99) %>%
+#       as_tibble() %>%
+#       mutate(preference = Observed - Null,
+#              preference.upper99 = Observed - Upper.99.CL,
+#              preference.lower99 = Observed - Lower.99.CL) %>%
+#       select(bb.sp = Consumer, ktype = Resource, everything()))) %>%
+#   
+#   # Unnest and add site data and bumble bee trait data
+#   unnest(null.model.tab) %>%
+#   left_join(site_data) %>%
+#   left_join(bb_traits) %>%
+#   
+#   # Calculate preference/total abundance per species
+#   group_by(site, bb.sp) %>%
+#   mutate(total.abund = sum(Observed)) %>%
+#   ungroup() %>%
+#   mutate(preference.std = preference/total.abund)
 
-  # Tabulate results (with test)
-  mutate(null.model.tab = map(null.model, function(x) 
-    test_interactions(x, signif.level = 0.99) %>%
-      as_tibble() %>%
-      mutate(preference = Observed - Null,
-             preference.upper99 = Observed - Upper.99.CL,
-             preference.lower99 = Observed - Lower.99.CL) %>%
-      select(bb.sp = Consumer, ktype = Resource, everything()))) %>%
+#write_rds(sitewise_null_models, "../output/nullmodel_ktype_sitewise.rds")
+#write_rds(sitewise_null_models, "./nullmodel_ktype_sitewise.rds") # hpc
+sitewise_null_models <- readRDS("../output/nullmodel_ktype_sitewise.rds")
 
-  # Unnest and add site data and bumble bee trait data
-  unnest(null.model.tab) %>%
-  left_join(site_data) %>%
-  left_join(bb_traits) %>%
+sitewise_null_models <- sitewise_null_models %>%
+  filter(total.abund > 0) %>% # for each species, drop sites where the species did not occur
+  unite(bb.ktype, c(bb.sp, ktype), remove = FALSE) %>%
+  mutate(ktype = factor(ktype),
+         bb.sp = factor(bb.sp),
+         bb.ktype = factor(bb.ktype))
+```
 
-  # Calculate preference/total abundance per species
-  group_by(site, bb.sp) %>%
-  mutate(total.abund = sum(Observed)) %>%
-  ungroup() %>%
-  mutate(preference.std = preference/total.abund)
+## Diet similarity vs. preference similarity
+```{r}
+prefs2 <- sitewise_null_models %>%
+  select(site, bb.sp, ktype, preference.std) %>%
+  unite(site.bb.sp, c(site, bb.sp)) %>%
+  pivot_wider(names_from = ktype, values_from = preference.std) %>%
+  column_to_rownames("site.bb.sp")
 
-write_rds(sitewise_null_models, "../output/nullmodel_ktype_sitewise.rds")
-sitewise_null_models <- readRDS("../output/nullmodel_ktype_sitewise.rds")
+visits.ktype <- null_tab1 %>%
+  select(bb.sp, ktype, Observed.std) %>%
+  pivot_wider(names_from = ktype, values_from = Observed.std) %>%
+  column_to_rownames("bb.sp")
+
+visits_diffmat <- vegdist(visits, method = "horn") %>%
+  broom::tidy() %>%
+  select(bb.sp1 = 1, bb.sp2 = 2, distance) %>%
+  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>%
+  select(bb.sp1, bb.sp2, distance, pbl.w1 = pbl.w) %>%
+  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>%
+  select(bb.sp1, bb.sp2, distance, pbl.w1, pbl.w2 = pbl.w) %>%
+  mutate(tongue.diff = abs(pbl.w1 - pbl.w2),
+         diet.similarity = 1 - distance) %>%
+  select(-distance)
+
+prefs_diffmat <- vegdist(prefs, method = "euclidean") %>%
+  broom::tidy() %>%
+  select(bb.sp1 = 1, bb.sp2 = 2, distance) %>%
+  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>%
+  select(bb.sp1, bb.sp2, distance, pbl.w1 = pbl.w) %>%
+  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>%
+  select(bb.sp1, bb.sp2, distance, pbl.w1, pbl.w2 = pbl.w) %>%
+  mutate(tongue.diff = abs(pbl.w1 - pbl.w2),
+         preference.similarity = 1 - distance) %>%
+  select(-distance)
+
+visits_prefs <- full_join(visits_diffmat, prefs_diffmat)
+
 ```
@@
 ```{r}
-ggplot(sitewise_null_models, aes(elev.mean, preference, color = ktype)) +
+# ktype preference x bb.sp ~ elevation
+ggplot(sitewise_null_models, aes(elev.mean, preference.std, color = ktype)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
+  facet_wrap(~bb.sp)
+
+ggplot(filter(sitewise_null_models, ktype == 1), aes(elev.mean, preference.std)) +
+  geom_point(aes(color = Test)) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
+  facet_wrap(~bb.sp)
+
+ggplot(filter(sitewise_null_models, ktype == 2), aes(elev.mean, preference.std)) +
+  geom_point(aes(color = Test)) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
+  facet_wrap(~bb.sp)
+
+ggplot(filter(sitewise_null_models, ktype == 3), aes(elev.mean, preference.std)) +
+  geom_point(aes(color = Test)) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~bb.sp)
+
+ggplot(filter(sitewise_null_models, ktype == 4), aes(elev.mean, preference.std)) +
+  geom_point(aes(color = Test)) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~bb.sp)
+
+ggplot(filter(sitewise_null_models, ktype == 5), aes(elev.mean, preference.std)) +
+  geom_point(aes(color = Test)) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~bb.sp)
+
+ggplot(filter(sitewise_null_models, ktype == 6), aes(elev.mean, preference.std)) +
+  geom_point(aes(color = Test)) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~bb.sp)
+
+ggplot(filter(sitewise_null_models, ktype == 7), aes(elev.mean, preference.std)) +
+  geom_point(aes(color = Test)) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~bb.sp)
+
+gam_k7 <- gamV(preference.std ~
+                 s(bb.sp, bs = "re") +
+                 s(elev.mean, by = bb.sp, k = 5),
+    data = filter(sitewise_null_models, ktype == 7))
+
+gam_bb.ktype <- bam(preference.std ~
+                      s(bb.ktype, bs = "re") +
+                      s(elev.mean, by = bb.ktype, k = 4),
+                    method = "fREML",
+                    discrete = TRUE,
+                    data = sitewise_null_models) %>% getViz()
+
+print(plot.gamViz(gam_bb.ktype, allTerms = TRUE), pages = 6)
+check.gamViz(gam_bb.ktype)
+summary(gam_bb.ktype)
+
+# ktype ~ elevation
+ggplot(sitewise_null_models, aes(elev.mean, preference.std)) +
+  geom_point(aes(color = Test)) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~ktype)
+
+ggplot(sitewise_null_models, aes(elev.mean, preference.std, color = pbl.w.class2)) +
   geom_point() +
   geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
-  facet_wrap(~bb.sp, scales = "free")
+  facet_wrap(~ktype)
 
-ggplot(sitewise_null_models, aes(elev.mean, SES, color = ktype)) +
-  geom_point() +
+# ktype ~ tongue length x site
+ggplot(filter(sitewise_null_models, ktype == 1), aes(pbl.w, preference.std)) +
+  geom_point(aes(color = Test)) +
   geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
-  facet_wrap(~bb.sp, scales = "free")
+  facet_wrap(~site)
 
-ggplot(sitewise_null_models, aes(elev.mean, preference.std, color = ktype)) +
-  geom_point() +
+ggplot(filter(sitewise_null_models, ktype == 2), aes(pbl.w, preference.std)) +
+  geom_point(aes(color = Test)) +
   geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
-  facet_wrap(~bb.sp, scales = "free")
+  facet_wrap(~site)
+
+ggplot(filter(sitewise_null_models, ktype == 3), aes(pbl.w, preference.std)) +
+  geom_point(aes(color = Test)) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~site)
+
+ggplot(filter(sitewise_null_models, ktype == 4), aes(pbl.w, preference.std)) +
+  geom_point(aes(color = Test)) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~site)
+
+ggplot(filter(sitewise_null_models, ktype == 5), aes(pbl.w, preference.std)) +
+  geom_point(aes(color = Test)) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~site)
+
+ggplot(filter(sitewise_null_models, ktype == 6), aes(pbl.w, preference.std)) +
+  geom_point(aes(color = Test)) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~site)
+
+ggplot(filter(sitewise_null_models, ktype == 7), aes(pbl.w, preference.std)) +
+  geom_point(aes(color = Test)) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~site)
 ```
+
+## Overall preference of each ktype ~ elevation
+```{r}
+
+```
+
+# Null model 3: bb.sp x plant genus, fully pooled
+```{r}
+# The "consumers" table input to generate_null_net()
+con3 <- visit_cover %>%
+  group_by(site, date, bb.sp, plant.genus) %>%
+  summarize(visits = sum(visits)) %>%
+  unite(id, c(site, date)) %>%
+  select(id, bb.sp, plant.genus, visits) %>%
+  pivot_wider(values_from = visits, names_from = plant.genus, values_fill = 0)
+
+# The "resources" table input to generate_null_net()
+res3 <- visit_cover %>%
+  group_by(site, date, bb.sp, plant.genus) %>%
+  summarize(flower.cover = sum(flower.cover)) %>%
+  group_by(site, date) %>%
+  mutate(perc.flower.cover = flower.cover/sum(flower.cover)) %>%
+  unite(id, c(site, date)) %>%
+  select(id, plant.genus, perc.flower.cover) %>%
+  distinct() %>%
+  pivot_wider(values_from = perc.flower.cover, names_from = plant.genus, values_fill = 0)
+
+# Generate null model
+null3 <- generate_null_net(consumers = con3[, 2:201],
+                           resources = res3[, 2:200],
+                           sims = 5000, # raised from default of 1000 (takes a while to run)
+                           data.type = "counts",
+                           c.samples = con3[, 1], # parsed by sample
+                           r.samples = res3[, 1])
+
+saveRDS(null3, file = "./nullmodel_plant.genus.rds") # hpc
+null3 <- readRDS("../output/nullmodel_ktype.rds")
+
+# Tabulate output and generate tests
+null_tab1 <- test_interactions(null1, signif.level = 0.99) %>% # make it a bit more conservative to account for multiple comparisons
+  as_tibble() %>%
+  select(bb.sp = Consumer, ktype = Resource, everything()) %>%
+  group_by(bb.sp) %>%
+  mutate(total.abund = sum(Observed), # total abundance per species
+         preference = Observed - Null, # a working definition of preference
+         preference.upper99 = Observed - Upper.99.CL, 
+         preference.lower99 = Observed - Lower.99.CL,
+
+         # If preference equals the difference between # observed visits and the null model prediction, then dividing preference by the total abundance of each species yields a standardized preference.
+Right?
+         preference.std = preference/total.abund, 
+         Observed.std = Observed/total.abund,
+         Upper.99.CL.std = Upper.99.CL/total.abund,
+         Lower.99.CL.std = Lower.99.CL/total.abund) %>%
+  left_join(bb_traits) %>%
+  ungroup() %>%
+  mutate(bb.sp = fct_reorder(factor(bb.sp), pbl.w))
+```
+
+# Null model 4: bb.sp x plant genus, fully pooled
+```{r}
+# The "consumers" table input to generate_null_net()
+con4 <- visit_cover %>%
+  group_by(site, date, bb.sp, plant.family) %>%
+  summarize(visits = sum(visits)) %>%
+  unite(id, c(site, date)) %>%
+  select(id, bb.sp, plant.family, visits) %>%
+  pivot_wider(values_from = visits, names_from = plant.family, values_fill = 0)
+
+# The "resources" table input to generate_null_net()
+res4 <- visit_cover %>%
+  group_by(site, date, bb.sp, plant.family) %>%
+  summarize(flower.cover = sum(flower.cover)) %>%
+  group_by(site, date) %>%
+  mutate(perc.flower.cover = flower.cover/sum(flower.cover)) %>%
+  unite(id, c(site, date)) %>%
+  select(id, plant.family, perc.flower.cover) %>%
+  distinct() %>%
+  pivot_wider(values_from = perc.flower.cover, names_from = plant.family, values_fill = 0)
+
+# Generate null model
+null4 <- generate_null_net(consumers = con4[, 2:56],
+                           resources = res4[, 2:55],
+                           sims = 5000, # raised from default of 1000 (takes a while to run)
+                           data.type = "counts",
+                           c.samples = con4[, 1], # parsed by sample
+                           r.samples = res4[, 1])
+
+saveRDS(null4, file = "./nullmodel_plant.family.rds") # hpc
+null3 <- readRDS("../output/nullmodel_ktype.rds")
+
+# Tabulate output and generate tests
+null_tab1 <- test_interactions(null1, signif.level = 0.99) %>% # make it a bit more conservative to account for multiple comparisons
+  as_tibble() %>%
+  select(bb.sp = Consumer, ktype = Resource, everything()) %>%
+  group_by(bb.sp) %>%
+  mutate(total.abund = sum(Observed), # total abundance per species
+         preference = Observed - Null, # a working definition of preference
+         preference.upper99 = Observed - Upper.99.CL, 
+         preference.lower99 = Observed - Lower.99.CL,
+
+         # If preference equals the difference between # observed visits and the null model prediction, then dividing preference by the total abundance of each species yields a standardized preference.
+Right?
+         preference.std = preference/total.abund, 
+         Observed.std = Observed/total.abund,
+         Upper.99.CL.std = Upper.99.CL/total.abund,
+         Lower.99.CL.std = Lower.99.CL/total.abund) %>%
+  left_join(bb_traits) %>%
+  ungroup() %>%
+  mutate(bb.sp = fct_reorder(factor(bb.sp), pbl.w))
+```

### 2021-04-20 code/ms2_niche_overlap.Rmd
null model analysis

--- 2021-04-06 code/ms2_bias_niche.Rmd
@@
 library(bipartite)
-library(brms)
-library(bayesplot)
-library(tidybayes)
 library(mgcv)
 library(mgcViz)
-library(DHARMa)
 library(modelr)
-library(ggdist)
 library(vegan)
@@
 # FL traits
-## family-level taxonomy
-fl_tax <- read_csv("../data/processed_data/floral_tax.csv")
+fl_traits <- read_csv("../data/processed_data/fl_traits.csv")
 
-## Kugler morphotypes (accessed through Bioflor)
-fl_ktype <- read_csv("../data/processed_data/floral_k_type.csv")
-
-## verbal description of Kugler morphotypes
-fl_ktype_key <- read_delim("../data/processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
-  dplyr::select(k.type = 1, description = 2)
-
-## join all floral trait tables
-fl_traits <- fl_tax %>%
-  full_join(fl_ktype) %>%
-  dplyr::select(plant.sp, plant.genus, plant.family, k.type, k.type.s, k.type.ss) 
-```
-
-## Align survey and visitation data with pseudodates
-```{r}
-# Site-dates where BB visitation was recorded (excluding dates where only males were recorded)
-visitation_samples <- read_csv("../data/processed_data/network.csv") %>%
-  filter(caste != "male") %>%
-  select(site, date) %>%
-  distinct()
-
-# Site dates where floral surveying was conducted
-survey_samples <- read_csv("../data/processed_data/floral_survey.csv") %>%
-  select(site, date) %>%
-  distinct()
-
-# Site-dates where visitation sampling and floral sampling coincided
-shared_dates <- inner_join(visitation_samples, survey_samples)
-
-# Site-dates where visitation was recorded WITHOUT corresponding survey data
-visitation_only <- anti_join(visitation_samples, survey_samples) %>%
-  mutate(date.v = date) %>%
-  select(-date)
-
-# This creates a pseudo date column for unmatched visitation site-dates that will match each visitation date
-# to its nearest survey data. For site t4, there is a 2-year mismatch; we will drop these site-dates later.
-pseudodates <- full_join(visitation_only, survey_samples) %>%
-  mutate(mismatch = abs(date.v - date)) %>%
-  group_by(site, date.v) %>%
-  filter(mismatch == min(mismatch)) %>%
-  rename(date = date.v, pseudo.date = date)
-```
-
-## Generate a data frame that represents visitation and survey abundance
-```{r}
-# BB visitation data
-visitation <- read_csv("../data/processed_data/network.csv") %>%
-  filter(caste != "male") %>% # we will ignore males 
-  filter(!bb.sp %in% c("humi")) %>% # humi too rare
-  group_by(site, date, bb.sp, plant.sp) %>%
-  summarize(visits = n()) %>% # visits per BB-FL pair
-  ungroup() %>%
-  left_join(pseudodates) %>% # Now we will replace unmatched dates with their pseudo dates
-  replace_na(replace = list(mismatch = 0)) %>% # set mismatch to zero for perfectly aligned dates
-  filter(mismatch <= 7) %>% # this gets rid of irreconcilable dates (like the t4 dates mentioned above)
-  mutate(date2 = if_else(is.na(pseudo.date), date, pseudo.date)) %>% # replace date with pseudodate for all site-dates lacking perfect alignments
-  select(-c(date, pseudo.date, mismatch)) %>%
-  rename(date = date2) %>%
-  group_by(site, date) %>%
-  complete(bb.sp, plant.sp, fill = list(visits = 0)) # generate all possible BB-FL combos for each site date and fill in zeros for visits
-
-# Floral cover data
-flower_cover <- read_csv("../data/processed_data/floral_survey.csv") %>%
-  filter(flower.cover > 0) %>% # ignore plants that were not blooming
-  group_by(site, date, plant.sp) %>%
-  summarize(flower.cover = sum(flower.cover))
-
-# Get shared dates after adding pseudodates. This will let us filter out survey dates that don't
-# have corresponding visitation sampling dates
-shared_dates_with_pseudo <- inner_join(visitation, flower_cover, by = c("site", "date")) %>%
-  select(site, date) %>%
-  distinct()
-
-# Join visitation and flower cover data
-
-## First we need a list of all BB spp. by site and date
-sp_by_site <- visitation %>% 
-  ungroup() %>%
-  select(site, date, bb.sp) %>%
-  distinct()
-
-## Then we need a list of unvisited plant.sp by site and date
-unvisited <- flower_cover %>% 
-  semi_join(shared_dates_with_pseudo) %>% # filter flower cover data to include only site-dates with visitation complement (real or pseudodates)
-  full_join(visitation) %>% # join to visitation data
-  filter(is.na(bb.sp)) %>% # select only the site-date-plants that were totally unvisited (and therefore have no BB spp to join to)
-  select(-bb.sp) %>% # drop bb.sp variable because I just want the plants
-  left_join(sp_by_site) %>% # join in the species list by site and date, thus populating the unvisited plants with all the BB spp recorded for each site-date
-  replace_na(replace = list(visits = 0)) # replace NAs (created in the line above) with zero
-
-## Now we can combine everything into a single data frame
-visit_cover <- flower_cover %>% 
-  semi_join(shared_dates_with_pseudo) %>% # filter flower cover data to include only site-dates with visitation complement (real or pseudodates)
-  full_join(visitation) %>% # join to visitation data
-  replace_na(replace = list(flower.cover = 0.01)) %>% # If a plant was visited but not surveyed, we will assign it the minimum non-zero flower cover value of 0.01.
-  filter(!is.na(bb.sp)) %>% # drop the univisited plants
-  bind_rows(unvisited) %>% # add them back, with BB spp and zeros
-  select(site, date, bb.sp, plant.sp, visits, flower.cover) %>%
-  group_by(site, date, bb.sp, plant.sp, flower.cover) %>% # In a few cases, the use of pseudodates causes two sampling period to converge on the same date, which creates apparent duplicate visitation data.
-We will handle this by averaging the visits.
-  summarize(visits = mean(visits)) %>%
-  left_join(fl_traits) %>% # add floral traits
-  ungroup() %>%
-  mutate(k.type.ss = factor(k.type.ss))
-
-## Get total visits by BB sp and site-date; this will be a measure of BB abundance that we can use after we've calculated d'.
-I could, of course, work this into the biaseval function, but I'll put that off for now.
-bb_abundance <- visit_cover %>%
-  group_by(site, date, bb.sp) %>%
-  summarize(bb.sp.abund = sum(visits)) %>%
-  group_by(site, date) %>%
-  mutate(bb.sp.prop.abund = bb.sp.abund/sum(bb.sp.abund))
-
-fl_abundance <- visit_cover %>%
-  group_by(site, date) %>%
-  summarize(total.flower.cover = sum(flower.cover))
-```
-
-# 1. Overall bias
-
-## 1.1 Calculate visitation probability per plant morphotype.
-Instead of keeping the plant species resolved, we will pool visitation by morphotype.
-The problem with not pooling by morphotype is that more species-rich morphotypes will be penalized simply because their visitation is diluted across more species.
-```{r}
-prop_visits_ktype <- visit_cover %>%
-  group_by(site, date, bb.sp, k.type.ss) %>%
-  summarize(flower.cover = sum(flower.cover), # get raw flower cover and visits by site*date*bb.sp*plant.sp
-            visits = sum(visits)) %>%
-  group_by(site, date) %>%
-  mutate(prop.flower.cover = flower.cover/sum(flower.cover))%>% # binary visits
-  group_by(site, date, bb.sp) %>%
-  mutate(prop.visits = visits/sum(visits),
-         bb.sp.abund = sum(visits)) %>% # abundance of each BB sp. per site-date
-  ungroup() %>%
-  left_join(bb_traits) %>% # add BB traits
-  mutate(ktype = factor(k.type.ss), # recode variables
-         tongue = factor(pbl.w.class2),
-         visits = as.integer(visits)) %>% # so we can run a Poisson model
-  left_join(site_data) %>% # add site data
-  ungroup() %>%
-  mutate(yday = yday(date),
-         elev.mean0 = scale(elev.mean),
-         yday0 = scale(yday),
-         prop.flower.cover0 = scale(prop.flower.cover),
-         perc.flower.cover = prop.flower.cover*100,
-         bb.sp.abund0 = scale(bb.sp.abund)) %>%
-  select(site, elev.mean, date, yday, bb.sp, ktype, visits, 
-         prop.visits, flower.cover, perc.flower.cover, bb.sp.abund, tongue,
-         elev.mean0, yday0, prop.flower.cover0, bb.sp.abund0) %>%
-  mutate(visits.bin = if_else(visits > 0, 1, 0))
-
-write_csv(prop_visits_ktype, "../hpc_packets/prop_visits_ktype.csv")
-
-prop_visits_ktype_pooled <- visit_cover %>%
-  group_by(site, date, k.type.ss) %>%
-  summarize(flower.cover = sum(flower.cover), # get raw flower cover and visits by site*date*bb.sp*plant.sp
-            visits = sum(visits)) %>%
-  group_by(site, date) %>%
-  mutate(prop.flower.cover = flower.cover/sum(flower.cover),
-         prop.visits = visits/sum(visits),
-         bb.abund = sum(visits),
-         visits = as.integer(visits)) %>%
-  mutate(ktype = factor(k.type.ss)) %>%
-  left_join(site_data) %>% # add site data
-  ungroup() %>%
-  mutate(yday = yday(date),
-         elev.mean0 = scale(elev.mean),
-         yday0 = scale(yday),
-         perc.flower.cover = prop.flower.cover*100,
-         prop.flower.cover0 = scale(prop.flower.cover),
-         bb.abund0 = scale(bb.abund)) %>%
-  select(site, elev.mean, date, yday, ktype, visits, 
-         prop.visits, bb.abund, perc.flower.cover,
-         elev.mean0, yday0, prop.flower.cover0, bb.abund0) %>%
-  mutate(visits.bin = if_else(visits > 0, 1, 0),
-         year = factor(year(date)))
-
-write_csv(prop_visits_ktype_pooled, "../hpc_packets/prop_visits_ktype_pool.csv")
-
-```
-
-## 1.2 Binomial multi-level regression models
-### 1.2.1 Choose priors
-```{r}
-# For starters, I'm just going to replace the flat prior on the fixed effects we a weakly regularizing prior (on the logit scale)
-brm_pvis_prior <- prior(normal(0, 1.5), 
-                        class = b) 
-```
-
-### 1.2.2 Call models
-All these run cleanly.
-Well, brm_pvis_01 throws maximum tree depth warnings, but these are just about efficiency, not validity.
-These warnings are evidently a consequence of increasing adapt_delta to 0.99, which succeeded in getting rid of the divergent transition warnings (those warnings are much more serious and pertain to the validity of the model).
-```{r}
-## Exploring the differential bias of bumble bee species and tongue length classes towards floral morphotypes
-### Bernoulli with binarized visitation
-brm_pvis_00 <- brm(as.integer(visits.bin) ~
-                    bb.sp.abund*perc.flower.cover + # fixed term: because more abundant flowers will tend to be more visited,and more abundance bees visit more
-                    (1 + perc.flower.cover | ktype:bb.sp) + # a random intercept and random slope for ktype:species; I think I need to model this before I can make conclusions about ktype:pbl.class
-                    (1 + perc.flower.cover | ktype:tongue) + # a random intercept and slope for ktype:tongue
-                    (1 | ktype:site), # a varying intercept and slope for ktype:site
-                  family = "bernoulli", 
-                  prior = brm_pvis_prior, # weakly regularizing
-                  file = "../output/brm_pvis_ktype_00",  # for running locally
-                  #file = "brm_pvis_ktype_00", # for running on the HPC
-                  chains = 4,
-                  #cores = 12,
-                  iter = 5000,
-                  data = filter(prop_visits_ktype, !ktype %in% c("0", "9", "10")))
-
-brm_pvis_01 <- brm(as.integer(visits.bin) ~
-                    bb.abund*perc.flower.cover + # fixed term: because more abundant flowers will tend to be more visited,and more abundance bees visit more
-                    (1 + perc.flower.cover | ktype) + # a random intercept and slope for ktype
-                    (1 | ktype:site),  # a varying intercept term for ktype:site
-                  family = "bernoulli", 
-                  prior = brm_pvis_prior, # weakly regularizing
-                  control = list(adapt_delta = 0.99), # the default setting yielded divergent transitions
-                  file = "../output/brm_pvis_ktype_01",  # for running locally
-                  #file = "brm_pvis_ktype_01", # for running on the HPC
-                  chains = 4,
-                  #cores = 12,
-                  iter = 5000,
-                  data = filter(prop_visits_ktype_pooled, !ktype %in% c("0", "9", "10")))
-
-# If I want to make inference about species visitation rates, I don't need the tongue-length class effect; I just need the site effect.
-But, when I want to make inference about the tongue length classes, I do need the species-level effect.
-brm_pvis_02 <- brm(as.integer(visits.bin) ~
-                    bb.sp.abund*perc.flower.cover + # fixed term: because more abundant flowers will tend to be more visited,and more abundance bees visit more
-                    (1 + perc.flower.cover | ktype:bb.sp) + # a random intercept and random slope for ktype:species; I think I need to model this before I can make conclusions about ktype:pbl.class
-                    (1 | ktype:site), # a varying intercept and slope for ktype:site
-                  family = "bernoulli", 
-                  prior = brm_pvis_prior, # weakly regularizing
-                  file = "../output/brm_pvis_ktype_02",  # for running locally
-                  #file = "brm_pvis_ktype_02", # for running on the HPC
-                  chains = 4,
-                  #cores = 12,
-                  iter = 5000, # Bulk ESS was a bit low wit 2000 iter
-                  data = filter(prop_visits_ktype, !ktype %in% c("0", "9", "10")))
-```
-
-## 1.3 Evaluate models
-```{r eval=FALSE, include=FALSE}
-# Summary
-summary(brm_pvis_00)
-summary(brm_pvis_01)
-summary(brm_pvis_02)
-
-# Posterior predictive checks
-pp_check(brm_pvis_00)
-pp_check(brm_pvis_01)
-pp_check(brm_pvis_02)
-
-# MCMC convergence
-plot(brm_pvis_00)
-plot(brm_pvis_01)
-plot(brm_pvis_02)
-
-# Bayes R2
-bayes_R2(brm_pvis_00)
-bayes_R2(brm_pvis_01)
-bayes_R2(brm_pvis_02)
-```
-
-## 1.4 Preliminary visualization
-```{r eval=FALSE, include=FALSE}
-get_variables(brm_pvis_00)
-conditional_effects(brm_pvis_00)
-mcmc_plot(brm_pvis_00, pars = c("^r_ktype:tongue.*Intercept]$"))
-mcmc_plot(brm_pvis_00, pars = c("^r_ktype:tongue.*perc.flower.cover]$"))
-mcmc_plot(brm_pvis_00, pars = c("^r_ktype:bb.sp.*Intercept]$"))
-mcmc_plot(brm_pvis_00, pars = c("^r_ktype:bb.sp.*perc.flower.cover]$"))
-mcmc_plot(brm_pvis_00, pars = c("^r_ktype:site.*Intercept]$"))
-mcmc_plot(brm_pvis_00, pars = c("^r_ktype:site.*perc.flower.cover]$"))
-
-get_variables(brm_pvis_01)
-conditional_effects(brm_pvis_01)
-mcmc_plot(brm_pvis_01, pars = c("^r_ktype*"))
-mcmc_plot(brm_pvis_01, pars = c("^r_ktype:site.*Intercept]$"))
-mcmc_plot(brm_pvis_01, pars = c("^r_ktype:site.*Intercept]$"))
-mcmc_plot(brm_pvis_01, pars = c("^r_ktype:site.*perc.flower.cover]$"))
-
-get_variables(brm_pvis_02)
-conditional_effects(brm_pvis_02)
-mcmc_plot(brm_pvis_02, pars = c("^r_ktype"))
-mcmc_plot(brm_pvis_02, pars = c("^r_ktype:bb.sp.*Intercept]$"))
-mcmc_plot(brm_pvis_02, pars = c("^r_ktype:site.*Intercept]$"))
-mcmc_plot(brm_pvis_02, pars = c("^r_ktype:bb.sp.*perc.flower.cover]$"))
-mcmc_plot(brm_pvis_02, pars = c("^r_ktype:site.*Intercept]$"))
-mcmc_plot(brm_pvis_02, pars = c("^r_ktype:site.*perc.flower.cover]$"))
-```
-
-## 1.5 Final visualization
-### 1.5.1 morphotype:tongue intercept
-```{r}
-brm_pvis_00_plotframe1 <- brm_pvis_00 %>%
-  spread_draws(b_Intercept, `r_ktype:tongue`[condition, term]) %>% # the back ticks are necessary to escape the :
-  rename(r_ktype.tongue = `r_ktype:tongue`) %>% # get rid of the problematic :
-  filter(term == "Intercept") %>% # drop slope terms for now
-  mutate(condition_mean = b_Intercept + r_ktype.tongue) %>% # get conditional mean
-  group_by(condition) %>% # group by tongue:morpotype pair
-  point_interval(condition_mean, .width = c(0.95, 0.66)) %>% # calculate credible intervals
-  mutate(condition_mean_trans = exp(condition_mean), # exponentiating the estimates gets us back to the response scale (negbinomial uses log link function)
-         .lower_trans = exp(.lower),
-         .upper_trans = exp(.upper)) %>% 
-separate(condition, c("ktype", "tongue"), sep = "_", remove = FALSE) %>% # separate condition term into ktype and tongue
-  mutate(ktype = case_when( # rename
-    ktype == "1" ~ "Disc (1)",
-    ktype == "2" ~ "Funnel (2)",
-    ktype == "3" ~ "Bell (3)",
-    ktype == "4" ~ "Stalk-disc (4)",
-    ktype == "5" ~ "Lip (5)",
-    ktype == "6" ~ "Flag (6)",
-    ktype == "7" ~ "Head (7)")
-  ) %>%
-  mutate(ktype = factor(ktype, levels = c("Disc (1)", "Funnel (2)", # relevel
-                                          "Bell (3)", "Stalk-disc (4)", "Lip (5)", 
-                                          "Flag (6)", "Head (7)"))) %>%
-  mutate(tongue = factor(tongue, levels = c("long", "med", "short")))
-
-ggplot(brm_pvis_00_plotframe1, aes(condition_mean_trans, rev(tongue), 
-                                 xmin = .lower_trans, xmax = .upper_trans,
-                                 color = tongue, fill = tongue)) +
-  geom_pointinterval() +
-  facet_wrap(~ktype, scales = "free") +
-  theme_light() +
-  labs(color = "Tongue length", x = "Visitation probability") +
-  guides(fill = FALSE) +
-  theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank())
-
-ggsave(file = "../output/brm_pvis_00_plot1.pdf", width = 6, height = 5)
-ggsave(file = "../output/brm_pvis_00_plot1.png", width = 6, height = 5)
-```
-
-### 1.5.2 morphotype:bb.sp intercept
-```{r}
-brm_pvis_02_plotframe2 <- brm_pvis_02 %>%
-  spread_draws(b_Intercept, `r_ktype:bb.sp`[condition, term]) %>% # the back ticks are necessary to escape the :
-  rename(r_ktype.bb.sp = `r_ktype:bb.sp`) %>% # get rid of the problematic :
-  filter(term == "Intercept") %>% # drop slope terms for now
-  mutate(condition_mean = b_Intercept + r_ktype.bb.sp) %>% # get conditional mean
-  group_by(condition) %>% # group by bb.sp:morpotype pair
-  point_interval(condition_mean, .width = c(0.95, 0.66)) %>% # calculate credible intervals
-  mutate(condition_mean_trans = inv_logit_scaled(condition_mean), # exponentiating the estimates gets us back to the response scale (negbinomial uses log link function)
-         .lower_trans = inv_logit_scaled(.lower),
-         .upper_trans = inv_logit_scaled(.upper)) %>% 
-separate(condition, c("ktype", "bb.sp"), sep = "_", remove = FALSE) %>% # separate condition term into ktype and bb.sp
-  mutate(ktype = case_when( # rename
-    ktype == "1" ~ "Disc (1)",
-    ktype == "2" ~ "Funnel (2)",
-    ktype == "3" ~ "Bell (3)",
-    ktype == "4" ~ "Stalk-disc (4)",
-    ktype == "5" ~ "Lip (5)",
-    ktype == "6" ~ "Flag (6)",
-    ktype == "7" ~ "Head (7)")
-  ) %>%
-   mutate(ktype = factor(ktype, levels = c("Disc (1)", "Funnel (2)", # relevel
-                                          "Bell (3)", "Stalk-disc (4)", "Lip (5)", 
-                                          "Flag (6)", "Head (7)"))) %>%
-  left_join(bb_traits) %>%
-  mutate(bb.sp = factor(bb.sp, levels = c("telu", "soro", "pyre", "psit", "prat", "lapi", "jone",
-                                           "wurf", "pasc", "muci", "mont", "hypn",
-                                           "mend", "hort", "gers")))
-
-ggplot(brm_pvis_02_plotframe2, aes(condition_mean_trans, bb.sp, 
-                                 xmin = .lower_trans, xmax = .upper_trans,
-                                 color = pbl.w.class2, fill = pbl.w.class2)) +
-  geom_pointinterval() +
-  facet_wrap(~ktype, scales = "free") +
-  theme_light() +
-  labs(color = "Tongue length", x = "Visitation probability", y = "Morphotype") +
-  guides(fill = FALSE)
-
-ggsave(file = "../output/brm_pvis_02_plot1.pdf", width = 8, height = 8)
-ggsave(file = "../output/brm_pvis_02_plot1.png", width = 8, height = 8)
-```
-
-### 1.5.3 morphotype:tongue slope
-```{r eval=FALSE, include=FALSE}
-brm_pvis_00_plotframe3 <- brm_pvis_00 %>%
-  spread_draws(b_perc.flower.cover, `r_ktype:tongue`[condition, term]) %>% # the back ticks are necessary to escape the :
-  rename(r_ktype.tongue = `r_ktype:tongue`) %>%
-  filter(term == "perc.flower.cover") %>%
-  mutate(condition_slope = b_perc.flower.cover + r_ktype.tongue) %>%
-  separate(condition, c("ktype", "tongue"), sep = "_", remove = FALSE) %>%
-  mutate(ktype = case_when(
-    ktype == "1" ~ "Disc (1)",
-    ktype == "2" ~ "Funnel (2)",
-    ktype == "3" ~ "Bell (3)",
-    ktype == "4" ~ "Stalk-disc (4)",
-    ktype == "5" ~ "Lip (5)",
-    ktype == "6" ~ "Flag (6)",
-    ktype == "7" ~ "Head (7)")
-  ) %>%
-  mutate(ktype = factor(ktype,
-                        levels = c("Disc (1)", "Funnel (2)", "Bell (3)", "Stalk-disc (4)",
-                                   "Lip (5)", "Flag (6)", "Head (7)")))
-
-ggplot(brm_pvis_00_plotframe3, aes(condition_slope, tongue, color = tongue)) +
-  stat_pointinterval() +
-  facet_grid(ktype~.) +
-  theme_light() +
-  labs(color = "Tongue length", x = "Slope of logit-likelihood") +
-  guides(fill = FALSE) +
-  theme(axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank())
-
-ggsave(file = "../output/brm_pvis_00_plot3.png", width = 7, height = 5.5)
-```
-
-### 1.5.3 morphotype intercept
-```{r}
-brm_pvis_01_plotframe1 <- brm_pvis_01 %>%
-  spread_draws(b_Intercept, r_ktype[condition, term]) %>% # the back ticks are necessary to escape the :
-  filter(term == "Intercept") %>% # drop slope terms for now
-  mutate(condition_mean = b_Intercept + r_ktype) %>% # get conditional mean
-  group_by(condition) %>% # group by tongue:morpotype pair
-  point_interval(condition_mean, .width = c(0.95, 0.66)) %>% # calculate credible intervals
-  mutate(condition_mean_trans = inv_logit_scaled(condition_mean), # exponentiating the estimates gets us back to the response scale (negbinomial uses log link function)
-         .lower_trans = inv_logit_scaled(.lower),
-         .upper_trans = inv_logit_scaled(.upper)) %>% 
-  mutate(ktype = case_when( # rename
-    condition == "1" ~ "Disc (1)",
-    condition == "2" ~ "Funnel (2)",
-    condition == "3" ~ "Bell (3)",
-    condition == "4" ~ "Stalk-disc (4)",
-    condition == "5" ~ "Lip (5)",
-    condition == "6" ~ "Flag (6)",
-    condition == "7" ~ "Head (7)")
-  ) %>%
-  mutate(ktype = factor(ktype, levels = c("Head (7)", "Flag (6)", # relevel
-                                          "Lip (5)", "Stalk-disc (4)", "Bell (3)", 
-                                          "Funnel (2)", "Disc (1)")))
-
-ggplot(brm_pvis_01_plotframe1, aes(condition_mean_trans, ktype, 
-                                 xmin = .lower_trans, xmax = .upper_trans)) +
-  geom_pointinterval() +
-  theme_light(16) +
-  labs(x = "Visitation probability", y= "Morphotype") +
-  guides(fill = FALSE)
-
-ggsave(file = "../output/brm_pvis_01_plot1.pdf", width = 6, height = 7)
-ggsave(file = "../output/brm_pvis_01_plot1.png", width = 6, height = 7)
-```
-
-### 1.5.3 morphotype slope
-```{r eval=FALSE, include=FALSE}
-brm_pvis_01_plotframe2 <- brm_pvis_01 %>%
-  spread_draws(b_perc.flower.cover, r_ktype[condition, term]) %>% # the back ticks are necessary to escape the :
-  filter(term == "perc.flower.cover") %>%
-  mutate(condition_slope = b_perc.flower.cover + r_ktype) %>%
-  mutate(ktype = case_when(
-    condition == "0" ~ "NA (0)",
-    condition == "1" ~ "Disc (1)",
-    condition == "2" ~ "Funnel (2)",
-    condition == "3" ~ "Bell (3)",
-    condition == "4" ~ "Stalk-disc (4)",
-    condition == "5" ~ "Lip (5)",
-    condition == "6" ~ "Flag (6)",
-    condition == "7" ~ "Head (7)")
-  ) %>%
-  mutate(ktype = factor(ktype,
-                        levels = c("Disc (1)", "Funnel (2)", "Bell (3)", "Stalk-disc (4)",
-                                   "Lip (5)", "Flag (6)", "Head (7)", "NA (0)")))
-
-ggplot(brm_pvis_01_plotframe2, aes(condition_slope, ktype)) +
-  stat_pointinterval() +
-  theme_light() +
-  labs(color = "Tongue length", x = "Slope of visitation probability ~ percent flower cover") +
-  guides(fill = FALSE)
-
-ggsave(file = "../output/brm_pvis_01_plot2.png", width = 7, height = 5.5)
-```
-
-## 1.6 Model 1 hypothesis tests
-```{r}
-get_variables(brm_pvis_01)
-
-hypothesis(brm_pvis_01, "r_ktype[1,Intercept] = r_ktype[4,Intercept]", class = "") 
-hypothesis(brm_pvis_01, "r_ktype[2,Intercept] = r_ktype[4,Intercept]", class = "")
-
-hypothesis(brm_pvis_01, "r_ktype[2,Intercept] = r_ktype[1,Intercept]", class = "")
-hypothesis(brm_pvis_01, "r_ktype[1,Intercept] = r_ktype[3,Intercept]", class = "") 
-hypothesis(brm_pvis_01, "r_ktype[2,Intercept] = r_ktype[3,Intercept]", class = "")
-hypothesis(brm_pvis_01, "r_ktype[4,Intercept] = r_ktype[3,Intercept]", class = "")
-
-hypothesis(brm_pvis_01, "r_ktype[5,Intercept] = r_ktype[6,Intercept]", class = "") 
-hypothesis(brm_pvis_01, "r_ktype[5,Intercept] = r_ktype[7,Intercept]", class = "") 
-```
-
-## 1.7 Model 2 hypothesis tests
-```{r}
-get_variables(brm_pvis_00)
-
-# Disc (group  1)
-hypothesis(brm_pvis_00, "r_ktype:tongue[1_short,Intercept] = r_ktype:tongue[1_med,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[1_short,Intercept] = r_ktype:tongue[1_long,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[1_med,Intercept] = r_ktype:tongue[1_long,Intercept]", class = "")
-
-# Funnel (group  2)
-hypothesis(brm_pvis_00, "r_ktype:tongue[2_short,Intercept] = r_ktype:tongue[2_med,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[2_short,Intercept] = r_ktype:tongue[2_long,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[2_med,Intercept] = r_ktype:tongue[2_long,Intercept]", class = "")
-
-# Bell (group  3)
-hypothesis(brm_pvis_00, "r_ktype:tongue[3_short,Intercept] = r_ktype:tongue[3_med,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[3_short,Intercept] = r_ktype:tongue[3_long,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[3_med,Intercept] = r_ktype:tongue[3_long,Intercept]", class = "")
-
-# Stalk-disc (group  4)
-hypothesis(brm_pvis_00, "r_ktype:tongue[4_short,Intercept] = r_ktype:tongue[4_med,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[4_short,Intercept] = r_ktype:tongue[4_long,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[4_med,Intercept] = r_ktype:tongue[4_long,Intercept]", class = "")
-
-# Lip (group  5)
-hypothesis(brm_pvis_00, "r_ktype:tongue[5_short,Intercept] = r_ktype:tongue[5_med,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[5_short,Intercept] = r_ktype:tongue[5_long,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[5_med,Intercept] = r_ktype:tongue[5_long,Intercept]", class = "")
-
-# Flag (group  6)
-hypothesis(brm_pvis_00, "r_ktype:tongue[6_short,Intercept] = r_ktype:tongue[6_med,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[6_short,Intercept] = r_ktype:tongue[6_long,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[6_med,Intercept] = r_ktype:tongue[6_long,Intercept]", class = "")
-
-# Head (group  7)
-hypothesis(brm_pvis_00, "r_ktype:tongue[7_short,Intercept] = r_ktype:tongue[7_med,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[7_short,Intercept] = r_ktype:tongue[7_long,Intercept]", class = "")
-hypothesis(brm_pvis_00, "r_ktype:tongue[7_med,Intercept] = r_ktype:tongue[7_long,Intercept]", class = "")
+# Visitation data
+visitation <- read_csv("../data/processed_data/visitation.csv")
 ```
@@
   x %>%
-    #group_by(bb.sp, !!criterion, site, date, year) %>% # group by site-year and BB-FL pair
     group_by(bb.sp, !!criterion, site, date) %>% # group by site-year and BB-FL pair
@@
     mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
-    #group_by(bb.sp, site, year, !!criterion) %>%
     group_by(bb.sp, site, !!criterion) %>%
@@
     pivot_wider(names_from = !!criterion, values_from = mean.prop, values_fill = 0) %>% # spread into the kind of array vegan likes
-    #group_by(site, year) %>%
     group_by(site) %>%
@@
            tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
-    #dplyr::select(site, year, tidy_dist) %>% # drop intermediate list-columns
     dplyr::select(site, tidy_dist) %>% # drop intermediate list-columns
@@
     left_join(site_data) %>% # add site data
-    # dplyr::select(site, year, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
     dplyr::select(site, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean, management) %>% # dplyr::select and rename vars
     left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
-    rename(subgenus1 = subgenus, tongue1 = pbl.w.class2) %>%
+    rename(subgenus1 = subgenus, tongue1 = pbl.w) %>%
     left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
-    rename(subgenus2 = subgenus, tongue2 = pbl.w.class2) %>%
+    rename(subgenus2 = subgenus, tongue2 = pbl.w) %>%
     dplyr::select(site, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
        tongue1, tongue2, elev.mean, management) %>%
-    mutate(tongue.diff = factor(if_else(tongue1 == tongue2, "same", "different")),
-           tongue.pair = factor(case_when(
-             tongue1 == "med" & tongue2 == "med" ~ "med:med",
-             tongue1 == "med" & tongue2 == "long" ~ "med:long",
-             tongue1 == "med" & tongue2 == "short" ~ "short:med",
-             tongue1 == "long" & tongue2 == "med" ~ "med:long",
-             tongue1 == "short" & tongue2 == "med" ~ "short:med",
-             tongue1 == "long" & tongue2 == "short" ~ "short:long",
-             tongue1 == "short" & tongue2 == "long" ~ "short:long",
-             tongue1 == "long" & tongue2 == "long" ~ "long:long",
-             tongue1 == "short" & tongue2 == "short" ~ "short:short",
-           )),
+    mutate(tongue.diff = abs(tongue1 - tongue2),
            subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different")) %>%
@@
 
+ggplot(niche_overlap_gen, aes(tongue.diff, niche.overlap)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~site)
+
+ggplot(niche_overlap_fam, aes(tongue.diff, niche.overlap)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~site)
+
+ggplot(niche_overlap_ktype, aes(tongue.diff, niche.overlap)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~site)
+
 # Collate across levels
@@
 
+write_csv(niche_overlap, "../output/niche_overlap.csv")
+
 # Pool across species pairs within level and site to get community mean niche overlap
@@
 
-# Isolate genus-level, then pool within dsp.pair and site
+# Isolate genus-level, then pool within sp.pair and site
 niche_overlap_sppair <-  niche_overlap %>%

### 2021-04-20 code/ms2_ktype_preference_nullmodel_analysis.Rmd
switched to proportional visitation data for null models

@@
 ```{r, warning=FALSE, include=FALSE, echo = FALSE}
-library(bipartite)
 library(econullnetr)
@@
 
-visit_cover <- read_csv("../data/processed_data/visit_cover.csv")
+# For setting up the HPC
+# prop_visits_ktype <- read_csv("./prop_visits_ktype.csv") %>%
+#   mutate(ktype = factor(ktype), # recode variables
+#          tongue = factor(tongue),
+#          visits = as.integer(visits))
+# 
+# prop_visits_ktype_pooled <- read_csv("./prop_visits_ktype_pool.csv") %>%
+#   mutate(ktype = factor(ktype), # recode variables
+#          visits = as.integer(visits))
+# 
+# bb_traits <- read_csv("./bb_traits.csv") %>%
+#   select(bb.sp, pbl.w, tongue = pbl.w.class2)
+# 
+# site_data <- read_csv("./site_data.csv") %>%
+#   dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean, elev.class, elev.class2)) %>%
+#   mutate(site = factor(site),
+#          tree_line = factor(tree_line))
 ```
@@
   unite(id, c(site, date)) %>%
-  select(id, bb.sp, ktype, visits) %>%
-  pivot_wider(values_from = visits, names_from = ktype, values_fill = 0) %>%
+  select(id, bb.sp, ktype, prop.visits) %>%
+  pivot_wider(values_from = prop.visits, names_from = ktype, values_fill = 0) %>%
   select(-c("0", "10", "9"))
@@
                            resources = res[, 2:8],
-                           sims = 5000, # raised from default of 1000 (takes a while to run)
-                           data.type = "counts",
+                           sims = 1000, # raised from default of 100 (takes a while to run)
+                           data.type = "quantities", # since we are using proportional visitation
                            c.samples = con[, 1], # parsed by sample
@@
 
-#saveRDS(null1, file = "../output/nullmodel_ktype.rds")
-null1 <- readRDS("../output/nullmodel_ktype.rds")
+saveRDS(null1, file = "./null1.rds")
+null1 <- readRDS("../output/null1.rds")
 
@@
   group_by(bb.sp) %>%
-  mutate(total.abund = sum(Observed), # total abundance per species
-         preference = Observed - Null, # a working definition of preference
+  mutate(preference = Observed - Null, # a working definition of preference
          preference.upper99 = Observed - Upper.99.CL, 
-         preference.lower99 = Observed - Lower.99.CL,
-
-         # If preference equals the difference between # observed visits and the null model prediction, then dividing preference by the total abundance of each species yields a standardized preference.
-Right?
-         preference.std = preference/total.abund, 
-         Observed.std = Observed/total.abund,
-         Upper.99.CL.std = Upper.99.CL/total.abund,
-         Lower.99.CL.std = Lower.99.CL/total.abund) %>%
+         preference.lower99 = Observed - Lower.99.CL) %>%
   left_join(bb_traits) %>%
@@
 ggsave("../output/null_model_ktypeXtongueclass_preference.png", height = 6, width = 7)
+
+ggplot(null_tab1, aes(pbl.w, SES)) +
+  geom_point(aes(color = Test)) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
+  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
+  facet_wrap(~ktype) +
+  labs(x = "Tongue length (mm)", y = "SES", 
+       color = "Test (p < 0.01)") +
+  theme_light()
+
+ggplot(null_tab1B, aes(pbl.w, preference)) +
+  geom_point(aes(color = Test)) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
+  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
+  facet_wrap(~ktype) +
+  labs(x = "Tongue length (mm)", y = "Standardized preference", 
+       color = "Test (p < 0.01)") +
+  theme_light()
 ```
@@
 con_nest <- prop_visits_ktype %>%
-  select(site, date, bb.sp, ktype, visits) %>%
-  pivot_wider(values_from = visits, names_from = ktype, values_fill = 0) %>%
+  select(site, date, bb.sp, ktype, prop.visits) %>%
+  pivot_wider(values_from = prop.visits, names_from = ktype, values_fill = 0) %>%
   select(-c("0", "10", "9")) %>%
@@
 ## RUN ON HPC!
-# sitewise_null_models <- conres %>%
-#   
-#   # Generate null models
-#   mutate(null.model = map2(bees, plants, function(x, y) 
-#     generate_null_net(consumers = x[, 2:9],
-#                       resources = y[, 2:8],
-#                       sims = 2000,
-#                       data.type = "counts",
-#                       c.samples = x[, 1],
-#                       r.samples = y[, 1]))) %>%
+null2 <- conres %>%
+
+  # Generate null models
+  mutate(null.model = map2(bees, plants, function(x, y)
+    generate_null_net(consumers = x[, 2:9],
+                      resources = y[, 2:8],
+                      sims = 1000, # raised from default of 100
+                      data.type = "quantities",
+                      c.samples = x[, 1],
+                      r.samples = y[, 1])))
+
+#write_rds(null2, "../output/nullmodel_ktype_sitewise.rds")
+write_rds(null2, "./null2.rds") # hpc
+
 # 
@@
 #   left_join(bb_traits) %>%
+#   filter(total.abund > 0) %>% # for each species, drop sites where the species did not occur
+#   unite(bb.ktype, c(bb.sp, ktype), remove = FALSE)
 # 
-#   # Calculate preference/total abundance per species
-#   group_by(site, bb.sp) %>%
-#   mutate(total.abund = sum(Observed)) %>%
-#   ungroup() %>%
-#   mutate(preference.std = preference/total.abund)
-
-#write_rds(sitewise_null_models, "../output/nullmodel_ktype_sitewise.rds")
-#write_rds(sitewise_null_models, "./nullmodel_ktype_sitewise.rds") # hpc
-sitewise_null_models <- readRDS("../output/nullmodel_ktype_sitewise.rds")
-
-sitewise_null_models <- sitewise_null_models %>%
-  filter(total.abund > 0) %>% # for each species, drop sites where the species did not occur
-  unite(bb.ktype, c(bb.sp, ktype), remove = FALSE) %>%
-  mutate(ktype = factor(ktype),
-         bb.sp = factor(bb.sp),
-         bb.ktype = factor(bb.ktype))
-```
-
-## Diet similarity vs. preference similarity
-```{r}
-prefs2 <- sitewise_null_models %>%
-  select(site, bb.sp, ktype, preference.std) %>%
-  unite(site.bb.sp, c(site, bb.sp)) %>%
-  pivot_wider(names_from = ktype, values_from = preference.std) %>%
-  column_to_rownames("site.bb.sp")
-
-visits.ktype <- null_tab1 %>%
-  select(bb.sp, ktype, Observed.std) %>%
-  pivot_wider(names_from = ktype, values_from = Observed.std) %>%
-  column_to_rownames("bb.sp")
-
-visits_diffmat <- vegdist(visits, method = "horn") %>%
-  broom::tidy() %>%
-  select(bb.sp1 = 1, bb.sp2 = 2, distance) %>%
-  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>%
-  select(bb.sp1, bb.sp2, distance, pbl.w1 = pbl.w) %>%
-  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>%
-  select(bb.sp1, bb.sp2, distance, pbl.w1, pbl.w2 = pbl.w) %>%
-  mutate(tongue.diff = abs(pbl.w1 - pbl.w2),
-         diet.similarity = 1 - distance) %>%
-  select(-distance)
-
-prefs_diffmat <- vegdist(prefs, method = "euclidean") %>%
-  broom::tidy() %>%
-  select(bb.sp1 = 1, bb.sp2 = 2, distance) %>%
-  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>%
-  select(bb.sp1, bb.sp2, distance, pbl.w1 = pbl.w) %>%
-  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>%
-  select(bb.sp1, bb.sp2, distance, pbl.w1, pbl.w2 = pbl.w) %>%
-  mutate(tongue.diff = abs(pbl.w1 - pbl.w2),
-         preference.similarity = 1 - distance) %>%
-  select(-distance)
-
-visits_prefs <- full_join(visits_diffmat, prefs_diffmat)
-
+# 
+# 
+# sitewise_null_models <- readRDS("../output/nullmodel_ktype_sitewise.rds") %>%
+#   mutate(ktype = factor(ktype),
+#          bb.sp = factor(bb.sp),
+#          bb.ktype = factor(bb.ktype))
 ```
@@
 ```{r}
+ggplot(sitewise_null_models, aes(pbl.w, preference.std)) +
+  geom_point(aes(color = Test)) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
+  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
+  facet_wrap(~ktype) +
+  labs(x = "Tongue length (mm)", y = "Standardized preference", 
+       color = "Test (p < 0.01)") +
+  theme_light()
+
+ggplot(sitewise_null_models, aes(pbl.w, SES)) +
+  geom_point(aes(color = Test)) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
+  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
+  facet_wrap(~ktype) +
+  labs(x = "Tongue length (mm)", y = "Standardized preference", 
+       color = "Test (p < 0.01)") +
+  theme_light()
+
 # ktype preference x bb.sp ~ elevation
@@
 
-## Overall preference of each ktype ~ elevation
+# Diet similarity vs. preference similarity
 ```{r}
+diet_similarity <- read_csv("../output/niche_overlap.csv") %>%
+  select(site, bb.sp1, bb.sp2, diet.similarity = niche.overlap, level) 
+# %>%
+#   pivot_wider(values_from = niche.overlap, names_from = level) %>%
+#   rename(genus.overlap = genus, family.overlap = family, morphotype.overlap = morphotype)
 
+pref_proc <- function(x) {
+  x %>%
+    pivot_wider(names_from = ktype, values_from = preference.std) %>%
+    column_to_rownames("bb.sp") %>%
+    vegdist(method = "euclidean") %>%
+    broom::tidy() %>%
+    select(bb.sp1 = 1, bb.sp2 = 2, distance)
+}
+
+preference_similarity_sitewise <- sitewise_null_models %>%
+  select(site, bb.sp, ktype, preference.std) %>%
+  group_by(site) %>%
+  nest() %>%
+  mutate(pdif = map(data, pref_proc)) %>%
+  unnest(pdif) %>%
+  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>%
+  select(bb.sp1, bb.sp2, distance, pbl.w1 = pbl.w) %>%
+  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>%
+  select(bb.sp1, bb.sp2, distance, pbl.w1, pbl.w2 = pbl.w) %>%
+  mutate(tongue.diff = abs(pbl.w1 - pbl.w2)) %>%
+  ungroup() %>%
+  mutate(distance.std = distance/max(distance),
+         pref.similarity = 1-distance.std)
+
+diet_preference <- full_join(preference_similarity_sitewise, diet_similarity) %>%
+  filter(!is.na(pref.similarity)) # the data set that was used to generate the niche overlap data includes slightly more data than the one used to generate the preference data due to a filtering step during the alignment of visitation data to floral.
+Only 10 rows are missing, so the difference is trivial.
+We can just drop the resulting NAs.
+
+ggplot(diet_preference, aes(pref.similarity, diet.similarity, color = level)) +
+  geom_point(alpha = 0.25) +
+  geom_smooth(method = "lm") +
+  facet_wrap(~site)
+
+ggplot(diet_preference, aes(tongue.diff, pref.similarity)) +
+  geom_point(alpha = 0.25) +
+  geom_smooth(method = "lm") +
+  facet_wrap(~site)
+
+ggplot(diet_preference, aes(tongue.diff, diet.similarity, color = level)) +
+  geom_point(alpha = 0.25) +
+  geom_smooth(method = "lm") +
+  facet_wrap(~site)
 ```
-
-# Null model 3: bb.sp x plant genus, fully pooled
-```{r}
-# The "consumers" table input to generate_null_net()
-con3 <- visit_cover %>%
-  group_by(site, date, bb.sp, plant.genus) %>%
-  summarize(visits = sum(visits)) %>%
-  unite(id, c(site, date)) %>%
-  select(id, bb.sp, plant.genus, visits) %>%
-  pivot_wider(values_from = visits, names_from = plant.genus, values_fill = 0)
-
-# The "resources" table input to generate_null_net()
-res3 <- visit_cover %>%
-  group_by(site, date, bb.sp, plant.genus) %>%
-  summarize(flower.cover = sum(flower.cover)) %>%
-  group_by(site, date) %>%
-  mutate(perc.flower.cover = flower.cover/sum(flower.cover)) %>%
-  unite(id, c(site, date)) %>%
-  select(id, plant.genus, perc.flower.cover) %>%
-  distinct() %>%
-  pivot_wider(values_from = perc.flower.cover, names_from = plant.genus, values_fill = 0)
-
-# Generate null model
-null3 <- generate_null_net(consumers = con3[, 2:201],
-                           resources = res3[, 2:200],
-                           sims = 5000, # raised from default of 1000 (takes a while to run)
-                           data.type = "counts",
-                           c.samples = con3[, 1], # parsed by sample
-                           r.samples = res3[, 1])
-
-saveRDS(null3, file = "./nullmodel_plant.genus.rds") # hpc
-null3 <- readRDS("../output/nullmodel_ktype.rds")
-
-# Tabulate output and generate tests
-null_tab1 <- test_interactions(null1, signif.level = 0.99) %>% # make it a bit more conservative to account for multiple comparisons
-  as_tibble() %>%
-  select(bb.sp = Consumer, ktype = Resource, everything()) %>%
-  group_by(bb.sp) %>%
-  mutate(total.abund = sum(Observed), # total abundance per species
-         preference = Observed - Null, # a working definition of preference
-         preference.upper99 = Observed - Upper.99.CL, 
-         preference.lower99 = Observed - Lower.99.CL,
-
-         # If preference equals the difference between # observed visits and the null model prediction, then dividing preference by the total abundance of each species yields a standardized preference.
-Right?
-         preference.std = preference/total.abund, 
-         Observed.std = Observed/total.abund,
-         Upper.99.CL.std = Upper.99.CL/total.abund,
-         Lower.99.CL.std = Lower.99.CL/total.abund) %>%
-  left_join(bb_traits) %>%
-  ungroup() %>%
-  mutate(bb.sp = fct_reorder(factor(bb.sp), pbl.w))
-```
-
-# Null model 4: bb.sp x plant genus, fully pooled
-```{r}
-# The "consumers" table input to generate_null_net()
-con4 <- visit_cover %>%
-  group_by(site, date, bb.sp, plant.family) %>%
-  summarize(visits = sum(visits)) %>%
-  unite(id, c(site, date)) %>%
-  select(id, bb.sp, plant.family, visits) %>%
-  pivot_wider(values_from = visits, names_from = plant.family, values_fill = 0)
-
-# The "resources" table input to generate_null_net()
-res4 <- visit_cover %>%
-  group_by(site, date, bb.sp, plant.family) %>%
-  summarize(flower.cover = sum(flower.cover)) %>%
-  group_by(site, date) %>%
-  mutate(perc.flower.cover = flower.cover/sum(flower.cover)) %>%
-  unite(id, c(site, date)) %>%
-  select(id, plant.family, perc.flower.cover) %>%
-  distinct() %>%
-  pivot_wider(values_from = perc.flower.cover, names_from = plant.family, values_fill = 0)
-
-# Generate null model
-null4 <- generate_null_net(consumers = con4[, 2:56],
-                           resources = res4[, 2:55],
-                           sims = 5000, # raised from default of 1000 (takes a while to run)
-                           data.type = "counts",
-                           c.samples = con4[, 1], # parsed by sample
-                           r.samples = res4[, 1])
-
-saveRDS(null4, file = "./nullmodel_plant.family.rds") # hpc
-null3 <- readRDS("../output/nullmodel_ktype.rds")
-
-# Tabulate output and generate tests
-null_tab1 <- test_interactions(null1, signif.level = 0.99) %>% # make it a bit more conservative to account for multiple comparisons
-  as_tibble() %>%
-  select(bb.sp = Consumer, ktype = Resource, everything()) %>%
-  group_by(bb.sp) %>%
-  mutate(total.abund = sum(Observed), # total abundance per species
-         preference = Observed - Null, # a working definition of preference
-         preference.upper99 = Observed - Upper.99.CL, 
-         preference.lower99 = Observed - Lower.99.CL,
-
-         # If preference equals the difference between # observed visits and the null model prediction, then dividing preference by the total abundance of each species yields a standardized preference.
-Right?
-         preference.std = preference/total.abund, 
-         Observed.std = Observed/total.abund,
-         Upper.99.CL.std = Upper.99.CL/total.abund,
-         Lower.99.CL.std = Lower.99.CL/total.abund) %>%
-  left_join(bb_traits) %>%
-  ungroup() %>%
-  mutate(bb.sp = fct_reorder(factor(bb.sp), pbl.w))
-```

### 2021-04-20 code/ms2_niche_overlap.Rmd
switched to proportional visitation data for null models

@@
 library(lubridate)
-
-rstan::rstan_options(auto_write = TRUE)
-options(mc.cores = parallel::detectCores())
 ```
@@
 visitation <- read_csv("../data/processed_data/visitation.csv")
+visit_cover <- read_csv("../data/processed_data/visitation.csv")
 ```
@@
 ```{r}
-# This one averages visitation and then calculates niche overlaps
+# This one averages visitation and then calculates niche overlaps using 1 - Horn-Morisita distance
 niche_overlapper <- function(x, criterion) {
@@
 niche_overlap_gen <- niche_overlapper(niche_data, plant.genus) %>%
-  mutate(level = rep("genus", n()),
-         niche.overlap.genus = niche.overlap)
+  mutate(level = rep("genus", n()))
 
 niche_overlap_fam <- niche_overlapper(niche_data, plant.family) %>%
-  mutate(level = rep("family", n()),
-         niche.overlap.family = niche.overlap)
+  mutate(level = rep("family", n()))
 
 niche_overlap_ktype <- niche_overlapper(niche_data, k.type.ss) %>%
-  mutate(level = rep("morphotype", n()),
-         niche.overlap.ktype = niche.overlap)
+  mutate(level = rep("morphotype", n()))
 
@@
                            niche_overlap_fam, 
-                           niche_overlap_ktype,) %>%
+                           niche_overlap_ktype) %>%
   mutate(level = factor(level, levels = c("family", "morphotype", "genus")),

### 2021-04-21 code/ms2_ktype_preference_nullmodel_analysis.Rmd
cleaned up null model script

--- 2021-04-14 code/ms2_ktype_preference_nullmodel_analysis.Rmd
@@
 library(lubridate)
+library(broom)
+library(vegan)
+library(ggvegan)
+library(ggrepel)
 ```
@@
          tree_line = factor(tree_line))
+
+# For setting up the HPC
+# prop_visits_ktype <- read_csv("./prop_visits_ktype.csv") %>%
+#   mutate(ktype = factor(ktype), # recode variables
+#          tongue = factor(tongue),
+#          visits = as.integer(visits))
+# 
+# prop_visits_ktype_pooled <- read_csv("./prop_visits_ktype_pool.csv") %>%
+#   mutate(ktype = factor(ktype), # recode variables
+#          visits = as.integer(visits))
+# 
+# bb_traits <- read_csv("./bb_traits.csv") %>%
+#   select(bb.sp, pbl.w, tongue = pbl.w.class2)
+# 
+# site_data <- read_csv("./site_data.csv") %>%
+#   dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean, elev.class, elev.class2)) %>%
+#   mutate(site = factor(site),
+#          tree_line = factor(tree_line))
 ```
 
-# Null model by species
+# Get number of presences by date and site-date for each species
+
+In null model 1, the calculation of total Observed vs.
+Null-predicted visitation involves the summing of observed and null-predicted visitation across the sampling unit of site-date.
+This causes the absolute values of total preference (totalObserved - totalNull) to vary with the number of site-dates that were summed, which inflates the absolute value of preference in abundant species relative to rare species.
+By dividing each species' preferences by the nuber of site-dates in which it was observed, we can standardize preference as per-site-date totalObserved - totalNull, which gets all species on a comparable scale and enables comparisons of preference across ktype or in response to tongue length to be meaningful.
+
+Similarly, in null model 2, the calculation of preference involves summing across the sampling unit of date within site, and we cna standardize preference by dividing by the number of dates each species was observed at each site.
+```{r}
+# Site-date presences for null model 1
+site_date_presences <- prop_visits_ktype %>%
+  unite(sitedate, site, date) %>%
+  group_by(bb.sp) %>%
+  summarize(presences = length(unique(sitedate)))
+
+# Date presences for null model 2
+date_presences <- prop_visits_ktype %>%
+  group_by(site, bb.sp) %>%
+  summarize(presences = length(unique(date))) 
+```
+
+# 1. Null model 1: bb.sp x ktype, fully pooled
 ```{r}
@@
 
-#saveRDS(null1, file = "../output/nullmodel_ktype.rds")
-null1 <- readRDS("../output/nullmodel_ktype.rds")
+#saveRDS(null1, file = "./null1.rds")
+null1 <- readRDS("../output/null1.rds")
 
@@
   group_by(bb.sp) %>%
-  mutate(total.abund = sum(Observed), # total abundance per species
-         preference = Observed - Null, # a working definition of preference
+  mutate(preference = Observed - Null, # a working definition of preference
          preference.upper99 = Observed - Upper.99.CL, 
-         preference.lower99 = Observed - Lower.99.CL,
-
-         # If preference equals the difference between # observed visits and the null model prediction, then dividing preference by the total abundance of each species yields a standardized preference.
-Right?
-         preference.std = preference/total.abund, 
-         Observed.std = Observed/total.abund,
-         Upper.99.CL.std = Upper.99.CL/total.abund,
-         Lower.99.CL.std = Lower.99.CL/total.abund) %>%
+         preference.lower99 = Observed - Lower.99.CL) %>%
   left_join(bb_traits) %>%
   ungroup() %>%
+  left_join(site_date_presences) %>%
+  mutate(preference.std = preference/presences, # standardize preference
+         observed.std = Observed/presences, # standardize visitation
+         lower99.std = Lower.99.CL/presences, # standardize confidence intervals
+         upper99.std = Upper.99.CL/presences) %>% 
   mutate(bb.sp = fct_reorder(factor(bb.sp), pbl.w))
 ```
-
-## Visualize network
+## 1.1 Standardized observed vs. null predictions
 ```{r}
-plot_bipartite(null1, signif.level = 0.99)
-```
-
-## Visualize species model
-```{r}
-# Standardized observed vs. null predictions
-## Species-wise
-ggplot(null_tab1, aes(Observed.std, ktype)) +
+# Species-wise
+ggplot(null_tab1, aes(observed.std, ktype)) +
   geom_segment(aes(y = ktype, yend = ktype, 
-                   x = Lower.99.CL.std, xend = Upper.99.CL.std), 
+                   x = lower99.std, xend = upper99.std), 
                color = "gray40", inherit.aes = FALSE) +
@@
 ## Morphotype-wise
-ggplot(null_tab1, aes(Observed.std, bb.sp)) +
+ggplot(null_tab1, aes(observed.std, bb.sp)) +
   geom_segment(aes(y = bb.sp, yend = bb.sp, 
-                   x = Lower.99.CL.std, xend = Upper.99.CL.std), 
+                   x = lower99.std, xend = upper99.std), 
                color = "gray40", inherit.aes = FALSE) +
   geom_point(aes(color = Test)) +
-  facet_wrap(~ktype) +
+  facet_wrap(~ktype, ncol = 4) +
   scale_y_discrete(limits = rev) +
-  scale_color_manual(values = c("gray40", "#00B6EB", "#F8766D")) +
+  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
   labs(x = "Standardized visits", y = "Bumble bee species", color = "Test (p < 0.01)") +
@@
 
-ggsave("../output/null_model_ktypewise_preference.pdf", height = 6, width = 7)
-ggsave("../output/null_model_ktypewise_preference.png", height = 6, width = 7)
+ggsave("../output/null_model_ktypewise_preference.pdf", height = 5, width = 7)
+ggsave("../output/null_model_ktypewise_preference.png", height = 5, width = 7)
+```
 
-# Standardized preference ~ morphotype
+## 1.2 Standardized preference by morphotype
+```{r}
 ggplot(null_tab1, aes(ktype, preference.std)) +
-  #geom_violin(fill = "#F8766D", color = "#F8766D") +
-  geom_violin(fill = "gray70", color = "gray70") +
-  geom_boxplot(fill = NA, color = "black") +
+  #geom_violin(fill = "gray70", color = "white", draw_quantiles = c(0.25, 0.5, 0.75)) +
+  geom_boxplot(fill = NA, color = "black", outlier.shape = 4, size = 0.35) +
+  geom_jitter(aes(color = Test), width = 0.3, alpha = 0.6) +
   theme_light(14) +
+  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
   labs(x = "Floral morphotype", 
@@
 ggsave("../output/null_model_ktype_preference.png", height = 6, width = 7)
+```
 
-# Standardized preference ~ morphotype * tongue class
-ggplot(null_tab1, aes(ktype, preference.std)) +
-  #geom_violin(fill = "#F8766D", color = "#F8766D") +
-  geom_violin(fill = "gray70", color = "gray70") +
-  geom_boxplot(fill = NA, color = "black") +
-  theme_light(14) +
-  labs(x = "Floral morphotype", 
-       y = "Standardized preference") +
-  facet_wrap(~tongue)
-
-ggsave("../output/null_model_ktypeXtongueclass_preference.pdf", height = 6, width = 7)
-ggsave("../output/null_model_ktypeXtongueclass_preference.png", height = 6, width = 7)
-
-# Standardized preference ~ tongue length
+## 1.3 Standardized preference by morphotype ~ tongue length
+### 1.3.1 Preliminary visualization
+```{r}
+# Standardized preference ~ morphotype * tongue length
 ggplot(null_tab1, aes(pbl.w, preference.std)) +
   geom_point(aes(color = Test)) +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5), color = "gray20", size = 0.5) +
-  facet_wrap(~ktype, scales = "free") +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
+  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
+  facet_wrap(~ktype) +
+  labs(x = "Tongue length (mm)", y = "Standardized preference", 
+       color = "Test (p < 0.01)") +
   theme_light()
@@
 
-## Model standardized preference ~ tongue length
+### 1.3.2 Model
 ```{r}
 gam1 <- gam(preference.std ~ 
-              #s(ktype, bs = "re") + 
               ktype + 
@@
 
-## Model standardized preference by morphotype
+## 1.4 Aggregate similarity of preferences and visitation across species
+
+### 1.4.1 Calculate similarities
 ```{r}
-library(brms)
+# Standardized preference matrix
+null1_prefs <- null_tab1 %>%
+  select(bb.sp, ktype, preference.std) %>%
+  pivot_wider(names_from = ktype, values_from = preference.std) %>%
+  column_to_rownames("bb.sp")
 
-x <- seq(-2, 2, by = .1)
-z <- seq(-20, 20, by = .1)
+# Standardized total visitation matrix
+null1_visits <- null_tab1 %>%
+  select(bb.sp, ktype, observed.std) %>%
+  pivot_wider(names_from = ktype, values_from = observed.std) %>%
+  column_to_rownames("bb.sp")
 
-# This seems like a good place to start
-y <- (dnorm(x, mean = 0, sd = 0.5))
-plot(x, y, type = "l")
+# Preference similarity
+null1_prefs_diff <- vegdist(prefs, method = "euclidean") %>% # most metrics don't work when matrix includes negative numbers
+  broom::tidy() %>%
+  select(bb.sp1 = 1, bb.sp2 = 2, distance) %>%
+  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>%
+  select(bb.sp1, bb.sp2, distance, pbl.w1 = pbl.w) %>%
+  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>%
+  select(bb.sp1, bb.sp2, distance, pbl.w1, pbl.w2 = pbl.w) %>%
+  mutate(tongue.diff = abs(pbl.w1 - pbl.w2),
+         preference.similarity = 1 - distance) %>%
+  select(-distance)
 
-ktype_brm <- brm(preference.std ~ ktype,
-                 prior = c(prior(normal(0, 0.5), class = b)),
-                 iter = 2000,
-                 chains = 4,
-                 data = null_tab1)
+# Visitation similarity
+null1_visits_diff <- vegdist(null1_visits, method = "horn") %>%
+  broom::tidy() %>%
+  select(bb.sp1 = 1, bb.sp2 = 2, distance) %>%
+  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>%
+  select(bb.sp1, bb.sp2, distance, pbl.w1 = pbl.w) %>%
+  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>%
+  select(bb.sp1, bb.sp2, distance, pbl.w1, pbl.w2 = pbl.w) %>%
+  mutate(tongue.diff = abs(pbl.w1 - pbl.w2),
+         diet.similarity = 1 - distance) %>%
+  select(-distance)
 
-conditional_effects(ktype_brm)
-
+null1_visits_prefs <- full_join(null1_prefs_diff, null1_visits_diff)
 ```
 
-# Null model by site
+### 1.4.2 How well does preference similarity predict (ktype-level) visitation similarity?
+Two interesting questions:
+1. Doe the relationship deviate from 1:1?
+2.
+How wide is the spread and is the relationship heteroskedastic?
+The wider the spread, the more visitation is affected by relative abundance.
+The tighter the spread, the more visitation is determined by selection bias independent of relative abundance.
+```{r}
+
+ggplot(null1_visits_prefs, aes(preference.similarity, diet.similarity)) +
+  geom_point() +
+  geom_smooth() +
+  geom_abline(slope = 1, color = "black", linetype = "dashed") +
+  theme_light()
+```
+
+### 1.4.3 How well does tongue length difference predict preference similarity?
+```{r}
+ggplot(null1_visits_prefs, aes(tongue.diff, preference.similarity)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 4))
+```
+
+### 1.4.4 Ordinate preferences
+```{r}
+# Call PCA
+pca_pref <- rda(prefs)
+
+# Extract tongue length data
+tongues <- bb_traits %>%
+  semi_join(null_tab1) %>%
+  arrange(bb.sp)
+
+# Ordisurf on tongue length
+ordisurf_pref <- ordisurf(pca_pref, tongues$pbl.w) 
+summary(ordisurf_pref)
+
+# Get variance explained
+pcoa_pref_pc1 <- round(pca_pref$CA$eig[[1]] / sum(pca_pref$CA$eig)*100)
+pcoa_pref_pc2 <- round(pca_pref$CA$eig[[2]] / sum(pca_pref$CA$eig)*100)
+
+# Fortify whole model
+pca_pref.fort <- fortify(pca_pref, axes = 1:2)
+
+# Fortify plants
+pca_pref.fort1 <- fortify(pca_pref, axes = 1:2) %>%
+  filter(Score == "species") %>%
+  rename(ktype = Label)
+
+# Fortify bees
+pca_pref.fort2 <- fortify(pca_pref, axes = 1:2) %>%
+  filter(Score == "sites") %>%
+  rename(bb.sp = Label) %>%
+  left_join(bb_traits)
+
+# Plot ordisurf
+
+## Plotting function based on https://oliviarata.wordpress.com/2014/07/17/ordinations-in-ggplot2-v2-ordisurf/
+surfplot <- function(x, y, min_score = 0) {
+
+  grid <- x$grid # extracts the ordisurf object
+  grid_exp <- expand.grid(x = grid$x, y = grid$y) # get x and ys
+  grid_exp$z <- as.vector(grid$z) # unravel the matrix for the z scores
+
+  contours <- data.frame(na.omit(grid_exp)) %>% # gets rid of the nas
+    as_tibble() %>%
+    rename(PC1 = x, PC2 = y)
+
+  scores <- y %>% 
+    filter((abs(PC1) > min_score | abs(PC2) > min_score) | Score == "BB.sp") %>%
+    mutate(max.score = max(PC1, PC2))
+
+  ggplot(scores, aes(PC1, PC2, pch = Score)) +
+    geom_contour(data = contours, 
+                 aes(PC1, PC2, z = z, color = stat(level)),
+                 size = 2, alpha = 0.35, inherit.aes = FALSE) +
+    #geom_point(size = 2) +
+    geom_text_repel(aes(PC1, PC2, label = Label), key_glyph = "rect") +
+    theme_light() +
+    scale_color_viridis_c(name = "Tongue length") +
+    xlab(paste("PC1 ", "(", pcoa_pref_pc1, "%", ")", sep = "")) +
+    ylab(paste("PC2 ", "(", pcoa_pref_pc2, "%", ")", sep = "")) +
+    coord_fixed() # Gavin Simpson says this is **really** important
+}
+
+## Plot
+surfplot(ordisurf_pref, pca_pref.fort)
+```
+
+# 2. Null model 2: bb.sp x ktype, by site
 ```{r}
@@
 
-# Create a big beautiful table of data and models
-sitewise_null_models <- conres %>%
+# Create a big beautiful table of data and models (run on HPC!)
+null2 <- conres %>%
 
@@
                       resources = y[, 2:8],
-                      sims = 5000,
-                      data.type = "counts",
+                      sims = 1000, # raised from default of 100
+                      data.type = "quantities",
                       c.samples = x[, 1],
-                      r.samples = y[, 1]))) %>%
+                      r.samples = y[, 1])))
+
+# write_rds(null2, "../output/nullmodel_ktype_sitewise.rds")
+# write_rds(null2, "./null2.rds") # hpc
+null2 <- readRDS("../output/null2.rds")
 
 # Tabulate results (with test)
+null_tab2 <- null2 %>%
   mutate(null.model.tab = map(null.model, function(x)
@@
   unnest(null.model.tab) %>%
+  select(site, bb.sp, ktype, Observed, Null, Lower.99.CL, Upper.99.CL, Test, 
+         SES, preference, preference.upper99, preference.lower99) %>%
   left_join(site_data) %>%
   left_join(bb_traits) %>%
-
-  # Calculate preference/total abundance per species
   group_by(site, bb.sp) %>%
-  mutate(total.abund = sum(Observed)) %>%
-  ungroup() %>%
-  mutate(preference.std = preference/total.abund)
-
-write_rds(sitewise_null_models, "../output/nullmodel_ktype_sitewise.rds")
-sitewise_null_models <- readRDS("../output/nullmodel_ktype_sitewise.rds")
+  mutate(total.observed = sum(Observed)) %>%
+  filter(total.observed > 0) %>% # for each species, drop sites where the species did not occur
+  unite(bb.ktype, c(bb.sp, ktype), remove = FALSE) %>%
+  mutate(ktype = factor(ktype),
+         bb.sp = factor(bb.sp),
+         bb.ktype = factor(bb.ktype)) %>%
+  left_join(date_presences) %>%
+  mutate(preference.std = preference/presences, # standardize preference
+         observed.std = Observed/presences, # standardize visitation
+         lower99.std = Lower.99.CL/presences, # standardize confidence intervals
+         upper.99.std = Upper.99.CL/presences) %>% 
+  mutate(bb.sp = factor(bb.sp, levels = c("jone", "pyre", "soro", "psit", # order by ascending tongue length
+                                          "telu", "lapi", "prat", "mont",
+                                          "wurf", "muci", "pasc", "hypn",
+                                          "mend", "hort", "gers")))
 ```
 
-## Visualize site model
+## 2.1 Lability of preference: Do BB species preferences for morphotypes vary with elevation?
+### 2.1.1 Preliminary visualization
 ```{r}
-ggplot(sitewise_null_models, aes(elev.mean, preference, color = ktype)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
-  facet_wrap(~bb.sp, scales = "free")
+ggplot(null_tab2, aes(elev.mean, preference.std)) +
+  geom_point(aes(color = Test)) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), color = "black") +
+  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
+  facet_grid(bb.sp ~ ktype, scales = "free") +
+  theme_light()
 
-ggplot(sitewise_null_models, aes(elev.mean, SES, color = ktype)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
-  facet_wrap(~bb.sp, scales = "free")
+ggplot(null_tab2, aes(elev.mean, preference.std)) +
+  geom_point(aes(color = Test)) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), color = "black") +
+  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
+  facet_grid(ktype ~ bb.sp) +
+  theme_light()
+```
+### 2.1.2 Model
+```{r}
+gam2 <- bam(preference.std ~
+              s(bb.ktype, bs = "re") +
+              s(elev.mean, by = bb.ktype, k = 4),
+            method = "fREML",
+            discrete = TRUE,
+            select = TRUE,
+            data = null_tab2) %>% getViz()
 
-ggplot(sitewise_null_models, aes(elev.mean, preference.std, color = ktype)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
-  facet_wrap(~bb.sp, scales = "free")
+print(plot.gamViz(gam2, allTerms = TRUE), pages = 6)
+check.gamViz(gam2)
+summary(gam2)
+
+# Extract significant smooths
+gam2_sig_smooths <- tidy(gam2) %>%
+  filter(p.value < 0.05 & term != "s(bb.ktype)") %>%
+  mutate(bb.ktype = str_extract(term, ".{6}$")) %>%
+  select(bb.ktype)
+
+# Extract predictions
+gam2_pred <- predict_gam(gam_bb.ktype) %>%
+  mutate(fit.plus.se = fit + (2*se.fit),
+         fit.minus.se = fit - (2*se.fit)) %>%
+  semi_join(gam2_sig_smooths) %>%
+  separate(bb.ktype, c("bb.sp", "ktype"))
+
+# Grab relevant subset of original data
+null_tab2_sub <- null_tab2 %>%
+  semi_join(gam2_sig_smooths)
+
+# Plot
+ggplot(gam2_pred, aes(elev.mean, fit)) +
+  geom_line() + # fit line
+  geom_point(data = null_tab2_sub, # Original data
+             aes(elev.mean, preference.std, color = Test),
+             inherit.aes = FALSE) +
+  geom_ribbon(aes(ymin = fit.minus.se, ymax = fit.plus.se), # Uncertainty
+              fill = "black", alpha = 0.2) +
+  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
+  facet_grid(bb.sp ~ ktype, scales = "free") +
+  labs(x = "Elevation (m.s.l.)", y = "Standardized preference",
+       color = "Test (p < 0.01)") +
+  theme_light()
 ```
+
+## 2.2 Diet similarity vs. preference similarity
+```{r}
+# Bring in niche overlap data generate in a different script
+diet_sim_sitewise <- read_csv("../output/niche_overlap.csv") %>%
+  select(site, bb.sp1, bb.sp2, diet.similarity = niche.overlap, level)
+
+# Function for generating a tidy preference-difference data frame
+pref_proc <- function(x) {
+  x %>%
+    pivot_wider(names_from = ktype, values_from = preference.std) %>%
+    column_to_rownames("bb.sp") %>%
+    vegdist(method = "euclidean") %>%
+    broom::tidy() %>%
+    select(bb.sp1 = 1, bb.sp2 = 2, distance)
+}
+
+# Preference similarity  
+pref_sim_sitewise <- null_tab2 %>%
+  select(site, bb.sp, ktype, preference.std) %>%
+  group_by(site) %>%
+  nest() %>%
+  mutate(pdif = map(data, pref_proc)) %>%
+  unnest(pdif) %>%
+  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>%
+  select(site, bb.sp1, bb.sp2, distance, pbl.w1 = pbl.w) %>%
+  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>%
+  select(site, bb.sp1, bb.sp2, distance, pbl.w1, pbl.w2 = pbl.w) %>%
+  mutate(tongue.diff = abs(pbl.w1 - pbl.w2)) %>%
+  ungroup() %>%
+  mutate(distance.std = distance/max(distance), # standardize relative to most distant pair
+         pref.similarity = 1-distance.std, # make positive (0,1)
+         pref.similarity2 = distance*(-1)) # alternatively
+
+# Join diet and preference similarity
+diet_pref_sitewise <- full_join(pref_sim_sitewise, diet_sim_sitewise) %>%
+  left_join(site_data) %>%
+  unite(sp.pair, c(bb.sp1, bb.sp2), remove = FALSE) %>%
+  mutate(sp.pair = factor(sp.pair)) %>%
+  filter(!is.na(pref.similarity)) # the data set that was used to generate the niche overlap data includes slightly more data than the one used to generate the preference data due to a filtering step during the alignment of visitation data to floral.
+Only 10 rows are missing, so the difference is trivial.
+We can just drop the resulting NAs.
+
+# How well does preference similarity predict diet similarity (and the levels of morphotype, family, and genus)? 
+ggplot(diet_pref_sitewise, aes(pref.similarity, diet.similarity, color = level)) +
+  geom_point(alpha = 0.25) +
+  geom_smooth(method = "glm", 
+              method.args = list(family = "binomial"))
+
+ggplot(diet_pref_sitewise, aes(pref.similarity, diet.similarity, color = level)) +
+  geom_point(alpha = 0.25) +
+  geom_smooth(method = "glm", 
+              method.args = list(family = "binomial")) +
+  facet_wrap(~site)
+
+# How well does tongue length difference predict preference similarity?
+ggplot(diet_preference, aes(tongue.diff, pref.similarity)) +
+  geom_point(alpha = 0.25) +
+  geom_smooth(method = "glm", 
+              method.args = list(family = "binomial"))
+
+ggplot(diet_preference, aes(tongue.diff, pref.similarity)) +
+  geom_point(alpha = 0.25) +
+  geom_smooth(method = "glm", 
+              method.args = list(family = "binomial")) +
+  facet_wrap(~site)
+
+# How well does tongue length difference predict diet similarity?
+ggplot(diet_preference, aes(tongue.diff, diet.similarity, color = level)) +
+  geom_point(alpha = 0.25) +
+  geom_smooth(method = "glm", 
+              method.args = list(family = "binomial"))
+
+ggplot(diet_preference, aes(tongue.diff, diet.similarity, color = level)) +
+  geom_point(alpha = 0.25) +
+  geom_smooth(method = "glm", 
+              method.args = list(family = "binomial")) +
+  facet_wrap(~site)
+
+# Preference similarity by elevation
+ggplot(diet_pref_sitewise, aes(elev.mean, pref.similarity)) +
+  geom_point(alpha = 0.25) +
+  geom_smooth(method = "glm", 
+              method.args = list(family = "binomial"))
+
+ggplot(diet_pref_sitewise, aes(elev.mean, pref.similarity)) +
+  geom_point(alpha = 0.25) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))
+
+# Diet similarity by elevation
+ggplot(diet_pref_sitewise, aes(elev.mean, diet.similarity, color = level)) +
+  geom_point(alpha = 0.25) +
+  geom_smooth(method = "glm", 
+              method.args = list(family = "binomial"))
+
+ggplot(diet_pref_sitewise, aes(elev.mean, diet.similarity, color = level)) +
+  geom_point(alpha = 0.25) +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))
+
+# Model
+gam3 <- bam(pref.similarity ~ 
+              s(sp.pair, bs = "re") +
+              s(elev.mean, k = 8),
+            family = betar(eps = 0.01),
+            method = "fREML",
+            discrete = TRUE,
+            select = TRUE,
+            data = diet_pref_sitewise) %>% getViz()
+
+check.gamViz(gam3)
+summary(gam3)
+plot.gamViz(gam3)
+
+gam4 <- bam(pref.similarity ~ 
+              #s(sp.pair, bs = "re") +
+              s(elev.mean,sp.pair, k = 5, bs = "fs"),
+            family = betar(eps = 0.01),
+            method = "fREML",
+            discrete = TRUE,
+            select = TRUE,
+            data = diet_pref_sitewise) %>% getViz()
+
+check.gamViz(gam4)
+summary(gam4)
+plot.gamViz(gam4)
+```

### 2021-04-28 Dryad_ms1/README.Rmd
switched to proportional visitation data for null models
resurrected lability analysis

@@
   3. `site` = site name
-  4. `snowcover` = whether site was snow-covered
+  4. `trap.time` = time interaction was observed
+  5. `caste` = bumble bee caste (queen, male, worker, NA = parasitic *Psithryus*)
+  6. `pollen` = whether the bee had visible pollen loads
+  7. `bb.sp` = abbreviated bumble bee species name
+  8. `bb.sp.lat` = full bumble bee species name
+  9. `plant.sp.abb` = abbreviated plant species name
   10. `plant.sp` = full plant species name
   11. `plant.genus` = plant genus name
-  12. `flower cover` = flower cover in m^2^ to the nearest 0.01 m^2^
   12. `day` = day of month
@@
   3. `site` = site name
-  4. `trap.time` = time interaction was observed
-  5. `caste` = bumble bee caste (queen, male, worker, NA = parasitic *Psithryus*)
-  6. `pollen` = whether the bee had visible pollen loads
-  7. `bb.sp` = abbreviated bumble bee species name
-  8. `bb.sp.lat` = full bumble bee species name
-  9. `plant.sp.abb` = abbreviated plant species name
+  4. `snowcover` = whether site was snow-covered
   10. `plant.sp` = full plant species name
   11. `plant.genus` = plant genus name
+  12. `flower cover` = flower cover in m^2^ to the nearest 0.01 m^2^
   12. `day` = day of month
@@
 
+## 4. `bb_traits.csv`
+
+### Description: Bumble bee trait data
+
+Clade and subgenus classifications are based on Williams et al (2008).
+
+Tongue length data are from Arbetman et al. (2017), Obeso (1992), and Durieux (2000)
+
+### Columns:
+ 1. `clade` = bumble bee clade: LF = long faced, SH = short faced, M = Mendacibombus
+ 2. `subgenus` = bumble bee subgenus
+ 3. `bb.sp.lat` = full bumble bee species name
+ 4. `bb.sp` = abbreviated bumble bee species name
+ 5. `pbl.w` = worker tongue length based primarily on Arbetman et al. (2017) and supplemented where necessary from Durieux (2000) and Obeso (1992)
+ 6. `pbl.w.class` = discretized worker tongue length
+ 7. `pbl.w.class2` = simplified discretized worker tongue length
+ 8. `pbl.w.ref` = reference from which pbl.w was gathered
+ 9. `Notes` = notes on pbl.w
+ 10. `pbl.w.durieux` = alternative set of tongue length based primarily on Durieux (2000) and supplemented where necessary from Arbetman et al. (2017) and Obeso (1992)
+ 11. `notes_durieux` = notes on pbl.w.durieux
+
+## 5. `fl_traits.csv`
+
+### Description: Kugler's (1970) floral morphotype classification for floral species
+
+For a description of Kugler's morphotype codes, see `kugler_key.tsv`.
+
+Data were accessed in April 2020 via the Bioflor plant trait database (Klotz et al. 2002). Missing data were manually supplemented as descrtibed in `fl_trait_proc.R`.
+
+### Columns:
+ 1. `plant.sp` = plant species
+ 2. `plant.genus` = plant genus
+ 3. `plant.family` = plant family
+ 4. `k.type` = Kugler morphotype
+ 5. `k.type.s` = Kugler morphotype simplified to 1 decimal place
+ 6. `k.type.ss` = Kugler morphotype simplified to 0 decimal places
+
+## 6. `kugler_key.tsv`
+
+### Description: A verbal description of Kugler's (1970) floral morphotype classification
+
+Data were accessed in April 2020 via the Bioflor plant trait database (Klotz et al. 2002).
+
+### Columns:
+ 1. `code` = Morphotype code
+ 2. `description` = Morphotype description
+
 # CODE FILES
@@
 ### Description: R script for running beta-diversity analyses.
+
+## 2. `fl_trait_proc.R`
+
+### Description: R script for processing raw BIOFLOR trait data and adding traits for missing species.
+
+# References
+
+Williams PH, Cameron SA, Hines HM, Cederberg B, Rasmont P. 2018. A simplified subgeneric classification of the bumblebees (genus Bombus). Apidologie, 39, pp. 46-74
+
+Arbetman MP, Gleiser G,Morales CL, Williams P, Aizen MA.
+2017 Global decline of bumblebees is phylogenetically structured and inversely related to species range size and pathogen incidence.
+Proc.
+R.
+Soc.
+B 284: 20170204.
+
+Obeso JR. 1992. Geographic distributions and community structure of Bumblebees in Northern Iberian Peninsula. Oecologia, 89, 244–252.
+
+Durieux E-A. 2000. Etude des choix floraux des bourdons (Hymenoptera, Apidae) de la commune d’eyne (france, Pyrénées-Orientales). PhD thesis, Université de Mons-Hainaut.
+
+Kugler, H. 1970. Blütenökologie. Gustav Fischer, Stuttgart.
+
+Klotz, S, Kühn I, Durka W. 2002. BIOLFLOR - eine datenbank zu biologisch-ökologischen merkmalen der gefäßpflanzen in deutschland. https://www.ufz.de/index.php?en=38567

### 2021-04-28 code/ms2_ktype_preference_nullmodel_analysis.Rmd
resurrected lability analysis

@@
 ```
+
+# 3. Null model 3: bb.sp x ktype, by site-date
+```{r}
+# Nest and join consumer and resource data frames
+
+# For HPC
+library(tidyverse)
+library(econullnetr)
+prop+visits_ktype <- read_csv("./prop_visits_ktype.csv")
+
+con_nest2 <- prop_visits_ktype %>%
+  select(site, date, bb.sp, ktype, prop.visits) %>%
+  pivot_wider(values_from = prop.visits, names_from = ktype, values_fill = 0) %>%
+  select(-c("0", "10", "9")) %>%
+  group_by(site, date) %>%
+  nest() %>%
+  rename(bees = data)
+
+res_nest2 <- prop_visits_ktype %>%
+  select(site, date, ktype, perc.flower.cover) %>%
+  distinct() %>%
+  pivot_wider(values_from = perc.flower.cover, names_from = ktype, values_fill = 0) %>%
+  select(-c("0", "10", "9")) %>%
+  group_by(site, date) %>%
+  nest() %>%
+  rename(plants = data)
+
+conres2 <- full_join(con_nest2, res_nest2)
+
+# Create a big beautiful table of data and models (run on HPC!)
+# This will probably take 8 hours to run
+null3 <- conres2 %>%
+
+  # Generate null models
+  mutate(null.model = map2(bees, plants, function(x, y)
+    generate_null_net(consumers = x[, 1:8],
+                      resources = y[, 1:7],
+                      sims = 1000, # raised from default of 100
+                      data.type = "quantities",
+                      prog.count = TRUE)))
+write_rds(null3, "./null3.rds") # hpc
+null3 <- readRDS("../output/null3.rds")
+
+# Tabulate results (with test)
+null_tab3 <- null3 %>%
+  mutate(null.model.tab = map(null.model, function(x)
+    test_interactions(x, signif.level = 0.99) %>%
+      as_tibble() %>%
+      mutate(preference = Observed - Null,
+             preference.upper99 = Observed - Upper.99.CL,
+             preference.lower99 = Observed - Lower.99.CL) %>%
+      select(bb.sp = Consumer, ktype = Resource, everything()))) %>%
+
+  # Unnest and add site data and bumble bee trait data
+  unnest(null.model.tab) %>%
+  select(site, date, bb.sp, ktype, Observed, Null, Lower.99.CL, Upper.99.CL, Test, 
+         SES, preference, preference.upper99, preference.lower99) %>%
+  left_join(site_data) %>%
+  left_join(bb_traits) %>%
+  group_by(site, date, bb.sp) %>%
+  mutate(total.observed = sum(Observed)) %>%
+  filter(total.observed > 0) %>% # for each species, drop sites where the species did not occur
+  filter(Observed != 0 | Null != 0) %>% # when both equal zero, it's probably because the ktype had zero abundance; in any case, it's meaningless to retain these, and they just cause zero-inflation
+  ungroup() %>%
+  unite(bb.ktype, c(bb.sp, ktype), remove = FALSE) %>%
+  mutate(ktype = factor(ktype),
+         site = factor(site),
+         yday = yday(date),
+         year = factor(year(date)),
+         bb.ktype = factor(bb.ktype),
+         bb.sp = factor(bb.sp, levels = c("jone", "pyre", "soro", "psit", # order by ascending tongue length
+                                          "telu", "lapi", "prat", "mont",
+                                          "wurf", "muci", "pasc", "hypn",
+                                          "mend", "hort", "gers")))
+
+gam3 <- bam(preference ~
+              bb.ktype +
+              s(site, by = bb.ktype, bs = "re") +
+              s(year, by = bb.ktype, bs = "re") +
+              te(yday, elev.mean, by = bb.ktype),
+            family = "gaussian",
+            method = "fREML",
+            discrete = TRUE,
+            data = filter(null_tab3, 
+                          bb.sp %in% c("pasc", "prat", "telu", 
+                                       "wurf", "soro", "hort") &
+                            ktype %in% c("3", "5", "6", "7"))) %>% 
+              getViz()
+
+check.gamViz(gam3)
+gam3_sum <- summary(gam3)
+plot.gamViz(gam3, allTerms = TRUE)
+
+hort_3 <- plot(sm(gam3, 49)) + 
+  labs(y = NULL, x = NULL) + 
+  theme(legend.position = "top") + 
+  theme_void(20)
+
+hort_5 <- plot(sm(gam3, 50)) + labs(y = NULL, x = NULL) + theme_void(20)
+hort_6 <- plot(sm(gam3, 51)) + labs(y = NULL, x = NULL) + theme_void(20)
+hort_7 <- plot(sm(gam3, 52)) + labs(y = NULL, x = NULL) + theme_void(20)
+
+pasc_3 <- plot(sm(gam3, 53)) + labs(y = NULL, x = NULL) + theme_void(20)
+pasc_5 <- plot(sm(gam3, 54)) + labs(y = NULL, x = NULL) + theme_void(20)
+pasc_6 <- plot(sm(gam3, 55)) + labs(y = NULL, x = NULL) + theme_void(20)
+pasc_7 <- plot(sm(gam3, 56)) + labs(y = NULL, x = NULL) + theme_void(20)
+
+prat_3 <- plot(sm(gam3, 57)) + labs(y = NULL, x = NULL) + theme_void(20)
+prat_5 <- plot(sm(gam3, 58)) + labs(y = NULL, x = NULL) + theme_void(20)
+prat_6 <- plot(sm(gam3, 59)) + labs(y = NULL, x = NULL) + theme_void(20)
+prat_7 <- plot(sm(gam3, 60)) + labs(y = NULL, x = NULL) + theme_void(20)
+
+soro_3 <- plot(sm(gam3, 61)) + labs(y = NULL, x = NULL) + theme_void(20)
+soro_5 <- plot(sm(gam3, 62)) + labs(y = NULL, x = NULL) + theme_void(20)
+soro_6 <- plot(sm(gam3, 63)) + labs(y = NULL, x = NULL) + theme_void(20)
+soro_7 <- plot(sm(gam3, 64)) + labs(y = NULL, x = NULL) + theme_void(20)
+
+telu_3 <- plot(sm(gam3, 65)) + labs(y = NULL, x = NULL) + theme_void(20)
+telu_5 <- plot(sm(gam3, 66)) + labs(y = NULL, x = NULL) + theme_void(20)
+telu_6 <- plot(sm(gam3, 67)) + labs(y = NULL, x = NULL) + theme_void(20)
+telu_7 <- plot(sm(gam3, 68)) + labs(y = NULL, x = NULL) + theme_void(20)
+
+wurf_3 <- plot(sm(gam3, 69)) + labs(y = NULL, x = NULL) + theme_void(20)
+wurf_5 <- plot(sm(gam3, 70)) + labs(y = NULL, x = NULL) + theme_void(20)
+wurf_6 <- plot(sm(gam3, 71)) + labs(y = NULL, x = NULL) + theme_void(20)
+wurf_7 <- plot(sm(gam3, 72)) + labs(y = NULL, x = NULL) + theme_void(20)
+
+pdf("../output/gam3_figures/hort_3.pdf") 
+hort_3 
+dev.off()
+
+pdf("../output/gam3_figures/hort_5.pdf") 
+hort_5 
+dev.off()
+
+pdf("../output/gam3_figures/hort_6.pdf") 
+hort_6 
+dev.off()
+
+pdf("../output/gam3_figures/hort_7.pdf") 
+hort_7 
+dev.off()
+
+pdf("../output/gam3_figures/pasc_3.pdf") 
+pasc_3 
+dev.off()
+
+pdf("../output/gam3_figures/pasc_5.pdf") 
+pasc_5 
+dev.off()
+
+pdf("../output/gam3_figures/pasc_6.pdf") 
+pasc_6 
+dev.off()
+
+pdf("../output/gam3_figures/pasc_7.pdf") 
+pasc_7 
+dev.off()
+
+pdf("../output/gam3_figures/prat_3.pdf") 
+prat_3 
+dev.off()
+
+pdf("../output/gam3_figures/prat_5.pdf") 
+prat_5 
+dev.off()
+
+pdf("../output/gam3_figures/prat_6.pdf") 
+prat_6 
+dev.off()
+
+pdf("../output/gam3_figures/prat_7.pdf") 
+prat_7 
+dev.off()
+
+pdf("../output/gam3_figures/soro_3.pdf") 
+soro_3 
+dev.off()
+
+pdf("../output/gam3_figures/soro_5.pdf") 
+soro_5 
+dev.off()
+
+pdf("../output/gam3_figures/soro_6.pdf") 
+soro_6 
+dev.off()
+
+pdf("../output/gam3_figures/soro_7.pdf") 
+soro_7 
+dev.off()
+
+pdf("../output/gam3_figures/telu_3.pdf") 
+telu_3 
+dev.off()
+
+pdf("../output/gam3_figures/telu_5.pdf") 
+telu_5 
+dev.off()
+
+pdf("../output/gam3_figures/telu_6.pdf") 
+telu_6 
+dev.off()
+
+pdf("../output/gam3_figures/telu_7.pdf") 
+telu_7 
+dev.off()
+
+pdf("../output/gam3_figures/wurf_3.pdf") 
+wurf_3 
+dev.off()
+
+pdf("../output/gam3_figures/wurf_5.pdf") 
+wurf_5 
+dev.off()
+
+pdf("../output/gam3_figures/wurf_6.pdf") 
+wurf_6 
+dev.off()
+
+pdf("../output/gam3_figures/wurf_7.pdf") 
+wurf_7 
+dev.off()
+```

### 2021-04-28 Dryad_ms1/fl_trait_proc.R

@@ -1,0 +1,131 @@
+# Packages
+library(tidyverse)
+
+# Floral trait data
+floral_k_type <-  read_delim("./data/raw_data/BioFlor_Kugler_classification.csv", delim = ";") %>%
+  dplyr::select(taxon = 1, k.type = 2) %>% 
+  na.omit() %>%
+  mutate(taxon = str_replace_all(taxon, c(" x " = " x"))) %>%
+  separate(taxon, c("genus", "species"), sep = " ") %>%
+  unite(plant.sp, genus, species, sep = " ", remove = FALSE) %>%
+  select(plant.sp, plant.genus = genus, k.type) %>%
+  mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"), # make simplified (and more simplified) k types
+         k.type.ss = str_extract(k.type, "[-+]?[0-9]*")) %>%
+  
+  # Fill in ktype data for plants present in my samples but missing from the Bioflor database
+  add_row(plant.sp = "Phyteuma orbiculare", plant.genus = "Phyteuma", 
+          k.type = "NA", k.type.s = "7.1", k.type.ss = "7") %>% # Following Neumayer and Paulus (1999)
+  add_row(plant.sp = "Phyteuma spicatum", plant.genus = "Phyteuma",
+          k.type = "NA", k.type.s = "7.1", k.type.ss = "7") %>% # Following Neumayer and Paulus (1999)
+  add_row(plant.sp = "Aconitum vulparia", plant.genus = "Aconitum",
+          k.type = "NA", k.type.s = "5.2", k.type.ss = "5") %>% # All scored Aconitum were classified as 5.2
+  add_row(plant.sp = "Cardamine enneaphyllos", plant.genus = "Cardamine", 
+          k.type = "NA", k.type.s = "1.2", k.type.ss = "1") %>% # All scored Cardamine were classified as 1.2
+  add_row(plant.sp = "Euphrasia picta", plant.genus = "Euphrasia", 
+          k.type = "NA", k.type.s = "5.1", k.type.ss = "5") %>% # All scored Euphrasia were classified as 5.1
+  add_row(plant.sp = "Euphrasia rostkoviana", plant.genus = "Euphrasia", 
+          k.type = "NA", k.type.s = "5.1", k.type.ss = "5") %>% # All scored Euphrasia were classified as 5.1
+  add_row(plant.sp = "Lamium galeobdolon", plant.genus = "Lamium", 
[…]

### 2021-06-09 code/ms2_data.Rmd
updated ms2

--- 2021-04-14 code/ms2_data.Rmd
@@
   filter(!bb.sp %in% c("humi")) %>% # humi too rare
-  group_by(site, date, bb.sp, plant.sp) %>%
-  summarize(visits = n()) %>% # visits per BB-FL pair
-  ungroup() %>%
   left_join(pseudodates) %>% # Now we will replace unmatched dates with their pseudo dates
@@
   group_by(site, date) %>%
-  complete(bb.sp, plant.sp, fill = list(visits = 0)) # generate all possible BB-FL combos for each site date and fill in zeros for visits
+  mutate(visited = rep(1, n())) %>% # binary visitation
+  select(site, date, bb.sp, plant.sp, visited) %>%
+  left_join(fl_traits) %>%
+  select(site, date, bb.sp, plant.sp, visited, plant.genus, plant.family, ktype = k.type.ss) %>%
+  unite(sample.id, site, date) %>%
+  mutate(sample.id = factor(sample.id),
+         bb.sp = factor(bb.sp), 
+         ktype = factor(ktype),
+         visited = as.integer(visited))
 
-# Floral cover data
-flower_cover <- read_csv("../data/processed_data/floral_survey.csv") %>%
+# abundance <- visitation %>%
+#   group_by(site, date, bb.sp) %>%
+#   summarize(abundance = sum(visited))
+
+# Floral cover data by ktype
+flower_cover_ktype <- read_csv("../data/processed_data/floral_survey.csv") %>%
   filter(flower.cover > 0) %>% # ignore plants that were not blooming
-  group_by(site, date, plant.sp) %>%
-  summarize(flower.cover = sum(flower.cover))
+  left_join(fl_traits) %>% # add traits
+  group_by(site, date, k.type.ss) %>% # sum cover by ktype
+  summarize(cover = sum(flower.cover)) %>%
+  complete(site, date, k.type.ss, fill = list(cover = 0)) %>% # fill in zeros for absent ktypes
+  select(site, date, ktype = k.type.ss, cover) %>%
+  unite(sample.id, site, date) %>%
+  mutate(sample.id = factor(sample.id),
+         ktype = factor(ktype)) %>%
+  semi_join(visitation, by = "sample.id")
+# 
+# # Join cover data to visitation data
+# visit_cover_ktype <- visitation %>%
+#   left_join(flower_cover_ktype)  %>%
+#   replace_na(replace = list(cover = 0.01)) %>% # If a ktype was visited but not surveyed, we will assign it the minimum non-zero flower cover value of 0.01.
+#   mutate(individual.id = row_number()) %>%
+#   mutate(ktype = case_when(
+#     ktype == 0 ~ "null",
+#     ktype == 1 ~ "disc",
+#     ktype == 2 ~ "funnel",
+#     ktype == 3 ~ "bell",
+#     ktype == 4 ~ "stalkdisc",
+#     ktype == 5 ~ "lip",
+#     ktype == 6 ~ "flag",
+#     ktype == 7 ~ "head",
+#     ktype == 9 ~ "brush"
+#   )) %>%
+#   unite(sample.id, c(site, date)) %>%
+#   mutate(ktype = factor(ktype, levels = c("null", "disc", "funnel", 
+#                                           "bell", "stalkdisc", "lip", 
+#                                           "flag", "head", "brush")),
+#          sample.id = factor(sample.id),
+#          bb.sp = factor(bb.sp),
+#          visited = as.integer(visited))
 
-# Get shared dates after adding pseudodates. This will let us filter out survey dates that don't
-# have corresponding visitation sampling dates
-shared_dates_with_pseudo <- inner_join(visitation, flower_cover, by = c("site", "date")) %>%
-  select(site, date) %>%
-  distinct()
-
-# Join visitation and flower cover data
-
-## First we need a list of all BB spp. by site and date
-sp_by_site <- visitation %>% 
-  ungroup() %>%
-  select(site, date, bb.sp) %>%
-  distinct()
-
-## Then we need a list of unvisited plant.sp by site and date
-unvisited <- flower_cover %>% 
-  semi_join(shared_dates_with_pseudo) %>% # filter flower cover data to include only site-dates with visitation complement (real or pseudodates)
-  full_join(visitation) %>% # join to visitation data
-  filter(is.na(bb.sp)) %>% # select only the site-date-plants that were totally unvisited (and therefore have no BB spp to join to)
-  select(-bb.sp) %>% # drop bb.sp variable because I just want the plants
-  left_join(sp_by_site) %>% # join in the species list by site and date, thus populating the unvisited plants with all the BB spp recorded for each site-date
-  replace_na(replace = list(visits = 0)) # replace NAs (created in the line above) with zero
-
-## Now we can combine everything into a single data frame
-visit_cover <- flower_cover %>% 
-  semi_join(shared_dates_with_pseudo) %>% # filter flower cover data to include only site-dates with visitation complement (real or pseudodates)
-  full_join(visitation) %>% # join to visitation data
-  replace_na(replace = list(flower.cover = 0.01)) %>% # If a plant was visited but not surveyed, we will assign it the minimum non-zero flower cover value of 0.01.
-  filter(!is.na(bb.sp)) %>% # drop the univisited plants
-  bind_rows(unvisited) %>% # add them back, with BB spp and zeros
-  select(site, date, bb.sp, plant.sp, visits, flower.cover) %>%
-  group_by(site, date, bb.sp, plant.sp, flower.cover) %>% # In a few cases, the use of pseudodates causes two sampling period to converge on the same date, which creates apparent duplicate visitation data.
-We will handle this by averaging the visits.
-  summarize(visits = mean(visits)) %>%
-  left_join(fl_traits) %>% # add floral traits
-  ungroup() %>%
-  mutate(k.type.ss = factor(k.type.ss))
-
-## Get total visits by BB sp and site-date; this will be a measure of BB abundance that we can use after we've calculated d'.
-I could, of course, work this into the biaseval function, but I'll put that off for now.
-bb_abundance <- visit_cover %>%
-  group_by(site, date, bb.sp) %>%
-  summarize(bb.sp.abund = sum(visits)) %>%
-  group_by(site, date) %>%
-  mutate(bb.sp.prop.abund = bb.sp.abund/sum(bb.sp.abund))
-
-fl_abundance <- visit_cover %>%
-  group_by(site, date) %>%
-  summarize(total.flower.cover = sum(flower.cover))
+# abundance2 <- visit_cover_ktype %>%
+#   group_by(site, date, bb.sp) %>%
+#   summarize(abundance = sum(visited))
 ```
 
-# 1. Overall bias
-
-## 1.1 Calculate visitation probability per plant morphotype.
-Instead of keeping the plant species resolved, we will pool visitation by morphotype.
-The problem with not pooling by morphotype is that more species-rich morphotypes will be penalized simply because their visitation is diluted across more species.
 ```{r}
-prop_visits_ktype <- visit_cover %>%
-  group_by(site, date, bb.sp, k.type.ss) %>%
-  summarize(flower.cover = sum(flower.cover), # get raw flower cover and visits by site*date*bb.sp*plant.sp
-            visits = sum(visits)) %>%
-  group_by(site, date) %>%
-  mutate(prop.flower.cover = flower.cover/sum(flower.cover))%>% # binary visits
-  group_by(site, date, bb.sp) %>%
-  mutate(prop.visits = visits/sum(visits),
-         bb.sp.abund = sum(visits)) %>% # abundance of each BB sp. per site-date
-  ungroup() %>%
-  left_join(bb_traits) %>% # add BB traits
-  mutate(ktype = factor(k.type.ss), # recode variables
-         tongue = factor(pbl.w.class2),
-         visits = as.integer(visits)) %>% # so we can run a Poisson model
-  left_join(site_data) %>% # add site data
-  ungroup() %>%
-  mutate(yday = yday(date),
-         elev.mean0 = scale(elev.mean),
-         yday0 = scale(yday),
-         prop.flower.cover0 = scale(prop.flower.cover),
-         perc.flower.cover = prop.flower.cover*100,
-         bb.sp.abund0 = scale(bb.sp.abund)) %>%
-  select(site, elev.mean, date, yday, bb.sp, ktype, visits, 
-         prop.visits, flower.cover, perc.flower.cover, bb.sp.abund, tongue,
-         elev.mean0, yday0, prop.flower.cover0, bb.sp.abund0) %>%
-  mutate(visits.bin = if_else(visits > 0, 1, 0))
-
-write_csv(prop_visits_ktype, "../data/processed_data/prop_visits_ktype.csv")
-
-prop_visits_ktype_pooled <- visit_cover %>%
-  group_by(site, date, k.type.ss) %>%
-  summarize(flower.cover = sum(flower.cover), # get raw flower cover and visits by site*date*bb.sp*plant.sp
-            visits = sum(visits)) %>%
-  group_by(site, date) %>%
-  mutate(prop.flower.cover = flower.cover/sum(flower.cover),
-         prop.visits = visits/sum(visits),
-         bb.abund = sum(visits),
-         visits = as.integer(visits)) %>%
-  mutate(ktype = factor(k.type.ss)) %>%
-  left_join(site_data) %>% # add site data
-  ungroup() %>%
-  mutate(yday = yday(date),
-         elev.mean0 = scale(elev.mean),
-         yday0 = scale(yday),
-         perc.flower.cover = prop.flower.cover*100,
-         prop.flower.cover0 = scale(prop.flower.cover),
-         bb.abund0 = scale(bb.abund)) %>%
-  select(site, elev.mean, date, yday, ktype, visits, 
-         prop.visits, bb.abund, perc.flower.cover,
-         elev.mean0, yday0, prop.flower.cover0, bb.abund0) %>%
-  mutate(visits.bin = if_else(visits > 0, 1, 0),
-         year = factor(year(date)))
-
-write_csv(prop_visits_ktype_pooled, "../data/processed_data/prop_visits_ktype_pool.csv")
+saveRDS(fl_traits, "../data/processed_data/fl_traits.rds")
+saveRDS(visitation, "../data/processed_data/visitation.rds")
+saveRDS(flower_cover_ktype, "../data/processed_data/flower_cover_ktype.rds")
+#saveRDS(visit_cover_ktype, "../data/processed_data/visit_cover_ktype.rds")
 ```

### 2021-06-09 code/ms2_nullmodels.Rmd
updated ms2

---
title: "visitation_bias"
author: "Doug Sponsler"
date: "12/14/2020"
output: html_document
---

# Preparations

## Packages 
```{r, warning=FALSE, include=FALSE, echo = FALSE}
library(econullnetr)
library(patchwork)
library(mgcv)
library(mgcViz)
library(tidymv)
library(tidytext)
library(tidyverse)
library(lubridate)
library(broom)
library(vegan)
library(ggvegan)
library(ggrepel)

Sys.setlocale(locale = "en_US.UTF-8")
```

## Load data
```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
#visit_cover_ktype <- readRDS("../data/processed_data/visit_cover_ktype.rds")

visitation <- readRDS("../data/processed_data/visitation.rds")

survey <- read_csv("../data/processed_data/floral_survey.csv")

flower_cover_ktype <- readRDS("../data/processed_data/flower_cover_ktype.rds")

bb_traits <- read_csv("../data/processed_data/bb_traits.csv") %>%
  select(bb.sp, pbl.w, tongue = pbl.w.class2)

site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean, elev.class, elev.class2)) %>% 
  mutate(site = factor(site),
         tree_line = factor(tree_line))

fl_traits <- readRDS("../data/processed_data/fl_traits.rds")

# For setting up the HPC
# 
# bb_traits <- read_csv("./bb_traits.csv") %>%
#   select(bb.sp, pbl.w, tongue = pbl.w.class2)
# 
# site_data <- read_csv("./site_data.csv") %>%
#   dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean, elev.class, elev.class2)) %>%
#   mutate(site = factor(site),
#          tree_line = factor(tree_line))
```

# 1. Introducing Kugler's morphotypes

## 1.1 Summarize visitation at genus level
```{r}
morphotypes <- visitation %>%
  group_by(plant.genus, ktype) %>%
  mutate(visits = sum(visited)) %>%
  mutate(ktype = case_when(
    ktype == 0 ~ "null (0)",
    ktype == 1 ~ "disc (1)",
    ktype == 2 ~ "funnel (2)",
    ktype == 3 ~ "bell (3)",
    ktype == 4 ~ "stalk-disc (4)",
    ktype == 5 ~ "lip (5)",
    ktype == 6 ~ "flag (6)",
    ktype == 7 ~ "head (7)",
    ktype == 9 ~ "brush (9)"
  )) %>%
  mutate(ktype = factor(ktype, levels = c("null (0)", "disc (1)", "funnel (2)", 
                                          "bell (3)", "stalk-disc (4)", "lip (5)", 
                                          "flag (6)", "head (7)", "brush (9)"))) %>%
  ungroup() %>%
  mutate(plant.genus = reorder_within(plant.genus, -visits, ktype))

ggplot(morphotypes, aes(plant.genus)) +
  geom_bar() +
  facet_wrap(~ktype, scales = "free") +
  labs(y = "Visits", x = "Genus") +
  theme_light(14) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.25, size = 7),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  scale_x_reordered()

ggsave("../ms_2/figures/morphotype_genera.pdf", height = 8, width = 9)
```

## 
1.2 Summarize visitation at family level
```{r}
morphotypes_fam <- visitation %>%
  group_by(plant.family, ktype) %>%
  mutate(visits = sum(visited)) %>%
  mutate(ktype = case_when(
    ktype == 0 ~ "null (0)",
    ktype == 1 ~ "disc (1)",
    ktype == 2 ~ "funnel (2)",
    ktype == 3 ~ "bell (3)",
    ktype == 4 ~ "stalk-disc (4)",
    ktype == 5 ~ "lip (5)",
    ktype == 6 ~ "flag (6)",
    ktype == 7 ~ "head (7)",
    ktype == 9 ~ "brush (9)"
  )) %>%
  mutate(ktype = factor(ktype, levels = c("null (0)", "disc (1)", "funnel (2)", 
                                          "bell (3)", "stalk-disc (4)", "lip (5)", 
                                          "flag (6)", "head (7)", "brush (9)"))) %>%
  ungroup() %>%
  mutate(plant.family = reorder_within(plant.family, -visits, ktype))

ggplot(morphotypes_fam, aes(plant.family)) +
  geom_bar() +
  facet_wrap(~ktype, scales = "free") +
  labs(y = "Visits", x = "Plant family") +
  theme_light(14) +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1, 
                                   vjust = 1, 
                                   size = 10),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  scale_x_reordered()

ggsave("../ms_2/figures/morphotype_families.pdf", height = 9, width = 9)

```

# 2. Null model 1: bb.sp x ktype, fully pooled
```{r}
# The "consumers" table input to generate_null_net()
con <- visitation %>%
  #filter(visited == 1) %>%
  mutate(individual.id = row_number()) %>%
  select(sample.id, individual.id, bb.sp, ktype, visited) %>%
  group_by(sample.id) %>%
  arrange(ktype) %>%
  pivot_wider(values_from = visited, names_from = ktype, values_fill = 0) %>%
  arrange(sample.id) %>%
  ungroup() %>%
  as.data.frame()

# The "resources" table input to generate_null_net()
res <- flower_cover_ktype %>%
  select(sample.id, ktype, cover) %>%
  group_by(sample.id) %>%
  arrange(ktype) %>%
  distinct() %>%
  pivot_wider(values_from = cover, names_from = ktype, values_fill = 0) %>%
  arrange(sample.id) %>%
  ungroup() %>%
  droplevels() %>%
  as.data.frame()

# Generate null model
null1 <- generate_null_net(consumers = con[, 3:12],
                           resources = res[, 2:10],
                           sims = 500, # raised from default of 100 (takes a while to run)
                           data.type = "names",
                           # c.samples = con$sample.id, # parsed by sample
                           # r.samples = res$sample.id,
                           c.samples = con[,1], # parsed by sample
                           r.samples = res[,1],
                           summary.type = "mean") # summarizing with the mean instead of the default sum standardizes across our bumble bee species that vary in total abundance,

saveRDS(null1, file = "../output/null1.rds")
null1 <- readRDS("../output/null1.rds")

# Tabulate output and generate tests
null_tab1 <- test_interactions(null1, signif.level = 0.99) %>% # make it a bit more conservative to account for multiple comparisons
  as_tibble() %>%
  select(bb.sp = Consumer, ktype = Resource, everything()) %>%
  group_by(bb.sp) %>%
  mutate(preference = Observed - Null, # a working definition of preference
         preference.upper99 = Observed - Upper.99.CL, 
         preference.lower99 = Observed - Lower.99.CL) %>%
  left_join(bb_traits) %>%
  ungroup() %>%
  mutate(bb.sp = fct_reorder(factor(bb.sp), pbl.w)) %>%
  mutate(ktype = case_when(
    ktype == 0 ~ "null",
    ktype == 1 ~ "disc",
    ktype == 2 ~ "funnel",
    ktype == 3 ~ "bell",
    ktype == 4 ~ "stalk-disc",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head",
    ktype == 9 ~ "brush"
  )) %>%
  mutate(ktype = factor(ktype, levels = c("null", "disc", "funnel", 
                                          "bell", "stalk-disc", "lip", 
                                          "flag", "head", "brush"))) %>%
  filter(!ktype %in% c("null", "brush")) %>% # type 0 were not reliably quantified (and rarely visited) and type 9 was vanishingly rare
  droplevels()
```

## 2.1 Observed vs. null predictions
```{r}
# Species-wise
ggplot(null_tab1, aes(Observed, ktype)) +
  geom_segment(aes(y = ktype, yend = ktype, 
                   x = Lower.99.CL, xend = Upper.99.CL), 
               color = "gray40", inherit.aes = FALSE) +
  geom_point(aes(color = Test)) +
  facet_wrap(~fct_rev(bb.sp)) +
  scale_y_discrete(limits = rev) +
  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  labs(x = "Mean proportional visitation", y = "Floral morphotype", color = "Test (p < 0.01)") +
  theme_light()

ggsave("../ms_2/figures/null_model_spwise_preference.pdf", height = 6, width = 7)
ggsave("../ms_2/figures/null_model_spwise_preference.png", height = 6, width = 7)

## Morphotype-wise
ggplot(null_tab1, aes(Observed, bb.sp)) +
  geom_segment(aes(y = bb.sp, yend = bb.sp, 
                   x = Lower.99.CL, xend = Upper.99.CL), 
               color = "gray40", inherit.aes = FALSE) +
  geom_point(aes(color = Test)) +
  facet_wrap(~ktype, ncol = 4) +
  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  labs(x = "Mean proportional visitation", y = "Bumble bee species", color = "Test (p < 0.01)") +
  theme_light()

ggsave("../ms_2/figures/null_model_ktypewise_preference.pdf", height = 5, width = 7)
ggsave("../ms_2/figures/null_model_ktypewise_preference.png", height = 5, width = 7)
```

## 2.2 Overall bias by morphotype
```{r}
ggplot(null_tab1, aes(ktype, preference)) +
  #geom_violin(fill = "gray70", color = "white", draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_boxplot(fill = NA, color = "black", outlier.shape = 4, size = 0.35) +
  geom_jitter(aes(color = Test), width = 0.3, alpha = 0.6) +
  theme_light(14) +
  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  labs(x = "Floral morphotype",
       y = "Bias")

ggsave("../ms_2/figures/null_model_ktype_preference.pdf", height = 6, width = 7)
ggsave("../ms_2/figures/null_model_ktype_preference.png", height = 6, width = 7)
```

## 2.3 Bias by morphotype ~ tongue length
### 2.3.1 Preliminary visualization
```{r}
# Standardized preference ~ morphotype * tongue length
ggplot(null_tab1, aes(pbl.w, preference)) +
  geom_point(aes(color = Test)) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  facet_wrap(~ktype) +
  labs(x = "Tongue length (mm)", y = "Preference", 
       color = "Test (p < 0.01)") +
  theme_light()
```

### 2.3.2 Model
```{r}
gam1 <- gam(preference ~ 
              ktype + 
              s(pbl.w, by = ktype, k = 5), 
            data = null_tab1) %>% getViz()

summary(gam1)
check.gamViz(gam1)
#plot.gamViz(gam1, allTerms = TRUE)

# Extract predictions
gam1_pred <- predict_gam(gam1) %>%
  mutate(fit.plus.se = fit + (2*se.fit),
         fit.minus.se = fit - (2*se.fit))

# Plot
ggplot(gam1_pred, aes(pbl.w, fit)) +
  geom_line() + # fit line
  geom_ribbon(aes(ymin = fit.minus.se, ymax = fit.plus.se), # Uncertainty
              fill = "black", alpha = 0.2) +
  geom_point(data = null_tab1, # Original data
             aes(pbl.w, preference, color = bb.sp), 
             inherit.aes = FALSE, alpha = 0.5) +
  #scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  facet_wrap(~ktype) +
  labs(x = "Tongue length (mm)", y = "Bias", 
       color = "Species") +
  theme_light()

ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference.pdf", height = 6, width = 7)
ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference.png", height = 6, width = 7)
```

## 2.4 Ordinate preferences

### 2.4.1 With all species
```{r}
# Bias matrix
null1_prefs <- null_tab1 %>%
  select(bb.sp, ktype, preference) %>%
  pivot_wider(names_from = ktype, values_from = preference) %>%
  column_to_rownames("bb.sp")

# Call PCA
pca_pref <- rda(null1_prefs)

# Extract tongue length data
tongues <- bb_traits %>%
  semi_join(null_tab1) %>%
  arrange(bb.sp)

# Ordisurf on tongue length
ordisurf_pref <- ordisurf(pca_pref, tongues$pbl.w) 
summary(ordisurf_pref)

# Get variance explained
pcoa_pref_pc1 <- round(pca_pref$CA$eig[[1]] / sum(pca_pref$CA$eig)*100)
pcoa_pref_pc2 <- round(pca_pref$CA$eig[[2]] / sum(pca_pref$CA$eig)*100)

# Fortify whole model
pca_pref.fort <- fortify(pca_pref, axes = 1:2) %>%
  # for plotting, I prefer ktype to be numerically coded
  mutate(Label = recode(Label,
                        "disc" = "1",
                        "funnel" = "2",
                        "bell" = "3",
                        "stalk-disc" = "4",
                        "lip" = "5",
                        "flag" = "6",
                        "head" = "7"))

# Fortify plants
pca_pref.fort1 <- fortify(pca_pref, axes = 1:2) %>%
  filter(Score == "species") %>%
  rename(ktype = Label) %>%
  # for plotting, I prefer ktype to be numerically coded
  mutate(ktype = recode(ktype,
                        "disc" = "1",
                        "funnel" = "2",
                        "bell" = "3",
                        "stalk-disc" = "4",
                        "lip" = "5",
                        "flag" = "6",
                        "head" = "7"))

# Fortify bees
pca_pref.fort2 <- fortify(pca_pref, axes = 1:2) %>%
  filter(Score == "sites") %>%
  rename(bb.sp = Label) %>%
  left_join(bb_traits)

# Plot ordisurf

## Plotting function based on https://oliviarata.wordpress.com/2014/07/17/ordinations-in-ggplot2-v2-ordisurf/
surfplot <- function(x, y, min_score = 0, pca1, pca2) {

  grid <- x$grid # extracts the ordisurf object
  grid_exp <- expand.grid(x = grid$x, y = grid$y) # get x and ys
  grid_exp$z <- as.vector(grid$z) # unravel the matrix for the z scores

  contours <- data.frame(na.omit(grid_exp)) %>% # gets rid of the nas
    as_tibble() %>%
    rename(PC1 = x, PC2 = y)

  scores <- y %>% 
    filter((abs(PC1) > min_score | abs(PC2) > min_score) | Score == "BB.sp") %>%
    mutate(max.score = max(PC1, PC2))

  ggplot(scores, aes(PC1, PC2, pch = Score)) +
    geom_contour(data = contours, 
                 aes(PC1, PC2, z = z, color = stat(level)),
                 size = 2, alpha = 0.4, inherit.aes = FALSE) +
    #geom_point(size = 2) +
    geom_text_repel(aes(PC1, PC2, label = Label), key_glyph = "rect") +
    theme_light() +
    scale_color_viridis_c(name = "Tongue length") +
    xlab(paste("PC1 ", "(", pca1, "%", ")", sep = "")) +
    ylab(paste("PC2 ", "(", pca2, "%", ")", sep = "")) +
    coord_fixed() # Gavin Simpson says this is **really** important
}

## Plot
bias_ord <- surfplot(ordisurf_pref, pca_pref.fort, 
         pca1 = pcoa_pref_pc1, pca2 = pcoa_pref_pc2)

ggsave("../ms_2/figures/preference_ordination.png", bias_ord, height = 5.5, width = 8)
ggsave("../ms_2/figures/preference_ordination.pdf", bias_ord, height = 5.5, width = 8)
```

### 2.4.2 Without B. gers, since it's a true specialist
```{r}
# Call PCA
pca_pref2 <- null1_prefs %>%
  slice(-1) %>% # drop B. gers
  rda()

# Extract tongue length data
tongues2 <- bb_traits %>%
  semi_join(null_tab1) %>%
  arrange(bb.sp) %>%
  filter(bb.sp != "gers")

# Ordisurf on tongue length
ordisurf_pref2 <- ordisurf(pca_pref2, tongues2$pbl.w) 
summary(ordisurf_pref2)

# Get variance explained
pcoa_pref2_pc1 <- round(pca_pref2$CA$eig[[1]] / sum(pca_pref2$CA$eig)*100)
pcoa_pref2_pc2 <- round(pca_pref2$CA$eig[[2]] / sum(pca_pref2$CA$eig)*100)

# Fortify whole model
pca_pref.fort2 <- fortify(pca_pref2, axes = 1:2) %>%
  # for plotting, I prefer ktype to be numerically coded
  mutate(Label = recode(Label,
                        "disc" = "1",
                        "funnel" = "2",
                        "bell" = "3",
                        "stalk-disc" = "4",
                        "lip" = "5",
                        "flag" = "6",
                        "head" = "7"))

# Fortify plants
pca_pref2.fort1 <- fortify(pca_pref2, axes = 1:2) %>%
  filter(Score == "species") %>%
  rename(ktype = Label) %>%
  # for plotting, I prefer ktype to be numerically coded
  mutate(ktype = recode(ktype,
                        "disc" = "1",
                        "funnel" = "2",
                        "bell" = "3",
                        "stalk-disc" = "4",
                        "lip" = "5",
                        "flag" = "6",
                        "head" = "7"))

# Fortify bees
pca_pref2.fort2 <- fortify(pca_pref2, axes = 1:2) %>%
  filter(Score == "sites") %>%
  rename(bb.sp = Label) %>%
  left_join(bb_traits)

# Plot ordisurf
surfplot(ordisurf_pref2, pca_pref.fort2, 
         pca1 = pcoa_pref2_pc1, pca2 = pcoa_pref2_pc2)
ggsave("../ms_2/figures/preference_ordination2.png")
ggsave("../ms_2/figures/preference_ordination2.pdf", width = 7, height = 7)
```

### 2.4.5 Cluster preferences
```{r}
clust1 <- null1_prefs %>%
  vegdist(method = "euclidean") %>%
  hclust(method = "complete")

pdf("../ms_2/figures/preference_clust.pdf", width = 5, height = 4)
plot(clust1, main = NULL, xlab = NULL, ylab = NULL)
dev.off()

clust2 <- null1_prefs %>%
  slice(-1) %>% # drop B. gers
  vegdist(method = "euclidean") %>%
  hclust(method = "complete")

pdf("../ms_2/figures/preference_clust2.pdf")
plot(clust2)
dev.off()
```

# 3. Null model 2: bb.sp x ktype, by site
```{r}
# Nest and join consumer and resource data frames
con2 <- visitation %>%
  mutate(individual.id = row_number()) %>%
  separate(sample.id, c("site", "date"), sep = "_") %>%
  select(site, date, individual.id, bb.sp, ktype, visited) %>%
  group_by(site, date) %>%
  arrange(ktype) %>%
  pivot_wider(values_from = visited, names_from = ktype, values_fill = 0) %>%
  group_by(site) %>%
  nest() %>%
  rename(bees = data)

res2 <- flower_cover_ktype %>%
  select(sample.id, ktype, cover) %>%
  separate(sample.id, c("site", "date"), sep = "_") %>%
  group_by(site, date) %>%
  arrange(ktype) %>%
  distinct() %>%
  pivot_wider(values_from = cover, names_from = ktype, values_fill = 0.01) %>%
  group_by(site) %>%
  nest() %>%
  rename(plants = data)

conres2 <- full_join(con2, res2)

# Create a big beautiful table of data and models (run on HPC!)
null2 <- conres2 %>%

  # Generate null models
  mutate(null.model = map2(bees, plants, function(x, y)
    generate_null_net(consumers = x[, 3:12],
                      resources = y[, 2:10],
                      sims = 1000, # raised from default of 100
                      data.type = "names",
                      c.samples = x[, 1],
                      r.samples = y[, 1],
                      summary.type = "mean")))

saveRDS(null2, "../output/null2.rds")
null2 <- readRDS("../output/null2.rds")

# Tabulate results (with test)
null_tab2 <- null2 %>%
  mutate(null.model.tab = map(null.model, function(x)
    test_interactions(x, signif.level = 0.99) %>%
      as_tibble() %>%
      mutate(preference = Observed - Null,
             preference.upper99 = Observed - Upper.99.CL,
             preference.lower99 = Observed - Lower.99.CL) %>%
      select(bb.sp = Consumer, ktype = Resource, everything()))) %>%

  # Unnest and add site data and bumble bee trait data
  unnest(null.model.tab) %>%
  select(site, bb.sp, ktype, Observed, Null, Lower.99.CL, Upper.99.CL, Test, 
         SES, preference, preference.upper99, preference.lower99) %>%
  left_join(site_data) %>%
  left_join(bb_traits) %>%
  group_by(site, bb.sp) %>%
  mutate(total.observed = sum(Observed)) %>%
  filter(total.observed > 0) %>% # for each species, drop sites where the species did not occur
  unite(bb.ktype, c(bb.sp, ktype), remove = FALSE) %>%
  mutate(ktype = factor(ktype),
         bb.sp = factor(bb.sp),
         bb.ktype = factor(bb.ktype)) %>%
  mutate(bb.sp = factor(bb.sp, levels = c("jone", "pyre", "soro", "psit", # order by ascending tongue length
                                          "telu", "lapi", "prat", "mont",
                                          "wurf", "muci", "pasc", "hypn",
                                          "mend", "hort", "gers"))) %>%
  mutate(ktype = case_when(
    ktype == 0 ~ "null",
    ktype == 1 ~ "disc",
    ktype == 2 ~ "funnel",
    ktype == 3 ~ "bell",
    ktype == 4 ~ "stalk-disc",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head",
    ktype == 9 ~ "brush"
  )) %>%
  mutate(ktype = factor(ktype, levels = c("null", "disc", "funnel", 
                                          "bell", "stalk-disc", "lip", 
                                          "flag", "head", "brush"))) %>%
  filter(!ktype %in% c("null", "brush")) %>% # type 0 were not reliably quantified (and rarely visited) and type 9 was vanishingly rare
  droplevels()
```

## 3.1 Lability of preference: Do BB species preferences for morphotypes vary with elevation?
### 3.1.1 Preliminary visualization
```{r}
ggplot(null_tab2, aes(elev.mean, preference)) +
  geom_point(aes(color = Test)) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), color = "black") +
  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  facet_grid(bb.sp ~ ktype, scales = "free") +
  theme_light()
```

### 3.1.2 Model
```{r}
gam2 <- bam(preference ~
              s(bb.ktype, bs = "re") +
              s(elev.mean, by = bb.ktype, k = 4),
            method = "fREML",
            discrete = TRUE,
            select = TRUE,
            data = null_tab2) %>% getViz()

print(plot.gamViz(gam2, allTerms = TRUE), pages = 6)
check.gamViz(gam2)
summary(gam2, re.test = FALSE)

# Extract significant smooths
gam2_sig_smooths <- tidy(gam2) %>%
  filter(p.value < 0.05 & term != "s(bb.ktype)") %>%
  mutate(bb.ktype = str_extract(term, ".{6}$")) %>%
  select(bb.ktype)

# Extract predictions
gam2_pred <- predict_gam(gam2) %>%
  mutate(fit.plus.se = fit + (2*se.fit),
         fit.minus.se = fit - (2*se.fit)) %>%
  semi_join(gam2_sig_smooths) %>%
  separate(bb.ktype, c("bb.sp", "ktype"))

# Grab relevant subset of original data
null_tab2_sub <- null_tab2 %>%
  semi_join(gam2_sig_smooths)

# Plot
ggplot(gam2_pred, aes(elev.mean, fit)) +
  geom_line() + # fit line
  geom_point(data = null_tab2_sub, # Original data
             aes(elev.mean, preference, color = Test),
             inherit.aes = FALSE) +
  geom_ribbon(aes(ymin = fit.minus.se, ymax = fit.plus.se), # Uncertainty
              fill = "black", alpha = 0.2) +
  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  facet_grid(bb.sp ~ ktype, scales = "free") +
  labs(x = "Elevation (m.s.l.)", y = "Standardized preference",
       color = "Test (p < 0.01)") +
  theme_light()
```

# 4. Null model 3: bb.sp x ktype, by site-date
```{r}
# Nest and join consumer and resource data frames

# For HPC
# library(tidyverse)
# library(econullnetr)
# flower_cover_ktype <- readRDS("./flower_cover_ktype.rds")
# visitation <- readRDS("./visitation.rds")

# The "consumers" table input to generate_null_net()
con3 <- visitation %>%
  mutate(individual.id = row_number()) %>%
  select(sample.id, individual.id, bb.sp, ktype, visited) %>%
  group_by(sample.id) %>%
  arrange(ktype) %>%
  pivot_wider(values_from = visited, names_from = ktype, values_fill = 0) %>%
  group_by(sample.id) %>%
  nest() %>%
  rename(bees = data)

# The "resources" table input to generate_null_net()
res3 <- flower_cover_ktype %>%
  select(sample.id, ktype, cover) %>%
  group_by(sample.id) %>%
  arrange(ktype) %>%
  distinct() %>%
  pivot_wider(values_from = cover, names_from = ktype, values_fill = 0) %>%
  group_by(sample.id) %>%
  nest() %>%
  rename(plants = data)

conres3 <- full_join(con3, res3) %>%
  separate(sample.id, c("site", "date"), sep = "_")

# Create a big beautiful table of data and models (run on HPC!)
# This will probably take 8 hours to run
null3 <- conres3 %>%

  # Generate null models
  mutate(null.model = map2(bees, plants, function(x, y)
    generate_null_net(consumers = x[, 2:11],
                      resources = y[, 1:9],
                      sims = 500, # raised from default of 100
                      data.type = "names",
                      prog.count = TRUE,
                      summary.type = "mean")))

#saveRDS(null3, "../output/null3.rds") # hpc

null3 <- readRDS("../output/null3.rds")

# Tabulate results (with test)
null_tab3 <- null3 %>%
  mutate(null.model.tab = map(null.model, function(x)
    test_interactions(x, signif.level = 0.95) %>%
      as_tibble() %>%
      mutate(preference = Observed - Null,
             preference.upper95 = Observed - Upper.95.CL,
             preference.lower95 = Observed - Lower.95.CL) %>%
      select(bb.sp = Consumer, ktype = Resource, everything()))) %>%

  # Unnest and add site data and bumble bee trait data
  unnest(null.model.tab) %>%
  select(site, date, bb.sp, ktype, Observed, Null, Lower.95.CL, Upper.95.CL, Test, 
         SES, preference, preference.upper95, preference.lower95) %>%
  left_join(site_data) %>%
  left_join(bb_traits) %>%
  group_by(site, date, bb.sp) %>%
  mutate(total.observed = sum(Observed)) %>% #There are a few cases in which total.observed does not equal unity. This is because bees occasionally visited plants in ktypes 0, 9, or 10, which we dropped from the data prior to null model calculation but after calculating proportional visitation. This is as it should be.
  ungroup() %>%
  filter(total.observed > 0) %>% # for each species, drop sites where the species did not occur
  filter(Observed != 0 | Null != 0) %>% # when both equal zero, it's probably because the ktype had zero abundance; in any case, it's meaningless to retain these, and they just cause zero-inflation; I checked, and every time Null = 0, Observed = 0 also. 
  unite(bb.ktype, c(bb.sp, ktype), remove = FALSE) %>%
  mutate(ktype = factor(ktype),
         site = factor(site),
         yday = yday(date),
         year = factor(year(date)),
         bb.ktype = factor(bb.ktype),
         bb.sp = factor(bb.sp, levels = c("jone", "pyre", "soro", "psit", # order by ascending tongue length
                                          "telu", "lapi", "prat", "mont",
                                          "wurf", "muci", "pasc", "hypn",
                                          "mend", "hort", "gers"))) %>%
  mutate(ktype = case_when(
    ktype == 0 ~ "null",
    ktype == 1 ~ "disc",
    ktype == 2 ~ "funnel",
    ktype == 3 ~ "bell",
    ktype == 4 ~ "stalk-disc",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head",
    ktype == 9 ~ "brush"
  )) %>%
  mutate(ktype = factor(ktype, levels = c("null", "disc", "funnel", 
                                          "bell", "stalk-disc", "lip", 
                                          "flag", "head", "brush"))) %>%
  filter(!ktype %in% c("null", "brush")) %>% # type 0 were not reliably quantified (and rarely visited) and type 9 was vanishingly rare
  droplevels()

ggplot(null_tab3, aes(ktype, preference)) +
  #geom_violin(fill = "gray70", color = "white", draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_boxplot(fill = NA, color = "black", outlier.shape = 4, size = 0.35) +
  geom_jitter(aes(color = Test), width = 0.3, alpha = 0.6) +
  theme_light(14) +
  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  labs(x = "Floral morphotype",
       y = "Bias")
```

## 4.1 Model preference lability

### 4.1.1 Model 1 (without abundance covariate)
#### 4.1.1.1 Call model
```{r}
gam3.1 <- bam(preference ~
                bb.ktype +
                s(site, by = bb.ktype, bs = "re") +
                s(year, by = bb.ktype, bs = "re") +
                te(yday, elev.mean, by = bb.ktype),
              family = "gaussian",
              method = "fREML",
              discrete = TRUE,
              data = filter(null_tab3, 
                            bb.sp %in% c("hort", "prat", "telu", 
                                         "wurf", "soro", "pasc") &
                              ktype %in% c("bell", "lip", "flag", "head"))) %>% 
                getViz()

#check.gamViz(gam3.1)
#gam3.1_sum <- summary(gam3.1, re.test = FALSE)
#plot.gamViz(gam3.1, allTerms = TRUE)
```

#### 4.1.1.2 Plot model
```{r}
# Tabulate model summary
gam3.1_s_table <- gam3.1_sum$s.table %>%
  as.data.frame() %>%
  rownames_to_column("term") %>%
  filter(str_detect(term, "te\\(")) %>%
  mutate(term = str_extract(term, ".{6}$")) %>%
  separate(term, c("bb.sp", "ktype")) %>%
  mutate(ktype = case_when(
    ktype == 3 ~ "bell",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head"
  )) %>%
  mutate(pval = signif(`p-value`, digits = 3)) %>%
  select(bb.sp, ktype, edf, pval) %>%
  as_tibble() %>%
  mutate(ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")))

# Extract plotting data.
hort_3_gam3.1 <- plot(sm(gam3.1, 49))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("bell", n()))

hort_5_gam3.1 <- plot(sm(gam3.1, 50))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("lip", n()))

hort_6_gam3.1 <- plot(sm(gam3.1, 51))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("flag", n()))

hort_7_gam3.1 <- plot(sm(gam3.1, 52))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("head", n()))

pasc_3_gam3.1 <- plot(sm(gam3.1, 53))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("bell", n()))

pasc_5_gam3.1 <- plot(sm(gam3.1, 54))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("lip", n()))

pasc_6_gam3.1 <- plot(sm(gam3.1, 55))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("flag", n()))

pasc_7_gam3.1 <- plot(sm(gam3.1, 56))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("head", n()))

prat_3_gam3.1 <- plot(sm(gam3.1, 57))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("bell", n()))

prat_5_gam3.1 <- plot(sm(gam3.1, 58))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("lip", n()))

prat_6_gam3.1 <- plot(sm(gam3.1, 59))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("flag", n()))

prat_7_gam3.1 <- plot(sm(gam3.1, 60))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("head", n()))

soro_3_gam3.1 <- plot(sm(gam3.1, 61))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("bell", n()))

soro_5_gam3.1 <- plot(sm(gam3.1, 62))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("lip", n()))

soro_6_gam3.1 <- plot(sm(gam3.1, 63))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("flag", n()))

soro_7_gam3.1 <- plot(sm(gam3.1, 64))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("head", n()))

telu_3_gam3.1 <- plot(sm(gam3.1, 65))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("bell", n()))

telu_5_gam3.1 <- plot(sm(gam3.1, 66))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("lip", n()))

telu_6_gam3.1 <- plot(sm(gam3.1, 67))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("flag", n()))

telu_7_gam3.1 <- plot(sm(gam3.1, 68))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("head", n()))

wurf_3_gam3.1 <- plot(sm(gam3.1, 69))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("bell", n()))

wurf_5_gam3.1 <- plot(sm(gam3.1, 70))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("lip", n()))

wurf_6_gam3.1 <- plot(sm(gam3.1, 71))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("flag", n()))

wurf_7_gam3.1 <- plot(sm(gam3.1, 72))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("head", n()))

plotting_data__gam3.1 <- bind_rows(hort_3_gam3.1, hort_5_gam3.1, hort_6_gam3.1, hort_7_gam3.1,
                           pasc_3_gam3.1, pasc_5_gam3.1, pasc_6_gam3.1, pasc_7_gam3.1,
                           prat_3_gam3.1, prat_5_gam3.1, prat_6_gam3.1, prat_7_gam3.1,
                           soro_3_gam3.1, soro_5_gam3.1, soro_6_gam3.1, soro_7_gam3.1,
                           telu_3_gam3.1, telu_5_gam3.1, telu_6_gam3.1, telu_7_gam3.1,
                           wurf_3_gam3.1, wurf_5_gam3.1, wurf_6_gam3.1, wurf_7_gam3.1) %>%
  group_by(bb.sp, ktype) %>%
  mutate(z.norm = scale(z, center = FALSE),
         ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")),
         date = as_date(x)) %>%
  ungroup() %>%
  left_join(bb_traits) %>%
  left_join(gam3.1_s_table) %>%
  select(z, z.norm, date, x, elevation = y, se, 
         bb.sp, ktype, pbl.w, edf, pval) %>%
  mutate(sig = if_else(pval < 0.05, TRUE, FALSE))

ggplot(plotting_data__gam3.1, aes(date, elevation, z = z.norm, fill = z.norm)) +
  geom_raster() +
  geom_contour(color = "black",
               size = 0.1,
               bins = 40,
               aes(linetype = sig)) + # by fixing bins, we scale contours to each plot while letting the color ramp represent differences in scale across plots; I think this is a good compromise
  facet_grid(reorder(bb.sp, -pbl.w) ~ ktype) +
  scale_linetype_manual(values = c("dotted", "solid"))+
  scale_fill_viridis_c(na.value = "transparent") +
  labs(x = NULL, y = "Elevation (m)", fill = "Normalized\neffect", linetype = "p < 0.05") +
  theme_light(12) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5))

ggsave("../ms_2/figures/preference_lability_gam.pdf", height = 5, width = 6.5)
ggsave("../ms_2/figures/preference_lability_gam.png", height = 5, width = 6.5)
```

### 4.2 Any residue of abundance?
Not just a residue. There is a strict mathematical constraint relating proportional abundance to inferred bias. A rare plant ca be strongly preferred but never strong avoided, while an abundant plant can be strongly avoided but never strongly preferred. Or, put more specifically, as relative abundance approaches zero, minimum bias approachs 0 and maximum bias approach 1. As relative abundance approaches 1, minimum bias approach -1 and maximum bias approaches zero. 
```{r}
ggplot(filter(null_tab3_join,
              bb.sp %in% c("hort", "prat", "telu", 
                           "wurf", "soro", "pasc") &
                ktype %in% c("bell", "lip", "flag", "head")), aes(prop.cover, preference)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm") +
  facet_grid(bb.sp~ktype) +
  ylim(c(-1, 1))

ggplot(filter(null_tab3_join,
              bb.sp %in% c("hort", "prat", "telu", 
                           "wurf", "soro", "pasc") &
                ktype %in% c("bell", "lip", "flag", "head")), aes(cover, preference)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm") +
  facet_grid(bb.sp~ktype, scales = "free") +
  ylim(c(-1, 1))

#########################################################
# Can we wiggle out of this problem via transformation? #
#########################################################
null_tab3.norm <- null_tab3_join %>%
  mutate(preference.norm = case_when(
    preference == 0 ~ 0,
    preference > 0 ~ preference/(1 - prop.cover),
    preference < 0 ~ (preference/(0 - prop.cover)) * -1
  ))

ggplot(filter(null_tab3.norm,
              bb.sp %in% c("hort", "prat", "telu", 
                           "wurf", "soro", "pasc") &
                ktype %in% c("bell", "lip", "flag", "head")), 
       aes(prop.cover, preference.norm)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm") +
  facet_grid(bb.sp~ktype) +
  ylim(c(-1, 1))

```

### 4.3 A binomial model: significant preference 0/1
One way to get around this problem is to model bias as a binary response, either significantly positive or not. 
#### 4.3.1 Call model
```{r}
null_tab3_bin <- null_tab3 %>%
  mutate(preference.bin = as.integer(if_else(Test == "Stronger", 1, 0)))

gam3.2 <- bam(preference.bin ~
                bb.ktype +
                s(site, by = bb.ktype, bs = "re") +
                s(year, by = bb.ktype, bs = "re") +
                te(yday, elev.mean, by = bb.ktype),
              family = "binomial",
              method = "fREML",
              discrete = TRUE,
              data = filter(null_tab3_bin, 
                            bb.sp %in% c("hort", "prat", "telu", 
                                         "wurf", "soro", "pasc") &
                              ktype %in% c("bell", "lip", "flag", "head"))) %>% 
                getViz()

check.gamViz(gam3.2)
gam3.2_sum <- summary(gam3.2, re.test = FALSE)
print(plot.gamViz(gam3.2, allTerms = TRUE), pages = 4)
```

#### 4.3.2 Plot model
```{r}
# Tabulate model summary
gam3.2_s_table <- gam3.2_sum$s.table %>%
  as.data.frame() %>%
  rownames_to_column("term") %>%
  filter(str_detect(term, "te\\(")) %>%
  mutate(term = str_extract(term, ".{6}$")) %>%
  separate(term, c("bb.sp", "ktype")) %>%
  mutate(ktype = case_when(
    ktype == 3 ~ "bell",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head"
  )) %>%
  mutate(pval = signif(`p-value`, digits = 3)) %>%
  select(bb.sp, ktype, edf, pval) %>%
  as_tibble() %>%
  mutate(ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")))

# Extract plotting data.
hort_3 <- plot(sm(gam3.2, 49))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("bell", n()))

hort_5 <- plot(sm(gam3.2, 50))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("lip", n()))

hort_6 <- plot(sm(gam3.2, 51))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("flag", n()))

hort_7 <- plot(sm(gam3.2, 52))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("head", n()))

pasc_3 <- plot(sm(gam3.2, 53))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("bell", n()))

pasc_5 <- plot(sm(gam3.2, 54))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("lip", n()))

pasc_6 <- plot(sm(gam3.2, 55))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("flag", n()))

pasc_7 <- plot(sm(gam3.2, 56))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("head", n()))

prat_3 <- plot(sm(gam3.2, 57))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("bell", n()))

prat_5 <- plot(sm(gam3.2, 58))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("lip", n()))

prat_6 <- plot(sm(gam3.2, 59))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("flag", n()))

prat_7 <- plot(sm(gam3.2, 60))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("head", n()))

soro_3 <- plot(sm(gam3.2, 61))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("bell", n()))

soro_5 <- plot(sm(gam3.2, 62))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("lip", n()))

soro_6 <- plot(sm(gam3.2, 63))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("flag", n()))

soro_7 <- plot(sm(gam3.2, 64))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("head", n()))

telu_3 <- plot(sm(gam3.2, 65))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("bell", n()))

telu_5 <- plot(sm(gam3.2, 66))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("lip", n()))

telu_6 <- plot(sm(gam3.2, 67))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("flag", n()))

telu_7 <- plot(sm(gam3.2, 68))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("head", n()))

wurf_3 <- plot(sm(gam3.2, 69))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("bell", n()))

wurf_5 <- plot(sm(gam3.2, 70))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("lip", n()))

wurf_6 <- plot(sm(gam3.2, 71))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("flag", n()))

wurf_7 <- plot(sm(gam3.2, 72))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("head", n()))

plotting_data <- bind_rows(hort_3, hort_5, hort_6, hort_7,
                           pasc_3, pasc_5, pasc_6, pasc_7,
                           prat_3, prat_5, prat_6, prat_7,
                           soro_3, soro_5, soro_6, soro_7,
                           telu_3, telu_5, telu_6, telu_7,
                           wurf_3, wurf_5, wurf_6, wurf_7) %>%
  group_by(bb.sp, ktype) %>%
  mutate(z.norm = scale(z, center = FALSE),
         ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")),
         date = as_date(x)) %>%
  ungroup() %>%
  left_join(bb_traits) %>%
  left_join(gam3.2_s_table) %>%
  select(z, z.norm, date, x, elevation = y, se, 
         bb.sp, ktype, pbl.w, edf, pval) %>%
  mutate(sig = if_else(pval < 0.05, TRUE, FALSE))

ggplot(plotting_data, aes(date, elevation, z = z.norm, fill = z.norm)) +
  geom_raster() +
  geom_contour(color = "black",
               size = 0.1,
               bins = 40,
               aes(linetype = sig)) + # by fixing bins, we scale contours to each plot while letting the color ramp represent differences in scale across plots; I think this is a good compromise
  facet_grid(reorder(bb.sp, -pbl.w) ~ ktype) +
  scale_linetype_manual(values = c("dotted", "solid"))+
  scale_fill_viridis_c(na.value = "transparent") +
  labs(x = NULL, y = "Elevation (m)", fill = "Normalized\neffect", linetype = "p < 0.05") +
  theme_light(12) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5))
```

### 4.4 A normalized model: proportion of maximum/minimum preference
#### 4.4.1 Call model
```{r}
gam3.3 <- bam(preference.norm ~
                bb.ktype +
                s(site, by = bb.ktype, bs = "re") +
                s(year, by = bb.ktype, bs = "re") +
                te(yday, elev.mean, by = bb.ktype),
              family = "gaussian",
              method = "fREML",
              discrete = TRUE,
              data = filter(null_tab3.norm, 
                            bb.sp %in% c("hort", "prat", "telu", 
                                         "wurf", "soro", "pasc") &
                              ktype %in% c("bell", "lip", "flag", "head"))) %>% 
                getViz()

check.gamViz(gam3.3)
gam3.3_sum <- summary(gam3.3, re.test = FALSE)
print(plot.gamViz(gam3.3, allTerms = TRUE), pages = 4)
```

#### 4.4.2 Plot model
```{r}
# Tabulate model summary
gam3.3_s_table <- gam3.3_sum$s.table %>%
  as.data.frame() %>%
  rownames_to_column("term") %>%
  filter(str_detect(term, "te\\(")) %>%
  mutate(term = str_extract(term, ".{6}$")) %>%
  separate(term, c("bb.sp", "ktype")) %>%
  mutate(ktype = case_when(
    ktype == 3 ~ "bell",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head"
  )) %>%
  mutate(pval = signif(`p-value`, digits = 3)) %>%
  select(bb.sp, ktype, edf, pval) %>%
  as_tibble() %>%
  mutate(ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")))

# Extract plotting data.
hort_3 <- plot(sm(gam3.3, 49))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("bell", n()))

hort_5 <- plot(sm(gam3.3, 50))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("lip", n()))

hort_6 <- plot(sm(gam3.3, 51))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("flag", n()))

hort_7 <- plot(sm(gam3.3, 52))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("head", n()))

pasc_3 <- plot(sm(gam3.3, 53))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("bell", n()))

pasc_5 <- plot(sm(gam3.3, 54))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("lip", n()))

pasc_6 <- plot(sm(gam3.3, 55))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("flag", n()))

pasc_7 <- plot(sm(gam3.3, 56))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("head", n()))

prat_3 <- plot(sm(gam3.3, 57))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("bell", n()))

prat_5 <- plot(sm(gam3.3, 58))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("lip", n()))

prat_6 <- plot(sm(gam3.3, 59))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("flag", n()))

prat_7 <- plot(sm(gam3.3, 60))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("head", n()))

soro_3 <- plot(sm(gam3.3, 61))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("bell", n()))

soro_5 <- plot(sm(gam3.3, 62))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("lip", n()))

soro_6 <- plot(sm(gam3.3, 63))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("flag", n()))

soro_7 <- plot(sm(gam3.3, 64))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("head", n()))

telu_3 <- plot(sm(gam3.3, 65))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("bell", n()))

telu_5 <- plot(sm(gam3.3, 66))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("lip", n()))

telu_6 <- plot(sm(gam3.3, 67))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("flag", n()))

telu_7 <- plot(sm(gam3.3, 68))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("head", n()))

wurf_3 <- plot(sm(gam3.3, 69))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("bell", n()))

wurf_5 <- plot(sm(gam3.3, 70))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("lip", n()))

wurf_6 <- plot(sm(gam3.3, 71))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("flag", n()))

wurf_7 <- plot(sm(gam3.3, 72))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("head", n()))

plotting_data <- bind_rows(hort_3, hort_5, hort_6, hort_7,
                           pasc_3, pasc_5, pasc_6, pasc_7,
                           prat_3, prat_5, prat_6, prat_7,
                           soro_3, soro_5, soro_6, soro_7,
                           telu_3, telu_5, telu_6, telu_7,
                           wurf_3, wurf_5, wurf_6, wurf_7) %>%
  group_by(bb.sp, ktype) %>%
  mutate(z.norm = scale(z, center = FALSE),
         ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")),
         date = as_date(x)) %>%
  ungroup() %>%
  left_join(bb_traits) %>%
  left_join(gam3.3_s_table) %>%
  select(z, z.norm, date, x, elevation = y, se, 
         bb.sp, ktype, pbl.w, edf, pval) %>%
  mutate(sig = if_else(pval < 0.05, TRUE, FALSE))

ggplot(plotting_data, aes(date, elevation, z = z.norm, fill = z.norm)) +
  geom_raster() +
  geom_contour(color = "black",
               size = 0.1,
               bins = 40,
               aes(linetype = sig)) + # by fixing bins, we scale contours to each plot while letting the color ramp represent differences in scale across plots; I think this is a good compromise
  facet_grid(reorder(bb.sp, -pbl.w) ~ ktype) +
  scale_linetype_manual(values = c("dotted", "solid"))+
  scale_fill_viridis_c(na.value = "transparent") +
  labs(x = NULL, y = "Elevation (m)", fill = "Normalized\neffect", linetype = "p < 0.05") +
  theme_light(12) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5))
```

### 4.1.2 Model 2 (with prop_abund as covariate)
```{r}
# With abundance covariate (interacting with bb-ktype pair)

## Add abundance data
abundance <- flower_cover_ktype %>%
  separate(sample.id, c("site", "date"), sep = "_") %>%
  group_by(site, date) %>%
  mutate(prop.cover = cover/sum(cover)) %>%
  mutate(ktype = case_when(
    ktype == 0 ~ "null",
    ktype == 1 ~ "disc",
    ktype == 2 ~ "funnel",
    ktype == 3 ~ "bell",
    ktype == 4 ~ "stalk-disc",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head",
    ktype == 9 ~ "brush"
  )) %>%
  mutate(ktype = factor(ktype))

null_tab3_join <- null_tab3 %>%
  left_join(abundance) %>%
  mutate(site = factor(site),
         year = factor(year(date)))

## Model with proportional cover data
gam3.2 <- bam(preference ~
                bb.ktype +
                #prop.cover*bb.ktype +
                cover:bb.ktype +
                s(site, by = bb.ktype, bs = "re") +
                s(year, by = bb.ktype, bs = "re") +
                te(yday, elev.mean, by = bb.ktype),
              family = "gaussian",
              method = "fREML",
              discrete = TRUE,
              data = filter(null_tab3_join, 
                            bb.sp %in% c("hort", "prat", "telu", 
                                         "wurf", "soro", "pasc") &
                              ktype %in% c("bell", "lip", "flag", "head"))) %>% 
                getViz()

#check.gamViz(gam3.2)
#gam3.2_corr_sum <- summary(gam3.2, re.test = FALSE)
#print(plot.gamViz(gam3.2, allTerms = TRUE), pages = 4)
```

Since we have a markedly negative effect of abundance on preference, due to the fact that an avoided resource will seem more avoided when it is abundant than when it is rare, even if the cognitive evaluation of the resource does not vary, perhaps it would be worth adding abundance as a covariate to the preference lability model. We do this in model 3.2.

### 4.1.3 Year-specific models
```{r}
# Year-specific models
gam3_2010 <- bam(preference ~
              bb.ktype +
              s(site, by = bb.ktype, bs = "re") +
              te(yday, elev.mean, by = bb.ktype),
            family = "gaussian",
            method = "fREML",
            discrete = TRUE,
            data = filter(null_tab3, 
                          year == 2010 &
                            bb.sp %in% c("hort", "prat", "telu", 
                                       "wurf", "soro", "pasc") &
                            ktype %in% c("3", "5", "6", "7"))) %>% 
              getViz()

check.gamViz(gam3_2010)
gam3_2010_sum <- summary(gam3_2010)
plot.gamViz(gam3_2010, allTerms = TRUE)

gam3_2011 <- bam(preference ~
              bb.ktype +
              s(site, by = bb.ktype, bs = "re") +
              te(yday, elev.mean, by = bb.ktype),
            family = "gaussian",
            method = "fREML",
            discrete = TRUE,
            data = filter(null_tab3, 
                          year == 2011 &
                            bb.sp %in% c("hort", "prat", "telu", 
                                       "wurf", "soro", "pasc") &
                            ktype %in% c("3", "5", "6", "7"))) %>% 
              getViz()

check.gamViz(gam3_2011)
gam3_2011_sum <- summary(gam3_2011)
plot.gamViz(gam3_2011, allTerms = TRUE)

gam3_2012 <- bam(preference ~
              bb.ktype +
              s(site, by = bb.ktype, bs = "re") +
              te(yday, elev.mean, by = bb.ktype),
            family = "gaussian",
            method = "fREML",
            discrete = TRUE,
            data = filter(null_tab3, 
                          year == 2012 &
                            bb.sp %in% c("hort", "prat", "telu", 
                                       "wurf", "soro", "pasc") &
                            ktype %in% c("3", "5", "6", "7"))) %>% 
              getViz()

check.gamViz(gam3_2012)
gam3_2012_sum <- summary(gam3_2012)
plot.gamViz(gam3_2012, allTerms = TRUE)
```

```{r}

gam3_corr_s_table <- gam3_corr_sum$s.table %>%
  as.data.frame() %>%
  rownames_to_column("term") %>%
  filter(str_detect(term, "te\\(")) %>%
  mutate(term = str_extract(term, ".{6}$")) %>%
  separate(term, c("bb.sp", "ktype")) %>%
  mutate(ktype = case_when(
    ktype == 3 ~ "bell",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head"
  )) %>%
  mutate(pval = signif(`p-value`, digits = 3)) %>%
  select(bb.sp, ktype, edf, pval) %>%
  as_tibble() %>%
  mutate(ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")))

# Extract plotting data.

hort_3_corr <- plot(sm(gam3_corr, 49))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("bell", n()))

hort_5_corr <- plot(sm(gam3_corr, 50))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("lip", n()))

hort_6_corr <- plot(sm(gam3_corr, 51))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("flag", n()))

hort_7_corr <- plot(sm(gam3_corr, 52))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("head", n()))

pasc_3_corr <- plot(sm(gam3_corr, 53))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("bell", n()))

pasc_5_corr <- plot(sm(gam3_corr, 54))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("lip", n()))

pasc_6_corr <- plot(sm(gam3_corr, 55))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("flag", n()))

pasc_7_corr <- plot(sm(gam3_corr, 56))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("head", n()))

prat_3_corr <- plot(sm(gam3_corr, 57))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("bell", n()))

prat_5_corr <- plot(sm(gam3_corr, 58))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("lip", n()))

prat_6_corr <- plot(sm(gam3_corr, 59))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("flag", n()))

prat_7_corr <- plot(sm(gam3_corr, 60))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("head", n()))

soro_3_corr <- plot(sm(gam3_corr, 61))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("bell", n()))

soro_5_corr <- plot(sm(gam3_corr, 62))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("lip", n()))

soro_6_corr <- plot(sm(gam3_corr, 63))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("flag", n()))

soro_7_corr <- plot(sm(gam3_corr, 64))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("head", n()))

telu_3_corr <- plot(sm(gam3_corr, 65))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("bell", n()))

telu_5_corr <- plot(sm(gam3_corr, 66))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("lip", n()))

telu_6_corr <- plot(sm(gam3_corr, 67))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("flag", n()))

telu_7_corr <- plot(sm(gam3_corr, 68))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("head", n()))

wurf_3_corr <- plot(sm(gam3_corr, 69))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("bell", n()))

wurf_5_corr <- plot(sm(gam3_corr, 70))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("lip", n()))

wurf_6_corr <- plot(sm(gam3_corr, 71))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("flag", n()))

wurf_7_corr <- plot(sm(gam3_corr, 72))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("head", n()))

plotting_data_corr <- bind_rows(hort_3_corr, hort_5_corr, hort_6_corr, hort_7_corr,
                           pasc_3_corr, pasc_5_corr, pasc_6_corr, pasc_7_corr,
                           prat_3_corr, prat_5_corr, prat_6_corr, prat_7_corr,
                           soro_3_corr, soro_5_corr, soro_6_corr, soro_7_corr,
                           telu_3_corr, telu_5_corr, telu_6_corr, telu_7_corr,
                           wurf_3_corr, wurf_5_corr, wurf_6_corr, wurf_7_corr) %>%
  group_by(bb.sp, ktype) %>%
  mutate(z.norm = scale(z, center = FALSE),
         ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")),
         date = as_date(x)) %>%
  ungroup() %>%
  left_join(bb_traits) %>%
  left_join(gam3_corr_s_table) %>%
  select(z, z.norm, date, x, elevation = y, se, 
         bb.sp, ktype, pbl.w, edf, pval) %>%
  filter(pval < 0.05)

ggplot(plotting_data_corr, aes(date, elevation, z = z.norm, fill = z.norm)) +
  geom_raster() +
  geom_contour(color = "black",
               size = 0.1,
               bins = 40) + # by fixing bins, we scale contours to each plot while letting the color ramp represent differences in scale across plots; I think this is a good compromise
  facet_grid(reorder(bb.sp, -pbl.w) ~ ktype) +
  scale_fill_viridis_c(na.value = "transparent") +
  labs(x = NULL, y = "Elevation (m)", fill = "Normalized\neffect") +
  theme_light(12) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5))

ggsave("../output/preference_lability_gam_corr.pdf", height = 5, width = 6.5)
```

# 5 Raw visitation rates
Null model analysis is potentially problematic. It is still wroth performing and presenting, but we should compare it to patterns in raw visitation to understand "preference" more completely. The problem is well-illustrated with an anecdote. Suppose I am shopping for Molly, and I need to find gluten-free flour. In a typical grocery store, gluten-free flour is hard to come by, so my selection of it would appear to be a strong preference --- after all, I passed up so many other options to get one of the few options I most preferred. If I went to a fancy healthfood store, though, gluten-free flours would probably make up a large proportion of available options, and my decision would seem consistent with a null model of random flour selection. In both cases, I was completely committed to choosing a gluten-free flour; I preferred it perfectly. But in one case, it looks like preference, and in the other it doesn't. This same problem applies to the use of null models to infer preference in bumble bees. If, for example, B. mucidus  is committed to foraging on flag flowers in all floristic contexts, the only patterns of preference we will find are artifacts of relative abundance, created precisely by the technique intended to correct for such artifacts. I think the best approach is to present both a null-model based analysis and a raw visitation analysis and try to say something about which is a more appropriate representation of bumble bee floral selection. Are they more like Balmford's students in a cafeteria deciding whether to opt for a vegetarian meal, or are they more like Molly shopping for g-free flour? Neither null model residuals --- which we will call "bias" --- nor raw visitation values can be interpreted unambiguously as "preference". "Preference" lies latent behind these observable patterns.

## 5.1 Overall visitation rates
```{r}
# Get mean proportional visitation by morphotype*bb.sp
visitation_prop <- visitation %>% 
  group_by(sample.id, bb.sp, ktype) %>%
  summarize(visits = n()) %>%
  group_by(sample.id, bb.sp) %>%
  mutate(prop.visits = visits/sum(visits)) %>%
  select(sample.id, bb.sp, ktype, prop.visits) %>%
  group_by(bb.sp) %>%
  nest() %>%
  mutate(fill = map(data, function(x) full_join(x, flower_cover_ktype) %>% replace_na(replace = list(prop.visits = 0)))) %>%
  unnest(fill) %>%
  select(-data) %>%
  group_by(bb.sp, ktype) %>%
  summarize(mean.prop.visits = mean(prop.visits)) %>%
  filter(ktype != 10) %>%
   mutate(ktype = case_when(
    ktype == 0 ~ "null",
    ktype == 1 ~ "disc",
    ktype == 2 ~ "funnel",
    ktype == 3 ~ "bell",
    ktype == 4 ~ "stalk-disc",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head",
    ktype == 9 ~ "brush"
  )) %>%
  mutate(ktype = factor(ktype, levels = c("null", "disc", "funnel", 
                                          "bell", "stalk-disc", "lip", 
                                          "flag", "head", "brush"))) %>%
  filter(!ktype %in% c("null", "brush")) %>% # type 0 were not reliably quantified (and rarely visited) and type 9 was vanishingly rare
  droplevels() %>%
  left_join(bb_traits)

null_model_overall <- ggplot(null_tab1, aes(ktype, preference)) +
  #geom_violin(fill = "gray70", color = "white", draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_boxplot(fill = NA, color = "black", outlier.shape = 4, size = 0.35) +
  geom_jitter(aes(color = Test), width = 0.3, alpha = 0.6) +
  theme_light(10) +
  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  labs(x = "Floral morphotype",
       y = "Bias")

raw_visits_overall <- ggplot(visitation_prop, aes(ktype, mean.prop.visits)) +
  geom_boxplot(fill = NA, color = "black", outlier.shape = 4, size = 0.35) +
  geom_jitter(width = 0.3, alpha = 0.5, color = "gray40") +
  theme_light(10) +
  labs(x = "Floral morphotype",
       y = "Mean proportional visitation")

ggsave("../ms_2/figures/raw_visitation_ktype.pdf", height = 6, width = 7)
ggsave("../ms_2/figures/raw_visitation_ktype.png", height = 6, width = 7)

overall_ktype <- (null_model_overall | raw_visits_overall)  +
  plot_annotation(tag_levels = 'A')

ggsave("../ms_2/figures/overall_ktype.pdf", overall_ktype, height = 4.5, width = 9.5)
ggsave("../ms_2/figures/overall_ktype.png", overall_ktype, height = 4.5, width = 9.5)
```

## 5.2 Visitation ~ tongue length
```{r}
gam1b <- gam(mean.prop.visits ~ 
              ktype + 
              s(pbl.w, by = ktype, k = 5),
             family = betar(eps = 0.001),
            data = visitation_prop) %>% getViz()

summary(gam1b)
check.gamViz(gam1b)
#plot.gamViz(gam1, allTerms = TRUE)

# Extract predictions
inv_logit <- function(x) {
  exp(x)/(1+exp(x))
}

gam1b_pred <- predict_gam(gam1b) %>%
  mutate(fit.plus.se = inv_logit(fit + (2*se.fit)),
         fit.minus.se = inv_logit(fit - (2*se.fit)),
         fit = inv_logit(fit))

raw_visits_pbl_GAM <- ggplot(gam1b_pred, aes(pbl.w, fit)) +
  geom_line() + # fit line
  geom_ribbon(aes(ymin = fit.minus.se, ymax = fit.plus.se), # Uncertainty
              fill = "black", alpha = 0.2) +
  geom_point(data = visitation_prop, # Original data
             aes(pbl.w, mean.prop.visits, color = reorder(bb.sp, pbl.w)), 
             inherit.aes = FALSE, alpha = 0.5) +
  facet_wrap(~ktype) +
  labs(x = "Tongue length (mm)", y = "Mean proportional visitation", 
       color = "Species") +
  scale_x_continuous(labels = scales::number_format(accuracy = 1)) +
  scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
  theme_light(10)

null_model_pbl_GAM <- ggplot(gam1_pred, aes(pbl.w, fit)) +
  geom_line() + # fit line
  geom_ribbon(aes(ymin = fit.minus.se, ymax = fit.plus.se), # Uncertainty
              fill = "black", alpha = 0.2) +
  geom_point(data = null_tab1, # Original data
             aes(pbl.w, preference, color = bb.sp), 
             inherit.aes = FALSE, alpha = 0.5) +
  #scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  facet_wrap(~ktype) +
  labs(x = "Tongue length (mm)", y = "Bias", 
       color = "Species") +
  scale_x_continuous(labels = scales::number_format(accuracy = 1)) +
  scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
  theme_light(10)

pbl_GAM_plot <- (null_model_pbl_GAM | raw_visits_pbl_GAM) + 
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A')

ggsave("../ms_2/figures/ktypeXtonguelength_raw_visitation.pdf", pbl_GAM_plot, height = 5, width = 8.5)
ggsave("../ms_2/figures/ktypeXtonguelength_raw_visitation.png", pbl_GAM_plot, height = 5, width =8.5)
```

## 5.3 Ordination of raw visitation
```{r}
# Visitation matrix
uncorrected_visits <- visitation_prop %>%
  select(bb.sp, ktype, mean.prop.visits) %>%
  pivot_wider(names_from = ktype, values_from = mean.prop.visits) %>%
  column_to_rownames("bb.sp")

# Call PCA
pcoa_visits <- capscale(uncorrected_visits ~ 1 , distance = "horn")

# Extract tongue length data
tongues2 <- bb_traits %>%
  semi_join(visitation_prop) %>%
  arrange(bb.sp)

# Ordisurf on tongue length
ordisurf_visits <- ordisurf(pca_visits, tongues2$pbl.w) 
summary(ordisurf_visits)

# Get variance explained
pcoa_visits_mds1 <- round(pcoa_visits$CA$eig[[1]] / sum(pcoa_visits$CA$eig)*100)
pcoa_visits_mds2 <- round(pcoa_visits$CA$eig[[2]] / sum(pcoa_visits$CA$eig)*100)

# Fortify whole model
pcoa_visits.fort <- fortify(pcoa_visits, axes = 1:2) %>%
  # for plotting, I prefer ktype to be numerically coded
  mutate(Label = recode(Label,
                        "disc" = "1",
                        "funnel" = "2",
                        "bell" = "3",
                        "stalk-disc" = "4",
                        "lip" = "5",
                        "flag" = "6",
                        "head" = "7"))

# Fortify plants
pcoa_visits.fort1 <- fortify(pcoa_visits, axes = 1:2) %>%
  filter(Score == "species") %>%
  rename(ktype = Label) %>%
  # for plotting, I prefer ktype to be numerically coded
  mutate(ktype = recode(ktype,
                        "disc" = "1",
                        "funnel" = "2",
                        "bell" = "3",
                        "stalk-disc" = "4",
                        "lip" = "5",
                        "flag" = "6",
                        "head" = "7"))

# Fortify bees
pcoa_visits.fort2 <- fortify(pcoa_visits, axes = 1:2) %>%
  filter(Score == "sites") %>%
  rename(bb.sp = Label) %>%
  left_join(bb_traits)

# Plot ordisurf

## Plotting function based on https://oliviarata.wordpress.com/2014/07/17/ordinations-in-ggplot2-v2-ordisurf/
surfplot2 <- function(x, y, min_score = 0, pca1, pca2) {

  grid <- x$grid # extracts the ordisurf object
  grid_exp <- expand.grid(x = grid$x, y = grid$y) # get x and ys
  grid_exp$z <- as.vector(grid$z) # unravel the matrix for the z scores

  contours <- data.frame(na.omit(grid_exp)) %>% # gets rid of the nas
    as_tibble() %>%
    rename(MDS1 = x, MDS2 = y)

  scores <- y %>% 
    filter((abs(MDS1) > min_score | abs(MDS2) > min_score) | Score == "BB.sp") %>%
    mutate(max.score = max(MDS1, MDS2))

  ggplot(scores, aes(MDS1, MDS2, pch = Score)) +
    geom_contour(data = contours, 
                 aes(MDS1, MDS2, z = z, color = stat(level)),
                 size = 2, alpha = 0.4, inherit.aes = FALSE) +
    #geom_point(size = 2) +
    geom_text_repel(aes(MDS1, MDS2, label = Label), key_glyph = "rect") +
    theme_light() +
    scale_color_viridis_c(name = "Tongue length") +
    xlab(paste("MDS1 ", "(", pca1, "%", ")", sep = "")) +
    ylab(paste("MDS2 ", "(", pca2, "%", ")", sep = "")) +
    coord_fixed() # Gavin Simpson says this is **really** important
}

## Plot
raw_visitation_ord <- surfplot2(ordisurf_visits, pcoa_visits.fort, 
         pca1 = pcoa_visits_mds1, pca2 = pcoa_visits_mds2)

ordination_plot <- (bias_ord | raw_visitation_ord) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A')

ggsave("../ms_2/figures/paired_ordination.png", ordination_plot, height = 6, width = 9.5)
ggsave("../ms_2/figures/paired_ordination.pdf", ordination_plot,  height = 6, width = 9.5)
```

## 5.4 Raw visitation ~ ktype x elevation x time
### 5.4.1 Call model
Binomial model works better than beta because of all the ones an zeros.
```{r}
visit_cover_sitedate <- visitation %>% 
  group_by(sample.id, bb.sp, ktype) %>%
  summarize(visits = n()) %>%
  group_by(sample.id, bb.sp) %>%
  mutate(prop.visits = visits/sum(visits)) %>%
  select(sample.id, bb.sp, ktype, prop.visits) %>%
  group_by(bb.sp) %>%
  nest() %>%
  mutate(fill = map(data, function(x) full_join(x, flower_cover_ktype) %>% replace_na(replace = list(prop.visits = 0)))) %>%
  unnest(fill) %>%
  select(-data) %>%
  filter(!ktype %in% c("0", "9")) %>% # type 0 were not reliably quantified (and rarely visited) and type 9 was vanishingly rare
  droplevels() %>%
  left_join(bb_traits) %>%
  separate(sample.id, c("site", "date"), sep = "_") %>%
  left_join(site_data) %>%
  unite(bb.ktype, c(bb.sp, ktype), remove = FALSE) %>%
  mutate(date = ymd(date),
         yday = yday(date),
         bb.ktype = factor(bb.ktype),
         site = factor(site),
         year = year(date))

# Binomial model
gam5 <- bam(prop.visits ~
              bb.ktype +
              s(site, by = bb.ktype, bs = "re") +
              s(year, by = bb.ktype, bs = "re") +
              te(yday, elev.mean, by = bb.ktype),
            #family = betar(eps = 0.001),
            family = "binomial",
            method = "fREML",
            discrete = TRUE,
            data = filter(visit_cover_sitedate, 
                          bb.sp %in% c("pasc", "prat", "telu", "hort", "soro", "wurf") &
                                         ktype %in% c("3", "5", "6", "7"))) %>% 
  getViz()

check.gamViz(gam5)
gam5_sum <- summary(gam5, re.test = FALSE) 
plot.gamViz(gam5)

# 
# # Beta model
# gam5b <- bam(prop.visits ~
#               bb.ktype +
#               s(site, by = bb.ktype, bs = "re") +
#               s(year, by = bb.ktype, bs = "re") +
#               te(yday, elev.mean, by = bb.ktype),
#             #family = betar(eps = 0.001),
#             family = betar(),
#             #family = "binomial",
#             method = "fREML",
#             discrete = TRUE,
#             data = filter(visit_cover_sitedate, 
#                           bb.sp %in% c("pasc", "prat", "telu", "hort", "soro", "wurf") &
#                                          ktype %in% c("bell", "lip", "flag", "head"))) %>% 
#   getViz()
# 
# check.gamViz(gam5b)
# summary(gam5b, re.test = FALSE) 
# plot.gamViz(gam5b)
```

### 5.4.2 Plot model
```{r}
# Tabulate model summary
gam5_s_table <- gam5_sum$s.table %>%
  as.data.frame() %>%
  rownames_to_column("term") %>%
  filter(str_detect(term, "te\\(")) %>%
  mutate(term = str_extract(term, ".{6}$")) %>%
  separate(term, c("bb.sp", "ktype")) %>%
  mutate(pval = signif(`p-value`, digits = 3)) %>%
  select(bb.sp, ktype, edf, pval) %>%
  as_tibble() %>%
  mutate(ktype = case_when(
    ktype == 3 ~ "bell",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head"  )) %>%
  mutate(ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")))

# Extract plotting data.
hort_3_gam5 <- plot(sm(gam5, 49))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("bell", n()))

hort_5_gam5 <- plot(sm(gam5, 50))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("lip", n()))

hort_6_gam5 <- plot(sm(gam5, 51))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("flag", n()))

hort_7_gam5 <- plot(sm(gam5, 52))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("head", n()))

pasc_3_gam5 <- plot(sm(gam5, 53))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("bell", n()))

pasc_5_gam5 <- plot(sm(gam5, 54))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("lip", n()))

pasc_6_gam5 <- plot(sm(gam5, 55))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("flag", n()))

pasc_7_gam5 <- plot(sm(gam5, 56))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("head", n()))

prat_3_gam5 <- plot(sm(gam5, 57))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("bell", n()))

prat_5_gam5 <- plot(sm(gam5, 58))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("lip", n()))

prat_6_gam5 <- plot(sm(gam5, 59))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("flag", n()))

prat_7_gam5 <- plot(sm(gam5, 60))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("head", n()))

soro_3_gam5 <- plot(sm(gam5, 61))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("bell", n()))

soro_5_gam5 <- plot(sm(gam5, 62))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("lip", n()))

soro_6_gam5 <- plot(sm(gam5, 63))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("flag", n()))

soro_7_gam5 <- plot(sm(gam5, 64))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("head", n()))

telu_3_gam5 <- plot(sm(gam5, 65))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("bell", n()))

telu_5_gam5 <- plot(sm(gam5, 66))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("lip", n()))

telu_6_gam5 <- plot(sm(gam5, 67))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("flag", n()))

telu_7_gam5 <- plot(sm(gam5, 68))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("head", n()))

wurf_3_gam5 <- plot(sm(gam5, 69))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("bell", n()))

wurf_5_gam5 <- plot(sm(gam5, 70))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("lip", n()))

wurf_6_gam5 <- plot(sm(gam5, 71))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("flag", n()))

wurf_7_gam5 <- plot(sm(gam5, 72))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("head", n()))

plotting_data_gam5 <- bind_rows(hort_3_gam5, hort_5_gam5, hort_6_gam5, hort_7_gam5,
                           pasc_3_gam5, pasc_5_gam5, pasc_6_gam5, pasc_7_gam5,
                           prat_3_gam5, prat_5_gam5, prat_6_gam5, prat_7_gam5,
                           soro_3_gam5, soro_5_gam5, soro_6_gam5, soro_7_gam5,
                           telu_3_gam5, telu_5_gam5, telu_6_gam5, telu_7_gam5,
                           wurf_3_gam5, wurf_5_gam5, wurf_6_gam5, wurf_7_gam5) %>%
  group_by(bb.sp, ktype) %>%
  mutate(z.norm = scale(z, center = FALSE),
         ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")),
         date = as_date(x)) %>%
  ungroup() %>%
  left_join(bb_traits) %>%
  left_join(gam5_s_table) %>%
  select(z, z.norm, date, x, elevation = y, se, 
         bb.sp, ktype, pbl.w, edf, pval) %>%
  mutate(sig = if_else(pval < 0.05, TRUE, FALSE))

ggplot(plotting_data_gam5, aes(date, elevation, z = z.norm, fill = z.norm)) +
  geom_raster() +
  geom_contour(color = "black",
               size = 0.1,
               bins = 40,
               aes(linetype = sig)) + # by fixing bins, we scale contours to each plot while letting the color ramp represent differences in scale across plots; I think this is a good compromise
  facet_grid(reorder(bb.sp, -pbl.w) ~ ktype) +
  scale_linetype_manual(values = c("dotted", "solid"))+
  scale_fill_viridis_c(na.value = "transparent") +
  labs(x = NULL, y = "Elevation (m)", fill = "Normalized\neffect", linetype = "p < 0.05") +
  theme_light(12) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5))

ggsave("../ms_2/figures/raw_visitation_gam.pdf", height = 5, width = 6.5)
ggsave("../ms_2/figures/raw_visitation_gam.png", height = 5, width = 6.5)
```

## 5.5 Raw flower cover
Beta seems to work a bit better for the flower cover, probably because it does not contain many ones or any zeros.

### 5.5.1 Call model
```{r}
flower_cover_ktype.prop <- flower_cover_ktype %>%
  group_by(sample.id) %>%
  mutate(prop.cover = cover/sum(cover)) %>%
  separate(sample.id, c("site", "date"), sep = "_") %>%
  mutate(date = ymd(date),
         yday = yday(date),
         ktype = factor(ktype),
         site = factor(site),
         year = year(date)) %>%
  filter(!ktype %in% c("10", "0", "9")) %>% # type 0 were not reliably quantified (and rarely visited) and type 9 was vanishingly rare
  droplevels() %>%
  left_join(site_data)

gam6 <- bam(prop.cover ~
              ktype +
              s(site, by = ktype, bs = "re") +
              s(year, by = ktype, bs = "re") +
              te(yday, elev.mean, by = ktype),
            #family = betar(eps = 0.001),
            family = "binomial",
            method = "fREML",
            discrete = TRUE,
            data = filter(flower_cover_ktype.prop, 
                            ktype %in% c("3", "5", "6", "7"))) %>% 
  getViz()

check.gamViz(gam6)
gam6_sum <- summary(gam6, re.test = FALSE) 
plot.gamViz(gam6)

gam6b <- bam(prop.cover ~
              ktype +
              s(site, by = ktype, bs = "re") +
              s(year, by = ktype, bs = "re") +
              te(yday, elev.mean, by = ktype),
            family = betar(),
            #family = "binomial",
            method = "fREML",
            discrete = TRUE,
            data = filter(flower_cover_ktype.prop, 
                            ktype %in% c("3", "5", "6", "7"))) %>% 
  getViz()

check.gamViz(gam6b)
gam6b_sum <- summary(gam6b, re.test = FALSE) 
plot.gamViz(gam6b)
```

### 5.5.2 Plot model
```{r}
# Tabulate model summary
gam6b_s_table <- gam6b_sum$s.table %>%
  as.data.frame() %>%
  rownames_to_column("term") %>%
  filter(str_detect(term, "te\\(")) %>%
  mutate(term = str_extract(term, ".{1}$")) %>%
  rename(ktype = term) %>%
  mutate(pval = signif(`p-value`, digits = 3)) %>%
  select(ktype, edf, pval) %>%
  as_tibble() %>%
  mutate(ktype = case_when(
    ktype == 3 ~ "bell",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head"  )) %>%
  mutate(ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")))

# Extract plotting data.
ktype_3_gam6b <- plot(sm(gam6b, 9))$ggObj[[1]] %>%
  mutate(ktype = rep("bell", n()))

ktype_5_gam6b <- plot(sm(gam6b, 10))$ggObj[[1]] %>%
  mutate(ktype = rep("lip", n()))

ktype_6_gam6b <- plot(sm(gam6b, 11))$ggObj[[1]] %>%
  mutate(ktype = rep("flag", n()))

ktype_7_gam6b <- plot(sm(gam6b, 12))$ggObj[[1]] %>%
  mutate(ktype = rep("head", n()))

plotting_data_gam6b <- bind_rows(ktype_3_gam6b, ktype_5_gam6b, ktype_6_gam6b, ktype_7_gam6b) %>%
  group_by(ktype) %>%
  mutate(z.norm = scale(z, center = FALSE),
         ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")),
         date = as_date(x)) %>%
  ungroup() %>%
  left_join(gam6b_s_table) %>%
  select(z, z.norm, date, x, elevation = y, se, 
         ktype, edf, pval) %>%
  mutate(sig = if_else(pval < 0.05, TRUE, FALSE))

ggplot(plotting_data_gam6b, aes(date, elevation, z = z.norm, fill = z.norm)) +
  geom_raster() +
  geom_contour(color = "black",
               size = 0.1,
               bins = 40) + # by fixing bins, we scale contours to each plot while letting the color ramp represent differences in scale across plots; I think this is a good compromise
  facet_wrap(~ktype) +
  scale_fill_viridis_c(na.value = "transparent") +
  labs(x = NULL, y = "Elevation (m)", fill = "Normalized\neffect", linetype = "p < 0.05") +
  theme_light(12) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5))

ggsave("../ms_2/figures/flower_cover_gam.pdf", height = 5, width = 6.5)
ggsave("../ms_2/figures/flower_cover_gam.png", height = 5, width = 6.5)
```

### 2021-06-09 ms_1/ms_1.Rmd
updated ms2

@@
 ## Cache
-# opts_chunk$set(cache = 2, cache.path = "output/cache/")
+# opts_chunk$set(cache = 2, cache.path = "ms_2/figures/cache/")
 
 ## Plots
-opts_chunk$set(fig.path = "../output/")
+opts_chunk$set(fig.path = "../ms_2/figures/")
 
@@
 ```{r fig1, echo = FALSE}
-knitr::include_graphics("../output/sites_GE.png")
+knitr::include_graphics("../ms_1/figures/sites_GE.png")
 ```
@@
 ```{r fig2, echo = FALSE}
-knitr::include_graphics("../output/gam_abund_grid_mod.png")
+knitr::include_graphics("../ms_1/figures/gam_abund_grid_mod.png")
 ```
@@
 ```{r fig3, echo = FALSE}
-knitr::include_graphics("../output/beta_stack.png")
+knitr::include_graphics("../ms_1/figures/beta_stack.png")
 ```
@@
 ```{r fig4, echo = FALSE}
-knitr::include_graphics("../output/gdm_combo.png")
+knitr::include_graphics("../ms_1/figures/gdm_combo.png")
 ```

### 2021-06-09 ms_2/ms_2.Rmd
updated ms2

@@
 ---
-title: "Bumble bees exhibit strong but flexible foraging biases through temporal and elevational gradients of floristic turnover"
+title: "Floral selectivity in mountain bumble bees and a test of the 'target use hypothesis'"
 
-author: Douglas B. Sponsler^1^*, Fabrice Requier^2^, Katharina Kallnik^1^, Alice Claßen^1^, A. Fabienne Maihoff^1^, Ingolf Steffan-Dewenter^1^
+author: Douglas B. Sponsler^1^*, Katharina Kallnik^1^, Fabrice Requier^2^, Alice Claßen^1^, A. Fabienne Maihoff^1^, Johanna Sieger^3^, Ingolf Steffan-Dewenter^1^
 
@@
 
-*Keywords*: optimal foraging, trait matching, pollinator, *Bombus*, alpine
+Alternative title: Mountain bumble bees exhibit strong but flexible foraging biases through temporal and elevational gradients of floristic turnover
+
+The process of resource selection by foraging animals has been a central theme of ecological research since the mid 20th century, particularly in the context of resource partitioning among sympatric species.
+Bumble bees as model system...
+Field studies have shown that floral choice in bumble bees is associated with the morphological fittedness between bumble bee tongue length to floral corolla depth.
+Morphological fittedness, however, is only a soft constraint on resource selection; bumble bees also tune their floral choices to learned patterns of reward, and they exhibit remarkable intelligence in learning the handling skills required to exploit difficult flower morphologies.
+
+Using a dataset consisting of more than 13,000 observations collected along a 1400 m elevation gradient over three consecutive years, we describe how bumble bee floral selection is constrained by morphology yet flexibly tuned to local conditions in elevation and time, with considerable idiosyncrasy in floral selection across species with similar tongue length.
+More results...
+
+Significance ...
+
+*Keywords*: foraging, trait matching, pollinator, *Bombus*, alpine
 
@@
 
-# INTRODUCTION
+# 1. INTRODUCTION
 
-<!-- Open with a discussion of the problem of foraging efficiently in a dynamic foraging environment, the balance of bias and flexibility, etc. -->
+The process of resource selection by animals has been an enduring theme of ecological research [@Johnson1980-co; @Beyer2010-vr].
+Methodological difficulties notwithstanding, the study of animals' resource choices is one of the central tasks of community ecology, where the selection and partitioning of resources are instrumental in community assembly, species coexistence, and ecosystem function [@Chesson2000-aw].
+Recently, questions of resource selection in ecological communities have taken on new urgency on account of anthropogenic climate and land use change, which threaten to disrupt existing patterns of species interactions and niche relationships [@Tylianakis2008-fi].
 
-Bumble bees are exceptionally versatile foragers whose ability to solve the puzzles posed by complex flower morphologies (Plowright and Laverty 1984), or indeed by human investigators (Alem et al.
-2016; Loukola et al.
-2017), renders almost any flower potentially accessible.
-Even when a gross mismatch between tongue length and corolla depth makes the nectar of a given flower inaccessible by ordinary means, some bumble bee species will resort readily to biting through the base of the flower to “rob” nectar, incidentally making the nectar available to other prospective foragers as well.
-Flower handling skills are, moreover, socially transmitted, amplifying the plasticity of individual behavior (Leadbeater and Chittka 2008; Loukola et al.
-2017).
-Nevertheless, the flexibility of bumble bee foraging is tuned to floral reward, which depends on both the intrinsic reward of a given flower and the searching/learning/handling costs associated with exploiting it.
-At an evolutionary level, the innate biases of bumble bees for certain floral traits likely reflect historic associations between floral traits and floral rewards (Raine and Chittka 2007).
-These innate biases, however, can be modulated or even nullified by patterns of reward learned during foraging (Kugler 1943; Maharaj et al.
-2019), which depend not only on intrinsic floral properties but also ecological modulators, such as the spatial distribution of floral patches or competition with other bumble bees (Ishii 2013; Plowright and Laverty 1984; Sowig 1989).
+Bumble bees (Hymentoptera: *Bombus* spp.) and their floral hosts provide a diverse yet tractable system in which to explore resource selection.
+Bumble bees are exceptionally versatile foragers whose ability to solve the puzzles posed by complex flower morphologies [@Plowright1984-lc] --- or, indeed, by human investigators [@Alem2016-rn; @Loukola2017-bt] --- renders almost any flower potentially accessible.
+Nevertheless, the flexibility of bumble bee foraging is tuned to floral reward, which depends on both the intrinsic reward of a given flower and the costs of exploiting it.
+These costs are closely related to the degree of morphological fittedness between bumble bees and their floral hosts, which consists principally in the matching of bumble bee tongue length to floral corolla depth.
+Bumble bee species can differ markedly in tongue length, and floral handling time increases when tongue length and corolla depth are poorly matched [@Inouye1980-mv; @Harder1983-bk], leading to a well-documented association between tongue length and floral choice in bumble bee communities [e.g.
+@Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn].
+This association between tongue length and floral niche is imperfect, though, and there can be considerable overlap in floral resource use between bumble bees of different tongue lengths [e.g.
+@Ranta1980-ok; @Goulson2008-oe].
+Moreover, bumble bee floral choice is also influenced by ecological modulators, such as the spatial distribution of floral patches or competition with other bumble bees [@Ishii2013-sm; @Plowright1984-lc; @Sowig1989-ld].
 
-The topic of floral bias in bumble bees has been approached via both laboratory experimentation (e.g.
-Kugler 1943; Raine and Chittka 2007) and observational field study (e.g.
-Heinrich 1976; Dave Goulson, Lye, and Darvill 2008).
-The former approach enables stronger causal inference with respect to the mechanisms of floral choice, but it suffers from all the usual limitations of biological realism associated with laboratory conditions.
-Observational field studies, however, tend to be anecdotal in scope, typically describing bumble bee foraging patterns in particular localities, over short time frames, and with limited floristic data to enable a distinction between true foraging bias and the neutral effects of floral abundance.
-The steep elevation gradients of mountain slopes provide an opportunity for more powerful observational study of bumble foraging because they generate broad gradients of environmental variation while controlling for confounding effects of geographic variation, such as genotypic differentiation within bumble bee species.
-When bumble bee visitation in such environments is sampled with sufficient spatial and temporal replication and complemented with quantitative floral surveying, true bias can be distinguished from the neutral effects of abundance, and variation in bias can be explored along elevational and temporal dimensions of floristic turnover.
+Understanding resource selection in bumble bee communities is more than just an exercise in ecological theory.
+Bumble bees are sensitive to climate and land use change [@Goulson2008-oe; @Soroye2020-zb], especially in the mountain ecosystems that host the world's richest bumble bee communities.
+Warming climates have already caused measurable upslope shifts in mountain bumble bees and their floral hosts [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc], which could introduce novel competitive interactions between bumble bee species that experience increased sympatry.
+At the same time, socioeconomic factors have led to a dramatic decline in traditional alpine grazing management, particularly in the European Alps, resulting in widespread pasture abandonment and reforestation of mountain meadows [@Tappeiner2006-jp; @Tasser2007-qo].
+The regression of meadows to forest entails a massive loss of floral resource availability, and even early stages of forest succession can cause declines in floral abundance and diversity, with corresponding effects on bumble bees [@Iserbyt2015-jc; @Walcher2017-kw; @Suzuki-Ohno2020-jv].
+In contrast, more accessible areas at the bases of mountains are undergoing agricultural intensification that displaces natural vegetation and favors grasses over forbs [@Tasser2002-om].
+Caught between adverse land use changes above and below and subject to climate-induced range shifts, mountain bumble bees are facing a future of reduced floral resource availability and altered sympatry, and the ability of bumble species to coexist under these changes will likely depend significantly on how they select and partition floral resources.
 
-In the present study, we focus on the question of bumble bee foraging bias with respect to floral morphology, and the lability of morphological biases across gradients of floristic variation.
-At the outset, we emphasize the use of the term “bias” in favor of the term “preference,” since the latter is laden with more assumptions and its inference under field conditions is problematic (Williams 2005).
-By “bias,” we mean simply an empirical disproportionality between the relative abundance of a given floral taxon or morphological class and its proportional visitation.
-The exact mechanism of this disproportionality --- whether an arbitrary “preference” on the part of the foraging bee, an unmeasured decoupling of relative abundance and rate of reward, an exogenous effect of competition, etc.
---- is beyond the scope of our discussion and probably best studied under laboratory conditions where strict factorial control is possible.
-We focus instead on the empirical phenomenon of bias, its variation, and its morphological correlations.
+In this study, we analyze a data set consisting of ~13,000 interactions between bumble bees and wildflowers recorded over 3 years along a 1400 m elevation gradient.
+Using these data in conjunction with complementary floral survey data, we explore the *selectivity* of the bumble bee community at the level of floral morphotype [@Kugler1970-gb].
 
-The meaning of “morphology,” however, is hardly more clear-cut than that of bias.
-Bumble bees and flowers vary in an array of interrelated morphological parameters, both within and between species, and the selection of relevant and measurable morphological metrics is nontrivial.
-For bumble bees, morphological variation consists principally (though not exclusively) of variation in tongue length.
-Tongue length in bumble bees corresponds functionally to corolla depth in flowers, and there is a well-established (though imperfect) association between tongue length and floral choice (e.g.
-Inouye 1980).
-Importantly, though, accessing floral reward involves more than just a match between tongue length and corolla depth.
-Such matching is entirely bypassed by nectar robbers, but even in the case of “legitimate” foraging, access to reward involves learning to manipulate floral structures that, in some floral morphotypes, are challenging irrespective of tongue length (Laverty 1994).
-For example, the characteristic Papilionaceous floral morphology of legumes requires a bee to separate tightly appressed keel petals to access anthers and nectaries (Westerkamp 1997), and the narrowly campanulate flowers of many Ericaceous flora pose a significant barrier to large-bodied visitors like bumble bees (Wallace 1977).
-Rather than considering various floral morphological traits independently, we rely on the synthesis of Kugler (1970), who developed a typology consisting of ten primary morphotypes and nested subdivision of each.
-Kugler’s typology was informed by laboratory studies of bumble bee flower choice (Kugler 1943), and there is a precedent for analyzing bumble bee visitation in terms of Kugler’s morphotypes (Neumayer and Paulus 1999; Schneller et al.
-2014).
+**This is a good place to get our terminological ducks in a row**
 
-<!-- Formulate this section more as a list of motivating questions; perhaps end with something that ties back into the more general introduction. -->
+We opt for the term *selectivity*, defined as the  (*sensu* @Beyer2010-vr) rather than the ...
+defined in this way, selectivity has the same conceptual meaning (though different mathematical derivation) as @Bluthgen2006-if *discrimination*.
 
-How do floral morphotypes compare in terms of total bumble bee visitation?
+We first compare the aggregate preferences of the bumble bee community across floral morphotypes and then model the preferences for each morphotype as a function of bumble bee tongue length.
+Then, by calculating preferences based on individual samples distributed through elevation and time, we model the variation of bumble bee preferences through temporal and elevational axes of floristic turnover, revealing the lability of bumble bee floral preferences.
+The balance of flexibility and constraint revealed in these analyses **Figure out how to finish this...**
 
-How does morphological bias vary across bumble bee species and tongue length classes?
+Our investigation also affords an opportunity to revisit an open question regarding the role of resource abundance as a driver of flower selection in bumble bees.
+In studies of resource selection, it is common to pose an abundance-only hypothesis in which foraging is random and selection merely a function of frequency of encounter, resulting in resource use patterns that mirror resource abundance patterns.
+For bumble bees, which are known to forage non-randomly, the abundance-only hypothesis can serve as a null-model against which observed use patterns can be compared to infer *selectivity*, i.e.
+resource use that is either positively or negatively disproportionate to resource abundance.
+It has also be proposed, however --- in what has been termed the *target use hypothesis* --- that bumble bees might select resources irrespective of abundance [@De_Valpine2013-be; @Harmon-Threatt2017-bz].
+This hypothesis, while obviously subject to boundary conditions of floral availability, is consistent with the observation that bumble bees sometimes persist in visiting favored flower species even when these resources are experimentally reduced to low abundance [@Heinrich1976-vb], and it was supported in a study comparing competing models of flower selection in *Bombus vosnesenskii* [@Harmon-Threatt2017-bz].
+If and when the *target use hypothesis* holds, uncorrected visitation rates are a better indication of preference than the residuals of an abundance-only null model, since the subtraction of presumed neutral effects of abundance when those effects do not, in fact, exist actually creates new abundance-dependent artifacts and obscures the true process of resource selection [@Harmon-Threatt2017-bz].
+On the other hand, failing to account for neutral effects of abundance when they do obtain similarly obscures the selection process underpinning foraging patterns [@Vazquez2003-qa; @Bluthgen2008-rp].
+In our study, we take advantage of the fact that there exists are well-established mechanistic basis for bumble bee flower choice, namely the compatibility between floral morphology and bumble bee tongue length.
+Taking this mechanism as a reference, we evaluate models of preference with and without abundance effects by comparing the explanatory power of bumble bee tongue length with respect to the preferences inferred under each model.
 
-Species' morphological biases give rise to patterns of diet (dis)similarity
+*"[. . .] preference  is  the  ratio  of  the  use  of  a  habitat  over  its availability, conditional on the availability of all habitats to the study animals." [@Aarts2008-fo]*
 
-How are the floral morphotypes distributed through elevation and time in our mountain ecosystem?
+One way we could handle our language and caveats is this.
+The kind of null model approach we used *does not* condition visitation to each focal resource on the abundances of all other resources, only on the proportional abundance of the focal resource.
+Following @Aarts2008-fo, we cannot then speak of *preference*.
+Instead, what we have is weaker concept of *selectivity* defined by Beyer2010-vr.
+We can, therefore, use the word *selectivity* after explaining its relation to preference in the strict sense.
+I do not want to go back and fit some crazy hierarchical model that conditions on the abundance of all morphotypes.
 
-For ubiquitously distributed bumble bee species, how does morphological bias vary through elevation and time?
+# 2. METHODS
 
-First, we compare the overall rate of bumble bee visitation across floral morphotypes.
-Then, we parse visitation rates by bumble bee species and explore the functional relationship between floral morphotype and bumble bee tongue length.
-These biases, together with the background effects of relative abundance, generate emergent patterns of diet (dis)similarity, which we present in the form of ordination analysis.
-Lastly, we analyze the lability of the morphological biases of individual bumble bee species through elevation and time, and compare these patterns of bias across bumble bee species and within floral morphotypes.
-The results of these analyses demonstrate both the strength of bumble bee floral bias, arising in part from morphological constraints, and the flexibility of bumble bee floral bias that enables them to forage adaptively in the dynamic floral communities of mountain landscapes.
-
-# METHODS
-
-## Field system
-
-<!-- I guess I'll need to tweak the wording of this section so that it is not a verbatim copy of the other manuscript.
-But it's okay if it recycles a lot of language because it is, after all, the same thing being described.
--->
-
-<!-- Describe general floristic and temperature patterns. Describe sampling in more detail. -->
+## 2.1 Field sampling
 
@@
 Fourteen of these pastures are extensively grazed by cattle or sheep, three are mowed for hay production, and eight have been abandoned throughout the last century and are no longer subject to any human management.
-The sites sampled in this study are the same as those used in a series of previous studies [@Hoiss2012-sn; @Hoiss2012-kr; @Hoiss2015-gv], but the data reported here are otherwise independent.
 
 Sampling consisted of repeated visits (hereafter, “visitation samples”) to each study site at approximately weekly intervals.
-Samples were only collected on dry days when the air temperature was at least 6°C.
-During each visit, bumble bees and their activity were recorded during a 50-minute transect walk.
-Bumblebees observed on or a given flower were counted as floral visitors.
-Bumble bee queens were identified to species level in the field, whereas workers and males were stored in individually labeled tubes in the freezer for later identification in the laboratory after Amiet (1996).
-While floral visitation by male bumble bees was recorded during visitation sampling, we chose to analyze only visitation by queens and workers.
-During observation, it was noted whether each bee was visibly carrying pollen at the time of its visit, but this provides at best an imperfect distinction between pollen and nectar foragers, and we opted not to differentiate between bees with and without pollen in our analyses.
-In conjunction with visitation observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.1 m^2^.
-Species identification followed @Lauber2007-ha and @Oberndorfer2001-ro.
+Sampling was only conducted during periods without rain and with an air temperature of at least 6°C.
+Bumble bees were recorded during one 50-minute transect walk per visit, and bumblebees observed on a given flower were counted as floral visitors.
+Queens were identified to species level in the field to avoid destructively sampling colony foundresses, but workers and males were collected and frozen for later identification in the laboratory.
+Species were identified following @Amiet1996-jz.
+While male bumble bees was recorded, we chose to analyze only the pooled visitation of by queens and workers.
+In conjunction with bumble bee observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.01 m^2^ and identified the plant species according to @Lauber2007-ha and @Oberndorfer2001-ro.
 
-While per-sample effort was standardized across all years, the total number of samples (i.e.
-site-dates) varied markedly,  with 153 in 2010, 212 in 2011, and 539 in 2012.
-The increased total sampling effort in 2012 was driven in part by the overall scarcity of bumble bees that year.
-Since the start of field sampling each year was determined by snowmelt, lower sites were sampled as early as April while higher sites could not be sampled until June or July.
-This resulted in a temporally staggered pattern of sampling in 2010 and 2012, while in 2011 a combination of early snowmelt and a late start to field activities resulted in sampling that was approximately aligned in time across the whole elevation gradient.
-See Appendix X for details of sampling patterns.
+While per-sample effort was standardized across all years, the total number of samples varied: 153 in 2010, 212 in 2011, and 539 in 2012.
+The start of field sampling each year was determined by snowmelt, so lower sites were sampled as early as April while higher sites could not be sampled until June or July.
+In 2010 and 2012, this resulted in a temporally staggered pattern of sampling, but a combination of early snowmelt and a late start to field activities in 2011 resulted in sampling that was approximately aligned in time across the whole elevation gradient.
+Due to these annual differences in sampling intensity and duration, we opted to pool data across years rather than include year effects in our models, which would multiply model complexity and be difficult to interpret.
+This means, however, that our analysis cannot detect year-to-year variation in floral preferences and niche overlap, and main effects cannot be distinguished from annual anomalies.
+With only three years of data, though, such annual inferences would be of little value, and we deemed them not worth the cost of added model complexity.
 
-## Data analysis
+The samples analyzed in this paper have been previously analyzed (Sponsler et al., submitted), but the analyses presented here are non-overlapping.
 
-### Data processing
+## 2.2 External data sources
 
-Bumble bee tongue-length data were obtained from the literature values compiled and summarized by Arbetman et al.
-(2017).
+### 2.2.1 Floral morphology
+
+Rather than considering various floral morphological traits independently, we relied on the synthesis of @Kugler1970-gb, who developed a typology consisting of ten primary morphotypes and nested subdivision of each.
+Kugler’s typology was informed by laboratory studies of bumble bee flower choice [@Kugler1943-uy], and there is a precedent for analyzing bumble bee visitation in terms of Kugler’s morphotypes [@Farnberger1996-xo; @Neumayer1999-il; @Schneller2014-ry].
+
+Kugler morphotype classifications of the plants in our study system were accessed via the BIOFLOR database [@Klotz2002-uc].
+Morphotypes were simplified to the primary classes numbered 0-10: 0 = null (includes wind-pollinated flora such as grasses), 1 = disc- and bowl-shaped flowers, 2 = funnel flowers, 3 = bell-shaped flowers, 4 = stalk-disc flowers, 5 = lip flowers, 6 = flag flowers, 7 = flower heads, 8 = spike flowers, 9 = brush flowers, 10 = trap flowers **(Figure 1)**.
+In a few cases, a plant species in our data set was not found in the BIOFLOR database.
+In most cases, morphotypes are constant at the genus level, and missing morphotypes were added based on the morphotypes found for congeners.
+In the case of the genus *Phyteuma*, we followed @Neumayer1999-il and classified it as a group 7 flower head.
+Morphotypes 9 and 10 were extremely rare in our study system and morphotype 8 was completely absent, so these morphotypes were omitted from all analyses.
+Plants belonging to morphotype 0 were typically not recorded during floral surveys due to their lack of conspicuous floral displays, and they were rarely visited by bumble bees, so this morphotype was also omitted from all analyses.
+
+### 2.2.2 Bumble bee tongue length
+
+Bumble bee tongue-length data (for the worker caste) were obtained from the literature values compiled and summarized by @Arbetman2017-xj.
 Values for *B.
 gerstaeckeri* and *B.
-mendax*, which were not included in Arbetman et al.
-(2017), were obtained from Obeso (1992) and Durieux (2000), respectively.
-We then binned bumble bee species into three tongue-length classes: short (< 7.5 mm), medium (7.5—10.0 mm), and long (>= 10.0 mm).
-Short-tongued species included *B.
-jonellus*, *B.
-lapidarius*, *B.
-pratorum*, *B.
-psithyrus*-group, *B.
-pyrenaeus*, and *B.
-terrestris/lucorum*.
-Medium-tongued species included *B.
-humilis*, *B.
-hypnorum*, *B.
-monticola*, *B.
-mucidus*, *B.
-pascuorum*, and *B.
-wurflenii*.
-Long-tongued species included *B.
-gerstaeckeri*, *B.
-hortorum*, and *B.
-mendax*.
+mendax*, which were not included in @Arbetman2017-xj, were obtained from @Obeso1992-yp and @Durieux2000-uk, respectively.
 
-Floral morphology was summarized using floral morphotypes of Kugler (1970), accessed via the BIOFLOR database (Klotz, Kühn, and Durka 2002).
-Morphotypes were simplified to the primary classes numbered 0-10: 0 = not applicable (includes wind-pollinated flora such as grasses), 1 = disc- and bowl-shaped flowers, 2 = funnel flowers, 3 = bell-shaped flowers, 4 = stalk-disc flowers, 5 = lip flowers, 6 = flag flowers, 7 = flower heads, 8 = spike flowers, 9 = brush flowers, 10 = trap flowers **(Figure 1)**.
-In a few cases, a plant species in our data set was not found in the BIOFLOR database.
-In most cases, morphotypes are constant at the genus level, and missing morphotypes were added based on the morphotypes found for congeners.
-The genus *Gentiana* consisted of both group 2 and group 4 representatives, so the missing species *Gentiana aspera* and *Gentiana ciliata* were classified based on visual comparison with their classified congeners.
-In the case of the genus *Phyteuma*, we followed Neumayer and Paulus (1999) and classified it as a group 7 flower head.
-Morphotypes 9 and 10 were extremely rare in our study system and were therefore omitted from all analyses.
-Plants belonging to morphotype 0 were typically not recorded during floral surveys due to their lack of conspicuous floral displays, so this morphotype was also omitted from all analyses.
+## 2.3 Data processing
 
@@
 In these cases, visitation samples were paired with the nearest available floral survey data within a maximum distance of 7 days.
-Visitation samples lacking corresponding floral survey data within 7 days were omitted from further analysis.
+Visitation samples lacking corresponding floral survey data within 7 days were omitted from further analysis, which resulted in dropping 40 site-dates out of a total of 895, leaving 855 site-dates included in our analyses.
 In a few cases, two consecutive visitation samples at a given site shared the same nearest floral survey sample; in these cases, the visitation samples were first pooled, then aligned to their shared nearest floral survey sample.
-In rare cases, a given floral species was recorded during visitation sampling but missed during floral surveying.
-These species were retroactively assigned the minimum floral area value of 0.01 m^2.
+In rare cases, a given floral species was recorded during visitation sampling but missed during floral surveying, and these species were retroactively assigned the minimum floral area value of 0.01 m^2^.
 
-Visitation data were extremely right-skewed because the majority of available flora received zero visits during a given sampling period.
-To facilitate model fitting, we created a binary version of visitation rate in which visited flora were assigned to 1 and unvisited flora to 0.
-We then modeled binary visitation using generalized models with binomial or Bernoulli distribution families.
-The conditional effects inferred from these models should, therefore, be interpreted as effects on visitation probability rather than visitation frequency.
+Having thus aligned visitation data and flower cover data, we summed visitation and floral cover across species within each floral morphotype.
+This yielded, for each of our 855 aligned samples, a cumulative value of flower cover for each floral morphotype and a cumulative value of visitation for each bumble- bee-morphotype pair.
 
-While our study spanned three years, we chose not to model year effects.
-Differences in sampling intensity and duration across years, for the reasons described above, make comparisons across years difficult to interpret, and we deemed these comparisons not worth the cost of added model complexity.
+## 2.4 Analysis of resource selection
 
-<!-- Should I be including year as a varying effect in these models? -->
+**Start this by working out my terminology.
+Are we saying preference or bias?
+And what about the distinction between preference, use, and selection?
+See @Johnson1980-co and @Beyer2010-vr as authorities on terminology.
+Whatever I choose, I then need to go revise the labels of the figures accordingly.
+I think we could make an argument for sticking wit the word "bias" to describe the inferences of our null model analysis, since we are allowing that the raw visitation patterns may be as good or better a measure of "preference".
+We don't actually find that, of course, but I think we should avoid calling the null model inferences "preference" as though it were a foregone conclusion that they represent preference while the uncorrected visitation rates do not.
+But preference, alas, is the word the litertaure seems to have settled on, so when I discuss my conclusions, I should talk about preference, comparing preference as inferred from bias and preference as inferred from uncorrected visitation rates.**
 
-<!-- State more explicityl how the models should be interpreted withr expect to bias -->
+*We will use the word "preference" in the phenomenological sense of @Beyer2010-vr to denote a disproportionality between resource availability and resource use while remaining agnostic as to the behavioral basis of this pattern.
+Or perhaps we will use the word bias but follow the same definition, since "bias" does not bear the teleological connotations of "preference".
+Double check that this is really how @Beyer2010-vr define preference.
+If so, it differs a bit from the more original definition of .*
 
-<!-- Report the effects of abundance that I am conditioning on -->
+A key challenge when interpreting interaction networks is to disentangle directed processes, such as a focal species' preference or avoidance of certain interaction partners, from the neutral effects of relative abundance [@Vazquez2003-qa; @Bluthgen2008-rp].
+For example, in the context of a bumble bee floral visitation network such as ours, abundant floral taxa should be expected to receive more visitation than rare taxa simply because they are more frequently encountered, and it would clearly be a mistake to interpret this as meaningful "preference" or "specialization" on the part of the bumble bees.
+One way to control for the neutral effects of relative abundance is to compare observed visitation patterns to the simulated visitation patterns of a null model in which plants are visited randomly in proportion to their relative abundance [@Vazquez2003-qa; @Bluthgen2008-rp].
+When constructing null models of interaction networks that are distributed through space and/or time, though, it is critical to respect the sampling unit of individual sites and dates, since only within this site-date sampling unit does floral visitation data correspond unequivocally to floral abundance data.
+Naive aggregation across sites and dates introduces forbidden links and decouples visitation and abundance [@Vaughan2018-gj].
 
-<!-- Perhaps show what happens if we remove the abundance interaction terms -->
+We constructed null models following @Vaughan2018-gj.
+First, resource selection is simulated for individual consumers --- in our case, individual bumble bees --- by according to a resource selection process in which the probability of a resource being chosen is proportional to its relative abundance [@Agusti2003-tt].
+By simulating the visitation of individual consumers rather than species in aggregate, the abundance of each resource, and therefore its probability of visitation, can vary across site-dates, thus preserving the within-sample correspondence between visitation and abundance.
+Simulated interactions are then summarized for each consumer species by taking the mean number of interactions with each resource across all individuals of the consumer species and all samples of the dataset.
+As this process is iterated, the result is a distribution of visitation rates of each consumer species to each resource under the null assumption that resource selection is determined only by resource abundance.
+The visitation distributions of the null model can then be compared to observed visitation rates to infer positive or negative bias of each consumer toward each resource.
+Since the mean of the null distribution can be interpreted as the expected value under null assumptions, the difference between the mean null value and the observed value can be interpreted as a measure of positive or negative bias [@Vaughan2018-gj].
+The statistical significance of this bias can be estimated as the proportion of simulated null values more extreme than the observed value.
 
-<!-- Mor edescriptive fgure captions -->
+We performed two rounds of null model analysis targeting different ecological questions.
+First, we followed the procedure described above, simulating our entire dataset in one modeling process to infer aggregate pattern of bias between bumble bee species and floral morphotypes.
+Then, we repeated the null model analysis for each site-date individually so that the bias between each bumble bee species and floral morphotype could be modeled as a variable through elevation and time.
+For both analyses, we used 500 iterations to achieve a robust sampling distribution, and we considered observed visitation rates to indicate significant bias if they fell outside the interval containing 99% of simulated values.
 
-<!-- Add something about model diagnostics -->
+Using the aggregate null model results, we modeled the biases toward each floral morphotype as a function of bumble bee tongue length using hierarchical generalized additive modeling (HGAM) with floral morphotype as a "by" smoothing factor [@Pedersen2019-kf].
+We then compared morphotype biases across bumble bee species by performing a principal components analysis and mapped bumble bee tongue length onto the ordination space by fitting a 2-dimensional GAM surface to the first two principal components.
 
-<!-- Consider adding year as a varying effect -->
-
-<!-- Figure 3 order by tongue length class -->
-
-<!-- Describe the differences in models in the Figure2 and 3 captions -->
-
-<!-- Explain filtering of ordination plot -->
-
-<!-- Year should really be included in Model 5, or just use 2012. -->
-
-<!-- Open each results paragraph with a summary sentence. -->
-
-### Model 1: Comparing overall visitation probability across floral morphotypes
-
-We first compared overall visitation probability across floral morphotypes.
-In this model, visitation for each sample was pooled across bumble bee species and within floral morphotypes prior to binarization.
-Visitation probability was then modeled using multilevel Bayesian regression with a Bernoulli distribution family (logit link function).
-*Bumble bee abundance* and *percent flower cover* (per morphotype) were included as a global interaction term, thus conditioning visitation probability on abundance and enabling the inference of bias.
-*Floral morphotype* was specified as a varying effect term with both a varying intercept and a varying slope with respect to *percent flower cover*, and a varying intercept term was also included for *site*.
-See Appendix X for the details of model specification and evaluation.
-
-### Model 2: Visitation probability as a function of floral morphotype and bumble bee tongue length
-
-To explore the functional underpinnings of bumble foraging bias, we analyzed visitation probability as a function of the interaction between floral morphology and bumble bee tongue length class.
-For this analysis, visitation was pooled within floral morphotype but not across bumble bee species.
-We then specified a Bayesian model of the same form as Model 1, but this time included a varying slope and intercept term for floral morphotype : bumble bee tongue length class and floral morphotype : bumble bee species.
-In this model, the relationship between floral morphotype and bumble bee tongue length class was the focus of inference, and the bumble bee species effect was added to avoid pseudoreplication arising from the grouping of observation within species.
-See Appendix X for the details of model specification and evaluation.
-
-### Model 3: Visitation probability as a function of floral morphotype and bumble bee species
-
-To focus on species-level morphotype biases, we removed the tongue length effect from Model 2 and re-fit the model considering only bumble bee species.
-See Appendix X for the details of model specification and evaluation.
-
-### Model 4: Aggregate diet ordination
-
-To visualize the net diet similarity between bumble bee species, we pooled visitation data across all samples within bumble bee species and plant species, then performed a principal coordinates analysis based on Horn-Morisita distance metric.
-To test the significance of tongue length class as a predictor of diet similarity, we performed a permutation test.
-
-### Model 5: The variation of morphological bias through elevation and time
-
-To explore the plasticity of floral biases, we first subsetted our data to include only the dominant bumble bee species (B.
-hortorum, B.
-pascuorum, B.
-pratorum, B.
-soroeensis, B.
-terrestris/lucorum, and B.
-wurflenii) and the most visited floral morphotypes (groups 3, 5, 6, and 7).
-We then constructed a generalized additive model (GAM) in which binary visitation was modeled as a tensor-product smooth of elevation and day-of-year, grouped by floral morphotype.
-This response was conditioned on linear interaction terms for bumble bee abundance : percent flower cover and percent flower cover : floral morphotype and a random intercept term for bumble bee species : floral morphotype.
-See Appendix X for the details of model specification and evaluation.
-
-## Software
-
-All analyses were conducted in `R` `r citep(citation())`.
-Bayesian multilevel models were implemented in Stan (Stan Development Team 2021) via the package `brms` `r citep(citation("brms"))`, and visualizations were constructed with the packages `tidybayes` `r citep(citation("tidybayes"))` and `ggdist` `r citep(citation("ggdist"))`.
-Ordination analysis and permutation testing were performed with the package `vegan` `r citep(citation("vegan"))` and visualized using the packages `ggplot2` `vegan` `r citep(citation("ggplot2"))` and `ggvegan` (Simpson 2019).
-GAMs were implemented with the package `mgcv` [@Wood_2017] and visualized with the package `mgcViz` [@Fasiolo2018aa].
-Data handling and visualization was performed with the `tidyverse` suite of R packages `r citep(citation("gdm"))`.
-Annotated R code is available in the Supplementary Material.
-
-# RESULTS
-
-## Summary of visitation and floral survey data
-
-<!-- Modify this section to describe how many observations/samples were dropped due to the non-alignment of visitation and survey data -->
-
-We recorded a total of 12,918 bumble- bee-flower interactions (excluding males) over the three years of our study.
-The metaweb across all sites and dates consisted of 16 bumble bee species (with *Bombus terrestric/lucorum* and *Bombus psithyrus* species groups lumped into species groups), 163 plant species (110 genera, 37 families), and 736 unique bumble- bee-plant interaction pairs.
-Five species --- *B.
+With the sample-wise null model results, we modeled the variation of bias between bumble bee species and floral morphotypes through elevation and time.
+For this analysis, we included only the six most abundant and widespread bumble bee species in our study system (*B.
+hortorum*, *B.
 pascuorum*, *B.
@@
 terrestris-lucorum*, and *B.
-wurflenii* --- accounted for the bulk of overall bumble bee abundance and were present over the entire elevation range of our sites **(Figure 1)**.
-*B.
-hortorum*, *B.
-jonellus*, and *B.
-psithyrus* (species-group) were widespread but at lower abundance.
-*B.
-monticola* occurred at moderate abundance but was rare or absent at sites below 1000 m.
-The remaining species were relatively rare in our study system.
-*B.
-mendax*, *B.
-mucidus*, and *B.
-pyrenaeus* occurred primarily above 1500 m, consistent with their known affinity for high elevation habitats [@Rasmont2010-ci].
-*B.
-gerstaeckeri*, *B.
-hypnorum*, *B.
-lapidarius* were sparsely distributed across mid-elevation sites.
-*B.
-humilis* was recorded only three times and only in one year, so we omitted it from all analyses.
+wurflenii*) to ensure sufficient data points through elevation and time, and we modeled their biases only toward the 4 most visited floral morphotypes (bell, lip, flag, and head flowers).
+Biases for each bumble- bee-morphotype pair were modeled together in a single HGAM with a tensor-product interaction smooth between elevation and time and bumble- bee-morphotype pair as a "by" smoothing factor.
 
-Floral surveying yielded a total of 352 plant species, representing 191 genera and 52 families.
-Of these, 155 species from 103 genera and 35 families were observed to be visited by bumble bees.
-Eight species --- *Rubus idaeus*, *Rosa canina*, *Juniperus communis*, *Larix decidua*, *Salix* sp., *Caltha palustris*, *Pulmonaria officinalis*, and *Rheum barbarum* --- were recorded in visitation data but not recorded during floral surveying.
-Each accounted for no more than 4 visits in total over the three years of our study, and they were omitted from the analysis of floral abundance.
+## 2.5 Software
 
-```{r fig1, echo = FALSE,  fig.cap = paste("Each floral morphotype is exemplified by a representative species common in our study system.
+All data processing and analyses were conducted in `R` [@RCoreTeam-2021aa].
+Data handling and visualization were performed with the `tidyverse` suite [@Wickham2019-aa].
+Null models were constructed using `econullnetr` [@Vaughan2018-gj].
+GAM analyses were performed with `mgcv` [@Wood_2017] and visualized with `mgcViz` [@Fasiolo2018aa] and `tidymv` [@Coretta2021-aa].
+PCA was performed with the `vegan` [@Oksanen_2019] and visualized using `ggplot2` [@Wickham2016-aa] and `ggvegan` [@Simpson2019aa].
+Annotated R code is available in the Appendix X.
+
+# 3. RESULTS
+
+## 3.1 Sampling overview
+
+```{r fig1, echo = FALSE,  fig.cap = paste("Visited families within each floral morphotype.
+Barplots show the total number of visits to each family across all bumble bee species and samples (note the rescaling of the y-axis across panels).
+Each floral morphotype is exemplified by a representative species common in our study system: 0 = Plantago lanceolata, 1 = Potentilla erecta, 2 = Gentiana asclepiadea, 3 = Campanula glomerata, 4 = Silene nutans, 5 = Ajuga reptans, 6 = Lotus corniculatus, 7 = Centaurea jacea, 8 = Salix retusa.
 Illustrations are modifications of public domain works obtained from www.plantillustrations.org.
 For details of original sources, see Appendix X.")}
-knitr::include_graphics("../output/ktype_abund_gam_mod.png")
+knitr::include_graphics("../ms_2/figures/morphotype_families_mod.png")
 ```
 
-## Overall visitation probability by floral morphotype
+## 3.2 Selection bias by species
 
-Lip (group 5) and flag (group 6) flowers had the highest overall visitation probability, with a conditional mean visitation probabilities of 0.36 (95% CI = 0.30–0.42) and 0.33 (95% CI = 0.27–0.39), respectively (Figure 1).
-In second place, head flowers (group 7) had a mean conditional visitation probability of 0.19 (95% CI = 0.16–0.24).
-Bell (group 3) flowers were a distant third with a conditional mean visitation probability of 0.092 (95% CI = 0.07–0.12), followed by stalk-disc (group 4), funnel (group 2), and disc flowers (group 1) with conditional visitation probabilities of 0.07 (95% CI = 0.05–0.09), 0.05 (95% CI = 0.04–0.07), and 0.05 (95% CI = 0.03–0.07), respectively.
-Unclassified (group 0) flowers, consisting mainly of wind-pollinated grasses, were the least-visited, with a conditional mean visitation probability of 0.03 (95% CI = 0.02–0.04).
-Our model explained approximately 35% of the total variance (97.5% CI = 34–36%).
-
-```{r fig2, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability by floral morphotype. Point-interval plots depict 66\\% and 95\\% CIs.")}
-knitr::include_graphics("../output/brm_pvis01_plot1.pdf")
+```{r fig2, echo = FALSE,  fig.cap = paste("Empirical visitation rates compared to null model simulations for each bumble bee species and floral morphotype.
+Gray bars depict 99\\% intervals of null model simulations and points depict observed visitation rates.
+Points are color-coded according to where they fall with respect to the 99\\% interval of the null model: red = above (positive bias), blue = below (negative bias), gray = within (nonsignificant bias).")}
+knitr::include_graphics("../ms_2/figures/null_model_spwise_preference.png")
 ```
 
-## Visitation probability by flower morphotype, bumble bee species, and bumble bee tongue length class
+## 3.3 Overall selection bias by morphotype
 
-Descriptions of tongue classes are based on Model 2 and refer to Figure 2, and descriptions of species are based on Model 3 and refer to Figure 3.
-Both models explained approximately 28% of total variance (97.5% CI = 27–29%).
-As in Model 1, visitation rates refer to conditional visitation probabilities.
-
-Disc (group 1) flowers were visited at similar rates across tongue length classes.
-At the species level, disc flowers were visited at relatively high rates by the short-tongued B.
-terrestris/lucorum and the medium-tongued Bombus monticola.
-Visitation to funnel (group 2) flowers was dominated by long-tongued species, an effect driven entirely by the Aconitum (group 2) specialist B.
-gerstaeckeri.
-Bell (group 3) flowers were visited at a higher rate by short-tongued species than by long-tongued species, with medium-tongued species exhibiting intermediate visitation rates.
-At the species-level, bell flowers were visited at highest probability by the short-tongued species B.
-jonellus, B.
-pyrenaeus, and B.
-soroeensis, and by the medium-tongued species B.
-hypnorum and B.
-monticola.
-Stalk-disc (group 4) flowers were visited at similar rates across all tongue length classes, and there was also broad overlap in visitation rates at the species level.
-Lip (group 5) flowers were also visited at similar rates across tongue length classes, but at the species level B.
-pascuorum exhibited an uniquely high preference for lip flowers.
-Flag (group 6) flowers were visited at a higher rate by medium-tongued species than by short-tongued species, with long-tongued species exhibiting intermediate visitation rates.
-Flag flowers generated strong separation in visitation rate at the species level, receiving very high visitation from the medium-tongued B.
-pascuorum and B.
-mucidus, moderately high visitation from the short-tongued B.
-terrestris/lucorum and B.
-lapidarius, the medium-tongued B.
-monticola, and the long-tongued B.
-mendax and B.
-hortorum, and low visitation by the short-tongued B.
-soroeensis, B.
-pyrenaeus, B.
-psithyrus, B.
-pratorum, and B.
-jonellus, the medium-tongued B.
-hypnorum, and the long-tongued B.
-gerstaeckeri.
-
-```{r fig3, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability by Kugler morphotype and bumble bee tongue length class.
-Point-interval plots depict 66\\% and 95\\% CIs.
-Letters indicate group differences based on evidence ratio at alpha = 0.05.
-Note that the range of the x-axis has been rescaled for each panel to emphasize the comparison of visitation probabilities across tongue length classes.")}
-knitr::include_graphics("../output/brm_pvis00_plot1.pdf")
+```{r fig3, echo = FALSE,  fig.cap = paste("Overall selection bias (A) and uncorrected visitation rate (B) of the bumble bee community toward each floral morphotype.
+Each point represents the bias/visitation of one bumble bee species toward the corresponding floral morphotype.
+The color of each point in panel A, retained from Figure 2, indicates whether a given observed bias is significantly (p < 0.01) positive or negative with respect to null model simulations.")}
+knitr::include_graphics("../ms_2/figures/overall_ktype.png")
 ```
 
-```{r fig4, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability by Kugler morphotype and bumble bee species.
-The tongue length class of each species is color-coded.
-Point-interval plots depict 66\\% and 95\\% CIs.
-Note that the range of the x-axis has been rescaled for each panel to emphasize the comparison of visitation probabilities across bumble bee species")}
-knitr::include_graphics("../output/brm_pvis02_plot1.pdf")
+## 3.4 Tongue length as a predictor of biases
+
+```{r fig4, echo = FALSE,  fig.cap = paste("Selection biases (A) and uncorrected visitation rates (B) for each morphotype regressed against bumble bee tongue length.
+Each point represents the bias/visitation of one bumble bee species toward the corresponding floral morphotype.
+Curves depict fitted GAM smooths with 95\\% confidence bands.
+Points are color-coded by species, and species are presented in the legend in order of increasing tongue length.")}
+knitr::include_graphics("../ms_2/figures/ktypeXtonguelength_preference.png")
 ```
 
-## Diet ordination
-
-Axis 1 accounted for 24% of total diet variance and primarily separated long- and medium-tongued species from short-tongued species (Figure 4).
-Long- and medium-tongued species scored low on Axis one and were associated with the lip flowers (group 5) Ajuga reptans, Rhinanthus glacialis, and Stachys alopecuros) and flag flowers (group 6) Anthyllis vulneraria, Lotus corniculatus, and Trifolium pratense.
-Short tongued species scored high on Axis one and were associated with the bell flowers (group 3) Campanula scheuchzeri, Erica carnea, and Geum rivale and the head flowers (group 7) Carduus defloratus, Centaurea jacea, and Phyteuma orbiculare.
-Axis 2 accounted for 19% of total variance and primarily explained variation within short-tongued species.
-Short-tongued species scoring high on Axis 2 were associated with Campanula scheuchzeri while those scoring low on Axis 2 were associated with Erica carnea.
-Short-tongued species scoring close to zero on Axis 2 also scored lower on Axis 1 (closer to medium- and long-tongued species) and were associated with head flowers.
-Tongue length class explained 22% of diet variance (Adonis permutation test, F = 1.68, p = 0.006).
-While short-tongued bees appeared most broadly spread in ordination space, there were no significant differences in dispersion across tongue length classes (F = 0.62, p = 0.55).
-
-```{r fig5, echo = FALSE,  fig.cap = paste(" Principal coordinates ordination of bumble bee visitation pooled across all samples.")}
-knitr::include_graphics("../output/pcoa_aggregate.png")
+```{r fig5, echo = FALSE,  fig.cap = paste("Ordination biplots of bumble bee species according to their selection biases (A) and uncorrected visitation rates (B) for each floral morphotype.
+Floral morphotypes are abbreviated with their numerical codes and illustrated as in Figure 1.
+Colored bands depict a 2-dimensional GAM smooth representing the relationship between the first two principal components and bumble bee tongue length.
+Morphotype illustrations are as described in Figure 1.")}
+knitr::include_graphics("../ms_2/figures/paired_ordination_mod.png")
 ```
 
-## The variation of morphological bias through elevation and time
+## 3.5 Lability of bias through elevation and time
 
-Model 5 explained 33.8% of total variance (adjusted R2 = 0.37).
-All tensor product smooths (Figure 5) were significant at p < 0.001 except for the one representing B.
-pratorum and flag (group 6), which was more weakly significant at p = 0.02.
-As in Models 1-3, the results of Model 5 should be interpreted in terms of conditional visitation probability.
-
-Bell (group 3) flowers were visited at highest rates early in the season and at low elevation by all species except B.
-hortorum, whose visitation rate to bell flowers peaked in mid-season and at mid-elevation.
-Notably, the medium-tongued B.
-pascuorum and B.
-wurflenii also exhibited relatively high visitation to bell flowers at mid-season and mid-elevation, while the visitation rates of the short-tongued B.
-pratorum, B.
-soroeensis.
-and B.
-terrestris lucorum were more concentrated at the early-season and low-elevation peak.
-
-Lip (group 5) flowers exhibited roughly the mirror image of the pattern of visitation rate seen in bell flowers, with peak visitation late in the season and at low elevation, but relatively high visitation extending toward mid-season and mid-elevation.
-This pattern was consistent for all bumble bee species.
-
-The pattern of visitation rates to flag (group 6) flowers fell generally along the lines of tongue length class.
-The short-tongued B.
-pratorum and B.
-soroeensis exhibited peak visitation early in the season and at low elevation.
-The long-tongued B.
-hortorum and the medium-tongued B.
-pascuorum and B.
-wurflenii exhibited peak visitation at mid-season, though B.
-pascuorum exhibited a bimodal pattern in elevation, with peaks around 1000 m and 1750 m, while the visitation rate of B.
-hortorum declined monotonically from high elevation to low.
-The short-tongued B.
-terrestis/lucorum broke the tongue-length pattern, exhibiting a diffuse pattern of visitation rate more similar to that of the long- and medium-tongued species than the other short-tongued species.
-
-Visitation rates for head (group 7) flowers were more or less bimodal for all species, with a first peak in mid-season and low elevation and a second peak in late-season and high elevation.
-
-```{r fig6, echo = FALSE,  fig.cap = paste("GAM analysis of bias as a function of the tensor-product interaction of elevation and day of year, parsed by bumble bee species and floral morphotype and conditioned on the abundance of each bumble bee species and the relative abundance of each floral morphotype.
-The color ramp of each smooth has been scaled to equal color range, thereby obscuring the absolute magnitude of effect.
-To avoid cluttering the figure, we have opted not to include a unique color ramp key for each plot, since the focus of our inference is on relative rather than absolute patterns.")}
-knitr::include_graphics("../output/ktype_lability_gam_mod.png")
+```{r fig6, echo = FALSE,  fig.cap = paste("Variation in bias through elevation and time modeled as HGAM surfaces.
+Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns.
+Significant (p < 0.05) smooths are depicted with solid contour lines and nonsignificant smooths with dashed contour lines.")}
+knitr::include_graphics("../ms_2/figures/preference_lability_gam.png")
+```
+```{r fig7, echo = FALSE,  fig.cap = paste("Variation in uncorrected visitation rates through elevation and time modeled as HGAM surfaces.
+Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns.
+Significant (p < 0.05) smooths are depicted with solid contour lines and nonsignificant smooths with dashed contour lines.")}
+knitr::include_graphics("../ms_2/figures/raw_visitation_gam.png")
 ```
 
-# DISCUSSION
+```{r fig8, echo = FALSE,  fig.cap = paste("Variation in flower cover through elevation and time modeled as HGAM surfaces.
+Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns.
+Significant (p < 0.05) smooths are depicted with solid contour lines and nonsignificant smooths with dashed contour lines.")}
+knitr::include_graphics("../ms_2/figures/flower_cover_gam.png")
+```
 
-## Foraging bias and floral morphology
+# 4. DISCUSSION
 
-Our study is not the first to compare bumble bee visitation rates across Kugler’s (1970) floral morphotypes (see Neumayer and Paulus 1999; Schneller et al.
-2014), but to our knowledge it is the first to enable the inference of bias across morphotypes by controlling statistically for the neutral effects of abundance.
+So, we would conclude that (A) abundance matters but (B) preference matters, too.
+We find support for a middle ground between the abundance-only and preference-only models.
+Need to acknowledge the potential for nonlinear "functional responses" of use to abundance [@Mysterud1998-rs].
 
-When bumble bee visitation is pooled across all species, it is clear that lip, flag, and head flowers dominate the floral community in terms of conditional visitation probability.
-Patterns of pooled visitation are, of course, driven disproportionately by the behavior of the most abundant bumble bee species, which in our system were B.
-hortorum, B.
-pascuorum, B.
-pratorum, B.
-psithyrus, B.
-soroeensis, B.
-terrestris/lucorum, and B.
-wurflenii.
-Visitation to lip and flag flowers was driven especially by the medium-tongued B.
-pascuorum, the most abundant bumble bee in our study system, and secondarily by the medium-tongued B.
-wurflenii.
-Visitation to head flowers was driven mainly by the short-tongued B.
-soroeensis and B.
-psithyrus.
+## 4. Caveats:
 
-The influence of tongue length on visitation bias was pronounced for some morphotypes but not for others.
-In the cases of disc, and stalk-disc, and unclassified flowers, visitation rates were uniformly low across all tongue length classes.
-The opposite was true of lip flowers, which were visited at similarly high rates by all tongue length classes.
-Bell flowers and head flowers were visited chiefly by short-tongued bees, consistent with the shallow corollas of these morphotypes.
-In contrast, funnel flowers, with their deep corollas were visited chiefly by long-tongued bees, though this effect was driven disproportionately by B.
-gerstaeckeri, which specializes on the funnel flowers of the genus Aconitum.
-Flag flowers were visited at the highest rate by medium-tongued species, and especially by B.
-pascuorum and B.
-mucidus.
+Inference of preference is sensitive to the scale at which availability is estimate [@Beyer2010-vr].
+My floral survey data estimates availability more or less at the meadow scale, right?
+But I don't know what is available outside the meadow.
+This problem is also discussed in [@Harmon-Threatt2017-bz], specifically in the context of bees and flowers.
 
-While the variation in floral bias across tongue length classes we observed largely comports with the established relationship between tongue length and corolla depth, variation in bias within tongue length classes demonstrates the insufficiency of simple morphological matching to explain bumble bee foraging patterns.
-Nectar robbing is one mechanism by which the relationship between tongue length and corolla depth might be relaxed, but its influence in our data is unclear.
-Of the bumble bees in our study system, B.
-terrestris/lucorum and especially B.
-wurflenii are known to engage frequently in nectar robbing (Brian 1954).
-In the case of B.
-wurflenii, however, its morphotype affinities did not differ markedly from those of other medium-tongued species, and it may be that the condition of medium-tongue length alone confers enough to foraging flexibility in our study system to dampen the divergent effects of nectar robbing.
-Nectar robbing might, however, account for the deviation of B.
-terrestris/lucorum from most other short-tongued species with respect to flag flowers.
-With the exception of B.
-lapidarius, all short-tongued species besides B.
-terrestris/lucorum eschewed flag flowers, while B.
-terrestris/lucorum visited them at rates comparable to medium- and long-tongued species.
-Indeed, B.
-terrestris/lucorum was, on the whole, an idiosyncratic forager, demonstrating a relatively high affinity for disc flowers and unclassified flowers (particularly Plantago lanceolata) that were largely ignored by other species.
-Perhaps the most striking divergence of floral bias within tongue length class, though, was the divisive effect of bell flowers on short-tongued bumble bees.
-B.
-pyrenaeus, B.
-soroeensis, B.
-pratorum, and B.
-jonellus visited bell flowers at high rates, while B.
-terrestris/lucorum, B.
-psithyrus, and B.
-lapidarius visited them relatively rarely.
-A similar split among the short-tongued species occurred with respect to head flowers, which were visited at high rates by B.
-soroeensis and B.
-psithyrus but at relatively low rates by other short-tongued species.
-There are no obvious morphological distinctions within short-tongued bumble bees that would account for these divergent affinities; ecological drivers such as competition or differences in spatial or temporal activity patterns are more likely explanations, as will be revisited below in our discussion of overall diet similarity.
+The sensitivity of "preference" to availability [@Johnson1980-co; @Beyer2010-vr].
 
-## Diet similarity
+*First, the estimates ofpreference  are  not  directly  interpretable  as  reflectingthe  intrinsic  behavioural  motivations  of  the  animal.Positive  (or  negative)  preference  for  a  given  habitatimplies only that the habitat is used more (or less) fre-quently than expected by chance,conditional upon thedefined sample of availability.
+Even  high  levels  of  useof a habitat can be associated with negative coefficientsif that habitat is common, and, conversely, low levelsof  use  can  be  associated  with  positive  coefficientsif the habitat is rare.
+The sign and magnitude of thecoefficient   cannot   be   interpreted   with   respect   tothe   behavioural   decision   the   animal   is   making.* [@Beyer2010-vr].
 
-Our finding that bumble bee diet ordination produces a primary axis corresponding to tongue length — and associated with affinity for legumes — and a secondary axis describing the dispersion of short-tongued species is consistent with studies in Poland (Dave Goulson, Lye, and Darvill 2008), France (Durieux 2000), and England (D.
-Goulson et al.
-2005) that analyzed bumble bee diet by similar methods, suggesting that this may be a general pattern among bumble bee communities, at least within temperate Europe.
-While our test of beta-dispersion did not reveal significant differences between tongue length classes, the spread of short-tongued bees along the secondary axis of our ordination, together with their divergence with respect to visitation rates on bell and head flowers described above, suggests that resource partitioning may be more pronounced among short-tongued bumble bees than among their medium- and long-tongued counterparts.
-It is tempting to invoke interspecific competition as a cause of the apparent niche partitioning among short-tongued bumble bees, in agreement with previous studies that have demonstrated the potential for severe competition between bumble bees of similar tongue length (e.g.
-Heinrich 1976; Inouye 1978; Pyke 1982).
-But the importance of competition in bumble bee communities is not universal (e.g.
-Ranta and Vepsäläinen 1981; Williams 1989), and in the rich wildflower meadows of our (sub)alpine system, bumble bee populations are more likely limited by harsh abiotic conditions than by food availablility.
-Alternatively, it may be that short-tongued bees in our study system differed from one another in their affinities for specific habitats (Schneller et al.
-2014) or spatial configurations of resource patches (Sowig 1989), and their divergent morphological biases arose secondarily from these effects.
+*The second consequence is that estimates of prefer-ence  based  on  different  samples  of  availability  (e.g.Design III inManlyet al.2002) are not directly com-parable.
+The  sign  and  magnitude  of  the  estimate  ofpreference  is  a  function  of  both  the  sample  of  useand the sample of availability.
+This creates an identifia-bility  problem  because  preference  can  change  as  afunction of changing use and/or changing availability.If   the   sample   of   availability   differs   between   twoestimates of preference, it is meaningless to draw infer-ences based on the difference in preference.
+Differentpreferences can arise from the same selection process,and   similar   preferences   can   arise   from   differentselection processes.* [@Beyer2010-vr].
 
-## The lability of bias through elevation and time
+This is exactly the problem I am facing.
+I think, though, I can present my "preference lability analysis" together with the raw visitation and availability data, discuss this limitation explicitly, and let the reader decide what to make of it.
+This is also discussed very clearly, and in the context of bumblebees, in @Harmon-Threatt2017-bz.
 
-While the intraspecific lability of bumble bee foraging biases has been demonstrated under laboratory conditions (Ings, Raine, and Chittka 2009; Maharaj et al.
-2019) and in a localized field study (Raine and Chittka 2007), ours is the first study, to our knowledge, to characterize the in situ lability of foraging biases within multiple species through elevational and temporal gradients of floristic turnover.
-In our analysis, it is evident that morphological biases in bumble bees, insofar as they are captured by the morphotypology used in our study, are not fixed but rather tuned to floristic (and possibly abiotic) context and therefore dynamic through elevation and time.
-Notably, bumble bee species differed from one another not just in their overall bias toward given morphotypes but also in when and where they were biased toward a given morphotype.
-This is especially evident in the case of flag flowers, for which positive bias was maximized in mid-season for B.
-hortorum, B.
-pascuorum, B.
-terrestris/lucorum, and B.
-wurflenii but in early-season for B.
-pratorum and B.
-soroeensis.
-This pattern corresponds to the overall biases of these species toward flag flowers, which were primary resources for B.
-hortorum, B.
-pascuorum, B.
-terrestris/lucorum, and B.
-wurflenii but only marginal resources for B.
-pratorum and B.
-soroeensis.
+*Given that preference is conditional on avail-ability,  that  availability  is  likely  to vary  in  space  andthat preference is likely to change as a function of avail-ability, there is little theoretical basis for believing thatpreferences estimated in one area will be good predic-tors  of  preference  in  others.* [@Beyer2010-vr].
 
-# CONCLUSIONS
+In answer to this, I can point to the largely agreeing results of other bumble bee studies.
 
-For bumble bees, floral bias is a labile product of innate preferences, the soft constraints of morphological compatibility, and the dynamic ecological context in which foraging occurs.
-By conditioning bumble bee visitation data upon concurrent floral abundance data across spatially and temporally replicated samples, we demonstrate the strength bumble bee floral bias, its functional underpinnings, its variation across species, and its lability within species in response to environmental variation.
-Such tuning of floral bias has obvious adaptive significance for bumble bees, which are characteristically — and most likely primordially — associated with mountain habitats (Williams 1985; Hines 2008), where steep temperature gradients and short growing seasons generate extreme spatial and temporal floral turnover.
+*Models of habitat preference provide a phenomenolo-gical  description  of  habitat  use,  conditional  on  thesample of availability.
+Although some of the simplify-ing  assumptions  upon  which  these  models  are  often based  may  be  biologically  unrealistic,  they  can  stillhave predictive power (Boyceet al.2002;Wienset al.2008).
+The degree, however, to which simple modelsof habitat preference help us to develop a mechanisticunderstanding of animal movement and distribution islimited.
+Mechanistic models are required to advanceour  biological  understanding  of  the  behavioural  pro-cesses  that  result  in  animal  movement,  distributionand habitat use.*
 
-# ACKNOWLEDGEMENTS
+Tongue length is my mechanistic model.
+And I think the observation that the mechanistic model better explains use-availability bias than raw selection indicates that the use-availability approach better captures the way bumble bees make foraging decisions.
 
-We thank D.J. McNeil for helpful conversations during data analysis.
+Need to acknowledge the potential for nonlinear "functional responses" of use to abundance [@Mysterud1998-rs].
+
+*The problem of conflating preference and niche partitioning:*
+
+Our study is set in the Berchtesgadener Alps of southeast Germany, a landscape of flower-rich mountain meadows embedded in a matrix of coniferous forest.
+This system is one of abundant floral resources but harsh abiotic conditions, so it is likely that competition for floral resources is low, or at least transient, and that bumble bee populations are limited instead by the shortness of the growing season, as has been suggested in the similar environment of the Eyne Valley of France [@Iserbyt2008-jn].
+Yet while *realized* competition between bumble bees in our system may be low, the question of *potential* competition is highly relevant because mountain ecosystems are currently subject to at least two powerful drivers of directed change.
+
+Resource usage is typically studied as a phenomenological level, with questions of mechanism relegated to speculations regarding the latent process of resource selection that lurks behind realized use patterns [@Beyer2010-vr].
+In our study, we evaluate the target use hypothesis by reference to a known mechanism of floral selection, namely bumble bee tongue length.
+
+# 5. CONCLUSIONS
+
+# ACKNOWLEDGMENTS
+
+We thank D.J. McNeil for helpful conversations during data analysis. O. Mitesser kindly provided a high-performance computing environment to expedite null modeling.
 

### 2021-06-09 Dryad_ms1/fl_trait_proc.R

@@ -1,131 +1,0 @@
-# Packages
-library(tidyverse)
-
-# Floral trait data
-floral_k_type <-  read_delim("./data/raw_data/BioFlor_Kugler_classification.csv", delim = ";") %>%
-  dplyr::select(taxon = 1, k.type = 2) %>% 
-  na.omit() %>%
-  mutate(taxon = str_replace_all(taxon, c(" x " = " x"))) %>%
-  separate(taxon, c("genus", "species"), sep = " ") %>%
-  unite(plant.sp, genus, species, sep = " ", remove = FALSE) %>%
-  select(plant.sp, plant.genus = genus, k.type) %>%
-  mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"), # make simplified (and more simplified) k types
-         k.type.ss = str_extract(k.type, "[-+]?[0-9]*")) %>%
-  
-  # Fill in ktype data for plants present in my samples but missing from the Bioflor database
-  add_row(plant.sp = "Phyteuma orbiculare", plant.genus = "Phyteuma", 
-          k.type = "NA", k.type.s = "7.1", k.type.ss = "7") %>% # Following Neumayer and Paulus (1999)
-  add_row(plant.sp = "Phyteuma spicatum", plant.genus = "Phyteuma",
-          k.type = "NA", k.type.s = "7.1", k.type.ss = "7") %>% # Following Neumayer and Paulus (1999)
-  add_row(plant.sp = "Aconitum vulparia", plant.genus = "Aconitum",
-          k.type = "NA", k.type.s = "5.2", k.type.ss = "5") %>% # All scored Aconitum were classified as 5.2
-  add_row(plant.sp = "Cardamine enneaphyllos", plant.genus = "Cardamine", 
-          k.type = "NA", k.type.s = "1.2", k.type.ss = "1") %>% # All scored Cardamine were classified as 1.2
-  add_row(plant.sp = "Euphrasia picta", plant.genus = "Euphrasia", 
-          k.type = "NA", k.type.s = "5.1", k.type.ss = "5") %>% # All scored Euphrasia were classified as 5.1
-  add_row(plant.sp = "Euphrasia rostkoviana", plant.genus = "Euphrasia", 
-          k.type = "NA", k.type.s = "5.1", k.type.ss = "5") %>% # All scored Euphrasia were classified as 5.1
-  add_row(plant.sp = "Lamium galeobdolon", plant.genus = "Lamium", 
[…]

### 2021-06-10 code/ms2_nullmodels.Rmd
finished intro and almost done methods and results

@@
 ```{r}
-ggplot(null_tab1, aes(ktype, preference)) +
+null_model_overall<- ggplot(null_tab1, aes(ktype, preference)) +
   #geom_violin(fill = "gray70", color = "white", draw_quantiles = c(0.25, 0.5, 0.75)) +
@@
   geom_jitter(aes(color = Test), width = 0.3, alpha = 0.6) +
-  theme_light(14) +
+  theme_light(10) +
   scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
   labs(x = "Floral morphotype",
-       y = "Bias")
-
-ggsave("../ms_2/figures/null_model_ktype_preference.pdf", height = 6, width = 7)
-ggsave("../ms_2/figures/null_model_ktype_preference.png", height = 6, width = 7)
+       y = "Selectivity")
+
+ggsave("../ms_2/figures/null_model_ktype_preference.pdf", null_model_overall, height = 6, width = 7)
+ggsave("../ms_2/figures/null_model_ktype_preference.png", null_model_overall, height = 6, width = 7)
 ```
@@
 
-### 2.3.2 Model
+### 2.3.2 Call model 
 ```{r}
@@
 
-summary(gam1)
+gam1_sum <- summary(gam1)
 check.gamViz(gam1)
 #plot.gamViz(gam1, allTerms = TRUE)
+```
+
+### 2.3.3 Plot
+```{r}
+# Tabulate model summary
+gam1_s_table <- gam1_sum$s.table %>%
+  as.data.frame() %>%
+  rownames_to_column("term") %>%
+  mutate(ktype = str_extract(term, "(?<=ktype).*$")) %>%
+  mutate(pval = signif(`p-value`, digits = 3)) %>%
+  select(ktype, edf, pval) %>%
+  as_tibble() %>%
+  mutate(ktype = factor(ktype))
 
@@
   mutate(fit.plus.se = fit + (2*se.fit),
-         fit.minus.se = fit - (2*se.fit))
+         fit.minus.se = fit - (2*se.fit)) %>%
+  left_join(gam1_s_table) %>%
+  mutate(sig = if_else(pval < 0.05, TRUE, FALSE)) %>%
+  mutate(sig = factor(sig, levels = c(TRUE, FALSE)))
 
 # Plot
-ggplot(gam1_pred, aes(pbl.w, fit)) +
-  geom_line() + # fit line
+null_model_pbl_GAM <- ggplot(gam1_pred, aes(pbl.w, fit)) +
+  geom_line(aes(linetype = sig)) + # fit line, linetype for significance
   geom_ribbon(aes(ymin = fit.minus.se, ymax = fit.plus.se), # Uncertainty
@@
              inherit.aes = FALSE, alpha = 0.5) +
-  #scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
   facet_wrap(~ktype) +
-  labs(x = "Tongue length (mm)", y = "Bias", 
-       color = "Species") +
-  theme_light()
-
-ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference.pdf", height = 6, width = 7)
-ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference.png", height = 6, width = 7)
+  labs(x = "Tongue length (mm)", y = "Selectivity", 
+       color = "Species", linetype = "p < 0.05") +
+  theme_light(10)
+
+ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference.pdf", null_model_pbl_GAM, height = 6, width = 7)
+ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference.png", null_model_pbl_GAM, height = 6, width = 7)
 ```
@@
 
-#check.gamViz(gam3.1)
-#gam3.1_sum <- summary(gam3.1, re.test = FALSE)
-#plot.gamViz(gam3.1, allTerms = TRUE)
+check.gamViz(gam3.1)
+gam3.1_sum <- summary(gam3.1, re.test = FALSE)
+plot.gamViz(gam3.1, allTerms = TRUE)
+
+gam3.1b <- bam(preference ~
+                bb.ktype +
+                s(site, by = bb.ktype, bs = "re") +
+                s(year, by = bb.ktype, bs = "re") +
+                te(yday, elev.mean, by = bb.ktype),
+              family = "gaussian",
+              method = "fREML",
+              discrete = TRUE,
+              data = filter(null_tab3_join, 
+                            bb.sp %in% c("hort", "prat", "telu", 
+                                         "wurf", "soro", "pasc") &
+                              ktype %in% c("bell", "lip", "flag", "head") &
+                              prop.cover < 0.5)) %>% 
+                getViz()
+
+check.gamViz(gam3.1b)
+gam3.1b_sum <- summary(gam3.1b, re.test = FALSE)
+plot.gamViz(gam3.1b, allTerms = TRUE)
 ```
@@
 ggsave("../ms_2/figures/preference_lability_gam.png", height = 5, width = 6.5)
+
+# And now having dropped the high abundance outliers
+gam3.1b_s_table <- gam3.1b_sum$s.table %>%
+  as.data.frame() %>%
+  rownames_to_column("term") %>%
+  filter(str_detect(term, "te\\(")) %>%
+  mutate(term = str_extract(term, ".{6}$")) %>%
+  separate(term, c("bb.sp", "ktype")) %>%
+  mutate(ktype = case_when(
+    ktype == 3 ~ "bell",
+    ktype == 5 ~ "lip",
+    ktype == 6 ~ "flag",
+    ktype == 7 ~ "head"
+  )) %>%
+  mutate(pval = signif(`p-value`, digits = 3)) %>%
+  select(bb.sp, ktype, edf, pval) %>%
+  as_tibble() %>%
+  mutate(ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")))
+
+# Extract plotting data.
+hort_3_gam3.1b <- plot(sm(gam3.1b, 49))$ggObj[[1]] %>%
+  mutate(bb.sp = rep("hort", n()),
+         ktype = rep("bell", n()))
+
+hort_5_gam3.1b <- plot(sm(gam3.1b, 50))$ggObj[[1]] %>%
+  mutate(bb.sp = rep("hort", n()),
+         ktype = rep("lip", n()))
+
+hort_6_gam3.1b <- plot(sm(gam3.1b, 51))$ggObj[[1]] %>%
+  mutate(bb.sp = rep("hort", n()),
+         ktype = rep("flag", n()))
+
+hort_7_gam3.1b <- plot(sm(gam3.1b, 52))$ggObj[[1]] %>%
+  mutate(bb.sp = rep("hort", n()),
+         ktype = rep("head", n()))
+
+pasc_3_gam3.1b <- plot(sm(gam3.1b, 53))$ggObj[[1]] %>%
+  mutate(bb.sp = rep("pasc", n()),
+         ktype = rep("bell", n()))
+
+pasc_5_gam3.1b <- plot(sm(gam3.1b, 54))$ggObj[[1]] %>%
+  mutate(bb.sp = rep("pasc", n()),
+         ktype = rep("lip", n()))
+
+pasc_6_gam3.1b <- plot(sm(gam3.1b, 55))$ggObj[[1]] %>%
+  mutate(bb.sp = rep("pasc", n()),
+         ktype = rep("flag", n()))
+
+pasc_7_gam3.1b <- plot(sm(gam3.1b, 56))$ggObj[[1]] %>%
+  mutate(bb.sp = rep("pasc", n()),
+         ktype = rep("head", n()))
+
+prat_3_gam3.1b <- plot(sm(gam3.1b, 57))$ggObj[[1]] %>%
+  mutate(bb.sp = rep("prat", n()),
+         ktype = rep("bell", n()))
+
+prat_5_gam3.1b <- plot(sm(gam3.1b, 58))$ggObj[[1]] %>%
+  mutate(bb.sp = rep("prat", n()),
+         ktype = rep("lip", n()))
+
+prat_6_gam3.1b <- plot(sm(gam3.1b, 59))$ggObj[[1]] %>%
+  mutate(bb.sp = rep("prat", n()),
+         ktype = rep("flag", n()))
+
+prat_7_gam3.1b <- plot(sm(gam3.1b, 60))$ggObj[[1]] %>%
+  mutate(bb.sp = rep("prat", n()),
+         ktype = rep("head", n()))
+
+soro_3_gam3.1b <- plot(sm(gam3.1b, 61))$ggObj[[1]] %>%
+  mutate(bb.sp = rep("soro", n()),
+         ktype = rep("bell", n()))
+
+soro_5_gam3.1b <- plot(sm(gam3.1b, 62))$ggObj[[1]] %>%
+  mutate(bb.sp = rep("soro", n()),
+         ktype = rep("lip", n()))
+
+soro_6_gam3.1b <- plot(sm(gam3.1b, 63))$ggObj[[1]] %>%
+  mutate(bb.sp = rep("soro", n()),
+         ktype = rep("flag", n()))
+
+soro_7_gam3.1b <- plot(sm(gam3.1b, 64))$ggObj[[1]] %>%
+  mutate(bb.sp = rep("soro", n()),
+         ktype = rep("head", n()))
+
+telu_3_gam3.1b <- plot(sm(gam3.1b, 65))$ggObj[[1]] %>%
+  mutate(bb.sp = rep("telu", n()),
+         ktype = rep("bell", n()))
+
+telu_5_gam3.1b <- plot(sm(gam3.1b, 66))$ggObj[[1]] %>%
+  mutate(bb.sp = rep("telu", n()),
+         ktype = rep("lip", n()))
+
+telu_6_gam3.1b <- plot(sm(gam3.1b, 67))$ggObj[[1]] %>%
+  mutate(bb.sp = rep("telu", n()),
+         ktype = rep("flag", n()))
+
+telu_7_gam3.1b <- plot(sm(gam3.1b, 68))$ggObj[[1]] %>%
+  mutate(bb.sp = rep("telu", n()),
+         ktype = rep("head", n()))
+
+wurf_3_gam3.1b <- plot(sm(gam3.1b, 69))$ggObj[[1]] %>%
+  mutate(bb.sp = rep("wurf", n()),
+         ktype = rep("bell", n()))
+
+wurf_5_gam3.1b <- plot(sm(gam3.1b, 70))$ggObj[[1]] %>%
+  mutate(bb.sp = rep("wurf", n()),
+         ktype = rep("lip", n()))
+
+wurf_6_gam3.1b <- plot(sm(gam3.1b, 71))$ggObj[[1]] %>%
+  mutate(bb.sp = rep("wurf", n()),
+         ktype = rep("flag", n()))
+
+wurf_7_gam3.1b <- plot(sm(gam3.1b, 72))$ggObj[[1]] %>%
+  mutate(bb.sp = rep("wurf", n()),
+         ktype = rep("head", n()))
+
+plotting_data__gam3.1b <- bind_rows(hort_3_gam3.1b, hort_5_gam3.1b, hort_6_gam3.1b, hort_7_gam3.1b,
+                           pasc_3_gam3.1b, pasc_5_gam3.1b, pasc_6_gam3.1b, pasc_7_gam3.1b,
+                           prat_3_gam3.1b, prat_5_gam3.1b, prat_6_gam3.1b, prat_7_gam3.1b,
+                           soro_3_gam3.1b, soro_5_gam3.1b, soro_6_gam3.1b, soro_7_gam3.1b,
+                           telu_3_gam3.1b, telu_5_gam3.1b, telu_6_gam3.1b, telu_7_gam3.1b,
+                           wurf_3_gam3.1b, wurf_5_gam3.1b, wurf_6_gam3.1b, wurf_7_gam3.1b) %>%
+  group_by(bb.sp, ktype) %>%
+  mutate(z.norm = scale(z, center = FALSE),
+         ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")),
+         date = as_date(x)) %>%
+  ungroup() %>%
+  left_join(bb_traits) %>%
+  left_join(gam3.1b_s_table) %>%
+  select(z, z.norm, date, x, elevation = y, se, 
+         bb.sp, ktype, pbl.w, edf, pval) %>%
+  mutate(sig = if_else(pval < 0.05, TRUE, FALSE))
+
+ggplot(plotting_data__gam3.1b, aes(date, elevation, z = z.norm, fill = z.norm)) +
+  geom_raster() +
+  geom_contour(color = "black",
+               size = 0.1,
+               bins = 40,
+               aes(linetype = sig)) + # by fixing bins, we scale contours to each plot while letting the color ramp represent differences in scale across plots; I think this is a good compromise
+  facet_grid(reorder(bb.sp, -pbl.w) ~ ktype) +
+  scale_linetype_manual(values = c("dotted", "solid"))+
+  scale_fill_viridis_c(na.value = "transparent") +
+  labs(x = NULL, y = "Elevation (m)", fill = "Normalized\neffect", linetype = "p < 0.05") +
+  theme_light(12) +
+  theme(panel.grid = element_blank(),
+        axis.text.x = element_text(angle = 90, vjust = 0.5))
+
+ggsave("../ms_2/figures/preference_lability_gam.pdf", height = 5, width = 6.5)
+ggsave("../ms_2/figures/preference_lability_gam.png", height = 5, width = 6.5)
 ```
@@
                            "wurf", "soro", "pasc") &
-                ktype %in% c("bell", "lip", "flag", "head")), aes(cover, preference)) +
+                ktype %in% c("bell", "lip", "flag", "head") &
+                prop.cover < 0.5), aes(prop.cover, preference)) +
   geom_point(alpha = 0.2) +
   geom_smooth(method = "lm") +
-  facet_grid(bb.sp~ktype, scales = "free") +
+  facet_grid(bb.sp~ktype) +
   ylim(c(-1, 1))
+
+ggplot(filter(abundance, ktype %in% c("bell", "lip", "flag", "head")), 
+       aes(prop.cover)) +
+  geom_histogram() +
+  facet_wrap(~ktype)
+
+ggplot(abundance, aes(prop.cover)) +
+  geom_histogram() +
+  facet_wrap(~ktype)
 
@@
   droplevels() %>%
-  left_join(bb_traits)
-
-null_model_overall <- ggplot(null_tab1, aes(ktype, preference)) +
-  #geom_violin(fill = "gray70", color = "white", draw_quantiles = c(0.25, 0.5, 0.75)) +
-  geom_boxplot(fill = NA, color = "black", outlier.shape = 4, size = 0.35) +
-  geom_jitter(aes(color = Test), width = 0.3, alpha = 0.6) +
-  theme_light(10) +
-  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
-  labs(x = "Floral morphotype",
-       y = "Bias")
+  left_join(bb_traits) %>%
+  ungroup() %>%
+  mutate(bb.sp = fct_reorder(factor(bb.sp), pbl.w))
+
+# null_model_overall <- ggplot(null_tab1, aes(ktype, preference)) +
+#   #geom_violin(fill = "gray70", color = "white", draw_quantiles = c(0.25, 0.5, 0.75)) +
+#   geom_boxplot(fill = NA, color = "black", outlier.shape = 4, size = 0.35) +
+#   geom_jitter(aes(color = Test), width = 0.3, alpha = 0.6) +
+#   theme_light(10) +
+#   scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
+#   labs(x = "Floral morphotype",
+#        y = "Bias")
 
@@
 
-summary(gam1b)
+gam1b_sum <- summary(gam1b)
 check.gamViz(gam1b)
@@
 
+# Tabulate model summary
+gam1b_s_table <- gam1b_sum$s.table %>%
+  as.data.frame() %>%
+  rownames_to_column("term") %>%
+  mutate(ktype = str_extract(term, "(?<=ktype).*$")) %>%
+  mutate(pval = signif(`p-value`, digits = 3)) %>%
+  select(ktype, edf, pval) %>%
+  as_tibble() %>%
+  mutate(ktype = factor(ktype))
+
+# Extract predictions
 gam1b_pred <- predict_gam(gam1b) %>%
@@
          fit.minus.se = inv_logit(fit - (2*se.fit)),
-         fit = inv_logit(fit))
-
+         fit = inv_logit(fit)) %>%
+  left_join(gam1b_s_table) %>%
+  mutate(sig = if_else(pval < 0.05, TRUE, FALSE)) %>%
+  mutate(sig = factor(sig, levels = c(TRUE, FALSE)))
+
+# Plot
 raw_visits_pbl_GAM <- ggplot(gam1b_pred, aes(pbl.w, fit)) +
-  geom_line() + # fit line
+  geom_line(aes(linetype = sig)) + # fit line, linetype for significance
   geom_ribbon(aes(ymin = fit.minus.se, ymax = fit.plus.se), # Uncertainty
@@
   geom_point(data = visitation_prop, # Original data
-             aes(pbl.w, mean.prop.visits, color = reorder(bb.sp, pbl.w)), 
+             aes(pbl.w, mean.prop.visits, color = bb.sp), 
              inherit.aes = FALSE, alpha = 0.5) +
@@
   labs(x = "Tongue length (mm)", y = "Mean proportional visitation", 
-       color = "Species") +
-  scale_x_continuous(labels = scales::number_format(accuracy = 1)) +
-  scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
-  theme_light(10)
-
-null_model_pbl_GAM <- ggplot(gam1_pred, aes(pbl.w, fit)) +
-  geom_line() + # fit line
-  geom_ribbon(aes(ymin = fit.minus.se, ymax = fit.plus.se), # Uncertainty
-              fill = "black", alpha = 0.2) +
-  geom_point(data = null_tab1, # Original data
-             aes(pbl.w, preference, color = bb.sp), 
-             inherit.aes = FALSE, alpha = 0.5) +
-  #scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
-  facet_wrap(~ktype) +
-  labs(x = "Tongue length (mm)", y = "Bias", 
-       color = "Species") +
-  scale_x_continuous(labels = scales::number_format(accuracy = 1)) +
-  scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
+       color = "Species", linetype = "p < 0.05") +
   theme_light(10)
@@
 
-ggsave("../ms_2/figures/ktypeXtonguelength_raw_visitation.pdf", pbl_GAM_plot, height = 5, width = 8.5)
-ggsave("../ms_2/figures/ktypeXtonguelength_raw_visitation.png", pbl_GAM_plot, height = 5, width =8.5)
+ggsave("../ms_2/figures/ktypeXtonguelength_paired.pdf", pbl_GAM_plot, height = 5, width = 9.5)
+ggsave("../ms_2/figures/ktypeXtonguelength_paired.png", pbl_GAM_plot, height = 5, width = 9.5)
 ```
@@
 # Ordisurf on tongue length
-ordisurf_visits <- ordisurf(pca_visits, tongues2$pbl.w) 
+ordisurf_visits <- ordisurf(pcoa_visits, tongues2$pbl.w) 
 summary(ordisurf_visits)

### 2021-06-10 ms_2/ms_2.Rmd
finished intro and almost done methods and results

@@
 The process of resource selection by animals has been an enduring theme of ecological research [@Johnson1980-co; @Beyer2010-vr].
-Methodological difficulties notwithstanding, the study of animals' resource choices is one of the central tasks of community ecology, where the selection and partitioning of resources are instrumental in community assembly, species coexistence, and ecosystem function [@Chesson2000-aw].
+Methodological difficulties notwithstanding, the study of animals' resource choices is one of the central tasks of community ecology, where the selection and partitioning of resources are instrumental in community assembly, species coexistence, and ecosystem function [@Chesson2000-aw; @Bascompte2007-ls].
 Recently, questions of resource selection in ecological communities have taken on new urgency on account of anthropogenic climate and land use change, which threaten to disrupt existing patterns of species interactions and niche relationships [@Tylianakis2008-fi].
@@
 Bumble bees (Hymentoptera: *Bombus* spp.) and their floral hosts provide a diverse yet tractable system in which to explore resource selection.
-Bumble bees are exceptionally versatile foragers whose ability to solve the puzzles posed by complex flower morphologies [@Plowright1984-lc] --- or, indeed, by human investigators [@Alem2016-rn; @Loukola2017-bt] --- renders almost any flower potentially accessible.
+Bumble bees are exceptionally versatile florivores whose ability to solve the puzzles posed by complex flower morphologies [@Plowright1984-lc] --- or, indeed, by human investigators [@Alem2016-rn; @Loukola2017-bt] --- renders almost any flower potentially accessible.
 Nevertheless, the flexibility of bumble bee foraging is tuned to floral reward, which depends on both the intrinsic reward of a given flower and the costs of exploiting it.
@@
 @Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn].
-This association between tongue length and floral niche is imperfect, though, and there can be considerable overlap in floral resource use between bumble bees of different tongue lengths [e.g.
+This association between tongue length and floral niche is imperfect, though, and there can be both convergence in floral resource use between bumble bees of different tongue length and divergence in floral resource use between bumble bees of similar tongue length [e.g.
 @Ranta1980-ok; @Goulson2008-oe].
@@
 
-Understanding resource selection in bumble bee communities is more than just an exercise in ecological theory.
+Floral choice in bumble bee communities can serve as a case study in the general phenomenon of animal resource selection, but it is also a matter of practical importance.
 Bumble bees are sensitive to climate and land use change [@Goulson2008-oe; @Soroye2020-zb], especially in the mountain ecosystems that host the world's richest bumble bee communities.
@@
 In contrast, more accessible areas at the bases of mountains are undergoing agricultural intensification that displaces natural vegetation and favors grasses over forbs [@Tasser2002-om].
-Caught between adverse land use changes above and below and subject to climate-induced range shifts, mountain bumble bees are facing a future of reduced floral resource availability and altered sympatry, and the ability of bumble species to coexist under these changes will likely depend significantly on how they select and partition floral resources.
+Caught between adverse land use changes above and below and subject to climate-induced range shifts, mountain bumble bees are facing a future of reduced floral resource availability and altered sympatry.
+The ability of bumble bee species to coexist under these changes will likely depend significantly on how they select and partition floral resources, and analogous considerations apply to their floral mutualists, many of which depend principally on bumble bee visitation for pollination service [@Goulson2010-lo].
 
@@
 Using these data in conjunction with complementary floral survey data, we explore the *selectivity* of the bumble bee community at the level of floral morphotype [@Kugler1970-gb].
+Following @Beyer2010-vr, we define *selectivity* simply as a disproportionality --- positive or negative --- between the relative abundance of a resource and its rate of use.
+This concept approaches the idea of resource *preference*, but inference of the latter, in an absolute sense, requires either experimental control of all resource abundances [@Johnson1980-co] or statistical techniques that condition visitation simultaneously on all resource abundances [@Aarts2008-fo].
+Given the scale and complexity of our observational study, we opt for simpler analytical approaches and restrict our inference to the phenomenon of selectivity, which can be understood as a metric of *relative* preference for a focal resource given unmodeled interactions with preferences for alternative resources.
+We deem these interactions beyond the scope of our study, but acknowledge that this requires us to remain partly agnostic as to the process underlying the empirical selectivity we describe.
 
-**This is a good place to get our terminological ducks in a row**
-
-We opt for the term *selectivity*, defined as the  (*sensu* @Beyer2010-vr) rather than the ...
-defined in this way, selectivity has the same conceptual meaning (though different mathematical derivation) as @Bluthgen2006-if *discrimination*.
-
-We first compare the aggregate preferences of the bumble bee community across floral morphotypes and then model the preferences for each morphotype as a function of bumble bee tongue length.
-Then, by calculating preferences based on individual samples distributed through elevation and time, we model the variation of bumble bee preferences through temporal and elevational axes of floristic turnover, revealing the lability of bumble bee floral preferences.
-The balance of flexibility and constraint revealed in these analyses **Figure out how to finish this...**
+We first compare the aggregate selectivity of the bumble bee community across floral morphotypes and then model the species-wise selectivity for each morphotype as a function of bumble bee tongue length.
+Then, by calculating selectivity based on individual samples distributed through elevation and time, we model the variation of bumble bee selectivity through temporal and elevational axes of floristic turnover, revealing the lability of bumble bee floral selectivity.
+Together, these analyses reveal which floral morphotypes receive the bulk of bumble bee visitation, how bumble bee species converge and diverge with respect to resource selection, and the temporal and elevation patterns of floral selectivity can be expected to be acted upon by the phenological and distributional distortions caused by climate and land use change.
 
@@
 In studies of resource selection, it is common to pose an abundance-only hypothesis in which foraging is random and selection merely a function of frequency of encounter, resulting in resource use patterns that mirror resource abundance patterns.
-For bumble bees, which are known to forage non-randomly, the abundance-only hypothesis can serve as a null-model against which observed use patterns can be compared to infer *selectivity*, i.e.
-resource use that is either positively or negatively disproportionate to resource abundance.
+For bumble bees, which are known to forage non-randomly, the abundance-only hypothesis can serve as a null-model against which observed use patterns can be compared to infer selectivity (often called "preference", following a *sensu lato* usage of the term) [@Vaughan2018-gj].
 It has also be proposed, however --- in what has been termed the *target use hypothesis* --- that bumble bees might select resources irrespective of abundance [@De_Valpine2013-be; @Harmon-Threatt2017-bz].
-This hypothesis, while obviously subject to boundary conditions of floral availability, is consistent with the observation that bumble bees sometimes persist in visiting favored flower species even when these resources are experimentally reduced to low abundance [@Heinrich1976-vb], and it was supported in a study comparing competing models of flower selection in *Bombus vosnesenskii* [@Harmon-Threatt2017-bz].
-If and when the *target use hypothesis* holds, uncorrected visitation rates are a better indication of preference than the residuals of an abundance-only null model, since the subtraction of presumed neutral effects of abundance when those effects do not, in fact, exist actually creates new abundance-dependent artifacts and obscures the true process of resource selection [@Harmon-Threatt2017-bz].
+This hypothesis, while obviously subject to boundary conditions of floral availability, is consistent with the anecdotal observation that bumble bees sometimes persist in visiting favored flower species even when these resources are experimentally reduced to low abundance [@Heinrich1976-vb], and it was supported in a study comparing competing models of flower selection in *Bombus vosnesenskii* [@Harmon-Threatt2017-bz].
+If and when the *target use hypothesis* holds, uncorrected visitation rates are a better index of preference (*sensu stricto*) than selectivity (i.e.
+the residuals of an abundance-only null model), since the subtraction of presumed neutral effects of abundance when those effects do not, in fact, exist actually creates new abundance-dependent artifacts and obscures the true process of resource selection [@Harmon-Threatt2017-bz].
 On the other hand, failing to account for neutral effects of abundance when they do obtain similarly obscures the selection process underpinning foraging patterns [@Vazquez2003-qa; @Bluthgen2008-rp].
-In our study, we take advantage of the fact that there exists are well-established mechanistic basis for bumble bee flower choice, namely the compatibility between floral morphology and bumble bee tongue length.
-Taking this mechanism as a reference, we evaluate models of preference with and without abundance effects by comparing the explanatory power of bumble bee tongue length with respect to the preferences inferred under each model.
-
-*"[. . .] preference  is  the  ratio  of  the  use  of  a  habitat  over  its availability, conditional on the availability of all habitats to the study animals." [@Aarts2008-fo]*
-
-One way we could handle our language and caveats is this.
-The kind of null model approach we used *does not* condition visitation to each focal resource on the abundances of all other resources, only on the proportional abundance of the focal resource.
-Following @Aarts2008-fo, we cannot then speak of *preference*.
-Instead, what we have is weaker concept of *selectivity* defined by Beyer2010-vr.
-We can, therefore, use the word *selectivity* after explaining its relation to preference in the strict sense.
-I do not want to go back and fit some crazy hierarchical model that conditions on the abundance of all morphotypes.
+In our study, we take advantage of the fact that there exists a well-established mechanistic basis for bumble bee flower choice, namely the compatibility between floral morphology and bumble bee tongue length.
+Taking this mechanism as a reference and exploiting the tongue-length diversity of the bumble bee species in our study system, we evaluate models of preference with and without abundance effects by comparing the explanatory power of bumble bee tongue length with respect to the preferences inferred under each model
 
@@
 
+The samples analyzed in this paper have been previously analyzed (Sponsler et al., submitted), but the analyses presented here are non-overlapping.
 Field sampling was conducted 2010-2012 in Berchtesgaden National Park (47.55°N, 12.92°E), located in the Northern Limestone Alps of southeast Germany.
-The landscape is composed of mountain pastures in a coniferous forest matrix.
-We selected 25 study sites (60 x 60 m) on pastures at elevations ranging from 641-2032 m above sea level.
-Fourteen of these pastures are extensively grazed by cattle or sheep, three are mowed for hay production, and eight have been abandoned throughout the last century and are no longer subject to any human management.
+The landscape is composed of a coniferous forest matrix embedded with pastures either currently or historically maintained by extensive (i.e.
+low intensity) agricultural management.
+We selected a total of 25 study sites (60 x 60 m) located on pastures along a ~1400 m elevation gradient ranging from 641-2032 m above sea level (m.s.l.).
+Fourteen of these pastures are grazed by cattle or sheep, three are mowed for hay production, and eight have been abandoned throughout the last century and are no longer subject to any human management.
 
-Sampling consisted of repeated visits (hereafter, “visitation samples”) to each study site at approximately weekly intervals.
+Sampling of bumble- bee-wildflower interactions consisted of repeated visits (hereafter, “visitation samples”) to each study site at approximately weekly intervals.
 Sampling was only conducted during periods without rain and with an air temperature of at least 6°C.
 Bumble bees were recorded during one 50-minute transect walk per visit, and bumblebees observed on a given flower were counted as floral visitors.
-Queens were identified to species level in the field to avoid destructively sampling colony foundresses, but workers and males were collected and frozen for later identification in the laboratory.
-Species were identified following @Amiet1996-jz.
-While male bumble bees was recorded, we chose to analyze only the pooled visitation of by queens and workers.
-In conjunction with bumble bee observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.01 m^2^ and identified the plant species according to @Lauber2007-ha and @Oberndorfer2001-ro.
-
+Queens were identified to species level in the field to avoid destructively sampling colony foundresses, but workers and males were collected and identified in the lab following @Amiet1996-jz.
+While male bumble bees were recorded, we chose to analyze only the pooled visitation of queens and workers.
+In conjunction with bumble bee observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.01 m^2^ (hereafter "floral survey samples") and identified the plant species according to @Lauber2007-ha and @Oberndorfer2001-ro.
 While per-sample effort was standardized across all years, the total number of samples varied: 153 in 2010, 212 in 2011, and 539 in 2012.
@@
 In 2010 and 2012, this resulted in a temporally staggered pattern of sampling, but a combination of early snowmelt and a late start to field activities in 2011 resulted in sampling that was approximately aligned in time across the whole elevation gradient.
-Due to these annual differences in sampling intensity and duration, we opted to pool data across years rather than include year effects in our models, which would multiply model complexity and be difficult to interpret.
-This means, however, that our analysis cannot detect year-to-year variation in floral preferences and niche overlap, and main effects cannot be distinguished from annual anomalies.
-With only three years of data, though, such annual inferences would be of little value, and we deemed them not worth the cost of added model complexity.
-
-The samples analyzed in this paper have been previously analyzed (Sponsler et al., submitted), but the analyses presented here are non-overlapping.
+See Appendix X for a full description of sampling patterns.
 
@@
 
-**Start this by working out my terminology.
-Are we saying preference or bias?
-And what about the distinction between preference, use, and selection?
-See @Johnson1980-co and @Beyer2010-vr as authorities on terminology.
-Whatever I choose, I then need to go revise the labels of the figures accordingly.
-I think we could make an argument for sticking wit the word "bias" to describe the inferences of our null model analysis, since we are allowing that the raw visitation patterns may be as good or better a measure of "preference".
-We don't actually find that, of course, but I think we should avoid calling the null model inferences "preference" as though it were a foregone conclusion that they represent preference while the uncorrected visitation rates do not.
-But preference, alas, is the word the litertaure seems to have settled on, so when I discuss my conclusions, I should talk about preference, comparing preference as inferred from bias and preference as inferred from uncorrected visitation rates.**
-
-*We will use the word "preference" in the phenomenological sense of @Beyer2010-vr to denote a disproportionality between resource availability and resource use while remaining agnostic as to the behavioral basis of this pattern.
-Or perhaps we will use the word bias but follow the same definition, since "bias" does not bear the teleological connotations of "preference".
-Double check that this is really how @Beyer2010-vr define preference.
-If so, it differs a bit from the more original definition of .*
-
-A key challenge when interpreting interaction networks is to disentangle directed processes, such as a focal species' preference or avoidance of certain interaction partners, from the neutral effects of relative abundance [@Vazquez2003-qa; @Bluthgen2008-rp].
-For example, in the context of a bumble bee floral visitation network such as ours, abundant floral taxa should be expected to receive more visitation than rare taxa simply because they are more frequently encountered, and it would clearly be a mistake to interpret this as meaningful "preference" or "specialization" on the part of the bumble bees.
-One way to control for the neutral effects of relative abundance is to compare observed visitation patterns to the simulated visitation patterns of a null model in which plants are visited randomly in proportion to their relative abundance [@Vazquez2003-qa; @Bluthgen2008-rp].
-When constructing null models of interaction networks that are distributed through space and/or time, though, it is critical to respect the sampling unit of individual sites and dates, since only within this site-date sampling unit does floral visitation data correspond unequivocally to floral abundance data.
+To quantify selectivity, we compared the proportional visitation of each morphotype (by each bumble bee species) with the rate of visitation expected under the null-hypothesis that visitation of a given resources is purely a function of its relative abundance [@Vazquez2003-qa; @Bluthgen2008-rp].
+When conducting a null model analysis of interaction networks that are distributed through space and/or time, it is critical to respect the sampling unit of individual sites and dates, since only within this site-date sampling unit does floral visitation data correspond unequivocally to floral abundance data.
 Naive aggregation across sites and dates introduces forbidden links and decouples visitation and abundance [@Vaughan2018-gj].
-
 We constructed null models following @Vaughan2018-gj.
-First, resource selection is simulated for individual consumers --- in our case, individual bumble bees --- by according to a resource selection process in which the probability of a resource being chosen is proportional to its relative abundance [@Agusti2003-tt].
+First, resource selection is simulated for individual consumers --- in our case, individual bumble bees --- according to a resource selection process in which the probability of a resource being chosen is proportional to its relative abundance [@Agusti2003-tt].
 By simulating the visitation of individual consumers rather than species in aggregate, the abundance of each resource, and therefore its probability of visitation, can vary across site-dates, thus preserving the within-sample correspondence between visitation and abundance.
@@
 As this process is iterated, the result is a distribution of visitation rates of each consumer species to each resource under the null assumption that resource selection is determined only by resource abundance.
-The visitation distributions of the null model can then be compared to observed visitation rates to infer positive or negative bias of each consumer toward each resource.
-Since the mean of the null distribution can be interpreted as the expected value under null assumptions, the difference between the mean null value and the observed value can be interpreted as a measure of positive or negative bias [@Vaughan2018-gj].
-The statistical significance of this bias can be estimated as the proportion of simulated null values more extreme than the observed value.
+The visitation distributions of the null model can then be compared to observed visitation rates to infer positive or negative selectivity of each consumer toward each resource.
+Since the mean of the null distribution can be interpreted as the expected value under null assumptions, the difference between the mean null value and the observed value can be interpreted as a magnitude of selectivity [@Vaughan2018-gj].
+The statistical significance of this selectivity can be estimated as the proportion of simulated null values more extreme than the observed value.
 
 We performed two rounds of null model analysis targeting different ecological questions.
-First, we followed the procedure described above, simulating our entire dataset in one modeling process to infer aggregate pattern of bias between bumble bee species and floral morphotypes.
-Then, we repeated the null model analysis for each site-date individually so that the bias between each bumble bee species and floral morphotype could be modeled as a variable through elevation and time.
-For both analyses, we used 500 iterations to achieve a robust sampling distribution, and we considered observed visitation rates to indicate significant bias if they fell outside the interval containing 99% of simulated values.
+First, we followed the procedure described above, simulating our entire dataset in one modeling process to infer aggregate pattern of selectivity between bumble bee species and floral morphotypes.
+Then, we repeated the null model analysis for each site-date individually so that the selectivity of each bumble bee species toward each floral morphotype could be modeled as a variable through elevation and time.
+For both analyses, we used 500 iterations to achieve a robust sampling distribution, and we considered observed visitation rates to indicate significant selectivity if they fell outside the interval containing 99% of simulated values.
 
-Using the aggregate null model results, we modeled the biases toward each floral morphotype as a function of bumble bee tongue length using hierarchical generalized additive modeling (HGAM) with floral morphotype as a "by" smoothing factor [@Pedersen2019-kf].
-We then compared morphotype biases across bumble bee species by performing a principal components analysis and mapped bumble bee tongue length onto the ordination space by fitting a 2-dimensional GAM surface to the first two principal components.
+Using the aggregate null model results, we modeled selectivity toward each floral morphotype as a function of bumble bee tongue length using hierarchical generalized additive modeling (HGAM), with floral morphotype specified as a smoothing factor [@Pedersen2019-kf].
+We then compared morphotype selectivity across bumble bee species by performing a principal components analysis, and we mapped bumble bee tongue length onto the ordination space by fitting a 2-dimensional GAM surface to the first two principal components.
 
-With the sample-wise null model results, we modeled the variation of bias between bumble bee species and floral morphotypes through elevation and time.
+With the sample-wise null model results, we modeled the variation of selectivity between bumble bee species and floral morphotypes through elevation and time.
 For this analysis, we included only the six most abundant and widespread bumble bee species in our study system (*B.
@@
 terrestris-lucorum*, and *B.
-wurflenii*) to ensure sufficient data points through elevation and time, and we modeled their biases only toward the 4 most visited floral morphotypes (bell, lip, flag, and head flowers).
-Biases for each bumble- bee-morphotype pair were modeled together in a single HGAM with a tensor-product interaction smooth between elevation and time and bumble- bee-morphotype pair as a "by" smoothing factor.
+wurflenii*) to ensure sufficient data points through elevation and time, and we modeled their selectivity only toward the 4 most visited floral morphotypes (bell, lip, flag, and head flowers).
+Selectivity for each bumble- bee-morphotype pair (hereafter "bb-fl pair") was modeled together in a single HGAM with a tensor-product interaction smooth between elevation and time, bb-fl pair as a smoothing factor and fixed intercept effect, and the interactions site * bb-fl pair and year * bb.fl pair as random intercept effects.
+
+An inherent problem with modeling sample-wise variation in selectivity is that the bounds of possible selectivity are strictly dependent on floral abundance [@@Beyer2010-vr].
+A morphotype present at very low relative abundance would, under the null model, be expected to receive very low visitation, which bounds potential negative selectivity near zero but allows positive selectivity to approach 1.
+Conversely, a morphotype present at very high relative abundance would receive very high visitation under the null model, and its potential positive selectivity would be bounded near zero and its potential negative selectivity near -1.
+A morphotype present at exactly 50% relative abundance would have selectivity values bounded at exactly (-0.5, 0.5).
+This clearly creates a degree of unavoidable confoundedness between selectivity and floral abundance.
+The sample-wise relative abundances of the 4 most visited floral morphotypes exhibited a strongly right-skewed distribution falling mainly below 0.5.
+To enable greater transparency of interpretation, therefore, we also fit an HGAM to the surveyed abundance of each floral morphotype so that it can be compared directly to the model of selectivity.
+
+<!-- To remove the effect of rare outliers in which relative abundance was high (and, therefore, potential selectivity low), we only included  -->
+
+The analyses described above were then repeated using uncorrected mean visitation rates rather than selectivity (i.e.
+null model residuals), thus representing the *target use hypothesis*.
+For the ordination of uncorrected mean visitation rates, we used principal coordinates analysis based on the Horn-Morisita distance rather than euclidean principal components analysis because the former is better suited for proportional composition data [@Borcard2018-gb].
 
@@
 
-```{r fig1, echo = FALSE,  fig.cap = paste("Visited families within each floral morphotype.
+A total of 12,918 bumble- bee-flower interactions were included in our final data set, including 16 bumble bee species (with *Bombus terrestric/lucorum* and *Bombus psithyrus* species lumped into species groups), 163 plant species (110 genera, 37 families), and 736 unique bumble- bee-plant interaction pairs.
+*B.
+pascuorum* ("pasc"), *B.
+pratorum* ("prat"), *B.
+soroeensis* ("soro"), *B.
+terrestris-lucorum* ("telu"), and *B.
+wurflenii* ("wurf") were the most abundant species, and they occurred at all sites.
+*B.
+hortorum* ("hort"), *B.
+jonellus* ("jone"),  *B.
+monticola* ("mont"), and *B.
+psithyrus* ("psit") occurred at moderate abundance.
+The remaining species --- *B.
+gerstaeckeri* ("gers"), *B.
+hypnorum* ("hypn"), *B.
+lapidarius* ("lapi"), *B.
+mendax* ("mend"), *B.
+mucidus* ("muci"), and *B.
+pyrenaeus* ("pyre") --- were relatively rare.
+*B.
+humilis* ("humi") was recorded only three times and only in one year, so we omitted it from all analyses.
+
+Floral surveying yielded a total of 352 plant species (191 genera, 52 families), of which 155 (103 genera, 35 families) were observed to be visited by bumble bees.
+Visited floral species included all morphotypes except spike (8) and trap (10) flowers **(Figure 1)**.
+Null (0) and brush (9) morphotypes were, however, excluded from analysis for the reasons given above.
+
+<!-- Check again and make sure all the numbers given here are still correct. -->
+
+```{r fig1, echo = FALSE,  fig.cap = paste("Visited floral morphotypes and their constituent plant families.
+Note that some families span more than on morphotype.
 Barplots show the total number of visits to each family across all bumble bee species and samples (note the rescaling of the y-axis across panels).
-Each floral morphotype is exemplified by a representative species common in our study system: 0 = Plantago lanceolata, 1 = Potentilla erecta, 2 = Gentiana asclepiadea, 3 = Campanula glomerata, 4 = Silene nutans, 5 = Ajuga reptans, 6 = Lotus corniculatus, 7 = Centaurea jacea, 8 = Salix retusa.
+Each floral morphotype is illustrated with a representative species common in our study system: 0 = Plantago lanceolata, 1 = Potentilla erecta, 2 = Gentiana asclepiadea, 3 = Campanula glomerata, 4 = Silene nutans, 5 = Ajuga reptans, 6 = Lotus corniculatus, 7 = Centaurea jacea, 8 = Salix retusa.
 Illustrations are modifications of public domain works obtained from www.plantillustrations.org.
@@
 
-## 3.2 Selection bias by species
+## 3.2 Aggregate selectivity and uncorrected visitation
+
+Selectivity was generally strong, with the visitation rates of most bumblebee species to most floral morphotypes falling significantly above or below the 99% interval of the null model **(Figure 2)**.
+Nonsignificant selectivity was observed mainly among the rarer bumblebee species because their small samples sizes lead to wide null model confidence intervals.
 
@@
 Gray bars depict 99\\% intervals of null model simulations and points depict observed visitation rates.
-Points are color-coded according to where they fall with respect to the 99\\% interval of the null model: red = above (positive bias), blue = below (negative bias), gray = within (nonsignificant bias).")}
+Points are color-coded according to where they fall with respect to the 99\\% interval of the null model: red = above (positive selectivity), blue = below (negative selectivity), gray = within (nonsignificant selectivity).
+The difference between the observed value and the middle of the null model interval can be interpreted as magnitude of selectivity.")}
 knitr::include_graphics("../ms_2/figures/null_model_spwise_preference.png")
@@
 
-## 3.3 Overall selection bias by morphotype
+When floral morphotypes are compared to one another in terms of selectivity, clear patterns emerge **(Figure 3A)**.
+All species exhibited negative selectivity toward disc flowers, while most species exhibited positive selectivity toward bell, lip, and flag flowers.
+Selectivity toward funnel and stalk-disc flowers tended to be neutral (with one strong positive outlier --- *B.
+gerstaeckeri* --- for funnel flowers), reflecting the fact that these morphotypes tended both to be rarely visited and to occur at low abundance.
+Selectivity toward head flowers varied widely from strongly positive to strongly negative.
 
-```{r fig3, echo = FALSE,  fig.cap = paste("Overall selection bias (A) and uncorrected visitation rate (B) of the bumble bee community toward each floral morphotype.
-Each point represents the bias/visitation of one bumble bee species toward the corresponding floral morphotype.
+Differences between floral morphotypes in terms of uncorrected mean visitation rates, reflecting the target use hypothesis, were less pronounced **(Figure 3B)**.
+Bell, lip, flag, and head flowers were visited at high rates by one or more bumble bee species, but median uncorrected mean visitation rates did not differ strongly across morphotypes.
+
+<!-- I could do some sort of linear mixed effects model on this I suppose, but I think it would violate assumptions due to the non-independence of selectivity (selectivity for one thing implies selectivity against others).
+I've already sort of done stats on this with the null model.
+-->
+
+<!-- Before we finalize this, check to make sure the uncorrected visitation rates were really prepared correctly and are not being deflated by a bunch of zeros that should'nt be there. -->
+
+```{r fig3, echo = FALSE,  fig.cap = paste("Selectivity (A) and uncorrected visitation rate (B) of the bumble bee species toward each floral morphotype.
+Each point represents the selectivity/visitation of one bumble bee species toward the corresponding floral morphotype.
 The color of each point in panel A, retained from Figure 2, indicates whether a given observed bias is significantly (p < 0.01) positive or negative with respect to null model simulations.")}
@@
 
-## 3.4 Tongue length as a predictor of biases
+## 3.4 Tongue length as a predictor of floral choice
 
-```{r fig4, echo = FALSE,  fig.cap = paste("Selection biases (A) and uncorrected visitation rates (B) for each morphotype regressed against bumble bee tongue length.
-Each point represents the bias/visitation of one bumble bee species toward the corresponding floral morphotype.
+Tongue length was a strong predictor of selectivity but not of uncorrected mean visitation rates **(Figure 4)**.
+Selectivity toward funnel, bell, lip, flag, and head flowers exhibited significant GAM curves in response bumble bee tongue length, and the model explained 77% of total deviance (adj.
+R^2^ = 0.728).
+For funnel and lip flowers, the response of selectivity to tongue length was positive and approximately linear.
+Selectivity toward bell flowers and head flowers exhibited approximately quadratic responses with selectivity peaking among short-tongued species and declining sharply with increasing tongue length before leveling off.
+Selectivity toward flag flowers exhibited a unimodal peak among medium-tongued species.
+In contrast, uncorrected mean visitation rates were significantly predicted by tongue length only for head flowers, which exhibited a weakly quadratic decline in selectivity from a peak among short-tongued species.
+The uncorrected mean visitation rate model explained only 25% of total deviance (adj.
+R^2^ = 0.053).
+
+```{r fig4, echo = FALSE,  fig.cap = paste("Selectivity (A) and uncorrected visitation rates (B) for each morphotype regressed against bumble bee tongue length.
+Each point represents the selectivity/visitation of one bumble bee species toward the corresponding floral morphotype.
 Curves depict fitted GAM smooths with 95\\% confidence bands.
+Solid lines indicate curves differing significantly (p < 0.05) from a flat line while dashed lines indicated nonsigificant curves.
 Points are color-coded by species, and species are presented in the legend in order of increasing tongue length.")}
-knitr::include_graphics("../ms_2/figures/ktypeXtonguelength_preference.png")
+knitr::include_graphics("../ms_2/figures/ktypeXtonguelength_paired.png")
 ```
 
-```{r fig5, echo = FALSE,  fig.cap = paste("Ordination biplots of bumble bee species according to their selection biases (A) and uncorrected visitation rates (B) for each floral morphotype.
+Interestingly, selectivity and uncorrected mean visitation rates produced broadly similar ordinations, each of which captured more than 80% of total variance **(Figure 5)**.
+In both ordinations, GAM smooths on first two axes also explained most of the variance in tongue length (selectivity mode: deviance explained = 94.7% (adj.
+R^2^ = 0.893), p < 0.001; uncorrected visitation model: deviance explained = 90.8% (adj.
+R^2^ = 0.859), p < 0.001).
+We stress, however, that the two ordinations are not directly comparable in statistical terms since they involve different underlying ordination techniques, necessitated by the fact that selectivity is bounded on the interval of (-1, 1) while mean visitation rate is bounded on the interval (0, 1).
+We present the ordinations primarily for the sake of visualizing the similarity between bumble bee species in terms of selectivity and uncorrected mean visitation rates toward the floral morphotypes considered.
+
+```{r fig5, echo = FALSE,  fig.cap = paste("Ordination biplots of bumble bee species according to their selectivity (A) and uncorrected visitation rates (B) for each floral morphotype.
 Floral morphotypes are abbreviated with their numerical codes and illustrated as in Figure 1.
@@
 
-## 3.5 Lability of bias through elevation and time
+## 3.5 Variation of selectivity, uncorrected visitation, and flower abundance through elevation and time
 
-```{r fig6, echo = FALSE,  fig.cap = paste("Variation in bias through elevation and time modeled as HGAM surfaces.
+We present the temporal and elevational variation of our data first in terms of floral abundance, then uncorrected visitation, and finally selectivity.
+This order of presentation is intended as check against over-interpretation of the variation in selectivity.
+
+The proportional abundance of the most visited floral morphotypes --- bell, lip, flag, and head flowers --- was strongly patterned in time and elevation **(Figure 6)**.
+Bell flowers peaked in proportional abundance early in the season and at mid-elevation, and lip flowers peaked late in the season and at low elevation.
+Flag flowers peaked in mid-season and exhibited two distinct elevational peaks, the lower at around 1000 m.s.l and the higher at around 1800 m.s.l.
+Head flowers exhibited a mid-season peak at low elevation and a broad late season peak at around 1600 m.s.l.
+
+```{r fig6, echo = FALSE,  fig.cap = paste("Variation in proportional flower cover through elevation and modeled as HGAM surfaces.
 Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns.
 Significant (p < 0.05) smooths are depicted with solid contour lines and nonsignificant smooths with dashed contour lines.")}
-knitr::include_graphics("../ms_2/figures/preference_lability_gam.png")
+knitr::include_graphics("../ms_2/figures/flower_cover_gam.png")
 ```
+
 ```{r fig7, echo = FALSE,  fig.cap = paste("Variation in uncorrected visitation rates through elevation and time modeled as HGAM surfaces.
@@
 
-```{r fig8, echo = FALSE,  fig.cap = paste("Variation in flower cover through elevation and time modeled as HGAM surfaces.
+```{r fig8, echo = FALSE,  fig.cap = paste("Variation in selectivity through elevation and time modeled as HGAM surfaces.
 Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns.
 Significant (p < 0.05) smooths are depicted with solid contour lines and nonsignificant smooths with dashed contour lines.")}
-knitr::include_graphics("../ms_2/figures/flower_cover_gam.png")
+knitr::include_graphics("../ms_2/figures/preference_lability_gam.png")
 ```

### 2021-06-11 code/ms2_nullmodels.Rmd
working on discussion

@@
 plot.gamViz(gam3.1, allTerms = TRUE)
-
-gam3.1b <- bam(preference ~
-                bb.ktype +
-                s(site, by = bb.ktype, bs = "re") +
-                s(year, by = bb.ktype, bs = "re") +
-                te(yday, elev.mean, by = bb.ktype),
-              family = "gaussian",
-              method = "fREML",
-              discrete = TRUE,
-              data = filter(null_tab3_join, 
-                            bb.sp %in% c("hort", "prat", "telu", 
-                                         "wurf", "soro", "pasc") &
-                              ktype %in% c("bell", "lip", "flag", "head") &
-                              prop.cover < 0.5)) %>% 
-                getViz()
-
-check.gamViz(gam3.1b)
-gam3.1b_sum <- summary(gam3.1b, re.test = FALSE)
-plot.gamViz(gam3.1b, allTerms = TRUE)
 ```
@@
 ggsave("../ms_2/figures/preference_lability_gam.png", height = 5, width = 6.5)
-
-# And now having dropped the high abundance outliers
-gam3.1b_s_table <- gam3.1b_sum$s.table %>%
-  as.data.frame() %>%
-  rownames_to_column("term") %>%
-  filter(str_detect(term, "te\\(")) %>%
-  mutate(term = str_extract(term, ".{6}$")) %>%
-  separate(term, c("bb.sp", "ktype")) %>%
-  mutate(ktype = case_when(
-    ktype == 3 ~ "bell",
-    ktype == 5 ~ "lip",
-    ktype == 6 ~ "flag",
-    ktype == 7 ~ "head"
-  )) %>%
-  mutate(pval = signif(`p-value`, digits = 3)) %>%
-  select(bb.sp, ktype, edf, pval) %>%
-  as_tibble() %>%
-  mutate(ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")))
-
-# Extract plotting data.
-hort_3_gam3.1b <- plot(sm(gam3.1b, 49))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("hort", n()),
-         ktype = rep("bell", n()))
-
-hort_5_gam3.1b <- plot(sm(gam3.1b, 50))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("hort", n()),
-         ktype = rep("lip", n()))
-
-hort_6_gam3.1b <- plot(sm(gam3.1b, 51))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("hort", n()),
-         ktype = rep("flag", n()))
-
-hort_7_gam3.1b <- plot(sm(gam3.1b, 52))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("hort", n()),
-         ktype = rep("head", n()))
-
-pasc_3_gam3.1b <- plot(sm(gam3.1b, 53))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("pasc", n()),
-         ktype = rep("bell", n()))
-
-pasc_5_gam3.1b <- plot(sm(gam3.1b, 54))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("pasc", n()),
-         ktype = rep("lip", n()))
-
-pasc_6_gam3.1b <- plot(sm(gam3.1b, 55))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("pasc", n()),
-         ktype = rep("flag", n()))
-
-pasc_7_gam3.1b <- plot(sm(gam3.1b, 56))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("pasc", n()),
-         ktype = rep("head", n()))
-
-prat_3_gam3.1b <- plot(sm(gam3.1b, 57))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("prat", n()),
-         ktype = rep("bell", n()))
-
-prat_5_gam3.1b <- plot(sm(gam3.1b, 58))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("prat", n()),
-         ktype = rep("lip", n()))
-
-prat_6_gam3.1b <- plot(sm(gam3.1b, 59))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("prat", n()),
-         ktype = rep("flag", n()))
-
-prat_7_gam3.1b <- plot(sm(gam3.1b, 60))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("prat", n()),
-         ktype = rep("head", n()))
-
-soro_3_gam3.1b <- plot(sm(gam3.1b, 61))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("soro", n()),
-         ktype = rep("bell", n()))
-
-soro_5_gam3.1b <- plot(sm(gam3.1b, 62))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("soro", n()),
-         ktype = rep("lip", n()))
-
-soro_6_gam3.1b <- plot(sm(gam3.1b, 63))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("soro", n()),
-         ktype = rep("flag", n()))
-
-soro_7_gam3.1b <- plot(sm(gam3.1b, 64))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("soro", n()),
-         ktype = rep("head", n()))
-
-telu_3_gam3.1b <- plot(sm(gam3.1b, 65))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("telu", n()),
-         ktype = rep("bell", n()))
-
-telu_5_gam3.1b <- plot(sm(gam3.1b, 66))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("telu", n()),
-         ktype = rep("lip", n()))
-
-telu_6_gam3.1b <- plot(sm(gam3.1b, 67))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("telu", n()),
-         ktype = rep("flag", n()))
-
-telu_7_gam3.1b <- plot(sm(gam3.1b, 68))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("telu", n()),
-         ktype = rep("head", n()))
-
-wurf_3_gam3.1b <- plot(sm(gam3.1b, 69))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("wurf", n()),
-         ktype = rep("bell", n()))
-
-wurf_5_gam3.1b <- plot(sm(gam3.1b, 70))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("wurf", n()),
-         ktype = rep("lip", n()))
-
-wurf_6_gam3.1b <- plot(sm(gam3.1b, 71))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("wurf", n()),
-         ktype = rep("flag", n()))
-
-wurf_7_gam3.1b <- plot(sm(gam3.1b, 72))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("wurf", n()),
-         ktype = rep("head", n()))
-
-plotting_data__gam3.1b <- bind_rows(hort_3_gam3.1b, hort_5_gam3.1b, hort_6_gam3.1b, hort_7_gam3.1b,
-                           pasc_3_gam3.1b, pasc_5_gam3.1b, pasc_6_gam3.1b, pasc_7_gam3.1b,
-                           prat_3_gam3.1b, prat_5_gam3.1b, prat_6_gam3.1b, prat_7_gam3.1b,
-                           soro_3_gam3.1b, soro_5_gam3.1b, soro_6_gam3.1b, soro_7_gam3.1b,
-                           telu_3_gam3.1b, telu_5_gam3.1b, telu_6_gam3.1b, telu_7_gam3.1b,
-                           wurf_3_gam3.1b, wurf_5_gam3.1b, wurf_6_gam3.1b, wurf_7_gam3.1b) %>%
-  group_by(bb.sp, ktype) %>%
-  mutate(z.norm = scale(z, center = FALSE),
-         ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")),
-         date = as_date(x)) %>%
-  ungroup() %>%
-  left_join(bb_traits) %>%
-  left_join(gam3.1b_s_table) %>%
-  select(z, z.norm, date, x, elevation = y, se, 
-         bb.sp, ktype, pbl.w, edf, pval) %>%
-  mutate(sig = if_else(pval < 0.05, TRUE, FALSE))
-
-ggplot(plotting_data__gam3.1b, aes(date, elevation, z = z.norm, fill = z.norm)) +
-  geom_raster() +
-  geom_contour(color = "black",
-               size = 0.1,
-               bins = 40,
-               aes(linetype = sig)) + # by fixing bins, we scale contours to each plot while letting the color ramp represent differences in scale across plots; I think this is a good compromise
-  facet_grid(reorder(bb.sp, -pbl.w) ~ ktype) +
-  scale_linetype_manual(values = c("dotted", "solid"))+
-  scale_fill_viridis_c(na.value = "transparent") +
-  labs(x = NULL, y = "Elevation (m)", fill = "Normalized\neffect", linetype = "p < 0.05") +
-  theme_light(12) +
-  theme(panel.grid = element_blank(),
-        axis.text.x = element_text(angle = 90, vjust = 0.5))
-
-ggsave("../ms_2/figures/preference_lability_gam.pdf", height = 5, width = 6.5)
-ggsave("../ms_2/figures/preference_lability_gam.png", height = 5, width = 6.5)
+```
+
+#### 4.1.1.3 Explore relationship between overall selectivity and the complexity of the elevation-time GAM
+```{r}
+edf_selectivity <- gam3.1_s_table %>%
+  left_join(null_tab1)
+
+ggplot(edf_selectivity, aes(preference, edf)) +
+  geom_smooth(method = "lm", color = "black") +
+  geom_point(aes(color = bb.sp)) +
+  facet_wrap(~ktype) +
+  labs(color = "Species", x = "Selectivity", y = "Effective degree of freedom") +
+  theme_light(14)
+
+ggsave("../ms_2/figures/edf_selectivity.png")
+ggsave("../ms_2/figures/edf_selectivity.pdf")
+
+ggplot(edf_selectivity, aes(preference, edf)) +
+  geom_smooth(method = "lm", color = "black") +
+  geom_point(aes(color = ktype)) +
+  facet_wrap(~bb.sp) +
+  labs(color = "Morphotype") +
+  theme_light(14)
+
 ```
@@
 
-### 4.3 A binomial model: significant preference 0/1
-One way to get around this problem is to model bias as a binary response, either significantly positive or not. 
-#### 4.3.1 Call model
-```{r}
-null_tab3_bin <- null_tab3 %>%
-  mutate(preference.bin = as.integer(if_else(Test == "Stronger", 1, 0)))
-
-gam3.2 <- bam(preference.bin ~
-                bb.ktype +
-                s(site, by = bb.ktype, bs = "re") +
-                s(year, by = bb.ktype, bs = "re") +
-                te(yday, elev.mean, by = bb.ktype),
-              family = "binomial",
-              method = "fREML",
-              discrete = TRUE,
-              data = filter(null_tab3_bin, 
-                            bb.sp %in% c("hort", "prat", "telu", 
-                                         "wurf", "soro", "pasc") &
-                              ktype %in% c("bell", "lip", "flag", "head"))) %>% 
-                getViz()
-
-check.gamViz(gam3.2)
-gam3.2_sum <- summary(gam3.2, re.test = FALSE)
-print(plot.gamViz(gam3.2, allTerms = TRUE), pages = 4)
-```
-
-#### 4.3.2 Plot model
-```{r}
-# Tabulate model summary
-gam3.2_s_table <- gam3.2_sum$s.table %>%
-  as.data.frame() %>%
-  rownames_to_column("term") %>%
-  filter(str_detect(term, "te\\(")) %>%
-  mutate(term = str_extract(term, ".{6}$")) %>%
-  separate(term, c("bb.sp", "ktype")) %>%
-  mutate(ktype = case_when(
-    ktype == 3 ~ "bell",
-    ktype == 5 ~ "lip",
-    ktype == 6 ~ "flag",
-    ktype == 7 ~ "head"
-  )) %>%
-  mutate(pval = signif(`p-value`, digits = 3)) %>%
-  select(bb.sp, ktype, edf, pval) %>%
-  as_tibble() %>%
-  mutate(ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")))
-
-# Extract plotting data.
-hort_3 <- plot(sm(gam3.2, 49))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("hort", n()),
-         ktype = rep("bell", n()))
-
-hort_5 <- plot(sm(gam3.2, 50))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("hort", n()),
-         ktype = rep("lip", n()))
-
-hort_6 <- plot(sm(gam3.2, 51))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("hort", n()),
-         ktype = rep("flag", n()))
-
-hort_7 <- plot(sm(gam3.2, 52))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("hort", n()),
-         ktype = rep("head", n()))
-
-pasc_3 <- plot(sm(gam3.2, 53))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("pasc", n()),
-         ktype = rep("bell", n()))
-
-pasc_5 <- plot(sm(gam3.2, 54))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("pasc", n()),
-         ktype = rep("lip", n()))
-
-pasc_6 <- plot(sm(gam3.2, 55))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("pasc", n()),
-         ktype = rep("flag", n()))
-
-pasc_7 <- plot(sm(gam3.2, 56))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("pasc", n()),
-         ktype = rep("head", n()))
-
-prat_3 <- plot(sm(gam3.2, 57))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("prat", n()),
-         ktype = rep("bell", n()))
-
-prat_5 <- plot(sm(gam3.2, 58))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("prat", n()),
-         ktype = rep("lip", n()))
-
-prat_6 <- plot(sm(gam3.2, 59))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("prat", n()),
-         ktype = rep("flag", n()))
-
-prat_7 <- plot(sm(gam3.2, 60))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("prat", n()),
-         ktype = rep("head", n()))
-
-soro_3 <- plot(sm(gam3.2, 61))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("soro", n()),
-         ktype = rep("bell", n()))
-
-soro_5 <- plot(sm(gam3.2, 62))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("soro", n()),
-         ktype = rep("lip", n()))
-
-soro_6 <- plot(sm(gam3.2, 63))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("soro", n()),
-         ktype = rep("flag", n()))
-
-soro_7 <- plot(sm(gam3.2, 64))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("soro", n()),
-         ktype = rep("head", n()))
-
-telu_3 <- plot(sm(gam3.2, 65))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("telu", n()),
-         ktype = rep("bell", n()))
-
-telu_5 <- plot(sm(gam3.2, 66))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("telu", n()),
-         ktype = rep("lip", n()))
-
-telu_6 <- plot(sm(gam3.2, 67))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("telu", n()),
-         ktype = rep("flag", n()))
-
-telu_7 <- plot(sm(gam3.2, 68))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("telu", n()),
-         ktype = rep("head", n()))
-
-wurf_3 <- plot(sm(gam3.2, 69))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("wurf", n()),
-         ktype = rep("bell", n()))
-
-wurf_5 <- plot(sm(gam3.2, 70))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("wurf", n()),
-         ktype = rep("lip", n()))
-
-wurf_6 <- plot(sm(gam3.2, 71))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("wurf", n()),
-         ktype = rep("flag", n()))
-
-wurf_7 <- plot(sm(gam3.2, 72))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("wurf", n()),
-         ktype = rep("head", n()))
-
-plotting_data <- bind_rows(hort_3, hort_5, hort_6, hort_7,
-                           pasc_3, pasc_5, pasc_6, pasc_7,
-                           prat_3, prat_5, prat_6, prat_7,
-                           soro_3, soro_5, soro_6, soro_7,
-                           telu_3, telu_5, telu_6, telu_7,
-                           wurf_3, wurf_5, wurf_6, wurf_7) %>%
-  group_by(bb.sp, ktype) %>%
-  mutate(z.norm = scale(z, center = FALSE),
-         ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")),
-         date = as_date(x)) %>%
-  ungroup() %>%
-  left_join(bb_traits) %>%
-  left_join(gam3.2_s_table) %>%
-  select(z, z.norm, date, x, elevation = y, se, 
-         bb.sp, ktype, pbl.w, edf, pval) %>%
-  mutate(sig = if_else(pval < 0.05, TRUE, FALSE))
-
-ggplot(plotting_data, aes(date, elevation, z = z.norm, fill = z.norm)) +
-  geom_raster() +
-  geom_contour(color = "black",
-               size = 0.1,
-               bins = 40,
-               aes(linetype = sig)) + # by fixing bins, we scale contours to each plot while letting the color ramp represent differences in scale across plots; I think this is a good compromise
-  facet_grid(reorder(bb.sp, -pbl.w) ~ ktype) +
-  scale_linetype_manual(values = c("dotted", "solid"))+
-  scale_fill_viridis_c(na.value = "transparent") +
-  labs(x = NULL, y = "Elevation (m)", fill = "Normalized\neffect", linetype = "p < 0.05") +
-  theme_light(12) +
-  theme(panel.grid = element_blank(),
-        axis.text.x = element_text(angle = 90, vjust = 0.5))
-```
-
-### 4.4 A normalized model: proportion of maximum/minimum preference
-#### 4.4.1 Call model
-```{r}
-gam3.3 <- bam(preference.norm ~
-                bb.ktype +
-                s(site, by = bb.ktype, bs = "re") +
-                s(year, by = bb.ktype, bs = "re") +
-                te(yday, elev.mean, by = bb.ktype),
-              family = "gaussian",
-              method = "fREML",
-              discrete = TRUE,
-              data = filter(null_tab3.norm, 
-                            bb.sp %in% c("hort", "prat", "telu", 
-                                         "wurf", "soro", "pasc") &
-                              ktype %in% c("bell", "lip", "flag", "head"))) %>% 
-                getViz()
-
-check.gamViz(gam3.3)
-gam3.3_sum <- summary(gam3.3, re.test = FALSE)
-print(plot.gamViz(gam3.3, allTerms = TRUE), pages = 4)
-```
-
-#### 4.4.2 Plot model
-```{r}
-# Tabulate model summary
-gam3.3_s_table <- gam3.3_sum$s.table %>%
-  as.data.frame() %>%
-  rownames_to_column("term") %>%
-  filter(str_detect(term, "te\\(")) %>%
-  mutate(term = str_extract(term, ".{6}$")) %>%
-  separate(term, c("bb.sp", "ktype")) %>%
-  mutate(ktype = case_when(
-    ktype == 3 ~ "bell",
-    ktype == 5 ~ "lip",
-    ktype == 6 ~ "flag",
-    ktype == 7 ~ "head"
-  )) %>%
-  mutate(pval = signif(`p-value`, digits = 3)) %>%
-  select(bb.sp, ktype, edf, pval) %>%
-  as_tibble() %>%
-  mutate(ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")))
-
-# Extract plotting data.
-hort_3 <- plot(sm(gam3.3, 49))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("hort", n()),
-         ktype = rep("bell", n()))
-
-hort_5 <- plot(sm(gam3.3, 50))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("hort", n()),
-         ktype = rep("lip", n()))
-
-hort_6 <- plot(sm(gam3.3, 51))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("hort", n()),
-         ktype = rep("flag", n()))
-
-hort_7 <- plot(sm(gam3.3, 52))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("hort", n()),
-         ktype = rep("head", n()))
-
-pasc_3 <- plot(sm(gam3.3, 53))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("pasc", n()),
-         ktype = rep("bell", n()))
-
-pasc_5 <- plot(sm(gam3.3, 54))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("pasc", n()),
-         ktype = rep("lip", n()))
-
-pasc_6 <- plot(sm(gam3.3, 55))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("pasc", n()),
-         ktype = rep("flag", n()))
-
-pasc_7 <- plot(sm(gam3.3, 56))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("pasc", n()),
-         ktype = rep("head", n()))
-
-prat_3 <- plot(sm(gam3.3, 57))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("prat", n()),
-         ktype = rep("bell", n()))
-
-prat_5 <- plot(sm(gam3.3, 58))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("prat", n()),
-         ktype = rep("lip", n()))
-
-prat_6 <- plot(sm(gam3.3, 59))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("prat", n()),
-         ktype = rep("flag", n()))
-
-prat_7 <- plot(sm(gam3.3, 60))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("prat", n()),
-         ktype = rep("head", n()))
-
-soro_3 <- plot(sm(gam3.3, 61))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("soro", n()),
-         ktype = rep("bell", n()))
-
-soro_5 <- plot(sm(gam3.3, 62))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("soro", n()),
-         ktype = rep("lip", n()))
-
-soro_6 <- plot(sm(gam3.3, 63))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("soro", n()),
-         ktype = rep("flag", n()))
-
-soro_7 <- plot(sm(gam3.3, 64))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("soro", n()),
-         ktype = rep("head", n()))
-
-telu_3 <- plot(sm(gam3.3, 65))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("telu", n()),
-         ktype = rep("bell", n()))
-
-telu_5 <- plot(sm(gam3.3, 66))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("telu", n()),
-         ktype = rep("lip", n()))
-
-telu_6 <- plot(sm(gam3.3, 67))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("telu", n()),
-         ktype = rep("flag", n()))
-
-telu_7 <- plot(sm(gam3.3, 68))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("telu", n()),
-         ktype = rep("head", n()))
-
-wurf_3 <- plot(sm(gam3.3, 69))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("wurf", n()),
-         ktype = rep("bell", n()))
-
-wurf_5 <- plot(sm(gam3.3, 70))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("wurf", n()),
-         ktype = rep("lip", n()))
-
-wurf_6 <- plot(sm(gam3.3, 71))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("wurf", n()),
-         ktype = rep("flag", n()))
-
-wurf_7 <- plot(sm(gam3.3, 72))$ggObj[[1]] %>%
-  mutate(bb.sp = rep("wurf", n()),
-         ktype = rep("head", n()))
-
-plotting_data <- bind_rows(hort_3, hort_5, hort_6, hort_7,
-                           pasc_3, pasc_5, pasc_6, pasc_7,
-                           prat_3, prat_5, prat_6, prat_7,
-                           soro_3, soro_5, soro_6, soro_7,
-                           telu_3, telu_5, telu_6, telu_7,
-                           wurf_3, wurf_5, wurf_6, wurf_7) %>%
-  group_by(bb.sp, ktype) %>%
-  mutate(z.norm = scale(z, center = FALSE),
-         ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")),
-         date = as_date(x)) %>%
-  ungroup() %>%
-  left_join(bb_traits) %>%
-  left_join(gam3.3_s_table) %>%
-  select(z, z.norm, date, x, elevation = y, se, 
-         bb.sp, ktype, pbl.w, edf, pval) %>%
-  mutate(sig = if_else(pval < 0.05, TRUE, FALSE))
-
-ggplot(plotting_data, aes(date, elevation, z = z.norm, fill = z.norm)) +
-  geom_raster() +
-  geom_contour(color = "black",
-               size = 0.1,
-               bins = 40,
-               aes(linetype = sig)) + # by fixing bins, we scale contours to each plot while letting the color ramp represent differences in scale across plots; I think this is a good compromise
-  facet_grid(reorder(bb.sp, -pbl.w) ~ ktype) +
-  scale_linetype_manual(values = c("dotted", "solid"))+
-  scale_fill_viridis_c(na.value = "transparent") +
-  labs(x = NULL, y = "Elevation (m)", fill = "Normalized\neffect", linetype = "p < 0.05") +
-  theme_light(12) +
-  theme(panel.grid = element_blank(),
-        axis.text.x = element_text(angle = 90, vjust = 0.5))
-```
-
-### 4.1.2 Model 2 (with prop_abund as covariate)
-```{r}
-# With abundance covariate (interacting with bb-ktype pair)
-
-## Add abundance data
-abundance <- flower_cover_ktype %>%
-  separate(sample.id, c("site", "date"), sep = "_") %>%
-  group_by(site, date) %>%
-  mutate(prop.cover = cover/sum(cover)) %>%
-  mutate(ktype = case_when(
-    ktype == 0 ~ "null",
-    ktype == 1 ~ "disc",
-    ktype == 2 ~ "funnel",
-    ktype == 3 ~ "bell",
-    ktype == 4 ~ "stalk-disc",
-    ktype == 5 ~ "lip",
-    ktype == 6 ~ "flag",
-    ktype == 7 ~ "head",
-    ktype == 9 ~ "brush"
-  )) %>%
-  mutate(ktype = factor(ktype))
-
-null_tab3_join <- null_tab3 %>%
-  left_join(abundance) %>%
-  mutate(site = factor(site),
-         year = factor(year(date)))
-
-## Model with proportional cover data
-gam3.2 <- bam(preference ~
-                bb.ktype +
-                #prop.cover*bb.ktype +
-                cover:bb.ktype +
-                s(site, by = bb.ktype, bs = "re") +
-                s(year, by = bb.ktype, bs = "re") +
-                te(yday, elev.mean, by = bb.ktype),
-              family = "gaussian",
-              method = "fREML",
-              discrete = TRUE,
-              data = filter(null_tab3_join, 
-                            bb.sp %in% c("hort", "prat", "telu", 
-                                         "wurf", "soro", "pasc") &
-                              ktype %in% c("bell", "lip", "flag", "head"))) %>% 
-                getViz()
-
-#check.gamViz(gam3.2)
-#gam3.2_corr_sum <- summary(gam3.2, re.test = FALSE)
-#print(plot.gamViz(gam3.2, allTerms = TRUE), pages = 4)
-```
-
-Since we have a markedly negative effect of abundance on preference, due to the fact that an avoided resource will seem more avoided when it is abundant than when it is rare, even if the cognitive evaluation of the resource does not vary, perhaps it would be worth adding abundance as a covariate to the preference lability model.
-We do this in model 3.2.
-
-### 4.1.3 Year-specific models
+### 4.3 Year-specific models
 ```{r}
@@
 
-# Binomial model
+# Binomial model (works better than beta, probably because of all the ones and zeros)
 gam5 <- bam(prop.visits ~
@@
 
+# Binomial model
 gam6 <- bam(prop.cover ~
@@
 
+# Beta model (seems to fit a bit better than the binomial model)
 gam6b <- bam(prop.cover ~

### 2021-06-11 ms_2/ms_2.Rmd
working on discussion

@@
 
-Alternative title: Mountain bumble bees exhibit strong but flexible foraging biases through temporal and elevational gradients of floristic turnover
+Understanding patterns of resource use by animals, and the underlying selection processes that govern these patterns, can clarify how ecological communities have assembled, how they currently function, and how they are likely to respond to future perturbations.
+Communities of bumble bees and their floral hosts provide a tractable and compelling system in which to study questions of resource use, and the answers to these questions have immediate relevance to the conservation of bumble bees and their mutualists under the forcings of climate and land use change, particularly in mountain ecosystems.
 
-The process of resource selection by foraging animals has been a central theme of ecological research since the mid 20th century, particularly in the context of resource partitioning among sympatric species.
-Bumble bees as model system...
-Field studies have shown that floral choice in bumble bees is associated with the morphological fittedness between bumble bee tongue length to floral corolla depth.
-Morphological fittedness, however, is only a soft constraint on resource selection; bumble bees also tune their floral choices to learned patterns of reward, and they exhibit remarkable intelligence in learning the handling skills required to exploit difficult flower morphologies.
-
-Using a dataset consisting of more than 13,000 observations collected along a 1400 m elevation gradient over three consecutive years, we describe how bumble bee floral selection is constrained by morphology yet flexibly tuned to local conditions in elevation and time, with considerable idiosyncrasy in floral selection across species with similar tongue length.
-More results...
+Using a dataset consisting of ~13,000 observations collected along a 1400 m elevation gradient over three consecutive years, we
 
@@
 Using these data in conjunction with complementary floral survey data, we explore the *selectivity* of the bumble bee community at the level of floral morphotype [@Kugler1970-gb].
-Following @Beyer2010-vr, we define *selectivity* simply as a disproportionality --- positive or negative --- between the relative abundance of a resource and its rate of use.
+Following @Beyer2010-vr, we define *selectivity* simply as any disproportionality --- positive or negative --- between the relative abundance of a resource and its rate of use.
 This concept approaches the idea of resource *preference*, but inference of the latter, in an absolute sense, requires either experimental control of all resource abundances [@Johnson1980-co] or statistical techniques that condition visitation simultaneously on all resource abundances [@Aarts2008-fo].
@@
 
-We first compare the aggregate selectivity of the bumble bee community across floral morphotypes and then model the species-wise selectivity for each morphotype as a function of bumble bee tongue length.
-Then, by calculating selectivity based on individual samples distributed through elevation and time, we model the variation of bumble bee selectivity through temporal and elevational axes of floristic turnover, revealing the lability of bumble bee floral selectivity.
-Together, these analyses reveal which floral morphotypes receive the bulk of bumble bee visitation, how bumble bee species converge and diverge with respect to resource selection, and the temporal and elevation patterns of floral selectivity can be expected to be acted upon by the phenological and distributional distortions caused by climate and land use change.
+The first objective of our study is to describe and explain the selectivity of the bumble bee community with respect to floral morphotype.
+Using null model analysis to remove neutral effects of resource abundance, we compare the selectivity of the bumble bee community across floral morphotypes and then model the species-wise selectivity for each morphotype as a function of bumble bee tongue length.
 
-Our investigation also affords an opportunity to revisit an open question regarding the role of resource abundance as a driver of flower selection in bumble bees.
+We then revisit an open question regarding the role of resource abundance as a driver of flower selection in bumble bees.
 In studies of resource selection, it is common to pose an abundance-only hypothesis in which foraging is random and selection merely a function of frequency of encounter, resulting in resource use patterns that mirror resource abundance patterns.
@@
 This hypothesis, while obviously subject to boundary conditions of floral availability, is consistent with the anecdotal observation that bumble bees sometimes persist in visiting favored flower species even when these resources are experimentally reduced to low abundance [@Heinrich1976-vb], and it was supported in a study comparing competing models of flower selection in *Bombus vosnesenskii* [@Harmon-Threatt2017-bz].
-If and when the *target use hypothesis* holds, uncorrected visitation rates are a better index of preference (*sensu stricto*) than selectivity (i.e.
+If and when the *target use hypothesis* holds, raw visitation rates (i.e.
+raw for abundance) are a better index of preference (*sensu stricto*) than selectivity (i.e.
 the residuals of an abundance-only null model), since the subtraction of presumed neutral effects of abundance when those effects do not, in fact, exist actually creates new abundance-dependent artifacts and obscures the true process of resource selection [@Harmon-Threatt2017-bz].
@@
 In our study, we take advantage of the fact that there exists a well-established mechanistic basis for bumble bee flower choice, namely the compatibility between floral morphology and bumble bee tongue length.
-Taking this mechanism as a reference and exploiting the tongue-length diversity of the bumble bee species in our study system, we evaluate models of preference with and without abundance effects by comparing the explanatory power of bumble bee tongue length with respect to the preferences inferred under each model
+Taking this mechanism as a reference and exploiting the tongue-length diversity of the bumble bee species in our study system, we evaluate models of preference with and without abundance effects by comparing the explanatory power of bumble bee tongue length with respect to the preferences inferred under each model.
+
+Finally, we model the variation of selectivity, raw visitation, and floral abundance through time and elevation.
+This analysis reveals the elevational and phenological dynamics of the underlying floral community as well as the lability of bumble bee foraging behavior.
+Exploring these patterns is the first step toward understanding how climate and land use change might distort the temporal and elevational structure of bumble- bee-wildflower mutualisms in mountain ecosystems.
 
@@
 
-To quantify selectivity, we compared the proportional visitation of each morphotype (by each bumble bee species) with the rate of visitation expected under the null-hypothesis that visitation of a given resources is purely a function of its relative abundance [@Vazquez2003-qa; @Bluthgen2008-rp].
+To quantify selectivity, we compared the proportional visitation of each morphotype (by each bumble bee species) with the rate of visitation expected under the null-hypothesis that visitation of a given resource is purely a function of its relative abundance [@Vazquez2003-qa; @Bluthgen2008-rp].
 When conducting a null model analysis of interaction networks that are distributed through space and/or time, it is critical to respect the sampling unit of individual sites and dates, since only within this site-date sampling unit does floral visitation data correspond unequivocally to floral abundance data.
@@
 
-We performed two rounds of null model analysis targeting different ecological questions.
-First, we followed the procedure described above, simulating our entire dataset in one modeling process to infer aggregate pattern of selectivity between bumble bee species and floral morphotypes.
-Then, we repeated the null model analysis for each site-date individually so that the selectivity of each bumble bee species toward each floral morphotype could be modeled as a variable through elevation and time.
-For both analyses, we used 500 iterations to achieve a robust sampling distribution, and we considered observed visitation rates to indicate significant selectivity if they fell outside the interval containing 99% of simulated values.
+We followed the procedure described above, simulating our entire dataset in one modeling process to infer aggregate pattern of selectivity between bumble bee species and floral morphotypes.
+We used 500 iterations to achieve a robust sampling distribution, and we considered observed visitation rates to indicate significant selectivity if they fell outside the interval containing 99% of simulated values.
+Then, using the results of this analysis, we modeled selectivity toward each floral morphotype as a function of bumble bee tongue length using hierarchical generalized additive modeling (HGAM), with floral morphotype specified as a smoothing factor [@Pedersen2019-kf].
+Finally, we compared morphotype selectivity across bumble bee species by performing a principal components analysis on the bumble bee species x floral morphotype selectivity matrix, and we mapped bumble bee tongue length onto the ordination space by fitting a 2-dimensional GAM surface to the first two principal components.
 
-Using the aggregate null model results, we modeled selectivity toward each floral morphotype as a function of bumble bee tongue length using hierarchical generalized additive modeling (HGAM), with floral morphotype specified as a smoothing factor [@Pedersen2019-kf].
-We then compared morphotype selectivity across bumble bee species by performing a principal components analysis, and we mapped bumble bee tongue length onto the ordination space by fitting a 2-dimensional GAM surface to the first two principal components.
+As a test of the *target use hypothesis*, we repeated the HGAM and ordination analyses described above using raw mean visitation rates rather than selectivity.
+For the ordination of raw mean visitation rates, we used principal coordinates analysis based on the Horn-Morisita distance rather than euclidean principal components analysis because the former is better suited for proportional composition data [@Borcard2018-gb].
 
-With the sample-wise null model results, we modeled the variation of selectivity between bumble bee species and floral morphotypes through elevation and time.
+To explore the temporal and elevational patterns in our data, we repeated the null model analysis described above but this time for each site-date individually, so that the selectivity of each bumble bee species toward each floral morphotype could be modeled as a variable through elevation and time.
 For this analysis, we included only the six most abundant and widespread bumble bee species in our study system (*B.
@@
 This clearly creates a degree of unavoidable confoundedness between selectivity and floral abundance.
-The sample-wise relative abundances of the 4 most visited floral morphotypes exhibited a strongly right-skewed distribution falling mainly below 0.5.
-To enable greater transparency of interpretation, therefore, we also fit an HGAM to the surveyed abundance of each floral morphotype so that it can be compared directly to the model of selectivity.
-
-<!-- To remove the effect of rare outliers in which relative abundance was high (and, therefore, potential selectivity low), we only included  -->
-
-The analyses described above were then repeated using uncorrected mean visitation rates rather than selectivity (i.e.
-null model residuals), thus representing the *target use hypothesis*.
-For the ordination of uncorrected mean visitation rates, we used principal coordinates analysis based on the Horn-Morisita distance rather than euclidean principal components analysis because the former is better suited for proportional composition data [@Borcard2018-gb].
+With this limitation in mind, we fit HGAMs of the same form to raw visitation rates and total flower cover of each floral morphotype, and we present the results together to be interpreted jointly.
 
@@
 
-## 3.2 Aggregate selectivity and uncorrected visitation
+## 3.2 Aggregate selectivity and raw visitation
 
@@
 
-Differences between floral morphotypes in terms of uncorrected mean visitation rates, reflecting the target use hypothesis, were less pronounced **(Figure 3B)**.
-Bell, lip, flag, and head flowers were visited at high rates by one or more bumble bee species, but median uncorrected mean visitation rates did not differ strongly across morphotypes.
+Differences between floral morphotypes in terms of raw mean visitation rates, reflecting the target use hypothesis, were less pronounced **(Figure 3B)**.
+Bell, lip, flag, and head flowers were visited at high rates by one or more bumble bee species, but median raw mean visitation rates did not differ strongly across morphotypes.
 
@@
 
-<!-- Before we finalize this, check to make sure the uncorrected visitation rates were really prepared correctly and are not being deflated by a bunch of zeros that should'nt be there. -->
+<!-- Before we finalize this, check to make sure the raw visitation rates were really prepared correctly and are not being deflated by a bunch of zeros that should'nt be there. -->
 
-```{r fig3, echo = FALSE,  fig.cap = paste("Selectivity (A) and uncorrected visitation rate (B) of the bumble bee species toward each floral morphotype.
+```{r fig3, echo = FALSE,  fig.cap = paste("Selectivity (A) and raw visitation rate (B) of the bumble bee species toward each floral morphotype.
 Each point represents the selectivity/visitation of one bumble bee species toward the corresponding floral morphotype.
@@
 
-Tongue length was a strong predictor of selectivity but not of uncorrected mean visitation rates **(Figure 4)**.
+Tongue length was a strong predictor of selectivity but not of raw mean visitation rates **(Figure 4)**.
 Selectivity toward funnel, bell, lip, flag, and head flowers exhibited significant GAM curves in response bumble bee tongue length, and the model explained 77% of total deviance (adj.
@@
 Selectivity toward flag flowers exhibited a unimodal peak among medium-tongued species.
-In contrast, uncorrected mean visitation rates were significantly predicted by tongue length only for head flowers, which exhibited a weakly quadratic decline in selectivity from a peak among short-tongued species.
-The uncorrected mean visitation rate model explained only 25% of total deviance (adj.
+In contrast, raw mean visitation rates were significantly predicted by tongue length only for head flowers, which exhibited a weakly quadratic decline in selectivity from a peak among short-tongued species.
+The raw mean visitation rate model explained only 25% of total deviance (adj.
 R^2^ = 0.053).
 
-```{r fig4, echo = FALSE,  fig.cap = paste("Selectivity (A) and uncorrected visitation rates (B) for each morphotype regressed against bumble bee tongue length.
+```{r fig4, echo = FALSE,  fig.cap = paste("Selectivity (A) and raw visitation rates (B) for each morphotype regressed against bumble bee tongue length.
 Each point represents the selectivity/visitation of one bumble bee species toward the corresponding floral morphotype.
@@
 
-Interestingly, selectivity and uncorrected mean visitation rates produced broadly similar ordinations, each of which captured more than 80% of total variance **(Figure 5)**.
+Interestingly, selectivity and raw mean visitation rates produced broadly similar ordinations, each of which captured more than 80% of total variance **(Figure 5)**.
 In both ordinations, GAM smooths on first two axes also explained most of the variance in tongue length (selectivity mode: deviance explained = 94.7% (adj.
-R^2^ = 0.893), p < 0.001; uncorrected visitation model: deviance explained = 90.8% (adj.
+R^2^ = 0.893), p < 0.001; raw visitation model: deviance explained = 90.8% (adj.
 R^2^ = 0.859), p < 0.001).
 We stress, however, that the two ordinations are not directly comparable in statistical terms since they involve different underlying ordination techniques, necessitated by the fact that selectivity is bounded on the interval of (-1, 1) while mean visitation rate is bounded on the interval (0, 1).
-We present the ordinations primarily for the sake of visualizing the similarity between bumble bee species in terms of selectivity and uncorrected mean visitation rates toward the floral morphotypes considered.
+We present the ordinations primarily for the sake of visualizing the similarity between bumble bee species in terms of selectivity and raw mean visitation rates toward the floral morphotypes considered.
 
-```{r fig5, echo = FALSE,  fig.cap = paste("Ordination biplots of bumble bee species according to their selectivity (A) and uncorrected visitation rates (B) for each floral morphotype.
+```{r fig5, echo = FALSE,  fig.cap = paste("Ordination biplots of bumble bee species according to their selectivity (A) and raw visitation rates (B) for each floral morphotype.
 Floral morphotypes are abbreviated with their numerical codes and illustrated as in Figure 1.
@@
 
-## 3.5 Variation of selectivity, uncorrected visitation, and flower abundance through elevation and time
-
-We present the temporal and elevational variation of our data first in terms of floral abundance, then uncorrected visitation, and finally selectivity.
-This order of presentation is intended as check against over-interpretation of the variation in selectivity.
+## 3.5 Variation of floral abundance, raw visitation, and selectivity through elevation and time
 
@@
 Head flowers exhibited a mid-season peak at low elevation and a broad late season peak at around 1600 m.s.l.
+All smooths were significant (p << 0.0001), and the model explained 64.3% of total deviance (adj.
+R^2^ = 0.554).
 
@@
 
-```{r fig7, echo = FALSE,  fig.cap = paste("Variation in uncorrected visitation rates through elevation and time modeled as HGAM surfaces.
+Patterns of raw visitation **(Figure 7)** generally mirrored patterns of proportional abundance **(Figure 6)** for each morphotype.
+Given the expected dependency between floral abundance and raw visitation, this is not surprising.
+There were, however, some notable deviations.
+For example, *B.
+soroeensis* exhibited a strong peak of visitation to bell flowers late in the year at the highest sites, outside the peak abundance of bell flowers.
+*B.
+pratorum* was unique in not exhibiting high visitation rate to lip flowers during their peak abundance late in the season and at low-elevation.
+Bumble bee species differed in the degree to which they exploited the mid- and high-elevation peaks of flag flower abundance; *B.
+hortorum* and *B.
+wurflenii* visited flag flowers heavily at high elevation, while *B.
+pascuorum* --- which also visited flag flowers heavily --- tended rather to exploit the low-elevation peak of abundance.
+The remaining species --- which, based on selectivity analysis, appear to have a low affinity for flag flowers **(Figure 2)** --- exhibited diffuse or nonsignficiant patterns of visitation rates that appear uncorrelated with the abundance of flag flowers.
+
+Smooths were significant (p < 0.05) for all bb-fl pairs except hort-bell, hort-head, pasc-bell, telu-head, and soro-flag. The model explained 33.3% of total deviance (adj. R^2^ = 0.272).
+
+```{r fig7, echo = FALSE,  fig.cap = paste("Variation in raw visitation rates through elevation and time modeled as HGAM surfaces.
 Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns.
@@
 ```
+
+Patterns of selectivity through elevation and time were more complex than those of raw visitation and, by design, more often differed from patterns of floral abundance **(Figure 8)**.
+In some cases, though, patterns of selectivity mirrored patterns of floral availability, suggesting that bumble bees may exhibit functional responses [@Mysterud1998-rs] to floral abundance not entirely accounted for by null models that assume a linear proportionality between abundance and visitation.
+
+One of the more salient patterns is that positive selection for flag flowers was strongly associated with high elevation, except among the species that tended to avoid flag flowers (*B.
+pratorum* and *B.
+soroeensis*) **(Figure 2)**.
+It is also apparent that, as a rule, selectivity is more dynamic where it is also higher overall.
+Bell flowers, which were positively selected overall by the short-tongued bumble bee species *B.
+pratorum*, "*B.
+terrestris-lucorum*, and *B.
+soroeensis* **(Figure 2)**, also exhibited the most complex patterns of selectivity in elevation and time for these bumble bee species.
+In contrast, the long and medium-tongued species *B.
+hortorum*, *B.
+pascuorum*, and *B.
+wurflenii*, which exhibited neutral or weakly positive overall selection toward bell flowers **(Figure 2)**, also show weak patterns of variation in selectivity toward bell flowers.
+Nearly the reverse pattern is evident for flag flowers, reflecting their association with longer-tongued bumble bee species.
+This relationship between the complexity of selectivity dynamics and overall selectivity can be formalized by plotting the effective degrees of freedom against the selectivity HGAM surface on the corresponding overall selectivity between each **(Figure 9)**.
+
+Smooths were significant (p < 0.05) for all bb-fl pairs except hort-head, prat-flag, telu-head, and soro-flag. The model explained 36.4% of total deviance (adj. R^2^ = 0.33).
 
@@
 
+```{r fig9, echo = FALSE,  fig.cap = paste("Complexity of selectivity dynamics through elevation and time as a function of overall selectivity.
+Linear best-fit lines are shown in black with gray bands indicating 95% confidence intervals.")}
+knitr::include_graphics("../ms_2/figures/edf_selectivity.png")
+```
+
 # 4. DISCUSSION
 
-So, we would conclude that (A) abundance matters but (B) preference matters, too.
-We find support for a middle ground between the abundance-only and preference-only models.
+Understanding patterns of resource use by animals, and the underlying selection processes that govern these patterns, can clarify how ecological communities have assembled, how they currently function, and how they are likely to respond to future perturbations.
+Communities of bumble bees and their floral hosts provide a tractable and compelling system in which to study questions of resource use, and the answers to these questions have immediate relevance to the conservation of bumble bees and their mutualists under the forcings of climate and land use change, particularly in mountain ecosystems.
+
+## 4.1 Floral abundance, floral selection, and the *target use hypothesis*
+
+It is well-known that bumble bees forage non-randomly accordingly to morphological constraints, innate preferences, and experienced rewards [e.g.
+Kugler1943-uy; @Brian1954-cy; @Inouye1978-qi; @Raine2007-nq; @Maharaj2019-wm].
+In this regard, the results of our null model analysis are as unsurprising as they are unambiguous; bumble bee resource selection almost never represents flora in unaltered proportion to their abundance.
+The *target use hypothesis* [@De_Valpine2013-be; @Harmon-Threatt2017-bz], however, extends this idea to the perfect antithesis of the random foraging model, positing that bumble bee foraging is a process of *pure* preference, completely irrespective of resource abundance.
+
+Given the well-established mechanistic link between bumble bee tongue length and floral selection [e.g.
+@Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn], our finding that bumble bee tongue length explained selectivity (i.e.
+null model residuals) much better than it explained raw visitation rates indicates that, in our study system, abundance and preference function jointly in bumble bee floral choice.
+This contradicts the *target use hypothesis* as a general rule of bumble bee foraging.
+It does not, however, mean that the *target use hypothesis* does not obtain under certain circumstances, such as when a highly preferred resource varies within a nonlimiting rage of availability.
+Moreover, our results suggest something akin to the *target use hypothesis* but formulated negatively with respect to resources that are *not selected*.
+Visitation rates to disc flowers in our system were very low despite the fact that they often occurred at high abundance, suggesting that while preference seems to interact with resource abundance, avoidance may not.
+In general, though, our findings support the hypothesis that bumble bee resource selection involves both the neutral effects of abundance and the directed effect of preference, which implies that analytical techniques designed to parse these effects, such as the null model analysis we employ, should be favored in studies of bumble bee foraging.
+Accordingly, we shall focus primarily on our calculations of selectivity, rather than raw visitation rates, as we interpret the other results of our study.
+
+## Patterns of selectivity
+
+We found that positive selection was focused on 3 floral morphotypes: bell, lip, and flag flowers.
+These results are broadly consistent with the aggregate visitation patterns previously reported by @Schneller2014-ry in the central Alps at Vorarlberg and Neumayer1999-il in the Austrian Alps ~55 km south of our study system, suggesting that our findings are at least regionally generalizeable.
+When considering these patterns from the perspective of pollination services to mountain flora, it is important to note that while bumble bees are the dominant Hymenopteran pollinators in our study system, flies (especially Syrphid hoverflies) are also abundant flower visitors in mountain ecosystems, and their visitation is mainly restricted to the open flower morphotypes (e.g.
+disc flowers) that receive less attention from bumble bees [@Neumayer1999-il].
+In the context of our research question, we have emphasized the differences in tongue length between bumble bee species, but all bumble bees have relatively long tongues compared to the overall pollinator community, and it is not surprising that their foraging tends to be focused on the deeper flower morphotypes.
+The process of floral resource partitioning *among* bumble bee species is nested within the broader process of floral partitioning *between* bumble bees and other flower visitors.
+
+The strong explanatory power of tongue length with respect to selectivity in our study corroborates the established link between tongue length and floral choice.
+*I think it does so more strongly than most other field studyies I've read.* It is important, though, to consider the species that appear as outliers with respect to the smooth dependency of selectivity on tongue length.
+Outliers worth talking about: telu-disc(+), muci-lip(-), muci-flag(+), soro/psit-head(+).
+
+## Similarity of selectivity across species, potential for competition
+
+While competition as a phenomenon arises from realized patterns of resource use, not the latent process of selection, insight into the process of selection may be a better predictor of competition.
+Use patterns alone are, in a sense, overfit to the observations at hand.
+
+## Variability of floral abundance, visitation, and selectivity through elevation and time; its relevance to climate and land use change
+
+Elevation-time patterns of visitation closely mirrored those of floral abundance, but the former were more diffuse; this may reflect the reluctance of bumble bees to switch from familiar floral taxa even when they become rarer [@Heinrich1976-vb] or less rewarding [I think I remember reading this somewhere...but there is also a paper showing that bumble bees switch more readily than honey bees].
+
+The late high elevation peak of visitation and selectivity between soro and bell flowers is probably driven by soro's peculiar affinity for Campanula.
+Talk about idiosyncracies not capture dby morphology alone.
+importance of other traits, particularly color.
+
+## Caveats:
+
+The problem of queens vs.
+workers...
+Probably the best way to fend off this object would be to re-run all our analyses with only workers and see if any thing changes qualitatively.
+This could all go in the SI.
+
 Need to acknowledge the potential for nonlinear "functional responses" of use to abundance [@Mysterud1998-rs].
-
-## 4. Caveats:
 
@@
 
+## What we did and why
+
+In this study, we investigated patterns of floral resource selection in a mountain bumble bee community with the goals of (1) identifying the most selected floral morphotypes, (2) evaluating the role of floral abundance in driving floral selection, (3)
+
 # ACKNOWLEDGMENTS

### 2021-06-15 code/ms2_data.Rmd
almost there...

@@
   mutate(visited = rep(1, n())) %>% # binary visitation
-  select(site, date, bb.sp, plant.sp, visited) %>%
+  select(site, date, bb.sp, caste, plant.sp, visited) %>%
   left_join(fl_traits) %>%
-  select(site, date, bb.sp, plant.sp, visited, plant.genus, plant.family, ktype = k.type.ss) %>%
+  select(site, date, bb.sp, caste, plant.sp, visited, plant.genus, plant.family, ktype = k.type.ss) %>%
   unite(sample.id, site, date) %>%

### 2021-06-15 code/ms2_nullmodels.Rmd
almost there...

@@
 
+visitation_workers <- visitation %>%
+  filter(caste == "worker")
+
 survey <- read_csv("../data/processed_data/floral_survey.csv")
@@
 fl_traits <- readRDS("../data/processed_data/fl_traits.rds")
+
+vis_spp <- visitation %>%
+  left_join(fl_traits) %>%
+  group_by(sample.id, bb.sp, plant.genus, ktype) %>%
+  summarize(visits = n()) %>%
+  group_by(bb.sp, sample.id) %>%
+  mutate(prop.visits = visits/sum(visits))
+
+view(filter(vis_spp, ktype == "1"))
+
+ggplot(filter(vis_spp, ))
 
@@
   droplevels()
+
+################################
+### Repeat, excluding queens ###
+################################
+
+# The "consumers" table input to generate_null_net()
+con_workers <- visitation_workers %>%
+  #filter(visited == 1) %>%
+  mutate(individual.id = row_number()) %>%
+  select(sample.id, individual.id, bb.sp, ktype, visited) %>%
+  group_by(sample.id) %>%
+  arrange(ktype) %>%
+  pivot_wider(values_from = visited, names_from = ktype, values_fill = 0) %>%
+  arrange(sample.id) %>%
+  ungroup() %>%
+  as.data.frame()
+
+# The "resources" table input to generate_null_net()
+res_workers <- flower_cover_ktype %>%
+  select(sample.id, ktype, cover) %>%
+  group_by(sample.id) %>%
+  arrange(ktype) %>%
+  distinct() %>%
+  pivot_wider(values_from = cover, names_from = ktype, values_fill = 0) %>%
+  ungroup() %>%
+  droplevels() %>%
+  as.data.frame() %>%
+  semi_join(con_workers, by = "sample.id") %>%
+  arrange(sample.id)
+
+# Generate null model
+null1_workers <- generate_null_net(consumers = con_workers[, 3:12],
+                           resources = res_workers[, 2:10],
+                           sims = 500, # raised from default of 100 (takes a while to run)
+                           data.type = "names",
+                           # c.samples = con$sample.id, # parsed by sample
+                           # r.samples = res$sample.id,
+                           c.samples = con_workers[,1], # parsed by sample
+                           r.samples = res_workers[,1],
+                           summary.type = "mean") # summarizing with the mean instead of the default sum standardizes across our bumble bee species that vary in total abundance,
+
+saveRDS(null1_workers, file = "../output/null1_workers.rds")
+null1_workers <- readRDS("../output/null1_workers.rds")
+
+# Tabulate output and generate tests
+null_tab1_workers <- test_interactions(null1_workers, signif.level = 0.99) %>% # make it a bit more conservative to account for multiple comparisons
+  as_tibble() %>%
+  select(bb.sp = Consumer, ktype = Resource, everything()) %>%
+  group_by(bb.sp) %>%
+  mutate(preference = Observed - Null, # a working definition of preference
+         preference.upper99 = Observed - Upper.99.CL, 
+         preference.lower99 = Observed - Lower.99.CL) %>%
+  left_join(bb_traits) %>%
+  ungroup() %>%
+  mutate(bb.sp = fct_reorder(factor(bb.sp), pbl.w)) %>%
+  mutate(ktype = case_when(
+    ktype == 0 ~ "null",
+    ktype == 1 ~ "disc",
+    ktype == 2 ~ "funnel",
+    ktype == 3 ~ "bell",
+    ktype == 4 ~ "stalk-disc",
+    ktype == 5 ~ "lip",
+    ktype == 6 ~ "flag",
+    ktype == 7 ~ "head",
+    ktype == 9 ~ "brush"
+  )) %>%
+  mutate(ktype = factor(ktype, levels = c("null", "disc", "funnel", 
+                                          "bell", "stalk-disc", "lip", 
+                                          "flag", "head", "brush"))) %>%
+  filter(!ktype %in% c("null", "brush")) %>% # type 0 were not reliably quantified (and rarely visited) and type 9 was vanishingly rare
+  droplevels()
 ```
@@
 ggsave("../ms_2/figures/null_model_ktypewise_preference.png", height = 5, width = 7)
+
+####################
+### Workers only ###
+####################
+
+ggplot(null_tab1_workers, aes(Observed, ktype)) +
+  geom_segment(aes(y = ktype, yend = ktype, 
+                   x = Lower.99.CL, xend = Upper.99.CL), 
+               color = "gray40", inherit.aes = FALSE) +
+  geom_point(aes(color = Test)) +
+  facet_wrap(~fct_rev(bb.sp)) +
+  scale_y_discrete(limits = rev) +
+  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
+  labs(x = "Mean proportional visitation", y = "Floral morphotype", color = "Test (p < 0.01)") +
+  theme_light()
+
+ggsave("../ms_2/figures/null_model_spwise_preference_workers.pdf", height = 6, width = 7)
+ggsave("../ms_2/figures/null_model_spwise_preference_workers.png", height = 6, width = 7)
 ```
@@
 ggsave("../ms_2/figures/null_model_ktype_preference.png", null_model_overall, height = 6, width = 7)
+
+####################
+### Workers only ###
+####################
+
+null_model_overall_workers <- ggplot(null_tab1_workers, aes(ktype, preference)) +
+  #geom_violin(fill = "gray70", color = "white", draw_quantiles = c(0.25, 0.5, 0.75)) +
+  geom_boxplot(fill = NA, color = "black", outlier.shape = 4, size = 0.35) +
+  geom_jitter(aes(color = Test), width = 0.3, alpha = 0.6) +
+  theme_light(10) +
+  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
+  labs(x = "Floral morphotype",
+       y = "Selectivity")
+
+ggsave("../ms_2/figures/null_model_ktype_preference_workers.pdf", null_model_overall, height = 6, width = 7)
+ggsave("../ms_2/figures/null_model_ktype_preference_workers.png", null_model_overall, height = 6, width = 7)
 ```
@@
 #plot.gamViz(gam1, allTerms = TRUE)
+
+####################
+### Workers only ###
+####################
+
+gam1_workers <- gam(preference ~ 
+              ktype + 
+              s(pbl.w, by = ktype, k = 5), 
+            data = null_tab1_workers) %>% getViz()
+
+gam1_sum_workers <- summary(gam1_workers)
+check.gamViz(gam1_workers)
+#plot.gamViz(gam1_workers, allTerms = TRUE)
 ```
@@
 ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference.png", null_model_pbl_GAM, height = 6, width = 7)
+
+####################
+### Workers only ###
+####################
+
+# Tabulate model summary
+gam1_s_table_workers <- gam1_sum_workers$s.table %>%
+  as.data.frame() %>%
+  rownames_to_column("term") %>%
+  mutate(ktype = str_extract(term, "(?<=ktype).*$")) %>%
+  mutate(pval = signif(`p-value`, digits = 3)) %>%
+  select(ktype, edf, pval) %>%
+  as_tibble() %>%
+  mutate(ktype = factor(ktype))
+
+# Extract predictions
+gam1_pred_workers <- predict_gam(gam1_workers) %>%
+  mutate(fit.plus.se = fit + (2*se.fit),
+         fit.minus.se = fit - (2*se.fit)) %>%
+  left_join(gam1_s_table_workers) %>%
+  mutate(sig = if_else(pval < 0.05, TRUE, FALSE)) %>%
+  mutate(sig = factor(sig, levels = c(TRUE, FALSE)))
+
+# Plot
+null_model_pbl_GAM_workers <- ggplot(gam1_pred_workers, aes(pbl.w, fit)) +
+  geom_line(aes(linetype = sig)) + # fit line, linetype for significance
+  geom_ribbon(aes(ymin = fit.minus.se, ymax = fit.plus.se), # Uncertainty
+              fill = "black", alpha = 0.2) +
+  geom_point(data = null_tab1_workers, # Original data
+             aes(pbl.w, preference, color = bb.sp), 
+             inherit.aes = FALSE, alpha = 0.5) +
+  facet_wrap(~ktype) +
+  labs(x = "Tongue length (mm)", y = "Selectivity", 
+       color = "Species", linetype = "p < 0.05") +
+  theme_light(10)
+
+ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference_workers.pdf", null_model_pbl_GAM, height = 6, width = 7)
+ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference_workers.png", null_model_pbl_GAM, height = 6, width = 7)
 ```

### 2021-06-15 ms_2/ms_2.Rmd
almost there...

@@
 While male bumble bees were recorded, we chose to analyze only the pooled visitation of queens and workers.
+
 In conjunction with bumble bee observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.01 m^2^ (hereafter "floral survey samples") and identified the plant species according to @Lauber2007-ha and @Oberndorfer2001-ro.
@@
 Then, using the results of this analysis, we modeled selectivity toward each floral morphotype as a function of bumble bee tongue length using hierarchical generalized additive modeling (HGAM), with floral morphotype specified as a smoothing factor [@Pedersen2019-kf].
+To verify that our inference of selectivity was not sensitive to our decision to pool queen and worker visitation, we recalculated the null model and tongue length HGAM using only data from workers.
 Finally, we compared morphotype selectivity across bumble bee species by performing a principal components analysis on the bumble bee species x floral morphotype selectivity matrix, and we mapped bumble bee tongue length onto the ordination space by fitting a 2-dimensional GAM surface to the first two principal components.
@@
 
+The omission of queens from the calculation of overall selectivity did not yield qualitatively different results, so we present only the results of the full data set.
+The results of the worker-only analysis are included in Appendix X.
+
 <!-- I could do some sort of linear mixed effects model on this I suppose, but I think it would violate assumptions due to the non-independence of selectivity (selectivity for one thing implies selectivity against others).
@@
 R^2^ = 0.053).
+
+Again, these results did not differ qualitatively when selectivity was based on the worker-only null model, so we present only the results of the full data set.
+The results of the worker-only analysis are included in Appendix X.
 
@@
 
-Understanding patterns of resource use by animals, and the underlying selection processes that govern these patterns, can clarify how ecological communities have assembled, how they currently function, and how they are likely to respond to future perturbations.
-Communities of bumble bees and their floral hosts provide a tractable and compelling system in which to study questions of resource use, and the answers to these questions have immediate relevance to the conservation of bumble bees and their mutualists under the forcings of climate and land use change, particularly in mountain ecosystems.
-
 ## 4.1 Floral abundance, floral selection, and the *target use hypothesis*
@@
 Kugler1943-uy; @Brian1954-cy; @Inouye1978-qi; @Raine2007-nq; @Maharaj2019-wm].
-In this regard, the results of our null model analysis are as unsurprising as they are unambiguous; bumble bee resource selection almost never represents flora in unaltered proportion to their abundance.
+In this regard, the results of our null model analysis are as unsurprising as they are unambiguous: bumble bee resource selection almost never represents flora in unaltered proportion to their abundance.
 The *target use hypothesis* [@De_Valpine2013-be; @Harmon-Threatt2017-bz], however, extends this idea to the perfect antithesis of the random foraging model, positing that bumble bee foraging is a process of *pure* preference, completely irrespective of resource abundance.
@@
 null model residuals) much better than it explained raw visitation rates indicates that, in our study system, abundance and preference function jointly in bumble bee floral choice.
-This contradicts the *target use hypothesis* as a general rule of bumble bee foraging.
+This casts doubt on the *target use hypothesis* as a general rule of bumble bee foraging.
 It does not, however, mean that the *target use hypothesis* does not obtain under certain circumstances, such as when a highly preferred resource varies within a nonlimiting rage of availability.
 Moreover, our results suggest something akin to the *target use hypothesis* but formulated negatively with respect to resources that are *not selected*.
-Visitation rates to disc flowers in our system were very low despite the fact that they often occurred at high abundance, suggesting that while preference seems to interact with resource abundance, avoidance may not.
+Visitation rates to disc flowers in our system, for example, were very low despite the fact that they often occurred at high abundance, suggesting that while preference seems to interact with resource abundance, avoidance may not.
 In general, though, our findings support the hypothesis that bumble bee resource selection involves both the neutral effects of abundance and the directed effect of preference, which implies that analytical techniques designed to parse these effects, such as the null model analysis we employ, should be favored in studies of bumble bee foraging.
-Accordingly, we shall focus primarily on our calculations of selectivity, rather than raw visitation rates, as we interpret the other results of our study.
+<!-- Accordingly, we shall focus primarily on our calculations of selectivity, rather than raw visitation rates, as we interpret the other results of our study. -->
 
-## Patterns of selectivity
+## 4.2 Overall patterns of selectivity
 
-We found that positive selection was focused on 3 floral morphotypes: bell, lip, and flag flowers.
-These results are broadly consistent with the aggregate visitation patterns previously reported by @Schneller2014-ry in the central Alps at Vorarlberg and Neumayer1999-il in the Austrian Alps ~55 km south of our study system, suggesting that our findings are at least regionally generalizeable.
-When considering these patterns from the perspective of pollination services to mountain flora, it is important to note that while bumble bees are the dominant Hymenopteran pollinators in our study system, flies (especially Syrphid hoverflies) are also abundant flower visitors in mountain ecosystems, and their visitation is mainly restricted to the open flower morphotypes (e.g.
+We found that positive selection was concentrated on 3 floral morphotypes: bell, lip, and flag flowers.
+These results are broadly consistent with the raw visitation patterns previously reported by @Schneller2014-ry in the central Alps at Vorarlberg, Neumayer1999-il in the Austrian Alps ~55 km south of our study system, and Farnberger1996-xo in the vicinity of Vienna, suggesting that our findings are at least regionally generalizeable.
+These neighboring studies tended to find a stronger role of head flowers than suggested by our selectivity analysis, but this is probably because they characterized foraging in terms of raw visitation rather than selectivity.
+Indeed, our raw visitation data similarly indicate a stronger role of head flowers, but our selectivity analysis reveals that while head flowers were highly visited, they tended to be under-visited relative to their abundance.
+
+When considering these patterns from the perspective of pollination services to mountain flora, it is important to note that while bumble bees are the dominant Hymenopteran pollinators in our study system, flies (Diptera), and especially the hover flies of the family Syrphidae, are also abundant flower visitors in mountain ecosystems, and their visitation is mainly restricted to the open flower morphotypes (e.g.
 disc flowers) that receive less attention from bumble bees [@Neumayer1999-il].
 In the context of our research question, we have emphasized the differences in tongue length between bumble bee species, but all bumble bees have relatively long tongues compared to the overall pollinator community, and it is not surprising that their foraging tends to be focused on the deeper flower morphotypes.
-The process of floral resource partitioning *among* bumble bee species is nested within the broader process of floral partitioning *between* bumble bees and other flower visitors.
+The process of floral resource partitioning *among* bumble bee species is nested within the process of floral partitioning *between* bumble bees and other flower visitors.
 
-The strong explanatory power of tongue length with respect to selectivity in our study corroborates the established link between tongue length and floral choice.
-*I think it does so more strongly than most other field studyies I've read.* It is important, though, to consider the species that appear as outliers with respect to the smooth dependency of selectivity on tongue length.
-Outliers worth talking about: telu-disc(+), muci-lip(-), muci-flag(+), soro/psit-head(+).
+While bell, lip, and flag flowers were the most selected morphotypes on average, selectivity toward these morphotypes varied broadly across bumble bee species.
+Moreover, while average selectivity toward funnel and head flowers was low, each received strongly positive selection from one or two bumble bee species.
+The majority of variation in selectivity toward floral morphotypes was explainable by bumble bee tongue length, with the more open bell and head morphotypes selected more positively by short-tongued bumble bee species and the deeper funnel, lip, and flag morphotypes selected more positively by medium- and long-tongued bumble bee species.
+The failure of tongue length to explain selectivity toward disc and stalk-disc flowers reflects the fact that these flowers were almost unanimously ignored by all bumble bee species.
+We cannot say from our data why disc and stalk-disc flowers received such consistently low visitation, but the pattern may be driven in part by the fact that some disc flowers offer pollen but little or no nectar; indeed, Kugler's (1970) typology recognizes this distinction with subdivisions of the disc morphotype into those with (class 1.2) and without (class 1.1) nectar rewards.
+A limitation of our study is that we could not reliably distinguish pollen- and nectar-foraging bumble bees, and it is likely that this distinction would shed further light on the mechanisms underlying the selection patterns we observed.
 
-## Similarity of selectivity across species, potential for competition
+The strong explanatory power of tongue length with respect to selectivity in our study corroborates the established link between tongue length and floral choice [e.g.
+@Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn].
+It is important, though, to consider the species that appear as outliers with respect to the smooth dependency of selectivity on tongue length.
+*B.
+terrestris/lucorum* showed comparatively high (though still negative) selectivity toward disc flowers.
+*B.
+mucidus* exhibited unusually low selectivity (though not significantly different from neutrality) toward lip flowers, complemented by unusually high selectivity toward flag flowers.
+*B.
+soroeensis* and the *B.
+psythyrus* species-group were anomalous in their strongly positive selection toward head flowers.
+Such unexplained variation in selectivity may indicate that traits not captured by Kugler's (1970) morphotypology are also important in governing floral selection by bumble bees.
+Indeed, one such trait --- floral color --- is known to be instrumental in the process of flower detection by foraging bumble bees [@Spaethe2001-bq], and some bumble bee species exhibit innate color preferences [@Raine2007-nq].
+This effect can also be seen in field studies, where bumble bees have been shown to over- or under-represent certain floral colors [e.g.
+@Teras1976-pi; @Neumayer1999-il].
+Moreover, since floral color can be correlated with floral morphology [@Teras1976-pi], it is possible that some of the selectivity attributed to morphology in our models could represent a latent effect of floral color.
 
-While competition as a phenomenon arises from realized patterns of resource use, not the latent process of selection, insight into the process of selection may be a better predictor of competition.
-Use patterns alone are, in a sense, overfit to the observations at hand.
+As mountain bumble bees can be expected to face a future of declining resource availability and altered elevational distributions, it is relevant to ask which species seem poised for the greatest resource conflict should competitive conditions arise.
+While any answer to this question should be regarded as speculative, considering the similarities of morphotype selectivity between species is a good place to start.
+As a general rule, it can be said that species of similar tongue length will be, on average, more likely to compete for resources than species of dissimilar tongue length.
+Our ordination analysis, though, suggests that there may be significant patterns of similarity and divergence even among species of similar tongue length.
+Among short-tongued species, *B.
+jonellus* and *B.
+pyrenaeus* cluster tightly around bell flowers, while *B.
+psithryus* (species group) and *B.
+soroeensis* cluster tightly around head flowers.
+Medium- and long-tongued species differ in their allocation of selectivity across funnel, lip, and flag flowers.
+*B.
+gerstaeckeri*, for example, is associated with lip and funnel flowers but eschews flag flowers, while *B.
+mucidus* does this opposite.
+With respect to the question of climate-induced competition, the case of *B.
+gerstaeckeri* deserves special consideration.
+A red-listed vulnerable species, *B.
+gerstaeckeri* is a pollen specilaist on the lip flowers of the genus *Aconitum* [@Ponchau2006-af; @Schneller2014-ry], and there is evidence that joint elevational shifts of bumble bees and floral resources over the past century have generated more intense exploitation of *Aconitum*, which may indicate nascent competitive conditions for *B.
+gerstaeckeri* [@Marshall2020-pc].
 
-## Variability of floral abundance, visitation, and selectivity through elevation and time; its relevance to climate and land use change
+Broaching the topic of competition raises an issue that can confound any observational study of resource selection: how to disentangle the endogenous process of resource preference from the exogenous process of niche partitioning driven by competition [@Pianka1974-qk].
+Empirical selectivity could reflect a latent process of resource preference, as we assume in our interpretation.
+Alternatively, though, visitation patterns and the selectivity inferred from them could be distorted by competitive forces.
+Intraspecific competition in a bumble bee community would tend to force individuals of a species to broaden their diet beyond what they would prefer under non-competitive conditions, while interspecific competition would be expected to cause a narrowing of diet breadth toward resources that can be exploited most efficiently.
+Thus, observed visitation patterns, even when calculated as abundance-relative selectivity, can only be interpreted in terms of *preference* to the extent that competition can be assumed to be negligible, and assumption must be justified on the basis of additional knowledge of the study system under consideration.
+Our study system in the Berchtesgadener Alps is one of abundant floral resources but harsh abiotic conditions, so it is likely that competition for floral resources is low, or at least transient, and that bumble bee populations are limited instead by the shortness of the growing season, as has been suggested in the similar environment of the Eyne Valley of France [@Iserbyt2008-jn].
+We cannot, however, exclude the possibility that competitive forces distort, to some degree, the visitation patterns we observed.
+Even when competitive conditions do not obtain universally in a given system, competition may occur in a spatially and/or temporally localized way [@Ranta1981-vx], perhaps associated with extreme climatic events that cause reductions in floral resource avaulability [@Rasmont2012-zq; @Thomson2016-gq; @Phillips2018-me].
+This calls to mind the insight of MacArthur1972-qz, that competition need not be frequent to have lasting ecological impacts, and it is likely that the "ghosts of competition past" [@Connell1980-dn] haunt the floral preferences exhibited by bumble bees today.
 
-Elevation-time patterns of visitation closely mirrored those of floral abundance, but the former were more diffuse; this may reflect the reluctance of bumble bees to switch from familiar floral taxa even when they become rarer [@Heinrich1976-vb] or less rewarding [I think I remember reading this somewhere...but there is also a paper showing that bumble bees switch more readily than honey bees].
+## 4.3 Variability of floral abundance, visitation, and selectivity through elevation and time
+
+*First, talk about the lability of selectivity, interesting difference between species, caveats.*
+
+General patterns
+
+While the floral choices of bumble bees have been studied extensively in both laboratory [e.g.
+@Kugler1943-uy; @Raine2007-nq; @Maharaj2019-wm] and field studies [e.g.
+@Heinrich1976-wc; @Teras1976-pi; @Goulson2008-zy], studies of the *dynamics* of bumble bee floral choice along ecological gradients are rare [e.g.
+@Miller-Struttmann2014-wz], and we are not aware of any that use null model controls to distinguish selectivity from the neutral effects of relative abundance.
+
+Interesting deviations
 
 The late high elevation peak of visitation and selectivity between soro and bell flowers is probably driven by soro's peculiar affinity for Campanula.
-Talk about idiosyncracies not capture dby morphology alone.
-importance of other traits, particularly color.
 
-## Caveats:
+We reiterate, however, the inherent problems of comparing selectivity --- understood as an index of the latent process of preference --- across different contexts of resource abundance.
+We recommend the cautious interpretation of selectivity patterns through elevation and time in conjunction with the underlying patterns of floral abundance and raw visitation.
 
-The problem of queens vs.
-workers...
-Probably the best way to fend off this object would be to re-run all our analyses with only workers and see if any thing changes qualitatively.
-This could all go in the SI.
+*Then, segue to climate and land use.*
 
-Need to acknowledge the potential for nonlinear "functional responses" of use to abundance [@Mysterud1998-rs].
+Biological patterns through elevation and time are acted upon by climate and land use change, and the consequence of this action is the generation of new biological patterns through elevation and time.
+A central task of ecology is to understand the patterns that will be acted upon and predict the ecological ramifications of the patterns that will be created.
+In this regard, it is significant that we find marked structure in the morphotypal composition.
+raw visitation, and selectivity of the floral community on which the bumble bees of the Berchtesgadener Alps depend.
+While our data permit only speculation as to the climate and land use driven processes at work in our system, our speculation can be nonetheless well-informed.
+Climate change over the last century has caused significant upslope movement [e.g.
+@Marshall2020-pc] and phenological shifts [e.g.
+@CaraDonna2014-nq] of floral resources for pollinators.
+It is likely that the current patterns of morphotype relative abundance that we observed differ from those that would have been observed a century ago, and in the coming decades warming climates will probably induce further shifts upslope [@Marshall2020-pc] and potentially complex shifts in the onset, peak, and end of flowering periods [@CaraDonna2014-nq].
+These effects of climate change will interact with the likely more rapid effects of land use change.
+In particular, the reforestation of montane meadows can be expected to erode the abundance of flowers from mid-elevation up to the tree line (in our system, around ~1500 m.s.l.).
+Above the tree line, however, floral communities should be relatively safe from reforestation since the advance of alpine tree lines tends to lag behind climate change [@Dullinger2004-qu].
+At lower elevations, agricultural intensification can be expected to suppress overall floral availability by favoring grasses over forbs and shift the composition of the community toward species that thrive under the altered soil conditions and disturbance regimes of intensive agriculture [@Tasser2002-om].
 
-Inference of preference is sensitive to the scale at which availability is estimate [@Beyer2010-vr].
-My floral survey data estimates availability more or less at the meadow scale, right?
-But I don't know what is available outside the meadow.
-This problem is also discussed in [@Harmon-Threatt2017-bz], specifically in the context of bees and flowers.
+## 4.4 Final caveats
 
-The sensitivity of "preference" to availability [@Johnson1980-co; @Beyer2010-vr].
-
-*First, the estimates ofpreference  are  not  directly  interpretable  as  reflectingthe  intrinsic  behavioural  motivations  of  the  animal.Positive  (or  negative)  preference  for  a  given  habitatimplies only that the habitat is used more (or less) fre-quently than expected by chance,conditional upon thedefined sample of availability.
-Even  high  levels  of  useof a habitat can be associated with negative coefficientsif that habitat is common, and, conversely, low levelsof  use  can  be  associated  with  positive  coefficientsif the habitat is rare.
-The sign and magnitude of thecoefficient   cannot   be   interpreted   with   respect   tothe   behavioural   decision   the   animal   is   making.* [@Beyer2010-vr].
-
-*The second consequence is that estimates of prefer-ence  based  on  different  samples  of  availability  (e.g.Design III inManlyet al.2002) are not directly com-parable.
-The  sign  and  magnitude  of  the  estimate  ofpreference  is  a  function  of  both  the  sample  of  useand the sample of availability.
-This creates an identifia-bility  problem  because  preference  can  change  as  afunction of changing use and/or changing availability.If   the   sample   of   availability   differs   between   twoestimates of preference, it is meaningless to draw infer-ences based on the difference in preference.
-Differentpreferences can arise from the same selection process,and   similar   preferences   can   arise   from   differentselection processes.* [@Beyer2010-vr].
-
-This is exactly the problem I am facing.
-I think, though, I can present my "preference lability analysis" together with the raw visitation and availability data, discuss this limitation explicitly, and let the reader decide what to make of it.
-This is also discussed very clearly, and in the context of bumblebees, in @Harmon-Threatt2017-bz.
-
-*Given that preference is conditional on avail-ability,  that  availability  is  likely  to vary  in  space  andthat preference is likely to change as a function of avail-ability, there is little theoretical basis for believing thatpreferences estimated in one area will be good predic-tors  of  preference  in  others.* [@Beyer2010-vr].
-
-In answer to this, I can point to the largely agreeing results of other bumble bee studies.
-
-*Models of habitat preference provide a phenomenolo-gical  description  of  habitat  use,  conditional  on  thesample of availability.
-Although some of the simplify-ing  assumptions  upon  which  these  models  are  often based  may  be  biologically  unrealistic,  they  can  stillhave predictive power (Boyceet al.2002;Wienset al.2008).
-The degree, however, to which simple modelsof habitat preference help us to develop a mechanisticunderstanding of animal movement and distribution islimited.
-Mechanistic models are required to advanceour  biological  understanding  of  the  behavioural  pro-cesses  that  result  in  animal  movement,  distributionand habitat use.*
-
-Tongue length is my mechanistic model.
-And I think the observation that the mechanistic model better explains use-availability bias than raw selection indicates that the use-availability approach better captures the way bumble bees make foraging decisions.
-
-Need to acknowledge the potential for nonlinear "functional responses" of use to abundance [@Mysterud1998-rs].
-
-*The problem of conflating preference and niche partitioning:*
-
-Our study is set in the Berchtesgadener Alps of southeast Germany, a landscape of flower-rich mountain meadows embedded in a matrix of coniferous forest.
-This system is one of abundant floral resources but harsh abiotic conditions, so it is likely that competition for floral resources is low, or at least transient, and that bumble bee populations are limited instead by the shortness of the growing season, as has been suggested in the similar environment of the Eyne Valley of France [@Iserbyt2008-jn].
-Yet while *realized* competition between bumble bees in our system may be low, the question of *potential* competition is highly relevant because mountain ecosystems are currently subject to at least two powerful drivers of directed change.
-
-Resource usage is typically studied as a phenomenological level, with questions of mechanism relegated to speculations regarding the latent process of resource selection that lurks behind realized use patterns [@Beyer2010-vr].
-In our study, we evaluate the target use hypothesis by reference to a known mechanism of floral selection, namely bumble bee tongue length.
+In addition to the qualifying remarks provided in the sections above, it is important to acknowledge some limitations of our study that should caution the reader against the over-interpretation of our results.
+First, the inference of selection as use-relative-to-abundance is sensitive to exactly *how* abundance is estimated [@Johnson1980-co; @Beyer2010-vr].
+Of particular concern in our study is the problem of spatial scale.
+By necessity, we measured floral abundance and visitation along transects *within* our study sites.
+Bumble bee foraging range, however, is much larger than our study sites, and it is possible that the relative abundances of floral morphotypes outside our study sites but within foraging range differ from our within-site estimates.
+Moreover, because bumble bees can forage across multiple meadow patches in a mountain forest matrix [@Mola2020-yc], there exists a prior selective process --- Johnson's (1980) "second-order selection" --- in which bumble bees choose among available foraging areas before selecting from resources within an area.
+Finally, to account for the neutral effects of abundance on floral use, we used a simple null model in which the probability of a resource being selected was assumed to be linearly proportional to the relative abundance of the resource.
+It is possible, however, for resource use to exhibit nonlinear "functional responses" @Mysterud1998-rs to resource abundance, which, relative to a linear null model, would give biased estimates of selectivity.
+Indeed, the fact that patterns of selectivity through elevation and time **(Figure 8)** partly retain of the pattern of resource abundance **(Figure 7)** suggests that there may be some nonlinear dependence of usage on abundance that cannot fully be removed by our linear null model.
 
@@
 
-## What we did and why
+Understanding patterns of resource use by animals, and the underlying selection processes that govern these patterns, can clarify how ecological communities have assembled in the past, how they currently function, and how they are likely to respond to future perturbations.
+Communities of bumble bees and their floral hosts provide a tractable and compelling system in which to study questions of resource use, and the answers to these questions have immediate relevance to the conservation of bumble bees and their mutualists under the forcings of climate and land use change, particularly in mountain ecosystems.
 
-In this study, we investigated patterns of floral resource selection in a mountain bumble bee community with the goals of (1) identifying the most selected floral morphotypes, (2) evaluating the role of floral abundance in driving floral selection, (3)
+<!-- Resource usage is typically studied as a phenomenological level, with questions of mechanism relegated to speculations regarding the latent process of resource selection that lurks behind realized use patterns [@Beyer2010-vr].
+In our study, we evaluate the target use hypothesis by reference to a known mechanism of floral selection, namely bumble bee tongue length.
+-->
 
@@
 
-We thank D.J. McNeil for helpful conversations during data analysis. O. Mitesser kindly provided a high-performance computing environment to expedite null modeling.
+We thank the National Park Berchtesgaden for their support and the permission to work in the national park, and we thank the owners of the study sites outside the national park for access to their land.
+The field work of this study was performed within the framework of the joint research center FORKAST as part of the ‘Bavarian Climate Programme 2020'.
+D.J.
+McNeil provided for helpful conversations during data analysis.
+O.
+Mitesser kindly provided a high-performance computing environment to expedite null modeling.
 

### 2021-06-16 ms_2/ms_2.Rmd
finalized new draft

@@
 ---
-title: "Floral selectivity in mountain bumble bees and a test of the 'target use hypothesis'"
+title: "Floral selection in mountain bumble bees and a test of the 'target use hypothesis'"
 
@@
 
-Understanding patterns of resource use by animals, and the underlying selection processes that govern these patterns, can clarify how ecological communities have assembled, how they currently function, and how they are likely to respond to future perturbations.
+Understanding patterns of resource use by animals, and the underlying selection processes that govern these patterns, can clarify how ecological communities have assembled in the past, how they currently function, and how they are likely to respond to future perturbations.
 Communities of bumble bees and their floral hosts provide a tractable and compelling system in which to study questions of resource use, and the answers to these questions have immediate relevance to the conservation of bumble bees and their mutualists under the forcings of climate and land use change, particularly in mountain ecosystems.
+Using a dataset consisting of ~13,000 interactions recorded along a 1400 m elevation gradient over three consecutive years, we analyze bumble bee resource selection in terms of floral morphotype.
+We find that most bumble bee species exhibit strong positive selection toward flowers of the bell, lip, and/or flag morphotypes.
+Species-level preferences for floral morphotypes were explained well by bumble bee tongue length, which is well-established a key matching trait that constrains bumble bee floral choice.
+Tongue length was a much stronger predictor of null-model-corrected visitation rates than of raw visitation rates, indicating that, contrary to the *target use hypothesis*, neutral effects of resource abundance function in conjunction with learned or innate floral biases to determine bumble bee floral choice.
+By modeling floral selection through elevation and time, we demonstrate strong dynamics of resource availability and resource selection, suggesting that bumble bees have evolved flexible morphotype preferences suited to the extreme spatiotemporal floristic turnover of mountain ecosystems.
+Patterns of resource availability and selection through elevation and time are likely to be distorted in coming decades by the simultaneous processes of climate and land use change, which together cause the displacement and redistribution of mountain floral communities.
 
-Using a dataset consisting of ~13,000 observations collected along a 1400 m elevation gradient over three consecutive years, we
-
-Significance ...
-
-*Keywords*: foraging, trait matching, pollinator, *Bombus*, alpine
+*Keywords*: foraging, trait matching, pollinator, *Bombus*, alpine, niche, resource selection
 
@@
 With this limitation in mind, we fit HGAMs of the same form to raw visitation rates and total flower cover of each floral morphotype, and we present the results together to be interpreted jointly.
+
+In addition to the qualifying remarks above, it is important to acknowledge some additional limitations of our methods.
+First, the inference of selection as use-relative-to-abundance is sensitive to exactly *how* abundance is estimated [@Johnson1980-co; @Beyer2010-vr].
+Of particular concern in our study is the problem of spatial scale.
+By necessity, we measured floral abundance and visitation along transects *within* our study sites.
+Bumble bee foraging range, however, is much larger than our study sites, and it is possible that the relative abundances of floral morphotypes outside our study sites but within foraging range differ from our within-site estimates.
+Moreover, because bumble bees can forage across multiple meadow patches in a mountain forest matrix [@Mola2020-yc], there exists a prior selective process --- Johnson's (1980) "second-order selection" --- in which bumble bees choose among available foraging areas before selecting from resources within an area.
+Finally, to account for the neutral effects of abundance on floral use, we used a simple null model in which the probability of a resource being selected was assumed to be linearly proportional to the relative abundance of the resource.
+It is possible, however, for resource use to exhibit nonlinear "functional responses" @Mysterud1998-rs to resource abundance, which, relative to a linear null model, would give biased estimates of selectivity.
 
@@
 Visitation rates to disc flowers in our system, for example, were very low despite the fact that they often occurred at high abundance, suggesting that while preference seems to interact with resource abundance, avoidance may not.
-In general, though, our findings support the hypothesis that bumble bee resource selection involves both the neutral effects of abundance and the directed effect of preference, which implies that analytical techniques designed to parse these effects, such as the null model analysis we employ, should be favored in studies of bumble bee foraging.
+In general, though, our findings support the hypothesis that bumble bee resource selection involves both the neutral effects of abundance and the directed effect of preference, which implies that analytical techniques designed to disentangle these effects, such as the null model analysis we employ, should be favored in studies of bumble bee foraging.
 <!-- Accordingly, we shall focus primarily on our calculations of selectivity, rather than raw visitation rates, as we interpret the other results of our study. -->
@@
 
-*First, talk about the lability of selectivity, interesting difference between species, caveats.*
+The intraspecific lability of bumble bee foraging preferences (*sensu stricto*) has been demonstrated under laboratory conditions [@Ings2009-vc; @Maharaj2019-wm]; our study reveals the *in situ* lability of selectivity for multiple bumble bee species through elevational and temporal gradients of floristic turnover.
+While we stress the limitations of interpreting empirical selectivity in terms of behavioral preference without controlling the joint abundances of all available resources [@Beyer2010-vr], interpreting the dynamics of selectivity in our study as the adaptive tuning of preference is consistent with the ecological hurdles bumble bees face as characteristically --- and most likely primordially --- mountain species [@Williams1985-ql; @Hines2008-er].
+Where steep temperature gradients and short growing seasons generate extreme spatiotemporal turnover of floral resources, adaptive flexibility of morphological preferences should be favored over fixed innate biases.
 
-General patterns
+At least three mechanisms could underlie the patterns of selectivity through elevation and time that we observed.
+To illustrate these hypotheses, we will take an as example the relatively strong selectivity dynamics of flag flowers in our system, which tended to be most preferred at high-elevation sites **(Figure 8)**.
+Our study does not enable the testing of these hypothesized mechanisms (and their constituent submechanisms), but we articulate them for the sake of clarity and as a guide for future investigations.
+Moreover, the hypotheses we propose not mutually exclusive, and it plausible that all three could obtain simultaneously in a given ecological context.
 
-While the floral choices of bumble bees have been studied extensively in both laboratory [e.g.
-@Kugler1943-uy; @Raine2007-nq; @Maharaj2019-wm] and field studies [e.g.
-@Heinrich1976-wc; @Teras1976-pi; @Goulson2008-zy], studies of the *dynamics* of bumble bee floral choice along ecological gradients are rare [e.g.
-@Miller-Struttmann2014-wz], and we are not aware of any that use null model controls to distinguish selectivity from the neutral effects of relative abundance.
+First, variation in selectivity toward a given morphotype could be driven by changes in the *intrinsic* reward of that focal morphotype.
+These changes might, in turn, be driven by (1) shifts in species composition of the focal morphotype (e.g.
+the occurrence of a particularly rewarding flag flower at high-elevation sites), (2) shifts in the reward levels of individual species (e.g.
+increased nectar production in the flag flower *Lotus corniculatus* under high elevation conditions), or (3) possibly changes in nutritional requirements of bumble bees (e.g.
+an increased need for concentrated nectar to support thermoregulation at colder ambient temperatures).
+The hypothesis is consistent with the interpretation of selectivity as an index of the latent process of absolute preference.
 
-Interesting deviations
+Second, selectivity toward a given morphotype could vary due to changes in the *relative* reward of the focal morphotype under changing background resource conditions.
+For example, it could be that flag flowers are selected more positively at high elevation because they must compete for visitation with highly attractive lip flowers at lower elevation.
+This hypothesis expresses the fundamental problem of inferring preference *sensu stricto* from selection, namely contingency of selection upon the joint availability of all resource options [@Johnson1980-co; Aarts2008-fo; @Beyer2010-vr].
 
-The late high elevation peak of visitation and selectivity between soro and bell flowers is probably driven by soro's peculiar affinity for Campanula.
+The two hypotheses above are both grounded in the intrinsic properties of the floral community, whether those of the focal resource of those of the background resources.
+It is also possible, however, that patterns of resource selection could vary due to competitive interactions within or between bumble bee species.
+For example, the increased selection of flag flowers at high elevation could be driven by competition release from *B.
+pascuorum*, which tends to be the most abundance bumble bee species at lower elevations but becomes relatively scarce above 1500 m (Sponsler et al, submitted).
+Under this hypothesis, an organism's resource selection is the product of exogenous constraint rather than endogenous preference.
+For the reasons discussed above, however, we consider it unlikely that competition plays a constitutive role in governing bumble bee resource selection in the Berchtesgadener Alps.
 
-We reiterate, however, the inherent problems of comparing selectivity --- understood as an index of the latent process of preference --- across different contexts of resource abundance.
+<!-- Patterns of selectivity through elevation and time toward major floral morphotypes tended to be similar across the six dominant bumble bee species we  -->
+
+<!-- Interesting deviations -->
+
+<!-- The late high elevation peak of visitation and selectivity between soro and bell flowers is probably driven by soro's peculiar affinity for Campanula.  -->
+
+<!-- Something about the apparent link between overall selectivity and its variability? -->
+
+<!-- <!-- We reiterate, however, the inherent problems of comparing selectivity --- understood as an index of the latent process of preference --- across different contexts of resource abundance.
 We recommend the cautious interpretation of selectivity patterns through elevation and time in conjunction with the underlying patterns of floral abundance and raw visitation.
-
-*Then, segue to climate and land use.*
+--> -->
 
@@
 A central task of ecology is to understand the patterns that will be acted upon and predict the ecological ramifications of the patterns that will be created.
-In this regard, it is significant that we find marked structure in the morphotypal composition.
-raw visitation, and selectivity of the floral community on which the bumble bees of the Berchtesgadener Alps depend.
+In this regard, it is significant that we find marked structure in the morphotypal composition, raw visitation, and selectivity of the floral community on which the bumble bees of the Berchtesgadener Alps depend.
 While our data permit only speculation as to the climate and land use driven processes at work in our system, our speculation can be nonetheless well-informed.
@@
 At lower elevations, agricultural intensification can be expected to suppress overall floral availability by favoring grasses over forbs and shift the composition of the community toward species that thrive under the altered soil conditions and disturbance regimes of intensive agriculture [@Tasser2002-om].
-
-## 4.4 Final caveats
-
-In addition to the qualifying remarks provided in the sections above, it is important to acknowledge some limitations of our study that should caution the reader against the over-interpretation of our results.
-First, the inference of selection as use-relative-to-abundance is sensitive to exactly *how* abundance is estimated [@Johnson1980-co; @Beyer2010-vr].
-Of particular concern in our study is the problem of spatial scale.
-By necessity, we measured floral abundance and visitation along transects *within* our study sites.
-Bumble bee foraging range, however, is much larger than our study sites, and it is possible that the relative abundances of floral morphotypes outside our study sites but within foraging range differ from our within-site estimates.
-Moreover, because bumble bees can forage across multiple meadow patches in a mountain forest matrix [@Mola2020-yc], there exists a prior selective process --- Johnson's (1980) "second-order selection" --- in which bumble bees choose among available foraging areas before selecting from resources within an area.
-Finally, to account for the neutral effects of abundance on floral use, we used a simple null model in which the probability of a resource being selected was assumed to be linearly proportional to the relative abundance of the resource.
-It is possible, however, for resource use to exhibit nonlinear "functional responses" @Mysterud1998-rs to resource abundance, which, relative to a linear null model, would give biased estimates of selectivity.
-Indeed, the fact that patterns of selectivity through elevation and time **(Figure 8)** partly retain of the pattern of resource abundance **(Figure 7)** suggests that there may be some nonlinear dependence of usage on abundance that cannot fully be removed by our linear null model.
-
-# 5. CONCLUSIONS
-
-Understanding patterns of resource use by animals, and the underlying selection processes that govern these patterns, can clarify how ecological communities have assembled in the past, how they currently function, and how they are likely to respond to future perturbations.
-Communities of bumble bees and their floral hosts provide a tractable and compelling system in which to study questions of resource use, and the answers to these questions have immediate relevance to the conservation of bumble bees and their mutualists under the forcings of climate and land use change, particularly in mountain ecosystems.
-
-<!-- Resource usage is typically studied as a phenomenological level, with questions of mechanism relegated to speculations regarding the latent process of resource selection that lurks behind realized use patterns [@Beyer2010-vr].
-In our study, we evaluate the target use hypothesis by reference to a known mechanism of floral selection, namely bumble bee tongue length.
--->
 

### 2021-07-15 code/ms2_nullmodels.Rmd
revision

@@
   scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
-  labs(x = "Mean proportional visitation", y = "Floral morphotype", color = "Test (p < 0.01)") +
+  labs(x = "Mean visitation rate", y = "Floral morphotype", color = "Test (p < 0.01)") +
   theme_light()
@@
 ggsave("../ms_2/figures/null_model_spwise_preference.png", height = 6, width = 7)
+
+# Species-wise (3-col)
+ggplot(null_tab1, aes(Observed, ktype)) +
+  geom_segment(aes(y = ktype, yend = ktype, 
+                   x = Lower.99.CL, xend = Upper.99.CL), 
+               color = "gray40", inherit.aes = FALSE) +
+  geom_point(aes(color = Test)) +
+  facet_wrap(~fct_rev(bb.sp), ncol = 3) +
+  scale_y_discrete(limits = rev) +
+  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
+  labs(x = "Mean visitation rate", y = "Floral morphotype", color = "Test (p < 0.01)") +
+  theme_light()
+
+ggsave("../ms_2/figures/null_model_spwise_preference_3col.pdf", height = 6, width = 5)
 
@@
   #geom_violin(fill = "gray70", color = "white", draw_quantiles = c(0.25, 0.5, 0.75)) +
-  geom_boxplot(fill = NA, color = "black", outlier.shape = 4, size = 0.35) +
+  geom_boxplot(fill = NA, color = "black", outlier.shape = NA, size = 0.35) +
   geom_jitter(aes(color = Test), width = 0.3, alpha = 0.6) +
@@
 raw_visits_overall <- ggplot(visitation_prop, aes(ktype, mean.prop.visits)) +
-  geom_boxplot(fill = NA, color = "black", outlier.shape = 4, size = 0.35) +
+  geom_boxplot(fill = NA, color = "black", outlier.shape = NA, size = 0.35) +
   geom_jitter(width = 0.3, alpha = 0.5, color = "gray40") +

### 2021-07-15 ms_2/ms_2_revision.Rmd
revision

--- 2021-06-16 ms_2/ms_2.Rmd
@@
 
-author: Douglas B. Sponsler^1^*, Katharina Kallnik^1^, Fabrice Requier^2^, Alice Claßen^1^, A. Fabienne Maihoff^1^, Johanna Sieger^3^, Ingolf Steffan-Dewenter^1^
+author: Douglas B. Sponsler^1^*, Katharina Kallnik^1^, Fabrice Requier^2^, Alice Classen^1^, A. Fabienne Maihoff^1^, Johanna Sieger^3^, Ingolf Steffan-Dewenter^1^
 
@@
 Understanding patterns of resource use by animals, and the underlying selection processes that govern these patterns, can clarify how ecological communities have assembled in the past, how they currently function, and how they are likely to respond to future perturbations.
-Communities of bumble bees and their floral hosts provide a tractable and compelling system in which to study questions of resource use, and the answers to these questions have immediate relevance to the conservation of bumble bees and their mutualists under the forcings of climate and land use change, particularly in mountain ecosystems.
-Using a dataset consisting of ~13,000 interactions recorded along a 1400 m elevation gradient over three consecutive years, we analyze bumble bee resource selection in terms of floral morphotype.
+Communities of bumble bees and their floral hosts provide a tractable and compelling system in which to study questions of resource use, and the answers to these questions have immediate relevance to the conservation of bumble bees and their mutualists under the forcings of climate and land use change.
+It is well-known that bumble bees select flowers disproportionately to their relative abundances, but the *target use hypothesis* proposes that bumble bees select flower types completely irrespective of their relative abundance.
+Using a dataset consisting of ~13,000 interactions recorded along a 1400 m elevation gradient over three consecutive years, we analyze bumble bee resource selection in terms of floral morphotype and compare our results to the predictions of the target use hypothesis.
 We find that most bumble bee species exhibit strong positive selection toward flowers of the bell, lip, and/or flag morphotypes.
-Species-level preferences for floral morphotypes were explained well by bumble bee tongue length, which is well-established a key matching trait that constrains bumble bee floral choice.
+Species-level preferences for floral morphotypes were explained well by bumble bee tongue length, a well-established matching trait that constrains bumble bee floral choice.
 Tongue length was a much stronger predictor of null-model-corrected visitation rates than of raw visitation rates, indicating that, contrary to the *target use hypothesis*, neutral effects of resource abundance function in conjunction with learned or innate floral biases to determine bumble bee floral choice.
@@
 
-The process of resource selection by animals has been an enduring theme of ecological research [@Johnson1980-co; @Beyer2010-vr].
-Methodological difficulties notwithstanding, the study of animals' resource choices is one of the central tasks of community ecology, where the selection and partitioning of resources are instrumental in community assembly, species coexistence, and ecosystem function [@Chesson2000-aw; @Bascompte2007-ls].
+The study of animals' resource choices is one of the central tasks of community ecology, where the selection and partitioning of resources are instrumental in community assembly, species coexistence, and ecosystem function [@Chesson2000-aw; @Bascompte2007-ls].
 Recently, questions of resource selection in ecological communities have taken on new urgency on account of anthropogenic climate and land use change, which threaten to disrupt existing patterns of species interactions and niche relationships [@Tylianakis2008-fi].
@@
 Bumble bees (Hymentoptera: *Bombus* spp.) and their floral hosts provide a diverse yet tractable system in which to explore resource selection.
-Bumble bees are exceptionally versatile florivores whose ability to solve the puzzles posed by complex flower morphologies [@Plowright1984-lc] --- or, indeed, by human investigators [@Alem2016-rn; @Loukola2017-bt] --- renders almost any flower potentially accessible.
+Bumble bees are exceptionally versatile florivores whose ability to solve the puzzles posed by complex flower morphologies [@Plowright1984-lc] --- or, indeed, by human researchers [@Alem2016-rn; @Loukola2017-bt] --- renders almost any flower potentially accessible.
 Nevertheless, the flexibility of bumble bee foraging is tuned to floral reward, which depends on both the intrinsic reward of a given flower and the costs of exploiting it.
@@
 @Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn].
-This association between tongue length and floral niche is imperfect, though, and there can be both convergence in floral resource use between bumble bees of different tongue length and divergence in floral resource use between bumble bees of similar tongue length [e.g.
-@Ranta1980-ok; @Goulson2008-oe].
-Moreover, bumble bee floral choice is also influenced by ecological modulators, such as the spatial distribution of floral patches or competition with other bumble bees [@Ishii2013-sm; @Plowright1984-lc; @Sowig1989-ld].
 
@@
 Warming climates have already caused measurable upslope shifts in mountain bumble bees and their floral hosts [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc], which could introduce novel competitive interactions between bumble bee species that experience increased sympatry.
-At the same time, socioeconomic factors have led to a dramatic decline in traditional alpine grazing management, particularly in the European Alps, resulting in widespread pasture abandonment and reforestation of mountain meadows [@Tappeiner2006-jp; @Tasser2007-qo].
-The regression of meadows to forest entails a massive loss of floral resource availability, and even early stages of forest succession can cause declines in floral abundance and diversity, with corresponding effects on bumble bees [@Iserbyt2015-jc; @Walcher2017-kw; @Suzuki-Ohno2020-jv].
-In contrast, more accessible areas at the bases of mountains are undergoing agricultural intensification that displaces natural vegetation and favors grasses over forbs [@Tasser2002-om].
+At the same time, socioeconomic factors have led to a dramatic decline in traditional alpine grazing management, particularly in the European Alps, resulting in widespread pasture abandonment and reforestation [@Tappeiner2006-jp; @Tasser2007-qo].
+The regression of pastures to forest entails a massive loss of floral resource availability, and even early stages of reforestation can cause declines in floral abundance and diversity, with corresponding effects on bumble bees [@Iserbyt2015-jc; @Walcher2017-kw; @Suzuki-Ohno2020-jv].
+In contrast, more accessible lowland areas are undergoing agricultural intensification that displaces natural vegetation and favors grasses over forbs [@Tasser2002-om].
 Caught between adverse land use changes above and below and subject to climate-induced range shifts, mountain bumble bees are facing a future of reduced floral resource availability and altered sympatry.
@@
 
-In this study, we analyze a data set consisting of ~13,000 interactions between bumble bees and wildflowers recorded over 3 years along a 1400 m elevation gradient.
+In studies of resource selection, it is common to pose a hypothesis in which foraging is random and resource selection is a linear function of resource abundance.
+For bumble bees, which are known to forage non-randomly, this abundance-only hypothesis can serve as a null-model against which observed use patterns can be compared to infer preference (*sensu-lato*) [@Vaughan2018-gj].
+It has also be proposed, however --- in what has been termed the *target use hypothesis* --- that floral selection by bumble bee is a process of *pure* preference, irrespective of resource abundance [@De_Valpine2013-be; @Harmon-Threatt2017-bz].
+This hypothesis has been supported in a study comparing competing models of flower selection in *Bombus vosnesenskii* [@Harmon-Threatt2017-bz], and it is consistent with the anecdotal observation that bumble bees sometimes persist in visiting favored flower species even when they are experimentally reduced to low abundance [@Heinrich1976-vb].
+If and when the *target use hypothesis* holds, raw visitation rates are a better index of preference than the residuals of an abundance-only null model, since the subtraction of presumed neutral effects of abundance when those effects do not really exist actually creates new abundance-dependent artifacts [@Harmon-Threatt2017-bz].
+On the other hand, failing to account for neutral effects of abundance when they do exist similarly obscures the true selection process underpinning observed foraging patterns [@Vazquez2003-qa; @Bluthgen2008-rp].
+
+In this study, we analyze ~13,000 interactions between bumble bees and wildflowers recorded over 3 years along a 1400 m elevation gradient.
 Using these data in conjunction with complementary floral survey data, we explore the *selectivity* of the bumble bee community at the level of floral morphotype [@Kugler1970-gb].
-Following @Beyer2010-vr, we define *selectivity* simply as any disproportionality --- positive or negative --- between the relative abundance of a resource and its rate of use.
+Following @Beyer2010-vr, we define *selectivity* as positive or negative disproportionality between the relative abundance of a resource and its rate of use.
 This concept approaches the idea of resource *preference*, but inference of the latter, in an absolute sense, requires either experimental control of all resource abundances [@Johnson1980-co] or statistical techniques that condition visitation simultaneously on all resource abundances [@Aarts2008-fo].
-Given the scale and complexity of our observational study, we opt for simpler analytical approaches and restrict our inference to the phenomenon of selectivity, which can be understood as a metric of *relative* preference for a focal resource given unmodeled interactions with preferences for alternative resources.
-We deem these interactions beyond the scope of our study, but acknowledge that this requires us to remain partly agnostic as to the process underlying the empirical selectivity we describe.
-
-The first objective of our study is to describe and explain the selectivity of the bumble bee community with respect to floral morphotype.
-Using null model analysis to remove neutral effects of resource abundance, we compare the selectivity of the bumble bee community across floral morphotypes and then model the species-wise selectivity for each morphotype as a function of bumble bee tongue length.
-
-We then revisit an open question regarding the role of resource abundance as a driver of flower selection in bumble bees.
-In studies of resource selection, it is common to pose an abundance-only hypothesis in which foraging is random and selection merely a function of frequency of encounter, resulting in resource use patterns that mirror resource abundance patterns.
-For bumble bees, which are known to forage non-randomly, the abundance-only hypothesis can serve as a null-model against which observed use patterns can be compared to infer selectivity (often called "preference", following a *sensu lato* usage of the term) [@Vaughan2018-gj].
-It has also be proposed, however --- in what has been termed the *target use hypothesis* --- that bumble bees might select resources irrespective of abundance [@De_Valpine2013-be; @Harmon-Threatt2017-bz].
-This hypothesis, while obviously subject to boundary conditions of floral availability, is consistent with the anecdotal observation that bumble bees sometimes persist in visiting favored flower species even when these resources are experimentally reduced to low abundance [@Heinrich1976-vb], and it was supported in a study comparing competing models of flower selection in *Bombus vosnesenskii* [@Harmon-Threatt2017-bz].
-If and when the *target use hypothesis* holds, raw visitation rates (i.e.
-raw for abundance) are a better index of preference (*sensu stricto*) than selectivity (i.e.
-the residuals of an abundance-only null model), since the subtraction of presumed neutral effects of abundance when those effects do not, in fact, exist actually creates new abundance-dependent artifacts and obscures the true process of resource selection [@Harmon-Threatt2017-bz].
-On the other hand, failing to account for neutral effects of abundance when they do obtain similarly obscures the selection process underpinning foraging patterns [@Vazquez2003-qa; @Bluthgen2008-rp].
-In our study, we take advantage of the fact that there exists a well-established mechanistic basis for bumble bee flower choice, namely the compatibility between floral morphology and bumble bee tongue length.
-Taking this mechanism as a reference and exploiting the tongue-length diversity of the bumble bee species in our study system, we evaluate models of preference with and without abundance effects by comparing the explanatory power of bumble bee tongue length with respect to the preferences inferred under each model.
-
-Finally, we model the variation of selectivity, raw visitation, and floral abundance through time and elevation.
-This analysis reveals the elevational and phenological dynamics of the underlying floral community as well as the lability of bumble bee foraging behavior.
+Given the scale and complexity of our observational study, we opt for simpler analytical approaches and restrict our inference to the phenomenon of selectivity, which can be understood as a metric of *relative* preference for a focal resource conditional on unmodeled interactions with preferences for alternative resources.
+Using null model analysis to remove neutral effects of resource abundance, we compare the selectivity of the bumble bee community across floral morphotypes and then model the species-wise selectivity toward each morphotype as a function of tongue length.
+Using this well-established mechanism as a point of reference, we test the *target use hypothesis* by comparing the explanatory power of bumble bee tongue length with respect to raw visitation rates vs.
+null model residuals (i.e.
+selectivity).
+Under the *target use hypothesis*, raw visitation rates should be better explained by bumble bee tongue length than null model residuals, while null model residuals should be better explained by tongue length if, contrary to the *target use hypothesis*, bumble bee resource selection is influenced by resource abundance.
+Finally, we model the variation of selectivity, raw visitation, and floral abundance through time and elevation, revealing the elevational and phenological dynamics of the underlying floral community as well as the lability of bumble bee foraging behavior.
 Exploring these patterns is the first step toward understanding how climate and land use change might distort the temporal and elevational structure of bumble- bee-wildflower mutualisms in mountain ecosystems.
@@
 
-The samples analyzed in this paper have been previously analyzed (Sponsler et al., submitted), but the analyses presented here are non-overlapping.
 Field sampling was conducted 2010-2012 in Berchtesgaden National Park (47.55°N, 12.92°E), located in the Northern Limestone Alps of southeast Germany.
@@
 In 2010 and 2012, this resulted in a temporally staggered pattern of sampling, but a combination of early snowmelt and a late start to field activities in 2011 resulted in sampling that was approximately aligned in time across the whole elevation gradient.
-See Appendix X for a full description of sampling patterns.
+See Sponsler et al.
+(submitted) for details of sampling design.
 
-## 2.2 External data sources
-
-### 2.2.1 Floral morphology
+## 2.2 Floral morphology and bumble bee tongue length
 
@@
 Kugler’s typology was informed by laboratory studies of bumble bee flower choice [@Kugler1943-uy], and there is a precedent for analyzing bumble bee visitation in terms of Kugler’s morphotypes [@Farnberger1996-xo; @Neumayer1999-il; @Schneller2014-ry].
-
 Kugler morphotype classifications of the plants in our study system were accessed via the BIOFLOR database [@Klotz2002-uc].
@@
 
-### 2.2.2 Bumble bee tongue length
-
 Bumble bee tongue-length data (for the worker caste) were obtained from the literature values compiled and summarized by @Arbetman2017-xj.
@@
 
-## 2.3 Data processing
+## 2.3 Data analysis
 
-Prior to analysis, we aligned each visitation sample with a corresponding floral survey sample.
+<!-- Prior to analysis, we aligned each visitation sample with a corresponding floral survey sample.
 In most cases, floral surveying was conducted jointly with bumble bee observation for each site and date.
@@
 In rare cases, a given floral species was recorded during visitation sampling but missed during floral surveying, and these species were retroactively assigned the minimum floral area value of 0.01 m^2^.
+-->
 
-Having thus aligned visitation data and flower cover data, we summed visitation and floral cover across species within each floral morphotype.
+<!-- Having thus aligned visitation data and flower cover data, we summed visitation and floral cover across species within each floral morphotype.
 This yielded, for each of our 855 aligned samples, a cumulative value of flower cover for each floral morphotype and a cumulative value of visitation for each bumble- bee-morphotype pair.
-
-## 2.4 Analysis of resource selection
+-->
 
 To quantify selectivity, we compared the proportional visitation of each morphotype (by each bumble bee species) with the rate of visitation expected under the null-hypothesis that visitation of a given resource is purely a function of its relative abundance [@Vazquez2003-qa; @Bluthgen2008-rp].
-When conducting a null model analysis of interaction networks that are distributed through space and/or time, it is critical to respect the sampling unit of individual sites and dates, since only within this site-date sampling unit does floral visitation data correspond unequivocally to floral abundance data.
-Naive aggregation across sites and dates introduces forbidden links and decouples visitation and abundance [@Vaughan2018-gj].
+When conducting a null model analysis of interaction networks that are distributed through space and/or time, it is critical to respect the sampling unit of individual sites and dates, since only within this site-date sampling unit does floral visitation data correspond unequivocally to floral abundance data [@Vaughan2018-gj].
 We constructed null models following @Vaughan2018-gj.
@@
 Then, using the results of this analysis, we modeled selectivity toward each floral morphotype as a function of bumble bee tongue length using hierarchical generalized additive modeling (HGAM), with floral morphotype specified as a smoothing factor [@Pedersen2019-kf].
+As a test of the *target use hypothesis*, we repeated the HGAM and ordination analyses described above using raw mean visitation rates rather than selectivity.
 To verify that our inference of selectivity was not sensitive to our decision to pool queen and worker visitation, we recalculated the null model and tongue length HGAM using only data from workers.
 Finally, we compared morphotype selectivity across bumble bee species by performing a principal components analysis on the bumble bee species x floral morphotype selectivity matrix, and we mapped bumble bee tongue length onto the ordination space by fitting a 2-dimensional GAM surface to the first two principal components.
-
-As a test of the *target use hypothesis*, we repeated the HGAM and ordination analyses described above using raw mean visitation rates rather than selectivity.
-For the ordination of raw mean visitation rates, we used principal coordinates analysis based on the Horn-Morisita distance rather than euclidean principal components analysis because the former is better suited for proportional composition data [@Borcard2018-gb].
 
@@
 
-An inherent problem with modeling sample-wise variation in selectivity is that the bounds of possible selectivity are strictly dependent on floral abundance [@@Beyer2010-vr].
+An inherent problem with modeling sample-wise variation in selectivity is that the bounds of possible selectivity are strictly dependent on floral abundance [@Beyer2010-vr].
 A morphotype present at very low relative abundance would, under the null model, be expected to receive very low visitation, which bounds potential negative selectivity near zero but allows positive selectivity to approach 1.
@@
 It is possible, however, for resource use to exhibit nonlinear "functional responses" @Mysterud1998-rs to resource abundance, which, relative to a linear null model, would give biased estimates of selectivity.
-
-## 2.5 Software
 
@@
 
-```{r fig5, echo = FALSE,  fig.cap = paste("Ordination biplots of bumble bee species according to their selectivity (A) and raw visitation rates (B) for each floral morphotype.
+```{r fig5, echo = FALSE,  fig.cap = paste("Principal components biplot of bumble bee species according to their selectivity toward each floral morphotype.
 Floral morphotypes are abbreviated with their numerical codes and illustrated as in Figure 1.
@@
 Morphotype illustrations are as described in Figure 1.")}
-knitr::include_graphics("../ms_2/figures/paired_ordination_mod.png")
+knitr::include_graphics("../ms_2/figures/preference_ordination_mod.png")
 ```
@@
 Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns.
-Significant (p < 0.05) smooths are depicted with solid contour lines and nonsignificant smooths with dashed contour lines.")}
+All smooths are significant (p < 0.05).")}
 knitr::include_graphics("../ms_2/figures/flower_cover_gam.png")
@@
 Nearly the reverse pattern is evident for flag flowers, reflecting their association with longer-tongued bumble bee species.
-This relationship between the complexity of selectivity dynamics and overall selectivity can be formalized by plotting the effective degrees of freedom against the selectivity HGAM surface on the corresponding overall selectivity between each **(Figure 9)**.
 
@@
 
-```{r fig9, echo = FALSE,  fig.cap = paste("Complexity of selectivity dynamics through elevation and time as a function of overall selectivity.
-Linear best-fit lines are shown in black with gray bands indicating 95% confidence intervals.")}
-knitr::include_graphics("../ms_2/figures/edf_selectivity.png")
-```
-
 # 4. DISCUSSION
@@
 It is well-known that bumble bees forage non-randomly accordingly to morphological constraints, innate preferences, and experienced rewards [e.g.
-Kugler1943-uy; @Brian1954-cy; @Inouye1978-qi; @Raine2007-nq; @Maharaj2019-wm].
+@Kugler1943-uy; @Brian1954-cy; @Inouye1978-qi; @Raine2007-nq; @Maharaj2019-wm].
 In this regard, the results of our null model analysis are as unsurprising as they are unambiguous: bumble bee resource selection almost never represents flora in unaltered proportion to their abundance.
@@
 
-When considering these patterns from the perspective of pollination services to mountain flora, it is important to note that while bumble bees are the dominant Hymenopteran pollinators in our study system, flies (Diptera), and especially the hover flies of the family Syrphidae, are also abundant flower visitors in mountain ecosystems, and their visitation is mainly restricted to the open flower morphotypes (e.g.
+When considering these patterns from the perspective of pollination services to mountain flora, it is important to note that while bumble bees are the dominant Hymenopteran pollinators in our study system, flies (Diptera), and especially the hover flies (Syrphidae), are also abundant flower visitors in mountain ecosystems, and their visitation is mainly restricted to the open flower morphotypes (e.g.
 disc flowers) that receive less attention from bumble bees [@Neumayer1999-il].
@@
 For example, it could be that flag flowers are selected more positively at high elevation because they must compete for visitation with highly attractive lip flowers at lower elevation.
-This hypothesis expresses the fundamental problem of inferring preference *sensu stricto* from selection, namely contingency of selection upon the joint availability of all resource options [@Johnson1980-co; Aarts2008-fo; @Beyer2010-vr].
+This hypothesis expresses the fundamental problem of inferring preference *sensu stricto* from selection, namely contingency of selection upon the joint availability of all resource options [@Johnson1980-co; @Aarts2008-fo; @Beyer2010-vr].
 

### 2021-07-15 ms_2/ms_2_revision2.Rmd
revision

--- 2021-07-15 ms_2/ms_2_revision.Rmd
@@
 ---
-title: "Floral selection in mountain bumble bees and a test of the 'target use hypothesis'"
+title: "Floral preferences of mountain bumble bees are driven by trait matching but labile through elevation and time"
 
@@
 
-Understanding patterns of resource use by animals, and the underlying selection processes that govern these patterns, can clarify how ecological communities have assembled in the past, how they currently function, and how they are likely to respond to future perturbations.
-Communities of bumble bees and their floral hosts provide a tractable and compelling system in which to study questions of resource use, and the answers to these questions have immediate relevance to the conservation of bumble bees and their mutualists under the forcings of climate and land use change.
-It is well-known that bumble bees select flowers disproportionately to their relative abundances, but the *target use hypothesis* proposes that bumble bees select flower types completely irrespective of their relative abundance.
-Using a dataset consisting of ~13,000 interactions recorded along a 1400 m elevation gradient over three consecutive years, we analyze bumble bee resource selection in terms of floral morphotype and compare our results to the predictions of the target use hypothesis.
-We find that most bumble bee species exhibit strong positive selection toward flowers of the bell, lip, and/or flag morphotypes.
-Species-level preferences for floral morphotypes were explained well by bumble bee tongue length, a well-established matching trait that constrains bumble bee floral choice.
-Tongue length was a much stronger predictor of null-model-corrected visitation rates than of raw visitation rates, indicating that, contrary to the *target use hypothesis*, neutral effects of resource abundance function in conjunction with learned or innate floral biases to determine bumble bee floral choice.
-By modeling floral selection through elevation and time, we demonstrate strong dynamics of resource availability and resource selection, suggesting that bumble bees have evolved flexible morphotype preferences suited to the extreme spatiotemporal floristic turnover of mountain ecosystems.
-Patterns of resource availability and selection through elevation and time are likely to be distorted in coming decades by the simultaneous processes of climate and land use change, which together cause the displacement and redistribution of mountain floral communities.
+Patterns of resource use by animals can clarify how ecological communities have assembled in the past, how they currently function, and how they are likely to respond to future perturbations.
+Using ~13,000 records of bumble- bee-wildflower interactions along a 1400 m elevation gradient, we find that the selection of floral morphotypes by bumble bees is driven both by the directed process of trait-matching and the neutral effects of floral abundance.
+Patterns of floral preference are, moreover, labile through time and elevation, which may be an adaptation to the extreme spatiotemporal floristic turnover of mountain ecosystems.
+The flexibility of bumble bee foraging will be tested in coming years as climate and land use change drive rapid changes in floral resource distribution in time and space, particularly in the mountain ecosystems that host the world's richest communities of bumble bees and their floral mutualists.
 
@@
 
-The study of animals' resource choices is one of the central tasks of community ecology, where the selection and partitioning of resources are instrumental in community assembly, species coexistence, and ecosystem function [@Chesson2000-aw; @Bascompte2007-ls].
-Recently, questions of resource selection in ecological communities have taken on new urgency on account of anthropogenic climate and land use change, which threaten to disrupt existing patterns of species interactions and niche relationships [@Tylianakis2008-fi].
+Patterns of resource selection and partitioning by animals are instrumental in community assembly, species coexistence, and ecosystem function [@Chesson2000-aw; @Bascompte2007-ls].
+Resource selection is fundamentally constrained by phenotypic compatibility (i.e.
+trait-matching) between resources and consumers, but it can also be influenced by the relative availability of resources [@Johnson1980-co] and modulated by exogenous drivers such as competition [@Pianka1974-qk].
+Understanding these interacting processes is key to anticipating the response of ecological communities to anthropogenic climate and land use change, which threaten to disrupt existing patterns of species interactions and niche relationships [@Tylianakis2008-fi].
 
-Bumble bees (Hymentoptera: *Bombus* spp.) and their floral hosts provide a diverse yet tractable system in which to explore resource selection.
-Bumble bees are exceptionally versatile florivores whose ability to solve the puzzles posed by complex flower morphologies [@Plowright1984-lc] --- or, indeed, by human researchers [@Alem2016-rn; @Loukola2017-bt] --- renders almost any flower potentially accessible.
-Nevertheless, the flexibility of bumble bee foraging is tuned to floral reward, which depends on both the intrinsic reward of a given flower and the costs of exploiting it.
-These costs are closely related to the degree of morphological fittedness between bumble bees and their floral hosts, which consists principally in the matching of bumble bee tongue length to floral corolla depth.
-Bumble bee species can differ markedly in tongue length, and floral handling time increases when tongue length and corolla depth are poorly matched [@Inouye1980-mv; @Harder1983-bk], leading to a well-documented association between tongue length and floral choice in bumble bee communities [e.g.
+Plant-pollinator networks are a special case of resource selection in which consumer fitness and resources fitness are positively correlated, causing the traits of plants and pollinators to converge toward compatibility with one another [e.g.
+@Newman2014-ke].
+The coexistence of multiple plant and pollinator species, however, requires niche differentiation [@Palmer2003-ih; @Johnson2019-su], which can be achieved by trait-based partitioning of interactions among pollinators and their floral hosts [@Willmer2011-ak].
+
+Bumble bees (Hymentoptera: *Bombus* spp.) and their floral hosts provide a diverse yet tractable system in which to explore resource selection in the context of plant-pollinator networks.
+Due to their adaptation to cold temperatures, bumble bees are sensitive to climate change [@Goulson2008-oe; @Soroye2020-zb], especially in the mountain ecosystems that host the world's richest bumble bee communities.
+Warming climates have already caused measurable upslope shifts in mountain bumble bees and their floral hosts [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc], which could introduce novel competitive interactions between bumble bee species that experience increased co-occurrence.
+At the same time, socioeconomic factors have led to widespread abandonment and reforestation of mountain meadows, particularly in the European Alps, which dramatically reduces floral resource availability and suppresses bumble bee diversity and abundance [@Tappeiner2006-jp; @Tasser2007-qo; @Walcher2017-kw].
+The ability of bumble bee species to coexist under these conditions of reshuffled distributions and declining floral resources will depend in part on how they select and partition floral resources, which will in turn affect the viability of their floral mutualists, many of which depend principally on bumble bee visitation for pollination [@Goulson2010-lo].
+
+Addressing the question of how bumble bees will select floral resources under novel conditions requires discernment of the generalizable mechanisms that underpin patterns of floral selection observed under current conditions.
+It is ordinarily assumed that the choice of resources by foraging animals is a joint function of trait-based preferences and the neutral effects of the relative abundances of available resources: the so-called "use-availability" model of resource selection [e.g.
+@Beyer2010-vr].
+The most well-studied mechanism of the preference component of bumble bees floral selection is trait-matching between bumble bee tongue length and floral corolla depth.
+Bumble bee species can differ markedly in tongue length, and floral handling time increases when tongue length and corolla depth are poorly matched [@Inouye1980-mv; @Harder1983-bk], leading to a well-documented association between tongue length and floral preferences in bumble bee communities [e.g.
 @Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn].
+Under the use-availability model, tongue-length-based floral preferences should be modulated by the relative abundance of available flora to generate observed patterns of floral selection.
+It has recently been proposed, however --- in what has been termed the *target use hypothesis* --- that floral selection by bumble bees is a process of *pure preference*, irrespective of resource abundance [@De_Valpine2013-be; @Harmon-Threatt2017-bz].
+If and when the *target use hypothesis* holds, observed visitation rates are a better index of preference than metrics designed to control for resource abundance (e.g.
+null model residuals), since the subtraction of presumed neutral effects of abundance when those effects do not really exist actually creates new abundance-dependent artifacts [@Harmon-Threatt2017-bz].
+On the other hand, failing to account for neutral effects of abundance when they do exist similarly obscures the general mechanisms behind observed foraging patterns [@Vazquez2003-qa; @Bluthgen2008-rp] that will determine bumble bee resource selection under novel conditions induced by climate and land use change.
 
-Floral choice in bumble bee communities can serve as a case study in the general phenomenon of animal resource selection, but it is also a matter of practical importance.
-Bumble bees are sensitive to climate and land use change [@Goulson2008-oe; @Soroye2020-zb], especially in the mountain ecosystems that host the world's richest bumble bee communities.
-Warming climates have already caused measurable upslope shifts in mountain bumble bees and their floral hosts [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc], which could introduce novel competitive interactions between bumble bee species that experience increased sympatry.
-At the same time, socioeconomic factors have led to a dramatic decline in traditional alpine grazing management, particularly in the European Alps, resulting in widespread pasture abandonment and reforestation [@Tappeiner2006-jp; @Tasser2007-qo].
-The regression of pastures to forest entails a massive loss of floral resource availability, and even early stages of reforestation can cause declines in floral abundance and diversity, with corresponding effects on bumble bees [@Iserbyt2015-jc; @Walcher2017-kw; @Suzuki-Ohno2020-jv].
-In contrast, more accessible lowland areas are undergoing agricultural intensification that displaces natural vegetation and favors grasses over forbs [@Tasser2002-om].
-Caught between adverse land use changes above and below and subject to climate-induced range shifts, mountain bumble bees are facing a future of reduced floral resource availability and altered sympatry.
-The ability of bumble bee species to coexist under these changes will likely depend significantly on how they select and partition floral resources, and analogous considerations apply to their floral mutualists, many of which depend principally on bumble bee visitation for pollination service [@Goulson2010-lo].
-
-In studies of resource selection, it is common to pose a hypothesis in which foraging is random and resource selection is a linear function of resource abundance.
-For bumble bees, which are known to forage non-randomly, this abundance-only hypothesis can serve as a null-model against which observed use patterns can be compared to infer preference (*sensu-lato*) [@Vaughan2018-gj].
-It has also be proposed, however --- in what has been termed the *target use hypothesis* --- that floral selection by bumble bee is a process of *pure* preference, irrespective of resource abundance [@De_Valpine2013-be; @Harmon-Threatt2017-bz].
-This hypothesis has been supported in a study comparing competing models of flower selection in *Bombus vosnesenskii* [@Harmon-Threatt2017-bz], and it is consistent with the anecdotal observation that bumble bees sometimes persist in visiting favored flower species even when they are experimentally reduced to low abundance [@Heinrich1976-vb].
-If and when the *target use hypothesis* holds, raw visitation rates are a better index of preference than the residuals of an abundance-only null model, since the subtraction of presumed neutral effects of abundance when those effects do not really exist actually creates new abundance-dependent artifacts [@Harmon-Threatt2017-bz].
-On the other hand, failing to account for neutral effects of abundance when they do exist similarly obscures the true selection process underpinning observed foraging patterns [@Vazquez2003-qa; @Bluthgen2008-rp].
-
-In this study, we analyze ~13,000 interactions between bumble bees and wildflowers recorded over 3 years along a 1400 m elevation gradient.
-Using these data in conjunction with complementary floral survey data, we explore the *selectivity* of the bumble bee community at the level of floral morphotype [@Kugler1970-gb].
-Following @Beyer2010-vr, we define *selectivity* as positive or negative disproportionality between the relative abundance of a resource and its rate of use.
-This concept approaches the idea of resource *preference*, but inference of the latter, in an absolute sense, requires either experimental control of all resource abundances [@Johnson1980-co] or statistical techniques that condition visitation simultaneously on all resource abundances [@Aarts2008-fo].
-Given the scale and complexity of our observational study, we opt for simpler analytical approaches and restrict our inference to the phenomenon of selectivity, which can be understood as a metric of *relative* preference for a focal resource conditional on unmodeled interactions with preferences for alternative resources.
-Using null model analysis to remove neutral effects of resource abundance, we compare the selectivity of the bumble bee community across floral morphotypes and then model the species-wise selectivity toward each morphotype as a function of tongue length.
-Using this well-established mechanism as a point of reference, we test the *target use hypothesis* by comparing the explanatory power of bumble bee tongue length with respect to raw visitation rates vs.
-null model residuals (i.e.
-selectivity).
-Under the *target use hypothesis*, raw visitation rates should be better explained by bumble bee tongue length than null model residuals, while null model residuals should be better explained by tongue length if, contrary to the *target use hypothesis*, bumble bee resource selection is influenced by resource abundance.
-Finally, we model the variation of selectivity, raw visitation, and floral abundance through time and elevation, revealing the elevational and phenological dynamics of the underlying floral community as well as the lability of bumble bee foraging behavior.
-Exploring these patterns is the first step toward understanding how climate and land use change might distort the temporal and elevational structure of bumble- bee-wildflower mutualisms in mountain ecosystems.
+In this study, we use ~13,000 records of bumble bee visits to wildflowers, collected over three years along a 1400 m elevation gradient, to explore three unanswered questions regarding resource selection by bumble bees and its response to climate and land use change.
+First, we ask how bumble bee floral preferences map onto a generalized floral morphotypology [@Kugler1970-gb].
+Then, using the well-established relationship between tongue length and floral selection as a point of reference, we test the *target use hypothesis* by comparing the explanatory power of bumble bee tongue length with respect to observed visitation rates vs.
+the residuals of a null model that assumes a linear dependency of selection on resource abundance.
+Under the *target use hypothesis*, tongue length should explain observed visitation rates better than null model residuals, while the opposite finding would indicate that, contrary to the *target use hypothesis*, bumble bee floral selection is influenced by relative abundance.
+Finally, we model the variation of bumble bee floral preference through time and elevation as a first step toward understanding how climate and land use change might distort the temporal and elevational structure of bumble- bee-wildflower mutualisms in mountain ecosystems.
 
@@
 We selected a total of 25 study sites (60 x 60 m) located on pastures along a ~1400 m elevation gradient ranging from 641-2032 m above sea level (m.s.l.).
-Fourteen of these pastures are grazed by cattle or sheep, three are mowed for hay production, and eight have been abandoned throughout the last century and are no longer subject to any human management.
+Fourteen of these pastures are grazed by cattle or sheep, three are mowed for hay production, and eight have been abandoned throughout the last century and are no longer subject to human management.
 
-Sampling of bumble- bee-wildflower interactions consisted of repeated visits (hereafter, “visitation samples”) to each study site at approximately weekly intervals.
+Bumble- bee-wildflower interactions were observed during repeated visits (hereafter, “visitation samples”) to each study site at approximately weekly intervals.
 Sampling was only conducted during periods without rain and with an air temperature of at least 6°C.
 Bumble bees were recorded during one 50-minute transect walk per visit, and bumblebees observed on a given flower were counted as floral visitors.
-Queens were identified to species level in the field to avoid destructively sampling colony foundresses, but workers and males were collected and identified in the lab following @Amiet1996-jz.
-While male bumble bees were recorded, we chose to analyze only the pooled visitation of queens and workers.
-
-In conjunction with bumble bee observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.01 m^2^ (hereafter "floral survey samples") and identified the plant species according to @Lauber2007-ha and @Oberndorfer2001-ro.
-While per-sample effort was standardized across all years, the total number of samples varied: 153 in 2010, 212 in 2011, and 539 in 2012.
-The start of field sampling each year was determined by snowmelt, so lower sites were sampled as early as April while higher sites could not be sampled until June or July.
-In 2010 and 2012, this resulted in a temporally staggered pattern of sampling, but a combination of early snowmelt and a late start to field activities in 2011 resulted in sampling that was approximately aligned in time across the whole elevation gradient.
+Queens were identified to species level in the field to avoid destructively sampling of colony foundresses, but workers and males were collected and identified in the lab following @Amiet1996-jz.
+For analysis, pooled the cryptic *Bombus lucorum/terrestris* species complex and the parasitic *Bombus* (*Psithyrus*) species, respectively, into species-groups.
+Males were recorded, we analyzed only the pooled visitation of queens and workers.
+In conjunction with each bumble bee visitation sample, we estimated the flower cover of herbaceous and shrubby plant species within each 60 x 60 m study plot to the nearest 0.01 m^2^ (hereafter "floral survey samples") and identified the plant species according to @Lauber2007-ha and @Oberndorfer2001-ro.
 See Sponsler et al.
@@
 
-Rather than considering various floral morphological traits independently, we relied on the synthesis of @Kugler1970-gb, who developed a typology consisting of ten primary morphotypes and nested subdivision of each.
+To summarize floral morphology, we relied on the synthesis of @Kugler1970-gb, who developed a typology consisting of ten primary morphotypes and nested subdivision of each.
 Kugler’s typology was informed by laboratory studies of bumble bee flower choice [@Kugler1943-uy], and there is a precedent for analyzing bumble bee visitation in terms of Kugler’s morphotypes [@Farnberger1996-xo; @Neumayer1999-il; @Schneller2014-ry].
@@
 
-<!-- Prior to analysis, we aligned each visitation sample with a corresponding floral survey sample.
-In most cases, floral surveying was conducted jointly with bumble bee observation for each site and date.
-Occasionally, though, floral surveying could not be completed on the same day as bumble bee observation, resulting in a small temporal offset between visitation samples and floral survey samples.
-In these cases, visitation samples were paired with the nearest available floral survey data within a maximum distance of 7 days.
-Visitation samples lacking corresponding floral survey data within 7 days were omitted from further analysis, which resulted in dropping 40 site-dates out of a total of 895, leaving 855 site-dates included in our analyses.
-In a few cases, two consecutive visitation samples at a given site shared the same nearest floral survey sample; in these cases, the visitation samples were first pooled, then aligned to their shared nearest floral survey sample.
-In rare cases, a given floral species was recorded during visitation sampling but missed during floral surveying, and these species were retroactively assigned the minimum floral area value of 0.01 m^2^.
--->
+For the purposes of our study, we define "preference" as the difference between the observed visitation rate to a given resource and visitation rate predicted by the relative abundance of that resource (preference can thus take negative values for under-visited resources).
+For more nuanced definitions of preference, we refer the reader to @Beyer2010-vr and references therein.
 
-<!-- Having thus aligned visitation data and flower cover data, we summed visitation and floral cover across species within each floral morphotype.
-This yielded, for each of our 855 aligned samples, a cumulative value of flower cover for each floral morphotype and a cumulative value of visitation for each bumble- bee-morphotype pair.
--->
-
-To quantify selectivity, we compared the proportional visitation of each morphotype (by each bumble bee species) with the rate of visitation expected under the null-hypothesis that visitation of a given resource is purely a function of its relative abundance [@Vazquez2003-qa; @Bluthgen2008-rp].
-When conducting a null model analysis of interaction networks that are distributed through space and/or time, it is critical to respect the sampling unit of individual sites and dates, since only within this site-date sampling unit does floral visitation data correspond unequivocally to floral abundance data [@Vaughan2018-gj].
-We constructed null models following @Vaughan2018-gj.
-First, resource selection is simulated for individual consumers --- in our case, individual bumble bees --- according to a resource selection process in which the probability of a resource being chosen is proportional to its relative abundance [@Agusti2003-tt].
+To quantify preference, we compared the proportional visitation of each morphotype (by each bumble bee species) with the rates of visitation expected under a null model constructed following @Vaughan2018-gj.
+In this modeling approach, resource selection is simulated for individual consumers --- in our case, individual bumble bees --- according to a resource selection process in which the probability of a resource being chosen is proportional to its relative abundance [@Agusti2003-tt].
 By simulating the visitation of individual consumers rather than species in aggregate, the abundance of each resource, and therefore its probability of visitation, can vary across site-dates, thus preserving the within-sample correspondence between visitation and abundance.
@@
 As this process is iterated, the result is a distribution of visitation rates of each consumer species to each resource under the null assumption that resource selection is determined only by resource abundance.
-The visitation distributions of the null model can then be compared to observed visitation rates to infer positive or negative selectivity of each consumer toward each resource.
-Since the mean of the null distribution can be interpreted as the expected value under null assumptions, the difference between the mean null value and the observed value can be interpreted as a magnitude of selectivity [@Vaughan2018-gj].
-The statistical significance of this selectivity can be estimated as the proportion of simulated null values more extreme than the observed value.
+The visitation distributions of the null model can then be compared to observed visitation rates to infer positive or negative preference of each consumer toward each resource.
+Since the mean of the null distribution can be interpreted as the expected value under null assumptions, the difference between the mean null value and the observed value can be interpreted as a magnitude of preference [@Vaughan2018-gj], and statistical significance can be estimated as the proportion of simulated null values more extreme than the observed value.
 
-We followed the procedure described above, simulating our entire dataset in one modeling process to infer aggregate pattern of selectivity between bumble bee species and floral morphotypes.
-We used 500 iterations to achieve a robust sampling distribution, and we considered observed visitation rates to indicate significant selectivity if they fell outside the interval containing 99% of simulated values.
-Then, using the results of this analysis, we modeled selectivity toward each floral morphotype as a function of bumble bee tongue length using hierarchical generalized additive modeling (HGAM), with floral morphotype specified as a smoothing factor [@Pedersen2019-kf].
-As a test of the *target use hypothesis*, we repeated the HGAM and ordination analyses described above using raw mean visitation rates rather than selectivity.
-To verify that our inference of selectivity was not sensitive to our decision to pool queen and worker visitation, we recalculated the null model and tongue length HGAM using only data from workers.
-Finally, we compared morphotype selectivity across bumble bee species by performing a principal components analysis on the bumble bee species x floral morphotype selectivity matrix, and we mapped bumble bee tongue length onto the ordination space by fitting a 2-dimensional GAM surface to the first two principal components.
+We followed the procedure described above, simulating our entire dataset in one modeling process to infer aggregate preference between bumble bee species and floral morphotypes.
+Using the results of this analysis, we modeled preference toward each floral morphotype as a function of bumble bee tongue length using hierarchical generalized additive modeling (HGAM) [@Pedersen2019-kf].
+As a test of the *target use hypothesis*, we repeated the HGAM analyses described above using observed visitation rates rather than preference.
+To verify that our inference of preference was not sensitive to our decision to pool queen and worker visitation, we recalculated the null model and tongue length HGAM using only data from workers.
+Finally, we compared morphotype preference across bumble bee species by performing a principal components analysis on the bumble bee species x floral morphotype preference matrix, and we mapped bumble bee tongue length onto the ordination space by fitting a 2-dimensional GAM surface to the first two principal components.
 
-To explore the temporal and elevational patterns in our data, we repeated the null model analysis described above but this time for each site-date individually, so that the selectivity of each bumble bee species toward each floral morphotype could be modeled as a variable through elevation and time.
+To explore temporal and elevational variation in floral preference, we repeated the null model analysis described above but this time for each site-date individually, so that the preference of each bumble bee species toward each floral morphotype could be modeled as a variable through elevation and time.
 For this analysis, we included only the six most abundant and widespread bumble bee species in our study system (*B.
@@
 terrestris-lucorum*, and *B.
-wurflenii*) to ensure sufficient data points through elevation and time, and we modeled their selectivity only toward the 4 most visited floral morphotypes (bell, lip, flag, and head flowers).
-Selectivity for each bumble- bee-morphotype pair (hereafter "bb-fl pair") was modeled together in a single HGAM with a tensor-product interaction smooth between elevation and time, bb-fl pair as a smoothing factor and fixed intercept effect, and the interactions site * bb-fl pair and year * bb.fl pair as random intercept effects.
+wurflenii*) to ensure sufficient data points through elevation and time, and we modeled their preference only toward the 4 most visited floral morphotypes (bell, lip, flag, and head flowers).
+Preference for each bumble- bee-morphotype pair (hereafter "bb-fl pair") was modeled together in a single HGAM with an interaction smooth between elevation and time.
 
-An inherent problem with modeling sample-wise variation in selectivity is that the bounds of possible selectivity are strictly dependent on floral abundance [@Beyer2010-vr].
-A morphotype present at very low relative abundance would, under the null model, be expected to receive very low visitation, which bounds potential negative selectivity near zero but allows positive selectivity to approach 1.
-Conversely, a morphotype present at very high relative abundance would receive very high visitation under the null model, and its potential positive selectivity would be bounded near zero and its potential negative selectivity near -1.
-A morphotype present at exactly 50% relative abundance would have selectivity values bounded at exactly (-0.5, 0.5).
-This clearly creates a degree of unavoidable confoundedness between selectivity and floral abundance.
-With this limitation in mind, we fit HGAMs of the same form to raw visitation rates and total flower cover of each floral morphotype, and we present the results together to be interpreted jointly.
-
-In addition to the qualifying remarks above, it is important to acknowledge some additional limitations of our methods.
-First, the inference of selection as use-relative-to-abundance is sensitive to exactly *how* abundance is estimated [@Johnson1980-co; @Beyer2010-vr].
-Of particular concern in our study is the problem of spatial scale.
-By necessity, we measured floral abundance and visitation along transects *within* our study sites.
-Bumble bee foraging range, however, is much larger than our study sites, and it is possible that the relative abundances of floral morphotypes outside our study sites but within foraging range differ from our within-site estimates.
-Moreover, because bumble bees can forage across multiple meadow patches in a mountain forest matrix [@Mola2020-yc], there exists a prior selective process --- Johnson's (1980) "second-order selection" --- in which bumble bees choose among available foraging areas before selecting from resources within an area.
-Finally, to account for the neutral effects of abundance on floral use, we used a simple null model in which the probability of a resource being selected was assumed to be linearly proportional to the relative abundance of the resource.
-It is possible, however, for resource use to exhibit nonlinear "functional responses" @Mysterud1998-rs to resource abundance, which, relative to a linear null model, would give biased estimates of selectivity.
+An inherent problem with modeling sample-wise variation in preference is that the bounds of preference are strictly dependent on floral abundance [@Beyer2010-vr].
+A morphotype present at very low relative abundance would, under the null model, be expected to receive very low visitation, which sets the lower bound of preference near zero (i.e.
+neutrality) but allows the upper bound to approach 1.
+Conversely, a morphotype present at very high relative abundance would receive very high visitation under the null model, and its upper bound of preference would fall near zero and its lower bound near -1.
+This clearly creates a degree of unavoidable confoundedness between preference and floral abundance.
+With this limitation in mind, we fit HGAMs of the same form to the relative abundance of each floral morphotype, and we present the results together to be interpreted jointly.
 
@@
 
-A total of 12,918 bumble- bee-flower interactions were included in our final data set, including 16 bumble bee species (with *Bombus terrestric/lucorum* and *Bombus psithyrus* species lumped into species groups), 163 plant species (110 genera, 37 families), and 736 unique bumble- bee-plant interaction pairs.
+A total of 12,918 bumble- bee-flower interactions were included in our final data set, including 16 bumble bee species, 163 plant species (110 genera, 37 families), and 736 unique bumble- bee-plant interaction pairs.
 *B.
@@
 Floral surveying yielded a total of 352 plant species (191 genera, 52 families), of which 155 (103 genera, 35 families) were observed to be visited by bumble bees.
-Visited floral species included all morphotypes except spike (8) and trap (10) flowers **(Figure 1)**.
+Visited floral species included all morphotypes except spike (8) and trap (10) flowers.
 Null (0) and brush (9) morphotypes were, however, excluded from analysis for the reasons given above.
@@
 
-```{r fig1, echo = FALSE,  fig.cap = paste("Visited floral morphotypes and their constituent plant families.
-Note that some families span more than on morphotype.
-Barplots show the total number of visits to each family across all bumble bee species and samples (note the rescaling of the y-axis across panels).
-Each floral morphotype is illustrated with a representative species common in our study system: 0 = Plantago lanceolata, 1 = Potentilla erecta, 2 = Gentiana asclepiadea, 3 = Campanula glomerata, 4 = Silene nutans, 5 = Ajuga reptans, 6 = Lotus corniculatus, 7 = Centaurea jacea, 8 = Salix retusa.
+## 3.2 Bumble bee preferences toward floral morphotypes
+
+Preference was generally strong, with the visitation rates of most bumblebee species to most floral morphotypes falling significantly above or below the 99% interval of the null model **(Figure 1)**.
+Nonsignificant preference was observed mainly among the rarer bumblebee species because their small samples sizes lead to wide null model confidence intervals.
+
+```{r fig1, echo = FALSE,  fig.cap = paste("Observed visitation rates compared to null model simulations for each bumble bee species and floral morphotype (A).
+Gray bars depict 99\\% intervals of null model simulations and points depict observed visitation rates.
+Points are color-coded according to where they fall with respect to the 99\\% interval of the null model: red = above (positive preference), blue = below (negative preference), gray = within (nonsignificant preference).
+The difference between the observed value and the center of the null model interval can be interpreted as magnitude of preference.
+Each floral morphotype is illustrated (B) with a representative species common in our study system: 1 = Potentilla erecta, 2 = Gentiana asclepiadea, 3 = Campanula glomerata, 4 = Silene nutans, 5 = Ajuga reptans, 6 = Lotus corniculatus, 7 = Centaurea jacea.
 Illustrations are modifications of public domain works obtained from www.plantillustrations.org.
 For details of original sources, see Appendix X.")}
-knitr::include_graphics("../ms_2/figures/morphotype_families_mod.png")
+knitr::include_graphics("../ms_2/figures/fig1.png")
 ```
 
-## 3.2 Aggregate selectivity and raw visitation
+When floral morphotypes are compared to one another in terms of preference, clear patterns emerge **(Figure 2A)**.
+All species exhibited negative preference toward disc flowers, while most species exhibited positive preference toward bell, lip, and flag flowers.
+preference toward funnel and stalk-disc flowers tended to be neutral (with one strong positive outlier --- *B.
+gerstaeckeri* --- for funnel flowers), reflecting the fact that these morphotypes tended both to be rarely visited and to occur at low abundance.
+preference toward head flowers varied widely from strongly positive to strongly negative.
 
-Selectivity was generally strong, with the visitation rates of most bumblebee species to most floral morphotypes falling significantly above or below the 99% interval of the null model **(Figure 2)**.
-Nonsignificant selectivity was observed mainly among the rarer bumblebee species because their small samples sizes lead to wide null model confidence intervals.
-
-```{r fig2, echo = FALSE,  fig.cap = paste("Empirical visitation rates compared to null model simulations for each bumble bee species and floral morphotype.
-Gray bars depict 99\\% intervals of null model simulations and points depict observed visitation rates.
-Points are color-coded according to where they fall with respect to the 99\\% interval of the null model: red = above (positive selectivity), blue = below (negative selectivity), gray = within (nonsignificant selectivity).
-The difference between the observed value and the middle of the null model interval can be interpreted as magnitude of selectivity.")}
-knitr::include_graphics("../ms_2/figures/null_model_spwise_preference.png")
-```
-
-When floral morphotypes are compared to one another in terms of selectivity, clear patterns emerge **(Figure 3A)**.
-All species exhibited negative selectivity toward disc flowers, while most species exhibited positive selectivity toward bell, lip, and flag flowers.
-Selectivity toward funnel and stalk-disc flowers tended to be neutral (with one strong positive outlier --- *B.
-gerstaeckeri* --- for funnel flowers), reflecting the fact that these morphotypes tended both to be rarely visited and to occur at low abundance.
-Selectivity toward head flowers varied widely from strongly positive to strongly negative.
-
-Differences between floral morphotypes in terms of raw mean visitation rates, reflecting the target use hypothesis, were less pronounced **(Figure 3B)**.
+Differences between floral morphotypes in terms of raw mean visitation rates, reflecting the target use hypothesis, were less pronounced **(Figure 2B)**.
 Bell, lip, flag, and head flowers were visited at high rates by one or more bumble bee species, but median raw mean visitation rates did not differ strongly across morphotypes.
 
-The omission of queens from the calculation of overall selectivity did not yield qualitatively different results, so we present only the results of the full data set.
-The results of the worker-only analysis are included in Appendix X.
+The omission of queens from the calculation of overall preference did not yield qualitatively different results (Appendix X).
 
-<!-- I could do some sort of linear mixed effects model on this I suppose, but I think it would violate assumptions due to the non-independence of selectivity (selectivity for one thing implies selectivity against others).
+<!-- I could do some sort of linear mixed effects model on this I suppose, but I think it would violate assumptions due to the non-independence of preference (preference for one thing implies preference against others).
 I've already sort of done stats on this with the null model.
@@
 
-```{r fig3, echo = FALSE,  fig.cap = paste("Selectivity (A) and raw visitation rate (B) of the bumble bee species toward each floral morphotype.
-Each point represents the selectivity/visitation of one bumble bee species toward the corresponding floral morphotype.
+```{r fig2, echo = FALSE,  fig.cap = paste("Preference (A) and observed visitation rate (B) of the bumble bee species toward each floral morphotype.
+Each point represents the preference/visitation of one bumble bee species toward the corresponding floral morphotype.
 The color of each point in panel A, retained from Figure 2, indicates whether a given observed bias is significantly (p < 0.01) positive or negative with respect to null model simulations.")}
@@
 
-Tongue length was a strong predictor of selectivity but not of raw mean visitation rates **(Figure 4)**.
-Selectivity toward funnel, bell, lip, flag, and head flowers exhibited significant GAM curves in response bumble bee tongue length, and the model explained 77% of total deviance (adj.
+Tongue length was a strong predictor of preference but not of observed visitation rates **(Figure 3)**.
+preference toward funnel, bell, lip, flag, and head flowers exhibited significant GAM curves in response bumble bee tongue length, and the model explained 77% of total deviance (adj.
 R^2^ = 0.728).
-For funnel and lip flowers, the response of selectivity to tongue length was positive and approximately linear.
-Selectivity toward bell flowers and head flowers exhibited approximately quadratic responses with selectivity peaking among short-tongued species and declining sharply with increasing tongue length before leveling off.
-Selectivity toward flag flowers exhibited a unimodal peak among medium-tongued species.
-In contrast, raw mean visitation rates were significantly predicted by tongue length only for head flowers, which exhibited a weakly quadratic decline in selectivity from a peak among short-tongued species.
-The raw mean visitation rate model explained only 25% of total deviance (adj.
+For funnel and lip flowers, preference increased linearly with tongue length.
+Preference toward bell flowers and head flowers exhibited approximately quadratic responses with preference peaking among short-tongued species and declining sharply with increasing tongue length before leveling off.
+preference toward flag flowers exhibited a unimodal peak among medium-tongued species.
+In contrast, observed visitation rates were significantly predicted by tongue length only for head flowers, which exhibited a weakly quadratic decline in preference from a peak among short-tongued species.
+The visitation rate model explained only 25% of total deviance (adj.
 R^2^ = 0.053).
 
-Again, these results did not differ qualitatively when selectivity was based on the worker-only null model, so we present only the results of the full data set.
-The results of the worker-only analysis are included in Appendix X.
+Again, these results did not differ qualitatively when preference was based on the worker-only null model (Appendix X).
 
-```{r fig4, echo = FALSE,  fig.cap = paste("Selectivity (A) and raw visitation rates (B) for each morphotype regressed against bumble bee tongue length.
-Each point represents the selectivity/visitation of one bumble bee species toward the corresponding floral morphotype.
+```{r fig3, echo = FALSE,  fig.cap = paste("preference (A) and observed visitation rates (B) for each morphotype regressed against bumble bee tongue length.
+Each point represents the preference/visitation of one bumble bee species toward the corresponding floral morphotype.
 Curves depict fitted GAM smooths with 95\\% confidence bands.
@@
 
-Interestingly, selectivity and raw mean visitation rates produced broadly similar ordinations, each of which captured more than 80% of total variance **(Figure 5)**.
-In both ordinations, GAM smooths on first two axes also explained most of the variance in tongue length (selectivity mode: deviance explained = 94.7% (adj.
-R^2^ = 0.893), p < 0.001; raw visitation model: deviance explained = 90.8% (adj.
-R^2^ = 0.859), p < 0.001).
-We stress, however, that the two ordinations are not directly comparable in statistical terms since they involve different underlying ordination techniques, necessitated by the fact that selectivity is bounded on the interval of (-1, 1) while mean visitation rate is bounded on the interval (0, 1).
-We present the ordinations primarily for the sake of visualizing the similarity between bumble bee species in terms of selectivity and raw mean visitation rates toward the floral morphotypes considered.
+PCA of preference captured more than 80% of total variance in the first two axes **(Figure 4)**, and the GAM surface fit to the firs two axes explained most of the variance in tongue length (deviance explained = 94.7%, adj.
+R^2^ = 0.893, p < 0.001).
 
-```{r fig5, echo = FALSE,  fig.cap = paste("Principal components biplot of bumble bee species according to their selectivity toward each floral morphotype.
+<!-- Need to add a bit of comment here about important patterns in the ordination. -->
+
+Perhaps the most striking pattern
+
+```{r fig4, echo = FALSE,  fig.cap = paste("Principal components biplot of bumble bee species according to their preference toward each floral morphotype.
 Floral morphotypes are abbreviated with their numerical codes and illustrated as in Figure 1.
@@
 
-## 3.5 Variation of floral abundance, raw visitation, and selectivity through elevation and time
+## 3.5 Variation of floral abundance, raw visitation, and preference through elevation and time
 
-The proportional abundance of the most visited floral morphotypes --- bell, lip, flag, and head flowers --- was strongly patterned in time and elevation **(Figure 6)**.
+The proportional abundance of the most visited floral morphotypes --- bell, lip, flag, and head flowers --- was strongly patterned in time and elevation **(Figure 5)**.
 Bell flowers peaked in proportional abundance early in the season and at mid-elevation, and lip flowers peaked late in the season and at low elevation.
@@
 
-```{r fig6, echo = FALSE,  fig.cap = paste("Variation in proportional flower cover through elevation and modeled as HGAM surfaces.
+```{r fig5, echo = FALSE,  fig.cap = paste("Variation in proportional flower cover through elevation and modeled as HGAM surfaces.
 Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns.
@@
 
-Patterns of raw visitation **(Figure 7)** generally mirrored patterns of proportional abundance **(Figure 6)** for each morphotype.
-Given the expected dependency between floral abundance and raw visitation, this is not surprising.
-There were, however, some notable deviations.
-For example, *B.
-soroeensis* exhibited a strong peak of visitation to bell flowers late in the year at the highest sites, outside the peak abundance of bell flowers.
-*B.
-pratorum* was unique in not exhibiting high visitation rate to lip flowers during their peak abundance late in the season and at low-elevation.
-Bumble bee species differed in the degree to which they exploited the mid- and high-elevation peaks of flag flower abundance; *B.
-hortorum* and *B.
-wurflenii* visited flag flowers heavily at high elevation, while *B.
-pascuorum* --- which also visited flag flowers heavily --- tended rather to exploit the low-elevation peak of abundance.
-The remaining species --- which, based on selectivity analysis, appear to have a low affinity for flag flowers **(Figure 2)** --- exhibited diffuse or nonsignficiant patterns of visitation rates that appear uncorrelated with the abundance of flag flowers.
-
-Smooths were significant (p < 0.05) for all bb-fl pairs except hort-bell, hort-head, pasc-bell, telu-head, and soro-flag. The model explained 33.3% of total deviance (adj. R^2^ = 0.272).
-
-```{r fig7, echo = FALSE,  fig.cap = paste("Variation in raw visitation rates through elevation and time modeled as HGAM surfaces.
-Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns.
-Significant (p < 0.05) smooths are depicted with solid contour lines and nonsignificant smooths with dashed contour lines.")}
-knitr::include_graphics("../ms_2/figures/raw_visitation_gam.png")
-```
-
-Patterns of selectivity through elevation and time were more complex than those of raw visitation and, by design, more often differed from patterns of floral abundance **(Figure 8)**.
-In some cases, though, patterns of selectivity mirrored patterns of floral availability, suggesting that bumble bees may exhibit functional responses [@Mysterud1998-rs] to floral abundance not entirely accounted for by null models that assume a linear proportionality between abundance and visitation.
+Patterns of preference through elevation and time were more complex than those of raw visitation and, by design, more often differed from patterns of floral abundance **(Figure 6)**.
+In some cases, though, patterns of preference mirrored patterns of floral availability, suggesting that bumble bees may exhibit functional responses [@Mysterud1998-rs] to floral abundance not entirely accounted for by null models that assume a linear proportionality between abundance and visitation.
 
@@
 pratorum* and *B.
-soroeensis*) **(Figure 2)**.
-It is also apparent that, as a rule, selectivity is more dynamic where it is also higher overall.
+soroeensis*) **(Figure 1)**.
+It is also apparent that, as a rule, preference is more dynamic where it is also higher overall.
 Bell flowers, which were positively selected overall by the short-tongued bumble bee species *B.
@@
 terrestris-lucorum*, and *B.
-soroeensis* **(Figure 2)**, also exhibited the most complex patterns of selectivity in elevation and time for these bumble bee species.
+soroeensis* **(Figure 1)**, also exhibited the most complex patterns of preference in elevation and time for these bumble bee species.
 In contrast, the long and medium-tongued species *B.
@@
 pascuorum*, and *B.
-wurflenii*, which exhibited neutral or weakly positive overall selection toward bell flowers **(Figure 2)**, also show weak patterns of variation in selectivity toward bell flowers.
+wurflenii*, which exhibited neutral or weakly positive overall selection toward bell flowers **(Figure 1)**, also show weak patterns of variation in preference toward bell flowers.
 Nearly the reverse pattern is evident for flag flowers, reflecting their association with longer-tongued bumble bee species.
@@
 
-```{r fig8, echo = FALSE,  fig.cap = paste("Variation in selectivity through elevation and time modeled as HGAM surfaces.
+```{r fig6, echo = FALSE,  fig.cap = paste("Variation in preference through elevation and time modeled as HGAM surfaces.
 Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns.
@@
 Given the well-established mechanistic link between bumble bee tongue length and floral selection [e.g.
-@Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn], our finding that bumble bee tongue length explained selectivity (i.e.
+@Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn], our finding that bumble bee tongue length explained preference (i.e.
 null model residuals) much better than it explained raw visitation rates indicates that, in our study system, abundance and preference function jointly in bumble bee floral choice.
@@
 In general, though, our findings support the hypothesis that bumble bee resource selection involves both the neutral effects of abundance and the directed effect of preference, which implies that analytical techniques designed to disentangle these effects, such as the null model analysis we employ, should be favored in studies of bumble bee foraging.
-<!-- Accordingly, we shall focus primarily on our calculations of selectivity, rather than raw visitation rates, as we interpret the other results of our study. -->
+<!-- Accordingly, we shall focus primarily on our calculations of preference, rather than raw visitation rates, as we interpret the other results of our study. -->
 
-## 4.2 Overall patterns of selectivity
+## 4.2 Overall patterns of preference
 
@@
 These results are broadly consistent with the raw visitation patterns previously reported by @Schneller2014-ry in the central Alps at Vorarlberg, Neumayer1999-il in the Austrian Alps ~55 km south of our study system, and Farnberger1996-xo in the vicinity of Vienna, suggesting that our findings are at least regionally generalizeable.
-These neighboring studies tended to find a stronger role of head flowers than suggested by our selectivity analysis, but this is probably because they characterized foraging in terms of raw visitation rather than selectivity.
-Indeed, our raw visitation data similarly indicate a stronger role of head flowers, but our selectivity analysis reveals that while head flowers were highly visited, they tended to be under-visited relative to their abundance.
+These neighboring studies tended to find a stronger role of head flowers than suggested by our preference analysis, but this is probably because they characterized foraging in terms of raw visitation rather than preference.
+Indeed, our raw visitation data similarly indicate a stronger role of head flowers, but our preference analysis reveals that while head flowers were highly visited, they tended to be under-visited relative to their abundance.
 
@@
 
-While bell, lip, and flag flowers were the most selected morphotypes on average, selectivity toward these morphotypes varied broadly across bumble bee species.
-Moreover, while average selectivity toward funnel and head flowers was low, each received strongly positive selection from one or two bumble bee species.
-The majority of variation in selectivity toward floral morphotypes was explainable by bumble bee tongue length, with the more open bell and head morphotypes selected more positively by short-tongued bumble bee species and the deeper funnel, lip, and flag morphotypes selected more positively by medium- and long-tongued bumble bee species.
-The failure of tongue length to explain selectivity toward disc and stalk-disc flowers reflects the fact that these flowers were almost unanimously ignored by all bumble bee species.
+While bell, lip, and flag flowers were the most selected morphotypes on average, preference toward these morphotypes varied broadly across bumble bee species.
+Moreover, while average preference toward funnel and head flowers was low, each received strongly positive selection from one or two bumble bee species.
+The majority of variation in preference toward floral morphotypes was explainable by bumble bee tongue length, with the more open bell and head morphotypes selected more positively by short-tongued bumble bee species and the deeper funnel, lip, and flag morphotypes selected more positively by medium- and long-tongued bumble bee species.
+The failure of tongue length to explain preference toward disc and stalk-disc flowers reflects the fact that these flowers were almost unanimously ignored by all bumble bee species.
 We cannot say from our data why disc and stalk-disc flowers received such consistently low visitation, but the pattern may be driven in part by the fact that some disc flowers offer pollen but little or no nectar; indeed, Kugler's (1970) typology recognizes this distinction with subdivisions of the disc morphotype into those with (class 1.2) and without (class 1.1) nectar rewards.
@@
 
-The strong explanatory power of tongue length with respect to selectivity in our study corroborates the established link between tongue length and floral choice [e.g.
+The strong explanatory power of tongue length with respect to preference in our study corroborates the established link between tongue length and floral choice [e.g.
 @Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn].
-It is important, though, to consider the species that appear as outliers with respect to the smooth dependency of selectivity on tongue length.
+It is important, though, to consider the species that appear as outliers with respect to the smooth dependency of preference on tongue length.
 *B.
-terrestris/lucorum* showed comparatively high (though still negative) selectivity toward disc flowers.
+terrestris/lucorum* showed comparatively high (though still negative) preference toward disc flowers.
 *B.
-mucidus* exhibited unusually low selectivity (though not significantly different from neutrality) toward lip flowers, complemented by unusually high selectivity toward flag flowers.
+mucidus* exhibited unusually low preference (though not significantly different from neutrality) toward lip flowers, complemented by unusually high preference toward flag flowers.
 *B.
@@
 psythyrus* species-group were anomalous in their strongly positive selection toward head flowers.
-Such unexplained variation in selectivity may indicate that traits not captured by Kugler's (1970) morphotypology are also important in governing floral selection by bumble bees.
+Such unexplained variation in preference may indicate that traits not captured by Kugler's (1970) morphotypology are also important in governing floral selection by bumble bees.
 Indeed, one such trait --- floral color --- is known to be instrumental in the process of flower detection by foraging bumble bees [@Spaethe2001-bq], and some bumble bee species exhibit innate color preferences [@Raine2007-nq].
@@
 @Teras1976-pi; @Neumayer1999-il].
-Moreover, since floral color can be correlated with floral morphology [@Teras1976-pi], it is possible that some of the selectivity attributed to morphology in our models could represent a latent effect of floral color.
+Moreover, since floral color can be correlated with floral morphology [@Teras1976-pi], it is possible that some of the preference attributed to morphology in our models could represent a latent effect of floral color.
 
 As mountain bumble bees can be expected to face a future of declining resource availability and altered elevational distributions, it is relevant to ask which species seem poised for the greatest resource conflict should competitive conditions arise.
-While any answer to this question should be regarded as speculative, considering the similarities of morphotype selectivity between species is a good place to start.
+While any answer to this question should be regarded as speculative, considering the similarities of morphotype preference between species is a good place to start.
 As a general rule, it can be said that species of similar tongue length will be, on average, more likely to compete for resources than species of dissimilar tongue length.
@@
 soroeensis* cluster tightly around head flowers.
-Medium- and long-tongued species differ in their allocation of selectivity across funnel, lip, and flag flowers.
+Medium- and long-tongued species differ in their allocation of preference across funnel, lip, and flag flowers.
 *B.
@@
 Broaching the topic of competition raises an issue that can confound any observational study of resource selection: how to disentangle the endogenous process of resource preference from the exogenous process of niche partitioning driven by competition [@Pianka1974-qk].
-Empirical selectivity could reflect a latent process of resource preference, as we assume in our interpretation.
-Alternatively, though, visitation patterns and the selectivity inferred from them could be distorted by competitive forces.
+Empirical preference could reflect a latent process of resource preference, as we assume in our interpretation.
+Alternatively, though, visitation patterns and the preference inferred from them could be distorted by competitive forces.
 Intraspecific competition in a bumble bee community would tend to force individuals of a species to broaden their diet beyond what they would prefer under non-competitive conditions, while interspecific competition would be expected to cause a narrowing of diet breadth toward resources that can be exploited most efficiently.
-Thus, observed visitation patterns, even when calculated as abundance-relative selectivity, can only be interpreted in terms of *preference* to the extent that competition can be assumed to be negligible, and assumption must be justified on the basis of additional knowledge of the study system under consideration.
+Thus, observed visitation patterns, even when calculated as abundance-relative preference, can only be interpreted in terms of *preference* to the extent that competition can be assumed to be negligible, and assumption must be justified on the basis of additional knowledge of the study system under consideration.
 Our study system in the Berchtesgadener Alps is one of abundant floral resources but harsh abiotic conditions, so it is likely that competition for floral resources is low, or at least transient, and that bumble bee populations are limited instead by the shortness of the growing season, as has been suggested in the similar environment of the Eyne Valley of France [@Iserbyt2008-jn].
@@
 
-## 4.3 Variability of floral abundance, visitation, and selectivity through elevation and time
+## 4.3 Variability of floral abundance, visitation, and preference through elevation and time
 
-The intraspecific lability of bumble bee foraging preferences (*sensu stricto*) has been demonstrated under laboratory conditions [@Ings2009-vc; @Maharaj2019-wm]; our study reveals the *in situ* lability of selectivity for multiple bumble bee species through elevational and temporal gradients of floristic turnover.
-While we stress the limitations of interpreting empirical selectivity in terms of behavioral preference without controlling the joint abundances of all available resources [@Beyer2010-vr], interpreting the dynamics of selectivity in our study as the adaptive tuning of preference is consistent with the ecological hurdles bumble bees face as characteristically --- and most likely primordially --- mountain species [@Williams1985-ql; @Hines2008-er].
+The intraspecific lability of bumble bee foraging preferences (*sensu stricto*) has been demonstrated under laboratory conditions [@Ings2009-vc; @Maharaj2019-wm]; our study reveals the *in situ* lability of preference for multiple bumble bee species through elevational and temporal gradients of floristic turnover.
+While we stress the limitations of interpreting empirical preference in terms of behavioral preference without controlling the joint abundances of all available resources [@Beyer2010-vr], interpreting the dynamics of preference in our study as the adaptive tuning of preference is consistent with the ecological hurdles bumble bees face as characteristically --- and most likely primordially --- mountain species [@Williams1985-ql; @Hines2008-er].
 Where steep temperature gradients and short growing seasons generate extreme spatiotemporal turnover of floral resources, adaptive flexibility of morphological preferences should be favored over fixed innate biases.
 
-At least three mechanisms could underlie the patterns of selectivity through elevation and time that we observed.
-To illustrate these hypotheses, we will take an as example the relatively strong selectivity dynamics of flag flowers in our system, which tended to be most preferred at high-elevation sites **(Figure 8)**.
+At least three mechanisms could underlie the patterns of preference through elevation and time that we observed.
+To illustrate these hypotheses, we will take an as example the relatively strong preference dynamics of flag flowers in our system, which tended to be most preferred at high-elevation sites **(Figure 8)**.
 Our study does not enable the testing of these hypothesized mechanisms (and their constituent submechanisms), but we articulate them for the sake of clarity and as a guide for future investigations.
@@
 
-First, variation in selectivity toward a given morphotype could be driven by changes in the *intrinsic* reward of that focal morphotype.
+First, variation in preference toward a given morphotype could be driven by changes in the *intrinsic* reward of that focal morphotype.
 These changes might, in turn, be driven by (1) shifts in species composition of the focal morphotype (e.g.
@@
 an increased need for concentrated nectar to support thermoregulation at colder ambient temperatures).
-The hypothesis is consistent with the interpretation of selectivity as an index of the latent process of absolute preference.
+The hypothesis is consistent with the interpretation of preference as an index of the latent process of absolute preference.
 
-Second, selectivity toward a given morphotype could vary due to changes in the *relative* reward of the focal morphotype under changing background resource conditions.
+Second, preference toward a given morphotype could vary due to changes in the *relative* reward of the focal morphotype under changing background resource conditions.
 For example, it could be that flag flowers are selected more positively at high elevation because they must compete for visitation with highly attractive lip flowers at lower elevation.
@@
 
-<!-- Patterns of selectivity through elevation and time toward major floral morphotypes tended to be similar across the six dominant bumble bee species we  -->
+<!-- Patterns of preference through elevation and time toward major floral morphotypes tended to be similar across the six dominant bumble bee species we  -->
 
@@
 
-<!-- The late high elevation peak of visitation and selectivity between soro and bell flowers is probably driven by soro's peculiar affinity for Campanula.  -->
+<!-- The late high elevation peak of visitation and preference between soro and bell flowers is probably driven by soro's peculiar affinity for Campanula.  -->
 
-<!-- Something about the apparent link between overall selectivity and its variability? -->
+<!-- Something about the apparent link between overall preference and its variability? -->
 
-<!-- <!-- We reiterate, however, the inherent problems of comparing selectivity --- understood as an index of the latent process of preference --- across different contexts of resource abundance.
-We recommend the cautious interpretation of selectivity patterns through elevation and time in conjunction with the underlying patterns of floral abundance and raw visitation.
+<!-- <!-- We reiterate, however, the inherent problems of comparing preference --- understood as an index of the latent process of preference --- across different contexts of resource abundance.
+We recommend the cautious interpretation of preference patterns through elevation and time in conjunction with the underlying patterns of floral abundance and raw visitation.
 --> -->
@@
 A central task of ecology is to understand the patterns that will be acted upon and predict the ecological ramifications of the patterns that will be created.
-In this regard, it is significant that we find marked structure in the morphotypal composition, raw visitation, and selectivity of the floral community on which the bumble bees of the Berchtesgadener Alps depend.
+In this regard, it is significant that we find marked structure in the morphotypal composition, raw visitation, and preference of the floral community on which the bumble bees of the Berchtesgadener Alps depend.
 While our data permit only speculation as to the climate and land use driven processes at work in our system, our speculation can be nonetheless well-informed.

### 2021-07-15 ms_2/ms_2_supp_methods.Rmd
revision

--- 2021-06-11 ms_2/ms_2.Rmd
@@
 ---
-title: "Floral selectivity in mountain bumble bees and a test of the 'target use hypothesis'"
+title: "Floral preferences of mountain bumble bees are driven by trait matching but labile through elevation and time"
 
-author: Douglas B. Sponsler^1^*, Katharina Kallnik^1^, Fabrice Requier^2^, Alice Claßen^1^, A. Fabienne Maihoff^1^, Johanna Sieger^3^, Ingolf Steffan-Dewenter^1^
+author: Douglas B. Sponsler^1^*, Katharina Kallnik^1^, Fabrice Requier^2^, Alice Classen^1^, A. Fabienne Maihoff^1^, Johanna Sieger^3^, Ingolf Steffan-Dewenter^1^
 
@@
 
-Understanding patterns of resource use by animals, and the underlying selection processes that govern these patterns, can clarify how ecological communities have assembled, how they currently function, and how they are likely to respond to future perturbations.
-Communities of bumble bees and their floral hosts provide a tractable and compelling system in which to study questions of resource use, and the answers to these questions have immediate relevance to the conservation of bumble bees and their mutualists under the forcings of climate and land use change, particularly in mountain ecosystems.
+Patterns of resource use by animals can clarify how ecological communities have assembled in the past, how they currently function, and how they are likely to respond to future perturbations.
+Using ~13,000 records of bumble- bee-wildflower interactions along a 1400 m elevation gradient, we find that the selection of floral morphotypes by bumble bees is driven both by the directed process of trait-matching and the neutral effects of floral abundance.
+Patterns of floral preference are, moreover, labile through time and elevation, which may be an adaptation to the extreme spatiotemporal floristic turnover of mountain ecosystems.
+The flexibility of bumble bee foraging will be tested in coming years as climate and land use change drive rapid changes in floral resource distribution in time and space, particularly in the mountain ecosystems that host the world's richest communities of bumble bees and their floral mutualists.
 
-Using a dataset consisting of ~13,000 observations collected along a 1400 m elevation gradient over three consecutive years, we
-
-Significance ...
-
-*Keywords*: foraging, trait matching, pollinator, *Bombus*, alpine
+*Keywords*: foraging, trait matching, pollinator, *Bombus*, alpine, niche, resource selection
 
@@
 
-# 1. INTRODUCTION
-
-The process of resource selection by animals has been an enduring theme of ecological research [@Johnson1980-co; @Beyer2010-vr].
-Methodological difficulties notwithstanding, the study of animals' resource choices is one of the central tasks of community ecology, where the selection and partitioning of resources are instrumental in community assembly, species coexistence, and ecosystem function [@Chesson2000-aw; @Bascompte2007-ls].
-Recently, questions of resource selection in ecological communities have taken on new urgency on account of anthropogenic climate and land use change, which threaten to disrupt existing patterns of species interactions and niche relationships [@Tylianakis2008-fi].
-
-Bumble bees (Hymentoptera: *Bombus* spp.) and their floral hosts provide a diverse yet tractable system in which to explore resource selection.
-Bumble bees are exceptionally versatile florivores whose ability to solve the puzzles posed by complex flower morphologies [@Plowright1984-lc] --- or, indeed, by human investigators [@Alem2016-rn; @Loukola2017-bt] --- renders almost any flower potentially accessible.
-Nevertheless, the flexibility of bumble bee foraging is tuned to floral reward, which depends on both the intrinsic reward of a given flower and the costs of exploiting it.
-These costs are closely related to the degree of morphological fittedness between bumble bees and their floral hosts, which consists principally in the matching of bumble bee tongue length to floral corolla depth.
-Bumble bee species can differ markedly in tongue length, and floral handling time increases when tongue length and corolla depth are poorly matched [@Inouye1980-mv; @Harder1983-bk], leading to a well-documented association between tongue length and floral choice in bumble bee communities [e.g.
-@Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn].
-This association between tongue length and floral niche is imperfect, though, and there can be both convergence in floral resource use between bumble bees of different tongue length and divergence in floral resource use between bumble bees of similar tongue length [e.g.
-@Ranta1980-ok; @Goulson2008-oe].
-Moreover, bumble bee floral choice is also influenced by ecological modulators, such as the spatial distribution of floral patches or competition with other bumble bees [@Ishii2013-sm; @Plowright1984-lc; @Sowig1989-ld].
-
-Floral choice in bumble bee communities can serve as a case study in the general phenomenon of animal resource selection, but it is also a matter of practical importance.
-Bumble bees are sensitive to climate and land use change [@Goulson2008-oe; @Soroye2020-zb], especially in the mountain ecosystems that host the world's richest bumble bee communities.
-Warming climates have already caused measurable upslope shifts in mountain bumble bees and their floral hosts [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc], which could introduce novel competitive interactions between bumble bee species that experience increased sympatry.
-At the same time, socioeconomic factors have led to a dramatic decline in traditional alpine grazing management, particularly in the European Alps, resulting in widespread pasture abandonment and reforestation of mountain meadows [@Tappeiner2006-jp; @Tasser2007-qo].
-The regression of meadows to forest entails a massive loss of floral resource availability, and even early stages of forest succession can cause declines in floral abundance and diversity, with corresponding effects on bumble bees [@Iserbyt2015-jc; @Walcher2017-kw; @Suzuki-Ohno2020-jv].
-In contrast, more accessible areas at the bases of mountains are undergoing agricultural intensification that displaces natural vegetation and favors grasses over forbs [@Tasser2002-om].
-Caught between adverse land use changes above and below and subject to climate-induced range shifts, mountain bumble bees are facing a future of reduced floral resource availability and altered sympatry.
-The ability of bumble bee species to coexist under these changes will likely depend significantly on how they select and partition floral resources, and analogous considerations apply to their floral mutualists, many of which depend principally on bumble bee visitation for pollination service [@Goulson2010-lo].
-
-In this study, we analyze a data set consisting of ~13,000 interactions between bumble bees and wildflowers recorded over 3 years along a 1400 m elevation gradient.
-Using these data in conjunction with complementary floral survey data, we explore the *selectivity* of the bumble bee community at the level of floral morphotype [@Kugler1970-gb].
-Following @Beyer2010-vr, we define *selectivity* simply as any disproportionality --- positive or negative --- between the relative abundance of a resource and its rate of use.
-This concept approaches the idea of resource *preference*, but inference of the latter, in an absolute sense, requires either experimental control of all resource abundances [@Johnson1980-co] or statistical techniques that condition visitation simultaneously on all resource abundances [@Aarts2008-fo].
-Given the scale and complexity of our observational study, we opt for simpler analytical approaches and restrict our inference to the phenomenon of selectivity, which can be understood as a metric of *relative* preference for a focal resource given unmodeled interactions with preferences for alternative resources.
-We deem these interactions beyond the scope of our study, but acknowledge that this requires us to remain partly agnostic as to the process underlying the empirical selectivity we describe.
-
-The first objective of our study is to describe and explain the selectivity of the bumble bee community with respect to floral morphotype.
-Using null model analysis to remove neutral effects of resource abundance, we compare the selectivity of the bumble bee community across floral morphotypes and then model the species-wise selectivity for each morphotype as a function of bumble bee tongue length.
-
-We then revisit an open question regarding the role of resource abundance as a driver of flower selection in bumble bees.
-In studies of resource selection, it is common to pose an abundance-only hypothesis in which foraging is random and selection merely a function of frequency of encounter, resulting in resource use patterns that mirror resource abundance patterns.
-For bumble bees, which are known to forage non-randomly, the abundance-only hypothesis can serve as a null-model against which observed use patterns can be compared to infer selectivity (often called "preference", following a *sensu lato* usage of the term) [@Vaughan2018-gj].
-It has also be proposed, however --- in what has been termed the *target use hypothesis* --- that bumble bees might select resources irrespective of abundance [@De_Valpine2013-be; @Harmon-Threatt2017-bz].
-This hypothesis, while obviously subject to boundary conditions of floral availability, is consistent with the anecdotal observation that bumble bees sometimes persist in visiting favored flower species even when these resources are experimentally reduced to low abundance [@Heinrich1976-vb], and it was supported in a study comparing competing models of flower selection in *Bombus vosnesenskii* [@Harmon-Threatt2017-bz].
-If and when the *target use hypothesis* holds, raw visitation rates (i.e.
-raw for abundance) are a better index of preference (*sensu stricto*) than selectivity (i.e.
-the residuals of an abundance-only null model), since the subtraction of presumed neutral effects of abundance when those effects do not, in fact, exist actually creates new abundance-dependent artifacts and obscures the true process of resource selection [@Harmon-Threatt2017-bz].
-On the other hand, failing to account for neutral effects of abundance when they do obtain similarly obscures the selection process underpinning foraging patterns [@Vazquez2003-qa; @Bluthgen2008-rp].
-In our study, we take advantage of the fact that there exists a well-established mechanistic basis for bumble bee flower choice, namely the compatibility between floral morphology and bumble bee tongue length.
-Taking this mechanism as a reference and exploiting the tongue-length diversity of the bumble bee species in our study system, we evaluate models of preference with and without abundance effects by comparing the explanatory power of bumble bee tongue length with respect to the preferences inferred under each model.
-
-Finally, we model the variation of selectivity, raw visitation, and floral abundance through time and elevation.
-This analysis reveals the elevational and phenological dynamics of the underlying floral community as well as the lability of bumble bee foraging behavior.
-Exploring these patterns is the first step toward understanding how climate and land use change might distort the temporal and elevational structure of bumble- bee-wildflower mutualisms in mountain ecosystems.
-
-# 2. METHODS
-
-## 2.1 Field sampling
-
-The samples analyzed in this paper have been previously analyzed (Sponsler et al., submitted), but the analyses presented here are non-overlapping.
-Field sampling was conducted 2010-2012 in Berchtesgaden National Park (47.55°N, 12.92°E), located in the Northern Limestone Alps of southeast Germany.
-The landscape is composed of a coniferous forest matrix embedded with pastures either currently or historically maintained by extensive (i.e.
-low intensity) agricultural management.
-We selected a total of 25 study sites (60 x 60 m) located on pastures along a ~1400 m elevation gradient ranging from 641-2032 m above sea level (m.s.l.).
-Fourteen of these pastures are grazed by cattle or sheep, three are mowed for hay production, and eight have been abandoned throughout the last century and are no longer subject to any human management.
-
-Sampling of bumble- bee-wildflower interactions consisted of repeated visits (hereafter, “visitation samples”) to each study site at approximately weekly intervals.
-Sampling was only conducted during periods without rain and with an air temperature of at least 6°C.
-Bumble bees were recorded during one 50-minute transect walk per visit, and bumblebees observed on a given flower were counted as floral visitors.
-Queens were identified to species level in the field to avoid destructively sampling colony foundresses, but workers and males were collected and identified in the lab following @Amiet1996-jz.
-While male bumble bees were recorded, we chose to analyze only the pooled visitation of queens and workers.
-In conjunction with bumble bee observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.01 m^2^ (hereafter "floral survey samples") and identified the plant species according to @Lauber2007-ha and @Oberndorfer2001-ro.
-While per-sample effort was standardized across all years, the total number of samples varied: 153 in 2010, 212 in 2011, and 539 in 2012.
-The start of field sampling each year was determined by snowmelt, so lower sites were sampled as early as April while higher sites could not be sampled until June or July.
-In 2010 and 2012, this resulted in a temporally staggered pattern of sampling, but a combination of early snowmelt and a late start to field activities in 2011 resulted in sampling that was approximately aligned in time across the whole elevation gradient.
-See Appendix X for a full description of sampling patterns.
-
-## 2.2 External data sources
-
-### 2.2.1 Floral morphology
-
-Rather than considering various floral morphological traits independently, we relied on the synthesis of @Kugler1970-gb, who developed a typology consisting of ten primary morphotypes and nested subdivision of each.
-Kugler’s typology was informed by laboratory studies of bumble bee flower choice [@Kugler1943-uy], and there is a precedent for analyzing bumble bee visitation in terms of Kugler’s morphotypes [@Farnberger1996-xo; @Neumayer1999-il; @Schneller2014-ry].
-
-Kugler morphotype classifications of the plants in our study system were accessed via the BIOFLOR database [@Klotz2002-uc].
-Morphotypes were simplified to the primary classes numbered 0-10: 0 = null (includes wind-pollinated flora such as grasses), 1 = disc- and bowl-shaped flowers, 2 = funnel flowers, 3 = bell-shaped flowers, 4 = stalk-disc flowers, 5 = lip flowers, 6 = flag flowers, 7 = flower heads, 8 = spike flowers, 9 = brush flowers, 10 = trap flowers **(Figure 1)**.
-In a few cases, a plant species in our data set was not found in the BIOFLOR database.
-In most cases, morphotypes are constant at the genus level, and missing morphotypes were added based on the morphotypes found for congeners.
-In the case of the genus *Phyteuma*, we followed @Neumayer1999-il and classified it as a group 7 flower head.
-Morphotypes 9 and 10 were extremely rare in our study system and morphotype 8 was completely absent, so these morphotypes were omitted from all analyses.
-Plants belonging to morphotype 0 were typically not recorded during floral surveys due to their lack of conspicuous floral displays, and they were rarely visited by bumble bees, so this morphotype was also omitted from all analyses.
-
-### 2.2.2 Bumble bee tongue length
-
-Bumble bee tongue-length data (for the worker caste) were obtained from the literature values compiled and summarized by @Arbetman2017-xj.
-Values for *B.
-gerstaeckeri* and *B.
-mendax*, which were not included in @Arbetman2017-xj, were obtained from @Obeso1992-yp and @Durieux2000-uk, respectively.
-
-## 2.3 Data processing
-
 Prior to analysis, we aligned each visitation sample with a corresponding floral survey sample.
@@
 
-## 2.4 Analysis of resource selection
+We used 500 iterations to achieve a robust sampling distribution, and we considered observed visitation rates to indicate significant selectivity if they fell outside the interval containing 99% of simulated values.
 
-To quantify selectivity, we compared the proportional visitation of each morphotype (by each bumble bee species) with the rate of visitation expected under the null-hypothesis that visitation of a given resource is purely a function of its relative abundance [@Vazquez2003-qa; @Bluthgen2008-rp].
-When conducting a null model analysis of interaction networks that are distributed through space and/or time, it is critical to respect the sampling unit of individual sites and dates, since only within this site-date sampling unit does floral visitation data correspond unequivocally to floral abundance data.
-Naive aggregation across sites and dates introduces forbidden links and decouples visitation and abundance [@Vaughan2018-gj].
-We constructed null models following @Vaughan2018-gj.
-First, resource selection is simulated for individual consumers --- in our case, individual bumble bees --- according to a resource selection process in which the probability of a resource being chosen is proportional to its relative abundance [@Agusti2003-tt].
-By simulating the visitation of individual consumers rather than species in aggregate, the abundance of each resource, and therefore its probability of visitation, can vary across site-dates, thus preserving the within-sample correspondence between visitation and abundance.
-Simulated interactions are then summarized for each consumer species by taking the mean number of interactions with each resource across all individuals of the consumer species and all samples of the dataset.
-As this process is iterated, the result is a distribution of visitation rates of each consumer species to each resource under the null assumption that resource selection is determined only by resource abundance.
-The visitation distributions of the null model can then be compared to observed visitation rates to infer positive or negative selectivity of each consumer toward each resource.
-Since the mean of the null distribution can be interpreted as the expected value under null assumptions, the difference between the mean null value and the observed value can be interpreted as a magnitude of selectivity [@Vaughan2018-gj].
-The statistical significance of this selectivity can be estimated as the proportion of simulated null values more extreme than the observed value.
-
-We followed the procedure described above, simulating our entire dataset in one modeling process to infer aggregate pattern of selectivity between bumble bee species and floral morphotypes.
-We used 500 iterations to achieve a robust sampling distribution, and we considered observed visitation rates to indicate significant selectivity if they fell outside the interval containing 99% of simulated values.
-Then, using the results of this analysis, we modeled selectivity toward each floral morphotype as a function of bumble bee tongue length using hierarchical generalized additive modeling (HGAM), with floral morphotype specified as a smoothing factor [@Pedersen2019-kf].
-Finally, we compared morphotype selectivity across bumble bee species by performing a principal components analysis on the bumble bee species x floral morphotype selectivity matrix, and we mapped bumble bee tongue length onto the ordination space by fitting a 2-dimensional GAM surface to the first two principal components.
-
-As a test of the *target use hypothesis*, we repeated the HGAM and ordination analyses described above using raw mean visitation rates rather than selectivity.
-For the ordination of raw mean visitation rates, we used principal coordinates analysis based on the Horn-Morisita distance rather than euclidean principal components analysis because the former is better suited for proportional composition data [@Borcard2018-gb].
-
-To explore the temporal and elevational patterns in our data, we repeated the null model analysis described above but this time for each site-date individually, so that the selectivity of each bumble bee species toward each floral morphotype could be modeled as a variable through elevation and time.
-For this analysis, we included only the six most abundant and widespread bumble bee species in our study system (*B.
-hortorum*, *B.
-pascuorum*, *B.
-pratorum*, *B.
-soroeensis*, *B.
-terrestris-lucorum*, and *B.
-wurflenii*) to ensure sufficient data points through elevation and time, and we modeled their selectivity only toward the 4 most visited floral morphotypes (bell, lip, flag, and head flowers).
-Selectivity for each bumble- bee-morphotype pair (hereafter "bb-fl pair") was modeled together in a single HGAM with a tensor-product interaction smooth between elevation and time, bb-fl pair as a smoothing factor and fixed intercept effect, and the interactions site * bb-fl pair and year * bb.fl pair as random intercept effects.
-
-An inherent problem with modeling sample-wise variation in selectivity is that the bounds of possible selectivity are strictly dependent on floral abundance [@@Beyer2010-vr].
-A morphotype present at very low relative abundance would, under the null model, be expected to receive very low visitation, which bounds potential negative selectivity near zero but allows positive selectivity to approach 1.
-Conversely, a morphotype present at very high relative abundance would receive very high visitation under the null model, and its potential positive selectivity would be bounded near zero and its potential negative selectivity near -1.
-A morphotype present at exactly 50% relative abundance would have selectivity values bounded at exactly (-0.5, 0.5).
-This clearly creates a degree of unavoidable confoundedness between selectivity and floral abundance.
-With this limitation in mind, we fit HGAMs of the same form to raw visitation rates and total flower cover of each floral morphotype, and we present the results together to be interpreted jointly.
-
-## 2.5 Software
-
-All data processing and analyses were conducted in `R` [@RCoreTeam-2021aa].
-Data handling and visualization were performed with the `tidyverse` suite [@Wickham2019-aa].
-Null models were constructed using `econullnetr` [@Vaughan2018-gj].
-GAM analyses were performed with `mgcv` [@Wood_2017] and visualized with `mgcViz` [@Fasiolo2018aa] and `tidymv` [@Coretta2021-aa].
-PCA was performed with the `vegan` [@Oksanen_2019] and visualized using `ggplot2` [@Wickham2016-aa] and `ggvegan` [@Simpson2019aa].
-Annotated R code is available in the Appendix X.
-
-# 3. RESULTS
-
-## 3.1 Sampling overview
-
-A total of 12,918 bumble- bee-flower interactions were included in our final data set, including 16 bumble bee species (with *Bombus terrestric/lucorum* and *Bombus psithyrus* species lumped into species groups), 163 plant species (110 genera, 37 families), and 736 unique bumble- bee-plant interaction pairs.
-*B.
-pascuorum* ("pasc"), *B.
-pratorum* ("prat"), *B.
-soroeensis* ("soro"), *B.
-terrestris-lucorum* ("telu"), and *B.
-wurflenii* ("wurf") were the most abundant species, and they occurred at all sites.
-*B.
-hortorum* ("hort"), *B.
-jonellus* ("jone"),  *B.
-monticola* ("mont"), and *B.
-psithyrus* ("psit") occurred at moderate abundance.
-The remaining species --- *B.
-gerstaeckeri* ("gers"), *B.
-hypnorum* ("hypn"), *B.
-lapidarius* ("lapi"), *B.
-mendax* ("mend"), *B.
-mucidus* ("muci"), and *B.
-pyrenaeus* ("pyre") --- were relatively rare.
-*B.
-humilis* ("humi") was recorded only three times and only in one year, so we omitted it from all analyses.
-
-Floral surveying yielded a total of 352 plant species (191 genera, 52 families), of which 155 (103 genera, 35 families) were observed to be visited by bumble bees.
-Visited floral species included all morphotypes except spike (8) and trap (10) flowers **(Figure 1)**.
-Null (0) and brush (9) morphotypes were, however, excluded from analysis for the reasons given above.
-
-<!-- Check again and make sure all the numbers given here are still correct. -->
-
-```{r fig1, echo = FALSE,  fig.cap = paste("Visited floral morphotypes and their constituent plant families.
-Note that some families span more than on morphotype.
-Barplots show the total number of visits to each family across all bumble bee species and samples (note the rescaling of the y-axis across panels).
-Each floral morphotype is illustrated with a representative species common in our study system: 0 = Plantago lanceolata, 1 = Potentilla erecta, 2 = Gentiana asclepiadea, 3 = Campanula glomerata, 4 = Silene nutans, 5 = Ajuga reptans, 6 = Lotus corniculatus, 7 = Centaurea jacea, 8 = Salix retusa.
-Illustrations are modifications of public domain works obtained from www.plantillustrations.org.
-For details of original sources, see Appendix X.")}
-knitr::include_graphics("../ms_2/figures/morphotype_families_mod.png")
-```
-
-## 3.2 Aggregate selectivity and raw visitation
-
-Selectivity was generally strong, with the visitation rates of most bumblebee species to most floral morphotypes falling significantly above or below the 99% interval of the null model **(Figure 2)**.
-Nonsignificant selectivity was observed mainly among the rarer bumblebee species because their small samples sizes lead to wide null model confidence intervals.
-
-```{r fig2, echo = FALSE,  fig.cap = paste("Empirical visitation rates compared to null model simulations for each bumble bee species and floral morphotype.
-Gray bars depict 99\\% intervals of null model simulations and points depict observed visitation rates.
-Points are color-coded according to where they fall with respect to the 99\\% interval of the null model: red = above (positive selectivity), blue = below (negative selectivity), gray = within (nonsignificant selectivity).
-The difference between the observed value and the middle of the null model interval can be interpreted as magnitude of selectivity.")}
-knitr::include_graphics("../ms_2/figures/null_model_spwise_preference.png")
-```
-
-When floral morphotypes are compared to one another in terms of selectivity, clear patterns emerge **(Figure 3A)**.
-All species exhibited negative selectivity toward disc flowers, while most species exhibited positive selectivity toward bell, lip, and flag flowers.
-Selectivity toward funnel and stalk-disc flowers tended to be neutral (with one strong positive outlier --- *B.
-gerstaeckeri* --- for funnel flowers), reflecting the fact that these morphotypes tended both to be rarely visited and to occur at low abundance.
-Selectivity toward head flowers varied widely from strongly positive to strongly negative.
-
-Differences between floral morphotypes in terms of raw mean visitation rates, reflecting the target use hypothesis, were less pronounced **(Figure 3B)**.
-Bell, lip, flag, and head flowers were visited at high rates by one or more bumble bee species, but median raw mean visitation rates did not differ strongly across morphotypes.
-
-<!-- I could do some sort of linear mixed effects model on this I suppose, but I think it would violate assumptions due to the non-independence of selectivity (selectivity for one thing implies selectivity against others).
-I've already sort of done stats on this with the null model.
--->
-
-<!-- Before we finalize this, check to make sure the raw visitation rates were really prepared correctly and are not being deflated by a bunch of zeros that should'nt be there. -->
-
-```{r fig3, echo = FALSE,  fig.cap = paste("Selectivity (A) and raw visitation rate (B) of the bumble bee species toward each floral morphotype.
-Each point represents the selectivity/visitation of one bumble bee species toward the corresponding floral morphotype.
-The color of each point in panel A, retained from Figure 2, indicates whether a given observed bias is significantly (p < 0.01) positive or negative with respect to null model simulations.")}
-knitr::include_graphics("../ms_2/figures/overall_ktype.png")
-```
-
-## 3.4 Tongue length as a predictor of floral choice
-
-Tongue length was a strong predictor of selectivity but not of raw mean visitation rates **(Figure 4)**.
-Selectivity toward funnel, bell, lip, flag, and head flowers exhibited significant GAM curves in response bumble bee tongue length, and the model explained 77% of total deviance (adj.
-R^2^ = 0.728).
-For funnel and lip flowers, the response of selectivity to tongue length was positive and approximately linear.
-Selectivity toward bell flowers and head flowers exhibited approximately quadratic responses with selectivity peaking among short-tongued species and declining sharply with increasing tongue length before leveling off.
-Selectivity toward flag flowers exhibited a unimodal peak among medium-tongued species.
-In contrast, raw mean visitation rates were significantly predicted by tongue length only for head flowers, which exhibited a weakly quadratic decline in selectivity from a peak among short-tongued species.
-The raw mean visitation rate model explained only 25% of total deviance (adj.
-R^2^ = 0.053).
-
-```{r fig4, echo = FALSE,  fig.cap = paste("Selectivity (A) and raw visitation rates (B) for each morphotype regressed against bumble bee tongue length.
-Each point represents the selectivity/visitation of one bumble bee species toward the corresponding floral morphotype.
-Curves depict fitted GAM smooths with 95\\% confidence bands.
-Solid lines indicate curves differing significantly (p < 0.05) from a flat line while dashed lines indicated nonsigificant curves.
-Points are color-coded by species, and species are presented in the legend in order of increasing tongue length.")}
-knitr::include_graphics("../ms_2/figures/ktypeXtonguelength_paired.png")
-```
-
-Interestingly, selectivity and raw mean visitation rates produced broadly similar ordinations, each of which captured more than 80% of total variance **(Figure 5)**.
-In both ordinations, GAM smooths on first two axes also explained most of the variance in tongue length (selectivity mode: deviance explained = 94.7% (adj.
-R^2^ = 0.893), p < 0.001; raw visitation model: deviance explained = 90.8% (adj.
-R^2^ = 0.859), p < 0.001).
-We stress, however, that the two ordinations are not directly comparable in statistical terms since they involve different underlying ordination techniques, necessitated by the fact that selectivity is bounded on the interval of (-1, 1) while mean visitation rate is bounded on the interval (0, 1).
-We present the ordinations primarily for the sake of visualizing the similarity between bumble bee species in terms of selectivity and raw mean visitation rates toward the floral morphotypes considered.
-
-```{r fig5, echo = FALSE,  fig.cap = paste("Ordination biplots of bumble bee species according to their selectivity (A) and raw visitation rates (B) for each floral morphotype.
-Floral morphotypes are abbreviated with their numerical codes and illustrated as in Figure 1.
-Colored bands depict a 2-dimensional GAM smooth representing the relationship between the first two principal components and bumble bee tongue length.
-Morphotype illustrations are as described in Figure 1.")}
-knitr::include_graphics("../ms_2/figures/paired_ordination_mod.png")
-```
-
-## 3.5 Variation of floral abundance, raw visitation, and selectivity through elevation and time
-
-The proportional abundance of the most visited floral morphotypes --- bell, lip, flag, and head flowers --- was strongly patterned in time and elevation **(Figure 6)**.
-Bell flowers peaked in proportional abundance early in the season and at mid-elevation, and lip flowers peaked late in the season and at low elevation.
-Flag flowers peaked in mid-season and exhibited two distinct elevational peaks, the lower at around 1000 m.s.l and the higher at around 1800 m.s.l.
-Head flowers exhibited a mid-season peak at low elevation and a broad late season peak at around 1600 m.s.l.
-All smooths were significant (p << 0.0001), and the model explained 64.3% of total deviance (adj.
-R^2^ = 0.554).
-
-```{r fig6, echo = FALSE,  fig.cap = paste("Variation in proportional flower cover through elevation and modeled as HGAM surfaces.
-Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns.
-Significant (p < 0.05) smooths are depicted with solid contour lines and nonsignificant smooths with dashed contour lines.")}
-knitr::include_graphics("../ms_2/figures/flower_cover_gam.png")
-```
+In addition to the qualifying remarks above, it is important to acknowledge some additional limitations of our methods.
+First, the inference of selection as use-relative-to-abundance is sensitive to exactly *how* abundance is estimated [@Johnson1980-co; @Beyer2010-vr].
+Of particular concern in our study is the problem of spatial scale.
+By necessity, we measured floral abundance and visitation along transects *within* our study sites.
+Bumble bee foraging range, however, is much larger than our study sites, and it is possible that the relative abundances of floral morphotypes outside our study sites but within foraging range differ from our within-site estimates.
+Moreover, because bumble bees can forage across multiple meadow patches in a mountain forest matrix [@Mola2020-yc], there exists a prior selective process --- Johnson's (1980) "second-order selection" --- in which bumble bees choose among available foraging areas before selecting from resources within an area.
+Finally, to account for the neutral effects of abundance on floral use, we used a simple null model in which the probability of a resource being selected was assumed to be linearly proportional to the relative abundance of the resource.
+It is possible, however, for resource use to exhibit nonlinear "functional responses" @Mysterud1998-rs to resource abundance, which, relative to a linear null model, would give biased estimates of preference.
 
@@
 pascuorum* --- which also visited flag flowers heavily --- tended rather to exploit the low-elevation peak of abundance.
-The remaining species --- which, based on selectivity analysis, appear to have a low affinity for flag flowers **(Figure 2)** --- exhibited diffuse or nonsignficiant patterns of visitation rates that appear uncorrelated with the abundance of flag flowers.
+The remaining species --- which, based on preference analysis, appear to have a low affinity for flag flowers **(Figure 2)** --- exhibited diffuse or nonsignficiant patterns of visitation rates that appear uncorrelated with the abundance of flag flowers.
 
@@
 
-Patterns of selectivity through elevation and time were more complex than those of raw visitation and, by design, more often differed from patterns of floral abundance **(Figure 8)**.
-In some cases, though, patterns of selectivity mirrored patterns of floral availability, suggesting that bumble bees may exhibit functional responses [@Mysterud1998-rs] to floral abundance not entirely accounted for by null models that assume a linear proportionality between abundance and visitation.
-
-One of the more salient patterns is that positive selection for flag flowers was strongly associated with high elevation, except among the species that tended to avoid flag flowers (*B.
-pratorum* and *B.
-soroeensis*) **(Figure 2)**.
-It is also apparent that, as a rule, selectivity is more dynamic where it is also higher overall.
-Bell flowers, which were positively selected overall by the short-tongued bumble bee species *B.
-pratorum*, "*B.
-terrestris-lucorum*, and *B.
-soroeensis* **(Figure 2)**, also exhibited the most complex patterns of selectivity in elevation and time for these bumble bee species.
-In contrast, the long and medium-tongued species *B.
-hortorum*, *B.
-pascuorum*, and *B.
-wurflenii*, which exhibited neutral or weakly positive overall selection toward bell flowers **(Figure 2)**, also show weak patterns of variation in selectivity toward bell flowers.
-Nearly the reverse pattern is evident for flag flowers, reflecting their association with longer-tongued bumble bee species.
-This relationship between the complexity of selectivity dynamics and overall selectivity can be formalized by plotting the effective degrees of freedom against the selectivity HGAM surface on the corresponding overall selectivity between each **(Figure 9)**.
-
-Smooths were significant (p < 0.05) for all bb-fl pairs except hort-head, prat-flag, telu-head, and soro-flag. The model explained 36.4% of total deviance (adj. R^2^ = 0.33).
-
-```{r fig8, echo = FALSE,  fig.cap = paste("Variation in selectivity through elevation and time modeled as HGAM surfaces.
-Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns.
-Significant (p < 0.05) smooths are depicted with solid contour lines and nonsignificant smooths with dashed contour lines.")}
-knitr::include_graphics("../ms_2/figures/preference_lability_gam.png")
-```
-
-```{r fig9, echo = FALSE,  fig.cap = paste("Complexity of selectivity dynamics through elevation and time as a function of overall selectivity.
-Linear best-fit lines are shown in black with gray bands indicating 95% confidence intervals.")}
-knitr::include_graphics("../ms_2/figures/edf_selectivity.png")
-```
-
-# 4. DISCUSSION
-
-Understanding patterns of resource use by animals, and the underlying selection processes that govern these patterns, can clarify how ecological communities have assembled, how they currently function, and how they are likely to respond to future perturbations.
-Communities of bumble bees and their floral hosts provide a tractable and compelling system in which to study questions of resource use, and the answers to these questions have immediate relevance to the conservation of bumble bees and their mutualists under the forcings of climate and land use change, particularly in mountain ecosystems.
-
-## 4.1 Floral abundance, floral selection, and the *target use hypothesis*
-
-It is well-known that bumble bees forage non-randomly accordingly to morphological constraints, innate preferences, and experienced rewards [e.g.
-Kugler1943-uy; @Brian1954-cy; @Inouye1978-qi; @Raine2007-nq; @Maharaj2019-wm].
-In this regard, the results of our null model analysis are as unsurprising as they are unambiguous; bumble bee resource selection almost never represents flora in unaltered proportion to their abundance.
-The *target use hypothesis* [@De_Valpine2013-be; @Harmon-Threatt2017-bz], however, extends this idea to the perfect antithesis of the random foraging model, positing that bumble bee foraging is a process of *pure* preference, completely irrespective of resource abundance.
-
-Given the well-established mechanistic link between bumble bee tongue length and floral selection [e.g.
-@Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn], our finding that bumble bee tongue length explained selectivity (i.e.
-null model residuals) much better than it explained raw visitation rates indicates that, in our study system, abundance and preference function jointly in bumble bee floral choice.
-This contradicts the *target use hypothesis* as a general rule of bumble bee foraging.
-It does not, however, mean that the *target use hypothesis* does not obtain under certain circumstances, such as when a highly preferred resource varies within a nonlimiting rage of availability.
-Moreover, our results suggest something akin to the *target use hypothesis* but formulated negatively with respect to resources that are *not selected*.
-Visitation rates to disc flowers in our system were very low despite the fact that they often occurred at high abundance, suggesting that while preference seems to interact with resource abundance, avoidance may not.
-In general, though, our findings support the hypothesis that bumble bee resource selection involves both the neutral effects of abundance and the directed effect of preference, which implies that analytical techniques designed to parse these effects, such as the null model analysis we employ, should be favored in studies of bumble bee foraging.
-Accordingly, we shall focus primarily on our calculations of selectivity, rather than raw visitation rates, as we interpret the other results of our study.
-
-## Patterns of selectivity
-
-We found that positive selection was focused on 3 floral morphotypes: bell, lip, and flag flowers.
-These results are broadly consistent with the aggregate visitation patterns previously reported by @Schneller2014-ry in the central Alps at Vorarlberg and Neumayer1999-il in the Austrian Alps ~55 km south of our study system, suggesting that our findings are at least regionally generalizeable.
-When considering these patterns from the perspective of pollination services to mountain flora, it is important to note that while bumble bees are the dominant Hymenopteran pollinators in our study system, flies (especially Syrphid hoverflies) are also abundant flower visitors in mountain ecosystems, and their visitation is mainly restricted to the open flower morphotypes (e.g.
-disc flowers) that receive less attention from bumble bees [@Neumayer1999-il].
-In the context of our research question, we have emphasized the differences in tongue length between bumble bee species, but all bumble bees have relatively long tongues compared to the overall pollinator community, and it is not surprising that their foraging tends to be focused on the deeper flower morphotypes.
-The process of floral resource partitioning *among* bumble bee species is nested within the broader process of floral partitioning *between* bumble bees and other flower visitors.
-
-The strong explanatory power of tongue length with respect to selectivity in our study corroborates the established link between tongue length and floral choice.
-*I think it does so more strongly than most other field studyies I've read.* It is important, though, to consider the species that appear as outliers with respect to the smooth dependency of selectivity on tongue length.
-Outliers worth talking about: telu-disc(+), muci-lip(-), muci-flag(+), soro/psit-head(+).
-
-## Similarity of selectivity across species, potential for competition
-
-While competition as a phenomenon arises from realized patterns of resource use, not the latent process of selection, insight into the process of selection may be a better predictor of competition.
-Use patterns alone are, in a sense, overfit to the observations at hand.
-
-## Variability of floral abundance, visitation, and selectivity through elevation and time; its relevance to climate and land use change
-
-Elevation-time patterns of visitation closely mirrored those of floral abundance, but the former were more diffuse; this may reflect the reluctance of bumble bees to switch from familiar floral taxa even when they become rarer [@Heinrich1976-vb] or less rewarding [I think I remember reading this somewhere...but there is also a paper showing that bumble bees switch more readily than honey bees].
-
-The late high elevation peak of visitation and selectivity between soro and bell flowers is probably driven by soro's peculiar affinity for Campanula.
-Talk about idiosyncracies not capture dby morphology alone.
-importance of other traits, particularly color.
-
-## Caveats:
-
-The problem of queens vs.
-workers...
-Probably the best way to fend off this object would be to re-run all our analyses with only workers and see if any thing changes qualitatively.
-This could all go in the SI.
-
-Need to acknowledge the potential for nonlinear "functional responses" of use to abundance [@Mysterud1998-rs].
-
-Inference of preference is sensitive to the scale at which availability is estimate [@Beyer2010-vr].
-My floral survey data estimates availability more or less at the meadow scale, right?
-But I don't know what is available outside the meadow.
-This problem is also discussed in [@Harmon-Threatt2017-bz], specifically in the context of bees and flowers.
-
-The sensitivity of "preference" to availability [@Johnson1980-co; @Beyer2010-vr].
-
-*First, the estimates ofpreference  are  not  directly  interpretable  as  reflectingthe  intrinsic  behavioural  motivations  of  the  animal.Positive  (or  negative)  preference  for  a  given  habitatimplies only that the habitat is used more (or less) fre-quently than expected by chance,conditional upon thedefined sample of availability.
-Even  high  levels  of  useof a habitat can be associated with negative coefficientsif that habitat is common, and, conversely, low levelsof  use  can  be  associated  with  positive  coefficientsif the habitat is rare.
-The sign and magnitude of thecoefficient   cannot   be   interpreted   with   respect   tothe   behavioural   decision   the   animal   is   making.* [@Beyer2010-vr].
-
-*The second consequence is that estimates of prefer-ence  based  on  different  samples  of  availability  (e.g.Design III inManlyet al.2002) are not directly com-parable.
-The  sign  and  magnitude  of  the  estimate  ofpreference  is  a  function  of  both  the  sample  of  useand the sample of availability.
-This creates an identifia-bility  problem  because  preference  can  change  as  afunction of changing use and/or changing availability.If   the   sample   of   availability   differs   between   twoestimates of preference, it is meaningless to draw infer-ences based on the difference in preference.
-Differentpreferences can arise from the same selection process,and   similar   preferences   can   arise   from   differentselection processes.* [@Beyer2010-vr].
-
-This is exactly the problem I am facing.
-I think, though, I can present my "preference lability analysis" together with the raw visitation and availability data, discuss this limitation explicitly, and let the reader decide what to make of it.
-This is also discussed very clearly, and in the context of bumblebees, in @Harmon-Threatt2017-bz.
-
-*Given that preference is conditional on avail-ability,  that  availability  is  likely  to vary  in  space  andthat preference is likely to change as a function of avail-ability, there is little theoretical basis for believing thatpreferences estimated in one area will be good predic-tors  of  preference  in  others.* [@Beyer2010-vr].
-
-In answer to this, I can point to the largely agreeing results of other bumble bee studies.
-
-*Models of habitat preference provide a phenomenolo-gical  description  of  habitat  use,  conditional  on  thesample of availability.
-Although some of the simplify-ing  assumptions  upon  which  these  models  are  often based  may  be  biologically  unrealistic,  they  can  stillhave predictive power (Boyceet al.2002;Wienset al.2008).
-The degree, however, to which simple modelsof habitat preference help us to develop a mechanisticunderstanding of animal movement and distribution islimited.
-Mechanistic models are required to advanceour  biological  understanding  of  the  behavioural  pro-cesses  that  result  in  animal  movement,  distributionand habitat use.*
-
-Tongue length is my mechanistic model.
-And I think the observation that the mechanistic model better explains use-availability bias than raw selection indicates that the use-availability approach better captures the way bumble bees make foraging decisions.
-
-Need to acknowledge the potential for nonlinear "functional responses" of use to abundance [@Mysterud1998-rs].
-
-*The problem of conflating preference and niche partitioning:*
-
-Our study is set in the Berchtesgadener Alps of southeast Germany, a landscape of flower-rich mountain meadows embedded in a matrix of coniferous forest.
-This system is one of abundant floral resources but harsh abiotic conditions, so it is likely that competition for floral resources is low, or at least transient, and that bumble bee populations are limited instead by the shortness of the growing season, as has been suggested in the similar environment of the Eyne Valley of France [@Iserbyt2008-jn].
-Yet while *realized* competition between bumble bees in our system may be low, the question of *potential* competition is highly relevant because mountain ecosystems are currently subject to at least two powerful drivers of directed change.
-
-Resource usage is typically studied as a phenomenological level, with questions of mechanism relegated to speculations regarding the latent process of resource selection that lurks behind realized use patterns [@Beyer2010-vr].
-In our study, we evaluate the target use hypothesis by reference to a known mechanism of floral selection, namely bumble bee tongue length.
-
-# 5. CONCLUSIONS
-
-## What we did and why
-
-In this study, we investigated patterns of floral resource selection in a mountain bumble bee community with the goals of (1) identifying the most selected floral morphotypes, (2) evaluating the role of floral abundance in driving floral selection, (3)
-
-# ACKNOWLEDGMENTS
-
-We thank D.J. McNeil for helpful conversations during data analysis. O. Mitesser kindly provided a high-performance computing environment to expedite null modeling.
-
 # REFERENCES

### 2021-08-25 code/ms2_nullmodels.Rmd
preped ms for Ecol Lett

@@
 
-view(filter(vis_spp, ktype == "1"))
-
-ggplot(filter(vis_spp, ))
-
-# For setting up the HPC
-# 
-# bb_traits <- read_csv("./bb_traits.csv") %>%
-#   select(bb.sp, pbl.w, tongue = pbl.w.class2)
-# 
-# site_data <- read_csv("./site_data.csv") %>%
-#   dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean, elev.class, elev.class2)) %>%
-#   mutate(site = factor(site),
-#          tree_line = factor(tree_line))
+fl_tax <- read_csv("../data/processed_data/floral_tax.csv")
 ```
@@
 ggsave("../ms_2/figures/null_model_spwise_preference.png", height = 6, width = 7)
-
-# Species-wise (3-col)
-ggplot(null_tab1, aes(Observed, ktype)) +
-  geom_segment(aes(y = ktype, yend = ktype, 
-                   x = Lower.99.CL, xend = Upper.99.CL), 
-               color = "gray40", inherit.aes = FALSE) +
-  geom_point(aes(color = Test)) +
-  facet_wrap(~fct_rev(bb.sp), ncol = 3) +
-  scale_y_discrete(limits = rev) +
-  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
-  labs(x = "Mean visitation rate", y = "Floral morphotype", color = "Test (p < 0.01)") +
-  theme_light()
-
-ggsave("../ms_2/figures/null_model_spwise_preference_3col.pdf", height = 6, width = 5)
-
-## Morphotype-wise
-ggplot(null_tab1, aes(Observed, bb.sp)) +
-  geom_segment(aes(y = bb.sp, yend = bb.sp, 
-                   x = Lower.99.CL, xend = Upper.99.CL), 
-               color = "gray40", inherit.aes = FALSE) +
-  geom_point(aes(color = Test)) +
-  facet_wrap(~ktype, ncol = 4) +
-  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
-  labs(x = "Mean proportional visitation", y = "Bumble bee species", color = "Test (p < 0.01)") +
-  theme_light()
-
-ggsave("../ms_2/figures/null_model_ktypewise_preference.pdf", height = 5, width = 7)
-ggsave("../ms_2/figures/null_model_ktypewise_preference.png", height = 5, width = 7)
+# 
+# # Species-wise (3-col)
+# ggplot(null_tab1, aes(Observed, ktype)) +
+#   geom_segment(aes(y = ktype, yend = ktype, 
+#                    x = Lower.99.CL, xend = Upper.99.CL), 
+#                color = "gray40", inherit.aes = FALSE) +
+#   geom_point(aes(color = Test)) +
+#   facet_wrap(~fct_rev(bb.sp), ncol = 3) +
+#   scale_y_discrete(limits = rev) +
+#   scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
+#   labs(x = "Mean visitation rate", y = "Floral morphotype", color = "Test (p < 0.01)") +
+#   theme_light()
+# 
+# ggsave("../ms_2/figures/null_model_spwise_preference_3col.pdf", height = 6, width = 5)
+# 
+# ## Morphotype-wise
+# ggplot(null_tab1, aes(Observed, bb.sp)) +
+#   geom_segment(aes(y = bb.sp, yend = bb.sp, 
+#                    x = Lower.99.CL, xend = Upper.99.CL), 
+#                color = "gray40", inherit.aes = FALSE) +
+#   geom_point(aes(color = Test)) +
+#   facet_wrap(~ktype, ncol = 4) +
+#   scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
+#   labs(x = "Mean proportional visitation", y = "Bumble bee species", color = "Test (p < 0.01)") +
+#   theme_light()
+# 
+# ggsave("../ms_2/figures/null_model_ktypewise_preference.pdf", height = 5, width = 7)
+# ggsave("../ms_2/figures/null_model_ktypewise_preference.png", height = 5, width = 7)
 
@@
 ```{r}
-null_model_overall<- ggplot(null_tab1, aes(ktype, preference)) +
+null_model_overall <- ggplot(null_tab1, aes(ktype, preference)) +
   #geom_violin(fill = "gray70", color = "white", draw_quantiles = c(0.25, 0.5, 0.75)) +
@@
   geom_jitter(aes(color = Test), width = 0.3, alpha = 0.6) +
-  theme_light(10) +
+  theme_light(12) +
   scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
   labs(x = "Floral morphotype",
-       y = "Selectivity")
+       y = "Preference")
 
@@
   facet_wrap(~ktype) +
-  labs(x = "Tongue length (mm)", y = "Selectivity", 
+  labs(x = "Tongue length (mm)", y = "Preference", 
        color = "Species", linetype = "p < 0.05") +
-  theme_light(10)
-
-ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference.pdf", null_model_pbl_GAM, height = 6, width = 7)
+  scale_x_continuous(breaks = c(7, 10, 13, 16)) +
+  theme_light(12)
+
+ggsave("../ms_2/figures/fig3.pdf", null_model_pbl_GAM, height = 6, width = 7)
 ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference.png", null_model_pbl_GAM, height = 6, width = 7)
@@
   droplevels()
-
-ggplot(null_tab3, aes(ktype, preference)) +
-  #geom_violin(fill = "gray70", color = "white", draw_quantiles = c(0.25, 0.5, 0.75)) +
-  geom_boxplot(fill = NA, color = "black", outlier.shape = 4, size = 0.35) +
-  geom_jitter(aes(color = Test), width = 0.3, alpha = 0.6) +
-  theme_light(14) +
-  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
-  labs(x = "Floral morphotype",
-       y = "Bias")
 ```
@@
 
-ggsave("../ms_2/figures/preference_lability_gam.pdf", height = 5, width = 6.5)
+ggsave("../ms_2/figures/fig5.pdf", height = 5, width = 6.5)
 ggsave("../ms_2/figures/preference_lability_gam.png", height = 5, width = 6.5)
@@
   geom_jitter(width = 0.3, alpha = 0.5, color = "gray40") +
-  theme_light(10) +
+  theme_light(12) +
   labs(x = "Floral morphotype",
@@
 
-ggsave("../ms_2/figures/overall_ktype.pdf", overall_ktype, height = 4.5, width = 9.5)
+ggsave("../ms_2/figures/fig2.pdf", overall_ktype, height = 4.5, width = 9.5)
 ggsave("../ms_2/figures/overall_ktype.png", overall_ktype, height = 4.5, width = 9.5)
@@
        color = "Species", linetype = "p < 0.05") +
-  theme_light(10)
+  scale_x_continuous(breaks = c(7, 10, 13, 16)) +
+  theme_light(12)
 
@@
 # Binomial model
-gam6 <- bam(prop.cover ~
-              ktype +
-              s(site, by = ktype, bs = "re") +
-              s(year, by = ktype, bs = "re") +
-              te(yday, elev.mean, by = ktype),
-            #family = betar(eps = 0.001),
-            family = "binomial",
-            method = "fREML",
-            discrete = TRUE,
-            data = filter(flower_cover_ktype.prop, 
-                            ktype %in% c("3", "5", "6", "7"))) %>% 
-  getViz()
-
-check.gamViz(gam6)
-gam6_sum <- summary(gam6, re.test = FALSE) 
-plot.gamViz(gam6)
+# gam6 <- bam(prop.cover ~
+#               ktype +
+#               s(site, by = ktype, bs = "re") +
+#               s(year, by = ktype, bs = "re") +
+#               te(yday, elev.mean, by = ktype),
+#             #family = betar(eps = 0.001),
+#             family = "binomial",
+#             method = "fREML",
+#             discrete = TRUE,
+#             data = filter(flower_cover_ktype.prop, 
+#                             ktype %in% c("3", "5", "6", "7"))) %>% 
+#   getViz()
+# 
+# check.gamViz(gam6)
+# gam6_sum <- summary(gam6, re.test = FALSE) 
+# plot.gamViz(gam6)
 
@@
 
-ggsave("../ms_2/figures/flower_cover_gam.pdf", height = 5, width = 6.5)
+ggsave("../ms_2/figures/fig6.pdf", height = 5, width = 6.5)
 ggsave("../ms_2/figures/flower_cover_gam.png", height = 5, width = 6.5)

### 2021-08-25 ms_2/ms_2_submission.Rmd
preped ms for Ecol Lett

--- 2021-07-15 ms_2/ms_2_revision2.Rmd
@@
 ---
-title: "Floral preferences of mountain bumble bees are driven by trait matching but labile through elevation and time"
-
-author: Douglas B. Sponsler^1^*, Katharina Kallnik^1^, Fabrice Requier^2^, Alice Classen^1^, A. Fabienne Maihoff^1^, Johanna Sieger^3^, Ingolf Steffan-Dewenter^1^
-
-# date: `r Sys.Date()`
-
-output:
-  bookdown::word_document2:
-    fig_caption: yes
-    fig_width: 7
-    fig_height: 7
-    df_print: kable
-    keep_md: yes
-    number_sections: no
-    toc: no
-    reference_docx: Word_template.docx
-
+title: "Floral preferences of mountain bumble bees are constrained by trait matching but flexible through elevation and season"
+header-includes:
+   - \usepackage{lineno}\linenumbers
+   - \usepackage{setspace}\doublespacing
+output: pdf_document
 bibliography: 
   - references.bib
-#  - knitcitations.bib
-csl: ecology.csl
-# csl: https://www.zotero.org/styles/ecology
-
+csl: ecologyletters.csl
 ---
 
-^1^ Department of Animal Ecology and Tropical Biology, Biocenter, University of Würzburg, Würzburg, Germany
+**Douglas Sponsler** (corresponding author)  
+University of Würzburg  
+Department of Animal Ecology and Tropical Biology  
+Biozentrum der Universität Würzburg C-018  
+Am Hubland  
+97074 Würzburg, Germany  
+Email: douglas.sponsler@uni-wuerzburg.de  
+Telephone: +49 0176 363 20273
 
-^2^ Université Paris-Saclay, CNRS, IRD, UMR Évolution, Génomes, Comportement et Écologie, 91198, Gif-sur-Yvette, France
+**Katharina Kallnik**  
+University of Würzburg  
+Department of Animal Ecology and Tropical Biology  
+97074 Würzburg, Germany  
+Email: katharina.kallnik@uni-wuerzburg.de
 
-^3^ Naturpark Frankenhöhe, Ansbach, Germany
+**Fabrice Requier**  
+Université Paris-Saclay  
+Évolution, Génomes, Comportement et Écologie  
+91198 Gif-sur-Yvette, France  
+Email: fabrice.requier@egce.cnrs-gif.fr
 
-`*` Corresponding author: douglas.sponsler@uni-wuerzburg.de
+**Alice Classen**  
+University of Würzburg  
+Department of Animal Ecology and Tropical Biology  
+97074 Würzburg, Germany  
+Email: alice.classen@uni-wuerzburg.de
+
+**A. Fabienne Maihoff**  
+University of Würzburg  
+Department of Animal Ecology and Tropical Biology  
+97074 Würzburg, Germany  
+Email: anne_fabienne.maihoff@uni-wuerzburg.de
+
+**Johanna Sieger**  
+Naturpark Frankenhöhe  
+91522 Ansbach, Germany   
+Email: sieger@lpv-mfr.de
+
+**Ingolf Steffan-Dewenter**  
+University of Würzburg  
+Department of Animal Ecology and Tropical Biology  
+97074 Würzburg, Germany  
+Email: ingolf.steffan@uni-wuerzburg.de
+
+\
+
+**AUTHORSHIP STATEMENT:** KK and ISD conceived the study.
+KK, JS, and ISD designed and carried out field work.
+DS analyzed the data and wrote the paper with substantial contributions from FR, AC, AFM, KK, and ISD.
+
+**DATA ACCESSIBILITY:** All data used in this study have been archived on DRYAD (https://datadryad.org/stash/dataset/doi:10.5061/dryad.s1rn8pk7f) and will be available upon article publication.
+
+**ARTICLE TYPE:** Letter
+
+**RUNNING TITLE:** Bumble bee floral preferences
+
+**KEYWORDS:** foraging, trait matching, pollinator, *Bombus*, alpine, niche, resource selection
+
+**SECTION LENGTHS:**  
+Abstract:  
+Main text:  
+References:  
+Figures:
+
+\newpage
 
@@
 Patterns of resource use by animals can clarify how ecological communities have assembled in the past, how they currently function, and how they are likely to respond to future perturbations.
-Using ~13,000 records of bumble- bee-wildflower interactions along a 1400 m elevation gradient, we find that the selection of floral morphotypes by bumble bees is driven both by the directed process of trait-matching and the neutral effects of floral abundance.
-Patterns of floral preference are, moreover, labile through time and elevation, which may be an adaptation to the extreme spatiotemporal floristic turnover of mountain ecosystems.
+Using ~13,000 records of bumble- bee-wildflower interactions along a 1400 m elevation gradient, we find that the selection of floral morphotypes by bumble bees is driven both by the directed process of trait-matching and the neutral effect of floral abundance.
+Patterns of floral preference also vary by elevation and season, which may be an adaptation to the strong spatiotemporal turnover of mountain floral communities.
 The flexibility of bumble bee foraging will be tested in coming years as climate and land use change drive rapid changes in floral resource distribution in time and space, particularly in the mountain ecosystems that host the world's richest communities of bumble bees and their floral mutualists.
 
-*Keywords*: foraging, trait matching, pollinator, *Bombus*, alpine, niche, resource selection
+\newpage
 
@@
 
-Patterns of resource selection and partitioning by animals are instrumental in community assembly, species coexistence, and ecosystem function [@Chesson2000-aw; @Bascompte2007-ls].
+Resource selection and partitioning by animals are instrumental in community assembly, species coexistence, and ecosystem function [@Chesson2000-aw; @Bascompte2007-ls].
 Resource selection is fundamentally constrained by phenotypic compatibility (i.e.
@@
 
-Plant-pollinator networks are a special case of resource selection in which consumer fitness and resources fitness are positively correlated, causing the traits of plants and pollinators to converge toward compatibility with one another [e.g.
+\
+
+<!-- Plant-pollinator networks are a special case of resource selection in which consumer fitness and resources fitness are positively correlated, causing the traits of plants and pollinators to converge toward compatibility with one another [e.g.
 @Newman2014-ke].
 The coexistence of multiple plant and pollinator species, however, requires niche differentiation [@Palmer2003-ih; @Johnson2019-su], which can be achieved by trait-based partitioning of interactions among pollinators and their floral hosts [@Willmer2011-ak].
+-->
 
@@
 Warming climates have already caused measurable upslope shifts in mountain bumble bees and their floral hosts [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc], which could introduce novel competitive interactions between bumble bee species that experience increased co-occurrence.
-At the same time, socioeconomic factors have led to widespread abandonment and reforestation of mountain meadows, particularly in the European Alps, which dramatically reduces floral resource availability and suppresses bumble bee diversity and abundance [@Tappeiner2006-jp; @Tasser2007-qo; @Walcher2017-kw].
+At the same time, socioeconomic factors have led to widespread abandonment and reforestation of mountain meadows, particularly in the European Alps, which reduces floral resource availability and suppresses bumble bee diversity and abundance [@Tappeiner2006-jp; @Tasser2007-qo; @Walcher2017-kw].
 The ability of bumble bee species to coexist under these conditions of reshuffled distributions and declining floral resources will depend in part on how they select and partition floral resources, which will in turn affect the viability of their floral mutualists, many of which depend principally on bumble bee visitation for pollination [@Goulson2010-lo].
+
+\
 
@@
 The most well-studied mechanism of the preference component of bumble bees floral selection is trait-matching between bumble bee tongue length and floral corolla depth.
-Bumble bee species can differ markedly in tongue length, and floral handling time increases when tongue length and corolla depth are poorly matched [@Inouye1980-mv; @Harder1983-bk], leading to a well-documented association between tongue length and floral preferences in bumble bee communities [e.g.
+Bumble bee species can differ markedly in tongue length, and floral handling time increases when tongue length and corolla depth are poorly matched [@Inouye1980-mv; @Harder1983-bk], leading to an association between tongue length and floral preferences in bumble bee communities [e.g.
 @Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn].
-Under the use-availability model, tongue-length-based floral preferences should be modulated by the relative abundance of available flora to generate observed patterns of floral selection.
+Under the use-availability model, tongue-length-based floral preferences should be modulated by the relative abundance of available resources to generate observed patterns of floral selection.
 It has recently been proposed, however --- in what has been termed the *target use hypothesis* --- that floral selection by bumble bees is a process of *pure preference*, irrespective of resource abundance [@De_Valpine2013-be; @Harmon-Threatt2017-bz].
@@
 
+\
+
 In this study, we use ~13,000 records of bumble bee visits to wildflowers, collected over three years along a 1400 m elevation gradient, to explore three unanswered questions regarding resource selection by bumble bees and its response to climate and land use change.
 First, we ask how bumble bee floral preferences map onto a generalized floral morphotypology [@Kugler1970-gb].
-Then, using the well-established relationship between tongue length and floral selection as a point of reference, we test the *target use hypothesis* by comparing the explanatory power of bumble bee tongue length with respect to observed visitation rates vs.
+Then, using the established relationship between tongue length and floral preference as a point of reference, we test the *target use hypothesis* by comparing the explanatory power of bumble bee tongue length with respect to observed visitation rates vs.
 the residuals of a null model that assumes a linear dependency of selection on resource abundance.
@@
 
+\
+
 Bumble- bee-wildflower interactions were observed during repeated visits (hereafter, “visitation samples”) to each study site at approximately weekly intervals.
@@
 (submitted) for details of sampling design.
+
+\
 
@@
 
+\
+
 Bumble bee tongue-length data (for the worker caste) were obtained from the literature values compiled and summarized by @Arbetman2017-xj.
@@
 
+\
+
 To quantify preference, we compared the proportional visitation of each morphotype (by each bumble bee species) with the rates of visitation expected under a null model constructed following @Vaughan2018-gj.
@@
 
+\
+
 We followed the procedure described above, simulating our entire dataset in one modeling process to infer aggregate preference between bumble bee species and floral morphotypes.
@@
 Finally, we compared morphotype preference across bumble bee species by performing a principal components analysis on the bumble bee species x floral morphotype preference matrix, and we mapped bumble bee tongue length onto the ordination space by fitting a 2-dimensional GAM surface to the first two principal components.
+
+\
 
@@
 
+\
+
 An inherent problem with modeling sample-wise variation in preference is that the bounds of preference are strictly dependent on floral abundance [@Beyer2010-vr].
@@
 With this limitation in mind, we fit HGAMs of the same form to the relative abundance of each floral morphotype, and we present the results together to be interpreted jointly.
+
+\
 
@@
 
-## 3.1 Sampling overview
+## 3.1 Bumble bee and floral communities
 
-A total of 12,918 bumble- bee-flower interactions were included in our final data set, including 16 bumble bee species, 163 plant species (110 genera, 37 families), and 736 unique bumble- bee-plant interaction pairs.
+A total of 13,293 bumble- bee-flower interactions were included in our final data set, comprising 15 bumble bee species, 163 plant species (110 genera, 37 families), and 730 unique bumble- bee-plant interaction pairs.
 *B.
@@
 humilis* ("humi") was recorded only three times and only in one year, so we omitted it from all analyses.
-
-Floral surveying yielded a total of 352 plant species (191 genera, 52 families), of which 155 (103 genera, 35 families) were observed to be visited by bumble bees.
+Floral surveying yielded a total of 354 plant species (194 genera, 53 families).
 Visited floral species included all morphotypes except spike (8) and trap (10) flowers.
 Null (0) and brush (9) morphotypes were, however, excluded from analysis for the reasons given above.
-
-<!-- Check again and make sure all the numbers given here are still correct. -->
 
@@
 Preference was generally strong, with the visitation rates of most bumblebee species to most floral morphotypes falling significantly above or below the 99% interval of the null model **(Figure 1)**.
-Nonsignificant preference was observed mainly among the rarer bumblebee species because their small samples sizes lead to wide null model confidence intervals.
+Nonsignificant preference was observed mainly among the rarer bumblebee species, likely because their small samples sizes lead to wide null model confidence intervals.
+Interestingly, bumble bees showed strong species-specific variation in their relative preferences of different floral morphotypes **(Figure 1)**.
+The omission of queens from the calculation of overall preference did not yield qualitatively different results (Appendix X).
 
-```{r fig1, echo = FALSE,  fig.cap = paste("Observed visitation rates compared to null model simulations for each bumble bee species and floral morphotype (A).
-Gray bars depict 99\\% intervals of null model simulations and points depict observed visitation rates.
-Points are color-coded according to where they fall with respect to the 99\\% interval of the null model: red = above (positive preference), blue = below (negative preference), gray = within (nonsignificant preference).
-The difference between the observed value and the center of the null model interval can be interpreted as magnitude of preference.
-Each floral morphotype is illustrated (B) with a representative species common in our study system: 1 = Potentilla erecta, 2 = Gentiana asclepiadea, 3 = Campanula glomerata, 4 = Silene nutans, 5 = Ajuga reptans, 6 = Lotus corniculatus, 7 = Centaurea jacea.
-Illustrations are modifications of public domain works obtained from www.plantillustrations.org.
-For details of original sources, see Appendix X.")}
-knitr::include_graphics("../ms_2/figures/fig1.png")
-```
+\
 
-When floral morphotypes are compared to one another in terms of preference, clear patterns emerge **(Figure 2A)**.
+When floral morphotypes are compared to one another in terms of preference across all bumble bee species, clear patterns emerge **(Figure 2A)**.
 All species exhibited negative preference toward disc flowers, while most species exhibited positive preference toward bell, lip, and flag flowers.
@@
 gerstaeckeri* --- for funnel flowers), reflecting the fact that these morphotypes tended both to be rarely visited and to occur at low abundance.
-preference toward head flowers varied widely from strongly positive to strongly negative.
+Preference toward head flowers varied widely across bumble bee species, ranging from strongly positive to strongly negative.
 
-Differences between floral morphotypes in terms of raw mean visitation rates, reflecting the target use hypothesis, were less pronounced **(Figure 2B)**.
-Bell, lip, flag, and head flowers were visited at high rates by one or more bumble bee species, but median raw mean visitation rates did not differ strongly across morphotypes.
+\
 
-The omission of queens from the calculation of overall preference did not yield qualitatively different results (Appendix X).
+Differences between floral morphotypes in terms of observed visitation rates, reflecting the target use hypothesis, were less pronounced **(Figure 2B)**.
+Bell, lip, flag, and head flowers were visited at high rates by one or more bumble bee species, but observed visitation rates did not differ strongly across morphotypes.
 
@@
 
-<!-- Before we finalize this, check to make sure the raw visitation rates were really prepared correctly and are not being deflated by a bunch of zeros that should'nt be there. -->
-
-```{r fig2, echo = FALSE,  fig.cap = paste("Preference (A) and observed visitation rate (B) of the bumble bee species toward each floral morphotype.
-Each point represents the preference/visitation of one bumble bee species toward the corresponding floral morphotype.
-The color of each point in panel A, retained from Figure 2, indicates whether a given observed bias is significantly (p < 0.01) positive or negative with respect to null model simulations.")}
-knitr::include_graphics("../ms_2/figures/overall_ktype.png")
-```
-
-## 3.4 Tongue length as a predictor of floral choice
+## 3.3 Tongue length as a predictor of floral choice
 
 Tongue length was a strong predictor of preference but not of observed visitation rates **(Figure 3)**.
-preference toward funnel, bell, lip, flag, and head flowers exhibited significant GAM curves in response bumble bee tongue length, and the model explained 77% of total deviance (adj.
+Preference toward funnel, bell, lip, flag, and head flowers exhibited significant GAM curves in response to bumble bee tongue length, and the model explained 77% of total deviance (adj.
 R^2^ = 0.728).
@@
 Preference toward bell flowers and head flowers exhibited approximately quadratic responses with preference peaking among short-tongued species and declining sharply with increasing tongue length before leveling off.
-preference toward flag flowers exhibited a unimodal peak among medium-tongued species.
-In contrast, observed visitation rates were significantly predicted by tongue length only for head flowers, which exhibited a weakly quadratic decline in preference from a peak among short-tongued species.
+Preference toward flag flowers exhibited a unimodal peak among medium-tongued species.
+In contrast, observed visitation rates were predicted by tongue length only for head flowers, which exhibited a weak decline in preference from a peak among short-tongued species.
 The visitation rate model explained only 25% of total deviance (adj.
 R^2^ = 0.053).
+Again, these results did not differ when preference was based on the worker-only null model (Appendix X).
 
-Again, these results did not differ qualitatively when preference was based on the worker-only null model (Appendix X).
+\
 
-```{r fig3, echo = FALSE,  fig.cap = paste("preference (A) and observed visitation rates (B) for each morphotype regressed against bumble bee tongue length.
-Each point represents the preference/visitation of one bumble bee species toward the corresponding floral morphotype.
-Curves depict fitted GAM smooths with 95\\% confidence bands.
-Solid lines indicate curves differing significantly (p < 0.05) from a flat line while dashed lines indicated nonsigificant curves.
-Points are color-coded by species, and species are presented in the legend in order of increasing tongue length.")}
-knitr::include_graphics("../ms_2/figures/ktypeXtonguelength_paired.png")
-```
+PCA of bumble bee floral preferences captured more than 80% of total variance in the first two axes **(Figure 4)**, and the GAM surface fit to the firs two axes explained most of the variance in tongue length (deviance explained = 94.7%, adj.
+R^2^ = 0.893, p < 0.001), reflecting the importance of trait matching between tongue length and floral morphotype as a mechanistic basis of preference.
+Longer tongue length was associated with deeper and more closed flower morphotypes (lip, funnel, flag), while shorter tongue length was associated with more shallow and open flower morphologies (disc, stalk-disc, bell, head).
+The strong affinity of *B.
+soroeensis* and *B.
+psithyrus* for head flowers is especially salient.
 
-PCA of preference captured more than 80% of total variance in the first two axes **(Figure 4)**, and the GAM surface fit to the firs two axes explained most of the variance in tongue length (deviance explained = 94.7%, adj.
-R^2^ = 0.893, p < 0.001).
+## 3.4 Variation of morphotype abundance and preference through elevation and season
 
-<!-- Need to add a bit of comment here about important patterns in the ordination. -->
-
-Perhaps the most striking pattern
-
-```{r fig4, echo = FALSE,  fig.cap = paste("Principal components biplot of bumble bee species according to their preference toward each floral morphotype.
-Floral morphotypes are abbreviated with their numerical codes and illustrated as in Figure 1.
-Colored bands depict a 2-dimensional GAM smooth representing the relationship between the first two principal components and bumble bee tongue length.
-Morphotype illustrations are as described in Figure 1.")}
-knitr::include_graphics("../ms_2/figures/preference_ordination_mod.png")
-```
-
-## 3.5 Variation of floral abundance, raw visitation, and preference through elevation and time
-
-The proportional abundance of the most visited floral morphotypes --- bell, lip, flag, and head flowers --- was strongly patterned in time and elevation **(Figure 5)**.
+The proportional abundance of the most visited floral morphotypes --- bell, lip, flag, and head flowers --- was strongly patterned by elevation and season **(Figure 5)**.
 Bell flowers peaked in proportional abundance early in the season and at mid-elevation, and lip flowers peaked late in the season and at low elevation.
@@
 
-```{r fig5, echo = FALSE,  fig.cap = paste("Variation in proportional flower cover through elevation and modeled as HGAM surfaces.
-Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns.
-All smooths are significant (p < 0.05).")}
-knitr::include_graphics("../ms_2/figures/flower_cover_gam.png")
-```
+\
 
-Patterns of preference through elevation and time were more complex than those of raw visitation and, by design, more often differed from patterns of floral abundance **(Figure 6)**.
+The morphotype preferences of bumble bee species were also strongly patterned through elevation and season.
 In some cases, though, patterns of preference mirrored patterns of floral availability, suggesting that bumble bees may exhibit functional responses [@Mysterud1998-rs] to floral abundance not entirely accounted for by null models that assume a linear proportionality between abundance and visitation.
-
-One of the more salient patterns is that positive selection for flag flowers was strongly associated with high elevation, except among the species that tended to avoid flag flowers (*B.
+Preference for flag flowers was strongly associated with high elevation, except among the species that tended to avoid flag flowers (*B.
 pratorum* and *B.
@@
 terrestris-lucorum*, and *B.
-soroeensis* **(Figure 1)**, also exhibited the most complex patterns of preference in elevation and time for these bumble bee species.
+soroeensis* **(Figure 1)**, also exhibited the most complex patterns of preference in elevation and time for these bumble bee species **(Figure 6)**.
 In contrast, the long and medium-tongued species *B.
@@
 pascuorum*, and *B.
-wurflenii*, which exhibited neutral or weakly positive overall selection toward bell flowers **(Figure 1)**, also show weak patterns of variation in preference toward bell flowers.
+wurflenii*, which exhibited neutral or weakly positive overall selection toward bell flowers **(Figure 1)**, also show weak patterns of variation in preference toward bell flowers **(Figure 6)**.
 Nearly the reverse pattern is evident for flag flowers, reflecting their association with longer-tongued bumble bee species.
 
-Smooths were significant (p < 0.05) for all bb-fl pairs except hort-head, prat-flag, telu-head, and soro-flag. The model explained 36.4% of total deviance (adj. R^2^ = 0.33).
+\
 
-```{r fig6, echo = FALSE,  fig.cap = paste("Variation in preference through elevation and time modeled as HGAM surfaces.
-Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns.
-Significant (p < 0.05) smooths are depicted with solid contour lines and nonsignificant smooths with dashed contour lines.")}
-knitr::include_graphics("../ms_2/figures/preference_lability_gam.png")
-```
+Smooths were significant (p < 0.05) for all bee-morphotype pairs except hort-head, prat-flag, telu-head, and soro-flag. The model explained 36.4% of total deviance (adj. R^2^ = 0.33).
 
@@
 
+Using ~13,000 bumble- bee-wildflower interactions, we characterized bumble bee floral selection in terms of generalized floral morphotypes and found evidence that bumble bee floral selection depends both on the directed process of morphotype preference and the neutral effect of floral abundance.
+Bumble bee species differed in their floral morphotype preferences, and these differences could be explained by trait-matching between floral morphotype and bumble bee tongue length.
+These preferences, however, varied by elevation and season, likely representing adaptive responses of bumble bee foraging behavior to elevational and seasonal shifts in floral resource composition, possibly modulated by competitive interactions among bumble bees species with overlapping resource preferences.
+
 ## 4.1 Floral abundance, floral selection, and the *target use hypothesis*
 
-It is well-known that bumble bees forage non-randomly accordingly to morphological constraints, innate preferences, and experienced rewards [e.g.
-@Kugler1943-uy; @Brian1954-cy; @Inouye1978-qi; @Raine2007-nq; @Maharaj2019-wm].
-In this regard, the results of our null model analysis are as unsurprising as they are unambiguous: bumble bee resource selection almost never represents flora in unaltered proportion to their abundance.
-The *target use hypothesis* [@De_Valpine2013-be; @Harmon-Threatt2017-bz], however, extends this idea to the perfect antithesis of the random foraging model, positing that bumble bee foraging is a process of *pure* preference, completely irrespective of resource abundance.
+Our finding that bumble bee tongue length explained preference (i.e.
+null model residuals) much better than it explained observed visitation rates indicates that, in our study system, abundance and preference function jointly in bumble bee floral selection.
+This casts doubt on the *target use hypothesis* as a general rule of bumble bee foraging, though it is possible that it still obtains under certain circumstances [e.g.
+@Harmon-Threatt2017-bz].
+In general, though, our findings support the hypothesis that bumble bee resource selection involves both the neutral effects of floral abundance and the directed effect of preference, meaning that analytical techniques designed to disentangle these effects, such as the null model method we employ, should be favored in studies of bumble bee foraging.
 
-Given the well-established mechanistic link between bumble bee tongue length and floral selection [e.g.
-@Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn], our finding that bumble bee tongue length explained preference (i.e.
-null model residuals) much better than it explained raw visitation rates indicates that, in our study system, abundance and preference function jointly in bumble bee floral choice.
-This casts doubt on the *target use hypothesis* as a general rule of bumble bee foraging.
-It does not, however, mean that the *target use hypothesis* does not obtain under certain circumstances, such as when a highly preferred resource varies within a nonlimiting rage of availability.
-Moreover, our results suggest something akin to the *target use hypothesis* but formulated negatively with respect to resources that are *not selected*.
-Visitation rates to disc flowers in our system, for example, were very low despite the fact that they often occurred at high abundance, suggesting that while preference seems to interact with resource abundance, avoidance may not.
-In general, though, our findings support the hypothesis that bumble bee resource selection involves both the neutral effects of abundance and the directed effect of preference, which implies that analytical techniques designed to disentangle these effects, such as the null model analysis we employ, should be favored in studies of bumble bee foraging.
-<!-- Accordingly, we shall focus primarily on our calculations of preference, rather than raw visitation rates, as we interpret the other results of our study. -->
+## 4.2 Community-level patterns of morphotype preferences
 
-## 4.2 Overall patterns of preference
+We found that the most preferred morphotypes were bell, lip, and flag flowers, which is broadly consistent with visitation patterns reported by previous studies in central Europe [@Schneller2014-ry; @Neumayer1999-il; Farnberger1996-xo], indicating that our findings are regionally generalizeable.
+These previous studies tended to find a stronger role of head flowers than indicated by our preference analysis, but this is probably because they characterized foraging in terms of simple visitation rates rather than null-model-inferred preference.
+We saw similarly high visitation rates toward head flowers, but our null-model analysis reveals that while head flowers were highly visited, they tended to be under-visited relative to their abundance.
 
-We found that positive selection was concentrated on 3 floral morphotypes: bell, lip, and flag flowers.
-These results are broadly consistent with the raw visitation patterns previously reported by @Schneller2014-ry in the central Alps at Vorarlberg, Neumayer1999-il in the Austrian Alps ~55 km south of our study system, and Farnberger1996-xo in the vicinity of Vienna, suggesting that our findings are at least regionally generalizeable.
-These neighboring studies tended to find a stronger role of head flowers than suggested by our preference analysis, but this is probably because they characterized foraging in terms of raw visitation rather than preference.
-Indeed, our raw visitation data similarly indicate a stronger role of head flowers, but our preference analysis reveals that while head flowers were highly visited, they tended to be under-visited relative to their abundance.
+\
 
-When considering these patterns from the perspective of pollination services to mountain flora, it is important to note that while bumble bees are the dominant Hymenopteran pollinators in our study system, flies (Diptera), and especially the hover flies (Syrphidae), are also abundant flower visitors in mountain ecosystems, and their visitation is mainly restricted to the open flower morphotypes (e.g.
-disc flowers) that receive less attention from bumble bees [@Neumayer1999-il].
-In the context of our research question, we have emphasized the differences in tongue length between bumble bee species, but all bumble bees have relatively long tongues compared to the overall pollinator community, and it is not surprising that their foraging tends to be focused on the deeper flower morphotypes.
-The process of floral resource partitioning *among* bumble bee species is nested within the process of floral partitioning *between* bumble bees and other flower visitors.
-
-While bell, lip, and flag flowers were the most selected morphotypes on average, preference toward these morphotypes varied broadly across bumble bee species.
-Moreover, while average preference toward funnel and head flowers was low, each received strongly positive selection from one or two bumble bee species.
+While bell, lip, and flag flowers were the most preferred morphotypes on average, preference toward these morphotypes varied broadly across bumble bee species.
+Moreover, while average preference toward funnel and head flowers was low, each was strongly preferred by one or two bumble bee species.
 The majority of variation in preference toward floral morphotypes was explainable by bumble bee tongue length, with the more open bell and head morphotypes selected more positively by short-tongued bumble bee species and the deeper funnel, lip, and flag morphotypes selected more positively by medium- and long-tongued bumble bee species.
@@
 A limitation of our study is that we could not reliably distinguish pollen- and nectar-foraging bumble bees, and it is likely that this distinction would shed further light on the mechanisms underlying the selection patterns we observed.
+
+\
 
@@
 *B.
-mucidus* exhibited unusually low preference (though not significantly different from neutrality) toward lip flowers, complemented by unusually high preference toward flag flowers.
+mucidus* exhibited unusually low preference toward lip flowers, complemented by unusually high preference toward flag flowers.
 *B.
@@
 Such unexplained variation in preference may indicate that traits not captured by Kugler's (1970) morphotypology are also important in governing floral selection by bumble bees.
-Indeed, one such trait --- floral color --- is known to be instrumental in the process of flower detection by foraging bumble bees [@Spaethe2001-bq], and some bumble bee species exhibit innate color preferences [@Raine2007-nq].
+FLoral color, in particular, is known to be instrumental in the process of flower detection by foraging bumble bees [@Spaethe2001-bq], and some bumble bee species exhibit innate color preferences [@Raine2007-nq].
 This effect can also be seen in field studies, where bumble bees have been shown to over- or under-represent certain floral colors [e.g.
@@
 
-As mountain bumble bees can be expected to face a future of declining resource availability and altered elevational distributions, it is relevant to ask which species seem poised for the greatest resource conflict should competitive conditions arise.
-While any answer to this question should be regarded as speculative, considering the similarities of morphotype preference between species is a good place to start.
-As a general rule, it can be said that species of similar tongue length will be, on average, more likely to compete for resources than species of dissimilar tongue length.
+\
+
+Species of similar tongue length will, on average, be more likely to compete for resources than species of dissimilar tongue length.
 Our ordination analysis, though, suggests that there may be significant patterns of similarity and divergence even among species of similar tongue length.
@@
 pyrenaeus* cluster tightly around bell flowers, while *B.
-psithryus* (species group) and *B.
+psithryus* and *B.
 soroeensis* cluster tightly around head flowers.
@@
 A red-listed vulnerable species, *B.
-gerstaeckeri* is a pollen specilaist on the lip flowers of the genus *Aconitum* [@Ponchau2006-af; @Schneller2014-ry], and there is evidence that joint elevational shifts of bumble bees and floral resources over the past century have generated more intense exploitation of *Aconitum*, which may indicate nascent competitive conditions for *B.
+gerstaeckeri* is a pollen specialist on the lip flowers of the genus *Aconitum* [@Ponchau2006-af; @Schneller2014-ry], and there is evidence that joint elevational shifts of bumble bees and floral resources over the past century have generated more intense exploitation of *Aconitum*, which may indicate nascent competitive conditions for *B.
 gerstaeckeri* [@Marshall2020-pc].
 
-Broaching the topic of competition raises an issue that can confound any observational study of resource selection: how to disentangle the endogenous process of resource preference from the exogenous process of niche partitioning driven by competition [@Pianka1974-qk].
-Empirical preference could reflect a latent process of resource preference, as we assume in our interpretation.
-Alternatively, though, visitation patterns and the preference inferred from them could be distorted by competitive forces.
-Intraspecific competition in a bumble bee community would tend to force individuals of a species to broaden their diet beyond what they would prefer under non-competitive conditions, while interspecific competition would be expected to cause a narrowing of diet breadth toward resources that can be exploited most efficiently.
-Thus, observed visitation patterns, even when calculated as abundance-relative preference, can only be interpreted in terms of *preference* to the extent that competition can be assumed to be negligible, and assumption must be justified on the basis of additional knowledge of the study system under consideration.
-Our study system in the Berchtesgadener Alps is one of abundant floral resources but harsh abiotic conditions, so it is likely that competition for floral resources is low, or at least transient, and that bumble bee populations are limited instead by the shortness of the growing season, as has been suggested in the similar environment of the Eyne Valley of France [@Iserbyt2008-jn].
-We cannot, however, exclude the possibility that competitive forces distort, to some degree, the visitation patterns we observed.
-Even when competitive conditions do not obtain universally in a given system, competition may occur in a spatially and/or temporally localized way [@Ranta1981-vx], perhaps associated with extreme climatic events that cause reductions in floral resource avaulability [@Rasmont2012-zq; @Thomson2016-gq; @Phillips2018-me].
-This calls to mind the insight of MacArthur1972-qz, that competition need not be frequent to have lasting ecological impacts, and it is likely that the "ghosts of competition past" [@Connell1980-dn] haunt the floral preferences exhibited by bumble bees today.
+\
+
+When considering community-level patterns of preference and visitation from the perspective of pollination services to mountain flora, it is important to note that while bumble bees are the dominant Hymenopteran pollinators in our study system, flies (Diptera) are also abundant flower visitors in mountain ecosystems, and their visitation is mainly restricted to the open flower morphotypes (e.g.
+disc flowers) that receive less attention from bumble bees [@Neumayer1999-il].
+In the context of our research question, we have emphasized the differences in tongue length between bumble bee species, but all bumble bees have relatively long tongues compared to the overall pollinator community, and it is not surprising that their foraging tends to be focused on the deeper and more closed flower morphotypes.
+The process of floral resource partitioning *among* bumble bee species is nested within the process of floral partitioning *between* bumble bees and other flower visitors.
 
@@
 
-The intraspecific lability of bumble bee foraging preferences (*sensu stricto*) has been demonstrated under laboratory conditions [@Ings2009-vc; @Maharaj2019-wm]; our study reveals the *in situ* lability of preference for multiple bumble bee species through elevational and temporal gradients of floristic turnover.
-While we stress the limitations of interpreting empirical preference in terms of behavioral preference without controlling the joint abundances of all available resources [@Beyer2010-vr], interpreting the dynamics of preference in our study as the adaptive tuning of preference is consistent with the ecological hurdles bumble bees face as characteristically --- and most likely primordially --- mountain species [@Williams1985-ql; @Hines2008-er].
-Where steep temperature gradients and short growing seasons generate extreme spatiotemporal turnover of floral resources, adaptive flexibility of morphological preferences should be favored over fixed innate biases.
+Intraspecific variation of bumble bee foraging preferences has been demonstrated under laboratory conditions [@Ings2009-vc; @Maharaj2019-wm]; our study reveals the *in situ* variation of preference for multiple bumble bee species through elevational and temporal gradients of floristic turnover.
 
-At least three mechanisms could underlie the patterns of preference through elevation and time that we observed.
-To illustrate these hypotheses, we will take an as example the relatively strong preference dynamics of flag flowers in our system, which tended to be most preferred at high-elevation sites **(Figure 8)**.
-Our study does not enable the testing of these hypothesized mechanisms (and their constituent submechanisms), but we articulate them for the sake of clarity and as a guide for future investigations.
-Moreover, the hypotheses we propose not mutually exclusive, and it plausible that all three could obtain simultaneously in a given ecological context.
+\
 
-First, variation in preference toward a given morphotype could be driven by changes in the *intrinsic* reward of that focal morphotype.
-These changes might, in turn, be driven by (1) shifts in species composition of the focal morphotype (e.g.
+At least three mechanisms could underlie the variation of preferences we observed.
+First, variation in preference toward a given morphotype could be driven by changes in the *intrinsic* reward of that focal morphotype, which could be caused by (1) shifts in species composition of the focal morphotype (e.g.
 the occurrence of a particularly rewarding flag flower at high-elevation sites), (2) shifts in the reward levels of individual species (e.g.
-increased nectar production in the flag flower *Lotus corniculatus* under high elevation conditions), or (3) possibly changes in nutritional requirements of bumble bees (e.g.
+increased nectar production in the flag flower *Lotus corniculatus* under high elevation conditions), or (3) changes in nutritional requirements of bumble bees (e.g.
 an increased need for concentrated nectar to support thermoregulation at colder ambient temperatures).
-The hypothesis is consistent with the interpretation of preference as an index of the latent process of absolute preference.
-
-Second, preference toward a given morphotype could vary due to changes in the *relative* reward of the focal morphotype under changing background resource conditions.
+Alternatively, preference toward a given morphotype could vary due to changes in the *relative* reward of the focal morphotype with respect to changing background resource conditions.
 For example, it could be that flag flowers are selected more positively at high elevation because they must compete for visitation with highly attractive lip flowers at lower elevation.
-This hypothesis expresses the fundamental problem of inferring preference *sensu stricto* from selection, namely contingency of selection upon the joint availability of all resource options [@Johnson1980-co; @Aarts2008-fo; @Beyer2010-vr].
-
-The two hypotheses above are both grounded in the intrinsic properties of the floral community, whether those of the focal resource of those of the background resources.
-It is also possible, however, that patterns of resource selection could vary due to competitive interactions within or between bumble bee species.
+Finally, it is possible that patterns of resource selection could vary due to competitive interactions within or between bumble bee species.
 For example, the increased selection of flag flowers at high elevation could be driven by competition release from *B.
@@
 Under this hypothesis, an organism's resource selection is the product of exogenous constraint rather than endogenous preference.
-For the reasons discussed above, however, we consider it unlikely that competition plays a constitutive role in governing bumble bee resource selection in the Berchtesgadener Alps.
 
-<!-- Patterns of preference through elevation and time toward major floral morphotypes tended to be similar across the six dominant bumble bee species we  -->
+## 4.4 Conclusions
 
-<!-- Interesting deviations -->
-
-<!-- The late high elevation peak of visitation and preference between soro and bell flowers is probably driven by soro's peculiar affinity for Campanula.  -->
-
-<!-- Something about the apparent link between overall preference and its variability? -->
-
-<!-- <!-- We reiterate, however, the inherent problems of comparing preference --- understood as an index of the latent process of preference --- across different contexts of resource abundance.
-We recommend the cautious interpretation of preference patterns through elevation and time in conjunction with the underlying patterns of floral abundance and raw visitation.
---> -->
-
-Biological patterns through elevation and time are acted upon by climate and land use change, and the consequence of this action is the generation of new biological patterns through elevation and time.
-A central task of ecology is to understand the patterns that will be acted upon and predict the ecological ramifications of the patterns that will be created.
-In this regard, it is significant that we find marked structure in the morphotypal composition, raw visitation, and preference of the floral community on which the bumble bees of the Berchtesgadener Alps depend.
-While our data permit only speculation as to the climate and land use driven processes at work in our system, our speculation can be nonetheless well-informed.
-Climate change over the last century has caused significant upslope movement [e.g.
-@Marshall2020-pc] and phenological shifts [e.g.
-@CaraDonna2014-nq] of floral resources for pollinators.
-It is likely that the current patterns of morphotype relative abundance that we observed differ from those that would have been observed a century ago, and in the coming decades warming climates will probably induce further shifts upslope [@Marshall2020-pc] and potentially complex shifts in the onset, peak, and end of flowering periods [@CaraDonna2014-nq].
-These effects of climate change will interact with the likely more rapid effects of land use change.
-In particular, the reforestation of montane meadows can be expected to erode the abundance of flowers from mid-elevation up to the tree line (in our system, around ~1500 m.s.l.).
-Above the tree line, however, floral communities should be relatively safe from reforestation since the advance of alpine tree lines tends to lag behind climate change [@Dullinger2004-qu].
-At lower elevations, agricultural intensification can be expected to suppress overall floral availability by favoring grasses over forbs and shift the composition of the community toward species that thrive under the altered soil conditions and disturbance regimes of intensive agriculture [@Tasser2002-om].
+Clarifying the patterns and processes of bumble floral selection sheds light on how bumble bee communities have evolved with respect to their mountain habitats and how they might respond to the incipient forcings of climate and land use change.
+The dynamics of preference in our study suggest adaptive tuning of floral choice consistent with the ecological hurdles that have shaped the evolution of bumble bees in the mountain habitats in which they arose and diversified [@Williams1985-ql; @Hines2008-er].
+Where steep temperature gradients and short growing seasons generate extreme spatiotemporal turnover of floral resources, adaptive flexibility of morphological preferences should be favored over fixed biases.
+As climate and land use change drive the redistribution of bumble bees and their floral mutualists, and likely an overall reduction in floral availability in mountain habitats, competition within bumble bee communities is likely to intensify.
+Whether this leads to competitive exclusion will depend on whether interspecific competition exceeds the stabilizing effects of interspecific competition.
+In this regard, the partial dependency of bumble bee floral selection on the neutral effects of floral abundance (in contrast to the *target use hypothesis*) can be expected to favor the coexistence of both bumble bee species and their floral host by increasing the likelihood that species will be self-limited by intraspecific competition [@Valdovinos2021-ya].
 
@@
 <div id ="refs"></div>
+
+# FIGURES
+**Figure 1**  
+```{r fig1, echo = FALSE}
+knitr::include_graphics("../ms_2/figures/fig1.pdf")
+```
+
+\newpage
+
+**Figure 2**  
+```{r fig2, echo = FALSE, out.width="100%"}
+knitr::include_graphics("../ms_2/figures/fig2.pdf")
+```
+
+\newpage
+
+**Figure 3**  
+```{r fig3, echo = FALSE, out.width="100%"}
+knitr::include_graphics("../ms_2/figures/fig3.pdf")
+```
+
+\newpage
+
+**Figure 4**  
+```{r fig4, echo = FALSE}
+knitr::include_graphics("../ms_2/figures/fig4.pdf")
+```
+
+\newpage
+
+**Figure 5**  
+```{r fig5, echo = FALSE}
+knitr::include_graphics("../ms_2/figures/fig5.pdf")
+```
+
+\newpage
+
+**Figure 6**  
+```{r fig6, echo = FALSE}
+knitr::include_graphics("../ms_2/figures/fig6.pdf")
+```
+
+\newpage
+
+# FIGURE CAPTIONS
+
+**Figure 1:** Observed visitation rates compared to null model simulations for each bumble bee species and floral morphotype (A).
+Gray bars depict 99% intervals of null model simulations and points depict observed visitation rates.
+Points are color-coded according to where they fall with respect to the 99% interval of the null model: red = above (positive preference), blue = below (negative preference), gray = within (nonsignificant preference).
+The difference between the observed value and the center of the null model interval can be interpreted as magnitude of preference.
+Each floral morphotype is illustrated (B) with a representative species common in our study system: 1 = *Potentilla erecta*, 2 = *Gentiana asclepiadea*, 3 = *Campanula glomerata*, 4 = *Silene nutans*, 5 = *Ajuga reptans*, 6 = *Lotus corniculatus*, 7 = *Centaurea jacea*.
+Illustrations are modifications of public domain works obtained from www.plantillustrations.org.
+For details of original sources, see Appendix X.
+
+\
+
+**Figure 2:** Preference (A) and observed visitation rate (B) of the bumble bee species toward each floral morphotype.
+Each point represents the preference/visitation of one bumble bee species toward the corresponding floral morphotype.
+The color of each point in panel A, retained from Figure 2, indicates whether a given observed bias is significantly (p < 0.01) positive or negative with respect to null model simulations.
+
+\
+
+**Figure 3:** Preference (A) and observed visitation rates (B) for each morphotype regressed against bumble bee tongue length.
+Each point represents the preference/visitation of one bumble bee species toward the corresponding floral morphotype.
+Curves depict fitted GAM smooths with 95\\% confidence bands.
+Solid lines indicate curves differing significantly (p < 0.05) from a flat line while dashed lines indicated nonsigificant curves.
+Points are color-coded by species, and species are presented in the legend in order of increasing tongue length.
+
+\
+
+**Figure 4:** Principal components biplot of bumble bee species according to their preference toward each floral morphotype.
+Floral morphotypes are abbreviated with their numerical codes and illustrated as in Figure 1.
+Colored bands depict a 2-dimensional GAM smooth representing the relationship between the first two principal components and bumble bee tongue length.
+Morphotype illustrations are as described in Figure 1.
+
+\
+
+**Figure 5:** Variation in proportional flower cover through elevation and season modeled as HGAM surfaces.
+Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns.
+All smooths are significant (p < 0.05).
+
+\
+
+**Figure 6:** Variation in preference through elevation and time modeled as HGAM surfaces.
+Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns.
+Significant (p < 0.05) smooths are depicted with solid contour lines and nonsignificant smooths with dashed contour lines.

### 2021-08-25 ms_2/ms_2_supp_methods.Rmd
preped ms for Ecol Lett

@@
 
+Patterns of preference through elevation and time were more complex than those of observed visitation and, by design, more often differed from patterns of floral abundance **(Figure 6)**.
+In some cases, though, patterns of preference mirrored patterns of floral availability, suggesting that bumble bees may exhibit functional responses [@Mysterud1998-rs] to floral abundance not entirely accounted for by null models that assume a linear proportionality between abundance and visitation.
+
 Smooths were significant (p < 0.05) for all bb-fl pairs except hort-bell, hort-head, pasc-bell, telu-head, and soro-flag. The model explained 33.3% of total deviance (adj. R^2^ = 0.272).
@@
 
+Broaching the topic of competition raises an issue that can confound any observational study of resource selection: how to disentangle the endogenous process of resource preference from the exogenous process of niche partitioning driven by competition [@Pianka1974-qk].
+Empirical preference could reflect a latent process of resource preference, as we assume in our interpretation.
+Alternatively, though, visitation patterns and the preference inferred from them could be distorted by competitive forces.
+Intraspecific competition in a bumble bee community would tend to force individuals of a species to broaden their diet beyond what they would prefer under non-competitive conditions, while interspecific competition would be expected to cause a narrowing of diet breadth toward resources that can be exploited most efficiently.
+Thus, observed visitation patterns, even when calculated as abundance-relative preference, can only be interpreted in terms of *preference* to the extent that competition can be assumed to be negligible, and assumption must be justified on the basis of additional knowledge of the study system under consideration.
+Our study system in the Berchtesgadener Alps is one of abundant floral resources but harsh abiotic conditions, so it is likely that competition for floral resources is low, or at least transient, and that bumble bee populations are limited instead by the shortness of the growing season, as has been suggested in the similar environment of the Eyne Valley of France [@Iserbyt2008-jn].
+We cannot, however, exclude the possibility that competitive forces distort, to some degree, the visitation patterns we observed.
+Even when competitive conditions do not obtain universally in a given system, competition may occur in a spatially and/or temporally localized way [@Ranta1981-vx], perhaps associated with extreme climatic events that cause reductions in floral resource avaulability [@Rasmont2012-zq; @Thomson2016-gq; @Phillips2018-me].
+This calls to mind the insight of MacArthur1972-qz, that competition need not be frequent to have lasting ecological impacts, and it is likely that the "ghosts of competition past" [@Connell1980-dn] haunt the floral preferences exhibited by bumble bees today.
+
 # REFERENCES

### 2021-09-02 code/ms1_abundance_v2.Rmd
beginning ms1 revision

--- 2021-04-14 code/ms1_abundance.Rmd
@@
 
-# 1. Prep work
-## 1.1 Prepare environment
+# 1. Prepare environment
+## 1.1 Load packages
 ```{r setup, include=FALSE, echo=FALSE}
@@
 ### BB-FL visitation data
+# net <- read_csv("../data/processed_data/network.csv") %>%
+#   filter(caste != "male",
+#          !bb.sp %in% c("humi", "hypn", "lapi")) %>% # too rare
+#   dplyr::select(site, date, bb.sp, plant.sp) # drop unused columns
+
 net <- read_csv("../data/processed_data/network.csv") %>%
-  filter(caste != "male",
-         !bb.sp %in% c("humi", "hypn", "lapi")) %>% # too rare
+  filter(caste != "male") %>% 
   dplyr::select(site, date, bb.sp, plant.sp) # drop unused columns
@@
 
-## 1.4 Calculate FL and BB abundance
+# 2. Process data
+## 2.1 Calculate transect-wise FL and BB abundance
 ```{r echo=FALSE, include=TRUE, message=FALSE}
+# Survey transects
+survey_transects_site <- survey %>%
+  group_by(site) %>%
+  summarize(survey.transects = length(unique(date)))
+
+# BB transects
+net_transects_site <- net %>%
+  group_by(site) %>%
+  summarize(net.transects = length(unique(date)))
+
 # Per-transect BB abundance
-## full dataset
 bb_abund_sp <- net %>%
@@
   left_join(fl_tax)
+
+### Bumble bee richness per site
+bb_rich_site <- net %>%
+  group_by(site) %>%
+  summarize(bb.rich = length(unique(bb.sp))) %>%
+  left_join(site_data) %>% # add site data
+  mutate(management = factor(management),
+         site = factor(site),
+         transect = factor(transect))
+
+### Peak and mean BB abundance per site
+bb_abund_site <- bb_abund_sp %>%
+  group_by(site, date) %>%
+  summarize(bb.total.abund = sum(bb.abund)) %>%
+  group_by(site) %>%
+  summarize(bb.peakabund = max(bb.total.abund)) %>%
+  left_join(site_data) %>% # add site data
+  mutate(management = factor(management),
+         site = factor(site),
+         transect = factor(transect))
+
+### Floral richness per site
+fl_rich_site <- survey %>%
+  group_by(site) %>%
+  summarize(fl.rich = length(unique(plant.sp))) %>%
+  left_join(site_data) %>% # add site data
+  mutate(management = factor(management),
+         site = factor(site),
+         transect = factor(transect))
+
+### Peak floral abundance per site
+fl_abund_site <- survey %>%
+  group_by(site, date) %>%
+  summarize(fl.abund = sum(flower.cover)) %>%
+  group_by(site) %>%
+  summarize(fl.peakabund = max(fl.abund),
+            fl.meanabund = mean(fl.abund),
+            fl.medabund = median(fl.abund)) %>%
+  left_join(site_data) %>% # add site data
+  mutate(management = factor(management),
+         site = factor(site),
+         transect = factor(transect))
+
+### Interaction richness per site
+interaction_rich_site <- net %>%
+  group_by(site) %>%
+  unite(interaction, c(bb.sp, plant.sp), sep = "_") %>%
+  summarize(interaction.rich = length(unique(interaction))) %>%
+  left_join(site_data) %>% # add site data
+  mutate(management = factor(management),
+         site = factor(site),
+         transect = factor(transect))
+
+rich_abund_collated <- bb_rich_site %>% 
+  full_join(fl_rich_site) %>%
+  full_join(bb_abund_site) %>% 
+  full_join(fl_abund_site) %>%
+  full_join(interaction_rich_site)
+
+rich_abund_collated.long <- rich_abund_collated %>%
+  pivot_longer(cols = c(bb.rich, fl.rich, bb.peakabund, 
+                        fl.peakabund, fl.meanabund, 
+                        fl.medabund, interaction.rich)) %>%
+  separate(name, c("group", "metric"), sep = "\\.") %>%
+  group_by(group, metric) %>%
+  mutate(z = scale(value))
+
+ggplot(rich_abund_collated.long, aes(elev.mean, z, color = group)) +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
+  facet_wrap(~metric) +
+  theme_light()
+
+# Exploring some bivariate relations
+
+ggplot(rich_abund_collated, aes(fl.rich, bb.rich)) +
+  geom_point() +
+  geom_smooth(method = "lm")
+
+ggplot(rich_abund_collated, aes(fl.peakabund, fl.rich)) +
+  geom_point() +
+  geom_smooth(method = "lm")
+
+ggplot(rich_abund_collated, aes(fl.peakabund, bb.peakabund)) +
+  geom_point() +
+  geom_smooth(method = "lm")
+
+ggplot(rich_abund_collated, aes(fl.peakabund, bb.rich)) +
+  geom_point() +
+  geom_smooth(method = "lm") 
 ```
 
-## 1.5 Calculate binary "preference" for each plant genus by each bumble bee species; score species 0 or 1 based on whether they were visited
-```{r message=FALSE}
-weights <- fl_abund_sp %>%
-  select(plant.sp, plant.genus) %>%
-  unique() %>%
-  left_join(select(net, bb.sp, plant.sp)) %>%
-  select(plant.genus, bb.sp) %>%
-  distinct() %>%
-  mutate(visited = rep(1, n())) %>%
-  pivot_wider(values_from = visited, names_from = bb.sp, values_fill = 0) %>%
-  pivot_longer(cols = 2:14, values_to = "visited", names_to = "bb.sp")
-
-weighted_fl_abund <- fl_abund_sp %>%
-  group_by(site, date, plant.genus) %>%
-  summarize(flower.cover = sum(fl.abund)) %>%
-  left_join(weights) %>%
-  mutate(weighted.flower.cover = flower.cover*visited) %>%
-  group_by(site, date, bb.sp) %>%
-  summarize(total.weighted.flower.cover = sum(weighted.flower.cover)) %>%
-  left_join(site_data) %>%
-  select(site, date, elev.mean, bb.sp, total.weighted.flower.cover) %>%
-  mutate(yday = yday(date),
-         bb.sp = factor(bb.sp),
-         site = factor(site),
-         year = factor(year(date)))
-```
+## 2.2 Summarize
 
@@
 ```
+
+## 2.3 Mean total floral cover (not broken down by bumble bee species) ~ elevation * time
+```{r}
+fl_abund_total <- fl_abund_sp %>%
+  group_by(site, date) %>%
+  summarize(fl.abund = sum(fl.abund)) %>%
+  left_join(site_data) %>%
+  mutate(yday = yday(date),
+         year = factor(year(date)),
+         site = factor(site),
+         management = factor(management))
+
+bb_abund_total <- net %>%
+  group_by(site, date) %>%
+  summarize(bb.abund = n()) %>%
+  left_join(site_data) %>%
+  mutate(yday = yday(date),
+         year = factor(year(date)),
+         site = factor(site),
+         management = factor(management))
+
+fl_rich_total <- fl_abund_sp %>%
+  filter(fl.abund > 0) %>%
+  group_by(site, date) %>%
+  summarize(fl.rich = length(unique(plant.sp))) %>%
+  left_join(site_data) %>%
+  mutate(yday = yday(date),
+         year = factor(year(date)),
+         site = factor(site),
+         management = factor(management))
+
+bb_rich_total <- net %>%
+  group_by(site, date) %>%
+  summarize(bb.rich = length(unique(bb.sp))) %>%
+  left_join(site_data) %>%
+  mutate(yday = yday(date),
+         year = factor(year(date)),
+         site = factor(site),
+         management = factor(management))
+
+fl_rich_total <- fl_abund_sp %>%
+  filter(fl.abund > 0) %>%
+  group_by(site, date) %>%
+  summarize(fl.rich = length(unique(plant.sp))) %>%
+  left_join(site_data) %>%
+  mutate(yday = yday(date),
+         year = factor(year(date)),
+         site = factor(site),
+         management = factor(management))
+
+ggplot(fl_abund_total, aes(fl.abund)) +
+  geom_histogram() +
+  facet_wrap(~site)
+
+ggplot(fl_rich_total, aes(fl.rich)) +
+  geom_histogram() +
+  facet_wrap(~site)
+
+### Abundance
+total_fl_abund_GAM <- gam(log(fl.abund + 0.01) ~ 
+                             year +
+                             management +
+                             s(site, bs = "re") +
+                             te(yday, elev.mean, by = year, bs=c("gp","tp")),
+                           family = "gaussian",
+                           method = "REML",
+                           data = fl_abund_total) %>% getViz()
+
+check.gamViz(total_fl_abund_GAM)
+summary(total_fl_abund_GAM)
+plot.gamViz(total_fl_abund_GAM, allTerms = TRUE)
+
+total_bb_abund_GAM <- gam(bb.abund ~ 
+                             year +
+                             management +
+                             s(site, bs = "re") +
+                             te(yday, elev.mean, by = year, bs=c("gp","tp")),
+                           family = "nb",
+                           method = "REML",
+                           data = bb_abund_total) %>% getViz()
+
+check.gamViz(total_bb_abund_GAM)
+summary(total_bb_abund_GAM)
+plot.gamViz(total_bb_abund_GAM, allTerms = TRUE)
+
+### Richness
+
+total_fl_rich_GAM <- gam(fl.rich ~ 
+                            year +
+                            management +
+                            s(site, bs = "re") +
+                            te(yday, elev.mean, by = year, bs=c("gp","tp")),
+                          family = "poisson",
+                          method = "REML",
+                          data = fl_rich_total) %>% getViz()
+
+check.gamViz(total_fl_rich_GAM)
+summary(total_fl_rich_GAM)
+plot.gamViz(total_fl_rich_GAM, allTerms = TRUE)
+
+total_bb_rich_GAM <- gam(bb.rich ~ 
+                            year +
+                            management +
+                            s(site, bs = "re") +
+                            te(yday, elev.mean, by = year, bs=c("gp","tp")),
+                          family = "nb",
+                          method = "REML",
+                          data = bb_rich_total) %>% getViz()
+
+check.gamViz(total_bb_rich_GAM)
+summary(total_bb_rich_GAM)
+plot.gamViz(total_bb_rich_GAM, allTerms = TRUE)
+```

### 2021-09-02 code/ms1_beta_diversity.Rmd
beginning ms1 revision

@@
   dplyr::select(-geometry)
-
-# Write to csv for archiving
-write_csv(site_data, "../Dryad_ms1/site_data.csv")
 ```
@@
   dplyr::select(site, date, plant.sp, flower.cover) 
-
-# Write to csv for archiving
-write_csv(net, "../Dryad_ms1//visitation_data.csv")
-write_csv(survey, "../Dryad_ms1//floral_survey_data.csv")
 ```
@@
 # Write to csv for archiving
-write_csv(beta_mit_data, "../Dryad_ms1/partitioned_beta_diversity")
+#write_csv(beta_mit_data, "../Dryad_ms1/partitioned_beta_diversity")
 ```
@@
   xlab("Elevation difference (m)") +
-  ylab("β-diversity") +
+  ylab("β-diversity (proportional species turnover)") +
   theme_light()
@@
 # Patchwork plot for comparing species beta-diversity with interaction beta-diversity
-png("../output/beta_stack.png", width = 6.5, height = 4.5, units = "in", res = 300)
+png("../ms_1/figures/beta_stack.png", width = 6.5, height = 4.5, units = "in", res = 300)
 (g1 | (g2 / g3)) + 
@@
 
-pdf("../output/beta_stack.pdf", width = 6.5, height = 4.5)
+pdf("../ms_1/figures/beta_stack.pdf", width = 6.5, height = 4.5)
 (g1 | (g2 / g3)) + 
@@
 sitepair_WN <- interaction_gdm_prep(net)
+sitepair_WNb <- interaction_gdm_prep2(net)
 
 # Run GDM analysis
-gdm_WN <- gdm(sitepair_WN, geo = FALSE)
+gdm_WN6 <- gdm(sitepair_WN, geo = FALSE, splines = c(6, 6, 6, 6, 6))
+gdm_WN6b <- gdm(sitepair_WNb, geo = FALSE, splines = c(6, 6, 6, 6, 6))
+
+gdm_WN6 <- gdm(sitepair_WN, geo = TRUE, splines = c(6, 6, 6, 6, 6, 6))
+
 summary(gdm_WN)
+
 gdm.varImp_WN <- gdm.varImp(sitepair_WN, geo = FALSE)
+gdm.varImp_WN <- gdm.varImp(sitepair_WN, geo = FALSE, splines = c(6, 6, 6, 6, 6))
+gdm.varImp_WN <- gdm.varImp(sitepair_WN, geo = FALSE, splines = c(6, 6, 6, 6))
+gdm.varImp_WN <- gdm.varImp(sitepair_WN, geo = FALSE, splines = c(6, 6, 6))
+
+# gdm.varImp_WN <- gdm.varImp(sitepair_WN, geo = FALSE, splines = c(6, 6, 6, 6))
 
 # Plot
-plot(gdm_WN)
+plot(gdm_WN6)
+plot(gdm_WN6b)
 
@@
 # Call GDMs
-gdm_survey <- gdm(sitepair_survey, geo = FALSE)
+gdm_survey10 <- gdm(sitepair_survey, geo = TRUE, splines = c(10, 10, 10, 10, 10, 10))
+gdm_survey6 <- gdm(sitepair_survey, geo = FALSE, splines = c(6, 6, 6, 6, 6))
+gdm_survey5 <- gdm(sitepair_survey, geo = TRUE, splines = c(5, 5, 5, 5, 5, 5))
+gdm_survey4 <- gdm(sitepair_survey, geo = TRUE, splines = c(4, 4, 4, 4, 4, 4))
+
 summary(gdm_survey)
-gdm.varImp_survey <- gdm.varImp(sitepair_survey, geo = FALSE)
+gdm.varImp_survey <- gdm.varImp(sitepair_survey, geo = FALSE, splines = c(6, 6, 6, 6, 6, 6))
 
 # Plot
-plot(gdm_survey)
+plot(gdm_survey10)
+plot(gdm_survey6)
+plot(gdm_survey5)
+plot(gdm_survey4)
 
@@
 # Call GDMs
-gdm_BB <- gdm(sitepair_BB, geo = FALSE)
+gdm_BB6 <- gdm(sitepair_BB, geo = FALSE, splines = c(6, 6, 6, 6, 6))
+gdm_BB5 <- gdm(sitepair_BB, geo = FALSE, splines = c(5, 5, 5, 5, 5, 5))
+gdm_BB4 <- gdm(sitepair_BB, geo = TRUE, splines = c(4, 4, 4, 4, 4, 4))
+
 summary(gdm_BB)
-gdm.varImp_BB <- gdm.varImp(sitepair_BB, geo = FALSE)
+gdm.varImp_BB <- gdm.varImp(sitepair_BB, geo = TRUE, splines = c(6, 6, 6, 6, 6, 6))
 
 # Plot
-plot(gdm_BB)
+plot(gdm_BB6)
+plot(gdm_BB5)
+plot(gdm_BB4)
 
@@
 
+### Fixed isplineExtract()
+```{r}
+##Extracts Ispline data from a gdm model
+isplineExtract2 <- function (model){
+  ###########################
+  #model = gdmOb
+  ###########################
+  ##error checking
+  ##checks to make sure a gdm model is given
+  if(!is(model, "gdm")){
+    stop("model argument must be of class = 'gdm'.")
+  }
+
+  ##Collects or sets simple data
+  options(warn.FPU = FALSE)
+  PSAMPLE <- 200
+  preddata <- rep(0, times = PSAMPLE)
+  pn <- model$predictors
+  nPreds <- length(pn)
+  yDat <- xDat <- matrix(0,PSAMPLE,nPreds)
+  colnames(yDat) <- colnames(xDat) <- pn
+  pmin <- 1
+  pmax <- PSAMPLE
+
+  ##cycles through each prodictor and fills the spline matrices
+  splineindex <- 1
+  for (i in 1:nPreds){ 
+    #i<-1
+    numsplines <- model$splines[i]
+    z <- .C("GetPredictorPlotData", 
+            pdata = as.double(preddata), 
+            as.integer(PSAMPLE), 
+            as.double(model$coefficients[splineindex:(splineindex + numsplines - 1)]), 
+            as.double(model$knots[splineindex:(splineindex + numsplines - 1)]), 
+            as.integer(numsplines),
+            PACKAGE = "gdm")
+    yDat[,i] <- z$pdata
+    pmin <- pmin + PSAMPLE
+    pmax <- pmax + PSAMPLE
+    xDat[,i] <-  seq(from=model$knots[[(i*numsplines)-(numsplines-1)]],to=model$knots[[(i*numsplines)]], length=PSAMPLE)
+    splineindex <- splineindex + numsplines
+  }
+
+  ##lists and returns matrices
+  outData <- list(x=xDat,y=yDat)
+  return(outData)
+}
+```
+
 ### Plot GDM splines
@@
 # Interaction model
-gdm_WN_splines1 <- isplineExtract(gdm_WN)$x %>%
+gdm_WN_splines1 <- isplineExtract2(gdm_WN6)$x %>%
   as.data.frame() %>%
-  select(elevation = elev.mean)
+  dplyr::select(elevation = elev.mean)
 
-gdm_WN_splines2 <- isplineExtract(gdm_WN)$y %>%
+gdm_WN_splines2 <- isplineExtract2(gdm_WN6)$y %>%
   as.data.frame() %>%
-  select(response = elev.mean)
+  dplyr::select(response = elev.mean)
 
@@
 # Flora model
-gdm_survey_splines1 <- isplineExtract(gdm_survey)$x %>%
+gdm_survey_splines1 <- isplineExtract2(gdm_survey6)$x %>%
   as.data.frame() %>%
-  select(elevation = elev.mean)
+  dplyr::select(elevation = elev.mean)
 
-gdm_survey_splines2 <- isplineExtract(gdm_survey)$y %>%
+gdm_survey_splines2 <- isplineExtract2(gdm_survey6)$y %>%
   as.data.frame() %>%
-  select(response = elev.mean)
+  dplyr::select(response = elev.mean)
 
@@
 # BB model
-gdm_BB_splines1 <- isplineExtract(gdm_BB)$x %>%
+gdm_BB_splines1 <- isplineExtract2(gdm_BB6)$x %>%
   as.data.frame() %>%
-  select(elevation = elev.mean)
+  dplyr::select(elevation = elev.mean)
 
-gdm_BB_splines2 <- isplineExtract(gdm_BB)$y %>%
+gdm_BB_splines2 <- isplineExtract2(gdm_BB6)$y %>%
   as.data.frame() %>%
-  select(response = elev.mean)
+  dplyr::select(response = elev.mean)
 
@@
 tree_line <- site_data %>%
-  select(elevation = elev.mean, tree_line)
+  dplyr::select(elevation = elev.mean, tree_line)
 
@@
 
-ggsave("../output/gdm_splines.pdf")
-ggsave("../output/gdm_splines.png")
+ggsave("../output/gdm_splines2.pdf")
+ggsave("../output/gdm_splines2.png")
 

### 2021-09-02 code/ms1_sampling.Rmd
beginning ms1 revision

--- 2021-04-14 ms_1/Appendix_S1.Rmd
@@
 ---
-title: "Appendix S1"
-output: pdf_document
+title: "sampling"
+output: html_document
 ---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+Sys.setlocale(locale = "en_US.UTF-8")
+```
 
@@
 library(lubridate)
+library(patchwork)
 library(tidyverse)
@@
 ```{r message=FALSE, warning=FALSE}
-ggplot(visitation_sampling, aes(x = "x")) +
+samplingplot1 <- ggplot(visitation_sampling, aes(x = year(date))) +
   geom_bar(fill = "gray60") +
-  ylab("Observations") +
-  facet_wrap(~year(date)) +
-  theme_light(12) +
-  theme(axis.title.x = element_blank(),
-        axis.text.x = element_blank(),
-        axis.ticks.x = element_blank()) +
-  ggtitle("Bumble bee observations per study year")
+  ylab("Total bumble bees\nobserved") +
+  theme_light(11) +
+  theme(axis.title.x = element_blank())
 
-ggplot(bb_obs_per_sample, aes(x = "x")) +
+samplingplot2 <- ggplot(bb_obs_per_sample, aes(x = year)) +
   geom_bar(fill = "gray60") +
-  ylab("Samples") +
-  facet_wrap(~year) +
-  theme_light(12) +
-  theme(axis.title.x = element_blank(),
-        axis.text.x = element_blank(),
-        axis.ticks.x = element_blank()) +
-  ggtitle("Sampling events (site/day) per year")
+  ylab("Total samples") +
+  theme_light(11) +
+  theme(axis.title.x = element_blank())
 
-ggplot(bb_obs_per_sample, aes(x = "x", y = obs)) +
+samplingplot3 <- ggplot(bb_obs_per_sample, aes(x = year, y = obs)) +
   geom_violin(draw_quantiles = 0.5, fill = "gray60", color = NA) +
@@
   #geom_density(outline.type = "full", position = "fill") +
-  ylab("Observations/sample") +
+  ylab("Bumble bees\nobserved / sample") +
   #xlab(NULL) +
-  facet_wrap(~year) +
-  theme_light(12) +
-  theme(axis.title.x = element_blank(),
-        axis.text.x = element_blank(),
-        axis.ticks.x = element_blank()) +
-  ggtitle("Bumble bee observations per sampling event")
+  theme_light(11) +
+  theme(axis.title.x = element_blank())
 
-ggplot(bb_obs_per_sample, aes(as_date(yday(date)), elev.mean/10)) +
+samplingplot4 <- ggplot(bb_obs_per_sample, aes(as_date(yday(date)), elev.mean/10)) +
   geom_point(size = 1, aes(alpha = obs)) +
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  scale_alpha_continuous(name = "Observations") +
+  scale_alpha_continuous(name = "Bumble bees\nobserved") +
   scale_x_date(date_labels = "%b", breaks = as_date(c("1970-04-01", "1970-05-01", "1970-06-01",
                                                     "1970-07-01", "1970-08-01", "1970-09-01"))) +
-  theme_light(12) +
+  theme_light(11) +
   labs(x = NULL,
-       y = "Elevation (m x 10)") +
-  ggtitle("Bumble bee observations through elevation and time")
+       y = "Elevation (m x 10)")
+
+samplingplot <- ((samplingplot1 / samplingplot2 / samplingplot3) | samplingplot4) +
+  plot_layout(widths = c(1, 3)) + 
+  plot_annotation(tag_levels = 'A')
+
+ggsave("../ms_1/figures/samplingplot.pdf", height = 6, width = 10.5)
+ggsave("../ms_1/figures/samplingplot.png", height = 6, width = 10.5)
 ```
-
-# Bumble bee abundance
-```{r message=FALSE, warning=FALSE}
-bb_range <- visitation_sampling %>%
-  left_join(site_data) %>%
-  group_by(site, date, elev.mean, tree_line, bb.sp) %>%
-  summarize(abund = n()) %>%
-  group_by(site, elev.mean, tree_line, bb.sp) %>%
-  summarize(abund = max(abund)) %>%
-  filter(abund > 0) %>%
-  group_by(bb.sp) %>%
-  mutate(elev.floor = min(elev.mean),
-         elev.ceiling = max(elev.mean),
-         elev.range = elev.ceiling - elev.floor,
-         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))
-
-ggplot(bb_range, aes(reorder(bb.sp, elev.med), elev.mean)) +
-  geom_line(size = 2, alpha = 0.25) +
-  geom_point(aes(size = abund, color = tree_line), alpha = 0.5) +
-  theme_light(14) +
-  scale_size_continuous(name = "Peak abundance",
-                        breaks = c(1, 10, 20, 40, 80)) +
-  scale_color_discrete(name = "Tree line") +
-  xlab("Bumble bee species") +
-  ylab("Elevation (m)") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  ggtitle("Range and peak abundance of each bumble bee sp.")
-
-# Overall bumble bee abundance
-bb_abund_sp <- visitation_sampling %>%
-  dplyr::select(site, date, bb.sp) %>%
-  group_by(site, date, bb.sp) %>% # 
-  summarize(bb.abund = n()) %>% # get per transect bb abundance
-  ungroup() %>%
-  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
-  semi_join(visitation_sampling, by = c("site", "date")) %>% # drop site-dates that did not exist in the data set
-  left_join(site_data) %>% # add site data
-  mutate(year = factor(year(date)),
-         yday = yday(date),
-         bb.sp = factor(bb.sp),
-         management = factor(management),
-         site = factor(site),
-         transect = factor(transect))
-
-bb_abund_sum <- bb_abund_sp %>%
-  group_by(site, bb.sp) %>%
-  summarize(cum.abund = sum(bb.abund),
-            mean.abund = mean(bb.abund)) %>%
-  group_by(site) %>%
-  mutate(prop.abund = cum.abund/sum(cum.abund)) %>%
-  group_by(bb.sp) %>%
-  mutate(mean.prop.abund = mean(prop.abund))
-
-ggplot(bb_abund_sum, aes(reorder(bb.sp, -mean.prop.abund), prop.abund)) +
-  geom_boxplot(fill = "gray85") +
-  theme_light() +
-  ylab("Proportional abundance") +
-  xlab("Bumble bee species") +
-  ggtitle("Relative abundances of bumble bee species")
-```
-
-# GAM diagnostics
-
-## Bumble bee abundance GAM
-```{r message=FALSE, warning=FALSE}
-bb_abund_gam <- readRDS("../output/gam_max_bb_abund_01.rds")
-summary(bb_abund_gam)
-check.gamViz
-```
-
-## Floral abundance GAM
-```{r message=FALSE, warning=FALSE}
-fl_abund_gam <- readRDS("../output/gam_mean_fl_abund_01.rds")
-summary(fl_abund_gam)
-check.gamViz(fl_abund_gam)
-```

### 2021-09-02 ms_1/Appendix_S1.Rmd
beginning ms1 revision

@@
 ---
-title: "Appendix S1"
-output: pdf_document
+title: "The tree line ecotone punctuates the abundance and β-diversity of plants, pollinators, and their interactions"
+subtitle: "Appendix S1: Analytical workflow"
+author: 
+  - Douglas B. Sponsler^[douglas.sponsler@uni-wuerzburg.de]
+  - Katharina Kallnik 
+  - Fabrice Requier 
+  - Alice Classen 
+  - A. Fabienne Maihoff
+  - Johanna Sieger
+  - Ingolf Steffan-Dewenter
+output: 
+  pdf_document:
+    toc: true
+    toc_depth: 3
+    latex_engine: xelatex
+#mainfont: Arial
 ---
@@
         axis.ticks.x = element_blank()) +
-  ggtitle("Sampling events (site/day) per year")
+  ggtitle("Samplings (site*day) per year")
 
@@
         axis.ticks.x = element_blank()) +
-  ggtitle("Bumble bee observations per sampling event")
+  ggtitle("Bumble bee observations per sample")
 
@@
 summary(bb_abund_gam)
-check.gamViz
+check.gamViz(bb_abund_gam)
 ```

### 2021-09-02 ms_1/ms_revision1.Rmd
beginning ms1 revision

--- 2021-06-09 ms_1/ms_1.Rmd
@@
 ---
-title: "The tree line ecotone punctuates the abundance and turnover of plants, pollinators, and their interactions"
+title: "The elevational structure of a mountain mutualism: modes and thresholds in the responses of bumble bees, wildflowers, and their interactions"
 
@@
 
-The structuring of biological communities along mountain slopes is complex, and elevational range shifts in response to climate change involve more than merely tracking suitable temperature envelopes.
-When species move, they do so in the context of biological communities, and the outcomes of these movements depend on how and to what extent biotic interactions are reordered.
-Bumble bees (Hymentopera: *Bombus* spp.) are cold-adapted species associated with mountain habitats, and they are already exhibiting upslope range shifts that are expected to result in habitat loss, novel competitive interactions, and the rewiring of pollination networks.
-Predicting and interpreting these shifts, however, requires an understanding of the current elevational patterns of bumble bees and their floral mutualists that are being acted upon by climate change.
-We recorded bumble- bee-flower interactions over three years along an 1400 m elevational gradient in the German Alps.
-Using nonlinear modeling, we analyze the elevational patterns at the nested levels of species abundance, species β-diversity, and interaction β-diversity.
-We demonstrate that the tree line ecotone is (1) a distributional interface between low/mid- and high-elevation bumble bee species, (2) a threshold above which floral resource availability sharply decreases, and (3) a zone of accelerated turnover of floral composition and bumble- bee-flower interactions.
-The implications of these findings extend beyond the particular case of bumble bees to demonstrate that linear elevational temperature gradients are ecologically punctuated, and the outcomes of climate-induced range shifts will depend on dynamics at the tree line ecotone.
-
 *Keywords*: climate, beta-diversity, elevation gradient, *Bombus*, bumble bee, mountain ecology, pollination network
@@
 
-Mountains are unique ecological theaters in which global climate dynamics interact with local climate gradation to shape the distributions of species and biotic interactions [@Telwala2013-sc; @CaraDonna2014-nq; @Miller-Struttmann2014-wz; Peters2019-km; @Rafferty2020-aj].
+1. Biological responses to environmental gradients
+
+Environmental gradients regulate the assembly of biological communities, and they have likely played a central role in generating and maintaining Earth's biological diversity [@Hutchinson1953-mo].
+Patterns of species- and community-level responses to environmental gradients can shed light on the underlying mechanisms that govern community assembly and function [e.g.
+@Beals1969-in].
+Such insight is important for anticipating and managing the effects of global ecological change that modify existing environmental gradients together with the species that inhabit them.
+
+2. The particular case of mountain elevation gradients
+
+Mountain slopes generate steep environmental gradients over short geographical distances, facilitating the assembly of biological communities with high species richness and turnover[@Rahbek2019-sw].
+The response of species richness and turnover to mountain elevation gradients
+
+*Species richness*
+
+Species richness along mountain slopes tends to show either a mid-elevation peak or a monotonic decline [@Rahbek1995-oj; @Guo2013-tu], though responses can vary considerably across taxonomic groups [e.g.
+@Peters2016-hf].
+
+*Species turnover*
+
+In some cases, continuous elevation gradients can generate punctuated shifts in species composition. as has been demonstrated in some mountain plant communities [@Beals1969-in; @Hemp2006-bw]
+
+Something about productivity ~ elevation
+
+Maybe something about abundance ~ elevation
+
+<!-- Vegetational Change Along Altitudinal Gradients; controversy regarding whether vegetation gradients are continuous or discrete, suggests something in between, consistent with the "threshold" that I see associated with the tree line.
+[@Beals1969-in] -->
+
+<!-- *A "though see" citation:* The ecological and floristic importance of the forest limit in phytogeographical zonations has remained un- questioned,   even though its  value as an important ecological boundary has  rarely been verified.
+An interesting question is therefore  whether  the  presence or absence of birch  forests in Aurland is associated with any abrupt discontinuity in floristic richness or composition ...
+Our results from Aurland suggest that the forest limit is  not associated with  any major floristic change,  either in species turnover or in species richness, but is  associated with some  change in overall species composition.
+Hofgaard (1997) similarly found little difference in beta diversity or floristic composition 300 m above or below the tree line in the central Norwegian mountains.
+-->
+
+3. From species to interactions: less studied, but amenable to the same tools used at the species-level [see references in @Simanonok2014-ou]
+
+While elevational patterns in gross summaries of ecological communities (e.g.
+species richness, net primary productivity) are valuable, potentially deeper insight can be gained by the more focused study of elevational patterns among guilds of species with known functional relationships, such as plants and herbivores or plants and pollinators.
+
+An elevational study of plants and insect herbivores [@Rasmann2014-oc]
+
+4. Mountain bumble bees and wildflowers
+  A. Opportunity to study responses not just at the level of species, but at the level of interactions [@Simanonok2014-ou; @Burkle2016-tq].
+  B. Significance of understanding the elevation structure of mountain BB-wildflower communities vis-a-vis climate change and meadow reforestation
+
+5. What we did
+
+Definitely need to discuss @Hoiss2012-sn and @Hoiss2012-kr and @Hoiss2015-gv
+
+The response-effect framework might be useful [@Suding2008-bl].
+
+<!-- First introduce the concept of ecological thresholds, nonlinear responses. Give some illustrative examples to show the general nature of the phenomenon.  -->
+
+Concept paper on ecological thresholds [@Groffman2006-ch].
+
+Threshold response of stream communities to *urbanization* gradient [@King2011-ry].
+
+Threshold response of anurans to proximity to highways [@Eigenbrod2009-zl].
+
+Nonlinear and threshold responses of nitrogen cycling to *aridity* gradient [@Wang2014-fh].
+
+Threshold response of marine fungal community composition to *salinity* gradient [@Rojas-Jimenez2019-rm].
+
+Threshold and nonlinear responses of marine benthic community to *disturbance* (sheer stress from waves, trawling, etc.) gradients [@Couce2020-hu].
+
+Threshold response of bat communities to habitat loss gradient [@Muylaert2016-rj].
+
+Mountains are unique ecological theaters in which global climate dynamics interact with local climate gradation to shape the distributions of species and biotic interactions [@Telwala2013-sc; @CaraDonna2014-nq; @Miller-Struttmann2014-wz; @Peters2019-km; @Rafferty2020-aj].
 The rate of recent climate change, however, is forcing species to adapt rapidly within the constraints of gene pools shaped by cooler climates [@Visser2008-ct].
@@
 Bumble bees (Hymenoptera: *Bombus* spp.) are classic model organisms in the field of community ecology, and they have more recently become conservation priorities due to well-documented population declines and vulnerability to climate change [@Goulson2008-oe; @Soroye2020-zb].
-Bumble bees are characteristically cold-adapted species [@Heinrich1994-cv], with peak abundance and diversity in mountain ranges and northern latitudes, where they are often the dominant guild of flower-visiting insects and the principal pollinators of entomophilous flora [@Williams1998-xi; @Goulson2010-lo].
+Bumble bees are characteristically cold-adapted [@Heinrich1994-cv], with peak abundance and diversity in mountain ranges and northern latitudes, where they are often the dominant guild of flower-visiting insects and the principal pollinators of entomophilous flora [@Williams1998-xi; @Goulson2010-lo].
 Upslope movement of bumble bee species in response to climate change has already been documented, and this is expected to cause the declines or extinction of some bumble bee species [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; Biella2017-ni; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc].
@@
 In the present study, set along a ~1400 m elevation gradient in the German Alps, we investigate elevational patterns in a community of bumble bees and wildflowers at three levels of biological organization.
-First, we ask how elevation shapes the abundances of bumble bee species and their floral resources.
+First, we ask how elevation shapes the abundance and richness of bumble bees and their floral resources.
 Then, we move from the level of individual species to that of community composition and explore the elevational structure of species turnover (i.e.
@@
 Bumblebees observed on a given flower were counted as floral visitors.
-Bumble bee queens were identified to species level in the field, while workers and males were stored in individually labeled tubes in the freezer for later identification in the laboratory after [@Amiet1996-jz].
-Floral visitation by male bumble bees was recorded during visitation sampling, but we chose to analyze only visitation by queens and workers; males are often found resting on flowers, so their "visits" cannot always be interpreted in the same way as those of queens and workers.
+Bumble bee queens were identified to species level in the field, while workers and males were stored in individually labeled tubes in the freezer for later identification in the laboratory according to [@Amiet1996-jz].
+Floral visitation by male bumble bees was recorded during visitation sampling, but we chose to analyze only visitation by queens and workers.
+Males are often found resting on flowers, particularly late in the year, so their presence on a flower cannot be interpreted reliably as a foraging visit.
 In conjunction with visitation observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.01 m^2^.
@@
 Although per-sample effort was standardized across all years, the total number of samples (i.e.
-site-dates) varied: 153 in 2010, 212 in 2011, and 539 in 2012.
+site-dates) varied: 153 in 2010, 212 in 2011, and 539 in 2012 **(Figure 2)**.
 Since the start of field sampling each year was determined by snowmelt, lower sites were sampled as early as April while higher sites could not be sampled until June or July.
 This resulted in a temporally staggered pattern of sampling in 2010 and 2012, while in 2011 a combination of early snowmelt and a late start to field activities resulted in sampling that was approximately aligned in time across the whole elevation gradient.
-See Appendix S1 for details of sampling patterns.
 
@@
 This approach was intended to dampen the effects of false non-detection, particularly for rare bumble bee species whose diet breadth would tend to be underestimated simply due to sparsity of observations [@Williams2005-yp].
-We treated the cryptic *B.
-terrestris/lucorum* species pair as a single morphospecies, since these two species cannot be visually distinguished, and we pooled representatives of the parasitic *Psithyrus* subgenus --- *B.
-barbutellus*, *B.
+We treated the cryptic *Bombus terrestris/lucorum* species pair as a single morphospecies, since these two species cannot be visually distinguished, and we pooled representatives of the parasitic *Psithyrus* subgenus --- *Bombus barbutellus*, *B.
 bohemicus*, *B.
@@
 psithryus*.
-*B.
-hypnorum*, *B.
+*Bombus hypnorum*, *B.
 humilis*, and *B.
@@
 To focus on elevational patterns, we collapsed the temporal variation of abundance within sites into summary values.
-For bumble bees, we modeled *peak* abundance at each site as a smooth response to `elevation`, with bumble bee `species` as a smoothing factor, `management` (grazing, mowing, or unmanaged) as a fixed effect, and `species:year` as a random intercept term (exploratory analysis indicated that species varied differently across years).
+For bumble bees, we modeled *peak* abundance at each site as a smooth response to `elevation`, with a group-level smoother for each bumble bee `species`, `management` (grazing, mowing, or unmanaged) as a fixed effect, and `species:year` as a random intercept term (exploratory analysis indicated that species varied differently across years).
 For floral resources, we modeled *mean* flower cover, weighted per bumble bee species (i.e.
@@
 To investigate the relationship between elevation and β-diversity more deeply, we performed a second analysis using generalized dissimilarity modeling (GDM) [@Ferrier2007-ji] to analyze the elevational variation in the β-diversity of bumble bees, flora, and (unpartitioned) interactions, respectively.
-In addition to estimating the relationship between differences in a response variable and differences in a predictor variables, GDM captures the slope of this relationship over the range of the predictor variable, revealing potential variation in the amount of change in the response induced by a given change in the predictor [@Ferrier2007-ji].
+In addition to estimating the relationship between differences in a response variable and differences in a predictor variable, GDM captures the slope of this relationship over the range of the predictor variable, revealing potential variation in the amount of change in the response induced by a given change in the predictor [@Ferrier2007-ji].
 This enabled us to ask whether the relationship between β-diversity and elevation exhibits thresholds or other nonlinearities.
@@
 The metaweb across all sites and dates consisted of 16 bumble bee species, 163 plant species (110 genera, 37 families), and 736 unique bumble- bee-plant interaction pairs.
-*B.
-pascuorum* ("pasc"), *B.
+*Bombus pascuorum* ("pasc"), *B.
 pratorum* ("prat"), *B.
@@
 gerstaeckeri* ("gers") were present at moderate to low abundance.
-*B.
-hypnorum*, *B.
+*Bombus hypnorum*, *B.
 humilis*, and *B.
@@
 HGAM analysis revealed that the response of peak abundance to elevation differed significantly (p ≤ 0.02) from a flat line for all bumble bee species **(Figure 2A)**.
-*B.
-mendax*, *B.
+*Bombus mendax*, *B.
 monticola*, *B.
@@
 pascuorum*, likely because its exceptionally long nesting cycle is incompatible with the short growing season at high elevations [@Neumayer2009-gx].
-*B.
-gerstaeckeri*, *B.
+*Bombus gerstaeckeri*, *B.
 jonellus*, and *B.
@@
 Our model explained 71% of total variance (adjusted R^2^ = 0.45).
-Neither mowing (p = 0.08) nor non-management (p = 0.95) had a significant effect on bumble bee abundance relative to the reference level of grasing management.
+Neither mowing (p = 0.08) nor non-management (p = 0.95) had a significant effect on bumble bee abundance relative to the reference level of grazing management.
 The random effect of species:year was also nonsignificant (p = 0.52).
@@
 Mean floral resource abundance likewise responded significantly (p << 0.0001) to elevation for all bumble bee species **(Figure 2B)**.
-*B.
-jonellus*, *B.
+*Bombus jonellus*, *B.
 mendax*, *B.
@@
 wurflenii* exhibited bimodal patterns of flora resource abundance, with peaks around 1000 and 1600 m.s.l, the latter peak corresponding closely to the location of the tree line.
-*B.
-pyrenaeus* exhibited a unimodal floral resource pattern peaking near 1600 m.s.l.
+*Bombus pyrenaeus* exhibited a unimodal floral resource pattern peaking near 1600 m.s.l.
 The remaining species --- *B.
@@
 Our model explained 67% of total variance (adjusted R^2^ = 0.41).
-Both lack of management (p << 0.0001) and mowing (p << 0.0001) had a significant positive effect on mean flower cover relative to the reference level of grasing management.
+Both lack of management (p << 0.0001) and mowing (p << 0.0001) had a significant positive effect on mean flower cover relative to the reference level of grazing management.
 The effect of year was also significant, with 2010 (p << 0.0001) and 2011 (p << 0.0001) exhibiting higher mean flower cover than 2012.
@@
 Floral β-diversity among our study sites was very high overall and responded steeply to elevation difference between sites **(Figure 3A)**.
-Even sites at similar elevation exhibited ~40% species turnover, and the most widely separated sites (elevation difference > 1250 m) differed by more than 85%.
+Even sites at similar elevation exhibited ~40% species turnover, and the most vertically distant sites (elevation difference > 1250 m) differed by more than 85%.
 Bumble bee β-diversity was, in comparison to floral β-diversity, both lower overall and less responsive to elevational difference between sites **(Figure 3A)**.
-Sites at similar elevation exhibited ~25% species turnover, and species turnover between the most widely separated sites remained less than 50%.
+Sites at similar elevation exhibited ~25% species turnover, and species turnover between the most vertically distant sites remained less than 50%.
 
-Total interaction β-diversity was >75% between sites at similar elevation and approached perfect dissimilarity in the most widely separated sites **(Figure 3B)**.
-Partitioning revealed that total interaction β-diversity was driven primarily by species turnover, which accounted for ~50% of total β-diversity for sites at similar elevation and ~90% of total β-diversity for the most widely separated sites.
-Species turnover consisted mainly of turnover in the floral community **(Figure 3C)**, though the joint turnover of plants and bumble bees accounted for ~25% of total species turnover between the most widely separated sites.
-Interaction rewiring was most significant between sites at similar elevation, where it accounted for ~30% of total β-diversity, but its share of total β-diversity declined to ~10% in the most widely separated sites **(Figure 3B)**.
+Total interaction β-diversity was >75% between sites at similar elevation and approached perfect dissimilarity in the most vertically distant sites **(Figure 3B)**.
+Partitioning revealed that total interaction β-diversity was driven primarily by species turnover, which accounted for ~50% of total β-diversity for sites at similar elevation and ~90% of total β-diversity for the most vertically distant sites.
+Species turnover consisted mainly of turnover in the floral community **(Figure 3C)**, though the joint turnover of plants and bumble bees accounted for ~25% of total species turnover between the most vertically distant sites.
+Interaction rewiring was most significant between sites at similar elevation, where it accounted for ~30% of total β-diversity, but its share of total β-diversity declined to ~10% in the most vertically distant sites **(Figure 3B)**.
 
@@
 The conditional effect of elevation was significant (p ~ 0) in all models, and its importance was 1-2 orders of magnitude higher that of sampling frequency and management **(Figure 4B)**.
-Nevertheless, the conditional effect of sampling frequency was significant (p ~ 0) in the interaction β-diversity model, and mowing management was significant (p ~ 0) in the floral β-diversity model.
+Nevertheless, the conditional effect of sampling frequency was significant (p = 0) in the interaction β-diversity model, and mowing management was significant (p = 0) in the floral β-diversity model.
 
@@
 Nevertheless, it highlights the behavioral challenge that bumble bees have presumably faced for the entirety of their evolutionary history as mountain florivores: how to forage efficiently in a diverse floral community that exhibits rapid spatial and temporal turnover.
-Indeed, the remarkable flexibility and intelligence of bumble bee foraging [@Heinrich1979-ep; @Loukola2017-bt], which have been interpreted as side-effects of their sociality [@Dukas1991-ig], could alternatively be interpreted as adaptations to precisely this problem of foraging in the context of extreme β-diversity.
 A fascinating question that could not be answered by our sampling approach is whether mountain bumble bees adaptively forage up- or downslope, as suggested by @Lundberg1980-qk.
@@
 
-While elevational patterns were the focus of our study, our analyses also revealed a negative conditional effect of grasing management relative to mowing or non-management.
-This is surprising because other studies have linked agricultural abandonment in the Alps (typically a transition from grasing to non-management) with reduced abundance and diversity of both flora and bumble bees [@Iserbyt2015-jc; @Walcher2017-kw].
-We caution, however, against the over-interpretation of the negative effect of grasing found in our study, since our methods did not focus on the characterization of land use or land use history.
+While elevational patterns were the focus of our study, our analyses also revealed a negative conditional effect of grazing management relative to mowing or non-management.
+This is surprising because other studies have linked agricultural abandonment in the Alps (typically a transition from grazing to non-management) with reduced abundance and diversity of both flora and bumble bees [@Iserbyt2015-jc; @Walcher2017-kw].
+We caution, however, against the over-interpretation of the negative effect of grazing found in our study, since our methods did not focus on the characterization of land use or land use history.
 
@@
 
-**Figure 2:** Conditional effects of elevation an dmanagement on the maximum abundance of each bumble bee species (A) and conditional effects of elevation, management, and year on mean flower cover weighted for each bumble bee species (B).
+**Figure 2:** Total bumble bee observations per year (A), total samples per year (B), bumble bee observations per sample per year (C), and the distribution of samples through elevation and time (D).
+Each sample represents a standardized transect walk for a given site and date, each observation represents a recorded interaction between a bumble bee and flower.
+
+**Figure 3:** Conditional effects of elevation and management on the maximum abundance of each bumble bee species (A) and conditional effects of elevation, management, and year on mean flower cover weighted for each bumble bee species (B).
 Effects are plotted on the link scale, and dashed lines depict 95\\% confidence intervals.
@@
 
-**Figure 3:** Species-level (A) and interaction-level (B, C) plotted against elevational difference.
+**Figure 4:** Species-level (A) and interaction-level (B, C) plotted against elevational difference.
 Each point represents the β-diversity (or β-diversity partition) between a pair of sites, and lines represent the overall relationship between β-diversity and elevation difference using binomial regression smooths.
@@
 
-**Figure 4:** GDM splines of species-level and interaction-level β-diversity in response to elevation (A) and tabulated model output showing variable importance and significance for elevation, sampling frequency, and site management.
+**Figure 5:** GDM splines of species-level and interaction-level β-diversity in response to elevation (A) and tabulated model output showing variable importance and significance for elevation, sampling frequency, and site management.
 The maximum height of each spline (A) represents the partial effect of elevation on β-diversity (its effect when covariates are held constant), and the shape of each spline represents to the rate of species turnover as it varies along the elevation gradient, with steeper parts of the curve indicating regions of the gradient over which species/interaction turnover is more rapid.
@@
 ```{r fig2, echo = FALSE}
-knitr::include_graphics("../ms_1/figures/gam_abund_grid_mod.png")
+knitr::include_graphics("../ms_1/figures/samplingplot.png")
 ```
@@
 ```{r fig3, echo = FALSE}
+knitr::include_graphics("../ms_1/figures/gam_abund_grid_mod.png")
+```
+
+\newpage
+
+**Figure 4**
+
+```{r fig4, echo = FALSE}
 knitr::include_graphics("../ms_1/figures/beta_stack.png")
@@
 
-**Figure4**
+**Figure5**
 
-```{r fig4, echo = FALSE}
+```{r fig5, echo = FALSE}
 knitr::include_graphics("../ms_1/figures/gdm_combo.png")

### 2021-09-02 ms_2/ms_2_oikos_part1.Rmd
beginning ms1 revision

--- 2021-08-25 ms_2/ms_2_submission.Rmd
@@
    - \usepackage{setspace}\doublespacing
-output: pdf_document
-bibliography: 
-  - references.bib
-csl: ecologyletters.csl
+output: 
+  pdf_document: 
+    keep_tex: yes
 ---
@@
 
-**ARTICLE TYPE:** Letter
+**ARTICLE TYPE:** Research
 
@@
 
-**KEYWORDS:** foraging, trait matching, pollinator, *Bombus*, alpine, niche, resource selection
-
 **SECTION LENGTHS:**  
-Abstract:  
-Main text:  
-References:  
-Figures:
-
-\newpage
-
-# ABSTRACT
-
-Patterns of resource use by animals can clarify how ecological communities have assembled in the past, how they currently function, and how they are likely to respond to future perturbations.
-Using ~13,000 records of bumble- bee-wildflower interactions along a 1400 m elevation gradient, we find that the selection of floral morphotypes by bumble bees is driven both by the directed process of trait-matching and the neutral effect of floral abundance.
-Patterns of floral preference also vary by elevation and season, which may be an adaptation to the strong spatiotemporal turnover of mountain floral communities.
-The flexibility of bumble bee foraging will be tested in coming years as climate and land use change drive rapid changes in floral resource distribution in time and space, particularly in the mountain ecosystems that host the world's richest communities of bumble bees and their floral mutualists.
-
-\newpage
-
-```{r setup, include=FALSE, cache=FALSE, message = FALSE}
-
-library("knitr")
-
-### Chunk options: see http://yihui.name/knitr/options/ ###
-## Text results
-opts_chunk$set(echo = FALSE, warning = TRUE, message = TRUE)
-
-## Code decoration
-opts_chunk$set(tidy = TRUE, comment = NA, highlight = TRUE)
-
-## Cache
-# opts_chunk$set(cache = 2, cache.path = "ms_2/figures/cache/")
-
-## Plots
-opts_chunk$set(fig.path = "../ms_2/figures/")
-
-```
-
-```{r knitcitations, cache = FALSE}
-library(knitcitations)
-cleanbib()   
-cite_options(citation_format = "pandoc")
-```
-
-# 1. INTRODUCTION
-
-Resource selection and partitioning by animals are instrumental in community assembly, species coexistence, and ecosystem function [@Chesson2000-aw; @Bascompte2007-ls].
-Resource selection is fundamentally constrained by phenotypic compatibility (i.e.
-trait-matching) between resources and consumers, but it can also be influenced by the relative availability of resources [@Johnson1980-co] and modulated by exogenous drivers such as competition [@Pianka1974-qk].
-Understanding these interacting processes is key to anticipating the response of ecological communities to anthropogenic climate and land use change, which threaten to disrupt existing patterns of species interactions and niche relationships [@Tylianakis2008-fi].
-
-\
-
-<!-- Plant-pollinator networks are a special case of resource selection in which consumer fitness and resources fitness are positively correlated, causing the traits of plants and pollinators to converge toward compatibility with one another [e.g.
-@Newman2014-ke].
-The coexistence of multiple plant and pollinator species, however, requires niche differentiation [@Palmer2003-ih; @Johnson2019-su], which can be achieved by trait-based partitioning of interactions among pollinators and their floral hosts [@Willmer2011-ak].
--->
-
-Bumble bees (Hymentoptera: *Bombus* spp.) and their floral hosts provide a diverse yet tractable system in which to explore resource selection in the context of plant-pollinator networks.
-Due to their adaptation to cold temperatures, bumble bees are sensitive to climate change [@Goulson2008-oe; @Soroye2020-zb], especially in the mountain ecosystems that host the world's richest bumble bee communities.
-Warming climates have already caused measurable upslope shifts in mountain bumble bees and their floral hosts [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc], which could introduce novel competitive interactions between bumble bee species that experience increased co-occurrence.
-At the same time, socioeconomic factors have led to widespread abandonment and reforestation of mountain meadows, particularly in the European Alps, which reduces floral resource availability and suppresses bumble bee diversity and abundance [@Tappeiner2006-jp; @Tasser2007-qo; @Walcher2017-kw].
-The ability of bumble bee species to coexist under these conditions of reshuffled distributions and declining floral resources will depend in part on how they select and partition floral resources, which will in turn affect the viability of their floral mutualists, many of which depend principally on bumble bee visitation for pollination [@Goulson2010-lo].
-
-\
-
-Addressing the question of how bumble bees will select floral resources under novel conditions requires discernment of the generalizable mechanisms that underpin patterns of floral selection observed under current conditions.
-It is ordinarily assumed that the choice of resources by foraging animals is a joint function of trait-based preferences and the neutral effects of the relative abundances of available resources: the so-called "use-availability" model of resource selection [e.g.
-@Beyer2010-vr].
-The most well-studied mechanism of the preference component of bumble bees floral selection is trait-matching between bumble bee tongue length and floral corolla depth.
-Bumble bee species can differ markedly in tongue length, and floral handling time increases when tongue length and corolla depth are poorly matched [@Inouye1980-mv; @Harder1983-bk], leading to an association between tongue length and floral preferences in bumble bee communities [e.g.
-@Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn].
-Under the use-availability model, tongue-length-based floral preferences should be modulated by the relative abundance of available resources to generate observed patterns of floral selection.
-It has recently been proposed, however --- in what has been termed the *target use hypothesis* --- that floral selection by bumble bees is a process of *pure preference*, irrespective of resource abundance [@De_Valpine2013-be; @Harmon-Threatt2017-bz].
-If and when the *target use hypothesis* holds, observed visitation rates are a better index of preference than metrics designed to control for resource abundance (e.g.
-null model residuals), since the subtraction of presumed neutral effects of abundance when those effects do not really exist actually creates new abundance-dependent artifacts [@Harmon-Threatt2017-bz].
-On the other hand, failing to account for neutral effects of abundance when they do exist similarly obscures the general mechanisms behind observed foraging patterns [@Vazquez2003-qa; @Bluthgen2008-rp] that will determine bumble bee resource selection under novel conditions induced by climate and land use change.
-
-\
-
-In this study, we use ~13,000 records of bumble bee visits to wildflowers, collected over three years along a 1400 m elevation gradient, to explore three unanswered questions regarding resource selection by bumble bees and its response to climate and land use change.
-First, we ask how bumble bee floral preferences map onto a generalized floral morphotypology [@Kugler1970-gb].
-Then, using the established relationship between tongue length and floral preference as a point of reference, we test the *target use hypothesis* by comparing the explanatory power of bumble bee tongue length with respect to observed visitation rates vs.
-the residuals of a null model that assumes a linear dependency of selection on resource abundance.
-Under the *target use hypothesis*, tongue length should explain observed visitation rates better than null model residuals, while the opposite finding would indicate that, contrary to the *target use hypothesis*, bumble bee floral selection is influenced by relative abundance.
-Finally, we model the variation of bumble bee floral preference through time and elevation as a first step toward understanding how climate and land use change might distort the temporal and elevational structure of bumble- bee-wildflower mutualisms in mountain ecosystems.
-
-# 2. METHODS
-
-## 2.1 Field sampling
-
-Field sampling was conducted 2010-2012 in Berchtesgaden National Park (47.55°N, 12.92°E), located in the Northern Limestone Alps of southeast Germany.
-The landscape is composed of a coniferous forest matrix embedded with pastures either currently or historically maintained by extensive (i.e.
-low intensity) agricultural management.
-We selected a total of 25 study sites (60 x 60 m) located on pastures along a ~1400 m elevation gradient ranging from 641-2032 m above sea level (m.s.l.).
-Fourteen of these pastures are grazed by cattle or sheep, three are mowed for hay production, and eight have been abandoned throughout the last century and are no longer subject to human management.
-
-\
-
-Bumble- bee-wildflower interactions were observed during repeated visits (hereafter, “visitation samples”) to each study site at approximately weekly intervals.
-Sampling was only conducted during periods without rain and with an air temperature of at least 6°C.
-Bumble bees were recorded during one 50-minute transect walk per visit, and bumblebees observed on a given flower were counted as floral visitors.
-Queens were identified to species level in the field to avoid destructively sampling of colony foundresses, but workers and males were collected and identified in the lab following @Amiet1996-jz.
-For analysis, pooled the cryptic *Bombus lucorum/terrestris* species complex and the parasitic *Bombus* (*Psithyrus*) species, respectively, into species-groups.
-Males were recorded, we analyzed only the pooled visitation of queens and workers.
-In conjunction with each bumble bee visitation sample, we estimated the flower cover of herbaceous and shrubby plant species within each 60 x 60 m study plot to the nearest 0.01 m^2^ (hereafter "floral survey samples") and identified the plant species according to @Lauber2007-ha and @Oberndorfer2001-ro.
-See Sponsler et al.
-(submitted) for details of sampling design.
-
-\
-
-## 2.2 Floral morphology and bumble bee tongue length
-
-To summarize floral morphology, we relied on the synthesis of @Kugler1970-gb, who developed a typology consisting of ten primary morphotypes and nested subdivision of each.
-Kugler’s typology was informed by laboratory studies of bumble bee flower choice [@Kugler1943-uy], and there is a precedent for analyzing bumble bee visitation in terms of Kugler’s morphotypes [@Farnberger1996-xo; @Neumayer1999-il; @Schneller2014-ry].
-Kugler morphotype classifications of the plants in our study system were accessed via the BIOFLOR database [@Klotz2002-uc].
-Morphotypes were simplified to the primary classes numbered 0-10: 0 = null (includes wind-pollinated flora such as grasses), 1 = disc- and bowl-shaped flowers, 2 = funnel flowers, 3 = bell-shaped flowers, 4 = stalk-disc flowers, 5 = lip flowers, 6 = flag flowers, 7 = flower heads, 8 = spike flowers, 9 = brush flowers, 10 = trap flowers **(Figure 1)**.
-In a few cases, a plant species in our data set was not found in the BIOFLOR database.
-In most cases, morphotypes are constant at the genus level, and missing morphotypes were added based on the morphotypes found for congeners.
-In the case of the genus *Phyteuma*, we followed @Neumayer1999-il and classified it as a group 7 flower head.
-Morphotypes 9 and 10 were extremely rare in our study system and morphotype 8 was completely absent, so these morphotypes were omitted from all analyses.
-Plants belonging to morphotype 0 were typically not recorded during floral surveys due to their lack of conspicuous floral displays, and they were rarely visited by bumble bees, so this morphotype was also omitted from all analyses.
-
-\
-
-Bumble bee tongue-length data (for the worker caste) were obtained from the literature values compiled and summarized by @Arbetman2017-xj.
-Values for *B.
-gerstaeckeri* and *B.
-mendax*, which were not included in @Arbetman2017-xj, were obtained from @Obeso1992-yp and @Durieux2000-uk, respectively.
-
-## 2.3 Data analysis
-
-For the purposes of our study, we define "preference" as the difference between the observed visitation rate to a given resource and visitation rate predicted by the relative abundance of that resource (preference can thus take negative values for under-visited resources).
-For more nuanced definitions of preference, we refer the reader to @Beyer2010-vr and references therein.
-
-\
-
-To quantify preference, we compared the proportional visitation of each morphotype (by each bumble bee species) with the rates of visitation expected under a null model constructed following @Vaughan2018-gj.
-In this modeling approach, resource selection is simulated for individual consumers --- in our case, individual bumble bees --- according to a resource selection process in which the probability of a resource being chosen is proportional to its relative abundance [@Agusti2003-tt].
-By simulating the visitation of individual consumers rather than species in aggregate, the abundance of each resource, and therefore its probability of visitation, can vary across site-dates, thus preserving the within-sample correspondence between visitation and abundance.
-Simulated interactions are then summarized for each consumer species by taking the mean number of interactions with each resource across all individuals of the consumer species and all samples of the dataset.
-As this process is iterated, the result is a distribution of visitation rates of each consumer species to each resource under the null assumption that resource selection is determined only by resource abundance.
-The visitation distributions of the null model can then be compared to observed visitation rates to infer positive or negative preference of each consumer toward each resource.
-Since the mean of the null distribution can be interpreted as the expected value under null assumptions, the difference between the mean null value and the observed value can be interpreted as a magnitude of preference [@Vaughan2018-gj], and statistical significance can be estimated as the proportion of simulated null values more extreme than the observed value.
-
-\
-
-We followed the procedure described above, simulating our entire dataset in one modeling process to infer aggregate preference between bumble bee species and floral morphotypes.
-Using the results of this analysis, we modeled preference toward each floral morphotype as a function of bumble bee tongue length using hierarchical generalized additive modeling (HGAM) [@Pedersen2019-kf].
-As a test of the *target use hypothesis*, we repeated the HGAM analyses described above using observed visitation rates rather than preference.
-To verify that our inference of preference was not sensitive to our decision to pool queen and worker visitation, we recalculated the null model and tongue length HGAM using only data from workers.
-Finally, we compared morphotype preference across bumble bee species by performing a principal components analysis on the bumble bee species x floral morphotype preference matrix, and we mapped bumble bee tongue length onto the ordination space by fitting a 2-dimensional GAM surface to the first two principal components.
-
-\
-
-To explore temporal and elevational variation in floral preference, we repeated the null model analysis described above but this time for each site-date individually, so that the preference of each bumble bee species toward each floral morphotype could be modeled as a variable through elevation and time.
-For this analysis, we included only the six most abundant and widespread bumble bee species in our study system (*B.
-hortorum*, *B.
-pascuorum*, *B.
-pratorum*, *B.
-soroeensis*, *B.
-terrestris-lucorum*, and *B.
-wurflenii*) to ensure sufficient data points through elevation and time, and we modeled their preference only toward the 4 most visited floral morphotypes (bell, lip, flag, and head flowers).
-Preference for each bumble- bee-morphotype pair (hereafter "bb-fl pair") was modeled together in a single HGAM with an interaction smooth between elevation and time.
-
-\
-
-An inherent problem with modeling sample-wise variation in preference is that the bounds of preference are strictly dependent on floral abundance [@Beyer2010-vr].
-A morphotype present at very low relative abundance would, under the null model, be expected to receive very low visitation, which sets the lower bound of preference near zero (i.e.
-neutrality) but allows the upper bound to approach 1.
-Conversely, a morphotype present at very high relative abundance would receive very high visitation under the null model, and its upper bound of preference would fall near zero and its lower bound near -1.
-This clearly creates a degree of unavoidable confoundedness between preference and floral abundance.
-With this limitation in mind, we fit HGAMs of the same form to the relative abundance of each floral morphotype, and we present the results together to be interpreted jointly.
-
-\
-
-All data processing and analyses were conducted in `R` [@RCoreTeam-2021aa].
-Data handling and visualization were performed with the `tidyverse` suite [@Wickham2019-aa].
-Null models were constructed using `econullnetr` [@Vaughan2018-gj].
-GAM analyses were performed with `mgcv` [@Wood_2017] and visualized with `mgcViz` [@Fasiolo2018aa] and `tidymv` [@Coretta2021-aa].
-PCA was performed with the `vegan` [@Oksanen_2019] and visualized using `ggplot2` [@Wickham2016-aa] and `ggvegan` [@Simpson2019aa].
-Annotated R code is available in the Appendix X.
-
-# 3. RESULTS
-
-## 3.1 Bumble bee and floral communities
-
-A total of 13,293 bumble- bee-flower interactions were included in our final data set, comprising 15 bumble bee species, 163 plant species (110 genera, 37 families), and 730 unique bumble- bee-plant interaction pairs.
-*B.
-pascuorum* ("pasc"), *B.
-pratorum* ("prat"), *B.
-soroeensis* ("soro"), *B.
-terrestris-lucorum* ("telu"), and *B.
-wurflenii* ("wurf") were the most abundant species, and they occurred at all sites.
-*B.
-hortorum* ("hort"), *B.
-jonellus* ("jone"),  *B.
-monticola* ("mont"), and *B.
-psithyrus* ("psit") occurred at moderate abundance.
-The remaining species --- *B.
-gerstaeckeri* ("gers"), *B.
-hypnorum* ("hypn"), *B.
-lapidarius* ("lapi"), *B.
-mendax* ("mend"), *B.
-mucidus* ("muci"), and *B.
-pyrenaeus* ("pyre") --- were relatively rare.
-*B.
-humilis* ("humi") was recorded only three times and only in one year, so we omitted it from all analyses.
-Floral surveying yielded a total of 354 plant species (194 genera, 53 families).
-Visited floral species included all morphotypes except spike (8) and trap (10) flowers.
-Null (0) and brush (9) morphotypes were, however, excluded from analysis for the reasons given above.
-
-## 3.2 Bumble bee preferences toward floral morphotypes
-
-Preference was generally strong, with the visitation rates of most bumblebee species to most floral morphotypes falling significantly above or below the 99% interval of the null model **(Figure 1)**.
-Nonsignificant preference was observed mainly among the rarer bumblebee species, likely because their small samples sizes lead to wide null model confidence intervals.
-Interestingly, bumble bees showed strong species-specific variation in their relative preferences of different floral morphotypes **(Figure 1)**.
-The omission of queens from the calculation of overall preference did not yield qualitatively different results (Appendix X).
-
-\
-
-When floral morphotypes are compared to one another in terms of preference across all bumble bee species, clear patterns emerge **(Figure 2A)**.
-All species exhibited negative preference toward disc flowers, while most species exhibited positive preference toward bell, lip, and flag flowers.
-preference toward funnel and stalk-disc flowers tended to be neutral (with one strong positive outlier --- *B.
-gerstaeckeri* --- for funnel flowers), reflecting the fact that these morphotypes tended both to be rarely visited and to occur at low abundance.
-Preference toward head flowers varied widely across bumble bee species, ranging from strongly positive to strongly negative.
-
-\
-
-Differences between floral morphotypes in terms of observed visitation rates, reflecting the target use hypothesis, were less pronounced **(Figure 2B)**.
-Bell, lip, flag, and head flowers were visited at high rates by one or more bumble bee species, but observed visitation rates did not differ strongly across morphotypes.
-
-<!-- I could do some sort of linear mixed effects model on this I suppose, but I think it would violate assumptions due to the non-independence of preference (preference for one thing implies preference against others).
-I've already sort of done stats on this with the null model.
--->
-
-## 3.3 Tongue length as a predictor of floral choice
-
-Tongue length was a strong predictor of preference but not of observed visitation rates **(Figure 3)**.
-Preference toward funnel, bell, lip, flag, and head flowers exhibited significant GAM curves in response to bumble bee tongue length, and the model explained 77% of total deviance (adj.
-R^2^ = 0.728).
-For funnel and lip flowers, preference increased linearly with tongue length.
-Preference toward bell flowers and head flowers exhibited approximately quadratic responses with preference peaking among short-tongued species and declining sharply with increasing tongue length before leveling off.
-Preference toward flag flowers exhibited a unimodal peak among medium-tongued species.
-In contrast, observed visitation rates were predicted by tongue length only for head flowers, which exhibited a weak decline in preference from a peak among short-tongued species.
-The visitation rate model explained only 25% of total deviance (adj.
-R^2^ = 0.053).
-Again, these results did not differ when preference was based on the worker-only null model (Appendix X).
-
-\
-
-PCA of bumble bee floral preferences captured more than 80% of total variance in the first two axes **(Figure 4)**, and the GAM surface fit to the firs two axes explained most of the variance in tongue length (deviance explained = 94.7%, adj.
-R^2^ = 0.893, p < 0.001), reflecting the importance of trait matching between tongue length and floral morphotype as a mechanistic basis of preference.
-Longer tongue length was associated with deeper and more closed flower morphotypes (lip, funnel, flag), while shorter tongue length was associated with more shallow and open flower morphologies (disc, stalk-disc, bell, head).
-The strong affinity of *B.
-soroeensis* and *B.
-psithyrus* for head flowers is especially salient.
-
-## 3.4 Variation of morphotype abundance and preference through elevation and season
-
-The proportional abundance of the most visited floral morphotypes --- bell, lip, flag, and head flowers --- was strongly patterned by elevation and season **(Figure 5)**.
-Bell flowers peaked in proportional abundance early in the season and at mid-elevation, and lip flowers peaked late in the season and at low elevation.
-Flag flowers peaked in mid-season and exhibited two distinct elevational peaks, the lower at around 1000 m.s.l and the higher at around 1800 m.s.l.
-Head flowers exhibited a mid-season peak at low elevation and a broad late season peak at around 1600 m.s.l.
-All smooths were significant (p << 0.0001), and the model explained 64.3% of total deviance (adj.
-R^2^ = 0.554).
-
-\
-
-The morphotype preferences of bumble bee species were also strongly patterned through elevation and season.
-In some cases, though, patterns of preference mirrored patterns of floral availability, suggesting that bumble bees may exhibit functional responses [@Mysterud1998-rs] to floral abundance not entirely accounted for by null models that assume a linear proportionality between abundance and visitation.
-Preference for flag flowers was strongly associated with high elevation, except among the species that tended to avoid flag flowers (*B.
-pratorum* and *B.
-soroeensis*) **(Figure 1)**.
-It is also apparent that, as a rule, preference is more dynamic where it is also higher overall.
-Bell flowers, which were positively selected overall by the short-tongued bumble bee species *B.
-pratorum*, "*B.
-terrestris-lucorum*, and *B.
-soroeensis* **(Figure 1)**, also exhibited the most complex patterns of preference in elevation and time for these bumble bee species **(Figure 6)**.
-In contrast, the long and medium-tongued species *B.
-hortorum*, *B.
-pascuorum*, and *B.
-wurflenii*, which exhibited neutral or weakly positive overall selection toward bell flowers **(Figure 1)**, also show weak patterns of variation in preference toward bell flowers **(Figure 6)**.
-Nearly the reverse pattern is evident for flag flowers, reflecting their association with longer-tongued bumble bee species.
-
-\
-
-Smooths were significant (p < 0.05) for all bee-morphotype pairs except hort-head, prat-flag, telu-head, and soro-flag. The model explained 36.4% of total deviance (adj. R^2^ = 0.33).
-
-# 4. DISCUSSION
-
-Using ~13,000 bumble- bee-wildflower interactions, we characterized bumble bee floral selection in terms of generalized floral morphotypes and found evidence that bumble bee floral selection depends both on the directed process of morphotype preference and the neutral effect of floral abundance.
-Bumble bee species differed in their floral morphotype preferences, and these differences could be explained by trait-matching between floral morphotype and bumble bee tongue length.
-These preferences, however, varied by elevation and season, likely representing adaptive responses of bumble bee foraging behavior to elevational and seasonal shifts in floral resource composition, possibly modulated by competitive interactions among bumble bees species with overlapping resource preferences.
-
-## 4.1 Floral abundance, floral selection, and the *target use hypothesis*
-
-Our finding that bumble bee tongue length explained preference (i.e.
-null model residuals) much better than it explained observed visitation rates indicates that, in our study system, abundance and preference function jointly in bumble bee floral selection.
-This casts doubt on the *target use hypothesis* as a general rule of bumble bee foraging, though it is possible that it still obtains under certain circumstances [e.g.
-@Harmon-Threatt2017-bz].
-In general, though, our findings support the hypothesis that bumble bee resource selection involves both the neutral effects of floral abundance and the directed effect of preference, meaning that analytical techniques designed to disentangle these effects, such as the null model method we employ, should be favored in studies of bumble bee foraging.
-
-## 4.2 Community-level patterns of morphotype preferences
-
-We found that the most preferred morphotypes were bell, lip, and flag flowers, which is broadly consistent with visitation patterns reported by previous studies in central Europe [@Schneller2014-ry; @Neumayer1999-il; Farnberger1996-xo], indicating that our findings are regionally generalizeable.
-These previous studies tended to find a stronger role of head flowers than indicated by our preference analysis, but this is probably because they characterized foraging in terms of simple visitation rates rather than null-model-inferred preference.
-We saw similarly high visitation rates toward head flowers, but our null-model analysis reveals that while head flowers were highly visited, they tended to be under-visited relative to their abundance.
-
-\
-
-While bell, lip, and flag flowers were the most preferred morphotypes on average, preference toward these morphotypes varied broadly across bumble bee species.
-Moreover, while average preference toward funnel and head flowers was low, each was strongly preferred by one or two bumble bee species.
-The majority of variation in preference toward floral morphotypes was explainable by bumble bee tongue length, with the more open bell and head morphotypes selected more positively by short-tongued bumble bee species and the deeper funnel, lip, and flag morphotypes selected more positively by medium- and long-tongued bumble bee species.
-The failure of tongue length to explain preference toward disc and stalk-disc flowers reflects the fact that these flowers were almost unanimously ignored by all bumble bee species.
-We cannot say from our data why disc and stalk-disc flowers received such consistently low visitation, but the pattern may be driven in part by the fact that some disc flowers offer pollen but little or no nectar; indeed, Kugler's (1970) typology recognizes this distinction with subdivisions of the disc morphotype into those with (class 1.2) and without (class 1.1) nectar rewards.
-A limitation of our study is that we could not reliably distinguish pollen- and nectar-foraging bumble bees, and it is likely that this distinction would shed further light on the mechanisms underlying the selection patterns we observed.
-
-\
-
-The strong explanatory power of tongue length with respect to preference in our study corroborates the established link between tongue length and floral choice [e.g.
-@Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn].
-It is important, though, to consider the species that appear as outliers with respect to the smooth dependency of preference on tongue length.
-*B.
-terrestris/lucorum* showed comparatively high (though still negative) preference toward disc flowers.
-*B.
-mucidus* exhibited unusually low preference toward lip flowers, complemented by unusually high preference toward flag flowers.
-*B.
-soroeensis* and the *B.
-psythyrus* species-group were anomalous in their strongly positive selection toward head flowers.
-Such unexplained variation in preference may indicate that traits not captured by Kugler's (1970) morphotypology are also important in governing floral selection by bumble bees.
-FLoral color, in particular, is known to be instrumental in the process of flower detection by foraging bumble bees [@Spaethe2001-bq], and some bumble bee species exhibit innate color preferences [@Raine2007-nq].
-This effect can also be seen in field studies, where bumble bees have been shown to over- or under-represent certain floral colors [e.g.
-@Teras1976-pi; @Neumayer1999-il].
-Moreover, since floral color can be correlated with floral morphology [@Teras1976-pi], it is possible that some of the preference attributed to morphology in our models could represent a latent effect of floral color.
-
-\
-
-Species of similar tongue length will, on average, be more likely to compete for resources than species of dissimilar tongue length.
-Our ordination analysis, though, suggests that there may be significant patterns of similarity and divergence even among species of similar tongue length.
-Among short-tongued species, *B.
-jonellus* and *B.
-pyrenaeus* cluster tightly around bell flowers, while *B.
-psithryus* and *B.
-soroeensis* cluster tightly around head flowers.
-Medium- and long-tongued species differ in their allocation of preference across funnel, lip, and flag flowers.
-*B.
-gerstaeckeri*, for example, is associated with lip and funnel flowers but eschews flag flowers, while *B.
-mucidus* does this opposite.
-With respect to the question of climate-induced competition, the case of *B.
-gerstaeckeri* deserves special consideration.
-A red-listed vulnerable species, *B.
-gerstaeckeri* is a pollen specialist on the lip flowers of the genus *Aconitum* [@Ponchau2006-af; @Schneller2014-ry], and there is evidence that joint elevational shifts of bumble bees and floral resources over the past century have generated more intense exploitation of *Aconitum*, which may indicate nascent competitive conditions for *B.
-gerstaeckeri* [@Marshall2020-pc].
-
-\
-
-When considering community-level patterns of preference and visitation from the perspective of pollination services to mountain flora, it is important to note that while bumble bees are the dominant Hymenopteran pollinators in our study system, flies (Diptera) are also abundant flower visitors in mountain ecosystems, and their visitation is mainly restricted to the open flower morphotypes (e.g.
-disc flowers) that receive less attention from bumble bees [@Neumayer1999-il].
-In the context of our research question, we have emphasized the differences in tongue length between bumble bee species, but all bumble bees have relatively long tongues compared to the overall pollinator community, and it is not surprising that their foraging tends to be focused on the deeper and more closed flower morphotypes.
-The process of floral resource partitioning *among* bumble bee species is nested within the process of floral partitioning *between* bumble bees and other flower visitors.
-
-## 4.3 Variability of floral abundance, visitation, and preference through elevation and time
-
-Intraspecific variation of bumble bee foraging preferences has been demonstrated under laboratory conditions [@Ings2009-vc; @Maharaj2019-wm]; our study reveals the *in situ* variation of preference for multiple bumble bee species through elevational and temporal gradients of floristic turnover.
-
-\
-
-At least three mechanisms could underlie the variation of preferences we observed.
-First, variation in preference toward a given morphotype could be driven by changes in the *intrinsic* reward of that focal morphotype, which could be caused by (1) shifts in species composition of the focal morphotype (e.g.
-the occurrence of a particularly rewarding flag flower at high-elevation sites), (2) shifts in the reward levels of individual species (e.g.
-increased nectar production in the flag flower *Lotus corniculatus* under high elevation conditions), or (3) changes in nutritional requirements of bumble bees (e.g.
-an increased need for concentrated nectar to support thermoregulation at colder ambient temperatures).
-Alternatively, preference toward a given morphotype could vary due to changes in the *relative* reward of the focal morphotype with respect to changing background resource conditions.
-For example, it could be that flag flowers are selected more positively at high elevation because they must compete for visitation with highly attractive lip flowers at lower elevation.
-Finally, it is possible that patterns of resource selection could vary due to competitive interactions within or between bumble bee species.
-For example, the increased selection of flag flowers at high elevation could be driven by competition release from *B.
-pascuorum*, which tends to be the most abundance bumble bee species at lower elevations but becomes relatively scarce above 1500 m (Sponsler et al, submitted).
-Under this hypothesis, an organism's resource selection is the product of exogenous constraint rather than endogenous preference.
-
-## 4.4 Conclusions
-
-Clarifying the patterns and processes of bumble floral selection sheds light on how bumble bee communities have evolved with respect to their mountain habitats and how they might respond to the incipient forcings of climate and land use change.
-The dynamics of preference in our study suggest adaptive tuning of floral choice consistent with the ecological hurdles that have shaped the evolution of bumble bees in the mountain habitats in which they arose and diversified [@Williams1985-ql; @Hines2008-er].
-Where steep temperature gradients and short growing seasons generate extreme spatiotemporal turnover of floral resources, adaptive flexibility of morphological preferences should be favored over fixed biases.
-As climate and land use change drive the redistribution of bumble bees and their floral mutualists, and likely an overall reduction in floral availability in mountain habitats, competition within bumble bee communities is likely to intensify.
-Whether this leads to competitive exclusion will depend on whether interspecific competition exceeds the stabilizing effects of interspecific competition.
-In this regard, the partial dependency of bumble bee floral selection on the neutral effects of floral abundance (in contrast to the *target use hypothesis*) can be expected to favor the coexistence of both bumble bee species and their floral host by increasing the likelihood that species will be self-limited by intraspecific competition [@Valdovinos2021-ya].
-
-# ACKNOWLEDGMENTS
-
-We thank the National Park Berchtesgaden for their support and the permission to work in the national park, and we thank the owners of the study sites outside the national park for access to their land.
-The field work of this study was performed within the framework of the joint research center FORKAST as part of the ‘Bavarian Climate Programme 2020'.
-D.J.
-McNeil provided for helpful conversations during data analysis.
-O.
-Mitesser kindly provided a high-performance computing environment to expedite null modeling.
-
-# REFERENCES
-
-<!-- ```{r write_citations, cache=FALSE, include=FALSE} -->
-<!-- write.bibtex(file = "knitcitations.bib") -->
-<!-- ``` -->
-
-<div id ="refs"></div>
-
-# FIGURES
-**Figure 1**  
-```{r fig1, echo = FALSE}
-knitr::include_graphics("../ms_2/figures/fig1.pdf")
-```
-
-\newpage
-
-**Figure 2**  
-```{r fig2, echo = FALSE, out.width="100%"}
-knitr::include_graphics("../ms_2/figures/fig2.pdf")
-```
-
-\newpage
-
-**Figure 3**  
-```{r fig3, echo = FALSE, out.width="100%"}
-knitr::include_graphics("../ms_2/figures/fig3.pdf")
-```
-
-\newpage
-
-**Figure 4**  
-```{r fig4, echo = FALSE}
-knitr::include_graphics("../ms_2/figures/fig4.pdf")
-```
-
-\newpage
-
-**Figure 5**  
-```{r fig5, echo = FALSE}
-knitr::include_graphics("../ms_2/figures/fig5.pdf")
-```
-
-\newpage
-
-**Figure 6**  
-```{r fig6, echo = FALSE}
-knitr::include_graphics("../ms_2/figures/fig6.pdf")
-```
-
-\newpage
-
-# FIGURE CAPTIONS
-
-**Figure 1:** Observed visitation rates compared to null model simulations for each bumble bee species and floral morphotype (A).
-Gray bars depict 99% intervals of null model simulations and points depict observed visitation rates.
-Points are color-coded according to where they fall with respect to the 99% interval of the null model: red = above (positive preference), blue = below (negative preference), gray = within (nonsignificant preference).
-The difference between the observed value and the center of the null model interval can be interpreted as magnitude of preference.
-Each floral morphotype is illustrated (B) with a representative species common in our study system: 1 = *Potentilla erecta*, 2 = *Gentiana asclepiadea*, 3 = *Campanula glomerata*, 4 = *Silene nutans*, 5 = *Ajuga reptans*, 6 = *Lotus corniculatus*, 7 = *Centaurea jacea*.
-Illustrations are modifications of public domain works obtained from www.plantillustrations.org.
-For details of original sources, see Appendix X.
-
-\
-
-**Figure 2:** Preference (A) and observed visitation rate (B) of the bumble bee species toward each floral morphotype.
-Each point represents the preference/visitation of one bumble bee species toward the corresponding floral morphotype.
-The color of each point in panel A, retained from Figure 2, indicates whether a given observed bias is significantly (p < 0.01) positive or negative with respect to null model simulations.
-
-\
-
-**Figure 3:** Preference (A) and observed visitation rates (B) for each morphotype regressed against bumble bee tongue length.
-Each point represents the preference/visitation of one bumble bee species toward the corresponding floral morphotype.
-Curves depict fitted GAM smooths with 95\\% confidence bands.
-Solid lines indicate curves differing significantly (p < 0.05) from a flat line while dashed lines indicated nonsigificant curves.
-Points are color-coded by species, and species are presented in the legend in order of increasing tongue length.
-
-\
-
-**Figure 4:** Principal components biplot of bumble bee species according to their preference toward each floral morphotype.
-Floral morphotypes are abbreviated with their numerical codes and illustrated as in Figure 1.
-Colored bands depict a 2-dimensional GAM smooth representing the relationship between the first two principal components and bumble bee tongue length.
-Morphotype illustrations are as described in Figure 1.
-
-\
-
-**Figure 5:** Variation in proportional flower cover through elevation and season modeled as HGAM surfaces.
-Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns.
-All smooths are significant (p < 0.05).
-
-\
-
-**Figure 6:** Variation in preference through elevation and time modeled as HGAM surfaces.
-Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns.
-Significant (p < 0.05) smooths are depicted with solid contour lines and nonsignificant smooths with dashed contour lines.
+Abstract: 141 words
+Main text: 4979 words
+References: 61  
+Figures: 6

### 2021-09-02 ms_2/ms_2_oikos_part2.Rmd
beginning ms1 revision

--- 2021-08-25 ms_2/ms_2_submission.Rmd
@@
    - \usepackage{setspace}\doublespacing
-output: pdf_document
+output: 
+  pdf_document: 
+    keep_tex: yes
 bibliography: 
   - references.bib
-csl: ecologyletters.csl
+csl: oikos.csl
 ---
-
-**Douglas Sponsler** (corresponding author)  
-University of Würzburg  
-Department of Animal Ecology and Tropical Biology  
-Biozentrum der Universität Würzburg C-018  
-Am Hubland  
-97074 Würzburg, Germany  
-Email: douglas.sponsler@uni-wuerzburg.de  
-Telephone: +49 0176 363 20273
-
-**Katharina Kallnik**  
-University of Würzburg  
-Department of Animal Ecology and Tropical Biology  
-97074 Würzburg, Germany  
-Email: katharina.kallnik@uni-wuerzburg.de
-
-**Fabrice Requier**  
-Université Paris-Saclay  
-Évolution, Génomes, Comportement et Écologie  
-91198 Gif-sur-Yvette, France  
-Email: fabrice.requier@egce.cnrs-gif.fr
-
-**Alice Classen**  
-University of Würzburg  
-Department of Animal Ecology and Tropical Biology  
-97074 Würzburg, Germany  
-Email: alice.classen@uni-wuerzburg.de
-
-**A. Fabienne Maihoff**  
-University of Würzburg  
-Department of Animal Ecology and Tropical Biology  
-97074 Würzburg, Germany  
-Email: anne_fabienne.maihoff@uni-wuerzburg.de
-
-**Johanna Sieger**  
-Naturpark Frankenhöhe  
-91522 Ansbach, Germany   
-Email: sieger@lpv-mfr.de
-
-**Ingolf Steffan-Dewenter**  
-University of Würzburg  
-Department of Animal Ecology and Tropical Biology  
-97074 Würzburg, Germany  
-Email: ingolf.steffan@uni-wuerzburg.de
-
-\
-
-**AUTHORSHIP STATEMENT:** KK and ISD conceived the study.
-KK, JS, and ISD designed and carried out field work.
-DS analyzed the data and wrote the paper with substantial contributions from FR, AC, AFM, KK, and ISD.
-
-**DATA ACCESSIBILITY:** All data used in this study have been archived on DRYAD (https://datadryad.org/stash/dataset/doi:10.5061/dryad.s1rn8pk7f) and will be available upon article publication.
-
-**ARTICLE TYPE:** Letter
-
-**RUNNING TITLE:** Bumble bee floral preferences
-
-**KEYWORDS:** foraging, trait matching, pollinator, *Bombus*, alpine, niche, resource selection
-
-**SECTION LENGTHS:**  
-Abstract:  
-Main text:  
-References:  
-Figures:
-
-\newpage
 
@@
 Patterns of floral preference also vary by elevation and season, which may be an adaptation to the strong spatiotemporal turnover of mountain floral communities.
-The flexibility of bumble bee foraging will be tested in coming years as climate and land use change drive rapid changes in floral resource distribution in time and space, particularly in the mountain ecosystems that host the world's richest communities of bumble bees and their floral mutualists.
+The flexibility of bumble bee foraging will be tested in coming years as climate and land use change alter floral resource distribution in time and space, particularly in the mountain ecosystems that host the world's richest communities of bumble bees and their floral mutualists.
+
+**KEYWORDS:** foraging, trait matching, pollinator, *Bombus*, niche, resource selection
 
@@
 
-<!-- Plant-pollinator networks are a special case of resource selection in which consumer fitness and resources fitness are positively correlated, causing the traits of plants and pollinators to converge toward compatibility with one another [e.g.
-@Newman2014-ke].
-The coexistence of multiple plant and pollinator species, however, requires niche differentiation [@Palmer2003-ih; @Johnson2019-su], which can be achieved by trait-based partitioning of interactions among pollinators and their floral hosts [@Willmer2011-ak].
--->
-
 Bumble bees (Hymentoptera: *Bombus* spp.) and their floral hosts provide a diverse yet tractable system in which to explore resource selection in the context of plant-pollinator networks.
@@
 @Beyer2010-vr].
-The most well-studied mechanism of the preference component of bumble bees floral selection is trait-matching between bumble bee tongue length and floral corolla depth.
-Bumble bee species can differ markedly in tongue length, and floral handling time increases when tongue length and corolla depth are poorly matched [@Inouye1980-mv; @Harder1983-bk], leading to an association between tongue length and floral preferences in bumble bee communities [e.g.
+The most well-studied mechanism of the preference component of bumble bee floral selection is trait-matching between bumble bee tongue length and floral corolla depth.
+Bumble bee species can differ markedly in tongue length, and floral handling time increases when tongue length and corolla depth are poorly matched [@Inouye1980-mv; @Harder1983-bk], leading to an association between tongue length and floral choice in bumble bee communities [e.g.
 @Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn].
@@
 
+\
+
 # 2. METHODS
@@
 We selected a total of 25 study sites (60 x 60 m) located on pastures along a ~1400 m elevation gradient ranging from 641-2032 m above sea level (m.s.l.).
-Fourteen of these pastures are grazed by cattle or sheep, three are mowed for hay production, and eight have been abandoned throughout the last century and are no longer subject to human management.
+Fourteen of these pastures are grazed by cattle or sheep, three are mowed for hay production, and eight have been abandoned during the last century and are no longer subject to human management.
 
@@
 
-Bumble- bee-wildflower interactions were observed during repeated visits (hereafter, “visitation samples”) to each study site at approximately weekly intervals.
+Bumble- bee-wildflower interactions were observed during repeated visits to each study site at approximately weekly intervals.
 Sampling was only conducted during periods without rain and with an air temperature of at least 6°C.
 Bumble bees were recorded during one 50-minute transect walk per visit, and bumblebees observed on a given flower were counted as floral visitors.
-Queens were identified to species level in the field to avoid destructively sampling of colony foundresses, but workers and males were collected and identified in the lab following @Amiet1996-jz.
-For analysis, pooled the cryptic *Bombus lucorum/terrestris* species complex and the parasitic *Bombus* (*Psithyrus*) species, respectively, into species-groups.
-Males were recorded, we analyzed only the pooled visitation of queens and workers.
-In conjunction with each bumble bee visitation sample, we estimated the flower cover of herbaceous and shrubby plant species within each 60 x 60 m study plot to the nearest 0.01 m^2^ (hereafter "floral survey samples") and identified the plant species according to @Lauber2007-ha and @Oberndorfer2001-ro.
+Queens were identified to species level in the field to avoid destructively sampling colony foundresses, but workers and males were collected and identified in the lab following @Amiet1996-jz.
+Prior to analysis, we pooled the cryptic *Bombus lucorum/terrestris* species complex and the parasitic *Bombus* (*Psithyrus*) species, respectively, into species-groups.
+Males were recorded, but we analyzed only the pooled visitation of queens and workers.
+In conjunction with each bumble bee visitation sample, we estimated the flower cover of herbaceous and shrubby plant species within each 60 x 60 m study plot to the nearest 0.01 m^2^ and identified the plant species according to @Lauber2007-ha and @Oberndorfer2001-ro.
 See Sponsler et al.
@@
 
-To summarize floral morphology, we relied on the synthesis of @Kugler1970-gb, who developed a typology consisting of ten primary morphotypes and nested subdivision of each.
+To summarize floral morphology, we relied on the synthesis of @Kugler1970-gb, who developed a typology consisting of ten primary morphotypes and nested subdivisions of each.
 Kugler’s typology was informed by laboratory studies of bumble bee flower choice [@Kugler1943-uy], and there is a precedent for analyzing bumble bee visitation in terms of Kugler’s morphotypes [@Farnberger1996-xo; @Neumayer1999-il; @Schneller2014-ry].
@@
 
+\
+
 ## 2.3 Data analysis
 
-For the purposes of our study, we define "preference" as the difference between the observed visitation rate to a given resource and visitation rate predicted by the relative abundance of that resource (preference can thus take negative values for under-visited resources).
+For the purposes of our study, we define "preference" as the difference between the observed visitation rate to a given resource and visitation rate predicted by a null model in which the probability of a resource being chosen is proportional to its relative abundance [@Agusti2003-tt].
+In other words, preferences are null model residuals, and thus take negative values for under-visited resources and positive values for over-visited resources.
 For more nuanced definitions of preference, we refer the reader to @Beyer2010-vr and references therein.
@@
 
-To quantify preference, we compared the proportional visitation of each morphotype (by each bumble bee species) with the rates of visitation expected under a null model constructed following @Vaughan2018-gj.
-In this modeling approach, resource selection is simulated for individual consumers --- in our case, individual bumble bees --- according to a resource selection process in which the probability of a resource being chosen is proportional to its relative abundance [@Agusti2003-tt].
-By simulating the visitation of individual consumers rather than species in aggregate, the abundance of each resource, and therefore its probability of visitation, can vary across site-dates, thus preserving the within-sample correspondence between visitation and abundance.
+We constructed null models following @Vaughan2018-gj.
+In this modeling approach, resource selection is simulated for individual consumers --- in our case, individual bumble bees.
+By simulating the visitation of individual consumers rather than species in aggregate, the abundance of each resource, and therefore its probability of visitation, can vary across samples (i.e.
+site-dates), thus preserving the within-sample correspondence between visitation and abundance.
 Simulated interactions are then summarized for each consumer species by taking the mean number of interactions with each resource across all individuals of the consumer species and all samples of the dataset.
@@
 
-We followed the procedure described above, simulating our entire dataset in one modeling process to infer aggregate preference between bumble bee species and floral morphotypes.
-Using the results of this analysis, we modeled preference toward each floral morphotype as a function of bumble bee tongue length using hierarchical generalized additive modeling (HGAM) [@Pedersen2019-kf].
+We followed the procedure described above, simulating our entire dataset in one modeling process to infer aggregate preferences of bumble bee species for each floral morphotype.
+Using the results of this analysis, we modeled preferences for each floral morphotype as a function of bumble bee tongue length with hierarchical generalized additive modeling (HGAM), following the *type I* effect structure described by @Pedersen2019-kf.
 As a test of the *target use hypothesis*, we repeated the HGAM analyses described above using observed visitation rates rather than preference.
 To verify that our inference of preference was not sensitive to our decision to pool queen and worker visitation, we recalculated the null model and tongue length HGAM using only data from workers.
-Finally, we compared morphotype preference across bumble bee species by performing a principal components analysis on the bumble bee species x floral morphotype preference matrix, and we mapped bumble bee tongue length onto the ordination space by fitting a 2-dimensional GAM surface to the first two principal components.
+Finally, we compared morphotype preferences across bumble bee species by performing a principal components analysis on the bumble bee species x floral morphotype preference matrix, and we mapped bumble bee tongue length onto the ordination space by fitting a 2-dimensional GAM surface to the first two principal components.
 
@@
 wurflenii*) to ensure sufficient data points through elevation and time, and we modeled their preference only toward the 4 most visited floral morphotypes (bell, lip, flag, and head flowers).
-Preference for each bumble- bee-morphotype pair (hereafter "bb-fl pair") was modeled together in a single HGAM with an interaction smooth between elevation and time.
+Preference for each bumble- bee-morphotype pair was modeled together in a single HGAM with an interaction smooth between elevation and time.
+We also included bee-morphotype pair as a fixed itnercept effect and the interactions between bee-morphotype pair and site and bee-morphotype pair and year as random effects.
 
@@
 Conversely, a morphotype present at very high relative abundance would receive very high visitation under the null model, and its upper bound of preference would fall near zero and its lower bound near -1.
-This clearly creates a degree of unavoidable confoundedness between preference and floral abundance.
+This creates some confoundedness between preference and floral abundance.
 With this limitation in mind, we fit HGAMs of the same form to the relative abundance of each floral morphotype, and we present the results together to be interpreted jointly.
@@
 PCA was performed with the `vegan` [@Oksanen_2019] and visualized using `ggplot2` [@Wickham2016-aa] and `ggvegan` [@Simpson2019aa].
-Annotated R code is available in the Appendix X.
+Annotated R code and complete output of our analytical workflow is available in Appendix 1.
+
+\
 
@@
 
+\
+
 ## 3.2 Bumble bee preferences toward floral morphotypes
 
-Preference was generally strong, with the visitation rates of most bumblebee species to most floral morphotypes falling significantly above or below the 99% interval of the null model **(Figure 1)**.
+Preferences were generally strong, with the visitation rates of most bumblebee species to most floral morphotypes falling significantly above or below the 99% interval of the null model **(Figure 1)**.
 Nonsignificant preference was observed mainly among the rarer bumblebee species, likely because their small samples sizes lead to wide null model confidence intervals.
-Interestingly, bumble bees showed strong species-specific variation in their relative preferences of different floral morphotypes **(Figure 1)**.
-The omission of queens from the calculation of overall preference did not yield qualitatively different results (Appendix X).
+Notably, bumble bees showed strong species-specific variation in their relative preferences of different floral morphotypes **(Figure 1)**.
+The omission of queens from the calculation of overall preference did not yield qualitatively different results (Appendix 1).
 
@@
 All species exhibited negative preference toward disc flowers, while most species exhibited positive preference toward bell, lip, and flag flowers.
-preference toward funnel and stalk-disc flowers tended to be neutral (with one strong positive outlier --- *B.
-gerstaeckeri* --- for funnel flowers), reflecting the fact that these morphotypes tended both to be rarely visited and to occur at low abundance.
-Preference toward head flowers varied widely across bumble bee species, ranging from strongly positive to strongly negative.
+Preference toward funnel and stalk-disc flowers tended to be neutral (with one strong positive outlier --- *B.
+gerstaeckeri* --- for funnel flowers), reflecting the fact that these morphotypes tended to be rarely visited while also occurring at low abundance.
+Preference toward head flowers varied widely across bumble bee species, ranging from strongly positive (*B.
+psithyrus*, *B.
+soroeensis*) to strongly negative (e.g.
+*B.
+gerstaeckeri*, *B.
+mendax*).
 
@@
 
-<!-- I could do some sort of linear mixed effects model on this I suppose, but I think it would violate assumptions due to the non-independence of preference (preference for one thing implies preference against others).
-I've already sort of done stats on this with the null model.
--->
+\
 
@@
 R^2^ = 0.053).
-Again, these results did not differ when preference was based on the worker-only null model (Appendix X).
+Again, these results did not differ when preference was based on the worker-only null model (Appendix 1).
 
@@
 
+\
+
 ## 3.4 Variation of morphotype abundance and preference through elevation and season
 
-The proportional abundance of the most visited floral morphotypes --- bell, lip, flag, and head flowers --- was strongly patterned by elevation and season **(Figure 5)**.
-Bell flowers peaked in proportional abundance early in the season and at mid-elevation, and lip flowers peaked late in the season and at low elevation.
+The relative abundance of the most visited floral morphotypes --- bell, lip, flag, and head flowers --- was strongly patterned by elevation and season **(Figure 5)**.
+Bell flowers peaked in relative abundance early in the season and at mid-elevation, and lip flowers peaked late in the season and at low elevation.
 Flag flowers peaked in mid-season and exhibited two distinct elevational peaks, the lower at around 1000 m.s.l and the higher at around 1800 m.s.l.
@@
 soroeensis*) **(Figure 1)**.
-It is also apparent that, as a rule, preference is more dynamic where it is also higher overall.
+It is also apparent that preference tends to be more dynamic where it is also higher overall.
 Bell flowers, which were positively selected overall by the short-tongued bumble bee species *B.
-pratorum*, "*B.
+pratorum*, *B.
 terrestris-lucorum*, and *B.
@@
 Nearly the reverse pattern is evident for flag flowers, reflecting their association with longer-tongued bumble bee species.
+Smooths were significant (p < 0.05) for all bee-morphotype pairs except hort-head, prat-flag, telu-head, and soro-flag, and the model explained 36.4% of total deviance (adj.
+R^2^ = 0.33).
 
 \
-
-Smooths were significant (p < 0.05) for all bee-morphotype pairs except hort-head, prat-flag, telu-head, and soro-flag. The model explained 36.4% of total deviance (adj. R^2^ = 0.33).
 
@@
 Bumble bee species differed in their floral morphotype preferences, and these differences could be explained by trait-matching between floral morphotype and bumble bee tongue length.
-These preferences, however, varied by elevation and season, likely representing adaptive responses of bumble bee foraging behavior to elevational and seasonal shifts in floral resource composition, possibly modulated by competitive interactions among bumble bees species with overlapping resource preferences.
+These preferences, however, varied by elevation and season, likely representing adaptive responses of bumble bee foraging behavior to elevational and seasonal shifts in floral resource composition.
+
+\
 
@@
 @Harmon-Threatt2017-bz].
-In general, though, our findings support the hypothesis that bumble bee resource selection involves both the neutral effects of floral abundance and the directed effect of preference, meaning that analytical techniques designed to disentangle these effects, such as the null model method we employ, should be favored in studies of bumble bee foraging.
+In general, though, our findings support the hypothesis that bumble bee resource selection involves both the neutral effects of floral abundance and the directed effect of preference, meaning that analytical techniques designed to disentangle these effects, such as the null modeling method we employ, should be applied in studies of bumble bee foraging.
+
+\
 
@@
 
-We found that the most preferred morphotypes were bell, lip, and flag flowers, which is broadly consistent with visitation patterns reported by previous studies in central Europe [@Schneller2014-ry; @Neumayer1999-il; Farnberger1996-xo], indicating that our findings are regionally generalizeable.
+We found that the most preferred morphotypes were bell, lip, and flag flowers, which is broadly consistent with visitation patterns reported by previous studies in central Europe [@Schneller2014-ry; @Neumayer1999-il; @Farnberger1996-xo], indicating that our findings are regionally generalizeable.
 These previous studies tended to find a stronger role of head flowers than indicated by our preference analysis, but this is probably because they characterized foraging in terms of simple visitation rates rather than null-model-inferred preference.
@@
 Moreover, while average preference toward funnel and head flowers was low, each was strongly preferred by one or two bumble bee species.
-The majority of variation in preference toward floral morphotypes was explainable by bumble bee tongue length, with the more open bell and head morphotypes selected more positively by short-tongued bumble bee species and the deeper funnel, lip, and flag morphotypes selected more positively by medium- and long-tongued bumble bee species.
-The failure of tongue length to explain preference toward disc and stalk-disc flowers reflects the fact that these flowers were almost unanimously ignored by all bumble bee species.
+The majority of variation in preference toward floral morphotypes was explainable by bumble bee tongue length, with the more open bell and head morphotypes preferred by short-tongued bumble bee species and the deeper funnel, lip, and flag morphotypes preferred by medium- and long-tongued bumble bee species.
+The failure of tongue length to explain preference toward disc and stalk-disc flowers reflects the fact that these flowers were almost unanimously ignored by the bumble bees in our study system.
 We cannot say from our data why disc and stalk-disc flowers received such consistently low visitation, but the pattern may be driven in part by the fact that some disc flowers offer pollen but little or no nectar; indeed, Kugler's (1970) typology recognizes this distinction with subdivisions of the disc morphotype into those with (class 1.2) and without (class 1.1) nectar rewards.
@@
 soroeensis* and the *B.
-psythyrus* species-group were anomalous in their strongly positive selection toward head flowers.
+psythyrus* species-group were anomalous in their high preference for head flowers.
 Such unexplained variation in preference may indicate that traits not captured by Kugler's (1970) morphotypology are also important in governing floral selection by bumble bees.
-FLoral color, in particular, is known to be instrumental in the process of flower detection by foraging bumble bees [@Spaethe2001-bq], and some bumble bee species exhibit innate color preferences [@Raine2007-nq].
+Floral color, in particular, is known to be instrumental in the process of flower detection by foraging bumble bees [@Spaethe2001-bq], and some bumble bee species exhibit innate color preferences [@Raine2007-nq].
 This effect can also be seen in field studies, where bumble bees have been shown to over- or under-represent certain floral colors [e.g.
@@
 
+\
+
 ## 4.3 Variability of floral abundance, visitation, and preference through elevation and time
@@
 
+\
+
 ## 4.4 Conclusions
@@
 As climate and land use change drive the redistribution of bumble bees and their floral mutualists, and likely an overall reduction in floral availability in mountain habitats, competition within bumble bee communities is likely to intensify.
-Whether this leads to competitive exclusion will depend on whether interspecific competition exceeds the stabilizing effects of interspecific competition.
-In this regard, the partial dependency of bumble bee floral selection on the neutral effects of floral abundance (in contrast to the *target use hypothesis*) can be expected to favor the coexistence of both bumble bee species and their floral host by increasing the likelihood that species will be self-limited by intraspecific competition [@Valdovinos2021-ya].
+Whether this leads to competitive exclusion will depend on whether interspecific competition exceeds the stabilizing effects of interspecific competition [@Hanski1995-kr].
+In this regard, the partial dependency of bumble bee floral selection on the neutral effects of floral abundance (in contrast to the *target use hypothesis*) can be expected to favor the coexistence of both bumble bee species and their floral host by increasing the likelihood that species will be self-limiting [@Valdovinos2021-ya].
+
+\
 
@@
 
+\newpage
+
 # REFERENCES
@@
 <div id ="refs"></div>
+
+\
 
@@
 Each floral morphotype is illustrated (B) with a representative species common in our study system: 1 = *Potentilla erecta*, 2 = *Gentiana asclepiadea*, 3 = *Campanula glomerata*, 4 = *Silene nutans*, 5 = *Ajuga reptans*, 6 = *Lotus corniculatus*, 7 = *Centaurea jacea*.
-Illustrations are modifications of public domain works obtained from www.plantillustrations.org.
-For details of original sources, see Appendix X.
+Illustrations are modifications of public domain works obtained from http://www.plantillustrations.org.
+For details of original sources, see Appendix 2.
 
@@
 Each point represents the preference/visitation of one bumble bee species toward the corresponding floral morphotype.
-The color of each point in panel A, retained from Figure 2, indicates whether a given observed bias is significantly (p < 0.01) positive or negative with respect to null model simulations.
+The color of each point in panel A, retained from Figure 2, indicates whether preference is significantly (p < 0.01) positive or negative with respect to null model simulations.
 
@@
 Colored bands depict a 2-dimensional GAM smooth representing the relationship between the first two principal components and bumble bee tongue length.
-Morphotype illustrations are as described in Figure 1.
 
@@
 
-**Figure 5:** Variation in proportional flower cover through elevation and season modeled as HGAM surfaces.
+**Figure 5:** Variation in relative floral abundance through elevation and season modeled as HGAM surfaces.
 Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns.

### 2021-09-02 ms_2/ms_2_revision2.Rmd
beginning ms1 revision

@@
 ---
-title: "Floral preferences of mountain bumble bees are driven by trait matching but labile through elevation and time"
+#title: "Floral preferences of mountain bumble bees are driven by trait matching but vary through elevation and season"
 
-author: Douglas B. Sponsler^1^*, Katharina Kallnik^1^, Fabrice Requier^2^, Alice Classen^1^, A. Fabienne Maihoff^1^, Johanna Sieger^3^, Ingolf Steffan-Dewenter^1^
+#author: Douglas B. Sponsler^1^*, Katharina Kallnik^1^, Fabrice Requier^2^, Alice Classen^1^, A. Fabienne Maihoff^1^, Johanna Sieger^3^, Ingolf Steffan-Dewenter^1^
 
@@
 #  - knitcitations.bib
-csl: ecology.csl
+csl: ecologyletters.csl
 # csl: https://www.zotero.org/styles/ecology
@@
 
-^1^ Department of Animal Ecology and Tropical Biology, Biocenter, University of Würzburg, Würzburg, Germany
+<!-- ^1^ Department of Animal Ecology and Tropical Biology, Biocenter, University of Würzburg, Würzburg, Germany -->
 
-^2^ Université Paris-Saclay, CNRS, IRD, UMR Évolution, Génomes, Comportement et Écologie, 91198, Gif-sur-Yvette, France
+<!-- ^2^ Université Paris-Saclay, CNRS, IRD, UMR Évolution, Génomes, Comportement et Écologie, 91198, Gif-sur-Yvette, France -->
 
-^3^ Naturpark Frankenhöhe, Ansbach, Germany
+<!-- ^3^ Naturpark Frankenhöhe, Ansbach, Germany -->
 
-`*` Corresponding author: douglas.sponsler@uni-wuerzburg.de
+**TITLE:** Floral preferences of mountain bumble bees are driven by trait matching but vary through elevation and season
+
+**AUTHORS:**
+
+**Douglas Sponsler** (corresponding author)  
+University of Würzburg  
+Department of Animal Ecology and Tropical Biology  
+Biozentrum der Universität Würzburg C-018  
+Am Hubland  
+97074 Würzburg, Germany  
+Email: douglas.sponsler@uni-wuerzburg.de  
+Telephone: +49 0176 363 20273
+
+**Katharina Kallnik**  
+University of Würzburg  
+Department of Animal Ecology and Tropical Biology  
+97074 Würzburg, Germany  
+Email: katharina.kallnik@uni-wuerzburg.de
+
+**Fabrice Requier**  
+Université Paris-Saclay  
+Évolution, Génomes, Comportement et Écologie  
+91198 Gif-sur-Yvette, France  
+Email: fabrice.requier@egce.cnrs-gif.fr
+
+**Alice Classen**  
+University of Würzburg  
+Department of Animal Ecology and Tropical Biology  
+97074 Würzburg, Germany  
+Email: alice.classen@uni-wuerzburg.de
+
+**A. Fabienne Maihoff**  
+University of Würzburg  
+Department of Animal Ecology and Tropical Biology  
+97074 Würzburg, Germany  
+Email: anne_fabienne.maihoff@uni-wuerzburg.de
+
+**Johanna Sieger**  
+Naturpark Frankenhöhe  
+91522 Ansbach, Germany   
+Email: sieger@lpv-mfr.de
+
+**Ingolf Steffan-Dewenter**  
+University of Würzburg  
+Department of Animal Ecology and Tropical Biology  
+97074 Würzburg, Germany  
+Email: ingolf.steffan@uni-wuerzburg.de
+
+**AUTHORSHIP STATEMENT:** KK and ISD conceived the study.
+KK, JS, and ISD designed and carried out field work.
+DS analyzed the data and wrote the paper with substantial contributions from FR, AC, AFM, KK, and ISD.
+
+**DATA ACCESSIBILITY:** All data used in this study have been archived on DRYAD (https://datadryad.org/stash/dataset/doi:10.5061/dryad.s1rn8pk7f) and will be available upon article publication.
+
+**ARTICLE TYPE:** Letter
+
+**RUNNING TITLE:** Bumble bee floral preferences
+
+**KEYWORDS:** foraging, trait matching, pollinator, *Bombus*, alpine, niche, resource selection
+
+**SECTION LENGTHS**
+
+  * Abstract:
+  * Main text
+  * References:
+  * Figures:
+
+\newpage
 
@@
 Patterns of resource use by animals can clarify how ecological communities have assembled in the past, how they currently function, and how they are likely to respond to future perturbations.
-Using ~13,000 records of bumble- bee-wildflower interactions along a 1400 m elevation gradient, we find that the selection of floral morphotypes by bumble bees is driven both by the directed process of trait-matching and the neutral effects of floral abundance.
-Patterns of floral preference are, moreover, labile through time and elevation, which may be an adaptation to the extreme spatiotemporal floristic turnover of mountain ecosystems.
+Using ~13,000 records of bumble- bee-wildflower interactions along a 1400 m elevation gradient, we find that preferences of bumble bees for certain floral morphotypes are driven both by the directed process of trait-matching and the neutral effect of floral abundance.
+Patterns of floral preference also vary by elevation and season, which may be an adaptation to the strong spatiotemporal turnover of mountain floral communities.
 The flexibility of bumble bee foraging will be tested in coming years as climate and land use change drive rapid changes in floral resource distribution in time and space, particularly in the mountain ecosystems that host the world's richest communities of bumble bees and their floral mutualists.
 
-*Keywords*: foraging, trait matching, pollinator, *Bombus*, alpine, niche, resource selection
+\newpage
 
@@
 
-Plant-pollinator networks are a special case of resource selection in which consumer fitness and resources fitness are positively correlated, causing the traits of plants and pollinators to converge toward compatibility with one another [e.g.
+<!-- Plant-pollinator networks are a special case of resource selection in which consumer fitness and resources fitness are positively correlated, causing the traits of plants and pollinators to converge toward compatibility with one another [e.g.
 @Newman2014-ke].
 The coexistence of multiple plant and pollinator species, however, requires niche differentiation [@Palmer2003-ih; @Johnson2019-su], which can be achieved by trait-based partitioning of interactions among pollinators and their floral hosts [@Willmer2011-ak].
+-->
 
@@
 Warming climates have already caused measurable upslope shifts in mountain bumble bees and their floral hosts [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc], which could introduce novel competitive interactions between bumble bee species that experience increased co-occurrence.
-At the same time, socioeconomic factors have led to widespread abandonment and reforestation of mountain meadows, particularly in the European Alps, which dramatically reduces floral resource availability and suppresses bumble bee diversity and abundance [@Tappeiner2006-jp; @Tasser2007-qo; @Walcher2017-kw].
+At the same time, socioeconomic factors have led to widespread abandonment and reforestation of mountain meadows, particularly in the European Alps, which reduces floral resource availability and suppresses bumble bee diversity and abundance [@Tappeiner2006-jp; @Tasser2007-qo; @Walcher2017-kw].
 The ability of bumble bee species to coexist under these conditions of reshuffled distributions and declining floral resources will depend in part on how they select and partition floral resources, which will in turn affect the viability of their floral mutualists, many of which depend principally on bumble bee visitation for pollination [@Goulson2010-lo].
@@
 First, we ask how bumble bee floral preferences map onto a generalized floral morphotypology [@Kugler1970-gb].
-Then, using the well-established relationship between tongue length and floral selection as a point of reference, we test the *target use hypothesis* by comparing the explanatory power of bumble bee tongue length with respect to observed visitation rates vs.
+Then, using the established relationship between tongue length and floral preference as a point of reference, we test the *target use hypothesis* by comparing the explanatory power of bumble bee tongue length with respect to observed visitation rates vs.
 the residuals of a null model that assumes a linear dependency of selection on resource abundance.
@@
 Preference was generally strong, with the visitation rates of most bumblebee species to most floral morphotypes falling significantly above or below the 99% interval of the null model **(Figure 1)**.
-Nonsignificant preference was observed mainly among the rarer bumblebee species because their small samples sizes lead to wide null model confidence intervals.
+Nonsignificant preference was observed mainly among the rarer bumblebee species, likely because their small samples sizes lead to wide null model confidence intervals.
+Interestingly, bumble bees showed strong species-specific variation in their relative preferences of different floral morphotypes **(Figure 1)**.
 
-```{r fig1, echo = FALSE,  fig.cap = paste("Observed visitation rates compared to null model simulations for each bumble bee species and floral morphotype (A).
-Gray bars depict 99\\% intervals of null model simulations and points depict observed visitation rates.
-Points are color-coded according to where they fall with respect to the 99\\% interval of the null model: red = above (positive preference), blue = below (negative preference), gray = within (nonsignificant preference).
-The difference between the observed value and the center of the null model interval can be interpreted as magnitude of preference.
-Each floral morphotype is illustrated (B) with a representative species common in our study system: 1 = Potentilla erecta, 2 = Gentiana asclepiadea, 3 = Campanula glomerata, 4 = Silene nutans, 5 = Ajuga reptans, 6 = Lotus corniculatus, 7 = Centaurea jacea.
-Illustrations are modifications of public domain works obtained from www.plantillustrations.org.
-For details of original sources, see Appendix X.")}
-knitr::include_graphics("../ms_2/figures/fig1.png")
-```
-
-When floral morphotypes are compared to one another in terms of preference, clear patterns emerge **(Figure 2A)**.
+When floral morphotypes are compared to one another in terms of preference across all bumble bee species, clear patterns emerge **(Figure 2A)**.
 All species exhibited negative preference toward disc flowers, while most species exhibited positive preference toward bell, lip, and flag flowers.
@@
 gerstaeckeri* --- for funnel flowers), reflecting the fact that these morphotypes tended both to be rarely visited and to occur at low abundance.
-preference toward head flowers varied widely from strongly positive to strongly negative.
+Preference toward head flowers varied widely across bumble bee species, ranging from strongly positive to strongly negative.
 
-Differences between floral morphotypes in terms of raw mean visitation rates, reflecting the target use hypothesis, were less pronounced **(Figure 2B)**.
-Bell, lip, flag, and head flowers were visited at high rates by one or more bumble bee species, but median raw mean visitation rates did not differ strongly across morphotypes.
+Differences between floral morphotypes in terms of observed visitation rates, reflecting the target use hypothesis, were less pronounced **(Figure 2B)**.
+Bell, lip, flag, and head flowers were visited at high rates by one or more bumble bee species, but observed visitation rates did not differ strongly across morphotypes.
 
@@
 
-<!-- Before we finalize this, check to make sure the raw visitation rates were really prepared correctly and are not being deflated by a bunch of zeros that should'nt be there. -->
-
-```{r fig2, echo = FALSE,  fig.cap = paste("Preference (A) and observed visitation rate (B) of the bumble bee species toward each floral morphotype.
-Each point represents the preference/visitation of one bumble bee species toward the corresponding floral morphotype.
-The color of each point in panel A, retained from Figure 2, indicates whether a given observed bias is significantly (p < 0.01) positive or negative with respect to null model simulations.")}
-knitr::include_graphics("../ms_2/figures/overall_ktype.png")
-```
-
 ## 3.4 Tongue length as a predictor of floral choice
@@
 Tongue length was a strong predictor of preference but not of observed visitation rates **(Figure 3)**.
-preference toward funnel, bell, lip, flag, and head flowers exhibited significant GAM curves in response bumble bee tongue length, and the model explained 77% of total deviance (adj.
+Preference toward funnel, bell, lip, flag, and head flowers exhibited significant GAM curves in response to bumble bee tongue length, and the model explained 77% of total deviance (adj.
 R^2^ = 0.728).
@@
 Preference toward bell flowers and head flowers exhibited approximately quadratic responses with preference peaking among short-tongued species and declining sharply with increasing tongue length before leveling off.
-preference toward flag flowers exhibited a unimodal peak among medium-tongued species.
-In contrast, observed visitation rates were significantly predicted by tongue length only for head flowers, which exhibited a weakly quadratic decline in preference from a peak among short-tongued species.
+Preference toward flag flowers exhibited a unimodal peak among medium-tongued species.
+In contrast, observed visitation rates were predicted by tongue length only for head flowers, which exhibited a weak decline in preference from a peak among short-tongued species.
 The visitation rate model explained only 25% of total deviance (adj.
 R^2^ = 0.053).
+Again, these results did not differ when preference was based on the worker-only null model (Appendix X).
 
-Again, these results did not differ qualitatively when preference was based on the worker-only null model (Appendix X).
+PCA of bumble bee floral preferences captured more than 80% of total variance in the first two axes **(Figure 4)**, and the GAM surface fit to the firs two axes explained most of the variance in tongue length (deviance explained = 94.7%, adj.
+R^2^ = 0.893, p < 0.001), reflecting the importance of trait matching between tongue length and floral morphotype as a mechanistic basis of preference.
+Longer tongue length was associated with deeper and more closed flower morphotypes (lip, funnel, flag), while shorter tongue length was associated with more shallow and open flower morphologies (disc, stalk-disc, bell, head).
+The strong affinity of *B.
+soroeensis* and *B.
+psithyrus* for head flowers is especially salient.
 
-```{r fig3, echo = FALSE,  fig.cap = paste("preference (A) and observed visitation rates (B) for each morphotype regressed against bumble bee tongue length.
-Each point represents the preference/visitation of one bumble bee species toward the corresponding floral morphotype.
-Curves depict fitted GAM smooths with 95\\% confidence bands.
-Solid lines indicate curves differing significantly (p < 0.05) from a flat line while dashed lines indicated nonsigificant curves.
-Points are color-coded by species, and species are presented in the legend in order of increasing tongue length.")}
-knitr::include_graphics("../ms_2/figures/ktypeXtonguelength_paired.png")
-```
+## 3.5 Variation of morphotype abundance and preference through elevation and season
 
-PCA of preference captured more than 80% of total variance in the first two axes **(Figure 4)**, and the GAM surface fit to the firs two axes explained most of the variance in tongue length (deviance explained = 94.7%, adj.
-R^2^ = 0.893, p < 0.001).
-
-<!-- Need to add a bit of comment here about important patterns in the ordination. -->
-
-Perhaps the most striking pattern
-
-```{r fig4, echo = FALSE,  fig.cap = paste("Principal components biplot of bumble bee species according to their preference toward each floral morphotype.
-Floral morphotypes are abbreviated with their numerical codes and illustrated as in Figure 1.
-Colored bands depict a 2-dimensional GAM smooth representing the relationship between the first two principal components and bumble bee tongue length.
-Morphotype illustrations are as described in Figure 1.")}
-knitr::include_graphics("../ms_2/figures/preference_ordination_mod.png")
-```
-
-## 3.5 Variation of floral abundance, raw visitation, and preference through elevation and time
-
-The proportional abundance of the most visited floral morphotypes --- bell, lip, flag, and head flowers --- was strongly patterned in time and elevation **(Figure 5)**.
+The proportional abundance of the most visited floral morphotypes --- bell, lip, flag, and head flowers --- was strongly patterned by elevation and season **(Figure 5)**.
 Bell flowers peaked in proportional abundance early in the season and at mid-elevation, and lip flowers peaked late in the season and at low elevation.
@@
 
-```{r fig5, echo = FALSE,  fig.cap = paste("Variation in proportional flower cover through elevation and modeled as HGAM surfaces.
-Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns.
-All smooths are significant (p < 0.05).")}
-knitr::include_graphics("../ms_2/figures/flower_cover_gam.png")
-```
-
-Patterns of preference through elevation and time were more complex than those of raw visitation and, by design, more often differed from patterns of floral abundance **(Figure 6)**.
+The morphotype preferences of bumble bee species were also strongly patterned through elevation and season.
 In some cases, though, patterns of preference mirrored patterns of floral availability, suggesting that bumble bees may exhibit functional responses [@Mysterud1998-rs] to floral abundance not entirely accounted for by null models that assume a linear proportionality between abundance and visitation.
-
-One of the more salient patterns is that positive selection for flag flowers was strongly associated with high elevation, except among the species that tended to avoid flag flowers (*B.
+Preference for flag flowers was strongly associated with high elevation, except among the species that tended to avoid flag flowers (*B.
 pratorum* and *B.
@@
 pascuorum*, and *B.
-wurflenii*, which exhibited neutral or weakly positive overall selection toward bell flowers **(Figure 1)**, also show weak patterns of variation in preference toward bell flowers.
+wurflenii*, which exhibited neutral or weakly positive overall selection toward bell flowers **(Figure 1)**, also show weak patterns of variation in preference toward bell flowers **(Figure 6)**.
 Nearly the reverse pattern is evident for flag flowers, reflecting their association with longer-tongued bumble bee species.
 
-Smooths were significant (p < 0.05) for all bb-fl pairs except hort-head, prat-flag, telu-head, and soro-flag. The model explained 36.4% of total deviance (adj. R^2^ = 0.33).
-
-```{r fig6, echo = FALSE,  fig.cap = paste("Variation in preference through elevation and time modeled as HGAM surfaces.
-Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns.
-Significant (p < 0.05) smooths are depicted with solid contour lines and nonsignificant smooths with dashed contour lines.")}
-knitr::include_graphics("../ms_2/figures/preference_lability_gam.png")
-```
+Smooths were significant (p < 0.05) for all bee-morphotype pairs except hort-head, prat-flag, telu-head, and soro-flag. The model explained 36.4% of total deviance (adj. R^2^ = 0.33).
 
@@
 
+Using ~13,000 bumble- bee-wildflower interactions, we characterized bumble bee floral selection in terms of generalized floral morphotypes and found evidence that bumble bee floral selection depends both on the directed process of morphotype preference and the neutral effect of floral abundance.
+Bumble bee species differed in their floral morphotype preferences, and these differences could be explained by trait-matching between floral morphotype and bumble bee tongue length.
+These preferences, however, varied by elevation and season, likely representing adaptive responses of bumble bee foraging behavior to elevational and seasonal shifts in floral resource composition, possibly modulated by competitive interactions among bumble bees species with overlapping resource preferences.
+
 ## 4.1 Floral abundance, floral selection, and the *target use hypothesis*
 
-It is well-known that bumble bees forage non-randomly accordingly to morphological constraints, innate preferences, and experienced rewards [e.g.
-@Kugler1943-uy; @Brian1954-cy; @Inouye1978-qi; @Raine2007-nq; @Maharaj2019-wm].
-In this regard, the results of our null model analysis are as unsurprising as they are unambiguous: bumble bee resource selection almost never represents flora in unaltered proportion to their abundance.
-The *target use hypothesis* [@De_Valpine2013-be; @Harmon-Threatt2017-bz], however, extends this idea to the perfect antithesis of the random foraging model, positing that bumble bee foraging is a process of *pure* preference, completely irrespective of resource abundance.
+Our finding that bumble bee tongue length explained preference (i.e.
+null model residuals) much better than it explained observed visitation rates indicates that, in our study system, abundance and preference function jointly in bumble bee floral selection.
+This casts doubt on the *target use hypothesis* as a general rule of bumble bee foraging, though it is possible that it still obtains under certain circumstances [e.g.
+@Harmon-Threatt2017-bz].
+In general, though, our findings support the hypothesis that bumble bee resource selection involves both the neutral effects of floral abundance and the directed effect of preference, meaning that analytical techniques designed to disentangle these effects, such as the null model method we employ, should be favored in studies of bumble bee foraging.
 
-Given the well-established mechanistic link between bumble bee tongue length and floral selection [e.g.
-@Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn], our finding that bumble bee tongue length explained preference (i.e.
-null model residuals) much better than it explained raw visitation rates indicates that, in our study system, abundance and preference function jointly in bumble bee floral choice.
-This casts doubt on the *target use hypothesis* as a general rule of bumble bee foraging.
-It does not, however, mean that the *target use hypothesis* does not obtain under certain circumstances, such as when a highly preferred resource varies within a nonlimiting rage of availability.
-Moreover, our results suggest something akin to the *target use hypothesis* but formulated negatively with respect to resources that are *not selected*.
-Visitation rates to disc flowers in our system, for example, were very low despite the fact that they often occurred at high abundance, suggesting that while preference seems to interact with resource abundance, avoidance may not.
-In general, though, our findings support the hypothesis that bumble bee resource selection involves both the neutral effects of abundance and the directed effect of preference, which implies that analytical techniques designed to disentangle these effects, such as the null model analysis we employ, should be favored in studies of bumble bee foraging.
-<!-- Accordingly, we shall focus primarily on our calculations of preference, rather than raw visitation rates, as we interpret the other results of our study. -->
+## 4.2 Community-level patterns of morphotype preferences
 
-## 4.2 Overall patterns of preference
+We found that the most preferred morphotypes were bell, lip, and flag flowers, which is broadly consistent with visitation patterns reported by previous studies in central Europe [@Schneller2014-ry; @Neumayer1999-il; Farnberger1996-xo], indicating that our findings are regionally generalizeable.
+These previous studies tended to find a stronger role of head flowers than indicated by our preference analysis, but this is probably because they characterized foraging in terms of simple visitation rates rather than null-model-inferred preference.
+We saw similarly high visitation rates toward head flowers, but our null-model analysis reveals that while head flowers were highly visited, they tended to be under-visited relative to their abundance.
 
-We found that positive selection was concentrated on 3 floral morphotypes: bell, lip, and flag flowers.
-These results are broadly consistent with the raw visitation patterns previously reported by @Schneller2014-ry in the central Alps at Vorarlberg, Neumayer1999-il in the Austrian Alps ~55 km south of our study system, and Farnberger1996-xo in the vicinity of Vienna, suggesting that our findings are at least regionally generalizeable.
-These neighboring studies tended to find a stronger role of head flowers than suggested by our preference analysis, but this is probably because they characterized foraging in terms of raw visitation rather than preference.
-Indeed, our raw visitation data similarly indicate a stronger role of head flowers, but our preference analysis reveals that while head flowers were highly visited, they tended to be under-visited relative to their abundance.
-
-When considering these patterns from the perspective of pollination services to mountain flora, it is important to note that while bumble bees are the dominant Hymenopteran pollinators in our study system, flies (Diptera), and especially the hover flies (Syrphidae), are also abundant flower visitors in mountain ecosystems, and their visitation is mainly restricted to the open flower morphotypes (e.g.
-disc flowers) that receive less attention from bumble bees [@Neumayer1999-il].
-In the context of our research question, we have emphasized the differences in tongue length between bumble bee species, but all bumble bees have relatively long tongues compared to the overall pollinator community, and it is not surprising that their foraging tends to be focused on the deeper flower morphotypes.
-The process of floral resource partitioning *among* bumble bee species is nested within the process of floral partitioning *between* bumble bees and other flower visitors.
-
-While bell, lip, and flag flowers were the most selected morphotypes on average, preference toward these morphotypes varied broadly across bumble bee species.
-Moreover, while average preference toward funnel and head flowers was low, each received strongly positive selection from one or two bumble bee species.
+While bell, lip, and flag flowers were the most preferred morphotypes on average, preference toward these morphotypes varied broadly across bumble bee species.
+Moreover, while average preference toward funnel and head flowers was low, each was strongly preferred by one or two bumble bee species.
 The majority of variation in preference toward floral morphotypes was explainable by bumble bee tongue length, with the more open bell and head morphotypes selected more positively by short-tongued bumble bee species and the deeper funnel, lip, and flag morphotypes selected more positively by medium- and long-tongued bumble bee species.
@@
 
-As mountain bumble bees can be expected to face a future of declining resource availability and altered elevational distributions, it is relevant to ask which species seem poised for the greatest resource conflict should competitive conditions arise.
-While any answer to this question should be regarded as speculative, considering the similarities of morphotype preference between species is a good place to start.
-As a general rule, it can be said that species of similar tongue length will be, on average, more likely to compete for resources than species of dissimilar tongue length.
+Species of similar tongue length will, on average, be more likely to compete for resources than species of dissimilar tongue length.
 Our ordination analysis, though, suggests that there may be significant patterns of similarity and divergence even among species of similar tongue length.
@@
 A red-listed vulnerable species, *B.
-gerstaeckeri* is a pollen specilaist on the lip flowers of the genus *Aconitum* [@Ponchau2006-af; @Schneller2014-ry], and there is evidence that joint elevational shifts of bumble bees and floral resources over the past century have generated more intense exploitation of *Aconitum*, which may indicate nascent competitive conditions for *B.
+gerstaeckeri* is a pollen specialist on the lip flowers of the genus *Aconitum* [@Ponchau2006-af; @Schneller2014-ry], and there is evidence that joint elevational shifts of bumble bees and floral resources over the past century have generated more intense exploitation of *Aconitum*, which may indicate nascent competitive conditions for *B.
 gerstaeckeri* [@Marshall2020-pc].
 
-Broaching the topic of competition raises an issue that can confound any observational study of resource selection: how to disentangle the endogenous process of resource preference from the exogenous process of niche partitioning driven by competition [@Pianka1974-qk].
-Empirical preference could reflect a latent process of resource preference, as we assume in our interpretation.
-Alternatively, though, visitation patterns and the preference inferred from them could be distorted by competitive forces.
-Intraspecific competition in a bumble bee community would tend to force individuals of a species to broaden their diet beyond what they would prefer under non-competitive conditions, while interspecific competition would be expected to cause a narrowing of diet breadth toward resources that can be exploited most efficiently.
-Thus, observed visitation patterns, even when calculated as abundance-relative preference, can only be interpreted in terms of *preference* to the extent that competition can be assumed to be negligible, and assumption must be justified on the basis of additional knowledge of the study system under consideration.
-Our study system in the Berchtesgadener Alps is one of abundant floral resources but harsh abiotic conditions, so it is likely that competition for floral resources is low, or at least transient, and that bumble bee populations are limited instead by the shortness of the growing season, as has been suggested in the similar environment of the Eyne Valley of France [@Iserbyt2008-jn].
-We cannot, however, exclude the possibility that competitive forces distort, to some degree, the visitation patterns we observed.
-Even when competitive conditions do not obtain universally in a given system, competition may occur in a spatially and/or temporally localized way [@Ranta1981-vx], perhaps associated with extreme climatic events that cause reductions in floral resource avaulability [@Rasmont2012-zq; @Thomson2016-gq; @Phillips2018-me].
-This calls to mind the insight of MacArthur1972-qz, that competition need not be frequent to have lasting ecological impacts, and it is likely that the "ghosts of competition past" [@Connell1980-dn] haunt the floral preferences exhibited by bumble bees today.
+When considering community-level patterns of preference and visitation from the perspective of pollination services to mountain flora, it is important to note that while bumble bees are the dominant Hymenopteran pollinators in our study system, flies (Diptera) are also abundant flower visitors in mountain ecosystems, and their visitation is mainly restricted to the open flower morphotypes (e.g.
+disc flowers) that receive less attention from bumble bees [@Neumayer1999-il].
+In the context of our research question, we have emphasized the differences in tongue length between bumble bee species, but all bumble bees have relatively long tongues compared to the overall pollinator community, and it is not surprising that their foraging tends to be focused on the deeper and more closed flower morphotypes.
+The process of floral resource partitioning *among* bumble bee species is nested within the process of floral partitioning *between* bumble bees and other flower visitors.
 
@@
 
-The intraspecific lability of bumble bee foraging preferences (*sensu stricto*) has been demonstrated under laboratory conditions [@Ings2009-vc; @Maharaj2019-wm]; our study reveals the *in situ* lability of preference for multiple bumble bee species through elevational and temporal gradients of floristic turnover.
-While we stress the limitations of interpreting empirical preference in terms of behavioral preference without controlling the joint abundances of all available resources [@Beyer2010-vr], interpreting the dynamics of preference in our study as the adaptive tuning of preference is consistent with the ecological hurdles bumble bees face as characteristically --- and most likely primordially --- mountain species [@Williams1985-ql; @Hines2008-er].
-Where steep temperature gradients and short growing seasons generate extreme spatiotemporal turnover of floral resources, adaptive flexibility of morphological preferences should be favored over fixed innate biases.
+Intraspecific variation of bumble bee foraging preferences has been demonstrated under laboratory conditions [@Ings2009-vc; @Maharaj2019-wm]; our study reveals the *in situ* variation of preference for multiple bumble bee species through elevational and temporal gradients of floristic turnover.
 
-At least three mechanisms could underlie the patterns of preference through elevation and time that we observed.
-To illustrate these hypotheses, we will take an as example the relatively strong preference dynamics of flag flowers in our system, which tended to be most preferred at high-elevation sites **(Figure 8)**.
-Our study does not enable the testing of these hypothesized mechanisms (and their constituent submechanisms), but we articulate them for the sake of clarity and as a guide for future investigations.
-Moreover, the hypotheses we propose not mutually exclusive, and it plausible that all three could obtain simultaneously in a given ecological context.
-
-First, variation in preference toward a given morphotype could be driven by changes in the *intrinsic* reward of that focal morphotype.
-These changes might, in turn, be driven by (1) shifts in species composition of the focal morphotype (e.g.
+At least three mechanisms could underlie the variation of preferences we observed.
+First, variation in preference toward a given morphotype could be driven by changes in the *intrinsic* reward of that focal morphotype, which could be caused by (1) shifts in species composition of the focal morphotype (e.g.
 the occurrence of a particularly rewarding flag flower at high-elevation sites), (2) shifts in the reward levels of individual species (e.g.
-increased nectar production in the flag flower *Lotus corniculatus* under high elevation conditions), or (3) possibly changes in nutritional requirements of bumble bees (e.g.
+increased nectar production in the flag flower *Lotus corniculatus* under high elevation conditions), or (3) changes in nutritional requirements of bumble bees (e.g.
 an increased need for concentrated nectar to support thermoregulation at colder ambient temperatures).
-The hypothesis is consistent with the interpretation of preference as an index of the latent process of absolute preference.
-
-Second, preference toward a given morphotype could vary due to changes in the *relative* reward of the focal morphotype under changing background resource conditions.
+Alternatively, preference toward a given morphotype could vary due to changes in the *relative* reward of the focal morphotype with respect to changing background resource conditions.
 For example, it could be that flag flowers are selected more positively at high elevation because they must compete for visitation with highly attractive lip flowers at lower elevation.
-This hypothesis expresses the fundamental problem of inferring preference *sensu stricto* from selection, namely contingency of selection upon the joint availability of all resource options [@Johnson1980-co; @Aarts2008-fo; @Beyer2010-vr].
-
-The two hypotheses above are both grounded in the intrinsic properties of the floral community, whether those of the focal resource of those of the background resources.
-It is also possible, however, that patterns of resource selection could vary due to competitive interactions within or between bumble bee species.
+Finally, it is possible that patterns of resource selection could vary due to competitive interactions within or between bumble bee species.
 For example, the increased selection of flag flowers at high elevation could be driven by competition release from *B.
@@
 <div id ="refs"></div>
+
+# FIGURES
+**Figure 1**
+```{r fig1, echo = FALSE}
+knitr::include_graphics("../ms_2/figures/fig1.png")
+```
+
+**Figure 2**
+```{r fig2, echo = FALSE}
+knitr::include_graphics("../ms_2/figures/overall_ktype.png")
+```
+
+**Figure 3**
+```{r fig3, echo = FALSE}
+knitr::include_graphics("../ms_2/figures/ktypeXtonguelength_paired.png")
+```
+
+**Figure 4**
+```{r fig4, echo = FALSE}
+knitr::include_graphics("../ms_2/figures/preference_ordination_mod.png")
+```
+
+**Figure 5**
+```{r fig5, echo = FALSE}
+knitr::include_graphics("../ms_2/figures/flower_cover_gam.png")
+```
+
+**Figure 6**
+```{r fig6, echo = FALSE}
+knitr::include_graphics("../ms_2/figures/preference_lability_gam.png")
+```
+
+# FIGURE CAPTIONS
+
+**Figure 1:** Observed visitation rates compared to null model simulations for each bumble bee species and floral morphotype (A).
+Gray bars depict 99% intervals of null model simulations and points depict observed visitation rates.
+Points are color-coded according to where they fall with respect to the 99% interval of the null model: red = above (positive preference), blue = below (negative preference), gray = within (nonsignificant preference).
+The difference between the observed value and the center of the null model interval can be interpreted as magnitude of preference.
+Each floral morphotype is illustrated (B) with a representative species common in our study system: 1 = *Potentilla erecta*, 2 = *Gentiana asclepiadea*, 3 = *Campanula glomerata*, 4 = *Silene nutans*, 5 = *Ajuga reptans*, 6 = *Lotus corniculatus*, 7 = *Centaurea jacea*.
+Illustrations are modifications of public domain works obtained from www.plantillustrations.org.
+For details of original sources, see Appendix X.
+
+**Figure 2:** Preference (A) and observed visitation rate (B) of the bumble bee species toward each floral morphotype.
+Each point represents the preference/visitation of one bumble bee species toward the corresponding floral morphotype.
+The color of each point in panel A, retained from Figure 2, indicates whether a given observed bias is significantly (p < 0.01) positive or negative with respect to null model simulations.
+
+**Figure 3:** Preference (A) and observed visitation rates (B) for each morphotype regressed against bumble bee tongue length.
+Each point represents the preference/visitation of one bumble bee species toward the corresponding floral morphotype.
+Curves depict fitted GAM smooths with 95\\% confidence bands.
+Solid lines indicate curves differing significantly (p < 0.05) from a flat line while dashed lines indicated nonsigificant curves.
+Points are color-coded by species, and species are presented in the legend in order of increasing tongue length.
+
+**Figure 4:** Principal components biplot of bumble bee species according to their preference toward each floral morphotype.
+Floral morphotypes are abbreviated with their numerical codes and illustrated as in Figure 1.
+Colored bands depict a 2-dimensional GAM smooth representing the relationship between the first two principal components and bumble bee tongue length.
+Morphotype illustrations are as described in Figure 1.
+
+**Figure 5:** Variation in proportional flower cover through elevation and season modeled as HGAM surfaces.
+Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns.
+All smooths are significant (p < 0.05).
+
+**Figure 6:** Variation in preference through elevation and time modeled as HGAM surfaces.
+Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns.
+Significant (p < 0.05) smooths are depicted with solid contour lines and nonsignificant smooths with dashed contour lines.

### 2021-09-22 code/ms1_beta_diversity.Rmd
analysis reboot

--- 2021-04-08 code/ms1_beta_diversity.Rmd
@@
 
-### Prepare environment
+# 1. Prepare environment
+## 1.1 Packages 
 ```{r message=FALSE, warning=FALSE}
 library(sf)
+library(magick)
 library(gdm)
@@
 library(ecodist)
-library(gt)
+library(flextable)
 library(tidyverse)
 library(lubridate)
-```
-
-### Define functions
+library(cowplot)
+```
+## 1.2 Define functions
+### 1.2.1 `sfc_as_cols`: convert geometry column back to x and y columns
 ```{r}
@@
 }
-
+```
+
+### 1.2.2 `web_prep`: convert networks to webs arrays
+```{r}
 # Function to convert networks to webs arrays
@@
 }
-
+```
+
+### 1.2.3 `betalinkr_set1`: betalinkr settings and conversion to tibble
+```{r}
 # Function for betalinkr settings and conversion to tibble()
@@
 }
-
+```
+
+### 1.2.4 `interaction_gdm_prep`: convert visitation data into site-pair table for GDM analysis
+```{r}
 # Function for converting visitation data into site-pair table for GDM analysis 
-
 interaction_gdm_prep <- function(x) {
@@
 }
-
+```
+
+### 1.2.5 `biodata_prep_survey`: prepare the floral survey biodata input to gdm's `formatsitepair()` function
+```{r}
 # Function for preparing the floral survey biodata input to gdm's formatsitepair function
@@
 }
-
+```
+
+### 1.2.6 `predData_prep_survey`: prepare the floral survey predictor data input to gdm's `formatsitepair()` function
+```{r}
 # Function for preparing the floral survey predictor data input to gdm's formatsitepair function
@@
 }
-
+```
+
+### 1.2.7 `biodata_prep_BB`: prepare the visitation biodata input to gdm's `formatsitepair()` function
+```{r}
 # Function for preparing the visitation biodata input to gdm's formatsitepair function
@@
 }
-
+```
+
+### 1.2.8 `predData_prep_BB`: prepare the visitation predictor data input to gdm's `formatsitepair()` function
+```{r}
 # Function for preparing the visitation predictor data input to gdm's formatsitepair function
@@
 }
-
+```
+
+### 1.2.9 `format_set`: setting for gdm's `formatsitepair()` function
+```{r}
 # Function with setting for gdm's formatsitepair function
@@
 } 
-
+```
+
+### 1.2.10 `gdm_tab`: tabulate `gdm.varImp()` output 
+```{r}
 # Function for tabulating gdm.varImp output 
@@
     rownames_to_column(var = "predictor") %>%
-    mutate(response = rep(response, n()),
-           Predictor = case_when(
+    mutate(model = rep(response, n()),
+           predictor = case_when(
              predictor == "elev.mean" ~ "elevation",
-             predictor == "samples" ~ "sampling freq.",
-             predictor == "mowing" ~ "mowing",
-             predictor == "grasing" ~ "grasing",
-             predictor == "none" ~ "none")) %>%
-    select(response, Predictor, Importance, pval) %>%
-    tibble()
-}
-```
-
-### Load site data
-```{r, include=FALSE, echo=FALSE}
+             predictor == "samples" ~ "samples",
+             predictor == "Geographic" ~ "distance")
+             # predictor == "mowing" ~ "mowing",
+             # predictor == "grasing" ~ "grazing",
+             # predictor == "none" ~ "none")
+           ) %>%
+    mutate(importance = round(importance, 2),
+           model.var.exp = round(model.var.exp, 2)) %>%
+    select(Model = model, Variable = predictor, 
+           "Importance [%]" = importance, Var.pval = pval, 
+           Model.pval = model.pval, "Model.dev.exp [%]" = model.var.exp) %>%
+    tibble() %>%
+    mutate(Variable = factor(Variable, levels = c("elevation", "distance", "samples"))) %>%
+    arrange(Variable)
+}
+```
+
+### 1.2.11 `isplineExtract2`: Fixes a bug in `isplineExtract()` that causes it to break when non-default number of splines are used
+```{r}
+##Extracts Ispline data from a gdm model
+isplineExtract2 <- function (model){
+  ###########################
+  #model = gdmOb
+  ###########################
+  ##error checking
+  ##checks to make sure a gdm model is given
+  if(!is(model, "gdm")){
+    stop("model argument must be of class = 'gdm'.")
+  }
+
+  ##Collects or sets simple data
+  options(warn.FPU = FALSE)
+  PSAMPLE <- 200
+  preddata <- rep(0, times = PSAMPLE)
+  pn <- model$predictors
+  nPreds <- length(pn)
+  yDat <- xDat <- matrix(0,PSAMPLE,nPreds)
+  colnames(yDat) <- colnames(xDat) <- pn
+  pmin <- 1
+  pmax <- PSAMPLE
+
+  ##cycles through each prodictor and fills the spline matrices
+  splineindex <- 1
+  for (i in 1:nPreds){ 
+    #i<-1
+    numsplines <- model$splines[i]
+    z <- .C("GetPredictorPlotData", 
+            pdata = as.double(preddata), 
+            as.integer(PSAMPLE), 
+            as.double(model$coefficients[splineindex:(splineindex + numsplines - 1)]), 
+            as.double(model$knots[splineindex:(splineindex + numsplines - 1)]), 
+            as.integer(numsplines),
+            PACKAGE = "gdm")
+    yDat[,i] <- z$pdata
+    pmin <- pmin + PSAMPLE
+    pmax <- pmax + PSAMPLE
+    xDat[,i] <-  seq(from=model$knots[[(i*numsplines)-(numsplines-1)]],to=model$knots[[(i*numsplines)]], length=PSAMPLE)
+    splineindex <- splineindex + numsplines
+  }
+
+  ##lists and returns matrices
+  outData <- list(x=xDat,y=yDat)
+  return(outData)
+}
+```
+
+### 1.2.12 `plotUncertainty2` Tweaked to call `isplineExtract2` instead of the broken version
+```{r}
+##########################################################################
+plotUncertainty2 <- function(spTable, sampleSites, bsIters, geo=FALSE, splines=NULL, 
+                           knots=NULL, splineCol="blue", errCol="grey80", 
+                           plot.linewidth=2.0, plot.layout=c(2,2), parallel=FALSE,
+                           cores=2){
+  ##A function to plot the uncertantiy of each variable from a GDM model
+  #################
+  #spTable <- sitePairTab          ##the input site-pair table to subsample from
+  #sampleSites <- 0.9     ##fraction of sites that should be retained from site pair table 
+  #bsIters <- 50       ##the number of times the site-pair table should be sampled
+  #geo <- F              ##rather or not the gdm model takes geography into account, see gdm
+  #splines <- NULL       ##splines gdm setting, see gdm
+  #knots <- NULL         ##knots gdm setting, see gdm
+  #splineCol <- "blue"    ##color of the center line
+  #errCol <- "grey80"        ##color of the uncertainty polygon
+  #plot.linewidth <- 2.0    ##line width of the center line
+  #plot.layout <- c(3,3)    ##number of plots per page
+  #parallel <- F            ##rather or not the sampling should happen in parallel processing, to speed it up
+  #cores <- 6               ##number of cores to if parallel processing
+  #################
+  ##function breaks and warnings
+  ##makes sure that table is a properly formated site-pair table
+  if(!is(spTable, "gdmData")){
+    warning("The spTable object is not of class 'gdmData'. See the formatsitepair function for help.")
+  }
+  ##checks to makes sure table is a matrix or table frame
+  if(!(is(spTable, "gdmData") | is(spTable, "matrix") | is(spTable, "data.frame"))){
+    stop("The spTable object must be of class 'gdmData', 'matrix', or 'data.frame'.")
+  }
+
+  ##sanity check on the data table
+  if(ncol(spTable) < 6){
+    stop("spTable object requires at least 6 columns: Observed, weights, s1.xCoord, s1.yCoord, s2.xCoord, s2.yCoord")
+  }  
+  if(nrow(spTable) < 1){
+    stop("spTable object has < 1 rows.")
+  }
+
+  ##checks that geo has either TRUE or FALSE
+  if(!(geo==TRUE | geo==FALSE)){
+    stop("geo argument must be either TRUE or FALSE")
+  }
+  ##makes sure splines is a numeric vector
+  if(is.null(splines)==FALSE & !is(splines, "numeric")){
+    stop("splines object must of of class = 'numeric'.")
+  }
+  ##checks knots inputs
+  if(is.null(knots)==FALSE & !is(knots, "numeric")){
+    stop("knots object must of of class = 'numeric'.")
+  }
+
+  ##checks that parallel has either TRUE or FALSE
+  if(!(parallel==TRUE | parallel==FALSE)){
+    stop("parallel argument must be either TRUE or FALSE")
+  }
+  ##makes sure that cores has a value when parallel is true
+  if(parallel==TRUE & is.null(cores)==TRUE){
+    stop("If parallel==TRUE, the number of cores must be specified")
+  }
+  ##makes sure that cores is a positive integer 
+  if((is.null(cores)==FALSE & is.numeric(cores)==FALSE) | cores<1){
+    stop("argument cores needs to be a positive integer")
+  }
+
+  ##makes sure that bsIters is a positive integer 
+  if((is.null(bsIters)==FALSE & is.numeric(bsIters)==FALSE) | bsIters<1){
+    stop("argument bsIters needs to be a positive integer")
+  }
+  ##makes sure sampleSites is a number
+  if(is.numeric(sampleSites)==FALSE){
+    stop("sampleSites must be a number between 0 and 1")
+  }
+  ##makes sure that sampleSites is between 0 and 1
+  if(sampleSites < 0){
+    stop("sampleSites must be a number between 0 and 1")
+  }
+  if(sampleSites > 1){
+    stop("sampleSites must be a number between 0 and 1")
+  }
+  if(sampleSites==0){
+    stop("a sampleSites value of 0 will remove all sites from the analysis (bad).")
+  }
+  ##double makes sure these values are integers, seems to truncate if not
+  cores <- as.integer(cores)
+  bsIters <- as.integer(bsIters)
+
+  ##assign k to prevent issues to cran checking
+  k <- NULL
+
+  ##makes copies of the site-pair table in order to randomly subsample each one differently
+  lstSP <- lapply(1:bsIters, function(i){spTable})
+
+  ##runs parallel if desired by the users
+  if(parallel==TRUE){
+    ##loads libraries
+    #require(foreach)
+    #require(doParallel)
+    #requireNamespace("foreach")
+    #requireNamespace("parallel")
+    #requireNamespace("doParallel")
+
+    ##sets cores
+    cl <- makeCluster(cores, outfile="")
+    registerDoParallel(cl)
+    ##first removes a number of sites according to input
+    subSamps <- foreach(k=1:length(lstSP), .verbose=F, .packages=c("gdm")) %dopar%
+                      removeSitesFromSitePair(lstSP[[k]], sampleSites=sampleSites)
+    ##models the subsamples
+    gdmMods <- foreach(k=1:length(subSamps), .verbose=F, .packages=c("gdm")) %dopar%
+                     #gdmMods <- try(foreach(k=1, .verbose=F, .packages=c("gdm")) %dopar%
+                     gdm(subSamps[[k]], geo=geo, splines=splines, knots=knots)
+    stopCluster(cl)
+  }else{
+    ##first removes a number of sites according to input
+    subSamps <- lapply(lstSP, removeSitesFromSitePair, sampleSites=sampleSites)
+    ##models the subsamples
+    gdmMods <- lapply(subSamps, gdm, geo=geo, splines=splines, knots=knots)
+  }
+
+  ##models the full gdm
+  fullGDMmodel <- gdm(spTable, geo=geo, splines=splines, knots=knots)
+
+  ##Extracts the splines for each model
+  exUncertSplines <- lapply(gdmMods, isplineExtract2)
+  fullGDMsplines <- isplineExtract2(fullGDMmodel)
+
+  ##get the names of the predictor variables
+  predVars <- colnames(exUncertSplines[[1]][[1]])
+
+  ##establish what plot layout to use
+  thisplot <- 0
+  one_page_per_plot <- FALSE
+  if ((plot.layout[1]==1) && (plot.layout[2]==1)){
+    one_page_per_plot <- TRUE
+  }else{
+    par(mfrow=plot.layout)
+  }
+
+  ##sets the plotting minimum and maximum y-values 
+  totalYmin <- Inf
+  totalYmax <- -Inf
+
+  ##determines the bounds of the plots
+  for(p in 1:length(predVars)){
+    #p=1
+    predV <- predVars[p]
+
+    ##gets the minimum and maximum, to set the ploting extent
+    for(nm in 1:length(exUncertSplines)){
+      #nm=1
+      selPlot <- exUncertSplines[[nm]]
+
+      spYmax <- max(selPlot[[2]][,predV])
+      spYmin <- min(selPlot[[2]][,predV])
+
+      totalYmax <- max(c(totalYmax, spYmax))
+      totalYmin <- min(c(totalYmin, spYmin))
+    }
+  }
+
+  ##plots by variable
+  for(p in 1:length(predVars)){
+    #p <- 1
+    predV <- predVars[p]
+
+    ##sets the plotting minimum and maximum x-values
+    totalXmin <- Inf
+    totalXmax <- -Inf
+    ##gets the minimum and maximum, to set the ploting extent
+    for(nm in 1:length(exUncertSplines)){
+      #nm=1
+      selPlot <- exUncertSplines[[nm]]
+
+      spXmax <- max(selPlot[[1]][,predV])
+      spXmin <- min(selPlot[[1]][,predV])
+
+      if(spXmax > totalXmax){totalXmax = spXmax}
+      if(spXmin < totalXmin){totalXmin = spXmin}
+    }
+
+    ##checks to make sure that there is some significance, if there is the the data is plotted
+    if(totalYmax!=0){
+      ##add mean of subsets to plot
+      plotX <- NULL
+      plotY <- NULL
+      byVarMatX <- NULL
+      byVarMatY <- NULL
+      ##create matrices based on the variable and its x and y location for each model iteration
+      for(nn in 1:length(exUncertSplines)){
+        #nn=1
+        plotX[[nn]] <- exUncertSplines[[nn]][[1]]
+        plotY[[nn]] <- exUncertSplines[[nn]][[2]]
+        byVarMatY <- cbind(byVarMatY, plotY[[nn]][,predV])
+        byVarMatX <- cbind(byVarMatX, plotX[[nn]][,predV])
+      }
+      ##gets spline data from the full gdm model
+      fullPlotX <- fullGDMsplines[[1]]
+      fullPlotX <- fullPlotX[,predV]
+      fullPlotY <- fullGDMsplines[[2]]
+      fullPlotY <- fullPlotY[,predV]
+
+      ##calculates the confidence intervals for plotting
+      sdX <- apply(as.matrix(byVarMatX), 1, sd)
+      sdY <- apply(as.matrix(byVarMatY), 1, sd)
+      highBoundX <- fullPlotX + sdX
+      lowBoundX <- fullPlotX - sdX
+      highBoundY <- fullPlotY + sdY
+      lowBoundY <- fullPlotY - sdY
+
+      ##collects the data to be used in the rug plot
+      if(predV=="Geographic"){
+        ##calculates unique eucildian distance between sites
+        rugData <- unique(sqrt(((spTable$s1.xCoord-spTable$s2.xCoord)^2)+((spTable$s1.yCoord-spTable$s2.yCoord)^2)))
+      }else{
+        ##gets unique values of variable data
+        varDat <- grep(predV, colnames(spTable))
+        rugData <- unique(c(spTable[,c(varDat[1])], spTable[,c(varDat[2])]))
+      }
+
+      ##plots one graph per page, unless specified otherwise
+      if (one_page_per_plot){
+        dev.new()
+        dev.next()
+      }else{
+        thisplot <- thisplot + 1
+        if(thisplot > (plot.layout[1] * plot.layout[2])){    			
+          thisplot <- 1
+          par(mfrow=plot.layout)	
+        }
+      }
+      #settings <- par(pars)
+      ##plots mean data line and polygon of uncertanty 
+      plot(NULL, xlim=c(totalXmin, totalXmax), ylim=c(totalYmin, totalYmax), xlab=predV, ylab="Partial Ecological Distance")
+      polygon(c(lowBoundX, rev(highBoundX)), c(lowBoundY, rev(highBoundY)), col=errCol, border=NA)
+      lines(fullPlotX, fullPlotY, col=splineCol, lwd=plot.linewidth)
+      rug(rugData)
+    }
+  }
+}
+##########################################################################
+```
+
+## 1.3 Load data
+```{r}
+net <- read_csv("../data/processed_data/network.csv") %>%
+  filter(caste  != "male") %>%
+  unite(site.year.day, site, year, dayofyear, sep = "_", remove = FALSE) %>% 
+  mutate(year = factor(year))
+
+survey <- read_csv("../data/processed_data/floral_survey.csv") %>%
+  filter(flower.cover > 0) %>%
+  semi_join(net, by = "plant.sp") %>% # Consider only visited plant spp. (anytime, anywhere, not specifically for a given site*date)
+  dplyr::select(site, date, plant.sp, flower.cover)
+
 site_data <- read_csv("../data/processed_data/site_data.csv") %>%
@@
          tree_line = factor(tree_line)) %>% 
-  semi_join(read_csv("../data/processed_data/network.csv"), by = "site") %>% # drop sites that were not sampled
+  semi_join(net, by = "site") %>% # drop sites that were not sampled
   st_as_sf(coords = c("lon", "lat")) %>%
@@
   dplyr::select(-geometry)
-
-# Write to csv for archiving
-write_csv(site_data, "../Dryad_ms1/site_data.csv")
-```
-
-### Load network and survey data
-```{r}
-net <- read_csv("../data/processed_data/network.csv") %>%
-  filter(caste  != "male") %>%
-  unite(site.year.day, site, year, dayofyear, sep = "_", remove = FALSE) %>% 
-  mutate(year = factor(year))
-
-survey <- read_csv("../data/processed_data/floral_survey.csv") %>%
-  filter(flower.cover > 0) %>%
-  semi_join(net, by = "plant.sp") %>% # Consider only visited plant spp. (anytime, anywhere, not specifically for a given site*date)
-  dplyr::select(site, date, plant.sp, flower.cover)
-
-# Write to csv for archiving
-write_csv(net, "../Dryad_ms1//visitation_data.csv")
-write_csv(survey, "../Dryad_ms1//floral_survey_data.csv")
-```
-
-### Get beta-diversity for BBs and plants (not their interactions)
+```
+
+## 1.4 Load images
+```{r}
+BB <- image_read("../ms_1/figures/phylopic_BB.png")
+FL <- image_read("../ms_1/figures/phylopic_FL.png")
+BBxFL <- image_read("../ms_1/figures/phylopic_BBxFL.png")
+```
+
+# 2. Process data
+## 2.1 BB and FL beta-diversity
 ```{r}
@@
   pivot_longer(cols = c(BB, FL), names_to = "group", values_to = "beta") %>%
-  filter(elev.diff >= 0) # For each site pair, one will be higher and the other lower; thus, when their order is reversed only the sign of the difference changes.
-Since no two sites have exactly the same elevation, this deduplicates the data by dropping the ordering of a given pair of sites (one of the two orderings) that yields a negative elevation difference while retaining the alternative ordering that yields a positive elevation difference.
+  filter(elev.diff >= 0)
+
+# For each site pair, one will be higher and the other lower; thus, when their order is reversed only the sign of the difference changes.
+Since no two sites have exactly the same elevation, `filter(elev.diff >= 0)` deduplicates the data by dropping the ordering of a given pair of sites (one of the two orderings) that yields a negative elevation difference while retaining the alternative ordering that yields a positive elevation difference.
 By >= 0, I retain comparisons of a site to itself.
@@
 
-# fl_beta_dist <- survey %>%
-#   group_by(site, plant.sp) %>%
-#   summarize(count = sum(flower.cover)) %>%
-#   pivot_wider(names_from = plant.sp, values_from = count) %>%
-#   replace(is.na(.), 0) %>%
-#   column_to_rownames("site") %>%
-#   vegdist(method = "jaccard", binary = TRUE, upper = TRUE) 
-# 
-# bb_beta_dist <- net %>%
-#   group_by(site, bb.sp) %>%
-#   summarize(count = n()) %>% 
-#   pivot_wider(names_from = bb.sp, values_from = count) %>%
-#   replace(is.na(.), 0) %>%
-#   column_to_rownames("site") %>%
-#   vegdist(method = "jaccard", binary = TRUE)
-```
-
-### Tabulate webs
+# Fun with betapart
+fl_betapart <- survey %>%
+  group_by(site, plant.sp) %>%
+  summarize(count = sum(flower.cover)) %>%
+  mutate(present = ifelse(count > 0, 1, 0)) %>%
+  select(-count) %>%
+  pivot_wider(names_from = plant.sp, values_from = present) %>%
+  replace(is.na(.), 0) %>%
+  column_to_rownames("site") %>%
+  beta.pair(index.family = "jaccard")
+
+fl_betapart.turnover <- fl_betapart$beta.jtu %>%
+  as.matrix() %>%
+  as.data.frame() %>%
+  rownames_to_column(var = "site1") %>%
+  pivot_longer(cols = -site1, names_to = "site2", values_to = "tu") %>%
+  filter(site1 != site2) %>%
+  left_join(site_data, by = c("site1" = "site")) %>%
+  dplyr::select(site1, site2, tu, tree_line1 = tree_line) %>%
+  left_join(site_data, by = c("site2" = "site")) %>%
+  dplyr::select(site1, site2, tu, tree_line1, tree_line2 = tree_line)
+
+fl_betapart.nestedness <- fl_betapart$beta.jne %>%
+  as.matrix() %>%
+  as.data.frame() %>%
+  rownames_to_column(var = "site1") %>%
+  pivot_longer(cols = -site1, names_to = "site2", values_to = "ne") %>%
+  filter(site1 != site2) %>%
+  left_join(site_data, by = c("site1" = "site")) %>%
+  dplyr::select(site1, site2, ne, tree_line1 = tree_line) %>%
+  left_join(site_data, by = c("site2" = "site")) %>%
+  dplyr::select(site1, site2, ne, tree_line1, tree_line2 = tree_line)
+
+fl_betapart.collate <- full_join(fl_betapart.turnover, fl_betapart.nestedness) %>%
+  mutate(tree_line.comp = case_when(
+    tree_line1 == "below" & tree_line2 == "below" ~ "bb",
+    tree_line1 == "above" & tree_line2 == "above" ~ "aa",
+    tree_line1 != tree_line2 ~ "ab"
+  )) %>%
+  pivot_longer(cols = c(tu, ne), values_to = "value", names_to = "var")
+
+ggplot(fl_betapart.collate, aes(var, value)) +
+  geom_boxplot() +
+  facet_wrap(~tree_line.comp)
+```
+
+## 3.2  Interaction beta-diversity and partitioning
 S	= beta_S, the dissimilarity in species composition
@@
 # Write to csv for archiving
-write_csv(beta_mit_data, "../Dryad_ms1/partitioned_beta_diversity")
-```
-
-### Plot interaction beta-diversity against elevation difference
+#write_csv(beta_mit_data, "../Dryad_ms1/partitioned_beta_diversity")
+```
+
+# 4. Visualize beta-diversity ~ elevation diff
 ```{r}
@@
 # BB + FL elevation diff
-g1 <- ggplot(filter(species_beta, 
+beta_plot1 <- ggplot(filter(species_beta, 
                     site1 != site2), # no within-site comparisons
@@
               se = FALSE) +
-  xlab("Elevation difference (m)") +
-  ylab("β-diversity") +
+  labs(x = NULL,
+       y = "Species turnover",
+       color = "Taxon") +
   theme_light()
@@
 # BB * FL elevation diff (aggregate)
-g2 <- ggplot(filter(beta_mit_data, 
+beta_plot2 <- ggplot(filter(beta_mit_data, 
                       site1 != site2 & # no within-site comparisons
@@
               se = FALSE) +
-  scale_color_manual(values = p1, name = "β partition") +
+  scale_color_manual(values = p1, name = "Partition") +
   xlab(NULL) +
-  ylab("Prop. β-diversity") +
+  ylab("Interaction turnover") +
   theme_light()
@@
 # BB * FL elevation diff (ST partition)
-g3 <- ggplot(filter(beta_mit_data, 
+beta_plot3 <- ggplot(filter(beta_mit_data, 
                       site1 != site2 & # no within-site comparisons
@@
   scale_color_manual(values = p2, name = "ST partition") +
-  xlab("Elevation difference (m)") +
+  xlab("Elevation difference [m]") +
   ylab("Prop. species turnover") +
@@
 # Patchwork plot for comparing species beta-diversity with interaction beta-diversity
-png("../output/beta_stack.png", width = 6.5, height = 4.5, units = "in", res = 300)
-(g1 | (g2 / g3)) + 
-  plot_annotation(tag_levels = c("A", "B", "C")) + 
-  theme_light() 
-dev.off()
-
-pdf("../output/beta_stack.pdf", width = 6.5, height = 4.5)
-(g1 | (g2 / g3)) + 
-  plot_annotation(tag_levels = c("A", "B", "C")) + 
-  theme_light() 
-dev.off()
-```
-
-### MRM analysis (not included in paper)
+beta_plot <- (beta_plot1 / beta_plot2 / beta_plot3) + 
+  plot_annotation(tag_levels = c("A", "B", "C"))
+# & 
+#   theme_classic()
+
+ggsave("../ms_1/figures/fig6.pdf", beta_plot, height = 8.5, width = 4.5)
+```
+
+# 5. GDM analysis
+
+## 5.1 Interactions
+```{r message=FALSE, warning=FALSE}
+# Prep sitepairs
+## GDM is not really built for handling categorical variables; when I include 
+## management in the models, they effect is weak, and I don't know how to
+## interpret the curves. So, I think it is better to drop the mgmt variables
+## from the models and just interpret the continuous effects.
+sitepair_WN <- interaction_gdm_prep(net) %>%
+  select(-c(s1.mowing, s1.grasing, s1.none, 
+            s2.mowing, s2.grasing, s2.none))
+
+# Run GDM analysis
+gdm_WN <- gdm(sitepair_WN, 
+              geo = TRUE, 
+              splines = c(6, 6, 6))
+
+summary(gdm_WN)
+
+gdm.varImp_WN <- gdm.varImp(sitepair_WN, 
+                            geo = TRUE,
+                            fullModelOnly = TRUE,
+                            splines = c(6, 6, 6),
+                            nPerm = 500)
+
+# Plot
+plot(gdm_WN)
+
+# Plot with confidence intervals 
+# pdf("../output/gdm_interactions.pdf", height = 2.5, width =  6.5)
+# plotUncertainty2(sitepair_WN, 
+#                 sampleSites = 0.7, 
+#                 bsIters = 100, 
+#                 geo  = TRUE,
+#                 splines = c(6, 6, 6),
+#                 splineCol = "black", 
+#                 plot.layout = c(1,3))
+# dev.off()
+```
+
+## 5.2 Bumble bees
+```{r}
+## GDM is not really built for handling categorical variables; when I include 
+## management in the models, they effect is weak, and I don't know how to
+## interpret the curves. So, I think it is better to drop the mgmt variables
+## from the models and just interpret the continuous effects.
+
+# Prep input for GDM
+bioData_BB <- biodata_prep_BB(net)
+predData_BB <- predData_prep_BB(net) %>%
+  select(-c(mowing, grasing, none))
+sitepair_BB <- format_set(bioData_BB, predData_BB)
+
+# Call GDMs
+gdm_BB <- gdm(sitepair_BB, 
+              geo = TRUE, 
+              splines = c(6, 6, 6))
+
+# Summarize
+summary(gdm_BB)
+
+# Determine variable importance and significance by permutation test
+gdm.varImp_BB <- gdm.varImp(sitepair_BB, 
+                            geo = TRUE,
+                            fullModelOnly = TRUE, 
+                            splines = c(6, 6, 6),
+                            nPerm = 500)
+
+# Preliminary plot
+plot(gdm_BB)
+
+# Plot with confidence intervals 
+# pdf("../output/gdm_BB.pdf", height = 2.5, width =  6.5)
+# plotUncertainty2(sitepair_BB, 
+#                  sampleSites = 0.7, 
+#                  bsIters = 100, 
+#                  geo  = TRUE,
+#                  splines = c(6, 6, 6),
+#                  splineCol = "black", 
+#                  plot.layout = c(1,3))
+# dev.off() 
+```
+
+## 5.3 Flora
+```{r}
+# Prep input for GDM
+bioData_survey <- biodata_prep_survey(survey)
+predData_survey <- predData_prep_survey(survey) %>%
+  select(-c(mowing, grasing, none))
+sitepair_survey <- format_set(bioData_survey, predData_survey)
+
+# Call GDM
+gdm_survey <- gdm(sitepair_survey, geo = TRUE, splines = c(6, 6, 6))
+
+# Summarize
+summary(gdm_survey)
+
+# Determine variable importance and significance by permutation test
+gdm.varImp_survey <- gdm.varImp(sitepair_survey, 
+                                geo = TRUE,
+                                fullModelOnly = TRUE,
+                                splines = c(6, 6, 6),
+                                nPerm = 500)
+
+# Preliminary plot
+plot(gdm_survey)
+
+# Plot with confidence intervals 
+# pdf("../output/gdm_survey.pdf", height = 2.5, width =  6.5)
+# plotUncertainty2(sitepair_survey, 
+#                  sampleSites = 0.7, 
+#                  bsIters = 100, 
+#                  geo  = TRUE,
+#                  splines = c(6, 6, 6),
+#                  splineCol = "black", 
+#                  plot.layout = c(1,3))
+# dev.off()
+```
+
+## 5.4 Tabulate `gdm.varImp()` results
+### 5.4.1 Interactions
+```{r}
+# Extract tables
+gdm.varImp_WN.tab1 <- gdm.varImp_WN[[1]] %>%
+  as.data.frame() %>%
+  dplyr::select(val = 1)
+
+gdm.varImp_WN.tab2 <- gdm.varImp_WN[[2]] %>%
+  as.data.frame() %>%
+  dplyr::select(importance = 1)
+
+gdm.varImp_WN.tab3 <- gdm.varImp_WN[[3]] %>%
+  as.data.frame() %>%
+  dplyr::select(pval = 1)
+
+# Assemble summary table
+gdm.varImp_WN.tab <- bind_cols(gdm.varImp_WN.tab2, gdm.varImp_WN.tab3) %>%
+  mutate(model.pval = rep(gdm.varImp_WN.tab1$val[3], n())) %>%
+  mutate(pval = as.character(pval),
+         pval = if_else(pval == "0", "<0.001", pval),
+         model.pval = as.character(model.pval),
+         model.pval = if_else(model.pval == "0", "<0.001", model.pval),
+         model.var.exp = rep(gdm.varImp_WN.tab1$val[2]), n()) %>%
+  gdm_tab("Interactions")
+```
+
+### 5.4.2 Bumble bees
+```{r}
+# Extract tables
+gdm.varImp_BB.tab1 <- gdm.varImp_BB[[1]] %>%
+  as.data.frame() %>%
+  dplyr::select(val = 1)
+
+gdm.varImp_BB.tab2 <- gdm.varImp_BB[[2]] %>%
+  as.data.frame() %>%
+  dplyr::select(importance = 1)
+
+gdm.varImp_BB.tab3 <- gdm.varImp_BB[[3]] %>%
+  as.data.frame() %>%
+  dplyr::select(pval = 1)
+
+# Assemble summary table
+gdm.varImp_BB.tab <- bind_cols(gdm.varImp_BB.tab2, gdm.varImp_BB.tab3) %>%
+  mutate(model.pval = rep(gdm.varImp_BB.tab1$val[3], n())) %>%
+  mutate(pval = as.character(pval),
+         pval = if_else(pval == "0", "<0.001", pval),
+         model.pval = as.character(model.pval),
+         model.pval = if_else(model.pval == "0", "<0.001", model.pval),
+         model.var.exp = rep(gdm.varImp_BB.tab1$val[2]), n()) %>%
+  gdm_tab("Bumble bees")
+```
+
+### 5.4.3 Flora
+```{r}
+# Extract tables
+gdm.varImp_survey.tab1 <- gdm.varImp_survey[[1]] %>%
+  as.data.frame() %>%
+  dplyr::select(val = 1)
+
+gdm.varImp_survey.tab2 <- gdm.varImp_survey[[2]] %>%
+  as.data.frame() %>%
+  dplyr::select(importance = 1)
+
+gdm.varImp_survey.tab3 <- gdm.varImp_survey[[3]] %>%
+  as.data.frame() %>%
+  dplyr::select(pval = 1)
+
+# Assemble summary table
+gdm.varImp_survey.tab <- bind_cols(gdm.varImp_survey.tab2, gdm.varImp_survey.tab3) %>%
+  mutate(model.pval = rep(gdm.varImp_survey.tab1$val[3], n())) %>%
+  mutate(pval = as.character(pval),
+         pval = if_else(pval == "0", "<0.001", pval),
+         model.pval = as.character(model.pval),
+         model.pval = if_else(model.pval == "0", "<0.001", model.pval),
+         model.var.exp = rep(gdm.varImp_survey.tab1$val[2]), n()) %>%
+  gdm_tab("Flora")
+```
+
+### 5.4.4 Merge and format
+```{r}
+colourer <- scales::col_numeric(
+  palette = c("transparent", "red"),
+  domain = c(0, 100))
+
+# Merge and format
+gdm.varImp.tab <- bind_rows(gdm.varImp_BB.tab,
+                            gdm.varImp_survey.tab, 
+                            gdm.varImp_WN.tab
+                            ) %>%
+  as_grouped_data(group = "Model") %>%
+  flextable::as_flextable() %>%
+  merge_v(j = c("Model.pval", "Model.dev.exp [%]")) %>%
+  valign(j = c("Model.pval", "Model.dev.exp [%]"), valign = "top") %>%
+  autofit() %>%
+  bg(j = "Importance [%]",
+     bg = colourer, part = "body")
+
+save_as_image(gdm.varImp.tab, "../ms_1/figures/table4.png")
+```
+
+## 5.5. Plot GDM splines
+```{r}
+# Interaction model
+gdm_WN_splines1 <- isplineExtract2(gdm_WN)$x %>%
+  as.data.frame() %>%
+  dplyr::select(elevation = elev.mean)
+
+gdm_WN_splines2 <- isplineExtract2(gdm_WN)$y %>%
+  as.data.frame() %>%
+  dplyr::select(response = elev.mean)
+
+gdm_WN_splines <- bind_cols(gdm_WN_splines1, gdm_WN_splines2) %>%
+  mutate(level = rep("Interactions"))
+
+# Flora model
+gdm_survey_splines1 <- isplineExtract2(gdm_survey)$x %>%
+  as.data.frame() %>%
+  dplyr::select(elevation = elev.mean)
+
+gdm_survey_splines2 <- isplineExtract2(gdm_survey)$y %>%
+  as.data.frame() %>%
+  dplyr::select(response = elev.mean)
+
+gdm_survey_splines <- bind_cols(gdm_survey_splines1, gdm_survey_splines2) %>%
+  mutate(level = rep("Flora"))
+
+# BB model
+gdm_BB_splines1 <- isplineExtract2(gdm_BB)$x %>%
+  as.data.frame() %>%
+  dplyr::select(elevation = elev.mean)
+
+gdm_BB_splines2 <- isplineExtract2(gdm_BB)$y %>%
+  as.data.frame() %>%
+  dplyr::select(response = elev.mean)
+
+gdm_BB_splines <- bind_cols(gdm_BB_splines1, gdm_BB_splines2) %>%
+  mutate(level = rep("Bumble bees"))
+
+# Combine
+gdm_splines <- bind_rows(gdm_WN_splines, gdm_survey_splines, gdm_BB_splines) %>%
+  mutate(Level = factor(level, levels = c("Interactions", "Flora", "Bumble bees"))) %>%
+  select(-level)
+
+tree_line <- site_data %>%
+  dplyr::select(elevation = elev.mean, tree_line)
+
+# Plot
+# gdm_plot_ <- ggplot(gdm_splines, aes(elevation, response, linetype = Level)) +
+#   geom_line(size = 0.75) +
+#   #geom_vline(xintercept = 1500) +
+#   geom_rug(data = tree_line, aes(elevation, color = tree_line), inherit.aes = FALSE) +
+#   labs(color = "Tree Line", x = "Elevation", y = "Conditional effect") +
+#   theme_classic(14) +
+#   guides(linetype = "none") +
+#   theme(legend.position = c(0.12, 0.85),
+#         plot.margin = margin(1, 3, 0.5, 0.5, "cm"),)
+# 
+# gdm_plot <- ggdraw() +
+#   draw_plot(gdm_plot_) +
+#   draw_image(BB, scale = 0.11, x = 0.34, y = -0.19) +
+#   draw_image(FL, scale = 0.15, x = 0.33, y = 0.15) +
+#   draw_image(BBxFL, scale = 0.18, x = 0.38, y = 0.41)
+
+#ggsave("../ms_1/figures/fig7.pdf", width = 7, height = 5)
+
+gdm_plot <- ggplot(gdm_splines, aes(elevation, response)) +
+  geom_line() +
+  geom_rug(data = tree_line, aes(elevation, color = tree_line), inherit.aes = FALSE) +
+  labs(color = "Tree Line", x = "Elevation", y = "Conditional effect") +
+  theme_light(12) +
+  guides(linetype = "none") +
+  facet_wrap(~Level, scales = "free", ncol = 1)
+
+ggsave("../ms_1/figures/fig7.pdf", gdm_plot, width = 4, height = 6.5)
+```
+
+# 6. MRM analysis (not included in paper)
 ```{r, warning=FALSE}
@@
              predictor == "mgmt.diff" ~ "management")) %>%
-    select(response, Predictor, b, pval) %>%
+    dplyr::select(response, Predictor, b, pval) %>%
     tibble()
@@
 
-### GDM: Interactions
-```{r message=FALSE, warning=FALSE}
-# Prep sitepairs
-sitepair_WN <- interaction_gdm_prep(net)
-
-# Run GDM analysis
-gdm_WN <- gdm(sitepair_WN, geo = FALSE)
-summary(gdm_WN)
-gdm.varImp_WN <- gdm.varImp(sitepair_WN, geo = FALSE)
-
-# Plot
-plot(gdm_WN)
-
-# Plot with confidence intervals 
-# pdf("../output/gdm_interactions.pdf", height = 2.5, width =  6.5)
-# plotUncertainty(sitepair.WN, sampleSites = 0.7, bsIters = 100, geo  = FALSE, 
-#                 splineCol = "black", plot.layout = c(1,3))
-# dev.off()
-```
-
-### GDM: Floral survey
-```{r}
-# Prep input for GDM
-bioData_survey <- biodata_prep_survey(survey)
-predData_survey <- predData_prep_survey(survey)
-sitepair_survey <- format_set(bioData_survey, predData_survey)
-
-# Call GDMs
-gdm_survey <- gdm(sitepair_survey, geo = FALSE)
-summary(gdm_survey)
-gdm.varImp_survey <- gdm.varImp(sitepair_survey, geo = FALSE)
-
-# Plot
-plot(gdm_survey)
-
-# Plot with confidence intervals (slow)
-# pdf("../output/gdm_flora.pdf", height = 2.5, width =  6.5)
-# plotUncertainty(sitepair_survey, sampleSites = 0.7, bsIters = 100, geo  = FALSE, 
-#                 splineCol = "#00BFC4", plot.layout = c(1,3))
-# dev.off()
-```
-
-### GDM: BB observations
-```{r}
-# Prep input for GDM
-bioData_BB <- biodata_prep_BB(net)
-predData_BB <- predData_prep_BB(net)
-sitepair_BB <- format_set(bioData_BB, predData_BB)
-
-# Call GDMs
-gdm_BB <- gdm(sitepair_BB, geo = FALSE)
-summary(gdm_BB)
-gdm.varImp_BB <- gdm.varImp(sitepair_BB, geo = FALSE)
-
-# Plot
-plot(gdm_BB)
-
-# Plot with confidence intervals (slow)
-# pdf("../output/gdm_BB.pdf", height = 2.5, width =  6.5)
-# plotUncertainty(sitepair_BB, sampleSites = 0.7, bsIters = 100, geo  = TRUE, 
-#                 splineCol = "#F8766D", plot.layout = c(1,3))
-# dev.off()
-```
-
-### Tabulate gdm.varImp results
-```{r}
-# Interaction model
-gdm.varImp_WN.tab2 <- gdm.varImp_WN[[2]] %>%
-  as.data.frame() %>%
-  select(Importance = 1)
-
-gdm.varImp_WN.tab3 <- gdm.varImp_WN[[3]] %>%
-  as.data.frame() %>%
-  select(pval = 1)
-
-gdm.varImp_WN.tab <- bind_cols(gdm.varImp_WN.tab2, gdm.varImp_WN.tab3) %>%
-  gdm_tab("Interactions")
-
-# Floral model
-gdm.varImp_survey.tab2 <- gdm.varImp_survey[[2]] %>%
-  as.data.frame() %>%
-  select(Importance = 1)
-
-gdm.varImp_survey.tab3 <- gdm.varImp_survey[[3]] %>%
-  as.data.frame() %>%
-  select(pval = 1)
-
-gdm.varImp_survey.tab <- bind_cols(gdm.varImp_survey.tab2, gdm.varImp_survey.tab3) %>%
-  gdm_tab("Flora")
-
-# BB model
-gdm.varImp_BB.tab2 <- gdm.varImp_BB[[2]] %>%
-  as.data.frame() %>%
-  select(Importance = 1)
-
-gdm.varImp_BB.tab3 <- gdm.varImp_BB[[3]] %>%
-  as.data.frame() %>%
-  select(pval = 1)
-
-gdm.varImp_BB.tab <- bind_cols(gdm.varImp_BB.tab2, gdm.varImp_BB.tab3) %>%
-  gdm_tab("Bumble bees")
-
-# Merge and format
-gdm.varImp.tab <- bind_rows(gdm.varImp_WN.tab, gdm.varImp_survey.tab, gdm.varImp_BB.tab) %>%
-  mutate(Importance = round(Importance, 2)) %>%
-  gt(groupname_col = "response") %>%
-  tab_style(
-    style = list(cell_text(weight = "bold")),
-    locations = list(cells_column_labels(everything()))
-  ) %>%
-  tab_style(
-    style = list(cell_text(style = "italic")),
-    locations = list(cells_row_groups(everything()))
-  ) %>%
-  data_color(
-    columns = vars(Importance),
-    colors = scales::col_numeric(
-      palette = c("white", "red"),
-      domain = c(-0, 56)), alpha = 0.75
-    ) %>%
-  tab_style(
-    style = list(
-      cell_text(weight = "bold")
-      ),
-    locations = cells_body(
-      columns = vars(pval),
-      rows = pval < 0.05
-    ))
-
-gtsave(gdm.varImp.tab, "../output/gdm_tab.pdf")
-```
-
-### Plot GDM splines
-```{r}
-# Interaction model
-gdm_WN_splines1 <- isplineExtract(gdm_WN)$x %>%
-  as.data.frame() %>%
-  select(elevation = elev.mean)
-
-gdm_WN_splines2 <- isplineExtract(gdm_WN)$y %>%
-  as.data.frame() %>%
-  select(response = elev.mean)
-
-gdm_WN_splines <- bind_cols(gdm_WN_splines1, gdm_WN_splines2) %>%
-  mutate(level = rep("Interactions"))
-
-# Flora model
-gdm_survey_splines1 <- isplineExtract(gdm_survey)$x %>%
-  as.data.frame() %>%
-  select(elevation = elev.mean)
-
-gdm_survey_splines2 <- isplineExtract(gdm_survey)$y %>%
-  as.data.frame() %>%
-  select(response = elev.mean)
-
-gdm_survey_splines <- bind_cols(gdm_survey_splines1, gdm_survey_splines2) %>%
-  mutate(level = rep("Flora"))
-
-# BB model
-gdm_BB_splines1 <- isplineExtract(gdm_BB)$x %>%
-  as.data.frame() %>%
-  select(elevation = elev.mean)
-
-gdm_BB_splines2 <- isplineExtract(gdm_BB)$y %>%
-  as.data.frame() %>%
-  select(response = elev.mean)
-
-gdm_BB_splines <- bind_cols(gdm_BB_splines1, gdm_BB_splines2) %>%
-  mutate(level = rep("Bumble bees"))
-
-# Combine
-gdm_splines <- bind_rows(gdm_WN_splines, gdm_survey_splines, gdm_BB_splines) %>%
-  mutate(Level = factor(level, levels = c("Interactions", "Flora", "Bumble bees")))
-
-tree_line <- site_data %>%
-  select(elevation = elev.mean, tree_line)
-
-# Plot
-ggplot(gdm_splines, aes(elevation, response, linetype = Level)) +
-  geom_line() +
-  #geom_vline(xintercept = 1500) +
-  geom_rug(data = tree_line, aes(elevation, color = tree_line), inherit.aes = FALSE) +
-  labs(color = "Tree Line", x = "Elevation", y = "Response") +
-  theme_light(14)
-
-ggsave("../output/gdm_splines.pdf")
-ggsave("../output/gdm_splines.png")
-
-```
+# Making plotUncertainty not suck
+```{r}
+test <- plotUncertainty3(sitepair_survey, 
+                sampleSites = 0.7, 
+                bsIters = 100, 
+                geo  = FALSE,
+                splines = c(6, 6, 6, 6, 6),
+                splineCol = "black", 
+                plot.layout = c(1,3))
+
+fullPlotX_list[p] <- tibble(fullPlotX)
+fullPlotY_list[p] <- tibble(fullPlotY)
+highBoundX_list[p] <- tibble(highBoundX)
+highBoundY_list[p] <- tibble(highBoundY)
+lowBoundX_list[p] <- tibble(lowBoundX)
+lowBoundY_list[p] <- tibble(lowBoundY)
+
+elev_tab <- tibble(test[[1]][[1]],
+                   test[[1]][[2]],
+                   test[[1]][[3]],
+                   test[[1]][[4]],
+                   test[[1]][[5]]) %>%
+  select(a = 1, b = 2, c = 3, d = 4, e = 5)
+
+ggplot(elev_tab, aes(a, b)) +
+  geom_line()
+
+ggplot(elev_tab, aes(a, c)) +
+  geom_line()
+
+ggplot(elev_tab, aes(a, d)) +
+  geom_line()
+
+ggplot(elev_tab, aes(a, e)) +
+  geom_line()
+
+ggplot(elev_tab, aes(b, c)) +
+  geom_line()
+
+##########################################################################
+plotUncertainty3 <- function(spTable, sampleSites, bsIters, geo=FALSE, splines=NULL, 
+                           knots=NULL, splineCol="blue", errCol="grey80", 
+                           plot.linewidth=2.0, plot.layout=c(2,2), parallel=FALSE,
+                           cores=2){
+  fullPlotX_list <- list()
+  fullPlotY_list <- list()
+  highBoundX_list <- list()
+  highBoundY_list <- list()
+  lowBoundX_list <- list()
+  lowBoundY_list <- list()
+
+  ##A function to plot the uncertantiy of each variable from a GDM model
+  #################
+  #spTable <- sitePairTab          ##the input site-pair table to subsample from
+  #sampleSites <- 0.9     ##fraction of sites that should be retained from site pair table 
+  #bsIters <- 50       ##the number of times the site-pair table should be sampled
+  #geo <- F              ##rather or not the gdm model takes geography into account, see gdm
+  #splines <- NULL       ##splines gdm setting, see gdm
+  #knots <- NULL         ##knots gdm setting, see gdm
+  #splineCol <- "blue"    ##color of the center line
+  #errCol <- "grey80"        ##color of the uncertainty polygon
+  #plot.linewidth <- 2.0    ##line width of the center line
+  #plot.layout <- c(3,3)    ##number of plots per page
+  #parallel <- F            ##rather or not the sampling should happen in parallel processing, to speed it up
+  #cores <- 6               ##number of cores to if parallel processing
+  #################
+  ##function breaks and warnings
+  ##makes sure that table is a properly formated site-pair table
+  if(!is(spTable, "gdmData")){
+    warning("The spTable object is not of class 'gdmData'. See the formatsitepair function for help.")
+  }
+  ##checks to makes sure table is a matrix or table frame
+  if(!(is(spTable, "gdmData") | is(spTable, "matrix") | is(spTable, "data.frame"))){
+    stop("The spTable object must be of class 'gdmData', 'matrix', or 'data.frame'.")
+  }
+
+  ##sanity check on the data table
+  if(ncol(spTable) < 6){
+    stop("spTable object requires at least 6 columns: Observed, weights, s1.xCoord, s1.yCoord, s2.xCoord, s2.yCoord")
+  }  
+  if(nrow(spTable) < 1){
+    stop("spTable object has < 1 rows.")
+  }
+
+  ##checks that geo has either TRUE or FALSE
+  if(!(geo==TRUE | geo==FALSE)){
+    stop("geo argument must be either TRUE or FALSE")
+  }
+  ##makes sure splines is a numeric vector
+  if(is.null(splines)==FALSE & !is(splines, "numeric")){
+    stop("splines object must of of class = 'numeric'.")
+  }
+  ##checks knots inputs
+  if(is.null(knots)==FALSE & !is(knots, "numeric")){
+    stop("knots object must of of class = 'numeric'.")
+  }
+
+  ##checks that parallel has either TRUE or FALSE
+  if(!(parallel==TRUE | parallel==FALSE)){
+    stop("parallel argument must be either TRUE or FALSE")
+  }
+  ##makes sure that cores has a value when parallel is true
+  if(parallel==TRUE & is.null(cores)==TRUE){
+    stop("If parallel==TRUE, the number of cores must be specified")
+  }
+  ##makes sure that cores is a positive integer 
+  if((is.null(cores)==FALSE & is.numeric(cores)==FALSE) | cores<1){
+    stop("argument cores needs to be a positive integer")
+  }
+
+  ##makes sure that bsIters is a positive integer 
+  if((is.null(bsIters)==FALSE & is.numeric(bsIters)==FALSE) | bsIters<1){
+    stop("argument bsIters needs to be a positive integer")
+  }
+  ##makes sure sampleSites is a number
+  if(is.numeric(sampleSites)==FALSE){
+    stop("sampleSites must be a number between 0 and 1")
+  }
+  ##makes sure that sampleSites is between 0 and 1
+  if(sampleSites < 0){
+    stop("sampleSites must be a number between 0 and 1")
+  }
+  if(sampleSites > 1){
+    stop("sampleSites must be a number between 0 and 1")
+  }
+  if(sampleSites==0){
+    stop("a sampleSites value of 0 will remove all sites from the analysis (bad).")
+  }
+  ##double makes sure these values are integers, seems to truncate if not
+  cores <- as.integer(cores)
+  bsIters <- as.integer(bsIters)
+
+  ##assign k to prevent issues to cran checking
+  k <- NULL
+
+  ##makes copies of the site-pair table in order to randomly subsample each one differently
+  lstSP <- lapply(1:bsIters, function(i){spTable})
+
+  ##runs parallel if desired by the users
+  if(parallel==TRUE){
+    ##loads libraries
+    #require(foreach)
+    #require(doParallel)
+    #requireNamespace("foreach")
+    #requireNamespace("parallel")
+    #requireNamespace("doParallel")
+
+    ##sets cores
+    cl <- makeCluster(cores, outfile="")
+    registerDoParallel(cl)
+    ##first removes a number of sites according to input
+    subSamps <- foreach(k=1:length(lstSP), .verbose=F, .packages=c("gdm")) %dopar%
+                      removeSitesFromSitePair(lstSP[[k]], sampleSites=sampleSites)
+    ##models the subsamples
+    gdmMods <- foreach(k=1:length(subSamps), .verbose=F, .packages=c("gdm")) %dopar%
+                     #gdmMods <- try(foreach(k=1, .verbose=F, .packages=c("gdm")) %dopar%
+                     gdm(subSamps[[k]], geo=geo, splines=splines, knots=knots)
+    stopCluster(cl)
+  }else{
+    ##first removes a number of sites according to input
+    subSamps <- lapply(lstSP, removeSitesFromSitePair, sampleSites=sampleSites)
+    ##models the subsamples
+    gdmMods <- lapply(subSamps, gdm, geo=geo, splines=splines, knots=knots)
+  }
+
+  ##models the full gdm
+  fullGDMmodel <- gdm(spTable, geo=geo, splines=splines, knots=knots)
+
+  ##Extracts the splines for each model
+  exUncertSplines <- lapply(gdmMods, isplineExtract2)
+  fullGDMsplines <- isplineExtract2(fullGDMmodel)
+
+  ##get the names of the predictor variables
+  predVars <- colnames(exUncertSplines[[1]][[1]])
+
+  ##establish what plot layout to use
+  thisplot <- 0
+  one_page_per_plot <- FALSE
+  if ((plot.layout[1]==1) && (plot.layout[2]==1)){
+    one_page_per_plot <- TRUE
+  }else{
+    par(mfrow=plot.layout)
+  }
+
+  ##sets the plotting minimum and maximum y-values 
+  totalYmin <- Inf
+  totalYmax <- -Inf
+
+  ##determines the bounds of the plots
+  for(p in 1:length(predVars)){
+    #p=1
+    predV <- predVars[p]
+
+    ##gets the minimum and maximum, to set the ploting extent
+    for(nm in 1:length(exUncertSplines)){
+      #nm=1
+      selPlot <- exUncertSplines[[nm]]
+
+      spYmax <- max(selPlot[[2]][,predV])
+      spYmin <- min(selPlot[[2]][,predV])
+
+      totalYmax <- max(c(totalYmax, spYmax))
+      totalYmin <- min(c(totalYmin, spYmin))
+    }
+  }
+
+  ##plots by variable
+  for(p in 1:length(predVars)){
+    #p <- 1
+    predV <- predVars[p]
+
+    ##sets the plotting minimum and maximum x-values
+    totalXmin <- Inf
+    totalXmax <- -Inf
+    ##gets the minimum and maximum, to set the ploting extent
+    for(nm in 1:length(exUncertSplines)){
+      #nm=1
+      selPlot <- exUncertSplines[[nm]]
+
+      spXmax <- max(selPlot[[1]][,predV])
+      spXmin <- min(selPlot[[1]][,predV])
+
+      if(spXmax > totalXmax){totalXmax = spXmax}
+      if(spXmin < totalXmin){totalXmin = spXmin}
+    }
+
+    ##checks to make sure that there is some significance, if there is the the data is plotted
+    if(totalYmax!=0){
+      ##add mean of subsets to plot
+      plotX <- NULL
+      plotY <- NULL
+      byVarMatX <- NULL
+      byVarMatY <- NULL
+      ##create matrices based on the variable and its x and y location for each model iteration
+      for(nn in 1:length(exUncertSplines)){
+        #nn=1
+        plotX[[nn]] <- exUncertSplines[[nn]][[1]]
+        plotY[[nn]] <- exUncertSplines[[nn]][[2]]
+        byVarMatY <- cbind(byVarMatY, plotY[[nn]][,predV])
+        byVarMatX <- cbind(byVarMatX, plotX[[nn]][,predV])
+      }
+      ##gets spline data from the full gdm model
+      fullPlotX <- fullGDMsplines[[1]]
+      fullPlotX <- fullPlotX[,predV]
+      fullPlotY <- fullGDMsplines[[2]]
+      fullPlotY <- fullPlotY[,predV]
+
+      ##calculates the confidence intervals for plotting
+      sdX <- apply(as.matrix(byVarMatX), 1, sd)
+      sdY <- apply(as.matrix(byVarMatY), 1, sd)
+      highBoundX <- fullPlotX + sdX
+      lowBoundX <- fullPlotX - sdX
+      highBoundY <- fullPlotY + sdY
+      lowBoundY <- fullPlotY - sdY
+
+      # outList[p] <- tibble(fullPlotX, fullPlotY,
+      #                      highBoundX, highBoundY,
+      #                      lowBoundX, lowBoundY)
+
+      fullPlotX_list[p] <- tibble(fullPlotX)
+      fullPlotY_list[p] <- tibble(fullPlotY)
+      highBoundX_list[p] <- tibble(highBoundX)
+      highBoundY_list[p] <- tibble(highBoundY)
+      lowBoundX_list[p] <- tibble(lowBoundX)
+      lowBoundY_list[p] <- tibble(lowBoundY)
+
+      # ##collects the data to be used in the rug plot
+      # if(predV=="Geographic"){
+      #   ##calculates unique eucildian distance between sites
+      #   rugData <- unique(sqrt(((spTable$s1.xCoord-spTable$s2.xCoord)^2)+((spTable$s1.yCoord-spTable$s2.yCoord)^2)))
+      # }else{
+      #   ##gets unique values of variable data
+      #   varDat <- grep(predV, colnames(spTable))
+      #   rugData <- unique(c(spTable[,c(varDat[1])], spTable[,c(varDat[2])]))
+      # }
+      # 
+      # ##plots one graph per page, unless specified otherwise
+      # if (one_page_per_plot){
+      #   dev.new()
+      #   dev.next()
+      # }else{
+      #   thisplot <- thisplot + 1
+      #   if(thisplot > (plot.layout[1] * plot.layout[2])){
+      #     thisplot <- 1
+      #     par(mfrow=plot.layout)
+      #   }
+      # }
+      # #settings <- par(pars)
+      # ##plots mean data line and polygon of uncertanty
+      # plot(NULL, xlim=c(totalXmin, totalXmax), ylim=c(totalYmin, totalYmax), xlab=predV, ylab="Partial Ecological Distance")
+      # polygon(c(lowBoundX, rev(highBoundX)), c(lowBoundY, rev(highBoundY)), col=errCol, border=NA)
+      # lines(fullPlotX, fullPlotY, col=splineCol, lwd=plot.linewidth)
+      # rug(rugData)
+    }
+  }
+  return(list(fullPlotX_list, fullPlotY_list,
+              highBoundX_list, highBoundY_list,
+              lowBoundX_list, lowBoundY_list))
+}
+##########################################################################
+```

### 2021-09-22 code/ms1_rich_abund.Rmd
analysis reboot

---
title: "Abundance/diversity GAM"
output: html_document
---

# 1. Prepare environment
## 1.1 Load packages
```{r setup, include=FALSE, echo=FALSE}
library(mgcv)
library(mgcViz)
library(magick)
library(patchwork)
library(tidymv)
library(vegan)
library(gratia)
library(tidyverse)
library(broom)
library(lubridate)
library(cowplot)
```

## 1.2 Load site data and floral taxonomy data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, 
                   temp.mean, elev.class, elev.class2)) %>% 
  mutate(site = factor(site),
         tree_line = factor(tree_line))

# FL traits
## family-level taxonomy
fl_tax <- read_csv("../data/processed_data/floral_tax.csv") 
```

## 1.3 Load network and survey data
```{r message=FALSE}
### BB-FL visitation data
net <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>% 
  dplyr::select(site, date, bb.sp, plant.sp) # drop unused columns

### Floral survey data
survey <- read_csv("../data/processed_data/floral_survey.csv") %>%
  # Consider only visited plant species (anytime anywhere, not specifically for a given site*date)
  semi_join(net, by = "plant.sp") %>% 
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  group_by(plant.sp) %>%
  mutate(present = case_when(
    max(flower.cover) > 0 ~ TRUE,
    max(flower.cover) == 0 ~ FALSE
  )) %>%
  filter(present == TRUE) # Ignore plants that never had any flower cover
```

## 1.4 Load images
```{r}
BB <- image_read("../ms_1/figures/phylopic_BB.png")
FL <- image_read("../ms_1/figures/phylopic_FL.png")
BBxFL <- image_read("../ms_1/figures/phylopic_BBxFL.png")
```

## 1.5 Define functions
```{r}
gam_tab1 <-function(x1, x2, x3, 
                    name1, name2, name3) {

  tab1.1 <- x1 %>%
    tidy(parametric = TRUE, conf.int = TRUE) %>%
    mutate(class = rep("parametric", n()),
           estimate = round(estimate, 3),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name1, n())
           )

  tab1.2 <- x1 %>%
    tidy(parametric = FALSE) %>%
    mutate(class = rep("smooth", n()),
           edf = round(edf, 2),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name1, n()))

  tab2.1 <- x2 %>%
    tidy(parametric = TRUE, conf.int = TRUE) %>%
    mutate(class = rep("parametric", n()),
           estimate = round(estimate, 3),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name2, n())
           )

  tab2.2 <- x2 %>%
    tidy(parametric = FALSE) %>%
    mutate(class = rep("smooth", n()),
           edf = round(edf, 2),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name2, n())
           )

  tab3.1 <- x3 %>%
    tidy(parametric = TRUE, conf.int = TRUE) %>%
    mutate(class = rep("parametric", n()),
           estimate = round(estimate, 3),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name3, n())
           )

  tab3.2 <- x3 %>%
    tidy(parametric = FALSE) %>%
    mutate(class = rep("smooth", n()),
           edf = round(edf, 2),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name3, n())
           )

  out <- bind_rows(tab1.1, tab1.2, tab2.1, tab2.2, tab3.1, tab3.2) %>%
    select(model, class, term, estimate, edf, statistic, p.value) %>%
    mutate(term = case_when(
      term == "(Intercept)" ~ "Intercept",
      term == "mgmtnone" ~ "mgmt:none",
      term == "mgmtmowing" ~ "mgmt:mowing",
      term == "year2011" ~ "year:2011",
      term == "year2012" ~ "year:2012",
      term == "s(site)" ~ "s(site)",
      term == "s(net.transects)" ~ "s(samples)",
      term == "s(survey.transects)" ~ "s(samples)",
      term == "s(elev.mean)" ~ "s(elevation)",
      term == "s(elev.mean):year2010" ~ "s(elevation:2010)",
      term == "s(elev.mean):year2011" ~ "s(elevation:2011)",
      term == "s(elev.mean):year2012" ~ "s(elevation:2012)"),
      sig = case_when(
        p.value < 0.001 ~ "***",
        p.value >= 0.001 & p.value < 0.01 ~ "**",
        p.value >= 0.01 & p.value < 0.05 ~ "*",
        p.value >= 0.05 ~ ""
      ),
      p.value = ifelse(p.value < 0.001, "<0.001", p.value),
      statistic = ifelse(statistic < 0.001, "<0.001", statistic),
      edf = ifelse(edf < 0.001, "<0.001", edf)) %>%
    flextable() %>%
    merge_v(j = c("model", "class")) %>%
    valign(j = c("model", "class"), valign = "top") %>%
    autofit()
}

gam_tab2 <-function(x1, x2, name1, name2) {

  tab1.1 <- x1 %>%
    tidy(parametric = TRUE, conf.int = TRUE) %>%
    mutate(class = rep("parametric", n()),
           estimate = round(estimate, 3),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name1, n())
           )

  tab1.2 <- x1 %>%
    tidy(parametric = FALSE) %>%
    mutate(class = rep("smooth", n()),
           edf = round(edf, 2),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name1, n()))

  tab2.1 <- x2 %>%
    tidy(parametric = TRUE, conf.int = TRUE) %>%
    mutate(class = rep("parametric", n()),
           estimate = round(estimate, 3),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name2, n())
           )

  tab2.2 <- x2 %>%
    tidy(parametric = FALSE) %>%
    mutate(class = rep("smooth", n()),
           edf = round(edf, 2),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name2, n())
           )

  out <- bind_rows(tab1.1, tab1.2, tab2.1, tab2.2) %>%
    select(model, class, term, estimate, edf, statistic, p.value) %>%
    mutate(term = case_when(
      term == "(Intercept)" ~ "Intercept",
      term == "mgmtnone" ~ "mgmt:none",
      term == "mgmtmowing" ~ "mgmt:mowing",
      term == "year2011" ~ "year:2011",
      term == "year2012" ~ "year:2012",
      term == "s(site)" ~ "s(site)",
      term == "s(net.transects)" ~ "s(samples)",
      term == "s(survey.transects)" ~ "s(samples)",
      term == "s(elev.mean)" ~ "s(elevation)",
      term == "s(elev.mean):year2010" ~ "s(elevation:2010)",
      term == "s(elev.mean):year2011" ~ "s(elevation:2011)",
      term == "s(elev.mean):year2012" ~ "s(elevation:2012)"),
      sig = case_when(
        p.value < 0.001 ~ "***",
        p.value >= 0.001 & p.value < 0.01 ~ "**",
        p.value >= 0.01 & p.value < 0.05 ~ "*",
        p.value >= 0.05 ~ ""
      ),
      p.value = ifelse(p.value < 0.001, "<0.001", p.value),
      statistic = ifelse(statistic < 0.001, "<0.001", statistic),
      edf = ifelse(edf < 0.001, "<0.001", edf)) %>%
    flextable() %>%
    merge_v(j = c("model", "class")) %>%
    valign(j = c("model", "class"), valign = "top") %>%
    autofit()
}
```

## 1.6 Identify plant species that were visited but not recorded in the survey
```{r message=FALSE}
vis_sp <- net %>%
  left_join(fl_tax) %>%
  select(plant.sp, plant.genus, plant.family) %>%
  unique()

surv_sp <- survey %>%
  left_join(fl_tax) %>%
  select(plant.sp, plant.genus, plant.family) %>%
  unique()

missing <- anti_join(vis_sp, surv_sp)
```

# 2. Process data
## 2.1 Calculate transect-wise FL and BB abundance and richness
```{r echo=FALSE, include=TRUE, message=FALSE}
# Survey transects per site
survey_transects_siteyear <- survey %>%
  mutate(year = factor(year(date)),
         site = factor(site)) %>%
  group_by(site, year) %>%
  summarize(survey.transects = length(unique(date)))

# BB transects per site
net_transects_siteyear <- net %>%
  mutate(year = factor(year(date)),
         site = factor(site)) %>%
  group_by(site, year) %>%
  summarize(net.transects = length(unique(date)))

# Per-transect BB abundance ~ sp
bb_abund_sp <- net %>%
  dplyr::select(site, date, bb.sp) %>%
  group_by(site, date, bb.sp) %>% # 
  summarize(bb.abund = n()) %>% # get per transect bb abundance
  ungroup() %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  semi_join(net, by = c("site", "date")) %>% # drop site-dates that did not exist in the data set
  left_join(site_data) %>% # add site data
  mutate(year = factor(year(date)),
         yday = yday(date),
         bb.sp = factor(bb.sp),
         mgmt = factor(management),
         site = factor(site),
         transect = factor(transect))

### Total BB abundance per transect
bb_abund <- bb_abund_sp %>%
  group_by(site, date) %>%
  summarize(bb.abund = sum(bb.abund)) %>%
  left_join(site_data) %>%
  mutate(site = factor(site),
         yday = yday(date),
         year = factor(year(date)),
         mgmt = factor(management)) %>%
  dplyr::select(-management)

### Per-transect floral abund by species
fl_abund_sp <- survey %>%
  group_by(site, date, plant.sp) %>% # group by plant.sp*transect
  summarize(fl.abund = sum(flower.cover)) %>% # get plant.sp*transect total flower cover; this step really shouldn't be required, but in a few cases, there is more than one abundance entry for the same species on a given transect; not sure how that happened, but it's in the original data; alternatively, I could drop the dups or average them. 
  group_by(site, date) %>%
  mutate(prop.fl.abund = fl.abund/sum(fl.abund)) %>% # get the proportional flower cover for each species.
  ungroup() %>%
  complete(site, date, plant.sp, fill = list(fl.abund = 0, prop.fl.abund = 0)) %>%
  semi_join(survey, by = c("site", "date")) %>%  # only include site-dates that actually occurred
  unique() %>%
  mutate(year = factor(year(date))) %>%
  left_join(fl_tax)

### Total floral abundance per transect
fl_abund <- fl_abund_sp %>%
  group_by(site, date) %>%
  summarize(fl.abund = sum(fl.abund)) %>%
  left_join(site_data) %>%
  mutate(site = factor(site),
         yday = yday(date),
         year = factor(year(date)),
         mgmt = factor(management)) %>%
  dplyr::select(-management) %>%
  mutate(year2010 = ifelse(year == "2010", 1, 0),
         year2011 = ifelse(year == "2011", 1, 0),
         year2012 = ifelse(year == "2012", 1, 0))
```

## 2.2 Calculate aggregate abundance and richness by site-year
```{r}
### Bumble bee richness per site-year
bb_rich_siteyear <- net %>%
  mutate(year = factor(year(date))) %>%
  group_by(site, year) %>%
  summarize(bb.rich = length(unique(bb.sp))) %>%
  left_join(site_data) %>% # add site data
  mutate(mgmt = factor(management),
         site = factor(site),
         transect = factor(transect)) %>%
  select(-management) %>%
  left_join(net_transects_siteyear)

### Peak BB abundance per site-year
bb_abund_siteyear <- bb_abund %>%
  group_by(site, year, elev.mean, mgmt) %>%
  summarize(bb.peakabund = max(bb.abund))

### Peak BB abundance by species per site-year
bb_abund_sp_siteyear <- bb_abund_sp %>%
  group_by(site, year, elev.mean, mgmt, bb.sp) %>%
  summarize(bb.peakabund = max(bb.abund))

### Floral richness per site-year
fl_rich_siteyear <- survey %>%
  mutate(year = factor(year(date))) %>%
  group_by(site, year) %>%
  summarize(fl.rich = length(unique(plant.sp))) %>%
  left_join(site_data) %>% # add site data
  mutate(mgmt = factor(management),
         site = factor(site),
         transect = factor(transect)) %>%
  select(-management) %>%
  left_join(survey_transects_siteyear)

### Peak floral abundance per site
fl_abund_siteyear <- fl_abund %>%
  group_by(site, year, elev.mean, mgmt) %>%
  summarize(fl.peakabund = max(fl.abund))

### Interaction richness per site
interaction_rich_siteyear <- net %>%
  mutate(year = factor(year(date))) %>%
  group_by(site, year) %>%
  unite(interaction, c(bb.sp, plant.sp), sep = "_") %>%
  summarize(interaction.rich = length(unique(interaction))) %>%
  left_join(site_data) %>% # add site data
  mutate(mgmt = factor(management),
         site = factor(site),
         transect = factor(transect)) %>%
  select(-management) %>%
  left_join(net_transects_siteyear)

### Get everything into a single data frame
rich_abund_siteyear <- bb_rich_siteyear %>% 
  full_join(fl_rich_siteyear) %>%
  full_join(bb_abund_siteyear) %>% 
  full_join(fl_abund_siteyear) %>%
  full_join(interaction_rich_siteyear)

### Long version for plotting
rich_abund_siteyear.long <- rich_abund_siteyear %>%
  pivot_longer(cols = c(bb.rich, fl.rich, bb.peakabund, 
                        fl.peakabund, interaction.rich)) %>%
  separate(name, c("group", "metric"), sep = "\\.") %>%
  group_by(group, metric) %>%
  mutate(z = scale(value))
```

# 3. Analyses

## 3.1 Preliminary visualization
```{r eval=FALSE, include=FALSE}
ggplot(rich_abund_siteyear.long, aes(elev.mean, z, color = group)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_grid(year~metric) +
  theme_classic()

ggplot(rich_abund_site.long, aes(z, fill = group)) +
  geom_histogram() +
  facet_wrap(~metric) +
  theme_classic()

# Exploring some bivariate relations

ggplot(rich_abund_site, aes(fl.rich, bb.rich)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(rich_abund_site, aes(fl.peakabund, fl.rich)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(rich_abund_site, aes(fl.peakabund, bb.peakabund)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(rich_abund_site, aes(fl.peakabund, bb.rich)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(rich_abund_site, aes(survey.transects, net.transects)) +
  geom_point() +
  geom_smooth(method = "lm") 
```

## 3.2 Model richness ~ elevation

Our interest in all these models is the global effect of elevation, but we need to account for the annual structure of our data. We do this by including year as a smoothing factor (interacting with elevation) that can deviate from the global effect of elevation. This represents the type GI model structure from Pedersen et al. (2019). What we will focus on is the "G" part of this model, i.e. the global elevation smoother, but I will include the year*elevation smooths in the Appendix.

It was a bit tricky to decide what distribution family to use for these models. The residuals are approximately gaussian, but richness data is count data, so theoretically we should use a zero-bounded distribution. Poisson would be a good start, but it is badly overdispersed in our models. Quasipoisson fixes that, and it gives results that are essentially identical to what we get with gaussian models. For some reason, a negative binomial model does not fix the dispersion problem.

The quasipoisson family also works well for the abundance models, and it can handle the continuous (i.e. non-integer) floral abundance data. So, we will stick with the quasipoisson family for all these models.

### 3.2.1 BB richness
```{r}
# Sanity check visualization
ggplot(bb_rich_siteyear, aes(elev.mean, bb.rich)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))  +
  facet_wrap(~year, ncol = 1) +
  theme_classic()

# Call model
gam1.1 <- gam(bb.rich ~ 
                mgmt +
                year +
                s(site, bs = "re") +
                s(net.transects) +
                s(elev.mean) +
                s(elev.mean, by = year, k = 3, m = 1),
              method = "REML",
              family = "quasipoisson",
              #select = TRUE,
              data = bb_rich_siteyear)

# Summary and model diagnostics
summary(gam1.1)
check.gamViz(getViz(gam1.1))

# Plot all conditional effects
draw(gam1.1, 
     residuals = TRUE) & theme_classic()

# Extract elevation effect
gam1.1_elev <- evaluate_smooth(gam1.1, smooth = "s(elev.mean)") %>%
  filter(is.na(by_variable)) %>% # we want only the global effect
  mutate(upper = est + (2*se),
         lower = est - (2*se))

# Extract year effect
gam1.1_year <- evaluate_parametric_term(gam1.1, term = "year") %>%
  mutate(upper = partial + (2*se),
         lower = partial - (2*se))

# Extract managment effect
gam1.1_mgmt <- evaluate_parametric_term(gam1.1, term = "mgmt") %>%
  mutate(upper = partial + (2*se),
         lower = partial - (2*se))

# Plot elevation effect
gam1.1_plot1 <- ggplot(gam1.1_elev, aes(elev.mean, est)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), # Uncertainty
              fill = "black", alpha = 0.2) +
  # geom_rug(data = semi_join(site_data, net), 
  #          aes(elev.mean, color = tree_line), 
  #          inherit.aes = FALSE) +
  theme_classic() +
  labs(y = "Conditional effect", x = "Elevation [m]", color = "Tree line")

# Plot year effect
gam1.1_plot2 <- ggplot(gam1.1_year, aes(value, partial)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  geom_point() +
  theme_classic() +
  labs(y = NULL, x = NULL)

# Plot management effect
gam1.1_plot3 <- ggplot(gam1.1_mgmt, aes(value, partial)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  geom_point() +
  theme_classic() +
  labs(y = NULL, x = NULL)

# Assemble
gam1.1_plot <- (gam1.1_plot1 | (gam1.1_plot2 / gam1.1_plot3)) 
```

### 3.2.2 FL richness
```{r}
# Sanity check visualization
ggplot(fl_rich_siteyear, aes(elev.mean, fl.rich)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))  +
  facet_wrap(~year, ncol = 1) +
  theme_classic()

# Call model
gam2.1 <- gam(fl.rich ~ 
                mgmt +
                year +
                s(site, bs = "re") +
                s(survey.transects) +
                s(elev.mean) +
                s(elev.mean, by = year, k = 3, m = 1),
              method = "REML",
              family = "quasipoisson",
              #select = TRUE,
              data = fl_rich_siteyear)

# Summary and model diagnostics
summary(gam2.1)
check.gamViz(getViz(gam2.1))

# Plot all conditional effects
draw(gam2.1, 
     residuals = TRUE) & theme_classic()

# Extract elevation effect
gam2.1_elev <- evaluate_smooth(gam2.1, smooth = "elev") %>%
  filter(is.na(by_variable)) %>% # we want only the global effect
  mutate(upper = est + (2*se),
         lower = est - (2*se))

# Extract year effect
gam2.1_year <- evaluate_parametric_term(gam2.1, term = "year") %>%
  mutate(upper = partial + (2*se),
         lower = partial - (2*se))

# Extract managment effect
gam2.1_mgmt <- evaluate_parametric_term(gam2.1, term = "mgmt") %>%
  mutate(upper = partial + (2*se),
         lower = partial - (2*se))

# Plot elevation effect
gam2.1_plot1 <- ggplot(gam2.1_elev, aes(elev.mean, est)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), # Uncertainty
              fill = "black", alpha = 0.2) +
  # geom_rug(data = semi_join(site_data, net), 
  #          aes(elev.mean, color = tree_line), 
  #          inherit.aes = FALSE) +
  theme_classic() +
  labs(y = "Conditional effect", x = "Elevation [m]", color = "Tree line")

# Plot year effect
gam2.1_plot2 <- ggplot(gam2.1_year, aes(value, partial)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  geom_point() +
  theme_classic() +
  labs(y = NULL, x = NULL)

# Plot management effect
gam2.1_plot3 <- ggplot(gam2.1_mgmt, aes(value, partial)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  geom_point() +
  theme_classic() +
  labs(y = NULL, x = NULL)

# Assemble
gam2.1_plot <- gam2.1_plot1 | (gam2.1_plot2 / gam2.1_plot3)
```

### 3.2.3 Interaction richness
```{r}
# Sanity check visualization
ggplot(interaction_rich_siteyear, aes(elev.mean, interaction.rich)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))  +
  facet_wrap(~year, ncol = 1) +
  theme_classic()

# Call model
gam3.1 <- gam(interaction.rich ~ 
                mgmt +
                year +
                s(site, bs = "re") +
                s(net.transects) +
                s(elev.mean) +
                s(elev.mean, by = year, k = 3, m = 1),
              method = "REML",
              family = "quasipoisson",
              #select = TRUE,
              data = interaction_rich_siteyear)

# Summary and model diagnostics
summary(gam3.1)
check.gamViz(getViz(gam3.1))

# Plot all conditional effects
draw(gam3.1, 
     residuals = TRUE) & theme_classic()

# Extract elevation effect
gam3.1_elev <- evaluate_smooth(gam3.1, smooth = "s(elev.mean)") %>%
  filter(is.na(by_variable)) %>% # we want only the global effect
  mutate(upper = est + (2*se),
         lower = est - (2*se))

# Extract year effect
gam3.1_year <- evaluate_parametric_term(gam3.1, term = "year") %>%
  mutate(upper = partial + (2*se),
         lower = partial - (2*se))

# Extract managment effect
gam3.1_mgmt <- evaluate_parametric_term(gam3.1, term = "mgmt") %>%
  mutate(upper = partial + (2*se),
         lower = partial - (2*se))

# Plot elevation effect
gam3.1_plot1 <- ggplot(gam3.1_elev, aes(elev.mean, est)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), # Uncertainty
              fill = "black", alpha = 0.2) +
  # geom_rug(data = semi_join(site_data, net), 
  #          aes(elev.mean, color = tree_line), 
  #          inherit.aes = FALSE) +
  theme_classic() +
  labs(y = "Conditional effect", x = "Elevation [m]", color = "Tree line")

# Plot year effect
gam3.1_plot2 <- ggplot(gam3.1_year, aes(value, partial)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  geom_point() +
  theme_classic() +
  labs(y = NULL, x = NULL)

# Plot management effect
gam3.1_plot3 <- ggplot(gam3.1_mgmt, aes(value, partial)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  geom_point() +
  theme_classic() +
  labs(y = NULL, x = NULL)

# Assemble
gam3.1_plot <- gam3.1_plot1 | (gam3.1_plot2 / gam3.1_plot3)

# Tabulate
library(flextable)
as_flextable(gam3.1)
```

### 3.1.4 Compile plots
```{r}
gam_plot_richness_ <- ( gam1.1_plot / gam2.1_plot / gam3.1_plot ) +
  #plot_annotation(tag_levels = 'A') +
  plot_layout(nrow = 3)

gam_plot_richness <- ggdraw() +
  draw_plot(gam_plot_richness_) +
  draw_image(BB, scale = 0.15, x = -0.18, y = 0.28) +
  draw_image(FL, scale = 0.125, x = -0.18, y = -0.025) +
  draw_image(BBxFL, scale = 0.25, x = -0.18, y = -0.38)

ggsave("../ms_1/figures/fig3.pdf", gam_plot_richness, width = 6.5, height = 8.5)
```

## 3.3 Model peak abundance ~ elevation
We will use the same model structure that we used for the richness models.

### 3.3.1 BB abundance
```{r}
ggplot(bb_abund_siteyear, aes(elev.mean, bb.peakabund)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))  +
  facet_wrap(~year, ncol = 1) +
  theme_classic()

# Call model
gam4.1 <- gam(bb.peakabund ~ 
                mgmt +
                year +
                s(site, bs = "re") +
                s(elev.mean) +
                s(elev.mean, by = year, k = 3, m = 2),
              method = "REML",
              family = "quasipoisson",
              data = bb_abund_siteyear)

# Summary and model diagnostics
summary(gam4.1)
check.gamViz(getViz(gam4.1))

# Plot all conditional effects
draw(gam4.1, 
     residuals = TRUE) & theme_classic()

# Extract elevation effect
gam4.1_elev <- evaluate_smooth(gam4.1, smooth = "s(elev.mean)") %>%
  filter(is.na(by_variable)) %>% # we want only the global effect
  mutate(upper = est + (2*se),
         lower = est - (2*se))

# Extract year effect
gam4.1_year <- evaluate_parametric_term(gam4.1, term = "year") %>%
  mutate(upper = partial + (2*se),
         lower = partial - (2*se))

# Extract managment effect
gam4.1_mgmt <- evaluate_parametric_term(gam4.1, term = "mgmt") %>%
  mutate(upper = partial + (2*se),
         lower = partial - (2*se))

# Plot elevation effect
gam4.1_plot1 <- ggplot(gam4.1_elev, aes(elev.mean, est)) +
  geom_line(linetype = "dashed") +
  geom_ribbon(aes(ymin = lower, ymax = upper), # Uncertainty
              fill = "black", alpha = 0.2) +
  # geom_rug(data = semi_join(site_data, net), 
  #          aes(elev.mean, color = tree_line), 
  #          inherit.aes = FALSE) +
  theme_classic() +
  labs(y = "Conditional effect", x = "Elevation [m]", color = "Tree line")

# Plot year effect
gam4.1_plot2 <- ggplot(gam4.1_year, aes(value, partial)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  geom_point() +
  theme_classic() +
  labs(y = NULL, x = NULL)

# Plot management effect
gam4.1_plot3 <- ggplot(gam4.1_mgmt, aes(value, partial)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  geom_point() +
  theme_classic() +
  labs(y = NULL, x = NULL)

# Assemble
gam4.1_plot <- gam4.1_plot1 | (gam4.1_plot2 / gam4.1_plot3)
```

### 3.3.2 FL abundance
```{r}
ggplot(fl_abund_siteyear, aes(elev.mean, fl.peakabund)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))  +
  facet_wrap(~year, ncol = 1) +
  theme_classic()

# Call model
gam5.1 <- gam(fl.peakabund ~ 
                mgmt +
                year +
                s(site, bs = "re") +
                s(elev.mean, k = 5) +
                s(elev.mean, by = year, k = 3, m = 2),
              method = "REML",
              family = "quasipoisson",
              data = fl_abund_siteyear)

# Summary and model diagnostics
summary(gam5.1)
check.gamViz(getViz(gam5.1))

# Plot all conditional effects
draw(gam5.1, 
     residuals = TRUE) & theme_classic()

# Extract elevation effect
gam5.1_elev <- evaluate_smooth(gam5.1, smooth = "s(elev.mean)") %>%
  filter(is.na(by_variable)) %>% # we want only the global effect
  mutate(upper = est + (2*se),
         lower = est - (2*se))

# Extract year effect
gam5.1_year <- evaluate_parametric_term(gam5.1, term = "year") %>%
  mutate(upper = partial + (2*se),
         lower = partial - (2*se))

# Extract managment effect
gam5.1_mgmt <- evaluate_parametric_term(gam5.1, term = "mgmt") %>%
  mutate(upper = partial + (2*se),
         lower = partial - (2*se))

# Plot elevation effect
gam5.1_plot1 <- ggplot(gam5.1_elev, aes(elev.mean, est)) +
  geom_line(linetype = "solid") +
  geom_ribbon(aes(ymin = lower, ymax = upper), # Uncertainty
              fill = "black", alpha = 0.2) +
  # geom_rug(data = semi_join(site_data, net), 
  #          aes(elev.mean, color = tree_line), 
  #          inherit.aes = FALSE) +
  theme_classic() +
  labs(y = "Conditional effect", x = "Elevation [m]", color = "Tree line")

# Plot year effect
gam5.1_plot2 <- ggplot(gam5.1_year, aes(value, partial)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  geom_point() +
  theme_classic() +
  labs(y = NULL, x = NULL)

# Plot management effect
gam5.1_plot3 <- ggplot(gam5.1_mgmt, aes(value, partial)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  geom_point() +
  theme_classic() +
  labs(y = NULL, x = NULL)

# Assemble
gam5.1_plot <- gam5.1_plot1 | (gam5.1_plot2 / gam5.1_plot3)
```

### 3.3.3 Compile plots
```{r}
gam_plot_abundance_ <- ( gam4.1_plot / gam5.1_plot) +
  #plot_annotation(tag_levels = 'A') +
  plot_layout(nrow = 2)

gam_plot_abundance <- ggdraw() +
  draw_plot(gam_plot_abundance_) +
  draw_image(BB, scale = 0.15, x = -0.18, y = 0.16) +
  draw_image(FL, scale = 0.2, x = -0.18, y = -0.3)

ggsave("../ms_1/figures/fig4.pdf", gam_plot_abundance, width = 6.5, height = 5.5)
```

### 3.3.4 BB sp-level abundance ~ elevation
For this model, the negative-binomial family works best. Maybe it handles zeros better than the quasipoisson.
```{r}
ggplot(bb_abund_sp_siteyear, aes(elev.mean, bb.peakabund, color = year)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp, scales = "free")

gam6.1 <- gam(bb.peakabund ~ 
                year +
                mgmt +
                s(site, bs = "re") +
                s(bb.sp, bs = "re") +
                s(elev.mean, by = bb.sp),
              method = "REML",
              family = "nb",
              select = FALSE, # I want to visualize even the non-significant smooths
              data = bb_abund_sp_siteyear)

# Summary and model diagnostics
summary(gam6.1)
check.gamViz(getViz(gam6.1))

# Plot all conditional effects
draw(gam6.1, 
     residuals = FALSE) & theme_classic()

# Extract p-values
gam6.1_pval <- gam6.1 %>%
  tidy()

# Extract elevation effect
gam6.1_elev <- evaluate_smooth(gam6.1, smooth = "s(elev.mean)") %>%
  mutate(upper = est + (2*se),
         lower = est - (2*se)) %>%
  group_by(bb.sp) %>%
  mutate(diff = last(est) - first(est)) %>%
  left_join(gam6.1_pval, by = c("smooth" = "term")) %>%
  mutate(sig = factor(ifelse(p.value < 0.05, "p < 0.05", "ns")))

# Plot elevation effect
gam6.1_plot1 <- ggplot(gam6.1_elev, aes(elev.mean, est)) +
  geom_line(aes(linetype = fct_rev(sig))) +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = diff), # Uncertainty
              alpha = 0.5) +
  scale_fill_gradient2(low = "#4dac26", mid = "gray75", high = "#d01c8b") +
  theme_light(12) +
  labs(y = "Conditional effect", x = "Elevation [m]", fill = "Slope",
       linetype = "Significance") +
  facet_wrap(~reorder(bb.sp, -diff)) +
  theme()

ggsave("../ms_1/figures/fig5.pdf", gam6.1_plot1, height = 6.5, width = 6.5)
```

### 3.3.5 Compile tabulated summaries
```{r}
gam_tab.rich <- gam_tab1(x1 = gam1.1, x2 = gam2.1, x3 = gam3.1, 
                   name1 = "Bumble bee richness", 
                   name2 = "Floral richness", 
                   name3 = "Interaction richness")

save_as_image(gam_tab.rich, "../ms_1/figures/table1.png")

gam_tab.abund <- gam_tab2(x1 = gam4.1, x2 = gam5.1,
                   name1 = "Bumble bee abundance", 
                   name2 = "Floral abundance")

save_as_image(gam_tab.abund, "../ms_1/figures/table2.png")

gam_tab.abund_sp1 <- gam6.1  %>%
  tidy(parametric = TRUE) %>%
  mutate(class = rep("parametric", n()),
         estimate = round(estimate, 3),
         statistic = round(statistic, 3),
         p.value = round(p.value, 3)
         )

gam_tab.abund_sp2 <- gam6.1 %>%
  tidy(parametric = FALSE) %>%
  mutate(class = rep("smooth", n()),
         edf = round(edf, 2),
         statistic = round(statistic, 3),
         p.value = round(p.value, 3))

gam_tab.abund_sp <- bind_rows(gam_tab.abund_sp1, gam_tab.abund_sp2) %>%
    select(class, term, estimate, edf, statistic, p.value) %>%
    mutate(term = case_when(
      term == "(Intercept)" ~ "Intercept",
      term == "mgmtnone" ~ "mgmt:none",
      term == "mgmtmowing" ~ "mgmt:mowing",
      term == "year2011" ~ "year:2011",
      term == "year2012" ~ "year:2012",
      term == "s(site)" ~ "s(site)",
      term == "s(bb.sp)" ~ "s(species)",
      term == "s(elev.mean):bb.spgers" ~ "s(elevation:gers)",
      term == "s(elev.mean):bb.sphort" ~ "s(elevation:hort)",
      term == "s(elev.mean):bb.sphumi" ~ "s(elevation:humi)",
      term == "s(elev.mean):bb.sphypn" ~ "s(elevation:hypn)",
      term == "s(elev.mean):bb.spjone" ~ "s(elevation:jone)",
      term == "s(elev.mean):bb.splapi" ~ "s(elevation:lapi)",
      term == "s(elev.mean):bb.spmend" ~ "s(elevation:mend)",
      term == "s(elev.mean):bb.spmont" ~ "s(elevation:mont)",
      term == "s(elev.mean):bb.spmuci" ~ "s(elevation:muci)",
      term == "s(elev.mean):bb.sppasc" ~ "s(elevation:pasc)",
      term == "s(elev.mean):bb.spprat" ~ "s(elevation:prat)",
      term == "s(elev.mean):bb.sppsit" ~ "s(elevation:psit)",
      term == "s(elev.mean):bb.sppyre" ~ "s(elevation:pyre)",
      term == "s(elev.mean):bb.spsoro" ~ "s(elevation:soro)",
      term == "s(elev.mean):bb.sptelu" ~ "s(elevation:telu)",
      term == "s(elev.mean):bb.spwurf" ~ "s(elevation:wurf)"),
      sig = case_when(
        p.value < 0.001 ~ "***",
        p.value >= 0.001 & p.value < 0.01 ~ "**",
        p.value >= 0.01 & p.value < 0.05 ~ "*",
        p.value >= 0.05 ~ ""
      ),
      p.value = ifelse(p.value < 0.001, "<0.001", p.value),
      statistic = ifelse(statistic < 0.001, "<0.001", statistic),
      edf = ifelse(edf < 0.001, "<0.001", edf)) %>%
    flextable() %>%
    merge_v(j = "class") %>%
    valign(j = "class", valign = "top") %>%
    autofit()

save_as_image(gam_tab.abund_sp, "../ms_1/figures/table3.png")
```

### 3.3.6 BB elevation distribution
```{r}
bb_range <- net %>%
  left_join(site_data) %>%
  group_by(site, date, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = max(abund)) %>%
  filter(abund > 0) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))

ggplot(bb_range, aes(reorder(bb.sp, elev.med), elev.mean)) +
  geom_line(size = 2, alpha = 0.25) +
  geom_point(aes(size = abund), alpha = 0.5) +
  theme_classic(14) +
  scale_size_continuous(name = "Peak abundance",
                        breaks = c(1, 10, 20, 40, 80)) +
  scale_color_discrete(name = "Tree line") +
  xlab("Bumble bee species") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave("../output/bb_range_ms1.pdf")
ggsave("../output/bb_range_ms1.png")

# Compare bumble bee species in terms of proportional abundance (unit = site)
bb_abund_sum <- bb_abund_sp %>%
  group_by(site, bb.sp) %>%
  summarize(cum.abund = sum(bb.abund),
            mean.abund = mean(bb.abund)) %>%
  group_by(site) %>%
  mutate(prop.abund = cum.abund/sum(cum.abund)) %>%
  group_by(bb.sp) %>%
  mutate(mean.prop.abund = mean(prop.abund))

ggplot(bb_abund_sum, aes(reorder(bb.sp, -mean.prop.abund), prop.abund)) +
  geom_boxplot(fill = "gray85") +
  theme_classic() +
  ylab("Proportional abundance") +
  xlab("Bumble bee species")

ggsave("../output/bb_abundsum.pdf")
ggsave("../output/bb_abundsum.png")
```

## 3.4 Model sample-wise abundance ~ elevation * time
### 3.4.1 BB abundance
```{r}
gam7.1 <- gam(bb.abund ~ 
                year +
                mgmt +
                s(site, bs = "re") +
                te(yday, elev.mean, bs=c("gp","tp")) +
                te(yday, elev.mean, by = year, bs = c("tp", "tp"), m = 1),
              family = "nb",
              method = "REML",
              select = TRUE,
              data = bb_abund)

check.gamViz(getViz(gam7.1))
summary(gam7.1)

draw(gam7.1, 
     parametric = TRUE, 
     residuals = TRUE) & theme_classic()
```

### 3.4.2 FL abundance
```{r}
gam8.1 <- gam(fl.abund ~ 
                year +
                mgmt +
                s(site, bs = "re") +
                te(yday, elev.mean, bs=c("gp","tp")) +
                te(yday, elev.mean, by = year, bs = c("tp", "tp"), m = 1),
              family = "quasipoisson",
              method = "REML",
              select = TRUE,
              data = fl_abund)

check.gamViz(getViz(gam8.1))
summary(gam8.1)

draw(gam8.1, 
     parametric = TRUE, 
     residuals = TRUE) & theme_classic()
```

### 2021-09-22 code/ms1_sampling.Rmd
analysis reboot

@@
 site_data <- read_csv("../data/processed_data/site_data.csv") %>%
+  semi_join(visitation_sampling) %>%
   dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean, elev.class, elev.class2)) %>% 
@@
   ylab("Total bumble bees\nobserved") +
-  theme_light(11) +
-  theme(axis.title.x = element_blank())
+  theme_light(13) +
+  theme(axis.title.x = element_blank(),
+        panel.grid.minor = element_blank(),
+        panel.grid.major = element_blank())
 
@@
   ylab("Total samples") +
-  theme_light(11) +
-  theme(axis.title.x = element_blank())
+  theme_light(13) +
+  theme(axis.title.x = element_blank(),
+        panel.grid.minor = element_blank(),
+        panel.grid.major = element_blank())
 
@@
   geom_boxplot(fill = NA) +
-  #geom_density(outline.type = "full", position = "fill") +
   ylab("Bumble bees\nobserved / sample") +
-  #xlab(NULL) +
-  theme_light(11) +
-  theme(axis.title.x = element_blank())
+  theme_light(13) +
+  theme(axis.title.x = element_blank(),
+        panel.grid.minor = element_blank(),
+        panel.grid.major = element_blank())
 
-samplingplot4 <- ggplot(bb_obs_per_sample, aes(as_date(yday(date)), elev.mean/10)) +
+samplingplot4 <- ggplot(bb_obs_per_sample, aes(as_date(yday(date)), elev.mean)) +
   geom_point(size = 1, aes(alpha = obs)) +
   facet_wrap(~year) +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
   scale_alpha_continuous(name = "Bumble bees\nobserved") +
@@
                                                     "1970-07-01", "1970-08-01", "1970-09-01"))) +
-  theme_light(11) +
+  theme_light(13) +
   labs(x = NULL,
-       y = "Elevation (m x 10)")
+       y = "Elevation [m]") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        panel.grid.minor = element_blank())
 
@@
 
-ggsave("../ms_1/figures/samplingplot.pdf", height = 6, width = 10.5)
-ggsave("../ms_1/figures/samplingplot.png", height = 6, width = 10.5)
+ggsave("../ms_1/figures/fig2.pdf", samplingplot, height = 6, width = 10.5)
 ```

### 2021-09-22 ms_1/ms_revision1.Rmd
analysis reboot

@@
 
-1. Biological responses to environmental gradients
-
 Environmental gradients regulate the assembly of biological communities, and they have likely played a central role in generating and maintaining Earth's biological diversity [@Hutchinson1953-mo].
@@
 @Beals1969-in].
-Such insight is important for anticipating and managing the effects of global ecological change that modify existing environmental gradients together with the species that inhabit them.
+In addition to its basic significance, such insight is important for anticipating and managing the effects of climate and land use change, which can modify existing environmental gradients and reorganize the biological communities that inhabit them.
 
-2. The particular case of mountain elevation gradients
+Mountain slopes generate steep environmental gradients over short geographical distances, facilitating the assembly of biological communities with high species richness and turnover [@Rahbek2019-sw].
+Perhaps the most salient environmental condition that changes along a mountain elevation gradient is air temperature, which declines according to a linear lapse rate of approximately -0.6°C per 100 m [@Rolland2003-my], but mountain slopes also generate variation in other conditions, including precipitation, atmospheric pressure, atmospheric turbulence, substrate characteristics, and exposure to solar radiation [@Korner1995-mh; @Hodkinson2005-af].
 
-Mountain slopes generate steep environmental gradients over short geographical distances, facilitating the assembly of biological communities with high species richness and turnover[@Rahbek2019-sw].
-The response of species richness and turnover to mountain elevation gradients
+Ecologists have long viewed mountain slopes as natural laboratories for the study of community assembly, and particular attention has been given to elevational patterns of species richness and composition.
+Species richness along mountain slopes tends to exhibit either a mid-elevation peak or a monotonic decline [@Rahbek1995-oj; @Guo2013-tu], though the exact shape of these responses can vary considerably across taxonomic groups [e.g.
+@Peters2016-hf].
+Species composition along mountain slopes, best studied in plant communities, can exhibit a range of patterns from smooth gradations to relatively abrupt transitions, with steeper slopes tending to exhibit more discrete patterns of species turnover [@Beals1969-in].
+For example, many mountains have a well-defined tree line ecotone, where plant communities exhibit accelerated species turnover and declining diversity as multiple abiotic conditions become more extreme away from the shelter of tree canopy [@Slatyer1992-ir; @Becker2007-cg; @Descombes2017-id; though see @Odland1999-vx].
 
-*Species richness*
-
-Species richness along mountain slopes tends to show either a mid-elevation peak or a monotonic decline [@Rahbek1995-oj; @Guo2013-tu], though responses can vary considerably across taxonomic groups [e.g.
-@Peters2016-hf].
-
-*Species turnover*
-
-In some cases, continuous elevation gradients can generate punctuated shifts in species composition. as has been demonstrated in some mountain plant communities [@Beals1969-in; @Hemp2006-bw]
-
-Something about productivity ~ elevation
-
-Maybe something about abundance ~ elevation
-
-<!-- Vegetational Change Along Altitudinal Gradients; controversy regarding whether vegetation gradients are continuous or discrete, suggests something in between, consistent with the "threshold" that I see associated with the tree line.
-[@Beals1969-in] -->
-
-<!-- *A "though see" citation:* The ecological and floristic importance of the forest limit in phytogeographical zonations has remained un- questioned,   even though its  value as an important ecological boundary has  rarely been verified.
-An interesting question is therefore  whether  the  presence or absence of birch  forests in Aurland is associated with any abrupt discontinuity in floristic richness or composition ...
-Our results from Aurland suggest that the forest limit is  not associated with  any major floristic change,  either in species turnover or in species richness, but is  associated with some  change in overall species composition.
-Hofgaard (1997) similarly found little difference in beta diversity or floristic composition 300 m above or below the tree line in the central Norwegian mountains.
+<!-- In most cases, plant species composition along a mountain slope varies more or less continuously but with nonlinearities or thresholds of species turnover, suggesting an interaction of processes more complex than a simple response to the linear decline of temperature with elevation [@Beals1969-in].
 -->
 
-3. From species to interactions: less studied, but amenable to the same tools used at the species-level [see references in @Simanonok2014-ou]
+<!-- The peaking of richness at mid-elevations might reflect intermediate environmental conditions that favor species richness, but this pattern has also been attributed to the *mid-domain effect*, which predicts mid-elevation peaks as a geometric effect of bounded ranges independent of environmental conditions [@Colwell2000-qf].
+Similarly, monotonic declines in richness with elevation could be explained by abiotic conditions becoming more stringent with elevation, but they are also consistent with a species-area effect driven by the geometric loss of surface area as a mountain's topography converges to a summit [@Rahbek1997-hj; though see @Elsen2015-gp].
+-->
 
-While elevational patterns in gross summaries of ecological communities (e.g.
-species richness, net primary productivity) are valuable, potentially deeper insight can be gained by the more focused study of elevational patterns among guilds of species with known functional relationships, such as plants and herbivores or plants and pollinators.
+<!-- When discrete transitions in species composition occur along a continuous environmental gradient, they can be interpreted as points where the fitness inequality between competing species reverses, implying that interspecific competition is a primary mechanism of community assembly [@Hutchinson1953-mo].
+-->
 
-An elevational study of plants and insect herbivores [@Rasmann2014-oc]
+While elevational patterns in overall species richness and composition are informative, the functional significance of mountain slopes can be explored more deeply by focusing on the responses of interacting guilds of species.
+For example, studies of plants and insect herbivores along mountain elevation gradients have revealed a positive correlation between elevation and herbivore niche-breadth, suggesting an extension of the classical *latitude-niche-breadth hypothesis* to the analogous case of elevational gradients [@Rasmann2018-gl].
+Similarly, plant-pollinator interactions along mountain slopes have been found to become less specialized as elevation increases, which may reflect a reciprocal relationship between the processes of competition and environmental filtering in community assembly [@Hoiss2015-gv; @Classen2020-nh].
+A unifying theme of such studies is the understanding of *species interactions* as ecological entities *per se*, dependent upon but not reducible to the species-level richness and composition of the interacting guilds.
 
-4. Mountain bumble bees and wildflowers
-  A. Opportunity to study responses not just at the level of species, but at the level of interactions [@Simanonok2014-ou; @Burkle2016-tq].
-  B. Significance of understanding the elevation structure of mountain BB-wildflower communities vis-a-vis climate change and meadow reforestation
+Bumble bees (Hymenoptera: *Bombus* spp.) and their floral hosts provide an excellent system in which to study elevational patterns at both the species- and interaction-level.
+Bumble bees are believed to have arisen in the mountains of Asia during the cooling climate of the Eocene-Oligocene boundary, and their subsequent spread and diversification along mountain corridors and into lowland habitats was likely driven by alternating range expansion during cooling periods and retreat to higher elevations during warming periods [@Hines2008-er; @Stewart2010-jn; @Martinet2018-rb].
+Today, they remain characteristically cold-adapted organisms [@Heinrich1994-cv], with peak abundance and diversity in mountain ranges and northern latitudes [@Williams1998-xi].
+In these habitats, they are often the dominant guild of flower-visiting insects and the principal pollinators of entomophilous flora [@Goulson2010-lo].
+As in the past, so now are mountain habitats critical theaters of ecological change for bumble bees.
+Warming climates are expected to cause habitat loss and potentially novel competitive pressures [@Marshall2020-pc; @Soroye2020-zb; though see @Guzman2021-my].
+At the same time, socioeconomic drivers are causing the reforestation of pastures on mountain slopes and the intensification of agriculture in lowlands and valleys, both of which entail a loss of floral resources for bumble bees [@Tasser2002-om; @Tappeiner2006-jp; @Tasser2007-qo; @Walcher2017-kw].
 
-5. What we did
-
-Definitely need to discuss @Hoiss2012-sn and @Hoiss2012-kr and @Hoiss2015-gv
-
-The response-effect framework might be useful [@Suding2008-bl].
-
-<!-- First introduce the concept of ecological thresholds, nonlinear responses. Give some illustrative examples to show the general nature of the phenomenon.  -->
-
-Concept paper on ecological thresholds [@Groffman2006-ch].
-
-Threshold response of stream communities to *urbanization* gradient [@King2011-ry].
-
-Threshold response of anurans to proximity to highways [@Eigenbrod2009-zl].
-
-Nonlinear and threshold responses of nitrogen cycling to *aridity* gradient [@Wang2014-fh].
-
-Threshold response of marine fungal community composition to *salinity* gradient [@Rojas-Jimenez2019-rm].
-
-Threshold and nonlinear responses of marine benthic community to *disturbance* (sheer stress from waves, trawling, etc.) gradients [@Couce2020-hu].
-
-Threshold response of bat communities to habitat loss gradient [@Muylaert2016-rj].
-
-Mountains are unique ecological theaters in which global climate dynamics interact with local climate gradation to shape the distributions of species and biotic interactions [@Telwala2013-sc; @CaraDonna2014-nq; @Miller-Struttmann2014-wz; @Peters2019-km; @Rafferty2020-aj].
-The rate of recent climate change, however, is forcing species to adapt rapidly within the constraints of gene pools shaped by cooler climates [@Visser2008-ct].
-As temperatures rise, species are expected to track suitable conditions to higher elevations, and such effects are already measurable across a broad range of taxa [@Lenoir2015-tj].
-Elevational range shifts, however, are not merely abstract translations along the linear lapse of temperature generated by mountain slopes; they are movements *in situ*, involving the advent, extinction, and reorganization of biological interactions [@Tylianakis2008-fi; @Blois2013-dz] as well as the concomitant variation of abiotic conditions other than temperature [@Korner1995-mh; @Hodkinson2005-af].
-Thus, when species move along continuous, linear climate gradients, they can experience punctuated and nonlinear ecological transitions.
-A salient example is the tree line ecotone that marks the elevational transition between forest and grassland.
-At this relatively discrete band within a linear elevational temperature gradient, plant communities exhibit rapid species turnover [@Descombes2017-id] and declining diversity [@Becker2007-cg] as abiotic conditions become more extreme away from the shelter of tree canopy [@Slatyer1992-ir].
-
-Bumble bees (Hymenoptera: *Bombus* spp.) are classic model organisms in the field of community ecology, and they have more recently become conservation priorities due to well-documented population declines and vulnerability to climate change [@Goulson2008-oe; @Soroye2020-zb].
-Bumble bees are characteristically cold-adapted [@Heinrich1994-cv], with peak abundance and diversity in mountain ranges and northern latitudes, where they are often the dominant guild of flower-visiting insects and the principal pollinators of entomophilous flora [@Williams1998-xi; @Goulson2010-lo].
-Upslope movement of bumble bee species in response to climate change has already been documented, and this is expected to cause the declines or extinction of some bumble bee species [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; Biella2017-ni; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc].
-From a purely geometric perspective, populations of bumble bee species already restricted to the highest elevation zones can be expected to suffer habitat shrinkage [@Dirnbock2011-jf], possibly accompanied by increasing genetic isolation [though see @Fijen2020-zg].
-But the outcomes of elevational range shifts on bumble bees also hinge upon the current elevational distributions of bumble bees and their floral resources that will be acted upon by climate change.
-The upward advance of lowland bumble bee species, for example, may introduce novel competitive pressures on highland species and rewire the pollination networks in which they participate [@Nagamitsu2010-wi; @Brosi2013-mu; @Ishii2013-sm], and elevational [@Lenoir2008-yn] and/or phenological [@Pyke2016-ll] shifts in the floral community could generate mismatches between bumble bee species and their floral hosts.
-When range shifts are considered in this complex biotic context, it is especially important to understand nonlinearities or thresholds in species-, community-, and interaction-level responses to elevational climate gradients, since these could produce dramatic ecological consequences that would not be predicted by geometric considerations alone.
-
-<!-- The outcomes of range shifts in mountain bumble bee and floral communities have obvious significance from a conservation perspective [@Soroye2020-zb], but they also provide a model system in which to study the basic process of community assembly under climate change, a phenomenon of both historical interest and prospective urgency.
--->
-In the present study, set along a ~1400 m elevation gradient in the German Alps, we investigate elevational patterns in a community of bumble bees and wildflowers at three levels of biological organization.
-First, we ask how elevation shapes the abundance and richness of bumble bees and their floral resources.
-Then, we move from the level of individual species to that of community composition and explore the elevational structure of species turnover (i.e.
-β-diversity) in bumble bees and flowering plants.
-Finally, we advance to the level of species interactions, investigating the turnover of bumble- bee-flower interaction partners through elevation.
-In each of these analyses, we employ flexible modeling techniques designed to capture nonlinear responses to elevation, with a particular emphasis on the tree line ecotone as a key threshold that structures the biological communities of mountains slopes.
+In this paper, we present a study based on ~13,000 interactions between bumble bees (Hymenoptera: *Bombus*) and their floral hosts recorded along a 1400 m elevation gradient, complemented by corresponding surveys of floral abundance.
+Using this dataset, we investigate elevational responses of richness, abundance, and β-diversity (i.e.
+species turnover), both at the level of species (bumble bees and flowering plants, respectively) and at the level of interactions (bumble- bee-flower pairs).
+Our results reveal nonlinear and, in some cases, contrasting elevational patterns, underscoring the fact that mountain slopes are complex ecological systems where multiple processes interact to produce biological responses.
+These patterns shed light on the basic ecology of the mountain mutualism between bumble bees and wildflowers, and they suggest how this mutualism might be reshaped by ongoing processes of climate and land use change.
 
@@
 
-The study was conducted 2010-2012 in Berchtesgaden National Park, located in the Northern Limestone Alps of southeast Germany (47.55°N, 12.92°E).
-The landscape is composed of mountain pastures mainly surrounded by coniferous forests.
+The study was conducted from 2010-2012 in Berchtesgaden National Park, located in the Northern Limestone Alps of southeast Germany (47.55°N, 12.92°E).
+The landscape is composed of mountain pastures surrounded mainly by coniferous forest.
 We selected 25 study sites (60 x 60 m) on mountain pastures at elevations ranging from 641-2032 meters above sea level (m.s.l.) **(Figure 1)**.
@@
 Samples were only collected during periods without rain when the air temperature was at least 6°C.
-During each visit, bumble bees and their activity were recorded during a 50-minute transect walk.
+During each visit, interactions between bumble bee and flowers were recorded during a 50-minute transect walk.
 Bumblebees observed on a given flower were counted as floral visitors.
@@
 Floral visitation by male bumble bees was recorded during visitation sampling, but we chose to analyze only visitation by queens and workers.
-Males are often found resting on flowers, particularly late in the year, so their presence on a flower cannot be interpreted reliably as a foraging visit.
+Males are often found resting on flowers, particularly late in the year, so their presence on a flower cannot necessarily be interpreted as a foraging visit.
 In conjunction with visitation observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.01 m^2^.
@@
 Although per-sample effort was standardized across all years, the total number of samples (i.e.
-site-dates) varied: 153 in 2010, 212 in 2011, and 539 in 2012 **(Figure 2)**.
+site-dates) and bumble bee observations varied across years **(Figure 2A-C)**.
 Since the start of field sampling each year was determined by snowmelt, lower sites were sampled as early as April while higher sites could not be sampled until June or July.
-This resulted in a temporally staggered pattern of sampling in 2010 and 2012, while in 2011 a combination of early snowmelt and a late start to field activities resulted in sampling that was approximately aligned in time across the whole elevation gradient.
+This resulted in a temporally staggered pattern of sampling in 2010 and 2012, while in 2011 a combination of early snowmelt and a late start to field activities resulted in sampling that was approximately aligned in time across the whole elevation gradient **(Figure 2D)**.
 
@@
 
-Bumble bee abundance was quantified as the total number of recorded floral visits per bumble bee species per site-date.
-To weight floral abundance based on the foraging preferences of each bumble bee species in our network, we scored each plant genus in our data set as either visited or not visited by each bumble bee species.
-We then calculated the total floral resource availability for each bumble bee species for each site-date by summing the observed flower cover of *visited* genera.
-We opted to work at the genus level based on the reasoning that if a bumble bee species visits one member of a given genus, other members of the same genus should also be considered potential floral hosts.
-This approach was intended to dampen the effects of false non-detection, particularly for rare bumble bee species whose diet breadth would tend to be underestimated simply due to sparsity of observations [@Williams2005-yp].
+We quantified bumble bee abundance as the total number of recorded floral visits per bumble bee species per site-date.
 We treated the cryptic *Bombus terrestris/lucorum* species pair as a single morphospecies, since these two species cannot be visually distinguished, and we pooled representatives of the parasitic *Psithyrus* subgenus --- *Bombus barbutellus*, *B.
@@
 
-Eight plant species --- *Rubus idaeus*, *Rosa canina*, *Juniperus communis*, *Larix decidua*, *Salix* sp., *Caltha palustris*, *Pulmonaria officinalis*, and *Rheum barbarum* --- were recorded in visitation data but not recorded during floral surveying.
-Each accounted for no more than 4 visits in total over the three years of our study, and they were omitted from the analysis of floral abundance.
+To focus our analysis on the subset if flowering plant species that are relevant to bumble bee foraging, we omitted from analysis all plant species that received no visits from bumble bees during the course of our study.
+Eight plant species --- *Rubus idaeus*, *Rosa canina*, *Juniperus communis*, *Larix decidua*, *Salix* sp., *Caltha palustris*, *Pulmonaria officinalis*, and *Rheum barbarum* --- were recorded as visited during interaction sampling but were missed during floral surveying.
+Each accounted for no more than 4 visits in total over the three years of our study, so they were also omitted from analysis.
 
-### Elevational distribution of bumble bees and floral resources
+### Elevational patterns of richness and abundance
 
-We analyzed the abundance of bumble bees and floral resources, respectively, using hierarchical generalized additive models (HGAMs) [@Wood2011-pg; @Pedersen2019-kf].
-To focus on elevational patterns, we collapsed the temporal variation of abundance within sites into summary values.
-For bumble bees, we modeled *peak* abundance at each site as a smooth response to `elevation`, with a group-level smoother for each bumble bee `species`, `management` (grazing, mowing, or unmanaged) as a fixed effect, and `species:year` as a random intercept term (exploratory analysis indicated that species varied differently across years).
-For floral resources, we modeled *mean* flower cover, weighted per bumble bee species (i.e.
-the sum of *visited* genera), as a smooth response to `elevation`, with `management` as a fixed effect, and `year` and bumble bee `species` as random intercept terms.
-For the bumble bee model, we used the negative binomial distribution family (log link function), and for the floral model we used a Tweedie family distribution (log link function).
-Both models follow the "type I" model form described in @Pedersen2019-kf.
+We analyzed the richness and abundance of bumble bees, flowering plants, and their interactions, respectively, using hierarchical generalized additive models (HGAMs) [@Wood2011-pg; @Pedersen2019-kf].
+To focus on elevational patterns, we collapsed within-year variation of richness and abundance at each site into annual summary values of cumulative richness and peak abundance.
+We then modeled these summary values, respectively, as smooth responses to elevation.
+Following the "GI" model form described by @Pedersen2019-kf, we specified both a global smoother and year-specific smoothers, with the latter understood as "random" deviations from the global effect.
+In each model, we also included `year` and `management` as fixed intercept effects, and we included `site` as a random intercept effect to avoid pseudoreplication due to repeated sampling of sites across years.
+In the richness models, we also added `number of transects` as a smoother to account for the effect of sampling intensity on observed richness.
+For all models, we used the quasipoisson (log-link) distribution family.
+
+To explore species-level elevational responses in bumble bees, we fit a model with a species-wise elevation smoother and no global smoother, corresponding to the "I" model form of @Pedersen2019-kf.
+`Year` and `management` were included as fixed intercept effects and `site` and `species` were included as random intercept effects.
+For this model, we used the negative binomial (log link) distribution family because it yielded a better residual distribution than the quasipoisson family.
+
+To facilitate clear interpretation with respect to our focal questions, the models above focus on elevational patterns of abundance by summarizing across seasonal time.
+For those interested in within-season abundance patterns and their interaction with elevation, more complex tensor-product GAMs are included in Appendix 1.
 
@@
 
-For our analyses of species and interaction β-diversity, we pooled all samples (including across years) within sites, since the sparse observations of individual sampling events tended to inflate β-diversity and its variance to a degree that impaired model fitting and interpretation.
-After pooling, we calculated species (separately for bumble bees and plants) and interaction β-diversity between all pairs of sites using the binary Jaccard index and partitioned total interaction β-diversity into its components of species turnover (changes in species composition) and interaction rewiring (interaction turnover despite species constancy) [@Novotny2009-my; @Poisot2012-fk].
-Only plant species that were visited at least once by a bumble bee over the course of the study were included in the calculation of floral β-diversity.
-
-<!-- and modeled each relationship using logistic matrix regression [@Goslee2009-ln].
-Site management difference (i.e.
-same or different management category) was included as a covariate to account for potential management effects, and the difference in number of sampling dates between sites was included as covariate to control for the potential confounding effects of sampling frequency.
--->
+For our analyses of species and interaction β-diversity, we pooled all samples by site irrespective of year.
+After pooling, we calculated β-diversity between all pairs of sites for bumble bees, flowering plants, and theirs interaction, respectively, using the binary Jaccard index.
+We then partitioned total interaction β-diversity into its components of species turnover (changes in species composition) and interaction rewiring (interaction turnover despite species constancy) [@Novotny2009-my; @Poisot2012-fk].
 
 As an initial analysis, we visualized the relationship between site-wise elevation difference and each metric of β-diversity, fitting each response with a binomial regression smooth.
-To investigate the relationship between elevation and β-diversity more deeply, we performed a second analysis using generalized dissimilarity modeling (GDM) [@Ferrier2007-ji] to analyze the elevational variation in the β-diversity of bumble bees, flora, and (unpartitioned) interactions, respectively.
+To investigate the relationship between elevation and β-diversity more deeply, we performed a second analysis using generalized dissimilarity modeling (GDM) [@Ferrier2007-ji] to analyze the elevational variation in the β-diversity of bumble bees, flora, and their (unpartitioned) interactions, respectively.
 In addition to estimating the relationship between differences in a response variable and differences in a predictor variable, GDM captures the slope of this relationship over the range of the predictor variable, revealing potential variation in the amount of change in the response induced by a given change in the predictor [@Ferrier2007-ji].
 This enabled us to ask whether the relationship between β-diversity and elevation exhibits thresholds or other nonlinearities.
-We also included site management in each model by encoding management classes as a series of binary dummy variables, following @Ferrier2007-ji.
-To control for potential confounding effects of the number of sampling days per site, sampling days was included as a covariate.
-The importance of the variables in each model were assessed by a permutation test in which variable importance is defined as the percent loss of explained deviance when a given variable is permuted in isolation [@Fitzpatrick-2021aa].
+To control for potential confounding effects of sampling intensity, sampling days per site was included as a covariate.
+We also included the geographic distances between sites as a covariate to account for potential spatial autocorrelation.
+The importance of the variables in each model were assessed by a permutation test [@Fitzpatrick-2021aa].
 
@@
 Data handling and visualization were performed with the `tidyverse` suite [@Wickham2019-aa].
-GAM analyses were performed with packages `mgcv` [@Wood_2017] and `mgcViz` [@Fasiolo2018aa].
+GAM analyses were performed with packages `mgcv` [@Wood_2017], `mgcViz` [@Fasiolo2018aa], and `gratia` [@Simpson2021-aa].
 β-diversity was calculated with the package `vegan` [@Oksanen_2019] and visualized using the packages `ggplot2` [@Wickham2016-aa] and `ggvegan` [@Simpson2019aa].
@@
 
-<!-- Matrix regression was performed with the package `ecodist` [@Goslee2007-aa]. -->
+# RESULTS
 
-# RESULTS
+**Check to make sure this summary is still true**
 
@@
 
-## Elevational distribution of bumble bees and floral resources
+## Elevational patterns of richness and abundance
 
-HGAM analysis revealed that the response of peak abundance to elevation differed significantly (p ≤ 0.02) from a flat line for all bumble bee species **(Figure 2A)**.
-*Bombus mendax*, *B.
-monticola*, *B.
-mucidus*, *B.
-pyrenaeus*, and *B.
-soroeensis* exhibited a positive response to elevation and peaked in abundance above the tree line, consistent with their known affinity for high elevation habitats [@Rasmont2010-ci].
-In contrast, the abundance of *B.
-hortorum*, *B.
-pascorum*, *B.
-psithyrus*, and *B.
-terrestris/lucorum* declined with elevation.
-This effect was especially pronounced in *B.
-pascuorum*, likely because its exceptionally long nesting cycle is incompatible with the short growing season at high elevations [@Neumayer2009-gx].
-*Bombus gerstaeckeri*, *B.
-jonellus*, and *B.
-wurflenii* peaked at mid-elevation below the tree line, while *B.
-pratorum* was unique in exhibiting a sharp peak of abundance *at* the tree line.
-Our model explained 71% of total variance (adjusted R^2^ = 0.45).
-Neither mowing (p = 0.08) nor non-management (p = 0.95) had a significant effect on bumble bee abundance relative to the reference level of grazing management.
-The random effect of species:year was also nonsignificant (p = 0.52).
-See Appendix S1 for tabulated model output and diagnostics.
+The response of bumble bee richness to elevation exhibited an approximately linear increase up to 1600-1700 m and then approximately leveled off **(Figure 3A)**.
+Neither mowing nor abandonment had a significant effect on abundance relative to the reference management level of grazing (p > 0.05), but richness was significantly lower in 2012 than in the reference year of 2010 (p = 0.04) **(Table 1)**.
+In contrast, the response of floral richness exhibited a clear unimodal pattern, peaking between 1300-1500 m **(Figure 3B)**.
+Management had no significant effect on floral richness (P > 0.1), but floral richness was significantly lower in 2011 (p < 0.001) and 2012 (p = 0.03) than in 2010.
+The response of interaction richness to elevation was intermediate to that of bumble bee and floral richness, peaking around 1500 m and leveling off at higher elevation **(Figure 3C)**.
+Management did not have a significant effect on interaction richness (p > 0.1), but interaction richness was significantly higher in 2011 (p < 0.05) and lower in 2012 (p < 0.001) relative to the reference level of 2010.
 
-Mean floral resource abundance likewise responded significantly (p << 0.0001) to elevation for all bumble bee species **(Figure 2B)**.
-*Bombus jonellus*, *B.
+Cumulative bumble bee abundance did not respond significantly to elevation (p > 0.1) or management (p > 0.1), but bumble bee abundance was significantly higher in 2011 than in the reference year of 2010 (p = 0.03) **(Figure 4A; Table 2)**.
+Floral abundance, however, peaked around 1000 and exhibited an approximately quadratic decline with increasing elevation (p = 0.003).
+Floral abundance was higher in abandoned meadows than in grazed meadows (p = 0.002) and lower in 2012 than in 2010 (p = 0.02).
+
+At the species level, bumble bees exhibited variable and sometimes divergent responses to elevation **(Figure 5)**.
+*Bombus pyrenaeus*, *B.
 mendax*, *B.
-monticola*, *B.
-mucidus*, *B.
-soroeensis*, and *B.
-wurflenii* exhibited bimodal patterns of flora resource abundance, with peaks around 1000 and 1600 m.s.l, the latter peak corresponding closely to the location of the tree line.
-*Bombus pyrenaeus* exhibited a unimodal floral resource pattern peaking near 1600 m.s.l.
-The remaining species --- *B.
-gerstaeckeri*, *B.
-hortorum*, *B.
-pascuorum*, *B.
-pratorum*, *B.
-psithyrus*, and *B.
-terrestris/lucorum* experienced linear or quadratic decline in floral resource abundance with increasing elevation.
-Floral resource abundance was low for all species at elevations above ~1600 m.s.l.
-Our model explained 67% of total variance (adjusted R^2^ = 0.41).
-Both lack of management (p << 0.0001) and mowing (p << 0.0001) had a significant positive effect on mean flower cover relative to the reference level of grazing management.
-The effect of year was also significant, with 2010 (p << 0.0001) and 2011 (p << 0.0001) exhibiting higher mean flower cover than 2012.
-See Appendix S1 for tabulated model output and diagnostics.
+monticola*, and *B.
+mucidus* exhibited approximately linear increases with elevation (p < 0.05).
+In contrast, *Bombus psithyrus* and *B.
+pascuorum* exhibited approximately linear decreases with elevation (p < 0.001).
+*Bombus pratorum* was unique in exhibiting a strongly unimodal response to elevation peaking around 1600 m (p < 0.001).
+The remaining species did not respond significantly to elevation (p > 0.05).
 
@@
 
-GDM analysis revealed a strongly nonlinear response of floral and interaction β-diversity to elevation **(Figure 4A)**.
-Turnover was high from ~600-1000 m, representing the transition from valley to lower slopes, and then leveled off from ~1000-1500 m.
-At around 1500 m --- the elevation of the tree line --- turnover accelerated sharply and remained steep for the remainder of the elevation gradient.
-In contrast, bumble bee turnover was modest and exhibited no marked thresholds.
-Our GDMs explained 65% of total deviance in interaction β-diversity, 74% of floral β-diversity, and 27% of bumble bee β-diversity.
-The conditional effect of elevation was significant (p ~ 0) in all models, and its importance was 1-2 orders of magnitude higher that of sampling frequency and management **(Figure 4B)**.
-Nevertheless, the conditional effect of sampling frequency was significant (p = 0) in the interaction β-diversity model, and mowing management was significant (p = 0) in the floral β-diversity model.
+Bumble bees, flora, and their interactions exhibited marked differences in their rates of β-diversity in response to elevation.
+**(Figure 4)**.
+Bumble bee turnover was relatively high from ~600-1100 m, briefly leveled off, and then accelerated again around 1250 m.
+Floral Turnover was high from ~600-1000 m, gradually leveled off from ~1000-1500 m, and then accelerated again around the tree line ecotone at ~1500 m.
+Interaction turnover mirrored the high rates of bumble bee and floral β-diversity from ~600-1000 m, but it exhibited a markedly different pattern above 1000 m, with relatively low turnover between 1000-1750 m and then a sharp acceleration from ~1750-2000 m.
+The conditional effect of elevation was significant (p < 0.001) in all models, and its importance was 4-20 time higher that of sampling intensity and geographic distance **(Table 3)**.
+Nevertheless, geographic distance was a significant predictor in all models (p < 0.001), and sampling intensity was a significant predictor in the bumble bee (p = 0.004) and interaction models (p < 0.001).
+Our GDMs explained 31% of total deviance in bumble bee β-diversity, 73% of floral β-diversity, and 71% of interactions β-diversity.
 
@@
 
-The core finding of our study, evident in each of our analyses, is that a linear elevation gradient can generate complex and nonlinear biological responses in a mountain bumble- bee-wildflower community in terms of both abundance and β-diversity, and the latter both at the level of community composition and at the level of species interactions.
-These responses are punctuated by the tree line ecotone, which constitutes a discrete ecological threshold along the linear elevation gradient of a mountain slope.
+## Big picture
 
-We found a clear distinction between bumble bee species with low- or mid-elevation abundance peaks (*B.
-gerstaeckeri*, *B.
-hortorum*, *B.
-jonellus*, *B.
-pascorum*, *B.
-pratorum*, *B.
-psithyrus*, *B.
-terrestris/lucorum*, *B.
-wurflenii*) and those with high-elevation abundance peaks (*B.
+The core finding of our study, evident in each of our analyses, is that a linear elevation gradient can generate complex and nonlinear biological responses in a mountain bumble- bee-wildflower community in terms of richness, abundance and β-diversity, and at the level of both species and interactions.
+
+The positive response of bumble bee richness to elevation sets bumble bees apart from most other taxa [@Rahbek1995-oj; @Guo2013-tu; @Peters2016-hf], underscoring their exceptional adaptedness to mountain environments.
+On higher mountains, the increase of bumble bee richness with elevation eventually gives way to a unimodal response, but peak richness is often found above 3000 m [e.g.
+@Williams1991-zu; @Streinzer2019-bz], ~1000 m above the highest peaks in the Berchtesgadener Alps.
+Thus, in our study system, the positive response of bumble bee richness to elevation encompasses the entire mountain slope, in agreement with the findings of @Minachilis2020-xd along the comparable elevation gradient (300---2700 m) of Mt.
+Olympus.
+Importantly, the positive response of bumble bee richness stands in sharp contrast to the monotonic decline of overall bee richness with elevation in the Berchtesgadener Alps [@Hoiss2012-sn], highlighting the special importance of bumble bees as pollinators where few other bees persist.
+
+While the response of overall bumble bee abundance to elevation was approximately flat, this aggregate pattern concealed divergent species-level responses.
+The strongly positive responses of *B.
 mendax*, *B.
@@
 mucidus*, *B.
-pyrenaeus*, and *B.
-soroeensis*).
-The tree line forms an important zone of overlap at which the declining abundance of low- and mid-elevation species intersects with the rising abundance of high-elevation species.
-The tree line was also important in the elevational structure of floral resource abundance.
-For some (mostly high-elevation) bumble bee species, the tree line marked a distinct peak of floral resource availability, often mirrored by a second peak at lower elevation (~1000 m.s.l.).
-It is notable that, for some of the high-elevation species that exhibited this bimodal pattern of floral resource abundance --- *B.
-mendax*, *B.
-monticola*, and *B.
-mucidus* --- the lower peak of floral resource abundance is almost entirely outside their elevational range.
-The tree line, therefore, probably functions as critical foraging habitat for high-elevation bumble bees.
-Above the tree line, we found that flora resource availability declines sharply as flower-rich grassland and shrubland transitions to sparse alpine scree.
-The observation that the upslope advancement of alpine tree lines often lags behind climate warming [@Dullinger2004-qu] suggests that high elevation bumble bees could move upslope faster than their floral resources and become increasingly confined to resource-poor habitat, though the opposite pattern could also occur if the climate-tracking of bumble bees is limited by other factors, such as nest site availability [@Marshall2020-pc].
-Several studies have found that the introduction or removal of bumble bee species from a community can have strong effects on the fitness of other species and on patterns of floral visitation [@Nagamitsu2010-wi; @Brosi2013-mu; @Ishii2013-sm].
-If elevational range shifts introduce new competitive pressures via the encroachment of low-elevation species into the historic ranges of high-elevation species, this effect can be expected to be most pronounced at the tree line.
-The significance of the tree line ecotone as a threshold of floral resource availability and as a zone of overlapping bumble bee distributions corroborates the recent findings of @Minachilis2020-xd in the Mediterranean system of Mount Olympus, indicating that these patterns are likely to generalize well across mountain ecosystems.
+pyrenaeus* are consistent with the known affinity of these species for high-elevation and/or high latitude habitats [@Rasmont2010-ci].
+At the other extreme, *B.
+pascuorum* and *B.
+psithyrus* exhibited a strongly negative responses to elevation.
+In the case of *B.
+pascuorum*, this pattern is likely driven by its exceptionally long nesting cycle, which is a liability under the short growing seasons associated with higher elevations [@Neumayer2009-gx].
+Growing season length may also limit the abundance of *B.
+psithyrus* at high elevation, since cuckoo bumble bees emerge later than the queens of their host species and must successfully usurp host nests before they can begin to raise brood [@Goulson2010-lo].
+*Bombus pratorum* was unique in exhibiting a strongly unimodal abundance pattern with a well-defined peak between 1500-1600 m, .
+@Iserbyt2008-jn describe a similar abundance pattern of *B.
+pratorum* in the Eyne Valley of the Pyrenees.
 
-The tree line not only punctuated patterns of abundance but also those of species-level and interaction-level β-diversity.
-While the turnover rate of bumble bee species was low across the whole elevation gradient and showed no response to the tree line, the turnover rate of the floral community was an order of magnitude higher and exhibited a marked inflection point at the tree line, above which it increased steeply.
-This floristic importance of the tree line has been noted in previous work [@Pellissier2010-kg; @Descombes2017-id], but our study provides a functional extension to this pattern, demonstrating that the tree line also marks a zone of rapid acceleration in the turnover of bumble- bee-flower *interactions*, evidently driven by the underlying turnover of the floral community.
+The roughly sigmoidal pattern of β-diversity rate in bumble bees across our elevation gradient, revealed by our GDM analysis, reflects the reciprocal abundance patterns of high- and low-elevation species described above as well as the occurrence of certain rarer species only at middle-elevation sites.
+It is important to note, though, that bumble bee turnover was low over all, and 8 out of the 16 species spanned the entire elevation gradient (Appendix 1).
 
-The contrast between the relative stability of the bumble bee community and the extreme dynamism of the floral community in our study is striking.
+Floral β-diversity also exhibited a sigmoidal pattern, but its inflection point clearly corresponds to the tree line ecotone at ~1500 m.
+This floristic importance of the tree line has been noted in previous work [e.g.
+@Pellissier2010-kg; @Descombes2017-id], and our study corroborates these findings specifically for the subset of flora visited by bumble bees.
+Interestingly, the tree line does not constitute a discrete transition to a new equilibrium, but rather a threshold at which turnover begins a process of acceleration that holds for the remainder of the elevation gradient.
+While the most obvious effect of the tree line is the relatively discrete transition from forest to scrub/grassland, the tree line also initiates a process of angiosperm species attrition that proceeds as a rate of about 40 species per 100 m [@Korner2003-sc], and the pattern of increasing turnover above the tree line mirrors the patterns of decreasing species richness **(Figure 3)** and floral abundance **(Figure 4)**.
+Thus, the bumble bees foraging above the tree line must cope with a floral community that not only turns over rapidly in composition but also declines steeply in diversity and abundance.
+
+We saw a noteworthy contrast between the relative stability of the bumble bee community and the extreme dynamism of the floral community.
 The dominance of total interaction β-diversity by floral species turnover in our system is consistent with the findings of @Simanonok2014-ou in the Rocky Mountains of North America, and it is perhaps unsurprising given the exceptional cold-hardiness of bumble bees and the sensitivity of plants to temperature gradients.
@@
 Bumble bees have large foraging ranges and have been shown both to track resources through space and time [@Devoto2014-ay] and to cross forest matrix to reach patches of foraging habitat [@Mola2020-yc].
-Assuming an average slope of 20° and a foraging range of 1 km, a bumble bee could travel up- or downslope by more than 340 m, thus spanning an elevation belt nearly 700 m wide.
-Such 3-dimensional foraging would enable bumble bees both to exploit the elevational turnover of floral species and to track preferred species through their elevationally staggered phenology, the latter constituting a sort of physiological time travel [@Van_Straalen1983-df].
+Assuming an average slope of 20° and a foraging range of 1 km, a bumble bee could travel up- or downslope by more than 340 m, thus spanning an elevation band nearly 700 m wide.
+Such 3-dimensional foraging would enable bumble bees both to exploit the elevational turnover of floral species and to track preferred species through their elevationally staggered phenology, the latter constituting a sort of physiological time-travel [@Van_Straalen1983-df].
 
-While elevational patterns were the focus of our study, our analyses also revealed a negative conditional effect of grazing management relative to mowing or non-management.
-This is surprising because other studies have linked agricultural abandonment in the Alps (typically a transition from grazing to non-management) with reduced abundance and diversity of both flora and bumble bees [@Iserbyt2015-jc; @Walcher2017-kw].
-We caution, however, against the over-interpretation of the negative effect of grazing found in our study, since our methods did not focus on the characterization of land use or land use history.
+<!-- Above the tree line, we found that flora resource availability declines sharply as flower-rich grassland and shrubland transitions to sparse alpine scree.
+The observation that the upslope advancement of alpine tree lines often lags behind climate warming [@Dullinger2004-qu] suggests that high elevation bumble bees could move upslope faster than their floral resources and become increasingly confined to resource-poor habitat, though the opposite pattern could also occur if the climate-tracking of bumble bees is limited by other factors, such as nest site availability [@Marshall2020-pc].
+-->
 
-The outcomes of range shifts in mountain bumble bee and floral communities have obvious significance from a conservation perspective [@Soroye2020-zb], but they also provide a model system in which to study the basic process of community assembly under climate change [@Hoiss2012-sn; @Classen2015-jc].
-The central finding of our study, that elevation gradients are characterized by complex and nonlinear responses of species, communities, and biotic interactions --- punctuated by the tree line --- is almost certainly not an idiosyncrasy of bumble bees and their floral hosts but rather a pattern emerging from fundamental physical constraints that should be expected to obtain broadly across the biotic communities of mountains worldwide.
-Species interactions along tree line ecotones, therefore, deserve special consideration in research and conservation management, since climate-induced range shifts at tree lines should be expected to have disproportionate ecological consequences relative to range shifts at other elevation bands.
+<!-- Several studies have found that the introduction or removal of bumble bee species from a community can have strong effects on the fitness of other species and on patterns of floral visitation [@Nagamitsu2010-wi; @Brosi2013-mu; @Ishii2013-sm].
+If elevational range shifts introduce new competitive pressures via the encroachment of low-elevation species into the historic ranges of high-elevation species .
+.
+.
+-->
+
+<!-- The outcomes of range shifts in mountain bumble bee and floral communities have obvious significance from a conservation perspective [@Soroye2020-zb], but they also provide a model system in which to study the basic process of community assembly under climate change [@Hoiss2012-sn; @Classen2015-jc].
+-->
 
@@
 McNeil provided helpful discussion of data analyses.
+M.
+Fitzpatrick kindly corresponded regarding GDM techniques.
 
@@
 <div id ="refs"></div>
+
+\newpage
+
+# TABLES
+
+```{r fig8, echo = FALSE}
+knitr::include_graphics("../ms_1/figures/table1.png")
+```
+
+**Table 1:** Tabular output of richness GAMs.
+Parametric model terms are summarized with an estimated intercept effect ("estimate") while smooth terms are summarized with their effective degrees of freedom ("edf"), the latter representing the complexity (i.e.
+"wiggliness") of the fitted spline.
+Elevation:year smoothers represent year-specific deviations from the global elevation smooths calculated across years.
+
+\newpage
+
+```{r fig9, echo = FALSE}
+knitr::include_graphics("../ms_1/figures/table2.png")
+```
+
+**Table 2:** Tabular output of abundance GAMs.
+Parametric model terms are summarized with an estimated intercept effect ("estimate") while smooth terms are summarized with their effective degrees of freedom ("edf"), the latter representing the complexity (i.e.
+"wiggliness") of the fitted spline.
+Elevation:year smoothers represent year-specific deviations from the global elevation smooths calculated across years.
+
+\newpage
+
+```{r fig10, echo = FALSE}
+knitr::include_graphics("../ms_1/figures/table3.png")
+```
+
+**Table 2:** Tabular output of species-level abundance GAMs.
+Parametric model terms are summarized with an estimated intercept effect ("estimate") while smooth terms are summarized with their effective degrees of freedom ("edf"), the latter representing the complexity (i.e.
+"wiggliness") of the fitted spline.
+
+\newpage
+
+```{r fig11, echo = FALSE}
+knitr::include_graphics("../ms_1/figures/table4.png")
+```
+
+**Table 3:** Summary of GDM results for each model (bumble bees, flora, and interactions, respectively).
+Variable importance represents the percent of explained deviance lost when the focal variable is permuted, and it is mapped to a color ramp to aid interpretation.
+The p-value of each variable (Var.pval) is based on comparing the deviance explained by the unpermuted model (Model.dev.exp) to the distribution of deviance explained when the focal variable is permuted.
+Similarly, the model p-value (Model.pval) is based on a comparison between the deviance explained by the unpermuted model and the distribution of deviance explained when the entire predictor matrix is permuted.
 
@@
 
-**Figure 3:** Conditional effects of elevation and management on the maximum abundance of each bumble bee species (A) and conditional effects of elevation, management, and year on mean flower cover weighted for each bumble bee species (B).
-Effects are plotted on the link scale, and dashed lines depict 95\\% confidence intervals.
-Rug plots depict the locations of sites along the elevation axis and are color-coded to indicate whether each site fell above or below the tree line.
+**Figure 3:** Conditional effects of elevation, year, and management on the richness of bumble bee species (A), floral species (B), and bumble- bee-flower interactions (C).
+Effects are plotted on the link scale, and shaded bands depict 95\\% confidence intervals.
+Bumble bee and plant illustrations were obtained from http://phylopic.org/; see Appendix 1 for details and licensing.
 
-**Figure 4:** Species-level (A) and interaction-level (B, C) plotted against elevational difference.
+**Figure 4:** Conditional effects of elevation, year, and management on the peak abundance of bumble bees (A) and flower cover (B).
+Effects are plotted on the link scale, and shaded bands depict 95\\% confidence intervals.
+The dashed line for the bumble bee abundance smooth indicates that this smooth did not differ significantly from a flat line (p > 0.1).
+
+**Figure 5:** Conditional effects of elevation on the peak abundance of each bumble bee species.
+Effects are plotted on the link scale, and shaded bands depict 95\\% confidence intervals.
+Confidence bands are colored to indicated the linear slope of the response curve, and species are plotted in order of decreasing slope.
+
+**Figure 6:** Species-level (A) and interaction-level (B, C) β-diversity plotted against elevational difference.
 Each point represents the β-diversity (or β-diversity partition) between a pair of sites, and lines represent the overall relationship between β-diversity and elevation difference using binomial regression smooths.
-Standard errors are not plotted because they would be misleading due to the non-independence inherent to distance matrix regression, but all regressions were significant (p < 0.01).
 Interaction β-diversity (B, C) is partitioned using Poisot et al.’s (2012) notation: WN = unpartitioned β-diversity, ST = β-diversity due to species turnover, OS =  β-diversity due to interaction rewiring, ST.h = β-diversity due to species turnover in the higher trophic level (bumble bees), ST.l = β-diversity due to species turnover in the lower trophic level (plants), and ST.lh =  β-diversity due to joint species turnover in higher and lower trophic levels.
 
-**Figure 5:** GDM splines of species-level and interaction-level β-diversity in response to elevation (A) and tabulated model output showing variable importance and significance for elevation, sampling frequency, and site management.
-The maximum height of each spline (A) represents the partial effect of elevation on β-diversity (its effect when covariates are held constant), and the shape of each spline represents to the rate of species turnover as it varies along the elevation gradient, with steeper parts of the curve indicating regions of the gradient over which species/interaction turnover is more rapid.
+**Figure 7:** GDM splines of species- and interaction-level β-diversity in response to elevation.
+The maximum height of each spline represents the conditional (i.e.
+partial) effect of elevation on β-diversity, and the shape of each spline represents to the rate of species turnover as it varies along the elevation gradient, with steeper parts of the curve indicating regions of the gradient over which species/interaction turnover is more rapid.
 Rug plots depict the locations of sites along the elevation axis and are color-coded to indicate whether each site fell above or below the tree line.
-Variable importance (B) is mapped to a color ramp, and p-values less than 0.05 are depicted in bold typeface.
-P-values of 0.00 arise from the discreteness of the permutation test and should be interpreted as arbitrarily small.
 
@@
 ```{r fig1, echo = FALSE}
-knitr::include_graphics("../ms_1/figures/sites_GE.png")
+knitr::include_graphics("../ms_1/figures/fig1.png")
 ```
@@
 ```{r fig2, echo = FALSE}
-knitr::include_graphics("../ms_1/figures/samplingplot.png")
+knitr::include_graphics("../ms_1/figures/fig2.pdf")
 ```
@@
 ```{r fig3, echo = FALSE}
-knitr::include_graphics("../ms_1/figures/gam_abund_grid_mod.png")
+knitr::include_graphics("../ms_1/figures/fig3.pdf")
 ```
@@
 ```{r fig4, echo = FALSE}
-knitr::include_graphics("../ms_1/figures/beta_stack.png")
+knitr::include_graphics("../ms_1/figures/fig4.pdf")
 ```
@@
 ```{r fig5, echo = FALSE}
-knitr::include_graphics("../ms_1/figures/gdm_combo.png")
+knitr::include_graphics("../ms_1/figures/fig5.pdf")
 ```
@@
 
+**Figure6**
+
+```{r fig6, echo = FALSE}
+knitr::include_graphics("../ms_1/figures/fig6.pdf")
+```
+
+\newpage
+
+**Figure7**
+
+```{r fig7, echo = FALSE}
+knitr::include_graphics("../ms_1/figures/fig7.pdf")
+```
+
 <!-- ```{r write_citations, cache=FALSE, include=FALSE} -->

### 2021-09-24 code/ms1_beta_diversity.Rmd
almost there...

@@
 Note that all this deduplication nonsense is not necessary for the betalinkr call because it deduplicates internally.
-
-# Fun with betapart
-fl_betapart <- survey %>%
-  group_by(site, plant.sp) %>%
-  summarize(count = sum(flower.cover)) %>%
-  mutate(present = ifelse(count > 0, 1, 0)) %>%
-  select(-count) %>%
-  pivot_wider(names_from = plant.sp, values_from = present) %>%
-  replace(is.na(.), 0) %>%
-  column_to_rownames("site") %>%
-  beta.pair(index.family = "jaccard")
-
-fl_betapart.turnover <- fl_betapart$beta.jtu %>%
-  as.matrix() %>%
-  as.data.frame() %>%
-  rownames_to_column(var = "site1") %>%
-  pivot_longer(cols = -site1, names_to = "site2", values_to = "tu") %>%
-  filter(site1 != site2) %>%
-  left_join(site_data, by = c("site1" = "site")) %>%
-  dplyr::select(site1, site2, tu, tree_line1 = tree_line) %>%
-  left_join(site_data, by = c("site2" = "site")) %>%
-  dplyr::select(site1, site2, tu, tree_line1, tree_line2 = tree_line)
-
-fl_betapart.nestedness <- fl_betapart$beta.jne %>%
-  as.matrix() %>%
-  as.data.frame() %>%
-  rownames_to_column(var = "site1") %>%
-  pivot_longer(cols = -site1, names_to = "site2", values_to = "ne") %>%
-  filter(site1 != site2) %>%
-  left_join(site_data, by = c("site1" = "site")) %>%
-  dplyr::select(site1, site2, ne, tree_line1 = tree_line) %>%
-  left_join(site_data, by = c("site2" = "site")) %>%
-  dplyr::select(site1, site2, ne, tree_line1, tree_line2 = tree_line)
-
-fl_betapart.collate <- full_join(fl_betapart.turnover, fl_betapart.nestedness) %>%
-  mutate(tree_line.comp = case_when(
-    tree_line1 == "below" & tree_line2 == "below" ~ "bb",
-    tree_line1 == "above" & tree_line2 == "above" ~ "aa",
-    tree_line1 != tree_line2 ~ "ab"
-  )) %>%
-  pivot_longer(cols = c(tu, ne), values_to = "value", names_to = "var")
-
-ggplot(fl_betapart.collate, aes(var, value)) +
-  geom_boxplot() +
-  facet_wrap(~tree_line.comp)
 ```

### 2021-09-24 ms_1/ms_revision1.Rmd
almost there...

@@
 ---
-title: "The elevational structure of a mountain mutualism: modes and thresholds in the responses of bumble bees, wildflowers, and their interactions"
+title: "The elevational structure of a mountain mutualism: richness, abundance, and β-diversity of bumble bees and their floral hosts"
 
@@
 # ABSTRACT
+
+Environmental gradients generate and maintain biodiversity on Earth.
+Mountain slopes are among the most pronounced terrestrial environmental gradients, and the elevational structure of species and their interactions can provide unique insight into the processes that govern community assembly and function in mountain ecosystems.
+
+We recorded bumble- bee-flower interactions over three years along an 1400 m elevational gradient in the German Alps.
+Using nonlinear modeling, we analyze the elevational patterns at the nested levels of species abundance, species β-diversity, and interaction β-diversity.
+
+While floral richness exhibited mid-elevation peak, bumble bee richness increased with elevation before leveling off at the highest sites, demonstrating the exceptional adaptation of these bees to cold temperatures and short growing seasons.
+At a species level, though, bumble bees exhibited divergent responses to elevation in terms of abundance, with a clear separation between species preferring low vs.
+high elevations.
 
@@
 As in the past, so now are mountain habitats critical theaters of ecological change for bumble bees.
-Warming climates are expected to cause habitat loss and potentially novel competitive pressures [@Marshall2020-pc; @Soroye2020-zb; though see @Guzman2021-my].
+Warming climates are expected to cause habitat loss and potentially novel competitive pressures [@Kerr2015-dg; @Soroye2020-zb; though see @Guzman2021-my].
 At the same time, socioeconomic drivers are causing the reforestation of pastures on mountain slopes and the intensification of agriculture in lowlands and valleys, both of which entail a loss of floral resources for bumble bees [@Tasser2002-om; @Tappeiner2006-jp; @Tasser2007-qo; @Walcher2017-kw].
@@
 
-The response of bumble bee richness to elevation exhibited an approximately linear increase up to 1600-1700 m and then approximately leveled off **(Figure 3A)**.
+The response of bumble bee richness to elevation exhibited an approximately linear increase up to 1600-1700 m and then leveled off and declined slightly at the upper limit of the gradient **(Figure 3A)**.
 Neither mowing nor abandonment had a significant effect on abundance relative to the reference management level of grazing (p > 0.05), but richness was significantly lower in 2012 than in the reference year of 2010 (p = 0.04) **(Table 1)**.
@@
 
-At the species level, bumble bees exhibited variable and sometimes divergent responses to elevation **(Figure 5)**.
+At the species level, bumble bees exhibited variable and sometimes divergent responses to elevation **(Figure 5; Table 3)**.
 *Bombus pyrenaeus*, *B.
@@
 monticola*, and *B.
-mucidus* exhibited approximately linear increases with elevation (p < 0.05).
+mucidus* exhibited approximately linear increases with elevation.
 In contrast, *Bombus psithyrus* and *B.
-pascuorum* exhibited approximately linear decreases with elevation (p < 0.001).
-*Bombus pratorum* was unique in exhibiting a strongly unimodal response to elevation peaking around 1600 m (p < 0.001).
+pascuorum* exhibited approximately linear decreases with elevation.
+*Bombus pratorum* was unique in exhibiting a strongly unimodal abundance pattern with a well-defined peak between 1500-1600 m, approximately at the tree line ecotone, similar to the abundance pattern of *B.
+pratorum* described by @Iserbyt2008-jn in the Eyne Valley of the Pyrenees.
 The remaining species did not respond significantly to elevation (p > 0.05).
@@
 Interaction turnover mirrored the high rates of bumble bee and floral β-diversity from ~600-1000 m, but it exhibited a markedly different pattern above 1000 m, with relatively low turnover between 1000-1750 m and then a sharp acceleration from ~1750-2000 m.
-The conditional effect of elevation was significant (p < 0.001) in all models, and its importance was 4-20 time higher that of sampling intensity and geographic distance **(Table 3)**.
+The conditional effect of elevation was significant (p < 0.001) in all models, and its importance was 4-20 time higher that of sampling intensity and geographic distance **(Table 4)**.
 Nevertheless, geographic distance was a significant predictor in all models (p < 0.001), and sampling intensity was a significant predictor in the bumble bee (p = 0.004) and interaction models (p < 0.001).
@@
 
-## Big picture
+Our study revealed strong and sometimes contrasting patterns of elevational structure in the richness, abundance, and β-diversity of mountain bumble bees and wildflowers.
+By extending our analysis to the level of species interactions, we show how the elevational patterns within bumble bee and wildflower communities are propagated to the functional relationships between these two guilds.
+These patterns suggest underlying processes governing the assembly of the bumble- bee-wildflower networks in mountain habitats, shedding light on how this mutualism has evolved in the past and how it will respond to the nascent forces of climate and land use change.
 
-The core finding of our study, evident in each of our analyses, is that a linear elevation gradient can generate complex and nonlinear biological responses in a mountain bumble- bee-wildflower community in terms of richness, abundance and β-diversity, and at the level of both species and interactions.
+While floral richness exhibited the expected unimodal response to elevation, bumble bee richness responded positive to elevation for nearly the whole gradient, showing only a slight declining trend at the highest sites.
+The positive response of bumble bee species richness to elevation sets bumble bees apart from most other taxa [@Rahbek1995-oj; @Guo2013-tu; @Peters2016-hf], underscoring their exceptional adaptation to mountain environments.
+On higher mountains, the increase of bumble bee richness with elevation eventually gives way to a unimodal response, but peak richness is often found above 3000 m [e.g.
+@Williams1991-zu; @Streinzer2019-bz], 1000 m above the highest peaks in the Berchtesgadener Alps.
+Thus, in our study system, the positive response of bumble bee richness to elevation encompasses almost the entire mountain slope, in agreement with the findings of @Minachilis2020-xd along a comparable elevation gradient (300-2700 m) at Mt.
+Olympus, Greece.
+The positive response of bumble bee richness stands in sharp contrast to the monotonic decline of overall bee richness with elevation in the Berchtesgadener Alps [@Hoiss2012-sn], highlighting the special importance of bumble bees as pollinators where few other bees persist.
 
-The positive response of bumble bee richness to elevation sets bumble bees apart from most other taxa [@Rahbek1995-oj; @Guo2013-tu; @Peters2016-hf], underscoring their exceptional adaptedness to mountain environments.
-On higher mountains, the increase of bumble bee richness with elevation eventually gives way to a unimodal response, but peak richness is often found above 3000 m [e.g.
-@Williams1991-zu; @Streinzer2019-bz], ~1000 m above the highest peaks in the Berchtesgadener Alps.
-Thus, in our study system, the positive response of bumble bee richness to elevation encompasses the entire mountain slope, in agreement with the findings of @Minachilis2020-xd along the comparable elevation gradient (300---2700 m) of Mt.
-Olympus.
-Importantly, the positive response of bumble bee richness stands in sharp contrast to the monotonic decline of overall bee richness with elevation in the Berchtesgadener Alps [@Hoiss2012-sn], highlighting the special importance of bumble bees as pollinators where few other bees persist.
-
-While the response of overall bumble bee abundance to elevation was approximately flat, this aggregate pattern concealed divergent species-level responses.
+The response of overall bumble bee abundance to elevation was approximately flat, but this aggregate pattern concealed divergent species-level responses.
 The strongly positive responses of *B.
@@
 mucidus*, *B.
-pyrenaeus* are consistent with the known affinity of these species for high-elevation and/or high latitude habitats [@Rasmont2010-ci].
+pyrenaeus* are consistent with the known affinity of these species for high-elevation habitats [@Rasmont2010-ci].
 At the other extreme, *B.
@@
 In the case of *B.
-pascuorum*, this pattern is likely driven by its exceptionally long nesting cycle, which is a liability under the short growing seasons associated with higher elevations [@Neumayer2009-gx].
-Growing season length may also limit the abundance of *B.
+pascuorum*, this pattern is likely driven by its exceptionally long nesting cycle, which limits its fitness in the short growing seasons of higher elevations [@Neumayer2009-gx].
+Growing season length may also explain the lower abundance of *B.
 psithyrus* at high elevation, since cuckoo bumble bees emerge later than the queens of their host species and must successfully usurp host nests before they can begin to raise brood [@Goulson2010-lo].
-*Bombus pratorum* was unique in exhibiting a strongly unimodal abundance pattern with a well-defined peak between 1500-1600 m, .
-@Iserbyt2008-jn describe a similar abundance pattern of *B.
-pratorum* in the Eyne Valley of the Pyrenees.
+Nevertheless, while the divergent responses of bumble bee species to elevation are informative, it is important to emphasize the general robustness of bumble bees to elevational variation.
+Nine out of the 16 species in our study system did not respond significantly to elevation, and 8 out of 16 were observed across the full span of the gradient.
 
-The roughly sigmoidal pattern of β-diversity rate in bumble bees across our elevation gradient, revealed by our GDM analysis, reflects the reciprocal abundance patterns of high- and low-elevation species described above as well as the occurrence of certain rarer species only at middle-elevation sites.
-It is important to note, though, that bumble bee turnover was low over all, and 8 out of the 16 species spanned the entire elevation gradient (Appendix 1).
-
-Floral β-diversity also exhibited a sigmoidal pattern, but its inflection point clearly corresponds to the tree line ecotone at ~1500 m.
-This floristic importance of the tree line has been noted in previous work [e.g.
-@Pellissier2010-kg; @Descombes2017-id], and our study corroborates these findings specifically for the subset of flora visited by bumble bees.
-Interestingly, the tree line does not constitute a discrete transition to a new equilibrium, but rather a threshold at which turnover begins a process of acceleration that holds for the remainder of the elevation gradient.
-While the most obvious effect of the tree line is the relatively discrete transition from forest to scrub/grassland, the tree line also initiates a process of angiosperm species attrition that proceeds as a rate of about 40 species per 100 m [@Korner2003-sc], and the pattern of increasing turnover above the tree line mirrors the patterns of decreasing species richness **(Figure 3)** and floral abundance **(Figure 4)**.
-Thus, the bumble bees foraging above the tree line must cope with a floral community that not only turns over rapidly in composition but also declines steeply in diversity and abundance.
-
-We saw a noteworthy contrast between the relative stability of the bumble bee community and the extreme dynamism of the floral community.
-The dominance of total interaction β-diversity by floral species turnover in our system is consistent with the findings of @Simanonok2014-ou in the Rocky Mountains of North America, and it is perhaps unsurprising given the exceptional cold-hardiness of bumble bees and the sensitivity of plants to temperature gradients.
+The relative stability of the bumble bee community in our study stands in stark contrast to the extreme dynamism of the floral community, as illustrated by the dominance of total interaction β-diversity by floral species turnover.
+This pattern is consistent with the findings of @Simanonok2014-ou in the Rocky Mountains of North America, and it is perhaps unsurprising given the exceptional cold-hardiness of bumble bees and the sensitivity of plants to temperature gradients.
 Nevertheless, it highlights the behavioral challenge that bumble bees have presumably faced for the entirety of their evolutionary history as mountain florivores: how to forage efficiently in a diverse floral community that exhibits rapid spatial and temporal turnover.
-A fascinating question that could not be answered by our sampling approach is whether mountain bumble bees adaptively forage up- or downslope, as suggested by @Lundberg1980-qk.
+A fascinating question that could not be answered by our sampling approach is whether mountain bumble bees actively exploit the elevational sensitivity of floral communities by adaptively foraging up- or downslope, as suggested by @Lundberg1980-qk.
 Bumble bees have large foraging ranges and have been shown both to track resources through space and time [@Devoto2014-ay] and to cross forest matrix to reach patches of foraging habitat [@Mola2020-yc].
@@
 
-<!-- Above the tree line, we found that flora resource availability declines sharply as flower-rich grassland and shrubland transitions to sparse alpine scree.
-The observation that the upslope advancement of alpine tree lines often lags behind climate warming [@Dullinger2004-qu] suggests that high elevation bumble bees could move upslope faster than their floral resources and become increasingly confined to resource-poor habitat, though the opposite pattern could also occur if the climate-tracking of bumble bees is limited by other factors, such as nest site availability [@Marshall2020-pc].
--->
+The roughly sigmoidal pattern of bumble bee β-diversity rate in response to elevation, revealed by our GDM analysis, probably reflects the reciprocal abundance patterns of high- and low-elevation species described above as well as the occurrence of certain rarer species only at middle-elevation sites.
+As discussed above, though, bumble bee turnover was low over all, demonstrating the robustness of bumble bees to the environmental variation encompassed by our elevation gradient.
+Floral β-diversity rate also exhibited a sigmoidal pattern, and its inflection point coincides with the tree line ecotone at ~1500 m.
+The floristic importance of the tree line has been noted in previous work [e.g.
+@Pellissier2010-kg; @Descombes2017-id], and our study corroborates these findings specifically for the subset of flora visited by bumble bees.
+Interestingly, the tree line did not constitute a discrete transition to a new equilibrium fo floral composition, but rather a threshold marking the beginning of a process of β-diversity acceleration that holds for the remainder of the elevation gradient.
+While the most obvious effect of the tree line may be the relatively discrete transition from forest to scrub/grassland, the tree line also initiates a process of angiosperm species attrition that proceeds at a rate of about 40 species per 100 m [@Korner2003-sc].
+Indeed, the accelerated β-diversity we observed above the tree line mirrors simultaneous patterns of decreasing floral richness and abundance **(Figures 3B, 4B)**.
 
-<!-- Several studies have found that the introduction or removal of bumble bee species from a community can have strong effects on the fitness of other species and on patterns of floral visitation [@Nagamitsu2010-wi; @Brosi2013-mu; @Ishii2013-sm].
-If elevational range shifts introduce new competitive pressures via the encroachment of low-elevation species into the historic ranges of high-elevation species .
-.
-.
--->
+In contrast, the rate of interaction β-diversity exhibited a late and very pronounced acceleration at ~1800 m, indicating that the two highest sites (at 1984 m and 2032 m) differed strongly in interaction composition from those ~200 m below them.
+At these highest sites, the growing season is very short and vegetation is sparse, giving way to bare rock and scree landscapes, and it may be that the last 200 m interval of our elevation gradient imposes the most stringent environmental filter.
+It is interesting that interaction β-diversity did not exhibit the inflection point at the tree line ecotone (~1500 m) that we saw for floral β-diversity, despite the fact that overall interaction β-diversity was driven primarily by floral β-diversity.
+A possible explanation for this pattern is that there could be an elevational lag in bumble bee floral selection such that most bumble bee species do not switch to supra-tree-line floral species when these species first appear, but rather maintain associations with flora whose ranges span the tree line until the extreme species turnover between 1800 and 2000 m forces them to adopt new floral hosts.
 
-<!-- The outcomes of range shifts in mountain bumble bee and floral communities have obvious significance from a conservation perspective [@Soroye2020-zb], but they also provide a model system in which to study the basic process of community assembly under climate change [@Hoiss2012-sn; @Classen2015-jc].
--->
+The poverty of floral resources at the highest elevations raises the question of why bumble bees choose to forage at such elevations at all.
+For some species, the selection of high elevation habitats may be influenced by thermal tolerances [@Oyen2016-ng], but that fails to explain the presence of elevation-generalist species (including the majority of the species in our study system) in resource-poor high-elevation habitats.
+It is possible that nest site limitation at lower elevations drives some bumble bees to establish nests at high elevation, and their ability to forage downslope is subsequently constrained by their flight range, which may be reduced due to the energetic costs of flight in cold temperatures [@Kovac2015-wy].
+It is also possible that high elevations, while sparse in absolute floral resources, offer an escape from more competitive conditions at lower elevations where forager density is higher [@Hoiss2015-gv; @Classen2020-nh].
+Similarly, higher elevations could also offer an escape from density-dependent parasite and pathogen pressure [@Parsche2018-ar].
+Finally, it is worth speculating that an affinity for open, high-elevation habitats might be an evolutionary legacy in mountain bumble bees, since the existence of flower-rich meadows in the forest matrix below the tree line is largely an artifact of human land use arising mainly in the last 500 years of bumble bees' evolutionary history [@Hejcman2013-ca].
 
-# ACKNOWLEDGEMENTS
+Mountains are unique ecological theaters in which the extrinsic forces of climate and land use change interact with intrinsic environmental gradation to shape the ecology of species and their interactions [@Peters2019-km].
+Given the elevational patterns of richness, abundance, and β-diversity among bumble bees and wildflowers in our study, what can be said about the response of these communities, and their functional relationships, to climate and land use change?
+One of the more troubling observations in our study is that the bulk of floral resources were located at low- to mid-elevation.
+If the process of mid-elevation pasture abandonment continues, the flower-rich meadows that now perforate the forest matrix will eventually close into flower-poor coniferous forest, and bumble bees will be pushed either into the more agriculturally intensified land at lower elevation or the relatively resource-poor open habitat above the tree line.
+This land- use-driven redistribution of bumble bees will interact with the concomitant process of climate warming, which has already generated measurable upslope shifts in some bumble bees species [@Ploquin2013-da; @Pyke2016-ll; @Marshall2020-pc], sometimes resulting in increased species richness at higher elevations [@Fourcade2019-ct].
+Under this scenario, the high-elevation bumble bees in the Berchtesgadener Alps can be expected to face increased densities of low-elevation or elevation-generalist congeners, and the introduction of competing bumble bee species in a given habitat can have strong effects both on the fitness of their congeners and on patterns of floral visitation [@Nagamitsu2010-wi; @Brosi2013-mu; @Ishii2013-sm].
+Nevertheless, the tolerance of bumble bees for a relatively broad climate envelope together with their long flight range may enable them to cope with climate and land use change better than their floral mutualists.
+The flowering plant community in our study system was strongly structured by elevation, and there is evidence that the upslope movement of plants in response to warming temperatures may outpace that of bumble bees [@Marshall2020-pc; though see @Pyke2016-ll].
+Moreover, it is the plant community that experiences the direct effects of pasture abandonment at mid-elevation and agricultural intensification at low-elevation.
+
+Finally, our study has focused on contemporary elevational patterns of mountain bumble bees and wildflowers, explored the processes that might regulate them, and related these processes to the nascent pressures of climate and land use change.
+Mountains are, however, inherently dynamic systems, and have been such throughout the co-evolutionary history of bumble bees and their floral mutualists.
+Both contemporary patterns and potential responses to global change processes would be elucidated by greater integration with the historical biogeography of bumble bees and wildflowers and their responses to previous episodes of global change [e.g.
+@Martinet2018-rb].
+
+# ACKNOWLEDGMENTS
 
@@
 
-**Table 2:** Tabular output of species-level abundance GAMs.
+**Table 3:** Tabular output of species-level abundance GAMs.
 Parametric model terms are summarized with an estimated intercept effect ("estimate") while smooth terms are summarized with their effective degrees of freedom ("edf"), the latter representing the complexity (i.e.
@@
 
-**Table 3:** Summary of GDM results for each model (bumble bees, flora, and interactions, respectively).
+**Table 4:** Summary of GDM results for each model (bumble bees, flora, and interactions, respectively).
 Variable importance represents the percent of explained deviance lost when the focal variable is permuted, and it is mapped to a color ramp to aid interpretation.

### 2021-09-27 code/ms1_rich_abund.Rmd
finished draft of revision + reviewer responses

@@
 ```
+
+## 3.5 Does flower cover predict BB abundance?
+```{r}
+ggplot(rich_abund_siteyear, aes(fl.peakabund, bb.peakabund)) +
+  geom_point()
+
+ggplot(rich_abund_siteyear, aes(log(fl.peakabund), log(bb.peakabund))) +
+  geom_point()
+
+ggplot(rich_abund_siteyear, aes(log(fl.peakabund), bb.peakabund)) +
+  geom_point()
+
+library(lme4)
+mod <- glm(bb.peakabund ~ log(fl.peakabund) + elev.mean + year,
+           data = rich_abund_siteyear)
+
+plot(mod)
+summary(mod)
+```

### 2021-09-27 ms_1/ms_revision1.Rmd
finished draft of revision + reviewer responses

@@
 Mountain slopes are among the most pronounced terrestrial environmental gradients, and the elevational structure of species and their interactions can provide unique insight into the processes that govern community assembly and function in mountain ecosystems.
-
 We recorded bumble- bee-flower interactions over three years along an 1400 m elevational gradient in the German Alps.
-Using nonlinear modeling, we analyze the elevational patterns at the nested levels of species abundance, species β-diversity, and interaction β-diversity.
-
+Using nonlinear modeling, we analyzed the elevational patterns at the nested levels of species abundance, species β-diversity, and interaction β-diversity.
 While floral richness exhibited mid-elevation peak, bumble bee richness increased with elevation before leveling off at the highest sites, demonstrating the exceptional adaptation of these bees to cold temperatures and short growing seasons.
@@
 high elevations.
+Overall interaction β-diversity was dominated by strong turnover in the floral community, which exhibited well-defined threshold of β-diversity rate at the tree line ecotone.
+Turnover of bumble bees along the elevation gradient was modest, with the highest rate of β-diversity occurring over the interval from low- to mid-elevation sites.
+Interaction β-diversity increased sharply at the upper extreme of the elevation gradient (1800-2000 m), an interval over which we also saw steep decline in floral richness and abundance, suggesting a zone of stringent environmental filtering.
+Ongoing changes in land use are likely to cause reforestation in resource-rich mid-elevation meadows, forcing bumble bees into agriculturally intensified lowlands or resource-poor highlands, and these effects will interact with climate-induced range shifts in bumble bees and their floral hosts.
+The contrast between the relative stability of bumble bees community and the dynamism of the floral community in the Berchtesgadener Alps suggests that the strongest effects of climate and land use change on mountain bumble bees will be indirect effects mediated by the responses of their floral hosts.
 
@@
 Warming climates are expected to cause habitat loss and potentially novel competitive pressures [@Kerr2015-dg; @Soroye2020-zb; though see @Guzman2021-my].
-At the same time, socioeconomic drivers are causing the reforestation of pastures on mountain slopes and the intensification of agriculture in lowlands and valleys, both of which entail a loss of floral resources for bumble bees [@Tasser2002-om; @Tappeiner2006-jp; @Tasser2007-qo; @Walcher2017-kw].
+At the same time, socioeconomic drivers are causing the reforestation of pastures on mountain slopes and the intensification of agriculture in lowlands, both of which entail a loss of floral resources for bumble bees [@Tasser2002-om; @Tappeiner2006-jp; @Tasser2007-qo; @Walcher2017-kw].
 
 In this paper, we present a study based on ~13,000 interactions between bumble bees (Hymenoptera: *Bombus*) and their floral hosts recorded along a 1400 m elevation gradient, complemented by corresponding surveys of floral abundance.
-Using this dataset, we investigate elevational responses of richness, abundance, and β-diversity (i.e.
-species turnover), both at the level of species (bumble bees and flowering plants, respectively) and at the level of interactions (bumble- bee-flower pairs).
+Using this dataset, we investigate elevational responses of richness, abundance, and β-diversity, both at the level of species (bumble bees and flowering plants, respectively) and at the level of interactions (bumble- bee-flower pairs).
 Our results reveal nonlinear and, in some cases, contrasting elevational patterns, underscoring the fact that mountain slopes are complex ecological systems where multiple processes interact to produce biological responses.
@@
 By extending our analysis to the level of species interactions, we show how the elevational patterns within bumble bee and wildflower communities are propagated to the functional relationships between these two guilds.
-These patterns suggest underlying processes governing the assembly of the bumble- bee-wildflower networks in mountain habitats, shedding light on how this mutualism has evolved in the past and how it will respond to the nascent forces of climate and land use change.
+These patterns suggest underlying processes governing the assembly of bumble- bee-wildflower networks in mountain habitats, shedding light on how this mutualism has evolved in the past and how it may respond to the nascent forces of climate and land use change.
 
-While floral richness exhibited the expected unimodal response to elevation, bumble bee richness responded positive to elevation for nearly the whole gradient, showing only a slight declining trend at the highest sites.
+While floral richness exhibited the expected unimodal response to elevation, bumble bee richness responded positively to elevation for nearly the whole gradient, showing only a slight declining trend at the highest sites.
 The positive response of bumble bee species richness to elevation sets bumble bees apart from most other taxa [@Rahbek1995-oj; @Guo2013-tu; @Peters2016-hf], underscoring their exceptional adaptation to mountain environments.
+In particular, it stands in sharp contrast to the monotonic decline of overall bee richness with elevation in the Berchtesgadener Alps [@Hoiss2012-sn], highlighting the special importance of bumble bees as pollinators where few other bees persist.
 On higher mountains, the increase of bumble bee richness with elevation eventually gives way to a unimodal response, but peak richness is often found above 3000 m [e.g.
-@Williams1991-zu; @Streinzer2019-bz], 1000 m above the highest peaks in the Berchtesgadener Alps.
-Thus, in our study system, the positive response of bumble bee richness to elevation encompasses almost the entire mountain slope, in agreement with the findings of @Minachilis2020-xd along a comparable elevation gradient (300-2700 m) at Mt.
+@Williams1991-zu; @Streinzer2019-bz].
+In the Berchtesgadener Alps, which reach just over 2000 m, the positive response of bumble bee richness to elevation encompasses almost the entire mountain slope, echoing the findings of @Minachilis2020-xd along a comparable elevation gradient (300-2700 m) at Mt.
 Olympus, Greece.
-The positive response of bumble bee richness stands in sharp contrast to the monotonic decline of overall bee richness with elevation in the Berchtesgadener Alps [@Hoiss2012-sn], highlighting the special importance of bumble bees as pollinators where few other bees persist.
 
@@
 At the other extreme, *B.
-pascuorum* and *B.
-psithyrus* exhibited a strongly negative responses to elevation.
+pascuorum* and the *B.
+psithyrus* species group exhibited strongly negative responses to elevation.
 In the case of *B.
@@
 Growing season length may also explain the lower abundance of *B.
-psithyrus* at high elevation, since cuckoo bumble bees emerge later than the queens of their host species and must successfully usurp host nests before they can begin to raise brood [@Goulson2010-lo].
-Nevertheless, while the divergent responses of bumble bee species to elevation are informative, it is important to emphasize the general robustness of bumble bees to elevational variation.
+psithyrus* at high elevation, since cuckoo bumble bees must emerge later than the queens of their host species and then successfully usurp host nests before they can begin to raise brood [@Goulson2010-lo].
+Nevertheless, while the divergent responses of bumble bee species to elevation are informative, it is important to emphasize the overall robustness of bumble bees to elevational variation.
 Nine out of the 16 species in our study system did not respond significantly to elevation, and 8 out of 16 were observed across the full span of the gradient.
@@
 
-The roughly sigmoidal pattern of bumble bee β-diversity rate in response to elevation, revealed by our GDM analysis, probably reflects the reciprocal abundance patterns of high- and low-elevation species described above as well as the occurrence of certain rarer species only at middle-elevation sites.
+The roughly sigmoidal pattern of bumble bee β-diversity rate in response to elevation, revealed by our GDM analysis, probably reflects both the reciprocal abundance patterns of high- and low-elevation species and the occurrence of certain rarer species only at middle-elevation sites.
 As discussed above, though, bumble bee turnover was low over all, demonstrating the robustness of bumble bees to the environmental variation encompassed by our elevation gradient.
@@
 @Pellissier2010-kg; @Descombes2017-id], and our study corroborates these findings specifically for the subset of flora visited by bumble bees.
-Interestingly, the tree line did not constitute a discrete transition to a new equilibrium fo floral composition, but rather a threshold marking the beginning of a process of β-diversity acceleration that holds for the remainder of the elevation gradient.
-While the most obvious effect of the tree line may be the relatively discrete transition from forest to scrub/grassland, the tree line also initiates a process of angiosperm species attrition that proceeds at a rate of about 40 species per 100 m [@Korner2003-sc].
-Indeed, the accelerated β-diversity we observed above the tree line mirrors simultaneous patterns of decreasing floral richness and abundance **(Figures 3B, 4B)**.
+Interestingly, the tree line did not constitute a discrete transition to a new equilibrium of species composition, but rather a threshold marking the beginning of a process of β-diversity acceleration that continued for the remainder of the elevation gradient.
+While the most obvious effect of the tree line may be the relatively discrete transition from forest to scrub/grassland, the tree line also initiates a process of angiosperm species attrition that proceeds at a rate of about 40 species per 100 m [@Korner2003-sc], and the accelerated β-diversity we observed above the tree line mirrors simultaneous patterns of decreasing floral richness and abundance **(Figures 3B, 4B)**.
 
-In contrast, the rate of interaction β-diversity exhibited a late and very pronounced acceleration at ~1800 m, indicating that the two highest sites (at 1984 m and 2032 m) differed strongly in interaction composition from those ~200 m below them.
+In contrast, the rate of interaction β-diversity exhibited a late and very pronounced acceleration at ~1800 m, indicating that the two highest sites at 1984 m and 2032 m differed strongly in interaction composition from the sites ~200 m below them.
 At these highest sites, the growing season is very short and vegetation is sparse, giving way to bare rock and scree landscapes, and it may be that the last 200 m interval of our elevation gradient imposes the most stringent environmental filter.
-It is interesting that interaction β-diversity did not exhibit the inflection point at the tree line ecotone (~1500 m) that we saw for floral β-diversity, despite the fact that overall interaction β-diversity was driven primarily by floral β-diversity.
+It is interesting that interaction β-diversity did not exhibit the inflection point at the tree line ecotone (~1500 m) that we saw for floral β-diversity, despite the fact that overall interaction β-diversity was driven primarily by the turnover of the floral community.
 A possible explanation for this pattern is that there could be an elevational lag in bumble bee floral selection such that most bumble bee species do not switch to supra-tree-line floral species when these species first appear, but rather maintain associations with flora whose ranges span the tree line until the extreme species turnover between 1800 and 2000 m forces them to adopt new floral hosts.
-
 The poverty of floral resources at the highest elevations raises the question of why bumble bees choose to forage at such elevations at all.
 For some species, the selection of high elevation habitats may be influenced by thermal tolerances [@Oyen2016-ng], but that fails to explain the presence of elevation-generalist species (including the majority of the species in our study system) in resource-poor high-elevation habitats.
-It is possible that nest site limitation at lower elevations drives some bumble bees to establish nests at high elevation, and their ability to forage downslope is subsequently constrained by their flight range, which may be reduced due to the energetic costs of flight in cold temperatures [@Kovac2015-wy].
-It is also possible that high elevations, while sparse in absolute floral resources, offer an escape from more competitive conditions at lower elevations where forager density is higher [@Hoiss2015-gv; @Classen2020-nh].
+It is possible that nest site limitation at lower elevations drives some bumble bees to establish nests at high elevation, and their ability to forage downslope is subsequently constrained by their flight range, which may be reduced due to the energetic costs of flight in cold environments [@Kovac2015-wy].
+It is also possible that high elevation landscapes, while sparse in absolute floral resources, offer an escape from more competitive conditions at lower elevations where forager density is higher [@Hoiss2015-gv; @Classen2020-nh].
 Similarly, higher elevations could also offer an escape from density-dependent parasite and pathogen pressure [@Parsche2018-ar].
@@
 Given the elevational patterns of richness, abundance, and β-diversity among bumble bees and wildflowers in our study, what can be said about the response of these communities, and their functional relationships, to climate and land use change?
-One of the more troubling observations in our study is that the bulk of floral resources were located at low- to mid-elevation.
 If the process of mid-elevation pasture abandonment continues, the flower-rich meadows that now perforate the forest matrix will eventually close into flower-poor coniferous forest, and bumble bees will be pushed either into the more agriculturally intensified land at lower elevation or the relatively resource-poor open habitat above the tree line.
 This land- use-driven redistribution of bumble bees will interact with the concomitant process of climate warming, which has already generated measurable upslope shifts in some bumble bees species [@Ploquin2013-da; @Pyke2016-ll; @Marshall2020-pc], sometimes resulting in increased species richness at higher elevations [@Fourcade2019-ct].
-Under this scenario, the high-elevation bumble bees in the Berchtesgadener Alps can be expected to face increased densities of low-elevation or elevation-generalist congeners, and the introduction of competing bumble bee species in a given habitat can have strong effects both on the fitness of their congeners and on patterns of floral visitation [@Nagamitsu2010-wi; @Brosi2013-mu; @Ishii2013-sm].
+Under this scenario, high-elevation bumble bees can be expected to face increased densities of low-elevation or elevation-generalist species, and the introduction of competing bumble bee species in a given habitat can have strong effects both on the fitness of their congeners and on patterns of floral visitation [@Nagamitsu2010-wi; @Brosi2013-mu; @Ishii2013-sm].
 Nevertheless, the tolerance of bumble bees for a relatively broad climate envelope together with their long flight range may enable them to cope with climate and land use change better than their floral mutualists.
 The flowering plant community in our study system was strongly structured by elevation, and there is evidence that the upslope movement of plants in response to warming temperatures may outpace that of bumble bees [@Marshall2020-pc; though see @Pyke2016-ll].
-Moreover, it is the plant community that experiences the direct effects of pasture abandonment at mid-elevation and agricultural intensification at low-elevation.
+Moreover, it is the plant community that experiences the direct effects of land use change.
+Thus, it may be that the strongest effects of climate and land use change on mountain bumble bees will be indirect effects mediated by changes in the floral community.
 
-Finally, our study has focused on contemporary elevational patterns of mountain bumble bees and wildflowers, explored the processes that might regulate them, and related these processes to the nascent pressures of climate and land use change.
-Mountains are, however, inherently dynamic systems, and have been such throughout the co-evolutionary history of bumble bees and their floral mutualists.
-Both contemporary patterns and potential responses to global change processes would be elucidated by greater integration with the historical biogeography of bumble bees and wildflowers and their responses to previous episodes of global change [e.g.
-@Martinet2018-rb].
+Finally, our study has focused on contemporary elevational patterns of mountain bumble bees and wildflowers, exploring the processes that might regulate them, and relating these processes to the nascent pressures of climate and land use change.
+Mountains, however, have been inherently dynamic systems throughout the co-evolutionary history of bumble bees and their floral mutualists [e.g.
+@Hines2008-er; @Martinet2018-rb].
+Both contemporary patterns and potential responses to global change processes would be elucidated by greater integration with the historical biogeography of bumble bees and wildflowers and their responses to previous episodes of global change.
 
@@
 
-**Figure 6:** Species-level (A) and interaction-level (B, C) β-diversity plotted against elevational difference.
+**Figure 6:** Species-level (A) and interaction-level (B, C) β-diversity (in terms of proportional turnover) plotted against elevational difference.
 Each point represents the β-diversity (or β-diversity partition) between a pair of sites, and lines represent the overall relationship between β-diversity and elevation difference using binomial regression smooths.

### 2021-11-05 code/ms1_beta_diversity.Rmd
updating just in case

@@
 ---
-title: "Beta Diversity through Elevation"
-output: html_document
+title: "Appendix 2: Beta-diversity analysis"
+output: 
+  pdf_document:
+    toc: true
+    toc_depth: 3
 ---
@@
 ### 1.2.1 `sfc_as_cols`: convert geometry column back to x and y columns
-```{r}
+```{r message=FALSE}
 # A function for converting a geometry column back to x and y columns
@@
 ### 1.2.2 `web_prep`: convert networks to webs arrays
-```{r}
+```{r message=FALSE}
 # Function to convert networks to webs arrays
@@
     ungroup() %>%
-    dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
+    dplyr::select(higher = bb.sp, 
+                  lower = plant.sp, 
+                  webID = site, freq) %>% # rename columns to match bipartite's expectations
     data.frame() %>% # convert to data frame
@@
 ### 1.2.3 `betalinkr_set1`: betalinkr settings and conversion to tibble
-```{r}
+```{r message=FALSE}
 # Function for betalinkr settings and conversion to tibble()
 betalinkr_set1 <- function(x) {
-  betalinkr_multi(x, partition.st = TRUE, partition.rr = TRUE, # partition both species turnover and richness/replacement
-                  binary = TRUE, # treating beta-diversity as binary is an appropriate complement to our abundance analysis
+  betalinkr_multi(x, partition.st = TRUE, 
+                  partition.rr = TRUE, # partition both species turnover and richness/replacement
+                  binary = TRUE, # treating beta-diversity as binary
                   index = "jaccard") %>% # the Jaccard index is best, I think, for binary analysis
@@
 ### 1.2.4 `interaction_gdm_prep`: convert visitation data into site-pair table for GDM analysis
-```{r}
+```{r message=FALSE}
 # Function for converting visitation data into site-pair table for GDM analysis 
@@
     mutate(present = rep(1, length(.[,1]))) %>% # we will treat abundance as binary
-    unite(interaction, c(bb.sp, plant.sp.abb), sep = "_") %>% # unite bb and fl into interaction "species", i.e. a unique bb-fl combo
-    pivot_wider(names_from = interaction, values_from = present) %>% # spread so that columns are interaction "species"
+    unite(interaction, 
+          c(bb.sp, plant.sp.abb), 
+          sep = "_") %>% # unite bb and fl into interaction "species", i.e. a unique bb-fl combo
+    pivot_wider(names_from = interaction, 
+                values_from = present) %>% # spread so that columns are interaction "species"
     replace(is.na(.), 0) %>% # replace NAs with absences
@@
   sitepair <- formatsitepair(spdat, 
-                             bioFormat = 1, # i.e. samples as rows, species as columns, site-dates as rows
+                             bioFormat = 1, # i.e. samples = rows, species = columns, site-dates = rows
                              siteColumn = "site", 
@@
 ### 1.2.5 `biodata_prep_survey`: prepare the floral survey biodata input to gdm's `formatsitepair()` function
-```{r}
+```{r message=FALSE}
 # Function for preparing the floral survey biodata input to gdm's formatsitepair function
@@
     summarize(cover = sum(flower.cover)) %>%
-    mutate(cover = rep(1, n())) %>% # this makes the data binary, which I think is what I want here; otherwise, just comment out
+    mutate(cover = rep(1, n())) %>% # this makes the data binary
     spread(plant.sp, cover) %>%
@@
 ### 1.2.6 `predData_prep_survey`: prepare the floral survey predictor data input to gdm's `formatsitepair()` function
-```{r}
+```{r message=FALSE}
 # Function for preparing the floral survey predictor data input to gdm's formatsitepair function
@@
 ### 1.2.7 `biodata_prep_BB`: prepare the visitation biodata input to gdm's `formatsitepair()` function
-```{r}
+```{r message=FALSE}
 # Function for preparing the visitation biodata input to gdm's formatsitepair function
@@
     summarize(count = n()) %>%
-    mutate(count = rep(1, n())) %>% # this makes the data binary, which I think is what I want here; otherwise, just comment out
+    mutate(count = rep(1, n())) %>% # this makes the data binary
     spread(bb.sp, count) %>%
@@
 ### 1.2.8 `predData_prep_BB`: prepare the visitation predictor data input to gdm's `formatsitepair()` function
-```{r}
+```{r message=FALSE}
 # Function for preparing the visitation predictor data input to gdm's formatsitepair function
@@
 ### 1.2.9 `format_set`: setting for gdm's `formatsitepair()` function
-```{r}
+```{r message=FALSE}
 # Function with setting for gdm's formatsitepair function
@@
                  predData = pred, abundance = FALSE) %>%
-  filter(s1.elev.mean != s2.elev.mean) # this avoids comparing repeated measures of the same site (taking advantage of the fact that each site has a unique elevation in our dataset)
+  filter(s1.elev.mean != s2.elev.mean) # this avoids comparing repeated measures 
+                                       # of the same site (taking advantage of the 
+                                       # fact that each site has a unique elevation 
+                                       # in our dataset)
 } 
@@
 ### 1.2.10 `gdm_tab`: tabulate `gdm.varImp()` output 
-```{r}
+```{r message=FALSE}
 # Function for tabulating gdm.varImp output 
@@
              predictor == "Geographic" ~ "distance")
-             # predictor == "mowing" ~ "mowing",
-             # predictor == "grasing" ~ "grazing",
-             # predictor == "none" ~ "none")
            ) %>%
@@
 ### 1.2.11 `isplineExtract2`: Fixes a bug in `isplineExtract()` that causes it to break when non-default number of splines are used
-```{r}
+```{r message=FALSE}
 ##Extracts Ispline data from a gdm model
@@
 ### 1.2.12 `plotUncertainty2` Tweaked to call `isplineExtract2` instead of the broken version
-```{r}
+```{r message=FALSE}
 ##########################################################################
@@
 ## 1.3 Load data
-```{r}
+```{r message=FALSE}
 net <- read_csv("../data/processed_data/network.csv") %>%
@@
   filter(flower.cover > 0) %>%
-  semi_join(net, by = "plant.sp") %>% # Consider only visited plant spp. (anytime, anywhere, not specifically for a given site*date)
+  semi_join(net, by = "plant.sp") %>% # Consider only visited plant spp. 
   dplyr::select(site, date, plant.sp, flower.cover)
@@
 ## 1.4 Load images
-```{r}
+```{r message=FALSE}
 BB <- image_read("../ms_1/figures/phylopic_BB.png")
@@
 ## 2.1 BB and FL beta-diversity
-```{r}
+```{r message=FALSE}
 # Bumble bee beta-diversity based on interaction sampling
@@
   column_to_rownames("site") %>%
-  vegdist(method = "jaccard", binary = TRUE) %>% # Jaccard and binary settings make this comparable to the way I'm running betalinkr
+  vegdist(method = "jaccard", 
+          binary = TRUE) %>% # Jaccard and binary settings make this comparable to the way I'm running betalinkr
   as.matrix() %>%
@@
 
-# For each site pair, one will be higher and the other lower; thus, when their order is reversed only the sign of the difference changes.
-Since no two sites have exactly the same elevation, `filter(elev.diff >= 0)` deduplicates the data by dropping the ordering of a given pair of sites (one of the two orderings) that yields a negative elevation difference while retaining the alternative ordering that yields a positive elevation difference.
-By >= 0, I retain comparisons of a site to itself.
-I drop (in the ggplot call) within-site comparisons when I plot against elevation difference, and I isolate within-site comparison when I plot against time.
-In the latter case, it is also necessary to deduplicate, since each pair of site dates occurs in A-B and B-A.
-For that, I filter (again, in the ggplot call) so that yday.diff > 0, similarly dropping the orderings that yield negative differences.
-Note that all this deduplication nonsense is not necessary for the betalinkr call because it deduplicates internally.
+# For each site pair, one will be higher and the other lower; thus, when their order is 
+# reversed only the sign of the difference changes. Since no two sites have exactly the 
+# same elevation, `filter(elev.diff >= 0)` deduplicates the data by dropping the ordering 
+# of a given pair of sites (one of the two orderings) that yields a negative elevation 
+# difference while retaining the alternative ordering that yields a positive elevation 
+# difference. By >= 0, I retain comparisons of a site to itself. I drop (in the ggplot call) 
+# within-site comparisons when I plot against elevation difference, and I isolate within-site 
+# comparison when I plot against time. In the latter case, it is also necessary to deduplicate, 
+# since each pair of site dates occurs in A-B and B-A. For that, I filter (again, in the 
+# ggplot call) so that yday.diff > 0, similarly dropping the orderings that yield negative 
+# differences. Note that all this deduplication nonsense is not necessary for the betalinkr
+# call because it deduplicates internally.
 ```
@@
                                        "OS.repl", "OS.rich")))
-
-# Write to csv for archiving
-#write_csv(beta_mit_data, "../Dryad_ms1/partitioned_beta_diversity")
 ```
@@
 # 4. Visualize beta-diversity ~ elevation diff
-```{r}
+```{r message=FALSE, warning=FALSE}
 # Set palette
@@
   plot_annotation(tag_levels = c("A", "B", "C"))
-# & 
-#   theme_classic()
 
@@
 ## 5.2 Bumble bees
-```{r}
-## GDM is not really built for handling categorical variables; when I include 
-## management in the models, they effect is weak, and I don't know how to
-## interpret the curves. So, I think it is better to drop the mgmt variables
-## from the models and just interpret the continuous effects.
-
+```{r message=FALSE, warning=FALSE}
 # Prep input for GDM
@@
 ## 5.3 Flora
-```{r}
+```{r message=FALSE, warning=FALSE}
 # Prep input for GDM
@@
 ### 5.4.1 Interactions
-```{r}
+```{r message=FALSE}
 # Extract tables
@@
 ### 5.4.2 Bumble bees
-```{r}
+```{r message=FALSE}
 # Extract tables
@@
 ### 5.4.3 Flora
-```{r}
+```{r message=FALSE}
 # Extract tables
@@
 ### 5.4.4 Merge and format
-```{r}
+```{r message=FALSE}
 colourer <- scales::col_numeric(
@@
   valign(j = c("Model.pval", "Model.dev.exp [%]"), valign = "top") %>%
-  autofit() %>%
-  bg(j = "Importance [%]",
-     bg = colourer, part = "body")
+  autofit() 
+# %>%
+#   bg(j = "Importance [%]",
+#      bg = colourer, part = "body")
+
+gdm.varImp.tab
 
@@
 ## 5.5. Plot GDM splines
-```{r}
+```{r message=FALSE}
 # Interaction model
@@
   dplyr::select(elevation = elev.mean, tree_line)
-
-# Plot
-# gdm_plot_ <- ggplot(gdm_splines, aes(elevation, response, linetype = Level)) +
-#   geom_line(size = 0.75) +
-#   #geom_vline(xintercept = 1500) +
-#   geom_rug(data = tree_line, aes(elevation, color = tree_line), inherit.aes = FALSE) +
-#   labs(color = "Tree Line", x = "Elevation", y = "Conditional effect") +
-#   theme_classic(14) +
-#   guides(linetype = "none") +
-#   theme(legend.position = c(0.12, 0.85),
-#         plot.margin = margin(1, 3, 0.5, 0.5, "cm"),)
-# 
-# gdm_plot <- ggdraw() +
-#   draw_plot(gdm_plot_) +
-#   draw_image(BB, scale = 0.11, x = 0.34, y = -0.19) +
-#   draw_image(FL, scale = 0.15, x = 0.33, y = 0.15) +
-#   draw_image(BBxFL, scale = 0.18, x = 0.38, y = 0.41)
-
-#ggsave("../ms_1/figures/fig7.pdf", width = 7, height = 5)
 
@@
 
-# 6. MRM analysis (not included in paper)
-```{r, warning=FALSE}
+<!-- # 6. MRM analysis (not included in paper) -->
+```{r eval=FALSE, warning=FALSE, include=FALSE}
 survey_sample_diff <- survey %>%
@@
 
-# Making plotUncertainty not suck
-```{r}
+<!-- # Making plotUncertainty not suck -->
+```{r eval=FALSE, include=FALSE}
 test <- plotUncertainty3(sitepair_survey, 

### 2021-11-05 code/ms1_rich_abund.Rmd
updating just in case

--- 2021-09-22 code/ms1_rich_abund.Rmd
@@
 ---
-title: "Abundance/diversity GAM"
-output: html_document
+title: "Appendix 1: Richness and abundance analysis"
+output: 
+  pdf_document:
+    toc: true
+    toc_depth: 3
 ---
@@
 ## 1.1 Load packages
-```{r setup, include=FALSE, echo=FALSE}
+```{r setup, message=FALSE}
 library(mgcv)
@@
 ## 1.2 Load site data and floral taxonomy data
-```{r, include=FALSE, echo=FALSE}
+```{r message=FALSE}
 site_data <- read_csv("../data/processed_data/site_data.csv") %>%
-  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, 
+  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, # drop unused columns
                    temp.mean, elev.class, elev.class2)) %>% 
@@
 
-# FL traits
-## family-level taxonomy
+# family-level taxonomy
 fl_tax <- read_csv("../data/processed_data/floral_tax.csv") 
@@
 survey <- read_csv("../data/processed_data/floral_survey.csv") %>%
-  # Consider only visited plant species (anytime anywhere, not specifically for a given site*date)
+  # Consider only visited plant species 
   semi_join(net, by = "plant.sp") %>% 
@@
 ## 1.4 Load images
-```{r}
+Bombus  
+  - Melissa Broussard
+  - http://phylopic.org/image/750eeb4f-7ca5-49ff-a814-27a12d0cb13a/ 
+  - https://creativecommons.org/licenses/by/3.0/
+
+Asteraceae
+  - http://phylopic.org/image/39335c0c-f879-4df6-90d5-e4f65d90d04e/
+  - The person who associated a work with this deed has dedicated the work to the public domain by waiving all of his or her rights to the work worldwide under copyright law, including all related and neighboring rights, to the extent allowed by law.
+
+```{r message=FALSE}
 BB <- image_read("../ms_1/figures/phylopic_BB.png")
@@
 ## 1.5 Define functions
-```{r}
+These are helper functions for tabulating the output of GAMs.
+```{r message=FALSE}
 gam_tab1 <-function(x1, x2, x3, 
@@
     mutate(class = rep("parametric", n()),
-           estimate = round(estimate, 3),
+           estimate = round(estimate, 2),
            statistic = round(statistic, 3),
            p.value = round(p.value, 3),
-           model = rep(name1, n())
+           lower95 = round(conf.low, 2),
+           upper95 = round(conf.high, 2),
+           Model = rep(name1, n())
            )
@@
            p.value = round(p.value, 3),
-           model = rep(name1, n()))
+           Model = rep(name1, n()))
 
@@
     mutate(class = rep("parametric", n()),
-           estimate = round(estimate, 3),
+           estimate = round(estimate, 2),
            statistic = round(statistic, 3),
            p.value = round(p.value, 3),
-           model = rep(name2, n())
+           lower95 = round(conf.low, 2),
+           upper95 = round(conf.high, 2),
+           Model = rep(name2, n())
            )
@@
            p.value = round(p.value, 3),
-           model = rep(name2, n())
+           Model = rep(name2, n())
            )
@@
     mutate(class = rep("parametric", n()),
-           estimate = round(estimate, 3),
+           estimate = round(estimate, 2),
            statistic = round(statistic, 3),
            p.value = round(p.value, 3),
-           model = rep(name3, n())
+           lower95 = round(conf.low, 2),
+           upper95 = round(conf.high, 2),
+           Model = rep(name3, n())
            )
@@
            p.value = round(p.value, 3),
-           model = rep(name3, n())
+           Model = rep(name3, n())
            )
@@
   out <- bind_rows(tab1.1, tab1.2, tab2.1, tab2.2, tab3.1, tab3.2) %>%
-    select(model, class, term, estimate, edf, statistic, p.value) %>%
+    select(Model, class, term, estimate, lower95, upper95, edf, p.value) %>%
     mutate(term = case_when(
@@
       p.value = ifelse(p.value < 0.001, "<0.001", p.value),
-      statistic = ifelse(statistic < 0.001, "<0.001", statistic),
+      #statistic = ifelse(statistic < 0.001, "<0.001", statistic),
       edf = ifelse(edf < 0.001, "<0.001", edf)) %>%
-    flextable() %>%
-    merge_v(j = c("model", "class")) %>%
-    valign(j = c("model", "class"), valign = "top") %>%
-    autofit()
+    as_grouped_data(group = "Model") %>%
+    as_flextable() %>%
+    # merge_v(j = c("Model", "class")) %>%
+    # valign(j = c("Model", "class"), valign = "top") %>%
+    merge_v(j = "class") %>%
+    valign(j = "class", valign = "top") %>%
+    autofit() %>%
+    fontsize(size = 12,
+             j = 1:2) %>%
+    fontsize(size = 12,
+             part = "header")
 }
@@
   out <- bind_rows(tab1.1, tab1.2, tab2.1, tab2.2) %>%
-    select(model, class, term, estimate, edf, statistic, p.value) %>%
+    select(Model, class, term, estimate, lower95, upper95, edf, p.value) %>%
     mutate(term = case_when(
@@
 ## 2.1 Calculate transect-wise FL and BB abundance and richness
-```{r echo=FALSE, include=TRUE, message=FALSE}
+```{r message=FALSE}
 # Survey transects per site
@@
   group_by(site, date, plant.sp) %>% # group by plant.sp*transect
-  summarize(fl.abund = sum(flower.cover)) %>% # get plant.sp*transect total flower cover; this step really shouldn't be required, but in a few cases, there is more than one abundance entry for the same species on a given transect; not sure how that happened, but it's in the original data; alternatively, I could drop the dups or average them.
+  summarize(fl.abund = sum(flower.cover)) %>% # get plant.sp*transect total flower cover; 
+                                              # this step really shouldn't be required, 
+                                              # but in a few cases, there is more than one 
+                                              # abundance entry for the same species on a given transect
   group_by(site, date) %>%
@@
 ## 2.2 Calculate aggregate abundance and richness by site-year
-```{r}
+```{r message=FALSE}
 ### Bumble bee richness per site-year
@@
 
-## 3.1 Preliminary visualization
-```{r eval=FALSE, include=FALSE}
-ggplot(rich_abund_siteyear.long, aes(elev.mean, z, color = group)) +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
-  facet_grid(year~metric) +
-  theme_classic()
-
-ggplot(rich_abund_site.long, aes(z, fill = group)) +
-  geom_histogram() +
-  facet_wrap(~metric) +
-  theme_classic()
-
-# Exploring some bivariate relations
-
-ggplot(rich_abund_site, aes(fl.rich, bb.rich)) +
-  geom_point() +
-  geom_smooth(method = "lm")
-
-ggplot(rich_abund_site, aes(fl.peakabund, fl.rich)) +
-  geom_point() +
-  geom_smooth(method = "lm")
-
-ggplot(rich_abund_site, aes(fl.peakabund, bb.peakabund)) +
-  geom_point() +
-  geom_smooth(method = "lm")
-
-ggplot(rich_abund_site, aes(fl.peakabund, bb.rich)) +
-  geom_point() +
-  geom_smooth(method = "lm")
-
-ggplot(rich_abund_site, aes(survey.transects, net.transects)) +
-  geom_point() +
-  geom_smooth(method = "lm") 
-```
-
-## 3.2 Model richness ~ elevation
+## 3.1 Model richness ~ elevation
 
@@
 
-### 3.2.1 BB richness
-```{r}
+### 3.1.1 BB richness
+```{r message=FALSE}
 # Sanity check visualization
@@
               fill = "black", alpha = 0.2) +
-  # geom_rug(data = semi_join(site_data, net), 
-  #          aes(elev.mean, color = tree_line), 
-  #          inherit.aes = FALSE) +
   theme_classic() +
@@
 
-### 3.2.2 FL richness
-```{r}
+### 3.1.2 FL richness
+```{r message=FALSE}
 # Sanity check visualization
@@
 
-### 3.2.3 Interaction richness
-```{r}
+### 3.1.3 Interaction richness
+```{r message=FALSE}
 # Sanity check visualization
@@
 ### 3.1.4 Compile plots
-```{r}
+```{r message=FALSE}
 gam_plot_richness_ <- ( gam1.1_plot / gam2.1_plot / gam3.1_plot ) +
-  #plot_annotation(tag_levels = 'A') +
+  plot_annotation(tag_levels = 'A') +
   plot_layout(nrow = 3)
@@
 
+gam_plot_richness_
+
 ggsave("../ms_1/figures/fig3.pdf", gam_plot_richness, width = 6.5, height = 8.5)
@@
 
-## 3.3 Model peak abundance ~ elevation
+## 3.2 Model peak abundance ~ elevation
 We will use the same model structure that we used for the richness models.
 
-### 3.3.1 BB abundance
-```{r}
+### 3.2.1 BB abundance
+```{r message=FALSE}
 ggplot(bb_abund_siteyear, aes(elev.mean, bb.peakabund)) +
@@
 
-### 3.3.2 FL abundance
-```{r}
+### 3.2.2 FL abundance
+```{r message=FALSE}
 ggplot(fl_abund_siteyear, aes(elev.mean, fl.peakabund)) +
@@
 
-### 3.3.3 Compile plots
-```{r}
+### 3.2.3 Compile plots
+```{r message=FALSE}
 gam_plot_abundance_ <- ( gam4.1_plot / gam5.1_plot) +
-  #plot_annotation(tag_levels = 'A') +
+  plot_annotation(tag_levels = 'A') +
   plot_layout(nrow = 2)
@@
 
+gam_plot_abundance_
+
 ggsave("../ms_1/figures/fig4.pdf", gam_plot_abundance, width = 6.5, height = 5.5)
@@
 
-### 3.3.4 BB sp-level abundance ~ elevation
+### 3.2.4 BB sp-level abundance ~ elevation
 For this model, the negative-binomial family works best. Maybe it handles zeros better than the quasipoisson.
-```{r}
+```{r message=FALSE}
 ggplot(bb_abund_sp_siteyear, aes(elev.mean, bb.peakabund, color = year)) +
@@
 
-### 3.3.5 Compile tabulated summaries
-```{r}
+## 3.3 Compile tabulated summaries
+```{r message=FALSE}
 gam_tab.rich <- gam_tab1(x1 = gam1.1, x2 = gam2.1, x3 = gam3.1, 
-                   name1 = "Bumble bee richness", 
-                   name2 = "Floral richness", 
-                   name3 = "Interaction richness")
+                   name1 = "Bumble bees", 
+                   name2 = "Flora", 
+                   name3 = "Interactions")
 
@@
 gam_tab.abund <- gam_tab2(x1 = gam4.1, x2 = gam5.1,
-                   name1 = "Bumble bee abundance", 
-                   name2 = "Floral abundance")
+                   name1 = "Bumble bees", 
+                   name2 = "Flora")
 
@@
 gam_tab.abund_sp1 <- gam6.1  %>%
-  tidy(parametric = TRUE) %>%
+  tidy(parametric = TRUE, conf.int = TRUE) %>%
   mutate(class = rep("parametric", n()),
-         estimate = round(estimate, 3),
-         statistic = round(statistic, 3),
-         p.value = round(p.value, 3)
+         estimate = round(estimate, 2),
+         statistic = round(statistic, 3),
+         p.value = round(p.value, 3),
+         lower95 = round(conf.low, 2),
+         upper95 = round(conf.high, 2),
          )
@@
 gam_tab.abund_sp <- bind_rows(gam_tab.abund_sp1, gam_tab.abund_sp2) %>%
-    select(class, term, estimate, edf, statistic, p.value) %>%
+    select(class, term, estimate, lower95, upper95, edf, p.value) %>%
     mutate(term = case_when(
@@
       p.value = ifelse(p.value < 0.001, "<0.001", p.value),
-      statistic = ifelse(statistic < 0.001, "<0.001", statistic),
+      #statistic = ifelse(statistic < 0.001, "<0.001", statistic),
       edf = ifelse(edf < 0.001, "<0.001", edf)) %>%
@@
     valign(j = "class", valign = "top") %>%
-    autofit()
+    autofit() %>%
+    fontsize(size = 12,
+             j = 1:2) %>%
+    fontsize(size = 12,
+             part = "header")
 
@@
 
-### 3.3.6 BB elevation distribution
-```{r}
+## 3.4 BB elevation distribution
+```{r message=FALSE}
 bb_range <- net %>%
@@
 
-## 3.4 Model sample-wise abundance ~ elevation * time
-### 3.4.1 BB abundance
-```{r}
+## 3.5 Model sample-wise abundance ~ elevation * time  
+These models are complex and potentially sensitive to sampling patterns; they should be interpreted cautiously.
+
+### 3.5.1 BB abundance
+```{r message=FALSE}
 gam7.1 <- gam(bb.abund ~ 
@@
 
-### 3.4.2 FL abundance
-```{r}
+### 3.5.2 FL abundance
+```{r message=FALSE}
 gam8.1 <- gam(fl.abund ~ 

### 2021-11-05 code/ms2_nullmodels.Rmd
updating just in case

@@
 
-ggsave("../ms_2/figures/null_model_ktype_preference.pdf", null_model_overall, height = 6, width = 7)
-ggsave("../ms_2/figures/null_model_ktype_preference.png", null_model_overall, height = 6, width = 7)
+ggsave("../ms_2/figures/fig2.pdf", null_model_overall, height = 4.5, width = 6)
+ggsave("../ms_2/figures/fig2.png", null_model_overall, height = 4.5, width = 6)
 
@@
 ggsave("../ms_2/figures/fig3.pdf", null_model_pbl_GAM, height = 6, width = 7)
-ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference.png", null_model_pbl_GAM, height = 6, width = 7)
+ggsave("../ms_2/figures/fig3.png", null_model_pbl_GAM, height = 6, width = 7)
 

### 2021-11-05 ms_1/ms_revision1.Rmd
updating just in case

@@
 ---
-title: "The elevational structure of a mountain mutualism: richness, abundance, and β-diversity of bumble bees and their floral hosts"
+title: "Contrasting patterns of richness, abundance, and turnover in mountain bumble bees and their floral hosts"
 
-author: Douglas B. Sponsler^1^*, Fabrice Requier^2^, Katharina Kallnik^1^, Alice Claßen^1^, A. Fabienne Maihoff^1^, Johanna Sieger^3^, Ingolf Steffan-Dewenter^1^
-
-# date: `r Sys.Date()`
+author: 
+  - Douglas B. Sponsler^1^*  
+  - Fabrice Requier^2^
+  - Katharina Kallnik^1^
+  - Alice Classen^1^
+  - A. Fabienne Maihoff^1^
+  - Johanna Sieger^3^
+  - Ingolf Steffan-Dewenter^1^
 
 output: 
-  bookdown::word_document2:
-    fig_caption: yes
-    fig_width: 7
-    fig_height: 7
-    df_print: kable
-    keep_md: yes
-    number_sections: no
-    toc: no
-    reference_docx: Word_template.docx
+  pdf_document: 
+    keep_tex: yes
+    latex_engine: xelatex 
+    includes:
+      in_header: preamble.tex
 
@@
   - references.bib
-#  - knitcitations.bib
 csl: ecology.csl
 # csl: https://www.zotero.org/styles/ecology
+---
 
----
+\fontsize{12}{22}
+\selectfont
 
@@
 
-Data are provided as private-for-peer review.
-This study also uses novel code.
-All data and code required to reproduce our analyses are archived on Dryad: https://datadryad.org/stash/share/M_4Yt9VVXTRm36FSfwwGzUs2OTXP9Fv52jjETdSu6ys
+This study also uses novel code. All data and code required to reproduce our analyses are archived on Dryad: https://doi.org/10.5061/dryad.s1rn8pk7f
+
+# Running Head: 
+Mountain plants and pollinators
 
@@
 We recorded bumble- bee-flower interactions over three years along an 1400 m elevational gradient in the German Alps.
-Using nonlinear modeling, we analyzed the elevational patterns at the nested levels of species abundance, species β-diversity, and interaction β-diversity.
+Using nonlinear modeling, we analyzed the elevational patterns at the nested levels of species abundance, species $\beta$-diversity, and interaction $\beta$-diversity.
 While floral richness exhibited mid-elevation peak, bumble bee richness increased with elevation before leveling off at the highest sites, demonstrating the exceptional adaptation of these bees to cold temperatures and short growing seasons.
-At a species level, though, bumble bees exhibited divergent responses to elevation in terms of abundance, with a clear separation between species preferring low vs.
+In terms of abundance, though, bumble bees exhibited divergent species-level responses to elevation, with a clear separation between species preferring low vs.
 high elevations.
-Overall interaction β-diversity was dominated by strong turnover in the floral community, which exhibited well-defined threshold of β-diversity rate at the tree line ecotone.
-Turnover of bumble bees along the elevation gradient was modest, with the highest rate of β-diversity occurring over the interval from low- to mid-elevation sites.
-Interaction β-diversity increased sharply at the upper extreme of the elevation gradient (1800-2000 m), an interval over which we also saw steep decline in floral richness and abundance, suggesting a zone of stringent environmental filtering.
-Ongoing changes in land use are likely to cause reforestation in resource-rich mid-elevation meadows, forcing bumble bees into agriculturally intensified lowlands or resource-poor highlands, and these effects will interact with climate-induced range shifts in bumble bees and their floral hosts.
-The contrast between the relative stability of bumble bees community and the dynamism of the floral community in the Berchtesgadener Alps suggests that the strongest effects of climate and land use change on mountain bumble bees will be indirect effects mediated by the responses of their floral hosts.
+Overall interaction $\beta$-diversity was dominated by strong turnover in the floral community, which exhibited well-defined threshold of $\beta$-diversity rate at the tree line ecotone.
+Interaction $\beta$-diversity increased sharply at the upper extreme of the elevation gradient (1800-2000 m), an interval over which we also saw steep decline in floral richness and abundance.
+Turnover of bumble bees along the elevation gradient was modest, with the highest rate of $\beta$-diversity occurring over the interval from low- to mid-elevation sites.
+The contrast between the relative robustness bumble bees and sensitivity of plants to the elevational gradient in our study suggests that the strongest effects of climate change on mountain bumble bees will be indirect effects mediated by the responses of their floral hosts.
+Similarly, ongoing changes in land use are likely to cause succession to forest in resource-rich mid-elevation meadows, which may force bumble bees to seek floral resources in agriculturally intensified lowlands or the sparse highlands above the tree line.
 
-*Keywords*: climate, beta-diversity, elevation gradient, *Bombus*, bumble bee, mountain ecology, pollination network
+**Keywords**: alpine plants, $\beta$-diversity, *Bombus*, climate,  elevation gradient, mountain ecology, pollination network
+
+\newpage
 
@@
 Warming climates are expected to cause habitat loss and potentially novel competitive pressures [@Kerr2015-dg; @Soroye2020-zb; though see @Guzman2021-my].
-At the same time, socioeconomic drivers are causing the reforestation of pastures on mountain slopes and the intensification of agriculture in lowlands, both of which entail a loss of floral resources for bumble bees [@Tasser2002-om; @Tappeiner2006-jp; @Tasser2007-qo; @Walcher2017-kw].
+At the same time, socioeconomic drivers are causing succession to forest in pastures on mountain slopes and the intensification of agriculture in lowlands, both of which entail a loss of floral resources for bumble bees [@Tasser2002-om; @Tappeiner2006-jp; @Tasser2007-qo; @Walcher2017-kw].
 
 In this paper, we present a study based on ~13,000 interactions between bumble bees (Hymenoptera: *Bombus*) and their floral hosts recorded along a 1400 m elevation gradient, complemented by corresponding surveys of floral abundance.
-Using this dataset, we investigate elevational responses of richness, abundance, and β-diversity, both at the level of species (bumble bees and flowering plants, respectively) and at the level of interactions (bumble- bee-flower pairs).
+Using this dataset, we investigate elevational responses of richness, abundance, and $\beta$-diversity, both at the level of species (bumble bees and flowering plants, respectively) and at the level of interactions (bumble- bee-flower pairs).
 Our results reveal nonlinear and, in some cases, contrasting elevational patterns, underscoring the fact that mountain slopes are complex ecological systems where multiple processes interact to produce biological responses.
@@
 Sampling in each year of the study consisted of repeated visits to each site at approximately weekly intervals.
-Samples were only collected during periods without rain when the air temperature was at least 6°C.
-During each visit, interactions between bumble bee and flowers were recorded during a 50-minute transect walk.
+Samples were only collected during periods without rain and when the air temperature was at least 6°C.
+We divided each 60 x 60 m study plot into 10 evenly-spaced 60 x 6 m transects.
+During each visit, the observer walked each transect in a zig-zag pattern for five minutes and recorded all interactions between bumble bees and flowers; thus, each sampling event consisted of a total of 50 minutes of observation.
 Bumblebees observed on a given flower were counted as floral visitors.
@@
 Males are often found resting on flowers, particularly late in the year, so their presence on a flower cannot necessarily be interpreted as a foraging visit.
-In conjunction with visitation observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.01 m^2^.
-Plant species identification followed @Lauber2007-ha and @Oberndorfer2001-ro.
+In conjunction with bumble bee observations, we estimated the flower cover of each herbaceous or shrubby plant species within the study plot to the nearest 0.01 m^2^; this was done by exhaustive visual inspection of the study plot rather than timed transect walks.
+To avoid observer bias, floral cover estimates were always performed by the same investigator (K.
+Kallnik).
+Plant species were identified following @Lauber2007-ha and @Oberndorfer2001-ro.
 
@@
 humilis*, and *B.
-lapidarius* were extremely rare in our samples, so we omitted them from abundance analyses but retained them for β-diversity analyses.
+lapidarius* were extremely rare in our samples, so we omitted them from abundance analyses but retained them for $\beta$-diversity analyses.
 
@@
 To facilitate clear interpretation with respect to our focal questions, the models above focus on elevational patterns of abundance by summarizing across seasonal time.
-For those interested in within-season abundance patterns and their interaction with elevation, more complex tensor-product GAMs are included in Appendix 1.
+For those interested in within-season abundance patterns and their interaction with elevation, more complex tensor-product GAMs are included in Appendix S1.
 
-### Species and interaction β-diversity
+### Species and interaction $\beta$-diversity
 
-For our analyses of species and interaction β-diversity, we pooled all samples by site irrespective of year.
-After pooling, we calculated β-diversity between all pairs of sites for bumble bees, flowering plants, and theirs interaction, respectively, using the binary Jaccard index.
-We then partitioned total interaction β-diversity into its components of species turnover (changes in species composition) and interaction rewiring (interaction turnover despite species constancy) [@Novotny2009-my; @Poisot2012-fk].
+For our analyses of species and interaction $\beta$-diversity, we pooled all samples by site irrespective of year.
+After pooling, we calculated $\beta$-diversity between all pairs of sites for bumble bees, flowering plants, and theirs interaction, respectively, using the binary Jaccard index.
+We then partitioned total interaction $\beta$-diversity into its components of species turnover (changes in species composition) and interaction rewiring (interaction turnover despite species constancy) [@Novotny2009-my; @Poisot2012-fk].
 
-As an initial analysis, we visualized the relationship between site-wise elevation difference and each metric of β-diversity, fitting each response with a binomial regression smooth.
-To investigate the relationship between elevation and β-diversity more deeply, we performed a second analysis using generalized dissimilarity modeling (GDM) [@Ferrier2007-ji] to analyze the elevational variation in the β-diversity of bumble bees, flora, and their (unpartitioned) interactions, respectively.
+As an initial analysis, we visualized the relationship between site-wise elevation difference and each metric of $\beta$-diversity, fitting each response with a binomial regression smooth.
+To investigate the relationship between elevation and $\beta$-diversity more deeply, we performed a second analysis using generalized dissimilarity modeling (GDM) [@Ferrier2007-ji] to analyze the elevational variation in the $\beta$-diversity of bumble bees, flora, and their (unpartitioned) interactions, respectively.
 In addition to estimating the relationship between differences in a response variable and differences in a predictor variable, GDM captures the slope of this relationship over the range of the predictor variable, revealing potential variation in the amount of change in the response induced by a given change in the predictor [@Ferrier2007-ji].
-This enabled us to ask whether the relationship between β-diversity and elevation exhibits thresholds or other nonlinearities.
+This enabled us to ask whether the relationship between $\beta$-diversity and elevation exhibits thresholds or other nonlinearities.
 To control for potential confounding effects of sampling intensity, sampling days per site was included as a covariate.
@@
 GAM analyses were performed with packages `mgcv` [@Wood_2017], `mgcViz` [@Fasiolo2018aa], and `gratia` [@Simpson2021-aa].
-β-diversity was calculated with the package `vegan` [@Oksanen_2019] and visualized using the packages `ggplot2` [@Wickham2016-aa] and `ggvegan` [@Simpson2019aa].
-Calculation and partitioning of interaction β-diversity were performed with the package `bipartite` [@Dormann2008-aa; Dormann2009-aa].
-GDM analysis of β-diversity was performed with the package `gdm` [@Fitzpatrick-2021aa].
-Data and code to reproduce these analyses are archived on Dryad: https://datadryad.org/stash/share/M_4Yt9VVXTRm36FSfwwGzUs2OTXP9Fv52jjETdSu6ys.
+$\beta$-diversity was calculated with the package `vegan` [@Oksanen_2019] and visualized using the packages `ggplot2` [@Wickham2016-aa] and `ggvegan` [@Simpson2019aa].
+Calculation and partitioning of interaction $\beta$-diversity were performed with the package `bipartite` [@Dormann2008-aa; Dormann2009-aa].
+GDM analysis of $\beta$-diversity was performed with the package `gdm` [@Fitzpatrick-2021aa].
+The full workflows of our analyses are available in Appendices S1 and S2, and the data are archived on Dryad: https://datadryad.org/stash/share/M_4Yt9VVXTRm36FSfwwGzUs2OTXP9Fv52jjETdSu6ys.
 
@@
 
-**Check to make sure this summary is still true**
-
-## Summary of visitation and floral survey data
+## Bee-flower interactions and floral survey data
 
@@
 The response of bumble bee richness to elevation exhibited an approximately linear increase up to 1600-1700 m and then leveled off and declined slightly at the upper limit of the gradient **(Figure 3A)**.
-Neither mowing nor abandonment had a significant effect on abundance relative to the reference management level of grazing (p > 0.05), but richness was significantly lower in 2012 than in the reference year of 2010 (p = 0.04) **(Table 1)**.
-In contrast, the response of floral richness exhibited a clear unimodal pattern, peaking between 1300-1500 m **(Figure 3B)**.
-Management had no significant effect on floral richness (P > 0.1), but floral richness was significantly lower in 2011 (p < 0.001) and 2012 (p = 0.03) than in 2010.
-The response of interaction richness to elevation was intermediate to that of bumble bee and floral richness, peaking around 1500 m and leveling off at higher elevation **(Figure 3C)**.
-Management did not have a significant effect on interaction richness (p > 0.1), but interaction richness was significantly higher in 2011 (p < 0.05) and lower in 2012 (p < 0.001) relative to the reference level of 2010.
+The year 2012 had a negative effect (p = 0.04) on bumble bee richness relative to the reference year 2010 **(Figure 3B)**, but neither mowing nor abandonment had a significant effect on bumble bee richness relative to the reference management level of grazing (p > 0.05)  **(Figure 3C; Table 1)** In contrast, the response of floral richness exhibited a clear unimodal pattern, peaking between 1300-1500 m **(Figure 3D)**.
+The years 2011 (p < 0.001) and 2012 (p < 0.03) had a negative effects on floral richness relative to the reference year 2010 **(Figure 3E)**, but again management had no significant effect (p > 0.1) **(Figure 3F; Table 1)**.
+The response of interaction richness to elevation was intermediate to that of bumble bee and floral richness, peaking around 1500 m and leveling off at higher elevation **(Figure 3G)**.
+The effect of year 2011 on interaction richness was positive (p < 0.05) and that of year 2012 negative (p < 0.001) **(Figure 3H)**, but management did not have a significant effect (p > 0.1) **(Figure 3I; Table 1)**.
 
-Cumulative bumble bee abundance did not respond significantly to elevation (p > 0.1) or management (p > 0.1), but bumble bee abundance was significantly higher in 2011 than in the reference year of 2010 (p = 0.03) **(Figure 4A; Table 2)**.
-Floral abundance, however, peaked around 1000 and exhibited an approximately quadratic decline with increasing elevation (p = 0.003).
-Floral abundance was higher in abandoned meadows than in grazed meadows (p = 0.002) and lower in 2012 than in 2010 (p = 0.02).
+Cumulative bumble bee abundance did not respond significantly to elevation (p > 0.1) **(Figure 4A)** or management (p > 0.1) **(Figure 4C)**, but the year 2011 had a positive effect (p = 0.03) on bumble bee abundance relative to the reference year of 2010 **(Figure 4B; Table 2)**.
+Floral abundance, however, peaked around 1000 and exhibited an approximately quadratic decline with increasing elevation (p = 0.003) **(Figure 4D)**.
+The effect of year 2012 was negative (p = 0.002) relative to the reference year 2010 (p = 0.02) **(Figure 4E)**, and the effect of abandoned meadows on floral abundance was positive compared to the reference level of grazed meadows **(Figure 4F; Table 2)**.
 
@@
 
-## Species and interaction β-diversity
+## Species and interaction $\beta$-diversity
 
-Floral β-diversity among our study sites was very high overall and responded steeply to elevation difference between sites **(Figure 3A)**.
+Floral $\beta$-diversity among our study sites was very high overall and responded steeply to elevation difference between sites **(Figure 3A)**.
 Even sites at similar elevation exhibited ~40% species turnover, and the most vertically distant sites (elevation difference > 1250 m) differed by more than 85%.
-Bumble bee β-diversity was, in comparison to floral β-diversity, both lower overall and less responsive to elevational difference between sites **(Figure 3A)**.
+Bumble bee $\beta$-diversity was, in comparison to floral $\beta$-diversity, both lower overall and less responsive to elevational difference between sites **(Figure 3A)**.
 Sites at similar elevation exhibited ~25% species turnover, and species turnover between the most vertically distant sites remained less than 50%.
 
-Total interaction β-diversity was >75% between sites at similar elevation and approached perfect dissimilarity in the most vertically distant sites **(Figure 3B)**.
-Partitioning revealed that total interaction β-diversity was driven primarily by species turnover, which accounted for ~50% of total β-diversity for sites at similar elevation and ~90% of total β-diversity for the most vertically distant sites.
+Total interaction $\beta$-diversity was >75% between sites at similar elevation and approached perfect dissimilarity in the most vertically distant sites **(Figure 3B)**.
+Partitioning revealed that total interaction $\beta$-diversity was driven primarily by species turnover, which accounted for ~50% of total $\beta$-diversity for sites at similar elevation and ~90% of total $\beta$-diversity for the most vertically distant sites.
 Species turnover consisted mainly of turnover in the floral community **(Figure 3C)**, though the joint turnover of plants and bumble bees accounted for ~25% of total species turnover between the most vertically distant sites.
-Interaction rewiring was most significant between sites at similar elevation, where it accounted for ~30% of total β-diversity, but its share of total β-diversity declined to ~10% in the most vertically distant sites **(Figure 3B)**.
+Interaction rewiring was most significant between sites at similar elevation, where it accounted for ~30% of total $\beta$-diversity, but its share of total $\beta$-diversity declined to ~10% in the most vertically distant sites **(Figure 3B)**.
 
-Bumble bees, flora, and their interactions exhibited marked differences in their rates of β-diversity in response to elevation.
+Bumble bees, flora, and their interactions exhibited marked differences in their rates of $\beta$-diversity in response to elevation.
 **(Figure 4)**.
@@
 Floral Turnover was high from ~600-1000 m, gradually leveled off from ~1000-1500 m, and then accelerated again around the tree line ecotone at ~1500 m.
-Interaction turnover mirrored the high rates of bumble bee and floral β-diversity from ~600-1000 m, but it exhibited a markedly different pattern above 1000 m, with relatively low turnover between 1000-1750 m and then a sharp acceleration from ~1750-2000 m.
+Interaction turnover mirrored the high rates of bumble bee and floral $\beta$-diversity from ~600-1000 m, but it exhibited a markedly different pattern above 1000 m, with relatively low turnover between 1000-1750 m and then a sharp acceleration from ~1750-2000 m.
 The conditional effect of elevation was significant (p < 0.001) in all models, and its importance was 4-20 time higher that of sampling intensity and geographic distance **(Table 4)**.
 Nevertheless, geographic distance was a significant predictor in all models (p < 0.001), and sampling intensity was a significant predictor in the bumble bee (p = 0.004) and interaction models (p < 0.001).
-Our GDMs explained 31% of total deviance in bumble bee β-diversity, 73% of floral β-diversity, and 71% of interactions β-diversity.
+Our GDMs explained 31% of total deviance in bumble bee $\beta$-diversity, 73% of floral $\beta$-diversity, and 71% of interactions $\beta$-diversity.
 
@@
 
-Our study revealed strong and sometimes contrasting patterns of elevational structure in the richness, abundance, and β-diversity of mountain bumble bees and wildflowers.
+Our study revealed strong and sometimes contrasting patterns of elevational structure in the richness, abundance, and $\beta$-diversity of mountain bumble bees and wildflowers.
 By extending our analysis to the level of species interactions, we show how the elevational patterns within bumble bee and wildflower communities are propagated to the functional relationships between these two guilds.
@@
 
-The relative stability of the bumble bee community in our study stands in stark contrast to the extreme dynamism of the floral community, as illustrated by the dominance of total interaction β-diversity by floral species turnover.
+The relative stability of the bumble bee community in our study stands in stark contrast to the extreme dynamism of the floral community, as illustrated by the dominance of total interaction $\beta$-diversity by floral species turnover.
 This pattern is consistent with the findings of @Simanonok2014-ou in the Rocky Mountains of North America, and it is perhaps unsurprising given the exceptional cold-hardiness of bumble bees and the sensitivity of plants to temperature gradients.
 Nevertheless, it highlights the behavioral challenge that bumble bees have presumably faced for the entirety of their evolutionary history as mountain florivores: how to forage efficiently in a diverse floral community that exhibits rapid spatial and temporal turnover.
-A fascinating question that could not be answered by our sampling approach is whether mountain bumble bees actively exploit the elevational sensitivity of floral communities by adaptively foraging up- or downslope, as suggested by @Lundberg1980-qk.
+A fascinating question that could not be answered by our sampling approach is to what extent mountain bumble bees actively exploit the elevational sensitivity of floral communities by adaptively foraging up- or downslope, as suggested by @Lundberg1980-qk.
 Bumble bees have large foraging ranges and have been shown both to track resources through space and time [@Devoto2014-ay] and to cross forest matrix to reach patches of foraging habitat [@Mola2020-yc].
@@
 
-The roughly sigmoidal pattern of bumble bee β-diversity rate in response to elevation, revealed by our GDM analysis, probably reflects both the reciprocal abundance patterns of high- and low-elevation species and the occurrence of certain rarer species only at middle-elevation sites.
+The roughly sigmoidal pattern of bumble bee $\beta$-diversity rate in response to elevation, revealed by our GDM analysis, probably reflects both the reciprocal abundance patterns of high- and low-elevation species and the occurrence of certain rarer species only at middle-elevation sites.
 As discussed above, though, bumble bee turnover was low over all, demonstrating the robustness of bumble bees to the environmental variation encompassed by our elevation gradient.
-Floral β-diversity rate also exhibited a sigmoidal pattern, and its inflection point coincides with the tree line ecotone at ~1500 m.
+Floral $\beta$-diversity rate also exhibited a sigmoidal pattern, and its inflection point coincides with the tree line ecotone at ~1500 m.
 The floristic importance of the tree line has been noted in previous work [e.g.
 @Pellissier2010-kg; @Descombes2017-id], and our study corroborates these findings specifically for the subset of flora visited by bumble bees.
-Interestingly, the tree line did not constitute a discrete transition to a new equilibrium of species composition, but rather a threshold marking the beginning of a process of β-diversity acceleration that continued for the remainder of the elevation gradient.
-While the most obvious effect of the tree line may be the relatively discrete transition from forest to scrub/grassland, the tree line also initiates a process of angiosperm species attrition that proceeds at a rate of about 40 species per 100 m [@Korner2003-sc], and the accelerated β-diversity we observed above the tree line mirrors simultaneous patterns of decreasing floral richness and abundance **(Figures 3B, 4B)**.
+Interestingly, the tree line did not constitute a discrete transition to a new equilibrium of species composition, but rather a threshold marking the beginning of a process of $\beta$-diversity acceleration that continued for the remainder of the elevation gradient.
+While the most obvious effect of the tree line may be the relatively discrete transition from forest to scrub/grassland, the tree line also initiates a process of angiosperm species attrition that proceeds at a rate of about 40 species per 100 m [@Korner2003-sc], and the accelerated $\beta$-diversity we observed above the tree line mirrors simultaneous patterns of decreasing floral richness and abundance **(Figures 3B, 4B)**.
 
-In contrast, the rate of interaction β-diversity exhibited a late and very pronounced acceleration at ~1800 m, indicating that the two highest sites at 1984 m and 2032 m differed strongly in interaction composition from the sites ~200 m below them.
+In contrast, the rate of interaction $\beta$-diversity exhibited a late and very pronounced acceleration at ~1800 m, indicating that the two highest sites at 1984 m and 2032 m differed strongly in interaction composition from the sites ~200 m below them.
 At these highest sites, the growing season is very short and vegetation is sparse, giving way to bare rock and scree landscapes, and it may be that the last 200 m interval of our elevation gradient imposes the most stringent environmental filter.
-It is interesting that interaction β-diversity did not exhibit the inflection point at the tree line ecotone (~1500 m) that we saw for floral β-diversity, despite the fact that overall interaction β-diversity was driven primarily by the turnover of the floral community.
+It is interesting that interaction $\beta$-diversity did not exhibit the inflection point at the tree line ecotone (~1500 m) that we saw for floral $\beta$-diversity, despite the fact that overall interaction $\beta$-diversity was driven primarily by the turnover of the floral community.
 A possible explanation for this pattern is that there could be an elevational lag in bumble bee floral selection such that most bumble bee species do not switch to supra-tree-line floral species when these species first appear, but rather maintain associations with flora whose ranges span the tree line until the extreme species turnover between 1800 and 2000 m forces them to adopt new floral hosts.
+
 The poverty of floral resources at the highest elevations raises the question of why bumble bees choose to forage at such elevations at all.
@@
 Mountains are unique ecological theaters in which the extrinsic forces of climate and land use change interact with intrinsic environmental gradation to shape the ecology of species and their interactions [@Peters2019-km].
-Given the elevational patterns of richness, abundance, and β-diversity among bumble bees and wildflowers in our study, what can be said about the response of these communities, and their functional relationships, to climate and land use change?
+Given the elevational patterns of richness, abundance, and $\beta$-diversity among bumble bees and wildflowers in our study, what can be said about the response of these communities, and their functional relationships, to climate and land use change?
 If the process of mid-elevation pasture abandonment continues, the flower-rich meadows that now perforate the forest matrix will eventually close into flower-poor coniferous forest, and bumble bees will be pushed either into the more agriculturally intensified land at lower elevation or the relatively resource-poor open habitat above the tree line.
@@
 
-Finally, our study has focused on contemporary elevational patterns of mountain bumble bees and wildflowers, exploring the processes that might regulate them, and relating these processes to the nascent pressures of climate and land use change.
+<!-- Finally, our study has focused on contemporary elevational patterns of mountain bumble bees and wildflowers, exploring the processes that might regulate them, and relating these processes to the nascent pressures of climate and land use change.
 Mountains, however, have been inherently dynamic systems throughout the co-evolutionary history of bumble bees and their floral mutualists [e.g.
@@
 Both contemporary patterns and potential responses to global change processes would be elucidated by greater integration with the historical biogeography of bumble bees and wildflowers and their responses to previous episodes of global change.
+-->
 
@@
 
-```{r fig8, echo = FALSE}
-knitr::include_graphics("../ms_1/figures/table1.png")
-```
-
 **Table 1:** Tabular output of richness GAMs.
@@
 
+```{r fig8, echo = FALSE, out.width="100%"}
+knitr::include_graphics("../ms_1/figures/table1.png")
+```
+
 \newpage
-
-```{r fig9, echo = FALSE}
-knitr::include_graphics("../ms_1/figures/table2.png")
-```
 
@@
 
+```{r fig9, echo = FALSE, out.width="100%"}
+knitr::include_graphics("../ms_1/figures/table2.png")
+```
+
 \newpage
-
-```{r fig10, echo = FALSE}
-knitr::include_graphics("../ms_1/figures/table3.png")
-```
 
@@
 
+```{r fig10, echo = FALSE, out.width="100%"}
+knitr::include_graphics("../ms_1/figures/table3.png")
+```
+
 \newpage
 
-```{r fig11, echo = FALSE}
+**Table 4:** Summary of GDM results for each model (bumble bees, flora, and interactions, respectively).
+Variable importance represents the percent of explained deviance lost when the focal variable is permuted.
+The p-value of each variable (Var.pval) is based on comparing the deviance explained by the unpermuted model (Model.dev.exp) to the distribution of deviance explained when the focal variable is permuted.
+Similarly, the model p-value (Model.pval) is based on a comparison between the deviance explained by the unpermuted model and the distribution of deviance explained when the entire predictor matrix is permuted.
+
+```{r fig11, echo = FALSE, out.width="100%"}
 knitr::include_graphics("../ms_1/figures/table4.png")
 ```
-
-**Table 4:** Summary of GDM results for each model (bumble bees, flora, and interactions, respectively).
-Variable importance represents the percent of explained deviance lost when the focal variable is permuted, and it is mapped to a color ramp to aid interpretation.
-The p-value of each variable (Var.pval) is based on comparing the deviance explained by the unpermuted model (Model.dev.exp) to the distribution of deviance explained when the focal variable is permuted.
-Similarly, the model p-value (Model.pval) is based on a comparison between the deviance explained by the unpermuted model and the distribution of deviance explained when the entire predictor matrix is permuted.
 
@@
 **Figure 3:** Conditional effects of elevation, year, and management on the richness of bumble bee species (A), floral species (B), and bumble- bee-flower interactions (C).
-Effects are plotted on the link scale, and shaded bands depict 95\\% confidence intervals.
-Bumble bee and plant illustrations were obtained from http://phylopic.org/; see Appendix 1 for details and licensing.
+Effects are plotted on the link scale, and shaded bands depict 95% confidence intervals.
+Bumble bee and plant illustrations were obtained from http://phylopic.org/; see Appendix S1 for details and licensing.
 
 **Figure 4:** Conditional effects of elevation, year, and management on the peak abundance of bumble bees (A) and flower cover (B).
-Effects are plotted on the link scale, and shaded bands depict 95\\% confidence intervals.
+Effects are plotted on the link scale, and shaded bands depict 95% confidence intervals.
 The dashed line for the bumble bee abundance smooth indicates that this smooth did not differ significantly from a flat line (p > 0.1).
@@
 **Figure 5:** Conditional effects of elevation on the peak abundance of each bumble bee species.
-Effects are plotted on the link scale, and shaded bands depict 95\\% confidence intervals.
+Effects are plotted on the link scale, and shaded bands depict 95% confidence intervals.
 Confidence bands are colored to indicated the linear slope of the response curve, and species are plotted in order of decreasing slope.
 
-**Figure 6:** Species-level (A) and interaction-level (B, C) β-diversity (in terms of proportional turnover) plotted against elevational difference.
-Each point represents the β-diversity (or β-diversity partition) between a pair of sites, and lines represent the overall relationship between β-diversity and elevation difference using binomial regression smooths.
-Interaction β-diversity (B, C) is partitioned using Poisot et al.’s (2012) notation: WN = unpartitioned β-diversity, ST = β-diversity due to species turnover, OS =  β-diversity due to interaction rewiring, ST.h = β-diversity due to species turnover in the higher trophic level (bumble bees), ST.l = β-diversity due to species turnover in the lower trophic level (plants), and ST.lh =  β-diversity due to joint species turnover in higher and lower trophic levels.
+**Figure 6:** Species-level (A) and interaction-level (B, C) $\beta$-diversity (in terms of proportional turnover) plotted against elevational difference.
+Each point represents the $\beta$-diversity (or $\beta$-diversity partition) between a pair of sites, and lines represent the overall relationship between $\beta$-diversity and elevation difference using binomial regression smooths.
+Interaction $\beta$-diversity (B, C) is partitioned using Poisot et al.’s (2012) notation: WN = unpartitioned $\beta$-diversity, ST = $\beta$-diversity due to species turnover, OS =  $\beta$-diversity due to interaction rewiring, ST.h = $\beta$-diversity due to species turnover in the higher trophic level (bumble bees), ST.l = $\beta$-diversity due to species turnover in the lower trophic level (plants), and ST.lh =  $\beta$-diversity due to joint species turnover in higher and lower trophic levels.
 
-**Figure 7:** GDM splines of species- and interaction-level β-diversity in response to elevation.
+**Figure 7:** GDM splines of species- and interaction-level $\beta$-diversity in response to elevation.
 The maximum height of each spline represents the conditional (i.e.
-partial) effect of elevation on β-diversity, and the shape of each spline represents to the rate of species turnover as it varies along the elevation gradient, with steeper parts of the curve indicating regions of the gradient over which species/interaction turnover is more rapid.
+partial) effect of elevation on $\beta$-diversity, and the shape of each spline represents to the rate of species turnover as it varies along the elevation gradient, with steeper parts of the curve indicating regions of the gradient over which species/interaction turnover is more rapid.
 Rug plots depict the locations of sites along the elevation axis and are color-coded to indicate whether each site fell above or below the tree line.
@@
 
-```{r fig1, echo = FALSE}
+```{r fig1, echo = FALSE, out.width="100%"}
 knitr::include_graphics("../ms_1/figures/fig1.png")
@@
 
-```{r fig2, echo = FALSE}
+```{r fig2, echo = FALSE, out.width="100%"}
 knitr::include_graphics("../ms_1/figures/fig2.pdf")

### 2021-11-05 ms_2/ms_2_oikos_part2.Rmd
updating just in case

@@
 ---
-title: "Floral preferences of mountain bumble bees are constrained by trait matching but flexible through elevation and season"
+title: "Floral preferences of mountain bumble bees are constrained by functional traits but flexible through elevation and season"
 header-includes:

### 2021-11-05 ms_2/ms_2_oikos_part2_rev.Rmd
updating just in case

--- 2021-11-05 ms_2/ms_2_oikos_part2.Rmd
@@
 # ABSTRACT
-
-Patterns of resource use by animals can clarify how ecological communities have assembled in the past, how they currently function, and how they are likely to respond to future perturbations.
-Using ~13,000 records of bumble- bee-wildflower interactions along a 1400 m elevation gradient, we find that the selection of floral morphotypes by bumble bees is driven both by the directed process of trait-matching and the neutral effect of floral abundance.
-Patterns of floral preference also vary by elevation and season, which may be an adaptation to the strong spatiotemporal turnover of mountain floral communities.
-The flexibility of bumble bee foraging will be tested in coming years as climate and land use change alter floral resource distribution in time and space, particularly in the mountain ecosystems that host the world's richest communities of bumble bees and their floral mutualists.
 
@@
 
-Resource selection and partitioning by animals are instrumental in community assembly, species coexistence, and ecosystem function [@Chesson2000-aw; @Bascompte2007-ls].
-Resource selection is fundamentally constrained by phenotypic compatibility (i.e.
-trait-matching) between resources and consumers, but it can also be influenced by the relative availability of resources [@Johnson1980-co] and modulated by exogenous drivers such as competition [@Pianka1974-qk].
-Understanding these interacting processes is key to anticipating the response of ecological communities to anthropogenic climate and land use change, which threaten to disrupt existing patterns of species interactions and niche relationships [@Tylianakis2008-fi].
+Resource selection by animals is instrumental in community assembly, species coexistence, and ecosystem function [@Chesson2000-aw; @Bascompte2007-ls].
+*Need another sentence in here* Understanding patterns of resource selection in biological communities is also key to anticipating the effects of climate and land use change, which can be expected to reorganize species interactions and niche relationships [@Tylianakis2008-fi].
 
-\
+Bumble bee (Hymentoptera: *Bombus* spp.) communities and their floral hosts raise interesting questions of resource selection and coexistence that have attracted the attention of ecologists for decades [e.g.
+@Heinrich1976-wc; @Inouye1977-kl].
+All bumble bee species are grossly similar in morphology and behavior, and all share the same general diet of floral nectar and pollen.
+The most important trait that distinguishes bumble bee species with respect to dietary niche is tongue length.
+Bumble bee species can differ markedly in tongue length, and floral handling time increases when tongue length and corolla depth are poorly matched [@Inouye1980-mv; @Harder1983-bk], leading to an association between tongue length and floral choice [e.g.
+@Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn].
+Under conditions of food limitation, the coexistence of bumble bee species should be limited by their similarity in tongue length, and hence in dietary niche.
+Based on field studies in the Rocky Mountains, @Inouye1977-kl concluded that, assuming food limitation, four bumble bee species can be expected to persist in any given locality: a long-tongued species, a medium-tongue species, a short-tongued species, and a nectar-robber (i.e.
+a species that accesses nectaries by biting through the corolla, thus bypassing the ordinary constraint of tongue length).
+In many places, though, considerably more than four bumble bee species coexist, indicating either that food is not limiting or that competitive conditions are too variable in time and space to be detected at the scale of field observations [e.g.
+@Ranta1981-vx; @Williams1989-vr; @Goulson2008-zy].
 
-Bumble bees (Hymentoptera: *Bombus* spp.) and their floral hosts provide a diverse yet tractable system in which to explore resource selection in the context of plant-pollinator networks.
-Due to their adaptation to cold temperatures, bumble bees are sensitive to climate change [@Goulson2008-oe; @Soroye2020-zb], especially in the mountain ecosystems that host the world's richest bumble bee communities.
-Warming climates have already caused measurable upslope shifts in mountain bumble bees and their floral hosts [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc], which could introduce novel competitive interactions between bumble bee species that experience increased co-occurrence.
-At the same time, socioeconomic factors have led to widespread abandonment and reforestation of mountain meadows, particularly in the European Alps, which reduces floral resource availability and suppresses bumble bee diversity and abundance [@Tappeiner2006-jp; @Tasser2007-qo; @Walcher2017-kw].
-The ability of bumble bee species to coexist under these conditions of reshuffled distributions and declining floral resources will depend in part on how they select and partition floral resources, which will in turn affect the viability of their floral mutualists, many of which depend principally on bumble bee visitation for pollination [@Goulson2010-lo].
+The most species rich bumble bee communities are found in the mountainous regions of Eurasia [@Williams1998-xi; @Hines2008-er], where the predictions of competitive exclusion are rarely realized [@Ranta1981-vx; @Goulson2008-zy].
+Nevertheless, climate and land use change in these regions have the potential to alter and intensify competitive relationships between bumble bee species.
+Warming climates have already caused measurable upslope shifts in some bumble bee species and their floral hosts [@Ploquin2013-da; @Pyke2016-ll; @Fourcade2019-ct; @Marshall2020-pc], which could introduce novel competitive interactions between bumble bee species that experience increased co-occurrence.
+At the same time, socioeconomic factors are causing widespread abandonment and reforestation of mountain meadows, particularly in the European Alps, which dramatically reduces floral resource abundance and diversity [@Tappeiner2006-jp; @Tasser2007-qo; @Walcher2017-kw].
+The consequences of reshuffled species distributions and declining floral resources for the coexistence of mountain bumble bees will depend in part on how bumble bee species select and partition available flora.
+Under competitive conditions, species coexistence is possible when intraspecific competition exceeds interspecific competition, and this stabilizing inequality is favored to the extent that species differ in dietary niche [@Abrams1977-zp; @Hanski1995-kr; @Chesson2000-aw].
+Estimating dietary niche overlap from field observations of resource use is problematic, though, because resource selection is usually a function not only of species' intrinsic preferences but also of the relative abundances of the available resources [e.g.
+@Beyer2010-vr].
+Diet overlap observed in a particular time and place may, therefore, be a poor predictor of diet overlap under different conditions of resource abundance.
+A more robust measure of competition potential between species can be obtained by isolating the process of resource preference from the neutral effects of resource abundance, and one way to do this is to compare observed patterns of resource selection to those predicted by a null model in which resources are selected at random in proportion to their relative abundance [@Vazquez2003-qa; @Bluthgen2008-rp].
 
-\
-
-Addressing the question of how bumble bees will select floral resources under novel conditions requires discernment of the generalizable mechanisms that underpin patterns of floral selection observed under current conditions.
-It is ordinarily assumed that the choice of resources by foraging animals is a joint function of trait-based preferences and the neutral effects of the relative abundances of available resources: the so-called "use-availability" model of resource selection [e.g.
-@Beyer2010-vr].
-The most well-studied mechanism of the preference component of bumble bee floral selection is trait-matching between bumble bee tongue length and floral corolla depth.
-Bumble bee species can differ markedly in tongue length, and floral handling time increases when tongue length and corolla depth are poorly matched [@Inouye1980-mv; @Harder1983-bk], leading to an association between tongue length and floral choice in bumble bee communities [e.g.
-@Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn].
-Under the use-availability model, tongue-length-based floral preferences should be modulated by the relative abundance of available resources to generate observed patterns of floral selection.
-It has recently been proposed, however --- in what has been termed the *target use hypothesis* --- that floral selection by bumble bees is a process of *pure preference*, irrespective of resource abundance [@De_Valpine2013-be; @Harmon-Threatt2017-bz].
-If and when the *target use hypothesis* holds, observed visitation rates are a better index of preference than metrics designed to control for resource abundance (e.g.
-null model residuals), since the subtraction of presumed neutral effects of abundance when those effects do not really exist actually creates new abundance-dependent artifacts [@Harmon-Threatt2017-bz].
-On the other hand, failing to account for neutral effects of abundance when they do exist similarly obscures the general mechanisms behind observed foraging patterns [@Vazquez2003-qa; @Bluthgen2008-rp] that will determine bumble bee resource selection under novel conditions induced by climate and land use change.
-
-\
-
-In this study, we use ~13,000 records of bumble bee visits to wildflowers, collected over three years along a 1400 m elevation gradient, to explore three unanswered questions regarding resource selection by bumble bees and its response to climate and land use change.
-First, we ask how bumble bee floral preferences map onto a generalized floral morphotypology [@Kugler1970-gb].
-Then, using the established relationship between tongue length and floral preference as a point of reference, we test the *target use hypothesis* by comparing the explanatory power of bumble bee tongue length with respect to observed visitation rates vs.
-the residuals of a null model that assumes a linear dependency of selection on resource abundance.
-Under the *target use hypothesis*, tongue length should explain observed visitation rates better than null model residuals, while the opposite finding would indicate that, contrary to the *target use hypothesis*, bumble bee floral selection is influenced by relative abundance.
-Finally, we model the variation of bumble bee floral preference through time and elevation as a first step toward understanding how climate and land use change might distort the temporal and elevational structure of bumble- bee-wildflower mutualisms in mountain ecosystems.
-
-\
+In this study, we explore the floral morphotype preferences of a mountain bumble bee community using ~13,000 records of bumble bee visits to wildflowers collected over three years along a 1400 m elevation gradient.
+First, we use null model analysis to compare floral morphotypes in terms of community-level preferences, identifying morphotypes that are consistently over- or under-visited relative to their abundance.
+Next, we compare morphotype preferences across bumble bee species and model species-level preferences as a function of tongue length.
+Then, we analyze preference similarity across species to identify species among which competition is most likely under conditions of resource limitation.
+Finally, we explore the flexibility of floral preferences in bumble bees by modeling the variation of morphotype preferences through elevation and time.
 
@@
 
-\
-
-Bumble- bee-wildflower interactions were observed during repeated visits to each study site at approximately weekly intervals.
+We recorded bumble- bee-wildflower interactions during repeated visits to each study site at approximately weekly intervals.
 Sampling was only conducted during periods without rain and with an air temperature of at least 6°C.
-Bumble bees were recorded during one 50-minute transect walk per visit, and bumblebees observed on a given flower were counted as floral visitors.
+Each 60 x 60 m study plot was divided into 10 evenly-spaced 60 x 6 m transects.
+During each visit, the observer walked each transect in a zig-zag pattern for five minutes and recorded all bumble bee visits to flowers, resulting in a total of 50 minutes of observation.
 Queens were identified to species level in the field to avoid destructively sampling colony foundresses, but workers and males were collected and identified in the lab following @Amiet1996-jz.
@@
 Males were recorded, but we analyzed only the pooled visitation of queens and workers.
-In conjunction with each bumble bee visitation sample, we estimated the flower cover of herbaceous and shrubby plant species within each 60 x 60 m study plot to the nearest 0.01 m^2^ and identified the plant species according to @Lauber2007-ha and @Oberndorfer2001-ro.
-See Sponsler et al.
-(submitted) for details of sampling design.
-
-\
+In conjunction with each bumble bee visitation sample, we visually estimated the flower cover of herbaceous and shrubby plant species within each 60 x 60 m study plot to the nearest 0.01 m^2^ and identified the plant species according to @Lauber2007-ha and @Oberndorfer2001-ro.
+Floral cover estimation was performed by exhaustive inspection of each study plot rather than timed transect walks, and the same investigator (K.
+Kallnik) performed all floral surveys to avoid observer bias.
 
@@
 
-\
-
 Bumble bee tongue-length data (for the worker caste) were obtained from the literature values compiled and summarized by @Arbetman2017-xj.
@@
 mendax*, which were not included in @Arbetman2017-xj, were obtained from @Obeso1992-yp and @Durieux2000-uk, respectively.
-
-\
 
@@
 For more nuanced definitions of preference, we refer the reader to @Beyer2010-vr and references therein.
-
-\
 
@@
 
-\
-
 We followed the procedure described above, simulating our entire dataset in one modeling process to infer aggregate preferences of bumble bee species for each floral morphotype.
 Using the results of this analysis, we modeled preferences for each floral morphotype as a function of bumble bee tongue length with hierarchical generalized additive modeling (HGAM), following the *type I* effect structure described by @Pedersen2019-kf.
-As a test of the *target use hypothesis*, we repeated the HGAM analyses described above using observed visitation rates rather than preference.
 To verify that our inference of preference was not sensitive to our decision to pool queen and worker visitation, we recalculated the null model and tongue length HGAM using only data from workers.
 Finally, we compared morphotype preferences across bumble bee species by performing a principal components analysis on the bumble bee species x floral morphotype preference matrix, and we mapped bumble bee tongue length onto the ordination space by fitting a 2-dimensional GAM surface to the first two principal components.
-
-\
 
@@
 Preference for each bumble- bee-morphotype pair was modeled together in a single HGAM with an interaction smooth between elevation and time.
-We also included bee-morphotype pair as a fixed itnercept effect and the interactions between bee-morphotype pair and site and bee-morphotype pair and year as random effects.
-
-\
+We also included bee-morphotype pair as a fixed intercept effect and the interactions between bee-morphotype pair and site and bee-morphotype pair and year as random effects.
 
@@
 
-\
-
 All data processing and analyses were conducted in `R` [@RCoreTeam-2021aa].
@@
 Annotated R code and complete output of our analytical workflow is available in Appendix 1.
-
-\
 
@@
 
-\
-
 ## 3.2 Bumble bee preferences toward floral morphotypes
@@
 The omission of queens from the calculation of overall preference did not yield qualitatively different results (Appendix 1).
-
-\
 
@@
 
-\
-
-Differences between floral morphotypes in terms of observed visitation rates, reflecting the target use hypothesis, were less pronounced **(Figure 2B)**.
-Bell, lip, flag, and head flowers were visited at high rates by one or more bumble bee species, but observed visitation rates did not differ strongly across morphotypes.
-
-\
-
 ## 3.3 Tongue length as a predictor of floral choice
 
-Tongue length was a strong predictor of preference but not of observed visitation rates **(Figure 3)**.
+Tongue length was a strong predictor of preference **(Figure 3)**.
 Preference toward funnel, bell, lip, flag, and head flowers exhibited significant GAM curves in response to bumble bee tongue length, and the model explained 77% of total deviance (adj.
@@
 Preference toward flag flowers exhibited a unimodal peak among medium-tongued species.
-In contrast, observed visitation rates were predicted by tongue length only for head flowers, which exhibited a weak decline in preference from a peak among short-tongued species.
-The visitation rate model explained only 25% of total deviance (adj.
-R^2^ = 0.053).
 Again, these results did not differ when preference was based on the worker-only null model (Appendix 1).
 
-\
-
 PCA of bumble bee floral preferences captured more than 80% of total variance in the first two axes **(Figure 4)**, and the GAM surface fit to the firs two axes explained most of the variance in tongue length (deviance explained = 94.7%, adj.
-R^2^ = 0.893, p < 0.001), reflecting the importance of trait matching between tongue length and floral morphotype as a mechanistic basis of preference.
+R^2^ = 0.893, p < 0.001), reflecting the importance of compatibility between tongue length and floral morphotype as a mechanistic basis of preference.
 Longer tongue length was associated with deeper and more closed flower morphotypes (lip, funnel, flag), while shorter tongue length was associated with more shallow and open flower morphologies (disc, stalk-disc, bell, head).
@@
 psithyrus* for head flowers is especially salient.
-
-\
 
@@
 R^2^ = 0.554).
-
-\
 
@@
 
-\
-
 # 4. DISCUSSION
 
-Using ~13,000 bumble- bee-wildflower interactions, we characterized bumble bee floral selection in terms of generalized floral morphotypes and found evidence that bumble bee floral selection depends both on the directed process of morphotype preference and the neutral effect of floral abundance.
-Bumble bee species differed in their floral morphotype preferences, and these differences could be explained by trait-matching between floral morphotype and bumble bee tongue length.
+Bumble bee species differed in their floral morphotype preferences, and these differences could be explained by bumble bee tongue length.
 These preferences, however, varied by elevation and season, likely representing adaptive responses of bumble bee foraging behavior to elevational and seasonal shifts in floral resource composition.
-
-\
-
-## 4.1 Floral abundance, floral selection, and the *target use hypothesis*
-
-Our finding that bumble bee tongue length explained preference (i.e.
-null model residuals) much better than it explained observed visitation rates indicates that, in our study system, abundance and preference function jointly in bumble bee floral selection.
-This casts doubt on the *target use hypothesis* as a general rule of bumble bee foraging, though it is possible that it still obtains under certain circumstances [e.g.
-@Harmon-Threatt2017-bz].
-In general, though, our findings support the hypothesis that bumble bee resource selection involves both the neutral effects of floral abundance and the directed effect of preference, meaning that analytical techniques designed to disentangle these effects, such as the null modeling method we employ, should be applied in studies of bumble bee foraging.
-
-\
-
-## 4.2 Community-level patterns of morphotype preferences
 
@@
 We saw similarly high visitation rates toward head flowers, but our null-model analysis reveals that while head flowers were highly visited, they tended to be under-visited relative to their abundance.
-
-\
 
@@
 A limitation of our study is that we could not reliably distinguish pollen- and nectar-foraging bumble bees, and it is likely that this distinction would shed further light on the mechanisms underlying the selection patterns we observed.
-
-\
 
@@
 Moreover, since floral color can be correlated with floral morphology [@Teras1976-pi], it is possible that some of the preference attributed to morphology in our models could represent a latent effect of floral color.
-
-\
 
@@
 
-\
-
 When considering community-level patterns of preference and visitation from the perspective of pollination services to mountain flora, it is important to note that while bumble bees are the dominant Hymenopteran pollinators in our study system, flies (Diptera) are also abundant flower visitors in mountain ecosystems, and their visitation is mainly restricted to the open flower morphotypes (e.g.
@@
 
-\
-
-## 4.3 Variability of floral abundance, visitation, and preference through elevation and time
-
 Intraspecific variation of bumble bee foraging preferences has been demonstrated under laboratory conditions [@Ings2009-vc; @Maharaj2019-wm]; our study reveals the *in situ* variation of preference for multiple bumble bee species through elevational and temporal gradients of floristic turnover.
-
-\
 
@@
 Under this hypothesis, an organism's resource selection is the product of exogenous constraint rather than endogenous preference.
-
-\
 
@@
 Mitesser kindly provided a high-performance computing environment to expedite null modeling.
+Two anonymous reviewers provided constructive feedback during manuscript revision.
 
@@
 
-**Figure 2:** Preference (A) and observed visitation rate (B) of the bumble bee species toward each floral morphotype.
-Each point represents the preference/visitation of one bumble bee species toward the corresponding floral morphotype.
+**Figure 2:** Preference of the bumble bee species toward each floral morphotype.
+Each point represents the preference of one bumble bee species toward the corresponding floral morphotype.
 The color of each point in panel A, retained from Figure 2, indicates whether preference is significantly (p < 0.01) positive or negative with respect to null model simulations.
@@
 
-**Figure 3:** Preference (A) and observed visitation rates (B) for each morphotype regressed against bumble bee tongue length.
-Each point represents the preference/visitation of one bumble bee species toward the corresponding floral morphotype.
-Curves depict fitted GAM smooths with 95\\% confidence bands.
+**Figure 3:** Preference for each morphotype regressed against bumble bee tongue length.
+Each point represents the preference of one bumble bee species toward the corresponding floral morphotype.
+Curves depict fitted GAM smooths with 95% confidence bands.
 Solid lines indicate curves differing significantly (p < 0.05) from a flat line while dashed lines indicated nonsigificant curves.

### 2021-11-10 ms_1/response_to_reviewers.Rmd
push to cloud just in case

---
title: "Response to reviewer comments"
output: pdf_document
# bibliography: 
  # - references.bib
csl: ecology.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Editor's summary

## Summary

Both reviewers liked many aspects of the manuscript but also had strong suggestions for improving it, especially for clarifying methods and the Introduction.  The two reviewers different on how to approach the introduction and I find myself falling on the side of Reviewer 2. Both I and Reviewer 2 (perhaps stated a bit differently) agree that the framing of the story needs to be rethought some.  Reviewer 2 sees it as a more ecological story of plant community response driving pattern of bumble bee interactions. I agree that the story is really more about patterns along a gradient rather than a paper about climate change and potential responses. This will mean rethinking the introduction and how the paper is set up, but I think will make it a more robust match to the data and still use the strong analytical approaches you have.

The reviewers felt that the methods need more clarification in several places. I felt overall you did a solid job of explaining what are some relatively novel approaches and these show promise for other similar types of data from different systems.

Below I have a several general broader comments/suggestions and then also detail specific comments chronologically from my own review.

## General comments

The set up of the introduction leads with climate change, elevational shifts etc. however, the data set and results are not really about this, but rather about how differential responses of flowering vegetation communities versus bees across an environmental gradient can produce thresholds in interaction pattern along such gradients.  I agree the could be (likely will be) implications for communities’ responses to climate change.  However, because the data do not look at shifts and changing distributions the framing of the introduction is mismatched.  The data on ecological gradients, niches, and emergent interaction beta-diversity seem plenty solid to me and I think recasting with that as the main story would be more in line with the data, then moving on to consider possible implications for climate change. The climate change implications could be developed mostly in the discussion.

**COMMENT**
A related point, I very much like the points about elevational gradients and mountains as valuable “ecological theaters”. There must be other ecological situations where similar gradients and discontinuities exist: coastal intertidal, estuary salinity gradients? If you were able to find other examples, I think this would broaden the relevance with the general ecological audience of Ecology and strengthen the paper.

**RESPONSE**
Thank you for your constructive critique of our framing. We have rewritten our introduction accordingly, focusing on the intrinsic ecology of environmental gradients rather than climate-induced range shifts. We agree that this is both a more natural framing of our data and a richer ecological story.

**COMMENT**
In the methods I wanted more detail of the sampling over sites and time- you could include a table of effort among dates and years etc. for the sites, or you could provide more detail in the text.

**RESPONSE**
In our original submission, we placed the detailed description of sampling patterns in Appendix S1. Based on your feedback, we have moved the description of sampling patterns into the main text, summarized in Figure 2.

**COMMENT**
In looking at Figure 2, if I am interpreting the figure correctly the lines are modeled standardized abundance for each species over the elevation gradient. The patterns of peak bumble bee abundance among panels looks consistent with a generally uniform turnover of species across elevation.  Its almost like a sort of elevational niche partitioning. Isn’t this something Pyke wrote about for Rocky Mountain bumble bees.  This seems like a good point to bring up and discuss. I am not sure how closely the point ties to Figure 4, but it seems consistent with the smooth line for bumble bees in that figure.

**RESPONSE**
We have modified both the visualization (now Figure 5) and discussion of our bumble bee abundance model to more clearly depict the differences in elevational responses across species and explore its potential causes.

A paragraph in the revised discussion reads:

"The response of overall bumble bee abundance to elevation was approximately flat, but this aggregate pattern concealed divergent species-level responses. The strongly positive responses of B. mendax, B. monticola, B. mucidus, B. pyrenaeus are consistent with the known affinity of these species for high-elevation habitats (Rasmont and Iserbyt 2010-2014). At the other extreme, B. pascuorum and the B. psithyrus species group exhibited strongly negative responses to elevation. In the case of B. pascuorum, this pattern is likely driven by its exceptionally long nesting cycle, which limits its fitness in the short growing seasons of higher elevations (Neumayer 2009). Growing season length may also explain the lower abundance of B. psithyrus at high elevation, since cuckoo bumble bees must emerge later than the queens of their host species and then successfully usurp host nests before they can begin to raise brood (Goulson 2010). Nevertheless, while the divergent responses of bumble bee species to elevation are informative, it is important to emphasize the overall robustness of bumble bees to elevational variation. Nine out of the 16 species in our study system did not respond significantly to elevation, and 8 out of 16 were observed across the full span of the gradient."

And we extend our discussion of these patterns to explore the question of why bubmle bees forage in resource-poor high-elevation habitats:

"The poverty of floral resources at the highest elevations raises the question of why bumble bees choose to forage at such elevations at all. For some species, the selection of high elevation habitats may be influenced by thermal tolerances (Oyen et al. 2016), but that fails to explain the presence of elevation-generalist species (including the majority of the species in our study system) in resource-poor high-elevation habitats. It is possible that nest site limitation at lower elevations drives some bumble bees to establish nests at high elevation, and their ability to forage downslope is subsequently constrained by their flight range, which may be reduced due to the energetic costs of flight in cold environments (Kovac et al. 2015). It is also possible that high elevation landscapes, while sparse in absolute floral resources, offer an escape from more competitive conditions at lower elevations where forager density is higher (Hoiss et al. 2015, Classen et al. 2020). Similarly, higher elevations could also offer an escape from density-dependent parasite and pathogen pressure (Parsche and Lattorff 2018). Finally, it is worth speculating that an affinity for open, high-elevation habitats might be an evolutionary legacy in mountain bumble bees, since the existence of flower-rich meadows in the forest matrix below the tree line is largely an artifact of human land use arising mainly in the last 500 years of bumble bees’ evolutionary history (Hejcman et al. 2013)."

## Specific comments

**COMMENT**:
Line 22  add “many of which”  somewhere around associated with mountains to describe bumble bees

**RESPONSE**:
Our revised abstract no longer contains this wording.

**COMMENT**
Line 33-34 this wording is awkward “to demonstrate”- doesn’t seem like quite the right word choice.

**RESPONSE**
Our revised abstract no longer contains this wording.

**COMMENT**
Line 46 abstract translations – wording here again seems odd.

**RESPONSE**
Our revised introduction no longer contains this wording.

**COMMENT**
Line 47-49 break into two sentences  this would improve clarity.  New sentence after the Blois et al citation. “The elevational gradients also involve variation in abiotic conditions other than temperature.”

**RESPONSE**
Our revised introduction no longer contains this wording.

**COMMENT**:
Line 60 delete species after cold-adapted.  As worded it implies that you are viewing bumble bees as a species.

**RESPONSE**
Section now reads: "Today, they remain characteristically cold-adapted organisms (Heinrich and Esch 1994)"

**COMMENT**
Line 71 replace the with their after hinge upon and delete of bumble bees after elevational distributions.

**RESPONSE**
Our revised introduction no longer contains this wording.

**COMMENT**
Line 80 I suggest that you spell out more clearly what you mean by geometric considerations, perhaps “area contractions”?

**RESPONSE**
Our revised introduction no longer contains this wording.

### Methods

**COMMENT**
I was struck by potentially unequal effort among sites.  I recognize that sampling within a site visit was consistent but different numbers of visits to sites within and among years would presumably create potential bias in detection among species and affect measures of abundance and richness.  These in turn could affect b-diversity. It would be helpful to see some coverage of this point including a reference to why it is not an issue (if indeed it isn’t).

**RESPONSE**
As described above, we have moved the description of sampling patterns into the main text, summarized in Figure 2. We account for the potential effect of sampling intensity (i.e. the total number of sampling transects per site) on observed beta-diversity by including sampling intensity as a covariate in our GDM analysis. The relevant line of the methods section reads:

"To control for potential confounding effects of sampling intensity, sampling days per site was included as a covariate."

**COMMENT**
Line 138 suggest changing sparsity of observations to sparse observations

**RESPONSE**
Our revised method section no longer contains this wording.

**COMMENT**
Lines 141-144.  Rarity alone doesn’t seem like a justification for omission.  These taxa would have little weight in such an assessment and so don’t need removing for that reason alone. If the issue was due to model fitting or something then I would suggest justifying it what way.  Otherwise, it seems odd to prioritize their inclusion in beta-diversity calculation but not elsewhere.

**RESPONSE**
In our revised manuscript, all models have been refit to include the previously omitted species. The main reason for omitting them in our initial analysis was that it was difficult to model species-specific floral resource availability for rare species. During our revision, we reconsidered the value of that analysis, since species did not exhibit strong differences in floral resource patterns anyway, and it added considerable complexity to our models. In our revised analysis, we model only aggregate floral resource availability, and we include all bumble bee species in our models of bumble bee richness, abundance, and beta-diversity.

Note that we have also opted to model peak rather than mean floral abundance. This decision did not change the resulting patterns very much, but we think as a matter of principle that modeling the mean of a log-normal distribution is problematic, and peak abundance seems a more natural summary for an essentially unimodal phenological process.

**COMMENT**
Line 143  write out genus name at the start of sentence.

**RESPONSE**
Done.

**COMMENT**
Line 182 This seems like a cool thing to look at and a powerful tool (although I am not really familiar with the approach).  It made me wonder why something similar wasn’t done to look at overall patterns of abundance and diversity for bees and flowers.

**RESPONSE**
In our revised manuscript, we include models of flower and bumble bee species richness. We opt for GAM analysis of these patterns, though, rather than GDM analysis. We chose GDM analysis of beta-diversity because beta-diversity is inherently a measure of change, and therefore best characterized in terms of rates.

**COMMENT**
One note about the beta-div partitioning analysis.  I believe that bi-partite package now has a way to do this quantitatively (new paper out by Jochen Freund), something you could consider, but I was also fine with it as is. You probably have seen this too.

**RESPONSE**
The paper you reference describes the function (and settings thereof) that we used in our analysis. The paper may have just come out, but it seems the method has been implemented in `bipartite` since the 2.14 release in early 2020.

**COMMENT**
Line 241 grazing in American English? I expect it would get changed in copy editing.

**RESPONSE**
We have replaced "grasing" with "grazing" throughout the manuscript.

**COMMENT**
Line 282 This sentence is awkward because its hard to know what you found differences in =peak abundance.  Also the next sentence is a pretty interesting statement, but I don’t think is quantitatively explored in the paper.  It would strengthen the argument about the special relevance of the tree line to do so. We expect species that are decline and those that are increase in abundance to have overlap in the middle, so not surprising, but your claim is interesting and more specific. Is there a way to assess if the tree line is indeed a particular zone.  It seems tricky to isolate because species actually appear to peak at different elevations so the slopes of their relative abundance will be different at each point in elevation.  Perhaps if the tree line was a point of highest absolute value in slopes among species or something?

**RESPONSE**
During revisions, we reconsidered our initial emphasis on the tree line ecotone, and we concluded that it was over-extended in the case of bumble bees (though not in the case of plants). In our revised manuscript, we do not describe the tree line ecotone as a special zone of turnover for bumble bees, and such an interpretation is not supported in our revised GDM analysis. We do, however, note that the tree line ecotone seems to be a significant peak of *B. pratorum* abundance. This section now reads:

"At the species level, bumble bees exhibited variable and sometimes divergent responses to elevation (Figure 5; Table 3). Bombus pyrenaeus, B. mendax, B. monticola, and B. mucidus exhibited approximately linear increases with elevation. In contrast, Bombus psithyrus and B. pascuorum exhibited approximately linear decreases with elevation. Bombus pratorum was unique in exhibiting a strongly unimodal abundance pattern with a well-defined peak between 1500-1600 m, approximately at the tree line ecotone, similar to the abundance pattern of B. pratorum described by Iserbyt et al. (2008) in the Eyne Valley of the Pyrenees. The remaining species did not respond significantly to elevation (p > 0.05)."

**COMMENT**
Line 290.  You say its notable that for some species one peak of flower resources falls outside their elevational zone.  Why would we expect it to be otherwise?  It would only be otherwise if for some reason bumble bee species were driving the distribution of their floral hosts.  Can you develop the point more clearly?

**RESPONSE**
In our revised manuscript, we decided to drop the species-specific model of floral resource availability, so this statement no longer appears in the paper.

**COMMENT**
Line 320 This is an interesting result but I wonder if it could be anything otherwise, so that the story is really about an elevational shift in plant distribution and turnover with bumble bees interactions following.  The bees themselves do NOT actually that obviously follow the same sort of discontinuity.  The paper sort of implies they do and that this tree line is a key place, but as I mentioned in the comment (line 282) that is not really demonstrated from abundance distributions (only 1 of 12 species shows such a pattern) and would be good to somehow quantify in the case of the bumble bees themselves.

**RESPONSE**
As described in our response to a comment above, we have stepped back from the emphasis on tree-line effect with respect to bumble bees. As you point out, it is actually more interesting to note the *differences* in response between plants and bumble bees, and how change in the latter are best understood as responses to changes in the former. This change of perspective is reflected in this paragraph of our discussion section:

"The roughly sigmoidal pattern of bumble bee beta-diversity rate in response to elevation, revealed by our GDM analysis, probably reflects both the reciprocal abundance patterns of high- and low-elevation species and the occurrence of certain rarer species only at middle-elevation sites. As discussed above, though, bumble bee turnover was low over all, demonstrating the robustness of bumble bees to the environmental variation encompassed by our elevation gradient. Floral beta-diversity rate also exhibited a sigmoidal pattern, and its inflection point coincides with the tree line ecotone at ~1500 m. The floristic importance of the tree line has been noted in previous work (e.g. Pellissier et al. 2010, Descombes et al. 2017), and our study corroborates these findings specifically for the subset of flora visited by bumble bees. Interestingly, the tree line did not constitute a discrete transition to a new equilibrium of species composition, but rather a threshold marking the beginning of a process of beta-diversity acceleration that continued for the remainder of the elevation gradient. While the most obvious effect of the tree line may be the relatively discrete transition from forest to scrub/grassland, the tree line also initiates a process of angiosperm species attrition that proceeds at a rate of about 40 species per 100 m (Körner 2003), and the accelerated beta-diversity we observed above the tree line mirrors simultaneous patterns of decreasing floral richness and abundance (Figures 3B, 4B)."

...and we pick up the again when we discuss the implications of our findings for the response of mountain bumble bees and wildflowers to climate and land use change:

"Mountains are unique ecological theaters in which the extrinsic forces of climate and land use change interact with intrinsic environmental gradation to shape the ecology of species and their interactions (Peters et al. 2019). Given the elevational patterns of richness, abundance, and beta-diversity among bumble bees and wildflowers in our study, what can be said about the response of these communities, and their functional relationships, to climate and land use change? If the process of mid-elevation pasture abandonment continues, the flower-rich meadows that now perforate the forest matrix will eventually close into flower-poor coniferous forest, and bumble bees will be pushed either into the more agriculturally intensified land at lower elevation or the relatively resource-poor open habitat above the tree line. This land- use-driven redistribution of bumble bees will interact with the concomitant process of climate warming, which has already generated measurable upslope shifts in some bumble bees species (Ploquin et al. 2013, Pyke et al. 2016, Marshall et al. 2020), sometimes resulting in increased species richness at higher elevations (Fourcade et al. 2019). Under this scenario, high-elevation bumble bees can be expected to face increased densities of low-elevation or elevation-generalist species, and the introduction of competing bumble bee species in a given habitat can have strong effects both on the fitness of their congeners and on patterns of floral visitation (Nagamitsu et al. 2010, Ishii 2013, Brosi and Briggs 2013). Nevertheless, the tolerance of bumble bees for a relatively broad climate envelope together with their long flight range may enable them to cope with climate and land use change better than their floral mutualists. The flowering plant community in our study system was strongly structured by elevation, and there is evidence that the upslope movement of plants in response to warming temperatures may outpace that of bumble bees (Marshall et al. 2020, though see Pyke et al. 2016). Moreover, it is the plant community that experiences the direct effects of land use change. Thus, it may be that the strongest effects of climate and land use change on mountain bumble bees will be indirect effects mediated by changes in the floral community."

**COMMENT**
Lines 327. This is an interesting point, but doesn’t seem unique to bumble bees.  Presumably other generalist bee species face similar issues in montane areas as would other mountain species with varied resources.  This would be worth addressing.  Also it seems a stretch to attribute this montane effect to bumble bees in general unless the speculation is that mountain environments have played a key role in the evolution of bumble bees as a group.  Are there data to support that point?

**RESPONSE**
Mountains have played a central role in the origin and diversification of *Bombus*, so we think it is wortwhile to point out the behavioral challenge that this entails. To strengthen this point, we have added to our introduction an expanded discussion of the role of mountains in bumble bee evolution:

"Bumble bees are believed to have arisen in the mountains of Asia during the cooling climate of the Eocene-Oligocene boundary, and their subsequent spread and diversification along mountain corridors and into lowland habitats was likely driven by alternating range expansion during cooling periods and retreat to higher elevations during warming periods (Hines 2008, Stewart et al. 2010, Martinet et al. 2018). Today, they remain characteristically cold-adapted organisms (Heinrich and Esch 1994), with peak abundance and diversity in mountain ranges and northern latitudes (Williams 1998). In these habitats, they are often the dominant guild of flower-visiting insects and the principal pollinators of entomophilous flora (Goulson 2010). As in the past, so now are mountain habitats critical theaters of ecological change for bumble bees. Warming climates are expected to cause habitat loss and potentially novel competitive pressures (Kerr et al. 2015, Soroye et al. 2020, though see Guzman et al. 2021). At the same time, socioeconomic drivers are causing the reforestation of pastures on mountain slopes and the intensification of agriculture in lowlands, both of which entail a loss of floral resources for bumble bees (Tasser and Tappeiner 2002, Tappeiner et al. 2006, Tasser et al. 2007, Walcher et al. 2017)."

**COMMENT**
The development of the implications of the research for climate change works better here in the discussion where it builds from the core findings, rather than as it is currently in the Introduction where it seems to motivate the research, but really isn’t targeted to address elevational shifts of bees, plants, etc.

**RESPONSE**
In the revised manuscript, our discussion of climate (and land use) change has been shifted mostly to the penultimate paragraph of the discussion (beginning with "Mountains are unique ecological theaters . . ."), which we quote in response to a previous comment.

\newpage

# Reviewer 1

## Summary

Here, the authors describe a unique study examining changes in bumble bee and plant community structure associated with an elevational gradient in SE Germany. They find evidence for relatively strong threshold effects of montane tree line on the abundance and composition (beta-diversity) of both plant, pollinator communities, as well as their interactions - suggesting that this region will be of importance given climate-change induced range shifts. The paper is very well written and analyzed. Although I am relatively novel user of the generalized modeling approach, the results appear robust and careful discussed. My suggesting are primarily limited to expanding some of the methodological details surrounding plant surveys, and would like to challenge the authors on one choice they make in the analysis.

The manuscript is very well-written, concise, and easy to follow. Only in a couple of spots do I recommend some tweaks.

## General comments

**COMMENT**
The analysis was very clever given the original design of the study may not have had this analysis in mind (looks to have been more about grazing impacts?)

**RESPONSE**
The elevation gradient was always the focus of this project, but management correlates with elevation in our study system (higher meadows have tended to be abandoned due to their lower accessibility), so it was necessary to keep track of management and account for it in our analysis. As it turns out, management does not seem to be a major driver of the responses we modeled, so it was significant in some cases. We have made this more explicit by visualizing the intercept effects of management along with the smooth effects of elevation and by including tabular output for all our models in the main text.

**COMMENT**
In reading the floral survey methods, I was left with an unclear idea of how exactly you measured the cover of floral resources in a 60x60m area. Was it a visual estimate of the entire area? Based on smaller subsets/quadrats? Were all floral species accounted for? Same for the bee transect walks - a bit more detail (meandering transect? 50m linear? etc...)

**RESPONSE**

We have added some details to the description of our sampling methods, which now reads:

"Sampling in each year of the study consisted of repeated visits to each site at approximately weekly intervals. Samples were only collected during periods without rain and when the air temperature was at least 6°C. We divided each 60 x 60 m study plot into 10 evenly-spaced 60 x 6 m transects. During each visit, the observer walked each transect in a zig-zag pattern for five minutes and recorded all interactions between bumble bees and flowers; thus, each sampling event consisted of a total of 50 minutes of observation. Bumblebees observed on a given flower were counted as floral visitors. Bumble bee queens were identified to species level in the field, while workers and males were stored in individually labeled tubes in the freezer for later identification in the laboratory according to (Amiet 1996). Floral visitation by male bumble bees was recorded during visitation sampling, but we chose to analyze only visitation by queens and workers. Males are often found resting on flowers, particularly late in the year, so their presence on a flower cannot necessarily be interpreted as a foraging visit. In conjunction with bumble bee observations, we estimated the flower cover of each herbaceous or shrubby plant species within the study plot to the nearest 0.01 m2; this was done by exhaustive visual inspection of the study plot rather than timed transect walks. To avoid observer bias, floral cover estimates were always performed by the same investigator (K. Kallnik). Plant species were identified following Lauber and Wagner (2007) and Oberndorfer (2001)."

**COMMENT**
In the analyses, total floral resource availability was calculated for each bumble bee species using only the visited floral genera. While I understand the authors logic for choosing this, I am left curious as to whether the results would change if the analyses was conducted on the entire set of floral resources? Also, in the method used, it wasn't clear to me whether the flower cover of genera used was site-specific, or if the genera used were for all sites for a given bumble bee species?

**RESPONSE**
During the process of revision, we decided that the small differences in species-level floral resource availability did not justify the added modeling complexity of treating floral resources in a species-specific manner. In our revised manuscript, we model only aggregate floral resource availability (though limited to the subset of floral species that were visited at least once).

**COMMENT**
The introduction is concise as it stands, but I feel that the two sections could be better integrated. I understand the focus on the tree line and use of bumble bees as a model system, but I feel that the two sections are disjointed relative to one another and could use a more effective transition to link them.

**RESPONSE**
We have rewritten the introduction significantly in our revised manuscript, and it no longer has the order of sections referred to in this comment.

**COMMENT**
Title: fine as is, but might benefit from a mention of climate? I know that wasn't directly tested here, but perhaps hinting of the climatic importance of this ecotone would help draw in readers interested in climate.

**RESPONSE**
Following the recommendation of the editor and Reviewer 2, we have downplayed climate in the framing of the paper, so we opt not to mention climate in the title. We have, however, changed the title to reflect the new framing of the paper, and it now reads: "The elevational structure of a mountain mutualism: richness, abundance, and beta-diversity of bumble bees and their floral hosts"

## Specific comments

**COMMENT**
Line 114: the sentence about male visitation is vague - they can still facilitate pollen movement despite not actively foraging on pollen. Perhaps say something more to the effect of "their foraging behavior is not directly comparable to queens and workers"

**RESPONSE**
This section now reads:

"Floral visitation by male bumble bees was recorded during visitation sampling, but we chose to analyze only visitation by queens and workers. Males are often found resting on flowers, particularly late in the year, so their presence on a flower cannot be interpreted reliably as a foraging visit."

While it is true that resting on flowers could still effect pollination, the emphasis of our study is on flowers as resources and bumble bees as foragers.

**COMMENT**
Line 147: Plants recorded during visitation but not during surveys makes me further long for how the plant surveys were conducted within the 60x60 m areas. How would these be missed? Were the 50min bee transect walks meandering or along a single defined path?

**RESPONSE**
As mentioned in a response to a previous comment, we have clarified the description of our sampling methods. The few cases in which a species was missed in a floral survey but observed in a bumble bee survey are probably attributable simply to observation error. In some cases, this may have been due to judgments made about observations occurring near the edge of the study plot. In any case, the rate of disagreement in our study was very low, and there is no reason to think it biases our inference.

**COMMENT**
Line 155: Management - this seems like perhaps it was the original purpose of data collection? It sort of stands out now as an oddity, so perhaps including some language that meadows were in different states of management therefore that state was included to account for any potential effects.

**RESPONSE**
In our revised manuscript, we discuss land use at greater depth. In the introduction, we note that:

"socioeconomic drivers are causing the reforestation of pastures on mountain slopes and the intensification of agriculture in lowlands, both of which entail a loss of floral resources for bumble bees (Tasser and Tappeiner 2002, Tappeiner et al. 2006, Tasser et al. 2007, Walcher et al. 2017)."

In the methods (also found in the original manuscript):
"Fourteen of these pastures are extensively grazed by cattle or sheep, 3 are mowed for hay production, and 8 have been abandoned throughout the last century and are no longer subject to any human management."

As mentioned in a response above, we treat the effects of management more explicitly in our revised results section. Then we pick up the theme again in the discussion section:

"And in the discussion:"If the process of mid-elevation pasture abandonment continues, the flower-rich meadows that now perforate the forest matrix will eventually close into flower-poor coniferous forest, and bumble bees will be pushed either into the more agriculturally intensified land at lower elevation or the relatively resource-poor open habitat above the tree line. This land- use-driven redistribution of bumble bees will interact with the concomitant process of climate warming, which has already generated measurable upslope shifts in some bumble bees species (Ploquin et al. 2013, Pyke et al. 2016, Marshall et al. 2020), sometimes resulting in increased species richness at higher elevations (Fourcade et al. 2019). Under this scenario, high-elevation bumble bees can be expected to face increased densities of low-elevation or elevation-generalist species, and the introduction of competing bumble bee species in a given habitat can have strong effects both on the fitness of their congeners and on patterns of floral visitation (Nagamitsu et al. 2010, Ishii 2013, Brosi and Briggs 2013). Nevertheless, the tolerance of bumble bees for a relatively broad climate envelope together with their long flight range may enable them to cope with climate and land use change better than their floral mutualists. The flowering plant community in our study system was strongly structured by elevation, and there is evidence that the upslope movement of plants in response to warming temperatures may outpace that of bumble bees (Marshall et al. 2020, though see Pyke et al. 2016). Moreover, it is the plant community that experiences the direct effects of land use change. Thus, it may be that the strongest effects of climate and land use change on mountain bumble bees will be indirect effects mediated by changes in the floral community."

**COMMENT**
Line 160: Year as a random term? Below n=5 threshold for number of random groups - I might suggest trying to model this as a fixed effect instead.

**RESPONSE**
In our revised analysis, we have respecified our models with year as a fixed term.

**COMMENT**
Line 229: Grasing? I think this should be grazing?

**RESPONSE**
We have replaced "grasing" with "grazing" throughout the manuscript.

**COMMENT**
Line 246 (and related to Fig 3A): I might change/add a parenthetical to fig. 3a axis that says "(percent species turnover)" as it's not immediately obvious (to me) what scale Beta diversity is on here.

**RESPONSE**
We have clarified in the figure caption that beta-diversity is depicted in terms of proportional species/interaction turnover.

**COMMENT**
Line 252: Instead of saying "most widely separated sites" I might suggest saying most vertically separated - widely to me implies 2-dimensional spatial separation, whereas here it's specially talking about elevation in the third dimension. If authors choose to change this, do so throughout the results/discussion.

**RESPONSE**
Agreed. Throughout the discussion, we have changed "widely separated" to "vertically distant."

**COMMENT**
Line 330: Super cool idea that excited me!

**RESPONSE**
In response to Reviewer 2's comment on this section, we decided to omit this speculation about the drivers of intelligence in bumble bees.

**COMMENT**
Line 359-360: Want to push back on this a bit: Certainly climate induced range shifts could have strong impacts along bumble bee species at elevation, but why disproportionate relative to those that may occur in lowland elevations? These species might not have any capacity to move upslope to escape hostile thermal/climatological conditions while at least those at elevation might be able to escape to some degree? To me, this read like those at elevation would be most impacted by climate, and I don't know if I necessarily agree with that...

**RESPONSE**
Our revised manuscript no longer focuses so exclusively on the tree line effect, and it no longer contains the language referred to in this comment.

\newpage

# Reviewer 2

## Summary

I had the chance to review the manuscript “The tree line ecotone punctuates the abundance and B-diversity of plants, pollinators, and their interactions” for consideration in the journal Ecology. Overall, I thought it was an interesting and very well-written paper. I think the manuscript could be improved by more strongly connecting the results from the different analyses to one another, as the story seems to me to be a more straightforward one than described – plant communities change drastically at tree line.

## General comments

**COMMENT**
The structure of the analyses with respect to how the authors used per-species bumble bee floral preference and how that interacts with the floral community, and interaction turnover. Does mean floral cover predict maximum bumble bee abundance for each species, separate from elevation?

**RESPONSE**
As described above, we have dropped the species-level analysis of floral resource availability. While there is an apparent correlation between bumble bee abundance and flower abundance in our data, it is not significant when elevation is included as a covariate. Since this question is not central to our paper, we opt not to include an additional model in our analysis.

## Specific comments

**COMMENT**
- Line 158: It isn’t immediately clear to me what “weighted per bumble bee species” means, and whether it interacts with the species random intercept.

**RESPONSE**
As described above, we no longer include the species-weighted model of foral abundance.

**COMMENT**
- Lines 231-244: Isn’t clear to me from this paragraph what “floral resource respond[ing] to elevation for bumble bee species” means, that is, what the analysis is specifically testing? Is this related to the species floral preference quantification, so change in abundance of floral species that each bumble bee species prefers as dependent on elevation? Please clarify.

**RESPONSE**
This language no longer appears in our revised manuscript.

**COMMENT**
Address in more detail the relationship between floral vs. bee species turnover and interaction species turnover, and the lack of turnover in the bee communities.

**RESPONSE**
Our revised discussion section treats this topic in much more detail:

"The roughly sigmoidal pattern of bumble bee beta-diversity rate in response to elevation, revealed by our GDM analysis, probably reflects both the reciprocal abundance patterns of high- and low-elevation species and the occurrence of certain rarer species only at middle-elevation sites. As discussed above, though, bumble bee turnover was low over all, demonstrating the robustness of bumble bees to the environmental variation encompassed by our elevation gradient. Floral beta-diversity rate also exhibited a sigmoidal pattern, and its inflection point coincides with the tree line ecotone at ~1500 m. The floristic importance of the tree line has been noted in previous work (e.g. Pellissier et al. 2010, Descombes et al. 2017), and our study corroborates these findings specifically for the subset of flora visited by bumble bees. Interestingly, the tree line did not constitute a discrete transition to a new equilibrium of species composition, but rather a threshold marking the beginning of a process of beta-diversity acceleration that continued for the remainder of the elevation gradient. While the most obvious effect of the tree line may be the relatively discrete transition from forest to scrub/grassland, the tree line also initiates a process of angiosperm species attrition that proceeds at a rate of about 40 species per 100 m (Körner 2003), and the accelerated beta-diversity we observed above the tree line mirrors simultaneous patterns of decreasing floral richness and abundance (Figures 3B, 4B).

In contrast, the rate of interaction beta-diversity exhibited a late and very pronounced acceleration at ~1800 m, indicating that the two highest sites at 1984 m and 2032 m differed strongly in interaction composition from the sites ~200 m below them. At these highest sites, the growing season is very short and vegetation is sparse, giving way to bare rock and scree landscapes, and it may be that the last 200 m interval of our elevation gradient imposes the most stringent environmental filter. It is interesting that interaction beta-diversity did not exhibit the inflection point at the tree line ecotone (~1500 m) that we saw for floral beta-diversity, despite the fact that overall interaction beta-diversity was driven primarily by the turnover of the floral community. A possible explanation for this pattern is that there could be an elevational lag in bumble bee floral selection such that most bumble bee species do not switch to supra-tree-line floral species when these species first appear, but rather maintain associations with flora whose ranges span the tree line until the extreme species turnover between 1800 and 2000 m forces them to adopt new floral hosts."

**COMMENT**
- Lines 263-264: Interaction beta diversity changes seem very tightly tied to floral changes?

**RESPONSE**
See response above, but note that the interpretation is more complex as a result of a correction made to our GDM analysis.

**COMMENT**
- Lines 285-287: But results for thresholds states this is not the case? (see lines 267-268). I would like to see the authors more directly address their results for changing abundance patterns for bumble bees vs. lack of species turnover for bumble bees.

**RESPONSE**
As discussed in responses to earlier comments, we have revised our GDM analysis and downplayed our initial emphasis on the tree line effect, except in the case of floral beta-diversity where the effect still holds. The original langue referred to in this comment no longer appear in our revised manuscript.

**COMMENT**
- Lines 320-321: This topic is important and unmentioned until here, maybe incorporate mention of it earlier e.g., first paragraph of discussion? Would also appreciate more discussion of the results of this study specifically, and the sampled bumble bee species’ biology/ecology as it relates to floral preference and abundance gradients.

**RESPONSE**
Our revised discussion section emphasizes the contrast between the elevational sensitivity of plants and the elevation robustness of bumble bees throughout. Before the referenced section, we note:

"Nevertheless, while the divergent responses of bumble bee species to elevation are informative, it is important to emphasize the overall robustness of bumble bees to elevational variation. Nine out of the 16 species in our study system did not respond significantly to elevation, and 8 out of 16 were observed across the full span of the gradient."

And later in the discussion we return to the theme:

"It is interesting that interaction beta-diversity did not exhibit the inflection point at the tree line ecotone (~1500 m) that we saw for floral beta-diversity, despite the fact that overall interaction beta-diversity was driven primarily by the turnover of the floral community. A possible explanation for this pattern is that there could be an elevational lag in bumble bee floral selection such that most bumble bee species do not switch to supra-tree-line floral species when these species first appear, but rather maintain associations with flora whose ranges span the tree line until the extreme species turnover between 1800 and 2000 m forces them to adopt new floral hosts."

"Nevertheless, the tolerance of bumble bees for a relatively broad climate envelope together with their long flight range may enable them to cope with climate and land use change better than their floral mutualists. The flowering plant community in our study system was strongly structured by elevation, and there is evidence that the upslope movement of plants in response to warming temperatures may outpace that of bumble bees (Marshall et al. 2020, though see Pyke et al. 2016). Moreover, it is the plant community that experiences the direct effects of land use change. Thus, it may be that the strongest effects of climate and land use change on mountain bumble bees will be indirect effects mediated by changes in the floral community."

**COMMENT**
Line 42: Citation formatting issue

**RESPONSE**
Fixed.

**COMMENT**
Lines 46-47: I find this sentence difficult to follow, as it is not immediately clear to me what “abstract translations along the linear lapse of temperature” means

**RESPONSE**
Our revised introduction no longer contains this wording.

**COMMENT**
Line 60: See recent paper Guzman et al. 2021 Biol. Conserv. critiquing methods in Soroye et al 2020, analytical biases suggest that their climate analyses require revisiting

**RESPONSE**
We have read the paper in question (which was very interesting), and we have added "though see Guzman et al. (2021)" to the in-text citation of Soroye et al. (2020). The details of this controversy, though, are beyond the scope of our paper.

**COMMENT**
Lines 65-67: Not all of these papers document upslope movement

**RESPONSE** 
The citations in this sentence originally referred not just to the claim of range shifts but to the claims of expected declines and extinctions. But this section has been rewritten in the revised introduction, and the comment no longer applies.

**COMMENT**
Lines 72-77: Long sentence, split into two

**RESPONSE**
Our revised introduction no longer contains this wording.

**COMMENT**
Line 95, 101,112, 129, 188, 215: Stylistic preference – use of passive voice throughout could be changed to active voice: e.g., was conducted, were classified, was recorded, was quantified, were assessed, were observed to be

**RESPONSE**
We have adopted a more active voice in our revised methods.

**COMMENT**
Lines 108-109: Include more detail re: the type of transect used, e.g., size, location within site, how was each sampled specifically?

**RESPONSE**
As mentioned in response to previous comments, we have added the relevant details the description of our sampling methods.

**COMMENT**
Line 112:  Can remove “after” (“later, after” repeated)

**RESPONSE**
"After" in this sentence was meant in the sense of "following" or "according to", referring to the identification key subsequently cited. But we agree that this could be more clearly worded, so we have replaced "after" with "according to."

**COMMENT**
Lines 143, 207, 218, 232: Stylistic preference --- I find it easier to follow if, for scientific species names, you start sentence each sentence with the full genus Bombus, can be difficult to read with shortened B.

**RESPONSE**
Agreed. We have made this change throughout the manuscript.

**COMMENT**
Line 155: Could you explain what “as a smoothing factor” means?

**RESPONSE**
We have replaced the term "smoothing factor" with the more explicit wording "a group-level smoother for each bumble bee species". This follows the terminology of @Pedersen2019-kf, and we refer the reader to their paper for a fuller description of HGAM methodology.

**COMMENT**
Line 160-162: Why these distributions?

**RESPONSE**
Richness (count data) and abundance are both zero-bounded, and therefore cannot (in most cases) be adequately modeled with the assumption of gaussian error distribution. The quasipoisson and negative binomial families used in our analyses are standard ways of dealing with this problem while accounting for the frequent issue of overdispersion when a simple poisson distirbution is assumed.

**COMMENT**
Line 178: What does unpartitioned mean in this context?

**RESPONSE**
"Unpartitioned" means total interaction B-diversity rather than one of the partitions thereof (Figure 3). Since this section is immediately preceded by a description of B-diversity partitioning, we think it is helpful to clarify that the interaction B-diversity analyzed using GDM is the total, unpartitioned interaction B-diversity.

**COMMENT**
Line 180: “Difference in a predictor variables” = variable (typo)

**RESPONSE**
Fixed.

**COMMENT**
Lines 206, 215: Are these the same data? Differ by 1 genera

**RESPONSE**
These are actually not the same data. The "metaweb" refers to the visitation data, which included 110 visited genera. The 111 genera mention in line 215 are the genera found during floral surveys but not necessarily visited.

**COMMENT**
Line 229, 241, 343, 345: Grazing

**RESPONSE**
We have replaced "grasing" with "grazing" throughout the manuscript.

**COMMENT**
Line 229-230: Per-species abundance didn’t change over time, then?

**RESPONSE**
In our revised models, which differ in their handling of annual variation, we saw significantly higher bumble bee abundance in 2011 than in 2010 or 2012.

**COMMENT**
Line 241: What is the reference level?

**RESPONSE**
In general, when a factor is included in a linear model as a fixed effect, one of its levels is treated as a baseline against which the other levels are compared. In our model, grazing was the baseline/reference level of the management variable.

**COMMENT**
Lines 270-273: I’m not familiar with p~0, maybe incorporate lines 554-555 in figure caption into text?

**RESPONSE**
We adopted this notation because the p-value actually given by the model was 0, but it is better to understand it as some very small number close to 0. A true zero p-value is ordinarily impossible, but the GDM package uses a permutation procedure to test significance, and it is possible for a permutation to return 0 cases of a simulated value more extreme than an observed value, thus yielding a p-value of 0. In the revised manuscript, though, we use < 0.001 instead of ~0.

**COMMENT**
Lines 328-331: Seems a bit outside your scope

**RESPONSE**
After considering the matter --- and despite R1's interest in this section --- we agree that this proposition probably stretches our inference too far. We have removed the lines in question.

**COMMENT**
Line 528: “Elevation an dmanagement” = and management (typo)

**RESPONSE**
Fixed.

**COMMENT**
Separate Fig 4B into a table

**RESPONSE**
Done

### 2021-11-10 ms_2/response_letter.Rmd
push to cloud just in case

---
title: "Response Letter"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary of changes

We thank our subject editor and reviewers for their constructive feedback on our manuscript. Our revision retains the core findings described in our original manuscript, but adopts a modified framing based on the comments we received. In particular, the subject editor and Reviewer 2 questioned whether our analysis can really be used as a comparison of the target-use and use-availability hypotheses of bumble bee resource selection. While we believe the logic of this comparison was sound, we acknowledge that it was weakened by the mechanistically vague connection between tongue length and floral morphotype, and we agree that a rigorous comparison of these hypotheses would require a different analytical approach (and, ideally, a different study design) than ours. Accordingly, we have rewritten our manuscript to avoid making explicit claims about the target-use or use-availability hypotheses and instead focus on interpreting the preferences we observed with respect to questions of competition and coexistence in mountain bumble bee communities. Other changes were focused mainly on clarifying our Methods section by describing our sampling methods more explicitly and explaining our use of Kugler's morphotypology to summarize trait variation in the floral community. Point-by-point responses to reviewer comments are detailed below.

# Response to reviewer comments

## Subject Editor

### Editor's summary
This is an interesting paper using a large dataset to ask about bumblebee preferences, and how those are structured by altitude and phenology. Both reviewers are overall happy with the methods, but rise some questions that are needed to clarify the paper. Please pay special attention to the question by R2 regarding the test to differentiate the target use hypothesis. While I think the conclusions are sound (the preference analysis shows strong preferences, but not completely deterministic, and hence plant abundance is likely to play a role too), I don't follow the rationale on how the comparison with observed visitation rates rather than preference tests for this hypothesis. Separating the contribution of neutral process (i.e. abundances) and preference (i.e. trait matching) should be tested in more detail, or alternatively, the authors can put less emphasis on testing the target use hypothesis and focus on the preferences and how they change across species, altitudes, and time periods.

### Response
Based on the feedback we received, we have rewritten our manuscript to avoid making explicit claims about the target use hypothesis. Instead, we put a stronger emphasis on the ecological consequences of resource preference, i.e. the relationship between diet overlap and species coexistence. This revision retains the core findings of our study, which both reviewers found compelling, but contextualizes them in a more modest framing that acknowledges the concerns our reviewers expressed.

## Reviewer 1

COMMENT: The description of how plant cover was estimated, especially at such a fine scale, is rather sparse. At lines 91-94, the main explanation is to look at “Sponsler et al.” but this manuscript is only submitted (and not included in references or otherwise attached), so there’s no way to evaluate if one thinks this is a fair way to estimate plant cover…because one cannot determine how it was measured. Even if Sponsler et al is published imminently, I still think at least a brief description (belt transect? Quadrats? Visual estimates? Whatever) should be included and then details can either be referenced in a supplement or external paper as long as its available.

RESPONSE: We have revised the description of our methods to be more complete, and we have omitted the reference to an unpublished manuscript. The relevant section now reads:

"We recorded bumble- bee-wildflower interactions during repeated visits to each study site at approximately weekly intervals. Sampling was only conducted during periods without rain and with an air temperature of at least 6°C. Each 60 x 60 m study plot was divided into 10 evenly-spaced 60 x 6 m transects. During each visit, the observer walked each transect in a zig-zag pattern for five minutes and recorded all bumble bee visits to flowers, resulting in a total of 50 minutes of observation [...] In conjunction with each bumble bee visitation sample, we visually estimated the flower cover of herbaceous and shrubby plant species within each 60 x 60 m study plot to the nearest 0.01 m^2^ and identified the plant species according to Lauber (2007) and Oberndorfer (2001). Floral cover estimation was performed by exhaustive inspection of each study plot rather than timed transect walks, and the same investigator (K. Kallnik) performed all floral surveys to avoid observer bias."

COMMENT: The manuscript does sort of assume that tongue length is the only thing driving bumble bee preference for flowers. Although this is a well-established premise, I think discussion on bumble bee selection of plants based on nutrition (see various studies by Anthony Vaudo) or phylogenetically conserved foraging preferences among subgenera (Wood et al. 2021 Journal of Applied Ecology “Global patterns in bumble bee pollen collection show phylogenetic conservation of diet”) is warranted. Perhaps it would fit nicely in section 4.1 (but anywhere, really). There is some brief discussion of colour preferences, but I think the work on protein:lipid ratios and phylogenetically conserved preferences for plant families is more well supported.

RESPONSE: This is a good point, we have added a sentence to the discussion to acknowledge the role of pollen nutrition and phylogenetic conservatism:

"Bumble bees floral choice is also influenced by pollen nutritional parameters (Vaudo et al. 2016a, Vaudo et al. 2016b), and floral preferences may be conserved in phylogenetic lineages within Bombus (Wood et al. 2021)."

COMMENT: L153-162 – I wonder if it induces any sort of bias into this analysis by selecting only the most common plants and animals? Perhaps just a brief justification here is warranted.

RESPONSE: As we note in the text, this decision was made "to ensure sufficient data points through elevation and time." It is hard to fit complex surface models with sparse data points, and the results are not likely to be trustworthy. We also wanted to limit the number of models to be presented and interpreted, since Figure 5 would quickly become unreadable with the addition of more rows or columns. It seemed natural, therefore, to focus our analysis on the most important floral morphotypes and the most abundant bumble bee species.

COMMENT: Figure 1: are the plots ordered in any particular manner? Is there a reason they’re non-alphabetical? It doesn’t really matter much, but wondering if there’s a reason for their order. Ah, after seeing Figure 3 it seems like they are in order of longest to shortest tongue length. I suggest reversing order to be consistent between figures

RESPONSE: We have reordered Figure 1 accordingly.

COMMENT: Figure legend 3 contains “95\%”

RESPONSE: Fixed.

COMMENT: For figures I would suggest just spelling out species names and using the names as indicated in Figure 1B for the floral morphs rather than numbers. It’s hard to keep track of the numbers and the numbers don’t seem to be meaningful (i.e. not ordered) as far as I can tell. Perhaps in Figure 4 the species names can be italicized and the floral morph names in plain font to distinguish

RESPONSE: This is a decision we wrestled with in the original preparation of our figures. In the end, we decided to stick with abbreviations for bumble bee species because in some cases printing the whole species name would result in cluttered plots or too-small fonts, and we thought it was best to adopt a consistent approach for all figures. In the case of Figure 4, we opted to use morphotype numbers rather than names to avoid cluttering the ordination space and to make a qualitative distinction between the names of bumble bee species and those of floral morphotypes. We realize that these decisions do not optimize all criteria of data visualization, but in our opinion, this was the best compromise.

COMMENT: I think the significance statement for figure 5 and 6 are reversed? There are no “dashed contour lines” in figure 6. Only in figure 5 are there dashed (i.e. nonsignificant) lines.

RESPONSE: In fact, the entire captions were reverse. Thanks for catching this.

COMMENT: L80-82 – if these management regimes are not tested, might be good to explicitly say so when mentioning them here

RESPONSE: We have removed the description of management at our sites, since it is not relevant to the analyses we perform.

COMMENT: L228 – “firs” -> “first”

RESPONSE: Corrected.

COMMENT: L274 – “applies” instead of “obtains”?

RESPONSE: This section no longer occurs in our manuscript.

## Reviewer 2

COMMENT: First of all, it is not elaborated by the authors why bumble bee floral morphotype preferences would be effected by their proboscis length. Apparently, this was their a-priori hypothesis, but clear aims of the research and/or hypotheses that are tested with the data are missing in the introduction. While the authors outline three questions regarding resource selection by bumble bees and its response to climate change at the end of the introduction, there is little scientific justification presented in the manuscript that enables the reader to connect individual findings with specific hypotheses. It would make sense to me that bumble bees would prefer those morphotypes on which they can forage most efficiently, and this might differ depending on their tong length. Foraging efficiency or bumble bee preference might also be related to floral traits such as nectar tube depth and nectar amount, or handling time which may correlate with floral morphotype.

RESPONSE: The reviewer's question about the relationship between tongue length and floral morphotype is a good one, and we have added a paragraph to our Methods section to clarify the logic behind our decision:

"Kugler’s typology was informed by laboratory studies of bumble bee flower choice (Kugler 1943), and there is a precedent for analyzing bumble bee visitation in terms of Kugler’s morphotypes (Farnberger and Paulus 1996, Neumayer and Paulus 1999, Schneller et al. 2014). It must be noted, though, that the use of a morphotypology in lieu of quantitative data on corolla depth (which are lacking for most species in our system) precludes strong statements about trait-matching, since there is no quantitative measure of fitness between a given morphotype and bumble bee tongue length. Nevertheless, the morphotypology we use captures qualitatively much of the information concerning the accessibility of floral rewards that would otherwise require quantitative measurement not only of corolla depth, but also of floral shape, openness, etc. Thus, it is reasonable to expect a functional link between morphotype and tongue length, even if this link is not mechanistically quantifiable."

We point out also that the point of our study was not to establish the role of tongue length in bumble bee floral choice, since this is already well know, but to use tongue length as means of explaining functionally the patterns of floral choice that we observed.

As to the matter of clarifying hypotheses, our revised manuscript includes a substantially rewritten introduction that lays out the purposes of our paper more clearly. Strictly speaking, this is not a matter of stating hypotheses, since our study is empirical in scope. In our revised introduction, we emphasize the role of floral selection in determining species coexistence under conditions of floral resource limitation, and we then lay out the questions concerning floral selection that we answer with our study:

"In this study, we explore the floral morphotype preferences of a mountain bumble bee community using ~13,000 records of bumble bee visits to wildflowers collected over three years along a 1400 m elevation gradient. First, we use null model analysis to compare floral morphotypes in terms of community-level preferences, identifying morphotypes that are consistently over- or under-visited relative to their abundance. Next, we compare morphotype preferences across bumble bee species and model species-level preferences as a function of tongue length. Then, we analyze preference similarity across species to identify species among which competition is most likely under conditions of resource limitation. Finally, we explore the flexibility of floral preferences in bumble bees by modeling the variation of morphotype preferences through elevation and time."

COMMENT: No data are provided regarding the plant species and their traits besides floral morphotype although they might be insightful for the interpretation of the results.

RESPONSE: We acknowledge this point, but our goal in this paper has been interpretable generalization. There is surely much more that could be learned by considering plant species individually or by generating quantitative data for traits relevant to bumble bee foraging, but that work is beyond the scope of our study. These limitations are acknowledged in the Discussion section, where we discuss information not represented by Kugler's morphotypology:

"Such unexplained variation in preference may indicate that traits not captured by Kugler’s (1970) morphotypology are also important in governing floral selection by bumble bees. Floral color, in particular, is known to be involved in the process of flower detection by foraging bumble bees (Spaethe et al. 2001), and some bumble bee species exhibit marked color preferences (e.g. Teräs 1976, Neumayer and Paulus 1999, Raine and Chittka 2007). Bumble bees floral choice is also influenced by pollen nutritional parameters (Vaudo et al. 2016a, Vaudo et al. 2016b), and floral preferences may be conserved in phylogenetic lineages within Bombus (Wood et al. 2021)."

COMMENT: Secondly, my main concern lies with another important aspect of their study. They use the established relationship between tongue length and floral preference as a point of reference, to test the target-use hypothesis by comparing the explanatory power of bumble bee tongue length with respect to observed visitation rates vs. preference. They assume that under the target-use hypothesis, tongue length should explain observed visitation rates better than preference (null model residuals), while the opposite finding would indicate that, contrary to the target-use hypothesis, bumble bee floral selection is influenced by relative abundance. In my opinion, you cannot test these two hypotheses based on whether preference or visitation rate is stronger or significant related to bumble bee tong length. Rather, you should compare the expected visitation rate, based on the assumptions of model, with the observed observation rate, independent of tongue length.

RESPONSE: Our comparison between the target-use and use-availability hypotheses was based on logic that we believe to be sound. If resource selection is influenced by neutral effects of abundance, failing to account for these effects will create artifacts in the inference of the underlying selection process. Similarly, if resource selection is *not* influenced by abundance, accounting for abundance will introduce artifacts that obscure the selection process. The selection process is, therefore, most apparent when the correct causal model is employed. Given a known selection process, in this case one mediated by tongue length, two causal models can be evaluated by which most clearly recovers the known process. That said, we grant that this is an unconventional sort of "back door" approach to hypothesis testing, and it relies on a vague mechanistic relationship between tongue length and floral morphotype (discussed above). Since our results unambiguously support what is already the general consensus on pollinator resource selection (i.e. the use-availability model), we decided that little was lost by removing the formal comparison of the use-availability and target-use hypotheses. Instead, we simply present and interpret the results of our null model (representing the use-availability hypothesis) with respect to ecological questions of diet overlap and species coexistence.

COMMENT: Thirdly, the authors claim that by modelling the variation of bumble bee floral preference through time and elevation their results provide a first step toward understanding how climate and land use change might distort the temporal and elevational structure of bumble bee – wildflower mutualisms in mountain ecosystems. In my opinion, although they include sites under different land use management, they did not take this into account in their analyses, therefore, they cannot make any statements about how land use change might impact the structure of bumble bee – wildflower interactions based on their results.

RESPONSE: We invoke climate and land use change in the framing of our study because these nascent processes can be expected to alter competitive relationships between bumble bee species, thus warranting the study of floral preference and diet overlap. These general claims are supported in the text by relevant references, and we do not make any specific claims about the effects of climate or land use change based on our data.

COMMENT: A minor comment is that, in my opinion, the terminology trait matching is a bit confusing. Matching implies that something is a good fit. Like how tong length matches the depth of the flowers that pollinators visit. In this paper, the authors use it to describe that the strength of bumble bee preference for a certain floral morphotype is strongly predicted by the length of their tongue. This doesn’t have anything to do with a good morphological fit. A good fit could also be that bumble bees have a foraging benefit from it, e.g., forage more efficiently on those flowers that they more strongly prefer. However, this study did not investigate that.

RESPONSE: This is a legitimate point. We have changed the title to: "Floral preferences of mountain bumble bees are constrained by functional traits but flexible through elevation and season." We have also removed all explicit language of trait matching from our paper, except for where discuss directly the limitations of using a morphotypology in lieu of quantitative trait data.

COMMENT: Overall, I agree that there is much that we still don’t understand about (bumble) bee foraging behavior and floral preference, and specifically how community composition and elevation affects this behavior. This paper provides new insights but could make an even stronger contribution if more insight would be provided in species composition and their traits at different study sites and elevation.

RESPONSE: Again, we agree that there is much more that could be learned with a more granular analysis of our data. Such an analysis is beyond the scope of the present paper, but our data are publicly available to anyone who want to dig deeper.

COMMENT: Abstract could provide more details on the specific results.

RESPONSE: We have rewritten our abstract substantially, both to add more information and to reflect the changes in our framing.

COMMENT: L10-13. "The flexibility of bumble bee foraging will be tested in coming years as climate and land use change alter floral resource distribution in time and space, particularly in the mountain ecosystems that host the world's richest communities of bumble bees and their floral mutualists." In my opinion, it would make the abstract (and discussion) stronger if the authors can connect the insights from their study (that bumble bee preferences is flexible) to this given fact. What are the implications of this flexibility in light of climate and land-use change?

RESPONSE: Our revised manuscript no longer contains this exact language, but the idea is still there. Climate and land use change, with their effects on species distributions and floral resource availability, can be expected to alter competitive relationships between bumble bee species. Thus, it is relevant to understand the patterns of preference that will determine competition under conditions of resource limitation. We have laid out this framing more clearly in our revised Introduction, and we pick it up again in the Discussion. When it comes to the direct relationship between climate / land use change and our data, however, we are really limited to informed (and, we believe, interesting) speculation, since we are talking about long temporal processes, and our data provide only a temporal snapshot of bumble bee floral selection.

COMMENT: There are different papers that shows matching between pollinator tongue length and floral nectar tube depth such as:

Martina, M., Klinkhamer, P.G.L, Waser, N.M., Stang, I. & van der Meijden, E. (2009) Size-specific interaction patterns and size matching in a plant –pollinator interaction web. Annals of Botany 103: 1459–1469.

Klumpers, S.G.T., Stang, M. & Klinkhamer, P.G.L. (2019). Foraging efficiency and size matching in a plant–pollinator community: the importance of sugar content and tongue length. Ecology Letters, 22: 469–479. - Shows that (bumble) bee preference is related to size matching between proboscis length and nectar tube depth and driven by foraging efficiency.

RESPONSE: We realize that trait-matching is a general phenomenon in plant-pollinator networks, and the papers mentioned here are good examples. Our study is concerned, though, with the particular case of bumble bees, so we focus just on a few of the most pertinent classical studies.

COMMENT: L79-82. "We selected a total of 25 study sites (60 x 60 m) located on pastures along a 80 ~1400 m elevation gradient ranging from 641-2032 m above sea level (m.s.l.), differing in management type (grazed by cattle or sheep, mowed for hay production, or abandoned during the last century and are no longer subject to human management)." -It is unclear (1) how the different sites are distributed along the elevational gradient, (2) whether all three management types occur at different elevations and (3) if not, whether this is taken into account in the analyses. If not, there might be an additional effect of land use, rather than just elevation.

RESPONSE: It is true that elevation and land use are confounded in our study system, but we are not claiming an *effect* of elevation *per se*; we're simply describing elevational patterns that could be attributed to various processes. Elevation *per se* doesn't really have a direct mechanistic relation to what we studied. Perhaps an argument could be made that temperature, which varies more or less linearly with elevation, could have some effect on bumble bee floral choice, as we mention briefly in the Discussion, but it is probably more accurate to understand the relationship between elevation and bumble bee foraging as mediated by a host of other variables, such as land use, species composition, etc. So, in a nutshell, we do not attribute the elevational patterns we see to a direct effect of elevation, so it is not a problem that elevation is confounded with land use (and lots of other variables). The point is that whatever changes when you go upslope, bumble bee preferences are, to some degree, responsive to it.

COMMENT: L84-85. Bumble- bee-wildflower interactions were observed during repeated visits to each study site at approximately weekly intervals. -Please eleborate during which period of the year these observations took place? E.g. from april to september.

RESPONSE: We have added the following line to the Methods section: "Sampling at each site began upon snowmelt in the spring and continued until the effective end of the bumble bee colony cycle in early September, thus encompassing approximately the full foraging season."

COMMENT: L86-87. Bumble bees were recorded during one 50-minute transect walk per visit, and bumblebees observed on a given flower were counted as floral visitors. -Be aware that when you sample using transect walks, you sample according to flower abundance.

RESPONSE: We are not clear on the reviewer's meaning here, but we have expanded our Methods section to describe our sampling approach more explicitly:

"We recorded bumble- bee-wildflower interactions during repeated visits to each study site at approximately weekly intervals. Sampling at each site began upon snowmelt in the spring and continued until the effective end of the bumble bee colony cycle in early September, thus encompassing approximately the full foraging season. Sampling was only conducted during periods without rain and with an air temperature of at least 6°C. Each 60 x 60 m study plot was divided into 10 evenly-spaced 60 x 6 m transects. During each visit, the observer walked each transect in a zig-zag pattern for five minutes and recorded all bumble bee visits to flowers, resulting in a total of 50 minutes of observation."

COMMENT: Several times, the authors refer to manuscripts that are not published yet. E.g. line 94: see Sponsler et al. (submitted) for details of sampling design. However, keep in mind that reviewers nor readers have access to these manuscripts. Rather elaborate on the sampling design in the supplementary materials.

RESPONSE: We have removed all references to this unpublished manuscript from our Methods section.

COMMENT: Did the number of plant species per floral morphotype differ among morphotypes and at different elevations? If so, can you elaborate on how this may not have effected your results.

RESPONSE: This is certainly an interesting question from a floristic perspective, but it seems a bit tangential to our paper. We acknowledge the point, though: there could be a "sampling effect" (in the biodiversity / ecosystem function sense of the term) whereby a more species-rich morphotype is more likely to include an especially attractive species. This possibility is effectively encompassed in our discussion of the mechanisms that might account for the elevation*time variation we observed in morphotype preference, where we suggest that patterns could be driven by changes in the species composition of a given morphotype (or of background morphotypes).

More problematic with respect to our conclusions is that the overall patterns of preference could be influenced by sampling effects (i.e., that more diverse morphotypes appear more attractive just because they have a better chance of having a dominant species). To check this, we added a brief analysis to Appendix 1, and the results give no reason to think that our conclusions are influenced by sampling effects. In fact, the most species-rich morphotype was "disc" (group 1), and this was the least preferred in our analysis. "Flag" flowers (group 6), in contrast, were relatively species-poor, but one of the most highly preferred.

COMMENT: Visitation rate is not defined. Often visitation rate is the number of visits per flower, but I got the impression that the authors has recorded the number of visits in total to a certain plant species. If I understood it correctly, the number of flowers is not assessed but rather the flower cover of herbaceous and shrubby plant species within each 60 x 60 m study plot to the nearest 0.01 m 2 . In my opinion, the number of flowers would be a better indicator of flower abundance and using flower cover rather than number of flowers might have affected the results. I am not sure what you mean by flower cover but e.g., plant species might have the same number of flowers, but some species have their flowers arrange along a stalk while others have their flowers arranged in a disk. Species that have their flowers arranged in a disk will have a larger flower cover than species that have their flowers arranged along a stalk. Also the size of the flowers might affect the percentage of flower cover.

RESPONSE: Most of the explicit discussion of visitation rate is no longer in the manuscript, since it was attached to the comparison of the target-use and use-availability hypotheses. But we calculated visitation rate simply as proportional visitation, i.e. visits to morphotype X divided by total visits. As for methods of estimating floral abundance, the approach we chose was a designed to give a reasonable estimate while enabling us to work quickly enough to cover all our sites. One can imagine more precise methods, such as capillary tube sampling of nectar, but our results are based on the methods that were practicable in our study.

COMMENT: When you present R^2^ values, please also include p-value and DF or N

RESPONSE: This comment seems to be referring to our presentation of the GAM for preference ~ tongue length. The point of mentioning the R^2^ of the model is to highlight that the smooth effect of tongue length * morphotype along with the intercept effect of morphotype very effectively account the variation in our data. There is no model-level p-value to report, and such a p-value would be trivial anyway (p-values are seldom otherwise). We could report p-values for each group-level smoother, as we indicate in Figure 3, but this would change nothing about the interpretation of the model. The null hypothesis in a GAM is that the smooth response is perfectly flat (i.e. that effective degrees of freedom = 0), which is approximately the case for disc and stalk-disc flowers (which, accordingly, have p-values > 0.05) and quite plainly not the case for all other morphotypes (which have p-value < 0.05) (Figure 3). The interpretation of our model hinges not so much on the "significance" of the smooth effects but their shape, and their shape is *per se* a matter of statistical inference, as controlled by the adapting smoothing algorithm of the GAM. Degrees of freedom are also a bit complicated with GAMs and not ordinarily reported in the way that we are accustomed to seeing for other types of models.

In a nutshell, we think we have provided a sufficient description of the model in the text for the reader to understand Figure 3 and the conclusions that we draw from it. But the full analytical workflow is available in Appendix 1 for anyone who wants to dig deeper, including the standard tabular summary of the model. We would be happy to add this tabular summary to the main text if the editor thinks that would be more appropriate.

COMMENT: L191-293. "Visited floral species included all morphotypes except spike (8) and trap (10) flowers. Null (0) and brush (9) morphotypes were, however, excluded from analysis for the reasons given above." -If I understand this correctly, you mean that spike, trap, null and brush flowers were excluded from the analyses, either because these were not frequently visited or not at all. This is already mentioned in the method section. I don’t see the need to mention this here again.

RESPONSE: We have removed this redundancy.

COMMENT: L230-232. "Longer tongue length was associated with deeper and more closed flower morphotypes (lip, funnel, flag), while shorter tongue length was associated with more shallow and open flower morphologies (disc, stalk-disc, bell, head)." -In my opinion, you can not make the generalization that lip, funnel and flag flowers have deeper flowers compared to disc, stalk-disc, bell and head shaped flowers. I have measured the nectar tube of flowers of legume species which were shorter than those of certain Asteraceae species. I do agree that lip, funnel and flag flowers are more closed compared to disc, stalk-disc, bell and head flowers.

RESPONSE: We have removed the words "deeper" and "shallower" from the referenced sentence.

COMMENT: Somehow it is a bit unclear what it is so new in the paper compared to published literature. That should be clearer from the beginning. E.g., are there previous studies that looked at pollinator preference (for certain floral morphotypes) and how foraging behavior differs among an elevational gradient?

RESPONSE: We hope the significance of our study is clearer in the revised Introduction and Discussion. We have made an effort throughout the paper to cite all relevant literature; if there are specific studies the reviewer feels are missing, we are open to considering them.

COMMENT: I am missing a discussion on what the implications of your results are for other pollination research areas, such as network structure, evolution, plant reproduction.

RESPONSE: These are interesting questions, but I'm not sure we have answers to them.

COMMENT: L264-268. "Bumble bee species differed in their floral morphotype preferences, and these differences could be explained by trait-matching between floral morphotype and bumble bee tongue length. These preferences, however, varied by elevation and season, likely representing adaptive responses of bumble bee foraging behavior to elevational and
seasonal shifts in floral resource composition." -Can this not also be an response to different pressure of resource competition? Comparing the proboscis length and floral morphotype distributions at different elevations and time might be insightful.

RESPONSE: In the Discussion, we discuss a range of mechanisms that might account for the variation in preference we observed, including the possible role of competition.

COMMENT: "Our finding that bumble bee tongue length explained preference (i.e. null model residuals) much better than it explained observed visitation rates indicates that, in our study system, abundance and preference function jointly in bumble bee floral selection." -I agree that your results show that bumble bee preference and thus floral selection is driven by proboscis length. However, I am not convinced that this can be concluded from the fact that bumble bee tongue length explained preference better than observed visitation rate (see my comments above).

RESPONSE: We have removed this section from the revised manuscript.

COMMENT: "In general, though, our findings support the hypothesis that bumble bee resource selection involves both the neutral effects of floral abundance and the directed effect of preference meaning" … -awkward. You just wrote the same sentence above.

RESPONSE: We have removed this section from the revised manuscript.

COMMENT: "We found that the most preferred morphotypes were bell, lip, and flag flowers, which is broadly consistent with visitation patterns reported by previous studies in central Europe (Farnberger and Paulus 1996, Neu283 mayer and Paulus 1999, Schneller et al. 2014). These previous studies tended to find a stronger role of head flowers than indicated by our preference analysis, but this is probably because they characterized foraging in terms of simple visitation rates rather than null-model-inferred preference." -It would be interested to know whether these studies include the same plant species. That might also explain why they found a stronger role of flower heads.

RESPONSE: True. There is definitely substantial overlap in plant species across these studies (including ours), but also significant differences. Neumayer's work, for example, was conducted at much higher elevation than ours, so it included a somewhat different set of plants and bees. Further elaboration on these studies seems beyond the scope of our paper, though.

COMMENT: L305-319. -Although the authors discuss reasons for why different bumble bee species might prefer different flowers, as mentioned before, I miss a discussion on why bumble bees with different tong lengths would prefer different floral morphotypes. In my opinion, this might be related to the reward accessibility or quantity.

RESPONSE: We clarify this in our revised Methods section, as described earlier.

COMMENT: L317-319. If color might explain why bumble bee species have different floral preferences, are there any studies that show that color preference differ among bumble bee species?

RESPONSE: This question has not be studied in depth, but the references we cite in the text include some observational comparisons of flower visitation across bumble bee species on the basis of color.

COMMENT: L347-348. "Intraspecific variation of bumble bee foraging preferences has been demonstrated under laboratory conditions (Ings et al. 2009, Maharaj et al. 2019)." -Again it would be relevant to cite Klumpers et al. 2019 here as they have shown that interspecific variation in foraging preference is related to proboscis length: (bumble) bees prefer those flowers of which the nectar tube depth matches the length of their proboscis.

RESPONSE: The line in question refers to *intra*specific variation in foraging preference; we would prefer not to revisit the theme of *inter*specific variation on the basis of tongue length, since that has already been discussed in our paper.

COMMENT: L351-366. Insight in which plant and bumble bee species and their traits occur at different elevations/time will help to disentangle which mechanism underly variation in bumble bee floral preference along an elevational gradient. You collected this data and could have a look.

RESPONSE: Again, we agree that there is a wealth of information to be found by going through our data species by species, but that would be material for a new paper, not something to be tacked on to the end of this one.

COMMENT: L353-357. "… which could be caused by (1) shifts in species composition of the focal morphotype (e.g. the occurrence of a particularly rewarding flag flower at high-elevation sites), (2) shifts in the reward levels of individual species (e.g. increased nectar production in the flag flower Lotus corniculatus under high elevation conditions), or (3) changes in nutritional requirements of bumble bees (e.g. an increased need for concentrated nectar to support thermoregulation at colder ambient temperatures)." - Please include references.

RESPONSE: This sentence presents a set of logical possibilities, each supported by a hypothetical scenario; nothing we say depends on an external source. We see, though, that some confusion may have been introduced by our suggestion that *Lotus corniculatus* might vary in nectar production with elevation. This is a purely hypothetical scenario, stated in the same subjunctive mode as the rest of the sentence, but we have reworded that clause to be more general and therefore more clearly subjunctive: "(e.g. increased nectar production in a particular flag flower under high elevation conditions)."

COMMENT: L362-366. "For example, the increased selection of flag flowers at high elevation could be driven by competition release from B. pascuorum, which tends to be the most abundance bumble bee species at lower elevations but becomes relatively scarce above 1500 m (Sponsler et al, submitted). Under this hypothesis, an organism’s resource selection is the product of exogenous constraint rather than endogenous preference." -Relevant literature: Brosi, B.J., Niezgoda, K. & Briggs, H.M. (2017) Experimental species removals impact the architecture of pollination networks. Biol. Lett. 13: 20170243. Has experimentally shown that the removal of a bumble bee species change the foraging behavior of other species.

RESPONSE: We have added a reference to an earlier paper from these authors (Brosi and Briggs, 2013) that is more directly relevant to the case of bumble bees.

COMMENT: L371-375. "The dynamics of preference in our study suggest adaptive tuning of floral choice consistent with the ecological hurdles that have shaped the evolution of bumble bees in the mountain habitats in which they arose and diversified (Williams 1985, Hines 2008). Where steep temperature gradients and short growing seasons generate extreme spatiotemporal turnover of floral resources, adaptive flexibility of morphological preferences should be favored over fixed biases." -I am not convinced that the flexibility of bumble bee preference is an adaptation to the spatiotemporal turnover of floral resources. Although bumble bees indeed occur in mountainous habitats, characterized with short growing seasons generate extreme spatiotemporal turnover of floral resources, they also occur in habitats where there is not such an extreme spatiotemporal turnover of floral resources.

RESPONSE: We have softened this statement somewhat in our revised conclusion, which reads:

"Clarifying the patterns and processes of bumble floral selection sheds light on how bumble bee communities have evolved with respect to their mountain habitats and how they might respond to the incipient forcings of climate and land use change. Perhaps the most salient finding of our study is the remarkable marriage of generalism and selectivity in bumble bee foraging ecology. While most bumble bees (with the exception of true specialists like *B. gerstaeckeri*) can and do visit most floral morphotypes, they allocate their visitation according to strong preferences; yet, these preferences are themselves flexible, evidently adapting to changes in ecological context. Such flexible selectivity among generalist foragers might make enable bumble bee communities to adaptively partition resources under varying competitive conditions, which could make them relatively robust to new competitive pressures introduced by species range shifts. Moreover, the dynamics of preference in our study suggest adaptive tuning of floral choice consistent with the ecological hurdles that have shaped the evolution of bumble bees in the mountain habitats in which they arose and diversified (Williams 1985, Hines 2008). Where steep temperature gradients and short growing seasons generate extreme spatiotemporal turnover of floral resources, adaptive flexibility of morphological preferences should be favored over fixed biases."

COMMENT: L381. The word ‘both’ can be deleted.

RESPONSE: This line no longer occurs in our revised manuscript.

### 2021-11-10 code/ms2_nullmodels.Rmd
push to cloud just in case

@@
   geom_point(aes(color = Test)) +
-  facet_wrap(~fct_rev(bb.sp)) +
+  #facet_wrap(~fct_rev(bb.sp)) +
+  facet_wrap(~bb.sp) +
   scale_y_discrete(limits = rev) +
@@
   vegdist(method = "euclidean") %>%
-  hclust(method = "complete")
+  hclust(method = "ward.D2")
 

### 2021-11-10 ms_2/ms_2_oikos_part2_rev.Rmd
push to cloud just in case

@@
 ---
-title: "Floral preferences of mountain bumble bees are constrained by functional traits but flexible through elevation and season"
+title: Floral preferences of mountain bumble bees are constrained by functional traits
+  but flexible through elevation and season
 header-includes:
@@
 output:
+  #word_document: default
   pdf_document:
     keep_tex: yes
-bibliography: 
-  - references.bib
+bibliography: references.bib
 csl: oikos.csl
@@
 
-**KEYWORDS:** foraging, trait matching, pollinator, *Bombus*, niche, resource selection
+Patterns of resource use by animals can clarify how ecological communities have assembled in the past, how they currently function, and how they are likely to respond to future perturbations.
+Bumble bees (Hymentoptera: *Bombus* spp.) and their floral hosts provide a diverse yet tractable system in which to explore resource selection in the context of plant-pollinator networks.
+Under conditions of resource limitation, the coexistence of bumble bee species should depend on dietary niche overlap.
+In this study, we report patterns and dynamics of floral morphotype preferences in a mountain bumble bee community based on ~13,000 observations of bumble bee floral visits recorded along a 1400 m elevation gradient.
+We found that bumble bees are highly selective generalists, rarely visiting floral morphotypes at the rates predicted by their relative abundances.
+Preferences also differed markedly across bumble bee species, and these differences were well-explained by variation in bumble bee tongue length, generating patterns of preference similarity that should be expected to predict competition under conditions of resource limitation.
+Within species, though, morphotype preferences varied by elevation and season, possibly representing adaptive flexibility in response to the high elevational and seasonal turnover of mountain floral communities.
+Patterns of resource partitioning among bumble bee communities may determine which species can coexist under the altered distributions of bumble bees and their floral hosts caused by climate and land use change.
+
+**KEYWORDS:** foraging, trait matching, pollinator, *Bombus*, niche, resource selection, competition, coexistence
 
@@
 Resource selection by animals is instrumental in community assembly, species coexistence, and ecosystem function [@Chesson2000-aw; @Bascompte2007-ls].
-*Need another sentence in here* Understanding patterns of resource selection in biological communities is also key to anticipating the effects of climate and land use change, which can be expected to reorganize species interactions and niche relationships [@Tylianakis2008-fi].
+Understanding patterns of resource selection in biological communities is also key to anticipating the effects of climate and land use change, which can be expected to reorganize species interactions and niche relationships [@Tylianakis2008-fi].
 
@@
 The most important trait that distinguishes bumble bee species with respect to dietary niche is tongue length.
-Bumble bee species can differ markedly in tongue length, and floral handling time increases when tongue length and corolla depth are poorly matched [@Inouye1980-mv; @Harder1983-bk], leading to an association between tongue length and floral choice [e.g.
+Bumble bee species can differ markedly in tongue length, and floral handling time increases when tongue length and corolla depth are poorly matched [@Inouye1980-mv; @Harder1983-bk], leading to an association between tongue length and floral choice in bumble bee communities [e.g.
 @Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn].
-Under conditions of food limitation, the coexistence of bumble bee species should be limited by their similarity in tongue length, and hence in dietary niche.
+Under conditions of food limitation, the coexistence of bumble bee species should be limited by their similarity in tongue length, insofar as it is a proxy for dietary niche.
 Based on field studies in the Rocky Mountains, @Inouye1977-kl concluded that, assuming food limitation, four bumble bee species can be expected to persist in any given locality: a long-tongued species, a medium-tongue species, a short-tongued species, and a nectar-robber (i.e.
 a species that accesses nectaries by biting through the corolla, thus bypassing the ordinary constraint of tongue length).
-In many places, though, considerably more than four bumble bee species coexist, indicating either that food is not limiting or that competitive conditions are too variable in time and space to be detected at the scale of field observations [e.g.
+In many places, though, considerably more than four bumble bee species coexist, often with overlapping tongue lengths, indicating either that food is often not limiting or that competitive conditions can be too variable in time and space to be detected at the scale of field observations [e.g.
 @Ranta1981-vx; @Williams1989-vr; @Goulson2008-zy].
 
-The most species rich bumble bee communities are found in the mountainous regions of Eurasia [@Williams1998-xi; @Hines2008-er], where the predictions of competitive exclusion are rarely realized [@Ranta1981-vx; @Goulson2008-zy].
+The most species rich bumble bee communities are found in the mountainous regions of Eurasia in which *Bombus* initially arose and diversified [@Williams1998-xi; @Hines2008-er] and where the predictions of competitive exclusion are rarely realized [@Ranta1981-vx; @Goulson2008-zy].
 Nevertheless, climate and land use change in these regions have the potential to alter and intensify competitive relationships between bumble bee species.
@@
 The consequences of reshuffled species distributions and declining floral resources for the coexistence of mountain bumble bees will depend in part on how bumble bee species select and partition available flora.
+
 Under competitive conditions, species coexistence is possible when intraspecific competition exceeds interspecific competition, and this stabilizing inequality is favored to the extent that species differ in dietary niche [@Abrams1977-zp; @Hanski1995-kr; @Chesson2000-aw].
@@
 We selected a total of 25 study sites (60 x 60 m) located on pastures along a ~1400 m elevation gradient ranging from 641-2032 m above sea level (m.s.l.).
-Fourteen of these pastures are grazed by cattle or sheep, three are mowed for hay production, and eight have been abandoned during the last century and are no longer subject to human management.
 
 We recorded bumble- bee-wildflower interactions during repeated visits to each study site at approximately weekly intervals.
+Sampling at each site began upon snowmelt in the spring and continued until the effective end of the bumble bee colony cycle in early September, thus encompassing approximately the full foraging season.
 Sampling was only conducted during periods without rain and with an air temperature of at least 6°C.
@@
 
-## 2.2 Floral morphology and bumble bee tongue length
+## 2.2 Bumble bee tongue length and floral morphology
+
+Bumble bee tongue-length data (for the worker caste) were obtained from the literature values compiled and summarized by @Arbetman2017-xj.
+Values for *B.
+gerstaeckeri* and *B.
+mendax*, which were not included in @Arbetman2017-xj, were obtained from @Obeso1992-yp and @Durieux2000-uk, respectively.
 
 To summarize floral morphology, we relied on the synthesis of @Kugler1970-gb, who developed a typology consisting of ten primary morphotypes and nested subdivisions of each.
-Kugler’s typology was informed by laboratory studies of bumble bee flower choice [@Kugler1943-uy], and there is a precedent for analyzing bumble bee visitation in terms of Kugler’s morphotypes [@Farnberger1996-xo; @Neumayer1999-il; @Schneller2014-ry].
-Kugler morphotype classifications of the plants in our study system were accessed via the BIOFLOR database [@Klotz2002-uc].
 Morphotypes were simplified to the primary classes numbered 0-10: 0 = null (includes wind-pollinated flora such as grasses), 1 = disc- and bowl-shaped flowers, 2 = funnel flowers, 3 = bell-shaped flowers, 4 = stalk-disc flowers, 5 = lip flowers, 6 = flag flowers, 7 = flower heads, 8 = spike flowers, 9 = brush flowers, 10 = trap flowers **(Figure 1)**.
@@
 Plants belonging to morphotype 0 were typically not recorded during floral surveys due to their lack of conspicuous floral displays, and they were rarely visited by bumble bees, so this morphotype was also omitted from all analyses.
+Kugler morphotype classifications of the plants in our study system were accessed via the BIOFLOR database [@Klotz2002-uc].
 
-Bumble bee tongue-length data (for the worker caste) were obtained from the literature values compiled and summarized by @Arbetman2017-xj.
-Values for *B.
-gerstaeckeri* and *B.
-mendax*, which were not included in @Arbetman2017-xj, were obtained from @Obeso1992-yp and @Durieux2000-uk, respectively.
+Kugler’s typology was informed by laboratory studies of bumble bee flower choice [@Kugler1943-uy], and there is a precedent for analyzing bumble bee visitation in terms of Kugler’s morphotypes [@Farnberger1996-xo; @Neumayer1999-il; @Schneller2014-ry].
+It must be noted, though, that the use of a morphotypology in lieu of quantitative data on corolla depth (which are lacking for most species in our system) precludes strong statements about trait-matching, since there is no quantitative measure of fitness between a given morphotype and bumble bee tongue length.
+Nevertheless, the morphotypology we use captures qualitatively much of the information concerning the accessibility of floral rewards that would otherwise require quantitative measurement not only of corolla depth, but also of floral shape, openness, etc.
+Thus, it is reasonable to expect a functional link between morphotype and tongue length, even if this link is not mechanistically quantifiable.
 
@@
 A total of 13,293 bumble- bee-flower interactions were included in our final data set, comprising 15 bumble bee species, 163 plant species (110 genera, 37 families), and 730 unique bumble- bee-plant interaction pairs.
+Most sites hosted 8 or more species per year (Appendix 1), suggesting that interspecific competition was not the main driver of community assembly.
 *B.
@@
 Floral surveying yielded a total of 354 plant species (194 genera, 53 families).
-Visited floral species included all morphotypes except spike (8) and trap (10) flowers.
-Null (0) and brush (9) morphotypes were, however, excluded from analysis for the reasons given above.
 
@@
 
-PCA of bumble bee floral preferences captured more than 80% of total variance in the first two axes **(Figure 4)**, and the GAM surface fit to the firs two axes explained most of the variance in tongue length (deviance explained = 94.7%, adj.
+PCA of bumble bee floral preferences captured more than 80% of total variance in the first two axes **(Figure 4)**, and the GAM surface fit to the first two axes explained most of the variance in tongue length (deviance explained = 94.7%, adj.
 R^2^ = 0.893, p < 0.001), reflecting the importance of compatibility between tongue length and floral morphotype as a mechanistic basis of preference.
-Longer tongue length was associated with deeper and more closed flower morphotypes (lip, funnel, flag), while shorter tongue length was associated with more shallow and open flower morphologies (disc, stalk-disc, bell, head).
+Longer tongue length was associated with more closed flower morphotypes (lip, funnel, flag), while shorter tongue length was associated with more open flower morphologies (disc, stalk-disc, bell, head).
 The strong affinity of *B.
@@
 
-Bumble bee species differed in their floral morphotype preferences, and these differences could be explained by bumble bee tongue length.
-These preferences, however, varied by elevation and season, likely representing adaptive responses of bumble bee foraging behavior to elevational and seasonal shifts in floral resource composition.
+Based on complementary bumble bee flower visitation and floral abundance data, we used null model analysis to reveal patterns and dynamics of floral morphotype preference in a mountain bumble bee community.
+Our results underscore that bumble bees are highly selective generalists, rarely visiting floral morphotypes at the rates predicted by their relative abundances.
+Floral morphotype preferences also differed markedly across bumble bee species, and these differences were well-explained by variation in bumble bee tongue length.
+Ordination of species preferences corroborated the role of tongue length in explaining preference and revealed the resulting patterns of preference similarity that should be expected to predict competition under conditions of resource limitation.
+Within species, though, morphotype preferences varied by elevation and season, possibly representing adaptive flexibility of foraging behavior in response to the high elevational and seasonal turnover of mountain floral communities.
 
-We found that the most preferred morphotypes were bell, lip, and flag flowers, which is broadly consistent with visitation patterns reported by previous studies in central Europe [@Schneller2014-ry; @Neumayer1999-il; @Farnberger1996-xo], indicating that our findings are regionally generalizeable.
+The most preferred morphotypes in our system were bell, lip, and flag flowers, which is broadly consistent with visitation patterns reported by previous studies in central Europe [@Schneller2014-ry; @Neumayer1999-il; @Farnberger1996-xo].
 These previous studies tended to find a stronger role of head flowers than indicated by our preference analysis, but this is probably because they characterized foraging in terms of simple visitation rates rather than null-model-inferred preference.
 We saw similarly high visitation rates toward head flowers, but our null-model analysis reveals that while head flowers were highly visited, they tended to be under-visited relative to their abundance.
+When considering these community-level patterns of preference and visitation from the perspective of pollination services to mountain flora, it is important to note that while bumble bees are the dominant Hymenopteran pollinators in our study system, flies (Diptera) are also abundant flower visitors in mountain ecosystems, and their visitation is mainly restricted to the open flower morphotypes (e.g.
+disc flowers) that receive less attention from bumble bees [@Neumayer1999-il].
+In the context of our research question, we have emphasized the differences in tongue length between bumble bee species, but all bumble bees have relatively long tongues compared to the overall pollinator community, and it is not surprising that their foraging tends to be focused on the deeper and more closed flower morphotypes.
+The process of floral resource partitioning *among* bumble bee species is nested within the process of floral partitioning *between* bumble bees and other flower visitors.
 
@@
 Moreover, while average preference toward funnel and head flowers was low, each was strongly preferred by one or two bumble bee species.
-The majority of variation in preference toward floral morphotypes was explainable by bumble bee tongue length, with the more open bell and head morphotypes preferred by short-tongued bumble bee species and the deeper funnel, lip, and flag morphotypes preferred by medium- and long-tongued bumble bee species.
+The majority of variation in preference toward floral morphotypes could be explained by bumble bee tongue length, with the more open bell and head morphotypes preferred by short-tongued bumble bee species and the more restrictive funnel, lip, and flag morphotypes preferred by medium- and long-tongued bumble bee species.
 The failure of tongue length to explain preference toward disc and stalk-disc flowers reflects the fact that these flowers were almost unanimously ignored by the bumble bees in our study system.
@@
 @Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn].
+Moreover, our study shows that bumble bee tongue length can predict floral choice even when a simple morphotypology is used in lieu of quantitative floral traits, which could streamline the prediction of floral choice in systems where quantitative floral trait data is missing.
 It is important, though, to consider the species that appear as outliers with respect to the smooth dependency of preference on tongue length.
@@
 Such unexplained variation in preference may indicate that traits not captured by Kugler's (1970) morphotypology are also important in governing floral selection by bumble bees.
-Floral color, in particular, is known to be instrumental in the process of flower detection by foraging bumble bees [@Spaethe2001-bq], and some bumble bee species exhibit innate color preferences [@Raine2007-nq].
-This effect can also be seen in field studies, where bumble bees have been shown to over- or under-represent certain floral colors [e.g.
-@Teras1976-pi; @Neumayer1999-il].
-Moreover, since floral color can be correlated with floral morphology [@Teras1976-pi], it is possible that some of the preference attributed to morphology in our models could represent a latent effect of floral color.
+Floral color, in particular, is known to be involved in the process of flower detection by foraging bumble bees [@Spaethe2001-bq], and some bumble bee species exhibit marked color preferences [e.g.
+@Teras1976-pi; @Neumayer1999-il; @Raine2007-nq].
+Bumble bees floral choice is also influenced by pollen nutritional parameters [@Vaudo2016-fp; @Vaudo2016-ur], and floral preferences may be conserved in phylogenetic lineages within *Bombus* [@Wood2021-eo].
 
@@
 gerstaeckeri*, for example, is associated with lip and funnel flowers but eschews flag flowers, while *B.
-mucidus* does this opposite.
+mucidus* does the opposite.
 With respect to the question of climate-induced competition, the case of *B.
@@
 
-When considering community-level patterns of preference and visitation from the perspective of pollination services to mountain flora, it is important to note that while bumble bees are the dominant Hymenopteran pollinators in our study system, flies (Diptera) are also abundant flower visitors in mountain ecosystems, and their visitation is mainly restricted to the open flower morphotypes (e.g.
-disc flowers) that receive less attention from bumble bees [@Neumayer1999-il].
-In the context of our research question, we have emphasized the differences in tongue length between bumble bee species, but all bumble bees have relatively long tongues compared to the overall pollinator community, and it is not surprising that their foraging tends to be focused on the deeper and more closed flower morphotypes.
-The process of floral resource partitioning *among* bumble bee species is nested within the process of floral partitioning *between* bumble bees and other flower visitors.
-
 Intraspecific variation of bumble bee foraging preferences has been demonstrated under laboratory conditions [@Ings2009-vc; @Maharaj2019-wm]; our study reveals the *in situ* variation of preference for multiple bumble bee species through elevational and temporal gradients of floristic turnover.
-
 At least three mechanisms could underlie the variation of preferences we observed.
@@
 the occurrence of a particularly rewarding flag flower at high-elevation sites), (2) shifts in the reward levels of individual species (e.g.
-increased nectar production in the flag flower *Lotus corniculatus* under high elevation conditions), or (3) changes in nutritional requirements of bumble bees (e.g.
+increased nectar production in a particular flag flower under high elevation conditions), or (3) changes in nutritional requirements of bumble bees (e.g.
 an increased need for concentrated nectar to support thermoregulation at colder ambient temperatures).
@@
 pascuorum*, which tends to be the most abundance bumble bee species at lower elevations but becomes relatively scarce above 1500 m (Sponsler et al, submitted).
-Under this hypothesis, an organism's resource selection is the product of exogenous constraint rather than endogenous preference.
-
-## 4.4 Conclusions
+Under this hypothesis, an organism's resource selection is the product of exogenous constraint rather than endogenous preference [e.g.
+@Brosi2013-mu].
+As we have noted previously, the high numbers of coexisting species at our study sites suggest that interspecific competition is currently not the main driver of community assembly in our system, but competition is likely to intensify as flower-rich abandoned meadows revert to flower-poor coniferous forest.
 
 Clarifying the patterns and processes of bumble floral selection sheds light on how bumble bee communities have evolved with respect to their mountain habitats and how they might respond to the incipient forcings of climate and land use change.
-The dynamics of preference in our study suggest adaptive tuning of floral choice consistent with the ecological hurdles that have shaped the evolution of bumble bees in the mountain habitats in which they arose and diversified [@Williams1985-ql; @Hines2008-er].
+Perhaps the most salient finding of our study is the remarkable marriage of generalism and selectivity in bumble bee foraging ecology.
+While most bumble bees (with the exception of true specialists like *B.
+gerstaeckeri*) can and do visit most floral morphotypes, they allocate their visitation according to strong preferences; yet, these preferences are themselves flexible, evidently adapting to changes in ecological context.
+Such flexible selectivity among generalist foragers might make enable bumble bee communities to adaptively partition resources under varying competitive conditions, which could make them relatively robust to new competitive pressures introduced by species range shifts.
+Moreover, the dynamics of preference in our study suggest adaptive tuning of floral choice consistent with the ecological hurdles that have shaped the evolution of bumble bees in the mountain habitats in which they arose and diversified [@Williams1985-ql; @Hines2008-er].
 Where steep temperature gradients and short growing seasons generate extreme spatiotemporal turnover of floral resources, adaptive flexibility of morphological preferences should be favored over fixed biases.
-As climate and land use change drive the redistribution of bumble bees and their floral mutualists, and likely an overall reduction in floral availability in mountain habitats, competition within bumble bee communities is likely to intensify.
-Whether this leads to competitive exclusion will depend on whether interspecific competition exceeds the stabilizing effects of interspecific competition [@Hanski1995-kr].
-In this regard, the partial dependency of bumble bee floral selection on the neutral effects of floral abundance (in contrast to the *target use hypothesis*) can be expected to favor the coexistence of both bumble bee species and their floral host by increasing the likelihood that species will be self-limiting [@Valdovinos2021-ya].
-
-\
 
@@
 
-**Figure 5:** Variation in relative floral abundance through elevation and season modeled as HGAM surfaces.
+**Figure 5:** Variation in preference through elevation and time modeled as HGAM surfaces.
 Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns.
-All smooths are significant (p < 0.05).
+Significant (p < 0.05) smooths are depicted with solid contour lines and nonsignificant smooths with dashed contour lines.
 
@@
 
-**Figure 6:** Variation in preference through elevation and time modeled as HGAM surfaces.
+**Figure 6:** Variation in relative floral abundance through elevation and season modeled as HGAM surfaces.
 Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns.
-Significant (p < 0.05) smooths are depicted with solid contour lines and nonsignificant smooths with dashed contour lines.
+All smooths are significant (p < 0.05).

### 2022-01-05 ms_1/response_to_reviewers2.rmd
started final revision

---
title: "Manuscript ECY21-0389.R1: Contrasting patterns of richness, abundance, and turnover in mountain bumble bees and their floral hosts"
subtitle: "Response to reviewer comments"
author: 
  - Douglas B. Sponsler^1^*  
  - Fabrice Requier^2^
  - Katharina Kallnik^1^
  - Alice Classen^1^
  - A. Fabienne Maihoff^1^
  - Johanna Sieger^3^
  - Ingolf Steffan-Dewenter^1^
output: pdf_document
bibliography:
- references.bib
csl: ecology.csl
# editor_options: 
#   markdown: 
#     wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Editor's comments

**COMMENT**\
I found the updated manuscript a refreshingly improved one. Its message now matches the data and retains novelty. The reviewers both agree, and we all appreciated your clear responses to comments. For me, the discussion really covers the various points well. I might put slightly more emphasis in the discussion on how the pattern of interaction diversity is really driven by plants not bees, but the development of the beta-diversity results was clear and quite interesting.

**RESPONSE**

**COMMENT**\
One point in the Discussion. You have some nice phenological data (highlighted in current Figure 2). Do you think it provides enough information to speculate that the bees could be neophobic as higher elevation species come into flower? You more or less say this in and around line 383-386. There are some others who have suggested related ideas and might help in supporting your point. Ogilvie and Thomson write about this in a paper on patch/site switching in bumble bees. I also wrote about this in a paper on Osmia and host mixing (Williams and Tepedino 2003). "First, foragers might have been initially insensitive to developing changes in resource density, and might have delayed forsaking relatively unproductive patches for more rewarding ones until differences were more marked (Heinrich, 1976;Thomson, 1988, 1996; Thomson et al., 1982). Such delays have been attributed to the influence of long-term memory of resource location and quality (Thomson, 1996)." I am not suggesting mine is something to cite, but there were references within that might be. In any case there are others who have suggested potential effects that might support your idea, which I like.

**RESPONSE**\

**COMMENT**\
I have two strong suggestions with respect to the figures of which there are relatively many.

1.  One of these (Figure 2) I would encourage you to put into supplement. Although this figure presents results, the way you use it is more like supplementary methods. This will relieve some of the extra space and complexity of the figures (most of which are already multi-panel.
2.  In both Figure 3 and 4 you have the main panels (Figure 3 A, D, and G Figure 4A,D) These are critical and nicely presented. The remaining panels focus on Year differences which are potentially interesting to some, but not at all a focus of the paper. They are presented in passing) but the results are not develop and not mentioned at all in the Discussion. I think you would better keep the reader's focus and probably strengthen the overall presentation by moving them to supplement or removing them all together. The information currently presented in the text is clear and not improved by the figures. I'd be happy with either moving to supplement or removing.

**RESPONSE**\

**COMMENT**\
Line 44 add OF before bumble bees

**RESPONSE**\
Done.

**COMMENT**\
Line 244 At period at end of the sentence after "Table 1)".

**RESPONSE**\
Done.

**COMMENT**\
Line 288 remove extra period

**RESPONSE**\
Done

**COMMENT**\
Line 388 Poverty isn't the best word choice here. I think you might mean "paucity" = lack of

**RESPONSE**\
Changed to "paucity."

------------------------------------------------------------------------

# Reviewer 1

Thank you to the authors for providing a thorough, detailed response to reviewer comments and completing substantial revisions to this manuscript. The quality, which I thought already high, has been dramatically improved with these revisions. The manuscript provides a clear, exciting, and novel look at montane bumble bee communities and their floral hosts. I look forward to citing this and employing the methods as I continue to use GAMs!

I have a few minor suggestions to further improve clarity:

**COMMENT**\
Line 175; The inclusion of management here sort of comes out of no where. I agree that accounting for it is crucial despite the non-significant results, but perhaps it can be broached just as controlling for a known factor of influence. That is to say, preface its description as a model term with something about why the authors have to include it. In fact, given the non-significant impact of management on the observed responses, is it necessary to discuss at length/include in figures? Seems a bit of a distraction from the otherwise clear story. If it stays in, change "grasing" in figures to "grazing".

**RESPONSE**\
This comment echoes the editor's recommendation to remove the secondary panels from Figures 3 and 4, which both reduces the attention given to management and streamlines the overall presentation of results. We have made this change, and we think it should satisfy the reviewer's concern also.

**COMMENT**\
Lines 198-212: I might move the purpose of GDM analyses up above the description, that is explain what GDM does first to preface why it makes sense to include this analysis (Perhaps move Lines 203-206 to Line 201).

**RESPONSE**\
We have rewritten this section as follows, and we agree with the reviewer that it reads better this way:

"As an initial analysis, we visualized the relationship between site-wise elevation difference and each metric of $\beta$-diversity, fitting each response with a binomial regression smooth. We then used generalized dissimilarity modeling (GDM) [@Ferrier2007-ji] to ask whether the relationship between $\beta$-diversity and elevation exhibits thresholds or other nonlinearities. In addition to estimating the relationship between differences in a response variable and differences in a predictor variable, GDM captures the slope of this relationship over the range of the predictor variable, revealing potential variation in the amount of change in the response induced by a given change in the predictor [@Ferrier2007-ji]. We fit separate GDMs for bumble bees, flora, and their (unpartitioned) interactions, respectively. To control for potential confounding effects of sampling intensity, we included sampling days per site as a covariate, and we included the geographic distances between sites as a covariate to account for potential spatial autocorrelation. The importance of the variables in each model were assessed by a permutation test [@Fitzpatrick-2021aa]."

**COMMENT**\
Line 210: Including geographical distance is a sort of bare-bones way to account for spatial autocorrelation. The fact that it was very significant in the authors models suggest there is some clear spatial clustering occurring, and I might suggest an explicit test of spatial autocorrelation to determine if the authors need to account for it using more robust methods.

**RESPONSE**\
The inclusion of geographical distances to account for spatial autocorrelation is actually one of the core features of GDM, since it was originally developed with spatial applications in mind [@Ferrier2007-ji]. There are certainly other ways to estimate and account for spatial autocorrelation, but the one we use is the standard way to do so within the GDM framework. And, while geographic distance was "significant" in our permutation tests, its effect was very small compared to the effect of elevation that we are interested in, so we don't think that the matter of spatial patterns in our data requires any further scrutiny.

**COMMENT**\
Fig 5: I would appreciate if the authors just used the full species names rather than the shorthand - there's plenty of room in the facet titles for them.

**RESPONSE**\
Done.

**COMMENT**\
Lines 270-300: Fig 3 should be Fig 6 and Fig 4 should be Fig 7.

**RESPONSE**\
Done.

**COMMENT**\
Figs. 6: I would appreciate more clear/meaningful legends that aren't shorthand/require looking down to the caption to decipher.

**RESPONSE**\

**COMMENT**\
Fig. 7: Perhaps include a vertical dark line (green?) at 1500m to indicate the approximate location of the tree line? Also, the y-axis here isn't super clear. I understand the GAM conditional effect, but what's the scale/units in this case? Conditional effect of rate of turnover?

**RESPONSE**\

**COMMENT**\
Line 295: "4-20 times higher than that of"

**RESPONSE**\

**COMMENT**\
Line 312: Or, could the bumble bee response be unimodal, but it's simply higher than the plants and this system just runs out of upper elevation to see the fall off that occurs with the plant richness?

**RESPONSE**\

------------------------------------------------------------------------

# Reviewer 2

I was happy to receive this manuscript again, and the authors have clearly put a lot of time and effort into addressing comments. I think the significant reframing makes the paper much easier to follow and the introduction is much improved. I also really enjoyed the additional framing of the results by the authors in the discussion, very insightful and interesting. I would enthusiastically recommend this for publication!

A few minor comments:

**COMMENTS**\
Could reduce specific description of results (Lines 23-33) in abstract and replace with summarized results, why they are meaningful, and end with importance for plants → bumble bees → mountain habitat theaters under climate change.

**RESPONSE**\

**COMMENT**\
Lines 23-24: "Exhibited [a] mid-elevation peak"

**RESPONSE**\

**COMMENT**\
Line 29: "Exhibited [a] well-defined threshold"

**RESPONSE**\

**COMMENT**\
Line 77: Could remove "have been found to"

**RESPONSE**\
Line 160: if = of

**COMMENT**\
Line 244: Missing a period after figure/table citation

**RESPONSE**\

**COMMENT**\
Line 288: Extra period before figure citation

**RESPONSE**\

**COMMENT**\
Line 290: Turnover shouldn't be capitalized

**RESPONSE**\

### 2022-01-05 code/ms1_beta_diversity.Rmd
started final revision

@@
 ---
-title: "Appendix 2: Beta-diversity analysis"
+title: "Appendix S3: Beta-diversity analysis"
 output: 
@@
          mean.elev = (elev1 + elev2)/2, mgmt.diff = if_else(mgmt1 == mgmt2, 0, 1)) %>%
-  pivot_longer(cols = c(BB, FL), names_to = "group", values_to = "beta") %>%
+  pivot_longer(cols = c(BB, FL), names_to = "metric", values_to = "value") %>%
   filter(elev.diff >= 0)
@@
          mgmt.diff = if_else(mgmt1 == mgmt2, 0, 1)) %>%
+  filter(site1 != site2) %>%# not interested in within-site comparisons
   gather(metric, value, -c(site1, site2, elev1, elev2, 
@@
                            x1, y1, x2, y2, mgmt1, mgmt2, mgmt.diff)) %>%
-  mutate(metric.class = case_when(
-    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
-    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
-    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
-    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"),
-    metric = factor(metric, levels = c("WN", "ST", "OS", "S", "ST.h", "ST.l", 
+  bind_rows(species_beta) %>%
+  filter(metric != "S") %>%
+  mutate(metric = factor(metric, 
+                         levels = c("OS", "ST", "WN", "ST.h", "ST.l", 
                                     "ST.lh", "WN.repl", "WN.rich", 
-                                       "OS.repl", "OS.rich")))
+                                    "OS.repl", "OS.rich", "BB", "FL"))) %>%
+  mutate(metric.name = case_when(
+    metric == "WN" ~ "interactions",
+    metric == "ST" ~ "sp. turnover",
+    metric == "OS" ~ "rewiring",
+    metric == "ST.h" ~ "BB turnover",
+    metric == "ST.l" ~ "FL turnover",
+    metric == "ST.lh" ~ "BBxFL turnover",
+    metric == "WN.repl" ~ "total (repl.)",
+    metric == "WN.rich" ~ "total (rich.)",
+    metric == "OS.repl" ~ "rewiring (repl.)",
+    metric == "OS.rich" ~ "rewiring (rich.)",
+    metric == "BB" ~ "bumble bees",
+    metric == "FL" ~ "floral hosts"),
+    metric.name = factor(metric.name, levels = c("sp. turnover", "interactions", 
+                                       "BB turnover", "FL turnover", "rewiring", 
+                                       "BBxFL turnover", "total (repl.)", "total (rich.)", 
+                                       "rewiring (repl.)", "rewiring (rich.)",
+                                       "bumble bees", "floral hosts")))
 ```
@@
 # Set palette
-p1 <- c("OS" = "#e7298a", "ST" = "#66a61e", "WN" = "black")
-
-p2 <- c("ST.h" = "#7570b3", "ST.l" = "#1b9e77", "ST.lh" = "#d95f02")
-
-# BB + FL elevation diff
-beta_plot1 <- ggplot(filter(species_beta, 
-                    site1 != site2), # no within-site comparisons
-             aes(elev.diff, beta, color = group)) +
+
+# Total turnover of bumble bees, floral hosts, and their interactions
+beta_plot1 <- ggplot(filter(beta_mit_data, 
+                    metric %in% c("WN", "BB", "FL")), 
+             aes(elev.diff, value, color = metric.name)) +
   geom_point(alpha = 0.25, pch = 16) +
@@
   labs(x = NULL,
-       y = "Species turnover",
-       color = "Taxon") +
-  theme_light()
-
-# BB * FL elevation diff (aggregate)
+       y = "Turnover",
+       color = "Level") +
+  theme_classic()
+
+# Partitioned interaction turnover
 beta_plot2 <- ggplot(filter(beta_mit_data, 
-                      site1 != site2 & # no within-site comparisons
-                      metric != "S" &
-                      metric.class == "aggregate"), 
-             aes(elev.diff, value, color = metric))  +
-  geom_point(alpha = 0.20, pch = 16) +
+                      metric %in% c("ST.l", "ST.h", "ST.lh", "OS")), 
+             aes(elev.diff, value, color = metric.name))  +
+  geom_point(alpha = 0.4, pch = 16) +
   geom_smooth(method = "glm", 
@@
               se = FALSE) +
-  scale_color_manual(values = p1, name = "Partition") +
-  xlab(NULL) +
-  ylab("Interaction turnover") +
-  theme_light()
-
-# BB * FL elevation diff (ST partition)
-beta_plot3 <- ggplot(filter(beta_mit_data, 
-                      site1 != site2 & # no within-site comparisons
-                      metric.class ==  "ST.partition"), 
-             aes(elev.diff, value, color = metric))  +
-  geom_point(alpha = 0.20, pch = 16) +
-  geom_smooth(method = "glm", 
-              method.args = list(family = "binomial"), 
-              se = FALSE) +
-  scale_color_manual(values = p2, name = "ST partition") +
+  scale_color_brewer(palette = "Set2", name = "Partition") +
+  ylab("Proportion of interaction turnover") +
   xlab("Elevation difference [m]") +
-  ylab("Prop. species turnover") +
-  theme_light()
+  theme_classic()
 
 # Patchwork plot for comparing species beta-diversity with interaction beta-diversity
-beta_plot <- (beta_plot1 / beta_plot2 / beta_plot3) + 
-  plot_annotation(tag_levels = c("A", "B", "C"))
-
-ggsave("../ms_1/figures/fig6.pdf", beta_plot, height = 8.5, width = 4.5)
+beta_plot <- (beta_plot1 / beta_plot2) + 
+  plot_annotation(tag_levels = c("A", "B"))
+
+ggsave("../ms_1/figures/fig5.pdf", beta_plot, height = 8.5, width = 5)
 ```
@@
 
-ggsave("../ms_1/figures/fig7.pdf", gdm_plot, width = 4, height = 6.5)
+ggsave("../ms_1/figures/fig6.pdf", gdm_plot, width = 4, height = 6.5)
 ```

### 2022-01-05 code/ms1_rich_abund.Rmd
started final revision

@@
 ---
-title: "Appendix 1: Richness and abundance analysis"
+title: "Appendix S2: Richness and abundance analysis"
 output: 
@@
 library(gratia)
+library(flextable)
 library(tidyverse)
@@
 
+gam1.1_plot1
+
 # Plot year effect
-gam1.1_plot2 <- ggplot(gam1.1_year, aes(value, partial)) +
+ggplot(gam1.1_year, aes(value, partial)) +
   geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
   geom_point() +
-  theme_classic() +
-  labs(y = NULL, x = NULL)
+  theme_classic()
 
 # Plot management effect
-gam1.1_plot3 <- ggplot(gam1.1_mgmt, aes(value, partial)) +
+ggplot(gam1.1_mgmt, aes(value, partial)) +
   geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
   geom_point() +
-  theme_classic() +
-  labs(y = NULL, x = NULL)
-
-# Assemble
-gam1.1_plot <- (gam1.1_plot1 | (gam1.1_plot2 / gam1.1_plot3)) 
+  theme_classic()
 ```
@@
 
+gam2.1_plot1
+
 # Plot year effect
-gam2.1_plot2 <- ggplot(gam2.1_year, aes(value, partial)) +
+ggplot(gam2.1_year, aes(value, partial)) +
   geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
   geom_point() +
-  theme_classic() +
-  labs(y = NULL, x = NULL)
+  theme_classic()
 
 # Plot management effect
-gam2.1_plot3 <- ggplot(gam2.1_mgmt, aes(value, partial)) +
+ggplot(gam2.1_mgmt, aes(value, partial)) +
   geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
   geom_point() +
-  theme_classic() +
-  labs(y = NULL, x = NULL)
-
-# Assemble
-gam2.1_plot <- gam2.1_plot1 | (gam2.1_plot2 / gam2.1_plot3)
+  theme_classic()
 ```
@@
 
+gam3.1_plot1
+
 # Plot year effect
-gam3.1_plot2 <- ggplot(gam3.1_year, aes(value, partial)) +
+ggplot(gam3.1_year, aes(value, partial)) +
   geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
   geom_point() +
-  theme_classic() +
-  labs(y = NULL, x = NULL)
+  theme_classic()
 
 # Plot management effect
-gam3.1_plot3 <- ggplot(gam3.1_mgmt, aes(value, partial)) +
+ggplot(gam3.1_mgmt, aes(value, partial)) +
   geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
   geom_point() +
-  theme_classic() +
-  labs(y = NULL, x = NULL)
-
-# Assemble
-gam3.1_plot <- gam3.1_plot1 | (gam3.1_plot2 / gam3.1_plot3)
-
-# Tabulate
-library(flextable)
-as_flextable(gam3.1)
+  theme_classic()
 ```
@@
 ```{r message=FALSE}
-gam_plot_richness_ <- ( gam1.1_plot / gam2.1_plot / gam3.1_plot ) +
+gam_plot_richness_ <- ( gam1.1_plot1 / gam2.1_plot1 / gam3.1_plot1 ) +
   plot_annotation(tag_levels = 'A') +
@@
   draw_plot(gam_plot_richness_) +
-  draw_image(BB, scale = 0.15, x = -0.18, y = 0.28) +
-  draw_image(FL, scale = 0.125, x = -0.18, y = -0.025) +
-  draw_image(BBxFL, scale = 0.25, x = -0.18, y = -0.38)
+  draw_image(BB, scale = 0.2, x = 0.07, y = 0.3) +
+  draw_image(FL, scale = 0.125, x = 0.07, y = -0.025) +
+  draw_image(BBxFL, scale = 0.35, x = 0.07, y = -0.35)
 
@@
 
-ggsave("../ms_1/figures/fig3.pdf", gam_plot_richness, width = 6.5, height = 8.5)
+ggsave("../ms_1/figures/fig2.pdf", gam_plot_richness, width = 4, height = 6.5)
 ```
@@
 
+gam4.1_plot1
+
 # Plot year effect
-gam4.1_plot2 <- ggplot(gam4.1_year, aes(value, partial)) +
+ggplot(gam4.1_year, aes(value, partial)) +
   geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
   geom_point() +
-  theme_classic() +
-  labs(y = NULL, x = NULL)
+  theme_classic()
 
 # Plot management effect
-gam4.1_plot3 <- ggplot(gam4.1_mgmt, aes(value, partial)) +
+ggplot(gam4.1_mgmt, aes(value, partial)) +
   geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
   geom_point() +
-  theme_classic() +
-  labs(y = NULL, x = NULL)
-
-# Assemble
-gam4.1_plot <- gam4.1_plot1 | (gam4.1_plot2 / gam4.1_plot3)
+  theme_classic()
 ```
@@
 
+gam5.1_plot1
+
 # Plot year effect
-gam5.1_plot2 <- ggplot(gam5.1_year, aes(value, partial)) +
+ggplot(gam5.1_year, aes(value, partial)) +
   geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
   geom_point() +
-  theme_classic() +
-  labs(y = NULL, x = NULL)
+  theme_classic()
 
 # Plot management effect
-gam5.1_plot3 <- ggplot(gam5.1_mgmt, aes(value, partial)) +
+ggplot(gam5.1_mgmt, aes(value, partial)) +
   geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
   geom_point() +
-  theme_classic() +
-  labs(y = NULL, x = NULL)
-
-# Assemble
-gam5.1_plot <- gam5.1_plot1 | (gam5.1_plot2 / gam5.1_plot3)
+  theme_classic()
 ```
@@
 ```{r message=FALSE}
-gam_plot_abundance_ <- ( gam4.1_plot / gam5.1_plot) +
+gam_plot_abundance_ <- ( gam4.1_plot1 / gam5.1_plot1 ) +
   plot_annotation(tag_levels = 'A') +
@@
   draw_plot(gam_plot_abundance_) +
-  draw_image(BB, scale = 0.15, x = -0.18, y = 0.16) +
-  draw_image(FL, scale = 0.2, x = -0.18, y = -0.3)
+  draw_image(BB, scale = 0.25, x = 0.07, y = 0.15) +
+  draw_image(FL, scale = 0.18, x = 0.07, y = -0.3)
 
@@
 
-ggsave("../ms_1/figures/fig4.pdf", gam_plot_abundance, width = 6.5, height = 5.5)
+ggsave("../ms_1/figures/fig3.pdf", gam_plot_abundance, width = 4, height = 6.5)
 ```
@@
   left_join(gam6.1_pval, by = c("smooth" = "term")) %>%
-  mutate(sig = factor(ifelse(p.value < 0.05, "p < 0.05", "ns")))
+  mutate(sig = factor(ifelse(p.value < 0.05, "p < 0.05", "ns"))) %>%
+  # as per reviewer's request for full species names
+  mutate(bb.sp.full = case_when(
+    bb.sp ==  "pyre" ~ "B. pyrenaeus",
+    bb.sp ==  "mend" ~ "B. mendax",
+    bb.sp ==  "mont" ~ "B. monticola",
+    bb.sp ==  "muci" ~ "B. mucidus",
+    bb.sp ==  "soro" ~ "B. soroeensis",
+    bb.sp ==  "wurf" ~ "B. wurflenii",
+    bb.sp ==  "gers" ~ "B. gerstaeckeri",
+    bb.sp ==  "lapi" ~ "B. lapidarius",
+    bb.sp ==  "jone" ~ "B. jonellus",
+    bb.sp ==  "telu" ~ "B. terrestris/lucorum",
+    bb.sp ==  "hort" ~ "B. hortorum",
+    bb.sp ==  "hypn" ~ "B. hypnorum",
+    bb.sp ==  "prat" ~ "B. pratorum",
+    bb.sp ==  "pasc" ~ "B. pascuorum",
+    bb.sp ==  "humi" ~ "B. humilis",
+    bb.sp ==  "psit" ~ "B. psithyrus"
+  ))
 
@@
   scale_fill_gradient2(low = "#4dac26", mid = "gray75", high = "#d01c8b") +
-  theme_light(12) +
+  theme_light(10) +
   labs(y = "Conditional effect", x = "Elevation [m]", fill = "Slope",
        linetype = "Significance") +
-  facet_wrap(~reorder(bb.sp, -diff)) +
-  theme()
-
-ggsave("../ms_1/figures/fig5.pdf", gam6.1_plot1, height = 6.5, width = 6.5)
+  facet_wrap(~reorder(bb.sp.full, -diff)) +
+  theme(strip.text = element_text(face = "italic", size = 8))
+
+ggsave("../ms_1/figures/fig4.pdf", gam6.1_plot1, height = 6.5, width = 6.5)
 ```
@@
 ## 3.3 Compile tabulated summaries
-```{r message=FALSE}
+```{r message=FALSE, warning=FALSE, eval=FALSE}
 gam_tab.rich <- gam_tab1(x1 = gam1.1, x2 = gam2.1, x3 = gam3.1, 
@@
 
-ggsave("../output/bb_range_ms1.pdf")
-ggsave("../output/bb_range_ms1.png")
-
 # Compare bumble bee species in terms of proportional abundance (unit = site)
@@
 
-ggsave("../output/bb_abundsum.pdf")
-ggsave("../output/bb_abundsum.png")
 ```

### 2022-01-05 code/ms1_sampling.Rmd
updating just in case
started final revision

@@
 ---
-title: "sampling"
-output: html_document
+title: "Appendix S1: Sampling patterns"
+output: 
+  pdf_document:
+    # toc: true
+    # toc_depth: 3
 ---
@@
   ylab("Total bumble bees\nobserved") +
-  theme_light(13) +
+  theme_light(10) +
   theme(axis.title.x = element_blank(),
@@
   ylab("Total samples") +
-  theme_light(13) +
+  theme_light(10) +
   theme(axis.title.x = element_blank(),
@@
   ylab("Bumble bees\nobserved / sample") +
-  theme_light(13) +
+  theme_light(10) +
   theme(axis.title.x = element_blank(),
@@
                                                     "1970-07-01", "1970-08-01", "1970-09-01"))) +
-  theme_light(13) +
+  theme_light(10) +
   labs(x = NULL,
@@
 
-samplingplot <- ((samplingplot1 / samplingplot2 / samplingplot3) | samplingplot4) +
-  plot_layout(widths = c(1, 3)) +
-  plot_annotation(tag_levels = 'A')
+# samplingplot <- ((samplingplot1 / samplingplot2 / samplingplot3) | samplingplot4) +
+#   plot_layout(widths = c(1, 3)) +
+#   plot_annotation(tag_levels = 'A')
 
-ggsave("../ms_1/figures/fig2.pdf", samplingplot, height = 6, width = 10.5)
+#ggsave("../ms_1/figures/fig2.pdf", samplingplot, height = 6, width = 10.5)
+
+((samplingplot1 | samplingplot2 | samplingplot3) / samplingplot4)
 ```

### 2022-01-05 ms_1/ms_revision2.rmd
started final revision

--- 2021-11-05 ms_1/ms_revision1.Rmd
@@
 title: "Contrasting patterns of richness, abundance, and turnover in mountain bumble bees and their floral hosts"
-
 author: 
@@
 
-This study also uses novel code. All data and code required to reproduce our analyses are archived on Dryad: https://doi.org/10.5061/dryad.s1rn8pk7f
+All data and code required to reproduce our analyses are archived on Dryad: https://doi.org/10.5061/dryad.s1rn8pk7f
 
@@
 We recorded bumble- bee-flower interactions over three years along an 1400 m elevational gradient in the German Alps.
-Using nonlinear modeling, we analyzed the elevational patterns at the nested levels of species abundance, species $\beta$-diversity, and interaction $\beta$-diversity.
+Using nonlinear modeling, we analyzed elevational patterns at the nested levels of species abundance, species $\beta$-diversity, and interaction $\beta$-diversity.
 While floral richness exhibited mid-elevation peak, bumble bee richness increased with elevation before leveling off at the highest sites, demonstrating the exceptional adaptation of these bees to cold temperatures and short growing seasons.
@@
 
-In this paper, we present a study based on ~13,000 interactions between bumble bees (Hymenoptera: *Bombus*) and their floral hosts recorded along a 1400 m elevation gradient, complemented by corresponding surveys of floral abundance.
+In this paper, we present a study based on ~13,000 interactions between bumble bees and their floral hosts recorded along a 1400 m elevation gradient, complemented by corresponding surveys of floral abundance.
 Using this dataset, we investigate elevational responses of richness, abundance, and $\beta$-diversity, both at the level of species (bumble bees and flowering plants, respectively) and at the level of interactions (bumble- bee-flower pairs).
@@
 Bumblebees observed on a given flower were counted as floral visitors.
-Bumble bee queens were identified to species level in the field, while workers and males were stored in individually labeled tubes in the freezer for later identification in the laboratory according to [@Amiet1996-jz].
+Bumble bee queens were identified to species level in the field, while workers and males were collected and stored in individually labeled tubes in the freezer for later identification in the laboratory according to @Amiet1996-jz.
 Floral visitation by male bumble bees was recorded during visitation sampling, but we chose to analyze only visitation by queens and workers.
 Males are often found resting on flowers, particularly late in the year, so their presence on a flower cannot necessarily be interpreted as a foraging visit.
-In conjunction with bumble bee observations, we estimated the flower cover of each herbaceous or shrubby plant species within the study plot to the nearest 0.01 m^2^; this was done by exhaustive visual inspection of the study plot rather than timed transect walks.
+In conjunction with bumble bee observations, we estimated the flower cover of each herbaceous or shrubby plant species within the study plot to the nearest 0.01 m^2^ by exhaustive visual inspection of the study plot.
 To avoid observer bias, floral cover estimates were always performed by the same investigator (K.
@@
 Although per-sample effort was standardized across all years, the total number of samples (i.e.
-site-dates) and bumble bee observations varied across years **(Figure 2A-C)**.
+site-dates) and of bumble bee observations varied across years.
 Since the start of field sampling each year was determined by snowmelt, lower sites were sampled as early as April while higher sites could not be sampled until June or July.
-This resulted in a temporally staggered pattern of sampling in 2010 and 2012, while in 2011 a combination of early snowmelt and a late start to field activities resulted in sampling that was approximately aligned in time across the whole elevation gradient **(Figure 2D)**.
+This resulted in a temporally staggered pattern of sampling in 2010 and 2012, while in 2011 a combination of early snowmelt and a late start to field activities resulted in sampling that was approximately aligned in time across the whole elevation gradient.
+See Appendix S1 for a detailed description of sampling patterns.
 
@@
 psithryus*.
-*Bombus hypnorum*, *B.
-humilis*, and *B.
-lapidarius* were extremely rare in our samples, so we omitted them from abundance analyses but retained them for $\beta$-diversity analyses.
 
@@
 As an initial analysis, we visualized the relationship between site-wise elevation difference and each metric of $\beta$-diversity, fitting each response with a binomial regression smooth.
-To investigate the relationship between elevation and $\beta$-diversity more deeply, we performed a second analysis using generalized dissimilarity modeling (GDM) [@Ferrier2007-ji] to analyze the elevational variation in the $\beta$-diversity of bumble bees, flora, and their (unpartitioned) interactions, respectively.
+We then used generalized dissimilarity modeling (GDM) [@Ferrier2007-ji] to ask whether the relationship between $\beta$-diversity and elevation exhibits thresholds or other nonlinearities.
 In addition to estimating the relationship between differences in a response variable and differences in a predictor variable, GDM captures the slope of this relationship over the range of the predictor variable, revealing potential variation in the amount of change in the response induced by a given change in the predictor [@Ferrier2007-ji].
-This enabled us to ask whether the relationship between $\beta$-diversity and elevation exhibits thresholds or other nonlinearities.
-To control for potential confounding effects of sampling intensity, sampling days per site was included as a covariate.
-We also included the geographic distances between sites as a covariate to account for potential spatial autocorrelation.
+We fit separate GDMs for bumble bees, flora, and their (unpartitioned) interactions, respectively.
+To control for potential confounding effects of sampling intensity, we included sampling days per site as a covariate, and we included the geographic distances between sites as a covariate to account for potential spatial autocorrelation.
 The importance of the variables in each model were assessed by a permutation test [@Fitzpatrick-2021aa].
@@
 
-The response of bumble bee richness to elevation exhibited an approximately linear increase up to 1600-1700 m and then leveled off and declined slightly at the upper limit of the gradient **(Figure 3A)**.
-The year 2012 had a negative effect (p = 0.04) on bumble bee richness relative to the reference year 2010 **(Figure 3B)**, but neither mowing nor abandonment had a significant effect on bumble bee richness relative to the reference management level of grazing (p > 0.05)  **(Figure 3C; Table 1)** In contrast, the response of floral richness exhibited a clear unimodal pattern, peaking between 1300-1500 m **(Figure 3D)**.
-The years 2011 (p < 0.001) and 2012 (p < 0.03) had a negative effects on floral richness relative to the reference year 2010 **(Figure 3E)**, but again management had no significant effect (p > 0.1) **(Figure 3F; Table 1)**.
-The response of interaction richness to elevation was intermediate to that of bumble bee and floral richness, peaking around 1500 m and leveling off at higher elevation **(Figure 3G)**.
-The effect of year 2011 on interaction richness was positive (p < 0.05) and that of year 2012 negative (p < 0.001) **(Figure 3H)**, but management did not have a significant effect (p > 0.1) **(Figure 3I; Table 1)**.
+The response of bumble bee richness to elevation exhibited an approximately linear increase up to 1600-1700 m and then leveled off and declined slightly at the upper limit of the gradient **(Figure 2A)**.
+The year 2012 had a negative effect (p = 0.04) on bumble bee richness relative to the reference year 2010 **(Figure 2B)**, but neither mowing nor abandonment had a significant effect on bumble bee richness relative to the reference management level of grazing (p > 0.05)  **(Figure 2C; Table 1)**.
+In contrast, the response of floral richness exhibited a clear unimodal pattern, peaking between 1300-1500 m **(Figure 2D)**.
+The years 2011 (p < 0.001) and 2012 (p < 0.03) had a negative effects on floral richness relative to the reference year 2010 **(Figure 2E)**, but again management had no significant effect (p > 0.1) **(Figure 2F; Table 1)**.
+The response of interaction richness to elevation was intermediate to that of bumble bee and floral richness, peaking around 1500 m and leveling off at higher elevation **(Figure 2G)**.
+The effect of year 2011 on interaction richness was positive (p < 0.05) and that of year 2012 negative (p < 0.001) **(Figure 2H)**, but management did not have a significant effect (p > 0.1) **(Figure 2I; Table 1)**.
 
-Cumulative bumble bee abundance did not respond significantly to elevation (p > 0.1) **(Figure 4A)** or management (p > 0.1) **(Figure 4C)**, but the year 2011 had a positive effect (p = 0.03) on bumble bee abundance relative to the reference year of 2010 **(Figure 4B; Table 2)**.
-Floral abundance, however, peaked around 1000 and exhibited an approximately quadratic decline with increasing elevation (p = 0.003) **(Figure 4D)**.
-The effect of year 2012 was negative (p = 0.002) relative to the reference year 2010 (p = 0.02) **(Figure 4E)**, and the effect of abandoned meadows on floral abundance was positive compared to the reference level of grazed meadows **(Figure 4F; Table 2)**.
+Cumulative bumble bee abundance did not respond significantly to elevation (p > 0.1) **(Figure 3A)** or management (p > 0.1) **(Figure 3C)**, but the year 2011 had a positive effect (p = 0.03) on bumble bee abundance relative to the reference year of 2010 **(Figure 3B; Table 2)**.
+Floral abundance, however, peaked around 1000 and exhibited an approximately quadratic decline with increasing elevation (p = 0.003) **(Figure 3D)**.
+The effect of year 2012 was negative (p = 0.002) relative to the reference year 2010 (p = 0.02) **(Figure 3E)**, and the effect of abandoned meadows on floral abundance was positive compared to the reference level of grazed meadows **(Figure 3F; Table 2)**.
 
-At the species level, bumble bees exhibited variable and sometimes divergent responses to elevation **(Figure 5; Table 3)**.
+At the species level, bumble bees exhibited variable and sometimes divergent responses to elevation **(Figure 4; Table 3)**.
 *Bombus pyrenaeus*, *B.
@@
 
-Floral $\beta$-diversity among our study sites was very high overall and responded steeply to elevation difference between sites **(Figure 3A)**.
+Floral $\beta$-diversity among our study sites was very high overall and responded steeply to elevation difference between sites **(Figure 5A)**.
 Even sites at similar elevation exhibited ~40% species turnover, and the most vertically distant sites (elevation difference > 1250 m) differed by more than 85%.
-Bumble bee $\beta$-diversity was, in comparison to floral $\beta$-diversity, both lower overall and less responsive to elevational difference between sites **(Figure 3A)**.
+Bumble bee $\beta$-diversity was, in comparison to floral $\beta$-diversity, both lower overall and less responsive to elevational difference between sites **(Figure 5A)**.
 Sites at similar elevation exhibited ~25% species turnover, and species turnover between the most vertically distant sites remained less than 50%.
 
-Total interaction $\beta$-diversity was >75% between sites at similar elevation and approached perfect dissimilarity in the most vertically distant sites **(Figure 3B)**.
+Total interaction $\beta$-diversity was >75% between sites at similar elevation and approached perfect dissimilarity in the most vertically distant sites **(Figure 5B)**.
 Partitioning revealed that total interaction $\beta$-diversity was driven primarily by species turnover, which accounted for ~50% of total $\beta$-diversity for sites at similar elevation and ~90% of total $\beta$-diversity for the most vertically distant sites.
-Species turnover consisted mainly of turnover in the floral community **(Figure 3C)**, though the joint turnover of plants and bumble bees accounted for ~25% of total species turnover between the most vertically distant sites.
-Interaction rewiring was most significant between sites at similar elevation, where it accounted for ~30% of total $\beta$-diversity, but its share of total $\beta$-diversity declined to ~10% in the most vertically distant sites **(Figure 3B)**.
+Species turnover consisted mainly of turnover in the floral community **(Figure 5C)**, though the joint turnover of plants and bumble bees accounted for ~25% of total species turnover between the most vertically distant sites.
+Interaction rewiring was most significant between sites at similar elevation, where it accounted for ~30% of total $\beta$-diversity, but its share of total $\beta$-diversity declined to ~10% in the most vertically distant sites **(Figure 5B)**.
 
-Bumble bees, flora, and their interactions exhibited marked differences in their rates of $\beta$-diversity in response to elevation.
-**(Figure 4)**.
+Bumble bees, flora, and their interactions exhibited marked differences in their rates of $\beta$-diversity in response to elevation **(Figure 6)**.
 Bumble bee turnover was relatively high from ~600-1100 m, briefly leveled off, and then accelerated again around 1250 m.
@@
 Interestingly, the tree line did not constitute a discrete transition to a new equilibrium of species composition, but rather a threshold marking the beginning of a process of $\beta$-diversity acceleration that continued for the remainder of the elevation gradient.
-While the most obvious effect of the tree line may be the relatively discrete transition from forest to scrub/grassland, the tree line also initiates a process of angiosperm species attrition that proceeds at a rate of about 40 species per 100 m [@Korner2003-sc], and the accelerated $\beta$-diversity we observed above the tree line mirrors simultaneous patterns of decreasing floral richness and abundance **(Figures 3B, 4B)**.
+While the most obvious effect of the tree line may be the relatively discrete transition from forest to scrub/grassland, the tree line also initiates a process of angiosperm species attrition that proceeds at a rate of about 40 species per 100 m [@Korner2003-sc], and the accelerated $\beta$-diversity we observed above the tree line mirrors simultaneous patterns of decreasing floral richness and abundance **(Figures 2B, 3B)**.
 
@@
 
-The poverty of floral resources at the highest elevations raises the question of why bumble bees choose to forage at such elevations at all.
+The paucity of floral resources at the highest elevations raises the question of why bumble bees choose to forage at such elevations at all.
 For some species, the selection of high elevation habitats may be influenced by thermal tolerances [@Oyen2016-ng], but that fails to explain the presence of elevation-generalist species (including the majority of the species in our study system) in resource-poor high-elevation habitats.
@@
 Similarly, higher elevations could also offer an escape from density-dependent parasite and pathogen pressure [@Parsche2018-ar].
-Finally, it is worth speculating that an affinity for open, high-elevation habitats might be an evolutionary legacy in mountain bumble bees, since the existence of flower-rich meadows in the forest matrix below the tree line is largely an artifact of human land use arising mainly in the last 500 years of bumble bees' evolutionary history [@Hejcman2013-ca].
+Finally, it is worth speculating that an affinity for open, high-elevation habitats might be an evolutionary legacy in mountain bumble bees, since the existence of flower-rich meadows in the forest matrix below the tree line is largely an artifact of human land use arising in the last 500 years of bumble bees' evolutionary history [@Hejcman2013-ca].
 
@@
 
-**Figure 2:** Total bumble bee observations per year (A), total samples per year (B), bumble bee observations per sample per year (C), and the distribution of samples through elevation and time (D).
-Each sample represents a standardized transect walk for a given site and date, each observation represents a recorded interaction between a bumble bee and flower.
-
-**Figure 3:** Conditional effects of elevation, year, and management on the richness of bumble bee species (A), floral species (B), and bumble- bee-flower interactions (C).
+**Figure 2:** Conditional effects of elevation, year, and management on the richness of bumble bee species (A), floral species (B), and bumble- bee-flower interactions (C).
 Effects are plotted on the link scale, and shaded bands depict 95% confidence intervals.
@@
 
-**Figure 4:** Conditional effects of elevation, year, and management on the peak abundance of bumble bees (A) and flower cover (B).
+**Figure 3:** Conditional effects of elevation, year, and management on the peak abundance of bumble bees (A) and flower cover (B).
 Effects are plotted on the link scale, and shaded bands depict 95% confidence intervals.
@@
 
-**Figure 5:** Conditional effects of elevation on the peak abundance of each bumble bee species.
+**Figure 4:** Conditional effects of elevation on the peak abundance of each bumble bee species.
 Effects are plotted on the link scale, and shaded bands depict 95% confidence intervals.
@@
 
-**Figure 6:** Species-level (A) and interaction-level (B, C) $\beta$-diversity (in terms of proportional turnover) plotted against elevational difference.
-Each point represents the $\beta$-diversity (or $\beta$-diversity partition) between a pair of sites, and lines represent the overall relationship between $\beta$-diversity and elevation difference using binomial regression smooths.
-Interaction $\beta$-diversity (B, C) is partitioned using Poisot et al.’s (2012) notation: WN = unpartitioned $\beta$-diversity, ST = $\beta$-diversity due to species turnover, OS =  $\beta$-diversity due to interaction rewiring, ST.h = $\beta$-diversity due to species turnover in the higher trophic level (bumble bees), ST.l = $\beta$-diversity due to species turnover in the lower trophic level (plants), and ST.lh =  $\beta$-diversity due to joint species turnover in higher and lower trophic levels.
+**Figure 5:** Species-level (A) and interaction-level (B, C) $\beta$-diversity (in terms of proportional turnover) plotted against elevational difference.
+Each point represents the turnover (or component of turnover) between a pair of sites, and lines represent the overall relationship between turnover and elevation difference using binomial regression smooths.
+Interaction turnover (B, C) is partitioned using Poisot et al.’s (2012) components but with more explicit nomenclature: total = unpartitioned interaction turnover, sp.
+turnover = interaction turnover due to species turnover, OS =  interaction turnover due to interaction rewiring, ST.h = interaction turnover due to species turnover in the higher trophic level (bumble bees), ST.l = $\beta$-diversity due to species turnover in the lower trophic level (plants), and ST.lh =  $\beta$-diversity due to joint species turnover in higher and lower trophic levels.
 
-**Figure 7:** GDM splines of species- and interaction-level $\beta$-diversity in response to elevation.
+**Figure 6:** GDM splines of species- and interaction-level $\beta$-diversity in response to elevation.
 The maximum height of each spline represents the conditional (i.e.
@@
 
-```{r fig2, echo = FALSE, out.width="100%"}
+```{r fig3, echo = FALSE}
 knitr::include_graphics("../ms_1/figures/fig2.pdf")
@@
 
-```{r fig3, echo = FALSE}
+```{r fig4, echo = FALSE}
 knitr::include_graphics("../ms_1/figures/fig3.pdf")
-```
-
-\newpage
-
-**Figure 4**
-
-```{r fig4, echo = FALSE}
-knitr::include_graphics("../ms_1/figures/fig4.pdf")
 ```
@@
 ```{r fig5, echo = FALSE}
-knitr::include_graphics("../ms_1/figures/fig5.pdf")
+knitr::include_graphics("../ms_1/figures/fig4.pdf")
 ```
@@
 ```{r fig6, echo = FALSE}
-knitr::include_graphics("../ms_1/figures/fig6.pdf")
+knitr::include_graphics("../ms_1/figures/fig5.pdf")
 ```
@@
 ```{r fig7, echo = FALSE}
-knitr::include_graphics("../ms_1/figures/fig7.pdf")
+knitr::include_graphics("../ms_1/figures/fig6.pdf")
 ```
